### 2021-08-04 main/main.Rmd
init paper with some outlines

---
title: A template for the *arxiv* style
authors:
  - name: H.Sherry Zhang
    department: Department of Econometrics and Business Statistics 
    affiliation: Monash University 
    location: Melbourne, Australia
    email:  huize.zhang@monash.edu
  - name: Dianne Cook
    department: Department of Econometrics and Business Statistics 
    affiliation: Monash University 
    location: Melbourne, Australia
    email:  dicook@monash.edu
  - name: Ursula Laa
    department: Institute of Statistics
    affiliation: University of Natural Resources and Life Sciences
    location: Vienna, Austria
    email:  ursula.laa@boku.ac.at  
  - name: Nicolas Langrené
    department: 34 Village Street, Docklands VIC 3008 Australia
    affiliation: CSIRO Data61 
    location: Melbourne, Australia
    email: nicolas.langrene@csiro.au
  - name: Patricia Menéndez
    department: Department of Econometrics and Business Statistics 
    affiliation: Monash University 
    location: Melbourne, Australia
    email:  patricia.menendez@monash.edu 
abstract: |
  Enter the text of your abstract here.
keywords:
  - blah
  - blee
  - bloo
  - these are optional and can be removed
bibliography: references.bib
biblio-style: unsrt
output: rticles::arxiv_article
---

# Introduction

Spatio-temporal data

Motivation: Many focus on spatio/ temporal solely -> spatio-temporal vector data structure for data analysis

sections

# Literature Review

Existing R pkgs for spatio/ temporal data

- `sp` + `spacetime` (time series data from package `xts`)

- `sf` 
  - simple features geometry 
  - only spatio
  - good for geographical but these geometry types i.e. multipolygon, can be hard to manipulate. Would be nice to have a vector data structure

- `stars`
  - a newer structure allow both spatio + temporal 
  - array

# A new data structure for spatio-temporal data

Intro to cubble:

- list-column: rowwise_df with temporal variables, including the time index, nested. 
  - Focus on spatio: those output per station

- long form: grouped_df 
  - Focus on temporal

Compatible with tidyverse manipulation and tsibble

# Examples

Daily climate data (prcp, tmax, and tmin) from RNOAA - lots of stations across Australia

An exploratory data analysis questions: What's the climate profile look like in Australia

  - General features: Any general trend/ fluctuation in prcp, tmax, and tmin?
  - Local features: Any station stands out from the crowd?

## Manipulation

  - data quality check: filter out stations have variables not properly recorded 
  - data summary: 
    - daily -> monthly/ weekly, 
    - summarise by mean for tmax/ tmin, sum for prcp
  -

## Graphics

Static + interactive -> tooltip to show additional information upon hovering

  - Where are those stations on the map? 
    - Mention mostly aero, airport, and lighthouse

# Summary

### 2021-08-04 main/main.tex
init paper with some outlines

\documentclass{article}

\usepackage{arxiv}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{lmodern}        % https://github.com/rstudio/rticles/issues/343
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{graphicx}

\title{A template for the \emph{arxiv} style}

\author{
    H.Sherry Zhang
   \\
    Department of Econometrics and Business Statistics \\
    Monash University \\
  Melbourne, Australia \\
  \texttt{\href{mailto:huize.zhang@monash.edu}{\nolinkurl{huize.zhang@monash.edu}}} \\
   \And
    Dianne Cook
   \\
    Department of Econometrics and Business Statistics \\
    Monash University \\
  Melbourne, Australia \\
  \texttt{\href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}} \\
   \And
    Ursula Laa
   \\
    Institute of Statistics \\
    University of Natural Resources and Life Sciences \\
  Vienna, Austria \\
  \texttt{\href{mailto:ursula.laa@boku.ac.at}{\nolinkurl{ursula.laa@boku.ac.at}}} \\
   \And
    Nicolas Langrené
   \\
    34 Village Street, Docklands VIC 3008 Australia \\
    CSIRO Data61 \\
  Melbourne, Australia \\
  \texttt{\href{mailto:nicolas.langrene@csiro.au}{\nolinkurl{nicolas.langrene@csiro.au}}} \\
   \And
    Patricia Menéndez
   \\
    Department of Econometrics and Business Statistics \\
    Monash University \\
  Melbourne, Australia \\
  \texttt{\href{mailto:patricia.menendez@monash.edu}{\nolinkurl{patricia.menendez@monash.edu}}} \\
  }

% Pandoc citation processing

\begin{document}
\maketitle

\def\tightlist{}

\begin{abstract}
Enter the text of your abstract here.
\end{abstract}

\keywords{
    blah
   \and
    blee
   \and
    bloo
   \and
    these are optional and can be removed
  }

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Spatio-temporal data

Motivation: Many focus on spatio/ temporal solely -\textgreater{}
spatio-temporal vector data structure for data analysis

sections

\hypertarget{literature-review}{%
\section{Literature Review}\label{literature-review}}

Existing R pkgs for spatio/ temporal data

\begin{itemize}
\item
  \texttt{sp} + \texttt{spacetime} (time series data from package
  \texttt{xts})
\item
  \texttt{sf}

  \begin{itemize}
  \tightlist
  \item
    simple features geometry
  \item
    only spatio
  \item
    good for geographical but these geometry types i.e.~multipolygon,
    can be hard to manipulate. Would be nice to have a vector data
    structure
  \end{itemize}
\item
  \texttt{stars}

  \begin{itemize}
  \tightlist
  \item
    a newer structure allow both spatio + temporal
  \item
    array
  \end{itemize}
\end{itemize}

\hypertarget{a-new-data-structure-for-spatio-temporal-data}{%
\section{A new data structure for spatio-temporal
data}\label{a-new-data-structure-for-spatio-temporal-data}}

Intro to cubble:

\begin{itemize}
\tightlist
\item
  list-column: rowwise\_df with temporal variables, including the time
  index, nested.

  \begin{itemize}
  \tightlist
  \item
    Focus on spatio: those output per station
  \end{itemize}
\item
  long form: grouped\_df

  \begin{itemize}
  \tightlist
  \item
    Focus on temporal
  \end{itemize}
\end{itemize}

Compatible with tidyverse manipulation and tsibble

\hypertarget{examples}{%
\section{Examples}\label{examples}}

Daily climate data (prcp, tmax, and tmin) from RNOAA - lots of stations
across Australia

An exploratory data analysis questions: What's the climate profile look
like in Australia

\begin{itemize}
\tightlist
\item
  General features: Any general trend/ fluctuation in prcp, tmax, and
  tmin?
\item
  Local features: Any station stands out from the crowd?
\end{itemize}

\hypertarget{manipulation}{%
\subsection{Manipulation}\label{manipulation}}

\begin{itemize}
\tightlist
\item
  data quality check: filter out stations have variables not properly
  recorded
\item
  data summary:

  \begin{itemize}
  \tightlist
  \item
    daily -\textgreater{} monthly/ weekly,
  \item
    summarise by mean for tmax/ tmin, sum for prcp
  \end{itemize}
\item
\end{itemize}

\hypertarget{graphics}{%
\subsection{Graphics}\label{graphics}}

Static + interactive -\textgreater{} tooltip to show additional
information upon hovering

\begin{itemize}
\tightlist
\item
  Where are those stations on the map?

  \begin{itemize}
  \tightlist
  \item
    Mention mostly aero, airport, and lighthouse
  \end{itemize}
\end{itemize}

\hypertarget{summary}{%
\section{Summary}\label{summary}}

\bibliographystyle{unsrt}
\bibliography{references.bib}

\end{document}

### 2021-08-04 main/main.Rmd
draft on existing spatio + temporal R pkgs

@@
 
+<!-- While satellite imagery can record information on a continuous spatial domain, other spatial information are recorded at a fixed set of locations characterised by latitude, longitude, and other spatial metadata.
+Examples of this type of data includes climate measures from weather observation stations and river level data recorded by electronic gauges.
+Analysing this data requires less considerations on the geometry type and map projection while more on how measures in these fixed locations changes across the time domain and whether these changes are related for adjacent locations.
+-->
+
 Motivation: Many focus on spatio/ temporal solely -> spatio-temporal vector data structure for data analysis
@@
 
-# Literature Review
+# Existing data structure for spatio and temporal data
 
-Existing R pkgs for spatio/ temporal data
+There has been a large class of implementations dedicated to processing the spatial data.
+This includes `sf` [@pebesma2018simple] and its precedent `sp` [@pebesma2005s] for ...
+and `raster` [@raster] and `terra` [@terra] for raster data.
+While these implementations specialised in geographic manipulations with different type of simple features, it doesn't incorporate a temporal dimension in the data structure.
+Project like `stars` [@stars] and `spacetime` [@spacetimebook] by R-Spatial allows for both space and time dimension for raster and vector data, but the underlying data structure is a multi-dimensional array, which could be difficult to operate for R users who are more familiar with the operation in 2D dataframe/ tibble.
 
-- `sp` + `spacetime` (time series data from package `xts`)
+<!-- - `sp` + `spacetime` (time series data from package `xts`) -->
 
-- `sf` 
-  - simple features geometry 
-  - only spatio
-  - good for geographical but these geometry types i.e. multipolygon, can be hard to manipulate. Would be nice to have a vector data structure
+<!-- - `sf`  -->
+<!--   - simple features geometry  -->
+<!--   - only spatio -->
+<!--   - good for geographical but these geometry types i.e. multipolygon, can be hard to manipulate. Would be nice to have a vector data structure -->
 
-- `stars`
-  - a newer structure allow both spatio + temporal 
-  - array
+<!-- - `stars` -->
+<!--   - a newer structure allow both spatio + temporal  -->
+<!--   - array -->
+
+In the temporal aspect, the `tsibble` [@tsibbles] structure and its tidyverts ecosystem have provided a [...
+] workflow to work with temporal data.
+In a tsibble structure, temporal data is characterised by `index` and `key` where `index` is the temporal identifier and `key` is the identifier for multiple series, which could be used as a spatio identifier.
+However, a tsibble object, by construction, always requires the `index` in its structure.
+This makes it less appealing for spatio-temporal data since the output of calculated spatio-specific variables (i.e.
+features of each series) don't have the time dimension.
+Analysts will either need to have an additional step to join this output to the original tsibble or operate with variables stored in two separate objects.
+In addition, the long form structure of a tsibble object means spatio variables (i.e.
+longitude, latitude, and features of each series if joined back to the tsibble) of each spatio identifier will be repetitively recorded at each timestamp.
+This repetition is unnecessary and would inflate the object size for long series.
 

### 2021-08-04 main/main.tex
draft on existing spatio + temporal R pkgs

@@
 % Pandoc citation processing
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+% for Pandoc 2.8 to 2.10.1
+\newenvironment{cslreferences}%
+  {}%
+  {\par}
+% For Pandoc 2.11+
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
+  % set entry spacing
+  \ifnum #2 > 0
+  \setlength{\parskip}{#2\baselineskip}
+  \fi
+ }%
+ {}
+\usepackage{calc} % for calculating minipage widths
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 
@@
 
-\hypertarget{literature-review}{%
-\section{Literature Review}\label{literature-review}}
+\hypertarget{existing-data-structure-for-spatio-and-temporal-data}{%
+\section{Existing data structure for spatio and temporal
+data}\label{existing-data-structure-for-spatio-and-temporal-data}}
 
-Existing R pkgs for spatio/ temporal data
+There has been a large class of implementations dedicated to processing
+the spatial data. This includes \texttt{sf} (E. J. Pebesma 2018) and its
+precedent \texttt{sp} (E. Pebesma and Bivand 2005) for \ldots{} and
+\texttt{raster} (Hijmans 2020) and \texttt{terra} (Hijmans 2021) for
+raster data. While these implementations specialised in geographic
+manipulations with different type of simple features, it doesn't
+incorporate a temporal dimension in the data structure. Project like
+\texttt{stars} (E. Pebesma 2021) and \texttt{spacetime} (Bivand,
+Pebesma, and Gomez-Rubio 2013) by R-Spatial allows for both space and
+time dimension for raster and vector data, but the underlying data
+structure is a multi-dimensional array, which could be difficult to
+operate for R users who are more familiar with the operation in 2D
+dataframe/ tibble.
 
-\begin{itemize}
-\item
-  \texttt{sp} + \texttt{spacetime} (time series data from package
-  \texttt{xts})
-\item
-  \texttt{sf}
-
-  \begin{itemize}
-  \tightlist
-  \item
-    simple features geometry
-  \item
-    only spatio
-  \item
-    good for geographical but these geometry types i.e.~multipolygon,
-    can be hard to manipulate. Would be nice to have a vector data
-    structure
-  \end{itemize}
-\item
-  \texttt{stars}
-
-  \begin{itemize}
-  \tightlist
-  \item
-    a newer structure allow both spatio + temporal
-  \item
-    array
-  \end{itemize}
-\end{itemize}
+In the temporal aspect, the \texttt{tsibble} (Wang, Cook, and Hyndman
+2020) structure and its tidyverts ecosystem have provided a {[}\ldots{}
+{]} workflow to work with temporal data. In a tsibble structure,
+temporal data is characterised by \texttt{index} and \texttt{key} where
+\texttt{index} is the temporal identifier and \texttt{key} is the
+identifier for multiple series, which could be used as a spatio
+identifier. However, a tsibble object, by construction, always requires
+the \texttt{index} in its structure. This makes it less appealing for
+spatio-temporal data since the output of calculated spatio-specific
+variables (i.e.~features of each series) don't have the time dimension.
+Analysts will either need to have an additional step to join this output
+to the original tsibble or operate with variables stored in two separate
+objects. In addition, the long form structure of a tsibble object means
+spatio variables (i.e.~longitude, latitude, and features of each series
+if joined back to the tsibble) of each spatio identifier will be
+repetitively recorded at each timestamp. This repetition is unnecessary
+and would inflate the object size for long series.
 
@@
 \hypertarget{summary}{%
-\section{Summary}\label{summary}}
+\section*{Summary}\label{summary}}
+\addcontentsline{toc}{section}{Summary}
+
+\hypertarget{refs}{}
+\begin{CSLReferences}{1}{0}
+\leavevmode\hypertarget{ref-spacetimebook}{}%
+Bivand, Roger S., Edzer Pebesma, and Virgilio Gomez-Rubio. 2013.
+\emph{Applied Spatial Data Analysis with {R}, Second Edition}. Springer,
+NY. \url{https://asdar-book.org/}.
+
+\leavevmode\hypertarget{ref-raster}{}%
+Hijmans, Robert J. 2020. \emph{Raster: Geographic Data Analysis and
+Modeling}. \url{https://CRAN.R-project.org/package=raster}.
+
+\leavevmode\hypertarget{ref-terra}{}%
+---------. 2021. \emph{Terra: Spatial Data Analysis}.
+\url{https://CRAN.R-project.org/package=terra}.
+
+\leavevmode\hypertarget{ref-stars}{}%
+Pebesma, Edzer. 2021. \emph{Stars: Spatiotemporal Arrays, Raster and
+Vector Data Cubes}. \url{https://CRAN.R-project.org/package=stars}.
+
+\leavevmode\hypertarget{ref-pebesma2018simple}{}%
+Pebesma, Edzer J. 2018. {``Simple Features for r: Standardized Support
+for Spatial Vector Data.''} \emph{R J.} 10 (1): 439.
+
+\leavevmode\hypertarget{ref-pebesma2005s}{}%
+Pebesma, Edzer, and Roger S Bivand. 2005. {``S Classes and Methods for
+Spatial Data: The Sp Package.''} \emph{R News} 5 (2): 9--13.
+
+\leavevmode\hypertarget{ref-tsibbles}{}%
+Wang, Earo, Dianne Cook, and Rob J Hyndman. 2020. {``A New Tidy Data
+Structure to Support Exploration and Modeling of Temporal Data.''}
+\emph{Journal of Computational and Graphical Statistics} 29 (3):
+466--78. \url{https://doi.org/10.1080/10618600.2019.1695624}.
+
+\end{CSLReferences}
 

### 2021-08-08 main/main.Rmd
add description of each section, a new section on to introduce manipulation/ vis with cubble

@@
 
-Spatio-temporal data
+Saptio-temporal data
+
+<!-- Spatio-temporal data requires the data to have both a time aspect that is long enough to be considered as time series data and a space aspect that spreads across the a certain geographical region.
+In the time aspect, we may expect to see trend and seasonality across different granularities and in the spatio aspect, similarity of patterns can be discovered.
+-->
+
+<!-- can see information of both dimension,  -->
+<!-- not sufficient implementation to manipulate and visualise spatio-temporal data. more focus solely on the spatio/ temporal aspect.  -->
+<!-- The purpose of this paper is to introduce a spatio-temporal vector data structure for data analysis in R -->
 
@@
 
-Motivation: Many focus on spatio/ temporal solely -> spatio-temporal vector data structure for data analysis
-
-sections
+The rest of the paper will be divided as follows: Section 2 reviews the existing data structure for spatio, temporal, and spatio-temporal data.
+Section 3 presents a new data structure for spatio-temporal data: cubble.
+Then the paper introduces the workflow of data manipulation and visualisation with the cubble structure in Section 4.
+Section 5 gives some examples on how common spatial and temporal manipulations are performed with cubble and how static and interactive visualisation help ot understand climate and [...] data.
 
@@
 
+# Manipulation and visualisation with `cubble`
+
+Mention different types of manipulation with cubble:
+
+  - `dplyr` support for cubble: 
+    - basic 5s: mutate, filter, summarise, select, arrange
+    - group and ungorup: group_by, ungroup
+    - slice family
+  - summarise missing stats
+
 # Examples

### 2021-08-08 main/main.tex
add description of each section, a new section on to introduce manipulation/ vis with cubble

@@
 
-Spatio-temporal data
+Saptio-temporal data
 
-Motivation: Many focus on spatio/ temporal solely -\textgreater{}
-spatio-temporal vector data structure for data analysis
-
-sections
+The rest of the paper will be divided as follows: Section 2 reviews the
+existing data structure for spatio, temporal, and spatio-temporal data.
+Section 3 presents a new data structure for spatio-temporal data:
+cubble. Then the paper introduces the workflow of data manipulation and
+visualisation with the cubble structure in Section 4. Section 5 gives
+some examples on how common spatial and temporal manipulations are
+performed with cubble and how static and interactive visualisation help
+ot understand climate and {[}\ldots{]} data.
 
@@
 Compatible with tidyverse manipulation and tsibble
+
+\hypertarget{manipulation-and-visualisation-with-cubble}{%
+\section{\texorpdfstring{Manipulation and visualisation with
+\texttt{cubble}}{Manipulation and visualisation with cubble}}\label{manipulation-and-visualisation-with-cubble}}
+
+Mention different types of manipulation with cubble:
+
+\begin{itemize}
+\tightlist
+\item
+  \texttt{dplyr} support for cubble:
+
+  \begin{itemize}
+  \tightlist
+  \item
+    basic 5s: mutate, filter, summarise, select, arrange
+  \item
+    group and ungorup: group\_by, ungroup
+  \item
+    slice family
+  \end{itemize}
+\item
+  summarise missing stats
+\end{itemize}
 

### 2021-08-09 main/main.Rmd
a draft of introduction

--- 2021-08-04 main/main.Rmd
@@
 
-Spatio-temporal data
+Spatio-temporal data record changes of variables [for a reasonable length of period and spread across geographical region].
+In this article, we consider spatio-temporal vector data, which are recorded in a fixed interval and are point based, characterised by longitude and latitude, in the spatial aspect.
+Examples of this type of data include the house price of a city or county, climate measures from weather stations in a country, and river level data from electronic gauges.
 
-<!-- While satellite imagery can record information on a continuous spatial domain, other spatial information are recorded at a fixed set of locations characterised by latitude, longitude, and other spatial metadata.
-Examples of this type of data includes climate measures from weather observation stations and river level data recorded by electronic gauges.
-Analysing this data requires less considerations on the geometry type and map projection while more on how measures in these fixed locations changes across the time domain and whether these changes are related for adjacent locations.
--->
+Analysing this type of data requires less considerations on the geographical geometry type and map projection but more on how measures in these fixed locations changes across the time domain and whether these changes are related for adjacent locations.
+For example, when nearby areas show patterns that are regular enough, visualising spatio-temporal data can 1) discover regional time series features, i.e.
+trend and seasonality, 2) find the Waldo sites from the crowd, and 3) see how correlation of nearby sites changes across time.
 
-Motivation: Many focus on spatio/ temporal solely -> spatio-temporal vector data structure for data analysis
+The main difficulty in visualising this type of data is to show information in both space and time dimension with the proper level of details without information overflow.
+This would sometimes require aggregating the time dimension into the proper level or slicing the data into a reasonable number of subset for display.
+In this sense, a data structure that regulates the manipulation spatio-temporal data will benefit the analysis workflow.
+While many implementations focus on manipulating and visualising pure spatial or temporal data, there are not sufficient tools to deal with spatio-temporal data.
+The purpose of this paper is to introduce a spatio-temporal vector data structure for data analysis in R.
 
-sections
+The rest of the paper will be divided as follows: Section 2 reviews the existing data structure for spatio, temporal, and spatio-temporal data.
+Section 3 presents a new data structure for spatio-temporal data: cubble.
+Then the paper introduces the workflow of data manipulation and visualisation with the cubble structure in Section 4.
+Section 5 gives some examples on how common spatial and temporal manipulations are performed with cubble and how static and interactive visualisation help to understand climate and [...] data.
 
@@
 
+# Manipulation and visualisation with `cubble`
+
+Mention different types of manipulation with cubble:
+
+  - `dplyr` support for cubble: 
+    - basic 5s: mutate, filter, summarise, select, arrange
+    - group and ungroup: group_by, ungroup
+    - slice family
+  - summarise missing stats
+
 # Examples

### 2021-08-09 main/main.tex
a draft of introduction

@@
 
-Saptio-temporal data
+Spatio-temporal data record changes of variables {[}for a reasonable
+length of period and spread across geographical region{]}. In this
+article, we consider spatio-temporal vector data, which are recorded in
+a fixed interval and are point based, characterised by longitude and
+latitude, in the spatial aspect. Examples of this type of data include
+the house price of a city or county, climate measures from weather
+stations in a country, and river level data from electronic gauges.
+
+Analysing this type of data requires less considerations on the
+geographical geometry type and map projection but more on how measures
+in these fixed locations changes across the time domain and whether
+these changes are related for adjacent locations. For example, when
+nearby areas show patterns that are regular enough, visualising
+spatio-temporal data can 1) discover regional time series features,
+i.e.~trend and seasonality, 2) find the Waldo sites from the crowd, and
+3) see how correlation of nearby sites changes across time.
+
+The main difficulty in visualising this type of data is to show
+information in both space and time dimension with the proper level of
+details without information overflow. This would sometimes require
+aggregating the time dimension into the proper level or slicing the data
+into a reasonable number of subset for display. In this sense, a data
+structure that regulates the manipulation spatio-temporal data will
+benefit the analysis workflow. While many implementations focus on
+manipulating and visualising pure spatial or temporal data, there are
+not sufficient tools to deal with spatio-temporal data. The purpose of
+this paper is to introduce a spatio-temporal vector data structure for
+data analysis in R.
 
@@
 performed with cubble and how static and interactive visualisation help
-ot understand climate and {[}\ldots{]} data.
+to understand climate and {[}\ldots{]} data.
 
@@
   \item
-    group and ungorup: group\_by, ungroup
+    group and ungroup: group\_by, ungroup
   \item

### 2021-08-10 main/main.Rmd
cubble structure para

@@
 ---
+
+```{r setup, echo = FALSE}
+knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
+                      fig.path = "figures/")
+library(cubble)
+```
 
@@
 
-Intro to cubble:
+In spatio-temporal data, a variable is usually either group-related or time-related.
+While time-related variables record variables that change in time, group-related variables are invariant to the time and have only one value per group.
+A cubble simplifies the workflow in spatio-temporal data through managing group and time-related variables separately in two forms: list-column and long form.
+The list column form:
 
-- list-column: rowwise_df with temporal variables, including the time index, nested. 
-  - Focus on spatio: those output per station
+  - defines each group in a row, 
+  - displays the group-related variables in columns, and 
+  - nests all the time-related variables into a column called `ts`.
 
-- long form: grouped_df 
-  - Focus on temporal
+In the long form,
 
-Compatible with tidyverse manipulation and tsibble
+  - each combination of group and timestamp occupies a row
+  - time-related variables are displayed, and 
+  - group-related variables are not explicitly displayed but can be accessed through the `meta` attribute
+
+Below are the how the list-column and long form look like for Australia climate data, which records daily precipitation, maximum and minimum temperature for 55 stations across Australia from 2015- 2020.
+
+\newpage
+
+```{r echo = FALSE}
+climate_small
+```
+
+```{r echo = FALSE}
+climate_small %>% zoom()
+```
 
 # Manipulation and visualisation with `cubble`
+
+## Cubble verbs
 

### 2021-08-10 main/main.tex
cubble structure para

@@
 
-Intro to cubble:
+In spatio-temporal data, a variable is usually either group-related or
+time-related. While time-related variables record variables that change
+in time, group-related variables are invariant to the time and have only
+one value per group. A cubble simplifies the workflow in spatio-temporal
+data through managing group and time-related variables separately in two
+forms: list-column and long form. The list column form:
 
@@
 \item
-  list-column: rowwise\_df with temporal variables, including the time
-  index, nested.
+  defines each group in a row,
+\item
+  displays the group-related variables in columns, and
+\item
+  nests all the time-related variables into a column called \texttt{ts}.
+\end{itemize}
+
+In the long form,
 
@@
 \item
-    Focus on spatio: those output per station
-  \end{itemize}
+  each combination of group and timestamp occupies a row
 \item
-  long form: grouped\_df
-
-  \begin{itemize}
-  \tightlist
+  time-related variables are displayed, and
 \item
-    Focus on temporal
-  \end{itemize}
+  group-related variables are not explicitly displayed but can be
+  accessed through the \texttt{meta} attribute
 \end{itemize}
 
-Compatible with tidyverse manipulation and tsibble
+Below are the how the list-column and long form look like for Australia
+climate data, which records daily precipitation, maximum and minimum
+temperature for 55 stations across Australia from 2015- 2020.
+
+\newpage
+
+\begin{verbatim}
+## # Cubble: station-wise: list-column
+## # Group:  station [55]
+## # Meta:   station [fct], lat [dbl], long [dbl], elevation [dbl], name [fct]
+##    station       lat  long elevation name                ts                  
+##    <fct>       <dbl> <dbl>     <dbl> <fct>               <list>              
+##  1 ASN00001019 -14.3  127.      23   kalumburu           <tbl_ts [2,147 x 4]>
+##  2 ASN00002012 -18.2  128.     422   halls creek airport <tbl_ts [2,191 x 4]>
+##  3 ASN00003003 -17.9  122.       7.4 broome airport      <tbl_ts [2,192 x 4]>
+##  4 ASN00006011 -24.9  114.       4   carnarvon airport   <tbl_ts [2,192 x 4]>
+##  5 ASN00009021 -31.9  116.      15.4 perth airport       <tbl_ts [2,192 x 4]>
+##  6 ASN00009193 -32.0  116.      43.1 rottnest island     <tbl_ts [2,192 x 4]>
+##  7 ASN00009518 -34.4  115.      13   cape leeuwin        <tbl_ts [2,184 x 4]>
+##  8 ASN00009789 -33.8  122.      25   esperance           <tbl_ts [2,187 x 4]>
+##  9 ASN00010286 -31.6  117.     217.  cunderdin airfield  <tbl_ts [2,190 x 4]>
+## 10 ASN00010917 -32.7  117.     275   wandering           <tbl_ts [2,192 x 4]>
+## # ... with 45 more rows
+\end{verbatim}
+
+\begin{verbatim}
+## # Cubble: time-wise: long form [tsibble]
+## # Group:  station [55]
+## # Meta:   station [fct], lat [dbl], long [dbl], elevation [dbl], name [fct]
+##    station     date        prcp  tmax  tmin
+##    <fct>       <date>     <dbl> <dbl> <dbl>
+##  1 ASN00001019 2015-01-01   164  31.5  25  
+##  2 ASN00001019 2015-01-02   124  31.3  25  
+##  3 ASN00001019 2015-01-03    50  29.2  24.7
+##  4 ASN00001019 2015-01-04   204  32.1  24.3
+##  5 ASN00001019 2015-01-05   412  31.4  24.6
+##  6 ASN00001019 2015-01-06   200  28.7  25.1
+##  7 ASN00001019 2015-01-07   822  30.2  24.5
+##  8 ASN00001019 2015-01-08    22  31    26.6
+##  9 ASN00001019 2015-01-09    62  32.7  26.2
+## 10 ASN00001019 2015-01-10   274  32.9  22.6
+## # ... with 120,311 more rows
+\end{verbatim}
 
@@
 \texttt{cubble}}{Manipulation and visualisation with cubble}}\label{manipulation-and-visualisation-with-cubble}}
+
+\hypertarget{cubble-verbs}{%
+\subsection{Cubble verbs}\label{cubble-verbs}}
 

### 2021-08-10 main/main.Rmd
para on review sf + stars

@@
 
-There has been a large class of implementations dedicated to processing the spatial data.
-This includes `sf` [@pebesma2018simple] and its precedent `sp` [@pebesma2005s] for ...
-and `raster` [@raster] and `terra` [@terra] for raster data.
-While these implementations specialised in geographic manipulations with different type of simple features, it doesn't incorporate a temporal dimension in the data structure.
-Project like `stars` [@stars] and `spacetime` [@spacetimebook] by R-Spatial allows for both space and time dimension for raster and vector data, but the underlying data structure is a multi-dimensional array, which could be difficult to operate for R users who are more familiar with the operation in 2D dataframe/ tibble.
+Below we review some structure for spatial, temporal, and spatio-temporal data.
 
-<!-- - `sp` + `spacetime` (time series data from package `xts`) -->
-
-<!-- - `sf`  -->
-<!--   - simple features geometry  -->
-<!--   - only spatio -->
-<!--   - good for geographical but these geometry types i.e. multipolygon, can be hard to manipulate. Would be nice to have a vector data structure -->
-
-<!-- - `stars` -->
-<!--   - a newer structure allow both spatio + temporal  -->
-<!--   - array -->
+Many spatial and spatio-temporal data structures have been developed by the R-spatial team for both raster and vector spatial data.
+For vector spatial data, which is the focus of this paper, `sf` [@pebesma2018simple] represents spatial vector information with simple features: points, lines, polygons and their multiples.
+Various `st_` function are designed to manipulate these features based on their geometric relationships.
+For spatio-temporal data,  `stars` [@stars] can represent both raster and vector data using multi-dimensional array.
+However, the underlying array structure can be difficult to operate for data analysts who are more familiar with a flat 2D data frame structure used by the tidyverse ecosystem.
 

### 2021-08-10 main/main.tex
para on review sf + stars

@@
 
-There has been a large class of implementations dedicated to processing
-the spatial data. This includes \texttt{sf} (E. J. Pebesma 2018) and its
-precedent \texttt{sp} (E. Pebesma and Bivand 2005) for \ldots{} and
-\texttt{raster} (Hijmans 2020) and \texttt{terra} (Hijmans 2021) for
-raster data. While these implementations specialised in geographic
-manipulations with different type of simple features, it doesn't
-incorporate a temporal dimension in the data structure. Project like
-\texttt{stars} (E. Pebesma 2021) and \texttt{spacetime} (Bivand,
-Pebesma, and Gomez-Rubio 2013) by R-Spatial allows for both space and
-time dimension for raster and vector data, but the underlying data
-structure is a multi-dimensional array, which could be difficult to
-operate for R users who are more familiar with the operation in 2D
-dataframe/ tibble.
+Below we review some structure for spatial, temporal, and
+spatio-temporal data.
+
+Many spatial and spatio-temporal data structures have been developed by
+the R-spatial team for both raster and vector spatial data. For vector
+spatial data, which is the focus of this paper, \texttt{sf} (E. J.
+Pebesma 2018) represents spatial vector information with simple
+features: points, lines, polygons and their multiples. Various
+\texttt{st\_} function are designed to manipulate these features based
+on their geometric relationships. For spatio-temporal data,
+\texttt{stars} (E. Pebesma 2021) can represent both raster and vector
+data using multi-dimensional array. However, the underlying array
+structure can be difficult to operate for data analysts who are more
+familiar with a flat 2D data frame structure used by the tidyverse
+ecosystem.
 
@@
 \begin{CSLReferences}{1}{0}
-\leavevmode\hypertarget{ref-spacetimebook}{}%
-Bivand, Roger S., Edzer Pebesma, and Virgilio Gomez-Rubio. 2013.
-\emph{Applied Spatial Data Analysis with {R}, Second Edition}. Springer,
-NY. \url{https://asdar-book.org/}.
-
-\leavevmode\hypertarget{ref-raster}{}%
-Hijmans, Robert J. 2020. \emph{Raster: Geographic Data Analysis and
-Modeling}. \url{https://CRAN.R-project.org/package=raster}.
-
-\leavevmode\hypertarget{ref-terra}{}%
----------. 2021. \emph{Terra: Spatial Data Analysis}.
-\url{https://CRAN.R-project.org/package=terra}.
-
 \leavevmode\hypertarget{ref-stars}{}%
@@
 for Spatial Vector Data.''} \emph{R J.} 10 (1): 439.
-
-\leavevmode\hypertarget{ref-pebesma2005s}{}%
-Pebesma, Edzer, and Roger S Bivand. 2005. {``S Classes and Methods for
-Spatial Data: The Sp Package.''} \emph{R News} 5 (2): 9--13.
 

### 2021-08-18 main/main.Rmd
more content in framework and examples

@@
 knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, 
-                      fig.path = "figures/")
+                      fig.path = "figures/",
+                      out.width = "100%", 
+                      out.height = "30%")
+```
+
+```{r echo = FALSE}
 library(cubble)
+library(dplyr)
+library(lubridate)
 ```
@@
 
-Spatio-temporal data record changes of variables [for a reasonable length of period and spread across geographical region].
+Spatio-temporal data record changes of variables in spatially separated regions across time.
 In this article, we consider spatio-temporal vector data, which are recorded in a fixed interval and are point based, characterised by longitude and latitude, in the spatial aspect.
@@
 
-In spatio-temporal data, a variable is usually either group-related or time-related.
-While time-related variables record variables that change in time, group-related variables are invariant to the time and have only one value per group.
-A cubble simplifies the workflow in spatio-temporal data through managing group and time-related variables separately in two forms: list-column and long form.
-The list column form:
+Spatio-temporal data can be thought of as characterised by three dimensions: group, time, and variable, which allows us to use a cube to represent the data.
+While group and time are can be seen as identifiers of the record, variables can be further divided into those that are group-related and time-related.
+Group-related variables are invariant to the time and have only one value per group while time-related variables change in time and requires both group and time to identify a record.
+Hence the cubic representation can be further simplified for those group-related variables as in the sketch \ref{fig:framework}, where group-related variables are squashed in the time dimension.
+
+```{r framework, echo = FALSE, fig.cap = "Cubic representation of spatio-temporal data.
+(1) shows the full cube, (2) simplifies teh cube by squashing the time dimension of time-invariant variables.
+With (2), a slice along the group dimension will affect both time-variant and invariant variables as in (3) while, a slice along the time will only affect the tim-variant variables as in (4)."}
+knitr::include_graphics(here::here("figures/cubble-framework-hand.jpeg"))
+```
+
+A cubble simplifies the workflow in spatio-temporal data through managing group and time-related variables separately in two forms: list-column and long form. The nested form:
 
@@
   - nests all the time-related variables into a column called `ts`.
+
+This form focuses on the time-invariant variable by squashing the time dimension
 
@@
 
-Below are the how the list-column and long form look like for Australia climate data, which records daily precipitation, maximum and minimum temperature for 55 stations across Australia from 2015- 2020.
-
-\newpage
+Below are the how the nested and long form look like for Australia climate data, which records daily precipitation, maximum and minimum temperature for 55 stations across Australia from 2015- 2020.
+Notice that each station forms a group in both forms and specifically, the nested  and long form have a underlying `rowwise_df` and `grouped_df` respectively.
 
@@
 ```{r echo = FALSE}
-climate_small %>% zoom()
+climate_small %>% stretch()
 ```
 
-# Manipulation and visualisation with `cubble`
+With a cubic framework on mind, different types of manipulation with cubble can be thought of as slicing the cube in various way.
+The table below shows how some `dplyr` verbs are mapped into the operation in a cubble.
+With the existing grouping on the station, additional groupping can be added with `group_by` and removed with `ungrouped`.
+[talk about why it is useful]
+
+```{r verbs, echo = FALSE, out.height = "50%", out.width = "100%", fig.cap = "sdfsdf"}
+knitr::include_graphics(here::here("figures/cubble-verbs.png"))
+```
 
@@
 
+\newpage
+
 # Examples
@@
 ## Manipulation
+
+### Mutate and filter
+
+In the first example, we want to only keep the stations that have `tmax` recorded in 2020.
+This requires first narrow down the records to those in 2020, determine if `tmax` is missing for each station, and then retain those stations that have `tmax` recorded.
+The year filtering is an operation on the time axis, so we start with the long form.
+Whether each station has `tmax` recorded is a result of each station, rather than of each time point, hence we need to switch to the nested form with `tamp()`.
+To calculate whether `tmax` is recorded, we mutate a column `tmax_missing` that takes `TRUE` if all the `tmax` in the nested list column `ts` are `NA` and `FALSE` otherwise.
+To get the stations that we want, we need another filter on `tmax_missing`.
+
+```{r}
+climate_small %>% 
+  stretch() %>% 
+  filter(year(date) == 2020) %>% 
+  tamp() %>% 
+  mutate(tmax_missing = ifelse(all(is.na(ts$tmax)), TRUE, FALSE)) %>% 
+  filter(!tmax_missing)
+```
+
+### Join
+
+Now we want to select the stations that have been registered with world meteorological organisation (WMO) and the dataset `station` has a column `wmo_id` that records this information.
+To do this task, we first need to join the `station` dataset with our climate dataset and then filter out those stations that don't have the WMO id.
+Since the join is by station rather than by time, we start with the nested form and write the exact same syntax of join and filter as with tidyverse.
+
+```{r}
+# join wmo_id for each station
+to_join <- station %>% select(id, wmo_id)
+out <- climate_small %>% 
+  left_join(to_join, by = c("station" = "id")) %>% 
+  filter(!is.na(wmo_id))
+out
+```
+
+Sometimes, we would like to have station-wise and time-wise variables in the same form (i.e.
+when plotting glyph maps).
+This can also be seen as a joining task, on the long form, with the dataset to join being the metadata.
+`migrate()` is a verb introduced as the shortcut for `left_join()` with a cubble's metadata and below is the comparison of the two syntaxes.
+
+```{r}
+out %>% 
+  stretch() %>% 
+  migrate(station, lat, long)
+```
 

### 2021-08-18 main/main.tex
more content in framework and examples

@@
 
+\usepackage{color}
+\usepackage{fancyvrb}
+\newcommand{\VerbBar}{|}
+\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
+\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
+% Add ',fontsize=\small' for more characters per line
+\usepackage{framed}
+\definecolor{shadecolor}{RGB}{248,248,248}
+\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
+\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
+\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
+\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\BuiltInTok}[1]{#1}
+\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
+\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
+\newcommand{\ExtensionTok}[1]{#1}
+\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ImportTok}[1]{#1}
+\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\NormalTok}[1]{#1}
+\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
+\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
+\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\RegionMarkerTok}[1]{#1}
+\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+
 % Pandoc citation processing
@@
 
-Spatio-temporal data record changes of variables {[}for a reasonable
-length of period and spread across geographical region{]}. In this
-article, we consider spatio-temporal vector data, which are recorded in
-a fixed interval and are point based, characterised by longitude and
-latitude, in the spatial aspect. Examples of this type of data include
-the house price of a city or county, climate measures from weather
-stations in a country, and river level data from electronic gauges.
+Spatio-temporal data record changes of variables in spatially separated
+regions across time. In this article, we consider spatio-temporal vector
+data, which are recorded in a fixed interval and are point based,
+characterised by longitude and latitude, in the spatial aspect. Examples
+of this type of data include the house price of a city or county,
+climate measures from weather stations in a country, and river level
+data from electronic gauges.
 
@@
 
-In spatio-temporal data, a variable is usually either group-related or
-time-related. While time-related variables record variables that change
-in time, group-related variables are invariant to the time and have only
-one value per group. A cubble simplifies the workflow in spatio-temporal
-data through managing group and time-related variables separately in two
-forms: list-column and long form. The list column form:
+Spatio-temporal data can be thought of as characterised by three
+dimensions: group, time, and variable, which allows us to use a cube to
+represent the data. While group and time are can be seen as identifiers
+of the record, variables can be further divided into those that are
+group-related and time-related. Group-related variables are invariant to
+the time and have only one value per group while time-related variables
+change in time and requires both group and time to identify a record.
+Hence the cubic representation can be further simplified for those
+group-related variables as in the sketch \ref{fig:framework}, where
+group-related variables are squashed in the time dimension.
+
+\begin{figure}
+\includegraphics[width=1\linewidth,height=0.3\textheight]{/Users/hzha400/Documents/PhD/research/paper-cubble/figures/cubble-framework-hand} \caption{Cubic representation of spatio-temporal data.
+(1) shows the full cube, (2) simplifies teh cube by squashing the time dimension of time-invariant variables.
+With (2), a slice along the group dimension will affect both time-variant and invariant variables as in (3) while, a slice along the time will only affect the tim-variant variables as in (4).}\label{fig:framework}
+\end{figure}
+
+A cubble simplifies the workflow in spatio-temporal data through
+managing group and time-related variables separately in two forms:
+list-column and long form. The nested form:
 
@@
 \end{itemize}
+
+This form focuses on the time-invariant variable by squashing the time
+dimension
 
@@
 
-Below are the how the list-column and long form look like for Australia
+Below are the how the nested and long form look like for Australia
 climate data, which records daily precipitation, maximum and minimum
-temperature for 55 stations across Australia from 2015- 2020.
-
-\newpage
+temperature for 55 stations across Australia from 2015- 2020. Notice
+that each station forms a group in both forms and specifically, the
+nested and long form have a underlying \texttt{rowwise\_df} and
+\texttt{grouped\_df} respectively.
 
 \begin{verbatim}
-## # Cubble: station-wise: list-column
+## # Cubble: station-wise: nested form
 ## # Group:  station [55]
-## # Meta:   station [fct], lat [dbl], long [dbl], elevation [dbl], name [fct]
 ##    station       lat  long elevation name                ts                  
@@
 ## # Group:  station [55]
-## # Meta:   station [fct], lat [dbl], long [dbl], elevation [dbl], name [fct]
 ##    station     date        prcp  tmax  tmin
@@
 
-\hypertarget{manipulation-and-visualisation-with-cubble}{%
-\section{\texorpdfstring{Manipulation and visualisation with
-\texttt{cubble}}{Manipulation and visualisation with cubble}}\label{manipulation-and-visualisation-with-cubble}}
+With a cubic framework on mind, different types of manipulation with
+cubble can be thought of as slicing the cube in various way. The table
+below shows how some \texttt{dplyr} verbs are mapped into the operation
+in a cubble. With the existing grouping on the station, additional
+groupping can be added with \texttt{group\_by} and removed with
+\texttt{ungrouped}. {[}talk about why it is useful{]}
+
+\begin{figure}
+\includegraphics[width=1\linewidth,height=0.5\textheight]{/Users/hzha400/Documents/PhD/research/paper-cubble/figures/cubble-verbs} \caption{sdfsdf}\label{fig:verbs}
+\end{figure}
 
@@
 
+\newpage
+
 \hypertarget{examples}{%
@@
 \subsection{Manipulation}\label{manipulation}}
+
+\hypertarget{mutate-and-filter}{%
+\subsubsection{Mutate and filter}\label{mutate-and-filter}}
+
+In the first example, we want to only keep the stations that have
+\texttt{tmax} recorded in 2020. This requires first narrow down the
+records to those in 2020, determine if \texttt{tmax} is missing for each
+station, and then retain those stations that have \texttt{tmax}
+recorded. The year filtering is an operation on the time axis, so we
+start with the long form. Whether each station has \texttt{tmax}
+recorded is a result of each station, rather than of each time point,
+hence we need to switch to the nested form with \texttt{tamp()}. To
+calculate whether \texttt{tmax} is recorded, we mutate a column
+\texttt{tmax\_missing} that takes \texttt{TRUE} if all the \texttt{tmax}
+in the nested list column \texttt{ts} are \texttt{NA} and \texttt{FALSE}
+otherwise. To get the stations that we want, we need another filter on
+\texttt{tmax\_missing}.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{climate\_small }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{stretch}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{filter}\NormalTok{(}\FunctionTok{year}\NormalTok{(date) }\SpecialCharTok{==} \DecValTok{2020}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{tamp}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{tmax\_missing =} \FunctionTok{ifelse}\NormalTok{(}\FunctionTok{all}\NormalTok{(}\FunctionTok{is.na}\NormalTok{(ts}\SpecialCharTok{$}\NormalTok{tmax)), }\ConstantTok{TRUE}\NormalTok{, }\ConstantTok{FALSE}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{filter}\NormalTok{(}\SpecialCharTok{!}\NormalTok{tmax\_missing)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## # Cubble: station-wise: nested form
+## # Group:  station [55]
+##    station      lat  long elevation name             ts             tmax_missing
+##    <fct>      <dbl> <dbl>     <dbl> <fct>            <list>         <lgl>       
+##  1 ASN000010~ -14.3  127.      23   kalumburu        <tbl_ts [366 ~ FALSE       
+##  2 ASN000020~ -18.2  128.     422   halls creek air~ <tbl_ts [366 ~ FALSE       
+##  3 ASN000030~ -17.9  122.       7.4 broome airport   <tbl_ts [366 ~ FALSE       
+##  4 ASN000060~ -24.9  114.       4   carnarvon airpo~ <tbl_ts [366 ~ FALSE       
+##  5 ASN000090~ -31.9  116.      15.4 perth airport    <tbl_ts [366 ~ FALSE       
+##  6 ASN000091~ -32.0  116.      43.1 rottnest island  <tbl_ts [366 ~ FALSE       
+##  7 ASN000095~ -34.4  115.      13   cape leeuwin     <tbl_ts [366 ~ FALSE       
+##  8 ASN000097~ -33.8  122.      25   esperance        <tbl_ts [361 ~ FALSE       
+##  9 ASN000102~ -31.6  117.     217.  cunderdin airfi~ <tbl_ts [366 ~ FALSE       
+## 10 ASN000109~ -32.7  117.     275   wandering        <tbl_ts [366 ~ FALSE       
+## # ... with 45 more rows
+\end{verbatim}
+
+\hypertarget{join}{%
+\subsubsection{Join}\label{join}}
+
+Now we want to select the stations that have been registered with world
+meteorological organisation (WMO) and the dataset \texttt{station} has a
+column \texttt{wmo\_id} that records this information. To do this task,
+we first need to join the \texttt{station} dataset with our climate
+dataset and then filter out those stations that don't have the WMO id.
+Since the join is by station rather than by time, we start with the
+nested form and write the exact same syntax of join and filter as with
+tidyverse.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# join wmo\_id for each station}
+\NormalTok{to\_join }\OtherTok{\textless{}{-}}\NormalTok{ station }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{select}\NormalTok{(id, wmo\_id)}
+\NormalTok{out }\OtherTok{\textless{}{-}}\NormalTok{ climate\_small }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{left\_join}\NormalTok{(to\_join, }\AttributeTok{by =} \FunctionTok{c}\NormalTok{(}\StringTok{"station"} \OtherTok{=} \StringTok{"id"}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{filter}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(wmo\_id))}
+\NormalTok{out}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## # Cubble: station-wise: nested form
+## # Group:  station [51]
+##    station       lat  long elevation name              ts                 wmo_id
+##    <fct>       <dbl> <dbl>     <dbl> <fct>             <list>              <dbl>
+##  1 ASN00001019 -14.3  127.      23   kalumburu         <tbl_ts [2,147 x ~  94100
+##  2 ASN00002012 -18.2  128.     422   halls creek airp~ <tbl_ts [2,191 x ~  94212
+##  3 ASN00003003 -17.9  122.       7.4 broome airport    <tbl_ts [2,192 x ~  94203
+##  4 ASN00006011 -24.9  114.       4   carnarvon airport <tbl_ts [2,192 x ~  94300
+##  5 ASN00009021 -31.9  116.      15.4 perth airport     <tbl_ts [2,192 x ~  94610
+##  6 ASN00009193 -32.0  116.      43.1 rottnest island   <tbl_ts [2,192 x ~  94602
+##  7 ASN00009518 -34.4  115.      13   cape leeuwin      <tbl_ts [2,184 x ~  94601
+##  8 ASN00009789 -33.8  122.      25   esperance         <tbl_ts [2,187 x ~  94638
+##  9 ASN00010286 -31.6  117.     217.  cunderdin airfie~ <tbl_ts [2,190 x ~  95625
+## 10 ASN00010917 -32.7  117.     275   wandering         <tbl_ts [2,192 x ~  95640
+## # ... with 41 more rows
+\end{verbatim}
+
+Sometimes, we would like to have station-wise and time-wise variables in
+the same form (i.e.~when plotting glyph maps). This can also be seen as
+a joining task, on the long form, with the dataset to join being the
+metadata. \texttt{migrate()} is a verb introduced as the shortcut for
+\texttt{left\_join()} with a cubble's metadata and below is the
+comparison of the two syntaxes.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{out }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{stretch}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{migrate}\NormalTok{(station, lat, long)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## # Cubble: time-wise: long form
+## # Group:  station [55]
+##    station     date        prcp  tmax  tmin   lat  long
+##    <fct>       <date>     <dbl> <dbl> <dbl> <dbl> <dbl>
+##  1 ASN00001019 2015-01-01   164  31.5  25   -14.3  127.
+##  2 ASN00001019 2015-01-02   124  31.3  25   -14.3  127.
+##  3 ASN00001019 2015-01-03    50  29.2  24.7 -14.3  127.
+##  4 ASN00001019 2015-01-04   204  32.1  24.3 -14.3  127.
+##  5 ASN00001019 2015-01-05   412  31.4  24.6 -14.3  127.
+##  6 ASN00001019 2015-01-06   200  28.7  25.1 -14.3  127.
+##  7 ASN00001019 2015-01-07   822  30.2  24.5 -14.3  127.
+##  8 ASN00001019 2015-01-08    22  31    26.6 -14.3  127.
+##  9 ASN00001019 2015-01-09    62  32.7  26.2 -14.3  127.
+## 10 ASN00001019 2015-01-10   274  32.9  22.6 -14.3  127.
+## # ... with 111,562 more rows
+\end{verbatim}
 

### 2021-10-12 main.Rmd
more data structure section, new create a cubble section, files to top level

--- 2021-08-18 main/main.Rmd
@@
 
+\newpage
+
 # Introduction
@@
 
+\newpage
+
 # A new data structure for spatio-temporal data
 
-Spatio-temporal data can be thought of as characterised by three dimensions: group, time, and variable, which allows us to use a cube to represent the data.
-While group and time are can be seen as identifiers of the record, variables can be further divided into those that are group-related and time-related.
-Group-related variables are invariant to the time and have only one value per group while time-related variables change in time and requires both group and time to identify a record.
-Hence the cubic representation can be further simplified for those group-related variables as in the sketch \ref{fig:framework}, where group-related variables are squashed in the time dimension.
+Spatio-temporal data don't usually come to the analysts as a whole piece.
+A way to look at these data is to divide it into spatial and temporal dimension with an ID that links between the two.
+The first row in Figure \ref{fig:cubble-diagram} illustrates this representation where in the spatial dimension, the data is characterised by `id`, `lat`, `long`.
+$V_s$ in the last column represents all the other site-wise variables, for example, elevation and full name etc.
+The temporal dimension, on the other hand, can be characterised by `id` and `t` with $V_t$ representing all the time-wise variables.
+In climate data, this could include precipitation, maximum or minimum temperature, and wind speed etc.
 
-```{r framework, echo = FALSE, fig.cap = "Cubic representation of spatio-temporal data.
-(1) shows the full cube, (2) simplifies teh cube by squashing the time dimension of time-invariant variables.
-With (2), a slice along the group dimension will affect both time-variant and invariant variables as in (3) while, a slice along the time will only affect the tim-variant variables as in (4)."}
-knitr::include_graphics(here::here("figures/cubble-framework-hand.jpeg"))
+```{r cubble-diagram, echo = FALSE, fig.cap="Cubble diagram", out.height="40%", fig.align='center'}
+knitr::include_graphics(here::here("figures/cubble-diagram/cubble-diagram.001.png"))
 ```
 
-A cubble simplifies the workflow in spatio-temporal data through managing group and time-related variables separately in two forms: list-column and long form. The nested form:
+To work with spatio-temporal data, analysts can choose to either work separately on each dimension or join the two sets together, however, each approach has its own problem: While is is natural to work separately on each sheet (since spatial and temporal operations usually don't overlap), analysts will need to manually keep the other data frame up to date.
+For example, the following pseudo code illustrates the scenario where once the spatial dataset is filtered for those within Victoria, the temporal dataset needs to be manually updated to reflect this spatial filter.
+
+```{r eval = FALSE}
+spatial_new <- spatial %>% filter(SITES_IN_VICTORIA)
+temporal_new <- temporal %>% filter(id %in% spatial_new$id)
+```
+
+If analysts choose to join the spatial and temporal data together, the joined dataset could be too large since each spatial variable will be repeated at each time stamp for each site.
+Also, recordings of the site ID from different data sources can be slightly different from each other, causing a painful checking and cleaning of site IDs before the join.
+
+A cubble, in essence, wires both dimensions in the spatio-temporal data into one object while provide two forms for manipulation the spatial and temporal dimension separately.
+
+When manipulating the spatial dimension it uses a nest form that:
 
@@
 
-This form focuses on the time-invariant variable by squashing the time dimension
-
-In the long form,
+When manipulating the temporal dimension, it uses the long form that:
 
@@
   - time-related variables are displayed, and 
-  - group-related variables are not explicitly displayed but can be accessed through the `meta` attribute
+  - group-related variables are not explicitly displayed but can be accessed through the `meta` attribute.
 
-Below are the how the nested and long form look like for Australia climate data, which records daily precipitation, maximum and minimum temperature for 55 stations across Australia from 2015- 2020.
-Notice that each station forms a group in both forms and specifically, the nested  and long form have a underlying `rowwise_df` and `grouped_df` respectively.
+\newpage
+
+# Create a cubble
+
+The creation of a cubble requires the site identifier (`key`), as well as the spatial (`coords`) and temporal (`index`) identifier.
+`climate_flat` is already a tibble and it uses `id` to identify each station, `date` as the time identifier, and `c(long, lat)` as the spatial identifier.
+To create a cubble for this data, use:
+
+```{r}
+climate_flat %>% as_cubble(key = id, index = date, coords = c(long, lat))
+```
+
+Most of the time, spatio-temporal data doesn't come into this form and analysts need to query the climate variables based on station metadata.
+For this type of task, one can structure a metadata into a tibble and use row-wise operator to query the climate variables into a nested list.
+As an example here we demonstrate the workflow to find the 5 closest stations to Melbourne.
+We first create a station data farme with the 5 target stations.
 
 ```{r echo = FALSE}
-climate_small
+latlon_ref <- tibble::tibble(id = "melbourne", lat = -37.8136, long = 144.9631)
+
+cand <- aus_climate %>% 
+  select(id, lat, long, elev, name, wmo_id) %>% 
+   as_tibble()
+
+calc_dist <- function(df_ref, df, n = 5){
+  long_ref <- df_ref$long
+  lat_ref <- df_ref$lat
+
+  df %>% 
+    mutate(long_ref = long_ref,
+           lat_ref=  lat_ref,
+           dist = rnoaa::meteo_spherical_distance(lat, long,lat_ref, long_ref)) %>% 
+    slice_min(dist, n = n) %>% 
+    mutate(city = df_ref$id) %>% 
+    select(-long_ref, -lat_ref)
+}
+
+(stations <- do.call("rbind", purrr::map(1:nrow(latlon_ref), 
+                             ~calc_dist(latlon_ref[.x, ], cand))))
+```
+
+We can query the climate information into a nested list named `ts` for each station with the `rowwise()` operator. To create a cubble, supply the same identifiers as with the first example.
+
+```{r eval = FALSE}
+sydmel_climate <- stations %>% 
+  rowwise() %>% 
+  mutate(ts = list(meteo_pull_monitors(id, 
+                                       date_min = "2020-01-01", 
+                                       date_max = "2020-12-31",
+                                       var = c("PRCP", "TMAX", "TMIN")) %>% 
+                     select(-id))) %>% 
+  as_cubble(key = id, index = date, coords = c(long, lat))
 ```
@@
 ```{r echo = FALSE}
-climate_small %>% stretch()
+(sydmel_climate <- stations %>% 
+   left_join(aus_climate) %>% 
+   as_cubble(key = id, index = date, coords = c(long, lat)))
 ```
 
-With a cubic framework on mind, different types of manipulation with cubble can be thought of as slicing the cube in various way.
+<!-- Below are the how the nested and long form look like for Australia climate data, which records daily precipitation, maximum and minimum temperature for 55 stations across Australia from 2015- 2020.
+Notice that each station forms a group in both forms and specifically, the nested  and long form have a underlying `rowwise_df` and `grouped_df` respectively.
+-->
+
+<!-- With a cubic framework on mind, different types of manipulation with cubble can be thought of as slicing the cube in various way.
 The table below shows how some `dplyr` verbs are mapped into the operation in a cubble.
 With the existing grouping on the station, additional groupping can be added with `group_by` and removed with `ungrouped`.
-[talk about why it is useful]
+[talk about why it is useful] -->
 
-```{r verbs, echo = FALSE, out.height = "50%", out.width = "100%", fig.cap = "sdfsdf"}
-knitr::include_graphics(here::here("figures/cubble-verbs.png"))
-```
+\newpage
 
-## Cubble verbs
+## Cubble operations
 
-Mention different types of manipulation with cubble:
+### Basics
 
-  - `dplyr` support for cubble: 
-    - basic 5s: mutate, filter, summarise, select, arrange
+- `stretch`: nest to long form
+- `tamp`: long to nest form
+- `migrate`: move selected spatial variables to the long form.
+- `add_dscrb_prct`: summary stats for missingness
+
+dplyr compatibility:
+
+  - mutate, filter, summarise, select, arrange
   - group and ungroup: group_by, ungroup
   - slice family
-  - summarise missing stats
+
+### Combine two cubbles
+
+  - match river and weather gauges data
+  - involve combining two cubbles
+  - join operations combine the two together by appending more rows but what we really want is to bind rows.
+  - bind rows also doesn't work since we want to bind only when there' s a matching????
+  - introduce bind_join
+
+### Hierarchical structure in cubble
+
+  - hierarchical is common. 
+  - Given examples. 
+  - Essence: switch between different levels
+  - introduce `switch_key`
 
@@
 ```{r}
-climate_small %>% 
+aus_climate %>% 
   stretch() %>% 
   filter(year(date) == 2020) %>% 
-  tamp() %>% 
-  mutate(tmax_missing = ifelse(all(is.na(ts$tmax)), TRUE, FALSE)) %>% 
-  filter(!tmax_missing)
+  tamp()
 ```
@@
 # join wmo_id for each station
-to_join <- station %>% select(id, wmo_id)
-out <- climate_small %>% 
-  left_join(to_join, by = c("station" = "id")) %>% 
-  filter(!is.na(wmo_id))
-out
+# to_join <- station %>% select(id, wmo_id)
+# out <- climate_small %>% 
+#   left_join(to_join, by = c("station" = "id")) %>% 
+#   filter(!is.na(wmo_id))
+# out
 ```
@@
 ```{r}
-out %>% 
+aus_climate %>% 
   stretch() %>% 
-  migrate(station, lat, long)
+  migrate(id, lat, long)
 ```

### 2021-10-12 main.tex
more data structure section, new create a cubble section, files to top level

--- 2021-08-10 main/main.tex
@@
 \usepackage{microtype}      % microtypography
+\usepackage{lipsum}
 \usepackage{graphicx}
@@
   }
+
+\usepackage{color}
+\usepackage{fancyvrb}
+\newcommand{\VerbBar}{|}
+\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
+\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
+% Add ',fontsize=\small' for more characters per line
+\usepackage{framed}
+\definecolor{shadecolor}{RGB}{248,248,248}
+\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
+\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
+\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
+\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\BuiltInTok}[1]{#1}
+\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
+\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
+\newcommand{\ExtensionTok}[1]{#1}
+\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ImportTok}[1]{#1}
+\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\NormalTok}[1]{#1}
+\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
+\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
+\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\RegionMarkerTok}[1]{#1}
+\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
 
@@
 
+\newpage
+
 \hypertarget{introduction}{%
@@
 
+\newpage
+
 \hypertarget{a-new-data-structure-for-spatio-temporal-data}{%
@@
 
-In spatio-temporal data, a variable is usually either group-related or
-time-related. While time-related variables record variables that change
-in time, group-related variables are invariant to the time and have only
-one value per group. A cubble simplifies the workflow in spatio-temporal
-data through managing group and time-related variables separately in two
-forms: list-column and long form. The list column form:
+Spatio-temporal data don't usually come to the analysts as a whole
+piece. A way to look at these data is to divide it into spatial and
+temporal dimension with an ID that links between the two. The first row
+in Figure \ref{fig:cubble-diagram} illustrates this representation where
+in the spatial dimension, the data is characterised by \texttt{id},
+\texttt{lat}, \texttt{long}. \(V_s\) in the last column represents all
+the other site-wise variables, for example, elevation and full name etc.
+The temporal dimension, on the other hand, can be characterised by
+\texttt{id} and \texttt{t} with \(V_t\) representing all the time-wise
+variables. In climate data, this could include precipitation, maximum or
+minimum temperature, and wind speed etc.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/cubble-diagram/cubble-diagram.001}
+
+}
+
+\caption{Cubble diagram}\label{fig:cubble-diagram}
+\end{figure}
+
+To work with spatio-temporal data, analysts can choose to either work
+separately on each dimension or join the two sets together, however,
+each approach has its own problem: While is is natural to work
+separately on each sheet (since spatial and temporal operations usually
+don't overlap), analysts will need to manually keep the other data frame
+up to date. For example, the following pseudo code illustrates the
+scenario where once the spatial dataset is filtered for those within
+Victoria, the temporal dataset needs to be manually updated to reflect
+this spatial filter.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{spatial\_new }\OtherTok{\textless{}{-}}\NormalTok{ spatial }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(SITES\_IN\_VICTORIA)}
+\NormalTok{temporal\_new }\OtherTok{\textless{}{-}}\NormalTok{ temporal }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(id }\SpecialCharTok{\%in\%}\NormalTok{ spatial\_new}\SpecialCharTok{$}\NormalTok{id)}
+\end{Highlighting}
+\end{Shaded}
+
+If analysts choose to join the spatial and temporal data together, the
+joined dataset could be too large since each spatial variable will be
+repeated at each time stamp for each site. Also, recordings of the site
+ID from different data sources can be slightly different from each
+other, causing a painful checking and cleaning of site IDs before the
+join.
+
+A cubble, in essence, wires both dimensions in the spatio-temporal data
+into one object while provide two forms for manipulation the spatial and
+temporal dimension separately.
+
+When manipulating the spatial dimension it uses a nest form that:
 
@@
 
-In the long form,
+When manipulating the temporal dimension, it uses the long form that:
 
@@
   group-related variables are not explicitly displayed but can be
-  accessed through the \texttt{meta} attribute
+  accessed through the \texttt{meta} attribute.
 \end{itemize}
-
-Below are the how the list-column and long form look like for Australia
-climate data, which records daily precipitation, maximum and minimum
-temperature for 55 stations across Australia from 2015- 2020.
 
@@
 
+\hypertarget{create-a-cubble}{%
+\section{Create a cubble}\label{create-a-cubble}}
+
+The creation of a cubble requires the site identifier (\texttt{key}), as
+well as the spatial (\texttt{coords}) and temporal (\texttt{index})
+identifier. \texttt{climate\_flat} is already a tibble and it uses
+\texttt{id} to identify each station, \texttt{date} as the time
+identifier, and \texttt{c(long,\ lat)} as the spatial identifier. To
+create a cubble for this data, use:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{climate\_flat }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{as\_cubble}\NormalTok{(}\AttributeTok{key =}\NormalTok{ id, }\AttributeTok{index =}\NormalTok{ date, }\AttributeTok{coords =} \FunctionTok{c}\NormalTok{(long, lat))}
+\end{Highlighting}
+\end{Shaded}
+
 \begin{verbatim}
-## # Cubble: station-wise: list-column
-## # Group:  station [55]
-## # Meta:   station [fct], lat [dbl], long [dbl], elevation [dbl], name [fct]
-##    station       lat  long elevation name                ts                  
-##    <fct>       <dbl> <dbl>     <dbl> <fct>               <list>              
-##  1 ASN00001019 -14.3  127.      23   kalumburu           <tbl_ts [2,147 x 4]>
-##  2 ASN00002012 -18.2  128.     422   halls creek airport <tbl_ts [2,191 x 4]>
-##  3 ASN00003003 -17.9  122.       7.4 broome airport      <tbl_ts [2,192 x 4]>
-##  4 ASN00006011 -24.9  114.       4   carnarvon airport   <tbl_ts [2,192 x 4]>
-##  5 ASN00009021 -31.9  116.      15.4 perth airport       <tbl_ts [2,192 x 4]>
-##  6 ASN00009193 -32.0  116.      43.1 rottnest island     <tbl_ts [2,192 x 4]>
-##  7 ASN00009518 -34.4  115.      13   cape leeuwin        <tbl_ts [2,184 x 4]>
-##  8 ASN00009789 -33.8  122.      25   esperance           <tbl_ts [2,187 x 4]>
-##  9 ASN00010286 -31.6  117.     217.  cunderdin airfield  <tbl_ts [2,190 x 4]>
-## 10 ASN00010917 -32.7  117.     275   wandering           <tbl_ts [2,192 x 4]>
-## # ... with 45 more rows
+## # Cubble: id-wise: nested form
+## # Key:    id [5]
+## # Leaves: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
+##   id            lat  long  elev name           wmo_id ts                
+##   <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>            
+## 1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tibble [366 x 4]>
+## 2 ASN00010311 -31.9  117. 179   york            94623 <tibble [366 x 4]>
+## 3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble [366 x 4]>
+## 4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble [366 x 4]>
+## 5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble [366 x 4]>
 \end{verbatim}
 
+Most of the time, spatio-temporal data doesn't come into this form and
+analysts need to query the climate variables based on station metadata.
+For this type of task, one can structure a metadata into a tibble and
+use row-wise operator to query the climate variables into a nested list.
+As an example here we demonstrate the workflow to find the 5 closest
+stations to Melbourne. We first create a station data farme with the 5
+target stations.
+
 \begin{verbatim}
-## # Cubble: time-wise: long form [tsibble]
-## # Group:  station [55]
-## # Meta:   station [fct], lat [dbl], long [dbl], elevation [dbl], name [fct]
-##    station     date        prcp  tmax  tmin
-##    <fct>       <date>     <dbl> <dbl> <dbl>
-##  1 ASN00001019 2015-01-01   164  31.5  25  
-##  2 ASN00001019 2015-01-02   124  31.3  25  
-##  3 ASN00001019 2015-01-03    50  29.2  24.7
-##  4 ASN00001019 2015-01-04   204  32.1  24.3
-##  5 ASN00001019 2015-01-05   412  31.4  24.6
-##  6 ASN00001019 2015-01-06   200  28.7  25.1
-##  7 ASN00001019 2015-01-07   822  30.2  24.5
-##  8 ASN00001019 2015-01-08    22  31    26.6
-##  9 ASN00001019 2015-01-09    62  32.7  26.2
-## 10 ASN00001019 2015-01-10   274  32.9  22.6
-## # ... with 120,311 more rows
+## # A tibble: 5 x 8
+##   id            lat  long  elev name                 wmo_id  dist city     
+##   <chr>       <dbl> <dbl> <dbl> <chr>                 <dbl> <dbl> <chr>    
+## 1 ASN00086038 -37.7  145.  78.4 essendon airport      95866  10.8 melbourne
+## 2 ASN00086282 -37.7  145. 113.  melbourne airport     94866  20.1 melbourne
+## 3 ASN00086077 -38.0  145.  12.1 moorabbin airport     94870  21.9 melbourne
+## 4 ASN00088162 -37.4  145. 528.  wallan (kilmore gap)  94860  48.1 melbourne
+## 5 ASN00087113 -38.0  144.  10.6 avalon airport        94854  48.8 melbourne
 \end{verbatim}
 
-\hypertarget{manipulation-and-visualisation-with-cubble}{%
-\section{\texorpdfstring{Manipulation and visualisation with
-\texttt{cubble}}{Manipulation and visualisation with cubble}}\label{manipulation-and-visualisation-with-cubble}}
+We can query the climate information into a nested list named
+\texttt{ts} for each station with the \texttt{rowwise()} operator. To
+create a cubble, supply the same identifiers as with the first example.
 
-\hypertarget{cubble-verbs}{%
-\subsection{Cubble verbs}\label{cubble-verbs}}
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{sydmel\_climate }\OtherTok{\textless{}{-}}\NormalTok{ stations }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{rowwise}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{ts =} \FunctionTok{list}\NormalTok{(}\FunctionTok{meteo\_pull\_monitors}\NormalTok{(id, }
+                                       \AttributeTok{date\_min =} \StringTok{"2020{-}01{-}01"}\NormalTok{, }
+                                       \AttributeTok{date\_max =} \StringTok{"2020{-}12{-}31"}\NormalTok{,}
+                                       \AttributeTok{var =} \FunctionTok{c}\NormalTok{(}\StringTok{"PRCP"}\NormalTok{, }\StringTok{"TMAX"}\NormalTok{, }\StringTok{"TMIN"}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
+                     \FunctionTok{select}\NormalTok{(}\SpecialCharTok{{-}}\NormalTok{id))) }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{as\_cubble}\NormalTok{(}\AttributeTok{key =}\NormalTok{ id, }\AttributeTok{index =}\NormalTok{ date, }\AttributeTok{coords =} \FunctionTok{c}\NormalTok{(long, lat))}
+\end{Highlighting}
+\end{Shaded}
 
-Mention different types of manipulation with cubble:
+\begin{verbatim}
+## # Cubble: id-wise: nested form
+## # Key:    id [5]
+## # Leaves: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
+##   id            lat  long  elev name                 wmo_id  dist city   ts     
+##   <chr>       <dbl> <dbl> <dbl> <chr>                 <dbl> <dbl> <chr>  <list> 
+## 1 ASN00086038 -37.7  145.  78.4 essendon airport      95866  10.8 melbo~ <tibbl~
+## 2 ASN00086282 -37.7  145. 113.  melbourne airport     94866  20.1 melbo~ <tibbl~
+## 3 ASN00086077 -38.0  145.  12.1 moorabbin airport     94870  21.9 melbo~ <tibbl~
+## 4 ASN00088162 -37.4  145. 528.  wallan (kilmore gap)  94860  48.1 melbo~ <tibbl~
+## 5 ASN00087113 -38.0  144.  10.6 avalon airport        94854  48.8 melbo~ <tibbl~
+\end{verbatim}
+
+\newpage
+
+\hypertarget{cubble-operations}{%
+\subsection{Cubble operations}\label{cubble-operations}}
+
+\hypertarget{basics}{%
+\subsubsection{Basics}\label{basics}}
 
@@
 \item
-  \texttt{dplyr} support for cubble:
+  \texttt{stretch}: nest to long form
+\item
+  \texttt{tamp}: long to nest form
+\item
+  \texttt{migrate}: move selected spatial variables to the long form.
+\item
+  \texttt{add\_dscrb\_prct}: summary stats for missingness
+\end{itemize}
+
+dplyr compatibility:
 
@@
 \item
-    basic 5s: mutate, filter, summarise, select, arrange
+  mutate, filter, summarise, select, arrange
 \item
@@
 \end{itemize}
+
+\hypertarget{combine-two-cubbles}{%
+\subsubsection{Combine two cubbles}\label{combine-two-cubbles}}
+
+\begin{itemize}
+\tightlist
 \item
-  summarise missing stats
+  match river and weather gauges data
+\item
+  involve combining two cubbles
+\item
+  join operations combine the two together by appending more rows but
+  what we really want is to bind rows.
+\item
+  bind rows also doesn't work since we want to bind only when there' s a
+  matching????
+\item
+  introduce bind\_join
 \end{itemize}
+
+\hypertarget{hierarchical-structure-in-cubble}{%
+\subsubsection{Hierarchical structure in
+cubble}\label{hierarchical-structure-in-cubble}}
+
+\begin{itemize}
+\tightlist
+\item
+  hierarchical is common.
+\item
+  Given examples.
+\item
+  Essence: switch between different levels
+\item
+  introduce \texttt{switch\_key}
+\end{itemize}
+
+\newpage
 
@@
 \subsection{Manipulation}\label{manipulation}}
+
+\hypertarget{mutate-and-filter}{%
+\subsubsection{Mutate and filter}\label{mutate-and-filter}}
+
+In the first example, we want to only keep the stations that have
+\texttt{tmax} recorded in 2020. This requires first narrow down the
+records to those in 2020, determine if \texttt{tmax} is missing for each
+station, and then retain those stations that have \texttt{tmax}
+recorded. The year filtering is an operation on the time axis, so we
+start with the long form. Whether each station has \texttt{tmax}
+recorded is a result of each station, rather than of each time point,
+hence we need to switch to the nested form with \texttt{tamp()}. To
+calculate whether \texttt{tmax} is recorded, we mutate a column
+\texttt{tmax\_missing} that takes \texttt{TRUE} if all the \texttt{tmax}
+in the nested list column \texttt{ts} are \texttt{NA} and \texttt{FALSE}
+otherwise. To get the stations that we want, we need another filter on
+\texttt{tmax\_missing}.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{aus\_climate }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{stretch}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{filter}\NormalTok{(}\FunctionTok{year}\NormalTok{(date) }\SpecialCharTok{==} \DecValTok{2020}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{tamp}\NormalTok{()}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## # Cubble: id-wise: nested form
+## # Key:    id [85]
+## # Leaves: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
+##    id            lat  long  elev name                   wmo_id ts               
+##    <chr>       <dbl> <dbl> <dbl> <chr>                   <dbl> <list>           
+##  1 ASN00009021 -31.9  116.  15.4 perth airport           94610 <tibble [366 x 4~
+##  2 ASN00010311 -31.9  117. 179   york                    94623 <tibble [366 x 4~
+##  3 ASN00010614 -32.9  117. 338   narrogin                94627 <tibble [366 x 4~
+##  4 ASN00014015 -12.4  131.  30.4 darwin airport          94120 <tibble [366 x 4~
+##  5 ASN00015131 -17.6  134. 220   elliott                 94236 <tibble [366 x 4~
+##  6 ASN00016065 -30.4  137.  76   andamooka               95660 <tibble [366 x 4~
+##  7 ASN00023034 -35.0  139.   2   adelaide airport        94672 <tibble [366 x 4~
+##  8 ASN00023373 -34.5  139. 275   nuriootpa viticultural  94681 <tibble [366 x 4~
+##  9 ASN00024024 -34.4  141.  30.1 loxton research centre  94682 <tibble [366 x 4~
+## 10 ASN00024048 -34.2  141.  31.5 renmark aero            95687 <tibble [366 x 4~
+## # ... with 75 more rows
+\end{verbatim}
+
+\hypertarget{join}{%
+\subsubsection{Join}\label{join}}
+
+Now we want to select the stations that have been registered with world
+meteorological organisation (WMO) and the dataset \texttt{station} has a
+column \texttt{wmo\_id} that records this information. To do this task,
+we first need to join the \texttt{station} dataset with our climate
+dataset and then filter out those stations that don't have the WMO id.
+Since the join is by station rather than by time, we start with the
+nested form and write the exact same syntax of join and filter as with
+tidyverse.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# join wmo\_id for each station}
+\CommentTok{\# to\_join \textless{}{-} station \%\textgreater{}\% select(id, wmo\_id)}
+\CommentTok{\# out \textless{}{-} climate\_small \%\textgreater{}\% }
+\CommentTok{\#   left\_join(to\_join, by = c("station" = "id")) \%\textgreater{}\% }
+\CommentTok{\#   filter(!is.na(wmo\_id))}
+\CommentTok{\# out}
+\end{Highlighting}
+\end{Shaded}
+
+Sometimes, we would like to have station-wise and time-wise variables in
+the same form (i.e.~when plotting glyph maps). This can also be seen as
+a joining task, on the long form, with the dataset to join being the
+metadata. \texttt{migrate()} is a verb introduced as the shortcut for
+\texttt{left\_join()} with a cubble's metadata and below is the
+comparison of the two syntaxes.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{aus\_climate }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{stretch}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{migrate}\NormalTok{(id, lat, long)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## # Cubble: time-wise: long form
+## # Key:    id [85]
+## # Leaves: id [chr], lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id
+## #   [dbl], prcp_missing [dbl], tmax_missing [dbl], tmin_missing [dbl]
+##    id          date        prcp  tmax  tmin   lat  long
+##    <chr>       <date>     <dbl> <dbl> <dbl> <dbl> <dbl>
+##  1 ASN00009021 2020-01-01     0   319   153 -31.9  116.
+##  2 ASN00009021 2020-01-02     0   249   164 -31.9  116.
+##  3 ASN00009021 2020-01-03     6   232   130 -31.9  116.
+##  4 ASN00009021 2020-01-04     0   284   124 -31.9  116.
+##  5 ASN00009021 2020-01-05     0   353   116 -31.9  116.
+##  6 ASN00009021 2020-01-06     0   348   131 -31.9  116.
+##  7 ASN00009021 2020-01-07     0   328   151 -31.9  116.
+##  8 ASN00009021 2020-01-08     0   304   174 -31.9  116.
+##  9 ASN00009021 2020-01-09     0   287   173 -31.9  116.
+## 10 ASN00009021 2020-01-10     0   326   158 -31.9  116.
+## # ... with 31,100 more rows
+\end{verbatim}
 

### 2021-10-12 data/create-data.R

@@ -1,0 +1,14 @@
+library(tidyverse)
+library(cubble)
+library(s2)
+library(ggplot2)
+state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)
+# plot_map(state_map) +
+#   geom_point(data = weatherdata::climate_large, aes(x = longitude, y = latitude))
+vic_map <- ozmaps::abs_ste %>% filter(NAME == "Victoria")
+vic_stations <- weatherdata::climate_large %>% 
+  rename(long = longitude, lat = latitude) %>% 
+  mutate(ll = s2_lnglat(long, lat)) %>% 
+  filter(s2_within(ll, vic_map)) 
+
+save(vic_stations, file = here::here("data/vic_stations.rda"))

### 2021-10-19 main.Rmd
some notes

@@
 
+\textcolor{red}{Under the third tidy data principle [@tidydata], these are two type of observational unit and should form different tables.
+An example (Table 8) is also given to demonstrate the billboard dataset where there are song variables and rank variables that changes each week.
+The other has mentioned the lack of tool to work directly with relational data and analysis will usually need to merge the dataset into one table, which could causes errors.
+These days more implementations have looked at relational data structure and in R, several good examples include, `tidygraph`[@tidygraph] for graph manipulation, `dm` [@dm] for relational data model.
+In the domain of spatio-temporal data, @spacetime has considered four types of spatio-temporal layout.
+`Spatstat` [@Spatstat] an implementation of point pattern.
+}
+
 \newpage
@@
 ```{r cubble-diagram, echo = FALSE, fig.cap="Cubble diagram", out.height="40%", fig.align='center'}
-knitr::include_graphics(here::here("figures/cubble-diagram/cubble-diagram.001.png"))
+knitr::include_graphics(here::here("figures/input-formats/input-formats.001.png"))
 ```
@@
 Most of the time, spatio-temporal data doesn't come into this form and analysts need to query the climate variables based on station metadata.
-For this type of task, one can structure a metadata into a tibble and use row-wise operator to query the climate variables into a nested list.
+\textcolor{red}{This is also a problem illustrated in Section 3.5 in @tidydata.
+Here we provide a structured way to query this data based on the row-wise operator and nested list.} For this type of task, one can structure a metadata into a tibble and use row-wise operator to query the climate variables into a nested list.
 As an example here we demonstrate the workflow to find the 5 closest stations to Melbourne.
-We first create a station data farme with the 5 target stations.
+We first create a station data frame with the 5 target stations.
 
@@
 ## Cubble operations
+
+```{r cubble-operations, echo = FALSE, fig.cap="Cubble operations", out.height="50%", fig.align='center'}
+knitr::include_graphics(here::here("figures/cubble-operations/cubble-operations.001.png"))
+```
 

### 2021-10-19 main.tex
some notes

@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/cubble-diagram/cubble-diagram.001}
+{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/input-formats/input-formats.001}
 
@@
 analysts need to query the climate variables based on station metadata.
+\textcolor{red}{This is also a problem illustrated in Section 3.5 in @tidydata. Here we provide a structured way to query this data based on the row-wise operator and nested list.}
 For this type of task, one can structure a metadata into a tibble and
@@
 As an example here we demonstrate the workflow to find the 5 closest
-stations to Melbourne. We first create a station data farme with the 5
+stations to Melbourne. We first create a station data frame with the 5
 target stations.
@@
 \subsection{Cubble operations}\label{cubble-operations}}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth,height=0.5\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/cubble-operations/cubble-operations.001}
+
+}
+
+\caption{Cubble operations}\label{fig:cubble-operations}
+\end{figure}
 

### 2021-11-20 main.Rmd
change to JSS format

@@
 ---
-title: A template for the *arxiv* style
-authors:
-  - name: H.Sherry Zhang
-    department: Department of Econometrics and Business Statistics 
+documentclass: jss
+author:
+  - name: H. Sherry Zhang
+    # department: Department of Econometrics and Business Statistics
     affiliation: Monash University
-    location: Melbourne, Australia
-    email:  huize.zhang@monash.edu
+    #location: Melbourne, Australia
+    email: \email{huize.zhang@monash.edu}
   - name: Dianne Cook
-    department: Department of Econometrics and Business Statistics 
-    affiliation: Monash University 
-    location: Melbourne, Australia
-    email:  dicook@monash.edu
+    # department: Department of Econometrics and Business Statistics
+    affiliation: 'Monash University \AND'
+    # location: Melbourne, Australia
+    # email:  dicook@monash.edu
   - name: Ursula Laa
-    department: Institute of Statistics
-    affiliation: University of Natural Resources and Life Sciences
-    location: Vienna, Austria
-    email:  ursula.laa@boku.ac.at  
+    # department: Institute of Statistics
+    affiliation: 'University of Natural Resources and Life Sciences \AND'
+    # location: Vienna, Austria
+    # email:  ursula.laa@boku.ac.at
   - name: Nicolas Langrené
-    department: 34 Village Street, Docklands VIC 3008 Australia
-    affiliation: CSIRO Data61 
-    location: Melbourne, Australia
-    email: nicolas.langrene@csiro.au
+    # department: 34 Village Street, Docklands VIC 3008 Australia
+    affiliation: 'CSIRO Data61 \AND'
+    # location: Melbourne, Australia
+    # email: nicolas.langrene@csiro.au
   - name: Patricia Menéndez
-    department: Department of Econometrics and Business Statistics 
-    affiliation: Monash University 
-    location: Melbourne, Australia
-    email:  patricia.menendez@monash.edu 
-abstract: |
-  Enter the text of your abstract here.
+    # department: Department of Econometrics and Business Statistics
+    affiliation: 'Monash University \AND'
+    # location: Melbourne, Australia
+    # email:  patricia.menendez@monash.edu
+title:
+  formatted: "\\pkg{cubble}: An R Package for Structuring Spatio-temporal Data"
+  # If you use tex in the formatted title, alsoge supply version without
+  plain:     "cubble: An R Package for Structuring Spatio-temporal Data"
+  # For running headers, if needed
+  # short:     "\\pkg{foo}: A Capitalized Title"
+abstract: >
+  The abstract of the article.
 keywords:
-  - blah
-  - blee
-  - bloo
-  - these are optional and can be removed
+  # at least one keyword must be supplied
+  formatted: [spatio-temporal data,  "\\proglang{R}"]
+  plain:     [spatio-temporal data, R]
+preamble: >
+  \usepackage{amsmath}
+output: rticles::jss_article
 bibliography: references.bib
-biblio-style: unsrt
-output: rticles::arxiv_article
 ---
@@
                       out.height = "30%")
+options(prompt = 'R> ', continue = '+ ')
 ```
@@
 cand <- aus_climate %>% 
-  select(id, lat, long, elev, name, wmo_id) %>% 
-   as_tibble()
+  as_tibble() %>% 
+  select(id, lat, long, elev, name, wmo_id)
 
@@
 
-```{r cubble-operations, echo = FALSE, fig.cap="Cubble operations", out.height="50%", fig.align='center'}
+```{r cubble-operations, echo = FALSE, fig.cap="Cubble operations", out.height="50%", out.width="90%", fig.align='center'}
 knitr::include_graphics(here::here("figures/cubble-operations/cubble-operations.001.png"))

### 2021-11-20 main.tex
change to JSS format

@@
-\documentclass{article}
+\documentclass[
+]{jss}
 
-\usepackage{arxiv}
+\usepackage[utf8]{inputenc}
 
-\usepackage[utf8]{inputenc} % allow utf-8 input
-\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
-\usepackage{lmodern}        % https://github.com/rstudio/rticles/issues/343
-\usepackage{hyperref}       % hyperlinks
-\usepackage{url}            % simple URL typesetting
-\usepackage{booktabs}       % professional-quality tables
-\usepackage{amsfonts}       % blackboard math symbols
-\usepackage{nicefrac}       % compact symbols for 1/2, etc.
-\usepackage{microtype}      % microtypography
-\usepackage{lipsum}
-\usepackage{graphicx}
-
-\title{A template for the \emph{arxiv} style}
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
 
 \author{
-    H.Sherry Zhang
-   \\
-    Department of Econometrics and Business Statistics \\
-    Monash University \\
-  Melbourne, Australia \\
-  \texttt{\href{mailto:huize.zhang@monash.edu}{\nolinkurl{huize.zhang@monash.edu}}} \\
-   \And
-    Dianne Cook
-   \\
-    Department of Econometrics and Business Statistics \\
-    Monash University \\
-  Melbourne, Australia \\
-  \texttt{\href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}} \\
-   \And
-    Ursula Laa
-   \\
-    Institute of Statistics \\
-    University of Natural Resources and Life Sciences \\
-  Vienna, Austria \\
-  \texttt{\href{mailto:ursula.laa@boku.ac.at}{\nolinkurl{ursula.laa@boku.ac.at}}} \\
-   \And
-    Nicolas Langrené
-   \\
-    34 Village Street, Docklands VIC 3008 Australia \\
-    CSIRO Data61 \\
-  Melbourne, Australia \\
-  \texttt{\href{mailto:nicolas.langrene@csiro.au}{\nolinkurl{nicolas.langrene@csiro.au}}} \\
-   \And
-    Patricia Menéndez
-   \\
-    Department of Econometrics and Business Statistics \\
-    Monash University \\
-  Melbourne, Australia \\
-  \texttt{\href{mailto:patricia.menendez@monash.edu}{\nolinkurl{patricia.menendez@monash.edu}}} \\
+H. Sherry Zhang\\Monash University \And Dianne Cook\\Monash University
+\AND Ursula Laa\\University of Natural Resources and Life Sciences
+\AND Nicolas Langrené\\CSIRO Data61 \AND Patricia
+Menéndez\\Monash University \AND
+}
+\title{\pkg{cubble}: An R Package for Structuring Spatio-temporal Data}
+
+\Plainauthor{H. Sherry Zhang, Dianne Cook, Ursula Laa, Nicolas
+Langrené, Patricia Menéndez}
+\Plaintitle{cubble: An R Package for Structuring Spatio-temporal Data}
+
+\Abstract{
+The abstract of the article.
 }
 
-\usepackage{color}
-\usepackage{fancyvrb}
-\newcommand{\VerbBar}{|}
-\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
-\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
-% Add ',fontsize=\small' for more characters per line
-\usepackage{framed}
-\definecolor{shadecolor}{RGB}{248,248,248}
-\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\BuiltInTok}[1]{#1}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
-\newcommand{\ExtensionTok}[1]{#1}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ImportTok}[1]{#1}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\NormalTok}[1]{#1}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\RegionMarkerTok}[1]{#1}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\Keywords{spatio-temporal data, \proglang{R}}
+\Plainkeywords{spatio-temporal data, R}
+
+%% publication information
+%% \Volume{50}
+%% \Issue{9}
+%% \Month{June}
+%% \Year{2012}
+%% \Submitdate{}
+%% \Acceptdate{2012-06-04}
+
+\Address{
+          }
 
 % Pandoc citation processing
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-% for Pandoc 2.8 to 2.10.1
-\newenvironment{cslreferences}%
-  {}%
-  {\par}
-% For Pandoc 2.11+
-\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
-  % set entry spacing
-  \ifnum #2 > 0
-  \setlength{\parskip}{#2\baselineskip}
-  \fi
- }%
- {}
-\usepackage{calc} % for calculating minipage widths
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
+
+% Pandoc header
+
+\usepackage{amsmath}
 
 \begin{document}
-\maketitle
-
-\def\tightlist{}
-
-\begin{abstract}
-Enter the text of your abstract here.
-\end{abstract}
-
-\keywords{
-    blah
-   \and
-    blee
-   \and
-    bloo
-   \and
-    these are optional and can be removed
-  }
 
@@
 the R-spatial team for both raster and vector spatial data. For vector
-spatial data, which is the focus of this paper, \texttt{sf} (E. J.
-Pebesma 2018) represents spatial vector information with simple
-features: points, lines, polygons and their multiples. Various
+spatial data, which is the focus of this paper, \texttt{sf}
+\citep{pebesma2018simple} represents spatial vector information with
+simple features: points, lines, polygons and their multiples. Various
 \texttt{st\_} function are designed to manipulate these features based
 on their geometric relationships. For spatio-temporal data,
-\texttt{stars} (E. Pebesma 2021) can represent both raster and vector
-data using multi-dimensional array. However, the underlying array
-structure can be difficult to operate for data analysts who are more
-familiar with a flat 2D data frame structure used by the tidyverse
-ecosystem.
+\texttt{stars} \citep{stars} can represent both raster and vector data
+using multi-dimensional array. However, the underlying array structure
+can be difficult to operate for data analysts who are more familiar with
+a flat 2D data frame structure used by the tidyverse ecosystem.
 
-In the temporal aspect, the \texttt{tsibble} (Wang, Cook, and Hyndman
-2020) structure and its tidyverts ecosystem have provided a {[}\ldots{}
-{]} workflow to work with temporal data. In a tsibble structure,
-temporal data is characterised by \texttt{index} and \texttt{key} where
-\texttt{index} is the temporal identifier and \texttt{key} is the
-identifier for multiple series, which could be used as a spatio
-identifier. However, a tsibble object, by construction, always requires
-the \texttt{index} in its structure. This makes it less appealing for
-spatio-temporal data since the output of calculated spatio-specific
-variables (i.e.~features of each series) don't have the time dimension.
-Analysts will either need to have an additional step to join this output
-to the original tsibble or operate with variables stored in two separate
-objects. In addition, the long form structure of a tsibble object means
-spatio variables (i.e.~longitude, latitude, and features of each series
-if joined back to the tsibble) of each spatio identifier will be
-repetitively recorded at each timestamp. This repetition is unnecessary
-and would inflate the object size for long series.
+In the temporal aspect, the \texttt{tsibble} \citep{tsibbles} structure
+and its tidyverts ecosystem have provided a {[}\ldots{} {]} workflow to
+work with temporal data. In a tsibble structure, temporal data is
+characterised by \texttt{index} and \texttt{key} where \texttt{index} is
+the temporal identifier and \texttt{key} is the identifier for multiple
+series, which could be used as a spatio identifier. However, a tsibble
+object, by construction, always requires the \texttt{index} in its
+structure. This makes it less appealing for spatio-temporal data since
+the output of calculated spatio-specific variables (i.e.~features of
+each series) don't have the time dimension. Analysts will either need to
+have an additional step to join this output to the original tsibble or
+operate with variables stored in two separate objects. In addition, the
+long form structure of a tsibble object means spatio variables
+(i.e.~longitude, latitude, and features of each series if joined back to
+the tsibble) of each spatio identifier will be repetitively recorded at
+each timestamp. This repetition is unnecessary and would inflate the
+object size for long series.
 
@@
 
+\begin{CodeChunk}
 \begin{figure}
@@
 
-\caption{Cubble diagram}\label{fig:cubble-diagram}
+\caption[Cubble diagram]{Cubble diagram}\label{fig:cubble-diagram}
 \end{figure}
+\end{CodeChunk}
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{spatial\_new }\OtherTok{\textless{}{-}}\NormalTok{ spatial }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(SITES\_IN\_VICTORIA)}
-\NormalTok{temporal\_new }\OtherTok{\textless{}{-}}\NormalTok{ temporal }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(id }\SpecialCharTok{\%in\%}\NormalTok{ spatial\_new}\SpecialCharTok{$}\NormalTok{id)}
-\end{Highlighting}
-\end{Shaded}
+\begin{CodeChunk}
+\begin{CodeInput}
+R> spatial_new <- spatial %>% filter(SITES_IN_VICTORIA)
+R> temporal_new <- temporal %>% filter(id %in% spatial_new$id)
+\end{CodeInput}
+\end{CodeChunk}
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{climate\_flat }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{as\_cubble}\NormalTok{(}\AttributeTok{key =}\NormalTok{ id, }\AttributeTok{index =}\NormalTok{ date, }\AttributeTok{coords =} \FunctionTok{c}\NormalTok{(long, lat))}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## # Cubble: id-wise: nested form
-## # Key:    id [5]
-## # Leaves: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
-##   id            lat  long  elev name           wmo_id ts                
-##   <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>            
-## 1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tibble [366 x 4]>
-## 2 ASN00010311 -31.9  117. 179   york            94623 <tibble [366 x 4]>
-## 3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble [366 x 4]>
-## 4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble [366 x 4]>
-## 5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble [366 x 4]>
-\end{verbatim}
+\begin{CodeChunk}
+\begin{CodeInput}
+R> climate_flat %>% as_cubble(key = id, index = date, coords = c(long, lat))
+\end{CodeInput}
+\begin{CodeOutput}
+# cubble:   id [5]: nested form
+# bbox:     [115.97, -32.94, 133.55, -12.42]- check gap on long and lat
+# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
+  id            lat  long  elev name           wmo_id ts                
+  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>            
+1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tibble [366 x 4]>
+2 ASN00010311 -31.9  117. 179   york            94623 <tibble [366 x 4]>
+3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble [366 x 4]>
+4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble [366 x 4]>
+5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble [366 x 4]>
+\end{CodeOutput}
+\end{CodeChunk}
 
@@
 
-\begin{verbatim}
-## # A tibble: 5 x 8
-##   id            lat  long  elev name                 wmo_id  dist city     
-##   <chr>       <dbl> <dbl> <dbl> <chr>                 <dbl> <dbl> <chr>    
-## 1 ASN00086038 -37.7  145.  78.4 essendon airport      95866  10.8 melbourne
-## 2 ASN00086282 -37.7  145. 113.  melbourne airport     94866  20.1 melbourne
-## 3 ASN00086077 -38.0  145.  12.1 moorabbin airport     94870  21.9 melbourne
-## 4 ASN00088162 -37.4  145. 528.  wallan (kilmore gap)  94860  48.1 melbourne
-## 5 ASN00087113 -38.0  144.  10.6 avalon airport        94854  48.8 melbourne
-\end{verbatim}
+\begin{CodeChunk}
+\begin{CodeOutput}
+# A tibble: 5 x 8
+  id            lat  long  elev name                 wmo_id  dist city     
+  <chr>       <dbl> <dbl> <dbl> <chr>                 <dbl> <dbl> <chr>    
+1 ASN00086038 -37.7  145.  78.4 essendon airport      95866  10.8 melbourne
+2 ASN00086282 -37.7  145. 113.  melbourne airport     94866  20.1 melbourne
+3 ASN00086077 -38.0  145.  12.1 moorabbin airport     94870  21.9 melbourne
+4 ASN00088162 -37.4  145. 528.  wallan (kilmore gap)  94860  48.1 melbourne
+5 ASN00087113 -38.0  144.  10.6 avalon airport        94854  48.8 melbourne
+\end{CodeOutput}
+\end{CodeChunk}
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{sydmel\_climate }\OtherTok{\textless{}{-}}\NormalTok{ stations }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{rowwise}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{ts =} \FunctionTok{list}\NormalTok{(}\FunctionTok{meteo\_pull\_monitors}\NormalTok{(id, }
-                                       \AttributeTok{date\_min =} \StringTok{"2020{-}01{-}01"}\NormalTok{, }
-                                       \AttributeTok{date\_max =} \StringTok{"2020{-}12{-}31"}\NormalTok{,}
-                                       \AttributeTok{var =} \FunctionTok{c}\NormalTok{(}\StringTok{"PRCP"}\NormalTok{, }\StringTok{"TMAX"}\NormalTok{, }\StringTok{"TMIN"}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
-                     \FunctionTok{select}\NormalTok{(}\SpecialCharTok{{-}}\NormalTok{id))) }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{as\_cubble}\NormalTok{(}\AttributeTok{key =}\NormalTok{ id, }\AttributeTok{index =}\NormalTok{ date, }\AttributeTok{coords =} \FunctionTok{c}\NormalTok{(long, lat))}
-\end{Highlighting}
-\end{Shaded}
+\begin{CodeChunk}
+\begin{CodeInput}
+R> sydmel_climate <- stations %>% 
++   rowwise() %>% 
++   mutate(ts = list(meteo_pull_monitors(id, 
++                                        date_min = "2020-01-01", 
++                                        date_max = "2020-12-31",
++                                        var = c("PRCP", "TMAX", "TMIN")) %>% 
++                      select(-id))) %>% 
++   as_cubble(key = id, index = date, coords = c(long, lat))
+\end{CodeInput}
+\end{CodeChunk}
 
-\begin{verbatim}
-## # Cubble: id-wise: nested form
-## # Key:    id [5]
-## # Leaves: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
-##   id            lat  long  elev name                 wmo_id  dist city   ts     
-##   <chr>       <dbl> <dbl> <dbl> <chr>                 <dbl> <dbl> <chr>  <list> 
-## 1 ASN00086038 -37.7  145.  78.4 essendon airport      95866  10.8 melbo~ <tibbl~
-## 2 ASN00086282 -37.7  145. 113.  melbourne airport     94866  20.1 melbo~ <tibbl~
-## 3 ASN00086077 -38.0  145.  12.1 moorabbin airport     94870  21.9 melbo~ <tibbl~
-## 4 ASN00088162 -37.4  145. 528.  wallan (kilmore gap)  94860  48.1 melbo~ <tibbl~
-## 5 ASN00087113 -38.0  144.  10.6 avalon airport        94854  48.8 melbo~ <tibbl~
-\end{verbatim}
+\begin{CodeChunk}
+\begin{CodeOutput}
+# cubble:   id [5]: nested form
+# bbox:     [144.47, -38.03, 145.1, -37.38]
+# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
+  id            lat  long  elev name                 wmo_id  dist city   ts     
+  <chr>       <dbl> <dbl> <dbl> <chr>                 <dbl> <dbl> <chr>  <list> 
+1 ASN00086038 -37.7  145.  78.4 essendon airport      95866  10.8 melbo~ <tibbl~
+2 ASN00086282 -37.7  145. 113.  melbourne airport     94866  20.1 melbo~ <tibbl~
+3 ASN00086077 -38.0  145.  12.1 moorabbin airport     94870  21.9 melbo~ <tibbl~
+4 ASN00088162 -37.4  145. 528.  wallan (kilmore gap)  94860  48.1 melbo~ <tibbl~
+5 ASN00087113 -38.0  144.  10.6 avalon airport        94854  48.8 melbo~ <tibbl~
+\end{CodeOutput}
+\end{CodeChunk}
 
@@
 
+\begin{CodeChunk}
 \begin{figure}
 
-{\centering \includegraphics[width=1\linewidth,height=0.5\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/cubble-operations/cubble-operations.001}
+{\centering \includegraphics[width=0.9\linewidth,height=0.5\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/cubble-operations/cubble-operations.001}
 
@@
 
-\caption{Cubble operations}\label{fig:cubble-operations}
+\caption[Cubble operations]{Cubble operations}\label{fig:cubble-operations}
 \end{figure}
+\end{CodeChunk}
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{aus\_climate }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{stretch}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{filter}\NormalTok{(}\FunctionTok{year}\NormalTok{(date) }\SpecialCharTok{==} \DecValTok{2020}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{tamp}\NormalTok{()}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## # Cubble: id-wise: nested form
-## # Key:    id [85]
-## # Leaves: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
-##    id            lat  long  elev name                   wmo_id ts               
-##    <chr>       <dbl> <dbl> <dbl> <chr>                   <dbl> <list>           
-##  1 ASN00009021 -31.9  116.  15.4 perth airport           94610 <tibble [366 x 4~
-##  2 ASN00010311 -31.9  117. 179   york                    94623 <tibble [366 x 4~
-##  3 ASN00010614 -32.9  117. 338   narrogin                94627 <tibble [366 x 4~
-##  4 ASN00014015 -12.4  131.  30.4 darwin airport          94120 <tibble [366 x 4~
-##  5 ASN00015131 -17.6  134. 220   elliott                 94236 <tibble [366 x 4~
-##  6 ASN00016065 -30.4  137.  76   andamooka               95660 <tibble [366 x 4~
-##  7 ASN00023034 -35.0  139.   2   adelaide airport        94672 <tibble [366 x 4~
-##  8 ASN00023373 -34.5  139. 275   nuriootpa viticultural  94681 <tibble [366 x 4~
-##  9 ASN00024024 -34.4  141.  30.1 loxton research centre  94682 <tibble [366 x 4~
-## 10 ASN00024048 -34.2  141.  31.5 renmark aero            95687 <tibble [366 x 4~
-## # ... with 75 more rows
-\end{verbatim}
+\begin{CodeChunk}
+\begin{CodeInput}
+R> aus_climate %>% 
++   stretch() %>% 
++   filter(year(date) == 2020) %>% 
++   tamp()
+\end{CodeInput}
+\begin{CodeOutput}
+# cubble:   id [85]: nested form
+# bbox:     [115.97, -43.32, 153.13, -12.42]- check gap on long
+# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
+   id            lat  long  elev name                   wmo_id ts               
+   <chr>       <dbl> <dbl> <dbl> <chr>                   <dbl> <list>           
+ 1 ASN00009021 -31.9  116.  15.4 perth airport           94610 <tibble [366 x 4~
+ 2 ASN00010311 -31.9  117. 179   york                    94623 <tibble [366 x 4~
+ 3 ASN00010614 -32.9  117. 338   narrogin                94627 <tibble [366 x 4~
+ 4 ASN00014015 -12.4  131.  30.4 darwin airport          94120 <tibble [366 x 4~
+ 5 ASN00015131 -17.6  134. 220   elliott                 94236 <tibble [366 x 4~
+ 6 ASN00016065 -30.4  137.  76   andamooka               95660 <tibble [366 x 4~
+ 7 ASN00023034 -35.0  139.   2   adelaide airport        94672 <tibble [366 x 4~
+ 8 ASN00023373 -34.5  139. 275   nuriootpa viticultural  94681 <tibble [366 x 4~
+ 9 ASN00024024 -34.4  141.  30.1 loxton research centre  94682 <tibble [366 x 4~
+10 ASN00024048 -34.2  141.  31.5 renmark aero            95687 <tibble [366 x 4~
+# ... with 75 more rows
+\end{CodeOutput}
+\end{CodeChunk}
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# join wmo\_id for each station}
-\CommentTok{\# to\_join \textless{}{-} station \%\textgreater{}\% select(id, wmo\_id)}
-\CommentTok{\# out \textless{}{-} climate\_small \%\textgreater{}\% }
-\CommentTok{\#   left\_join(to\_join, by = c("station" = "id")) \%\textgreater{}\% }
-\CommentTok{\#   filter(!is.na(wmo\_id))}
-\CommentTok{\# out}
-\end{Highlighting}
-\end{Shaded}
+\begin{CodeChunk}
+\begin{CodeInput}
+R> # join wmo_id for each station
+R> # to_join <- station %>% select(id, wmo_id)
+R> # out <- climate_small %>% 
+R> #   left_join(to_join, by = c("station" = "id")) %>% 
+R> #   filter(!is.na(wmo_id))
+R> # out
+\end{CodeInput}
+\end{CodeChunk}
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{aus\_climate }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{stretch}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{migrate}\NormalTok{(id, lat, long)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## # Cubble: time-wise: long form
-## # Key:    id [85]
-## # Leaves: id [chr], lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id
-## #   [dbl], prcp_missing [dbl], tmax_missing [dbl], tmin_missing [dbl]
-##    id          date        prcp  tmax  tmin   lat  long
-##    <chr>       <date>     <dbl> <dbl> <dbl> <dbl> <dbl>
-##  1 ASN00009021 2020-01-01     0   319   153 -31.9  116.
-##  2 ASN00009021 2020-01-02     0   249   164 -31.9  116.
-##  3 ASN00009021 2020-01-03     6   232   130 -31.9  116.
-##  4 ASN00009021 2020-01-04     0   284   124 -31.9  116.
-##  5 ASN00009021 2020-01-05     0   353   116 -31.9  116.
-##  6 ASN00009021 2020-01-06     0   348   131 -31.9  116.
-##  7 ASN00009021 2020-01-07     0   328   151 -31.9  116.
-##  8 ASN00009021 2020-01-08     0   304   174 -31.9  116.
-##  9 ASN00009021 2020-01-09     0   287   173 -31.9  116.
-## 10 ASN00009021 2020-01-10     0   326   158 -31.9  116.
-## # ... with 31,100 more rows
-\end{verbatim}
+\begin{CodeChunk}
+\begin{CodeInput}
+R> aus_climate %>% 
++   stretch() %>% 
++   migrate(id, lat, long)
+\end{CodeInput}
+\begin{CodeOutput}
+# cubble:  date, id [85]: long form
+# bbox:    [115.97, -43.32, 153.13, -12.42]- check gap on long
+# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]
+   id          date        prcp  tmax  tmin   lat  long
+   <chr>       <date>     <dbl> <dbl> <dbl> <dbl> <dbl>
+ 1 ASN00009021 2020-01-01     0  31.9  15.3 -31.9  116.
+ 2 ASN00009021 2020-01-02     0  24.9  16.4 -31.9  116.
+ 3 ASN00009021 2020-01-03     6  23.2  13   -31.9  116.
+ 4 ASN00009021 2020-01-04     0  28.4  12.4 -31.9  116.
+ 5 ASN00009021 2020-01-05     0  35.3  11.6 -31.9  116.
+ 6 ASN00009021 2020-01-06     0  34.8  13.1 -31.9  116.
+ 7 ASN00009021 2020-01-07     0  32.8  15.1 -31.9  116.
+ 8 ASN00009021 2020-01-08     0  30.4  17.4 -31.9  116.
+ 9 ASN00009021 2020-01-09     0  28.7  17.3 -31.9  116.
+10 ASN00009021 2020-01-10     0  32.6  15.8 -31.9  116.
+# ... with 31,100 more rows
+\end{CodeOutput}
+\end{CodeChunk}
 
@@
 
-\hypertarget{summary}{%
-\section*{Summary}\label{summary}}
-\addcontentsline{toc}{section}{Summary}
-
-\hypertarget{refs}{}
-\begin{CSLReferences}{1}{0}
-\leavevmode\hypertarget{ref-stars}{}%
-Pebesma, Edzer. 2021. \emph{Stars: Spatiotemporal Arrays, Raster and
-Vector Data Cubes}. \url{https://CRAN.R-project.org/package=stars}.
-
-\leavevmode\hypertarget{ref-pebesma2018simple}{}%
-Pebesma, Edzer J. 2018. {``Simple Features for r: Standardized Support
-for Spatial Vector Data.''} \emph{R J.} 10 (1): 439.
-
-\leavevmode\hypertarget{ref-tsibbles}{}%
-Wang, Earo, Dianne Cook, and Rob J Hyndman. 2020. {``A New Tidy Data
-Structure to Support Exploration and Modeling of Temporal Data.''}
-\emph{Journal of Computational and Graphical Statistics} 29 (3):
-466--78. \url{https://doi.org/10.1080/10618600.2019.1695624}.
-
-\end{CSLReferences}
-
-\bibliographystyle{unsrt}
+\renewcommand\refname{Summary}
 \bibliography{references.bib}

### 2021-11-22 main.Rmd
restructure the paper

@@
 
+**Spatio-temporal data + examples**  
 Spatio-temporal data record changes of variables in spatially separated regions across time.
-In this article, we consider spatio-temporal vector data, which are recorded in a fixed interval and are point based, characterised by longitude and latitude, in the spatial aspect.
-Examples of this type of data include the house price of a city or county, climate measures from weather stations in a country, and river level data from electronic gauges.
+In this article, we consider spatio-temporal vector data, which are recorded in a fixed interval.
+Examples of this type of data include the house price of a city or county, climate measures from weather stations in a country, and river level data from electronic gauges \textcolor{red}{more examples here}.
+\newline
 
+<!-- and are point based, characterised by longitude and latitude, in the spatial aspect -->
+
+**point to common feature: spatial level variables + time level variables relates to Table 13 in Tidy data paper**  
+Usually, this type of spatio-temporal data don't come in as a single table.
+Tidy data principle (...) prescribes each type of observational unit to form a table.
+This would suggest two tables to store the data, one for spatial-level and  one for temporal-level.
+The Table 13 in the tidy data paper present a structure like this and the author argues the lack of tools to work with relational data.
+\newline
+
+**My proposal**  
+Recent software development in R has proposed several relational data structure: `tidygraph`[@tidygraph] for graph manipulation, `dm` [@dm] for relational data model, while spatio-temporal data could benefit from having its own relational data structure.
+In this paper, we propose a tidy data structure for vector spaio-temporal data.
+\newline
+
+**Section division**  
+The rest of the paper will be divided as follows: Section 2 reviews the existing data structure for spatio, temporal, and spatio-temporal data.
+Section 3 presents a new data structure for spatio-temporal data: cubble.
+Then the paper introduces the workflow of data manipulation and visualisation with the cubble structure in Section 4.
+Section 5 gives some examples on how common spatial and temporal manipulations are performed with cubble and how static and interactive visualisation help to understand climate and [...] data.
+\newline
+
+\newpage
+
+# Existing data structure for spatio and temporal data
+
+**Existing packages**  
+Many data structures have been proposed for spatial and temporal data, but not many for spatio-temporal data.
+One of the reason could be the inherent different levels of information make it inefficient to store in the same table.
+`spacetime`[@spacetime] proposed four space-time layouts: Full grid (STF), sparse grid(STS), irregular (STI), and trajectory (STT) based on underlying spatial structure `sp` [@sp] and temporal structure `xts`[@xts].
+`spatstat` [@spatstat] implements a `ppp` class for point pattern data.
+More recent package `stars` [@stars] uses a spatiotemporal array to store the data and the array structure has its influence from `cubelyr` [@cubelyr], a dplyr data cube backend.
+\newline
+
+**Criteria of a data structure we want**
+
+With recent development in the R community, `sf` [@sf] and `tsibble`[@tsibble] have replaced `sp` and `xts` to be the convention structure for spatial and temporal data.
+One reason for their popularity is its integration with tidyverse ecosystem, making them intuitive and easy to adopt.
+For spatio-temporal data, we hope to build a data structure has the following features:
+
+  1) A data structure that handles spatial and temporal dimension in a relational structure. This derives from the 3rd tidy data principal.
+  2) An intuitive and easy to use interface that fits into the tidyverse ecosystem, and
+  3) Compatibility with the latest spatial and temporal data structure. This would give users the flexibility to use work from existing packages.
+
+\newpage
+# Workflow/ Data pipeline
+
+Now we introduce cubble in a data pipeline structure (review the concept of data pipeline? at least articulate its importance - automation, predictable of output)
+
+## Spatio-temporal data in the wild
+
+**Format of st in the wild**  
 Analysing this type of data requires less considerations on the geographical geometry type and map projection but more on how measures in these fixed locations changes across the time domain and whether these changes are related for adjacent locations.
@@
 trend and seasonality, 2) find the Waldo sites from the crowd, and 3) see how correlation of nearby sites changes across time.
+\newline
 
+**The main difficulty and challenge**  
 The main difficulty in visualising this type of data is to show information in both space and time dimension with the proper level of details without information overflow.
@@
 
-The rest of the paper will be divided as follows: Section 2 reviews the existing data structure for spatio, temporal, and spatio-temporal data.
-Section 3 presents a new data structure for spatio-temporal data: cubble.
-Then the paper introduces the workflow of data manipulation and visualisation with the cubble structure in Section 4.
-Section 5 gives some examples on how common spatial and temporal manipulations are performed with cubble and how static and interactive visualisation help to understand climate and [...] data.
+## Cubble
 
-# Existing data structure for spatio and temporal data
+**We form a cubble by defining abc... - Nested and long form**  
+When manipulating the spatial dimension it uses a nest form that:
 
-Below we review some structure for spatial, temporal, and spatio-temporal data.
+  - defines each group in a row,
+  - displays the group-related variables in columns, and
+  - nests all the time-related variables into a column called `ts`.
+
+When manipulating the temporal dimension, it uses the long form that:
+
+  - each combination of group and timestamp occupies a row
+  - time-related variables are displayed, and
+  - group-related variables are not explicitly displayed but can be accessed through the `meta` attribute.
+
+### stretch
+
+*By default build a nested cubble, stretch to long, create spatial attributes*
+
+### tamp
+
+*Tamp to nested, use the spatial attributes*
+
+### Support on hierarchical structure
+
+## Tidyverse compatibility
+
+*list supported tidyverse functions*
+
+## Others
+
+\newpage
+
+# Examples
+
+## Australia precipitation pattern in 2020
+
+Forming a cubble + basic tidyverse verbs - Vig 2
+Aggregation  - Vig 4
+
+## Matching precipitation and river level in Victria water gauges
+
+Matching  - Vig 3
+
+# Conclusion
+
+\newpage
+
+# Old stuff
 
@@
 This repetition is unnecessary and would inflate the object size for long series.
-
-\textcolor{red}{Under the third tidy data principle [@tidydata], these are two type of observational unit and should form different tables.
-An example (Table 8) is also given to demonstrate the billboard dataset where there are song variables and rank variables that changes each week.
-The other has mentioned the lack of tool to work directly with relational data and analysis will usually need to merge the dataset into one table, which could causes errors.
-These days more implementations have looked at relational data structure and in R, several good examples include, `tidygraph`[@tidygraph] for graph manipulation, `dm` [@dm] for relational data model.
-In the domain of spatio-temporal data, @spacetime has considered four types of spatio-temporal layout.
-`Spatstat` [@Spatstat] an implementation of point pattern.
-}
-
-\newpage
 
@@
 
-<!-- Below are the how the nested and long form look like for Australia climate data, which records daily precipitation, maximum and minimum temperature for 55 stations across Australia from 2015- 2020.
+Below are the how the nested and long form look like for Australia climate data, which records daily precipitation, maximum and minimum temperature for 55 stations across Australia from 2015- 2020.
 Notice that each station forms a group in both forms and specifically, the nested  and long form have a underlying `rowwise_df` and `grouped_df` respectively.
--->
 
-<!-- With a cubic framework on mind, different types of manipulation with cubble can be thought of as slicing the cube in various way.
+With a cubic framework on mind, different types of manipulation with cubble can be thought of as slicing the cube in various way.
 The table below shows how some `dplyr` verbs are mapped into the operation in a cubble.
 With the existing grouping on the station, additional groupping can be added with `group_by` and removed with `ungrouped`.
-[talk about why it is useful] -->
+[talk about why it is useful]
 
@@
 
-\newpage
-
 # Examples
@@
   - Local features: Any station stands out from the crowd?
-
-## Manipulation
-
-### Mutate and filter
-
-In the first example, we want to only keep the stations that have `tmax` recorded in 2020.
-This requires first narrow down the records to those in 2020, determine if `tmax` is missing for each station, and then retain those stations that have `tmax` recorded.
-The year filtering is an operation on the time axis, so we start with the long form.
-Whether each station has `tmax` recorded is a result of each station, rather than of each time point, hence we need to switch to the nested form with `tamp()`.
-To calculate whether `tmax` is recorded, we mutate a column `tmax_missing` that takes `TRUE` if all the `tmax` in the nested list column `ts` are `NA` and `FALSE` otherwise.
-To get the stations that we want, we need another filter on `tmax_missing`.
-
-```{r}
-aus_climate %>% 
-  stretch() %>% 
-  filter(year(date) == 2020) %>% 
-  tamp()
-```
-
-### Join
-
-Now we want to select the stations that have been registered with world meteorological organisation (WMO) and the dataset `station` has a column `wmo_id` that records this information.
-To do this task, we first need to join the `station` dataset with our climate dataset and then filter out those stations that don't have the WMO id.
-Since the join is by station rather than by time, we start with the nested form and write the exact same syntax of join and filter as with tidyverse.
-
-```{r}
-# join wmo_id for each station
-# to_join <- station %>% select(id, wmo_id)
-# out <- climate_small %>% 
-#   left_join(to_join, by = c("station" = "id")) %>% 
-#   filter(!is.na(wmo_id))
-# out
-```
-
-Sometimes, we would like to have station-wise and time-wise variables in the same form (i.e.
-when plotting glyph maps).
-This can also be seen as a joining task, on the long form, with the dataset to join being the metadata.
-`migrate()` is a verb introduced as the shortcut for `left_join()` with a cubble's metadata and below is the comparison of the two syntaxes.
-
-```{r}
-aus_climate %>% 
-  stretch() %>% 
-  migrate(id, lat, long)
-```
-
-  - data quality check: filter out stations have variables not properly recorded 
-  - data summary: 
-    - daily -> monthly/ weekly, 
-    - summarise by mean for tmax/ tmin, sum for prcp
-  -
-
-## Graphics
-
-Static + interactive -> tooltip to show additional information upon hovering
-
-  - Where are those stations on the map? 
-    - Mention mostly aero, airport, and lighthouse
-
-# Summary

### 2021-11-22 main.tex
restructure the paper

@@
 
+\textbf{Spatio-temporal data + examples}\\
 Spatio-temporal data record changes of variables in spatially separated
 regions across time. In this article, we consider spatio-temporal vector
-data, which are recorded in a fixed interval and are point based,
-characterised by longitude and latitude, in the spatial aspect. Examples
-of this type of data include the house price of a city or county,
-climate measures from weather stations in a country, and river level
-data from electronic gauges.
+data, which are recorded in a fixed interval. Examples of this type of
+data include the house price of a city or county, climate measures from
+weather stations in a country, and river level data from electronic
+gauges \textcolor{red}{more examples here}. \newline
 
+\textbf{point to common feature: spatial level variables + time level
+variables relates to Table 13 in Tidy data paper}\\
+Usually, this type of spatio-temporal data don't come in as a single
+table. Tidy data principle (\ldots) prescribes each type of
+observational unit to form a table. This would suggest two tables to
+store the data, one for spatial-level and one for temporal-level. The
+Table 13 in the tidy data paper present a structure like this and the
+author argues the lack of tools to work with relational data. \newline
+
+\textbf{My proposal}\\
+Recent software development in R has proposed several relational data
+structure: \texttt{tidygraph}\citep{tidygraph} for graph manipulation,
+\texttt{dm} \citep{dm} for relational data model, while spatio-temporal
+data could benefit from having its own relational data structure. In
+this paper, we propose a tidy data structure for vector spaio-temporal
+data. \newline
+
+\textbf{Section division}\\
+The rest of the paper will be divided as follows: Section 2 reviews the
+existing data structure for spatio, temporal, and spatio-temporal data.
+Section 3 presents a new data structure for spatio-temporal data:
+cubble. Then the paper introduces the workflow of data manipulation and
+visualisation with the cubble structure in Section 4. Section 5 gives
+some examples on how common spatial and temporal manipulations are
+performed with cubble and how static and interactive visualisation help
+to understand climate and {[}\ldots{]} data. \newline
+
+\newpage
+
+\hypertarget{existing-data-structure-for-spatio-and-temporal-data}{%
+\section{Existing data structure for spatio and temporal
+data}\label{existing-data-structure-for-spatio-and-temporal-data}}
+
+\textbf{Existing packages}\\
+Many data structures have been proposed for spatial and temporal data,
+but not many for spatio-temporal data. One of the reason could be the
+inherent different levels of information make it inefficient to store in
+the same table. \texttt{spacetime}\citep{spacetime} proposed four
+space-time layouts: Full grid (STF), sparse grid(STS), irregular (STI),
+and trajectory (STT) based on underlying spatial structure \texttt{sp}
+\citep{sp} and temporal structure \texttt{xts}\citep{xts}.
+\texttt{spatstat} \citep{spatstat} implements a \texttt{ppp} class for
+point pattern data. More recent package \texttt{stars} \citep{stars}
+uses a spatiotemporal array to store the data and the array structure
+has its influence from \texttt{cubelyr} \citep{cubelyr}, a dplyr data
+cube backend. \newline
+
+\textbf{Criteria of a data structure we want}
+
+With recent development in the R community, \texttt{sf} \citep{sf} and
+\texttt{tsibble}\citep{tsibble} have replaced \texttt{sp} and
+\texttt{xts} to be the convention structure for spatial and temporal
+data. One reason for their popularity is its integration with tidyverse
+ecosystem, making them intuitive and easy to adopt. For spatio-temporal
+data, we hope to build a data structure has the following features:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi})}
+\tightlist
+\item
+  A data structure that handles spatial and temporal dimension in a
+  relational structure. This derives from the 3rd tidy data principal.
+\item
+  An intuitive and easy to use interface that fits into the tidyverse
+  ecosystem, and
+\item
+  Compatibility with the latest spatial and temporal data structure.
+  This would give users the flexibility to use work from existing
+  packages.
+\end{enumerate}
+
+\newpage
+
+\hypertarget{workflow-data-pipeline}{%
+\section{Workflow/ Data pipeline}\label{workflow-data-pipeline}}
+
+Now we introduce cubble in a data pipeline structure (review the concept
+of data pipeline? at least articulate its importance - automation,
+predictable of output)
+
+\hypertarget{spatio-temporal-data-in-the-wild}{%
+\subsection{Spatio-temporal data in the
+wild}\label{spatio-temporal-data-in-the-wild}}
+
+\textbf{Format of st in the wild}\\
 Analysing this type of data requires less considerations on the
@@
 i.e.~trend and seasonality, 2) find the Waldo sites from the crowd, and
-3) see how correlation of nearby sites changes across time.
+3) see how correlation of nearby sites changes across time. \newline
 
+\textbf{The main difficulty and challenge}\\
 The main difficulty in visualising this type of data is to show
@@
 
-The rest of the paper will be divided as follows: Section 2 reviews the
-existing data structure for spatio, temporal, and spatio-temporal data.
-Section 3 presents a new data structure for spatio-temporal data:
-cubble. Then the paper introduces the workflow of data manipulation and
-visualisation with the cubble structure in Section 4. Section 5 gives
-some examples on how common spatial and temporal manipulations are
-performed with cubble and how static and interactive visualisation help
-to understand climate and {[}\ldots{]} data.
+\hypertarget{cubble}{%
+\subsection{Cubble}\label{cubble}}
 
-\hypertarget{existing-data-structure-for-spatio-and-temporal-data}{%
-\section{Existing data structure for spatio and temporal
-data}\label{existing-data-structure-for-spatio-and-temporal-data}}
+\textbf{We form a cubble by defining abc\ldots{} - Nested and long
+form}\\
+When manipulating the spatial dimension it uses a nest form that:
 
-Below we review some structure for spatial, temporal, and
-spatio-temporal data.
+\begin{itemize}
+\tightlist
+\item
+  defines each group in a row,
+\item
+  displays the group-related variables in columns, and
+\item
+  nests all the time-related variables into a column called \texttt{ts}.
+\end{itemize}
+
+When manipulating the temporal dimension, it uses the long form that:
+
+\begin{itemize}
+\tightlist
+\item
+  each combination of group and timestamp occupies a row
+\item
+  time-related variables are displayed, and
+\item
+  group-related variables are not explicitly displayed but can be
+  accessed through the \texttt{meta} attribute.
+\end{itemize}
+
+\hypertarget{stretch}{%
+\subsubsection{stretch}\label{stretch}}
+
+\emph{By default build a nested cubble, stretch to long, create spatial
+attributes}
+
+\hypertarget{tamp}{%
+\subsubsection{tamp}\label{tamp}}
+
+\emph{Tamp to nested, use the spatial attributes}
+
+\hypertarget{support-on-hierarchical-structure}{%
+\subsubsection{Support on hierarchical
+structure}\label{support-on-hierarchical-structure}}
+
+\hypertarget{tidyverse-compatibility}{%
+\subsection{Tidyverse compatibility}\label{tidyverse-compatibility}}
+
+\emph{list supported tidyverse functions}
+
+\hypertarget{others}{%
+\subsection{Others}\label{others}}
+
+\newpage
+
+\hypertarget{examples}{%
+\section{Examples}\label{examples}}
+
+\hypertarget{australia-precipitation-pattern-in-2020}{%
+\subsection{Australia precipitation pattern in
+2020}\label{australia-precipitation-pattern-in-2020}}
+
+Forming a cubble + basic tidyverse verbs - Vig 2 Aggregation - Vig 4
+
+\hypertarget{matching-precipitation-and-river-level-in-victria-water-gauges}{%
+\subsection{Matching precipitation and river level in Victria water
+gauges}\label{matching-precipitation-and-river-level-in-victria-water-gauges}}
+
+Matching - Vig 3
+
+\hypertarget{conclusion}{%
+\section{Conclusion}\label{conclusion}}
+
+\newpage
+
+\hypertarget{old-stuff}{%
+\section{Old stuff}\label{old-stuff}}
 
@@
 object size for long series.
-
-\textcolor{red}{Under the third tidy data principle [@tidydata], these are two type of observational unit and should form different tables.
-An example (Table 8) is also given to demonstrate the billboard dataset where there are song variables and rank variables that changes each week.
-The other has mentioned the lack of tool to work directly with relational data and analysis will usually need to merge the dataset into one table, which could causes errors.
-These days more implementations have looked at relational data structure and in R, several good examples include, `tidygraph`[@tidygraph] for graph manipulation, `dm` [@dm] for relational data model.
-In the domain of spatio-temporal data, @spacetime has considered four types of spatio-temporal layout.
-`Spatstat` [@Spatstat] an implementation of point pattern.
-}
-
-\newpage
 
@@
 
+Below are the how the nested and long form look like for Australia
+climate data, which records daily precipitation, maximum and minimum
+temperature for 55 stations across Australia from 2015- 2020. Notice
+that each station forms a group in both forms and specifically, the
+nested and long form have a underlying \texttt{rowwise\_df} and
+\texttt{grouped\_df} respectively.
+
+With a cubic framework on mind, different types of manipulation with
+cubble can be thought of as slicing the cube in various way. The table
+below shows how some \texttt{dplyr} verbs are mapped into the operation
+in a cubble. With the existing grouping on the station, additional
+groupping can be added with \texttt{group\_by} and removed with
+\texttt{ungrouped}. {[}talk about why it is useful{]}
+
 \newpage
@@
 
-\newpage
-
-\hypertarget{examples}{%
-\section{Examples}\label{examples}}
+\hypertarget{examples-1}{%
+\section{Examples}\label{examples-1}}
 
@@
 
-\hypertarget{manipulation}{%
-\subsection{Manipulation}\label{manipulation}}
-
-\hypertarget{mutate-and-filter}{%
-\subsubsection{Mutate and filter}\label{mutate-and-filter}}
-
-In the first example, we want to only keep the stations that have
-\texttt{tmax} recorded in 2020. This requires first narrow down the
-records to those in 2020, determine if \texttt{tmax} is missing for each
-station, and then retain those stations that have \texttt{tmax}
-recorded. The year filtering is an operation on the time axis, so we
-start with the long form. Whether each station has \texttt{tmax}
-recorded is a result of each station, rather than of each time point,
-hence we need to switch to the nested form with \texttt{tamp()}. To
-calculate whether \texttt{tmax} is recorded, we mutate a column
-\texttt{tmax\_missing} that takes \texttt{TRUE} if all the \texttt{tmax}
-in the nested list column \texttt{ts} are \texttt{NA} and \texttt{FALSE}
-otherwise. To get the stations that we want, we need another filter on
-\texttt{tmax\_missing}.
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> aus_climate %>% 
-+   stretch() %>% 
-+   filter(year(date) == 2020) %>% 
-+   tamp()
-\end{CodeInput}
-\begin{CodeOutput}
-# cubble:   id [85]: nested form
-# bbox:     [115.97, -43.32, 153.13, -12.42]- check gap on long
-# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
-   id            lat  long  elev name                   wmo_id ts               
-   <chr>       <dbl> <dbl> <dbl> <chr>                   <dbl> <list>           
- 1 ASN00009021 -31.9  116.  15.4 perth airport           94610 <tibble [366 x 4~
- 2 ASN00010311 -31.9  117. 179   york                    94623 <tibble [366 x 4~
- 3 ASN00010614 -32.9  117. 338   narrogin                94627 <tibble [366 x 4~
- 4 ASN00014015 -12.4  131.  30.4 darwin airport          94120 <tibble [366 x 4~
- 5 ASN00015131 -17.6  134. 220   elliott                 94236 <tibble [366 x 4~
- 6 ASN00016065 -30.4  137.  76   andamooka               95660 <tibble [366 x 4~
- 7 ASN00023034 -35.0  139.   2   adelaide airport        94672 <tibble [366 x 4~
- 8 ASN00023373 -34.5  139. 275   nuriootpa viticultural  94681 <tibble [366 x 4~
- 9 ASN00024024 -34.4  141.  30.1 loxton research centre  94682 <tibble [366 x 4~
-10 ASN00024048 -34.2  141.  31.5 renmark aero            95687 <tibble [366 x 4~
-# ... with 75 more rows
-\end{CodeOutput}
-\end{CodeChunk}
-
-\hypertarget{join}{%
-\subsubsection{Join}\label{join}}
-
-Now we want to select the stations that have been registered with world
-meteorological organisation (WMO) and the dataset \texttt{station} has a
-column \texttt{wmo\_id} that records this information. To do this task,
-we first need to join the \texttt{station} dataset with our climate
-dataset and then filter out those stations that don't have the WMO id.
-Since the join is by station rather than by time, we start with the
-nested form and write the exact same syntax of join and filter as with
-tidyverse.
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> # join wmo_id for each station
-R> # to_join <- station %>% select(id, wmo_id)
-R> # out <- climate_small %>% 
-R> #   left_join(to_join, by = c("station" = "id")) %>% 
-R> #   filter(!is.na(wmo_id))
-R> # out
-\end{CodeInput}
-\end{CodeChunk}
-
-Sometimes, we would like to have station-wise and time-wise variables in
-the same form (i.e.~when plotting glyph maps). This can also be seen as
-a joining task, on the long form, with the dataset to join being the
-metadata. \texttt{migrate()} is a verb introduced as the shortcut for
-\texttt{left\_join()} with a cubble's metadata and below is the
-comparison of the two syntaxes.
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> aus_climate %>% 
-+   stretch() %>% 
-+   migrate(id, lat, long)
-\end{CodeInput}
-\begin{CodeOutput}
-# cubble:  date, id [85]: long form
-# bbox:    [115.97, -43.32, 153.13, -12.42]- check gap on long
-# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]
-   id          date        prcp  tmax  tmin   lat  long
-   <chr>       <date>     <dbl> <dbl> <dbl> <dbl> <dbl>
- 1 ASN00009021 2020-01-01     0  31.9  15.3 -31.9  116.
- 2 ASN00009021 2020-01-02     0  24.9  16.4 -31.9  116.
- 3 ASN00009021 2020-01-03     6  23.2  13   -31.9  116.
- 4 ASN00009021 2020-01-04     0  28.4  12.4 -31.9  116.
- 5 ASN00009021 2020-01-05     0  35.3  11.6 -31.9  116.
- 6 ASN00009021 2020-01-06     0  34.8  13.1 -31.9  116.
- 7 ASN00009021 2020-01-07     0  32.8  15.1 -31.9  116.
- 8 ASN00009021 2020-01-08     0  30.4  17.4 -31.9  116.
- 9 ASN00009021 2020-01-09     0  28.7  17.3 -31.9  116.
-10 ASN00009021 2020-01-10     0  32.6  15.8 -31.9  116.
-# ... with 31,100 more rows
-\end{CodeOutput}
-\end{CodeChunk}
-
-\begin{itemize}
-\tightlist
-\item
-  data quality check: filter out stations have variables not properly
-  recorded
-\item
-  data summary:
-
-  \begin{itemize}
-  \tightlist
-  \item
-    daily -\textgreater{} monthly/ weekly,
-  \item
-    summarise by mean for tmax/ tmin, sum for prcp
-  \end{itemize}
-\item
-\end{itemize}
-
-\hypertarget{graphics}{%
-\subsection{Graphics}\label{graphics}}
-
-Static + interactive -\textgreater{} tooltip to show additional
-information upon hovering
-
-\begin{itemize}
-\tightlist
-\item
-  Where are those stations on the map?
-
-  \begin{itemize}
-  \tightlist
-  \item
-    Mention mostly aero, airport, and lighthouse
-  \end{itemize}
-\end{itemize}
-
-\renewcommand\refname{Summary}
 \bibliography{references.bib}

### 2021-11-23 main.Rmd
add more content

@@
 **point to common feature: spatial level variables + time level variables relates to Table 13 in Tidy data paper**  
-Usually, this type of spatio-temporal data don't come in as a single table.
-Tidy data principle (...) prescribes each type of observational unit to form a table.
+Usually, this type of spatio-temporal data does not necessarily come in as a single table.
+Tidy data principle [@tidydata] prescribes each type of observational unit to form a table.
 This would suggest two tables to store the data, one for spatial-level and one for temporal-level.
-The Table 13 in the tidy data paper present a structure like this and the author argues the lack of tools to work with relational data.
+The Table 13 in @tidydata presents a structure like this and the author argues the lack of tools to work with relational data like such.
 \newline
@@
 **My proposal**  
-Recent software development in R has proposed several relational data structure: `tidygraph`[@tidygraph] for graph manipulation, `dm` [@dm] for relational data model, while spatio-temporal data could benefit from having its own relational data structure.
+Recent software development in R has proposed several relational data structure: `tidygraph`[@tidygraph] for graph manipulation, `dm` [@dm] for relational data model, while spatio-temporal data could benefit from having its own relational data structure that identifies the essential spati0-temporal element: site identifier, time identifier, geographic coordinates.
 In this paper, we propose a tidy data structure for vector spaio-temporal data.
@@
 
-# Existing data structure for spatio and temporal data
+# Existing data structure for spatio and temporal data {#lit-review}
 
@@
   2) An intuitive and easy to use interface that fits into the tidyverse ecosystem, and
-  3) Compatibility with the latest spatial and temporal data structure. This would give users the flexibility to use work from existing packages.
+  3) Compatibility with the latest spatial and temporal data structure. This would give users the flexibility to use the work in existing packages.
 
 \newpage
-# Workflow/ Data pipeline
-
-Now we introduce cubble in a data pipeline structure (review the concept of data pipeline? at least articulate its importance - automation, predictable of output)
+# Cubble data structure
 
@@
 
-**Format of st in the wild**  
-Analysing this type of data requires less considerations on the geographical geometry type and map projection but more on how measures in these fixed locations changes across the time domain and whether these changes are related for adjacent locations.
-For example, when nearby areas show patterns that are regular enough, visualising spatio-temporal data can 1) discover regional time series features, i.e.
-trend and seasonality, 2) find the Waldo sites from the crowd, and 3) see how correlation of nearby sites changes across time.
-\newline
+Spatio-temporal data don't usually come to the analysts in various forms.
+Figure \ref{fig:cubble-diagram} groups them into four categories.
+Despite whichever form the data is in, all the forms have a group identifier, `key`, a time column, `t`, and variables that either various across group, i.e.
+`lon`, `lat`, $C_s$, or across time, i.e.
+$V_t$.
+These identifiers will be the building blocks for the data structure introduced below.
 
-**The main difficulty and challenge**  
-The main difficulty in visualising this type of data is to show information in both space and time dimension with the proper level of details without information overflow.
-This would sometimes require aggregating the time dimension into the proper level or slicing the data into a reasonable number of subset for display.
-In this sense, a data structure that regulates the manipulation spatio-temporal data will benefit the analysis workflow.
-While many implementations focus on manipulating and visualising pure spatial or temporal data, there are not sufficient tools to deal with spatio-temporal data.
-The purpose of this paper is to introduce a spatio-temporal vector data structure for data analysis in R.
+```{r cubble-diagram, echo = FALSE, fig.cap="Illustration of incoming data formats for spatio-temporal data.
+(1) Data comes in as a single table; (2) Separate tables for spatial and temporal variables; (3) A single table with all the parameters used to query the database and a separate table for queried data; and (4) Cubical data in array or NetCDF format.", fig.align='center', out.height = "15%"}
+knitr::include_graphics(here::here("figures/input-formats/input-formats.001.png"))
+```
 
-## Cubble
+## A new data structure: cubble
 
-**We form a cubble by defining abc... - Nested and long form**  
-When manipulating the spatial dimension it uses a nest form that:
+A cubble, in essence, wires the spatial and temporal dimension into a single object while provides two forms for manipulation each dimension separately.
+
+In the nest form, Figure \ref{fig:def-cubble}a), a cubble:
 
@@
 
-When manipulating the temporal dimension, it uses the long form that:
+In the long form, Figure \ref{fig:def-cubble}b), a cubble:
 
-  - each combination of group and timestamp occupies a row
-  - time-related variables are displayed, and
-  - group-related variables are not explicitly displayed but can be accessed through the `meta` attribute.
+  - defines the combination of group and timestamp as a row
+  - displays time-related variables, along with the group identifier in columns, and
+  - stores group-related variables as a attributes called `spatial`.
 
-### stretch
+```{r def-cubble, echo = FALSE, fig.align="center", out.height="30%", fig.cap = "Illustration of nested and long cubble."}
+knitr::include_graphics(here::here("figures/def-cubble/def-cubble.001.png"))
+```
 
-*By default build a nested cubble, stretch to long, create spatial attributes*
+\newpage
+## Create a cubble in the nested form
 
-### tamp
+To create a cubble, you can use `as_cubble` with proper supply of `key`, `index`, and `coords`.
+[`key` defines whether a variable is spatial or temperal by looking at whether it varies across `key`.
+For example, `lat` is a spatial variable since for one `id` there's only one `lat` value.
+`index` and `coords` prescribes the temporal index and spaital coordinates used in the data.
+These are common variables that should present in the point pattern spatio-temporal data.] The cubble created by default is in the nested form.
+\newline
 
-*Tamp to nested, use the spatial attributes*
+```{r}
+(cubble_nested <- climate_flat %>% 
+  as_cubble(key = id, index = date, coords = c("long", "lat")))
+```
 
-### Support on hierarchical structure
+In the `cubble` header, you can read the key variable + number, bbox, and also the name of variable nested in the `ts` column.
+Here in this example, the temporal variables are precipitation, `prcp`, maximum temporature, `tmax`, and minimum temperature, `tmin`.
 
-## Tidyverse compatibility
+In this structure, the nested form is suitable for manipulating the spatial dimension of the data and this would include manipulating variables that 1) are inherently spatial and 2) summarise the temporal variables by `key`.
+Under the hood, the nested cubble is built on top of the rowwise dataframe (`rowwise_df`).
+This is design to simplify the code when working with the temporal variables - they are nested into the list-column `ts`.
 
-*list supported tidyverse functions*
+## Stretch a nested cubble into the long form
 
-## Others
+The verb `stretch()` switch the cubble from the nested form into a long form.
+Under the hood, it first extracts the spatial variables into a separate tibble to store in the attribute `spatial` and then unnests the `ts` column to show the temporal content:
+
+```{r}
+(cubble_long <- cubble_nested %>% stretch(ts))
+```
+
+Notice here that the third line in the header is changed to reflect the spatial variables stored. This is a format suitable for computing time-wise variables.
+
+## Tamp a long cubble back to the nested form
+
+Manipulation on the spatial and temporal dimension can be an iterative process.
+Many times, we may decide to go back to the nested form after some temporal manipulation.
+The verb to switch a long cubble back to the nested form is `tamp()`:
+
+```{r}
+(cubble_back <- cubble_long %>% tamp())
+```
+
+## Migrate spatial variables to a long cubble
+
+As an output to be supplied to further visualisation or modelling, analysts would usually like the spatial and temporal variables to be in the same table.
+`migrate()` moves the spatial variables in the attribute `spatial` into the long form cubble.
+
+```{r}
+(cubble_long %>% migrate(long, lat))
+```
+
+## Support on hierarchical structure
+
+`switch_key()`
+
+## Integrating into a tidy workflow
+
+Building from an underlying `tbl_df` structure, it is natural to implement methods available in `dplyr` to `cubble`.
+Supported methods in the `cubble` with `dplyr` generics includes: `mutate`, `filter`, `summarise`, `select`, `arrange`, `rename`, `left_join`, `group_by`, `ungroup`, and the slice family (`slice_head`, `slice_tail`, `slice_sample`, `slice_min` and `slice_max`) \textcolor{red}{into a table}.
+
+`cubble` is also compatible with `tsibble` in the sense that the original list-column can be a `tbl_ts` object.
+Duplicates and gaps shoudl be first checked before structuring the data into a cubble.
+If the input data is a `tsibble` object, the long form cubble is also a `tsibble` where users can directly apply time series operations.
 
@@
 
-Spatio-temporal data don't usually come to the analysts as a whole piece.
-A way to look at these data is to divide it into spatial and temporal dimension with an ID that links between the two.
-The first row in Figure \ref{fig:cubble-diagram} illustrates this representation where in the spatial dimension, the data is characterised by `id`, `lat`, `long`.
-$V_s$ in the last column represents all the other site-wise variables, for example, elevation and full name etc.
-The temporal dimension, on the other hand, can be characterised by `id` and `t` with $V_t$ representing all the time-wise variables.
-In climate data, this could include precipitation, maximum or minimum temperature, and wind speed etc.
-
-```{r cubble-diagram, echo = FALSE, fig.cap="Cubble diagram", out.height="40%", fig.align='center'}
-knitr::include_graphics(here::here("figures/input-formats/input-formats.001.png"))
-```
+**The main difficulty and challenge**  
+The main difficulty in visualising this type of data is to show information in both space and time dimension with the proper level of details without information overflow.
+This would sometimes require aggregating the time dimension into the proper level or slicing the data into a reasonable number of subset for display.
+In this sense, a data structure that regulates the manipulation spatio-temporal data will benefit the analysis workflow.
+While many implementations focus on manipulating and visualising pure spatial or temporal data, there are not sufficient tools to deal with spatio-temporal data.
+The purpose of this paper is to introduce a spatio-temporal vector data structure for data analysis in R.
 
@@
 Also, recordings of the site ID from different data sources can be slightly different from each other, causing a painful checking and cleaning of site IDs before the join.
-
-A cubble, in essence, wires both dimensions in the spatio-temporal data into one object while provide two forms for manipulation the spatial and temporal dimension separately.
-
-When manipulating the spatial dimension it uses a nest form that:
-
-  - defines each group in a row,
-  - displays the group-related variables in columns, and
-  - nests all the time-related variables into a column called `ts`.
-
-When manipulating the temporal dimension, it uses the long form that:
-
-  - each combination of group and timestamp occupies a row
-  - time-related variables are displayed, and
-  - group-related variables are not explicitly displayed but can be accessed through the `meta` attribute.
 

### 2021-11-23 main.tex
add more content

@@
 variables relates to Table 13 in Tidy data paper}\\
-Usually, this type of spatio-temporal data don't come in as a single
-table. Tidy data principle (\ldots) prescribes each type of
-observational unit to form a table. This would suggest two tables to
-store the data, one for spatial-level and one for temporal-level. The
-Table 13 in the tidy data paper present a structure like this and the
-author argues the lack of tools to work with relational data. \newline
+Usually, this type of spatio-temporal data does not necessarily come in
+as a single table. Tidy data principle \citep{tidydata} prescribes each
+type of observational unit to form a table. This would suggest two
+tables to store the data, one for spatial-level and one for
+temporal-level. The Table 13 in \citet{tidydata} presents a structure
+like this and the author argues the lack of tools to work with
+relational data like such. \newline
 
@@
 \texttt{dm} \citep{dm} for relational data model, while spatio-temporal
-data could benefit from having its own relational data structure. In
-this paper, we propose a tidy data structure for vector spaio-temporal
-data. \newline
+data could benefit from having its own relational data structure that
+identifies the essential spati0-temporal element: site identifier, time
+identifier, geographic coordinates. In this paper, we propose a tidy
+data structure for vector spaio-temporal data. \newline
 
@@
 
-\hypertarget{existing-data-structure-for-spatio-and-temporal-data}{%
+\hypertarget{lit-review}{%
 \section{Existing data structure for spatio and temporal
-data}\label{existing-data-structure-for-spatio-and-temporal-data}}
+data}\label{lit-review}}
 
@@
   Compatibility with the latest spatial and temporal data structure.
-  This would give users the flexibility to use work from existing
+  This would give users the flexibility to use the work in existing
   packages.
@@
 
-Now we introduce cubble in a data pipeline structure (review the concept
-of data pipeline? at least articulate its importance - automation,
-predictable of output)
-
 \hypertarget{spatio-temporal-data-in-the-wild}{%
@@
 
-\textbf{Format of st in the wild}\\
-Analysing this type of data requires less considerations on the
-geographical geometry type and map projection but more on how measures
-in these fixed locations changes across the time domain and whether
-these changes are related for adjacent locations. For example, when
-nearby areas show patterns that are regular enough, visualising
-spatio-temporal data can 1) discover regional time series features,
-i.e.~trend and seasonality, 2) find the Waldo sites from the crowd, and
-3) see how correlation of nearby sites changes across time. \newline
+Spatio-temporal data don't usually come to the analysts in various
+forms. Figure \ref{fig:cubble-diagram} groups them into four categories.
+Despite whichever form the data is in, all the forms have a group
+identifier, \texttt{key}, a time column, \texttt{t}, and variables that
+either various across group, i.e.~\texttt{lon}, \texttt{lat}, \(C_s\),
+or across time, i.e.~\(V_t\). These identifiers will be the building
+blocks for the data structure introduced below.
 
-\textbf{The main difficulty and challenge}\\
-The main difficulty in visualising this type of data is to show
-information in both space and time dimension with the proper level of
-details without information overflow. This would sometimes require
-aggregating the time dimension into the proper level or slicing the data
-into a reasonable number of subset for display. In this sense, a data
-structure that regulates the manipulation spatio-temporal data will
-benefit the analysis workflow. While many implementations focus on
-manipulating and visualising pure spatial or temporal data, there are
-not sufficient tools to deal with spatio-temporal data. The purpose of
-this paper is to introduce a spatio-temporal vector data structure for
-data analysis in R.
+\begin{CodeChunk}
+\begin{figure}
 
-\hypertarget{cubble}{%
-\subsection{Cubble}\label{cubble}}
+{\centering \includegraphics[width=1\linewidth,height=0.15\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/input-formats/input-formats.001}
 
-\textbf{We form a cubble by defining abc\ldots{} - Nested and long
-form}\\
-When manipulating the spatial dimension it uses a nest form that:
+}
+
+\caption[Illustration of incoming data formats for spatio-temporal data]{Illustration of incoming data formats for spatio-temporal data.
+(1) Data comes in as a single table; (2) Separate tables for spatial and temporal variables; (3) A single table with all the parameters used to query the database and a separate table for queried data; and (4) Cubical data in array or NetCDF format.}\label{fig:cubble-diagram}
+\end{figure}
+\end{CodeChunk}
+
+\hypertarget{a-new-data-structure-cubble}{%
+\subsection{A new data structure:
+cubble}\label{a-new-data-structure-cubble}}
+
+A cubble, in essence, wires the spatial and temporal dimension into a
+single object while provides two forms for manipulation each dimension
+separately.
+
+In the nest form, Figure \ref{fig:def-cubble}a), a cubble:
 
@@
 
-When manipulating the temporal dimension, it uses the long form that:
+In the long form, Figure \ref{fig:def-cubble}b), a cubble:
 
@@
 \item
-  each combination of group and timestamp occupies a row
+  defines the combination of group and timestamp as a row
 \item
-  time-related variables are displayed, and
+  displays time-related variables, along with the group identifier in
+  columns, and
 \item
-  group-related variables are not explicitly displayed but can be
-  accessed through the \texttt{meta} attribute.
+  stores group-related variables as a attributes called
+  \texttt{spatial}.
 \end{itemize}
 
-\hypertarget{stretch}{%
-\subsubsection{stretch}\label{stretch}}
+\begin{CodeChunk}
+\begin{figure}
 
-\emph{By default build a nested cubble, stretch to long, create spatial
-attributes}
+{\centering \includegraphics[width=1\linewidth,height=0.3\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/def-cubble/def-cubble.001}
 
-\hypertarget{tamp}{%
-\subsubsection{tamp}\label{tamp}}
+}
 
-\emph{Tamp to nested, use the spatial attributes}
+\caption[Illustration of nested and long cubble]{Illustration of nested and long cubble.}\label{fig:def-cubble}
+\end{figure}
+\end{CodeChunk}
+
+\newpage
+
+\hypertarget{create-a-cubble-in-the-nested-form}{%
+\subsection{Create a cubble in the nested
+form}\label{create-a-cubble-in-the-nested-form}}
+
+To create a cubble, you can use \texttt{as\_cubble} with proper supply
+of \texttt{key}, \texttt{index}, and \texttt{coords}. {[}\texttt{key}
+defines whether a variable is spatial or temperal by looking at whether
+it varies across \texttt{key}. For example, \texttt{lat} is a spatial
+variable since for one \texttt{id} there's only one \texttt{lat} value.
+\texttt{index} and \texttt{coords} prescribes the temporal index and
+spaital coordinates used in the data. These are common variables that
+should present in the point pattern spatio-temporal data.{]} The cubble
+created by default is in the nested form. \newline
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> (cubble_nested <- climate_flat %>% 
++   as_cubble(key = id, index = date, coords = c("long", "lat")))
+\end{CodeInput}
+\begin{CodeOutput}
+# cubble:   id [5]: nested form
+# bbox:     [115.97, -32.94, 133.55, -12.42]- check gap on long and lat
+# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
+  id            lat  long  elev name           wmo_id ts                
+  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>            
+1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tibble [366 x 4]>
+2 ASN00010311 -31.9  117. 179   york            94623 <tibble [366 x 4]>
+3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble [366 x 4]>
+4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble [366 x 4]>
+5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble [366 x 4]>
+\end{CodeOutput}
+\end{CodeChunk}
+
+In the \texttt{cubble} header, you can read the key variable + number,
+bbox, and also the name of variable nested in the \texttt{ts} column.
+Here in this example, the temporal variables are precipitation,
+\texttt{prcp}, maximum temporature, \texttt{tmax}, and minimum
+temperature, \texttt{tmin}.
+
+In this structure, the nested form is suitable for manipulating the
+spatial dimension of the data and this would include manipulating
+variables that 1) are inherently spatial and 2) summarise the temporal
+variables by \texttt{key}. Under the hood, the nested cubble is built on
+top of the rowwise dataframe (\texttt{rowwise\_df}). This is design to
+simplify the code when working with the temporal variables - they are
+nested into the list-column \texttt{ts}.
+
+\hypertarget{stretch-a-nested-cubble-into-the-long-form}{%
+\subsection{Stretch a nested cubble into the long
+form}\label{stretch-a-nested-cubble-into-the-long-form}}
+
+The verb \texttt{stretch()} switch the cubble from the nested form into
+a long form. Under the hood, it first extracts the spatial variables
+into a separate tibble to store in the attribute \texttt{spatial} and
+then unnests the \texttt{ts} column to show the temporal content:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> (cubble_long <- cubble_nested %>% stretch(ts))
+\end{CodeInput}
+\begin{CodeOutput}
+# cubble:  date, id [5]: long form
+# bbox:    [115.97, -32.94, 133.55, -12.42]- check gap on long and lat
+# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]
+   id          date        prcp  tmax  tmin
+   <chr>       <date>     <dbl> <dbl> <dbl>
+ 1 ASN00009021 2020-01-01     0  31.9  15.3
+ 2 ASN00009021 2020-01-02     0  24.9  16.4
+ 3 ASN00009021 2020-01-03     6  23.2  13  
+ 4 ASN00009021 2020-01-04     0  28.4  12.4
+ 5 ASN00009021 2020-01-05     0  35.3  11.6
+ 6 ASN00009021 2020-01-06     0  34.8  13.1
+ 7 ASN00009021 2020-01-07     0  32.8  15.1
+ 8 ASN00009021 2020-01-08     0  30.4  17.4
+ 9 ASN00009021 2020-01-09     0  28.7  17.3
+10 ASN00009021 2020-01-10     0  32.6  15.8
+# ... with 1,820 more rows
+\end{CodeOutput}
+\end{CodeChunk}
+
+Notice here that the third line in the header is changed to reflect the
+spatial variables stored. This is a format suitable for computing
+time-wise variables.
+
+\hypertarget{tamp-a-long-cubble-back-to-the-nested-form}{%
+\subsection{Tamp a long cubble back to the nested
+form}\label{tamp-a-long-cubble-back-to-the-nested-form}}
+
+Manipulation on the spatial and temporal dimension can be an iterative
+process. Many times, we may decide to go back to the nested form after
+some temporal manipulation. The verb to switch a long cubble back to the
+nested form is \texttt{tamp()}:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> (cubble_back <- cubble_long %>% tamp())
+\end{CodeInput}
+\begin{CodeOutput}
+# cubble:   id [5]: nested form
+# bbox:     [115.97, -32.94, 133.55, -12.42]- check gap on long and lat
+# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
+  id            lat  long  elev name           wmo_id ts                
+  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>            
+1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tibble [366 x 4]>
+2 ASN00010311 -31.9  117. 179   york            94623 <tibble [366 x 4]>
+3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble [366 x 4]>
+4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble [366 x 4]>
+5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble [366 x 4]>
+\end{CodeOutput}
+\end{CodeChunk}
+
+\hypertarget{migrate-spatial-variables-to-a-long-cubble}{%
+\subsection{Migrate spatial variables to a long
+cubble}\label{migrate-spatial-variables-to-a-long-cubble}}
+
+As an output to be supplied to further visualisation or modelling,
+analysts would usually like the spatial and temporal variables to be in
+the same table. \texttt{migrate()} moves the spatial variables in the
+attribute \texttt{spatial} into the long form cubble.
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> (cubble_long %>% migrate(long, lat))
+\end{CodeInput}
+\begin{CodeOutput}
+# cubble:  date, id [5]: long form
+# bbox:    [115.97, -32.94, 133.55, -12.42]- check gap on long and lat
+# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]
+   id          date        prcp  tmax  tmin  long   lat
+   <chr>       <date>     <dbl> <dbl> <dbl> <dbl> <dbl>
+ 1 ASN00009021 2020-01-01     0  31.9  15.3  116. -31.9
+ 2 ASN00009021 2020-01-02     0  24.9  16.4  116. -31.9
+ 3 ASN00009021 2020-01-03     6  23.2  13    116. -31.9
+ 4 ASN00009021 2020-01-04     0  28.4  12.4  116. -31.9
+ 5 ASN00009021 2020-01-05     0  35.3  11.6  116. -31.9
+ 6 ASN00009021 2020-01-06     0  34.8  13.1  116. -31.9
+ 7 ASN00009021 2020-01-07     0  32.8  15.1  116. -31.9
+ 8 ASN00009021 2020-01-08     0  30.4  17.4  116. -31.9
+ 9 ASN00009021 2020-01-09     0  28.7  17.3  116. -31.9
+10 ASN00009021 2020-01-10     0  32.6  15.8  116. -31.9
+# ... with 1,820 more rows
+\end{CodeOutput}
+\end{CodeChunk}
 
 \hypertarget{support-on-hierarchical-structure}{%
-\subsubsection{Support on hierarchical
+\subsection{Support on hierarchical
 structure}\label{support-on-hierarchical-structure}}
 
-\hypertarget{tidyverse-compatibility}{%
-\subsection{Tidyverse compatibility}\label{tidyverse-compatibility}}
+\texttt{switch\_key()}
 
-\emph{list supported tidyverse functions}
+\hypertarget{compatibility-with-existing-software}{%
+\subsection{Compatibility with existing
+software}\label{compatibility-with-existing-software}}
 
-\hypertarget{others}{%
-\subsection{Others}\label{others}}
+Building from an underlying \texttt{tbl\_df} structure, it is natural to
+implement methods available in \texttt{dplyr} to \texttt{cubble}.
+Supported methods in the \texttt{cubble} with \texttt{dplyr} generics
+includes: \texttt{mutate}, \texttt{filter}, \texttt{summarise},
+\texttt{select}, \texttt{arrange}, \texttt{rename}, \texttt{left\_join},
+\texttt{group\_by}, \texttt{ungroup}, and the slice family
+(\texttt{slice\_head}, \texttt{slice\_tail}, \texttt{slice\_sample},
+\texttt{slice\_min} and \texttt{slice\_max})
+\textcolor{red}{into a table}.
+
+\texttt{cubble} is also compatible with \texttt{tsibble} in the sense
+that the original list-column can be a \texttt{tbl\_ts} object.
+Duplicates and gaps shoudl be first checked before structuring the data
+into a cubble. If the input data is a \texttt{tsibble} object, the long
+form cubble is also a \texttt{tsibble} where users can directly apply
+time series operations.
 
@@
 
-Spatio-temporal data don't usually come to the analysts as a whole
-piece. A way to look at these data is to divide it into spatial and
-temporal dimension with an ID that links between the two. The first row
-in Figure \ref{fig:cubble-diagram} illustrates this representation where
-in the spatial dimension, the data is characterised by \texttt{id},
-\texttt{lat}, \texttt{long}. \(V_s\) in the last column represents all
-the other site-wise variables, for example, elevation and full name etc.
-The temporal dimension, on the other hand, can be characterised by
-\texttt{id} and \texttt{t} with \(V_t\) representing all the time-wise
-variables. In climate data, this could include precipitation, maximum or
-minimum temperature, and wind speed etc.
-
-\begin{CodeChunk}
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/input-formats/input-formats.001}
-
-}
-
-\caption[Cubble diagram]{Cubble diagram}\label{fig:cubble-diagram}
-\end{figure}
-\end{CodeChunk}
+\textbf{The main difficulty and challenge}\\
+The main difficulty in visualising this type of data is to show
+information in both space and time dimension with the proper level of
+details without information overflow. This would sometimes require
+aggregating the time dimension into the proper level or slicing the data
+into a reasonable number of subset for display. In this sense, a data
+structure that regulates the manipulation spatio-temporal data will
+benefit the analysis workflow. While many implementations focus on
+manipulating and visualising pure spatial or temporal data, there are
+not sufficient tools to deal with spatio-temporal data. The purpose of
+this paper is to introduce a spatio-temporal vector data structure for
+data analysis in R.
 
@@
 join.
-
-A cubble, in essence, wires both dimensions in the spatio-temporal data
-into one object while provide two forms for manipulation the spatial and
-temporal dimension separately.
-
-When manipulating the spatial dimension it uses a nest form that:
-
-\begin{itemize}
-\tightlist
-\item
-  defines each group in a row,
-\item
-  displays the group-related variables in columns, and
-\item
-  nests all the time-related variables into a column called \texttt{ts}.
-\end{itemize}
-
-When manipulating the temporal dimension, it uses the long form that:
-
-\begin{itemize}
-\tightlist
-\item
-  each combination of group and timestamp occupies a row
-\item
-  time-related variables are displayed, and
-\item
-  group-related variables are not explicitly displayed but can be
-  accessed through the \texttt{meta} attribute.
-\end{itemize}
 

### 2021-11-25 main.Rmd
re-structure introduction and section 2 WIP

@@
 
+**Motivation**
+
+In spatio-temporal data analysis, there's a need for a new spato-temporal data structure.
+The requirement for such a tool is necessary given the ubiquity of spatio-temporal data in the wild.
+Example of this includes house price of a city or county, climate station data from the government.
+The ireland wind data from
+
+**Existing packages**
+
+Many data structures have been proposed for spatial and temporal data, but not many for spatio-temporal data.
+One of the reason could be the inherent different levels of information make it inefficient to store in the same table.
+`spacetime`[@spacetime] proposed four space-time layouts: Full grid (STF), sparse grid(STS), irregular (STI), and trajectory (STT) based on underlying spatial structure `sp` [@sp] and temporal structure `xts`[@xts].
+`spatstat` [@spatstat] implements a `ppp` class for point pattern data.
+More recent package `stars` [@stars] uses a spatiotemporal array to store the data and the array structure has its influence from `cubelyr` [@cubelyr], a dplyr data cube backend.
+\newline
+
+**Our new data structure for spatio-temporal data**
+
+In this paper, a new spatio-temporal data structure, `cubble`, is introduced.
+`cubble` implements a relational data structure that splits the data into its spatial and temporal component.
+With this structure, users can separately manipulate the spaital or temporal component of the data while keep the two dimension linked.
+The software is available from the Comprehensive R Archive Network (CRAN) at [CRAN link].
+\newline
+
+**Section division**
+
+The rest of the paper will be divided as follows: Section 2 reviews the existing data structure for spatio, temporal, and spatio-temporal data.
+Section 3 presents a new data structure for spatio-temporal data: cubble.
+Then the paper introduces the workflow of data manipulation and visualisation with the cubble structure in Section 4.
+Section 5 gives some examples on how common spatial and temporal manipulations are performed with cubble and how static and interactive visualisation help to understand climate and [...] data.
+\newline
+
+**Criteria of a data structure we want**
+
+With recent development in the R community, `sf` [@sf] and `tsibble`[@tsibble] have replaced `sp` and `xts` to be the convention structure for spatial and temporal data.
+One reason for their popularity is its integration with tidyverse ecosystem, making them intuitive and easy to adopt.
+For spatio-temporal data, we hope to build a data structure has the following features:
+
+  1) A data structure that handles spatial and temporal dimension in a relational structure. This derives from the 3rd tidy data principal.
+  2) An intuitive and easy to use interface that fits into the tidyverse ecosystem, and
+  3) Compatibility with the latest spatial and temporal data structure. This would give users the flexibility to use the work in existing packages.
+
 **Spatio-temporal data + examples**  
@@
 Recent software development in R has proposed several relational data structure: `tidygraph`[@tidygraph] for graph manipulation, `dm` [@dm] for relational data model, while spatio-temporal data could benefit from having its own relational data structure that identifies the essential spati0-temporal element: site identifier, time identifier, geographic coordinates.
-In this paper, we propose a tidy data structure for vector spaio-temporal data.
-\newline
-
-**Section division**  
-The rest of the paper will be divided as follows: Section 2 reviews the existing data structure for spatio, temporal, and spatio-temporal data.
-Section 3 presents a new data structure for spatio-temporal data: cubble.
-Then the paper introduces the workflow of data manipulation and visualisation with the cubble structure in Section 4.
-Section 5 gives some examples on how common spatial and temporal manipulations are performed with cubble and how static and interactive visualisation help to understand climate and [...] data.
 \newline
@@
 # Existing data structure for spatio and temporal data {#lit-review}
-
-**Existing packages**  
-Many data structures have been proposed for spatial and temporal data, but not many for spatio-temporal data.
-One of the reason could be the inherent different levels of information make it inefficient to store in the same table.
-`spacetime`[@spacetime] proposed four space-time layouts: Full grid (STF), sparse grid(STS), irregular (STI), and trajectory (STT) based on underlying spatial structure `sp` [@sp] and temporal structure `xts`[@xts].
-`spatstat` [@spatstat] implements a `ppp` class for point pattern data.
-More recent package `stars` [@stars] uses a spatiotemporal array to store the data and the array structure has its influence from `cubelyr` [@cubelyr], a dplyr data cube backend.
-\newline
-
-**Criteria of a data structure we want**
-
-With recent development in the R community, `sf` [@sf] and `tsibble`[@tsibble] have replaced `sp` and `xts` to be the convention structure for spatial and temporal data.
-One reason for their popularity is its integration with tidyverse ecosystem, making them intuitive and easy to adopt.
-For spatio-temporal data, we hope to build a data structure has the following features:
-
-  1) A data structure that handles spatial and temporal dimension in a relational structure. This derives from the 3rd tidy data principal.
-  2) An intuitive and easy to use interface that fits into the tidyverse ecosystem, and
-  3) Compatibility with the latest spatial and temporal data structure. This would give users the flexibility to use the work in existing packages.
 

### 2021-11-25 main.tex
re-structure introduction and section 2 WIP

@@
 
-\textbf{Spatio-temporal data + examples}\\
-Spatio-temporal data record changes of variables in spatially separated
-regions across time. In this article, we consider spatio-temporal vector
-data, which are recorded in a fixed interval. Examples of this type of
-data include the house price of a city or county, climate measures from
-weather stations in a country, and river level data from electronic
-gauges \textcolor{red}{more examples here}. \newline
+\textbf{Motivation}
 
-\textbf{point to common feature: spatial level variables + time level
-variables relates to Table 13 in Tidy data paper}\\
-Usually, this type of spatio-temporal data does not necessarily come in
-as a single table. Tidy data principle \citep{tidydata} prescribes each
-type of observational unit to form a table. This would suggest two
-tables to store the data, one for spatial-level and one for
-temporal-level. The Table 13 in \citet{tidydata} presents a structure
-like this and the author argues the lack of tools to work with
-relational data like such. \newline
+In spatio-temporal data analysis, there's a need for a new
+spato-temporal data structure. The requirement for such a tool is
+necessary given the ubiquity of spatio-temporal data in the wild.
+Example of this includes house price of a city or county, climate
+station data from the government,
 
-\textbf{My proposal}\\
-Recent software development in R has proposed several relational data
-structure: \texttt{tidygraph}\citep{tidygraph} for graph manipulation,
-\texttt{dm} \citep{dm} for relational data model, while spatio-temporal
-data could benefit from having its own relational data structure that
-identifies the essential spati0-temporal element: site identifier, time
-identifier, geographic coordinates. In this paper, we propose a tidy
-data structure for vector spaio-temporal data. \newline
+\textbf{Existing packages}
 
-\textbf{Section division}\\
-The rest of the paper will be divided as follows: Section 2 reviews the
-existing data structure for spatio, temporal, and spatio-temporal data.
-Section 3 presents a new data structure for spatio-temporal data:
-cubble. Then the paper introduces the workflow of data manipulation and
-visualisation with the cubble structure in Section 4. Section 5 gives
-some examples on how common spatial and temporal manipulations are
-performed with cubble and how static and interactive visualisation help
-to understand climate and {[}\ldots{]} data. \newline
-
-\newpage
-
-\hypertarget{lit-review}{%
-\section{Existing data structure for spatio and temporal
-data}\label{lit-review}}
-
-\textbf{Existing packages}\\
 Many data structures have been proposed for spatial and temporal data,
@@
 cube backend. \newline
+
+\textbf{Our new data structure for spatio-temporal data}
+
+In this paper, a new spatio-temporal data structure, \texttt{cubble}, is
+introduced. \texttt{cubble} implements a relational data structure that
+splits the data into its spatial and temporal component. With this
+structure, users can separately manipulate the spaital or temporal
+component of the data while keep the two dimension linked. The software
+is available from the Comprehensive R Archive Network (CRAN) at {[}CRAN
+link{]}. \newline
+
+\textbf{Section division}
+
+The rest of the paper will be divided as follows: Section 2 reviews the
+existing data structure for spatio, temporal, and spatio-temporal data.
+Section 3 presents a new data structure for spatio-temporal data:
+cubble. Then the paper introduces the workflow of data manipulation and
+visualisation with the cubble structure in Section 4. Section 5 gives
+some examples on how common spatial and temporal manipulations are
+performed with cubble and how static and interactive visualisation help
+to understand climate and {[}\ldots{]} data. \newline
 
@@
 
+\textbf{Spatio-temporal data + examples}\\
+Spatio-temporal data record changes of variables in spatially separated
+regions across time. In this article, we consider spatio-temporal vector
+data, which are recorded in a fixed interval. Examples of this type of
+data include the house price of a city or county, climate measures from
+weather stations in a country, and river level data from electronic
+gauges \textcolor{red}{more examples here}. \newline
+
+\textbf{point to common feature: spatial level variables + time level
+variables relates to Table 13 in Tidy data paper}\\
+Usually, this type of spatio-temporal data does not necessarily come in
+as a single table. Tidy data principle \citep{tidydata} prescribes each
+type of observational unit to form a table. This would suggest two
+tables to store the data, one for spatial-level and one for
+temporal-level. The Table 13 in \citet{tidydata} presents a structure
+like this and the author argues the lack of tools to work with
+relational data like such. \newline
+
+\textbf{My proposal}\\
+Recent software development in R has proposed several relational data
+structure: \texttt{tidygraph}\citep{tidygraph} for graph manipulation,
+\texttt{dm} \citep{dm} for relational data model, while spatio-temporal
+data could benefit from having its own relational data structure that
+identifies the essential spati0-temporal element: site identifier, time
+identifier, geographic coordinates. \newline
+
 \newpage
 
-\hypertarget{workflow-data-pipeline}{%
-\section{Workflow/ Data pipeline}\label{workflow-data-pipeline}}
+\hypertarget{lit-review}{%
+\section{Existing data structure for spatio and temporal
+data}\label{lit-review}}
+
+\newpage
+
+\hypertarget{cubble-data-structure}{%
+\section{Cubble data structure}\label{cubble-data-structure}}
 
@@
 
-\hypertarget{compatibility-with-existing-software}{%
-\subsection{Compatibility with existing
-software}\label{compatibility-with-existing-software}}
+\hypertarget{integrating-into-a-tidy-workflow}{%
+\subsection{Integrating into a tidy
+workflow}\label{integrating-into-a-tidy-workflow}}
 

### 2021-11-25 script.R
use cubble to analysis ireland wind data

@@ -1,0 +1,29 @@
+library(gstat)
+library(tidyverse)
+library(cubble)
+
+wind_temporal <- wind %>% 
+  pivot_longer(RPT:MAL, names_to = "code", values_to = "wind_speed") %>% 
+  mutate(date = lubridate::make_date(year + 1900, month, day)) %>% 
+  select(date, code, wind_speed)
+
+wind.loc$y = as.numeric(sp::char2dms(as.character(wind.loc[["Latitude"]])))
+wind.loc$x = as.numeric(sp::char2dms(as.character(wind.loc[["Longitude"]])))
+coordinates(wind.loc) = ~x+y
+
+wind_spatial <- wind.loc %>% 
+  as_tibble() %>% 
+  rename(long = x, lat = y, station = Station, code = Code) %>% 
+  select(station, code, long, lat)
+  
+wind_cubble <- wind_temporal %>% left_join(wind_spatial) %>% 
+  as_cubble(key = code, index = date, coords = c("long", "lat"))
+
+wind_cubble %>% ggplot(aes(x = long, y = lat)) + geom_point()
+ireland <- rnaturalearth::ne_countries(scale = 50, type = "countries", 
+                   country = "Ireland", 
+                   returnclass = "sf")
+plot_map(ireland) + 
+  geom_point(data = wind_cubble, aes(x = long, y = lat), shape = "+", color = "red", size = 5) + 
+  geom_text(data = wind_cubble, aes(x = long, y = lat-0.05, label = station), size= 2)
+

### 2021-11-26 main.Rmd
a second sketch of introduction

@@
 
-In spatio-temporal data analysis, there's a need for a new spato-temporal data structure.
-The requirement for such a tool is necessary given the ubiquity of spatio-temporal data in the wild.
-Example of this includes house price of a city or county, climate station data from the government.
-The ireland wind data from
+Many data structures have been proposed for spatial (`sf` by @sf) and temporal (`tsibble` by @tsibble) data in the R community, while less has been done for spatio-temporal data.
+The lack of such tools could potentially  because analysts usually treat the spatial and temporal dimension separately, without realising the need to create a new data structure for spatio-temporal data.
+While this approach fulfills the third tidy data principal to store each type of observational unit in its own table [@tidydata], analysts will always need to manually join the station summary of temporal results into the spatial table or combine the two tables into one for downstream analysis.
+This additional step doesn't add new operations but can be error prone.
+\newline
 
@@
 
-Many data structures have been proposed for spatial and temporal data, but not many for spatio-temporal data.
-One of the reason could be the inherent different levels of information make it inefficient to store in the same table.
-`spacetime`[@spacetime] proposed four space-time layouts: Full grid (STF), sparse grid(STS), irregular (STI), and trajectory (STT) based on underlying spatial structure `sp` [@sp] and temporal structure `xts`[@xts].
+Currently, available spatio-temporal data structure in R includes: `spacetime`[@spacetime], which has proposed four space-time layouts: Full grid (STF), sparse grid(STS), irregular (STI), and trajectory (STT).
+The data structure it uses are based on `sp` [@sp] for spatial and `xts`[@xts] for temporal, both of which has been replaced by more recent implementations.
 `spatstat` [@spatstat] implements a `ppp` class for point pattern data.
-More recent package `stars` [@stars] uses a spatiotemporal array to store the data and the array structure has its influence from `cubelyr` [@cubelyr], a dplyr data cube backend.
+More recent package `stars` [@stars] uses a spatio-temporal array to store the data and the array structure has its influence from `cubelyr` [@cubelyr], a dplyr data cube backend.
+None of these packages provide [....] and this motivates a new spatio-temporal data structure that can smoothly switch between the spatial and temporal table for manipulation of each dimension, while also take care of combining the two dimension together.
+The requirement for such a tool is important given the ubiquity of spatio-temporal vector data in the wild.
+Example of this includes [house price of a city or county], climate station data from Beureau of Meteorology (BoM), the Ireland wind data from `gstat`, air pollution data.
 \newline
@@
 
+##################################################################
+
 **Criteria of a data structure we want**
@@
 
-**Spatio-temporal data + examples**  
-Spatio-temporal data record changes of variables in spatially separated regions across time.
-In this article, we consider spatio-temporal vector data, which are recorded in a fixed interval.
-Examples of this type of data include the house price of a city or county, climate measures from weather stations in a country, and river level data from electronic gauges \textcolor{red}{more examples here}.
-\newline
+ * In this article, we consider spatio-temporal vector data, which are recorded in a fixed interval.
 

### 2021-11-26 main.tex
a second sketch of introduction

@@
 
-In spatio-temporal data analysis, there's a need for a new
-spato-temporal data structure. The requirement for such a tool is
-necessary given the ubiquity of spatio-temporal data in the wild.
-Example of this includes house price of a city or county, climate
-station data from the government,
+Many data structures have been proposed for spatial (\texttt{sf} by
+\citet{sf}) and temporal (\texttt{tsibble} by \citet{tsibble}) data in
+the R community, while less has been done for spatio-temporal data. The
+lack of such tools could potentially because analysts usually treat the
+spatial and temporal dimension separately, without realising the need to
+create a new data structure for spatio-temporal data. While this
+approach fulfills the third tidy data principal to store each type of
+observational unit in its own table \citep{tidydata}, analysts will
+always need to manually join the station summary of temporal results
+into the spatial table or combine the two tables into one for downstream
+analysis. This additional step doesn't add new operations but can be
+error prone. \newline
 
@@
 
-Many data structures have been proposed for spatial and temporal data,
-but not many for spatio-temporal data. One of the reason could be the
-inherent different levels of information make it inefficient to store in
-the same table. \texttt{spacetime}\citep{spacetime} proposed four
-space-time layouts: Full grid (STF), sparse grid(STS), irregular (STI),
-and trajectory (STT) based on underlying spatial structure \texttt{sp}
-\citep{sp} and temporal structure \texttt{xts}\citep{xts}.
+Currently, available spatio-temporal data structure in R includes:
+\texttt{spacetime}\citep{spacetime}, which has proposed four space-time
+layouts: Full grid (STF), sparse grid(STS), irregular (STI), and
+trajectory (STT). The data structure it uses are based on \texttt{sp}
+\citep{sp} for spatial and \texttt{xts}\citep{xts} for temporal, both of
+which has been replaced by more recent implementations.
 \texttt{spatstat} \citep{spatstat} implements a \texttt{ppp} class for
 point pattern data. More recent package \texttt{stars} \citep{stars}
-uses a spatiotemporal array to store the data and the array structure
+uses a spatio-temporal array to store the data and the array structure
 has its influence from \texttt{cubelyr} \citep{cubelyr}, a dplyr data
-cube backend. \newline
+cube backend. None of these packages provide {[}\ldots.{]} and this
+motivates a new spatio-temporal data structure that can smoothly switch
+between the spatial and temporal table for manipulation of each
+dimension, while also take care of combining the two dimension together.
+The requirement for such a tool is important given the ubiquity of
+spatio-temporal vector data in the wild. Example of this includes
+{[}house price of a city or county{]}, climate station data from Beureau
+of Meteorology (BoM), the Ireland wind data from \texttt{gstat}, air
+pollution data. \newline
 
@@
 
-\textbf{Spatio-temporal data + examples}\\
-Spatio-temporal data record changes of variables in spatially separated
-regions across time. In this article, we consider spatio-temporal vector
-data, which are recorded in a fixed interval. Examples of this type of
-data include the house price of a city or county, climate measures from
-weather stations in a country, and river level data from electronic
-gauges \textcolor{red}{more examples here}. \newline
+\begin{itemize}
+\tightlist
+\item
+  In this article, we consider spatio-temporal vector data, which are
+  recorded in a fixed interval.
+\end{itemize}
 

### 2021-11-28 main.Rmd
introduction revised

--- 2021-11-23 main.Rmd
@@
 
-**Spatio-temporal data + examples**  
-Spatio-temporal data record changes of variables in spatially separated regions across time.
-In this article, we consider spatio-temporal vector data, which are recorded in a fixed interval.
-Examples of this type of data include the house price of a city or county, climate measures from weather stations in a country, and river level data from electronic gauges \textcolor{red}{more examples here}.
+**Motivation**
+
+Many data structures have been proposed for spatial (`sf` by @sf) and temporal (`tsibble` by @tsibble) data in the R community, while less has been done for spatio-temporal data.
+The lack of such tools could potentially  because analysts usually treat the spatial and temporal dimension pf the data separately, without realising the need to create a new data structure.
+While this approach follows the third tidy data principal [@tidydata] (*Each type of observational unit forms a table*), analysts always need to manually join results from different observational units or combining multiple tables into one for downstream analysis.
+This additional step doesn't add new operations into the data but can be error prone.
 \newline
 
-<!-- and are point based, characterised by longitude and latitude, in the spatial aspect -->
+**Existing packages**
 
-**point to common feature: spatial level variables + time level variables relates to Table 13 in Tidy data paper**  
-Usually, this type of spatio-temporal data does not necessarily come in as a single table.
-Tidy data principle [@tidydata] prescribes each type of observational unit to form a table.
-This would suggest two tables to store the data, one for spatial-level and one for temporal-level.
-The Table 13 in @tidydata presents a structure like this and the author argues the lack of tools to work with relational data like such.
+Currently, available spatio-temporal data structure in R includes: `spacetime`[@spacetime], which proposes four space-time layouts: Full grid (STF), sparse grid(STS), irregular (STI), and trajectory (STT).
+The data structure it uses is based on `sp` [@sp] and `xts`[@xts], both of which has been replaced by more recent implementations.
+`spatstat` [@spatstat] implements a `ppp` class for point pattern data; and more recent, `stars` [@stars] implements a spatio-temporal array with the dplyr's data cube structure `cubelyr` [@cubelyr] as its backend.
+While these implementations either store spatial and temporal variables all in a single table, hence duplicate the spatial variables for each temporal unit; or split them into two separate tables that has the problem of manually joining, mentioned in the previously.
+None of these packages enjoy both the benefits of being able to separate manipulation in the two dimensions while also keep the data object as a whole.
+This create a gap in the software development.
+The requirement for such a tool is important given the ubiquity of spatio-temporal vector data in the wild: the Ireland wind data from `gstat` is an classic example data that splits variables into spatial (`wind.loc`) and temporal (`wind`) dimension;Bureau of Meteorology (BoM) provides climate observations that are widely applied in agriculture and ecology study; air pollution data.
 \newline
 
-**My proposal**  
-Recent software development in R has proposed several relational data structure: `tidygraph`[@tidygraph] for graph manipulation, `dm` [@dm] for relational data model, while spatio-temporal data could benefit from having its own relational data structure that identifies the essential spati0-temporal element: site identifier, time identifier, geographic coordinates.
-In this paper, we propose a tidy data structure for vector spaio-temporal data.
+**Our new data structure for spatio-temporal data**  
+This paper describes the implementation of a new spatio-temporal data structure: `cubble`.
+`cubble` implements a relational data structure that uses two forms to manage the switch between spatial and temporal dimension.
+With this structure, users can manipulate the spatial or temporal dimension separately, while leaves the linking of two dimensions to `cubble`.
+The software is available from the Comprehensive R Archive Network (CRAN) at [CRAN link].
 \newline
@@
 **Section division**
-The rest of the paper will be divided as follows: Section 2 reviews the existing data structure for spatio, temporal, and spatio-temporal data.
+
+The rest of the paper will be divided as follows: [complete when the paper structure is more solid]
+
+<!-- Section 2 reviews the existing data structure for spatio, temporal, and spatio-temporal data.
 Section 3 presents a new data structure for spatio-temporal data: cubble.
@@
 Section 5 gives some examples on how common spatial and temporal manipulations are performed with cubble and how static and interactive visualisation help to understand climate and [...] data.
-\newline
-
-\newpage
-
-# Existing data structure for spatio and temporal data {#lit-review}
-
-**Existing packages**  
-Many data structures have been proposed for spatial and temporal data, but not many for spatio-temporal data.
-One of the reason could be the inherent different levels of information make it inefficient to store in the same table.
-`spacetime`[@spacetime] proposed four space-time layouts: Full grid (STF), sparse grid(STS), irregular (STI), and trajectory (STT) based on underlying spatial structure `sp` [@sp] and temporal structure `xts`[@xts].
-`spatstat` [@spatstat] implements a `ppp` class for point pattern data.
-More recent package `stars` [@stars] uses a spatiotemporal array to store the data and the array structure has its influence from `cubelyr` [@cubelyr], a dplyr data cube backend.
-\newline
-
-**Criteria of a data structure we want**
-
-With recent development in the R community, `sf` [@sf] and `tsibble`[@tsibble] have replaced `sp` and `xts` to be the convention structure for spatial and temporal data.
-One reason for their popularity is its integration with tidyverse ecosystem, making them intuitive and easy to adopt.
-For spatio-temporal data, we hope to build a data structure has the following features:
-
-  1) A data structure that handles spatial and temporal dimension in a relational structure. This derives from the 3rd tidy data principal.
-  2) An intuitive and easy to use interface that fits into the tidyverse ecosystem, and
-  3) Compatibility with the latest spatial and temporal data structure. This would give users the flexibility to use the work in existing packages.
+\newline -->
 

### 2021-11-28 main.tex
introduction revised

--- 2021-11-25 main.tex
@@
 
-In spatio-temporal data analysis, there's a need for a new
-spato-temporal data structure. The requirement for such a tool is
-necessary given the ubiquity of spatio-temporal data in the wild.
-Example of this includes house price of a city or county, climate
-station data from the government,
+Many data structures have been proposed for spatial (\texttt{sf} by
+\citet{sf}) and temporal (\texttt{tsibble} by \citet{tsibble}) data in
+the R community, while less has been done for spatio-temporal data. The
+lack of such tools could potentially because analysts usually treat the
+spatial and temporal dimension pf the data separately, without realising
+the need to create a new data structure. While this approach follows the
+third tidy data principal \citep{tidydata} (\emph{Each type of
+observational unit forms a table}), analysts always need to manually
+join results from different observational units or combining multiple
+tables into one for downstream analysis. This additional step doesn't
+add new operations into the data but can be error prone. \newline
 
@@
 
-Many data structures have been proposed for spatial and temporal data,
-but not many for spatio-temporal data. One of the reason could be the
-inherent different levels of information make it inefficient to store in
-the same table. \texttt{spacetime}\citep{spacetime} proposed four
-space-time layouts: Full grid (STF), sparse grid(STS), irregular (STI),
-and trajectory (STT) based on underlying spatial structure \texttt{sp}
-\citep{sp} and temporal structure \texttt{xts}\citep{xts}.
-\texttt{spatstat} \citep{spatstat} implements a \texttt{ppp} class for
-point pattern data. More recent package \texttt{stars} \citep{stars}
-uses a spatiotemporal array to store the data and the array structure
-has its influence from \texttt{cubelyr} \citep{cubelyr}, a dplyr data
-cube backend. \newline
+Currently, available spatio-temporal data structure in R includes:
+\texttt{spacetime}\citep{spacetime}, which proposes four space-time
+layouts: Full grid (STF), sparse grid(STS), irregular (STI), and
+trajectory (STT). The data structure it uses is based on \texttt{sp}
+\citep{sp} and \texttt{xts}\citep{xts}, both of which has been replaced
+by more recent implementations. \texttt{spatstat} \citep{spatstat}
+implements a \texttt{ppp} class for point pattern data; and more recent,
+\texttt{stars} \citep{stars} implements a spatio-temporal array with the
+dplyr's data cube structure \texttt{cubelyr} \citep{cubelyr} as its
+backend. While these implementations either store spatial and temporal
+variables all in a single table, hence duplicate the spatial variables
+for each temporal unit; or split them into two separate tables that has
+the problem of manually joining, mentioned in the previously. None of
+these packages enjoy both the benefits of being able to separate
+manipulation in the two dimensions while also keep the data object as a
+whole. This create a gap in the software development. The requirement
+for such a tool is important given the ubiquity of spatio-temporal
+vector data in the wild: the Ireland wind data from \texttt{gstat} is an
+classic example data that splits variables into spatial
+(\texttt{wind.loc}) and temporal (\texttt{wind}) dimension;Bureau of
+Meteorology (BoM) provides climate observations that are widely applied
+in agriculture and ecology study; air pollution data. \newline
 
-\textbf{Our new data structure for spatio-temporal data}
-
-In this paper, a new spatio-temporal data structure, \texttt{cubble}, is
-introduced. \texttt{cubble} implements a relational data structure that
-splits the data into its spatial and temporal component. With this
-structure, users can separately manipulate the spaital or temporal
-component of the data while keep the two dimension linked. The software
-is available from the Comprehensive R Archive Network (CRAN) at {[}CRAN
-link{]}. \newline
+\textbf{Our new data structure for spatio-temporal data}\\
+This paper describes the implementation of a new spatio-temporal data
+structure: \texttt{cubble}. \texttt{cubble} implements a relational data
+structure that uses two forms to manage the switch between spatial and
+temporal dimension. With this structure, users can manipulate the
+spatial or temporal dimension separately, while leaves the linking of
+two dimensions to \texttt{cubble}. The software is available from the
+Comprehensive R Archive Network (CRAN) at {[}CRAN link{]}. \newline
 
@@
 
-The rest of the paper will be divided as follows: Section 2 reviews the
-existing data structure for spatio, temporal, and spatio-temporal data.
-Section 3 presents a new data structure for spatio-temporal data:
-cubble. Then the paper introduces the workflow of data manipulation and
-visualisation with the cubble structure in Section 4. Section 5 gives
-some examples on how common spatial and temporal manipulations are
-performed with cubble and how static and interactive visualisation help
-to understand climate and {[}\ldots{]} data. \newline
-
-\textbf{Criteria of a data structure we want}
-
-With recent development in the R community, \texttt{sf} \citep{sf} and
-\texttt{tsibble}\citep{tsibble} have replaced \texttt{sp} and
-\texttt{xts} to be the convention structure for spatial and temporal
-data. One reason for their popularity is its integration with tidyverse
-ecosystem, making them intuitive and easy to adopt. For spatio-temporal
-data, we hope to build a data structure has the following features:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi})}
-\tightlist
-\item
-  A data structure that handles spatial and temporal dimension in a
-  relational structure. This derives from the 3rd tidy data principal.
-\item
-  An intuitive and easy to use interface that fits into the tidyverse
-  ecosystem, and
-\item
-  Compatibility with the latest spatial and temporal data structure.
-  This would give users the flexibility to use the work in existing
-  packages.
-\end{enumerate}
-
-\textbf{Spatio-temporal data + examples}\\
-Spatio-temporal data record changes of variables in spatially separated
-regions across time. In this article, we consider spatio-temporal vector
-data, which are recorded in a fixed interval. Examples of this type of
-data include the house price of a city or county, climate measures from
-weather stations in a country, and river level data from electronic
-gauges \textcolor{red}{more examples here}. \newline
-
-\textbf{point to common feature: spatial level variables + time level
-variables relates to Table 13 in Tidy data paper}\\
-Usually, this type of spatio-temporal data does not necessarily come in
-as a single table. Tidy data principle \citep{tidydata} prescribes each
-type of observational unit to form a table. This would suggest two
-tables to store the data, one for spatial-level and one for
-temporal-level. The Table 13 in \citet{tidydata} presents a structure
-like this and the author argues the lack of tools to work with
-relational data like such. \newline
-
-\textbf{My proposal}\\
-Recent software development in R has proposed several relational data
-structure: \texttt{tidygraph}\citep{tidygraph} for graph manipulation,
-\texttt{dm} \citep{dm} for relational data model, while spatio-temporal
-data could benefit from having its own relational data structure that
-identifies the essential spati0-temporal element: site identifier, time
-identifier, geographic coordinates. \newline
-
-\newpage
-
-\hypertarget{lit-review}{%
-\section{Existing data structure for spatio and temporal
-data}\label{lit-review}}
+The rest of the paper will be divided as follows: {[}complete when the
+paper structure is more solid{]}
 

### 2021-11-28 main.Rmd
editing on intro to cubble

@@
   \usepackage{amsmath}
+  \usepackage{array}
 output: rticles::jss_article
@@
 This create a gap in the software development.
-The requirement for such a tool is important given the ubiquity of spatio-temporal vector data in the wild: the Ireland wind data from `gstat` is an classic example data that splits variables into spatial (`wind.loc`) and temporal (`wind`) dimension;Bureau of Meteorology (BoM) provides climate observations that are widely applied in agriculture and ecology study; air pollution data.
+The requirement for such a tool is important given the ubiquity of spatio-temporal vector data in the wild: the Ireland wind data from `gstat` is an classic example data that splits variables into spatial (`wind.loc`) and temporal (`wind`) dimension; Bureau of Meteorology (BoM) provides climate observations that are widely applied in agriculture and ecology study; air pollution data.
 \newline
@@
 \newpage
-# Cubble data structure
+# The cubble package
 
-## Spatio-temporal data in the wild
-
-Spatio-temporal data don't usually come to the analysts in various forms.
-Figure \ref{fig:cubble-diagram} groups them into four categories.
-Despite whichever form the data is in, all the forms have a group identifier, `key`, a time column, `t`, and variables that either various across group, i.e.
-`lon`, `lat`, $C_s$, or across time, i.e.
-$V_t$.
-These identifiers will be the building blocks for the data structure introduced below.
+Spatio-temporal data usually come in various forms and Figure \ref{fig:cubble-diagram} shows four examples of this.
+No matter whichever form the data is in, there are always some common components shared by these data.
+A spatial identifier (`id` in the diagram) identifies each site.
+The temporal identifier (`t` in the diagram) [...].
+Coordinates, comprising of latitude and longitude, are commonly used variables for point pattern data.These identifiers will be the building blocks for the data structure introduced below.
+For other variables, those invariant at each time stamp are spatial variables and those differ are temporal variables.
 
@@
 
-## A new data structure: cubble
+In a cubble, there are two forms: 1) nested form, for manipulating the spatial dimension, and 2) long form, for manipulating the temporal dimension.
+Figure \ref{fig:def-cubble} sketches the two forms along with the associated attributes.
+A variable identifies by the spatial identifies can come from manipulating spatial variables itself, or summary of temporal variables.
+The nested cubble is best suited to work with this type of operation, since it defines each spatial unit as a row.
+The spatial variables are directly displayed in columns.
+Temporal variables are nested in a column called `ts` and the underlying rowwise dataframe uses a `group` attributes to ensure each row  is in its own group.
 
-A cubble, in essence, wires the spatial and temporal dimension into a single object while provides two forms for manipulation each dimension separately.
+<!-- In this structure, the nested form is suitable for manipulating the spatial dimension of the data and this would include manipulating variables that 1) are inherently spatial and 2) summarise the temporal variables by `key`.
+Under the hood, the nested cubble is built on top of the rowwise dataframe (`rowwise_df`).
+This is design to simplify the code when working with the temporal variables - they are nested into the list-column `ts`.
+-->
 
-In the nest form, Figure \ref{fig:def-cubble}a), a cubble:
-
-  - defines each group in a row,
-  - displays the group-related variables in columns, and
-  - nests all the time-related variables into a column called `ts`.
-
-In the long form, Figure \ref{fig:def-cubble}b), a cubble:
-
-  - defines the combination of group and timestamp as a row
-  - displays time-related variables, along with the group identifier in columns, and
-  - stores group-related variables as a attributes called `spatial`.
+Temporal operations are suited to be performed in the long cubble as each row is defines as the combination of spatial and temporal identifier.
+This is also the structure that `tsibble` adopts.
+Temporal variables are directly displayed.
+To avoid repeating the same spatial at each temporal unit, all the spatial variables, along with the spatial identifier, are stored as a `spatial` attributes.
+This information is used when switching back to the nested form.
 
@@
 
-To create a cubble, you can use `as_cubble` with proper supply of `key`, `index`, and `coords`.
-[`key` defines whether a variable is spatial or temperal by looking at whether it varies across `key`.
-For example, `lat` is a spatial variable since for one `id` there's only one `lat` value.
-`index` and `coords` prescribes the temporal index and spaital coordinates used in the data.
-These are common variables that should present in the point pattern spatio-temporal data.] The cubble created by default is in the nested form.
-\newline
+To use functionalities from cubble, data analysts first need to create a cubble.
+`as_cubble` create a `cubble` by supplying the three key components: `key` as the spatial identifier; `index` as the temporal identifier; and a vector of `coords` in the order of longitude first and then latitude.
+The use of `key` and `index` follows the naming convention in `tsibble`.
+The cubble created by default is in the nested form.
+Below is an example of creating a cubble:  \newline
 
 ```{r}
-(cubble_nested <- climate_flat %>% 
+(cubble_nested <- cubble::climate_flat %>% 
   as_cubble(key = id, index = date, coords = c("long", "lat")))
@@
 
-In the `cubble` header, you can read the key variable + number, bbox, and also the name of variable nested in the `ts` column.
-Here in this example, the temporal variables are precipitation, `prcp`, maximum temporature, `tmax`, and minimum temperature, `tmin`.
-
-In this structure, the nested form is suitable for manipulating the spatial dimension of the data and this would include manipulating variables that 1) are inherently spatial and 2) summarise the temporal variables by `key`.
-Under the hood, the nested cubble is built on top of the rowwise dataframe (`rowwise_df`).
-This is design to simplify the code when working with the temporal variables - they are nested into the list-column `ts`.
+In the `cubble` header, you can read the name of the `key` variable, bbox, and also the name of variable nested in the `ts` column.
+In this example, the spatial identifier is `id` and the number in the bracket means there are 5 unique `id` in this dataset.
+The bbox in the second row gives the range of the coordinates.
+The temporal variables are all nested in the `ts` column, but it could be useful to know the name these variables.
+The third row in the cubble header shows these names and in this example this includes: precipitation, `prcp`, maximum temporature, `tmax`, and minimum temperature, `tmin`.
 
@@
 
+From a created cubble in the nested form, analysts may want to directly manipulate the temporal variables.
+This would require switching to the long form using `stretch()`.
 The verb `stretch()` switch the cubble from the nested form into a long form.
@@
 
+## Support on interactive graphics
+
 ## Integrating into a tidy workflow
 
-Building from an underlying `tbl_df` structure, it is natural to implement methods available in `dplyr` to `cubble`.
-Supported methods in the `cubble` with `dplyr` generics includes: `mutate`, `filter`, `summarise`, `select`, `arrange`, `rename`, `left_join`, `group_by`, `ungroup`, and the slice family (`slice_head`, `slice_tail`, `slice_sample`, `slice_min` and `slice_max`) \textcolor{red}{into a table}.
+Building from an underlying `tbl_df` structure, it is natural to implement methods available in `dplyr` to `cubble`. Supported methods in the `cubble` with `dplyr` generics includes:
+
+\begin{center}
+\begin{tabular}{ | m{5em} | m{15cm}| } 
+\textbf{mutate} \\
+\textbf{filter}\\
+\textbf{summarise} \\
+\textbf{select} \\
+\textbf{arrange} \\
+\textbf{rename} \\
+\textbf{left\_join} \\
+\textbf{group\_by} \\
+\textbf{ungroup}\\
+slice family & \textbf{slice\_head}, \textbf{slice\_tail}, \textbf{slice\_sample}, \textbf{slice\_min} and \textbf{slice\_max} \\
+\end{tabular}
+\end{center}
 

### 2021-11-28 main.tex
editing on intro to cubble

@@
 
-\usepackage{amsmath}
+\usepackage{amsmath} \usepackage{array}
 
@@
 classic example data that splits variables into spatial
-(\texttt{wind.loc}) and temporal (\texttt{wind}) dimension;Bureau of
+(\texttt{wind.loc}) and temporal (\texttt{wind}) dimension; Bureau of
 Meteorology (BoM) provides climate observations that are widely applied
@@
 
-\hypertarget{cubble-data-structure}{%
-\section{Cubble data structure}\label{cubble-data-structure}}
+\hypertarget{the-cubble-package}{%
+\section{The cubble package}\label{the-cubble-package}}
 
-\hypertarget{spatio-temporal-data-in-the-wild}{%
-\subsection{Spatio-temporal data in the
-wild}\label{spatio-temporal-data-in-the-wild}}
-
-Spatio-temporal data don't usually come to the analysts in various
-forms. Figure \ref{fig:cubble-diagram} groups them into four categories.
-Despite whichever form the data is in, all the forms have a group
-identifier, \texttt{key}, a time column, \texttt{t}, and variables that
-either various across group, i.e.~\texttt{lon}, \texttt{lat}, \(C_s\),
-or across time, i.e.~\(V_t\). These identifiers will be the building
-blocks for the data structure introduced below.
+Spatio-temporal data usually come in various forms and Figure
+\ref{fig:cubble-diagram} shows four examples of this. No matter
+whichever form the data is in, there are always some common components
+shared by these data. A spatial identifier (\texttt{id} in the diagram)
+identifies each site. The temporal identifier (\texttt{t} in the
+diagram) {[}\ldots{]}. Coordinates, comprising of latitude and
+longitude, are commonly used variables for point pattern data.These
+identifiers will be the building blocks for the data structure
+introduced below. For other variables, those invariant at each time
+stamp are spatial variables and those differ are temporal variables.
 
@@
 
-\hypertarget{a-new-data-structure-cubble}{%
-\subsection{A new data structure:
-cubble}\label{a-new-data-structure-cubble}}
+In a cubble, there are two forms: 1) nested form, for manipulating the
+spatial dimension, and 2) long form, for manipulating the temporal
+dimension. Figure \ref{fig:def-cubble} sketches the two forms along with
+the associated attributes. A variable identifies by the spatial
+identifies can come from manipulating spatial variables itself, or
+summary of temporal variables. The nested cubble is best suited to work
+with this type of operation, since it defines each spatial unit as a
+row. The spatial variables are directly displayed in columns. Temporal
+variables are nested in a column called \texttt{ts} and the underlying
+rowwise dataframe uses a \texttt{group} attributes to ensure each row is
+in its own group.
 
-A cubble, in essence, wires the spatial and temporal dimension into a
-single object while provides two forms for manipulation each dimension
-separately.
-
-In the nest form, Figure \ref{fig:def-cubble}a), a cubble:
-
-\begin{itemize}
-\tightlist
-\item
-  defines each group in a row,
-\item
-  displays the group-related variables in columns, and
-\item
-  nests all the time-related variables into a column called \texttt{ts}.
-\end{itemize}
-
-In the long form, Figure \ref{fig:def-cubble}b), a cubble:
-
-\begin{itemize}
-\tightlist
-\item
-  defines the combination of group and timestamp as a row
-\item
-  displays time-related variables, along with the group identifier in
-  columns, and
-\item
-  stores group-related variables as a attributes called
-  \texttt{spatial}.
-\end{itemize}
+Temporal operations are suited to be performed in the long cubble as
+each row is defines as the combination of spatial and temporal
+identifier. This is also the structure that \texttt{tsibble} adopts.
+Temporal variables are directly displayed. To avoid repeating the same
+spatial at each temporal unit, all the spatial variables, along with the
+spatial identifier, are stored as a \texttt{spatial} attributes. This
+information is used when switching back to the nested form.
 
@@
 
-To create a cubble, you can use \texttt{as\_cubble} with proper supply
-of \texttt{key}, \texttt{index}, and \texttt{coords}. {[}\texttt{key}
-defines whether a variable is spatial or temperal by looking at whether
-it varies across \texttt{key}. For example, \texttt{lat} is a spatial
-variable since for one \texttt{id} there's only one \texttt{lat} value.
-\texttt{index} and \texttt{coords} prescribes the temporal index and
-spaital coordinates used in the data. These are common variables that
-should present in the point pattern spatio-temporal data.{]} The cubble
-created by default is in the nested form. \newline
+To use functionalities from cubble, data analysts first need to create a
+cubble. \texttt{as\_cubble} create a \texttt{cubble} by supplying the
+three key components: \texttt{key} as the spatial identifier;
+\texttt{index} as the temporal identifier; and a vector of
+\texttt{coords} in the order of longitude first and then latitude. The
+use of \texttt{key} and \texttt{index} follows the naming convention in
+\texttt{tsibble}. The cubble created by default is in the nested form.
+Below is an example of creating a cubble: \newline
 
@@
 \begin{CodeInput}
-R> (cubble_nested <- climate_flat %>% 
+R> (cubble_nested <- cubble::climate_flat %>% 
 +   as_cubble(key = id, index = date, coords = c("long", "lat")))
@@
 
-In the \texttt{cubble} header, you can read the key variable + number,
-bbox, and also the name of variable nested in the \texttt{ts} column.
-Here in this example, the temporal variables are precipitation,
-\texttt{prcp}, maximum temporature, \texttt{tmax}, and minimum
-temperature, \texttt{tmin}.
-
-In this structure, the nested form is suitable for manipulating the
-spatial dimension of the data and this would include manipulating
-variables that 1) are inherently spatial and 2) summarise the temporal
-variables by \texttt{key}. Under the hood, the nested cubble is built on
-top of the rowwise dataframe (\texttt{rowwise\_df}). This is design to
-simplify the code when working with the temporal variables - they are
-nested into the list-column \texttt{ts}.
+In the \texttt{cubble} header, you can read the name of the \texttt{key}
+variable, bbox, and also the name of variable nested in the \texttt{ts}
+column. In this example, the spatial identifier is \texttt{id} and the
+number in the bracket means there are 5 unique \texttt{id} in this
+dataset. The bbox in the second row gives the range of the coordinates.
+The temporal variables are all nested in the \texttt{ts} column, but it
+could be useful to know the name these variables. The third row in the
+cubble header shows these names and in this example this includes:
+precipitation, \texttt{prcp}, maximum temporature, \texttt{tmax}, and
+minimum temperature, \texttt{tmin}.
 
@@
 
-The verb \texttt{stretch()} switch the cubble from the nested form into
-a long form. Under the hood, it first extracts the spatial variables
-into a separate tibble to store in the attribute \texttt{spatial} and
-then unnests the \texttt{ts} column to show the temporal content:
+From a created cubble in the nested form, analysts may want to directly
+manipulate the temporal variables. This would require switching to the
+long form using \texttt{stretch()}. The verb \texttt{stretch()} switch
+the cubble from the nested form into a long form. Under the hood, it
+first extracts the spatial variables into a separate tibble to store in
+the attribute \texttt{spatial} and then unnests the \texttt{ts} column
+to show the temporal content:
 
@@
 Supported methods in the \texttt{cubble} with \texttt{dplyr} generics
-includes: \texttt{mutate}, \texttt{filter}, \texttt{summarise},
-\texttt{select}, \texttt{arrange}, \texttt{rename}, \texttt{left\_join},
-\texttt{group\_by}, \texttt{ungroup}, and the slice family
-(\texttt{slice\_head}, \texttt{slice\_tail}, \texttt{slice\_sample},
-\texttt{slice\_min} and \texttt{slice\_max})
-\textcolor{red}{into a table}.
+includes:
+
+\begin{center}
+\begin{tabular}{ | m{5em} | m{15cm}| } 
+\textbf{mutate} \\
+\textbf{filter}\\
+\textbf{summarise} \\
+\textbf{select} \\
+\textbf{arrange} \\
+\textbf{rename} \\
+\textbf{left\_join} \\
+\textbf{group\_by} \\
+\textbf{ungroup}\\
+slice family & \textbf{slice\_head}, \textbf{slice\_tail}, \textbf{slice\_sample}, \textbf{slice\_min} and \textbf{slice\_max} \\
+\end{tabular}
+\end{center}
 

### 2021-11-28 main.Rmd
add in examples from vignettes

@@
 ```{r setup, echo = FALSE}
-knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, 
+knitr::opts_chunk$set(
+  echo = TRUE,
+  warning = FALSE,
+  message = FALSE,
   fig.path = "figures/",
-                      out.width = "100%", 
-                      out.height = "30%")
-options(prompt = 'R> ', continue = '+ ')
+  fig.height = 10
+)
+options(prompt = "R> ", continue = "+ ")
 ```
@@
 library(cubble)
+library(forcats)
+library(stringr)
 library(dplyr)
 library(lubridate)
+library(ggplot2)
+library(GGally)
+library(plotly)
+library(crosstalk)
+library(patchwork)
 ```
@@
 
-```{r cubble-diagram, echo = FALSE, fig.cap="Illustration of incoming data formats for spatio-temporal data.
+```{r cubble-diagram, echo = FALSE, out.width = "100%", out.height = "30%", fig.cap="Illustration of incoming data formats for spatio-temporal data.
 (1) Data comes in as a single table; (2) Separate tables for spatial and temporal variables; (3) A single table with all the parameters used to query the database and a separate table for queried data; and (4) Cubical data in array or NetCDF format.", fig.align='center', out.height = "15%"}
@@
 
-```{r def-cubble, echo = FALSE, fig.align="center", out.height="30%", fig.cap = "Illustration of nested and long cubble."}
+```{r def-cubble, echo = FALSE, fig.align="center", out.height="30%", out.width = "100%", fig.cap = "Illustration of nested and long cubble."}
 knitr::include_graphics(here::here("figures/def-cubble/def-cubble.001.png"))
@@
 
+This vignette introduces how to perform spatial and temporal manipulate in a cubble with dplyr verbs.
+We will illustrate with `weatherdata::historical_tmax` data, which have the historical maximum temperature recorded for Australian stations with the earliest dating back to 1859.
+
+### Spatial manipulation
+
+`historical_tmax` is already in a nested cubble format, which is suitable for station-wise manipulation.
+`rnoaa` construct the station id by prefix `ASN00` to the Bureau of Meteorology (BOM) station number.
+In  [BOM's numbering system](http://www.bom.gov.au/climate/cdo/about/site-num.shtml), the 2nd and 3rd digit denotes the state a station is located in and this is equivalent to the 7th to 8th digit in our string.
+We can mutate/ filter station in a particular state based on this information and here we add a column called `state_id` and filter out the ones in New South Wales and Victoria (ranging from 46 to 90).
+
+```{r}
+tmax_nested <- weatherdata::historical_tmax %>%
+  mutate(state_id = stringr::str_sub(id, 7, 8)) %>%
+  filter(between(state_id, 46, 90))
+```
+
+### Temporal manipulation
+
+There are some operations in the time dimension we would like to make:
+
+  - Extract observations in a particular period, say, those in 1971 to 1975 and 2016 to 2020. This can be used to compare the historical and recent climate. 
+  - Summarise daily records into monthly to remove sparsity
+
+These time dimension operations can be computed in the long form and `stretch()` converts a nested cubble to a long cubble.
+From a long cubble, we can write the exact  same dplyr codes to complete the two tasks:
+
+```{r}
+tmax_long <- tmax_nested %>%
+  stretch() %>%
+  filter(lubridate::year(date) %in% c(1971:1975, 2016:2020)) %>%
+  mutate(
+    month = lubridate::month(date),
+    group = as.factor(ifelse(
+      lubridate::year(date) > 2015,
+      "2016 ~ 2020",
+      "1971 ~ 1975"
+    ))
+  ) %>%
+  group_by(month, group) %>%
+  summarise(tmax = mean(tmax, na.rm = TRUE))
+```
+
+### Back to spatial
+
+A data quality issue with the `rnoaa` data is that while it records the first and last year recorded of each series,  it doesn't report the period of missingness.
+For example, station `ASN00047048` starts it first record in 1957, pauses for a period from 1963 to 1990, and then resumes it recording till today.
+Out of the `r nrow(tmax_nested)` stations in New South Wales and Victoria, 7 stations have this issue and we would like to remove those stations from the comparison.
+
+Again, this is a station-wise operation and to convert back from the long cubble to a nested one, use `tamp()`.
+Here we keep the stations with 24 observations (12 months for both periods) after the monthly aggregation.
+
+```{r}
+tmax_nested2 <- tmax_long %>%
+  tamp() %>%
+  filter(nrow(ts) == 24)
+```
+
+In some visualisation, we may need information from both spatial and temporal dimension.
+One example of this is a glyph map, where spatial variables, i.e.
+`longitude` and `latitude`, are used to construct the major axes and temporal variables, i.e.
+`month` and `tmax`, are used to construct the minor axes.
+This requires these variables to be in the same table, rather than in different forms.
+In cubble, you can append spatial variables that are invariant to the `key` withe `migrate()`
+
+```{r}
+tmax_long2 <- tmax_nested2 %>%
+  stretch() %>%
+  migrate(latitude, longitude)
+```
+
+### Glyph map
+
+Now its time for the glyph map!
+
+```{r}
+gly <- glyphs(
+  tmax_long2,
+  x_major = "longitude",
+  y_major = "latitude",
+  x_minor = "month",
+  y_minor = "tmax",
+  height = 0.6,
+  width = 1
+)
+
+nsw_vic <- ozmaps::abs_ste %>%
+  filter(NAME %in% c("Victoria", "New South Wales"))
+
+plot_map(nsw_vic, fill = "transparent") +
+  geom_path(data = gly, aes(gx, gy,
+    group = interaction(id, group),
+    color = group
+  )) +
+  scale_color_brewer(palette = "Dark2") + 
+  theme_bw() + 
+  coord_sf(xlim = c(140, 155)) + 
+  theme(legend.position = "bottom")
+```
+
+## Spatial and temporal aggregation
+
+On top of `aus_climate` data in this package, `climate_full` from package [`weatherdata`](https://github.com/huizezhang-sherry/weatherdata) has daily climate data of `r nrow(weatherdata::climate_full)` Australia stations from 2016 to 2020.
+This is where these stations locate in an Australia map:
+
+This is a lot of stations to look at for one climate variable and they can't all fit into a glyph map.
+What we can do is to group stations into clusters and look at the aggregated series in the glyph map.
+In this vignette, I will introduce how to perform spatial (and temporal) aggregation using hierarchical data structure in cubble.
+
+First let's add a new column `ll` for calculating distance matrix in the next section and aggregate the daily precipitation into weekly measure.
+These steps should be familiar from the vignette *Data manipulation with cubble*.
+
+```{r}
+station_nested <- weatherdata::climate_full %>%
+  mutate(ll = s2::s2_lnglat(long, lat)) %>%
+  stretch() %>%
+  mutate(wk = lubridate::week(date)) %>%
+  group_by(wk) %>%
+  summarise(prcp = sum(prcp, na.rm = TRUE)) %>%
+  tamp()
+```
+
+### Hierarchical data structure
+
+Imposing a clustering structure can be thought of as building a hierarchical structure where stations are nested within clusters.
+As an example to illustrate here, we use a kmean clustering algorithm based on the distance matrix and specify the number of centers to be 20.
+More complex algorithms can also be used for more complex problem, as long as a mapping from each station id to the cluster id can be constructed.
+We then join this data to our station data:
+
+```{r}
+dist_raw <- scale(s2::s2_distance_matrix(station_nested$ll, station_nested$ll))
+
+cluster_res <- tibble(
+  id = station_nested$id,
+  cluster = kmeans(dist_raw, centers = 20, nstart = 50)$cluster
+)
+
+station_nested <- station_nested %>%
+  left_join(cluster_res)
+```
+
+One thing we hope to do with the cluster is to find the coordinates of the centroid.
+These are variables variant to the station but invariant to the cluster and it would be nice to have a function that structure each cluster as a row.
+`switch_key()` is the function that does this: it lets you to specify a new key, say `cluster` and nests all spatial variables variant to `cluster` into a column.
+Temporal observations from different stations while within the same cluster are bound in the nested column `ts`.
+
+```{r}
+cluster_nested <- station_nested %>%
+  switch_key(cluster)
+```
+
+This structure makes it easy to compute cluster level variable, although there are a few steps to calculate the centroid coordinates: we need to find the convex hull that wraps around the cluster, make it a polygon, find the centroid of the polygon and finally, extract the x and y coordinate of each centroid:
+
+```{r}
+cluster_nested <- cluster_nested %>%
+  mutate(
+    chull = list(chull(.val$long, .val$lat)),
+    ll_cluster = sf::st_as_sfc(
+      s2::s2_make_polygon(c(.val$long[chull]),
+        c(.val$lat[chull]),
+        oriented = FALSE
+      )
+    ),
+    centroid = s2::s2_centroid(ll_cluster),
+    cent_long = s2::s2_x(centroid),
+    cent_lat = s2::s2_y(centroid)
+  )
+```
+
+After we have got `cluster_nested`, spatial and temporal data at both levels can be easily obtained.
+If we use `station` and `cluster` prefix to denote the two levels and `nested` and `long` for whether the data shows the spatial or temporal dimension, the relationship among the four datasets can be illustrated in the following workflow:
+
+Start with the original `station_nested`, `stretch()` expands the `ts` column with each station (`id`) forming a group and attach variables invariant to `id` as an attribute.
+`switch_key()` changes the `key` from `id` to `cluster` and nests all the spatial variables that variant to `cluster`.
+`stretch()` `cluster_nested` will store variables that are invariant to `cluster` as a tibble in the attribute.
+
+Now we can obtain the aggregated series as described in the workflow diagram above and construct the glyph map with `GGally::glyphs()`:
+
+```{r}
+cluster_long <- cluster_nested %>%
+  stretch(ts) %>%
+  group_by(wk) %>%
+  summarise(prcp = sum(prcp, na.rm = TRUE)) %>%
+  migrate(cent_long, cent_lat)
+
+gly <- GGally::glyphs(cluster_long,
+  x_major = "cent_long", x_minor = "wk",
+  y_major = "cent_lat", y_minor = "prcp",
+  height = 2, width = 4
+)
+
+state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)
+plot_map(state_map) +
+  geom_text(
+    data = cluster_nested,
+    aes(x = cent_long, y = cent_lat, label = cluster)
+  ) +
+  geom_path(
+    data = gly,
+    aes(x = gx, y = gy, group = gid)
+  )
+```
+
+We can also look at the precipitation of each individual station within the same cluster:
+
+```{r fig.height = 5, fig.cap = "shtishisa"}
+station_long <- station_nested %>%
+  stretch(ts) %>%
+  migrate(cluster)
+station_long %>%
+  ggplot(aes(x = wk, y = prcp, group = id)) +
+  geom_line(alpha = .3) +
+  facet_wrap(vars(cluster), scales = "free_y", ncol = 4) +
+  theme_bw()
+```
+
+Lastly, there is one series on Tasmania island standing out from others, lets look at where it is:
+
+```{r}
+# this part still needs some fixing
+tas_latlong <- station_nested %>%
+  filter(lat < -40) %>%
+  mutate(p = max(ts$prcp)) %>%
+  strip_rowwise() %>%
+  filter(p == max(p))
+
+state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)
+plot_map(state_map) +
+  geom_point(data = station_nested, aes(x = long, y = lat), size = 0.5) +
+  geom_sf(data = cluster_nested, aes(geometry = ll_cluster), fill = "transparent") +
+  geom_point(data = tas_latlong, aes(x = long, y = lat), col = "red")
+```
+
 ## Matching precipitation and river level in Victria water gauges
 
-Matching  - Vig 3
+The water level data comes from [Bureau of Meteorology](http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full) and has a copy in `weatherdata`.
+Here we extract the water course level and add a column annotate this data of type `river`.
+For the rainfall data, we will still use the `weatherdata::climate_full`, filtering for Victorian stations in 2020 should be pretty familiar by now.
+Again, we first look at where these stations are on the map first:
+
+```{r}
+river <- weatherdata::water %>%
+  stretch() %>%
+  select(date, Water_course_level) %>%
+  tamp() %>%
+  mutate(type = "river")
+
+climate <- weatherdata::climate_full %>%
+  filter(between(stringr::str_sub(id, 7, 8), 76, 90)) %>%
+  stretch() %>%
+  filter(lubridate::year(date) == 2020) %>%
+  tamp() %>%
+  mutate(type = "climate")
+
+vic_map <- rmapshaper::ms_simplify(ozmaps::abs_ste %>% filter(NAME == "Victoria"))
+plot_map(vic_map) +
+  geom_point(
+    data = dplyr::bind_rows(river, climate),
+    aes(x = long, y = lat, color = type)
+  ) +
+  scale_color_brewer(palette = "Dark2") +
+  theme_bw()
+```
+
+### Theory
+
+Temporal matching checks how spatially matched pairs align temporally. We use the following chart to illustrate how the temporal matching works:
+
+```{r echo = FALSE, fig.height=3, fig.cap = "shtishisaasdf"}
+set.seed(123)
+dt <- tibble(
+  id = factor(c(rep("A", 31), rep("a", 31)), levels = c("A", "a")),
+  date = rep(1:31, 2),
+  val = c(
+    c(rnorm(5), 10, rnorm(7), 5, rnorm(8), 7, rnorm(8)),
+    c(
+      rnorm(6, sd = 0.5), 7, rnorm(7, sd = 0.5), rnorm(6, sd = 0.5),
+      4, rnorm(5, sd = 0.5), 6, rnorm(4, sd = 0.5)
+    )
+  )
+) %>% mutate(val = ifelse(val < 0, 0, val))
+
+circle <- tibble(
+  x = c(6, 14, 23, 7, 21, 27),
+  y = c(10, 5, 7, 7, 4, 6),
+  id = factor(c(rep("A", 3), rep("a", 3)), levels = c("A", "a")),
+  xend = x + 5
+)
+
+errorbar <- bind_rows(
+  circle %>% filter(id == "A"),
+  circle %>% filter(id == "A") %>% mutate(id = "a")
+) %>%
+  mutate(id = factor(id, c("A", "a")))
+
+fallin <- tibble(
+  id = factor(rep("a", 2), c("A", "a")),
+  x = c(7, 27),
+  y = -0.5,
+  label = "\U2714"
+)
+
+ggplot() +
+  geom_line(data = dt, aes(x = date, y = val, group = id), color = "grey60") +
+  geom_vline(data = circle, aes(xintercept = x), linetype = "longdash", color = "grey60") +
+  geom_errorbar(data = errorbar, aes(xmin = x, xmax = xend, y = -0.5), width = 0.5) +
+  geom_point(data = circle, aes(x = x, y = y), color = "brown") +
+  geom_text(data = fallin, aes(x = x, y = y, label = label), color = "brown", size = 5) +
+  facet_wrap(vars(id), nrow = 2) +
+  scale_y_continuous(breaks = seq(0, 10, 2)) +
+  scale_x_continuous(breaks = seq(1, 31, 1)) +
+  theme_bw() +
+  theme(panel.grid = element_blank()) +
+  labs(x = "Time", y = "Value")
+```
+
+For each spatially matched pair, say `A` and `a`, we first find the largest `n` points in each series, colored in brown points here.
+Here we use the largest three but you can tune this number by `temporal_n_highest`.
+Then we construct the interval of the largest points from one series and see how many points, from the other series, fall into the intervals.
+The series used to construct the interval is controlled by `temporal_independent` and the window size by `temporal_window` with a default of 5.
+
+In this illustration, we construct the interval based on series `A` and two of the three peaks from `a` falls into this interval at Time 7 and 27.
+
+### Rainfall translates into river level
+
+There's another mandatory argument that hasn't been introduced above: `temporal_var_to_match`.
+This argument controls the variable to match and it needs to appear in both the `major` and `minor` set.
+In the water level matching example, we match the variable `Water_course_level` from `river` to `prcp` from `climate`, hence need to manually rename one of them to match the other, here we rename `Water_course_level` to `prcp` in `river`:
+
+```{r}
+river <- river %>%
+  stretch() %>%
+  rename(prcp = Water_course_level) %>%
+  tamp()
+```
+
+Now we use `match_sites()` to first pair the weather stations with the river gauges spatially and then apply the temporal matching on `prcp`.
+We will construct the interval based on peaks in `climate` since we would expect a lag effect for precipitation to flow into the river and cause a raise in river level, hence `temporal_independent = climate`.
+We select the 30 highest peak from the series to construct the match by setting `temporal_n_highest = 30`.
+This is a tuning parameter and you can start with 10% of the points of one series (here we have daily data for a year, 10% is roughly 30 points).
+`temporal_min_match` filters out pairs don't have enough match and to return all the pairs, set `temporal_min_match` to `0`.
+
+```{r}
+res <- match_sites(river, climate,
+  temporal_var_to_match = prcp,
+  temporal_independent = climate,  
+  temporal_n_highest = 30,
+  temporal_min_match = 15
+)
+
+res
+```
+
+The output from temporal matching is also a cubble, with additional column `.dist` and `.group` inherent from spatial matching and `n_match` for the number of matched temporal peaks.
+Then you can use this output to plot the location of match or to look at the series:
+```{r out.width="100%", fig.cap="..."}
+p1 <- plot_map(vic_map) +
+  geom_point(
+    data = res,
+    aes(x = long, y = lat, color = type)
+  ) +
+  ggrepel::geom_label_repel(
+    data = res %>% filter(type == "river"),
+    aes(x = long, y = lat, label = .group)
+  ) +
+  scale_color_brewer(palette = "Dark2") +
+  ggtitle("Victoria") +
+  theme_minimal() +
+  theme(
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    legend.position = "bottom"
+  )
+
+res_long <- res %>%
+  stretch(ts) %>%
+  migrate(.group, type) %>%
+  mutate(prcp = scale(prcp)[, 1]) %>%
+  switch_key(.group) %>%
+  migrate(type)
+
+p2 <- res_long %>%
+  ggplot(aes(
+    x = date, y = prcp,
+    group = id, color = type
+  )) +
+  geom_line() +
+  facet_wrap(vars(.group)) +
+  scale_color_brewer(palette = "Dark2", guide = "none") +
+  theme_bw() +
+  labs(x = "week") +
+  scale_x_date(date_labels = "%b")
+
+p1 | p2
+```
+
+## Interative graphic with cubble
+
+```{r eval = FALSE}
+set.seed(123)
+tmax_data <- weatherdata::climate_full %>%
+  slice_sample(n = 50) %>%
+  stretch() %>%
+  mutate(ym = tsibble::yearmonth(date)) %>%
+  group_by(ym) %>%
+  summarise(
+    tmax = mean(tmax, na.rm = TRUE),
+    tmin = mean(tmin, na.rm = TRUE),
+    prcp = sum(prcp, na.rm = TRUE)
+  ) %>%
+  ungroup(ym) %>%
+  mutate(
+    year = as.factor(year(ym)),
+    month = as.factor(month(ym)),
+    prcp = sqrt(prcp + 0.001)
+  ) %>%
+  tamp() %>%
+  mutate(median_tmax = median(ts$tmax, na.rm = TRUE)) %>%
+  stretch() %>%
+  migrate(median_tmax, long, lat, name) %>%
+  highlight_key(~id)
+
+p1 <- tmax_data %>%
+  ggplot(aes(
+    x = month, y = tmax, group = interaction(year, id),
+    color = median_tmax, label = name
+  )) +
+  geom_line() +
+  facet_wrap(vars(fct_reorder(name, -median_tmax))) +
+  theme(legend.position = "none")
+
+state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)
+p2 <- tmax_data %>%
+  ggplot() +
+  geom_sf(data = state_map, aes(geometry = geometry)) +
+  geom_point(aes(x = long, y = lat, color = median_tmax, label = name)) +
+  theme_void()
+
+bscols(
+  ggplotly(p1, width = 800, height = 800, tooltip = "label") %>%
+    highlight(on = "plotly_selected", off = "plotly_deselect", color = "red"),
+  ggplotly(p2, width = 800, height = 800, tooltip = "label") %>%
+    highlight(on = "plotly_selected", off = "plotly_deselect", color = "red")
+)
+```
 
@@
 
-calc_dist <- function(df_ref, df, n = 5){
+calc_dist <- function(df_ref, df, n = 5) {
   long_ref <- df_ref$long
@@
   df %>%
-    mutate(long_ref = long_ref,
-           lat_ref=  lat_ref,
-           dist = rnoaa::meteo_spherical_distance(lat, long,lat_ref, long_ref)) %>%
+    mutate(
+      long_ref = long_ref,
+      lat_ref = lat_ref,
+      dist = rnoaa::meteo_spherical_distance(lat, long, lat_ref, long_ref)
+    ) %>%
     slice_min(dist, n = n) %>%
@@
 
-(stations <- do.call("rbind", purrr::map(1:nrow(latlon_ref),
-                             ~calc_dist(latlon_ref[.x, ], cand))))
+(stations <- do.call("rbind", purrr::map(
+  1:nrow(latlon_ref),
+  ~ calc_dist(latlon_ref[.x, ], cand)
+)))
 ```
@@
     date_max = "2020-12-31",
-                                       var = c("PRCP", "TMAX", "TMIN")) %>%
+    var = c("PRCP", "TMAX", "TMIN")
+  ) %>%
     select(-id))) %>%

### 2021-11-28 main.tex
add in examples from vignettes

@@
 
+\hypertarget{support-on-interactive-graphics}{%
+\subsection{Support on interactive
+graphics}\label{support-on-interactive-graphics}}
+
 \hypertarget{integrating-into-a-tidy-workflow}{%
@@
 
+This vignette introduces how to perform spatial and temporal manipulate
+in a cubble with dplyr verbs. We will illustrate with
+\texttt{weatherdata::historical\_tmax} data, which have the historical
+maximum temperature recorded for Australian stations with the earliest
+dating back to 1859.
+
+\hypertarget{spatial-manipulation}{%
+\subsubsection{Spatial manipulation}\label{spatial-manipulation}}
+
+\texttt{historical\_tmax} is already in a nested cubble format, which is
+suitable for station-wise manipulation. \texttt{rnoaa} construct the
+station id by prefix \texttt{ASN00} to the Bureau of Meteorology (BOM)
+station number. In
+\href{http://www.bom.gov.au/climate/cdo/about/site-num.shtml}{BOM's
+numbering system}, the 2nd and 3rd digit denotes the state a station is
+located in and this is equivalent to the 7th to 8th digit in our string.
+We can mutate/ filter station in a particular state based on this
+information and here we add a column called \texttt{state\_id} and
+filter out the ones in New South Wales and Victoria (ranging from 46 to
+90).
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> tmax_nested <- weatherdata::historical_tmax %>%
++   mutate(state_id = stringr::str_sub(id, 7, 8)) %>%
++   filter(between(state_id, 46, 90))
+\end{CodeInput}
+\end{CodeChunk}
+
+\hypertarget{temporal-manipulation}{%
+\subsubsection{Temporal manipulation}\label{temporal-manipulation}}
+
+There are some operations in the time dimension we would like to make:
+
+\begin{itemize}
+\tightlist
+\item
+  Extract observations in a particular period, say, those in 1971 to
+  1975 and 2016 to 2020. This can be used to compare the historical and
+  recent climate.
+\item
+  Summarise daily records into monthly to remove sparsity
+\end{itemize}
+
+These time dimension operations can be computed in the long form and
+\texttt{stretch()} converts a nested cubble to a long cubble. From a
+long cubble, we can write the exact same dplyr codes to complete the two
+tasks:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> tmax_long <- tmax_nested %>%
++   stretch() %>%
++   filter(lubridate::year(date) %in% c(1971:1975, 2016:2020)) %>%
++   mutate(
++     month = lubridate::month(date),
++     group = as.factor(ifelse(
++       lubridate::year(date) > 2015,
++       "2016 ~ 2020",
++       "1971 ~ 1975"
++     ))
++   ) %>%
++   group_by(month, group) %>%
++   summarise(tmax = mean(tmax, na.rm = TRUE))
+\end{CodeInput}
+\end{CodeChunk}
+
+\hypertarget{back-to-spatial}{%
+\subsubsection{Back to spatial}\label{back-to-spatial}}
+
+A data quality issue with the \texttt{rnoaa} data is that while it
+records the first and last year recorded of each series, it doesn't
+report the period of missingness. For example, station
+\texttt{ASN00047048} starts it first record in 1957, pauses for a period
+from 1963 to 1990, and then resumes it recording till today. Out of the
+77 stations in New South Wales and Victoria, 7 stations have this issue
+and we would like to remove those stations from the comparison.
+
+Again, this is a station-wise operation and to convert back from the
+long cubble to a nested one, use \texttt{tamp()}. Here we keep the
+stations with 24 observations (12 months for both periods) after the
+monthly aggregation.
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> tmax_nested2 <- tmax_long %>%
++   tamp() %>%
++   filter(nrow(ts) == 24)
+\end{CodeInput}
+\end{CodeChunk}
+
+In some visualisation, we may need information from both spatial and
+temporal dimension. One example of this is a glyph map, where spatial
+variables, i.e.~\texttt{longitude} and \texttt{latitude}, are used to
+construct the major axes and temporal variables, i.e.~\texttt{month} and
+\texttt{tmax}, are used to construct the minor axes. This requires these
+variables to be in the same table, rather than in different forms. In
+cubble, you can append spatial variables that are invariant to the
+\texttt{key} withe \texttt{migrate()}
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> tmax_long2 <- tmax_nested2 %>%
++   stretch() %>%
++   migrate(latitude, longitude)
+\end{CodeInput}
+\end{CodeChunk}
+
+\hypertarget{glyph-map}{%
+\subsubsection{Glyph map}\label{glyph-map}}
+
+Now its time for the glyph map!
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> gly <- glyphs(
++   tmax_long2,
++   x_major = "longitude",
++   y_major = "latitude",
++   x_minor = "month",
++   y_minor = "tmax",
++   height = 0.6,
++   width = 1
++ )
+R> 
+R> nsw_vic <- ozmaps::abs_ste %>%
++   filter(NAME %in% c("Victoria", "New South Wales"))
+R> 
+R> plot_map(nsw_vic, fill = "transparent") +
++   geom_path(data = gly, aes(gx, gy,
++     group = interaction(id, group),
++     color = group
++   )) +
++   scale_color_brewer(palette = "Dark2") + 
++   theme_bw() + 
++   coord_sf(xlim = c(140, 155)) + 
++   theme(legend.position = "bottom")
+\end{CodeInput}
+
+\begin{center}\includegraphics{figures/unnamed-chunk-10-1} \end{center}
+
+\end{CodeChunk}
+
+\hypertarget{spatial-and-temporal-aggregation}{%
+\subsection{Spatial and temporal
+aggregation}\label{spatial-and-temporal-aggregation}}
+
+On top of \texttt{aus\_climate} data in this package,
+\texttt{climate\_full} from package
+\href{https://github.com/huizezhang-sherry/weatherdata}{\texttt{weatherdata}}
+has daily climate data of 639 Australia stations from 2016 to 2020. This
+is where these stations locate in an Australia map:
+
+This is a lot of stations to look at for one climate variable and they
+can't all fit into a glyph map. What we can do is to group stations into
+clusters and look at the aggregated series in the glyph map. In this
+vignette, I will introduce how to perform spatial (and temporal)
+aggregation using hierarchical data structure in cubble.
+
+First let's add a new column \texttt{ll} for calculating distance matrix
+in the next section and aggregate the daily precipitation into weekly
+measure. These steps should be familiar from the vignette \emph{Data
+manipulation with cubble}.
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> station_nested <- weatherdata::climate_full %>%
++   mutate(ll = s2::s2_lnglat(long, lat)) %>%
++   stretch() %>%
++   mutate(wk = lubridate::week(date)) %>%
++   group_by(wk) %>%
++   summarise(prcp = sum(prcp, na.rm = TRUE)) %>%
++   tamp()
+\end{CodeInput}
+\end{CodeChunk}
+
+\hypertarget{hierarchical-data-structure}{%
+\subsubsection{Hierarchical data
+structure}\label{hierarchical-data-structure}}
+
+Imposing a clustering structure can be thought of as building a
+hierarchical structure where stations are nested within clusters. As an
+example to illustrate here, we use a kmean clustering algorithm based on
+the distance matrix and specify the number of centers to be 20. More
+complex algorithms can also be used for more complex problem, as long as
+a mapping from each station id to the cluster id can be constructed. We
+then join this data to our station data:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> dist_raw <- scale(s2::s2_distance_matrix(station_nested$ll, station_nested$ll))
+R> 
+R> cluster_res <- tibble(
++   id = station_nested$id,
++   cluster = kmeans(dist_raw, centers = 20, nstart = 50)$cluster
++ )
+R> 
+R> station_nested <- station_nested %>%
++   left_join(cluster_res)
+\end{CodeInput}
+\end{CodeChunk}
+
+One thing we hope to do with the cluster is to find the coordinates of
+the centroid. These are variables variant to the station but invariant
+to the cluster and it would be nice to have a function that structure
+each cluster as a row. \texttt{switch\_key()} is the function that does
+this: it lets you to specify a new key, say \texttt{cluster} and nests
+all spatial variables variant to \texttt{cluster} into a column.
+Temporal observations from different stations while within the same
+cluster are bound in the nested column \texttt{ts}.
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> cluster_nested <- station_nested %>%
++   switch_key(cluster)
+\end{CodeInput}
+\end{CodeChunk}
+
+This structure makes it easy to compute cluster level variable, although
+there are a few steps to calculate the centroid coordinates: we need to
+find the convex hull that wraps around the cluster, make it a polygon,
+find the centroid of the polygon and finally, extract the x and y
+coordinate of each centroid:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> cluster_nested <- cluster_nested %>%
++   mutate(
++     chull = list(chull(.val$long, .val$lat)),
++     ll_cluster = sf::st_as_sfc(
++       s2::s2_make_polygon(c(.val$long[chull]),
++         c(.val$lat[chull]),
++         oriented = FALSE
++       )
++     ),
++     centroid = s2::s2_centroid(ll_cluster),
++     cent_long = s2::s2_x(centroid),
++     cent_lat = s2::s2_y(centroid)
++   )
+\end{CodeInput}
+\end{CodeChunk}
+
+After we have got \texttt{cluster\_nested}, spatial and temporal data at
+both levels can be easily obtained. If we use \texttt{station} and
+\texttt{cluster} prefix to denote the two levels and \texttt{nested} and
+\texttt{long} for whether the data shows the spatial or temporal
+dimension, the relationship among the four datasets can be illustrated
+in the following workflow:
+
+Start with the original \texttt{station\_nested}, \texttt{stretch()}
+expands the \texttt{ts} column with each station (\texttt{id}) forming a
+group and attach variables invariant to \texttt{id} as an attribute.
+\texttt{switch\_key()} changes the \texttt{key} from \texttt{id} to
+\texttt{cluster} and nests all the spatial variables that variant to
+\texttt{cluster}. \texttt{stretch()} \texttt{cluster\_nested} will store
+variables that are invariant to \texttt{cluster} as a tibble in the
+attribute.
+
+Now we can obtain the aggregated series as described in the workflow
+diagram above and construct the glyph map with
+\texttt{GGally::glyphs()}:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> cluster_long <- cluster_nested %>%
++   stretch(ts) %>%
++   group_by(wk) %>%
++   summarise(prcp = sum(prcp, na.rm = TRUE)) %>%
++   migrate(cent_long, cent_lat)
+R> 
+R> gly <- GGally::glyphs(cluster_long,
++   x_major = "cent_long", x_minor = "wk",
++   y_major = "cent_lat", y_minor = "prcp",
++   height = 2, width = 4
++ )
+R> 
+R> state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)
+R> plot_map(state_map) +
++   geom_text(
++     data = cluster_nested,
++     aes(x = cent_long, y = cent_lat, label = cluster)
++   ) +
++   geom_path(
++     data = gly,
++     aes(x = gx, y = gy, group = gid)
++   )
+\end{CodeInput}
+
+\begin{center}\includegraphics{figures/unnamed-chunk-15-1} \end{center}
+
+\end{CodeChunk}
+
+We can also look at the precipitation of each individual station within
+the same cluster:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> station_long <- station_nested %>%
++   stretch(ts) %>%
++   migrate(cluster)
+R> station_long %>%
++   ggplot(aes(x = wk, y = prcp, group = id)) +
++   geom_line(alpha = .3) +
++   facet_wrap(vars(cluster), scales = "free_y", ncol = 4) +
++   theme_bw()
+\end{CodeInput}
+\begin{figure}
+
+{\centering \includegraphics{figures/unnamed-chunk-16-1}
+
+}
+
+\caption[shtishisa]{shtishisa}\label{fig:unnamed-chunk-16}
+\end{figure}
+\end{CodeChunk}
+
+Lastly, there is one series on Tasmania island standing out from others,
+lets look at where it is:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> # this part still needs some fixing
+R> tas_latlong <- station_nested %>%
++   filter(lat < -40) %>%
++   mutate(p = max(ts$prcp)) %>%
++   strip_rowwise() %>%
++   filter(p == max(p))
+R> 
+R> state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)
+R> plot_map(state_map) +
++   geom_point(data = station_nested, aes(x = long, y = lat), size = 0.5) +
++   geom_sf(data = cluster_nested, aes(geometry = ll_cluster), fill = "transparent") +
++   geom_point(data = tas_latlong, aes(x = long, y = lat), col = "red")
+\end{CodeInput}
+
+\begin{center}\includegraphics{figures/unnamed-chunk-17-1} \end{center}
+
+\end{CodeChunk}
+
 \hypertarget{matching-precipitation-and-river-level-in-victria-water-gauges}{%
@@
 
-Matching - Vig 3
+The water level data comes from
+\href{http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full}{Bureau
+of Meteorology} and has a copy in \texttt{weatherdata}. Here we extract
+the water course level and add a column annotate this data of type
+\texttt{river}. For the rainfall data, we will still use the
+\texttt{weatherdata::climate\_full}, filtering for Victorian stations in
+2020 should be pretty familiar by now. Again, we first look at where
+these stations are on the map first:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> river <- weatherdata::water %>%
++   stretch() %>%
++   select(date, Water_course_level) %>%
++   tamp() %>%
++   mutate(type = "river")
+R> 
+R> climate <- weatherdata::climate_full %>%
++   filter(between(stringr::str_sub(id, 7, 8), 76, 90)) %>%
++   stretch() %>%
++   filter(lubridate::year(date) == 2020) %>%
++   tamp() %>%
++   mutate(type = "climate")
+R> 
+R> vic_map <- rmapshaper::ms_simplify(ozmaps::abs_ste %>% filter(NAME == "Victoria"))
+R> plot_map(vic_map) +
++   geom_point(
++     data = dplyr::bind_rows(river, climate),
++     aes(x = long, y = lat, color = type)
++   ) +
++   scale_color_brewer(palette = "Dark2") +
++   theme_bw()
+\end{CodeInput}
+
+\begin{center}\includegraphics{figures/unnamed-chunk-18-1} \end{center}
+
+\end{CodeChunk}
+
+\hypertarget{theory}{%
+\subsubsection{Theory}\label{theory}}
+
+Temporal matching checks how spatially matched pairs align temporally.
+We use the following chart to illustrate how the temporal matching
+works:
+
+\begin{CodeChunk}
+\begin{figure}
+
+{\centering \includegraphics{figures/unnamed-chunk-19-1}
+
+}
+
+\caption[shtishisaasdf]{shtishisaasdf}\label{fig:unnamed-chunk-19}
+\end{figure}
+\end{CodeChunk}
+
+For each spatially matched pair, say \texttt{A} and \texttt{a}, we first
+find the largest \texttt{n} points in each series, colored in brown
+points here. Here we use the largest three but you can tune this number
+by \texttt{temporal\_n\_highest}. Then we construct the interval of the
+largest points from one series and see how many points, from the other
+series, fall into the intervals. The series used to construct the
+interval is controlled by \texttt{temporal\_independent} and the window
+size by \texttt{temporal\_window} with a default of 5.
+
+In this illustration, we construct the interval based on series
+\texttt{A} and two of the three peaks from \texttt{a} falls into this
+interval at Time 7 and 27.
+
+\hypertarget{rainfall-translates-into-river-level}{%
+\subsubsection{Rainfall translates into river
+level}\label{rainfall-translates-into-river-level}}
+
+There's another mandatory argument that hasn't been introduced above:
+\texttt{temporal\_var\_to\_match}. This argument controls the variable
+to match and it needs to appear in both the \texttt{major} and
+\texttt{minor} set. In the water level matching example, we match the
+variable \texttt{Water\_course\_level} from \texttt{river} to
+\texttt{prcp} from \texttt{climate}, hence need to manually rename one
+of them to match the other, here we rename \texttt{Water\_course\_level}
+to \texttt{prcp} in \texttt{river}:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> river <- river %>%
++   stretch() %>%
++   rename(prcp = Water_course_level) %>%
++   tamp()
+\end{CodeInput}
+\end{CodeChunk}
+
+Now we use \texttt{match\_sites()} to first pair the weather stations
+with the river gauges spatially and then apply the temporal matching on
+\texttt{prcp}. We will construct the interval based on peaks in
+\texttt{climate} since we would expect a lag effect for precipitation to
+flow into the river and cause a raise in river level, hence
+\texttt{temporal\_independent\ =\ climate}. We select the 30 highest
+peak from the series to construct the match by setting
+\texttt{temporal\_n\_highest\ =\ 30}. This is a tuning parameter and you
+can start with 10\% of the points of one series (here we have daily data
+for a year, 10\% is roughly 30 points). \texttt{temporal\_min\_match}
+filters out pairs don't have enough match and to return all the pairs,
+set \texttt{temporal\_min\_match} to \texttt{0}.
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> res <- match_sites(river, climate,
++   temporal_var_to_match = prcp,
++   temporal_independent = climate,  
++   temporal_n_highest = 30,
++   temporal_min_match = 15
++ )
+R> 
+R> res
+\end{CodeInput}
+\begin{CodeOutput}
+# cubble:   id [8]: nested form
+# bbox:     [144.52, -37.73, 146.06, -36.55]
+# temporal: date [date], prcp [dbl]
+  id          name                lat  long type  ts        .dist .group n_match
+  <chr>       <chr>             <dbl> <dbl> <chr> <list>    <dbl>  <int>   <int>
+1 405234      SEVEN CREEKS @ D~ -36.9  146. river <tibble ~  6.15      5      21
+2 ASN00082042 strathbogie       -36.8  146. clim~ <tibble ~  6.15      5      21
+3 404207      HOLLAND CREEK @ ~ -36.6  146. river <tibble ~  8.54     10      21
+4 ASN00082170 benalla airport   -36.6  146. clim~ <tibble ~  8.54     10      21
+5 230200      MARIBYRNONG RIVE~ -37.7  145. river <tibble ~  6.17      6      19
+6 ASN00086038 essendon airport  -37.7  145. clim~ <tibble ~  6.17      6      19
+7 406213      CAMPASPE RIVER @~ -37.0  145. river <tibble ~  1.84      1      18
+8 ASN00088051 redesdale         -37.0  145. clim~ <tibble ~  1.84      1      18
+\end{CodeOutput}
+\end{CodeChunk}
+
+The output from temporal matching is also a cubble, with additional
+column \texttt{.dist} and \texttt{.group} inherent from spatial matching
+and \texttt{n\_match} for the number of matched temporal peaks. Then you
+can use this output to plot the location of match or to look at the
+series:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> p1 <- plot_map(vic_map) +
++   geom_point(
++     data = res,
++     aes(x = long, y = lat, color = type)
++   ) +
++   ggrepel::geom_label_repel(
++     data = res %>% filter(type == "river"),
++     aes(x = long, y = lat, label = .group)
++   ) +
++   scale_color_brewer(palette = "Dark2") +
++   ggtitle("Victoria") +
++   theme_minimal() +
++   theme(
++     panel.grid.major = element_blank(),
++     panel.grid.minor = element_blank(),
++     legend.position = "bottom"
++   )
+R> 
+R> res_long <- res %>%
++   stretch(ts) %>%
++   migrate(.group, type) %>%
++   mutate(prcp = scale(prcp)[, 1]) %>%
++   switch_key(.group) %>%
++   migrate(type)
+R> 
+R> p2 <- res_long %>%
++   ggplot(aes(
++     x = date, y = prcp,
++     group = id, color = type
++   )) +
++   geom_line() +
++   facet_wrap(vars(.group)) +
++   scale_color_brewer(palette = "Dark2", guide = "none") +
++   theme_bw() +
++   labs(x = "week") +
++   scale_x_date(date_labels = "%b")
+R> 
+R> p1 | p2
+\end{CodeInput}
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{figures/unnamed-chunk-22-1}
+
+}
+
+\caption[..]{...}\label{fig:unnamed-chunk-22}
+\end{figure}
+\end{CodeChunk}
+
+\hypertarget{interative-graphic-with-cubble}{%
+\subsection{Interative graphic with
+cubble}\label{interative-graphic-with-cubble}}
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> set.seed(123)
+R> tmax_data <- weatherdata::climate_full %>%
++   slice_sample(n = 50) %>%
++   stretch() %>%
++   mutate(ym = tsibble::yearmonth(date)) %>%
++   group_by(ym) %>%
++   summarise(
++     tmax = mean(tmax, na.rm = TRUE),
++     tmin = mean(tmin, na.rm = TRUE),
++     prcp = sum(prcp, na.rm = TRUE)
++   ) %>%
++   ungroup(ym) %>%
++   mutate(
++     year = as.factor(year(ym)),
++     month = as.factor(month(ym)),
++     prcp = sqrt(prcp + 0.001)
++   ) %>%
++   tamp() %>%
++   mutate(median_tmax = median(ts$tmax, na.rm = TRUE)) %>%
++   stretch() %>%
++   migrate(median_tmax, long, lat, name) %>%
++   highlight_key(~id)
+R> 
+R> p1 <- tmax_data %>%
++   ggplot(aes(
++     x = month, y = tmax, group = interaction(year, id),
++     color = median_tmax, label = name
++   )) +
++   geom_line() +
++   facet_wrap(vars(fct_reorder(name, -median_tmax))) +
++   theme(legend.position = "none")
+R> 
+R> state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)
+R> p2 <- tmax_data %>%
++   ggplot() +
++   geom_sf(data = state_map, aes(geometry = geometry)) +
++   geom_point(aes(x = long, y = lat, color = median_tmax, label = name)) +
++   theme_void()
+R> 
+R> bscols(
++   ggplotly(p1, width = 800, height = 800, tooltip = "label") %>%
++     highlight(on = "plotly_selected", off = "plotly_deselect", color = "red"),
++   ggplotly(p2, width = 800, height = 800, tooltip = "label") %>%
++     highlight(on = "plotly_selected", off = "plotly_deselect", color = "red")
++ )
+\end{CodeInput}
+\end{CodeChunk}
 
@@
 +     date_max = "2020-12-31",
-+                                        var = c("PRCP", "TMAX", "TMIN")) %>%
++     var = c("PRCP", "TMAX", "TMIN")
++   ) %>%
 +     select(-id))) %>%

### 2021-11-28 main.Rmd
editing before creating a cubble in Section 2

@@
   message = FALSE,
+  cache = TRUE,
   fig.path = "figures/",
@@
 Many data structures have been proposed for spatial (`sf` by @sf) and temporal (`tsibble` by @tsibble) data in the R community, while less has been done for spatio-temporal data.
-The lack of such tools could potentially  because analysts usually treat the spatial and temporal dimension pf the data separately, without realising the need to create a new data structure.
+The lack of such tools could potentially  because analysts usually treat the spatial and temporal dimension of the data separately, without realising the need to create a new data structure.
 While this approach follows the third tidy data principal [@tidydata] (*Each type of observational unit forms a table*), analysts always need to manually join results from different observational units or combining multiple tables into one for downstream analysis.
@@
 Spatio-temporal data usually come in various forms and Figure \ref{fig:cubble-diagram} shows four examples of this.
-No matter whichever form the data is in, there are always some common components shared by these data.
-A spatial identifier (`id` in the diagram) identifies each site.
-The temporal identifier (`t` in the diagram) [...].
-Coordinates, comprising of latitude and longitude, are commonly used variables for point pattern data.These identifiers will be the building blocks for the data structure introduced below.
-For other variables, those invariant at each time stamp are spatial variables and those differ are temporal variables.
+No matter which form the data is in, these formats share some common components that characterise spatio-temporal data.
+A spatial identifier (`id` in the diagram) is the unique identifier of each site.
+The temporal identifier (`t` in the diagram) prescribes the time stamp each site is recorded.
+Coordinates, comprising of latitude and longitude (`lon` and `lat` in the diagram), locates each site on the map.
+These identifiers will be the building blocks for the data structure introduced below.
+Other variables in the data can be categorised into two groups: spatial variables that are invariant at each time stamp for every site, i.e.
+the name or code of the weather station and temporal variables that varies with time.
 
@@
 
-In a cubble, there are two forms: 1) nested form, for manipulating the spatial dimension, and 2) long form, for manipulating the temporal dimension.
-Figure \ref{fig:def-cubble} sketches the two forms along with the associated attributes.
-A variable identifies by the spatial identifies can come from manipulating spatial variables itself, or summary of temporal variables.
-The nested cubble is best suited to work with this type of operation, since it defines each spatial unit as a row.
-The spatial variables are directly displayed in columns.
-Temporal variables are nested in a column called `ts` and the underlying rowwise dataframe uses a `group` attributes to ensure each row  is in its own group.
+In a cubble, there are two forms: nested form and long form, and Figure \ref{fig:def-cubble} sketches the two forms along with the associated attributes.
+The decision on which form to use is output-oriented, meaning analysts need to first think about whether the output of a particular operation is identified only by the spatial identifier, or a combination of spatial and temporal identifier.
+The nested cubble is suitable for working with operations that are only identified by site and this type of operation can be a pure manipulation of spatial variables, or a summary of temporal variables by site (i.e.
+the output of counting the number of raining day is only identified by sites and hence should be performed with the nested form).
+Underneath the nested form, a cubble is built from a row-wise dataframe (`rowwise_df`) where each site forms a separate group.
+This structure simplifies the calculation that involves temporal variables by avoiding the use of `map` syntax when working with list-column.
 
-<!-- In this structure, the nested form is suitable for manipulating the spatial dimension of the data and this would include manipulating variables that 1) are inherently spatial and 2) summarise the temporal variables by `key`.
-Under the hood, the nested cubble is built on top of the rowwise dataframe (`rowwise_df`).
-This is design to simplify the code when working with the temporal variables - they are nested into the list-column `ts`.
--->
-
-Temporal operations are suited to be performed in the long cubble as each row is defines as the combination of spatial and temporal identifier.
-This is also the structure that `tsibble` adopts.
-Temporal variables are directly displayed.
-To avoid repeating the same spatial at each temporal unit, all the spatial variables, along with the spatial identifier, are stored as a `spatial` attributes.
-This information is used when switching back to the nested form.
+For those operations whose output involves both a spatial and temporal dimension, long form should be used.
+The long form is identified by both the spatial and temporal identifier and adopts a grouped dataframe (`grouped_df`) to forms each site as a group.
+Spatial variables are stored separately in a `tibble` as an special attribute of the long cubble.
+This design avoids repeating the spatial variables at each time stamp while not dropping information from spatial variables.
 
@@
 `as_cubble` create a `cubble` by supplying the three key components: `key` as the spatial identifier; `index` as the temporal identifier; and a vector of `coords` in the order of longitude first and then latitude.
-The use of `key` and `index` follows the naming convention in `tsibble`.
+The naming of `key` and `index` follows the convention in the `tsibble` package.
 The cubble created by default is in the nested form.
@@
 The temporal variables are all nested in the `ts` column, but it could be useful to know the name these variables.
-The third row in the cubble header shows these names and in this example this includes: precipitation, `prcp`, maximum temporature, `tmax`, and minimum temperature, `tmin`.
+The third row in the cubble header shows these names and in this example this includes: precipitation, `prcp`, maximum temperature, `tmax`, and minimum temperature, `tmin`.
 
@@
 `cubble` is also compatible with `tsibble` in the sense that the original list-column can be a `tbl_ts` object.
-Duplicates and gaps shoudl be first checked before structuring the data into a cubble.
+Duplicates and gaps should be first checked before structuring the data into a cubble.
 If the input data is a `tsibble` object, the long form cubble is also a `tsibble` where users can directly apply time series operations.
@@
 This requires these variables to be in the same table, rather than in different forms.
-In cubble, you can append spatial variables that are invariant to the `key` withe `migrate()`
+In cubble, you can append spatial variables that are invariant to the `key` with `migrate()`
 
@@
 Imposing a clustering structure can be thought of as building a hierarchical structure where stations are nested within clusters.
-As an example to illustrate here, we use a kmean clustering algorithm based on the distance matrix and specify the number of centers to be 20.
+As an example to illustrate here, we use a kmean clustering algorithm based on the distance matrix and specify the number of centres to be 20.
 More complex algorithms can also be used for more complex problem, as long as a mapping from each station id to the cluster id can be constructed.
@@
 The table below shows how some `dplyr` verbs are mapped into the operation in a cubble.
-With the existing grouping on the station, additional groupping can be added with `group_by` and removed with `ungrouped`.
+With the existing grouping on the station, additional grouping can be added with `group_by` and removed with `ungrouped`.
 [talk about why it is useful]

### 2021-11-28 main.tex
editing before creating a cubble in Section 2

@@
 lack of such tools could potentially because analysts usually treat the
-spatial and temporal dimension pf the data separately, without realising
+spatial and temporal dimension of the data separately, without realising
 the need to create a new data structure. While this approach follows the
@@
 Spatio-temporal data usually come in various forms and Figure
-\ref{fig:cubble-diagram} shows four examples of this. No matter
-whichever form the data is in, there are always some common components
-shared by these data. A spatial identifier (\texttt{id} in the diagram)
-identifies each site. The temporal identifier (\texttt{t} in the
-diagram) {[}\ldots{]}. Coordinates, comprising of latitude and
-longitude, are commonly used variables for point pattern data.These
-identifiers will be the building blocks for the data structure
-introduced below. For other variables, those invariant at each time
-stamp are spatial variables and those differ are temporal variables.
+\ref{fig:cubble-diagram} shows four examples of this. No matter which
+form the data is in, these formats share some common components that
+characterise spatio-temporal data. A spatial identifier (\texttt{id} in
+the diagram) is the unique identifier of each site. The temporal
+identifier (\texttt{t} in the diagram) prescribes the time stamp each
+site is recorded. Coordinates, comprising of latitude and longitude
+(\texttt{lon} and \texttt{lat} in the diagram), locates each site on the
+map. These identifiers will be the building blocks for the data
+structure introduced below. Other variables in the data can be
+categorised into two groups: spatial variables that are invariant at
+each time stamp for every site, i.e.~the name or code of the weather
+station and temporal variables that varies with time.
 
@@
 
-In a cubble, there are two forms: 1) nested form, for manipulating the
-spatial dimension, and 2) long form, for manipulating the temporal
-dimension. Figure \ref{fig:def-cubble} sketches the two forms along with
-the associated attributes. A variable identifies by the spatial
-identifies can come from manipulating spatial variables itself, or
-summary of temporal variables. The nested cubble is best suited to work
-with this type of operation, since it defines each spatial unit as a
-row. The spatial variables are directly displayed in columns. Temporal
-variables are nested in a column called \texttt{ts} and the underlying
-rowwise dataframe uses a \texttt{group} attributes to ensure each row is
-in its own group.
-
-Temporal operations are suited to be performed in the long cubble as
-each row is defines as the combination of spatial and temporal
-identifier. This is also the structure that \texttt{tsibble} adopts.
-Temporal variables are directly displayed. To avoid repeating the same
-spatial at each temporal unit, all the spatial variables, along with the
-spatial identifier, are stored as a \texttt{spatial} attributes. This
-information is used when switching back to the nested form.
+In a cubble, there are two forms: nested form and long form, and Figure
+\ref{fig:def-cubble} sketches the two forms along with the associated
+attributes. The decision on which form to use is output-oriented,
+meaning analysts need to first think about whether the output of a
+particular operation is identified only by the spatial identifier, or a
+combination of spatial and temporal identifier. The nested cubble is
+suitable for working with operations that are only identified by site
+and this type of operation can be a pure manipulation of spatial
+variables, or a summary of temporal variables by site (i.e.~the output
+of counting the number of raining day is only identified by sites and
+hence should be performed with the nested form). Underneath the nested
+form, a cubble is built from a row-wise dataframe (\texttt{rowwise\_df})
+where each site forms a separate group. This structure simplifies the
+calculation that involves temporal variables by avoiding the use of
+\texttt{map} syntax when working with list-column.
+
+For those operations whose output involves both a spatial and temporal
+dimension, long form should be used. The long form is identified by both
+the spatial and temporal identifier and adopts a grouped dataframe
+(\texttt{grouped\_df}) to forms each site as a group. Spatial variables
+are stored separately in a \texttt{tibble} as an special attribute of
+the long cubble. This design avoids repeating the spatial variables at
+each time stamp while not dropping information from spatial variables.
 
@@
 \texttt{coords} in the order of longitude first and then latitude. The
-use of \texttt{key} and \texttt{index} follows the naming convention in
-\texttt{tsibble}. The cubble created by default is in the nested form.
-Below is an example of creating a cubble: \newline
+naming of \texttt{key} and \texttt{index} follows the convention in the
+\texttt{tsibble} package. The cubble created by default is in the nested
+form. Below is an example of creating a cubble: \newline
 
@@
 cubble header shows these names and in this example this includes:
-precipitation, \texttt{prcp}, maximum temporature, \texttt{tmax}, and
+precipitation, \texttt{prcp}, maximum temperature, \texttt{tmax}, and
 minimum temperature, \texttt{tmin}.
@@
 that the original list-column can be a \texttt{tbl\_ts} object.
-Duplicates and gaps shoudl be first checked before structuring the data
+Duplicates and gaps should be first checked before structuring the data
 into a cubble. If the input data is a \texttt{tsibble} object, the long
@@
 cubble, you can append spatial variables that are invariant to the
-\texttt{key} withe \texttt{migrate()}
+\texttt{key} with \texttt{migrate()}
 
@@
 example to illustrate here, we use a kmean clustering algorithm based on
-the distance matrix and specify the number of centers to be 20. More
+the distance matrix and specify the number of centres to be 20. More
 complex algorithms can also be used for more complex problem, as long as
@@
 in a cubble. With the existing grouping on the station, additional
-groupping can be added with \texttt{group\_by} and removed with
+grouping can be added with \texttt{group\_by} and removed with
 \texttt{ungrouped}. {[}talk about why it is useful{]}

### 2021-11-28 script.R

@@ -2,6 +2,8 @@ library(gstat)
 library(tidyverse)
 library(cubble)
 
+data(wind)
+
 wind_temporal <- wind %>% 
   pivot_longer(RPT:MAL, names_to = "code", values_to = "wind_speed") %>% 
   mutate(date = lubridate::make_date(year + 1900, month, day)) %>% 

### 2021-11-29 main.Rmd
tweak section 2 more + some cleaning up

@@
 
-\newpage
 ## Create a cubble in the nested form
@@
 
-In the `cubble` header, you can read the name of the `key` variable, bbox, and also the name of variable nested in the `ts` column.
-In this example, the spatial identifier is `id` and the number in the bracket means there are 5 unique `id` in this dataset.
+There are a few information in the `cubble` header: the name of the `key` variable, bbox, and also the name of variable nested in the `ts` column.
+In this example, each site is identifier is `id` and the number in the bracket means there are 5 unique `id` in this dataset.
 The bbox in the second row gives the range of the coordinates.
@@
 
-From a created cubble in the nested form, analysts may want to directly manipulate the temporal variables.
-This would require switching to the long form using `stretch()`.
-The verb `stretch()` switch the cubble from the nested form into a long form.
-Under the hood, it first extracts the spatial variables into a separate tibble to store in the attribute `spatial` and then unnests the `ts` column to show the temporal content:
+The long cubble is suitable to manipulate the time dimension of the data.
+The function `stretch()` switches the nested cubble into the long cubble by first extracts all the spatial variables into a separate tibble and store in the `spatial` attribute and then unnests the `ts` column:
 
@@
 
-Notice here that the third line in the header is changed to reflect the spatial variables stored. This is a format suitable for computing time-wise variables.
+Notice here that the third line in the header now shows the name of  spatial variables rather than the temporal variables.
 
@@
 Manipulation on the spatial and temporal dimension can be an iterative process.
-Many times, we may decide to go back to the nested form after some temporal manipulation.
-The verb to switch a long cubble back to the nested form is `tamp()`:
+Many times, analysts will need to go back and forth between the nested and long cubble.
+The `stretch()` function introduced in the previous section switches a nested cubble into a long cubble and function `tamp()` is its inverse function to switch a long cubble back to the nested cubble:
 
@@
 
-As an output to be supplied to further visualisation or modelling, analysts would usually like the spatial and temporal variables to be in the same table.
-`migrate()` moves the spatial variables in the attribute `spatial` into the long form cubble.
+As a final data output for modelling or visualisation, spatio-temporal data is usually expected to be in a single table.
+Function `migrate()` moves the spatial variables from the `spatial` attribute into the long cubble:
 
@@
 
-## Support on hierarchical structure
-
-`switch_key()`
-
-## Support on interactive graphics
-
-## Integrating into a tidy workflow
+In this workflow described above, data objects come into cubble in the nested form, then various operations on the spatial and temporal dimension can go back and forth between the nested and long form, and finally, the data will come out of cubble in the long form for further modelling or visualisation.
 
@@
 \begin{tabular}{ | m{5em} | m{15cm}| } 
-\textbf{mutate} \\
-\textbf{filter}\\
-\textbf{summarise} \\
-\textbf{select} \\
-\textbf{arrange} \\
-\textbf{rename} \\
-\textbf{left\_join} \\
-\textbf{group\_by} \\
-\textbf{ungroup}\\
+basics & \textbf{mutate}, \textbf{filter}, \textbf{summarise}, \textbf{select}, \textbf{arrange}, \textbf{rename}, \textbf{left\_join} \\
+grouping &  \textbf{group\_by}, \textbf{ungroup}\\
 slice family & \textbf{slice\_head}, \textbf{slice\_tail}, \textbf{slice\_sample}, \textbf{slice\_min} and \textbf{slice\_max} \\
@@
 If the input data is a `tsibble` object, the long form cubble is also a `tsibble` where users can directly apply time series operations.
+
+# Advanced features/ considerations
+
+## Hierarchical structure
+
+`switch_key()`
+
+## Interactive graphics
+
+## Data fusion and matching
 
@@
 
-```{r}
+```{r echo = FALSE}
 gly <- glyphs(
@@
 
-```{r}
+```{r echo = FALSE}
 cluster_long <- cluster_nested %>%
@@
 
-```{r fig.height = 5, fig.cap = "shtishisa"}
+```{r fig.height = 5, fig.cap = "shtishisa", echo = FALSE}
 station_long <- station_nested %>%
@@
 
-```{r}
+```{r echo = FALSE}
 # this part still needs some fixing
@@
 
-```{r}
+```{r echo = FALSE}
 river <- weatherdata::water %>%
@@
 Then you can use this output to plot the location of match or to look at the series:
-```{r out.width="100%", fig.cap="..."}
+```{r echo = FALSE, out.width="100%", fig.cap="..."}
 p1 <- plot_map(vic_map) +
@@
 set.seed(123)
-tmax_data <- weatherdata::climate_full %>%
+dt <- weatherdata::climate_full %>%
   slice_sample(n = 50) %>% 
+  mutate(name = str_replace(name, "\\s*\\([^\\)]+\\)", ""))
+
+tmax_data <- dt %>%
   stretch() %>%
@@
   highlight_key(~id)
+```
 
+```{r eval = FALSE, echo = FALSE}
 p1 <- tmax_data %>%
@@
   geom_line() +
+  scale_x_discrete(breaks = seq(1, 12, 3)) + 
   facet_wrap(vars(fct_reorder(name, -median_tmax))) +
@@
 
-Many spatial and spatio-temporal data structures have been developed by the R-spatial team for both raster and vector spatial data.
-For vector spatial data, which is the focus of this paper, `sf` [@pebesma2018simple] represents spatial vector information with simple features: points, lines, polygons and their multiples.
-Various `st_` function are designed to manipulate these features based on their geometric relationships.
-For spatio-temporal data,  `stars` [@stars] can represent both raster and vector data using multi-dimensional array.
-However, the underlying array structure can be difficult to operate for data analysts who are more familiar with a flat 2D data frame structure used by the tidyverse ecosystem.
-
 In the temporal aspect, the `tsibble` [@tsibbles] structure and its tidyverts ecosystem have provided a [...
@@
 This repetition is unnecessary and would inflate the object size for long series.
-
-# A new data structure for spatio-temporal data
-
-**The main difficulty and challenge**  
-The main difficulty in visualising this type of data is to show information in both space and time dimension with the proper level of details without information overflow.
-This would sometimes require aggregating the time dimension into the proper level or slicing the data into a reasonable number of subset for display.
-In this sense, a data structure that regulates the manipulation spatio-temporal data will benefit the analysis workflow.
-While many implementations focus on manipulating and visualising pure spatial or temporal data, there are not sufficient tools to deal with spatio-temporal data.
-The purpose of this paper is to introduce a spatio-temporal vector data structure for data analysis in R.
-
-To work with spatio-temporal data, analysts can choose to either work separately on each dimension or join the two sets together, however, each approach has its own problem: While is is natural to work separately on each sheet (since spatial and temporal operations usually don't overlap), analysts will need to manually keep the other data frame up to date.
-For example, the following pseudo code illustrates the scenario where once the spatial dataset is filtered for those within Victoria, the temporal dataset needs to be manually updated to reflect this spatial filter.
-
-```{r eval = FALSE}
-spatial_new <- spatial %>% filter(SITES_IN_VICTORIA)
-temporal_new <- temporal %>% filter(id %in% spatial_new$id)
-```
-
-If analysts choose to join the spatial and temporal data together, the joined dataset could be too large since each spatial variable will be repeated at each time stamp for each site.
-Also, recordings of the site ID from different data sources can be slightly different from each other, causing a painful checking and cleaning of site IDs before the join.
-
-\newpage
 
@@
 ```
-
-### Basics
-
-- `stretch`: nest to long form
-- `tamp`: long to nest form
-- `migrate`: move selected spatial variables to the long form.
-- `add_dscrb_prct`: summary stats for missingness
-
-dplyr compatibility:
-
-  - mutate, filter, summarise, select, arrange
-  - group and ungroup: group_by, ungroup
-  - slice family
-
-### Combine two cubbles
-
-  - match river and weather gauges data
-  - involve combining two cubbles
-  - join operations combine the two together by appending more rows but what we really want is to bind rows.
-  - bind rows also doesn't work since we want to bind only when there' s a matching????
-  - introduce bind_join
-
-### Hierarchical structure in cubble
-
-  - hierarchical is common.
-  - Given examples.
-  - Essence: switch between different levels
-  - introduce `switch_key`
-
-# Examples
-
-Daily climate data (prcp, tmax, and tmin) from RNOAA - lots of stations across Australia
-
-An exploratory data analysis questions: What's the climate profile look like in Australia
-
-  - General features: Any general trend/ fluctuation in prcp, tmax, and tmin?
-  - Local features: Any station stands out from the crowd?

### 2021-11-29 main.tex
tweak section 2 more + some cleaning up

@@
 
-\newpage
-
 \hypertarget{create-a-cubble-in-the-nested-form}{%
@@
 
-In the \texttt{cubble} header, you can read the name of the \texttt{key}
-variable, bbox, and also the name of variable nested in the \texttt{ts}
-column. In this example, the spatial identifier is \texttt{id} and the
-number in the bracket means there are 5 unique \texttt{id} in this
-dataset. The bbox in the second row gives the range of the coordinates.
-The temporal variables are all nested in the \texttt{ts} column, but it
-could be useful to know the name these variables. The third row in the
-cubble header shows these names and in this example this includes:
-precipitation, \texttt{prcp}, maximum temperature, \texttt{tmax}, and
-minimum temperature, \texttt{tmin}.
+There are a few information in the \texttt{cubble} header: the name of
+the \texttt{key} variable, bbox, and also the name of variable nested in
+the \texttt{ts} column. In this example, each site is identifier is
+\texttt{id} and the number in the bracket means there are 5 unique
+\texttt{id} in this dataset. The bbox in the second row gives the range
+of the coordinates. The temporal variables are all nested in the
+\texttt{ts} column, but it could be useful to know the name these
+variables. The third row in the cubble header shows these names and in
+this example this includes: precipitation, \texttt{prcp}, maximum
+temperature, \texttt{tmax}, and minimum temperature, \texttt{tmin}.
 
@@
 
-From a created cubble in the nested form, analysts may want to directly
-manipulate the temporal variables. This would require switching to the
-long form using \texttt{stretch()}. The verb \texttt{stretch()} switch
-the cubble from the nested form into a long form. Under the hood, it
-first extracts the spatial variables into a separate tibble to store in
-the attribute \texttt{spatial} and then unnests the \texttt{ts} column
-to show the temporal content:
+The long cubble is suitable to manipulate the time dimension of the
+data. The function \texttt{stretch()} switches the nested cubble into
+the long cubble by first extracts all the spatial variables into a
+separate tibble and store in the \texttt{spatial} attribute and then
+unnests the \texttt{ts} column:
 
@@
 
-Notice here that the third line in the header is changed to reflect the
-spatial variables stored. This is a format suitable for computing
-time-wise variables.
+Notice here that the third line in the header now shows the name of
+spatial variables rather than the temporal variables.
 
@@
 Manipulation on the spatial and temporal dimension can be an iterative
-process. Many times, we may decide to go back to the nested form after
-some temporal manipulation. The verb to switch a long cubble back to the
-nested form is \texttt{tamp()}:
+process. Many times, analysts will need to go back and forth between the
+nested and long cubble. The \texttt{stretch()} function introduced in
+the previous section switches a nested cubble into a long cubble and
+function \texttt{tamp()} is its inverse function to switch a long cubble
+back to the nested cubble:
 
@@
 
-As an output to be supplied to further visualisation or modelling,
-analysts would usually like the spatial and temporal variables to be in
-the same table. \texttt{migrate()} moves the spatial variables in the
-attribute \texttt{spatial} into the long form cubble.
+As a final data output for modelling or visualisation, spatio-temporal
+data is usually expected to be in a single table. Function
+\texttt{migrate()} moves the spatial variables from the \texttt{spatial}
+attribute into the long cubble:
 
@@
 
-\hypertarget{support-on-hierarchical-structure}{%
-\subsection{Support on hierarchical
-structure}\label{support-on-hierarchical-structure}}
-
-\texttt{switch\_key()}
-
-\hypertarget{support-on-interactive-graphics}{%
-\subsection{Support on interactive
-graphics}\label{support-on-interactive-graphics}}
-
-\hypertarget{integrating-into-a-tidy-workflow}{%
-\subsection{Integrating into a tidy
-workflow}\label{integrating-into-a-tidy-workflow}}
+In this workflow described above, data objects come into cubble in the
+nested form, then various operations on the spatial and temporal
+dimension can go back and forth between the nested and long form, and
+finally, the data will come out of cubble in the long form for further
+modelling or visualisation.
 
@@
 time series operations.
+
+\hypertarget{advanced-features-considerations}{%
+\section{Advanced features/
+considerations}\label{advanced-features-considerations}}
+
+\hypertarget{hierarchical-structure}{%
+\subsection{Hierarchical structure}\label{hierarchical-structure}}
+
+\texttt{switch\_key()}
+
+\hypertarget{interactive-graphics}{%
+\subsection{Interactive graphics}\label{interactive-graphics}}
+
+\hypertarget{data-fusion-and-matching}{%
+\subsection{Data fusion and matching}\label{data-fusion-and-matching}}
 
@@
 \begin{CodeChunk}
-\begin{CodeInput}
-R> gly <- glyphs(
-+   tmax_long2,
-+   x_major = "longitude",
-+   y_major = "latitude",
-+   x_minor = "month",
-+   y_minor = "tmax",
-+   height = 0.6,
-+   width = 1
-+ )
-R> 
-R> nsw_vic <- ozmaps::abs_ste %>%
-+   filter(NAME %in% c("Victoria", "New South Wales"))
-R> 
-R> plot_map(nsw_vic, fill = "transparent") +
-+   geom_path(data = gly, aes(gx, gy,
-+     group = interaction(id, group),
-+     color = group
-+   )) +
-+   scale_color_brewer(palette = "Dark2") + 
-+   theme_bw() + 
-+   coord_sf(xlim = c(140, 155)) + 
-+   theme(legend.position = "bottom")
-\end{CodeInput}
 
@@
 \begin{CodeChunk}
-\begin{CodeInput}
-R> cluster_long <- cluster_nested %>%
-+   stretch(ts) %>%
-+   group_by(wk) %>%
-+   summarise(prcp = sum(prcp, na.rm = TRUE)) %>%
-+   migrate(cent_long, cent_lat)
-R> 
-R> gly <- GGally::glyphs(cluster_long,
-+   x_major = "cent_long", x_minor = "wk",
-+   y_major = "cent_lat", y_minor = "prcp",
-+   height = 2, width = 4
-+ )
-R> 
-R> state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)
-R> plot_map(state_map) +
-+   geom_text(
-+     data = cluster_nested,
-+     aes(x = cent_long, y = cent_lat, label = cluster)
-+   ) +
-+   geom_path(
-+     data = gly,
-+     aes(x = gx, y = gy, group = gid)
-+   )
-\end{CodeInput}
 
@@
 \begin{CodeChunk}
-\begin{CodeInput}
-R> station_long <- station_nested %>%
-+   stretch(ts) %>%
-+   migrate(cluster)
-R> station_long %>%
-+   ggplot(aes(x = wk, y = prcp, group = id)) +
-+   geom_line(alpha = .3) +
-+   facet_wrap(vars(cluster), scales = "free_y", ncol = 4) +
-+   theme_bw()
-\end{CodeInput}
 \begin{figure}
@@
 \begin{CodeChunk}
-\begin{CodeInput}
-R> # this part still needs some fixing
-R> tas_latlong <- station_nested %>%
-+   filter(lat < -40) %>%
-+   mutate(p = max(ts$prcp)) %>%
-+   strip_rowwise() %>%
-+   filter(p == max(p))
-R> 
-R> state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)
-R> plot_map(state_map) +
-+   geom_point(data = station_nested, aes(x = long, y = lat), size = 0.5) +
-+   geom_sf(data = cluster_nested, aes(geometry = ll_cluster), fill = "transparent") +
-+   geom_point(data = tas_latlong, aes(x = long, y = lat), col = "red")
-\end{CodeInput}
 
@@
 \begin{CodeChunk}
-\begin{CodeInput}
-R> river <- weatherdata::water %>%
-+   stretch() %>%
-+   select(date, Water_course_level) %>%
-+   tamp() %>%
-+   mutate(type = "river")
-R> 
-R> climate <- weatherdata::climate_full %>%
-+   filter(between(stringr::str_sub(id, 7, 8), 76, 90)) %>%
-+   stretch() %>%
-+   filter(lubridate::year(date) == 2020) %>%
-+   tamp() %>%
-+   mutate(type = "climate")
-R> 
-R> vic_map <- rmapshaper::ms_simplify(ozmaps::abs_ste %>% filter(NAME == "Victoria"))
-R> plot_map(vic_map) +
-+   geom_point(
-+     data = dplyr::bind_rows(river, climate),
-+     aes(x = long, y = lat, color = type)
-+   ) +
-+   scale_color_brewer(palette = "Dark2") +
-+   theme_bw()
-\end{CodeInput}
 
@@
 \begin{CodeChunk}
-\begin{CodeInput}
-R> p1 <- plot_map(vic_map) +
-+   geom_point(
-+     data = res,
-+     aes(x = long, y = lat, color = type)
-+   ) +
-+   ggrepel::geom_label_repel(
-+     data = res %>% filter(type == "river"),
-+     aes(x = long, y = lat, label = .group)
-+   ) +
-+   scale_color_brewer(palette = "Dark2") +
-+   ggtitle("Victoria") +
-+   theme_minimal() +
-+   theme(
-+     panel.grid.major = element_blank(),
-+     panel.grid.minor = element_blank(),
-+     legend.position = "bottom"
-+   )
-R> 
-R> res_long <- res %>%
-+   stretch(ts) %>%
-+   migrate(.group, type) %>%
-+   mutate(prcp = scale(prcp)[, 1]) %>%
-+   switch_key(.group) %>%
-+   migrate(type)
-R> 
-R> p2 <- res_long %>%
-+   ggplot(aes(
-+     x = date, y = prcp,
-+     group = id, color = type
-+   )) +
-+   geom_line() +
-+   facet_wrap(vars(.group)) +
-+   scale_color_brewer(palette = "Dark2", guide = "none") +
-+   theme_bw() +
-+   labs(x = "week") +
-+   scale_x_date(date_labels = "%b")
-R> 
-R> p1 | p2
-\end{CodeInput}
 \begin{figure}
@@
 R> set.seed(123)
-R> tmax_data <- weatherdata::climate_full %>%
+R> dt <- weatherdata::climate_full %>%
 +   slice_sample(n = 50) %>% 
++   mutate(name = str_replace(name, "\\s*\\([^\\)]+\\)", ""))
+R>  
+R> tmax_data <- dt %>%
 +   stretch() %>%
@@
 +   highlight_key(~id)
-R> 
-R> p1 <- tmax_data %>%
-+   ggplot(aes(
-+     x = month, y = tmax, group = interaction(year, id),
-+     color = median_tmax, label = name
-+   )) +
-+   geom_line() +
-+   facet_wrap(vars(fct_reorder(name, -median_tmax))) +
-+   theme(legend.position = "none")
-R> 
-R> state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)
-R> p2 <- tmax_data %>%
-+   ggplot() +
-+   geom_sf(data = state_map, aes(geometry = geometry)) +
-+   geom_point(aes(x = long, y = lat, color = median_tmax, label = name)) +
-+   theme_void()
-R> 
-R> bscols(
-+   ggplotly(p1, width = 800, height = 800, tooltip = "label") %>%
-+     highlight(on = "plotly_selected", off = "plotly_deselect", color = "red"),
-+   ggplotly(p2, width = 800, height = 800, tooltip = "label") %>%
-+     highlight(on = "plotly_selected", off = "plotly_deselect", color = "red")
-+ )
 \end{CodeInput}
@@
 \section{Old stuff}\label{old-stuff}}
-
-Many spatial and spatio-temporal data structures have been developed by
-the R-spatial team for both raster and vector spatial data. For vector
-spatial data, which is the focus of this paper, \texttt{sf}
-\citep{pebesma2018simple} represents spatial vector information with
-simple features: points, lines, polygons and their multiples. Various
-\texttt{st\_} function are designed to manipulate these features based
-on their geometric relationships. For spatio-temporal data,
-\texttt{stars} \citep{stars} can represent both raster and vector data
-using multi-dimensional array. However, the underlying array structure
-can be difficult to operate for data analysts who are more familiar with
-a flat 2D data frame structure used by the tidyverse ecosystem.
 
@@
 
-\hypertarget{a-new-data-structure-for-spatio-temporal-data}{%
-\section{A new data structure for spatio-temporal
-data}\label{a-new-data-structure-for-spatio-temporal-data}}
-
-\textbf{The main difficulty and challenge}\\
-The main difficulty in visualising this type of data is to show
-information in both space and time dimension with the proper level of
-details without information overflow. This would sometimes require
-aggregating the time dimension into the proper level or slicing the data
-into a reasonable number of subset for display. In this sense, a data
-structure that regulates the manipulation spatio-temporal data will
-benefit the analysis workflow. While many implementations focus on
-manipulating and visualising pure spatial or temporal data, there are
-not sufficient tools to deal with spatio-temporal data. The purpose of
-this paper is to introduce a spatio-temporal vector data structure for
-data analysis in R.
-
-To work with spatio-temporal data, analysts can choose to either work
-separately on each dimension or join the two sets together, however,
-each approach has its own problem: While is is natural to work
-separately on each sheet (since spatial and temporal operations usually
-don't overlap), analysts will need to manually keep the other data frame
-up to date. For example, the following pseudo code illustrates the
-scenario where once the spatial dataset is filtered for those within
-Victoria, the temporal dataset needs to be manually updated to reflect
-this spatial filter.
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> spatial_new <- spatial %>% filter(SITES_IN_VICTORIA)
-R> temporal_new <- temporal %>% filter(id %in% spatial_new$id)
-\end{CodeInput}
-\end{CodeChunk}
-
-If analysts choose to join the spatial and temporal data together, the
-joined dataset could be too large since each spatial variable will be
-repeated at each time stamp for each site. Also, recordings of the site
-ID from different data sources can be slightly different from each
-other, causing a painful checking and cleaning of site IDs before the
-join.
-
-\newpage
-
 \hypertarget{create-a-cubble}{%
@@
 
-\hypertarget{basics}{%
-\subsubsection{Basics}\label{basics}}
-
-\begin{itemize}
-\tightlist
-\item
-  \texttt{stretch}: nest to long form
-\item
-  \texttt{tamp}: long to nest form
-\item
-  \texttt{migrate}: move selected spatial variables to the long form.
-\item
-  \texttt{add\_dscrb\_prct}: summary stats for missingness
-\end{itemize}
-
-dplyr compatibility:
-
-\begin{itemize}
-\tightlist
-\item
-  mutate, filter, summarise, select, arrange
-\item
-  group and ungroup: group\_by, ungroup
-\item
-  slice family
-\end{itemize}
-
-\hypertarget{combine-two-cubbles}{%
-\subsubsection{Combine two cubbles}\label{combine-two-cubbles}}
-
-\begin{itemize}
-\tightlist
-\item
-  match river and weather gauges data
-\item
-  involve combining two cubbles
-\item
-  join operations combine the two together by appending more rows but
-  what we really want is to bind rows.
-\item
-  bind rows also doesn't work since we want to bind only when there' s a
-  matching????
-\item
-  introduce bind\_join
-\end{itemize}
-
-\hypertarget{hierarchical-structure-in-cubble}{%
-\subsubsection{Hierarchical structure in
-cubble}\label{hierarchical-structure-in-cubble}}
-
-\begin{itemize}
-\tightlist
-\item
-  hierarchical is common.
-\item
-  Given examples.
-\item
-  Essence: switch between different levels
-\item
-  introduce \texttt{switch\_key}
-\end{itemize}
-
-\hypertarget{examples-1}{%
-\section{Examples}\label{examples-1}}
-
-Daily climate data (prcp, tmax, and tmin) from RNOAA - lots of stations
-across Australia
-
-An exploratory data analysis questions: What's the climate profile look
-like in Australia
-
-\begin{itemize}
-\tightlist
-\item
-  General features: Any general trend/ fluctuation in prcp, tmax, and
-  tmin?
-\item
-  Local features: Any station stands out from the crowd?
-\end{itemize}
-
 \bibliography{references.bib}

### 2021-11-29 script.R

@@ -29,3 +29,40 @@ plot_map(ireland) +
   geom_point(data = wind_cubble, aes(x = long, y = lat), shape = "+", color = "red", size = 5) + 
   geom_text(data = wind_cubble, aes(x = long, y = lat-0.05, label = station), size= 2)
 
+######################################
+
+nc <- ncdf4::nc_open(file.choose())
+temp <- as_cubble(nc, vars = "tp")
+
+aus <- temp %>% 
+  stretch() %>% 
+  filter(between(lubridate::date(time), as.Date("2021-11-17"), as.Date("2021-11-23"))) %>% 
+  tamp() %>% 
+  mutate(long = floor(long), lat = floor(lat), id2 = paste0(long, lat)) %>% 
+  switch_key(id2) %>% 
+  filter(s2::s2_within(s2::s2_lnglat(long, lat), sf::st_union(state_map))) 
+
+out <- aus %>% 
+  stretch(ts) %>% 
+  mutate(tp = tp * 39 * 31) %>% 
+  group_by(id) %>% 
+  summarise(tp = sum(tp, na.rm = TRUE)) %>% 
+  ungroup(id) %>% 
+  summarise(tp = mean(tp, na.rm = TRUE))
+
+res <- out %>%
+  migrate(long, lat) %>% 
+  distinct()
+
+state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3) 
+
+state_string <- state_map %>% sf::st_cast("MULTILINESTRING") 
+res %>% 
+  ggplot() +
+  geom_tile(aes(x = long, y = lat, fill = tp)) +
+  geom_sf(data = state_string , alpha = 0.5, fill = "transparent") +
+  theme_bw() + 
+  colorspace::scale_fill_continuous_sequential(
+    "Purple", breaks = seq(0, 150, 25)
+    )
+

### 2021-12-03 main.Rmd
more tweaks of text on examples

@@
 
-**Our new data structure for spatio-temporal data**  
+**Our new data structure for spatio-temporal data**\
 This paper describes the implementation of a new spatio-temporal data structure: `cubble`.
@@
 \newpage
+
 # The cubble package
@@
 
+```{=tex}
 \begin{center}
@@
 \end{center}
-
+```
 `cubble` is also compatible with `tsibble` in the sense that the original list-column can be a `tbl_ts` object.
@@
 
+## Data fusion and matching
+
 ## Interactive graphics
 
-## Data fusion and matching
+Interactive graphics give users the chance to manipulate the graphes for additional information.
+This is especially useful for spatio-temporal data since it would be interesting to explore the correlation between nearby sites on the map or compare between clusters.
+Underlying each interactive graphic system is a pipeline workflow that communicates the user action down the workflow to [...].
+Various such workflow has been proposed in the 90s literature, including [...].
+Current general purpose R implementation on iterative graphics includes: plotly [@plotly] for [...]; ggiraph [@ggiraph] for [...]; crosstalk[@crosstalk] for [...].
+Specifically for spatio-temporal data, there is tmap [@map], leftlet [@leaflet], mapview [@mapview].
+
+some common interactive actions include zoom and panning, tooltip/ highlight, and more advanced cross-linking.
 
@@
 
-Forming a cubble + basic tidyverse verbs - Vig 2
-Aggregation  - Vig 4
+Forming a cubble + basic tidyverse verbs - Vig 2 Aggregation - Vig 4
 
@@
 Then you can use this output to plot the location of match or to look at the series:
+
 ```{r echo = FALSE, out.width="100%", fig.cap="..."}
@@
 
+## Spatio-temporal data with NetCDF
+
+```{r}
+nc <- ncdf4::nc_open(here::here("data/adaptor.mars.internal-1638344933.932894-4194-11-a5d35ac3-b36d-4982-afa5-a5dd193a6718.nc"))
+temp <- as_cubble(nc, vars = "d2m")
+vic_map <- ozmaps::abs_ste %>% filter(NAME == "Victoria")
+
+res <- temp %>% 
+  stretch(ts) %>% 
+  mutate(d2m = d2m - 273.15) %>% 
+  filter(lubridate::hour(time) == 12) %>% 
+  group_by(id) %>% 
+  summarise(d2m = mean(d2m, na.rm= TRUE)) %>%
+  migrate(long, lat)
+
+vic_string <- vic_map %>% sf::st_cast("MULTILINESTRING")
+
+res %>% 
+  ggplot() +
+  geom_tile(aes(x = long, y = lat, fill = d2m)) +
+  geom_sf(data = vic_string , alpha = 0.5, fill = "transparent") +
+  theme_bw() + 
+  colorspace::scale_fill_continuous_sequential("Mint")
+
+```
+
 ## Interative graphic with cubble
 
-```{r eval = FALSE}
-set.seed(123)
-dt <- weatherdata::climate_full %>%
-  slice_sample(n = 50) %>% 
-  mutate(name = str_replace(name, "\\s*\\([^\\)]+\\)", ""))
-
-tmax_data <- dt %>%
-  stretch() %>%
-  mutate(ym = tsibble::yearmonth(date)) %>%
-  group_by(ym) %>%
-  summarise(
-    tmax = mean(tmax, na.rm = TRUE),
-    tmin = mean(tmin, na.rm = TRUE),
-    prcp = sum(prcp, na.rm = TRUE)
-  ) %>%
-  ungroup(ym) %>%
-  mutate(
-    year = as.factor(year(ym)),
-    month = as.factor(month(ym)),
-    prcp = sqrt(prcp + 0.001)
-  ) %>%
-  tamp() %>%
-  mutate(median_tmax = median(ts$tmax, na.rm = TRUE)) %>%
-  stretch() %>%
-  migrate(median_tmax, long, lat, name) %>%
-  highlight_key(~id)
-```
-
-```{r eval = FALSE, echo = FALSE}
-p1 <- tmax_data %>%
-  ggplot(aes(
-    x = month, y = tmax, group = interaction(year, id),
-    color = median_tmax, label = name
-  )) +
-  geom_line() +
-  scale_x_discrete(breaks = seq(1, 12, 3)) + 
-  facet_wrap(vars(fct_reorder(name, -median_tmax))) +
-  theme(legend.position = "none")
-
-state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)
-p2 <- tmax_data %>%
-  ggplot() +
-  geom_sf(data = state_map, aes(geometry = geometry)) +
-  geom_point(aes(x = long, y = lat, color = median_tmax, label = name)) +
-  theme_void()
-
-bscols(
-  ggplotly(p1, width = 800, height = 800, tooltip = "label") %>%
-    highlight(on = "plotly_selected", off = "plotly_deselect", color = "red"),
-  ggplotly(p2, width = 800, height = 800, tooltip = "label") %>%
-    highlight(on = "plotly_selected", off = "plotly_deselect", color = "red")
-)
+```{r echo = FALSE, out.width="100%", out.height="30%"}
+knitr::include_graphics(here::here("figures/linking.png"))
 ```

### 2021-12-03 main.tex
more tweaks of text on examples

@@
 
+\hypertarget{spatio-temporal-data-with-netcdf}{%
+\subsection{Spatio-temporal data with
+NetCDF}\label{spatio-temporal-data-with-netcdf}}
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> nc <- ncdf4::nc_open(here::here("data/adaptor.mars.internal-1638344933.932894-4194-11-a5d35ac3-b36d-4982-afa5-a5dd193a6718.nc"))
+R> temp <- as_cubble(nc, vars = "d2m")
+R> vic_map <- ozmaps::abs_ste %>% filter(NAME == "Victoria")
+R> 
+R> res <- temp %>% 
++   stretch(ts) %>% 
++   mutate(d2m = d2m - 273.15) %>% 
++   filter(lubridate::hour(time) == 12) %>% 
++   group_by(id) %>% 
++   summarise(d2m = mean(d2m, na.rm= TRUE)) %>%
++   migrate(long, lat)
+R> 
+R> vic_string <- vic_map %>% sf::st_cast("MULTILINESTRING") 
+R> 
+R> res %>% 
++   ggplot() +
++   geom_tile(aes(x = long, y = lat, fill = d2m)) +
++   geom_sf(data = vic_string , alpha = 0.5, fill = "transparent") +
++   theme_bw() + 
++   colorspace::scale_fill_continuous_sequential("Mint")
+\end{CodeInput}
+
+\begin{center}\includegraphics{figures/unnamed-chunk-23-1} \end{center}
+
+\end{CodeChunk}
+
 \hypertarget{interative-graphic-with-cubble}{%
@@
 \begin{CodeChunk}
-\begin{CodeInput}
-R> set.seed(123)
-R> dt <- weatherdata::climate_full %>%
-+   slice_sample(n = 50) %>% 
-+   mutate(name = str_replace(name, "\\s*\\([^\\)]+\\)", ""))
-R>  
-R> tmax_data <- dt %>%
-+   stretch() %>%
-+   mutate(ym = tsibble::yearmonth(date)) %>%
-+   group_by(ym) %>%
-+   summarise(
-+     tmax = mean(tmax, na.rm = TRUE),
-+     tmin = mean(tmin, na.rm = TRUE),
-+     prcp = sum(prcp, na.rm = TRUE)
-+   ) %>%
-+   ungroup(ym) %>%
-+   mutate(
-+     year = as.factor(year(ym)),
-+     month = as.factor(month(ym)),
-+     prcp = sqrt(prcp + 0.001)
-+   ) %>%
-+   tamp() %>%
-+   mutate(median_tmax = median(ts$tmax, na.rm = TRUE)) %>%
-+   stretch() %>%
-+   migrate(median_tmax, long, lat, name) %>%
-+   highlight_key(~id)
-\end{CodeInput}
+
+\begin{center}\includegraphics[width=1\linewidth,height=0.3\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/linking} \end{center}
+
 \end{CodeChunk}

### 2021-12-04 main.Rmd
a croase draft for the interactive graphic section

@@
 Interactive graphics give users the chance to manipulate the graphes for additional information.
-This is especially useful for spatio-temporal data since it would be interesting to explore the correlation between nearby sites on the map or compare between clusters.
+This is especially useful for spatio-temporal data since users can zoom or pan the map to see different areas; use tooltips or popups to query more detailed information about a particular point; or explore points in multiple facets through highlight and cross-linking.
+
+Current general purpose R implementation on iterative graphics includes: plotly [@plotly], leaflet [@leaflet], and mapdeck [@mapdeck] uses the underlying javascript library plotly.js, leaflet.js, and mapbox.js; ggiraph [@ggiraph] for [...]; crosstalk [@crosstalk] has developed a cross linking between html widgets.
+There's also more R implement that produces leaflet extension, i.e.
+tmap [@map], mapview [@mapview], leafpop [@leafpop], and leafgl [@leafgl].
+
 Underlying each interactive graphic system is a pipeline workflow that communicates the user action down the workflow to [...].
 Various such workflow has been proposed in the 90s literature, including [...].
-Current general purpose R implementation on iterative graphics includes: plotly [@plotly] for [...]; ggiraph [@ggiraph] for [...]; crosstalk[@crosstalk] for [...].
-Specifically for spatio-temporal data, there is tmap [@map], leftlet [@leaflet], mapview [@mapview].
+These designs do the linking as follows: when an user action is captured, the system searches for the unique identifiers of that observation, and fetches the queried information with the same unique identifiers.
 
-some common interactive actions include zoom and panning, tooltip/ highlight, and more advanced cross-linking.
+With spatial temporal data, a common type of user action is to link between the time series plotted with the site on the map.
+Figure \ref{fig:interactive-workflow} shows how does interactivities work with graphics created by two forms in a cubble.
+With cubble, the time series plot and site map can be created with the long and nested form separately.
+The unique identifier shared by both the nested and long form in a cubble is the `key` variable.
+Once a user action is captured, the system will look for the corresponding key variable, which will highlight the observations in both forms.
+Then graphics will be updated with the highlighted observations.
 
@@
 
+```{r echo = FALSE, out.width="50%", out.height="30%"}
+knitr::include_graphics(here::here("figures/popup.png"))
+```
+
 # Conclusion

### 2021-12-04 main.tex
a croase draft for the interactive graphic section

@@
 
+\hypertarget{data-fusion-and-matching}{%
+\subsection{Data fusion and matching}\label{data-fusion-and-matching}}
+
 \hypertarget{interactive-graphics}{%
@@
 
-\hypertarget{data-fusion-and-matching}{%
-\subsection{Data fusion and matching}\label{data-fusion-and-matching}}
+Interactive graphics give users the chance to manipulate the graphes for
+additional information. This is especially useful for spatio-temporal
+data since users can zoom or pan the map to see different areas; use
+tooltips or popups to query more detailed information about a particular
+point; or explore points in multiple facets through highlight and
+cross-linking.
+
+Current general purpose R implementation on iterative graphics includes:
+plotly \citep{plotly}, leaflet \citep{leaflet}, and mapdeck
+\citep{mapdeck} uses the underlying javascript library plotly.js,
+leaflet.js, and mapbox.js; ggiraph \citep{ggiraph} for {[}\ldots{]};
+crosstalk \citep{crosstalk} has developed a cross linking between html
+widgets. There's also more R implement that produces leaflet extension,
+i.e.~tmap \citep{map}, mapview \citep{mapview}, leafpop \citep{leafpop},
+and leafgl \citep{leafgl}.
+
+Underlying each interactive graphic system is a pipeline workflow that
+communicates the user action down the workflow to {[}\ldots{]}. Various
+such workflow has been proposed in the 90s literature, including
+{[}\ldots{]}. These designs do the linking as follows: when an user
+action is captured, the system searches for the unique identifiers of
+that observation, and fetches the queried information with the same
+unique identifiers.
+
+With spatial temporal data, a common type of user action is to link
+between the time series plotted with the site on the map. Figure
+\ref{fig:interactive-workflow} shows how does interactivities work with
+graphics created by two forms in a cubble. With cubble, the time series
+plot and site map can be created with the long and nested form
+separately. The unique identifier shared by both the nested and long
+form in a cubble is the \texttt{key} variable. Once a user action is
+captured, the system will look for the corresponding key variable, which
+will highlight the observations in both forms. Then graphics will be
+updated with the highlighted observations.
 
@@
 
+\begin{CodeChunk}
+
+\begin{center}\includegraphics[width=0.5\linewidth,height=0.3\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/popup} \end{center}
+
+\end{CodeChunk}
+
 \hypertarget{conclusion}{%

### 2021-12-05 main.Rmd
interactive graphics section before polishing

@@
 
-Many data structures have been proposed for spatial (`sf` by @sf) and temporal (`tsibble` by @tsibble) data in the R community, while less has been done for spatio-temporal data.
+Many data structures have been proposed for spatial (\pkg{sf} by @sf) and temporal (\pkg{tsibble} by @tsibble) data in the R community, while less has been done for spatio-temporal data.
 The lack of such tools could potentially because analysts usually treat the spatial and temporal dimension of the data separately, without realising the need to create a new data structure.
@@
 
-Currently, available spatio-temporal data structure in R includes: `spacetime`[@spacetime], which proposes four space-time layouts: Full grid (STF), sparse grid(STS), irregular (STI), and trajectory (STT).
-The data structure it uses is based on `sp` [@sp] and `xts`[@xts], both of which has been replaced by more recent implementations.
-`spatstat` [@spatstat] implements a `ppp` class for point pattern data; and more recent, `stars` [@stars] implements a spatio-temporal array with the dplyr's data cube structure `cubelyr` [@cubelyr] as its backend.
+Currently, available spatio-temporal data structure in R includes: \pkg{spacetime} [@spacetime], which proposes four space-time layouts: Full grid (STF), sparse grid(STS), irregular (STI), and trajectory (STT).
+The data structure it uses is based on \pkg{sp} [@sp] and \pkg{xts} [@xts], both of which has been replaced by more recent implementations.
+\pkg{spatstat} [@spatstat] implements a `ppp` class for point pattern data; and more recent, \pkg{stars} [@stars] implements a spatio-temporal array with the dplyr's data cube structure \pkg{cubelyr} [@cubelyr] as its backend.
 While these implementations either store spatial and temporal variables all in a single table, hence duplicate the spatial variables for each temporal unit; or split them into two separate tables that has the problem of manually joining, mentioned in the previously.
@@
 This create a gap in the software development.
-The requirement for such a tool is important given the ubiquity of spatio-temporal vector data in the wild: the Ireland wind data from `gstat` is an classic example data that splits variables into spatial (`wind.loc`) and temporal (`wind`) dimension; Bureau of Meteorology (BoM) provides climate observations that are widely applied in agriculture and ecology study; air pollution data.
+The requirement for such a tool is important given the ubiquity of spatio-temporal vector data in the wild: the Ireland wind data from \pkg{gstat} is an classic example data that splits variables into spatial (`wind.loc`) and temporal (`wind`) dimension; Bureau of Meteorology (BoM) provides climate observations that are widely applied in agriculture and ecology study; air pollution data.
 \newline
@@
 **Our new data structure for spatio-temporal data**\
-This paper describes the implementation of a new spatio-temporal data structure: `cubble`.
-`cubble` implements a relational data structure that uses two forms to manage the switch between spatial and temporal dimension.
-With this structure, users can manipulate the spatial or temporal dimension separately, while leaves the linking of two dimensions to `cubble`.
+This paper describes the implementation of a new spatio-temporal data structure: \pkg{cubble}.
+\pkg{cubble} implements a relational data structure that uses two forms to manage the switch between spatial and temporal dimension.
+With this structure, users can manipulate the spatial or temporal dimension separately, while leaves the linking of two dimensions to \pkg{cubble\pkg{.
 The software is available from the Comprehensive R Archive Network (CRAN) at [CRAN link].
@@
 
-```{r cubble-diagram, echo = FALSE, out.width = "100%", out.height = "30%", fig.cap="Illustration of incoming data formats for spatio-temporal data.
+```{r cubble-diagram, echo = FALSE, out.width = "100%", fig.cap="Illustration of incoming data formats for spatio-temporal data.
 (1) Data comes in as a single table; (2) Separate tables for spatial and temporal variables; (3) A single table with all the parameters used to query the database and a separate table for queried data; and (4) Cubical data in array or NetCDF format.", fig.align='center', out.height = "15%"}
-knitr::include_graphics(here::here("figures/input-formats/input-formats.001.png"))
+knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.001.png"))
 ```
@@
 The long form is identified by both the spatial and temporal identifier and adopts a grouped dataframe (`grouped_df`) to forms each site as a group.
-Spatial variables are stored separately in a `tibble` as an special attribute of the long cubble.
+Spatial variables are stored separately in a \pkg{tibble} as an special attribute of the long cubble.
 This design avoids repeating the spatial variables at each time stamp while not dropping information from spatial variables.
 
-```{r def-cubble, echo = FALSE, fig.align="center", out.height="30%", out.width = "100%", fig.cap = "Illustration of nested and long cubble."}
-knitr::include_graphics(here::here("figures/def-cubble/def-cubble.001.png"))
+```{r def-cubble, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "Illustration of nested and long cubble."}
+knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.002.png"))
 ```
@@
 To use functionalities from cubble, data analysts first need to create a cubble.
-`as_cubble` create a `cubble` by supplying the three key components: `key` as the spatial identifier; `index` as the temporal identifier; and a vector of `coords` in the order of longitude first and then latitude.
-The naming of `key` and `index` follows the convention in the `tsibble` package.
+`as_cubble` create a \pkg{cubble} by supplying the three key components: `key` as the spatial identifier; `index` as the temporal identifier; and a vector of `coords` in the order of longitude first and then latitude.
+The naming of `key` and `index` follows the convention in the \pkg{tsibble} package.
 The cubble created by default is in the nested form.
@@
 
-There are a few information in the `cubble` header: the name of the `key` variable, bbox, and also the name of variable nested in the `ts` column.
+There are a few information in the \pkg{cubble} header: the name of the `key` variable, bbox, and also the name of variable nested in the `ts` column.
 In this example, each site is identifier is `id` and the number in the bracket means there are 5 unique `id` in this dataset.
@@
 ```
-`cubble` is also compatible with `tsibble` in the sense that the original list-column can be a `tbl_ts` object.
+\pkg{cubble} is also compatible with \pkg{tsibble} in the sense that the original list-column can be a `tbl_ts` object.
 Duplicates and gaps should be first checked before structuring the data into a cubble.
-If the input data is a `tsibble` object, the long form cubble is also a `tsibble` where users can directly apply time series operations.
+If the input data is a \pkg{tsibble} object, the long form cubble is also a \pkg{tsibble} where users can directly apply time series operations.
 
@@
 Interactive graphics give users the chance to manipulate the graphes for additional information.
-This is especially useful for spatio-temporal data since users can zoom or pan the map to see different areas; use tooltips or popups to query more detailed information about a particular point; or explore points in multiple facets through highlight and cross-linking.
+This is especially useful for spatio-temporal data since users can zoom or pan the map to see more details; use tooltips or popups to query more information about a graphic element; or explore points in multiple facets through highlight and cross-linking.
 
-Current general purpose R implementation on iterative graphics includes: plotly [@plotly], leaflet [@leaflet], and mapdeck [@mapdeck] uses the underlying javascript library plotly.js, leaflet.js, and mapbox.js; ggiraph [@ggiraph] for [...]; crosstalk [@crosstalk] has developed a cross linking between html widgets.
-There's also more R implement that produces leaflet extension, i.e.
-tmap [@map], mapview [@mapview], leafpop [@leafpop], and leafgl [@leafgl].
+Current general purpose R implementations on interactive graphics include: \pkg{plotly} [@plotly], \pkg{leaflet} [@leaflet] that uses the underlying javascript library to realise various interactive actions; \pkg{ggiraph} [@ggiraph] implements tooltip, self-linking, and customised actions specified through \\proglang{javascripts}; \pkg{crosstalk} [@crosstalk] implements brushed linking between htmlwidgets.
+There's also more R implements that produces leaflet extension, i.e.
+\pkg{tmap} [@tmap], \pkg{mapview} [@mapview], and \pkg{leafpop} [@leafpop].
 
-Underlying each interactive graphic system is a pipeline workflow that communicates the user action down the workflow to [...].
-Various such workflow has been proposed in the 90s literature, including [...].
-These designs do the linking as follows: when an user action is captured, the system searches for the unique identifiers of that observation, and fetches the queried information with the same unique identifiers.
+While many graphic implementations present worked examples to illustrate the usage of the package, few documents the design framework that transforms the raw data step-by-step to the final view on the screen.
+This step is called data pipeline and some early work in building a data pipeline includes @buja1988elements, @buja1996interactive, and  @sutherland2000orca.
+@wickham2009plumbing argues that *this pipeline must be presented, in some form, in all graphic software*.
+For interactive graphics, the pipeline also needs to include how this workflow is updated to reflect a captured user action.
+Some good examples of this includes [@xie2014reactive].
+maybe one sentence on @cheng2016enabling.
+In @wickham2009plumbing, they hope to implement a workflow where subseting of the data can be moved to a later stage.
+This would enable the plot to display different subsets of the data.
+@xie2014reactive describes the machanism of category linking with multiple sources through a key variable.
 
+Here we will describe brushed linking between the two forms in cubble.
 With spatial temporal data, a common type of user action is to link between the time series plotted with the site on the map.
@@
 Then graphics will be updated with the highlighted observations.
+
+```{r interactive-workflow, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "demon interactivity"}
+knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.003.png"))
+```
+
+<!-- communicates the user action down the workflow to [...].
+Various such workflow has been proposed in the 90s literature, including [...].
+These designs do the linking as follows: when an user action is)) captured, the system searches for the unique identifiers of that observation, and fetches the queried information with the same unique identifiers.
+-->
+
+linking machenism in crosstalk and leaflet
 
@@
 ```{r cubble-operations, echo = FALSE, fig.cap="Cubble operations", out.height="50%", out.width="90%", fig.align='center'}
-knitr::include_graphics(here::here("figures/cubble-operations/cubble-operations.001.png"))
+knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.004.png"))
 ```

### 2021-12-05 main.tex
interactive graphics section before polishing

@@
 
-Many data structures have been proposed for spatial (\texttt{sf} by
-\citet{sf}) and temporal (\texttt{tsibble} by \citet{tsibble}) data in
-the R community, while less has been done for spatio-temporal data. The
-lack of such tools could potentially because analysts usually treat the
+Many data structures have been proposed for spatial (\pkg{sf} by
+\citet{sf}) and temporal (\pkg{tsibble} by \citet{tsibble}) data in the
+R community, while less has been done for spatio-temporal data. The lack
+of such tools could potentially because analysts usually treat the
 spatial and temporal dimension of the data separately, without realising
@@
 Currently, available spatio-temporal data structure in R includes:
-\texttt{spacetime}\citep{spacetime}, which proposes four space-time
+\pkg{spacetime} \citep{spacetime}, which proposes four space-time
 layouts: Full grid (STF), sparse grid(STS), irregular (STI), and
-trajectory (STT). The data structure it uses is based on \texttt{sp}
-\citep{sp} and \texttt{xts}\citep{xts}, both of which has been replaced
-by more recent implementations. \texttt{spatstat} \citep{spatstat}
-implements a \texttt{ppp} class for point pattern data; and more recent,
-\texttt{stars} \citep{stars} implements a spatio-temporal array with the
-dplyr's data cube structure \texttt{cubelyr} \citep{cubelyr} as its
+trajectory (STT). The data structure it uses is based on \pkg{sp}
+\citep{sp} and \pkg{xts} \citep{xts}, both of which has been replaced by
+more recent implementations. \pkg{spatstat} \citep{spatstat} implements
+a \texttt{ppp} class for point pattern data; and more recent,
+\pkg{stars} \citep{stars} implements a spatio-temporal array with the
+dplyr's data cube structure \pkg{cubelyr} \citep{cubelyr} as its
 backend. While these implementations either store spatial and temporal
@@
 for such a tool is important given the ubiquity of spatio-temporal
-vector data in the wild: the Ireland wind data from \texttt{gstat} is an
+vector data in the wild: the Ireland wind data from \pkg{gstat} is an
 classic example data that splits variables into spatial
@@
 This paper describes the implementation of a new spatio-temporal data
-structure: \texttt{cubble}. \texttt{cubble} implements a relational data
+structure: \pkg{cubble}. \pkg{cubble} implements a relational data
 structure that uses two forms to manage the switch between spatial and
@@
 spatial or temporal dimension separately, while leaves the linking of
-two dimensions to \texttt{cubble}. The software is available from the
-Comprehensive R Archive Network (CRAN) at {[}CRAN link{]}. \newline
+two dimensions to \textbackslash pkg\{cubble\textbackslash pkg\{. The
+software is available from the Comprehensive R Archive Network (CRAN) at
+{[}CRAN link{]}. \newline
 
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.15\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/input-formats/input-formats.001}
+{\centering \includegraphics[width=1\linewidth,height=0.15\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.001}
 
@@
 (\texttt{grouped\_df}) to forms each site as a group. Spatial variables
-are stored separately in a \texttt{tibble} as an special attribute of
-the long cubble. This design avoids repeating the spatial variables at
-each time stamp while not dropping information from spatial variables.
+are stored separately in a \pkg{tibble} as an special attribute of the
+long cubble. This design avoids repeating the spatial variables at each
+time stamp while not dropping information from spatial variables.
 
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.3\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/def-cubble/def-cubble.001}
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.002}
 
@@
 To use functionalities from cubble, data analysts first need to create a
-cubble. \texttt{as\_cubble} create a \texttt{cubble} by supplying the
-three key components: \texttt{key} as the spatial identifier;
-\texttt{index} as the temporal identifier; and a vector of
-\texttt{coords} in the order of longitude first and then latitude. The
-naming of \texttt{key} and \texttt{index} follows the convention in the
-\texttt{tsibble} package. The cubble created by default is in the nested
-form. Below is an example of creating a cubble: \newline
+cubble. \texttt{as\_cubble} create a \pkg{cubble} by supplying the three
+key components: \texttt{key} as the spatial identifier; \texttt{index}
+as the temporal identifier; and a vector of \texttt{coords} in the order
+of longitude first and then latitude. The naming of \texttt{key} and
+\texttt{index} follows the convention in the \pkg{tsibble} package. The
+cubble created by default is in the nested form. Below is an example of
+creating a cubble: \newline
 
@@
 
-There are a few information in the \texttt{cubble} header: the name of
-the \texttt{key} variable, bbox, and also the name of variable nested in
-the \texttt{ts} column. In this example, each site is identifier is
+There are a few information in the \pkg{cubble} header: the name of the
+\texttt{key} variable, bbox, and also the name of variable nested in the
+\texttt{ts} column. In this example, each site is identifier is
 \texttt{id} and the number in the bracket means there are 5 unique
@@
 
-\texttt{cubble} is also compatible with \texttt{tsibble} in the sense
-that the original list-column can be a \texttt{tbl\_ts} object.
-Duplicates and gaps should be first checked before structuring the data
-into a cubble. If the input data is a \texttt{tsibble} object, the long
-form cubble is also a \texttt{tsibble} where users can directly apply
-time series operations.
+\pkg{cubble} is also compatible with \pkg{tsibble} in the sense that the
+original list-column can be a \texttt{tbl\_ts} object. Duplicates and
+gaps should be first checked before structuring the data into a cubble.
+If the input data is a \pkg{tsibble} object, the long form cubble is
+also a \pkg{tsibble} where users can directly apply time series
+operations.
 
@@
 Current general purpose R implementation on iterative graphics includes:
-plotly \citep{plotly}, leaflet \citep{leaflet}, and mapdeck
-\citep{mapdeck} uses the underlying javascript library plotly.js,
-leaflet.js, and mapbox.js; ggiraph \citep{ggiraph} for {[}\ldots{]};
-crosstalk \citep{crosstalk} has developed a cross linking between html
-widgets. There's also more R implement that produces leaflet extension,
-i.e.~tmap \citep{map}, mapview \citep{mapview}, leafpop \citep{leafpop},
-and leafgl \citep{leafgl}.
+\pkg{plotly} \citep{plotly}, \pkg{leaflet} \citep{leaflet}, and
+\pkg{mapdeck} \citep{mapdeck} uses the underlying javascript library
+\pkg{plotly.js}, \pkg{leaflet.js}, and \pkg{mapbox.js}; \pkg{ggiraph}
+\citep{ggiraph} for {[}\ldots{]}; \pkg{crosstalk} \citep{crosstalk} has
+developed a cross linking between html widgets. There's also more R
+implement that produces leaflet extension, i.e.~\pkg{tmap} \citep{tmap},
+\pkg{mapview} \citep{mapview}, and \pkg{leafpop} \citep{leafpop}.
 
-Underlying each interactive graphic system is a pipeline workflow that
-communicates the user action down the workflow to {[}\ldots{]}. Various
-such workflow has been proposed in the 90s literature, including
-{[}\ldots{]}. These designs do the linking as follows: when an user
-action is captured, the system searches for the unique identifiers of
-that observation, and fetches the queried information with the same
-unique identifiers.
+While many graphic implementations present the design framework and
+presents worked examples, few demonstrates how the design framework
+transforms the raw data step-by-step to the final view on the screen.
+This step is called data pipeline and some early work in building a data
+pipeline includes \citet{buja1988elements}, \citet{buja1996interactive},
+and \citet{sutherland2000orca}. \citet{wickham2009plumbing} argues that
+\emph{this pipeline must be presented, in some form, in all graphic
+software}. For interactive graphics, the pipeline also needs to include
+how this workflow is updated to reflect a captured user action. Some
+good examples of this includes \citep{xie2014reactive}. maybe one
+sentence on \citet{cheng2016enabling}. In \citet{wickham2009plumbing},
+they hope to implement a workflow where subseting of the data can be
+moved to a later stage. This would enable the plot to display different
+subsets of the data. \citet{xie2014reactive} describes the machanism of
+category linking with multiple sources through a key variable.
 
+Here we will describe brushed linking between the two forms in cubble.
 With spatial temporal data, a common type of user action is to link
@@
 updated with the highlighted observations.
+
+\begin{CodeChunk}
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.003}
+
+}
+
+\caption[demon interactivity]{demon interactivity}\label{fig:interactive-workflow}
+\end{figure}
+\end{CodeChunk}
+
+linking machenism in crosstalk and leaflet
 
@@
 
-{\centering \includegraphics[width=0.9\linewidth,height=0.5\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/cubble-operations/cubble-operations.001}
+{\centering \includegraphics[width=0.9\linewidth,height=0.5\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.004}
 

### 2021-12-06 main.Rmd
interactive graphic section finishes

@@
 
-**Our new data structure for spatio-temporal data**\
+**Our new data structure for spatio-temporal data**
 This paper describes the implementation of a new spatio-temporal data structure: \pkg{cubble}.
@@
 
-Interactive graphics give users the chance to manipulate the graphes for additional information.
-This is especially useful for spatio-temporal data since users can zoom or pan the map to see more details; use tooltips or popups to query more information about a graphic element; or explore points in multiple facets through highlight and cross-linking.
+Interactive graphics can listen to users' actions on the plot to provide additional information that facilitates data exploration.
+This is a useful technique for spatio-temporal data since users can zoom or pan the map to view the local and global structure of the map; use tooltips or popups to query more information about a graphic element; or highlight points to explore its linked views in other plots.
+In the R community, many implementations have been developed to connect \proglang{R} to \proglang{javascript} to create interactive graphics.
+In relation to spatio, temporal, and spatio-temporal data, the general purpose packages \pkg{plotly} [@plotly] and \pkg{leaflet} [@leaflet] realise various interactive actions through their corresponding javascript libraries.
+\pkg{crosstalk} [@crosstalk] and \pkg{tsibbletalk} [@tsibbletalk] implement brushed linking between htmlwidgets.
+\pkg{ggiraph} [@ggiraph] enables tooltip, self-linking, and customised actions specified through its own \proglang{javascript}.
 
-Current general purpose R implementations on interactive graphics include: \pkg{plotly} [@plotly], \pkg{leaflet} [@leaflet] that uses the underlying javascript library to realise various interactive actions; \pkg{ggiraph} [@ggiraph] implements tooltip, self-linking, and customised actions specified through \\proglang{javascripts}; \pkg{crosstalk} [@crosstalk] implements brushed linking between htmlwidgets.
-There's also more R implements that produces leaflet extension, i.e.
-\pkg{tmap} [@tmap], \pkg{mapview} [@mapview], and \pkg{leafpop} [@leafpop].
+While many graphic implementations present worked examples to illustrate the usage of the package, few documents the underlying pipeline that transforms the raw data step-by-step to the final view on the screen.
+There have been some early work in building the data pipeline for (interactive) graphics [@buja1988elements; @buja1996interactive;  @sutherland2000orca] and more recent discussions include @wickham2009plumbing, @xie2014reactive, and @cheng2016enabling.
 
-While many graphic implementations present worked examples to illustrate the usage of the package, few documents the design framework that transforms the raw data step-by-step to the final view on the screen.
-This step is called data pipeline and some early work in building a data pipeline includes @buja1988elements, @buja1996interactive, and  @sutherland2000orca.
-@wickham2009plumbing argues that *this pipeline must be presented, in some form, in all graphic software*.
-For interactive graphics, the pipeline also needs to include how this workflow is updated to reflect a captured user action.
-Some good examples of this includes [@xie2014reactive].
-maybe one sentence on @cheng2016enabling.
-In @wickham2009plumbing, they hope to implement a workflow where subseting of the data can be moved to a later stage.
-This would enable the plot to display different subsets of the data.
-@xie2014reactive describes the machanism of category linking with multiple sources through a key variable.
-
-Here we will describe brushed linking between the two forms in cubble.
-With spatial temporal data, a common type of user action is to link between the time series plotted with the site on the map.
-Figure \ref{fig:interactive-workflow} shows how does interactivities work with graphics created by two forms in a cubble.
-With cubble, the time series plot and site map can be created with the long and nested form separately.
-The unique identifier shared by both the nested and long form in a cubble is the `key` variable.
-Once a user action is captured, the system will look for the corresponding key variable, which will highlight the observations in both forms.
-Then graphics will be updated with the highlighted observations.
+Figure \ref{fig:interactive-workflow} shows how cross-linking works with the two forms in a cubble.
+The data pipeline flows from the *Augmented Data* to *Transformation*, and then *Subset*, to finally, *Plot*.
+This data pipeline, proposed by @wickham2009plumbing, places the *Subset* stage as the last step before *Plot* so that each plot can be made separately from different subsets.
+This design fits naturally with cubble where the spatial map is solely made from the nested form and the time series plot can be created from the long form.
+When a user action is captured either from the map or the time series, the relevant row will be highlighted in the data and all the observations that shares the same `key` variable (`id`), in both forms, will then be highlighted.
+This type of linking is called categorical linking [@xie2014reactive], which is a one-to-many linking through highlighting observations within the same category.
 
@@
 -->
-
-linking machenism in crosstalk and leaflet
 

### 2021-12-06 main.tex
interactive graphic section finishes

@@
 
-\textbf{Our new data structure for spatio-temporal data}\\
-This paper describes the implementation of a new spatio-temporal data
-structure: \pkg{cubble}. \pkg{cubble} implements a relational data
-structure that uses two forms to manage the switch between spatial and
-temporal dimension. With this structure, users can manipulate the
-spatial or temporal dimension separately, while leaves the linking of
-two dimensions to \textbackslash pkg\{cubble\textbackslash pkg\{. The
+\textbf{Our new data structure for spatio-temporal data} This paper
+describes the implementation of a new spatio-temporal data structure:
+\pkg{cubble}. \pkg{cubble} implements a relational data structure that
+uses two forms to manage the switch between spatial and temporal
+dimension. With this structure, users can manipulate the spatial or
+temporal dimension separately, while leaves the linking of two
+dimensions to \textbackslash pkg\{cubble\textbackslash pkg\{. The
 software is available from the Comprehensive R Archive Network (CRAN) at
@@
 
-Interactive graphics give users the chance to manipulate the graphes for
-additional information. This is especially useful for spatio-temporal
-data since users can zoom or pan the map to see different areas; use
-tooltips or popups to query more detailed information about a particular
-point; or explore points in multiple facets through highlight and
-cross-linking.
-
-Current general purpose R implementation on iterative graphics includes:
-\pkg{plotly} \citep{plotly}, \pkg{leaflet} \citep{leaflet}, and
-\pkg{mapdeck} \citep{mapdeck} uses the underlying javascript library
-\pkg{plotly.js}, \pkg{leaflet.js}, and \pkg{mapbox.js}; \pkg{ggiraph}
-\citep{ggiraph} for {[}\ldots{]}; \pkg{crosstalk} \citep{crosstalk} has
-developed a cross linking between html widgets. There's also more R
-implement that produces leaflet extension, i.e.~\pkg{tmap} \citep{tmap},
-\pkg{mapview} \citep{mapview}, and \pkg{leafpop} \citep{leafpop}.
-
-While many graphic implementations present the design framework and
-presents worked examples, few demonstrates how the design framework
+Interactive graphics can listen to users' actions on the plot to provide
+additional information that facilitates data exploration. This is a
+useful technique for spatio-temporal data since users can zoom or pan
+the map to view the local and global structure of the map; use tooltips
+or popups to query more information about a graphic element; or
+highlight points to explore its linked views in other plots. In the R
+community, many implementations have been developed to connect
+\proglang{R} to \proglang{javascript} to create interactive graphics. In
+relation to spatio, temporal, and spatio-temporal data, the general
+purpose packages \pkg{plotly} \citep{plotly} and \pkg{leaflet}
+\citep{leaflet} realise various interactive actions through their
+corresponding javascript libraries. \pkg{crosstalk} \citep{crosstalk}
+and \pkg{tsibbletalk} \citep{tsibbletalk} implement brushed linking
+between htmlwidgets. \pkg{ggiraph} \citep{ggiraph} enables tooltip,
+self-linking, and customised actions specified through its own
+\proglang{javascript}.
+
+While many graphic implementations present worked examples to illustrate
+the usage of the package, few documents the underlying pipeline that
 transforms the raw data step-by-step to the final view on the screen.
-This step is called data pipeline and some early work in building a data
-pipeline includes \citet{buja1988elements}, \citet{buja1996interactive},
-and \citet{sutherland2000orca}. \citet{wickham2009plumbing} argues that
-\emph{this pipeline must be presented, in some form, in all graphic
-software}. For interactive graphics, the pipeline also needs to include
-how this workflow is updated to reflect a captured user action. Some
-good examples of this includes \citep{xie2014reactive}. maybe one
-sentence on \citet{cheng2016enabling}. In \citet{wickham2009plumbing},
-they hope to implement a workflow where subseting of the data can be
-moved to a later stage. This would enable the plot to display different
-subsets of the data. \citet{xie2014reactive} describes the machanism of
-category linking with multiple sources through a key variable.
-
-Here we will describe brushed linking between the two forms in cubble.
-With spatial temporal data, a common type of user action is to link
-between the time series plotted with the site on the map. Figure
-\ref{fig:interactive-workflow} shows how does interactivities work with
-graphics created by two forms in a cubble. With cubble, the time series
-plot and site map can be created with the long and nested form
-separately. The unique identifier shared by both the nested and long
-form in a cubble is the \texttt{key} variable. Once a user action is
-captured, the system will look for the corresponding key variable, which
-will highlight the observations in both forms. Then graphics will be
-updated with the highlighted observations.
+There have been some early work in building the data pipeline for
+(interactive) graphics
+\citep{buja1988elements, buja1996interactive, sutherland2000orca} and
+more recent discussions include \citet{wickham2009plumbing},
+\citet{xie2014reactive}, and \citet{cheng2016enabling}.
+
+Figure \ref{fig:interactive-workflow} shows how cross-linking works with
+the two forms in a cubble. The data pipeline flows from the
+\emph{Augmented Data} to \emph{Transformation}, and then \emph{Subset},
+to finally, \emph{Plot}. This data pipeline, proposed by
+\citet{wickham2009plumbing}, places the \emph{Subset} stage as the last
+step before \emph{Plot} so that each plot can be made separately from
+different subsets. This design fits naturally with cubble where the
+spatial map is solely made from the nested form and the time series plot
+can be created from the long form. When a user action is captured either
+from the map or the time series, the relevant row will be highlighted in
+the data and all the observations that shares the same \texttt{key}
+variable (\texttt{id}), in both forms, will then be highlighted. This
+type of linking is called categorical linking \citep{xie2014reactive},
+which is a one-to-many linking through highlighting observations within
+the same category.
 
@@
 \end{CodeChunk}
-
-linking machenism in crosstalk and leaflet
 

### 2021-12-06 main.Rmd
small tweak on section heading in examples, fix x axis label

@@
 
-Forming a cubble + basic tidyverse verbs - Vig 2 Aggregation - Vig 4
+### Glyph map
 
@@
 We will illustrate with `weatherdata::historical_tmax` data, which have the historical maximum temperature recorded for Australian stations with the earliest dating back to 1859.
-
-### Spatial manipulation
 
@@
 ```
-
-### Temporal manipulation
 
@@
 ```
-
-### Back to spatial
 
@@
 ```
-
-### Glyph map
 
@@
 
-## Spatial and temporal aggregation
+### Aggregation
 
@@
 ```
-
-### Hierarchical data structure
 
@@
 
-## Matching precipitation and river level in Victria water gauges
+## River level data in Victria water gauges
 
@@
 ```
-
-### Theory
 
@@
 In this illustration, we construct the interval based on series `A` and two of the three peaks from `a` falls into this interval at Time 7 and 27.
-
-### Rainfall translates into river level
 
@@
 
-## Spatio-temporal data with NetCDF
+## ERA5: climate reanalysis data
 
-```{r}
+```{r eval = FALSE}
 nc <- ncdf4::nc_open(here::here("data/adaptor.mars.internal-1638344933.932894-4194-11-a5d35ac3-b36d-4982-afa5-a5dd193a6718.nc"))
-temp <- as_cubble(nc, vars = "d2m")
+
+nc <- ncdf4::nc_open(here::here("data/tas_Amon_ACCESS-CM2_ssp585_r1i1p1f1_gn_20150116-20201216_v20191108.nc"))
+temp <- as_cubble(nc, vars = "tas")
 vic_map <- ozmaps::abs_ste %>% filter(NAME == "Victoria")

### 2021-12-06 main.tex
small tweak on section heading in examples, fix x axis label

@@
 
-Forming a cubble + basic tidyverse verbs - Vig 2 Aggregation - Vig 4
+\hypertarget{glyph-map}{%
+\subsubsection{Glyph map}\label{glyph-map}}
 
@@
 dating back to 1859.
-
-\hypertarget{spatial-manipulation}{%
-\subsubsection{Spatial manipulation}\label{spatial-manipulation}}
 
@@
 \end{CodeChunk}
-
-\hypertarget{temporal-manipulation}{%
-\subsubsection{Temporal manipulation}\label{temporal-manipulation}}
 
@@
 \end{CodeChunk}
-
-\hypertarget{back-to-spatial}{%
-\subsubsection{Back to spatial}\label{back-to-spatial}}
 
@@
 
-\hypertarget{glyph-map}{%
-\subsubsection{Glyph map}\label{glyph-map}}
-
 Now its time for the glyph map!
@@
 
-\hypertarget{spatial-and-temporal-aggregation}{%
-\subsection{Spatial and temporal
-aggregation}\label{spatial-and-temporal-aggregation}}
+\hypertarget{aggregation}{%
+\subsubsection{Aggregation}\label{aggregation}}
 
@@
 \end{CodeChunk}
-
-\hypertarget{hierarchical-data-structure}{%
-\subsubsection{Hierarchical data
-structure}\label{hierarchical-data-structure}}
 
@@
 
-\hypertarget{matching-precipitation-and-river-level-in-victria-water-gauges}{%
-\subsection{Matching precipitation and river level in Victria water
-gauges}\label{matching-precipitation-and-river-level-in-victria-water-gauges}}
+\hypertarget{river-level-data-in-victria-water-gauges}{%
+\subsection{River level data in Victria water
+gauges}\label{river-level-data-in-victria-water-gauges}}
 
@@
 \end{CodeChunk}
-
-\hypertarget{theory}{%
-\subsubsection{Theory}\label{theory}}
 
@@
 interval at Time 7 and 27.
-
-\hypertarget{rainfall-translates-into-river-level}{%
-\subsubsection{Rainfall translates into river
-level}\label{rainfall-translates-into-river-level}}
 
@@
 
-\hypertarget{spatio-temporal-data-with-netcdf}{%
-\subsection{Spatio-temporal data with
-NetCDF}\label{spatio-temporal-data-with-netcdf}}
+\hypertarget{era5-climate-reanalysis-data}{%
+\subsection{ERA5: climate reanalysis
+data}\label{era5-climate-reanalysis-data}}
 
@@
 R> nc <- ncdf4::nc_open(here::here("data/adaptor.mars.internal-1638344933.932894-4194-11-a5d35ac3-b36d-4982-afa5-a5dd193a6718.nc"))
-R> temp <- as_cubble(nc, vars = "d2m")
+R> 
+R> nc <- ncdf4::nc_open(here::here("data/tas_Amon_ACCESS-CM2_ssp585_r1i1p1f1_gn_20150116-20201216_v20191108.nc"))
+R> temp <- as_cubble(nc, vars = "tas")
 R> vic_map <- ozmaps::abs_ste %>% filter(NAME == "Victoria")
@@
 \end{CodeInput}
-
-\begin{center}\includegraphics{figures/unnamed-chunk-23-1} \end{center}
-
 \end{CodeChunk}

### 2021-12-06 scripts/popup.R

@@ -17,7 +17,8 @@ long <- weatherdata::climate_full %>%
   ) %>%
   ungroup(ym) %>%
   mutate(month = lubridate::month(ym),
-         year = as.factor(lubridate::year(ym))) %>% 
+         year = as.factor(lubridate::year(ym)),
+         dummy_date = as.Date(glue::glue("2021-{month}-01"))) %>% 
   migrate(name)
 
 nested <- long %>% tamp()
@@ -32,11 +33,12 @@ for (i in 1:length(df_id)) {
   
   p[[i]] <- dt %>% 
     ggplot(aes(
-      x = month, y = tmax, group = year,
+      x = dummy_date, y = tmax, group = year,
       color = year, label = name
     )) +
     geom_point(size = 0.5) + 
     geom_line() +
+    scale_x_date(date_labels = "%b") + 
     scale_color_brewer(palette = "Reds",direction = 1) + 
     theme_bw() + 
     labs(x = "Month", title = name[i]) 

### 2021-12-10 main.Rmd
rearrange example 1 WIP

@@
 
-### Glyph map
+### Basic manipulation
 
-This vignette introduces how to perform spatial and temporal manipulate in a cubble with dplyr verbs.
-We will illustrate with `weatherdata::historical_tmax` data, which have the historical maximum temperature recorded for Australian stations with the earliest dating back to 1859.
+Bureau of Meteorology (BoM) collects climate variables.
+This variables can be openly queried via `rnoaa`.
+`weatherdata` provides a readily available copy of several data.
+`weatherdata::historical_tmax` provides historical maximum temperature recorded for Australian stations with the earliest dating back to 1859 and this can be used to understand the change of maximum temperature from a past period to nowadays.
 
-`historical_tmax` is already in a nested cubble format, which is suitable for station-wise manipulation.
+```{r code1-workflow, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "demon code"}
+knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.004.png"))
+```
+
+<!-- `historical_tmax` is already in a nested cubble format, which is suitable for station-wise manipulation.
 `rnoaa` construct the station id by prefix `ASN00` to the Bureau of Meteorology (BOM) station number.
@@
 We can mutate/ filter station in a particular state based on this information and here we add a column called `state_id` and filter out the ones in New South Wales and Victoria (ranging from 46 to 90).
+-->
 
-```{r}
+<!-- There are some operations in the time dimension we would like to make: -->
+
+<!-- -   Extract observations in a particular period, say, those in 1971 to 1975 and 2016 to 2020. This can be used to compare the historical and recent climate. -->
+<!-- -   Summarise daily records into monthly to remove sparsity -->
+
+<!-- These time dimension operations can be computed in the long form and `stretch()` converts a nested cubble to a long cubble.
+From a long cubble, we can write the exact same dplyr codes to complete the two tasks: -->
+
+<!-- A data quality issue with the `rnoaa` data is that while it records the first and last year recorded of each series, it doesn't report the period of missingness.
+For example, station `ASN00047048` starts it first record in 1957, pauses for a period from 1963 to 1990, and then resumes it recording till today.
+Out of the `nrow(tmax_nested)` stations in New South Wales and Victoria, 7 stations have this issue and we would like to remove those stations from the comparison.
+-->
+
+<!-- Again, this is a station-wise operation and to convert back from the long cubble to a nested one, use `tamp()`.
+Here we keep the stations with 24 observations (12 months for both periods) after the monthly aggregation.
+-->
+
+<!-- In some visualisation, we may need information from both spatial and temporal dimension.
+One example of this is a glyph map, where spatial variables, i.e.
+`longitude` and `latitude`, are used to construct the major axes and temporal variables, i.e.
+`month` and `tmax`, are used to construct the minor axes.
+This requires these variables to be in the same table, rather than in different forms.
+In cubble, you can append spatial variables that are invariant to the `key` with `migrate()` -->
+
+```{r echo = FALSE}
 tmax_nested <- weatherdata::historical_tmax %>%
-  mutate(state_id = stringr::str_sub(id, 7, 8)) %>%
-  filter(between(state_id, 46, 90))
-```
+  filter(between(stringr::str_sub(id, 7, 8), 46, 90))
 
-There are some operations in the time dimension we would like to make:
-
--   Extract observations in a particular period, say, those in 1971 to 1975 and 2016 to 2020. This can be used to compare the historical and recent climate.
--   Summarise daily records into monthly to remove sparsity
-
-These time dimension operations can be computed in the long form and `stretch()` converts a nested cubble to a long cubble.
-From a long cubble, we can write the exact same dplyr codes to complete the two tasks:
-
-```{r}
 tmax_long <- tmax_nested %>%
@@
   summarise(tmax = mean(tmax, na.rm = TRUE))
-```
 
-A data quality issue with the `rnoaa` data is that while it records the first and last year recorded of each series, it doesn't report the period of missingness.
-For example, station `ASN00047048` starts it first record in 1957, pauses for a period from 1963 to 1990, and then resumes it recording till today.
-Out of the `r nrow(tmax_nested)` stations in New South Wales and Victoria, 7 stations have this issue and we would like to remove those stations from the comparison.
-
-Again, this is a station-wise operation and to convert back from the long cubble to a nested one, use `tamp()`.
-Here we keep the stations with 24 observations (12 months for both periods) after the monthly aggregation.
-
-```{r}
 tmax_nested2 <- tmax_long %>%
@@
   filter(nrow(ts) == 24)
-```
 
-In some visualisation, we may need information from both spatial and temporal dimension.
-One example of this is a glyph map, where spatial variables, i.e.
-`longitude` and `latitude`, are used to construct the major axes and temporal variables, i.e.
-`month` and `tmax`, are used to construct the minor axes.
-This requires these variables to be in the same table, rather than in different forms.
-In cubble, you can append spatial variables that are invariant to the `key` with `migrate()`
-
-```{r}
 tmax_long2 <- tmax_nested2 %>%
@@
   migrate(latitude, longitude)
-```
 
-Now its time for the glyph map!
-
-```{r echo = FALSE}
-gly <- glyphs(
-  tmax_long2,
-  x_major = "longitude",
-  y_major = "latitude",
-  x_minor = "month",
-  y_minor = "tmax",
-  height = 0.6,
-  width = 1
-)
+gly <- glyphs(tmax_long2, "longitude", "month", "latitude", "tmax",
+              height = 0.6, width = 1)
 
@@
 plot_map(nsw_vic, fill = "transparent") +
-  geom_path(data = gly, aes(gx, gy,
-    group = interaction(id, group),
-    color = group
-  )) +
+  geom_path(data = gly, 
+            aes(gx, gy, group = interaction(id, group), 
+                color = group)) +
   scale_color_brewer(palette = "Dark2") + 

### 2021-12-10 main.tex
rearrange example 1 WIP

@@
 
-\hypertarget{glyph-map}{%
-\subsubsection{Glyph map}\label{glyph-map}}
+\hypertarget{basic-manipulation}{%
+\subsubsection{Basic manipulation}\label{basic-manipulation}}
 
-This vignette introduces how to perform spatial and temporal manipulate
-in a cubble with dplyr verbs. We will illustrate with
-\texttt{weatherdata::historical\_tmax} data, which have the historical
-maximum temperature recorded for Australian stations with the earliest
-dating back to 1859.
-
-\texttt{historical\_tmax} is already in a nested cubble format, which is
-suitable for station-wise manipulation. \texttt{rnoaa} construct the
-station id by prefix \texttt{ASN00} to the Bureau of Meteorology (BOM)
-station number. In
-\href{http://www.bom.gov.au/climate/cdo/about/site-num.shtml}{BOM's
-numbering system}, the 2nd and 3rd digit denotes the state a station is
-located in and this is equivalent to the 7th to 8th digit in our string.
-We can mutate/ filter station in a particular state based on this
-information and here we add a column called \texttt{state\_id} and
-filter out the ones in New South Wales and Victoria (ranging from 46 to
-90).
+Bureau of Meteorology (BoM) collects climate variables. This variables
+can be openly queried via \texttt{rnoaa}. \texttt{weatherdata} provides
+a readily available copy of several data.
+\texttt{weatherdata::historical\_tmax} provides historical maximum
+temperature recorded for Australian stations with the earliest dating
+back to 1859 and this can be used to understand the change of maximum
+temperature from a past period to nowadays.
 
 \begin{CodeChunk}
-\begin{CodeInput}
-R> tmax_nested <- weatherdata::historical_tmax %>%
-+   mutate(state_id = stringr::str_sub(id, 7, 8)) %>%
-+   filter(between(state_id, 46, 90))
-\end{CodeInput}
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.004}
+
+}
+
+\caption[demon code]{demon code}\label{fig:code1-workflow}
+\end{figure}
 \end{CodeChunk}
-
-There are some operations in the time dimension we would like to make:
-
-\begin{itemize}
-\tightlist
-\item
-  Extract observations in a particular period, say, those in 1971 to
-  1975 and 2016 to 2020. This can be used to compare the historical and
-  recent climate.
-\item
-  Summarise daily records into monthly to remove sparsity
-\end{itemize}
-
-These time dimension operations can be computed in the long form and
-\texttt{stretch()} converts a nested cubble to a long cubble. From a
-long cubble, we can write the exact same dplyr codes to complete the two
-tasks:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> tmax_long <- tmax_nested %>%
-+   stretch() %>%
-+   filter(lubridate::year(date) %in% c(1971:1975, 2016:2020)) %>%
-+   mutate(
-+     month = lubridate::month(date),
-+     group = as.factor(ifelse(
-+       lubridate::year(date) > 2015,
-+       "2016 ~ 2020",
-+       "1971 ~ 1975"
-+     ))
-+   ) %>%
-+   group_by(month, group) %>%
-+   summarise(tmax = mean(tmax, na.rm = TRUE))
-\end{CodeInput}
-\end{CodeChunk}
-
-A data quality issue with the \texttt{rnoaa} data is that while it
-records the first and last year recorded of each series, it doesn't
-report the period of missingness. For example, station
-\texttt{ASN00047048} starts it first record in 1957, pauses for a period
-from 1963 to 1990, and then resumes it recording till today. Out of the
-77 stations in New South Wales and Victoria, 7 stations have this issue
-and we would like to remove those stations from the comparison.
-
-Again, this is a station-wise operation and to convert back from the
-long cubble to a nested one, use \texttt{tamp()}. Here we keep the
-stations with 24 observations (12 months for both periods) after the
-monthly aggregation.
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> tmax_nested2 <- tmax_long %>%
-+   tamp() %>%
-+   filter(nrow(ts) == 24)
-\end{CodeInput}
-\end{CodeChunk}
-
-In some visualisation, we may need information from both spatial and
-temporal dimension. One example of this is a glyph map, where spatial
-variables, i.e.~\texttt{longitude} and \texttt{latitude}, are used to
-construct the major axes and temporal variables, i.e.~\texttt{month} and
-\texttt{tmax}, are used to construct the minor axes. This requires these
-variables to be in the same table, rather than in different forms. In
-cubble, you can append spatial variables that are invariant to the
-\texttt{key} with \texttt{migrate()}
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> tmax_long2 <- tmax_nested2 %>%
-+   stretch() %>%
-+   migrate(latitude, longitude)
-\end{CodeInput}
-\end{CodeChunk}
-
-Now its time for the glyph map!
 
@@
 
-\begin{center}\includegraphics{figures/unnamed-chunk-10-1} \end{center}
+\begin{center}\includegraphics{figures/unnamed-chunk-6-1} \end{center}
 
@@
 
-\begin{center}\includegraphics{figures/unnamed-chunk-15-1} \end{center}
+\begin{center}\includegraphics{figures/unnamed-chunk-11-1} \end{center}
 
@@
 
-{\centering \includegraphics{figures/unnamed-chunk-16-1}
+{\centering \includegraphics{figures/unnamed-chunk-12-1}
 
@@
 
-\caption[shtishisa]{shtishisa}\label{fig:unnamed-chunk-16}
+\caption[shtishisa]{shtishisa}\label{fig:unnamed-chunk-12}
 \end{figure}
@@
 
-\begin{center}\includegraphics{figures/unnamed-chunk-17-1} \end{center}
+\begin{center}\includegraphics{figures/unnamed-chunk-13-1} \end{center}
 
@@
 
-\begin{center}\includegraphics{figures/unnamed-chunk-18-1} \end{center}
+\begin{center}\includegraphics{figures/unnamed-chunk-14-1} \end{center}
 
@@
 
-{\centering \includegraphics{figures/unnamed-chunk-19-1}
+{\centering \includegraphics{figures/unnamed-chunk-15-1}
 
@@
 
-\caption[shtishisaasdf]{shtishisaasdf}\label{fig:unnamed-chunk-19}
+\caption[shtishisaasdf]{shtishisaasdf}\label{fig:unnamed-chunk-15}
 \end{figure}
@@
 
-{\centering \includegraphics[width=1\linewidth]{figures/unnamed-chunk-22-1}
+{\centering \includegraphics[width=1\linewidth]{figures/unnamed-chunk-18-1}
 
@@
 
-\caption[..]{...}\label{fig:unnamed-chunk-22}
+\caption[..]{...}\label{fig:unnamed-chunk-18}
 \end{figure}

### 2021-12-12 main.Rmd
work on writing hierarical structure and example

@@
 knitr::opts_chunk$set(
-  echo = TRUE,
+  echo = FALSE,
   warning = FALSE,
@@
 
-`switch_key()`
+Imposing a clustering structure can be thought of as building a hierarchical structure where stations are nested within clusters.
+This can be useful when there's intrinsic nesting structure in the data (i.e.
+country nested in the continent, county nested in the state) or there's some clustering.
+When we have access to both level, `switch_key()` is the function to re-structure the data as one cluster per row.
+Temporal observations from different stations while within the same cluster are bound in the nested column `ts`.
+
+One thing we hope to do with the cluster is to find the coordinates of the centroid.
+These are variables variant to the station but invariant to the cluster and it would be nice to have a function that structure each cluster as a row.
+`switch_key()` is the function that does this: it lets you to specify a new key, say `cluster` and nests all spatial variables variant to `cluster` into a column.
+Spatial variables are all nested inside a new column `.val`.
+Temporal observations from different stations while within the same cluster are bound in the nested column `ts`.
+
+This structure makes it easy to compute cluster level variable, for example, the convex hull and the centroid coordinate of each cluster.
+These can be amended into the nested form with function `get_centroid`.
+
+After we have got `cluster_nested`, spatial and temporal data at both levels can be easily obtained.
+Figure \ref{fig:hierachical} illustrate the relationship between the long and nested form cubble at both site and cluster level.
+More description on this.
+Start with the original `station_nested`, `stretch()` expands the `ts` column with each station (`id`) forming a group and attach variables invariant to `id` as an attribute.
+`switch_key()` changes the `key` from `id` to `cluster` and nests all the spatial variables that variant to `cluster`.
+`stretch()` `cluster_nested` will store variables that are invariant to `cluster` as a tibble in the attribute.
+
+this fit into the remaining data pipeline.
 
@@
 This variables can be openly queried via `rnoaa`.
-`weatherdata` provides a readily available copy of several data.
-`weatherdata::historical_tmax` provides historical maximum temperature recorded for Australian stations with the earliest dating back to 1859 and this can be used to understand the change of maximum temperature from a past period to nowadays.
+`weatherdata` provides a readily available copy of several data.`weatherdata::historical_tmax` provides historical maximum temperature recorded for Australian stations with the earliest dating back to 1859 and this can be used to understand the change of maximum temperature from a past period to nowadays.
+Figure \ref{fig:glyphmap-workflow} shows the steps involved from the initial data to the final glyph map in Figure \ref{fig:glyphmap}.
 
-```{r code1-workflow, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "demon code"}
+```{r glyphmap-workflow, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "demon code"}
 knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.004.png"))
@@
 
-```{r echo = FALSE}
+```{r echo = FALSE, eval = FALSE}
 tmax_nested <- weatherdata::historical_tmax %>%
@@
 nsw_vic <- ozmaps::abs_ste %>%
-  filter(NAME %in% c("Victoria", "New South Wales"))
+  filter(NAME %in% c("Victoria", "New South Wales")) #
 
@@
   theme_bw() + 
-  coord_sf(xlim = c(140, 155)) + 
+  coord_sf(xlim = c(141, 154)) + 
   theme(legend.position = "bottom")
+
+ggsave(filename = "figures/basic-manipulation.png", width = 10, height = 10)
+```
+
+```{r glyphmap, out.width="100%", out.height="70%", fig.cap="asdfasd"}
+knitr::include_graphics("figures/basic-manipulation.png")
 ```
@@
 
-On top of `aus_climate` data in this package, `climate_full` from package [`weatherdata`](https://github.com/huizezhang-sherry/weatherdata) has daily climate data of `r nrow(weatherdata::climate_full)` Australia stations from 2016 to 2020.
-This is where these stations locate in an Australia map:
+In figure \ref{fig:glyphmap}, there are a few stations around Sydney (151E, 34S) and New castle (152E, 33S).
+Aggregating them into averages or other statistics would give a better picture.
+This can be done with cubble through a hierarchical structure, where the cluster indicator is one level above the individual station.
 
-This is a lot of stations to look at for one climate variable and they can't all fit into a glyph map.
-What we can do is to group stations into clusters and look at the aggregated series in the glyph map.
-In this vignette, I will introduce how to perform spatial (and temporal) aggregation using hierarchical data structure in cubble.
+In Figure \ref{fig:aggregation}
 
-First let's add a new column `ll` for calculating distance matrix in the next section and aggregate the daily precipitation into weekly measure.
-These steps should be familiar from the vignette *Data manipulation with cubble*.
+The `weatherdata::climate_full` data has daily climate data of `r nrow(weatherdata::climate_full)` Australia stations from 2016 to 2020.
 
-```{r}
-station_nested <- weatherdata::climate_full %>%
-  mutate(ll = s2::s2_lnglat(long, lat)) %>%
-  stretch() %>%
-  mutate(wk = lubridate::week(date)) %>%
-  group_by(wk) %>%
-  summarise(prcp = sum(prcp, na.rm = TRUE)) %>%
-  tamp()
-```
+calcluate distance matrix
+As an example to illustrate here, a kmean clustering algorithm based on the distance matrix is used and the number of centres is set to 20.
+More complex algorithms can also be used for more complex problem, as long as a mapping from each station id to the cluster id can be constructed.
 
-Imposing a clustering structure can be thought of as building a hierarchical structure where stations are nested within clusters.
-As an example to illustrate here, we use a kmean clustering algorithm based on the distance matrix and specify the number of centres to be 20.
-More complex algorithms can also be used for more complex problem, as long as a mapping from each station id to the cluster id can be constructed.
-We then join this data to our station data:
+<!-- Now we can obtain the aggregated series as described in the workflow diagram above and construct the glyph map with `GGally::glyphs()`: -->
 
-```{r}
-dist_raw <- scale(s2::s2_distance_matrix(station_nested$ll, station_nested$ll))
+<!-- We can also look at the precipitation of each individual station within the same cluster: -->
 
-cluster_res <- tibble(
-  id = station_nested$id,
-  cluster = kmeans(dist_raw, centers = 20, nstart = 50)$cluster
-)
+<!-- Lastly, there is one series on Tasmania island standing out from others, lets look at where it is: -->
 
-station_nested <- station_nested %>%
-  left_join(cluster_res)
-```
-
-One thing we hope to do with the cluster is to find the coordinates of the centroid.
-These are variables variant to the station but invariant to the cluster and it would be nice to have a function that structure each cluster as a row.
-`switch_key()` is the function that does this: it lets you to specify a new key, say `cluster` and nests all spatial variables variant to `cluster` into a column.
-Temporal observations from different stations while within the same cluster are bound in the nested column `ts`.
-
-```{r}
-cluster_nested <- station_nested %>%
-  switch_key(cluster)
-```
-
-This structure makes it easy to compute cluster level variable, although there are a few steps to calculate the centroid coordinates: we need to find the convex hull that wraps around the cluster, make it a polygon, find the centroid of the polygon and finally, extract the x and y coordinate of each centroid:
-
-```{r}
-cluster_nested <- cluster_nested %>%
-  mutate(
-    chull = list(chull(.val$long, .val$lat)),
-    ll_cluster = sf::st_as_sfc(
-      s2::s2_make_polygon(c(.val$long[chull]),
-        c(.val$lat[chull]),
-        oriented = FALSE
-      )
-    ),
-    centroid = s2::s2_centroid(ll_cluster),
-    cent_long = s2::s2_x(centroid),
-    cent_lat = s2::s2_y(centroid)
-  )
-```
-
-After we have got `cluster_nested`, spatial and temporal data at both levels can be easily obtained.
-If we use `station` and `cluster` prefix to denote the two levels and `nested` and `long` for whether the data shows the spatial or temporal dimension, the relationship among the four datasets can be illustrated in the following workflow:
-
-Start with the original `station_nested`, `stretch()` expands the `ts` column with each station (`id`) forming a group and attach variables invariant to `id` as an attribute.
-`switch_key()` changes the `key` from `id` to `cluster` and nests all the spatial variables that variant to `cluster`.
-`stretch()` `cluster_nested` will store variables that are invariant to `cluster` as a tibble in the attribute.
-
-Now we can obtain the aggregated series as described in the workflow diagram above and construct the glyph map with `GGally::glyphs()`:
-
-```{r echo = FALSE}
-cluster_long <- cluster_nested %>%
-  stretch(ts) %>%
-  group_by(wk) %>%
-  summarise(prcp = sum(prcp, na.rm = TRUE)) %>%
-  migrate(cent_long, cent_lat)
-
-gly <- GGally::glyphs(cluster_long,
-  x_major = "cent_long", x_minor = "wk",
-  y_major = "cent_lat", y_minor = "prcp",
-  height = 2, width = 4
-)
-
-state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)
-plot_map(state_map) +
-  geom_text(
-    data = cluster_nested,
-    aes(x = cent_long, y = cent_lat, label = cluster)
-  ) +
-  geom_path(
-    data = gly,
-    aes(x = gx, y = gy, group = gid)
-  )
-```
-
-We can also look at the precipitation of each individual station within the same cluster:
-
-```{r fig.height = 5, fig.cap = "shtishisa", echo = FALSE}
-station_long <- station_nested %>%
-  stretch(ts) %>%
-  migrate(cluster)
-station_long %>%
-  ggplot(aes(x = wk, y = prcp, group = id)) +
-  geom_line(alpha = .3) +
-  facet_wrap(vars(cluster), scales = "free_y", ncol = 4) +
-  theme_bw()
-```
-
-Lastly, there is one series on Tasmania island standing out from others, lets look at where it is:
-
-```{r echo = FALSE}
-# this part still needs some fixing
-tas_latlong <- station_nested %>%
-  filter(lat < -40) %>%
-  mutate(p = max(ts$prcp)) %>%
-  strip_rowwise() %>%
-  filter(p == max(p))
-
-state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)
-plot_map(state_map) +
-  geom_point(data = station_nested, aes(x = long, y = lat), size = 0.5) +
-  geom_sf(data = cluster_nested, aes(geometry = ll_cluster), fill = "transparent") +
-  geom_point(data = tas_latlong, aes(x = long, y = lat), col = "red")
+```{r aggregation, out.width="100%", out.height="90%", fig.cap="sdfasdf"}
+knitr::include_graphics("figures/aggregation.png")
 ```

### 2021-12-12 main.tex
work on writing hierarical structure and example

@@
 \begin{CodeChunk}
-\begin{CodeInput}
-R> (cubble_nested <- cubble::climate_flat %>%
-+   as_cubble(key = id, index = date, coords = c("long", "lat")))
-\end{CodeInput}
 \begin{CodeOutput}
@@
 \begin{CodeChunk}
-\begin{CodeInput}
-R> (cubble_long <- cubble_nested %>% stretch(ts))
-\end{CodeInput}
 \begin{CodeOutput}
@@
 \begin{CodeChunk}
-\begin{CodeInput}
-R> (cubble_back <- cubble_long %>% tamp())
-\end{CodeInput}
 \begin{CodeOutput}
@@
 \begin{CodeChunk}
-\begin{CodeInput}
-R> (cubble_long %>% migrate(long, lat))
-\end{CodeInput}
 \begin{CodeOutput}
@@
 
-\texttt{switch\_key()}
+Imposing a clustering structure can be thought of as building a
+hierarchical structure where stations are nested within clusters. This
+can be useful when there's intrinsic nesting structure in the data
+(i.e.~country nested in the continent, county nested in the state) or
+there's some clustering. When we have access to both level,
+\texttt{switch\_key()} is the function to re-structure the data as one
+cluster per row. Temporal observations from different stations while
+within the same cluster are bound in the nested column \texttt{ts}.
+
+One thing we hope to do with the cluster is to find the coordinates of
+the centroid. These are variables variant to the station but invariant
+to the cluster and it would be nice to have a function that structure
+each cluster as a row. \texttt{switch\_key()} is the function that does
+this: it lets you to specify a new key, say \texttt{cluster} and nests
+all spatial variables variant to \texttt{cluster} into a column. Spatial
+variables are all nested inside a new column \texttt{.val}. Temporal
+observations from different stations while within the same cluster are
+bound in the nested column \texttt{ts}.
+
+This structure makes it easy to compute cluster level variable, for
+example, the convex hull and the centroid coordinate of each cluster.
+These can be amended into the nested form with function
+\texttt{get\_centroid}.
+
+After we have got \texttt{cluster\_nested}, spatial and temporal data at
+both levels can be easily obtained. Figure \ref{fig:hierachical}
+illustrate the relationship between the long and nested form cubble at
+both site and cluster level. More description on this. Start with the
+original \texttt{station\_nested}, \texttt{stretch()} expands the
+\texttt{ts} column with each station (\texttt{id}) forming a group and
+attach variables invariant to \texttt{id} as an attribute.
+\texttt{switch\_key()} changes the \texttt{key} from \texttt{id} to
+\texttt{cluster} and nests all the spatial variables that variant to
+\texttt{cluster}. \texttt{stretch()} \texttt{cluster\_nested} will store
+variables that are invariant to \texttt{cluster} as a tibble in the
+attribute.
+
+this fit into the remaining data pipeline.
 
@@
 can be openly queried via \texttt{rnoaa}. \texttt{weatherdata} provides
-a readily available copy of several data.
-\texttt{weatherdata::historical\_tmax} provides historical maximum
+a readily available copy of several
+data.\texttt{weatherdata::historical\_tmax} provides historical maximum
 temperature recorded for Australian stations with the earliest dating
 back to 1859 and this can be used to understand the change of maximum
-temperature from a past period to nowadays.
+temperature from a past period to nowadays. Figure
+\ref{fig:glyphmap-workflow} shows the steps involved from the initial
+data to the final glyph map in Figure \ref{fig:glyphmap}.
 
@@
 
-\caption[demon code]{demon code}\label{fig:code1-workflow}
+\caption[demon code]{demon code}\label{fig:glyphmap-workflow}
 \end{figure}
@@
 \begin{CodeChunk}
+\begin{figure}
 
-\begin{center}\includegraphics{figures/unnamed-chunk-6-1} \end{center}
+{\centering \includegraphics[width=1\linewidth,height=0.7\textheight]{figures/basic-manipulation}
 
+}
+
+\caption[asdfasd]{asdfasd}\label{fig:glyphmap}
+\end{figure}
 \end{CodeChunk}
@@
 
-On top of \texttt{aus\_climate} data in this package,
-\texttt{climate\_full} from package
-\href{https://github.com/huizezhang-sherry/weatherdata}{\texttt{weatherdata}}
-has daily climate data of 639 Australia stations from 2016 to 2020. This
-is where these stations locate in an Australia map:
+In figure \ref{fig:glyphmap}, there are a few stations around Sydney
+(151E, 34S) and New castle (152E, 33S). Aggregating them into averages
+or other statistics would give a better picture. This can be done with
+cubble through a hierarchical structure, where the cluster indicator is
+one level above the individual station.
 
-This is a lot of stations to look at for one climate variable and they
-can't all fit into a glyph map. What we can do is to group stations into
-clusters and look at the aggregated series in the glyph map. In this
-vignette, I will introduce how to perform spatial (and temporal)
-aggregation using hierarchical data structure in cubble.
+In Figure \ref{fig:aggregation}
 
-First let's add a new column \texttt{ll} for calculating distance matrix
-in the next section and aggregate the daily precipitation into weekly
-measure. These steps should be familiar from the vignette \emph{Data
-manipulation with cubble}.
+The \texttt{weatherdata::climate\_full} data has daily climate data of
+639 Australia stations from 2016 to 2020.
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> station_nested <- weatherdata::climate_full %>%
-+   mutate(ll = s2::s2_lnglat(long, lat)) %>%
-+   stretch() %>%
-+   mutate(wk = lubridate::week(date)) %>%
-+   group_by(wk) %>%
-+   summarise(prcp = sum(prcp, na.rm = TRUE)) %>%
-+   tamp()
-\end{CodeInput}
-\end{CodeChunk}
-
-Imposing a clustering structure can be thought of as building a
-hierarchical structure where stations are nested within clusters. As an
-example to illustrate here, we use a kmean clustering algorithm based on
-the distance matrix and specify the number of centres to be 20. More
-complex algorithms can also be used for more complex problem, as long as
-a mapping from each station id to the cluster id can be constructed. We
-then join this data to our station data:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> dist_raw <- scale(s2::s2_distance_matrix(station_nested$ll, station_nested$ll))
-R> 
-R> cluster_res <- tibble(
-+   id = station_nested$id,
-+   cluster = kmeans(dist_raw, centers = 20, nstart = 50)$cluster
-+ )
-R> 
-R> station_nested <- station_nested %>%
-+   left_join(cluster_res)
-\end{CodeInput}
-\end{CodeChunk}
-
-One thing we hope to do with the cluster is to find the coordinates of
-the centroid. These are variables variant to the station but invariant
-to the cluster and it would be nice to have a function that structure
-each cluster as a row. \texttt{switch\_key()} is the function that does
-this: it lets you to specify a new key, say \texttt{cluster} and nests
-all spatial variables variant to \texttt{cluster} into a column.
-Temporal observations from different stations while within the same
-cluster are bound in the nested column \texttt{ts}.
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> cluster_nested <- station_nested %>%
-+   switch_key(cluster)
-\end{CodeInput}
-\end{CodeChunk}
-
-This structure makes it easy to compute cluster level variable, although
-there are a few steps to calculate the centroid coordinates: we need to
-find the convex hull that wraps around the cluster, make it a polygon,
-find the centroid of the polygon and finally, extract the x and y
-coordinate of each centroid:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> cluster_nested <- cluster_nested %>%
-+   mutate(
-+     chull = list(chull(.val$long, .val$lat)),
-+     ll_cluster = sf::st_as_sfc(
-+       s2::s2_make_polygon(c(.val$long[chull]),
-+         c(.val$lat[chull]),
-+         oriented = FALSE
-+       )
-+     ),
-+     centroid = s2::s2_centroid(ll_cluster),
-+     cent_long = s2::s2_x(centroid),
-+     cent_lat = s2::s2_y(centroid)
-+   )
-\end{CodeInput}
-\end{CodeChunk}
-
-After we have got \texttt{cluster\_nested}, spatial and temporal data at
-both levels can be easily obtained. If we use \texttt{station} and
-\texttt{cluster} prefix to denote the two levels and \texttt{nested} and
-\texttt{long} for whether the data shows the spatial or temporal
-dimension, the relationship among the four datasets can be illustrated
-in the following workflow:
-
-Start with the original \texttt{station\_nested}, \texttt{stretch()}
-expands the \texttt{ts} column with each station (\texttt{id}) forming a
-group and attach variables invariant to \texttt{id} as an attribute.
-\texttt{switch\_key()} changes the \texttt{key} from \texttt{id} to
-\texttt{cluster} and nests all the spatial variables that variant to
-\texttt{cluster}. \texttt{stretch()} \texttt{cluster\_nested} will store
-variables that are invariant to \texttt{cluster} as a tibble in the
-attribute.
-
-Now we can obtain the aggregated series as described in the workflow
-diagram above and construct the glyph map with
-\texttt{GGally::glyphs()}:
-
-\begin{CodeChunk}
-
-\begin{center}\includegraphics{figures/unnamed-chunk-11-1} \end{center}
-
-\end{CodeChunk}
-
-We can also look at the precipitation of each individual station within
-the same cluster:
+calcluate distance matrix As an example to illustrate here, a kmean
+clustering algorithm based on the distance matrix is used and the number
+of centres is set to 20. More complex algorithms can also be used for
+more complex problem, as long as a mapping from each station id to the
+cluster id can be constructed.
 
@@
 
-{\centering \includegraphics{figures/unnamed-chunk-12-1}
+{\centering \includegraphics[width=1\linewidth,height=0.9\textheight]{figures/aggregation}
 
@@
 
-\caption[shtishisa]{shtishisa}\label{fig:unnamed-chunk-12}
+\caption[sdfasdf]{sdfasdf}\label{fig:aggregation}
 \end{figure}
-\end{CodeChunk}
-
-Lastly, there is one series on Tasmania island standing out from others,
-lets look at where it is:
-
-\begin{CodeChunk}
-
-\begin{center}\includegraphics{figures/unnamed-chunk-13-1} \end{center}
-
 \end{CodeChunk}
@@
 
-\begin{center}\includegraphics{figures/unnamed-chunk-14-1} \end{center}
+\begin{center}\includegraphics{figures/unnamed-chunk-7-1} \end{center}
 
@@
 
-{\centering \includegraphics{figures/unnamed-chunk-15-1}
+{\centering \includegraphics{figures/unnamed-chunk-8-1}
 
@@
 
-\caption[shtishisaasdf]{shtishisaasdf}\label{fig:unnamed-chunk-15}
+\caption[shtishisaasdf]{shtishisaasdf}\label{fig:unnamed-chunk-8}
 \end{figure}
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> river <- river %>%
-+   stretch() %>%
-+   rename(prcp = Water_course_level) %>%
-+   tamp()
-\end{CodeInput}
-\end{CodeChunk}
-
 Now we use \texttt{match\_sites()} to first pair the weather stations
@@
 \begin{CodeChunk}
-\begin{CodeInput}
-R> res <- match_sites(river, climate,
-+   temporal_var_to_match = prcp,
-+   temporal_independent = climate,  
-+   temporal_n_highest = 30,
-+   temporal_min_match = 15
-+ )
-R> 
-R> res
-\end{CodeInput}
 \begin{CodeOutput}
@@
 
-{\centering \includegraphics[width=1\linewidth]{figures/unnamed-chunk-18-1}
+{\centering \includegraphics[width=1\linewidth]{figures/unnamed-chunk-11-1}
 
@@
 
-\caption[..]{...}\label{fig:unnamed-chunk-18}
+\caption[..]{...}\label{fig:unnamed-chunk-11}
 \end{figure}
@@
 data}\label{era5-climate-reanalysis-data}}
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> nc <- ncdf4::nc_open(here::here("data/adaptor.mars.internal-1638344933.932894-4194-11-a5d35ac3-b36d-4982-afa5-a5dd193a6718.nc"))
-R> 
-R> nc <- ncdf4::nc_open(here::here("data/tas_Amon_ACCESS-CM2_ssp585_r1i1p1f1_gn_20150116-20201216_v20191108.nc"))
-R> temp <- as_cubble(nc, vars = "tas")
-R> vic_map <- ozmaps::abs_ste %>% filter(NAME == "Victoria")
-R> 
-R> res <- temp %>% 
-+   stretch(ts) %>% 
-+   mutate(d2m = d2m - 273.15) %>% 
-+   filter(lubridate::hour(time) == 12) %>% 
-+   group_by(id) %>% 
-+   summarise(d2m = mean(d2m, na.rm= TRUE)) %>%
-+   migrate(long, lat)
-R> 
-R> vic_string <- vic_map %>% sf::st_cast("MULTILINESTRING") 
-R> 
-R> res %>% 
-+   ggplot() +
-+   geom_tile(aes(x = long, y = lat, fill = d2m)) +
-+   geom_sf(data = vic_string , alpha = 0.5, fill = "transparent") +
-+   theme_bw() + 
-+   colorspace::scale_fill_continuous_sequential("Mint")
-\end{CodeInput}
-\end{CodeChunk}
 
@@
 \begin{CodeChunk}
-\begin{CodeInput}
-R> climate_flat %>% as_cubble(key = id, index = date, coords = c(long, lat))
-\end{CodeInput}
 \begin{CodeOutput}
@@
 create a cubble, supply the same identifiers as with the first example.
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> sydmel_climate <- stations %>%
-+   rowwise() %>%
-+   mutate(ts = list(meteo_pull_monitors(id,
-+     date_min = "2020-01-01",
-+     date_max = "2020-12-31",
-+     var = c("PRCP", "TMAX", "TMIN")
-+   ) %>%
-+     select(-id))) %>%
-+   as_cubble(key = id, index = date, coords = c(long, lat))
-\end{CodeInput}
-\end{CodeChunk}
 

### 2021-12-12 scripts/example-aggregation.R

@@ -1,0 +1,56 @@
+coords <- cbind(weatherdata::climate_full$long, weatherdata::climate_full$lat)
+dist_raw <- geosphere::distm(coords, coords)
+
+station_nested <- weatherdata::climate_full %>% 
+  strip_rowwise() %>% 
+  mutate(cluster = kmeans(dist_raw,centers = 20, nstart = 500)$cluster)
+
+cluster_nested <- station_nested %>%
+  switch_key(cluster) %>% 
+  get_centroid()
+
+station_long <- cluster_nested %>% 
+  stretch(ts) %>% 
+  mutate(wk = lubridate::week(date)) %>% 
+  group_by(wk, id) %>% 
+  summarise(prcp = sum(prcp, na.rm = TRUE))
+
+cluster_long <- station_long %>% 
+  ungroup(id)
+  summarise(prcp = mean(prcp, na.rm = TRUE)) %>% 
+  migrate(cent_long, cent_lat)
+
+gly <- GGally::glyphs(cluster_long,
+                      x_major = "cent_long", x_minor = "wk",
+                      y_major = "cent_lat", y_minor = "prcp",
+                      height = 2, width = 4
+)
+
+state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)
+p1 <- plot_map(state_map) +
+  geom_text(data = cluster_nested, 
+            aes(x = cent_long, y = cent_lat, label = cluster)) +
+  geom_path(data = gly,aes(x = gx, y = gy, group = gid))
+
+
+p2 <- station_long %>%
+  ggplot(aes(x = wk, y = prcp, group = id)) +
+  geom_line(alpha = .3) +
+  facet_wrap(vars(cluster), scales = "free_y", ncol = 4) +
+  theme_bw()
+
+tas_latlong <- station_long %>% 
+  ungroup(wk, id) %>% 
+  filter(cluster == 15) %>% 
+  filter(prcp == max(prcp)) %>% 
+  tamp()
+
+p3 <- plot_map(state_map) +
+  geom_point(data = station_nested, aes(x = long, y = lat), size = 0.5) +
+  ggforce::geom_mark_hull(data = cluster_nested %>% tidyr::unnest(hull),
+                          expand = 0, radius = 0,
+                          aes(x = long, y = lat, group = cluster)) + 
+  geom_point(data = tas_latlong, aes(x = long, y = lat), col = "red")
+
[…]

### 2021-12-13 main.Rmd
update example diagram on linking, new interactie graphic example, writing up

@@
 
-Interactive graphics can listen to users' actions on the plot to provide additional information that facilitates data exploration.
+<!-- Interactive graphics can listen to users' actions on the plot to provide additional information that facilitates data exploration.
 This is a useful technique for spatio-temporal data since users can zoom or pan the map to view the local and global structure of the map; use tooltips or popups to query more information about a graphic element; or highlight points to explore its linked views in other plots.
@@
 \pkg{ggiraph} [@ggiraph] enables tooltip, self-linking, and customised actions specified through its own \proglang{javascript}.
+-->
 
-While many graphic implementations present worked examples to illustrate the usage of the package, few documents the underlying pipeline that transforms the raw data step-by-step to the final view on the screen.
+<!-- While many graphic implementations present worked examples to illustrate the usage of the package, few documents the underlying pipeline that transforms the raw data step-by-step to the final view on the screen.
 There have been some early work in building the data pipeline for (interactive) graphics [@buja1988elements; @buja1996interactive;  @sutherland2000orca] and more recent discussions include @wickham2009plumbing, @xie2014reactive, and @cheng2016enabling.
+-->
 
-Figure \ref{fig:interactive-workflow} shows how cross-linking works with the two forms in a cubble.
-The data pipeline flows from the *Augmented Data* to *Transformation*, and then *Subset*, to finally, *Plot*.
-This data pipeline, proposed by @wickham2009plumbing, places the *Subset* stage as the last step before *Plot* so that each plot can be made separately from different subsets.
-This design fits naturally with cubble where the spatial map is solely made from the nested form and the time series plot can be created from the long form.
-When a user action is captured either from the map or the time series, the relevant row will be highlighted in the data and all the observations that shares the same `key` variable (`id`), in both forms, will then be highlighted.
-This type of linking is called categorical linking [@xie2014reactive], which is a one-to-many linking through highlighting observations within the same category.
+The cubble structure fits in naturally with the interactive graphic pipeline discussed in the literature [@buja1988elements; @buja1996interactive;  @sutherland2000orca; @xie2014reactive; @cheng2016enabling].
+Diagram \ref{fig:interactive-workflow} illustrates how linking works with the two forms in a cubble, where a time series plot is created with the long cubble and a map is created with the nested cubble.
+When a user action is captured from the map, the site will be activated in the nested cubble.
+Then, the nested cubble will communicate to the long cubble to activate all the observations with the same `id`.
+The long cubble will then highlight the  activated series in the time series plot.
+
+The linking is also available from the time series plot to the map.
+The selection on the time series is through selecting the point on the time series and once a point is selected, it will be activated in the long cubble.
+All the observations that share the same `id`,  either in the long and nested cubble, are then activated.
+This includes other points in the same time series in the long cubble and the corresponding observation of site in the nested cubble.
+These activated observations will then being reflected in the updated plots and Diagram \ref{fig:interactive-workflow-2} in the Appendix illustrates this process.
 
@@
 ```
-
-<!-- communicates the user action down the workflow to [...].
-Various such workflow has been proposed in the 90s literature, including [...].
-These designs do the linking as follows: when an user action is)) captured, the system searches for the unique identifiers of that observation, and fetches the queried information with the same unique identifiers.
--->
 
@@
 
+With a larger dataset, interactive graphics can be helpful to explore these data from multiple perspective.
+Two examples are presented here to show how interactivity can be used to explore the climate data across Australia.
+
 ```{r echo = FALSE, out.width="100%", out.height="30%"}
@@
 # Conclusion
+
+# Appendix
+
+```{r interactive-workflow-2, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "demon interactivity"}
+knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.007.png"))
+```
 

### 2021-12-13 main.tex
update example diagram on linking, new interactie graphic example, writing up

@@
 
-Interactive graphics can listen to users' actions on the plot to provide
-additional information that facilitates data exploration. This is a
-useful technique for spatio-temporal data since users can zoom or pan
-the map to view the local and global structure of the map; use tooltips
-or popups to query more information about a graphic element; or
-highlight points to explore its linked views in other plots. In the R
-community, many implementations have been developed to connect
-\proglang{R} to \proglang{javascript} to create interactive graphics. In
-relation to spatio, temporal, and spatio-temporal data, the general
-purpose packages \pkg{plotly} \citep{plotly} and \pkg{leaflet}
-\citep{leaflet} realise various interactive actions through their
-corresponding javascript libraries. \pkg{crosstalk} \citep{crosstalk}
-and \pkg{tsibbletalk} \citep{tsibbletalk} implement brushed linking
-between htmlwidgets. \pkg{ggiraph} \citep{ggiraph} enables tooltip,
-self-linking, and customised actions specified through its own
-\proglang{javascript}.
+The cubble structure fits in naturally with the interactive graphic
+pipeline discussed in the literature
+\citep{buja1988elements, buja1996interactive, sutherland2000orca, xie2014reactive, cheng2016enabling}.
+Diagram \ref{fig:interactive-workflow} illustrates how linking works
+with the two forms in a cubble, where a time series plot is created with
+the long cubble and a map is created with the nested cubble. When a user
+action is captured from the map, the site will be activated in the
+nested cubble. Then, the nested cubble will communicate to the long
+cubble to activate all the observations with the same \texttt{id}. The
+long cubble will then highlight the activated series in the time series
+plot.
 
-While many graphic implementations present worked examples to illustrate
-the usage of the package, few documents the underlying pipeline that
-transforms the raw data step-by-step to the final view on the screen.
-There have been some early work in building the data pipeline for
-(interactive) graphics
-\citep{buja1988elements, buja1996interactive, sutherland2000orca} and
-more recent discussions include \citet{wickham2009plumbing},
-\citet{xie2014reactive}, and \citet{cheng2016enabling}.
-
-Figure \ref{fig:interactive-workflow} shows how cross-linking works with
-the two forms in a cubble. The data pipeline flows from the
-\emph{Augmented Data} to \emph{Transformation}, and then \emph{Subset},
-to finally, \emph{Plot}. This data pipeline, proposed by
-\citet{wickham2009plumbing}, places the \emph{Subset} stage as the last
-step before \emph{Plot} so that each plot can be made separately from
-different subsets. This design fits naturally with cubble where the
-spatial map is solely made from the nested form and the time series plot
-can be created from the long form. When a user action is captured either
-from the map or the time series, the relevant row will be highlighted in
-the data and all the observations that shares the same \texttt{key}
-variable (\texttt{id}), in both forms, will then be highlighted. This
-type of linking is called categorical linking \citep{xie2014reactive},
-which is a one-to-many linking through highlighting observations within
-the same category.
+The linking is also available from the time series plot to the map. The
+selection on the time series is through selecting the point on the time
+series and once a point is selected, it will be activated in the long
+cubble. All the observations that share the same \texttt{id}, either in
+the long and nested cubble, are then activated. This includes other
+points in the same time series in the long cubble and the corresponding
+observation of site in the nested cubble. These activated observations
+will then being reflected in the updated plots and Diagram
+\ref{fig:interactive-workflow-2} in the Appendix illustrates this
+process.
 
@@
 
+With a larger dataset, interactive graphics can be helpful to explore
+these data from multiple perspective. Two examples are presented here to
+show how interactivity can be used to explore the climate data across
+Australia.
+
 \begin{CodeChunk}
@@
 \section{Conclusion}\label{conclusion}}
+
+\hypertarget{appendix}{%
+\section{Appendix}\label{appendix}}
+
+\begin{CodeChunk}
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.007}
+
+}
+
+\caption[demon interactivity]{demon interactivity}\label{fig:interactive-workflow-2}
+\end{figure}
+\end{CodeChunk}
 

### 2021-12-13 scripts/demo-interactive.R

@@ -6,14 +6,20 @@ library(tsibble)
 set.seed(123)
 dt <- weatherdata::climate_full %>% slice_sample(n = 50)
 
-
-
+###########################
+# from map to ts
 state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)
 plot_map(state_map) +
   geom_point(data = dt, aes(x = long, y = lat)) + 
-  geom_point(data = dt %>% filter(id == "ASN00003057"), aes(x = long, y = lat), color = "red") 
+  geom_point(data = dt %>% filter(id == "ASN00003057"), aes(x = long, y = lat), 
+             color = "red", size =5) 
 
-ggsave(filename = "figures/map.png")
+ggsave(filename = "figures/illu-highlighted-map.png")
+
+plot_map(state_map) +
+  geom_point(data = dt, aes(x = long, y = lat)) 
+  
+ggsave(filename = "figures/illu-default-map.png")
 
 long <- dt %>% 
   stretch() %>% 
@@ -23,14 +29,39 @@ long <- dt %>%
   mutate(dummy_date = as.Date(glue::glue("2021-{month}-01")))
 
 highlight_line <- long %>% filter(id == "ASN00003057")
+#highlight_point <- highlight_line %>% filter(month == 9)
+long %>% 
+  ggplot(aes(x = dummy_date, y = tmax, group = id)) +
+  #geom_point(size = 0.3) + 
+  geom_line() + 
+  geom_line(data = highlight_line, color = "red", size = 1.5) + 
+  #geom_point(data = highlight_point, color = "red") + 
+  scale_x_date(date_labels = "%b") + 
+  xlab("Month")
+  
+ggsave(filename = "figures/illu-highlighted-long.png", width = 5, height = 3)
+
+long %>% 
+  ggplot(aes(x = dummy_date, y = tmax, group = id)) +
+  geom_line() + 
+  scale_x_date(date_labels = "%b") + 
+  xlab("Month")
+
+ggsave(filename = "figures/illu-default-long.png", width = 5, height = 3)
+
+###########################
+# from ts to map
+
+highlight_line <- long %>% filter(id == "ASN00003057")
 highlight_point <- highlight_line %>% filter(month == 9)
 long %>% 
   ggplot(aes(x = dummy_date, y = tmax, group = id)) +
   geom_point(size = 0.3) + 
   geom_line() + 
-  geom_line(data = highlight_line, color = "red") + 
[…]

### 2021-12-13 scripts/example-interactive.R

@@ -1,0 +1,129 @@
+library(cubble)
+library(dplyr)
+library(ggplot2)
+library(leaflet)
+library(leafpop)
+
+clean <-  weatherdata::climate_full %>%
+  stretch() %>%
+  mutate(month = month(date),
+         diff = tmax - tmin) %>% 
+  group_by(month) %>% 
+  summarise(tmax = mean(tmax, na.rm = TRUE),
+            diff = mean(diff, na.rm = TRUE)) %>% 
+  ungroup(month) %>% 
+  tamp() %>%
+  mutate(diff_winter = ts %>% filter(month %in% c(6,7,8)) %>% pull(diff) %>% mean(na.rm = TRUE),
+         diff_summer = ts %>% filter(month %in% c(12, 1,2)) %>% pull(diff) %>% mean(na.rm = TRUE),
+         diff_year = diff_winter - diff_summer) %>% 
+  stretch() %>% 
+  mutate(dummy_date = as.Date(glue::glue("2021-{month}-01"))) %>%
+  migrate(name) 
+
+long <- clean %>% 
+  SharedData$new(~id, group = "cubble")
+
+nested <- clean %>% tamp() %>% 
+  SharedData$new(~id, group = "cubble")
+
+
+p1 <- nested %>% 
+  ggplot(aes(x = diff_summer, y = diff_winter, color = diff_year, label = name)) + 
+  geom_point() + 
+  geom_abline(slope = 1, intercept = 0, linetype = "dashed") + 
+  colorspace::scale_color_continuous_diverging("Purple-Green", mid = 0, l2 = 80) + 
+  xlim(0, 25) + 
+  ylim(0, 25) +
+  theme_bw() + 
+  theme(legend.position = "none") + 
+  xlab("temperature difference in summer") + 
+  ylab("temperature difference in winter") 
+
+state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)
+p2 <- nested %>%
+  ggplot() +
+  geom_sf(data = state_map, aes(geometry = geometry), fill = "transparent") +
+  geom_point(aes(x = long, y = lat, label = name, color = diff_year)) +
+  colorspace::scale_color_continuous_diverging(
+    "Purple-Green", 
+    mid = 0, l2 = 80,
+    name = "winter-summer difference") + 
+  theme_bw() + 
+  theme(panel.grid.major = element_blank(),
+        legend.position = "bottom")
+
+out <- bscols(
[…]

### 2021-12-13 scripts/script.R

@@ -56,6 +56,9 @@ res %>%
 
 
 ######################################
+set.seed(123)
+dt <- weatherdata::climate_full %>% slice_sample(n = 200)
+
 df <- dt %>%
   stretch() %>%
   mutate(month = month(date)) %>%
@@ -86,7 +89,7 @@ p1 <- long %>%
   geom_point(size = 0.5) + 
   geom_line() +
   scale_x_date(date_labels = "%b") + 
-  scale_color_distiller(palette = "YlOrRd",direction = 1) + 
+  scale_color_distiller(palette = "Reds",direction = 1) + 
   theme_bw() + 
   theme(legend.position = "none") + 
   xlab("Month")
@@ -96,17 +99,17 @@ p2 <- nested %>%
   ggplot() +
   geom_sf(data = state_map, aes(geometry = geometry)) +
   geom_point(aes(x = long, y = lat, color = median_tmax, label = name)) +
-  scale_color_distiller(palette = "YlOrRd", direction = 1) + 
-  theme_void()
+  scale_color_distiller(palette = "Reds", direction = 1) + 
+  theme_bw() + 
+  theme(panel.grid.major = element_blank())
 
 out <- bscols(
   ggplotly(p1, width = 800, height = 800, tooltip = "label") %>%
     highlight(on = "plotly_selected", off = "plotly_deselect", color = "red"),
-  ggplotly(p2, width = 800, height = 800, tooltip = "label") 
-  %>%
+  ggplotly(p2, width = 900, height = 800, tooltip = "label")%>%
     highlight(on = "plotly_selected", off = "plotly_deselect", color = "red")
 )
 
 htmltools::save_html(out, file = "figures/linking.html")
 webshot2::webshot("figures/linking.html", "figures/linking.png", vwidth = 1800, vheight = 2000)
-```
+

### 2021-12-16 main.Rmd
clean up

@@
 )
-options(prompt = "R> ", continue = "+ ")
+options(prompt = "R> ", continue = "+ ", 
+        tibble.print_max = 5, tibble.print_min = 5)
 ```
@@
 
-**Motivation**
-
 Many data structures have been proposed for spatial (\pkg{sf} by @sf) and temporal (\pkg{tsibble} by @tsibble) data in the R community, while less has been done for spatio-temporal data.
@@
 \newline
-
-**Existing packages**
 
@@
 
-**Our new data structure for spatio-temporal data**
 This paper describes the implementation of a new spatio-temporal data structure: \pkg{cubble}.
 \pkg{cubble} implements a relational data structure that uses two forms to manage the switch between spatial and temporal dimension.
-With this structure, users can manipulate the spatial or temporal dimension separately, while leaves the linking of two dimensions to \pkg{cubble\pkg{.
+With this structure, users can manipulate the spatial or temporal dimension separately, while leaves the linking of two dimensions to \pkg{cubble}.
 The software is available from the Comprehensive R Archive Network (CRAN) at [CRAN link].
 \newline
-
-**Section division**
 
@@
 
-Spatio-temporal data usually come in various forms and Figure \ref{fig:cubble-diagram} shows four examples of this.
+Spatio-temporal data usually come in various forms and Figure \ref{fig:illu-input} shows four examples of this.
 No matter which form the data is in, these formats share some common components that characterise spatio-temporal data.
@@
 
-```{r cubble-diagram, echo = FALSE, out.width = "100%", fig.cap="Illustration of incoming data formats for spatio-temporal data.
+```{r illu-input, echo = FALSE, out.width = "100%", fig.cap="Illustration of incoming data formats for spatio-temporal data.
 (1) Data comes in as a single table; (2) Separate tables for spatial and temporal variables; (3) A single table with all the parameters used to query the database and a separate table for queried data; and (4) Cubical data in array or NetCDF format.", fig.align='center', out.height = "15%"}
@@
 
-In a cubble, there are two forms: nested form and long form, and Figure \ref{fig:def-cubble} sketches the two forms along with the associated attributes.
+In a cubble, there are two forms: nested form and long form, and Figure \ref{fig:illu-cubble} sketches the two forms along with the associated attributes.
 The decision on which form to use is output-oriented, meaning analysts need to first think about whether the output of a particular operation is identified only by the spatial identifier, or a combination of spatial and temporal identifier.
@@
 
-```{r def-cubble, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "Illustration of nested and long cubble."}
+```{r illu-cubble, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "Illustration of nested and long cubble."}
 knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.002.png"))
@@
 After we have got `cluster_nested`, spatial and temporal data at both levels can be easily obtained.
-Figure \ref{fig:hierachical} illustrate the relationship between the long and nested form cubble at both site and cluster level.
+Figure \ref{fig:illu-hier} illustrate the relationship between the long and nested form cubble at both site and cluster level.
 More description on this.
@@
 
+```{r illu-hier, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "Hierarchical structure"}
+knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.003.png"))
+```
+
 ## Data fusion and matching
+
+Temporal matching checks how spatially matched pairs align temporally. We use the following chart to illustrate how the temporal matching works:
+
+For each spatially matched pair, say `A` and `a`, we first find the largest `n` points in each series, colored in brown points here.
+Here we use the largest three but you can tune this number by `temporal_n_highest`.
+Then we construct the interval of the largest points from one series and see how many points, from the other series, fall into the intervals.
+The series used to construct the interval is controlled by `temporal_independent` and the window size by `temporal_window` with a default of 5.
+
+In this illustration, we construct the interval based on series `A` and two of the three peaks from `a` falls into this interval at Time 7 and 27.
+
+There's another mandatory argument that hasn't been introduced above: `temporal_var_to_match`.
+This argument controls the variable to match and it needs to appear in both the `major` and `minor` set.
+In the water level matching example, we match the variable `Water_course_level` from `river` to `prcp` from `climate`, hence need to manually rename one of them to match the other, here we rename `Water_course_level` to `prcp` in `river`:
+
+Figure \ref{fig:illu-matching}
+
+```{r illu-matching, out.width="100%", out.height="30%", fig.cap="sdfasdf"}
+knitr::include_graphics("figures/illu-matching.png")
+```
 
@@
 The cubble structure fits in naturally with the interactive graphic pipeline discussed in the literature [@buja1988elements; @buja1996interactive;  @sutherland2000orca; @xie2014reactive; @cheng2016enabling].
-Diagram \ref{fig:interactive-workflow} illustrates how linking works with the two forms in a cubble, where a time series plot is created with the long cubble and a map is created with the nested cubble.
+Diagram \ref{fig:illu-interactive} illustrates how linking works with the two forms in a cubble, where a time series plot is created with the long cubble and a map is created with the nested cubble.
 When a user action is captured from the map, the site will be activated in the nested cubble.
@@
 This includes other points in the same time series in the long cubble and the corresponding observation of site in the nested cubble.
-These activated observations will then being reflected in the updated plots and Diagram \ref{fig:interactive-workflow-2} in the Appendix illustrates this process.
+These activated observations will then being reflected in the updated plots and Diagram \ref{fig:illu-interactive-2} in the Appendix illustrates this process.
 
-```{r interactive-workflow, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "demon interactivity"}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.003.png"))
+```{r illu-interactive, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "demon interactivity"}
+knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.004.png"))
 ```
@@
 This variables can be openly queried via `rnoaa`.
-`weatherdata` provides a readily available copy of several data.`weatherdata::historical_tmax` provides historical maximum temperature recorded for Australian stations with the earliest dating back to 1859 and this can be used to understand the change of maximum temperature from a past period to nowadays.
-Figure \ref{fig:glyphmap-workflow} shows the steps involved from the initial data to the final glyph map in Figure \ref{fig:glyphmap}.
+`weatherdata` provides a readily available copy of several data.
+`weatherdata::historical_tmax` provides historical maximum temperature recorded for Australian stations with the earliest dating back to 1859 and this can be used to understand the change of maximum temperature from a past period to nowadays.
 
-```{r glyphmap-workflow, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "demon code"}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.004.png"))
+Figure \ref{fig:basic-manip}.
+
+```{r eval=  FALSE, echo = TRUE}
+tmax <- weatherdata::historical_tmax %>%
+  # filter out VIC and NSW stations
+  filter(between(stringr::str_sub(id, 7, 8), 46, 90)) %>%
+
+  # filter out observations in between 1971-1975 and 2016 - 2020
+  stretch() %>%
+  filter(lubridate::year(date) %in% c(1971:1975, 2016:2020)) %>%
+  mutate(month = lubridate::month(date), 
+         group = as.factor(ifelse(lubridate::year(date) > 2015, 
+                                  "2016 ~ 2020", "1971 ~ 1975"))) %>%
+  group_by(month, group) %>%
+  # summarise tmax into monthly mean for both periods
+  summarise(tmax = mean(tmax, na.rm = TRUE)) %>%
+
+  # filter out stations that have complete records in both periods 
+  tamp() %>%
+  filter(nrow(ts) == 24) %>%
+
+  # migrate longitude and latitude to the long form for making the glyph map
+  stretch() %>%
+  migrate(latitude, longitude)
+
+gly <- glyphs(tmax, "longitude", "month", "latitude", "tmax",
+              height = 0.6, width = 1)
+
+nsw_vic <- ozmaps::abs_ste %>% filter(NAME %in% c("Victoria", "New South Wales"))
+
+ggplot() + 
+  geom_sf(data = nsw_vic, fill = "transparent") + 
+  geom_path(data = gly, 
+            aes(gx, gy, group = interaction(id, group), color = group)) +
+  scale_color_brewer(palette = "Dark2") + 
+  theme_bw() + 
+  coord_sf(xlim = c(140, 155)) + 
+  theme(legend.position = "bottom")
+```
+
+```{r basic-manip, out.width="100%", out.height="70%", fig.cap="asdfasd"}
+knitr::include_graphics("figures/basic-manip.png")
 ```
@@
 
-```{r echo = FALSE, eval = FALSE}
-tmax_nested <- weatherdata::historical_tmax %>%
-  filter(between(stringr::str_sub(id, 7, 8), 46, 90))
-
-tmax_long <- tmax_nested %>%
-  stretch() %>%
-  filter(lubridate::year(date) %in% c(1971:1975, 2016:2020)) %>%
-  mutate(
-    month = lubridate::month(date),
-    group = as.factor(ifelse(
-      lubridate::year(date) > 2015,
-      "2016 ~ 2020",
-      "1971 ~ 1975"
-    ))
-  ) %>%
-  group_by(month, group) %>%
-  summarise(tmax = mean(tmax, na.rm = TRUE))
-
-tmax_nested2 <- tmax_long %>%
-  tamp() %>%
-  filter(nrow(ts) == 24)
-
-tmax_long2 <- tmax_nested2 %>%
-  stretch() %>%
-  migrate(latitude, longitude)
-
-gly <- glyphs(tmax_long2, "longitude", "month", "latitude", "tmax",
-              height = 0.6, width = 1)
-
-nsw_vic <- ozmaps::abs_ste %>%
-  filter(NAME %in% c("Victoria", "New South Wales")) #
-
-plot_map(nsw_vic, fill = "transparent") +
-  geom_path(data = gly, 
-            aes(gx, gy, group = interaction(id, group), 
-                color = group)) +
-  scale_color_brewer(palette = "Dark2") + 
-  theme_bw() + 
-  coord_sf(xlim = c(141, 154)) + 
-  theme(legend.position = "bottom")
-
-ggsave(filename = "figures/basic-manipulation.png", width = 10, height = 10)
-```
-
-```{r glyphmap, out.width="100%", out.height="70%", fig.cap="asdfasd"}
-knitr::include_graphics("figures/basic-manipulation.png")
-```
-
 ### Aggregation
 
-In figure \ref{fig:glyphmap}, there are a few stations around Sydney (151E, 34S) and New castle (152E, 33S).
+In figure \ref{fig:basic-manip}, there are a few stations around Sydney (151E, 34S) and New castle (152E, 33S).
 Aggregating them into averages or other statistics would give a better picture.
@@
 
-In Figure \ref{fig:aggregation}
+In Figure \ref{fig:basic-agg}
 
@@
 
-```{r aggregation, out.width="100%", out.height="90%", fig.cap="sdfasdf"}
-knitr::include_graphics("figures/aggregation.png")
+```{r basic-agg, out.width="100%", out.height="90%", fig.cap="sdfasdf"}
+knitr::include_graphics("figures/basic-agg.png")
 ```
@@
 
-```{r echo = FALSE}
-river <- weatherdata::water %>%
-  stretch() %>%
-  select(date, Water_course_level) %>%
-  tamp() %>%
-  mutate(type = "river")
-
-climate <- weatherdata::climate_full %>%
-  filter(between(stringr::str_sub(id, 7, 8), 76, 90)) %>%
-  stretch() %>%
-  filter(lubridate::year(date) == 2020) %>%
-  tamp() %>%
-  mutate(type = "climate")
-
-vic_map <- rmapshaper::ms_simplify(ozmaps::abs_ste %>% filter(NAME == "Victoria"))
-plot_map(vic_map) +
-  geom_point(
-    data = dplyr::bind_rows(river, climate),
-    aes(x = long, y = lat, color = type)
-  ) +
-  scale_color_brewer(palette = "Dark2") +
-  theme_bw()
-```
-
-Temporal matching checks how spatially matched pairs align temporally. We use the following chart to illustrate how the temporal matching works:
-
-```{r echo = FALSE, fig.height=3, fig.cap = "shtishisaasdf"}
-set.seed(123)
-dt <- tibble(
-  id = factor(c(rep("A", 31), rep("a", 31)), levels = c("A", "a")),
-  date = rep(1:31, 2),
-  val = c(
-    c(rnorm(5), 10, rnorm(7), 5, rnorm(8), 7, rnorm(8)),
-    c(
-      rnorm(6, sd = 0.5), 7, rnorm(7, sd = 0.5), rnorm(6, sd = 0.5),
-      4, rnorm(5, sd = 0.5), 6, rnorm(4, sd = 0.5)
-    )
-  )
-) %>% mutate(val = ifelse(val < 0, 0, val))
-
-circle <- tibble(
-  x = c(6, 14, 23, 7, 21, 27),
-  y = c(10, 5, 7, 7, 4, 6),
-  id = factor(c(rep("A", 3), rep("a", 3)), levels = c("A", "a")),
-  xend = x + 5
-)
-
-errorbar <- bind_rows(
-  circle %>% filter(id == "A"),
-  circle %>% filter(id == "A") %>% mutate(id = "a")
-) %>%
-  mutate(id = factor(id, c("A", "a")))
-
-fallin <- tibble(
-  id = factor(rep("a", 2), c("A", "a")),
-  x = c(7, 27),
-  y = -0.5,
-  label = "\U2714"
-)
-
-ggplot() +
-  geom_line(data = dt, aes(x = date, y = val, group = id), color = "grey60") +
-  geom_vline(data = circle, aes(xintercept = x), linetype = "longdash", color = "grey60") +
-  geom_errorbar(data = errorbar, aes(xmin = x, xmax = xend, y = -0.5), width = 0.5) +
-  geom_point(data = circle, aes(x = x, y = y), color = "brown") +
-  geom_text(data = fallin, aes(x = x, y = y, label = label), color = "brown", size = 5) +
-  facet_wrap(vars(id), nrow = 2) +
-  scale_y_continuous(breaks = seq(0, 10, 2)) +
-  scale_x_continuous(breaks = seq(1, 31, 1)) +
-  theme_bw() +
-  theme(panel.grid = element_blank()) +
-  labs(x = "Time", y = "Value")
-```
-
-For each spatially matched pair, say `A` and `a`, we first find the largest `n` points in each series, colored in brown points here.
-Here we use the largest three but you can tune this number by `temporal_n_highest`.
-Then we construct the interval of the largest points from one series and see how many points, from the other series, fall into the intervals.
-The series used to construct the interval is controlled by `temporal_independent` and the window size by `temporal_window` with a default of 5.
-
-In this illustration, we construct the interval based on series `A` and two of the three peaks from `a` falls into this interval at Time 7 and 27.
-
-There's another mandatory argument that hasn't been introduced above: `temporal_var_to_match`.
-This argument controls the variable to match and it needs to appear in both the `major` and `minor` set.
-In the water level matching example, we match the variable `Water_course_level` from `river` to `prcp` from `climate`, hence need to manually rename one of them to match the other, here we rename `Water_course_level` to `prcp` in `river`:
-
-```{r}
-river <- river %>%
-  stretch() %>%
-  rename(prcp = Water_course_level) %>%
-  tamp()
-```
-
 Now we use `match_sites()` to first pair the weather stations with the river gauges spatially and then apply the temporal matching on `prcp`.
@@
 This is a tuning parameter and you can start with 10% of the points of one series (here we have daily data for a year, 10% is roughly 30 points).
-`temporal_min_match` filters out pairs don't have enough match and to return all the pairs, set `temporal_min_match` to `0`.
-
-```{r}
-res <- match_sites(river, climate,
-  temporal_var_to_match = prcp,
-  temporal_independent = climate,  
-  temporal_n_highest = 30,
-  temporal_min_match = 15
-)
-
-res
-```
+`temporal_min_match` filters out pairs don't have enough match and to return all the pairs, set `temporal_min_match` to `0`.``
 
@@
 
-```{r echo = FALSE, out.width="100%", fig.cap="..."}
-p1 <- plot_map(vic_map) +
-  geom_point(
-    data = res,
-    aes(x = long, y = lat, color = type)
-  ) +
-  ggrepel::geom_label_repel(
-    data = res %>% filter(type == "river"),
-    aes(x = long, y = lat, label = .group)
-  ) +
-  scale_color_brewer(palette = "Dark2") +
-  ggtitle("Victoria") +
-  theme_minimal() +
-  theme(
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank(),
-    legend.position = "bottom"
-  )
-
-res_long <- res %>%
-  stretch(ts) %>%
-  migrate(.group, type) %>%
-  mutate(prcp = scale(prcp)[, 1]) %>%
-  switch_key(.group) %>%
-  migrate(type)
-
-p2 <- res_long %>%
-  ggplot(aes(
-    x = date, y = prcp,
-    group = id, color = type
-  )) +
-  geom_line() +
-  facet_wrap(vars(.group)) +
-  scale_color_brewer(palette = "Dark2", guide = "none") +
-  theme_bw() +
-  labs(x = "week") +
-  scale_x_date(date_labels = "%b")
-
-p1 | p2
+```{r matching, out.width="100%", out.height="30%", fig.cap="asdfasd"}
+knitr::include_graphics("figures/matching.png")
 ```
@@
 ## ERA5: climate reanalysis data
-
-```{r eval = FALSE}
-nc <- ncdf4::nc_open(here::here("data/adaptor.mars.internal-1638344933.932894-4194-11-a5d35ac3-b36d-4982-afa5-a5dd193a6718.nc"))
-
-nc <- ncdf4::nc_open(here::here("data/tas_Amon_ACCESS-CM2_ssp585_r1i1p1f1_gn_20150116-20201216_v20191108.nc"))
-temp <- as_cubble(nc, vars = "tas")
-vic_map <- ozmaps::abs_ste %>% filter(NAME == "Victoria")
-
-res <- temp %>% 
-  stretch(ts) %>% 
-  mutate(d2m = d2m - 273.15) %>% 
-  filter(lubridate::hour(time) == 12) %>% 
-  group_by(id) %>% 
-  summarise(d2m = mean(d2m, na.rm= TRUE)) %>%
-  migrate(long, lat)
-
-vic_string <- vic_map %>% sf::st_cast("MULTILINESTRING")
-
-res %>% 
-  ggplot() +
-  geom_tile(aes(x = long, y = lat, fill = d2m)) +
-  geom_sf(data = vic_string , alpha = 0.5, fill = "transparent") +
-  theme_bw() + 
-  colorspace::scale_fill_continuous_sequential("Mint")
-
-```
 
@@
 
-```{r echo = FALSE, out.width="100%", out.height="30%"}
+Figure \ref{fig:interactive-linking} 
+Figure \ref{fig:interactive-popup}
+
+```{r interactive-linking, echo = FALSE, out.width="100%", out.height="30%", fig.cap = "sdfasd"}
 knitr::include_graphics(here::here("figures/linking.png"))
@@
 
-```{r echo = FALSE, out.width="50%", out.height="30%"}
+```{r interactive-popup, echo = FALSE, out.width="50%", out.height="30%", fig.cap = "sdfasd"}
 knitr::include_graphics(here::here("figures/popup.png"))
@@
 
-```{r interactive-workflow-2, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "demon interactivity"}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.007.png"))
+```{r illu-interactive-2, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "demon interactivity"}
+knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.005.png"))
 ```
@@
 
-# Old stuff
+<!-- # Old stuff -->
 
-In the temporal aspect, the `tsibble` [@tsibbles] structure and its tidyverts ecosystem have provided a [...
+<!-- In the temporal aspect, the `tsibble` [@tsibbles] structure and its tidyverts ecosystem have provided a [...
 ] workflow to work with temporal data.
@@
 This repetition is unnecessary and would inflate the object size for long series.
+-->
 
-# Create a cubble
+<!-- # Create a cubble -->
 
-The creation of a cubble requires the site identifier (`key`), as well as the spatial (`coords`) and temporal (`index`) identifier.
+<!-- The creation of a cubble requires the site identifier (`key`), as well as the spatial (`coords`) and temporal (`index`) identifier.
 `climate_flat` is already a tibble and it uses `id` to identify each station, `date` as the time identifier, and `c(long, lat)` as the spatial identifier.
-To create a cubble for this data, use:
+To create a cubble for this data, use: -->
 
-```{r}
-climate_flat %>% as_cubble(key = id, index = date, coords = c(long, lat))
-```
+<!-- ```{r} -->
+<!-- climate_flat %>% as_cubble(key = id, index = date, coords = c(long, lat)) -->
+<!-- ``` -->
 
-Most of the time, spatio-temporal data doesn't come into this form and analysts need to query the climate variables based on station metadata.
+<!-- Most of the time, spatio-temporal data doesn't come into this form and analysts need to query the climate variables based on station metadata.
 \textcolor{red}{This is also a problem illustrated in Section 3.5 in @tidydata.
@@
 We first create a station data frame with the 5 target stations.
+-->
 
-```{r echo = FALSE}
-latlon_ref <- tibble::tibble(id = "melbourne", lat = -37.8136, long = 144.9631)
+<!-- ```{r echo = FALSE} -->
+<!-- latlon_ref <- tibble::tibble(id = "melbourne", lat = -37.8136, long = 144.9631) -->
 
-cand <- aus_climate %>%
-  as_tibble() %>%
-  select(id, lat, long, elev, name, wmo_id)
+<!-- cand <- aus_climate %>% -->
+<!--   as_tibble() %>% -->
+<!--   select(id, lat, long, elev, name, wmo_id) -->
 
-calc_dist <- function(df_ref, df, n = 5) {
-  long_ref <- df_ref$long
-  lat_ref <- df_ref$lat
+<!-- calc_dist <- function(df_ref, df, n = 5) { -->
+<!--   long_ref <- df_ref$long -->
+<!--   lat_ref <- df_ref$lat -->
 
-  df %>%
-    mutate(
-      long_ref = long_ref,
-      lat_ref = lat_ref,
-      dist = rnoaa::meteo_spherical_distance(lat, long, lat_ref, long_ref)
-    ) %>%
-    slice_min(dist, n = n) %>%
-    mutate(city = df_ref$id) %>%
-    select(-long_ref, -lat_ref)
-}
+<!--   df %>% -->
+<!--     mutate( -->
+<!--       long_ref = long_ref, -->
+<!--       lat_ref = lat_ref, -->
+<!--       dist = rnoaa::meteo_spherical_distance(lat, long, lat_ref, long_ref) -->
+<!--     ) %>% -->
+<!--     slice_min(dist, n = n) %>% -->
+<!--     mutate(city = df_ref$id) %>% -->
+<!--     select(-long_ref, -lat_ref) -->
+<!-- } -->
 
-(stations <- do.call("rbind", purrr::map(
-  1:nrow(latlon_ref),
-  ~ calc_dist(latlon_ref[.x, ], cand)
-)))
-```
+<!-- (stations <- do.call("rbind", purrr::map( -->
+<!--   1:nrow(latlon_ref), -->
+<!--   ~ calc_dist(latlon_ref[.x, ], cand) -->
+<!-- ))) -->
+<!-- ``` -->
 
-We can query the climate information into a nested list named `ts` for each station with the `rowwise()` operator. To create a cubble, supply the same identifiers as with the first example.
+<!-- We can query the climate information into a nested list named `ts` for each station with the `rowwise()` operator. To create a cubble, supply the same identifiers as with the first example. -->
 
-```{r eval = FALSE}
-sydmel_climate <- stations %>%
-  rowwise() %>%
-  mutate(ts = list(meteo_pull_monitors(id,
-    date_min = "2020-01-01",
-    date_max = "2020-12-31",
-    var = c("PRCP", "TMAX", "TMIN")
-  ) %>%
-    select(-id))) %>%
-  as_cubble(key = id, index = date, coords = c(long, lat))
-```
+<!-- ```{r eval = FALSE} -->
+<!-- sydmel_climate <- stations %>% -->
+<!--   rowwise() %>% -->
+<!--   mutate(ts = list(meteo_pull_monitors(id, -->
+<!--     date_min = "2020-01-01", -->
+<!--     date_max = "2020-12-31", -->
+<!--     var = c("PRCP", "TMAX", "TMIN") -->
+<!--   ) %>% -->
+<!--     select(-id))) %>% -->
+<!--   as_cubble(key = id, index = date, coords = c(long, lat)) -->
+<!-- ``` -->
 
-```{r echo = FALSE}
-(sydmel_climate <- stations %>%
-  left_join(aus_climate) %>%
-  as_cubble(key = id, index = date, coords = c(long, lat)))
-```
+<!-- ```{r echo = FALSE} -->
+<!-- (sydmel_climate <- stations %>% -->
+<!--   left_join(aus_climate) %>% -->
+<!--   as_cubble(key = id, index = date, coords = c(long, lat))) -->
+<!-- ``` -->
 
-Below are the how the nested and long form look like for Australia climate data, which records daily precipitation, maximum and minimum temperature for 55 stations across Australia from 2015- 2020.
+<!-- Below are the how the nested and long form look like for Australia climate data, which records daily precipitation, maximum and minimum temperature for 55 stations across Australia from 2015- 2020.
 Notice that each station forms a group in both forms and specifically, the nested and long form have a underlying `rowwise_df` and `grouped_df` respectively.
+-->
 
-With a cubic framework on mind, different types of manipulation with cubble can be thought of as slicing the cube in various way.
+<!-- With a cubic framework on mind, different types of manipulation with cubble can be thought of as slicing the cube in various way.
 The table below shows how some `dplyr` verbs are mapped into the operation in a cubble.
 With the existing grouping on the station, additional grouping can be added with `group_by` and removed with `ungrouped`.
-[talk about why it is useful]
+[talk about why it is useful] -->
 
-\newpage
+<!-- \newpage -->
 
-## Cubble operations
+<!-- ## Cubble operations -->
 
-```{r cubble-operations, echo = FALSE, fig.cap="Cubble operations", out.height="50%", out.width="90%", fig.align='center'}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.004.png"))
-```
+<!-- ```{r cubble-operations, echo = FALSE, fig.cap="Cubble operations", out.height="50%", out.width="90%", fig.align='center'} -->
+<!-- knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.004.png")) -->
+<!-- ``` -->

### 2021-12-16 main.tex
clean up

@@
 
-\textbf{Motivation}
-
 Many data structures have been proposed for spatial (\pkg{sf} by
@@
 add new operations into the data but can be error prone. \newline
-
-\textbf{Existing packages}
 
@@
 
-\textbf{Our new data structure for spatio-temporal data} This paper
-describes the implementation of a new spatio-temporal data structure:
-\pkg{cubble}. \pkg{cubble} implements a relational data structure that
-uses two forms to manage the switch between spatial and temporal
-dimension. With this structure, users can manipulate the spatial or
-temporal dimension separately, while leaves the linking of two
-dimensions to \textbackslash pkg\{cubble\textbackslash pkg\{. The
-software is available from the Comprehensive R Archive Network (CRAN) at
-{[}CRAN link{]}. \newline
-
-\textbf{Section division}
+This paper describes the implementation of a new spatio-temporal data
+structure: \pkg{cubble}. \pkg{cubble} implements a relational data
+structure that uses two forms to manage the switch between spatial and
+temporal dimension. With this structure, users can manipulate the
+spatial or temporal dimension separately, while leaves the linking of
+two dimensions to \pkg{cubble}. The software is available from the
+Comprehensive R Archive Network (CRAN) at {[}CRAN link{]}. \newline
 
@@
 Spatio-temporal data usually come in various forms and Figure
-\ref{fig:cubble-diagram} shows four examples of this. No matter which
-form the data is in, these formats share some common components that
+\ref{fig:illu-input} shows four examples of this. No matter which form
+the data is in, these formats share some common components that
 characterise spatio-temporal data. A spatial identifier (\texttt{id} in
@@
 \caption[Illustration of incoming data formats for spatio-temporal data]{Illustration of incoming data formats for spatio-temporal data.
-(1) Data comes in as a single table; (2) Separate tables for spatial and temporal variables; (3) A single table with all the parameters used to query the database and a separate table for queried data; and (4) Cubical data in array or NetCDF format.}\label{fig:cubble-diagram}
+(1) Data comes in as a single table; (2) Separate tables for spatial and temporal variables; (3) A single table with all the parameters used to query the database and a separate table for queried data; and (4) Cubical data in array or NetCDF format.}\label{fig:illu-input}
 \end{figure}
@@
 In a cubble, there are two forms: nested form and long form, and Figure
-\ref{fig:def-cubble} sketches the two forms along with the associated
+\ref{fig:illu-cubble} sketches the two forms along with the associated
 attributes. The decision on which form to use is output-oriented,
@@
 
-\caption[Illustration of nested and long cubble]{Illustration of nested and long cubble.}\label{fig:def-cubble}
+\caption[Illustration of nested and long cubble]{Illustration of nested and long cubble.}\label{fig:illu-cubble}
 \end{figure}
@@
 5 ASN00009021 2020-01-05     0  35.3  11.6
- 6 ASN00009021 2020-01-06     0  34.8  13.1
- 7 ASN00009021 2020-01-07     0  32.8  15.1
- 8 ASN00009021 2020-01-08     0  30.4  17.4
- 9 ASN00009021 2020-01-09     0  28.7  17.3
-10 ASN00009021 2020-01-10     0  32.6  15.8
-# ... with 1,820 more rows
+# ... with 1,825 more rows
 \end{CodeOutput}
@@
 5 ASN00009021 2020-01-05     0  35.3  11.6  116. -31.9
- 6 ASN00009021 2020-01-06     0  34.8  13.1  116. -31.9
- 7 ASN00009021 2020-01-07     0  32.8  15.1  116. -31.9
- 8 ASN00009021 2020-01-08     0  30.4  17.4  116. -31.9
- 9 ASN00009021 2020-01-09     0  28.7  17.3  116. -31.9
-10 ASN00009021 2020-01-10     0  32.6  15.8  116. -31.9
-# ... with 1,820 more rows
+# ... with 1,825 more rows
 \end{CodeOutput}
@@
 After we have got \texttt{cluster\_nested}, spatial and temporal data at
-both levels can be easily obtained. Figure \ref{fig:hierachical}
+both levels can be easily obtained. Figure \ref{fig:illu-hier}
 illustrate the relationship between the long and nested form cubble at
@@
 
-\hypertarget{data-fusion-and-matching}{%
-\subsection{Data fusion and matching}\label{data-fusion-and-matching}}
-
-\hypertarget{interactive-graphics}{%
-\subsection{Interactive graphics}\label{interactive-graphics}}
-
-The cubble structure fits in naturally with the interactive graphic
-pipeline discussed in the literature
-\citep{buja1988elements, buja1996interactive, sutherland2000orca, xie2014reactive, cheng2016enabling}.
-Diagram \ref{fig:interactive-workflow} illustrates how linking works
-with the two forms in a cubble, where a time series plot is created with
-the long cubble and a map is created with the nested cubble. When a user
-action is captured from the map, the site will be activated in the
-nested cubble. Then, the nested cubble will communicate to the long
-cubble to activate all the observations with the same \texttt{id}. The
-long cubble will then highlight the activated series in the time series
-plot.
-
-The linking is also available from the time series plot to the map. The
-selection on the time series is through selecting the point on the time
-series and once a point is selected, it will be activated in the long
-cubble. All the observations that share the same \texttt{id}, either in
-the long and nested cubble, are then activated. This includes other
-points in the same time series in the long cubble and the corresponding
-observation of site in the nested cubble. These activated observations
-will then being reflected in the updated plots and Diagram
-\ref{fig:interactive-workflow-2} in the Appendix illustrates this
-process.
-
 \begin{CodeChunk}
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.003}
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.003}
 
@@
 
-\caption[demon interactivity]{demon interactivity}\label{fig:interactive-workflow}
+\caption[Hierarchical structure]{Hierarchical structure}\label{fig:illu-hier}
 \end{figure}
@@
 
-\newpage
-
-\hypertarget{examples}{%
-\section{Examples}\label{examples}}
-
-\hypertarget{australia-precipitation-pattern-in-2020}{%
-\subsection{Australia precipitation pattern in
-2020}\label{australia-precipitation-pattern-in-2020}}
-
-\hypertarget{basic-manipulation}{%
-\subsubsection{Basic manipulation}\label{basic-manipulation}}
-
-Bureau of Meteorology (BoM) collects climate variables. This variables
-can be openly queried via \texttt{rnoaa}. \texttt{weatherdata} provides
-a readily available copy of several
-data.\texttt{weatherdata::historical\_tmax} provides historical maximum
-temperature recorded for Australian stations with the earliest dating
-back to 1859 and this can be used to understand the change of maximum
-temperature from a past period to nowadays. Figure
-\ref{fig:glyphmap-workflow} shows the steps involved from the initial
-data to the final glyph map in Figure \ref{fig:glyphmap}.
-
-\begin{CodeChunk}
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.004}
-
-}
-
-\caption[demon code]{demon code}\label{fig:glyphmap-workflow}
-\end{figure}
-\end{CodeChunk}
-
-\begin{CodeChunk}
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth,height=0.7\textheight]{figures/basic-manipulation}
-
-}
-
-\caption[asdfasd]{asdfasd}\label{fig:glyphmap}
-\end{figure}
-\end{CodeChunk}
-
-\hypertarget{aggregation}{%
-\subsubsection{Aggregation}\label{aggregation}}
-
-In figure \ref{fig:glyphmap}, there are a few stations around Sydney
-(151E, 34S) and New castle (152E, 33S). Aggregating them into averages
-or other statistics would give a better picture. This can be done with
-cubble through a hierarchical structure, where the cluster indicator is
-one level above the individual station.
-
-In Figure \ref{fig:aggregation}
-
-The \texttt{weatherdata::climate\_full} data has daily climate data of
-639 Australia stations from 2016 to 2020.
-
-calcluate distance matrix As an example to illustrate here, a kmean
-clustering algorithm based on the distance matrix is used and the number
-of centres is set to 20. More complex algorithms can also be used for
-more complex problem, as long as a mapping from each station id to the
-cluster id can be constructed.
-
-\begin{CodeChunk}
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth,height=0.9\textheight]{figures/aggregation}
-
-}
-
-\caption[sdfasdf]{sdfasdf}\label{fig:aggregation}
-\end{figure}
-\end{CodeChunk}
-
-\hypertarget{river-level-data-in-victria-water-gauges}{%
-\subsection{River level data in Victria water
-gauges}\label{river-level-data-in-victria-water-gauges}}
-
-The water level data comes from
-\href{http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full}{Bureau
-of Meteorology} and has a copy in \texttt{weatherdata}. Here we extract
-the water course level and add a column annotate this data of type
-\texttt{river}. For the rainfall data, we will still use the
-\texttt{weatherdata::climate\_full}, filtering for Victorian stations in
-2020 should be pretty familiar by now. Again, we first look at where
-these stations are on the map first:
-
-\begin{CodeChunk}
-
-\begin{center}\includegraphics{figures/unnamed-chunk-7-1} \end{center}
-
-\end{CodeChunk}
+\hypertarget{data-fusion-and-matching}{%
+\subsection{Data fusion and matching}\label{data-fusion-and-matching}}
 
@@
 works:
-
-\begin{CodeChunk}
-\begin{figure}
-
-{\centering \includegraphics{figures/unnamed-chunk-8-1}
-
-}
-
-\caption[shtishisaasdf]{shtishisaasdf}\label{fig:unnamed-chunk-8}
-\end{figure}
-\end{CodeChunk}
 
@@
 
+Figure \ref{fig:illu-matching}
+
+\begin{CodeChunk}
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth,height=0.3\textheight]{figures/illu-matching}
+
+}
+
+\caption[sdfasdf]{sdfasdf}\label{fig:illu-matching}
+\end{figure}
+\end{CodeChunk}
+
+\hypertarget{interactive-graphics}{%
+\subsection{Interactive graphics}\label{interactive-graphics}}
+
+The cubble structure fits in naturally with the interactive graphic
+pipeline discussed in the literature
+\citep{buja1988elements, buja1996interactive, sutherland2000orca, xie2014reactive, cheng2016enabling}.
+Diagram \ref{fig:illu-interactive} illustrates how linking works with
+the two forms in a cubble, where a time series plot is created with the
+long cubble and a map is created with the nested cubble. When a user
+action is captured from the map, the site will be activated in the
+nested cubble. Then, the nested cubble will communicate to the long
+cubble to activate all the observations with the same \texttt{id}. The
+long cubble will then highlight the activated series in the time series
+plot.
+
+The linking is also available from the time series plot to the map. The
+selection on the time series is through selecting the point on the time
+series and once a point is selected, it will be activated in the long
+cubble. All the observations that share the same \texttt{id}, either in
+the long and nested cubble, are then activated. This includes other
+points in the same time series in the long cubble and the corresponding
+observation of site in the nested cubble. These activated observations
+will then being reflected in the updated plots and Diagram
+\ref{fig:illu-interactive-2} in the Appendix illustrates this process.
+
+\begin{CodeChunk}
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.004}
+
+}
+
+\caption[demon interactivity]{demon interactivity}\label{fig:illu-interactive}
+\end{figure}
+\end{CodeChunk}
+
+\newpage
+
+\hypertarget{examples}{%
+\section{Examples}\label{examples}}
+
+\hypertarget{australia-precipitation-pattern-in-2020}{%
+\subsection{Australia precipitation pattern in
+2020}\label{australia-precipitation-pattern-in-2020}}
+
+\hypertarget{basic-manipulation}{%
+\subsubsection{Basic manipulation}\label{basic-manipulation}}
+
+Bureau of Meteorology (BoM) collects climate variables. This variables
+can be openly queried via \texttt{rnoaa}. \texttt{weatherdata} provides
+a readily available copy of several data.
+\texttt{weatherdata::historical\_tmax} provides historical maximum
+temperature recorded for Australian stations with the earliest dating
+back to 1859 and this can be used to understand the change of maximum
+temperature from a past period to nowadays.
+
+Figure \ref{fig:basic-manip}.
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> tmax <- weatherdata::historical_tmax %>%
++   # filter out VIC and NSW stations
++   filter(between(stringr::str_sub(id, 7, 8), 46, 90)) %>% 
++   
++   # filter out observations in between 1971-1975 and 2016 - 2020
++   stretch() %>%
++   filter(lubridate::year(date) %in% c(1971:1975, 2016:2020)) %>%
++   mutate(month = lubridate::month(date), 
++          group = as.factor(ifelse(lubridate::year(date) > 2015, 
++                                   "2016 ~ 2020", "1971 ~ 1975"))) %>%
++   group_by(month, group) %>%
++   # summarise tmax into monthly mean for both periods
++   summarise(tmax = mean(tmax, na.rm = TRUE)) %>% 
++   
++   # filter out stations that have complete records in both periods 
++   tamp() %>%
++   filter(nrow(ts) == 24) %>% 
++   
++   # migrate longitude and latitude to the long form for making the glyph map
++   stretch() %>%
++   migrate(latitude, longitude)
+R> 
+R> gly <- glyphs(tmax, "longitude", "month", "latitude", "tmax",
++               height = 0.6, width = 1)
+R> 
+R> nsw_vic <- ozmaps::abs_ste %>% filter(NAME %in% c("Victoria", "New South Wales"))
+R> 
+R> ggplot() + 
++   geom_sf(data = nsw_vic, fill = "transparent") + 
++   geom_path(data = gly, 
++             aes(gx, gy, group = interaction(id, group), color = group)) +
++   scale_color_brewer(palette = "Dark2") + 
++   theme_bw() + 
++   coord_sf(xlim = c(140, 155)) + 
++   theme(legend.position = "bottom")
+\end{CodeInput}
+\end{CodeChunk}
+
+\begin{CodeChunk}
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth,height=0.7\textheight]{figures/basic-manip}
+
+}
+
+\caption[asdfasd]{asdfasd}\label{fig:basic-manip}
+\end{figure}
+\end{CodeChunk}
+
+\hypertarget{aggregation}{%
+\subsubsection{Aggregation}\label{aggregation}}
+
+In figure \ref{fig:basic-manip}, there are a few stations around Sydney
+(151E, 34S) and New castle (152E, 33S). Aggregating them into averages
+or other statistics would give a better picture. This can be done with
+cubble through a hierarchical structure, where the cluster indicator is
+one level above the individual station.
+
+In Figure \ref{fig:basic-agg}
+
+The \texttt{weatherdata::climate\_full} data has daily climate data of
+639 Australia stations from 2016 to 2020.
+
+calcluate distance matrix As an example to illustrate here, a kmean
+clustering algorithm based on the distance matrix is used and the number
+of centres is set to 20. More complex algorithms can also be used for
+more complex problem, as long as a mapping from each station id to the
+cluster id can be constructed.
+
+\begin{CodeChunk}
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth,height=0.9\textheight]{figures/basic-agg}
+
+}
+
+\caption[sdfasdf]{sdfasdf}\label{fig:basic-agg}
+\end{figure}
+\end{CodeChunk}
+
+\hypertarget{river-level-data-in-victria-water-gauges}{%
+\subsection{River level data in Victria water
+gauges}\label{river-level-data-in-victria-water-gauges}}
+
+The water level data comes from
+\href{http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full}{Bureau
+of Meteorology} and has a copy in \texttt{weatherdata}. Here we extract
+the water course level and add a column annotate this data of type
+\texttt{river}. For the rainfall data, we will still use the
+\texttt{weatherdata::climate\_full}, filtering for Victorian stations in
+2020 should be pretty familiar by now. Again, we first look at where
+these stations are on the map first:
+
 Now we use \texttt{match\_sites()} to first pair the weather stations
@@
 filters out pairs don't have enough match and to return all the pairs,
-set \texttt{temporal\_min\_match} to \texttt{0}.
-
-\begin{CodeChunk}
-\begin{CodeOutput}
-# cubble:   id [8]: nested form
-# bbox:     [144.52, -37.73, 146.06, -36.55]
-# temporal: date [date], prcp [dbl]
-  id          name                lat  long type  ts        .dist .group n_match
-  <chr>       <chr>             <dbl> <dbl> <chr> <list>    <dbl>  <int>   <int>
-1 405234      SEVEN CREEKS @ D~ -36.9  146. river <tibble ~  6.15      5      21
-2 ASN00082042 strathbogie       -36.8  146. clim~ <tibble ~  6.15      5      21
-3 404207      HOLLAND CREEK @ ~ -36.6  146. river <tibble ~  8.54     10      21
-4 ASN00082170 benalla airport   -36.6  146. clim~ <tibble ~  8.54     10      21
-5 230200      MARIBYRNONG RIVE~ -37.7  145. river <tibble ~  6.17      6      19
-6 ASN00086038 essendon airport  -37.7  145. clim~ <tibble ~  6.17      6      19
-7 406213      CAMPASPE RIVER @~ -37.0  145. river <tibble ~  1.84      1      18
-8 ASN00088051 redesdale         -37.0  145. clim~ <tibble ~  1.84      1      18
-\end{CodeOutput}
-\end{CodeChunk}
+set \texttt{temporal\_min\_match} to \texttt{0}.``
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{figures/unnamed-chunk-11-1}
+{\centering \includegraphics[width=1\linewidth,height=0.3\textheight]{figures/matching}
 
@@
 
-\caption[..]{...}\label{fig:unnamed-chunk-11}
+\caption[asdfasd]{asdfasd}\label{fig:matching}
 \end{figure}
@@
 
+Figure \ref{fig:interactive-linking} Figure \ref{fig:interactive-popup}
+
 \begin{CodeChunk}
+\begin{figure}
 
-\begin{center}\includegraphics[width=1\linewidth,height=0.3\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/linking} \end{center}
+{\centering \includegraphics[width=1\linewidth,height=0.3\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/linking}
 
+}
+
+\caption[sdfasd]{sdfasd}\label{fig:interactive-linking}
+\end{figure}
 \end{CodeChunk}
@@
 \begin{CodeChunk}
+\begin{figure}
 
-\begin{center}\includegraphics[width=0.5\linewidth,height=0.3\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/popup} \end{center}
+{\centering \includegraphics[width=0.5\linewidth,height=0.3\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/popup}
 
+}
+
+\caption[sdfasd]{sdfasd}\label{fig:interactive-popup}
+\end{figure}
 \end{CodeChunk}
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.007}
+{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.005}
 
@@
 
-\caption[demon interactivity]{demon interactivity}\label{fig:interactive-workflow-2}
+\caption[demon interactivity]{demon interactivity}\label{fig:illu-interactive-2}
 \end{figure}
@@
 
-\hypertarget{old-stuff}{%
-\section{Old stuff}\label{old-stuff}}
-
-In the temporal aspect, the \texttt{tsibble} \citep{tsibbles} structure
-and its tidyverts ecosystem have provided a {[}\ldots{} {]} workflow to
-work with temporal data. In a tsibble structure, temporal data is
-characterised by \texttt{index} and \texttt{key} where \texttt{index} is
-the temporal identifier and \texttt{key} is the identifier for multiple
-series, which could be used as a spatio identifier. However, a tsibble
-object, by construction, always requires the \texttt{index} in its
-structure. This makes it less appealing for spatio-temporal data since
-the output of calculated spatio-specific variables (i.e.~features of
-each series) don't have the time dimension. Analysts will either need to
-have an additional step to join this output to the original tsibble or
-operate with variables stored in two separate objects. In addition, the
-long form structure of a tsibble object means spatio variables
-(i.e.~longitude, latitude, and features of each series if joined back to
-the tsibble) of each spatio identifier will be repetitively recorded at
-each timestamp. This repetition is unnecessary and would inflate the
-object size for long series.
-
-\hypertarget{create-a-cubble}{%
-\section{Create a cubble}\label{create-a-cubble}}
-
-The creation of a cubble requires the site identifier (\texttt{key}), as
-well as the spatial (\texttt{coords}) and temporal (\texttt{index})
-identifier. \texttt{climate\_flat} is already a tibble and it uses
-\texttt{id} to identify each station, \texttt{date} as the time
-identifier, and \texttt{c(long,\ lat)} as the spatial identifier. To
-create a cubble for this data, use:
-
-\begin{CodeChunk}
-\begin{CodeOutput}
-# cubble:   id [5]: nested form
-# bbox:     [115.97, -32.94, 133.55, -12.42]- check gap on long and lat
-# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
-  id            lat  long  elev name           wmo_id ts                
-  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>            
-1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tibble [366 x 4]>
-2 ASN00010311 -31.9  117. 179   york            94623 <tibble [366 x 4]>
-3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble [366 x 4]>
-4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble [366 x 4]>
-5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble [366 x 4]>
-\end{CodeOutput}
-\end{CodeChunk}
-
-Most of the time, spatio-temporal data doesn't come into this form and
-analysts need to query the climate variables based on station metadata.
-\textcolor{red}{This is also a problem illustrated in Section 3.5 in @tidydata. Here we provide a structured way to query this data based on the row-wise operator and nested list.}
-For this type of task, one can structure a metadata into a tibble and
-use row-wise operator to query the climate variables into a nested list.
-As an example here we demonstrate the workflow to find the 5 closest
-stations to Melbourne. We first create a station data frame with the 5
-target stations.
-
-\begin{CodeChunk}
-\begin{CodeOutput}
-# A tibble: 5 x 8
-  id            lat  long  elev name                 wmo_id  dist city     
-  <chr>       <dbl> <dbl> <dbl> <chr>                 <dbl> <dbl> <chr>    
-1 ASN00086038 -37.7  145.  78.4 essendon airport      95866  10.8 melbourne
-2 ASN00086282 -37.7  145. 113.  melbourne airport     94866  20.1 melbourne
-3 ASN00086077 -38.0  145.  12.1 moorabbin airport     94870  21.9 melbourne
-4 ASN00088162 -37.4  145. 528.  wallan (kilmore gap)  94860  48.1 melbourne
-5 ASN00087113 -38.0  144.  10.6 avalon airport        94854  48.8 melbourne
-\end{CodeOutput}
-\end{CodeChunk}
-
-We can query the climate information into a nested list named
-\texttt{ts} for each station with the \texttt{rowwise()} operator. To
-create a cubble, supply the same identifiers as with the first example.
-
-\begin{CodeChunk}
-\begin{CodeOutput}
-# cubble:   id [5]: nested form
-# bbox:     [144.47, -38.03, 145.1, -37.38]
-# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
-  id            lat  long  elev name                 wmo_id  dist city   ts     
-  <chr>       <dbl> <dbl> <dbl> <chr>                 <dbl> <dbl> <chr>  <list> 
-1 ASN00086038 -37.7  145.  78.4 essendon airport      95866  10.8 melbo~ <tibbl~
-2 ASN00086282 -37.7  145. 113.  melbourne airport     94866  20.1 melbo~ <tibbl~
-3 ASN00086077 -38.0  145.  12.1 moorabbin airport     94870  21.9 melbo~ <tibbl~
-4 ASN00088162 -37.4  145. 528.  wallan (kilmore gap)  94860  48.1 melbo~ <tibbl~
-5 ASN00087113 -38.0  144.  10.6 avalon airport        94854  48.8 melbo~ <tibbl~
-\end{CodeOutput}
-\end{CodeChunk}
-
-Below are the how the nested and long form look like for Australia
-climate data, which records daily precipitation, maximum and minimum
-temperature for 55 stations across Australia from 2015- 2020. Notice
-that each station forms a group in both forms and specifically, the
-nested and long form have a underlying \texttt{rowwise\_df} and
-\texttt{grouped\_df} respectively.
-
-With a cubic framework on mind, different types of manipulation with
-cubble can be thought of as slicing the cube in various way. The table
-below shows how some \texttt{dplyr} verbs are mapped into the operation
-in a cubble. With the existing grouping on the station, additional
-grouping can be added with \texttt{group\_by} and removed with
-\texttt{ungrouped}. {[}talk about why it is useful{]}
-
-\newpage
-
-\hypertarget{cubble-operations}{%
-\subsection{Cubble operations}\label{cubble-operations}}
-
-\begin{CodeChunk}
-\begin{figure}
-
-{\centering \includegraphics[width=0.9\linewidth,height=0.5\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.004}
-
-}
-
-\caption[Cubble operations]{Cubble operations}\label{fig:cubble-operations}
-\end{figure}
-\end{CodeChunk}
-
 \bibliography{references.bib}

### 2021-12-16 main.Rmd
interactive graphics

@@
 
-```{r interactive-linking, echo = FALSE, out.width="100%", out.height="30%", fig.cap = "sdfasd"}
+```{r interactive-linking, echo = FALSE, out.width="100%", out.height="25%", fig.cap = "sdfasd"}
 knitr::include_graphics(here::here("figures/linking.png"))
@@
 
-```{r interactive-popup, echo = FALSE, out.width="50%", out.height="30%", fig.cap = "sdfasd"}
-knitr::include_graphics(here::here("figures/popup.png"))
+```{r interactive-popup, echo = FALSE, out.width="50%", out.height="30%", fig.cap = "sdfasd", fig.show='hold'}
+knitr::include_graphics(here::here("figures/popup-tas.png"))
+knitr::include_graphics(here::here("figures/popup-mid.png"))
+knitr::include_graphics(here::here("figures/popup-north.png"))
 ```

### 2021-12-16 main.tex
interactive graphics

@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.3\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/linking}
+{\centering \includegraphics[width=1\linewidth,height=0.25\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/linking}
 
@@
 
-{\centering \includegraphics[width=0.5\linewidth,height=0.3\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/popup}
+{\centering \includegraphics[width=0.5\linewidth,height=0.3\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/popup-tas} \includegraphics[width=0.5\linewidth,height=0.3\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/popup-mid} \includegraphics[width=0.5\linewidth,height=0.3\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/popup-north}
 

### 2021-12-17 main.Rmd
sketch out intch out interactive graphic example section

--- 2021-12-16 main.Rmd
@@
 
+\newpage
+
 ## Interative graphic with cubble
@@
 Two examples are presented here to show how interactivity can be used to explore the climate data across Australia.
+Again, `weatherdata::climate_full` will be used and we hope to include more than one time variable here.
+The difference between the maximum and minimum temperature of the day, the diurnal range, can be directly calculated from the data and it can have large impact on the wine industry (lol this is going far).
+This would be useful to compare the temperature difference across stations and to compare how this temperature difference differs throughout the year.
 
-Figure \ref{fig:interactive-linking} 
-Figure \ref{fig:interactive-popup}
+Temperature difference is a time-wise variable and should be computed in the long form cubble. The variables are averaged over the 2016 to 2020 period by month to stabilise the result.
 
-```{r interactive-linking, echo = FALSE, out.width="100%", out.height="30%", fig.cap = "sdfasd"}
-knitr::include_graphics(here::here("figures/linking.png"))
+```{r eval = FALSE, echo = TRUE}
+res <- weatherdata::climate_full %>%
+  # filter(!id %in% c("ASN00067033", "ASN00072091", 
+  #                   "ASN00059040", "ASN00097053")) %>%
+  stretch() %>%
+  mutate(month = lubridate::month(date),
+         diff = tmax - tmin) %>%
+  group_by(month) %>%
+  summarise(tmax = mean(tmax, na.rm = TRUE),
+            diff = mean(diff, na.rm = TRUE)) %>%
+  ungroup(month)
+
 ```
 
-```{r interactive-popup, echo = FALSE, out.width="50%", out.height="30%", fig.cap = "sdfasd"}
-knitr::include_graphics(here::here("figures/popup.png"))
+Several variables can be useful when comparing the seasonal effect of the temperature differences, for example, the difference of average temperature difference between summer and winter can tell [...].
+This is a station-wise measure and hence computed in the nested form.
+
+```{r eval = FALSE, echo = TRUE}
+res <- res %>% 
+  tamp() %>%
+  mutate(
+    diff_winter = ts %>% filter(month %in% c(6, 7, 8)) %>%
+      pull(diff) %>% mean(na.rm = TRUE),
+    diff_summer = ts %>% filter(month %in% c(12, 1, 2)) %>%
+      pull(diff) %>% mean(na.rm = TRUE),
+    diff_year = diff_winter - diff_summer
+  )
+```
+
+The result from the above manipulation can be directly supplied to make interactive plots and Figure \ref{fig:interactive-linking} is an example made with \pkg{plotly} and \pkg{crosstalk}.
+For example, the second row results from selecting the highest mean maximum temperature in July and this corresponds to Kalumburu where there is a larger winter (June to August) temperature difference than summer (December to February).
+The third row selects one of the stations (Birdsville Airport) with a large average temperature difference in the interior desert region on the map and the linked view from the time series shows that this maximum and minimum temperature difference is constant throughout the year.
+The last row selects the lowest maximum temperature in July on the temperature band and surprisingly, the corresponding station doesn't fall into the Tasmania island but Mountain Hotham in Victoria.
+
+```{r interactive-linking, echo = FALSE, out.width="100%", out.height="25%", fig.cap = "Screenshots of the mean daily temperature difference averaged by month from 2016 to 2020 at various locations in Australia.
+Crossed linking between the map and the temperature band is implemented in both directions for exploring the temperature pattern interactively.
+The first row is the general appearance of the graph and the next three rows are examples from various selection to show different temperature patterns in Australia.
+The html version can be found at ...", fig.show='hold'}
+knitr::include_graphics(here::here("figures/linking.png"))
+knitr::include_graphics(here::here("figures/linking-north.png"))
+knitr::include_graphics(here::here("figures/linking-mid.png"))
+knitr::include_graphics(here::here("figures/linking-south.png"))
+```
+
+Cubble is also compatible with \pkg{leaflet} and with leaflet extension package \pkg{leafpop}, the temperature band can be displayed upon clicking on the station on the map.
+Figure \ref{fig:interactive-popup} shows more examples of temperature pattern that can be found in Australia.
+With the same latitude as Kalumburu but in the Pacific coastline, Cooktown airport shows a similarly high but constant temperature throughout the year.
+With a slightly higher latitude than the mountain Hotham, flat land in Melbourne airport shows a less extreme temperature band.
+Lastly, the Cape Grim Baps, in the north-west tip of the Tasmania island shows a uniformly low and thin temperature band.
+
+```{r interactive-popup, echo = FALSE, out.width="50%", out.height="30%", fig.cap = "Screenshots of the diurnal temperature range averaged by month from 2016 to 2020 at various locations in Australia.
+The leaflet popup allows users to click on a single station for its temperature band as a small plot in the popup window.
+The html version can be found at ...", fig.show='hold'}
+knitr::include_graphics(here::here("figures/popup-north.png"))
+knitr::include_graphics(here::here("figures/popup-vic.png"))
+knitr::include_graphics(here::here("figures/popup-tas.png"))
 ```

### 2021-12-17 main.tex
sketch out intch out interactive graphic example section

--- 2021-12-16 main.tex
@@
 
+\newpage
+
 \hypertarget{interative-graphic-with-cubble}{%
@@
 show how interactivity can be used to explore the climate data across
-Australia.
+Australia. Again, \texttt{weatherdata::climate\_full} will be used and
+we hope to include more than one time variable here. The difference
+between the maximum and minimum temperature of the day, the diurnal
+range, can be directly calculated from the data and it can have large
+impact on the wine industry (lol this is going far). This would be
+useful to compare the temperature difference across stations and to
+compare how this temperature difference differs throughout the year.
 
-Figure \ref{fig:interactive-linking} Figure \ref{fig:interactive-popup}
+Temperature difference is a time-wise variable and should be computed in
+the long form cubble. The variables are averaged over the 2016 to 2020
+period by month to stabilise the result.
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> res <- weatherdata::climate_full %>%
++   # filter(!id %in% c("ASN00067033", "ASN00072091", 
++   #                   "ASN00059040", "ASN00097053")) %>%
++   stretch() %>%
++   mutate(month = lubridate::month(date),
++          diff = tmax - tmin) %>%
++   group_by(month) %>%
++   summarise(tmax = mean(tmax, na.rm = TRUE),
++             diff = mean(diff, na.rm = TRUE)) %>%
++   ungroup(month)
+\end{CodeInput}
+\end{CodeChunk}
+
+Several variables can be useful when comparing the seasonal effect of
+the temperature differences, for example, the difference of average
+temperature difference between summer and winter can tell {[}\ldots{]}.
+This is a station-wise measure and hence computed in the nested form.
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> res <- res %>% 
++   tamp() %>%
++   mutate(
++     diff_winter = ts %>% filter(month %in% c(6, 7, 8)) %>%
++       pull(diff) %>% mean(na.rm = TRUE),
++     diff_summer = ts %>% filter(month %in% c(12, 1, 2)) %>%
++       pull(diff) %>% mean(na.rm = TRUE),
++     diff_year = diff_winter - diff_summer
++   )
+\end{CodeInput}
+\end{CodeChunk}
+
+The result from the above manipulation can be directly supplied to make
+interactive plots and Figure \ref{fig:interactive-linking} is an example
+made with \pkg{plotly} and \pkg{crosstalk}. For example, the second row
+results from selecting the highest mean maximum temperature in July and
+this corresponds to Kalumburu where there is a larger winter (June to
+August) temperature difference than summer (December to February). The
+third row selects one of the stations (Birdsville Airport) with a large
+average temperature difference in the interior desert region on the map
+and the linked view from the time series shows that this maximum and
+minimum temperature difference is constant throughout the year. The last
+row selects the lowest maximum temperature in July on the temperature
+band and surprisingly, the corresponding station doesn't fall into the
+Tasmania island but Mountain Hotham in Victoria.
 
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.3\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/linking}
+{\centering \includegraphics[width=1\linewidth,height=0.25\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/linking} \includegraphics[width=1\linewidth,height=0.25\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/linking-north} \includegraphics[width=1\linewidth,height=0.25\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/linking-mid} \includegraphics[width=1\linewidth,height=0.25\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/linking-south}
 
@@
 
-\caption[sdfasd]{sdfasd}\label{fig:interactive-linking}
+\caption[Screenshots of the mean daily temperature difference averaged by month from 2016 to 2020 at various locations in Australia]{Screenshots of the mean daily temperature difference averaged by month from 2016 to 2020 at various locations in Australia.
+Crossed linking between the map and the temperature band is implemented in both directions for exploring the temperature pattern interactively.
+The first row is the general appearance of the graph and the next three rows are examples from various selection to show different temperature patterns in Australia.
+The html version can be found at ...}\label{fig:interactive-linking}
 \end{figure}
 \end{CodeChunk}
+
+Cubble is also compatible with \pkg{leaflet} and with leaflet extension
+package \pkg{leafpop}, the temperature band can be displayed upon
+clicking on the station on the map. Figure \ref{fig:interactive-popup}
+shows more examples of temperature pattern that can be found in
+Australia. With the same latitude as Kalumburu but in the Pacific
+coastline, Cooktown airport shows a similarly high but constant
+temperature throughout the year. With a slightly higher latitude than
+the mountain Hotham, flat land in Melbourne airport shows a less extreme
+temperature band. Lastly, the Cape Grim Baps, in the north-west tip of
+the Tasmania island shows a uniformly low and thin temperature band.
 
@@
 
-{\centering \includegraphics[width=0.5\linewidth,height=0.3\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/popup}
+{\centering \includegraphics[width=0.5\linewidth,height=0.3\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/popup-north} \includegraphics[width=0.5\linewidth,height=0.3\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/popup-vic} \includegraphics[width=0.5\linewidth,height=0.3\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/popup-tas}
 
@@
 
-\caption[sdfasd]{sdfasd}\label{fig:interactive-popup}
+\caption[Screenshots of the diurnal temperature range averaged by month from 2016 to 2020 at various locations in Australia]{Screenshots of the diurnal temperature range averaged by month from 2016 to 2020 at various locations in Australia.
+The leaflet popup allows users to click on a single station for its temperature band as a small plot in the popup window.
+The html version can be found at ...}\label{fig:interactive-popup}
 \end{figure}

### 2021-12-17 scripts/interactive-linking.R

@@ -4,30 +4,29 @@ library(crosstalk)
 library(plotly)
 
 clean <-  weatherdata::climate_full %>%
-  filter(!id %in% c("ASN00067033", "ASN00072091")) %>% 
+  filter(!id %in% c("ASN00067033", "ASN00072091", "ASN00059040", "ASN00097053")) %>%
   stretch() %>%
   mutate(month = lubridate::month(date),
-         diff = tmax - tmin) %>% 
-  group_by(month) %>% 
+         diff = tmax - tmin,
+         avg = (tmax + tmin)/2) %>%
+  group_by(month) %>%
   summarise(tmax = mean(tmax, na.rm = TRUE),
-            diff = mean(diff, na.rm = TRUE)) %>% 
-  ungroup(month) %>% 
+            tmin = mean(tmin, na.rm = TRUE),
+            diff = mean(diff, na.rm = TRUE),
+            avg = mean(avg, na.rm =TRUE)) %>%
+  ungroup(month) %>%
+  mutate(dummy_date = as.Date(glue::glue("2021-{month}-01"))) %>% 
   tamp() %>%
-  mutate(a = nrow(ts)) %>% 
-  filter(a == 12) %>% 
-  mutate(diff_winter = ts %>% filter(month %in% c(6,7,8)) %>% pull(diff) %>% mean(na.rm = TRUE),
-         diff_summer = ts %>% filter(month %in% c(12, 1,2)) %>% pull(diff) %>% mean(na.rm = TRUE),
-         diff_year = diff_winter - diff_summer) %>% 
-  stretch() %>% 
-  mutate(dummy_date = as.Date(glue::glue("2021-{month}-01"))) %>%
-  migrate(name) 
+  mutate(
+    diff_winter = ts %>% filter(month %in% c(6, 7, 8)) %>% pull(diff) %>% mean(na.rm = TRUE),
+    diff_summer = ts %>% filter(month %in% c(12, 1, 2)) %>% pull(diff) %>% mean(na.rm = TRUE),
+    diff_year = diff_winter - diff_summer,
+    annual = ts %>% filter(month %in% c(7, 1)) %>% pull(avg) %>% diff() %>% abs(),
+    diurnal = ts %>% pull(diff) %>% mean(na.rm = TRUE)
+  )
 
-long <- clean %>% 
+nested <- clean %>% 
   SharedData$new(~id, group = "cubble")
-
-nested <- clean %>% tamp() %>% as_tibble() %>% 
[…]

### 2021-12-17 scripts/interactive-linking2.R

@@ -1,0 +1,55 @@
+dt <- clean %>% tamp() 
+
+dt %>% 
+  ggplot() +
+  geom_sf(data = state_map, aes(geometry = geometry), fill = "transparent") +
+  geom_point(aes(x = long, y = lat, label = name, color = annual)) +
+  colorspace::scale_color_continuous_sequential("Red") + 
+  theme_bw() + 
+  theme(panel.grid.major = element_blank(),
+        legend.position = "bottom")
+
+nested <- clean %>% 
+  SharedData$new(~id, group = "cubble")
+
+long <- clean %>% stretch() %>% 
+  SharedData$new(~id, group = "cubble")
+
+
+p1 <- nested %>% 
+  ggplot() +
+  geom_sf(data = state_map, aes(geometry = geometry), fill = "transparent") +
+  geom_point(aes(x = long, y = lat,color = diurnal, label = name)) +
+  colorspace::scale_color_continuous_sequential(
+    "Oslo",
+    name = "Avg. temp. diff.") + 
+  labs(x = "Longitude", y = "Latitude") + 
+  coord_sf() + 
+  theme_bw() + 
+  theme(panel.grid.major = element_blank(),
+        legend.position = "bottom")
+
+p2 <- long %>% 
+  ggplot() +
+  geom_ribbon(aes(x = dummy_date, ymin = tmin, ymax = tmax, group = id), 
+              color = "grey30", fill = "grey70", alpha = 0.2) +
+  geom_point(aes(x = dummy_date, y = tmax), size = 0.1) + 
+  scale_x_date(date_labels = "%b", date_breaks = "1 month") + 
+  labs(x = "Month", y = "Temperature") + 
+  theme_bw() + 
+  theme(panel.grid.major = element_blank(),
+        legend.position = "bottom")
+
+# More adjustment:
+# - lighter color on the selected temperature band
+# - wider width for ts and narrow for the map
+# - color correspondence in the popup map
+
+out <- bscols(
+  ggplotly(p1, width = 900, height = 700, tooltip = "label") %>%
+    highlight(on = "plotly_selected", off = "plotly_deselect"),
+  ggplotly(p2, width = 850, height = 700, tooltip = "label") %>% 
+    highlight(on = "plotly_selected", off = "plotly_deselect", color = "grey30"),
+  widths = c(6, 6)
+)
+

### 2021-12-17 scripts/interactive-popup.R

@@ -35,7 +35,8 @@ for (i in 1:length(df_id)) {
     theme_bw() + 
     scale_x_date(date_labels = "%b", date_breaks = "1 month") + 
     labs(x = "Month", y = "Temperature") + 
-    ylim(-10, 50)
+    ylim(-10, 50) + 
+    ggtitle(paste0(stringr::str_to_title(unique(dt$name))))
   
   i <- i + 1
   

### 2021-12-20 main.Rmd
basic manip sexctionon almost done:

@@
 
-## Australia precipitation pattern in 2020
+## Australia historical maximum temperature
 
-### Basic manipulation
+Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world and the dataset `weatherdata::historical_tmax` extracts the historical maximum temperature recorded for `r nrow(weatherdata::historical_tmax)` Australian stations.
+The data `historical_tmax` is already presented as a cubble, with `id` as the key, `date` as the index, and `c("longitude", "latitude")` as the coordinates.
+Other variables include `elevation`, `name`, `wmo_id`, `first_year`, and  `last_year` in the nested form and `tmax` in the long form.
+This example compares the maximum temperature in two periods: 1971 - 1975 and 2016 - 2020 for stations in Victoria and New South Wales.
 
-Bureau of Meteorology (BoM) collects climate variables.
-This variables can be openly queried via `rnoaa`.
-`weatherdata` provides a readily available copy of several data.
-`weatherdata::historical_tmax` provides historical maximum temperature recorded for Australian stations with the earliest dating back to 1859 and this can be used to understand the change of maximum temperature from a past period to nowadays.
+Stations in the two states can be subsetted on the station number: Australia GHCN station number starts with `ASN00` and followed by the [Bureau of Meteorology (BOM) station number](http://www.bom.gov.au/climate/cdo/about/site-num.shtml), where the 2nd and 3rd digit (7th and 8th in the GHCN number) denote the state a station belongs to.
+New South Wales stations start from 46 to 75 and Victoria stations then follow from 76 to 90.
+Extracting Victoria and New South Wales stations is a filter operation in the spatial dimension and hence is operated in the nested form:
 
-Figure \ref{fig:basic-manip}.
+```{r eval = FALSE, echo = TRUE}
+tmax <- weatherdata::historical_tmax %>%
+  filter(between(stringr::str_sub(id, 7, 8), 46, 90))
+```
+
+The five year window is chosen to remove the effect of a particular year and the historical period of 1971 - 1975 is used since all the stations have records from 1970.
+This following chunk filters on records in the two study periods and summarises the maximum temperature into monthly measure.
+`stretch()` is used to convert the nested form cubble into the long form for these operations:
 
 ```{r eval=  FALSE, echo = TRUE}
-tmax <- weatherdata::historical_tmax %>%
-  # filter out VIC and NSW stations
-  filter(between(stringr::str_sub(id, 7, 8), 46, 90)) %>%
-
-  # filter out observations in between 1971-1975 and 2016 - 2020
+tmax <- tmax %>% 
   stretch() %>%
@@
   group_by(month, group) %>%
-  # summarise tmax into monthly mean for both periods
-  summarise(tmax = mean(tmax, na.rm = TRUE)) %>%
+  summarise(tmax = mean(tmax, na.rm = TRUE))
+```
 
-  # filter out stations that have complete records in both periods 
+A data quality issue with GHCN data is that while the first and last year of each series is provided, years missing in this period is not reported.
+There are a few stations which don't have records during 1971 - 1975 and these stations are filtered out by examining whether the summarised `tmax` has a total of 24 months.
+This is again a station-wise operation and is operated in the nested form, which is switched to from the long form by  `tamp()`:
+
+```{r eval = FALSE, echo = TRUE}
+tmax <- tmax %>% 
   tamp() %>%
-  filter(nrow(ts) == 24) %>%
+  filter(nrow(ts) == 24) 
+```
 
-  # migrate longitude and latitude to the long form for making the glyph map
+Lastly, to create a glyph map, both the major (`longitude`, `latitude`) and minor (`month`, `tmax`) coordinates need to be in the same table.
+Spatial variables can be moved to the long form with `migrate()`:
+
+```{r eval = FALSE, echo = TRUE}  
+tmax <- tmax %>%   
   stretch() %>%
   migrate(latitude, longitude)
-
-gly <- glyphs(tmax, "longitude", "month", "latitude", "tmax",
-              height = 0.6, width = 1)
-
-nsw_vic <- ozmaps::abs_ste %>% filter(NAME %in% c("Victoria", "New South Wales"))
-
-ggplot() + 
-  geom_sf(data = nsw_vic, fill = "transparent") + 
-  geom_path(data = gly, 
-            aes(gx, gy, group = interaction(id, group), color = group)) +
-  scale_color_brewer(palette = "Dark2") + 
-  theme_bw() + 
-  coord_sf(xlim = c(140, 155)) + 
-  theme(legend.position = "bottom")
 ```
 
-```{r basic-manip, out.width="100%", out.height="70%", fig.cap="asdfasd"}
+Figure \ref{fig:basic-manip} shows the glyph map made with the data after the wangling above.
+One issue with this map is that the similar pattern shown from a few nearby stations around Sydney (151E, 34S) and New castle (152E, 33S) can be distracting.
+Aggregation is a useful technique to observe the general pattern of a collection of series and will be the topic of the next example.
+
+```{r basic-manip, out.width="100%", out.height="70%", fig.cap="Glyph map of the mean maximum temperature by month for Victoria and New South Wales weather stations.
+On the top left corner is a detailed legend for station Cobar highlighted in the black box.
+Compared to 1971 - 1975,  the period 2016 - 2020 sees an increase of mean maximum temperature in spring and summer in Australia (end and beginning of the year).
+A larger increase during the first three months of the year is observed at stations with a lower latitude."}
 knitr::include_graphics("figures/basic-manip.png")
@@
 
-<!-- `historical_tmax` is already in a nested cubble format, which is suitable for station-wise manipulation.
-`rnoaa` construct the station id by prefix `ASN00` to the Bureau of Meteorology (BOM) station number.
-In [BOM's numbering system](http://www.bom.gov.au/climate/cdo/about/site-num.shtml), the 2nd and 3rd digit denotes the state a station is located in and this is equivalent to the 7th to 8th digit in our string.
-We can mutate/ filter station in a particular state based on this information and here we add a column called `state_id` and filter out the ones in New South Wales and Victoria (ranging from 46 to 90).
--->
+## Australia precipitation pattern in 2020
 
-<!-- There are some operations in the time dimension we would like to make: -->
-
-<!-- -   Extract observations in a particular period, say, those in 1971 to 1975 and 2016 to 2020. This can be used to compare the historical and recent climate. -->
-<!-- -   Summarise daily records into monthly to remove sparsity -->
-
-<!-- These time dimension operations can be computed in the long form and `stretch()` converts a nested cubble to a long cubble.
-From a long cubble, we can write the exact same dplyr codes to complete the two tasks: -->
-
-<!-- A data quality issue with the `rnoaa` data is that while it records the first and last year recorded of each series, it doesn't report the period of missingness.
-For example, station `ASN00047048` starts it first record in 1957, pauses for a period from 1963 to 1990, and then resumes it recording till today.
-Out of the `nrow(tmax_nested)` stations in New South Wales and Victoria, 7 stations have this issue and we would like to remove those stations from the comparison.
--->
-
-<!-- Again, this is a station-wise operation and to convert back from the long cubble to a nested one, use `tamp()`.
-Here we keep the stations with 24 observations (12 months for both periods) after the monthly aggregation.
--->
-
-<!-- In some visualisation, we may need information from both spatial and temporal dimension.
-One example of this is a glyph map, where spatial variables, i.e.
-`longitude` and `latitude`, are used to construct the major axes and temporal variables, i.e.
-`month` and `tmax`, are used to construct the minor axes.
-This requires these variables to be in the same table, rather than in different forms.
-In cubble, you can append spatial variables that are invariant to the `key` with `migrate()` -->
-
-### Aggregation
-
-In figure \ref{fig:basic-manip}, there are a few stations around Sydney (151E, 34S) and New castle (152E, 33S).
+In Figure \ref{fig:basic-agg}.
 Aggregating them into averages or other statistics would give a better picture.
 This can be done with cubble through a hierarchical structure, where the cluster indicator is one level above the individual station.
-
-In Figure \ref{fig:basic-agg}
 

### 2021-12-20 main.tex
basic manip sexctionon almost done:

@@
 
-\hypertarget{australia-precipitation-pattern-in-2020}{%
-\subsection{Australia precipitation pattern in
-2020}\label{australia-precipitation-pattern-in-2020}}
+\hypertarget{australia-historical-maximum-temperature}{%
+\subsection{Australia historical maximum
+temperature}\label{australia-historical-maximum-temperature}}
 
-\hypertarget{basic-manipulation}{%
-\subsubsection{Basic manipulation}\label{basic-manipulation}}
+Global Historical Climatology Network (GHCN) provides daily climate
+measures from stations across the world and the dataset
+\texttt{weatherdata::historical\_tmax} extracts the historical maximum
+temperature recorded for 236 Australian stations. The data
+\texttt{historical\_tmax} is already presented as a cubble, with
+\texttt{id} as the key, \texttt{date} as the index, and
+\texttt{c("longitude",\ "latitude")} as the coordinates. Other variables
+include \texttt{elevation}, \texttt{name}, \texttt{wmo\_id},
+\texttt{first\_year}, and \texttt{last\_year} in the nested form and
+\texttt{tmax} in the long form. This example compares the maximum
+temperature in two periods: 1971 - 1975 and 2016 - 2020 for stations in
+Victoria and New South Wales.
 
-Bureau of Meteorology (BoM) collects climate variables. This variables
-can be openly queried via \texttt{rnoaa}. \texttt{weatherdata} provides
-a readily available copy of several data.
-\texttt{weatherdata::historical\_tmax} provides historical maximum
-temperature recorded for Australian stations with the earliest dating
-back to 1859 and this can be used to understand the change of maximum
-temperature from a past period to nowadays.
-
-Figure \ref{fig:basic-manip}.
+Stations in the two states can be subsetted on the station number:
+Australia GHCN station number starts with \texttt{ASN00} and followed by
+the \href{http://www.bom.gov.au/climate/cdo/about/site-num.shtml}{Bureau
+of Meteorology (BOM) station number}, where the 2nd and 3rd digit (7th
+and 8th in the GHCN number) denote the state a station belongs to. New
+South Wales stations start from 46 to 75 and Victoria stations then
+follow from 76 to 90. Extracting Victoria and New South Wales stations
+is a filter operation in the spatial dimension and hence is operated in
+the nested form:
 
@@
 R> tmax <- weatherdata::historical_tmax %>%
-+   # filter out VIC and NSW stations
-+   filter(between(stringr::str_sub(id, 7, 8), 46, 90)) %>% 
-+   
-+   # filter out observations in between 1971-1975 and 2016 - 2020
++   filter(between(stringr::str_sub(id, 7, 8), 46, 90))
+\end{CodeInput}
+\end{CodeChunk}
+
+The five year window is chosen to remove the effect of a particular year
+and the historical period of 1971 - 1975 is used since all the stations
+have records from 1970. This following chunk filters on records in the
+two study periods and summarises the maximum temperature into monthly
+measure. \texttt{stretch()} is used to convert the nested form cubble
+into the long form for these operations:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> tmax <- tmax %>% 
 +   stretch() %>%
@@
 +   group_by(month, group) %>%
-+   # summarise tmax into monthly mean for both periods
-+   summarise(tmax = mean(tmax, na.rm = TRUE)) %>% 
-+   
-+   # filter out stations that have complete records in both periods 
++   summarise(tmax = mean(tmax, na.rm = TRUE))
+\end{CodeInput}
+\end{CodeChunk}
+
+A data quality issue with GHCN data is that while the first and last
+year of each series is provided, years missing in this period is not
+reported. There are a few stations which don't have records during 1971
+- 1975 and these stations are filtered out by examining whether the
+summarised \texttt{tmax} has a total of 24 months. This is again a
+station-wise operation and is operated in the nested form, which is
+switched to from the long form by \texttt{tamp()}:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> tmax <- tmax %>% 
 +   tamp() %>%
-+   filter(nrow(ts) == 24) %>% 
-+   
-+   # migrate longitude and latitude to the long form for making the glyph map
++   filter(nrow(ts) == 24) 
+\end{CodeInput}
+\end{CodeChunk}
+
+Lastly, to create a glyph map, both the major (\texttt{longitude},
+\texttt{latitude}) and minor (\texttt{month}, \texttt{tmax}) coordinates
+need to be in the same table. Spatial variables can be moved to the long
+form with \texttt{migrate()}:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> tmax <- tmax %>%   
 +   stretch() %>%
 +   migrate(latitude, longitude)
-R> 
-R> gly <- glyphs(tmax, "longitude", "month", "latitude", "tmax",
-+               height = 0.6, width = 1)
-R> 
-R> nsw_vic <- ozmaps::abs_ste %>% filter(NAME %in% c("Victoria", "New South Wales"))
-R> 
-R> ggplot() + 
-+   geom_sf(data = nsw_vic, fill = "transparent") + 
-+   geom_path(data = gly, 
-+             aes(gx, gy, group = interaction(id, group), color = group)) +
-+   scale_color_brewer(palette = "Dark2") + 
-+   theme_bw() + 
-+   coord_sf(xlim = c(140, 155)) + 
-+   theme(legend.position = "bottom")
 \end{CodeInput}
 \end{CodeChunk}
+
+Figure \ref{fig:basic-manip} shows the glyph map made with the data
+after the wangling above. One issue with this map is that the similar
+pattern shown from a few nearby stations around Sydney (151E, 34S) and
+New castle (152E, 33S) can be distracting. Aggregation is a useful
+technique to observe the general pattern of a collection of series and
+will be the topic of the next example.
 
@@
 
-\caption[asdfasd]{asdfasd}\label{fig:basic-manip}
+\caption[Glyph map of the mean maximum temperature by month for Victoria and New South Wales weather stations]{Glyph map of the mean maximum temperature by month for Victoria and New South Wales weather stations.
+On the top left corner is a detailed legend for station Cobar highlighted in the black box.
+Compared to 1971 - 1975,  the period 2016 - 2020 sees an increase of mean maximum temperature in spring and summer in Australia (end and beginning of the year).
+A larger increase during the first three months of the year is observed at stations with a lower latitude.}\label{fig:basic-manip}
 \end{figure}
@@
 
-\hypertarget{aggregation}{%
-\subsubsection{Aggregation}\label{aggregation}}
+\hypertarget{australia-precipitation-pattern-in-2020}{%
+\subsection{Australia precipitation pattern in
+2020}\label{australia-precipitation-pattern-in-2020}}
 
-In figure \ref{fig:basic-manip}, there are a few stations around Sydney
-(151E, 34S) and New castle (152E, 33S). Aggregating them into averages
-or other statistics would give a better picture. This can be done with
-cubble through a hierarchical structure, where the cluster indicator is
-one level above the individual station.
-
-In Figure \ref{fig:basic-agg}
+In Figure \ref{fig:basic-agg}. Aggregating them into averages or other
+statistics would give a better picture. This can be done with cubble
+through a hierarchical structure, where the cluster indicator is one
+level above the individual station.
 

### 2021-12-20 scripts/basic-manip.R

@@ -1,3 +1,5 @@
+library(tidyverse)
+library(GGally)
 tmax <- weatherdata::historical_tmax %>%
   # filter out VIC and NSW stations
   filter(between(stringr::str_sub(id, 7, 8), 46, 90)) %>% 
@@ -21,17 +23,55 @@ tmax <- weatherdata::historical_tmax %>%
   migrate(latitude, longitude)
 
 gly <- glyphs(tmax, "longitude", "month", "latitude", "tmax",
-              height = 0.6, width = 1)
+              height = 0.5, width = 1)
 
 nsw_vic <- ozmaps::abs_ste %>% filter(NAME %in% c("Victoria", "New South Wales"))
 
-ggplot() + 
-  geom_sf(data = nsw_vic, fill = "transparent") + 
+p1 <- ggplot() + 
+  geom_sf(data = nsw_vic, fill = "transparent", color = "grey", linetype = "dotted") + 
   geom_path(data = gly, 
             aes(gx, gy, group = interaction(id, group), color = group)) +
   scale_color_brewer(palette = "Dark2") + 
   theme_bw() + 
-  coord_sf(xlim = c(140, 155)) + 
-  theme(legend.position = "bottom")
+  coord_sf(xlim = c(141, 154)) + 
+  theme(legend.position = "bottom") +
+  labs(x = "Longitude", y = "Latitude")
+
+#################
+# inset
+tmax %>% filter(id == "ASN00048027") %>%
+  ggplot(aes(x = month,
+             y = tmax,
+             color = group)) +
+  geom_line(size = 1.5) +
+  scale_color_brewer(palette = "Dark2", guide = "none") +
+  scale_x_continuous(breaks = seq(1, 12, 1)) +
+  labs(x = "Month", y  = "Temperature", title = "ASN00048027: Cobar") +
+  theme_bw() +
+  theme(
+    aspect.ratio = 0.3,
+    axis.text = element_text(size = 20),
+    title =  element_text(size = 20)
+  )
+
+#ggsave(filename = here::here("figures/basic-manip-inset.png"), width = 7)
+
+box_df <- tmax %>% tamp() %>% filter(id == "ASN00048027")
+single <-
+  tibble::tibble(img = here::here("figures/basic-manip-inset.png"))
+p1 +
+  geom_rect(
+    data = box_df,
+    aes(
+      xmin = longitude - 0.6,
+      xmax = longitude + 0.6,
+      ymin = latitude - 0.12,
[…]

### 2021-12-20 main.Rmd
more writing and updates on the interactive graphic section

@@
 
-With a larger dataset, interactive graphics can be helpful to explore these data from multiple perspective.
-Two examples are presented here to show how interactivity can be used to explore the climate data across Australia.
-Again, `weatherdata::climate_full` will be used and we hope to include more than one time variable here.
-The difference between the maximum and minimum temperature of the day, the diurnal range, can be directly calculated from the data and it can have large impact on the wine industry (lol this is going far).
-This would be useful to compare the temperature difference across stations and to compare how this temperature difference differs throughout the year.
+To visualise climate variables from all the stations in Australia, interactive graphics can be helpful.
+Here a plotly and a leaflet example is given to explore the yearly diurnal temperature range pattern in Australia.
+This would be useful to see how the average temperature range differs in various latitudes in Australia, whether there's a summer or winter effect, any coastline, mountain or desert topological considerations that would affect the temperature.
 
-Temperature difference is a time-wise variable and should be computed in the long form cubble. The variables are averaged over the 2016 to 2020 period by month to stabilise the result.
+Again, `weatherdata::climate_full` will be first processed to produce summarised monthly maximum and minimum temperature across 2016 to 2020.
+Then a shared data with the same group is created from the nested and long cubble to enable the cross-talking via the common column `id`.
+The map and temperature band is created separately with the two shared data objects and combined together with `crosstalk::bscols`.
+A sketch of the code to produce Figure \ref{fig:interactive-linking} is presented below:
 
 ```{r eval = FALSE, echo = TRUE}
-res <- weatherdata::climate_full %>%
-  # filter(!id %in% c("ASN00067033", "ASN00072091", 
-  #                   "ASN00059040", "ASN00097053")) %>%
-  stretch() %>%
-  mutate(month = lubridate::month(date),
-         diff = tmax - tmin) %>%
-  group_by(month) %>%
-  summarise(tmax = mean(tmax, na.rm = TRUE),
-            diff = mean(diff, na.rm = TRUE)) %>%
-  ungroup(month)
+# data processing
+clean <- weatherdata::climate_full %>% ...
+
+# created grouped instance for crosstalk
+nested <- clean %>% SharedData$new(~id, group = "cubble")
+long <- stretch(clean) %>% SharedData$new(~id, group = "cubble")
+
+# create the map with SharedData nested, the time series with SharedData long
+p1 <- nested %>% ...
+p2 <- long %>% ...
+
+# Combine p1 and p2
+crosstalk::bscols(plotly::ggplotly(p1), plotly::ggplotly(p2), ...)
 
@@
 
-Several variables can be useful when comparing the seasonal effect of the temperature differences, for example, the difference of average temperature difference between summer and winter can tell [...].
-This is a station-wise measure and hence computed in the nested form.
+In Figure \ref{fig:interactive-linking}, the first row shows the default plot and when a selection is made, relevant component from both plots will be highlighted.
+For example, in the second row, the highest mean maximum temperature in July is selected on the time series and this corresponds to Kalumburu on the map where there is a larger diurnal temperature range in winter (June to August) than summer (December to February).
+The third row selects a station (Birdsville Airport) in the interior desert region that has a large average temperature difference on the map and the linked view shows a wide but constant diurnal temperature range throughout the year.
+The last row selects the lowest maximum temperature in July on the right plot and surprisingly, this links to the Thredbo airport at the boarder of Victoria and New South Wales, rather than any station in the Tasmania island!
 
-```{r eval = FALSE, echo = TRUE}
-res <- res %>% 
-  tamp() %>%
-  mutate(
-    diff_winter = ts %>% filter(month %in% c(6, 7, 8)) %>%
-      pull(diff) %>% mean(na.rm = TRUE),
-    diff_summer = ts %>% filter(month %in% c(12, 1, 2)) %>%
-      pull(diff) %>% mean(na.rm = TRUE),
-    diff_year = diff_winter - diff_summer
-  )
-```
-
-The result from the above manipulation can be directly supplied to make interactive plots and Figure \ref{fig:interactive-linking} is an example made with \pkg{plotly} and \pkg{crosstalk}.
-For example, the second row results from selecting the highest mean maximum temperature in July and this corresponds to Kalumburu where there is a larger winter (June to August) temperature difference than summer (December to February).
-The third row selects one of the stations (Birdsville Airport) with a large average temperature difference in the interior desert region on the map and the linked view from the time series shows that this maximum and minimum temperature difference is constant throughout the year.
-The last row selects the lowest maximum temperature in July on the temperature band and surprisingly, the corresponding station doesn't fall into the Tasmania island but Mountain Hotham in Victoria.
-
-```{r interactive-linking, echo = FALSE, out.width="100%", out.height="25%", fig.cap = "Screenshots of the mean daily temperature difference averaged by month from 2016 to 2020 at various locations in Australia.
+```{r interactive-linking, echo = FALSE, out.width="100%", out.height="25%", fig.retina = 2, dpi = 300, fig.cap = "Screenshots of the mean daily temperature difference averaged by month from 2016 to 2020 at various locations in Australia.
 Crossed linking between the map and the temperature band is implemented in both directions for exploring the temperature pattern interactively.
@@
 
-Cubble is also compatible with \pkg{leaflet} and with leaflet extension package \pkg{leafpop}, the temperature band can be displayed upon clicking on the station on the map.
+With leaflet popup, the temperature range can be displayed as a small plot upon clicking on the station on the map.
+This would require first creating all the popup plot separately in a vector and then add these plots to a leaflet map with `leafpop::addPopupGraphs()`:
+
+```{r eval = FALSE, echo = TRUE}
+# create subplots for each station
+df_id <- unique(clean$id)
+p <- map(1:length(df_id), function(i){
+  dt <- clean %>% filter(id == df_id[i])
+  dt %>% ...
+})
+
+# inset subplots as a popup in leaflet
+leaflet(clean) %>% 
+  addTiles() %>% 
+  addCircleMarkers(group = "a", ...) %>% 
+  leafpop::addPopupGraphs(graph = p, ...)
+
+```
+
 Figure \ref{fig:interactive-popup} shows more examples of temperature pattern that can be found in Australia.
-With the same latitude as Kalumburu but in the Pacific coastline, Cooktown airport shows a similarly high but constant temperature throughout the year.
-With a slightly higher latitude than the mountain Hotham, flat land in Melbourne airport shows a less extreme temperature band.
-Lastly, the Cape Grim Baps, in the north-west tip of the Tasmania island shows a uniformly low and thin temperature band.
+With the same latitude as Kalumburu but in the Pacific coastline, Cooktown airport shows a similarly high but constant temperature range throughout the year.
+With a slightly higher latitude than the Thredbo airport , Melbourne airport shows a less extreme temperature pattern Lastly, the Cape Grim Baps, in the north-west tip of the Tasmania island shows low and narrow temperature band, constantly betwee 10 and 20 degree throughout the year.
 
-```{r interactive-popup, echo = FALSE, out.width="50%", out.height="30%", fig.cap = "Screenshots of the diurnal temperature range averaged by month from 2016 to 2020 at various locations in Australia.
+```{r interactive-popup, echo = FALSE, out.width="45%", out.height="30%", fig.retina = 2, dpi = 300, fig.cap = "Screenshots of the diurnal temperature range averaged by month from 2016 to 2020 at various locations in Australia.
 The leaflet popup allows users to click on a single station for its temperature band as a small plot in the popup window.

### 2021-12-20 main.tex
more writing and updates on the interactive graphic section

@@
 
-With a larger dataset, interactive graphics can be helpful to explore
-these data from multiple perspective. Two examples are presented here to
-show how interactivity can be used to explore the climate data across
-Australia. Again, \texttt{weatherdata::climate\_full} will be used and
-we hope to include more than one time variable here. The difference
-between the maximum and minimum temperature of the day, the diurnal
-range, can be directly calculated from the data and it can have large
-impact on the wine industry (lol this is going far). This would be
-useful to compare the temperature difference across stations and to
-compare how this temperature difference differs throughout the year.
+To visualise climate variables from all the stations in Australia,
+interactive graphics can be helpful. Here a plotly and a leaflet example
+is given to explore the yearly diurnal temperature range pattern in
+Australia. This would be useful to see how the average temperature range
+differs in various latitudes in Australia, whether there's a summer or
+winter effect, any coastline, mountain or desert topological
+considerations that would affect the temperature.
 
-Temperature difference is a time-wise variable and should be computed in
-the long form cubble. The variables are averaged over the 2016 to 2020
-period by month to stabilise the result.
+Again, \texttt{weatherdata::climate\_full} will be first processed to
+produce summarised monthly maximum and minimum temperature across 2016
+to 2020. Then a shared data with the same group is created from the
+nested and long cubble to enable the cross-talking via the common column
+\texttt{id}. The map and temperature band is created separately with the
+two shared data objects and combined together with
+\texttt{crosstalk::bscols}. A sketch of the code to produce Figure
+\ref{fig:interactive-linking} is presented below:
 
@@
 \begin{CodeInput}
-R> res <- weatherdata::climate_full %>%
-+   # filter(!id %in% c("ASN00067033", "ASN00072091", 
-+   #                   "ASN00059040", "ASN00097053")) %>%
-+   stretch() %>%
-+   mutate(month = lubridate::month(date),
-+          diff = tmax - tmin) %>%
-+   group_by(month) %>%
-+   summarise(tmax = mean(tmax, na.rm = TRUE),
-+             diff = mean(diff, na.rm = TRUE)) %>%
-+   ungroup(month)
+R> # data processing
+R> clean <- weatherdata::climate_full %>% ...
+R> 
+R> # created grouped instance for crosstalk
+R> nested <- clean %>% SharedData$new(~id, group = "cubble")
+R> long <- stretch(clean) %>% SharedData$new(~id, group = "cubble")
+R> 
+R> # create the map with SharedData nested, the time series with SharedData long
+R> p1 <- nested %>% ...
+R> p2 <- long %>% ...
+R> 
+R> # Combine p1 and p2
+R> crosstalk::bscols(plotly::ggplotly(p1), plotly::ggplotly(p2), ...)
 \end{CodeInput}
@@
 
-Several variables can be useful when comparing the seasonal effect of
-the temperature differences, for example, the difference of average
-temperature difference between summer and winter can tell {[}\ldots{]}.
-This is a station-wise measure and hence computed in the nested form.
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> res <- res %>% 
-+   tamp() %>%
-+   mutate(
-+     diff_winter = ts %>% filter(month %in% c(6, 7, 8)) %>%
-+       pull(diff) %>% mean(na.rm = TRUE),
-+     diff_summer = ts %>% filter(month %in% c(12, 1, 2)) %>%
-+       pull(diff) %>% mean(na.rm = TRUE),
-+     diff_year = diff_winter - diff_summer
-+   )
-\end{CodeInput}
-\end{CodeChunk}
-
-The result from the above manipulation can be directly supplied to make
-interactive plots and Figure \ref{fig:interactive-linking} is an example
-made with \pkg{plotly} and \pkg{crosstalk}. For example, the second row
-results from selecting the highest mean maximum temperature in July and
-this corresponds to Kalumburu where there is a larger winter (June to
-August) temperature difference than summer (December to February). The
-third row selects one of the stations (Birdsville Airport) with a large
-average temperature difference in the interior desert region on the map
-and the linked view from the time series shows that this maximum and
-minimum temperature difference is constant throughout the year. The last
-row selects the lowest maximum temperature in July on the temperature
-band and surprisingly, the corresponding station doesn't fall into the
-Tasmania island but Mountain Hotham in Victoria.
+In Figure \ref{fig:interactive-linking}, the first row shows the default
+plot and when a selection is made, relevant component from both plots
+will be highlighted. For example, in the second row, the highest mean
+maximum temperature in July is selected on the time series and this
+corresponds to Kalumburu on the map where there is a larger diurnal
+temperature range in winter (June to August) than summer (December to
+February). The third row selects a station (Birdsville Airport) in the
+interior desert region that has a large average temperature difference
+on the map and the linked view shows a wide but constant diurnal
+temperature range throughout the year. The last row selects the lowest
+maximum temperature in July on the right plot and surprisingly, this
+links to the Thredbo airport at the boarder of Victoria and New South
+Wales, rather than any station in the Tasmania island!
 
@@
 
-Cubble is also compatible with \pkg{leaflet} and with leaflet extension
-package \pkg{leafpop}, the temperature band can be displayed upon
-clicking on the station on the map. Figure \ref{fig:interactive-popup}
-shows more examples of temperature pattern that can be found in
-Australia. With the same latitude as Kalumburu but in the Pacific
-coastline, Cooktown airport shows a similarly high but constant
-temperature throughout the year. With a slightly higher latitude than
-the mountain Hotham, flat land in Melbourne airport shows a less extreme
-temperature band. Lastly, the Cape Grim Baps, in the north-west tip of
-the Tasmania island shows a uniformly low and thin temperature band.
+With leaflet popup, the temperature range can be displayed as a small
+plot upon clicking on the station on the map. This would require first
+creating all the popup plot separately in a vector and then add these
+plots to a leaflet map with \texttt{leafpop::addPopupGraphs()}:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> # create subplots for each station
+R> df_id <- unique(clean$id)
+R> p <- map(1:length(df_id), function(i){
++   dt <- clean %>% filter(id == df_id[i])
++   dt %>% ...
++ })
+R> 
+R> # inset subplots as a popup in leaflet
+R> leaflet(clean) %>% 
++   addTiles() %>% 
++   addCircleMarkers(group = "a", ...) %>% 
++   leafpop::addPopupGraphs(graph = p, ...)
+\end{CodeInput}
+\end{CodeChunk}
+
+Figure \ref{fig:interactive-popup} shows more examples of temperature
+pattern that can be found in Australia. With the same latitude as
+Kalumburu but in the Pacific coastline, Cooktown airport shows a
+similarly high but constant temperature range throughout the year. With
+a slightly higher latitude than the Thredbo airport , Melbourne airport
+shows a less extreme temperature pattern Lastly, the Cape Grim Baps, in
+the north-west tip of the Tasmania island shows low and narrow
+temperature band, constantly betwee 10 and 20 degree throughout the
+year.
 
@@
 
-{\centering \includegraphics[width=0.5\linewidth,height=0.3\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/popup-north} \includegraphics[width=0.5\linewidth,height=0.3\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/popup-vic} \includegraphics[width=0.5\linewidth,height=0.3\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/popup-tas}
+{\centering \includegraphics[width=0.45\linewidth,height=0.3\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/popup-north} \includegraphics[width=0.45\linewidth,height=0.3\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/popup-vic} \includegraphics[width=0.45\linewidth,height=0.3\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/popup-tas}
 

### 2021-12-20 scripts/interactive-linking2.R

@@ -1,5 +1,5 @@
-dt <- clean %>% tamp() 
-
+dt <- clean 
+state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)
 dt %>% 
   ggplot() +
   geom_sf(data = state_map, aes(geometry = geometry), fill = "transparent") +
@@ -31,9 +31,10 @@ p1 <- nested %>%
 
 p2 <- long %>% 
   ggplot() +
-  geom_ribbon(aes(x = dummy_date, ymin = tmin, ymax = tmax, group = id), 
-              color = "grey30", fill = "grey70", alpha = 0.2) +
-  geom_point(aes(x = dummy_date, y = tmax), size = 0.1) + 
+  # geom_ribbon(aes(x = dummy_date, ymin = tmin, ymax = tmax, group = id), 
+  #             color = "grey30", fill = "grey30", alpha = 0.2) +
+  geom_line(aes(x = dummy_date, y = tmax, group = id), size = 0.1) + 
+  geom_line(aes(x = dummy_date, y = tmin, group = id), size = 0.1) + 
   scale_x_date(date_labels = "%b", date_breaks = "1 month") + 
   labs(x = "Month", y = "Temperature") + 
   theme_bw() + 
@@ -48,8 +49,8 @@ p2 <- long %>%
 out <- bscols(
   ggplotly(p1, width = 900, height = 700, tooltip = "label") %>%
     highlight(on = "plotly_selected", off = "plotly_deselect"),
-  ggplotly(p2, width = 850, height = 700, tooltip = "label") %>% 
-    highlight(on = "plotly_selected", off = "plotly_deselect", color = "grey30"),
-  widths = c(6, 6)
+  ggplotly(p2, width = 1200, height = 700, tooltip = "label") %>% 
+    highlight(on = "plotly_selected", off = "plotly_deselect", color = "red"),
+  widths = c(5, 5)
 )
 

### 2021-12-20 scripts/interactive-popup.R

@@ -12,6 +12,7 @@ clean <-  weatherdata::climate_full %>%
   summarise(tmax = mean(tmax, na.rm = TRUE),
             diff = mean(diff, na.rm = TRUE)) %>% 
   ungroup(month) %>% 
+  mutate(dummy_date = as.Date(glue::glue("2021-{month}-01"))) %>% 
   tamp() %>%
   mutate(a = nrow(ts)) %>% 
   filter(a == 12) %>% 
@@ -24,7 +25,19 @@ clean <-  weatherdata::climate_full %>%
 
 i <- 1
 df_id <- clean$id %>% unique()
-p <- vector("list", length(df_id))
+p <- purrr::map(1:length(df_id), function(i){
+  dt <- clean %>% filter(id == df_id[i])
+  dt %>% 
+    ggplot(aes(x = dummy_date, ymin = tmax - diff, ymax = tmax, group = id)) + 
+    geom_ribbon(color = "grey50", fill = "grey50", alpha = 0.5) + 
+    theme_bw() + 
+    scale_x_date(date_labels = "%b", date_breaks = "1 month") + 
+    labs(x = "Month", y = "Temperature") + 
+    ylim(-10, 50) + 
+    ggtitle(paste0(stringr::str_to_title(unique(dt$name))))
+})
+
+
 for (i in 1:length(df_id)) {
   dt <- clean %>% 
     filter(id == df_id[i]) 
@@ -44,8 +57,8 @@ for (i in 1:length(df_id)) {
 
 nested <- tamp(clean) 
 domain <- nested$diff_year
-pal <- colorNumeric(colorspace::diverging_hcl(palette = "Purple-Green", 
-                                              n = 11, l2 = 80),
+pal <- colorNumeric(colorspace::sequential_hcl(palette = "Oslo", 
+                                              n = 11),
                     domain = domain)
 
 out <- leaflet(nested) %>% 

### 2022-01-07 1-intro.Rmd
play aroudnd with child rmd to organising sections

--- 2021-12-20 main.Rmd
@@
 ---
-documentclass: jss
-author:
-  - name: H. Sherry Zhang
-    # department: Department of Econometrics and Business Statistics
-    affiliation: Monash University
-    #location: Melbourne, Australia
-    email: \email{huize.zhang@monash.edu}
-  - name: Dianne Cook
-    # department: Department of Econometrics and Business Statistics
-    affiliation: 'Monash University \AND'
-    # location: Melbourne, Australia
-    # email:  dicook@monash.edu
-  - name: Ursula Laa
-    # department: Institute of Statistics
-    affiliation: 'University of Natural Resources and Life Sciences \AND'
-    # location: Vienna, Austria
-    # email:  ursula.laa@boku.ac.at
-  - name: Nicolas Langrené
-    # department: 34 Village Street, Docklands VIC 3008 Australia
-    affiliation: 'CSIRO Data61 \AND'
-    # location: Melbourne, Australia
-    # email: nicolas.langrene@csiro.au
-  - name: Patricia Menéndez
-    # department: Department of Econometrics and Business Statistics
-    affiliation: 'Monash University \AND'
-    # location: Melbourne, Australia
-    # email:  patricia.menendez@monash.edu
-title:
-  formatted: "\\pkg{cubble}: An R Package for Structuring Spatio-temporal Data"
-  # If you use tex in the formatted title, alsoge supply version without
-  plain:     "cubble: An R Package for Structuring Spatio-temporal Data"
-  # For running headers, if needed
-  # short:     "\\pkg{foo}: A Capitalized Title"
-abstract: >
-  The abstract of the article.
-keywords:
-  # at least one keyword must be supplied
-  formatted: [spatio-temporal data,  "\\proglang{R}"]
-  plain:     [spatio-temporal data, R]
-preamble: >
-  \usepackage{amsmath}
-  \usepackage{array}
-output: rticles::jss_article
-bibliography: references.bib
+title: "1-intro"
+output: pdf_document
 ---
-
-```{r setup, echo = FALSE}
-knitr::opts_chunk$set(
-  echo = FALSE,
-  warning = FALSE,
-  message = FALSE,
-  cache = TRUE,
-  fig.path = "figures/",
-  fig.height = 10
-)
-options(prompt = "R> ", continue = "+ ", 
-        tibble.print_max = 5, tibble.print_min = 5)
-```
-
-```{r echo = FALSE}
-library(cubble)
-library(forcats)
-library(stringr)
-library(dplyr)
-library(lubridate)
-library(ggplot2)
-library(GGally)
-library(plotly)
-library(crosstalk)
-library(patchwork)
-```
-
-\newpage
 
@@
 \newline -->
-
-\newpage
-
-# The cubble package
-
-Spatio-temporal data usually come in various forms and Figure \ref{fig:illu-input} shows four examples of this.
-No matter which form the data is in, these formats share some common components that characterise spatio-temporal data.
-A spatial identifier (`id` in the diagram) is the unique identifier of each site.
-The temporal identifier (`t` in the diagram) prescribes the time stamp each site is recorded.
-Coordinates, comprising of latitude and longitude (`lon` and `lat` in the diagram), locates each site on the map.
-These identifiers will be the building blocks for the data structure introduced below.
-Other variables in the data can be categorised into two groups: spatial variables that are invariant at each time stamp for every site, i.e.
-the name or code of the weather station and temporal variables that varies with time.
-
-```{r illu-input, echo = FALSE, out.width = "100%", fig.cap="Illustration of incoming data formats for spatio-temporal data.
-(1) Data comes in as a single table; (2) Separate tables for spatial and temporal variables; (3) A single table with all the parameters used to query the database and a separate table for queried data; and (4) Cubical data in array or NetCDF format.", fig.align='center', out.height = "15%"}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.001.png"))
-```
-
-In a cubble, there are two forms: nested form and long form, and Figure \ref{fig:illu-cubble} sketches the two forms along with the associated attributes.
-The decision on which form to use is output-oriented, meaning analysts need to first think about whether the output of a particular operation is identified only by the spatial identifier, or a combination of spatial and temporal identifier.
-The nested cubble is suitable for working with operations that are only identified by site and this type of operation can be a pure manipulation of spatial variables, or a summary of temporal variables by site (i.e.
-the output of counting the number of raining day is only identified by sites and hence should be performed with the nested form).
-Underneath the nested form, a cubble is built from a row-wise dataframe (`rowwise_df`) where each site forms a separate group.
-This structure simplifies the calculation that involves temporal variables by avoiding the use of `map` syntax when working with list-column.
-
-For those operations whose output involves both a spatial and temporal dimension, long form should be used.
-The long form is identified by both the spatial and temporal identifier and adopts a grouped dataframe (`grouped_df`) to forms each site as a group.
-Spatial variables are stored separately in a \pkg{tibble} as an special attribute of the long cubble.
-This design avoids repeating the spatial variables at each time stamp while not dropping information from spatial variables.
-
-```{r illu-cubble, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "Illustration of nested and long cubble."}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.002.png"))
-```
-
-## Create a cubble in the nested form
-
-To use functionalities from cubble, data analysts first need to create a cubble.
-`as_cubble` create a \pkg{cubble} by supplying the three key components: `key` as the spatial identifier; `index` as the temporal identifier; and a vector of `coords` in the order of longitude first and then latitude.
-The naming of `key` and `index` follows the convention in the \pkg{tsibble} package.
-The cubble created by default is in the nested form.
-Below is an example of creating a cubble: \newline
-
-```{r}
-(cubble_nested <- cubble::climate_flat %>%
-  as_cubble(key = id, index = date, coords = c("long", "lat")))
-```
-
-There are a few information in the \pkg{cubble} header: the name of the `key` variable, bbox, and also the name of variable nested in the `ts` column.
-In this example, each site is identifier is `id` and the number in the bracket means there are 5 unique `id` in this dataset.
-The bbox in the second row gives the range of the coordinates.
-The temporal variables are all nested in the `ts` column, but it could be useful to know the name these variables.
-The third row in the cubble header shows these names and in this example this includes: precipitation, `prcp`, maximum temperature, `tmax`, and minimum temperature, `tmin`.
-
-## Stretch a nested cubble into the long form
-
-The long cubble is suitable to manipulate the time dimension of the data.
-The function `stretch()` switches the nested cubble into the long cubble by first extracts all the spatial variables into a separate tibble and store in the `spatial` attribute and then unnests the `ts` column:
-
-```{r}
-(cubble_long <- cubble_nested %>% stretch(ts))
-```
-
-Notice here that the third line in the header now shows the name of spatial variables rather than the temporal variables.
-
-## Tamp a long cubble back to the nested form
-
-Manipulation on the spatial and temporal dimension can be an iterative process.
-Many times, analysts will need to go back and forth between the nested and long cubble.
-The `stretch()` function introduced in the previous section switches a nested cubble into a long cubble and function `tamp()` is its inverse function to switch a long cubble back to the nested cubble:
-
-```{r}
-(cubble_back <- cubble_long %>% tamp())
-```
-
-## Migrate spatial variables to a long cubble
-
-As a final data output for modelling or visualisation, spatio-temporal data is usually expected to be in a single table.
-Function `migrate()` moves the spatial variables from the `spatial` attribute into the long cubble:
-
-```{r}
-(cubble_long %>% migrate(long, lat))
-```
-
-In this workflow described above, data objects come into cubble in the nested form, then various operations on the spatial and temporal dimension can go back and forth between the nested and long form, and finally, the data will come out of cubble in the long form for further modelling or visualisation.
-
-Building from an underlying `tbl_df` structure, it is natural to implement methods available in `dplyr` to `cubble`. Supported methods in the `cubble` with `dplyr` generics includes:
-
-```{=tex}
-\begin{center}
-\begin{tabular}{ | m{5em} | m{15cm}| } 
-basics & \textbf{mutate}, \textbf{filter}, \textbf{summarise}, \textbf{select}, \textbf{arrange}, \textbf{rename}, \textbf{left\_join} \\
-grouping &  \textbf{group\_by}, \textbf{ungroup}\\
-slice family & \textbf{slice\_head}, \textbf{slice\_tail}, \textbf{slice\_sample}, \textbf{slice\_min} and \textbf{slice\_max} \\
-\end{tabular}
-\end{center}
-```
-\pkg{cubble} is also compatible with \pkg{tsibble} in the sense that the original list-column can be a `tbl_ts` object.
-Duplicates and gaps should be first checked before structuring the data into a cubble.
-If the input data is a \pkg{tsibble} object, the long form cubble is also a \pkg{tsibble} where users can directly apply time series operations.
-
-# Advanced features/ considerations
-
-## Hierarchical structure
-
-Imposing a clustering structure can be thought of as building a hierarchical structure where stations are nested within clusters.
-This can be useful when there's intrinsic nesting structure in the data (i.e.
-country nested in the continent, county nested in the state) or there's some clustering.
-When we have access to both level, `switch_key()` is the function to re-structure the data as one cluster per row.
-Temporal observations from different stations while within the same cluster are bound in the nested column `ts`.
-
-One thing we hope to do with the cluster is to find the coordinates of the centroid.
-These are variables variant to the station but invariant to the cluster and it would be nice to have a function that structure each cluster as a row.
-`switch_key()` is the function that does this: it lets you to specify a new key, say `cluster` and nests all spatial variables variant to `cluster` into a column.
-Spatial variables are all nested inside a new column `.val`.
-Temporal observations from different stations while within the same cluster are bound in the nested column `ts`.
-
-This structure makes it easy to compute cluster level variable, for example, the convex hull and the centroid coordinate of each cluster.
-These can be amended into the nested form with function `get_centroid`.
-
-After we have got `cluster_nested`, spatial and temporal data at both levels can be easily obtained.
-Figure \ref{fig:illu-hier} illustrate the relationship between the long and nested form cubble at both site and cluster level.
-More description on this.
-Start with the original `station_nested`, `stretch()` expands the `ts` column with each station (`id`) forming a group and attach variables invariant to `id` as an attribute.
-`switch_key()` changes the `key` from `id` to `cluster` and nests all the spatial variables that variant to `cluster`.
-`stretch()` `cluster_nested` will store variables that are invariant to `cluster` as a tibble in the attribute.
-
-this fit into the remaining data pipeline.
-
-```{r illu-hier, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "Hierarchical structure"}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.003.png"))
-```
-
-## Data fusion and matching
-
-Temporal matching checks how spatially matched pairs align temporally. We use the following chart to illustrate how the temporal matching works:
-
-For each spatially matched pair, say `A` and `a`, we first find the largest `n` points in each series, colored in brown points here.
-Here we use the largest three but you can tune this number by `temporal_n_highest`.
-Then we construct the interval of the largest points from one series and see how many points, from the other series, fall into the intervals.
-The series used to construct the interval is controlled by `temporal_independent` and the window size by `temporal_window` with a default of 5.
-
-In this illustration, we construct the interval based on series `A` and two of the three peaks from `a` falls into this interval at Time 7 and 27.
-
-There's another mandatory argument that hasn't been introduced above: `temporal_var_to_match`.
-This argument controls the variable to match and it needs to appear in both the `major` and `minor` set.
-In the water level matching example, we match the variable `Water_course_level` from `river` to `prcp` from `climate`, hence need to manually rename one of them to match the other, here we rename `Water_course_level` to `prcp` in `river`:
-
-Figure \ref{fig:illu-matching}
-
-```{r illu-matching, out.width="100%", out.height="30%", fig.cap="sdfasdf"}
-knitr::include_graphics("figures/illu-matching.png")
-```
-
-## Interactive graphics
-
-<!-- Interactive graphics can listen to users' actions on the plot to provide additional information that facilitates data exploration.
-This is a useful technique for spatio-temporal data since users can zoom or pan the map to view the local and global structure of the map; use tooltips or popups to query more information about a graphic element; or highlight points to explore its linked views in other plots.
-In the R community, many implementations have been developed to connect \proglang{R} to \proglang{javascript} to create interactive graphics.
-In relation to spatio, temporal, and spatio-temporal data, the general purpose packages \pkg{plotly} [@plotly] and \pkg{leaflet} [@leaflet] realise various interactive actions through their corresponding javascript libraries.
-\pkg{crosstalk} [@crosstalk] and \pkg{tsibbletalk} [@tsibbletalk] implement brushed linking between htmlwidgets.
-\pkg{ggiraph} [@ggiraph] enables tooltip, self-linking, and customised actions specified through its own \proglang{javascript}.
--->
-
-<!-- While many graphic implementations present worked examples to illustrate the usage of the package, few documents the underlying pipeline that transforms the raw data step-by-step to the final view on the screen.
-There have been some early work in building the data pipeline for (interactive) graphics [@buja1988elements; @buja1996interactive;  @sutherland2000orca] and more recent discussions include @wickham2009plumbing, @xie2014reactive, and @cheng2016enabling.
--->
-
-The cubble structure fits in naturally with the interactive graphic pipeline discussed in the literature [@buja1988elements; @buja1996interactive;  @sutherland2000orca; @xie2014reactive; @cheng2016enabling].
-Diagram \ref{fig:illu-interactive} illustrates how linking works with the two forms in a cubble, where a time series plot is created with the long cubble and a map is created with the nested cubble.
-When a user action is captured from the map, the site will be activated in the nested cubble.
-Then, the nested cubble will communicate to the long cubble to activate all the observations with the same `id`.
-The long cubble will then highlight the  activated series in the time series plot.
-
-The linking is also available from the time series plot to the map.
-The selection on the time series is through selecting the point on the time series and once a point is selected, it will be activated in the long cubble.
-All the observations that share the same `id`,  either in the long and nested cubble, are then activated.
-This includes other points in the same time series in the long cubble and the corresponding observation of site in the nested cubble.
-These activated observations will then being reflected in the updated plots and Diagram \ref{fig:illu-interactive-2} in the Appendix illustrates this process.
-
-```{r illu-interactive, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "demon interactivity"}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.004.png"))
-```
-
-\newpage
-
-# Examples
-
-## Australia historical maximum temperature
-
-Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world and the dataset `weatherdata::historical_tmax` extracts the historical maximum temperature recorded for `r nrow(weatherdata::historical_tmax)` Australian stations.
-The data `historical_tmax` is already presented as a cubble, with `id` as the key, `date` as the index, and `c("longitude", "latitude")` as the coordinates.
-Other variables include `elevation`, `name`, `wmo_id`, `first_year`, and  `last_year` in the nested form and `tmax` in the long form.
-This example compares the maximum temperature in two periods: 1971 - 1975 and 2016 - 2020 for stations in Victoria and New South Wales.
-
-Stations in the two states can be subsetted on the station number: Australia GHCN station number starts with `ASN00` and followed by the [Bureau of Meteorology (BOM) station number](http://www.bom.gov.au/climate/cdo/about/site-num.shtml), where the 2nd and 3rd digit (7th and 8th in the GHCN number) denote the state a station belongs to.
-New South Wales stations start from 46 to 75 and Victoria stations then follow from 76 to 90.
-Extracting Victoria and New South Wales stations is a filter operation in the spatial dimension and hence is operated in the nested form:
-
-```{r eval = FALSE, echo = TRUE}
-tmax <- weatherdata::historical_tmax %>%
-  filter(between(stringr::str_sub(id, 7, 8), 46, 90))
-```
-
-The five year window is chosen to remove the effect of a particular year and the historical period of 1971 - 1975 is used since all the stations have records from 1970.
-This following chunk filters on records in the two study periods and summarises the maximum temperature into monthly measure.
-`stretch()` is used to convert the nested form cubble into the long form for these operations:
-
-```{r eval=  FALSE, echo = TRUE}
-tmax <- tmax %>% 
-  stretch() %>%
-  filter(lubridate::year(date) %in% c(1971:1975, 2016:2020)) %>%
-  mutate(month = lubridate::month(date), 
-         group = as.factor(ifelse(lubridate::year(date) > 2015, 
-                                  "2016 ~ 2020", "1971 ~ 1975"))) %>%
-  group_by(month, group) %>%
-  summarise(tmax = mean(tmax, na.rm = TRUE))
-```
-
-A data quality issue with GHCN data is that while the first and last year of each series is provided, years missing in this period is not reported.
-There are a few stations which don't have records during 1971 - 1975 and these stations are filtered out by examining whether the summarised `tmax` has a total of 24 months.
-This is again a station-wise operation and is operated in the nested form, which is switched to from the long form by  `tamp()`:
-
-```{r eval = FALSE, echo = TRUE}
-tmax <- tmax %>% 
-  tamp() %>%
-  filter(nrow(ts) == 24) 
-```
-
-Lastly, to create a glyph map, both the major (`longitude`, `latitude`) and minor (`month`, `tmax`) coordinates need to be in the same table.
-Spatial variables can be moved to the long form with `migrate()`:
-
-```{r eval = FALSE, echo = TRUE}  
-tmax <- tmax %>%   
-  stretch() %>%
-  migrate(latitude, longitude)
-```
-
-Figure \ref{fig:basic-manip} shows the glyph map made with the data after the wangling above.
-One issue with this map is that the similar pattern shown from a few nearby stations around Sydney (151E, 34S) and New castle (152E, 33S) can be distracting.
-Aggregation is a useful technique to observe the general pattern of a collection of series and will be the topic of the next example.
-
-```{r basic-manip, out.width="100%", out.height="70%", fig.cap="Glyph map of the mean maximum temperature by month for Victoria and New South Wales weather stations.
-On the top left corner is a detailed legend for station Cobar highlighted in the black box.
-Compared to 1971 - 1975,  the period 2016 - 2020 sees an increase of mean maximum temperature in spring and summer in Australia (end and beginning of the year).
-A larger increase during the first three months of the year is observed at stations with a lower latitude."}
-knitr::include_graphics("figures/basic-manip.png")
-```
-
-## Australia precipitation pattern in 2020
-
-In Figure \ref{fig:basic-agg}.
-Aggregating them into averages or other statistics would give a better picture.
-This can be done with cubble through a hierarchical structure, where the cluster indicator is one level above the individual station.
-
-The `weatherdata::climate_full` data has daily climate data of `r nrow(weatherdata::climate_full)` Australia stations from 2016 to 2020.
-
-calcluate distance matrix
-As an example to illustrate here, a kmean clustering algorithm based on the distance matrix is used and the number of centres is set to 20.
-More complex algorithms can also be used for more complex problem, as long as a mapping from each station id to the cluster id can be constructed.
-
-<!-- Now we can obtain the aggregated series as described in the workflow diagram above and construct the glyph map with `GGally::glyphs()`: -->
-
-<!-- We can also look at the precipitation of each individual station within the same cluster: -->
-
-<!-- Lastly, there is one series on Tasmania island standing out from others, lets look at where it is: -->
-
-```{r basic-agg, out.width="100%", out.height="90%", fig.cap="sdfasdf"}
-knitr::include_graphics("figures/basic-agg.png")
-```
-
-## River level data in Victria water gauges
-
-The water level data comes from [Bureau of Meteorology](http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full) and has a copy in `weatherdata`.
-Here we extract the water course level and add a column annotate this data of type `river`.
-For the rainfall data, we will still use the `weatherdata::climate_full`, filtering for Victorian stations in 2020 should be pretty familiar by now.
-Again, we first look at where these stations are on the map first:
-
-Now we use `match_sites()` to first pair the weather stations with the river gauges spatially and then apply the temporal matching on `prcp`.
-We will construct the interval based on peaks in `climate` since we would expect a lag effect for precipitation to flow into the river and cause a raise in river level, hence `temporal_independent = climate`.
-We select the 30 highest peak from the series to construct the match by setting `temporal_n_highest = 30`.
-This is a tuning parameter and you can start with 10% of the points of one series (here we have daily data for a year, 10% is roughly 30 points).
-`temporal_min_match` filters out pairs don't have enough match and to return all the pairs, set `temporal_min_match` to `0`.``
-
-The output from temporal matching is also a cubble, with additional column `.dist` and `.group` inherent from spatial matching and `n_match` for the number of matched temporal peaks.
-Then you can use this output to plot the location of match or to look at the series:
-
-```{r matching, out.width="100%", out.height="30%", fig.cap="asdfasd"}
-knitr::include_graphics("figures/matching.png")
-```
-
-## ERA5: climate reanalysis data
-
-\newpage
-
-## Interative graphic with cubble
-
-With a larger dataset, interactive graphics can be helpful to explore these data from multiple perspective.
-Two examples are presented here to show how interactivity can be used to explore the climate data across Australia.
-Again, `weatherdata::climate_full` will be used and we hope to include more than one time variable here.
-The difference between the maximum and minimum temperature of the day, the diurnal range, can be directly calculated from the data and it can have large impact on the wine industry (lol this is going far).
-This would be useful to compare the temperature difference across stations and to compare how this temperature difference differs throughout the year.
-
-Temperature difference is a time-wise variable and should be computed in the long form cubble. The variables are averaged over the 2016 to 2020 period by month to stabilise the result.
-
-```{r eval = FALSE, echo = TRUE}
-res <- weatherdata::climate_full %>%
-  # filter(!id %in% c("ASN00067033", "ASN00072091", 
-  #                   "ASN00059040", "ASN00097053")) %>%
-  stretch() %>%
-  mutate(month = lubridate::month(date),
-         diff = tmax - tmin) %>%
-  group_by(month) %>%
-  summarise(tmax = mean(tmax, na.rm = TRUE),
-            diff = mean(diff, na.rm = TRUE)) %>%
-  ungroup(month)
-
-```
-
-Several variables can be useful when comparing the seasonal effect of the temperature differences, for example, the difference of average temperature difference between summer and winter can tell [...].
-This is a station-wise measure and hence computed in the nested form.
-
-```{r eval = FALSE, echo = TRUE}
-res <- res %>% 
-  tamp() %>%
-  mutate(
-    diff_winter = ts %>% filter(month %in% c(6, 7, 8)) %>%
-      pull(diff) %>% mean(na.rm = TRUE),
-    diff_summer = ts %>% filter(month %in% c(12, 1, 2)) %>%
-      pull(diff) %>% mean(na.rm = TRUE),
-    diff_year = diff_winter - diff_summer
-  )
-```
-
-The result from the above manipulation can be directly supplied to make interactive plots and Figure \ref{fig:interactive-linking} is an example made with \pkg{plotly} and \pkg{crosstalk}.
-For example, the second row results from selecting the highest mean maximum temperature in July and this corresponds to Kalumburu where there is a larger winter (June to August) temperature difference than summer (December to February).
-The third row selects one of the stations (Birdsville Airport) with a large average temperature difference in the interior desert region on the map and the linked view from the time series shows that this maximum and minimum temperature difference is constant throughout the year.
-The last row selects the lowest maximum temperature in July on the temperature band and surprisingly, the corresponding station doesn't fall into the Tasmania island but Mountain Hotham in Victoria.
-
-```{r interactive-linking, echo = FALSE, out.width="100%", out.height="25%", fig.cap = "Screenshots of the mean daily temperature difference averaged by month from 2016 to 2020 at various locations in Australia.
-Crossed linking between the map and the temperature band is implemented in both directions for exploring the temperature pattern interactively.
-The first row is the general appearance of the graph and the next three rows are examples from various selection to show different temperature patterns in Australia.
-The html version can be found at ...", fig.show='hold'}
-knitr::include_graphics(here::here("figures/linking.png"))
-knitr::include_graphics(here::here("figures/linking-north.png"))
-knitr::include_graphics(here::here("figures/linking-mid.png"))
-knitr::include_graphics(here::here("figures/linking-south.png"))
-```
-
-Cubble is also compatible with \pkg{leaflet} and with leaflet extension package \pkg{leafpop}, the temperature band can be displayed upon clicking on the station on the map.
-Figure \ref{fig:interactive-popup} shows more examples of temperature pattern that can be found in Australia.
-With the same latitude as Kalumburu but in the Pacific coastline, Cooktown airport shows a similarly high but constant temperature throughout the year.
-With a slightly higher latitude than the mountain Hotham, flat land in Melbourne airport shows a less extreme temperature band.
-Lastly, the Cape Grim Baps, in the north-west tip of the Tasmania island shows a uniformly low and thin temperature band.
-
-```{r interactive-popup, echo = FALSE, out.width="50%", out.height="30%", fig.cap = "Screenshots of the diurnal temperature range averaged by month from 2016 to 2020 at various locations in Australia.
-The leaflet popup allows users to click on a single station for its temperature band as a small plot in the popup window.
-The html version can be found at ...", fig.show='hold'}
-knitr::include_graphics(here::here("figures/popup-north.png"))
-knitr::include_graphics(here::here("figures/popup-vic.png"))
-knitr::include_graphics(here::here("figures/popup-tas.png"))
-```
-
-# Conclusion
-
-# Appendix
-
-```{r illu-interactive-2, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "demon interactivity"}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.005.png"))
-```
-
-\newpage
-
-<!-- # Old stuff -->
-
-<!-- In the temporal aspect, the `tsibble` [@tsibbles] structure and its tidyverts ecosystem have provided a [...
-] workflow to work with temporal data.
-In a tsibble structure, temporal data is characterised by `index` and `key` where `index` is the temporal identifier and `key` is the identifier for multiple series, which could be used as a spatio identifier.
-However, a tsibble object, by construction, always requires the `index` in its structure.
-This makes it less appealing for spatio-temporal data since the output of calculated spatio-specific variables (i.e.
-features of each series) don't have the time dimension.
-Analysts will either need to have an additional step to join this output to the original tsibble or operate with variables stored in two separate objects.
-In addition, the long form structure of a tsibble object means spatio variables (i.e.
-longitude, latitude, and features of each series if joined back to the tsibble) of each spatio identifier will be repetitively recorded at each timestamp.
-This repetition is unnecessary and would inflate the object size for long series.
--->
-
-<!-- # Create a cubble -->
-
-<!-- The creation of a cubble requires the site identifier (`key`), as well as the spatial (`coords`) and temporal (`index`) identifier.
-`climate_flat` is already a tibble and it uses `id` to identify each station, `date` as the time identifier, and `c(long, lat)` as the spatial identifier.
-To create a cubble for this data, use: -->
-
-<!-- ```{r} -->
-<!-- climate_flat %>% as_cubble(key = id, index = date, coords = c(long, lat)) -->
-<!-- ``` -->
-
-<!-- Most of the time, spatio-temporal data doesn't come into this form and analysts need to query the climate variables based on station metadata.
-\textcolor{red}{This is also a problem illustrated in Section 3.5 in @tidydata.
-Here we provide a structured way to query this data based on the row-wise operator and nested list.} For this type of task, one can structure a metadata into a tibble and use row-wise operator to query the climate variables into a nested list.
-As an example here we demonstrate the workflow to find the 5 closest stations to Melbourne.
-We first create a station data frame with the 5 target stations.
--->
-
-<!-- ```{r echo = FALSE} -->
-<!-- latlon_ref <- tibble::tibble(id = "melbourne", lat = -37.8136, long = 144.9631) -->
-
-<!-- cand <- aus_climate %>% -->
-<!--   as_tibble() %>% -->
-<!--   select(id, lat, long, elev, name, wmo_id) -->
-
-<!-- calc_dist <- function(df_ref, df, n = 5) { -->
-<!--   long_ref <- df_ref$long -->
-<!--   lat_ref <- df_ref$lat -->
-
-<!--   df %>% -->
-<!--     mutate( -->
-<!--       long_ref = long_ref, -->
-<!--       lat_ref = lat_ref, -->
-<!--       dist = rnoaa::meteo_spherical_distance(lat, long, lat_ref, long_ref) -->
-<!--     ) %>% -->
-<!--     slice_min(dist, n = n) %>% -->
-<!--     mutate(city = df_ref$id) %>% -->
-<!--     select(-long_ref, -lat_ref) -->
-<!-- } -->
-
-<!-- (stations <- do.call("rbind", purrr::map( -->
-<!--   1:nrow(latlon_ref), -->
-<!--   ~ calc_dist(latlon_ref[.x, ], cand) -->
-<!-- ))) -->
-<!-- ``` -->
-
-<!-- We can query the climate information into a nested list named `ts` for each station with the `rowwise()` operator. To create a cubble, supply the same identifiers as with the first example. -->
-
-<!-- ```{r eval = FALSE} -->
-<!-- sydmel_climate <- stations %>% -->
-<!--   rowwise() %>% -->
-<!--   mutate(ts = list(meteo_pull_monitors(id, -->
-<!--     date_min = "2020-01-01", -->
-<!--     date_max = "2020-12-31", -->
-<!--     var = c("PRCP", "TMAX", "TMIN") -->
-<!--   ) %>% -->
-<!--     select(-id))) %>% -->
-<!--   as_cubble(key = id, index = date, coords = c(long, lat)) -->
-<!-- ``` -->
-
-<!-- ```{r echo = FALSE} -->
-<!-- (sydmel_climate <- stations %>% -->
-<!--   left_join(aus_climate) %>% -->
-<!--   as_cubble(key = id, index = date, coords = c(long, lat))) -->
-<!-- ``` -->
-
-<!-- Below are the how the nested and long form look like for Australia climate data, which records daily precipitation, maximum and minimum temperature for 55 stations across Australia from 2015- 2020.
-Notice that each station forms a group in both forms and specifically, the nested and long form have a underlying `rowwise_df` and `grouped_df` respectively.
--->
-
-<!-- With a cubic framework on mind, different types of manipulation with cubble can be thought of as slicing the cube in various way.
-The table below shows how some `dplyr` verbs are mapped into the operation in a cubble.
-With the existing grouping on the station, additional grouping can be added with `group_by` and removed with `ungrouped`.
-[talk about why it is useful] -->
-
-<!-- \newpage -->
-
-<!-- ## Cubble operations -->
-
-<!-- ```{r cubble-operations, echo = FALSE, fig.cap="Cubble operations", out.height="50%", out.width="90%", fig.align='center'} -->
-<!-- knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.004.png")) -->
-<!-- ``` -->

### 2022-01-07 2-the-cubble-pkg.Rmd
play aroudnd with child rmd to organising sections

--- 2021-12-20 main.Rmd
@@
 ---
-documentclass: jss
-author:
-  - name: H. Sherry Zhang
-    # department: Department of Econometrics and Business Statistics
-    affiliation: Monash University
-    #location: Melbourne, Australia
-    email: \email{huize.zhang@monash.edu}
-  - name: Dianne Cook
-    # department: Department of Econometrics and Business Statistics
-    affiliation: 'Monash University \AND'
-    # location: Melbourne, Australia
-    # email:  dicook@monash.edu
-  - name: Ursula Laa
-    # department: Institute of Statistics
-    affiliation: 'University of Natural Resources and Life Sciences \AND'
-    # location: Vienna, Austria
-    # email:  ursula.laa@boku.ac.at
-  - name: Nicolas Langrené
-    # department: 34 Village Street, Docklands VIC 3008 Australia
-    affiliation: 'CSIRO Data61 \AND'
-    # location: Melbourne, Australia
-    # email: nicolas.langrene@csiro.au
-  - name: Patricia Menéndez
-    # department: Department of Econometrics and Business Statistics
-    affiliation: 'Monash University \AND'
-    # location: Melbourne, Australia
-    # email:  patricia.menendez@monash.edu
-title:
-  formatted: "\\pkg{cubble}: An R Package for Structuring Spatio-temporal Data"
-  # If you use tex in the formatted title, alsoge supply version without
-  plain:     "cubble: An R Package for Structuring Spatio-temporal Data"
-  # For running headers, if needed
-  # short:     "\\pkg{foo}: A Capitalized Title"
-abstract: >
-  The abstract of the article.
-keywords:
-  # at least one keyword must be supplied
-  formatted: [spatio-temporal data,  "\\proglang{R}"]
-  plain:     [spatio-temporal data, R]
-preamble: >
-  \usepackage{amsmath}
-  \usepackage{array}
-output: rticles::jss_article
-bibliography: references.bib
+title: "2-the-cubble-pkg"
+output: pdf_document
 ---
-
-```{r setup, echo = FALSE}
-knitr::opts_chunk$set(
-  echo = FALSE,
-  warning = FALSE,
-  message = FALSE,
-  cache = TRUE,
-  fig.path = "figures/",
-  fig.height = 10
-)
-options(prompt = "R> ", continue = "+ ", 
-        tibble.print_max = 5, tibble.print_min = 5)
-```
-
-```{r echo = FALSE}
-library(cubble)
-library(forcats)
-library(stringr)
-library(dplyr)
-library(lubridate)
-library(ggplot2)
-library(GGally)
-library(plotly)
-library(crosstalk)
-library(patchwork)
-```
-
-\newpage
-
-# Introduction
-
-Many data structures have been proposed for spatial (\pkg{sf} by @sf) and temporal (\pkg{tsibble} by @tsibble) data in the R community, while less has been done for spatio-temporal data.
-The lack of such tools could potentially because analysts usually treat the spatial and temporal dimension of the data separately, without realising the need to create a new data structure.
-While this approach follows the third tidy data principal [@tidydata] (*Each type of observational unit forms a table*), analysts always need to manually join results from different observational units or combining multiple tables into one for downstream analysis.
-This additional step doesn't add new operations into the data but can be error prone.
-\newline
-
-Currently, available spatio-temporal data structure in R includes: \pkg{spacetime} [@spacetime], which proposes four space-time layouts: Full grid (STF), sparse grid(STS), irregular (STI), and trajectory (STT).
-The data structure it uses is based on \pkg{sp} [@sp] and \pkg{xts} [@xts], both of which has been replaced by more recent implementations.
-\pkg{spatstat} [@spatstat] implements a `ppp` class for point pattern data; and more recent, \pkg{stars} [@stars] implements a spatio-temporal array with the dplyr's data cube structure \pkg{cubelyr} [@cubelyr] as its backend.
-While these implementations either store spatial and temporal variables all in a single table, hence duplicate the spatial variables for each temporal unit; or split them into two separate tables that has the problem of manually joining, mentioned in the previously.
-None of these packages enjoy both the benefits of being able to separate manipulation in the two dimensions while also keep the data object as a whole.
-This create a gap in the software development.
-The requirement for such a tool is important given the ubiquity of spatio-temporal vector data in the wild: the Ireland wind data from \pkg{gstat} is an classic example data that splits variables into spatial (`wind.loc`) and temporal (`wind`) dimension; Bureau of Meteorology (BoM) provides climate observations that are widely applied in agriculture and ecology study; air pollution data.
-\newline
-
-This paper describes the implementation of a new spatio-temporal data structure: \pkg{cubble}.
-\pkg{cubble} implements a relational data structure that uses two forms to manage the switch between spatial and temporal dimension.
-With this structure, users can manipulate the spatial or temporal dimension separately, while leaves the linking of two dimensions to \pkg{cubble}.
-The software is available from the Comprehensive R Archive Network (CRAN) at [CRAN link].
-\newline
-
-The rest of the paper will be divided as follows: [complete when the paper structure is more solid]
-
-<!-- Section 2 reviews the existing data structure for spatio, temporal, and spatio-temporal data.
-Section 3 presents a new data structure for spatio-temporal data: cubble.
-Then the paper introduces the workflow of data manipulation and visualisation with the cubble structure in Section 4.
-Section 5 gives some examples on how common spatial and temporal manipulations are performed with cubble and how static and interactive visualisation help to understand climate and [...] data.
-\newline -->
-
-\newpage
 
@@
 If the input data is a \pkg{tsibble} object, the long form cubble is also a \pkg{tsibble} where users can directly apply time series operations.
-
-# Advanced features/ considerations
-
-## Hierarchical structure
-
-Imposing a clustering structure can be thought of as building a hierarchical structure where stations are nested within clusters.
-This can be useful when there's intrinsic nesting structure in the data (i.e.
-country nested in the continent, county nested in the state) or there's some clustering.
-When we have access to both level, `switch_key()` is the function to re-structure the data as one cluster per row.
-Temporal observations from different stations while within the same cluster are bound in the nested column `ts`.
-
-One thing we hope to do with the cluster is to find the coordinates of the centroid.
-These are variables variant to the station but invariant to the cluster and it would be nice to have a function that structure each cluster as a row.
-`switch_key()` is the function that does this: it lets you to specify a new key, say `cluster` and nests all spatial variables variant to `cluster` into a column.
-Spatial variables are all nested inside a new column `.val`.
-Temporal observations from different stations while within the same cluster are bound in the nested column `ts`.
-
-This structure makes it easy to compute cluster level variable, for example, the convex hull and the centroid coordinate of each cluster.
-These can be amended into the nested form with function `get_centroid`.
-
-After we have got `cluster_nested`, spatial and temporal data at both levels can be easily obtained.
-Figure \ref{fig:illu-hier} illustrate the relationship between the long and nested form cubble at both site and cluster level.
-More description on this.
-Start with the original `station_nested`, `stretch()` expands the `ts` column with each station (`id`) forming a group and attach variables invariant to `id` as an attribute.
-`switch_key()` changes the `key` from `id` to `cluster` and nests all the spatial variables that variant to `cluster`.
-`stretch()` `cluster_nested` will store variables that are invariant to `cluster` as a tibble in the attribute.
-
-this fit into the remaining data pipeline.
-
-```{r illu-hier, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "Hierarchical structure"}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.003.png"))
-```
-
-## Data fusion and matching
-
-Temporal matching checks how spatially matched pairs align temporally. We use the following chart to illustrate how the temporal matching works:
-
-For each spatially matched pair, say `A` and `a`, we first find the largest `n` points in each series, colored in brown points here.
-Here we use the largest three but you can tune this number by `temporal_n_highest`.
-Then we construct the interval of the largest points from one series and see how many points, from the other series, fall into the intervals.
-The series used to construct the interval is controlled by `temporal_independent` and the window size by `temporal_window` with a default of 5.
-
-In this illustration, we construct the interval based on series `A` and two of the three peaks from `a` falls into this interval at Time 7 and 27.
-
-There's another mandatory argument that hasn't been introduced above: `temporal_var_to_match`.
-This argument controls the variable to match and it needs to appear in both the `major` and `minor` set.
-In the water level matching example, we match the variable `Water_course_level` from `river` to `prcp` from `climate`, hence need to manually rename one of them to match the other, here we rename `Water_course_level` to `prcp` in `river`:
-
-Figure \ref{fig:illu-matching}
-
-```{r illu-matching, out.width="100%", out.height="30%", fig.cap="sdfasdf"}
-knitr::include_graphics("figures/illu-matching.png")
-```
-
-## Interactive graphics
-
-<!-- Interactive graphics can listen to users' actions on the plot to provide additional information that facilitates data exploration.
-This is a useful technique for spatio-temporal data since users can zoom or pan the map to view the local and global structure of the map; use tooltips or popups to query more information about a graphic element; or highlight points to explore its linked views in other plots.
-In the R community, many implementations have been developed to connect \proglang{R} to \proglang{javascript} to create interactive graphics.
-In relation to spatio, temporal, and spatio-temporal data, the general purpose packages \pkg{plotly} [@plotly] and \pkg{leaflet} [@leaflet] realise various interactive actions through their corresponding javascript libraries.
-\pkg{crosstalk} [@crosstalk] and \pkg{tsibbletalk} [@tsibbletalk] implement brushed linking between htmlwidgets.
-\pkg{ggiraph} [@ggiraph] enables tooltip, self-linking, and customised actions specified through its own \proglang{javascript}.
--->
-
-<!-- While many graphic implementations present worked examples to illustrate the usage of the package, few documents the underlying pipeline that transforms the raw data step-by-step to the final view on the screen.
-There have been some early work in building the data pipeline for (interactive) graphics [@buja1988elements; @buja1996interactive;  @sutherland2000orca] and more recent discussions include @wickham2009plumbing, @xie2014reactive, and @cheng2016enabling.
--->
-
-The cubble structure fits in naturally with the interactive graphic pipeline discussed in the literature [@buja1988elements; @buja1996interactive;  @sutherland2000orca; @xie2014reactive; @cheng2016enabling].
-Diagram \ref{fig:illu-interactive} illustrates how linking works with the two forms in a cubble, where a time series plot is created with the long cubble and a map is created with the nested cubble.
-When a user action is captured from the map, the site will be activated in the nested cubble.
-Then, the nested cubble will communicate to the long cubble to activate all the observations with the same `id`.
-The long cubble will then highlight the  activated series in the time series plot.
-
-The linking is also available from the time series plot to the map.
-The selection on the time series is through selecting the point on the time series and once a point is selected, it will be activated in the long cubble.
-All the observations that share the same `id`,  either in the long and nested cubble, are then activated.
-This includes other points in the same time series in the long cubble and the corresponding observation of site in the nested cubble.
-These activated observations will then being reflected in the updated plots and Diagram \ref{fig:illu-interactive-2} in the Appendix illustrates this process.
-
-```{r illu-interactive, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "demon interactivity"}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.004.png"))
-```
-
-\newpage
-
-# Examples
-
-## Australia historical maximum temperature
-
-Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world and the dataset `weatherdata::historical_tmax` extracts the historical maximum temperature recorded for `r nrow(weatherdata::historical_tmax)` Australian stations.
-The data `historical_tmax` is already presented as a cubble, with `id` as the key, `date` as the index, and `c("longitude", "latitude")` as the coordinates.
-Other variables include `elevation`, `name`, `wmo_id`, `first_year`, and  `last_year` in the nested form and `tmax` in the long form.
-This example compares the maximum temperature in two periods: 1971 - 1975 and 2016 - 2020 for stations in Victoria and New South Wales.
-
-Stations in the two states can be subsetted on the station number: Australia GHCN station number starts with `ASN00` and followed by the [Bureau of Meteorology (BOM) station number](http://www.bom.gov.au/climate/cdo/about/site-num.shtml), where the 2nd and 3rd digit (7th and 8th in the GHCN number) denote the state a station belongs to.
-New South Wales stations start from 46 to 75 and Victoria stations then follow from 76 to 90.
-Extracting Victoria and New South Wales stations is a filter operation in the spatial dimension and hence is operated in the nested form:
-
-```{r eval = FALSE, echo = TRUE}
-tmax <- weatherdata::historical_tmax %>%
-  filter(between(stringr::str_sub(id, 7, 8), 46, 90))
-```
-
-The five year window is chosen to remove the effect of a particular year and the historical period of 1971 - 1975 is used since all the stations have records from 1970.
-This following chunk filters on records in the two study periods and summarises the maximum temperature into monthly measure.
-`stretch()` is used to convert the nested form cubble into the long form for these operations:
-
-```{r eval=  FALSE, echo = TRUE}
-tmax <- tmax %>% 
-  stretch() %>%
-  filter(lubridate::year(date) %in% c(1971:1975, 2016:2020)) %>%
-  mutate(month = lubridate::month(date), 
-         group = as.factor(ifelse(lubridate::year(date) > 2015, 
-                                  "2016 ~ 2020", "1971 ~ 1975"))) %>%
-  group_by(month, group) %>%
-  summarise(tmax = mean(tmax, na.rm = TRUE))
-```
-
-A data quality issue with GHCN data is that while the first and last year of each series is provided, years missing in this period is not reported.
-There are a few stations which don't have records during 1971 - 1975 and these stations are filtered out by examining whether the summarised `tmax` has a total of 24 months.
-This is again a station-wise operation and is operated in the nested form, which is switched to from the long form by  `tamp()`:
-
-```{r eval = FALSE, echo = TRUE}
-tmax <- tmax %>% 
-  tamp() %>%
-  filter(nrow(ts) == 24) 
-```
-
-Lastly, to create a glyph map, both the major (`longitude`, `latitude`) and minor (`month`, `tmax`) coordinates need to be in the same table.
-Spatial variables can be moved to the long form with `migrate()`:
-
-```{r eval = FALSE, echo = TRUE}  
-tmax <- tmax %>%   
-  stretch() %>%
-  migrate(latitude, longitude)
-```
-
-Figure \ref{fig:basic-manip} shows the glyph map made with the data after the wangling above.
-One issue with this map is that the similar pattern shown from a few nearby stations around Sydney (151E, 34S) and New castle (152E, 33S) can be distracting.
-Aggregation is a useful technique to observe the general pattern of a collection of series and will be the topic of the next example.
-
-```{r basic-manip, out.width="100%", out.height="70%", fig.cap="Glyph map of the mean maximum temperature by month for Victoria and New South Wales weather stations.
-On the top left corner is a detailed legend for station Cobar highlighted in the black box.
-Compared to 1971 - 1975,  the period 2016 - 2020 sees an increase of mean maximum temperature in spring and summer in Australia (end and beginning of the year).
-A larger increase during the first three months of the year is observed at stations with a lower latitude."}
-knitr::include_graphics("figures/basic-manip.png")
-```
-
-## Australia precipitation pattern in 2020
-
-In Figure \ref{fig:basic-agg}.
-Aggregating them into averages or other statistics would give a better picture.
-This can be done with cubble through a hierarchical structure, where the cluster indicator is one level above the individual station.
-
-The `weatherdata::climate_full` data has daily climate data of `r nrow(weatherdata::climate_full)` Australia stations from 2016 to 2020.
-
-calcluate distance matrix
-As an example to illustrate here, a kmean clustering algorithm based on the distance matrix is used and the number of centres is set to 20.
-More complex algorithms can also be used for more complex problem, as long as a mapping from each station id to the cluster id can be constructed.
-
-<!-- Now we can obtain the aggregated series as described in the workflow diagram above and construct the glyph map with `GGally::glyphs()`: -->
-
-<!-- We can also look at the precipitation of each individual station within the same cluster: -->
-
-<!-- Lastly, there is one series on Tasmania island standing out from others, lets look at where it is: -->
-
-```{r basic-agg, out.width="100%", out.height="90%", fig.cap="sdfasdf"}
-knitr::include_graphics("figures/basic-agg.png")
-```
-
-## River level data in Victria water gauges
-
-The water level data comes from [Bureau of Meteorology](http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full) and has a copy in `weatherdata`.
-Here we extract the water course level and add a column annotate this data of type `river`.
-For the rainfall data, we will still use the `weatherdata::climate_full`, filtering for Victorian stations in 2020 should be pretty familiar by now.
-Again, we first look at where these stations are on the map first:
-
-Now we use `match_sites()` to first pair the weather stations with the river gauges spatially and then apply the temporal matching on `prcp`.
-We will construct the interval based on peaks in `climate` since we would expect a lag effect for precipitation to flow into the river and cause a raise in river level, hence `temporal_independent = climate`.
-We select the 30 highest peak from the series to construct the match by setting `temporal_n_highest = 30`.
-This is a tuning parameter and you can start with 10% of the points of one series (here we have daily data for a year, 10% is roughly 30 points).
-`temporal_min_match` filters out pairs don't have enough match and to return all the pairs, set `temporal_min_match` to `0`.``
-
-The output from temporal matching is also a cubble, with additional column `.dist` and `.group` inherent from spatial matching and `n_match` for the number of matched temporal peaks.
-Then you can use this output to plot the location of match or to look at the series:
-
-```{r matching, out.width="100%", out.height="30%", fig.cap="asdfasd"}
-knitr::include_graphics("figures/matching.png")
-```
-
-## ERA5: climate reanalysis data
-
-\newpage
-
-## Interative graphic with cubble
-
-With a larger dataset, interactive graphics can be helpful to explore these data from multiple perspective.
-Two examples are presented here to show how interactivity can be used to explore the climate data across Australia.
-Again, `weatherdata::climate_full` will be used and we hope to include more than one time variable here.
-The difference between the maximum and minimum temperature of the day, the diurnal range, can be directly calculated from the data and it can have large impact on the wine industry (lol this is going far).
-This would be useful to compare the temperature difference across stations and to compare how this temperature difference differs throughout the year.
-
-Temperature difference is a time-wise variable and should be computed in the long form cubble. The variables are averaged over the 2016 to 2020 period by month to stabilise the result.
-
-```{r eval = FALSE, echo = TRUE}
-res <- weatherdata::climate_full %>%
-  # filter(!id %in% c("ASN00067033", "ASN00072091", 
-  #                   "ASN00059040", "ASN00097053")) %>%
-  stretch() %>%
-  mutate(month = lubridate::month(date),
-         diff = tmax - tmin) %>%
-  group_by(month) %>%
-  summarise(tmax = mean(tmax, na.rm = TRUE),
-            diff = mean(diff, na.rm = TRUE)) %>%
-  ungroup(month)
-
-```
-
-Several variables can be useful when comparing the seasonal effect of the temperature differences, for example, the difference of average temperature difference between summer and winter can tell [...].
-This is a station-wise measure and hence computed in the nested form.
-
-```{r eval = FALSE, echo = TRUE}
-res <- res %>% 
-  tamp() %>%
-  mutate(
-    diff_winter = ts %>% filter(month %in% c(6, 7, 8)) %>%
-      pull(diff) %>% mean(na.rm = TRUE),
-    diff_summer = ts %>% filter(month %in% c(12, 1, 2)) %>%
-      pull(diff) %>% mean(na.rm = TRUE),
-    diff_year = diff_winter - diff_summer
-  )
-```
-
-The result from the above manipulation can be directly supplied to make interactive plots and Figure \ref{fig:interactive-linking} is an example made with \pkg{plotly} and \pkg{crosstalk}.
-For example, the second row results from selecting the highest mean maximum temperature in July and this corresponds to Kalumburu where there is a larger winter (June to August) temperature difference than summer (December to February).
-The third row selects one of the stations (Birdsville Airport) with a large average temperature difference in the interior desert region on the map and the linked view from the time series shows that this maximum and minimum temperature difference is constant throughout the year.
-The last row selects the lowest maximum temperature in July on the temperature band and surprisingly, the corresponding station doesn't fall into the Tasmania island but Mountain Hotham in Victoria.
-
-```{r interactive-linking, echo = FALSE, out.width="100%", out.height="25%", fig.cap = "Screenshots of the mean daily temperature difference averaged by month from 2016 to 2020 at various locations in Australia.
-Crossed linking between the map and the temperature band is implemented in both directions for exploring the temperature pattern interactively.
-The first row is the general appearance of the graph and the next three rows are examples from various selection to show different temperature patterns in Australia.
-The html version can be found at ...", fig.show='hold'}
-knitr::include_graphics(here::here("figures/linking.png"))
-knitr::include_graphics(here::here("figures/linking-north.png"))
-knitr::include_graphics(here::here("figures/linking-mid.png"))
-knitr::include_graphics(here::here("figures/linking-south.png"))
-```
-
-Cubble is also compatible with \pkg{leaflet} and with leaflet extension package \pkg{leafpop}, the temperature band can be displayed upon clicking on the station on the map.
-Figure \ref{fig:interactive-popup} shows more examples of temperature pattern that can be found in Australia.
-With the same latitude as Kalumburu but in the Pacific coastline, Cooktown airport shows a similarly high but constant temperature throughout the year.
-With a slightly higher latitude than the mountain Hotham, flat land in Melbourne airport shows a less extreme temperature band.
-Lastly, the Cape Grim Baps, in the north-west tip of the Tasmania island shows a uniformly low and thin temperature band.
-
-```{r interactive-popup, echo = FALSE, out.width="50%", out.height="30%", fig.cap = "Screenshots of the diurnal temperature range averaged by month from 2016 to 2020 at various locations in Australia.
-The leaflet popup allows users to click on a single station for its temperature band as a small plot in the popup window.
-The html version can be found at ...", fig.show='hold'}
-knitr::include_graphics(here::here("figures/popup-north.png"))
-knitr::include_graphics(here::here("figures/popup-vic.png"))
-knitr::include_graphics(here::here("figures/popup-tas.png"))
-```
-
-# Conclusion
-
-# Appendix
-
-```{r illu-interactive-2, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "demon interactivity"}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.005.png"))
-```
-
-\newpage
-
-<!-- # Old stuff -->
-
-<!-- In the temporal aspect, the `tsibble` [@tsibbles] structure and its tidyverts ecosystem have provided a [...
-] workflow to work with temporal data.
-In a tsibble structure, temporal data is characterised by `index` and `key` where `index` is the temporal identifier and `key` is the identifier for multiple series, which could be used as a spatio identifier.
-However, a tsibble object, by construction, always requires the `index` in its structure.
-This makes it less appealing for spatio-temporal data since the output of calculated spatio-specific variables (i.e.
-features of each series) don't have the time dimension.
-Analysts will either need to have an additional step to join this output to the original tsibble or operate with variables stored in two separate objects.
-In addition, the long form structure of a tsibble object means spatio variables (i.e.
-longitude, latitude, and features of each series if joined back to the tsibble) of each spatio identifier will be repetitively recorded at each timestamp.
-This repetition is unnecessary and would inflate the object size for long series.
--->
-
-<!-- # Create a cubble -->
-
-<!-- The creation of a cubble requires the site identifier (`key`), as well as the spatial (`coords`) and temporal (`index`) identifier.
-`climate_flat` is already a tibble and it uses `id` to identify each station, `date` as the time identifier, and `c(long, lat)` as the spatial identifier.
-To create a cubble for this data, use: -->
-
-<!-- ```{r} -->
-<!-- climate_flat %>% as_cubble(key = id, index = date, coords = c(long, lat)) -->
-<!-- ``` -->
-
-<!-- Most of the time, spatio-temporal data doesn't come into this form and analysts need to query the climate variables based on station metadata.
-\textcolor{red}{This is also a problem illustrated in Section 3.5 in @tidydata.
-Here we provide a structured way to query this data based on the row-wise operator and nested list.} For this type of task, one can structure a metadata into a tibble and use row-wise operator to query the climate variables into a nested list.
-As an example here we demonstrate the workflow to find the 5 closest stations to Melbourne.
-We first create a station data frame with the 5 target stations.
--->
-
-<!-- ```{r echo = FALSE} -->
-<!-- latlon_ref <- tibble::tibble(id = "melbourne", lat = -37.8136, long = 144.9631) -->
-
-<!-- cand <- aus_climate %>% -->
-<!--   as_tibble() %>% -->
-<!--   select(id, lat, long, elev, name, wmo_id) -->
-
-<!-- calc_dist <- function(df_ref, df, n = 5) { -->
-<!--   long_ref <- df_ref$long -->
-<!--   lat_ref <- df_ref$lat -->
-
-<!--   df %>% -->
-<!--     mutate( -->
-<!--       long_ref = long_ref, -->
-<!--       lat_ref = lat_ref, -->
-<!--       dist = rnoaa::meteo_spherical_distance(lat, long, lat_ref, long_ref) -->
-<!--     ) %>% -->
-<!--     slice_min(dist, n = n) %>% -->
-<!--     mutate(city = df_ref$id) %>% -->
-<!--     select(-long_ref, -lat_ref) -->
-<!-- } -->
-
-<!-- (stations <- do.call("rbind", purrr::map( -->
-<!--   1:nrow(latlon_ref), -->
-<!--   ~ calc_dist(latlon_ref[.x, ], cand) -->
-<!-- ))) -->
-<!-- ``` -->
-
-<!-- We can query the climate information into a nested list named `ts` for each station with the `rowwise()` operator. To create a cubble, supply the same identifiers as with the first example. -->
-
-<!-- ```{r eval = FALSE} -->
-<!-- sydmel_climate <- stations %>% -->
-<!--   rowwise() %>% -->
-<!--   mutate(ts = list(meteo_pull_monitors(id, -->
-<!--     date_min = "2020-01-01", -->
-<!--     date_max = "2020-12-31", -->
-<!--     var = c("PRCP", "TMAX", "TMIN") -->
-<!--   ) %>% -->
-<!--     select(-id))) %>% -->
-<!--   as_cubble(key = id, index = date, coords = c(long, lat)) -->
-<!-- ``` -->
-
-<!-- ```{r echo = FALSE} -->
-<!-- (sydmel_climate <- stations %>% -->
-<!--   left_join(aus_climate) %>% -->
-<!--   as_cubble(key = id, index = date, coords = c(long, lat))) -->
-<!-- ``` -->
-
-<!-- Below are the how the nested and long form look like for Australia climate data, which records daily precipitation, maximum and minimum temperature for 55 stations across Australia from 2015- 2020.
-Notice that each station forms a group in both forms and specifically, the nested and long form have a underlying `rowwise_df` and `grouped_df` respectively.
--->
-
-<!-- With a cubic framework on mind, different types of manipulation with cubble can be thought of as slicing the cube in various way.
-The table below shows how some `dplyr` verbs are mapped into the operation in a cubble.
-With the existing grouping on the station, additional grouping can be added with `group_by` and removed with `ungrouped`.
-[talk about why it is useful] -->
-
-<!-- \newpage -->
-
-<!-- ## Cubble operations -->
-
-<!-- ```{r cubble-operations, echo = FALSE, fig.cap="Cubble operations", out.height="50%", out.width="90%", fig.align='center'} -->
-<!-- knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.004.png")) -->
-<!-- ``` -->

### 2022-01-07 3-advanced.Rmd
play aroudnd with child rmd to organising sections

--- 2021-12-20 main.Rmd
@@
 ---
-documentclass: jss
-author:
-  - name: H. Sherry Zhang
-    # department: Department of Econometrics and Business Statistics
-    affiliation: Monash University
-    #location: Melbourne, Australia
-    email: \email{huize.zhang@monash.edu}
-  - name: Dianne Cook
-    # department: Department of Econometrics and Business Statistics
-    affiliation: 'Monash University \AND'
-    # location: Melbourne, Australia
-    # email:  dicook@monash.edu
-  - name: Ursula Laa
-    # department: Institute of Statistics
-    affiliation: 'University of Natural Resources and Life Sciences \AND'
-    # location: Vienna, Austria
-    # email:  ursula.laa@boku.ac.at
-  - name: Nicolas Langrené
-    # department: 34 Village Street, Docklands VIC 3008 Australia
-    affiliation: 'CSIRO Data61 \AND'
-    # location: Melbourne, Australia
-    # email: nicolas.langrene@csiro.au
-  - name: Patricia Menéndez
-    # department: Department of Econometrics and Business Statistics
-    affiliation: 'Monash University \AND'
-    # location: Melbourne, Australia
-    # email:  patricia.menendez@monash.edu
-title:
-  formatted: "\\pkg{cubble}: An R Package for Structuring Spatio-temporal Data"
-  # If you use tex in the formatted title, alsoge supply version without
-  plain:     "cubble: An R Package for Structuring Spatio-temporal Data"
-  # For running headers, if needed
-  # short:     "\\pkg{foo}: A Capitalized Title"
-abstract: >
-  The abstract of the article.
-keywords:
-  # at least one keyword must be supplied
-  formatted: [spatio-temporal data,  "\\proglang{R}"]
-  plain:     [spatio-temporal data, R]
-preamble: >
-  \usepackage{amsmath}
-  \usepackage{array}
-output: rticles::jss_article
-bibliography: references.bib
+title: "3-advanced"
+output: pdf_document
 ---
-
-```{r setup, echo = FALSE}
-knitr::opts_chunk$set(
-  echo = FALSE,
-  warning = FALSE,
-  message = FALSE,
-  cache = TRUE,
-  fig.path = "figures/",
-  fig.height = 10
-)
-options(prompt = "R> ", continue = "+ ", 
-        tibble.print_max = 5, tibble.print_min = 5)
-```
-
-```{r echo = FALSE}
-library(cubble)
-library(forcats)
-library(stringr)
-library(dplyr)
-library(lubridate)
-library(ggplot2)
-library(GGally)
-library(plotly)
-library(crosstalk)
-library(patchwork)
-```
-
-\newpage
-
-# Introduction
-
-Many data structures have been proposed for spatial (\pkg{sf} by @sf) and temporal (\pkg{tsibble} by @tsibble) data in the R community, while less has been done for spatio-temporal data.
-The lack of such tools could potentially because analysts usually treat the spatial and temporal dimension of the data separately, without realising the need to create a new data structure.
-While this approach follows the third tidy data principal [@tidydata] (*Each type of observational unit forms a table*), analysts always need to manually join results from different observational units or combining multiple tables into one for downstream analysis.
-This additional step doesn't add new operations into the data but can be error prone.
-\newline
-
-Currently, available spatio-temporal data structure in R includes: \pkg{spacetime} [@spacetime], which proposes four space-time layouts: Full grid (STF), sparse grid(STS), irregular (STI), and trajectory (STT).
-The data structure it uses is based on \pkg{sp} [@sp] and \pkg{xts} [@xts], both of which has been replaced by more recent implementations.
-\pkg{spatstat} [@spatstat] implements a `ppp` class for point pattern data; and more recent, \pkg{stars} [@stars] implements a spatio-temporal array with the dplyr's data cube structure \pkg{cubelyr} [@cubelyr] as its backend.
-While these implementations either store spatial and temporal variables all in a single table, hence duplicate the spatial variables for each temporal unit; or split them into two separate tables that has the problem of manually joining, mentioned in the previously.
-None of these packages enjoy both the benefits of being able to separate manipulation in the two dimensions while also keep the data object as a whole.
-This create a gap in the software development.
-The requirement for such a tool is important given the ubiquity of spatio-temporal vector data in the wild: the Ireland wind data from \pkg{gstat} is an classic example data that splits variables into spatial (`wind.loc`) and temporal (`wind`) dimension; Bureau of Meteorology (BoM) provides climate observations that are widely applied in agriculture and ecology study; air pollution data.
-\newline
-
-This paper describes the implementation of a new spatio-temporal data structure: \pkg{cubble}.
-\pkg{cubble} implements a relational data structure that uses two forms to manage the switch between spatial and temporal dimension.
-With this structure, users can manipulate the spatial or temporal dimension separately, while leaves the linking of two dimensions to \pkg{cubble}.
-The software is available from the Comprehensive R Archive Network (CRAN) at [CRAN link].
-\newline
-
-The rest of the paper will be divided as follows: [complete when the paper structure is more solid]
-
-<!-- Section 2 reviews the existing data structure for spatio, temporal, and spatio-temporal data.
-Section 3 presents a new data structure for spatio-temporal data: cubble.
-Then the paper introduces the workflow of data manipulation and visualisation with the cubble structure in Section 4.
-Section 5 gives some examples on how common spatial and temporal manipulations are performed with cubble and how static and interactive visualisation help to understand climate and [...] data.
-\newline -->
-
-\newpage
-
-# The cubble package
-
-Spatio-temporal data usually come in various forms and Figure \ref{fig:illu-input} shows four examples of this.
-No matter which form the data is in, these formats share some common components that characterise spatio-temporal data.
-A spatial identifier (`id` in the diagram) is the unique identifier of each site.
-The temporal identifier (`t` in the diagram) prescribes the time stamp each site is recorded.
-Coordinates, comprising of latitude and longitude (`lon` and `lat` in the diagram), locates each site on the map.
-These identifiers will be the building blocks for the data structure introduced below.
-Other variables in the data can be categorised into two groups: spatial variables that are invariant at each time stamp for every site, i.e.
-the name or code of the weather station and temporal variables that varies with time.
-
-```{r illu-input, echo = FALSE, out.width = "100%", fig.cap="Illustration of incoming data formats for spatio-temporal data.
-(1) Data comes in as a single table; (2) Separate tables for spatial and temporal variables; (3) A single table with all the parameters used to query the database and a separate table for queried data; and (4) Cubical data in array or NetCDF format.", fig.align='center', out.height = "15%"}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.001.png"))
-```
-
-In a cubble, there are two forms: nested form and long form, and Figure \ref{fig:illu-cubble} sketches the two forms along with the associated attributes.
-The decision on which form to use is output-oriented, meaning analysts need to first think about whether the output of a particular operation is identified only by the spatial identifier, or a combination of spatial and temporal identifier.
-The nested cubble is suitable for working with operations that are only identified by site and this type of operation can be a pure manipulation of spatial variables, or a summary of temporal variables by site (i.e.
-the output of counting the number of raining day is only identified by sites and hence should be performed with the nested form).
-Underneath the nested form, a cubble is built from a row-wise dataframe (`rowwise_df`) where each site forms a separate group.
-This structure simplifies the calculation that involves temporal variables by avoiding the use of `map` syntax when working with list-column.
-
-For those operations whose output involves both a spatial and temporal dimension, long form should be used.
-The long form is identified by both the spatial and temporal identifier and adopts a grouped dataframe (`grouped_df`) to forms each site as a group.
-Spatial variables are stored separately in a \pkg{tibble} as an special attribute of the long cubble.
-This design avoids repeating the spatial variables at each time stamp while not dropping information from spatial variables.
-
-```{r illu-cubble, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "Illustration of nested and long cubble."}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.002.png"))
-```
-
-## Create a cubble in the nested form
-
-To use functionalities from cubble, data analysts first need to create a cubble.
-`as_cubble` create a \pkg{cubble} by supplying the three key components: `key` as the spatial identifier; `index` as the temporal identifier; and a vector of `coords` in the order of longitude first and then latitude.
-The naming of `key` and `index` follows the convention in the \pkg{tsibble} package.
-The cubble created by default is in the nested form.
-Below is an example of creating a cubble: \newline
-
-```{r}
-(cubble_nested <- cubble::climate_flat %>%
-  as_cubble(key = id, index = date, coords = c("long", "lat")))
-```
-
-There are a few information in the \pkg{cubble} header: the name of the `key` variable, bbox, and also the name of variable nested in the `ts` column.
-In this example, each site is identifier is `id` and the number in the bracket means there are 5 unique `id` in this dataset.
-The bbox in the second row gives the range of the coordinates.
-The temporal variables are all nested in the `ts` column, but it could be useful to know the name these variables.
-The third row in the cubble header shows these names and in this example this includes: precipitation, `prcp`, maximum temperature, `tmax`, and minimum temperature, `tmin`.
-
-## Stretch a nested cubble into the long form
-
-The long cubble is suitable to manipulate the time dimension of the data.
-The function `stretch()` switches the nested cubble into the long cubble by first extracts all the spatial variables into a separate tibble and store in the `spatial` attribute and then unnests the `ts` column:
-
-```{r}
-(cubble_long <- cubble_nested %>% stretch(ts))
-```
-
-Notice here that the third line in the header now shows the name of spatial variables rather than the temporal variables.
-
-## Tamp a long cubble back to the nested form
-
-Manipulation on the spatial and temporal dimension can be an iterative process.
-Many times, analysts will need to go back and forth between the nested and long cubble.
-The `stretch()` function introduced in the previous section switches a nested cubble into a long cubble and function `tamp()` is its inverse function to switch a long cubble back to the nested cubble:
-
-```{r}
-(cubble_back <- cubble_long %>% tamp())
-```
-
-## Migrate spatial variables to a long cubble
-
-As a final data output for modelling or visualisation, spatio-temporal data is usually expected to be in a single table.
-Function `migrate()` moves the spatial variables from the `spatial` attribute into the long cubble:
-
-```{r}
-(cubble_long %>% migrate(long, lat))
-```
-
-In this workflow described above, data objects come into cubble in the nested form, then various operations on the spatial and temporal dimension can go back and forth between the nested and long form, and finally, the data will come out of cubble in the long form for further modelling or visualisation.
-
-Building from an underlying `tbl_df` structure, it is natural to implement methods available in `dplyr` to `cubble`. Supported methods in the `cubble` with `dplyr` generics includes:
-
-```{=tex}
-\begin{center}
-\begin{tabular}{ | m{5em} | m{15cm}| } 
-basics & \textbf{mutate}, \textbf{filter}, \textbf{summarise}, \textbf{select}, \textbf{arrange}, \textbf{rename}, \textbf{left\_join} \\
-grouping &  \textbf{group\_by}, \textbf{ungroup}\\
-slice family & \textbf{slice\_head}, \textbf{slice\_tail}, \textbf{slice\_sample}, \textbf{slice\_min} and \textbf{slice\_max} \\
-\end{tabular}
-\end{center}
-```
-\pkg{cubble} is also compatible with \pkg{tsibble} in the sense that the original list-column can be a `tbl_ts` object.
-Duplicates and gaps should be first checked before structuring the data into a cubble.
-If the input data is a \pkg{tsibble} object, the long form cubble is also a \pkg{tsibble} where users can directly apply time series operations.
 
@@
 ```
-
-\newpage
-
-# Examples
-
-## Australia historical maximum temperature
-
-Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world and the dataset `weatherdata::historical_tmax` extracts the historical maximum temperature recorded for `r nrow(weatherdata::historical_tmax)` Australian stations.
-The data `historical_tmax` is already presented as a cubble, with `id` as the key, `date` as the index, and `c("longitude", "latitude")` as the coordinates.
-Other variables include `elevation`, `name`, `wmo_id`, `first_year`, and  `last_year` in the nested form and `tmax` in the long form.
-This example compares the maximum temperature in two periods: 1971 - 1975 and 2016 - 2020 for stations in Victoria and New South Wales.
-
-Stations in the two states can be subsetted on the station number: Australia GHCN station number starts with `ASN00` and followed by the [Bureau of Meteorology (BOM) station number](http://www.bom.gov.au/climate/cdo/about/site-num.shtml), where the 2nd and 3rd digit (7th and 8th in the GHCN number) denote the state a station belongs to.
-New South Wales stations start from 46 to 75 and Victoria stations then follow from 76 to 90.
-Extracting Victoria and New South Wales stations is a filter operation in the spatial dimension and hence is operated in the nested form:
-
-```{r eval = FALSE, echo = TRUE}
-tmax <- weatherdata::historical_tmax %>%
-  filter(between(stringr::str_sub(id, 7, 8), 46, 90))
-```
-
-The five year window is chosen to remove the effect of a particular year and the historical period of 1971 - 1975 is used since all the stations have records from 1970.
-This following chunk filters on records in the two study periods and summarises the maximum temperature into monthly measure.
-`stretch()` is used to convert the nested form cubble into the long form for these operations:
-
-```{r eval=  FALSE, echo = TRUE}
-tmax <- tmax %>% 
-  stretch() %>%
-  filter(lubridate::year(date) %in% c(1971:1975, 2016:2020)) %>%
-  mutate(month = lubridate::month(date), 
-         group = as.factor(ifelse(lubridate::year(date) > 2015, 
-                                  "2016 ~ 2020", "1971 ~ 1975"))) %>%
-  group_by(month, group) %>%
-  summarise(tmax = mean(tmax, na.rm = TRUE))
-```
-
-A data quality issue with GHCN data is that while the first and last year of each series is provided, years missing in this period is not reported.
-There are a few stations which don't have records during 1971 - 1975 and these stations are filtered out by examining whether the summarised `tmax` has a total of 24 months.
-This is again a station-wise operation and is operated in the nested form, which is switched to from the long form by  `tamp()`:
-
-```{r eval = FALSE, echo = TRUE}
-tmax <- tmax %>% 
-  tamp() %>%
-  filter(nrow(ts) == 24) 
-```
-
-Lastly, to create a glyph map, both the major (`longitude`, `latitude`) and minor (`month`, `tmax`) coordinates need to be in the same table.
-Spatial variables can be moved to the long form with `migrate()`:
-
-```{r eval = FALSE, echo = TRUE}  
-tmax <- tmax %>%   
-  stretch() %>%
-  migrate(latitude, longitude)
-```
-
-Figure \ref{fig:basic-manip} shows the glyph map made with the data after the wangling above.
-One issue with this map is that the similar pattern shown from a few nearby stations around Sydney (151E, 34S) and New castle (152E, 33S) can be distracting.
-Aggregation is a useful technique to observe the general pattern of a collection of series and will be the topic of the next example.
-
-```{r basic-manip, out.width="100%", out.height="70%", fig.cap="Glyph map of the mean maximum temperature by month for Victoria and New South Wales weather stations.
-On the top left corner is a detailed legend for station Cobar highlighted in the black box.
-Compared to 1971 - 1975,  the period 2016 - 2020 sees an increase of mean maximum temperature in spring and summer in Australia (end and beginning of the year).
-A larger increase during the first three months of the year is observed at stations with a lower latitude."}
-knitr::include_graphics("figures/basic-manip.png")
-```
-
-## Australia precipitation pattern in 2020
-
-In Figure \ref{fig:basic-agg}.
-Aggregating them into averages or other statistics would give a better picture.
-This can be done with cubble through a hierarchical structure, where the cluster indicator is one level above the individual station.
-
-The `weatherdata::climate_full` data has daily climate data of `r nrow(weatherdata::climate_full)` Australia stations from 2016 to 2020.
-
-calcluate distance matrix
-As an example to illustrate here, a kmean clustering algorithm based on the distance matrix is used and the number of centres is set to 20.
-More complex algorithms can also be used for more complex problem, as long as a mapping from each station id to the cluster id can be constructed.
-
-<!-- Now we can obtain the aggregated series as described in the workflow diagram above and construct the glyph map with `GGally::glyphs()`: -->
-
-<!-- We can also look at the precipitation of each individual station within the same cluster: -->
-
-<!-- Lastly, there is one series on Tasmania island standing out from others, lets look at where it is: -->
-
-```{r basic-agg, out.width="100%", out.height="90%", fig.cap="sdfasdf"}
-knitr::include_graphics("figures/basic-agg.png")
-```
-
-## River level data in Victria water gauges
-
-The water level data comes from [Bureau of Meteorology](http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full) and has a copy in `weatherdata`.
-Here we extract the water course level and add a column annotate this data of type `river`.
-For the rainfall data, we will still use the `weatherdata::climate_full`, filtering for Victorian stations in 2020 should be pretty familiar by now.
-Again, we first look at where these stations are on the map first:
-
-Now we use `match_sites()` to first pair the weather stations with the river gauges spatially and then apply the temporal matching on `prcp`.
-We will construct the interval based on peaks in `climate` since we would expect a lag effect for precipitation to flow into the river and cause a raise in river level, hence `temporal_independent = climate`.
-We select the 30 highest peak from the series to construct the match by setting `temporal_n_highest = 30`.
-This is a tuning parameter and you can start with 10% of the points of one series (here we have daily data for a year, 10% is roughly 30 points).
-`temporal_min_match` filters out pairs don't have enough match and to return all the pairs, set `temporal_min_match` to `0`.``
-
-The output from temporal matching is also a cubble, with additional column `.dist` and `.group` inherent from spatial matching and `n_match` for the number of matched temporal peaks.
-Then you can use this output to plot the location of match or to look at the series:
-
-```{r matching, out.width="100%", out.height="30%", fig.cap="asdfasd"}
-knitr::include_graphics("figures/matching.png")
-```
-
-## ERA5: climate reanalysis data
-
-\newpage
-
-## Interative graphic with cubble
-
-With a larger dataset, interactive graphics can be helpful to explore these data from multiple perspective.
-Two examples are presented here to show how interactivity can be used to explore the climate data across Australia.
-Again, `weatherdata::climate_full` will be used and we hope to include more than one time variable here.
-The difference between the maximum and minimum temperature of the day, the diurnal range, can be directly calculated from the data and it can have large impact on the wine industry (lol this is going far).
-This would be useful to compare the temperature difference across stations and to compare how this temperature difference differs throughout the year.
-
-Temperature difference is a time-wise variable and should be computed in the long form cubble. The variables are averaged over the 2016 to 2020 period by month to stabilise the result.
-
-```{r eval = FALSE, echo = TRUE}
-res <- weatherdata::climate_full %>%
-  # filter(!id %in% c("ASN00067033", "ASN00072091", 
-  #                   "ASN00059040", "ASN00097053")) %>%
-  stretch() %>%
-  mutate(month = lubridate::month(date),
-         diff = tmax - tmin) %>%
-  group_by(month) %>%
-  summarise(tmax = mean(tmax, na.rm = TRUE),
-            diff = mean(diff, na.rm = TRUE)) %>%
-  ungroup(month)
-
-```
-
-Several variables can be useful when comparing the seasonal effect of the temperature differences, for example, the difference of average temperature difference between summer and winter can tell [...].
-This is a station-wise measure and hence computed in the nested form.
-
-```{r eval = FALSE, echo = TRUE}
-res <- res %>% 
-  tamp() %>%
-  mutate(
-    diff_winter = ts %>% filter(month %in% c(6, 7, 8)) %>%
-      pull(diff) %>% mean(na.rm = TRUE),
-    diff_summer = ts %>% filter(month %in% c(12, 1, 2)) %>%
-      pull(diff) %>% mean(na.rm = TRUE),
-    diff_year = diff_winter - diff_summer
-  )
-```
-
-The result from the above manipulation can be directly supplied to make interactive plots and Figure \ref{fig:interactive-linking} is an example made with \pkg{plotly} and \pkg{crosstalk}.
-For example, the second row results from selecting the highest mean maximum temperature in July and this corresponds to Kalumburu where there is a larger winter (June to August) temperature difference than summer (December to February).
-The third row selects one of the stations (Birdsville Airport) with a large average temperature difference in the interior desert region on the map and the linked view from the time series shows that this maximum and minimum temperature difference is constant throughout the year.
-The last row selects the lowest maximum temperature in July on the temperature band and surprisingly, the corresponding station doesn't fall into the Tasmania island but Mountain Hotham in Victoria.
-
-```{r interactive-linking, echo = FALSE, out.width="100%", out.height="25%", fig.cap = "Screenshots of the mean daily temperature difference averaged by month from 2016 to 2020 at various locations in Australia.
-Crossed linking between the map and the temperature band is implemented in both directions for exploring the temperature pattern interactively.
-The first row is the general appearance of the graph and the next three rows are examples from various selection to show different temperature patterns in Australia.
-The html version can be found at ...", fig.show='hold'}
-knitr::include_graphics(here::here("figures/linking.png"))
-knitr::include_graphics(here::here("figures/linking-north.png"))
-knitr::include_graphics(here::here("figures/linking-mid.png"))
-knitr::include_graphics(here::here("figures/linking-south.png"))
-```
-
-Cubble is also compatible with \pkg{leaflet} and with leaflet extension package \pkg{leafpop}, the temperature band can be displayed upon clicking on the station on the map.
-Figure \ref{fig:interactive-popup} shows more examples of temperature pattern that can be found in Australia.
-With the same latitude as Kalumburu but in the Pacific coastline, Cooktown airport shows a similarly high but constant temperature throughout the year.
-With a slightly higher latitude than the mountain Hotham, flat land in Melbourne airport shows a less extreme temperature band.
-Lastly, the Cape Grim Baps, in the north-west tip of the Tasmania island shows a uniformly low and thin temperature band.
-
-```{r interactive-popup, echo = FALSE, out.width="50%", out.height="30%", fig.cap = "Screenshots of the diurnal temperature range averaged by month from 2016 to 2020 at various locations in Australia.
-The leaflet popup allows users to click on a single station for its temperature band as a small plot in the popup window.
-The html version can be found at ...", fig.show='hold'}
-knitr::include_graphics(here::here("figures/popup-north.png"))
-knitr::include_graphics(here::here("figures/popup-vic.png"))
-knitr::include_graphics(here::here("figures/popup-tas.png"))
-```
-
-# Conclusion
-
-# Appendix
-
-```{r illu-interactive-2, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "demon interactivity"}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.005.png"))
-```
-
-\newpage
-
-<!-- # Old stuff -->
-
-<!-- In the temporal aspect, the `tsibble` [@tsibbles] structure and its tidyverts ecosystem have provided a [...
-] workflow to work with temporal data.
-In a tsibble structure, temporal data is characterised by `index` and `key` where `index` is the temporal identifier and `key` is the identifier for multiple series, which could be used as a spatio identifier.
-However, a tsibble object, by construction, always requires the `index` in its structure.
-This makes it less appealing for spatio-temporal data since the output of calculated spatio-specific variables (i.e.
-features of each series) don't have the time dimension.
-Analysts will either need to have an additional step to join this output to the original tsibble or operate with variables stored in two separate objects.
-In addition, the long form structure of a tsibble object means spatio variables (i.e.
-longitude, latitude, and features of each series if joined back to the tsibble) of each spatio identifier will be repetitively recorded at each timestamp.
-This repetition is unnecessary and would inflate the object size for long series.
--->
-
-<!-- # Create a cubble -->
-
-<!-- The creation of a cubble requires the site identifier (`key`), as well as the spatial (`coords`) and temporal (`index`) identifier.
-`climate_flat` is already a tibble and it uses `id` to identify each station, `date` as the time identifier, and `c(long, lat)` as the spatial identifier.
-To create a cubble for this data, use: -->
-
-<!-- ```{r} -->
-<!-- climate_flat %>% as_cubble(key = id, index = date, coords = c(long, lat)) -->
-<!-- ``` -->
-
-<!-- Most of the time, spatio-temporal data doesn't come into this form and analysts need to query the climate variables based on station metadata.
-\textcolor{red}{This is also a problem illustrated in Section 3.5 in @tidydata.
-Here we provide a structured way to query this data based on the row-wise operator and nested list.} For this type of task, one can structure a metadata into a tibble and use row-wise operator to query the climate variables into a nested list.
-As an example here we demonstrate the workflow to find the 5 closest stations to Melbourne.
-We first create a station data frame with the 5 target stations.
--->
-
-<!-- ```{r echo = FALSE} -->
-<!-- latlon_ref <- tibble::tibble(id = "melbourne", lat = -37.8136, long = 144.9631) -->
-
-<!-- cand <- aus_climate %>% -->
-<!--   as_tibble() %>% -->
-<!--   select(id, lat, long, elev, name, wmo_id) -->
-
-<!-- calc_dist <- function(df_ref, df, n = 5) { -->
-<!--   long_ref <- df_ref$long -->
-<!--   lat_ref <- df_ref$lat -->
-
-<!--   df %>% -->
-<!--     mutate( -->
-<!--       long_ref = long_ref, -->
-<!--       lat_ref = lat_ref, -->
-<!--       dist = rnoaa::meteo_spherical_distance(lat, long, lat_ref, long_ref) -->
-<!--     ) %>% -->
-<!--     slice_min(dist, n = n) %>% -->
-<!--     mutate(city = df_ref$id) %>% -->
-<!--     select(-long_ref, -lat_ref) -->
-<!-- } -->
-
-<!-- (stations <- do.call("rbind", purrr::map( -->
-<!--   1:nrow(latlon_ref), -->
-<!--   ~ calc_dist(latlon_ref[.x, ], cand) -->
-<!-- ))) -->
-<!-- ``` -->
-
-<!-- We can query the climate information into a nested list named `ts` for each station with the `rowwise()` operator. To create a cubble, supply the same identifiers as with the first example. -->
-
-<!-- ```{r eval = FALSE} -->
-<!-- sydmel_climate <- stations %>% -->
-<!--   rowwise() %>% -->
-<!--   mutate(ts = list(meteo_pull_monitors(id, -->
-<!--     date_min = "2020-01-01", -->
-<!--     date_max = "2020-12-31", -->
-<!--     var = c("PRCP", "TMAX", "TMIN") -->
-<!--   ) %>% -->
-<!--     select(-id))) %>% -->
-<!--   as_cubble(key = id, index = date, coords = c(long, lat)) -->
-<!-- ``` -->
-
-<!-- ```{r echo = FALSE} -->
-<!-- (sydmel_climate <- stations %>% -->
-<!--   left_join(aus_climate) %>% -->
-<!--   as_cubble(key = id, index = date, coords = c(long, lat))) -->
-<!-- ``` -->
-
-<!-- Below are the how the nested and long form look like for Australia climate data, which records daily precipitation, maximum and minimum temperature for 55 stations across Australia from 2015- 2020.
-Notice that each station forms a group in both forms and specifically, the nested and long form have a underlying `rowwise_df` and `grouped_df` respectively.
--->
-
-<!-- With a cubic framework on mind, different types of manipulation with cubble can be thought of as slicing the cube in various way.
-The table below shows how some `dplyr` verbs are mapped into the operation in a cubble.
-With the existing grouping on the station, additional grouping can be added with `group_by` and removed with `ungrouped`.
-[talk about why it is useful] -->
-
-<!-- \newpage -->
-
-<!-- ## Cubble operations -->
-
-<!-- ```{r cubble-operations, echo = FALSE, fig.cap="Cubble operations", out.height="50%", out.width="90%", fig.align='center'} -->
-<!-- knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.004.png")) -->
-<!-- ``` -->

### 2022-01-07 4-1-basic.Rmd
play aroudnd with child rmd to organising sections

--- 2021-12-20 main.Rmd
@@
 ---
-documentclass: jss
-author:
-  - name: H. Sherry Zhang
-    # department: Department of Econometrics and Business Statistics
-    affiliation: Monash University
-    #location: Melbourne, Australia
-    email: \email{huize.zhang@monash.edu}
-  - name: Dianne Cook
-    # department: Department of Econometrics and Business Statistics
-    affiliation: 'Monash University \AND'
-    # location: Melbourne, Australia
-    # email:  dicook@monash.edu
-  - name: Ursula Laa
-    # department: Institute of Statistics
-    affiliation: 'University of Natural Resources and Life Sciences \AND'
-    # location: Vienna, Austria
-    # email:  ursula.laa@boku.ac.at
-  - name: Nicolas Langrené
-    # department: 34 Village Street, Docklands VIC 3008 Australia
-    affiliation: 'CSIRO Data61 \AND'
-    # location: Melbourne, Australia
-    # email: nicolas.langrene@csiro.au
-  - name: Patricia Menéndez
-    # department: Department of Econometrics and Business Statistics
-    affiliation: 'Monash University \AND'
-    # location: Melbourne, Australia
-    # email:  patricia.menendez@monash.edu
-title:
-  formatted: "\\pkg{cubble}: An R Package for Structuring Spatio-temporal Data"
-  # If you use tex in the formatted title, alsoge supply version without
-  plain:     "cubble: An R Package for Structuring Spatio-temporal Data"
-  # For running headers, if needed
-  # short:     "\\pkg{foo}: A Capitalized Title"
-abstract: >
-  The abstract of the article.
-keywords:
-  # at least one keyword must be supplied
-  formatted: [spatio-temporal data,  "\\proglang{R}"]
-  plain:     [spatio-temporal data, R]
-preamble: >
-  \usepackage{amsmath}
-  \usepackage{array}
-output: rticles::jss_article
-bibliography: references.bib
+title: "4-examples"
+output: html_document
 ---
-
-```{r setup, echo = FALSE}
-knitr::opts_chunk$set(
-  echo = FALSE,
-  warning = FALSE,
-  message = FALSE,
-  cache = TRUE,
-  fig.path = "figures/",
-  fig.height = 10
-)
-options(prompt = "R> ", continue = "+ ", 
-        tibble.print_max = 5, tibble.print_min = 5)
-```
-
-```{r echo = FALSE}
-library(cubble)
-library(forcats)
-library(stringr)
-library(dplyr)
-library(lubridate)
-library(ggplot2)
-library(GGally)
-library(plotly)
-library(crosstalk)
-library(patchwork)
-```
-
-\newpage
-
-# Introduction
-
-Many data structures have been proposed for spatial (\pkg{sf} by @sf) and temporal (\pkg{tsibble} by @tsibble) data in the R community, while less has been done for spatio-temporal data.
-The lack of such tools could potentially because analysts usually treat the spatial and temporal dimension of the data separately, without realising the need to create a new data structure.
-While this approach follows the third tidy data principal [@tidydata] (*Each type of observational unit forms a table*), analysts always need to manually join results from different observational units or combining multiple tables into one for downstream analysis.
-This additional step doesn't add new operations into the data but can be error prone.
-\newline
-
-Currently, available spatio-temporal data structure in R includes: \pkg{spacetime} [@spacetime], which proposes four space-time layouts: Full grid (STF), sparse grid(STS), irregular (STI), and trajectory (STT).
-The data structure it uses is based on \pkg{sp} [@sp] and \pkg{xts} [@xts], both of which has been replaced by more recent implementations.
-\pkg{spatstat} [@spatstat] implements a `ppp` class for point pattern data; and more recent, \pkg{stars} [@stars] implements a spatio-temporal array with the dplyr's data cube structure \pkg{cubelyr} [@cubelyr] as its backend.
-While these implementations either store spatial and temporal variables all in a single table, hence duplicate the spatial variables for each temporal unit; or split them into two separate tables that has the problem of manually joining, mentioned in the previously.
-None of these packages enjoy both the benefits of being able to separate manipulation in the two dimensions while also keep the data object as a whole.
-This create a gap in the software development.
-The requirement for such a tool is important given the ubiquity of spatio-temporal vector data in the wild: the Ireland wind data from \pkg{gstat} is an classic example data that splits variables into spatial (`wind.loc`) and temporal (`wind`) dimension; Bureau of Meteorology (BoM) provides climate observations that are widely applied in agriculture and ecology study; air pollution data.
-\newline
-
-This paper describes the implementation of a new spatio-temporal data structure: \pkg{cubble}.
-\pkg{cubble} implements a relational data structure that uses two forms to manage the switch between spatial and temporal dimension.
-With this structure, users can manipulate the spatial or temporal dimension separately, while leaves the linking of two dimensions to \pkg{cubble}.
-The software is available from the Comprehensive R Archive Network (CRAN) at [CRAN link].
-\newline
-
-The rest of the paper will be divided as follows: [complete when the paper structure is more solid]
-
-<!-- Section 2 reviews the existing data structure for spatio, temporal, and spatio-temporal data.
-Section 3 presents a new data structure for spatio-temporal data: cubble.
-Then the paper introduces the workflow of data manipulation and visualisation with the cubble structure in Section 4.
-Section 5 gives some examples on how common spatial and temporal manipulations are performed with cubble and how static and interactive visualisation help to understand climate and [...] data.
-\newline -->
-
-\newpage
-
-# The cubble package
-
-Spatio-temporal data usually come in various forms and Figure \ref{fig:illu-input} shows four examples of this.
-No matter which form the data is in, these formats share some common components that characterise spatio-temporal data.
-A spatial identifier (`id` in the diagram) is the unique identifier of each site.
-The temporal identifier (`t` in the diagram) prescribes the time stamp each site is recorded.
-Coordinates, comprising of latitude and longitude (`lon` and `lat` in the diagram), locates each site on the map.
-These identifiers will be the building blocks for the data structure introduced below.
-Other variables in the data can be categorised into two groups: spatial variables that are invariant at each time stamp for every site, i.e.
-the name or code of the weather station and temporal variables that varies with time.
-
-```{r illu-input, echo = FALSE, out.width = "100%", fig.cap="Illustration of incoming data formats for spatio-temporal data.
-(1) Data comes in as a single table; (2) Separate tables for spatial and temporal variables; (3) A single table with all the parameters used to query the database and a separate table for queried data; and (4) Cubical data in array or NetCDF format.", fig.align='center', out.height = "15%"}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.001.png"))
-```
-
-In a cubble, there are two forms: nested form and long form, and Figure \ref{fig:illu-cubble} sketches the two forms along with the associated attributes.
-The decision on which form to use is output-oriented, meaning analysts need to first think about whether the output of a particular operation is identified only by the spatial identifier, or a combination of spatial and temporal identifier.
-The nested cubble is suitable for working with operations that are only identified by site and this type of operation can be a pure manipulation of spatial variables, or a summary of temporal variables by site (i.e.
-the output of counting the number of raining day is only identified by sites and hence should be performed with the nested form).
-Underneath the nested form, a cubble is built from a row-wise dataframe (`rowwise_df`) where each site forms a separate group.
-This structure simplifies the calculation that involves temporal variables by avoiding the use of `map` syntax when working with list-column.
-
-For those operations whose output involves both a spatial and temporal dimension, long form should be used.
-The long form is identified by both the spatial and temporal identifier and adopts a grouped dataframe (`grouped_df`) to forms each site as a group.
-Spatial variables are stored separately in a \pkg{tibble} as an special attribute of the long cubble.
-This design avoids repeating the spatial variables at each time stamp while not dropping information from spatial variables.
-
-```{r illu-cubble, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "Illustration of nested and long cubble."}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.002.png"))
-```
-
-## Create a cubble in the nested form
-
-To use functionalities from cubble, data analysts first need to create a cubble.
-`as_cubble` create a \pkg{cubble} by supplying the three key components: `key` as the spatial identifier; `index` as the temporal identifier; and a vector of `coords` in the order of longitude first and then latitude.
-The naming of `key` and `index` follows the convention in the \pkg{tsibble} package.
-The cubble created by default is in the nested form.
-Below is an example of creating a cubble: \newline
-
-```{r}
-(cubble_nested <- cubble::climate_flat %>%
-  as_cubble(key = id, index = date, coords = c("long", "lat")))
-```
-
-There are a few information in the \pkg{cubble} header: the name of the `key` variable, bbox, and also the name of variable nested in the `ts` column.
-In this example, each site is identifier is `id` and the number in the bracket means there are 5 unique `id` in this dataset.
-The bbox in the second row gives the range of the coordinates.
-The temporal variables are all nested in the `ts` column, but it could be useful to know the name these variables.
-The third row in the cubble header shows these names and in this example this includes: precipitation, `prcp`, maximum temperature, `tmax`, and minimum temperature, `tmin`.
-
-## Stretch a nested cubble into the long form
-
-The long cubble is suitable to manipulate the time dimension of the data.
-The function `stretch()` switches the nested cubble into the long cubble by first extracts all the spatial variables into a separate tibble and store in the `spatial` attribute and then unnests the `ts` column:
-
-```{r}
-(cubble_long <- cubble_nested %>% stretch(ts))
-```
-
-Notice here that the third line in the header now shows the name of spatial variables rather than the temporal variables.
-
-## Tamp a long cubble back to the nested form
-
-Manipulation on the spatial and temporal dimension can be an iterative process.
-Many times, analysts will need to go back and forth between the nested and long cubble.
-The `stretch()` function introduced in the previous section switches a nested cubble into a long cubble and function `tamp()` is its inverse function to switch a long cubble back to the nested cubble:
-
-```{r}
-(cubble_back <- cubble_long %>% tamp())
-```
-
-## Migrate spatial variables to a long cubble
-
-As a final data output for modelling or visualisation, spatio-temporal data is usually expected to be in a single table.
-Function `migrate()` moves the spatial variables from the `spatial` attribute into the long cubble:
-
-```{r}
-(cubble_long %>% migrate(long, lat))
-```
-
-In this workflow described above, data objects come into cubble in the nested form, then various operations on the spatial and temporal dimension can go back and forth between the nested and long form, and finally, the data will come out of cubble in the long form for further modelling or visualisation.
-
-Building from an underlying `tbl_df` structure, it is natural to implement methods available in `dplyr` to `cubble`. Supported methods in the `cubble` with `dplyr` generics includes:
-
-```{=tex}
-\begin{center}
-\begin{tabular}{ | m{5em} | m{15cm}| } 
-basics & \textbf{mutate}, \textbf{filter}, \textbf{summarise}, \textbf{select}, \textbf{arrange}, \textbf{rename}, \textbf{left\_join} \\
-grouping &  \textbf{group\_by}, \textbf{ungroup}\\
-slice family & \textbf{slice\_head}, \textbf{slice\_tail}, \textbf{slice\_sample}, \textbf{slice\_min} and \textbf{slice\_max} \\
-\end{tabular}
-\end{center}
-```
-\pkg{cubble} is also compatible with \pkg{tsibble} in the sense that the original list-column can be a `tbl_ts` object.
-Duplicates and gaps should be first checked before structuring the data into a cubble.
-If the input data is a \pkg{tsibble} object, the long form cubble is also a \pkg{tsibble} where users can directly apply time series operations.
-
-# Advanced features/ considerations
-
-## Hierarchical structure
-
-Imposing a clustering structure can be thought of as building a hierarchical structure where stations are nested within clusters.
-This can be useful when there's intrinsic nesting structure in the data (i.e.
-country nested in the continent, county nested in the state) or there's some clustering.
-When we have access to both level, `switch_key()` is the function to re-structure the data as one cluster per row.
-Temporal observations from different stations while within the same cluster are bound in the nested column `ts`.
-
-One thing we hope to do with the cluster is to find the coordinates of the centroid.
-These are variables variant to the station but invariant to the cluster and it would be nice to have a function that structure each cluster as a row.
-`switch_key()` is the function that does this: it lets you to specify a new key, say `cluster` and nests all spatial variables variant to `cluster` into a column.
-Spatial variables are all nested inside a new column `.val`.
-Temporal observations from different stations while within the same cluster are bound in the nested column `ts`.
-
-This structure makes it easy to compute cluster level variable, for example, the convex hull and the centroid coordinate of each cluster.
-These can be amended into the nested form with function `get_centroid`.
-
-After we have got `cluster_nested`, spatial and temporal data at both levels can be easily obtained.
-Figure \ref{fig:illu-hier} illustrate the relationship between the long and nested form cubble at both site and cluster level.
-More description on this.
-Start with the original `station_nested`, `stretch()` expands the `ts` column with each station (`id`) forming a group and attach variables invariant to `id` as an attribute.
-`switch_key()` changes the `key` from `id` to `cluster` and nests all the spatial variables that variant to `cluster`.
-`stretch()` `cluster_nested` will store variables that are invariant to `cluster` as a tibble in the attribute.
-
-this fit into the remaining data pipeline.
-
-```{r illu-hier, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "Hierarchical structure"}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.003.png"))
-```
-
-## Data fusion and matching
-
-Temporal matching checks how spatially matched pairs align temporally. We use the following chart to illustrate how the temporal matching works:
-
-For each spatially matched pair, say `A` and `a`, we first find the largest `n` points in each series, colored in brown points here.
-Here we use the largest three but you can tune this number by `temporal_n_highest`.
-Then we construct the interval of the largest points from one series and see how many points, from the other series, fall into the intervals.
-The series used to construct the interval is controlled by `temporal_independent` and the window size by `temporal_window` with a default of 5.
-
-In this illustration, we construct the interval based on series `A` and two of the three peaks from `a` falls into this interval at Time 7 and 27.
-
-There's another mandatory argument that hasn't been introduced above: `temporal_var_to_match`.
-This argument controls the variable to match and it needs to appear in both the `major` and `minor` set.
-In the water level matching example, we match the variable `Water_course_level` from `river` to `prcp` from `climate`, hence need to manually rename one of them to match the other, here we rename `Water_course_level` to `prcp` in `river`:
-
-Figure \ref{fig:illu-matching}
-
-```{r illu-matching, out.width="100%", out.height="30%", fig.cap="sdfasdf"}
-knitr::include_graphics("figures/illu-matching.png")
-```
-
-## Interactive graphics
-
-<!-- Interactive graphics can listen to users' actions on the plot to provide additional information that facilitates data exploration.
-This is a useful technique for spatio-temporal data since users can zoom or pan the map to view the local and global structure of the map; use tooltips or popups to query more information about a graphic element; or highlight points to explore its linked views in other plots.
-In the R community, many implementations have been developed to connect \proglang{R} to \proglang{javascript} to create interactive graphics.
-In relation to spatio, temporal, and spatio-temporal data, the general purpose packages \pkg{plotly} [@plotly] and \pkg{leaflet} [@leaflet] realise various interactive actions through their corresponding javascript libraries.
-\pkg{crosstalk} [@crosstalk] and \pkg{tsibbletalk} [@tsibbletalk] implement brushed linking between htmlwidgets.
-\pkg{ggiraph} [@ggiraph] enables tooltip, self-linking, and customised actions specified through its own \proglang{javascript}.
--->
-
-<!-- While many graphic implementations present worked examples to illustrate the usage of the package, few documents the underlying pipeline that transforms the raw data step-by-step to the final view on the screen.
-There have been some early work in building the data pipeline for (interactive) graphics [@buja1988elements; @buja1996interactive;  @sutherland2000orca] and more recent discussions include @wickham2009plumbing, @xie2014reactive, and @cheng2016enabling.
--->
-
-The cubble structure fits in naturally with the interactive graphic pipeline discussed in the literature [@buja1988elements; @buja1996interactive;  @sutherland2000orca; @xie2014reactive; @cheng2016enabling].
-Diagram \ref{fig:illu-interactive} illustrates how linking works with the two forms in a cubble, where a time series plot is created with the long cubble and a map is created with the nested cubble.
-When a user action is captured from the map, the site will be activated in the nested cubble.
-Then, the nested cubble will communicate to the long cubble to activate all the observations with the same `id`.
-The long cubble will then highlight the  activated series in the time series plot.
-
-The linking is also available from the time series plot to the map.
-The selection on the time series is through selecting the point on the time series and once a point is selected, it will be activated in the long cubble.
-All the observations that share the same `id`,  either in the long and nested cubble, are then activated.
-This includes other points in the same time series in the long cubble and the corresponding observation of site in the nested cubble.
-These activated observations will then being reflected in the updated plots and Diagram \ref{fig:illu-interactive-2} in the Appendix illustrates this process.
-
-```{r illu-interactive, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "demon interactivity"}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.004.png"))
-```
-
-\newpage
 
@@
 ```
-
-## Australia precipitation pattern in 2020
-
-In Figure \ref{fig:basic-agg}.
-Aggregating them into averages or other statistics would give a better picture.
-This can be done with cubble through a hierarchical structure, where the cluster indicator is one level above the individual station.
-
-The `weatherdata::climate_full` data has daily climate data of `r nrow(weatherdata::climate_full)` Australia stations from 2016 to 2020.
-
-calcluate distance matrix
-As an example to illustrate here, a kmean clustering algorithm based on the distance matrix is used and the number of centres is set to 20.
-More complex algorithms can also be used for more complex problem, as long as a mapping from each station id to the cluster id can be constructed.
-
-<!-- Now we can obtain the aggregated series as described in the workflow diagram above and construct the glyph map with `GGally::glyphs()`: -->
-
-<!-- We can also look at the precipitation of each individual station within the same cluster: -->
-
-<!-- Lastly, there is one series on Tasmania island standing out from others, lets look at where it is: -->
-
-```{r basic-agg, out.width="100%", out.height="90%", fig.cap="sdfasdf"}
-knitr::include_graphics("figures/basic-agg.png")
-```
-
-## River level data in Victria water gauges
-
-The water level data comes from [Bureau of Meteorology](http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full) and has a copy in `weatherdata`.
-Here we extract the water course level and add a column annotate this data of type `river`.
-For the rainfall data, we will still use the `weatherdata::climate_full`, filtering for Victorian stations in 2020 should be pretty familiar by now.
-Again, we first look at where these stations are on the map first:
-
-Now we use `match_sites()` to first pair the weather stations with the river gauges spatially and then apply the temporal matching on `prcp`.
-We will construct the interval based on peaks in `climate` since we would expect a lag effect for precipitation to flow into the river and cause a raise in river level, hence `temporal_independent = climate`.
-We select the 30 highest peak from the series to construct the match by setting `temporal_n_highest = 30`.
-This is a tuning parameter and you can start with 10% of the points of one series (here we have daily data for a year, 10% is roughly 30 points).
-`temporal_min_match` filters out pairs don't have enough match and to return all the pairs, set `temporal_min_match` to `0`.``
-
-The output from temporal matching is also a cubble, with additional column `.dist` and `.group` inherent from spatial matching and `n_match` for the number of matched temporal peaks.
-Then you can use this output to plot the location of match or to look at the series:
-
-```{r matching, out.width="100%", out.height="30%", fig.cap="asdfasd"}
-knitr::include_graphics("figures/matching.png")
-```
-
-## ERA5: climate reanalysis data
-
-\newpage
-
-## Interative graphic with cubble
-
-With a larger dataset, interactive graphics can be helpful to explore these data from multiple perspective.
-Two examples are presented here to show how interactivity can be used to explore the climate data across Australia.
-Again, `weatherdata::climate_full` will be used and we hope to include more than one time variable here.
-The difference between the maximum and minimum temperature of the day, the diurnal range, can be directly calculated from the data and it can have large impact on the wine industry (lol this is going far).
-This would be useful to compare the temperature difference across stations and to compare how this temperature difference differs throughout the year.
-
-Temperature difference is a time-wise variable and should be computed in the long form cubble. The variables are averaged over the 2016 to 2020 period by month to stabilise the result.
-
-```{r eval = FALSE, echo = TRUE}
-res <- weatherdata::climate_full %>%
-  # filter(!id %in% c("ASN00067033", "ASN00072091", 
-  #                   "ASN00059040", "ASN00097053")) %>%
-  stretch() %>%
-  mutate(month = lubridate::month(date),
-         diff = tmax - tmin) %>%
-  group_by(month) %>%
-  summarise(tmax = mean(tmax, na.rm = TRUE),
-            diff = mean(diff, na.rm = TRUE)) %>%
-  ungroup(month)
-
-```
-
-Several variables can be useful when comparing the seasonal effect of the temperature differences, for example, the difference of average temperature difference between summer and winter can tell [...].
-This is a station-wise measure and hence computed in the nested form.
-
-```{r eval = FALSE, echo = TRUE}
-res <- res %>% 
-  tamp() %>%
-  mutate(
-    diff_winter = ts %>% filter(month %in% c(6, 7, 8)) %>%
-      pull(diff) %>% mean(na.rm = TRUE),
-    diff_summer = ts %>% filter(month %in% c(12, 1, 2)) %>%
-      pull(diff) %>% mean(na.rm = TRUE),
-    diff_year = diff_winter - diff_summer
-  )
-```
-
-The result from the above manipulation can be directly supplied to make interactive plots and Figure \ref{fig:interactive-linking} is an example made with \pkg{plotly} and \pkg{crosstalk}.
-For example, the second row results from selecting the highest mean maximum temperature in July and this corresponds to Kalumburu where there is a larger winter (June to August) temperature difference than summer (December to February).
-The third row selects one of the stations (Birdsville Airport) with a large average temperature difference in the interior desert region on the map and the linked view from the time series shows that this maximum and minimum temperature difference is constant throughout the year.
-The last row selects the lowest maximum temperature in July on the temperature band and surprisingly, the corresponding station doesn't fall into the Tasmania island but Mountain Hotham in Victoria.
-
-```{r interactive-linking, echo = FALSE, out.width="100%", out.height="25%", fig.cap = "Screenshots of the mean daily temperature difference averaged by month from 2016 to 2020 at various locations in Australia.
-Crossed linking between the map and the temperature band is implemented in both directions for exploring the temperature pattern interactively.
-The first row is the general appearance of the graph and the next three rows are examples from various selection to show different temperature patterns in Australia.
-The html version can be found at ...", fig.show='hold'}
-knitr::include_graphics(here::here("figures/linking.png"))
-knitr::include_graphics(here::here("figures/linking-north.png"))
-knitr::include_graphics(here::here("figures/linking-mid.png"))
-knitr::include_graphics(here::here("figures/linking-south.png"))
-```
-
-Cubble is also compatible with \pkg{leaflet} and with leaflet extension package \pkg{leafpop}, the temperature band can be displayed upon clicking on the station on the map.
-Figure \ref{fig:interactive-popup} shows more examples of temperature pattern that can be found in Australia.
-With the same latitude as Kalumburu but in the Pacific coastline, Cooktown airport shows a similarly high but constant temperature throughout the year.
-With a slightly higher latitude than the mountain Hotham, flat land in Melbourne airport shows a less extreme temperature band.
-Lastly, the Cape Grim Baps, in the north-west tip of the Tasmania island shows a uniformly low and thin temperature band.
-
-```{r interactive-popup, echo = FALSE, out.width="50%", out.height="30%", fig.cap = "Screenshots of the diurnal temperature range averaged by month from 2016 to 2020 at various locations in Australia.
-The leaflet popup allows users to click on a single station for its temperature band as a small plot in the popup window.
-The html version can be found at ...", fig.show='hold'}
-knitr::include_graphics(here::here("figures/popup-north.png"))
-knitr::include_graphics(here::here("figures/popup-vic.png"))
-knitr::include_graphics(here::here("figures/popup-tas.png"))
-```
-
-# Conclusion
-
-# Appendix
-
-```{r illu-interactive-2, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "demon interactivity"}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.005.png"))
-```
-
-\newpage
-
-<!-- # Old stuff -->
-
-<!-- In the temporal aspect, the `tsibble` [@tsibbles] structure and its tidyverts ecosystem have provided a [...
-] workflow to work with temporal data.
-In a tsibble structure, temporal data is characterised by `index` and `key` where `index` is the temporal identifier and `key` is the identifier for multiple series, which could be used as a spatio identifier.
-However, a tsibble object, by construction, always requires the `index` in its structure.
-This makes it less appealing for spatio-temporal data since the output of calculated spatio-specific variables (i.e.
-features of each series) don't have the time dimension.
-Analysts will either need to have an additional step to join this output to the original tsibble or operate with variables stored in two separate objects.
-In addition, the long form structure of a tsibble object means spatio variables (i.e.
-longitude, latitude, and features of each series if joined back to the tsibble) of each spatio identifier will be repetitively recorded at each timestamp.
-This repetition is unnecessary and would inflate the object size for long series.
--->
-
-<!-- # Create a cubble -->
-
-<!-- The creation of a cubble requires the site identifier (`key`), as well as the spatial (`coords`) and temporal (`index`) identifier.
-`climate_flat` is already a tibble and it uses `id` to identify each station, `date` as the time identifier, and `c(long, lat)` as the spatial identifier.
-To create a cubble for this data, use: -->
-
-<!-- ```{r} -->
-<!-- climate_flat %>% as_cubble(key = id, index = date, coords = c(long, lat)) -->
-<!-- ``` -->
-
-<!-- Most of the time, spatio-temporal data doesn't come into this form and analysts need to query the climate variables based on station metadata.
-\textcolor{red}{This is also a problem illustrated in Section 3.5 in @tidydata.
-Here we provide a structured way to query this data based on the row-wise operator and nested list.} For this type of task, one can structure a metadata into a tibble and use row-wise operator to query the climate variables into a nested list.
-As an example here we demonstrate the workflow to find the 5 closest stations to Melbourne.
-We first create a station data frame with the 5 target stations.
--->
-
-<!-- ```{r echo = FALSE} -->
-<!-- latlon_ref <- tibble::tibble(id = "melbourne", lat = -37.8136, long = 144.9631) -->
-
-<!-- cand <- aus_climate %>% -->
-<!--   as_tibble() %>% -->
-<!--   select(id, lat, long, elev, name, wmo_id) -->
-
-<!-- calc_dist <- function(df_ref, df, n = 5) { -->
-<!--   long_ref <- df_ref$long -->
-<!--   lat_ref <- df_ref$lat -->
-
-<!--   df %>% -->
-<!--     mutate( -->
-<!--       long_ref = long_ref, -->
-<!--       lat_ref = lat_ref, -->
-<!--       dist = rnoaa::meteo_spherical_distance(lat, long, lat_ref, long_ref) -->
-<!--     ) %>% -->
-<!--     slice_min(dist, n = n) %>% -->
-<!--     mutate(city = df_ref$id) %>% -->
-<!--     select(-long_ref, -lat_ref) -->
-<!-- } -->
-
-<!-- (stations <- do.call("rbind", purrr::map( -->
-<!--   1:nrow(latlon_ref), -->
-<!--   ~ calc_dist(latlon_ref[.x, ], cand) -->
-<!-- ))) -->
-<!-- ``` -->
-
-<!-- We can query the climate information into a nested list named `ts` for each station with the `rowwise()` operator. To create a cubble, supply the same identifiers as with the first example. -->
-
-<!-- ```{r eval = FALSE} -->
-<!-- sydmel_climate <- stations %>% -->
-<!--   rowwise() %>% -->
-<!--   mutate(ts = list(meteo_pull_monitors(id, -->
-<!--     date_min = "2020-01-01", -->
-<!--     date_max = "2020-12-31", -->
-<!--     var = c("PRCP", "TMAX", "TMIN") -->
-<!--   ) %>% -->
-<!--     select(-id))) %>% -->
-<!--   as_cubble(key = id, index = date, coords = c(long, lat)) -->
-<!-- ``` -->
-
-<!-- ```{r echo = FALSE} -->
-<!-- (sydmel_climate <- stations %>% -->
-<!--   left_join(aus_climate) %>% -->
-<!--   as_cubble(key = id, index = date, coords = c(long, lat))) -->
-<!-- ``` -->
-
-<!-- Below are the how the nested and long form look like for Australia climate data, which records daily precipitation, maximum and minimum temperature for 55 stations across Australia from 2015- 2020.
-Notice that each station forms a group in both forms and specifically, the nested and long form have a underlying `rowwise_df` and `grouped_df` respectively.
--->
-
-<!-- With a cubic framework on mind, different types of manipulation with cubble can be thought of as slicing the cube in various way.
-The table below shows how some `dplyr` verbs are mapped into the operation in a cubble.
-With the existing grouping on the station, additional grouping can be added with `group_by` and removed with `ungrouped`.
-[talk about why it is useful] -->
-
-<!-- \newpage -->
-
-<!-- ## Cubble operations -->
-
-<!-- ```{r cubble-operations, echo = FALSE, fig.cap="Cubble operations", out.height="50%", out.width="90%", fig.align='center'} -->
-<!-- knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.004.png")) -->
-<!-- ``` -->

### 2022-01-07 4-2-agg.Rmd
play aroudnd with child rmd to organising sections

--- 2021-12-20 main.Rmd
@@
 ---
-documentclass: jss
-author:
-  - name: H. Sherry Zhang
-    # department: Department of Econometrics and Business Statistics
-    affiliation: Monash University
-    #location: Melbourne, Australia
-    email: \email{huize.zhang@monash.edu}
-  - name: Dianne Cook
-    # department: Department of Econometrics and Business Statistics
-    affiliation: 'Monash University \AND'
-    # location: Melbourne, Australia
-    # email:  dicook@monash.edu
-  - name: Ursula Laa
-    # department: Institute of Statistics
-    affiliation: 'University of Natural Resources and Life Sciences \AND'
-    # location: Vienna, Austria
-    # email:  ursula.laa@boku.ac.at
-  - name: Nicolas Langrené
-    # department: 34 Village Street, Docklands VIC 3008 Australia
-    affiliation: 'CSIRO Data61 \AND'
-    # location: Melbourne, Australia
-    # email: nicolas.langrene@csiro.au
-  - name: Patricia Menéndez
-    # department: Department of Econometrics and Business Statistics
-    affiliation: 'Monash University \AND'
-    # location: Melbourne, Australia
-    # email:  patricia.menendez@monash.edu
-title:
-  formatted: "\\pkg{cubble}: An R Package for Structuring Spatio-temporal Data"
-  # If you use tex in the formatted title, alsoge supply version without
-  plain:     "cubble: An R Package for Structuring Spatio-temporal Data"
-  # For running headers, if needed
-  # short:     "\\pkg{foo}: A Capitalized Title"
-abstract: >
-  The abstract of the article.
-keywords:
-  # at least one keyword must be supplied
-  formatted: [spatio-temporal data,  "\\proglang{R}"]
-  plain:     [spatio-temporal data, R]
-preamble: >
-  \usepackage{amsmath}
-  \usepackage{array}
-output: rticles::jss_article
-bibliography: references.bib
+title: "4-2-aggre"
+output: pdf_document
 ---
-
-```{r setup, echo = FALSE}
-knitr::opts_chunk$set(
-  echo = FALSE,
-  warning = FALSE,
-  message = FALSE,
-  cache = TRUE,
-  fig.path = "figures/",
-  fig.height = 10
-)
-options(prompt = "R> ", continue = "+ ", 
-        tibble.print_max = 5, tibble.print_min = 5)
-```
-
-```{r echo = FALSE}
-library(cubble)
-library(forcats)
-library(stringr)
-library(dplyr)
-library(lubridate)
-library(ggplot2)
-library(GGally)
-library(plotly)
-library(crosstalk)
-library(patchwork)
-```
-
-\newpage
-
-# Introduction
-
-Many data structures have been proposed for spatial (\pkg{sf} by @sf) and temporal (\pkg{tsibble} by @tsibble) data in the R community, while less has been done for spatio-temporal data.
-The lack of such tools could potentially because analysts usually treat the spatial and temporal dimension of the data separately, without realising the need to create a new data structure.
-While this approach follows the third tidy data principal [@tidydata] (*Each type of observational unit forms a table*), analysts always need to manually join results from different observational units or combining multiple tables into one for downstream analysis.
-This additional step doesn't add new operations into the data but can be error prone.
-\newline
-
-Currently, available spatio-temporal data structure in R includes: \pkg{spacetime} [@spacetime], which proposes four space-time layouts: Full grid (STF), sparse grid(STS), irregular (STI), and trajectory (STT).
-The data structure it uses is based on \pkg{sp} [@sp] and \pkg{xts} [@xts], both of which has been replaced by more recent implementations.
-\pkg{spatstat} [@spatstat] implements a `ppp` class for point pattern data; and more recent, \pkg{stars} [@stars] implements a spatio-temporal array with the dplyr's data cube structure \pkg{cubelyr} [@cubelyr] as its backend.
-While these implementations either store spatial and temporal variables all in a single table, hence duplicate the spatial variables for each temporal unit; or split them into two separate tables that has the problem of manually joining, mentioned in the previously.
-None of these packages enjoy both the benefits of being able to separate manipulation in the two dimensions while also keep the data object as a whole.
-This create a gap in the software development.
-The requirement for such a tool is important given the ubiquity of spatio-temporal vector data in the wild: the Ireland wind data from \pkg{gstat} is an classic example data that splits variables into spatial (`wind.loc`) and temporal (`wind`) dimension; Bureau of Meteorology (BoM) provides climate observations that are widely applied in agriculture and ecology study; air pollution data.
-\newline
-
-This paper describes the implementation of a new spatio-temporal data structure: \pkg{cubble}.
-\pkg{cubble} implements a relational data structure that uses two forms to manage the switch between spatial and temporal dimension.
-With this structure, users can manipulate the spatial or temporal dimension separately, while leaves the linking of two dimensions to \pkg{cubble}.
-The software is available from the Comprehensive R Archive Network (CRAN) at [CRAN link].
-\newline
-
-The rest of the paper will be divided as follows: [complete when the paper structure is more solid]
-
-<!-- Section 2 reviews the existing data structure for spatio, temporal, and spatio-temporal data.
-Section 3 presents a new data structure for spatio-temporal data: cubble.
-Then the paper introduces the workflow of data manipulation and visualisation with the cubble structure in Section 4.
-Section 5 gives some examples on how common spatial and temporal manipulations are performed with cubble and how static and interactive visualisation help to understand climate and [...] data.
-\newline -->
-
-\newpage
-
-# The cubble package
-
-Spatio-temporal data usually come in various forms and Figure \ref{fig:illu-input} shows four examples of this.
-No matter which form the data is in, these formats share some common components that characterise spatio-temporal data.
-A spatial identifier (`id` in the diagram) is the unique identifier of each site.
-The temporal identifier (`t` in the diagram) prescribes the time stamp each site is recorded.
-Coordinates, comprising of latitude and longitude (`lon` and `lat` in the diagram), locates each site on the map.
-These identifiers will be the building blocks for the data structure introduced below.
-Other variables in the data can be categorised into two groups: spatial variables that are invariant at each time stamp for every site, i.e.
-the name or code of the weather station and temporal variables that varies with time.
-
-```{r illu-input, echo = FALSE, out.width = "100%", fig.cap="Illustration of incoming data formats for spatio-temporal data.
-(1) Data comes in as a single table; (2) Separate tables for spatial and temporal variables; (3) A single table with all the parameters used to query the database and a separate table for queried data; and (4) Cubical data in array or NetCDF format.", fig.align='center', out.height = "15%"}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.001.png"))
-```
-
-In a cubble, there are two forms: nested form and long form, and Figure \ref{fig:illu-cubble} sketches the two forms along with the associated attributes.
-The decision on which form to use is output-oriented, meaning analysts need to first think about whether the output of a particular operation is identified only by the spatial identifier, or a combination of spatial and temporal identifier.
-The nested cubble is suitable for working with operations that are only identified by site and this type of operation can be a pure manipulation of spatial variables, or a summary of temporal variables by site (i.e.
-the output of counting the number of raining day is only identified by sites and hence should be performed with the nested form).
-Underneath the nested form, a cubble is built from a row-wise dataframe (`rowwise_df`) where each site forms a separate group.
-This structure simplifies the calculation that involves temporal variables by avoiding the use of `map` syntax when working with list-column.
-
-For those operations whose output involves both a spatial and temporal dimension, long form should be used.
-The long form is identified by both the spatial and temporal identifier and adopts a grouped dataframe (`grouped_df`) to forms each site as a group.
-Spatial variables are stored separately in a \pkg{tibble} as an special attribute of the long cubble.
-This design avoids repeating the spatial variables at each time stamp while not dropping information from spatial variables.
-
-```{r illu-cubble, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "Illustration of nested and long cubble."}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.002.png"))
-```
-
-## Create a cubble in the nested form
-
-To use functionalities from cubble, data analysts first need to create a cubble.
-`as_cubble` create a \pkg{cubble} by supplying the three key components: `key` as the spatial identifier; `index` as the temporal identifier; and a vector of `coords` in the order of longitude first and then latitude.
-The naming of `key` and `index` follows the convention in the \pkg{tsibble} package.
-The cubble created by default is in the nested form.
-Below is an example of creating a cubble: \newline
-
-```{r}
-(cubble_nested <- cubble::climate_flat %>%
-  as_cubble(key = id, index = date, coords = c("long", "lat")))
-```
-
-There are a few information in the \pkg{cubble} header: the name of the `key` variable, bbox, and also the name of variable nested in the `ts` column.
-In this example, each site is identifier is `id` and the number in the bracket means there are 5 unique `id` in this dataset.
-The bbox in the second row gives the range of the coordinates.
-The temporal variables are all nested in the `ts` column, but it could be useful to know the name these variables.
-The third row in the cubble header shows these names and in this example this includes: precipitation, `prcp`, maximum temperature, `tmax`, and minimum temperature, `tmin`.
-
-## Stretch a nested cubble into the long form
-
-The long cubble is suitable to manipulate the time dimension of the data.
-The function `stretch()` switches the nested cubble into the long cubble by first extracts all the spatial variables into a separate tibble and store in the `spatial` attribute and then unnests the `ts` column:
-
-```{r}
-(cubble_long <- cubble_nested %>% stretch(ts))
-```
-
-Notice here that the third line in the header now shows the name of spatial variables rather than the temporal variables.
-
-## Tamp a long cubble back to the nested form
-
-Manipulation on the spatial and temporal dimension can be an iterative process.
-Many times, analysts will need to go back and forth between the nested and long cubble.
-The `stretch()` function introduced in the previous section switches a nested cubble into a long cubble and function `tamp()` is its inverse function to switch a long cubble back to the nested cubble:
-
-```{r}
-(cubble_back <- cubble_long %>% tamp())
-```
-
-## Migrate spatial variables to a long cubble
-
-As a final data output for modelling or visualisation, spatio-temporal data is usually expected to be in a single table.
-Function `migrate()` moves the spatial variables from the `spatial` attribute into the long cubble:
-
-```{r}
-(cubble_long %>% migrate(long, lat))
-```
-
-In this workflow described above, data objects come into cubble in the nested form, then various operations on the spatial and temporal dimension can go back and forth between the nested and long form, and finally, the data will come out of cubble in the long form for further modelling or visualisation.
-
-Building from an underlying `tbl_df` structure, it is natural to implement methods available in `dplyr` to `cubble`. Supported methods in the `cubble` with `dplyr` generics includes:
-
-```{=tex}
-\begin{center}
-\begin{tabular}{ | m{5em} | m{15cm}| } 
-basics & \textbf{mutate}, \textbf{filter}, \textbf{summarise}, \textbf{select}, \textbf{arrange}, \textbf{rename}, \textbf{left\_join} \\
-grouping &  \textbf{group\_by}, \textbf{ungroup}\\
-slice family & \textbf{slice\_head}, \textbf{slice\_tail}, \textbf{slice\_sample}, \textbf{slice\_min} and \textbf{slice\_max} \\
-\end{tabular}
-\end{center}
-```
-\pkg{cubble} is also compatible with \pkg{tsibble} in the sense that the original list-column can be a `tbl_ts` object.
-Duplicates and gaps should be first checked before structuring the data into a cubble.
-If the input data is a \pkg{tsibble} object, the long form cubble is also a \pkg{tsibble} where users can directly apply time series operations.
-
-# Advanced features/ considerations
-
-## Hierarchical structure
-
-Imposing a clustering structure can be thought of as building a hierarchical structure where stations are nested within clusters.
-This can be useful when there's intrinsic nesting structure in the data (i.e.
-country nested in the continent, county nested in the state) or there's some clustering.
-When we have access to both level, `switch_key()` is the function to re-structure the data as one cluster per row.
-Temporal observations from different stations while within the same cluster are bound in the nested column `ts`.
-
-One thing we hope to do with the cluster is to find the coordinates of the centroid.
-These are variables variant to the station but invariant to the cluster and it would be nice to have a function that structure each cluster as a row.
-`switch_key()` is the function that does this: it lets you to specify a new key, say `cluster` and nests all spatial variables variant to `cluster` into a column.
-Spatial variables are all nested inside a new column `.val`.
-Temporal observations from different stations while within the same cluster are bound in the nested column `ts`.
-
-This structure makes it easy to compute cluster level variable, for example, the convex hull and the centroid coordinate of each cluster.
-These can be amended into the nested form with function `get_centroid`.
-
-After we have got `cluster_nested`, spatial and temporal data at both levels can be easily obtained.
-Figure \ref{fig:illu-hier} illustrate the relationship between the long and nested form cubble at both site and cluster level.
-More description on this.
-Start with the original `station_nested`, `stretch()` expands the `ts` column with each station (`id`) forming a group and attach variables invariant to `id` as an attribute.
-`switch_key()` changes the `key` from `id` to `cluster` and nests all the spatial variables that variant to `cluster`.
-`stretch()` `cluster_nested` will store variables that are invariant to `cluster` as a tibble in the attribute.
-
-this fit into the remaining data pipeline.
-
-```{r illu-hier, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "Hierarchical structure"}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.003.png"))
-```
-
-## Data fusion and matching
-
-Temporal matching checks how spatially matched pairs align temporally. We use the following chart to illustrate how the temporal matching works:
-
-For each spatially matched pair, say `A` and `a`, we first find the largest `n` points in each series, colored in brown points here.
-Here we use the largest three but you can tune this number by `temporal_n_highest`.
-Then we construct the interval of the largest points from one series and see how many points, from the other series, fall into the intervals.
-The series used to construct the interval is controlled by `temporal_independent` and the window size by `temporal_window` with a default of 5.
-
-In this illustration, we construct the interval based on series `A` and two of the three peaks from `a` falls into this interval at Time 7 and 27.
-
-There's another mandatory argument that hasn't been introduced above: `temporal_var_to_match`.
-This argument controls the variable to match and it needs to appear in both the `major` and `minor` set.
-In the water level matching example, we match the variable `Water_course_level` from `river` to `prcp` from `climate`, hence need to manually rename one of them to match the other, here we rename `Water_course_level` to `prcp` in `river`:
-
-Figure \ref{fig:illu-matching}
-
-```{r illu-matching, out.width="100%", out.height="30%", fig.cap="sdfasdf"}
-knitr::include_graphics("figures/illu-matching.png")
-```
-
-## Interactive graphics
-
-<!-- Interactive graphics can listen to users' actions on the plot to provide additional information that facilitates data exploration.
-This is a useful technique for spatio-temporal data since users can zoom or pan the map to view the local and global structure of the map; use tooltips or popups to query more information about a graphic element; or highlight points to explore its linked views in other plots.
-In the R community, many implementations have been developed to connect \proglang{R} to \proglang{javascript} to create interactive graphics.
-In relation to spatio, temporal, and spatio-temporal data, the general purpose packages \pkg{plotly} [@plotly] and \pkg{leaflet} [@leaflet] realise various interactive actions through their corresponding javascript libraries.
-\pkg{crosstalk} [@crosstalk] and \pkg{tsibbletalk} [@tsibbletalk] implement brushed linking between htmlwidgets.
-\pkg{ggiraph} [@ggiraph] enables tooltip, self-linking, and customised actions specified through its own \proglang{javascript}.
--->
-
-<!-- While many graphic implementations present worked examples to illustrate the usage of the package, few documents the underlying pipeline that transforms the raw data step-by-step to the final view on the screen.
-There have been some early work in building the data pipeline for (interactive) graphics [@buja1988elements; @buja1996interactive;  @sutherland2000orca] and more recent discussions include @wickham2009plumbing, @xie2014reactive, and @cheng2016enabling.
--->
-
-The cubble structure fits in naturally with the interactive graphic pipeline discussed in the literature [@buja1988elements; @buja1996interactive;  @sutherland2000orca; @xie2014reactive; @cheng2016enabling].
-Diagram \ref{fig:illu-interactive} illustrates how linking works with the two forms in a cubble, where a time series plot is created with the long cubble and a map is created with the nested cubble.
-When a user action is captured from the map, the site will be activated in the nested cubble.
-Then, the nested cubble will communicate to the long cubble to activate all the observations with the same `id`.
-The long cubble will then highlight the  activated series in the time series plot.
-
-The linking is also available from the time series plot to the map.
-The selection on the time series is through selecting the point on the time series and once a point is selected, it will be activated in the long cubble.
-All the observations that share the same `id`,  either in the long and nested cubble, are then activated.
-This includes other points in the same time series in the long cubble and the corresponding observation of site in the nested cubble.
-These activated observations will then being reflected in the updated plots and Diagram \ref{fig:illu-interactive-2} in the Appendix illustrates this process.
-
-```{r illu-interactive, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "demon interactivity"}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.004.png"))
-```
-
-\newpage
-
-# Examples
-
-## Australia historical maximum temperature
-
-Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world and the dataset `weatherdata::historical_tmax` extracts the historical maximum temperature recorded for `r nrow(weatherdata::historical_tmax)` Australian stations.
-The data `historical_tmax` is already presented as a cubble, with `id` as the key, `date` as the index, and `c("longitude", "latitude")` as the coordinates.
-Other variables include `elevation`, `name`, `wmo_id`, `first_year`, and  `last_year` in the nested form and `tmax` in the long form.
-This example compares the maximum temperature in two periods: 1971 - 1975 and 2016 - 2020 for stations in Victoria and New South Wales.
-
-Stations in the two states can be subsetted on the station number: Australia GHCN station number starts with `ASN00` and followed by the [Bureau of Meteorology (BOM) station number](http://www.bom.gov.au/climate/cdo/about/site-num.shtml), where the 2nd and 3rd digit (7th and 8th in the GHCN number) denote the state a station belongs to.
-New South Wales stations start from 46 to 75 and Victoria stations then follow from 76 to 90.
-Extracting Victoria and New South Wales stations is a filter operation in the spatial dimension and hence is operated in the nested form:
-
-```{r eval = FALSE, echo = TRUE}
-tmax <- weatherdata::historical_tmax %>%
-  filter(between(stringr::str_sub(id, 7, 8), 46, 90))
-```
-
-The five year window is chosen to remove the effect of a particular year and the historical period of 1971 - 1975 is used since all the stations have records from 1970.
-This following chunk filters on records in the two study periods and summarises the maximum temperature into monthly measure.
-`stretch()` is used to convert the nested form cubble into the long form for these operations:
-
-```{r eval=  FALSE, echo = TRUE}
-tmax <- tmax %>% 
-  stretch() %>%
-  filter(lubridate::year(date) %in% c(1971:1975, 2016:2020)) %>%
-  mutate(month = lubridate::month(date), 
-         group = as.factor(ifelse(lubridate::year(date) > 2015, 
-                                  "2016 ~ 2020", "1971 ~ 1975"))) %>%
-  group_by(month, group) %>%
-  summarise(tmax = mean(tmax, na.rm = TRUE))
-```
-
-A data quality issue with GHCN data is that while the first and last year of each series is provided, years missing in this period is not reported.
-There are a few stations which don't have records during 1971 - 1975 and these stations are filtered out by examining whether the summarised `tmax` has a total of 24 months.
-This is again a station-wise operation and is operated in the nested form, which is switched to from the long form by  `tamp()`:
-
-```{r eval = FALSE, echo = TRUE}
-tmax <- tmax %>% 
-  tamp() %>%
-  filter(nrow(ts) == 24) 
-```
-
-Lastly, to create a glyph map, both the major (`longitude`, `latitude`) and minor (`month`, `tmax`) coordinates need to be in the same table.
-Spatial variables can be moved to the long form with `migrate()`:
-
-```{r eval = FALSE, echo = TRUE}  
-tmax <- tmax %>%   
-  stretch() %>%
-  migrate(latitude, longitude)
-```
-
-Figure \ref{fig:basic-manip} shows the glyph map made with the data after the wangling above.
-One issue with this map is that the similar pattern shown from a few nearby stations around Sydney (151E, 34S) and New castle (152E, 33S) can be distracting.
-Aggregation is a useful technique to observe the general pattern of a collection of series and will be the topic of the next example.
-
-```{r basic-manip, out.width="100%", out.height="70%", fig.cap="Glyph map of the mean maximum temperature by month for Victoria and New South Wales weather stations.
-On the top left corner is a detailed legend for station Cobar highlighted in the black box.
-Compared to 1971 - 1975,  the period 2016 - 2020 sees an increase of mean maximum temperature in spring and summer in Australia (end and beginning of the year).
-A larger increase during the first three months of the year is observed at stations with a lower latitude."}
-knitr::include_graphics("figures/basic-manip.png")
-```
 
@@
 ```
-
-## River level data in Victria water gauges
-
-The water level data comes from [Bureau of Meteorology](http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full) and has a copy in `weatherdata`.
-Here we extract the water course level and add a column annotate this data of type `river`.
-For the rainfall data, we will still use the `weatherdata::climate_full`, filtering for Victorian stations in 2020 should be pretty familiar by now.
-Again, we first look at where these stations are on the map first:
-
-Now we use `match_sites()` to first pair the weather stations with the river gauges spatially and then apply the temporal matching on `prcp`.
-We will construct the interval based on peaks in `climate` since we would expect a lag effect for precipitation to flow into the river and cause a raise in river level, hence `temporal_independent = climate`.
-We select the 30 highest peak from the series to construct the match by setting `temporal_n_highest = 30`.
-This is a tuning parameter and you can start with 10% of the points of one series (here we have daily data for a year, 10% is roughly 30 points).
-`temporal_min_match` filters out pairs don't have enough match and to return all the pairs, set `temporal_min_match` to `0`.``
-
-The output from temporal matching is also a cubble, with additional column `.dist` and `.group` inherent from spatial matching and `n_match` for the number of matched temporal peaks.
-Then you can use this output to plot the location of match or to look at the series:
-
-```{r matching, out.width="100%", out.height="30%", fig.cap="asdfasd"}
-knitr::include_graphics("figures/matching.png")
-```
-
-## ERA5: climate reanalysis data
-
-\newpage
-
-## Interative graphic with cubble
-
-With a larger dataset, interactive graphics can be helpful to explore these data from multiple perspective.
-Two examples are presented here to show how interactivity can be used to explore the climate data across Australia.
-Again, `weatherdata::climate_full` will be used and we hope to include more than one time variable here.
-The difference between the maximum and minimum temperature of the day, the diurnal range, can be directly calculated from the data and it can have large impact on the wine industry (lol this is going far).
-This would be useful to compare the temperature difference across stations and to compare how this temperature difference differs throughout the year.
-
-Temperature difference is a time-wise variable and should be computed in the long form cubble. The variables are averaged over the 2016 to 2020 period by month to stabilise the result.
-
-```{r eval = FALSE, echo = TRUE}
-res <- weatherdata::climate_full %>%
-  # filter(!id %in% c("ASN00067033", "ASN00072091", 
-  #                   "ASN00059040", "ASN00097053")) %>%
-  stretch() %>%
-  mutate(month = lubridate::month(date),
-         diff = tmax - tmin) %>%
-  group_by(month) %>%
-  summarise(tmax = mean(tmax, na.rm = TRUE),
-            diff = mean(diff, na.rm = TRUE)) %>%
-  ungroup(month)
-
-```
-
-Several variables can be useful when comparing the seasonal effect of the temperature differences, for example, the difference of average temperature difference between summer and winter can tell [...].
-This is a station-wise measure and hence computed in the nested form.
-
-```{r eval = FALSE, echo = TRUE}
-res <- res %>% 
-  tamp() %>%
-  mutate(
-    diff_winter = ts %>% filter(month %in% c(6, 7, 8)) %>%
-      pull(diff) %>% mean(na.rm = TRUE),
-    diff_summer = ts %>% filter(month %in% c(12, 1, 2)) %>%
-      pull(diff) %>% mean(na.rm = TRUE),
-    diff_year = diff_winter - diff_summer
-  )
-```
-
-The result from the above manipulation can be directly supplied to make interactive plots and Figure \ref{fig:interactive-linking} is an example made with \pkg{plotly} and \pkg{crosstalk}.
-For example, the second row results from selecting the highest mean maximum temperature in July and this corresponds to Kalumburu where there is a larger winter (June to August) temperature difference than summer (December to February).
-The third row selects one of the stations (Birdsville Airport) with a large average temperature difference in the interior desert region on the map and the linked view from the time series shows that this maximum and minimum temperature difference is constant throughout the year.
-The last row selects the lowest maximum temperature in July on the temperature band and surprisingly, the corresponding station doesn't fall into the Tasmania island but Mountain Hotham in Victoria.
-
-```{r interactive-linking, echo = FALSE, out.width="100%", out.height="25%", fig.cap = "Screenshots of the mean daily temperature difference averaged by month from 2016 to 2020 at various locations in Australia.
-Crossed linking between the map and the temperature band is implemented in both directions for exploring the temperature pattern interactively.
-The first row is the general appearance of the graph and the next three rows are examples from various selection to show different temperature patterns in Australia.
-The html version can be found at ...", fig.show='hold'}
-knitr::include_graphics(here::here("figures/linking.png"))
-knitr::include_graphics(here::here("figures/linking-north.png"))
-knitr::include_graphics(here::here("figures/linking-mid.png"))
-knitr::include_graphics(here::here("figures/linking-south.png"))
-```
-
-Cubble is also compatible with \pkg{leaflet} and with leaflet extension package \pkg{leafpop}, the temperature band can be displayed upon clicking on the station on the map.
-Figure \ref{fig:interactive-popup} shows more examples of temperature pattern that can be found in Australia.
-With the same latitude as Kalumburu but in the Pacific coastline, Cooktown airport shows a similarly high but constant temperature throughout the year.
-With a slightly higher latitude than the mountain Hotham, flat land in Melbourne airport shows a less extreme temperature band.
-Lastly, the Cape Grim Baps, in the north-west tip of the Tasmania island shows a uniformly low and thin temperature band.
-
-```{r interactive-popup, echo = FALSE, out.width="50%", out.height="30%", fig.cap = "Screenshots of the diurnal temperature range averaged by month from 2016 to 2020 at various locations in Australia.
-The leaflet popup allows users to click on a single station for its temperature band as a small plot in the popup window.
-The html version can be found at ...", fig.show='hold'}
-knitr::include_graphics(here::here("figures/popup-north.png"))
-knitr::include_graphics(here::here("figures/popup-vic.png"))
-knitr::include_graphics(here::here("figures/popup-tas.png"))
-```
-
-# Conclusion
-
-# Appendix
-
-```{r illu-interactive-2, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "demon interactivity"}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.005.png"))
-```
-
-\newpage
-
-<!-- # Old stuff -->
-
-<!-- In the temporal aspect, the `tsibble` [@tsibbles] structure and its tidyverts ecosystem have provided a [...
-] workflow to work with temporal data.
-In a tsibble structure, temporal data is characterised by `index` and `key` where `index` is the temporal identifier and `key` is the identifier for multiple series, which could be used as a spatio identifier.
-However, a tsibble object, by construction, always requires the `index` in its structure.
-This makes it less appealing for spatio-temporal data since the output of calculated spatio-specific variables (i.e.
-features of each series) don't have the time dimension.
-Analysts will either need to have an additional step to join this output to the original tsibble or operate with variables stored in two separate objects.
-In addition, the long form structure of a tsibble object means spatio variables (i.e.
-longitude, latitude, and features of each series if joined back to the tsibble) of each spatio identifier will be repetitively recorded at each timestamp.
-This repetition is unnecessary and would inflate the object size for long series.
--->
-
-<!-- # Create a cubble -->
-
-<!-- The creation of a cubble requires the site identifier (`key`), as well as the spatial (`coords`) and temporal (`index`) identifier.
-`climate_flat` is already a tibble and it uses `id` to identify each station, `date` as the time identifier, and `c(long, lat)` as the spatial identifier.
-To create a cubble for this data, use: -->
-
-<!-- ```{r} -->
-<!-- climate_flat %>% as_cubble(key = id, index = date, coords = c(long, lat)) -->
-<!-- ``` -->
-
-<!-- Most of the time, spatio-temporal data doesn't come into this form and analysts need to query the climate variables based on station metadata.
-\textcolor{red}{This is also a problem illustrated in Section 3.5 in @tidydata.
-Here we provide a structured way to query this data based on the row-wise operator and nested list.} For this type of task, one can structure a metadata into a tibble and use row-wise operator to query the climate variables into a nested list.
-As an example here we demonstrate the workflow to find the 5 closest stations to Melbourne.
-We first create a station data frame with the 5 target stations.
--->
-
-<!-- ```{r echo = FALSE} -->
-<!-- latlon_ref <- tibble::tibble(id = "melbourne", lat = -37.8136, long = 144.9631) -->
-
-<!-- cand <- aus_climate %>% -->
-<!--   as_tibble() %>% -->
-<!--   select(id, lat, long, elev, name, wmo_id) -->
-
-<!-- calc_dist <- function(df_ref, df, n = 5) { -->
-<!--   long_ref <- df_ref$long -->
-<!--   lat_ref <- df_ref$lat -->
-
-<!--   df %>% -->
-<!--     mutate( -->
-<!--       long_ref = long_ref, -->
-<!--       lat_ref = lat_ref, -->
-<!--       dist = rnoaa::meteo_spherical_distance(lat, long, lat_ref, long_ref) -->
-<!--     ) %>% -->
-<!--     slice_min(dist, n = n) %>% -->
-<!--     mutate(city = df_ref$id) %>% -->
-<!--     select(-long_ref, -lat_ref) -->
-<!-- } -->
-
-<!-- (stations <- do.call("rbind", purrr::map( -->
-<!--   1:nrow(latlon_ref), -->
-<!--   ~ calc_dist(latlon_ref[.x, ], cand) -->
-<!-- ))) -->
-<!-- ``` -->
-
-<!-- We can query the climate information into a nested list named `ts` for each station with the `rowwise()` operator. To create a cubble, supply the same identifiers as with the first example. -->
-
-<!-- ```{r eval = FALSE} -->
-<!-- sydmel_climate <- stations %>% -->
-<!--   rowwise() %>% -->
-<!--   mutate(ts = list(meteo_pull_monitors(id, -->
-<!--     date_min = "2020-01-01", -->
-<!--     date_max = "2020-12-31", -->
-<!--     var = c("PRCP", "TMAX", "TMIN") -->
-<!--   ) %>% -->
-<!--     select(-id))) %>% -->
-<!--   as_cubble(key = id, index = date, coords = c(long, lat)) -->
-<!-- ``` -->
-
-<!-- ```{r echo = FALSE} -->
-<!-- (sydmel_climate <- stations %>% -->
-<!--   left_join(aus_climate) %>% -->
-<!--   as_cubble(key = id, index = date, coords = c(long, lat))) -->
-<!-- ``` -->
-
-<!-- Below are the how the nested and long form look like for Australia climate data, which records daily precipitation, maximum and minimum temperature for 55 stations across Australia from 2015- 2020.
-Notice that each station forms a group in both forms and specifically, the nested and long form have a underlying `rowwise_df` and `grouped_df` respectively.
--->
-
-<!-- With a cubic framework on mind, different types of manipulation with cubble can be thought of as slicing the cube in various way.
-The table below shows how some `dplyr` verbs are mapped into the operation in a cubble.
-With the existing grouping on the station, additional grouping can be added with `group_by` and removed with `ungrouped`.
-[talk about why it is useful] -->
-
-<!-- \newpage -->
-
-<!-- ## Cubble operations -->
-
-<!-- ```{r cubble-operations, echo = FALSE, fig.cap="Cubble operations", out.height="50%", out.width="90%", fig.align='center'} -->
-<!-- knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.004.png")) -->
-<!-- ``` -->

### 2022-01-07 4-3-water-level.Rmd
play aroudnd with child rmd to organising sections

--- 2021-12-20 main.Rmd
@@
 ---
-documentclass: jss
-author:
-  - name: H. Sherry Zhang
-    # department: Department of Econometrics and Business Statistics
-    affiliation: Monash University
-    #location: Melbourne, Australia
-    email: \email{huize.zhang@monash.edu}
-  - name: Dianne Cook
-    # department: Department of Econometrics and Business Statistics
-    affiliation: 'Monash University \AND'
-    # location: Melbourne, Australia
-    # email:  dicook@monash.edu
-  - name: Ursula Laa
-    # department: Institute of Statistics
-    affiliation: 'University of Natural Resources and Life Sciences \AND'
-    # location: Vienna, Austria
-    # email:  ursula.laa@boku.ac.at
-  - name: Nicolas Langrené
-    # department: 34 Village Street, Docklands VIC 3008 Australia
-    affiliation: 'CSIRO Data61 \AND'
-    # location: Melbourne, Australia
-    # email: nicolas.langrene@csiro.au
-  - name: Patricia Menéndez
-    # department: Department of Econometrics and Business Statistics
-    affiliation: 'Monash University \AND'
-    # location: Melbourne, Australia
-    # email:  patricia.menendez@monash.edu
-title:
-  formatted: "\\pkg{cubble}: An R Package for Structuring Spatio-temporal Data"
-  # If you use tex in the formatted title, alsoge supply version without
-  plain:     "cubble: An R Package for Structuring Spatio-temporal Data"
-  # For running headers, if needed
-  # short:     "\\pkg{foo}: A Capitalized Title"
-abstract: >
-  The abstract of the article.
-keywords:
-  # at least one keyword must be supplied
-  formatted: [spatio-temporal data,  "\\proglang{R}"]
-  plain:     [spatio-temporal data, R]
-preamble: >
-  \usepackage{amsmath}
-  \usepackage{array}
-output: rticles::jss_article
-bibliography: references.bib
+title: "4-3-water-level"
+output: html_document
 ---
-
-```{r setup, echo = FALSE}
-knitr::opts_chunk$set(
-  echo = FALSE,
-  warning = FALSE,
-  message = FALSE,
-  cache = TRUE,
-  fig.path = "figures/",
-  fig.height = 10
-)
-options(prompt = "R> ", continue = "+ ", 
-        tibble.print_max = 5, tibble.print_min = 5)
-```
-
-```{r echo = FALSE}
-library(cubble)
-library(forcats)
-library(stringr)
-library(dplyr)
-library(lubridate)
-library(ggplot2)
-library(GGally)
-library(plotly)
-library(crosstalk)
-library(patchwork)
-```
-
-\newpage
-
-# Introduction
-
-Many data structures have been proposed for spatial (\pkg{sf} by @sf) and temporal (\pkg{tsibble} by @tsibble) data in the R community, while less has been done for spatio-temporal data.
-The lack of such tools could potentially because analysts usually treat the spatial and temporal dimension of the data separately, without realising the need to create a new data structure.
-While this approach follows the third tidy data principal [@tidydata] (*Each type of observational unit forms a table*), analysts always need to manually join results from different observational units or combining multiple tables into one for downstream analysis.
-This additional step doesn't add new operations into the data but can be error prone.
-\newline
-
-Currently, available spatio-temporal data structure in R includes: \pkg{spacetime} [@spacetime], which proposes four space-time layouts: Full grid (STF), sparse grid(STS), irregular (STI), and trajectory (STT).
-The data structure it uses is based on \pkg{sp} [@sp] and \pkg{xts} [@xts], both of which has been replaced by more recent implementations.
-\pkg{spatstat} [@spatstat] implements a `ppp` class for point pattern data; and more recent, \pkg{stars} [@stars] implements a spatio-temporal array with the dplyr's data cube structure \pkg{cubelyr} [@cubelyr] as its backend.
-While these implementations either store spatial and temporal variables all in a single table, hence duplicate the spatial variables for each temporal unit; or split them into two separate tables that has the problem of manually joining, mentioned in the previously.
-None of these packages enjoy both the benefits of being able to separate manipulation in the two dimensions while also keep the data object as a whole.
-This create a gap in the software development.
-The requirement for such a tool is important given the ubiquity of spatio-temporal vector data in the wild: the Ireland wind data from \pkg{gstat} is an classic example data that splits variables into spatial (`wind.loc`) and temporal (`wind`) dimension; Bureau of Meteorology (BoM) provides climate observations that are widely applied in agriculture and ecology study; air pollution data.
-\newline
-
-This paper describes the implementation of a new spatio-temporal data structure: \pkg{cubble}.
-\pkg{cubble} implements a relational data structure that uses two forms to manage the switch between spatial and temporal dimension.
-With this structure, users can manipulate the spatial or temporal dimension separately, while leaves the linking of two dimensions to \pkg{cubble}.
-The software is available from the Comprehensive R Archive Network (CRAN) at [CRAN link].
-\newline
-
-The rest of the paper will be divided as follows: [complete when the paper structure is more solid]
-
-<!-- Section 2 reviews the existing data structure for spatio, temporal, and spatio-temporal data.
-Section 3 presents a new data structure for spatio-temporal data: cubble.
-Then the paper introduces the workflow of data manipulation and visualisation with the cubble structure in Section 4.
-Section 5 gives some examples on how common spatial and temporal manipulations are performed with cubble and how static and interactive visualisation help to understand climate and [...] data.
-\newline -->
-
-\newpage
-
-# The cubble package
-
-Spatio-temporal data usually come in various forms and Figure \ref{fig:illu-input} shows four examples of this.
-No matter which form the data is in, these formats share some common components that characterise spatio-temporal data.
-A spatial identifier (`id` in the diagram) is the unique identifier of each site.
-The temporal identifier (`t` in the diagram) prescribes the time stamp each site is recorded.
-Coordinates, comprising of latitude and longitude (`lon` and `lat` in the diagram), locates each site on the map.
-These identifiers will be the building blocks for the data structure introduced below.
-Other variables in the data can be categorised into two groups: spatial variables that are invariant at each time stamp for every site, i.e.
-the name or code of the weather station and temporal variables that varies with time.
-
-```{r illu-input, echo = FALSE, out.width = "100%", fig.cap="Illustration of incoming data formats for spatio-temporal data.
-(1) Data comes in as a single table; (2) Separate tables for spatial and temporal variables; (3) A single table with all the parameters used to query the database and a separate table for queried data; and (4) Cubical data in array or NetCDF format.", fig.align='center', out.height = "15%"}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.001.png"))
-```
-
-In a cubble, there are two forms: nested form and long form, and Figure \ref{fig:illu-cubble} sketches the two forms along with the associated attributes.
-The decision on which form to use is output-oriented, meaning analysts need to first think about whether the output of a particular operation is identified only by the spatial identifier, or a combination of spatial and temporal identifier.
-The nested cubble is suitable for working with operations that are only identified by site and this type of operation can be a pure manipulation of spatial variables, or a summary of temporal variables by site (i.e.
-the output of counting the number of raining day is only identified by sites and hence should be performed with the nested form).
-Underneath the nested form, a cubble is built from a row-wise dataframe (`rowwise_df`) where each site forms a separate group.
-This structure simplifies the calculation that involves temporal variables by avoiding the use of `map` syntax when working with list-column.
-
-For those operations whose output involves both a spatial and temporal dimension, long form should be used.
-The long form is identified by both the spatial and temporal identifier and adopts a grouped dataframe (`grouped_df`) to forms each site as a group.
-Spatial variables are stored separately in a \pkg{tibble} as an special attribute of the long cubble.
-This design avoids repeating the spatial variables at each time stamp while not dropping information from spatial variables.
-
-```{r illu-cubble, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "Illustration of nested and long cubble."}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.002.png"))
-```
-
-## Create a cubble in the nested form
-
-To use functionalities from cubble, data analysts first need to create a cubble.
-`as_cubble` create a \pkg{cubble} by supplying the three key components: `key` as the spatial identifier; `index` as the temporal identifier; and a vector of `coords` in the order of longitude first and then latitude.
-The naming of `key` and `index` follows the convention in the \pkg{tsibble} package.
-The cubble created by default is in the nested form.
-Below is an example of creating a cubble: \newline
-
-```{r}
-(cubble_nested <- cubble::climate_flat %>%
-  as_cubble(key = id, index = date, coords = c("long", "lat")))
-```
-
-There are a few information in the \pkg{cubble} header: the name of the `key` variable, bbox, and also the name of variable nested in the `ts` column.
-In this example, each site is identifier is `id` and the number in the bracket means there are 5 unique `id` in this dataset.
-The bbox in the second row gives the range of the coordinates.
-The temporal variables are all nested in the `ts` column, but it could be useful to know the name these variables.
-The third row in the cubble header shows these names and in this example this includes: precipitation, `prcp`, maximum temperature, `tmax`, and minimum temperature, `tmin`.
-
-## Stretch a nested cubble into the long form
-
-The long cubble is suitable to manipulate the time dimension of the data.
-The function `stretch()` switches the nested cubble into the long cubble by first extracts all the spatial variables into a separate tibble and store in the `spatial` attribute and then unnests the `ts` column:
-
-```{r}
-(cubble_long <- cubble_nested %>% stretch(ts))
-```
-
-Notice here that the third line in the header now shows the name of spatial variables rather than the temporal variables.
-
-## Tamp a long cubble back to the nested form
-
-Manipulation on the spatial and temporal dimension can be an iterative process.
-Many times, analysts will need to go back and forth between the nested and long cubble.
-The `stretch()` function introduced in the previous section switches a nested cubble into a long cubble and function `tamp()` is its inverse function to switch a long cubble back to the nested cubble:
-
-```{r}
-(cubble_back <- cubble_long %>% tamp())
-```
-
-## Migrate spatial variables to a long cubble
-
-As a final data output for modelling or visualisation, spatio-temporal data is usually expected to be in a single table.
-Function `migrate()` moves the spatial variables from the `spatial` attribute into the long cubble:
-
-```{r}
-(cubble_long %>% migrate(long, lat))
-```
-
-In this workflow described above, data objects come into cubble in the nested form, then various operations on the spatial and temporal dimension can go back and forth between the nested and long form, and finally, the data will come out of cubble in the long form for further modelling or visualisation.
-
-Building from an underlying `tbl_df` structure, it is natural to implement methods available in `dplyr` to `cubble`. Supported methods in the `cubble` with `dplyr` generics includes:
-
-```{=tex}
-\begin{center}
-\begin{tabular}{ | m{5em} | m{15cm}| } 
-basics & \textbf{mutate}, \textbf{filter}, \textbf{summarise}, \textbf{select}, \textbf{arrange}, \textbf{rename}, \textbf{left\_join} \\
-grouping &  \textbf{group\_by}, \textbf{ungroup}\\
-slice family & \textbf{slice\_head}, \textbf{slice\_tail}, \textbf{slice\_sample}, \textbf{slice\_min} and \textbf{slice\_max} \\
-\end{tabular}
-\end{center}
-```
-\pkg{cubble} is also compatible with \pkg{tsibble} in the sense that the original list-column can be a `tbl_ts` object.
-Duplicates and gaps should be first checked before structuring the data into a cubble.
-If the input data is a \pkg{tsibble} object, the long form cubble is also a \pkg{tsibble} where users can directly apply time series operations.
-
-# Advanced features/ considerations
-
-## Hierarchical structure
-
-Imposing a clustering structure can be thought of as building a hierarchical structure where stations are nested within clusters.
-This can be useful when there's intrinsic nesting structure in the data (i.e.
-country nested in the continent, county nested in the state) or there's some clustering.
-When we have access to both level, `switch_key()` is the function to re-structure the data as one cluster per row.
-Temporal observations from different stations while within the same cluster are bound in the nested column `ts`.
-
-One thing we hope to do with the cluster is to find the coordinates of the centroid.
-These are variables variant to the station but invariant to the cluster and it would be nice to have a function that structure each cluster as a row.
-`switch_key()` is the function that does this: it lets you to specify a new key, say `cluster` and nests all spatial variables variant to `cluster` into a column.
-Spatial variables are all nested inside a new column `.val`.
-Temporal observations from different stations while within the same cluster are bound in the nested column `ts`.
-
-This structure makes it easy to compute cluster level variable, for example, the convex hull and the centroid coordinate of each cluster.
-These can be amended into the nested form with function `get_centroid`.
-
-After we have got `cluster_nested`, spatial and temporal data at both levels can be easily obtained.
-Figure \ref{fig:illu-hier} illustrate the relationship between the long and nested form cubble at both site and cluster level.
-More description on this.
-Start with the original `station_nested`, `stretch()` expands the `ts` column with each station (`id`) forming a group and attach variables invariant to `id` as an attribute.
-`switch_key()` changes the `key` from `id` to `cluster` and nests all the spatial variables that variant to `cluster`.
-`stretch()` `cluster_nested` will store variables that are invariant to `cluster` as a tibble in the attribute.
-
-this fit into the remaining data pipeline.
-
-```{r illu-hier, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "Hierarchical structure"}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.003.png"))
-```
-
-## Data fusion and matching
-
-Temporal matching checks how spatially matched pairs align temporally. We use the following chart to illustrate how the temporal matching works:
-
-For each spatially matched pair, say `A` and `a`, we first find the largest `n` points in each series, colored in brown points here.
-Here we use the largest three but you can tune this number by `temporal_n_highest`.
-Then we construct the interval of the largest points from one series and see how many points, from the other series, fall into the intervals.
-The series used to construct the interval is controlled by `temporal_independent` and the window size by `temporal_window` with a default of 5.
-
-In this illustration, we construct the interval based on series `A` and two of the three peaks from `a` falls into this interval at Time 7 and 27.
-
-There's another mandatory argument that hasn't been introduced above: `temporal_var_to_match`.
-This argument controls the variable to match and it needs to appear in both the `major` and `minor` set.
-In the water level matching example, we match the variable `Water_course_level` from `river` to `prcp` from `climate`, hence need to manually rename one of them to match the other, here we rename `Water_course_level` to `prcp` in `river`:
-
-Figure \ref{fig:illu-matching}
-
-```{r illu-matching, out.width="100%", out.height="30%", fig.cap="sdfasdf"}
-knitr::include_graphics("figures/illu-matching.png")
-```
-
-## Interactive graphics
-
-<!-- Interactive graphics can listen to users' actions on the plot to provide additional information that facilitates data exploration.
-This is a useful technique for spatio-temporal data since users can zoom or pan the map to view the local and global structure of the map; use tooltips or popups to query more information about a graphic element; or highlight points to explore its linked views in other plots.
-In the R community, many implementations have been developed to connect \proglang{R} to \proglang{javascript} to create interactive graphics.
-In relation to spatio, temporal, and spatio-temporal data, the general purpose packages \pkg{plotly} [@plotly] and \pkg{leaflet} [@leaflet] realise various interactive actions through their corresponding javascript libraries.
-\pkg{crosstalk} [@crosstalk] and \pkg{tsibbletalk} [@tsibbletalk] implement brushed linking between htmlwidgets.
-\pkg{ggiraph} [@ggiraph] enables tooltip, self-linking, and customised actions specified through its own \proglang{javascript}.
--->
-
-<!-- While many graphic implementations present worked examples to illustrate the usage of the package, few documents the underlying pipeline that transforms the raw data step-by-step to the final view on the screen.
-There have been some early work in building the data pipeline for (interactive) graphics [@buja1988elements; @buja1996interactive;  @sutherland2000orca] and more recent discussions include @wickham2009plumbing, @xie2014reactive, and @cheng2016enabling.
--->
-
-The cubble structure fits in naturally with the interactive graphic pipeline discussed in the literature [@buja1988elements; @buja1996interactive;  @sutherland2000orca; @xie2014reactive; @cheng2016enabling].
-Diagram \ref{fig:illu-interactive} illustrates how linking works with the two forms in a cubble, where a time series plot is created with the long cubble and a map is created with the nested cubble.
-When a user action is captured from the map, the site will be activated in the nested cubble.
-Then, the nested cubble will communicate to the long cubble to activate all the observations with the same `id`.
-The long cubble will then highlight the  activated series in the time series plot.
-
-The linking is also available from the time series plot to the map.
-The selection on the time series is through selecting the point on the time series and once a point is selected, it will be activated in the long cubble.
-All the observations that share the same `id`,  either in the long and nested cubble, are then activated.
-This includes other points in the same time series in the long cubble and the corresponding observation of site in the nested cubble.
-These activated observations will then being reflected in the updated plots and Diagram \ref{fig:illu-interactive-2} in the Appendix illustrates this process.
-
-```{r illu-interactive, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "demon interactivity"}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.004.png"))
-```
-
-\newpage
-
-# Examples
-
-## Australia historical maximum temperature
-
-Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world and the dataset `weatherdata::historical_tmax` extracts the historical maximum temperature recorded for `r nrow(weatherdata::historical_tmax)` Australian stations.
-The data `historical_tmax` is already presented as a cubble, with `id` as the key, `date` as the index, and `c("longitude", "latitude")` as the coordinates.
-Other variables include `elevation`, `name`, `wmo_id`, `first_year`, and  `last_year` in the nested form and `tmax` in the long form.
-This example compares the maximum temperature in two periods: 1971 - 1975 and 2016 - 2020 for stations in Victoria and New South Wales.
-
-Stations in the two states can be subsetted on the station number: Australia GHCN station number starts with `ASN00` and followed by the [Bureau of Meteorology (BOM) station number](http://www.bom.gov.au/climate/cdo/about/site-num.shtml), where the 2nd and 3rd digit (7th and 8th in the GHCN number) denote the state a station belongs to.
-New South Wales stations start from 46 to 75 and Victoria stations then follow from 76 to 90.
-Extracting Victoria and New South Wales stations is a filter operation in the spatial dimension and hence is operated in the nested form:
-
-```{r eval = FALSE, echo = TRUE}
-tmax <- weatherdata::historical_tmax %>%
-  filter(between(stringr::str_sub(id, 7, 8), 46, 90))
-```
-
-The five year window is chosen to remove the effect of a particular year and the historical period of 1971 - 1975 is used since all the stations have records from 1970.
-This following chunk filters on records in the two study periods and summarises the maximum temperature into monthly measure.
-`stretch()` is used to convert the nested form cubble into the long form for these operations:
-
-```{r eval=  FALSE, echo = TRUE}
-tmax <- tmax %>% 
-  stretch() %>%
-  filter(lubridate::year(date) %in% c(1971:1975, 2016:2020)) %>%
-  mutate(month = lubridate::month(date), 
-         group = as.factor(ifelse(lubridate::year(date) > 2015, 
-                                  "2016 ~ 2020", "1971 ~ 1975"))) %>%
-  group_by(month, group) %>%
-  summarise(tmax = mean(tmax, na.rm = TRUE))
-```
-
-A data quality issue with GHCN data is that while the first and last year of each series is provided, years missing in this period is not reported.
-There are a few stations which don't have records during 1971 - 1975 and these stations are filtered out by examining whether the summarised `tmax` has a total of 24 months.
-This is again a station-wise operation and is operated in the nested form, which is switched to from the long form by  `tamp()`:
-
-```{r eval = FALSE, echo = TRUE}
-tmax <- tmax %>% 
-  tamp() %>%
-  filter(nrow(ts) == 24) 
-```
-
-Lastly, to create a glyph map, both the major (`longitude`, `latitude`) and minor (`month`, `tmax`) coordinates need to be in the same table.
-Spatial variables can be moved to the long form with `migrate()`:
-
-```{r eval = FALSE, echo = TRUE}  
-tmax <- tmax %>%   
-  stretch() %>%
-  migrate(latitude, longitude)
-```
-
-Figure \ref{fig:basic-manip} shows the glyph map made with the data after the wangling above.
-One issue with this map is that the similar pattern shown from a few nearby stations around Sydney (151E, 34S) and New castle (152E, 33S) can be distracting.
-Aggregation is a useful technique to observe the general pattern of a collection of series and will be the topic of the next example.
-
-```{r basic-manip, out.width="100%", out.height="70%", fig.cap="Glyph map of the mean maximum temperature by month for Victoria and New South Wales weather stations.
-On the top left corner is a detailed legend for station Cobar highlighted in the black box.
-Compared to 1971 - 1975,  the period 2016 - 2020 sees an increase of mean maximum temperature in spring and summer in Australia (end and beginning of the year).
-A larger increase during the first three months of the year is observed at stations with a lower latitude."}
-knitr::include_graphics("figures/basic-manip.png")
-```
-
-## Australia precipitation pattern in 2020
-
-In Figure \ref{fig:basic-agg}.
-Aggregating them into averages or other statistics would give a better picture.
-This can be done with cubble through a hierarchical structure, where the cluster indicator is one level above the individual station.
-
-The `weatherdata::climate_full` data has daily climate data of `r nrow(weatherdata::climate_full)` Australia stations from 2016 to 2020.
-
-calcluate distance matrix
-As an example to illustrate here, a kmean clustering algorithm based on the distance matrix is used and the number of centres is set to 20.
-More complex algorithms can also be used for more complex problem, as long as a mapping from each station id to the cluster id can be constructed.
-
-<!-- Now we can obtain the aggregated series as described in the workflow diagram above and construct the glyph map with `GGally::glyphs()`: -->
-
-<!-- We can also look at the precipitation of each individual station within the same cluster: -->
-
-<!-- Lastly, there is one series on Tasmania island standing out from others, lets look at where it is: -->
-
-```{r basic-agg, out.width="100%", out.height="90%", fig.cap="sdfasdf"}
-knitr::include_graphics("figures/basic-agg.png")
-```
 
@@
 ```
-
-## ERA5: climate reanalysis data
-
-\newpage
-
-## Interative graphic with cubble
-
-With a larger dataset, interactive graphics can be helpful to explore these data from multiple perspective.
-Two examples are presented here to show how interactivity can be used to explore the climate data across Australia.
-Again, `weatherdata::climate_full` will be used and we hope to include more than one time variable here.
-The difference between the maximum and minimum temperature of the day, the diurnal range, can be directly calculated from the data and it can have large impact on the wine industry (lol this is going far).
-This would be useful to compare the temperature difference across stations and to compare how this temperature difference differs throughout the year.
-
-Temperature difference is a time-wise variable and should be computed in the long form cubble. The variables are averaged over the 2016 to 2020 period by month to stabilise the result.
-
-```{r eval = FALSE, echo = TRUE}
-res <- weatherdata::climate_full %>%
-  # filter(!id %in% c("ASN00067033", "ASN00072091", 
-  #                   "ASN00059040", "ASN00097053")) %>%
-  stretch() %>%
-  mutate(month = lubridate::month(date),
-         diff = tmax - tmin) %>%
-  group_by(month) %>%
-  summarise(tmax = mean(tmax, na.rm = TRUE),
-            diff = mean(diff, na.rm = TRUE)) %>%
-  ungroup(month)
-
-```
-
-Several variables can be useful when comparing the seasonal effect of the temperature differences, for example, the difference of average temperature difference between summer and winter can tell [...].
-This is a station-wise measure and hence computed in the nested form.
-
-```{r eval = FALSE, echo = TRUE}
-res <- res %>% 
-  tamp() %>%
-  mutate(
-    diff_winter = ts %>% filter(month %in% c(6, 7, 8)) %>%
-      pull(diff) %>% mean(na.rm = TRUE),
-    diff_summer = ts %>% filter(month %in% c(12, 1, 2)) %>%
-      pull(diff) %>% mean(na.rm = TRUE),
-    diff_year = diff_winter - diff_summer
-  )
-```
-
-The result from the above manipulation can be directly supplied to make interactive plots and Figure \ref{fig:interactive-linking} is an example made with \pkg{plotly} and \pkg{crosstalk}.
-For example, the second row results from selecting the highest mean maximum temperature in July and this corresponds to Kalumburu where there is a larger winter (June to August) temperature difference than summer (December to February).
-The third row selects one of the stations (Birdsville Airport) with a large average temperature difference in the interior desert region on the map and the linked view from the time series shows that this maximum and minimum temperature difference is constant throughout the year.
-The last row selects the lowest maximum temperature in July on the temperature band and surprisingly, the corresponding station doesn't fall into the Tasmania island but Mountain Hotham in Victoria.
-
-```{r interactive-linking, echo = FALSE, out.width="100%", out.height="25%", fig.cap = "Screenshots of the mean daily temperature difference averaged by month from 2016 to 2020 at various locations in Australia.
-Crossed linking between the map and the temperature band is implemented in both directions for exploring the temperature pattern interactively.
-The first row is the general appearance of the graph and the next three rows are examples from various selection to show different temperature patterns in Australia.
-The html version can be found at ...", fig.show='hold'}
-knitr::include_graphics(here::here("figures/linking.png"))
-knitr::include_graphics(here::here("figures/linking-north.png"))
-knitr::include_graphics(here::here("figures/linking-mid.png"))
-knitr::include_graphics(here::here("figures/linking-south.png"))
-```
-
-Cubble is also compatible with \pkg{leaflet} and with leaflet extension package \pkg{leafpop}, the temperature band can be displayed upon clicking on the station on the map.
-Figure \ref{fig:interactive-popup} shows more examples of temperature pattern that can be found in Australia.
-With the same latitude as Kalumburu but in the Pacific coastline, Cooktown airport shows a similarly high but constant temperature throughout the year.
-With a slightly higher latitude than the mountain Hotham, flat land in Melbourne airport shows a less extreme temperature band.
-Lastly, the Cape Grim Baps, in the north-west tip of the Tasmania island shows a uniformly low and thin temperature band.
-
-```{r interactive-popup, echo = FALSE, out.width="50%", out.height="30%", fig.cap = "Screenshots of the diurnal temperature range averaged by month from 2016 to 2020 at various locations in Australia.
-The leaflet popup allows users to click on a single station for its temperature band as a small plot in the popup window.
-The html version can be found at ...", fig.show='hold'}
-knitr::include_graphics(here::here("figures/popup-north.png"))
-knitr::include_graphics(here::here("figures/popup-vic.png"))
-knitr::include_graphics(here::here("figures/popup-tas.png"))
-```
-
-# Conclusion
-
-# Appendix
-
-```{r illu-interactive-2, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "demon interactivity"}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.005.png"))
-```
-
-\newpage
-
-<!-- # Old stuff -->
-
-<!-- In the temporal aspect, the `tsibble` [@tsibbles] structure and its tidyverts ecosystem have provided a [...
-] workflow to work with temporal data.
-In a tsibble structure, temporal data is characterised by `index` and `key` where `index` is the temporal identifier and `key` is the identifier for multiple series, which could be used as a spatio identifier.
-However, a tsibble object, by construction, always requires the `index` in its structure.
-This makes it less appealing for spatio-temporal data since the output of calculated spatio-specific variables (i.e.
-features of each series) don't have the time dimension.
-Analysts will either need to have an additional step to join this output to the original tsibble or operate with variables stored in two separate objects.
-In addition, the long form structure of a tsibble object means spatio variables (i.e.
-longitude, latitude, and features of each series if joined back to the tsibble) of each spatio identifier will be repetitively recorded at each timestamp.
-This repetition is unnecessary and would inflate the object size for long series.
--->
-
-<!-- # Create a cubble -->
-
-<!-- The creation of a cubble requires the site identifier (`key`), as well as the spatial (`coords`) and temporal (`index`) identifier.
-`climate_flat` is already a tibble and it uses `id` to identify each station, `date` as the time identifier, and `c(long, lat)` as the spatial identifier.
-To create a cubble for this data, use: -->
-
-<!-- ```{r} -->
-<!-- climate_flat %>% as_cubble(key = id, index = date, coords = c(long, lat)) -->
-<!-- ``` -->
-
-<!-- Most of the time, spatio-temporal data doesn't come into this form and analysts need to query the climate variables based on station metadata.
-\textcolor{red}{This is also a problem illustrated in Section 3.5 in @tidydata.
-Here we provide a structured way to query this data based on the row-wise operator and nested list.} For this type of task, one can structure a metadata into a tibble and use row-wise operator to query the climate variables into a nested list.
-As an example here we demonstrate the workflow to find the 5 closest stations to Melbourne.
-We first create a station data frame with the 5 target stations.
--->
-
-<!-- ```{r echo = FALSE} -->
-<!-- latlon_ref <- tibble::tibble(id = "melbourne", lat = -37.8136, long = 144.9631) -->
-
-<!-- cand <- aus_climate %>% -->
-<!--   as_tibble() %>% -->
-<!--   select(id, lat, long, elev, name, wmo_id) -->
-
-<!-- calc_dist <- function(df_ref, df, n = 5) { -->
-<!--   long_ref <- df_ref$long -->
-<!--   lat_ref <- df_ref$lat -->
-
-<!--   df %>% -->
-<!--     mutate( -->
-<!--       long_ref = long_ref, -->
-<!--       lat_ref = lat_ref, -->
-<!--       dist = rnoaa::meteo_spherical_distance(lat, long, lat_ref, long_ref) -->
-<!--     ) %>% -->
-<!--     slice_min(dist, n = n) %>% -->
-<!--     mutate(city = df_ref$id) %>% -->
-<!--     select(-long_ref, -lat_ref) -->
-<!-- } -->
-
-<!-- (stations <- do.call("rbind", purrr::map( -->
-<!--   1:nrow(latlon_ref), -->
-<!--   ~ calc_dist(latlon_ref[.x, ], cand) -->
-<!-- ))) -->
-<!-- ``` -->
-
-<!-- We can query the climate information into a nested list named `ts` for each station with the `rowwise()` operator. To create a cubble, supply the same identifiers as with the first example. -->
-
-<!-- ```{r eval = FALSE} -->
-<!-- sydmel_climate <- stations %>% -->
-<!--   rowwise() %>% -->
-<!--   mutate(ts = list(meteo_pull_monitors(id, -->
-<!--     date_min = "2020-01-01", -->
-<!--     date_max = "2020-12-31", -->
-<!--     var = c("PRCP", "TMAX", "TMIN") -->
-<!--   ) %>% -->
-<!--     select(-id))) %>% -->
-<!--   as_cubble(key = id, index = date, coords = c(long, lat)) -->
-<!-- ``` -->
-
-<!-- ```{r echo = FALSE} -->
-<!-- (sydmel_climate <- stations %>% -->
-<!--   left_join(aus_climate) %>% -->
-<!--   as_cubble(key = id, index = date, coords = c(long, lat))) -->
-<!-- ``` -->
-
-<!-- Below are the how the nested and long form look like for Australia climate data, which records daily precipitation, maximum and minimum temperature for 55 stations across Australia from 2015- 2020.
-Notice that each station forms a group in both forms and specifically, the nested and long form have a underlying `rowwise_df` and `grouped_df` respectively.
--->
-
-<!-- With a cubic framework on mind, different types of manipulation with cubble can be thought of as slicing the cube in various way.
-The table below shows how some `dplyr` verbs are mapped into the operation in a cubble.
-With the existing grouping on the station, additional grouping can be added with `group_by` and removed with `ungrouped`.
-[talk about why it is useful] -->
-
-<!-- \newpage -->
-
-<!-- ## Cubble operations -->
-
-<!-- ```{r cubble-operations, echo = FALSE, fig.cap="Cubble operations", out.height="50%", out.width="90%", fig.align='center'} -->
-<!-- knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.004.png")) -->
-<!-- ``` -->

### 2022-01-07 4-4-netcdf.Rmd
play aroudnd with child rmd to organising sections

--- 2021-12-20 main.Rmd
@@
 ---
-documentclass: jss
-author:
-  - name: H. Sherry Zhang
-    # department: Department of Econometrics and Business Statistics
-    affiliation: Monash University
-    #location: Melbourne, Australia
-    email: \email{huize.zhang@monash.edu}
-  - name: Dianne Cook
-    # department: Department of Econometrics and Business Statistics
-    affiliation: 'Monash University \AND'
-    # location: Melbourne, Australia
-    # email:  dicook@monash.edu
-  - name: Ursula Laa
-    # department: Institute of Statistics
-    affiliation: 'University of Natural Resources and Life Sciences \AND'
-    # location: Vienna, Austria
-    # email:  ursula.laa@boku.ac.at
-  - name: Nicolas Langrené
-    # department: 34 Village Street, Docklands VIC 3008 Australia
-    affiliation: 'CSIRO Data61 \AND'
-    # location: Melbourne, Australia
-    # email: nicolas.langrene@csiro.au
-  - name: Patricia Menéndez
-    # department: Department of Econometrics and Business Statistics
-    affiliation: 'Monash University \AND'
-    # location: Melbourne, Australia
-    # email:  patricia.menendez@monash.edu
-title:
-  formatted: "\\pkg{cubble}: An R Package for Structuring Spatio-temporal Data"
-  # If you use tex in the formatted title, alsoge supply version without
-  plain:     "cubble: An R Package for Structuring Spatio-temporal Data"
-  # For running headers, if needed
-  # short:     "\\pkg{foo}: A Capitalized Title"
-abstract: >
-  The abstract of the article.
-keywords:
-  # at least one keyword must be supplied
-  formatted: [spatio-temporal data,  "\\proglang{R}"]
-  plain:     [spatio-temporal data, R]
-preamble: >
-  \usepackage{amsmath}
-  \usepackage{array}
-output: rticles::jss_article
-bibliography: references.bib
+title: "4-4-netcdf"
+output: pdf_document
 ---
-
-```{r setup, echo = FALSE}
-knitr::opts_chunk$set(
-  echo = FALSE,
-  warning = FALSE,
-  message = FALSE,
-  cache = TRUE,
-  fig.path = "figures/",
-  fig.height = 10
-)
-options(prompt = "R> ", continue = "+ ", 
-        tibble.print_max = 5, tibble.print_min = 5)
-```
-
-```{r echo = FALSE}
-library(cubble)
-library(forcats)
-library(stringr)
-library(dplyr)
-library(lubridate)
-library(ggplot2)
-library(GGally)
-library(plotly)
-library(crosstalk)
-library(patchwork)
-```
-
-\newpage
-
-# Introduction
-
-Many data structures have been proposed for spatial (\pkg{sf} by @sf) and temporal (\pkg{tsibble} by @tsibble) data in the R community, while less has been done for spatio-temporal data.
-The lack of such tools could potentially because analysts usually treat the spatial and temporal dimension of the data separately, without realising the need to create a new data structure.
-While this approach follows the third tidy data principal [@tidydata] (*Each type of observational unit forms a table*), analysts always need to manually join results from different observational units or combining multiple tables into one for downstream analysis.
-This additional step doesn't add new operations into the data but can be error prone.
-\newline
-
-Currently, available spatio-temporal data structure in R includes: \pkg{spacetime} [@spacetime], which proposes four space-time layouts: Full grid (STF), sparse grid(STS), irregular (STI), and trajectory (STT).
-The data structure it uses is based on \pkg{sp} [@sp] and \pkg{xts} [@xts], both of which has been replaced by more recent implementations.
-\pkg{spatstat} [@spatstat] implements a `ppp` class for point pattern data; and more recent, \pkg{stars} [@stars] implements a spatio-temporal array with the dplyr's data cube structure \pkg{cubelyr} [@cubelyr] as its backend.
-While these implementations either store spatial and temporal variables all in a single table, hence duplicate the spatial variables for each temporal unit; or split them into two separate tables that has the problem of manually joining, mentioned in the previously.
-None of these packages enjoy both the benefits of being able to separate manipulation in the two dimensions while also keep the data object as a whole.
-This create a gap in the software development.
-The requirement for such a tool is important given the ubiquity of spatio-temporal vector data in the wild: the Ireland wind data from \pkg{gstat} is an classic example data that splits variables into spatial (`wind.loc`) and temporal (`wind`) dimension; Bureau of Meteorology (BoM) provides climate observations that are widely applied in agriculture and ecology study; air pollution data.
-\newline
-
-This paper describes the implementation of a new spatio-temporal data structure: \pkg{cubble}.
-\pkg{cubble} implements a relational data structure that uses two forms to manage the switch between spatial and temporal dimension.
-With this structure, users can manipulate the spatial or temporal dimension separately, while leaves the linking of two dimensions to \pkg{cubble}.
-The software is available from the Comprehensive R Archive Network (CRAN) at [CRAN link].
-\newline
-
-The rest of the paper will be divided as follows: [complete when the paper structure is more solid]
-
-<!-- Section 2 reviews the existing data structure for spatio, temporal, and spatio-temporal data.
-Section 3 presents a new data structure for spatio-temporal data: cubble.
-Then the paper introduces the workflow of data manipulation and visualisation with the cubble structure in Section 4.
-Section 5 gives some examples on how common spatial and temporal manipulations are performed with cubble and how static and interactive visualisation help to understand climate and [...] data.
-\newline -->
-
-\newpage
-
-# The cubble package
-
-Spatio-temporal data usually come in various forms and Figure \ref{fig:illu-input} shows four examples of this.
-No matter which form the data is in, these formats share some common components that characterise spatio-temporal data.
-A spatial identifier (`id` in the diagram) is the unique identifier of each site.
-The temporal identifier (`t` in the diagram) prescribes the time stamp each site is recorded.
-Coordinates, comprising of latitude and longitude (`lon` and `lat` in the diagram), locates each site on the map.
-These identifiers will be the building blocks for the data structure introduced below.
-Other variables in the data can be categorised into two groups: spatial variables that are invariant at each time stamp for every site, i.e.
-the name or code of the weather station and temporal variables that varies with time.
-
-```{r illu-input, echo = FALSE, out.width = "100%", fig.cap="Illustration of incoming data formats for spatio-temporal data.
-(1) Data comes in as a single table; (2) Separate tables for spatial and temporal variables; (3) A single table with all the parameters used to query the database and a separate table for queried data; and (4) Cubical data in array or NetCDF format.", fig.align='center', out.height = "15%"}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.001.png"))
-```
-
-In a cubble, there are two forms: nested form and long form, and Figure \ref{fig:illu-cubble} sketches the two forms along with the associated attributes.
-The decision on which form to use is output-oriented, meaning analysts need to first think about whether the output of a particular operation is identified only by the spatial identifier, or a combination of spatial and temporal identifier.
-The nested cubble is suitable for working with operations that are only identified by site and this type of operation can be a pure manipulation of spatial variables, or a summary of temporal variables by site (i.e.
-the output of counting the number of raining day is only identified by sites and hence should be performed with the nested form).
-Underneath the nested form, a cubble is built from a row-wise dataframe (`rowwise_df`) where each site forms a separate group.
-This structure simplifies the calculation that involves temporal variables by avoiding the use of `map` syntax when working with list-column.
-
-For those operations whose output involves both a spatial and temporal dimension, long form should be used.
-The long form is identified by both the spatial and temporal identifier and adopts a grouped dataframe (`grouped_df`) to forms each site as a group.
-Spatial variables are stored separately in a \pkg{tibble} as an special attribute of the long cubble.
-This design avoids repeating the spatial variables at each time stamp while not dropping information from spatial variables.
-
-```{r illu-cubble, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "Illustration of nested and long cubble."}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.002.png"))
-```
-
-## Create a cubble in the nested form
-
-To use functionalities from cubble, data analysts first need to create a cubble.
-`as_cubble` create a \pkg{cubble} by supplying the three key components: `key` as the spatial identifier; `index` as the temporal identifier; and a vector of `coords` in the order of longitude first and then latitude.
-The naming of `key` and `index` follows the convention in the \pkg{tsibble} package.
-The cubble created by default is in the nested form.
-Below is an example of creating a cubble: \newline
-
-```{r}
-(cubble_nested <- cubble::climate_flat %>%
-  as_cubble(key = id, index = date, coords = c("long", "lat")))
-```
-
-There are a few information in the \pkg{cubble} header: the name of the `key` variable, bbox, and also the name of variable nested in the `ts` column.
-In this example, each site is identifier is `id` and the number in the bracket means there are 5 unique `id` in this dataset.
-The bbox in the second row gives the range of the coordinates.
-The temporal variables are all nested in the `ts` column, but it could be useful to know the name these variables.
-The third row in the cubble header shows these names and in this example this includes: precipitation, `prcp`, maximum temperature, `tmax`, and minimum temperature, `tmin`.
-
-## Stretch a nested cubble into the long form
-
-The long cubble is suitable to manipulate the time dimension of the data.
-The function `stretch()` switches the nested cubble into the long cubble by first extracts all the spatial variables into a separate tibble and store in the `spatial` attribute and then unnests the `ts` column:
-
-```{r}
-(cubble_long <- cubble_nested %>% stretch(ts))
-```
-
-Notice here that the third line in the header now shows the name of spatial variables rather than the temporal variables.
-
-## Tamp a long cubble back to the nested form
-
-Manipulation on the spatial and temporal dimension can be an iterative process.
-Many times, analysts will need to go back and forth between the nested and long cubble.
-The `stretch()` function introduced in the previous section switches a nested cubble into a long cubble and function `tamp()` is its inverse function to switch a long cubble back to the nested cubble:
-
-```{r}
-(cubble_back <- cubble_long %>% tamp())
-```
-
-## Migrate spatial variables to a long cubble
-
-As a final data output for modelling or visualisation, spatio-temporal data is usually expected to be in a single table.
-Function `migrate()` moves the spatial variables from the `spatial` attribute into the long cubble:
-
-```{r}
-(cubble_long %>% migrate(long, lat))
-```
-
-In this workflow described above, data objects come into cubble in the nested form, then various operations on the spatial and temporal dimension can go back and forth between the nested and long form, and finally, the data will come out of cubble in the long form for further modelling or visualisation.
-
-Building from an underlying `tbl_df` structure, it is natural to implement methods available in `dplyr` to `cubble`. Supported methods in the `cubble` with `dplyr` generics includes:
-
-```{=tex}
-\begin{center}
-\begin{tabular}{ | m{5em} | m{15cm}| } 
-basics & \textbf{mutate}, \textbf{filter}, \textbf{summarise}, \textbf{select}, \textbf{arrange}, \textbf{rename}, \textbf{left\_join} \\
-grouping &  \textbf{group\_by}, \textbf{ungroup}\\
-slice family & \textbf{slice\_head}, \textbf{slice\_tail}, \textbf{slice\_sample}, \textbf{slice\_min} and \textbf{slice\_max} \\
-\end{tabular}
-\end{center}
-```
-\pkg{cubble} is also compatible with \pkg{tsibble} in the sense that the original list-column can be a `tbl_ts` object.
-Duplicates and gaps should be first checked before structuring the data into a cubble.
-If the input data is a \pkg{tsibble} object, the long form cubble is also a \pkg{tsibble} where users can directly apply time series operations.
-
-# Advanced features/ considerations
-
-## Hierarchical structure
-
-Imposing a clustering structure can be thought of as building a hierarchical structure where stations are nested within clusters.
-This can be useful when there's intrinsic nesting structure in the data (i.e.
-country nested in the continent, county nested in the state) or there's some clustering.
-When we have access to both level, `switch_key()` is the function to re-structure the data as one cluster per row.
-Temporal observations from different stations while within the same cluster are bound in the nested column `ts`.
-
-One thing we hope to do with the cluster is to find the coordinates of the centroid.
-These are variables variant to the station but invariant to the cluster and it would be nice to have a function that structure each cluster as a row.
-`switch_key()` is the function that does this: it lets you to specify a new key, say `cluster` and nests all spatial variables variant to `cluster` into a column.
-Spatial variables are all nested inside a new column `.val`.
-Temporal observations from different stations while within the same cluster are bound in the nested column `ts`.
-
-This structure makes it easy to compute cluster level variable, for example, the convex hull and the centroid coordinate of each cluster.
-These can be amended into the nested form with function `get_centroid`.
-
-After we have got `cluster_nested`, spatial and temporal data at both levels can be easily obtained.
-Figure \ref{fig:illu-hier} illustrate the relationship between the long and nested form cubble at both site and cluster level.
-More description on this.
-Start with the original `station_nested`, `stretch()` expands the `ts` column with each station (`id`) forming a group and attach variables invariant to `id` as an attribute.
-`switch_key()` changes the `key` from `id` to `cluster` and nests all the spatial variables that variant to `cluster`.
-`stretch()` `cluster_nested` will store variables that are invariant to `cluster` as a tibble in the attribute.
-
-this fit into the remaining data pipeline.
-
-```{r illu-hier, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "Hierarchical structure"}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.003.png"))
-```
-
-## Data fusion and matching
-
-Temporal matching checks how spatially matched pairs align temporally. We use the following chart to illustrate how the temporal matching works:
-
-For each spatially matched pair, say `A` and `a`, we first find the largest `n` points in each series, colored in brown points here.
-Here we use the largest three but you can tune this number by `temporal_n_highest`.
-Then we construct the interval of the largest points from one series and see how many points, from the other series, fall into the intervals.
-The series used to construct the interval is controlled by `temporal_independent` and the window size by `temporal_window` with a default of 5.
-
-In this illustration, we construct the interval based on series `A` and two of the three peaks from `a` falls into this interval at Time 7 and 27.
-
-There's another mandatory argument that hasn't been introduced above: `temporal_var_to_match`.
-This argument controls the variable to match and it needs to appear in both the `major` and `minor` set.
-In the water level matching example, we match the variable `Water_course_level` from `river` to `prcp` from `climate`, hence need to manually rename one of them to match the other, here we rename `Water_course_level` to `prcp` in `river`:
-
-Figure \ref{fig:illu-matching}
-
-```{r illu-matching, out.width="100%", out.height="30%", fig.cap="sdfasdf"}
-knitr::include_graphics("figures/illu-matching.png")
-```
-
-## Interactive graphics
-
-<!-- Interactive graphics can listen to users' actions on the plot to provide additional information that facilitates data exploration.
-This is a useful technique for spatio-temporal data since users can zoom or pan the map to view the local and global structure of the map; use tooltips or popups to query more information about a graphic element; or highlight points to explore its linked views in other plots.
-In the R community, many implementations have been developed to connect \proglang{R} to \proglang{javascript} to create interactive graphics.
-In relation to spatio, temporal, and spatio-temporal data, the general purpose packages \pkg{plotly} [@plotly] and \pkg{leaflet} [@leaflet] realise various interactive actions through their corresponding javascript libraries.
-\pkg{crosstalk} [@crosstalk] and \pkg{tsibbletalk} [@tsibbletalk] implement brushed linking between htmlwidgets.
-\pkg{ggiraph} [@ggiraph] enables tooltip, self-linking, and customised actions specified through its own \proglang{javascript}.
--->
-
-<!-- While many graphic implementations present worked examples to illustrate the usage of the package, few documents the underlying pipeline that transforms the raw data step-by-step to the final view on the screen.
-There have been some early work in building the data pipeline for (interactive) graphics [@buja1988elements; @buja1996interactive;  @sutherland2000orca] and more recent discussions include @wickham2009plumbing, @xie2014reactive, and @cheng2016enabling.
--->
-
-The cubble structure fits in naturally with the interactive graphic pipeline discussed in the literature [@buja1988elements; @buja1996interactive;  @sutherland2000orca; @xie2014reactive; @cheng2016enabling].
-Diagram \ref{fig:illu-interactive} illustrates how linking works with the two forms in a cubble, where a time series plot is created with the long cubble and a map is created with the nested cubble.
-When a user action is captured from the map, the site will be activated in the nested cubble.
-Then, the nested cubble will communicate to the long cubble to activate all the observations with the same `id`.
-The long cubble will then highlight the  activated series in the time series plot.
-
-The linking is also available from the time series plot to the map.
-The selection on the time series is through selecting the point on the time series and once a point is selected, it will be activated in the long cubble.
-All the observations that share the same `id`,  either in the long and nested cubble, are then activated.
-This includes other points in the same time series in the long cubble and the corresponding observation of site in the nested cubble.
-These activated observations will then being reflected in the updated plots and Diagram \ref{fig:illu-interactive-2} in the Appendix illustrates this process.
-
-```{r illu-interactive, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "demon interactivity"}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.004.png"))
-```
-
-\newpage
-
-# Examples
-
-## Australia historical maximum temperature
-
-Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world and the dataset `weatherdata::historical_tmax` extracts the historical maximum temperature recorded for `r nrow(weatherdata::historical_tmax)` Australian stations.
-The data `historical_tmax` is already presented as a cubble, with `id` as the key, `date` as the index, and `c("longitude", "latitude")` as the coordinates.
-Other variables include `elevation`, `name`, `wmo_id`, `first_year`, and  `last_year` in the nested form and `tmax` in the long form.
-This example compares the maximum temperature in two periods: 1971 - 1975 and 2016 - 2020 for stations in Victoria and New South Wales.
-
-Stations in the two states can be subsetted on the station number: Australia GHCN station number starts with `ASN00` and followed by the [Bureau of Meteorology (BOM) station number](http://www.bom.gov.au/climate/cdo/about/site-num.shtml), where the 2nd and 3rd digit (7th and 8th in the GHCN number) denote the state a station belongs to.
-New South Wales stations start from 46 to 75 and Victoria stations then follow from 76 to 90.
-Extracting Victoria and New South Wales stations is a filter operation in the spatial dimension and hence is operated in the nested form:
-
-```{r eval = FALSE, echo = TRUE}
-tmax <- weatherdata::historical_tmax %>%
-  filter(between(stringr::str_sub(id, 7, 8), 46, 90))
-```
-
-The five year window is chosen to remove the effect of a particular year and the historical period of 1971 - 1975 is used since all the stations have records from 1970.
-This following chunk filters on records in the two study periods and summarises the maximum temperature into monthly measure.
-`stretch()` is used to convert the nested form cubble into the long form for these operations:
-
-```{r eval=  FALSE, echo = TRUE}
-tmax <- tmax %>% 
-  stretch() %>%
-  filter(lubridate::year(date) %in% c(1971:1975, 2016:2020)) %>%
-  mutate(month = lubridate::month(date), 
-         group = as.factor(ifelse(lubridate::year(date) > 2015, 
-                                  "2016 ~ 2020", "1971 ~ 1975"))) %>%
-  group_by(month, group) %>%
-  summarise(tmax = mean(tmax, na.rm = TRUE))
-```
-
-A data quality issue with GHCN data is that while the first and last year of each series is provided, years missing in this period is not reported.
-There are a few stations which don't have records during 1971 - 1975 and these stations are filtered out by examining whether the summarised `tmax` has a total of 24 months.
-This is again a station-wise operation and is operated in the nested form, which is switched to from the long form by  `tamp()`:
-
-```{r eval = FALSE, echo = TRUE}
-tmax <- tmax %>% 
-  tamp() %>%
-  filter(nrow(ts) == 24) 
-```
-
-Lastly, to create a glyph map, both the major (`longitude`, `latitude`) and minor (`month`, `tmax`) coordinates need to be in the same table.
-Spatial variables can be moved to the long form with `migrate()`:
-
-```{r eval = FALSE, echo = TRUE}  
-tmax <- tmax %>%   
-  stretch() %>%
-  migrate(latitude, longitude)
-```
-
-Figure \ref{fig:basic-manip} shows the glyph map made with the data after the wangling above.
-One issue with this map is that the similar pattern shown from a few nearby stations around Sydney (151E, 34S) and New castle (152E, 33S) can be distracting.
-Aggregation is a useful technique to observe the general pattern of a collection of series and will be the topic of the next example.
-
-```{r basic-manip, out.width="100%", out.height="70%", fig.cap="Glyph map of the mean maximum temperature by month for Victoria and New South Wales weather stations.
-On the top left corner is a detailed legend for station Cobar highlighted in the black box.
-Compared to 1971 - 1975,  the period 2016 - 2020 sees an increase of mean maximum temperature in spring and summer in Australia (end and beginning of the year).
-A larger increase during the first three months of the year is observed at stations with a lower latitude."}
-knitr::include_graphics("figures/basic-manip.png")
-```
-
-## Australia precipitation pattern in 2020
-
-In Figure \ref{fig:basic-agg}.
-Aggregating them into averages or other statistics would give a better picture.
-This can be done with cubble through a hierarchical structure, where the cluster indicator is one level above the individual station.
-
-The `weatherdata::climate_full` data has daily climate data of `r nrow(weatherdata::climate_full)` Australia stations from 2016 to 2020.
-
-calcluate distance matrix
-As an example to illustrate here, a kmean clustering algorithm based on the distance matrix is used and the number of centres is set to 20.
-More complex algorithms can also be used for more complex problem, as long as a mapping from each station id to the cluster id can be constructed.
-
-<!-- Now we can obtain the aggregated series as described in the workflow diagram above and construct the glyph map with `GGally::glyphs()`: -->
-
-<!-- We can also look at the precipitation of each individual station within the same cluster: -->
-
-<!-- Lastly, there is one series on Tasmania island standing out from others, lets look at where it is: -->
-
-```{r basic-agg, out.width="100%", out.height="90%", fig.cap="sdfasdf"}
-knitr::include_graphics("figures/basic-agg.png")
-```
-
-## River level data in Victria water gauges
-
-The water level data comes from [Bureau of Meteorology](http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full) and has a copy in `weatherdata`.
-Here we extract the water course level and add a column annotate this data of type `river`.
-For the rainfall data, we will still use the `weatherdata::climate_full`, filtering for Victorian stations in 2020 should be pretty familiar by now.
-Again, we first look at where these stations are on the map first:
-
-Now we use `match_sites()` to first pair the weather stations with the river gauges spatially and then apply the temporal matching on `prcp`.
-We will construct the interval based on peaks in `climate` since we would expect a lag effect for precipitation to flow into the river and cause a raise in river level, hence `temporal_independent = climate`.
-We select the 30 highest peak from the series to construct the match by setting `temporal_n_highest = 30`.
-This is a tuning parameter and you can start with 10% of the points of one series (here we have daily data for a year, 10% is roughly 30 points).
-`temporal_min_match` filters out pairs don't have enough match and to return all the pairs, set `temporal_min_match` to `0`.``
-
-The output from temporal matching is also a cubble, with additional column `.dist` and `.group` inherent from spatial matching and `n_match` for the number of matched temporal peaks.
-Then you can use this output to plot the location of match or to look at the series:
-
-```{r matching, out.width="100%", out.height="30%", fig.cap="asdfasd"}
-knitr::include_graphics("figures/matching.png")
-```
 
@@
 \newpage
-
-## Interative graphic with cubble
-
-With a larger dataset, interactive graphics can be helpful to explore these data from multiple perspective.
-Two examples are presented here to show how interactivity can be used to explore the climate data across Australia.
-Again, `weatherdata::climate_full` will be used and we hope to include more than one time variable here.
-The difference between the maximum and minimum temperature of the day, the diurnal range, can be directly calculated from the data and it can have large impact on the wine industry (lol this is going far).
-This would be useful to compare the temperature difference across stations and to compare how this temperature difference differs throughout the year.
-
-Temperature difference is a time-wise variable and should be computed in the long form cubble. The variables are averaged over the 2016 to 2020 period by month to stabilise the result.
-
-```{r eval = FALSE, echo = TRUE}
-res <- weatherdata::climate_full %>%
-  # filter(!id %in% c("ASN00067033", "ASN00072091", 
-  #                   "ASN00059040", "ASN00097053")) %>%
-  stretch() %>%
-  mutate(month = lubridate::month(date),
-         diff = tmax - tmin) %>%
-  group_by(month) %>%
-  summarise(tmax = mean(tmax, na.rm = TRUE),
-            diff = mean(diff, na.rm = TRUE)) %>%
-  ungroup(month)
-
-```
-
-Several variables can be useful when comparing the seasonal effect of the temperature differences, for example, the difference of average temperature difference between summer and winter can tell [...].
-This is a station-wise measure and hence computed in the nested form.
-
-```{r eval = FALSE, echo = TRUE}
-res <- res %>% 
-  tamp() %>%
-  mutate(
-    diff_winter = ts %>% filter(month %in% c(6, 7, 8)) %>%
-      pull(diff) %>% mean(na.rm = TRUE),
-    diff_summer = ts %>% filter(month %in% c(12, 1, 2)) %>%
-      pull(diff) %>% mean(na.rm = TRUE),
-    diff_year = diff_winter - diff_summer
-  )
-```
-
-The result from the above manipulation can be directly supplied to make interactive plots and Figure \ref{fig:interactive-linking} is an example made with \pkg{plotly} and \pkg{crosstalk}.
-For example, the second row results from selecting the highest mean maximum temperature in July and this corresponds to Kalumburu where there is a larger winter (June to August) temperature difference than summer (December to February).
-The third row selects one of the stations (Birdsville Airport) with a large average temperature difference in the interior desert region on the map and the linked view from the time series shows that this maximum and minimum temperature difference is constant throughout the year.
-The last row selects the lowest maximum temperature in July on the temperature band and surprisingly, the corresponding station doesn't fall into the Tasmania island but Mountain Hotham in Victoria.
-
-```{r interactive-linking, echo = FALSE, out.width="100%", out.height="25%", fig.cap = "Screenshots of the mean daily temperature difference averaged by month from 2016 to 2020 at various locations in Australia.
-Crossed linking between the map and the temperature band is implemented in both directions for exploring the temperature pattern interactively.
-The first row is the general appearance of the graph and the next three rows are examples from various selection to show different temperature patterns in Australia.
-The html version can be found at ...", fig.show='hold'}
-knitr::include_graphics(here::here("figures/linking.png"))
-knitr::include_graphics(here::here("figures/linking-north.png"))
-knitr::include_graphics(here::here("figures/linking-mid.png"))
-knitr::include_graphics(here::here("figures/linking-south.png"))
-```
-
-Cubble is also compatible with \pkg{leaflet} and with leaflet extension package \pkg{leafpop}, the temperature band can be displayed upon clicking on the station on the map.
-Figure \ref{fig:interactive-popup} shows more examples of temperature pattern that can be found in Australia.
-With the same latitude as Kalumburu but in the Pacific coastline, Cooktown airport shows a similarly high but constant temperature throughout the year.
-With a slightly higher latitude than the mountain Hotham, flat land in Melbourne airport shows a less extreme temperature band.
-Lastly, the Cape Grim Baps, in the north-west tip of the Tasmania island shows a uniformly low and thin temperature band.
-
-```{r interactive-popup, echo = FALSE, out.width="50%", out.height="30%", fig.cap = "Screenshots of the diurnal temperature range averaged by month from 2016 to 2020 at various locations in Australia.
-The leaflet popup allows users to click on a single station for its temperature band as a small plot in the popup window.
-The html version can be found at ...", fig.show='hold'}
-knitr::include_graphics(here::here("figures/popup-north.png"))
-knitr::include_graphics(here::here("figures/popup-vic.png"))
-knitr::include_graphics(here::here("figures/popup-tas.png"))
-```
-
-# Conclusion
-
-# Appendix
-
-```{r illu-interactive-2, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "demon interactivity"}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.005.png"))
-```
-
-\newpage
-
-<!-- # Old stuff -->
-
-<!-- In the temporal aspect, the `tsibble` [@tsibbles] structure and its tidyverts ecosystem have provided a [...
-] workflow to work with temporal data.
-In a tsibble structure, temporal data is characterised by `index` and `key` where `index` is the temporal identifier and `key` is the identifier for multiple series, which could be used as a spatio identifier.
-However, a tsibble object, by construction, always requires the `index` in its structure.
-This makes it less appealing for spatio-temporal data since the output of calculated spatio-specific variables (i.e.
-features of each series) don't have the time dimension.
-Analysts will either need to have an additional step to join this output to the original tsibble or operate with variables stored in two separate objects.
-In addition, the long form structure of a tsibble object means spatio variables (i.e.
-longitude, latitude, and features of each series if joined back to the tsibble) of each spatio identifier will be repetitively recorded at each timestamp.
-This repetition is unnecessary and would inflate the object size for long series.
--->
-
-<!-- # Create a cubble -->
-
-<!-- The creation of a cubble requires the site identifier (`key`), as well as the spatial (`coords`) and temporal (`index`) identifier.
-`climate_flat` is already a tibble and it uses `id` to identify each station, `date` as the time identifier, and `c(long, lat)` as the spatial identifier.
-To create a cubble for this data, use: -->
-
-<!-- ```{r} -->
-<!-- climate_flat %>% as_cubble(key = id, index = date, coords = c(long, lat)) -->
-<!-- ``` -->
-
-<!-- Most of the time, spatio-temporal data doesn't come into this form and analysts need to query the climate variables based on station metadata.
-\textcolor{red}{This is also a problem illustrated in Section 3.5 in @tidydata.
-Here we provide a structured way to query this data based on the row-wise operator and nested list.} For this type of task, one can structure a metadata into a tibble and use row-wise operator to query the climate variables into a nested list.
-As an example here we demonstrate the workflow to find the 5 closest stations to Melbourne.
-We first create a station data frame with the 5 target stations.
--->
-
-<!-- ```{r echo = FALSE} -->
-<!-- latlon_ref <- tibble::tibble(id = "melbourne", lat = -37.8136, long = 144.9631) -->
-
-<!-- cand <- aus_climate %>% -->
-<!--   as_tibble() %>% -->
-<!--   select(id, lat, long, elev, name, wmo_id) -->
-
-<!-- calc_dist <- function(df_ref, df, n = 5) { -->
-<!--   long_ref <- df_ref$long -->
-<!--   lat_ref <- df_ref$lat -->
-
-<!--   df %>% -->
-<!--     mutate( -->
-<!--       long_ref = long_ref, -->
-<!--       lat_ref = lat_ref, -->
-<!--       dist = rnoaa::meteo_spherical_distance(lat, long, lat_ref, long_ref) -->
-<!--     ) %>% -->
-<!--     slice_min(dist, n = n) %>% -->
-<!--     mutate(city = df_ref$id) %>% -->
-<!--     select(-long_ref, -lat_ref) -->
-<!-- } -->
-
-<!-- (stations <- do.call("rbind", purrr::map( -->
-<!--   1:nrow(latlon_ref), -->
-<!--   ~ calc_dist(latlon_ref[.x, ], cand) -->
-<!-- ))) -->
-<!-- ``` -->
-
-<!-- We can query the climate information into a nested list named `ts` for each station with the `rowwise()` operator. To create a cubble, supply the same identifiers as with the first example. -->
-
-<!-- ```{r eval = FALSE} -->
-<!-- sydmel_climate <- stations %>% -->
-<!--   rowwise() %>% -->
-<!--   mutate(ts = list(meteo_pull_monitors(id, -->
-<!--     date_min = "2020-01-01", -->
-<!--     date_max = "2020-12-31", -->
-<!--     var = c("PRCP", "TMAX", "TMIN") -->
-<!--   ) %>% -->
-<!--     select(-id))) %>% -->
-<!--   as_cubble(key = id, index = date, coords = c(long, lat)) -->
-<!-- ``` -->
-
-<!-- ```{r echo = FALSE} -->
-<!-- (sydmel_climate <- stations %>% -->
-<!--   left_join(aus_climate) %>% -->
-<!--   as_cubble(key = id, index = date, coords = c(long, lat))) -->
-<!-- ``` -->
-
-<!-- Below are the how the nested and long form look like for Australia climate data, which records daily precipitation, maximum and minimum temperature for 55 stations across Australia from 2015- 2020.
-Notice that each station forms a group in both forms and specifically, the nested and long form have a underlying `rowwise_df` and `grouped_df` respectively.
--->
-
-<!-- With a cubic framework on mind, different types of manipulation with cubble can be thought of as slicing the cube in various way.
-The table below shows how some `dplyr` verbs are mapped into the operation in a cubble.
-With the existing grouping on the station, additional grouping can be added with `group_by` and removed with `ungrouped`.
-[talk about why it is useful] -->
-
-<!-- \newpage -->
-
-<!-- ## Cubble operations -->
-
-<!-- ```{r cubble-operations, echo = FALSE, fig.cap="Cubble operations", out.height="50%", out.width="90%", fig.align='center'} -->
-<!-- knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.004.png")) -->
-<!-- ``` -->

### 2022-01-07 4-5-interactive.Rmd
play aroudnd with child rmd to organising sections

--- 2021-12-20 main.Rmd
@@
 ---
-documentclass: jss
-author:
-  - name: H. Sherry Zhang
-    # department: Department of Econometrics and Business Statistics
-    affiliation: Monash University
-    #location: Melbourne, Australia
-    email: \email{huize.zhang@monash.edu}
-  - name: Dianne Cook
-    # department: Department of Econometrics and Business Statistics
-    affiliation: 'Monash University \AND'
-    # location: Melbourne, Australia
-    # email:  dicook@monash.edu
-  - name: Ursula Laa
-    # department: Institute of Statistics
-    affiliation: 'University of Natural Resources and Life Sciences \AND'
-    # location: Vienna, Austria
-    # email:  ursula.laa@boku.ac.at
-  - name: Nicolas Langrené
-    # department: 34 Village Street, Docklands VIC 3008 Australia
-    affiliation: 'CSIRO Data61 \AND'
-    # location: Melbourne, Australia
-    # email: nicolas.langrene@csiro.au
-  - name: Patricia Menéndez
-    # department: Department of Econometrics and Business Statistics
-    affiliation: 'Monash University \AND'
-    # location: Melbourne, Australia
-    # email:  patricia.menendez@monash.edu
-title:
-  formatted: "\\pkg{cubble}: An R Package for Structuring Spatio-temporal Data"
-  # If you use tex in the formatted title, alsoge supply version without
-  plain:     "cubble: An R Package for Structuring Spatio-temporal Data"
-  # For running headers, if needed
-  # short:     "\\pkg{foo}: A Capitalized Title"
-abstract: >
-  The abstract of the article.
-keywords:
-  # at least one keyword must be supplied
-  formatted: [spatio-temporal data,  "\\proglang{R}"]
-  plain:     [spatio-temporal data, R]
-preamble: >
-  \usepackage{amsmath}
-  \usepackage{array}
-output: rticles::jss_article
-bibliography: references.bib
+title: "4-5-interactive"
+output: html_document
 ---
-
-```{r setup, echo = FALSE}
-knitr::opts_chunk$set(
-  echo = FALSE,
-  warning = FALSE,
-  message = FALSE,
-  cache = TRUE,
-  fig.path = "figures/",
-  fig.height = 10
-)
-options(prompt = "R> ", continue = "+ ", 
-        tibble.print_max = 5, tibble.print_min = 5)
-```
-
-```{r echo = FALSE}
-library(cubble)
-library(forcats)
-library(stringr)
-library(dplyr)
-library(lubridate)
-library(ggplot2)
-library(GGally)
-library(plotly)
-library(crosstalk)
-library(patchwork)
-```
-
-\newpage
-
-# Introduction
-
-Many data structures have been proposed for spatial (\pkg{sf} by @sf) and temporal (\pkg{tsibble} by @tsibble) data in the R community, while less has been done for spatio-temporal data.
-The lack of such tools could potentially because analysts usually treat the spatial and temporal dimension of the data separately, without realising the need to create a new data structure.
-While this approach follows the third tidy data principal [@tidydata] (*Each type of observational unit forms a table*), analysts always need to manually join results from different observational units or combining multiple tables into one for downstream analysis.
-This additional step doesn't add new operations into the data but can be error prone.
-\newline
-
-Currently, available spatio-temporal data structure in R includes: \pkg{spacetime} [@spacetime], which proposes four space-time layouts: Full grid (STF), sparse grid(STS), irregular (STI), and trajectory (STT).
-The data structure it uses is based on \pkg{sp} [@sp] and \pkg{xts} [@xts], both of which has been replaced by more recent implementations.
-\pkg{spatstat} [@spatstat] implements a `ppp` class for point pattern data; and more recent, \pkg{stars} [@stars] implements a spatio-temporal array with the dplyr's data cube structure \pkg{cubelyr} [@cubelyr] as its backend.
-While these implementations either store spatial and temporal variables all in a single table, hence duplicate the spatial variables for each temporal unit; or split them into two separate tables that has the problem of manually joining, mentioned in the previously.
-None of these packages enjoy both the benefits of being able to separate manipulation in the two dimensions while also keep the data object as a whole.
-This create a gap in the software development.
-The requirement for such a tool is important given the ubiquity of spatio-temporal vector data in the wild: the Ireland wind data from \pkg{gstat} is an classic example data that splits variables into spatial (`wind.loc`) and temporal (`wind`) dimension; Bureau of Meteorology (BoM) provides climate observations that are widely applied in agriculture and ecology study; air pollution data.
-\newline
-
-This paper describes the implementation of a new spatio-temporal data structure: \pkg{cubble}.
-\pkg{cubble} implements a relational data structure that uses two forms to manage the switch between spatial and temporal dimension.
-With this structure, users can manipulate the spatial or temporal dimension separately, while leaves the linking of two dimensions to \pkg{cubble}.
-The software is available from the Comprehensive R Archive Network (CRAN) at [CRAN link].
-\newline
-
-The rest of the paper will be divided as follows: [complete when the paper structure is more solid]
-
-<!-- Section 2 reviews the existing data structure for spatio, temporal, and spatio-temporal data.
-Section 3 presents a new data structure for spatio-temporal data: cubble.
-Then the paper introduces the workflow of data manipulation and visualisation with the cubble structure in Section 4.
-Section 5 gives some examples on how common spatial and temporal manipulations are performed with cubble and how static and interactive visualisation help to understand climate and [...] data.
-\newline -->
-
-\newpage
-
-# The cubble package
-
-Spatio-temporal data usually come in various forms and Figure \ref{fig:illu-input} shows four examples of this.
-No matter which form the data is in, these formats share some common components that characterise spatio-temporal data.
-A spatial identifier (`id` in the diagram) is the unique identifier of each site.
-The temporal identifier (`t` in the diagram) prescribes the time stamp each site is recorded.
-Coordinates, comprising of latitude and longitude (`lon` and `lat` in the diagram), locates each site on the map.
-These identifiers will be the building blocks for the data structure introduced below.
-Other variables in the data can be categorised into two groups: spatial variables that are invariant at each time stamp for every site, i.e.
-the name or code of the weather station and temporal variables that varies with time.
-
-```{r illu-input, echo = FALSE, out.width = "100%", fig.cap="Illustration of incoming data formats for spatio-temporal data.
-(1) Data comes in as a single table; (2) Separate tables for spatial and temporal variables; (3) A single table with all the parameters used to query the database and a separate table for queried data; and (4) Cubical data in array or NetCDF format.", fig.align='center', out.height = "15%"}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.001.png"))
-```
-
-In a cubble, there are two forms: nested form and long form, and Figure \ref{fig:illu-cubble} sketches the two forms along with the associated attributes.
-The decision on which form to use is output-oriented, meaning analysts need to first think about whether the output of a particular operation is identified only by the spatial identifier, or a combination of spatial and temporal identifier.
-The nested cubble is suitable for working with operations that are only identified by site and this type of operation can be a pure manipulation of spatial variables, or a summary of temporal variables by site (i.e.
-the output of counting the number of raining day is only identified by sites and hence should be performed with the nested form).
-Underneath the nested form, a cubble is built from a row-wise dataframe (`rowwise_df`) where each site forms a separate group.
-This structure simplifies the calculation that involves temporal variables by avoiding the use of `map` syntax when working with list-column.
-
-For those operations whose output involves both a spatial and temporal dimension, long form should be used.
-The long form is identified by both the spatial and temporal identifier and adopts a grouped dataframe (`grouped_df`) to forms each site as a group.
-Spatial variables are stored separately in a \pkg{tibble} as an special attribute of the long cubble.
-This design avoids repeating the spatial variables at each time stamp while not dropping information from spatial variables.
-
-```{r illu-cubble, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "Illustration of nested and long cubble."}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.002.png"))
-```
-
-## Create a cubble in the nested form
-
-To use functionalities from cubble, data analysts first need to create a cubble.
-`as_cubble` create a \pkg{cubble} by supplying the three key components: `key` as the spatial identifier; `index` as the temporal identifier; and a vector of `coords` in the order of longitude first and then latitude.
-The naming of `key` and `index` follows the convention in the \pkg{tsibble} package.
-The cubble created by default is in the nested form.
-Below is an example of creating a cubble: \newline
-
-```{r}
-(cubble_nested <- cubble::climate_flat %>%
-  as_cubble(key = id, index = date, coords = c("long", "lat")))
-```
-
-There are a few information in the \pkg{cubble} header: the name of the `key` variable, bbox, and also the name of variable nested in the `ts` column.
-In this example, each site is identifier is `id` and the number in the bracket means there are 5 unique `id` in this dataset.
-The bbox in the second row gives the range of the coordinates.
-The temporal variables are all nested in the `ts` column, but it could be useful to know the name these variables.
-The third row in the cubble header shows these names and in this example this includes: precipitation, `prcp`, maximum temperature, `tmax`, and minimum temperature, `tmin`.
-
-## Stretch a nested cubble into the long form
-
-The long cubble is suitable to manipulate the time dimension of the data.
-The function `stretch()` switches the nested cubble into the long cubble by first extracts all the spatial variables into a separate tibble and store in the `spatial` attribute and then unnests the `ts` column:
-
-```{r}
-(cubble_long <- cubble_nested %>% stretch(ts))
-```
-
-Notice here that the third line in the header now shows the name of spatial variables rather than the temporal variables.
-
-## Tamp a long cubble back to the nested form
-
-Manipulation on the spatial and temporal dimension can be an iterative process.
-Many times, analysts will need to go back and forth between the nested and long cubble.
-The `stretch()` function introduced in the previous section switches a nested cubble into a long cubble and function `tamp()` is its inverse function to switch a long cubble back to the nested cubble:
-
-```{r}
-(cubble_back <- cubble_long %>% tamp())
-```
-
-## Migrate spatial variables to a long cubble
-
-As a final data output for modelling or visualisation, spatio-temporal data is usually expected to be in a single table.
-Function `migrate()` moves the spatial variables from the `spatial` attribute into the long cubble:
-
-```{r}
-(cubble_long %>% migrate(long, lat))
-```
-
-In this workflow described above, data objects come into cubble in the nested form, then various operations on the spatial and temporal dimension can go back and forth between the nested and long form, and finally, the data will come out of cubble in the long form for further modelling or visualisation.
-
-Building from an underlying `tbl_df` structure, it is natural to implement methods available in `dplyr` to `cubble`. Supported methods in the `cubble` with `dplyr` generics includes:
-
-```{=tex}
-\begin{center}
-\begin{tabular}{ | m{5em} | m{15cm}| } 
-basics & \textbf{mutate}, \textbf{filter}, \textbf{summarise}, \textbf{select}, \textbf{arrange}, \textbf{rename}, \textbf{left\_join} \\
-grouping &  \textbf{group\_by}, \textbf{ungroup}\\
-slice family & \textbf{slice\_head}, \textbf{slice\_tail}, \textbf{slice\_sample}, \textbf{slice\_min} and \textbf{slice\_max} \\
-\end{tabular}
-\end{center}
-```
-\pkg{cubble} is also compatible with \pkg{tsibble} in the sense that the original list-column can be a `tbl_ts` object.
-Duplicates and gaps should be first checked before structuring the data into a cubble.
-If the input data is a \pkg{tsibble} object, the long form cubble is also a \pkg{tsibble} where users can directly apply time series operations.
-
-# Advanced features/ considerations
-
-## Hierarchical structure
-
-Imposing a clustering structure can be thought of as building a hierarchical structure where stations are nested within clusters.
-This can be useful when there's intrinsic nesting structure in the data (i.e.
-country nested in the continent, county nested in the state) or there's some clustering.
-When we have access to both level, `switch_key()` is the function to re-structure the data as one cluster per row.
-Temporal observations from different stations while within the same cluster are bound in the nested column `ts`.
-
-One thing we hope to do with the cluster is to find the coordinates of the centroid.
-These are variables variant to the station but invariant to the cluster and it would be nice to have a function that structure each cluster as a row.
-`switch_key()` is the function that does this: it lets you to specify a new key, say `cluster` and nests all spatial variables variant to `cluster` into a column.
-Spatial variables are all nested inside a new column `.val`.
-Temporal observations from different stations while within the same cluster are bound in the nested column `ts`.
-
-This structure makes it easy to compute cluster level variable, for example, the convex hull and the centroid coordinate of each cluster.
-These can be amended into the nested form with function `get_centroid`.
-
-After we have got `cluster_nested`, spatial and temporal data at both levels can be easily obtained.
-Figure \ref{fig:illu-hier} illustrate the relationship between the long and nested form cubble at both site and cluster level.
-More description on this.
-Start with the original `station_nested`, `stretch()` expands the `ts` column with each station (`id`) forming a group and attach variables invariant to `id` as an attribute.
-`switch_key()` changes the `key` from `id` to `cluster` and nests all the spatial variables that variant to `cluster`.
-`stretch()` `cluster_nested` will store variables that are invariant to `cluster` as a tibble in the attribute.
-
-this fit into the remaining data pipeline.
-
-```{r illu-hier, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "Hierarchical structure"}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.003.png"))
-```
-
-## Data fusion and matching
-
-Temporal matching checks how spatially matched pairs align temporally. We use the following chart to illustrate how the temporal matching works:
-
-For each spatially matched pair, say `A` and `a`, we first find the largest `n` points in each series, colored in brown points here.
-Here we use the largest three but you can tune this number by `temporal_n_highest`.
-Then we construct the interval of the largest points from one series and see how many points, from the other series, fall into the intervals.
-The series used to construct the interval is controlled by `temporal_independent` and the window size by `temporal_window` with a default of 5.
-
-In this illustration, we construct the interval based on series `A` and two of the three peaks from `a` falls into this interval at Time 7 and 27.
-
-There's another mandatory argument that hasn't been introduced above: `temporal_var_to_match`.
-This argument controls the variable to match and it needs to appear in both the `major` and `minor` set.
-In the water level matching example, we match the variable `Water_course_level` from `river` to `prcp` from `climate`, hence need to manually rename one of them to match the other, here we rename `Water_course_level` to `prcp` in `river`:
-
-Figure \ref{fig:illu-matching}
-
-```{r illu-matching, out.width="100%", out.height="30%", fig.cap="sdfasdf"}
-knitr::include_graphics("figures/illu-matching.png")
-```
-
-## Interactive graphics
-
-<!-- Interactive graphics can listen to users' actions on the plot to provide additional information that facilitates data exploration.
-This is a useful technique for spatio-temporal data since users can zoom or pan the map to view the local and global structure of the map; use tooltips or popups to query more information about a graphic element; or highlight points to explore its linked views in other plots.
-In the R community, many implementations have been developed to connect \proglang{R} to \proglang{javascript} to create interactive graphics.
-In relation to spatio, temporal, and spatio-temporal data, the general purpose packages \pkg{plotly} [@plotly] and \pkg{leaflet} [@leaflet] realise various interactive actions through their corresponding javascript libraries.
-\pkg{crosstalk} [@crosstalk] and \pkg{tsibbletalk} [@tsibbletalk] implement brushed linking between htmlwidgets.
-\pkg{ggiraph} [@ggiraph] enables tooltip, self-linking, and customised actions specified through its own \proglang{javascript}.
--->
-
-<!-- While many graphic implementations present worked examples to illustrate the usage of the package, few documents the underlying pipeline that transforms the raw data step-by-step to the final view on the screen.
-There have been some early work in building the data pipeline for (interactive) graphics [@buja1988elements; @buja1996interactive;  @sutherland2000orca] and more recent discussions include @wickham2009plumbing, @xie2014reactive, and @cheng2016enabling.
--->
-
-The cubble structure fits in naturally with the interactive graphic pipeline discussed in the literature [@buja1988elements; @buja1996interactive;  @sutherland2000orca; @xie2014reactive; @cheng2016enabling].
-Diagram \ref{fig:illu-interactive} illustrates how linking works with the two forms in a cubble, where a time series plot is created with the long cubble and a map is created with the nested cubble.
-When a user action is captured from the map, the site will be activated in the nested cubble.
-Then, the nested cubble will communicate to the long cubble to activate all the observations with the same `id`.
-The long cubble will then highlight the  activated series in the time series plot.
-
-The linking is also available from the time series plot to the map.
-The selection on the time series is through selecting the point on the time series and once a point is selected, it will be activated in the long cubble.
-All the observations that share the same `id`,  either in the long and nested cubble, are then activated.
-This includes other points in the same time series in the long cubble and the corresponding observation of site in the nested cubble.
-These activated observations will then being reflected in the updated plots and Diagram \ref{fig:illu-interactive-2} in the Appendix illustrates this process.
-
-```{r illu-interactive, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "demon interactivity"}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.004.png"))
-```
-
-\newpage
-
-# Examples
-
-## Australia historical maximum temperature
-
-Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world and the dataset `weatherdata::historical_tmax` extracts the historical maximum temperature recorded for `r nrow(weatherdata::historical_tmax)` Australian stations.
-The data `historical_tmax` is already presented as a cubble, with `id` as the key, `date` as the index, and `c("longitude", "latitude")` as the coordinates.
-Other variables include `elevation`, `name`, `wmo_id`, `first_year`, and  `last_year` in the nested form and `tmax` in the long form.
-This example compares the maximum temperature in two periods: 1971 - 1975 and 2016 - 2020 for stations in Victoria and New South Wales.
-
-Stations in the two states can be subsetted on the station number: Australia GHCN station number starts with `ASN00` and followed by the [Bureau of Meteorology (BOM) station number](http://www.bom.gov.au/climate/cdo/about/site-num.shtml), where the 2nd and 3rd digit (7th and 8th in the GHCN number) denote the state a station belongs to.
-New South Wales stations start from 46 to 75 and Victoria stations then follow from 76 to 90.
-Extracting Victoria and New South Wales stations is a filter operation in the spatial dimension and hence is operated in the nested form:
-
-```{r eval = FALSE, echo = TRUE}
-tmax <- weatherdata::historical_tmax %>%
-  filter(between(stringr::str_sub(id, 7, 8), 46, 90))
-```
-
-The five year window is chosen to remove the effect of a particular year and the historical period of 1971 - 1975 is used since all the stations have records from 1970.
-This following chunk filters on records in the two study periods and summarises the maximum temperature into monthly measure.
-`stretch()` is used to convert the nested form cubble into the long form for these operations:
-
-```{r eval=  FALSE, echo = TRUE}
-tmax <- tmax %>% 
-  stretch() %>%
-  filter(lubridate::year(date) %in% c(1971:1975, 2016:2020)) %>%
-  mutate(month = lubridate::month(date), 
-         group = as.factor(ifelse(lubridate::year(date) > 2015, 
-                                  "2016 ~ 2020", "1971 ~ 1975"))) %>%
-  group_by(month, group) %>%
-  summarise(tmax = mean(tmax, na.rm = TRUE))
-```
-
-A data quality issue with GHCN data is that while the first and last year of each series is provided, years missing in this period is not reported.
-There are a few stations which don't have records during 1971 - 1975 and these stations are filtered out by examining whether the summarised `tmax` has a total of 24 months.
-This is again a station-wise operation and is operated in the nested form, which is switched to from the long form by  `tamp()`:
-
-```{r eval = FALSE, echo = TRUE}
-tmax <- tmax %>% 
-  tamp() %>%
-  filter(nrow(ts) == 24) 
-```
-
-Lastly, to create a glyph map, both the major (`longitude`, `latitude`) and minor (`month`, `tmax`) coordinates need to be in the same table.
-Spatial variables can be moved to the long form with `migrate()`:
-
-```{r eval = FALSE, echo = TRUE}  
-tmax <- tmax %>%   
-  stretch() %>%
-  migrate(latitude, longitude)
-```
-
-Figure \ref{fig:basic-manip} shows the glyph map made with the data after the wangling above.
-One issue with this map is that the similar pattern shown from a few nearby stations around Sydney (151E, 34S) and New castle (152E, 33S) can be distracting.
-Aggregation is a useful technique to observe the general pattern of a collection of series and will be the topic of the next example.
-
-```{r basic-manip, out.width="100%", out.height="70%", fig.cap="Glyph map of the mean maximum temperature by month for Victoria and New South Wales weather stations.
-On the top left corner is a detailed legend for station Cobar highlighted in the black box.
-Compared to 1971 - 1975,  the period 2016 - 2020 sees an increase of mean maximum temperature in spring and summer in Australia (end and beginning of the year).
-A larger increase during the first three months of the year is observed at stations with a lower latitude."}
-knitr::include_graphics("figures/basic-manip.png")
-```
-
-## Australia precipitation pattern in 2020
-
-In Figure \ref{fig:basic-agg}.
-Aggregating them into averages or other statistics would give a better picture.
-This can be done with cubble through a hierarchical structure, where the cluster indicator is one level above the individual station.
-
-The `weatherdata::climate_full` data has daily climate data of `r nrow(weatherdata::climate_full)` Australia stations from 2016 to 2020.
-
-calcluate distance matrix
-As an example to illustrate here, a kmean clustering algorithm based on the distance matrix is used and the number of centres is set to 20.
-More complex algorithms can also be used for more complex problem, as long as a mapping from each station id to the cluster id can be constructed.
-
-<!-- Now we can obtain the aggregated series as described in the workflow diagram above and construct the glyph map with `GGally::glyphs()`: -->
-
-<!-- We can also look at the precipitation of each individual station within the same cluster: -->
-
-<!-- Lastly, there is one series on Tasmania island standing out from others, lets look at where it is: -->
-
-```{r basic-agg, out.width="100%", out.height="90%", fig.cap="sdfasdf"}
-knitr::include_graphics("figures/basic-agg.png")
-```
-
-## River level data in Victria water gauges
-
-The water level data comes from [Bureau of Meteorology](http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full) and has a copy in `weatherdata`.
-Here we extract the water course level and add a column annotate this data of type `river`.
-For the rainfall data, we will still use the `weatherdata::climate_full`, filtering for Victorian stations in 2020 should be pretty familiar by now.
-Again, we first look at where these stations are on the map first:
-
-Now we use `match_sites()` to first pair the weather stations with the river gauges spatially and then apply the temporal matching on `prcp`.
-We will construct the interval based on peaks in `climate` since we would expect a lag effect for precipitation to flow into the river and cause a raise in river level, hence `temporal_independent = climate`.
-We select the 30 highest peak from the series to construct the match by setting `temporal_n_highest = 30`.
-This is a tuning parameter and you can start with 10% of the points of one series (here we have daily data for a year, 10% is roughly 30 points).
-`temporal_min_match` filters out pairs don't have enough match and to return all the pairs, set `temporal_min_match` to `0`.``
-
-The output from temporal matching is also a cubble, with additional column `.dist` and `.group` inherent from spatial matching and `n_match` for the number of matched temporal peaks.
-Then you can use this output to plot the location of match or to look at the series:
-
-```{r matching, out.width="100%", out.height="30%", fig.cap="asdfasd"}
-knitr::include_graphics("figures/matching.png")
-```
-
-## ERA5: climate reanalysis data
-
-\newpage
-
 ## Interative graphic with cubble
@@
 ```
-
-# Conclusion
-
-# Appendix
-
-```{r illu-interactive-2, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "demon interactivity"}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.005.png"))
-```
-
-\newpage
-
-<!-- # Old stuff -->
-
-<!-- In the temporal aspect, the `tsibble` [@tsibbles] structure and its tidyverts ecosystem have provided a [...
-] workflow to work with temporal data.
-In a tsibble structure, temporal data is characterised by `index` and `key` where `index` is the temporal identifier and `key` is the identifier for multiple series, which could be used as a spatio identifier.
-However, a tsibble object, by construction, always requires the `index` in its structure.
-This makes it less appealing for spatio-temporal data since the output of calculated spatio-specific variables (i.e.
-features of each series) don't have the time dimension.
-Analysts will either need to have an additional step to join this output to the original tsibble or operate with variables stored in two separate objects.
-In addition, the long form structure of a tsibble object means spatio variables (i.e.
-longitude, latitude, and features of each series if joined back to the tsibble) of each spatio identifier will be repetitively recorded at each timestamp.
-This repetition is unnecessary and would inflate the object size for long series.
--->
-
-<!-- # Create a cubble -->
-
-<!-- The creation of a cubble requires the site identifier (`key`), as well as the spatial (`coords`) and temporal (`index`) identifier.
-`climate_flat` is already a tibble and it uses `id` to identify each station, `date` as the time identifier, and `c(long, lat)` as the spatial identifier.
-To create a cubble for this data, use: -->
-
-<!-- ```{r} -->
-<!-- climate_flat %>% as_cubble(key = id, index = date, coords = c(long, lat)) -->
-<!-- ``` -->
-
-<!-- Most of the time, spatio-temporal data doesn't come into this form and analysts need to query the climate variables based on station metadata.
-\textcolor{red}{This is also a problem illustrated in Section 3.5 in @tidydata.
-Here we provide a structured way to query this data based on the row-wise operator and nested list.} For this type of task, one can structure a metadata into a tibble and use row-wise operator to query the climate variables into a nested list.
-As an example here we demonstrate the workflow to find the 5 closest stations to Melbourne.
-We first create a station data frame with the 5 target stations.
--->
-
-<!-- ```{r echo = FALSE} -->
-<!-- latlon_ref <- tibble::tibble(id = "melbourne", lat = -37.8136, long = 144.9631) -->
-
-<!-- cand <- aus_climate %>% -->
-<!--   as_tibble() %>% -->
-<!--   select(id, lat, long, elev, name, wmo_id) -->
-
-<!-- calc_dist <- function(df_ref, df, n = 5) { -->
-<!--   long_ref <- df_ref$long -->
-<!--   lat_ref <- df_ref$lat -->
-
-<!--   df %>% -->
-<!--     mutate( -->
-<!--       long_ref = long_ref, -->
-<!--       lat_ref = lat_ref, -->
-<!--       dist = rnoaa::meteo_spherical_distance(lat, long, lat_ref, long_ref) -->
-<!--     ) %>% -->
-<!--     slice_min(dist, n = n) %>% -->
-<!--     mutate(city = df_ref$id) %>% -->
-<!--     select(-long_ref, -lat_ref) -->
-<!-- } -->
-
-<!-- (stations <- do.call("rbind", purrr::map( -->
-<!--   1:nrow(latlon_ref), -->
-<!--   ~ calc_dist(latlon_ref[.x, ], cand) -->
-<!-- ))) -->
-<!-- ``` -->
-
-<!-- We can query the climate information into a nested list named `ts` for each station with the `rowwise()` operator. To create a cubble, supply the same identifiers as with the first example. -->
-
-<!-- ```{r eval = FALSE} -->
-<!-- sydmel_climate <- stations %>% -->
-<!--   rowwise() %>% -->
-<!--   mutate(ts = list(meteo_pull_monitors(id, -->
-<!--     date_min = "2020-01-01", -->
-<!--     date_max = "2020-12-31", -->
-<!--     var = c("PRCP", "TMAX", "TMIN") -->
-<!--   ) %>% -->
-<!--     select(-id))) %>% -->
-<!--   as_cubble(key = id, index = date, coords = c(long, lat)) -->
-<!-- ``` -->
-
-<!-- ```{r echo = FALSE} -->
-<!-- (sydmel_climate <- stations %>% -->
-<!--   left_join(aus_climate) %>% -->
-<!--   as_cubble(key = id, index = date, coords = c(long, lat))) -->
-<!-- ``` -->
-
-<!-- Below are the how the nested and long form look like for Australia climate data, which records daily precipitation, maximum and minimum temperature for 55 stations across Australia from 2015- 2020.
-Notice that each station forms a group in both forms and specifically, the nested and long form have a underlying `rowwise_df` and `grouped_df` respectively.
--->
-
-<!-- With a cubic framework on mind, different types of manipulation with cubble can be thought of as slicing the cube in various way.
-The table below shows how some `dplyr` verbs are mapped into the operation in a cubble.
-With the existing grouping on the station, additional grouping can be added with `group_by` and removed with `ungrouped`.
-[talk about why it is useful] -->
-
-<!-- \newpage -->
-
-<!-- ## Cubble operations -->
-
-<!-- ```{r cubble-operations, echo = FALSE, fig.cap="Cubble operations", out.height="50%", out.width="90%", fig.align='center'} -->
-<!-- knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.004.png")) -->
-<!-- ``` -->

### 2022-01-07 5-conclusion.Rmd
play aroudnd with child rmd to organising sections

--- 2021-12-20 main.Rmd
@@
 ---
-documentclass: jss
-author:
-  - name: H. Sherry Zhang
-    # department: Department of Econometrics and Business Statistics
-    affiliation: Monash University
-    #location: Melbourne, Australia
-    email: \email{huize.zhang@monash.edu}
-  - name: Dianne Cook
-    # department: Department of Econometrics and Business Statistics
-    affiliation: 'Monash University \AND'
-    # location: Melbourne, Australia
-    # email:  dicook@monash.edu
-  - name: Ursula Laa
-    # department: Institute of Statistics
-    affiliation: 'University of Natural Resources and Life Sciences \AND'
-    # location: Vienna, Austria
-    # email:  ursula.laa@boku.ac.at
-  - name: Nicolas Langrené
-    # department: 34 Village Street, Docklands VIC 3008 Australia
-    affiliation: 'CSIRO Data61 \AND'
-    # location: Melbourne, Australia
-    # email: nicolas.langrene@csiro.au
-  - name: Patricia Menéndez
-    # department: Department of Econometrics and Business Statistics
-    affiliation: 'Monash University \AND'
-    # location: Melbourne, Australia
-    # email:  patricia.menendez@monash.edu
-title:
-  formatted: "\\pkg{cubble}: An R Package for Structuring Spatio-temporal Data"
-  # If you use tex in the formatted title, alsoge supply version without
-  plain:     "cubble: An R Package for Structuring Spatio-temporal Data"
-  # For running headers, if needed
-  # short:     "\\pkg{foo}: A Capitalized Title"
-abstract: >
-  The abstract of the article.
-keywords:
-  # at least one keyword must be supplied
-  formatted: [spatio-temporal data,  "\\proglang{R}"]
-  plain:     [spatio-temporal data, R]
-preamble: >
-  \usepackage{amsmath}
-  \usepackage{array}
-output: rticles::jss_article
-bibliography: references.bib
+title: "5-conclusion"
+output: pdf_document
 ---
-
-```{r setup, echo = FALSE}
-knitr::opts_chunk$set(
-  echo = FALSE,
-  warning = FALSE,
-  message = FALSE,
-  cache = TRUE,
-  fig.path = "figures/",
-  fig.height = 10
-)
-options(prompt = "R> ", continue = "+ ", 
-        tibble.print_max = 5, tibble.print_min = 5)
-```
-
-```{r echo = FALSE}
-library(cubble)
-library(forcats)
-library(stringr)
-library(dplyr)
-library(lubridate)
-library(ggplot2)
-library(GGally)
-library(plotly)
-library(crosstalk)
-library(patchwork)
-```
-
-\newpage
-
-# Introduction
-
-Many data structures have been proposed for spatial (\pkg{sf} by @sf) and temporal (\pkg{tsibble} by @tsibble) data in the R community, while less has been done for spatio-temporal data.
-The lack of such tools could potentially because analysts usually treat the spatial and temporal dimension of the data separately, without realising the need to create a new data structure.
-While this approach follows the third tidy data principal [@tidydata] (*Each type of observational unit forms a table*), analysts always need to manually join results from different observational units or combining multiple tables into one for downstream analysis.
-This additional step doesn't add new operations into the data but can be error prone.
-\newline
-
-Currently, available spatio-temporal data structure in R includes: \pkg{spacetime} [@spacetime], which proposes four space-time layouts: Full grid (STF), sparse grid(STS), irregular (STI), and trajectory (STT).
-The data structure it uses is based on \pkg{sp} [@sp] and \pkg{xts} [@xts], both of which has been replaced by more recent implementations.
-\pkg{spatstat} [@spatstat] implements a `ppp` class for point pattern data; and more recent, \pkg{stars} [@stars] implements a spatio-temporal array with the dplyr's data cube structure \pkg{cubelyr} [@cubelyr] as its backend.
-While these implementations either store spatial and temporal variables all in a single table, hence duplicate the spatial variables for each temporal unit; or split them into two separate tables that has the problem of manually joining, mentioned in the previously.
-None of these packages enjoy both the benefits of being able to separate manipulation in the two dimensions while also keep the data object as a whole.
-This create a gap in the software development.
-The requirement for such a tool is important given the ubiquity of spatio-temporal vector data in the wild: the Ireland wind data from \pkg{gstat} is an classic example data that splits variables into spatial (`wind.loc`) and temporal (`wind`) dimension; Bureau of Meteorology (BoM) provides climate observations that are widely applied in agriculture and ecology study; air pollution data.
-\newline
-
-This paper describes the implementation of a new spatio-temporal data structure: \pkg{cubble}.
-\pkg{cubble} implements a relational data structure that uses two forms to manage the switch between spatial and temporal dimension.
-With this structure, users can manipulate the spatial or temporal dimension separately, while leaves the linking of two dimensions to \pkg{cubble}.
-The software is available from the Comprehensive R Archive Network (CRAN) at [CRAN link].
-\newline
-
-The rest of the paper will be divided as follows: [complete when the paper structure is more solid]
-
-<!-- Section 2 reviews the existing data structure for spatio, temporal, and spatio-temporal data.
-Section 3 presents a new data structure for spatio-temporal data: cubble.
-Then the paper introduces the workflow of data manipulation and visualisation with the cubble structure in Section 4.
-Section 5 gives some examples on how common spatial and temporal manipulations are performed with cubble and how static and interactive visualisation help to understand climate and [...] data.
-\newline -->
-
-\newpage
-
-# The cubble package
-
-Spatio-temporal data usually come in various forms and Figure \ref{fig:illu-input} shows four examples of this.
-No matter which form the data is in, these formats share some common components that characterise spatio-temporal data.
-A spatial identifier (`id` in the diagram) is the unique identifier of each site.
-The temporal identifier (`t` in the diagram) prescribes the time stamp each site is recorded.
-Coordinates, comprising of latitude and longitude (`lon` and `lat` in the diagram), locates each site on the map.
-These identifiers will be the building blocks for the data structure introduced below.
-Other variables in the data can be categorised into two groups: spatial variables that are invariant at each time stamp for every site, i.e.
-the name or code of the weather station and temporal variables that varies with time.
-
-```{r illu-input, echo = FALSE, out.width = "100%", fig.cap="Illustration of incoming data formats for spatio-temporal data.
-(1) Data comes in as a single table; (2) Separate tables for spatial and temporal variables; (3) A single table with all the parameters used to query the database and a separate table for queried data; and (4) Cubical data in array or NetCDF format.", fig.align='center', out.height = "15%"}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.001.png"))
-```
-
-In a cubble, there are two forms: nested form and long form, and Figure \ref{fig:illu-cubble} sketches the two forms along with the associated attributes.
-The decision on which form to use is output-oriented, meaning analysts need to first think about whether the output of a particular operation is identified only by the spatial identifier, or a combination of spatial and temporal identifier.
-The nested cubble is suitable for working with operations that are only identified by site and this type of operation can be a pure manipulation of spatial variables, or a summary of temporal variables by site (i.e.
-the output of counting the number of raining day is only identified by sites and hence should be performed with the nested form).
-Underneath the nested form, a cubble is built from a row-wise dataframe (`rowwise_df`) where each site forms a separate group.
-This structure simplifies the calculation that involves temporal variables by avoiding the use of `map` syntax when working with list-column.
-
-For those operations whose output involves both a spatial and temporal dimension, long form should be used.
-The long form is identified by both the spatial and temporal identifier and adopts a grouped dataframe (`grouped_df`) to forms each site as a group.
-Spatial variables are stored separately in a \pkg{tibble} as an special attribute of the long cubble.
-This design avoids repeating the spatial variables at each time stamp while not dropping information from spatial variables.
-
-```{r illu-cubble, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "Illustration of nested and long cubble."}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.002.png"))
-```
-
-## Create a cubble in the nested form
-
-To use functionalities from cubble, data analysts first need to create a cubble.
-`as_cubble` create a \pkg{cubble} by supplying the three key components: `key` as the spatial identifier; `index` as the temporal identifier; and a vector of `coords` in the order of longitude first and then latitude.
-The naming of `key` and `index` follows the convention in the \pkg{tsibble} package.
-The cubble created by default is in the nested form.
-Below is an example of creating a cubble: \newline
-
-```{r}
-(cubble_nested <- cubble::climate_flat %>%
-  as_cubble(key = id, index = date, coords = c("long", "lat")))
-```
-
-There are a few information in the \pkg{cubble} header: the name of the `key` variable, bbox, and also the name of variable nested in the `ts` column.
-In this example, each site is identifier is `id` and the number in the bracket means there are 5 unique `id` in this dataset.
-The bbox in the second row gives the range of the coordinates.
-The temporal variables are all nested in the `ts` column, but it could be useful to know the name these variables.
-The third row in the cubble header shows these names and in this example this includes: precipitation, `prcp`, maximum temperature, `tmax`, and minimum temperature, `tmin`.
-
-## Stretch a nested cubble into the long form
-
-The long cubble is suitable to manipulate the time dimension of the data.
-The function `stretch()` switches the nested cubble into the long cubble by first extracts all the spatial variables into a separate tibble and store in the `spatial` attribute and then unnests the `ts` column:
-
-```{r}
-(cubble_long <- cubble_nested %>% stretch(ts))
-```
-
-Notice here that the third line in the header now shows the name of spatial variables rather than the temporal variables.
-
-## Tamp a long cubble back to the nested form
-
-Manipulation on the spatial and temporal dimension can be an iterative process.
-Many times, analysts will need to go back and forth between the nested and long cubble.
-The `stretch()` function introduced in the previous section switches a nested cubble into a long cubble and function `tamp()` is its inverse function to switch a long cubble back to the nested cubble:
-
-```{r}
-(cubble_back <- cubble_long %>% tamp())
-```
-
-## Migrate spatial variables to a long cubble
-
-As a final data output for modelling or visualisation, spatio-temporal data is usually expected to be in a single table.
-Function `migrate()` moves the spatial variables from the `spatial` attribute into the long cubble:
-
-```{r}
-(cubble_long %>% migrate(long, lat))
-```
-
-In this workflow described above, data objects come into cubble in the nested form, then various operations on the spatial and temporal dimension can go back and forth between the nested and long form, and finally, the data will come out of cubble in the long form for further modelling or visualisation.
-
-Building from an underlying `tbl_df` structure, it is natural to implement methods available in `dplyr` to `cubble`. Supported methods in the `cubble` with `dplyr` generics includes:
-
-```{=tex}
-\begin{center}
-\begin{tabular}{ | m{5em} | m{15cm}| } 
-basics & \textbf{mutate}, \textbf{filter}, \textbf{summarise}, \textbf{select}, \textbf{arrange}, \textbf{rename}, \textbf{left\_join} \\
-grouping &  \textbf{group\_by}, \textbf{ungroup}\\
-slice family & \textbf{slice\_head}, \textbf{slice\_tail}, \textbf{slice\_sample}, \textbf{slice\_min} and \textbf{slice\_max} \\
-\end{tabular}
-\end{center}
-```
-\pkg{cubble} is also compatible with \pkg{tsibble} in the sense that the original list-column can be a `tbl_ts` object.
-Duplicates and gaps should be first checked before structuring the data into a cubble.
-If the input data is a \pkg{tsibble} object, the long form cubble is also a \pkg{tsibble} where users can directly apply time series operations.
-
-# Advanced features/ considerations
-
-## Hierarchical structure
-
-Imposing a clustering structure can be thought of as building a hierarchical structure where stations are nested within clusters.
-This can be useful when there's intrinsic nesting structure in the data (i.e.
-country nested in the continent, county nested in the state) or there's some clustering.
-When we have access to both level, `switch_key()` is the function to re-structure the data as one cluster per row.
-Temporal observations from different stations while within the same cluster are bound in the nested column `ts`.
-
-One thing we hope to do with the cluster is to find the coordinates of the centroid.
-These are variables variant to the station but invariant to the cluster and it would be nice to have a function that structure each cluster as a row.
-`switch_key()` is the function that does this: it lets you to specify a new key, say `cluster` and nests all spatial variables variant to `cluster` into a column.
-Spatial variables are all nested inside a new column `.val`.
-Temporal observations from different stations while within the same cluster are bound in the nested column `ts`.
-
-This structure makes it easy to compute cluster level variable, for example, the convex hull and the centroid coordinate of each cluster.
-These can be amended into the nested form with function `get_centroid`.
-
-After we have got `cluster_nested`, spatial and temporal data at both levels can be easily obtained.
-Figure \ref{fig:illu-hier} illustrate the relationship between the long and nested form cubble at both site and cluster level.
-More description on this.
-Start with the original `station_nested`, `stretch()` expands the `ts` column with each station (`id`) forming a group and attach variables invariant to `id` as an attribute.
-`switch_key()` changes the `key` from `id` to `cluster` and nests all the spatial variables that variant to `cluster`.
-`stretch()` `cluster_nested` will store variables that are invariant to `cluster` as a tibble in the attribute.
-
-this fit into the remaining data pipeline.
-
-```{r illu-hier, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "Hierarchical structure"}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.003.png"))
-```
-
-## Data fusion and matching
-
-Temporal matching checks how spatially matched pairs align temporally. We use the following chart to illustrate how the temporal matching works:
-
-For each spatially matched pair, say `A` and `a`, we first find the largest `n` points in each series, colored in brown points here.
-Here we use the largest three but you can tune this number by `temporal_n_highest`.
-Then we construct the interval of the largest points from one series and see how many points, from the other series, fall into the intervals.
-The series used to construct the interval is controlled by `temporal_independent` and the window size by `temporal_window` with a default of 5.
-
-In this illustration, we construct the interval based on series `A` and two of the three peaks from `a` falls into this interval at Time 7 and 27.
-
-There's another mandatory argument that hasn't been introduced above: `temporal_var_to_match`.
-This argument controls the variable to match and it needs to appear in both the `major` and `minor` set.
-In the water level matching example, we match the variable `Water_course_level` from `river` to `prcp` from `climate`, hence need to manually rename one of them to match the other, here we rename `Water_course_level` to `prcp` in `river`:
-
-Figure \ref{fig:illu-matching}
-
-```{r illu-matching, out.width="100%", out.height="30%", fig.cap="sdfasdf"}
-knitr::include_graphics("figures/illu-matching.png")
-```
-
-## Interactive graphics
-
-<!-- Interactive graphics can listen to users' actions on the plot to provide additional information that facilitates data exploration.
-This is a useful technique for spatio-temporal data since users can zoom or pan the map to view the local and global structure of the map; use tooltips or popups to query more information about a graphic element; or highlight points to explore its linked views in other plots.
-In the R community, many implementations have been developed to connect \proglang{R} to \proglang{javascript} to create interactive graphics.
-In relation to spatio, temporal, and spatio-temporal data, the general purpose packages \pkg{plotly} [@plotly] and \pkg{leaflet} [@leaflet] realise various interactive actions through their corresponding javascript libraries.
-\pkg{crosstalk} [@crosstalk] and \pkg{tsibbletalk} [@tsibbletalk] implement brushed linking between htmlwidgets.
-\pkg{ggiraph} [@ggiraph] enables tooltip, self-linking, and customised actions specified through its own \proglang{javascript}.
--->
-
-<!-- While many graphic implementations present worked examples to illustrate the usage of the package, few documents the underlying pipeline that transforms the raw data step-by-step to the final view on the screen.
-There have been some early work in building the data pipeline for (interactive) graphics [@buja1988elements; @buja1996interactive;  @sutherland2000orca] and more recent discussions include @wickham2009plumbing, @xie2014reactive, and @cheng2016enabling.
--->
-
-The cubble structure fits in naturally with the interactive graphic pipeline discussed in the literature [@buja1988elements; @buja1996interactive;  @sutherland2000orca; @xie2014reactive; @cheng2016enabling].
-Diagram \ref{fig:illu-interactive} illustrates how linking works with the two forms in a cubble, where a time series plot is created with the long cubble and a map is created with the nested cubble.
-When a user action is captured from the map, the site will be activated in the nested cubble.
-Then, the nested cubble will communicate to the long cubble to activate all the observations with the same `id`.
-The long cubble will then highlight the  activated series in the time series plot.
-
-The linking is also available from the time series plot to the map.
-The selection on the time series is through selecting the point on the time series and once a point is selected, it will be activated in the long cubble.
-All the observations that share the same `id`,  either in the long and nested cubble, are then activated.
-This includes other points in the same time series in the long cubble and the corresponding observation of site in the nested cubble.
-These activated observations will then being reflected in the updated plots and Diagram \ref{fig:illu-interactive-2} in the Appendix illustrates this process.
-
-```{r illu-interactive, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "demon interactivity"}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.004.png"))
-```
-
-\newpage
-
-# Examples
-
-## Australia historical maximum temperature
-
-Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world and the dataset `weatherdata::historical_tmax` extracts the historical maximum temperature recorded for `r nrow(weatherdata::historical_tmax)` Australian stations.
-The data `historical_tmax` is already presented as a cubble, with `id` as the key, `date` as the index, and `c("longitude", "latitude")` as the coordinates.
-Other variables include `elevation`, `name`, `wmo_id`, `first_year`, and  `last_year` in the nested form and `tmax` in the long form.
-This example compares the maximum temperature in two periods: 1971 - 1975 and 2016 - 2020 for stations in Victoria and New South Wales.
-
-Stations in the two states can be subsetted on the station number: Australia GHCN station number starts with `ASN00` and followed by the [Bureau of Meteorology (BOM) station number](http://www.bom.gov.au/climate/cdo/about/site-num.shtml), where the 2nd and 3rd digit (7th and 8th in the GHCN number) denote the state a station belongs to.
-New South Wales stations start from 46 to 75 and Victoria stations then follow from 76 to 90.
-Extracting Victoria and New South Wales stations is a filter operation in the spatial dimension and hence is operated in the nested form:
-
-```{r eval = FALSE, echo = TRUE}
-tmax <- weatherdata::historical_tmax %>%
-  filter(between(stringr::str_sub(id, 7, 8), 46, 90))
-```
-
-The five year window is chosen to remove the effect of a particular year and the historical period of 1971 - 1975 is used since all the stations have records from 1970.
-This following chunk filters on records in the two study periods and summarises the maximum temperature into monthly measure.
-`stretch()` is used to convert the nested form cubble into the long form for these operations:
-
-```{r eval=  FALSE, echo = TRUE}
-tmax <- tmax %>% 
-  stretch() %>%
-  filter(lubridate::year(date) %in% c(1971:1975, 2016:2020)) %>%
-  mutate(month = lubridate::month(date), 
-         group = as.factor(ifelse(lubridate::year(date) > 2015, 
-                                  "2016 ~ 2020", "1971 ~ 1975"))) %>%
-  group_by(month, group) %>%
-  summarise(tmax = mean(tmax, na.rm = TRUE))
-```
-
-A data quality issue with GHCN data is that while the first and last year of each series is provided, years missing in this period is not reported.
-There are a few stations which don't have records during 1971 - 1975 and these stations are filtered out by examining whether the summarised `tmax` has a total of 24 months.
-This is again a station-wise operation and is operated in the nested form, which is switched to from the long form by  `tamp()`:
-
-```{r eval = FALSE, echo = TRUE}
-tmax <- tmax %>% 
-  tamp() %>%
-  filter(nrow(ts) == 24) 
-```
-
-Lastly, to create a glyph map, both the major (`longitude`, `latitude`) and minor (`month`, `tmax`) coordinates need to be in the same table.
-Spatial variables can be moved to the long form with `migrate()`:
-
-```{r eval = FALSE, echo = TRUE}  
-tmax <- tmax %>%   
-  stretch() %>%
-  migrate(latitude, longitude)
-```
-
-Figure \ref{fig:basic-manip} shows the glyph map made with the data after the wangling above.
-One issue with this map is that the similar pattern shown from a few nearby stations around Sydney (151E, 34S) and New castle (152E, 33S) can be distracting.
-Aggregation is a useful technique to observe the general pattern of a collection of series and will be the topic of the next example.
-
-```{r basic-manip, out.width="100%", out.height="70%", fig.cap="Glyph map of the mean maximum temperature by month for Victoria and New South Wales weather stations.
-On the top left corner is a detailed legend for station Cobar highlighted in the black box.
-Compared to 1971 - 1975,  the period 2016 - 2020 sees an increase of mean maximum temperature in spring and summer in Australia (end and beginning of the year).
-A larger increase during the first three months of the year is observed at stations with a lower latitude."}
-knitr::include_graphics("figures/basic-manip.png")
-```
-
-## Australia precipitation pattern in 2020
-
-In Figure \ref{fig:basic-agg}.
-Aggregating them into averages or other statistics would give a better picture.
-This can be done with cubble through a hierarchical structure, where the cluster indicator is one level above the individual station.
-
-The `weatherdata::climate_full` data has daily climate data of `r nrow(weatherdata::climate_full)` Australia stations from 2016 to 2020.
-
-calcluate distance matrix
-As an example to illustrate here, a kmean clustering algorithm based on the distance matrix is used and the number of centres is set to 20.
-More complex algorithms can also be used for more complex problem, as long as a mapping from each station id to the cluster id can be constructed.
-
-<!-- Now we can obtain the aggregated series as described in the workflow diagram above and construct the glyph map with `GGally::glyphs()`: -->
-
-<!-- We can also look at the precipitation of each individual station within the same cluster: -->
-
-<!-- Lastly, there is one series on Tasmania island standing out from others, lets look at where it is: -->
-
-```{r basic-agg, out.width="100%", out.height="90%", fig.cap="sdfasdf"}
-knitr::include_graphics("figures/basic-agg.png")
-```
-
-## River level data in Victria water gauges
-
-The water level data comes from [Bureau of Meteorology](http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full) and has a copy in `weatherdata`.
-Here we extract the water course level and add a column annotate this data of type `river`.
-For the rainfall data, we will still use the `weatherdata::climate_full`, filtering for Victorian stations in 2020 should be pretty familiar by now.
-Again, we first look at where these stations are on the map first:
-
-Now we use `match_sites()` to first pair the weather stations with the river gauges spatially and then apply the temporal matching on `prcp`.
-We will construct the interval based on peaks in `climate` since we would expect a lag effect for precipitation to flow into the river and cause a raise in river level, hence `temporal_independent = climate`.
-We select the 30 highest peak from the series to construct the match by setting `temporal_n_highest = 30`.
-This is a tuning parameter and you can start with 10% of the points of one series (here we have daily data for a year, 10% is roughly 30 points).
-`temporal_min_match` filters out pairs don't have enough match and to return all the pairs, set `temporal_min_match` to `0`.``
-
-The output from temporal matching is also a cubble, with additional column `.dist` and `.group` inherent from spatial matching and `n_match` for the number of matched temporal peaks.
-Then you can use this output to plot the location of match or to look at the series:
-
-```{r matching, out.width="100%", out.height="30%", fig.cap="asdfasd"}
-knitr::include_graphics("figures/matching.png")
-```
-
-## ERA5: climate reanalysis data
-
-\newpage
-
-## Interative graphic with cubble
-
-With a larger dataset, interactive graphics can be helpful to explore these data from multiple perspective.
-Two examples are presented here to show how interactivity can be used to explore the climate data across Australia.
-Again, `weatherdata::climate_full` will be used and we hope to include more than one time variable here.
-The difference between the maximum and minimum temperature of the day, the diurnal range, can be directly calculated from the data and it can have large impact on the wine industry (lol this is going far).
-This would be useful to compare the temperature difference across stations and to compare how this temperature difference differs throughout the year.
-
-Temperature difference is a time-wise variable and should be computed in the long form cubble. The variables are averaged over the 2016 to 2020 period by month to stabilise the result.
-
-```{r eval = FALSE, echo = TRUE}
-res <- weatherdata::climate_full %>%
-  # filter(!id %in% c("ASN00067033", "ASN00072091", 
-  #                   "ASN00059040", "ASN00097053")) %>%
-  stretch() %>%
-  mutate(month = lubridate::month(date),
-         diff = tmax - tmin) %>%
-  group_by(month) %>%
-  summarise(tmax = mean(tmax, na.rm = TRUE),
-            diff = mean(diff, na.rm = TRUE)) %>%
-  ungroup(month)
-
-```
-
-Several variables can be useful when comparing the seasonal effect of the temperature differences, for example, the difference of average temperature difference between summer and winter can tell [...].
-This is a station-wise measure and hence computed in the nested form.
-
-```{r eval = FALSE, echo = TRUE}
-res <- res %>% 
-  tamp() %>%
-  mutate(
-    diff_winter = ts %>% filter(month %in% c(6, 7, 8)) %>%
-      pull(diff) %>% mean(na.rm = TRUE),
-    diff_summer = ts %>% filter(month %in% c(12, 1, 2)) %>%
-      pull(diff) %>% mean(na.rm = TRUE),
-    diff_year = diff_winter - diff_summer
-  )
-```
-
-The result from the above manipulation can be directly supplied to make interactive plots and Figure \ref{fig:interactive-linking} is an example made with \pkg{plotly} and \pkg{crosstalk}.
-For example, the second row results from selecting the highest mean maximum temperature in July and this corresponds to Kalumburu where there is a larger winter (June to August) temperature difference than summer (December to February).
-The third row selects one of the stations (Birdsville Airport) with a large average temperature difference in the interior desert region on the map and the linked view from the time series shows that this maximum and minimum temperature difference is constant throughout the year.
-The last row selects the lowest maximum temperature in July on the temperature band and surprisingly, the corresponding station doesn't fall into the Tasmania island but Mountain Hotham in Victoria.
-
-```{r interactive-linking, echo = FALSE, out.width="100%", out.height="25%", fig.cap = "Screenshots of the mean daily temperature difference averaged by month from 2016 to 2020 at various locations in Australia.
-Crossed linking between the map and the temperature band is implemented in both directions for exploring the temperature pattern interactively.
-The first row is the general appearance of the graph and the next three rows are examples from various selection to show different temperature patterns in Australia.
-The html version can be found at ...", fig.show='hold'}
-knitr::include_graphics(here::here("figures/linking.png"))
-knitr::include_graphics(here::here("figures/linking-north.png"))
-knitr::include_graphics(here::here("figures/linking-mid.png"))
-knitr::include_graphics(here::here("figures/linking-south.png"))
-```
-
-Cubble is also compatible with \pkg{leaflet} and with leaflet extension package \pkg{leafpop}, the temperature band can be displayed upon clicking on the station on the map.
-Figure \ref{fig:interactive-popup} shows more examples of temperature pattern that can be found in Australia.
-With the same latitude as Kalumburu but in the Pacific coastline, Cooktown airport shows a similarly high but constant temperature throughout the year.
-With a slightly higher latitude than the mountain Hotham, flat land in Melbourne airport shows a less extreme temperature band.
-Lastly, the Cape Grim Baps, in the north-west tip of the Tasmania island shows a uniformly low and thin temperature band.
-
-```{r interactive-popup, echo = FALSE, out.width="50%", out.height="30%", fig.cap = "Screenshots of the diurnal temperature range averaged by month from 2016 to 2020 at various locations in Australia.
-The leaflet popup allows users to click on a single station for its temperature band as a small plot in the popup window.
-The html version can be found at ...", fig.show='hold'}
-knitr::include_graphics(here::here("figures/popup-north.png"))
-knitr::include_graphics(here::here("figures/popup-vic.png"))
-knitr::include_graphics(here::here("figures/popup-tas.png"))
-```
-
 # Conclusion
@@
 ```
-
-\newpage
-
-<!-- # Old stuff -->
-
-<!-- In the temporal aspect, the `tsibble` [@tsibbles] structure and its tidyverts ecosystem have provided a [...
-] workflow to work with temporal data.
-In a tsibble structure, temporal data is characterised by `index` and `key` where `index` is the temporal identifier and `key` is the identifier for multiple series, which could be used as a spatio identifier.
-However, a tsibble object, by construction, always requires the `index` in its structure.
-This makes it less appealing for spatio-temporal data since the output of calculated spatio-specific variables (i.e.
-features of each series) don't have the time dimension.
-Analysts will either need to have an additional step to join this output to the original tsibble or operate with variables stored in two separate objects.
-In addition, the long form structure of a tsibble object means spatio variables (i.e.
-longitude, latitude, and features of each series if joined back to the tsibble) of each spatio identifier will be repetitively recorded at each timestamp.
-This repetition is unnecessary and would inflate the object size for long series.
--->
-
-<!-- # Create a cubble -->
-
-<!-- The creation of a cubble requires the site identifier (`key`), as well as the spatial (`coords`) and temporal (`index`) identifier.
-`climate_flat` is already a tibble and it uses `id` to identify each station, `date` as the time identifier, and `c(long, lat)` as the spatial identifier.
-To create a cubble for this data, use: -->
-
-<!-- ```{r} -->
-<!-- climate_flat %>% as_cubble(key = id, index = date, coords = c(long, lat)) -->
-<!-- ``` -->
-
-<!-- Most of the time, spatio-temporal data doesn't come into this form and analysts need to query the climate variables based on station metadata.
-\textcolor{red}{This is also a problem illustrated in Section 3.5 in @tidydata.
-Here we provide a structured way to query this data based on the row-wise operator and nested list.} For this type of task, one can structure a metadata into a tibble and use row-wise operator to query the climate variables into a nested list.
-As an example here we demonstrate the workflow to find the 5 closest stations to Melbourne.
-We first create a station data frame with the 5 target stations.
--->
-
-<!-- ```{r echo = FALSE} -->
-<!-- latlon_ref <- tibble::tibble(id = "melbourne", lat = -37.8136, long = 144.9631) -->
-
-<!-- cand <- aus_climate %>% -->
-<!--   as_tibble() %>% -->
-<!--   select(id, lat, long, elev, name, wmo_id) -->
-
-<!-- calc_dist <- function(df_ref, df, n = 5) { -->
-<!--   long_ref <- df_ref$long -->
-<!--   lat_ref <- df_ref$lat -->
-
-<!--   df %>% -->
-<!--     mutate( -->
-<!--       long_ref = long_ref, -->
-<!--       lat_ref = lat_ref, -->
-<!--       dist = rnoaa::meteo_spherical_distance(lat, long, lat_ref, long_ref) -->
-<!--     ) %>% -->
-<!--     slice_min(dist, n = n) %>% -->
-<!--     mutate(city = df_ref$id) %>% -->
-<!--     select(-long_ref, -lat_ref) -->
-<!-- } -->
-
-<!-- (stations <- do.call("rbind", purrr::map( -->
-<!--   1:nrow(latlon_ref), -->
-<!--   ~ calc_dist(latlon_ref[.x, ], cand) -->
-<!-- ))) -->
-<!-- ``` -->
-
-<!-- We can query the climate information into a nested list named `ts` for each station with the `rowwise()` operator. To create a cubble, supply the same identifiers as with the first example. -->
-
-<!-- ```{r eval = FALSE} -->
-<!-- sydmel_climate <- stations %>% -->
-<!--   rowwise() %>% -->
-<!--   mutate(ts = list(meteo_pull_monitors(id, -->
-<!--     date_min = "2020-01-01", -->
-<!--     date_max = "2020-12-31", -->
-<!--     var = c("PRCP", "TMAX", "TMIN") -->
-<!--   ) %>% -->
-<!--     select(-id))) %>% -->
-<!--   as_cubble(key = id, index = date, coords = c(long, lat)) -->
-<!-- ``` -->
-
-<!-- ```{r echo = FALSE} -->
-<!-- (sydmel_climate <- stations %>% -->
-<!--   left_join(aus_climate) %>% -->
-<!--   as_cubble(key = id, index = date, coords = c(long, lat))) -->
-<!-- ``` -->
-
-<!-- Below are the how the nested and long form look like for Australia climate data, which records daily precipitation, maximum and minimum temperature for 55 stations across Australia from 2015- 2020.
-Notice that each station forms a group in both forms and specifically, the nested and long form have a underlying `rowwise_df` and `grouped_df` respectively.
--->
-
-<!-- With a cubic framework on mind, different types of manipulation with cubble can be thought of as slicing the cube in various way.
-The table below shows how some `dplyr` verbs are mapped into the operation in a cubble.
-With the existing grouping on the station, additional grouping can be added with `group_by` and removed with `ungrouped`.
-[talk about why it is useful] -->
-
-<!-- \newpage -->
-
-<!-- ## Cubble operations -->
-
-<!-- ```{r cubble-operations, echo = FALSE, fig.cap="Cubble operations", out.height="50%", out.width="90%", fig.align='center'} -->
-<!-- knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.004.png")) -->
-<!-- ``` -->

### 2022-01-07 main.Rmd
play aroudnd with child rmd to organising sections

--- 2021-12-20 main.Rmd
@@
 
-# Introduction
+```{r child = "1-intro.Rmd"}
 
-Many data structures have been proposed for spatial (\pkg{sf} by @sf) and temporal (\pkg{tsibble} by @tsibble) data in the R community, while less has been done for spatio-temporal data.
-The lack of such tools could potentially because analysts usually treat the spatial and temporal dimension of the data separately, without realising the need to create a new data structure.
-While this approach follows the third tidy data principal [@tidydata] (*Each type of observational unit forms a table*), analysts always need to manually join results from different observational units or combining multiple tables into one for downstream analysis.
-This additional step doesn't add new operations into the data but can be error prone.
-\newline
-
-Currently, available spatio-temporal data structure in R includes: \pkg{spacetime} [@spacetime], which proposes four space-time layouts: Full grid (STF), sparse grid(STS), irregular (STI), and trajectory (STT).
-The data structure it uses is based on \pkg{sp} [@sp] and \pkg{xts} [@xts], both of which has been replaced by more recent implementations.
-\pkg{spatstat} [@spatstat] implements a `ppp` class for point pattern data; and more recent, \pkg{stars} [@stars] implements a spatio-temporal array with the dplyr's data cube structure \pkg{cubelyr} [@cubelyr] as its backend.
-While these implementations either store spatial and temporal variables all in a single table, hence duplicate the spatial variables for each temporal unit; or split them into two separate tables that has the problem of manually joining, mentioned in the previously.
-None of these packages enjoy both the benefits of being able to separate manipulation in the two dimensions while also keep the data object as a whole.
-This create a gap in the software development.
-The requirement for such a tool is important given the ubiquity of spatio-temporal vector data in the wild: the Ireland wind data from \pkg{gstat} is an classic example data that splits variables into spatial (`wind.loc`) and temporal (`wind`) dimension; Bureau of Meteorology (BoM) provides climate observations that are widely applied in agriculture and ecology study; air pollution data.
-\newline
-
-This paper describes the implementation of a new spatio-temporal data structure: \pkg{cubble}.
-\pkg{cubble} implements a relational data structure that uses two forms to manage the switch between spatial and temporal dimension.
-With this structure, users can manipulate the spatial or temporal dimension separately, while leaves the linking of two dimensions to \pkg{cubble}.
-The software is available from the Comprehensive R Archive Network (CRAN) at [CRAN link].
-\newline
-
-The rest of the paper will be divided as follows: [complete when the paper structure is more solid]
-
-<!-- Section 2 reviews the existing data structure for spatio, temporal, and spatio-temporal data.
-Section 3 presents a new data structure for spatio-temporal data: cubble.
-Then the paper introduces the workflow of data manipulation and visualisation with the cubble structure in Section 4.
-Section 5 gives some examples on how common spatial and temporal manipulations are performed with cubble and how static and interactive visualisation help to understand climate and [...] data.
-\newline -->
-
+```
 \newpage
 
-# The cubble package
+```{r child = "2-the-cubble-pkg.Rmd"}
 
-Spatio-temporal data usually come in various forms and Figure \ref{fig:illu-input} shows four examples of this.
-No matter which form the data is in, these formats share some common components that characterise spatio-temporal data.
-A spatial identifier (`id` in the diagram) is the unique identifier of each site.
-The temporal identifier (`t` in the diagram) prescribes the time stamp each site is recorded.
-Coordinates, comprising of latitude and longitude (`lon` and `lat` in the diagram), locates each site on the map.
-These identifiers will be the building blocks for the data structure introduced below.
-Other variables in the data can be categorised into two groups: spatial variables that are invariant at each time stamp for every site, i.e.
-the name or code of the weather station and temporal variables that varies with time.
-
-```{r illu-input, echo = FALSE, out.width = "100%", fig.cap="Illustration of incoming data formats for spatio-temporal data.
-(1) Data comes in as a single table; (2) Separate tables for spatial and temporal variables; (3) A single table with all the parameters used to query the database and a separate table for queried data; and (4) Cubical data in array or NetCDF format.", fig.align='center', out.height = "15%"}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.001.png"))
-```
-
-In a cubble, there are two forms: nested form and long form, and Figure \ref{fig:illu-cubble} sketches the two forms along with the associated attributes.
-The decision on which form to use is output-oriented, meaning analysts need to first think about whether the output of a particular operation is identified only by the spatial identifier, or a combination of spatial and temporal identifier.
-The nested cubble is suitable for working with operations that are only identified by site and this type of operation can be a pure manipulation of spatial variables, or a summary of temporal variables by site (i.e.
-the output of counting the number of raining day is only identified by sites and hence should be performed with the nested form).
-Underneath the nested form, a cubble is built from a row-wise dataframe (`rowwise_df`) where each site forms a separate group.
-This structure simplifies the calculation that involves temporal variables by avoiding the use of `map` syntax when working with list-column.
-
-For those operations whose output involves both a spatial and temporal dimension, long form should be used.
-The long form is identified by both the spatial and temporal identifier and adopts a grouped dataframe (`grouped_df`) to forms each site as a group.
-Spatial variables are stored separately in a \pkg{tibble} as an special attribute of the long cubble.
-This design avoids repeating the spatial variables at each time stamp while not dropping information from spatial variables.
-
-```{r illu-cubble, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "Illustration of nested and long cubble."}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.002.png"))
-```
-
-## Create a cubble in the nested form
-
-To use functionalities from cubble, data analysts first need to create a cubble.
-`as_cubble` create a \pkg{cubble} by supplying the three key components: `key` as the spatial identifier; `index` as the temporal identifier; and a vector of `coords` in the order of longitude first and then latitude.
-The naming of `key` and `index` follows the convention in the \pkg{tsibble} package.
-The cubble created by default is in the nested form.
-Below is an example of creating a cubble: \newline
-
-```{r}
-(cubble_nested <- cubble::climate_flat %>%
-  as_cubble(key = id, index = date, coords = c("long", "lat")))
-```
-
-There are a few information in the \pkg{cubble} header: the name of the `key` variable, bbox, and also the name of variable nested in the `ts` column.
-In this example, each site is identifier is `id` and the number in the bracket means there are 5 unique `id` in this dataset.
-The bbox in the second row gives the range of the coordinates.
-The temporal variables are all nested in the `ts` column, but it could be useful to know the name these variables.
-The third row in the cubble header shows these names and in this example this includes: precipitation, `prcp`, maximum temperature, `tmax`, and minimum temperature, `tmin`.
-
-## Stretch a nested cubble into the long form
-
-The long cubble is suitable to manipulate the time dimension of the data.
-The function `stretch()` switches the nested cubble into the long cubble by first extracts all the spatial variables into a separate tibble and store in the `spatial` attribute and then unnests the `ts` column:
-
-```{r}
-(cubble_long <- cubble_nested %>% stretch(ts))
-```
-
-Notice here that the third line in the header now shows the name of spatial variables rather than the temporal variables.
-
-## Tamp a long cubble back to the nested form
-
-Manipulation on the spatial and temporal dimension can be an iterative process.
-Many times, analysts will need to go back and forth between the nested and long cubble.
-The `stretch()` function introduced in the previous section switches a nested cubble into a long cubble and function `tamp()` is its inverse function to switch a long cubble back to the nested cubble:
-
-```{r}
-(cubble_back <- cubble_long %>% tamp())
-```
-
-## Migrate spatial variables to a long cubble
-
-As a final data output for modelling or visualisation, spatio-temporal data is usually expected to be in a single table.
-Function `migrate()` moves the spatial variables from the `spatial` attribute into the long cubble:
-
-```{r}
-(cubble_long %>% migrate(long, lat))
-```
-
-In this workflow described above, data objects come into cubble in the nested form, then various operations on the spatial and temporal dimension can go back and forth between the nested and long form, and finally, the data will come out of cubble in the long form for further modelling or visualisation.
-
-Building from an underlying `tbl_df` structure, it is natural to implement methods available in `dplyr` to `cubble`. Supported methods in the `cubble` with `dplyr` generics includes:
-
-```{=tex}
-\begin{center}
-\begin{tabular}{ | m{5em} | m{15cm}| } 
-basics & \textbf{mutate}, \textbf{filter}, \textbf{summarise}, \textbf{select}, \textbf{arrange}, \textbf{rename}, \textbf{left\_join} \\
-grouping &  \textbf{group\_by}, \textbf{ungroup}\\
-slice family & \textbf{slice\_head}, \textbf{slice\_tail}, \textbf{slice\_sample}, \textbf{slice\_min} and \textbf{slice\_max} \\
-\end{tabular}
-\end{center}
-```
-\pkg{cubble} is also compatible with \pkg{tsibble} in the sense that the original list-column can be a `tbl_ts` object.
-Duplicates and gaps should be first checked before structuring the data into a cubble.
-If the input data is a \pkg{tsibble} object, the long form cubble is also a \pkg{tsibble} where users can directly apply time series operations.
-
-# Advanced features/ considerations
-
-## Hierarchical structure
-
-Imposing a clustering structure can be thought of as building a hierarchical structure where stations are nested within clusters.
-This can be useful when there's intrinsic nesting structure in the data (i.e.
-country nested in the continent, county nested in the state) or there's some clustering.
-When we have access to both level, `switch_key()` is the function to re-structure the data as one cluster per row.
-Temporal observations from different stations while within the same cluster are bound in the nested column `ts`.
-
-One thing we hope to do with the cluster is to find the coordinates of the centroid.
-These are variables variant to the station but invariant to the cluster and it would be nice to have a function that structure each cluster as a row.
-`switch_key()` is the function that does this: it lets you to specify a new key, say `cluster` and nests all spatial variables variant to `cluster` into a column.
-Spatial variables are all nested inside a new column `.val`.
-Temporal observations from different stations while within the same cluster are bound in the nested column `ts`.
-
-This structure makes it easy to compute cluster level variable, for example, the convex hull and the centroid coordinate of each cluster.
-These can be amended into the nested form with function `get_centroid`.
-
-After we have got `cluster_nested`, spatial and temporal data at both levels can be easily obtained.
-Figure \ref{fig:illu-hier} illustrate the relationship between the long and nested form cubble at both site and cluster level.
-More description on this.
-Start with the original `station_nested`, `stretch()` expands the `ts` column with each station (`id`) forming a group and attach variables invariant to `id` as an attribute.
-`switch_key()` changes the `key` from `id` to `cluster` and nests all the spatial variables that variant to `cluster`.
-`stretch()` `cluster_nested` will store variables that are invariant to `cluster` as a tibble in the attribute.
-
-this fit into the remaining data pipeline.
-
-```{r illu-hier, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "Hierarchical structure"}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.003.png"))
-```
-
-## Data fusion and matching
-
-Temporal matching checks how spatially matched pairs align temporally. We use the following chart to illustrate how the temporal matching works:
-
-For each spatially matched pair, say `A` and `a`, we first find the largest `n` points in each series, colored in brown points here.
-Here we use the largest three but you can tune this number by `temporal_n_highest`.
-Then we construct the interval of the largest points from one series and see how many points, from the other series, fall into the intervals.
-The series used to construct the interval is controlled by `temporal_independent` and the window size by `temporal_window` with a default of 5.
-
-In this illustration, we construct the interval based on series `A` and two of the three peaks from `a` falls into this interval at Time 7 and 27.
-
-There's another mandatory argument that hasn't been introduced above: `temporal_var_to_match`.
-This argument controls the variable to match and it needs to appear in both the `major` and `minor` set.
-In the water level matching example, we match the variable `Water_course_level` from `river` to `prcp` from `climate`, hence need to manually rename one of them to match the other, here we rename `Water_course_level` to `prcp` in `river`:
-
-Figure \ref{fig:illu-matching}
-
-```{r illu-matching, out.width="100%", out.height="30%", fig.cap="sdfasdf"}
-knitr::include_graphics("figures/illu-matching.png")
-```
-
-## Interactive graphics
-
-<!-- Interactive graphics can listen to users' actions on the plot to provide additional information that facilitates data exploration.
-This is a useful technique for spatio-temporal data since users can zoom or pan the map to view the local and global structure of the map; use tooltips or popups to query more information about a graphic element; or highlight points to explore its linked views in other plots.
-In the R community, many implementations have been developed to connect \proglang{R} to \proglang{javascript} to create interactive graphics.
-In relation to spatio, temporal, and spatio-temporal data, the general purpose packages \pkg{plotly} [@plotly] and \pkg{leaflet} [@leaflet] realise various interactive actions through their corresponding javascript libraries.
-\pkg{crosstalk} [@crosstalk] and \pkg{tsibbletalk} [@tsibbletalk] implement brushed linking between htmlwidgets.
-\pkg{ggiraph} [@ggiraph] enables tooltip, self-linking, and customised actions specified through its own \proglang{javascript}.
--->
-
-<!-- While many graphic implementations present worked examples to illustrate the usage of the package, few documents the underlying pipeline that transforms the raw data step-by-step to the final view on the screen.
-There have been some early work in building the data pipeline for (interactive) graphics [@buja1988elements; @buja1996interactive;  @sutherland2000orca] and more recent discussions include @wickham2009plumbing, @xie2014reactive, and @cheng2016enabling.
--->
-
-The cubble structure fits in naturally with the interactive graphic pipeline discussed in the literature [@buja1988elements; @buja1996interactive;  @sutherland2000orca; @xie2014reactive; @cheng2016enabling].
-Diagram \ref{fig:illu-interactive} illustrates how linking works with the two forms in a cubble, where a time series plot is created with the long cubble and a map is created with the nested cubble.
-When a user action is captured from the map, the site will be activated in the nested cubble.
-Then, the nested cubble will communicate to the long cubble to activate all the observations with the same `id`.
-The long cubble will then highlight the  activated series in the time series plot.
-
-The linking is also available from the time series plot to the map.
-The selection on the time series is through selecting the point on the time series and once a point is selected, it will be activated in the long cubble.
-All the observations that share the same `id`,  either in the long and nested cubble, are then activated.
-This includes other points in the same time series in the long cubble and the corresponding observation of site in the nested cubble.
-These activated observations will then being reflected in the updated plots and Diagram \ref{fig:illu-interactive-2} in the Appendix illustrates this process.
-
-```{r illu-interactive, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "demon interactivity"}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.004.png"))
 ```
@@
 
-# Examples
+```{r child = "3-advanced.Rmd"}
 
-## Australia historical maximum temperature
-
-Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world and the dataset `weatherdata::historical_tmax` extracts the historical maximum temperature recorded for `r nrow(weatherdata::historical_tmax)` Australian stations.
-The data `historical_tmax` is already presented as a cubble, with `id` as the key, `date` as the index, and `c("longitude", "latitude")` as the coordinates.
-Other variables include `elevation`, `name`, `wmo_id`, `first_year`, and  `last_year` in the nested form and `tmax` in the long form.
-This example compares the maximum temperature in two periods: 1971 - 1975 and 2016 - 2020 for stations in Victoria and New South Wales.
-
-Stations in the two states can be subsetted on the station number: Australia GHCN station number starts with `ASN00` and followed by the [Bureau of Meteorology (BOM) station number](http://www.bom.gov.au/climate/cdo/about/site-num.shtml), where the 2nd and 3rd digit (7th and 8th in the GHCN number) denote the state a station belongs to.
-New South Wales stations start from 46 to 75 and Victoria stations then follow from 76 to 90.
-Extracting Victoria and New South Wales stations is a filter operation in the spatial dimension and hence is operated in the nested form:
-
-```{r eval = FALSE, echo = TRUE}
-tmax <- weatherdata::historical_tmax %>%
-  filter(between(stringr::str_sub(id, 7, 8), 46, 90))
-```
-
-The five year window is chosen to remove the effect of a particular year and the historical period of 1971 - 1975 is used since all the stations have records from 1970.
-This following chunk filters on records in the two study periods and summarises the maximum temperature into monthly measure.
-`stretch()` is used to convert the nested form cubble into the long form for these operations:
-
-```{r eval=  FALSE, echo = TRUE}
-tmax <- tmax %>% 
-  stretch() %>%
-  filter(lubridate::year(date) %in% c(1971:1975, 2016:2020)) %>%
-  mutate(month = lubridate::month(date), 
-         group = as.factor(ifelse(lubridate::year(date) > 2015, 
-                                  "2016 ~ 2020", "1971 ~ 1975"))) %>%
-  group_by(month, group) %>%
-  summarise(tmax = mean(tmax, na.rm = TRUE))
-```
-
-A data quality issue with GHCN data is that while the first and last year of each series is provided, years missing in this period is not reported.
-There are a few stations which don't have records during 1971 - 1975 and these stations are filtered out by examining whether the summarised `tmax` has a total of 24 months.
-This is again a station-wise operation and is operated in the nested form, which is switched to from the long form by  `tamp()`:
-
-```{r eval = FALSE, echo = TRUE}
-tmax <- tmax %>% 
-  tamp() %>%
-  filter(nrow(ts) == 24) 
-```
-
-Lastly, to create a glyph map, both the major (`longitude`, `latitude`) and minor (`month`, `tmax`) coordinates need to be in the same table.
-Spatial variables can be moved to the long form with `migrate()`:
-
-```{r eval = FALSE, echo = TRUE}  
-tmax <- tmax %>%   
-  stretch() %>%
-  migrate(latitude, longitude)
-```
-
-Figure \ref{fig:basic-manip} shows the glyph map made with the data after the wangling above.
-One issue with this map is that the similar pattern shown from a few nearby stations around Sydney (151E, 34S) and New castle (152E, 33S) can be distracting.
-Aggregation is a useful technique to observe the general pattern of a collection of series and will be the topic of the next example.
-
-```{r basic-manip, out.width="100%", out.height="70%", fig.cap="Glyph map of the mean maximum temperature by month for Victoria and New South Wales weather stations.
-On the top left corner is a detailed legend for station Cobar highlighted in the black box.
-Compared to 1971 - 1975,  the period 2016 - 2020 sees an increase of mean maximum temperature in spring and summer in Australia (end and beginning of the year).
-A larger increase during the first three months of the year is observed at stations with a lower latitude."}
-knitr::include_graphics("figures/basic-manip.png")
-```
-
-## Australia precipitation pattern in 2020
-
-In Figure \ref{fig:basic-agg}.
-Aggregating them into averages or other statistics would give a better picture.
-This can be done with cubble through a hierarchical structure, where the cluster indicator is one level above the individual station.
-
-The `weatherdata::climate_full` data has daily climate data of `r nrow(weatherdata::climate_full)` Australia stations from 2016 to 2020.
-
-calcluate distance matrix
-As an example to illustrate here, a kmean clustering algorithm based on the distance matrix is used and the number of centres is set to 20.
-More complex algorithms can also be used for more complex problem, as long as a mapping from each station id to the cluster id can be constructed.
-
-<!-- Now we can obtain the aggregated series as described in the workflow diagram above and construct the glyph map with `GGally::glyphs()`: -->
-
-<!-- We can also look at the precipitation of each individual station within the same cluster: -->
-
-<!-- Lastly, there is one series on Tasmania island standing out from others, lets look at where it is: -->
-
-```{r basic-agg, out.width="100%", out.height="90%", fig.cap="sdfasdf"}
-knitr::include_graphics("figures/basic-agg.png")
-```
-
-## River level data in Victria water gauges
-
-The water level data comes from [Bureau of Meteorology](http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full) and has a copy in `weatherdata`.
-Here we extract the water course level and add a column annotate this data of type `river`.
-For the rainfall data, we will still use the `weatherdata::climate_full`, filtering for Victorian stations in 2020 should be pretty familiar by now.
-Again, we first look at where these stations are on the map first:
-
-Now we use `match_sites()` to first pair the weather stations with the river gauges spatially and then apply the temporal matching on `prcp`.
-We will construct the interval based on peaks in `climate` since we would expect a lag effect for precipitation to flow into the river and cause a raise in river level, hence `temporal_independent = climate`.
-We select the 30 highest peak from the series to construct the match by setting `temporal_n_highest = 30`.
-This is a tuning parameter and you can start with 10% of the points of one series (here we have daily data for a year, 10% is roughly 30 points).
-`temporal_min_match` filters out pairs don't have enough match and to return all the pairs, set `temporal_min_match` to `0`.``
-
-The output from temporal matching is also a cubble, with additional column `.dist` and `.group` inherent from spatial matching and `n_match` for the number of matched temporal peaks.
-Then you can use this output to plot the location of match or to look at the series:
-
-```{r matching, out.width="100%", out.height="30%", fig.cap="asdfasd"}
-knitr::include_graphics("figures/matching.png")
-```
-
-## ERA5: climate reanalysis data
-
-\newpage
-
-## Interative graphic with cubble
-
-With a larger dataset, interactive graphics can be helpful to explore these data from multiple perspective.
-Two examples are presented here to show how interactivity can be used to explore the climate data across Australia.
-Again, `weatherdata::climate_full` will be used and we hope to include more than one time variable here.
-The difference between the maximum and minimum temperature of the day, the diurnal range, can be directly calculated from the data and it can have large impact on the wine industry (lol this is going far).
-This would be useful to compare the temperature difference across stations and to compare how this temperature difference differs throughout the year.
-
-Temperature difference is a time-wise variable and should be computed in the long form cubble. The variables are averaged over the 2016 to 2020 period by month to stabilise the result.
-
-```{r eval = FALSE, echo = TRUE}
-res <- weatherdata::climate_full %>%
-  # filter(!id %in% c("ASN00067033", "ASN00072091", 
-  #                   "ASN00059040", "ASN00097053")) %>%
-  stretch() %>%
-  mutate(month = lubridate::month(date),
-         diff = tmax - tmin) %>%
-  group_by(month) %>%
-  summarise(tmax = mean(tmax, na.rm = TRUE),
-            diff = mean(diff, na.rm = TRUE)) %>%
-  ungroup(month)
-
-```
-
-Several variables can be useful when comparing the seasonal effect of the temperature differences, for example, the difference of average temperature difference between summer and winter can tell [...].
-This is a station-wise measure and hence computed in the nested form.
-
-```{r eval = FALSE, echo = TRUE}
-res <- res %>% 
-  tamp() %>%
-  mutate(
-    diff_winter = ts %>% filter(month %in% c(6, 7, 8)) %>%
-      pull(diff) %>% mean(na.rm = TRUE),
-    diff_summer = ts %>% filter(month %in% c(12, 1, 2)) %>%
-      pull(diff) %>% mean(na.rm = TRUE),
-    diff_year = diff_winter - diff_summer
-  )
-```
-
-The result from the above manipulation can be directly supplied to make interactive plots and Figure \ref{fig:interactive-linking} is an example made with \pkg{plotly} and \pkg{crosstalk}.
-For example, the second row results from selecting the highest mean maximum temperature in July and this corresponds to Kalumburu where there is a larger winter (June to August) temperature difference than summer (December to February).
-The third row selects one of the stations (Birdsville Airport) with a large average temperature difference in the interior desert region on the map and the linked view from the time series shows that this maximum and minimum temperature difference is constant throughout the year.
-The last row selects the lowest maximum temperature in July on the temperature band and surprisingly, the corresponding station doesn't fall into the Tasmania island but Mountain Hotham in Victoria.
-
-```{r interactive-linking, echo = FALSE, out.width="100%", out.height="25%", fig.cap = "Screenshots of the mean daily temperature difference averaged by month from 2016 to 2020 at various locations in Australia.
-Crossed linking between the map and the temperature band is implemented in both directions for exploring the temperature pattern interactively.
-The first row is the general appearance of the graph and the next three rows are examples from various selection to show different temperature patterns in Australia.
-The html version can be found at ...", fig.show='hold'}
-knitr::include_graphics(here::here("figures/linking.png"))
-knitr::include_graphics(here::here("figures/linking-north.png"))
-knitr::include_graphics(here::here("figures/linking-mid.png"))
-knitr::include_graphics(here::here("figures/linking-south.png"))
-```
-
-Cubble is also compatible with \pkg{leaflet} and with leaflet extension package \pkg{leafpop}, the temperature band can be displayed upon clicking on the station on the map.
-Figure \ref{fig:interactive-popup} shows more examples of temperature pattern that can be found in Australia.
-With the same latitude as Kalumburu but in the Pacific coastline, Cooktown airport shows a similarly high but constant temperature throughout the year.
-With a slightly higher latitude than the mountain Hotham, flat land in Melbourne airport shows a less extreme temperature band.
-Lastly, the Cape Grim Baps, in the north-west tip of the Tasmania island shows a uniformly low and thin temperature band.
-
-```{r interactive-popup, echo = FALSE, out.width="50%", out.height="30%", fig.cap = "Screenshots of the diurnal temperature range averaged by month from 2016 to 2020 at various locations in Australia.
-The leaflet popup allows users to click on a single station for its temperature band as a small plot in the popup window.
-The html version can be found at ...", fig.show='hold'}
-knitr::include_graphics(here::here("figures/popup-north.png"))
-knitr::include_graphics(here::here("figures/popup-vic.png"))
-knitr::include_graphics(here::here("figures/popup-tas.png"))
-```
-
-# Conclusion
-
-# Appendix
-
-```{r illu-interactive-2, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "demon interactivity"}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.005.png"))
 ```
@@
 
-<!-- # Old stuff -->
+```{r child = "4-1-basic.Rmd"}
 
-<!-- In the temporal aspect, the `tsibble` [@tsibbles] structure and its tidyverts ecosystem have provided a [...
-] workflow to work with temporal data.
-In a tsibble structure, temporal data is characterised by `index` and `key` where `index` is the temporal identifier and `key` is the identifier for multiple series, which could be used as a spatio identifier.
-However, a tsibble object, by construction, always requires the `index` in its structure.
-This makes it less appealing for spatio-temporal data since the output of calculated spatio-specific variables (i.e.
-features of each series) don't have the time dimension.
-Analysts will either need to have an additional step to join this output to the original tsibble or operate with variables stored in two separate objects.
-In addition, the long form structure of a tsibble object means spatio variables (i.e.
-longitude, latitude, and features of each series if joined back to the tsibble) of each spatio identifier will be repetitively recorded at each timestamp.
-This repetition is unnecessary and would inflate the object size for long series.
--->
+```
 
-<!-- # Create a cubble -->
+```{r child = "4-2-agg.Rmd"}
 
-<!-- The creation of a cubble requires the site identifier (`key`), as well as the spatial (`coords`) and temporal (`index`) identifier.
-`climate_flat` is already a tibble and it uses `id` to identify each station, `date` as the time identifier, and `c(long, lat)` as the spatial identifier.
-To create a cubble for this data, use: -->
+```
 
-<!-- ```{r} -->
-<!-- climate_flat %>% as_cubble(key = id, index = date, coords = c(long, lat)) -->
-<!-- ``` -->
+```{r child = "4-3-water-level.Rmd"}
 
-<!-- Most of the time, spatio-temporal data doesn't come into this form and analysts need to query the climate variables based on station metadata.
-\textcolor{red}{This is also a problem illustrated in Section 3.5 in @tidydata.
-Here we provide a structured way to query this data based on the row-wise operator and nested list.} For this type of task, one can structure a metadata into a tibble and use row-wise operator to query the climate variables into a nested list.
-As an example here we demonstrate the workflow to find the 5 closest stations to Melbourne.
-We first create a station data frame with the 5 target stations.
--->
+```
 
-<!-- ```{r echo = FALSE} -->
-<!-- latlon_ref <- tibble::tibble(id = "melbourne", lat = -37.8136, long = 144.9631) -->
+```{r child = "4-4-netcdf.Rmd"}
 
-<!-- cand <- aus_climate %>% -->
-<!--   as_tibble() %>% -->
-<!--   select(id, lat, long, elev, name, wmo_id) -->
+```
 
-<!-- calc_dist <- function(df_ref, df, n = 5) { -->
-<!--   long_ref <- df_ref$long -->
-<!--   lat_ref <- df_ref$lat -->
+```{r child = "4-5-interactive.Rmd"}
 
-<!--   df %>% -->
-<!--     mutate( -->
-<!--       long_ref = long_ref, -->
-<!--       lat_ref = lat_ref, -->
-<!--       dist = rnoaa::meteo_spherical_distance(lat, long, lat_ref, long_ref) -->
-<!--     ) %>% -->
-<!--     slice_min(dist, n = n) %>% -->
-<!--     mutate(city = df_ref$id) %>% -->
-<!--     select(-long_ref, -lat_ref) -->
-<!-- } -->
+```
 
-<!-- (stations <- do.call("rbind", purrr::map( -->
-<!--   1:nrow(latlon_ref), -->
-<!--   ~ calc_dist(latlon_ref[.x, ], cand) -->
-<!-- ))) -->
-<!-- ``` -->
+```{r child = "5-conclusion.Rmd"}
 
-<!-- We can query the climate information into a nested list named `ts` for each station with the `rowwise()` operator. To create a cubble, supply the same identifiers as with the first example. -->
-
-<!-- ```{r eval = FALSE} -->
-<!-- sydmel_climate <- stations %>% -->
-<!--   rowwise() %>% -->
-<!--   mutate(ts = list(meteo_pull_monitors(id, -->
-<!--     date_min = "2020-01-01", -->
-<!--     date_max = "2020-12-31", -->
-<!--     var = c("PRCP", "TMAX", "TMIN") -->
-<!--   ) %>% -->
-<!--     select(-id))) %>% -->
-<!--   as_cubble(key = id, index = date, coords = c(long, lat)) -->
-<!-- ``` -->
-
-<!-- ```{r echo = FALSE} -->
-<!-- (sydmel_climate <- stations %>% -->
-<!--   left_join(aus_climate) %>% -->
-<!--   as_cubble(key = id, index = date, coords = c(long, lat))) -->
-<!-- ``` -->
-
-<!-- Below are the how the nested and long form look like for Australia climate data, which records daily precipitation, maximum and minimum temperature for 55 stations across Australia from 2015- 2020.
-Notice that each station forms a group in both forms and specifically, the nested and long form have a underlying `rowwise_df` and `grouped_df` respectively.
--->
-
-<!-- With a cubic framework on mind, different types of manipulation with cubble can be thought of as slicing the cube in various way.
-The table below shows how some `dplyr` verbs are mapped into the operation in a cubble.
-With the existing grouping on the station, additional grouping can be added with `group_by` and removed with `ungrouped`.
-[talk about why it is useful] -->
-
-<!-- \newpage -->
-
-<!-- ## Cubble operations -->
-
-<!-- ```{r cubble-operations, echo = FALSE, fig.cap="Cubble operations", out.height="50%", out.width="90%", fig.align='center'} -->
-<!-- knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.004.png")) -->
-<!-- ``` -->
+```

### 2022-01-13 4-5-interactive.Rmd
interactive graphic crosstalk done

@@
 
-To visualise climate variables from all the stations in Australia, interactive graphics can be helpful.
-Here a plotly and a leaflet example is given to explore the yearly diurnal temperature range pattern in Australia.
-This would be useful to see how the average temperature range differs in various latitudes in Australia, whether there's a summer or winter effect, any coastline, mountain or desert topological considerations that would affect the temperature.
+With spatio-temporal data, users may wish to make plots to learn the spatial distribution of a variable or to find patterns such as trend or seasonality in the time series.
+Combining this two types of plot with interactivity let users to link between points on the map and their corresponding time series and explore the spatial and temporal dimension of the data simultaneously.
+Below is an example that describes the process of building an interactive graphic with `cubble` and `crosstalk`.
 
-Again, `weatherdata::climate_full` will be first processed to produce summarised monthly maximum and minimum temperature across 2016 to 2020.
-Then a shared data with the same group is created from the nested and long cubble to enable the cross-talking via the common column `id`.
-The map and temperature band is created separately with the two shared data objects and combined together with `crosstalk::bscols`.
-A sketch of the code to produce Figure \ref{fig:interactive-linking} is presented below:
+Starting with the original data, some pre-processing may be required to summarise the data before the visualisation.
+This example explores the variation of monthly temperature range across `r nrow(weatherdata::climate_full)` weather stations in Australia and with `weatherdata::climate_full`, daily temperature range is first calculated as the difference between `tmax` and `tmin`.
+The three daily variables are then averaged into month over 2016 - 2020 in the long form.
+Variance of the temperature difference is then calculated for each station in the nested form.
+Then, a `SharedData` object is constructed for each form of the cubble and the same `group` argument ensures the cross-linking of the two forms via the common `id` column.
+The spatial map and time series plot are then made with each `SharedData` objects separately.
+In this example, stations on the Australia map, made from the nested form, are coloured by the calculated variance and a ribbon band is constructed using the long form cubble to show the maximum and minimum temperature of each station across month.
+With a different dataset, users are free to calculate any per station measure in the nested form or to make any time-wise summary of the data in the long form to customise the spatial or temporal view.
+And the cross-linking is always safeguarded by the shared `id` column embedded in the cubble structure.
+Below is the pseudo code that outlines the process to construct an interactive graphic described above:
 
 ```{r eval = FALSE, echo = TRUE}
-# data processing
+# data pre-processing
 clean <- weatherdata::climate_full %>% ...
 
-# created grouped instance for crosstalk
+# created SharedData instance for crosstalk
 nested <- clean %>% SharedData$new(~id, group = "cubble")
@@
 
-# create the map with SharedData nested, the time series with SharedData long
+# create the spatial and temporal view each with a ShareData instance
 p1 <- nested %>% ...
@@
 
-In Figure \ref{fig:interactive-linking}, the first row shows the default plot and when a selection is made, relevant component from both plots will be highlighted.
-For example, in the second row, the highest mean maximum temperature in July is selected on the time series and this corresponds to Kalumburu on the map where there is a larger diurnal temperature range in winter (June to August) than summer (December to February).
-The third row selects a station (Birdsville Airport) in the interior desert region that has a large average temperature difference on the map and the linked view shows a wide but constant diurnal temperature range throughout the year.
-The last row selects the lowest maximum temperature in July on the right plot and surprisingly, this links to the Thredbo airport at the boarder of Victoria and New South Wales, rather than any station in the Tasmania island!
+In Figure \ref{fig:interactive-linking}, the first row shows the initial view of the interactive graphic.
+Overall, most regions in Australia have low variance of temperature range across different months while the north-west coastline, bottom of South Australia, and Victoria stands out with larger monthly changes.
+In the second row, Mount Elizabeth is selected on the map given its dark colour and this links to the ribbon on the right where there is a larger temperature range is presented in Australian winter (June to August).
+The third row selects the Grampian station in Victoria and the linked ribbon shows an opposite wider range in Australian summer period (December to February).
+The last row selects point with the lowest minimum temperature in August in the ribbon plot and surprisingly, this links to the Thredbo airport in the Victoria and New South Wales border, rather than somewhere in the Tasmania island!
 
-```{r interactive-linking, echo = FALSE, out.width="100%", out.height="25%", fig.retina = 2, dpi = 300, fig.cap = "Screenshots of the mean daily temperature difference averaged by month from 2016 to 2020 at various locations in Australia.
+```{r interactive-linking, echo = FALSE, out.width="100%", out.height="20%", fig.retina = 2, dpi = 300, fig.cap = "Screenshots of the mean daily temperature difference averaged by month from 2016 to 2020 at various locations in Australia.
 Crossed linking between the map and the temperature band is implemented in both directions for exploring the temperature pattern interactively.

### 2022-01-13 main.tex
play aroudnd with child rmd to organising sections
interactive graphic crosstalk done

@@
 operations.
+
+\newpage
 
@@
 
-To visualise climate variables from all the stations in Australia,
-interactive graphics can be helpful. Here a plotly and a leaflet example
-is given to explore the yearly diurnal temperature range pattern in
-Australia. This would be useful to see how the average temperature range
-differs in various latitudes in Australia, whether there's a summer or
-winter effect, any coastline, mountain or desert topological
-considerations that would affect the temperature.
+With spatio-temporal data, users may wish to make plots to learn the
+spatial distribution of a variable or to find patterns such as trend or
+seasonality in the time series. Combining this two types of plot with
+interactivity let users to link between points on the map and their
+corresponding time series and explore the spatial and temporal dimension
+of the data simultaneously. Below is an example that describes the
+process of building an interactive graphic with \texttt{cubble} and
+\texttt{crosstalk}.
 
-Again, \texttt{weatherdata::climate\_full} will be first processed to
-produce summarised monthly maximum and minimum temperature across 2016
-to 2020. Then a shared data with the same group is created from the
-nested and long cubble to enable the cross-talking via the common column
-\texttt{id}. The map and temperature band is created separately with the
-two shared data objects and combined together with
-\texttt{crosstalk::bscols}. A sketch of the code to produce Figure
-\ref{fig:interactive-linking} is presented below:
+Starting with the original data, some pre-processing may be required to
+summarise the data before the visualisation. This example explores the
+variation of monthly temperature range across 639 weather stations in
+Australia and with \texttt{weatherdata::climate\_full}, daily
+temperature range is first calculated as the difference between
+\texttt{tmax} and \texttt{tmin}. The three daily variables are then
+averaged into month over 2016 - 2020 in the long form. Variance of the
+temperature difference is then calculated for each station in the nested
+form. Then, a \texttt{SharedData} object is constructed for each form of
+the cubble and the same \texttt{group} argument ensures the
+cross-linking of the two forms via the common \texttt{id} column. The
+spatial map and time series plot are then made with each
+\texttt{SharedData} objects separately. In this example, stations on the
+Australia map, made from the nested form, are coloured by the calculated
+variance and a ribbon band is constructed using the long form cubble to
+show the maximum and minimum temperature of each station across month.
+With a different dataset, users are free to calculate any per station
+measure in the nested form or to make any time-wise summary of the data
+in the long form to customise the spatial or temporal view. And the
+cross-linking is always safeguarded by the shared \texttt{id} column
+embedded in the cubble structure. Below is the pseudo code that outlines
+the process to construct an interactive graphic described above:
 
@@
 \begin{CodeInput}
-R> # data processing
+R> # data pre-processing
 R> clean <- weatherdata::climate_full %>% ...
 R> 
-R> # created grouped instance for crosstalk
+R> # created SharedData instance for crosstalk
 R> nested <- clean %>% SharedData$new(~id, group = "cubble")
@@
 R> 
-R> # create the map with SharedData nested, the time series with SharedData long
+R> # create the spatial and temporal view each with a ShareData instance
 R> p1 <- nested %>% ...
@@
 
-In Figure \ref{fig:interactive-linking}, the first row shows the default
-plot and when a selection is made, relevant component from both plots
-will be highlighted. For example, in the second row, the highest mean
-maximum temperature in July is selected on the time series and this
-corresponds to Kalumburu on the map where there is a larger diurnal
-temperature range in winter (June to August) than summer (December to
-February). The third row selects a station (Birdsville Airport) in the
-interior desert region that has a large average temperature difference
-on the map and the linked view shows a wide but constant diurnal
-temperature range throughout the year. The last row selects the lowest
-maximum temperature in July on the right plot and surprisingly, this
-links to the Thredbo airport at the boarder of Victoria and New South
-Wales, rather than any station in the Tasmania island!
+In Figure \ref{fig:interactive-linking}, the first row shows the initial
+view of the interactive graphic. Overall, most regions in Australia have
+low variance of temperature range across different months while the
+north-west coastline, bottom of South Australia, and Victoria stands out
+with larger monthly changes. In the second row, Mount Elizabeth is
+selected on the map given its dark colour and this links to the ribbon
+on the right where there is a larger temperature range is presented in
+Australian winter (June to August). The third row selects the Grampian
+station in Victoria and the linked ribbon shows an opposite wider range
+in Australian summer period (December to February). The last row selects
+point with the lowest minimum temperature in August in the ribbon plot
+and surprisingly, this links to the Thredbo airport in the Victoria and
+New South Wales border, rather than somewhere in the Tasmania island!
 
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.25\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/linking} \includegraphics[width=1\linewidth,height=0.25\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/linking-north} \includegraphics[width=1\linewidth,height=0.25\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/linking-mid} \includegraphics[width=1\linewidth,height=0.25\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/linking-south}
+{\centering \includegraphics[width=1\linewidth,height=0.2\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/linking} \includegraphics[width=1\linewidth,height=0.2\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/linking-north} \includegraphics[width=1\linewidth,height=0.2\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/linking-mid} \includegraphics[width=1\linewidth,height=0.2\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/linking-south}
 
@@
 
-\newpage
-
 \bibliography{references.bib}

### 2022-01-13 scripts/interactive-linking.R

@@ -3,65 +3,64 @@ library(tidyverse)
 library(crosstalk)
 library(plotly)
 
-clean <-  weatherdata::climate_full %>%
-  filter(!id %in% c("ASN00067033", "ASN00072091", "ASN00059040", "ASN00097053")) %>%
-  stretch() %>%
-  mutate(month = lubridate::month(date),
-         diff = tmax - tmin,
-         avg = (tmax + tmin)/2) %>%
+clean <- weatherdata::climate_full %>%
+  filter(!id %in% c("ASN00067033", "ASN00072091", "ASN00059040", 
+                    "ASN00097053", "ASN00040856", "ASN00015548")) %>%
+  stretch() %>% 
+  mutate(month = lubridate::month(date)) %>%
   group_by(month) %>%
   summarise(tmax = mean(tmax, na.rm = TRUE),
             tmin = mean(tmin, na.rm = TRUE),
-            diff = mean(diff, na.rm = TRUE),
-            avg = mean(avg, na.rm =TRUE)) %>%
+            diff = mean(tmax - tmin, na.rm = TRUE)) %>% 
   ungroup(month) %>%
   mutate(dummy_date = as.Date(glue::glue("2021-{month}-01"))) %>% 
-  tamp() %>%
-  mutate(
-    diff_winter = ts %>% filter(month %in% c(6, 7, 8)) %>% pull(diff) %>% mean(na.rm = TRUE),
-    diff_summer = ts %>% filter(month %in% c(12, 1, 2)) %>% pull(diff) %>% mean(na.rm = TRUE),
-    diff_year = diff_winter - diff_summer,
-    annual = ts %>% filter(month %in% c(7, 1)) %>% pull(avg) %>% diff() %>% abs(),
-    diurnal = ts %>% pull(diff) %>% mean(na.rm = TRUE)
-  )
+  tamp() %>% 
+  mutate(temp_diff_var = ts %>% pull(diff) %>% var(na.rm = TRUE))
+  
 
 nested <- clean %>% 
   SharedData$new(~id, group = "cubble")
 
-p1 <- nested %>%  
-  ggplot(aes(x = diff_summer, y = diff_winter, color = diff_year, label = name)) + 
-  geom_point() + 
-  geom_abline(slope = 1, intercept = 0, linetype = "dashed") + 
-  colorspace::scale_color_continuous_diverging("Purple-Green", mid = 0, l2 = 80) + 
[…]

### 2022-01-13 scripts/interactive-linking2.R

@@ -1,56 +1,0 @@
-dt <- clean 
-state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)
-dt %>% 
-  ggplot() +
-  geom_sf(data = state_map, aes(geometry = geometry), fill = "transparent") +
-  geom_point(aes(x = long, y = lat, label = name, color = annual)) +
-  colorspace::scale_color_continuous_sequential("Red") + 
-  theme_bw() + 
-  theme(panel.grid.major = element_blank(),
-        legend.position = "bottom")
-
-nested <- clean %>% 
-  SharedData$new(~id, group = "cubble")
-
-long <- clean %>% stretch() %>% 
-  SharedData$new(~id, group = "cubble")
-
-
-p1 <- nested %>% 
-  ggplot() +
-  geom_sf(data = state_map, aes(geometry = geometry), fill = "transparent") +
-  geom_point(aes(x = long, y = lat,color = diurnal, label = name)) +
-  colorspace::scale_color_continuous_sequential(
-    "Oslo",
-    name = "Avg. temp. diff.") + 
-  labs(x = "Longitude", y = "Latitude") + 
-  coord_sf() + 
-  theme_bw() + 
-  theme(panel.grid.major = element_blank(),
-        legend.position = "bottom")
-
-p2 <- long %>% 
-  ggplot() +
-  # geom_ribbon(aes(x = dummy_date, ymin = tmin, ymax = tmax, group = id), 
-  #             color = "grey30", fill = "grey30", alpha = 0.2) +
-  geom_line(aes(x = dummy_date, y = tmax, group = id), size = 0.1) + 
-  geom_line(aes(x = dummy_date, y = tmin, group = id), size = 0.1) + 
-  scale_x_date(date_labels = "%b", date_breaks = "1 month") + 
-  labs(x = "Month", y = "Temperature") + 
-  theme_bw() + 
-  theme(panel.grid.major = element_blank(),
-        legend.position = "bottom")
-
-# More adjustment:
-# - lighter color on the selected temperature band
-# - wider width for ts and narrow for the map
-# - color correspondence in the popup map
-
-out <- bscols(
-  ggplotly(p1, width = 900, height = 700, tooltip = "label") %>%
-    highlight(on = "plotly_selected", off = "plotly_deselect"),
-  ggplotly(p2, width = 1200, height = 700, tooltip = "label") %>% 
[…]

### 2022-01-15 4-5-interactive.Rmd
interactive graphic sectionsection done

@@
 The third row selects the Grampian station in Victoria and the linked ribbon shows an opposite wider range in Australian summer period (December to February).
-The last row selects point with the lowest minimum temperature in August in the ribbon plot and surprisingly, this links to the Thredbo airport in the Victoria and New South Wales border, rather than somewhere in the Tasmania island!
+The last row selects point with the lowest minimum temperature in August in the ribbon plot and surprisingly, this links to the Thredbo Airport in the Victoria and New South Wales border, rather than somewhere in the Tasmania island!
 
-```{r interactive-linking, echo = FALSE, out.width="100%", out.height="20%", fig.retina = 2, dpi = 300, fig.cap = "Screenshots of the mean daily temperature difference averaged by month from 2016 to 2020 at various locations in Australia.
-Crossed linking between the map and the temperature band is implemented in both directions for exploring the temperature pattern interactively.
-The first row is the general appearance of the graph and the next three rows are examples from various selection to show different temperature patterns in Australia.
-The html version can be found at ...", fig.show='hold'}
+```{r interactive-linking, echo = FALSE, out.width="100%", out.height="20%", fig.retina = 2, dpi = 300, fig.cap = "Screenshots of variance of monthly temperature range from 2016 to 2020 in Australia.
+Crossed linking between the map and the temperature band is implemented in both directions to explore the temperature pattern interactively.
+The first row is the initial view of the graphic and the latter three show the interactive view after selecting on the map (row 2 and 3) and on the temperature band (row 4).", fig.show='hold'}
 knitr::include_graphics(here::here("figures/linking.png"))
@@
 
-With leaflet popup, the temperature range can be displayed as a small plot upon clicking on the station on the map.
-This would require first creating all the popup plot separately in a vector and then add these plots to a leaflet map with `leafpop::addPopupGraphs()`:
+With leaflet popup, the temperature range can be displayed as a small subplot upon clicking on the map.
+This would require first creating the popup plots from the long form cubble as a vector and then add these plots to a leaflet map, created from the nested cubble, with `leafpop::addPopupGraphs()`:
 
 ```{r eval = FALSE, echo = TRUE}
-# create subplots for each station
+# data pre-processing
+clean <- weatherdata::climate_full %>% ...
+
+# use the long form to create subplots for each station
 df_id <- unique(clean$id)
@@
   dt <- clean %>% filter(id == df_id[i])
-  dt %>% ...
+  ggplot(dt) %>% ...
 })
 
-# inset subplots as a popup in leaflet
-leaflet(clean) %>% 
+# create nested form leaflet map with temperature band as subplots 
+nested <- tamp(clean)
+leaflet(nested) %>% 
   addTiles() %>% 
@@
 
-Figure \ref{fig:interactive-popup} shows more examples of temperature pattern that can be found in Australia.
-With the same latitude as Kalumburu but in the Pacific coastline, Cooktown airport shows a similarly high but constant temperature range throughout the year.
-With a slightly higher latitude than the Thredbo airport , Melbourne airport shows a less extreme temperature pattern Lastly, the Cape Grim Baps, in the north-west tip of the Tasmania island shows low and narrow temperature band, constantly betwee 10 and 20 degree throughout the year.
+Figure \ref{fig:interactive-popup} shows the same content as Figure \ref{fig:interactive-linking} but made with leaflet and popups.
+A few sites with low temperature variance are selected and these sites still show different patterns: Lady Elloit Island in the Pacific coastline shows a thin temperature band of 5 degree; Birdsville Airport, in the middle of Australia desert, shows a wide temperature range of 10 degree throughout the year; and King Island Airport, in the north-west tip of Tasmania island shows a similar narrow range as Lady Elloits but a maximum and minimum temperature in general.
+This would suggest variance on itself may not be enough to summarise the temperature variation in Australia and other summary statistics, for example, mean of temperature range or maximum temperature itself, could also be useful to find temperature pattern in Australia.
 
-```{r interactive-popup, echo = FALSE, out.width="45%", out.height="30%", fig.retina = 2, dpi = 300, fig.cap = "Screenshots of the diurnal temperature range averaged by month from 2016 to 2020 at various locations in Australia.
-The leaflet popup allows users to click on a single station for its temperature band as a small plot in the popup window.
-The html version can be found at ...", fig.show='hold'}
-knitr::include_graphics(here::here("figures/popup-north.png"))
-knitr::include_graphics(here::here("figures/popup-vic.png"))
-knitr::include_graphics(here::here("figures/popup-tas.png"))
+```{r interactive-popup, echo = FALSE, out.width="45%", out.height="25%", fig.retina = 2, dpi = 300, fig.cap = "Screenshots of variance of monthly temperature range from 2016 to 2020 in Australia.
+Upon clicking a single station on the leaftlet up, the temperature band will be shown as a subplot in the popup box.", fig.show='hold'}
+knitr::include_graphics(here::here("figures/popup.png"))
+knitr::include_graphics(here::here("figures/popup-right.png"))
+knitr::include_graphics(here::here("figures/popup-mid.png"))
+knitr::include_graphics(here::here("figures/popup-south.png"))
 ```

### 2022-01-15 main.tex
interactive graphic sectionsection done

@@
 point with the lowest minimum temperature in August in the ribbon plot
-and surprisingly, this links to the Thredbo airport in the Victoria and
+and surprisingly, this links to the Thredbo Airport in the Victoria and
 New South Wales border, rather than somewhere in the Tasmania island!
@@
 
-\caption[Screenshots of the mean daily temperature difference averaged by month from 2016 to 2020 at various locations in Australia]{Screenshots of the mean daily temperature difference averaged by month from 2016 to 2020 at various locations in Australia.
-Crossed linking between the map and the temperature band is implemented in both directions for exploring the temperature pattern interactively.
-The first row is the general appearance of the graph and the next three rows are examples from various selection to show different temperature patterns in Australia.
-The html version can be found at ...}\label{fig:interactive-linking}
+\caption[Screenshots of variance of monthly temperature range from 2016 to 2020 in Australia]{Screenshots of variance of monthly temperature range from 2016 to 2020 in Australia.
+Crossed linking between the map and the temperature band is implemented in both directions to explore the temperature pattern interactively.
+The first row is the initial view of the graphic and the latter three show the interactive view after selecting on the map (row 2 and 3) and on the temperature band (row 4).}\label{fig:interactive-linking}
 \end{figure}
@@
 With leaflet popup, the temperature range can be displayed as a small
-plot upon clicking on the station on the map. This would require first
-creating all the popup plot separately in a vector and then add these
-plots to a leaflet map with \texttt{leafpop::addPopupGraphs()}:
+subplot upon clicking on the map. This would require first creating the
+popup plots from the long form cubble as a vector and then add these
+plots to a leaflet map, created from the nested cubble, with
+\texttt{leafpop::addPopupGraphs()}:
 
@@
 \begin{CodeInput}
-R> # create subplots for each station
+R> # data pre-processing
+R> clean <- weatherdata::climate_full %>% ...
+R> 
+R> # use the long form to create subplots for each station
 R> df_id <- unique(clean$id)
@@
 +   dt <- clean %>% filter(id == df_id[i])
-+   dt %>% ...
++   ggplot(dt) %>% ...
 + })
 R> 
-R> # inset subplots as a popup in leaflet
-R> leaflet(clean) %>% 
+R> # create nested form leaflet map with temperature band as subplots 
+R> nested <- tamp(clean)
+R> leaflet(nested) %>% 
 +   addTiles() %>% 
@@
 
-Figure \ref{fig:interactive-popup} shows more examples of temperature
-pattern that can be found in Australia. With the same latitude as
-Kalumburu but in the Pacific coastline, Cooktown airport shows a
-similarly high but constant temperature range throughout the year. With
-a slightly higher latitude than the Thredbo airport , Melbourne airport
-shows a less extreme temperature pattern Lastly, the Cape Grim Baps, in
-the north-west tip of the Tasmania island shows low and narrow
-temperature band, constantly betwee 10 and 20 degree throughout the
-year.
+Figure \ref{fig:interactive-popup} shows the same content as Figure
+\ref{fig:interactive-linking} but made with leaflet and popups. A few
+sites with low temperature variance are selected and these sites still
+show different patterns: Lady Elloit Island in the Pacific coastline
+shows a thin temperature band of 5 degree; Birdsville Airport, in the
+middle of Australia desert, shows a wide temperature range of 10 degree
+throughout the year; and King Island Airport, in the north-west tip of
+Tasmania island shows a similar narrow range as Lady Elloits but a
+maximum and minimum temperature in general. This would suggest variance
+on itself may not be enough to summarise the temperature variation in
+Australia and other summary statistics, for example, mean of temperature
+range or maximum temperature itself, could also be useful to find
+temperature pattern in Australia.
 
@@
 
-{\centering \includegraphics[width=0.45\linewidth,height=0.3\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/popup-north} \includegraphics[width=0.45\linewidth,height=0.3\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/popup-vic} \includegraphics[width=0.45\linewidth,height=0.3\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/popup-tas}
+{\centering \includegraphics[width=0.45\linewidth,height=0.25\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/popup} \includegraphics[width=0.45\linewidth,height=0.25\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/popup-right} \includegraphics[width=0.45\linewidth,height=0.25\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/popup-mid} \includegraphics[width=0.45\linewidth,height=0.25\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/popup-south}
 
@@
 
-\caption[Screenshots of the diurnal temperature range averaged by month from 2016 to 2020 at various locations in Australia]{Screenshots of the diurnal temperature range averaged by month from 2016 to 2020 at various locations in Australia.
-The leaflet popup allows users to click on a single station for its temperature band as a small plot in the popup window.
-The html version can be found at ...}\label{fig:interactive-popup}
+\caption[Screenshots of variance of monthly temperature range from 2016 to 2020 in Australia]{Screenshots of variance of monthly temperature range from 2016 to 2020 in Australia.
+Upon clicking a single station on the leaftlet up, the temperature band will be shown as a subplot in the popup box.}\label{fig:interactive-popup}
 \end{figure}

### 2022-01-15 scripts/interactive-popup.R

@@ -4,31 +4,26 @@ library(leaflet)
 library(leafpop)
 
 clean <-  weatherdata::climate_full %>%
-  filter(!id %in% c("ASN00067033", "ASN00072091")) %>% 
+  filter(!id %in% c("ASN00067033", "ASN00072091", "ASN00059040", 
+                    "ASN00097053", "ASN00040856", "ASN00015548")) %>%
   stretch() %>%
-  mutate(month = lubridate::month(date),
-         diff = tmax - tmin) %>% 
+  mutate(month = lubridate::month(date)) %>% 
   group_by(month) %>% 
   summarise(tmax = mean(tmax, na.rm = TRUE),
-            diff = mean(diff, na.rm = TRUE)) %>% 
+            tmin = mean(tmin, na.rm = TRUE),
+            diff = mean(tmax - tmin, na.rm = TRUE)) %>% 
   ungroup(month) %>% 
-  mutate(dummy_date = as.Date(glue::glue("2021-{month}-01"))) %>% 
   tamp() %>%
-  mutate(a = nrow(ts)) %>% 
-  filter(a == 12) %>% 
-  mutate(diff_winter = ts %>% filter(month %in% c(6,7,8)) %>% pull(diff) %>% mean(na.rm = TRUE),
-         diff_summer = ts %>% filter(month %in% c(12, 1,2)) %>% pull(diff) %>% mean(na.rm = TRUE),
-         diff_year = diff_winter - diff_summer) %>% 
+  mutate(temp_diff_var = ts %>% pull(diff) %>% var(na.rm = TRUE)) %>% 
   stretch() %>% 
-  mutate(dummy_date = as.Date(glue::glue("2021-{month}-01"))) %>%
-  migrate(name) 
-
-i <- 1
+  mutate(dummy_date = as.Date(glue::glue("2021-{month}-01"))) %>% 
+  migrate(name)
+  
 df_id <- clean$id %>% unique()
 p <- purrr::map(1:length(df_id), function(i){
   dt <- clean %>% filter(id == df_id[i])
   dt %>% 
-    ggplot(aes(x = dummy_date, ymin = tmax - diff, ymax = tmax, group = id)) + 
+    ggplot(aes(x = dummy_date, ymin = tmin, ymax = tmax, group = id)) + 
     geom_ribbon(color = "grey50", fill = "grey50", alpha = 0.5) + 
     theme_bw() + 
     scale_x_date(date_labels = "%b", date_breaks = "1 month") + 
[…]

### 2022-01-17 3-advanced.Rmd
section on glyph map

@@
 ```
+
+## Glyph map
+
+Glyph map [@Wickham2012-yr] plots the time series as single glyph on the map. 
+<!-- Variation of glyph maps has been developed to visualise uncertainty in **Visumap** [@Lucchesi2021] and deal with irregular local grid [@Beecham2021-vk].  -->
+In \proglang{R}, `GGally` implements the glyph map through the `glyphs()` function, which outputs a data frame with calculated position (`gx`, `gy`, `gid`) of each point on the time series given the major and minor xy variable using linear algebra (Equation 1 and 2 in @Wickham2012-yr).
+The data can then be piped into `ggplot` to create the glyph map:
+
+```{r eval = FALSE, echo = TRUE}
+library(ggplot2)
+gly <- glyphs(data, 
+              x_major = ..., x_minor = ..., 
+              y_major = ..., y_minor = ..., ...)
+
+ggplot(gly, aes(gx, gy, group = gid)) + 
+  geom_path() 
+```
+
+While this calculation can be seen as part of the `setup_data()` in a `ggproto`.
+A re-implementation of the glyph map as `GeomGlyph` has been made in `cubble` to create the glyph map with `geom_glyph`:
+
+```{r eval = FALSE, echo = TRUE}
+ggplot(data = data) +
+  geom_glyph(aes(x_major = ..., x_minor = ..., 
+                 y_major = ..., y_minor = ...))
+```
+
+Polar glyph map can be specify as a parameter, `polar = TRUE`, in the `geom_glyph()`, along with `width` and `height` in either absolute or relative value.
+Global and local scale can also be controlled by the parameter `global_rescale`, which default to `TRUE`.
+Reference box and line can be added with separate `geom_glyph_box()` and `geom_glyph_line()`.

### 2022-01-17 main.tex
section on glyph map

@@
 
+\hypertarget{glyph-map}{%
+\subsection{Glyph map}\label{glyph-map}}
+
+Glyph map \citep{Wickham2012-yr} plots the time series as single glyph
+on the map. In \proglang{R}, \texttt{GGally} implements the glyph map
+through the \texttt{glyphs()} function, which outputs a data frame with
+calculated position (\texttt{gx}, \texttt{gy}, \texttt{gid}) of each
+point on the time series given the major and minor xy variable using
+linear algebra (Equation 1 and 2 in \citet{Wickham2012-yr}). The data
+can then be piped into \texttt{ggplot} to create the glyph map:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> library(ggplot2)
+R> gly <- glyphs(data, 
++               x_major = ..., x_minor = ..., 
++               y_major = ..., y_minor = ..., ...)
+R> 
+R> ggplot(gly, aes(gx, gy, group = gid)) + 
++   geom_path() 
+\end{CodeInput}
+\end{CodeChunk}
+
+While this calculation can be seen as part of the \texttt{setup\_data()}
+in a \texttt{ggproto}. A re-implementation of the glyph map as
+\texttt{GeomGlyph} has been made in \texttt{cubble} to create the glyph
+map with \texttt{geom\_glyph}:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> ggplot(data = data) +
++   geom_glyph(aes(x_major = ..., x_minor = ..., 
++                  y_major = ..., y_minor = ...))
+\end{CodeInput}
+\end{CodeChunk}
+
+Polar glyph map can be specify as a parameter, \texttt{polar\ =\ TRUE},
+in the \texttt{geom\_glyph()}, along with \texttt{width} and
+\texttt{height} in either absolute or relative value. Global and local
+scale can also be controlled by the parameter \texttt{global\_rescale},
+which default to \texttt{TRUE}. Reference box and line can be added with
+separate \texttt{geom\_glyph\_box()} and \texttt{geom\_glyph\_line()}.
+
 \newpage

### 2022-01-18 4-5-interactive.Rmd
quick fix after the meeting

@@
 
-```{r interactive-linking, echo = FALSE, out.width="100%", out.height="20%", fig.retina = 2, dpi = 300, fig.cap = "Screenshots of variance of monthly temperature range from 2016 to 2020 in Australia.
-Crossed linking between the map and the temperature band is implemented in both directions to explore the temperature pattern interactively.
-The first row is the initial view of the graphic and the latter three show the interactive view after selecting on the map (row 2 and 3) and on the temperature band (row 4).", fig.show='hold'}
+```{r interactive-linking, echo = FALSE, out.width="100%", out.height="20%", fig.retina = 2, dpi = 300, fig.cap = "Exploring temperature variation using linkning of a map and seasonal display.
+Each row is a screen dump of the process.
+The top row shows all locations and all temperature profiels.
+Selecting a location with high variance on the map produces the plot in the second row.
+The maximum nad minimum temperature is shown using a ribbon.", fig.show='hold'}
 knitr::include_graphics(here::here("figures/linking.png"))
@@
 Figure \ref{fig:interactive-popup} shows the same content as Figure \ref{fig:interactive-linking} but made with leaflet and popups.
-A few sites with low temperature variance are selected and these sites still show different patterns: Lady Elloit Island in the Pacific coastline shows a thin temperature band of 5 degree; Birdsville Airport, in the middle of Australia desert, shows a wide temperature range of 10 degree throughout the year; and King Island Airport, in the north-west tip of Tasmania island shows a similar narrow range as Lady Elloits but a maximum and minimum temperature in general.
-This would suggest variance on itself may not be enough to summarise the temperature variation in Australia and other summary statistics, for example, mean of temperature range or maximum temperature itself, could also be useful to find temperature pattern in Australia.
 
@@
 Upon clicking a single station on the leaftlet up, the temperature band will be shown as a subplot in the popup box.", fig.show='hold'}
-knitr::include_graphics(here::here("figures/popup.png"))
-knitr::include_graphics(here::here("figures/popup-right.png"))
 knitr::include_graphics(here::here("figures/popup-mid.png")) 
-knitr::include_graphics(here::here("figures/popup-south.png"))
 ```

### 2022-01-18 main.tex
quick fix after the meeting

@@
 
-\caption[Screenshots of variance of monthly temperature range from 2016 to 2020 in Australia]{Screenshots of variance of monthly temperature range from 2016 to 2020 in Australia.
-Crossed linking between the map and the temperature band is implemented in both directions to explore the temperature pattern interactively.
-The first row is the initial view of the graphic and the latter three show the interactive view after selecting on the map (row 2 and 3) and on the temperature band (row 4).}\label{fig:interactive-linking}
+\caption[Exploring temperature variation using linkning of a map and seasonal display]{Exploring temperature variation using linkning of a map and seasonal display.
+Each row is a screen dump of the process.
+The top row shows all locations and all temperature profiels.
+Selecting a location with high variance on the map produces the plot in the second row.
+The maximum nad minimum temperature is shown using a ribbon.}\label{fig:interactive-linking}
 \end{figure}
@@
 Figure \ref{fig:interactive-popup} shows the same content as Figure
-\ref{fig:interactive-linking} but made with leaflet and popups. A few
-sites with low temperature variance are selected and these sites still
-show different patterns: Lady Elloit Island in the Pacific coastline
-shows a thin temperature band of 5 degree; Birdsville Airport, in the
-middle of Australia desert, shows a wide temperature range of 10 degree
-throughout the year; and King Island Airport, in the north-west tip of
-Tasmania island shows a similar narrow range as Lady Elloits but a
-maximum and minimum temperature in general. This would suggest variance
-on itself may not be enough to summarise the temperature variation in
-Australia and other summary statistics, for example, mean of temperature
-range or maximum temperature itself, could also be useful to find
-temperature pattern in Australia.
+\ref{fig:interactive-linking} but made with leaflet and popups.
 
@@
 
-{\centering \includegraphics[width=0.45\linewidth,height=0.25\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/popup} \includegraphics[width=0.45\linewidth,height=0.25\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/popup-right} \includegraphics[width=0.45\linewidth,height=0.25\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/popup-mid} \includegraphics[width=0.45\linewidth,height=0.25\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/popup-south}
+{\centering \includegraphics[width=0.45\linewidth,height=0.25\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/popup-mid}
 

### 2022-01-18 scripts/interactive-linking.R

@@ -40,7 +40,9 @@ p1 <- nested %>%
 
 p2 <- long %>% 
   ggplot(aes(x = dummy_date, group = id, fill = temp_diff_var, color = temp_diff_var)) +
-  geom_ribbon(aes(ymin = tmin, ymax = tmax), size = 0.1) + 
+  geom_ribbon(aes(ymin = tmin, ymax = tmax), size = 0.1, alpha = 0.3) + 
+  geom_line(aes(y = tmin), alpha = 0.1) + 
+  geom_line(aes(y = tmax), alpha = 0.1) + 
   geom_point(aes(y = tmax), size = 0.1) + 
   geom_point(aes(y = tmin), size = 0.1) + 
   colorspace::scale_fill_continuous_sequential(
@@ -59,7 +61,7 @@ out <- bscols(
   ggplotly(p1, width = 900, height = 700, tooltip = "label") %>%
     highlight(on = "plotly_selected", off = "plotly_deselect", opacityDim = 0.1),
   ggplotly(p2, width = 1200, height = 700, tooltip = "label") %>% 
-    highlight(on = "plotly_selected", off = "plotly_deselect", opacityDim = 0.01),
+    highlight(on = "plotly_selected", off = "plotly_deselect", opacityDim = 0.012),
   widths = c(5, 5)
 )
 

### 2022-01-19 3-advanced.Rmd
polish hierarchical structure section

@@
 
-Imposing a clustering structure can be thought of as building a hierarchical structure where stations are nested within clusters.
-This can be useful when there's intrinsic nesting structure in the data (i.e.
-country nested in the continent, county nested in the state) or there's some clustering.
-When we have access to both level, `switch_key()` is the function to re-structure the data as one cluster per row.
-Temporal observations from different stations while within the same cluster are bound in the nested column `ts`.
+Spatial locations can have further grouping structure either in nature (i.e.
+countries are nested in continents), or calculated by some clustering algorithms.
+Rather than investigate the read from individual locations, group character summarised from multiple locations can give a clearer picture of the local dynamic.
+In cubble, `switch_key()` can be used to create a new level of grouping of spatial location by specifying a clustering variable.
+Figure \ref{fig:illu-hier} illustrates the relationship of cubbles at site and cluster level, in both the long and nested form.
+By specifying `cluster_nested <- station_nested %>% switch_key(key = cluster)`, the cubble re-defines the cubble key from `id` in `station_nested` to `cluster` in `cluster_nested`.
+All the spatial variables variant to `cluster` are now nested into a `.val` column and cluster level variables can be computed in the same fashion as station level variables in `station_nested`.
+Following the same principal of `stretch`, the long form cluster level cubble expands the time series variables with both indices: `cluster` and `id`.
+All the cluster level variables are stored separately for recovering the nested from from the long form.
 
-One thing we hope to do with the cluster is to find the coordinates of the centroid.
-These are variables variant to the station but invariant to the cluster and it would be nice to have a function that structure each cluster as a row.
-`switch_key()` is the function that does this: it lets you to specify a new key, say `cluster` and nests all spatial variables variant to `cluster` into a column.
-Spatial variables are all nested inside a new column `.val`.
-Temporal observations from different stations while within the same cluster are bound in the nested column `ts`.
-
-This structure makes it easy to compute cluster level variable, for example, the convex hull and the centroid coordinate of each cluster.
-These can be amended into the nested form with function `get_centroid`.
-
-After we have got `cluster_nested`, spatial and temporal data at both levels can be easily obtained.
-Figure \ref{fig:illu-hier} illustrate the relationship between the long and nested form cubble at both site and cluster level.
-More description on this.
-Start with the original `station_nested`, `stretch()` expands the `ts` column with each station (`id`) forming a group and attach variables invariant to `id` as an attribute.
-`switch_key()` changes the `key` from `id` to `cluster` and nests all the spatial variables that variant to `cluster`.
-`stretch()` `cluster_nested` will store variables that are invariant to `cluster` as a tibble in the attribute.
-
-this fit into the remaining data pipeline.
-
-```{r illu-hier, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "Hierarchical structure"}
+```{r illu-hier, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "Relationship between the station and cluster level cubble in both long and nested form."}
 knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.003.png"))

### 2022-01-19 main.tex
polish hierarchical structure section

@@
 
-Imposing a clustering structure can be thought of as building a
-hierarchical structure where stations are nested within clusters. This
-can be useful when there's intrinsic nesting structure in the data
-(i.e.~country nested in the continent, county nested in the state) or
-there's some clustering. When we have access to both level,
-\texttt{switch\_key()} is the function to re-structure the data as one
-cluster per row. Temporal observations from different stations while
-within the same cluster are bound in the nested column \texttt{ts}.
-
-One thing we hope to do with the cluster is to find the coordinates of
-the centroid. These are variables variant to the station but invariant
-to the cluster and it would be nice to have a function that structure
-each cluster as a row. \texttt{switch\_key()} is the function that does
-this: it lets you to specify a new key, say \texttt{cluster} and nests
-all spatial variables variant to \texttt{cluster} into a column. Spatial
-variables are all nested inside a new column \texttt{.val}. Temporal
-observations from different stations while within the same cluster are
-bound in the nested column \texttt{ts}.
-
-This structure makes it easy to compute cluster level variable, for
-example, the convex hull and the centroid coordinate of each cluster.
-These can be amended into the nested form with function
-\texttt{get\_centroid}.
-
-After we have got \texttt{cluster\_nested}, spatial and temporal data at
-both levels can be easily obtained. Figure \ref{fig:illu-hier}
-illustrate the relationship between the long and nested form cubble at
-both site and cluster level. More description on this. Start with the
-original \texttt{station\_nested}, \texttt{stretch()} expands the
-\texttt{ts} column with each station (\texttt{id}) forming a group and
-attach variables invariant to \texttt{id} as an attribute.
-\texttt{switch\_key()} changes the \texttt{key} from \texttt{id} to
-\texttt{cluster} and nests all the spatial variables that variant to
-\texttt{cluster}. \texttt{stretch()} \texttt{cluster\_nested} will store
-variables that are invariant to \texttt{cluster} as a tibble in the
-attribute.
-
-this fit into the remaining data pipeline.
+Spatial locations can have further grouping structure either in nature
+(i.e.~countries are nested in continents), or calculated by some
+clustering algorithms. Rather than investigate the read from individual
+locations, group character summarised from multiple locations can give a
+clearer picture of the local dynamic. In cubble, \texttt{switch\_key()}
+can be used to create a new level of grouping of spatial location by
+specifying a clustering variable. Figure \ref{fig:illu-hier} illustrates
+the relationship of cubbles at site and cluster level, in both the long
+and nested form. By specifying
+\texttt{cluster\_nested\ \textless{}-\ station\_nested\ \%\textgreater{}\%\ switch\_key(key\ =\ cluster)},
+the cubble re-defines the cubble key from \texttt{id} in
+\texttt{station\_nested} to \texttt{cluster} in
+\texttt{cluster\_nested}. All the spatial variables variant to
+\texttt{cluster} is now nested into a \texttt{.val} column and cluster
+level variables can be computed in the same fashion as station level
+variables in \texttt{station\_nested}. Following the same principal of
+\texttt{stretch}, the long form cluster level cubble expands the time
+series variables with both indices: \texttt{cluster} and \texttt{id}.
+All the cluster level variables are stored separately for recovering the
+nested from from the long form.
 
@@
 
-\caption[Hierarchical structure]{Hierarchical structure}\label{fig:illu-hier}
+\caption[Relationship between the station and cluster level cubble in both long and nested form]{Relationship between the station and cluster level cubble in both long and nested form.}\label{fig:illu-hier}
 \end{figure}

### 2022-01-19 4-2-agg.Rmd
aggregation example WIP

@@
 
-In Figure \ref{fig:basic-agg}.
-Aggregating them into averages or other statistics would give a better picture.
-This can be done with cubble through a hierarchical structure, where the cluster indicator is one level above the individual station.
+With recent climate, there are `r nrow(weatherdata::climate_full)` stations in Australia that records daily maximum and minimum temperature and precipitation.
+While it would be too much series to show on a glyph map, stations close to each other should have similar precipitation pattern.
+Aggregation can be helpful here to cluster series into groups and visualise them with glyph map.
 
-The `weatherdata::climate_full` data has daily climate data of `r nrow(weatherdata::climate_full)` Australia stations from 2016 to 2020.
+A simple kmean algorithm is used based on the distance matrix to cluster the stations into 20 groups. 
+This creates `station_nested` as a station level nested cubble with a cluster column indicating which group each station belongs to.
+More complex algorithms can be used for more complex problem, as long as a mapping from each station id to the cluster id can be constructed.
 
-calcluate distance matrix
-As an example to illustrate here, a kmean clustering algorithm based on the distance matrix is used and the number of centres is set to 20.
-More complex algorithms can also be used for more complex problem, as long as a mapping from each station id to the cluster id can be constructed.
+```{r}
+load(here::here("data/station_nested.rda"))
+station_nested
+```
 
-<!-- Now we can obtain the aggregated series as described in the workflow diagram above and construct the glyph map with `GGally::glyphs()`: -->
+To create a group level cubble, use `switch_key()` with the new key variable, `cluster`:
 
-<!-- We can also look at the precipitation of each individual station within the same cluster: -->
+```{r echo = TRUE}
+cluster_nested <- station_nested %>% switch_key(cluster) 
+```
 
-<!-- Lastly, there is one series on Tasmania island standing out from others, lets look at where it is: -->
+With the group level cubble, it is useful to compute the centroid of each cluster as the location of group level glyph map.
+`get_centroid()` is a shortcut for doing so or you can manually find the convex hull, its centroid, and extract the longitude and latitude:
+
+```{r eval = FALSE, echo = TRUE}
+cluster_nested <- cluster_nested %>% 
+  get_centroid()
+```
+
+Long form cubble at both levels can be access through stretching the nested form. Here weekly averages of the precipitation is calculated:
+
+```{r eval = FALSE, echo = TRUE}
+station_long <- station_nested %>% 
+  stretch(ts) %>% 
+  mutate(wk = lubridate::week(date)) %>% 
+  group_by(wk) %>% 
+  summarise(prcp = sum(prcp, na.rm = TRUE))
+
+cluster_long <- cluster_nested %>% 
+  stretch(ts) %>% 
+  mutate(wk = lubridate::week(date)) %>% 
+  group_by(wk) %>% 
+  summarise(prcp = mean(prcp, na.rm = TRUE)) %>% 
+  migrate(cent_long, cent_lat)
+```
+
+With the four cubbles: `station_nested`, `station_long`, `cluster_nested`, and `cluster_long`, all the information at both levels are accessible.
+This allows us to make glyph map on the cluster level, looking at station composition of each cluster, and plot the individual series to check for consistency within groups.
+Figure \ref{fig:basic-agg} composes of the three plots together to show the precipitation in Australia.
 

### 2022-01-19 main.tex
aggregation example WIP

@@
 \texttt{cluster\_nested}. All the spatial variables variant to
-\texttt{cluster} is now nested into a \texttt{.val} column and cluster
+\texttt{cluster} are now nested into a \texttt{.val} column and cluster
 level variables can be computed in the same fashion as station level
@@
 
-In Figure \ref{fig:basic-agg}. Aggregating them into averages or other
-statistics would give a better picture. This can be done with cubble
-through a hierarchical structure, where the cluster indicator is one
-level above the individual station.
+With recent climate, there are 639 stations in Australia that records
+daily maximum and minimum temperature and precipitation. While it would
+be too much series to show on a glyph map, stations close to each other
+should have similar precipitation pattern. Aggregation can be helpful
+here to cluster series into groups and visualise them with glyph map.
 
-The \texttt{weatherdata::climate\_full} data has daily climate data of
-639 Australia stations from 2016 to 2020.
+A simple kmean algorithm is used based on the distance matrix to cluster
+the stations into 20 groups. This creates \texttt{station\_nested} as a
+station level nested cubble with a cluster column indicating which group
+each station belongs to. More complex algorithms can be used for more
+complex problem, as long as a mapping from each station id to the
+cluster id can be constructed.
 
-calcluate distance matrix As an example to illustrate here, a kmean
-clustering algorithm based on the distance matrix is used and the number
-of centres is set to 20. More complex algorithms can also be used for
-more complex problem, as long as a mapping from each station id to the
-cluster id can be constructed.
+\begin{CodeChunk}
+\begin{CodeOutput}
+# cubble:   id [639]: nested form
+# bbox:     [113.53, -43.66, 153.64, -10.05]
+# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
+  id            lat  long  elev name             wmo_id ts               cluster
+  <chr>       <dbl> <dbl> <dbl> <chr>             <dbl> <list>             <int>
+1 ASN00001006 -15.5  128.   3.8 wyndham aero      95214 <tibble [1,827 ~       3
+2 ASN00001007 -13.8  126.   6   troughton island  94102 <tibble [1,827 ~       3
+3 ASN00001018 -16.4  126. 546   mount elizabeth   94211 <tibble [1,460 ~      16
+4 ASN00001019 -14.3  127.  23   kalumburu         94100 <tibble [1,827 ~       3
+5 ASN00001020 -14.1  126.  51   truscott          95101 <tibble [1,827 ~       3
+# ... with 634 more rows
+\end{CodeOutput}
+\end{CodeChunk}
+
+To create a group level cubble, use \texttt{switch\_key()} with the new
+key variable, \texttt{cluster}:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> cluster_nested <- station_nested %>% switch_key(cluster) 
+\end{CodeInput}
+\end{CodeChunk}
+
+With the group level cubble, it is useful to compute the centroid of
+each cluster as the location of group level glyph map.
+\texttt{get\_centroid()} is a shortcut for doing so or you can manually
+find the convex hull, its centroid, and extract the longitude and
+latitude:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> cluster_nested <- cluster_nested %>% 
++   get_centroid()
+\end{CodeInput}
+\end{CodeChunk}
+
+Long form cubble at both levels can be access through stretching the
+nested form. Here weekly averages of the precipitation is calculated:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> station_long <- station_nested %>% 
++   stretch(ts) %>% 
++   mutate(wk = lubridate::week(date)) %>% 
++   group_by(wk) %>% 
++   summarise(prcp = sum(prcp, na.rm = TRUE))
+R> 
+R> cluster_long <- cluster_nested %>% 
++   stretch(ts) %>% 
++   mutate(wk = lubridate::week(date)) %>% 
++   group_by(wk) %>% 
++   summarise(prcp = mean(prcp, na.rm = TRUE)) %>% 
++   migrate(cent_long, cent_lat)
+\end{CodeInput}
+\end{CodeChunk}
+
+With the four cubbles: \texttt{station\_nested}, \texttt{station\_long},
+\texttt{cluster\_nested}, and \texttt{cluster\_long}, all the
+information at both levels are accessible. This allows us to make glyph
+map on the cluster level, looking at station composition of each
+cluster, and plot the individual series to check for consistency within
+groups. Figure \ref{fig:basic-agg} composes of the three plots together
+to show the precipitation in Australia.
 

### 2022-01-19 scripts/basic-agg.R

@@ -1,51 +1,56 @@
 library(tidyverse)
 library(cubble)
+library(patchwork)
 coords <- cbind(weatherdata::climate_full$long, weatherdata::climate_full$lat)
 dist_raw <- geosphere::distm(coords, coords)
 
+set.seed(123)
 station_nested <- weatherdata::climate_full %>% 
   strip_rowwise() %>% 
   mutate(cluster = kmeans(dist_raw,centers = 20, nstart = 500)$cluster)
+save(station_nested, file = here::here("data/station_nested.rda"))
 
 cluster_nested <- station_nested %>%
   switch_key(cluster) %>% 
   get_centroid()
 
-station_long <- cluster_nested %>% 
+station_long <- station_nested %>% 
   stretch(ts) %>% 
   mutate(wk = lubridate::week(date)) %>% 
-  group_by(wk, id) %>% 
+  group_by(wk) %>% 
   summarise(prcp = sum(prcp, na.rm = TRUE))
 
-cluster_long <- station_long %>% 
-  ungroup(id) %>% 
+cluster_long <- cluster_nested %>% 
+  stretch(ts) %>% 
+  mutate(wk = lubridate::week(date)) %>% 
+  group_by(wk) %>% 
   summarise(prcp = mean(prcp, na.rm = TRUE)) %>% 
   migrate(cent_long, cent_lat)
-
-gly <- GGally::glyphs(cluster_long,
-                      x_major = "cent_long", x_minor = "wk",
-                      y_major = "cent_lat", y_minor = "prcp",
-                      height = 2, width = 4
-)
 
 state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)
 p1 <- plot_map(state_map) +
   geom_text(data = cluster_nested, 
             aes(x = cent_long, y = cent_lat, label = cluster)) +
-  geom_path(data = gly,aes(x = gx, y = gy, group = gid))
+  geom_glyph(data = cluster_long, 
+             aes(x_major = cent_long, x_minor = wk, y_major = cent_lat, y_minor = prcp), 
+             height = 2, width = 4)
 
+tas_latlong <- station_long %>% 
+  ungroup(wk, id) %>% 
+  filter(cluster == 7) %>% 
+  filter(prcp == max(prcp)) %>% 
+  tamp()
+
+tas_id <- tas_latlong %>% pull(id)
[…]

### 2022-01-19 scripts/basic-agg.R
improve aggregation plot

@@ -18,46 +18,60 @@ station_long <- station_nested %>%
   stretch(ts) %>% 
   mutate(wk = lubridate::week(date)) %>% 
   group_by(wk) %>% 
-  summarise(prcp = sum(prcp, na.rm = TRUE))
+  summarise(prcp = sum(prcp, na.rm = TRUE)) %>% 
+  migrate(cluster)
 
 cluster_long <- cluster_nested %>% 
   stretch(ts) %>% 
   mutate(wk = lubridate::week(date)) %>% 
-  group_by(wk) %>% 
+  group_by(id, wk) %>% 
+  summarise(prcp = sum(prcp, na.rm = TRUE)) %>% 
+  ungroup(id) %>% 
   summarise(prcp = mean(prcp, na.rm = TRUE)) %>% 
   migrate(cent_long, cent_lat)
 
 state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)
-p1 <- plot_map(state_map) +
+
+ggplot_smooth <- cluster_long %>% 
+  ggplot() +
+  geom_smooth(aes(x = wk, y = prcp, group = cluster)) 
+
+smoother <- layer_data(ggplot_smooth) %>% 
+  left_join(cluster_long %>% select(cluster, cent_long, cent_lat), by = c("group" = "cluster"))
+
+p1 <- ggplot(data = smoother, 
+             aes(x_minor = x, y_minor = y, x_major = cent_long, y_major = cent_lat)) + 
+  geom_sf(data = state_map, inherit.aes = FALSE,
+          color = "grey80", alpha = 0.4, linetype = 3) + 
   geom_text(data = cluster_nested, 
-            aes(x = cent_long, y = cent_lat, label = cluster)) +
-  geom_glyph(data = cluster_long, 
-             aes(x_major = cent_long, x_minor = wk, y_major = cent_lat, y_minor = prcp), 
-             height = 2, width = 4)
+            aes(x = cent_long, y = cent_lat, label = cluster), inherit.aes = FALSE) +
+  geom_glyph(height = 2, width = 4) + 
+  theme_void()
 
-tas_latlong <- station_long %>% 
-  ungroup(wk, id) %>% 
-  filter(cluster == 7) %>% 
-  filter(prcp == max(prcp)) %>% 
-  tamp()
+# tas_latlong <- station_long %>% 
+#   filter(cluster == 7) %>% 
+#   switch_key(cluster) %>% 
+#   filter(prcp == max(prcp)) %>% 
[…]

### 2022-01-27 4-3-water-level.Rmd
clean up water level example code and a draft on its example section

@@
 
-The water level data comes from [Bureau of Meteorology](http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full) and has a copy in `weatherdata`.
-Here we extract the water course level and add a column annotate this data of type `river`.
-For the rainfall data, we will still use the `weatherdata::climate_full`, filtering for Victorian stations in 2020 should be pretty familiar by now.
-Again, we first look at where these stations are on the map first:
+Water data is collected by the [Bureau of Meteorology](http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full).
+This includes ...
+water gauges, variables collected.
+In particular, water level will interactive with the climate variable precipitation as rains fall into the river will raise the water level.
+Figure \ref{fig:matching-map} shows the available weather station and water gauges in Victoria, Melbourne.
 
-Now we use `match_sites()` to first pair the weather stations with the river gauges spatially and then apply the temporal matching on `prcp`.
+```{r echo = TRUE}
+climate <- weatherdata::climate_full %>%
+  filter(between(stringr::str_sub(id, 7, 8), 76, 90)) %>%
+  stretch() %>%
+  filter(lubridate::year(date) == 2020) %>%
+  tamp() %>%
+  mutate(type = "climate")
+
+river <- weatherdata::water %>%
+  stretch() %>%
+  select(date, Water_course_level) %>%
+  rename(prcp = Water_course_level) %>% 
+  tamp() %>%
+  mutate(type = "river")
+```
+
+```{r matching-map, out.width="100%", fig.cap="asdfasd"}
+knitr::include_graphics("figures/matching-map.png")
+```
+
+From the map, a few water gauges and weather stations are close to each other and we may be able to observe a spontaneous increase in the precipitation and water level from the measures, potential with a little delay in the river gauges.
+`match_sites()` matches one source of data with another source in both spatial and temporal dimension in a cubble.
+
+<!-- Now we use `match_sites()` to first pair the weather stations with the river gauges spatially and then apply the temporal matching on `prcp`.
 We will construct the interval based on peaks in `climate` since we would expect a lag effect for precipitation to flow into the river and cause a raise in river level, hence `temporal_independent = climate`.
@@
 This is a tuning parameter and you can start with 10% of the points of one series (here we have daily data for a year, 10% is roughly 30 points).
-`temporal_min_match` filters out pairs don't have enough match and to return all the pairs, set `temporal_min_match` to `0`.``
+`temporal_min_match` filters out pairs don't have enough match and to return all the pairs, set `temporal_min_match` to `0`.`` -->
 
-The output from temporal matching is also a cubble, with additional column `.dist` and `.group` inherent from spatial matching and `n_match` for the number of matched temporal peaks.
-Then you can use this output to plot the location of match or to look at the series:
+```{r echo = TRUE}
+res <- match_sites(river, climate,
+                   temporal_var_to_match = prcp,
+                   temporal_independent = climate,  
+                   temporal_n_highest = 30,
+                   temporal_min_match = 15)
+```
 
-```{r matching, out.width="100%", out.height="30%", fig.cap="asdfasd"}
+The output from temporal matching is also a cubble, with additional column `dist` and `group` produced from spatial matching and `n_match` from the number of peak matched temporally.
+
+```{r}
+res
+```
+
+Then plots can be made to inspect the matched pairs on the map or to view the co-movement of the two series:
+
+```{r matching, out.width="100%", fig.cap="asdfasd"}
 knitr::include_graphics("figures/matching.png")
 ```
+
+With the chosen parameter, four pairs of stations stands out, which all locates in the middle Victoria and the concurrent increase of precipitation and water level can be observed.

### 2022-01-27 main.tex
improve aggregation plot
clean up water level example code and a draft on its example section

@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.9\textheight]{figures/basic-agg}
+{\centering \includegraphics[width=1\linewidth]{figures/basic-agg}
 
@@
 
-\caption[sdfasdf]{sdfasdf}\label{fig:basic-agg}
+\caption[Aggregated precipitation in Australia]{Aggregated precipitation in Australia.}\label{fig:basic-agg}
 \end{figure}
@@
 
-The water level data comes from
+Water data is collected by the
 \href{http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full}{Bureau
-of Meteorology} and has a copy in \texttt{weatherdata}. Here we extract
-the water course level and add a column annotate this data of type
-\texttt{river}. For the rainfall data, we will still use the
-\texttt{weatherdata::climate\_full}, filtering for Victorian stations in
-2020 should be pretty familiar by now. Again, we first look at where
-these stations are on the map first:
+of Meteorology}. This includes \ldots{} water gauges, variables
+collected. In particular, water level will interactive with the climate
+variable precipitation as rains fall into the river will raise the water
+level. Figure \ref{fig:matching-map} shows the available weather station
+and water gauges in Victoria, Melbourne.
 
-Now we use \texttt{match\_sites()} to first pair the weather stations
-with the river gauges spatially and then apply the temporal matching on
-\texttt{prcp}. We will construct the interval based on peaks in
-\texttt{climate} since we would expect a lag effect for precipitation to
-flow into the river and cause a raise in river level, hence
-\texttt{temporal\_independent\ =\ climate}. We select the 30 highest
-peak from the series to construct the match by setting
-\texttt{temporal\_n\_highest\ =\ 30}. This is a tuning parameter and you
-can start with 10\% of the points of one series (here we have daily data
-for a year, 10\% is roughly 30 points). \texttt{temporal\_min\_match}
-filters out pairs don't have enough match and to return all the pairs,
-set \texttt{temporal\_min\_match} to \texttt{0}.``
-
-The output from temporal matching is also a cubble, with additional
-column \texttt{.dist} and \texttt{.group} inherent from spatial matching
-and \texttt{n\_match} for the number of matched temporal peaks. Then you
-can use this output to plot the location of match or to look at the
-series:
+\begin{CodeChunk}
+\begin{CodeInput}
+R> climate <- weatherdata::climate_full %>%
++   filter(between(stringr::str_sub(id, 7, 8), 76, 90)) %>%
++   stretch() %>%
++   filter(lubridate::year(date) == 2020) %>%
++   tamp() %>%
++   mutate(type = "climate")
+R> 
+R> river <- weatherdata::water %>%
++   stretch() %>%
++   select(date, Water_course_level) %>%
++   rename(prcp = Water_course_level) %>% 
++   tamp() %>%
++   mutate(type = "river")
+\end{CodeInput}
+\end{CodeChunk}
 
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.3\textheight]{figures/matching}
+{\centering \includegraphics[width=1\linewidth]{figures/matching-map}
+
+}
+
+\caption[asdfasd]{asdfasd}\label{fig:matching-map}
+\end{figure}
+\end{CodeChunk}
+
+From the map, a few water gauges and weather stations are close to each
+other and we may be able to observe a spontaneous increase in the
+precipitation and water level from the measures, potential with a little
+delay in the river gauges. \texttt{match\_sites()} matches one source of
+data with another source in both spatial and temporal dimension in a
+cubble.
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> res <- match_sites(river, climate,
++                    temporal_var_to_match = prcp,
++                    temporal_independent = climate,  
++                    temporal_n_highest = 30,
++                    temporal_min_match = 15)
+\end{CodeInput}
+\end{CodeChunk}
+
+The output from temporal matching is also a cubble, with additional
+column \texttt{dist} and \texttt{group} produced from spatial matching
+and \texttt{n\_match} from the number of peak matched temporally.
+
+\begin{CodeChunk}
+\begin{CodeOutput}
+# cubble:   id [8]: nested form
+# bbox:     [144.52, -37.73, 146.06, -36.55]
+# temporal: date [date], prcp [dbl]
+  id          name                lat  long ts         type   dist group n_match
+  <chr>       <chr>             <dbl> <dbl> <list>     <chr> <dbl> <int>   <int>
+1 405234      SEVEN CREEKS @ D~ -36.9  146. <tibble [~ river  6.15     5      21
+2 ASN00082042 strathbogie       -36.8  146. <tibble [~ clim~  6.15     5      21
+3 404207      HOLLAND CREEK @ ~ -36.6  146. <tibble [~ river  8.54    10      21
+4 ASN00082170 benalla airport   -36.6  146. <tibble [~ clim~  8.54    10      21
+5 230200      MARIBYRNONG RIVE~ -37.7  145. <tibble [~ river  6.17     6      19
+# ... with 3 more rows
+\end{CodeOutput}
+\end{CodeChunk}
+
+Then plots can be made to inspect the matched pairs on the map or to
+view the co-movement of the two series:
+
+\begin{CodeChunk}
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{figures/matching}
 
@@
 \end{CodeChunk}
+
+With the chosen parameter, four pairs of stations stands out, which all
+locates in the middle Victoria and the concurrent increase of
+precipitation and water level can be observed.
 

### 2022-01-27 scripts/matching.R

@@ -1,11 +1,6 @@
 library(tidyverse)
 library(cubble)
 library(patchwork)
-river <- weatherdata::water %>%
-  stretch() %>%
-  select(date, Water_course_level) %>%
-  tamp() %>%
-  mutate(type = "river")
 
 climate <- weatherdata::climate_full %>%
   filter(between(stringr::str_sub(id, 7, 8), 76, 90)) %>%
@@ -14,64 +9,56 @@ climate <- weatherdata::climate_full %>%
   tamp() %>%
   mutate(type = "climate")
 
-vic_map <- rmapshaper::ms_simplify(ozmaps::abs_ste %>% filter(NAME == "Victoria"))
-plot_map(vic_map) +
-  geom_point(
-    data = dplyr::bind_rows(river, climate),
-    aes(x = long, y = lat, color = type)
-  ) +
-  scale_color_brewer(palette = "Dark2") +
-  theme_bw()
-
-river <- river %>%
+river <- weatherdata::water %>%
   stretch() %>%
-  rename(prcp = Water_course_level) %>%
-  tamp()
+  select(date, Water_course_level) %>%
+  rename(prcp = Water_course_level) %>% 
+  tamp() %>%
+  mutate(type = "river")
+
+vic_map <- rmapshaper::ms_simplify(ozmaps::abs_ste %>% filter(NAME == "Victoria"))  
+plot_map(vic_map) +
+  geom_point(data = dplyr::bind_rows(river, climate), 
+             aes(x = long, y = lat, color = type)) + 
+  scale_color_brewer(palette = "Dark2") +
+  theme_bw() + 
+  theme(legend.position = "bottom")
+ggsave(filename = "figures/matching-map.png", width = 10, height = 5)  
 
 res <- match_sites(river, climate,
                    temporal_var_to_match = prcp,
                    temporal_independent = climate,  
                    temporal_n_highest = 30,
-                   temporal_min_match = 15
-)
+                   temporal_min_match = 15)
 
 p1 <- plot_map(vic_map) +
-  geom_point(
-    data = res,
-    aes(x = long, y = lat, color = type)
-  ) +
+  geom_point(data = res, aes(x = long, y = lat, color = type)) +
   ggrepel::geom_label_repel(
-    data = res %>% filter(type == "river"),
-    aes(x = long, y = lat, label = .group)
-  ) +
[…]

### 2022-01-28 3-advanced.Rmd
some draft on the theory section of matching

@@
 
-Temporal matching checks how spatially matched pairs align temporally. We use the following chart to illustrate how the temporal matching works:
+Data fusion is a process that consider combining multiple multiple data sources. 
+widely used in remote sensing
 
-For each spatially matched pair, say `A` and `a`, we first find the largest `n` points in each series, colored in brown points here.
-Here we use the largest three but you can tune this number by `temporal_n_highest`.
-Then we construct the interval of the largest points from one series and see how many points, from the other series, fall into the intervals.
-The series used to construct the interval is controlled by `temporal_independent` and the window size by `temporal_window` with a default of 5.
-
-In this illustration, we construct the interval based on series `A` and two of the three peaks from `a` falls into this interval at Time 7 and 27.
-
-There's another mandatory argument that hasn't been introduced above: `temporal_var_to_match`.
-This argument controls the variable to match and it needs to appear in both the `major` and `minor` set.
-In the water level matching example, we match the variable `Water_course_level` from `river` to `prcp` from `climate`, hence need to manually rename one of them to match the other, here we rename `Water_course_level` to `prcp` in `river`:
-
-Figure \ref{fig:illu-matching}
+Meteorological and hydrological data have been matched in the applied studies in the literature to understand drought, ...
+[ref].
+To be a valid pair of matching, the matched pair from different data sources need to be spatially close to each other and temporally [...].
+[
+can also think about matching as a form of correlation between series from a statistics point of view.] Cubble has `match_sites()` to match data from multiple sources spatially and temporally.
+`match_sites()` first matches the two data source spatially through computing the distance [more details on spatial matching - keep one, max dist] and then computes the number of peaks in the two time series that fall within in a fixed length window.
+Figure \ref{fig:illu-matching} illustrates the temporal matching.
+With temporal matching, a `temporal_var_to_match` needs to be first specified.
+Another parameter needs to be specified is `temporal_independent` which determines the series for construct the interval.
+In Figure \ref{fig:illu-matching}, series A moves before series a and an interval is constructed after the three highest peaks.
+Then the highest three peak from series a is identified and two of them falls into the interval constructed.
+There are two matches here at Time 5 and 27.
 
@@
 ```
+
+Several arguments are available in `match_sites()` to adjust the matching:
+
+ - `spatial_n_keep`: the number of matching to keep
+ - `spatial_dist_max`: the maximum distance allowed between matched pair
+ - `temporal_n_highest`: the number of highest peak used for temporal matching
+ - `temporal_window`: the temporal window 
+ - `temporal_min_match`: the minimum number of peak matching for temporal matching
 

### 2022-01-28 main.tex
some draft on the theory section of matching

@@
 
-Temporal matching checks how spatially matched pairs align temporally.
-We use the following chart to illustrate how the temporal matching
-works:
+Data fusion is a process that consider combining multiple multiple data
+sources. widely used in remote sensing
 
-For each spatially matched pair, say \texttt{A} and \texttt{a}, we first
-find the largest \texttt{n} points in each series, colored in brown
-points here. Here we use the largest three but you can tune this number
-by \texttt{temporal\_n\_highest}. Then we construct the interval of the
-largest points from one series and see how many points, from the other
-series, fall into the intervals. The series used to construct the
-interval is controlled by \texttt{temporal\_independent} and the window
-size by \texttt{temporal\_window} with a default of 5.
-
-In this illustration, we construct the interval based on series
-\texttt{A} and two of the three peaks from \texttt{a} falls into this
-interval at Time 7 and 27.
-
-There's another mandatory argument that hasn't been introduced above:
-\texttt{temporal\_var\_to\_match}. This argument controls the variable
-to match and it needs to appear in both the \texttt{major} and
-\texttt{minor} set. In the water level matching example, we match the
-variable \texttt{Water\_course\_level} from \texttt{river} to
-\texttt{prcp} from \texttt{climate}, hence need to manually rename one
-of them to match the other, here we rename \texttt{Water\_course\_level}
-to \texttt{prcp} in \texttt{river}:
-
-Figure \ref{fig:illu-matching}
+Meteorological and hydrological data have been matched in the applied
+studies in the literature to understand drought, \ldots{} {[}ref{]}. To
+be a valid pair of matching, the matched pair from different data
+sources need to be spatially close to each other and temporally
+{[}\ldots{]}. {[} can also think about matching as a form of correlation
+between series from a statistics point of view.{]} Cubble has
+\texttt{match\_sites()} to match data from multiple sources spatially
+and temporally. \texttt{match\_sites()} first matches the two data
+source spatially through computing the distance {[}more details on
+spatial matching - keep one, max dist{]} and then computes the number of
+peaks in the two time series that fall within in a fixed length window.
+Figure \ref{fig:illu-matching} illustrates the temporal matching. With
+temporal matching, a \texttt{temporal\_var\_to\_match} needs to be first
+specified. Another parameter needs to be specified is
+\texttt{temporal\_independent} which determines the series for construct
+the interval. In Figure \ref{fig:illu-matching}, series A moves before
+series a and an interval is constructed after the three highest peaks.
+Then the highest three peak from series a is identified and two of them
+falls into the interval constructed. There are two matches here at Time
+5 and 27.
 
@@
 \end{CodeChunk}
+
+Several arguments are available in \texttt{match\_sites()} to adjust the
+matching:
+
+\begin{itemize}
+\tightlist
+\item
+  \texttt{spatial\_n\_keep}: the number of matching to keep
+\item
+  \texttt{spatial\_dist\_max}: the maximum distance allowed between
+  matched pair
+\item
+  \texttt{temporal\_n\_highest}: the number of highest peak used for
+  temporal matching
+\item
+  \texttt{temporal\_window}: the temporal window
+\item
+  \texttt{temporal\_min\_match}: the minimum number of peak matching for
+  temporal matching
+\end{itemize}
 

### 2022-01-28 scripts/matching.R

@@ -46,7 +46,7 @@ p1 <- plot_map(vic_map) +
 res_long <- res %>%
   stretch(ts) %>%
   migrate(group, type) %>%
-  mutate(prcp = scale(prcp)[, 1])
+  mutate(prcp = (prcp - min(prcp, na.rm = TRUE))/ (max(prcp, na.rm = TRUE) - min(prcp, na.rm = TRUE))) 
 
 p2 <- res_long %>% 
   ggplot(aes(x = date, y = prcp, color = type, group = id)) +

### 2022-01-29 3-advanced.Rmd
more work on matching in the advanced feature section

--- 2022-01-19 3-advanced.Rmd
@@
 
-Temporal matching checks how spatially matched pairs align temporally. We use the following chart to illustrate how the temporal matching works:
+Data fusion is a process that consider combining multiple multiple data sources and it has been applied widely in the remote sensing community.
 
-For each spatially matched pair, say `A` and `a`, we first find the largest `n` points in each series, colored in brown points here.
-Here we use the largest three but you can tune this number by `temporal_n_highest`.
-Then we construct the interval of the largest points from one series and see how many points, from the other series, fall into the intervals.
-The series used to construct the interval is controlled by `temporal_independent` and the window size by `temporal_window` with a default of 5.
+Meteorological and hydrological data have been matched in the applied studies in the literature to understand drought, ...
+[ref].
+To be a valid pair of matching, the matched pair from different data sources need to
 
-In this illustration, we construct the interval based on series `A` and two of the three peaks from `a` falls into this interval at Time 7 and 27.
+  - be spatially close to each other, and 
+  - have similar temporal movement
 
-There's another mandatory argument that hasn't been introduced above: `temporal_var_to_match`.
-This argument controls the variable to match and it needs to appear in both the `major` and `minor` set.
-In the water level matching example, we match the variable `Water_course_level` from `river` to `prcp` from `climate`, hence need to manually rename one of them to match the other, here we rename `Water_course_level` to `prcp` in `river`:
+`match_sites()` from `cubble` matches data based on these two criteria.
+It first matches the two data sources spatially through computing the pairwise distance based on latitude and longitude.
+Pairs that pass the spatial matching are then matched temporally through computing the number of matched peaks within a fixed length window.
+Figure \ref{fig:illu-matching} illustrate this temporal matching in more details.
+Given two series `A` and `a`, 3 peaks have been picked in each series.
+An interval, with default length of 5, is constructed for each peak in series `A` and the peaks in series `a` are tested against whether they fall into the any of the intervals.
+In this illustration, there are 2 matches for these two series.
+Several arguments are available in `match_sites()` to fine-tune the matching:
 
-Figure \ref{fig:illu-matching}
+<!-- [can also think about matching as a form of correlation between series from a statistics point of view.] -->
 
-```{r illu-matching, out.width="100%", out.height="30%", fig.cap="sdfasdf"}
+```{r illu-matching, out.width="100%", fig.cap="An illustration of temporal matching in cubble.
+Three highest peaks are identified in each series and intervals are constructed on series A.
+Two peaks in series a fall into the intervals and hence the two series are considered to have 2 matches."}
 knitr::include_graphics("figures/illu-matching.png")
 ```
+
+ - `spatial_n_keep`: the number of spatial match for each site to keep
+ - `spatial_dist_max`: the maximum distance allowed for a matched pair 
+ - `temporal_n_highest`: the number of peak used - 3 in the example above
+ - `temporal_window`: the length of the interval - 5 in the example above
+ - `temporal_min_match`: the minimum number of matched peak for a valid matched pair
 

### 2022-01-29 main.tex
more work on matching in the advanced feature section

--- 2022-01-27 main.tex
@@
 
-Temporal matching checks how spatially matched pairs align temporally.
-We use the following chart to illustrate how the temporal matching
-works:
+Data fusion is a process that consider combining multiple multiple data
+sources and it has been applied widely in the remote sensing community.
 
-For each spatially matched pair, say \texttt{A} and \texttt{a}, we first
-find the largest \texttt{n} points in each series, colored in brown
-points here. Here we use the largest three but you can tune this number
-by \texttt{temporal\_n\_highest}. Then we construct the interval of the
-largest points from one series and see how many points, from the other
-series, fall into the intervals. The series used to construct the
-interval is controlled by \texttt{temporal\_independent} and the window
-size by \texttt{temporal\_window} with a default of 5.
+Meteorological and hydrological data have been matched in the applied
+studies in the literature to understand drought, \ldots{} {[}ref{]}. To
+be a valid pair of matching, the matched pair from different data
+sources need to
 
-In this illustration, we construct the interval based on series
-\texttt{A} and two of the three peaks from \texttt{a} falls into this
-interval at Time 7 and 27.
+\begin{itemize}
+\tightlist
+\item
+  be spatially close to each other, and
+\item
+  have similar temporal movement
+\end{itemize}
 
-There's another mandatory argument that hasn't been introduced above:
-\texttt{temporal\_var\_to\_match}. This argument controls the variable
-to match and it needs to appear in both the \texttt{major} and
-\texttt{minor} set. In the water level matching example, we match the
-variable \texttt{Water\_course\_level} from \texttt{river} to
-\texttt{prcp} from \texttt{climate}, hence need to manually rename one
-of them to match the other, here we rename \texttt{Water\_course\_level}
-to \texttt{prcp} in \texttt{river}:
-
-Figure \ref{fig:illu-matching}
+\texttt{match\_sites()} from \texttt{cubble} matches data based on these
+two criteria. It first matches the two data sources spatially through
+computing the pairwise distance based on latitude and longitude. Pairs
+that pass the spatial matching are then matched temporally through
+computing the number of matched peaks within a fixed length window.
+Figure \ref{fig:illu-matching} illustrate this temporal matching in more
+details. Given two series \texttt{A} and \texttt{a}, 3 peaks have been
+picked in each series. An interval, with default length of 5, is
+constructed for each peak in series \texttt{A} and the peaks in series
+\texttt{a} are tested against whether they fall into the any of the
+intervals. In this illustration, there are 2 matches for these two
+series. Several arguments are available in \texttt{match\_sites()} to
+fine-tune the matching:
 
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.3\textheight]{figures/illu-matching}
+{\centering \includegraphics[width=1\linewidth]{figures/illu-matching}
 
@@
 
-\caption[sdfasdf]{sdfasdf}\label{fig:illu-matching}
+\caption[An illustration of temporal matching in cubble]{An illustration of temporal matching in cubble.
+Three highest peaks are identified in each series and intervals are constructed on series A.
+Two peaks in series a fall into the intervals and hence the two series are considered to have 2 matches.}\label{fig:illu-matching}
 \end{figure}
 \end{CodeChunk}
+
+\begin{itemize}
+\tightlist
+\item
+  \texttt{spatial\_n\_keep}: the number of spatial match for each site
+  to keep
+\item
+  \texttt{spatial\_dist\_max}: the maximum distance allowed for a
+  matched pair
+\item
+  \texttt{temporal\_n\_highest}: the number of peak used - 3 in the
+  example above
+\item
+  \texttt{temporal\_window}: the length of the interval - 5 in the
+  example above
+\item
+  \texttt{temporal\_min\_match}: the minimum number of matched peak for
+  a valid matched pair
+\end{itemize}
 

### 2022-01-29 scripts/illu-matching.R

@@ -16,8 +16,9 @@ circle <- tibble(
   x = c(6, 14, 23, 7, 21, 27),
   y = c(10, 5, 7, 7, 4, 6),
   id = factor(c(rep("A", 3), rep("a", 3)), levels = c("A", "a")),
-  xend = x + 5
-)
+  xend = x + 5,
+  match = factor(c("yes", "no", rep("yes", 2), "no", "yes"), levels = c("yes", "no"))
+) 
 
 errorbar <- bind_rows(
   circle %>% filter(id == "A"),
@@ -25,24 +26,19 @@ errorbar <- bind_rows(
 ) %>%
   mutate(id = factor(id, c("A", "a")))
 
-fallin <- tibble(
-  id = factor(rep("a", 2), c("A", "a")),
-  x = c(7, 27),
-  y = -0.5,
-  label = "\U2714"
-)
-
 ggplot() +
-  geom_line(data = dt, aes(x = date, y = val, group = id), color = "grey60") +
-  geom_vline(data = circle, aes(xintercept = x), linetype = "longdash", color = "grey60") +
-  geom_errorbar(data = errorbar, aes(xmin = x, xmax = xend, y = -0.5), width = 0.5) +
-  geom_point(data = circle, aes(x = x, y = y), color = "brown") +
-  geom_text(data = fallin, aes(x = x, y = y, label = label), color = "brown", size = 5) +
+  geom_rect(data = errorbar, aes(xmin = x, xmax = xend, ymin = 0, ymax = 10), fill = "grey90") +
+  geom_line(data = dt, aes(x = date, y = val, group = id), color = "black") +
+  geom_vline(data = circle %>% filter(id == "A"), 
+             aes(xintercept = x), linetype = "longdash", color = "grey10", lwd = 0.2) +
+  geom_point(data = circle, aes(x = x, y = y, color = match), size = 3) +
   facet_wrap(vars(id), nrow = 2) +
   scale_y_continuous(breaks = seq(0, 10, 2)) +
   scale_x_continuous(breaks = seq(1, 31, 1)) +
+  scale_color_brewer(palette = "Dark2") + 
   theme_bw() +
-  theme(panel.grid = element_blank()) +
+  theme(panel.grid = element_blank(),
+        legend.position = "bottom") +
   labs(x = "Time", y = "Value")
 
 ggsave(filename = "figures/illu-matching.png", width = 10, height = 5)

### 2022-01-30 4-2-agg.Rmd
improve aggregation plot
aggregation example done

--- 2022-01-07 4-2-agg.Rmd
@@
 
-In Figure \ref{fig:basic-agg}.
-Aggregating them into averages or other statistics would give a better picture.
-This can be done with cubble through a hierarchical structure, where the cluster indicator is one level above the individual station.
+There are `r nrow(weatherdata::climate_full)` stations that record daily precipitation in Australia and these series will overlay on top of each other in a glyph map.
+Aggregation can be helpful to group series into clusters before visualising the cluster with glyph map.
+This example shows how to organise data at both level with the cubble function `switch_key()`.
 
-The `weatherdata::climate_full` data has daily climate data of `r nrow(weatherdata::climate_full)` Australia stations from 2016 to 2020.
+A simple kmean algorithm based on the distance matrix is used here to create 20 clusters.
+This creates `station_nested` as a station level nested cubble with a cluster column indicating which group each station belongs to.
+More complex algorithms can be used for other problem, as long as there is a mapping from each station to a cluster.
 
-calcluate distance matrix
-As an example to illustrate here, a kmean clustering algorithm based on the distance matrix is used and the number of centres is set to 20.
-More complex algorithms can also be used for more complex problem, as long as a mapping from each station id to the cluster id can be constructed.
+```{r}
+load(here::here("data/station_nested.rda"))
+```
 
-<!-- Now we can obtain the aggregated series as described in the workflow diagram above and construct the glyph map with `GGally::glyphs()`: -->
+```{r echo = TRUE}
+station_nested
+```
 
-<!-- We can also look at the precipitation of each individual station within the same cluster: -->
+To create a group level cubble, use `switch_key()` with the new key variable, `cluster`:
 
-<!-- Lastly, there is one series on Tasmania island standing out from others, lets look at where it is: -->
+```{r echo = TRUE}
+cluster_nested <- station_nested %>% switch_key(cluster) 
+```
 
-```{r basic-agg, out.width="100%", out.height="90%", fig.cap="sdfasdf"}
+With the group level cubble, `get_centroid()` is useful to compute the centroid of each cluster, which will be used as the major axis for the glyph map later:
+
+```{r eval = FALSE, echo = TRUE}
+cluster_nested <- cluster_nested %>% get_centroid()
+```
+
+Long form cubble at both levels can be access through stretching the nested form and operations applicable to long form cubble are still available.
+With access to both station and cluster level cubbles, various plots can be made to understand the cluster.
+Figure \ref{fig:basic-agg} shows two example plots that can be made with this data: subplot A makes a glyph map with cluster level cubble and subplot B inspects the station membership of each cluster using the nested station level cubble.
+
+```{r basic-agg, out.width="100%", fig.cap="Profile of aggregated precipitation from 639 weather stations in Australia.
+Subplot A shows the glyph map of the weekly averaged precipitation of each cluster.
+The group number of printed in the middle of y minor axis and can be used as a reference line to read the magnitude of precipitation.
+Subplot B shows the station membership of each cluster"}
 knitr::include_graphics("figures/basic-agg.png")

### 2022-01-30 main.tex
aggregation example done

@@
 
-With recent climate, there are 639 stations in Australia that records
-daily maximum and minimum temperature and precipitation. While it would
-be too much series to show on a glyph map, stations close to each other
-should have similar precipitation pattern. Aggregation can be helpful
-here to cluster series into groups and visualise them with glyph map.
+There are 639 stations that record daily precipitation in Australia and
+these series will overlay on top of each other in a glyph map.
+Aggregation can be helpful to group series into clusters before
+visualising the cluster with glyph map. This example shows how to
+organise data at both level with the cubble function
+\texttt{switch\_key()}.
 
-A simple kmean algorithm is used based on the distance matrix to cluster
-the stations into 20 groups. This creates \texttt{station\_nested} as a
-station level nested cubble with a cluster column indicating which group
-each station belongs to. More complex algorithms can be used for more
-complex problem, as long as a mapping from each station id to the
-cluster id can be constructed.
+A simple kmean algorithm based on the distance matrix is used here to
+create 20 clusters. This creates \texttt{station\_nested} as a station
+level nested cubble with a cluster column indicating which group each
+station belongs to. More complex algorithms can be used for other
+problem, as long as there is a mapping from each station to a cluster.
 
 \begin{CodeChunk}
+\begin{CodeInput}
+R> station_nested
+\end{CodeInput}
 \begin{CodeOutput}
@@
 
-With the group level cubble, it is useful to compute the centroid of
-each cluster as the location of group level glyph map.
-\texttt{get\_centroid()} is a shortcut for doing so or you can manually
-find the convex hull, its centroid, and extract the longitude and
-latitude:
+With the group level cubble, \texttt{get\_centroid()} is useful to
+compute the centroid of each cluster, which will be used as the major
+axis for the glyph map later:
 
@@
 \begin{CodeInput}
-R> cluster_nested <- cluster_nested %>% 
-+   get_centroid()
+R> cluster_nested <- cluster_nested %>% get_centroid()
 \end{CodeInput}
@@
 Long form cubble at both levels can be access through stretching the
-nested form. Here weekly averages of the precipitation is calculated:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> station_long <- station_nested %>% 
-+   stretch(ts) %>% 
-+   mutate(wk = lubridate::week(date)) %>% 
-+   group_by(wk) %>% 
-+   summarise(prcp = sum(prcp, na.rm = TRUE))
-R> 
-R> cluster_long <- cluster_nested %>% 
-+   stretch(ts) %>% 
-+   mutate(wk = lubridate::week(date)) %>% 
-+   group_by(wk) %>% 
-+   summarise(prcp = mean(prcp, na.rm = TRUE)) %>% 
-+   migrate(cent_long, cent_lat)
-\end{CodeInput}
-\end{CodeChunk}
-
-With the four cubbles: \texttt{station\_nested}, \texttt{station\_long},
-\texttt{cluster\_nested}, and \texttt{cluster\_long}, all the
-information at both levels are accessible. This allows us to make glyph
-map on the cluster level, looking at station composition of each
-cluster, and plot the individual series to check for consistency within
-groups. Figure \ref{fig:basic-agg} composes of the three plots together
-to show the precipitation in Australia.
+nested form and operations applicable to long form cubble are still
+available. With access to both station and cluster level cubbles,
+various plots can be made to understand the cluster. Figure
+\ref{fig:basic-agg} shows two example plots that can be made with this
+data: subplot A makes a glyph map with cluster level cubble and subplot
+B inspects the station membership of each cluster using the nested
+station level cubble.
 
@@
 
-\caption[Aggregated precipitation in Australia]{Aggregated precipitation in Australia.}\label{fig:basic-agg}
+\caption[Profile of aggregated precipitation from 639 weather stations in Australia]{Profile of aggregated precipitation from 639 weather stations in Australia.
+Subplot A shows the glyph map of the weekly averaged precipitation of each cluster.
+The group number of printed in the middle of y minor axis and can be used as a reference line to read the magnitude of precipitation.
+Subplot B shows the station membership of each cluster}\label{fig:basic-agg}
 \end{figure}

### 2022-01-30 4-3-water-level.Rmd
matching example done :

--- 2022-01-07 4-3-water-level.Rmd
@@
 
-The water level data comes from [Bureau of Meteorology](http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full) and has a copy in `weatherdata`.
-Here we extract the water course level and add a column annotate this data of type `river`.
-For the rainfall data, we will still use the `weatherdata::climate_full`, filtering for Victorian stations in 2020 should be pretty familiar by now.
-Again, we first look at where these stations are on the map first:
+Bureau of Meteorology also collects [water data](http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full) from river gauges, which includes electrical conductivity, turbidity, water course discharge, water course level, and water temperature.
+In particular, water level will interactive with precipitation from the climate data since rainfall will raise the water level in the river.
+Figure \ref{fig:matching-map} gives the location of available weather station and water gauges in Victoria.
 
-Now we use `match_sites()` to first pair the weather stations with the river gauges spatially and then apply the temporal matching on `prcp`.
-We will construct the interval based on peaks in `climate` since we would expect a lag effect for precipitation to flow into the river and cause a raise in river level, hence `temporal_independent = climate`.
-We select the 30 highest peak from the series to construct the match by setting `temporal_n_highest = 30`.
-This is a tuning parameter and you can start with 10% of the points of one series (here we have daily data for a year, 10% is roughly 30 points).
-`temporal_min_match` filters out pairs don't have enough match and to return all the pairs, set `temporal_min_match` to `0`.``
+```{r eval = FALSE}
+climate <- weatherdata::climate_full %>%
+  filter(between(stringr::str_sub(id, 7, 8), 76, 90)) %>%
+  stretch() %>%
+  filter(lubridate::year(date) == 2020) %>%
+  tamp() %>%
+  mutate(type = "climate")
 
-The output from temporal matching is also a cubble, with additional column `.dist` and `.group` inherent from spatial matching and `n_match` for the number of matched temporal peaks.
-Then you can use this output to plot the location of match or to look at the series:
+river <- weatherdata::water %>%
+  stretch() %>%
+  select(date, Water_course_level) %>%
+  rename(prcp = Water_course_level) %>% 
+  tamp() %>%
+  mutate(type = "river")
+```
 
-```{r matching, out.width="100%", out.height="30%", fig.cap="asdfasd"}
+```{r matching-map, out.width="100%", fig.cap="Location of climate station and river gauges in Victoria, Australia."}
+knitr::include_graphics("figures/matching-map.png")
+```
+
+From the map, a few water gauges and weather stations are close to each other and there is the potential that fluctuation of the water level can reflect the precipitation measured by the climate station.
+As introduced in Section 3.2, `match_sites()` can be used to match one source of data with another source in a cubble.
+Here `prcp` in `climate` will be matched to `Water_course_level` in `river`, which is specified in the argument `temporal_var_to_match`.
+`temporal_independent` controls the dataset that will be used to construct the interval (see the illustration in section 3.2 for details on this).
+Here the goal is to see if precipitation will be reflected by the water level in the river and hence a lagged effect of water level on precipitation.
+This will put precipitation data, `climate`, as the independent.
+Given there is one year worth of data, the number of peak (`temporal_n_highest`) to consider is slightly raised from a default 20 to 30.
+`temporal_min_match` filters out pairs don't have enough match and to return all the pairs, set `temporal_min_match` to `0`
+
+```{r echo = TRUE}
+res <- match_sites(river, climate,
+                   temporal_var_to_match = prcp,
+                   temporal_independent = climate,  
+                   temporal_n_highest = 30,
+                   temporal_min_match = 15)
+```
+
+The output from matching is also a cubble, with additional column `dist` and `group` produced from spatial matching and `n_match` from the temporal matching.
+
+```{r}
+res %>% print(n = 8)
+```
+
+Figure \ref{fig:matching} plots the matched pairs on the map or to view the matched series and with the chosen parameter, there are four pairs of matches, which all locates in the middle Victoria and the concurrent increase of precipitation and water level can be observed.
+
+```{r matching, out.width="100%", fig.cap="Matched weather stations and river gauges on the map (A) and across time (B).
+Precipitation and water level have been standardised between 0 and 1 to be displayed on the same scale.
+The increases in the precipitation is reflected well by the water level."}
 knitr::include_graphics("figures/matching.png")

### 2022-01-30 main.tex
matching example done :

@@
 
-Water data is collected by the
-\href{http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full}{Bureau
-of Meteorology}. This includes \ldots{} water gauges, variables
-collected. In particular, water level will interactive with the climate
-variable precipitation as rains fall into the river will raise the water
-level. Figure \ref{fig:matching-map} shows the available weather station
-and water gauges in Victoria, Melbourne.
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> climate <- weatherdata::climate_full %>%
-+   filter(between(stringr::str_sub(id, 7, 8), 76, 90)) %>%
-+   stretch() %>%
-+   filter(lubridate::year(date) == 2020) %>%
-+   tamp() %>%
-+   mutate(type = "climate")
-R> 
-R> river <- weatherdata::water %>%
-+   stretch() %>%
-+   select(date, Water_course_level) %>%
-+   rename(prcp = Water_course_level) %>% 
-+   tamp() %>%
-+   mutate(type = "river")
-\end{CodeInput}
-\end{CodeChunk}
+Bureau of Meteorology also collects
+\href{http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full}{water
+data} from river gauges, which includes electrical conductivity,
+turbidity, water course discharge, water course level, and water
+temperature. In particular, water level will interactive with
+precipitation from the climate data since rainfall will raise the water
+level in the river. Figure \ref{fig:matching-map} gives the location of
+available weather station and water gauges in Victoria.
 
@@
 
-\caption[asdfasd]{asdfasd}\label{fig:matching-map}
+\caption[Location of climate station and river gauges in Victoria, Australia]{Location of climate station and river gauges in Victoria, Australia.}\label{fig:matching-map}
 \end{figure}
@@
 From the map, a few water gauges and weather stations are close to each
-other and we may be able to observe a spontaneous increase in the
-precipitation and water level from the measures, potential with a little
-delay in the river gauges. \texttt{match\_sites()} matches one source of
-data with another source in both spatial and temporal dimension in a
-cubble.
+other and there is the potential that fluctuation of the water level can
+reflect the precipitation measured by the climate station. As introduced
+in Section 3.2, \texttt{match\_sites()} can be used to match one source
+of data with another source in a cubble. Here \texttt{prcp} in
+\texttt{climate} will be matched to \texttt{Water\_course\_level} in
+\texttt{river}, which is specified in the argument
+\texttt{temporal\_var\_to\_match}. \texttt{temporal\_independent}
+controls the dataset that will be used to construct the interval (see
+the illustration in section 3.2 for details on this). Here the goal is
+to see if precipitation will be reflected by the water level in the
+river and hence a lagged effect of water level on precipitation. This
+will put precipitation data, \texttt{climate}, as the independent. Given
+there is one year worth of data, the number of peak
+(\texttt{temporal\_n\_highest}) to consider is slightly raised from a
+default 20 to 30. \texttt{temporal\_min\_match} filters out pairs don't
+have enough match and to return all the pairs, set
+\texttt{temporal\_min\_match} to \texttt{0}
 
@@
 
-The output from temporal matching is also a cubble, with additional
-column \texttt{dist} and \texttt{group} produced from spatial matching
-and \texttt{n\_match} from the number of peak matched temporally.
+The output from matching is also a cubble, with additional column
+\texttt{dist} and \texttt{group} produced from spatial matching and
+\texttt{n\_match} from the temporal matching.
 
@@
 5 230200      MARIBYRNONG RIVE~ -37.7  145. <tibble [~ river  6.17     6      19
-# ... with 3 more rows
+6 ASN00086038 essendon airport  -37.7  145. <tibble [~ clim~  6.17     6      19
+7 406213      CAMPASPE RIVER @~ -37.0  145. <tibble [~ river  1.84     1      18
+8 ASN00088051 redesdale         -37.0  145. <tibble [~ clim~  1.84     1      18
 \end{CodeOutput}
@@
 
-Then plots can be made to inspect the matched pairs on the map or to
-view the co-movement of the two series:
+Figure \ref{fig:matching} plots the matched pairs on the map or to view
+the matched series and with the chosen parameter, there are four pairs
+of matches, which all locates in the middle Victoria and the concurrent
+increase of precipitation and water level can be observed.
 
@@
 
-\caption[asdfasd]{asdfasd}\label{fig:matching}
+\caption[Matched weather stations and river gauges on the map (A) and across time (B)]{Matched weather stations and river gauges on the map (A) and across time (B).
+Precipitation and water level have been standardised between 0 and 1 to be displayed on the same scale.
+The increases in the precipitation is reflected well by the water level.}\label{fig:matching}
 \end{figure}
 \end{CodeChunk}
-
-With the chosen parameter, four pairs of stations stands out, which all
-locates in the middle Victoria and the concurrent increase of
-precipitation and water level can be observed.
 

### 2022-01-30 scripts/matching.R

@@ -22,7 +22,8 @@ plot_map(vic_map) +
              aes(x = long, y = lat, color = type)) + 
   scale_color_brewer(palette = "Dark2") +
   theme_bw() + 
-  theme(legend.position = "bottom")
+  theme(legend.position = "bottom") + 
+  labs(x = "Longitude", y = "Latitude")
 ggsave(filename = "figures/matching-map.png", width = 10, height = 5)  
 
 res <- match_sites(river, climate,
@@ -58,7 +59,9 @@ p2 <- res_long %>%
   scale_x_date(date_labels = "%b") + 
   labs(x = "Week", y = "Precipitation/ water level")
 
-(p1 | p2) + patchwork::plot_layout(guides = "collect") &
+(p1 | p2) + 
+  patchwork::plot_layout(guides = "collect") + 
+  plot_annotation(tag_levels = 'A')&
   theme(legend.position = "bottom")
 
-ggsave(filename = "figures/matching.png", width = 8, height = 4)  
+ggsave(filename = "figures/matching.png", width = 10, height = 5)  

### 2022-01-31 4-5-interactive.Rmd
update the interactive linking figure

@@
 
-```{r interactive-linking, echo = FALSE, out.width="100%", out.height="20%", fig.retina = 2, dpi = 300, fig.cap = "Exploring temperature variation using linkning of a map and seasonal display.
+```{r interactive-linking, echo = FALSE, out.width="100%", out.height="23%", fig.retina = 2, dpi = 300, fig.cap = "Exploring temperature variation using linking of a map and seasonal display.
 Each row is a screen dump of the process.
-The top row shows all locations and all temperature profiels.
+The top row shows all locations and all temperature profiles.
 Selecting a location with high variance on the map produces the plot in the second row.
-The maximum nad minimum temperature is shown using a ribbon.", fig.show='hold'}
+The maximum nad minimum temperature is shown using a ribbon.
+The bottom row first selects the lowest temperature in Auguest in the seasonal display.
+A location in the Tasmania Island is then selected and surprisingly, it is actually warmer than the Thredbo AWS.", fig.show='hold'}
 knitr::include_graphics(here::here("figures/linking.png"))
 knitr::include_graphics(here::here("figures/linking-north.png"))
-knitr::include_graphics(here::here("figures/linking-mid.png"))
-knitr::include_graphics(here::here("figures/linking-south.png"))
+knitr::include_graphics(here::here("figures/linking-lower.png"))
+#knitr::include_graphics(here::here("figures/linking-south.png"))
 ```

### 2022-01-31 main.tex
update the interactive linking figure

@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.2\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/linking} \includegraphics[width=1\linewidth,height=0.2\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/linking-north} \includegraphics[width=1\linewidth,height=0.2\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/linking-mid} \includegraphics[width=1\linewidth,height=0.2\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/linking-south}
+{\centering \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/linking} \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/linking-north} \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/linking-lower}
 
@@
 
-\caption[Exploring temperature variation using linkning of a map and seasonal display]{Exploring temperature variation using linkning of a map and seasonal display.
+\caption[Exploring temperature variation using linking of a map and seasonal display]{Exploring temperature variation using linking of a map and seasonal display.
 Each row is a screen dump of the process.
-The top row shows all locations and all temperature profiels.
+The top row shows all locations and all temperature profiles.
 Selecting a location with high variance on the map produces the plot in the second row.
-The maximum nad minimum temperature is shown using a ribbon.}\label{fig:interactive-linking}
+The maximum nad minimum temperature is shown using a ribbon.
+The bottom row first selects the lowest temperature in Auguest in the seasonal display.
+A location in the Tasmania Island is then selected and surprisingly, it is actually warmer than the Thredbo AWS.}\label{fig:interactive-linking}
 \end{figure}

### 2022-02-02 main.tex
netcdf example with era5 data

@@
 \begin{CodeInput}
-R> res <- match_sites(river, climate,
-+                    temporal_var_to_match = prcp,
-+                    temporal_independent = climate,  
+R> res <- match_sites(
++   river, climate,
++   temporal_var_to_match = c("Water_course_level" = "prcp"),
++   temporal_independent = "prcp",  
 +   temporal_n_highest = 30,
-+                    temporal_min_match = 15)
++   temporal_min_match = 15
++ )
 \end{CodeInput}
@@
 \newpage
+
+Figure \ref{fig:netcdf}
+
+\begin{CodeChunk}
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/netcdf}
+
+}
+
+\caption[sdf]{sdf}\label{fig:netcdf}
+\end{figure}
+\end{CodeChunk}
 
@@
 The bottom row first selects the lowest temperature in Auguest in the seasonal display.
-A location in the Tasmania Island is then selected and surprisingly, it is actually warmer than the Thredbo AWS.}\label{fig:interactive-linking}
+A location in the Tasmania Island is then selected and it is actually warmer than the Thredbo AWS.}\label{fig:interactive-linking}
 \end{figure}

### 2022-02-02 scripts/netcdf.R

@@ -1,27 +1,59 @@
 library(tidyverse)
 library(cubble)
+library(ecmwfr)
 
-nc <- ncdf4::nc_open(here::here("data/adaptor.mars.internal-1638344933.932894-4194-11-a5d35ac3-b36d-4982-afa5-a5dd193a6718.nc"))
-temp <- as_cubble(nc, vars = "d2m")
-vic_map <- ozmaps::abs_ste %>% filter(NAME == "Victoria")
+my_request <- list(
+  product_type = "reanalysis",
+  format = "netcdf",
+  pressure_level = "10",
+  year = "2002",
+  month = c("09", "10"),
+  day = c("04", "22", "26", "30"),
+  time = "00:00",
+  area = c(-5, -180, -90, 180),
+  variable = c("specific_humidity", 'geopotential', 'potential_vorticity', 'temperature'),
+  dataset_short_name = "reanalysis-era5-pressure-levels",
+  target = "era5-pressure.nc"
+)
 
-out <- temp %>% 
-  stretch(ts) %>% 
-  mutate(d2m = d2m - 273.15) %>% 
-  filter(lubridate::hour(time) == 12) %>% 
-  group_by(id) %>% 
-  summarise(d2m = mean(d2m, na.rm= TRUE))
+wf_request(
+  user = "113339",
+  request = my_request,
+  transfer = TRUE,
+  path = "~/Documents/research/paper-cubble/data/")
 
-res <- out %>%
-  migrate(long, lat)
+raw <- ncdf4::nc_open(file.choose())
+dt <- as_cubble(raw, vars = c("q", "z", "pv", "t"))
 
-vic_string <- vic_map %>% sf::st_cast("MULTILINESTRING") 
+date <- c("2002-09-22", "2002-09-26", "2002-09-30", "2002-10-04") %>% as.Date()
+res <- dt %>% 
+  stretch() %>% 
+  filter(lubridate::date(time) %in% date) %>%
+  migrate(long, lat) %>% 
+  mutate(q =  q* 10^6,
+         mq = z + pv / t)
+  
+box = c(xmin = -180, ymin = -90, xmax = 180, ymax = -5)
+sf::sf_use_s2(FALSE)
+country <- rnaturalearth::ne_coastline("small", returnclass = "sf") %>% 
+  sf::st_crop(box) %>% 
+  sf::st_cast("MULTILINESTRING")
 
 res %>% 
   ggplot() +
-  geom_tile(aes(x = long, y = lat, fill = d2m)) +
[…]

### 2022-02-02 4-3-water-level.Rmd
update matching syntax

@@
 
-```{r eval = FALSE}
+```{r}
 climate <- weatherdata::climate_full %>%
@@
 river <- weatherdata::water %>%
-  stretch() %>%
-  select(date, Water_course_level) %>%
-  rename(prcp = Water_course_level) %>% 
-  tamp() %>%
   mutate(type = "river")
@@
 ```{r echo = TRUE}
-res <- match_sites(river, climate,
-                   temporal_var_to_match = prcp,
-                   temporal_independent = climate,  
+res <- match_sites(
+  river, climate,
+  temporal_var_to_match = c("Water_course_level" = "prcp"),
+  temporal_independent = "prcp",  
   temporal_n_highest = 30,
-                   temporal_min_match = 15)
+  temporal_min_match = 15
+)
 ```

### 2022-02-02 4-4-netcdf.Rmd
netcdf example with era5 data
some content in the netcdf exmaple section

@@
 
-\newpage
+NetCDF (Network Common Data Form) is a data format commonly used in climatology community to deliver global mapping of atmosphere, ocean, and land.
+It has two major components: dimension to define the spatiotemporal grid (longitude, latitude, and time) and variable which populates the defined grid.
+Attributes are usually associated with dimension and variable in the NetCDF format data and a [metadata convention for climate and forecast](http://cfconventions.org/) has been designed to standardise this data format.
+Current packages on manipulating NetCDF data in R include a high-level R interface: `ncdf4`[@ncdf4], a low-level interface that calls C-interface: `RNetCDF`[@rnetcdf; @michna2013rnetcdf], and a tidyverse implementation: `tidync`[@tidync].
+
+Cubble provides an `as_cubble()` method to coerce the `ncdf4` class from the `ncdf4` package into a `cubble`. It maps each combination of latitude and longitude into a unit.
+
+```{r echo = TRUE, eval = FALSE}
+# read in the .nc file as a ncdf4 class
+raw <- ncdf4::nc_open(here::here("data/era5-pressure.nc"))
+
+# convert the variable q and z in the ncdf4 into a cubble
+dt <- as_cubble(raw, vars = c("q", "z"))
+```
+
+Since each combination of latitude and longitude grid is regarded as a site, the number of site quickly goes up when mapping on a global scale.
+
+  - For a southern hemisphere bounding box of [-5, -180, -90, 180], 491,040 id, 8 timestamp, 15MB -> 11 sec
+  - For an Australia bounding box of [-12, 115, -43, 153], 19125 id, 1,827 times tamp, 133 MB -> 1 mins .
+
+This shows a computation on the time dimension is cheaper than the spatial dimension.
+Argument `long_range` and `lat_range` are available to subset to a looser grid when reading in global level resolution.
+`era5-pressure.nc` has 0.25 degree resolution and to read in the data in whole degree, use:
+
+```{r eval = FALSE, echo = TRUE}
+dt <- as_cubble(raw, vars = c("q", "z"), 
+                long_range = seq(-180, 180, 1), 
+                lat_range = seq(-90, 90, 1))
+```
+
+Here an example in Figure \ref{fig:netcdf} is given to reproduce Figure 19 in @hersbach2020era5 from the ERA5 data.
+ERA5 data [@hersbach2020era5] is the latest reanalysis of global atmosphere, land surface, and ocean waves from 1950 onwards and is available in the NetCDF format.
+It can be directly downloaded from the [Copernicus Climate Data Store (CDS)](https://cds.climate.copernicus.eu/cdsapp#!/dataset/reanalysis-era5-pressure-levels?tab=overview) website or programmatically via R package `ecmwfr`[@ecwmfr].
+Variable Specific humidity and geopotential are queried on the 10 hPa pressure level for four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04.
+Readers who are interested in the details of  can refer to @simmons2020global and @simmons2005ecmwf.
+
+```{r netcdf, out.width="100%", fig.cap = "sdf"}
+knitr::include_graphics(here::here("figures/netcdf.png"))
+```

### 2022-02-02 main.tex
some content in the netcdf exmaple section

@@
 
-\newpage
+NetCDF (Network Common Data Form) is a data format commonly used in
+climatology community to deliver global mapping of atmosphere, ocean,
+and land. It has two major components: dimension to define the
+spatiotemporal grid (longitude, latitude, and time) and variable which
+populates the defined grid. Attributes are usually associated with
+dimension and variable in the NetCDF format data and a
+\href{http://cfconventions.org/}{metadata convention for climate and
+forecast} has been designed to standardise this data format. Current
+packages on manipulating NetCDF data in R include a high-level R
+interface: \texttt{ncdf4}\citep{ncdf4}, a low-level interface that calls
+C-interface: \texttt{RNetCDF}\citep{rnetcdf, michna2013rnetcdf}, and a
+tidyverse implementation: \texttt{tidync}\citep{tidync}.
 
-Figure \ref{fig:netcdf}
+Cubble provides an \texttt{as\_cubble()} method to coerce the
+\texttt{ncdf4} class from the \texttt{ncdf4} package into a
+\texttt{cubble}. It maps each combination of latitude and longitude into
+a unit.
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> # read in the .nc file as a ncdf4 class
+R> raw <- ncdf4::nc_open(here::here("data/era5-pressure.nc"))
+R> 
+R> # convert the variable q and z in the ncdf4 into a cubble
+R> dt <- as_cubble(raw, vars = c("q", "z"))
+\end{CodeInput}
+\end{CodeChunk}
+
+Since each combination of latitude and longitude grid is regarded as a
+site, the number of site quickly goes up when mapping on a global scale.
+
+\begin{itemize}
+\tightlist
+\item
+  For a southern hemisphere bounding box of {[}-5, -180, -90, 180{]},
+  491,040 id, 8 timestamp, 15MB -\textgreater{} 11 sec
+\item
+  For an Australia bounding box of {[}-12, 115, -43, 153{]}, 19125 id,
+  1,827 times tamp, 133 MB -\textgreater{} 1 mins .
+\end{itemize}
+
+This shows a computation on the time dimension is cheaper than the
+spatial dimension. Argument \texttt{long\_range} and \texttt{lat\_range}
+are available to subset to a looser grid when reading in global level
+resolution. \texttt{era5-pressure.nc} has 0.25 degree resolution and to
+read in the data in whole degree, use:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> dt <- as_cubble(raw, vars = c("q", "z"), 
++                 long_range = seq(-180, 180, 1), 
++                 lat_range = seq(-90, 90, 1))
+\end{CodeInput}
+\end{CodeChunk}
+
+Here an example in Figure \ref{fig:netcdf} is given to reproduce Figure
+19 in \citet{hersbach2020era5} from the ERA5 data. ERA5 data
+\citep{hersbach2020era5} is the latest reanalysis of global atmosphere,
+land surface, and ocean waves from 1950 onwards and is available in the
+NetCDF format. It can be directly downloaded from the
+\href{https://cds.climate.copernicus.eu/cdsapp\#!/dataset/reanalysis-era5-pressure-levels?tab=overview}{Copernicus
+Climate Data Store (CDS)} website or programmatically via R package
+\texttt{ecmwfr}\citep{ecwmfr}. Variable Specific humidity and
+geopotential are queried on the 10 hPa pressure level for four dates:
+2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04. Readers who are
+interested in the details of can refer to \citet{simmons2020global} and
+\citet{simmons2005ecmwf}.
 
@@
 The bottom row first selects the lowest temperature in Auguest in the seasonal display.
-A location in the Tasmania Island is then selected and it is actually warmer than the Thredbo AWS.}\label{fig:interactive-linking}
+A location in the Tasmania Island is then selected and surprisingly, it is actually warmer than the Thredbo AWS.}\label{fig:interactive-linking}
 \end{figure}

### 2022-02-03 4-4-netcdf.Rmd
netcdf4 section done

@@
 
-Cubble provides an `as_cubble()` method to coerce the `ncdf4` class from the `ncdf4` package into a `cubble`. It maps each combination of latitude and longitude into a unit.
+Cubble provides an `as_cubble()` method to coerce the `ncdf4` class from the `ncdf4` package into a `cubble`.
+It maps each combination of longitude and latitude into an `id` (the `key`) and nests time and variables in the nested form:
 
@@
 
-Since each combination of latitude and longitude grid is regarded as a site, the number of site quickly goes up when mapping on a global scale.
-
-  - For a southern hemisphere bounding box of [-5, -180, -90, 180], 491,040 id, 8 timestamp, 15MB -> 11 sec
-  - For an Australia bounding box of [-12, 115, -43, 153], 19125 id, 1,827 times tamp, 133 MB -> 1 mins .
-
-This shows a computation on the time dimension is cheaper than the spatial dimension.
-Argument `long_range` and `lat_range` are available to subset to a looser grid when reading in global level resolution.
-`era5-pressure.nc` has 0.25 degree resolution and to read in the data in whole degree, use:
+Memory limit with Netcdf data in cubble depends on longitude grid point x latitude grid point x time grid point x number of variable.
+Cubble can handle slightly more than 300 x 300 (longitude x longitude) grid points for 3 variables in one year and spatial grid can be reduced to trade for longer time period and more variables.
+A 300 by 300 spatial grid can be a bounding box of [100, -80, 180, 0] at 0.25 degree resolution or global bounding box [-180, -90, 180, -90] at 1 degree resolution.
+Subsetting longitude and latitude grid is available through `long_range` and `lat_range` if the Netcdf file has finer resolution than needed.
 
 ```{r eval = FALSE, echo = TRUE}
-dt <- as_cubble(raw, vars = c("q", "z"), 
+# Assume `my_ncdf` has bounding box of [-180, -90, 180, -90] 
+# at 0.25 degree resolution and subset it to have 
+# 1 degree resolution:
+dt <- as_cubble(my_ncdf, vars = c("q", "z"), 
                 long_range = seq(-180, 180, 1), 
@@
 
-Here an example in Figure \ref{fig:netcdf} is given to reproduce Figure 19 in @hersbach2020era5 from the ERA5 data.
+Figure \ref{fig:netcdf} gives an example of reproducing Figure 19 in @hersbach2020era5 using ERA5 data.
 ERA5 data [@hersbach2020era5] is the latest reanalysis of global atmosphere, land surface, and ocean waves from 1950 onwards and is available in the NetCDF format.
 It can be directly downloaded from the [Copernicus Climate Data Store (CDS)](https://cds.climate.copernicus.eu/cdsapp#!/dataset/reanalysis-era5-pressure-levels?tab=overview) website or programmatically via R package `ecmwfr`[@ecwmfr].
-Variable Specific humidity and geopotential are queried on the 10 hPa pressure level for four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04.
+Variable Specific humidity and geopotential are queried on the 10 hPa pressure level for four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04 adn once downloaded, the data can be read into a cubble using the code above.
 Readers who are interested in the details of can refer to @simmons2020global and @simmons2005ecmwf.
 
-```{r netcdf, out.width="100%", fig.cap = "sdf"}
+```{r netcdf, out.width="100%", fig.cap = "A reproduction of the second row of Figure 19 in  Hersbach (2020)."}
 knitr::include_graphics(here::here("figures/netcdf.png"))

### 2022-02-03 main.tex
netcdf4 section done

@@
 \texttt{ncdf4} class from the \texttt{ncdf4} package into a
-\texttt{cubble}. It maps each combination of latitude and longitude into
-a unit.
+\texttt{cubble}. It maps each combination of longitude and latitude into
+an \texttt{id} (the \texttt{key}) and nests time and variables in the
+nested form:
 
@@
 
-Since each combination of latitude and longitude grid is regarded as a
-site, the number of site quickly goes up when mapping on a global scale.
-
-\begin{itemize}
-\tightlist
-\item
-  For a southern hemisphere bounding box of {[}-5, -180, -90, 180{]},
-  491,040 id, 8 timestamp, 15MB -\textgreater{} 11 sec
-\item
-  For an Australia bounding box of {[}-12, 115, -43, 153{]}, 19125 id,
-  1,827 times tamp, 133 MB -\textgreater{} 1 mins .
-\end{itemize}
-
-This shows a computation on the time dimension is cheaper than the
-spatial dimension. Argument \texttt{long\_range} and \texttt{lat\_range}
-are available to subset to a looser grid when reading in global level
-resolution. \texttt{era5-pressure.nc} has 0.25 degree resolution and to
-read in the data in whole degree, use:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> dt <- as_cubble(raw, vars = c("q", "z"), 
+Memory limit with Netcdf data in cubble depends on longitude grid point
+x latitude grid point x time grid point x number of variable. Cubble can
+handle slightly more than 300 x 300 (longitude x longitude) grid points
+for 3 variables in one year and spatial grid can be reduced to trade for
+longer time period and more variables. A 300 by 300 spatial grid can be
+a bounding box of {[}100, -80, 180, 0{]} at 0.25 degree resolution or
+global bounding box {[}-180, -90, 180, -90{]} at 1 degree resolution.
+Subsetting longitude and latitude grid is available through
+\texttt{long\_range} and \texttt{lat\_range} if the Netcdf file has
+finer resolution than needed.
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> # Assume `my_ncdf` has bounding box of [-180, -90, 180, -90] 
+R> # at 0.25 degree resolution and subset it to have 
+R> # 1 degree resolution:
+R> dt <- as_cubble(my_ncdf, vars = c("q", "z"), 
 +                 long_range = seq(-180, 180, 1), 
@@
 
-Here an example in Figure \ref{fig:netcdf} is given to reproduce Figure
-19 in \citet{hersbach2020era5} from the ERA5 data. ERA5 data
+Figure \ref{fig:netcdf} gives an example of reproducing Figure 19 in
+\citet{hersbach2020era5} using ERA5 data. ERA5 data
 \citep{hersbach2020era5} is the latest reanalysis of global atmosphere,
@@
 geopotential are queried on the 10 hPa pressure level for four dates:
-2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04. Readers who are
+2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04 adn once downloaded,
+the data can be read into a cubble using the code above. Readers who are
 interested in the details of can refer to \citet{simmons2020global} and
@@
 
-\caption[sdf]{sdf}\label{fig:netcdf}
+\caption[A reproduction of the second row of Figure 19 in  Hersbach (2020)]{A reproduction of the second row of Figure 19 in  Hersbach (2020).}\label{fig:netcdf}
 \end{figure}

### 2022-02-03 scripts/netcdf.R

@@ -11,7 +11,7 @@ my_request <- list(
   day = c("04", "22", "26", "30"),
   time = "00:00",
   area = c(-5, -180, -90, 180),
-  variable = c("specific_humidity", 'geopotential', 'potential_vorticity', 'temperature'),
+  variable = c("specific_humidity", 'geopotential'),
   dataset_short_name = "reanalysis-era5-pressure-levels",
   target = "era5-pressure.nc"
 )
@@ -22,16 +22,15 @@ wf_request(
   transfer = TRUE,
   path = "~/Documents/research/paper-cubble/data/")
 
-raw <- ncdf4::nc_open(file.choose())
-dt <- as_cubble(raw, vars = c("q", "z", "pv", "t"))
+raw <- ncdf4::nc_open(here::here("data/era5-pressure.nc"))
+dt <- as_cubble(raw, vars = c("q", "z"))
 
 date <- c("2002-09-22", "2002-09-26", "2002-09-30", "2002-10-04") %>% as.Date()
 res <- dt %>% 
   stretch() %>% 
   filter(lubridate::date(time) %in% date) %>%
   migrate(long, lat) %>% 
-  mutate(q =  q* 10^6,
-         mq = z + pv / t)
+  mutate(q =  q* 10^6)
   
 box = c(xmin = -180, ymin = -90, xmax = 180, ymax = -5)
 sf::sf_use_s2(FALSE)

### 2022-02-03 3-advanced.Rmd
data fusion + matching

@@
 
-Data fusion is a process that consider combining multiple multiple data sources and it has been applied widely in the remote sensing community.
-
-Meteorological and hydrological data have been matched in the applied studies in the literature to understand drought, ...
-[ref].
 To be a valid pair of matching, the matched pair from different data sources need to
@@
  - `temporal_min_match`: the minimum number of matched peak for a valid matched pair
+
+This functionality can be seen as a matching [@stuart2010matching; @mcintosh2018using] exercise in the spatio-temporal domain to pair up nearby observatoins.
+It can also be considered as a form of data fusion [@castanedo2013review; @cocchi2019data], where data from multiple sources are combined together.
 

### 2022-02-03 main.tex
data fusion + matching

@@
 
-Data fusion is a process that consider combining multiple multiple data
-sources and it has been applied widely in the remote sensing community.
-
-Meteorological and hydrological data have been matched in the applied
-studies in the literature to understand drought, \ldots{} {[}ref{]}. To
-be a valid pair of matching, the matched pair from different data
+To be a valid pair of matching, the matched pair from different data
 sources need to
@@
 \end{itemize}
+
+This functionality can be seen as a matching
+\citep{stuart2010matching, mcintosh2018using} exercise in the
+spatio-temporal domain to pair up nearby observatoins. It can also be
+considered as a form of data fusion
+\citep{castanedo2013review, cocchi2019data}, where data from multiple
+sources are combined together.
 

### 2022-02-04 1-intro.Rmd
section division

@@
 
-The rest of the paper will be divided as follows: [complete when the paper structure is more solid]
-
-<!-- Section 2 reviews the existing data structure for spatio, temporal, and spatio-temporal data.
-Section 3 presents a new data structure for spatio-temporal data: cubble.
-Then the paper introduces the workflow of data manipulation and visualisation with the cubble structure in Section 4.
-Section 5 gives some examples on how common spatial and temporal manipulations are performed with cubble and how static and interactive visualisation help to understand climate and [...] data.
-\newline -->
+The rest of the paper will be divided as follows: Section 2 presents the workflow of data manipulation in cubble.
+Section 3 explains how cubble deals with some advanced considerations  including data with hierarchical structure, data matching, and how cubble fits with existing static and interactive visualisation tools.
+Section 4 gives examples of the features introduced in the previous two sections with climate and hydrology data.
+An example on how cubble handle Netcdf data is also provided.
+Section 5 concludes the paper

### 2022-02-04 2-the-cubble-pkg.Rmd
make notes and small changes till section 4.2

@@
 Spatio-temporal data usually come in various forms and Figure \ref{fig:illu-input} shows four examples of this.
-No matter which form the data is in, these formats share some common components that characterise spatio-temporal data.
+No matter which form the data is in, these formats share some common components that characterise the spatio-temporal data.
 A spatial identifier (`id` in the diagram) is the unique identifier of each site.
@@
 Other variables in the data can be categorised into two groups: spatial variables that are invariant at each time stamp for every site, i.e.
-the name or code of the weather station and temporal variables that varies with time.
+the name or code of the weather station, and temporal variables that varies with time.
 
 ```{r illu-input, echo = FALSE, out.width = "100%", fig.cap="Illustration of incoming data formats for spatio-temporal data.
-(1) Data comes in as a single table; (2) Separate tables for spatial and temporal variables; (3) A single table with all the parameters used to query the database and a separate table for queried data; and (4) Cubical data in array or NetCDF format.", fig.align='center', out.height = "15%"}
+(1) Data comes in as a single table; (2) Separate tables for spatial and temporal variables; (3) A single table with all the parameters used to query the database and a separate table for queried data; and (4) Cubical data in array or NetCDF format.", fig.align='center', out.height = "40%"}
 knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.001.png"))
@@
 The decision on which form to use is output-oriented, meaning analysts need to first think about whether the output of a particular operation is identified only by the spatial identifier, or a combination of spatial and temporal identifier.
-The nested cubble is suitable for working with operations that are only identified by site and this type of operation can be a pure manipulation of spatial variables, or a summary of temporal variables by site (i.e.
+The nested cubble is suitable for working with operations that are only identified by the site and this type of operation can be a pure manipulation of spatial variables, or a summary of temporal variables by site (i.e.
 the output of counting the number of raining day is only identified by sites and hence should be performed with the nested form).
@@
 
-To use functionalities from cubble, data analysts first need to create a cubble.
-`as_cubble` create a \pkg{cubble} by supplying the three key components: `key` as the spatial identifier; `index` as the temporal identifier; and a vector of `coords` in the order of longitude first and then latitude.
+\textcolor{red}{To use functionalities from cubble, data analysts first need to create a cubble.} `as_cubble()` creates a \pkg{cubble} by supplying the three key components: `key` as the spatial identifier; `index` as the temporal identifier; and a vector of `coords` in the order of longitude first and then latitude.
 The naming of `key` and `index` follows the convention in the \pkg{tsibble} package.
@@
 
-```{r}
-(cubble_nested <- cubble::climate_flat %>%
-  as_cubble(key = id, index = date, coords = c("long", "lat")))
+```{r echo = TRUE}
+cubble_nested <- cubble::climate_flat %>%
+  as_cubble(key = id, index = date, coords = c("long", "lat"))
+cubble_nested
 ```
 
-There are a few information in the \pkg{cubble} header: the name of the `key` variable, bbox, and also the name of variable nested in the `ts` column.
-In this example, each site is identifier is `id` and the number in the bracket means there are 5 unique `id` in this dataset.
-The bbox in the second row gives the range of the coordinates.
-The temporal variables are all nested in the `ts` column, but it could be useful to know the name these variables.
-The third row in the cubble header shows these names and in this example this includes: precipitation, `prcp`, maximum temperature, `tmax`, and minimum temperature, `tmin`.
+There are a few information in the \pkg{cubble} header: the name of the `key` variable (`id`) and its number (5), bounding box (`[115.97, -32.94, 133.55, -12.42]`), and also the name of variable nested in the `ts` column with its type (`date [date], prcp [dbl], tmax [dbl], tmin [dbl]`).
 
@@
 
-```{r}
-(cubble_long <- cubble_nested %>% stretch(ts))
+```{r echo = TRUE}
+cubble_long <- cubble_nested %>% stretch(ts)
+cubble_long
 ```
 
-Notice here that the third line in the header now shows the name of spatial variables rather than the temporal variables.
+Notice here that the third line in the header now shows the name and type of spatial variables (`lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]`).
 
@@
 Many times, analysts will need to go back and forth between the nested and long cubble.
-The `stretch()` function introduced in the previous section switches a nested cubble into a long cubble and function `tamp()` is its inverse function to switch a long cubble back to the nested cubble:
+`tamp()` inverses the `stretch()` that introduced in the previous section, to switch from a long cubble to the nested cubble:
 
-```{r}
-(cubble_back <- cubble_long %>% tamp())
+```{r echo = TRUE}
+cubble_back <- cubble_long %>% tamp()
+cubble_back
 ```
@@
 
-As a final data output for modelling or visualisation, spatio-temporal data is usually expected to be in a single table.
-Function `migrate()` moves the spatial variables from the `spatial` attribute into the long cubble:
+\textcolor{red}{As a final data output for modelling or visualisation, spatio-temporal data is usually expected to be in a single table.} Function `migrate()` moves the spatial variables from the `spatial` attribute into the long cubble:
 
 ```{r}
-(cubble_long %>% migrate(long, lat))
+cubble_long %>% migrate(long, lat)
 ```
 
-In this workflow described above, data objects come into cubble in the nested form, then various operations on the spatial and temporal dimension can go back and forth between the nested and long form, and finally, the data will come out of cubble in the long form for further modelling or visualisation.
+<!-- In the workflow described above, data objects come into cubble in the nested form, then various operations on the spatial and temporal dimension can go back and forth between the nested and long form, and finally, the data will come out of cubble in the long form for further modelling or visualisation.
+-->
 
@@
 ```
-\pkg{cubble} is also compatible with \pkg{tsibble} in the sense that the original list-column can be a `tbl_ts` object.
-Duplicates and gaps should be first checked before structuring the data into a cubble.
-If the input data is a \pkg{tsibble} object, the long form cubble is also a \pkg{tsibble} where users can directly apply time series operations.
+
+\pkg{cubble} is also compatible with \pkg{tsibble} where each element list-column in the nested form can be a `tbl_ts` object.
+\textcolor{red}{Duplicates and gaps should be first checked before structuring the data into a cubble.} If the input data is a \pkg{tsibble} object, the long form cubble is also a \pkg{tsibble} where users can directly apply time series operations.
+\textcolor{red}{more experiment with tsibble}

### 2022-02-04 3-advanced.Rmd
make notes and small changes till section 4.2

@@
 
-Spatial locations can have further grouping structure either in nature (i.e.
-countries are nested in continents), or calculated by some clustering algorithms.
-Rather than investigate the read from individual locations, group character summarised from multiple locations can give a clearer picture of the local dynamic.
+Spatial locations can have grouping structure that is either inherent to the data  (i.e.
+countries are nested in continents), or formed by clustering.
+Rather than analysing variables in the site level, summarised variables in the cluster level can give a clearer picture of \textcolor{red}{local dynamic}.
 In cubble, `switch_key()` can be used to create a new level of grouping of spatial location by specifying a clustering variable.
 Figure \ref{fig:illu-hier} illustrates the relationship of cubbles at site and cluster level, in both the long and nested form.
+\textcolor{red}{unify the use of site/ station}.
 By specifying `cluster_nested <- station_nested %>% switch_key(key = cluster)`, the cubble re-defines the cubble key from `id` in `station_nested` to `cluster` in `cluster_nested`.
@@
 
-```{r illu-hier, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "Relationship between the station and cluster level cubble in both long and nested form."}
+```{r illu-hier, echo = FALSE, fig.align="center",out.height="40%", out.width = "100%", fig.cap = "Relationship between the station and cluster level cubble in both long and nested form."}
 knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.003.png"))
@@
 
-To be a valid pair of matching, the matched pair from different data sources need to
+\textcolor{red}{refine the beginning.} To be a valid pair of matching, the matched pair from different data sources need to
 
@@
 
-The cubble structure fits in naturally with the interactive graphic pipeline discussed in the literature [@buja1988elements; @buja1996interactive;  @sutherland2000orca; @xie2014reactive; @cheng2016enabling].
-Diagram \ref{fig:illu-interactive} illustrates how linking works with the two forms in a cubble, where a time series plot is created with the long cubble and a map is created with the nested cubble.
-When a user action is captured from the map, the site will be activated in the nested cubble.
-Then, the nested cubble will communicate to the long cubble to activate all the observations with the same `id`.
-The long cubble will then highlight the  activated series in the time series plot.
+Cubble fits in naturally with the interactive graphic pipeline discussed in the literature [@buja1988elements; @buja1996interactive;  @sutherland2000orca; @xie2014reactive; @cheng2016enabling].
+Diagram \ref{fig:illu-interactive} illustrates how linking works from the map to the time series with the two forms of a cubble.
+The map and time series plot is associated with the nested or long cubble respectively and when a user action is captured on the map, the site will be activated in the nested cubble (left).
+The nested cubble will communicate to the long cubble to activate all the observations with the same `id` (middle).
+The long cubble will then highlight the activated series in the time series plot (right).
 
 The linking is also available from the time series plot to the map.
-The selection on the time series is through selecting the point on the time series and once a point is selected, it will be activated in the long cubble.
-All the observations that share the same `id`,  either in the long and nested cubble, are then activated.
-This includes other points in the same time series in the long cubble and the corresponding observation of site in the nested cubble.
+The selection(s) on the time series is through selecting the point(s) on the time series and once a point is selected, it will be activated in the long cubble.
+All the observations that share the same `id` are then activated and this includes other points in the same time series in the long cubble and the corresponding observation of site in the nested cubble.
 These activated observations will then being reflected in the updated plots and Diagram \ref{fig:illu-interactive-2} in the Appendix illustrates this process.
@@
 
-Glyph map [@Wickham2012-yr] plots the time series as single glyph on the map. 
-<!-- Variation of glyph maps has been developed to visualise uncertainty in **Visumap** [@Lucchesi2021] and deal with irregular local grid [@Beecham2021-vk].  -->
-In \proglang{R}, `GGally` implements the glyph map through the `glyphs()` function, which outputs a data frame with calculated position (`gx`, `gy`, `gid`) of each point on the time series given the major and minor xy variable using linear algebra (Equation 1 and 2 in @Wickham2012-yr).
-The data can then be piped into `ggplot` to create the glyph map:
+Glyph map [@Wickham2012-yr] plots the time series as single glyph on the map and in \proglang{R}, `GGally` implements the glyph map through the `glyphs()` function.
+The function construct a data frame with calculated position (`gx`, `gy`, `gid`) of each point on the time series using linear algebra (Equation 1 and 2 in @Wickham2012-yr).
+The data can then be piped into `ggplot` to create the glyph map as:
 
@@
 
-While this calculation can be seen as part of the `setup_data()` in a `ggproto`.
-A re-implementation of the glyph map as `GeomGlyph` has been made in `cubble` to create the glyph map with `geom_glyph`:
+A re-implementation of the glyph map as a ggproto `GeomGlyph` has been made in the `cubble` package so that the glyph map can be created with `geom_glyph()`:
 
@@
 Polar glyph map can be specify as a parameter, `polar = TRUE`, in the `geom_glyph()`, along with `width` and `height` in either absolute or relative value.
-Global and local scale can also be controlled by the parameter `global_rescale`, which default to `TRUE`.
+Global and local scale can be controlled by the parameter `global_rescale`, which default to `TRUE` for global scaling.
 Reference box and line can be added with separate `geom_glyph_box()` and `geom_glyph_line()`.

### 2022-02-04 4-1-basic.Rmd
make notes and small changes till section 4.2

@@
 
-Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world and the dataset `weatherdata::historical_tmax` extracts the historical maximum temperature recorded for `r nrow(weatherdata::historical_tmax)` Australian stations.
+Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world and the dataset `weatherdata::historical_tmax` extracts the maximum temperature for `r nrow(weatherdata::historical_tmax)` Australian stations.
+All the stations start recording from year `r range(weatherdata::historical_tmax$first_year)[2]` with the longest back to year `r range(weatherdata::historical_tmax$first_year)[1]`.
 The data `historical_tmax` is already presented as a cubble, with `id` as the key, `date` as the index, and `c("longitude", "latitude")` as the coordinates.
-Other variables include `elevation`, `name`, `wmo_id`, `first_year`, and  `last_year` in the nested form and `tmax` in the long form.
 This example compares the maximum temperature in two periods: 1971 - 1975 and 2016 - 2020 for stations in Victoria and New South Wales.
 
-Stations in the two states can be subsetted on the station number: Australia GHCN station number starts with `ASN00` and followed by the [Bureau of Meteorology (BOM) station number](http://www.bom.gov.au/climate/cdo/about/site-num.shtml), where the 2nd and 3rd digit (7th and 8th in the GHCN number) denote the state a station belongs to.
-New South Wales stations start from 46 to 75 and Victoria stations then follow from 76 to 90.
-Extracting Victoria and New South Wales stations is a filter operation in the spatial dimension and hence is operated in the nested form:
+Stations in the two states can be subsetted on the station number: Australia GHCN station number starts with "ASN00" and followed by the [Bureau of Meteorology (BOM) station number](http://www.bom.gov.au/climate/cdo/about/site-num.shtml), where the 2nd and 3rd digit (7th and 8th in the GHCN number) define the state of the station.
+New South Wales stations start from 46 to 75 and Victoria stations follow from 76 to 90.
+Filtering Victoria and New South Wales stations is an operation in the spatial dimension and hence uses the nested form:
 
@@
 
-The five year window is chosen to remove the effect of a particular year and the historical period of 1971 - 1975 is used since all the stations have records from 1970.
-This following chunk filters on records in the two study periods and summarises the maximum temperature into monthly measure.
-`stretch()` is used to convert the nested form cubble into the long form for these operations:
+Filtering for the period 1971: 1975 and 2016:2020 is an operation on the time dimension and the nested cubble needs to be switched to the long cubble by `stretch()`:
 
@@
   stretch() %>%
-  filter(lubridate::year(date) %in% c(1971:1975, 2016:2020)) %>%
+  filter(lubridate::year(date) %in% c(1971:1975, 2016:2020)) 
+```
+
+A monthly average is used for both periods to smooth the maximum temperature and it is again an operation on the time dimension:
+
+```{r eval=  FALSE, echo = TRUE}
+tmax <- tmax %>%
   mutate(month = lubridate::month(date), 
@@
 
-A data quality issue with GHCN data is that while the first and last year of each series is provided, years missing in this period is not reported.
-There are a few stations which don't have records during 1971 - 1975 and these stations are filtered out by examining whether the summarised `tmax` has a total of 24 months.
-This is again a station-wise operation and is operated in the nested form, which is switched to from the long form by  `tamp()`:
+There are a few stations which don't have records during 1971 - 1975 and further investigation shows that while the first and last year of each series is recorded, the missing years in this period is not reported.
+These stations are filtered out by examining whether the summarised time series has a total of 24 months.
+The long cubble needs to be switched to the nested form for this spatial operation using `tamp()`:
 
 ```{r eval = FALSE, echo = TRUE}
-tmax <- tmax %>% 
-  tamp() %>%
-  filter(nrow(ts) == 24) 
+tmax <- tmax %>% tamp() %>% filter(nrow(ts) == 24) 
 ```
@@
 ```{r eval = FALSE, echo = TRUE}  
-tmax <- tmax %>%   
-  stretch() %>%
-  migrate(latitude, longitude)
+tmax <- tmax %>% stretch() %>% migrate(latitude, longitude)
 ```
 
-Figure \ref{fig:basic-manip} shows the glyph map made with the data after the wangling above.
-One issue with this map is that the similar pattern shown from a few nearby stations around Sydney (151E, 34S) and New castle (152E, 33S) can be distracting.
-Aggregation is a useful technique to observe the general pattern of a collection of series and will be the topic of the next example.
+`tmax` can then be supplied to `geom_glyph()` for the glyph map in Figure \ref{fig:basic-manip} with a station inset on the top left corner:
 
-```{r basic-manip, out.width="100%", out.height="70%", fig.cap="Glyph map of the mean maximum temperature by month for Victoria and New South Wales weather stations.
-On the top left corner is a detailed legend for station Cobar highlighted in the black box.
-Compared to 1971 - 1975,  the period 2016 - 2020 sees an increase of mean maximum temperature in spring and summer in Australia (end and beginning of the year).
-A larger increase during the first three months of the year is observed at stations with a lower latitude."}
+```{r eval = FALSE, echo = TRUE}
+nsw_vic <- ozmaps::abs_ste %>% 
+  filter(NAME %in% c("Victoria", "New South Wales"))
+
+ggplot() + 
+  geom_sf(data = nsw_vic, 
+          fill = "transparent", color = "grey", linetype = "dotted") + 
+  geom_glyph(data = tmax, 
+             aes(x_major = longitude, x_minor = month, 
+                 y_major = latitude, y_minor = tmax,
+                 group = interaction(id, group), color = group),
+             width = 1, height = 0.5) +
+  ...
+```
+
+```{r basic-manip, out.width="100%", out.height="70%", fig.cap="A glyph map of the mean maximum temperature by month for Victoria and New South Wales weather stations in Australia.
+On the top left corner is an insetted plot of station Cobar, which is highlighted in the black box."}
 knitr::include_graphics("figures/basic-manip.png")

### 2022-02-04 4-2-agg.Rmd
make notes and small changes till section 4.2

@@
 
-There are `r nrow(weatherdata::climate_full)` stations that record daily precipitation in Australia and these series will overlay on top of each other in a glyph map.
+In the previous example, there has already been some overlapping of the glyphs for a few stations near (151E, 34S) and (152E, 33S) and this will be a problem when mapping more stations in the national level.
 Aggregation can be helpful to group series into clusters before visualising the cluster with glyph map.
-This example shows how to organise data at both level with the cubble function `switch_key()`.
 
+This example shows how to organise data at both level with `switch_key()`.
+The dataset used is `weatherdata::climate_full`, which records daily precipitation and maximum/ minimum temperature for `r nrow(weatherdata::climate_full)` stations in Australia from 2016 to 2020.
 A simple kmean algorithm based on the distance matrix is used here to create 20 clusters.
-This creates `station_nested` as a station level nested cubble with a cluster column indicating which group each station belongs to.
+This creates `station_nested` as a station level nested cubble with a cluster column indicating the group each station belongs to.
 More complex algorithms can be used for other problem, as long as there is a mapping from each station to a cluster.
@@
 
-```{r echo = TRUE}
-station_nested
+```{r echo = TRUE, eval = FALSE}
+station_nested <- weatherdata::climate_full %>% 
+  mutate(cluster = ...)
 ```

### 2022-02-04 main.Rmd
make notes and small changes till section 4.2

@@
 abstract: >
-  The abstract of the article.
+  Sptio-temporal data naturally has two levels of observations: site and time level.
+[problem].
+In this paper, a new data structure, `cubble`, is proposed for manipulating and visualising spatio-temporal data.
 keywords:
   # at least one keyword must be supplied
-  formatted: [spatio-temporal data,  "\\proglang{R}"]
-  plain:     [spatio-temporal data, R]
+  formatted: [spatio-temporal data,  "\\proglang{R}", climate data]
+  plain:     [spatio-temporal data, R, climate data]
 preamble: >
@@
   \usepackage{array}
+  \usepackage{xcolor}
 output: rticles::jss_article

### 2022-02-04 main.tex
make notes and small changes till section 4.2

@@
 \Abstract{
-The abstract of the article.
-}
-
-\Keywords{spatio-temporal data, \proglang{R}}
-\Plainkeywords{spatio-temporal data, R}
+Sptio-temporal data naturally has two levels of observations: site and
+time level. {[}problem{]}. In this paper, a new data structure,
+\texttt{cubble}, is proposed for manipulating and visualising
+spatio-temporal data.
+}
+
+\Keywords{spatio-temporal data, \proglang{R}, climate data}
+\Plainkeywords{spatio-temporal data, R, climate data}
 
@@
 
-\usepackage{amsmath} \usepackage{array}
+\usepackage{amsmath} \usepackage{array} \usepackage{xcolor}
 
@@
 
-The rest of the paper will be divided as follows: {[}complete when the
-paper structure is more solid{]}
+The rest of the paper will be divided as follows: Section 2 presents the
+workflow of data manipulation in cubble. Section 3 explains how cubble
+deals with some advanced considerations including data with hierarchical
+structure, data matching, and how cubble fits with existing static and
+interactive visualisation tools. Section 4 gives examples of the
+features introduced in the previous two sections with climate and
+hydrology data. An example on how cubble handle Netcdf data is also
+provided. Section 5 concludes the paper
 
@@
 the data is in, these formats share some common components that
-characterise spatio-temporal data. A spatial identifier (\texttt{id} in
-the diagram) is the unique identifier of each site. The temporal
+characterise the spatio-temporal data. A spatial identifier (\texttt{id}
+in the diagram) is the unique identifier of each site. The temporal
 identifier (\texttt{t} in the diagram) prescribes the time stamp each
@@
 each time stamp for every site, i.e.~the name or code of the weather
-station and temporal variables that varies with time.
-
-\begin{CodeChunk}
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth,height=0.15\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.001}
+station, and temporal variables that varies with time.
+
+\begin{CodeChunk}
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.001}
 
@@
 combination of spatial and temporal identifier. The nested cubble is
-suitable for working with operations that are only identified by site
-and this type of operation can be a pure manipulation of spatial
+suitable for working with operations that are only identified by the
+site and this type of operation can be a pure manipulation of spatial
 variables, or a summary of temporal variables by site (i.e.~the output
@@
 
-To use functionalities from cubble, data analysts first need to create a
-cubble. \texttt{as\_cubble} create a \pkg{cubble} by supplying the three
-key components: \texttt{key} as the spatial identifier; \texttt{index}
-as the temporal identifier; and a vector of \texttt{coords} in the order
-of longitude first and then latitude. The naming of \texttt{key} and
+\textcolor{red}{To use functionalities from cubble, data analysts first need to create a cubble.}
+\texttt{as\_cubble()} creates a \pkg{cubble} by supplying the three key
+components: \texttt{key} as the spatial identifier; \texttt{index} as
+the temporal identifier; and a vector of \texttt{coords} in the order of
+longitude first and then latitude. The naming of \texttt{key} and
 \texttt{index} follows the convention in the \pkg{tsibble} package. The
@@
 \begin{CodeChunk}
+\begin{CodeInput}
+R> cubble_nested <- cubble::climate_flat %>%
++   as_cubble(key = id, index = date, coords = c("long", "lat"))
+R> cubble_nested
+\end{CodeInput}
 \begin{CodeOutput}
@@
 There are a few information in the \pkg{cubble} header: the name of the
-\texttt{key} variable, bbox, and also the name of variable nested in the
-\texttt{ts} column. In this example, each site is identifier is
-\texttt{id} and the number in the bracket means there are 5 unique
-\texttt{id} in this dataset. The bbox in the second row gives the range
-of the coordinates. The temporal variables are all nested in the
-\texttt{ts} column, but it could be useful to know the name these
-variables. The third row in the cubble header shows these names and in
-this example this includes: precipitation, \texttt{prcp}, maximum
-temperature, \texttt{tmax}, and minimum temperature, \texttt{tmin}.
+\texttt{key} variable (\texttt{id}) and its number (5), bounding box
+(\texttt{{[}115.97,\ -32.94,\ 133.55,\ -12.42{]}}), and also the name of
+variable nested in the \texttt{ts} column with its type
+(\texttt{date\ {[}date{]},\ prcp\ {[}dbl{]},\ tmax\ {[}dbl{]},\ tmin\ {[}dbl{]}}).
 
@@
 \begin{CodeChunk}
+\begin{CodeInput}
+R> cubble_long <- cubble_nested %>% stretch(ts)
+R> cubble_long
+\end{CodeInput}
 \begin{CodeOutput}
@@
 
-Notice here that the third line in the header now shows the name of
-spatial variables rather than the temporal variables.
+Notice here that the third line in the header now shows the name and
+type of spatial variables
+(\texttt{lat\ {[}dbl{]},\ long\ {[}dbl{]},\ elev\ {[}dbl{]},\ name\ {[}chr{]},\ wmo\_id\ {[}dbl{]}}).
 
@@
 process. Many times, analysts will need to go back and forth between the
-nested and long cubble. The \texttt{stretch()} function introduced in
-the previous section switches a nested cubble into a long cubble and
-function \texttt{tamp()} is its inverse function to switch a long cubble
-back to the nested cubble:
-
-\begin{CodeChunk}
+nested and long cubble. \texttt{tamp()} inverses the \texttt{stretch()}
+that introduced in the previous section, to switch from a long cubble to
+the nested cubble:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> cubble_back <- cubble_long %>% tamp()
+R> cubble_back
+\end{CodeInput}
 \begin{CodeOutput}
@@
 
-As a final data output for modelling or visualisation, spatio-temporal
-data is usually expected to be in a single table. Function
-\texttt{migrate()} moves the spatial variables from the \texttt{spatial}
-attribute into the long cubble:
+\textcolor{red}{As a final data output for modelling or visualisation, spatio-temporal data is usually expected to be in a single table.}
+Function \texttt{migrate()} moves the spatial variables from the
+\texttt{spatial} attribute into the long cubble:
 
@@
 
-In this workflow described above, data objects come into cubble in the
-nested form, then various operations on the spatial and temporal
-dimension can go back and forth between the nested and long form, and
-finally, the data will come out of cubble in the long form for further
-modelling or visualisation.
-
 Building from an underlying \texttt{tbl\_df} structure, it is natural to
@@
 
-\pkg{cubble} is also compatible with \pkg{tsibble} in the sense that the
-original list-column can be a \texttt{tbl\_ts} object. Duplicates and
-gaps should be first checked before structuring the data into a cubble.
+\pkg{cubble} is also compatible with \pkg{tsibble} where each element
+list-column in the nested form can be a \texttt{tbl\_ts} object.
+\textcolor{red}{Duplicates and gaps should be first checked before structuring the data into a cubble.}
 If the input data is a \pkg{tsibble} object, the long form cubble is
 also a \pkg{tsibble} where users can directly apply time series
-operations.
+operations. \textcolor{red}{more experiment with tsibble}
 
@@
 
-Spatial locations can have further grouping structure either in nature
-(i.e.~countries are nested in continents), or calculated by some
-clustering algorithms. Rather than investigate the read from individual
-locations, group character summarised from multiple locations can give a
-clearer picture of the local dynamic. In cubble, \texttt{switch\_key()}
-can be used to create a new level of grouping of spatial location by
-specifying a clustering variable. Figure \ref{fig:illu-hier} illustrates
-the relationship of cubbles at site and cluster level, in both the long
-and nested form. By specifying
+Spatial locations can have grouping structure that is either inherent to
+the data (i.e.~countries are nested in continents), or formed by
+clustering. Rather than analysing variables in the site level,
+summarised variables in the cluster level can give a clearer picture of
+\textcolor{red}{local dynamic}. In cubble, \texttt{switch\_key()} can be
+used to create a new level of grouping of spatial location by specifying
+a clustering variable. Figure \ref{fig:illu-hier} illustrates the
+relationship of cubbles at site and cluster level, in both the long and
+nested form. \textcolor{red}{unify the use of site/ station}. By
+specifying
 \texttt{cluster\_nested\ \textless{}-\ station\_nested\ \%\textgreater{}\%\ switch\_key(key\ =\ cluster)},
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.003}
+{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.003}
 
@@
 
-To be a valid pair of matching, the matched pair from different data
-sources need to
+\textcolor{red}{refine the beginning.} To be a valid pair of matching,
+the matched pair from different data sources need to
 
@@
 
-The cubble structure fits in naturally with the interactive graphic
-pipeline discussed in the literature
+Cubble fits in naturally with the interactive graphic pipeline discussed
+in the literature
 \citep{buja1988elements, buja1996interactive, sutherland2000orca, xie2014reactive, cheng2016enabling}.
-Diagram \ref{fig:illu-interactive} illustrates how linking works with
-the two forms in a cubble, where a time series plot is created with the
-long cubble and a map is created with the nested cubble. When a user
-action is captured from the map, the site will be activated in the
-nested cubble. Then, the nested cubble will communicate to the long
-cubble to activate all the observations with the same \texttt{id}. The
-long cubble will then highlight the activated series in the time series
-plot.
+Diagram \ref{fig:illu-interactive} illustrates how linking works from
+the map to the time series with the two forms of a cubble. The map and
+time series plot is associated with the nested or long cubble
+respectively and when a user action is captured on the map, the site
+will be activated in the nested cubble (left). The nested cubble will
+communicate to the long cubble to activate all the observations with the
+same \texttt{id} (middle). The long cubble will then highlight the
+activated series in the time series plot (right).
 
 The linking is also available from the time series plot to the map. The
-selection on the time series is through selecting the point on the time
-series and once a point is selected, it will be activated in the long
-cubble. All the observations that share the same \texttt{id}, either in
-the long and nested cubble, are then activated. This includes other
-points in the same time series in the long cubble and the corresponding
-observation of site in the nested cubble. These activated observations
-will then being reflected in the updated plots and Diagram
-\ref{fig:illu-interactive-2} in the Appendix illustrates this process.
+selection(s) on the time series is through selecting the point(s) on the
+time series and once a point is selected, it will be activated in the
+long cubble. All the observations that share the same \texttt{id} are
+then activated and this includes other points in the same time series in
+the long cubble and the corresponding observation of site in the nested
+cubble. These activated observations will then being reflected in the
+updated plots and Diagram \ref{fig:illu-interactive-2} in the Appendix
+illustrates this process.
 
@@
 Glyph map \citep{Wickham2012-yr} plots the time series as single glyph
-on the map. In \proglang{R}, \texttt{GGally} implements the glyph map
-through the \texttt{glyphs()} function, which outputs a data frame with
-calculated position (\texttt{gx}, \texttt{gy}, \texttt{gid}) of each
-point on the time series given the major and minor xy variable using
-linear algebra (Equation 1 and 2 in \citet{Wickham2012-yr}). The data
-can then be piped into \texttt{ggplot} to create the glyph map:
+on the map and in \proglang{R}, \texttt{GGally} implements the glyph map
+through the \texttt{glyphs()} function. The function construct a data
+frame with calculated position (\texttt{gx}, \texttt{gy}, \texttt{gid})
+of each point on the time series using linear algebra (Equation 1 and 2
+in \citet{Wickham2012-yr}). The data can then be piped into
+\texttt{ggplot} to create the glyph map as:
 
@@
 
-While this calculation can be seen as part of the \texttt{setup\_data()}
-in a \texttt{ggproto}. A re-implementation of the glyph map as
-\texttt{GeomGlyph} has been made in \texttt{cubble} to create the glyph
-map with \texttt{geom\_glyph}:
+A re-implementation of the glyph map as a ggproto \texttt{GeomGlyph} has
+been made in the \texttt{cubble} package so that the glyph map can be
+created with \texttt{geom\_glyph()}:
 
@@
 \texttt{height} in either absolute or relative value. Global and local
-scale can also be controlled by the parameter \texttt{global\_rescale},
-which default to \texttt{TRUE}. Reference box and line can be added with
-separate \texttt{geom\_glyph\_box()} and \texttt{geom\_glyph\_line()}.
+scale can be controlled by the parameter \texttt{global\_rescale}, which
+default to \texttt{TRUE} for global scaling. Reference box and line can
+be added with separate \texttt{geom\_glyph\_box()} and
+\texttt{geom\_glyph\_line()}.
 
@@
 measures from stations across the world and the dataset
-\texttt{weatherdata::historical\_tmax} extracts the historical maximum
-temperature recorded for 236 Australian stations. The data
+\texttt{weatherdata::historical\_tmax} extracts the maximum temperature
+for 236 Australian stations. All the stations start recording from year
+1969 with the longest back to year 1859. The data
 \texttt{historical\_tmax} is already presented as a cubble, with
 \texttt{id} as the key, \texttt{date} as the index, and
-\texttt{c("longitude",\ "latitude")} as the coordinates. Other variables
-include \texttt{elevation}, \texttt{name}, \texttt{wmo\_id},
-\texttt{first\_year}, and \texttt{last\_year} in the nested form and
-\texttt{tmax} in the long form. This example compares the maximum
-temperature in two periods: 1971 - 1975 and 2016 - 2020 for stations in
-Victoria and New South Wales.
+\texttt{c("longitude",\ "latitude")} as the coordinates. This example
+compares the maximum temperature in two periods: 1971 - 1975 and 2016 -
+2020 for stations in Victoria and New South Wales.
 
 Stations in the two states can be subsetted on the station number:
-Australia GHCN station number starts with \texttt{ASN00} and followed by
-the \href{http://www.bom.gov.au/climate/cdo/about/site-num.shtml}{Bureau
-of Meteorology (BOM) station number}, where the 2nd and 3rd digit (7th
-and 8th in the GHCN number) denote the state a station belongs to. New
-South Wales stations start from 46 to 75 and Victoria stations then
-follow from 76 to 90. Extracting Victoria and New South Wales stations
-is a filter operation in the spatial dimension and hence is operated in
-the nested form:
+Australia GHCN station number starts with ``ASN00'' and followed by the
+\href{http://www.bom.gov.au/climate/cdo/about/site-num.shtml}{Bureau of
+Meteorology (BOM) station number}, where the 2nd and 3rd digit (7th and
+8th in the GHCN number) define the state of the station. New South Wales
+stations start from 46 to 75 and Victoria stations follow from 76 to 90.
+Filtering Victoria and New South Wales stations is an operation in the
+spatial dimension and hence uses the nested form:
 
@@
 
-The five year window is chosen to remove the effect of a particular year
-and the historical period of 1971 - 1975 is used since all the stations
-have records from 1970. This following chunk filters on records in the
-two study periods and summarises the maximum temperature into monthly
-measure. \texttt{stretch()} is used to convert the nested form cubble
-into the long form for these operations:
+Filtering for the period 1971: 1975 and 2016:2020 is an operation on the
+time dimension and the nested cubble needs to be switched to the long
+cubble by \texttt{stretch()}:
 
@@
 +   stretch() %>%
-+   filter(lubridate::year(date) %in% c(1971:1975, 2016:2020)) %>%
++   filter(lubridate::year(date) %in% c(1971:1975, 2016:2020)) 
+\end{CodeInput}
+\end{CodeChunk}
+
+A monthly average is used for both periods to smooth the maximum
+temperature and it is again an operation on the time dimension:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> tmax <- tmax %>%
 +   mutate(month = lubridate::month(date), 
@@
 
-A data quality issue with GHCN data is that while the first and last
-year of each series is provided, years missing in this period is not
-reported. There are a few stations which don't have records during 1971
-- 1975 and these stations are filtered out by examining whether the
-summarised \texttt{tmax} has a total of 24 months. This is again a
-station-wise operation and is operated in the nested form, which is
-switched to from the long form by \texttt{tamp()}:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> tmax <- tmax %>% 
-+   tamp() %>%
-+   filter(nrow(ts) == 24) 
+There are a few stations which don't have records during 1971 - 1975 and
+further investigation shows that while the first and last year of each
+series is recorded, the missing years in this period is not reported.
+These stations are filtered out by examining whether the summarised time
+series has a total of 24 months. The long cubble needs to be switched to
+the nested form for this spatial operation using \texttt{tamp()}:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> tmax <- tmax %>% tamp() %>% filter(nrow(ts) == 24) 
 \end{CodeInput}
@@
 \begin{CodeInput}
-R> tmax <- tmax %>%   
-+   stretch() %>%
-+   migrate(latitude, longitude)
-\end{CodeInput}
-\end{CodeChunk}
-
-Figure \ref{fig:basic-manip} shows the glyph map made with the data
-after the wangling above. One issue with this map is that the similar
-pattern shown from a few nearby stations around Sydney (151E, 34S) and
-New castle (152E, 33S) can be distracting. Aggregation is a useful
-technique to observe the general pattern of a collection of series and
-will be the topic of the next example.
+R> tmax <- tmax %>% stretch() %>% migrate(latitude, longitude)
+\end{CodeInput}
+\end{CodeChunk}
+
+\texttt{tmax} can then be supplied to \texttt{geom\_glyph()} for the
+glyph map in Figure \ref{fig:basic-manip} with a station inset on the
+top left corner:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> nsw_vic <- ozmaps::abs_ste %>% 
++   filter(NAME %in% c("Victoria", "New South Wales"))
+R> 
+R> ggplot() + 
++   geom_sf(data = nsw_vic, 
++           fill = "transparent", color = "grey", linetype = "dotted") + 
++   geom_glyph(data = tmax, 
++              aes(x_major = longitude, x_minor = month, 
++                  y_major = latitude, y_minor = tmax,
++                  group = interaction(id, group), color = group),
++              width = 1, height = 0.5) +
++   ...
+\end{CodeInput}
+\end{CodeChunk}
 
@@
 
-\caption[Glyph map of the mean maximum temperature by month for Victoria and New South Wales weather stations]{Glyph map of the mean maximum temperature by month for Victoria and New South Wales weather stations.
-On the top left corner is a detailed legend for station Cobar highlighted in the black box.
-Compared to 1971 - 1975,  the period 2016 - 2020 sees an increase of mean maximum temperature in spring and summer in Australia (end and beginning of the year).
-A larger increase during the first three months of the year is observed at stations with a lower latitude.}\label{fig:basic-manip}
+\caption[A glyph map of the mean maximum temperature by month for Victoria and New South Wales weather stations in Australia]{A glyph map of the mean maximum temperature by month for Victoria and New South Wales weather stations in Australia.
+On the top left corner is an insetted plot of station Cobar, which is highlighted in the black box.}\label{fig:basic-manip}
 \end{figure}
@@
 
-There are 639 stations that record daily precipitation in Australia and
-these series will overlay on top of each other in a glyph map.
+In the previous example, there has already been some overlapping of the
+glyphs for a few stations near (151E, 34S) and (152E, 33S) and this will
+be a problem when mapping more stations in the national level.
 Aggregation can be helpful to group series into clusters before
 visualising the cluster with glyph map. This example shows how to
-organise data at both level with the cubble function
-\texttt{switch\_key()}.
+organise data at both level with \texttt{switch\_key()}. The dataset
+used is \texttt{weatherdata::climate\_full}, which records daily
+precipitation and maximum/ minimum temperature for 639 stations in
+Australia from 2016 to 2020.
 
@@
 create 20 clusters. This creates \texttt{station\_nested} as a station
-level nested cubble with a cluster column indicating which group each
+level nested cubble with a cluster column indicating the group each
 station belongs to. More complex algorithms can be used for other
@@
 \begin{CodeInput}
-R> station_nested
-\end{CodeInput}
-\begin{CodeOutput}
-# cubble:   id [639]: nested form
-# bbox:     [113.53, -43.66, 153.64, -10.05]
-# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
-  id            lat  long  elev name             wmo_id ts               cluster
-  <chr>       <dbl> <dbl> <dbl> <chr>             <dbl> <list>             <int>
-1 ASN00001006 -15.5  128.   3.8 wyndham aero      95214 <tibble [1,827 ~       3
-2 ASN00001007 -13.8  126.   6   troughton island  94102 <tibble [1,827 ~       3
-3 ASN00001018 -16.4  126. 546   mount elizabeth   94211 <tibble [1,460 ~      16
-4 ASN00001019 -14.3  127.  23   kalumburu         94100 <tibble [1,827 ~       3
-5 ASN00001020 -14.1  126.  51   truscott          95101 <tibble [1,827 ~       3
-# ... with 634 more rows
-\end{CodeOutput}
+R> station_nested <- weatherdata::climate_full %>% 
++   mutate(cluster = ...)
+\end{CodeInput}
 \end{CodeChunk}
@@
 # bbox:     [144.52, -37.73, 146.06, -36.55]
-# temporal: date [date], prcp [dbl]
-  id          name                lat  long ts         type   dist group n_match
-  <chr>       <chr>             <dbl> <dbl> <list>     <chr> <dbl> <int>   <int>
-1 405234      SEVEN CREEKS @ D~ -36.9  146. <tibble [~ river  6.15     5      21
-2 ASN00082042 strathbogie       -36.8  146. <tibble [~ clim~  6.15     5      21
-3 404207      HOLLAND CREEK @ ~ -36.6  146. <tibble [~ river  8.54    10      21
-4 ASN00082170 benalla airport   -36.6  146. <tibble [~ clim~  8.54    10      21
-5 230200      MARIBYRNONG RIVE~ -37.7  145. <tibble [~ river  6.17     6      19
-6 ASN00086038 essendon airport  -37.7  145. <tibble [~ clim~  6.17     6      19
-7 406213      CAMPASPE RIVER @~ -37.0  145. <tibble [~ river  1.84     1      18
-8 ASN00088051 redesdale         -37.0  145. <tibble [~ clim~  1.84     1      18
+# temporal: date [date], matched_var [dbl]
+  id          name                lat  long type    dist group ts        n_match
+  <chr>       <chr>             <dbl> <dbl> <chr>  <dbl> <int> <list>      <int>
+1 405234      SEVEN CREEKS @ D~ -36.9  146. river   6.15     5 <tibble ~      21
+2 404207      HOLLAND CREEK @ ~ -36.6  146. river   8.54    10 <tibble ~      21
+3 ASN00082042 strathbogie       -36.8  146. clima~  6.15     5 <tibble ~      21
+4 ASN00082170 benalla airport   -36.6  146. clima~  8.54    10 <tibble ~      21
+5 230200      MARIBYRNONG RIVE~ -37.7  145. river   6.17     6 <tibble ~      19
+6 ASN00086038 essendon airport  -37.7  145. clima~  6.17     6 <tibble ~      19
+7 406213      CAMPASPE RIVER @~ -37.0  145. river   1.84     1 <tibble ~      18
+8 ASN00088051 redesdale         -37.0  145. clima~  1.84     1 <tibble ~      18
 \end{CodeOutput}

### 2022-02-04 data/create-data.R

@@ -1,14 +1,0 @@
-library(tidyverse)
-library(cubble)
-library(s2)
-library(ggplot2)
-state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)
-# plot_map(state_map) +
-#   geom_point(data = weatherdata::climate_large, aes(x = longitude, y = latitude))
-vic_map <- ozmaps::abs_ste %>% filter(NAME == "Victoria")
-vic_stations <- weatherdata::climate_large %>% 
-  rename(long = longitude, lat = latitude) %>% 
-  mutate(ll = s2_lnglat(long, lat)) %>% 
-  filter(s2_within(ll, vic_map)) 
-
-save(vic_stations, file = here::here("data/vic_stations.rda"))

### 2022-02-04 scripts/basic-agg.R

@@ -48,30 +48,11 @@ p1 <- ggplot(data = smoother,
   geom_glyph(height = 2, width = 4) + 
   theme_void()
 
-# tas_latlong <- station_long %>% 
-#   filter(cluster == 7) %>% 
-#   switch_key(cluster) %>% 
-#   filter(prcp == max(prcp)) %>% 
-#   tamp()
-# 
-# tas_id <- tas_latlong %>% pull(id)
-# tas_long <- station_long %>% filter(id == tas_id)
-
-# p2 <- station_long %>%
-#   ggplot(aes(x = wk, y = prcp, group = id)) +
-#   geom_line(size = 0.1) +
-#   geom_smooth(se = FALSE, inherit.aes = FALSE, aes(x = wk, y = prcp)) + 
-#   geom_line(data = tas_long, color = "red") + 
-#   facet_wrap(vars(cluster), scales = "free_y", ncol = 4) +
-#   theme_bw()
-
-
 p3 <- plot_map(state_map) +
   geom_point(data = station_nested, aes(x = long, y = lat), size = 0.5) +
   ggforce::geom_mark_hull(data = cluster_nested %>% tidyr::unnest(hull),
                           expand = 0, radius = 0,
                           aes(x = long, y = lat, group = cluster)) 
-  #geom_point(data = tas_latlong, aes(x = long, y = lat), col = "red")
 
 (p1 | p3) + plot_annotation(tag_levels = 'A')
 ggsave(filename = "figures/basic-agg.png", width = 15, height = 7)

### 2022-02-04 scripts/basic-manip.R

@@ -22,20 +22,22 @@ tmax <- weatherdata::historical_tmax %>%
   stretch() %>%
   migrate(latitude, longitude)
 
-gly <- glyphs(tmax, "longitude", "month", "latitude", "tmax",
-              height = 0.5, width = 1)
 
 nsw_vic <- ozmaps::abs_ste %>% filter(NAME %in% c("Victoria", "New South Wales"))
 
 p1 <- ggplot() + 
   geom_sf(data = nsw_vic, fill = "transparent", color = "grey", linetype = "dotted") + 
-  geom_path(data = gly, 
-            aes(gx, gy, group = interaction(id, group), color = group)) +
+  geom_glyph(data = tmax, 
+             aes(x_major = longitude, x_minor = month, 
+                 y_major = latitude, y_minor = tmax,
+                 group = interaction(id, group), color = group),
+             width = 1, height = 0.5) +
   scale_color_brewer(palette = "Dark2") + 
   theme_bw() + 
   coord_sf(xlim = c(141, 154)) + 
   theme(legend.position = "bottom") +
   labs(x = "Longitude", y = "Latitude")
+
 
 #################
 # inset

### 2022-02-05 1-intro.Rmd
modify introduction section

@@
 
-Many data structures have been proposed for spatial (\pkg{sf} by @sf) and temporal (\pkg{tsibble} by @tsibble) data in the R community, while less has been done for spatio-temporal data.
+Spatio-temporal data concerns variables in the spatial coordinates across time.
+Some variables are invariant to time and can be solely identified by its spatial location while other time varying variables need both the spatial and temporal identifier to identify.
+Manipulating data in these two levels is usually done separately and the third tidy data principal [@tidydata] recommends each type of observational unit to form a table.
+Another feature for a spatio-temporal data structure to have is that variables created from subsequent operations can be automatically added to the existing table with the same level.
+This is better than an output that index the variables with the identifiers since otherwise analysts will need to manually join this output with existing variables in the same level.
+
+<!-- Many data structures have been proposed for spatial (\pkg{sf} by @sf) and temporal (\pkg{tsibble} by @tsibble) data in the R community, while less has been done for spatio-temporal data.
 The lack of such tools could potentially because analysts usually treat the spatial and temporal dimension of the data separately, without realising the need to create a new data structure.
@@
 This additional step doesn't add new operations into the data but can be error prone.
-\newline
+\newline -->
 
@@
 While these implementations either store spatial and temporal variables all in a single table, hence duplicate the spatial variables for each temporal unit; or split them into two separate tables that has the problem of manually joining, mentioned in the previously.
-None of these packages enjoy both the benefits of being able to separate manipulation in the two dimensions while also keep the data object as a whole.
-This create a gap in the software development.
-The requirement for such a tool is important given the ubiquity of spatio-temporal vector data in the wild: the Ireland wind data from \pkg{gstat} is an classic example data that splits variables into spatial (`wind.loc`) and temporal (`wind`) dimension; Bureau of Meteorology (BoM) provides climate observations that are widely applied in agriculture and ecology study; air pollution data.
+This creates a gap in the software development.
+The requirement for such a tool is important given the ubiquity of spatio-temporal vector data in the wild: the Ireland wind data from \pkg{gstat} is an classic example data that splits variables into its spatial (`wind.loc`) and temporal (`wind`) dimension; Bureau of Meteorology (BoM) provides climate observations widely applied in agriculture and ecology study.
 \newline
@@
 An example on how cubble handle Netcdf data is also provided.
-Section 5 concludes the paper
+Section 5 concludes the paper.

### 2022-02-05 main.tex
modify introduction section

@@
 
-Many data structures have been proposed for spatial (\pkg{sf} by
-\citet{sf}) and temporal (\pkg{tsibble} by \citet{tsibble}) data in the
-R community, while less has been done for spatio-temporal data. The lack
-of such tools could potentially because analysts usually treat the
-spatial and temporal dimension of the data separately, without realising
-the need to create a new data structure. While this approach follows the
-third tidy data principal \citep{tidydata} (\emph{Each type of
-observational unit forms a table}), analysts always need to manually
-join results from different observational units or combining multiple
-tables into one for downstream analysis. This additional step doesn't
-add new operations into the data but can be error prone. \newline
+Spatio-temporal data concerns variables in the spatial coordinates
+across time. Some variables are invariant to time and can be solely
+identified by its spatial location while other time varying variables
+need both the spatial and temporal identifier to identify. Manipulating
+data in these two levels is usually done separately and the third tidy
+data principal \citep{tidydata} recommends each type of observational
+unit to form a table. Another feature for a spatio-temporal data
+structure to have is that variables created from subsequent operations
+can be automatically added to the existing table with the same level.
+This is better than an output that index the variables with the
+identifiers since otherwise analysts will need to manually join this
+output with existing variables in the same level.
 
@@
 for each temporal unit; or split them into two separate tables that has
-the problem of manually joining, mentioned in the previously. None of
-these packages enjoy both the benefits of being able to separate
-manipulation in the two dimensions while also keep the data object as a
-whole. This create a gap in the software development. The requirement
-for such a tool is important given the ubiquity of spatio-temporal
-vector data in the wild: the Ireland wind data from \pkg{gstat} is an
-classic example data that splits variables into spatial
-(\texttt{wind.loc}) and temporal (\texttt{wind}) dimension; Bureau of
-Meteorology (BoM) provides climate observations that are widely applied
-in agriculture and ecology study; air pollution data. \newline
+the problem of manually joining, mentioned in the previously. This
+creates a gap in the software development. The requirement for such a
+tool is important given the ubiquity of spatio-temporal vector data in
+the wild: the Ireland wind data from \pkg{gstat} is an classic example
+data that splits variables into its spatial (\texttt{wind.loc}) and
+temporal (\texttt{wind}) dimension; Bureau of Meteorology (BoM) provides
+climate observations widely applied in agriculture and ecology study.
+\newline
 
@@
 hydrology data. An example on how cubble handle Netcdf data is also
-provided. Section 5 concludes the paper
+provided. Section 5 concludes the paper.
 
@@
 Aggregation can be helpful to group series into clusters before
-visualising the cluster with glyph map. This example shows how to
-organise data at both level with \texttt{switch\_key()}. The dataset
-used is \texttt{weatherdata::climate\_full}, which records daily
-precipitation and maximum/ minimum temperature for 639 stations in
-Australia from 2016 to 2020.
-
-A simple kmean algorithm based on the distance matrix is used here to
-create 20 clusters. This creates \texttt{station\_nested} as a station
-level nested cubble with a cluster column indicating the group each
-station belongs to. More complex algorithms can be used for other
+visualising the cluster with glyph map.
+
+This example shows how to organise data at both level with
+\texttt{switch\_key()}. The dataset used is
+\texttt{weatherdata::climate\_full}, which records daily precipitation
+and maximum/ minimum temperature for 639 stations in Australia from 2016
+to 2020. A simple kmean algorithm based on the distance matrix is used
+here to create 20 clusters. This creates \texttt{station\_nested} as a
+station level nested cubble with a cluster column indicating the group
+each station belongs to. More complex algorithms can be used for other
 problem, as long as there is a mapping from each station to a cluster.

### 2022-02-05 5-conclusion.Rmd
chnage the affiliation of nicolas and add CSIRO as Ackno

@@
 
+# Acknowledgement
+
+This work is funded by the Commonwealth Scientific and Industrial Research Organisation (CSIRO) Data61 Scholarship and started while Nicolas Langrené was affiliated with CSIRO's Data61.
+The article is created using `knitr` [@knitr] and `rmarkdown` [@rmarkdown] in R.
+The source code for reproducing this paper can be found [here](https://github.com/huizezhang-sherry/paper-cubble).
+
 # Appendix

### 2022-02-05 main.Rmd
chnage the affiliation of nicolas and add CSIRO as Ackno

@@
     # department: 34 Village Street, Docklands VIC 3008 Australia
-    affiliation: 'CSIRO Data61 \AND'
+    affiliation: 'BNU-HKBU United International College \AND'
     # location: Melbourne, Australia

### 2022-02-05 main.tex
chnage the affiliation of nicolas and add CSIRO as Ackno

@@
 \AND Ursula Laa\\University of Natural Resources and Life Sciences
-\AND Nicolas Langrené\\CSIRO Data61 \AND Patricia
-Menéndez\\Monash University \AND
+\AND Nicolas Langrené\\BNU-HKBU United International College
+\AND Patricia Menéndez\\Monash University \AND
 }
@@
 
+\hypertarget{acknowledgement}{%
+\section{Acknowledgement}\label{acknowledgement}}
+
+This work is funded by the Commonwealth Scientific and Industrial
+Research Organisation (CSIRO) Data61 Scholarship and started while
+Nicolas Langrené was affiliated with CSIRO's Data61. The article is
+created using \texttt{knitr} and \texttt{rmarkdown} in R. The source
+code for reproducing this paper can be found
+\href{https://github.com/huizezhang-sherry/paper-cubble}{here}.
+
 \hypertarget{appendix}{%

### 2022-02-05 5-conclusion.Rmd
conclusion and affiliation

@@
 
+This paper describes the \proglang{R} package `cubble` for manipulating and visualising spatio-temporal data.
+A new data structure, that builds from the `rowwise_df` and `grouped_df` class in the tidyverse ecosystem, is proposed to connect the spatial and temporal variables in the spatio-temporal data.
+This design frees the data analysts from spending time on organising variables of different observational units.
+The data structure is also flexible to the techniques and packages analysts use to analyse the data, for example, in the matching example in section 4.3, users are free to use algorithms from another pacakge to cluster stations.
+Further development and maintenance of the package involves responding to changes in the tidyverse packages that `cubble` imports, in particular, `tibble`, `tidyr`, and `dplyr`.
+Another area for further development is to extend `cubble` to other spatial objects other than points.
+
 # Acknowledgement
@@
 The article is created using `knitr` [@knitr] and `rmarkdown` [@rmarkdown] in R.
-The source code for reproducing this paper can be found [here](https://github.com/huizezhang-sherry/paper-cubble).
+The source code for reproducing this paper can be found at: https://github.com/huizezhang-sherry/paper-cubble.
 

### 2022-02-05 main.Rmd
conclusion and affiliation

--- 2022-02-04 main.Rmd
@@
   - name: H. Sherry Zhang
-    # department: Department of Econometrics and Business Statistics
     affiliation: Monash University
-    #location: Melbourne, Australia
+    address: 21 Chancellors Walk, Clayton VIC 3800 Australia
+    department: Department of Econometrics and Business Statistics
     email: \email{huize.zhang@monash.edu}
   - name: Dianne Cook
-    # department: Department of Econometrics and Business Statistics
     affiliation: 'Monash University \AND'
+    address: 21 Chancellors Walk, Clayton VIC 3800 Australia
     # location: Melbourne, Australia
-    # email:  dicook@monash.edu
+    email:  dicook@monash.edu
   - name: Ursula Laa
-    # department: Institute of Statistics
     affiliation: 'University of Natural Resources and Life Sciences \AND'
-    # location: Vienna, Austria
-    # email:  ursula.laa@boku.ac.at
+    address: Gregor-Mendel-Straße 33, 1180 Wien, Austria
+    email:  ursula.laa@boku.ac.at
   - name: Nicolas Langrené
-    # department: 34 Village Street, Docklands VIC 3008 Australia
-    affiliation: 'CSIRO Data61 \AND'
-    # location: Melbourne, Australia
-    # email: nicolas.langrene@csiro.au
+    affiliation: 'BNU-HKBU United International College \AND'
+    addresss: 2000 Jintong Road, Tangjiawan, Zhuhai, Guangdong Province, China
+    email: nicolas@langrene.com
   - name: Patricia Menéndez
-    # department: Department of Econometrics and Business Statistics
     affiliation: 'Monash University \AND'
-    # location: Melbourne, Australia
-    # email:  patricia.menendez@monash.edu
+    address: 21 Chancellors Walk, Clayton VIC 3800 Australia
+    email:  patricia.menendez@monash.edu
 title:

### 2022-02-05 main.tex
conclusion and affiliation

@@
 \Address{
+    H. Sherry Zhang\\
+    Monash University\\
+    21 Chancellors Walk, Clayton VIC 3800 Australia\\
+  E-mail: \email{huize.zhang@monash.edu}\\
+
+      Dianne Cook\\
+    Monash University\\
+    21 Chancellors Walk, Clayton VIC 3800 Australia\\
+  E-mail: \href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}\\
+
+      Ursula Laa\\
+    University of Natural Resources and Life Sciences\\
+    Gregor-Mendel-Straße 33, 1180 Wien, Austria\\
+  E-mail: \href{mailto:ursula.laa@boku.ac.at}{\nolinkurl{ursula.laa@boku.ac.at}}\\
+
+        Patricia Menéndez\\
+    Monash University\\
+    21 Chancellors Walk, Clayton VIC 3800 Australia\\
+  E-mail: \href{mailto:patricia.menendez@monash.edu}{\nolinkurl{patricia.menendez@monash.edu}}\\
+
   }
@@
 
+This paper describes the \proglang{R} package \texttt{cubble} for
+manipulating and visualising spatio-temporal data. A new data structure,
+that builds from the \texttt{rowwise\_df} and \texttt{grouped\_df} class
+in the tidyverse ecosystem, is proposed to connect the spatial and
+temporal variables in the spatio-temporal data. This design frees the
+data analysts from spending time on organising variables of different
+observational units. The data structure is also flexible to the
+techniques and packages analysts use to analyse the data, for example,
+in the matching example in section 4.3, users are free to use algorithms
+from another pacakge to cluster stations. Further development and
+maintenance of the package involves responding to changes in the
+tidyverse packages that \texttt{cubble} imports, in particular,
+\texttt{tibble}, \texttt{tidyr}, and \texttt{dplyr}. Another area for
+further development is to extend \texttt{cubble} to other spatial
+objects other than points.
+
 \hypertarget{acknowledgement}{%
@@
 Nicolas Langrené was affiliated with CSIRO's Data61. The article is
-created using \texttt{knitr} and \texttt{rmarkdown} in R. The source
-code for reproducing this paper can be found
-\href{https://github.com/huizezhang-sherry/paper-cubble}{here}.
+created using \texttt{knitr} \citep{knitr} and \texttt{rmarkdown}
+\citep{rmarkdown} in R. The source code for reproducing this paper can
+be found at: \url{https://github.com/huizezhang-sherry/paper-cubble}.
 

### 2022-02-05 main.Rmd
small fix

@@
     affiliation: 'BNU-HKBU United International College \AND'
-    addresss: 2000 Jintong Road, Tangjiawan, Zhuhai, Guangdong Province, China
+    address: 2000 Jintong Road, Tangjiawan, Zhuhai, Guangdong Province, China
     email: nicolas@langrene.com

### 2022-02-05 main.Rmd
more keywords

@@
   # at least one keyword must be supplied
-  formatted: [spatio-temporal data,  "\\proglang{R}", climate data]
-  plain:     [spatio-temporal data, R, climate data]
+  formatted: [spatio temporal,  "\\proglang{R}", climate weather station, data analysis]
+  plain:     [spatio temporal, R, climate weather station, data analysis]
 preamble: >

### 2022-02-05 3-advanced.Rmd
add caption for figure 5

@@
 
-```{r illu-interactive, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "demon interactivity"}
+```{r illu-interactive, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "An illustration of the data model under interactive graphics with cubble.
+The line plot and map is made separately with the long and nested cubble.
+When a station is selected on the map (left), the corresponding row in the nested cubble will be activated.
+This will link to all the rows with the same id in the long cubble (middle) and update the line plot (right)."}
 knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.004.png"))

### 2022-02-05 main.tex
small fix
add caption for figure 5

@@
 
-\Keywords{spatio-temporal data, \proglang{R}, climate data}
-\Plainkeywords{spatio-temporal data, R, climate data}
+\Keywords{spatio temporal, \proglang{R}, climate weather station, data
+analysis}
+\Plainkeywords{spatio temporal, R, climate weather station, data
+analysis}
 
@@
   E-mail: \href{mailto:ursula.laa@boku.ac.at}{\nolinkurl{ursula.laa@boku.ac.at}}\\
+
+      Nicolas Langrené\\
+    BNU-HKBU United International College\\
+    2000 Jintong Road, Tangjiawan, Zhuhai, Guangdong Province, China\\
+  E-mail: \href{mailto:nicolas@langrene.com}{\nolinkurl{nicolas@langrene.com}}\\
 
@@
 
-\caption[demon interactivity]{demon interactivity}\label{fig:illu-interactive}
+\caption[An illustration of the data model under interactive graphics with cubble]{An illustration of the data model under interactive graphics with cubble.
+The line plot and map is made separately with the long and nested cubble.
+When a station is selected on the map (left), the corresponding row in the nested cubble will be activated.
+This will link to all the rows with the same id in the long cubble (middle) and update the line plot (right).}\label{fig:illu-interactive}
 \end{figure}

### 2022-02-06 2-the-cubble-pkg.Rmd
fix marked red sentences labelled for myseyself

@@
 
-\textcolor{red}{To use functionalities from cubble, data analysts first need to create a cubble.} `as_cubble()` creates a \pkg{cubble} by supplying the three key components: `key` as the spatial identifier; `index` as the temporal identifier; and a vector of `coords` in the order of longitude first and then latitude.
+`as_cubble()` creates a \pkg{cubble} by supplying the three key components: `key` as the spatial identifier; `index` as the temporal identifier; and a vector of `coords` in the order of longitude first and then latitude.
 The naming of `key` and `index` follows the convention in the \pkg{tsibble} package.
 The cubble created by default is in the nested form.
-Below is an example of creating a cubble: \newline
+Below is an example of creating a cubble:
 
@@
 cubble_nested <- cubble::climate_flat %>%
-  as_cubble(key = id, index = date, coords = c("long", "lat"))
+  as_cubble(key = id, index = date, coords = c(long, lat))
 cubble_nested
@@
 
-There are a few information in the \pkg{cubble} header: the name of the `key` variable (`id`) and its number (5), bounding box (`[115.97, -32.94, 133.55, -12.42]`), and also the name of variable nested in the `ts` column with its type (`date [date], prcp [dbl], tmax [dbl], tmin [dbl]`).
+There are a few information in the \pkg{cubble} header: the name of the `key` variable (`id`) and its number (5), bounding box (`[115.97, -32.94, 133.55, -12.42]`), and also the name of variable nested in the `ts` column with its type: `date [date], prcp [dbl], tmax [dbl], tmin [dbl]`.
 
@@
 
-Notice here that the third line in the header now shows the name and type of spatial variables (`lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]`).
+Notice here that the third line in the header now shows the name and type of spatial variables: `lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]`.
 
@@
 
-\textcolor{red}{As a final data output for modelling or visualisation, spatio-temporal data is usually expected to be in a single table.} Function `migrate()` moves the spatial variables from the `spatial` attribute into the long cubble:
+Some functions require time invariant variables to be in the same table as the time varying ones.
+In cubble, this can be done through `migrate()`, which moves the time invariant variables into the long cubble:
 
-```{r}
-cubble_long %>% migrate(long, lat)
+```{r echo = TRUE}
+cubble_mig <- cubble_long %>% migrate(long, lat)
+cubble_mig
 ```
 
-<!-- In the workflow described above, data objects come into cubble in the nested form, then various operations on the spatial and temporal dimension can go back and forth between the nested and long form, and finally, the data will come out of cubble in the long form for further modelling or visualisation.
--->
+Notice that these added time invariant variables will not be remembered the cubble and will disappear if switched to the nested form and then switched back.
+Hence they should most likely been used at the last step of the analysis:
 
-Building from an underlying `tbl_df` structure, it is natural to implement methods available in `dplyr` to `cubble`. Supported methods in the `cubble` with `dplyr` generics includes:
-
-```{=tex}
-\begin{center}
-\begin{tabular}{ | m{5em} | m{15cm}| } 
-basics & \textbf{mutate}, \textbf{filter}, \textbf{summarise}, \textbf{select}, \textbf{arrange}, \textbf{rename}, \textbf{left\_join} \\
-grouping &  \textbf{group\_by}, \textbf{ungroup}\\
-slice family & \textbf{slice\_head}, \textbf{slice\_tail}, \textbf{slice\_sample}, \textbf{slice\_min} and \textbf{slice\_max} \\
-\end{tabular}
-\end{center}
+```{r echo = TRUE}
+cubble_mig %>% tamp() %>% stretch()
 ```
 
-\pkg{cubble} is also compatible with \pkg{tsibble} where each element list-column in the nested form can be a `tbl_ts` object.
-\textcolor{red}{Duplicates and gaps should be first checked before structuring the data into a cubble.} If the input data is a \pkg{tsibble} object, the long form cubble is also a \pkg{tsibble} where users can directly apply time series operations.
-\textcolor{red}{more experiment with tsibble}
+## Compatibility with existing packages
+
+Building from an `tbl_df` class, `cubble` has implemented methods for `dplyr` generics, which includes:
+
+  - basic: mutate, filter, summarise, select, arrange, rename, left_join
+  - grouping: group_by, ungroup
+  - slice family: slice_head, slice_tail, slice_sample, slice_min, slice_max
+
+\pkg{cubble} is also compatible with \pkg{tsibble}. When creating a cubble from a tsibble, Only the `coords` argument need to be specified:
+
+```{r echo = TRUE}
+# casting a tsibble into cubble
+dt <- climate_flat %>% 
+  tsibble::as_tsibble(key = id, index = date) %>% 
+  cubble::as_cubble(coords = c(long, lat))
+dt
+```
+
+In the created nested form, each element of the list-column `ts` is of `tbl_ts` class so that existing operations on the tsibble class can still be applied:
+
+```{r echo = TRUE}
+# add station-based features in the nested form.
+dt %>% 
+  mutate(fabletools::features(ts, tmax, list(tmax_mean = mean, tmax_var = var)))
+```

### 2022-02-06 3-advanced.Rmd
fix marked red sentences labelled for myseyself

@@
 countries are nested in continents), or formed by clustering.
-Rather than analysing variables in the site level, summarised variables in the cluster level can give a clearer picture of \textcolor{red}{local dynamic}.
+Rather than analysing variables in the site level, summarised variables in the cluster level can give a crisper picture of local areas.
 In cubble, `switch_key()` can be used to create a new level of grouping of spatial location by specifying a clustering variable.
-Figure \ref{fig:illu-hier} illustrates the relationship of cubbles at site and cluster level, in both the long and nested form.
-\textcolor{red}{unify the use of site/ station}.
+Figure \ref{fig:illu-hier} illustrates the relationship of cubbles at station and cluster level, in both the long and nested form.
 By specifying `cluster_nested <- station_nested %>% switch_key(key = cluster)`, the cubble re-defines the cubble key from `id` in `station_nested` to `cluster` in `cluster_nested`.

### 2022-02-06 4-1-basic.Rmd
fix marked red sentences labelled for myseyself

@@
 All the stations start recording from year `r range(weatherdata::historical_tmax$first_year)[2]` with the longest back to year `r range(weatherdata::historical_tmax$first_year)[1]`.
-The data `historical_tmax` is already presented as a cubble, with `id` as the key, `date` as the index, and `c("longitude", "latitude")` as the coordinates.
+The data `historical_tmax` is already presented as a cubble, with `id` as the key, `date` as the index, and `c(longitude, latitude)` as the coordinates.
 This example compares the maximum temperature in two periods: 1971 - 1975 and 2016 - 2020 for stations in Victoria and New South Wales.
 
-Stations in the two states can be subsetted on the station number: Australia GHCN station number starts with "ASN00" and followed by the [Bureau of Meteorology (BOM) station number](http://www.bom.gov.au/climate/cdo/about/site-num.shtml), where the 2nd and 3rd digit (7th and 8th in the GHCN number) define the state of the station.
+Stations in the two states can be subsetted by the station number: Australia GHCN station number starts with "ASN00" and followed by the [Bureau of Meteorology (BOM) station number](http://www.bom.gov.au/climate/cdo/about/site-num.shtml), where the 2nd and 3rd digit (7th and 8th in the GHCN number) define the state of the station.
 New South Wales stations start from 46 to 75 and Victoria stations follow from 76 to 90.
@@
 
-Filtering for the period 1971: 1975 and 2016:2020 is an operation on the time dimension and the nested cubble needs to be switched to the long cubble by `stretch()`:
+Filtering for the period 1971 ~ 1975 and 2016 ~ 2020 is an operation on the time dimension and the nested cubble needs to be switched to the long cubble by `stretch()`:
 
@@
 tmax <- tmax %>%
-  mutate(month = lubridate::month(date), 
+  group_by(month = lubridate::month(date), 
          group = as.factor(ifelse(lubridate::year(date) > 2015, 
                                   "2016 ~ 2020", "1971 ~ 1975"))) %>%
-  group_by(month, group) %>%
   summarise(tmax = mean(tmax, na.rm = TRUE))

### 2022-02-06 main.tex
fix marked red sentences labelled for myseyself

@@
 
-\newpage
-
 \hypertarget{the-cubble-package}{%
@@
 
-\textcolor{red}{To use functionalities from cubble, data analysts first need to create a cubble.}
 \texttt{as\_cubble()} creates a \pkg{cubble} by supplying the three key
@@
 R> cubble_nested <- cubble::climate_flat %>%
-+   as_cubble(key = id, index = date, coords = c("long", "lat"))
++   as_cubble(key = id, index = date, coords = c(long, lat))
 R> cubble_nested
@@
 Notice here that the third line in the header now shows the name and
-type of spatial variables
-(\texttt{lat\ {[}dbl{]},\ long\ {[}dbl{]},\ elev\ {[}dbl{]},\ name\ {[}chr{]},\ wmo\_id\ {[}dbl{]}}).
+type of spatial variables:
+\texttt{lat\ {[}dbl{]},\ long\ {[}dbl{]},\ elev\ {[}dbl{]},\ name\ {[}chr{]},\ wmo\_id\ {[}dbl{]}}.
 
@@
 
-\textcolor{red}{As a final data output for modelling or visualisation, spatio-temporal data is usually expected to be in a single table.}
-Function \texttt{migrate()} moves the spatial variables from the
-\texttt{spatial} attribute into the long cubble:
-
-\begin{CodeChunk}
+Some functions require time invariant variables to be in the same table
+as the time varying ones. In cubble, this can be done through
+\texttt{migrate()}, which moves the time invariant variables into the
+long cubble:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> cubble_mig <- cubble_long %>% migrate(long, lat)
+R> cubble_mig
+\end{CodeInput}
 \begin{CodeOutput}
@@
 
-Building from an underlying \texttt{tbl\_df} structure, it is natural to
-implement methods available in \texttt{dplyr} to \texttt{cubble}.
-Supported methods in the \texttt{cubble} with \texttt{dplyr} generics
-includes:
-
-\begin{center}
-\begin{tabular}{ | m{5em} | m{15cm}| } 
-basics & \textbf{mutate}, \textbf{filter}, \textbf{summarise}, \textbf{select}, \textbf{arrange}, \textbf{rename}, \textbf{left\_join} \\
-grouping &  \textbf{group\_by}, \textbf{ungroup}\\
-slice family & \textbf{slice\_head}, \textbf{slice\_tail}, \textbf{slice\_sample}, \textbf{slice\_min} and \textbf{slice\_max} \\
-\end{tabular}
-\end{center}
-
-\pkg{cubble} is also compatible with \pkg{tsibble} where each element
-list-column in the nested form can be a \texttt{tbl\_ts} object.
-\textcolor{red}{Duplicates and gaps should be first checked before structuring the data into a cubble.}
-If the input data is a \pkg{tsibble} object, the long form cubble is
-also a \pkg{tsibble} where users can directly apply time series
-operations. \textcolor{red}{more experiment with tsibble}
-
-\newpage
+Notice that these added time invariant variables will not be remembered
+the cubble and will disappear if switched to the nested form and then
+switched back. Hence they should most likely been used at the last step
+of the analysis:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> cubble_mig %>% tamp() %>% stretch()
+\end{CodeInput}
+\begin{CodeOutput}
+# cubble:  date, id [5]: long form
+# bbox:    [115.97, -32.94, 133.55, -12.42]- check gap on long and lat
+# spatial: long [dbl], lat [dbl], elev [dbl], name [chr], wmo_id [dbl]
+  id          date        prcp  tmax  tmin
+  <chr>       <date>     <dbl> <dbl> <dbl>
+1 ASN00009021 2020-01-01     0  31.9  15.3
+2 ASN00009021 2020-01-02     0  24.9  16.4
+3 ASN00009021 2020-01-03     6  23.2  13  
+4 ASN00009021 2020-01-04     0  28.4  12.4
+5 ASN00009021 2020-01-05     0  35.3  11.6
+# ... with 1,825 more rows
+\end{CodeOutput}
+\end{CodeChunk}
+
+\hypertarget{compatibility-with-existing-packages}{%
+\subsection{Compatibility with existing
+packages}\label{compatibility-with-existing-packages}}
+
+Building from an \texttt{tbl\_df} class, \texttt{cubble} has implemented
+methods for \texttt{dplyr} generics, which includes:
+
+\begin{itemize}
+\tightlist
+\item
+  basic: mutate, filter, summarise, select, arrange, rename, left\_join
+\item
+  grouping: group\_by, ungroup
+\item
+  slice family: slice\_head, slice\_tail, slice\_sample, slice\_min,
+  slice\_max
+\end{itemize}
+
+\pkg{cubble} is also compatible with \pkg{tsibble}. When creating a
+cubble from a tsibble, Only the \texttt{coords} argument need to be
+specified. In the created nested form, each element of the list-column
+\texttt{ts} is of \texttt{tbl\_ts} class so that existing operations on
+the tsibble class can still be applied:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> # casting a tsibble into cubble
+R> dt <- climate_flat %>% 
++   tsibble::as_tsibble(key = id, index = date) %>% 
++   cubble::as_cubble(coords = c(long, lat))
+R> dt
+\end{CodeInput}
+\begin{CodeOutput}
+# cubble:   id [5]: nested form
+# bbox:     [115.97, -32.94, 133.55, -12.42]- check gap on long and lat
+# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
+  id            lat  long  elev name           wmo_id ts                
+  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>            
+1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tbl_ts [366 x 4]>
+2 ASN00010311 -31.9  117. 179   york            94623 <tbl_ts [366 x 4]>
+3 ASN00010614 -32.9  117. 338   narrogin        94627 <tbl_ts [366 x 4]>
+4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tbl_ts [366 x 4]>
+5 ASN00015131 -17.6  134. 220   elliott         94236 <tbl_ts [366 x 4]>
+\end{CodeOutput}
+\begin{CodeInput}
+R> # add station-based features in the nested form.
+R> dt %>% 
++   mutate(fabletools::features(ts, tmax, list(tmax_mean = mean, tmax_var = var)))
+\end{CodeInput}
+\begin{CodeOutput}
+# cubble:   id [5]: nested form
+# bbox:     [115.97, -32.94, 133.55, -12.42]- check gap on long and lat
+# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
+  id            lat  long  elev name           wmo_id ts      tmax_mean tmax_var
+  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>      <dbl>    <dbl>
+1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tbl_t~      25.7    38.6 
+2 ASN00010311 -31.9  117. 179   york            94623 <tbl_t~      26.2    51.1 
+3 ASN00010614 -32.9  117. 338   narrogin        94627 <tbl_t~      23.7    45.4 
+4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tbl_t~      33.1     3.02
+5 ASN00015131 -17.6  134. 220   elliott         94236 <tbl_t~      34.6    24.7 
+\end{CodeOutput}
+\end{CodeChunk}
 
@@
 clustering. Rather than analysing variables in the site level,
-summarised variables in the cluster level can give a clearer picture of
-\textcolor{red}{local dynamic}. In cubble, \texttt{switch\_key()} can be
-used to create a new level of grouping of spatial location by specifying
-a clustering variable. Figure \ref{fig:illu-hier} illustrates the
-relationship of cubbles at site and cluster level, in both the long and
-nested form. \textcolor{red}{unify the use of site/ station}. By
-specifying
+summarised variables in the cluster level can give a crisper picture of
+local areas. In cubble, \texttt{switch\_key()} can be used to create a
+new level of grouping of spatial location by specifying a clustering
+variable. Figure \ref{fig:illu-hier} illustrates the relationship of
+cubbles at station and cluster level, in both the long and nested form.
+By specifying
 \texttt{cluster\_nested\ \textless{}-\ station\_nested\ \%\textgreater{}\%\ switch\_key(key\ =\ cluster)},
@@
 
-\newpage
-
 \hypertarget{examples}{%
@@
 \texttt{id} as the key, \texttt{date} as the index, and
-\texttt{c("longitude",\ "latitude")} as the coordinates. This example
+\texttt{c(longitude,\ latitude)} as the coordinates. This example
 compares the maximum temperature in two periods: 1971 - 1975 and 2016 -
@@
 
-Stations in the two states can be subsetted on the station number:
+Stations in the two states can be subsetted by the station number:
 Australia GHCN station number starts with ``ASN00'' and followed by the
@@
 
-Filtering for the period 1971: 1975 and 2016:2020 is an operation on the
-time dimension and the nested cubble needs to be switched to the long
-cubble by \texttt{stretch()}:
+Filtering for the period 1971 \textasciitilde{} 1975 and 2016
+\textasciitilde{} 2020 is an operation on the time dimension and the
+nested cubble needs to be switched to the long cubble by
+\texttt{stretch()}:
 
@@
 R> tmax <- tmax %>%
-+   mutate(month = lubridate::month(date), 
++   group_by(month = lubridate::month(date), 
 +          group = as.factor(ifelse(lubridate::year(date) > 2015, 
 +                                   "2016 ~ 2020", "1971 ~ 1975"))) %>%
-+   group_by(month, group) %>%
 +   summarise(tmax = mean(tmax, na.rm = TRUE))

### 2022-02-06 scripts/basic-manip.R

@@ -7,10 +7,9 @@ tmax <- weatherdata::historical_tmax %>%
   # filter out observations in between 1971-1975 and 2016 - 2020
   stretch() %>%
   filter(lubridate::year(date) %in% c(1971:1975, 2016:2020)) %>%
-  mutate(month = lubridate::month(date), 
+  group_by(month = lubridate::month(date), 
          group = as.factor(ifelse(lubridate::year(date) > 2015, 
                                   "2016 ~ 2020", "1971 ~ 1975"))) %>%
-  group_by(month, group) %>%
   # summarise tmax into monthly mean for both periods
   summarise(tmax = mean(tmax, na.rm = TRUE)) %>% 
   

### 2022-02-06 2-the-cubble-pkg.Rmd
all the  -> \code{CODE} as per jss  guideline

@@
 No matter which form the data is in, these formats share some common components that characterise the spatio-temporal data.
-A spatial identifier (`id` in the diagram) is the unique identifier of each site.
-The temporal identifier (`t` in the diagram) prescribes the time stamp each site is recorded.
-Coordinates, comprising of latitude and longitude (`lon` and `lat` in the diagram), locates each site on the map.
+A spatial identifier (\code{id} in the diagram) is the unique identifier of each site.
+The temporal identifier (\code{t} in the diagram) prescribes the time stamp each site is recorded.
+Coordinates, comprising of latitude and longitude (\code{lat} and \code{(long)} in the diagram), locates each site on the map.
 These identifiers will be the building blocks for the data structure introduced below.
@@
 the output of counting the number of raining day is only identified by sites and hence should be performed with the nested form).
-Underneath the nested form, a cubble is built from a row-wise dataframe (`rowwise_df`) where each site forms a separate group.
-This structure simplifies the calculation that involves temporal variables by avoiding the use of `map` syntax when working with list-column.
+Underneath the nested form, a cubble is built from a row-wise dataframe (\code{rowwise_df}) where each site forms a separate group.
+This structure simplifies the calculation that involves temporal variables by avoiding the use of \code{map} syntax when working with list-column.
 
@@
 
-`as_cubble()` creates a \pkg{cubble} by supplying the three key components: `key` as the spatial identifier; `index` as the temporal identifier; and a vector of `coords` in the order of longitude first and then latitude.
-The naming of `key` and `index` follows the convention in the \pkg{tsibble} package.
+`as_cubble()} creates a \pkg{cubble} by supplying the three key components: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order of longitude first and then latitude.
+The naming of \code{key} and \code{index} follows the convention in the \pkg{tsibble} package.
 The cubble created by default is in the nested form.
@@
 
-There are a few information in the \pkg{cubble} header: the name of the `key` variable (`id`) and its number (5), bounding box (`[115.97, -32.94, 133.55, -12.42]`), and also the name of variable nested in the `ts` column with its type: `date [date], prcp [dbl], tmax [dbl], tmin [dbl]`.
+There are a few information in the \pkg{cubble} header: the name of the \code{key} variable (`id`) and its number (5), bounding box (`[115.97, -32.94, 133.55, -12.42]`), and also the name of variable nested in the \code{ts} column with its type: \code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.
 
@@
 The long cubble is suitable to manipulate the time dimension of the data.
-The function `stretch()` switches the nested cubble into the long cubble by first extracts all the spatial variables into a separate tibble and store in the `spatial` attribute and then unnests the `ts` column:
+The function \code{stretch()} switches the nested cubble into the long cubble by first extracts all the spatial variables into a separate tibble and store in the \code{spatial} attribute and then unnests the \code{ts} column:
 
@@
 
-Notice here that the third line in the header now shows the name and type of spatial variables: `lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]`.
+Notice here that the third line in the header now shows the name and type of spatial variables: \code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]}.
 
@@
 Many times, analysts will need to go back and forth between the nested and long cubble.
-`tamp()` inverses the `stretch()` that introduced in the previous section, to switch from a long cubble to the nested cubble:
+\code{tamp()} inverses the \code{stretch()} that introduced in the previous section, to switch from a long cubble to the nested cubble:
 
@@
 Some functions require time invariant variables to be in the same table as the time varying ones.
-In cubble, this can be done through `migrate()`, which moves the time invariant variables into the long cubble:
+In cubble, this can be done through \code{migrate()}, which moves the time invariant variables into the long cubble:
 
@@
 
-Building from an `tbl_df` class, `cubble` has implemented methods for `dplyr` generics, which includes:
+Building from an \code{tbl_df} class, \pkg{cubble} has implemented methods for \pkg{dplyr} generics, which includes:
 
@@
 
-\pkg{cubble} is also compatible with \pkg{tsibble}. When creating a cubble from a tsibble, Only the `coords` argument need to be specified:
+\pkg{cubble} is also compatible with \pkg{tsibble}. When creating a cubble from a tsibble, Only the \code{coords} argument need to be specified:
 
@@
 
-In the created nested form, each element of the list-column `ts` is of `tbl_ts` class so that existing operations on the tsibble class can still be applied:
+In the created nested form, each element of the list-column \code{ts} is of \code{tbl_ts} class so that existing operations on the tsibble class can still be applied:
 

### 2022-02-06 3-advanced.Rmd
all the  -> \code{CODE} as per jss  guideline

@@
 Rather than analysing variables in the site level, summarised variables in the cluster level can give a crisper picture of local areas.
-In cubble, `switch_key()` can be used to create a new level of grouping of spatial location by specifying a clustering variable.
+In cubble, \code{switch_key()} can be used to create a new level of grouping of spatial location by specifying a clustering variable.
 Figure \ref{fig:illu-hier} illustrates the relationship of cubbles at station and cluster level, in both the long and nested form.
-By specifying `cluster_nested <- station_nested %>% switch_key(key = cluster)`, the cubble re-defines the cubble key from `id` in `station_nested` to `cluster` in `cluster_nested`.
-All the spatial variables variant to `cluster` are now nested into a `.val` column and cluster level variables can be computed in the same fashion as station level variables in `station_nested`.
-Following the same principal of `stretch`, the long form cluster level cubble expands the time series variables with both indices: `cluster` and `id`.
+By specifying \code{cluster_nested <- station_nested \%>\% switch_key(key = cluster)}, the cubble re-defines the cubble key from \code{id} in \code{station_nested} to \code{cluster} in \code{cluster_nested}.
+All the spatial variables variant to \code{cluster} are now nested into a \code{.val} column and cluster level variables can be computed in the same fashion as station level variables in \code{station_nested}.
+Following the same principal of \code{stretch}.
+the long form cluster level cubble expands the time series variables with both indices: \code{cluster} and \code{id}.
 All the cluster level variables are stored separately for recovering the nested from from the long form.
@@
 
-`match_sites()` from `cubble` matches data based on these two criteria.
+`match_sites()} from \pkg{cubble} matches data based on these two criteria.
 It first matches the two data sources spatially through computing the pairwise distance based on latitude and longitude.
@@
 Figure \ref{fig:illu-matching} illustrate this temporal matching in more details.
-Given two series `A` and `a`, 3 peaks have been picked in each series.
-An interval, with default length of 5, is constructed for each peak in series `A` and the peaks in series `a` are tested against whether they fall into the any of the intervals.
+Given two series \code{A} and \code{a}.
+3 peaks have been picked in each series.
+An interval, with default length of 5, is constructed for each peak in series \code{A} and the peaks in series \code{a} are tested against whether they fall into the any of the intervals.
 In this illustration, there are 2 matches for these two series.
-Several arguments are available in `match_sites()` to fine-tune the matching:
+Several arguments are available in \code{match_sites()} to fine-tune the matching:
 
@@
 
- - `spatial_n_keep`: the number of spatial match for each site to keep
- - `spatial_dist_max`: the maximum distance allowed for a matched pair 
- - `temporal_n_highest`: the number of peak used - 3 in the example above
- - `temporal_window`: the length of the interval - 5 in the example above
- - `temporal_min_match`: the minimum number of matched peak for a valid matched pair
+ - \code{spatial_n_keep}. the number of spatial match for each site to keep
+ - \code{spatial_dist_max}. the maximum distance allowed for a matched pair 
+ - \code{temporal_n_highest}. the number of peak used - 3 in the example above
+ - \code{temporal_window}. the length of the interval - 5 in the example above
+ - \code{temporal_min_match}. the minimum number of matched peak for a valid matched pair
 
@@
 The map and time series plot is associated with the nested or long cubble respectively and when a user action is captured on the map, the site will be activated in the nested cubble (left).
-The nested cubble will communicate to the long cubble to activate all the observations with the same `id` (middle).
+The nested cubble will communicate to the long cubble to activate all the observations with the same \code{id} (middle).
 The long cubble will then highlight the activated series in the time series plot (right).
@@
 The selection(s) on the time series is through selecting the point(s) on the time series and once a point is selected, it will be activated in the long cubble.
-All the observations that share the same `id` are then activated and this includes other points in the same time series in the long cubble and the corresponding observation of site in the nested cubble.
+All the observations that share the same \code{id} are then activated and this includes other points in the same time series in the long cubble and the corresponding observation of site in the nested cubble.
 These activated observations will then being reflected in the updated plots and Diagram \ref{fig:illu-interactive-2} in the Appendix illustrates this process.
@@
 
-Glyph map [@Wickham2012-yr] plots the time series as single glyph on the map and in \proglang{R}, `GGally` implements the glyph map through the `glyphs()` function.
-The function construct a data frame with calculated position (`gx`, `gy`, `gid`) of each point on the time series using linear algebra (Equation 1 and 2 in @Wickham2012-yr).
-The data can then be piped into `ggplot` to create the glyph map as:
+Glyph map [@Wickham2012-yr] plots the time series as single glyph on the map and in \proglang{R}, \pkg{GGally} implements the glyph map through the \code{glyphs()} function.
+The function construct a data frame with calculated position (`gx}.
+\code{gy}.
+\code{gid}.
+of each point on the time series using linear algebra (Equation 1 and 2 in @Wickham2012-yr).
+The data can then be piped into \code{ggplot} to create the glyph map as:
 
@@
 
-A re-implementation of the glyph map as a ggproto `GeomGlyph` has been made in the `cubble` package so that the glyph map can be created with `geom_glyph()`:
+A re-implementation of the glyph map as a ggproto \code{GeomGlyph} has been made in the \pkg{cubble} package so that the glyph map can be created with \code{geom_glyph()}:
 
@@
 
-Polar glyph map can be specify as a parameter, `polar = TRUE`, in the `geom_glyph()`, along with `width` and `height` in either absolute or relative value.
-Global and local scale can be controlled by the parameter `global_rescale`, which default to `TRUE` for global scaling.
-Reference box and line can be added with separate `geom_glyph_box()` and `geom_glyph_line()`.
+Polar glyph map can be specify as a parameter, \code{polar = TRUE}.
+in the \code{geom_glyph()}.
+along with \code{width} and \code{height} in either absolute or relative value.
+Global and local scale can be controlled by the parameter \code{global_rescale}.
+which default to \code{TRUE} for global scaling.
+Reference box and line can be added with separate \code{geom_glyph_box()} and \code{geom_glyph_line()`.

### 2022-02-06 4-1-basic.Rmd
all the  -> \code{CODE} as per jss  guideline

@@
 
-Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world and the dataset `weatherdata::historical_tmax` extracts the maximum temperature for `r nrow(weatherdata::historical_tmax)` Australian stations.
+Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world and the dataset \code{weatherdata::historical_tmax} extracts the maximum temperature for `r nrow(weatherdata::historical_tmax)` Australian stations.
 All the stations start recording from year `r range(weatherdata::historical_tmax$first_year)[2]` with the longest back to year `r range(weatherdata::historical_tmax$first_year)[1]`.
-The data `historical_tmax` is already presented as a cubble, with `id` as the key, `date` as the index, and `c(longitude, latitude)` as the coordinates.
+The data \code{historical_tmax} is already presented as a cubble, with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates.
 This example compares the maximum temperature in two periods: 1971 - 1975 and 2016 - 2020 for stations in Victoria and New South Wales.
@@
 
-Filtering for the period 1971 ~ 1975 and 2016 ~ 2020 is an operation on the time dimension and the nested cubble needs to be switched to the long cubble by `stretch()`:
+Filtering for the period 1971 ~ 1975 and 2016 ~ 2020 is an operation on the time dimension and the nested cubble needs to be switched to the long cubble by \code{stretch()}:
 
@@
 These stations are filtered out by examining whether the summarised time series has a total of 24 months.
-The long cubble needs to be switched to the nested form for this spatial operation using `tamp()`:
+The long cubble needs to be switched to the nested form for this spatial operation using \code{tamp()}:
 
@@
 
-Lastly, to create a glyph map, both the major (`longitude`, `latitude`) and minor (`month`, `tmax`) coordinates need to be in the same table.
-Spatial variables can be moved to the long form with `migrate()`:
+Lastly, to create a glyph map, both the major (\code{longitude}, \code{latitude}) and minor (\code{month}, \code{tmax}) coordinates need to be in the same table.
+Spatial variables can be moved to the long form with \code{migrate()}:
 
@@
 
-`tmax` can then be supplied to `geom_glyph()` for the glyph map in Figure \ref{fig:basic-manip} with a station inset on the top left corner:
+\code{tmax} can then be supplied to \code{geom_glyph()} for the glyph map in Figure \ref{fig:basic-manip} with a station inset on the top left corner:
 

### 2022-02-06 4-2-agg.Rmd
all the  -> \code{CODE} as per jss  guideline

@@
 
-This example shows how to organise data at both level with `switch_key()`.
-The dataset used is `weatherdata::climate_full`, which records daily precipitation and maximum/ minimum temperature for `r nrow(weatherdata::climate_full)` stations in Australia from 2016 to 2020.
+This example shows how to organise data at both level with \code{switch_key()}.
+The dataset used is \code{weatherdata::climate_full}, which records daily precipitation and maximum/ minimum temperature for \code{r nrow(weatherdata::climate_full)} stations in Australia from 2016 to 2020.
 A simple kmean algorithm based on the distance matrix is used here to create 20 clusters.
-This creates `station_nested` as a station level nested cubble with a cluster column indicating the group each station belongs to.
+This creates \code{station_nested} as a station level nested cubble with a cluster column indicating the group each station belongs to.
 More complex algorithms can be used for other problem, as long as there is a mapping from each station to a cluster.
@@
 
-To create a group level cubble, use `switch_key()` with the new key variable, `cluster`:
+To create a group level cubble, use \code{switch_key()} with the new key variable, \code{cluster}:
 
@@
 
-With the group level cubble, `get_centroid()` is useful to compute the centroid of each cluster, which will be used as the major axis for the glyph map later:
+With the group level cubble, \code{get_centroid()} is useful to compute the centroid of each cluster, which will be used as the major axis for the glyph map later:
 

### 2022-02-06 4-3-water-level.Rmd
all the  -> \code{CODE} as per jss  guideline

@@
 From the map, a few water gauges and weather stations are close to each other and there is the potential that fluctuation of the water level can reflect the precipitation measured by the climate station.
-As introduced in Section 3.2, `match_sites()` can be used to match one source of data with another source in a cubble.
-Here `prcp` in `climate` will be matched to `Water_course_level` in `river`, which is specified in the argument `temporal_var_to_match`.
-`temporal_independent` controls the dataset that will be used to construct the interval (see the illustration in section 3.2 for details on this).
+As introduced in Section 3.2, \code{match_sites()} can be used to match one source of data with another source in a cubble.
+Here \code{prcp} in \code{climate} will be matched to \code{Water_course_level} in \code{river}, which is specified in the argument \code{temporal_var_to_match}.
+\code{temporal_independent} controls the dataset that will be used to construct the interval (see the illustration in section 3.2 for details on this).
 Here the goal is to see if precipitation will be reflected by the water level in the river and hence a lagged effect of water level on precipitation.
-This will put precipitation data, `climate`, as the independent.
-Given there is one year worth of data, the number of peak (`temporal_n_highest`) to consider is slightly raised from a default 20 to 30.
-`temporal_min_match` filters out pairs don't have enough match and to return all the pairs, set `temporal_min_match` to `0`
+This will put precipitation data, \code{climate}, as the independent.
+Given there is one year worth of data, the number of peak (\code{temporal_n_highest}) to consider is slightly raised from a default 20 to 30.
+\code{temporal_min_match} filters out pairs don't have enough match and to return all the pairs, set \code{temporal_min_match} to 0.
 
@@
 
-The output from matching is also a cubble, with additional column `dist` and `group` produced from spatial matching and `n_match` from the temporal matching.
+The output from matching is also a cubble, with additional column \code{dist} and \code{group} produced from spatial matching and \code{n_match} from the temporal matching.
 

### 2022-02-06 4-4-netcdf.Rmd
all the  -> \code{CODE} as per jss  guideline

@@
 Attributes are usually associated with dimension and variable in the NetCDF format data and a [metadata convention for climate and forecast](http://cfconventions.org/) has been designed to standardise this data format.
-Current packages on manipulating NetCDF data in R include a high-level R interface: `ncdf4`[@ncdf4], a low-level interface that calls C-interface: `RNetCDF`[@rnetcdf; @michna2013rnetcdf], and a tidyverse implementation: `tidync`[@tidync].
+Current packages on manipulating NetCDF data in R include a high-level R interface: \pkg{ncdf4} [@ncdf4], a low-level interface that calls C-interface: \pkg{RNetCDF} [@rnetcdf; @michna2013rnetcdf], and a tidyverse implementation: \pkg{tidync} [@tidync].
 
-Cubble provides an `as_cubble()` method to coerce the `ncdf4` class from the `ncdf4` package into a `cubble`.
-It maps each combination of longitude and latitude into an `id` (the `key`) and nests time and variables in the nested form:
+Cubble provides an \code{as_cubble()} method to coerce the \code{ncdf4} class from the \pkg{ncdf4} package into a \code{cubble}.
+It maps each combination of longitude and latitude into an \code{id} (the \code{key}) and nests time and variables in the nested form:
 
@@
 A 300 by 300 spatial grid can be a bounding box of [100, -80, 180, 0] at 0.25 degree resolution or global bounding box [-180, -90, 180, -90] at 1 degree resolution.
-Subsetting longitude and latitude grid is available through `long_range` and `lat_range` if the Netcdf file has finer resolution than needed.
+Subsetting longitude and latitude grid is available through \code{long_range} and \code{lat_range} if the Netcdf file has finer resolution than needed.
 
 ```{r eval = FALSE, echo = TRUE}
-# Assume `my_ncdf` has bounding box of [-180, -90, 180, -90] 
+# Assume \code{my_ncdf} has bounding box of [-180, -90, 180, -90] 
 # at 0.25 degree resolution and subset it to have 
@@
 ERA5 data [@hersbach2020era5] is the latest reanalysis of global atmosphere, land surface, and ocean waves from 1950 onwards and is available in the NetCDF format.
-It can be directly downloaded from the [Copernicus Climate Data Store (CDS)](https://cds.climate.copernicus.eu/cdsapp#!/dataset/reanalysis-era5-pressure-levels?tab=overview) website or programmatically via R package `ecmwfr`[@ecwmfr].
-Variable Specific humidity and geopotential are queried on the 10 hPa pressure level for four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04 adn once downloaded, the data can be read into a cubble using the code above.
+It can be directly downloaded from the [Copernicus Climate Data Store (CDS)](https://cds.climate.copernicus.eu/cdsapp#!/dataset/reanalysis-era5-pressure-levels?tab=overview) website or programmatically via R package \pkg{ecmwfr} [@ecwmfr].
+Variable Specific humidity and geopotential are queried on the 10 hPa pressure level for four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04 and once downloaded, the data can be read into a cubble using the code above.
 Readers who are interested in the details of can refer to @simmons2020global and @simmons2005ecmwf.

### 2022-02-06 4-5-interactive.Rmd
all the  -> \code{CODE} as per jss  guideline

@@
 Combining this two types of plot with interactivity let users to link between points on the map and their corresponding time series and explore the spatial and temporal dimension of the data simultaneously.
-Below is an example that describes the process of building an interactive graphic with `cubble` and `crosstalk`.
+Below is an example that describes the process of building an interactive graphic with \pkg{cubble} and \pkg{crosstalk{.
 
 Starting with the original data, some pre-processing may be required to summarise the data before the visualisation.
-This example explores the variation of monthly temperature range across `r nrow(weatherdata::climate_full)` weather stations in Australia and with `weatherdata::climate_full`, daily temperature range is first calculated as the difference between `tmax` and `tmin`.
+This example explores the variation of monthly temperature range across `r nrow(weatherdata::climate_full)` weather stations in Australia and with `weatherdata::climate_full`, daily temperature range is first calculated as the difference between \code{tmax} and \code{tmin}.
 The three daily variables are then averaged into month over 2016 - 2020 in the long form.
 Variance of the temperature difference is then calculated for each station in the nested form.
-Then, a `SharedData` object is constructed for each form of the cubble and the same `group` argument ensures the cross-linking of the two forms via the common `id` column.
-The spatial map and time series plot are then made with each `SharedData` objects separately.
+Then, a \code{SharedData} object is constructed for each form of the cubble and the same \code{group} argument ensures the cross-linking of the two forms via the common \code{id} column.
+The spatial map and time series plot are then made with each \code{SharedData} objects separately.
 In this example, stations on the Australia map, made from the nested form, are coloured by the calculated variance and a ribbon band is constructed using the long form cubble to show the maximum and minimum temperature of each station across month.
 With a different dataset, users are free to calculate any per station measure in the nested form or to make any time-wise summary of the data in the long form to customise the spatial or temporal view.
-And the cross-linking is always safeguarded by the shared `id` column embedded in the cubble structure.
+And the cross-linking is always safeguarded by the shared \code{id} column embedded in the cubble structure.
 Below is the pseudo code that outlines the process to construct an interactive graphic described above:
@@
 With leaflet popup, the temperature range can be displayed as a small subplot upon clicking on the map.
-This would require first creating the popup plots from the long form cubble as a vector and then add these plots to a leaflet map, created from the nested cubble, with `leafpop::addPopupGraphs()`:
+This would require first creating the popup plots from the long form cubble as a vector and then add these plots to a leaflet map, created from the nested cubble, with \code{leafpop::addPopupGraphs()}:
 

### 2022-02-06 main.Rmd
fix marked red sentences labelled for myseyself
all the  -> \code{CODE} as per jss  guideline

@@
     address: 21 Chancellors Walk, Clayton VIC 3800 Australia
-    department: Department of Econometrics and Business Statistics
     email: \email{huize.zhang@monash.edu}
@@
 ```
-\newpage
 
@@
 
-\newpage
-
 ```{r child = "3-advanced.Rmd"}
@@
 ```
-
-\newpage
 

### 2022-02-06 main.tex
all the  -> \code{CODE} as per jss  guideline

@@
 climate observations widely applied in agriculture and ecology study.
-\newline
 
@@
 two dimensions to \pkg{cubble}. The software is available from the
-Comprehensive R Archive Network (CRAN) at {[}CRAN link{]}. \newline
+Comprehensive R Archive Network (CRAN) at {[}CRAN link{]}.
 
@@
 the data is in, these formats share some common components that
-characterise the spatio-temporal data. A spatial identifier (\texttt{id}
+characterise the spatio-temporal data. A spatial identifier (\code{id}
 in the diagram) is the unique identifier of each site. The temporal
-identifier (\texttt{t} in the diagram) prescribes the time stamp each
-site is recorded. Coordinates, comprising of latitude and longitude
-(\texttt{lon} and \texttt{lat} in the diagram), locates each site on the
+identifier (\code{t} in the diagram) prescribes the time stamp each site
+is recorded. Coordinates, comprising of latitude and longitude
+(\code{lat} and \code{(long)} in the diagram), locates each site on the
 map. These identifiers will be the building blocks for the data
@@
 hence should be performed with the nested form). Underneath the nested
-form, a cubble is built from a row-wise dataframe (\texttt{rowwise\_df})
+form, a cubble is built from a row-wise dataframe (\code{rowwise_df})
 where each site forms a separate group. This structure simplifies the
 calculation that involves temporal variables by avoiding the use of
-\texttt{map} syntax when working with list-column.
+\code{map} syntax when working with list-column.
 
@@
 
-\texttt{as\_cubble()} creates a \pkg{cubble} by supplying the three key
-components: \texttt{key} as the spatial identifier; \texttt{index} as
-the temporal identifier; and a vector of \texttt{coords} in the order of
-longitude first and then latitude. The naming of \texttt{key} and
-\texttt{index} follows the convention in the \pkg{tsibble} package. The
+`as\_cubble()\} creates a \pkg{cubble} by supplying the three key
+components: \code{key} as the spatial identifier; \code{index} as the
+temporal identifier; and a vector of \code{coords} in the order of
+longitude first and then latitude. The naming of \code{key} and
+\code{index} follows the convention in the \pkg{tsibble} package. The
 cubble created by default is in the nested form. Below is an example of
-creating a cubble: \newline
+creating a cubble:
 
@@
 There are a few information in the \pkg{cubble} header: the name of the
-\texttt{key} variable (\texttt{id}) and its number (5), bounding box
+\code{key} variable (\texttt{id}) and its number (5), bounding box
 (\texttt{{[}115.97,\ -32.94,\ 133.55,\ -12.42{]}}), and also the name of
-variable nested in the \texttt{ts} column with its type
-(\texttt{date\ {[}date{]},\ prcp\ {[}dbl{]},\ tmax\ {[}dbl{]},\ tmin\ {[}dbl{]}}).
+variable nested in the \code{ts} column with its type:
+\code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.
 
@@
 The long cubble is suitable to manipulate the time dimension of the
-data. The function \texttt{stretch()} switches the nested cubble into
-the long cubble by first extracts all the spatial variables into a
-separate tibble and store in the \texttt{spatial} attribute and then
-unnests the \texttt{ts} column:
+data. The function \code{stretch()} switches the nested cubble into the
+long cubble by first extracts all the spatial variables into a separate
+tibble and store in the \code{spatial} attribute and then unnests the
+\code{ts} column:
 
@@
 type of spatial variables:
-\texttt{lat\ {[}dbl{]},\ long\ {[}dbl{]},\ elev\ {[}dbl{]},\ name\ {[}chr{]},\ wmo\_id\ {[}dbl{]}}.
+\code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]}.
 
@@
 process. Many times, analysts will need to go back and forth between the
-nested and long cubble. \texttt{tamp()} inverses the \texttt{stretch()}
-that introduced in the previous section, to switch from a long cubble to
-the nested cubble:
+nested and long cubble. \code{tamp()} inverses the \code{stretch()} that
+introduced in the previous section, to switch from a long cubble to the
+nested cubble:
 
@@
 as the time varying ones. In cubble, this can be done through
-\texttt{migrate()}, which moves the time invariant variables into the
-long cubble:
+\code{migrate()}, which moves the time invariant variables into the long
+cubble:
 
@@
 
-Building from an \texttt{tbl\_df} class, \texttt{cubble} has implemented
-methods for \texttt{dplyr} generics, which includes:
+Building from an \code{tbl_df} class, \pkg{cubble} has implemented
+methods for \pkg{dplyr} generics, which includes:
 
@@
 \pkg{cubble} is also compatible with \pkg{tsibble}. When creating a
-cubble from a tsibble, Only the \texttt{coords} argument need to be
-specified. In the created nested form, each element of the list-column
-\texttt{ts} is of \texttt{tbl\_ts} class so that existing operations on
-the tsibble class can still be applied:
+cubble from a tsibble, Only the \code{coords} argument need to be
+specified:
 
@@
 \end{CodeOutput}
+\end{CodeChunk}
+
+In the created nested form, each element of the list-column \code{ts} is
+of \code{tbl_ts} class so that existing operations on the tsibble class
+can still be applied:
+
+\begin{CodeChunk}
 \begin{CodeInput}
@@
 summarised variables in the cluster level can give a crisper picture of
-local areas. In cubble, \texttt{switch\_key()} can be used to create a
-new level of grouping of spatial location by specifying a clustering
+local areas. In cubble, \code{switch_key()} can be used to create a new
+level of grouping of spatial location by specifying a clustering
 variable. Figure \ref{fig:illu-hier} illustrates the relationship of
@@
 By specifying
-\texttt{cluster\_nested\ \textless{}-\ station\_nested\ \%\textgreater{}\%\ switch\_key(key\ =\ cluster)},
-the cubble re-defines the cubble key from \texttt{id} in
-\texttt{station\_nested} to \texttt{cluster} in
-\texttt{cluster\_nested}. All the spatial variables variant to
-\texttt{cluster} are now nested into a \texttt{.val} column and cluster
-level variables can be computed in the same fashion as station level
-variables in \texttt{station\_nested}. Following the same principal of
-\texttt{stretch}, the long form cluster level cubble expands the time
-series variables with both indices: \texttt{cluster} and \texttt{id}.
-All the cluster level variables are stored separately for recovering the
-nested from from the long form.
+\code{cluster_nested <- station_nested \%>\% switch_key(key = cluster)},
+the cubble re-defines the cubble key from \code{id} in
+\code{station_nested} to \code{cluster} in \code{cluster_nested}. All
+the spatial variables variant to \code{cluster} are now nested into a
+\code{.val} column and cluster level variables can be computed in the
+same fashion as station level variables in \code{station_nested}.
+Following the same principal of \code{stretch}. the long form cluster
+level cubble expands the time series variables with both indices:
+\code{cluster} and \code{id}. All the cluster level variables are stored
+separately for recovering the nested from from the long form.
 
@@
 
-\texttt{match\_sites()} from \texttt{cubble} matches data based on these
-two criteria. It first matches the two data sources spatially through
+`match\_sites()\} from \pkg{cubble} matches data based on these two
+criteria. It first matches the two data sources spatially through
 computing the pairwise distance based on latitude and longitude. Pairs
@@
 Figure \ref{fig:illu-matching} illustrate this temporal matching in more
-details. Given two series \texttt{A} and \texttt{a}, 3 peaks have been
+details. Given two series \code{A} and \code{a}. 3 peaks have been
 picked in each series. An interval, with default length of 5, is
-constructed for each peak in series \texttt{A} and the peaks in series
-\texttt{a} are tested against whether they fall into the any of the
+constructed for each peak in series \code{A} and the peaks in series
+\code{a} are tested against whether they fall into the any of the
 intervals. In this illustration, there are 2 matches for these two
-series. Several arguments are available in \texttt{match\_sites()} to
+series. Several arguments are available in \code{match_sites()} to
 fine-tune the matching:
@@
 \item
-  \texttt{spatial\_n\_keep}: the number of spatial match for each site
-  to keep
+  \code{spatial_n_keep}. the number of spatial match for each site to
+  keep
 \item
-  \texttt{spatial\_dist\_max}: the maximum distance allowed for a
-  matched pair
+  \code{spatial_dist_max}. the maximum distance allowed for a matched
+  pair
 \item
-  \texttt{temporal\_n\_highest}: the number of peak used - 3 in the
-  example above
+  \code{temporal_n_highest}. the number of peak used - 3 in the example
+  above
 \item
-  \texttt{temporal\_window}: the length of the interval - 5 in the
-  example above
+  \code{temporal_window}. the length of the interval - 5 in the example
+  above
 \item
-  \texttt{temporal\_min\_match}: the minimum number of matched peak for
-  a valid matched pair
+  \code{temporal_min_match}. the minimum number of matched peak for a
+  valid matched pair
 \end{itemize}
@@
 communicate to the long cubble to activate all the observations with the
-same \texttt{id} (middle). The long cubble will then highlight the
+same \code{id} (middle). The long cubble will then highlight the
 activated series in the time series plot (right).
@@
 time series and once a point is selected, it will be activated in the
-long cubble. All the observations that share the same \texttt{id} are
-then activated and this includes other points in the same time series in
-the long cubble and the corresponding observation of site in the nested
+long cubble. All the observations that share the same \code{id} are then
+activated and this includes other points in the same time series in the
+long cubble and the corresponding observation of site in the nested
 cubble. These activated observations will then being reflected in the
@@
 Glyph map \citep{Wickham2012-yr} plots the time series as single glyph
-on the map and in \proglang{R}, \texttt{GGally} implements the glyph map
-through the \texttt{glyphs()} function. The function construct a data
-frame with calculated position (\texttt{gx}, \texttt{gy}, \texttt{gid})
-of each point on the time series using linear algebra (Equation 1 and 2
-in \citet{Wickham2012-yr}). The data can then be piped into
-\texttt{ggplot} to create the glyph map as:
+on the map and in \proglang{R}, \pkg{GGally} implements the glyph map
+through the \code{glyphs()} function. The function construct a data
+frame with calculated position (`gx\}. \code{gy}. \code{gid}. of each
+point on the time series using linear algebra (Equation 1 and 2 in
+\citet{Wickham2012-yr}). The data can then be piped into \code{ggplot}
+to create the glyph map as:
 
@@
 
-A re-implementation of the glyph map as a ggproto \texttt{GeomGlyph} has
-been made in the \texttt{cubble} package so that the glyph map can be
-created with \texttt{geom\_glyph()}:
+A re-implementation of the glyph map as a ggproto \code{GeomGlyph} has
+been made in the \pkg{cubble} package so that the glyph map can be
+created with \code{geom_glyph()}:
 
@@
 
-Polar glyph map can be specify as a parameter, \texttt{polar\ =\ TRUE},
-in the \texttt{geom\_glyph()}, along with \texttt{width} and
-\texttt{height} in either absolute or relative value. Global and local
-scale can be controlled by the parameter \texttt{global\_rescale}, which
-default to \texttt{TRUE} for global scaling. Reference box and line can
-be added with separate \texttt{geom\_glyph\_box()} and
-\texttt{geom\_glyph\_line()}.
+Polar glyph map can be specify as a parameter, \code{polar = TRUE}. in
+the \code{geom_glyph()}. along with \code{width} and \code{height} in
+either absolute or relative value. Global and local scale can be
+controlled by the parameter \code{global_rescale}. which default to
+\code{TRUE} for global scaling. Reference box and line can be added with
+separate \code{geom_glyph_box()} and
+\textbackslash code\{geom\_glyph\_line()`.
 
@@
 measures from stations across the world and the dataset
-\texttt{weatherdata::historical\_tmax} extracts the maximum temperature
-for 236 Australian stations. All the stations start recording from year
-1969 with the longest back to year 1859. The data
-\texttt{historical\_tmax} is already presented as a cubble, with
-\texttt{id} as the key, \texttt{date} as the index, and
-\texttt{c(longitude,\ latitude)} as the coordinates. This example
-compares the maximum temperature in two periods: 1971 - 1975 and 2016 -
-2020 for stations in Victoria and New South Wales.
+\code{weatherdata::historical_tmax} extracts the maximum temperature for
+236 Australian stations. All the stations start recording from year 1969
+with the longest back to year 1859. The data \code{historical_tmax} is
+already presented as a cubble, with \code{id} as the key, \code{date} as
+the index, and \code{c(longitude, latitude)} as the coordinates. This
+example compares the maximum temperature in two periods: 1971 - 1975 and
+2016 - 2020 for stations in Victoria and New South Wales.
 
@@
 nested cubble needs to be switched to the long cubble by
-\texttt{stretch()}:
+\code{stretch()}:
 
@@
 series has a total of 24 months. The long cubble needs to be switched to
-the nested form for this spatial operation using \texttt{tamp()}:
+the nested form for this spatial operation using \code{tamp()}:
 
@@
 
-Lastly, to create a glyph map, both the major (\texttt{longitude},
-\texttt{latitude}) and minor (\texttt{month}, \texttt{tmax}) coordinates
-need to be in the same table. Spatial variables can be moved to the long
-form with \texttt{migrate()}:
+Lastly, to create a glyph map, both the major (\code{longitude},
+\code{latitude}) and minor (\code{month}, \code{tmax}) coordinates need
+to be in the same table. Spatial variables can be moved to the long form
+with \code{migrate()}:
 
@@
 
-\texttt{tmax} can then be supplied to \texttt{geom\_glyph()} for the
-glyph map in Figure \ref{fig:basic-manip} with a station inset on the
-top left corner:
+\code{tmax} can then be supplied to \code{geom_glyph()} for the glyph
+map in Figure \ref{fig:basic-manip} with a station inset on the top left
+corner:
 
@@
 This example shows how to organise data at both level with
-\texttt{switch\_key()}. The dataset used is
-\texttt{weatherdata::climate\_full}, which records daily precipitation
-and maximum/ minimum temperature for 639 stations in Australia from 2016
+\code{switch_key()}. The dataset used is
+\code{weatherdata::climate_full}, which records daily precipitation and
+maximum/ minimum temperature for
+\code{r nrow(weatherdata::climate_full)} stations in Australia from 2016
 to 2020. A simple kmean algorithm based on the distance matrix is used
-here to create 20 clusters. This creates \texttt{station\_nested} as a
+here to create 20 clusters. This creates \code{station_nested} as a
 station level nested cubble with a cluster column indicating the group
@@
 
-To create a group level cubble, use \texttt{switch\_key()} with the new
-key variable, \texttt{cluster}:
+To create a group level cubble, use \code{switch_key()} with the new key
+variable, \code{cluster}:
 
@@
 
-With the group level cubble, \texttt{get\_centroid()} is useful to
-compute the centroid of each cluster, which will be used as the major
-axis for the glyph map later:
+With the group level cubble, \code{get_centroid()} is useful to compute
+the centroid of each cluster, which will be used as the major axis for
+the glyph map later:
 
@@
 reflect the precipitation measured by the climate station. As introduced
-in Section 3.2, \texttt{match\_sites()} can be used to match one source
-of data with another source in a cubble. Here \texttt{prcp} in
-\texttt{climate} will be matched to \texttt{Water\_course\_level} in
-\texttt{river}, which is specified in the argument
-\texttt{temporal\_var\_to\_match}. \texttt{temporal\_independent}
-controls the dataset that will be used to construct the interval (see
-the illustration in section 3.2 for details on this). Here the goal is
-to see if precipitation will be reflected by the water level in the
-river and hence a lagged effect of water level on precipitation. This
-will put precipitation data, \texttt{climate}, as the independent. Given
-there is one year worth of data, the number of peak
-(\texttt{temporal\_n\_highest}) to consider is slightly raised from a
-default 20 to 30. \texttt{temporal\_min\_match} filters out pairs don't
-have enough match and to return all the pairs, set
-\texttt{temporal\_min\_match} to \texttt{0}
+in Section 3.2, \code{match_sites()} can be used to match one source of
+data with another source in a cubble. Here \code{prcp} in \code{climate}
+will be matched to \code{Water_course_level} in \code{river}, which is
+specified in the argument \code{temporal_var_to_match}.
+\code{temporal_independent} controls the dataset that will be used to
+construct the interval (see the illustration in section 3.2 for details
+on this). Here the goal is to see if precipitation will be reflected by
+the water level in the river and hence a lagged effect of water level on
+precipitation. This will put precipitation data, \code{climate}, as the
+independent. Given there is one year worth of data, the number of peak
+(\code{temporal_n_highest}) to consider is slightly raised from a
+default 20 to 30. \code{temporal_min_match} filters out pairs don't have
+enough match and to return all the pairs, set \code{temporal_min_match}
+to 0.
 
@@
 The output from matching is also a cubble, with additional column
-\texttt{dist} and \texttt{group} produced from spatial matching and
-\texttt{n\_match} from the temporal matching.
+\code{dist} and \code{group} produced from spatial matching and
+\code{n_match} from the temporal matching.
 
@@
 packages on manipulating NetCDF data in R include a high-level R
-interface: \texttt{ncdf4}\citep{ncdf4}, a low-level interface that calls
-C-interface: \texttt{RNetCDF}\citep{rnetcdf, michna2013rnetcdf}, and a
-tidyverse implementation: \texttt{tidync}\citep{tidync}.
-
-Cubble provides an \texttt{as\_cubble()} method to coerce the
-\texttt{ncdf4} class from the \texttt{ncdf4} package into a
-\texttt{cubble}. It maps each combination of longitude and latitude into
-an \texttt{id} (the \texttt{key}) and nests time and variables in the
-nested form:
+interface: \pkg{ncdf4} \citep{ncdf4}, a low-level interface that calls
+C-interface: \pkg{RNetCDF} \citep{rnetcdf, michna2013rnetcdf}, and a
+tidyverse implementation: \pkg{tidync} \citep{tidync}.
+
+Cubble provides an \code{as_cubble()} method to coerce the \code{ncdf4}
+class from the \pkg{ncdf4} package into a \code{cubble}. It maps each
+combination of longitude and latitude into an \code{id} (the \code{key})
+and nests time and variables in the nested form:
 
@@
 Subsetting longitude and latitude grid is available through
-\texttt{long\_range} and \texttt{lat\_range} if the Netcdf file has
-finer resolution than needed.
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> # Assume `my_ncdf` has bounding box of [-180, -90, 180, -90] 
+\code{long_range} and \code{lat_range} if the Netcdf file has finer
+resolution than needed.
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> # Assume \code{my_ncdf} has bounding box of [-180, -90, 180, -90] 
 R> # at 0.25 degree resolution and subset it to have 
@@
 Climate Data Store (CDS)} website or programmatically via R package
-\texttt{ecmwfr}\citep{ecwmfr}. Variable Specific humidity and
-geopotential are queried on the 10 hPa pressure level for four dates:
-2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04 adn once downloaded,
-the data can be read into a cubble using the code above. Readers who are
-interested in the details of can refer to \citet{simmons2020global} and
+\pkg{ecmwfr} \citep{ecwmfr}. Variable Specific humidity and geopotential
+are queried on the 10 hPa pressure level for four dates: 2002-09-22,
+2002-09-26, 2002-09-30, and 2002-10-04 and once downloaded, the data can
+be read into a cubble using the code above. Readers who are interested
+in the details of can refer to \citet{simmons2020global} and
 \citet{simmons2005ecmwf}.
@@
 of the data simultaneously. Below is an example that describes the
-process of building an interactive graphic with \texttt{cubble} and
-\texttt{crosstalk}.
+process of building an interactive graphic with \pkg{cubble} and
+\textbackslash pkg\{crosstalk\{.
 
@@
 temperature range is first calculated as the difference between
-\texttt{tmax} and \texttt{tmin}. The three daily variables are then
-averaged into month over 2016 - 2020 in the long form. Variance of the
+\code{tmax} and \code{tmin}. The three daily variables are then averaged
+into month over 2016 - 2020 in the long form. Variance of the
 temperature difference is then calculated for each station in the nested
-form. Then, a \texttt{SharedData} object is constructed for each form of
-the cubble and the same \texttt{group} argument ensures the
-cross-linking of the two forms via the common \texttt{id} column. The
-spatial map and time series plot are then made with each
-\texttt{SharedData} objects separately. In this example, stations on the
-Australia map, made from the nested form, are coloured by the calculated
-variance and a ribbon band is constructed using the long form cubble to
-show the maximum and minimum temperature of each station across month.
-With a different dataset, users are free to calculate any per station
-measure in the nested form or to make any time-wise summary of the data
-in the long form to customise the spatial or temporal view. And the
-cross-linking is always safeguarded by the shared \texttt{id} column
-embedded in the cubble structure. Below is the pseudo code that outlines
-the process to construct an interactive graphic described above:
+form. Then, a \code{SharedData} object is constructed for each form of
+the cubble and the same \code{group} argument ensures the cross-linking
+of the two forms via the common \code{id} column. The spatial map and
+time series plot are then made with each \code{SharedData} objects
+separately. In this example, stations on the Australia map, made from
+the nested form, are coloured by the calculated variance and a ribbon
+band is constructed using the long form cubble to show the maximum and
+minimum temperature of each station across month. With a different
+dataset, users are free to calculate any per station measure in the
+nested form or to make any time-wise summary of the data in the long
+form to customise the spatial or temporal view. And the cross-linking is
+always safeguarded by the shared \code{id} column embedded in the cubble
+structure. Below is the pseudo code that outlines the process to
+construct an interactive graphic described above:
 
@@
 plots to a leaflet map, created from the nested cubble, with
-\texttt{leafpop::addPopupGraphs()}:
+\code{leafpop::addPopupGraphs()}:
 
@@
 Nicolas Langrené was affiliated with CSIRO's Data61. The article is
-created using \texttt{knitr} \citep{knitr} and \texttt{rmarkdown}
+created using \pkg{knitr} \citep{knitr} and \pkg{rmarkdown}
 \citep{rmarkdown} in R. The source code for reproducing this paper can

### 2022-02-07 main.Rmd
abstract

@@
 abstract: >
-  Sptio-temporal data naturally has two levels of observations: site and time level.
-[problem].
-In this paper, a new data structure, `cubble`, is proposed for manipulating and visualising spatio-temporal data.
+  Spatio-temporal variables can be time invariant or not and this divides them into those can be identified only with the spatial identifier and those cross-identified by the spatial and temporal identifier.
+In this paper, a new data structure, \code{cubble}, is proposed for manipulating and visualising spatio-temporal data.
+The new data structure proposes two forms for manipulating the spatial and temporal dimension separately while providing a smooth switch between the two forms to link the two dimensions.
+Advanced considerations are given to illustrate data with hierarchical structure, data matching, and how \pkg{cubble} fits with interactive graphics pipeline.
+Examples are given to analysing Australia meteorology data, river level data, and climate reanalysis (ERA5).
 keywords:

### 2022-02-07 main.tex
abstract

@@
 \Abstract{
-Sptio-temporal data naturally has two levels of observations: site and
-time level. {[}problem{]}. In this paper, a new data structure,
-\texttt{cubble}, is proposed for manipulating and visualising
-spatio-temporal data.
+Spatio-temporal variables can be time invariant or not and this divides
+them into those can be identified only with the spatial identifier and
+those cross-identified by the spatial and temporal identifier. In this
+paper, a new data structure, \code{cubble}, is proposed for manipulating
+and visualising spatio-temporal data. The new data structure proposes
+two forms for manipulating the spatial and temporal dimension separately
+while providing a smooth switch between the two forms to link the two
+dimensions. Advanced considerations are given to illustrate data with
+hierarchical structure, data matching, and how \pkg{cubble} fits with
+interactive graphics pipeline. Examples are given to analysing Australia
+meteorology data, river level data, and climate reanalysis (ERA5).
 }

### 2022-02-07 2-the-cubble-pkg.Rmd
add some sentences on data cube and fix \code{...} rendering

@@
 Spatio-temporal data usually come in various forms and Figure \ref{fig:illu-input} shows four examples of this.
-No matter which form the data is in, these formats share some common components that characterise the spatio-temporal data.
-A spatial identifier (\code{id} in the diagram) is the unique identifier of each site.
-The temporal identifier (\code{t} in the diagram) prescribes the time stamp each site is recorded.
-Coordinates, comprising of latitude and longitude (\code{lat} and \code{(long)} in the diagram), locates each site on the map.
-These identifiers will be the building blocks for the data structure introduced below.
-Other variables in the data can be categorised into two groups: spatial variables that are invariant at each time stamp for every site, i.e.
-the name or code of the weather station, and temporal variables that varies with time.
+These four examples show some common characters of the spatio-temporal data: 1) a spatial identifier, that uniquely identifies each site, 2) a temporal identifier, that prescribes the time stamp each site is recorded, 3) Coordinates, comprising of latitude and longitude, that locate each site on the map.
+These identifiers will be the building blocks for cubble.
+Spatio-temporal data can also be thought of as a data cube with three axes: id, time, and variable.
+Notice that in the time-variable face, time invariant variables share the same value across all the time stamps while time varying variables have values varying across time.
+This heterogeneity will be used to split the data into two forms in cubble.
 
@@
 
-`as_cubble()} creates a \pkg{cubble} by supplying the three key components: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order of longitude first and then latitude.
+\code{as_cubble()} creates a cubble by supplying the three key components: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order of longitude first and then latitude.
 The naming of \code{key} and \code{index} follows the convention in the \pkg{tsibble} package.
@@
 
-There are a few information in the \pkg{cubble} header: the name of the \code{key} variable (`id`) and its number (5), bounding box (`[115.97, -32.94, 133.55, -12.42]`), and also the name of variable nested in the \code{ts} column with its type: \code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.
+There are a few information in the cubble header: the name of the \code{key} variable (`id`) and its number (5), bounding box (`[115.97, -32.94, 133.55, -12.42]`), and also the name of variable nested in the \code{ts} column with its type: \code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.
 
@@
 
-Building from an \code{tbl_df} class, \pkg{cubble} has implemented methods for \pkg{dplyr} generics, which includes:
+Building from an \code{tbl_df} class, cubble has implemented methods for \pkg{dplyr} generics, which includes:
 

### 2022-02-07 3-advanced.Rmd
add some sentences on data cube and fix \code{...} rendering

@@
 
-`match_sites()} from \pkg{cubble} matches data based on these two criteria.
+\code{match_sites()} from \pkg{cubble} matches data based on these two criteria.
 It first matches the two data sources spatially through computing the pairwise distance based on latitude and longitude.
@@
 Glyph map [@Wickham2012-yr] plots the time series as single glyph on the map and in \proglang{R}, \pkg{GGally} implements the glyph map through the \code{glyphs()} function.
-The function construct a data frame with calculated position (`gx}.
+The function construct a data frame with calculated position (\code{gx}.
 \code{gy}.
@@
 which default to \code{TRUE} for global scaling.
-Reference box and line can be added with separate \code{geom_glyph_box()} and \code{geom_glyph_line()`.
+Reference box and line can be added with separate \code{geom_glyph_box()} and \code{geom_glyph_line()}.

### 2022-02-07 4-4-netcdf.Rmd
add some sentences on data cube and fix \code{...} rendering

@@
 ```{r eval = FALSE, echo = TRUE}
-# Assume \code{my_ncdf} has bounding box of [-180, -90, 180, -90] 
+# Assume my_ncdf has bounding box of [-180, -90, 180, -90] 
 # at 0.25 degree resolution and subset it to have 

### 2022-02-07 main.tex
add some sentences on data cube and fix \code{...} rendering

@@
 Spatio-temporal data usually come in various forms and Figure
-\ref{fig:illu-input} shows four examples of this. No matter which form
-the data is in, these formats share some common components that
-characterise the spatio-temporal data. A spatial identifier (\code{id}
-in the diagram) is the unique identifier of each site. The temporal
-identifier (\code{t} in the diagram) prescribes the time stamp each site
-is recorded. Coordinates, comprising of latitude and longitude
-(\code{lat} and \code{(long)} in the diagram), locates each site on the
-map. These identifiers will be the building blocks for the data
-structure introduced below. Other variables in the data can be
-categorised into two groups: spatial variables that are invariant at
-each time stamp for every site, i.e.~the name or code of the weather
-station, and temporal variables that varies with time.
+\ref{fig:illu-input} shows four examples of this. These four examples
+show some common characters of the spatio-temporal data: 1) a spatial
+identifier, that uniquely identifies each site, 2) a temporal
+identifier, that prescribes the time stamp each site is recorded, 3)
+Coordinates, comprising of latitude and longitude, that locate each site
+on the map. These identifiers will be the building blocks for cubble.
+Spatio-temporal data can also be thought of as a data cube with three
+axes: id, time, and variable. Notice that in the time-variable face,
+time invariant variables share the same value across all the time stamps
+while time varying variables have values varying across time. This
+heterogeneity will be used to split the data into two forms in cubble.
 
@@
 
-`as\_cubble()\} creates a \pkg{cubble} by supplying the three key
+\code{as_cubble()} creates a cubble by supplying the three key
 components: \code{key} as the spatial identifier; \code{index} as the
@@
 
-There are a few information in the \pkg{cubble} header: the name of the
+There are a few information in the cubble header: the name of the
 \code{key} variable (\texttt{id}) and its number (5), bounding box
@@
 
-Building from an \code{tbl_df} class, \pkg{cubble} has implemented
-methods for \pkg{dplyr} generics, which includes:
+Building from an \code{tbl_df} class, cubble has implemented methods for
+\pkg{dplyr} generics, which includes:
 
@@
 
-`match\_sites()\} from \pkg{cubble} matches data based on these two
+\code{match_sites()} from \pkg{cubble} matches data based on these two
 criteria. It first matches the two data sources spatially through
@@
 through the \code{glyphs()} function. The function construct a data
-frame with calculated position (`gx\}. \code{gy}. \code{gid}. of each
-point on the time series using linear algebra (Equation 1 and 2 in
+frame with calculated position (\code{gx}. \code{gy}. \code{gid}. of
+each point on the time series using linear algebra (Equation 1 and 2 in
 \citet{Wickham2012-yr}). The data can then be piped into \code{ggplot}
@@
 \code{TRUE} for global scaling. Reference box and line can be added with
-separate \code{geom_glyph_box()} and
-\textbackslash code\{geom\_glyph\_line()`.
+separate \code{geom_glyph_box()} and \code{geom_glyph_line()}.
 
@@
 \begin{CodeInput}
-R> # Assume \code{my_ncdf} has bounding box of [-180, -90, 180, -90] 
+R> # Assume my_ncdf has bounding box of [-180, -90, 180, -90] 
 R> # at 0.25 degree resolution and subset it to have 
@@
 process of building an interactive graphic with \pkg{cubble} and
-\textbackslash pkg\{crosstalk\{.
+\pkg{crosstalk}.
 

### 2022-02-07 2-the-cubble-pkg.Rmd
go through all the fig captions

@@
 
-```{r illu-input, echo = FALSE, out.width = "100%", fig.cap="Illustration of incoming data formats for spatio-temporal data.
-(1) Data comes in as a single table; (2) Separate tables for spatial and temporal variables; (3) A single table with all the parameters used to query the database and a separate table for queried data; and (4) Cubical data in array or NetCDF format.", fig.align='center', out.height = "40%"}
+```{r illu-input, echo = FALSE, out.width = "100%", fig.cap="An illustration of different spatio-temporal data format: (1) a single table for all the variables; (2) separate tables for spatial and temporal variables; (3) a table of metadata used to query the database and a separate table for queried data; and (4) an array data structure of cube or hypercube.", fig.align='center', out.height = "40%"}
 knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.001.png"))
@@
 
-```{r illu-cubble, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "Illustration of nested and long cubble."}
+```{r illu-cubble, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "An illustration of the nested and long form in cubble.
+The nested form defines each site in a row and nests the time varying variables into a single column \\code{ts}.
+The long form cubble uses \\code{id} and \\code{t} to identify each row and store the time invariant variables as an attribute, \\code{spatial}."}
 knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.002.png"))

### 2022-02-07 3-advanced.Rmd
go through all the fig captions

@@
 
-```{r illu-hier, echo = FALSE, fig.align="center",out.height="40%", out.width = "100%", fig.cap = "Relationship between the station and cluster level cubble in both long and nested form."}
+```{r illu-hier, echo = FALSE, fig.align="center",out.height="40%", out.width = "100%", fig.cap = "An illustration of the original and cluster level cubble in the nested form long form for hierarchical structure data.
+\\code{switch\\_key()} changes the station level cubble into a cluster level cubble and both can be stretched into the long form."}
 knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.003.png"))
@@
 ```{r illu-matching, out.width="100%", fig.cap="An illustration of temporal matching in cubble.
-Three highest peaks are identified in each series and intervals are constructed on series A.
-Two peaks in series a fall into the intervals and hence the two series are considered to have 2 matches."}
+Three highest peaks are identified in each series and intervals are constructed on series \\code{A}.
+Two peaks in series \\code{a} fall into the intervals and hence the two series are considered to have 2 matches."}
 knitr::include_graphics("figures/illu-matching.png")
@@
 ```{r illu-interactive, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "An illustration of the data model under interactive graphics with cubble.
-The line plot and map is made separately with the long and nested cubble.
+The line plot and the map is made separately with the long and nested cubble.
 When a station is selected on the map (left), the corresponding row in the nested cubble will be activated.

### 2022-02-07 4-1-basic.Rmd
go through all the fig captions

@@
 
-```{r basic-manip, out.width="100%", out.height="70%", fig.cap="A glyph map of the mean maximum temperature by month for Victoria and New South Wales weather stations in Australia.
-On the top left corner is an insetted plot of station Cobar, which is highlighted in the black box."}
+```{r basic-manip, out.width="100%", out.height="70%", fig.cap="A glyph map of the average maximum temperature by month of Victoria and New South Wales weather stations in Australia.
+On the top left corner is an insetted plot of station Cobar highlighted in the black box."}
 knitr::include_graphics("figures/basic-manip.png")

### 2022-02-07 4-2-agg.Rmd
go through all the fig captions

@@
 This example shows how to organise data at both level with \code{switch_key()}.
-The dataset used is \code{weatherdata::climate_full}, which records daily precipitation and maximum/ minimum temperature for \code{r nrow(weatherdata::climate_full)} stations in Australia from 2016 to 2020.
+The dataset used is \code{weatherdata::climate_full}, which records daily precipitation and maximum/ minimum temperature for `r nrow(weatherdata::climate_full)` stations in Australia from 2016 to 2020.
 A simple kmean algorithm based on the distance matrix is used here to create 20 clusters.
@@
 Subplot A shows the glyph map of the weekly averaged precipitation of each cluster.
-The group number of printed in the middle of y minor axis and can be used as a reference line to read the magnitude of precipitation.
-Subplot B shows the station membership of each cluster"}
+The group number of printed in the middle of y minor axis and can be used as a reference line to read the magnitude.
+Subplot B shows the station membership of each cluster."}
 knitr::include_graphics("figures/basic-agg.png")

### 2022-02-07 4-3-water-level.Rmd
go through all the fig captions

@@
 
-```{r matching-map, out.width="100%", fig.cap="Location of climate station and river gauges in Victoria, Australia."}
+```{r matching-map, out.width="100%", fig.cap="Location of weather stations and river gauges in Victoria, Australia."}
 knitr::include_graphics("figures/matching-map.png")
@@
 Precipitation and water level have been standardised between 0 and 1 to be displayed on the same scale.
-The increases in the precipitation is reflected well by the water level."}
+The increases in the precipitation is reflected by the water level."}
 knitr::include_graphics("figures/matching.png")

### 2022-02-07 4-4-netcdf.Rmd
go through all the fig captions

@@
 NetCDF (Network Common Data Form) is a data format commonly used in climatology community to deliver global mapping of atmosphere, ocean, and land.
-It has two major components: dimension to define the spatiotemporal grid (longitude, latitude, and time) and variable which populates the defined grid.
+It has two major components: dimension to define the spatio-temporal grid (longitude, latitude, and time) and variable which populates the defined grid.
 Attributes are usually associated with dimension and variable in the NetCDF format data and a [metadata convention for climate and forecast](http://cfconventions.org/) has been designed to standardise this data format.
@@
 
-```{r netcdf, out.width="100%", fig.cap = "A reproduction of the second row of Figure 19 in  Hersbach (2020)."}
+```{r netcdf, out.width="100%", fig.cap = "A reproduction of the second row (ERA5 data) of Figure 19 in  Hersbach et al (2020)."}
 knitr::include_graphics(here::here("figures/netcdf.png"))

### 2022-02-07 4-5-interactive.Rmd
add some sentences on data cube and fix \code{...} rendering
go through all the fig captions

@@
 Combining this two types of plot with interactivity let users to link between points on the map and their corresponding time series and explore the spatial and temporal dimension of the data simultaneously.
-Below is an example that describes the process of building an interactive graphic with \pkg{cubble} and \pkg{crosstalk{.
+Below is an example that describes the process of building an interactive graphic with \pkg{cubble} and \pkg{crosstalk}.
 
@@
 The bottom row first selects the lowest temperature in Auguest in the seasonal display.
-A location in the Tasmania Island is then selected and surprisingly, it is actually warmer than the Thredbo AWS.", fig.show='hold'}
+A location in the Tasmania Island is then selected to compare the temperature variation with Thredbo AWS.", fig.show='hold'}
 knitr::include_graphics(here::here("figures/linking.png"))
@@
 
-```{r interactive-popup, echo = FALSE, out.width="45%", out.height="25%", fig.retina = 2, dpi = 300, fig.cap = "Screenshots of variance of monthly temperature range from 2016 to 2020 in Australia.
-Upon clicking a single station on the leaftlet up, the temperature band will be shown as a subplot in the popup box.", fig.show='hold'}
+```{r interactive-popup, echo = FALSE, out.width="45%", out.height="25%", fig.retina = 2, dpi = 300, fig.cap = "Same as FIgure 11 with thetemperature variation shown as a popup in the leaflet map.", fig.show='hold'}
 knitr::include_graphics(here::here("figures/popup-mid.png")) 

### 2022-02-07 5-conclusion.Rmd
all the  -> \code{CODE} as per jss  guideline
go through all the fig captions

--- 2022-02-05 5-conclusion.Rmd
@@
 
+This paper describes the \proglang{R} package \pkg{cubble} for manipulating and visualising spatio-temporal data.
+A new data structure, that builds from the \code{rowwise_df} and \code{grouped_df} class in the tidyverse ecosystem, is proposed to connect the spatial and temporal variables in the spatio-temporal data.
+This design frees the data analysts from spending time on organising variables of different observational units.
+The data structure is also flexible to the techniques and packages analysts use to analyse the data, for example, in the matching example in section 4.3, users are free to use algorithms from another pacakge to cluster stations.
+Further development and maintenance of the package involves responding to changes in the tidyverse packages that \pkg{cubble} imports, in particular, \pkg{tibble}, \pkg{tidyr}, and \pkg{dplyr}.
+Another area for further development is to extend \pkg{cubble} to other spatial objects other than points.
+
 # Acknowledgement
@@
 This work is funded by the Commonwealth Scientific and Industrial Research Organisation (CSIRO) Data61 Scholarship and started while Nicolas Langrené was affiliated with CSIRO's Data61.
-The article is created using `knitr` [@knitr] and `rmarkdown` [@rmarkdown] in R.
-The source code for reproducing this paper can be found [here](https://github.com/huizezhang-sherry/paper-cubble).
+The article is created using \pkg{knitr} [@knitr] and \pkg{rmarkdown} [@rmarkdown] in R.
+The source code for reproducing this paper can be found at: https://github.com/huizezhang-sherry/paper-cubble.
 
@@
 
-```{r illu-interactive-2, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "demon interactivity"}
+```{r illu-interactive-2, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "An illustration of the data model under interactive graphics with cubble.
+When a point on the time series is selected, the corresponding row in the long cubble will be activated.
+This will link to all the rows with the same id in the long cubble and the row in the nested cubble with the smae id (middle).
+Both plots will be updated with the full line selected and the point highlighted on the map (right)."}
 knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.005.png"))

### 2022-02-07 main.tex
go through all the fig captions

@@
 
-\caption[Illustration of incoming data formats for spatio-temporal data]{Illustration of incoming data formats for spatio-temporal data.
-(1) Data comes in as a single table; (2) Separate tables for spatial and temporal variables; (3) A single table with all the parameters used to query the database and a separate table for queried data; and (4) Cubical data in array or NetCDF format.}\label{fig:illu-input}
+\caption[An illustration of different spatio-temporal data format]{An illustration of different spatio-temporal data format: (1) a single table for all the variables; (2) separate tables for spatial and temporal variables; (3) a table of metadata used to query the database and a separate table for queried data; and (4) an array data structure of cube or hypercube.}\label{fig:illu-input}
 \end{figure}
@@
 
-\caption[Illustration of nested and long cubble]{Illustration of nested and long cubble.}\label{fig:illu-cubble}
+\caption{An illustration of the nested and long form in cubble.
+The nested form defines each site in a row and nests the time varying variables into a single column \code{ts}.
+The long form cubble uses \code{id} and \code{t} to identify each row and store the time invariant variables as an attribute, \code{spatial}.}\label{fig:illu-cubble}
 \end{figure}
@@
 
-\caption[Relationship between the station and cluster level cubble in both long and nested form]{Relationship between the station and cluster level cubble in both long and nested form.}\label{fig:illu-hier}
+\caption[An illustration of the original and cluster level cubble in the nested form long form for hierarchical structure data]{An illustration of the original and cluster level cubble in the nested form long form for hierarchical structure data.
+\code{switch\_key()} changes the station level cubble into a cluster level cubble and both can be stretched into the long form.}\label{fig:illu-hier}
 \end{figure}
@@
 
-\caption[An illustration of temporal matching in cubble]{An illustration of temporal matching in cubble.
-Three highest peaks are identified in each series and intervals are constructed on series A.
-Two peaks in series a fall into the intervals and hence the two series are considered to have 2 matches.}\label{fig:illu-matching}
+\caption{An illustration of temporal matching in cubble.
+Three highest peaks are identified in each series and intervals are constructed on series \code{A}.
+Two peaks in series \code{a} fall into the intervals and hence the two series are considered to have 2 matches.}\label{fig:illu-matching}
 \end{figure}
@@
 \caption[An illustration of the data model under interactive graphics with cubble]{An illustration of the data model under interactive graphics with cubble.
-The line plot and map is made separately with the long and nested cubble.
+The line plot and the map is made separately with the long and nested cubble.
 When a station is selected on the map (left), the corresponding row in the nested cubble will be activated.
@@
 
-\caption[A glyph map of the mean maximum temperature by month for Victoria and New South Wales weather stations in Australia]{A glyph map of the mean maximum temperature by month for Victoria and New South Wales weather stations in Australia.
-On the top left corner is an insetted plot of station Cobar, which is highlighted in the black box.}\label{fig:basic-manip}
+\caption[A glyph map of the average maximum temperature by month of Victoria and New South Wales weather stations in Australia]{A glyph map of the average maximum temperature by month of Victoria and New South Wales weather stations in Australia.
+On the top left corner is an insetted plot of station Cobar highlighted in the black box.}\label{fig:basic-manip}
 \end{figure}
@@
 \code{weatherdata::climate_full}, which records daily precipitation and
-maximum/ minimum temperature for
-\code{r nrow(weatherdata::climate_full)} stations in Australia from 2016
-to 2020. A simple kmean algorithm based on the distance matrix is used
-here to create 20 clusters. This creates \code{station_nested} as a
-station level nested cubble with a cluster column indicating the group
-each station belongs to. More complex algorithms can be used for other
+maximum/ minimum temperature for 639 stations in Australia from 2016 to
+2020. A simple kmean algorithm based on the distance matrix is used here
+to create 20 clusters. This creates \code{station_nested} as a station
+level nested cubble with a cluster column indicating the group each
+station belongs to. More complex algorithms can be used for other
 problem, as long as there is a mapping from each station to a cluster.
@@
 Subplot A shows the glyph map of the weekly averaged precipitation of each cluster.
-The group number of printed in the middle of y minor axis and can be used as a reference line to read the magnitude of precipitation.
-Subplot B shows the station membership of each cluster}\label{fig:basic-agg}
+The group number of printed in the middle of y minor axis and can be used as a reference line to read the magnitude.
+Subplot B shows the station membership of each cluster.}\label{fig:basic-agg}
 \end{figure}
@@
 
-\caption[Location of climate station and river gauges in Victoria, Australia]{Location of climate station and river gauges in Victoria, Australia.}\label{fig:matching-map}
+\caption[Location of weather stations and river gauges in Victoria, Australia]{Location of weather stations and river gauges in Victoria, Australia.}\label{fig:matching-map}
 \end{figure}
@@
 Precipitation and water level have been standardised between 0 and 1 to be displayed on the same scale.
-The increases in the precipitation is reflected well by the water level.}\label{fig:matching}
+The increases in the precipitation is reflected by the water level.}\label{fig:matching}
 \end{figure}
@@
 and land. It has two major components: dimension to define the
-spatiotemporal grid (longitude, latitude, and time) and variable which
+spatio-temporal grid (longitude, latitude, and time) and variable which
 populates the defined grid. Attributes are usually associated with
@@
 
-\caption[A reproduction of the second row of Figure 19 in  Hersbach (2020)]{A reproduction of the second row of Figure 19 in  Hersbach (2020).}\label{fig:netcdf}
+\caption[A reproduction of the second row (ERA5 data) of Figure 19 in  Hersbach et al (2020)]{A reproduction of the second row (ERA5 data) of Figure 19 in  Hersbach et al (2020).}\label{fig:netcdf}
 \end{figure}
@@
 The bottom row first selects the lowest temperature in Auguest in the seasonal display.
-A location in the Tasmania Island is then selected and surprisingly, it is actually warmer than the Thredbo AWS.}\label{fig:interactive-linking}
+A location in the Tasmania Island is then selected to compare the temperature variation with Thredbo AWS.}\label{fig:interactive-linking}
 \end{figure}
@@
 
-\caption[Screenshots of variance of monthly temperature range from 2016 to 2020 in Australia]{Screenshots of variance of monthly temperature range from 2016 to 2020 in Australia.
-Upon clicking a single station on the leaftlet up, the temperature band will be shown as a subplot in the popup box.}\label{fig:interactive-popup}
+\caption[Same as FIgure 11 with thetemperature variation shown as a popup in the leaflet map]{Same as FIgure 11 with thetemperature variation shown as a popup in the leaflet map.}\label{fig:interactive-popup}
 \end{figure}
@@
 
-This paper describes the \proglang{R} package \texttt{cubble} for
+This paper describes the \proglang{R} package \pkg{cubble} for
 manipulating and visualising spatio-temporal data. A new data structure,
-that builds from the \texttt{rowwise\_df} and \texttt{grouped\_df} class
-in the tidyverse ecosystem, is proposed to connect the spatial and
-temporal variables in the spatio-temporal data. This design frees the
-data analysts from spending time on organising variables of different
+that builds from the \code{rowwise_df} and \code{grouped_df} class in
+the tidyverse ecosystem, is proposed to connect the spatial and temporal
+variables in the spatio-temporal data. This design frees the data
+analysts from spending time on organising variables of different
 observational units. The data structure is also flexible to the
@@
 maintenance of the package involves responding to changes in the
-tidyverse packages that \texttt{cubble} imports, in particular,
-\texttt{tibble}, \texttt{tidyr}, and \texttt{dplyr}. Another area for
-further development is to extend \texttt{cubble} to other spatial
-objects other than points.
+tidyverse packages that \pkg{cubble} imports, in particular,
+\pkg{tibble}, \pkg{tidyr}, and \pkg{dplyr}. Another area for further
+development is to extend \pkg{cubble} to other spatial objects other
+than points.
 
@@
 
-\caption[demon interactivity]{demon interactivity}\label{fig:illu-interactive-2}
+\caption[An illustration of the data model under interactive graphics with cubble]{An illustration of the data model under interactive graphics with cubble.
+When a point on the time series is selected, the corresponding row in the long cubble will be activated.
+This will link to all the rows with the same id in the long cubble and the row in the nested cubble with the smae id (middle).
+Both plots will be updated with the full line selected and the point highlighted on the map (right).}\label{fig:illu-interactive-2}
 \end{figure}

### 2022-02-07 1-intro.Rmd
all the  -> \code{CODE} as per jss  guideline
spelling

@@
 The data structure it uses is based on \pkg{sp} [@sp] and \pkg{xts} [@xts], both of which has been replaced by more recent implementations.
-\pkg{spatstat} [@spatstat] implements a `ppp` class for point pattern data; and more recent, \pkg{stars} [@stars] implements a spatio-temporal array with the dplyr's data cube structure \pkg{cubelyr} [@cubelyr] as its backend.
+\pkg{spatstat} [@spatstat] implements a `ppp` class for point pattern data; and more recent, \pkg{stars} [@stars] implements a spatio-temporal array with the dplyr's data cube structure \pkg{cubelyr} [@cubelyr] as its back-end.
 While these implementations either store spatial and temporal variables all in a single table, hence duplicate the spatial variables for each temporal unit; or split them into two separate tables that has the problem of manually joining, mentioned in the previously.
@@
 The requirement for such a tool is important given the ubiquity of spatio-temporal vector data in the wild: the Ireland wind data from \pkg{gstat} is an classic example data that splits variables into its spatial (`wind.loc`) and temporal (`wind`) dimension; Bureau of Meteorology (BoM) provides climate observations widely applied in agriculture and ecology study.
-\newline
 
@@
 The software is available from the Comprehensive R Archive Network (CRAN) at [CRAN link].
-\newline
 
@@
 Section 4 gives examples of the features introduced in the previous two sections with climate and hydrology data.
-An example on how cubble handle Netcdf data is also provided.
+An example on how cubble handle NetCDF data is also provided.
 Section 5 concludes the paper.

### 2022-02-07 3-advanced.Rmd
spelling

@@
 
-This functionality can be seen as a matching [@stuart2010matching; @mcintosh2018using] exercise in the spatio-temporal domain to pair up nearby observatoins.
+This functionality can be seen as a matching [@stuart2010matching; @mcintosh2018using] exercise in the spatio-temporal domain to pair up nearby observations.
 It can also be considered as a form of data fusion [@castanedo2013review; @cocchi2019data], where data from multiple sources are combined together.

### 2022-02-07 4-4-netcdf.Rmd
spelling

@@
 
-Memory limit with Netcdf data in cubble depends on longitude grid point x latitude grid point x time grid point x number of variable.
+Memory limit with NetCDF data in cubble depends on longitude grid point x latitude grid point x time grid point x number of variable.
 Cubble can handle slightly more than 300 x 300 (longitude x longitude) grid points for 3 variables in one year and spatial grid can be reduced to trade for longer time period and more variables.
 A 300 by 300 spatial grid can be a bounding box of [100, -80, 180, 0] at 0.25 degree resolution or global bounding box [-180, -90, 180, -90] at 1 degree resolution.
-Subsetting longitude and latitude grid is available through \code{long_range} and \code{lat_range} if the Netcdf file has finer resolution than needed.
+Subsetting longitude and latitude grid is available through \code{long_range} and \code{lat_range} if the NetCDF file has finer resolution than needed.
 

### 2022-02-07 5-conclusion.Rmd
spelling

@@
 This design frees the data analysts from spending time on organising variables of different observational units.
-The data structure is also flexible to the techniques and packages analysts use to analyse the data, for example, in the matching example in section 4.3, users are free to use algorithms from another pacakge to cluster stations.
+The data structure is also flexible to the techniques and packages analysts use to analyse the data, for example, in the matching example in section 4.3, users are free to use algorithms from another package to cluster stations.
 Further development and maintenance of the package involves responding to changes in the tidyverse packages that \pkg{cubble} imports, in particular, \pkg{tibble}, \pkg{tidyr}, and \pkg{dplyr}.

### 2022-02-07 main.tex
spelling

@@
 dplyr's data cube structure \pkg{cubelyr} \citep{cubelyr} as its
-backend. While these implementations either store spatial and temporal
+back-end. While these implementations either store spatial and temporal
 variables all in a single table, hence duplicate the spatial variables
@@
 features introduced in the previous two sections with climate and
-hydrology data. An example on how cubble handle Netcdf data is also
+hydrology data. An example on how cubble handle NetCDF data is also
 provided. Section 5 concludes the paper.
@@
 \citep{stuart2010matching, mcintosh2018using} exercise in the
-spatio-temporal domain to pair up nearby observatoins. It can also be
+spatio-temporal domain to pair up nearby observations. It can also be
 considered as a form of data fusion
@@
 
-Memory limit with Netcdf data in cubble depends on longitude grid point
+Memory limit with NetCDF data in cubble depends on longitude grid point
 x latitude grid point x time grid point x number of variable. Cubble can
@@
 Subsetting longitude and latitude grid is available through
-\code{long_range} and \code{lat_range} if the Netcdf file has finer
+\code{long_range} and \code{lat_range} if the NetCDF file has finer
 resolution than needed.
@@
 in the matching example in section 4.3, users are free to use algorithms
-from another pacakge to cluster stations. Further development and
+from another package to cluster stations. Further development and
 maintenance of the package involves responding to changes in the

### 2022-02-07 1-intro.Rmd
small changes in read-through

@@
 Spatio-temporal data concerns variables in the spatial coordinates across time.
-Some variables are invariant to time and can be solely identified by its spatial location while other time varying variables need both the spatial and temporal identifier to identify.
+Some variables are invariant to time and can be solely identified by its spatial location while other time varying variables need both the spatial and temporal identifiers to identify.
 Manipulating data in these two levels is usually done separately and the third tidy data principal [@tidydata] recommends each type of observational unit to form a table.
@@
 
-Currently, available spatio-temporal data structure in R includes: \pkg{spacetime} [@spacetime], which proposes four space-time layouts: Full grid (STF), sparse grid(STS), irregular (STI), and trajectory (STT).
-The data structure it uses is based on \pkg{sp} [@sp] and \pkg{xts} [@xts], both of which has been replaced by more recent implementations.
-\pkg{spatstat} [@spatstat] implements a `ppp` class for point pattern data; and more recent, \pkg{stars} [@stars] implements a spatio-temporal array with the dplyr's data cube structure \pkg{cubelyr} [@cubelyr] as its back-end.
-While these implementations either store spatial and temporal variables all in a single table, hence duplicate the spatial variables for each temporal unit; or split them into two separate tables that has the problem of manually joining, mentioned in the previously.
+Currently, available spatio-temporal data structures in R includes: \pkg{spacetime} [@spacetime], which proposes four space-time layouts: Full grid (STF), sparse grid(STS), irregular (STI), and trajectory (STT).
+\pkg{spacetime} is built on \pkg{sp} [@sp] and \pkg{xts} [@xts], both of which have been replaced by more recent implementations.
+\pkg{spatstat} [@spatstat] implements a `ppp` class for point pattern data; and more recent, \pkg{stars} [@stars] implements a spatio-temporal for vector and raster spatio-temporal data.
+While these implementations either store spatial and temporal variables all in a single table, hence duplicate the spatial variables at each temporal unit; or split them into two separate tables that analysts need to manually join the data constantly when doing the analysis.
 This creates a gap in the software development.
-The requirement for such a tool is important given the ubiquity of spatio-temporal vector data in the wild: the Ireland wind data from \pkg{gstat} is an classic example data that splits variables into its spatial (`wind.loc`) and temporal (`wind`) dimension; Bureau of Meteorology (BoM) provides climate observations widely applied in agriculture and ecology study.
+The requirement for such a tool is important given the ubiquity of spatio-temporal vector data in the wild: climate observations from weather stations, air quality variables from air monitoring stations, or any variables that are measured at fixed location across time.
 
 This paper describes the implementation of a new spatio-temporal data structure: \pkg{cubble}.
-\pkg{cubble} implements a relational data structure that uses two forms to manage the switch between spatial and temporal dimension.
+\pkg{Cubble} implements a data structure that uses two forms to manage the switch between spatial and temporal dimension.
 With this structure, users can manipulate the spatial or temporal dimension separately, while leaves the linking of two dimensions to \pkg{cubble}.
@@
 
-The rest of the paper will be divided as follows: Section 2 presents the workflow of data manipulation in cubble.
+The rest of the paper is divided as follows: Section 2 presents the workflow of data manipulation in cubble.
 Section 3 explains how cubble deals with some advanced considerations  including data with hierarchical structure, data matching, and how cubble fits with existing static and interactive visualisation tools.

### 2022-02-07 2-the-cubble-pkg.Rmd
small changes in read-through

@@
 
-Spatio-temporal data usually come in various forms and Figure \ref{fig:illu-input} shows four examples of this.
-These four examples show some common characters of the spatio-temporal data: 1) a spatial identifier, that uniquely identifies each site, 2) a temporal identifier, that prescribes the time stamp each site is recorded, 3) Coordinates, comprising of latitude and longitude, that locate each site on the map.
+Spatio-temporal data usually come to the analysts in various forms and Figure \ref{fig:illu-input} shows four examples of this.
+These four examples show the common characters of the spatio-temporal data: 1) a spatial identifier that uniquely identifies each site, 2) a temporal identifier that prescribes the time stamp each site is recorded, 3) Coordinates, comprising of latitude and longitude, that locate the site on the map.
 These identifiers will be the building blocks for cubble.
-Spatio-temporal data can also be thought of as a data cube with three axes: id, time, and variable.
-Notice that in the time-variable face, time invariant variables share the same value across all the time stamps while time varying variables have values varying across time.
+Another way to structure the spatio-temporal data is to think of it as a data cube with three axes: id, time, and variable.
+Notice that in the time-variable face of this cube, time invariant variables would share the same value across all the time stamps, while time varying variables have values varying across time.
+(Maybe add a cube diagram here).
 This heterogeneity will be used to split the data into two forms in cubble.
@@
 
-In a cubble, there are two forms: nested form and long form, and Figure \ref{fig:illu-cubble} sketches the two forms along with the associated attributes.
-The decision on which form to use is output-oriented, meaning analysts need to first think about whether the output of a particular operation is identified only by the spatial identifier, or a combination of spatial and temporal identifier.
-The nested cubble is suitable for working with operations that are only identified by the site and this type of operation can be a pure manipulation of spatial variables, or a summary of temporal variables by site (i.e.
-the output of counting the number of raining day is only identified by sites and hence should be performed with the nested form).
-Underneath the nested form, a cubble is built from a row-wise dataframe (\code{rowwise_df}) where each site forms a separate group.
-This structure simplifies the calculation that involves temporal variables by avoiding the use of \code{map} syntax when working with list-column.
+In a cubble, there are two forms: nested form and long form, and Figure \ref{fig:illu-cubble} sketches the two forms with the associated attributes.
+The decision on which form to use in data manipulation is output-oriented, meaning analysts need to first think about whether the output of an operation is identified only by the spatial identifier, or a combination of spatial and temporal identifier.
+The nested cubble is suitable for working with operations that are only identified by the site and this type of operation can be a pure manipulation of time invariant variables, or an operation that summarises time varying variables into site.
+Underneath the nested form, a cubble is built from a \code{rowwise_df} data frame where each site forms a separate group.
+This structure simplifies the calculation that involves temporal variables by avoiding the use of \code{purrr::map()} syntax when working with list-column.
 
-For those operations whose output involves both a spatial and temporal dimension, long form should be used.
-The long form is identified by both the spatial and temporal identifier and adopts a grouped dataframe (`grouped_df`) to forms each site as a group.
-Spatial variables are stored separately in a \pkg{tibble} as an special attribute of the long cubble.
+For those operations whose output involves both a spatial and temporal identifier, long form should be used.
+The long form uses a  \code{`grouped_df`} data frame to forms all the time of a site as a group.
+Time invariant variables are stored separately as an special attribute of the long cubble.
 This design avoids repeating the spatial variables at each time stamp while not dropping information from spatial variables.
@@
 ## Create a cubble in the nested form
-
-\code{as_cubble()} creates a cubble by supplying the three key components: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order of longitude first and then latitude.
+A cubble can be created with \code{as_cubble()} by supplying the three key components: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order of longitude and latitude.
 The naming of \code{key} and \code{index} follows the convention in the \pkg{tsibble} package.
-The cubble created by default is in the nested form.
-Below is an example of creating a cubble:
+Below is an example of creating a cubble and by default, the cubble created is in the nested form:
 
@@
 
-There are a few information in the cubble header: the name of the \code{key} variable (`id`) and its number (5), bounding box (`[115.97, -32.94, 133.55, -12.42]`), and also the name of variable nested in the \code{ts} column with its type: \code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.
+There are a few information in the cubble header: the name of the \code{key} variable (`id`) and its number (5), the bounding box: \code{[115.97, -32.94, 133.55, -12.42]} , and also the name of variable nested in the \code{ts} column with its type: \code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.
 
@@
 
-The long cubble is suitable to manipulate the time dimension of the data.
-The function \code{stretch()} switches the nested cubble into the long cubble by first extracts all the spatial variables into a separate tibble and store in the \code{spatial} attribute and then unnests the \code{ts} column:
+The long cubble is suitable to manipulate the time dimension of the data and  \code{stretch()} switches the nested cubble into the long cubble.
+It first extracts all the spatial variables into a separate tibble and then unnests the \code{ts} column.
+The spatial variables are then attached as an attributes of the long cubble and can be accessed through `spatial()`
 
@@
 cubble_long
+
+spatial(cubble_long)
 ```
 
-Notice here that the third line in the header now shows the name and type of spatial variables: \code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]}.
+Notice that the third line in the header now shows the name and type of spatial variables: \code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]}.
 
@@
 
-Manipulation on the spatial and temporal dimension can be an iterative process.
-Many times, analysts will need to go back and forth between the nested and long cubble.
-\code{tamp()} inverses the \code{stretch()} that introduced in the previous section, to switch from a long cubble to the nested cubble:
+Manipulation on the spatial and temporal dimension can be an iterative process and analysts may need to go back and forth between the nested and long cubble.
+\code{tamp()}, an inverse of \code{stretch()}, switches a long cubble to a nested cubble:
 
@@
 
-Some functions require time invariant variables to be in the same table as the time varying ones.
+Analysts may need the time invariant/ varying variable to be in the same table to apply certain function.
 In cubble, this can be done through \code{migrate()}, which moves the time invariant variables into the long cubble:
@@
 
-Notice that these added time invariant variables will not be remembered the cubble and will disappear if switched to the nested form and then switched back.
-Hence they should most likely been used at the last step of the analysis:
+Notice that these added time invariant variables will not be remembered by cubble and will disappear if switched to the nested form and then switched back.
+Hence, \code{migrat} should usually be used as the last step of the analysis:
 
@@
 
-In the created nested form, each element of the list-column \code{ts} is of \code{tbl_ts} class so that existing operations on the tsibble class can still be applied:
+In the created nested form, each element of the list-column \code{ts} is of \code{tbl_ts} class and existing operations on the tsibble class can still be applied:
 

### 2022-02-07 3-advanced.Rmd
small changes in read-through

@@
 
-Spatial locations can have grouping structure that is either inherent to the data  (i.e.
-countries are nested in continents), or formed by clustering.
+Spatial locations can have grouping structure either inherent to the data or formed by clustering.
 Rather than analysing variables in the site level, summarised variables in the cluster level can give a crisper picture of local areas.
-In cubble, \code{switch_key()} can be used to create a new level of grouping of spatial location by specifying a clustering variable.
+In cubble, \code{switch_key()} can be used to create a new level of grouping of spatial locations by specifying a clustering variable.
 Figure \ref{fig:illu-hier} illustrates the relationship of cubbles at station and cluster level, in both the long and nested form.
@@
 All the spatial variables variant to \code{cluster} are now nested into a \code{.val} column and cluster level variables can be computed in the same fashion as station level variables in \code{station_nested}.
-Following the same principal of \code{stretch}.
-the long form cluster level cubble expands the time series variables with both indices: \code{cluster} and \code{id}.
-All the cluster level variables are stored separately for recovering the nested from from the long form.
 
@@
 
-\textcolor{red}{refine the beginning.} To be a valid pair of matching, the matched pair from different data sources need to
+Cubble provides functionality for matching sites from two data sources. To be a valid pair of matching, the matched pair from different data sources need to:
 
@@
 
-\code{match_sites()} from \pkg{cubble} matches data based on these two criteria.
-It first matches the two data sources spatially through computing the pairwise distance based on latitude and longitude.
+\code{match_sites()} matches data based on these two criteria: it first matches the two data sources spatially through computing the pairwise distance on latitude and longitude.
 Pairs that pass the spatial matching are then matched temporally through computing the number of matched peaks within a fixed length window.
-Figure \ref{fig:illu-matching} illustrate this temporal matching in more details.
-Given two series \code{A} and \code{a}.
-3 peaks have been picked in each series.
+Figure \ref{fig:illu-matching} illustrates this temporal matching in more details.
+Given two series \code{A} and \code{a}, 3 peaks have been picked in each series.
 An interval, with default length of 5, is constructed for each peak in series \code{A} and the peaks in series \code{a} are tested against whether they fall into the any of the intervals.
@@
 Several arguments are available in \code{match_sites()} to fine-tune the matching:
-
-<!-- [can also think about matching as a form of correlation between series from a statistics point of view.] -->
 
@@
 
- - \code{spatial_n_keep}. the number of spatial match for each site to keep
- - \code{spatial_dist_max}. the maximum distance allowed for a matched pair 
- - \code{temporal_n_highest}. the number of peak used - 3 in the example above
- - \code{temporal_window}. the length of the interval - 5 in the example above
- - \code{temporal_min_match}. the minimum number of matched peak for a valid matched pair
+ - \code{spatial_n_keep}: the number of spatial match for each site to keep
+ - \code{spatial_dist_max}: the maximum distance allowed for a matched pair 
+ - \code{temporal_n_highest}: the number of peak used - 3 in the example above
+ - \code{temporal_window}: the length of the interval - 5 in the example above
+ - \code{temporal_min_match}: the minimum number of matched peak for a valid matched pair
 
-This functionality can be seen as a matching [@stuart2010matching; @mcintosh2018using] exercise in the spatio-temporal domain to pair up nearby observations.
+This functionality can be seen as a matching exercise [@stuart2010matching; @mcintosh2018using] in the spatio-temporal domain to pair up nearby observations.
 It can also be considered as a form of data fusion [@castanedo2013review; @cocchi2019data], where data from multiple sources are combined together.
@@
 Cubble fits in naturally with the interactive graphic pipeline discussed in the literature [@buja1988elements; @buja1996interactive;  @sutherland2000orca; @xie2014reactive; @cheng2016enabling].
-Diagram \ref{fig:illu-interactive} illustrates how linking works from the map to the time series with the two forms of a cubble.
-The map and time series plot is associated with the nested or long cubble respectively and when a user action is captured on the map, the site will be activated in the nested cubble (left).
+Diagram \ref{fig:illu-interactive} illustrates how linking works from the map to the time series in cubble.
+The map and time series plot is associated with the nested or long cubble, respectively, and when a user action is captured on the map, the site will be activated in the nested cubble (left).
 The nested cubble will communicate to the long cubble to activate all the observations with the same \code{id} (middle).
@@
 Glyph map [@Wickham2012-yr] plots the time series as single glyph on the map and in \proglang{R}, \pkg{GGally} implements the glyph map through the \code{glyphs()} function.
-The function construct a data frame with calculated position (\code{gx}.
+The function constructs a data frame with calculated position (\code{gx}.
 \code{gy}.
-\code{gid}.
-of each point on the time series using linear algebra (Equation 1 and 2 in @Wickham2012-yr).
+\code{gid}) of each point on the time series using linear algebra (Equation 1 and 2 in @Wickham2012-yr).
 The data can then be piped into \code{ggplot} to create the glyph map as:
@@
 
-A re-implementation of the glyph map as a ggproto \code{GeomGlyph} has been made in the \pkg{cubble} package so that the glyph map can be created with \code{geom_glyph()}:
+A re-implementation of the glyph map as a ggproto, \code{GeomGlyph}, has been made in the \pkg{cubble} package and now the glyph map can be created with \code{geom_glyph()}:
 
@@
 
+Some useful controls over the glyph map is also available in the \code{geom_glyph()} implementation.
 Polar glyph map can be specify as a parameter, \code{polar = TRUE}.
-in the \code{geom_glyph()}.
-along with \code{width} and \code{height} in either absolute or relative value.
+in the \code{geom_glyph()}, along with \code{width} and \code{height} in either absolute or relative value.
 Global and local scale can be controlled by the parameter \code{global_rescale}.

### 2022-02-07 4-1-basic.Rmd
small changes in read-through

@@
 
-Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world and the dataset \code{weatherdata::historical_tmax} extracts the maximum temperature for `r nrow(weatherdata::historical_tmax)` Australian stations.
-All the stations start recording from year `r range(weatherdata::historical_tmax$first_year)[2]` with the longest back to year `r range(weatherdata::historical_tmax$first_year)[1]`.
-The data \code{historical_tmax} is already presented as a cubble, with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates.
+Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world.
+The dataset \code{weatherdata::historical_tmax} extracts the maximum temperature for `r nrow(weatherdata::historical_tmax)` Australian stations from GHCN with starting from year `r range(weatherdata::historical_tmax$first_year)[2]`.
+\code{weatherdata::historical_tmax} is already in a cubble, with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates.
 This example compares the maximum temperature in two periods: 1971 - 1975 and 2016 - 2020 for stations in Victoria and New South Wales.
@@
 
-There are a few stations which don't have records during 1971 - 1975 and further investigation shows that while the first and last year of each series is recorded, the missing years in this period is not reported.
+A few stations don't have records during 1971 - 1975 and further investigation shows that while the first and last year of each series is recorded, the missing years in this period is not reported.
 These stations are filtered out by examining whether the summarised time series has a total of 24 months.

### 2022-02-07 4-2-agg.Rmd
small changes in read-through

@@
 Aggregation can be helpful to group series into clusters before visualising the cluster with glyph map.
+This example shows how to organise data at both level with \code{switch_key()}.
 
-This example shows how to organise data at both level with \code{switch_key()}.
-The dataset used is \code{weatherdata::climate_full}, which records daily precipitation and maximum/ minimum temperature for `r nrow(weatherdata::climate_full)` stations in Australia from 2016 to 2020.
-A simple kmean algorithm based on the distance matrix is used here to create 20 clusters.
+\code{weatherdata::climate_full}, also extracted from GHCN, records daily precipitation and maximum/ minimum temperature for `r nrow(weatherdata::climate_full)` stations in Australia from 2016 to 2020.
+A simple kmean algorithm based on the distance matrix is used to create 20 clusters.
 This creates \code{station_nested} as a station level nested cubble with a cluster column indicating the group each station belongs to.
@@
 
-Long form cubble at both levels can be access through stretching the nested form and operations applicable to long form cubble are still available.
-With access to both station and cluster level cubbles, various plots can be made to understand the cluster.
-Figure \ref{fig:basic-agg} shows two example plots that can be made with this data: subplot A makes a glyph map with cluster level cubble and subplot B inspects the station membership of each cluster using the nested station level cubble.
+Long form cubble at both levels can be accessed through stretching the nested form and with access to both station and cluster level cubbles, various plots can be made to understand the cluster.
+Figure \ref{fig:basic-agg} shows two example plots that can be made with this data: subplot A is a glyph map made with the cluster level cubble in the long form and subplot B inspects the station membership of each cluster using the station level cubble in the nested form.
 

### 2022-02-07 4-3-water-level.Rmd
small changes in read-through

@@
 
-Bureau of Meteorology also collects [water data](http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full) from river gauges, which includes electrical conductivity, turbidity, water course discharge, water course level, and water temperature.
+Bureau of Meteorology collects [water data](http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full) from river gauges and this includes variables: electrical conductivity, turbidity, water course discharge, water course level, and water temperature.
 In particular, water level will interactive with precipitation from the climate data since rainfall will raise the water level in the river.
@@
 
-From the map, a few water gauges and weather stations are close to each other and there is the potential that fluctuation of the water level can reflect the precipitation measured by the climate station.
+From the map, a few water gauges and weather stations are close to each other and the fluctuation of the water level could be matched up with precipitation measured by the climate station.
 As introduced in Section 3.2, \code{match_sites()} can be used to match one source of data with another source in a cubble.
-Here \code{prcp} in \code{climate} will be matched to \code{Water_course_level} in \code{river}, which is specified in the argument \code{temporal_var_to_match}.
-\code{temporal_independent} controls the dataset that will be used to construct the interval (see the illustration in section 3.2 for details on this).
-Here the goal is to see if precipitation will be reflected by the water level in the river and hence a lagged effect of water level on precipitation.
-This will put precipitation data, \code{climate}, as the independent.
-Given there is one year worth of data, the number of peak (\code{temporal_n_highest}) to consider is slightly raised from a default 20 to 30.
-\code{temporal_min_match} filters out pairs don't have enough match and to return all the pairs, set \code{temporal_min_match} to 0.
+Here \code{Water_course_level} in \code{river} will be matched to \code{prcp} in \code{climate} in 2020.
+The two datasets need to be specified as the first two arguments and the variable to match can be specified in \code{temporal_var_to_match} using the \code{by} syntax in \code{join}.
+\code{temporal_independent} controls the variable used to construct the interval and the goal here is to see if precipitation will be reflected by the water level in the river.
+This puts precipitation \code{prcp}, as the independent.
+Given there is one year worth of data, the number of peak (\code{temporal_n_highest}) to consider is slightly raised from a default 20 to 30 and \code{temporal_min_match} is raised correspondingly.
+To return all the pairs of the match, \code{temporal_min_match} can be set to 0.
 
@@
 
-The output from matching is also a cubble, with additional column \code{dist} and \code{group} produced from spatial matching and \code{n_match} from the temporal matching.
+The output from matching is also a cubble, with additional column \code{dist} and \code{group} produced from spatial matching and \code{n_match} from temporal matching.
 

### 2022-02-07 4-4-netcdf.Rmd
small changes in read-through

@@
 
-NetCDF (Network Common Data Form) is a data format commonly used in climatology community to deliver global mapping of atmosphere, ocean, and land.
-It has two major components: dimension to define the spatio-temporal grid (longitude, latitude, and time) and variable which populates the defined grid.
-Attributes are usually associated with dimension and variable in the NetCDF format data and a [metadata convention for climate and forecast](http://cfconventions.org/) has been designed to standardise this data format.
-Current packages on manipulating NetCDF data in R include a high-level R interface: \pkg{ncdf4} [@ncdf4], a low-level interface that calls C-interface: \pkg{RNetCDF} [@rnetcdf; @michna2013rnetcdf], and a tidyverse implementation: \pkg{tidync} [@tidync].
+NetCDF (Network Common Data Form) is a data format commonly used in the climatology community to deliver global mapping of atmosphere, ocean, and land.
+It has two major components: *dimension* for defining the spatio-temporal grid (longitude, latitude, and time) and *variable* that populates the defined grid.
+Attributes are usually associated with dimension and variable in the NetCDF format data and a [metadata convention for climate and forecast](http://cfconventions.org/) has been designed to standardise the format of the attributes.
+Current packages for manipulating NetCDF data in R include a high-level R interface: \pkg{ncdf4} [@ncdf4], a low-level interface that calls C-interface: \pkg{RNetCDF} [@rnetcdf; @michna2013rnetcdf], and a tidyverse implementation: \pkg{tidync} [@tidync].
 
 Cubble provides an \code{as_cubble()} method to coerce the \code{ncdf4} class from the \pkg{ncdf4} package into a \code{cubble}.
-It maps each combination of longitude and latitude into an \code{id} (the \code{key}) and nests time and variables in the nested form:
+It maps each combination of longitude and latitude into an \code{id} as the \code{key}:
 
@@
 Cubble can handle slightly more than 300 x 300 (longitude x longitude) grid points for 3 variables in one year and spatial grid can be reduced to trade for longer time period and more variables.
-A 300 by 300 spatial grid can be a bounding box of [100, -80, 180, 0] at 0.25 degree resolution or global bounding box [-180, -90, 180, -90] at 1 degree resolution.
+A 300 by 300 spatial grid can be a bounding box of [100, -80, 180, 0] at 0.25 degree resolution or a global bounding box [-180, -90, 180, -90] at 1 degree resolution.
 Subsetting longitude and latitude grid is available through \code{long_range} and \code{lat_range} if the NetCDF file has finer resolution than needed.
@@
 Figure \ref{fig:netcdf} gives an example of reproducing Figure 19 in @hersbach2020era5 using ERA5 data.
-ERA5 data [@hersbach2020era5] is the latest reanalysis of global atmosphere, land surface, and ocean waves from 1950 onwards and is available in the NetCDF format.
-It can be directly downloaded from the [Copernicus Climate Data Store (CDS)](https://cds.climate.copernicus.eu/cdsapp#!/dataset/reanalysis-era5-pressure-levels?tab=overview) website or programmatically via R package \pkg{ecmwfr} [@ecwmfr].
-Variable Specific humidity and geopotential are queried on the 10 hPa pressure level for four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04 and once downloaded, the data can be read into a cubble using the code above.
-Readers who are interested in the details of can refer to @simmons2020global and @simmons2005ecmwf.
+ERA5 data [@hersbach2020era5] is the latest reanalysis of global atmosphere, land surface, and ocean waves from 1950 onwards and is available in the NetCDF format from European Centre for Medium-Range Weather Forecasts (ECMWF).
+The data can be directly downloaded from [Copernicus Climate Data Store (CDS)](https://cds.climate.copernicus.eu/cdsapp#!/dataset/reanalysis-era5-pressure-levels?tab=overview) website or programmatically via an R package \pkg{ecmwfr} [@ecwmfr].
+Variable *specific humidity* and *geopotential* are queried on the 10 hPa pressure level on four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04.
+Once downloaded, the data can be read into a cubble as:
+
+```{r eval = FALSE, echo = TRUE}
+raw <- ncdf4::nc_open(here::here("data/era5-pressure.nc"))
+dt <- as_cubble(raw, vars = c("q", "z"))
+```
+
+Readers who are interested in the analysis of the figure can refer to @simmons2020global and @simmons2005ecmwf for more details.
 

### 2022-02-07 4-5-interactive.Rmd
small changes in read-through

@@
 
-With spatio-temporal data, users may wish to make plots to learn the spatial distribution of a variable or to find patterns such as trend or seasonality in the time series.
-Combining this two types of plot with interactivity let users to link between points on the map and their corresponding time series and explore the spatial and temporal dimension of the data simultaneously.
-Below is an example that describes the process of building an interactive graphic with \pkg{cubble} and \pkg{crosstalk}.
+With spatio-temporal data, users may wish to make plots to learn the spatial distribution of a variable, or to find patterns, such as trend or seasonality, in the time series.
+Combining this two types of plot with interactivity let users to link between points on the map and the corresponding time series to explore the spatial and temporal dimension of the data simultaneously.
+Below is an example that describes the process of building an interactive graphic with \pkg{cubble} and \pkg{crosstalk} The example explores the variation of monthly temperature range with \code{weatherdata::climate_full} data.
 
-Starting with the original data, some pre-processing may be required to summarise the data before the visualisation.
-This example explores the variation of monthly temperature range across `r nrow(weatherdata::climate_full)` weather stations in Australia and with `weatherdata::climate_full`, daily temperature range is first calculated as the difference between \code{tmax} and \code{tmin}.
-The three daily variables are then averaged into month over 2016 - 2020 in the long form.
-Variance of the temperature difference is then calculated for each station in the nested form.
-Then, a \code{SharedData} object is constructed for each form of the cubble and the same \code{group} argument ensures the cross-linking of the two forms via the common \code{id} column.
+The temperature range is calculated as the difference between \code{tmax} and \code{tmin} and its monthly average over 2016 - 2020 is taken before calculating the variance.
+A \code{SharedData} object is constructed for each form of the cubble and the same \code{group} argument ensures the cross-linking of the two forms via the common \code{id} column.
 The spatial map and time series plot are then made with each \code{SharedData} objects separately.
@@
 With a different dataset, users are free to calculate any per station measure in the nested form or to make any time-wise summary of the data in the long form to customise the spatial or temporal view.
-And the cross-linking is always safeguarded by the shared \code{id} column embedded in the cubble structure.
+The cross-linking between the two plots is always safeguarded by the shared \code{id} column embedded in the cubble structure.
 Below is the pseudo code that outlines the process to construct an interactive graphic described above:
@@
 In Figure \ref{fig:interactive-linking}, the first row shows the initial view of the interactive graphic.
-Overall, most regions in Australia have low variance of temperature range across different months while the north-west coastline, bottom of South Australia, and Victoria stands out with larger monthly changes.
-In the second row, Mount Elizabeth is selected on the map given its dark colour and this links to the ribbon on the right where there is a larger temperature range is presented in Australian winter (June to August).
-The third row selects the Grampian station in Victoria and the linked ribbon shows an opposite wider range in Australian summer period (December to February).
-The last row selects point with the lowest minimum temperature in August in the ribbon plot and surprisingly, this links to the Thredbo Airport in the Victoria and New South Wales border, rather than somewhere in the Tasmania island!
+On the map, most regions in Australia have low variance of temperature range while the north-west coastline, bottom of South Australia, and Victoria stands out with larger monthly changes.
+In the second row, Mount Elizabeth is selected on the map given its high variance colour on the initial map and this links to the ribbon on the right.
+The third row the lowest temperature in August and this corresponds to Thredbo AWS in the Victoria and New South Wales border.
+Another station in the Tasmania island is selected on the map to cross compare with Thredbo AWS.
 
@@
 
-With leaflet popup, the temperature range can be displayed as a small subplot upon clicking on the map.
-This would require first creating the popup plots from the long form cubble as a vector and then add these plots to a leaflet map, created from the nested cubble, with \code{leafpop::addPopupGraphs()}:
+This plot can also be made using \pkg{cubble} and \pkg{leaflet} where the temperature range can be displayed as a small subplot upon clicking on the map.
+This would require first creating the popup plots from the long form cubble as a vector and then add these plots to a leaflet map created from the nested cubble, with \code{leafpop::addPopupGraphs()}:
 
@@
 
-Figure \ref{fig:interactive-popup} shows the same content as Figure \ref{fig:interactive-linking} but made with leaflet and popups.
+Figure \ref{fig:interactive-popup} shows Figure \ref{fig:interactive-linking} made made with leaflet and popups.
 
-```{r interactive-popup, echo = FALSE, out.width="45%", out.height="25%", fig.retina = 2, dpi = 300, fig.cap = "Same as FIgure 11 with thetemperature variation shown as a popup in the leaflet map.", fig.show='hold'}
+```{r interactive-popup, echo = FALSE, out.width="45%", out.height="25%", fig.retina = 2, dpi = 300, fig.cap = "Same as Figure 11 with thetemperature variation shown as a popup in the leaflet map.", fig.show='hold'}
 knitr::include_graphics(here::here("figures/popup-mid.png")) 

### 2022-02-07 5-conclusion.Rmd
small changes in read-through

@@
 
-This paper describes the \proglang{R} package \pkg{cubble} for manipulating and visualising spatio-temporal data.
-A new data structure, that builds from the \code{rowwise_df} and \code{grouped_df} class in the tidyverse ecosystem, is proposed to connect the spatial and temporal variables in the spatio-temporal data.
+This paper describes an \proglang{R} package \pkg{cubble} for manipulating and visualising spatio-temporal data.
+A new data structure, \code{cubble} that builds from the \code{rowwise_df} and \code{grouped_df} class in the tidyverse ecosystem, is proposed to connect the time invariant and varying variables in the spatio-temporal data.
 This design frees the data analysts from spending time on organising variables of different observational units.
 The data structure is also flexible to the techniques and packages analysts use to analyse the data, for example, in the matching example in section 4.3, users are free to use algorithms from another package to cluster stations.
+
 Further development and maintenance of the package involves responding to changes in the tidyverse packages that \pkg{cubble} imports, in particular, \pkg{tibble}, \pkg{tidyr}, and \pkg{dplyr}.

### 2022-02-07 main.Rmd
small changes in read-through

--- 2022-02-06 main.Rmd
@@
 abstract: >
-  Sptio-temporal data naturally has two levels of observations: site and time level.
-[problem].
-In this paper, a new data structure, `cubble`, is proposed for manipulating and visualising spatio-temporal data.
+  Spatio-temporal variables can be divided into time invariant ones that can be identified by the spatial identifier, or the time varying ones that cross-identified by the spatial and temporal identifiers.
+In this paper, a new data structure, \code{cubble}, is proposed for manipulating and visualising spatio-temporal data.
+The new data structure uses two forms to organise time invariant/ varying variables so that the spatial and temporal dimension of the data can be manipulated separately while keep synchronised.
+Advanced considerations are given to illustrate data with hierarchical structure, data matching, and how cubble fits with the interactive graphics pipeline.
+Examples are given to analysing Australia meteorology data, river level data, and climate reanalysis (ERA5).
 keywords:

### 2022-02-07 main.tex
small changes in read-through

@@
 \Abstract{
-Spatio-temporal variables can be time invariant or not and this divides
-them into those can be identified only with the spatial identifier and
-those cross-identified by the spatial and temporal identifier. In this
+Spatio-temporal variables can be divided into time invariant ones that
+can be identified by the spatial identifier, or the time varying ones
+that cross-identified by the spatial and temporal identifiers. In this
 paper, a new data structure, \code{cubble}, is proposed for manipulating
-and visualising spatio-temporal data. The new data structure proposes
-two forms for manipulating the spatial and temporal dimension separately
-while providing a smooth switch between the two forms to link the two
-dimensions. Advanced considerations are given to illustrate data with
-hierarchical structure, data matching, and how \pkg{cubble} fits with
+and visualising spatio-temporal data. The new data structure uses two
+forms to organise time invariant/ varying variables so that the spatial
+and temporal dimension of the data can be manipulated separately while
+keep synchronised. Advanced considerations are given to illustrate data
+with hierarchical structure, data matching, and how cubble fits with the
 interactive graphics pipeline. Examples are given to analysing Australia
@@
 identified by its spatial location while other time varying variables
-need both the spatial and temporal identifier to identify. Manipulating
+need both the spatial and temporal identifiers to identify. Manipulating
 data in these two levels is usually done separately and the third tidy
@@
 
-Currently, available spatio-temporal data structure in R includes:
+Currently, available spatio-temporal data structures in R includes:
 \pkg{spacetime} \citep{spacetime}, which proposes four space-time
 layouts: Full grid (STF), sparse grid(STS), irregular (STI), and
-trajectory (STT). The data structure it uses is based on \pkg{sp}
-\citep{sp} and \pkg{xts} \citep{xts}, both of which has been replaced by
-more recent implementations. \pkg{spatstat} \citep{spatstat} implements
-a \texttt{ppp} class for point pattern data; and more recent,
-\pkg{stars} \citep{stars} implements a spatio-temporal array with the
-dplyr's data cube structure \pkg{cubelyr} \citep{cubelyr} as its
-back-end. While these implementations either store spatial and temporal
-variables all in a single table, hence duplicate the spatial variables
-for each temporal unit; or split them into two separate tables that has
-the problem of manually joining, mentioned in the previously. This
-creates a gap in the software development. The requirement for such a
-tool is important given the ubiquity of spatio-temporal vector data in
-the wild: the Ireland wind data from \pkg{gstat} is an classic example
-data that splits variables into its spatial (\texttt{wind.loc}) and
-temporal (\texttt{wind}) dimension; Bureau of Meteorology (BoM) provides
-climate observations widely applied in agriculture and ecology study.
+trajectory (STT). \pkg{spacetime} is built on \pkg{sp} \citep{sp} and
+\pkg{xts} \citep{xts}, both of which have been replaced by more recent
+implementations. \pkg{spatstat} \citep{spatstat} implements a
+\texttt{ppp} class for point pattern data; and more recent, \pkg{stars}
+\citep{stars} implements a spatio-temporal for vector and raster
+spatio-temporal data. While these implementations either store spatial
+and temporal variables all in a single table, hence duplicate the
+spatial variables at each temporal unit; or split them into two separate
+tables that analysts need to manually join the data constantly when
+doing the analysis. This creates a gap in the software development. The
+requirement for such a tool is important given the ubiquity of
+spatio-temporal vector data in the wild: climate observations from
+weather stations, air quality variables from air monitoring stations, or
+any variables that are measured at fixed location across time.
 
 This paper describes the implementation of a new spatio-temporal data
-structure: \pkg{cubble}. \pkg{cubble} implements a relational data
-structure that uses two forms to manage the switch between spatial and
-temporal dimension. With this structure, users can manipulate the
-spatial or temporal dimension separately, while leaves the linking of
-two dimensions to \pkg{cubble}. The software is available from the
+structure: \pkg{cubble}. \pkg{Cubble} implements a data structure that
+uses two forms to manage the switch between spatial and temporal
+dimension. With this structure, users can manipulate the spatial or
+temporal dimension separately, while leaves the linking of two
+dimensions to \pkg{cubble}. The software is available from the
 Comprehensive R Archive Network (CRAN) at {[}CRAN link{]}.
 
-The rest of the paper will be divided as follows: Section 2 presents the
+The rest of the paper is divided as follows: Section 2 presents the
 workflow of data manipulation in cubble. Section 3 explains how cubble
@@
 
-Spatio-temporal data usually come in various forms and Figure
-\ref{fig:illu-input} shows four examples of this. These four examples
-show some common characters of the spatio-temporal data: 1) a spatial
-identifier, that uniquely identifies each site, 2) a temporal
-identifier, that prescribes the time stamp each site is recorded, 3)
-Coordinates, comprising of latitude and longitude, that locate each site
+Spatio-temporal data usually come to the analysts in various forms and
+Figure \ref{fig:illu-input} shows four examples of this. These four
+examples show the common characters of the spatio-temporal data: 1) a
+spatial identifier that uniquely identifies each site, 2) a temporal
+identifier that prescribes the time stamp each site is recorded, 3)
+Coordinates, comprising of latitude and longitude, that locate the site
 on the map. These identifiers will be the building blocks for cubble.
-Spatio-temporal data can also be thought of as a data cube with three
-axes: id, time, and variable. Notice that in the time-variable face,
-time invariant variables share the same value across all the time stamps
-while time varying variables have values varying across time. This
+Another way to structure the spatio-temporal data is to think of it as a
+data cube with three axes: id, time, and variable. Notice that in the
+time-variable face of this cube, time invariant variables would share
+the same value across all the time stamps, while time varying variables
+have values varying across time. (Maybe add a cube diagram here). This
 heterogeneity will be used to split the data into two forms in cubble.
@@
 In a cubble, there are two forms: nested form and long form, and Figure
-\ref{fig:illu-cubble} sketches the two forms along with the associated
-attributes. The decision on which form to use is output-oriented,
-meaning analysts need to first think about whether the output of a
-particular operation is identified only by the spatial identifier, or a
-combination of spatial and temporal identifier. The nested cubble is
+\ref{fig:illu-cubble} sketches the two forms with the associated
+attributes. The decision on which form to use in data manipulation is
+output-oriented, meaning analysts need to first think about whether the
+output of an operation is identified only by the spatial identifier, or
+a combination of spatial and temporal identifier. The nested cubble is
 suitable for working with operations that are only identified by the
-site and this type of operation can be a pure manipulation of spatial
-variables, or a summary of temporal variables by site (i.e.~the output
-of counting the number of raining day is only identified by sites and
-hence should be performed with the nested form). Underneath the nested
-form, a cubble is built from a row-wise dataframe (\code{rowwise_df})
-where each site forms a separate group. This structure simplifies the
-calculation that involves temporal variables by avoiding the use of
-\code{map} syntax when working with list-column.
+site and this type of operation can be a pure manipulation of time
+invariant variables, or an operation that summarises time varying
+variables into site. Underneath the nested form, a cubble is built from
+a \code{rowwise_df} data frame where each site forms a separate group.
+This structure simplifies the calculation that involves temporal
+variables by avoiding the use of \code{purrr::map()} syntax when working
+with list-column.
 
 For those operations whose output involves both a spatial and temporal
-dimension, long form should be used. The long form is identified by both
-the spatial and temporal identifier and adopts a grouped dataframe
-(\texttt{grouped\_df}) to forms each site as a group. Spatial variables
-are stored separately in a \pkg{tibble} as an special attribute of the
-long cubble. This design avoids repeating the spatial variables at each
-time stamp while not dropping information from spatial variables.
+identifier, long form should be used. The long form uses a
+\code{`grouped_df`} data frame to forms all the time of a site as a
+group. Time invariant variables are stored separately as an special
+attribute of the long cubble. This design avoids repeating the spatial
+variables at each time stamp while not dropping information from spatial
+variables.
 
@@
 
-\code{as_cubble()} creates a cubble by supplying the three key
-components: \code{key} as the spatial identifier; \code{index} as the
-temporal identifier; and a vector of \code{coords} in the order of
-longitude first and then latitude. The naming of \code{key} and
-\code{index} follows the convention in the \pkg{tsibble} package. The
-cubble created by default is in the nested form. Below is an example of
-creating a cubble:
+A cubble can be created with \code{as_cubble()} by supplying the three
+key components: \code{key} as the spatial identifier; \code{index} as
+the temporal identifier; and a vector of \code{coords} in the order of
+longitude and latitude. The naming of \code{key} and \code{index}
+follows the convention in the \pkg{tsibble} package. Below is an example
+of creating a cubble and by default, the cubble created is in the nested
+form:
 
@@
 There are a few information in the cubble header: the name of the
-\code{key} variable (\texttt{id}) and its number (5), bounding box
-(\texttt{{[}115.97,\ -32.94,\ 133.55,\ -12.42{]}}), and also the name of
-variable nested in the \code{ts} column with its type:
+\code{key} variable (\texttt{id}) and its number (5), the bounding box:
+\code{[115.97, -32.94, 133.55, -12.42]} , and also the name of variable
+nested in the \code{ts} column with its type:
 \code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.
@@
 
-The long cubble is suitable to manipulate the time dimension of the
-data. The function \code{stretch()} switches the nested cubble into the
-long cubble by first extracts all the spatial variables into a separate
-tibble and store in the \code{spatial} attribute and then unnests the
-\code{ts} column:
+The long cubble is suitable to manipulate the time dimension of the data
+and \code{stretch()} switches the nested cubble into the long cubble. It
+first extracts all the spatial variables into a separate tibble and then
+unnests the \code{ts} column. The spatial variables are then attached as
+an attributes of the long cubble and can be accessed through
+\texttt{spatial()}
 
@@
 # cubble:  date, id [5]: long form
-# bbox:    [115.97, -32.94, 133.55, -12.42]- check gap on long and lat
+# bbox:    [115.97, -32.94, 133.55, -12.42]
 # spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]
@@
 \end{CodeOutput}
-\end{CodeChunk}
-
-Notice here that the third line in the header now shows the name and
-type of spatial variables:
+\begin{CodeInput}
+R> spatial(cubble_long)
+\end{CodeInput}
+\begin{CodeOutput}
+# A tibble: 5 x 6
+  id            lat  long  elev name           wmo_id
+  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl>
+1 ASN00009021 -31.9  116.  15.4 perth airport   94610
+2 ASN00010311 -31.9  117. 179   york            94623
+3 ASN00010614 -32.9  117. 338   narrogin        94627
+4 ASN00014015 -12.4  131.  30.4 darwin airport  94120
+5 ASN00015131 -17.6  134. 220   elliott         94236
+\end{CodeOutput}
+\end{CodeChunk}
+
+Notice that the third line in the header now shows the name and type of
+spatial variables:
 \code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]}.
@@
 Manipulation on the spatial and temporal dimension can be an iterative
-process. Many times, analysts will need to go back and forth between the
-nested and long cubble. \code{tamp()} inverses the \code{stretch()} that
-introduced in the previous section, to switch from a long cubble to the
-nested cubble:
+process and analysts may need to go back and forth between the nested
+and long cubble. \code{tamp()}, an inverse of \code{stretch()}, switches
+a long cubble to a nested cubble:
 
@@
 
-Some functions require time invariant variables to be in the same table
-as the time varying ones. In cubble, this can be done through
+Analysts may need the time invariant/ varying variable to be in the same
+table to apply certain function. In cubble, this can be done through
 \code{migrate()}, which moves the time invariant variables into the long
@@
 Notice that these added time invariant variables will not be remembered
-the cubble and will disappear if switched to the nested form and then
-switched back. Hence they should most likely been used at the last step
-of the analysis:
+by cubble and will disappear if switched to the nested form and then
+switched back. Hence, \code{migrat} should usually be used as the last
+step of the analysis:
 
@@
 In the created nested form, each element of the list-column \code{ts} is
-of \code{tbl_ts} class so that existing operations on the tsibble class
-can still be applied:
+of \code{tbl_ts} class and existing operations on the tsibble class can
+still be applied:
 
@@
 
-Spatial locations can have grouping structure that is either inherent to
-the data (i.e.~countries are nested in continents), or formed by
-clustering. Rather than analysing variables in the site level,
-summarised variables in the cluster level can give a crisper picture of
-local areas. In cubble, \code{switch_key()} can be used to create a new
-level of grouping of spatial location by specifying a clustering
-variable. Figure \ref{fig:illu-hier} illustrates the relationship of
-cubbles at station and cluster level, in both the long and nested form.
-By specifying
+Spatial locations can have grouping structure either inherent to the
+data or formed by clustering. Rather than analysing variables in the
+site level, summarised variables in the cluster level can give a crisper
+picture of local areas. In cubble, \code{switch_key()} can be used to
+create a new level of grouping of spatial locations by specifying a
+clustering variable. Figure \ref{fig:illu-hier} illustrates the
+relationship of cubbles at station and cluster level, in both the long
+and nested form. By specifying
 \code{cluster_nested <- station_nested \%>\% switch_key(key = cluster)},
@@
 same fashion as station level variables in \code{station_nested}.
-Following the same principal of \code{stretch}. the long form cluster
-level cubble expands the time series variables with both indices:
-\code{cluster} and \code{id}. All the cluster level variables are stored
-separately for recovering the nested from from the long form.
 
@@
 
-\textcolor{red}{refine the beginning.} To be a valid pair of matching,
-the matched pair from different data sources need to
+Cubble provides functionality for matching sites from two data sources.
+To be a valid pair of matching, the matched pair from different data
+sources need to:
 
@@
 
-\code{match_sites()} from \pkg{cubble} matches data based on these two
-criteria. It first matches the two data sources spatially through
-computing the pairwise distance based on latitude and longitude. Pairs
-that pass the spatial matching are then matched temporally through
-computing the number of matched peaks within a fixed length window.
-Figure \ref{fig:illu-matching} illustrate this temporal matching in more
-details. Given two series \code{A} and \code{a}. 3 peaks have been
-picked in each series. An interval, with default length of 5, is
-constructed for each peak in series \code{A} and the peaks in series
-\code{a} are tested against whether they fall into the any of the
-intervals. In this illustration, there are 2 matches for these two
-series. Several arguments are available in \code{match_sites()} to
-fine-tune the matching:
+\code{match_sites()} matches data based on these two criteria: it first
+matches the two data sources spatially through computing the pairwise
+distance on latitude and longitude. Pairs that pass the spatial matching
+are then matched temporally through computing the number of matched
+peaks within a fixed length window. Figure \ref{fig:illu-matching}
+illustrates this temporal matching in more details. Given two series
+\code{A} and \code{a}, 3 peaks have been picked in each series. An
+interval, with default length of 5, is constructed for each peak in
+series \code{A} and the peaks in series \code{a} are tested against
+whether they fall into the any of the intervals. In this illustration,
+there are 2 matches for these two series. Several arguments are
+available in \code{match_sites()} to fine-tune the matching:
 
@@
 \item
-  \code{spatial_n_keep}. the number of spatial match for each site to
+  \code{spatial_n_keep}: the number of spatial match for each site to
   keep
 \item
-  \code{spatial_dist_max}. the maximum distance allowed for a matched
+  \code{spatial_dist_max}: the maximum distance allowed for a matched
   pair
 \item
-  \code{temporal_n_highest}. the number of peak used - 3 in the example
+  \code{temporal_n_highest}: the number of peak used - 3 in the example
   above
 \item
-  \code{temporal_window}. the length of the interval - 5 in the example
+  \code{temporal_window}: the length of the interval - 5 in the example
   above
 \item
-  \code{temporal_min_match}. the minimum number of matched peak for a
+  \code{temporal_min_match}: the minimum number of matched peak for a
   valid matched pair
@@
 
-This functionality can be seen as a matching
-\citep{stuart2010matching, mcintosh2018using} exercise in the
-spatio-temporal domain to pair up nearby observations. It can also be
-considered as a form of data fusion
-\citep{castanedo2013review, cocchi2019data}, where data from multiple
-sources are combined together.
+This functionality can be seen as a matching exercise
+\citep{stuart2010matching, mcintosh2018using} in the spatio-temporal
+domain to pair up nearby observations. It can also be considered as a
+form of data fusion \citep{castanedo2013review, cocchi2019data}, where
+data from multiple sources are combined together.
 
@@
 Diagram \ref{fig:illu-interactive} illustrates how linking works from
-the map to the time series with the two forms of a cubble. The map and
-time series plot is associated with the nested or long cubble
-respectively and when a user action is captured on the map, the site
-will be activated in the nested cubble (left). The nested cubble will
-communicate to the long cubble to activate all the observations with the
-same \code{id} (middle). The long cubble will then highlight the
-activated series in the time series plot (right).
+the map to the time series in cubble. The map and time series plot is
+associated with the nested or long cubble, respectively, and when a user
+action is captured on the map, the site will be activated in the nested
+cubble (left). The nested cubble will communicate to the long cubble to
+activate all the observations with the same \code{id} (middle). The long
+cubble will then highlight the activated series in the time series plot
+(right).
 
@@
 on the map and in \proglang{R}, \pkg{GGally} implements the glyph map
-through the \code{glyphs()} function. The function construct a data
-frame with calculated position (\code{gx}. \code{gy}. \code{gid}. of
+through the \code{glyphs()} function. The function constructs a data
+frame with calculated position (\code{gx}. \code{gy}. \code{gid}) of
 each point on the time series using linear algebra (Equation 1 and 2 in
@@
 
-A re-implementation of the glyph map as a ggproto \code{GeomGlyph} has
-been made in the \pkg{cubble} package so that the glyph map can be
+A re-implementation of the glyph map as a ggproto, \code{GeomGlyph}, has
+been made in the \pkg{cubble} package and now the glyph map can be
 created with \code{geom_glyph()}:
@@
 
-Polar glyph map can be specify as a parameter, \code{polar = TRUE}. in
-the \code{geom_glyph()}. along with \code{width} and \code{height} in
-either absolute or relative value. Global and local scale can be
-controlled by the parameter \code{global_rescale}. which default to
-\code{TRUE} for global scaling. Reference box and line can be added with
-separate \code{geom_glyph_box()} and \code{geom_glyph_line()}.
+Some useful controls over the glyph map is also available in the
+\code{geom_glyph()} implementation. Polar glyph map can be specify as a
+parameter, \code{polar = TRUE}. in the \code{geom_glyph()}, along with
+\code{width} and \code{height} in either absolute or relative value.
+Global and local scale can be controlled by the parameter
+\code{global_rescale}. which default to \code{TRUE} for global scaling.
+Reference box and line can be added with separate
+\code{geom_glyph_box()} and \code{geom_glyph_line()}.
 
@@
 Global Historical Climatology Network (GHCN) provides daily climate
-measures from stations across the world and the dataset
+measures from stations across the world. The dataset
 \code{weatherdata::historical_tmax} extracts the maximum temperature for
-236 Australian stations. All the stations start recording from year 1969
-with the longest back to year 1859. The data \code{historical_tmax} is
-already presented as a cubble, with \code{id} as the key, \code{date} as
-the index, and \code{c(longitude, latitude)} as the coordinates. This
-example compares the maximum temperature in two periods: 1971 - 1975 and
-2016 - 2020 for stations in Victoria and New South Wales.
+236 Australian stations from GHCN with starting from year 1969.
+\code{weatherdata::historical_tmax} is already in a cubble, with
+\code{id} as the key, \code{date} as the index, and
+\code{c(longitude, latitude)} as the coordinates. This example compares
+the maximum temperature in two periods: 1971 - 1975 and 2016 - 2020 for
+stations in Victoria and New South Wales.
 
@@
 
-There are a few stations which don't have records during 1971 - 1975 and
-further investigation shows that while the first and last year of each
-series is recorded, the missing years in this period is not reported.
-These stations are filtered out by examining whether the summarised time
+A few stations don't have records during 1971 - 1975 and further
+investigation shows that while the first and last year of each series is
+recorded, the missing years in this period is not reported. These
+stations are filtered out by examining whether the summarised time
 series has a total of 24 months. The long cubble needs to be switched to
@@
 Aggregation can be helpful to group series into clusters before
-visualising the cluster with glyph map.
-
-This example shows how to organise data at both level with
-\code{switch_key()}. The dataset used is
-\code{weatherdata::climate_full}, which records daily precipitation and
-maximum/ minimum temperature for 639 stations in Australia from 2016 to
-2020. A simple kmean algorithm based on the distance matrix is used here
-to create 20 clusters. This creates \code{station_nested} as a station
-level nested cubble with a cluster column indicating the group each
-station belongs to. More complex algorithms can be used for other
-problem, as long as there is a mapping from each station to a cluster.
+visualising the cluster with glyph map. This example shows how to
+organise data at both level with \code{switch_key()}.
+
+\code{weatherdata::climate_full}, also extracted from GHCN, records
+daily precipitation and maximum/ minimum temperature for 639 stations in
+Australia from 2016 to 2020. A simple kmean algorithm based on the
+distance matrix is used to create 20 clusters. This creates
+\code{station_nested} as a station level nested cubble with a cluster
+column indicating the group each station belongs to. More complex
+algorithms can be used for other problem, as long as there is a mapping
+from each station to a cluster.
 
@@
 
-Long form cubble at both levels can be access through stretching the
-nested form and operations applicable to long form cubble are still
-available. With access to both station and cluster level cubbles,
+Long form cubble at both levels can be accessed through stretching the
+nested form and with access to both station and cluster level cubbles,
 various plots can be made to understand the cluster. Figure
 \ref{fig:basic-agg} shows two example plots that can be made with this
-data: subplot A makes a glyph map with cluster level cubble and subplot
-B inspects the station membership of each cluster using the nested
-station level cubble.
+data: subplot A is a glyph map made with the cluster level cubble in the
+long form and subplot B inspects the station membership of each cluster
+using the station level cubble in the nested form.
 
@@
 
-Bureau of Meteorology also collects
+Bureau of Meteorology collects
 \href{http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full}{water
-data} from river gauges, which includes electrical conductivity,
-turbidity, water course discharge, water course level, and water
-temperature. In particular, water level will interactive with
+data} from river gauges and this includes variables: electrical
+conductivity, turbidity, water course discharge, water course level, and
+water temperature. In particular, water level will interactive with
 precipitation from the climate data since rainfall will raise the water
@@
 From the map, a few water gauges and weather stations are close to each
-other and there is the potential that fluctuation of the water level can
-reflect the precipitation measured by the climate station. As introduced
-in Section 3.2, \code{match_sites()} can be used to match one source of
-data with another source in a cubble. Here \code{prcp} in \code{climate}
-will be matched to \code{Water_course_level} in \code{river}, which is
-specified in the argument \code{temporal_var_to_match}.
-\code{temporal_independent} controls the dataset that will be used to
-construct the interval (see the illustration in section 3.2 for details
-on this). Here the goal is to see if precipitation will be reflected by
-the water level in the river and hence a lagged effect of water level on
-precipitation. This will put precipitation data, \code{climate}, as the
-independent. Given there is one year worth of data, the number of peak
+other and the fluctuation of the water level could be matched up with
+precipitation measured by the climate station. As introduced in Section
+3.2, \code{match_sites()} can be used to match one source of data with
+another source in a cubble. Here \code{Water_course_level} in
+\code{river} will be matched to \code{prcp} in \code{climate} in 2020.
+The two datasets need to be specified as the first two arguments and the
+variable to match can be specified in \code{temporal_var_to_match} using
+the \code{by} syntax in \code{join}. \code{temporal_independent}
+controls the variable used to construct the interval and the goal here
+is to see if precipitation will be reflected by the water level in the
+river. This puts precipitation \code{prcp}, as the independent. Given
+there is one year worth of data, the number of peak
 (\code{temporal_n_highest}) to consider is slightly raised from a
-default 20 to 30. \code{temporal_min_match} filters out pairs don't have
-enough match and to return all the pairs, set \code{temporal_min_match}
-to 0.
+default 20 to 30 and \code{temporal_min_match} is raised
+correspondingly. To return all the pairs of the match,
+\code{temporal_min_match} can be set to 0.
 
@@
 \code{dist} and \code{group} produced from spatial matching and
-\code{n_match} from the temporal matching.
+\code{n_match} from temporal matching.
 
@@
 
-NetCDF (Network Common Data Form) is a data format commonly used in
+NetCDF (Network Common Data Form) is a data format commonly used in the
 climatology community to deliver global mapping of atmosphere, ocean,
-and land. It has two major components: dimension to define the
-spatio-temporal grid (longitude, latitude, and time) and variable which
-populates the defined grid. Attributes are usually associated with
+and land. It has two major components: \emph{dimension} for defining the
+spatio-temporal grid (longitude, latitude, and time) and \emph{variable}
+that populates the defined grid. Attributes are usually associated with
 dimension and variable in the NetCDF format data and a
 \href{http://cfconventions.org/}{metadata convention for climate and
-forecast} has been designed to standardise this data format. Current
-packages on manipulating NetCDF data in R include a high-level R
-interface: \pkg{ncdf4} \citep{ncdf4}, a low-level interface that calls
+forecast} has been designed to standardise the format of the attributes.
+Current packages for manipulating NetCDF data in R include a high-level
+R interface: \pkg{ncdf4} \citep{ncdf4}, a low-level interface that calls
 C-interface: \pkg{RNetCDF} \citep{rnetcdf, michna2013rnetcdf}, and a
@@
 class from the \pkg{ncdf4} package into a \code{cubble}. It maps each
-combination of longitude and latitude into an \code{id} (the \code{key})
-and nests time and variables in the nested form:
+combination of longitude and latitude into an \code{id} as the
+\code{key}:
 
@@
 longer time period and more variables. A 300 by 300 spatial grid can be
-a bounding box of {[}100, -80, 180, 0{]} at 0.25 degree resolution or
+a bounding box of {[}100, -80, 180, 0{]} at 0.25 degree resolution or a
 global bounding box {[}-180, -90, 180, -90{]} at 1 degree resolution.
@@
 land surface, and ocean waves from 1950 onwards and is available in the
-NetCDF format. It can be directly downloaded from the
+NetCDF format from European Centre for Medium-Range Weather Forecasts
+(ECMWF). The data can be directly downloaded from
 \href{https://cds.climate.copernicus.eu/cdsapp\#!/dataset/reanalysis-era5-pressure-levels?tab=overview}{Copernicus
-Climate Data Store (CDS)} website or programmatically via R package
-\pkg{ecmwfr} \citep{ecwmfr}. Variable Specific humidity and geopotential
-are queried on the 10 hPa pressure level for four dates: 2002-09-22,
-2002-09-26, 2002-09-30, and 2002-10-04 and once downloaded, the data can
-be read into a cubble using the code above. Readers who are interested
-in the details of can refer to \citet{simmons2020global} and
-\citet{simmons2005ecmwf}.
+Climate Data Store (CDS)} website or programmatically via an R package
+\pkg{ecmwfr} \citep{ecwmfr}. Variable \emph{specific humidity} and
+\emph{geopotential} are queried on the 10 hPa pressure level on four
+dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04. Once
+downloaded, the data can be read into a cubble as:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> raw <- ncdf4::nc_open(here::here("data/era5-pressure.nc"))
+R> dt <- as_cubble(raw, vars = c("q", "z"))
+\end{CodeInput}
+\end{CodeChunk}
+
+Readers who are interested in the analysis of the figure can refer to
+\citet{simmons2020global} and \citet{simmons2005ecmwf} for more details.
 
@@
 With spatio-temporal data, users may wish to make plots to learn the
-spatial distribution of a variable or to find patterns such as trend or
-seasonality in the time series. Combining this two types of plot with
-interactivity let users to link between points on the map and their
-corresponding time series and explore the spatial and temporal dimension
+spatial distribution of a variable, or to find patterns, such as trend
+or seasonality, in the time series. Combining this two types of plot
+with interactivity let users to link between points on the map and the
+corresponding time series to explore the spatial and temporal dimension
 of the data simultaneously. Below is an example that describes the
 process of building an interactive graphic with \pkg{cubble} and
-\pkg{crosstalk}.
-
-Starting with the original data, some pre-processing may be required to
-summarise the data before the visualisation. This example explores the
-variation of monthly temperature range across 639 weather stations in
-Australia and with \texttt{weatherdata::climate\_full}, daily
-temperature range is first calculated as the difference between
-\code{tmax} and \code{tmin}. The three daily variables are then averaged
-into month over 2016 - 2020 in the long form. Variance of the
-temperature difference is then calculated for each station in the nested
-form. Then, a \code{SharedData} object is constructed for each form of
-the cubble and the same \code{group} argument ensures the cross-linking
-of the two forms via the common \code{id} column. The spatial map and
-time series plot are then made with each \code{SharedData} objects
-separately. In this example, stations on the Australia map, made from
-the nested form, are coloured by the calculated variance and a ribbon
-band is constructed using the long form cubble to show the maximum and
-minimum temperature of each station across month. With a different
-dataset, users are free to calculate any per station measure in the
-nested form or to make any time-wise summary of the data in the long
-form to customise the spatial or temporal view. And the cross-linking is
-always safeguarded by the shared \code{id} column embedded in the cubble
+\pkg{crosstalk} The example explores the variation of monthly
+temperature range with \code{weatherdata::climate_full} data.
+
+The temperature range is calculated as the difference between
+\code{tmax} and \code{tmin} and its monthly average over 2016 - 2020 is
+taken before calculating the variance. A \code{SharedData} object is
+constructed for each form of the cubble and the same \code{group}
+argument ensures the cross-linking of the two forms via the common
+\code{id} column. The spatial map and time series plot are then made
+with each \code{SharedData} objects separately. In this example,
+stations on the Australia map, made from the nested form, are coloured
+by the calculated variance and a ribbon band is constructed using the
+long form cubble to show the maximum and minimum temperature of each
+station across month. With a different dataset, users are free to
+calculate any per station measure in the nested form or to make any
+time-wise summary of the data in the long form to customise the spatial
+or temporal view. The cross-linking between the two plots is always
+safeguarded by the shared \code{id} column embedded in the cubble
 structure. Below is the pseudo code that outlines the process to
@@
 In Figure \ref{fig:interactive-linking}, the first row shows the initial
-view of the interactive graphic. Overall, most regions in Australia have
-low variance of temperature range across different months while the
-north-west coastline, bottom of South Australia, and Victoria stands out
-with larger monthly changes. In the second row, Mount Elizabeth is
-selected on the map given its dark colour and this links to the ribbon
-on the right where there is a larger temperature range is presented in
-Australian winter (June to August). The third row selects the Grampian
-station in Victoria and the linked ribbon shows an opposite wider range
-in Australian summer period (December to February). The last row selects
-point with the lowest minimum temperature in August in the ribbon plot
-and surprisingly, this links to the Thredbo Airport in the Victoria and
-New South Wales border, rather than somewhere in the Tasmania island!
+view of the interactive graphic. On the map, most regions in Australia
+have low variance of temperature range while the north-west coastline,
+bottom of South Australia, and Victoria stands out with larger monthly
+changes. In the second row, Mount Elizabeth is selected on the map given
+its high variance colour on the initial map and this links to the ribbon
+on the right. The third row the lowest temperature in August and this
+corresponds to Thredbo AWS in the Victoria and New South Wales border.
+Another station in the Tasmania island is selected on the map to cross
+compare with Thredbo AWS.
 
@@
 
-With leaflet popup, the temperature range can be displayed as a small
-subplot upon clicking on the map. This would require first creating the
-popup plots from the long form cubble as a vector and then add these
-plots to a leaflet map, created from the nested cubble, with
-\code{leafpop::addPopupGraphs()}:
+This plot can also be made using \pkg{cubble} and \pkg{leaflet} where
+the temperature range can be displayed as a small subplot upon clicking
+on the map. This would require first creating the popup plots from the
+long form cubble as a vector and then add these plots to a leaflet map
+created from the nested cubble, with \code{leafpop::addPopupGraphs()}:
 
@@
 
-Figure \ref{fig:interactive-popup} shows the same content as Figure
-\ref{fig:interactive-linking} but made with leaflet and popups.
+Figure \ref{fig:interactive-popup} shows Figure
+\ref{fig:interactive-linking} made made with leaflet and popups.
 
@@
 
-\caption[Same as FIgure 11 with thetemperature variation shown as a popup in the leaflet map]{Same as FIgure 11 with thetemperature variation shown as a popup in the leaflet map.}\label{fig:interactive-popup}
+\caption[Same as Figure 11 with thetemperature variation shown as a popup in the leaflet map]{Same as Figure 11 with thetemperature variation shown as a popup in the leaflet map.}\label{fig:interactive-popup}
 \end{figure}
@@
 
-This paper describes the \proglang{R} package \pkg{cubble} for
+This paper describes an \proglang{R} package \pkg{cubble} for
 manipulating and visualising spatio-temporal data. A new data structure,
-that builds from the \code{rowwise_df} and \code{grouped_df} class in
-the tidyverse ecosystem, is proposed to connect the spatial and temporal
-variables in the spatio-temporal data. This design frees the data
-analysts from spending time on organising variables of different
-observational units. The data structure is also flexible to the
-techniques and packages analysts use to analyse the data, for example,
-in the matching example in section 4.3, users are free to use algorithms
-from another package to cluster stations. Further development and
-maintenance of the package involves responding to changes in the
-tidyverse packages that \pkg{cubble} imports, in particular,
-\pkg{tibble}, \pkg{tidyr}, and \pkg{dplyr}. Another area for further
-development is to extend \pkg{cubble} to other spatial objects other
-than points.
+\code{cubble} that builds from the \code{rowwise_df} and
+\code{grouped_df} class in the tidyverse ecosystem, is proposed to
+connect the time invariant and varying variables in the spatio-temporal
+data. This design frees the data analysts from spending time on
+organising variables of different observational units. The data
+structure is also flexible to the techniques and packages analysts use
+to analyse the data, for example, in the matching example in section
+4.3, users are free to use algorithms from another package to cluster
+stations. Further development and maintenance of the package involves
+responding to changes in the tidyverse packages that \pkg{cubble}
+imports, in particular, \pkg{tibble}, \pkg{tidyr}, and \pkg{dplyr}.
+Another area for further development is to extend \pkg{cubble} to other
+spatial objects other than points.
 

### 2022-02-07 2-the-cubble-pkg.Rmd
some changes as per meeting today

@@
 Spatio-temporal data usually come to the analysts in various forms and Figure \ref{fig:illu-input} shows four examples of this.
+The first three are variations of tabular data that uses a single table (1) or multiple tables (2 & 3) while the fourth uses a cube or hyper-cube to to represent the data.
+An example of the fourth type is the NetCDF (Network Common Data Form), commonly used in the climatology community to deliver global mapping of atmosphere, ocean, and land.
+
 These four examples show the common characters of the spatio-temporal data: 1) a spatial identifier that uniquely identifies each site, 2) a temporal identifier that prescribes the time stamp each site is recorded, 3) Coordinates, comprising of latitude and longitude, that locate the site on the map.
@@
 For those operations whose output involves both a spatial and temporal identifier, long form should be used.
-The long form uses a  \code{`grouped_df`} data frame to forms all the time of a site as a group.
+The long form uses a  \code{grouped_df} data frame to forms all the time of a site as a group.
 Time invariant variables are stored separately as an special attribute of the long cubble.
@@
 
-There are a few information in the cubble header: the name of the \code{key} variable (`id`) and its number (5), the bounding box: \code{[115.97, -32.94, 133.55, -12.42]} , and also the name of variable nested in the \code{ts} column with its type: \code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.
+There are a few information in the cubble header: the name of the \code{key} variable: \code{id} and its number: 5, the bounding box: \code{[115.97, -32.94, 133.55, -12.42]} , and also the name of variable nested in the \code{ts} column with its type: \code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.
 
@@
 It first extracts all the spatial variables into a separate tibble and then unnests the \code{ts} column.
-The spatial variables are then attached as an attributes of the long cubble and can be accessed through `spatial()`
+The spatial variables are then attached as an attributes of the long cubble and can be accessed through `spatial()`:
 
@@
 
-Analysts may need the time invariant/ varying variable to be in the same table to apply certain function.
+Analysts may need the time invariant and varying variable to be in the same table to apply certain function.
 In cubble, this can be done through \code{migrate()}, which moves the time invariant variables into the long cubble:
@@
 Notice that these added time invariant variables will not be remembered by cubble and will disappear if switched to the nested form and then switched back.
-Hence, \code{migrat} should usually be used as the last step of the analysis:
+Hence, \code{migrate()} should usually be used as the last step of the analysis:
 
@@
 
+### dplyr
+
 Building from an \code{tbl_df} class, cubble has implemented methods for \pkg{dplyr} generics, which includes:
@@
 
-\pkg{cubble} is also compatible with \pkg{tsibble}. When creating a cubble from a tsibble, Only the \code{coords} argument need to be specified:
+### tsibble
+
+Cubble is also compatible with \pkg{tsibble}. When creating a cubble from a tsibble, Only the \code{coords} argument needs to be specified:
 
@@
 ```
+
+### sf
+
+\textcolor{red}{add content}
+
+### netcdf
+
+NetCDF data has two main components: *dimension* for defining the spatio-temporal grid (longitude, latitude, and time) and *variable* that populates the defined grid.
+Attributes are usually associated with dimension and variable in the NetCDF format data and a [metadata convention for climate and forecast](http://cfconventions.org/) has been designed to standardise the format of the attributes.
+A few packages in R exists for manipulating NetCDF data and this includes a high-level R interface: \pkg{ncdf4} [@ncdf4], a low-level interface that calls C-interface: \pkg{RNetCDF} [@rnetcdf; @michna2013rnetcdf], and a tidyverse implementation: \pkg{tidync} [@tidync].
+
+Cubble provides an \code{as_cubble()} method to coerce the \code{ncdf4} class from the \pkg{ncdf4} package into a \code{cubble}.
+It maps each combination of longitude and latitude into an \code{id} as the \code{key}:
+
+```{r echo = TRUE, eval = FALSE}
+# read in the .nc file as a ncdf4 class
+raw <- ncdf4::nc_open(here::here("data/era5-pressure.nc"))
+
+# convert the variable q and z in the ncdf4 into a cubble
+dt <- as_cubble(raw, vars = c("q", "z"))
+```
+
+Memory limit with NetCDF data in cubble depends on longitude grid point x latitude grid point x time grid point x number of variable.
+Cubble can handle slightly more than 300 x 300 (longitude x longitude) grid points for 3 variables in one year and spatial grid can be reduced to trade for longer time period and more variables.
+A 300 by 300 spatial grid can be a bounding box of [100, -80, 180, 0] at 0.25 degree resolution or a global bounding box [-180, -90, 180, -90] at 1 degree resolution.
+Subsetting longitude and latitude grid is available through \code{long_range} and \code{lat_range} if the NetCDF file has finer resolution than needed.
+
+```{r eval = FALSE, echo = TRUE}
+# Assume my_ncdf has bounding box of [-180, -90, 180, -90] 
+# at 0.25 degree resolution and subset it to have 
+# 1 degree resolution:
+dt <- as_cubble(my_ncdf, vars = c("q", "z"), 
+                long_range = seq(-180, 180, 1), 
+                lat_range = seq(-90, 90, 1))
+```

### 2022-02-07 4-3-water-level.Rmd
some changes as per meeting today

@@
 
-## River level data in Victria water gauges
+## River level data in Victoria water gauges
 
@@
 Here \code{Water_course_level} in \code{river} will be matched to \code{prcp} in \code{climate} in 2020.
-The two datasets need to be specified as the first two arguments and the variable to match can be specified in \code{temporal_var_to_match} using the \code{by} syntax in \code{join}.
+The two datasets need to be specified as the first two arguments and the variable to match can be specified in \code{temporal_by} using the \code{by} syntax in \code{join}.
 \code{temporal_independent} controls the variable used to construct the interval and the goal here is to see if precipitation will be reflected by the water level in the river.
@@
   river, climate,
-  temporal_var_to_match = c("Water_course_level" = "prcp"),
+  temporal_by = c("Water_course_level" = "prcp"),
   temporal_independent = "prcp",  

### 2022-02-07 4-4-netcdf.Rmd
some changes as per meeting today

@@
 
-NetCDF (Network Common Data Form) is a data format commonly used in the climatology community to deliver global mapping of atmosphere, ocean, and land.
-It has two major components: *dimension* for defining the spatio-temporal grid (longitude, latitude, and time) and *variable* that populates the defined grid.
-Attributes are usually associated with dimension and variable in the NetCDF format data and a [metadata convention for climate and forecast](http://cfconventions.org/) has been designed to standardise the format of the attributes.
-Current packages for manipulating NetCDF data in R include a high-level R interface: \pkg{ncdf4} [@ncdf4], a low-level interface that calls C-interface: \pkg{RNetCDF} [@rnetcdf; @michna2013rnetcdf], and a tidyverse implementation: \pkg{tidync} [@tidync].
-
-Cubble provides an \code{as_cubble()} method to coerce the \code{ncdf4} class from the \pkg{ncdf4} package into a \code{cubble}.
-It maps each combination of longitude and latitude into an \code{id} as the \code{key}:
-
-```{r echo = TRUE, eval = FALSE}
-# read in the .nc file as a ncdf4 class
-raw <- ncdf4::nc_open(here::here("data/era5-pressure.nc"))
-
-# convert the variable q and z in the ncdf4 into a cubble
-dt <- as_cubble(raw, vars = c("q", "z"))
-```
-
-Memory limit with NetCDF data in cubble depends on longitude grid point x latitude grid point x time grid point x number of variable.
-Cubble can handle slightly more than 300 x 300 (longitude x longitude) grid points for 3 variables in one year and spatial grid can be reduced to trade for longer time period and more variables.
-A 300 by 300 spatial grid can be a bounding box of [100, -80, 180, 0] at 0.25 degree resolution or a global bounding box [-180, -90, 180, -90] at 1 degree resolution.
-Subsetting longitude and latitude grid is available through \code{long_range} and \code{lat_range} if the NetCDF file has finer resolution than needed.
-
-```{r eval = FALSE, echo = TRUE}
-# Assume my_ncdf has bounding box of [-180, -90, 180, -90] 
-# at 0.25 degree resolution and subset it to have 
-# 1 degree resolution:
-dt <- as_cubble(my_ncdf, vars = c("q", "z"), 
-                long_range = seq(-180, 180, 1), 
-                lat_range = seq(-90, 90, 1))
-```
-
-Figure \ref{fig:netcdf} gives an example of reproducing Figure 19 in @hersbach2020era5 using ERA5 data.
 ERA5 data [@hersbach2020era5] is the latest reanalysis of global atmosphere, land surface, and ocean waves from 1950 onwards and is available in the NetCDF format from European Centre for Medium-Range Weather Forecasts (ECMWF).
 The data can be directly downloaded from [Copernicus Climate Data Store (CDS)](https://cds.climate.copernicus.eu/cdsapp#!/dataset/reanalysis-era5-pressure-levels?tab=overview) website or programmatically via an R package \pkg{ecmwfr} [@ecwmfr].
-Variable *specific humidity* and *geopotential* are queried on the 10 hPa pressure level on four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04.
+The \code{era5-pressure} data contains variable *specific humidity* and *geopotential* on the 10 hPa pressure level on four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04.
 Once downloaded, the data can be read into a cubble as:
@@
 
-Readers who are interested in the analysis of the figure can refer to @simmons2020global and @simmons2005ecmwf for more details.
+Figure \ref{fig:netcdf} reproduces the ERA5 data row of Figure 19 in @hersbach2020era5.
+It shows the southern polar vortex splits into two on 2002-09-26 and further splits into four on 2002-10-04 in the stratosphere.
+Readers interested in the analysis of this figure can refer to @hersbach2020era5, @simmons2020global and @simmons2005ecmwf for more details.
 

### 2022-02-07 main.Rmd
some changes as per meeting today

@@
 In this paper, a new data structure, \code{cubble}, is proposed for manipulating and visualising spatio-temporal data.
-The new data structure uses two forms to organise time invariant/ varying variables so that the spatial and temporal dimension of the data can be manipulated separately while keep synchronised.
+The new data structure uses two forms to organise time invariant and varying variables so that the spatial and temporal dimension of the data can be manipulated separately while keep synchronised.
 Advanced considerations are given to illustrate data with hierarchical structure, data matching, and how cubble fits with the interactive graphics pipeline.

### 2022-02-07 main.tex
some changes as per meeting today

@@
 and visualising spatio-temporal data. The new data structure uses two
-forms to organise time invariant/ varying variables so that the spatial
-and temporal dimension of the data can be manipulated separately while
-keep synchronised. Advanced considerations are given to illustrate data
-with hierarchical structure, data matching, and how cubble fits with the
-interactive graphics pipeline. Examples are given to analysing Australia
-meteorology data, river level data, and climate reanalysis (ERA5).
+forms to organise time invariant and varying variables so that the
+spatial and temporal dimension of the data can be manipulated separately
+while keep synchronised. Advanced considerations are given to illustrate
+data with hierarchical structure, data matching, and how cubble fits
+with the interactive graphics pipeline. Examples are given to analysing
+Australia meteorology data, river level data, and climate reanalysis
+(ERA5).
 }
@@
 Spatio-temporal data usually come to the analysts in various forms and
-Figure \ref{fig:illu-input} shows four examples of this. These four
-examples show the common characters of the spatio-temporal data: 1) a
-spatial identifier that uniquely identifies each site, 2) a temporal
-identifier that prescribes the time stamp each site is recorded, 3)
-Coordinates, comprising of latitude and longitude, that locate the site
-on the map. These identifiers will be the building blocks for cubble.
-Another way to structure the spatio-temporal data is to think of it as a
-data cube with three axes: id, time, and variable. Notice that in the
-time-variable face of this cube, time invariant variables would share
-the same value across all the time stamps, while time varying variables
-have values varying across time. (Maybe add a cube diagram here). This
-heterogeneity will be used to split the data into two forms in cubble.
+Figure \ref{fig:illu-input} shows four examples of this. The first three
+are variations of tabular data that uses a single table (1) or multiple
+tables (2 \& 3) while the fourth uses a cube or hyper-cube to to
+represent the data. An example of the fourth type is the NetCDF (Network
+Common Data Form), commonly used in the climatology community to deliver
+global mapping of atmosphere, ocean, and land.
+
+These four examples show the common characters of the spatio-temporal
+data: 1) a spatial identifier that uniquely identifies each site, 2) a
+temporal identifier that prescribes the time stamp each site is
+recorded, 3) Coordinates, comprising of latitude and longitude, that
+locate the site on the map. These identifiers will be the building
+blocks for cubble. Another way to structure the spatio-temporal data is
+to think of it as a data cube with three axes: id, time, and variable.
+Notice that in the time-variable face of this cube, time invariant
+variables would share the same value across all the time stamps, while
+time varying variables have values varying across time. (Maybe add a
+cube diagram here). This heterogeneity will be used to split the data
+into two forms in cubble.
 
@@
 identifier, long form should be used. The long form uses a
-\code{`grouped_df`} data frame to forms all the time of a site as a
-group. Time invariant variables are stored separately as an special
-attribute of the long cubble. This design avoids repeating the spatial
-variables at each time stamp while not dropping information from spatial
+\code{grouped_df} data frame to forms all the time of a site as a group.
+Time invariant variables are stored separately as an special attribute
+of the long cubble. This design avoids repeating the spatial variables
+at each time stamp while not dropping information from spatial
 variables.
@@
 There are a few information in the cubble header: the name of the
-\code{key} variable (\texttt{id}) and its number (5), the bounding box:
+\code{key} variable: \code{id} and its number: 5, the bounding box:
 \code{[115.97, -32.94, 133.55, -12.42]} , and also the name of variable
@@
 an attributes of the long cubble and can be accessed through
-\texttt{spatial()}
+\texttt{spatial()}:
 
@@
 
-Analysts may need the time invariant/ varying variable to be in the same
-table to apply certain function. In cubble, this can be done through
-\code{migrate()}, which moves the time invariant variables into the long
-cubble:
+Analysts may need the time invariant and varying variable to be in the
+same table to apply certain function. In cubble, this can be done
+through \code{migrate()}, which moves the time invariant variables into
+the long cubble:
 
@@
 by cubble and will disappear if switched to the nested form and then
-switched back. Hence, \code{migrat} should usually be used as the last
-step of the analysis:
+switched back. Hence, \code{migrate()} should usually be used as the
+last step of the analysis:
 
@@
 
+\hypertarget{dplyr}{%
+\subsubsection{dplyr}\label{dplyr}}
+
 Building from an \code{tbl_df} class, cubble has implemented methods for
@@
 
-\pkg{cubble} is also compatible with \pkg{tsibble}. When creating a
-cubble from a tsibble, Only the \code{coords} argument need to be
-specified:
+\hypertarget{tsibble}{%
+\subsubsection{tsibble}\label{tsibble}}
+
+Cubble is also compatible with \pkg{tsibble}. When creating a cubble
+from a tsibble, Only the \code{coords} argument needs to be specified:
 
@@
 
-\hypertarget{advanced-features-considerations}{%
-\section{Advanced features/
-considerations}\label{advanced-features-considerations}}
+\hypertarget{sf}{%
+\subsubsection{sf}\label{sf}}
+
+\textcolor{red}{add content}
+
+\hypertarget{netcdf}{%
+\subsubsection{netcdf}\label{netcdf}}
+
+NetCDF data has two main components: \emph{dimension} for defining the
+spatio-temporal grid (longitude, latitude, and time) and \emph{variable}
+that populates the defined grid. Attributes are usually associated with
+dimension and variable in the NetCDF format data and a
+\href{http://cfconventions.org/}{metadata convention for climate and
+forecast} has been designed to standardise the format of the attributes.
+A few packages in R exists for manipulating NetCDF data and this
+includes a high-level R interface: \pkg{ncdf4} \citep{ncdf4}, a
+low-level interface that calls C-interface: \pkg{RNetCDF}
+\citep{rnetcdf, michna2013rnetcdf}, and a tidyverse implementation:
+\pkg{tidync} \citep{tidync}.
+
+Cubble provides an \code{as_cubble()} method to coerce the \code{ncdf4}
+class from the \pkg{ncdf4} package into a \code{cubble}. It maps each
+combination of longitude and latitude into an \code{id} as the
+\code{key}:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> # read in the .nc file as a ncdf4 class
+R> raw <- ncdf4::nc_open(here::here("data/era5-pressure.nc"))
+R> 
+R> # convert the variable q and z in the ncdf4 into a cubble
+R> dt <- as_cubble(raw, vars = c("q", "z"))
+\end{CodeInput}
+\end{CodeChunk}
+
+Memory limit with NetCDF data in cubble depends on longitude grid point
+x latitude grid point x time grid point x number of variable. Cubble can
+handle slightly more than 300 x 300 (longitude x longitude) grid points
+for 3 variables in one year and spatial grid can be reduced to trade for
+longer time period and more variables. A 300 by 300 spatial grid can be
+a bounding box of {[}100, -80, 180, 0{]} at 0.25 degree resolution or a
+global bounding box {[}-180, -90, 180, -90{]} at 1 degree resolution.
+Subsetting longitude and latitude grid is available through
+\code{long_range} and \code{lat_range} if the NetCDF file has finer
+resolution than needed.
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> # Assume my_ncdf has bounding box of [-180, -90, 180, -90] 
+R> # at 0.25 degree resolution and subset it to have 
+R> # 1 degree resolution:
+R> dt <- as_cubble(my_ncdf, vars = c("q", "z"), 
++                 long_range = seq(-180, 180, 1), 
++                 lat_range = seq(-90, 90, 1))
+\end{CodeInput}
+\end{CodeChunk}
+
+\hypertarget{other-features-and-considerations}{%
+\section{Other features and
+considerations}\label{other-features-and-considerations}}
 
@@
 
-\hypertarget{glyph-map}{%
-\subsection{Glyph map}\label{glyph-map}}
+\hypertarget{spatio-temporal-transformations}{%
+\subsection{Spatio-temporal
+transformations}\label{spatio-temporal-transformations}}
+
+\textcolor{red}{Glyph map + transforming on time (calendar plot) + projections (linear combination).}
 
@@
 
-\hypertarget{river-level-data-in-victria-water-gauges}{%
-\subsection{River level data in Victria water
-gauges}\label{river-level-data-in-victria-water-gauges}}
+\hypertarget{river-level-data-in-victoria-water-gauges}{%
+\subsection{River level data in Victoria water
+gauges}\label{river-level-data-in-victoria-water-gauges}}
 
@@
 The two datasets need to be specified as the first two arguments and the
-variable to match can be specified in \code{temporal_var_to_match} using
-the \code{by} syntax in \code{join}. \code{temporal_independent}
-controls the variable used to construct the interval and the goal here
-is to see if precipitation will be reflected by the water level in the
-river. This puts precipitation \code{prcp}, as the independent. Given
-there is one year worth of data, the number of peak
-(\code{temporal_n_highest}) to consider is slightly raised from a
-default 20 to 30 and \code{temporal_min_match} is raised
-correspondingly. To return all the pairs of the match,
-\code{temporal_min_match} can be set to 0.
+variable to match can be specified in \code{temporal_by} using the
+\code{by} syntax in \code{join}. \code{temporal_independent} controls
+the variable used to construct the interval and the goal here is to see
+if precipitation will be reflected by the water level in the river. This
+puts precipitation \code{prcp}, as the independent. Given there is one
+year worth of data, the number of peak (\code{temporal_n_highest}) to
+consider is slightly raised from a default 20 to 30 and
+\code{temporal_min_match} is raised correspondingly. To return all the
+pairs of the match, \code{temporal_min_match} can be set to 0.
 
@@
 +   river, climate,
-+   temporal_var_to_match = c("Water_course_level" = "prcp"),
++   temporal_by = c("Water_course_level" = "prcp"),
 +   temporal_independent = "prcp",  
@@
 
-NetCDF (Network Common Data Form) is a data format commonly used in the
-climatology community to deliver global mapping of atmosphere, ocean,
-and land. It has two major components: \emph{dimension} for defining the
-spatio-temporal grid (longitude, latitude, and time) and \emph{variable}
-that populates the defined grid. Attributes are usually associated with
-dimension and variable in the NetCDF format data and a
-\href{http://cfconventions.org/}{metadata convention for climate and
-forecast} has been designed to standardise the format of the attributes.
-Current packages for manipulating NetCDF data in R include a high-level
-R interface: \pkg{ncdf4} \citep{ncdf4}, a low-level interface that calls
-C-interface: \pkg{RNetCDF} \citep{rnetcdf, michna2013rnetcdf}, and a
-tidyverse implementation: \pkg{tidync} \citep{tidync}.
-
-Cubble provides an \code{as_cubble()} method to coerce the \code{ncdf4}
-class from the \pkg{ncdf4} package into a \code{cubble}. It maps each
-combination of longitude and latitude into an \code{id} as the
-\code{key}:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> # read in the .nc file as a ncdf4 class
-R> raw <- ncdf4::nc_open(here::here("data/era5-pressure.nc"))
-R> 
-R> # convert the variable q and z in the ncdf4 into a cubble
-R> dt <- as_cubble(raw, vars = c("q", "z"))
-\end{CodeInput}
-\end{CodeChunk}
-
-Memory limit with NetCDF data in cubble depends on longitude grid point
-x latitude grid point x time grid point x number of variable. Cubble can
-handle slightly more than 300 x 300 (longitude x longitude) grid points
-for 3 variables in one year and spatial grid can be reduced to trade for
-longer time period and more variables. A 300 by 300 spatial grid can be
-a bounding box of {[}100, -80, 180, 0{]} at 0.25 degree resolution or a
-global bounding box {[}-180, -90, 180, -90{]} at 1 degree resolution.
-Subsetting longitude and latitude grid is available through
-\code{long_range} and \code{lat_range} if the NetCDF file has finer
-resolution than needed.
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> # Assume my_ncdf has bounding box of [-180, -90, 180, -90] 
-R> # at 0.25 degree resolution and subset it to have 
-R> # 1 degree resolution:
-R> dt <- as_cubble(my_ncdf, vars = c("q", "z"), 
-+                 long_range = seq(-180, 180, 1), 
-+                 lat_range = seq(-90, 90, 1))
-\end{CodeInput}
-\end{CodeChunk}
-
-Figure \ref{fig:netcdf} gives an example of reproducing Figure 19 in
-\citet{hersbach2020era5} using ERA5 data. ERA5 data
-\citep{hersbach2020era5} is the latest reanalysis of global atmosphere,
-land surface, and ocean waves from 1950 onwards and is available in the
-NetCDF format from European Centre for Medium-Range Weather Forecasts
-(ECMWF). The data can be directly downloaded from
+ERA5 data \citep{hersbach2020era5} is the latest reanalysis of global
+atmosphere, land surface, and ocean waves from 1950 onwards and is
+available in the NetCDF format from European Centre for Medium-Range
+Weather Forecasts (ECMWF). The data can be directly downloaded from
 \href{https://cds.climate.copernicus.eu/cdsapp\#!/dataset/reanalysis-era5-pressure-levels?tab=overview}{Copernicus
 Climate Data Store (CDS)} website or programmatically via an R package
-\pkg{ecmwfr} \citep{ecwmfr}. Variable \emph{specific humidity} and
-\emph{geopotential} are queried on the 10 hPa pressure level on four
-dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04. Once
-downloaded, the data can be read into a cubble as:
+\pkg{ecmwfr} \citep{ecwmfr}. The \code{era5-pressure} data contains
+variable \emph{specific humidity} and \emph{geopotential} on the 10 hPa
+pressure level on four dates: 2002-09-22, 2002-09-26, 2002-09-30, and
+2002-10-04. Once downloaded, the data can be read into a cubble as:
 
@@
 
-Readers who are interested in the analysis of the figure can refer to
-\citet{simmons2020global} and \citet{simmons2005ecmwf} for more details.
+Figure \ref{fig:netcdf} reproduces the ERA5 data row of Figure 19 in
+\citet{hersbach2020era5}. It shows the southern polar vortex splits into
+two on 2002-09-26 and further splits into four on 2002-10-04 in the
+stratosphere. Readers interested in the analysis of this figure can
+refer to \citet{hersbach2020era5}, \citet{simmons2020global} and
+\citet{simmons2005ecmwf} for more details.
 
@@
 4.3, users are free to use algorithms from another package to cluster
-stations. Further development and maintenance of the package involves
-responding to changes in the tidyverse packages that \pkg{cubble}
-imports, in particular, \pkg{tibble}, \pkg{tidyr}, and \pkg{dplyr}.
-Another area for further development is to extend \pkg{cubble} to other
-spatial objects other than points.
+stations.
+
+Further development and maintenance of the package involves responding
+to changes in the tidyverse packages that \pkg{cubble} imports, in
+particular, \pkg{tibble}, \pkg{tidyr}, and \pkg{dplyr}. Another area for
+further development is to extend \pkg{cubble} to other spatial objects
+other than points.
 

### 2022-02-08 2-the-cubble-pkg.Rmd
add transformation of calendar plot and projection

@@
 
-### sf
+### sf and s2
 
-\textcolor{red}{add content}
+Cubble is compatible with simple feature geometry list-column (\code{sfc}) in \pkg{sf} and  \code{s2_lnglat} vector in \pkg{S2} in the nested form.
+After creating a cubble, analysts can add the geometry vectors as needed to perform spatial operations:
+
+```{r echo = TRUE}
+library(s2)
+# create a cubble
+cb <- climate_flat %>% 
+  cubble::as_cubble(key = id, index = date, coords = c(long, lat))
+
+# western australia map
+wa <- ozmaps::abs_ste %>% filter(NAME == "Western Australia")
+
+# create a `s2_lnglat` vector based on long and lat and filter stations in wa
+cb %>%
+  mutate(ll = s2_lnglat(long, lat)) %>% 
+  filter(s2_within(ll, wa))
+```
+
+```{r echo = TRUE}
+library(sf)
+# simplify australia map
+aus <- ozmaps::abs_ste %>% rmapshaper::ms_simplify(keep = 0.01)
+
+dt <- cb %>%
+  mutate(
+    # create `sfc` column based on long and lat
+    ll = st_sfc(
+      purrr::map2(long, lat, ~st_point(c(.x, .y))),
+      crs = st_crs(aus)),
+    # append state multi-polygon based on the `sfc` created
+    state = aus$geometry[st_within(ll, aus, sparse = FALSE)],
+    # adopt a different projection: lambert conformal conic (EPSG:3112)
+    state = st_transform(state, crs = "EPSG:3112"))
+
+dt
+```
 

### 2022-02-08 3-advanced.Rmd
some changes as per meeting today
add transformation of calendar plot and projection

@@
 
-# Advanced features/ considerations
+# Other features and considerations
 
@@
 
-## Glyph map
+## Spatio-temporal transformations
 
-Glyph map [@Wickham2012-yr] plots the time series as single glyph on the map and in \proglang{R}, \pkg{GGally} implements the glyph map through the \code{glyphs()} function.
+Spatio-temporal transformation is a useful technique to extract information form spatio-temporal data.
+Glyph maps [@Wickham2012-yr] transform the time coordinates into the space coordinates to plot the time series of different locations on the map.
+Calendar plots [@wang2020calendar] reconstructs time into calendar-based grid to discover weekday and weekend pattern.
+Projection, or linear combination, of variables summarises multivariate information into lower dimension to further digest.
+This section elaborates on the glyph map.
+
+In \proglang{R}, \pkg{GGally} implements glyph maps through the \code{glyphs()} function.
 The function constructs a data frame with calculated position (\code{gx}.

### 2022-02-08 main.tex
add transformation of calendar plot and projection

@@
 
-\hypertarget{sf}{%
-\subsubsection{sf}\label{sf}}
-
-\textcolor{red}{add content}
+\hypertarget{sf-and-s2}{%
+\subsubsection{sf and s2}\label{sf-and-s2}}
+
+Cubble is compatible with simple feature geometry list-column
+(\code{sfc}) in \pkg{sf} and \code{s2_lnglat} vector in \pkg{S2} in the
+nested form. After creating a cubble, analysts can add the geometry
+vectors as needed to perform spatial operations:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> library(s2)
+R> # create a cubble
+R> cb <- climate_flat %>% 
++   cubble::as_cubble(key = id, index = date, coords = c(long, lat))
+R> 
+R> # western australia map
+R> wa <- ozmaps::abs_ste %>% filter(NAME == "Western Australia")
+R> 
+R> # create a `s2_lnglat` vector based on long and lat and filter stations in wa
+R> cb %>%
++   mutate(ll = s2_lnglat(long, lat)) %>% 
++   filter(s2_within(ll, wa))
+\end{CodeInput}
+\begin{CodeOutput}
+# cubble:   id [3]: nested form
+# bbox:     [115.97, -32.94, 117.18, -31.89]
+# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
+  id            lat  long  elev name          wmo_id ts                 ll      
+  <chr>       <dbl> <dbl> <dbl> <chr>          <dbl> <list>             <s2_lng>
+1 ASN00009021 -31.9  116.  15.4 perth airport  94610 <tibble [366 x 4]> (115.97~
+2 ASN00010311 -31.9  117. 179   york           94623 <tibble [366 x 4]> (116.76~
+3 ASN00010614 -32.9  117. 338   narrogin       94627 <tibble [366 x 4]> (117.17~
+\end{CodeOutput}
+\end{CodeChunk}
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> library(sf)
+R> # simplify australia map
+R> aus <- ozmaps::abs_ste %>% rmapshaper::ms_simplify(keep = 0.01)
+R> 
+R> dt <- cb %>%
++   mutate(
++     # create `sfc` column based on long and lat
++     ll = st_sfc(
++       purrr::map2(long, lat, ~st_point(c(.x, .y))),
++       crs = st_crs(aus)),
++     # append state multi-polygon based on the `sfc` created
++     state = aus$geometry[st_within(ll, aus, sparse = FALSE)],
++     # adopt a different projection: lambert conformal conic (EPSG:3112)
++     state = st_transform(state, crs = "EPSG:3112"))
+R> 
+R> dt
+\end{CodeInput}
+\begin{CodeOutput}
+# cubble:   id [5]: nested form
+# bbox:     [115.97, -32.94, 133.55, -12.42]
+# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
+  id            lat  long  elev name         wmo_id ts                        ll
+  <chr>       <dbl> <dbl> <dbl> <chr>         <dbl> <list>           <POINT [°]>
+1 ASN00009021 -31.9  116.  15.4 perth airpo~  94610 <tibble~ (115.9764 -31.9275)
+2 ASN00010311 -31.9  117. 179   york          94623 <tibble~  (116.765 -31.8997)
+3 ASN00010614 -32.9  117. 338   narrogin      94627 <tibble~ (117.1797 -32.9342)
+4 ASN00014015 -12.4  131.  30.4 darwin airp~  94120 <tibble~ (130.8925 -12.4239)
+5 ASN00015131 -17.6  134. 220   elliott       94236 <tibble~ (133.5407 -17.5521)
+# ... with 1 more variable: state <MULTIPOLYGON [m]>
+\end{CodeOutput}
+\end{CodeChunk}
 
@@
 
-\textcolor{red}{Glyph map + transforming on time (calendar plot) + projections (linear combination).}
-
-Glyph map \citep{Wickham2012-yr} plots the time series as single glyph
-on the map and in \proglang{R}, \pkg{GGally} implements the glyph map
-through the \code{glyphs()} function. The function constructs a data
-frame with calculated position (\code{gx}. \code{gy}. \code{gid}) of
-each point on the time series using linear algebra (Equation 1 and 2 in
+Spatio-temporal transformation is a useful technique to extract
+information form spatio-temporal data. Glyph maps \citep{Wickham2012-yr}
+transform the time coordinates into the space coordinates to plot the
+time series of different locations on the map. Calendar plots
+\citep{wang2020calendar} reconstructs time into calendar-based grid to
+discover weekday and weekend pattern. Projection, or linear combination,
+of variables summarises multivariate information into lower dimension to
+further digest. This section elaborates on the glyph map.
+
+In \proglang{R}, \pkg{GGally} implements glyph maps through the
+\code{glyphs()} function. The function constructs a data frame with
+calculated position (\code{gx}. \code{gy}. \code{gid}) of each point on
+the time series using linear algebra (Equation 1 and 2 in
 \citet{Wickham2012-yr}). The data can then be piped into \code{ggplot}

### 2022-02-08 main.tex
clean up bibbib

@@
 domain to pair up nearby observations. It can also be considered as a
-form of data fusion \citep{castanedo2013review, cocchi2019data}, where
-data from multiple sources are combined together.
+form of data fusion \citep{cocchi2019data}, where data from multiple
+sources are combined together.
 
@@
 Figure \ref{fig:interactive-popup} shows Figure
-\ref{fig:interactive-linking} made made with leaflet and popups.
+\ref{fig:interactive-linking} made made with leaflet and popups
+\citep{leafpop}.
 
@@
 
+\newpage
+
 \hypertarget{appendix}{%
@@
 
+\newpage
+
 \bibliography{references.bib}

### 2022-02-13 4-3-water-level.Rmd
a few small changes as per writing the vignette

@@
 This puts precipitation \code{prcp}, as the independent.
-Given there is one year worth of data, the number of peak (\code{temporal_n_highest}) to consider is slightly raised from a default 20 to 30 and \code{temporal_min_match} is raised correspondingly.
-To return all the pairs of the match, \code{temporal_min_match} can be set to 0.
+Given there is one year worth of data, the number of peak (\code{temporal_n_highest}) to consider is slightly raised from a default 20 to 30 and \code{temporal_min_match} is raised accordingly To return all the pairs of the match, \code{temporal_min_match} can be set to 0.
 
@@
 
-Figure \ref{fig:matching} plots the matched pairs on the map or to view the matched series and with the chosen parameter, there are four pairs of matches, which all locates in the middle Victoria and the concurrent increase of precipitation and water level can be observed.
+Figure \ref{fig:matching} plots the matched pairs on the map or to view the matched series.
+There are four pairs of matches, which all locates in the middle Victoria and the concurrent increase of precipitation and water level can be observed.
 

### 2022-02-13 main.tex
minor al layout adjustment
a few small changes as per writing the vignette

@@
 consider is slightly raised from a default 20 to 30 and
-\code{temporal_min_match} is raised correspondingly. To return all the
-pairs of the match, \code{temporal_min_match} can be set to 0.
+\code{temporal_min_match} is raised accordingly To return all the pairs
+of the match, \code{temporal_min_match} can be set to 0.
 
@@
 Figure \ref{fig:matching} plots the matched pairs on the map or to view
-the matched series and with the chosen parameter, there are four pairs
-of matches, which all locates in the middle Victoria and the concurrent
-increase of precipitation and water level can be observed.
+the matched series. There are four pairs of matches, which all locates
+in the middle Victoria and the concurrent increase of precipitation and
+water level can be observed.
 
@@
 
+\newpage
+
 \hypertarget{acknowledgement}{%
@@
 
-\newpage
-
 \hypertarget{appendix}{%

### 2022-02-13 scripts/basic-agg.R

@@ -24,9 +24,9 @@ station_long <- station_nested %>%
 cluster_long <- cluster_nested %>% 
   stretch(ts) %>% 
   mutate(wk = lubridate::week(date)) %>% 
-  group_by(id, wk) %>% 
-  summarise(prcp = sum(prcp, na.rm = TRUE)) %>% 
-  ungroup(id) %>% 
+  group_by(id, wk) %>%
+  summarise(prcp = sum(prcp, na.rm = TRUE)) %>%
+  ungroup(id) %>%
   summarise(prcp = mean(prcp, na.rm = TRUE)) %>% 
   migrate(cent_long, cent_lat)
 
@@ -34,7 +34,7 @@ state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)
 
 ggplot_smooth <- cluster_long %>% 
   ggplot() +
-  geom_smooth(aes(x = wk, y = prcp, group = cluster)) 
+  geom_smooth(aes(x = wk, y = prcp, group = cluster), span = 0.4) 
 
 smoother <- layer_data(ggplot_smooth) %>% 
   left_join(cluster_long %>% select(cluster, cent_long, cent_lat), by = c("group" = "cluster"))
@@ -48,11 +48,14 @@ p1 <- ggplot(data = smoother,
   geom_glyph(height = 2, width = 4) + 
   theme_void()
 
-p3 <- plot_map(state_map) +
+p3 <- ggplot() + 
+  geom_sf(data = state_map, inherit.aes = FALSE,
+          color = "grey80", alpha = 0.4, linetype = 3) + 
   geom_point(data = station_nested, aes(x = long, y = lat), size = 0.5) +
   ggforce::geom_mark_hull(data = cluster_nested %>% tidyr::unnest(hull),
                           expand = 0, radius = 0,
-                          aes(x = long, y = lat, group = cluster)) 
+                          aes(x = long, y = lat, group = cluster)) + 
+  theme_void()
 
 (p1 | p3) + plot_annotation(tag_levels = 'A')
 ggsave(filename = "figures/basic-agg.png", width = 15, height = 7)

### 2022-02-13 scripts/matching.R

@@ -15,14 +15,11 @@ vic_map <- rmapshaper::ms_simplify(ozmaps::abs_ste %>% filter(NAME == "Victoria"
 plot_map(vic_map) +
   geom_point(data = dplyr::bind_rows(river, climate), 
              aes(x = long, y = lat, color = type)) + 
-  scale_color_brewer(palette = "Dark2") +
-  theme_bw() + 
-  theme(legend.position = "bottom") + 
-  labs(x = "Longitude", y = "Latitude")
+  scale_color_brewer(palette = "Dark2") 
 ggsave(filename = "figures/matching-map.png", width = 10, height = 5)  
 
 res <- match_sites(river, climate,
-                   temporal_var_to_match = c("Water_course_level" = "prcp"),
+                   temporal_by = c("Water_course_level" = "prcp"),
                    temporal_independent = "prcp",  
                    temporal_n_highest = 30,
                    temporal_min_match = 15)
@@ -31,13 +28,7 @@ p1 <- plot_map(vic_map) +
   geom_point(data = res, aes(x = long, y = lat, color = type)) +
   ggrepel::geom_label_repel(
     data = res %>% filter(type == "river"), aes(x = long, y = lat, label = group)) +
-  scale_color_brewer(palette = "Dark2") +
-  theme_minimal() +
-  theme(
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank()
-  ) + 
-  labs(x = "Longitude", y = "Latitude")
+  scale_color_brewer(palette = "Dark2") 
 
 res_long <- res %>%
   stretch(ts) %>%
@@ -51,9 +42,8 @@ p2 <- res_long %>%
   facet_wrap(vars(group)) +
   scale_color_brewer(palette = "Dark2", guide = "none") +
   theme_bw() +
-  labs(x = "week") +
   scale_x_date(date_labels = "%b") + 
-  labs(x = "Week", y = "Precipitation/ water level")
+  labs(x = "Date", y = "Precipitation/ water level")
 
 (p1 | p2) + 
   patchwork::plot_layout(guides = "collect") + 

### 2022-02-15 1-intro.Rmd
merge general information of spatio-temporal data  in intro and section 2

@@
 
-Spatio-temporal data concerns variables in the spatial coordinates across time.
-Some variables are invariant to time and can be solely identified by its spatial location while other time varying variables need both the spatial and temporal identifiers to identify.
-Manipulating data in these two levels is usually done separately and the third tidy data principal [@tidydata] recommends each type of observational unit to form a table.
+Spatio-temporal data concerns variables in the space across time.
+\textcolor{red}{Elaborate on variable, space, and time}.
+If the time interval is regular and each site has fixed point location, it can also be thought of as a data cube with three axes: spatial identifier, temporal identifier, and variable.
+In this cubical layout, variables can be those that are invariant to time, i.e.
+the longitude and latitude coordinates, or those that varying across time.
+While conceptually the data format is relatively simple, the actual data object that comes to the analysts can be in various forms in practice.
+Figure \ref{fig:illu-input} shows four examples where the first one combines the time invariant and varying variables into a single table through duplicating the time invariant variables; 2) splits those variables into two separate sheets; 3) ...; and lastly 4) considers the array-based data, like NetCDF (Network Common Data Form) data, which is commonly used in earth science, e.g.
+climatology, meteorology, and oceanography, among others,  to organise multivariate data in a spatio-temporal grid.
+
+\textcolor{red}{think through a cube of long x lat x time vs. id x time x variable}
+
+```{r illu-input, echo = FALSE, out.width = "100%", fig.cap="An illustration of different spatio-temporal data format: (1) a single table for all the variables; (2) separate tables for spatial and temporal variables; (3) a table of metadata used to query the database and a separate table for queried data; and (4) an array data structure of cube or hypercube.", fig.align='center', out.height = "40%"}
+knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.001.png"))
+```
+
+<!-- Manipulating data in these two levels is usually done separately and the third tidy data principal [@tidydata] recommends each type of observational unit to form a table.
 Another feature for a spatio-temporal data structure to have is that variables created from subsequent operations can be automatically added to the existing table with the same level.
 This is better than an output that index the variables with the identifiers since otherwise analysts will need to manually join this output with existing variables in the same level.
+-->
 
@@
 
-Currently, available spatio-temporal data structures in R includes: \pkg{spacetime} [@spacetime], which proposes four space-time layouts: Full grid (STF), sparse grid(STS), irregular (STI), and trajectory (STT).
+\textcolor{red}{more beginning like "working with such data different authors ..."} Currently, there's various different spatio-temporal data structures in R includes: \pkg{spacetime} [@spacetime], which proposes four space-time layouts: Full grid (STF), sparse grid(STS), irregular (STI), and trajectory (STT).
 \pkg{spacetime} is built on \pkg{sp} [@sp] and \pkg{xts} [@xts], both of which have been replaced by more recent implementations.
 \pkg{spatstat} [@spatstat] implements a `ppp` class for point pattern data; and more recent, \pkg{stars} [@stars] implements a spatio-temporal for vector and raster spatio-temporal data.
-While these implementations either store spatial and temporal variables all in a single table, hence duplicate the spatial variables at each temporal unit; or split them into two separate tables that analysts need to manually join the data constantly when doing the analysis.
-This creates a gap in the software development.
+
+In spatio-temporal data analysis, analysts will receive one of the various data formats depends on the data provider and need to manually rearrange the data to fit into one of the data object in \proglang{R} before starting any analysis on the data itself.
+Looking at spatial or temporal data analysis, data objects like \pkg{sf} [@sf] and \pkg{tsibble} [@tsibble] have smoothed the data analysis in these two domains while there has not yet been a spatio-temporal data structure that makes it easy to explore spatio-temporal data and this creates a gap in the software development.
 The requirement for such a tool is important given the ubiquity of spatio-temporal vector data in the wild: climate observations from weather stations, air quality variables from air monitoring stations, or any variables that are measured at fixed location across time.
+
+<!-- While these implementations either store spatial and temporal variables all in a single table, hence duplicate the spatial variables at each temporal unit; or split them into two separate tables that analysts need to manually join the data constantly when doing the analysis.
+\textcolor{red}{There has been data structure proposed for spatial (\pkg{sf} by @sf) and temporal (\pkg{tsibble} by @tsibble) data in the R community, while less has been done for spatio-temporal data.}  -->
 

### 2022-02-15 2-the-cubble-pkg.Rmd
a few small changes as per writing the vignette
merge general information of spatio-temporal data  in intro and section 2

@@
 
-Spatio-temporal data usually come to the analysts in various forms and Figure \ref{fig:illu-input} shows four examples of this.
-The first three are variations of tabular data that uses a single table (1) or multiple tables (2 & 3) while the fourth uses a cube or hyper-cube to to represent the data.
-An example of the fourth type is the NetCDF (Network Common Data Form), commonly used in the climatology community to deliver global mapping of atmosphere, ocean, and land.
-
-These four examples show the common characters of the spatio-temporal data: 1) a spatial identifier that uniquely identifies each site, 2) a temporal identifier that prescribes the time stamp each site is recorded, 3) Coordinates, comprising of latitude and longitude, that locate the site on the map.
-These identifiers will be the building blocks for cubble.
-Another way to structure the spatio-temporal data is to think of it as a data cube with three axes: id, time, and variable.
-Notice that in the time-variable face of this cube, time invariant variables would share the same value across all the time stamps, while time varying variables have values varying across time.
-(Maybe add a cube diagram here).
-This heterogeneity will be used to split the data into two forms in cubble.
-
-```{r illu-input, echo = FALSE, out.width = "100%", fig.cap="An illustration of different spatio-temporal data format: (1) a single table for all the variables; (2) separate tables for spatial and temporal variables; (3) a table of metadata used to query the database and a separate table for queried data; and (4) an array data structure of cube or hypercube.", fig.align='center', out.height = "40%"}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.001.png"))
-```
-
-In a cubble, there are two forms: nested form and long form, and Figure \ref{fig:illu-cubble} sketches the two forms with the associated attributes.
+\textcolor{red}{more beginning} In cubble data can represented in two formats: nested form and long form, and Figure \ref{fig:illu-cubble} sketches the two forms with the associated attributes.
 The decision on which form to use in data manipulation is output-oriented, meaning analysts need to first think about whether the output of an operation is identified only by the spatial identifier, or a combination of spatial and temporal identifier.
@@
 ## Create a cubble in the nested form
-A cubble can be created with \code{as_cubble()} by supplying the three key components: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order of longitude and latitude.
+
+\textcolor{red}{more beginning} A cubble can be created with \code{as_cubble()} by supplying the three key components: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order of longitude and latitude.
 The naming of \code{key} and \code{index} follows the convention in the \pkg{tsibble} package.
-Below is an example of creating a cubble and by default, the cubble created is in the nested form:
+\textcolor{red}{give some context of the data about to be used: a subset of a larger data etc} Below is an example of creating a cubble and by default, the cubble created is in the nested form:
 
@@
 
+\textcolor{red}{A beginning sentence that summarises the subsection}. 
+\textcolor{red}{Given more details and examples when going through each subsection}
+
 ### dplyr
@@
 
-Cubble is also compatible with \pkg{tsibble}. When creating a cubble from a tsibble, Only the \code{coords} argument needs to be specified:
+Cubble is also compatible with \pkg{tsibble}. When creating a cubble from a tsibble, only the \code{coords} argument needs to be specified:
 

### 2022-02-15 3-advanced.Rmd
clean up bibbib
merge general information of spatio-temporal data  in intro and section 2

@@
 
-Cubble provides functionality for matching sites from two data sources. To be a valid pair of matching, the matched pair from different data sources need to:
+\textcolor{red}{more beginning} Cubble provides functionality for matching sites from two data sources. To be a valid pair of matching, the matched pair from different data sources need to:
 
@@
 This functionality can be seen as a matching exercise [@stuart2010matching; @mcintosh2018using] in the spatio-temporal domain to pair up nearby observations.
-It can also be considered as a form of data fusion [@castanedo2013review; @cocchi2019data], where data from multiple sources are combined together.
+It can also be considered as a form of data fusion [@cocchi2019data], where data from multiple sources are combined together.
 

### 2022-02-15 main.tex
merge general information of spatio-temporal data  in intro and section 2

@@
 
-Spatio-temporal data concerns variables in the spatial coordinates
-across time. Some variables are invariant to time and can be solely
-identified by its spatial location while other time varying variables
-need both the spatial and temporal identifiers to identify. Manipulating
-data in these two levels is usually done separately and the third tidy
-data principal \citep{tidydata} recommends each type of observational
-unit to form a table. Another feature for a spatio-temporal data
-structure to have is that variables created from subsequent operations
-can be automatically added to the existing table with the same level.
-This is better than an output that index the variables with the
-identifiers since otherwise analysts will need to manually join this
-output with existing variables in the same level.
-
-Currently, available spatio-temporal data structures in R includes:
-\pkg{spacetime} \citep{spacetime}, which proposes four space-time
-layouts: Full grid (STF), sparse grid(STS), irregular (STI), and
-trajectory (STT). \pkg{spacetime} is built on \pkg{sp} \citep{sp} and
-\pkg{xts} \citep{xts}, both of which have been replaced by more recent
-implementations. \pkg{spatstat} \citep{spatstat} implements a
+Spatio-temporal data concerns variables in the space across time.
+\textcolor{red}{Elaborate on variable, space, and time}. If the time
+interval is regular and each site has fixed point location, it can also
+be thought of as a data cube with three axes: spatial identifier,
+temporal identifier, and variable. In this cubical layout, variables can
+be those that are invariant to time, i.e.~the longitude and latitude
+coordinates, or those that varying across time. While conceptually the
+data format is relatively simple, the actual data object that comes to
+the analysts can be in various forms in practice. Figure
+\ref{fig:illu-input} shows four examples where the first one combines
+the time invariant and varying variables into a single table through
+duplicating the time invariant variables; 2) splits those variables into
+two separate sheets; 3) \ldots; and lastly 4) considers the array-based
+data, like NetCDF (Network Common Data Form) data, which is commonly
+used in earth science, e.g.~climatology, meteorology, and oceanography,
+among others, to organise multivariate data in a spatio-temporal grid.
+
+\textcolor{red}{think through a cube of long x lat x time vs. id x time x variable}
+
+\begin{CodeChunk}
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.001}
+
+}
+
+\caption[An illustration of different spatio-temporal data format]{An illustration of different spatio-temporal data format: (1) a single table for all the variables; (2) separate tables for spatial and temporal variables; (3) a table of metadata used to query the database and a separate table for queried data; and (4) an array data structure of cube or hypercube.}\label{fig:illu-input}
+\end{figure}
+\end{CodeChunk}
+
+\textcolor{red}{more beginning like "working with such data different authors ..."}
+Currently, there's various different spatio-temporal data structures in
+R includes: \pkg{spacetime} \citep{spacetime}, which proposes four
+space-time layouts: Full grid (STF), sparse grid(STS), irregular (STI),
+and trajectory (STT). \pkg{spacetime} is built on \pkg{sp} \citep{sp}
+and \pkg{xts} \citep{xts}, both of which have been replaced by more
+recent implementations. \pkg{spatstat} \citep{spatstat} implements a
 \texttt{ppp} class for point pattern data; and more recent, \pkg{stars}
 \citep{stars} implements a spatio-temporal for vector and raster
-spatio-temporal data. While these implementations either store spatial
-and temporal variables all in a single table, hence duplicate the
-spatial variables at each temporal unit; or split them into two separate
-tables that analysts need to manually join the data constantly when
-doing the analysis. This creates a gap in the software development. The
-requirement for such a tool is important given the ubiquity of
-spatio-temporal vector data in the wild: climate observations from
-weather stations, air quality variables from air monitoring stations, or
-any variables that are measured at fixed location across time.
+spatio-temporal data.
+
+In spatio-temporal data analysis, analysts will receive one of the
+various data formats depends on the data provider and need to manually
+rearrange the data to fit into one of the data object in \proglang{R}
+before starting any analysis on the data itself. Looking at spatial or
+temporal data analysis, data objects like \pkg{sf} \citep{sf} and
+\pkg{tsibble} \citep{tsibble} have smoothed the data analysis in these
+two domains while there has not yet been a spatio-temporal data
+structure that makes it easy to explore spatio-temporal data and this
+creates a gap in the software development. The requirement for such a
+tool is important given the ubiquity of spatio-temporal vector data in
+the wild: climate observations from weather stations, air quality
+variables from air monitoring stations, or any variables that are
+measured at fixed location across time.
 
@@
 
-Spatio-temporal data usually come to the analysts in various forms and
-Figure \ref{fig:illu-input} shows four examples of this. The first three
-are variations of tabular data that uses a single table (1) or multiple
-tables (2 \& 3) while the fourth uses a cube or hyper-cube to to
-represent the data. An example of the fourth type is the NetCDF (Network
-Common Data Form), commonly used in the climatology community to deliver
-global mapping of atmosphere, ocean, and land.
-
-These four examples show the common characters of the spatio-temporal
-data: 1) a spatial identifier that uniquely identifies each site, 2) a
-temporal identifier that prescribes the time stamp each site is
-recorded, 3) Coordinates, comprising of latitude and longitude, that
-locate the site on the map. These identifiers will be the building
-blocks for cubble. Another way to structure the spatio-temporal data is
-to think of it as a data cube with three axes: id, time, and variable.
-Notice that in the time-variable face of this cube, time invariant
-variables would share the same value across all the time stamps, while
-time varying variables have values varying across time. (Maybe add a
-cube diagram here). This heterogeneity will be used to split the data
-into two forms in cubble.
-
-\begin{CodeChunk}
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.001}
-
-}
-
-\caption[An illustration of different spatio-temporal data format]{An illustration of different spatio-temporal data format: (1) a single table for all the variables; (2) separate tables for spatial and temporal variables; (3) a table of metadata used to query the database and a separate table for queried data; and (4) an array data structure of cube or hypercube.}\label{fig:illu-input}
-\end{figure}
-\end{CodeChunk}
-
-In a cubble, there are two forms: nested form and long form, and Figure
-\ref{fig:illu-cubble} sketches the two forms with the associated
-attributes. The decision on which form to use in data manipulation is
-output-oriented, meaning analysts need to first think about whether the
-output of an operation is identified only by the spatial identifier, or
-a combination of spatial and temporal identifier. The nested cubble is
-suitable for working with operations that are only identified by the
-site and this type of operation can be a pure manipulation of time
-invariant variables, or an operation that summarises time varying
-variables into site. Underneath the nested form, a cubble is built from
-a \code{rowwise_df} data frame where each site forms a separate group.
-This structure simplifies the calculation that involves temporal
-variables by avoiding the use of \code{purrr::map()} syntax when working
-with list-column.
+\textcolor{red}{more beginning} In cubble data can represented in two
+formats: nested form and long form, and Figure \ref{fig:illu-cubble}
+sketches the two forms with the associated attributes. The decision on
+which form to use in data manipulation is output-oriented, meaning
+analysts need to first think about whether the output of an operation is
+identified only by the spatial identifier, or a combination of spatial
+and temporal identifier. The nested cubble is suitable for working with
+operations that are only identified by the site and this type of
+operation can be a pure manipulation of time invariant variables, or an
+operation that summarises time varying variables into site. Underneath
+the nested form, a cubble is built from a \code{rowwise_df} data frame
+where each site forms a separate group. This structure simplifies the
+calculation that involves temporal variables by avoiding the use of
+\code{purrr::map()} syntax when working with list-column.
 
@@
 
-A cubble can be created with \code{as_cubble()} by supplying the three
-key components: \code{key} as the spatial identifier; \code{index} as
-the temporal identifier; and a vector of \code{coords} in the order of
-longitude and latitude. The naming of \code{key} and \code{index}
-follows the convention in the \pkg{tsibble} package. Below is an example
-of creating a cubble and by default, the cubble created is in the nested
-form:
+\textcolor{red}{more beginning} A cubble can be created with
+\code{as_cubble()} by supplying the three key components: \code{key} as
+the spatial identifier; \code{index} as the temporal identifier; and a
+vector of \code{coords} in the order of longitude and latitude. The
+naming of \code{key} and \code{index} follows the convention in the
+\pkg{tsibble} package.
+\textcolor{red}{give some context of the data about to be used: a subset of a larger data etc}
+Below is an example of creating a cubble and by default, the cubble
+created is in the nested form:
 
@@
 
+\textcolor{red}{A beginning sentence that summarises the subsection}.
+\textcolor{red}{Given more details and examples when going through each subsection}
+
 \hypertarget{dplyr}{%
@@
 Cubble is also compatible with \pkg{tsibble}. When creating a cubble
-from a tsibble, Only the \code{coords} argument needs to be specified:
+from a tsibble, only the \code{coords} argument needs to be specified:
 
@@
 
-Cubble provides functionality for matching sites from two data sources.
-To be a valid pair of matching, the matched pair from different data
-sources need to:
+\textcolor{red}{more beginning} Cubble provides functionality for
+matching sites from two data sources. To be a valid pair of matching,
+the matched pair from different data sources need to:
 

### 2022-02-16 2-the-cubble-pkg.Rmd
flesh out section 2.1 - 2.4

@@
 The nested cubble is suitable for working with operations that are only identified by the site and this type of operation can be a pure manipulation of time invariant variables, or an operation that summarises time varying variables into site.
-Underneath the nested form, a cubble is built from a \code{rowwise_df} data frame where each site forms a separate group.
+Underneath the nested form, a cubble is built from a \code{rowwise_df} class where each site forms a separate group.
 This structure simplifies the calculation that involves temporal variables by avoiding the use of \code{purrr::map()} syntax when working with list-column.
@@
 For those operations whose output involves both a spatial and temporal identifier, long form should be used.
-The long form uses a  \code{grouped_df} data frame to forms all the time of a site as a group.
+The long form uses a  \code{grouped_df} class to forms all the time of a site as a group.
 Time invariant variables are stored separately as an special attribute of the long cubble.
@@
 
-\textcolor{red}{more beginning} A cubble can be created with \code{as_cubble()} by supplying the three key components: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order of longitude and latitude.
-The naming of \code{key} and \code{index} follows the convention in the \pkg{tsibble} package.
-\textcolor{red}{give some context of the data about to be used: a subset of a larger data etc} Below is an example of creating a cubble and by default, the cubble created is in the nested form:
+To use functions in the `cubble` package, an analyst will need to first turn the data object into a \code{cubble} class.
+\code{as_cubble()} does this through supplying the three key components: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order of longitude and latitude.
+The use of \code{key} and \code{index} follows the convention in the \pkg{tsibble} package.
+The cubble created by default is in the nested form.
+
+Before the formal examples in Section \ref{examples}, each function introduced in this section will be demonstrated with a short example with a data that contains 5 Australia weather stations, a subset of hundreds of weather stations in Australia.
+This dataset contains spatial information of each station: station id, latitude, longitude, elevation, station name and World Meteorology Organisation ID and also daily temporal information: date, maximum and minimum temperature and precipitation for 2020.
+\code{climate_flat} stores this data in format 1) in Figure \ref{fig:illu-input} and the code below creates a cubble out of \code{climate_flat} with \code{id} as the key, \code{date} as the index, and \code{c(long, lat)} as the coordinates:
 
@@
 
-There are a few information in the cubble header: the name of the \code{key} variable: \code{id} and its number: 5, the bounding box: \code{[115.97, -32.94, 133.55, -12.42]} , and also the name of variable nested in the \code{ts} column with its type: \code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.
+There are a few information in the cubble header: the name of the \code{key} variable: \code{id} and its unique number: 5, the bounding box: \code{[115.97, -32.94, 133.55, -12.42]} , and also the name of variable nested in the \code{ts} column with its type: \code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.
 
@@
 
-The long cubble is suitable to manipulate the time dimension of the data and  \code{stretch()} switches the nested cubble into the long cubble.
-It first extracts all the spatial variables into a separate tibble and then unnests the \code{ts} column.
-The spatial variables are then attached as an attributes of the long cubble and can be accessed through `spatial()`:
+The nested format is convenient for those operations whose output doesn't contain a time dimension, for those that are identified by both the spatial and temporal identifier, a long cubble is more suitable.
+Analysts will need to switch the nested cubble into a long cubble for such operation and \code{stretch()} is designed for this purpose: it first extracts all the spatial variables into a separate tibble and then unnests the \code{ts} column.
+This code switches the nested cubble just created into its long form:
 
@@
 cubble_long
-
-spatial(cubble_long)
 ```
@@
 Notice that the third line in the header now shows the name and type of spatial variables: \code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]}.
+The spatial variables are still part of the data object but as a \code{spatial} attribute:
+
+```{r}
+attr(cubble_long, "spatial")
+```
 
@@
 
-Analysts may need the time invariant and varying variable to be in the same table to apply certain function.
-In cubble, this can be done through \code{migrate()}, which moves the time invariant variables into the long cubble:
+Sometimes, analysts may need the spatial and temporal variable to be in the same table to apply some variable transformation.
+Cubble allows this operation through \code{migrate()}, which  moves the specified spatial variables into the long cubble:
 
@@
 
-Notice that these added time invariant variables will not be remembered by cubble and will disappear if switched to the nested form and then switched back.
-Hence, \code{migrate()} should usually be used as the last step of the analysis:
+This operation should generally be used in the last step of the analysis since it is a temporary operation, meaning meaning these added spatial variables will disappear if switched to the nested form and then switched back:
 
@@
 
-\textcolor{red}{A beginning sentence that summarises the subsection}. 
-\textcolor{red}{Given more details and examples when going through each subsection}
+The last 4 subsections introduces operations specific to the \code{cubble} class and this section will demonstrate how the \code{cubble} class interacts with existing packages, specifically, \code{dplyr}, \code{tsibble}, \code{sf}, and \code{netcdf4}.
 

### 2022-02-16 main.tex
flesh out section 2.1 - 2.4

@@
 operation that summarises time varying variables into site. Underneath
-the nested form, a cubble is built from a \code{rowwise_df} data frame
-where each site forms a separate group. This structure simplifies the
+the nested form, a cubble is built from a \code{rowwise_df} class where
+each site forms a separate group. This structure simplifies the
 calculation that involves temporal variables by avoiding the use of
@@
 identifier, long form should be used. The long form uses a
-\code{grouped_df} data frame to forms all the time of a site as a group.
-Time invariant variables are stored separately as an special attribute
-of the long cubble. This design avoids repeating the spatial variables
-at each time stamp while not dropping information from spatial
-variables.
+\code{grouped_df} class to forms all the time of a site as a group. Time
+invariant variables are stored separately as an special attribute of the
+long cubble. This design avoids repeating the spatial variables at each
+time stamp while not dropping information from spatial variables.
 
@@
 
-\textcolor{red}{more beginning} A cubble can be created with
-\code{as_cubble()} by supplying the three key components: \code{key} as
-the spatial identifier; \code{index} as the temporal identifier; and a
-vector of \code{coords} in the order of longitude and latitude. The
-naming of \code{key} and \code{index} follows the convention in the
-\pkg{tsibble} package.
-\textcolor{red}{give some context of the data about to be used: a subset of a larger data etc}
-Below is an example of creating a cubble and by default, the cubble
-created is in the nested form:
+To use functions in the \texttt{cubble} package, an analyst will need to
+first turn the data object into a \code{cubble} class.
+\code{as_cubble()} does this through supplying the three key components:
+\code{key} as the spatial identifier; \code{index} as the temporal
+identifier; and a vector of \code{coords} in the order of longitude and
+latitude. The use of \code{key} and \code{index} follows the convention
+in the \pkg{tsibble} package. The cubble created by default is in the
+nested form.
+
+Before the formal examples in Section \ref{examples}, each function
+introduced in this section will be demonstrated with a short example
+with a data that contains 5 Australia weather stations, a subset of
+hundreds of weather stations in Australia. This dataset contains spatial
+information of each station: station id, latitude, longitude, elevation,
+station name and World Meteorology Organisation ID and also daily
+temporal information: date, maximum and minimum temperature and
+precipitation for 2020. \code{climate_flat} stores this data in format
+1) in Figure \ref{fig:illu-input} and the code below creates a cubble
+out of \code{climate_flat} with \code{id} as the key, \code{date} as the
+index, and \code{c(long, lat)} as the coordinates:
 
@@
 # cubble:   id [5]: nested form
-# bbox:     [115.97, -32.94, 133.55, -12.42]- check gap on long and lat
+# bbox:     [115.97, -32.94, 133.55, -12.42]
 # temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
@@
 There are a few information in the cubble header: the name of the
-\code{key} variable: \code{id} and its number: 5, the bounding box:
-\code{[115.97, -32.94, 133.55, -12.42]} , and also the name of variable
-nested in the \code{ts} column with its type:
+\code{key} variable: \code{id} and its unique number: 5, the bounding
+box: \code{[115.97, -32.94, 133.55, -12.42]} , and also the name of
+variable nested in the \code{ts} column with its type:
 \code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.
@@
 
-The long cubble is suitable to manipulate the time dimension of the data
-and \code{stretch()} switches the nested cubble into the long cubble. It
+The nested format is convenient for those operations whose output
+doesn't contain a time dimension, for those that are identified by both
+the spatial and temporal identifier, a long cubble is more suitable.
+Analysts will need to switch the nested cubble into a long cubble for
+such operation and \code{stretch()} is designed for this purpose: it
 first extracts all the spatial variables into a separate tibble and then
-unnests the \code{ts} column. The spatial variables are then attached as
-an attributes of the long cubble and can be accessed through
-\texttt{spatial()}:
+unnests the \code{ts} column. This code switches the nested cubble just
+created into its long form:
 
@@
 \end{CodeOutput}
-\begin{CodeInput}
-R> spatial(cubble_long)
-\end{CodeInput}
+\end{CodeChunk}
+
+Notice that the third line in the header now shows the name and type of
+spatial variables:
+\code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]}. The
+spatial variables are still part of the data object but as a
+\code{spatial} attribute:
+
+\begin{CodeChunk}
 \begin{CodeOutput}
@@
 
-Notice that the third line in the header now shows the name and type of
-spatial variables:
-\code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]}.
-
 \hypertarget{tamp-a-long-cubble-back-to-the-nested-form}{%
@@
 
-Analysts may need the time invariant and varying variable to be in the
-same table to apply certain function. In cubble, this can be done
-through \code{migrate()}, which moves the time invariant variables into
-the long cubble:
+Sometimes, analysts may need the spatial and temporal variable to be in
+the same table to apply some variable transformation. Cubble allows this
+operation through \code{migrate()}, which moves the specified spatial
+variables into the long cubble:
 
@@
 
-Notice that these added time invariant variables will not be remembered
-by cubble and will disappear if switched to the nested form and then
-switched back. Hence, \code{migrate()} should usually be used as the
-last step of the analysis:
+This operation should generally be used in the last step of the analysis
+since it is a temporary operation, meaning meaning these added spatial
+variables will disappear if switched to the nested form and then
+switched back:
 
@@
 # cubble:  date, id [5]: long form
-# bbox:    [115.97, -32.94, 133.55, -12.42]- check gap on long and lat
+# bbox:    [115.97, -32.94, 133.55, -12.42]
 # spatial: long [dbl], lat [dbl], elev [dbl], name [chr], wmo_id [dbl]
@@
 
-\textcolor{red}{A beginning sentence that summarises the subsection}.
-\textcolor{red}{Given more details and examples when going through each subsection}
+The last 4 subsections introduces operations specific to the
+\code{cubble} class and this section will demonstrate how the
+\code{cubble} class interacts with existing packages, specifically,
+\code{dplyr}, \code{tsibble}, \code{sf}, and \code{netcdf4}.
 

### 2022-02-16 2-the-cubble-pkg.Rmd
flesh out more in the beginning of sec 2 and compatibility

@@
 
-\textcolor{red}{more beginning} In cubble data can represented in two formats: nested form and long form, and Figure \ref{fig:illu-cubble} sketches the two forms with the associated attributes.
+In this section, the two formats cubble uses to arrange spatio-temporal data will be presented Then operations will be introduced, with short examples, to illustrate how to work with these two formats.
+Lastly, a subsection will be dedicated to how existing packages, in spatial and temporal analysis, fit in with cubble.
+
+In cubble, data can represented in two formats: nested form and long form, and Figure \ref{fig:illu-cubble} sketches the two forms with the associated attributes.
 The decision on which form to use in data manipulation is output-oriented, meaning analysts need to first think about whether the output of an operation is identified only by the spatial identifier, or a combination of spatial and temporal identifier.
@@
 \code{as_cubble()} does this through supplying the three key components: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order of longitude and latitude.
-The use of \code{key} and \code{index} follows the convention in the \pkg{tsibble} package.
+The use of \code{key} and \code{index} follows the design in the \pkg{tsibble} package.
 The cubble created by default is in the nested form.
 
-Before the formal examples in Section \ref{examples}, each function introduced in this section will be demonstrated with a short example with a data that contains 5 Australia weather stations, a subset of hundreds of weather stations in Australia.
+Before the formal examples in Section \ref{examples}, each function introduced in this section will be demonstrated with a short example with a data that contains five Australia weather stations, a subset of hundreds of weather stations in Australia.
 This dataset contains spatial information of each station: station id, latitude, longitude, elevation, station name and World Meteorology Organisation ID and also daily temporal information: date, maximum and minimum temperature and precipitation for 2020.
@@
 ```{r echo = TRUE}
-cubble_long <- cubble_nested %>% stretch(ts)
+cubble_long <- cubble_nested %>% stretch()
 cubble_long
@@
 
-```{r}
+```{r echo = TRUE}
 attr(cubble_long, "spatial")
@@
 
-The last 4 subsections introduces operations specific to the \code{cubble} class and this section will demonstrate how the \code{cubble} class interacts with existing packages, specifically, \code{dplyr}, \code{tsibble}, \code{sf}, and \code{netcdf4}.
+The last four subsections introduce operations specific to the \code{cubble} class and this section will demonstrate how the \code{cubble} class interacts with existing packages, specifically, \code{dplyr}, \code{tsibble}, \code{sf} )\code{s2}), and \code{netcdf4}.
 
@@
 
-Building from an \code{tbl_df} class, cubble has implemented methods for \pkg{dplyr} generics, which includes:
+The \code{dplyr} package has provided many tools for common data wrangling tasks and these operations are useful in the spatio-temporal context.
+\code{cubble} provides its method for the following \code{dplyr} verbs to wrangle data in both the nested and long form:
 
-  - basic: mutate, filter, summarise, select, arrange, rename, left_join
-  - grouping: group_by, ungroup
-  - slice family: slice_head, slice_tail, slice_sample, slice_min, slice_max
+ > `mutate`, `filter`, `summarise`, `select`, `arrange`, `rename`, `left_join`, and the slice family (`slice_head`, `slice_tail`, `slice_sample`, `slice_min`, `slice_max`)
+
+\textcolor{red}{more work on groupby} The grouping pair, \code{group_by} and \code{ungroup}, can only be used in the long form to define grouping on the time but not in the nested form.
 
@@
 
-Cubble is also compatible with \pkg{tsibble}. When creating a cubble from a tsibble, only the \code{coords} argument needs to be specified:
+Tsibble is a temporal data structure that uses \code{index} and \code{key} to identify the time and different series.
+Cubble can be seen as a natural extension of tsibble for spatio-temporal data with an additional coordinates component.
+The use of two forms avoids the duplicates of spatial variables at each time stamp and hence, in the visualisation, avoid drawing duplicated points with the same coordinate.
+To cast a \code{tsibble} into a \code{cubble}, only the \code{coords} argument needs to be supplied:
 
 ```{r echo = TRUE}
-# casting a tsibble into cubble
-dt <- climate_flat %>% 
-  tsibble::as_tsibble(key = id, index = date) %>% 
+# demonstrate with a tsibble created from climate_flat
+raw <- climate_flat %>% 
+  tsibble::as_tsibble(key = id, index = date)
+
+dt <-  raw %>% 
   cubble::as_cubble(coords = c(long, lat))
@@
 
-In the created nested form, each element of the list-column \code{ts} is of \code{tbl_ts} class and existing operations on the tsibble class can still be applied:
+In the created nested form, each element of the list-column \code{ts} is of \code{tbl_ts} class and operations available to the tsibble class is still valid under cubble.
+For example, the code below calculate two features of the maximum temperature:
 
@@
 
-Cubble is compatible with simple feature geometry list-column (\code{sfc}) in \pkg{sf} and  \code{s2_lnglat} vector in \pkg{S2} in the nested form.
-After creating a cubble, analysts can add the geometry vectors as needed to perform spatial operations:
+As a spatial data object, \code{sf} introduces a simple feature geometry list-column (\code{sfc}) to the data frame and provides spatial operations on various geometry types (\code{POINT}, \code{LINESTRING}, \code{POLYGON}, \code{MULTIPOLYGON} etc).
+While these operations are valueable for spatio-temporal data analysis, an \code{sf} object usually don't contain temporal variables and hence can't be directly cast into a \code{cubble}.
+However, \code{cubble} can contain \code{sfc} columns in the nested form and spatial operations built on the \code{sfc} class are available.
+The following example shows how to create an \code{sfc} column of \code{POINT} type based on latitude and longitude in cubble.
+The state \code{MULTIPOLYGON} is then added based on spatial operation \code{st_within} before making a coordinate transformation.
 
 ```{r echo = TRUE}
-library(s2)
+library(sf)
 # create a cubble
@@
 
-# western australia map
-wa <- ozmaps::abs_ste %>% filter(NAME == "Western Australia")
-
-# create a `s2_lnglat` vector based on long and lat and filter stations in wa
-cb %>%
-  mutate(ll = s2_lnglat(long, lat)) %>% 
-  filter(s2_within(ll, wa))
-```
-
-```{r echo = TRUE}
-library(sf)
-# simplify australia map
-aus <- ozmaps::abs_ste %>% rmapshaper::ms_simplify(keep = 0.01)
+aus <- ozmaps::abs_ste
 
@@
       crs = st_crs(aus)),
+
     # append state multi-polygon based on the `sfc` created
     state = aus$geometry[st_within(ll, aus, sparse = FALSE)],
+
     # adopt a different projection: lambert conformal conic (EPSG:3112)
-    state = st_transform(state, crs = "EPSG:3112"))
+    state = st_transform(state, crs = "EPSG:3112")
+    )
 
 dt
+```
+
+An \pkg{s2} \code{lnglat} vector can similarly be created in cubble for using \code{s2}-prefixed functions:
+
+```{r echo = TRUE}
+library(s2)
+# western australia map
+wa <- ozmaps::abs_ste %>% filter(NAME == "Western Australia")
+
+# mutate a `s2_lnglat` vector on `cb` created in the last chunk
+cb %>%
+  mutate(ll = s2_lnglat(long, lat)) %>% 
+  filter(s2_within(ll, wa))
 ```

### 2022-02-16 main.tex
flesh out more in the beginning of sec 2 and compatibility

@@
 
-\textcolor{red}{more beginning} In cubble data can represented in two
-formats: nested form and long form, and Figure \ref{fig:illu-cubble}
-sketches the two forms with the associated attributes. The decision on
-which form to use in data manipulation is output-oriented, meaning
-analysts need to first think about whether the output of an operation is
-identified only by the spatial identifier, or a combination of spatial
-and temporal identifier. The nested cubble is suitable for working with
-operations that are only identified by the site and this type of
-operation can be a pure manipulation of time invariant variables, or an
-operation that summarises time varying variables into site. Underneath
-the nested form, a cubble is built from a \code{rowwise_df} class where
-each site forms a separate group. This structure simplifies the
-calculation that involves temporal variables by avoiding the use of
-\code{purrr::map()} syntax when working with list-column.
+In this section, the two formats cubble uses to arrange spatio-temporal
+data will be presented Then operations will be introduced, with short
+examples, to illustrate how to work with these two formats. Lastly, a
+subsection will be dedicated to how existing packages, in spatial and
+temporal analysis, fit in with cubble.
+
+In cubble, data can represented in two formats: nested form and long
+form, and Figure \ref{fig:illu-cubble} sketches the two forms with the
+associated attributes. The decision on which form to use in data
+manipulation is output-oriented, meaning analysts need to first think
+about whether the output of an operation is identified only by the
+spatial identifier, or a combination of spatial and temporal identifier.
+The nested cubble is suitable for working with operations that are only
+identified by the site and this type of operation can be a pure
+manipulation of time invariant variables, or an operation that
+summarises time varying variables into site. Underneath the nested form,
+a cubble is built from a \code{rowwise_df} class where each site forms a
+separate group. This structure simplifies the calculation that involves
+temporal variables by avoiding the use of \code{purrr::map()} syntax
+when working with list-column.
 
@@
 identifier; and a vector of \code{coords} in the order of longitude and
-latitude. The use of \code{key} and \code{index} follows the convention
-in the \pkg{tsibble} package. The cubble created by default is in the
+latitude. The use of \code{key} and \code{index} follows the design in
+the \pkg{tsibble} package. The cubble created by default is in the
 nested form.
@@
 introduced in this section will be demonstrated with a short example
-with a data that contains 5 Australia weather stations, a subset of
+with a data that contains five Australia weather stations, a subset of
 hundreds of weather stations in Australia. This dataset contains spatial
@@
 \begin{CodeInput}
-R> cubble_long <- cubble_nested %>% stretch(ts)
+R> cubble_long <- cubble_nested %>% stretch()
 R> cubble_long
@@
 \begin{CodeChunk}
+\begin{CodeInput}
+R> attr(cubble_long, "spatial")
+\end{CodeInput}
 \begin{CodeOutput}
@@
 
-The last 4 subsections introduces operations specific to the
+The last four subsections introduce operations specific to the
 \code{cubble} class and this section will demonstrate how the
 \code{cubble} class interacts with existing packages, specifically,
-\code{dplyr}, \code{tsibble}, \code{sf}, and \code{netcdf4}.
+\code{dplyr}, \code{tsibble}, \code{sf} )\code{s2}), and \code{netcdf4}.
 
@@
 
-Building from an \code{tbl_df} class, cubble has implemented methods for
-\pkg{dplyr} generics, which includes:
-
-\begin{itemize}
-\tightlist
-\item
-  basic: mutate, filter, summarise, select, arrange, rename, left\_join
-\item
-  grouping: group\_by, ungroup
-\item
-  slice family: slice\_head, slice\_tail, slice\_sample, slice\_min,
-  slice\_max
-\end{itemize}
+The \code{dplyr} package has provided many tools for common data
+wrangling tasks and these operations are useful in the spatio-temporal
+context. \code{cubble} provides its method for the following
+\code{dplyr} verbs to wrangle data in both the nested and long form:
+
+\begin{quote}
+\texttt{mutate}, \texttt{filter}, \texttt{summarise}, \texttt{select},
+\texttt{arrange}, \texttt{rename}, \texttt{left\_join}, and the slice
+family (\texttt{slice\_head}, \texttt{slice\_tail},
+\texttt{slice\_sample}, \texttt{slice\_min}, \texttt{slice\_max})
+\end{quote}
+
+\textcolor{red}{more work on groupby} The grouping pair, \code{group_by}
+and \code{ungroup}, can only be used in the long form to define grouping
+on the time but not in the nested form.
 
@@
 
-Cubble is also compatible with \pkg{tsibble}. When creating a cubble
-from a tsibble, only the \code{coords} argument needs to be specified:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> # casting a tsibble into cubble
-R> dt <- climate_flat %>% 
-+   tsibble::as_tsibble(key = id, index = date) %>% 
+Tsibble is a temporal data structure that uses \code{index} and
+\code{key} to identify the time and different series. Cubble can be seen
+as a natural extension of tsibble for spatio-temporal data with an
+additional coordinates component. The use of two forms avoids the
+duplicates of spatial variables at each time stamp and hence, in the
+visualisation, avoid drawing duplicated points with the same coordinate.
+To cast a \code{tsibble} into a \code{cubble}, only the \code{coords}
+argument needs to be supplied:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> # demonstrate with a tsibble created from climate_flat
+R> raw <- climate_flat %>% 
++   tsibble::as_tsibble(key = id, index = date) 
+R> 
+R> dt <-  raw %>% 
 +   cubble::as_cubble(coords = c(long, lat))
@@
 In the created nested form, each element of the list-column \code{ts} is
-of \code{tbl_ts} class and existing operations on the tsibble class can
-still be applied:
+of \code{tbl_ts} class and operations available to the tsibble class is
+still valid under cubble. For example, the code below calculate two
+features of the maximum temperature:
 
@@
 
-Cubble is compatible with simple feature geometry list-column
-(\code{sfc}) in \pkg{sf} and \code{s2_lnglat} vector in \pkg{S2} in the
-nested form. After creating a cubble, analysts can add the geometry
-vectors as needed to perform spatial operations:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> library(s2)
+As a spatial data object, \code{sf} introduces a simple feature geometry
+list-column (\code{sfc}) to the data frame and provides spatial
+operations on various geometry types (\code{POINT}, \code{LINESTRING},
+\code{POLYGON}, \code{MULTIPOLYGON} etc). While these operations are
+valueable for spatio-temporal data analysis, an \code{sf} object usually
+don't contain temporal variables and hence can't be directly cast into a
+\code{cubble}. However, \code{cubble} can contain \code{sfc} columns in
+the nested form and spatial operations built on the \code{sfc} class are
+available. The following example shows how to create an \code{sfc}
+column of \code{POINT} type based on latitude and longitude in cubble.
+The state \code{MULTIPOLYGON} is then added based on spatial operation
+\code{st_within} before making a coordinate transformation.
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> library(sf)
 R> # create a cubble
@@
 R> 
-R> # western australia map
-R> wa <- ozmaps::abs_ste %>% filter(NAME == "Western Australia")
-R> 
-R> # create a `s2_lnglat` vector based on long and lat and filter stations in wa
-R> cb %>%
-+   mutate(ll = s2_lnglat(long, lat)) %>% 
-+   filter(s2_within(ll, wa))
-\end{CodeInput}
-\begin{CodeOutput}
-# cubble:   id [3]: nested form
-# bbox:     [115.97, -32.94, 117.18, -31.89]
-# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
-  id            lat  long  elev name          wmo_id ts                 ll      
-  <chr>       <dbl> <dbl> <dbl> <chr>          <dbl> <list>             <s2_lng>
-1 ASN00009021 -31.9  116.  15.4 perth airport  94610 <tibble [366 x 4]> (115.97~
-2 ASN00010311 -31.9  117. 179   york           94623 <tibble [366 x 4]> (116.76~
-3 ASN00010614 -32.9  117. 338   narrogin       94627 <tibble [366 x 4]> (117.17~
-\end{CodeOutput}
-\end{CodeChunk}
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> library(sf)
-R> # simplify australia map
-R> aus <- ozmaps::abs_ste %>% rmapshaper::ms_simplify(keep = 0.01)
+R> aus <- ozmaps::abs_ste
 R> 
@@
 +       crs = st_crs(aus)),
++     
 +     # append state multi-polygon based on the `sfc` created
 +     state = aus$geometry[st_within(ll, aus, sparse = FALSE)],
++     
 +     # adopt a different projection: lambert conformal conic (EPSG:3112)
-+     state = st_transform(state, crs = "EPSG:3112"))
++     state = st_transform(state, crs = "EPSG:3112")
++     )
 R> 
@@
 # ... with 1 more variable: state <MULTIPOLYGON [m]>
+\end{CodeOutput}
+\end{CodeChunk}
+
+An \pkg{s2} \code{lnglat} vector can similarly be created in cubble for
+using \code{s2}-prefixed functions:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> library(s2)
+R> # western australia map
+R> wa <- ozmaps::abs_ste %>% filter(NAME == "Western Australia")
+R> 
+R> # mutate a `s2_lnglat` vector on `cb` created in the last chunk
+R> cb %>%
++   mutate(ll = s2_lnglat(long, lat)) %>% 
++   filter(s2_within(ll, wa))
+\end{CodeInput}
+\begin{CodeOutput}
+# cubble:   id [3]: nested form
+# bbox:     [115.97, -32.94, 117.18, -31.89]
+# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
+  id            lat  long  elev name          wmo_id ts                 ll      
+  <chr>       <dbl> <dbl> <dbl> <chr>          <dbl> <list>             <s2_lng>
+1 ASN00009021 -31.9  116.  15.4 perth airport  94610 <tibble [366 x 4]> (115.97~
+2 ASN00010311 -31.9  117. 179   york           94623 <tibble [366 x 4]> (116.76~
+3 ASN00010614 -32.9  117. 338   narrogin       94627 <tibble [366 x 4]> (117.17~
 \end{CodeOutput}

### 2022-02-17 2-the-cubble-pkg.Rmd
go through writing made yesterday

@@
 
-In this section, the two formats cubble uses to arrange spatio-temporal data will be presented Then operations will be introduced, with short examples, to illustrate how to work with these two formats.
-Lastly, a subsection will be dedicated to how existing packages, in spatial and temporal analysis, fit in with cubble.
+This section will present two formats that cubble uses to arrange spatio-temporal data.
+Main cubble functions will then be introduced to illustrate how to work with these two formats in short examples.
+Lastly, a subsection will dedicate to how existing packages, in spatial and temporal analysis, fit in with cubble.
 
@@
 
-To use functions in the `cubble` package, an analyst will need to first turn the data object into a \code{cubble} class.
+To use functions in the `cubble` package, an analyst will first need to turn the data object into a \code{cubble} class.
 \code{as_cubble()} does this through supplying the three key components: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order of longitude and latitude.
-The use of \code{key} and \code{index} follows the design in the \pkg{tsibble} package.
-The cubble created by default is in the nested form.
+The use of \code{key} and \code{index} follows the design in the \pkg{tsibble} package and the cubble created by default is in the nested form.
 
-Before the formal examples in Section \ref{examples}, each function introduced in this section will be demonstrated with a short example with a data that contains five Australia weather stations, a subset of hundreds of weather stations in Australia.
-This dataset contains spatial information of each station: station id, latitude, longitude, elevation, station name and World Meteorology Organisation ID and also daily temporal information: date, maximum and minimum temperature and precipitation for 2020.
+Before the formal examples in Section \ref{examples}, each function introduced in this section will be accompanied by a small example.
+The data used for the small examples is a subset of a larger climate weather station data used in the formal examples and has been simplified to include only five weather stations.
+It contains spatial information of each station: station id, latitude, longitude, elevation, station name and World Meteorology Organisation ID and also daily temporal information of date, maximum and minimum temperature and precipitation for 2020.
 \code{climate_flat} stores this data in format 1) in Figure \ref{fig:illu-input} and the code below creates a cubble out of \code{climate_flat} with \code{id} as the key, \code{date} as the index, and \code{c(long, lat)} as the coordinates:
@@
 
-The nested format is convenient for those operations whose output doesn't contain a time dimension, for those that are identified by both the spatial and temporal identifier, a long cubble is more suitable.
-Analysts will need to switch the nested cubble into a long cubble for such operation and \code{stretch()} is designed for this purpose: it first extracts all the spatial variables into a separate tibble and then unnests the \code{ts} column.
-This code switches the nested cubble just created into its long form:
+The nested format is convenient for those operations whose output doesn't contain a time dimension, for those outputs that are cross-identified by the spatial and temporal identifier, a long cubble needs to be used.
+The function \code{stretch()} is designed to switch the cubble from the nested form to the long form.
+This code shows how to switch the nested cubble just created into its long form:
 
@@
 
-Notice that the third line in the header now shows the name and type of spatial variables: \code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]}.
-The spatial variables are still part of the data object but as a \code{spatial} attribute:
+Notice that the third line in the header now shows the name and type of spatial variables: \code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]} and these variables are now stored as a \code{spatial} attribute of the data:
 
@@
 
-Sometimes, analysts may need the spatial and temporal variable to be in the same table to apply some variable transformation.
+Sometimes, analysts may need to apply some variable transformation that involves both the spatial and temporal variable.
+An example of this is the transformation of temporal variables into the spatial dimension in glyph maps(which will be elaborated in section \ref{st_transformation}).
 Cubble allows this operation through \code{migrate()}, which moves the specified spatial variables into the long cubble:
@@
 
-This operation should generally be used in the last step of the analysis since it is a temporary operation, meaning meaning these added spatial variables will disappear if switched to the nested form and then switched back:
+This function should generally be used in the last step of the analysis since it is a temporary operation, meaning these added spatial variables are not stored in the long form and will disappear if switched to the nested form and then switched back:
 
@@
 
-The last four subsections introduce operations specific to the \code{cubble} class and this section will demonstrate how the \code{cubble} class interacts with existing packages, specifically, \code{dplyr}, \code{tsibble}, \code{sf} )\code{s2}), and \code{netcdf4}.
+The previous four subsections have introduced operations specific to the \code{cubble} class and this section will demonstrate how the \code{cubble} class interacts with existing packages commonly used in spatial and temporal analysis, specifically, \code{dplyr}, \code{tsibble}, \code{sf} (\code{s2}), and \code{netcdf4}.
 
@@
 
-The \code{dplyr} package has provided many tools for common data wrangling tasks and these operations are useful in the spatio-temporal context.
-\code{cubble} provides its method for the following \code{dplyr} verbs to wrangle data in both the nested and long form:
+The \code{dplyr} package has provided many tools for data wrangling tasks and these operations are useful in the spatio-temporal context.
+\code{cubble} provides methods that support the following \code{dplyr} verbs in both the nested and long form:
 
@@
 
-Tsibble is a temporal data structure that uses \code{index} and \code{key} to identify the time and different series.
-Cubble can be seen as a natural extension of tsibble for spatio-temporal data with an additional coordinates component.
-The use of two forms avoids the duplicates of spatial variables at each time stamp and hence, in the visualisation, avoid drawing duplicated points with the same coordinate.
-To cast a \code{tsibble} into a \code{cubble}, only the \code{coords} argument needs to be supplied:
+\code{tsibble} is a temporal data structure that uses \code{index} and \code{key} to identify the time and different series and \code{cubble} can be seen as a natural extension of \code{tsibble} for spatio-temporal data with an additional coordinates component and using two forms to arrange variables.
+This makes it easy to cast a \code{tsibble} into a \code{cubble} as only the \code{coords} argument needs to be supplied:
 
@@
 
-In the created nested form, each element of the list-column \code{ts} is of \code{tbl_ts} class and operations available to the tsibble class is still valid under cubble.
-For example, the code below calculate two features of the maximum temperature:
+In the nested cubble created, each element in the list-column \code{ts} is of \code{tbl_ts} class and operations available to the tsibble class is still valid under cubble.
+For example, the code below calculates two features of the maximum temperature:
 
@@
 
-As a spatial data object, \code{sf} introduces a simple feature geometry list-column (\code{sfc}) to the data frame and provides spatial operations on various geometry types (\code{POINT}, \code{LINESTRING}, \code{POLYGON}, \code{MULTIPOLYGON} etc).
-While these operations are valueable for spatio-temporal data analysis, an \code{sf} object usually don't contain temporal variables and hence can't be directly cast into a \code{cubble}.
-However, \code{cubble} can contain \code{sfc} columns in the nested form and spatial operations built on the \code{sfc} class are available.
-The following example shows how to create an \code{sfc} column of \code{POINT} type based on latitude and longitude in cubble.
-The state \code{MULTIPOLYGON} is then added based on spatial operation \code{st_within} before making a coordinate transformation.
+As a spatial data object, \code{sf} creates a simple feature geometry list-column (\code{sfc}) in the data frame to provide spatial operations on various geometry types (\code{POINT}, \code{LINESTRING}, \code{POLYGON}, \code{MULTIPOLYGON}, etc).
+These spatial operations are also valuable for spatio-temporal data analysis, but an \code{sf} object *usually* don't contain temporal variables.
+This means \code{sf} can't be directly cast into a \code{cubble}, however, \code{cubble} does support \code{sfc} columns in the nested form and spatial operations applied to the \code{sfc} column in \code{sf} can still be applied to the \code{sfc} column in a cubble.
+The following example shows how to create an \code{sfc} column of \code{POINT} type from latitude and longitude in cubble.
+Then \code{sf::st_within} is used to add the state \code{MULTIPOLYGON} of each weather station before a coordinate transformation is made.
 
@@
 
-An \pkg{s2} \code{lnglat} vector can similarly be created in cubble for using \code{s2}-prefixed functions:
+An \pkg{s2} \code{lnglat} vector can similarly be created as an \code{sfc} in cubble before using any \code{s2}-prefixed function:
 

### 2022-02-17 main.tex
go through writing made yesterday

@@
 
-In this section, the two formats cubble uses to arrange spatio-temporal
-data will be presented Then operations will be introduced, with short
-examples, to illustrate how to work with these two formats. Lastly, a
-subsection will be dedicated to how existing packages, in spatial and
+This section will present two formats that cubble uses to arrange
+spatio-temporal data. Main cubble functions will then be introduced to
+illustrate how to work with these two formats in short examples. Lastly,
+a subsection will dedicate to how existing packages, in spatial and
 temporal analysis, fit in with cubble.
@@
 
-To use functions in the \texttt{cubble} package, an analyst will need to
-first turn the data object into a \code{cubble} class.
+To use functions in the \texttt{cubble} package, an analyst will first
+need to turn the data object into a \code{cubble} class.
 \code{as_cubble()} does this through supplying the three key components:
@@
 latitude. The use of \code{key} and \code{index} follows the design in
-the \pkg{tsibble} package. The cubble created by default is in the
+the \pkg{tsibble} package and the cubble created by default is in the
 nested form.
@@
 Before the formal examples in Section \ref{examples}, each function
-introduced in this section will be demonstrated with a short example
-with a data that contains five Australia weather stations, a subset of
-hundreds of weather stations in Australia. This dataset contains spatial
-information of each station: station id, latitude, longitude, elevation,
-station name and World Meteorology Organisation ID and also daily
-temporal information: date, maximum and minimum temperature and
-precipitation for 2020. \code{climate_flat} stores this data in format
-1) in Figure \ref{fig:illu-input} and the code below creates a cubble
-out of \code{climate_flat} with \code{id} as the key, \code{date} as the
-index, and \code{c(long, lat)} as the coordinates:
+introduced in this section will be accompanied by a small example. The
+data used for the small examples is a subset of a larger climate weather
+station data used in the formal examples and has been simplified to
+include only five weather stations. It contains spatial information of
+each station: station id, latitude, longitude, elevation, station name
+and World Meteorology Organisation ID and also daily temporal
+information of date, maximum and minimum temperature and precipitation
+for 2020. \code{climate_flat} stores this data in format 1) in Figure
+\ref{fig:illu-input} and the code below creates a cubble out of
+\code{climate_flat} with \code{id} as the key, \code{date} as the index,
+and \code{c(long, lat)} as the coordinates:
 
@@
 The nested format is convenient for those operations whose output
-doesn't contain a time dimension, for those that are identified by both
-the spatial and temporal identifier, a long cubble is more suitable.
-Analysts will need to switch the nested cubble into a long cubble for
-such operation and \code{stretch()} is designed for this purpose: it
-first extracts all the spatial variables into a separate tibble and then
-unnests the \code{ts} column. This code switches the nested cubble just
-created into its long form:
+doesn't contain a time dimension, for those outputs that are
+cross-identified by the spatial and temporal identifier, a long cubble
+needs to be used. The function \code{stretch()} is designed to switch
+the cubble from the nested form to the long form. This code shows how to
+switch the nested cubble just created into its long form:
 
@@
 spatial variables:
-\code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]}. The
-spatial variables are still part of the data object but as a
-\code{spatial} attribute:
+\code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]} and
+these variables are now stored as a \code{spatial} attribute of the
+data:
 
@@
 
-Sometimes, analysts may need the spatial and temporal variable to be in
-the same table to apply some variable transformation. Cubble allows this
-operation through \code{migrate()}, which moves the specified spatial
-variables into the long cubble:
+Sometimes, analysts may need to apply some variable transformation that
+involves both the spatial and temporal variable. An example of this is
+the transformation of temporal variables into the spatial dimension in
+glyph maps(which will be elaborated in section \ref{st_transformation}).
+Cubble allows this operation through \code{migrate()}, which moves the
+specified spatial variables into the long cubble:
 
@@
 
-This operation should generally be used in the last step of the analysis
-since it is a temporary operation, meaning meaning these added spatial
-variables will disappear if switched to the nested form and then
-switched back:
+This function should generally be used in the last step of the analysis
+since it is a temporary operation, meaning these added spatial variables
+are not stored in the long form and will disappear if switched to the
+nested form and then switched back:
 
@@
 
-The last four subsections introduce operations specific to the
+The previous four subsections have introduced operations specific to the
 \code{cubble} class and this section will demonstrate how the
-\code{cubble} class interacts with existing packages, specifically,
-\code{dplyr}, \code{tsibble}, \code{sf} )\code{s2}), and \code{netcdf4}.
+\code{cubble} class interacts with existing packages commonly used in
+spatial and temporal analysis, specifically, \code{dplyr},
+\code{tsibble}, \code{sf} (\code{s2}), and \code{netcdf4}.
 
@@
 
-The \code{dplyr} package has provided many tools for common data
-wrangling tasks and these operations are useful in the spatio-temporal
-context. \code{cubble} provides its method for the following
-\code{dplyr} verbs to wrangle data in both the nested and long form:
+The \code{dplyr} package has provided many tools for data wrangling
+tasks and these operations are useful in the spatio-temporal context.
+\code{cubble} provides methods that support the following \code{dplyr}
+verbs in both the nested and long form:
 
@@
 
-Tsibble is a temporal data structure that uses \code{index} and
-\code{key} to identify the time and different series. Cubble can be seen
-as a natural extension of tsibble for spatio-temporal data with an
-additional coordinates component. The use of two forms avoids the
-duplicates of spatial variables at each time stamp and hence, in the
-visualisation, avoid drawing duplicated points with the same coordinate.
-To cast a \code{tsibble} into a \code{cubble}, only the \code{coords}
-argument needs to be supplied:
+\code{tsibble} is a temporal data structure that uses \code{index} and
+\code{key} to identify the time and different series and \code{cubble}
+can be seen as a natural extension of \code{tsibble} for spatio-temporal
+data with an additional coordinates component and using two forms to
+arrange variables. This makes it easy to cast a \code{tsibble} into a
+\code{cubble} as only the \code{coords} argument needs to be supplied:
 
@@
 
-In the created nested form, each element of the list-column \code{ts} is
-of \code{tbl_ts} class and operations available to the tsibble class is
-still valid under cubble. For example, the code below calculate two
+In the nested cubble created, each element in the list-column \code{ts}
+is of \code{tbl_ts} class and operations available to the tsibble class
+is still valid under cubble. For example, the code below calculates two
 features of the maximum temperature:
@@
 
-As a spatial data object, \code{sf} introduces a simple feature geometry
-list-column (\code{sfc}) to the data frame and provides spatial
-operations on various geometry types (\code{POINT}, \code{LINESTRING},
-\code{POLYGON}, \code{MULTIPOLYGON} etc). While these operations are
-valueable for spatio-temporal data analysis, an \code{sf} object usually
-don't contain temporal variables and hence can't be directly cast into a
-\code{cubble}. However, \code{cubble} can contain \code{sfc} columns in
-the nested form and spatial operations built on the \code{sfc} class are
-available. The following example shows how to create an \code{sfc}
-column of \code{POINT} type based on latitude and longitude in cubble.
-The state \code{MULTIPOLYGON} is then added based on spatial operation
-\code{st_within} before making a coordinate transformation.
+As a spatial data object, \code{sf} creates a simple feature geometry
+list-column (\code{sfc}) in the data frame to provide spatial operations
+on various geometry types (\code{POINT}, \code{LINESTRING},
+\code{POLYGON}, \code{MULTIPOLYGON}, etc). These spatial operations are
+also valuable for spatio-temporal data analysis, but an \code{sf} object
+\emph{usually} don't contain temporal variables. This means \code{sf}
+can't be directly cast into a \code{cubble}, however, \code{cubble} does
+support \code{sfc} columns in the nested form and spatial operations
+applied to the \code{sfc} column in \code{sf} can still be applied to
+the \code{sfc} column in a cubble. The following example shows how to
+create an \code{sfc} column of \code{POINT} type from latitude and
+longitude in cubble. Then \code{sf::st_within} is used to add the state
+\code{MULTIPOLYGON} of each weather station before a coordinate
+transformation is made.
 
@@
 
-An \pkg{s2} \code{lnglat} vector can similarly be created in cubble for
-using \code{s2}-prefixed functions:
+An \pkg{s2} \code{lnglat} vector can similarly be created as an
+\code{sfc} in cubble before using any \code{s2}-prefixed function:
 
@@
 
-\hypertarget{spatio-temporal-transformations}{%
-\subsection{Spatio-temporal
-transformations}\label{spatio-temporal-transformations}}
+\hypertarget{st_transformation}{%
+\subsection{Spatio-temporal transformations}\label{st_transformation}}
 

### 2022-02-17 1-intro.Rmd
more refine on intro

--- 2022-02-07 1-intro.Rmd
@@
 
-Spatio-temporal data concerns variables in the spatial coordinates across time.
-Some variables are invariant to time and can be solely identified by its spatial location while other time varying variables need both the spatial and temporal identifiers to identify.
-Manipulating data in these two levels is usually done separately and the third tidy data principal [@tidydata] recommends each type of observational unit to form a table.
+Spatio-temporal data concerns variables observed in the space across time and this paper focuses on the spatio-temporal vector data with fixed locations and regular time interval.
+This type of data can naturally been considered as a data cube and here a cube with three axes: spatial identifier, temporal identifier, and variable can be used to conceptualise spatio-temporal data.
+In this cubic formation, space only occupy one dimension, rather than two, to avoid the needs of hypercube for multivariate spatio-temporal data.
+
+While conceptually the data format is relatively simple, the actual data object that comes to the analysts in the analysis can be in various forms.
+Figure \ref{fig:illu-input} shows four examples of wild spatio-temporal data: 1) combines the time invariant and varying variables into a single table through duplicating the time invariant variables; 2) splits those variables into a separate spatial table and a temporal table; 3) ...; and lastly 4) is array-based data, like NetCDF (Network Common Data Form) data, which is commonly used in earth science, e.g.
+climatology, meteorology, and oceanography, among others, to organise multivariate data in a spatio-temporal grid.
+
+```{r illu-input, echo = FALSE, out.width = "100%", fig.cap="An illustration of different spatio-temporal data format: (1) a single table for all the variables; (2) separate tables for spatial and temporal variables; (3) a table of metadata used to query the database and a separate table for queried data; and (4) an array data structure.", fig.align='center', out.height = "40%"}
+knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.001.png"))
+```
+
+<!-- Manipulating data in these two levels is usually done separately and the third tidy data principal [@tidydata] recommends each type of observational unit to form a table.
 Another feature for a spatio-temporal data structure to have is that variables created from subsequent operations can be automatically added to the existing table with the same level.
 This is better than an output that index the variables with the identifiers since otherwise analysts will need to manually join this output with existing variables in the same level.
+-->
 
@@
 
-Currently, available spatio-temporal data structures in R includes: \pkg{spacetime} [@spacetime], which proposes four space-time layouts: Full grid (STF), sparse grid(STS), irregular (STI), and trajectory (STT).
-\pkg{spacetime} is built on \pkg{sp} [@sp] and \pkg{xts} [@xts], both of which have been replaced by more recent implementations.
-\pkg{spatstat} [@spatstat] implements a `ppp` class for point pattern data; and more recent, \pkg{stars} [@stars] implements a spatio-temporal for vector and raster spatio-temporal data.
-While these implementations either store spatial and temporal variables all in a single table, hence duplicate the spatial variables at each temporal unit; or split them into two separate tables that analysts need to manually join the data constantly when doing the analysis.
-This creates a gap in the software development.
+Different authors have proposed various data objects to work with spatial temporal data in \proglang{R} and this includes: \pkg{spacetime} [@spacetime] is built on the \pkg{sp} [@sp] and \pkg{xts} [@xts] class and both of which have been replaced by more recent spatial and temporal data objects.
+\pkg{spatstat} [@spatstat] implements a `ppp` class for analysing point pattern data; and more recent, \pkg{stars} [@stars] has implemented an spatio-temporal array for wrangling both vector and raster data.
+
+\textcolor{red}{Think again on the motivation and why it is hard to work existing tools.} In spatio-temporal data analysis, analysts will receive one of the various data formats depends on the data provider and need to manually rearrange the data to fit into one of the data object in \proglang{R} before starting any analysis on the data itself.
+
+Looking at spatial or temporal data analysis, data objects like \pkg{sf} [@sf] and \pkg{tsibble} [@tsibble] have smoothed the data analysis in these two domains while there has not yet been a spatio-temporal data structure that makes it easy to explore spatio-temporal data and this creates a gap in the software development.
 The requirement for such a tool is important given the ubiquity of spatio-temporal vector data in the wild: climate observations from weather stations, air quality variables from air monitoring stations, or any variables that are measured at fixed location across time.
+
+<!-- While these implementations either store spatial and temporal variables all in a single table, hence duplicate the spatial variables at each temporal unit; or split them into two separate tables that analysts need to manually join the data constantly when doing the analysis.
+\textcolor{red}{There has been data structure proposed for spatial (\pkg{sf} by @sf) and temporal (\pkg{tsibble} by @tsibble) data in the R community, while less has been done for spatio-temporal data.}  -->
 

### 2022-02-17 main.tex
more refine on intro

@@
 
-Spatio-temporal data concerns variables in the space across time.
-\textcolor{red}{Elaborate on variable, space, and time}. If the time
-interval is regular and each site has fixed point location, it can also
-be thought of as a data cube with three axes: spatial identifier,
-temporal identifier, and variable. In this cubical layout, variables can
-be those that are invariant to time, i.e.~the longitude and latitude
-coordinates, or those that varying across time. While conceptually the
-data format is relatively simple, the actual data object that comes to
-the analysts can be in various forms in practice. Figure
-\ref{fig:illu-input} shows four examples where the first one combines
-the time invariant and varying variables into a single table through
-duplicating the time invariant variables; 2) splits those variables into
-two separate sheets; 3) \ldots; and lastly 4) considers the array-based
-data, like NetCDF (Network Common Data Form) data, which is commonly
-used in earth science, e.g.~climatology, meteorology, and oceanography,
-among others, to organise multivariate data in a spatio-temporal grid.
-
-\textcolor{red}{think through a cube of long x lat x time vs. id x time x variable}
+Spatio-temporal data concerns variables observed in the space across
+time and this paper focuses on the spatio-temporal vector data with
+fixed locations and regular time interval. This type of data can
+naturally been considered as a data cube and here a cube with three
+axes: spatial identifier, temporal identifier, and variable can be used
+to conceptualise spatio-temporal data. In this cubic formation, space
+only occupy one dimension, rather than two, to avoid the needs of
+hypercube for multivariate spatio-temporal data.
+
+While conceptually the data format is relatively simple, the actual data
+object that comes to the analysts in the analysis can be in various
+forms. Figure \ref{fig:illu-input} shows four examples of wild
+spatio-temporal data: 1) combines the time invariant and varying
+variables into a single table through duplicating the time invariant
+variables; 2) splits those variables into a separate spatial table and a
+temporal table; 3) \ldots; and lastly 4) is array-based data, like
+NetCDF (Network Common Data Form) data, which is commonly used in earth
+science, e.g.~climatology, meteorology, and oceanography, among others,
+to organise multivariate data in a spatio-temporal grid.
 
@@
 
-\caption[An illustration of different spatio-temporal data format]{An illustration of different spatio-temporal data format: (1) a single table for all the variables; (2) separate tables for spatial and temporal variables; (3) a table of metadata used to query the database and a separate table for queried data; and (4) an array data structure of cube or hypercube.}\label{fig:illu-input}
+\caption[An illustration of different spatio-temporal data format]{An illustration of different spatio-temporal data format: (1) a single table for all the variables; (2) separate tables for spatial and temporal variables; (3) a table of metadata used to query the database and a separate table for queried data; and (4) an array data structure.}\label{fig:illu-input}
 \end{figure}
@@
 
-\textcolor{red}{more beginning like "working with such data different authors ..."}
-Currently, there's various different spatio-temporal data structures in
-R includes: \pkg{spacetime} \citep{spacetime}, which proposes four
-space-time layouts: Full grid (STF), sparse grid(STS), irregular (STI),
-and trajectory (STT). \pkg{spacetime} is built on \pkg{sp} \citep{sp}
-and \pkg{xts} \citep{xts}, both of which have been replaced by more
-recent implementations. \pkg{spatstat} \citep{spatstat} implements a
-\texttt{ppp} class for point pattern data; and more recent, \pkg{stars}
-\citep{stars} implements a spatio-temporal for vector and raster
-spatio-temporal data.
-
+Different authors have proposed various data objects to work with
+spatial temporal data in \proglang{R} and this includes: \pkg{spacetime}
+\citep{spacetime} is built on the \pkg{sp} \citep{sp} and \pkg{xts}
+\citep{xts} class and both of which have been replaced by more recent
+spatial and temporal data objects. \pkg{spatstat} \citep{spatstat}
+implements a \texttt{ppp} class for analysing point pattern data; and
+more recent, \pkg{stars} \citep{stars} has implemented an
+spatio-temporal array for wrangling both vector and raster data.
+
+\textcolor{red}{Think again on the motivation and why it is hard to work existing tools.}
 In spatio-temporal data analysis, analysts will receive one of the
@@
 rearrange the data to fit into one of the data object in \proglang{R}
-before starting any analysis on the data itself. Looking at spatial or
-temporal data analysis, data objects like \pkg{sf} \citep{sf} and
-\pkg{tsibble} \citep{tsibble} have smoothed the data analysis in these
-two domains while there has not yet been a spatio-temporal data
-structure that makes it easy to explore spatio-temporal data and this
-creates a gap in the software development. The requirement for such a
-tool is important given the ubiquity of spatio-temporal vector data in
-the wild: climate observations from weather stations, air quality
-variables from air monitoring stations, or any variables that are
-measured at fixed location across time.
+before starting any analysis on the data itself.
+
+Looking at spatial or temporal data analysis, data objects like \pkg{sf}
+\citep{sf} and \pkg{tsibble} \citep{tsibble} have smoothed the data
+analysis in these two domains while there has not yet been a
+spatio-temporal data structure that makes it easy to explore
+spatio-temporal data and this creates a gap in the software development.
+The requirement for such a tool is important given the ubiquity of
+spatio-temporal vector data in the wild: climate observations from
+weather stations, air quality variables from air monitoring stations, or
+any variables that are measured at fixed location across time.
 

### 2022-02-20 1-intro.Rmd
further refine the motivation for a new spatio-temporal data structure

@@
 In this cubic formation, space only occupy one dimension, rather than two, to avoid the needs of hypercube for multivariate spatio-temporal data.
-
 While conceptually the data format is relatively simple, the actual data object that comes to the analysts in the analysis can be in various forms.
@@
 
-Different authors have proposed various data objects to work with spatial temporal data in \proglang{R} and this includes: \pkg{spacetime} [@spacetime] is built on the \pkg{sp} [@sp] and \pkg{xts} [@xts] class and both of which have been replaced by more recent spatial and temporal data objects.
+Different authors have proposed various data objects to work with spatial temporal data in \proglang{R}.
+\pkg{spacetime} [@spacetime] is built on the \pkg{sp} [@sp] and \pkg{xts} [@xts] class and both of which have been replaced by more recent spatial and temporal data objects.
 \pkg{spatstat} [@spatstat] implements a `ppp` class for analysing point pattern data; and more recent, \pkg{stars} [@stars] has implemented an spatio-temporal array for wrangling both vector and raster data.
 
-\textcolor{red}{Think again on the motivation and why it is hard to work existing tools.} In spatio-temporal data analysis, analysts will receive one of the various data formats depends on the data provider and need to manually rearrange the data to fit into one of the data object in \proglang{R} before starting any analysis on the data itself.
+None of these three formats in Figure \ref{fig:illu-input} is not particularly convenient to work with.
+To augment a one-site-per-row summary with the original spatial variables in a single table format (example 1 in Figure \ref{fig:illu-input}), analysts will need to first make a distinct one-site-per-row summary of the original spatial variables before the join.
+This step adds few value to the analyais but is a tedious step that first select all the spatial variables and then find the distinct row of each.
+In another type of operation that filter the sites based on per-site summary, intermediate objects, like the per-station summary and the filtered station id vector, will need be created as standalone objects in the environment.
+The spirit of using pipe reduces the intermediate data objects that are unnecessarily created in the analysis while in this case, these only-used-once data objects are hard to be incorporated into a piping workflow.
+In addition, if the data is organised as a separate spatial and temporal table as example 2 in Figure \ref{fig:illu-input}, the filter on site ID will need to performed twice in each table to keep them both up-to-date.
+With the array data structure, although it can easily store and wrangle large spatio-temporal data, the 3D or higher array can be less intuitive than a 2D tabular data in data wrangling.
 
-Looking at spatial or temporal data analysis, data objects like \pkg{sf} [@sf] and \pkg{tsibble} [@tsibble] have smoothed the data analysis in these two domains while there has not yet been a spatio-temporal data structure that makes it easy to explore spatio-temporal data and this creates a gap in the software development.
+Data objects like \pkg{sf} [@sf] and \pkg{tsibble} [@tsibble] have smoothed the data analysis in spatial or temporal data analysis, while there has not yet been a spatio-temporal data structure that makes it easy to explore spatio-temporal data and this creates a gap in the software development.
 The requirement for such a tool is important given the ubiquity of spatio-temporal vector data in the wild: climate observations from weather stations, air quality variables from air monitoring stations, or any variables that are measured at fixed location across time.
@@
 
-This paper describes the implementation of a new spatio-temporal data structure: \pkg{cubble}.
-\pkg{Cubble} implements a data structure that uses two forms to manage the switch between spatial and temporal dimension.
-With this structure, users can manipulate the spatial or temporal dimension separately, while leaves the linking of two dimensions to \pkg{cubble}.
+This paper describes the a new tabular data object:  \pkg{cubble} for wrangling spatio-temporal data.
+\pkg{Cubble} overcomes the inconvenience in data wrangling mentioned above through two forms: a nested form to organise spatial variables and a long form to organise temporal variables.
+With this structure, analysts can explore the spatio-temporal data with less efforts spend on organising variables at different observational units.
 The software is available from the Comprehensive R Archive Network (CRAN) at [CRAN link].

### 2022-02-20 main.tex
further refine the motivation for a new spatio-temporal data structure

@@
 only occupy one dimension, rather than two, to avoid the needs of
-hypercube for multivariate spatio-temporal data.
-
-While conceptually the data format is relatively simple, the actual data
-object that comes to the analysts in the analysis can be in various
-forms. Figure \ref{fig:illu-input} shows four examples of wild
-spatio-temporal data: 1) combines the time invariant and varying
-variables into a single table through duplicating the time invariant
-variables; 2) splits those variables into a separate spatial table and a
-temporal table; 3) \ldots; and lastly 4) is array-based data, like
-NetCDF (Network Common Data Form) data, which is commonly used in earth
-science, e.g.~climatology, meteorology, and oceanography, among others,
-to organise multivariate data in a spatio-temporal grid.
+hypercube for multivariate spatio-temporal data. While conceptually the
+data format is relatively simple, the actual data object that comes to
+the analysts in the analysis can be in various forms. Figure
+\ref{fig:illu-input} shows four examples of wild spatio-temporal data:
+1) combines the time invariant and varying variables into a single table
+through duplicating the time invariant variables; 2) splits those
+variables into a separate spatial table and a temporal table; 3) \ldots;
+and lastly 4) is array-based data, like NetCDF (Network Common Data
+Form) data, which is commonly used in earth science, e.g.~climatology,
+meteorology, and oceanography, among others, to organise multivariate
+data in a spatio-temporal grid.
 
@@
 Different authors have proposed various data objects to work with
-spatial temporal data in \proglang{R} and this includes: \pkg{spacetime}
-\citep{spacetime} is built on the \pkg{sp} \citep{sp} and \pkg{xts}
-\citep{xts} class and both of which have been replaced by more recent
-spatial and temporal data objects. \pkg{spatstat} \citep{spatstat}
-implements a \texttt{ppp} class for analysing point pattern data; and
-more recent, \pkg{stars} \citep{stars} has implemented an
-spatio-temporal array for wrangling both vector and raster data.
-
-\textcolor{red}{Think again on the motivation and why it is hard to work existing tools.}
-In spatio-temporal data analysis, analysts will receive one of the
-various data formats depends on the data provider and need to manually
-rearrange the data to fit into one of the data object in \proglang{R}
-before starting any analysis on the data itself.
-
-Looking at spatial or temporal data analysis, data objects like \pkg{sf}
-\citep{sf} and \pkg{tsibble} \citep{tsibble} have smoothed the data
-analysis in these two domains while there has not yet been a
-spatio-temporal data structure that makes it easy to explore
-spatio-temporal data and this creates a gap in the software development.
-The requirement for such a tool is important given the ubiquity of
-spatio-temporal vector data in the wild: climate observations from
-weather stations, air quality variables from air monitoring stations, or
-any variables that are measured at fixed location across time.
-
-This paper describes the implementation of a new spatio-temporal data
-structure: \pkg{cubble}. \pkg{Cubble} implements a data structure that
-uses two forms to manage the switch between spatial and temporal
-dimension. With this structure, users can manipulate the spatial or
-temporal dimension separately, while leaves the linking of two
-dimensions to \pkg{cubble}. The software is available from the
+spatial temporal data in \proglang{R}. \pkg{spacetime} \citep{spacetime}
+is built on the \pkg{sp} \citep{sp} and \pkg{xts} \citep{xts} class and
+both of which have been replaced by more recent spatial and temporal
+data objects. \pkg{spatstat} \citep{spatstat} implements a \texttt{ppp}
+class for analysing point pattern data; and more recent, \pkg{stars}
+\citep{stars} has implemented an spatio-temporal array for wrangling
+both vector and raster data.
+
+None of these three formats in Figure \ref{fig:illu-input} is not
+particularly convenient to work with. To augment a one-site-per-row
+summary with the original spatial variables in a single table format
+(example 1 in Figure \ref{fig:illu-input}), analysts will need to first
+make a distinct one-site-per-row summary of the original spatial
+variables before the join. This step adds few value to the analyais but
+is a tedious step that first select all the spatial variables and then
+find the distinct row of each. In another type of operation that filter
+the sites based on per-site summary, intermediate objects, like the
+per-station summary and the filtered station id vector, will need be
+created as standalone objects in the environment. The spirit of using
+pipe reduces the intermediate data objects that are unnecessarily
+created in the analysis while in this case, these only-used-once data
+objects are hard to be incorporated into a piping workflow. In addition,
+if the data is organised as a separate spatial and temporal table as
+example 2 in Figure \ref{fig:illu-input}, the filter on site ID will
+need to performed twice in each table to keep them both up-to-date. With
+the array data structure, although it can easily store and wrangle large
+spatio-temporal data, the 3D or higher array can be less intuitive than
+a 2D tabular data in data wrangling.
+
+Data objects like \pkg{sf} \citep{sf} and \pkg{tsibble} \citep{tsibble}
+have smoothed the data analysis in spatial or temporal data analysis,
+while there has not yet been a spatio-temporal data structure that makes
+it easy to explore spatio-temporal data and this creates a gap in the
+software development. The requirement for such a tool is important given
+the ubiquity of spatio-temporal vector data in the wild: climate
+observations from weather stations, air quality variables from air
+monitoring stations, or any variables that are measured at fixed
+location across time.
+
+This paper describes the a new tabular data object: \pkg{cubble} for
+wrangling spatio-temporal data. \pkg{Cubble} overcomes the inconvenience
+in data wrangling mentioned above through a nested form to organise
+spatial variables and a long form to organise temporal variables. With
+this structure, analysts can explore the spatio-temporal data with less
+efforts spend on \ldots{} The software is available from the
 Comprehensive R Archive Network (CRAN) at {[}CRAN link{]}.

### 2022-02-21 1-intro.Rmd
intro done

--- 2022-02-17 1-intro.Rmd
@@
 
-Spatio-temporal data concerns variables observed in the space across time and this paper focuses on the spatio-temporal vector data with fixed locations and regular time interval.
-This type of data can naturally been considered as a data cube and here a cube with three axes: spatial identifier, temporal identifier, and variable can be used to conceptualise spatio-temporal data.
-In this cubic formation, space only occupy one dimension, rather than two, to avoid the needs of hypercube for multivariate spatio-temporal data.
-
-While conceptually the data format is relatively simple, the actual data object that comes to the analysts in the analysis can be in various forms.
-Figure \ref{fig:illu-input} shows four examples of wild spatio-temporal data: 1) combines the time invariant and varying variables into a single table through duplicating the time invariant variables; 2) splits those variables into a separate spatial table and a temporal table; 3) ...; and lastly 4) is array-based data, like NetCDF (Network Common Data Form) data, which is commonly used in earth science, e.g.
+Spatio-temporal data concerns variables observed in the space across time and this paper focuses on spatio-temporal vector data with fixed locations and regular time interval.
+This type of data can naturally been considered as a data cube with three axes: spatial identifier, temporal identifier, and variable.
+In this cubic formation, space only occupies one axis, rather than two, to avoid hypercubes for multivariate spatio-temporal data.
+While conceptually the data format is relatively simple, the data objects coming to the analysts can vary in forms in the analysis.
+Figure \ref{fig:illu-input} shows four examples of wild spatio-temporal data: 1) a single table that includes both the spatial and temporal variables; 2) separate tables for spatial and temporal variables; 3) array-based data, i.e.
+NetCDF (Network Common Data Form) data, which is commonly used in earth science, e.g.
 climatology, meteorology, and oceanography, among others, to organise multivariate data in a spatio-temporal grid.
 
-```{r illu-input, echo = FALSE, out.width = "100%", fig.cap="An illustration of different spatio-temporal data format: (1) a single table for all the variables; (2) separate tables for spatial and temporal variables; (3) a table of metadata used to query the database and a separate table for queried data; and (4) an array data structure.", fig.align='center', out.height = "40%"}
+```{r illu-input, echo = FALSE, out.width = "100%", fig.cap="An illustration of different spatio-temporal data layout: (1) a single table for both spatial and temporal variables; (2) separate tables for spatial and temporal variables; and (3) an array data structure that populates variables into a gridded longitude-latitude-time (hyper)cube.", fig.align='center', out.height = "40%"}
 knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.001.png"))
@@
 
-Different authors have proposed various data objects to work with spatial temporal data in \proglang{R} and this includes: \pkg{spacetime} [@spacetime] is built on the \pkg{sp} [@sp] and \pkg{xts} [@xts] class and both of which have been replaced by more recent spatial and temporal data objects.
-\pkg{spatstat} [@spatstat] implements a `ppp` class for analysing point pattern data; and more recent, \pkg{stars} [@stars] has implemented an spatio-temporal array for wrangling both vector and raster data.
+None of these three formats in Figure \ref{fig:illu-input} is not particularly convenient to work with.
+In the single table format, to avoid plotting duplicated points of the same sites on the map, analysts always need to create a spatial variables table from the original single table.
+This can be a tedious step if needs to performed many times during the analysis.
+In another type of operation that filter the sites based on certain statistics, intermediate objects, like the per-station summary of the statistic and the filtered station ID vector, will need be created as standalone objects in the environment.
+The spirit of using pipe reduces the intermediate data objects that are unnecessarily created in the analysis while in this case, these only-used-once data objects can hardly be incorporated into a piping workflow in the analysis.
+In addition, if the data is organised as a separate spatial and temporal table as format 2 in Figure \ref{fig:illu-input}, the filter on site ID will need to performed twice in each table to keep them both up-to-date.
+With the array data structure, although it can easily store and wrangle large spatio-temporal data, the 3D or higher-D array can be less intuitive than a 2D tabular data structure for some analysts in data wrangling.
 
-\textcolor{red}{Think again on the motivation and why it is hard to work existing tools.} In spatio-temporal data analysis, analysts will receive one of the various data formats depends on the data provider and need to manually rearrange the data to fit into one of the data object in \proglang{R} before starting any analysis on the data itself.
-
-Looking at spatial or temporal data analysis, data objects like \pkg{sf} [@sf] and \pkg{tsibble} [@tsibble] have smoothed the data analysis in these two domains while there has not yet been a spatio-temporal data structure that makes it easy to explore spatio-temporal data and this creates a gap in the software development.
-The requirement for such a tool is important given the ubiquity of spatio-temporal vector data in the wild: climate observations from weather stations, air quality variables from air monitoring stations, or any variables that are measured at fixed location across time.
+In the \proglang{R} community, different data objects have been proposed to work with spatio-temporal data.
+\pkg{spacetime} [@spacetime] is built on class \pkg{sp} [@sp] and \pkg{xts} [@xts] and both of which have been replaced by more recent spatial and temporal data objects.
+\pkg{spatstat} [@spatstat] implements a `ppp` class for analysing point pattern data; and more recent, \pkg{stars} [@stars] has implemented a spatio-temporal array for wrangling both vector and raster data.
+While these data objects are useful for its designed tasks, there has not yet been a data object that makes it easy to explore spatio-temporal data.
+Data objects like \pkg{sf} [@sf] and \pkg{tsibble} [@tsibble] are good examples of this in the spatial and temporal data analysis and hence there is a gap in software development for spatio-temporal data.
+The requirement for such a tool is important given the ubiquity of spatio-temporal vector data in the wild: climate observations from weather stations, air quality variables from air monitoring stations, or any data collected at fixed location across time.
 
@@
 
-This paper describes the implementation of a new spatio-temporal data structure: \pkg{cubble}.
-\pkg{Cubble} implements a data structure that uses two forms to manage the switch between spatial and temporal dimension.
-With this structure, users can manipulate the spatial or temporal dimension separately, while leaves the linking of two dimensions to \pkg{cubble}.
+This paper describes a new tabular data object:  \pkg{cubble} for wrangling spatio-temporal data.
+\pkg{Cubble} overcomes the inconvenience in data wrangling mentioned above through wiring the separate spatial and temporal table into a single data object.
+With the new data object, analysts can spend less time organising spatial and temporal variables at different observational units and focus more on exploring the data itself.
 The software is available from the Comprehensive R Archive Network (CRAN) at [CRAN link].
 
-The rest of the paper is divided as follows: Section 2 presents the workflow of data manipulation in cubble.
-Section 3 explains how cubble deals with some advanced considerations  including data with hierarchical structure, data matching, and how cubble fits with existing static and interactive visualisation tools.
-Section 4 gives examples of the features introduced in the previous two sections with climate and hydrology data.
-An example on how cubble handle NetCDF data is also provided.
+The rest of the paper is divided as follows: Section 2 presents the main design and functionality of \pkg{cubble}.
+Section 3 explains how cubble deals with more advanced considerations including data with hierarchical structure, data matching, how cubble fits with existing static and interactive visualisation tools, and spatio-temporal data transformation.
+Section 4 uses Australia weather station data and river level data as examples to demonstrate the use with \pkg{cubble}.
+An example on how \pkg{cubble} handles NetCDF data is also provided.
 Section 5 concludes the paper.

### 2022-02-21 main.tex
intro done

--- 2022-02-17 main.tex
@@
 Spatio-temporal data concerns variables observed in the space across
-time and this paper focuses on the spatio-temporal vector data with
-fixed locations and regular time interval. This type of data can
-naturally been considered as a data cube and here a cube with three
-axes: spatial identifier, temporal identifier, and variable can be used
-to conceptualise spatio-temporal data. In this cubic formation, space
-only occupy one dimension, rather than two, to avoid the needs of
-hypercube for multivariate spatio-temporal data.
-
-While conceptually the data format is relatively simple, the actual data
-object that comes to the analysts in the analysis can be in various
-forms. Figure \ref{fig:illu-input} shows four examples of wild
-spatio-temporal data: 1) combines the time invariant and varying
-variables into a single table through duplicating the time invariant
-variables; 2) splits those variables into a separate spatial table and a
-temporal table; 3) \ldots; and lastly 4) is array-based data, like
-NetCDF (Network Common Data Form) data, which is commonly used in earth
-science, e.g.~climatology, meteorology, and oceanography, among others,
-to organise multivariate data in a spatio-temporal grid.
+time and this paper focuses on spatio-temporal vector data with fixed
+locations and regular time interval. This type of data can naturally
+been considered as a data cube with three axes: spatial identifier,
+temporal identifier, and variable. In this cubic formation, space only
+occupies one axis, rather than two, to avoid hypercubes for multivariate
+spatio-temporal data. While conceptually the data format is relatively
+simple, the data objects coming to the analysts can vary in forms in the
+analysis. Figure \ref{fig:illu-input} shows four examples of wild
+spatio-temporal data: 1) a single table that includes both the spatial
+and temporal variables; 2) separate tables for spatial and temporal
+variables; 3) array-based data, i.e.~NetCDF (Network Common Data Form)
+data, which is commonly used in earth science, e.g.~climatology,
+meteorology, and oceanography, among others, to organise multivariate
+data in a spatio-temporal grid.
 
@@
 
-\caption[An illustration of different spatio-temporal data format]{An illustration of different spatio-temporal data format: (1) a single table for all the variables; (2) separate tables for spatial and temporal variables; (3) a table of metadata used to query the database and a separate table for queried data; and (4) an array data structure.}\label{fig:illu-input}
+\caption[An illustration of different spatio-temporal data layout]{An illustration of different spatio-temporal data layout: (1) a single table for both spatial and temporal variables; (2) separate tables for spatial and temporal variables; and (3) an array data structure that populates variables into a gridded longitude-latitude-time (hyper)cube.}\label{fig:illu-input}
 \end{figure}
@@
 
-Different authors have proposed various data objects to work with
-spatial temporal data in \proglang{R} and this includes: \pkg{spacetime}
-\citep{spacetime} is built on the \pkg{sp} \citep{sp} and \pkg{xts}
-\citep{xts} class and both of which have been replaced by more recent
-spatial and temporal data objects. \pkg{spatstat} \citep{spatstat}
-implements a \texttt{ppp} class for analysing point pattern data; and
-more recent, \pkg{stars} \citep{stars} has implemented an
-spatio-temporal array for wrangling both vector and raster data.
-
-\textcolor{red}{Think again on the motivation and why it is hard to work existing tools.}
-In spatio-temporal data analysis, analysts will receive one of the
-various data formats depends on the data provider and need to manually
-rearrange the data to fit into one of the data object in \proglang{R}
-before starting any analysis on the data itself.
-
-Looking at spatial or temporal data analysis, data objects like \pkg{sf}
-\citep{sf} and \pkg{tsibble} \citep{tsibble} have smoothed the data
-analysis in these two domains while there has not yet been a
-spatio-temporal data structure that makes it easy to explore
-spatio-temporal data and this creates a gap in the software development.
-The requirement for such a tool is important given the ubiquity of
-spatio-temporal vector data in the wild: climate observations from
-weather stations, air quality variables from air monitoring stations, or
-any variables that are measured at fixed location across time.
-
-This paper describes the implementation of a new spatio-temporal data
-structure: \pkg{cubble}. \pkg{Cubble} implements a data structure that
-uses two forms to manage the switch between spatial and temporal
-dimension. With this structure, users can manipulate the spatial or
-temporal dimension separately, while leaves the linking of two
-dimensions to \pkg{cubble}. The software is available from the
-Comprehensive R Archive Network (CRAN) at {[}CRAN link{]}.
-
-The rest of the paper is divided as follows: Section 2 presents the
-workflow of data manipulation in cubble. Section 3 explains how cubble
-deals with some advanced considerations including data with hierarchical
-structure, data matching, and how cubble fits with existing static and
-interactive visualisation tools. Section 4 gives examples of the
-features introduced in the previous two sections with climate and
-hydrology data. An example on how cubble handle NetCDF data is also
-provided. Section 5 concludes the paper.
+None of these three formats in Figure \ref{fig:illu-input} is not
+particularly convenient to work with. In the single table format, to
+avoid plotting duplicated points of the same sites on the map, analysts
+always need to create a spatial variables table from the original single
+table. This can be a tedious step if needs to performed many times
+during the analysis. In another type of operation that filter the sites
+based on certain statistics, intermediate objects, like the per-station
+summary of the statistic and the filtered station ID vector, will need
+be created as standalone objects in the environment. The spirit of using
+pipe reduces the intermediate data objects that are unnecessarily
+created in the analysis while in this case, these only-used-once data
+objects can hardly be incorporated into a piping workflow in the
+analysis. In addition, if the data is organised as a separate spatial
+and temporal table as format 2 in Figure \ref{fig:illu-input}, the
+filter on site ID will need to performed twice in each table to keep
+them both up-to-date. With the array data structure, although it can
+easily store and wrangle large spatio-temporal data, the 3D or higher-D
+array can be less intuitive than a 2D tabular data structure for some
+analysts in data wrangling.
+
+In the \proglang{R} community, different data objects have been proposed
+to work with spatio-temporal data. \pkg{spacetime} \citep{spacetime} is
+built on class \pkg{sp} \citep{sp} and \pkg{xts} \citep{xts} and both of
+which have been replaced by more recent spatial and temporal data
+objects. \pkg{spatstat} \citep{spatstat} implements a \texttt{ppp} class
+for analysing point pattern data; and more recent, \pkg{stars}
+\citep{stars} has implemented a spatio-temporal array for wrangling both
+vector and raster data. While these data objects are useful for its
+designed tasks, there has not yet been a data object that makes it easy
+to explore spatio-temporal data. Data objects like \pkg{sf} \citep{sf}
+and \pkg{tsibble} \citep{tsibble} are good examples of this in the
+spatial and temporal data analysis and hence there is a gap in software
+development for spatio-temporal data. The requirement for such a tool is
+important given the ubiquity of spatio-temporal vector data in the wild:
+climate observations from weather stations, air quality variables from
+air monitoring stations, or any data collected at fixed location across
+time.
+
+This paper describes a new tabular data object: \pkg{cubble} for
+wrangling spatio-temporal data. \pkg{Cubble} overcomes the inconvenience
+in data wrangling mentioned above through wiring the separate spatial
+and temporal table into a single data object. With the new data object,
+analysts can spend less time organising spatial and temporal variables
+at different observational units and focus more on exploring the data
+itself. The software is available from the Comprehensive R Archive
+Network (CRAN) at {[}CRAN link{]}.
+
+The rest of the paper is divided as follows: Section 2 presents the main
+design and functionality of \pkg{cubble}. Section 3 explains how cubble
+deals with more advanced considerations including data with hierarchical
+structure, data matching, how cubble fits with existing static and
+interactive visualisation tools, and spatio-temporal data
+transformation. Section 4 uses Australia weather station data and river
+level data as examples to demonstrate the use with \pkg{cubble}. An
+example on how \pkg{cubble} handles NetCDF data is also provided.
+Section 5 concludes the paper.
 
@@
 
-In cubble, data can represented in two formats: nested form and long
-form, and Figure \ref{fig:illu-cubble} sketches the two forms with the
-associated attributes. The decision on which form to use in data
-manipulation is output-oriented, meaning analysts need to first think
-about whether the output of an operation is identified only by the
+In cubble, a single data object can be arranged into two forms: nested
+form and long form, and Figure \ref{fig:illu-cubble} sketches the two
+forms with the associated attributes. The decision on which form to use
+in data wrangling is output-oriented, meaning analysts need to first
+think about whether the output of an operation is identified only by the
 spatial identifier, or a combination of spatial and temporal identifier.

### 2022-02-21 main.Rmd
more work on abstract

--- 2022-02-07 main.Rmd
@@
 abstract: >
-  Spatio-temporal variables can be time invariant or not and this divides them into those can be identified only with the spatial identifier and those cross-identified by the spatial and temporal identifier.
-In this paper, a new data structure, \code{cubble}, is proposed for manipulating and visualising spatio-temporal data.
-The new data structure proposes two forms for manipulating the spatial and temporal dimension separately while providing a smooth switch between the two forms to link the two dimensions.
-Advanced considerations are given to illustrate data with hierarchical structure, data matching, and how \pkg{cubble} fits with interactive graphics pipeline.
+  Exploring spatio-temporal data can sometimes be discouraging given the various different formats spatio-temporal data can present themselves in practice.
+Moreover, the two observational units naturally exist in spatio-temporal data can cost analysts many time and efforts on organising variables and results and switching between a single table and multiple tables in the analysis.
+In this paper, a new data structure, \code{cubble}, is proposed for wrangling and visualising spatio-temporal data.
+The new data structure uses a single data object to organise spatial and temporal variables in two forms so that they can be wrangled separately while keep syncronised.
+Advanced considerations are given to illustrate data with hierarchical structure, data matching, how cubble fits with the interactive graphics pipeline, and spatio-temporal transformation.
 Examples are given to analysing Australia meteorology data, river level data, and climate reanalysis (ERA5).

### 2022-02-21 main.tex
more work on abstract

@@
 \Abstract{
-Spatio-temporal variables can be divided into time invariant ones that
-can be identified by the spatial identifier, or the time varying ones
-that cross-identified by the spatial and temporal identifiers. In this
-paper, a new data structure, \code{cubble}, is proposed for manipulating
-and visualising spatio-temporal data. The new data structure uses two
-forms to organise time invariant and varying variables so that the
-spatial and temporal dimension of the data can be manipulated separately
-while keep synchronised. Advanced considerations are given to illustrate
-data with hierarchical structure, data matching, and how cubble fits
-with the interactive graphics pipeline. Examples are given to analysing
+Exploring spatio-temporal data can sometimes be discouraging given the
+various different formats spatio-temporal data can present themselves in
+practice. Moreover, the two observational units naturally exist in
+spatio-temporal data can cost analysts many time and efforts on
+organising variables and results and switching between a single table
+and multiple tables in the analysis. In this paper, a new data
+structure, \code{cubble}, is proposed for wrangling and visualising
+spatio-temporal data. The new data structure uses a single data object
+to organise spatial and temporal variables in two forms so that they can
+be wrangled separately while keep syncronised. Advanced considerations
+are given to illustrate data with hierarchical structure, data matching,
+how cubble fits with the interactive graphics pipeline, and
+spatio-temporal transformation. Examples are given to analysing
 Australia meteorology data, river level data, and climate reanalysis

### 2022-02-21 2-the-cubble-pkg.Rmd
work on beginning of matching

@@
 
-In cubble, data can represented in two formats: nested form and long form, and Figure \ref{fig:illu-cubble} sketches the two forms with the associated attributes.
-The decision on which form to use in data manipulation is output-oriented, meaning analysts need to first think about whether the output of an operation is identified only by the spatial identifier, or a combination of spatial and temporal identifier.
+In cubble, a single data object can be arranged into two forms: nested form and long form, and Figure \ref{fig:illu-cubble} sketches the two forms with the associated attributes.
+The decision on which form to use in data wrangling is output-oriented, meaning analysts need to first think about whether the output of an operation is identified only by the spatial identifier, or a combination of spatial and temporal identifier.
 The nested cubble is suitable for working with operations that are only identified by the site and this type of operation can be a pure manipulation of time invariant variables, or an operation that summarises time varying variables into site.
@@
  > `mutate`, `filter`, `summarise`, `select`, `arrange`, `rename`, `left_join`, and the slice family (`slice_head`, `slice_tail`, `slice_sample`, `slice_min`, `slice_max`)
-
-\textcolor{red}{more work on groupby} The grouping pair, \code{group_by} and \code{ungroup}, can only be used in the long form to define grouping on the time but not in the nested form.
 

### 2022-02-21 3-advanced.Rmd
go through writing made yesterday
work on beginning of matching

--- 2022-02-08 3-advanced.Rmd
@@
 
-Cubble provides functionality for matching sites from two data sources. To be a valid pair of matching, the matched pair from different data sources need to:
-
-  - be spatially close to each other, and 
-  - have similar temporal movement
-
-\code{match_sites()} matches data based on these two criteria: it first matches the two data sources spatially through computing the pairwise distance on latitude and longitude.
+One task that may interest analysts in spatio-temporal data is to find how similar the time series from nearby sites are.
+This can be seen as a matching problem [@stuart2010matching; @mcintosh2018using] that pairs up similar similar time series in nearby locations or a data fusing exercise that merges data collected from different sources [@cocchi2019data].
+\code{match_sites()} in \pkg{cubble} provides a simple algorithm for this task.
+The algorithm first matches the two data sources spatially through computing the pairwise distance on latitude and longitude.
 Pairs that pass the spatial matching are then matched temporally through computing the number of matched peaks within a fixed length window.
@@
  - \code{temporal_min_match}: the minimum number of matched peak for a valid matched pair
-
-This functionality can be seen as a matching exercise [@stuart2010matching; @mcintosh2018using] in the spatio-temporal domain to pair up nearby observations.
-It can also be considered as a form of data fusion [@castanedo2013review; @cocchi2019data], where data from multiple sources are combined together.
 
@@
 
-## Spatio-temporal transformations
+## Spatio-temporal transformations {#st_transformation}
 

### 2022-02-21 main.tex
work on beginning of matching

@@
 \end{quote}
-
-\textcolor{red}{more work on groupby} The grouping pair, \code{group_by}
-and \code{ungroup}, can only be used in the long form to define grouping
-on the time but not in the nested form.
 
@@
 
-\textcolor{red}{more beginning} Cubble provides functionality for
-matching sites from two data sources. To be a valid pair of matching,
-the matched pair from different data sources need to:
-
-\begin{itemize}
-\tightlist
-\item
-  be spatially close to each other, and
-\item
-  have similar temporal movement
-\end{itemize}
-
-\code{match_sites()} matches data based on these two criteria: it first
-matches the two data sources spatially through computing the pairwise
-distance on latitude and longitude. Pairs that pass the spatial matching
-are then matched temporally through computing the number of matched
-peaks within a fixed length window. Figure \ref{fig:illu-matching}
-illustrates this temporal matching in more details. Given two series
-\code{A} and \code{a}, 3 peaks have been picked in each series. An
-interval, with default length of 5, is constructed for each peak in
-series \code{A} and the peaks in series \code{a} are tested against
-whether they fall into the any of the intervals. In this illustration,
-there are 2 matches for these two series. Several arguments are
-available in \code{match_sites()} to fine-tune the matching:
+One task that may interest analysts in spatio-temporal data is to find
+how similar the time series from nearby sites are. This can be seen as a
+matching problem \citep{stuart2010matching, mcintosh2018using} that
+pairs up similar similar time series in nearby locations or a data
+fusing exercise that merges data collected from different sources
+\citep{cocchi2019data}. \code{match_sites()} in \pkg{cubble} provides a
+simple algorithm for this task. The algorithm first matches the two data
+sources spatially through computing the pairwise distance on latitude
+and longitude. Pairs that pass the spatial matching are then matched
+temporally through computing the number of matched peaks within a fixed
+length window. Figure \ref{fig:illu-matching} illustrates this temporal
+matching in more details. Given two series \code{A} and \code{a}, 3
+peaks have been picked in each series. An interval, with default length
+of 5, is constructed for each peak in series \code{A} and the peaks in
+series \code{a} are tested against whether they fall into the any of the
+intervals. In this illustration, there are 2 matches for these two
+series. Several arguments are available in \code{match_sites()} to
+fine-tune the matching:
 
@@
 
-This functionality can be seen as a matching exercise
-\citep{stuart2010matching, mcintosh2018using} in the spatio-temporal
-domain to pair up nearby observations. It can also be considered as a
-form of data fusion \citep{cocchi2019data}, where data from multiple
-sources are combined together.
-
 \hypertarget{interactive-graphics}{%

### 2022-02-21 1-intro.Rmd
notes of another round of changes to make in the introduction

@@
 
-Spatio-temporal data concerns variables observed in the space across time and this paper focuses on spatio-temporal vector data with fixed locations and regular time interval.
+Spatio-temporal data involves variables observed in the space across time and this paper focuses on spatio-temporal vector data with fixed locations and regular time interval.
 This type of data can naturally been considered as a data cube with three axes: spatial identifier, temporal identifier, and variable.
@@
 
-```{r illu-input, echo = FALSE, out.width = "100%", fig.cap="An illustration of different spatio-temporal data layout: (1) a single table for both spatial and temporal variables; (2) separate tables for spatial and temporal variables; and (3) an array data structure that populates variables into a gridded longitude-latitude-time (hyper)cube.", fig.align='center', out.height = "40%"}
+```{r illu-input, echo = FALSE, out.width = "100%", fig.cap="An illustration of different spatio-temporal data layout: (1) a single table for both spatial and temporal variables; (2) separate tables for spatial and temporal variables; and (3) an array data structure that populates variables into a gridded longitude-latitude-time (hyper)cube.", fig.align='center'}
 knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.001.png"))
@@
 \newline -->
+
+\textcolor{red}{start with "we need something to do exploratory data analysis for spatio-temporal data", then why it is hard to work with. This section needs to be abstract a bit further.}
 

### 2022-02-21 main.Rmd
notes of another round of changes to make in the introduction

@@
 abstract: >
-  Exploring spatio-temporal data can sometimes be discouraging given the various different formats spatio-temporal data can present themselves in practice.
-Moreover, the two observational units naturally exist in spatio-temporal data can cost analysts many time and efforts on organising variables and results and switching between a single table and multiple tables in the analysis.
+
+  \textcolor{red}{start with describing the data: spatial and temporal, then what's the problem, then propose cubble.}
+
+  \textcolor{red}{Exploring spatio-temporal data can sometimes be discouraging given the various different formats spatio-temporal data can present themselves in practice}.
+Moreover, the two observational units naturally exist in spatio-temporal data \textcolor{red}{can cost analysts many time and efforts} on organising variables and results and switching between a single table and multiple tables in the analysis.
 In this paper, a new data structure, \code{cubble}, is proposed for wrangling and visualising spatio-temporal data.
@@
   # at least one keyword must be supplied
-  formatted: [spatio temporal,  "\\proglang{R}", climate weather station, data analysis]
-  plain:     [spatio temporal, R, climate weather station, data analysis]
+  formatted: [spatial, temporal, spatio temporal, "\\proglang{R}", environmental data, exploratory data analysis]
+  plain:     [spatial, temporal, spatio temporal, R, environmental data, exploratory data analysis]
 preamble: >

### 2022-02-21 main.tex
notes of another round of changes to make in the introduction

@@
 \Abstract{
-Exploring spatio-temporal data can sometimes be discouraging given the
-various different formats spatio-temporal data can present themselves in
-practice. Moreover, the two observational units naturally exist in
-spatio-temporal data can cost analysts many time and efforts on
+\textcolor{red}{start with describing the data: spatial and temporal, then what's the problem, then propose cubble.}
+\textcolor{red}{Exploring spatio-temporal data can sometimes be discouraging given the various different formats spatio-temporal data can present themselves in practice}.
+Moreover, the two observational units naturally exist in spatio-temporal
+data \textcolor{red}{can cost analysts many time and efforts} on
 organising variables and results and switching between a single table
@@
 
-\Keywords{spatio temporal, \proglang{R}, climate weather station, data
-analysis}
-\Plainkeywords{spatio temporal, R, climate weather station, data
-analysis}
+\Keywords{spatial, temporal, spatio
+temporal, \proglang{R}, environmental data, exploratory data analysis}
+\Plainkeywords{spatial, temporal, spatio temporal, R, environmental
+data, exploratory data analysis}
 
@@
 
-Spatio-temporal data concerns variables observed in the space across
+Spatio-temporal data involves variables observed in the space across
 time and this paper focuses on spatio-temporal vector data with fixed
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.001}
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.001}
 
@@
 \end{CodeChunk}
+
+\textcolor{red}{start with "we need something to do exploratory data analysis for spatio-temporal data", then why it is hard to work with. This section needs to be abstract a bit further.}
 

### 2022-02-23 main.Rmd
refine abstract again

@@
 abstract: >
-
-  \textcolor{red}{start with describing the data: spatial and temporal, then what's the problem, then propose cubble.}
-
-  \textcolor{red}{Exploring spatio-temporal data can sometimes be discouraging given the various different formats spatio-temporal data can present themselves in practice}.
-Moreover, the two observational units naturally exist in spatio-temporal data \textcolor{red}{can cost analysts many time and efforts} on organising variables and results and switching between a single table and multiple tables in the analysis.
-In this paper, a new data structure, \code{cubble}, is proposed for wrangling and visualising spatio-temporal data.
-The new data structure uses a single data object to organise spatial and temporal variables in two forms so that they can be wrangled separately while keep syncronised.
-Advanced considerations are given to illustrate data with hierarchical structure, data matching, how cubble fits with the interactive graphics pipeline, and spatio-temporal transformation.
-Examples are given to analysing Australia meteorology data, river level data, and climate reanalysis (ERA5).
+  Spatio-temporal data have two types of variables: spatial variables have fixed value across time in the same location while temporal variables need both the location and timestamp to identify its value.
+These variables can be organised in two separate tables, each focusing on one aspect of the data, or a single table that combines all the variables together.
+For exploratory data analysis and visualisation, analysts needs both formats for different types of exploration, while it can be tedious to constantly split and join variables in order to cast the data into the desireable shape.
+In this paper, a new data structure, \code{cubble}, is proposed for exploring spatio-temporal data.
+The new data structure organises spatial and temporal variables as two forms of a single data object so that they can be wrangled separately while keep syncronised.
+Considerations are given to data with hierarchical structure, data matching and explain how \pkg{cubble} fits in with the interactive graphics pipeline and spatio-temporal transformation.
+Examples are given to analysing Australia meteorology data, river level data, and climate reanalysis (ERA5) data.
 keywords:

### 2022-02-23 main.tex
refine abstract again

@@
 \usepackage[utf8]{inputenc}
-
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
 
@@
 \Abstract{
-\textcolor{red}{start with describing the data: spatial and temporal, then what's the problem, then propose cubble.}
-\textcolor{red}{Exploring spatio-temporal data can sometimes be discouraging given the various different formats spatio-temporal data can present themselves in practice}.
-Moreover, the two observational units naturally exist in spatio-temporal
-data \textcolor{red}{can cost analysts many time and efforts} on
-organising variables and results and switching between a single table
-and multiple tables in the analysis. In this paper, a new data
-structure, \code{cubble}, is proposed for wrangling and visualising
-spatio-temporal data. The new data structure uses a single data object
-to organise spatial and temporal variables in two forms so that they can
-be wrangled separately while keep syncronised. Advanced considerations
-are given to illustrate data with hierarchical structure, data matching,
-how cubble fits with the interactive graphics pipeline, and
-spatio-temporal transformation. Examples are given to analysing
-Australia meteorology data, river level data, and climate reanalysis
-(ERA5).
+Spatio-temporal data have two types of variables: spatial variables have
+fixed value across time in the same location while temporal variables
+need both the location and timestamp to identify its value. These
+variables can be organised in two separate tables, each focusing on one
+aspect of the data, or a single table that combines all the variables
+together. For exploratory data analysis and visualisation, analysts
+needs both formats for different types of exploration, while it can be
+tedious to constantly split and join variables in order to cast the data
+into the desireable shape. In this paper, a new data structure,
+\code{cubble}, is proposed for exploring spatio-temporal data. The new
+data structure organises spatial and temporal variables as two forms of
+a single data object so that they can be wrangled separately while keep
+syncronised. Considerations are given to data with hierarchical
+structure, data matching and explain how \pkg{cubble} fits in with the
+interactive graphics pipeline and spatio-temporal transformation.
+Examples are given to analysing Australia meteorology data, river level
+data, and climate reanalysis (ERA5) data.
 }
@@
 
-% Pandoc citation processing
-
-% Pandoc header
+% tightlist command for lists without linebreak
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
 

### 2022-02-24 main.Rmd
refine abstract again

--- 2022-02-21 main.Rmd
@@
 abstract: >
-
-  \textcolor{red}{start with describing the data: spatial and temporal, then what's the problem, then propose cubble.}
-
-  \textcolor{red}{Exploring spatio-temporal data can sometimes be discouraging given the various different formats spatio-temporal data can present themselves in practice}.
-Moreover, the two observational units naturally exist in spatio-temporal data \textcolor{red}{can cost analysts many time and efforts} on organising variables and results and switching between a single table and multiple tables in the analysis.
-In this paper, a new data structure, \code{cubble}, is proposed for wrangling and visualising spatio-temporal data.
-The new data structure uses a single data object to organise spatial and temporal variables in two forms so that they can be wrangled separately while keep syncronised.
-Advanced considerations are given to illustrate data with hierarchical structure, data matching, how cubble fits with the interactive graphics pipeline, and spatio-temporal transformation.
-Examples are given to analysing Australia meteorology data, river level data, and climate reanalysis (ERA5).
+  Spatio-temporal data have two types of variables: spatial variables have fixed value across time for the same location while temporal variables need both the location and timestamp to identify its value.
+These variables can be either organised in two separate tables, each focusing on one aspect of the data, or a single table that combines the two types of variables together.
+Both formats are useful in exploratory data analysis and visualisation, however it can be tedious to constantly switch between the two formats during the analysis and spend time casting the data into the desireable shape.
+In this paper, a new data structure, \code{cubble}, is proposed to ease the transformation between the two formats in spatio-temporal data analysis.
+The new data structure organises spatial and temporal variables as two forms of a single data object so that they can be wrangled separately while syncronised as a whole piece.
+More advanced topics discussed include handling data with hierarchical structure, matching data from multiple sources and using cubble in interactive graphics and spatio-temporal transformation.
+Examples are given using Australia climate weather station data, river level data, and climate reanalysis (ERA5) data to demonstrate the usage of the package.
 keywords:

### 2022-02-24 main.tex
refine abstract again

@@
 Spatio-temporal data have two types of variables: spatial variables have
-fixed value across time in the same location while temporal variables
+fixed value across time for the same location while temporal variables
 need both the location and timestamp to identify its value. These
-variables can be organised in two separate tables, each focusing on one
-aspect of the data, or a single table that combines all the variables
-together. For exploratory data analysis and visualisation, analysts
-needs both formats for different types of exploration, while it can be
-tedious to constantly split and join variables in order to cast the data
-into the desireable shape. In this paper, a new data structure,
-\code{cubble}, is proposed for exploring spatio-temporal data. The new
-data structure organises spatial and temporal variables as two forms of
-a single data object so that they can be wrangled separately while keep
-syncronised. Considerations are given to data with hierarchical
-structure, data matching and explain how \pkg{cubble} fits in with the
-interactive graphics pipeline and spatio-temporal transformation.
-Examples are given to analysing Australia meteorology data, river level
-data, and climate reanalysis (ERA5) data.
+variables can be either organised in two separate tables, each focusing
+on one aspect of the data, or a single table that combines the two types
+of variables together. Both formats are useful in exploratory data
+analysis and visualisation, however it can be tedious to constantly
+switch between the two formats during the analysis and spend time
+casting the data into the desireable shape. In this paper, a new data
+structure, \code{cubble}, is proposed to ease the transformation between
+the two formats in spatio-temporal data analysis. The new data structure
+organises spatial and temporal variables as two forms of a single data
+object so that they can be wrangled separately while syncronised as a
+whole piece. More advanced topics discussed include handling data with
+hierarchical structure, matching data from multiple sources and using
+cubble in interactive graphics and spatio-temporal transformation.
+Examples are given using Australia climate weather station data, river
+level data, and climate reanalysis (ERA5) data to demonstrate the usage
+of the package.
 }

### 2022-02-27 1-intro.Rmd
chunk in changes

@@
 While conceptually the data format is relatively simple, the data objects coming to the analysts can vary in forms in the analysis.
-Figure \ref{fig:illu-input} shows four examples of wild spatio-temporal data: 1) a single table that includes both the spatial and temporal variables; 2) separate tables for spatial and temporal variables; 3) array-based data, i.e.
+Figure \ref{fig:illu-input} shows three examples of wild spatio-temporal data: 1) a single table that includes both the spatial and temporal variables; 2) separate tables for spatial and temporal variables; 3) array-based data, i.e.
 NetCDF (Network Common Data Form) data, which is commonly used in earth science, e.g.
@@
 
-\textcolor{red}{start with "we need something to do exploratory data analysis for spatio-temporal data", then why it is hard to work with. This section needs to be abstract a bit further.}
+While the data is enriched the information in both the spatial and temporal dimension, it also brings variables in two different observational units, which poses the problem of organising data for various operations in exploratory data analysis, visualisation, and modelling.
+The single table format is convenient for spatio-temporal operations, however, it makes duplicates of spatial variables and is not convenient for spatial operations.
+The two tables format allows separate spatial and temporal operations on each table while analysts need to work with two data objects simultaneously and dedicate efforts to make sure the sites in both tables are matched up.
+This data organising issue can also raise during the analysis, for example, a spatial summary of the temporal or spatio-temporal data naturally introduces an additional spatial table and makes the data in the separate table format.
+With the current tools, at each step, analysts need to decide on the format to operate on, perform the operation, and organise the result with existing data.
 
-None of these three formats in Figure \ref{fig:illu-input} is not particularly convenient to work with.
+<!-- This would require analysts to constantly switch among different format at each step.
+All these difficulties are rooted in the two different observational units of spatial and temporal variables and this poses difficulties to organise spatio-temporal data in a way that operates both on a single spatial or temporal dimension and the combined spatio-temporal dimension.
+-->
+
+<!-- None of these three formats in Figure \ref{fig:illu-input} is not particularly convenient to work with.
 In the single table format, to avoid plotting duplicated points of the same sites on the map, analysts always need to create a spatial variables table from the original single table.
@@
 With the array data structure, although it can easily store and wrangle large spatio-temporal data, the 3D or higher-D array can be less intuitive than a 2D tabular data structure for some analysts in data wrangling.
+-->
 
-In the \proglang{R} community, different data objects have been proposed to work with spatio-temporal data.
+In the \proglang{R} community, different data objects have been proposed to work with spatio-temporal data when the specific task is known, while few have the capability to freely explore the data for exploratory data analysis.
 \pkg{spacetime} [@spacetime] is built on class \pkg{sp} [@sp] and \pkg{xts} [@xts] and both of which have been replaced by more recent spatial and temporal data objects.
-\pkg{spatstat} [@spatstat] implements a `ppp` class for analysing point pattern data; and more recent, \pkg{stars} [@stars] has implemented a spatio-temporal array for wrangling both vector and raster data.
-While these data objects are useful for its designed tasks, there has not yet been a data object that makes it easy to explore spatio-temporal data.
-Data objects like \pkg{sf} [@sf] and \pkg{tsibble} [@tsibble] are good examples of this in the spatial and temporal data analysis and hence there is a gap in software development for spatio-temporal data.
+A more implementation, \pkg{stars} [@stars], uses a spatio-temporal array for wrangling both vector and raster data, but for raw data already in tabular format, need to convert the data into the array structure, also convert at vis.
+While there has not yet been a data object that makes it easy to explore spatio-temporal data that overcomes the difficulties of wrangling spatio-temporal data, this motivates developing a new tool that hopefully can solve this problem.
 The requirement for such a tool is important given the ubiquity of spatio-temporal vector data in the wild: climate observations from weather stations, air quality variables from air monitoring stations, or any data collected at fixed location across time.
+The requirement can also comes from spatial data analysts wanting to incorporate a temporal dimension in their data or vice versa.
 
-<!-- While these implementations either store spatial and temporal variables all in a single table, hence duplicate the spatial variables at each temporal unit; or split them into two separate tables that analysts need to manually join the data constantly when doing the analysis.
-\textcolor{red}{There has been data structure proposed for spatial (\pkg{sf} by @sf) and temporal (\pkg{tsibble} by @tsibble) data in the R community, while less has been done for spatio-temporal data.}  -->
-
-This paper describes a new tabular data object:  \pkg{cubble} for wrangling spatio-temporal data.
-\pkg{Cubble} overcomes the inconvenience in data wrangling mentioned above through wiring the separate spatial and temporal table into a single data object.
-With the new data object, analysts can spend less time organising spatial and temporal variables at different observational units and focus more on exploring the data itself.
+This paper proposes an \proglang{R} package, \pkg{cubble}, for exploratory data analysis in spatio-temporal data.
+The proposed new data structure, also called \code{cubble}, overcomes the inconvenience in data wrangling mentioned above through wiring the separate spatial and temporal table into a single data object.
+With the new data object, analysts can spend less time organising spatial and temporal variables and focus more on the exploratory data analysis itself.
 The software is available from the Comprehensive R Archive Network (CRAN) at [CRAN link].

### 2022-02-27 main.Rmd
chunk in changes

@@
 abstract: >
-  Spatio-temporal data have two types of variables: spatial variables have fixed value across time for the same location while temporal variables need both the location and timestamp to identify its value.
-These variables can be either organised in two separate tables, each focusing on one aspect of the data, or a single table that combines the two types of variables together.
-Both formats are useful in exploratory data analysis and visualisation, however it can be tedious to constantly switch between the two formats during the analysis and spend time casting the data into the desireable shape.
-In this paper, a new data structure, \code{cubble}, is proposed to ease the transformation between the two formats in spatio-temporal data analysis.
+  Spatio-temporal data refers to measurements taken across space over time.
+In practise, spatial and temporal data many times need to be combined in order to form a spatio-temporal data set.
+However, these two types of data are collected at different observational units and this poses a problem on which format to use for different data wrangling tasks.
+Analysts can either organised the data in a single table format that combines all the variables together, or in two separate tables, each focusing on the spatial or temporal aspect of the data.
+Both formats are useful in exploratory data analysis, visualisation, and, modelling, however it can be labourious to constantly switch between the two formats during the analysis and write code to cast the data into the desireable shape.
+This paper presents a new \proglang{R} package, \pkg{cubble}, that provides a new data sturcture to ease the transformation between the two formats in spatio-temporal data.
 The new data structure organises spatial and temporal variables as two forms of a single data object so that they can be wrangled separately while syncronised as a whole piece.
-More advanced topics discussed include handling data with hierarchical structure, matching data from multiple sources and using cubble in interactive graphics and spatio-temporal transformation.
-Examples are given using Australia climate weather station data, river level data, and climate reanalysis (ERA5) data to demonstrate the usage of the package.
+\pkg{Cubble} also provides capability for handling data with hierarchical structure, matching data from multiple sources, constructing interactive graphics and performing spatio-temporal transformation.
+The paper will also demonstrate \pkg{cubble} with examples using Australia climate weather stations, river level data, and climate reanalysis (ERA5) data.
 keywords:

### 2022-02-27 main.tex
chunk in changes

@@
 \Abstract{
-Spatio-temporal data have two types of variables: spatial variables have
-fixed value across time for the same location while temporal variables
-need both the location and timestamp to identify its value. These
-variables can be either organised in two separate tables, each focusing
-on one aspect of the data, or a single table that combines the two types
-of variables together. Both formats are useful in exploratory data
-analysis and visualisation, however it can be tedious to constantly
-switch between the two formats during the analysis and spend time
-casting the data into the desireable shape. In this paper, a new data
-structure, \code{cubble}, is proposed to ease the transformation between
-the two formats in spatio-temporal data analysis. The new data structure
-organises spatial and temporal variables as two forms of a single data
-object so that they can be wrangled separately while syncronised as a
-whole piece. More advanced topics discussed include handling data with
-hierarchical structure, matching data from multiple sources and using
-cubble in interactive graphics and spatio-temporal transformation.
-Examples are given using Australia climate weather station data, river
-level data, and climate reanalysis (ERA5) data to demonstrate the usage
-of the package.
+Spatio-temporal data refers to measurements taken across space over
+time. In practise, spatial and temporal data many times need to be
+combined in order to form a spatio-temporal data set. However, these two
+types of data are collected at different observational units and this
+poses a problem on which format to use for different data wrangling
+tasks. Analysts can either organised the data in a single table format
+that combines all the variables together, or in two separate tables,
+each focusing on the spatial or temporal aspect of the data. Both
+formats are useful in exploratory data analysis, visualisation, and,
+modelling, however it can be labourious to constantly switch between the
+two formats during the analysis and write code to cast the data into the
+desireable shape. This paper presents a new \proglang{R} package,
+\pkg{cubble}, that provides a new data sturcture to ease the
+transformation between the two formats in spatio-temporal data. The new
+data structure organises spatial and temporal variables as two forms of
+a single data object so that they can be wrangled separately while
+syncronised as a whole piece. \pkg{Cubble} also provides capability for
+handling data with hierarchical structure, matching data from multiple
+sources, constructing interactive graphics and performing
+spatio-temporal transformation. The paper will also demonstrate
+\pkg{cubble} with examples using Australia climate weather stations,
+river level data, and climate reanalysis (ERA5) data.
 }
@@
 simple, the data objects coming to the analysts can vary in forms in the
-analysis. Figure \ref{fig:illu-input} shows four examples of wild
+analysis. Figure \ref{fig:illu-input} shows three examples of wild
 spatio-temporal data: 1) a single table that includes both the spatial
@@
 
-\textcolor{red}{start with "we need something to do exploratory data analysis for spatio-temporal data", then why it is hard to work with. This section needs to be abstract a bit further.}
-
-None of these three formats in Figure \ref{fig:illu-input} is not
-particularly convenient to work with. In the single table format, to
-avoid plotting duplicated points of the same sites on the map, analysts
-always need to create a spatial variables table from the original single
-table. This can be a tedious step if needs to performed many times
-during the analysis. In another type of operation that filter the sites
-based on certain statistics, intermediate objects, like the per-station
-summary of the statistic and the filtered station ID vector, will need
-be created as standalone objects in the environment. The spirit of using
-pipe reduces the intermediate data objects that are unnecessarily
-created in the analysis while in this case, these only-used-once data
-objects can hardly be incorporated into a piping workflow in the
-analysis. In addition, if the data is organised as a separate spatial
-and temporal table as format 2 in Figure \ref{fig:illu-input}, the
-filter on site ID will need to performed twice in each table to keep
-them both up-to-date. With the array data structure, although it can
-easily store and wrangle large spatio-temporal data, the 3D or higher-D
-array can be less intuitive than a 2D tabular data structure for some
-analysts in data wrangling.
+While the data is enriched the information in both the spatial and
+temporal dimension, it also brings variables in two different
+observational units, which poses the problem of organising data for
+various operations in exploratory data analysis, visualisation, and
+modelling. The single table format is convenient for spatio-temporal
+operations, however, it makes duplicates of spatial variables and is not
+convenient for spatial operations. The two tables format allows separate
+spatial and temporal operations on each table while analysts need to
+work with two data objects simultaneously and dedicate efforts to make
+sure the sites in both tables are matched up. This data organising issue
+can also raise during the analysis, for example, a spatial summary of
+the temporal or spatio-temporal data naturally introduces an additional
+spatial table and makes the data in the separate table format. With the
+current tools, at each step, analysts need to decide on the format to
+operate on, perform the operation, and organise the result with existing
+data.
 
 In the \proglang{R} community, different data objects have been proposed
-to work with spatio-temporal data. \pkg{spacetime} \citep{spacetime} is
-built on class \pkg{sp} \citep{sp} and \pkg{xts} \citep{xts} and both of
-which have been replaced by more recent spatial and temporal data
-objects. \pkg{spatstat} \citep{spatstat} implements a \texttt{ppp} class
-for analysing point pattern data; and more recent, \pkg{stars}
-\citep{stars} has implemented a spatio-temporal array for wrangling both
-vector and raster data. While these data objects are useful for its
-designed tasks, there has not yet been a data object that makes it easy
-to explore spatio-temporal data. Data objects like \pkg{sf} \citep{sf}
-and \pkg{tsibble} \citep{tsibble} are good examples of this in the
-spatial and temporal data analysis and hence there is a gap in software
-development for spatio-temporal data. The requirement for such a tool is
-important given the ubiquity of spatio-temporal vector data in the wild:
-climate observations from weather stations, air quality variables from
-air monitoring stations, or any data collected at fixed location across
-time.
-
-This paper describes a new tabular data object: \pkg{cubble} for
-wrangling spatio-temporal data. \pkg{Cubble} overcomes the inconvenience
-in data wrangling mentioned above through wiring the separate spatial
-and temporal table into a single data object. With the new data object,
+to work with spatio-temporal data when the specific task is known, while
+few have the capability to freely explore the data for exploratory data
+analysis. \pkg{spacetime} \citep{spacetime} is built on class \pkg{sp}
+\citep{sp} and \pkg{xts} \citep{xts} and both of which have been
+replaced by more recent spatial and temporal data objects. A more
+implementation, \pkg{stars} \citep{stars}, uses a spatio-temporal array
+for wrangling both vector and raster data, but for raw data already in
+tabular format, need to convert the data into the array structure, also
+convert at vis. While there has not yet been a data object that makes it
+easy to explore spatio-temporal data that overcomes the difficulties of
+wrangling spatio-temporal data, this motivates developing a new tool
+that hopefully can solve this problem. The requirement for such a tool
+is important given the ubiquity of spatio-temporal vector data in the
+wild: climate observations from weather stations, air quality variables
+from air monitoring stations, or any data collected at fixed location
+across time. The requirement can also comes from spatial data analysts
+wanting to incorporate a temporal dimension in their data or vice versa.
+
+This paper proposes an \proglang{R} package, \pkg{cubble}, for
+exploratory data analysis in spatio-temporal data. The proposed new data
+structure, also called \code{cubble}, overcomes the inconvenience in
+data wrangling mentioned above through wiring the separate spatial and
+temporal table into a single data object. With the new data object,
 analysts can spend less time organising spatial and temporal variables
-at different observational units and focus more on exploring the data
-itself. The software is available from the Comprehensive R Archive
-Network (CRAN) at {[}CRAN link{]}.
+and focus more on the exploratory data analysis itself. The software is
+available from the Comprehensive R Archive Network (CRAN) at {[}CRAN
+link{]}.
 

### 2022-02-28 1-intro.Rmd
proofread abstract adn intro

@@
 
-Spatio-temporal data involves variables observed in the space across time and this paper focuses on spatio-temporal vector data with fixed locations and regular time interval.
-This type of data can naturally been considered as a data cube with three axes: spatial identifier, temporal identifier, and variable.
+Spatio-temporal data involves variables observed in the space across time and this paper focuses on spatio-temporal vector data with fixed locations and regular time intervals.
+This type of data can naturally be considered as a data cube with three axes: spatial identifier, temporal identifier, and variable.
 In this cubic formation, space only occupies one axis, rather than two, to avoid hypercubes for multivariate spatio-temporal data.
@@
 
-While the data is enriched the information in both the spatial and temporal dimension, it also brings variables in two different observational units, which poses the problem of organising data for various operations in exploratory data analysis, visualisation, and modelling.
-The single table format is convenient for spatio-temporal operations, however, it makes duplicates of spatial variables and is not convenient for spatial operations.
-The two tables format allows separate spatial and temporal operations on each table while analysts need to work with two data objects simultaneously and dedicate efforts to make sure the sites in both tables are matched up.
-This data organising issue can also raise during the analysis, for example, a spatial summary of the temporal or spatio-temporal data naturally introduces an additional spatial table and makes the data in the separate table format.
-With the current tools, at each step, analysts need to decide on the format to operate on, perform the operation, and organise the result with existing data.
+While incorporating both spatial and temporal information enriches the data in the analysis, spatio-temporal data can sometimes be hard to work with given its two observational units in the variables.
+The single table format in Figure \ref{fig:illu-input} is convenient for spatio-temporal operations that involve both types of variables, however, it makes duplicates of spatial variables and is not ideal when only looking at the spatial aspect of the data.
+The two tables format allows separate operations on spatial and temporal variables while analysts need to work with two data objects simultaneously and join them together into the single table format for spatio-temporal operations.
+Neither of the two formats is desirable at the same time for spatial, temporal, and spatio-temporal operations.
+This issue implies analysts need to constantly think about the data format and arrange variables into the proper shape before the actual operations, which can be a laborious step when switching between the two formats back and forth several times during the analysis.
 
@@
 
-In the \proglang{R} community, different data objects have been proposed to work with spatio-temporal data when the specific task is known, while few have the capability to freely explore the data for exploratory data analysis.
-\pkg{spacetime} [@spacetime] is built on class \pkg{sp} [@sp] and \pkg{xts} [@xts] and both of which have been replaced by more recent spatial and temporal data objects.
-A more implementation, \pkg{stars} [@stars], uses a spatio-temporal array for wrangling both vector and raster data, but for raw data already in tabular format, need to convert the data into the array structure, also convert at vis.
-While there has not yet been a data object that makes it easy to explore spatio-temporal data that overcomes the difficulties of wrangling spatio-temporal data, this motivates developing a new tool that hopefully can solve this problem.
-The requirement for such a tool is important given the ubiquity of spatio-temporal vector data in the wild: climate observations from weather stations, air quality variables from air monitoring stations, or any data collected at fixed location across time.
-The requirement can also comes from spatial data analysts wanting to incorporate a temporal dimension in their data or vice versa.
+In the \proglang{R} community, many data objects have been proposed to work with specific spatio-temporal tasks, while few can freely explore the data.
+\pkg{spacetime} [@spacetime] is an attempt for such a data object, but the spatial and temporal class it built on have now been gradually replaced by more recent implementations.
+The \pkg{stars} [@stars] package uses an array to wrangle spatio-temporal data, but the dense array it uses can quickly exhaust the memory for data in the irregular spatial grid.
+While there has not yet been a data object that makes it easy to wrangle spatio-temporal data, there is a strong need for such tools given the ubiquitous spatial and temporal data collected in the society, i.e., climate observations from weather stations, air quality variables from air monitoring stations.
+The demand for such tools can also come from spatial data analysts wishing to incorporate temporal data in their analysis or vice versa.
 
-This paper proposes an \proglang{R} package, \pkg{cubble}, for exploratory data analysis in spatio-temporal data.
-The proposed new data structure, also called \code{cubble}, overcomes the inconvenience in data wrangling mentioned above through wiring the separate spatial and temporal table into a single data object.
-With the new data object, analysts can spend less time organising spatial and temporal variables and focus more on the exploratory data analysis itself.
+This paper presents a new \proglang{R} package, \pkg{cubble}, to overcome the problem with organising variables in different levels when working with spatio-temporal data.
+In the package a new data structure, also called \code{cubble}, is proposed to organise spatial and temporal variables as two forms of a single data object so that they can be wrangled separately while synchronised as a whole piece.
+With the new data object, analysts can spend less time writing codes to organise spatial and temporal variables and focus more on the exploratory data analysis itself.
 The software is available from the Comprehensive R Archive Network (CRAN) at [CRAN link].

### 2022-02-28 main.Rmd
proofread abstract adn intro

@@
   Spatio-temporal data refers to measurements taken across space over time.
-In practise, spatial and temporal data many times need to be combined in order to form a spatio-temporal data set.
-However, these two types of data are collected at different observational units and this poses a problem on which format to use for different data wrangling tasks.
-Analysts can either organised the data in a single table format that combines all the variables together, or in two separate tables, each focusing on the spatial or temporal aspect of the data.
-Both formats are useful in exploratory data analysis, visualisation, and, modelling, however it can be labourious to constantly switch between the two formats during the analysis and write code to cast the data into the desireable shape.
-This paper presents a new \proglang{R} package, \pkg{cubble}, that provides a new data sturcture to ease the transformation between the two formats in spatio-temporal data.
-The new data structure organises spatial and temporal variables as two forms of a single data object so that they can be wrangled separately while syncronised as a whole piece.
-\pkg{Cubble} also provides capability for handling data with hierarchical structure, matching data from multiple sources, constructing interactive graphics and performing spatio-temporal transformation.
-The paper will also demonstrate \pkg{cubble} with examples using Australia climate weather stations, river level data, and climate reanalysis (ERA5) data.
+In practice, spatial and temporal data many times need to be combined in order to form a spatio-temporal data set.
+However, these two types of data are collected at different observational units and this poses a problem on how to structure the data for different wrangling tasks.
+Analysts can either organise the data in a single table format that combines all the variables, or in two separate tables, each focusing on the spatial or temporal aspect of the data.
+Both formats are useful in exploratory data analysis, visualisation, and, modelling, however, it can be labourious to constantly switch between the two formats during the analysis and write code to cast the data into the desirable shape.
+This paper presents a new \proglang{R} package, \pkg{cubble}, that provides a new data structure to ease the conversion between the two formats in spatio-temporal data.
+The new data structure organises spatial and temporal variables as two forms of a single data object so that they can be wrangled separately while synchronised as a whole piece.
+\pkg{Cubble} also provides the capability for handling data with hierarchical structure, matching data from multiple sources, constructing interactive graphics, and performing spatio-temporal transformation.
+The paper will demonstrate \pkg{cubble} with examples using Australian climate weather stations, river level data, and climate reanalysis (ERA5) data.
 keywords:

### 2022-02-28 main.tex
proofread abstract adn intro

@@
 Spatio-temporal data refers to measurements taken across space over
-time. In practise, spatial and temporal data many times need to be
+time. In practice, spatial and temporal data many times need to be
 combined in order to form a spatio-temporal data set. However, these two
 types of data are collected at different observational units and this
-poses a problem on which format to use for different data wrangling
-tasks. Analysts can either organised the data in a single table format
-that combines all the variables together, or in two separate tables,
-each focusing on the spatial or temporal aspect of the data. Both
-formats are useful in exploratory data analysis, visualisation, and,
-modelling, however it can be labourious to constantly switch between the
-two formats during the analysis and write code to cast the data into the
-desireable shape. This paper presents a new \proglang{R} package,
-\pkg{cubble}, that provides a new data sturcture to ease the
-transformation between the two formats in spatio-temporal data. The new
-data structure organises spatial and temporal variables as two forms of
-a single data object so that they can be wrangled separately while
-syncronised as a whole piece. \pkg{Cubble} also provides capability for
-handling data with hierarchical structure, matching data from multiple
-sources, constructing interactive graphics and performing
-spatio-temporal transformation. The paper will also demonstrate
-\pkg{cubble} with examples using Australia climate weather stations,
-river level data, and climate reanalysis (ERA5) data.
+poses a problem on how to structure the data for different wrangling
+tasks. Analysts can either organise the data in a single table format
+that combines all the variables, or in two separate tables, each
+focusing on the spatial or temporal aspect of the data. Both formats are
+useful in exploratory data analysis, visualisation, and, modelling,
+however, it can be labourious to constantly switch between the two
+formats during the analysis and write code to cast the data into the
+desirable shape. This paper presents a new \proglang{R} package,
+\pkg{cubble}, that provides a new data structure to ease the conversion
+between the two formats in spatio-temporal data. The new data structure
+organises spatial and temporal variables as two forms of a single data
+object so that they can be wrangled separately while synchronised as a
+whole piece. \pkg{Cubble} also provides the capability for handling data
+with hierarchical structure, matching data from multiple sources,
+constructing interactive graphics, and performing spatio-temporal
+transformation. The paper will demonstrate \pkg{cubble} with examples
+using Australian climate weather stations, river level data, and climate
+reanalysis (ERA5) data.
 }
@@
 time and this paper focuses on spatio-temporal vector data with fixed
-locations and regular time interval. This type of data can naturally
-been considered as a data cube with three axes: spatial identifier,
-temporal identifier, and variable. In this cubic formation, space only
-occupies one axis, rather than two, to avoid hypercubes for multivariate
+locations and regular time intervals. This type of data can naturally be
+considered as a data cube with three axes: spatial identifier, temporal
+identifier, and variable. In this cubic formation, space only occupies
+one axis, rather than two, to avoid hypercubes for multivariate
 spatio-temporal data. While conceptually the data format is relatively
@@
 
-While the data is enriched the information in both the spatial and
-temporal dimension, it also brings variables in two different
-observational units, which poses the problem of organising data for
-various operations in exploratory data analysis, visualisation, and
-modelling. The single table format is convenient for spatio-temporal
-operations, however, it makes duplicates of spatial variables and is not
-convenient for spatial operations. The two tables format allows separate
-spatial and temporal operations on each table while analysts need to
-work with two data objects simultaneously and dedicate efforts to make
-sure the sites in both tables are matched up. This data organising issue
-can also raise during the analysis, for example, a spatial summary of
-the temporal or spatio-temporal data naturally introduces an additional
-spatial table and makes the data in the separate table format. With the
-current tools, at each step, analysts need to decide on the format to
-operate on, perform the operation, and organise the result with existing
-data.
-
-In the \proglang{R} community, different data objects have been proposed
-to work with spatio-temporal data when the specific task is known, while
-few have the capability to freely explore the data for exploratory data
-analysis. \pkg{spacetime} \citep{spacetime} is built on class \pkg{sp}
-\citep{sp} and \pkg{xts} \citep{xts} and both of which have been
-replaced by more recent spatial and temporal data objects. A more
-implementation, \pkg{stars} \citep{stars}, uses a spatio-temporal array
-for wrangling both vector and raster data, but for raw data already in
-tabular format, need to convert the data into the array structure, also
-convert at vis. While there has not yet been a data object that makes it
-easy to explore spatio-temporal data that overcomes the difficulties of
-wrangling spatio-temporal data, this motivates developing a new tool
-that hopefully can solve this problem. The requirement for such a tool
-is important given the ubiquity of spatio-temporal vector data in the
-wild: climate observations from weather stations, air quality variables
-from air monitoring stations, or any data collected at fixed location
-across time. The requirement can also comes from spatial data analysts
-wanting to incorporate a temporal dimension in their data or vice versa.
-
-This paper proposes an \proglang{R} package, \pkg{cubble}, for
-exploratory data analysis in spatio-temporal data. The proposed new data
-structure, also called \code{cubble}, overcomes the inconvenience in
-data wrangling mentioned above through wiring the separate spatial and
-temporal table into a single data object. With the new data object,
-analysts can spend less time organising spatial and temporal variables
-and focus more on the exploratory data analysis itself. The software is
-available from the Comprehensive R Archive Network (CRAN) at {[}CRAN
-link{]}.
+While incorporating both spatial and temporal information enriches the
+data in the analysis, spatio-temporal data can sometimes be hard to work
+with given its two observational units in the variables. The single
+table format in Figure \ref{fig:illu-input} is convenient for
+spatio-temporal operations that involve both types of variables,
+however, it makes duplicates of spatial variables and is not ideal when
+only looking at the spatial aspect of the data. The two tables format
+allows separate operations on spatial and temporal variables while
+analysts need to work with two data objects simultaneously and join them
+together into the single table format for spatio-temporal operations.
+Neither of the two formats is desirable at the same time for spatial,
+temporal, and spatio-temporal operations. This issue implies analysts
+need to constantly think about the data format and arrange variables
+into the proper shape before the actual operations, which can be a
+laborious step when switching between the two formats back and forth
+several times during the analysis.
+
+In the \proglang{R} community, many data objects have been proposed to
+work with specific spatio-temporal tasks, while few can freely explore
+the data. \pkg{spacetime} \citep{spacetime} is an attempt for such a
+data object, but the spatial and temporal class it built on have now
+been gradually replaced by more recent implementations. The \pkg{stars}
+\citep{stars} package uses an array to wrangle spatio-temporal data, but
+the dense array it uses can quickly exhaust the memory for data in the
+irregular spatial grid. While there has not yet been a data object that
+makes it easy to wrangle spatio-temporal data, there is a strong need
+for such tools given the ubiquitous spatial and temporal data collected
+in the society, i.e., climate observations from weather stations, air
+quality variables from air monitoring stations. The demand for such
+tools can also come from spatial data analysts wishing to incorporate
+temporal data in their analysis or vice versa.
+
+This paper presents a new \proglang{R} package, \pkg{cubble}, to
+overcome the problem with organising variables in different levels when
+working with spatio-temporal data. In the package a new data structure,
+also called \code{cubble}, is proposed to organise spatial and temporal
+variables as two forms of a single data object so that they can be
+wrangled separately while synchronised as a whole piece. With the new
+data object, analysts can spend less time writing codes to organise
+spatial and temporal variables and focus more on the exploratory data
+analysis itself. The software is available from the Comprehensive R
+Archive Network (CRAN) at {[}CRAN link{]}.
 

### 2022-03-01 main.Rmd
Update main.Rmd
minor polishing

@@
     address: 2000 Jintong Road, Tangjiawan, Zhuhai, Guangdong Province, China
-    email: nicolas@langrene.com
+    email: nicolaslangrene@uic.edu.cn
   - name: Patricia Menéndez
@@
   Spatio-temporal data refers to measurements taken across space over time.
-In practice, spatial and temporal data many times need to be combined in order to form a spatio-temporal data set.
-However, these two types of data are collected at different observational units and this poses a problem on how to structure the data for different wrangling tasks.
+In practice, spatial and temporal data often need to be combined in order to form a spatio-temporal data set.
+However, different types of data might be collected at different observational units, which poses a problem on how to structure the data for different wrangling tasks.
 Analysts can either organise the data in a single table format that combines all the variables, or in two separate tables, each focusing on the spatial or temporal aspect of the data.
-Both formats are useful in exploratory data analysis, visualisation, and, modelling, however, it can be labourious to constantly switch between the two formats during the analysis and write code to cast the data into the desirable shape.
+Both formats are useful in exploratory data analysis, visualisation, and, modelling; however, it can be laborious to constantly switch between the two formats during the analysis and write code to cast the data into the desirable shape.
 This paper presents a new \proglang{R} package, \pkg{cubble}, that provides a new data structure to ease the conversion between the two formats in spatio-temporal data.
@@
 \pkg{Cubble} also provides the capability for handling data with hierarchical structure, matching data from multiple sources, constructing interactive graphics, and performing spatio-temporal transformation.
-The paper will demonstrate \pkg{cubble} with examples using Australian climate weather stations, river level data, and climate reanalysis (ERA5) data.
+The paper demonstrates these capabilities \pkg{cubble} with examples using Australian climate weather stations, river level data, and climate reanalysis (ERA5) data.
 keywords:

### 2022-03-01 1-intro.Rmd
Update 1-intro.Rmd
minor polishing

@@
 Spatio-temporal data involves variables observed in the space across time and this paper focuses on spatio-temporal vector data with fixed locations and regular time intervals.
-This type of data can naturally be considered as a data cube with three axes: spatial identifier, temporal identifier, and variable.
+This type of data can naturally be considered as a data cube with three axes: spatial identifier, temporal identifier, and variables.
 In this cubic formation, space only occupies one axis, rather than two, to avoid hypercubes for multivariate spatio-temporal data.
@@
 
-```{r illu-input, echo = FALSE, out.width = "100%", fig.cap="An illustration of different spatio-temporal data layout: (1) a single table for both spatial and temporal variables; (2) separate tables for spatial and temporal variables; and (3) an array data structure that populates variables into a gridded longitude-latitude-time (hyper)cube.", fig.align='center'}
+```{r illu-input, echo = FALSE, out.width = "100%", fig.cap="An illustration of different spatio-temporal data layouts: (1) a single table for both spatial and temporal variables; (2) separate tables for spatial and temporal variables; and (3) an array data structure that populates variables into a gridded longitude-latitude-time (hyper)cube.", fig.align='center'}
 knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.001.png"))
@@
 In the \proglang{R} community, many data objects have been proposed to work with specific spatio-temporal tasks, while few can freely explore the data.
-\pkg{spacetime} [@spacetime] is an attempt for such a data object, but the spatial and temporal class it built on have now been gradually replaced by more recent implementations.
-The \pkg{stars} [@stars] package uses an array to wrangle spatio-temporal data, but the dense array it uses can quickly exhaust the memory for data in the irregular spatial grid.
+\pkg{spacetime} [@spacetime] is an attempt for such a data object, but the spatial and temporal classes it is built on have now been gradually replaced by more recent implementations.
+The \pkg{stars} [@stars] package uses an array to wrangle spatio-temporal data, but the dense array it uses can quickly exhaust the memory for data in the case of irregular spatial grid.
 While there has not yet been a data object that makes it easy to wrangle spatio-temporal data, there is a strong need for such tools given the ubiquitous spatial and temporal data collected in the society, i.e., climate observations from weather stations, air quality variables from air monitoring stations.
@@
 In the package a new data structure, also called \code{cubble}, is proposed to organise spatial and temporal variables as two forms of a single data object so that they can be wrangled separately while synchronised as a whole piece.
-With the new data object, analysts can spend less time writing codes to organise spatial and temporal variables and focus more on the exploratory data analysis itself.
+With this new data object, analysts can spend less time writing codes to organise spatial and temporal variables and focus more on the exploratory data analysis itself.
 The software is available from the Comprehensive R Archive Network (CRAN) at [CRAN link].
@@
 Section 3 explains how cubble deals with more advanced considerations including data with hierarchical structure, data matching, how cubble fits with existing static and interactive visualisation tools, and spatio-temporal data transformation.
-Section 4 uses Australia weather station data and river level data as examples to demonstrate the use with \pkg{cubble}.
+Section 4 uses Australia weather station data and river level data as examples to demonstrate the use of \pkg{cubble}.
 An example on how \pkg{cubble} handles NetCDF data is also provided.

### 2022-03-01 2-the-cubble-pkg.Rmd
Update 2-the-cubble-pkg.Rmd
minor polishing

@@
 This section will present two formats that cubble uses to arrange spatio-temporal data.
-Main cubble functions will then be introduced to illustrate how to work with these two formats in short examples.
-Lastly, a subsection will dedicate to how existing packages, in spatial and temporal analysis, fit in with cubble.
+The main cubble functions will then be introduced to illustrate how to work with these two formats in short examples.
+Lastly, a subsection will be dedicated to how existing packages, in spatial and temporal analysis, fit in with cubble.
 
-In cubble, a single data object can be arranged into two forms: nested form and long form, and Figure \ref{fig:illu-cubble} sketches the two forms with the associated attributes.
+In cubble, a single data object can be arranged into two forms: nested form and long form.
+Figure \ref{fig:illu-cubble} sketches the two forms with the associated attributes.
 The decision on which form to use in data wrangling is output-oriented, meaning analysts need to first think about whether the output of an operation is identified only by the spatial identifier, or a combination of spatial and temporal identifier.
@@
 The nested form defines each site in a row and nests the time varying variables into a single column \\code{ts}.
-The long form cubble uses \\code{id} and \\code{t} to identify each row and store the time invariant variables as an attribute, \\code{spatial}."}
+The long form cubble uses \\code{id} and \\code{t} to identify each row and stores the time invariant variables as an attribute, \\code{spatial}."}
 knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.002.png"))
@@
 
-The nested format is convenient for those operations whose output doesn't contain a time dimension, for those outputs that are cross-identified by the spatial and temporal identifier, a long cubble needs to be used.
+The nested format is convenient for those operations whose output does not contain a time dimension.
+For those outputs that are cross-identified by the spatial and temporal identifier, a long cubble needs to be used.
 The function \code{stretch()} is designed to switch the cubble from the nested form to the long form.
@@
 Sometimes, analysts may need to apply some variable transformation that involves both the spatial and temporal variable.
-An example of this is the transformation of temporal variables into the spatial dimension in glyph maps(which will be elaborated in section \ref{st_transformation}).
+An example of this is the transformation of temporal variables into the spatial dimension in glyph maps (which will be elaborated in section \ref{st_transformation}).
 Cubble allows this operation through \code{migrate()}, which moves the specified spatial variables into the long cubble:
@@
 
-\code{tsibble} is a temporal data structure that uses \code{index} and \code{key} to identify the time and different series and \code{cubble} can be seen as a natural extension of \code{tsibble} for spatio-temporal data with an additional coordinates component and using two forms to arrange variables.
+\code{tsibble} is a temporal data structure that uses \code{index} and \code{key} to identify the time and different series.
+\code{cubble} can be seen as a natural extension of \code{tsibble} for spatio-temporal data with an additional coordinates component and using two forms to arrange variables.
 This makes it easy to cast a \code{tsibble} into a \code{cubble} as only the \code{coords} argument needs to be supplied:
@@
 ```{r echo = TRUE}
-# demonstrate with a tsibble created from climate_flat
+# example with a tsibble created from climate_flat
 raw <- climate_flat %>%
@@
 As a spatial data object, \code{sf} creates a simple feature geometry list-column (\code{sfc}) in the data frame to provide spatial operations on various geometry types (\code{POINT}, \code{LINESTRING}, \code{POLYGON}, \code{MULTIPOLYGON}, etc).
-These spatial operations are also valuable for spatio-temporal data analysis, but an \code{sf} object *usually* don't contain temporal variables.
-This means \code{sf} can't be directly cast into a \code{cubble}, however, \code{cubble} does support \code{sfc} columns in the nested form and spatial operations applied to the \code{sfc} column in \code{sf} can still be applied to the \code{sfc} column in a cubble.
+These spatial operations are also valuable for spatio-temporal data analysis, but an \code{sf} object *usually* does not contain temporal variables.
+This means \code{sf} cannot be directly cast into a \code{cubble}, however, \code{cubble} does support \code{sfc} columns in the nested form and spatial operations applied to the \code{sfc} column in \code{sf} can still be applied to the \code{sfc} column in a cubble.
 The following example shows how to create an \code{sfc} column of \code{POINT} type from latitude and longitude in cubble.
@@
 library(s2)
-# western australia map
+# Western Australia map
 wa <- ozmaps::abs_ste %>% filter(NAME == "Western Australia")
@@
 
-Memory limit with NetCDF data in cubble depends on longitude grid point x latitude grid point x time grid point x number of variable.
-Cubble can handle slightly more than 300 x 300 (longitude x longitude) grid points for 3 variables in one year and spatial grid can be reduced to trade for longer time period and more variables.
+The memory limit with NetCDF data in cubble depends on the number of longitude grid points times the number of latitude grid points times the number of time grid points times the number of variables.
+Cubble can handle slightly more than 300 x 300 (longitude x longitude) grid points for 3 variables in one year.
+The spatial grid can be reduced to trade for longer time periods and more variables.
 A 300 by 300 spatial grid can be a bounding box of [100, -80, 180, 0] at 0.25 degree resolution or a global bounding box [-180, -90, 180, -90] at 1 degree resolution.
-Subsetting longitude and latitude grid is available through \code{long_range} and \code{lat_range} if the NetCDF file has finer resolution than needed.
+Subsetting longitude and latitude grids is available through \code{long_range} and \code{lat_range} if the NetCDF file has finer resolution than needed.
 

### 2022-03-01 main.Rmd
major changes of abstract after talking to Di

--- 2022-02-28 main.Rmd
@@
 abstract: >
-  Spatio-temporal data refers to measurements taken across space over time.
-In practice, spatial and temporal data many times need to be combined in order to form a spatio-temporal data set.
-However, these two types of data are collected at different observational units and this poses a problem on how to structure the data for different wrangling tasks.
-Analysts can either organise the data in a single table format that combines all the variables, or in two separate tables, each focusing on the spatial or temporal aspect of the data.
-Both formats are useful in exploratory data analysis, visualisation, and, modelling, however, it can be labourious to constantly switch between the two formats during the analysis and write code to cast the data into the desirable shape.
-This paper presents a new \proglang{R} package, \pkg{cubble}, that provides a new data structure to ease the conversion between the two formats in spatio-temporal data.
-The new data structure organises spatial and temporal variables as two forms of a single data object so that they can be wrangled separately while synchronised as a whole piece.
-\pkg{Cubble} also provides the capability for handling data with hierarchical structure, matching data from multiple sources, constructing interactive graphics, and performing spatio-temporal transformation.
+  Spatio-temporal data refer to measurements taken across space and time.
+In practice, spatio-temporal data can be decomposed into a spatial and temporal component: at one time, we would select a spatial location and inspect the temporal trend; at other time, we might select one or multiple time value(s) and explore the spatial distribution.
+Ideally, we could make multiple maps and multiple time series to explore these together, however, doing all of these actions is complicated when data arrive fragmented in multiple objects.
+To make it easy to do all these tasks, ideally spatial and temporal variables are in a single data object that we can slice and dice in different ways to conduct different visualisations.
+This work suggests a new data structure, \code{cubble}, that uses a nested form and a long form to organise spatial and temporal variables in a single data object.
+The new data structure ensures that data in the two forms are synchronised while been flexible to explore the two aspect of spatio-temporal data.
+It also provides capability for handling data with hierarchical structure, matching data from multiple sources, constructing interactive graphics, and performing spatio-temporal transformation.
 The paper will demonstrate \pkg{cubble} with examples using Australian climate weather stations, river level data, and climate reanalysis (ERA5) data.

### 2022-03-01 main.tex
major changes of abstract after talking to Di

@@
 \Abstract{
-Spatio-temporal data refers to measurements taken across space over
-time. In practice, spatial and temporal data many times need to be
-combined in order to form a spatio-temporal data set. However, these two
-types of data are collected at different observational units and this
-poses a problem on how to structure the data for different wrangling
-tasks. Analysts can either organise the data in a single table format
-that combines all the variables, or in two separate tables, each
-focusing on the spatial or temporal aspect of the data. Both formats are
-useful in exploratory data analysis, visualisation, and, modelling,
-however, it can be labourious to constantly switch between the two
-formats during the analysis and write code to cast the data into the
-desirable shape. This paper presents a new \proglang{R} package,
-\pkg{cubble}, that provides a new data structure to ease the conversion
-between the two formats in spatio-temporal data. The new data structure
-organises spatial and temporal variables as two forms of a single data
-object so that they can be wrangled separately while synchronised as a
-whole piece. \pkg{Cubble} also provides the capability for handling data
-with hierarchical structure, matching data from multiple sources,
-constructing interactive graphics, and performing spatio-temporal
-transformation. The paper will demonstrate \pkg{cubble} with examples
-using Australian climate weather stations, river level data, and climate
-reanalysis (ERA5) data.
+Spatio-temporal data refer to measurements taken across space and time.
+In practice, spatio-temporal data can be decomposed into a spatial and
+temporal component: at one time, we would select a spatial location and
+inspect the temporal trend; at other time, we might select one or
+multiple time value(s) and explore the spatial distribution. Ideally, we
+could make multiple maps and multiple time series to explore these
+together, however, doing all of these actions is complicated when data
+arrive fragmented in multiple objects. To make it easy to do all these
+tasks, ideally spatial and temporal variables are in a single data
+object that we can slice and dice in different ways to conduct different
+visualisations. This work suggests a new data structure, \code{cubble},
+that uses a nested form and a long form to organise spatial and temporal
+variables in a single data object. The new data structure ensures that
+data in the two forms are synchronised while been flexible to explore
+the two aspect of spatio-temporal data. It also provides capability for
+handling data with hierarchical structure, matching data from multiple
+sources, constructing interactive graphics, and performing
+spatio-temporal transformation. The paper will demonstrate \pkg{cubble}
+with examples using Australian climate weather stations, river level
+data, and climate reanalysis (ERA5) data.
 }
@@
     2000 Jintong Road, Tangjiawan, Zhuhai, Guangdong Province, China\\
-  E-mail: \href{mailto:nicolas@langrene.com}{\nolinkurl{nicolas@langrene.com}}\\
+  E-mail: \href{mailto:nicolaslangrene@uic.edu.cn}{\nolinkurl{nicolaslangrene@uic.edu.cn}}\\
 
@@
 considered as a data cube with three axes: spatial identifier, temporal
-identifier, and variable. In this cubic formation, space only occupies
+identifier, and variables. In this cubic formation, space only occupies
 one axis, rather than two, to avoid hypercubes for multivariate
@@
 
-\caption[An illustration of different spatio-temporal data layout]{An illustration of different spatio-temporal data layout: (1) a single table for both spatial and temporal variables; (2) separate tables for spatial and temporal variables; and (3) an array data structure that populates variables into a gridded longitude-latitude-time (hyper)cube.}\label{fig:illu-input}
+\caption[An illustration of different spatio-temporal data layouts]{An illustration of different spatio-temporal data layouts: (1) a single table for both spatial and temporal variables; (2) separate tables for spatial and temporal variables; and (3) an array data structure that populates variables into a gridded longitude-latitude-time (hyper)cube.}\label{fig:illu-input}
 \end{figure}
@@
 the data. \pkg{spacetime} \citep{spacetime} is an attempt for such a
-data object, but the spatial and temporal class it built on have now
-been gradually replaced by more recent implementations. The \pkg{stars}
-\citep{stars} package uses an array to wrangle spatio-temporal data, but
-the dense array it uses can quickly exhaust the memory for data in the
-irregular spatial grid. While there has not yet been a data object that
-makes it easy to wrangle spatio-temporal data, there is a strong need
-for such tools given the ubiquitous spatial and temporal data collected
-in the society, i.e., climate observations from weather stations, air
-quality variables from air monitoring stations. The demand for such
-tools can also come from spatial data analysts wishing to incorporate
-temporal data in their analysis or vice versa.
+data object, but the spatial and temporal classes it is built on have
+now been gradually replaced by more recent implementations. The
+\pkg{stars} \citep{stars} package uses an array to wrangle
+spatio-temporal data, but the dense array it uses can quickly exhaust
+the memory for data in the case of irregular spatial grid. While there
+has not yet been a data object that makes it easy to wrangle
+spatio-temporal data, there is a strong need for such tools given the
+ubiquitous spatial and temporal data collected in the society, i.e.,
+climate observations from weather stations, air quality variables from
+air monitoring stations. The demand for such tools can also come from
+spatial data analysts wishing to incorporate temporal data in their
+analysis or vice versa.
 
@@
 variables as two forms of a single data object so that they can be
-wrangled separately while synchronised as a whole piece. With the new
+wrangled separately while synchronised as a whole piece. With this new
 data object, analysts can spend less time writing codes to organise
@@
 transformation. Section 4 uses Australia weather station data and river
-level data as examples to demonstrate the use with \pkg{cubble}. An
+level data as examples to demonstrate the use of \pkg{cubble}. An
 example on how \pkg{cubble} handles NetCDF data is also provided.

### 2022-03-26 2-0-cube.Rmd
compare array with cubble, different formats spatial and temporal data are stored

---
title: "2-0-cube"
output: pdf_document
---

# Conceptual framework

Multivariate spatio-temporal data can be conceptualised using a cubical data model with reference to variables, time and space. This naturally motivates the usage of array to represent and store spatio-temporal data, as evident by satellite imageries, large climate models, among others. The work by @lu_multidimensional_2018 has discussed the suitability of this representation through array algebra and practical spatio-temporal tasks and software implementation has been seen in many R packages: \pkg{raster} [@raster], \pkg{terra} [@terra], and \pkg{stars} [@stars].

While it is efficient to represent spatio-temporal data in an array, there are situation where array [is not the best solution]. mutate in the array is not clearly define (how to add a column vector)

Also, array is not the best solution for a large subset of data, specifically, the vector spatio-temporal data. With vector data, spatial and temporal variables are commonly collected in separate 2D tables and there is usually a location ID to link the spatial and temporal information. Sometimes, the geometrical shape of the location may not present and an additional linking is needed to incorporate the map data. With the multiple pieces in 2D data frame format, there should be a solution to organise all these data from multiple sources in a 2D structure, without needing to convert it into higher dimensional representation.

Reliable software and data structure has been proposed in the spatial and temporal domain, while a direct extension to spatio-temporal data is not ideal. The main issue here lays in the format that spatial and temporal data is stored. For spatial data, \pkg{sf} [@sf] covers a wide range of geometrical operations and is the format most map data stored. When incorporating the time dimension, a time wide format, introduced in @spacetime, is usually adopted. This treats time as variable name rather than value and makes it inconvenient to wrangle time, as per the tidy data principle [@tidydata]. In the temporal domain, \pkg{tsibble} [@tsibble] follows the tidy data principle to arrange the time in the long format, while when adding spatial variables, a long format implies repeating variables at each time and this can make a huge difference when the repetition is on large geometrical shapes with long time period.

### 2022-03-26 main.tex
compare array with cubble, different formats spatial and temporal data are stored

@@
 
+\hypertarget{conceptual-framework}{%
+\section{Conceptual framework}\label{conceptual-framework}}
+
+Multivariate spatio-temporal data can be conceptualised using a cubical
+data model with reference to variables, time and space. This naturally
+motivates the usage of array to represent and store spatio-temporal
+data, as evident by satellite imageries, large climate models, among
+others. The work by \citet{lu_multidimensional_2018} has discussed the
+suitability of this representation through array algebra and practical
+spatio-temporal tasks and software implementation has been seen in many
+R packages: \pkg{raster} \citep{raster}, \pkg{terra} \citep{terra}, and
+\pkg{stars} \citep{stars}.
+
+While it is efficient to represent spatio-temporal data in an array,
+there are situation where array {[}is not the best solution{]}.
+Sometimes, one may want to create a new
+
+Also, array is not the best solution for a large subset of data,
+specifically, the vector spatio-temporal data. With vector data, spatial
+and temporal variables are commonly collected in separate 2D tables and
+there is usually a location ID to link the spatial and temporal
+information. Sometimes, the geometrical shape of the location may not
+present and an additional linking is needed to incorporate the map data.
+With the multiple pieces in 2D data frame format, there should be a
+solution to organise all these data from multiple sources in a 2D
+structure, without needing to convert it into higher dimensional
+representation.
+
+Reliable software and data structure has been proposed in the spatial
+and temporal domain, while a direct extension to spatio-temporal data is
+not ideal. The main issue here lays in the format that spatial and
+temporal data is stored. For spatial data, \pkg{sf} \citep{sf} covers a
+wide range of geometrical operations and is the format most map data
+stored. When incorporating the time dimension, a time wide format,
+introduced in \citet{spacetime}, is usually adopted. This treats time as
+variable name rather than value and makes it inconvenient to wrangle
+time, as per the tidy data principle \citep{tidydata}. In the temporal
+domain, \pkg{tsibble} \citep{tsibble} follows the tidy data principle to
+arrange the time in the long format, while when adding spatial
+variables, a long format implies repeating variables at each time and
+this can make a huge difference when the repetition is on large
+geometrical shapes with long time period.
+
+However, with all the benefit of conceptualising the data into a cube,
+the cubical structure is not particular convenient for, using the
+tidyverse word, mutating new variables in the data object. This is
+because an array is a higher dimensional object than the 2D data frame
+and adding a new column doesn't clearly define how a vector should be
+added to an (multidimensional) array.
+
+The cubble package is a collection of tools to work with vector
+spatio-temporal data from multiple sources.
+
 \hypertarget{the-cubble-package}{%
@@
 This section will present two formats that cubble uses to arrange
-spatio-temporal data. Main cubble functions will then be introduced to
-illustrate how to work with these two formats in short examples. Lastly,
-a subsection will dedicate to how existing packages, in spatial and
-temporal analysis, fit in with cubble.
+spatio-temporal data. The main cubble functions will then be introduced
+to illustrate how to work with these two formats in short examples.
+Lastly, a subsection will be dedicated to how existing packages, in
+spatial and temporal analysis, fit in with cubble.
 
 In cubble, a single data object can be arranged into two forms: nested
-form and long form, and Figure \ref{fig:illu-cubble} sketches the two
-forms with the associated attributes. The decision on which form to use
-in data wrangling is output-oriented, meaning analysts need to first
-think about whether the output of an operation is identified only by the
+form and long form. Figure \ref{fig:illu-cubble} sketches the two forms
+with the associated attributes. The decision on which form to use in
+data wrangling is output-oriented, meaning analysts need to first think
+about whether the output of an operation is identified only by the
 spatial identifier, or a combination of spatial and temporal identifier.
@@
 The nested form defines each site in a row and nests the time varying variables into a single column \code{ts}.
-The long form cubble uses \code{id} and \code{t} to identify each row and store the time invariant variables as an attribute, \code{spatial}.}\label{fig:illu-cubble}
+The long form cubble uses \code{id} and \code{t} to identify each row and stores the time invariant variables as an attribute, \code{spatial}.}\label{fig:illu-cubble}
 \end{figure}
@@
 
-The nested format is convenient for those operations whose output
-doesn't contain a time dimension, for those outputs that are
+The nested format is convenient for those operations whose output does
+not contain a time dimension. For those outputs that are
 cross-identified by the spatial and temporal identifier, a long cubble
@@
 the transformation of temporal variables into the spatial dimension in
-glyph maps(which will be elaborated in section \ref{st_transformation}).
-Cubble allows this operation through \code{migrate()}, which moves the
-specified spatial variables into the long cubble:
+glyph maps (which will be elaborated in section
+\ref{st_transformation}). Cubble allows this operation through
+\code{migrate()}, which moves the specified spatial variables into the
+long cubble:
 
@@
 # bbox:    [115.97, -32.94, 133.55, -12.42]
-# spatial: long [dbl], lat [dbl], elev [dbl], name [chr], wmo_id [dbl]
+# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]
   id          date        prcp  tmax  tmin
@@
 \code{tsibble} is a temporal data structure that uses \code{index} and
-\code{key} to identify the time and different series and \code{cubble}
-can be seen as a natural extension of \code{tsibble} for spatio-temporal
+\code{key} to identify the time and different series. \code{cubble} can
+be seen as a natural extension of \code{tsibble} for spatio-temporal
 data with an additional coordinates component and using two forms to
@@
 \begin{CodeInput}
-R> # demonstrate with a tsibble created from climate_flat
+R> # example with a tsibble created from climate_flat
 R> raw <- climate_flat %>%
@@
   <chr>       <dbl> <dbl> <dbl> <chr>          <dbl> <list>       <dbl>    <dbl>
-1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tbl_t~      25.7    38.6 
-2 ASN00010311 -31.9  117. 179   york            94623 <tbl_t~      26.2    51.1 
-3 ASN00010614 -32.9  117. 338   narrogin        94627 <tbl_t~      23.7    45.4 
-4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tbl_t~      33.1     3.02
-5 ASN00015131 -17.6  134. 220   elliott         94236 <tbl_t~      34.6    24.7 
+1 ASN00009021 -31.9  116.  15.4 perth airport  94610 <tbl_ts>      25.7    38.6 
+2 ASN00010311 -31.9  117. 179   york           94623 <tbl_ts>      26.2    51.1 
+3 ASN00010614 -32.9  117. 338   narrogin       94627 <tbl_ts>      23.7    45.4 
+4 ASN00014015 -12.4  131.  30.4 darwin airpo~  94120 <tbl_ts>      33.1     3.02
+5 ASN00015131 -17.6  134. 220   elliott        94236 <tbl_ts>      34.6    24.7 
 \end{CodeOutput}
@@
 also valuable for spatio-temporal data analysis, but an \code{sf} object
-\emph{usually} don't contain temporal variables. This means \code{sf}
-can't be directly cast into a \code{cubble}, however, \code{cubble} does
-support \code{sfc} columns in the nested form and spatial operations
-applied to the \code{sfc} column in \code{sf} can still be applied to
-the \code{sfc} column in a cubble. The following example shows how to
-create an \code{sfc} column of \code{POINT} type from latitude and
-longitude in cubble. Then \code{sf::st_within} is used to add the state
-\code{MULTIPOLYGON} of each weather station before a coordinate
-transformation is made.
+\emph{usually} does not contain temporal variables. This means \code{sf}
+cannot be directly cast into a \code{cubble}, however, \code{cubble}
+does support \code{sfc} columns in the nested form and spatial
+operations applied to the \code{sfc} column in \code{sf} can still be
+applied to the \code{sfc} column in a cubble. The following example
+shows how to create an \code{sfc} column of \code{POINT} type from
+latitude and longitude in cubble. Then \code{sf::st_within} is used to
+add the state \code{MULTIPOLYGON} of each weather station before a
+coordinate transformation is made.
 
@@
   <chr>       <dbl> <dbl> <dbl> <chr>   <dbl> <list>           <POINT [°]>
-1 ASN00009021 -31.9  116.  15.4 perth airpo~  94610 <tibble~ (115.9764 -31.9275)
-2 ASN00010311 -31.9  117. 179   york          94623 <tibble~  (116.765 -31.8997)
-3 ASN00010614 -32.9  117. 338   narrogin      94627 <tibble~ (117.1797 -32.9342)
-4 ASN00014015 -12.4  131.  30.4 darwin airp~  94120 <tibble~ (130.8925 -12.4239)
-5 ASN00015131 -17.6  134. 220   elliott       94236 <tibble~ (133.5407 -17.5521)
+1 ASN00009021 -31.9  116.  15.4 perth~  94610 <tibble> (115.9764 -31.9275)
+2 ASN00010311 -31.9  117. 179   york    94623 <tibble>  (116.765 -31.8997)
+3 ASN00010614 -32.9  117. 338   narro~  94627 <tibble> (117.1797 -32.9342)
+4 ASN00014015 -12.4  131.  30.4 darwi~  94120 <tibble> (130.8925 -12.4239)
+5 ASN00015131 -17.6  134. 220   ellio~  94236 <tibble> (133.5407 -17.5521)
 # ... with 1 more variable: state <MULTIPOLYGON [m]>
@@
 R> library(s2)
-R> # western australia map
+R> # Western Australia map
 R> wa <- ozmaps::abs_ste %>% filter(NAME == "Western Australia")
@@
 
-Memory limit with NetCDF data in cubble depends on longitude grid point
-x latitude grid point x time grid point x number of variable. Cubble can
+The memory limit with NetCDF data in cubble depends on the number of
+longitude grid points times the number of latitude grid points times the
+number of time grid points times the number of variables. Cubble can
 handle slightly more than 300 x 300 (longitude x longitude) grid points
-for 3 variables in one year and spatial grid can be reduced to trade for
-longer time period and more variables. A 300 by 300 spatial grid can be
-a bounding box of {[}100, -80, 180, 0{]} at 0.25 degree resolution or a
-global bounding box {[}-180, -90, 180, -90{]} at 1 degree resolution.
-Subsetting longitude and latitude grid is available through
-\code{long_range} and \code{lat_range} if the NetCDF file has finer
-resolution than needed.
+for 3 variables in one year. The spatial grid can be reduced to trade
+for longer time periods and more variables. A 300 by 300 spatial grid
+can be a bounding box of {[}100, -80, 180, 0{]} at 0.25 degree
+resolution or a global bounding box {[}-180, -90, 180, -90{]} at 1
+degree resolution. Subsetting longitude and latitude grids is available
+through \code{long_range} and \code{lat_range} if the NetCDF file has
+finer resolution than needed.
 
@@
   <chr>       <chr>               <dbl> <dbl> <chr> <dbl> <int> <list>     <int>
-1 405234      SEVEN CREEKS @ D~ -36.9  146. river   6.15     5 <tibble ~      21
-2 404207      HOLLAND CREEK @ ~ -36.6  146. river   8.54    10 <tibble ~      21
-3 ASN00082042 strathbogie       -36.8  146. clima~  6.15     5 <tibble ~      21
-4 ASN00082170 benalla airport   -36.6  146. clima~  8.54    10 <tibble ~      21
-5 230200      MARIBYRNONG RIVE~ -37.7  145. river   6.17     6 <tibble ~      19
-6 ASN00086038 essendon airport  -37.7  145. clima~  6.17     6 <tibble ~      19
-7 406213      CAMPASPE RIVER @~ -37.0  145. river   1.84     1 <tibble ~      18
-8 ASN00088051 redesdale         -37.0  145. clima~  1.84     1 <tibble ~      18
+1 405234      SEVEN CREEKS @ D/S~ -36.9  146. river  6.15     5 <tibble>      21
+2 404207      HOLLAND CREEK @ KE~ -36.6  146. river  8.54    10 <tibble>      21
+3 ASN00082042 strathbogie         -36.8  146. clim~  6.15     5 <tibble>      21
+4 ASN00082170 benalla airport     -36.6  146. clim~  8.54    10 <tibble>      21
+5 230200      MARIBYRNONG RIVER ~ -37.7  145. river  6.17     6 <tibble>      19
+6 ASN00086038 essendon airport    -37.7  145. clim~  6.17     6 <tibble>      19
+7 406213      CAMPASPE RIVER @ R~ -37.0  145. river  1.84     1 <tibble>      18
+8 ASN00088051 redesdale           -37.0  145. clim~  1.84     1 <tibble>      18
 \end{CodeOutput}

### 2022-03-31 2-0-cube.Rmd
draft on conceptual cube and new challenges

@@
 
-# Conceptual framework
+# Conceptual framework: spatio-temporal cube
 
@@
 This naturally motivates the usage of array to represent and store spatio-temporal data, as evident by satellite imageries, large climate models, among others.
-The work by @lu_multidimensional_2018 has discussed the suitability of this representation through array algebra and practical spatio-temporal tasks and software implementation has been seen in many R packages: \pkg{raster} [@raster], \pkg{terra} [@terra], and \pkg{stars} [@stars].
+The work by @lu_multidimensional_2018 has discussed the suitability of this representation through array algebra and practical spatio-temporal tasks.
 
-While it is efficient to represent spatio-temporal data in an array, there are situation where array [is not the best solution]. mutate in the array is not clearly define (how to add a column vector)
+The cubical conceptual framework provides a generalisation to the operation can be done on the data for visualisation.
+A paper by @bach_review_2014 has reviewed the temporal data visualisation based on space-time cube operations.
+Notice that the term space-time cube in their article "does not need to involve spatial data", but referring to "an abstract 2D substrate that is used to visualize data at a specific time".
+Despite its main focus is on temporal data, the mindset of abstracting out data representation to construct visualisation is applicable to our spatio-temporal data manipulation and visualisation.
 
-Also, array is not the best solution for a large subset of data, specifically, the vector spatio-temporal data.
+<!-- and software implementation has been seen in many R packages: \pkg{raster} [@raster], \pkg{terra} [@terra], and \pkg{stars} [@stars]. -->
+
+# Existing work and new challenges
+
+**what has been done and the new challenges**
+
+The \pkg{spacetime} paper [@spacetime] has set the foundation for wrangling spatio-temporal data in \proglang{R}.
+It proposes four layouts (STF: full grid layout, STS: sparse grid layout, STI: irregular layout, and STT: trajectory) to represent spatio-temporal data.
+The implementation uses class \pkg{sp} to represent spatial data and class \pkg{xts} for temporal data.
+In recent years, new data structures have appeared to replace and improve on these two classes to represent spatial and temporal data.
+The package \pkg{sf} has adopted the simple feature access [@jr_herring_opengis_2011] to represent spatial information using POINT, MULTIPOINT, LINESTRING, POLYGON, MULTILINESTRING, MULTIPOLYGON, and GEOMETRYCOLLECTION.
+For temporal data, the package \pkg{tsibble} adopts the long form to explicitly record the time stamp as a column rather than an attribute of the data.
+These changes in the spatial and temporal community have yet been reflected in the spatio-temporal data analysis.
+The @spacetime paper also summarises three formats spatio-temporal data are usually recorded: time-wide, space-wide, and long format.
+With the concept of tidy data [@tidydata], the long form has been increasingly considered as the tidy format.
+The main advantages of it over the wide format is that the latter treat either time or location value as variable names (Section 3.1 in @tidydata), making them difficult to be wrangled.
+While temporal data can be benefited from a long format, it brings repetition of spatial variables in the combined spatio-temporal data and this can be a huge inefficiency for large MULTIPOLOYGON over long time period.
+This suggests the need to look for other format for spatio-temporal data.
+
+<!-- For spatial data, \pkg{sf} [@sf] covers a wide range of geometrical operations and is the format most map data stored.
+When incorporating the time dimension, a time wide format, introduced in @spacetime, is usually adopted.
+This treats time as variable name rather than value and makes it inconvenient to wrangle time, as per the tidy data principle [@tidydata].
+In the temporal domain, \pkg{tsibble} [@tsibble] follows the tidy data principle to arrange the time in the long format, while when adding spatial variables, a long format implies repeating variables at each time and this can make a huge difference when the repetition is on large geometrical shapes with long time period.
+-->
+
+**two challenages: multiple data sources + large spatio-temporal object**
+In additional to this, there are more challenges in manipulation spatio-temporal data.
+Different variables are recorded and collected in multiple tables, for example, an areal data would map data for a collection of areas, geo-scientific data would record locations and other meta data, and the temporal variablem indexed by both location and time are stored in the long form table.
+To work with these data, analysts need to link the geo-scientific data with the temporal variables and work out how they are contained in the area.
+This would require merging all three sources together and can often go wrong because of the slightest unmatched of linking variable in different sources.
+Another challenge analysts face when wrangling spatio-temporal data is with the simple feature access representation of spatial data.
+While they have become the standard for recording large and complex spatial geometry and precise operation on the spatial coordinates, with data analysis, the detail is more than enough.
+A data analysts don't necessarily need to have the every nitty gritty detail of the location and a map with sufficient information to outline the shape of the location would be enough for the propose of visualisation.
+With a large and detailed geometrical object, the map would take longer to render and analysts always need to first simplify it into a more manageable size.
+Also, while the simple feature access structure the coordinates into a list column, analysts would sometimes like to unpack it into two columns of longitude and latitude to work with.
+They should be provided with the option to do so.
+
+**A newer implementation**
+The \pkg{stars} has been an implementation to represent spatio-temporal data with array.
+While it is efficient to represent spatio-temporal data in an array, there are situation where array [is not the best solution].
+mutate in the array is not clearly define (how to add a column vector).
+Also, it doesn't solve the problem with multiple data sources as it may take a large amount of effort to shape multivariate spatio-temporal data into high-dimensional cube before the analysis starts and they should be provided with option to work with 2D tables.
+
+<!-- Also, array is not the best solution for a large subset of data, specifically, the vector spatio-temporal data.
 With vector data, spatial and temporal variables are commonly collected in separate 2D tables and there is usually a location ID to link the spatial and temporal information.
@@
 With the multiple pieces in 2D data frame format, there should be a solution to organise all these data from multiple sources in a 2D structure, without needing to convert it into higher dimensional representation.
-
-Reliable software and data structure has been proposed in the spatial and temporal domain, while a direct extension to spatio-temporal data is not ideal.
-The main issue here lays in the format that spatial and temporal data is stored.
-For spatial data, \pkg{sf} [@sf] covers a wide range of geometrical operations and is the format most map data stored.
-When incorporating the time dimension, a time wide format, introduced in @spacetime, is usually adopted.
-This treats time as variable name rather than value and makes it inconvenient to wrangle time, as per the tidy data principle [@tidydata].
-In the temporal domain, \pkg{tsibble} [@tsibble] follows the tidy data principle to arrange the time in the long format, while when adding spatial variables, a long format implies repeating variables at each time and this can make a huge difference when the repetition is on large geometrical shapes with long time period.
+-->

### 2022-03-31 main.tex
draft on conceptual cube and new challenges

@@
 
-\hypertarget{conceptual-framework}{%
-\section{Conceptual framework}\label{conceptual-framework}}
+\hypertarget{conceptual-framework-spatio-temporal-cube}{%
+\section{Conceptual framework: spatio-temporal
+cube}\label{conceptual-framework-spatio-temporal-cube}}
 
@@
 suitability of this representation through array algebra and practical
-spatio-temporal tasks and software implementation has been seen in many
-R packages: \pkg{raster} \citep{raster}, \pkg{terra} \citep{terra}, and
-\pkg{stars} \citep{stars}.
-
-While it is efficient to represent spatio-temporal data in an array,
-there are situation where array {[}is not the best solution{]}.
-Sometimes, one may want to create a new
-
-Also, array is not the best solution for a large subset of data,
-specifically, the vector spatio-temporal data. With vector data, spatial
-and temporal variables are commonly collected in separate 2D tables and
-there is usually a location ID to link the spatial and temporal
-information. Sometimes, the geometrical shape of the location may not
-present and an additional linking is needed to incorporate the map data.
-With the multiple pieces in 2D data frame format, there should be a
-solution to organise all these data from multiple sources in a 2D
-structure, without needing to convert it into higher dimensional
-representation.
-
-Reliable software and data structure has been proposed in the spatial
-and temporal domain, while a direct extension to spatio-temporal data is
-not ideal. The main issue here lays in the format that spatial and
-temporal data is stored. For spatial data, \pkg{sf} \citep{sf} covers a
-wide range of geometrical operations and is the format most map data
-stored. When incorporating the time dimension, a time wide format,
-introduced in \citet{spacetime}, is usually adopted. This treats time as
-variable name rather than value and makes it inconvenient to wrangle
-time, as per the tidy data principle \citep{tidydata}. In the temporal
-domain, \pkg{tsibble} \citep{tsibble} follows the tidy data principle to
-arrange the time in the long format, while when adding spatial
-variables, a long format implies repeating variables at each time and
-this can make a huge difference when the repetition is on large
-geometrical shapes with long time period.
-
-However, with all the benefit of conceptualising the data into a cube,
-the cubical structure is not particular convenient for, using the
-tidyverse word, mutating new variables in the data object. This is
-because an array is a higher dimensional object than the 2D data frame
-and adding a new column doesn't clearly define how a vector should be
-added to an (multidimensional) array.
-
-The cubble package is a collection of tools to work with vector
-spatio-temporal data from multiple sources.
+spatio-temporal tasks.
+
+The cubical conceptual framework provides a generalisation to the
+operation can be done on the data for visualisation. A paper by
+\citet{bach_review_2014} has reviewed the temporal data visualisation
+based on space-time cube operations. Notice that the term space-time
+cube in their article ``does not need to involve spatial data'', but
+referring to ``an abstract 2D substrate that is used to visualize data
+at a specific time''. Despite its main focus is on temporal data, the
+mindset of abstracting out data representation to construct
+visualisation is applicable to our spatio-temporal data manipulation and
+visualisation.
+
+\hypertarget{existing-work-and-new-challenges}{%
+\section{Existing work and new
+challenges}\label{existing-work-and-new-challenges}}
+
+\textbf{what has been done and the new challenges}
+
+The \pkg{spacetime} paper \citep{spacetime} has set the foundation for
+wrangling spatio-temporal data in \proglang{R}. It proposes four layouts
+(STF: full grid layout, STS: sparse grid layout, STI: irregular layout,
+and STT: trajectory) to represent spatio-temporal data. The
+implementation uses class \pkg{sp} to represent spatial data and class
+\pkg{xts} for temporal data. In recent years, new data structures have
+appeared to replace and improve on these two classes to represent
+spatial and temporal data. The package \pkg{sf} has adopted the simple
+feature access \citep{jr_herring_opengis_2011} to represent spatial
+information using POINT, MULTIPOINT, LINESTRING, POLYGON,
+MULTILINESTRING, MULTIPOLYGON, and GEOMETRYCOLLECTION. For temporal
+data, the package \pkg{tsibble} adopts the long form to explicitly
+record the time stamp as a column rather than an attribute of the data.
+These changes in the spatial and temporal community have yet been
+reflected in the spatio-temporal data analysis. The \citet{spacetime}
+paper also summarises three formats spatio-temporal data are usually
+recorded: time-wide, space-wide, and long format. With the concept of
+tidy data \citep{tidydata}, the long form has been increasingly
+considered as the tidy format. The main advantages of it over the wide
+format is that the latter treat either time or location value as
+variable names (Section 3.1 in \citet{tidydata}), making them difficult
+to be wrangled. While temporal data can be benefited from a long format,
+it brings repetition of spatial variables in the combined
+spatio-temporal data and this can be a huge inefficiency for large
+MULTIPOLOYGON over long time period. This suggests the need to look for
+other format for spatio-temporal data.
+
+\textbf{two challenages: multiple data sources + large spatio-temporal
+object} In additional to this, there are more challenges in manipulation
+spatio-temporal data. Different variables are recorded and collected in
+multiple tables, for example, an areal data would map data for a
+collection of areas, geo-scientific data would record locations and
+other meta data, and the temporal variablem indexed by both location and
+time are stored in the long form table. To work with these data,
+analysts need to link the geo-scientific data with the temporal
+variables and work out how they are contained in the area. This would
+require merging all three sources together and can often go wrong
+because of the slightest unmatched of linking variable in different
+sources. Another challenge analysts face when wrangling spatio-temporal
+data is with the simple feature access representation of spatial data.
+While they have become the standard for recording large and complex
+spatial geometry and precise operation on the spatial coordinates, with
+data analysis, the detail is more than enough. A data analysts don't
+necessarily need to have the every nitty gritty detail of the location
+and a map with sufficient information to outline the shape of the
+location would be enough for the propose of visualisation. With a large
+and detailed geometrical object, the map would take longer to render and
+analysts always need to first simplify it into a more manageable size.
+Also, while the simple feature access structure the coordinates into a
+list column, analysts would sometimes like to unpack it into two columns
+of longitude and latitude to work with. They should be provided with the
+option to do so.
+
+\textbf{A newer implementation} The \pkg{stars} has been an
+implementation to represent spatio-temporal data with array. While it is
+efficient to represent spatio-temporal data in an array, there are
+situation where array {[}is not the best solution{]}. mutate in the
+array is not clearly define (how to add a column vector). Also, it
+doesn't solve the problem with multiple data sources as it may take a
+large amount of effort to shape multivariate spatio-temporal data into
+high-dimensional cube before the analysis starts and they should be
+provided with option to work with 2D tables.
 

### 2022-04-01 2-0-cube.Rmd
refine new challenges

--- 2022-03-26 2-0-cube.Rmd
@@
 This naturally motivates the usage of array to represent and store spatio-temporal data, as evident by satellite imageries, large climate models, among others.
-The work by @lu_multidimensional_2018 has discussed the suitability of this representation through array algebra and practical spatio-temporal tasks and software implementation has been seen in many R packages: \pkg{raster} [@raster], \pkg{terra} [@terra], and \pkg{stars} [@stars].
+The work by @lu_multidimensional_2018 has discussed the suitability of this representation through array algebra and practical spatio-temporal tasks.
 
-While it is efficient to represent spatio-temporal data in an array, there are situation where array [is not the best solution]. mutate in the array is not clearly define (how to add a column vector)
+The cubical conceptual framework provides a generalisation to the operation can be done on the data for visualisation.
+A paper by @bach_review_2014 has reviewed the temporal data visualisation based on space-time cube operations.
+Notice that the term space-time cube in their article "does not need to involve spatial data", but referring to "an abstract 2D substrate that is used to visualize data at a specific time".
+Despite its main focus is on temporal data, the mindset of abstracting out data representation to construct visualisation is applicable to our spatio-temporal data manipulation and visualisation.
 
-Also, array is not the best solution for a large subset of data, specifically, the vector spatio-temporal data.
+<!-- and software implementation has been seen in many R packages: \pkg{raster} [@raster], \pkg{terra} [@terra], and \pkg{stars} [@stars]. -->
+
+# Existing work and new challenges
+
+<!-- Spatio-temporal software should echo with the advance in spatial and temporal data analysis.
+This section will start with the \pkg{spacetime} paper [@spacetime] that sets the foundation for wrangling spatio-temporal data in \proglang{R}, and then describe some emerging concepts and software in data analysis since the paper publishes and what remains as challenges.
+-->
+
+To represent spatio-temporal data, @spacetime proposes four layouts: full grid layout (STF),  sparse grid layout (STS), irregular layout (STI), and trajectory (STT).
+These layouts are implemented in the \pkg{spacetime} package with underlying class \pkg{sp} [@sp] and \pkg{xts} [@xts] for handling spatial and temporal variables.
+While these layouts and classes are still being used and well-maintained today, they are not compatible with the \pkg{tidyverse} ecosystem [@tidyverse], which has a huge influence on data analysis in \proglang{R}, including in the spatial and temporal communities.
+The package \pkg{sf} [@sf] uses data frame objects to represent spatial data with an additional \code{geometry} column dedicating to feature geometries [@jr_herring_opengis_2011] (e.g.
+POINT, MULTIPOINT, POLYGON, MULTIPOLYGON etc).
+For temporal data, \pkg{tsibble} [@tsibble] builds on top of the \pkg{tibble} [@tibble] object and explicitly includes date and time in columns as variables, rather than data attributes.
+
+Underlying tidyverse is the concept of tidy data [@tidydata], which considers how variables should be structured for analysis.
+In the three formats (time-wide, space-wide, and long) summarised in the @spacetime paper, only the long form is considered as tidy.
+The main advantages of it over the time-wide and space-wide format is that the latter treat either time or location as variable names (Section 3.1 in @tidydata) rather than values of variables, making them difficult to be wrangled.
+One problem with the long format is that it can be inefficient to store feature geometries, especially for large MULTIPOLYGON over long time period.
+This motivates a new tidyverse compatible data object to work with spatio-temporal data.
+
+<!-- For spatial data, \pkg{sf} [@sf] covers a wide range of geometrical operations and is the format most map data stored.
+When incorporating the time dimension, a time wide format, introduced in @spacetime, is usually adopted.
+This treats time as variable name rather than value and makes it inconvenient to wrangle time, as per the tidy data principle [@tidydata].
+In the temporal domain, \pkg{tsibble} [@tsibble] follows the tidy data principle to arrange the time in the long format, while when adding spatial variables, a long format implies repeating variables at each time and this can make a huge difference when the repetition is on large geometrical shapes with long time period.
+-->
+
+**two challenages: multiple data sources + large spatio-temporal object** 
+In additional to the shift into more tidyverse workflow for working with data, spatio-temporal data have some unique challenges.
+The first one being the split of variables in multiple tables.
+Often the time, spatio-temporal data is not presented in a single data object, but split into fragmented pieces from different sources.
+For example, map data recorded the shape of a collection of area of interest is provided by the government website; the geo-scientific data is recorded with longitude and latitude coordinate of locations in the area; and temporal variables indexed by both location and time need to be further queried using location data.
+For spatio-temporal data analysis, merging all these sources can be challenging and joining these tables can go wrong because of the slightest unmatched of the linking variable from different sources.
+
+Another challenge when analysing spatio-temporal data is on the simple feature representation of spatial data.
+While the simple feature is able to handle large and complex spatial geometry and perform precise spatial operations, data analysts usually do not have a large expectation on the map accuracy and would prefer a simplified map with quick rendering for visualisation.
+Also, the feature geometry column in \pkg{sf} wraps the spatial coordinates in a list column, sometimes, analysts may like to unpack them into longitude and latitude to work with.
+A smooth transition with the coordinate columns and and simple feature geometry column would satisfy analysts from both ends.
+
+With some of the aforementioned issues recognised, the \pkg{stars} package has been a new implementation to handle spatio-temporal data conceptually as data cubes and implemented as arrays.
+The package depends on the \pkg{sf} package and has allowed more access to tidyverse verbs in the workflow.
+However, for vector spatio-temporal data, major difficulties to work with \pkg{stars} is first to cast the variables from multiple tables into a star object and then keep the operations in the high-dimensional cube.
+Given that most table mentioned before collect variables in a 2D table, analysts could benefit from a structure to allow them work in 2D tables.
+
+<!-- Also, array is not the best solution for a large subset of data, specifically, the vector spatio-temporal data.
 With vector data, spatial and temporal variables are commonly collected in separate 2D tables and there is usually a location ID to link the spatial and temporal information.

### 2022-04-01 main.tex
refine new challenges

--- 2022-03-26 main.tex
@@
 suitability of this representation through array algebra and practical
-spatio-temporal tasks and software implementation has been seen in many
-R packages: \pkg{raster} \citep{raster}, \pkg{terra} \citep{terra}, and
-\pkg{stars} \citep{stars}.
-
-While it is efficient to represent spatio-temporal data in an array,
-there are situation where array {[}is not the best solution{]}.
-Sometimes, one may want to create a new
-
-Also, array is not the best solution for a large subset of data,
-specifically, the vector spatio-temporal data. With vector data, spatial
-and temporal variables are commonly collected in separate 2D tables and
-there is usually a location ID to link the spatial and temporal
-information. Sometimes, the geometrical shape of the location may not
-present and an additional linking is needed to incorporate the map data.
-With the multiple pieces in 2D data frame format, there should be a
-solution to organise all these data from multiple sources in a 2D
-structure, without needing to convert it into higher dimensional
-representation.
-
-Reliable software and data structure has been proposed in the spatial
-and temporal domain, while a direct extension to spatio-temporal data is
-not ideal. The main issue here lays in the format that spatial and
-temporal data is stored. For spatial data, \pkg{sf} \citep{sf} covers a
-wide range of geometrical operations and is the format most map data
-stored. When incorporating the time dimension, a time wide format,
-introduced in \citet{spacetime}, is usually adopted. This treats time as
-variable name rather than value and makes it inconvenient to wrangle
-time, as per the tidy data principle \citep{tidydata}. In the temporal
-domain, \pkg{tsibble} \citep{tsibble} follows the tidy data principle to
-arrange the time in the long format, while when adding spatial
-variables, a long format implies repeating variables at each time and
-this can make a huge difference when the repetition is on large
-geometrical shapes with long time period.
-
-However, with all the benefit of conceptualising the data into a cube,
-the cubical structure is not particular convenient for, using the
-tidyverse word, mutating new variables in the data object. This is
-because an array is a higher dimensional object than the 2D data frame
-and adding a new column doesn't clearly define how a vector should be
-added to an (multidimensional) array.
-
-The cubble package is a collection of tools to work with vector
-spatio-temporal data from multiple sources.
+spatio-temporal tasks.
+
+The cubical conceptual framework provides a generalisation to the
+operation can be done on the data for visualisation. A paper by
+\citet{bach_review_2014} has reviewed the temporal data visualisation
+based on space-time cube operations. Notice that the term space-time
+cube in their article ``does not need to involve spatial data'', but
+referring to ``an abstract 2D substrate that is used to visualize data
+at a specific time''. Despite its main focus is on temporal data, the
+mindset of abstracting out data representation to construct
+visualisation is applicable to our spatio-temporal data manipulation and
+visualisation.
+
+\hypertarget{existing-work-and-new-challenges}{%
+\section{Existing work and new
+challenges}\label{existing-work-and-new-challenges}}
+
+To represent spatio-temporal data, \citet{spacetime} proposes four
+layouts: full grid layout (STF), sparse grid layout (STS), irregular
+layout (STI), and trajectory (STT). These layouts are implemented in the
+\pkg{spacetime} package with underlying class \pkg{sp} \citep{sp} and
+\pkg{xts} \citep{xts} for handling spatial and temporal variables. While
+these layouts and classes are still being used and well-maintained
+today, they are not compatible with the \pkg{tidyverse} ecosystem
+\citep{tidyverse}, which has a huge influence on data analysis in
+\proglang{R}, including in the spatial and temporal communities. The
+package \pkg{sf} \citep{sf} uses data frame objects to represent spatial
+data with an additional \code{geometry} column dedicating to feature
+geometries \citep{jr_herring_opengis_2011} (e.g.~POINT, MULTIPOINT,
+POLYGON, MULTIPOLYGON etc). For temporal data, \pkg{tsibble}
+\citep{tsibble} builds on top of the \pkg{tibble} \citep{tibble} object
+and explicitly includes date and time in columns as variables, rather
+than data attributes.
+
+Underlying tidyverse is the concept of tidy data \citep{tidydata}, which
+considers how variables should be structured for analysis. In the three
+formats (time-wide, space-wide, and long) summarised in the
+\citet{spacetime} paper, only the long form is considered as tidy. The
+main advantages of it over the time-wide and space-wide format is that
+the latter treat either time or location as variable names (Section 3.1
+in \citet{tidydata}) rather than values of variables, making them
+difficult to be wrangled. One problem with the long format is that it
+can be inefficient to store feature geometries, especially for large
+MULTIPOLYGON over long time period. This motivates a new tidyverse
+compatible data object to work with spatio-temporal data.
+
+\textbf{two challenages: multiple data sources + large spatio-temporal
+object} In additional to the shift into more tidyverse workflow for
+working with data, spatio-temporal data have some unique challenges. The
+first one being the split of variables in multiple tables. Often the
+time, spatio-temporal data is not presented in a single data object, but
+split into fragmented pieces from different sources. For example, map
+data recorded the shape of a collection of area of interest is provided
+by the government website; the geo-scientific data is recorded with
+longitude and latitude coordinate of locations in the area; and temporal
+variables indexed by both location and time need to be further queried
+using location data. For spatio-temporal data analysis, merging all
+these sources can be challenging and joining these tables can go wrong
+because of the slightest unmatched of the linking variable from
+different sources.
+
+Another challenge when analysing spatio-temporal data is on the simple
+feature representation of spatial data. While the simple feature is able
+to handle large and complex spatial geometry and perform precise spatial
+operations, data analysts usually do not have a large expectation on the
+map accuracy and would prefer a simplified map with quick rendering for
+visualisation. Also, the feature geometry column in \pkg{sf} wraps the
+spatial coordinates in a list column, sometimes, analysts may like to
+unpack them into longitude and latitude to work with. A smooth
+transition with the coordinate columns and and simple feature geometry
+column would satisfy analysts from both ends.
+
+With some of the aforementioned issues recognised, the \pkg{stars}
+package has been a new implementation to handle spatio-temporal data
+conceptually as data cubes and implemented as arrays. The package
+depends on the \pkg{sf} package and has allowed more access to tidyverse
+verbs in the workflow. However, for vector spatio-temporal data, major
+difficulties to work with \pkg{stars} is first to cast the variables
+from multiple tables into a star object and then keep the operations in
+the high-dimensional cube. Given that most table mentioned before
+collect variables in a 2D table, analysts could benefit from a structure
+to allow them work in 2D tables.
 

### 2022-04-17 2-the-cubble-pkg.Rmd
small changes

@@
 cubble_mig %>% tamp() %>% stretch()
+```
+
+## Why not just use \pkg{dplyr::nest_by}
+
+```{r eval = FALSE}
+library(cubble)
+library(tidyverse)
+left_join(cubble::stations, cubble::climate) %>% 
+  dplyr::nest_by(id, lat,  long,  elev, name, wmo_id) %>% 
+  tidyr::unnest() %>% 
+  group_by(........, geometry)
+```
+
+```{r eval = FALSE}
+as_cubble(list(spatial = stations, temporal = climate),
+          key = id, index = date, coords = c(long, lat))
+```
+
+```{r eval = FALSE}
+usmap::us_map("states")
 ```

### 2022-04-17 4-5-interactive.Rmd
clean up bibbib
small changes

@@
 ---
-## Interative graphic with cubble
+## Interative graphic
 
@@
 
-Figure \ref{fig:interactive-popup} shows Figure \ref{fig:interactive-linking} made made with leaflet and popups.
+Figure \ref{fig:interactive-popup} shows Figure \ref{fig:interactive-linking} made made with leaflet and popups [@leafpop].
 

### 2022-04-17 scripts/basic-manip.R

@@ -33,7 +33,7 @@ p1 <- ggplot() +
              width = 1, height = 0.5) +
   scale_color_brewer(palette = "Dark2") + 
   theme_bw() + 
-  coord_sf(xlim = c(141, 154)) + 
+  coord_sf(xlim = c(141, 154), ylim = c(-36, -29)) + 
   theme(legend.position = "bottom") +
   labs(x = "Longitude", y = "Latitude")
 

### 2022-04-17 2-the-cubble-pkg.Rmd
update codes with the latest cubble

--- 2022-03-01 2-the-cubble-pkg.Rmd
@@
 ```{r echo = TRUE}
-cubble_long <- cubble_nested %>% stretch()
+cubble_long <- cubble_nested %>% face_temporal()
 cubble_long
@@
 ```{r echo = TRUE}
-cubble_back <- cubble_long %>% tamp()
+cubble_back <- cubble_long %>% face_spatial()
 cubble_back
@@
 An example of this is the transformation of temporal variables into the spatial dimension in glyph maps (which will be elaborated in section \ref{st_transformation}).
-Cubble allows this operation through \code{migrate()}, which moves the specified spatial variables into the long cubble:
+Cubble allows this operation through \code{unfold()}, which moves the specified spatial variables into the long cubble:
 
 ```{r echo = TRUE}
-cubble_mig <- cubble_long %>% migrate(long, lat)
+cubble_mig <- cubble_long %>% unfold(long, lat)
 cubble_mig
@@
 ```{r echo = TRUE}
-cubble_mig %>% tamp() %>% stretch()
+cubble_mig %>% face_spatial() %>% face_temporal()
 ```
+
+<!-- ## Why not just use dplyrnest_by -->
+
+<!-- ```{r eval = FALSE} -->
+<!-- library(cubble) -->
+<!-- library(tidyverse) -->
+<!-- left_join(cubble::stations, cubble::climate) %>%  -->
+<!--   dplyr::nest_by(id, lat,  long,  elev, name, wmo_id) %>%  -->
+<!--   tidyr::unnest() %>%  -->
+<!--   group_by(........, geometry) -->
+<!-- ``` -->
+
+<!-- ```{r eval = FALSE} -->
+<!-- as_cubble(list(spatial = stations, temporal = climate), -->
+<!--           key = id, index = date, coords = c(long, lat)) -->
+<!-- ``` -->
+
+<!-- ```{r eval = FALSE} -->
+<!-- usmap::us_map("states") -->
+<!-- ``` -->
 

### 2022-04-17 4-1-basic.Rmd
flesh out section 2.1 - 2.4
update codes with the latest cubble

@@
 
-# Examples
+# Examples {#examples}
 
@@
 tmax <- tmax %>% 
-  stretch() %>%
+  face_temporal() %>%
   filter(lubridate::year(date) %in% c(1971:1975, 2016:2020)) 
@@
 These stations are filtered out by examining whether the summarised time series has a total of 24 months.
-The long cubble needs to be switched to the nested form for this spatial operation using \code{tamp()}:
+The long cubble needs to be switched to the nested form for this spatial operation using \code{face_spatial()}:
 
 ```{r eval = FALSE, echo = TRUE}
-tmax <- tmax %>% tamp() %>% filter(nrow(ts) == 24) 
+tmax <- tmax %>% face_spatial() %>% filter(nrow(ts) == 24) 
 ```
@@
 ```{r eval = FALSE, echo = TRUE}  
-tmax <- tmax %>% stretch() %>% migrate(latitude, longitude)
+tmax <- tmax %>% face_temporal() %>% unfold(latitude, longitude)
 ```

### 2022-04-17 4-3-water-level.Rmd
update codes with the latest cubble

@@
   filter(between(stringr::str_sub(id, 7, 8), 76, 90)) %>%
-  stretch() %>%
+  face_temporal() %>%
   filter(lubridate::year(date) == 2020) %>%
-  tamp() %>%
+  face_spatial() %>%
   mutate(type = "climate")
 
-river <- weatherdata::water %>%
-  mutate(type = "river")
+river <- river %>% mutate(type = "river")
 ```

### 2022-04-17 4-5-interactive.Rmd
update codes with the latest cubble

@@
 nested <- clean %>% SharedData$new(~id, group = "cubble")
-long <- stretch(clean) %>% SharedData$new(~id, group = "cubble")
+long <- face_temporal(clean) %>% SharedData$new(~id, group = "cubble")
 
@@
 # create nested form leaflet map with temperature band as subplots 
-nested <- tamp(clean)
+nested <- face_spatial(clean)
 leaflet(nested) %>% 

### 2022-04-17 main.tex
small changes
update codes with the latest cubble

@@
 ]{jss}
+
+%% recommended packages
+\usepackage{orcidlink,thumbpdf,lmodern}
 
@@
 \begin{CodeInput}
-R> cubble_long <- cubble_nested %>% stretch()
+R> cubble_long <- cubble_nested %>% face_temporal()
 R> cubble_long
@@
 \begin{CodeInput}
-R> cubble_back <- cubble_long %>% tamp()
+R> cubble_back <- cubble_long %>% face_spatial()
 R> cubble_back
@@
 \ref{st_transformation}). Cubble allows this operation through
-\code{migrate()}, which moves the specified spatial variables into the
+\code{unfold()}, which moves the specified spatial variables into the
 long cubble:
@@
 \begin{CodeInput}
-R> cubble_mig <- cubble_long %>% migrate(long, lat)
+R> cubble_mig <- cubble_long %>% unfold(long, lat)
 R> cubble_mig
@@
 \begin{CodeInput}
-R> cubble_mig %>% tamp() %>% stretch()
+R> cubble_mig %>% face_spatial() %>% face_temporal()
 \end{CodeInput}
@@
 R> tmax <- tmax %>% 
-+   stretch() %>%
++   face_temporal() %>%
 +   filter(lubridate::year(date) %in% c(1971:1975, 2016:2020)) 
@@
 series has a total of 24 months. The long cubble needs to be switched to
-the nested form for this spatial operation using \code{tamp()}:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> tmax <- tmax %>% tamp() %>% filter(nrow(ts) == 24) 
+the nested form for this spatial operation using \code{face_spatial()}:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> tmax <- tmax %>% face_spatial() %>% filter(nrow(ts) == 24) 
 \end{CodeInput}
@@
 \begin{CodeInput}
-R> tmax <- tmax %>% stretch() %>% migrate(latitude, longitude)
+R> tmax <- tmax %>% face_temporal() %>% unfold(latitude, longitude)
 \end{CodeInput}
@@
 
-\hypertarget{interative-graphic-with-cubble}{%
-\subsection{Interative graphic with
-cubble}\label{interative-graphic-with-cubble}}
+\hypertarget{interative-graphic}{%
+\subsection{Interative graphic}\label{interative-graphic}}
 
@@
 R> nested <- clean %>% SharedData$new(~id, group = "cubble")
-R> long <- stretch(clean) %>% SharedData$new(~id, group = "cubble")
+R> long <- face_temporal(clean) %>% SharedData$new(~id, group = "cubble")
 R> 
@@
 R> # create nested form leaflet map with temperature band as subplots 
-R> nested <- tamp(clean)
+R> nested <- face_spatial(clean)
 R> leaflet(nested) %>% 

### 2022-04-17 scripts/basic-agg.R

@@ -1,34 +1,32 @@
 library(tidyverse)
 library(cubble)
 library(patchwork)
-coords <- cbind(weatherdata::climate_full$long, weatherdata::climate_full$lat)
+coords <- cbind(prcp_aus$long, prcp_aus$lat)
 dist_raw <- geosphere::distm(coords, coords)
 
+station_long <- prcp_aus |> 
+  face_temporal(ts) |> 
+  mutate(wk = lubridate::week(date)) |> 
+  group_by(wk) |> 
+  summarise(prcp = sum(prcp, na.rm = TRUE)) |> 
+  unfold(cluster)
+
 set.seed(123)
-station_nested <- weatherdata::climate_full %>% 
+station_nested <- station_long %>% 
+  face_spatial() |> 
   strip_rowwise() %>% 
   mutate(cluster = kmeans(dist_raw,centers = 20, nstart = 500)$cluster)
-save(station_nested, file = here::here("data/station_nested.rda"))
+#save(station_nested, file = here::here("data/station_nested.rda"))
 
 cluster_nested <- station_nested %>%
   switch_key(cluster) %>% 
   get_centroid()
 
-station_long <- station_nested %>% 
-  stretch(ts) %>% 
-  mutate(wk = lubridate::week(date)) %>% 
-  group_by(wk) %>% 
-  summarise(prcp = sum(prcp, na.rm = TRUE)) %>% 
-  migrate(cluster)
-
 cluster_long <- cluster_nested %>% 
-  stretch(ts) %>% 
-  mutate(wk = lubridate::week(date)) %>% 
-  group_by(id, wk) %>%
-  summarise(prcp = sum(prcp, na.rm = TRUE)) %>%
-  ungroup(id) %>%
+  face_temporal(ts) %>% 
+  group_by(wk) %>%
   summarise(prcp = mean(prcp, na.rm = TRUE)) %>% 
-  migrate(cent_long, cent_lat)
+  unfold(cent_long, cent_lat)
 
 state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)
 

### 2022-04-17 scripts/basic-manip.R

@@ -1,11 +1,10 @@
 library(tidyverse)
-library(GGally)
 tmax <- weatherdata::historical_tmax %>%
   # filter out VIC and NSW stations
   filter(between(stringr::str_sub(id, 7, 8), 46, 90)) %>% 
   
   # filter out observations in between 1971-1975 and 2016 - 2020
-  stretch() %>%
+  face_temporal() %>%
   filter(lubridate::year(date) %in% c(1971:1975, 2016:2020)) %>%
   group_by(month = lubridate::month(date), 
          group = as.factor(ifelse(lubridate::year(date) > 2015, 
@@ -14,12 +13,12 @@ tmax <- weatherdata::historical_tmax %>%
   summarise(tmax = mean(tmax, na.rm = TRUE)) %>% 
   
   # filter out stations that have complete records in both periods 
-  tamp() %>%
+  face_spatial() %>%
   filter(nrow(ts) == 24) %>% 
   
-  # migrate longitude and latitude to the long form for making the glyph map
-  stretch() %>%
-  migrate(latitude, longitude)
+  # unfold longitude and latitude to the long form for making the glyph map
+  face_temporal() %>%
+  unfold(latitude, longitude)
 
 
 nsw_vic <- ozmaps::abs_ste %>% filter(NAME %in% c("Victoria", "New South Wales"))
@@ -57,7 +56,7 @@ tmax %>% filter(id == "ASN00048027") %>%
 
 #ggsave(filename = here::here("figures/basic-manip-inset.png"), width = 7)
 
-box_df <- tmax %>% tamp() %>% filter(id == "ASN00048027")
+box_df <- tmax %>% face_spatial() %>% filter(id == "ASN00048027")
 single <-
   tibble::tibble(img = here::here("figures/basic-manip-inset.png"))
 p1 +
@@ -73,6 +72,6 @@ p1 +
     color = "black"
   ) +
   ggimg::geom_point_img(data = single,
-                        aes(x = 143, y = -29, img = img), size = 6)
+                        aes(x = 143.7, y = -30, img = img), size = 6)
 
 ggsave(filename = "figures/basic-manip.png", width = 10, height = 10)

### 2022-04-17 scripts/illu-interactive.R

@@ -4,7 +4,7 @@ library(ggplot2)
 library(tsibble)
 
 set.seed(123)
-dt <- weatherdata::climate_full %>% slice_sample(n = 50)
+dt <- prcp_aus %>% slice_sample(n = 50)
 
 ###########################
 # from map to ts
@@ -22,7 +22,7 @@ plot_map(state_map) +
 ggsave(filename = "figures/keynotes-figures/illu-default-map.png")
 
 long <- dt %>% 
-  stretch() %>% 
+  face_temporal() %>% 
   mutate(month = lubridate::month(date)) %>% 
   group_by(month) %>% 
   summarise(tmax = mean(tmax, na.rm = TRUE)) %>% 

### 2022-04-17 scripts/interactive-linking.R

@@ -6,7 +6,7 @@ library(plotly)
 clean <- weatherdata::climate_full %>%
   filter(!id %in% c("ASN00067033", "ASN00072091", "ASN00059040", 
                     "ASN00097053", "ASN00040856", "ASN00015548")) %>%
-  stretch() %>% 
+  face_temporal() %>% 
   mutate(month = lubridate::month(date)) %>%
   group_by(month) %>%
   summarise(tmax = mean(tmax, na.rm = TRUE),
@@ -14,14 +14,14 @@ clean <- weatherdata::climate_full %>%
             diff = mean(tmax - tmin, na.rm = TRUE)) %>% 
   ungroup(month) %>%
   mutate(dummy_date = as.Date(glue::glue("2021-{month}-01"))) %>% 
-  tamp() %>% 
+  face_spatial() %>% 
   mutate(temp_diff_var = ts %>% pull(diff) %>% var(na.rm = TRUE))
   
 
 nested <- clean %>% 
   SharedData$new(~id, group = "cubble")
 
-long <- clean %>% stretch() %>% migrate(temp_diff_var) %>% arrange(temp_diff_var) %>% 
+long <- clean %>% face_temporal() %>% unfold(temp_diff_var) %>% arrange(temp_diff_var) %>% 
   SharedData$new(~id, group = "cubble")
 
 state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)

### 2022-04-17 scripts/interactive-popup.R

@@ -6,18 +6,18 @@ library(leafpop)
 clean <-  weatherdata::climate_full %>%
   filter(!id %in% c("ASN00067033", "ASN00072091", "ASN00059040", 
                     "ASN00097053", "ASN00040856", "ASN00015548")) %>%
-  stretch() %>%
+  face_temporal() %>%
   mutate(month = lubridate::month(date)) %>% 
   group_by(month) %>% 
   summarise(tmax = mean(tmax, na.rm = TRUE),
             tmin = mean(tmin, na.rm = TRUE),
             diff = mean(tmax - tmin, na.rm = TRUE)) %>% 
   ungroup(month) %>% 
-  tamp() %>%
+  face_spatial() %>%
   mutate(temp_diff_var = ts %>% pull(diff) %>% var(na.rm = TRUE)) %>% 
-  stretch() %>% 
+  face_temporal() %>% 
   mutate(dummy_date = as.Date(glue::glue("2021-{month}-01"))) %>% 
-  migrate(name)
+  unfold(name)
   
 df_id <- clean$id %>% unique()
 p <- purrr::map(1:length(df_id), function(i){
@@ -32,7 +32,7 @@ p <- purrr::map(1:length(df_id), function(i){
     ggtitle(paste0(stringr::str_to_title(unique(dt$name))))
 })
 
-nested <- tamp(clean) 
+nested <- face_spatial(clean) 
 domain <- nested$temp_diff_var
 pal <- colorNumeric(
   colorspace::sequential_hcl(

### 2022-04-17 scripts/matching.R

@@ -2,19 +2,23 @@ library(tidyverse)
 library(cubble)
 library(patchwork)
 
-climate <- weatherdata::climate_full %>%
+climate <- prcp_aus %>%
   filter(between(stringr::str_sub(id, 7, 8), 76, 90)) %>%
-  stretch() %>%
+  face_temporal() %>%
   filter(lubridate::year(date) == 2020) %>%
-  tamp() %>%
+  face_spatial() %>%
   mutate(type = "climate")
 
-river <- weatherdata::water %>% mutate(type = "river")
+river <- river %>% mutate(type = "river")
 
 vic_map <- rmapshaper::ms_simplify(ozmaps::abs_ste %>% filter(NAME == "Victoria"))  
-plot_map(vic_map) +
+ggplot() + 
+  geom_sf(data = vic_map, fill = "transparent", color = "grey") + 
   geom_point(data = dplyr::bind_rows(river, climate), 
              aes(x = long, y = lat, color = type)) + 
+  ggplot2::theme_bw() +
+  ggplot2::theme(legend.position = "bottom") +
+  ggplot2::labs(x = "Longitude", y = "Latitude") + 
   scale_color_brewer(palette = "Dark2") 
 ggsave(filename = "figures/matching-map.png", width = 10, height = 5)  
 
@@ -24,15 +28,19 @@ res <- match_sites(river, climate,
                    temporal_n_highest = 30,
                    temporal_min_match = 15)
 
-p1 <- plot_map(vic_map) +
+p1 <-ggplot() + 
+  geom_sf(data = vic_map, fill = "transparent", color = "grey") + 
   geom_point(data = res, aes(x = long, y = lat, color = type)) +
   ggrepel::geom_label_repel(
     data = res %>% filter(type == "river"), aes(x = long, y = lat, label = group)) +
-  scale_color_brewer(palette = "Dark2") 
+  scale_color_brewer(palette = "Dark2")  + 
+  ggplot2::theme_bw() +
+  ggplot2::theme(legend.position = "bottom") +
+  ggplot2::labs(x = "Longitude", y = "Latitude")
 
 res_long <- res %>%
-  stretch(ts) %>%
-  migrate(group, type) %>%
+  face_temporal(ts) %>%
+  unfold(group, type) %>%
   rename(prcp = matched_var) %>% 
   mutate(prcp = (prcp - min(prcp, na.rm = TRUE))/ (max(prcp, na.rm = TRUE) - min(prcp, na.rm = TRUE))) 
 

### 2022-04-17 scripts/netcdf.R

@@ -27,9 +27,9 @@ dt <- as_cubble(raw, vars = c("q", "z"))
 
 date <- c("2002-09-22", "2002-09-26", "2002-09-30", "2002-10-04") %>% as.Date()
 res <- dt %>% 
-  stretch() %>% 
+  face_temporal() %>% 
   filter(lubridate::date(time) %in% date) %>%
-  migrate(long, lat) %>% 
+  unfold(long, lat) %>% 
   mutate(q =  q* 10^6)
   
 box = c(xmin = -180, ymin = -90, xmax = 180, ymax = -5)

### 2022-04-17 2-the-cubble-pkg.Rmd
%>% to |>

@@
 ```{r echo = TRUE}
-cubble_nested <- cubble::climate_flat %>%
+cubble_nested <- cubble::climate_flat |>
   as_cubble(key = id, index = date, coords = c(long, lat))
@@
 ```{r echo = TRUE}
-cubble_long <- cubble_nested %>% face_temporal()
+cubble_long <- cubble_nested |> face_temporal()
 cubble_long
@@
 ```{r echo = TRUE}
-cubble_back <- cubble_long %>% face_spatial()
+cubble_back <- cubble_long |> face_spatial()
 cubble_back
@@
 ```{r echo = TRUE}
-cubble_mig <- cubble_long %>% unfold(long, lat)
+cubble_mig <- cubble_long |> unfold(long, lat)
 cubble_mig
@@
 ```{r echo = TRUE}
-cubble_mig %>% face_spatial() %>% face_temporal()
+cubble_mig |> face_spatial() |> face_temporal()
 ```
@@
 <!-- library(tidyverse) -->
-<!-- left_join(cubble::stations, cubble::climate) %>%  -->
-<!--   dplyr::nest_by(id, lat,  long,  elev, name, wmo_id) %>%  -->
-<!--   tidyr::unnest() %>%  -->
+<!-- left_join(cubble::stations, cubble::climate) |>  -->
+<!--   dplyr::nest_by(id, lat,  long,  elev, name, wmo_id) |>  -->
+<!--   tidyr::unnest() |>  -->
 <!--   group_by(........, geometry) -->
@@
 # example with a tsibble created from climate_flat
-raw <- climate_flat %>%
+raw <- climate_flat |>
   tsibble::as_tsibble(key = id, index = date)
 
-dt <-  raw %>%
+dt <-  raw |>
   cubble::as_cubble(coords = c(long, lat))
@@
 # add station-based features in the nested form.
-dt %>%
+dt |>
   mutate(fabletools::features(ts, tmax, list(tmax_mean = mean, tmax_var = var)))
@@
 # create a cubble
-cb <- climate_flat %>%
+cb <- climate_flat |>
   cubble::as_cubble(key = id, index = date, coords = c(long, lat))
@@
 
-dt <- cb %>%
+dt <- cb |>
   mutate(
@@
 # Western Australia map
-wa <- ozmaps::abs_ste %>% filter(NAME == "Western Australia")
+wa <- ozmaps::abs_ste |> filter(NAME == "Western Australia")
 
 # mutate a `s2_lnglat` vector on `cb` created in the last chunk
-cb %>%
-  mutate(ll = s2_lnglat(long, lat)) %>%
+cb |>
+  mutate(ll = s2_lnglat(long, lat)) |>
   filter(s2_within(ll, wa))

### 2022-04-17 3-advanced.Rmd
%>% to |>

@@
 Two peaks in series \\code{a} fall into the intervals and hence the two series are considered to have 2 matches."}
-knitr::include_graphics("figures/illu-matching.png")
+knitr::include_graphics(here::here("figures/illu-matching.png"))
 ```

### 2022-04-17 4-1-basic.Rmd
%>% to |>

@@
 ```{r eval = FALSE, echo = TRUE}
-tmax <- weatherdata::historical_tmax %>%
+tmax <- weatherdata::historical_tmax |>
   filter(between(stringr::str_sub(id, 7, 8), 46, 90))
@@
 ```{r eval=  FALSE, echo = TRUE}
-tmax <- tmax %>% 
-  face_temporal() %>%
+tmax <- tmax |> 
+  face_temporal() |>
   filter(lubridate::year(date) %in% c(1971:1975, 2016:2020)) 
@@
 ```{r eval=  FALSE, echo = TRUE}
-tmax <- tmax %>%
+tmax <- tmax |>
   group_by(month = lubridate::month(date), 
          group = as.factor(ifelse(lubridate::year(date) > 2015, 
-                                  "2016 ~ 2020", "1971 ~ 1975"))) %>%
+                                  "2016 ~ 2020", "1971 ~ 1975"))) |>
   summarise(tmax = mean(tmax, na.rm = TRUE))
@@
 ```{r eval = FALSE, echo = TRUE}
-tmax <- tmax %>% face_spatial() %>% filter(nrow(ts) == 24) 
+tmax <- tmax |> face_spatial() |> filter(nrow(ts) == 24) 
 ```
@@
 ```{r eval = FALSE, echo = TRUE}  
-tmax <- tmax %>% face_temporal() %>% unfold(latitude, longitude)
+tmax <- tmax |> face_temporal() |> unfold(latitude, longitude)
 ```
@@
 ```{r eval = FALSE, echo = TRUE}
-nsw_vic <- ozmaps::abs_ste %>% 
+nsw_vic <- ozmaps::abs_ste |> 
   filter(NAME %in% c("Victoria", "New South Wales"))
@@
 On the top left corner is an insetted plot of station Cobar highlighted in the black box."}
-knitr::include_graphics("figures/basic-manip.png")
+knitr::include_graphics(here::here("figures/basic-manip.png"))
 ```

### 2022-04-17 4-2-agg.Rmd
%>% to |>

@@
 ```{r echo = TRUE, eval = FALSE}
-station_nested <- weatherdata::climate_full %>% 
+station_nested <- weatherdata::climate_full |> 
   mutate(cluster = ...)
@@
 ```{r echo = TRUE}
-cluster_nested <- station_nested %>% switch_key(cluster) 
+cluster_nested <- station_nested |> switch_key(cluster) 
 ```
@@
 ```{r eval = FALSE, echo = TRUE}
-cluster_nested <- cluster_nested %>% get_centroid()
+cluster_nested <- cluster_nested |> get_centroid()
 ```
@@
 Subplot B shows the station membership of each cluster."}
-knitr::include_graphics("figures/basic-agg.png")
+knitr::include_graphics(here::here("figures/basic-agg.png"))
 ```

### 2022-04-17 4-3-water-level.Rmd
%>% to |>

--- 2022-02-13 4-3-water-level.Rmd
@@
 ```{r}
-climate <- weatherdata::climate_full %>%
-  filter(between(stringr::str_sub(id, 7, 8), 76, 90)) %>%
-  stretch() %>%
-  filter(lubridate::year(date) == 2020) %>%
-  tamp() %>%
+climate <- weatherdata::climate_full |>
+  filter(between(stringr::str_sub(id, 7, 8), 76, 90)) |>
+  face_temporal() |>
+  filter(lubridate::year(date) == 2020) |>
+  face_spatial() |>
   mutate(type = "climate")
 
-river <- weatherdata::water %>%
-  mutate(type = "river")
+river <- river |> mutate(type = "river")
 ```
@@
 ```{r matching-map, out.width="100%", fig.cap="Location of weather stations and river gauges in Victoria, Australia."}
-knitr::include_graphics("figures/matching-map.png")
+knitr::include_graphics(here::here("figures/matching-map.png"))
 ```
@@
 ```{r}
-res %>% print(n = 8)
+res |> print(n = 8)
 ```
@@
 The increases in the precipitation is reflected by the water level."}
-knitr::include_graphics("figures/matching.png")
+knitr::include_graphics(here::here("figures/matching.png"))
 ```

### 2022-04-17 4-5-interactive.Rmd
%>% to |>

--- 2022-04-17 4-5-interactive.Rmd
@@
 
-```{r eval = FALSE, echo = TRUE}
+``` r
 # data pre-processing
-clean <- weatherdata::climate_full %>% ...
+clean <- weatherdata::climate_full |> ...
 
 # created SharedData instance for crosstalk
-nested <- clean %>% SharedData$new(~id, group = "cubble")
-long <- stretch(clean) %>% SharedData$new(~id, group = "cubble")
+nested <- clean |> SharedData$new(~id, group = "cubble")
+long <- face_temporal(clean) |> SharedData$new(~id, group = "cubble")
 
 # create the spatial and temporal view each with a ShareData instance
-p1 <- nested %>% ...
-p2 <- long %>% ...
+p1 <- nested |> ...
+p2 <- long |> ...
 
@@
 
-```{r eval = FALSE, echo = TRUE}
+``` r
 # data pre-processing
-clean <- weatherdata::climate_full %>% ...
+clean <- weatherdata::climate_full |> ...
 
@@
 p <- map(1:length(df_id), function(i){
-  dt <- clean %>% filter(id == df_id[i])
-  ggplot(dt) %>% ...
+  dt <- clean |> filter(id == df_id[i])
+  ggplot(dt) |> ...
 })
@@
 # create nested form leaflet map with temperature band as subplots
-nested <- tamp(clean)
-leaflet(nested) %>% 
-  addTiles() %>% 
-  addCircleMarkers(group = "a", ...) %>% 
+nested <- face_spatial(clean)
+leaflet(nested) |>
+  addTiles() |>
+  addCircleMarkers(group = "a", ...) |>
   leafpop::addPopupGraphs(graph = p, ...)

### 2022-04-17 main.tex
%>% to |>

@@
 
+% Pandoc syntax highlighting
+\usepackage{color}
+\usepackage{fancyvrb}
+\newcommand{\VerbBar}{|}
+\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
+\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
+% Add ',fontsize=\small' for more characters per line
+\usepackage{framed}
+\definecolor{shadecolor}{RGB}{248,248,248}
+\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
+\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
+\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
+\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\BuiltInTok}[1]{#1}
+\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
+\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
+\newcommand{\ExtensionTok}[1]{#1}
+\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ImportTok}[1]{#1}
+\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\NormalTok}[1]{#1}
+\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
+\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
+\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\RegionMarkerTok}[1]{#1}
+\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+
 % tightlist command for lists without linebreak
@@
 \begin{CodeInput}
-R> cubble_nested <- cubble::climate_flat %>%
+R> cubble_nested <- cubble::climate_flat |>
 +   as_cubble(key = id, index = date, coords = c(long, lat))
@@
 \begin{CodeInput}
-R> cubble_long <- cubble_nested %>% face_temporal()
+R> cubble_long <- cubble_nested |> face_temporal()
 R> cubble_long
@@
 \begin{CodeInput}
-R> cubble_back <- cubble_long %>% face_spatial()
+R> cubble_back <- cubble_long |> face_spatial()
 R> cubble_back
@@
 \begin{CodeInput}
-R> cubble_mig <- cubble_long %>% unfold(long, lat)
+R> cubble_mig <- cubble_long |> unfold(long, lat)
 R> cubble_mig
@@
 \begin{CodeInput}
-R> cubble_mig %>% face_spatial() %>% face_temporal()
+R> cubble_mig |> face_spatial() |> face_temporal()
 \end{CodeInput}
@@
 R> # example with a tsibble created from climate_flat
-R> raw <- climate_flat %>%
+R> raw <- climate_flat |>
 +   tsibble::as_tsibble(key = id, index = date)
 R> 
-R> dt <-  raw %>%
+R> dt <-  raw |>
 +   cubble::as_cubble(coords = c(long, lat))
@@
 R> # add station-based features in the nested form.
-R> dt %>%
+R> dt |>
 +   mutate(fabletools::features(ts, tmax, list(tmax_mean = mean, tmax_var = var)))
@@
 R> # create a cubble
-R> cb <- climate_flat %>%
+R> cb <- climate_flat |>
 +   cubble::as_cubble(key = id, index = date, coords = c(long, lat))
@@
 R> 
-R> dt <- cb %>%
+R> dt <- cb |>
 +   mutate(
@@
 R> # Western Australia map
-R> wa <- ozmaps::abs_ste %>% filter(NAME == "Western Australia")
+R> wa <- ozmaps::abs_ste |> filter(NAME == "Western Australia")
 R> 
 R> # mutate a `s2_lnglat` vector on `cb` created in the last chunk
-R> cb %>%
-+   mutate(ll = s2_lnglat(long, lat)) %>%
+R> cb |>
++   mutate(ll = s2_lnglat(long, lat)) |>
 +   filter(s2_within(ll, wa))
@@
 \begin{CodeInput}
-R> tmax <- weatherdata::historical_tmax %>%
+R> tmax <- weatherdata::historical_tmax |>
 +   filter(between(stringr::str_sub(id, 7, 8), 46, 90))
@@
 \begin{CodeInput}
-R> tmax <- tmax %>% 
-+   face_temporal() %>%
+R> tmax <- tmax |> 
++   face_temporal() |>
 +   filter(lubridate::year(date) %in% c(1971:1975, 2016:2020)) 
@@
 \begin{CodeInput}
-R> tmax <- tmax %>%
+R> tmax <- tmax |>
 +   group_by(month = lubridate::month(date), 
 +          group = as.factor(ifelse(lubridate::year(date) > 2015, 
-+                                   "2016 ~ 2020", "1971 ~ 1975"))) %>%
++                                   "2016 ~ 2020", "1971 ~ 1975"))) |>
 +   summarise(tmax = mean(tmax, na.rm = TRUE))
@@
 \begin{CodeInput}
-R> tmax <- tmax %>% face_spatial() %>% filter(nrow(ts) == 24) 
+R> tmax <- tmax |> face_spatial() |> filter(nrow(ts) == 24) 
 \end{CodeInput}
@@
 \begin{CodeInput}
-R> tmax <- tmax %>% face_temporal() %>% unfold(latitude, longitude)
+R> tmax <- tmax |> face_temporal() |> unfold(latitude, longitude)
 \end{CodeInput}
@@
 \begin{CodeInput}
-R> nsw_vic <- ozmaps::abs_ste %>% 
+R> nsw_vic <- ozmaps::abs_ste |> 
 +   filter(NAME %in% c("Victoria", "New South Wales"))
@@
 \begin{CodeInput}
-R> station_nested <- weatherdata::climate_full %>% 
+R> station_nested <- weatherdata::climate_full |> 
 +   mutate(cluster = ...)
@@
 \begin{CodeInput}
-R> cluster_nested <- station_nested %>% switch_key(cluster) 
+R> cluster_nested <- station_nested |> switch_key(cluster) 
 \end{CodeInput}
@@
 \begin{CodeInput}
-R> cluster_nested <- cluster_nested %>% get_centroid()
+R> cluster_nested <- cluster_nested |> get_centroid()
 \end{CodeInput}
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> # data pre-processing
-R> clean <- weatherdata::climate_full %>% ...
-R> 
-R> # created SharedData instance for crosstalk
-R> nested <- clean %>% SharedData$new(~id, group = "cubble")
-R> long <- face_temporal(clean) %>% SharedData$new(~id, group = "cubble")
-R> 
-R> # create the spatial and temporal view each with a ShareData instance
-R> p1 <- nested %>% ...
-R> p2 <- long %>% ...
-R> 
-R> # Combine p1 and p2
-R> crosstalk::bscols(plotly::ggplotly(p1), plotly::ggplotly(p2), ...)
-\end{CodeInput}
-\end{CodeChunk}
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# data pre{-}processing}
+\NormalTok{clean }\OtherTok{\textless{}{-}}\NormalTok{ weatherdata}\SpecialCharTok{::}\NormalTok{climate\_full }\SpecialCharTok{|\textgreater{}}\NormalTok{ ...}
+
+\CommentTok{\# created SharedData instance for crosstalk}
+\NormalTok{nested }\OtherTok{\textless{}{-}}\NormalTok{ clean }\SpecialCharTok{|\textgreater{}}\NormalTok{ SharedData}\SpecialCharTok{$}\FunctionTok{new}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{id, }\AttributeTok{group =} \StringTok{"cubble"}\NormalTok{)}
+\NormalTok{long }\OtherTok{\textless{}{-}} \FunctionTok{face\_temporal}\NormalTok{(clean) }\SpecialCharTok{|\textgreater{}}\NormalTok{ SharedData}\SpecialCharTok{$}\FunctionTok{new}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{id, }\AttributeTok{group =} \StringTok{"cubble"}\NormalTok{)}
+
+\CommentTok{\# create the spatial and temporal view each with a ShareData instance}
+\NormalTok{p1 }\OtherTok{\textless{}{-}}\NormalTok{ nested }\SpecialCharTok{|\textgreater{}}\NormalTok{ ...}
+\NormalTok{p2 }\OtherTok{\textless{}{-}}\NormalTok{ long }\SpecialCharTok{|\textgreater{}}\NormalTok{ ...}
+
+\CommentTok{\# Combine p1 and p2}
+\NormalTok{crosstalk}\SpecialCharTok{::}\FunctionTok{bscols}\NormalTok{(plotly}\SpecialCharTok{::}\FunctionTok{ggplotly}\NormalTok{(p1), plotly}\SpecialCharTok{::}\FunctionTok{ggplotly}\NormalTok{(p2), ...)}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> # data pre-processing
-R> clean <- weatherdata::climate_full %>% ...
-R> 
-R> # use the long form to create subplots for each station
-R> df_id <- unique(clean$id)
-R> p <- map(1:length(df_id), function(i){
-+   dt <- clean %>% filter(id == df_id[i])
-+   ggplot(dt) %>% ...
-+ })
-R> 
-R> # create nested form leaflet map with temperature band as subplots 
-R> nested <- face_spatial(clean)
-R> leaflet(nested) %>% 
-+   addTiles() %>% 
-+   addCircleMarkers(group = "a", ...) %>% 
-+   leafpop::addPopupGraphs(graph = p, ...)
-\end{CodeInput}
-\end{CodeChunk}
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# data pre{-}processing}
+\NormalTok{clean }\OtherTok{\textless{}{-}}\NormalTok{ weatherdata}\SpecialCharTok{::}\NormalTok{climate\_full }\SpecialCharTok{|\textgreater{}}\NormalTok{ ...}
+
+\CommentTok{\# use the long form to create subplots for each station}
+\NormalTok{df\_id }\OtherTok{\textless{}{-}} \FunctionTok{unique}\NormalTok{(clean}\SpecialCharTok{$}\NormalTok{id)}
+\NormalTok{p }\OtherTok{\textless{}{-}} \FunctionTok{map}\NormalTok{(}\DecValTok{1}\SpecialCharTok{:}\FunctionTok{length}\NormalTok{(df\_id), }\ControlFlowTok{function}\NormalTok{(i)\{}
+\NormalTok{  dt }\OtherTok{\textless{}{-}}\NormalTok{ clean }\SpecialCharTok{|\textgreater{}} \FunctionTok{filter}\NormalTok{(id }\SpecialCharTok{==}\NormalTok{ df\_id[i])}
+  \FunctionTok{ggplot}\NormalTok{(dt) }\SpecialCharTok{|\textgreater{}}\NormalTok{ ...}
+\NormalTok{\})}
+
+\CommentTok{\# create nested form leaflet map with temperature band as subplots}
+\NormalTok{nested }\OtherTok{\textless{}{-}} \FunctionTok{face\_spatial}\NormalTok{(clean)}
+\FunctionTok{leaflet}\NormalTok{(nested) }\SpecialCharTok{|\textgreater{}}
+  \FunctionTok{addTiles}\NormalTok{() }\SpecialCharTok{|\textgreater{}}
+  \FunctionTok{addCircleMarkers}\NormalTok{(}\AttributeTok{group =} \StringTok{"a"}\NormalTok{, ...) }\SpecialCharTok{|\textgreater{}}
+\NormalTok{  leafpop}\SpecialCharTok{::}\FunctionTok{addPopupGraphs}\NormalTok{(}\AttributeTok{graph =}\NormalTok{ p, ...)}
+\end{Highlighting}
+\end{Shaded}
 

### 2022-04-17 scripts/basic-agg.R

@@ -12,30 +12,30 @@ station_long <- prcp_aus |>
   unfold(cluster)
 
 set.seed(123)
-station_nested <- station_long %>% 
+station_nested <- station_long |> 
   face_spatial() |> 
-  strip_rowwise() %>% 
+  strip_rowwise() |> 
   mutate(cluster = kmeans(dist_raw,centers = 20, nstart = 500)$cluster)
 #save(station_nested, file = here::here("data/station_nested.rda"))
 
-cluster_nested <- station_nested %>%
-  switch_key(cluster) %>% 
+cluster_nested <- station_nested |>
+  switch_key(cluster) |> 
   get_centroid()
 
-cluster_long <- cluster_nested %>% 
-  face_temporal(ts) %>% 
-  group_by(wk) %>%
-  summarise(prcp = mean(prcp, na.rm = TRUE)) %>% 
+cluster_long <- cluster_nested |> 
+  face_temporal(ts) |> 
+  group_by(wk) |>
+  summarise(prcp = mean(prcp, na.rm = TRUE)) |> 
   unfold(cent_long, cent_lat)
 
 state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)
 
-ggplot_smooth <- cluster_long %>% 
+ggplot_smooth <- cluster_long |> 
   ggplot() +
   geom_smooth(aes(x = wk, y = prcp, group = cluster), span = 0.4) 
 
-smoother <- layer_data(ggplot_smooth) %>% 
-  left_join(cluster_long %>% select(cluster, cent_long, cent_lat), by = c("group" = "cluster"))
+smoother <- layer_data(ggplot_smooth) |> 
+  left_join(cluster_long |> select(cluster, cent_long, cent_lat), by = c("group" = "cluster"))
 
 p1 <- ggplot(data = smoother, 
              aes(x_minor = x, y_minor = y, x_major = cent_long, y_major = cent_lat)) + 
@@ -50,7 +50,7 @@ p3 <- ggplot() +
   geom_sf(data = state_map, inherit.aes = FALSE,
           color = "grey80", alpha = 0.4, linetype = 3) + 
   geom_point(data = station_nested, aes(x = long, y = lat), size = 0.5) +
-  ggforce::geom_mark_hull(data = cluster_nested %>% tidyr::unnest(hull),
+  ggforce::geom_mark_hull(data = cluster_nested |> tidyr::unnest(hull),
                           expand = 0, radius = 0,
[…]

### 2022-04-17 scripts/basic-manip.R

@@ -1,27 +1,27 @@
 library(tidyverse)
-tmax <- weatherdata::historical_tmax %>%
+tmax <- weatherdata::historical_tmax |>
   # filter out VIC and NSW stations
-  filter(between(stringr::str_sub(id, 7, 8), 46, 90)) %>% 
+  filter(between(stringr::str_sub(id, 7, 8), 46, 90)) |> 
   
   # filter out observations in between 1971-1975 and 2016 - 2020
-  face_temporal() %>%
-  filter(lubridate::year(date) %in% c(1971:1975, 2016:2020)) %>%
+  face_temporal() |>
+  filter(lubridate::year(date) %in% c(1971:1975, 2016:2020)) |>
   group_by(month = lubridate::month(date), 
          group = as.factor(ifelse(lubridate::year(date) > 2015, 
-                                  "2016 ~ 2020", "1971 ~ 1975"))) %>%
+                                  "2016 ~ 2020", "1971 ~ 1975"))) |>
   # summarise tmax into monthly mean for both periods
-  summarise(tmax = mean(tmax, na.rm = TRUE)) %>% 
+  summarise(tmax = mean(tmax, na.rm = TRUE)) |> 
   
   # filter out stations that have complete records in both periods 
-  face_spatial() %>%
-  filter(nrow(ts) == 24) %>% 
+  face_spatial() |>
+  filter(nrow(ts) == 24) |> 
   
   # unfold longitude and latitude to the long form for making the glyph map
-  face_temporal() %>%
+  face_temporal() |>
   unfold(latitude, longitude)
 
 
-nsw_vic <- ozmaps::abs_ste %>% filter(NAME %in% c("Victoria", "New South Wales"))
+nsw_vic <- ozmaps::abs_ste |> filter(NAME %in% c("Victoria", "New South Wales"))
 
 p1 <- ggplot() + 
   geom_sf(data = nsw_vic, fill = "transparent", color = "grey", linetype = "dotted") + 
@@ -39,7 +39,7 @@ p1 <- ggplot() +
 
 #################
 # inset
-tmax %>% filter(id == "ASN00048027") %>%
+tmax |> filter(id == "ASN00048027") |>
   ggplot(aes(x = month,
              y = tmax,
              color = group)) +
@@ -56,7 +56,7 @@ tmax %>% filter(id == "ASN00048027") %>%
 
 #ggsave(filename = here::here("figures/basic-manip-inset.png"), width = 7)
 
[…]

### 2022-04-17 scripts/illu-interactive.R

@@ -4,14 +4,14 @@ library(ggplot2)
 library(tsibble)
 
 set.seed(123)
-dt <- prcp_aus %>% slice_sample(n = 50)
+dt <- prcp_aus |> slice_sample(n = 50)
 
 ###########################
 # from map to ts
 state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)
 plot_map(state_map) +
   geom_point(data = dt, aes(x = long, y = lat)) + 
-  geom_point(data = dt %>% filter(id == "ASN00003057"), aes(x = long, y = lat), 
+  geom_point(data = dt |> filter(id == "ASN00003057"), aes(x = long, y = lat), 
              color = "red", size =5) 
 
 ggsave(filename = "figures/keynotes-figures/illu-highlighted-map.png")
@@ -21,16 +21,16 @@ plot_map(state_map) +
   
 ggsave(filename = "figures/keynotes-figures/illu-default-map.png")
 
-long <- dt %>% 
-  face_temporal() %>% 
-  mutate(month = lubridate::month(date)) %>% 
-  group_by(month) %>% 
-  summarise(tmax = mean(tmax, na.rm = TRUE)) %>% 
+long <- dt |> 
+  face_temporal() |> 
+  mutate(month = lubridate::month(date)) |> 
+  group_by(month) |> 
+  summarise(tmax = mean(tmax, na.rm = TRUE)) |> 
   mutate(dummy_date = as.Date(glue::glue("2021-{month}-01")))
 
-highlight_line <- long %>% filter(id == "ASN00003057")
-#highlight_point <- highlight_line %>% filter(month == 9)
-long %>% 
+highlight_line <- long |> filter(id == "ASN00003057")
+#highlight_point <- highlight_line |> filter(month == 9)
+long |> 
   ggplot(aes(x = dummy_date, y = tmax, group = id)) +
   #geom_point(size = 0.3) + 
   geom_line() + 
@@ -41,7 +41,7 @@ long %>%
   
 ggsave(filename = "figures/keynotes-figures/illu-highlighted-long.png", width = 5, height = 3)
 
-long %>% 
+long |> 
   ggplot(aes(x = dummy_date, y = tmax, group = id)) +
   geom_line() + 
   scale_x_date(date_labels = "%b") + 
@@ -52,9 +52,9 @@ ggsave(filename = "figures/keynotes-figures/illu-default-long.png", width = 5, h
 ###########################
 # from ts to map
 
[…]

### 2022-04-17 scripts/illu-matching.R

@@ -10,7 +10,7 @@ dt <- tibble(
       4, rnorm(5, sd = 0.5), 6, rnorm(4, sd = 0.5)
     )
   )
-) %>% mutate(val = ifelse(val < 0, 0, val))
+) |> mutate(val = ifelse(val < 0, 0, val))
 
 circle <- tibble(
   x = c(6, 14, 23, 7, 21, 27),
@@ -21,15 +21,15 @@ circle <- tibble(
 ) 
 
 errorbar <- bind_rows(
-  circle %>% filter(id == "A"),
-  circle %>% filter(id == "A") %>% mutate(id = "a")
-) %>%
+  circle |> filter(id == "A"),
+  circle |> filter(id == "A") |> mutate(id = "a")
+) |>
   mutate(id = factor(id, c("A", "a")))
 
 ggplot() +
   geom_rect(data = errorbar, aes(xmin = x, xmax = xend, ymin = 0, ymax = 10), fill = "grey90") +
   geom_line(data = dt, aes(x = date, y = val, group = id), color = "black") +
-  geom_vline(data = circle %>% filter(id == "A"), 
+  geom_vline(data = circle |> filter(id == "A"), 
              aes(xintercept = x), linetype = "longdash", color = "grey10", lwd = 0.2) +
   geom_point(data = circle, aes(x = x, y = y, color = match), size = 3) +
   facet_wrap(vars(id), nrow = 2) +

### 2022-04-17 scripts/interactive-linking.R

@@ -3,29 +3,29 @@ library(tidyverse)
 library(crosstalk)
 library(plotly)
 
-clean <- weatherdata::climate_full %>%
+clean <- weatherdata::climate_full |>
   filter(!id %in% c("ASN00067033", "ASN00072091", "ASN00059040", 
-                    "ASN00097053", "ASN00040856", "ASN00015548")) %>%
-  face_temporal() %>% 
-  mutate(month = lubridate::month(date)) %>%
-  group_by(month) %>%
+                    "ASN00097053", "ASN00040856", "ASN00015548")) |>
+  face_temporal() |> 
+  mutate(month = lubridate::month(date)) |>
+  group_by(month) |>
   summarise(tmax = mean(tmax, na.rm = TRUE),
             tmin = mean(tmin, na.rm = TRUE),
-            diff = mean(tmax - tmin, na.rm = TRUE)) %>% 
-  ungroup(month) %>%
-  mutate(dummy_date = as.Date(glue::glue("2021-{month}-01"))) %>% 
-  face_spatial() %>% 
-  mutate(temp_diff_var = ts %>% pull(diff) %>% var(na.rm = TRUE))
+            diff = mean(tmax - tmin, na.rm = TRUE)) |> 
+  ungroup(month) |>
+  mutate(dummy_date = as.Date(glue::glue("2021-{month}-01"))) |> 
+  face_spatial() |> 
+  mutate(temp_diff_var = ts |> pull(diff) |> var(na.rm = TRUE))
   
 
-nested <- clean %>% 
+nested <- clean |> 
   SharedData$new(~id, group = "cubble")
 
-long <- clean %>% face_temporal() %>% unfold(temp_diff_var) %>% arrange(temp_diff_var) %>% 
+long <- clean |> face_temporal() |> unfold(temp_diff_var) |> arrange(temp_diff_var) |> 
   SharedData$new(~id, group = "cubble")
 
 state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)
-p1 <- nested %>% 
+p1 <- nested |> 
   ggplot() +
   geom_sf(data = state_map, aes(geometry = geometry), fill = "transparent", color = "grey50") +
   geom_point(aes(x = long, y = lat,color = temp_diff_var, label = name)) +
@@ -38,7 +38,7 @@ p1 <- nested %>%
   theme(panel.grid.major = element_blank(),
         legend.position = "none")
 
[…]

### 2022-04-17 scripts/interactive-popup.R

@@ -3,26 +3,26 @@ library(tidyverse)
 library(leaflet)
 library(leafpop)
 
-clean <-  weatherdata::climate_full %>%
+clean <-  weatherdata::climate_full |>
   filter(!id %in% c("ASN00067033", "ASN00072091", "ASN00059040", 
-                    "ASN00097053", "ASN00040856", "ASN00015548")) %>%
-  face_temporal() %>%
-  mutate(month = lubridate::month(date)) %>% 
-  group_by(month) %>% 
+                    "ASN00097053", "ASN00040856", "ASN00015548")) |>
+  face_temporal() |>
+  mutate(month = lubridate::month(date)) |> 
+  group_by(month) |> 
   summarise(tmax = mean(tmax, na.rm = TRUE),
             tmin = mean(tmin, na.rm = TRUE),
-            diff = mean(tmax - tmin, na.rm = TRUE)) %>% 
-  ungroup(month) %>% 
-  face_spatial() %>%
-  mutate(temp_diff_var = ts %>% pull(diff) %>% var(na.rm = TRUE)) %>% 
-  face_temporal() %>% 
-  mutate(dummy_date = as.Date(glue::glue("2021-{month}-01"))) %>% 
+            diff = mean(tmax - tmin, na.rm = TRUE)) |> 
+  ungroup(month) |> 
+  face_spatial() |>
+  mutate(temp_diff_var = ts |> pull(diff) |> var(na.rm = TRUE)) |> 
+  face_temporal() |> 
+  mutate(dummy_date = as.Date(glue::glue("2021-{month}-01"))) |> 
   unfold(name)
   
-df_id <- clean$id %>% unique()
+df_id <- clean$id |> unique()
 p <- purrr::map(1:length(df_id), function(i){
-  dt <- clean %>% filter(id == df_id[i])
-  dt %>% 
+  dt <- clean |> filter(id == df_id[i])
+  dt |> 
     ggplot(aes(x = dummy_date, ymin = tmin, ymax = tmax, group = id)) + 
     geom_ribbon(color = "grey50", fill = "grey50", alpha = 0.5) + 
     theme_bw() + 
@@ -40,9 +40,9 @@ pal <- colorNumeric(
     c2 = 40, l2= 85, c1 = 20, l1 = 30),
   domain = domain)
 
-out <- leaflet(nested) %>% 
-  addTiles() %>% 
[…]

### 2022-04-17 scripts/matching.R

@@ -2,16 +2,16 @@ library(tidyverse)
 library(cubble)
 library(patchwork)
 
-climate <- prcp_aus %>%
-  filter(between(stringr::str_sub(id, 7, 8), 76, 90)) %>%
-  face_temporal() %>%
-  filter(lubridate::year(date) == 2020) %>%
-  face_spatial() %>%
+climate <- prcp_aus |>
+  filter(between(stringr::str_sub(id, 7, 8), 76, 90)) |>
+  face_temporal() |>
+  filter(lubridate::year(date) == 2020) |>
+  face_spatial() |>
   mutate(type = "climate")
 
-river <- river %>% mutate(type = "river")
+river <- river |> mutate(type = "river")
 
-vic_map <- rmapshaper::ms_simplify(ozmaps::abs_ste %>% filter(NAME == "Victoria"))  
+vic_map <- rmapshaper::ms_simplify(ozmaps::abs_ste |> filter(NAME == "Victoria"))  
 ggplot() + 
   geom_sf(data = vic_map, fill = "transparent", color = "grey") + 
   geom_point(data = dplyr::bind_rows(river, climate), 
@@ -32,19 +32,19 @@ p1 <-ggplot() +
   geom_sf(data = vic_map, fill = "transparent", color = "grey") + 
   geom_point(data = res, aes(x = long, y = lat, color = type)) +
   ggrepel::geom_label_repel(
-    data = res %>% filter(type == "river"), aes(x = long, y = lat, label = group)) +
+    data = res |> filter(type == "river"), aes(x = long, y = lat, label = group)) +
   scale_color_brewer(palette = "Dark2")  + 
   ggplot2::theme_bw() +
   ggplot2::theme(legend.position = "bottom") +
   ggplot2::labs(x = "Longitude", y = "Latitude")
 
-res_long <- res %>%
-  face_temporal(ts) %>%
-  unfold(group, type) %>%
-  rename(prcp = matched_var) %>% 
+res_long <- res |>
+  face_temporal(ts) |>
+  unfold(group, type) |>
+  rename(prcp = matched_var) |> 
   mutate(prcp = (prcp - min(prcp, na.rm = TRUE))/ (max(prcp, na.rm = TRUE) - min(prcp, na.rm = TRUE))) 
 
-p2 <- res_long %>% 
+p2 <- res_long |> 
   ggplot(aes(x = date, y = prcp, color = type, group = id)) +
   geom_line(alpha = 0.8) +
   facet_wrap(vars(group)) +

### 2022-04-17 scripts/netcdf.R

@@ -25,20 +25,20 @@ wf_request(
 raw <- ncdf4::nc_open(here::here("data/era5-pressure.nc"))
 dt <- as_cubble(raw, vars = c("q", "z"))
 
-date <- c("2002-09-22", "2002-09-26", "2002-09-30", "2002-10-04") %>% as.Date()
-res <- dt %>% 
-  face_temporal() %>% 
-  filter(lubridate::date(time) %in% date) %>%
-  unfold(long, lat) %>% 
+date <- c("2002-09-22", "2002-09-26", "2002-09-30", "2002-10-04") |> as.Date()
+res <- dt |> 
+  face_temporal() |> 
+  filter(lubridate::date(time) %in% date) |>
+  unfold(long, lat) |> 
   mutate(q =  q* 10^6)
   
 box = c(xmin = -180, ymin = -90, xmax = 180, ymax = -5)
 sf::sf_use_s2(FALSE)
-country <- rnaturalearth::ne_coastline("small", returnclass = "sf") %>% 
-  sf::st_crop(box) %>% 
+country <- rnaturalearth::ne_coastline("small", returnclass = "sf") |> 
+  sf::st_crop(box) |> 
   sf::st_cast("MULTILINESTRING")
 
-res %>% 
+res |> 
   ggplot() +
   geom_point(aes(x = long, y = lat, color = q)) +
   geom_contour(data = res, aes(x = long, y = lat, z = z),

### 2022-04-17 arxiv/cubble-arxiv.Rmd
arxiv version

--- 2022-03-01 main.Rmd
@@
 ---
-documentclass: jss
-author:
+title: cubble An R Package for Structuring Spatio-temporal Data
+authors:
   - name: H. Sherry Zhang
     affiliation: Monash University
-    address: 21 Chancellors Walk, Clayton VIC 3800 Australia
-    email: \email{huize.zhang@monash.edu}
+    location: 21 Chancellors Walk, Clayton VIC 3800 Australia
+    email: huize.zhang@monash.edu
   - name: Dianne Cook
-    affiliation: 'Monash University \AND'
-    address: 21 Chancellors Walk, Clayton VIC 3800 Australia
-    # location: Melbourne, Australia
+    affiliation: Monash University
+    location: 21 Chancellors Walk, Clayton VIC 3800 Australia
     email: dicook@monash.edu
   - name: Ursula Laa
-    affiliation: 'University of Natural Resources and Life Sciences \AND'
-    address: Gregor-Mendel-Straße 33, 1180 Wien, Austria
+    affiliation: University of Natural Resources and Life Sciences
+    location: Gregor-Mendel-Straße 33, 1180 Wien, Austria
     email:  ursula.laa@boku.ac.at
   - name: Nicolas Langrené
-    affiliation: 'BNU-HKBU United International College \AND'
-    address: 2000 Jintong Road, Tangjiawan, Zhuhai, Guangdong Province, China
+    affiliation: BNU-HKBU United International College
+    location: 2000 Jintong Road, Tangjiawan, Zhuhai, Guangdong Province, China
     email: nicolaslangrene@uic.edu.cn
   - name: Patricia Menéndez
-    affiliation: 'Monash University \AND'
-    address: 21 Chancellors Walk, Clayton VIC 3800 Australia
+    affiliation: Monash University
+    location: 21 Chancellors Walk, Clayton VIC 3800 Australia
     email:  patricia.menendez@monash.edu
-title:
-  formatted: "\\pkg{cubble}: An R Package for Structuring Spatio-temporal Data"
-  # If you use tex in the formatted title, alsoge supply version without
-  plain:     "cubble: An R Package for Structuring Spatio-temporal Data"
-  # For running headers, if needed
-  # short:     "\\pkg{foo}: A Capitalized Title"
-abstract: >
+abstract: |
     Spatio-temporal data refer to measurements taken across space and time.
@@
 keywords:
-  # at least one keyword must be supplied
-  formatted: [spatial, temporal, spatio temporal, "\\proglang{R}", environmental data, exploratory data analysis]
-  plain:     [spatial, temporal, spatio temporal, R, environmental data, exploratory data analysis]
-preamble: >
-  \usepackage{amsmath}
-  \usepackage{array}
-  \usepackage{xcolor}
-output: rticles::jss_article
-bibliography: references.bib
+- spatial, temporal, spatio temporal, R, environmental data, exploratory data analysis
+bibliography: "../references.bib"
+cls: "../jss.cls"
+bst: "../jss.bst"
+output: rticles::arxiv_article
+link-citations: true
 ---
-
 ```{r setup, echo = FALSE}
@@
   cache = TRUE,
-  fig.path = "figures/",
-  fig.height = 10
+  fig.path = here::here("figures/"),
+  fig.height = 10, 
+  root.dir = here::here()
 )
@@
 
-```{r echo = FALSE}
+```{r}
 library(cubble)
-library(forcats)
-library(stringr)
 library(dplyr)
-library(lubridate)
-library(ggplot2)
-library(GGally)
-library(plotly)
-library(crosstalk)
-library(patchwork)
 ```
@@
 
-```{r child = "1-intro.Rmd"}
+```{r child = here::here("1-intro.Rmd")}
 
@@
 
-```{r child = "2-the-cubble-pkg.Rmd"}
+```{r child = here::here("2-0-cube.Rmd")}
 
@@
 
-```{r child = "3-advanced.Rmd"}
+```{r child = here::here("2-the-cubble-pkg.Rmd")}
 
@@
 
-```{r child = "4-1-basic.Rmd"}
+```{r child = here::here("3-advanced.Rmd")}
 
@@
 
-```{r child = "4-2-agg.Rmd"}
+```{r child = here::here("4-1-basic.Rmd")}
 
@@
 
-```{r child = "4-3-water-level.Rmd"}
+```{r child = here::here("4-2-agg.Rmd")}
 
@@
 
-```{r child = "4-4-netcdf.Rmd"}
+```{r child = here::here("4-3-water-level.Rmd")}
 
@@
 
-```{r child = "4-5-interactive.Rmd"}
+```{r child = here::here("4-4-netcdf.Rmd")}
 
@@
 
-```{r child = "5-conclusion.Rmd"}
+```{r child = here::here("4-5-interactive.Rmd")}
 
 ```
+
+```{r child = here::here("5-conclusion.Rmd")}
+
+```

### 2022-04-17 arxiv/cubble-arxiv.tex
arxiv version

--- 2022-04-17 main.tex
@@
-\documentclass[
-]{jss}
-
-%% recommended packages
-\usepackage{orcidlink,thumbpdf,lmodern}
-
-\usepackage[utf8]{inputenc}
+\documentclass{article}
+
+\usepackage{arxiv}
+
+\usepackage[utf8]{inputenc} % allow utf-8 input
+\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
+\usepackage{lmodern}        % https://github.com/rstudio/rticles/issues/343
+\usepackage{hyperref}       % hyperlinks
+\usepackage{url}            % simple URL typesetting
+\usepackage{booktabs}       % professional-quality tables
+\usepackage{amsfonts}       % blackboard math symbols
+\usepackage{nicefrac}       % compact symbols for 1/2, etc.
+\usepackage{microtype}      % microtypography
+\usepackage{graphicx}
+
+\title{cubble An R Package for Structuring Spatio-temporal Data}
 
 \author{
-H. Sherry Zhang\\Monash University \And Dianne Cook\\Monash University
-\AND Ursula Laa\\University of Natural Resources and Life Sciences
-\AND Nicolas Langrené\\BNU-HKBU United International College
-\AND Patricia Menéndez\\Monash University \AND
-}
-\title{\pkg{cubble}: An R Package for Structuring Spatio-temporal Data}
-
-\Plainauthor{H. Sherry Zhang, Dianne Cook, Ursula Laa, Nicolas
-Langrené, Patricia Menéndez}
-\Plaintitle{cubble: An R Package for Structuring Spatio-temporal Data}
-
-\Abstract{
-Spatio-temporal data refer to measurements taken across space and time.
-In practice, spatio-temporal data can be decomposed into a spatial and
-temporal component: at one time, we would select a spatial location and
-inspect the temporal trend; at other time, we might select one or
-multiple time value(s) and explore the spatial distribution. Ideally, we
-could make multiple maps and multiple time series to explore these
-together, however, doing all of these actions is complicated when data
-arrive fragmented in multiple objects. To make it easy to do all these
-tasks, ideally spatial and temporal variables are in a single data
-object that we can slice and dice in different ways to conduct different
-visualisations. This work suggests a new data structure, \code{cubble},
-that uses a nested form and a long form to organise spatial and temporal
-variables in a single data object. The new data structure ensures that
-data in the two forms are synchronised while been flexible to explore
-the two aspect of spatio-temporal data. It also provides capability for
-handling data with hierarchical structure, matching data from multiple
-sources, constructing interactive graphics, and performing
-spatio-temporal transformation. The paper will demonstrate \pkg{cubble}
-with examples using Australian climate weather stations, river level
-data, and climate reanalysis (ERA5) data.
-}
-
-\Keywords{spatial, temporal, spatio
-temporal, \proglang{R}, environmental data, exploratory data analysis}
-\Plainkeywords{spatial, temporal, spatio temporal, R, environmental
-data, exploratory data analysis}
-
-%% publication information
-%% \Volume{50}
-%% \Issue{9}
-%% \Month{June}
-%% \Year{2012}
-%% \Submitdate{}
-%% \Acceptdate{2012-06-04}
-
-\Address{
-    H. Sherry Zhang\\
-    Monash University\\
-    21 Chancellors Walk, Clayton VIC 3800 Australia\\
-  E-mail: \email{huize.zhang@monash.edu}\\
-
-      Dianne Cook\\
-    Monash University\\
-    21 Chancellors Walk, Clayton VIC 3800 Australia\\
-  E-mail: \href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}\\
-
-      Ursula Laa\\
-    University of Natural Resources and Life Sciences\\
-    Gregor-Mendel-Straße 33, 1180 Wien, Austria\\
-  E-mail: \href{mailto:ursula.laa@boku.ac.at}{\nolinkurl{ursula.laa@boku.ac.at}}\\
-
-      Nicolas Langrené\\
-    BNU-HKBU United International College\\
-    2000 Jintong Road, Tangjiawan, Zhuhai, Guangdong Province, China\\
-  E-mail: \href{mailto:nicolaslangrene@uic.edu.cn}{\nolinkurl{nicolaslangrene@uic.edu.cn}}\\
-
-      Patricia Menéndez\\
-    Monash University\\
-    21 Chancellors Walk, Clayton VIC 3800 Australia\\
-  E-mail: \href{mailto:patricia.menendez@monash.edu}{\nolinkurl{patricia.menendez@monash.edu}}\\
-
+    H. Sherry Zhang
+   \\
+    Monash University \\
+  21 Chancellors Walk, Clayton VIC 3800 Australia \\
+  \texttt{\href{mailto:huize.zhang@monash.edu}{\nolinkurl{huize.zhang@monash.edu}}} \\
+   \And
+    Dianne Cook
+   \\
+    Monash University \\
+  21 Chancellors Walk, Clayton VIC 3800 Australia \\
+  \texttt{\href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}} \\
+   \And
+    Ursula Laa
+   \\
+    University of Natural Resources and Life Sciences \\
+  Gregor-Mendel-Straße 33, 1180 Wien, Austria \\
+  \texttt{\href{mailto:ursula.laa@boku.ac.at}{\nolinkurl{ursula.laa@boku.ac.at}}} \\
+   \And
+    Nicolas Langrené
+   \\
+    BNU-HKBU United International College \\
+  2000 Jintong Road, Tangjiawan, Zhuhai, Guangdong Province, China \\
+  \texttt{\href{mailto:nicolaslangrene@uic.edu.cn}{\nolinkurl{nicolaslangrene@uic.edu.cn}}} \\
+   \And
+    Patricia Menéndez
+   \\
+    Monash University \\
+  21 Chancellors Walk, Clayton VIC 3800 Australia \\
+  \texttt{\href{mailto:patricia.menendez@monash.edu}{\nolinkurl{patricia.menendez@monash.edu}}} \\
   }
@@
 
-\usepackage{amsmath} \usepackage{array} \usepackage{xcolor}
+% Pandoc citation processing
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newlength{\cslentryspacingunit} % times entry-spacing
+\setlength{\cslentryspacingunit}{\parskip}
+% for Pandoc 2.8 to 2.10.1
+\newenvironment{cslreferences}%
+  {}%
+  {\par}
+% For Pandoc 2.11+
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1
+  \let\oldpar\par
+  \def\par{\hangindent=\cslhangindent\oldpar}
+  \fi
+  % set entry spacing
+  \setlength{\parskip}{#2\cslentryspacingunit}
+ }%
+ {}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 
 \begin{document}
+\maketitle
+
+\begin{abstract}
+Spatio-temporal data refer to measurements taken across space and time.
+In practice, spatio-temporal data can be decomposed into a spatial and
+temporal component: at one time, we would select a spatial location and
+inspect the temporal trend; at other time, we might select one or
+multiple time value(s) and explore the spatial distribution. Ideally, we
+could make multiple maps and multiple time series to explore these
+together, however, doing all of these actions is complicated when data
+arrive fragmented in multiple objects. To make it easy to do all these
+tasks, ideally spatial and temporal variables are in a single data
+object that we can slice and dice in different ways to conduct different
+visualisations. This work suggests a new data structure, \code{cubble},
+that uses a nested form and a long form to organise spatial and temporal
+variables in a single data object. The new data structure ensures that
+data in the two forms are synchronised while been flexible to explore
+the two aspect of spatio-temporal data. It also provides capability for
+handling data with hierarchical structure, matching data from multiple
+sources, constructing interactive graphics, and performing
+spatio-temporal transformation. The paper will demonstrate \pkg{cubble}
+with examples using Australian climate weather stations, river level
+data, and climate reanalysis (ERA5) data.
+\end{abstract}
+
+\keywords{
+    spatial, temporal, spatio temporal, R, environmental data,
+    exploratory data analysis
+  }
 
@@
 
-\begin{CodeChunk}
 \begin{figure}
@@
 
-\caption[An illustration of different spatio-temporal data layouts]{An illustration of different spatio-temporal data layouts: (1) a single table for both spatial and temporal variables; (2) separate tables for spatial and temporal variables; and (3) an array data structure that populates variables into a gridded longitude-latitude-time (hyper)cube.}\label{fig:illu-input}
+\caption{An illustration of different spatio-temporal data layouts: (1) a single table for both spatial and temporal variables; (2) separate tables for spatial and temporal variables; and (3) an array data structure that populates variables into a gridded longitude-latitude-time (hyper)cube.}\label{fig:illu-input}
 \end{figure}
-\end{CodeChunk}
 
@@
 work with specific spatio-temporal tasks, while few can freely explore
-the data. \pkg{spacetime} \citep{spacetime} is an attempt for such a
-data object, but the spatial and temporal classes it is built on have
-now been gradually replaced by more recent implementations. The
-\pkg{stars} \citep{stars} package uses an array to wrangle
-spatio-temporal data, but the dense array it uses can quickly exhaust
-the memory for data in the case of irregular spatial grid. While there
-has not yet been a data object that makes it easy to wrangle
+the data. \pkg{spacetime} (\protect\hyperlink{ref-spacetime}{E. Pebesma
+2012}) is an attempt for such a data object, but the spatial and
+temporal classes it is built on have now been gradually replaced by more
+recent implementations. The \pkg{stars}
+(\protect\hyperlink{ref-stars}{E. Pebesma 2021}) package uses an array
+to wrangle spatio-temporal data, but the dense array it uses can quickly
+exhaust the memory for data in the case of irregular spatial grid. While
+there has not yet been a data object that makes it easy to wrangle
 spatio-temporal data, there is a strong need for such tools given the
@@
 data, as evident by satellite imageries, large climate models, among
-others. The work by \citet{lu_multidimensional_2018} has discussed the
-suitability of this representation through array algebra and practical
-spatio-temporal tasks.
+others. The work by Lu, Appel, and Pebesma
+(\protect\hyperlink{ref-lu_multidimensional_2018}{2018}) has discussed
+the suitability of this representation through array algebra and
+practical spatio-temporal tasks.
 
 The cubical conceptual framework provides a generalisation to the
-operation can be done on the data for visualisation. A paper by
-\citet{bach_review_2014} has reviewed the temporal data visualisation
-based on space-time cube operations. Notice that the term space-time
-cube in their article ``does not need to involve spatial data'', but
-referring to ``an abstract 2D substrate that is used to visualize data
-at a specific time''. Despite its main focus is on temporal data, the
-mindset of abstracting out data representation to construct
-visualisation is applicable to our spatio-temporal data manipulation and
-visualisation.
+operation can be done on the data for visualisation. A paper by Bach et
+al. (\protect\hyperlink{ref-bach_review_2014}{2014}) has reviewed the
+temporal data visualisation based on space-time cube operations. Notice
+that the term space-time cube in their article ``does not need to
+involve spatial data'', but referring to ``an abstract 2D substrate that
+is used to visualize data at a specific time''. Despite its main focus
+is on temporal data, the mindset of abstracting out data representation
+to construct visualisation is applicable to our spatio-temporal data
+manipulation and visualisation.
 
@@
 
-To represent spatio-temporal data, \citet{spacetime} proposes four
-layouts: full grid layout (STF), sparse grid layout (STS), irregular
-layout (STI), and trajectory (STT). These layouts are implemented in the
-\pkg{spacetime} package with underlying class \pkg{sp} \citep{sp} and
-\pkg{xts} \citep{xts} for handling spatial and temporal variables. While
-these layouts and classes are still being used and well-maintained
-today, they are not compatible with the \pkg{tidyverse} ecosystem
-\citep{tidyverse}, which has a huge influence on data analysis in
-\proglang{R}, including in the spatial and temporal communities. The
-package \pkg{sf} \citep{sf} uses data frame objects to represent spatial
-data with an additional \code{geometry} column dedicating to feature
-geometries \citep{jr_herring_opengis_2011} (e.g.~POINT, MULTIPOINT,
-POLYGON, MULTIPOLYGON etc). For temporal data, \pkg{tsibble}
-\citep{tsibble} builds on top of the \pkg{tibble} \citep{tibble} object
-and explicitly includes date and time in columns as variables, rather
-than data attributes.
-
-Underlying tidyverse is the concept of tidy data \citep{tidydata}, which
-considers how variables should be structured for analysis. In the three
-formats (time-wide, space-wide, and long) summarised in the
-\citet{spacetime} paper, only the long form is considered as tidy. The
-main advantages of it over the time-wide and space-wide format is that
-the latter treat either time or location as variable names (Section 3.1
-in \citet{tidydata}) rather than values of variables, making them
-difficult to be wrangled. One problem with the long format is that it
-can be inefficient to store feature geometries, especially for large
-MULTIPOLYGON over long time period. This motivates a new tidyverse
-compatible data object to work with spatio-temporal data.
+To represent spatio-temporal data, E. Pebesma
+(\protect\hyperlink{ref-spacetime}{2012}) proposes four layouts: full
+grid layout (STF), sparse grid layout (STS), irregular layout (STI), and
+trajectory (STT). These layouts are implemented in the \pkg{spacetime}
+package with underlying class \pkg{sp} (\protect\hyperlink{ref-sp}{E.
+Pebesma and Bivand 2005}) and \pkg{xts}
+(\protect\hyperlink{ref-xts}{Ryan and Ulrich 2020}) for handling spatial
+and temporal variables. While these layouts and classes are still being
+used and well-maintained today, they are not compatible with the
+\pkg{tidyverse} ecosystem (\protect\hyperlink{ref-tidyverse}{Wickham et
+al. 2019}), which has a huge influence on data analysis in \proglang{R},
+including in the spatial and temporal communities. The package \pkg{sf}
+(\protect\hyperlink{ref-sf}{E. J. Pebesma 2018}) uses data frame objects
+to represent spatial data with an additional \code{geometry} column
+dedicating to feature geometries
+(\protect\hyperlink{ref-jr_herring_opengis_2011}{J.R. Herring 2011})
+(e.g.~POINT, MULTIPOINT, POLYGON, MULTIPOLYGON etc). For temporal data,
+\pkg{tsibble} (\protect\hyperlink{ref-tsibble}{Wang, Cook, and Hyndman
+2020a}) builds on top of the \pkg{tibble}
+(\protect\hyperlink{ref-tibble}{Müller and Wickham 2021}) object and
+explicitly includes date and time in columns as variables, rather than
+data attributes.
+
+Underlying tidyverse is the concept of tidy data
+(\protect\hyperlink{ref-tidydata}{Wickham 2014}), which considers how
+variables should be structured for analysis. In the three formats
+(time-wide, space-wide, and long) summarised in the E. Pebesma
+(\protect\hyperlink{ref-spacetime}{2012}) paper, only the long form is
+considered as tidy. The main advantages of it over the time-wide and
+space-wide format is that the latter treat either time or location as
+variable names (Section 3.1 in Wickham
+(\protect\hyperlink{ref-tidydata}{2014})) rather than values of
+variables, making them difficult to be wrangled. One problem with the
+long format is that it can be inefficient to store feature geometries,
+especially for large MULTIPOLYGON over long time period. This motivates
+a new tidyverse compatible data object to work with spatio-temporal
+data.
 
@@
 \end{figure}
-\end{CodeChunk}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> cubble_nested <- cubble::climate_flat |>
-+   as_cubble(key = id, index = date, coords = c(long, lat))
-R> cubble_nested
-\end{CodeInput}
-\begin{CodeOutput}
-# cubble:   id [5]: nested form
-# bbox:     [115.97, -32.94, 133.55, -12.42]
-# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
-  id            lat  long  elev name           wmo_id ts                
-  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>            
-1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tibble [366 x 4]>
-2 ASN00010311 -31.9  117. 179   york            94623 <tibble [366 x 4]>
-3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble [366 x 4]>
-4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble [366 x 4]>
-5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble [366 x 4]>
-\end{CodeOutput}
-\end{CodeChunk}
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{cubble\_nested }\OtherTok{\textless{}{-}}\NormalTok{ cubble}\SpecialCharTok{::}\NormalTok{climate\_flat }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
+  \FunctionTok{as\_cubble}\NormalTok{(}\AttributeTok{key =}\NormalTok{ id, }\AttributeTok{index =}\NormalTok{ date, }\AttributeTok{coords =} \FunctionTok{c}\NormalTok{(long, lat))}
+\NormalTok{cubble\_nested}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## # cubble:   id [5]: nested form
+## # bbox:     [115.97, -32.94, 133.55, -12.42]
+## # temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
+##   id            lat  long  elev name           wmo_id ts                
+##   <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>            
+## 1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tibble [366 x 4]>
+## 2 ASN00010311 -31.9  117. 179   york            94623 <tibble [366 x 4]>
+## 3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble [366 x 4]>
+## 4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble [366 x 4]>
+## 5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble [366 x 4]>
+\end{verbatim}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> cubble_long <- cubble_nested |> face_temporal()
-R> cubble_long
-\end{CodeInput}
-\begin{CodeOutput}
-# cubble:  date, id [5]: long form
-# bbox:    [115.97, -32.94, 133.55, -12.42]
-# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]
-  id          date        prcp  tmax  tmin
-  <chr>       <date>     <dbl> <dbl> <dbl>
-1 ASN00009021 2020-01-01     0  31.9  15.3
-2 ASN00009021 2020-01-02     0  24.9  16.4
-3 ASN00009021 2020-01-03     6  23.2  13  
-4 ASN00009021 2020-01-04     0  28.4  12.4
-5 ASN00009021 2020-01-05     0  35.3  11.6
-# ... with 1,825 more rows
-\end{CodeOutput}
-\end{CodeChunk}
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{cubble\_long }\OtherTok{\textless{}{-}}\NormalTok{ cubble\_nested }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{face\_temporal}\NormalTok{()}
+\NormalTok{cubble\_long}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## # cubble:  date, id [5]: long form
+## # bbox:    [115.97, -32.94, 133.55, -12.42]
+## # spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]
+##   id          date        prcp  tmax  tmin
+##   <chr>       <date>     <dbl> <dbl> <dbl>
+## 1 ASN00009021 2020-01-01     0  31.9  15.3
+## 2 ASN00009021 2020-01-02     0  24.9  16.4
+## 3 ASN00009021 2020-01-03     6  23.2  13  
+## 4 ASN00009021 2020-01-04     0  28.4  12.4
+## 5 ASN00009021 2020-01-05     0  35.3  11.6
+## # ... with 1,825 more rows
+\end{verbatim}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> attr(cubble_long, "spatial")
-\end{CodeInput}
-\begin{CodeOutput}
-# A tibble: 5 x 6
-  id            lat  long  elev name           wmo_id
-  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl>
-1 ASN00009021 -31.9  116.  15.4 perth airport   94610
-2 ASN00010311 -31.9  117. 179   york            94623
-3 ASN00010614 -32.9  117. 338   narrogin        94627
-4 ASN00014015 -12.4  131.  30.4 darwin airport  94120
-5 ASN00015131 -17.6  134. 220   elliott         94236
-\end{CodeOutput}
-\end{CodeChunk}
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{attr}\NormalTok{(cubble\_long, }\StringTok{"spatial"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## # A tibble: 5 x 6
+##   id            lat  long  elev name           wmo_id
+##   <chr>       <dbl> <dbl> <dbl> <chr>           <dbl>
+## 1 ASN00009021 -31.9  116.  15.4 perth airport   94610
+## 2 ASN00010311 -31.9  117. 179   york            94623
+## 3 ASN00010614 -32.9  117. 338   narrogin        94627
+## 4 ASN00014015 -12.4  131.  30.4 darwin airport  94120
+## 5 ASN00015131 -17.6  134. 220   elliott         94236
+\end{verbatim}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> cubble_back <- cubble_long |> face_spatial()
-R> cubble_back
-\end{CodeInput}
-\begin{CodeOutput}
-# cubble:   id [5]: nested form
-# bbox:     [115.97, -32.94, 133.55, -12.42]
-# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
-  id            lat  long  elev name           wmo_id ts                
-  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>            
-1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tibble [366 x 4]>
-2 ASN00010311 -31.9  117. 179   york            94623 <tibble [366 x 4]>
-3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble [366 x 4]>
-4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble [366 x 4]>
-5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble [366 x 4]>
-\end{CodeOutput}
-\end{CodeChunk}
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{cubble\_back }\OtherTok{\textless{}{-}}\NormalTok{ cubble\_long }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{face\_spatial}\NormalTok{()}
+\NormalTok{cubble\_back}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## # cubble:   id [5]: nested form
+## # bbox:     [115.97, -32.94, 133.55, -12.42]
+## # temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
+##   id            lat  long  elev name           wmo_id ts                
+##   <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>            
+## 1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tibble [366 x 4]>
+## 2 ASN00010311 -31.9  117. 179   york            94623 <tibble [366 x 4]>
+## 3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble [366 x 4]>
+## 4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble [366 x 4]>
+## 5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble [366 x 4]>
+\end{verbatim}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> cubble_mig <- cubble_long |> unfold(long, lat)
-R> cubble_mig
-\end{CodeInput}
-\begin{CodeOutput}
-# cubble:  date, id [5]: long form
-# bbox:    [115.97, -32.94, 133.55, -12.42]
-# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]
-  id          date        prcp  tmax  tmin  long   lat
-  <chr>       <date>     <dbl> <dbl> <dbl> <dbl> <dbl>
-1 ASN00009021 2020-01-01     0  31.9  15.3  116. -31.9
-2 ASN00009021 2020-01-02     0  24.9  16.4  116. -31.9
-3 ASN00009021 2020-01-03     6  23.2  13    116. -31.9
-4 ASN00009021 2020-01-04     0  28.4  12.4  116. -31.9
-5 ASN00009021 2020-01-05     0  35.3  11.6  116. -31.9
-# ... with 1,825 more rows
-\end{CodeOutput}
-\end{CodeChunk}
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{cubble\_mig }\OtherTok{\textless{}{-}}\NormalTok{ cubble\_long }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{unfold}\NormalTok{(long, lat)}
+\NormalTok{cubble\_mig}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## # cubble:  date, id [5]: long form
+## # bbox:    [115.97, -32.94, 133.55, -12.42]
+## # spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]
+##   id          date        prcp  tmax  tmin  long   lat
+##   <chr>       <date>     <dbl> <dbl> <dbl> <dbl> <dbl>
+## 1 ASN00009021 2020-01-01     0  31.9  15.3  116. -31.9
+## 2 ASN00009021 2020-01-02     0  24.9  16.4  116. -31.9
+## 3 ASN00009021 2020-01-03     6  23.2  13    116. -31.9
+## 4 ASN00009021 2020-01-04     0  28.4  12.4  116. -31.9
+## 5 ASN00009021 2020-01-05     0  35.3  11.6  116. -31.9
+## # ... with 1,825 more rows
+\end{verbatim}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> cubble_mig |> face_spatial() |> face_temporal()
-\end{CodeInput}
-\begin{CodeOutput}
-# cubble:  date, id [5]: long form
-# bbox:    [115.97, -32.94, 133.55, -12.42]
-# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]
-  id          date        prcp  tmax  tmin
-  <chr>       <date>     <dbl> <dbl> <dbl>
-1 ASN00009021 2020-01-01     0  31.9  15.3
-2 ASN00009021 2020-01-02     0  24.9  16.4
-3 ASN00009021 2020-01-03     6  23.2  13  
-4 ASN00009021 2020-01-04     0  28.4  12.4
-5 ASN00009021 2020-01-05     0  35.3  11.6
-# ... with 1,825 more rows
-\end{CodeOutput}
-\end{CodeChunk}
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{cubble\_mig }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{face\_spatial}\NormalTok{() }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{face\_temporal}\NormalTok{()}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## # cubble:  date, id [5]: long form
+## # bbox:    [115.97, -32.94, 133.55, -12.42]
+## # spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]
+##   id          date        prcp  tmax  tmin
+##   <chr>       <date>     <dbl> <dbl> <dbl>
+## 1 ASN00009021 2020-01-01     0  31.9  15.3
+## 2 ASN00009021 2020-01-02     0  24.9  16.4
+## 3 ASN00009021 2020-01-03     6  23.2  13  
+## 4 ASN00009021 2020-01-04     0  28.4  12.4
+## 5 ASN00009021 2020-01-05     0  35.3  11.6
+## # ... with 1,825 more rows
+\end{verbatim}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> # example with a tsibble created from climate_flat
-R> raw <- climate_flat |>
-+   tsibble::as_tsibble(key = id, index = date)
-R> 
-R> dt <-  raw |>
-+   cubble::as_cubble(coords = c(long, lat))
-R> dt
-\end{CodeInput}
-\begin{CodeOutput}
-# cubble:   id [5]: nested form
-# bbox:     [115.97, -32.94, 133.55, -12.42]
-# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
-  id            lat  long  elev name           wmo_id ts                
-  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>            
-1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tbl_ts [366 x 4]>
-2 ASN00010311 -31.9  117. 179   york            94623 <tbl_ts [366 x 4]>
-3 ASN00010614 -32.9  117. 338   narrogin        94627 <tbl_ts [366 x 4]>
-4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tbl_ts [366 x 4]>
-5 ASN00015131 -17.6  134. 220   elliott         94236 <tbl_ts [366 x 4]>
-\end{CodeOutput}
-\end{CodeChunk}
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# example with a tsibble created from climate\_flat}
+\NormalTok{raw }\OtherTok{\textless{}{-}}\NormalTok{ climate\_flat }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
+\NormalTok{  tsibble}\SpecialCharTok{::}\FunctionTok{as\_tsibble}\NormalTok{(}\AttributeTok{key =}\NormalTok{ id, }\AttributeTok{index =}\NormalTok{ date)}
+
+\NormalTok{dt }\OtherTok{\textless{}{-}}\NormalTok{  raw }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
+\NormalTok{  cubble}\SpecialCharTok{::}\FunctionTok{as\_cubble}\NormalTok{(}\AttributeTok{coords =} \FunctionTok{c}\NormalTok{(long, lat))}
+\NormalTok{dt}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## # cubble:   id [5]: nested form
+## # bbox:     [115.97, -32.94, 133.55, -12.42]
+## # temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
+##   id            lat  long  elev name           wmo_id ts                
+##   <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>            
+## 1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tbl_ts [366 x 4]>
+## 2 ASN00010311 -31.9  117. 179   york            94623 <tbl_ts [366 x 4]>
+## 3 ASN00010614 -32.9  117. 338   narrogin        94627 <tbl_ts [366 x 4]>
+## 4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tbl_ts [366 x 4]>
+## 5 ASN00015131 -17.6  134. 220   elliott         94236 <tbl_ts [366 x 4]>
+\end{verbatim}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> # add station-based features in the nested form.
-R> dt |>
-+   mutate(fabletools::features(ts, tmax, list(tmax_mean = mean, tmax_var = var)))
-\end{CodeInput}
-\begin{CodeOutput}
-# cubble:   id [5]: nested form
-# bbox:     [115.97, -32.94, 133.55, -12.42]
-# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
-  id            lat  long  elev name          wmo_id ts       tmax_mean tmax_var
-  <chr>       <dbl> <dbl> <dbl> <chr>          <dbl> <list>       <dbl>    <dbl>
-1 ASN00009021 -31.9  116.  15.4 perth airport  94610 <tbl_ts>      25.7    38.6 
-2 ASN00010311 -31.9  117. 179   york           94623 <tbl_ts>      26.2    51.1 
-3 ASN00010614 -32.9  117. 338   narrogin       94627 <tbl_ts>      23.7    45.4 
-4 ASN00014015 -12.4  131.  30.4 darwin airpo~  94120 <tbl_ts>      33.1     3.02
-5 ASN00015131 -17.6  134. 220   elliott        94236 <tbl_ts>      34.6    24.7 
-\end{CodeOutput}
-\end{CodeChunk}
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# add station{-}based features in the nested form.}
+\NormalTok{dt }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
+  \FunctionTok{mutate}\NormalTok{(fabletools}\SpecialCharTok{::}\FunctionTok{features}\NormalTok{(ts, tmax, }\FunctionTok{list}\NormalTok{(}\AttributeTok{tmax\_mean =}\NormalTok{ mean, }\AttributeTok{tmax\_var =}\NormalTok{ var)))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## # cubble:   id [5]: nested form
+## # bbox:     [115.97, -32.94, 133.55, -12.42]
+## # temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
+##   id            lat  long  elev name          wmo_id ts       tmax_mean tmax_var
+##   <chr>       <dbl> <dbl> <dbl> <chr>          <dbl> <list>       <dbl>    <dbl>
+## 1 ASN00009021 -31.9  116.  15.4 perth airport  94610 <tbl_ts>      25.7    38.6 
+## 2 ASN00010311 -31.9  117. 179   york           94623 <tbl_ts>      26.2    51.1 
+## 3 ASN00010614 -32.9  117. 338   narrogin       94627 <tbl_ts>      23.7    45.4 
+## 4 ASN00014015 -12.4  131.  30.4 darwin airpo~  94120 <tbl_ts>      33.1     3.02
+## 5 ASN00015131 -17.6  134. 220   elliott        94236 <tbl_ts>      34.6    24.7
+\end{verbatim}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> library(sf)
-R> # create a cubble
-R> cb <- climate_flat |>
-+   cubble::as_cubble(key = id, index = date, coords = c(long, lat))
-R> 
-R> aus <- ozmaps::abs_ste
-R> 
-R> dt <- cb |>
-+   mutate(
-+     # create `sfc` column based on long and lat
-+     ll = st_sfc(
-+       purrr::map2(long, lat, ~st_point(c(.x, .y))),
-+       crs = st_crs(aus)),
-+ 
-+     # append state multi-polygon based on the `sfc` created
-+     state = aus$geometry[st_within(ll, aus, sparse = FALSE)],
-+ 
-+     # adopt a different projection: lambert conformal conic (EPSG:3112)
-+     state = st_transform(state, crs = "EPSG:3112")
-+     )
-R> 
-R> dt
-\end{CodeInput}
-\begin{CodeOutput}
-# cubble:   id [5]: nested form
-# bbox:     [115.97, -32.94, 133.55, -12.42]
-# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
-  id            lat  long  elev name   wmo_id ts                        ll
-  <chr>       <dbl> <dbl> <dbl> <chr>   <dbl> <list>           <POINT [°]>
-1 ASN00009021 -31.9  116.  15.4 perth~  94610 <tibble> (115.9764 -31.9275)
-2 ASN00010311 -31.9  117. 179   york    94623 <tibble>  (116.765 -31.8997)
-3 ASN00010614 -32.9  117. 338   narro~  94627 <tibble> (117.1797 -32.9342)
-4 ASN00014015 -12.4  131.  30.4 darwi~  94120 <tibble> (130.8925 -12.4239)
-5 ASN00015131 -17.6  134. 220   ellio~  94236 <tibble> (133.5407 -17.5521)
-# ... with 1 more variable: state <MULTIPOLYGON [m]>
-\end{CodeOutput}
-\end{CodeChunk}
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{library}\NormalTok{(sf)}
+\CommentTok{\# create a cubble}
+\NormalTok{cb }\OtherTok{\textless{}{-}}\NormalTok{ climate\_flat }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
+\NormalTok{  cubble}\SpecialCharTok{::}\FunctionTok{as\_cubble}\NormalTok{(}\AttributeTok{key =}\NormalTok{ id, }\AttributeTok{index =}\NormalTok{ date, }\AttributeTok{coords =} \FunctionTok{c}\NormalTok{(long, lat))}
+
+\NormalTok{aus }\OtherTok{\textless{}{-}}\NormalTok{ ozmaps}\SpecialCharTok{::}\NormalTok{abs\_ste}
+
+\NormalTok{dt }\OtherTok{\textless{}{-}}\NormalTok{ cb }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
+  \FunctionTok{mutate}\NormalTok{(}
+    \CommentTok{\# create \textasciigrave{}sfc\textasciigrave{} column based on long and lat}
+    \AttributeTok{ll =} \FunctionTok{st\_sfc}\NormalTok{(}
+\NormalTok{      purrr}\SpecialCharTok{::}\FunctionTok{map2}\NormalTok{(long, lat, }\SpecialCharTok{\textasciitilde{}}\FunctionTok{st\_point}\NormalTok{(}\FunctionTok{c}\NormalTok{(.x, .y))),}
+      \AttributeTok{crs =} \FunctionTok{st\_crs}\NormalTok{(aus)),}
+
+    \CommentTok{\# append state multi{-}polygon based on the \textasciigrave{}sfc\textasciigrave{} created}
+    \AttributeTok{state =}\NormalTok{ aus}\SpecialCharTok{$}\NormalTok{geometry[}\FunctionTok{st\_within}\NormalTok{(ll, aus, }\AttributeTok{sparse =} \ConstantTok{FALSE}\NormalTok{)],}
+
+    \CommentTok{\# adopt a different projection: lambert conformal conic (EPSG:3112)}
+    \AttributeTok{state =} \FunctionTok{st\_transform}\NormalTok{(state, }\AttributeTok{crs =} \StringTok{"EPSG:3112"}\NormalTok{)}
+\NormalTok{    )}
+
+\NormalTok{dt}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## # cubble:   id [5]: nested form
+## # bbox:     [115.97, -32.94, 133.55, -12.42]
+## # temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
+##   id            lat  long  elev name   wmo_id ts                        ll
+##   <chr>       <dbl> <dbl> <dbl> <chr>   <dbl> <list>           <POINT [°]>
+## 1 ASN00009021 -31.9  116.  15.4 perth~  94610 <tibble> (115.9764 -31.9275)
+## 2 ASN00010311 -31.9  117. 179   york    94623 <tibble>  (116.765 -31.8997)
+## 3 ASN00010614 -32.9  117. 338   narro~  94627 <tibble> (117.1797 -32.9342)
+## 4 ASN00014015 -12.4  131.  30.4 darwi~  94120 <tibble> (130.8925 -12.4239)
+## 5 ASN00015131 -17.6  134. 220   ellio~  94236 <tibble> (133.5407 -17.5521)
+## # ... with 1 more variable: state <MULTIPOLYGON [m]>
+\end{verbatim}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> library(s2)
-R> # Western Australia map
-R> wa <- ozmaps::abs_ste |> filter(NAME == "Western Australia")
-R> 
-R> # mutate a `s2_lnglat` vector on `cb` created in the last chunk
-R> cb |>
-+   mutate(ll = s2_lnglat(long, lat)) |>
-+   filter(s2_within(ll, wa))
-\end{CodeInput}
-\begin{CodeOutput}
-# cubble:   id [3]: nested form
-# bbox:     [115.97, -32.94, 117.18, -31.89]
-# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
-  id            lat  long  elev name          wmo_id ts                 ll      
-  <chr>       <dbl> <dbl> <dbl> <chr>          <dbl> <list>             <s2_lng>
-1 ASN00009021 -31.9  116.  15.4 perth airport  94610 <tibble [366 x 4]> (115.97~
-2 ASN00010311 -31.9  117. 179   york           94623 <tibble [366 x 4]> (116.76~
-3 ASN00010614 -32.9  117. 338   narrogin       94627 <tibble [366 x 4]> (117.17~
-\end{CodeOutput}
-\end{CodeChunk}
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{library}\NormalTok{(s2)}
+\CommentTok{\# Western Australia map}
+\NormalTok{wa }\OtherTok{\textless{}{-}}\NormalTok{ ozmaps}\SpecialCharTok{::}\NormalTok{abs\_ste }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{filter}\NormalTok{(NAME }\SpecialCharTok{==} \StringTok{"Western Australia"}\NormalTok{)}
+
+\CommentTok{\# mutate a \textasciigrave{}s2\_lnglat\textasciigrave{} vector on \textasciigrave{}cb\textasciigrave{} created in the last chunk}
+\NormalTok{cb }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{ll =} \FunctionTok{s2\_lnglat}\NormalTok{(long, lat)) }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
+  \FunctionTok{filter}\NormalTok{(}\FunctionTok{s2\_within}\NormalTok{(ll, wa))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## # cubble:   id [3]: nested form
+## # bbox:     [115.97, -32.94, 117.18, -31.89]
+## # temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
+##   id            lat  long  elev name          wmo_id ts                 ll      
+##   <chr>       <dbl> <dbl> <dbl> <chr>          <dbl> <list>             <s2_lng>
+## 1 ASN00009021 -31.9  116.  15.4 perth airport  94610 <tibble [366 x 4]> (115.97~
+## 2 ASN00010311 -31.9  117. 179   york           94623 <tibble [366 x 4]> (116.76~
+## 3 ASN00010614 -32.9  117. 338   narrogin       94627 <tibble [366 x 4]> (117.17~
+\end{verbatim}
 
@@
 A few packages in R exists for manipulating NetCDF data and this
-includes a high-level R interface: \pkg{ncdf4} \citep{ncdf4}, a
-low-level interface that calls C-interface: \pkg{RNetCDF}
-\citep{rnetcdf, michna2013rnetcdf}, and a tidyverse implementation:
-\pkg{tidync} \citep{tidync}.
+includes a high-level R interface: \pkg{ncdf4}
+(\protect\hyperlink{ref-ncdf4}{Pierce 2019}), a low-level interface that
+calls C-interface: \pkg{RNetCDF} (\protect\hyperlink{ref-rnetcdf}{Michna
+and Woods 2021}, \protect\hyperlink{ref-michna2013rnetcdf}{2013}), and a
+tidyverse implementation: \pkg{tidync}
+(\protect\hyperlink{ref-tidync}{Sumner 2020}).
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> # read in the .nc file as a ncdf4 class
-R> raw <- ncdf4::nc_open(here::here("data/era5-pressure.nc"))
-R> 
-R> # convert the variable q and z in the ncdf4 into a cubble
-R> dt <- as_cubble(raw, vars = c("q", "z"))
-\end{CodeInput}
-\end{CodeChunk}
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# read in the .nc file as a ncdf4 class}
+\NormalTok{raw }\OtherTok{\textless{}{-}}\NormalTok{ ncdf4}\SpecialCharTok{::}\FunctionTok{nc\_open}\NormalTok{(here}\SpecialCharTok{::}\FunctionTok{here}\NormalTok{(}\StringTok{"data/era5{-}pressure.nc"}\NormalTok{))}
+
+\CommentTok{\# convert the variable q and z in the ncdf4 into a cubble}
+\NormalTok{dt }\OtherTok{\textless{}{-}} \FunctionTok{as\_cubble}\NormalTok{(raw, }\AttributeTok{vars =} \FunctionTok{c}\NormalTok{(}\StringTok{"q"}\NormalTok{, }\StringTok{"z"}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> # Assume my_ncdf has bounding box of [-180, -90, 180, -90]
-R> # at 0.25 degree resolution and subset it to have
-R> # 1 degree resolution:
-R> dt <- as_cubble(my_ncdf, vars = c("q", "z"),
-+                 long_range = seq(-180, 180, 1),
-+                 lat_range = seq(-90, 90, 1))
-\end{CodeInput}
-\end{CodeChunk}
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# Assume my\_ncdf has bounding box of [{-}180, {-}90, 180, {-}90]}
+\CommentTok{\# at 0.25 degree resolution and subset it to have}
+\CommentTok{\# 1 degree resolution:}
+\NormalTok{dt }\OtherTok{\textless{}{-}} \FunctionTok{as\_cubble}\NormalTok{(my\_ncdf, }\AttributeTok{vars =} \FunctionTok{c}\NormalTok{(}\StringTok{"q"}\NormalTok{, }\StringTok{"z"}\NormalTok{),}
+                \AttributeTok{long\_range =} \FunctionTok{seq}\NormalTok{(}\SpecialCharTok{{-}}\DecValTok{180}\NormalTok{, }\DecValTok{180}\NormalTok{, }\DecValTok{1}\NormalTok{),}
+                \AttributeTok{lat\_range =} \FunctionTok{seq}\NormalTok{(}\SpecialCharTok{{-}}\DecValTok{90}\NormalTok{, }\DecValTok{90}\NormalTok{, }\DecValTok{1}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\caption[An illustration of the original and cluster level cubble in the nested form long form for hierarchical structure data]{An illustration of the original and cluster level cubble in the nested form long form for hierarchical structure data.
+\caption{An illustration of the original and cluster level cubble in the nested form long form for hierarchical structure data.
 \code{switch\_key()} changes the station level cubble into a cluster level cubble and both can be stretched into the long form.}\label{fig:illu-hier}
 \end{figure}
-\end{CodeChunk}
 
@@
 how similar the time series from nearby sites are. This can be seen as a
-matching problem \citep{stuart2010matching, mcintosh2018using} that
-pairs up similar similar time series in nearby locations or a data
+matching problem (\protect\hyperlink{ref-stuart2010matching}{Stuart
+2010}; \protect\hyperlink{ref-mcintosh2018using}{McIntosh et al. 2018})
+that pairs up similar similar time series in nearby locations or a data
 fusing exercise that merges data collected from different sources
-\citep{cocchi2019data}. \code{match_sites()} in \pkg{cubble} provides a
-simple algorithm for this task. The algorithm first matches the two data
-sources spatially through computing the pairwise distance on latitude
-and longitude. Pairs that pass the spatial matching are then matched
-temporally through computing the number of matched peaks within a fixed
-length window. Figure \ref{fig:illu-matching} illustrates this temporal
-matching in more details. Given two series \code{A} and \code{a}, 3
-peaks have been picked in each series. An interval, with default length
-of 5, is constructed for each peak in series \code{A} and the peaks in
-series \code{a} are tested against whether they fall into the any of the
+(\protect\hyperlink{ref-cocchi2019data}{Cocchi 2019}).
+\code{match_sites()} in \pkg{cubble} provides a simple algorithm for
+this task. The algorithm first matches the two data sources spatially
+through computing the pairwise distance on latitude and longitude. Pairs
+that pass the spatial matching are then matched temporally through
+computing the number of matched peaks within a fixed length window.
+Figure \ref{fig:illu-matching} illustrates this temporal matching in
+more details. Given two series \code{A} and \code{a}, 3 peaks have been
+picked in each series. An interval, with default length of 5, is
+constructed for each peak in series \code{A} and the peaks in series
+\code{a} are tested against whether they fall into the any of the
 intervals. In this illustration, there are 2 matches for these two
@@
 
-\begin{CodeChunk}
 \begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{figures/illu-matching}
-
-}
-
-\caption{An illustration of temporal matching in cubble.
+\includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/illu-matching} \caption{An illustration of temporal matching in cubble.
 Three highest peaks are identified in each series and intervals are constructed on series \code{A}.
@@
 Cubble fits in naturally with the interactive graphic pipeline discussed
-in the literature
-\citep{buja1988elements, buja1996interactive, sutherland2000orca, xie2014reactive, cheng2016enabling}.
-Diagram \ref{fig:illu-interactive} illustrates how linking works from
-the map to the time series in cubble. The map and time series plot is
-associated with the nested or long cubble, respectively, and when a user
-action is captured on the map, the site will be activated in the nested
-cubble (left). The nested cubble will communicate to the long cubble to
-activate all the observations with the same \code{id} (middle). The long
-cubble will then highlight the activated series in the time series plot
-(right).
+in the literature (\protect\hyperlink{ref-buja1988elements}{Buja,
+Asimov, and Hurley 1988};
+\protect\hyperlink{ref-buja1996interactive}{Buja, Cook, and Swayne
+1996}; \protect\hyperlink{ref-sutherland2000orca}{Sutherland et al.
+2000}; \protect\hyperlink{ref-xie2014reactive}{Xie, Hofmann, and Cheng
+2014}; \protect\hyperlink{ref-cheng2016enabling}{Cheng, Cook, and
+Hofmann 2016}). Diagram \ref{fig:illu-interactive} illustrates how
+linking works from the map to the time series in cubble. The map and
+time series plot is associated with the nested or long cubble,
+respectively, and when a user action is captured on the map, the site
+will be activated in the nested cubble (left). The nested cubble will
+communicate to the long cubble to activate all the observations with the
+same \code{id} (middle). The long cubble will then highlight the
+activated series in the time series plot (right).
 
@@
 
-\caption[An illustration of the data model under interactive graphics with cubble]{An illustration of the data model under interactive graphics with cubble.
+\caption{An illustration of the data model under interactive graphics with cubble.
 The line plot and the map is made separately with the long and nested cubble.
@@
 Spatio-temporal transformation is a useful technique to extract
-information form spatio-temporal data. Glyph maps \citep{Wickham2012-yr}
-transform the time coordinates into the space coordinates to plot the
-time series of different locations on the map. Calendar plots
-\citep{wang2020calendar} reconstructs time into calendar-based grid to
-discover weekday and weekend pattern. Projection, or linear combination,
-of variables summarises multivariate information into lower dimension to
-further digest. This section elaborates on the glyph map.
+information form spatio-temporal data. Glyph maps
+(\protect\hyperlink{ref-Wickham2012-yr}{Wickham et al. 2012}) transform
+the time coordinates into the space coordinates to plot the time series
+of different locations on the map. Calendar plots
+(\protect\hyperlink{ref-wang2020calendar}{Wang, Cook, and Hyndman
+2020b}) reconstructs time into calendar-based grid to discover weekday
+and weekend pattern. Projection, or linear combination, of variables
+summarises multivariate information into lower dimension to further
+digest. This section elaborates on the glyph map.
 
@@
 calculated position (\code{gx}. \code{gy}. \code{gid}) of each point on
-the time series using linear algebra (Equation 1 and 2 in
-\citet{Wickham2012-yr}). The data can then be piped into \code{ggplot}
-to create the glyph map as:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> library(ggplot2)
-R> gly <- glyphs(data, 
-+               x_major = ..., x_minor = ..., 
-+               y_major = ..., y_minor = ..., ...)
-R> 
-R> ggplot(gly, aes(gx, gy, group = gid)) + 
-+   geom_path() 
-\end{CodeInput}
-\end{CodeChunk}
+the time series using linear algebra (Equation 1 and 2 in Wickham et al.
+(\protect\hyperlink{ref-Wickham2012-yr}{2012})). The data can then be
+piped into \code{ggplot} to create the glyph map as:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{library}\NormalTok{(ggplot2)}
+\NormalTok{gly }\OtherTok{\textless{}{-}} \FunctionTok{glyphs}\NormalTok{(data, }
+              \AttributeTok{x\_major =}\NormalTok{ ..., }\AttributeTok{x\_minor =}\NormalTok{ ..., }
+              \AttributeTok{y\_major =}\NormalTok{ ..., }\AttributeTok{y\_minor =}\NormalTok{ ..., ...)}
+
+\FunctionTok{ggplot}\NormalTok{(gly, }\FunctionTok{aes}\NormalTok{(gx, gy, }\AttributeTok{group =}\NormalTok{ gid)) }\SpecialCharTok{+} 
+  \FunctionTok{geom\_path}\NormalTok{() }
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> ggplot(data = data) +
-+   geom_glyph(aes(x_major = ..., x_minor = ..., 
-+                  y_major = ..., y_minor = ...))
-\end{CodeInput}
-\end{CodeChunk}
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{ggplot}\NormalTok{(}\AttributeTok{data =}\NormalTok{ data) }\SpecialCharTok{+}
+  \FunctionTok{geom\_glyph}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x\_major =}\NormalTok{ ..., }\AttributeTok{x\_minor =}\NormalTok{ ..., }
+                 \AttributeTok{y\_major =}\NormalTok{ ..., }\AttributeTok{y\_minor =}\NormalTok{ ...))}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> tmax <- weatherdata::historical_tmax |>
-+   filter(between(stringr::str_sub(id, 7, 8), 46, 90))
-\end{CodeInput}
-\end{CodeChunk}
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ weatherdata}\SpecialCharTok{::}\NormalTok{historical\_tmax }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
+  \FunctionTok{filter}\NormalTok{(}\FunctionTok{between}\NormalTok{(stringr}\SpecialCharTok{::}\FunctionTok{str\_sub}\NormalTok{(id, }\DecValTok{7}\NormalTok{, }\DecValTok{8}\NormalTok{), }\DecValTok{46}\NormalTok{, }\DecValTok{90}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> tmax <- tmax |> 
-+   face_temporal() |>
-+   filter(lubridate::year(date) %in% c(1971:1975, 2016:2020)) 
-\end{CodeInput}
-\end{CodeChunk}
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|}\ErrorTok{\textgreater{}} 
+  \FunctionTok{face\_temporal}\NormalTok{() }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
+  \FunctionTok{filter}\NormalTok{(lubridate}\SpecialCharTok{::}\FunctionTok{year}\NormalTok{(date) }\SpecialCharTok{\%in\%} \FunctionTok{c}\NormalTok{(}\DecValTok{1971}\SpecialCharTok{:}\DecValTok{1975}\NormalTok{, }\DecValTok{2016}\SpecialCharTok{:}\DecValTok{2020}\NormalTok{)) }
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> tmax <- tmax |>
-+   group_by(month = lubridate::month(date), 
-+          group = as.factor(ifelse(lubridate::year(date) > 2015, 
-+                                   "2016 ~ 2020", "1971 ~ 1975"))) |>
-+   summarise(tmax = mean(tmax, na.rm = TRUE))
-\end{CodeInput}
-\end{CodeChunk}
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
+  \FunctionTok{group\_by}\NormalTok{(}\AttributeTok{month =}\NormalTok{ lubridate}\SpecialCharTok{::}\FunctionTok{month}\NormalTok{(date), }
+         \AttributeTok{group =} \FunctionTok{as.factor}\NormalTok{(}\FunctionTok{ifelse}\NormalTok{(lubridate}\SpecialCharTok{::}\FunctionTok{year}\NormalTok{(date) }\SpecialCharTok{\textgreater{}} \DecValTok{2015}\NormalTok{, }
+                                  \StringTok{"2016 \textasciitilde{} 2020"}\NormalTok{, }\StringTok{"1971 \textasciitilde{} 1975"}\NormalTok{))) }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
+  \FunctionTok{summarise}\NormalTok{(}\AttributeTok{tmax =} \FunctionTok{mean}\NormalTok{(tmax, }\AttributeTok{na.rm =} \ConstantTok{TRUE}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> tmax <- tmax |> face_spatial() |> filter(nrow(ts) == 24) 
-\end{CodeInput}
-\end{CodeChunk}
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{face\_spatial}\NormalTok{() }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{filter}\NormalTok{(}\FunctionTok{nrow}\NormalTok{(ts) }\SpecialCharTok{==} \DecValTok{24}\NormalTok{) }
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> tmax <- tmax |> face_temporal() |> unfold(latitude, longitude)
-\end{CodeInput}
-\end{CodeChunk}
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{face\_temporal}\NormalTok{() }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{unfold}\NormalTok{(latitude, longitude)}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> nsw_vic <- ozmaps::abs_ste |> 
-+   filter(NAME %in% c("Victoria", "New South Wales"))
-R> 
-R> ggplot() + 
-+   geom_sf(data = nsw_vic, 
-+           fill = "transparent", color = "grey", linetype = "dotted") + 
-+   geom_glyph(data = tmax, 
-+              aes(x_major = longitude, x_minor = month, 
-+                  y_major = latitude, y_minor = tmax,
-+                  group = interaction(id, group), color = group),
-+              width = 1, height = 0.5) +
-+   ...
-\end{CodeInput}
-\end{CodeChunk}
-
-\begin{CodeChunk}
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{nsw\_vic }\OtherTok{\textless{}{-}}\NormalTok{ ozmaps}\SpecialCharTok{::}\NormalTok{abs\_ste }\SpecialCharTok{|}\ErrorTok{\textgreater{}} 
+  \FunctionTok{filter}\NormalTok{(NAME }\SpecialCharTok{\%in\%} \FunctionTok{c}\NormalTok{(}\StringTok{"Victoria"}\NormalTok{, }\StringTok{"New South Wales"}\NormalTok{))}
+
+\FunctionTok{ggplot}\NormalTok{() }\SpecialCharTok{+} 
+  \FunctionTok{geom\_sf}\NormalTok{(}\AttributeTok{data =}\NormalTok{ nsw\_vic, }
+          \AttributeTok{fill =} \StringTok{"transparent"}\NormalTok{, }\AttributeTok{color =} \StringTok{"grey"}\NormalTok{, }\AttributeTok{linetype =} \StringTok{"dotted"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{geom\_glyph}\NormalTok{(}\AttributeTok{data =}\NormalTok{ tmax, }
+             \FunctionTok{aes}\NormalTok{(}\AttributeTok{x\_major =}\NormalTok{ longitude, }\AttributeTok{x\_minor =}\NormalTok{ month, }
+                 \AttributeTok{y\_major =}\NormalTok{ latitude, }\AttributeTok{y\_minor =}\NormalTok{ tmax,}
+                 \AttributeTok{group =} \FunctionTok{interaction}\NormalTok{(id, group), }\AttributeTok{color =}\NormalTok{ group),}
+             \AttributeTok{width =} \DecValTok{1}\NormalTok{, }\AttributeTok{height =} \FloatTok{0.5}\NormalTok{) }\SpecialCharTok{+}
+\NormalTok{  ...}
+\end{Highlighting}
+\end{Shaded}
+
 \begin{figure}
-
-{\centering \includegraphics[width=1\linewidth,height=0.7\textheight]{figures/basic-manip}
-
-}
-
-\caption[A glyph map of the average maximum temperature by month of Victoria and New South Wales weather stations in Australia]{A glyph map of the average maximum temperature by month of Victoria and New South Wales weather stations in Australia.
+\includegraphics[width=1\linewidth,height=0.7\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/basic-manip} \caption{A glyph map of the average maximum temperature by month of Victoria and New South Wales weather stations in Australia.
 On the top left corner is an insetted plot of station Cobar highlighted in the black box.}\label{fig:basic-manip}
 \end{figure}
-\end{CodeChunk}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> station_nested <- weatherdata::climate_full |> 
-+   mutate(cluster = ...)
-\end{CodeInput}
-\end{CodeChunk}
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{station\_nested }\OtherTok{\textless{}{-}}\NormalTok{ weatherdata}\SpecialCharTok{::}\NormalTok{climate\_full }\SpecialCharTok{|}\ErrorTok{\textgreater{}} 
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{cluster =}\NormalTok{ ...)}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> cluster_nested <- station_nested |> switch_key(cluster) 
-\end{CodeInput}
-\end{CodeChunk}
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{cluster\_nested }\OtherTok{\textless{}{-}}\NormalTok{ station\_nested }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{switch\_key}\NormalTok{(cluster) }
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> cluster_nested <- cluster_nested |> get_centroid()
-\end{CodeInput}
-\end{CodeChunk}
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{cluster\_nested }\OtherTok{\textless{}{-}}\NormalTok{ cluster\_nested }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{get\_centroid}\NormalTok{()}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\begin{CodeChunk}
 \begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{figures/basic-agg}
-
-}
-
-\caption[Profile of aggregated precipitation from 639 weather stations in Australia]{Profile of aggregated precipitation from 639 weather stations in Australia.
+\includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/basic-agg} \caption{Profile of aggregated precipitation from 639 weather stations in Australia.
 Subplot A shows the glyph map of the weekly averaged precipitation of each cluster.
@@
 
-\begin{CodeChunk}
 \begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{figures/matching-map}
-
-}
-
-\caption[Location of weather stations and river gauges in Victoria, Australia]{Location of weather stations and river gauges in Victoria, Australia.}\label{fig:matching-map}
+\includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/matching-map} \caption{Location of weather stations and river gauges in Victoria, Australia.}\label{fig:matching-map}
 \end{figure}
-\end{CodeChunk}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> res <- match_sites(
-+   river, climate,
-+   temporal_by = c("Water_course_level" = "prcp"),
-+   temporal_independent = "prcp",  
-+   temporal_n_highest = 30,
-+   temporal_min_match = 15
-+ )
-\end{CodeInput}
-\end{CodeChunk}
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{res }\OtherTok{\textless{}{-}} \FunctionTok{match\_sites}\NormalTok{(}
+\NormalTok{  river, climate,}
+  \AttributeTok{temporal\_by =} \FunctionTok{c}\NormalTok{(}\StringTok{"Water\_course\_level"} \OtherTok{=} \StringTok{"prcp"}\NormalTok{),}
+  \AttributeTok{temporal\_independent =} \StringTok{"prcp"}\NormalTok{,  }
+  \AttributeTok{temporal\_n\_highest =} \DecValTok{30}\NormalTok{,}
+  \AttributeTok{temporal\_min\_match =} \DecValTok{15}
+\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeOutput}
-# cubble:   id [8]: nested form
-# bbox:     [144.52, -37.73, 146.06, -36.55]
-# temporal: date [date], matched_var [dbl]
-  id          name                  lat  long type   dist group ts       n_match
-  <chr>       <chr>               <dbl> <dbl> <chr> <dbl> <int> <list>     <int>
-1 405234      SEVEN CREEKS @ D/S~ -36.9  146. river  6.15     5 <tibble>      21
-2 404207      HOLLAND CREEK @ KE~ -36.6  146. river  8.54    10 <tibble>      21
-3 ASN00082042 strathbogie         -36.8  146. clim~  6.15     5 <tibble>      21
-4 ASN00082170 benalla airport     -36.6  146. clim~  8.54    10 <tibble>      21
-5 230200      MARIBYRNONG RIVER ~ -37.7  145. river  6.17     6 <tibble>      19
-6 ASN00086038 essendon airport    -37.7  145. clim~  6.17     6 <tibble>      19
-7 406213      CAMPASPE RIVER @ R~ -37.0  145. river  1.84     1 <tibble>      18
-8 ASN00088051 redesdale           -37.0  145. clim~  1.84     1 <tibble>      18
-\end{CodeOutput}
-\end{CodeChunk}
+\begin{verbatim}
+## # cubble:   id [8]: nested form
+## # bbox:     [144.52, -37.73, 146.06, -36.55]
+## # temporal: date [date], matched_var [dbl]
+##   id          name                  lat  long type   dist group ts       n_match
+##   <chr>       <chr>               <dbl> <dbl> <chr> <dbl> <int> <list>     <int>
+## 1 405234      SEVEN CREEKS @ D/S~ -36.9  146. river  6.15     5 <tibble>      21
+## 2 404207      HOLLAND CREEK @ KE~ -36.6  146. river  8.54    10 <tibble>      21
+## 3 ASN00082042 strathbogie         -36.8  146. clim~  6.15     5 <tibble>      21
+## 4 ASN00082170 benalla airport     -36.6  146. clim~  8.54    10 <tibble>      21
+## 5 230200      MARIBYRNONG RIVER ~ -37.7  145. river  6.17     6 <tibble>      19
+## 6 ASN00086038 essendon airport    -37.7  145. clim~  6.17     6 <tibble>      19
+## 7 406213      CAMPASPE RIVER @ R~ -37.0  145. river  1.84     1 <tibble>      18
+## 8 ASN00088051 redesdale           -37.0  145. clim~  1.84     1 <tibble>      18
+\end{verbatim}
 
@@
 
-\begin{CodeChunk}
 \begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{figures/matching}
-
-}
-
-\caption[Matched weather stations and river gauges on the map (A) and across time (B)]{Matched weather stations and river gauges on the map (A) and across time (B).
+\includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/matching} \caption{Matched weather stations and river gauges on the map (A) and across time (B).
 Precipitation and water level have been standardised between 0 and 1 to be displayed on the same scale.
@@
 
-ERA5 data \citep{hersbach2020era5} is the latest reanalysis of global
-atmosphere, land surface, and ocean waves from 1950 onwards and is
-available in the NetCDF format from European Centre for Medium-Range
-Weather Forecasts (ECMWF). The data can be directly downloaded from
+ERA5 data (\protect\hyperlink{ref-hersbach2020era5}{Hersbach et al.
+2020}) is the latest reanalysis of global atmosphere, land surface, and
+ocean waves from 1950 onwards and is available in the NetCDF format from
+European Centre for Medium-Range Weather Forecasts (ECMWF). The data can
+be directly downloaded from
 \href{https://cds.climate.copernicus.eu/cdsapp\#!/dataset/reanalysis-era5-pressure-levels?tab=overview}{Copernicus
 Climate Data Store (CDS)} website or programmatically via an R package
-\pkg{ecmwfr} \citep{ecwmfr}. The \code{era5-pressure} data contains
-variable \emph{specific humidity} and \emph{geopotential} on the 10 hPa
-pressure level on four dates: 2002-09-22, 2002-09-26, 2002-09-30, and
-2002-10-04. Once downloaded, the data can be read into a cubble as:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> raw <- ncdf4::nc_open(here::here("data/era5-pressure.nc"))
-R> dt <- as_cubble(raw, vars = c("q", "z"))
-\end{CodeInput}
-\end{CodeChunk}
+\pkg{ecmwfr} (\protect\hyperlink{ref-ecwmfr}{Hufkens, Stauffer, and
+Campitelli 2019}). The \code{era5-pressure} data contains variable
+\emph{specific humidity} and \emph{geopotential} on the 10 hPa pressure
+level on four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04.
+Once downloaded, the data can be read into a cubble as:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{raw }\OtherTok{\textless{}{-}}\NormalTok{ ncdf4}\SpecialCharTok{::}\FunctionTok{nc\_open}\NormalTok{(here}\SpecialCharTok{::}\FunctionTok{here}\NormalTok{(}\StringTok{"data/era5{-}pressure.nc"}\NormalTok{))}
+\NormalTok{dt }\OtherTok{\textless{}{-}} \FunctionTok{as\_cubble}\NormalTok{(raw, }\AttributeTok{vars =} \FunctionTok{c}\NormalTok{(}\StringTok{"q"}\NormalTok{, }\StringTok{"z"}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
 
 Figure \ref{fig:netcdf} reproduces the ERA5 data row of Figure 19 in
-\citet{hersbach2020era5}. It shows the southern polar vortex splits into
-two on 2002-09-26 and further splits into four on 2002-10-04 in the
-stratosphere. Readers interested in the analysis of this figure can
-refer to \citet{hersbach2020era5}, \citet{simmons2020global} and
-\citet{simmons2005ecmwf} for more details.
-
-\begin{CodeChunk}
+Hersbach et al. (\protect\hyperlink{ref-hersbach2020era5}{2020}). It
+shows the southern polar vortex splits into two on 2002-09-26 and
+further splits into four on 2002-10-04 in the stratosphere. Readers
+interested in the analysis of this figure can refer to Hersbach et al.
+(\protect\hyperlink{ref-hersbach2020era5}{2020}), Simmons et al.
+(\protect\hyperlink{ref-simmons2020global}{2020}) and Simmons et al.
+(\protect\hyperlink{ref-simmons2005ecmwf}{2005}) for more details.
+
 \begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/netcdf}
-
-}
-
-\caption[A reproduction of the second row (ERA5 data) of Figure 19 in  Hersbach et al (2020)]{A reproduction of the second row (ERA5 data) of Figure 19 in  Hersbach et al (2020).}\label{fig:netcdf}
+\includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/netcdf} \caption{A reproduction of the second row (ERA5 data) of Figure 19 in  Hersbach et al (2020).}\label{fig:netcdf}
 \end{figure}
-\end{CodeChunk}
 
@@
 \CommentTok{\# data pre{-}processing}
-\NormalTok{clean }\OtherTok{\textless{}{-}}\NormalTok{ weatherdata}\SpecialCharTok{::}\NormalTok{climate\_full }\SpecialCharTok{|\textgreater{}}\NormalTok{ ...}
+\NormalTok{clean }\OtherTok{\textless{}{-}}\NormalTok{ weatherdata}\SpecialCharTok{::}\NormalTok{climate\_full }\SpecialCharTok{|}\ErrorTok{\textgreater{}}\NormalTok{ ...}
 
 \CommentTok{\# created SharedData instance for crosstalk}
-\NormalTok{nested }\OtherTok{\textless{}{-}}\NormalTok{ clean }\SpecialCharTok{|\textgreater{}}\NormalTok{ SharedData}\SpecialCharTok{$}\FunctionTok{new}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{id, }\AttributeTok{group =} \StringTok{"cubble"}\NormalTok{)}
-\NormalTok{long }\OtherTok{\textless{}{-}} \FunctionTok{face\_temporal}\NormalTok{(clean) }\SpecialCharTok{|\textgreater{}}\NormalTok{ SharedData}\SpecialCharTok{$}\FunctionTok{new}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{id, }\AttributeTok{group =} \StringTok{"cubble"}\NormalTok{)}
+\NormalTok{nested }\OtherTok{\textless{}{-}}\NormalTok{ clean }\SpecialCharTok{|}\ErrorTok{\textgreater{}}\NormalTok{ SharedData}\SpecialCharTok{$}\FunctionTok{new}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{id, }\AttributeTok{group =} \StringTok{"cubble"}\NormalTok{)}
+\NormalTok{long }\OtherTok{\textless{}{-}} \FunctionTok{face\_temporal}\NormalTok{(clean) }\SpecialCharTok{|}\ErrorTok{\textgreater{}}\NormalTok{ SharedData}\SpecialCharTok{$}\FunctionTok{new}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{id, }\AttributeTok{group =} \StringTok{"cubble"}\NormalTok{)}
 
 \CommentTok{\# create the spatial and temporal view each with a ShareData instance}
-\NormalTok{p1 }\OtherTok{\textless{}{-}}\NormalTok{ nested }\SpecialCharTok{|\textgreater{}}\NormalTok{ ...}
-\NormalTok{p2 }\OtherTok{\textless{}{-}}\NormalTok{ long }\SpecialCharTok{|\textgreater{}}\NormalTok{ ...}
+\NormalTok{p1 }\OtherTok{\textless{}{-}}\NormalTok{ nested }\SpecialCharTok{|}\ErrorTok{\textgreater{}}\NormalTok{ ...}
+\NormalTok{p2 }\OtherTok{\textless{}{-}}\NormalTok{ long }\SpecialCharTok{|}\ErrorTok{\textgreater{}}\NormalTok{ ...}
 
@@
 
-\begin{CodeChunk}
 \begin{figure}
-
-{\centering \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/linking} \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/linking-north} \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/linking-lower}
-
-}
-
-\caption[Exploring temperature variation using linking of a map and seasonal display]{Exploring temperature variation using linking of a map and seasonal display.
+\includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/linking} \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/linking-north} \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/linking-lower} \caption{Exploring temperature variation using linking of a map and seasonal display.
 Each row is a screen dump of the process.
@@
 \CommentTok{\# data pre{-}processing}
-\NormalTok{clean }\OtherTok{\textless{}{-}}\NormalTok{ weatherdata}\SpecialCharTok{::}\NormalTok{climate\_full }\SpecialCharTok{|\textgreater{}}\NormalTok{ ...}
+\NormalTok{clean }\OtherTok{\textless{}{-}}\NormalTok{ weatherdata}\SpecialCharTok{::}\NormalTok{climate\_full }\SpecialCharTok{|}\ErrorTok{\textgreater{}}\NormalTok{ ...}
 
@@
 \NormalTok{p }\OtherTok{\textless{}{-}} \FunctionTok{map}\NormalTok{(}\DecValTok{1}\SpecialCharTok{:}\FunctionTok{length}\NormalTok{(df\_id), }\ControlFlowTok{function}\NormalTok{(i)\{}
-\NormalTok{  dt }\OtherTok{\textless{}{-}}\NormalTok{ clean }\SpecialCharTok{|\textgreater{}} \FunctionTok{filter}\NormalTok{(id }\SpecialCharTok{==}\NormalTok{ df\_id[i])}
-  \FunctionTok{ggplot}\NormalTok{(dt) }\SpecialCharTok{|\textgreater{}}\NormalTok{ ...}
+\NormalTok{  dt }\OtherTok{\textless{}{-}}\NormalTok{ clean }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{filter}\NormalTok{(id }\SpecialCharTok{==}\NormalTok{ df\_id[i])}
+  \FunctionTok{ggplot}\NormalTok{(dt) }\SpecialCharTok{|}\ErrorTok{\textgreater{}}\NormalTok{ ...}
 \NormalTok{\})}
@@
 \NormalTok{nested }\OtherTok{\textless{}{-}} \FunctionTok{face\_spatial}\NormalTok{(clean)}
-\FunctionTok{leaflet}\NormalTok{(nested) }\SpecialCharTok{|\textgreater{}}
-  \FunctionTok{addTiles}\NormalTok{() }\SpecialCharTok{|\textgreater{}}
-  \FunctionTok{addCircleMarkers}\NormalTok{(}\AttributeTok{group =} \StringTok{"a"}\NormalTok{, ...) }\SpecialCharTok{|\textgreater{}}
+\FunctionTok{leaflet}\NormalTok{(nested) }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
+  \FunctionTok{addTiles}\NormalTok{() }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
+  \FunctionTok{addCircleMarkers}\NormalTok{(}\AttributeTok{group =} \StringTok{"a"}\NormalTok{, ...) }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
 \NormalTok{  leafpop}\SpecialCharTok{::}\FunctionTok{addPopupGraphs}\NormalTok{(}\AttributeTok{graph =}\NormalTok{ p, ...)}
@@
 \ref{fig:interactive-linking} made made with leaflet and popups
-\citep{leafpop}.
-
-\begin{CodeChunk}
+(\protect\hyperlink{ref-leafpop}{Appelhans and Detsch 2021}).
+
 \begin{figure}
-
-{\centering \includegraphics[width=0.45\linewidth,height=0.25\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/popup-mid}
-
-}
-
-\caption[Same as Figure 11 with thetemperature variation shown as a popup in the leaflet map]{Same as Figure 11 with thetemperature variation shown as a popup in the leaflet map.}\label{fig:interactive-popup}
+\includegraphics[width=0.45\linewidth,height=0.25\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/popup-mid} \caption{Same as Figure 11 with thetemperature variation shown as a popup in the leaflet map.}\label{fig:interactive-popup}
 \end{figure}
-\end{CodeChunk}
 
@@
 Nicolas Langrené was affiliated with CSIRO's Data61. The article is
-created using \pkg{knitr} \citep{knitr} and \pkg{rmarkdown}
-\citep{rmarkdown} in R. The source code for reproducing this paper can
-be found at: \url{https://github.com/huizezhang-sherry/paper-cubble}.
+created using \pkg{knitr} (\protect\hyperlink{ref-knitr}{Xie 2015}) and
+\pkg{rmarkdown} (\protect\hyperlink{ref-rmarkdown}{Xie, Allaire, and
+Grolemund 2018}) in R. The source code for reproducing this paper can be
+found at: \url{https://github.com/huizezhang-sherry/paper-cubble}.
 
@@
 
-\caption[An illustration of the data model under interactive graphics with cubble]{An illustration of the data model under interactive graphics with cubble.
+\caption{An illustration of the data model under interactive graphics with cubble.
 When a point on the time series is selected, the corresponding row in the long cubble will be activated.
@@
 
-\bibliography{references.bib}
+\hypertarget{refs}{}
+\begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-leafpop}{}}%
+Appelhans, Tim, and Florian Detsch. 2021. \emph{\pkg{leafpop}: Include
+Tables, Images and Graphs in Leaflet Pop-Ups}.
+\url{https://CRAN.R-project.org/package=leafpop}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-bach_review_2014}{}}%
+Bach, Benjamin, Pierre Dragicevic, Dragicevic Archambault, Christophe
+Hurter, and Sheelagh Carpendale. 2014. {``A {Review} of {Temporal}
+{Data} {Visualizations} {Based} on {Space}-{Time} {Cube}
+{Operations}.''} \emph{Eurographics Conference on Visualization}, 19.
+\url{https://hal.inria.fr/hal-01006140/}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-buja1988elements}{}}%
+Buja, Andreas, Daniel Asimov, and Catherine Hurley. 1988. {``Elements of
+a Viewing Pipeline.''} \emph{Dynamic Graphics Statistics}, 277.
+
+\leavevmode\vadjust pre{\hypertarget{ref-buja1996interactive}{}}%
+Buja, Andreas, Dianne Cook, and Deborah F Swayne. 1996. {``Interactive
+High-Dimensional Data Visualization.''} \emph{Journal of Computational
+and Graphical Statistics} 5 (1): 78--99.
+\url{https://doi.org/10.2307/1390754}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-cheng2016enabling}{}}%
+Cheng, Xiaoyue, Dianne Cook, and Heike Hofmann. 2016. {``Enabling
+Interactivity on Displays of Multivariate Time Series and Longitudinal
+Data.''} \emph{Journal of Computational and Graphical Statistics} 25
+(4): 1057--76. \url{https://doi.org/10.1080/10618600.2015.1105749}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-cocchi2019data}{}}%
+Cocchi, Marina. 2019. \emph{Data Fusion Methodology and Applications}.
+Elsevier.
+
+\leavevmode\vadjust pre{\hypertarget{ref-hersbach2020era5}{}}%
+Hersbach, Hans, Bill Bell, Paul Berrisford, Shoji Hirahara, András
+Horányi, Joaquín Muñoz-Sabater, Julien Nicolas, et al. 2020. {``The Era5
+Global Reanalysis.''} \emph{Quarterly Journal of the Royal
+Meteorological Society} 146 (730): 1999--2049.
+
+\leavevmode\vadjust pre{\hypertarget{ref-ecwmfr}{}}%
+Hufkens, Koen, Reto Stauffer, and Elio Campitelli. 2019. {``The
+\pkg{ecwmfr} Package: An Interface to ECMWF API Endpoints.''}
+\url{https://doi.org/10.5281/zenodo.2647541}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-jr_herring_opengis_2011}{}}%
+J.R. Herring. 2011. {``{OpenGIS}® {Implementation} {Standard} for
+{Geographic} Information - {Simple} Feature Access - {Part} 1: {Common}
+Architecture.''} Open Geospatial Consortium Inc.
+\href{http://portal.\%20opengeospatial.org/files/?artifact_id=25355}{http://portal.
+opengeospatial.org/files/?artifact\_id=25355}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-lu_multidimensional_2018}{}}%
+Lu, Meng, Marius Appel, and Edzer Pebesma. 2018. {``Multidimensional
+{Arrays} for {Analysing} {Geoscientific} {Data}.''} \emph{ISPRS
+International Journal of Geo-Information} 7 (8): 313.
+\url{https://doi.org/10.3390/ijgi7080313}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-mcintosh2018using}{}}%
+McIntosh, Avery I, Helen E Jenkins, Laura F White, Marinus Barnard, Dana
+R Thomson, Tania Dolby, John Simpson, et al. 2018. {``Using Routinely
+Collected Laboratory Data to Identify High Rifampicin-Resistant
+Tuberculosis Burden Communities in the Western Cape Province, South
+Africa: A Retrospective Spatiotemporal Analysis.''} \emph{PLoS Medicine}
+15 (8): e1002638.
+
+\leavevmode\vadjust pre{\hypertarget{ref-michna2013rnetcdf}{}}%
+Michna, Pavel, and Milton Woods. 2013. {``\pkg{RNetCDF}: A Package for
+Reading and Writing NetCDF Datasets.''} \emph{The R Journal} 5 (2):
+29--36.
+
+\leavevmode\vadjust pre{\hypertarget{ref-rnetcdf}{}}%
+---------. 2021. \emph{\pkg{RNetCDF}: Interface to 'NetCDF' Datasets}.
+\url{https://CRAN.R-project.org/package=RNetCDF}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-tibble}{}}%
+Müller, Kirill, and Hadley Wickham. 2021. \emph{Tibble: Simple Data
+Frames}. \url{https://CRAN.R-project.org/package=tibble}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-spacetime}{}}%
+Pebesma, Edzer. 2012. {``\pkg{spacetime}: Spatio-Temporal Data in r.''}
+\emph{Journal of Statistical Software} 51 (7): 1--30.
+\url{https://doi.org/10.18637/jss.v051.i07}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-stars}{}}%
+---------. 2021. \emph{\pkg{stars}: Spatiotemporal Arrays, Raster and
+Vector Data Cubes}. \url{https://CRAN.R-project.org/package=stars}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-sf}{}}%
+Pebesma, Edzer J. 2018. {``Simple Features for \proglang{R}:
+Standardized Support for Spatial Vector Data.''} \emph{R Journal} 10
+(1): 439.
+
+\leavevmode\vadjust pre{\hypertarget{ref-sp}{}}%
+Pebesma, Edzer, and Roger S Bivand. 2005. {``S Classes and Methods for
+Spatial Data: The \pkg{sp} Package.''} \emph{R News} 5 (2): 9--13.
+
+\leavevmode\vadjust pre{\hypertarget{ref-ncdf4}{}}%
+Pierce, David. 2019. \emph{\pkg{ncdf4}: Interface to Unidata netCDF
+(Version 4 or Earlier) Format Data Files}.
+\url{https://CRAN.R-project.org/package=ncdf4}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-xts}{}}%
+Ryan, Jeffrey A., and Joshua M. Ulrich. 2020. \emph{\pkg{xts}:
+eXtensible Time Series}. \url{https://CRAN.R-project.org/package=xts}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-simmons2005ecmwf}{}}%
+Simmons, Adrian, Mariano Hortal, Graeme Kelly, Anthony McNally, Agathe
+Untch, and Sakari Uppala. 2005. {``ECMWF Analyses and Forecasts of
+Stratospheric Winter Polar Vortex Breakup: September 2002 in the
+Southern Hemisphere and Related Events.''} \emph{Journal of the
+Atmospheric Sciences} 62 (3): 668--89.
+\url{https://doi.org/10.1175/JAS-3322.1}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-simmons2020global}{}}%
+Simmons, Adrian, Cornel Soci, Julien Nicolas, Bill Bell, P. Berrisford,
+Rossana Dragani, Johannes Flemming, et al. 2020. {``Global Stratospheric
+Temperature Bias and Other Stratospheric Aspects of Era5 and Era5.1,''}
+no. 859 (January). \url{https://doi.org/10.21957/rcxqfmg0}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-stuart2010matching}{}}%
+Stuart, Elizabeth A. 2010. {``Matching Methods for Causal Inference: A
+Review and a Look Forward.''} \emph{Statistical Science} 25 (1): 1.
+
+\leavevmode\vadjust pre{\hypertarget{ref-tidync}{}}%
+Sumner, Michael. 2020. \emph{\pkg{tidync}: A Tidy Approach to 'NetCDF'
+Data Exploration and Extraction}.
+\url{https://CRAN.R-project.org/package=tidync}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-sutherland2000orca}{}}%
+Sutherland, Peter, Anthony Rossini, Thomas Lumley, Nicholas Lewin-Koh,
+Julie Dickerson, Zach Cox, and Dianne Cook. 2000. {``\pkg{Orca}: A
+Visualization Toolkit for High-Dimensional Data.''} \emph{Journal of
+Computational and Graphical Statistics} 9 (3): 509--29.
+\url{https://www.tandfonline.com/doi/abs/10.1080/10618600.2000.10474896}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-tsibble}{}}%
+Wang, Earo, Dianne Cook, and Rob J Hyndman. 2020a. {``A New Tidy Data
+Structure to Support Exploration and Modeling of Temporal Data.''}
+\emph{Journal of Computational and Graphical Statistics} 29 (3):
+466--78. \url{https://doi.org/10.1080/10618600.2019.1695624}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-wang2020calendar}{}}%
+---------. 2020b. {``Calendar-Based Graphics for Visualizing People's
+Daily Schedules.''} \emph{Journal of Computational and Graphical
+Statistics} 29 (3): 490--502.
+
+\leavevmode\vadjust pre{\hypertarget{ref-tidydata}{}}%
+Wickham, Hadley. 2014. {``Tidy Data.''} \emph{Journal of Statistical
+Software} 59 (10): 1--23. \url{https://doi.org/10.18637/jss.v059.i10}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-tidyverse}{}}%
+Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy
+D'Agostino McGowan, Romain François, Garrett Grolemund, et al. 2019.
+{``Welcome to the {tidyverse}.''} \emph{Journal of Open Source Software}
+4 (43): 1686. \url{https://doi.org/10.21105/joss.01686}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Wickham2012-yr}{}}%
+Wickham, Hadley, Heike Hofmann, Charlotte Wickham, and Dianne Cook.
+2012. {``Glyph-Maps for Visually Exploring Temporal Patterns in Climate
+Data and Models.''} \emph{Environmetrics} 23 (5): 382--93.
+
+\leavevmode\vadjust pre{\hypertarget{ref-knitr}{}}%
+Xie, Yihui. 2015. \emph{Dynamic Documents with \proglang{R} and
+\pkg{knitr}}. 2nd ed. Boca Raton, Florida: Chapman; Hall/CRC.
+\url{https://yihui.name/knitr/}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-rmarkdown}{}}%
+Xie, Yihui, J. J. Allaire, and Garrett Grolemund. 2018. \emph{R
+Markdown: The Definitive Guide}. Boca Raton, Florida: Chapman; Hall/CRC.
+\url{https://bookdown.org/yihui/rmarkdown}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-xie2014reactive}{}}%
+Xie, Yihui, Heike Hofmann, and Xiaoyue Cheng. 2014. {``{Reactive
+Programming for Interactive Graphics}.''} \emph{Statistical Science} 29
+(2): 201--13. \url{https://doi.org/10.1214/14-STS477}.
+
+\end{CSLReferences}
+
+\bibliographystyle{unsrt}
+\bibliography{../references.bib}
 

### 2022-04-17 jss/cubble-jss.Rmd
notes on package install and rendering

--- 2022-03-01 main.Rmd
@@
 library(cubble)
-library(forcats)
-library(stringr)
 library(dplyr)
-library(lubridate)
-library(ggplot2)
-library(GGally)
-library(plotly)
-library(crosstalk)
-library(patchwork)
 ```

### 2022-04-17 jss/cubble-jss.tex
notes on package install and rendering

--- 2022-04-17 main.tex
@@
 
-{\centering \includegraphics[width=1\linewidth]{figures/illu-matching}
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/illu-matching}
 
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.7\textheight]{figures/basic-manip}
+{\centering \includegraphics[width=1\linewidth,height=0.7\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/basic-manip}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{figures/basic-agg}
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/basic-agg}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{figures/matching-map}
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/matching-map}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{figures/matching}
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/matching}
 

### 2022-04-18 1-intro.Rmd
a new skeleton of intro incorporating various past writing

@@
 
-Spatio-temporal data involves variables observed in the space across time and this paper focuses on spatio-temporal vector data with fixed locations and regular time intervals.
-This type of data can naturally be considered as a data cube with three axes: spatial identifier, temporal identifier, and variables.
+Spatio-temporal data has a spatial component referring to the location of each observation, a temporal component that is recorded at regular or irregular time intervals, and multiple variables measured at each spatial and temporal value.
+In data collection, spatio-temporal data is less often collected as a whole piece but fragmented spatial and temporal data for analysts to ensemble into spatio-temporal data.
+With the [integrated] data, we hope to fix the time to explore the spatial distribution, or fix the spaital location to explore temporal trends.
+Ideally, we would be able to carry out each of these tasks easily, and even accommodate complex explorations encompassing both simultaneously.
+These are all underpinned by a [flexible] data structure that can represent spatio-temporal data.
+
+The *Representing data* section of the CRAN task view [ref] lists existing spatio-temporal data representation.
+Among them, @spacetime proposes four layouts: full grid layout, sparse grid layout, irregular layout, and trajectory.
+These layouts are implemented in the \pkg{spacetime} package with underlying class \pkg{sp} [@sp] and \pkg{xts} [@xts] for handling spatial and temporal variables.
+The \pkg{stars} package has been a recent implementation to handle spatio-temporal data in an array.
+The package depends on the \pkg{sf} package and has allowed more access to tidyverse verbs in the workflow.
+
+What remains to be difficult when wrangling spatio-temporal data is that the raw data is not naturally structured in any one of the representation mentioned above and the first difficulty is to cast the data, potentially from multiple sources into one of the data representation.
+Some examples of these includes map data recorded the shape of a collection of area of interest is provided by the government website; geo-scientific data recorded with longitude and latitude coordinate of locations in the area; and temporal variables indexed by both location and time for the temporal aspect.
+For spatio-temporal data analysis, merging all these sources can be challenging and joining these tables can go wrong because of the slightest unmatched of the linking variable from different sources.
+
+Another issue is how the concept of tidy data should be applied to spatio-temporal data.
+The tidy data concept advocates the data to be shaped into 1) one row per observation; 2) one column per variable; and 3) one type of data per table.
+The long form data is preferred over wide data given the downstream data cleaning and visualisation software i.e.
+\pkg{dplyr} and \pkg{ggplot2}.
+However, the long form can be inefficient to store feature geometries, especially for large multipolygon over long time period, i.e.
+sub-daily time series data.
+
+This paper presents a new \proglang{R} package, \pkg{cubble}, to [overcome the problem] with organising variables in different levels when working with spatio-temporal data.
+The data structure focuses on connecting to the upstream data collection and is suitable for temporal data collected at the fixed location over time.
+Among the four spacetime layouts in @spacetime, it can be full grid layout, sparse grid layout, or irregular layout, but trajectory is not within the scope of this work.
+In the package a new data structure, also called \code{cubble}, is proposed to organise spatial and temporal variables as two forms of a single data object so that they can be wrangled separately while synchronised as a whole piece.
+With this new data object, analysts can spend less time writing codes to organise spatial and temporal variables and focus more on the exploratory data analysis itself.
+The software is available from the Comprehensive R Archive Network (CRAN) at [CRAN link].
+
+The rest of the paper is divided as follows: Section \@ref(cube) discusses the spatio-temporal cube to conceptualise spatio-temporal data.
+Section \@ref(cubble) presents the main design and functionality of \pkg{cubble}.
+Section \@ref(others) explains how cubble deals with more advanced considerations including data with hierarchical structure, data matching, how cubble fits with existing static and interactive visualisation tools, and spatio-temporal data transformation.
+Section \@ref(examples) uses Australia weather station data and river level data as examples to demonstrate the use of \pkg{cubble}.
+An example on how \pkg{cubble} handles NetCDF data is also provided.
+Section \@ref(conclude) concludes the paper.
+
+<!-- there is a strong need for such tools given the ubiquitous spatial and temporal data collected in the society, i.e., climate observations from weather stations, air quality variables from air monitoring stations.
+The demand for such tools can also come from spatial data analysts wishing to incorporate temporal data in their analysis or vice versa.
+-->
+
+<!-- This type of data can naturally be considered as a data cube with three axes: spatial identifier, temporal identifier, and variables.
 In this cubic formation, space only occupies one axis, rather than two, to avoid hypercubes for multivariate spatio-temporal data.
@@
 climatology, meteorology, and oceanography, among others, to organise multivariate data in a spatio-temporal grid.
+-->
 
 ```{r illu-input, echo = FALSE, out.width = "100%", fig.cap="An illustration of different spatio-temporal data layouts: (1) a single table for both spatial and temporal variables; (2) separate tables for spatial and temporal variables; and (3) an array data structure that populates variables into a gridded longitude-latitude-time (hyper)cube.", fig.align='center'}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.001.png"))
+#knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.001.png"))
 ```
@@
 
-While incorporating both spatial and temporal information enriches the data in the analysis, spatio-temporal data can sometimes be hard to work with given its two observational units in the variables.
+<!-- While incorporating both spatial and temporal information enriches the data in the analysis, spatio-temporal data can sometimes be hard to work with given its two observational units in the variables.
 The single table format in Figure \ref{fig:illu-input} is convenient for spatio-temporal operations that involve both types of variables, however, it makes duplicates of spatial variables and is not ideal when only looking at the spatial aspect of the data.
@@
 This issue implies analysts need to constantly think about the data format and arrange variables into the proper shape before the actual operations, which can be a laborious step when switching between the two formats back and forth several times during the analysis.
+-->
 
@@
 
-In the \proglang{R} community, many data objects have been proposed to work with specific spatio-temporal tasks, while few can freely explore the data.
+<!-- The package \pkg{sf} [@sf] uses data frame objects to represent spatial data with an additional \code{geometry} column dedicating to feature geometries [@jr_herring_opengis_2011] (e.g.
+POINT, MULTIPOINT, POLYGON, MULTIPOLYGON etc).
+-->
+
+<!-- For temporal data, \pkg{tsibble} [@tsibble] builds on top of the \pkg{tibble} [@tibble] object and explicitly includes date and time in columns as variables, rather than data attributes.  -->
+
+<!-- Underlying tidyverse is the concept of tidy data [@tidydata], which considers how variables should be structured for analysis.
+In the three formats (time-wide, space-wide, and long) summarised in the @spacetime paper, only the long form is considered as tidy.
+The main advantages of it over the time-wide and space-wide format is that the latter treat either time or location as variable names (Section 3.1 in @tidydata) rather than values of variables, making them difficult to be wrangled.
+One problem with the long format is that it can be inefficient to store feature geometries, especially for large MULTIPOLYGON over long time period.
+-->
+
+<!-- For spatial data, \pkg{sf} [@sf] covers a wide range of geometrical operations and is the format most map data stored.
+When incorporating the time dimension, a time wide format, introduced in @spacetime, is usually adopted.
+This treats time as variable name rather than value and makes it inconvenient to wrangle time, as per the tidy data principle [@tidydata].
+In the temporal domain, \pkg{tsibble} [@tsibble] follows the tidy data principle to arrange the time in the long format, while when adding spatial variables, a long format implies repeating variables at each time and this can make a huge difference when the repetition is on large geometrical shapes with long time period.
+-->
+
+<!-- In the \proglang{R} community, many data objects have been proposed to work with specific spatio-temporal tasks, while few can freely explore the data.
 \pkg{spacetime} [@spacetime] is an attempt for such a data object, but the spatial and temporal classes it is built on have now been gradually replaced by more recent implementations.
 The \pkg{stars} [@stars] package uses an array to wrangle spatio-temporal data, but the dense array it uses can quickly exhaust the memory for data in the case of irregular spatial grid.
-While there has not yet been a data object that makes it easy to wrangle spatio-temporal data, there is a strong need for such tools given the ubiquitous spatial and temporal data collected in the society, i.e., climate observations from weather stations, air quality variables from air monitoring stations.
-The demand for such tools can also come from spatial data analysts wishing to incorporate temporal data in their analysis or vice versa.
-
-This paper presents a new \proglang{R} package, \pkg{cubble}, to overcome the problem with organising variables in different levels when working with spatio-temporal data.
-In the package a new data structure, also called \code{cubble}, is proposed to organise spatial and temporal variables as two forms of a single data object so that they can be wrangled separately while synchronised as a whole piece.
-With this new data object, analysts can spend less time writing codes to organise spatial and temporal variables and focus more on the exploratory data analysis itself.
-The software is available from the Comprehensive R Archive Network (CRAN) at [CRAN link].
-
-The rest of the paper is divided as follows: Section 2 presents the main design and functionality of \pkg{cubble}.
-Section 3 explains how cubble deals with more advanced considerations including data with hierarchical structure, data matching, how cubble fits with existing static and interactive visualisation tools, and spatio-temporal data transformation.
-Section 4 uses Australia weather station data and river level data as examples to demonstrate the use of \pkg{cubble}.
-An example on how \pkg{cubble} handles NetCDF data is also provided.
-Section 5 concludes the paper.
+-->

### 2022-04-18 2-0-cube.Rmd
a new skeleton of intro incorporating various past writing

--- 2022-03-26 2-0-cube.Rmd
@@
 
-# Conceptual framework
+<!-- and software implementation has been seen in many R packages: \pkg{raster} [@raster], \pkg{terra} [@terra], and \pkg{stars} [@stars]. -->
+
+<!-- # Existing work and new challenges -->
+
+<!-- Spatio-temporal software should echo with the advance in spatial and temporal data analysis.
+This section will start with the \pkg{spacetime} paper [@spacetime] that sets the foundation for wrangling spatio-temporal data in \proglang{R}, and then describe some emerging concepts and software in data analysis since the paper publishes and what remains as challenges.
+-->
+
+<!-- Also, array is not the best solution for a large subset of data, specifically, the vector spatio-temporal data.
+With vector data, spatial and temporal variables are commonly collected in separate 2D tables and there is usually a location ID to link the spatial and temporal information.
+Sometimes, the geometrical shape of the location may not present and an additional linking is needed to incorporate the map data.
+With the multiple pieces in 2D data frame format, there should be a solution to organise all these data from multiple sources in a 2D structure, without needing to convert it into higher dimensional representation.
+-->
+
+# Conceptual framework: spatio-temporal cube {#cube}
 
@@
 This naturally motivates the usage of array to represent and store spatio-temporal data, as evident by satellite imageries, large climate models, among others.
-The work by @lu_multidimensional_2018 has discussed the suitability of this representation through array algebra and practical spatio-temporal tasks and software implementation has been seen in many R packages: \pkg{raster} [@raster], \pkg{terra} [@terra], and \pkg{stars} [@stars].
+The work by @lu_multidimensional_2018 has discussed the suitability of this representation through array algebra and practical spatio-temporal tasks.
 
-While it is efficient to represent spatio-temporal data in an array, there are situation where array [is not the best solution]. mutate in the array is not clearly define (how to add a column vector)
-
-Also, array is not the best solution for a large subset of data, specifically, the vector spatio-temporal data.
-With vector data, spatial and temporal variables are commonly collected in separate 2D tables and there is usually a location ID to link the spatial and temporal information.
-Sometimes, the geometrical shape of the location may not present and an additional linking is needed to incorporate the map data.
-With the multiple pieces in 2D data frame format, there should be a solution to organise all these data from multiple sources in a 2D structure, without needing to convert it into higher dimensional representation.
-
-Reliable software and data structure has been proposed in the spatial and temporal domain, while a direct extension to spatio-temporal data is not ideal.
-The main issue here lays in the format that spatial and temporal data is stored.
-For spatial data, \pkg{sf} [@sf] covers a wide range of geometrical operations and is the format most map data stored.
-When incorporating the time dimension, a time wide format, introduced in @spacetime, is usually adopted.
-This treats time as variable name rather than value and makes it inconvenient to wrangle time, as per the tidy data principle [@tidydata].
-In the temporal domain, \pkg{tsibble} [@tsibble] follows the tidy data principle to arrange the time in the long format, while when adding spatial variables, a long format implies repeating variables at each time and this can make a huge difference when the repetition is on large geometrical shapes with long time period.
+The cubical conceptual framework provides a generalisation to the operation can be done on the data for visualisation.
+A paper by @bach_review_2014 has reviewed the temporal data visualisation based on space-time cube operations.
+Notice that the term space-time cube in their article "does not need to involve spatial data", but referring to "an abstract 2D substrate that is used to visualize data at a specific time".
+Despite its main focus is on temporal data, the mindset of abstracting out data representation to construct visualisation is applicable to our spatio-temporal data manipulation and visualisation.

### 2022-04-18 5-conclusion.Rmd
clean up bibbib
minor al layout adjustment
a new skeleton of intro incorporating various past writing

@@
 ---
-# Conclusion
+# Conclusion {#conclude}
 
@@
 Further development and maintenance of the package involves responding to changes in the tidyverse packages that \pkg{cubble} imports, in particular, \pkg{tibble}, \pkg{tidyr}, and \pkg{dplyr}.
-Another area for further development is to extend \pkg{cubble} to other spatial objects other than points.
+In the spatial aspect, the simple feature representation wraps spatial coordinates in a list column, for some complex spatial operations, while sometimes, analysts may want to unpack them into longitude and latitude columns to work from.
+Another extension to \pkg{cubble} is to build a smooth transition between the coordinate columns and simple feature geometry column.
+
+\newpage
 
@@
 ```
+
+\newpage

### 2022-04-18 jss/cubble-jss.tex
a new skeleton of intro incorporating various past writing

@@
 \author{
-H. Sherry Zhang\\Monash University \And Dianne Cook\\Monash University
-\AND Ursula Laa\\University of Natural Resources and Life Sciences
-\AND Nicolas Langrené\\BNU-HKBU United International College
-\AND Patricia Menéndez\\Monash University \AND
+H.
+Sherry Zhang\\Monash University \And Dianne Cook\\Monash University \AND Ursula Laa\\University of Natural Resources and Life Sciences \AND Nicolas Langrené\\BNU-HKBU United International College \AND Patricia Menéndez\\Monash University \AND
 }
@@
 
-\Plainauthor{H. Sherry Zhang, Dianne Cook, Ursula Laa, Nicolas
-Langrené, Patricia Menéndez}
+\Plainauthor{H. Sherry Zhang, Dianne Cook, Ursula Laa, Nicolas Langrené, Patricia Menéndez}
 \Plaintitle{cubble: An R Package for Structuring Spatio-temporal Data}
@@
 Spatio-temporal data refer to measurements taken across space and time.
-In practice, spatio-temporal data can be decomposed into a spatial and
-temporal component: at one time, we would select a spatial location and
-inspect the temporal trend; at other time, we might select one or
-multiple time value(s) and explore the spatial distribution. Ideally, we
-could make multiple maps and multiple time series to explore these
-together, however, doing all of these actions is complicated when data
-arrive fragmented in multiple objects. To make it easy to do all these
-tasks, ideally spatial and temporal variables are in a single data
-object that we can slice and dice in different ways to conduct different
-visualisations. This work suggests a new data structure, \code{cubble},
-that uses a nested form and a long form to organise spatial and temporal
-variables in a single data object. The new data structure ensures that
-data in the two forms are synchronised while been flexible to explore
-the two aspect of spatio-temporal data. It also provides capability for
-handling data with hierarchical structure, matching data from multiple
-sources, constructing interactive graphics, and performing
-spatio-temporal transformation. The paper will demonstrate \pkg{cubble}
-with examples using Australian climate weather stations, river level
-data, and climate reanalysis (ERA5) data.
-}
-
-\Keywords{spatial, temporal, spatio
-temporal, \proglang{R}, environmental data, exploratory data analysis}
-\Plainkeywords{spatial, temporal, spatio temporal, R, environmental
-data, exploratory data analysis}
+In practice, spatio-temporal data can be decomposed into a spatial and temporal component: at one time, we would select a spatial location and inspect the temporal trend; at other time, we might select one or multiple time value(s) and explore the spatial distribution.
+Ideally, we could make multiple maps and multiple time series to explore these together, however, doing all of these actions is complicated when data arrive fragmented in multiple objects.
+To make it easy to do all these tasks, ideally spatial and temporal variables are in a single data object that we can slice and dice in different ways to conduct different visualisations.
+This work suggests a new data structure, \code{cubble}, that uses a nested form and a long form to organise spatial and temporal variables in a single data object.
+The new data structure ensures that data in the two forms are synchronised while been flexible to explore the two aspect of spatio-temporal data.
+It also provides capability for handling data with hierarchical structure, matching data from multiple sources, constructing interactive graphics, and performing spatio-temporal transformation.
+The paper will demonstrate \pkg{cubble} with examples using Australian climate weather stations, river level data, and climate reanalysis (ERA5) data.
+}
+
+\Keywords{spatial, temporal, spatio temporal, \proglang{R}, environmental data, exploratory data analysis}
+\Plainkeywords{spatial, temporal, spatio temporal, R, environmental data, exploratory data analysis}
 
@@
 
+% From pandoc table feature
+\usepackage{longtable,booktabs,array}
+\usepackage{calc} % for calculating minipage widths
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
+
 \usepackage{amsmath} \usepackage{array} \usepackage{xcolor}
@@
 
-Spatio-temporal data involves variables observed in the space across
-time and this paper focuses on spatio-temporal vector data with fixed
-locations and regular time intervals. This type of data can naturally be
-considered as a data cube with three axes: spatial identifier, temporal
-identifier, and variables. In this cubic formation, space only occupies
-one axis, rather than two, to avoid hypercubes for multivariate
-spatio-temporal data. While conceptually the data format is relatively
-simple, the data objects coming to the analysts can vary in forms in the
-analysis. Figure \ref{fig:illu-input} shows three examples of wild
-spatio-temporal data: 1) a single table that includes both the spatial
-and temporal variables; 2) separate tables for spatial and temporal
-variables; 3) array-based data, i.e.~NetCDF (Network Common Data Form)
-data, which is commonly used in earth science, e.g.~climatology,
-meteorology, and oceanography, among others, to organise multivariate
-data in a spatio-temporal grid.
-
-\begin{CodeChunk}
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.001}
-
-}
-
-\caption[An illustration of different spatio-temporal data layouts]{An illustration of different spatio-temporal data layouts: (1) a single table for both spatial and temporal variables; (2) separate tables for spatial and temporal variables; and (3) an array data structure that populates variables into a gridded longitude-latitude-time (hyper)cube.}\label{fig:illu-input}
-\end{figure}
-\end{CodeChunk}
-
-While incorporating both spatial and temporal information enriches the
-data in the analysis, spatio-temporal data can sometimes be hard to work
-with given its two observational units in the variables. The single
-table format in Figure \ref{fig:illu-input} is convenient for
-spatio-temporal operations that involve both types of variables,
-however, it makes duplicates of spatial variables and is not ideal when
-only looking at the spatial aspect of the data. The two tables format
-allows separate operations on spatial and temporal variables while
-analysts need to work with two data objects simultaneously and join them
-together into the single table format for spatio-temporal operations.
-Neither of the two formats is desirable at the same time for spatial,
-temporal, and spatio-temporal operations. This issue implies analysts
-need to constantly think about the data format and arrange variables
-into the proper shape before the actual operations, which can be a
-laborious step when switching between the two formats back and forth
-several times during the analysis.
-
-In the \proglang{R} community, many data objects have been proposed to
-work with specific spatio-temporal tasks, while few can freely explore
-the data. \pkg{spacetime} \citep{spacetime} is an attempt for such a
-data object, but the spatial and temporal classes it is built on have
-now been gradually replaced by more recent implementations. The
-\pkg{stars} \citep{stars} package uses an array to wrangle
-spatio-temporal data, but the dense array it uses can quickly exhaust
-the memory for data in the case of irregular spatial grid. While there
-has not yet been a data object that makes it easy to wrangle
-spatio-temporal data, there is a strong need for such tools given the
-ubiquitous spatial and temporal data collected in the society, i.e.,
-climate observations from weather stations, air quality variables from
-air monitoring stations. The demand for such tools can also come from
-spatial data analysts wishing to incorporate temporal data in their
-analysis or vice versa.
-
-This paper presents a new \proglang{R} package, \pkg{cubble}, to
-overcome the problem with organising variables in different levels when
-working with spatio-temporal data. In the package a new data structure,
-also called \code{cubble}, is proposed to organise spatial and temporal
-variables as two forms of a single data object so that they can be
-wrangled separately while synchronised as a whole piece. With this new
-data object, analysts can spend less time writing codes to organise
-spatial and temporal variables and focus more on the exploratory data
-analysis itself. The software is available from the Comprehensive R
-Archive Network (CRAN) at {[}CRAN link{]}.
-
-The rest of the paper is divided as follows: Section 2 presents the main
-design and functionality of \pkg{cubble}. Section 3 explains how cubble
-deals with more advanced considerations including data with hierarchical
-structure, data matching, how cubble fits with existing static and
-interactive visualisation tools, and spatio-temporal data
-transformation. Section 4 uses Australia weather station data and river
-level data as examples to demonstrate the use of \pkg{cubble}. An
-example on how \pkg{cubble} handles NetCDF data is also provided.
-Section 5 concludes the paper.
-
-\hypertarget{conceptual-framework-spatio-temporal-cube}{%
-\section{Conceptual framework: spatio-temporal
-cube}\label{conceptual-framework-spatio-temporal-cube}}
-
-Multivariate spatio-temporal data can be conceptualised using a cubical
-data model with reference to variables, time and space. This naturally
-motivates the usage of array to represent and store spatio-temporal
-data, as evident by satellite imageries, large climate models, among
-others. The work by \citet{lu_multidimensional_2018} has discussed the
-suitability of this representation through array algebra and practical
-spatio-temporal tasks.
-
-The cubical conceptual framework provides a generalisation to the
-operation can be done on the data for visualisation. A paper by
-\citet{bach_review_2014} has reviewed the temporal data visualisation
-based on space-time cube operations. Notice that the term space-time
-cube in their article ``does not need to involve spatial data'', but
-referring to ``an abstract 2D substrate that is used to visualize data
-at a specific time''. Despite its main focus is on temporal data, the
-mindset of abstracting out data representation to construct
-visualisation is applicable to our spatio-temporal data manipulation and
-visualisation.
-
-\hypertarget{existing-work-and-new-challenges}{%
-\section{Existing work and new
-challenges}\label{existing-work-and-new-challenges}}
-
-To represent spatio-temporal data, \citet{spacetime} proposes four
-layouts: full grid layout (STF), sparse grid layout (STS), irregular
-layout (STI), and trajectory (STT). These layouts are implemented in the
-\pkg{spacetime} package with underlying class \pkg{sp} \citep{sp} and
-\pkg{xts} \citep{xts} for handling spatial and temporal variables. While
-these layouts and classes are still being used and well-maintained
-today, they are not compatible with the \pkg{tidyverse} ecosystem
-\citep{tidyverse}, which has a huge influence on data analysis in
-\proglang{R}, including in the spatial and temporal communities. The
-package \pkg{sf} \citep{sf} uses data frame objects to represent spatial
-data with an additional \code{geometry} column dedicating to feature
-geometries \citep{jr_herring_opengis_2011} (e.g.~POINT, MULTIPOINT,
-POLYGON, MULTIPOLYGON etc). For temporal data, \pkg{tsibble}
-\citep{tsibble} builds on top of the \pkg{tibble} \citep{tibble} object
-and explicitly includes date and time in columns as variables, rather
-than data attributes.
-
-Underlying tidyverse is the concept of tidy data \citep{tidydata}, which
-considers how variables should be structured for analysis. In the three
-formats (time-wide, space-wide, and long) summarised in the
-\citet{spacetime} paper, only the long form is considered as tidy. The
-main advantages of it over the time-wide and space-wide format is that
-the latter treat either time or location as variable names (Section 3.1
-in \citet{tidydata}) rather than values of variables, making them
-difficult to be wrangled. One problem with the long format is that it
-can be inefficient to store feature geometries, especially for large
-MULTIPOLYGON over long time period. This motivates a new tidyverse
-compatible data object to work with spatio-temporal data.
-
-\textbf{two challenages: multiple data sources + large spatio-temporal
-object} In additional to the shift into more tidyverse workflow for
-working with data, spatio-temporal data have some unique challenges. The
-first one being the split of variables in multiple tables. Often the
-time, spatio-temporal data is not presented in a single data object, but
-split into fragmented pieces from different sources. For example, map
-data recorded the shape of a collection of area of interest is provided
-by the government website; the geo-scientific data is recorded with
-longitude and latitude coordinate of locations in the area; and temporal
-variables indexed by both location and time need to be further queried
-using location data. For spatio-temporal data analysis, merging all
-these sources can be challenging and joining these tables can go wrong
-because of the slightest unmatched of the linking variable from
-different sources.
-
-Another challenge when analysing spatio-temporal data is on the simple
-feature representation of spatial data. While the simple feature is able
-to handle large and complex spatial geometry and perform precise spatial
-operations, data analysts usually do not have a large expectation on the
-map accuracy and would prefer a simplified map with quick rendering for
-visualisation. Also, the feature geometry column in \pkg{sf} wraps the
-spatial coordinates in a list column, sometimes, analysts may like to
-unpack them into longitude and latitude to work with. A smooth
-transition with the coordinate columns and and simple feature geometry
-column would satisfy analysts from both ends.
-
-With some of the aforementioned issues recognised, the \pkg{stars}
-package has been a new implementation to handle spatio-temporal data
-conceptually as data cubes and implemented as arrays. The package
-depends on the \pkg{sf} package and has allowed more access to tidyverse
-verbs in the workflow. However, for vector spatio-temporal data, major
-difficulties to work with \pkg{stars} is first to cast the variables
-from multiple tables into a star object and then keep the operations in
-the high-dimensional cube. Given that most table mentioned before
-collect variables in a 2D table, analysts could benefit from a structure
-to allow them work in 2D tables.
-
-\hypertarget{the-cubble-package}{%
-\section{The cubble package}\label{the-cubble-package}}
-
-This section will present two formats that cubble uses to arrange
-spatio-temporal data. The main cubble functions will then be introduced
-to illustrate how to work with these two formats in short examples.
-Lastly, a subsection will be dedicated to how existing packages, in
-spatial and temporal analysis, fit in with cubble.
-
-In cubble, a single data object can be arranged into two forms: nested
-form and long form. Figure \ref{fig:illu-cubble} sketches the two forms
-with the associated attributes. The decision on which form to use in
-data wrangling is output-oriented, meaning analysts need to first think
-about whether the output of an operation is identified only by the
-spatial identifier, or a combination of spatial and temporal identifier.
-The nested cubble is suitable for working with operations that are only
-identified by the site and this type of operation can be a pure
-manipulation of time invariant variables, or an operation that
-summarises time varying variables into site. Underneath the nested form,
-a cubble is built from a \code{rowwise_df} class where each site forms a
-separate group. This structure simplifies the calculation that involves
-temporal variables by avoiding the use of \code{purrr::map()} syntax
-when working with list-column.
-
-For those operations whose output involves both a spatial and temporal
-identifier, long form should be used. The long form uses a
-\code{grouped_df} class to forms all the time of a site as a group. Time
-invariant variables are stored separately as an special attribute of the
-long cubble. This design avoids repeating the spatial variables at each
-time stamp while not dropping information from spatial variables.
+Spatio-temporal data has a spatial component referring to the location of each observation, a temporal component that is recorded at regular or irregular time intervals, and multiple variables measured at each spatial and temporal value.
+In data collection, spatio-temporal data is less often collected as a whole piece but fragmented spatial and temporal data for analysts to ensemble into spatio-temporal data.
+With the {[}integrated{]} data, we hope to fix the time to explore the spatial distribution, or fix the spaital location to explore temporal trends.
+Ideally, we would be able to carry out each of these tasks easily, and even accommodate complex explorations encompassing both simultaneously.
+These are all underpinned by a {[}flexible{]} data structure that can represent spatio-temporal data.
+
+The \emph{Representing data} section of the CRAN task view {[}ref{]} lists existing spatio-temporal data representation.
+Among them, \citet{spacetime} proposes four layouts: full grid layout, sparse grid layout, irregular layout, and trajectory.
+These layouts are implemented in the \pkg{spacetime} package with underlying class \pkg{sp} \citep{sp} and \pkg{xts} \citep{xts} for handling spatial and temporal variables.
+The \pkg{stars} package has been a recent implementation to handle spatio-temporal data in an array.
+The package depends on the \pkg{sf} package and has allowed more access to tidyverse verbs in the workflow.
+
+What remains to be difficult when wrangling spatio-temporal data is that the raw data is not naturally structured in any one of the representation mentioned above and the first difficulty is to cast the data, potentially from multiple sources into one of the data representation.
+Some examples of these includes map data recorded the shape of a collection of area of interest is provided by the government website; geo-scientific data recorded with longitude and latitude coordinate of locations in the area; and temporal variables indexed by both location and time for the temporal aspect.
+For spatio-temporal data analysis, merging all these sources can be challenging and joining these tables can go wrong because of the slightest unmatched of the linking variable from different sources.
+
+Another issue is how the concept of tidy data should be applied to spatio-temporal data.
+The tidy data concept advocates the data to be shaped into 1) one row per observation; 2) one column per variable; and 3) one type of data per table.
+The long form data is preferred over wide data given the downstream data cleaning and visualisation software i.e.~\pkg{dplyr} and \pkg{ggplot2}.
+However, the long form can be inefficient to store feature geometries, especially for large multipolygon over long time period, i.e.~sub-daily time series data.
+
+This paper presents a new \proglang{R} package, \pkg{cubble}, to {[}overcome the problem{]} with organising variables in different levels when working with spatio-temporal data.
+The data structure focuses on connecting to the upstream data collection and is suitable for temporal data collected at the fixed location over time.
+Among the four spacetime layouts in \citet{spacetime}, it can be full grid layout, sparse grid layout, or irregular layout, but trajectory is not within the scope of this work.
+In the package a new data structure, also called \code{cubble}, is proposed to organise spatial and temporal variables as two forms of a single data object so that they can be wrangled separately while synchronised as a whole piece.
+With this new data object, analysts can spend less time writing codes to organise spatial and temporal variables and focus more on the exploratory data analysis itself.
+The software is available from the Comprehensive R Archive Network (CRAN) at {[}CRAN link{]}.
+
+The rest of the paper is divided as follows: Section \ref{cube} discusses the spatio-temporal cube to conceptualise spatio-temporal data.
+Section \ref{cubble} presents the main design and functionality of \pkg{cubble}.
+Section \ref{others} explains how cubble deals with more advanced considerations including data with hierarchical structure, data matching, how cubble fits with existing static and interactive visualisation tools, and spatio-temporal data transformation.
+Section \ref{examples} uses Australia weather station data and river level data as examples to demonstrate the use of \pkg{cubble}.
+An example on how \pkg{cubble} handles NetCDF data is also provided.
+Section \ref{conclude} concludes the paper.
+
+\hypertarget{cube}{%
+\section{Conceptual framework: spatio-temporal cube}\label{cube}}
+
+Multivariate spatio-temporal data can be conceptualised using a cubical data model with reference to variables, time and space.
+This naturally motivates the usage of array to represent and store spatio-temporal data, as evident by satellite imageries, large climate models, among others.
+The work by \citet{lu_multidimensional_2018} has discussed the suitability of this representation through array algebra and practical spatio-temporal tasks.
+
+The cubical conceptual framework provides a generalisation to the operation can be done on the data for visualisation.
+A paper by \citet{bach_review_2014} has reviewed the temporal data visualisation based on space-time cube operations.
+Notice that the term space-time cube in their article ``does not need to involve spatial data'', but referring to ``an abstract 2D substrate that is used to visualize data at a specific time''.
+Despite its main focus is on temporal data, the mindset of abstracting out data representation to construct visualisation is applicable to our spatio-temporal data manipulation and visualisation.
+
+\hypertarget{cubble}{%
+\section{The cubble package}\label{cubble}}
+
+This section will present two formats that cubble uses to arrange spatio-temporal data.
+The main cubble functions will then be introduced to illustrate how to work with these two formats in short examples.
+Lastly, a subsection will be dedicated to how existing packages, in spatial and temporal analysis, fit in with cubble.
+
+In cubble, a single data object can be arranged into two forms: nested form and long form.
+Figure \ref{fig:illu-cubble} sketches the two forms with the associated attributes.
+The decision on which form to use in data wrangling is output-oriented, meaning analysts need to first think about whether the output of an operation is identified only by the spatial identifier, or a combination of spatial and temporal identifier.
+The nested cubble is suitable for working with operations that are only identified by the site and this type of operation can be a pure manipulation of time invariant variables, or an operation that summarises time varying variables into site.
+Underneath the nested form, a cubble is built from a \code{rowwise_df} class where each site forms a separate group.
+This structure simplifies the calculation that involves temporal variables by avoiding the use of \code{purrr::map()} syntax when working with list-column.
+
+For those operations whose output involves both a spatial and temporal identifier, long form should be used.
+The long form uses a \code{grouped_df} class to forms all the time of a site as a group.
+Time invariant variables are stored separately as an special attribute of the long cubble.
+This design avoids repeating the spatial variables at each time stamp while not dropping information from spatial variables.
 
@@
 \hypertarget{create-a-cubble-in-the-nested-form}{%
-\subsection{Create a cubble in the nested
-form}\label{create-a-cubble-in-the-nested-form}}
-
-To use functions in the \texttt{cubble} package, an analyst will first
-need to turn the data object into a \code{cubble} class.
-\code{as_cubble()} does this through supplying the three key components:
-\code{key} as the spatial identifier; \code{index} as the temporal
-identifier; and a vector of \code{coords} in the order of longitude and
-latitude. The use of \code{key} and \code{index} follows the design in
-the \pkg{tsibble} package and the cubble created by default is in the
-nested form.
-
-Before the formal examples in Section \ref{examples}, each function
-introduced in this section will be accompanied by a small example. The
-data used for the small examples is a subset of a larger climate weather
-station data used in the formal examples and has been simplified to
-include only five weather stations. It contains spatial information of
-each station: station id, latitude, longitude, elevation, station name
-and World Meteorology Organisation ID and also daily temporal
-information of date, maximum and minimum temperature and precipitation
-for 2020. \code{climate_flat} stores this data in format 1) in Figure
-\ref{fig:illu-input} and the code below creates a cubble out of
-\code{climate_flat} with \code{id} as the key, \code{date} as the index,
-and \code{c(long, lat)} as the coordinates:
+\subsection{Create a cubble in the nested form}\label{create-a-cubble-in-the-nested-form}}
+
+To use functions in the \texttt{cubble} package, an analyst will first need to turn the data object into a \code{cubble} class.
+\code{as_cubble()} does this through supplying the three key components: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order of longitude and latitude.
+The use of \code{key} and \code{index} follows the design in the \pkg{tsibble} package and the cubble created by default is in the nested form.
+
+Before the formal examples in Section \ref{examples}, each function introduced in this section will be accompanied by a small example.
+The data used for the small examples is a subset of a larger climate weather station data used in the formal examples and has been simplified to include only five weather stations.
+It contains spatial information of each station: station id, latitude, longitude, elevation, station name and World Meteorology Organisation ID and also daily temporal information of date, maximum and minimum temperature and precipitation for 2020.
+\code{climate_flat} stores this data in format 1) in Figure \ref{fig:illu-input} and the code below creates a cubble out of \code{climate_flat} with \code{id} as the key, \code{date} as the index, and \code{c(long, lat)} as the coordinates:
 
@@
 
-There are a few information in the cubble header: the name of the
-\code{key} variable: \code{id} and its unique number: 5, the bounding
-box: \code{[115.97, -32.94, 133.55, -12.42]} , and also the name of
-variable nested in the \code{ts} column with its type:
-\code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.
+There are a few information in the cubble header: the name of the \code{key} variable: \code{id} and its unique number: 5, the bounding box: \code{[115.97, -32.94, 133.55, -12.42]} , and also the name of variable nested in the \code{ts} column with its type: \code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.
 
 \hypertarget{stretch-a-nested-cubble-into-the-long-form}{%
-\subsection{Stretch a nested cubble into the long
-form}\label{stretch-a-nested-cubble-into-the-long-form}}
-
-The nested format is convenient for those operations whose output does
-not contain a time dimension. For those outputs that are
-cross-identified by the spatial and temporal identifier, a long cubble
-needs to be used. The function \code{stretch()} is designed to switch
-the cubble from the nested form to the long form. This code shows how to
-switch the nested cubble just created into its long form:
+\subsection{Stretch a nested cubble into the long form}\label{stretch-a-nested-cubble-into-the-long-form}}
+
+The nested format is convenient for those operations whose output does not contain a time dimension.
+For those outputs that are cross-identified by the spatial and temporal identifier, a long cubble needs to be used.
+The function \code{stretch()} is designed to switch the cubble from the nested form to the long form.
+This code shows how to switch the nested cubble just created into its long form:
 
@@
 
-Notice that the third line in the header now shows the name and type of
-spatial variables:
-\code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]} and
-these variables are now stored as a \code{spatial} attribute of the
-data:
+Notice that the third line in the header now shows the name and type of spatial variables: \code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]} and these variables are now stored as a \code{spatial} attribute of the data:
 
@@
 \hypertarget{tamp-a-long-cubble-back-to-the-nested-form}{%
-\subsection{Tamp a long cubble back to the nested
-form}\label{tamp-a-long-cubble-back-to-the-nested-form}}
-
-Manipulation on the spatial and temporal dimension can be an iterative
-process and analysts may need to go back and forth between the nested
-and long cubble. \code{tamp()}, an inverse of \code{stretch()}, switches
-a long cubble to a nested cubble:
+\subsection{Tamp a long cubble back to the nested form}\label{tamp-a-long-cubble-back-to-the-nested-form}}
+
+Manipulation on the spatial and temporal dimension can be an iterative process and analysts may need to go back and forth between the nested and long cubble.
+\code{tamp()}, an inverse of \code{stretch()}, switches a long cubble to a nested cubble:
 
@@
 \hypertarget{migrate-spatial-variables-to-a-long-cubble}{%
-\subsection{Migrate spatial variables to a long
-cubble}\label{migrate-spatial-variables-to-a-long-cubble}}
-
-Sometimes, analysts may need to apply some variable transformation that
-involves both the spatial and temporal variable. An example of this is
-the transformation of temporal variables into the spatial dimension in
-glyph maps (which will be elaborated in section
-\ref{st_transformation}). Cubble allows this operation through
-\code{unfold()}, which moves the specified spatial variables into the
-long cubble:
+\subsection{Migrate spatial variables to a long cubble}\label{migrate-spatial-variables-to-a-long-cubble}}
+
+Sometimes, analysts may need to apply some variable transformation that involves both the spatial and temporal variable.
+An example of this is the transformation of temporal variables into the spatial dimension in glyph maps (which will be elaborated in section \ref{st_transformation}).
+Cubble allows this operation through \code{unfold()}, which moves the specified spatial variables into the long cubble:
 
@@
 
-This function should generally be used in the last step of the analysis
-since it is a temporary operation, meaning these added spatial variables
-are not stored in the long form and will disappear if switched to the
-nested form and then switched back:
+This function should generally be used in the last step of the analysis since it is a temporary operation, meaning these added spatial variables are not stored in the long form and will disappear if switched to the nested form and then switched back:
 
@@
 \hypertarget{compatibility-with-existing-packages}{%
-\subsection{Compatibility with existing
-packages}\label{compatibility-with-existing-packages}}
-
-The previous four subsections have introduced operations specific to the
-\code{cubble} class and this section will demonstrate how the
-\code{cubble} class interacts with existing packages commonly used in
-spatial and temporal analysis, specifically, \code{dplyr},
-\code{tsibble}, \code{sf} (\code{s2}), and \code{netcdf4}.
+\subsection{Compatibility with existing packages}\label{compatibility-with-existing-packages}}
+
+The previous four subsections have introduced operations specific to the \code{cubble} class and this section will demonstrate how the \code{cubble} class interacts with existing packages commonly used in spatial and temporal analysis, specifically, \code{dplyr}, \code{tsibble}, \code{sf} (\code{s2}), and \code{netcdf4}.
 
@@
 
-The \code{dplyr} package has provided many tools for data wrangling
-tasks and these operations are useful in the spatio-temporal context.
-\code{cubble} provides methods that support the following \code{dplyr}
-verbs in both the nested and long form:
+The \code{dplyr} package has provided many tools for data wrangling tasks and these operations are useful in the spatio-temporal context.
+\code{cubble} provides methods that support the following \code{dplyr} verbs in both the nested and long form:
 
 \begin{quote}
-\texttt{mutate}, \texttt{filter}, \texttt{summarise}, \texttt{select},
-\texttt{arrange}, \texttt{rename}, \texttt{left\_join}, and the slice
-family (\texttt{slice\_head}, \texttt{slice\_tail},
-\texttt{slice\_sample}, \texttt{slice\_min}, \texttt{slice\_max})
+\texttt{mutate}, \texttt{filter}, \texttt{summarise}, \texttt{select}, \texttt{arrange}, \texttt{rename}, \texttt{left\_join}, and the slice family (\texttt{slice\_head}, \texttt{slice\_tail}, \texttt{slice\_sample}, \texttt{slice\_min}, \texttt{slice\_max})
 \end{quote}
@@
 
-\code{tsibble} is a temporal data structure that uses \code{index} and
-\code{key} to identify the time and different series. \code{cubble} can
-be seen as a natural extension of \code{tsibble} for spatio-temporal
-data with an additional coordinates component and using two forms to
-arrange variables. This makes it easy to cast a \code{tsibble} into a
-\code{cubble} as only the \code{coords} argument needs to be supplied:
+\code{tsibble} is a temporal data structure that uses \code{index} and \code{key} to identify the time and different series.
+\code{cubble} can be seen as a natural extension of \code{tsibble} for spatio-temporal data with an additional coordinates component and using two forms to arrange variables.
+This makes it easy to cast a \code{tsibble} into a \code{cubble} as only the \code{coords} argument needs to be supplied:
 
@@
 
-In the nested cubble created, each element in the list-column \code{ts}
-is of \code{tbl_ts} class and operations available to the tsibble class
-is still valid under cubble. For example, the code below calculates two
-features of the maximum temperature:
+In the nested cubble created, each element in the list-column \code{ts} is of \code{tbl_ts} class and operations available to the tsibble class is still valid under cubble.
+For example, the code below calculates two features of the maximum temperature:
 
@@
 
-As a spatial data object, \code{sf} creates a simple feature geometry
-list-column (\code{sfc}) in the data frame to provide spatial operations
-on various geometry types (\code{POINT}, \code{LINESTRING},
-\code{POLYGON}, \code{MULTIPOLYGON}, etc). These spatial operations are
-also valuable for spatio-temporal data analysis, but an \code{sf} object
-\emph{usually} does not contain temporal variables. This means \code{sf}
-cannot be directly cast into a \code{cubble}, however, \code{cubble}
-does support \code{sfc} columns in the nested form and spatial
-operations applied to the \code{sfc} column in \code{sf} can still be
-applied to the \code{sfc} column in a cubble. The following example
-shows how to create an \code{sfc} column of \code{POINT} type from
-latitude and longitude in cubble. Then \code{sf::st_within} is used to
-add the state \code{MULTIPOLYGON} of each weather station before a
-coordinate transformation is made.
+As a spatial data object, \code{sf} creates a simple feature geometry list-column (\code{sfc}) in the data frame to provide spatial operations on various geometry types (\code{POINT}, \code{LINESTRING}, \code{POLYGON}, \code{MULTIPOLYGON}, etc).
+These spatial operations are also valuable for spatio-temporal data analysis, but an \code{sf} object \emph{usually} does not contain temporal variables.
+This means \code{sf} cannot be directly cast into a \code{cubble}, however, \code{cubble} does support \code{sfc} columns in the nested form and spatial operations applied to the \code{sfc} column in \code{sf} can still be applied to the \code{sfc} column in a cubble.
+The following example shows how to create an \code{sfc} column of \code{POINT} type from latitude and longitude in cubble.
+Then \code{sf::st_within} is used to add the state \code{MULTIPOLYGON} of each weather station before a coordinate transformation is made.
 
@@
 
-An \pkg{s2} \code{lnglat} vector can similarly be created as an
-\code{sfc} in cubble before using any \code{s2}-prefixed function:
+An \pkg{s2} \code{lnglat} vector can similarly be created as an \code{sfc} in cubble before using any \code{s2}-prefixed function:
 
@@
 
-NetCDF data has two main components: \emph{dimension} for defining the
-spatio-temporal grid (longitude, latitude, and time) and \emph{variable}
-that populates the defined grid. Attributes are usually associated with
-dimension and variable in the NetCDF format data and a
-\href{http://cfconventions.org/}{metadata convention for climate and
-forecast} has been designed to standardise the format of the attributes.
-A few packages in R exists for manipulating NetCDF data and this
-includes a high-level R interface: \pkg{ncdf4} \citep{ncdf4}, a
-low-level interface that calls C-interface: \pkg{RNetCDF}
-\citep{rnetcdf, michna2013rnetcdf}, and a tidyverse implementation:
-\pkg{tidync} \citep{tidync}.
-
-Cubble provides an \code{as_cubble()} method to coerce the \code{ncdf4}
-class from the \pkg{ncdf4} package into a \code{cubble}. It maps each
-combination of longitude and latitude into an \code{id} as the
-\code{key}:
+NetCDF data has two main components: \emph{dimension} for defining the spatio-temporal grid (longitude, latitude, and time) and \emph{variable} that populates the defined grid.
+Attributes are usually associated with dimension and variable in the NetCDF format data and a \href{http://cfconventions.org/}{metadata convention for climate and forecast} has been designed to standardise the format of the attributes.
+A few packages in R exists for manipulating NetCDF data and this includes a high-level R interface: \pkg{ncdf4} \citep{ncdf4}, a low-level interface that calls C-interface: \pkg{RNetCDF} \citep{rnetcdf, michna2013rnetcdf}, and a tidyverse implementation: \pkg{tidync} \citep{tidync}.
+
+Cubble provides an \code{as_cubble()} method to coerce the \code{ncdf4} class from the \pkg{ncdf4} package into a \code{cubble}.
+It maps each combination of longitude and latitude into an \code{id} as the \code{key}:
 
@@
 
-The memory limit with NetCDF data in cubble depends on the number of
-longitude grid points times the number of latitude grid points times the
-number of time grid points times the number of variables. Cubble can
-handle slightly more than 300 x 300 (longitude x longitude) grid points
-for 3 variables in one year. The spatial grid can be reduced to trade
-for longer time periods and more variables. A 300 by 300 spatial grid
-can be a bounding box of {[}100, -80, 180, 0{]} at 0.25 degree
-resolution or a global bounding box {[}-180, -90, 180, -90{]} at 1
-degree resolution. Subsetting longitude and latitude grids is available
-through \code{long_range} and \code{lat_range} if the NetCDF file has
-finer resolution than needed.
+The memory limit with NetCDF data in cubble depends on the number of longitude grid points times the number of latitude grid points times the number of time grid points times the number of variables.
+Cubble can handle slightly more than 300 x 300 (longitude x longitude) grid points for 3 variables in one year.
+The spatial grid can be reduced to trade for longer time periods and more variables.
+A 300 by 300 spatial grid can be a bounding box of {[}100, -80, 180, 0{]} at 0.25 degree resolution or a global bounding box {[}-180, -90, 180, -90{]} at 1 degree resolution.
+Subsetting longitude and latitude grids is available through \code{long_range} and \code{lat_range} if the NetCDF file has finer resolution than needed.
 
@@
 
-\hypertarget{other-features-and-considerations}{%
-\section{Other features and
-considerations}\label{other-features-and-considerations}}
+\hypertarget{others}{%
+\section{Other features and considerations}\label{others}}
 
@@
 
-Spatial locations can have grouping structure either inherent to the
-data or formed by clustering. Rather than analysing variables in the
-site level, summarised variables in the cluster level can give a crisper
-picture of local areas. In cubble, \code{switch_key()} can be used to
-create a new level of grouping of spatial locations by specifying a
-clustering variable. Figure \ref{fig:illu-hier} illustrates the
-relationship of cubbles at station and cluster level, in both the long
-and nested form. By specifying
-\code{cluster_nested <- station_nested \%>\% switch_key(key = cluster)},
-the cubble re-defines the cubble key from \code{id} in
-\code{station_nested} to \code{cluster} in \code{cluster_nested}. All
-the spatial variables variant to \code{cluster} are now nested into a
-\code{.val} column and cluster level variables can be computed in the
-same fashion as station level variables in \code{station_nested}.
+Spatial locations can have grouping structure either inherent to the data or formed by clustering.
+Rather than analysing variables in the site level, summarised variables in the cluster level can give a crisper picture of local areas.
+In cubble, \code{switch_key()} can be used to create a new level of grouping of spatial locations by specifying a clustering variable.
+Figure \ref{fig:illu-hier} illustrates the relationship of cubbles at station and cluster level, in both the long and nested form.
+By specifying \code{cluster_nested <- station_nested \%>\% switch_key(key = cluster)}, the cubble re-defines the cubble key from \code{id} in \code{station_nested} to \code{cluster} in \code{cluster_nested}.
+All the spatial variables variant to \code{cluster} are now nested into a \code{.val} column and cluster level variables can be computed in the same fashion as station level variables in \code{station_nested}.
 
@@
 
-One task that may interest analysts in spatio-temporal data is to find
-how similar the time series from nearby sites are. This can be seen as a
-matching problem \citep{stuart2010matching, mcintosh2018using} that
-pairs up similar similar time series in nearby locations or a data
-fusing exercise that merges data collected from different sources
-\citep{cocchi2019data}. \code{match_sites()} in \pkg{cubble} provides a
-simple algorithm for this task. The algorithm first matches the two data
-sources spatially through computing the pairwise distance on latitude
-and longitude. Pairs that pass the spatial matching are then matched
-temporally through computing the number of matched peaks within a fixed
-length window. Figure \ref{fig:illu-matching} illustrates this temporal
-matching in more details. Given two series \code{A} and \code{a}, 3
-peaks have been picked in each series. An interval, with default length
-of 5, is constructed for each peak in series \code{A} and the peaks in
-series \code{a} are tested against whether they fall into the any of the
-intervals. In this illustration, there are 2 matches for these two
-series. Several arguments are available in \code{match_sites()} to
-fine-tune the matching:
+One task that may interest analysts in spatio-temporal data is to find how similar the time series from nearby sites are.
+This can be seen as a matching problem \citep{stuart2010matching, mcintosh2018using} that pairs up similar similar time series in nearby locations or a data fusing exercise that merges data collected from different sources \citep{cocchi2019data}.
+\code{match_sites()} in \pkg{cubble} provides a simple algorithm for this task.
+The algorithm first matches the two data sources spatially through computing the pairwise distance on latitude and longitude.
+Pairs that pass the spatial matching are then matched temporally through computing the number of matched peaks within a fixed length window.
+Figure \ref{fig:illu-matching} illustrates this temporal matching in more details.
+Given two series \code{A} and \code{a}, 3 peaks have been picked in each series.
+An interval, with default length of 5, is constructed for each peak in series \code{A} and the peaks in series \code{a} are tested against whether they fall into the any of the intervals.
+In this illustration, there are 2 matches for these two series.
+Several arguments are available in \code{match_sites()} to fine-tune the matching:
 
@@
 \item
-  \code{spatial_n_keep}: the number of spatial match for each site to
-  keep
+  \code{spatial_n_keep}: the number of spatial match for each site to keep
 \item
-  \code{spatial_dist_max}: the maximum distance allowed for a matched
-  pair
+  \code{spatial_dist_max}: the maximum distance allowed for a matched pair
 \item
-  \code{temporal_n_highest}: the number of peak used - 3 in the example
-  above
+  \code{temporal_n_highest}: the number of peak used - 3 in the example above
 \item
-  \code{temporal_window}: the length of the interval - 5 in the example
-  above
+  \code{temporal_window}: the length of the interval - 5 in the example above
 \item
-  \code{temporal_min_match}: the minimum number of matched peak for a
-  valid matched pair
+  \code{temporal_min_match}: the minimum number of matched peak for a valid matched pair
 \end{itemize}
@@
 
-Cubble fits in naturally with the interactive graphic pipeline discussed
-in the literature
-\citep{buja1988elements, buja1996interactive, sutherland2000orca, xie2014reactive, cheng2016enabling}.
-Diagram \ref{fig:illu-interactive} illustrates how linking works from
-the map to the time series in cubble. The map and time series plot is
-associated with the nested or long cubble, respectively, and when a user
-action is captured on the map, the site will be activated in the nested
-cubble (left). The nested cubble will communicate to the long cubble to
-activate all the observations with the same \code{id} (middle). The long
-cubble will then highlight the activated series in the time series plot
-(right).
-
-The linking is also available from the time series plot to the map. The
-selection(s) on the time series is through selecting the point(s) on the
-time series and once a point is selected, it will be activated in the
-long cubble. All the observations that share the same \code{id} are then
-activated and this includes other points in the same time series in the
-long cubble and the corresponding observation of site in the nested
-cubble. These activated observations will then being reflected in the
-updated plots and Diagram \ref{fig:illu-interactive-2} in the Appendix
-illustrates this process.
+Cubble fits in naturally with the interactive graphic pipeline discussed in the literature \citep{buja1988elements, buja1996interactive, sutherland2000orca, xie2014reactive, cheng2016enabling}.
+Diagram \ref{fig:illu-interactive} illustrates how linking works from the map to the time series in cubble.
+The map and time series plot is associated with the nested or long cubble, respectively, and when a user action is captured on the map, the site will be activated in the nested cubble (left).
+The nested cubble will communicate to the long cubble to activate all the observations with the same \code{id} (middle).
+The long cubble will then highlight the activated series in the time series plot (right).
+
+The linking is also available from the time series plot to the map.
+The selection(s) on the time series is through selecting the point(s) on the time series and once a point is selected, it will be activated in the long cubble.
+All the observations that share the same \code{id} are then activated and this includes other points in the same time series in the long cubble and the corresponding observation of site in the nested cubble.
+These activated observations will then being reflected in the updated plots and Diagram \ref{fig:illu-interactive-2} in the Appendix illustrates this process.
 
@@
 
-Spatio-temporal transformation is a useful technique to extract
-information form spatio-temporal data. Glyph maps \citep{Wickham2012-yr}
-transform the time coordinates into the space coordinates to plot the
-time series of different locations on the map. Calendar plots
-\citep{wang2020calendar} reconstructs time into calendar-based grid to
-discover weekday and weekend pattern. Projection, or linear combination,
-of variables summarises multivariate information into lower dimension to
-further digest. This section elaborates on the glyph map.
-
-In \proglang{R}, \pkg{GGally} implements glyph maps through the
-\code{glyphs()} function. The function constructs a data frame with
-calculated position (\code{gx}. \code{gy}. \code{gid}) of each point on
-the time series using linear algebra (Equation 1 and 2 in
-\citet{Wickham2012-yr}). The data can then be piped into \code{ggplot}
-to create the glyph map as:
+Spatio-temporal transformation is a useful technique to extract information form spatio-temporal data.
+Glyph maps \citep{Wickham2012-yr} transform the time coordinates into the space coordinates to plot the time series of different locations on the map.
+Calendar plots \citep{wang2020calendar} reconstructs time into calendar-based grid to discover weekday and weekend pattern.
+Projection, or linear combination, of variables summarises multivariate information into lower dimension to further digest.
+This section elaborates on the glyph map.
+
+In \proglang{R}, \pkg{GGally} implements glyph maps through the \code{glyphs()} function.
+The function constructs a data frame with calculated position (\code{gx}.
+\code{gy}.
+\code{gid}) of each point on the time series using linear algebra (Equation 1 and 2 in \citet{Wickham2012-yr}).
+The data can then be piped into \code{ggplot} to create the glyph map as:
 
@@
 
-A re-implementation of the glyph map as a ggproto, \code{GeomGlyph}, has
-been made in the \pkg{cubble} package and now the glyph map can be
-created with \code{geom_glyph()}:
+A re-implementation of the glyph map as a ggproto, \code{GeomGlyph}, has been made in the \pkg{cubble} package and now the glyph map can be created with \code{geom_glyph()}:
 
@@
 
-Some useful controls over the glyph map is also available in the
-\code{geom_glyph()} implementation. Polar glyph map can be specify as a
-parameter, \code{polar = TRUE}. in the \code{geom_glyph()}, along with
-\code{width} and \code{height} in either absolute or relative value.
-Global and local scale can be controlled by the parameter
-\code{global_rescale}. which default to \code{TRUE} for global scaling.
-Reference box and line can be added with separate
-\code{geom_glyph_box()} and \code{geom_glyph_line()}.
+Some useful controls over the glyph map is also available in the \code{geom_glyph()} implementation.
+Polar glyph map can be specify as a parameter, \code{polar = TRUE}.
+in the \code{geom_glyph()}, along with \code{width} and \code{height} in either absolute or relative value.
+Global and local scale can be controlled by the parameter \code{global_rescale}.
+which default to \code{TRUE} for global scaling.
+Reference box and line can be added with separate \code{geom_glyph_box()} and \code{geom_glyph_line()}.
 
@@
 \hypertarget{australia-historical-maximum-temperature}{%
-\subsection{Australia historical maximum
-temperature}\label{australia-historical-maximum-temperature}}
-
-Global Historical Climatology Network (GHCN) provides daily climate
-measures from stations across the world. The dataset
-\code{weatherdata::historical_tmax} extracts the maximum temperature for
-236 Australian stations from GHCN with starting from year 1969.
-\code{weatherdata::historical_tmax} is already in a cubble, with
-\code{id} as the key, \code{date} as the index, and
-\code{c(longitude, latitude)} as the coordinates. This example compares
-the maximum temperature in two periods: 1971 - 1975 and 2016 - 2020 for
-stations in Victoria and New South Wales.
-
-Stations in the two states can be subsetted by the station number:
-Australia GHCN station number starts with ``ASN00'' and followed by the
-\href{http://www.bom.gov.au/climate/cdo/about/site-num.shtml}{Bureau of
-Meteorology (BOM) station number}, where the 2nd and 3rd digit (7th and
-8th in the GHCN number) define the state of the station. New South Wales
-stations start from 46 to 75 and Victoria stations follow from 76 to 90.
-Filtering Victoria and New South Wales stations is an operation in the
-spatial dimension and hence uses the nested form:
+\subsection{Australia historical maximum temperature}\label{australia-historical-maximum-temperature}}
+
+Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world.
+The dataset \code{weatherdata::historical_tmax} extracts the maximum temperature for 236 Australian stations from GHCN with starting from year 1969.
+\code{weatherdata::historical_tmax} is already in a cubble, with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates.
+This example compares the maximum temperature in two periods: 1971 - 1975 and 2016 - 2020 for stations in Victoria and New South Wales.
+
+Stations in the two states can be subsetted by the station number: Australia GHCN station number starts with ``ASN00'' and followed by the \href{http://www.bom.gov.au/climate/cdo/about/site-num.shtml}{Bureau of Meteorology (BOM) station number}, where the 2nd and 3rd digit (7th and 8th in the GHCN number) define the state of the station.
+New South Wales stations start from 46 to 75 and Victoria stations follow from 76 to 90.
+Filtering Victoria and New South Wales stations is an operation in the spatial dimension and hence uses the nested form:
 
@@
 
-Filtering for the period 1971 \textasciitilde{} 1975 and 2016
-\textasciitilde{} 2020 is an operation on the time dimension and the
-nested cubble needs to be switched to the long cubble by
-\code{stretch()}:
+Filtering for the period 1971 \textasciitilde{} 1975 and 2016 \textasciitilde{} 2020 is an operation on the time dimension and the nested cubble needs to be switched to the long cubble by \code{stretch()}:
 
@@
 
-A monthly average is used for both periods to smooth the maximum
-temperature and it is again an operation on the time dimension:
+A monthly average is used for both periods to smooth the maximum temperature and it is again an operation on the time dimension:
 
@@
 
-A few stations don't have records during 1971 - 1975 and further
-investigation shows that while the first and last year of each series is
-recorded, the missing years in this period is not reported. These
-stations are filtered out by examining whether the summarised time
-series has a total of 24 months. The long cubble needs to be switched to
-the nested form for this spatial operation using \code{face_spatial()}:
+A few stations don't have records during 1971 - 1975 and further investigation shows that while the first and last year of each series is recorded, the missing years in this period is not reported.
+These stations are filtered out by examining whether the summarised time series has a total of 24 months.
+The long cubble needs to be switched to the nested form for this spatial operation using \code{face_spatial()}:
 
@@
 
-Lastly, to create a glyph map, both the major (\code{longitude},
-\code{latitude}) and minor (\code{month}, \code{tmax}) coordinates need
-to be in the same table. Spatial variables can be moved to the long form
-with \code{migrate()}:
+Lastly, to create a glyph map, both the major (\code{longitude}, \code{latitude}) and minor (\code{month}, \code{tmax}) coordinates need to be in the same table.
+Spatial variables can be moved to the long form with \code{migrate()}:
 
@@
 
-\code{tmax} can then be supplied to \code{geom_glyph()} for the glyph
-map in Figure \ref{fig:basic-manip} with a station inset on the top left
-corner:
+\code{tmax} can then be supplied to \code{geom_glyph()} for the glyph map in Figure \ref{fig:basic-manip} with a station inset on the top left corner:
 
@@
 \hypertarget{australia-precipitation-pattern-in-2020}{%
-\subsection{Australia precipitation pattern in
-2020}\label{australia-precipitation-pattern-in-2020}}
-
-In the previous example, there has already been some overlapping of the
-glyphs for a few stations near (151E, 34S) and (152E, 33S) and this will
-be a problem when mapping more stations in the national level.
-Aggregation can be helpful to group series into clusters before
-visualising the cluster with glyph map. This example shows how to
-organise data at both level with \code{switch_key()}.
-
-\code{weatherdata::climate_full}, also extracted from GHCN, records
-daily precipitation and maximum/ minimum temperature for 639 stations in
-Australia from 2016 to 2020. A simple kmean algorithm based on the
-distance matrix is used to create 20 clusters. This creates
-\code{station_nested} as a station level nested cubble with a cluster
-column indicating the group each station belongs to. More complex
-algorithms can be used for other problem, as long as there is a mapping
-from each station to a cluster.
+\subsection{Australia precipitation pattern in 2020}\label{australia-precipitation-pattern-in-2020}}
+
+In the previous example, there has already been some overlapping of the glyphs for a few stations near (151E, 34S) and (152E, 33S) and this will be a problem when mapping more stations in the national level.
+Aggregation can be helpful to group series into clusters before visualising the cluster with glyph map.
+This example shows how to organise data at both level with \code{switch_key()}.
+
+\code{weatherdata::climate_full}, also extracted from GHCN, records daily precipitation and maximum/ minimum temperature for 639 stations in Australia from 2016 to 2020.
+A simple kmean algorithm based on the distance matrix is used to create 20 clusters.
+This creates \code{station_nested} as a station level nested cubble with a cluster column indicating the group each station belongs to.
+More complex algorithms can be used for other problem, as long as there is a mapping from each station to a cluster.
 
@@
 
-To create a group level cubble, use \code{switch_key()} with the new key
-variable, \code{cluster}:
+To create a group level cubble, use \code{switch_key()} with the new key variable, \code{cluster}:
 
@@
 
-With the group level cubble, \code{get_centroid()} is useful to compute
-the centroid of each cluster, which will be used as the major axis for
-the glyph map later:
+With the group level cubble, \code{get_centroid()} is useful to compute the centroid of each cluster, which will be used as the major axis for the glyph map later:
 
@@
 
-Long form cubble at both levels can be accessed through stretching the
-nested form and with access to both station and cluster level cubbles,
-various plots can be made to understand the cluster. Figure
-\ref{fig:basic-agg} shows two example plots that can be made with this
-data: subplot A is a glyph map made with the cluster level cubble in the
-long form and subplot B inspects the station membership of each cluster
-using the station level cubble in the nested form.
+Long form cubble at both levels can be accessed through stretching the nested form and with access to both station and cluster level cubbles, various plots can be made to understand the cluster.
+Figure \ref{fig:basic-agg} shows two example plots that can be made with this data: subplot A is a glyph map made with the cluster level cubble in the long form and subplot B inspects the station membership of each cluster using the station level cubble in the nested form.
 
@@
 \hypertarget{river-level-data-in-victoria-water-gauges}{%
-\subsection{River level data in Victoria water
-gauges}\label{river-level-data-in-victoria-water-gauges}}
-
-Bureau of Meteorology collects
-\href{http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full}{water
-data} from river gauges and this includes variables: electrical
-conductivity, turbidity, water course discharge, water course level, and
-water temperature. In particular, water level will interactive with
-precipitation from the climate data since rainfall will raise the water
-level in the river. Figure \ref{fig:matching-map} gives the location of
-available weather station and water gauges in Victoria.
+\subsection{River level data in Victoria water gauges}\label{river-level-data-in-victoria-water-gauges}}
+
+Bureau of Meteorology collects \href{http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full}{water data} from river gauges and this includes variables: electrical conductivity, turbidity, water course discharge, water course level, and water temperature.
+In particular, water level will interactive with precipitation from the climate data since rainfall will raise the water level in the river.
+Figure \ref{fig:matching-map} gives the location of available weather station and water gauges in Victoria.
 
@@
 
-From the map, a few water gauges and weather stations are close to each
-other and the fluctuation of the water level could be matched up with
-precipitation measured by the climate station. As introduced in Section
-3.2, \code{match_sites()} can be used to match one source of data with
-another source in a cubble. Here \code{Water_course_level} in
-\code{river} will be matched to \code{prcp} in \code{climate} in 2020.
-The two datasets need to be specified as the first two arguments and the
-variable to match can be specified in \code{temporal_by} using the
-\code{by} syntax in \code{join}. \code{temporal_independent} controls
-the variable used to construct the interval and the goal here is to see
-if precipitation will be reflected by the water level in the river. This
-puts precipitation \code{prcp}, as the independent. Given there is one
-year worth of data, the number of peak (\code{temporal_n_highest}) to
-consider is slightly raised from a default 20 to 30 and
-\code{temporal_min_match} is raised accordingly To return all the pairs
-of the match, \code{temporal_min_match} can be set to 0.
+From the map, a few water gauges and weather stations are close to each other and the fluctuation of the water level could be matched up with precipitation measured by the climate station.
+As introduced in Section 3.2, \code{match_sites()} can be used to match one source of data with another source in a cubble.
+Here \code{Water_course_level} in \code{river} will be matched to \code{prcp} in \code{climate} in 2020.
+The two datasets need to be specified as the first two arguments and the variable to match can be specified in \code{temporal_by} using the \code{by} syntax in \code{join}.
+\code{temporal_independent} controls the variable used to construct the interval and the goal here is to see if precipitation will be reflected by the water level in the river.
+This puts precipitation \code{prcp}, as the independent.
+Given there is one year worth of data, the number of peak (\code{temporal_n_highest}) to consider is slightly raised from a default 20 to 30 and \code{temporal_min_match} is raised accordingly To return all the pairs of the match, \code{temporal_min_match} can be set to 0.
 
@@
 
-The output from matching is also a cubble, with additional column
-\code{dist} and \code{group} produced from spatial matching and
-\code{n_match} from temporal matching.
+The output from matching is also a cubble, with additional column \code{dist} and \code{group} produced from spatial matching and \code{n_match} from temporal matching.
 
@@
 
-Figure \ref{fig:matching} plots the matched pairs on the map or to view
-the matched series. There are four pairs of matches, which all locates
-in the middle Victoria and the concurrent increase of precipitation and
-water level can be observed.
+Figure \ref{fig:matching} plots the matched pairs on the map or to view the matched series.
+There are four pairs of matches, which all locates in the middle Victoria and the concurrent increase of precipitation and water level can be observed.
 
@@
 \hypertarget{era5-climate-reanalysis-data}{%
-\subsection{ERA5: climate reanalysis
-data}\label{era5-climate-reanalysis-data}}
-
-ERA5 data \citep{hersbach2020era5} is the latest reanalysis of global
-atmosphere, land surface, and ocean waves from 1950 onwards and is
-available in the NetCDF format from European Centre for Medium-Range
-Weather Forecasts (ECMWF). The data can be directly downloaded from
-\href{https://cds.climate.copernicus.eu/cdsapp\#!/dataset/reanalysis-era5-pressure-levels?tab=overview}{Copernicus
-Climate Data Store (CDS)} website or programmatically via an R package
-\pkg{ecmwfr} \citep{ecwmfr}. The \code{era5-pressure} data contains
-variable \emph{specific humidity} and \emph{geopotential} on the 10 hPa
-pressure level on four dates: 2002-09-22, 2002-09-26, 2002-09-30, and
-2002-10-04. Once downloaded, the data can be read into a cubble as:
+\subsection{ERA5: climate reanalysis data}\label{era5-climate-reanalysis-data}}
+
+ERA5 data \citep{hersbach2020era5} is the latest reanalysis of global atmosphere, land surface, and ocean waves from 1950 onwards and is available in the NetCDF format from European Centre for Medium-Range Weather Forecasts (ECMWF).
+The data can be directly downloaded from \href{https://cds.climate.copernicus.eu/cdsapp\#!/dataset/reanalysis-era5-pressure-levels?tab=overview}{Copernicus Climate Data Store (CDS)} website or programmatically via an R package \pkg{ecmwfr} \citep{ecwmfr}.
+The \code{era5-pressure} data contains variable \emph{specific humidity} and \emph{geopotential} on the 10 hPa pressure level on four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04.
+Once downloaded, the data can be read into a cubble as:
 
@@
 
-Figure \ref{fig:netcdf} reproduces the ERA5 data row of Figure 19 in
-\citet{hersbach2020era5}. It shows the southern polar vortex splits into
-two on 2002-09-26 and further splits into four on 2002-10-04 in the
-stratosphere. Readers interested in the analysis of this figure can
-refer to \citet{hersbach2020era5}, \citet{simmons2020global} and
-\citet{simmons2005ecmwf} for more details.
+Figure \ref{fig:netcdf} reproduces the ERA5 data row of Figure 19 in \citet{hersbach2020era5}.
+It shows the southern polar vortex splits into two on 2002-09-26 and further splits into four on 2002-10-04 in the stratosphere.
+Readers interested in the analysis of this figure can refer to \citet{hersbach2020era5}, \citet{simmons2020global} and \citet{simmons2005ecmwf} for more details.
 
@@
 
-With spatio-temporal data, users may wish to make plots to learn the
-spatial distribution of a variable, or to find patterns, such as trend
-or seasonality, in the time series. Combining this two types of plot
-with interactivity let users to link between points on the map and the
-corresponding time series to explore the spatial and temporal dimension
-of the data simultaneously. Below is an example that describes the
-process of building an interactive graphic with \pkg{cubble} and
-\pkg{crosstalk} The example explores the variation of monthly
-temperature range with \code{weatherdata::climate_full} data.
-
-The temperature range is calculated as the difference between
-\code{tmax} and \code{tmin} and its monthly average over 2016 - 2020 is
-taken before calculating the variance. A \code{SharedData} object is
-constructed for each form of the cubble and the same \code{group}
-argument ensures the cross-linking of the two forms via the common
-\code{id} column. The spatial map and time series plot are then made
-with each \code{SharedData} objects separately. In this example,
-stations on the Australia map, made from the nested form, are coloured
-by the calculated variance and a ribbon band is constructed using the
-long form cubble to show the maximum and minimum temperature of each
-station across month. With a different dataset, users are free to
-calculate any per station measure in the nested form or to make any
-time-wise summary of the data in the long form to customise the spatial
-or temporal view. The cross-linking between the two plots is always
-safeguarded by the shared \code{id} column embedded in the cubble
-structure. Below is the pseudo code that outlines the process to
-construct an interactive graphic described above:
+With spatio-temporal data, users may wish to make plots to learn the spatial distribution of a variable, or to find patterns, such as trend or seasonality, in the time series.
+Combining this two types of plot with interactivity let users to link between points on the map and the corresponding time series to explore the spatial and temporal dimension of the data simultaneously.
+Below is an example that describes the process of building an interactive graphic with \pkg{cubble} and \pkg{crosstalk} The example explores the variation of monthly temperature range with \code{weatherdata::climate_full} data.
+
+The temperature range is calculated as the difference between \code{tmax} and \code{tmin} and its monthly average over 2016 - 2020 is taken before calculating the variance.
+A \code{SharedData} object is constructed for each form of the cubble and the same \code{group} argument ensures the cross-linking of the two forms via the common \code{id} column.
+The spatial map and time series plot are then made with each \code{SharedData} objects separately.
+In this example, stations on the Australia map, made from the nested form, are coloured by the calculated variance and a ribbon band is constructed using the long form cubble to show the maximum and minimum temperature of each station across month.
+With a different dataset, users are free to calculate any per station measure in the nested form or to make any time-wise summary of the data in the long form to customise the spatial or temporal view.
+The cross-linking between the two plots is always safeguarded by the shared \code{id} column embedded in the cubble structure.
+Below is the pseudo code that outlines the process to construct an interactive graphic described above:
 
@@
 
-In Figure \ref{fig:interactive-linking}, the first row shows the initial
-view of the interactive graphic. On the map, most regions in Australia
-have low variance of temperature range while the north-west coastline,
-bottom of South Australia, and Victoria stands out with larger monthly
-changes. In the second row, Mount Elizabeth is selected on the map given
-its high variance colour on the initial map and this links to the ribbon
-on the right. The third row the lowest temperature in August and this
-corresponds to Thredbo AWS in the Victoria and New South Wales border.
-Another station in the Tasmania island is selected on the map to cross
-compare with Thredbo AWS.
+In Figure \ref{fig:interactive-linking}, the first row shows the initial view of the interactive graphic.
+On the map, most regions in Australia have low variance of temperature range while the north-west coastline, bottom of South Australia, and Victoria stands out with larger monthly changes.
+In the second row, Mount Elizabeth is selected on the map given its high variance colour on the initial map and this links to the ribbon on the right.
+The third row the lowest temperature in August and this corresponds to Thredbo AWS in the Victoria and New South Wales border.
+Another station in the Tasmania island is selected on the map to cross compare with Thredbo AWS.
 
@@
 
-This plot can also be made using \pkg{cubble} and \pkg{leaflet} where
-the temperature range can be displayed as a small subplot upon clicking
-on the map. This would require first creating the popup plots from the
-long form cubble as a vector and then add these plots to a leaflet map
-created from the nested cubble, with \code{leafpop::addPopupGraphs()}:
+This plot can also be made using \pkg{cubble} and \pkg{leaflet} where the temperature range can be displayed as a small subplot upon clicking on the map.
+This would require first creating the popup plots from the long form cubble as a vector and then add these plots to a leaflet map created from the nested cubble, with \code{leafpop::addPopupGraphs()}:
 
@@
 
-Figure \ref{fig:interactive-popup} shows Figure
-\ref{fig:interactive-linking} made made with leaflet and popups
-\citep{leafpop}.
+Figure \ref{fig:interactive-popup} shows Figure \ref{fig:interactive-linking} made made with leaflet and popups \citep{leafpop}.
 
@@
 
-\hypertarget{conclusion}{%
-\section{Conclusion}\label{conclusion}}
-
-This paper describes an \proglang{R} package \pkg{cubble} for
-manipulating and visualising spatio-temporal data. A new data structure,
-\code{cubble} that builds from the \code{rowwise_df} and
-\code{grouped_df} class in the tidyverse ecosystem, is proposed to
-connect the time invariant and varying variables in the spatio-temporal
-data. This design frees the data analysts from spending time on
-organising variables of different observational units. The data
-structure is also flexible to the techniques and packages analysts use
-to analyse the data, for example, in the matching example in section
-4.3, users are free to use algorithms from another package to cluster
-stations.
-
-Further development and maintenance of the package involves responding
-to changes in the tidyverse packages that \pkg{cubble} imports, in
-particular, \pkg{tibble}, \pkg{tidyr}, and \pkg{dplyr}. Another area for
-further development is to extend \pkg{cubble} to other spatial objects
-other than points.
+\hypertarget{conclude}{%
+\section{Conclusion}\label{conclude}}
+
+This paper describes an \proglang{R} package \pkg{cubble} for manipulating and visualising spatio-temporal data.
+A new data structure, \code{cubble} that builds from the \code{rowwise_df} and \code{grouped_df} class in the tidyverse ecosystem, is proposed to connect the time invariant and varying variables in the spatio-temporal data.
+This design frees the data analysts from spending time on organising variables of different observational units.
+The data structure is also flexible to the techniques and packages analysts use to analyse the data, for example, in the matching example in section 4.3, users are free to use algorithms from another package to cluster stations.
+
+Further development and maintenance of the package involves responding to changes in the tidyverse packages that \pkg{cubble} imports, in particular, \pkg{tibble}, \pkg{tidyr}, and \pkg{dplyr}.
+In the spatial aspect, the simple feature representation wraps spatial coordinates in a list column, for some complex spatial operations, while sometimes, analysts may want to unpack them into longitude and latitude columns to work from.
+Another extension to \pkg{cubble} is to build a smooth transition between the coordinate columns and simple feature geometry column.
 
@@
 
-This work is funded by the Commonwealth Scientific and Industrial
-Research Organisation (CSIRO) Data61 Scholarship and started while
-Nicolas Langrené was affiliated with CSIRO's Data61. The article is
-created using \pkg{knitr} \citep{knitr} and \pkg{rmarkdown}
-\citep{rmarkdown} in R. The source code for reproducing this paper can
-be found at: \url{https://github.com/huizezhang-sherry/paper-cubble}.
+This work is funded by the Commonwealth Scientific and Industrial Research Organisation (CSIRO) Data61 Scholarship and started while Nicolas Langrené was affiliated with CSIRO's Data61.
+The article is created using \pkg{knitr} \citep{knitr} and \pkg{rmarkdown} \citep{rmarkdown} in R.
+The source code for reproducing this paper can be found at: \url{https://github.com/huizezhang-sherry/paper-cubble}.
 
@@
 
-\bibliography{references.bib}
+\bibliography{../references.bib}
 

### 2022-04-19 1-intro.Rmd
some fix to intro

@@
 Spatio-temporal data has a spatial component referring to the location of each observation, a temporal component that is recorded at regular or irregular time intervals, and multiple variables measured at each spatial and temporal value.
-In data collection, spatio-temporal data is less often collected as a whole piece but fragmented spatial and temporal data for analysts to ensemble into spatio-temporal data.
-With the [integrated] data, we hope to fix the time to explore the spatial distribution, or fix the spaital location to explore temporal trends.
-Ideally, we would be able to carry out each of these tasks easily, and even accommodate complex explorations encompassing both simultaneously.
-These are all underpinned by a [flexible] data structure that can represent spatio-temporal data.
+The analysis of spatio-temporal data can fix the time to explore the spatial distribution, fix the spatial location to explore temporal trends, or dynamically explore the space and time simultaneously.
+For these tasks to be carried out easily, there is a need for a spatio-temporal data representation that underpins various operations.
 
@@
 
-What remains to be difficult when wrangling spatio-temporal data is that the raw data is not naturally structured in any one of the representation mentioned above and the first difficulty is to cast the data, potentially from multiple sources into one of the data representation.
-Some examples of these includes map data recorded the shape of a collection of area of interest is provided by the government website; geo-scientific data recorded with longitude and latitude coordinate of locations in the area; and temporal variables indexed by both location and time for the temporal aspect.
+What remains to be difficult when wrangling spatio-temporal data is that the raw data is not naturally structured in any one of the representation above and the first task is to cast the data into a class.
+Also, in data collection, spatio-temporal data is less often collected as a whole piece but fragmented spatial and temporal data for analysts to ensemble.
+For example, map data records the shape of a collection of area of interest is provided by official government website; geo-scientific data records the longitude and latitude coordinate of locations in the area along with other meta data relates to the location; and temporal variables indexed by both location and time.
 For spatio-temporal data analysis, merging all these sources can be challenging and joining these tables can go wrong because of the slightest unmatched of the linking variable from different sources.
 
-Another issue is how the concept of tidy data should be applied to spatio-temporal data.
+Another issue is how the concept of tidy data [@tidydata] should be applied to spatio-temporal data.
 The tidy data concept advocates the data to be shaped into 1) one row per observation; 2) one column per variable; and 3) one type of data per table.
-The long form data is preferred over wide data given the downstream data cleaning and visualisation software i.e.
-\pkg{dplyr} and \pkg{ggplot2}.
-However, the long form can be inefficient to store feature geometries, especially for large multipolygon over long time period, i.e.
-sub-daily time series data.
+The long form data is preferred over wide data given the downstream software i.e.
+\pkg{dplyr} and \pkg{ggplot2} for data wrangling and visualisation.
+However, the long form can be inefficient to store feature geometries, especially for large multipolygon over long time period.
+This poses the question on how to arrange spatial and temporal variables that makes analysing spatio-temporal data easier.
 

### 2022-04-19 4-1-basic.Rmd
add example to cast covid data into cubble

@@
 # Examples {#examples}
+
+## Victoria covid data
+
+This example illustrate casting different sources into a cubble object.
+Victoria government of Australia provides daily COVID case number by date, source, and local government area (LGA).
+This data can be useful for visualising COVID spread when combined with LGA map information, which is available from Australian Bureau of Statistics.
+Variable \code{covid} and \code{lga} show the first 5 rows of these two data:
+
+```{r eval = FALSE}
+library(tidyverse)
+#https://www.coronavirus.vic.gov.au/victorian-coronavirus-covid-19-data
+raw <- read_csv(here::here("data/NCOV_COVID_Cases_by_LGA_Source_20220324.csv"), 
+                  col_names = c("date", "postcode", "source", "lga"), skip = 1)
+
+covid <- raw %>%
+  count(date, lga, source, postcode) %>%
+  filter(date >= as.Date("2022-01-01")) %>% 
+  filter(source == "Contact with a confirmed case") %>% 
+  group_by(date, lga, source) %>%
+  count() %>%
+  ungroup(date) %>%
+  mutate(roll_mean = zoo::rollmean(n, k = 14, fill = NA)) %>% 
+  tsibble::as_tsibble(key = lga, index = date) %>% 
+  tsibble::fill_gaps(.full = TRUE)
+save(covid, file = here::here("data/covid.rda"))
+```
+
+```{r}
+load(here::here("data/covid.rda"))
+lga <- absmapsdata::lga2018 %>%
+  rename(lga = lga_name_2018) %>% 
+  filter(state_name_2016 == "Victoria") %>% 
+  rmapshaper::ms_simplify()
+```
+
+```{r echo = TRUE}
+covid |> head(5)
+lga |> head(5)
+```
+
+To cast these two data into a cubble object, supply them as a list with name \code{spatial} and \code{temporal} and other arguments (\code{key}, \code{index}, and \code{coords}).
+\code{as_cubble()} will automatically check the match of spatial and temporal information in the two datasets and warnings will emit if a location has missing spatial or temporal information.
+
+```{r echo = TRUE, warning = TRUE, message= TRUE}
+cb <- as_cubble(list(spatial = lga, temporal = covid),
+                key = lga, index = date, coords = c(cent_long, cent_lat))
+```
+
+Here, some locations have been detected with this issue and you can use \code{output = "unmatch"} to check on these locations:
+
+```{r echo = TRUE}
+pair <- as_cubble(list(spatial = lga, temporal = covid),
+                key = lga, index = date, coords = c(cent_long, cent_lat),
+                output = "unmatch")
+
+pair
+```
+
+\pkg{cubble} will make an attempt to pair the unmatched sites as well as showing unpaired locations. These can be useful information to help clean up the data.
+
+```{r echo = TRUE}
+lga <- lga %>% 
+  mutate(lga = ifelse(lga == "Kingston (C) (Vic.)", "Kingston (C)", lga),
+         lga = ifelse(lga == "Latrobe (C) (Vic.)", "Latrobe (C)", lga)) %>%
+  filter(!lga %in% pair$others$spatial)
+
+covid <- covid %>% filter(!lga %in% pair$others$temporal)
+```
+
+After cleaning, you can cast the data into cubble with the same syntax without warnings:
+
+```{r echo = TRUE}
+cb <- as_cubble(data = list(spatial = lga, temporal = covid),
+                key = lga, index = date, coords = c(cent_long, cent_lat))
+```
 

### 2022-04-19 jss/cubble-jss.Rmd
a new skeleton of intro incorporating various past writing
add example to cast covid data into cubble

@@
   \usepackage{xcolor}
-output: rticles::jss_article
-bibliography: references.bib
+output: 
+  bookdown::pdf_book:
+    base_format: rticles::jss_article
+bibliography: ../references.bib
 ---
@@
 )
+
 options(prompt = "R> ", continue = "+ ",

### 2022-04-19 jss/cubble-jss.tex
add example to cast covid data into cubble

@@
 Spatio-temporal data has a spatial component referring to the location of each observation, a temporal component that is recorded at regular or irregular time intervals, and multiple variables measured at each spatial and temporal value.
-In data collection, spatio-temporal data is less often collected as a whole piece but fragmented spatial and temporal data for analysts to ensemble into spatio-temporal data.
-With the {[}integrated{]} data, we hope to fix the time to explore the spatial distribution, or fix the spaital location to explore temporal trends.
-Ideally, we would be able to carry out each of these tasks easily, and even accommodate complex explorations encompassing both simultaneously.
-These are all underpinned by a {[}flexible{]} data structure that can represent spatio-temporal data.
+The analysis of spatio-temporal data can fix the time to explore the spatial distribution, fix the spatial location to explore temporal trends, or dynamically explore the space and time simultaneously.
+For these tasks to be carried out easily, there is a need for a spatio-temporal data representation that underpins various operations.
 
@@
 
-What remains to be difficult when wrangling spatio-temporal data is that the raw data is not naturally structured in any one of the representation mentioned above and the first difficulty is to cast the data, potentially from multiple sources into one of the data representation.
-Some examples of these includes map data recorded the shape of a collection of area of interest is provided by the government website; geo-scientific data recorded with longitude and latitude coordinate of locations in the area; and temporal variables indexed by both location and time for the temporal aspect.
+What remains to be difficult when wrangling spatio-temporal data is that the raw data is not naturally structured in any one of the representation above and the first task is to cast the data into a class.
+Also, in data collection, spatio-temporal data is less often collected as a whole piece but fragmented spatial and temporal data for analysts to ensemble.
+For example, map data records the shape of a collection of area of interest is provided by official government website; geo-scientific data records the longitude and latitude coordinate of locations in the area along with other meta data relates to the location; and temporal variables indexed by both location and time.
 For spatio-temporal data analysis, merging all these sources can be challenging and joining these tables can go wrong because of the slightest unmatched of the linking variable from different sources.
 
-Another issue is how the concept of tidy data should be applied to spatio-temporal data.
+Another issue is how the concept of tidy data \citep{tidydata} should be applied to spatio-temporal data.
 The tidy data concept advocates the data to be shaped into 1) one row per observation; 2) one column per variable; and 3) one type of data per table.
-The long form data is preferred over wide data given the downstream data cleaning and visualisation software i.e.~\pkg{dplyr} and \pkg{ggplot2}.
-However, the long form can be inefficient to store feature geometries, especially for large multipolygon over long time period, i.e.~sub-daily time series data.
+The long form data is preferred over wide data given the downstream software i.e.~\pkg{dplyr} and \pkg{ggplot2} for data wrangling and visualisation.
+However, the long form can be inefficient to store feature geometries, especially for large multipolygon over long time period.
+This poses the question on how to arrange spatial and temporal variables that makes analysing spatio-temporal data easier.
 
@@
 \end{CodeChunk}
+
+\hypertarget{why-not-just-use-dplyrnestby}{%
+\subsection{Why not just use dplyrnestby}\label{why-not-just-use-dplyrnestby}}
 
@@
 
+\hypertarget{victoria-covid-data}{%
+\subsection{Victoria covid data}\label{victoria-covid-data}}
+
+This example illustrate casting different sources into a cubble object.
+Victoria government of Australia provides daily COVID case number by date, source, and local government area (LGA).
+This data can be useful for visualising COVID spread when combined with LGA map information, which is available from Australian Bureau of Statistics.
+Variable \code{covid} and \code{lga} show the first 5 rows of these two data:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> covid |> head(5)
+\end{CodeInput}
+\begin{CodeOutput}
+# A tsibble: 5 x 5 [1D]
+# Key:       lga [1]
+# Groups:    lga, source [1]
+  date       lga        source                            n roll_mean
+  <date>     <chr>      <chr>                         <int>     <dbl>
+1 2022-01-01 Alpine (S) Contact with a confirmed case     1        NA
+2 2022-01-02 Alpine (S) Contact with a confirmed case     2        NA
+3 2022-01-03 Alpine (S) Contact with a confirmed case     4        NA
+4 2022-01-04 Alpine (S) Contact with a confirmed case     4        NA
+5 2022-01-05 Alpine (S) Contact with a confirmed case     2        NA
+\end{CodeOutput}
+\begin{CodeInput}
+R> lga |> head(5)
+\end{CodeInput}
+\begin{CodeOutput}
+Simple feature collection with 5 features and 7 fields
+Geometry type: MULTIPOLYGON
+Dimension:     XY
+Bounding box:  xmin: 142.3643 ymin: -38.67876 xmax: 147.386 ymax: -36.39269
+Geodetic CRS:  WGS 84
+  lga_code_2018            lga state_code_2016 state_name_2016 areasqkm_2018
+1         20110     Alpine (S)               2        Victoria     4788.1568
+2         20260    Ararat (RC)               2        Victoria     4211.1171
+3         20570   Ballarat (C)               2        Victoria      739.0321
+4         20660    Banyule (C)               2        Victoria       62.5402
+5         20740 Bass Coast (S)               2        Victoria      865.8095
+  cent_long  cent_lat                       geometry
+1  146.9742 -36.85357 MULTIPOLYGON (((146.7258 -3...
+2  142.8432 -37.47271 MULTIPOLYGON (((143.1807 -3...
+3  143.7815 -37.49286 MULTIPOLYGON (((143.6622 -3...
+4  145.0851 -37.73043 MULTIPOLYGON (((145.1357 -3...
+5  145.5581 -38.50730 MULTIPOLYGON (((145.5207 -3...
+\end{CodeOutput}
+\end{CodeChunk}
+
+To cast these two data into a cubble object, supply them as a list with name \code{spatial} and \code{temporal} and other arguments (\code{key}, \code{index}, and \code{coords}).
+\code{as_cubble()} will automatically check the match of spatial and temporal information in the two datasets and warnings will emit if a location has missing spatial or temporal information.
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> cb <- as_cubble(list(spatial = lga, temporal = covid),
++                 key = lga, index = date, coords = c(cent_long, cent_lat))
+\end{CodeInput}
+\begin{CodeOutput}
+! Some sites in the temporal table don't have corresponding spatial information
+\end{CodeOutput}
+\begin{CodeOutput}
+! Some sites in the spatial table don't have corresponding temporal information
+\end{CodeOutput}
+\begin{CodeOutput}
+! Use argument `output = "unmatch"` to check on the unmatched key
+\end{CodeOutput}
+\end{CodeChunk}
+
+Here, some locations have been detected with this issue and you can use \code{output = "unmatch"} to check on these locations:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> pair <- as_cubble(list(spatial = lga, temporal = covid),
++                 key = lga, index = date, coords = c(cent_long, cent_lat),
++                 output = "unmatch")
+R> 
+R> pair
+\end{CodeInput}
+\begin{CodeOutput}
+$paired
+# A tibble: 2 x 2
+  spatial             temporal    
+  <chr>               <chr>       
+1 Kingston (C) (Vic.) Kingston (C)
+2 Latrobe (C) (Vic.)  Latrobe (C)
+
+$others
+$others$temporal
+[1] "Interstate"       "Overseas"         "Queenscliffe (B)" "Unknown"
+
+$others$spatial
+character(0)
+\end{CodeOutput}
+\end{CodeChunk}
+
+\pkg{cubble} will make an attempt to pair the unmatched sites as well as showing unpaired locations. These can be useful information to help clean up the data.
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> lga <- lga %>% 
++   mutate(lga = ifelse(lga == "Kingston (C) (Vic.)", "Kingston (C)", lga),
++          lga = ifelse(lga == "Latrobe (C) (Vic.)", "Latrobe (C)", lga)) %>%
++   filter(!lga %in% pair$others$spatial)
+R> 
+R> covid <- covid %>% filter(!lga %in% pair$others$temporal)
+\end{CodeInput}
+\end{CodeChunk}
+
+After cleaning, you can cast the data into cubble with the same syntax without warnings:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> cb <- as_cubble(data = list(spatial = lga, temporal = covid),
++                 key = lga, index = date, coords = c(cent_long, cent_lat))
+\end{CodeInput}
+\end{CodeChunk}
+
 \hypertarget{australia-historical-maximum-temperature}{%

### 2022-04-20 2-the-cubble-pkg.Rmd
a new skeleton of intro incorporating various past writing
why not nest_by

@@
 
-# The cubble package
+# The cubble package {#cubble}
 
@@
 
-<!-- ## Why not just use dplyrnest_by -->
+<!-- https://www.jstatsoft.org/style#my-latex-paper-does-not-compile-when-there-is-jss-markup-in-section-titles-what-should-i-do -->
+\subsection[Why not just use]{Why not just use \code{dplyr::nest\_by()}}
 
-<!-- ```{r eval = FALSE} -->
+It may naturally raise the question that why bother a new data structure since the nested form can already be created by the \pkg{dplyr} function \code{nest_by()}?
+While \code{nest_by()} and \code{unnest()} can mimic the nested form in cubble, cubble can be seen as an attempt to arrange variables from different observational units into a single object, specifically for spatio-temporal data.
+While `nest_by()` solves this problem on the spatial side, it doesn't solve the temporal side.
+For temporal data, a combined representation with spatial variables is useful for spatio-temporal operations but its duplication can be a burden to operate on.
+Also, spatial variables are dropped when summarised and you will need to either specified all of them in the \code{group_by()} before \code{summarise}, or add a joining after \code{summarise()}.
+
+<!-- ```{r} -->
+<!-- library(dplyr) -->
+<!-- vars <- c(id, lat, long, elev, name, wmo_id) -->
+<!-- left_join(cubble::stations, cubble::climate) |>  -->
+<!--   group_by(id, lat, long, elev, name, wmo_id) |>  -->
+<!--   summarise(tmax_mean = mean(tmax)) -->
+
+<!-- cubble::climate |>  -->
+<!--   mutate(month = lubridate::month(date)) |>  -->
+<!--   group_by(id, month) |>  -->
+<!--   summarise(tmax_mean = mean(tmax)) |>  -->
+<!--   ungroup() |>  -->
+<!--   left_join(cubble::stations) |>  -->
+<!--   nest_by(id, lat, long, elev, name, wmo_id) -->
+
+<!-- ``` --
+
+<!-- ``` r -->
 <!-- library(cubble) -->
@@
 <!--   group_by(........, geometry) -->
-<!-- ``` -->
-
-<!-- ```{r eval = FALSE} -->
-<!-- as_cubble(list(spatial = stations, temporal = climate), -->
-<!--           key = id, index = date, coords = c(long, lat)) -->
-<!-- ``` -->
-
-<!-- ```{r eval = FALSE} -->
-<!-- usmap::us_map("states") -->
 <!-- ``` -->

### 2022-04-20 jss/cubble-jss.tex
why not nest_by

@@
 
-\hypertarget{why-not-just-use-dplyrnestby}{%
-\subsection{Why not just use dplyrnestby}\label{why-not-just-use-dplyrnestby}}
+\subsection[Why not just use]{Why not just use \code{dplyr::nest\_by()}}
+
+It may naturally raise the question that why bother a new data structure since the nested form can already be created by the \pkg{dplyr} function \code{nest_by()}?
+While \code{nest_by()} and \code{unnest()} can mimic the nested form in cubble, cubble can be seen as an attempt to arrange variables from different observational units into a single object, specifically for spatio-temporal data.
+While \texttt{nest\_by()} solves this problem on the spatial side, it doesn't solve the temporal side.
+For temporal data, a combined representation with spatial variables is useful for spatio-temporal operations but its duplication can be a burden to operate on.
+Also, spatial variables are dropped when summarised and you will need to either specified all of them in the \code{group_by()} before \code{summarise}, or add a joining after \code{summarise()}.
 

### 2022-04-20 2-the-cubble-pkg.Rmd
work on section 3 the cubble package

@@
 
-This section will present two formats that cubble uses to arrange spatio-temporal data.
-The main cubble functions will then be introduced to illustrate how to work with these two formats in short examples.
-Lastly, a subsection will be dedicated to how existing packages, in spatial and temporal analysis, fit in with cubble.
+This section will first introduce the core functions in the cubble class: \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and \code{unfold()} from subsection \@ref(create) to \@ref(unfold).
+The next subsection addresses why bother a new class before subsection \@ref(compact) finishing on the compatibility of cubble with existing packages in spatial and temporal analysis.
 
-In cubble, a single data object can be arranged into two forms: nested form and long form.
-Figure \ref{fig:illu-cubble} sketches the two forms with the associated attributes.
-The decision on which form to use in data wrangling is output-oriented, meaning analysts need to first think about whether the output of an operation is identified only by the spatial identifier, or a combination of spatial and temporal identifier.
-The nested cubble is suitable for working with operations that are only identified by the site and this type of operation can be a pure manipulation of time invariant variables, or an operation that summarises time varying variables into site.
-Underneath the nested form, a cubble is built from a \code{rowwise_df} class where each site forms a separate group.
-This structure simplifies the calculation that involves temporal variables by avoiding the use of \code{purrr::map()} syntax when working with list-column.
+Each core function is introduced with a short example accompanied and the data `climate_flat` is used throughout this section.
+`climate_flat` contains five weather stations in Australia with spatial information of each station: station id, latitude, longitude, elevation, station name and World Meteorology Organisation ID and also daily temporal information of date, maximum and minimum temperature and precipitation for 2020.
+Below prints out the first five rows of `climate_flat`:
 
-For those operations whose output involves both a spatial and temporal identifier, long form should be used.
-The long form uses a \code{grouped_df} class to forms all the time of a site as a group.
-Time invariant variables are stored separately as an special attribute of the long cubble.
-This design avoids repeating the spatial variables at each time stamp while not dropping information from spatial variables.
-
-```{r illu-cubble, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "An illustration of the nested and long form in cubble.
-The nested form defines each site in a row and nests the time varying variables into a single column \\code{ts}.
-The long form cubble uses \\code{id} and \\code{t} to identify each row and stores the time invariant variables as an attribute, \\code{spatial}."}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.002.png"))
+```{r}
+cubble::climate_flat 
 ```
 
-## Create a cubble in the nested form
+## Create a cubble in the nested form {#create}
 
-To use functions in the `cubble` package, an analyst will first need to turn the data object into a \code{cubble} class.
-\code{as_cubble()} does this through supplying the three key components: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order of longitude and latitude.
-The use of \code{key} and \code{index} follows the design in the \pkg{tsibble} package and the cubble created by default is in the nested form.
-
-Before the formal examples in Section \ref{examples}, each function introduced in this section will be accompanied by a small example.
-The data used for the small examples is a subset of a larger climate weather station data used in the formal examples and has been simplified to include only five weather stations.
-It contains spatial information of each station: station id, latitude, longitude, elevation, station name and World Meteorology Organisation ID and also daily temporal information of date, maximum and minimum temperature and precipitation for 2020.
-\code{climate_flat} stores this data in format 1) in Figure \ref{fig:illu-input} and the code below creates a cubble out of \code{climate_flat} with \code{id} as the key, \code{date} as the index, and \code{c(long, lat)} as the coordinates:
+As described in the introduction, spatio-temporal data can come in various formats and shapes.
+A cubble can be created from various raw data, including tibble and its variates multiple tables (an example provided in section \@ref(covid)), and netcdf (detailed in section \@ref(netcdf)).
+The function \code{as_cubble()} is used to create a cubble with three additional components: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order of longitude and latitude.
+Argument \code{key} and \code{index} follow the wording in \pkg{tsibble} to define temporal order and multiple series while \code{coords} is a space specific argument to locate each site in two columns.
+The code below creates a cubble out of \code{climate_flat} with \code{id} as the key, \code{date} as the index, and \code{c(long, lat)} as the coordinates:
 
 ```{r echo = TRUE}
-cubble_nested <- cubble::climate_flat |>
+cubble_nested <- climate_flat |>
   as_cubble(key = id, index = date, coords = c(long, lat))
@@
 
-There are a few information in the cubble header: the name of the \code{key} variable: \code{id} and its unique number: 5, the bounding box: \code{[115.97, -32.94, 133.55, -12.42]} , and also the name of variable nested in the \code{ts} column with its type: \code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.
+The cubble header provides some information about this data.
+\code{id} is the variable name to identify each location and there are five unique locations.
+The bounding box is \code{[115.97, -32.94, 133.55, -12.42]} and the names of variable nested in the \code{ts} column, along with its type, are \code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.
 
-## Stretch a nested cubble into the long form
+The created cubble is built from a \code{rowwise_df} class where each row forms a group and all the temporal variables are nested in a list column, hence it is also called the nested cubble.
+The rowwise structure makes it simpler to calculate on the list using \code{mutate()} and avoid the \code{map()} when working with list column.
+For example calculating the number of raining day can be done by:
+
+```{r}
+cubble_nested |> 
+  mutate(rain_day = sum(ts$prcp != 0))
+```
+
+## Switch from the nested cubble into the long form
 
@@
 For those outputs that are cross-identified by the spatial and temporal identifier, a long cubble needs to be used.
-The function \code{stretch()} is designed to switch the cubble from the nested form to the long form.
-This code shows how to switch the nested cubble just created into its long form:
+The function \code{face_temporal()} switches the cubble from the nested form to the long form and the first row in Figure \ref{fig:face} sketches the function in a diagram.
+This code shows how to switch the cubble just created into its long form:
 
@@
 
-## Tamp a long cubble back to the nested form
+The long form is built form class \code{groupped_df} where all the observations from the same sites forms a group.
+
+```{r face, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "An illustration of function \\code{face\\_temporal} and \\code{face\\_spatial} in cubble.
+In the first row, \\code{face\\_temporal} switches a cubble from the nested form into the long form and the focus has switched from the spatial aspect (the side face) to the temporal aspect (the front face).
+In the second row, \\code{face\\_spatial} switches a cubble back to the nested form from te long form and shift focus back to the spatial aspect."}
+knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.008.png"))
+```
+
+## Switch a long cubble back to the nested form
 
 Manipulation on the spatial and temporal dimension can be an iterative process and analysts may need to go back and forth between the nested and long cubble.
-\code{tamp()}, an inverse of \code{stretch()}, switches a long cubble to a nested cubble:
+\code{face_spatial()}, the inverse of \code{face_temporal()}, switches the long cubble back to the nested form, as sketched in the second row in Figure \ref{fig:face}:
 
@@
 
-## Migrate spatial variables to a long cubble
+## Flatten spatial variables into the long cubble {#unfold}
 
 Sometimes, analysts may need to apply some variable transformation that involves both the spatial and temporal variable.
-An example of this is the transformation of temporal variables into the spatial dimension in glyph maps (which will be elaborated in section \ref{st_transformation}).
-Cubble allows this operation through \code{unfold()}, which moves the specified spatial variables into the long cubble:
+An example of this is the transformation of temporal variables into the spatial dimension in glyph maps, which will be elaborated in section \ref{st_transformation}.
+This type of operations can be seen as flattening, or *unfolding*, the cube into a 2D data frame.
+Here the function \code{unfold()} moves the spatial variables \code{long} and \code{lat} into the long cubble:
 
 ```{r echo = TRUE}
-cubble_mig <- cubble_long |> unfold(long, lat)
-cubble_mig
+cubble_unfold <- cubble_long |> unfold(long, lat)
+cubble_unfold
 ```
@@
 ```{r echo = TRUE}
-cubble_mig |> face_spatial() |> face_temporal()
+cubble_unfold |> face_spatial() |> face_temporal()
 ```
@@
 
-## Compatibility with existing packages
+## Compatibility with existing packages {#compact}
 
@@
 \code{tsibble} is a temporal data structure that uses \code{index} and \code{key} to identify the time and different series.
-\code{cubble} can be seen as a natural extension of \code{tsibble} for spatio-temporal data with an additional coordinates component and using two forms to arrange variables.
+\code{cubble} can be seen as following the same vein as \code{tsibble} for spatio-temporal data.
 This makes it easy to cast a \code{tsibble} into a \code{cubble} as only the \code{coords} argument needs to be supplied:
@@
 # example with a tsibble created from climate_flat
-raw <- climate_flat |>
-  tsibble::as_tsibble(key = id, index = date)
-
-dt <-  raw |>
-  cubble::as_cubble(coords = c(long, lat))
+raw <- climate_flat |> tsibble::as_tsibble(key = id, index = date)
+dt <-  raw |> cubble::as_cubble(coords = c(long, lat))
 dt
@@
 # add station-based features in the nested form.
-dt |>
-  mutate(fabletools::features(ts, tmax, list(tmax_mean = mean, tmax_var = var)))
+dt |> mutate(fabletools::features(ts, tmax, list(tmax_mean = mean, tmax_var = var)))
 ```
@@
 
-### netcdf
+### netcdf {#netcdf}
 

### 2022-04-20 arxiv/cubble-arxiv.Rmd
work on section 3 the cubble package

@@
 bst: "../jss.bst"
-output: rticles::arxiv_article
+preamble: |
+  \newcommand{\sherry}[1]{\textcolor{magenta}{#1}}
+output: 
+  bookdown::pdf_book:
+    base_format: rticles::arxiv_article
 link-citations: true

### 2022-04-20 arxiv/cubble-arxiv.tex
work on section 3 the cubble package

@@
 
+% From pandoc table feature
+\usepackage{longtable,booktabs,array}
+\usepackage{calc} % for calculating minipage widths
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
+
 % Pandoc citation processing
@@
 Spatio-temporal data refer to measurements taken across space and time.
-In practice, spatio-temporal data can be decomposed into a spatial and
-temporal component: at one time, we would select a spatial location and
-inspect the temporal trend; at other time, we might select one or
-multiple time value(s) and explore the spatial distribution. Ideally, we
-could make multiple maps and multiple time series to explore these
-together, however, doing all of these actions is complicated when data
-arrive fragmented in multiple objects. To make it easy to do all these
-tasks, ideally spatial and temporal variables are in a single data
-object that we can slice and dice in different ways to conduct different
-visualisations. This work suggests a new data structure, \code{cubble},
-that uses a nested form and a long form to organise spatial and temporal
-variables in a single data object. The new data structure ensures that
-data in the two forms are synchronised while been flexible to explore
-the two aspect of spatio-temporal data. It also provides capability for
-handling data with hierarchical structure, matching data from multiple
-sources, constructing interactive graphics, and performing
-spatio-temporal transformation. The paper will demonstrate \pkg{cubble}
-with examples using Australian climate weather stations, river level
-data, and climate reanalysis (ERA5) data.
+In practice, spatio-temporal data can be decomposed into a spatial and temporal component: at one time, we would select a spatial location and inspect the temporal trend; at other time, we might select one or multiple time value(s) and explore the spatial distribution.
+Ideally, we could make multiple maps and multiple time series to explore these together, however, doing all of these actions is complicated when data arrive fragmented in multiple objects.
+To make it easy to do all these tasks, ideally spatial and temporal variables are in a single data object that we can slice and dice in different ways to conduct different visualisations.
+This work suggests a new data structure, \code{cubble}, that uses a nested form and a long form to organise spatial and temporal variables in a single data object.
+The new data structure ensures that data in the two forms are synchronised while been flexible to explore the two aspect of spatio-temporal data.
+It also provides capability for handling data with hierarchical structure, matching data from multiple sources, constructing interactive graphics, and performing spatio-temporal transformation.
+The paper will demonstrate \pkg{cubble} with examples using Australian climate weather stations, river level data, and climate reanalysis (ERA5) data.
 \end{abstract}
@@
 \keywords{
-    spatial, temporal, spatio temporal, R, environmental data,
-    exploratory data analysis
+    spatial, temporal, spatio temporal, R, environmental data, exploratory data analysis
   }
@@
 
-Spatio-temporal data involves variables observed in the space across
-time and this paper focuses on spatio-temporal vector data with fixed
-locations and regular time intervals. This type of data can naturally be
-considered as a data cube with three axes: spatial identifier, temporal
-identifier, and variables. In this cubic formation, space only occupies
-one axis, rather than two, to avoid hypercubes for multivariate
-spatio-temporal data. While conceptually the data format is relatively
-simple, the data objects coming to the analysts can vary in forms in the
-analysis. Figure \ref{fig:illu-input} shows three examples of wild
-spatio-temporal data: 1) a single table that includes both the spatial
-and temporal variables; 2) separate tables for spatial and temporal
-variables; 3) array-based data, i.e.~NetCDF (Network Common Data Form)
-data, which is commonly used in earth science, e.g.~climatology,
-meteorology, and oceanography, among others, to organise multivariate
-data in a spatio-temporal grid.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.001}
-
-}
-
-\caption{An illustration of different spatio-temporal data layouts: (1) a single table for both spatial and temporal variables; (2) separate tables for spatial and temporal variables; and (3) an array data structure that populates variables into a gridded longitude-latitude-time (hyper)cube.}\label{fig:illu-input}
-\end{figure}
-
-While incorporating both spatial and temporal information enriches the
-data in the analysis, spatio-temporal data can sometimes be hard to work
-with given its two observational units in the variables. The single
-table format in Figure \ref{fig:illu-input} is convenient for
-spatio-temporal operations that involve both types of variables,
-however, it makes duplicates of spatial variables and is not ideal when
-only looking at the spatial aspect of the data. The two tables format
-allows separate operations on spatial and temporal variables while
-analysts need to work with two data objects simultaneously and join them
-together into the single table format for spatio-temporal operations.
-Neither of the two formats is desirable at the same time for spatial,
-temporal, and spatio-temporal operations. This issue implies analysts
-need to constantly think about the data format and arrange variables
-into the proper shape before the actual operations, which can be a
-laborious step when switching between the two formats back and forth
-several times during the analysis.
-
-In the \proglang{R} community, many data objects have been proposed to
-work with specific spatio-temporal tasks, while few can freely explore
-the data. \pkg{spacetime} (\protect\hyperlink{ref-spacetime}{E. Pebesma
-2012}) is an attempt for such a data object, but the spatial and
-temporal classes it is built on have now been gradually replaced by more
-recent implementations. The \pkg{stars}
-(\protect\hyperlink{ref-stars}{E. Pebesma 2021}) package uses an array
-to wrangle spatio-temporal data, but the dense array it uses can quickly
-exhaust the memory for data in the case of irregular spatial grid. While
-there has not yet been a data object that makes it easy to wrangle
-spatio-temporal data, there is a strong need for such tools given the
-ubiquitous spatial and temporal data collected in the society, i.e.,
-climate observations from weather stations, air quality variables from
-air monitoring stations. The demand for such tools can also come from
-spatial data analysts wishing to incorporate temporal data in their
-analysis or vice versa.
-
-This paper presents a new \proglang{R} package, \pkg{cubble}, to
-overcome the problem with organising variables in different levels when
-working with spatio-temporal data. In the package a new data structure,
-also called \code{cubble}, is proposed to organise spatial and temporal
-variables as two forms of a single data object so that they can be
-wrangled separately while synchronised as a whole piece. With this new
-data object, analysts can spend less time writing codes to organise
-spatial and temporal variables and focus more on the exploratory data
-analysis itself. The software is available from the Comprehensive R
-Archive Network (CRAN) at {[}CRAN link{]}.
-
-The rest of the paper is divided as follows: Section 2 presents the main
-design and functionality of \pkg{cubble}. Section 3 explains how cubble
-deals with more advanced considerations including data with hierarchical
-structure, data matching, how cubble fits with existing static and
-interactive visualisation tools, and spatio-temporal data
-transformation. Section 4 uses Australia weather station data and river
-level data as examples to demonstrate the use of \pkg{cubble}. An
-example on how \pkg{cubble} handles NetCDF data is also provided.
-Section 5 concludes the paper.
-
-\hypertarget{conceptual-framework-spatio-temporal-cube}{%
-\section{Conceptual framework: spatio-temporal
-cube}\label{conceptual-framework-spatio-temporal-cube}}
-
-Multivariate spatio-temporal data can be conceptualised using a cubical
-data model with reference to variables, time and space. This naturally
-motivates the usage of array to represent and store spatio-temporal
-data, as evident by satellite imageries, large climate models, among
-others. The work by Lu, Appel, and Pebesma
-(\protect\hyperlink{ref-lu_multidimensional_2018}{2018}) has discussed
-the suitability of this representation through array algebra and
-practical spatio-temporal tasks.
-
-The cubical conceptual framework provides a generalisation to the
-operation can be done on the data for visualisation. A paper by Bach et
-al. (\protect\hyperlink{ref-bach_review_2014}{2014}) has reviewed the
-temporal data visualisation based on space-time cube operations. Notice
-that the term space-time cube in their article ``does not need to
-involve spatial data'', but referring to ``an abstract 2D substrate that
-is used to visualize data at a specific time''. Despite its main focus
-is on temporal data, the mindset of abstracting out data representation
-to construct visualisation is applicable to our spatio-temporal data
-manipulation and visualisation.
-
-\hypertarget{existing-work-and-new-challenges}{%
-\section{Existing work and new
-challenges}\label{existing-work-and-new-challenges}}
-
-To represent spatio-temporal data, E. Pebesma
-(\protect\hyperlink{ref-spacetime}{2012}) proposes four layouts: full
-grid layout (STF), sparse grid layout (STS), irregular layout (STI), and
-trajectory (STT). These layouts are implemented in the \pkg{spacetime}
-package with underlying class \pkg{sp} (\protect\hyperlink{ref-sp}{E.
-Pebesma and Bivand 2005}) and \pkg{xts}
-(\protect\hyperlink{ref-xts}{Ryan and Ulrich 2020}) for handling spatial
-and temporal variables. While these layouts and classes are still being
-used and well-maintained today, they are not compatible with the
-\pkg{tidyverse} ecosystem (\protect\hyperlink{ref-tidyverse}{Wickham et
-al. 2019}), which has a huge influence on data analysis in \proglang{R},
-including in the spatial and temporal communities. The package \pkg{sf}
-(\protect\hyperlink{ref-sf}{E. J. Pebesma 2018}) uses data frame objects
-to represent spatial data with an additional \code{geometry} column
-dedicating to feature geometries
-(\protect\hyperlink{ref-jr_herring_opengis_2011}{J.R. Herring 2011})
-(e.g.~POINT, MULTIPOINT, POLYGON, MULTIPOLYGON etc). For temporal data,
-\pkg{tsibble} (\protect\hyperlink{ref-tsibble}{Wang, Cook, and Hyndman
-2020a}) builds on top of the \pkg{tibble}
-(\protect\hyperlink{ref-tibble}{Müller and Wickham 2021}) object and
-explicitly includes date and time in columns as variables, rather than
-data attributes.
-
-Underlying tidyverse is the concept of tidy data
-(\protect\hyperlink{ref-tidydata}{Wickham 2014}), which considers how
-variables should be structured for analysis. In the three formats
-(time-wide, space-wide, and long) summarised in the E. Pebesma
-(\protect\hyperlink{ref-spacetime}{2012}) paper, only the long form is
-considered as tidy. The main advantages of it over the time-wide and
-space-wide format is that the latter treat either time or location as
-variable names (Section 3.1 in Wickham
-(\protect\hyperlink{ref-tidydata}{2014})) rather than values of
-variables, making them difficult to be wrangled. One problem with the
-long format is that it can be inefficient to store feature geometries,
-especially for large MULTIPOLYGON over long time period. This motivates
-a new tidyverse compatible data object to work with spatio-temporal
-data.
-
-\textbf{two challenages: multiple data sources + large spatio-temporal
-object} In additional to the shift into more tidyverse workflow for
-working with data, spatio-temporal data have some unique challenges. The
-first one being the split of variables in multiple tables. Often the
-time, spatio-temporal data is not presented in a single data object, but
-split into fragmented pieces from different sources. For example, map
-data recorded the shape of a collection of area of interest is provided
-by the government website; the geo-scientific data is recorded with
-longitude and latitude coordinate of locations in the area; and temporal
-variables indexed by both location and time need to be further queried
-using location data. For spatio-temporal data analysis, merging all
-these sources can be challenging and joining these tables can go wrong
-because of the slightest unmatched of the linking variable from
-different sources.
-
-Another challenge when analysing spatio-temporal data is on the simple
-feature representation of spatial data. While the simple feature is able
-to handle large and complex spatial geometry and perform precise spatial
-operations, data analysts usually do not have a large expectation on the
-map accuracy and would prefer a simplified map with quick rendering for
-visualisation. Also, the feature geometry column in \pkg{sf} wraps the
-spatial coordinates in a list column, sometimes, analysts may like to
-unpack them into longitude and latitude to work with. A smooth
-transition with the coordinate columns and and simple feature geometry
-column would satisfy analysts from both ends.
-
-With some of the aforementioned issues recognised, the \pkg{stars}
-package has been a new implementation to handle spatio-temporal data
-conceptually as data cubes and implemented as arrays. The package
-depends on the \pkg{sf} package and has allowed more access to tidyverse
-verbs in the workflow. However, for vector spatio-temporal data, major
-difficulties to work with \pkg{stars} is first to cast the variables
-from multiple tables into a star object and then keep the operations in
-the high-dimensional cube. Given that most table mentioned before
-collect variables in a 2D table, analysts could benefit from a structure
-to allow them work in 2D tables.
-
-\hypertarget{the-cubble-package}{%
-\section{The cubble package}\label{the-cubble-package}}
-
-This section will present two formats that cubble uses to arrange
-spatio-temporal data. The main cubble functions will then be introduced
-to illustrate how to work with these two formats in short examples.
-Lastly, a subsection will be dedicated to how existing packages, in
-spatial and temporal analysis, fit in with cubble.
-
-In cubble, a single data object can be arranged into two forms: nested
-form and long form. Figure \ref{fig:illu-cubble} sketches the two forms
-with the associated attributes. The decision on which form to use in
-data wrangling is output-oriented, meaning analysts need to first think
-about whether the output of an operation is identified only by the
-spatial identifier, or a combination of spatial and temporal identifier.
-The nested cubble is suitable for working with operations that are only
-identified by the site and this type of operation can be a pure
-manipulation of time invariant variables, or an operation that
-summarises time varying variables into site. Underneath the nested form,
-a cubble is built from a \code{rowwise_df} class where each site forms a
-separate group. This structure simplifies the calculation that involves
-temporal variables by avoiding the use of \code{purrr::map()} syntax
-when working with list-column.
-
-For those operations whose output involves both a spatial and temporal
-identifier, long form should be used. The long form uses a
-\code{grouped_df} class to forms all the time of a site as a group. Time
-invariant variables are stored separately as an special attribute of the
-long cubble. This design avoids repeating the spatial variables at each
-time stamp while not dropping information from spatial variables.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.002}
-
-}
-
-\caption{An illustration of the nested and long form in cubble.
-The nested form defines each site in a row and nests the time varying variables into a single column \code{ts}.
-The long form cubble uses \code{id} and \code{t} to identify each row and stores the time invariant variables as an attribute, \code{spatial}.}\label{fig:illu-cubble}
-\end{figure}
-
-\hypertarget{create-a-cubble-in-the-nested-form}{%
-\subsection{Create a cubble in the nested
-form}\label{create-a-cubble-in-the-nested-form}}
-
-To use functions in the \texttt{cubble} package, an analyst will first
-need to turn the data object into a \code{cubble} class.
-\code{as_cubble()} does this through supplying the three key components:
-\code{key} as the spatial identifier; \code{index} as the temporal
-identifier; and a vector of \code{coords} in the order of longitude and
-latitude. The use of \code{key} and \code{index} follows the design in
-the \pkg{tsibble} package and the cubble created by default is in the
-nested form.
-
-Before the formal examples in Section \ref{examples}, each function
-introduced in this section will be accompanied by a small example. The
-data used for the small examples is a subset of a larger climate weather
-station data used in the formal examples and has been simplified to
-include only five weather stations. It contains spatial information of
-each station: station id, latitude, longitude, elevation, station name
-and World Meteorology Organisation ID and also daily temporal
-information of date, maximum and minimum temperature and precipitation
-for 2020. \code{climate_flat} stores this data in format 1) in Figure
-\ref{fig:illu-input} and the code below creates a cubble out of
-\code{climate_flat} with \code{id} as the key, \code{date} as the index,
-and \code{c(long, lat)} as the coordinates:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{cubble\_nested }\OtherTok{\textless{}{-}}\NormalTok{ cubble}\SpecialCharTok{::}\NormalTok{climate\_flat }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
+Spatio-temporal data has a spatial component referring to the location of each observation, a temporal component that is recorded at regular or irregular time intervals, and may have multiple variables measured at each spatial and temporal value.
+The analysis of spatio-temporal data can fix the time to explore the spatial distribution, fix the spatial location to explore temporal trends, or dynamically explore the space and time simultaneously.
+For these tasks to be carried out easily, there is a need for a spatio-temporal data representation that underpins various operations.
+
+The \emph{Representing data} section of the SpatioTemproal CRAN task view (\protect\hyperlink{ref-ctvspatiotemporal}{Edzer Pebesma 2022}) lists existing spatio-temporal data representation.
+Among them, E.
+Pebesma (\protect\hyperlink{ref-spacetime}{2012}) proposes four layouts: full grid layout, sparse grid layout, irregular layout, and trajectory.
+These layouts are implemented in the \pkg{spacetime} (\protect\hyperlink{ref-spacetime}{E.
+Pebesma 2012}) package with the underlying spatial and temporal class: \pkg{sp} (\protect\hyperlink{ref-sp}{E.
+Pebesma and Bivand 2005}) and \pkg{xts} (\protect\hyperlink{ref-xts}{Ryan and Ulrich 2020}).
+The \pkg{stars} (\protect\hyperlink{ref-stars}{E.
+Pebesma 2021}) package has been a recent implementation to handle both raster and vector spatio-temporal data using array.
+It depends on the newer spatial package \pkg{sf} (\protect\hyperlink{ref-sf}{E.
+J.
+Pebesma 2018}) and interfaces with the widely-used tidyverse verbs.
+
+{[}till here{]} One issue with wrangling spatio-temporal data is that the raw data does not naturally structured in any one of the representation above.
+Also, in data collection, spatio-temporal data is less often collected as a whole piece but fragmented spatial and temporal data for analysts to ensemble.
+For example, map data records the shape of a collection of area of interest is provided by official government website; geo-scientific data records the longitude and latitude coordinate of locations in the area along with other meta data relates to the location; and temporal variables indexed by both location and time.
+For spatio-temporal data analysis, merging all these sources can be challenging and joining these tables can go wrong because of the slightest unmatched of the linking variable from different sources.
+
+Another issue is how the concept of tidy data (\protect\hyperlink{ref-tidydata}{Wickham 2014}) should be applied to spatio-temporal data.
+The tidy data concept advocates the data to be shaped into 1) one row per observation; 2) one column per variable; and 3) one type of data per table.
+The long form data is preferred over wide data given the downstream software i.e.~\pkg{dplyr} and \pkg{ggplot2} for data wrangling and visualisation.
+However, the long form can be inefficient to store feature geometries, especially for large multipolygon over long time period.
+This poses the question on how to arrange spatial and temporal variables that makes analysing spatio-temporal data easier.
+
+This paper presents a new \proglang{R} package, \pkg{cubble}, to {[}overcome the problem{]} with organising variables in different levels when working with spatio-temporal data.
+The data structure focuses on connecting to the upstream data collection and is suitable for temporal data collected at the fixed location over time.
+Among the four spacetime layouts in E.
+Pebesma (\protect\hyperlink{ref-spacetime}{2012}), it can be full grid layout, sparse grid layout, or irregular layout, but trajectory is not within the scope of this work.
+In the package a new data structure, also called \code{cubble}, is proposed to organise spatial and temporal variables as two forms of a single data object so that they can be wrangled separately while synchronised as a whole piece.
+With this new data object, analysts can spend less time writing codes to organise spatial and temporal variables and focus more on the exploratory data analysis itself.
+The software is available from the Comprehensive R Archive Network (CRAN) at {[}CRAN link{]}.
+
+The rest of the paper is divided as follows: Section \ref{cube} discusses the spatio-temporal cube to conceptualise spatio-temporal data.
+Section \ref{cubble} presents the main design and functionality of \pkg{cubble}.
+Section \ref{others} explains how cubble deals with more advanced considerations including data with hierarchical structure, data matching, how cubble fits with existing static and interactive visualisation tools, and spatio-temporal data transformation.
+Section \ref{examples} uses Australia weather station data and river level data as examples to demonstrate the use of \pkg{cubble}.
+An example on how \pkg{cubble} handles NetCDF data is also provided.
+Section \ref{conclude} concludes the paper.
+
+\hypertarget{cube}{%
+\section{Conceptual framework: spatio-temporal cube}\label{cube}}
+
+Spatio-temporal data can be conceptualised using a cubical data model with reference to variables, time and space.
+This naturally motivates the usage of array to represent and store spatio-temporal data, as evident by satellite imageries, large climate models, among others.
+The work by Lu, Appel, and Pebesma (\protect\hyperlink{ref-lu_multidimensional_2018}{2018}) has discussed the suitability of this representation through array algebra and practical spatio-temporal tasks.
+\textcolor{magenta}{SZ: is this too direct?}
+
+The cubical conceptual framework provides a generalisation to the operation can be done on the data for visualisation.
+A paper by Bach et al.
+(\protect\hyperlink{ref-bach_review_2014}{2014}) has reviewed the temporal data visualisation based on space-time cube operations.
+Notice that the term space-time cube in their article ``does not need to involve spatial data'', but referring to ``an abstract 2D substrate that is used to visualize data at a specific time''.
+Despite its main focus is on temporal data, the mindset of abstracting out data representation to construct visualisation is applicable to our spatio-temporal data manipulation and visualisation.
+
+\hypertarget{cubble}{%
+\section{The cubble package}\label{cubble}}
+
+This section will first introduce the core functions in the cubble class: \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and \code{unfold()} from subsection \ref{create} to \ref{unfold}.
+The next subsection addresses why bother a new class before subsection \ref{compact} finishing on the compatibility of cubble with existing packages in spatial and temporal analysis.
+
+Each core function is introduced with a short example accompanied and the data \texttt{climate\_flat} is used throughout this section.
+\texttt{climate\_flat} contains five weather stations in Australia with spatial information of each station: station id, latitude, longitude, elevation, station name and World Meteorology Organisation ID and also daily temporal information of date, maximum and minimum temperature and precipitation for 2020.
+Below prints out the first five rows of \texttt{climate\_flat}:
+
+\begin{verbatim}
+## # A tibble: 1,830 x 10
+##   id            lat  long  elev name         wmo_id date        prcp  tmax  tmin
+##   <chr>       <dbl> <dbl> <dbl> <chr>         <dbl> <date>     <dbl> <dbl> <dbl>
+## 1 ASN00009021 -31.9  116.  15.4 perth airpo~  94610 2020-01-01     0  31.9  15.3
+## 2 ASN00009021 -31.9  116.  15.4 perth airpo~  94610 2020-01-02     0  24.9  16.4
+## 3 ASN00009021 -31.9  116.  15.4 perth airpo~  94610 2020-01-03     6  23.2  13  
+## 4 ASN00009021 -31.9  116.  15.4 perth airpo~  94610 2020-01-04     0  28.4  12.4
+## 5 ASN00009021 -31.9  116.  15.4 perth airpo~  94610 2020-01-05     0  35.3  11.6
+## # ... with 1,825 more rows
+\end{verbatim}
+
+\hypertarget{create}{%
+\subsection{Create a cubble in the nested form}\label{create}}
+
+As described in the introduction, spatio-temporal data can come in various formats and shapes.
+A cubble can be created from various raw data, including tibble and its variates multiple tables (an example provided in section \ref{covid}), and netcdf (detailed in section \ref{netcdf}).
+The function \code{as_cubble()} is used to create a cubble with three additional components: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order of longitude and latitude.
+Argument \code{key} and \code{index} follow the wording in \pkg{tsibble} to define temporal order and multiple series while \code{coords} is a space specific argument to locate each site in two columns.
+The code below creates a cubble out of \code{climate_flat} with \code{id} as the key, \code{date} as the index, and \code{c(long, lat)} as the coordinates:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{cubble\_nested }\OtherTok{\textless{}{-}}\NormalTok{ climate\_flat }\SpecialCharTok{|\textgreater{}}
   \FunctionTok{as\_cubble}\NormalTok{(}\AttributeTok{key =}\NormalTok{ id, }\AttributeTok{index =}\NormalTok{ date, }\AttributeTok{coords =} \FunctionTok{c}\NormalTok{(long, lat))}
@@
 
-There are a few information in the cubble header: the name of the
-\code{key} variable: \code{id} and its unique number: 5, the bounding
-box: \code{[115.97, -32.94, 133.55, -12.42]} , and also the name of
-variable nested in the \code{ts} column with its type:
-\code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.
-
-\hypertarget{stretch-a-nested-cubble-into-the-long-form}{%
-\subsection{Stretch a nested cubble into the long
-form}\label{stretch-a-nested-cubble-into-the-long-form}}
-
-The nested format is convenient for those operations whose output does
-not contain a time dimension. For those outputs that are
-cross-identified by the spatial and temporal identifier, a long cubble
-needs to be used. The function \code{stretch()} is designed to switch
-the cubble from the nested form to the long form. This code shows how to
-switch the nested cubble just created into its long form:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{cubble\_long }\OtherTok{\textless{}{-}}\NormalTok{ cubble\_nested }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{face\_temporal}\NormalTok{()}
+The cubble header provides some information about this data.
+\code{id} is the variable name to identify each location and there are five unique locations.
+The bounding box is \code{[115.97, -32.94, 133.55, -12.42]} and the names of variable nested in the \code{ts} column, along with its type, are \code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.
+
+The created cubble is built from a \code{rowwise_df} class where each row forms a group and all the temporal variables are nested in a list column, hence it is also called the nested cubble.
+The rowwise structure makes it simpler to calculate on the list using \code{mutate()} and avoid the \code{map()} when working with list column.
+For example calculating the number of raining day can be done by:
+
+\begin{verbatim}
+## # cubble:   id [5]: nested form
+## # bbox:     [115.97, -32.94, 133.55, -12.42]
+## # temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
+##   id            lat  long  elev name           wmo_id ts       rain_day
+##   <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>      <int>
+## 1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tibble>      104
+## 2 ASN00010311 -31.9  117. 179   york            94623 <tibble>       89
+## 3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble>       90
+## 4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble>      106
+## 5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble>       63
+\end{verbatim}
+
+\hypertarget{switch-from-the-nested-cubble-into-the-long-form}{%
+\subsection{Switch from the nested cubble into the long form}\label{switch-from-the-nested-cubble-into-the-long-form}}
+
+The nested format is convenient for those operations whose output does not contain a time dimension.
+For those outputs that are cross-identified by the spatial and temporal identifier, a long cubble needs to be used.
+The function \code{face_temporal()} switches the cubble from the nested form to the long form and the first row in Figure \ref{fig:face} sketches the function in a diagram.
+This code shows how to switch the cubble just created into its long form:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{cubble\_long }\OtherTok{\textless{}{-}}\NormalTok{ cubble\_nested }\SpecialCharTok{|\textgreater{}} \FunctionTok{face\_temporal}\NormalTok{()}
 \NormalTok{cubble\_long}
@@
 
-\hypertarget{tamp-a-long-cubble-back-to-the-nested-form}{%
-\subsection{Tamp a long cubble back to the nested
-form}\label{tamp-a-long-cubble-back-to-the-nested-form}}
-
-Manipulation on the spatial and temporal dimension can be an iterative
-process and analysts may need to go back and forth between the nested
-and long cubble. \code{tamp()}, an inverse of \code{stretch()}, switches
-a long cubble to a nested cubble:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{cubble\_back }\OtherTok{\textless{}{-}}\NormalTok{ cubble\_long }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{face\_spatial}\NormalTok{()}
+The long form is built form class \code{groupped_df} where all the observations from the same sites forms a group.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.008}
+
+}
+
+\caption{An illustration of function \code{face\_temporal} and \code{face\_spatial} in cubble.
+In the first row, \code{face\_temporal} switches a cubble from the nested form into the long form and the focus has switched from the spatial aspect (the side face) to the temporal aspect (the front face).
+In the second row, \code{face\_spatial} switches a cubble back to the nested form from te long form and shift focus back to the spatial aspect.}\label{fig:face}
+\end{figure}
+
+\hypertarget{switch-a-long-cubble-back-to-the-nested-form}{%
+\subsection{Switch a long cubble back to the nested form}\label{switch-a-long-cubble-back-to-the-nested-form}}
+
+Manipulation on the spatial and temporal dimension can be an iterative process and analysts may need to go back and forth between the nested and long cubble.
+\code{face_spatial()}, the inverse of \code{face_temporal()}, switches the long cubble back to the nested form, as sketched in the second row in Figure \ref{fig:face}:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{cubble\_back }\OtherTok{\textless{}{-}}\NormalTok{ cubble\_long }\SpecialCharTok{|\textgreater{}} \FunctionTok{face\_spatial}\NormalTok{()}
 \NormalTok{cubble\_back}
@@
 
-\hypertarget{migrate-spatial-variables-to-a-long-cubble}{%
-\subsection{Migrate spatial variables to a long
-cubble}\label{migrate-spatial-variables-to-a-long-cubble}}
-
-Sometimes, analysts may need to apply some variable transformation that
-involves both the spatial and temporal variable. An example of this is
-the transformation of temporal variables into the spatial dimension in
-glyph maps (which will be elaborated in section
-\ref{st_transformation}). Cubble allows this operation through
-\code{unfold()}, which moves the specified spatial variables into the
-long cubble:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{cubble\_mig }\OtherTok{\textless{}{-}}\NormalTok{ cubble\_long }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{unfold}\NormalTok{(long, lat)}
-\NormalTok{cubble\_mig}
+\hypertarget{unfold}{%
+\subsection{Flatten spatial variables into the long cubble}\label{unfold}}
+
+Sometimes, analysts may need to apply some variable transformation that involves both the spatial and temporal variable.
+An example of this is the transformation of temporal variables into the spatial dimension in glyph maps, which will be elaborated in section \ref{st_transformation}.
+This type of operations can be seen as flattening, or \emph{unfolding}, the cube into a 2D data frame.
+Here the function \code{unfold()} moves the spatial variables \code{long} and \code{lat} into the long cubble:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{cubble\_unfold }\OtherTok{\textless{}{-}}\NormalTok{ cubble\_long }\SpecialCharTok{|\textgreater{}} \FunctionTok{unfold}\NormalTok{(long, lat)}
+\NormalTok{cubble\_unfold}
 \end{Highlighting}
@@
 
-This function should generally be used in the last step of the analysis
-since it is a temporary operation, meaning these added spatial variables
-are not stored in the long form and will disappear if switched to the
-nested form and then switched back:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{cubble\_mig }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{face\_spatial}\NormalTok{() }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{face\_temporal}\NormalTok{()}
+This function should generally be used in the last step of the analysis since it is a temporary operation, meaning these added spatial variables are not stored in the long form and will disappear if switched to the nested form and then switched back:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{cubble\_unfold }\SpecialCharTok{|\textgreater{}} \FunctionTok{face\_spatial}\NormalTok{() }\SpecialCharTok{|\textgreater{}} \FunctionTok{face\_temporal}\NormalTok{()}
 \end{Highlighting}
@@
 
-\hypertarget{compatibility-with-existing-packages}{%
-\subsection{Compatibility with existing
-packages}\label{compatibility-with-existing-packages}}
-
-The previous four subsections have introduced operations specific to the
-\code{cubble} class and this section will demonstrate how the
-\code{cubble} class interacts with existing packages commonly used in
-spatial and temporal analysis, specifically, \code{dplyr},
-\code{tsibble}, \code{sf} (\code{s2}), and \code{netcdf4}.
+\subsection[Why not just use]{Why not just use \code{dplyr::nest\_by()}}
+
+It may naturally raise the question that why bother a new data structure since the nested form can already be created by the \pkg{dplyr} function \code{nest_by()}?
+While \code{nest_by()} and \code{unnest()} can mimic the nested form in cubble, cubble can be seen as an attempt to arrange variables from different observational units into a single object, specifically for spatio-temporal data.
+While \texttt{nest\_by()} solves this problem on the spatial side, it doesn't solve the temporal side.
+For temporal data, a combined representation with spatial variables is useful for spatio-temporal operations but its duplication can be a burden to operate on.
+Also, spatial variables are dropped when summarised and you will need to either specified all of them in the \code{group_by()} before \code{summarise}, or add a joining after \code{summarise()}.
+
+\hypertarget{compact}{%
+\subsection{Compatibility with existing packages}\label{compact}}
+
+The previous four subsections have introduced operations specific to the \code{cubble} class and this section will demonstrate how the \code{cubble} class interacts with existing packages commonly used in spatial and temporal analysis, specifically, \code{dplyr}, \code{tsibble}, \code{sf} (\code{s2}), and \code{netcdf4}.
 
@@
 
-\code{tsibble} is a temporal data structure that uses \code{index} and
-\code{key} to identify the time and different series. \code{cubble} can
-be seen as a natural extension of \code{tsibble} for spatio-temporal
-data with an additional coordinates component and using two forms to
-arrange variables. This makes it easy to cast a \code{tsibble} into a
-\code{cubble} as only the \code{coords} argument needs to be supplied:
+\code{tsibble} is a temporal data structure that uses \code{index} and \code{key} to identify the time and different series.
+\code{cubble} can be seen as following the same vein as \code{tsibble} for spatio-temporal data.
+This makes it easy to cast a \code{tsibble} into a \code{cubble} as only the \code{coords} argument needs to be supplied:
 
@@
 \CommentTok{\# example with a tsibble created from climate\_flat}
-\NormalTok{raw }\OtherTok{\textless{}{-}}\NormalTok{ climate\_flat }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
-\NormalTok{  tsibble}\SpecialCharTok{::}\FunctionTok{as\_tsibble}\NormalTok{(}\AttributeTok{key =}\NormalTok{ id, }\AttributeTok{index =}\NormalTok{ date)}
-
-\NormalTok{dt }\OtherTok{\textless{}{-}}\NormalTok{  raw }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
-\NormalTok{  cubble}\SpecialCharTok{::}\FunctionTok{as\_cubble}\NormalTok{(}\AttributeTok{coords =} \FunctionTok{c}\NormalTok{(long, lat))}
+\NormalTok{raw }\OtherTok{\textless{}{-}}\NormalTok{ climate\_flat }\SpecialCharTok{|\textgreater{}}\NormalTok{ tsibble}\SpecialCharTok{::}\FunctionTok{as\_tsibble}\NormalTok{(}\AttributeTok{key =}\NormalTok{ id, }\AttributeTok{index =}\NormalTok{ date)}
+\NormalTok{dt }\OtherTok{\textless{}{-}}\NormalTok{  raw }\SpecialCharTok{|\textgreater{}}\NormalTok{ cubble}\SpecialCharTok{::}\FunctionTok{as\_cubble}\NormalTok{(}\AttributeTok{coords =} \FunctionTok{c}\NormalTok{(long, lat))}
 \NormalTok{dt}
@@
 \CommentTok{\# add station{-}based features in the nested form.}
-\NormalTok{dt }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
-  \FunctionTok{mutate}\NormalTok{(fabletools}\SpecialCharTok{::}\FunctionTok{features}\NormalTok{(ts, tmax, }\FunctionTok{list}\NormalTok{(}\AttributeTok{tmax\_mean =}\NormalTok{ mean, }\AttributeTok{tmax\_var =}\NormalTok{ var)))}
+\NormalTok{dt }\SpecialCharTok{|\textgreater{}} \FunctionTok{mutate}\NormalTok{(fabletools}\SpecialCharTok{::}\FunctionTok{features}\NormalTok{(ts, tmax, }\FunctionTok{list}\NormalTok{(}\AttributeTok{tmax\_mean =}\NormalTok{ mean, }\AttributeTok{tmax\_var =}\NormalTok{ var)))}
 \end{Highlighting}
@@
 \CommentTok{\# create a cubble}
-\NormalTok{cb }\OtherTok{\textless{}{-}}\NormalTok{ climate\_flat }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
+\NormalTok{cb }\OtherTok{\textless{}{-}}\NormalTok{ climate\_flat }\SpecialCharTok{|\textgreater{}}
 \NormalTok{  cubble}\SpecialCharTok{::}\FunctionTok{as\_cubble}\NormalTok{(}\AttributeTok{key =}\NormalTok{ id, }\AttributeTok{index =}\NormalTok{ date, }\AttributeTok{coords =} \FunctionTok{c}\NormalTok{(long, lat))}
@@
 
-\NormalTok{dt }\OtherTok{\textless{}{-}}\NormalTok{ cb }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
+\NormalTok{dt }\OtherTok{\textless{}{-}}\NormalTok{ cb }\SpecialCharTok{|\textgreater{}}
   \FunctionTok{mutate}\NormalTok{(}
@@
 \CommentTok{\# Western Australia map}
-\NormalTok{wa }\OtherTok{\textless{}{-}}\NormalTok{ ozmaps}\SpecialCharTok{::}\NormalTok{abs\_ste }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{filter}\NormalTok{(NAME }\SpecialCharTok{==} \StringTok{"Western Australia"}\NormalTok{)}
+\NormalTok{wa }\OtherTok{\textless{}{-}}\NormalTok{ ozmaps}\SpecialCharTok{::}\NormalTok{abs\_ste }\SpecialCharTok{|\textgreater{}} \FunctionTok{filter}\NormalTok{(NAME }\SpecialCharTok{==} \StringTok{"Western Australia"}\NormalTok{)}
 
 \CommentTok{\# mutate a \textasciigrave{}s2\_lnglat\textasciigrave{} vector on \textasciigrave{}cb\textasciigrave{} created in the last chunk}
-\NormalTok{cb }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{ll =} \FunctionTok{s2\_lnglat}\NormalTok{(long, lat)) }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
+\NormalTok{cb }\SpecialCharTok{|\textgreater{}}
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{ll =} \FunctionTok{s2\_lnglat}\NormalTok{(long, lat)) }\SpecialCharTok{|\textgreater{}}
   \FunctionTok{filter}\NormalTok{(}\FunctionTok{s2\_within}\NormalTok{(ll, wa))}
@@
 
-NetCDF data has two main components: \emph{dimension} for defining the
-spatio-temporal grid (longitude, latitude, and time) and \emph{variable}
-that populates the defined grid. Attributes are usually associated with
-dimension and variable in the NetCDF format data and a
-\href{http://cfconventions.org/}{metadata convention for climate and
-forecast} has been designed to standardise the format of the attributes.
-A few packages in R exists for manipulating NetCDF data and this
-includes a high-level R interface: \pkg{ncdf4}
-(\protect\hyperlink{ref-ncdf4}{Pierce 2019}), a low-level interface that
-calls C-interface: \pkg{RNetCDF} (\protect\hyperlink{ref-rnetcdf}{Michna
-and Woods 2021}, \protect\hyperlink{ref-michna2013rnetcdf}{2013}), and a
-tidyverse implementation: \pkg{tidync}
-(\protect\hyperlink{ref-tidync}{Sumner 2020}).
-
-Cubble provides an \code{as_cubble()} method to coerce the \code{ncdf4}
-class from the \pkg{ncdf4} package into a \code{cubble}. It maps each
-combination of longitude and latitude into an \code{id} as the
-\code{key}:
+NetCDF data has two main components: \emph{dimension} for defining the spatio-temporal grid (longitude, latitude, and time) and \emph{variable} that populates the defined grid.
+Attributes are usually associated with dimension and variable in the NetCDF format data and a \href{http://cfconventions.org/}{metadata convention for climate and forecast} has been designed to standardise the format of the attributes.
+A few packages in R exists for manipulating NetCDF data and this includes a high-level R interface: \pkg{ncdf4} (\protect\hyperlink{ref-ncdf4}{Pierce 2019}), a low-level interface that calls C-interface: \pkg{RNetCDF} (\protect\hyperlink{ref-rnetcdf}{Michna and Woods 2021}, \protect\hyperlink{ref-michna2013rnetcdf}{2013}), and a tidyverse implementation: \pkg{tidync} (\protect\hyperlink{ref-tidync}{Sumner 2020}).
+
+Cubble provides an \code{as_cubble()} method to coerce the \code{ncdf4} class from the \pkg{ncdf4} package into a \code{cubble}.
+It maps each combination of longitude and latitude into an \code{id} as the \code{key}:
 
@@
 
-One task that may interest analysts in spatio-temporal data is to find
-how similar the time series from nearby sites are. This can be seen as a
-matching problem (\protect\hyperlink{ref-stuart2010matching}{Stuart
-2010}; \protect\hyperlink{ref-mcintosh2018using}{McIntosh et al. 2018})
-that pairs up similar similar time series in nearby locations or a data
-fusing exercise that merges data collected from different sources
-(\protect\hyperlink{ref-cocchi2019data}{Cocchi 2019}).
-\code{match_sites()} in \pkg{cubble} provides a simple algorithm for
-this task. The algorithm first matches the two data sources spatially
-through computing the pairwise distance on latitude and longitude. Pairs
-that pass the spatial matching are then matched temporally through
-computing the number of matched peaks within a fixed length window.
-Figure \ref{fig:illu-matching} illustrates this temporal matching in
-more details. Given two series \code{A} and \code{a}, 3 peaks have been
-picked in each series. An interval, with default length of 5, is
-constructed for each peak in series \code{A} and the peaks in series
-\code{a} are tested against whether they fall into the any of the
-intervals. In this illustration, there are 2 matches for these two
-series. Several arguments are available in \code{match_sites()} to
-fine-tune the matching:
+One task that may interest analysts in spatio-temporal data is to find how similar the time series from nearby sites are.
+This can be seen as a matching problem (\protect\hyperlink{ref-stuart2010matching}{Stuart 2010}; \protect\hyperlink{ref-mcintosh2018using}{McIntosh et al.
+2018}) that pairs up similar similar time series in nearby locations or a data fusing exercise that merges data collected from different sources (\protect\hyperlink{ref-cocchi2019data}{Cocchi 2019}).
+\code{match_sites()} in \pkg{cubble} provides a simple algorithm for this task.
+The algorithm first matches the two data sources spatially through computing the pairwise distance on latitude and longitude.
+Pairs that pass the spatial matching are then matched temporally through computing the number of matched peaks within a fixed length window.
+Figure \ref{fig:illu-matching} illustrates this temporal matching in more details.
+Given two series \code{A} and \code{a}, 3 peaks have been picked in each series.
+An interval, with default length of 5, is constructed for each peak in series \code{A} and the peaks in series \code{a} are tested against whether they fall into the any of the intervals.
+In this illustration, there are 2 matches for these two series.
+Several arguments are available in \code{match_sites()} to fine-tune the matching:
 
@@
 
-Cubble fits in naturally with the interactive graphic pipeline discussed
-in the literature (\protect\hyperlink{ref-buja1988elements}{Buja,
-Asimov, and Hurley 1988};
-\protect\hyperlink{ref-buja1996interactive}{Buja, Cook, and Swayne
-1996}; \protect\hyperlink{ref-sutherland2000orca}{Sutherland et al.
-2000}; \protect\hyperlink{ref-xie2014reactive}{Xie, Hofmann, and Cheng
-2014}; \protect\hyperlink{ref-cheng2016enabling}{Cheng, Cook, and
-Hofmann 2016}). Diagram \ref{fig:illu-interactive} illustrates how
-linking works from the map to the time series in cubble. The map and
-time series plot is associated with the nested or long cubble,
-respectively, and when a user action is captured on the map, the site
-will be activated in the nested cubble (left). The nested cubble will
-communicate to the long cubble to activate all the observations with the
-same \code{id} (middle). The long cubble will then highlight the
-activated series in the time series plot (right).
-
-The linking is also available from the time series plot to the map. The
-selection(s) on the time series is through selecting the point(s) on the
-time series and once a point is selected, it will be activated in the
-long cubble. All the observations that share the same \code{id} are then
-activated and this includes other points in the same time series in the
-long cubble and the corresponding observation of site in the nested
-cubble. These activated observations will then being reflected in the
-updated plots and Diagram \ref{fig:illu-interactive-2} in the Appendix
-illustrates this process.
+Cubble fits in naturally with the interactive graphic pipeline discussed in the literature (\protect\hyperlink{ref-buja1988elements}{Buja, Asimov, and Hurley 1988}; \protect\hyperlink{ref-buja1996interactive}{Buja, Cook, and Swayne 1996}; \protect\hyperlink{ref-sutherland2000orca}{Sutherland et al.
+2000}; \protect\hyperlink{ref-xie2014reactive}{Xie, Hofmann, and Cheng 2014}; \protect\hyperlink{ref-cheng2016enabling}{Cheng, Cook, and Hofmann 2016}).
+Diagram \ref{fig:illu-interactive} illustrates how linking works from the map to the time series in cubble.
+The map and time series plot is associated with the nested or long cubble, respectively, and when a user action is captured on the map, the site will be activated in the nested cubble (left).
+The nested cubble will communicate to the long cubble to activate all the observations with the same \code{id} (middle).
+The long cubble will then highlight the activated series in the time series plot (right).
+
+The linking is also available from the time series plot to the map.
+The selection(s) on the time series is through selecting the point(s) on the time series and once a point is selected, it will be activated in the long cubble.
+All the observations that share the same \code{id} are then activated and this includes other points in the same time series in the long cubble and the corresponding observation of site in the nested cubble.
+These activated observations will then being reflected in the updated plots and Diagram \ref{fig:illu-interactive-2} in the Appendix illustrates this process.
 
@@
 
-Spatio-temporal transformation is a useful technique to extract
-information form spatio-temporal data. Glyph maps
-(\protect\hyperlink{ref-Wickham2012-yr}{Wickham et al. 2012}) transform
-the time coordinates into the space coordinates to plot the time series
-of different locations on the map. Calendar plots
-(\protect\hyperlink{ref-wang2020calendar}{Wang, Cook, and Hyndman
-2020b}) reconstructs time into calendar-based grid to discover weekday
-and weekend pattern. Projection, or linear combination, of variables
-summarises multivariate information into lower dimension to further
-digest. This section elaborates on the glyph map.
-
-In \proglang{R}, \pkg{GGally} implements glyph maps through the
-\code{glyphs()} function. The function constructs a data frame with
-calculated position (\code{gx}. \code{gy}. \code{gid}) of each point on
-the time series using linear algebra (Equation 1 and 2 in Wickham et al.
-(\protect\hyperlink{ref-Wickham2012-yr}{2012})). The data can then be
-piped into \code{ggplot} to create the glyph map as:
+Spatio-temporal transformation is a useful technique to extract information form spatio-temporal data.
+Glyph maps (\protect\hyperlink{ref-Wickham2012-yr}{Wickham et al.
+2012}) transform the time coordinates into the space coordinates to plot the time series of different locations on the map.
+Calendar plots (\protect\hyperlink{ref-wang2020calendar}{Wang, Cook, and Hyndman 2020}) reconstructs time into calendar-based grid to discover weekday and weekend pattern.
+Projection, or linear combination, of variables summarises multivariate information into lower dimension to further digest.
+This section elaborates on the glyph map.
+
+In \proglang{R}, \pkg{GGally} implements glyph maps through the \code{glyphs()} function.
+The function constructs a data frame with calculated position (\code{gx}.
+\code{gy}.
+\code{gid}) of each point on the time series using linear algebra (Equation 1 and 2 in Wickham et al.
+(\protect\hyperlink{ref-Wickham2012-yr}{2012})).
+The data can then be piped into \code{ggplot} to create the glyph map as:
 
@@
 
+\hypertarget{covid}{%
+\subsection{Victoria covid data}\label{covid}}
+
+This example illustrate casting different sources into a cubble object.
+Victoria government of Australia provides daily COVID case number by date, source, and local government area (LGA).
+This data can be useful for visualising COVID spread when combined with LGA map information, which is available from Australian Bureau of Statistics.
+Variable \code{covid} and \code{lga} show the first 5 rows of these two data:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{covid }\SpecialCharTok{|\textgreater{}} \FunctionTok{head}\NormalTok{(}\DecValTok{5}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## # A tsibble: 5 x 5 [1D]
+## # Key:       lga [1]
+## # Groups:    lga, source [1]
+##   date       lga        source                            n roll_mean
+##   <date>     <chr>      <chr>                         <int>     <dbl>
+## 1 2022-01-01 Alpine (S) Contact with a confirmed case     1        NA
+## 2 2022-01-02 Alpine (S) Contact with a confirmed case     2        NA
+## 3 2022-01-03 Alpine (S) Contact with a confirmed case     4        NA
+## 4 2022-01-04 Alpine (S) Contact with a confirmed case     4        NA
+## 5 2022-01-05 Alpine (S) Contact with a confirmed case     2        NA
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{lga }\SpecialCharTok{|\textgreater{}} \FunctionTok{head}\NormalTok{(}\DecValTok{5}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## Simple feature collection with 5 features and 7 fields
+## Geometry type: MULTIPOLYGON
+## Dimension:     XY
+## Bounding box:  xmin: 142.3643 ymin: -38.67876 xmax: 147.386 ymax: -36.39269
+## Geodetic CRS:  WGS 84
+##   lga_code_2018            lga state_code_2016 state_name_2016 areasqkm_2018
+## 1         20110     Alpine (S)               2        Victoria     4788.1568
+## 2         20260    Ararat (RC)               2        Victoria     4211.1171
+## 3         20570   Ballarat (C)               2        Victoria      739.0321
+## 4         20660    Banyule (C)               2        Victoria       62.5402
+## 5         20740 Bass Coast (S)               2        Victoria      865.8095
+##   cent_long  cent_lat                       geometry
+## 1  146.9742 -36.85357 MULTIPOLYGON (((146.7258 -3...
+## 2  142.8432 -37.47271 MULTIPOLYGON (((143.1807 -3...
+## 3  143.7815 -37.49286 MULTIPOLYGON (((143.6622 -3...
+## 4  145.0851 -37.73043 MULTIPOLYGON (((145.1357 -3...
+## 5  145.5581 -38.50730 MULTIPOLYGON (((145.5207 -3...
+\end{verbatim}
+
+To cast these two data into a cubble object, supply them as a list with name \code{spatial} and \code{temporal} and other arguments (\code{key}, \code{index}, and \code{coords}).
+\code{as_cubble()} will automatically check the match of spatial and temporal information in the two datasets and warnings will emit if a location has missing spatial or temporal information.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{cb }\OtherTok{\textless{}{-}} \FunctionTok{as\_cubble}\NormalTok{(}\FunctionTok{list}\NormalTok{(}\AttributeTok{spatial =}\NormalTok{ lga, }\AttributeTok{temporal =}\NormalTok{ covid),}
+                \AttributeTok{key =}\NormalTok{ lga, }\AttributeTok{index =}\NormalTok{ date, }\AttributeTok{coords =} \FunctionTok{c}\NormalTok{(cent\_long, cent\_lat))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## ! Some sites in the temporal table don't have corresponding spatial information
+\end{verbatim}
+
+\begin{verbatim}
+## ! Some sites in the spatial table don't have corresponding temporal information
+\end{verbatim}
+
+\begin{verbatim}
+## ! Use argument `output = "unmatch"` to check on the unmatched key
+\end{verbatim}
+
+Here, some locations have been detected with this issue and you can use \code{output = "unmatch"} to check on these locations:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{pair }\OtherTok{\textless{}{-}} \FunctionTok{as\_cubble}\NormalTok{(}\FunctionTok{list}\NormalTok{(}\AttributeTok{spatial =}\NormalTok{ lga, }\AttributeTok{temporal =}\NormalTok{ covid),}
+                \AttributeTok{key =}\NormalTok{ lga, }\AttributeTok{index =}\NormalTok{ date, }\AttributeTok{coords =} \FunctionTok{c}\NormalTok{(cent\_long, cent\_lat),}
+                \AttributeTok{output =} \StringTok{"unmatch"}\NormalTok{)}
+
+\NormalTok{pair}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## $paired
+## # A tibble: 2 x 2
+##   spatial             temporal    
+##   <chr>               <chr>       
+## 1 Kingston (C) (Vic.) Kingston (C)
+## 2 Latrobe (C) (Vic.)  Latrobe (C) 
+## 
+## $others
+## $others$temporal
+## [1] "Interstate"       "Overseas"         "Queenscliffe (B)" "Unknown"         
+## 
+## $others$spatial
+## character(0)
+\end{verbatim}
+
+\pkg{cubble} will make an attempt to pair the unmatched sites as well as showing unpaired locations. These can be useful information to help clean up the data.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{lga }\OtherTok{\textless{}{-}}\NormalTok{ lga }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{lga =} \FunctionTok{ifelse}\NormalTok{(lga }\SpecialCharTok{==} \StringTok{"Kingston (C) (Vic.)"}\NormalTok{, }\StringTok{"Kingston (C)"}\NormalTok{, lga),}
+         \AttributeTok{lga =} \FunctionTok{ifelse}\NormalTok{(lga }\SpecialCharTok{==} \StringTok{"Latrobe (C) (Vic.)"}\NormalTok{, }\StringTok{"Latrobe (C)"}\NormalTok{, lga)) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{filter}\NormalTok{(}\SpecialCharTok{!}\NormalTok{lga }\SpecialCharTok{\%in\%}\NormalTok{ pair}\SpecialCharTok{$}\NormalTok{others}\SpecialCharTok{$}\NormalTok{spatial)}
+
+\NormalTok{covid }\OtherTok{\textless{}{-}}\NormalTok{ covid }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(}\SpecialCharTok{!}\NormalTok{lga }\SpecialCharTok{\%in\%}\NormalTok{ pair}\SpecialCharTok{$}\NormalTok{others}\SpecialCharTok{$}\NormalTok{temporal)}
+\end{Highlighting}
+\end{Shaded}
+
+After cleaning, you can cast the data into cubble with the same syntax without warnings:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{cb }\OtherTok{\textless{}{-}} \FunctionTok{as\_cubble}\NormalTok{(}\AttributeTok{data =} \FunctionTok{list}\NormalTok{(}\AttributeTok{spatial =}\NormalTok{ lga, }\AttributeTok{temporal =}\NormalTok{ covid),}
+                \AttributeTok{key =}\NormalTok{ lga, }\AttributeTok{index =}\NormalTok{ date, }\AttributeTok{coords =} \FunctionTok{c}\NormalTok{(cent\_long, cent\_lat))}
+\end{Highlighting}
+\end{Shaded}
+
 \hypertarget{australia-historical-maximum-temperature}{%
-\subsection{Australia historical maximum
-temperature}\label{australia-historical-maximum-temperature}}
-
-Global Historical Climatology Network (GHCN) provides daily climate
-measures from stations across the world. The dataset
-\code{weatherdata::historical_tmax} extracts the maximum temperature for
-236 Australian stations from GHCN with starting from year 1969.
-\code{weatherdata::historical_tmax} is already in a cubble, with
-\code{id} as the key, \code{date} as the index, and
-\code{c(longitude, latitude)} as the coordinates. This example compares
-the maximum temperature in two periods: 1971 - 1975 and 2016 - 2020 for
-stations in Victoria and New South Wales.
-
-Stations in the two states can be subsetted by the station number:
-Australia GHCN station number starts with ``ASN00'' and followed by the
-\href{http://www.bom.gov.au/climate/cdo/about/site-num.shtml}{Bureau of
-Meteorology (BOM) station number}, where the 2nd and 3rd digit (7th and
-8th in the GHCN number) define the state of the station. New South Wales
-stations start from 46 to 75 and Victoria stations follow from 76 to 90.
-Filtering Victoria and New South Wales stations is an operation in the
-spatial dimension and hence uses the nested form:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ weatherdata}\SpecialCharTok{::}\NormalTok{historical\_tmax }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
+\subsection{Australia historical maximum temperature}\label{australia-historical-maximum-temperature}}
+
+Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world.
+The dataset \code{weatherdata::historical_tmax} extracts the maximum temperature for 236 Australian stations from GHCN with starting from year 1969.
+\code{weatherdata::historical_tmax} is already in a cubble, with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates.
+This example compares the maximum temperature in two periods: 1971 - 1975 and 2016 - 2020 for stations in Victoria and New South Wales.
+
+Stations in the two states can be subsetted by the station number: Australia GHCN station number starts with ``ASN00'' and followed by the \href{http://www.bom.gov.au/climate/cdo/about/site-num.shtml}{Bureau of Meteorology (BOM) station number}, where the 2nd and 3rd digit (7th and 8th in the GHCN number) define the state of the station.
+New South Wales stations start from 46 to 75 and Victoria stations follow from 76 to 90.
+Filtering Victoria and New South Wales stations is an operation in the spatial dimension and hence uses the nested form:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ weatherdata}\SpecialCharTok{::}\NormalTok{historical\_tmax }\SpecialCharTok{|\textgreater{}}
   \FunctionTok{filter}\NormalTok{(}\FunctionTok{between}\NormalTok{(stringr}\SpecialCharTok{::}\FunctionTok{str\_sub}\NormalTok{(id, }\DecValTok{7}\NormalTok{, }\DecValTok{8}\NormalTok{), }\DecValTok{46}\NormalTok{, }\DecValTok{90}\NormalTok{))}
@@
 
-Filtering for the period 1971 \textasciitilde{} 1975 and 2016
-\textasciitilde{} 2020 is an operation on the time dimension and the
-nested cubble needs to be switched to the long cubble by
-\code{stretch()}:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|}\ErrorTok{\textgreater{}} 
-  \FunctionTok{face\_temporal}\NormalTok{() }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
+Filtering for the period 1971 \textasciitilde{} 1975 and 2016 \textasciitilde{} 2020 is an operation on the time dimension and the nested cubble needs to be switched to the long cubble by \code{stretch()}:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|\textgreater{}} 
+  \FunctionTok{face\_temporal}\NormalTok{() }\SpecialCharTok{|\textgreater{}}
   \FunctionTok{filter}\NormalTok{(lubridate}\SpecialCharTok{::}\FunctionTok{year}\NormalTok{(date) }\SpecialCharTok{\%in\%} \FunctionTok{c}\NormalTok{(}\DecValTok{1971}\SpecialCharTok{:}\DecValTok{1975}\NormalTok{, }\DecValTok{2016}\SpecialCharTok{:}\DecValTok{2020}\NormalTok{)) }
@@
 
-A monthly average is used for both periods to smooth the maximum
-temperature and it is again an operation on the time dimension:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
+A monthly average is used for both periods to smooth the maximum temperature and it is again an operation on the time dimension:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|\textgreater{}}
   \FunctionTok{group\_by}\NormalTok{(}\AttributeTok{month =}\NormalTok{ lubridate}\SpecialCharTok{::}\FunctionTok{month}\NormalTok{(date), }
          \AttributeTok{group =} \FunctionTok{as.factor}\NormalTok{(}\FunctionTok{ifelse}\NormalTok{(lubridate}\SpecialCharTok{::}\FunctionTok{year}\NormalTok{(date) }\SpecialCharTok{\textgreater{}} \DecValTok{2015}\NormalTok{, }
-                                  \StringTok{"2016 \textasciitilde{} 2020"}\NormalTok{, }\StringTok{"1971 \textasciitilde{} 1975"}\NormalTok{))) }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
+                                  \StringTok{"2016 \textasciitilde{} 2020"}\NormalTok{, }\StringTok{"1971 \textasciitilde{} 1975"}\NormalTok{))) }\SpecialCharTok{|\textgreater{}}
   \FunctionTok{summarise}\NormalTok{(}\AttributeTok{tmax =} \FunctionTok{mean}\NormalTok{(tmax, }\AttributeTok{na.rm =} \ConstantTok{TRUE}\NormalTok{))}
@@
 
-A few stations don't have records during 1971 - 1975 and further
-investigation shows that while the first and last year of each series is
-recorded, the missing years in this period is not reported. These
-stations are filtered out by examining whether the summarised time
-series has a total of 24 months. The long cubble needs to be switched to
-the nested form for this spatial operation using \code{face_spatial()}:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{face\_spatial}\NormalTok{() }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{filter}\NormalTok{(}\FunctionTok{nrow}\NormalTok{(ts) }\SpecialCharTok{==} \DecValTok{24}\NormalTok{) }
-\end{Highlighting}
-\end{Shaded}
-
-Lastly, to create a glyph map, both the major (\code{longitude},
-\code{latitude}) and minor (\code{month}, \code{tmax}) coordinates need
-to be in the same table. Spatial variables can be moved to the long form
-with \code{migrate()}:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{face\_temporal}\NormalTok{() }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{unfold}\NormalTok{(latitude, longitude)}
-\end{Highlighting}
-\end{Shaded}
-
-\code{tmax} can then be supplied to \code{geom_glyph()} for the glyph
-map in Figure \ref{fig:basic-manip} with a station inset on the top left
-corner:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{nsw\_vic }\OtherTok{\textless{}{-}}\NormalTok{ ozmaps}\SpecialCharTok{::}\NormalTok{abs\_ste }\SpecialCharTok{|}\ErrorTok{\textgreater{}} 
+A few stations don't have records during 1971 - 1975 and further investigation shows that while the first and last year of each series is recorded, the missing years in this period is not reported.
+These stations are filtered out by examining whether the summarised time series has a total of 24 months.
+The long cubble needs to be switched to the nested form for this spatial operation using \code{face_spatial()}:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|\textgreater{}} \FunctionTok{face\_spatial}\NormalTok{() }\SpecialCharTok{|\textgreater{}} \FunctionTok{filter}\NormalTok{(}\FunctionTok{nrow}\NormalTok{(ts) }\SpecialCharTok{==} \DecValTok{24}\NormalTok{) }
+\end{Highlighting}
+\end{Shaded}
+
+Lastly, to create a glyph map, both the major (\code{longitude}, \code{latitude}) and minor (\code{month}, \code{tmax}) coordinates need to be in the same table.
+Spatial variables can be moved to the long form with \code{migrate()}:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|\textgreater{}} \FunctionTok{face\_temporal}\NormalTok{() }\SpecialCharTok{|\textgreater{}} \FunctionTok{unfold}\NormalTok{(latitude, longitude)}
+\end{Highlighting}
+\end{Shaded}
+
+\code{tmax} can then be supplied to \code{geom_glyph()} for the glyph map in Figure \ref{fig:basic-manip} with a station inset on the top left corner:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{nsw\_vic }\OtherTok{\textless{}{-}}\NormalTok{ ozmaps}\SpecialCharTok{::}\NormalTok{abs\_ste }\SpecialCharTok{|\textgreater{}} 
   \FunctionTok{filter}\NormalTok{(NAME }\SpecialCharTok{\%in\%} \FunctionTok{c}\NormalTok{(}\StringTok{"Victoria"}\NormalTok{, }\StringTok{"New South Wales"}\NormalTok{))}
@@
 \hypertarget{australia-precipitation-pattern-in-2020}{%
-\subsection{Australia precipitation pattern in
-2020}\label{australia-precipitation-pattern-in-2020}}
-
-In the previous example, there has already been some overlapping of the
-glyphs for a few stations near (151E, 34S) and (152E, 33S) and this will
-be a problem when mapping more stations in the national level.
-Aggregation can be helpful to group series into clusters before
-visualising the cluster with glyph map. This example shows how to
-organise data at both level with \code{switch_key()}.
-
-\code{weatherdata::climate_full}, also extracted from GHCN, records
-daily precipitation and maximum/ minimum temperature for 639 stations in
-Australia from 2016 to 2020. A simple kmean algorithm based on the
-distance matrix is used to create 20 clusters. This creates
-\code{station_nested} as a station level nested cubble with a cluster
-column indicating the group each station belongs to. More complex
-algorithms can be used for other problem, as long as there is a mapping
-from each station to a cluster.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{station\_nested }\OtherTok{\textless{}{-}}\NormalTok{ weatherdata}\SpecialCharTok{::}\NormalTok{climate\_full }\SpecialCharTok{|}\ErrorTok{\textgreater{}} 
+\subsection{Australia precipitation pattern in 2020}\label{australia-precipitation-pattern-in-2020}}
+
+In the previous example, there has already been some overlapping of the glyphs for a few stations near (151E, 34S) and (152E, 33S) and this will be a problem when mapping more stations in the national level.
+Aggregation can be helpful to group series into clusters before visualising the cluster with glyph map.
+This example shows how to organise data at both level with \code{switch_key()}.
+
+\code{weatherdata::climate_full}, also extracted from GHCN, records daily precipitation and maximum/ minimum temperature for 639 stations in Australia from 2016 to 2020.
+A simple kmean algorithm based on the distance matrix is used to create 20 clusters.
+This creates \code{station_nested} as a station level nested cubble with a cluster column indicating the group each station belongs to.
+More complex algorithms can be used for other problem, as long as there is a mapping from each station to a cluster.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{station\_nested }\OtherTok{\textless{}{-}}\NormalTok{ weatherdata}\SpecialCharTok{::}\NormalTok{climate\_full }\SpecialCharTok{|\textgreater{}} 
   \FunctionTok{mutate}\NormalTok{(}\AttributeTok{cluster =}\NormalTok{ ...)}
@@
 
-To create a group level cubble, use \code{switch_key()} with the new key
-variable, \code{cluster}:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{cluster\_nested }\OtherTok{\textless{}{-}}\NormalTok{ station\_nested }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{switch\_key}\NormalTok{(cluster) }
-\end{Highlighting}
-\end{Shaded}
-
-With the group level cubble, \code{get_centroid()} is useful to compute
-the centroid of each cluster, which will be used as the major axis for
-the glyph map later:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{cluster\_nested }\OtherTok{\textless{}{-}}\NormalTok{ cluster\_nested }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{get\_centroid}\NormalTok{()}
-\end{Highlighting}
-\end{Shaded}
-
-Long form cubble at both levels can be accessed through stretching the
-nested form and with access to both station and cluster level cubbles,
-various plots can be made to understand the cluster. Figure
-\ref{fig:basic-agg} shows two example plots that can be made with this
-data: subplot A is a glyph map made with the cluster level cubble in the
-long form and subplot B inspects the station membership of each cluster
-using the station level cubble in the nested form.
+To create a group level cubble, use \code{switch_key()} with the new key variable, \code{cluster}:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{cluster\_nested }\OtherTok{\textless{}{-}}\NormalTok{ station\_nested }\SpecialCharTok{|\textgreater{}} \FunctionTok{switch\_key}\NormalTok{(cluster) }
+\end{Highlighting}
+\end{Shaded}
+
+With the group level cubble, \code{get_centroid()} is useful to compute the centroid of each cluster, which will be used as the major axis for the glyph map later:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{cluster\_nested }\OtherTok{\textless{}{-}}\NormalTok{ cluster\_nested }\SpecialCharTok{|\textgreater{}} \FunctionTok{get\_centroid}\NormalTok{()}
+\end{Highlighting}
+\end{Shaded}
+
+Long form cubble at both levels can be accessed through stretching the nested form and with access to both station and cluster level cubbles, various plots can be made to understand the cluster.
+Figure \ref{fig:basic-agg} shows two example plots that can be made with this data: subplot A is a glyph map made with the cluster level cubble in the long form and subplot B inspects the station membership of each cluster using the station level cubble in the nested form.
 
@@
 \hypertarget{era5-climate-reanalysis-data}{%
-\subsection{ERA5: climate reanalysis
-data}\label{era5-climate-reanalysis-data}}
+\subsection{ERA5: climate reanalysis data}\label{era5-climate-reanalysis-data}}
 
 ERA5 data (\protect\hyperlink{ref-hersbach2020era5}{Hersbach et al.
-2020}) is the latest reanalysis of global atmosphere, land surface, and
-ocean waves from 1950 onwards and is available in the NetCDF format from
-European Centre for Medium-Range Weather Forecasts (ECMWF). The data can
-be directly downloaded from
-\href{https://cds.climate.copernicus.eu/cdsapp\#!/dataset/reanalysis-era5-pressure-levels?tab=overview}{Copernicus
-Climate Data Store (CDS)} website or programmatically via an R package
-\pkg{ecmwfr} (\protect\hyperlink{ref-ecwmfr}{Hufkens, Stauffer, and
-Campitelli 2019}). The \code{era5-pressure} data contains variable
-\emph{specific humidity} and \emph{geopotential} on the 10 hPa pressure
-level on four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04.
+2020}) is the latest reanalysis of global atmosphere, land surface, and ocean waves from 1950 onwards and is available in the NetCDF format from European Centre for Medium-Range Weather Forecasts (ECMWF).
+The data can be directly downloaded from \href{https://cds.climate.copernicus.eu/cdsapp\#!/dataset/reanalysis-era5-pressure-levels?tab=overview}{Copernicus Climate Data Store (CDS)} website or programmatically via an R package \pkg{ecmwfr} (\protect\hyperlink{ref-ecwmfr}{Hufkens, Stauffer, and Campitelli 2019}).
+The \code{era5-pressure} data contains variable \emph{specific humidity} and \emph{geopotential} on the 10 hPa pressure level on four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04.
 Once downloaded, the data can be read into a cubble as:
@@
 
-Figure \ref{fig:netcdf} reproduces the ERA5 data row of Figure 19 in
-Hersbach et al. (\protect\hyperlink{ref-hersbach2020era5}{2020}). It
-shows the southern polar vortex splits into two on 2002-09-26 and
-further splits into four on 2002-10-04 in the stratosphere. Readers
-interested in the analysis of this figure can refer to Hersbach et al.
+Figure \ref{fig:netcdf} reproduces the ERA5 data row of Figure 19 in Hersbach et al.
+(\protect\hyperlink{ref-hersbach2020era5}{2020}).
+It shows the southern polar vortex splits into two on 2002-09-26 and further splits into four on 2002-10-04 in the stratosphere.
+Readers interested in the analysis of this figure can refer to Hersbach et al.
 (\protect\hyperlink{ref-hersbach2020era5}{2020}), Simmons et al.
@@
 \CommentTok{\# data pre{-}processing}
-\NormalTok{clean }\OtherTok{\textless{}{-}}\NormalTok{ weatherdata}\SpecialCharTok{::}\NormalTok{climate\_full }\SpecialCharTok{|}\ErrorTok{\textgreater{}}\NormalTok{ ...}
+\NormalTok{clean }\OtherTok{\textless{}{-}}\NormalTok{ weatherdata}\SpecialCharTok{::}\NormalTok{climate\_full }\SpecialCharTok{|\textgreater{}}\NormalTok{ ...}
 
 \CommentTok{\# created SharedData instance for crosstalk}
-\NormalTok{nested }\OtherTok{\textless{}{-}}\NormalTok{ clean }\SpecialCharTok{|}\ErrorTok{\textgreater{}}\NormalTok{ SharedData}\SpecialCharTok{$}\FunctionTok{new}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{id, }\AttributeTok{group =} \StringTok{"cubble"}\NormalTok{)}
-\NormalTok{long }\OtherTok{\textless{}{-}} \FunctionTok{face\_temporal}\NormalTok{(clean) }\SpecialCharTok{|}\ErrorTok{\textgreater{}}\NormalTok{ SharedData}\SpecialCharTok{$}\FunctionTok{new}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{id, }\AttributeTok{group =} \StringTok{"cubble"}\NormalTok{)}
+\NormalTok{nested }\OtherTok{\textless{}{-}}\NormalTok{ clean }\SpecialCharTok{|\textgreater{}}\NormalTok{ SharedData}\SpecialCharTok{$}\FunctionTok{new}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{id, }\AttributeTok{group =} \StringTok{"cubble"}\NormalTok{)}
+\NormalTok{long }\OtherTok{\textless{}{-}} \FunctionTok{face\_temporal}\NormalTok{(clean) }\SpecialCharTok{|\textgreater{}}\NormalTok{ SharedData}\SpecialCharTok{$}\FunctionTok{new}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{id, }\AttributeTok{group =} \StringTok{"cubble"}\NormalTok{)}
 
 \CommentTok{\# create the spatial and temporal view each with a ShareData instance}
-\NormalTok{p1 }\OtherTok{\textless{}{-}}\NormalTok{ nested }\SpecialCharTok{|}\ErrorTok{\textgreater{}}\NormalTok{ ...}
-\NormalTok{p2 }\OtherTok{\textless{}{-}}\NormalTok{ long }\SpecialCharTok{|}\ErrorTok{\textgreater{}}\NormalTok{ ...}
+\NormalTok{p1 }\OtherTok{\textless{}{-}}\NormalTok{ nested }\SpecialCharTok{|\textgreater{}}\NormalTok{ ...}
+\NormalTok{p2 }\OtherTok{\textless{}{-}}\NormalTok{ long }\SpecialCharTok{|\textgreater{}}\NormalTok{ ...}
 
@@
 \CommentTok{\# data pre{-}processing}
-\NormalTok{clean }\OtherTok{\textless{}{-}}\NormalTok{ weatherdata}\SpecialCharTok{::}\NormalTok{climate\_full }\SpecialCharTok{|}\ErrorTok{\textgreater{}}\NormalTok{ ...}
+\NormalTok{clean }\OtherTok{\textless{}{-}}\NormalTok{ weatherdata}\SpecialCharTok{::}\NormalTok{climate\_full }\SpecialCharTok{|\textgreater{}}\NormalTok{ ...}
 
@@
 \NormalTok{p }\OtherTok{\textless{}{-}} \FunctionTok{map}\NormalTok{(}\DecValTok{1}\SpecialCharTok{:}\FunctionTok{length}\NormalTok{(df\_id), }\ControlFlowTok{function}\NormalTok{(i)\{}
-\NormalTok{  dt }\OtherTok{\textless{}{-}}\NormalTok{ clean }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{filter}\NormalTok{(id }\SpecialCharTok{==}\NormalTok{ df\_id[i])}
-  \FunctionTok{ggplot}\NormalTok{(dt) }\SpecialCharTok{|}\ErrorTok{\textgreater{}}\NormalTok{ ...}
+\NormalTok{  dt }\OtherTok{\textless{}{-}}\NormalTok{ clean }\SpecialCharTok{|\textgreater{}} \FunctionTok{filter}\NormalTok{(id }\SpecialCharTok{==}\NormalTok{ df\_id[i])}
+  \FunctionTok{ggplot}\NormalTok{(dt) }\SpecialCharTok{|\textgreater{}}\NormalTok{ ...}
 \NormalTok{\})}
@@
 \NormalTok{nested }\OtherTok{\textless{}{-}} \FunctionTok{face\_spatial}\NormalTok{(clean)}
-\FunctionTok{leaflet}\NormalTok{(nested) }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
-  \FunctionTok{addTiles}\NormalTok{() }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
-  \FunctionTok{addCircleMarkers}\NormalTok{(}\AttributeTok{group =} \StringTok{"a"}\NormalTok{, ...) }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
+\FunctionTok{leaflet}\NormalTok{(nested) }\SpecialCharTok{|\textgreater{}}
+  \FunctionTok{addTiles}\NormalTok{() }\SpecialCharTok{|\textgreater{}}
+  \FunctionTok{addCircleMarkers}\NormalTok{(}\AttributeTok{group =} \StringTok{"a"}\NormalTok{, ...) }\SpecialCharTok{|\textgreater{}}
 \NormalTok{  leafpop}\SpecialCharTok{::}\FunctionTok{addPopupGraphs}\NormalTok{(}\AttributeTok{graph =}\NormalTok{ p, ...)}
@@
 
-Figure \ref{fig:interactive-popup} shows Figure
-\ref{fig:interactive-linking} made made with leaflet and popups
-(\protect\hyperlink{ref-leafpop}{Appelhans and Detsch 2021}).
+Figure \ref{fig:interactive-popup} shows Figure \ref{fig:interactive-linking} made made with leaflet and popups (\protect\hyperlink{ref-leafpop}{Appelhans and Detsch 2021}).
 
@@
 
-This work is funded by the Commonwealth Scientific and Industrial
-Research Organisation (CSIRO) Data61 Scholarship and started while
-Nicolas Langrené was affiliated with CSIRO's Data61. The article is
-created using \pkg{knitr} (\protect\hyperlink{ref-knitr}{Xie 2015}) and
-\pkg{rmarkdown} (\protect\hyperlink{ref-rmarkdown}{Xie, Allaire, and
-Grolemund 2018}) in R. The source code for reproducing this paper can be
-found at: \url{https://github.com/huizezhang-sherry/paper-cubble}.
+This work is funded by the Commonwealth Scientific and Industrial Research Organisation (CSIRO) Data61 Scholarship and started while Nicolas Langrené was affiliated with CSIRO's Data61.
+The article is created using \pkg{knitr} (\protect\hyperlink{ref-knitr}{Xie 2015}) and \pkg{rmarkdown} (\protect\hyperlink{ref-rmarkdown}{Xie, Allaire, and Grolemund 2018}) in R.
+The source code for reproducing this paper can be found at: \url{https://github.com/huizezhang-sherry/paper-cubble}.
 
@@
 \leavevmode\vadjust pre{\hypertarget{ref-leafpop}{}}%
-Appelhans, Tim, and Florian Detsch. 2021. \emph{\pkg{leafpop}: Include
-Tables, Images and Graphs in Leaflet Pop-Ups}.
-\url{https://CRAN.R-project.org/package=leafpop}.
+Appelhans, Tim, and Florian Detsch. 2021. \emph{\pkg{leafpop}: Include Tables, Images and Graphs in Leaflet Pop-Ups}. \url{https://CRAN.R-project.org/package=leafpop}.
 
 \leavevmode\vadjust pre{\hypertarget{ref-bach_review_2014}{}}%
-Bach, Benjamin, Pierre Dragicevic, Dragicevic Archambault, Christophe
-Hurter, and Sheelagh Carpendale. 2014. {``A {Review} of {Temporal}
-{Data} {Visualizations} {Based} on {Space}-{Time} {Cube}
-{Operations}.''} \emph{Eurographics Conference on Visualization}, 19.
+Bach, Benjamin, Pierre Dragicevic, Dragicevic Archambault, Christophe Hurter, and Sheelagh Carpendale.
+2014.
+{``A {Review} of {Temporal} {Data} {Visualizations} {Based} on {Space}-{Time} {Cube} {Operations}.''} \emph{Eurographics Conference on Visualization}, 19.
 \url{https://hal.inria.fr/hal-01006140/}.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-buja1988elements}{}}%
-Buja, Andreas, Daniel Asimov, and Catherine Hurley. 1988. {``Elements of
-a Viewing Pipeline.''} \emph{Dynamic Graphics Statistics}, 277.
+Buja, Andreas, Daniel Asimov, and Catherine Hurley. 1988. {``Elements of a Viewing Pipeline.''} \emph{Dynamic Graphics Statistics}, 277.
 
 \leavevmode\vadjust pre{\hypertarget{ref-buja1996interactive}{}}%
-Buja, Andreas, Dianne Cook, and Deborah F Swayne. 1996. {``Interactive
-High-Dimensional Data Visualization.''} \emph{Journal of Computational
-and Graphical Statistics} 5 (1): 78--99.
+Buja, Andreas, Dianne Cook, and Deborah F Swayne.
+1996.
+{``Interactive High-Dimensional Data Visualization.''} \emph{Journal of Computational and Graphical Statistics} 5 (1): 78--99.
 \url{https://doi.org/10.2307/1390754}.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-cheng2016enabling}{}}%
-Cheng, Xiaoyue, Dianne Cook, and Heike Hofmann. 2016. {``Enabling
-Interactivity on Displays of Multivariate Time Series and Longitudinal
-Data.''} \emph{Journal of Computational and Graphical Statistics} 25
-(4): 1057--76. \url{https://doi.org/10.1080/10618600.2015.1105749}.
+Cheng, Xiaoyue, Dianne Cook, and Heike Hofmann.
+2016.
+{``Enabling Interactivity on Displays of Multivariate Time Series and Longitudinal Data.''} \emph{Journal of Computational and Graphical Statistics} 25 (4): 1057--76.
+\url{https://doi.org/10.1080/10618600.2015.1105749}.
 
 \leavevmode\vadjust pre{\hypertarget{ref-cocchi2019data}{}}%
-Cocchi, Marina. 2019. \emph{Data Fusion Methodology and Applications}.
-Elsevier.
+Cocchi, Marina. 2019. \emph{Data Fusion Methodology and Applications}. Elsevier.
+
+\leavevmode\vadjust pre{\hypertarget{ref-ctvspatiotemporal}{}}%
+Edzer Pebesma, Roger Bivand. 2022. {``{CRAN} Task View: Handling and Analyzing Spatio-Temporal Data.''} \url{https://CRAN.R-project.org/view=SpatioTemporal}.
 
 \leavevmode\vadjust pre{\hypertarget{ref-hersbach2020era5}{}}%
-Hersbach, Hans, Bill Bell, Paul Berrisford, Shoji Hirahara, András
-Horányi, Joaquín Muñoz-Sabater, Julien Nicolas, et al. 2020. {``The Era5
-Global Reanalysis.''} \emph{Quarterly Journal of the Royal
-Meteorological Society} 146 (730): 1999--2049.
+Hersbach, Hans, Bill Bell, Paul Berrisford, Shoji Hirahara, András Horányi, Joaquín Muñoz-Sabater, Julien Nicolas, et al.
+2020.
+{``The Era5 Global Reanalysis.''} \emph{Quarterly Journal of the Royal Meteorological Society} 146 (730): 1999--2049.
 
 \leavevmode\vadjust pre{\hypertarget{ref-ecwmfr}{}}%
-Hufkens, Koen, Reto Stauffer, and Elio Campitelli. 2019. {``The
-\pkg{ecwmfr} Package: An Interface to ECMWF API Endpoints.''}
-\url{https://doi.org/10.5281/zenodo.2647541}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-jr_herring_opengis_2011}{}}%
-J.R. Herring. 2011. {``{OpenGIS}® {Implementation} {Standard} for
-{Geographic} Information - {Simple} Feature Access - {Part} 1: {Common}
-Architecture.''} Open Geospatial Consortium Inc.
-\href{http://portal.\%20opengeospatial.org/files/?artifact_id=25355}{http://portal.
-opengeospatial.org/files/?artifact\_id=25355}.
+Hufkens, Koen, Reto Stauffer, and Elio Campitelli. 2019. {``The \pkg{ecwmfr} Package: An Interface to ECMWF API Endpoints.''} \url{https://doi.org/10.5281/zenodo.2647541}.
 
 \leavevmode\vadjust pre{\hypertarget{ref-lu_multidimensional_2018}{}}%
-Lu, Meng, Marius Appel, and Edzer Pebesma. 2018. {``Multidimensional
-{Arrays} for {Analysing} {Geoscientific} {Data}.''} \emph{ISPRS
-International Journal of Geo-Information} 7 (8): 313.
+Lu, Meng, Marius Appel, and Edzer Pebesma.
+2018.
+{``Multidimensional {Arrays} for {Analysing} {Geoscientific} {Data}.''} \emph{ISPRS International Journal of Geo-Information} 7 (8): 313.
 \url{https://doi.org/10.3390/ijgi7080313}.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-mcintosh2018using}{}}%
-McIntosh, Avery I, Helen E Jenkins, Laura F White, Marinus Barnard, Dana
-R Thomson, Tania Dolby, John Simpson, et al. 2018. {``Using Routinely
-Collected Laboratory Data to Identify High Rifampicin-Resistant
-Tuberculosis Burden Communities in the Western Cape Province, South
-Africa: A Retrospective Spatiotemporal Analysis.''} \emph{PLoS Medicine}
-15 (8): e1002638.
+McIntosh, Avery I, Helen E Jenkins, Laura F White, Marinus Barnard, Dana R Thomson, Tania Dolby, John Simpson, et al.
+2018.
+{``Using Routinely Collected Laboratory Data to Identify High Rifampicin-Resistant Tuberculosis Burden Communities in the Western Cape Province, South Africa: A Retrospective Spatiotemporal Analysis.''} \emph{PLoS Medicine} 15 (8): e1002638.
 
 \leavevmode\vadjust pre{\hypertarget{ref-michna2013rnetcdf}{}}%
-Michna, Pavel, and Milton Woods. 2013. {``\pkg{RNetCDF}: A Package for
-Reading and Writing NetCDF Datasets.''} \emph{The R Journal} 5 (2):
-29--36.
+Michna, Pavel, and Milton Woods. 2013. {``\pkg{RNetCDF}: A Package for Reading and Writing NetCDF Datasets.''} \emph{The R Journal} 5 (2): 29--36.
 
 \leavevmode\vadjust pre{\hypertarget{ref-rnetcdf}{}}%
----------. 2021. \emph{\pkg{RNetCDF}: Interface to 'NetCDF' Datasets}.
-\url{https://CRAN.R-project.org/package=RNetCDF}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-tibble}{}}%
-Müller, Kirill, and Hadley Wickham. 2021. \emph{Tibble: Simple Data
-Frames}. \url{https://CRAN.R-project.org/package=tibble}.
+---------. 2021. \emph{\pkg{RNetCDF}: Interface to 'NetCDF' Datasets}. \url{https://CRAN.R-project.org/package=RNetCDF}.
 
 \leavevmode\vadjust pre{\hypertarget{ref-spacetime}{}}%
-Pebesma, Edzer. 2012. {``\pkg{spacetime}: Spatio-Temporal Data in r.''}
-\emph{Journal of Statistical Software} 51 (7): 1--30.
-\url{https://doi.org/10.18637/jss.v051.i07}.
+Pebesma, Edzer. 2012. {``\pkg{spacetime}: Spatio-Temporal Data in r.''} \emph{Journal of Statistical Software} 51 (7): 1--30. \url{https://doi.org/10.18637/jss.v051.i07}.
 
 \leavevmode\vadjust pre{\hypertarget{ref-stars}{}}%
----------. 2021. \emph{\pkg{stars}: Spatiotemporal Arrays, Raster and
-Vector Data Cubes}. \url{https://CRAN.R-project.org/package=stars}.
+---------. 2021. \emph{\pkg{stars}: Spatiotemporal Arrays, Raster and Vector Data Cubes}. \url{https://CRAN.R-project.org/package=stars}.
 
 \leavevmode\vadjust pre{\hypertarget{ref-sf}{}}%
-Pebesma, Edzer J. 2018. {``Simple Features for \proglang{R}:
-Standardized Support for Spatial Vector Data.''} \emph{R Journal} 10
-(1): 439.
+Pebesma, Edzer J. 2018. {``Simple Features for \proglang{R}: Standardized Support for Spatial Vector Data.''} \emph{R Journal} 10 (1): 439.
 
 \leavevmode\vadjust pre{\hypertarget{ref-sp}{}}%
-Pebesma, Edzer, and Roger S Bivand. 2005. {``S Classes and Methods for
-Spatial Data: The \pkg{sp} Package.''} \emph{R News} 5 (2): 9--13.
+Pebesma, Edzer, and Roger S Bivand. 2005. {``S Classes and Methods for Spatial Data: The \pkg{sp} Package.''} \emph{R News} 5 (2): 9--13.
 
 \leavevmode\vadjust pre{\hypertarget{ref-ncdf4}{}}%
-Pierce, David. 2019. \emph{\pkg{ncdf4}: Interface to Unidata netCDF
-(Version 4 or Earlier) Format Data Files}.
-\url{https://CRAN.R-project.org/package=ncdf4}.
+Pierce, David. 2019. \emph{\pkg{ncdf4}: Interface to Unidata netCDF (Version 4 or Earlier) Format Data Files}. \url{https://CRAN.R-project.org/package=ncdf4}.
 
 \leavevmode\vadjust pre{\hypertarget{ref-xts}{}}%
-Ryan, Jeffrey A., and Joshua M. Ulrich. 2020. \emph{\pkg{xts}:
-eXtensible Time Series}. \url{https://CRAN.R-project.org/package=xts}.
+Ryan, Jeffrey A., and Joshua M. Ulrich. 2020. \emph{\pkg{xts}: eXtensible Time Series}. \url{https://CRAN.R-project.org/package=xts}.
 
 \leavevmode\vadjust pre{\hypertarget{ref-simmons2005ecmwf}{}}%
-Simmons, Adrian, Mariano Hortal, Graeme Kelly, Anthony McNally, Agathe
-Untch, and Sakari Uppala. 2005. {``ECMWF Analyses and Forecasts of
-Stratospheric Winter Polar Vortex Breakup: September 2002 in the
-Southern Hemisphere and Related Events.''} \emph{Journal of the
-Atmospheric Sciences} 62 (3): 668--89.
+Simmons, Adrian, Mariano Hortal, Graeme Kelly, Anthony McNally, Agathe Untch, and Sakari Uppala.
+2005.
+{``ECMWF Analyses and Forecasts of Stratospheric Winter Polar Vortex Breakup: September 2002 in the Southern Hemisphere and Related Events.''} \emph{Journal of the Atmospheric Sciences} 62 (3): 668--89.
 \url{https://doi.org/10.1175/JAS-3322.1}.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-simmons2020global}{}}%
-Simmons, Adrian, Cornel Soci, Julien Nicolas, Bill Bell, P. Berrisford,
-Rossana Dragani, Johannes Flemming, et al. 2020. {``Global Stratospheric
-Temperature Bias and Other Stratospheric Aspects of Era5 and Era5.1,''}
-no. 859 (January). \url{https://doi.org/10.21957/rcxqfmg0}.
+Simmons, Adrian, Cornel Soci, Julien Nicolas, Bill Bell, P.
+Berrisford, Rossana Dragani, Johannes Flemming, et al.
+2020.
+{``Global Stratospheric Temperature Bias and Other Stratospheric Aspects of Era5 and Era5.1,''} no.
+859 (January).
+\url{https://doi.org/10.21957/rcxqfmg0}.
 
 \leavevmode\vadjust pre{\hypertarget{ref-stuart2010matching}{}}%
-Stuart, Elizabeth A. 2010. {``Matching Methods for Causal Inference: A
-Review and a Look Forward.''} \emph{Statistical Science} 25 (1): 1.
+Stuart, Elizabeth A. 2010. {``Matching Methods for Causal Inference: A Review and a Look Forward.''} \emph{Statistical Science} 25 (1): 1.
 
 \leavevmode\vadjust pre{\hypertarget{ref-tidync}{}}%
-Sumner, Michael. 2020. \emph{\pkg{tidync}: A Tidy Approach to 'NetCDF'
-Data Exploration and Extraction}.
-\url{https://CRAN.R-project.org/package=tidync}.
+Sumner, Michael. 2020. \emph{\pkg{tidync}: A Tidy Approach to 'NetCDF' Data Exploration and Extraction}. \url{https://CRAN.R-project.org/package=tidync}.
 
 \leavevmode\vadjust pre{\hypertarget{ref-sutherland2000orca}{}}%
-Sutherland, Peter, Anthony Rossini, Thomas Lumley, Nicholas Lewin-Koh,
-Julie Dickerson, Zach Cox, and Dianne Cook. 2000. {``\pkg{Orca}: A
-Visualization Toolkit for High-Dimensional Data.''} \emph{Journal of
-Computational and Graphical Statistics} 9 (3): 509--29.
+Sutherland, Peter, Anthony Rossini, Thomas Lumley, Nicholas Lewin-Koh, Julie Dickerson, Zach Cox, and Dianne Cook.
+2000.
+{``\pkg{Orca}: A Visualization Toolkit for High-Dimensional Data.''} \emph{Journal of Computational and Graphical Statistics} 9 (3): 509--29.
 \url{https://www.tandfonline.com/doi/abs/10.1080/10618600.2000.10474896}.
 
-\leavevmode\vadjust pre{\hypertarget{ref-tsibble}{}}%
-Wang, Earo, Dianne Cook, and Rob J Hyndman. 2020a. {``A New Tidy Data
-Structure to Support Exploration and Modeling of Temporal Data.''}
-\emph{Journal of Computational and Graphical Statistics} 29 (3):
-466--78. \url{https://doi.org/10.1080/10618600.2019.1695624}.
-
 \leavevmode\vadjust pre{\hypertarget{ref-wang2020calendar}{}}%
----------. 2020b. {``Calendar-Based Graphics for Visualizing People's
-Daily Schedules.''} \emph{Journal of Computational and Graphical
-Statistics} 29 (3): 490--502.
+Wang, Earo, Dianne Cook, and Rob J Hyndman. 2020. {``Calendar-Based Graphics for Visualizing People's Daily Schedules.''} \emph{Journal of Computational and Graphical Statistics} 29 (3): 490--502.
 
 \leavevmode\vadjust pre{\hypertarget{ref-tidydata}{}}%
-Wickham, Hadley. 2014. {``Tidy Data.''} \emph{Journal of Statistical
-Software} 59 (10): 1--23. \url{https://doi.org/10.18637/jss.v059.i10}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-tidyverse}{}}%
-Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy
-D'Agostino McGowan, Romain François, Garrett Grolemund, et al. 2019.
-{``Welcome to the {tidyverse}.''} \emph{Journal of Open Source Software}
-4 (43): 1686. \url{https://doi.org/10.21105/joss.01686}.
+Wickham, Hadley. 2014. {``Tidy Data.''} \emph{Journal of Statistical Software} 59 (10): 1--23. \url{https://doi.org/10.18637/jss.v059.i10}.
 
 \leavevmode\vadjust pre{\hypertarget{ref-Wickham2012-yr}{}}%
-Wickham, Hadley, Heike Hofmann, Charlotte Wickham, and Dianne Cook.
-2012. {``Glyph-Maps for Visually Exploring Temporal Patterns in Climate
-Data and Models.''} \emph{Environmetrics} 23 (5): 382--93.
+Wickham, Hadley, Heike Hofmann, Charlotte Wickham, and Dianne Cook. 2012. {``Glyph-Maps for Visually Exploring Temporal Patterns in Climate Data and Models.''} \emph{Environmetrics} 23 (5): 382--93.
 
 \leavevmode\vadjust pre{\hypertarget{ref-knitr}{}}%
-Xie, Yihui. 2015. \emph{Dynamic Documents with \proglang{R} and
-\pkg{knitr}}. 2nd ed. Boca Raton, Florida: Chapman; Hall/CRC.
-\url{https://yihui.name/knitr/}.
+Xie, Yihui. 2015. \emph{Dynamic Documents with \proglang{R} and \pkg{knitr}}. 2nd ed. Boca Raton, Florida: Chapman; Hall/CRC. \url{https://yihui.name/knitr/}.
 
 \leavevmode\vadjust pre{\hypertarget{ref-rmarkdown}{}}%
-Xie, Yihui, J. J. Allaire, and Garrett Grolemund. 2018. \emph{R
-Markdown: The Definitive Guide}. Boca Raton, Florida: Chapman; Hall/CRC.
-\url{https://bookdown.org/yihui/rmarkdown}.
+Xie, Yihui, J. J. Allaire, and Garrett Grolemund. 2018. \emph{R Markdown: The Definitive Guide}. Boca Raton, Florida: Chapman; Hall/CRC. \url{https://bookdown.org/yihui/rmarkdown}.
 
 \leavevmode\vadjust pre{\hypertarget{ref-xie2014reactive}{}}%
-Xie, Yihui, Heike Hofmann, and Xiaoyue Cheng. 2014. {``{Reactive
-Programming for Interactive Graphics}.''} \emph{Statistical Science} 29
-(2): 201--13. \url{https://doi.org/10.1214/14-STS477}.
+Xie, Yihui, Heike Hofmann, and Xiaoyue Cheng. 2014. {``{Reactive Programming for Interactive Graphics}.''} \emph{Statistical Science} 29 (2): 201--13. \url{https://doi.org/10.1214/14-STS477}.
 

### 2022-04-20 jss/cubble-jss.tex
also knit the jss version

@@
 
-\usepackage{amsmath} \usepackage{array} \usepackage{xcolor}
+\usepackage{amsmath} \usepackage{array} \usepackage{xcolor} \newcommand{\sherry}[1]{\textcolor{magenta}{#1}}
 
@@
 
-Spatio-temporal data has a spatial component referring to the location of each observation, a temporal component that is recorded at regular or irregular time intervals, and multiple variables measured at each spatial and temporal value.
+Spatio-temporal data has a spatial component referring to the location of each observation, a temporal component that is recorded at regular or irregular time intervals, and may have multiple variables measured at each spatial and temporal value.
 The analysis of spatio-temporal data can fix the time to explore the spatial distribution, fix the spatial location to explore temporal trends, or dynamically explore the space and time simultaneously.
@@
 
-The \emph{Representing data} section of the CRAN task view {[}ref{]} lists existing spatio-temporal data representation.
+The \emph{Representing data} section of the SpatioTemproal CRAN task view \citep{ctvspatiotemporal} lists existing spatio-temporal data representation.
 Among them, \citet{spacetime} proposes four layouts: full grid layout, sparse grid layout, irregular layout, and trajectory.
-These layouts are implemented in the \pkg{spacetime} package with underlying class \pkg{sp} \citep{sp} and \pkg{xts} \citep{xts} for handling spatial and temporal variables.
-The \pkg{stars} package has been a recent implementation to handle spatio-temporal data in an array.
-The package depends on the \pkg{sf} package and has allowed more access to tidyverse verbs in the workflow.
-
-What remains to be difficult when wrangling spatio-temporal data is that the raw data is not naturally structured in any one of the representation above and the first task is to cast the data into a class.
+These layouts are implemented in the \pkg{spacetime} \citep{spacetime} package with the underlying spatial and temporal class: \pkg{sp} \citep{sp} and \pkg{xts} \citep{xts}.
+The \pkg{stars} \citep{stars} package has been a recent implementation to handle both raster and vector spatio-temporal data using array.
+It depends on the newer spatial package \pkg{sf} \citep{sf} and interfaces with the widely-used tidyverse verbs.
+
+{[}till here{]} One issue with wrangling spatio-temporal data is that the raw data does not naturally structured in any one of the representation above.
 Also, in data collection, spatio-temporal data is less often collected as a whole piece but fragmented spatial and temporal data for analysts to ensemble.
@@
 
-Multivariate spatio-temporal data can be conceptualised using a cubical data model with reference to variables, time and space.
+Spatio-temporal data can be conceptualised using a cubical data model with reference to variables, time and space.
 This naturally motivates the usage of array to represent and store spatio-temporal data, as evident by satellite imageries, large climate models, among others.
 The work by \citet{lu_multidimensional_2018} has discussed the suitability of this representation through array algebra and practical spatio-temporal tasks.
+\textcolor{magenta}{SZ: is this too direct?}
 
@@
 
-This section will present two formats that cubble uses to arrange spatio-temporal data.
-The main cubble functions will then be introduced to illustrate how to work with these two formats in short examples.
-Lastly, a subsection will be dedicated to how existing packages, in spatial and temporal analysis, fit in with cubble.
-
-In cubble, a single data object can be arranged into two forms: nested form and long form.
-Figure \ref{fig:illu-cubble} sketches the two forms with the associated attributes.
-The decision on which form to use in data wrangling is output-oriented, meaning analysts need to first think about whether the output of an operation is identified only by the spatial identifier, or a combination of spatial and temporal identifier.
-The nested cubble is suitable for working with operations that are only identified by the site and this type of operation can be a pure manipulation of time invariant variables, or an operation that summarises time varying variables into site.
-Underneath the nested form, a cubble is built from a \code{rowwise_df} class where each site forms a separate group.
-This structure simplifies the calculation that involves temporal variables by avoiding the use of \code{purrr::map()} syntax when working with list-column.
-
-For those operations whose output involves both a spatial and temporal identifier, long form should be used.
-The long form uses a \code{grouped_df} class to forms all the time of a site as a group.
-Time invariant variables are stored separately as an special attribute of the long cubble.
-This design avoids repeating the spatial variables at each time stamp while not dropping information from spatial variables.
-
-\begin{CodeChunk}
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.002}
-
-}
-
-\caption{An illustration of the nested and long form in cubble.
-The nested form defines each site in a row and nests the time varying variables into a single column \code{ts}.
-The long form cubble uses \code{id} and \code{t} to identify each row and stores the time invariant variables as an attribute, \code{spatial}.}\label{fig:illu-cubble}
-\end{figure}
-\end{CodeChunk}
-
-\hypertarget{create-a-cubble-in-the-nested-form}{%
-\subsection{Create a cubble in the nested form}\label{create-a-cubble-in-the-nested-form}}
-
-To use functions in the \texttt{cubble} package, an analyst will first need to turn the data object into a \code{cubble} class.
-\code{as_cubble()} does this through supplying the three key components: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order of longitude and latitude.
-The use of \code{key} and \code{index} follows the design in the \pkg{tsibble} package and the cubble created by default is in the nested form.
-
-Before the formal examples in Section \ref{examples}, each function introduced in this section will be accompanied by a small example.
-The data used for the small examples is a subset of a larger climate weather station data used in the formal examples and has been simplified to include only five weather stations.
-It contains spatial information of each station: station id, latitude, longitude, elevation, station name and World Meteorology Organisation ID and also daily temporal information of date, maximum and minimum temperature and precipitation for 2020.
-\code{climate_flat} stores this data in format 1) in Figure \ref{fig:illu-input} and the code below creates a cubble out of \code{climate_flat} with \code{id} as the key, \code{date} as the index, and \code{c(long, lat)} as the coordinates:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> cubble_nested <- cubble::climate_flat |>
+This section will first introduce the core functions in the cubble class: \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and \code{unfold()} from subsection \ref{create} to \ref{unfold}.
+The next subsection addresses why bother a new class before subsection \ref{compact} finishing on the compatibility of cubble with existing packages in spatial and temporal analysis.
+
+Each core function is introduced with a short example accompanied and the data \texttt{climate\_flat} is used throughout this section.
+\texttt{climate\_flat} contains five weather stations in Australia with spatial information of each station: station id, latitude, longitude, elevation, station name and World Meteorology Organisation ID and also daily temporal information of date, maximum and minimum temperature and precipitation for 2020.
+Below prints out the first five rows of \texttt{climate\_flat}:
+
+\begin{CodeChunk}
+\begin{CodeOutput}
+# A tibble: 1,830 x 10
+  id            lat  long  elev name         wmo_id date        prcp  tmax  tmin
+  <chr>       <dbl> <dbl> <dbl> <chr>         <dbl> <date>     <dbl> <dbl> <dbl>
+1 ASN00009021 -31.9  116.  15.4 perth airpo~  94610 2020-01-01     0  31.9  15.3
+2 ASN00009021 -31.9  116.  15.4 perth airpo~  94610 2020-01-02     0  24.9  16.4
+3 ASN00009021 -31.9  116.  15.4 perth airpo~  94610 2020-01-03     6  23.2  13  
+4 ASN00009021 -31.9  116.  15.4 perth airpo~  94610 2020-01-04     0  28.4  12.4
+5 ASN00009021 -31.9  116.  15.4 perth airpo~  94610 2020-01-05     0  35.3  11.6
+# ... with 1,825 more rows
+\end{CodeOutput}
+\end{CodeChunk}
+
+\hypertarget{create}{%
+\subsection{Create a cubble in the nested form}\label{create}}
+
+As described in the introduction, spatio-temporal data can come in various formats and shapes.
+A cubble can be created from various raw data, including tibble and its variates multiple tables (an example provided in section \ref{covid}), and netcdf (detailed in section \ref{netcdf}).
+The function \code{as_cubble()} is used to create a cubble with three additional components: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order of longitude and latitude.
+Argument \code{key} and \code{index} follow the wording in \pkg{tsibble} to define temporal order and multiple series while \code{coords} is a space specific argument to locate each site in two columns.
+The code below creates a cubble out of \code{climate_flat} with \code{id} as the key, \code{date} as the index, and \code{c(long, lat)} as the coordinates:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> cubble_nested <- climate_flat |>
 +   as_cubble(key = id, index = date, coords = c(long, lat))
@@
 
-There are a few information in the cubble header: the name of the \code{key} variable: \code{id} and its unique number: 5, the bounding box: \code{[115.97, -32.94, 133.55, -12.42]} , and also the name of variable nested in the \code{ts} column with its type: \code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.
-
-\hypertarget{stretch-a-nested-cubble-into-the-long-form}{%
-\subsection{Stretch a nested cubble into the long form}\label{stretch-a-nested-cubble-into-the-long-form}}
+The cubble header provides some information about this data.
+\code{id} is the variable name to identify each location and there are five unique locations.
+The bounding box is \code{[115.97, -32.94, 133.55, -12.42]} and the names of variable nested in the \code{ts} column, along with its type, are \code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.
+
+The created cubble is built from a \code{rowwise_df} class where each row forms a group and all the temporal variables are nested in a list column, hence it is also called the nested cubble.
+The rowwise structure makes it simpler to calculate on the list using \code{mutate()} and avoid the \code{map()} when working with list column.
+For example calculating the number of raining day can be done by:
+
+\begin{CodeChunk}
+\begin{CodeOutput}
+# cubble:   id [5]: nested form
+# bbox:     [115.97, -32.94, 133.55, -12.42]
+# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
+  id            lat  long  elev name           wmo_id ts       rain_day
+  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>      <int>
+1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tibble>      104
+2 ASN00010311 -31.9  117. 179   york            94623 <tibble>       89
+3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble>       90
+4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble>      106
+5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble>       63
+\end{CodeOutput}
+\end{CodeChunk}
+
+\hypertarget{switch-from-the-nested-cubble-into-the-long-form}{%
+\subsection{Switch from the nested cubble into the long form}\label{switch-from-the-nested-cubble-into-the-long-form}}
 
@@
 
-\hypertarget{tamp-a-long-cubble-back-to-the-nested-form}{%
-\subsection{Tamp a long cubble back to the nested form}\label{tamp-a-long-cubble-back-to-the-nested-form}}
+The long form is built form class \code{groupped_df} where all the observations from the same sites forms a group.
+
+\begin{CodeChunk}
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.008}
+
+}
+
+\caption{An illustration of function \code{face\_temporal} and \code{face\_spatial} in cubble.
+In the first row, \code{face\_temporal} switches a cubble from the nested form into the long form and the focus has switched from the spatial aspect (the side face) to the temporal aspect (the front face).
+In the second row, \code{face\_spatial} switches a cubble back to the nested form from te long form and shift focus back to the spatial aspect.}\label{fig:face}
+\end{figure}
+\end{CodeChunk}
+
+\hypertarget{switch-a-long-cubble-back-to-the-nested-form}{%
+\subsection{Switch a long cubble back to the nested form}\label{switch-a-long-cubble-back-to-the-nested-form}}
 
 Manipulation on the spatial and temporal dimension can be an iterative process and analysts may need to go back and forth between the nested and long cubble.
-\code{tamp()}, an inverse of \code{stretch()}, switches a long cubble to a nested cubble:
+\code{face_spatial()}, the inverse of \code{face_temporal()}, switches the long cubble back to the nested form, as sketched in the second row in Figure \ref{fig:face}:
 
@@
 
-\hypertarget{migrate-spatial-variables-to-a-long-cubble}{%
-\subsection{Migrate spatial variables to a long cubble}\label{migrate-spatial-variables-to-a-long-cubble}}
+\hypertarget{unfold}{%
+\subsection{Flatten spatial variables into the long cubble}\label{unfold}}
 
 Sometimes, analysts may need to apply some variable transformation that involves both the spatial and temporal variable.
-An example of this is the transformation of temporal variables into the spatial dimension in glyph maps (which will be elaborated in section \ref{st_transformation}).
-Cubble allows this operation through \code{unfold()}, which moves the specified spatial variables into the long cubble:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> cubble_mig <- cubble_long |> unfold(long, lat)
-R> cubble_mig
+An example of this is the transformation of temporal variables into the spatial dimension in glyph maps, which will be elaborated in section \ref{st_transformation}.
+This type of operations can be seen as flattening, or \emph{unfolding}, the cube into a 2D data frame.
+Here the function \code{unfold()} moves the spatial variables \code{long} and \code{lat} into the long cubble:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> cubble_unfold <- cubble_long |> unfold(long, lat)
+R> cubble_unfold
 \end{CodeInput}
@@
 \begin{CodeInput}
-R> cubble_mig |> face_spatial() |> face_temporal()
+R> cubble_unfold |> face_spatial() |> face_temporal()
 \end{CodeInput}
@@
 
-\hypertarget{compatibility-with-existing-packages}{%
-\subsection{Compatibility with existing packages}\label{compatibility-with-existing-packages}}
+\hypertarget{compact}{%
+\subsection{Compatibility with existing packages}\label{compact}}
 
@@
 R> # example with a tsibble created from climate_flat
-R> raw <- climate_flat |>
-+   tsibble::as_tsibble(key = id, index = date)
-R> 
-R> dt <-  raw |>
-+   cubble::as_cubble(coords = c(long, lat))
+R> raw <- climate_flat |> tsibble::as_tsibble(key = id, index = date)
+R> dt <-  raw |> cubble::as_cubble(coords = c(long, lat))
 R> dt
@@
 R> # add station-based features in the nested form.
-R> dt |>
-+   mutate(fabletools::features(ts, tmax, list(tmax_mean = mean, tmax_var = var)))
+R> dt |> mutate(fabletools::features(ts, tmax, list(tmax_mean = mean, tmax_var = var)))
 \end{CodeInput}
@@
 
-\hypertarget{victoria-covid-data}{%
-\subsection{Victoria covid data}\label{victoria-covid-data}}
+\hypertarget{covid}{%
+\subsection{Victoria covid data}\label{covid}}
 

### 2022-04-21 2-0-cube.Rmd
draft on cube

@@
 
-Multivariate spatio-temporal data can be conceptualised using a cubical data model with reference to variables, time and space.
-This naturally motivates the usage of array to represent and store spatio-temporal data, as evident by satellite imageries, large climate models, among others.
-The work by @lu_multidimensional_2018 has discussed the suitability of this representation through array algebra and practical spatio-temporal tasks.
-
-The cubical conceptual framework provides a generalisation to the operation can be done on the data for visualisation.
-A paper by @bach_review_2014 has reviewed the temporal data visualisation based on space-time cube operations.
+Spatio-temporal data can be conceptualised using a cubical data model with three axis: variable, time, and space.
+This abstraction can be a useful tool to map operations on spatio-temporal data to operations on the cube.
+@lu_multidimensional_2018 has shown an example of how array operations can be mapped onto the cube including select, scale, reduce, rearrange, and compute.
+The cubical abstract can also be used to generalise visualisation, where @bach_review_2014 has reviewed the temporal data visualisation based on space-time cube operations.
 Notice that the term space-time cube in their article "does not need to involve spatial data", but referring to "an abstract 2D substrate that is used to visualize data at a specific time".
 Despite its main focus is on temporal data, the mindset of abstracting out data representation to construct visualisation is applicable to our spatio-temporal data manipulation and visualisation.
+
+While a cubical conceptual data model is neat to conceptually view the spatio-temporal data, data wrangling is not convenient in a cubical representation or arrays.
+There are a few reasons for this: 1) Some variables only have the spatial dimension and an array representation will need to handle the duplication issue mentioned above.
+2) While arrays are efficient for computation on numerical variables, spatio-temporal data contains more than just numerical variables.
+Wrangling on character strings and specific datetime classes are not easy with arrays.
+3) The 2D operation of creating a new variable, i.e.
+`mutate()`, is not clear defined in an array.
+There are three different ways to add a 1D vector to a cubical array and this doesn't precisely define how the new variable should be added.

### 2022-04-21 arxiv/cubble-arxiv.tex
draft on cube

@@
 The tidy data concept advocates the data to be shaped into 1) one row per observation; 2) one column per variable; and 3) one type of data per table.
-The long form data is preferred over wide data given the downstream software i.e.~\pkg{dplyr} and \pkg{ggplot2} for data wrangling and visualisation.
+The long form data is preferred over wide data given the downstream software i.e.~\pkg{dplyr} (\protect\hyperlink{ref-dplyr}{Wickham et al.
+2022}) and \pkg{ggplot2} (\protect\hyperlink{ref-ggplot2}{Wickham 2016}) for data wrangling and visualisation.
 However, the long form can be inefficient to store feature geometries, especially for large multipolygon over long time period.
@@
 
-Spatio-temporal data can be conceptualised using a cubical data model with reference to variables, time and space.
-This naturally motivates the usage of array to represent and store spatio-temporal data, as evident by satellite imageries, large climate models, among others.
-The work by Lu, Appel, and Pebesma (\protect\hyperlink{ref-lu_multidimensional_2018}{2018}) has discussed the suitability of this representation through array algebra and practical spatio-temporal tasks.
-\textcolor{magenta}{SZ: is this too direct?}
-
-The cubical conceptual framework provides a generalisation to the operation can be done on the data for visualisation.
-A paper by Bach et al.
+Spatio-temporal data can be conceptualised using a cubical data model with three axis: variable, time, and space.
+This abstraction can be a useful tool to map operations on spatio-temporal data to operations on the cube.
+Lu, Appel, and Pebesma (\protect\hyperlink{ref-lu_multidimensional_2018}{2018}) has shown an example of how array operations can be mapped onto the cube including select, scale, reduce, rearrange, and compute.
+The cubical abstract can also be used to generalise visualisation, where Bach et al.
 (\protect\hyperlink{ref-bach_review_2014}{2014}) has reviewed the temporal data visualisation based on space-time cube operations.
@@
 Despite its main focus is on temporal data, the mindset of abstracting out data representation to construct visualisation is applicable to our spatio-temporal data manipulation and visualisation.
+
+While a cubical conceptual data model is neat to conceptually view the spatio-temporal data, data wrangling is not convenient in a cubical representation or arrays.
+There are a few reasons for this: 1) Some variables only have the spatial dimension and an array representation will need to handle the duplication issue mentioned above.
+2) While arrays are efficient for computation on numerical variables, spatio-temporal data contains more than just numerical variables.
+Wrangling on character strings and specific datetime classes are not easy with arrays.
+3) The 2D operation of creating a new variable, i.e.~\texttt{mutate()}, is not clear defined in an array.
+There are three different ways to add a 1D vector to a cubical array and this doesn't precisely define how the new variable should be added.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-ggplot2}{}}%
+---------. 2016. \emph{Ggplot2: Elegant Graphics for Data Analysis}. Springer-Verlag New York. \url{https://ggplot2.tidyverse.org}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-dplyr}{}}%
+Wickham, Hadley, Romain François, Lionel Henry, and Kirill Müller. 2022. \emph{Dplyr: A Grammar of Data Manipulation}. \url{https://CRAN.R-project.org/package=dplyr}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-Wickham2012-yr}{}}%

### 2022-04-21 1-intro.Rmd
introduction done

@@
 
-Spatio-temporal data has a spatial component referring to the location of each observation, a temporal component that is recorded at regular or irregular time intervals, and multiple variables measured at each spatial and temporal value.
-The analysis of spatio-temporal data can fix the time to explore the spatial distribution, fix the spatial location to explore temporal trends, or dynamically explore the space and time simultaneously.
-For these tasks to be carried out easily, there is a need for a spatio-temporal data representation that underpins various operations.
+Spatio-temporal data has a spatial component referring to the location of each observation, a temporal component that is recorded at regular or irregular time intervals, and may have multiple variables measured at each spatial and temporal value.
+With spatio-temporal data, one can fix the time to explore the spatial distribution, fix the spatial location to explore temporal trends, or dynamically explore the space and time simultaneously.
+Underlying all these tasks is a data representation that enables the spatio-temporal data to be sliced and diced in different ways.
 
-The *Representing data* section of the CRAN task view [ref] lists existing spatio-temporal data representation.
-Among them, @spacetime proposes four layouts: full grid layout, sparse grid layout, irregular layout, and trajectory.
-These layouts are implemented in the \pkg{spacetime} package with underlying class \pkg{sp} [@sp] and \pkg{xts} [@xts] for handling spatial and temporal variables.
-The \pkg{stars} package has been a recent implementation to handle spatio-temporal data in an array.
-The package depends on the \pkg{sf} package and has allowed more access to tidyverse verbs in the workflow.
+The SpatioTemproal CRAN task view [@ctvspatiotemporal] has a section *Representing data* that lists existing spatio-temporal data representation in \proglang{R}.
+Among them, @spacetime proposes four spatio-temporal layouts: full grid layout, sparse grid layout, irregular layout, and trajectory.
+These layouts are implemented in the \pkg{spacetime} [@spacetime] package with the underlying spatial and temporal class: \pkg{sp} [@sp] and \pkg{xts} [@xts].
+The \pkg{stars} [@stars] package has been a recent implementation to handle both raster and vector data using spatio-temporal arrays.
+It interfaces with \pkg{sf} [@sf], which adopts the simple features representation of spatial data, and \pkg{tidyverse} [@tidyverse], a suite of tools to wrangle and visualise data in \proglang{R}.
 
-What remains to be difficult when wrangling spatio-temporal data is that the raw data is not naturally structured in any one of the representation above and the first task is to cast the data into a class.
-Also, in data collection, spatio-temporal data is less often collected as a whole piece but fragmented spatial and temporal data for analysts to ensemble.
-For example, map data records the shape of a collection of area of interest is provided by official government website; geo-scientific data records the longitude and latitude coordinate of locations in the area along with other meta data relates to the location; and temporal variables indexed by both location and time.
-For spatio-temporal data analysis, merging all these sources can be challenging and joining these tables can go wrong because of the slightest unmatched of the linking variable from different sources.
+Still, the data representation for spatio-temporal data can be further improved and there are two reasons for this.
+Firstly, the raw data sourced in the wild is less often presented in any one of the representations above, and fitting the raw data into one of the representations can sometimes be not easy.
+More often, spatio-temporal data is collected in separate 2D tables and analysts need to ensemble them into a whole piece before exploring the data.
+Examples of components of spatio-temporal data can be 1) map data records the shape of a collection of area of interest is provided by the official government websites; 2) geo-scientific data records the longitude and latitude coordinates of locations in the area along with other meta data relates to the location, and; 3) temporal variables indexed by both location and time.
 
-Another issue is how the concept of tidy data [@tidydata] should be applied to spatio-temporal data.
-The tidy data concept advocates the data to be shaped into 1) one row per observation; 2) one column per variable; and 3) one type of data per table.
+<!-- For spatio-temporal data analysis, merging all these sources can be challenging and joining these tables can go wrong because of the slightest unmatched of the linking variable from different sources.
+-->
+
+The other reason relates to how the concept of tidy data [@tidydata] should be applied to spatio-temporal data.
+The tidy data concept advocates the data to be shaped into 1) one row per observation; 2) one column per variable, and; 3) one type of data per table.
 The long form data is preferred over wide data given the downstream software i.e.
-\pkg{dplyr} and \pkg{ggplot2} for data wrangling and visualisation.
-However, the long form can be inefficient to store feature geometries, especially for large multipolygon over long time period.
-This poses the question on how to arrange spatial and temporal variables that makes analysing spatio-temporal data easier.
+\pkg{dplyr} [@dplyr] and \pkg{ggplot2} [@ggplot2] for data wrangling and visualisation.
+However, the long form can be inefficient to store feature geometries, especially for large multipolygon over daily or sub-daily data over years, which are extensively collected and handled in time series analysis.
+This poses the question of how to arrange spatial and temporal variables that makes analysing spatio-temporal data easier.
 
-This paper presents a new \proglang{R} package, \pkg{cubble}, to [overcome the problem] with organising variables in different levels when working with spatio-temporal data.
-The data structure focuses on connecting to the upstream data collection and is suitable for temporal data collected at the fixed location over time.
-Among the four spacetime layouts in @spacetime, it can be full grid layout, sparse grid layout, or irregular layout, but trajectory is not within the scope of this work.
-In the package a new data structure, also called \code{cubble}, is proposed to organise spatial and temporal variables as two forms of a single data object so that they can be wrangled separately while synchronised as a whole piece.
-With this new data object, analysts can spend less time writing codes to organise spatial and temporal variables and focus more on the exploratory data analysis itself.
+This paper presents a new \proglang{R} package, \pkg{cubble}, for working with a new spatio-temporal data representation, which addresses the two issues mentioned above.
+In the package, a new data structure, also called \code{cubble}, is proposed to organise spatial and temporal variables as two forms of a single data object so that they can be wrangled separately while synchronised as a whole piece.
+Among the four spacetime layouts in @spacetime, \pkg{cubble} can be applied to full grid layout, sparse grid layout, or irregular layout, but not trajectory, which is outside the scope of this work.
 The software is available from the Comprehensive R Archive Network (CRAN) at [CRAN link].
@@
 Section \@ref(cubble) presents the main design and functionality of \pkg{cubble}.
-Section \@ref(others) explains how cubble deals with more advanced considerations including data with hierarchical structure, data matching, how cubble fits with existing static and interactive visualisation tools, and spatio-temporal data transformation.
+Section \@ref(others) explains how cubble deals with more advanced considerations including data with the hierarchical structure, data matching, how cubble fits with existing static and interactive visualisation tools, and spatio-temporal data transformation.
 Section \@ref(examples) uses Australia weather station data and river level data as examples to demonstrate the use of \pkg{cubble}.
-An example on how \pkg{cubble} handles NetCDF data is also provided.
+An example of how \pkg{cubble} handles NetCDF data is also provided.
 Section \@ref(conclude) concludes the paper.

### 2022-04-21 arxiv/cubble-arxiv.tex
introduction done

@@
 Spatio-temporal data has a spatial component referring to the location of each observation, a temporal component that is recorded at regular or irregular time intervals, and may have multiple variables measured at each spatial and temporal value.
-The analysis of spatio-temporal data can fix the time to explore the spatial distribution, fix the spatial location to explore temporal trends, or dynamically explore the space and time simultaneously.
-For these tasks to be carried out easily, there is a need for a spatio-temporal data representation that underpins various operations.
-
-The \emph{Representing data} section of the SpatioTemproal CRAN task view (\protect\hyperlink{ref-ctvspatiotemporal}{Edzer Pebesma 2022}) lists existing spatio-temporal data representation.
+With spatio-temporal data, one can fix the time to explore the spatial distribution, fix the spatial location to explore temporal trends, or dynamically explore the space and time simultaneously.
+Underlying all these tasks is a data representation that enables the spatio-temporal data to be sliced and diced in different ways.
+
+The SpatioTemproal CRAN task view (\protect\hyperlink{ref-ctvspatiotemporal}{Edzer Pebesma 2022}) has a section \emph{Representing data} that lists existing spatio-temporal data representation in \proglang{R}.
 Among them, E.
-Pebesma (\protect\hyperlink{ref-spacetime}{2012}) proposes four layouts: full grid layout, sparse grid layout, irregular layout, and trajectory.
+Pebesma (\protect\hyperlink{ref-spacetime}{2012}) proposes four spatio-temporal layouts: full grid layout, sparse grid layout, irregular layout, and trajectory.
 These layouts are implemented in the \pkg{spacetime} (\protect\hyperlink{ref-spacetime}{E.
@@
 The \pkg{stars} (\protect\hyperlink{ref-stars}{E.
-Pebesma 2021}) package has been a recent implementation to handle both raster and vector spatio-temporal data using array.
-It depends on the newer spatial package \pkg{sf} (\protect\hyperlink{ref-sf}{E.
+Pebesma 2021}) package has been a recent implementation to handle both raster and vector data using spatio-temporal arrays.
+It interfaces with \pkg{sf} (\protect\hyperlink{ref-sf}{E.
 J.
-Pebesma 2018}) and interfaces with the widely-used tidyverse verbs.
-
-{[}till here{]} One issue with wrangling spatio-temporal data is that the raw data does not naturally structured in any one of the representation above.
-Also, in data collection, spatio-temporal data is less often collected as a whole piece but fragmented spatial and temporal data for analysts to ensemble.
-For example, map data records the shape of a collection of area of interest is provided by official government website; geo-scientific data records the longitude and latitude coordinate of locations in the area along with other meta data relates to the location; and temporal variables indexed by both location and time.
-For spatio-temporal data analysis, merging all these sources can be challenging and joining these tables can go wrong because of the slightest unmatched of the linking variable from different sources.
-
-Another issue is how the concept of tidy data (\protect\hyperlink{ref-tidydata}{Wickham 2014}) should be applied to spatio-temporal data.
-The tidy data concept advocates the data to be shaped into 1) one row per observation; 2) one column per variable; and 3) one type of data per table.
+Pebesma 2018}), which adopts the simple features representation of spatial data, and \pkg{tidyverse} (\protect\hyperlink{ref-tidyverse}{Wickham et al.
+2019}), a suite of tools to wrangle and visualise data in \proglang{R}.
+
+Still, the data representation for spatio-temporal data can be further improved and there are two reasons for this.
+Firstly, the raw data sourced in the wild is less often presented in any one of the representations above, and fitting the raw data into one of the representations can sometimes be not easy.
+More often, spatio-temporal data is collected in separate 2D tables and analysts need to ensemble them into a whole piece before exploring the data.
+Examples of components of spatio-temporal data can be 1) map data records the shape of a collection of area of interest is provided by the official government websites; 2) geo-scientific data records the longitude and latitude coordinates of locations in the area along with other meta data relates to the location, and; 3) temporal variables indexed by both location and time.
+
+The other reason relates to how the concept of tidy data (\protect\hyperlink{ref-tidydata}{Wickham 2014}) should be applied to spatio-temporal data.
+The tidy data concept advocates the data to be shaped into 1) one row per observation; 2) one column per variable, and; 3) one type of data per table.
 The long form data is preferred over wide data given the downstream software i.e.~\pkg{dplyr} (\protect\hyperlink{ref-dplyr}{Wickham et al.
 2022}) and \pkg{ggplot2} (\protect\hyperlink{ref-ggplot2}{Wickham 2016}) for data wrangling and visualisation.
-However, the long form can be inefficient to store feature geometries, especially for large multipolygon over long time period.
-This poses the question on how to arrange spatial and temporal variables that makes analysing spatio-temporal data easier.
-
-This paper presents a new \proglang{R} package, \pkg{cubble}, to {[}overcome the problem{]} with organising variables in different levels when working with spatio-temporal data.
-The data structure focuses on connecting to the upstream data collection and is suitable for temporal data collected at the fixed location over time.
+However, the long form can be inefficient to store feature geometries, especially for large multipolygon over daily or sub-daily data over years, which are extensively collected and handled in time series analysis.
+This poses the question of how to arrange spatial and temporal variables that makes analysing spatio-temporal data easier.
+
+This paper presents a new \proglang{R} package, \pkg{cubble}, for working with a new spatio-temporal data representation, which addresses the two issues mentioned above.
+In the package, a new data structure, also called \code{cubble}, is proposed to organise spatial and temporal variables as two forms of a single data object so that they can be wrangled separately while synchronised as a whole piece.
 Among the four spacetime layouts in E.
-Pebesma (\protect\hyperlink{ref-spacetime}{2012}), it can be full grid layout, sparse grid layout, or irregular layout, but trajectory is not within the scope of this work.
-In the package a new data structure, also called \code{cubble}, is proposed to organise spatial and temporal variables as two forms of a single data object so that they can be wrangled separately while synchronised as a whole piece.
-With this new data object, analysts can spend less time writing codes to organise spatial and temporal variables and focus more on the exploratory data analysis itself.
+Pebesma (\protect\hyperlink{ref-spacetime}{2012}), \pkg{cubble} can be applied to full grid layout, sparse grid layout, or irregular layout, but not trajectory, which is outside the scope of this work.
 The software is available from the Comprehensive R Archive Network (CRAN) at {[}CRAN link{]}.
@@
 Section \ref{cubble} presents the main design and functionality of \pkg{cubble}.
-Section \ref{others} explains how cubble deals with more advanced considerations including data with hierarchical structure, data matching, how cubble fits with existing static and interactive visualisation tools, and spatio-temporal data transformation.
+Section \ref{others} explains how cubble deals with more advanced considerations including data with the hierarchical structure, data matching, how cubble fits with existing static and interactive visualisation tools, and spatio-temporal data transformation.
 Section \ref{examples} uses Australia weather station data and river level data as examples to demonstrate the use of \pkg{cubble}.
-An example on how \pkg{cubble} handles NetCDF data is also provided.
+An example of how \pkg{cubble} handles NetCDF data is also provided.
 Section \ref{conclude} concludes the paper.
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-tidyverse}{}}%
+Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy D'Agostino McGowan, Romain François, Garrett Grolemund, et al.
+2019.
+{``Welcome to the {tidyverse}.''} \emph{Journal of Open Source Software} 4 (43): 1686.
+\url{https://doi.org/10.21105/joss.01686}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-dplyr}{}}%

### 2022-04-21 2-the-cubble-pkg.Rmd
before section 3.4 + update diagram

@@
 
-This section will first introduce the core functions in the cubble class: \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and \code{unfold()} from subsection \@ref(create) to \@ref(unfold).
-The next subsection addresses why bother a new class before subsection \@ref(compact) finishing on the compatibility of cubble with existing packages in spatial and temporal analysis.
+This section will first introduce the core functions in the cubble class: \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and \code{unfold()} in subsection \@ref(create) to \@ref(unfold).
+The next subsection addresses the necessity of creating a new class.
+Subsection \@ref(compact) will then show the compatibility of cubble with existing packages in spatial and temporal analysis.
 
@@
 
-## Create a cubble in the nested form {#create}
+## Create a cubble {#create}
 
-As described in the introduction, spatio-temporal data can come in various formats and shapes.
-A cubble can be created from various raw data, including tibble and its variates multiple tables (an example provided in section \@ref(covid)), and netcdf (detailed in section \@ref(netcdf)).
+Spatio-temporal data can come in various formats and shapes and a cubble can be created from various raw data.
+This includes tibble and its variates multiple tables (an example provided in section \@ref(covid)), and netcdf (detailed in section \@ref(netcdf)).
 The function \code{as_cubble()} is used to create a cubble with three additional components: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order of longitude and latitude.
-Argument \code{key} and \code{index} follow the wording in \pkg{tsibble} to define temporal order and multiple series while \code{coords} is a space specific argument to locate each site in two columns.
+Argument \code{key} and \code{index} follow the wording in \pkg{tsibble} to describe the temporal order and multiple series while \code{coords} is a space specific argument to locate each site in two columns.
 The code below creates a cubble out of \code{climate_flat} with \code{id} as the key, \code{date} as the index, and \code{c(long, lat)} as the coordinates:
@@
 \code{id} is the variable name to identify each location and there are five unique locations.
-The bounding box is \code{[115.97, -32.94, 133.55, -12.42]} and the names of variable nested in the \code{ts} column, along with its type, are \code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.
+The bounding box is \code{[115.97, -32.94, 133.55, -12.42]}.
+In the third row is the the name and type of variable nested in the \code{ts} column and here it includes \code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.
 
-The created cubble is built from a \code{rowwise_df} class where each row forms a group and all the temporal variables are nested in a list column, hence it is also called the nested cubble.
-The rowwise structure makes it simpler to calculate on the list using \code{mutate()} and avoid the \code{map()} when working with list column.
-For example calculating the number of raining day can be done by:
+The created cubble is built from a \code{rowwise_df} class where each row forms a group.
+All the temporal variables are nested in a list column, hence it is also called the nested cubble.
+The rowwise structure makes it simpler to calculate on the list using the \code{mutate()} syntax, which is simplier than the {\code{map()} when working with list column.
+For example, calculating the number of raining day can be done by:
 
-```{r}
+```{r echo = TRUE}
 cubble_nested |> 
@@
 
-## Switch from the nested cubble into the long form
+## Change focus to the time-variable face
 
-The nested format is convenient for those operations whose output does not contain a time dimension.
-For those outputs that are cross-identified by the spatial and temporal identifier, a long cubble needs to be used.
-The function \code{face_temporal()} switches the cubble from the nested form to the long form and the first row in Figure \ref{fig:face} sketches the function in a diagram.
-This code shows how to switch the cubble just created into its long form:
+The nested form can be used for those operations whose output is only indexed by the spatial identifier (\code{key}), while for those outputs cross-identified by the spatial and temporal identifier (\code{key} and \code{index}), the nested form becomes inadequate.
+\pkg{cubble} also provides a long form, which expands the \code{ts} column and temporarily "hide" the spatial variables.
+The function \code{face_temporal()} is used to switch a nested cubble into a long one and the first row in Figure \ref{fig:face} illustrate this operation where the focus of the cube now changes from the site-variable face to the time-variable face.
+This code switches the cubble just created into its long form:
 
@@
 
-Notice that the third line in the header now shows the name and type of spatial variables: \code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]} and these variables are now stored as a \code{spatial} attribute of the data:
-
-```{r echo = TRUE}
-attr(cubble_long, "spatial")
-```
-
-The long form is built form class \code{groupped_df} where all the observations from the same sites forms a group.
+The first line in the header now shows the cubble is in its long form and the third line has been changed to show the name and type of spatial variables: \code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]}.
+Unlike the nested form buildin form a \code{rowwise_df}, the long cubble is built form class \code{groupped_df} where all the observations from the same sites forms a group.
 
@@
 
-## Switch a long cubble back to the nested form
+## Change focus back to the site-variable face
 
-Manipulation on the spatial and temporal dimension can be an iterative process and analysts may need to go back and forth between the nested and long cubble.
-\code{face_spatial()}, the inverse of \code{face_temporal()}, switches the long cubble back to the nested form, as sketched in the second row in Figure \ref{fig:face}:
+Wrangling spatio-temproal data can be an iterative process on the spatial and temporal dimension.
+Switching the focus back to the site-variable face can be accomplished by the function \code{face_spatial()}, which is the inverse of \code{face_temporal()}.
+The second row of Figure \ref{fig:face} illustrates  the function and below is an example on the climate data:
 
@@
 cubble_back
+identical(cubble_nested, cubble_back)
 ```
@@
 
-This function should generally be used in the last step of the analysis since it is a temporary operation, meaning these added spatial variables are not stored in the long form and will disappear if switched to the nested form and then switched back:
+This function should generally be used in the last step of the analysis since it is a temporary operation, meaning these added spatial variables will disappear if switched to the nested form and then switched back:
 
@@
 
-It may naturally raise the question that why bother a new data structure since the nested form can already be created by the \pkg{dplyr} function \code{nest_by()}?
+It may naturally raise the question that why bother a new data structure since the nested form can already be created by \code{dplyr::nest_by()}?
 While \code{nest_by()} and \code{unnest()} can mimic the nested form in cubble, cubble can be seen as an attempt to arrange variables from different observational units into a single object, specifically for spatio-temporal data.

### 2022-04-21 arxiv/cubble-arxiv.tex
before section 3.4 + update diagram

@@
 
-This section will first introduce the core functions in the cubble class: \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and \code{unfold()} from subsection \ref{create} to \ref{unfold}.
-The next subsection addresses why bother a new class before subsection \ref{compact} finishing on the compatibility of cubble with existing packages in spatial and temporal analysis.
+This section will first introduce the core functions in the cubble class: \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and \code{unfold()} in subsection \ref{create} to \ref{unfold}.
+The next subsection addresses the necessity of creating a new class.
+Subsection \ref{compact} will then show the compatibility of cubble with existing packages in spatial and temporal analysis.
 
@@
 \hypertarget{create}{%
-\subsection{Create a cubble in the nested form}\label{create}}
-
-As described in the introduction, spatio-temporal data can come in various formats and shapes.
-A cubble can be created from various raw data, including tibble and its variates multiple tables (an example provided in section \ref{covid}), and netcdf (detailed in section \ref{netcdf}).
+\subsection{Create a cubble}\label{create}}
+
+Spatio-temporal data can come in various formats and shapes and a cubble can be created from various raw data.
+This includes tibble and its variates multiple tables (an example provided in section \ref{covid}), and netcdf (detailed in section \ref{netcdf}).
 The function \code{as_cubble()} is used to create a cubble with three additional components: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order of longitude and latitude.
-Argument \code{key} and \code{index} follow the wording in \pkg{tsibble} to define temporal order and multiple series while \code{coords} is a space specific argument to locate each site in two columns.
+Argument \code{key} and \code{index} follow the wording in \pkg{tsibble} to describe the temporal order and multiple series while \code{coords} is a space specific argument to locate each site in two columns.
 The code below creates a cubble out of \code{climate_flat} with \code{id} as the key, \code{date} as the index, and \code{c(long, lat)} as the coordinates:
@@
 \code{id} is the variable name to identify each location and there are five unique locations.
-The bounding box is \code{[115.97, -32.94, 133.55, -12.42]} and the names of variable nested in the \code{ts} column, along with its type, are \code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.
-
-The created cubble is built from a \code{rowwise_df} class where each row forms a group and all the temporal variables are nested in a list column, hence it is also called the nested cubble.
-The rowwise structure makes it simpler to calculate on the list using \code{mutate()} and avoid the \code{map()} when working with list column.
-For example calculating the number of raining day can be done by:
+The bounding box is \code{[115.97, -32.94, 133.55, -12.42]}.
+In the third row is the the name and type of variable nested in the \code{ts} column and here it includes \code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.
+
+The created cubble is built from a \code{rowwise_df} class where each row forms a group.
+All the temporal variables are nested in a list column, hence it is also called the nested cubble.
+The rowwise structure makes it simpler to calculate on the list using the \code{mutate()} syntax, which is simplier than the \{\code{map()} when working with list column.
+For example, calculating the number of raining day can be done by:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{cubble\_nested }\SpecialCharTok{|\textgreater{}} 
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{rain\_day =} \FunctionTok{sum}\NormalTok{(ts}\SpecialCharTok{$}\NormalTok{prcp }\SpecialCharTok{!=} \DecValTok{0}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\hypertarget{switch-from-the-nested-cubble-into-the-long-form}{%
-\subsection{Switch from the nested cubble into the long form}\label{switch-from-the-nested-cubble-into-the-long-form}}
-
-The nested format is convenient for those operations whose output does not contain a time dimension.
-For those outputs that are cross-identified by the spatial and temporal identifier, a long cubble needs to be used.
-The function \code{face_temporal()} switches the cubble from the nested form to the long form and the first row in Figure \ref{fig:face} sketches the function in a diagram.
-This code shows how to switch the cubble just created into its long form:
+\hypertarget{change-focus-to-the-time-variable-face}{%
+\subsection{Change focus to the time-variable face}\label{change-focus-to-the-time-variable-face}}
+
+The nested form can be used for those operations whose output is only indexed by the spatial identifier (\code{key}), while for those outputs cross-identified by the spatial and temporal identifier (\code{key} and \code{index}), the nested form becomes inadequate.
+\pkg{cubble} also provides a long form, which expands the \code{ts} column and temporarily ``hide'' the spatial variables.
+The function \code{face_temporal()} is used to switch a nested cubble into a long one and the first row in Figure \ref{fig:face} illustrate this operation where the focus of the cube now changes from the site-variable face to the time-variable face.
+This code switches the cubble just created into its long form:
 
@@
 
-Notice that the third line in the header now shows the name and type of spatial variables: \code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]} and these variables are now stored as a \code{spatial} attribute of the data:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\FunctionTok{attr}\NormalTok{(cubble\_long, }\StringTok{"spatial"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## # A tibble: 5 x 6
-##   id            lat  long  elev name           wmo_id
-##   <chr>       <dbl> <dbl> <dbl> <chr>           <dbl>
-## 1 ASN00009021 -31.9  116.  15.4 perth airport   94610
-## 2 ASN00010311 -31.9  117. 179   york            94623
-## 3 ASN00010614 -32.9  117. 338   narrogin        94627
-## 4 ASN00014015 -12.4  131.  30.4 darwin airport  94120
-## 5 ASN00015131 -17.6  134. 220   elliott         94236
-\end{verbatim}
-
-The long form is built form class \code{groupped_df} where all the observations from the same sites forms a group.
+The first line in the header now shows the cubble is in its long form and the third line has been changed to show the name and type of spatial variables: \code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]}.
+Unlike the nested form buildin form a \code{rowwise_df}, the long cubble is built form class \code{groupped_df} where all the observations from the same sites forms a group.
 
@@
 
-\hypertarget{switch-a-long-cubble-back-to-the-nested-form}{%
-\subsection{Switch a long cubble back to the nested form}\label{switch-a-long-cubble-back-to-the-nested-form}}
-
-Manipulation on the spatial and temporal dimension can be an iterative process and analysts may need to go back and forth between the nested and long cubble.
-\code{face_spatial()}, the inverse of \code{face_temporal()}, switches the long cubble back to the nested form, as sketched in the second row in Figure \ref{fig:face}:
+\hypertarget{change-focus-back-to-the-site-variable-face}{%
+\subsection{Change focus back to the site-variable face}\label{change-focus-back-to-the-site-variable-face}}
+
+Wrangling spatio-temproal data can be an iterative process on the spatial and temporal dimension.
+Switching the focus back to the site-variable face can be accomplished by the function \code{face_spatial()}, which is the inverse of \code{face_temporal()}.
+The second row of Figure \ref{fig:face} illustrates the function and below is an example on the climate data:
 
@@
 ## 5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble [366 x 4]>
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{identical}\NormalTok{(cubble\_nested, cubble\_back)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## [1] TRUE
 \end{verbatim}

### 2022-04-23 2-0-cube.Rmd
section 2 done

--- 2022-04-18 2-0-cube.Rmd
@@
 
-Multivariate spatio-temporal data can be conceptualised using a cubical data model with reference to variables, time and space.
-This naturally motivates the usage of array to represent and store spatio-temporal data, as evident by satellite imageries, large climate models, among others.
-The work by @lu_multidimensional_2018 has discussed the suitability of this representation through array algebra and practical spatio-temporal tasks.
-
-The cubical conceptual framework provides a generalisation to the operation can be done on the data for visualisation.
-A paper by @bach_review_2014 has reviewed the temporal data visualisation based on space-time cube operations.
+Spatio-temporal data can be conceptualised using a cubical data model with three axis.
+This abstraction can be useful to generalise operations or even visualisation: @lu_multidimensional_2018 shows an example of how array operations (select, scale, reduce, rearrange, and compute) can be mapped onto the cube; @bach_review_2014 reviews the temporal data visualisation based on space-time cube operations.
 Notice that the term space-time cube in their article "does not need to involve spatial data", but referring to "an abstract 2D substrate that is used to visualize data at a specific time".
 Despite its main focus is on temporal data, the mindset of abstracting out data representation to construct visualisation is applicable to our spatio-temporal data manipulation and visualisation.
+
+While a cubical conceptual data model is neat to conceptually view the spatio-temporal data, a 3D data representation, or an array,  may not be convenient for data wrangling.
+There are two reasons for this: 1) While arrays are efficient for computation on numerical values, spatio-temporal data contains more than just numerical variables.
+Wrangling on character strings and specific datetime classes are common tasks for spatio-temporal data but are not easy with arrays.
+2) The `mutate()` syntax for creating new variables is no longer valid since the operation can no longer define how a created 1D vector should be added to the array.
+
+There has been many cubes defined using time, latitude, and longitude, which can be seen as taking snapshots of the space and stack on time.
+Our spatio-temporal cube is defined slightly different with three axes: time, site, and variable  as Figure \ref{fig:cube-def}.
+The time axis is the same across the two formation, site encompasses latitude and longitude into a single column identifier, while variables are stacked on this defined space-time canvas.
+This notion is adopted to avoid the use of hypercube when describing multivariate spatio-temporal data.
+
+```{r cube-def, echo = FALSE, fig.align="center", out.width = "50%", fig.cap = "An ilustration of the conceputalised spatio-temporal cube with three axies: time, site, and variable."}
+knitr::include_graphics(here::here("figures/cube-def.png"))
+```

### 2022-04-23 arxiv/cubble-arxiv.tex
section 2 done

@@
 
-Spatio-temporal data can be conceptualised using a cubical data model with three axis: variable, time, and space.
-This abstraction can be a useful tool to map operations on spatio-temporal data to operations on the cube.
-Lu, Appel, and Pebesma (\protect\hyperlink{ref-lu_multidimensional_2018}{2018}) has shown an example of how array operations can be mapped onto the cube including select, scale, reduce, rearrange, and compute.
-The cubical abstract can also be used to generalise visualisation, where Bach et al.
-(\protect\hyperlink{ref-bach_review_2014}{2014}) has reviewed the temporal data visualisation based on space-time cube operations.
+Spatio-temporal data can be conceptualised using a cubical data model with three axis.
+This abstraction can be useful to generalise operations or even visualisation: Lu, Appel, and Pebesma (\protect\hyperlink{ref-lu_multidimensional_2018}{2018}) shows an example of how array operations (select, scale, reduce, rearrange, and compute) can be mapped onto the cube; Bach et al.
+(\protect\hyperlink{ref-bach_review_2014}{2014}) reviews the temporal data visualisation based on space-time cube operations.
 Notice that the term space-time cube in their article ``does not need to involve spatial data'', but referring to ``an abstract 2D substrate that is used to visualize data at a specific time''.
@@
 
-While a cubical conceptual data model is neat to conceptually view the spatio-temporal data, data wrangling is not convenient in a cubical representation or arrays.
-There are a few reasons for this: 1) Some variables only have the spatial dimension and an array representation will need to handle the duplication issue mentioned above.
-2) While arrays are efficient for computation on numerical variables, spatio-temporal data contains more than just numerical variables.
-Wrangling on character strings and specific datetime classes are not easy with arrays.
-3) The 2D operation of creating a new variable, i.e.~\texttt{mutate()}, is not clear defined in an array.
-There are three different ways to add a 1D vector to a cubical array and this doesn't precisely define how the new variable should be added.
+While a cubical conceptual data model is neat to conceptually view the spatio-temporal data, a 3D data representation, or an array, may not be convenient for data wrangling.
+There are two reasons for this: 1) While arrays are efficient for computation on numerical values, spatio-temporal data contains more than just numerical variables.
+Wrangling on character strings and specific datetime classes are common tasks for spatio-temporal data but are not easy with arrays.
+2) The \texttt{mutate()} syntax for creating new variables is no longer valid since the operation can no longer define how a created 1D vector should be added to the array.
+
+There has been many cubes defined using time, latitude, and longitude, which can be seen as taking snapshots of the space and stack on time.
+Our spatio-temporal cube is defined slightly different with three axes: time, site, and variable as Figure \ref{fig:cube-def}.
+The time axis is the same across the two formation, site encompasses latitude and longitude into a single column identifier, while variables are stacked on this defined space-time canvas.
+This notion is adopted to avoid the use of hypercube when describing multivariate spatio-temporal data.
+
+\begin{figure}
+
+{\centering \includegraphics[width=0.5\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/cube-def}
+
+}
+
+\caption{An ilustration of the conceputalised spatio-temporal cube with three axies: time, site, and variable.}\label{fig:cube-def}
+\end{figure}
 

### 2022-04-23 2-the-cubble-pkg.Rmd
section 3 done

@@
 
-It may naturally raise the question that why bother a new data structure since the nested form can already be created by \code{dplyr::nest_by()}?
-While \code{nest_by()} and \code{unnest()} can mimic the nested form in cubble, cubble can be seen as an attempt to arrange variables from different observational units into a single object, specifically for spatio-temporal data.
-While `nest_by()` solves this problem on the spatial side, it doesn't solve the temporal side.
-For temporal data, a combined representation with spatial variables is useful for spatio-temporal operations but its duplication can be a burden to operate on.
-Also, spatial variables are dropped when summarised and you will need to either specified all of them in the \code{group_by()} before \code{summarise}, or add a joining after \code{summarise()}.
+Some readers may raise the question why a new data structure rather than using \code{dplyr::nest_by()} on the combined data to directly create a list-column.
+While \code{nest_by()} and its pair \code{unnest()} can mimic the nested form in cubble, cubble can be seen as an attempt to arrange variables from different observational units into a single object, specifically for spatio-temporal data.
+While `nest_by()` can recreate the nested form, it can't create a clean slate for temporal variables as the long form.
+For temporal variables, carrying the duplicated spatial variables in the combined form is not convenient to work with.
+This is because these spatial variables will lost after \code{summarise()} and one will need to subset the spatial variables again from the origin data and make a left join, which tends to create unnecessary intermediate products in the workflow.
+
+```{r eval = FALSE, echo = TRUE}
+# operation on the long form w/o cubble
+# cubble workflow
+dt |> 
+  as_cubble(...) |> 
+  face_temporal() |> 
+  group_by() |> 
+  summarise() |> 
+  unfold(...)
+
+# tidyverse workflow
+temporal <- dt |> 
+  group_by(...) |> 
+  summarise(...)
+
+spatial <- dt |> select(....)
+
+spatial |> left_join(temporal)
+```
 
@@
 As a spatial data object, \code{sf} creates a simple feature geometry list-column (\code{sfc}) in the data frame to provide spatial operations on various geometry types (\code{POINT}, \code{LINESTRING}, \code{POLYGON}, \code{MULTIPOLYGON}, etc).
-These spatial operations are also valuable for spatio-temporal data analysis, but an \code{sf} object *usually* does not contain temporal variables.
+An \code{sf} object can be supplied as the spatial table, as part of the raw data, along with a temporal table, to create a cubble.
+Once formed in a cubble, methods for the \code{sfc} class can be applied on the nested form.
+An example of this is shown in \@ref(covid), which also handles the case when the site in the two tables don't match exactly.
+A spatial data object with \code{s2} vector can also be the input of the spatial table.
+
+<!-- These spatial operations are also valuable for spatio-temporal data analysis, but an \code{sf} object *usually* does not contain temporal variables.
 This means \code{sf} cannot be directly cast into a \code{cubble}, however, \code{cubble} does support \code{sfc} columns in the nested form and spatial operations applied to the \code{sfc} column in \code{sf} can still be applied to the \code{sfc} column in a cubble.
@@
 Then \code{sf::st_within} is used to add the state \code{MULTIPOLYGON} of each weather station before a coordinate transformation is made.
+-->
 
-```{r echo = TRUE}
-library(sf)
-# create a cubble
-cb <- climate_flat |>
-  cubble::as_cubble(key = id, index = date, coords = c(long, lat))
+<!-- ```{r echo = TRUE} -->
+<!-- library(sf) -->
+<!-- # create a cubble -->
+<!-- cb <- climate_flat |> -->
+<!--   cubble::as_cubble(key = id, index = date, coords = c(long, lat)) -->
 
-aus <- ozmaps::abs_ste
+<!-- aus <- ozmaps::abs_ste -->
 
-dt <- cb |>
-  mutate(
-    # create `sfc` column based on long and lat
-    ll = st_sfc(
-      purrr::map2(long, lat, ~st_point(c(.x, .y))),
-      crs = st_crs(aus)),
+<!-- dt <- cb |> -->
+<!--   mutate( -->
+<!--     # create `sfc` column based on long and lat -->
+<!--     ll = st_sfc( -->
+<!--       purrr::map2(long, lat, ~st_point(c(.x, .y))), -->
+<!--       crs = st_crs(aus)), -->
 
-    # append state multi-polygon based on the `sfc` created
-    state = aus$geometry[st_within(ll, aus, sparse = FALSE)],
+<!--     # append state multi-polygon based on the `sfc` created -->
+<!--     state = aus$geometry[st_within(ll, aus, sparse = FALSE)], -->
 
-    # adopt a different projection: lambert conformal conic (EPSG:3112)
-    state = st_transform(state, crs = "EPSG:3112")
-    )
+<!--     # adopt a different projection: lambert conformal conic (EPSG:3112) -->
+<!--     state = st_transform(state, crs = "EPSG:3112") -->
+<!--     ) -->
 
-dt
-```
+<!-- dt -->
+<!-- ``` -->
 
-An \pkg{s2} \code{lnglat} vector can similarly be created as an \code{sfc} in cubble before using any \code{s2}-prefixed function:
+<!-- An \pkg{s2} \code{lnglat} vector can similarly be created as an \code{sfc} in cubble before using any \code{s2}-prefixed function: -->
 
-```{r echo = TRUE}
-library(s2)
-# Western Australia map
-wa <- ozmaps::abs_ste |> filter(NAME == "Western Australia")
+<!-- ```{r echo = TRUE} -->
+<!-- library(s2) -->
+<!-- # Western Australia map -->
+<!-- wa <- ozmaps::abs_ste |> filter(NAME == "Western Australia") -->
 
-# mutate a `s2_lnglat` vector on `cb` created in the last chunk
-cb |>
-  mutate(ll = s2_lnglat(long, lat)) |>
-  filter(s2_within(ll, wa))
-```
+<!-- # mutate a `s2_lnglat` vector on `cb` created in the last chunk -->
+<!-- cb |> -->
+<!--   mutate(ll = s2_lnglat(long, lat)) |> -->
+<!--   filter(s2_within(ll, wa)) -->
+<!-- ``` -->
 
@@
 
-NetCDF data has two main components: *dimension* for defining the spatio-temporal grid (longitude, latitude, and time) and *variable* that populates the defined grid.
+NetCDF data is another format that is commonly used for storing spatio-temporal data.
+It has two main components: *dimension* for defining the spatio-temporal grid (longitude, latitude, and time) and *variable* that populates the defined grid.
 Attributes are usually associated with dimension and variable in the NetCDF format data and a [metadata convention for climate and forecast](http://cfconventions.org/) has been designed to standardise the format of the attributes.
@@
 
-The memory limit with NetCDF data in cubble depends on the number of longitude grid points times the number of latitude grid points times the number of time grid points times the number of variables.
-Cubble can handle slightly more than 300 x 300 (longitude x longitude) grid points for 3 variables in one year.
-The spatial grid can be reduced to trade for longer time periods and more variables.
+The memory limit with NetCDF data in cubble depends on the number of longitude grid points x the number of latitude grid points x the number of time grid points x the number of variables.
+Cubble can handle slightly more than 300 x 300 (longitude x longitude) grid points for daily 3 variables in one year.
 A 300 by 300 spatial grid can be a bounding box of [100, -80, 180, 0] at 0.25 degree resolution or a global bounding box [-180, -90, 180, -90] at 1 degree resolution.
-Subsetting longitude and latitude grids is available through \code{long_range} and \code{lat_range} if the NetCDF file has finer resolution than needed.
+The spatial grid can also be subsetted for to trade for longer time periods and more variables through \code{long_range} and \code{lat_range} if the NetCDF file has finer resolution than needed:
 

### 2022-04-23 arxiv/cubble-arxiv.tex
section 3 done

@@
 
-It may naturally raise the question that why bother a new data structure since the nested form can already be created by \code{dplyr::nest_by()}?
-While \code{nest_by()} and \code{unnest()} can mimic the nested form in cubble, cubble can be seen as an attempt to arrange variables from different observational units into a single object, specifically for spatio-temporal data.
-While \texttt{nest\_by()} solves this problem on the spatial side, it doesn't solve the temporal side.
-For temporal data, a combined representation with spatial variables is useful for spatio-temporal operations but its duplication can be a burden to operate on.
-Also, spatial variables are dropped when summarised and you will need to either specified all of them in the \code{group_by()} before \code{summarise}, or add a joining after \code{summarise()}.
+Some readers may raise the question why a new data structure rather than using \code{dplyr::nest_by()} on the combined data to directly create a list-column.
+While \code{nest_by()} and its pair \code{unnest()} can mimic the nested form in cubble, cubble can be seen as an attempt to arrange variables from different observational units into a single object, specifically for spatio-temporal data.
+While \texttt{nest\_by()} can recreate the nested form, it can't create a clean slate for temporal variables as the long form.
+For temporal variables, carrying the duplicated spatial variables in the combined form is not convenient to work with.
+This is because these spatial variables will lost after \code{summarise()} and one will need to subset the spatial variables again from the origin data and make a left join, which tends to create unnecessary intermediate products in the workflow.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# operation on the long form w/o cubble}
+\CommentTok{\# cubble workflow}
+\NormalTok{dt }\SpecialCharTok{|\textgreater{}} 
+  \FunctionTok{as\_cubble}\NormalTok{(...) }\SpecialCharTok{|\textgreater{}} 
+  \FunctionTok{face\_temporal}\NormalTok{() }\SpecialCharTok{|\textgreater{}} 
+  \FunctionTok{group\_by}\NormalTok{() }\SpecialCharTok{|\textgreater{}} 
+  \FunctionTok{summarise}\NormalTok{() }\SpecialCharTok{|\textgreater{}} 
+  \FunctionTok{unfold}\NormalTok{(...)}
+
+\CommentTok{\# tidyverse workflow}
+\NormalTok{temporal }\OtherTok{\textless{}{-}}\NormalTok{ dt }\SpecialCharTok{|\textgreater{}} 
+  \FunctionTok{group\_by}\NormalTok{(...) }\SpecialCharTok{|\textgreater{}} 
+  \FunctionTok{summarise}\NormalTok{(...)}
+
+\NormalTok{spatial }\OtherTok{\textless{}{-}}\NormalTok{ dt }\SpecialCharTok{|\textgreater{}} \FunctionTok{select}\NormalTok{(....)}
+
+\NormalTok{spatial }\SpecialCharTok{|\textgreater{}} \FunctionTok{left\_join}\NormalTok{(temporal)}
+\end{Highlighting}
+\end{Shaded}
 
@@
 As a spatial data object, \code{sf} creates a simple feature geometry list-column (\code{sfc}) in the data frame to provide spatial operations on various geometry types (\code{POINT}, \code{LINESTRING}, \code{POLYGON}, \code{MULTIPOLYGON}, etc).
-These spatial operations are also valuable for spatio-temporal data analysis, but an \code{sf} object \emph{usually} does not contain temporal variables.
-This means \code{sf} cannot be directly cast into a \code{cubble}, however, \code{cubble} does support \code{sfc} columns in the nested form and spatial operations applied to the \code{sfc} column in \code{sf} can still be applied to the \code{sfc} column in a cubble.
-The following example shows how to create an \code{sfc} column of \code{POINT} type from latitude and longitude in cubble.
-Then \code{sf::st_within} is used to add the state \code{MULTIPOLYGON} of each weather station before a coordinate transformation is made.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\FunctionTok{library}\NormalTok{(sf)}
-\CommentTok{\# create a cubble}
-\NormalTok{cb }\OtherTok{\textless{}{-}}\NormalTok{ climate\_flat }\SpecialCharTok{|\textgreater{}}
-\NormalTok{  cubble}\SpecialCharTok{::}\FunctionTok{as\_cubble}\NormalTok{(}\AttributeTok{key =}\NormalTok{ id, }\AttributeTok{index =}\NormalTok{ date, }\AttributeTok{coords =} \FunctionTok{c}\NormalTok{(long, lat))}
-
-\NormalTok{aus }\OtherTok{\textless{}{-}}\NormalTok{ ozmaps}\SpecialCharTok{::}\NormalTok{abs\_ste}
-
-\NormalTok{dt }\OtherTok{\textless{}{-}}\NormalTok{ cb }\SpecialCharTok{|\textgreater{}}
-  \FunctionTok{mutate}\NormalTok{(}
-    \CommentTok{\# create \textasciigrave{}sfc\textasciigrave{} column based on long and lat}
-    \AttributeTok{ll =} \FunctionTok{st\_sfc}\NormalTok{(}
-\NormalTok{      purrr}\SpecialCharTok{::}\FunctionTok{map2}\NormalTok{(long, lat, }\SpecialCharTok{\textasciitilde{}}\FunctionTok{st\_point}\NormalTok{(}\FunctionTok{c}\NormalTok{(.x, .y))),}
-      \AttributeTok{crs =} \FunctionTok{st\_crs}\NormalTok{(aus)),}
-
-    \CommentTok{\# append state multi{-}polygon based on the \textasciigrave{}sfc\textasciigrave{} created}
-    \AttributeTok{state =}\NormalTok{ aus}\SpecialCharTok{$}\NormalTok{geometry[}\FunctionTok{st\_within}\NormalTok{(ll, aus, }\AttributeTok{sparse =} \ConstantTok{FALSE}\NormalTok{)],}
-
-    \CommentTok{\# adopt a different projection: lambert conformal conic (EPSG:3112)}
-    \AttributeTok{state =} \FunctionTok{st\_transform}\NormalTok{(state, }\AttributeTok{crs =} \StringTok{"EPSG:3112"}\NormalTok{)}
-\NormalTok{    )}
-
-\NormalTok{dt}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## # cubble:   id [5]: nested form
-## # bbox:     [115.97, -32.94, 133.55, -12.42]
-## # temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
-##   id            lat  long  elev name   wmo_id ts                        ll
-##   <chr>       <dbl> <dbl> <dbl> <chr>   <dbl> <list>           <POINT [°]>
-## 1 ASN00009021 -31.9  116.  15.4 perth~  94610 <tibble> (115.9764 -31.9275)
-## 2 ASN00010311 -31.9  117. 179   york    94623 <tibble>  (116.765 -31.8997)
-## 3 ASN00010614 -32.9  117. 338   narro~  94627 <tibble> (117.1797 -32.9342)
-## 4 ASN00014015 -12.4  131.  30.4 darwi~  94120 <tibble> (130.8925 -12.4239)
-## 5 ASN00015131 -17.6  134. 220   ellio~  94236 <tibble> (133.5407 -17.5521)
-## # ... with 1 more variable: state <MULTIPOLYGON [m]>
-\end{verbatim}
-
-An \pkg{s2} \code{lnglat} vector can similarly be created as an \code{sfc} in cubble before using any \code{s2}-prefixed function:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\FunctionTok{library}\NormalTok{(s2)}
-\CommentTok{\# Western Australia map}
-\NormalTok{wa }\OtherTok{\textless{}{-}}\NormalTok{ ozmaps}\SpecialCharTok{::}\NormalTok{abs\_ste }\SpecialCharTok{|\textgreater{}} \FunctionTok{filter}\NormalTok{(NAME }\SpecialCharTok{==} \StringTok{"Western Australia"}\NormalTok{)}
-
-\CommentTok{\# mutate a \textasciigrave{}s2\_lnglat\textasciigrave{} vector on \textasciigrave{}cb\textasciigrave{} created in the last chunk}
-\NormalTok{cb }\SpecialCharTok{|\textgreater{}}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{ll =} \FunctionTok{s2\_lnglat}\NormalTok{(long, lat)) }\SpecialCharTok{|\textgreater{}}
-  \FunctionTok{filter}\NormalTok{(}\FunctionTok{s2\_within}\NormalTok{(ll, wa))}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## # cubble:   id [3]: nested form
-## # bbox:     [115.97, -32.94, 117.18, -31.89]
-## # temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
-##   id            lat  long  elev name          wmo_id ts                 ll      
-##   <chr>       <dbl> <dbl> <dbl> <chr>          <dbl> <list>             <s2_lng>
-## 1 ASN00009021 -31.9  116.  15.4 perth airport  94610 <tibble [366 x 4]> (115.97~
-## 2 ASN00010311 -31.9  117. 179   york           94623 <tibble [366 x 4]> (116.76~
-## 3 ASN00010614 -32.9  117. 338   narrogin       94627 <tibble [366 x 4]> (117.17~
-\end{verbatim}
+An \code{sf} object can be supplied as the spatial table, as part of the raw data, along with a temporal table, to create a cubble.
+Once formed in a cubble, methods for the \code{sfc} class can be applied on the nested form.
+An example of this is shown in \ref{covid}, which also handles the case when the site in the two tables don't match exactly.
+A spatial data object with \code{s2} vector can also be the input of the spatial table.
 
@@
 
-NetCDF data has two main components: \emph{dimension} for defining the spatio-temporal grid (longitude, latitude, and time) and \emph{variable} that populates the defined grid.
+NetCDF data is another format that is commonly used for storing spatio-temporal data.
+It has two main components: \emph{dimension} for defining the spatio-temporal grid (longitude, latitude, and time) and \emph{variable} that populates the defined grid.
 Attributes are usually associated with dimension and variable in the NetCDF format data and a \href{http://cfconventions.org/}{metadata convention for climate and forecast} has been designed to standardise the format of the attributes.
@@
 
-The memory limit with NetCDF data in cubble depends on the number of longitude grid points times the number of latitude grid points times the number of time grid points times the number of variables.
-Cubble can handle slightly more than 300 x 300 (longitude x longitude) grid points for 3 variables in one year.
-The spatial grid can be reduced to trade for longer time periods and more variables.
+The memory limit with NetCDF data in cubble depends on the number of longitude grid points x the number of latitude grid points x the number of time grid points x the number of variables.
+Cubble can handle slightly more than 300 x 300 (longitude x longitude) grid points for daily 3 variables in one year.
 A 300 by 300 spatial grid can be a bounding box of {[}100, -80, 180, 0{]} at 0.25 degree resolution or a global bounding box {[}-180, -90, 180, -90{]} at 1 degree resolution.
-Subsetting longitude and latitude grids is available through \code{long_range} and \code{lat_range} if the NetCDF file has finer resolution than needed.
+The spatial grid can also be subsetted for to trade for longer time periods and more variables through \code{long_range} and \code{lat_range} if the NetCDF file has finer resolution than needed:
 

### 2022-04-23 2-0-cube.Rmd
spell check

@@
 
-Spatio-temporal data can be conceptualised using a cubical data model with three axis.
+Spatio-temporal data can be conceptualised using a cubical data model with three axes.
 This abstraction can be useful to generalise operations or even visualisation: @lu_multidimensional_2018 shows an example of how array operations (select, scale, reduce, rearrange, and compute) can be mapped onto the cube; @bach_review_2014 reviews the temporal data visualisation based on space-time cube operations.
-Notice that the term space-time cube in their article "does not need to involve spatial data", but referring to "an abstract 2D substrate that is used to visualize data at a specific time".
-Despite its main focus is on temporal data, the mindset of abstracting out data representation to construct visualisation is applicable to our spatio-temporal data manipulation and visualisation.
+Notice that the term space-time cube in their article "does not need to involve spatial data", but refers to "an abstract 2D substrate that is used to visualize data at a specific time".
+Despite its main focus being on temporal data, the mindset of abstracting out data representation to construct visualisation is applicable to our spatio-temporal data manipulation and visualisation.
 
 While a cubical conceptual data model is neat to conceptually view the spatio-temporal data, a 3D data representation, or an array,  may not be convenient for data wrangling.
-There are two reasons for this: 1) While arrays are efficient for computation on numerical values, spatio-temporal data contains more than just numerical variables.
+There are two reasons for this: 1) While arrays are efficient for the computation on numerical values, spatio-temporal data contains more than just numerical variables.
 Wrangling on character strings and specific datetime classes are common tasks for spatio-temporal data but are not easy with arrays.
@@
 
-There has been many cubes defined using time, latitude, and longitude, which can be seen as taking snapshots of the space and stack on time.
-Our spatio-temporal cube is defined slightly different with three axes: time, site, and variable  as Figure \ref{fig:cube-def}.
-The time axis is the same across the two formation, site encompasses latitude and longitude into a single column identifier, while variables are stacked on this defined space-time canvas.
+There have been many cubes defined using time, latitude, and longitude, which can be seen as taking snapshots of the space and stacking on time.
+Our spatio-temporal cube is defined slightly differently with three axes: time, site, and variable as Figure \ref{fig:cube-def}.
+The time axis is the same across the two formations, the site encompasses latitude and longitude into a single column identifier, while variables are stacked on this defined space-time canvas.
 This notion is adopted to avoid the use of hypercube when describing multivariate spatio-temporal data.

### 2022-04-23 2-the-cubble-pkg.Rmd
spell check

@@
 Each core function is introduced with a short example accompanied and the data `climate_flat` is used throughout this section.
-`climate_flat` contains five weather stations in Australia with spatial information of each station: station id, latitude, longitude, elevation, station name and World Meteorology Organisation ID and also daily temporal information of date, maximum and minimum temperature and precipitation for 2020.
+`climate_flat` contains five weather stations in Australia with spatial information of each station: station id, latitude, longitude, elevation, station name, and World Meteorology Organisation ID and also daily temporal information of date, maximum and minimum temperature and precipitation for 2020.
 Below prints out the first five rows of `climate_flat`:
@@
 All the temporal variables are nested in a list column, hence it is also called the nested cubble.
-The rowwise structure makes it simpler to calculate on the list using the \code{mutate()} syntax, which is simplier than the {\code{map()} when working with list column.
-For example, calculating the number of raining day can be done by:
+The rowwise structure makes it simpler to calculate on the list using the \code{mutate()} syntax, which is simpler than the {\code{map()} when working with list column.
+For example, calculating the number of raining days can be done by:
 
@@
 \pkg{cubble} also provides a long form, which expands the \code{ts} column and temporarily "hide" the spatial variables.
-The function \code{face_temporal()} is used to switch a nested cubble into a long one and the first row in Figure \ref{fig:face} illustrate this operation where the focus of the cube now changes from the site-variable face to the time-variable face.
+The function \code{face_temporal()} is used to switch a nested cubble into a long one and the first row in Figure \ref{fig:face} illustrates this operation where the focus of the cube now changes from the site-variable face to the time-variable face.
 This code switches the cubble just created into its long form:
@@
 The first line in the header now shows the cubble is in its long form and the third line has been changed to show the name and type of spatial variables: \code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]}.
-Unlike the nested form buildin form a \code{rowwise_df}, the long cubble is built form class \code{groupped_df} where all the observations from the same sites forms a group.
+Unlike the nested form, the long cubble is built form class \code{groupped_df} where all the observations from the same sites form a group.
 
@@
 
-Wrangling spatio-temproal data can be an iterative process on the spatial and temporal dimension.
+Wrangling spatio-temporal data can be an iterative process in the spatial and temporal dimensions.
 Switching the focus back to the site-variable face can be accomplished by the function \code{face_spatial()}, which is the inverse of \code{face_temporal()}.
@@
 
-Sometimes, analysts may need to apply some variable transformation that involves both the spatial and temporal variable.
+Sometimes, analysts may need to apply some variable transformation that involves both the spatial and temporal variables.
 An example of this is the transformation of temporal variables into the spatial dimension in glyph maps, which will be elaborated in section \ref{st_transformation}.
-This type of operations can be seen as flattening, or *unfolding*, the cube into a 2D data frame.
+This type of operation can be seen as flattening, or *unfolding*, the cube into a 2D data frame.
 Here the function \code{unfold()} moves the spatial variables \code{long} and \code{lat} into the long cubble:
@@
 For temporal variables, carrying the duplicated spatial variables in the combined form is not convenient to work with.
-This is because these spatial variables will lost after \code{summarise()} and one will need to subset the spatial variables again from the origin data and make a left join, which tends to create unnecessary intermediate products in the workflow.
+This is because these spatial variables will lose after \code{summarise()} and one will need to subset the spatial variables again from the original data and make a left join, which tends to create unnecessary intermediate products in the workflow.
 
@@
 An \code{sf} object can be supplied as the spatial table, as part of the raw data, along with a temporal table, to create a cubble.
-Once formed in a cubble, methods for the \code{sfc} class can be applied on the nested form.
-An example of this is shown in \@ref(covid), which also handles the case when the site in the two tables don't match exactly.
+Once formed in a cubble, methods for the \code{sfc} class can be applied in the nested form.
+An example of this is shown in \@ref(covid), which also handles the case when the sites in the two tables don't match exactly.
 A spatial data object with \code{s2} vector can also be the input of the spatial table.
@@
 It has two main components: *dimension* for defining the spatio-temporal grid (longitude, latitude, and time) and *variable* that populates the defined grid.
-Attributes are usually associated with dimension and variable in the NetCDF format data and a [metadata convention for climate and forecast](http://cfconventions.org/) has been designed to standardise the format of the attributes.
+Attributes are usually associated with dimensions and variables in the NetCDF format data and a [metadata convention for climate and forecast](http://cfconventions.org/) has been designed to standardise the format of the attributes.
 A few packages in R exists for manipulating NetCDF data and this includes a high-level R interface: \pkg{ncdf4} [@ncdf4], a low-level interface that calls C-interface: \pkg{RNetCDF} [@rnetcdf; @michna2013rnetcdf], and a tidyverse implementation: \pkg{tidync} [@tidync].
@@
 A 300 by 300 spatial grid can be a bounding box of [100, -80, 180, 0] at 0.25 degree resolution or a global bounding box [-180, -90, 180, -90] at 1 degree resolution.
-The spatial grid can also be subsetted for to trade for longer time periods and more variables through \code{long_range} and \code{lat_range} if the NetCDF file has finer resolution than needed:
+The spatial grid can also be subsetted to trade for longer time periods and more variables through \code{long_range} and \code{lat_range} if the NetCDF file has finer resolution than needed:
 

### 2022-04-23 arxiv/cubble-arxiv.tex
spell check

@@
 
-Spatio-temporal data can be conceptualised using a cubical data model with three axis.
+Spatio-temporal data can be conceptualised using a cubical data model with three axes.
 This abstraction can be useful to generalise operations or even visualisation: Lu, Appel, and Pebesma (\protect\hyperlink{ref-lu_multidimensional_2018}{2018}) shows an example of how array operations (select, scale, reduce, rearrange, and compute) can be mapped onto the cube; Bach et al.
 (\protect\hyperlink{ref-bach_review_2014}{2014}) reviews the temporal data visualisation based on space-time cube operations.
-Notice that the term space-time cube in their article ``does not need to involve spatial data'', but referring to ``an abstract 2D substrate that is used to visualize data at a specific time''.
-Despite its main focus is on temporal data, the mindset of abstracting out data representation to construct visualisation is applicable to our spatio-temporal data manipulation and visualisation.
+Notice that the term space-time cube in their article ``does not need to involve spatial data'', but refers to ``an abstract 2D substrate that is used to visualize data at a specific time''.
+Despite its main focus being on temporal data, the mindset of abstracting out data representation to construct visualisation is applicable to our spatio-temporal data manipulation and visualisation.
 
 While a cubical conceptual data model is neat to conceptually view the spatio-temporal data, a 3D data representation, or an array, may not be convenient for data wrangling.
-There are two reasons for this: 1) While arrays are efficient for computation on numerical values, spatio-temporal data contains more than just numerical variables.
+There are two reasons for this: 1) While arrays are efficient for the computation on numerical values, spatio-temporal data contains more than just numerical variables.
 Wrangling on character strings and specific datetime classes are common tasks for spatio-temporal data but are not easy with arrays.
@@
 
-There has been many cubes defined using time, latitude, and longitude, which can be seen as taking snapshots of the space and stack on time.
-Our spatio-temporal cube is defined slightly different with three axes: time, site, and variable as Figure \ref{fig:cube-def}.
-The time axis is the same across the two formation, site encompasses latitude and longitude into a single column identifier, while variables are stacked on this defined space-time canvas.
+There have been many cubes defined using time, latitude, and longitude, which can be seen as taking snapshots of the space and stacking on time.
+Our spatio-temporal cube is defined slightly differently with three axes: time, site, and variable as Figure \ref{fig:cube-def}.
+The time axis is the same across the two formations, the site encompasses latitude and longitude into a single column identifier, while variables are stacked on this defined space-time canvas.
 This notion is adopted to avoid the use of hypercube when describing multivariate spatio-temporal data.
@@
 Each core function is introduced with a short example accompanied and the data \texttt{climate\_flat} is used throughout this section.
-\texttt{climate\_flat} contains five weather stations in Australia with spatial information of each station: station id, latitude, longitude, elevation, station name and World Meteorology Organisation ID and also daily temporal information of date, maximum and minimum temperature and precipitation for 2020.
+\texttt{climate\_flat} contains five weather stations in Australia with spatial information of each station: station id, latitude, longitude, elevation, station name, and World Meteorology Organisation ID and also daily temporal information of date, maximum and minimum temperature and precipitation for 2020.
 Below prints out the first five rows of \texttt{climate\_flat}:
@@
 All the temporal variables are nested in a list column, hence it is also called the nested cubble.
-The rowwise structure makes it simpler to calculate on the list using the \code{mutate()} syntax, which is simplier than the \{\code{map()} when working with list column.
-For example, calculating the number of raining day can be done by:
+The rowwise structure makes it simpler to calculate on the list using the \code{mutate()} syntax, which is simpler than the \{\code{map()} when working with list column.
+For example, calculating the number of raining days can be done by:
 
@@
 \pkg{cubble} also provides a long form, which expands the \code{ts} column and temporarily ``hide'' the spatial variables.
-The function \code{face_temporal()} is used to switch a nested cubble into a long one and the first row in Figure \ref{fig:face} illustrate this operation where the focus of the cube now changes from the site-variable face to the time-variable face.
+The function \code{face_temporal()} is used to switch a nested cubble into a long one and the first row in Figure \ref{fig:face} illustrates this operation where the focus of the cube now changes from the site-variable face to the time-variable face.
 This code switches the cubble just created into its long form:
@@
 
-Sometimes, analysts may need to apply some variable transformation that involves both the spatial and temporal variable.
+Sometimes, analysts may need to apply some variable transformation that involves both the spatial and temporal variables.
 An example of this is the transformation of temporal variables into the spatial dimension in glyph maps, which will be elaborated in section \ref{st_transformation}.
-This type of operations can be seen as flattening, or \emph{unfolding}, the cube into a 2D data frame.
+This type of operation can be seen as flattening, or \emph{unfolding}, the cube into a 2D data frame.
 Here the function \code{unfold()} moves the spatial variables \code{long} and \code{lat} into the long cubble:
@@
 An \code{sf} object can be supplied as the spatial table, as part of the raw data, along with a temporal table, to create a cubble.
-Once formed in a cubble, methods for the \code{sfc} class can be applied on the nested form.
-An example of this is shown in \ref{covid}, which also handles the case when the site in the two tables don't match exactly.
+Once formed in a cubble, methods for the \code{sfc} class can be applied in the nested form.
+An example of this is shown in \ref{covid}, which also handles the case when the sites in the two tables don't match exactly.
 A spatial data object with \code{s2} vector can also be the input of the spatial table.
@@
 It has two main components: \emph{dimension} for defining the spatio-temporal grid (longitude, latitude, and time) and \emph{variable} that populates the defined grid.
-Attributes are usually associated with dimension and variable in the NetCDF format data and a \href{http://cfconventions.org/}{metadata convention for climate and forecast} has been designed to standardise the format of the attributes.
+Attributes are usually associated with dimensions and variables in the NetCDF format data and a \href{http://cfconventions.org/}{metadata convention for climate and forecast} has been designed to standardise the format of the attributes.
 A few packages in R exists for manipulating NetCDF data and this includes a high-level R interface: \pkg{ncdf4} (\protect\hyperlink{ref-ncdf4}{Pierce 2019}), a low-level interface that calls C-interface: \pkg{RNetCDF} (\protect\hyperlink{ref-rnetcdf}{Michna and Woods 2021}, \protect\hyperlink{ref-michna2013rnetcdf}{2013}), and a tidyverse implementation: \pkg{tidync} (\protect\hyperlink{ref-tidync}{Sumner 2020}).
@@
 A 300 by 300 spatial grid can be a bounding box of {[}100, -80, 180, 0{]} at 0.25 degree resolution or a global bounding box {[}-180, -90, 180, -90{]} at 1 degree resolution.
-The spatial grid can also be subsetted for to trade for longer time periods and more variables through \code{long_range} and \code{lat_range} if the NetCDF file has finer resolution than needed:
+The spatial grid can also be subsetted to trade for longer time periods and more variables through \code{long_range} and \code{lat_range} if the NetCDF file has finer resolution than needed:
 
@@
 
-Spatial locations can have grouping structure either inherent to the data or formed by clustering.
+Spatial locations can have grouping structure either inherent in the data or formed by clustering.
 Rather than analysing variables in the site level, summarised variables in the cluster level can give a crisper picture of local areas.

### 2022-04-23 3-advanced.Rmd
a new skeleton of intro incorporating various past writing
polish wording

@@
 
-# Other features and considerations
+# Other features and considerations {#others}
 
@@
 One task that may interest analysts in spatio-temporal data is to find how similar the time series from nearby sites are.
-This can be seen as a matching problem [@stuart2010matching; @mcintosh2018using] that pairs up similar similar time series in nearby locations or a data fusing exercise that merges data collected from different sources [@cocchi2019data].
+This can be seen as a matching problem [@stuart2010matching; @mcintosh2018using] that pairs up similar time series in nearby locations or a data fusing exercise that merges data collected from different sources [@cocchi2019data].
 \code{match_sites()} in \pkg{cubble} provides a simple algorithm for this task.
@@
 In \proglang{R}, \pkg{GGally} implements glyph maps through the \code{glyphs()} function.
-The function constructs a data frame with calculated position (\code{gx}.
-\code{gy}.
+The function constructs a data frame with calculated position (\code{gx}, \code{gy}.
 \code{gid}) of each point on the time series using linear algebra (Equation 1 and 2 in @Wickham2012-yr).

### 2022-04-23 4-1-basic.Rmd
work on section 3 the cubble package
polish wording

@@
 
-## Victoria covid data
+## Victoria covid data {#covid}
 
-This example illustrate casting different sources into a cubble object.
 Victoria government of Australia provides daily COVID case number by date, source, and local government area (LGA).
-This data can be useful for visualising COVID spread when combined with LGA map information, which is available from Australian Bureau of Statistics.
-Variable \code{covid} and \code{lga} show the first 5 rows of these two data:
+This data can be used to visualise COVID spread when combined with map information on LGA, available from the Australian Bureau of Statistics.
+The first five rows of both data are printed below:
 
@@
 
-To cast these two data into a cubble object, supply them as a list with name \code{spatial} and \code{temporal} and other arguments (\code{key}, \code{index}, and \code{coords}).
-\code{as_cubble()} will automatically check the match of spatial and temporal information in the two datasets and warnings will emit if a location has missing spatial or temporal information.
+Cubble can be created with separate spatial and temporal table in a list, with other arguments (\code{key}, \code{index}, and \code{coords}) introduced in section \@ref(create).
+\code{as_cubble()} will automatically check the match of site in both tables emit a warning if a location has missing spatial or temporal information.
 
@@
 
-\pkg{cubble} will make an attempt to pair the unmatched sites as well as showing unpaired locations. These can be useful information to help clean up the data.
+\pkg{cubble} will attempt to pair the unmatched sites as well as those that cannot be paired. These can be helpful to clean up the data before creating the cubble again:
 
@@
 covid <- covid %>% filter(!lga %in% pair$others$temporal)
-```
 
-After cleaning, you can cast the data into cubble with the same syntax without warnings:
-
-```{r echo = TRUE}
 cb <- as_cubble(data = list(spatial = lga, temporal = covid),
@@
 
-## Australia historical maximum temperature
+## Australian historical maximum temperature
 
@@
 
-Stations in the two states can be subsetted by the station number: Australia GHCN station number starts with "ASN00" and followed by the [Bureau of Meteorology (BOM) station number](http://www.bom.gov.au/climate/cdo/about/site-num.shtml), where the 2nd and 3rd digit (7th and 8th in the GHCN number) define the state of the station.
-New South Wales stations start from 46 to 75 and Victoria stations follow from 76 to 90.
+Stations in the two states can be subsetted by the station number: Australia GHCN station number starts with "ASN00" and is followed by the [Bureau of Meteorology (BOM) station number](http://www.bom.gov.au/climate/cdo/about/site-num.shtml).
+The second and third digits of BOM station number (7th and 8th in the GHCN number) define the state of the station, where 46 - 75 are New South Wales stations and 76 - 90 are Victoria stations.
 Filtering Victoria and New South Wales stations is an operation in the spatial dimension and hence uses the nested form:
@@
 
-A few stations don't have records during 1971 - 1975 and further investigation shows that while the first and last year of each series is recorded, the missing years in this period is not reported.
-These stations are filtered out by examining whether the summarised time series has a total of 24 months.
+A few stations do not have records during 1971 - 1975 and further investigation shows that while the first and last year of each series is recorded, the missing years in this period are not reported.
+These stations are filtered out by examining whether the summarised time series has 24 months.
 The long cubble needs to be switched to the nested form for this spatial operation using \code{face_spatial()}:

### 2022-04-23 4-2-agg.Rmd
polish wording

@@
 
-In the previous example, there has already been some overlapping of the glyphs for a few stations near (151E, 34S) and (152E, 33S) and this will be a problem when mapping more stations in the national level.
-Aggregation can be helpful to group series into clusters before visualising the cluster with glyph map.
-This example shows how to organise data at both level with \code{switch_key()}.
+In the previous example, there has already been some overlapping of the glyphs for a few stations near (151E, 34S) and (152E, 33S), which is a problem when mapping more stations at the national level.
+Aggregation can be helpful in grouping series into clusters before visualising the cluster with a glyph map.
+This example shows how to organise data at both levels with \code{switch_key()}.
 
@@
 A simple kmean algorithm based on the distance matrix is used to create 20 clusters.
-This creates \code{station_nested} as a station level nested cubble with a cluster column indicating the group each station belongs to.
-More complex algorithms can be used for other problem, as long as there is a mapping from each station to a cluster.
+The dataset \code{station_nested} is a nested cubble with a cluster column indicating the group each station belongs to.
+More complex algorithms can be used for other problems, as long as there is a mapping from each station to a cluster.
 
@@
 Long form cubble at both levels can be accessed through stretching the nested form and with access to both station and cluster level cubbles, various plots can be made to understand the cluster.
-Figure \ref{fig:basic-agg} shows two example plots that can be made with this data: subplot A is a glyph map made with the cluster level cubble in the long form and subplot B inspects the station membership of each cluster using the station level cubble in the nested form.
+Figure \ref{fig:basic-agg} shows two example plots that can be made with this data.
+Subplot A is a glyph map made with the cluster level cubble in the long form and subplot B inspects the station membership of each cluster using the station level cubble in the nested form.
 

### 2022-04-23 4-3-water-level.Rmd
polish wording

@@
 
-Bureau of Meteorology collects [water data](http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full) from river gauges and this includes variables: electrical conductivity, turbidity, water course discharge, water course level, and water temperature.
-In particular, water level will interactive with precipitation from the climate data since rainfall will raise the water level in the river.
-Figure \ref{fig:matching-map} gives the location of available weather station and water gauges in Victoria.
+Bureau of Meteorology collects [water data](http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full) from river gauges and this includes variables: electrical conductivity, turbidity, watercourse discharge, watercourse level, and water temperature.
+In particular, water level will interact with precipitation from the climate data since rainfall will raise the water level in the river.
+Figure \ref{fig:matching-map} gives the location of available weather stations and water gauges in Victoria.
 
@@
 The two datasets need to be specified as the first two arguments and the variable to match can be specified in \code{temporal_by} using the \code{by} syntax in \code{join}.
-\code{temporal_independent} controls the variable used to construct the interval and the goal here is to see if precipitation will be reflected by the water level in the river.
+\code{temporal_independent} controls the variable used to construct the interval and the goal here is to see if the water level in the river will reflect precipitation.
 This puts precipitation \code{prcp}, as the independent.
-Given there is one year worth of data, the number of peak (\code{temporal_n_highest}) to consider is slightly raised from a default 20 to 30 and \code{temporal_min_match} is raised accordingly To return all the pairs of the match, \code{temporal_min_match} can be set to 0.
+Given there is one year's worth of data, the number of peaks (\code{temporal_n_highest}) to consider is slightly raised from a default of 20 to 30 and \code{temporal_min_match} is raised accordingly To return all the pairs of the match, \code{temporal_min_match} can be set to 0.
 
@@
 Figure \ref{fig:matching} plots the matched pairs on the map or to view the matched series.
-There are four pairs of matches, which all locates in the middle Victoria and the concurrent increase of precipitation and water level can be observed.
+There are four pairs of matches, all located in the middle of Victoria and the concurrent increase in precipitation and water level can be observed.
 
@@
 Precipitation and water level have been standardised between 0 and 1 to be displayed on the same scale.
-The increases in the precipitation is reflected by the water level."}
+The water level reflects the increase in precipitation."}
 knitr::include_graphics(here::here("figures/matching.png"))

### 2022-04-23 4-5-interactive.Rmd
polish wording

@@
 ---
-## Interative graphic
 
-With spatio-temporal data, users may wish to make plots to learn the spatial distribution of a variable, or to find patterns, such as trend or seasonality, in the time series.
-Combining this two types of plot with interactivity let users to link between points on the map and the corresponding time series to explore the spatial and temporal dimension of the data simultaneously.
-Below is an example that describes the process of building an interactive graphic with \pkg{cubble} and \pkg{crosstalk} The example explores the variation of monthly temperature range with \code{weatherdata::climate_full} data.
+## Interactive graphic
+
+With spatio-temporal data, users may wish to make plots to learn the spatial distribution of a variable or to find patterns, such as trend or seasonality, in the time series.
+Combining these two types of plots with interactivity lets users link between points on the map and the corresponding time series to explore the spatial and temporal dimensions of the data simultaneously.
+Below is an example that describes the process of building an interactive graphic with \pkg{cubble} and \pkg{crosstalk}.
+The example explores the variation of monthly temperature range with \code{weatherdata::climate_full} data.
 
@@
 A \code{SharedData} object is constructed for each form of the cubble and the same \code{group} argument ensures the cross-linking of the two forms via the common \code{id} column.
-The spatial map and time series plot are then made with each \code{SharedData} objects separately.
-In this example, stations on the Australia map, made from the nested form, are coloured by the calculated variance and a ribbon band is constructed using the long form cubble to show the maximum and minimum temperature of each station across month.
-With a different dataset, users are free to calculate any per station measure in the nested form or to make any time-wise summary of the data in the long form to customise the spatial or temporal view.
+The spatial map and time series plot are then made with each \code{SharedData} object separately.
+In this example, stations on the Australia map, made from the nested form, are coloured by the calculated variance.
+A ribbon band is constructed using the long form cubble to show each station's  maximum and minimum temperature across month.
+With a different dataset, users can calculate other per station measure in the nested form or make other time-wise summary of the data in the long form to customise the spatial or temporal view.
 The cross-linking between the two plots is always safeguarded by the shared \code{id} column embedded in the cubble structure.
-Below is the pseudo code that outlines the process to construct an interactive graphic described above:
+Below is the pseudo-code that outlines the process to construct an interactive graphic described above:
 
@@
 In Figure \ref{fig:interactive-linking}, the first row shows the initial view of the interactive graphic.
-On the map, most regions in Australia have low variance of temperature range while the north-west coastline, bottom of South Australia, and Victoria stands out with larger monthly changes.
+Most regions in Australia have low variance of temperature range On the map, while the north-west coastline, bottom of South Australia, and Victoria stand out with larger monthly changes.
 In the second row, Mount Elizabeth is selected on the map given its high variance colour on the initial map and this links to the ribbon on the right.
-The third row the lowest temperature in August and this corresponds to Thredbo AWS in the Victoria and New South Wales border.
+The third row selects the lowest temperature in August and this corresponds to Thredbo AWS on the Victoria and New South Wales border.
 Another station in the Tasmania island is selected on the map to cross compare with Thredbo AWS.
@@
 
-This plot can also be made using \pkg{cubble} and \pkg{leaflet} where the temperature range can be displayed as a small subplot upon clicking on the map.
-This would require first creating the popup plots from the long form cubble as a vector and then add these plots to a leaflet map created from the nested cubble, with \code{leafpop::addPopupGraphs()}:
+This plot can also be made using \pkg{cubble} and \pkg{leaflet} where the temperature range is displayed as a small subplot upon clicking on the map.
+The procedure involves first creating the popup plots from the long form cubble as a vector and then adding these plots to a leaflet map created from the nested cubble, with \code{leafpop::addPopupGraphs()}:
 
@@
 
-```{r interactive-popup, echo = FALSE, out.width="45%", out.height="25%", fig.retina = 2, dpi = 300, fig.cap = "Same as Figure 11 with thetemperature variation shown as a popup in the leaflet map.", fig.show='hold'}
+```{r interactive-popup, fig.align='center', echo = FALSE, out.width="45%", out.height="25%", fig.retina = 2, dpi = 300, fig.cap = "Same as Figure 11 with thetemperature variation shown as a popup in the leaflet map.", fig.show='hold'}
 knitr::include_graphics(here::here("figures/popup-mid.png")) 

### 2022-04-23 arxiv/cubble-arxiv.tex
polish wording

@@
 This can be seen as a matching problem (\protect\hyperlink{ref-stuart2010matching}{Stuart 2010}; \protect\hyperlink{ref-mcintosh2018using}{McIntosh et al.
-2018}) that pairs up similar similar time series in nearby locations or a data fusing exercise that merges data collected from different sources (\protect\hyperlink{ref-cocchi2019data}{Cocchi 2019}).
+2018}) that pairs up similar time series in nearby locations or a data fusing exercise that merges data collected from different sources (\protect\hyperlink{ref-cocchi2019data}{Cocchi 2019}).
 \code{match_sites()} in \pkg{cubble} provides a simple algorithm for this task.
@@
 In \proglang{R}, \pkg{GGally} implements glyph maps through the \code{glyphs()} function.
-The function constructs a data frame with calculated position (\code{gx}.
-\code{gy}.
+The function constructs a data frame with calculated position (\code{gx}, \code{gy}.
 \code{gid}) of each point on the time series using linear algebra (Equation 1 and 2 in Wickham et al.
@@
 
-This example illustrate casting different sources into a cubble object.
 Victoria government of Australia provides daily COVID case number by date, source, and local government area (LGA).
-This data can be useful for visualising COVID spread when combined with LGA map information, which is available from Australian Bureau of Statistics.
-Variable \code{covid} and \code{lga} show the first 5 rows of these two data:
+This data can be used to visualise COVID spread when combined with map information on LGA, available from the Australian Bureau of Statistics.
+The first five rows of both data are printed below:
 
@@
 
-To cast these two data into a cubble object, supply them as a list with name \code{spatial} and \code{temporal} and other arguments (\code{key}, \code{index}, and \code{coords}).
-\code{as_cubble()} will automatically check the match of spatial and temporal information in the two datasets and warnings will emit if a location has missing spatial or temporal information.
+Cubble can be created with separate spatial and temporal table in a list, with other arguments (\code{key}, \code{index}, and \code{coords}) introduced in section \ref{create}.
+\code{as_cubble()} will automatically check the match of site in both tables emit a warning if a location has missing spatial or temporal information.
 
@@
 \NormalTok{covid }\OtherTok{\textless{}{-}}\NormalTok{ covid }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(}\SpecialCharTok{!}\NormalTok{lga }\SpecialCharTok{\%in\%}\NormalTok{ pair}\SpecialCharTok{$}\NormalTok{others}\SpecialCharTok{$}\NormalTok{temporal)}
-\end{Highlighting}
-\end{Shaded}
-
-After cleaning, you can cast the data into cubble with the same syntax without warnings:
-
-\begin{Shaded}
-\begin{Highlighting}[]
+
 \NormalTok{cb }\OtherTok{\textless{}{-}} \FunctionTok{as\_cubble}\NormalTok{(}\AttributeTok{data =} \FunctionTok{list}\NormalTok{(}\AttributeTok{spatial =}\NormalTok{ lga, }\AttributeTok{temporal =}\NormalTok{ covid),}
@@
 
-\hypertarget{australia-historical-maximum-temperature}{%
-\subsection{Australia historical maximum temperature}\label{australia-historical-maximum-temperature}}
+\hypertarget{australian-historical-maximum-temperature}{%
+\subsection{Australian historical maximum temperature}\label{australian-historical-maximum-temperature}}
 
@@
 
-Stations in the two states can be subsetted by the station number: Australia GHCN station number starts with ``ASN00'' and followed by the \href{http://www.bom.gov.au/climate/cdo/about/site-num.shtml}{Bureau of Meteorology (BOM) station number}, where the 2nd and 3rd digit (7th and 8th in the GHCN number) define the state of the station.
-New South Wales stations start from 46 to 75 and Victoria stations follow from 76 to 90.
+Stations in the two states can be subsetted by the station number: Australia GHCN station number starts with ``ASN00'' and is followed by the \href{http://www.bom.gov.au/climate/cdo/about/site-num.shtml}{Bureau of Meteorology (BOM) station number}.
+The second and third digits of BOM station number (7th and 8th in the GHCN number) define the state of the station, where 46 - 75 are New South Wales stations and 76 - 90 are Victoria stations.
 Filtering Victoria and New South Wales stations is an operation in the spatial dimension and hence uses the nested form:
@@
 
-Bureau of Meteorology collects \href{http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full}{water data} from river gauges and this includes variables: electrical conductivity, turbidity, water course discharge, water course level, and water temperature.
-In particular, water level will interactive with precipitation from the climate data since rainfall will raise the water level in the river.
-Figure \ref{fig:matching-map} gives the location of available weather station and water gauges in Victoria.
+Bureau of Meteorology collects \href{http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full}{water data} from river gauges and this includes variables: electrical conductivity, turbidity, watercourse discharge, watercourse level, and water temperature.
+In particular, water level will interact with precipitation from the climate data since rainfall will raise the water level in the river.
+Figure \ref{fig:matching-map} gives the location of available weather stations and water gauges in Victoria.
 
@@
 Precipitation and water level have been standardised between 0 and 1 to be displayed on the same scale.
-The increases in the precipitation is reflected by the water level.}\label{fig:matching}
+The water level reflects the increase in precipitation.}\label{fig:matching}
 \end{figure}
@@
 (\protect\hyperlink{ref-hersbach2020era5}{2020}), Simmons et al.
-(\protect\hyperlink{ref-simmons2020global}{2020}) and Simmons et al.
+(\protect\hyperlink{ref-simmons2020global}{2020}), and Simmons et al.
 (\protect\hyperlink{ref-simmons2005ecmwf}{2005}) for more details.
@@
 
-\hypertarget{interative-graphic}{%
-\subsection{Interative graphic}\label{interative-graphic}}
-
-With spatio-temporal data, users may wish to make plots to learn the spatial distribution of a variable, or to find patterns, such as trend or seasonality, in the time series.
-Combining this two types of plot with interactivity let users to link between points on the map and the corresponding time series to explore the spatial and temporal dimension of the data simultaneously.
-Below is an example that describes the process of building an interactive graphic with \pkg{cubble} and \pkg{crosstalk} The example explores the variation of monthly temperature range with \code{weatherdata::climate_full} data.
+\hypertarget{interactive-graphic}{%
+\subsection{Interactive graphic}\label{interactive-graphic}}
+
+With spatio-temporal data, users may wish to make plots to learn the spatial distribution of a variable or to find patterns, such as trend or seasonality, in the time series.
+Combining these two types of plots with interactivity lets users link between points on the map and the corresponding time series to explore the spatial and temporal dimensions of the data simultaneously.
+Below is an example that describes the process of building an interactive graphic with \pkg{cubble} and \pkg{crosstalk}.
+The example explores the variation of monthly temperature range with \code{weatherdata::climate_full} data.
 
@@
 A \code{SharedData} object is constructed for each form of the cubble and the same \code{group} argument ensures the cross-linking of the two forms via the common \code{id} column.
-The spatial map and time series plot are then made with each \code{SharedData} objects separately.
-In this example, stations on the Australia map, made from the nested form, are coloured by the calculated variance and a ribbon band is constructed using the long form cubble to show the maximum and minimum temperature of each station across month.
-With a different dataset, users are free to calculate any per station measure in the nested form or to make any time-wise summary of the data in the long form to customise the spatial or temporal view.
+The spatial map and time series plot are then made with each \code{SharedData} object separately.
+In this example, stations on the Australia map, made from the nested form, are coloured by the calculated variance.
+A ribbon band is constructed using the long form cubble to show each station's maximum and minimum temperature across month.
+With a different dataset, users can calculate other per station measure in the nested form or make other time-wise summary of the data in the long form to customise the spatial or temporal view.
 The cross-linking between the two plots is always safeguarded by the shared \code{id} column embedded in the cubble structure.
-Below is the pseudo code that outlines the process to construct an interactive graphic described above:
+Below is the pseudo-code that outlines the process to construct an interactive graphic described above:
 
@@
 \begin{figure}
-\includegraphics[width=0.45\linewidth,height=0.25\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/popup-mid} \caption{Same as Figure 11 with thetemperature variation shown as a popup in the leaflet map.}\label{fig:interactive-popup}
+
+{\centering \includegraphics[width=0.45\linewidth,height=0.25\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/popup-mid}
+
+}
+
+\caption{Same as Figure 11 with thetemperature variation shown as a popup in the leaflet map.}\label{fig:interactive-popup}
 \end{figure}

### 2022-04-23 1-intro.Rmd
more poilsh on wording

@@
 With spatio-temporal data, one can fix the time to explore the spatial distribution, fix the spatial location to explore temporal trends, or dynamically explore the space and time simultaneously.
-Underlying all these tasks is a data representation that enables the spatio-temporal data to be sliced and diced in different ways.
+Underlying all these tasks is a data representation that enables the spatio-temporal data to be sliced and diced differently.
 
@@
 Still, the data representation for spatio-temporal data can be further improved and there are two reasons for this.
-Firstly, the raw data sourced in the wild is less often presented in any one of the representations above, and fitting the raw data into one of the representations can sometimes be not easy.
-More often, spatio-temporal data is collected in separate 2D tables and analysts need to ensemble them into a whole piece before exploring the data.
-Examples of components of spatio-temporal data can be 1) map data records the shape of a collection of area of interest is provided by the official government websites; 2) geo-scientific data records the longitude and latitude coordinates of locations in the area along with other meta data relates to the location, and; 3) temporal variables indexed by both location and time.
+Firstly, the raw data sourced in the wild is less often presented in any one of the representations above, and fitting the raw data into a data object can sometimes be not easy.
+More often, spatio-temporal data are collected in separate 2D tables and analysts need to ensemble them into a whole piece before exploring the data.
+Examples of components of spatio-temporal data can be 1) map data records the shape of a collection of area of interest; 2) geo-scientific data records the longitude and latitude coordinates of locations in the area along with other meta data relates to the location, and; 3) temporal variables indexed by both location and time.
 
@@
 
-The other reason relates to how the concept of tidy data [@tidydata] should be applied to spatio-temporal data.
-The tidy data concept advocates the data to be shaped into 1) one row per observation; 2) one column per variable, and; 3) one type of data per table.
+The other reason is how tidy data [@tidydata] should be applied to spatio-temporal data.
+The tidy data concept advocates the data to be structured into 1) one row per observation, 2) one column per variable, and 3) one type of data per table.
 The long form data is preferred over wide data given the downstream software i.e.
@@
 Section \@ref(cubble) presents the main design and functionality of \pkg{cubble}.
-Section \@ref(others) explains how cubble deals with more advanced considerations including data with the hierarchical structure, data matching, how cubble fits with existing static and interactive visualisation tools, and spatio-temporal data transformation.
+Section \@ref(others) explains how cubble deals with more advanced considerations, including data with the hierarchical structure, data matching, how cubble fits with existing static and interactive visualisation tools, and spatio-temporal data transformation.
 Section \@ref(examples) uses Australia weather station data and river level data as examples to demonstrate the use of \pkg{cubble}.

### 2022-04-23 2-0-cube.Rmd
more poilsh on wording

@@
 Spatio-temporal data can be conceptualised using a cubical data model with three axes.
-This abstraction can be useful to generalise operations or even visualisation: @lu_multidimensional_2018 shows an example of how array operations (select, scale, reduce, rearrange, and compute) can be mapped onto the cube; @bach_review_2014 reviews the temporal data visualisation based on space-time cube operations.
+This abstraction can be useful for generalising operations and visualisation: @lu_multidimensional_2018 shows how array operations (select, scale, reduce, rearrange, and compute) can be mapped onto the cube; @bach_review_2014 reviews the temporal data visualisation based on space-time cube operations.
 Notice that the term space-time cube in their article "does not need to involve spatial data", but refers to "an abstract 2D substrate that is used to visualize data at a specific time".
-Despite its main focus being on temporal data, the mindset of abstracting out data representation to construct visualisation is applicable to our spatio-temporal data manipulation and visualisation.
+Despite its main focus being on temporal data, the mindset of abstracting out data representation to construct visualisation still applies to our spatio-temporal data manipulation and visualisation.
 
-While a cubical conceptual data model is neat to conceptually view the spatio-temporal data, a 3D data representation, or an array,  may not be convenient for data wrangling.
+While a cubical data model is neat to view the spatio-temporal data conceptually, a 3D data representation, or an array,  may not be convenient for data wrangling.
 There are two reasons for this: 1) While arrays are efficient for the computation on numerical values, spatio-temporal data contains more than just numerical variables.
-Wrangling on character strings and specific datetime classes are common tasks for spatio-temporal data but are not easy with arrays.
+For example, wrangling on character strings and specific datetime classes is common for spatio-temporal data but may not be easy done with array objects.
 2) The `mutate()` syntax for creating new variables is no longer valid since the operation can no longer define how a created 1D vector should be added to the array.
 
-There have been many cubes defined using time, latitude, and longitude, which can be seen as taking snapshots of the space and stacking on time.
+The most common data cube is defined using the three axes: time, latitude, and longitude, which can be seen as taking snapshots of the space and stacking on time.
 Our spatio-temporal cube is defined slightly differently with three axes: time, site, and variable as Figure \ref{fig:cube-def}.
 The time axis is the same across the two formations, the site encompasses latitude and longitude into a single column identifier, while variables are stacked on this defined space-time canvas.
-This notion is adopted to avoid the use of hypercube when describing multivariate spatio-temporal data.
+This notion is adopted to avoid using hypercube when describing multivariate spatio-temporal data.
 

### 2022-04-23 3-advanced.Rmd
more poilsh on wording

@@
 One task that may interest analysts in spatio-temporal data is to find how similar the time series from nearby sites are.
-This can be seen as a matching problem [@stuart2010matching; @mcintosh2018using] that pairs up similar time series in nearby locations or a data fusing exercise that merges data collected from different sources [@cocchi2019data].
+This problem can be seen as a matching problem [@stuart2010matching; @mcintosh2018using] that pairs up similar time series in nearby locations or a data fusing exercise that merges data collected from different sources [@cocchi2019data].
 \code{match_sites()} in \pkg{cubble} provides a simple algorithm for this task.
-The algorithm first matches the two data sources spatially through computing the pairwise distance on latitude and longitude.
+The algorithm first matches the two data sources spatially by computing the pairwise distance on latitude and longitude.
 Pairs that pass the spatial matching are then matched temporally through computing the number of matched peaks within a fixed length window.
 Figure \ref{fig:illu-matching} illustrates this temporal matching in more details.
-Given two series \code{A} and \code{a}, 3 peaks have been picked in each series.
-An interval, with default length of 5, is constructed for each peak in series \code{A} and the peaks in series \code{a} are tested against whether they fall into the any of the intervals.
-In this illustration, there are 2 matches for these two series.
+Given two series \code{A} and \code{a}, three peaks have been picked in each series.
+An interval, with default length of five, is constructed for each peak in series \code{A} and the peaks in series \code{a} are tested against whether they fall into the any of the intervals.
+In this illustration, there are two matches for these two series.
 Several arguments are available in \code{match_sites()} to fine-tune the matching:
@@
 
-Cubble fits in naturally with the interactive graphic pipeline discussed in the literature [@buja1988elements; @buja1996interactive;  @sutherland2000orca; @xie2014reactive; @cheng2016enabling].
+The linking structure of two forms in cubble fits naturally with the interactive graphic pipeline discussed in the literature [@buja1988elements; @buja1996interactive;  @sutherland2000orca; @xie2014reactive; @cheng2016enabling].
 Diagram \ref{fig:illu-interactive} illustrates how linking works from the map to the time series in cubble.
@@
 
-Spatio-temporal transformation is a useful technique to extract information form spatio-temporal data.
+Spatio-temporal transformation is a useful technique to extract information from spatio-temporal data.
 Glyph maps [@Wickham2012-yr] transform the time coordinates into the space coordinates to plot the time series of different locations on the map.
-Calendar plots [@wang2020calendar] reconstructs time into calendar-based grid to discover weekday and weekend pattern.
+Calendar plots [@wang2020calendar] reconstructs time into a calendar-based grid to discover weekday and weekend pattern.
 Projection, or linear combination, of variables summarises multivariate information into lower dimension to further digest.

### 2022-04-23 arxiv/cubble-arxiv.Rmd
more poilsh on wording

@@
 In practice, spatio-temporal data can be decomposed into a spatial and temporal component: at one time, we would select a spatial location and inspect the temporal trend; at other time, we might select one or multiple time value(s) and explore the spatial distribution.
-Ideally, we could make multiple maps and multiple time series to explore these together, however, doing all of these actions is complicated when data arrive fragmented in multiple objects.
-To make it easy to do all these tasks, ideally spatial and temporal variables are in a single data object that we can slice and dice in different ways to conduct different visualisations.
+Ideally, we could make multiple maps and multiple time series to explore these together.
+However, doing all of these actions is complicated when data arrive fragmented in multiple objects.
+To make it easy to do all these tasks, ideally, spatial and temporal variables are in a single data object that we can slice and dice in different ways to conduct different visualisations.
 This work suggests a new data structure, \code{cubble}, that uses a nested form and a long form to organise spatial and temporal variables in a single data object.
-The new data structure ensures that data in the two forms are synchronised while been flexible to explore the two aspect of spatio-temporal data.
-It also provides capability for handling data with hierarchical structure, matching data from multiple sources, constructing interactive graphics, and performing spatio-temporal transformation.
+The new data structure ensures that data in the two forms are synchronised while being flexible to explore the two aspects of spatio-temporal data.
+It also provides the capability to handle data with hierarchical structure, match data from multiple sources, construct interactive graphics, and perform spatio-temporal transformation.
 The paper will demonstrate \pkg{cubble} with examples using Australian climate weather stations, river level data, and climate reanalysis (ERA5) data.

### 2022-04-23 arxiv/cubble-arxiv.tex
more poilsh on wording

@@
 Still, the data representation for spatio-temporal data can be further improved and there are two reasons for this.
-Firstly, the raw data sourced in the wild is less often presented in any one of the representations above, and fitting the raw data into one of the representations can sometimes be not easy.
-More often, spatio-temporal data is collected in separate 2D tables and analysts need to ensemble them into a whole piece before exploring the data.
-Examples of components of spatio-temporal data can be 1) map data records the shape of a collection of area of interest is provided by the official government websites; 2) geo-scientific data records the longitude and latitude coordinates of locations in the area along with other meta data relates to the location, and; 3) temporal variables indexed by both location and time.
-
-The other reason relates to how the concept of tidy data (\protect\hyperlink{ref-tidydata}{Wickham 2014}) should be applied to spatio-temporal data.
-The tidy data concept advocates the data to be shaped into 1) one row per observation; 2) one column per variable, and; 3) one type of data per table.
+Firstly, the raw data sourced in the wild is less often presented in any one of the representations above, and fitting the raw data into a data object can sometimes be not easy.
+More often, spatio-temporal data are collected in separate 2D tables and analysts need to ensemble them into a whole piece before exploring the data.
+Examples of components of spatio-temporal data can be 1) map data records the shape of a collection of area of interest; 2) geo-scientific data records the longitude and latitude coordinates of locations in the area along with other meta data relates to the location, and; 3) temporal variables indexed by both location and time.
+
+The other reason is how tidy data (\protect\hyperlink{ref-tidydata}{Wickham 2014}) should be applied to spatio-temporal data.
+The tidy data concept advocates the data to be structured into 1) one row per observation, 2) one column per variable, and 3) one type of data per table.
 The long form data is preferred over wide data given the downstream software i.e.~\pkg{dplyr} (\protect\hyperlink{ref-dplyr}{Wickham et al.
@@
 Section \ref{cubble} presents the main design and functionality of \pkg{cubble}.
-Section \ref{others} explains how cubble deals with more advanced considerations including data with the hierarchical structure, data matching, how cubble fits with existing static and interactive visualisation tools, and spatio-temporal data transformation.
+Section \ref{others} explains how cubble deals with more advanced considerations, including data with the hierarchical structure, data matching, how cubble fits with existing static and interactive visualisation tools, and spatio-temporal data transformation.
 Section \ref{examples} uses Australia weather station data and river level data as examples to demonstrate the use of \pkg{cubble}.
@@
 Spatio-temporal data can be conceptualised using a cubical data model with three axes.
-This abstraction can be useful to generalise operations or even visualisation: Lu, Appel, and Pebesma (\protect\hyperlink{ref-lu_multidimensional_2018}{2018}) shows an example of how array operations (select, scale, reduce, rearrange, and compute) can be mapped onto the cube; Bach et al.
+This abstraction can be useful for generalising operations and visualisation: Lu, Appel, and Pebesma (\protect\hyperlink{ref-lu_multidimensional_2018}{2018}) shows how array operations (select, scale, reduce, rearrange, and compute) can be mapped onto the cube; Bach et al.
 (\protect\hyperlink{ref-bach_review_2014}{2014}) reviews the temporal data visualisation based on space-time cube operations.
 Notice that the term space-time cube in their article ``does not need to involve spatial data'', but refers to ``an abstract 2D substrate that is used to visualize data at a specific time''.
-Despite its main focus being on temporal data, the mindset of abstracting out data representation to construct visualisation is applicable to our spatio-temporal data manipulation and visualisation.
-
-While a cubical conceptual data model is neat to conceptually view the spatio-temporal data, a 3D data representation, or an array, may not be convenient for data wrangling.
+Despite its main focus being on temporal data, the mindset of abstracting out data representation to construct visualisation still applies to our spatio-temporal data manipulation and visualisation.
+
+While a cubical data model is neat to view the spatio-temporal data conceptually, a 3D data representation, or an array, may not be convenient for data wrangling.
 There are two reasons for this: 1) While arrays are efficient for the computation on numerical values, spatio-temporal data contains more than just numerical variables.
-Wrangling on character strings and specific datetime classes are common tasks for spatio-temporal data but are not easy with arrays.
+For example, wrangling on character strings and specific datetime classes is common for spatio-temporal data but may not be easy done with array objects.
 2) The \texttt{mutate()} syntax for creating new variables is no longer valid since the operation can no longer define how a created 1D vector should be added to the array.
 
-There have been many cubes defined using time, latitude, and longitude, which can be seen as taking snapshots of the space and stacking on time.
+The most common data cube is defined using the three axes: time, latitude, and longitude, which can be seen as taking snapshots of the space and stacking on time.
 Our spatio-temporal cube is defined slightly differently with three axes: time, site, and variable as Figure \ref{fig:cube-def}.
 The time axis is the same across the two formations, the site encompasses latitude and longitude into a single column identifier, while variables are stacked on this defined space-time canvas.
-This notion is adopted to avoid the use of hypercube when describing multivariate spatio-temporal data.
+This notion is adopted to avoid using hypercube when describing multivariate spatio-temporal data.
 
@@
 One task that may interest analysts in spatio-temporal data is to find how similar the time series from nearby sites are.
-This can be seen as a matching problem (\protect\hyperlink{ref-stuart2010matching}{Stuart 2010}; \protect\hyperlink{ref-mcintosh2018using}{McIntosh et al.
+This problem can be seen as a matching problem (\protect\hyperlink{ref-stuart2010matching}{Stuart 2010}; \protect\hyperlink{ref-mcintosh2018using}{McIntosh et al.
 2018}) that pairs up similar time series in nearby locations or a data fusing exercise that merges data collected from different sources (\protect\hyperlink{ref-cocchi2019data}{Cocchi 2019}).
 \code{match_sites()} in \pkg{cubble} provides a simple algorithm for this task.
-The algorithm first matches the two data sources spatially through computing the pairwise distance on latitude and longitude.
+The algorithm first matches the two data sources spatially by computing the pairwise distance on latitude and longitude.
 Pairs that pass the spatial matching are then matched temporally through computing the number of matched peaks within a fixed length window.
 Figure \ref{fig:illu-matching} illustrates this temporal matching in more details.
-Given two series \code{A} and \code{a}, 3 peaks have been picked in each series.
-An interval, with default length of 5, is constructed for each peak in series \code{A} and the peaks in series \code{a} are tested against whether they fall into the any of the intervals.
-In this illustration, there are 2 matches for these two series.
+Given two series \code{A} and \code{a}, three peaks have been picked in each series.
+An interval, with default length of five, is constructed for each peak in series \code{A} and the peaks in series \code{a} are tested against whether they fall into the any of the intervals.
+In this illustration, there are two matches for these two series.
 Several arguments are available in \code{match_sites()} to fine-tune the matching:
@@
 
-Cubble fits in naturally with the interactive graphic pipeline discussed in the literature (\protect\hyperlink{ref-buja1988elements}{Buja, Asimov, and Hurley 1988}; \protect\hyperlink{ref-buja1996interactive}{Buja, Cook, and Swayne 1996}; \protect\hyperlink{ref-sutherland2000orca}{Sutherland et al.
+The linking structure of two forms in cubble fits naturally with the interactive graphic pipeline discussed in the literature (\protect\hyperlink{ref-buja1988elements}{Buja, Asimov, and Hurley 1988}; \protect\hyperlink{ref-buja1996interactive}{Buja, Cook, and Swayne 1996}; \protect\hyperlink{ref-sutherland2000orca}{Sutherland et al.
 2000}; \protect\hyperlink{ref-xie2014reactive}{Xie, Hofmann, and Cheng 2014}; \protect\hyperlink{ref-cheng2016enabling}{Cheng, Cook, and Hofmann 2016}).
@@
 
-Spatio-temporal transformation is a useful technique to extract information form spatio-temporal data.
+Spatio-temporal transformation is a useful technique to extract information from spatio-temporal data.
 Glyph maps (\protect\hyperlink{ref-Wickham2012-yr}{Wickham et al.
 2012}) transform the time coordinates into the space coordinates to plot the time series of different locations on the map.
-Calendar plots (\protect\hyperlink{ref-wang2020calendar}{Wang, Cook, and Hyndman 2020}) reconstructs time into calendar-based grid to discover weekday and weekend pattern.
+Calendar plots (\protect\hyperlink{ref-wang2020calendar}{Wang, Cook, and Hyndman 2020}) reconstructs time into a calendar-based grid to discover weekday and weekend pattern.
 Projection, or linear combination, of variables summarises multivariate information into lower dimension to further digest.

### 2022-04-26 1-intro.Rmd
minor fixes
minor fixes Sections 1 to 3

@@
 More often, spatio-temporal data are collected in separate 2D tables and analysts need to ensemble them into a whole piece before exploring the data.
-Examples of components of spatio-temporal data can be 1) map data records the shape of a collection of area of interest; 2) geo-scientific data records the longitude and latitude coordinates of locations in the area along with other meta data relates to the location, and; 3) temporal variables indexed by both location and time.
+Examples of components of spatio-temporal data can be 1) map data recording the shape of a collection of area of interest; 2) geo-scientific data recording the longitude and latitude coordinates of locations in the area along with other metadata related to the location, and; 3) temporal variables indexed by both location and time.
 
@@
 However, the long form can be inefficient to store feature geometries, especially for large multipolygon over daily or sub-daily data over years, which are extensively collected and handled in time series analysis.
-This poses the question of how to arrange spatial and temporal variables that makes analysing spatio-temporal data easier.
+This poses the question of how to arrange spatial and temporal variables in a way that would make analysing spatio-temporal data easier.
 
@@
 
-The rest of the paper is divided as follows: Section \@ref(cube) discusses the spatio-temporal cube to conceptualise spatio-temporal data.
+The rest of the paper is divided as follows: Section \@ref(cube) discusses the proposed spatio-temporal cube to conceptualise spatio-temporal data.
 Section \@ref(cubble) presents the main design and functionality of \pkg{cubble}.

### 2022-04-26 2-0-cube.Rmd
minor fixes
minor fixes Sections 1 to 3

@@
 There are two reasons for this: 1) While arrays are efficient for the computation on numerical values, spatio-temporal data contains more than just numerical variables.
-For example, wrangling on character strings and specific datetime classes is common for spatio-temporal data but may not be easy done with array objects.
+For example, wrangling on character strings and specific datetime classes is common for spatio-temporal data but may not be easily done with array objects.
 2) The `mutate()` syntax for creating new variables is no longer valid since the operation can no longer define how a created 1D vector should be added to the array.
@@
 The time axis is the same across the two formations, the site encompasses latitude and longitude into a single column identifier, while variables are stacked on this defined space-time canvas.
-This notion is adopted to avoid using hypercube when describing multivariate spatio-temporal data.
+This notion is adopted to avoid using hypercubes when describing multivariate spatio-temporal data.
 
-```{r cube-def, echo = FALSE, fig.align="center", out.width = "50%", fig.cap = "An ilustration of the conceputalised spatio-temporal cube with three axies: time, site, and variable."}
+```{r cube-def, echo = FALSE, fig.align="center", out.width = "50%", fig.cap = "An illustration of the conceptualised spatio-temporal cube with three axes: time, site, and variable."}
 knitr::include_graphics(here::here("figures/cube-def.png"))

### 2022-04-26 2-the-cubble-pkg.Rmd
minor fixes
minor fixes Sections 1 to 3

@@
 
-Each core function is introduced with a short example accompanied and the data `climate_flat` is used throughout this section.
-`climate_flat` contains five weather stations in Australia with spatial information of each station: station id, latitude, longitude, elevation, station name, and World Meteorology Organisation ID and also daily temporal information of date, maximum and minimum temperature and precipitation for 2020.
-Below prints out the first five rows of `climate_flat`:
+Each core function is introduced accompanied with a short example and the data `climate_flat` is used throughout this section.
+`climate_flat` contains five weather stations in Australia with spatial information of each station: station id, latitude, longitude, elevation, station name, World Meteorology Organisation ID and also daily temporal information of date, maximum and minimum temperature and precipitation for 2020.
+The first five rows of `climate_flat` are shown below:
 
@@
 Spatio-temporal data can come in various formats and shapes and a cubble can be created from various raw data.
-This includes tibble and its variates multiple tables (an example provided in section \@ref(covid)), and netcdf (detailed in section \@ref(netcdf)).
+This includes tibble and its variates multiple tables (an example provided in section \@ref(covid)), and netCDF (detailed in section \@ref(netcdf)).
 The function \code{as_cubble()} is used to create a cubble with three additional components: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order of longitude and latitude.
-Argument \code{key} and \code{index} follow the wording in \pkg{tsibble} to describe the temporal order and multiple series while \code{coords} is a space specific argument to locate each site in two columns.
+The arguments \code{key} and \code{index} follow the wording in \pkg{tsibble} to describe the temporal order and multiple series while \code{coords} is a space specific argument to locate each site in two columns.
 The code below creates a cubble out of \code{climate_flat} with \code{id} as the key, \code{date} as the index, and \code{c(long, lat)} as the coordinates:
@@
 The bounding box is \code{[115.97, -32.94, 133.55, -12.42]}.
-In the third row is the the name and type of variable nested in the \code{ts} column and here it includes \code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.
+In the third row is the name and type of variable nested in the \code{ts} column and here it includes \code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.
 
@@
 The nested form can be used for those operations whose output is only indexed by the spatial identifier (\code{key}), while for those outputs cross-identified by the spatial and temporal identifier (\code{key} and \code{index}), the nested form becomes inadequate.
-\pkg{cubble} also provides a long form, which expands the \code{ts} column and temporarily "hide" the spatial variables.
+\pkg{cubble} also provides a long form, which expands the \code{ts} column and temporarily "hides" the spatial variables.
 The function \code{face_temporal()} is used to switch a nested cubble into a long one and the first row in Figure \ref{fig:face} illustrates this operation where the focus of the cube now changes from the site-variable face to the time-variable face.
@@
 The first line in the header now shows the cubble is in its long form and the third line has been changed to show the name and type of spatial variables: \code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]}.
-Unlike the nested form, the long cubble is built form class \code{groupped_df} where all the observations from the same sites form a group.
+Unlike the nested form, the long cubble is built from class \code{groupped_df} where all the observations from the same sites form a group.
 
@@
 In the first row, \\code{face\\_temporal} switches a cubble from the nested form into the long form and the focus has switched from the spatial aspect (the side face) to the temporal aspect (the front face).
-In the second row, \\code{face\\_spatial} switches a cubble back to the nested form from te long form and shift focus back to the spatial aspect."}
+In the second row, \\code{face\\_spatial} switches a cubble back to the nested form from the long form and shifts focus back to the spatial aspect."}
 knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.008.png"))
@@
 <!-- https://www.jstatsoft.org/style#my-latex-paper-does-not-compile-when-there-is-jss-markup-in-section-titles-what-should-i-do -->
-\subsection[Why not just use]{Why not just use \code{dplyr::nest\_by()}}
+\subsection[Why not just use]{Why not just use \code{dplyr::nest\_by()}?}
 
@@
 While \code{nest_by()} and its pair \code{unnest()} can mimic the nested form in cubble, cubble can be seen as an attempt to arrange variables from different observational units into a single object, specifically for spatio-temporal data.
-While `nest_by()` can recreate the nested form, it can't create a clean slate for temporal variables as the long form.
+While `nest_by()` can recreate the nested form, it cannot create a clean slate for temporal variables as the long form.
 For temporal variables, carrying the duplicated spatial variables in the combined form is not convenient to work with.
-This is because these spatial variables will lose after \code{summarise()} and one will need to subset the spatial variables again from the original data and make a left join, which tends to create unnecessary intermediate products in the workflow.
+This is because these spatial variables will be lost after \code{summarise()} and one will need to subset the spatial variables again from the original data and make a left join, which tends to create unnecessary intermediate products in the workflow.
 
@@
 Once formed in a cubble, methods for the \code{sfc} class can be applied in the nested form.
-An example of this is shown in \@ref(covid), which also handles the case when the sites in the two tables don't match exactly.
+An example of this is shown in \@ref(covid), which also handles the case when the sites in the two tables do not match exactly.
 A spatial data object with \code{s2} vector can also be the input of the spatial table.
@@
 Attributes are usually associated with dimensions and variables in the NetCDF format data and a [metadata convention for climate and forecast](http://cfconventions.org/) has been designed to standardise the format of the attributes.
-A few packages in R exists for manipulating NetCDF data and this includes a high-level R interface: \pkg{ncdf4} [@ncdf4], a low-level interface that calls C-interface: \pkg{RNetCDF} [@rnetcdf; @michna2013rnetcdf], and a tidyverse implementation: \pkg{tidync} [@tidync].
+A few packages in R exist for manipulating NetCDF data and these include a high-level R interface: \pkg{ncdf4} [@ncdf4], a low-level interface that calls a C-interface: \pkg{RNetCDF} [@rnetcdf; @michna2013rnetcdf], and a tidyverse implementation: \pkg{tidync} [@tidync].
 
@@
 
-The memory limit with NetCDF data in cubble depends on the number of longitude grid points x the number of latitude grid points x the number of time grid points x the number of variables.
-Cubble can handle slightly more than 300 x 300 (longitude x longitude) grid points for daily 3 variables in one year.
+The memory limit with NetCDF data in cubble depends on the number of longitude grid points $\times$ the number of latitude grid points $\times$ the number of time grid points $\times$ the number of variables.
+Cubble can handle slightly more than 300 $\times$ 300 (longitude $\times$ longitude) grid points for three daily variables in one year.
 A 300 by 300 spatial grid can be a bounding box of [100, -80, 180, 0] at 0.25 degree resolution or a global bounding box [-180, -90, 180, -90] at 1 degree resolution.
@@
 ```{r eval = FALSE, echo = TRUE}
-# Assume my_ncdf has bounding box of [-180, -90, 180, -90]
+# Assume my_ncdf has a bounding box of [-180, -90, 180, -90]
 # at 0.25 degree resolution and subset it to have

### 2022-04-26 3-advanced.Rmd
minor fixes
minor fixes Sections 1 to 3

@@
 
-Spatial locations can have grouping structure either inherent in the data or formed by clustering.
+Spatial locations can have grouping structures either inherent in the data or formed by clustering.
 Rather than analysing variables in the site level, summarised variables in the cluster level can give a crisper picture of local areas.
@@
 Figure \ref{fig:illu-hier} illustrates the relationship of cubbles at station and cluster level, in both the long and nested form.
-By specifying \code{cluster_nested <- station_nested \%>\% switch_key(key = cluster)}, the cubble re-defines the cubble key from \code{id} in \code{station_nested} to \code{cluster} in \code{cluster_nested}.
+By specifying \code{cluster_nested <- station_nested \%>\% switch_key(key = cluster)}, the cubble redefines the cubble key from the \code{id} in \code{station_nested} to \code{cluster} in \code{cluster_nested}.
 All the spatial variables variant to \code{cluster} are now nested into a \code{.val} column and cluster level variables can be computed in the same fashion as station level variables in \code{station_nested}.
@@
 Given two series \code{A} and \code{a}, three peaks have been picked in each series.
-An interval, with default length of five, is constructed for each peak in series \code{A} and the peaks in series \code{a} are tested against whether they fall into the any of the intervals.
+An interval, with default length of five, is constructed for each peak in series \code{A} and the peaks in series \code{a} are tested against whether they fall into any of the intervals.
 In this illustration, there are two matches for these two series.
@@
 Three highest peaks are identified in each series and intervals are constructed on series \\code{A}.
-Two peaks in series \\code{a} fall into the intervals and hence the two series are considered to have 2 matches."}
+Two peaks in series \\code{a} fall into the intervals and hence the two series are considered to have two matches."}
 knitr::include_graphics(here::here("figures/illu-matching.png"))
@@
  - \code{temporal_window}: the length of the interval - 5 in the example above
- - \code{temporal_min_match}: the minimum number of matched peak for a valid matched pair
+ - \code{temporal_min_match}: the minimum number of matched peaks for a valid matched pair
 
@@
 The linking structure of two forms in cubble fits naturally with the interactive graphic pipeline discussed in the literature [@buja1988elements; @buja1996interactive;  @sutherland2000orca; @xie2014reactive; @cheng2016enabling].
-Diagram \ref{fig:illu-interactive} illustrates how linking works from the map to the time series in cubble.
+The diagram \ref{fig:illu-interactive} illustrates how linking works from the map to the time series in cubble.
 The map and time series plot is associated with the nested or long cubble, respectively, and when a user action is captured on the map, the site will be activated in the nested cubble (left).
@@
 All the observations that share the same \code{id} are then activated and this includes other points in the same time series in the long cubble and the corresponding observation of site in the nested cubble.
-These activated observations will then being reflected in the updated plots and Diagram \ref{fig:illu-interactive-2} in the Appendix illustrates this process.
+These activated observations will then be reflected in the updated plots and Diagram \ref{fig:illu-interactive-2} in the Appendix illustrates this process.
 
@@
 Glyph maps [@Wickham2012-yr] transform the time coordinates into the space coordinates to plot the time series of different locations on the map.
-Calendar plots [@wang2020calendar] reconstructs time into a calendar-based grid to discover weekday and weekend pattern.
+Calendar plots [@wang2020calendar] reconstruct time into a calendar-based grid to discover weekday and weekend pattern.
 Projection, or linear combination, of variables summarises multivariate information into lower dimension to further digest.
@@
 
-A re-implementation of the glyph map as a ggproto, \code{GeomGlyph}, has been made in the \pkg{cubble} package and now the glyph map can be created with \code{geom_glyph()}:
+A reimplementation of the glyph map as a ggproto, \code{GeomGlyph}, has been made in the \pkg{cubble} package and now the glyph map can be created with \code{geom_glyph()}:
 
@@
 
-Some useful controls over the glyph map is also available in the \code{geom_glyph()} implementation.
-Polar glyph map can be specify as a parameter, \code{polar = TRUE}.
-in the \code{geom_glyph()}, along with \code{width} and \code{height} in either absolute or relative value.
-Global and local scale can be controlled by the parameter \code{global_rescale}.
-which default to \code{TRUE} for global scaling.
+Some useful controls over the glyph map are also available in the \code{geom_glyph()} implementation.
+Polar glyph map can be specified as a parameter \code{polar = TRUE} in the \code{geom_glyph()}, along with \code{width} and \code{height} in either absolute or relative value.
+Global and local scale can be controlled by the parameter \code{global_rescale} which default to \code{TRUE} for global scaling.
 Reference box and line can be added with separate \code{geom_glyph_box()} and \code{geom_glyph_line()}.

### 2022-04-26 arxiv/cubble-arxiv.Rmd
minor fixes
minor fixes Sections 1 to 3

@@
 ---
-title: cubble An R Package for Structuring Spatio-temporal Data
+title: cubble: An R Package for Structuring Spatio-temporal Data
 authors:
@@
     Spatio-temporal data refer to measurements taken across space and time.
-In practice, spatio-temporal data can be decomposed into a spatial and temporal component: at one time, we would select a spatial location and inspect the temporal trend; at other time, we might select one or multiple time value(s) and explore the spatial distribution.
-Ideally, we could make multiple maps and multiple time series to explore these together.
+In practice, spatio-temporal data can be decomposed into a spatial and temporal component: at one time, one can select a spatial location and inspect the temporal trend; at other time, one might select one or multiple time value(s) and explore the spatial distribution.
+Ideally, multiple maps and multiple time series could be made to explore these together.
 However, doing all of these actions is complicated when data arrive fragmented in multiple objects.
-To make it easy to do all these tasks, ideally, spatial and temporal variables are in a single data object that we can slice and dice in different ways to conduct different visualisations.
+To make it easy to do all these tasks, ideally, spatial and temporal variables should be in a single data object that we can slice and dice in different ways to conduct different visualisations.
 This work suggests a new data structure, \code{cubble}, that uses a nested form and a long form to organise spatial and temporal variables in a single data object.
@@
 It also provides the capability to handle data with hierarchical structure, match data from multiple sources, construct interactive graphics, and perform spatio-temporal transformation.
-The paper will demonstrate \pkg{cubble} with examples using Australian climate weather stations, river level data, and climate reanalysis (ERA5) data.
+The paper will demonstrate the use of \pkg{cubble} with examples using Australian climate weather stations, river level data, and climate reanalysis (ERA5) data.
 keywords:
-- spatial, temporal, spatio temporal, R, environmental data, exploratory data analysis
+- spatial, temporal, spatio temporal, R, exploratory data analysis, environmental data, climate data
 bibliography: "../references.bib"

### 2022-04-26 1-intro.Rmd
Some rewriting and comments added

@@
 
-The SpatioTemproal CRAN task view [@ctvspatiotemporal] has a section *Representing data* that lists existing spatio-temporal data representation in \proglang{R}.
+The SpatioTemproal CRAN task view [@ctvspatiotemporal] has a section *Representing data* that lists existing spatio-temporal data representations in \proglang{R}.
 Among them, @spacetime proposes four spatio-temporal layouts: full grid layout, sparse grid layout, irregular layout, and trajectory.
-These layouts are implemented in the \pkg{spacetime} [@spacetime] package with the underlying spatial and temporal class: \pkg{sp} [@sp] and \pkg{xts} [@xts].
+These layouts are implemented in the \pkg{spacetime} [@spacetime] package with the underlying spatial and temporal classes \pkg{sp} [@sp] and \pkg{xts} [@xts].
 The \pkg{stars} [@stars] package has been a recent implementation to handle both raster and vector data using spatio-temporal arrays.
@@
 Still, the data representation for spatio-temporal data can be further improved and there are two reasons for this.
-Firstly, the raw data sourced in the wild is less often presented in any one of the representations above, and fitting the raw data into a data object can sometimes be not easy.
-More often, spatio-temporal data are collected in separate 2D tables and analysts need to ensemble them into a whole piece before exploring the data.
-Examples of components of spatio-temporal data can be 1) map data recording the shape of a collection of area of interest; 2) geo-scientific data recording the longitude and latitude coordinates of locations in the area along with other metadata related to the location, and; 3) temporal variables indexed by both location and time.
+Firstly, the raw data sourced in the wild is less often presented in any one of the representations above, and fitting the raw data into a data object can sometimes be difficult.
+More often, spatio-temporal data are collected in separate 2D tables and analysts need to assemble them into a whole piece before exploring the data.
+Examples of components of spatio-temporal data can be 1) map data recording the shape of a collection of areas of interest; 2) geo-scientific data recording the longitude and latitude coordinates of locations in the area along with other metadata related to the location, and; 3) temporal variables indexed by both location and time.
 
@@
 
-The other reason is how tidy data [@tidydata] should be applied to spatio-temporal data.
-The tidy data concept advocates the data to be structured into 1) one row per observation, 2) one column per variable, and 3) one type of data per table.
+The other reason is how tidy data concepts [@tidydata] should be applied to spatio-temporal data.
+Tidy data should be structured into 1) one row per observation, 2) one column per variable, and 3) one type of data per table.
 The long form data is preferred over wide data given the downstream software i.e.
 \pkg{dplyr} [@dplyr] and \pkg{ggplot2} [@ggplot2] for data wrangling and visualisation.
-However, the long form can be inefficient to store feature geometries, especially for large multipolygon over daily or sub-daily data over years, which are extensively collected and handled in time series analysis.
+However, the long form can be inefficient to store feature geometries, especially for large multipolygons for daily or sub-daily data over years, which are extensively collected and handled in time series analysis.
 This poses the question of how to arrange spatial and temporal variables in a way that would make analysing spatio-temporal data easier.
 
-This paper presents a new \proglang{R} package, \pkg{cubble}, for working with a new spatio-temporal data representation, which addresses the two issues mentioned above.
-In the package, a new data structure, also called \code{cubble}, is proposed to organise spatial and temporal variables as two forms of a single data object so that they can be wrangled separately while synchronised as a whole piece.
+This paper presents a new \proglang{R} package, \pkg{cubble} which addresses the two issues mentioned above.
+In the package, a new data structure, also called \code{cubble}, is proposed to organise spatial and temporal variables as two forms of a single data object so that they can be wrangled separately while being kept synchronised.
 Among the four spacetime layouts in @spacetime, \pkg{cubble} can be applied to full grid layout, sparse grid layout, or irregular layout, but not trajectory, which is outside the scope of this work.
@@
 
-The rest of the paper is divided as follows: Section \@ref(cube) discusses the proposed spatio-temporal cube to conceptualise spatio-temporal data.
+The rest of the paper is divided as follows: Section \@ref(cube) introduces the proposed cube structure as a way to conceptualise multivariate spatio-temporal data.
 Section \@ref(cubble) presents the main design and functionality of \pkg{cubble}.
 Section \@ref(others) explains how cubble deals with more advanced considerations, including data with the hierarchical structure, data matching, how cubble fits with existing static and interactive visualisation tools, and spatio-temporal data transformation.
-Section \@ref(examples) uses Australia weather station data and river level data as examples to demonstrate the use of \pkg{cubble}.
+Section \@ref(examples) uses Australian weather station data and river level data as examples to demonstrate the use of \pkg{cubble}.
 An example of how \pkg{cubble} handles NetCDF data is also provided.

### 2022-04-26 2-0-cube.Rmd
Some rewriting and comments added

@@
 
-Spatio-temporal data can be conceptualised using a cubical data model with three axes.
+\textcolor{blue}{Ursula: I have rearranged this section somewhat, please check!}
+
+Spatio-temporal data can be conceptualised using a cubical data model with three axes, typically time, latitude and longitude.
 This abstraction can be useful for generalising operations and visualisation: @lu_multidimensional_2018 shows how array operations (select, scale, reduce, rearrange, and compute) can be mapped onto the cube; @bach_review_2014 reviews the temporal data visualisation based on space-time cube operations.
@@
 Despite its main focus being on temporal data, the mindset of abstracting out data representation to construct visualisation still applies to our spatio-temporal data manipulation and visualisation.
+
+The most common data cube using the three axes (time, latitude, longitude)  could be considered as taking snapshots of the space and stacking on time.
+Here we will define the spatio-temporal cube slightly differently with the three axes (time, site, variable), as illustrated in Figure \ref{fig:cube-def}.
+The time axis is the same in both versions, while the site axis now captures both latitude and longitude.
+Finally, variables are stacked on this space-time canvas, with one observation per site and time point.
+This notion is adopted to avoid using hypercubes when describing multivariate spatio-temporal data.
 
@@
 2) The `mutate()` syntax for creating new variables is no longer valid since the operation can no longer define how a created 1D vector should be added to the array.
-
-The most common data cube is defined using the three axes: time, latitude, and longitude, which can be seen as taking snapshots of the space and stacking on time.
-Our spatio-temporal cube is defined slightly differently with three axes: time, site, and variable as Figure \ref{fig:cube-def}.
-The time axis is the same across the two formations, the site encompasses latitude and longitude into a single column identifier, while variables are stacked on this defined space-time canvas.
-This notion is adopted to avoid using hypercubes when describing multivariate spatio-temporal data.
+\textcolor{blue}{Ursula: this seems too specific, you should be able to make the same argument without referencing the mutate function?}
 

### 2022-04-26 2-the-cubble-pkg.Rmd
Some rewriting and comments added

@@
 
-This section will first introduce the core functions in the cubble class: \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and \code{unfold()} in subsection \@ref(create) to \@ref(unfold).
+This section will first introduce the core functions in the cubble class: \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and \code{unfold()} in subsection \@ref(create) - \@ref(unfold).
 The next subsection addresses the necessity of creating a new class.
@@
 
-Spatio-temporal data can come in various formats and shapes and a cubble can be created from various raw data.
-This includes tibble and its variates multiple tables (an example provided in section \@ref(covid)), and netCDF (detailed in section \@ref(netcdf)).
-The function \code{as_cubble()} is used to create a cubble with three additional components: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order of longitude and latitude.
-The arguments \code{key} and \code{index} follow the wording in \pkg{tsibble} to describe the temporal order and multiple series while \code{coords} is a space specific argument to locate each site in two columns.
+Spatio-temporal data can come in various formats and shapes and a cubble can be created from various types of raw data.
+This includes tibble and its variates in multiple tables (an example is provided in section \@ref(covid)), and netCDF (detailed in section \@ref(netcdf)).
+The function \code{as_cubble()} is used to create a cubble with three additional inputs: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order  (longitude, latitude).
+The arguments \code{key} and \code{index} follow the wording in \pkg{tsibble} to describe the temporal order and multiple series while \code{coords} specifies the spatial location of each site.
 The code below creates a cubble out of \code{climate_flat} with \code{id} as the key, \code{date} as the index, and \code{c(long, lat)} as the coordinates:
@@
 The bounding box is \code{[115.97, -32.94, 133.55, -12.42]}.
-In the third row is the name and type of variable nested in the \code{ts} column and here it includes \code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.
+The third row shows the name and type of all variables nested in the \code{ts} column and here it includes \code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.
 
-The created cubble is built from a \code{rowwise_df} class where each row forms a group.
+The resulting cubble is built from a \code{rowwise_df} class where each row forms a group.
 All the temporal variables are nested in a list column, hence it is also called the nested cubble.
-The rowwise structure makes it simpler to calculate on the list using the \code{mutate()} syntax, which is simpler than the {\code{map()} when working with list column.
+The rowwise structure makes it simpler to calculate on the list using the \code{mutate()} syntax, which is simpler than the \code{map()} when working with list column.
 For example, calculating the number of raining days can be done by:
@@
 
-## Change focus to the time-variable face
+## Change focus by facing the time-variables
 
-The nested form can be used for those operations whose output is only indexed by the spatial identifier (\code{key}), while for those outputs cross-identified by the spatial and temporal identifier (\code{key} and \code{index}), the nested form becomes inadequate.
+The nested form can be used for those operations where the output is only indexed by the spatial identifier (\code{key}), but becomes inadequate when outputs need both a spatial and a temporal identifier (\code{key} and \code{index}).
 \pkg{cubble} also provides a long form, which expands the \code{ts} column and temporarily "hides" the spatial variables.
@@
 
-## Flatten spatial variables into the long cubble {#unfold}
+## Unfold spatial variables into the long cubble {#unfold}
 
@@
 
+\textcolor{blue}{Ursula: I feel like this doesn't get the message accross, the main point seems to be that with cubble we can switch and never have to add information back in "by-hand" (with some join function), maybe make the point without refering to specific functions first and then explain the example?}
+
 Some readers may raise the question why a new data structure rather than using \code{dplyr::nest_by()} on the combined data to directly create a list-column.
-While \code{nest_by()} and its pair \code{unnest()} can mimic the nested form in cubble, cubble can be seen as an attempt to arrange variables from different observational units into a single object, specifically for spatio-temporal data.
-While `nest_by()` can recreate the nested form, it cannot create a clean slate for temporal variables as the long form.
+While \code{nest_by()} and its inverse \code{unnest()} can mimic the nested form in cubble, cubble can be seen as an attempt to arrange variables from different observational units into a single object, specifically for spatio-temporal data.
+While `nest_by()` can recreate the nested form, it cannot create a clean slate for temporal variables as used in the long form.
 For temporal variables, carrying the duplicated spatial variables in the combined form is not convenient to work with.
@@
 
-The previous four subsections have introduced operations specific to the \code{cubble} class and this section will demonstrate how the \code{cubble} class interacts with existing packages commonly used in spatial and temporal analysis, specifically, \code{dplyr}, \code{tsibble}, \code{sf} (\code{s2}), and \code{netcdf4}.
+This section will demonstrate how the \code{cubble} class interacts with existing packages commonly used in spatial and temporal analysis, specifically, \code{dplyr}, \code{tsibble}, \code{sf} (\code{s2}), and \code{netcdf4}.
 
@@
 
-In the nested cubble created, each element in the list-column \code{ts} is of \code{tbl_ts} class and operations available to the tsibble class is still valid under cubble.
+In the nested cubble created, each element in the list-column \code{ts} is of \code{tbl_ts} class and operations available to the tsibble class are still valid within the cubble.
 For example, the code below calculates two features of the maximum temperature:
@@
 As a spatial data object, \code{sf} creates a simple feature geometry list-column (\code{sfc}) in the data frame to provide spatial operations on various geometry types (\code{POINT}, \code{LINESTRING}, \code{POLYGON}, \code{MULTIPOLYGON}, etc).
-An \code{sf} object can be supplied as the spatial table, as part of the raw data, along with a temporal table, to create a cubble.
-Once formed in a cubble, methods for the \code{sfc} class can be applied in the nested form.
-An example of this is shown in \@ref(covid), which also handles the case when the sites in the two tables do not match exactly.
-A spatial data object with \code{s2} vector can also be the input of the spatial table.
+When creating a cubble an \code{sf} object can be supplied as the spatial table.
+Once included in a cubble, methods for the \code{sfc} class can be applied in the nested form.
+An example of this is shown in \@ref(covid), which also handles the case when the site identifiers in the two tables do not match exactly.
+A spatial data object with \code{s2} vector can also be the input for the spatial table.
 

### 2022-04-26 3-advanced.Rmd
Some rewriting and comments added

@@
 
-Spatial locations can have grouping structures either inherent in the data or formed by clustering.
+Spatial locations can have grouping structures either inherent in the data or obtained by clustering.
 Rather than analysing variables in the site level, summarised variables in the cluster level can give a crisper picture of local areas.
@@
  - \code{spatial_dist_max}: the maximum distance allowed for a matched pair
- - \code{temporal_n_highest}: the number of peak used - 3 in the example above
+ - \code{temporal_n_highest}: the number of peaks used - 3 in the example above
  - \code{temporal_window}: the length of the interval - 5 in the example above
@@
 The linking structure of two forms in cubble fits naturally with the interactive graphic pipeline discussed in the literature [@buja1988elements; @buja1996interactive;  @sutherland2000orca; @xie2014reactive; @cheng2016enabling].
-The diagram \ref{fig:illu-interactive} illustrates how linking works from the map to the time series in cubble.
+Diagram \ref{fig:illu-interactive} illustrates how linking works from the map to the time series in cubble.
 The map and time series plot is associated with the nested or long cubble, respectively, and when a user action is captured on the map, the site will be activated in the nested cubble (left).
@@
 
-Spatio-temporal transformation is a useful technique to extract information from spatio-temporal data.
-Glyph maps [@Wickham2012-yr] transform the time coordinates into the space coordinates to plot the time series of different locations on the map.
+Different types of transformations can be useful to extract new information from spatio-temporal data.
+Glyph maps [@Wickham2012-yr] transform the time coordinates into space coordinates to plot the time series of different locations on the map.
 Calendar plots [@wang2020calendar] reconstruct time into a calendar-based grid to discover weekday and weekend pattern.

### 2022-04-26 4-1-basic.Rmd
Some rewriting and comments added

@@
 
-Cubble can be created with separate spatial and temporal table in a list, with other arguments (\code{key}, \code{index}, and \code{coords}) introduced in section \@ref(create).
-\code{as_cubble()} will automatically check the match of site in both tables emit a warning if a location has missing spatial or temporal information.
+Cubble can be created with separate spatial and temporal tables in a list, with other arguments (\code{key}, \code{index}, and \code{coords}) introduced in section \@ref(create).
+\code{as_cubble()} will automatically check the match of sites in both tables and emit a warning if a location has missing spatial or temporal information.
 
@@
 Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world.
-The dataset \code{weatherdata::historical_tmax} extracts the maximum temperature for `r nrow(weatherdata::historical_tmax)` Australian stations from GHCN with starting from year `r range(weatherdata::historical_tmax$first_year)[2]`.
+The dataset \code{weatherdata::historical_tmax} extracts the maximum temperature for `r nrow(weatherdata::historical_tmax)` Australian stations from GHCN, starting from year `r range(weatherdata::historical_tmax$first_year)[2]`.
 \code{weatherdata::historical_tmax} is already in a cubble, with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates.
@@
 Lastly, to create a glyph map, both the major (\code{longitude}, \code{latitude}) and minor (\code{month}, \code{tmax}) coordinates need to be in the same table.
-Spatial variables can be moved to the long form with \code{migrate()}:
+Spatial variables can be moved to the long form with \code{unfold()}:
 

### 2022-04-26 4-2-agg.Rmd
Some rewriting and comments added

@@
 
-In the previous example, there has already been some overlapping of the glyphs for a few stations near (151E, 34S) and (152E, 33S), which is a problem when mapping more stations at the national level.
+In the previous example, there has been some overlap of glyphs when stations are near each other, (151E, 34S) and (152E, 33S), which is a problem when mapping more stations at the national level.
 Aggregation can be helpful in grouping series into clusters before visualising the cluster with a glyph map.
@@
 
-Long form cubble at both levels can be accessed through stretching the nested form and with access to both station and cluster level cubbles, various plots can be made to understand the cluster.
+Long form cubbles at both levels can be accessed through stretching the nested form and with access to both station and cluster level cubbles, various plots can be made to understand the cluster.
 Figure \ref{fig:basic-agg} shows two example plots that can be made with this data.
@@
 Subplot A shows the glyph map of the weekly averaged precipitation of each cluster.
-The group number of printed in the middle of y minor axis and can be used as a reference line to read the magnitude.
+The group number is printed in the middle of y minor axis and can be used as a reference line to read the magnitude.
 Subplot B shows the station membership of each cluster."}

### 2022-04-26 4-3-water-level.Rmd
Some rewriting and comments added

@@
 Bureau of Meteorology collects [water data](http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full) from river gauges and this includes variables: electrical conductivity, turbidity, watercourse discharge, watercourse level, and water temperature.
-In particular, water level will interact with precipitation from the climate data since rainfall will raise the water level in the river.
+In particular, we expect that water levels will interact with precipitation from the climate data since rainfall will raise the water level in the river.
 Figure \ref{fig:matching-map} gives the location of available weather stations and water gauges in Victoria.
@@
 
-From the map, a few water gauges and weather stations are close to each other and the fluctuation of the water level could be matched up with precipitation measured by the climate station.
+From the map we can see that a few water gauges and weather stations are close to each other and the fluctuation of the water level could be matched up with precipitation measured by the climate station.
 As introduced in Section 3.2, \code{match_sites()} can be used to match one source of data with another source in a cubble.
@@
 \code{temporal_independent} controls the variable used to construct the interval and the goal here is to see if the water level in the river will reflect precipitation.
-This puts precipitation \code{prcp}, as the independent.
+This puts precipitation \code{prcp}, as the independent variable.
 Given there is one year's worth of data, the number of peaks (\code{temporal_n_highest}) to consider is slightly raised from a default of 20 to 30 and \code{temporal_min_match} is raised accordingly To return all the pairs of the match, \code{temporal_min_match} can be set to 0.
@@
 
-Figure \ref{fig:matching} plots the matched pairs on the map or to view the matched series.
-There are four pairs of matches, all located in the middle of Victoria and the concurrent increase in precipitation and water level can be observed.
+Figure \ref{fig:matching} shows the matched pairs on the map and a view the matched time series.
+There are four pairs of matches, all located in the middle of Victoria, for which the concurrent increase in precipitation and water level can be observed.
 

### 2022-04-26 4-5-interactive.Rmd
Some rewriting and comments added

@@
 In Figure \ref{fig:interactive-linking}, the first row shows the initial view of the interactive graphic.
-Most regions in Australia have low variance of temperature range On the map, while the north-west coastline, bottom of South Australia, and Victoria stand out with larger monthly changes.
+Most regions in Australia have low variance of temperature range, while the north-west coastline, bottom of South Australia, and Victoria stand out with larger monthly changes.
 In the second row, Mount Elizabeth is selected on the map given its high variance colour on the initial map and this links to the ribbon on the right.
@@
 
-Figure \ref{fig:interactive-popup} shows Figure \ref{fig:interactive-linking} made made with leaflet and popups [@leafpop].
+Figure \ref{fig:interactive-popup} shows a reproduction of Figure \ref{fig:interactive-linking} made  with leaflet and popups [@leafpop].
 
-```{r interactive-popup, fig.align='center', echo = FALSE, out.width="45%", out.height="25%", fig.retina = 2, dpi = 300, fig.cap = "Same as Figure 11 with thetemperature variation shown as a popup in the leaflet map.", fig.show='hold'}
+```{r interactive-popup, fig.align='center', echo = FALSE, out.width="45%", out.height="25%", fig.retina = 2, dpi = 300, fig.cap = "Same as Figure 11 with the temperature variation shown as a popup in the leaflet map.", fig.show='hold'}
 knitr::include_graphics(here::here("figures/popup-mid.png")) 

### 2022-04-26 arxiv/cubble-arxiv.Rmd
Some rewriting and comments added

@@
 ---
-title: cubble: An R Package for Structuring Spatio-temporal Data
+title: "cubble: An R Package for Structuring Spatio-temporal Data"
 authors:
@@
     Spatio-temporal data refer to measurements taken across space and time.
-In practice, spatio-temporal data can be decomposed into a spatial and temporal component: at one time, one can select a spatial location and inspect the temporal trend; at other time, one might select one or multiple time value(s) and explore the spatial distribution.
+In practice, spatio-temporal data can be decomposed into a spatial and temporal component: one can select a spatial location and inspect the temporal trend; or, one might select one or multiple time value(s) and explore the spatial distribution.
 Ideally, multiple maps and multiple time series could be made to explore these together.

### 2022-04-26 arxiv/cubble-arxiv.tex
Some rewriting and comments added

@@
 
-\title{cubble An R Package for Structuring Spatio-temporal Data}
+\title{cubble: An R Package for Structuring Spatio-temporal Data}
 
@@
 Spatio-temporal data refer to measurements taken across space and time.
-In practice, spatio-temporal data can be decomposed into a spatial and temporal component: at one time, we would select a spatial location and inspect the temporal trend; at other time, we might select one or multiple time value(s) and explore the spatial distribution.
-Ideally, we could make multiple maps and multiple time series to explore these together.
+In practice, spatio-temporal data can be decomposed into a spatial and temporal component: one can select a spatial location and inspect the temporal trend; or, one might select one or multiple time value(s) and explore the spatial distribution.
+Ideally, multiple maps and multiple time series could be made to explore these together.
 However, doing all of these actions is complicated when data arrive fragmented in multiple objects.
-To make it easy to do all these tasks, ideally, spatial and temporal variables are in a single data object that we can slice and dice in different ways to conduct different visualisations.
+To make it easy to do all these tasks, ideally, spatial and temporal variables should be in a single data object that we can slice and dice in different ways to conduct different visualisations.
 This work suggests a new data structure, \code{cubble}, that uses a nested form and a long form to organise spatial and temporal variables in a single data object.
@@
 It also provides the capability to handle data with hierarchical structure, match data from multiple sources, construct interactive graphics, and perform spatio-temporal transformation.
-The paper will demonstrate \pkg{cubble} with examples using Australian climate weather stations, river level data, and climate reanalysis (ERA5) data.
+The paper will demonstrate the use of \pkg{cubble} with examples using Australian climate weather stations, river level data, and climate reanalysis (ERA5) data.
 \end{abstract}
@@
 \keywords{
-    spatial, temporal, spatio temporal, R, environmental data, exploratory data analysis
+    spatial, temporal, spatio temporal, R, exploratory data analysis, environmental data, climate data
   }
@@
 
-The SpatioTemproal CRAN task view (\protect\hyperlink{ref-ctvspatiotemporal}{Edzer Pebesma 2022}) has a section \emph{Representing data} that lists existing spatio-temporal data representation in \proglang{R}.
-Among them, E.
-Pebesma (\protect\hyperlink{ref-spacetime}{2012}) proposes four spatio-temporal layouts: full grid layout, sparse grid layout, irregular layout, and trajectory.
+The SpatioTemproal CRAN task view (\protect\hyperlink{ref-ctvspatiotemporal}{Edzer Pebesma 2022}) has a section \emph{Representing data} that lists existing spatio-temporal data representations in \proglang{R}.
+Among them, \protect\hyperlink{ref-spacetime}{E.
+Pebesma} (\protect\hyperlink{ref-spacetime}{2012}) proposes four spatio-temporal layouts: full grid layout, sparse grid layout, irregular layout, and trajectory.
 These layouts are implemented in the \pkg{spacetime} (\protect\hyperlink{ref-spacetime}{E.
-Pebesma 2012}) package with the underlying spatial and temporal class: \pkg{sp} (\protect\hyperlink{ref-sp}{E.
+Pebesma 2012}) package with the underlying spatial and temporal classes \pkg{sp} (\protect\hyperlink{ref-sp}{E.
 Pebesma and Bivand 2005}) and \pkg{xts} (\protect\hyperlink{ref-xts}{Ryan and Ulrich 2020}).
@@
 Still, the data representation for spatio-temporal data can be further improved and there are two reasons for this.
-Firstly, the raw data sourced in the wild is less often presented in any one of the representations above, and fitting the raw data into a data object can sometimes be not easy.
-More often, spatio-temporal data are collected in separate 2D tables and analysts need to ensemble them into a whole piece before exploring the data.
-Examples of components of spatio-temporal data can be 1) map data records the shape of a collection of area of interest; 2) geo-scientific data records the longitude and latitude coordinates of locations in the area along with other meta data relates to the location, and; 3) temporal variables indexed by both location and time.
-
-The other reason is how tidy data (\protect\hyperlink{ref-tidydata}{Wickham 2014}) should be applied to spatio-temporal data.
-The tidy data concept advocates the data to be structured into 1) one row per observation, 2) one column per variable, and 3) one type of data per table.
+Firstly, the raw data sourced in the wild is less often presented in any one of the representations above, and fitting the raw data into a data object can sometimes be difficult.
+More often, spatio-temporal data are collected in separate 2D tables and analysts need to assemble them into a whole piece before exploring the data.
+Examples of components of spatio-temporal data can be 1) map data recording the shape of a collection of areas of interest; 2) geo-scientific data recording the longitude and latitude coordinates of locations in the area along with other metadata related to the location, and; 3) temporal variables indexed by both location and time.
+
+The other reason is how tidy data concepts (\protect\hyperlink{ref-tidydata}{Wickham 2014}) should be applied to spatio-temporal data.
+Tidy data should be structured into 1) one row per observation, 2) one column per variable, and 3) one type of data per table.
 The long form data is preferred over wide data given the downstream software i.e.~\pkg{dplyr} (\protect\hyperlink{ref-dplyr}{Wickham et al.
 2022}) and \pkg{ggplot2} (\protect\hyperlink{ref-ggplot2}{Wickham 2016}) for data wrangling and visualisation.
-However, the long form can be inefficient to store feature geometries, especially for large multipolygon over daily or sub-daily data over years, which are extensively collected and handled in time series analysis.
-This poses the question of how to arrange spatial and temporal variables that makes analysing spatio-temporal data easier.
-
-This paper presents a new \proglang{R} package, \pkg{cubble}, for working with a new spatio-temporal data representation, which addresses the two issues mentioned above.
-In the package, a new data structure, also called \code{cubble}, is proposed to organise spatial and temporal variables as two forms of a single data object so that they can be wrangled separately while synchronised as a whole piece.
-Among the four spacetime layouts in E.
-Pebesma (\protect\hyperlink{ref-spacetime}{2012}), \pkg{cubble} can be applied to full grid layout, sparse grid layout, or irregular layout, but not trajectory, which is outside the scope of this work.
+However, the long form can be inefficient to store feature geometries, especially for large multipolygons for daily or sub-daily data over years, which are extensively collected and handled in time series analysis.
+This poses the question of how to arrange spatial and temporal variables in a way that would make analysing spatio-temporal data easier.
+
+This paper presents a new \proglang{R} package, \pkg{cubble} which addresses the two issues mentioned above.
+In the package, a new data structure, also called \code{cubble}, is proposed to organise spatial and temporal variables as two forms of a single data object so that they can be wrangled separately while being kept synchronised.
+Among the four spacetime layouts in \protect\hyperlink{ref-spacetime}{E.
+Pebesma} (\protect\hyperlink{ref-spacetime}{2012}), \pkg{cubble} can be applied to full grid layout, sparse grid layout, or irregular layout, but not trajectory, which is outside the scope of this work.
 The software is available from the Comprehensive R Archive Network (CRAN) at {[}CRAN link{]}.
 
-The rest of the paper is divided as follows: Section \ref{cube} discusses the spatio-temporal cube to conceptualise spatio-temporal data.
+The rest of the paper is divided as follows: Section \ref{cube} introduces the proposed cube structure as a way to conceptualise multivariate spatio-temporal data.
 Section \ref{cubble} presents the main design and functionality of \pkg{cubble}.
 Section \ref{others} explains how cubble deals with more advanced considerations, including data with the hierarchical structure, data matching, how cubble fits with existing static and interactive visualisation tools, and spatio-temporal data transformation.
-Section \ref{examples} uses Australia weather station data and river level data as examples to demonstrate the use of \pkg{cubble}.
+Section \ref{examples} uses Australian weather station data and river level data as examples to demonstrate the use of \pkg{cubble}.
 An example of how \pkg{cubble} handles NetCDF data is also provided.
@@
 
-Spatio-temporal data can be conceptualised using a cubical data model with three axes.
-This abstraction can be useful for generalising operations and visualisation: Lu, Appel, and Pebesma (\protect\hyperlink{ref-lu_multidimensional_2018}{2018}) shows how array operations (select, scale, reduce, rearrange, and compute) can be mapped onto the cube; Bach et al.
-(\protect\hyperlink{ref-bach_review_2014}{2014}) reviews the temporal data visualisation based on space-time cube operations.
-Notice that the term space-time cube in their article ``does not need to involve spatial data'', but refers to ``an abstract 2D substrate that is used to visualize data at a specific time''.
-Despite its main focus being on temporal data, the mindset of abstracting out data representation to construct visualisation still applies to our spatio-temporal data manipulation and visualisation.
+\textcolor{blue}{Ursula: I have rearranged this section somewhat, please check!}
+
+Spatio-temporal data can be conceptualised using a cubical data model with three axes, typically time, latitude and longitude.
+This abstraction can be useful for generalising operations and visualisation: \protect\hyperlink{ref-lu_multidimensional_2018}{Lu, Appel, and Pebesma} (\protect\hyperlink{ref-lu_multidimensional_2018}{2018}) shows how array operations (select, scale, reduce, rearrange, and compute) can be mapped onto the cube; \protect\hyperlink{ref-bach_review_2014}{Bach et al.} (\protect\hyperlink{ref-bach_review_2014}{2014}) reviews the temporal data visualisation based on space-time cube operations.
+Notice that the term space-time cube in their article ``does not need to involve spatial data,'' but refers to ``an abstract 2D substrate that is used to visualize data at a specific time.'' Despite its main focus being on temporal data, the mindset of abstracting out data representation to construct visualisation still applies to our spatio-temporal data manipulation and visualisation.
+
+The most common data cube using the three axes (time, latitude, longitude) could be considered as taking snapshots of the space and stacking on time.
+Here we will define the spatio-temporal cube slightly differently with the three axes (time, site, variable), as illustrated in Figure \ref{fig:cube-def}.
+The time axis is the same in both versions, while the site axis now captures both latitude and longitude.
+Finally, variables are stacked on this space-time canvas, with one observation per site and time point.
+This notion is adopted to avoid using hypercubes when describing multivariate spatio-temporal data.
 
@@
 There are two reasons for this: 1) While arrays are efficient for the computation on numerical values, spatio-temporal data contains more than just numerical variables.
-For example, wrangling on character strings and specific datetime classes is common for spatio-temporal data but may not be easy done with array objects.
+For example, wrangling on character strings and specific datetime classes is common for spatio-temporal data but may not be easily done with array objects.
 2) The \texttt{mutate()} syntax for creating new variables is no longer valid since the operation can no longer define how a created 1D vector should be added to the array.
-
-The most common data cube is defined using the three axes: time, latitude, and longitude, which can be seen as taking snapshots of the space and stacking on time.
-Our spatio-temporal cube is defined slightly differently with three axes: time, site, and variable as Figure \ref{fig:cube-def}.
-The time axis is the same across the two formations, the site encompasses latitude and longitude into a single column identifier, while variables are stacked on this defined space-time canvas.
-This notion is adopted to avoid using hypercube when describing multivariate spatio-temporal data.
+\textcolor{blue}{Ursula: this seems too specific, you should be able to make the same argument without referencing the mutate function?}
 
@@
 
-{\centering \includegraphics[width=0.5\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/cube-def}
+{\centering \includegraphics[width=0.5\linewidth]{/Users/ursulalaa/papers/paper-cubble/figures/cube-def}
 
@@
 
-\caption{An ilustration of the conceputalised spatio-temporal cube with three axies: time, site, and variable.}\label{fig:cube-def}
+\caption{An illustration of the conceptualised spatio-temporal cube with three axes: time, site, and variable.}\label{fig:cube-def}
 \end{figure}
@@
 
-This section will first introduce the core functions in the cubble class: \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and \code{unfold()} in subsection \ref{create} to \ref{unfold}.
+This section will first introduce the core functions in the cubble class: \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and \code{unfold()} in subsection \ref{create} - \ref{unfold}.
 The next subsection addresses the necessity of creating a new class.
@@
 
-Each core function is introduced with a short example accompanied and the data \texttt{climate\_flat} is used throughout this section.
-\texttt{climate\_flat} contains five weather stations in Australia with spatial information of each station: station id, latitude, longitude, elevation, station name, and World Meteorology Organisation ID and also daily temporal information of date, maximum and minimum temperature and precipitation for 2020.
-Below prints out the first five rows of \texttt{climate\_flat}:
+Each core function is introduced accompanied with a short example and the data \texttt{climate\_flat} is used throughout this section.
+\texttt{climate\_flat} contains five weather stations in Australia with spatial information of each station: station id, latitude, longitude, elevation, station name, World Meteorology Organisation ID and also daily temporal information of date, maximum and minimum temperature and precipitation for 2020.
+The first five rows of \texttt{climate\_flat} are shown below:
 
@@
 
-Spatio-temporal data can come in various formats and shapes and a cubble can be created from various raw data.
-This includes tibble and its variates multiple tables (an example provided in section \ref{covid}), and netcdf (detailed in section \ref{netcdf}).
-The function \code{as_cubble()} is used to create a cubble with three additional components: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order of longitude and latitude.
-Argument \code{key} and \code{index} follow the wording in \pkg{tsibble} to describe the temporal order and multiple series while \code{coords} is a space specific argument to locate each site in two columns.
+Spatio-temporal data can come in various formats and shapes and a cubble can be created from various types of raw data.
+This includes tibble and its variates in multiple tables (an example is provided in section \ref{covid}), and netCDF (detailed in section \ref{netcdf}).
+The function \code{as_cubble()} is used to create a cubble with three additional inputs: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order (longitude, latitude).
+The arguments \code{key} and \code{index} follow the wording in \pkg{tsibble} to describe the temporal order and multiple series while \code{coords} specifies the spatial location of each site.
 The code below creates a cubble out of \code{climate_flat} with \code{id} as the key, \code{date} as the index, and \code{c(long, lat)} as the coordinates:
@@
 \begin{Highlighting}[]
-\NormalTok{cubble\_nested }\OtherTok{\textless{}{-}}\NormalTok{ climate\_flat }\SpecialCharTok{|\textgreater{}}
+\NormalTok{cubble\_nested }\OtherTok{\textless{}{-}}\NormalTok{ climate\_flat }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
   \FunctionTok{as\_cubble}\NormalTok{(}\AttributeTok{key =}\NormalTok{ id, }\AttributeTok{index =}\NormalTok{ date, }\AttributeTok{coords =} \FunctionTok{c}\NormalTok{(long, lat))}
@@
 The bounding box is \code{[115.97, -32.94, 133.55, -12.42]}.
-In the third row is the the name and type of variable nested in the \code{ts} column and here it includes \code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.
-
-The created cubble is built from a \code{rowwise_df} class where each row forms a group.
+The third row shows the name and type of all variables nested in the \code{ts} column and here it includes \code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.
+
+The resulting cubble is built from a \code{rowwise_df} class where each row forms a group.
 All the temporal variables are nested in a list column, hence it is also called the nested cubble.
-The rowwise structure makes it simpler to calculate on the list using the \code{mutate()} syntax, which is simpler than the \{\code{map()} when working with list column.
+The rowwise structure makes it simpler to calculate on the list using the \code{mutate()} syntax, which is simpler than the \code{map()} when working with list column.
 For example, calculating the number of raining days can be done by:
@@
 \begin{Highlighting}[]
-\NormalTok{cubble\_nested }\SpecialCharTok{|\textgreater{}} 
+\NormalTok{cubble\_nested }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
   \FunctionTok{mutate}\NormalTok{(}\AttributeTok{rain\_day =} \FunctionTok{sum}\NormalTok{(ts}\SpecialCharTok{$}\NormalTok{prcp }\SpecialCharTok{!=} \DecValTok{0}\NormalTok{))}
@@
 
-\hypertarget{change-focus-to-the-time-variable-face}{%
-\subsection{Change focus to the time-variable face}\label{change-focus-to-the-time-variable-face}}
-
-The nested form can be used for those operations whose output is only indexed by the spatial identifier (\code{key}), while for those outputs cross-identified by the spatial and temporal identifier (\code{key} and \code{index}), the nested form becomes inadequate.
-\pkg{cubble} also provides a long form, which expands the \code{ts} column and temporarily ``hide'' the spatial variables.
+\hypertarget{change-focus-by-facing-the-time-variables}{%
+\subsection{Change focus by facing the time-variables}\label{change-focus-by-facing-the-time-variables}}
+
+The nested form can be used for those operations where the output is only indexed by the spatial identifier (\code{key}), but becomes inadequate when outputs need both a spatial and a temporal identifier (\code{key} and \code{index}).
+\pkg{cubble} also provides a long form, which expands the \code{ts} column and temporarily ``hides'' the spatial variables.
 The function \code{face_temporal()} is used to switch a nested cubble into a long one and the first row in Figure \ref{fig:face} illustrates this operation where the focus of the cube now changes from the site-variable face to the time-variable face.
@@
 \begin{Highlighting}[]
-\NormalTok{cubble\_long }\OtherTok{\textless{}{-}}\NormalTok{ cubble\_nested }\SpecialCharTok{|\textgreater{}} \FunctionTok{face\_temporal}\NormalTok{()}
+\NormalTok{cubble\_long }\OtherTok{\textless{}{-}}\NormalTok{ cubble\_nested }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{face\_temporal}\NormalTok{()}
 \NormalTok{cubble\_long}
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.008}
+{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-cubble/figures/diagram-keynotes/diagram-keynotes.008}
 
@@
 In the first row, \code{face\_temporal} switches a cubble from the nested form into the long form and the focus has switched from the spatial aspect (the side face) to the temporal aspect (the front face).
-In the second row, \code{face\_spatial} switches a cubble back to the nested form from te long form and shift focus back to the spatial aspect.}\label{fig:face}
+In the second row, \code{face\_spatial} switches a cubble back to the nested form from the long form and shifts focus back to the spatial aspect.}\label{fig:face}
 \end{figure}
@@
 \begin{Highlighting}[]
-\NormalTok{cubble\_back }\OtherTok{\textless{}{-}}\NormalTok{ cubble\_long }\SpecialCharTok{|\textgreater{}} \FunctionTok{face\_spatial}\NormalTok{()}
+\NormalTok{cubble\_back }\OtherTok{\textless{}{-}}\NormalTok{ cubble\_long }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{face\_spatial}\NormalTok{()}
 \NormalTok{cubble\_back}
@@
 \hypertarget{unfold}{%
-\subsection{Flatten spatial variables into the long cubble}\label{unfold}}
+\subsection{Unfold spatial variables into the long cubble}\label{unfold}}
 
@@
 \begin{Highlighting}[]
-\NormalTok{cubble\_unfold }\OtherTok{\textless{}{-}}\NormalTok{ cubble\_long }\SpecialCharTok{|\textgreater{}} \FunctionTok{unfold}\NormalTok{(long, lat)}
+\NormalTok{cubble\_unfold }\OtherTok{\textless{}{-}}\NormalTok{ cubble\_long }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{unfold}\NormalTok{(long, lat)}
 \NormalTok{cubble\_unfold}
@@
 \begin{Highlighting}[]
-\NormalTok{cubble\_unfold }\SpecialCharTok{|\textgreater{}} \FunctionTok{face\_spatial}\NormalTok{() }\SpecialCharTok{|\textgreater{}} \FunctionTok{face\_temporal}\NormalTok{()}
+\NormalTok{cubble\_unfold }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{face\_spatial}\NormalTok{() }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{face\_temporal}\NormalTok{()}
 \end{Highlighting}
@@
 
-\subsection[Why not just use]{Why not just use \code{dplyr::nest\_by()}}
+\subsection[Why not just use]{Why not just use \code{dplyr::nest\_by()}?}
+
+\textcolor{blue}{Ursula: I feel like this doesn't get the message accross, the main point seems to be that with cubble we can switch and never have to add information back in "by-hand" (with some join function), maybe make the point without refering to specific functions first and then explain the example?}
 
 Some readers may raise the question why a new data structure rather than using \code{dplyr::nest_by()} on the combined data to directly create a list-column.
-While \code{nest_by()} and its pair \code{unnest()} can mimic the nested form in cubble, cubble can be seen as an attempt to arrange variables from different observational units into a single object, specifically for spatio-temporal data.
-While \texttt{nest\_by()} can recreate the nested form, it can't create a clean slate for temporal variables as the long form.
+While \code{nest_by()} and its inverse \code{unnest()} can mimic the nested form in cubble, cubble can be seen as an attempt to arrange variables from different observational units into a single object, specifically for spatio-temporal data.
+While \texttt{nest\_by()} can recreate the nested form, it cannot create a clean slate for temporal variables as used in the long form.
 For temporal variables, carrying the duplicated spatial variables in the combined form is not convenient to work with.
-This is because these spatial variables will lose after \code{summarise()} and one will need to subset the spatial variables again from the original data and make a left join, which tends to create unnecessary intermediate products in the workflow.
+This is because these spatial variables will be lost after \code{summarise()} and one will need to subset the spatial variables again from the original data and make a left join, which tends to create unnecessary intermediate products in the workflow.
 
@@
 \CommentTok{\# cubble workflow}
-\NormalTok{dt }\SpecialCharTok{|\textgreater{}} 
-  \FunctionTok{as\_cubble}\NormalTok{(...) }\SpecialCharTok{|\textgreater{}} 
-  \FunctionTok{face\_temporal}\NormalTok{() }\SpecialCharTok{|\textgreater{}} 
-  \FunctionTok{group\_by}\NormalTok{() }\SpecialCharTok{|\textgreater{}} 
-  \FunctionTok{summarise}\NormalTok{() }\SpecialCharTok{|\textgreater{}} 
+\NormalTok{dt }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
+  \FunctionTok{as\_cubble}\NormalTok{(...) }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
+  \FunctionTok{face\_temporal}\NormalTok{() }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
+  \FunctionTok{group\_by}\NormalTok{() }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
+  \FunctionTok{summarise}\NormalTok{() }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
   \FunctionTok{unfold}\NormalTok{(...)}
@@
 \CommentTok{\# tidyverse workflow}
-\NormalTok{temporal }\OtherTok{\textless{}{-}}\NormalTok{ dt }\SpecialCharTok{|\textgreater{}} 
-  \FunctionTok{group\_by}\NormalTok{(...) }\SpecialCharTok{|\textgreater{}} 
+\NormalTok{temporal }\OtherTok{\textless{}{-}}\NormalTok{ dt }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
+  \FunctionTok{group\_by}\NormalTok{(...) }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
   \FunctionTok{summarise}\NormalTok{(...)}
 
-\NormalTok{spatial }\OtherTok{\textless{}{-}}\NormalTok{ dt }\SpecialCharTok{|\textgreater{}} \FunctionTok{select}\NormalTok{(....)}
-
-\NormalTok{spatial }\SpecialCharTok{|\textgreater{}} \FunctionTok{left\_join}\NormalTok{(temporal)}
+\NormalTok{spatial }\OtherTok{\textless{}{-}}\NormalTok{ dt }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{select}\NormalTok{(....)}
+
+\NormalTok{spatial }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{left\_join}\NormalTok{(temporal)}
 \end{Highlighting}
@@
 \CommentTok{\# example with a tsibble created from climate\_flat}
-\NormalTok{raw }\OtherTok{\textless{}{-}}\NormalTok{ climate\_flat }\SpecialCharTok{|\textgreater{}}\NormalTok{ tsibble}\SpecialCharTok{::}\FunctionTok{as\_tsibble}\NormalTok{(}\AttributeTok{key =}\NormalTok{ id, }\AttributeTok{index =}\NormalTok{ date)}
-\NormalTok{dt }\OtherTok{\textless{}{-}}\NormalTok{  raw }\SpecialCharTok{|\textgreater{}}\NormalTok{ cubble}\SpecialCharTok{::}\FunctionTok{as\_cubble}\NormalTok{(}\AttributeTok{coords =} \FunctionTok{c}\NormalTok{(long, lat))}
+\NormalTok{raw }\OtherTok{\textless{}{-}}\NormalTok{ climate\_flat }\SpecialCharTok{|}\ErrorTok{\textgreater{}}\NormalTok{ tsibble}\SpecialCharTok{::}\FunctionTok{as\_tsibble}\NormalTok{(}\AttributeTok{key =}\NormalTok{ id, }\AttributeTok{index =}\NormalTok{ date)}
+\NormalTok{dt }\OtherTok{\textless{}{-}}\NormalTok{  raw }\SpecialCharTok{|}\ErrorTok{\textgreater{}}\NormalTok{ cubble}\SpecialCharTok{::}\FunctionTok{as\_cubble}\NormalTok{(}\AttributeTok{coords =} \FunctionTok{c}\NormalTok{(long, lat))}
 \NormalTok{dt}
@@
 \CommentTok{\# add station{-}based features in the nested form.}
-\NormalTok{dt }\SpecialCharTok{|\textgreater{}} \FunctionTok{mutate}\NormalTok{(fabletools}\SpecialCharTok{::}\FunctionTok{features}\NormalTok{(ts, tmax, }\FunctionTok{list}\NormalTok{(}\AttributeTok{tmax\_mean =}\NormalTok{ mean, }\AttributeTok{tmax\_var =}\NormalTok{ var)))}
+\NormalTok{dt }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{mutate}\NormalTok{(fabletools}\SpecialCharTok{::}\FunctionTok{features}\NormalTok{(ts, tmax, }\FunctionTok{list}\NormalTok{(}\AttributeTok{tmax\_mean =}\NormalTok{ mean, }\AttributeTok{tmax\_var =}\NormalTok{ var)))}
 \end{Highlighting}
@@
 As a spatial data object, \code{sf} creates a simple feature geometry list-column (\code{sfc}) in the data frame to provide spatial operations on various geometry types (\code{POINT}, \code{LINESTRING}, \code{POLYGON}, \code{MULTIPOLYGON}, etc).
-An \code{sf} object can be supplied as the spatial table, as part of the raw data, along with a temporal table, to create a cubble.
-Once formed in a cubble, methods for the \code{sfc} class can be applied in the nested form.
-An example of this is shown in \ref{covid}, which also handles the case when the sites in the two tables don't match exactly.
-A spatial data object with \code{s2} vector can also be the input of the spatial table.
+When creating a cubble an \code{sf} object can be supplied as the spatial table.
+Once included in a cubble, methods for the \code{sfc} class can be applied in the nested form.
+An example of this is shown in \ref{covid}, which also handles the case when the site identifiers in the two tables do not match exactly.
+A spatial data object with \code{s2} vector can also be the input for the spatial table.
 
@@
 Attributes are usually associated with dimensions and variables in the NetCDF format data and a \href{http://cfconventions.org/}{metadata convention for climate and forecast} has been designed to standardise the format of the attributes.
-A few packages in R exists for manipulating NetCDF data and this includes a high-level R interface: \pkg{ncdf4} (\protect\hyperlink{ref-ncdf4}{Pierce 2019}), a low-level interface that calls C-interface: \pkg{RNetCDF} (\protect\hyperlink{ref-rnetcdf}{Michna and Woods 2021}, \protect\hyperlink{ref-michna2013rnetcdf}{2013}), and a tidyverse implementation: \pkg{tidync} (\protect\hyperlink{ref-tidync}{Sumner 2020}).
+A few packages in R exist for manipulating NetCDF data and these include a high-level R interface: \pkg{ncdf4} (\protect\hyperlink{ref-ncdf4}{Pierce 2019}), a low-level interface that calls a C-interface: \pkg{RNetCDF} (\protect\hyperlink{ref-rnetcdf}{Michna and Woods 2021}, \protect\hyperlink{ref-michna2013rnetcdf}{2013}), and a tidyverse implementation: \pkg{tidync} (\protect\hyperlink{ref-tidync}{Sumner 2020}).
 
@@
 
-The memory limit with NetCDF data in cubble depends on the number of longitude grid points x the number of latitude grid points x the number of time grid points x the number of variables.
-Cubble can handle slightly more than 300 x 300 (longitude x longitude) grid points for daily 3 variables in one year.
+The memory limit with NetCDF data in cubble depends on the number of longitude grid points \(\times\) the number of latitude grid points \(\times\) the number of time grid points \(\times\) the number of variables.
+Cubble can handle slightly more than 300 \(\times\) 300 (longitude \(\times\) longitude) grid points for three daily variables in one year.
 A 300 by 300 spatial grid can be a bounding box of {[}100, -80, 180, 0{]} at 0.25 degree resolution or a global bounding box {[}-180, -90, 180, -90{]} at 1 degree resolution.
@@
 \begin{Highlighting}[]
-\CommentTok{\# Assume my\_ncdf has bounding box of [{-}180, {-}90, 180, {-}90]}
+\CommentTok{\# Assume my\_ncdf has a bounding box of [{-}180, {-}90, 180, {-}90]}
 \CommentTok{\# at 0.25 degree resolution and subset it to have}
@@
 
-Spatial locations can have grouping structure either inherent in the data or formed by clustering.
+Spatial locations can have grouping structures either inherent in the data or obtained by clustering.
 Rather than analysing variables in the site level, summarised variables in the cluster level can give a crisper picture of local areas.
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.003}
+{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/ursulalaa/papers/paper-cubble/figures/diagram-keynotes/diagram-keynotes.003}
 
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/illu-matching} \caption{An illustration of temporal matching in cubble.
+\includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-cubble/figures/illu-matching} \caption{An illustration of temporal matching in cubble.
 Three highest peaks are identified in each series and intervals are constructed on series \code{A}.
-Two peaks in series \code{a} fall into the intervals and hence the two series are considered to have 2 matches.}\label{fig:illu-matching}
+Two peaks in series \code{a} fall into the intervals and hence the two series are considered to have two matches.}\label{fig:illu-matching}
 \end{figure}
@@
 \item
-  \code{temporal_n_highest}: the number of peak used - 3 in the example above
+  \code{temporal_n_highest}: the number of peaks used - 3 in the example above
 \item
@@
 \item
-  \code{temporal_min_match}: the minimum number of matched peak for a valid matched pair
+  \code{temporal_min_match}: the minimum number of matched peaks for a valid matched pair
 \end{itemize}
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.004}
+{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/ursulalaa/papers/paper-cubble/figures/diagram-keynotes/diagram-keynotes.004}
 
@@
 
-Spatio-temporal transformation is a useful technique to extract information from spatio-temporal data.
+Different types of transformations can be useful to extract new information from spatio-temporal data.
 Glyph maps (\protect\hyperlink{ref-Wickham2012-yr}{Wickham et al.
-2012}) transform the time coordinates into the space coordinates to plot the time series of different locations on the map.
-Calendar plots (\protect\hyperlink{ref-wang2020calendar}{Wang, Cook, and Hyndman 2020}) reconstructs time into a calendar-based grid to discover weekday and weekend pattern.
+2012}) transform the time coordinates into space coordinates to plot the time series of different locations on the map.
+Calendar plots (\protect\hyperlink{ref-wang2020calendar}{Wang, Cook, and Hyndman 2020}) reconstruct time into a calendar-based grid to discover weekday and weekend pattern.
 Projection, or linear combination, of variables summarises multivariate information into lower dimension to further digest.
@@
 The function constructs a data frame with calculated position (\code{gx}, \code{gy}.
-\code{gid}) of each point on the time series using linear algebra (Equation 1 and 2 in Wickham et al.
-(\protect\hyperlink{ref-Wickham2012-yr}{2012})).
+\code{gid}) of each point on the time series using linear algebra (Equation 1 and 2 in \protect\hyperlink{ref-Wickham2012-yr}{Wickham et al.} (\protect\hyperlink{ref-Wickham2012-yr}{2012})).
 The data can then be piped into \code{ggplot} to create the glyph map as:
@@
 \FunctionTok{library}\NormalTok{(ggplot2)}
-\NormalTok{gly }\OtherTok{\textless{}{-}} \FunctionTok{glyphs}\NormalTok{(data, }
-              \AttributeTok{x\_major =}\NormalTok{ ..., }\AttributeTok{x\_minor =}\NormalTok{ ..., }
+\NormalTok{gly }\OtherTok{\textless{}{-}} \FunctionTok{glyphs}\NormalTok{(data,}
+              \AttributeTok{x\_major =}\NormalTok{ ..., }\AttributeTok{x\_minor =}\NormalTok{ ...,}
               \AttributeTok{y\_major =}\NormalTok{ ..., }\AttributeTok{y\_minor =}\NormalTok{ ..., ...)}
@@
 \FunctionTok{ggplot}\NormalTok{(gly, }\FunctionTok{aes}\NormalTok{(gx, gy, }\AttributeTok{group =}\NormalTok{ gid)) }\SpecialCharTok{+}
-  \FunctionTok{geom\_path}\NormalTok{() }
-\end{Highlighting}
-\end{Shaded}
-
-A re-implementation of the glyph map as a ggproto, \code{GeomGlyph}, has been made in the \pkg{cubble} package and now the glyph map can be created with \code{geom_glyph()}:
+  \FunctionTok{geom\_path}\NormalTok{()}
+\end{Highlighting}
+\end{Shaded}
+
+A reimplementation of the glyph map as a ggproto, \code{GeomGlyph}, has been made in the \pkg{cubble} package and now the glyph map can be created with \code{geom_glyph()}:
 
@@
 \FunctionTok{ggplot}\NormalTok{(}\AttributeTok{data =}\NormalTok{ data) }\SpecialCharTok{+}
-  \FunctionTok{geom\_glyph}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x\_major =}\NormalTok{ ..., }\AttributeTok{x\_minor =}\NormalTok{ ..., }
+  \FunctionTok{geom\_glyph}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x\_major =}\NormalTok{ ..., }\AttributeTok{x\_minor =}\NormalTok{ ...,}
                  \AttributeTok{y\_major =}\NormalTok{ ..., }\AttributeTok{y\_minor =}\NormalTok{ ...))}
@@
 \begin{Highlighting}[]
-\NormalTok{covid }\SpecialCharTok{|\textgreater{}} \FunctionTok{head}\NormalTok{(}\DecValTok{5}\NormalTok{)}
+\NormalTok{covid }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{head}\NormalTok{(}\DecValTok{5}\NormalTok{)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{lga }\SpecialCharTok{|\textgreater{}} \FunctionTok{head}\NormalTok{(}\DecValTok{5}\NormalTok{)}
+\NormalTok{lga }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{head}\NormalTok{(}\DecValTok{5}\NormalTok{)}
 \end{Highlighting}
@@
 
-Cubble can be created with separate spatial and temporal table in a list, with other arguments (\code{key}, \code{index}, and \code{coords}) introduced in section \ref{create}.
-\code{as_cubble()} will automatically check the match of site in both tables emit a warning if a location has missing spatial or temporal information.
+Cubble can be created with separate spatial and temporal tables in a list, with other arguments (\code{key}, \code{index}, and \code{coords}) introduced in section \ref{create}.
+\code{as_cubble()} will automatically check the match of sites in both tables and emit a warning if a location has missing spatial or temporal information.
 
@@
 Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world.
-The dataset \code{weatherdata::historical_tmax} extracts the maximum temperature for 236 Australian stations from GHCN with starting from year 1969.
+The dataset \code{weatherdata::historical_tmax} extracts the maximum temperature for 236 Australian stations from GHCN, starting from year 1969.
 \code{weatherdata::historical_tmax} is already in a cubble, with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates.
@@
 \begin{Highlighting}[]
-\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ weatherdata}\SpecialCharTok{::}\NormalTok{historical\_tmax }\SpecialCharTok{|\textgreater{}}
+\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ weatherdata}\SpecialCharTok{::}\NormalTok{historical\_tmax }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
   \FunctionTok{filter}\NormalTok{(}\FunctionTok{between}\NormalTok{(stringr}\SpecialCharTok{::}\FunctionTok{str\_sub}\NormalTok{(id, }\DecValTok{7}\NormalTok{, }\DecValTok{8}\NormalTok{), }\DecValTok{46}\NormalTok{, }\DecValTok{90}\NormalTok{))}
@@
 \begin{Highlighting}[]
-\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|\textgreater{}} 
-  \FunctionTok{face\_temporal}\NormalTok{() }\SpecialCharTok{|\textgreater{}}
+\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|}\ErrorTok{\textgreater{}} 
+  \FunctionTok{face\_temporal}\NormalTok{() }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
   \FunctionTok{filter}\NormalTok{(lubridate}\SpecialCharTok{::}\FunctionTok{year}\NormalTok{(date) }\SpecialCharTok{\%in\%} \FunctionTok{c}\NormalTok{(}\DecValTok{1971}\SpecialCharTok{:}\DecValTok{1975}\NormalTok{, }\DecValTok{2016}\SpecialCharTok{:}\DecValTok{2020}\NormalTok{)) }
@@
 \begin{Highlighting}[]
-\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|\textgreater{}}
+\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
   \FunctionTok{group\_by}\NormalTok{(}\AttributeTok{month =}\NormalTok{ lubridate}\SpecialCharTok{::}\FunctionTok{month}\NormalTok{(date), }
          \AttributeTok{group =} \FunctionTok{as.factor}\NormalTok{(}\FunctionTok{ifelse}\NormalTok{(lubridate}\SpecialCharTok{::}\FunctionTok{year}\NormalTok{(date) }\SpecialCharTok{\textgreater{}} \DecValTok{2015}\NormalTok{, }
-                                  \StringTok{"2016 \textasciitilde{} 2020"}\NormalTok{, }\StringTok{"1971 \textasciitilde{} 1975"}\NormalTok{))) }\SpecialCharTok{|\textgreater{}}
+                                  \StringTok{"2016 \textasciitilde{} 2020"}\NormalTok{, }\StringTok{"1971 \textasciitilde{} 1975"}\NormalTok{))) }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
   \FunctionTok{summarise}\NormalTok{(}\AttributeTok{tmax =} \FunctionTok{mean}\NormalTok{(tmax, }\AttributeTok{na.rm =} \ConstantTok{TRUE}\NormalTok{))}
@@
 \begin{Highlighting}[]
-\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|\textgreater{}} \FunctionTok{face\_spatial}\NormalTok{() }\SpecialCharTok{|\textgreater{}} \FunctionTok{filter}\NormalTok{(}\FunctionTok{nrow}\NormalTok{(ts) }\SpecialCharTok{==} \DecValTok{24}\NormalTok{) }
+\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{face\_spatial}\NormalTok{() }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{filter}\NormalTok{(}\FunctionTok{nrow}\NormalTok{(ts) }\SpecialCharTok{==} \DecValTok{24}\NormalTok{) }
 \end{Highlighting}
@@
 Lastly, to create a glyph map, both the major (\code{longitude}, \code{latitude}) and minor (\code{month}, \code{tmax}) coordinates need to be in the same table.
-Spatial variables can be moved to the long form with \code{migrate()}:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|\textgreater{}} \FunctionTok{face\_temporal}\NormalTok{() }\SpecialCharTok{|\textgreater{}} \FunctionTok{unfold}\NormalTok{(latitude, longitude)}
+Spatial variables can be moved to the long form with \code{unfold()}:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{face\_temporal}\NormalTok{() }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{unfold}\NormalTok{(latitude, longitude)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{nsw\_vic }\OtherTok{\textless{}{-}}\NormalTok{ ozmaps}\SpecialCharTok{::}\NormalTok{abs\_ste }\SpecialCharTok{|\textgreater{}} 
+\NormalTok{nsw\_vic }\OtherTok{\textless{}{-}}\NormalTok{ ozmaps}\SpecialCharTok{::}\NormalTok{abs\_ste }\SpecialCharTok{|}\ErrorTok{\textgreater{}} 
   \FunctionTok{filter}\NormalTok{(NAME }\SpecialCharTok{\%in\%} \FunctionTok{c}\NormalTok{(}\StringTok{"Victoria"}\NormalTok{, }\StringTok{"New South Wales"}\NormalTok{))}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth,height=0.7\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/basic-manip} \caption{A glyph map of the average maximum temperature by month of Victoria and New South Wales weather stations in Australia.
+\includegraphics[width=1\linewidth,height=0.7\textheight]{/Users/ursulalaa/papers/paper-cubble/figures/basic-manip} \caption{A glyph map of the average maximum temperature by month of Victoria and New South Wales weather stations in Australia.
 On the top left corner is an insetted plot of station Cobar highlighted in the black box.}\label{fig:basic-manip}
@@
 \begin{Highlighting}[]
-\NormalTok{station\_nested }\OtherTok{\textless{}{-}}\NormalTok{ weatherdata}\SpecialCharTok{::}\NormalTok{climate\_full }\SpecialCharTok{|\textgreater{}} 
+\NormalTok{station\_nested }\OtherTok{\textless{}{-}}\NormalTok{ weatherdata}\SpecialCharTok{::}\NormalTok{climate\_full }\SpecialCharTok{|}\ErrorTok{\textgreater{}} 
   \FunctionTok{mutate}\NormalTok{(}\AttributeTok{cluster =}\NormalTok{ ...)}
@@
 \begin{Highlighting}[]
-\NormalTok{cluster\_nested }\OtherTok{\textless{}{-}}\NormalTok{ station\_nested }\SpecialCharTok{|\textgreater{}} \FunctionTok{switch\_key}\NormalTok{(cluster) }
+\NormalTok{cluster\_nested }\OtherTok{\textless{}{-}}\NormalTok{ station\_nested }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{switch\_key}\NormalTok{(cluster) }
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{cluster\_nested }\OtherTok{\textless{}{-}}\NormalTok{ cluster\_nested }\SpecialCharTok{|\textgreater{}} \FunctionTok{get\_centroid}\NormalTok{()}
-\end{Highlighting}
-\end{Shaded}
-
-Long form cubble at both levels can be accessed through stretching the nested form and with access to both station and cluster level cubbles, various plots can be made to understand the cluster.
+\NormalTok{cluster\_nested }\OtherTok{\textless{}{-}}\NormalTok{ cluster\_nested }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{get\_centroid}\NormalTok{()}
+\end{Highlighting}
+\end{Shaded}
+
+Long form cubbles at both levels can be accessed through stretching the nested form and with access to both station and cluster level cubbles, various plots can be made to understand the cluster.
 Figure \ref{fig:basic-agg} shows two example plots that can be made with this data.
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/basic-agg} \caption{Profile of aggregated precipitation from 639 weather stations in Australia.
+\includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-cubble/figures/basic-agg} \caption{Profile of aggregated precipitation from 639 weather stations in Australia.
 Subplot A shows the glyph map of the weekly averaged precipitation of each cluster.
-The group number of printed in the middle of y minor axis and can be used as a reference line to read the magnitude.
+The group number is printed in the middle of y minor axis and can be used as a reference line to read the magnitude.
 Subplot B shows the station membership of each cluster.}\label{fig:basic-agg}
@@
 Bureau of Meteorology collects \href{http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full}{water data} from river gauges and this includes variables: electrical conductivity, turbidity, watercourse discharge, watercourse level, and water temperature.
-In particular, water level will interact with precipitation from the climate data since rainfall will raise the water level in the river.
+In particular, we expect that water levels will interact with precipitation from the climate data since rainfall will raise the water level in the river.
 Figure \ref{fig:matching-map} gives the location of available weather stations and water gauges in Victoria.
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/matching-map} \caption{Location of weather stations and river gauges in Victoria, Australia.}\label{fig:matching-map}
+\includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-cubble/figures/matching-map} \caption{Location of weather stations and river gauges in Victoria, Australia.}\label{fig:matching-map}
 \end{figure}
 
-From the map, a few water gauges and weather stations are close to each other and the fluctuation of the water level could be matched up with precipitation measured by the climate station.
+From the map we can see that a few water gauges and weather stations are close to each other and the fluctuation of the water level could be matched up with precipitation measured by the climate station.
 As introduced in Section 3.2, \code{match_sites()} can be used to match one source of data with another source in a cubble.
@@
 
-Figure \ref{fig:matching} plots the matched pairs on the map or to view the matched series.
-There are four pairs of matches, all located in the middle of Victoria and the concurrent increase in precipitation and water level can be observed.
+Figure \ref{fig:matching} shows the matched pairs on the map and a view the matched time series.
+There are four pairs of matches, all located in the middle of Victoria, for which the concurrent increase in precipitation and water level can be observed.
 
 \begin{figure}
-\includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/matching} \caption{Matched weather stations and river gauges on the map (A) and across time (B).
+\includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-cubble/figures/matching} \caption{Matched weather stations and river gauges on the map (A) and across time (B).
 Precipitation and water level have been standardised between 0 and 1 to be displayed on the same scale.
@@
 
-Figure \ref{fig:netcdf} reproduces the ERA5 data row of Figure 19 in Hersbach et al.
-(\protect\hyperlink{ref-hersbach2020era5}{2020}).
+Figure \ref{fig:netcdf} reproduces the ERA5 data row of Figure 19 in \protect\hyperlink{ref-hersbach2020era5}{Hersbach et al.} (\protect\hyperlink{ref-hersbach2020era5}{2020}).
 It shows the southern polar vortex splits into two on 2002-09-26 and further splits into four on 2002-10-04 in the stratosphere.
-Readers interested in the analysis of this figure can refer to Hersbach et al.
-(\protect\hyperlink{ref-hersbach2020era5}{2020}), Simmons et al.
-(\protect\hyperlink{ref-simmons2020global}{2020}), and Simmons et al.
-(\protect\hyperlink{ref-simmons2005ecmwf}{2005}) for more details.
+Readers interested in the analysis of this figure can refer to \protect\hyperlink{ref-hersbach2020era5}{Hersbach et al.} (\protect\hyperlink{ref-hersbach2020era5}{2020}), \protect\hyperlink{ref-simmons2020global}{Simmons et al.} (\protect\hyperlink{ref-simmons2020global}{2020}), and \protect\hyperlink{ref-simmons2005ecmwf}{Simmons et al.} (\protect\hyperlink{ref-simmons2005ecmwf}{2005}) for more details.
 
 \begin{figure}
-\includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/netcdf} \caption{A reproduction of the second row (ERA5 data) of Figure 19 in  Hersbach et al (2020).}\label{fig:netcdf}
+\includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-cubble/figures/netcdf} \caption{A reproduction of the second row (ERA5 data) of Figure 19 in  Hersbach et al (2020).}\label{fig:netcdf}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/linking} \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/linking-north} \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/linking-lower} \caption{Exploring temperature variation using linking of a map and seasonal display.
+\includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/ursulalaa/papers/paper-cubble/figures/linking} \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/ursulalaa/papers/paper-cubble/figures/linking-north} \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/ursulalaa/papers/paper-cubble/figures/linking-lower} \caption{Exploring temperature variation using linking of a map and seasonal display.
 Each row is a screen dump of the process.
@@
 
-Figure \ref{fig:interactive-popup} shows Figure \ref{fig:interactive-linking} made made with leaflet and popups (\protect\hyperlink{ref-leafpop}{Appelhans and Detsch 2021}).
+Figure \ref{fig:interactive-popup} shows a reproduction of Figure \ref{fig:interactive-linking} made with leaflet and popups (\protect\hyperlink{ref-leafpop}{Appelhans and Detsch 2021}).
 
@@
 
-{\centering \includegraphics[width=0.45\linewidth,height=0.25\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/popup-mid}
+{\centering \includegraphics[width=0.45\linewidth,height=0.25\textheight]{/Users/ursulalaa/papers/paper-cubble/figures/popup-mid}
 
@@
 
-\caption{Same as Figure 11 with thetemperature variation shown as a popup in the leaflet map.}\label{fig:interactive-popup}
+\caption{Same as Figure 11 with the temperature variation shown as a popup in the leaflet map.}\label{fig:interactive-popup}
 \end{figure}
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.005}
+{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/ursulalaa/papers/paper-cubble/figures/diagram-keynotes/diagram-keynotes.005}
 
@@
 When a point on the time series is selected, the corresponding row in the long cubble will be activated.
-This will link to all the rows with the same id in the long cubble and the row in the nested cubble with the smae id (middle).
+This will link to all the rows with the same id in the long cubble and the row in the nested cubble with the same id (middle).
 Both plots will be updated with the full line selected and the point highlighted on the map (right).}\label{fig:illu-interactive-2}
@@
 \begin{CSLReferences}{1}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-leafpop}{}}%
-Appelhans, Tim, and Florian Detsch. 2021. \emph{\pkg{leafpop}: Include Tables, Images and Graphs in Leaflet Pop-Ups}. \url{https://CRAN.R-project.org/package=leafpop}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-bach_review_2014}{}}%
+\leavevmode\hypertarget{ref-leafpop}{}%
+Appelhans, Tim, and Florian Detsch. 2021. \emph{{leafpop}: Include Tables, Images and Graphs in Leaflet Pop-Ups}. \url{https://CRAN.R-project.org/package=leafpop}.
+
+\leavevmode\hypertarget{ref-bach_review_2014}{}%
 Bach, Benjamin, Pierre Dragicevic, Dragicevic Archambault, Christophe Hurter, and Sheelagh Carpendale.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-buja1988elements}{}}%
+\leavevmode\hypertarget{ref-buja1988elements}{}%
 Buja, Andreas, Daniel Asimov, and Catherine Hurley. 1988. {``Elements of a Viewing Pipeline.''} \emph{Dynamic Graphics Statistics}, 277.
 
-\leavevmode\vadjust pre{\hypertarget{ref-buja1996interactive}{}}%
+\leavevmode\hypertarget{ref-buja1996interactive}{}%
 Buja, Andreas, Dianne Cook, and Deborah F Swayne.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-cheng2016enabling}{}}%
+\leavevmode\hypertarget{ref-cheng2016enabling}{}%
 Cheng, Xiaoyue, Dianne Cook, and Heike Hofmann.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-cocchi2019data}{}}%
+\leavevmode\hypertarget{ref-cocchi2019data}{}%
 Cocchi, Marina. 2019. \emph{Data Fusion Methodology and Applications}. Elsevier.
 
-\leavevmode\vadjust pre{\hypertarget{ref-ctvspatiotemporal}{}}%
+\leavevmode\hypertarget{ref-ctvspatiotemporal}{}%
 Edzer Pebesma, Roger Bivand. 2022. {``{CRAN} Task View: Handling and Analyzing Spatio-Temporal Data.''} \url{https://CRAN.R-project.org/view=SpatioTemporal}.
 
-\leavevmode\vadjust pre{\hypertarget{ref-hersbach2020era5}{}}%
+\leavevmode\hypertarget{ref-hersbach2020era5}{}%
 Hersbach, Hans, Bill Bell, Paul Berrisford, Shoji Hirahara, András Horányi, Joaquín Muñoz-Sabater, Julien Nicolas, et al.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-ecwmfr}{}}%
-Hufkens, Koen, Reto Stauffer, and Elio Campitelli. 2019. {``The \pkg{ecwmfr} Package: An Interface to ECMWF API Endpoints.''} \url{https://doi.org/10.5281/zenodo.2647541}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-lu_multidimensional_2018}{}}%
+\leavevmode\hypertarget{ref-ecwmfr}{}%
+Hufkens, Koen, Reto Stauffer, and Elio Campitelli. 2019. {``The {ecwmfr} Package: An Interface to ECMWF API Endpoints.''} \url{https://doi.org/10.5281/zenodo.2647541}.
+
+\leavevmode\hypertarget{ref-lu_multidimensional_2018}{}%
 Lu, Meng, Marius Appel, and Edzer Pebesma.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-mcintosh2018using}{}}%
+\leavevmode\hypertarget{ref-mcintosh2018using}{}%
 McIntosh, Avery I, Helen E Jenkins, Laura F White, Marinus Barnard, Dana R Thomson, Tania Dolby, John Simpson, et al.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-michna2013rnetcdf}{}}%
-Michna, Pavel, and Milton Woods. 2013. {``\pkg{RNetCDF}: A Package for Reading and Writing NetCDF Datasets.''} \emph{The R Journal} 5 (2): 29--36.
-
-\leavevmode\vadjust pre{\hypertarget{ref-rnetcdf}{}}%
----------. 2021. \emph{\pkg{RNetCDF}: Interface to 'NetCDF' Datasets}. \url{https://CRAN.R-project.org/package=RNetCDF}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-spacetime}{}}%
-Pebesma, Edzer. 2012. {``\pkg{spacetime}: Spatio-Temporal Data in r.''} \emph{Journal of Statistical Software} 51 (7): 1--30. \url{https://doi.org/10.18637/jss.v051.i07}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-stars}{}}%
----------. 2021. \emph{\pkg{stars}: Spatiotemporal Arrays, Raster and Vector Data Cubes}. \url{https://CRAN.R-project.org/package=stars}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-sf}{}}%
-Pebesma, Edzer J. 2018. {``Simple Features for \proglang{R}: Standardized Support for Spatial Vector Data.''} \emph{R Journal} 10 (1): 439.
-
-\leavevmode\vadjust pre{\hypertarget{ref-sp}{}}%
-Pebesma, Edzer, and Roger S Bivand. 2005. {``S Classes and Methods for Spatial Data: The \pkg{sp} Package.''} \emph{R News} 5 (2): 9--13.
-
-\leavevmode\vadjust pre{\hypertarget{ref-ncdf4}{}}%
-Pierce, David. 2019. \emph{\pkg{ncdf4}: Interface to Unidata netCDF (Version 4 or Earlier) Format Data Files}. \url{https://CRAN.R-project.org/package=ncdf4}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-xts}{}}%
-Ryan, Jeffrey A., and Joshua M. Ulrich. 2020. \emph{\pkg{xts}: eXtensible Time Series}. \url{https://CRAN.R-project.org/package=xts}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-simmons2005ecmwf}{}}%
+\leavevmode\hypertarget{ref-michna2013rnetcdf}{}%
+Michna, Pavel, and Milton Woods. 2013. {``{RNetCDF}: A Package for Reading and Writing NetCDF Datasets.''} \emph{The R Journal} 5 (2): 29--36.
+
+\leavevmode\hypertarget{ref-rnetcdf}{}%
+---------. 2021. \emph{{RNetCDF}: Interface to 'NetCDF' Datasets}. \url{https://CRAN.R-project.org/package=RNetCDF}.
+
+\leavevmode\hypertarget{ref-spacetime}{}%
+Pebesma, Edzer. 2012. {``{spacetime}: Spatio-Temporal Data in r.''} \emph{Journal of Statistical Software} 51 (7): 1--30. \url{https://doi.org/10.18637/jss.v051.i07}.
+
+\leavevmode\hypertarget{ref-stars}{}%
+---------. 2021. \emph{{stars}: Spatiotemporal Arrays, Raster and Vector Data Cubes}. \url{https://CRAN.R-project.org/package=stars}.
+
+\leavevmode\hypertarget{ref-sf}{}%
+Pebesma, Edzer J. 2018. {``Simple Features for {R}: Standardized Support for Spatial Vector Data.''} \emph{R Journal} 10 (1): 439.
+
+\leavevmode\hypertarget{ref-sp}{}%
+Pebesma, Edzer, and Roger S Bivand. 2005. {``S Classes and Methods for Spatial Data: The {sp} Package.''} \emph{R News} 5 (2): 9--13.
+
+\leavevmode\hypertarget{ref-ncdf4}{}%
+Pierce, David. 2019. \emph{{ncdf4}: Interface to Unidata netCDF (Version 4 or Earlier) Format Data Files}. \url{https://CRAN.R-project.org/package=ncdf4}.
+
+\leavevmode\hypertarget{ref-xts}{}%
+Ryan, Jeffrey A., and Joshua M. Ulrich. 2020. \emph{{xts}: eXtensible Time Series}. \url{https://CRAN.R-project.org/package=xts}.
+
+\leavevmode\hypertarget{ref-simmons2005ecmwf}{}%
 Simmons, Adrian, Mariano Hortal, Graeme Kelly, Anthony McNally, Agathe Untch, and Sakari Uppala.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-simmons2020global}{}}%
+\leavevmode\hypertarget{ref-simmons2020global}{}%
 Simmons, Adrian, Cornel Soci, Julien Nicolas, Bill Bell, P.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-stuart2010matching}{}}%
+\leavevmode\hypertarget{ref-stuart2010matching}{}%
 Stuart, Elizabeth A. 2010. {``Matching Methods for Causal Inference: A Review and a Look Forward.''} \emph{Statistical Science} 25 (1): 1.
 
-\leavevmode\vadjust pre{\hypertarget{ref-tidync}{}}%
-Sumner, Michael. 2020. \emph{\pkg{tidync}: A Tidy Approach to 'NetCDF' Data Exploration and Extraction}. \url{https://CRAN.R-project.org/package=tidync}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-sutherland2000orca}{}}%
+\leavevmode\hypertarget{ref-tidync}{}%
+Sumner, Michael. 2020. \emph{{tidync}: A Tidy Approach to 'NetCDF' Data Exploration and Extraction}. \url{https://CRAN.R-project.org/package=tidync}.
+
+\leavevmode\hypertarget{ref-sutherland2000orca}{}%
 Sutherland, Peter, Anthony Rossini, Thomas Lumley, Nicholas Lewin-Koh, Julie Dickerson, Zach Cox, and Dianne Cook.
 2000.
-{``\pkg{Orca}: A Visualization Toolkit for High-Dimensional Data.''} \emph{Journal of Computational and Graphical Statistics} 9 (3): 509--29.
+{``{Orca}: A Visualization Toolkit for High-Dimensional Data.''} \emph{Journal of Computational and Graphical Statistics} 9 (3): 509--29.
 \url{https://www.tandfonline.com/doi/abs/10.1080/10618600.2000.10474896}.
 
-\leavevmode\vadjust pre{\hypertarget{ref-wang2020calendar}{}}%
+\leavevmode\hypertarget{ref-wang2020calendar}{}%
 Wang, Earo, Dianne Cook, and Rob J Hyndman. 2020. {``Calendar-Based Graphics for Visualizing People's Daily Schedules.''} \emph{Journal of Computational and Graphical Statistics} 29 (3): 490--502.
 
-\leavevmode\vadjust pre{\hypertarget{ref-tidydata}{}}%
+\leavevmode\hypertarget{ref-tidydata}{}%
 Wickham, Hadley. 2014. {``Tidy Data.''} \emph{Journal of Statistical Software} 59 (10): 1--23. \url{https://doi.org/10.18637/jss.v059.i10}.
 
-\leavevmode\vadjust pre{\hypertarget{ref-ggplot2}{}}%
+\leavevmode\hypertarget{ref-ggplot2}{}%
 ---------. 2016. \emph{Ggplot2: Elegant Graphics for Data Analysis}. Springer-Verlag New York. \url{https://ggplot2.tidyverse.org}.
 
-\leavevmode\vadjust pre{\hypertarget{ref-tidyverse}{}}%
+\leavevmode\hypertarget{ref-tidyverse}{}%
 Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy D'Agostino McGowan, Romain François, Garrett Grolemund, et al.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-dplyr}{}}%
+\leavevmode\hypertarget{ref-dplyr}{}%
 Wickham, Hadley, Romain François, Lionel Henry, and Kirill Müller. 2022. \emph{Dplyr: A Grammar of Data Manipulation}. \url{https://CRAN.R-project.org/package=dplyr}.
 
-\leavevmode\vadjust pre{\hypertarget{ref-Wickham2012-yr}{}}%
+\leavevmode\hypertarget{ref-Wickham2012-yr}{}%
 Wickham, Hadley, Heike Hofmann, Charlotte Wickham, and Dianne Cook. 2012. {``Glyph-Maps for Visually Exploring Temporal Patterns in Climate Data and Models.''} \emph{Environmetrics} 23 (5): 382--93.
 
-\leavevmode\vadjust pre{\hypertarget{ref-knitr}{}}%
-Xie, Yihui. 2015. \emph{Dynamic Documents with \proglang{R} and \pkg{knitr}}. 2nd ed. Boca Raton, Florida: Chapman; Hall/CRC. \url{https://yihui.name/knitr/}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-rmarkdown}{}}%
+\leavevmode\hypertarget{ref-knitr}{}%
+Xie, Yihui. 2015. \emph{Dynamic Documents with {R} and {knitr}}. 2nd ed. Boca Raton, Florida: Chapman; Hall/CRC. \url{https://yihui.name/knitr/}.
+
+\leavevmode\hypertarget{ref-rmarkdown}{}%
 Xie, Yihui, J. J. Allaire, and Garrett Grolemund. 2018. \emph{R Markdown: The Definitive Guide}. Boca Raton, Florida: Chapman; Hall/CRC. \url{https://bookdown.org/yihui/rmarkdown}.
 
-\leavevmode\vadjust pre{\hypertarget{ref-xie2014reactive}{}}%
+\leavevmode\hypertarget{ref-xie2014reactive}{}%
 Xie, Yihui, Heike Hofmann, and Xiaoyue Cheng. 2014. {``{Reactive Programming for Interactive Graphics}.''} \emph{Statistical Science} 29 (2): 201--13. \url{https://doi.org/10.1214/14-STS477}.

### 2022-04-26 4-4-netcdf.Rmd
polish wording
minor fixes
minor fixes sections 4.4 and 5

@@
 
-ERA5 data [@hersbach2020era5] is the latest reanalysis of global atmosphere, land surface, and ocean waves from 1950 onwards and is available in the NetCDF format from European Centre for Medium-Range Weather Forecasts (ECMWF).
-The data can be directly downloaded from [Copernicus Climate Data Store (CDS)](https://cds.climate.copernicus.eu/cdsapp#!/dataset/reanalysis-era5-pressure-levels?tab=overview) website or programmatically via an R package \pkg{ecmwfr} [@ecwmfr].
-The \code{era5-pressure} data contains variable *specific humidity* and *geopotential* on the 10 hPa pressure level on four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04.
+The ERA5 dataset [@hersbach2020era5] is the latest reanalysis of global atmosphere, land surface, and ocean waves from 1950 onwards and is available in the NetCDF format from the European Centre for Medium-Range Weather Forecasts (ECMWF).
+The data can be directly downloaded from [Copernicus Climate Data Store (CDS)](https://cds.climate.copernicus.eu/cdsapp#!/dataset/reanalysis-era5-pressure-levels?tab=overview) website or programmatically via the R package \pkg{ecmwfr} [@ecwmfr].
+The \code{era5-pressure} data contains the *specific humidity* and *geopotential* variables on the 10 hPa pressure level on four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04.
 Once downloaded, the data can be read into a cubble as:
@@
 It shows the southern polar vortex splits into two on 2002-09-26 and further splits into four on 2002-10-04 in the stratosphere.
-Readers interested in the analysis of this figure can refer to @hersbach2020era5, @simmons2020global and @simmons2005ecmwf for more details.
+Readers interested in the analysis of this figure can refer to @hersbach2020era5, @simmons2020global, and @simmons2005ecmwf for more details.
 

### 2022-04-26 5-conclusion.Rmd
Some rewriting and comments added
minor fixes
minor fixes sections 4.4 and 5

@@
 
-This work is funded by the Commonwealth Scientific and Industrial Research Organisation (CSIRO) Data61 Scholarship and started while Nicolas Langrené was affiliated with CSIRO's Data61.
+This work is funded by a Commonwealth Scientific and Industrial Research Organisation (CSIRO) Data61 Scholarship and started while Nicolas Langrené was affiliated with CSIRO's Data61.
 The article is created using \pkg{knitr} [@knitr] and \pkg{rmarkdown} [@rmarkdown] in R.
@@
 When a point on the time series is selected, the corresponding row in the long cubble will be activated.
-This will link to all the rows with the same id in the long cubble and the row in the nested cubble with the smae id (middle).
+This will link to all the rows with the same id in the long cubble and the row in the nested cubble with the same id (middle).
 Both plots will be updated with the full line selected and the point highlighted on the map (right)."}

### 2022-04-27 arxiv/cubble-arxiv.Rmd
revised abstract

@@
 abstract: |
-    Spatio-temporal data refer to measurements taken across space and time.
-In practice, spatio-temporal data can be decomposed into a spatial and temporal component: one can select a spatial location and inspect the temporal trend; or, one might select one or multiple time value(s) and explore the spatial distribution.
-Ideally, multiple maps and multiple time series could be made to explore these together.
-However, doing all of these actions is complicated when data arrive fragmented in multiple objects.
-To make it easy to do all these tasks, ideally, spatial and temporal variables should be in a single data object that we can slice and dice in different ways to conduct different visualisations.
-This work suggests a new data structure, \code{cubble}, that uses a nested form and a long form to organise spatial and temporal variables in a single data object.
-The new data structure ensures that data in the two forms are synchronised while being flexible to explore the two aspects of spatio-temporal data.
-It also provides the capability to handle data with hierarchical structure, match data from multiple sources, construct interactive graphics, and perform spatio-temporal transformation.
-The paper will demonstrate the use of \pkg{cubble} with examples using Australian climate weather stations, river level data, and climate reanalysis (ERA5) data.
+    Multivariate spatio-temporal data refers to multiple measurements taken across space and time.
+For many analyses, components can be separately studied: for example, to explore temporal trend of one variable for a single spatial location, or spatial distribution of one variable for one time might be modelled.
+However, at some level it is important to analyse all aspects, like temporal trend of multiple variables across all locations.
+To make it easy to do all these tasks, we introduce a new data structure, \code{cubble}, with a suite of functions enabling easy slicing and dicing on the different components, for visual and numerical explorations.
+A \code{cubble} ensures that all the components of the data are synchronised while providing flexibility for analysis, and also the capability to handle hierarchical spatial and temporal structure.
+The tools are illustrated with examples showing visualisation of Australian climate data, merging climate and river level data sources, reproducing a climate reanalysis (ERA5), and creating interactive graphics.
 keywords:

### 2022-04-27 4-1-basic.Rmd
fix merge conflict

@@
 
-Victoria government of Australia provides daily COVID case number by date, source, and local government area (LGA).
+The Victoria State Government in Australia provides daily COVID case numbers by date, source, and local government area (LGA).
 This data can be used to visualise COVID spread when combined with map information on LGA, available from the Australian Bureau of Statistics.
@@
 Cubble can be created with separate spatial and temporal tables in a list, with other arguments (\code{key}, \code{index}, and \code{coords}) introduced in section \@ref(create).
-\code{as_cubble()} will automatically check the match of sites in both tables and emit a warning if a location has missing spatial or temporal information.
+\code{as_cubble()} will automatically check the matching of sites in both tables and emit a warning if a location has missing spatial or temporal information.
 
@@
 
-Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world.
-The dataset \code{weatherdata::historical_tmax} extracts the maximum temperature for `r nrow(weatherdata::historical_tmax)` Australian stations from GHCN, starting from year `r range(weatherdata::historical_tmax$first_year)[2]`.
+The Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world.
+The dataset \code{weatherdata::historical_tmax} extracts the maximum temperature for `r nrow(weatherdata::historical_tmax)` Australian stations from the GHCN starting from year `r range(weatherdata::historical_tmax$first_year)[2]`.
 \code{weatherdata::historical_tmax} is already in a cubble, with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates.
-This example compares the maximum temperature in two periods: 1971 - 1975 and 2016 - 2020 for stations in Victoria and New South Wales.
+This example compares the maximum temperature in two periods: 1971-1975 and 2016-2020 for stations in Victoria and New South Wales.
 
-Stations in the two states can be subsetted by the station number: Australia GHCN station number starts with "ASN00" and is followed by the [Bureau of Meteorology (BOM) station number](http://www.bom.gov.au/climate/cdo/about/site-num.shtml).
-The second and third digits of BOM station number (7th and 8th in the GHCN number) define the state of the station, where 46 - 75 are New South Wales stations and 76 - 90 are Victoria stations.
-Filtering Victoria and New South Wales stations is an operation in the spatial dimension and hence uses the nested form:
+Stations in the two states can be subsetted by the station number: the Australian GHCN station numbers start with "ASN00" and are followed by the [Bureau of Meteorology (BOM) station number](http://www.bom.gov.au/climate/cdo/about/site-num.shtml).
+The second and third digits of the BOM station number (7th and 8th in the GHCN number) define the state of the station; the range 46-75 corresponds to New South Wales stations and the range 76-90 corresponds to Victorian stations.
+Filtering Victoria and New South Wales stations is a spatial operation and hence uses the nested form:
 
@@
 
-Filtering for the period 1971 ~ 1975 and 2016 ~ 2020 is an operation on the time dimension and the nested cubble needs to be switched to the long cubble by \code{stretch()}:
+Filtering for the time periods 1971-1975 and 2016-2020 is a time operation and the nested cubble needs to be switched to the long cubble form by \code{stretch()}:
 
@@
 
-A monthly average is used for both periods to smooth the maximum temperature and it is again an operation on the time dimension:
+A monthly average is used for both periods to smooth the maximum temperature, which is another time operation:
 
@@
 
-A few stations do not have records during 1971 - 1975 and further investigation shows that while the first and last year of each series is recorded, the missing years in this period are not reported.
+A few stations do not have records during the time period 1971-1975 and further investigation shows that while the first and last year of each series is recorded, the missing years in this period are not reported.
 These stations are filtered out by examining whether the summarised time series has 24 months.

### 2022-04-27 4-2-agg.Rmd
fix merge conflict

--- 2022-04-23 4-2-agg.Rmd
@@
 In the previous example, there has already been some overlapping of the glyphs for a few stations near (151E, 34S) and (152E, 33S), which is a problem when mapping more stations at the national level.
-Aggregation can be helpful in grouping series into clusters before visualising the cluster with a glyph map.
+Aggregation can be helpful in grouping series into clusters before visualising the clusters with a glyph map.
 This example shows how to organise data at both levels with \code{switch_key()}.
 
-\code{weatherdata::climate_full}, also extracted from GHCN, records daily precipitation and maximum/ minimum temperature for `r nrow(weatherdata::climate_full)` stations in Australia from 2016 to 2020.
-A simple kmean algorithm based on the distance matrix is used to create 20 clusters.
-The dataset \code{station_nested} is a nested cubble with a cluster column indicating the group each station belongs to.
-More complex algorithms can be used for other problems, as long as there is a mapping from each station to a cluster.
+\code{weatherdata::climate_full}, also extracted from the GHCN, records daily precipitation and maximum/minimum temperature for `r nrow(weatherdata::climate_full)` stations in Australia from 2016 to 2020.
+A simple $k$-means algorithm based on the distance matrix is used to create 20 clusters.
+The dataset \code{station_nested} is a nested cubble with a cluster column indicating the group to which each station belongs.
+More advanced clustering algorithms can be used for other applications, as long as there is a mapping from each station to a cluster.
 
@@
 
-To create a group level cubble, use \code{switch_key()} with the new key variable, \code{cluster}:
+To create a group-level cubble, use \code{switch_key()} with the new key variable, \code{cluster}:
 
@@
 
-With the group level cubble, \code{get_centroid()} is useful to compute the centroid of each cluster, which will be used as the major axis for the glyph map later:
+With the group-level cubble, \code{get_centroid()} is useful to compute the centroid of each cluster, which will be used as the major axis for the glyph map later:
 
@@
 
-Long form cubble at both levels can be accessed through stretching the nested form and with access to both station and cluster level cubbles, various plots can be made to understand the cluster.
+Long form cubble at both levels can be accessed through stretching the nested form and with access to both station and cluster-level cubbles, various plots can be made to understand the cluster.
 Figure \ref{fig:basic-agg} shows two example plots that can be made with this data.
@@
 Subplot A shows the glyph map of the weekly averaged precipitation of each cluster.
-The group number of printed in the middle of y minor axis and can be used as a reference line to read the magnitude.
+The group number is printed in the middle of the y minor axis and can be used as a reference line to read the magnitude.
 Subplot B shows the station membership of each cluster."}

### 2022-04-27 4-3-water-level.Rmd
fix merge conflict

--- 2022-04-23 4-3-water-level.Rmd
@@
 
-Bureau of Meteorology collects [water data](http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full) from river gauges and this includes variables: electrical conductivity, turbidity, watercourse discharge, watercourse level, and water temperature.
+The Bureau of Meteorology collects [water data](http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full) from river gauges.
+The collected variables include: electrical conductivity, turbidity, watercourse discharge, watercourse level, and water temperature.
 In particular, water level will interact with precipitation from the climate data since rainfall will raise the water level in the river.
@@
 
-From the map, a few water gauges and weather stations are close to each other and the fluctuation of the water level could be matched up with precipitation measured by the climate station.
-As introduced in Section 3.2, \code{match_sites()} can be used to match one source of data with another source in a cubble.
+From the map, one can see that a few water gauges and weather stations are close to each other and the fluctuation of the water level could be matched up with the precipitation measured by the nearby climate station.
+As introduced in Section \@ref(matching), \code{match_sites()} can be used to match one source of data with another source in a cubble.
 Here \code{Water_course_level} in \code{river} will be matched to \code{prcp} in \code{climate} in 2020.
@@
 \code{temporal_independent} controls the variable used to construct the interval and the goal here is to see if the water level in the river will reflect precipitation.
-This puts precipitation \code{prcp}, as the independent.
-Given there is one year's worth of data, the number of peaks (\code{temporal_n_highest}) to consider is slightly raised from a default of 20 to 30 and \code{temporal_min_match} is raised accordingly To return all the pairs of the match, \code{temporal_min_match} can be set to 0.
+This puts precipitation \code{prcp}, as the independent variable.
+Given there is one year's worth of data, the number of peaks (\code{temporal_n_highest}) to consider is slightly raised from a default of 20 to 30 and \code{temporal_min_match} is raised accordingly.
+To return all the pairs of the match, \code{temporal_min_match} can be set to 0.
 
@@
 
-Figure \ref{fig:matching} plots the matched pairs on the map or to view the matched series.
-There are four pairs of matches, all located in the middle of Victoria and the concurrent increase in precipitation and water level can be observed.
+Figure \ref{fig:matching} plots the matched pairs on the map.
+There are four pairs of matches, all located in the middle of Victoria and the expected concurrent increase in precipitation and water level can be observed.
 

### 2022-04-27 4-5-interactive.Rmd
fix merge conflict

@@
 
-The temperature range is calculated as the difference between \code{tmax} and \code{tmin} and its monthly average over 2016 - 2020 is taken before calculating the variance.
+The temperature range is calculated as the difference between \code{tmax} and \code{tmin} and its monthly average over 2016-2020 is taken before calculating the variance.
 A \code{SharedData} object is constructed for each form of the cubble and the same \code{group} argument ensures the cross-linking of the two forms via the common \code{id} column.
@@
 Selecting a location with high variance on the map produces the plot in the second row.
-The maximum nad minimum temperature is shown using a ribbon.
-The bottom row first selects the lowest temperature in Auguest in the seasonal display.
+The maximum and minimum temperature is shown using a ribbon.
+The bottom row first selects the lowest temperature in August in the seasonal display.
 A location in the Tasmania Island is then selected to compare the temperature variation with Thredbo AWS.", fig.show='hold'}
@@
 
-Figure \ref{fig:interactive-popup} shows a reproduction of Figure \ref{fig:interactive-linking} made  with leaflet and popups [@leafpop].
+Figure \ref{fig:interactive-popup} shows Figure \ref{fig:interactive-linking} made with leaflet and popups [@leafpop].
 

### 2022-04-27 arxiv/cubble-arxiv.tex
fix merge conflict

--- 2022-04-23 arxiv/cubble-arxiv.tex
@@
 \begin{abstract}
-Spatio-temporal data refer to measurements taken across space and time.
-In practice, spatio-temporal data can be decomposed into a spatial and temporal component: at one time, we would select a spatial location and inspect the temporal trend; at other time, we might select one or multiple time value(s) and explore the spatial distribution.
-Ideally, we could make multiple maps and multiple time series to explore these together.
-However, doing all of these actions is complicated when data arrive fragmented in multiple objects.
-To make it easy to do all these tasks, ideally, spatial and temporal variables are in a single data object that we can slice and dice in different ways to conduct different visualisations.
-This work suggests a new data structure, \code{cubble}, that uses a nested form and a long form to organise spatial and temporal variables in a single data object.
-The new data structure ensures that data in the two forms are synchronised while being flexible to explore the two aspects of spatio-temporal data.
-It also provides the capability to handle data with hierarchical structure, match data from multiple sources, construct interactive graphics, and perform spatio-temporal transformation.
-The paper will demonstrate \pkg{cubble} with examples using Australian climate weather stations, river level data, and climate reanalysis (ERA5) data.
+Multivariate spatio-temporal data refers to multiple measurements taken across space and time.
+For many analyses, components can be separately studied: for example, to explore temporal trend of one variable for a single spatial location, or spatial distribution of one variable for one time might be modelled.
+However, at some level it is important to analyse all aspects, like temporal trend of multiple variables across all locations.
+To make it easy to do all these tasks, we introduce a new data structure, \code{cubble}, with a suite of functions enabling easy slicing and dicing on the different components, for visual and numerical explorations.
+A \code{cubble} ensures that all the components of the data are synchronised while providing flexibility for analysis, and also the capability to handle hierarchical spatial and temporal structure.
+The tools are illustrated with examples showing visualisation of Australian climate data, merging climate and river level data sources, reproducing a climate reanalysis (ERA5), and creating interactive graphics.
 \end{abstract}
@@
 
-The SpatioTemproal CRAN task view (\protect\hyperlink{ref-ctvspatiotemporal}{Edzer Pebesma 2022}) has a section \emph{Representing data} that lists existing spatio-temporal data representation in \proglang{R}.
+The SpatioTemproal CRAN task view (\protect\hyperlink{ref-ctvspatiotemporal}{Edzer Pebesma 2022}) has a section \emph{Representing data} that lists existing spatio-temporal data representations in \proglang{R}.
 Among them, E.
@@
 These layouts are implemented in the \pkg{spacetime} (\protect\hyperlink{ref-spacetime}{E.
-Pebesma 2012}) package with the underlying spatial and temporal class: \pkg{sp} (\protect\hyperlink{ref-sp}{E.
+Pebesma 2012}) package with the underlying spatial and temporal classes \pkg{sp} (\protect\hyperlink{ref-sp}{E.
 Pebesma and Bivand 2005}) and \pkg{xts} (\protect\hyperlink{ref-xts}{Ryan and Ulrich 2020}).
@@
 Still, the data representation for spatio-temporal data can be further improved and there are two reasons for this.
-Firstly, the raw data sourced in the wild is less often presented in any one of the representations above, and fitting the raw data into a data object can sometimes be not easy.
-More often, spatio-temporal data are collected in separate 2D tables and analysts need to ensemble them into a whole piece before exploring the data.
-Examples of components of spatio-temporal data can be 1) map data records the shape of a collection of area of interest; 2) geo-scientific data records the longitude and latitude coordinates of locations in the area along with other meta data relates to the location, and; 3) temporal variables indexed by both location and time.
-
-The other reason is how tidy data (\protect\hyperlink{ref-tidydata}{Wickham 2014}) should be applied to spatio-temporal data.
-The tidy data concept advocates the data to be structured into 1) one row per observation, 2) one column per variable, and 3) one type of data per table.
+Firstly, the raw data sourced in the wild is less often presented in any one of the representations above, and fitting the raw data into a data object can sometimes be difficult.
+More often, spatio-temporal data are collected in separate 2D tables and analysts need to assemble them into a whole piece before exploring the data.
+Examples of components of spatio-temporal data can be 1) map data recording the shape of a collection of areas of interest; 2) geo-scientific data recording the longitude and latitude coordinates of locations in the area along with other metadata related to the location, and; 3) temporal variables indexed by both location and time.
+
+The other reason is how tidy data concepts (\protect\hyperlink{ref-tidydata}{Wickham 2014}) should be applied to spatio-temporal data.
+Tidy data should be structured into 1) one row per observation, 2) one column per variable, and 3) one type of data per table.
 The long form data is preferred over wide data given the downstream software i.e.~\pkg{dplyr} (\protect\hyperlink{ref-dplyr}{Wickham et al.
 2022}) and \pkg{ggplot2} (\protect\hyperlink{ref-ggplot2}{Wickham 2016}) for data wrangling and visualisation.
-However, the long form can be inefficient to store feature geometries, especially for large multipolygon over daily or sub-daily data over years, which are extensively collected and handled in time series analysis.
-This poses the question of how to arrange spatial and temporal variables that makes analysing spatio-temporal data easier.
-
-This paper presents a new \proglang{R} package, \pkg{cubble}, for working with a new spatio-temporal data representation, which addresses the two issues mentioned above.
-In the package, a new data structure, also called \code{cubble}, is proposed to organise spatial and temporal variables as two forms of a single data object so that they can be wrangled separately while synchronised as a whole piece.
+However, the long form can be inefficient to store feature geometries, especially for large multipolygons for daily or sub-daily data over years, which are extensively collected and handled in time series analysis.
+This poses the question of how to arrange spatial and temporal variables in a way that would make analysing spatio-temporal data easier.
+
+This paper presents a new \proglang{R} package, \pkg{cubble} which addresses the two issues mentioned above.
+In the package, a new data structure, also called \code{cubble}, is proposed to organise spatial and temporal variables as two forms of a single data object so that they can be wrangled separately while being kept synchronised.
 Among the four spacetime layouts in E.
@@
 
-The rest of the paper is divided as follows: Section \ref{cube} discusses the spatio-temporal cube to conceptualise spatio-temporal data.
+The rest of the paper is divided as follows: Section \ref{cube} introduces the proposed cube structure as a way to conceptualise multivariate spatio-temporal data.
 Section \ref{cubble} presents the main design and functionality of \pkg{cubble}.
 Section \ref{others} explains how cubble deals with more advanced considerations, including data with the hierarchical structure, data matching, how cubble fits with existing static and interactive visualisation tools, and spatio-temporal data transformation.
-Section \ref{examples} uses Australia weather station data and river level data as examples to demonstrate the use of \pkg{cubble}.
+Section \ref{examples} uses Australian weather station data and river level data as examples to demonstrate the use of \pkg{cubble}.
 An example of how \pkg{cubble} handles NetCDF data is also provided.
@@
 
-Spatio-temporal data can be conceptualised using a cubical data model with three axes.
+\textcolor{blue}{Ursula: I have rearranged this section somewhat, please check!}
+
+Spatio-temporal data can be conceptualised using a cubical data model with three axes, typically time, latitude and longitude.
 This abstraction can be useful for generalising operations and visualisation: Lu, Appel, and Pebesma (\protect\hyperlink{ref-lu_multidimensional_2018}{2018}) shows how array operations (select, scale, reduce, rearrange, and compute) can be mapped onto the cube; Bach et al.
@@
 
+The most common data cube using the three axes (time, latitude, longitude) could be considered as taking snapshots of the space and stacking on time.
+Here we will define the spatio-temporal cube slightly differently with the three axes (time, site, variable), as illustrated in Figure \ref{fig:cube-def}.
+The time axis is the same in both versions, while the site axis now captures both latitude and longitude.
+Finally, variables are stacked on this space-time canvas, with one observation per site and time point.
+This notion is adopted to avoid using hypercubes when describing multivariate spatio-temporal data.
+
 While a cubical data model is neat to view the spatio-temporal data conceptually, a 3D data representation, or an array, may not be convenient for data wrangling.
 There are two reasons for this: 1) While arrays are efficient for the computation on numerical values, spatio-temporal data contains more than just numerical variables.
-For example, wrangling on character strings and specific datetime classes is common for spatio-temporal data but may not be easy done with array objects.
+For example, wrangling on character strings and specific datetime classes is common for spatio-temporal data but may not be easily done with array objects.
 2) The \texttt{mutate()} syntax for creating new variables is no longer valid since the operation can no longer define how a created 1D vector should be added to the array.
-
-The most common data cube is defined using the three axes: time, latitude, and longitude, which can be seen as taking snapshots of the space and stacking on time.
-Our spatio-temporal cube is defined slightly differently with three axes: time, site, and variable as Figure \ref{fig:cube-def}.
-The time axis is the same across the two formations, the site encompasses latitude and longitude into a single column identifier, while variables are stacked on this defined space-time canvas.
-This notion is adopted to avoid using hypercube when describing multivariate spatio-temporal data.
+\textcolor{blue}{Ursula: this seems too specific, you should be able to make the same argument without referencing the mutate function?}
 
@@
 Three highest peaks are identified in each series and intervals are constructed on series \code{A}.
-Two peaks in series \code{a} fall into the intervals and hence the two series are considered to have 2 matches.}\label{fig:illu-matching}
+Two peaks in series \code{a} fall into the intervals and hence the two series are considered to have two matches.}\label{fig:illu-matching}
 \end{figure}
@@
 
-Cubble can be created with separate spatial and temporal table in a list, with other arguments (\code{key}, \code{index}, and \code{coords}) introduced in section \ref{create}.
-\code{as_cubble()} will automatically check the match of site in both tables emit a warning if a location has missing spatial or temporal information.
+Cubble can be created with separate spatial and temporal tables in a list, with other arguments (\code{key}, \code{index}, and \code{coords}) introduced in section \ref{create}.
+\code{as_cubble()} will automatically check the matching of sites in both tables and emit a warning if a location has missing spatial or temporal information.
 
@@
 
-Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world.
-The dataset \code{weatherdata::historical_tmax} extracts the maximum temperature for 236 Australian stations from GHCN with starting from year 1969.
+The Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world.
+The dataset \code{weatherdata::historical_tmax} extracts the maximum temperature for 236 Australian stations from the GHCN starting from year 1969.
 \code{weatherdata::historical_tmax} is already in a cubble, with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates.
-This example compares the maximum temperature in two periods: 1971 - 1975 and 2016 - 2020 for stations in Victoria and New South Wales.
-
-Stations in the two states can be subsetted by the station number: Australia GHCN station number starts with ``ASN00'' and is followed by the \href{http://www.bom.gov.au/climate/cdo/about/site-num.shtml}{Bureau of Meteorology (BOM) station number}.
-The second and third digits of BOM station number (7th and 8th in the GHCN number) define the state of the station, where 46 - 75 are New South Wales stations and 76 - 90 are Victoria stations.
-Filtering Victoria and New South Wales stations is an operation in the spatial dimension and hence uses the nested form:
+This example compares the maximum temperature in two periods: 1971-1975 and 2016-2020 for stations in Victoria and New South Wales.
+
+Stations in the two states can be subsetted by the station number: the Australian GHCN station numbers start with ``ASN00'' and are followed by the \href{http://www.bom.gov.au/climate/cdo/about/site-num.shtml}{Bureau of Meteorology (BOM) station number}.
+The second and third digits of the BOM station number (7th and 8th in the GHCN number) define the state of the station; the range 46-75 corresponds to New South Wales stations and the range 76-90 corresponds to Victorian stations.
+Filtering Victoria and New South Wales stations is a spatial operation and hence uses the nested form:
 
@@
 
-Filtering for the period 1971 \textasciitilde{} 1975 and 2016 \textasciitilde{} 2020 is an operation on the time dimension and the nested cubble needs to be switched to the long cubble by \code{stretch()}:
+Filtering for the time periods 1971-1975 and 2016-2020 is a time operation and the nested cubble needs to be switched to the long cubble form by \code{stretch()}:
 
@@
   \FunctionTok{face\_temporal}\NormalTok{() }\SpecialCharTok{|\textgreater{}}
-  \FunctionTok{filter}\NormalTok{(lubridate}\SpecialCharTok{::}\FunctionTok{year}\NormalTok{(date) }\SpecialCharTok{\%in\%} \FunctionTok{c}\NormalTok{(}\DecValTok{1971}\SpecialCharTok{:}\DecValTok{1975}\NormalTok{, }\DecValTok{2016}\SpecialCharTok{:}\DecValTok{2020}\NormalTok{)) }
-\end{Highlighting}
-\end{Shaded}
-
-A monthly average is used for both periods to smooth the maximum temperature and it is again an operation on the time dimension:
+  \FunctionTok{filter}\NormalTok{(lubridate}\SpecialCharTok{::}\FunctionTok{year}\NormalTok{(date) }\SpecialCharTok{\%in\%} \FunctionTok{c}\NormalTok{(}\DecValTok{1971}\SpecialCharTok{:}\DecValTok{1975}\NormalTok{, }\DecValTok{2016}\SpecialCharTok{:}\DecValTok{2020}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+A monthly average is used for both periods to smooth the maximum temperature, which is another time operation:
 
@@
 \NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|\textgreater{}}
-  \FunctionTok{group\_by}\NormalTok{(}\AttributeTok{month =}\NormalTok{ lubridate}\SpecialCharTok{::}\FunctionTok{month}\NormalTok{(date), }
-         \AttributeTok{group =} \FunctionTok{as.factor}\NormalTok{(}\FunctionTok{ifelse}\NormalTok{(lubridate}\SpecialCharTok{::}\FunctionTok{year}\NormalTok{(date) }\SpecialCharTok{\textgreater{}} \DecValTok{2015}\NormalTok{, }
+  \FunctionTok{group\_by}\NormalTok{(}\AttributeTok{month =}\NormalTok{ lubridate}\SpecialCharTok{::}\FunctionTok{month}\NormalTok{(date),}
+         \AttributeTok{group =} \FunctionTok{as.factor}\NormalTok{(}\FunctionTok{ifelse}\NormalTok{(lubridate}\SpecialCharTok{::}\FunctionTok{year}\NormalTok{(date) }\SpecialCharTok{\textgreater{}} \DecValTok{2015}\NormalTok{,}
                                   \StringTok{"2016 \textasciitilde{} 2020"}\NormalTok{, }\StringTok{"1971 \textasciitilde{} 1975"}\NormalTok{))) }\SpecialCharTok{|\textgreater{}}
@@
 \begin{Highlighting}[]
-\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|\textgreater{}} \FunctionTok{face\_spatial}\NormalTok{() }\SpecialCharTok{|\textgreater{}} \FunctionTok{filter}\NormalTok{(}\FunctionTok{nrow}\NormalTok{(ts) }\SpecialCharTok{==} \DecValTok{24}\NormalTok{) }
+\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|\textgreater{}} \FunctionTok{face\_spatial}\NormalTok{() }\SpecialCharTok{|\textgreater{}} \FunctionTok{filter}\NormalTok{(}\FunctionTok{nrow}\NormalTok{(ts) }\SpecialCharTok{==} \DecValTok{24}\NormalTok{)}
 \end{Highlighting}
@@
 \FunctionTok{ggplot}\NormalTok{() }\SpecialCharTok{+}
-  \FunctionTok{geom\_sf}\NormalTok{(}\AttributeTok{data =}\NormalTok{ nsw\_vic, }
+  \FunctionTok{geom\_sf}\NormalTok{(}\AttributeTok{data =}\NormalTok{ nsw\_vic,}
           \AttributeTok{fill =} \StringTok{"transparent"}\NormalTok{, }\AttributeTok{color =} \StringTok{"grey"}\NormalTok{, }\AttributeTok{linetype =} \StringTok{"dotted"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{geom\_glyph}\NormalTok{(}\AttributeTok{data =}\NormalTok{ tmax, }
-             \FunctionTok{aes}\NormalTok{(}\AttributeTok{x\_major =}\NormalTok{ longitude, }\AttributeTok{x\_minor =}\NormalTok{ month, }
+  \FunctionTok{geom\_glyph}\NormalTok{(}\AttributeTok{data =}\NormalTok{ tmax,}
+             \FunctionTok{aes}\NormalTok{(}\AttributeTok{x\_major =}\NormalTok{ longitude, }\AttributeTok{x\_minor =}\NormalTok{ month,}
                  \AttributeTok{y\_major =}\NormalTok{ latitude, }\AttributeTok{y\_minor =}\NormalTok{ tmax,}
@@
 In the previous example, there has already been some overlapping of the glyphs for a few stations near (151E, 34S) and (152E, 33S), which is a problem when mapping more stations at the national level.
-Aggregation can be helpful in grouping series into clusters before visualising the cluster with a glyph map.
+Aggregation can be helpful in grouping series into clusters before visualising the clusters with a glyph map.
 This example shows how to organise data at both levels with \code{switch_key()}.
 
-\code{weatherdata::climate_full}, also extracted from GHCN, records daily precipitation and maximum/ minimum temperature for 639 stations in Australia from 2016 to 2020.
-A simple kmean algorithm based on the distance matrix is used to create 20 clusters.
-The dataset \code{station_nested} is a nested cubble with a cluster column indicating the group each station belongs to.
-More complex algorithms can be used for other problems, as long as there is a mapping from each station to a cluster.
+\code{weatherdata::climate_full}, also extracted from the GHCN, records daily precipitation and maximum/minimum temperature for 639 stations in Australia from 2016 to 2020.
+A simple \(k\)-means algorithm based on the distance matrix is used to create 20 clusters.
+The dataset \code{station_nested} is a nested cubble with a cluster column indicating the group to which each station belongs.
+More advanced clustering algorithms can be used for other applications, as long as there is a mapping from each station to a cluster.
 
@@
 
-To create a group level cubble, use \code{switch_key()} with the new key variable, \code{cluster}:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{cluster\_nested }\OtherTok{\textless{}{-}}\NormalTok{ station\_nested }\SpecialCharTok{|\textgreater{}} \FunctionTok{switch\_key}\NormalTok{(cluster) }
-\end{Highlighting}
-\end{Shaded}
-
-With the group level cubble, \code{get_centroid()} is useful to compute the centroid of each cluster, which will be used as the major axis for the glyph map later:
+To create a group-level cubble, use \code{switch_key()} with the new key variable, \code{cluster}:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{cluster\_nested }\OtherTok{\textless{}{-}}\NormalTok{ station\_nested }\SpecialCharTok{|\textgreater{}} \FunctionTok{switch\_key}\NormalTok{(cluster)}
+\end{Highlighting}
+\end{Shaded}
+
+With the group-level cubble, \code{get_centroid()} is useful to compute the centroid of each cluster, which will be used as the major axis for the glyph map later:
 
@@
 Subplot A shows the glyph map of the weekly averaged precipitation of each cluster.
-The group number of printed in the middle of y minor axis and can be used as a reference line to read the magnitude.
+The group number is printed in the middle of the y minor axis and can be used as a reference line to read the magnitude.
 Subplot B shows the station membership of each cluster.}\label{fig:basic-agg}
@@
 
-Bureau of Meteorology collects \href{http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full}{water data} from river gauges and this includes variables: electrical conductivity, turbidity, watercourse discharge, watercourse level, and water temperature.
+The Bureau of Meteorology collects \href{http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full}{water data} from river gauges.
+The collected variables include: electrical conductivity, turbidity, watercourse discharge, watercourse level, and water temperature.
 In particular, water level will interact with precipitation from the climate data since rainfall will raise the water level in the river.
@@
 
-From the map, a few water gauges and weather stations are close to each other and the fluctuation of the water level could be matched up with precipitation measured by the climate station.
+From the map, one can see that a few water gauges and weather stations are close to each other and the fluctuation of the water level could be matched up with the precipitation measured by the nearby climate station.
 As introduced in Section 3.2, \code{match_sites()} can be used to match one source of data with another source in a cubble.
@@
 
-ERA5 data (\protect\hyperlink{ref-hersbach2020era5}{Hersbach et al.
-2020}) is the latest reanalysis of global atmosphere, land surface, and ocean waves from 1950 onwards and is available in the NetCDF format from European Centre for Medium-Range Weather Forecasts (ECMWF).
-The data can be directly downloaded from \href{https://cds.climate.copernicus.eu/cdsapp\#!/dataset/reanalysis-era5-pressure-levels?tab=overview}{Copernicus Climate Data Store (CDS)} website or programmatically via an R package \pkg{ecmwfr} (\protect\hyperlink{ref-ecwmfr}{Hufkens, Stauffer, and Campitelli 2019}).
-The \code{era5-pressure} data contains variable \emph{specific humidity} and \emph{geopotential} on the 10 hPa pressure level on four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04.
+The ERA5 dataset (\protect\hyperlink{ref-hersbach2020era5}{Hersbach et al.
+2020}) is the latest reanalysis of global atmosphere, land surface, and ocean waves from 1950 onwards and is available in the NetCDF format from the European Centre for Medium-Range Weather Forecasts (ECMWF).
+The data can be directly downloaded from \href{https://cds.climate.copernicus.eu/cdsapp\#!/dataset/reanalysis-era5-pressure-levels?tab=overview}{Copernicus Climate Data Store (CDS)} website or programmatically via the R package \pkg{ecmwfr} (\protect\hyperlink{ref-ecwmfr}{Hufkens, Stauffer, and Campitelli 2019}).
+The \code{era5-pressure} data contains the \emph{specific humidity} and \emph{geopotential} variables on the 10 hPa pressure level on four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04.
 Once downloaded, the data can be read into a cubble as:
@@
 Selecting a location with high variance on the map produces the plot in the second row.
-The maximum nad minimum temperature is shown using a ribbon.
-The bottom row first selects the lowest temperature in Auguest in the seasonal display.
+The maximum and minimum temperature is shown using a ribbon.
+The bottom row first selects the lowest temperature in August in the seasonal display.
 A location in the Tasmania Island is then selected to compare the temperature variation with Thredbo AWS.}\label{fig:interactive-linking}
@@
 
-Figure \ref{fig:interactive-popup} shows Figure \ref{fig:interactive-linking} made made with leaflet and popups (\protect\hyperlink{ref-leafpop}{Appelhans and Detsch 2021}).
+Figure \ref{fig:interactive-popup} shows Figure \ref{fig:interactive-linking} made with leaflet and popups (\protect\hyperlink{ref-leafpop}{Appelhans and Detsch 2021}).
 
@@
 
-This work is funded by the Commonwealth Scientific and Industrial Research Organisation (CSIRO) Data61 Scholarship and started while Nicolas Langrené was affiliated with CSIRO's Data61.
+This work is funded by a Commonwealth Scientific and Industrial Research Organisation (CSIRO) Data61 Scholarship and started while Nicolas Langrené was affiliated with CSIRO's Data61.
 The article is created using \pkg{knitr} (\protect\hyperlink{ref-knitr}{Xie 2015}) and \pkg{rmarkdown} (\protect\hyperlink{ref-rmarkdown}{Xie, Allaire, and Grolemund 2018}) in R.

### 2022-04-27 arxiv/cubble-arxiv.Rmd
Update abstract

@@
     Multivariate spatio-temporal data refers to multiple measurements taken across space and time.
-For many analyses, components can be separately studied: for example, to explore temporal trend of one variable for a single spatial location, or spatial distribution of one variable for one time might be modelled.
-However, at some level it is important to analyse all aspects, like temporal trend of multiple variables across all locations.
-To make it easy to do all these tasks, we introduce a new data structure, \code{cubble}, with a suite of functions enabling easy slicing and dicing on the different components, for visual and numerical explorations.
-A \code{cubble} ensures that all the components of the data are synchronised while providing flexibility for analysis, and also the capability to handle hierarchical spatial and temporal structure.
-The tools are illustrated with examples showing visualisation of Australian climate data, merging climate and river level data sources, reproducing a climate reanalysis (ERA5), and creating interactive graphics.
+For many analyses, spatial and time components can be separately studied: for example, to explore the temporal trend of one variable for a single spatial location, or to model the spatial distribution of one variable at a given time.
+However for some studies, it is important to analyse different aspects of the spatio-temporal data simultaneouly, like for instance instance, temporal trends of multiple variables across locations.
+In order to facilitate the study of different sections or combinations of spatio-temporal data, we introduce a new data structure, \code{cubble}, with a suite of functions enabling easy slicing and dicing on the different components spatio-temporal components, for visual and numerical explorations while facilitating data wrangling for modelling.
+The propose \code{cubble} structure ensures that all the components of the data are easy to access and manipulate while providing flexibility for data analysis.
+The \code{cubble} structure and the tools provided in the \code{cubble} R package equip users with the  capability to handle hierarchical spatial and temporal structures.
+\color{green}{The tools are illustrated with examples of Australian climate data, merging climate and river level data sources, reproducing a climate reanalysis (ERA5), and creating interactive graphics.}
 keywords:

### 2022-04-27 arxiv/cubble-arxiv.tex
Update abstract

--- 2022-04-26 arxiv/cubble-arxiv.tex
@@
 \begin{abstract}
-Spatio-temporal data refer to measurements taken across space and time.
-In practice, spatio-temporal data can be decomposed into a spatial and temporal component: one can select a spatial location and inspect the temporal trend; or, one might select one or multiple time value(s) and explore the spatial distribution.
-Ideally, multiple maps and multiple time series could be made to explore these together.
-However, doing all of these actions is complicated when data arrive fragmented in multiple objects.
-To make it easy to do all these tasks, ideally, spatial and temporal variables should be in a single data object that we can slice and dice in different ways to conduct different visualisations.
-This work suggests a new data structure, \code{cubble}, that uses a nested form and a long form to organise spatial and temporal variables in a single data object.
-The new data structure ensures that data in the two forms are synchronised while being flexible to explore the two aspects of spatio-temporal data.
-It also provides the capability to handle data with hierarchical structure, match data from multiple sources, construct interactive graphics, and perform spatio-temporal transformation.
-The paper will demonstrate the use of \pkg{cubble} with examples using Australian climate weather stations, river level data, and climate reanalysis (ERA5) data.
+Multivariate spatio-temporal data refers to multiple measurements taken across space and time.
+For many analyses, spatial and time components can be separately studied: for example, to explore the temporal trend of one variable for a single spatial location, or to model the spatial distribution of one variable at a given time.
+However for some studies, it is important to analyse different aspects of the spatio-temporal data simultaneouly, like for instance instance, temporal trends of multiple variables across locations.
+In order to facilitate the study of different sections or combinations of spatio-temporal data, we introduce a new data structure, \code{cubble}, with a suite of functions enabling easy slicing and dicing on the different components spatio-temporal components, for visual and numerical explorations while facilitating data wrangling for modelling.
+The propose \code{cubble} structure ensures that all the components of the data are easy to access and manipulate while providing flexibility for data analysis.
+The \code{cubble} structure and the tools provided in the \code{cubble} R package equip users with the capability to handle hierarchical spatial and temporal structures.
+\color{green}{The tools are illustrated with examples of Australian climate data, merging climate and river level data sources, reproducing a climate reanalysis (ERA5), and creating interactive graphics.}
 \end{abstract}
@@
 
-{\centering \includegraphics[width=0.5\linewidth]{/Users/ursulalaa/papers/paper-cubble/figures/cube-def}
+{\centering \includegraphics[width=0.5\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/cube-def}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-cubble/figures/diagram-keynotes/diagram-keynotes.008}
+{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/diagram-keynotes/diagram-keynotes.008}
 
@@
 \begin{verbatim}
-## [1] TRUE
+## [1] FALSE
 \end{verbatim}
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/ursulalaa/papers/paper-cubble/figures/diagram-keynotes/diagram-keynotes.003}
+{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/diagram-keynotes/diagram-keynotes.003}
 
@@
 
-\hypertarget{data-fusion-and-matching}{%
-\subsection{Data fusion and matching}\label{data-fusion-and-matching}}
+\hypertarget{matching}{%
+\subsection{Data fusion and matching}\label{matching}}
 
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-cubble/figures/illu-matching} \caption{An illustration of temporal matching in cubble.
+\includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/illu-matching} \caption{An illustration of temporal matching in cubble.
 Three highest peaks are identified in each series and intervals are constructed on series \code{A}.
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/ursulalaa/papers/paper-cubble/figures/diagram-keynotes/diagram-keynotes.004}
+{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/diagram-keynotes/diagram-keynotes.004}
 
@@
 Cubble can be created with separate spatial and temporal tables in a list, with other arguments (\code{key}, \code{index}, and \code{coords}) introduced in section \ref{create}.
-\code{as_cubble()} will automatically check the match of sites in both tables and emit a warning if a location has missing spatial or temporal information.
+\code{as_cubble()} will automatically check the matching of sites in both tables and emit a warning if a location has missing spatial or temporal information.
 
@@
 
-Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world.
-The dataset \code{weatherdata::historical_tmax} extracts the maximum temperature for 236 Australian stations from GHCN, starting from year 1969.
+The Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world.
+The dataset \code{weatherdata::historical_tmax} extracts the maximum temperature for 236 Australian stations from the GHCN starting from year 1969.
 \code{weatherdata::historical_tmax} is already in a cubble, with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates.
-This example compares the maximum temperature in two periods: 1971 - 1975 and 2016 - 2020 for stations in Victoria and New South Wales.
-
-Stations in the two states can be subsetted by the station number: Australia GHCN station number starts with ``ASN00'' and is followed by the \href{http://www.bom.gov.au/climate/cdo/about/site-num.shtml}{Bureau of Meteorology (BOM) station number}.
-The second and third digits of BOM station number (7th and 8th in the GHCN number) define the state of the station, where 46 - 75 are New South Wales stations and 76 - 90 are Victoria stations.
-Filtering Victoria and New South Wales stations is an operation in the spatial dimension and hence uses the nested form:
+This example compares the maximum temperature in two periods: 1971-1975 and 2016-2020 for stations in Victoria and New South Wales.
+
+Stations in the two states can be subsetted by the station number: the Australian GHCN station numbers start with ``ASN00'' and are followed by the \href{http://www.bom.gov.au/climate/cdo/about/site-num.shtml}{Bureau of Meteorology (BOM) station number}.
+The second and third digits of the BOM station number (7th and 8th in the GHCN number) define the state of the station; the range 46-75 corresponds to New South Wales stations and the range 76-90 corresponds to Victorian stations.
+Filtering Victoria and New South Wales stations is a spatial operation and hence uses the nested form:
 
@@
   \FunctionTok{face\_temporal}\NormalTok{() }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
-  \FunctionTok{filter}\NormalTok{(lubridate}\SpecialCharTok{::}\FunctionTok{year}\NormalTok{(date) }\SpecialCharTok{\%in\%} \FunctionTok{c}\NormalTok{(}\DecValTok{1971}\SpecialCharTok{:}\DecValTok{1975}\NormalTok{, }\DecValTok{2016}\SpecialCharTok{:}\DecValTok{2020}\NormalTok{)) }
-\end{Highlighting}
-\end{Shaded}
-
-A monthly average is used for both periods to smooth the maximum temperature and it is again an operation on the time dimension:
+  \FunctionTok{filter}\NormalTok{(lubridate}\SpecialCharTok{::}\FunctionTok{year}\NormalTok{(date) }\SpecialCharTok{\%in\%} \FunctionTok{c}\NormalTok{(}\DecValTok{1971}\SpecialCharTok{:}\DecValTok{1975}\NormalTok{, }\DecValTok{2016}\SpecialCharTok{:}\DecValTok{2020}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+A monthly average is used for both periods to smooth the maximum temperature, which is another time operation:
 
@@
 \NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
-  \FunctionTok{group\_by}\NormalTok{(}\AttributeTok{month =}\NormalTok{ lubridate}\SpecialCharTok{::}\FunctionTok{month}\NormalTok{(date), }
-         \AttributeTok{group =} \FunctionTok{as.factor}\NormalTok{(}\FunctionTok{ifelse}\NormalTok{(lubridate}\SpecialCharTok{::}\FunctionTok{year}\NormalTok{(date) }\SpecialCharTok{\textgreater{}} \DecValTok{2015}\NormalTok{, }
+  \FunctionTok{group\_by}\NormalTok{(}\AttributeTok{month =}\NormalTok{ lubridate}\SpecialCharTok{::}\FunctionTok{month}\NormalTok{(date),}
+         \AttributeTok{group =} \FunctionTok{as.factor}\NormalTok{(}\FunctionTok{ifelse}\NormalTok{(lubridate}\SpecialCharTok{::}\FunctionTok{year}\NormalTok{(date) }\SpecialCharTok{\textgreater{}} \DecValTok{2015}\NormalTok{,}
                                   \StringTok{"2016 \textasciitilde{} 2020"}\NormalTok{, }\StringTok{"1971 \textasciitilde{} 1975"}\NormalTok{))) }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
@@
 \begin{Highlighting}[]
-\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{face\_spatial}\NormalTok{() }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{filter}\NormalTok{(}\FunctionTok{nrow}\NormalTok{(ts) }\SpecialCharTok{==} \DecValTok{24}\NormalTok{) }
+\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{face\_spatial}\NormalTok{() }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{filter}\NormalTok{(}\FunctionTok{nrow}\NormalTok{(ts) }\SpecialCharTok{==} \DecValTok{24}\NormalTok{)}
 \end{Highlighting}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth,height=0.7\textheight]{/Users/ursulalaa/papers/paper-cubble/figures/basic-manip} \caption{A glyph map of the average maximum temperature by month of Victoria and New South Wales weather stations in Australia.
+\includegraphics[width=1\linewidth,height=0.7\textheight]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/basic-manip} \caption{A glyph map of the average maximum temperature by month of Victoria and New South Wales weather stations in Australia.
 On the top left corner is an insetted plot of station Cobar highlighted in the black box.}\label{fig:basic-manip}
@@
 
-In the previous example, there has been some overlap of glyphs when stations are near each other, (151E, 34S) and (152E, 33S), which is a problem when mapping more stations at the national level.
-Aggregation can be helpful in grouping series into clusters before visualising the cluster with a glyph map.
+In the previous example, there has already been some overlapping of the glyphs for a few stations near (151E, 34S) and (152E, 33S), which is a problem when mapping more stations at the national level.
+Aggregation can be helpful in grouping series into clusters before visualising the clusters with a glyph map.
 This example shows how to organise data at both levels with \code{switch_key()}.
 
-\code{weatherdata::climate_full}, also extracted from GHCN, records daily precipitation and maximum/ minimum temperature for 639 stations in Australia from 2016 to 2020.
-A simple kmean algorithm based on the distance matrix is used to create 20 clusters.
-The dataset \code{station_nested} is a nested cubble with a cluster column indicating the group each station belongs to.
-More complex algorithms can be used for other problems, as long as there is a mapping from each station to a cluster.
+\code{weatherdata::climate_full}, also extracted from the GHCN, records daily precipitation and maximum/minimum temperature for 639 stations in Australia from 2016 to 2020.
+A simple \(k\)-means algorithm based on the distance matrix is used to create 20 clusters.
+The dataset \code{station_nested} is a nested cubble with a cluster column indicating the group to which each station belongs.
+More advanced clustering algorithms can be used for other applications, as long as there is a mapping from each station to a cluster.
 
@@
 
-To create a group level cubble, use \code{switch_key()} with the new key variable, \code{cluster}:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{cluster\_nested }\OtherTok{\textless{}{-}}\NormalTok{ station\_nested }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{switch\_key}\NormalTok{(cluster) }
-\end{Highlighting}
-\end{Shaded}
-
-With the group level cubble, \code{get_centroid()} is useful to compute the centroid of each cluster, which will be used as the major axis for the glyph map later:
+To create a group-level cubble, use \code{switch_key()} with the new key variable, \code{cluster}:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{cluster\_nested }\OtherTok{\textless{}{-}}\NormalTok{ station\_nested }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{switch\_key}\NormalTok{(cluster)}
+\end{Highlighting}
+\end{Shaded}
+
+With the group-level cubble, \code{get_centroid()} is useful to compute the centroid of each cluster, which will be used as the major axis for the glyph map later:
 
@@
 
-Long form cubbles at both levels can be accessed through stretching the nested form and with access to both station and cluster level cubbles, various plots can be made to understand the cluster.
+Long form cubble at both levels can be accessed through stretching the nested form and with access to both station and cluster-level cubbles, various plots can be made to understand the cluster.
 Figure \ref{fig:basic-agg} shows two example plots that can be made with this data.
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-cubble/figures/basic-agg} \caption{Profile of aggregated precipitation from 639 weather stations in Australia.
+\includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/basic-agg} \caption{Profile of aggregated precipitation from 639 weather stations in Australia.
 Subplot A shows the glyph map of the weekly averaged precipitation of each cluster.
-The group number is printed in the middle of y minor axis and can be used as a reference line to read the magnitude.
+The group number is printed in the middle of the y minor axis and can be used as a reference line to read the magnitude.
 Subplot B shows the station membership of each cluster.}\label{fig:basic-agg}
@@
 
-Bureau of Meteorology collects \href{http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full}{water data} from river gauges and this includes variables: electrical conductivity, turbidity, watercourse discharge, watercourse level, and water temperature.
-In particular, we expect that water levels will interact with precipitation from the climate data since rainfall will raise the water level in the river.
+The Bureau of Meteorology collects \href{http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full}{water data} from river gauges.
+The collected variables include: electrical conductivity, turbidity, watercourse discharge, watercourse level, and water temperature.
+In particular, water level will interact with precipitation from the climate data since rainfall will raise the water level in the river.
 Figure \ref{fig:matching-map} gives the location of available weather stations and water gauges in Victoria.
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-cubble/figures/matching-map} \caption{Location of weather stations and river gauges in Victoria, Australia.}\label{fig:matching-map}
+\includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/matching-map} \caption{Location of weather stations and river gauges in Victoria, Australia.}\label{fig:matching-map}
 \end{figure}
 
-From the map we can see that a few water gauges and weather stations are close to each other and the fluctuation of the water level could be matched up with precipitation measured by the climate station.
-As introduced in Section 3.2, \code{match_sites()} can be used to match one source of data with another source in a cubble.
+From the map, one can see that a few water gauges and weather stations are close to each other and the fluctuation of the water level could be matched up with the precipitation measured by the nearby climate station.
+As introduced in Section \ref{matching}, \code{match_sites()} can be used to match one source of data with another source in a cubble.
 Here \code{Water_course_level} in \code{river} will be matched to \code{prcp} in \code{climate} in 2020.
@@
 This puts precipitation \code{prcp}, as the independent variable.
-Given there is one year's worth of data, the number of peaks (\code{temporal_n_highest}) to consider is slightly raised from a default of 20 to 30 and \code{temporal_min_match} is raised accordingly To return all the pairs of the match, \code{temporal_min_match} can be set to 0.
+Given there is one year's worth of data, the number of peaks (\code{temporal_n_highest}) to consider is slightly raised from a default of 20 to 30 and \code{temporal_min_match} is raised accordingly.
+To return all the pairs of the match, \code{temporal_min_match} can be set to 0.
 
@@
 
-Figure \ref{fig:matching} shows the matched pairs on the map and a view the matched time series.
-There are four pairs of matches, all located in the middle of Victoria, for which the concurrent increase in precipitation and water level can be observed.
+Figure \ref{fig:matching} plots the matched pairs on the map.
+There are four pairs of matches, all located in the middle of Victoria and the expected concurrent increase in precipitation and water level can be observed.
 
 \begin{figure}
-\includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-cubble/figures/matching} \caption{Matched weather stations and river gauges on the map (A) and across time (B).
+\includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/matching} \caption{Matched weather stations and river gauges on the map (A) and across time (B).
 Precipitation and water level have been standardised between 0 and 1 to be displayed on the same scale.
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-cubble/figures/netcdf} \caption{A reproduction of the second row (ERA5 data) of Figure 19 in  Hersbach et al (2020).}\label{fig:netcdf}
+\includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/netcdf} \caption{A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020).}\label{fig:netcdf}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/ursulalaa/papers/paper-cubble/figures/linking} \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/ursulalaa/papers/paper-cubble/figures/linking-north} \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/ursulalaa/papers/paper-cubble/figures/linking-lower} \caption{Exploring temperature variation using linking of a map and seasonal display.
+\includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/linking} \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/linking-north} \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/linking-lower} \caption{Exploring temperature variation using linking of a map and seasonal display.
 Each row is a screen dump of the process.
@@
 
-Figure \ref{fig:interactive-popup} shows a reproduction of Figure \ref{fig:interactive-linking} made with leaflet and popups (\protect\hyperlink{ref-leafpop}{Appelhans and Detsch 2021}).
+Figure \ref{fig:interactive-popup} shows Figure \ref{fig:interactive-linking} made with leaflet and popups (\protect\hyperlink{ref-leafpop}{Appelhans and Detsch 2021}).
 
@@
 
-{\centering \includegraphics[width=0.45\linewidth,height=0.25\textheight]{/Users/ursulalaa/papers/paper-cubble/figures/popup-mid}
+{\centering \includegraphics[width=0.45\linewidth,height=0.25\textheight]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/popup-mid}
 
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/ursulalaa/papers/paper-cubble/figures/diagram-keynotes/diagram-keynotes.005}
+{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/diagram-keynotes/diagram-keynotes.005}
 

### 2022-04-27 1-intro.Rmd
Introduction sections changes and comments

@@
 
-Spatio-temporal data has a spatial component referring to the location of each observation, a temporal component that is recorded at regular or irregular time intervals, and may have multiple variables measured at each spatial and temporal value.
-With spatio-temporal data, one can fix the time to explore the spatial distribution, fix the spatial location to explore temporal trends, or dynamically explore the space and time simultaneously.
-Underlying all these tasks is a data representation that enables the spatio-temporal data to be sliced and diced differently.
+Spatio-temporal data has a spatial component referring to the location of each observation and a temporal component that is recorded at regular or irregular time intervals.
+It may also include multiple variables measured at each spatial and temporal values.
+With spatio-temporal data, one can fix the time to explore the spatial features of the data, fix the spatial location/s to explore temporal aspects, or dynamically explore the space and time simultaneously.
+<!-- Underlying all these tasks, there is a data representation that enables the spatio-temporal data to be sliced and diced differently.  -->
+In order to computationally explore the spatial, temporal and spatio-temporal faces of such data, the data needs to be stored and represented under a specific data object that allows the user to query, group and dissect all the data faces.**
 
-The SpatioTemproal CRAN task view [@ctvspatiotemporal] has a section *Representing data* that lists existing spatio-temporal data representations in \proglang{R}.
+The SpatioTemporal CRAN task view [@ctvspatiotemporal] gathers information about R packages designed for spatio-temporal data and it has a section on *Representing data* that lists existing spatio-temporal data representations used in \proglang{R}.
 Among them, @spacetime proposes four spatio-temporal layouts: full grid layout, sparse grid layout, irregular layout, and trajectory.
-These layouts are implemented in the \pkg{spacetime} [@spacetime] package with the underlying spatial and temporal classes \pkg{sp} [@sp] and \pkg{xts} [@xts].
+These layouts are implemented in the \pkg{spacetime} [@spacetime] package via the underlying spatial and temporal classes \pkg{sp} [@sp] and \pkg{xts} [@xts].
 The \pkg{stars} [@stars] package has been a recent implementation to handle both raster and vector data using spatio-temporal arrays.
@@
 
+**The paragraph above needs a bit more flesh**
+
 Still, the data representation for spatio-temporal data can be further improved and there are two reasons for this.
-Firstly, the raw data sourced in the wild is less often presented in any one of the representations above, and fitting the raw data into a data object can sometimes be difficult.
+Firstly, the raw data sourced in the wild is less often presented in any one of the representations above (Which representations - in the previous paragraph there are more than two?), and fitting the raw data into a data object can sometimes be difficult.
 More often, spatio-temporal data are collected in separate 2D tables and analysts need to assemble them into a whole piece before exploring the data.
 Examples of components of spatio-temporal data can be 1) map data recording the shape of a collection of areas of interest; 2) geo-scientific data recording the longitude and latitude coordinates of locations in the area along with other metadata related to the location, and; 3) temporal variables indexed by both location and time.
+
+**I would aligned the examples of spatio-temporal data with those discuss in the st books. See Paula Moraga for example**
 
@@
 
-The other reason is how tidy data concepts [@tidydata] should be applied to spatio-temporal data.
+The other reason is about tidy data concepts [@tidydata] and how they should be applied to spatio-temporal data.
 Tidy data should be structured into 1) one row per observation, 2) one column per variable, and 3) one type of data per table.
@@
 \pkg{dplyr} [@dplyr] and \pkg{ggplot2} [@ggplot2] for data wrangling and visualisation.
-However, the long form can be inefficient to store feature geometries, especially for large multipolygons for daily or sub-daily data over years, which are extensively collected and handled in time series analysis.
-This poses the question of how to arrange spatial and temporal variables in a way that would make analysing spatio-temporal data easier.
+However, the long form can be inefficient to store feature geometries, especially for large multipolygons for hourly, daily or sub-daily periods over years, which are extensively collected and handled in time series analysis.
+This poses the question of how to arrange spatial and temporal variables in a way that would make data wrangling, visualizing and analysing spatio-temporal data easier.
 
 This paper presents a new \proglang{R} package, \pkg{cubble} which addresses the two issues mentioned above.
-In the package, a new data structure, also called \code{cubble}, is proposed to organise spatial and temporal variables as two forms of a single data object so that they can be wrangled separately while being kept synchronised.
-Among the four spacetime layouts in @spacetime, \pkg{cubble} can be applied to full grid layout, sparse grid layout, or irregular layout, but not trajectory, which is outside the scope of this work.
+In the package, a new data structure, also called \code{cubble}, is proposed to organise spatial and temporal variables as two forms of a single data object so that they can be wrangled separately or combined while being kept synchronised.
+Among the four spacetime layouts in @spacetime, \pkg{cubble} can be applied to full di layout, sparse grid layout, or irregular layout, but not trajectory, which is outside the scope of this work.
 The software is available from the Comprehensive R Archive Network (CRAN) at [CRAN link].
+**At this point what is cubble is adding with respect the other packages it is not  clear to me**
 
-The rest of the paper is divided as follows: Section \@ref(cube) introduces the proposed cube structure as a way to conceptualise multivariate spatio-temporal data.
+The rest of the paper is organized as follows: Section \@ref(cube) introduces the proposed cube structure as a way to conceptualise multivariate spatio-temporal data.
 Section \@ref(cubble) presents the main design and functionality of \pkg{cubble}.
-Section \@ref(others) explains how cubble deals with more advanced considerations, including data with the hierarchical structure, data matching, how cubble fits with existing static and interactive visualisation tools, and spatio-temporal data transformation.
+Section \@ref(others) explains how cubble deals with more advanced considerations, including data with hierarchical structure, data matching, how cubble fits with existing static and interactive visualisation tools, and spatio-temporal data transformation.
 Section \@ref(examples) uses Australian weather station data and river level data as examples to demonstrate the use of \pkg{cubble}.

### 2022-04-27 arxiv/cubble-arxiv.tex
Introduction sections changes and comments

@@
 
-Spatio-temporal data has a spatial component referring to the location of each observation, a temporal component that is recorded at regular or irregular time intervals, and may have multiple variables measured at each spatial and temporal value.
-With spatio-temporal data, one can fix the time to explore the spatial distribution, fix the spatial location to explore temporal trends, or dynamically explore the space and time simultaneously.
-Underlying all these tasks is a data representation that enables the spatio-temporal data to be sliced and diced differently.
-
-The SpatioTemproal CRAN task view (\protect\hyperlink{ref-ctvspatiotemporal}{Edzer Pebesma 2022}) has a section \emph{Representing data} that lists existing spatio-temporal data representations in \proglang{R}.
+Spatio-temporal data has a spatial component referring to the location of each observation and a temporal component that is recorded at regular or irregular time intervals.
+It may also include multiple variables measured at each spatial and temporal values.
+With spatio-temporal data, one can fix the time to explore the spatial features of the data, fix the spatial location/s to explore temporal aspects, or dynamically explore the space and time simultaneously.
+In order to computationally explore the spatial, temporal and spatio-temporal faces of such data, the data needs to be stored and represented under a specific data object that allows the user to query, group and dissect all the data faces.**
+
+The SpatioTemporal CRAN task view (\protect\hyperlink{ref-ctvspatiotemporal}{Edzer Pebesma 2022}) gathers information about R packages designed for spatio-temporal data and it has a section on \emph{Representing data} that lists existing spatio-temporal data representations used in \proglang{R}.
 Among them, \protect\hyperlink{ref-spacetime}{E.
@@
 These layouts are implemented in the \pkg{spacetime} (\protect\hyperlink{ref-spacetime}{E.
-Pebesma 2012}) package with the underlying spatial and temporal classes \pkg{sp} (\protect\hyperlink{ref-sp}{E.
+Pebesma 2012}) package via the underlying spatial and temporal classes \pkg{sp} (\protect\hyperlink{ref-sp}{E.
 Pebesma and Bivand 2005}) and \pkg{xts} (\protect\hyperlink{ref-xts}{Ryan and Ulrich 2020}).
@@
 
+\textbf{The paragraph above needs a bit more flesh}
+
 Still, the data representation for spatio-temporal data can be further improved and there are two reasons for this.
-Firstly, the raw data sourced in the wild is less often presented in any one of the representations above, and fitting the raw data into a data object can sometimes be difficult.
+Firstly, the raw data sourced in the wild is less often presented in any one of the representations above (Which representations - in the previous paragraph there are more than two?), and fitting the raw data into a data object can sometimes be difficult.
 More often, spatio-temporal data are collected in separate 2D tables and analysts need to assemble them into a whole piece before exploring the data.
@@
 
-The other reason is how tidy data concepts (\protect\hyperlink{ref-tidydata}{Wickham 2014}) should be applied to spatio-temporal data.
+\textbf{I would aligned the examples of spatio-temporal data with those discuss in the st books. See Paula Moraga for example}
+
+The other reason is about tidy data concepts (\protect\hyperlink{ref-tidydata}{Wickham 2014}) and how they should be applied to spatio-temporal data.
 Tidy data should be structured into 1) one row per observation, 2) one column per variable, and 3) one type of data per table.
@@
 2022}) and \pkg{ggplot2} (\protect\hyperlink{ref-ggplot2}{Wickham 2016}) for data wrangling and visualisation.
-However, the long form can be inefficient to store feature geometries, especially for large multipolygons for daily or sub-daily data over years, which are extensively collected and handled in time series analysis.
-This poses the question of how to arrange spatial and temporal variables in a way that would make analysing spatio-temporal data easier.
+However, the long form can be inefficient to store feature geometries, especially for large multipolygons for hourly, daily or sub-daily periods over years, which are extensively collected and handled in time series analysis.
+This poses the question of how to arrange spatial and temporal variables in a way that would make data wrangling, visualizing and analysing spatio-temporal data easier.
 
 This paper presents a new \proglang{R} package, \pkg{cubble} which addresses the two issues mentioned above.
-In the package, a new data structure, also called \code{cubble}, is proposed to organise spatial and temporal variables as two forms of a single data object so that they can be wrangled separately while being kept synchronised.
+In the package, a new data structure, also called \code{cubble}, is proposed to organise spatial and temporal variables as two forms of a single data object so that they can be wrangled separately or combined while being kept synchronised.
 Among the four spacetime layouts in \protect\hyperlink{ref-spacetime}{E.
-Pebesma} (\protect\hyperlink{ref-spacetime}{2012}), \pkg{cubble} can be applied to full grid layout, sparse grid layout, or irregular layout, but not trajectory, which is outside the scope of this work.
+Pebesma} (\protect\hyperlink{ref-spacetime}{2012}), \pkg{cubble} can be applied to full di layout, sparse grid layout, or irregular layout, but not trajectory, which is outside the scope of this work.
 The software is available from the Comprehensive R Archive Network (CRAN) at {[}CRAN link{]}.
-
-The rest of the paper is divided as follows: Section \ref{cube} introduces the proposed cube structure as a way to conceptualise multivariate spatio-temporal data.
+\textbf{At this point what is cubble is adding with respect the other packages it is not clear to me}
+
+The rest of the paper is organized as follows: Section \ref{cube} introduces the proposed cube structure as a way to conceptualise multivariate spatio-temporal data.
 Section \ref{cubble} presents the main design and functionality of \pkg{cubble}.
-Section \ref{others} explains how cubble deals with more advanced considerations, including data with the hierarchical structure, data matching, how cubble fits with existing static and interactive visualisation tools, and spatio-temporal data transformation.
+Section \ref{others} explains how cubble deals with more advanced considerations, including data with hierarchical structure, data matching, how cubble fits with existing static and interactive visualisation tools, and spatio-temporal data transformation.
 Section \ref{examples} uses Australian weather station data and river level data as examples to demonstrate the use of \pkg{cubble}.

### 2022-04-27 2-0-cube.Rmd
Comments and modifications 2-0-cube section

@@
 Spatio-temporal data can be conceptualised using a cubical data model with three axes, typically time, latitude and longitude.
-This abstraction can be useful for generalising operations and visualisation: @lu_multidimensional_2018 shows how array operations (select, scale, reduce, rearrange, and compute) can be mapped onto the cube; @bach_review_2014 reviews the temporal data visualisation based on space-time cube operations.
+This abstraction can be useful for generalising operations and visualisation purposes: @lu_multidimensional_2018 shows how array operations (select, scale, reduce, rearrange, and compute) can be mapped onto the cube; @bach_review_2014 reviews the temporal data visualisation based on space-time cube operations.
 Notice that the term space-time cube in their article "does not need to involve spatial data", but refers to "an abstract 2D substrate that is used to visualize data at a specific time".
@@
 Finally, variables are stacked on this space-time canvas, with one observation per site and time point.
-This notion is adopted to avoid using hypercubes when describing multivariate spatio-temporal data.
+This notion is adopted to avoid using hypercubes when describing multivariate spatio-temporal data and is the workhorse behind \pkg{cubble}.
 
@@
 \textcolor{blue}{Ursula: this seems too specific, you should be able to make the same argument without referencing the mutate function?}
+\textcolor{green}{How about the other functions from dplyr?, the same principle would apply right?}
 

### 2022-04-27 arxiv/cubble-arxiv.tex
Comments and modifications 2-0-cube section

@@
 Spatio-temporal data can be conceptualised using a cubical data model with three axes, typically time, latitude and longitude.
-This abstraction can be useful for generalising operations and visualisation: \protect\hyperlink{ref-lu_multidimensional_2018}{Lu, Appel, and Pebesma} (\protect\hyperlink{ref-lu_multidimensional_2018}{2018}) shows how array operations (select, scale, reduce, rearrange, and compute) can be mapped onto the cube; \protect\hyperlink{ref-bach_review_2014}{Bach et al.} (\protect\hyperlink{ref-bach_review_2014}{2014}) reviews the temporal data visualisation based on space-time cube operations.
+This abstraction can be useful for generalising operations and visualisation purposes: \protect\hyperlink{ref-lu_multidimensional_2018}{Lu, Appel, and Pebesma} (\protect\hyperlink{ref-lu_multidimensional_2018}{2018}) shows how array operations (select, scale, reduce, rearrange, and compute) can be mapped onto the cube; \protect\hyperlink{ref-bach_review_2014}{Bach et al.} (\protect\hyperlink{ref-bach_review_2014}{2014}) reviews the temporal data visualisation based on space-time cube operations.
 Notice that the term space-time cube in their article ``does not need to involve spatial data,'' but refers to ``an abstract 2D substrate that is used to visualize data at a specific time.'' Despite its main focus being on temporal data, the mindset of abstracting out data representation to construct visualisation still applies to our spatio-temporal data manipulation and visualisation.

### 2022-04-27 2-the-cubble-pkg.Rmd
Few comments on 2-the-cubble-pkg section

@@
 
-Each core function is introduced accompanied with a short example and the data `climate_flat` is used throughout this section.
-`climate_flat` contains five weather stations in Australia with spatial information of each station: station id, latitude, longitude, elevation, station name, World Meteorology Organisation ID and also daily temporal information of date, maximum and minimum temperature and precipitation for 2020.
+Each core \pkg{cubble} function is introduced and accompanied with a short example using the data `climate_flat`  throughout this section.
+`climate_flat` contains data on five weather stations in Australia including station id, latitude, longitude, elevation, station name, \textcolor{green}{World Meteorology Organisation ID and also information on the record's date (recorded daily) together with the maximum and minimum temperature and precipitation for 2020}.
 The first five rows of `climate_flat` are shown below:

### 2022-04-27 arxiv/cubble-arxiv.tex
Few comments on 2-the-cubble-pkg section

@@
 
-Each core function is introduced accompanied with a short example and the data \texttt{climate\_flat} is used throughout this section.
-\texttt{climate\_flat} contains five weather stations in Australia with spatial information of each station: station id, latitude, longitude, elevation, station name, World Meteorology Organisation ID and also daily temporal information of date, maximum and minimum temperature and precipitation for 2020.
+Each core \pkg{cubble} function is introduced and accompanied with a short example using the data \texttt{climate\_flat} throughout this section.
+\texttt{climate\_flat} contains five weather stations in Australia with spatial information of each station: station id, latitude, longitude, elevation, station name, \textcolor{green}{World Meteorology Organisation ID and also information on the record's date (recorded daily) together with the maximum and minimum temperature and precipitation for 2020}.
 The first five rows of \texttt{climate\_flat} are shown below:

### 2022-04-27 1-intro.Rmd
misc small changes

@@
 <!-- Underlying all these tasks, there is a data representation that enables the spatio-temporal data to be sliced and diced differently.  -->
-In order to computationally explore the spatial, temporal and spatio-temporal faces of such data, the data needs to be stored and represented under a specific data object that allows the user to query, group and dissect all the data faces.**
+In order to computationally explore the spatial, temporal and spatio-temporal faces of such data, the data needs to be stored and represented under a specific data object that allows the user to query, group and dissect all the data faces.
 
@@
 Among the four spacetime layouts in @spacetime, \pkg{cubble} can be applied to full di layout, sparse grid layout, or irregular layout, but not trajectory, which is outside the scope of this work.
-The software is available from the Comprehensive R Archive Network (CRAN) at [CRAN link].
+The software is available from the Comprehensive R Archive Network (CRAN) at https://CRAN.R-project.org/package=cubble.
 **At this point what is cubble is adding with respect the other packages it is not  clear to me**

### 2022-04-27 2-0-cube.Rmd
misc small changes

@@
 
-\textcolor{blue}{Ursula: I have rearranged this section somewhat, please check!}
-
 Spatio-temporal data can be conceptualised using a cubical data model with three axes, typically time, latitude and longitude.
@@
 Finally, variables are stacked on this space-time canvas, with one observation per site and time point.
-This notion is adopted to avoid using hypercubes when describing multivariate spatio-temporal data and is the workhorse behind \pkg{cubble}.
+This notion is adopted to avoid using hypercubes when describing multivariate spatio-temporal data and is the conceptual framework behind \pkg{cubble}.
 
@@
 For example, wrangling on character strings and specific datetime classes is common for spatio-temporal data but may not be easily done with array objects.
-2) The `mutate()` syntax for creating new variables is no longer valid since the operation can no longer define how a created 1D vector should be added to the array.
-\textcolor{blue}{Ursula: this seems too specific, you should be able to make the same argument without referencing the mutate function?}
-\textcolor{green}{How about the other functions from dplyr?, the same principle would apply right?}
+2) Creating new variables in an array is more complex than in a data frame since it now needs to be defined in a 2D space, rather than a 1D vector, to be added into the array.
 

### 2022-04-27 3-advanced.Rmd
fix merge conflict
misc small changes

@@
 
-## Data fusion and matching
+## Data fusion and matching {#matching}
 
@@
 In \proglang{R}, \pkg{GGally} implements glyph maps through the \code{glyphs()} function.
-The function constructs a data frame with calculated position (\code{gx}, \code{gy}.
-\code{gid}) of each point on the time series using linear algebra (Equation 1 and 2 in @Wickham2012-yr).
+The function constructs a data frame with calculated position (\code{gx}, \code{gy}, \code{gid}) of each point on the time series using linear algebra (Equation 1 and 2 in @Wickham2012-yr).
 The data can then be piped into \code{ggplot} to create the glyph map as:

### 2022-04-27 4-1-basic.Rmd
misc small changes

@@
 load(here::here("data/covid.rda"))
-lga <- absmapsdata::lga2018 %>%
+
+lga <- strayr::read_absmap("lga2018") %>%
   rename(lga = lga_name_2018) %>%
-  filter(state_name_2016 == "Victoria") %>%
-  rmapshaper::ms_simplify()
+  filter(state_name_2016 == "Victoria") 
 ```

### 2022-04-27 5-conclusion.Rmd
misc small changes

@@
 
-This paper describes an \proglang{R} package \pkg{cubble} for manipulating and visualising spatio-temporal data.
-A new data structure, \code{cubble} that builds from the \code{rowwise_df} and \code{grouped_df} class in the tidyverse ecosystem, is proposed to connect the time invariant and varying variables in the spatio-temporal data.
+This paper describes an \proglang{R} package \pkg{cubble} for organising, manipulating and visualising spatio-temporal data.
+A new data structure, \code{cubble}, with the nested and long form, is proposed to connect the time invariant and varying variables in the spatio-temporal data.
 This design frees the data analysts from spending time on organising variables of different observational units.
@@
 
-Further development and maintenance of the package involves responding to changes in the tidyverse packages that \pkg{cubble} imports, in particular, \pkg{tibble}, \pkg{tidyr}, and \pkg{dplyr}.
+Future directions of the package involves responding to changes in the tidyverse packages that \pkg{cubble} imports, in particular, \pkg{tibble}, \pkg{tidyr}, and \pkg{dplyr}.
 In the spatial aspect, the simple feature representation wraps spatial coordinates in a list column, for some complex spatial operations, while sometimes, analysts may want to unpack them into longitude and latitude columns to work from.

### 2022-04-27 arxiv/cubble-arxiv.Rmd
misc small changes

@@
 ---
-title: "cubble: An R Package for Structuring Spatio-temporal Data"
+title: "Wrangling multivariate spatio-temporal data with the R package cubble"
 authors:
@@
 For many analyses, spatial and time components can be separately studied: for example, to explore the temporal trend of one variable for a single spatial location, or to model the spatial distribution of one variable at a given time.
-However for some studies, it is important to analyse different aspects of the spatio-temporal data simultaneouly, like for instance instance, temporal trends of multiple variables across locations.
+However for some studies, it is important to analyse different aspects of the spatio-temporal data simultaneouly, like for instance, temporal trends of multiple variables across locations.
 In order to facilitate the study of different sections or combinations of spatio-temporal data, we introduce a new data structure, \code{cubble}, with a suite of functions enabling easy slicing and dicing on the different components spatio-temporal components, for visual and numerical explorations while facilitating data wrangling for modelling.
-The propose \code{cubble} structure ensures that all the components of the data are easy to access and manipulate while providing flexibility for data analysis.
+The proposed \code{cubble} structure ensures that all the components of the data are easy to access and manipulate while providing flexibility for data analysis.
 The \code{cubble} structure and the tools provided in the \code{cubble} R package equip users with the capability to handle hierarchical spatial and temporal structures.
-\color{green}{The tools are illustrated with examples of Australian climate data, merging climate and river level data sources, reproducing a climate reanalysis (ERA5), and creating interactive graphics.}
+The tools are illustrated with examples of Australian climate data, merging climate and river level data sources, reproducing a climate reanalysis (ERA5), and creating interactive graphics.
 keywords:

### 2022-04-27 arxiv/cubble-arxiv.tex
misc small changes

--- 2022-04-27 arxiv/cubble-arxiv.tex
@@
 
-\title{cubble: An R Package for Structuring Spatio-temporal Data}
+\title{Wrangling multivariate spatio-temporal data with the R package cubble}
 
@@
 For many analyses, spatial and time components can be separately studied: for example, to explore the temporal trend of one variable for a single spatial location, or to model the spatial distribution of one variable at a given time.
-However for some studies, it is important to analyse different aspects of the spatio-temporal data simultaneouly, like for instance instance, temporal trends of multiple variables across locations.
+However for some studies, it is important to analyse different aspects of the spatio-temporal data simultaneouly, like for instance, temporal trends of multiple variables across locations.
 In order to facilitate the study of different sections or combinations of spatio-temporal data, we introduce a new data structure, \code{cubble}, with a suite of functions enabling easy slicing and dicing on the different components spatio-temporal components, for visual and numerical explorations while facilitating data wrangling for modelling.
-The propose \code{cubble} structure ensures that all the components of the data are easy to access and manipulate while providing flexibility for data analysis.
+The proposed \code{cubble} structure ensures that all the components of the data are easy to access and manipulate while providing flexibility for data analysis.
 The \code{cubble} structure and the tools provided in the \code{cubble} R package equip users with the capability to handle hierarchical spatial and temporal structures.
-\color{green}{The tools are illustrated with examples of Australian climate data, merging climate and river level data sources, reproducing a climate reanalysis (ERA5), and creating interactive graphics.}
+The tools are illustrated with examples of Australian climate data, merging climate and river level data sources, reproducing a climate reanalysis (ERA5), and creating interactive graphics.
 \end{abstract}
@@
 With spatio-temporal data, one can fix the time to explore the spatial features of the data, fix the spatial location/s to explore temporal aspects, or dynamically explore the space and time simultaneously.
-In order to computationally explore the spatial, temporal and spatio-temporal faces of such data, the data needs to be stored and represented under a specific data object that allows the user to query, group and dissect all the data faces.**
+In order to computationally explore the spatial, temporal and spatio-temporal faces of such data, the data needs to be stored and represented under a specific data object that allows the user to query, group and dissect all the data faces.
 
 The SpatioTemporal CRAN task view (\protect\hyperlink{ref-ctvspatiotemporal}{Edzer Pebesma 2022}) gathers information about R packages designed for spatio-temporal data and it has a section on \emph{Representing data} that lists existing spatio-temporal data representations used in \proglang{R}.
-Among them, \protect\hyperlink{ref-spacetime}{E.
-Pebesma} (\protect\hyperlink{ref-spacetime}{2012}) proposes four spatio-temporal layouts: full grid layout, sparse grid layout, irregular layout, and trajectory.
+Among them, E.
+Pebesma (\protect\hyperlink{ref-spacetime}{2012}) proposes four spatio-temporal layouts: full grid layout, sparse grid layout, irregular layout, and trajectory.
 These layouts are implemented in the \pkg{spacetime} (\protect\hyperlink{ref-spacetime}{E.
@@
 In the package, a new data structure, also called \code{cubble}, is proposed to organise spatial and temporal variables as two forms of a single data object so that they can be wrangled separately or combined while being kept synchronised.
-Among the four spacetime layouts in \protect\hyperlink{ref-spacetime}{E.
-Pebesma} (\protect\hyperlink{ref-spacetime}{2012}), \pkg{cubble} can be applied to full di layout, sparse grid layout, or irregular layout, but not trajectory, which is outside the scope of this work.
-The software is available from the Comprehensive R Archive Network (CRAN) at {[}CRAN link{]}.
+Among the four spacetime layouts in E.
+Pebesma (\protect\hyperlink{ref-spacetime}{2012}), \pkg{cubble} can be applied to full di layout, sparse grid layout, or irregular layout, but not trajectory, which is outside the scope of this work.
+The software is available from the Comprehensive R Archive Network (CRAN) at \url{https://CRAN.R-project.org/package=cubble}.
 \textbf{At this point what is cubble is adding with respect the other packages it is not clear to me}
@@
 
-\textcolor{blue}{Ursula: I have rearranged this section somewhat, please check!}
-
 Spatio-temporal data can be conceptualised using a cubical data model with three axes, typically time, latitude and longitude.
-This abstraction can be useful for generalising operations and visualisation: \protect\hyperlink{ref-lu_multidimensional_2018}{Lu, Appel, and Pebesma} (\protect\hyperlink{ref-lu_multidimensional_2018}{2018}) shows how array operations (select, scale, reduce, rearrange, and compute) can be mapped onto the cube; \protect\hyperlink{ref-bach_review_2014}{Bach et al.} (\protect\hyperlink{ref-bach_review_2014}{2014}) reviews the temporal data visualisation based on space-time cube operations.
-Notice that the term space-time cube in their article ``does not need to involve spatial data,'' but refers to ``an abstract 2D substrate that is used to visualize data at a specific time.'' Despite its main focus being on temporal data, the mindset of abstracting out data representation to construct visualisation still applies to our spatio-temporal data manipulation and visualisation.
+This abstraction can be useful for generalising operations and visualisation purposes: Lu, Appel, and Pebesma (\protect\hyperlink{ref-lu_multidimensional_2018}{2018}) shows how array operations (select, scale, reduce, rearrange, and compute) can be mapped onto the cube; Bach et al.
+(\protect\hyperlink{ref-bach_review_2014}{2014}) reviews the temporal data visualisation based on space-time cube operations.
+Notice that the term space-time cube in their article ``does not need to involve spatial data'', but refers to ``an abstract 2D substrate that is used to visualize data at a specific time''.
+Despite its main focus being on temporal data, the mindset of abstracting out data representation to construct visualisation still applies to our spatio-temporal data manipulation and visualisation.
 
@@
 Finally, variables are stacked on this space-time canvas, with one observation per site and time point.
-This notion is adopted to avoid using hypercubes when describing multivariate spatio-temporal data.
+This notion is adopted to avoid using hypercubes when describing multivariate spatio-temporal data and is the conceptual framework behind \pkg{cubble}.
 
@@
 For example, wrangling on character strings and specific datetime classes is common for spatio-temporal data but may not be easily done with array objects.
-2) The \texttt{mutate()} syntax for creating new variables is no longer valid since the operation can no longer define how a created 1D vector should be added to the array.
-\textcolor{blue}{Ursula: this seems too specific, you should be able to make the same argument without referencing the mutate function?}
+2) Creating new variables in an array is more complex than in a data frame since it now needs to be defined in a 2D space, rather than a 1D vector, to be added into the array.
 
@@
 
-{\centering \includegraphics[width=0.5\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/cube-def}
+{\centering \includegraphics[width=0.5\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/cube-def}
 
@@
 
-Each core function is introduced accompanied with a short example and the data \texttt{climate\_flat} is used throughout this section.
+Each core \pkg{cubble} function is introduced and accompanied with a short example using the data \texttt{climate\_flat} throughout this section.
 \texttt{climate\_flat} contains five weather stations in Australia with spatial information of each station: station id, latitude, longitude, elevation, station name, World Meteorology Organisation ID and also daily temporal information of date, maximum and minimum temperature and precipitation for 2020.
@@
 \begin{Highlighting}[]
-\NormalTok{cubble\_nested }\OtherTok{\textless{}{-}}\NormalTok{ climate\_flat }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
+\NormalTok{cubble\_nested }\OtherTok{\textless{}{-}}\NormalTok{ climate\_flat }\SpecialCharTok{|\textgreater{}}
   \FunctionTok{as\_cubble}\NormalTok{(}\AttributeTok{key =}\NormalTok{ id, }\AttributeTok{index =}\NormalTok{ date, }\AttributeTok{coords =} \FunctionTok{c}\NormalTok{(long, lat))}
@@
 \begin{Highlighting}[]
-\NormalTok{cubble\_nested }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
+\NormalTok{cubble\_nested }\SpecialCharTok{|\textgreater{}}
   \FunctionTok{mutate}\NormalTok{(}\AttributeTok{rain\_day =} \FunctionTok{sum}\NormalTok{(ts}\SpecialCharTok{$}\NormalTok{prcp }\SpecialCharTok{!=} \DecValTok{0}\NormalTok{))}
@@
 \begin{Highlighting}[]
-\NormalTok{cubble\_long }\OtherTok{\textless{}{-}}\NormalTok{ cubble\_nested }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{face\_temporal}\NormalTok{()}
+\NormalTok{cubble\_long }\OtherTok{\textless{}{-}}\NormalTok{ cubble\_nested }\SpecialCharTok{|\textgreater{}} \FunctionTok{face\_temporal}\NormalTok{()}
 \NormalTok{cubble\_long}
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/diagram-keynotes/diagram-keynotes.008}
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.008}
 
@@
 \begin{Highlighting}[]
-\NormalTok{cubble\_back }\OtherTok{\textless{}{-}}\NormalTok{ cubble\_long }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{face\_spatial}\NormalTok{()}
+\NormalTok{cubble\_back }\OtherTok{\textless{}{-}}\NormalTok{ cubble\_long }\SpecialCharTok{|\textgreater{}} \FunctionTok{face\_spatial}\NormalTok{()}
 \NormalTok{cubble\_back}
@@
 \begin{verbatim}
-## [1] FALSE
+## [1] TRUE
 \end{verbatim}
@@
 \begin{Highlighting}[]
-\NormalTok{cubble\_unfold }\OtherTok{\textless{}{-}}\NormalTok{ cubble\_long }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{unfold}\NormalTok{(long, lat)}
+\NormalTok{cubble\_unfold }\OtherTok{\textless{}{-}}\NormalTok{ cubble\_long }\SpecialCharTok{|\textgreater{}} \FunctionTok{unfold}\NormalTok{(long, lat)}
 \NormalTok{cubble\_unfold}
@@
 \begin{Highlighting}[]
-\NormalTok{cubble\_unfold }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{face\_spatial}\NormalTok{() }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{face\_temporal}\NormalTok{()}
+\NormalTok{cubble\_unfold }\SpecialCharTok{|\textgreater{}} \FunctionTok{face\_spatial}\NormalTok{() }\SpecialCharTok{|\textgreater{}} \FunctionTok{face\_temporal}\NormalTok{()}
 \end{Highlighting}
@@
 \CommentTok{\# cubble workflow}
-\NormalTok{dt }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
-  \FunctionTok{as\_cubble}\NormalTok{(...) }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
-  \FunctionTok{face\_temporal}\NormalTok{() }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
-  \FunctionTok{group\_by}\NormalTok{() }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
-  \FunctionTok{summarise}\NormalTok{() }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
+\NormalTok{dt }\SpecialCharTok{|\textgreater{}}
+  \FunctionTok{as\_cubble}\NormalTok{(...) }\SpecialCharTok{|\textgreater{}}
+  \FunctionTok{face\_temporal}\NormalTok{() }\SpecialCharTok{|\textgreater{}}
+  \FunctionTok{group\_by}\NormalTok{() }\SpecialCharTok{|\textgreater{}}
+  \FunctionTok{summarise}\NormalTok{() }\SpecialCharTok{|\textgreater{}}
   \FunctionTok{unfold}\NormalTok{(...)}
@@
 \CommentTok{\# tidyverse workflow}
-\NormalTok{temporal }\OtherTok{\textless{}{-}}\NormalTok{ dt }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
-  \FunctionTok{group\_by}\NormalTok{(...) }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
+\NormalTok{temporal }\OtherTok{\textless{}{-}}\NormalTok{ dt }\SpecialCharTok{|\textgreater{}}
+  \FunctionTok{group\_by}\NormalTok{(...) }\SpecialCharTok{|\textgreater{}}
   \FunctionTok{summarise}\NormalTok{(...)}
 
-\NormalTok{spatial }\OtherTok{\textless{}{-}}\NormalTok{ dt }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{select}\NormalTok{(....)}
-
-\NormalTok{spatial }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{left\_join}\NormalTok{(temporal)}
+\NormalTok{spatial }\OtherTok{\textless{}{-}}\NormalTok{ dt }\SpecialCharTok{|\textgreater{}} \FunctionTok{select}\NormalTok{(....)}
+
+\NormalTok{spatial }\SpecialCharTok{|\textgreater{}} \FunctionTok{left\_join}\NormalTok{(temporal)}
 \end{Highlighting}
@@
 \CommentTok{\# example with a tsibble created from climate\_flat}
-\NormalTok{raw }\OtherTok{\textless{}{-}}\NormalTok{ climate\_flat }\SpecialCharTok{|}\ErrorTok{\textgreater{}}\NormalTok{ tsibble}\SpecialCharTok{::}\FunctionTok{as\_tsibble}\NormalTok{(}\AttributeTok{key =}\NormalTok{ id, }\AttributeTok{index =}\NormalTok{ date)}
-\NormalTok{dt }\OtherTok{\textless{}{-}}\NormalTok{  raw }\SpecialCharTok{|}\ErrorTok{\textgreater{}}\NormalTok{ cubble}\SpecialCharTok{::}\FunctionTok{as\_cubble}\NormalTok{(}\AttributeTok{coords =} \FunctionTok{c}\NormalTok{(long, lat))}
+\NormalTok{raw }\OtherTok{\textless{}{-}}\NormalTok{ climate\_flat }\SpecialCharTok{|\textgreater{}}\NormalTok{ tsibble}\SpecialCharTok{::}\FunctionTok{as\_tsibble}\NormalTok{(}\AttributeTok{key =}\NormalTok{ id, }\AttributeTok{index =}\NormalTok{ date)}
+\NormalTok{dt }\OtherTok{\textless{}{-}}\NormalTok{  raw }\SpecialCharTok{|\textgreater{}}\NormalTok{ cubble}\SpecialCharTok{::}\FunctionTok{as\_cubble}\NormalTok{(}\AttributeTok{coords =} \FunctionTok{c}\NormalTok{(long, lat))}
 \NormalTok{dt}
@@
 \CommentTok{\# add station{-}based features in the nested form.}
-\NormalTok{dt }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{mutate}\NormalTok{(fabletools}\SpecialCharTok{::}\FunctionTok{features}\NormalTok{(ts, tmax, }\FunctionTok{list}\NormalTok{(}\AttributeTok{tmax\_mean =}\NormalTok{ mean, }\AttributeTok{tmax\_var =}\NormalTok{ var)))}
+\NormalTok{dt }\SpecialCharTok{|\textgreater{}} \FunctionTok{mutate}\NormalTok{(fabletools}\SpecialCharTok{::}\FunctionTok{features}\NormalTok{(ts, tmax, }\FunctionTok{list}\NormalTok{(}\AttributeTok{tmax\_mean =}\NormalTok{ mean, }\AttributeTok{tmax\_var =}\NormalTok{ var)))}
 \end{Highlighting}
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/diagram-keynotes/diagram-keynotes.003}
+{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.003}
 
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/illu-matching} \caption{An illustration of temporal matching in cubble.
+\includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/illu-matching} \caption{An illustration of temporal matching in cubble.
 Three highest peaks are identified in each series and intervals are constructed on series \code{A}.
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/diagram-keynotes/diagram-keynotes.004}
+{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.004}
 
@@
 In \proglang{R}, \pkg{GGally} implements glyph maps through the \code{glyphs()} function.
-The function constructs a data frame with calculated position (\code{gx}, \code{gy}.
-\code{gid}) of each point on the time series using linear algebra (Equation 1 and 2 in \protect\hyperlink{ref-Wickham2012-yr}{Wickham et al.} (\protect\hyperlink{ref-Wickham2012-yr}{2012})).
+The function constructs a data frame with calculated position (\code{gx}, \code{gy}, \code{gid}) of each point on the time series using linear algebra (Equation 1 and 2 in Wickham et al.
+(\protect\hyperlink{ref-Wickham2012-yr}{2012})).
 The data can then be piped into \code{ggplot} to create the glyph map as:
@@
 \begin{Highlighting}[]
-\NormalTok{covid }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{head}\NormalTok{(}\DecValTok{5}\NormalTok{)}
+\NormalTok{covid }\SpecialCharTok{|\textgreater{}} \FunctionTok{head}\NormalTok{(}\DecValTok{5}\NormalTok{)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{lga }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{head}\NormalTok{(}\DecValTok{5}\NormalTok{)}
+\NormalTok{lga }\SpecialCharTok{|\textgreater{}} \FunctionTok{head}\NormalTok{(}\DecValTok{5}\NormalTok{)}
 \end{Highlighting}
@@
 ## $others$temporal
-## [1] "Interstate"       "Overseas"         "Queenscliffe (B)" "Unknown"         
+## [1] "Interstate" "Overseas"   "Unknown"   
 ## 
 ## $others$spatial
-## character(0)
+## [1] "No usual address (Vic.)"               
+## [2] "Migratory - Offshore - Shipping (Vic.)"
 \end{verbatim}
@@
 \begin{Highlighting}[]
-\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ weatherdata}\SpecialCharTok{::}\NormalTok{historical\_tmax }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
+\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ weatherdata}\SpecialCharTok{::}\NormalTok{historical\_tmax }\SpecialCharTok{|\textgreater{}}
   \FunctionTok{filter}\NormalTok{(}\FunctionTok{between}\NormalTok{(stringr}\SpecialCharTok{::}\FunctionTok{str\_sub}\NormalTok{(id, }\DecValTok{7}\NormalTok{, }\DecValTok{8}\NormalTok{), }\DecValTok{46}\NormalTok{, }\DecValTok{90}\NormalTok{))}
@@
 \begin{Highlighting}[]
-\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
-  \FunctionTok{face\_temporal}\NormalTok{() }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
+\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|\textgreater{}}
+  \FunctionTok{face\_temporal}\NormalTok{() }\SpecialCharTok{|\textgreater{}}
   \FunctionTok{filter}\NormalTok{(lubridate}\SpecialCharTok{::}\FunctionTok{year}\NormalTok{(date) }\SpecialCharTok{\%in\%} \FunctionTok{c}\NormalTok{(}\DecValTok{1971}\SpecialCharTok{:}\DecValTok{1975}\NormalTok{, }\DecValTok{2016}\SpecialCharTok{:}\DecValTok{2020}\NormalTok{))}
@@
 \begin{Highlighting}[]
-\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
+\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|\textgreater{}}
   \FunctionTok{group\_by}\NormalTok{(}\AttributeTok{month =}\NormalTok{ lubridate}\SpecialCharTok{::}\FunctionTok{month}\NormalTok{(date),}
          \AttributeTok{group =} \FunctionTok{as.factor}\NormalTok{(}\FunctionTok{ifelse}\NormalTok{(lubridate}\SpecialCharTok{::}\FunctionTok{year}\NormalTok{(date) }\SpecialCharTok{\textgreater{}} \DecValTok{2015}\NormalTok{,}
-                                  \StringTok{"2016 \textasciitilde{} 2020"}\NormalTok{, }\StringTok{"1971 \textasciitilde{} 1975"}\NormalTok{))) }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
+                                  \StringTok{"2016 \textasciitilde{} 2020"}\NormalTok{, }\StringTok{"1971 \textasciitilde{} 1975"}\NormalTok{))) }\SpecialCharTok{|\textgreater{}}
   \FunctionTok{summarise}\NormalTok{(}\AttributeTok{tmax =} \FunctionTok{mean}\NormalTok{(tmax, }\AttributeTok{na.rm =} \ConstantTok{TRUE}\NormalTok{))}
@@
 \begin{Highlighting}[]
-\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{face\_spatial}\NormalTok{() }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{filter}\NormalTok{(}\FunctionTok{nrow}\NormalTok{(ts) }\SpecialCharTok{==} \DecValTok{24}\NormalTok{)}
+\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|\textgreater{}} \FunctionTok{face\_spatial}\NormalTok{() }\SpecialCharTok{|\textgreater{}} \FunctionTok{filter}\NormalTok{(}\FunctionTok{nrow}\NormalTok{(ts) }\SpecialCharTok{==} \DecValTok{24}\NormalTok{)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{face\_temporal}\NormalTok{() }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{unfold}\NormalTok{(latitude, longitude)}
+\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|\textgreater{}} \FunctionTok{face\_temporal}\NormalTok{() }\SpecialCharTok{|\textgreater{}} \FunctionTok{unfold}\NormalTok{(latitude, longitude)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{nsw\_vic }\OtherTok{\textless{}{-}}\NormalTok{ ozmaps}\SpecialCharTok{::}\NormalTok{abs\_ste }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
+\NormalTok{nsw\_vic }\OtherTok{\textless{}{-}}\NormalTok{ ozmaps}\SpecialCharTok{::}\NormalTok{abs\_ste }\SpecialCharTok{|\textgreater{}}
   \FunctionTok{filter}\NormalTok{(NAME }\SpecialCharTok{\%in\%} \FunctionTok{c}\NormalTok{(}\StringTok{"Victoria"}\NormalTok{, }\StringTok{"New South Wales"}\NormalTok{))}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth,height=0.7\textheight]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/basic-manip} \caption{A glyph map of the average maximum temperature by month of Victoria and New South Wales weather stations in Australia.
+\includegraphics[width=1\linewidth,height=0.7\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/basic-manip} \caption{A glyph map of the average maximum temperature by month of Victoria and New South Wales weather stations in Australia.
 On the top left corner is an insetted plot of station Cobar highlighted in the black box.}\label{fig:basic-manip}
@@
 \begin{Highlighting}[]
-\NormalTok{station\_nested }\OtherTok{\textless{}{-}}\NormalTok{ weatherdata}\SpecialCharTok{::}\NormalTok{climate\_full }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
+\NormalTok{station\_nested }\OtherTok{\textless{}{-}}\NormalTok{ weatherdata}\SpecialCharTok{::}\NormalTok{climate\_full }\SpecialCharTok{|\textgreater{}}
   \FunctionTok{mutate}\NormalTok{(}\AttributeTok{cluster =}\NormalTok{ ...)}
@@
 \begin{Highlighting}[]
-\NormalTok{cluster\_nested }\OtherTok{\textless{}{-}}\NormalTok{ station\_nested }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{switch\_key}\NormalTok{(cluster)}
+\NormalTok{cluster\_nested }\OtherTok{\textless{}{-}}\NormalTok{ station\_nested }\SpecialCharTok{|\textgreater{}} \FunctionTok{switch\_key}\NormalTok{(cluster)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{cluster\_nested }\OtherTok{\textless{}{-}}\NormalTok{ cluster\_nested }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{get\_centroid}\NormalTok{()}
+\NormalTok{cluster\_nested }\OtherTok{\textless{}{-}}\NormalTok{ cluster\_nested }\SpecialCharTok{|\textgreater{}} \FunctionTok{get\_centroid}\NormalTok{()}
 \end{Highlighting}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/basic-agg} \caption{Profile of aggregated precipitation from 639 weather stations in Australia.
+\includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/basic-agg} \caption{Profile of aggregated precipitation from 639 weather stations in Australia.
 Subplot A shows the glyph map of the weekly averaged precipitation of each cluster.
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/matching-map} \caption{Location of weather stations and river gauges in Victoria, Australia.}\label{fig:matching-map}
+\includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/matching-map} \caption{Location of weather stations and river gauges in Victoria, Australia.}\label{fig:matching-map}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/matching} \caption{Matched weather stations and river gauges on the map (A) and across time (B).
+\includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/matching} \caption{Matched weather stations and river gauges on the map (A) and across time (B).
 Precipitation and water level have been standardised between 0 and 1 to be displayed on the same scale.
@@
 
-Figure \ref{fig:netcdf} reproduces the ERA5 data row of Figure 19 in \protect\hyperlink{ref-hersbach2020era5}{Hersbach et al.} (\protect\hyperlink{ref-hersbach2020era5}{2020}).
+Figure \ref{fig:netcdf} reproduces the ERA5 data row of Figure 19 in Hersbach et al.
+(\protect\hyperlink{ref-hersbach2020era5}{2020}).
 It shows the southern polar vortex splits into two on 2002-09-26 and further splits into four on 2002-10-04 in the stratosphere.
-Readers interested in the analysis of this figure can refer to \protect\hyperlink{ref-hersbach2020era5}{Hersbach et al.} (\protect\hyperlink{ref-hersbach2020era5}{2020}), \protect\hyperlink{ref-simmons2020global}{Simmons et al.} (\protect\hyperlink{ref-simmons2020global}{2020}), and \protect\hyperlink{ref-simmons2005ecmwf}{Simmons et al.} (\protect\hyperlink{ref-simmons2005ecmwf}{2005}) for more details.
+Readers interested in the analysis of this figure can refer to Hersbach et al.
+(\protect\hyperlink{ref-hersbach2020era5}{2020}), Simmons et al.
+(\protect\hyperlink{ref-simmons2020global}{2020}), and Simmons et al.
+(\protect\hyperlink{ref-simmons2005ecmwf}{2005}) for more details.
 
 \begin{figure}
-\includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/netcdf} \caption{A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020).}\label{fig:netcdf}
+\includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/netcdf} \caption{A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020).}\label{fig:netcdf}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/linking} \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/linking-north} \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/linking-lower} \caption{Exploring temperature variation using linking of a map and seasonal display.
+\includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/linking} \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/linking-north} \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/linking-lower} \caption{Exploring temperature variation using linking of a map and seasonal display.
 Each row is a screen dump of the process.
@@
 
-{\centering \includegraphics[width=0.45\linewidth,height=0.25\textheight]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/popup-mid}
+{\centering \includegraphics[width=0.45\linewidth,height=0.25\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/popup-mid}
 
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/diagram-keynotes/diagram-keynotes.005}
+{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.005}
 
@@
 \begin{CSLReferences}{1}{0}
-\leavevmode\hypertarget{ref-leafpop}{}%
-Appelhans, Tim, and Florian Detsch. 2021. \emph{{leafpop}: Include Tables, Images and Graphs in Leaflet Pop-Ups}. \url{https://CRAN.R-project.org/package=leafpop}.
-
-\leavevmode\hypertarget{ref-bach_review_2014}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-leafpop}{}}%
+Appelhans, Tim, and Florian Detsch. 2021. \emph{\pkg{leafpop}: Include Tables, Images and Graphs in Leaflet Pop-Ups}. \url{https://CRAN.R-project.org/package=leafpop}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-bach_review_2014}{}}%
 Bach, Benjamin, Pierre Dragicevic, Dragicevic Archambault, Christophe Hurter, and Sheelagh Carpendale.
@@
 
-\leavevmode\hypertarget{ref-buja1988elements}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-buja1988elements}{}}%
 Buja, Andreas, Daniel Asimov, and Catherine Hurley. 1988. {``Elements of a Viewing Pipeline.''} \emph{Dynamic Graphics Statistics}, 277.
 
-\leavevmode\hypertarget{ref-buja1996interactive}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-buja1996interactive}{}}%
 Buja, Andreas, Dianne Cook, and Deborah F Swayne.
@@
 
-\leavevmode\hypertarget{ref-cheng2016enabling}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-cheng2016enabling}{}}%
 Cheng, Xiaoyue, Dianne Cook, and Heike Hofmann.
@@
 
-\leavevmode\hypertarget{ref-cocchi2019data}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-cocchi2019data}{}}%
 Cocchi, Marina. 2019. \emph{Data Fusion Methodology and Applications}. Elsevier.
 
-\leavevmode\hypertarget{ref-ctvspatiotemporal}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-ctvspatiotemporal}{}}%
 Edzer Pebesma, Roger Bivand. 2022. {``{CRAN} Task View: Handling and Analyzing Spatio-Temporal Data.''} \url{https://CRAN.R-project.org/view=SpatioTemporal}.
 
-\leavevmode\hypertarget{ref-hersbach2020era5}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-hersbach2020era5}{}}%
 Hersbach, Hans, Bill Bell, Paul Berrisford, Shoji Hirahara, András Horányi, Joaquín Muñoz-Sabater, Julien Nicolas, et al.
@@
 
-\leavevmode\hypertarget{ref-ecwmfr}{}%
-Hufkens, Koen, Reto Stauffer, and Elio Campitelli. 2019. {``The {ecwmfr} Package: An Interface to ECMWF API Endpoints.''} \url{https://doi.org/10.5281/zenodo.2647541}.
-
-\leavevmode\hypertarget{ref-lu_multidimensional_2018}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-ecwmfr}{}}%
+Hufkens, Koen, Reto Stauffer, and Elio Campitelli. 2019. {``The \pkg{ecwmfr} Package: An Interface to ECMWF API Endpoints.''} \url{https://doi.org/10.5281/zenodo.2647541}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-lu_multidimensional_2018}{}}%
 Lu, Meng, Marius Appel, and Edzer Pebesma.
@@
 
-\leavevmode\hypertarget{ref-mcintosh2018using}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-mcintosh2018using}{}}%
 McIntosh, Avery I, Helen E Jenkins, Laura F White, Marinus Barnard, Dana R Thomson, Tania Dolby, John Simpson, et al.
@@
 
-\leavevmode\hypertarget{ref-michna2013rnetcdf}{}%
-Michna, Pavel, and Milton Woods. 2013. {``{RNetCDF}: A Package for Reading and Writing NetCDF Datasets.''} \emph{The R Journal} 5 (2): 29--36.
-
-\leavevmode\hypertarget{ref-rnetcdf}{}%
----------. 2021. \emph{{RNetCDF}: Interface to 'NetCDF' Datasets}. \url{https://CRAN.R-project.org/package=RNetCDF}.
-
-\leavevmode\hypertarget{ref-spacetime}{}%
-Pebesma, Edzer. 2012. {``{spacetime}: Spatio-Temporal Data in r.''} \emph{Journal of Statistical Software} 51 (7): 1--30. \url{https://doi.org/10.18637/jss.v051.i07}.
-
-\leavevmode\hypertarget{ref-stars}{}%
----------. 2021. \emph{{stars}: Spatiotemporal Arrays, Raster and Vector Data Cubes}. \url{https://CRAN.R-project.org/package=stars}.
-
-\leavevmode\hypertarget{ref-sf}{}%
-Pebesma, Edzer J. 2018. {``Simple Features for {R}: Standardized Support for Spatial Vector Data.''} \emph{R Journal} 10 (1): 439.
-
-\leavevmode\hypertarget{ref-sp}{}%
-Pebesma, Edzer, and Roger S Bivand. 2005. {``S Classes and Methods for Spatial Data: The {sp} Package.''} \emph{R News} 5 (2): 9--13.
-
-\leavevmode\hypertarget{ref-ncdf4}{}%
-Pierce, David. 2019. \emph{{ncdf4}: Interface to Unidata netCDF (Version 4 or Earlier) Format Data Files}. \url{https://CRAN.R-project.org/package=ncdf4}.
-
-\leavevmode\hypertarget{ref-xts}{}%
-Ryan, Jeffrey A., and Joshua M. Ulrich. 2020. \emph{{xts}: eXtensible Time Series}. \url{https://CRAN.R-project.org/package=xts}.
-
-\leavevmode\hypertarget{ref-simmons2005ecmwf}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-michna2013rnetcdf}{}}%
+Michna, Pavel, and Milton Woods. 2013. {``\pkg{RNetCDF}: A Package for Reading and Writing NetCDF Datasets.''} \emph{The R Journal} 5 (2): 29--36.
+
+\leavevmode\vadjust pre{\hypertarget{ref-rnetcdf}{}}%
+---------. 2021. \emph{\pkg{RNetCDF}: Interface to 'NetCDF' Datasets}. \url{https://CRAN.R-project.org/package=RNetCDF}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-spacetime}{}}%
+Pebesma, Edzer. 2012. {``\pkg{spacetime}: Spatio-Temporal Data in r.''} \emph{Journal of Statistical Software} 51 (7): 1--30. \url{https://doi.org/10.18637/jss.v051.i07}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-stars}{}}%
+---------. 2021. \emph{\pkg{stars}: Spatiotemporal Arrays, Raster and Vector Data Cubes}. \url{https://CRAN.R-project.org/package=stars}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-sf}{}}%
+Pebesma, Edzer J. 2018. {``Simple Features for \proglang{R}: Standardized Support for Spatial Vector Data.''} \emph{R Journal} 10 (1): 439.
+
+\leavevmode\vadjust pre{\hypertarget{ref-sp}{}}%
+Pebesma, Edzer, and Roger S Bivand. 2005. {``S Classes and Methods for Spatial Data: The \pkg{sp} Package.''} \emph{R News} 5 (2): 9--13.
+
+\leavevmode\vadjust pre{\hypertarget{ref-ncdf4}{}}%
+Pierce, David. 2019. \emph{\pkg{ncdf4}: Interface to Unidata netCDF (Version 4 or Earlier) Format Data Files}. \url{https://CRAN.R-project.org/package=ncdf4}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-xts}{}}%
+Ryan, Jeffrey A., and Joshua M. Ulrich. 2020. \emph{\pkg{xts}: eXtensible Time Series}. \url{https://CRAN.R-project.org/package=xts}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-simmons2005ecmwf}{}}%
 Simmons, Adrian, Mariano Hortal, Graeme Kelly, Anthony McNally, Agathe Untch, and Sakari Uppala.
@@
 
-\leavevmode\hypertarget{ref-simmons2020global}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-simmons2020global}{}}%
 Simmons, Adrian, Cornel Soci, Julien Nicolas, Bill Bell, P.
@@
 
-\leavevmode\hypertarget{ref-stuart2010matching}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-stuart2010matching}{}}%
 Stuart, Elizabeth A. 2010. {``Matching Methods for Causal Inference: A Review and a Look Forward.''} \emph{Statistical Science} 25 (1): 1.
 
-\leavevmode\hypertarget{ref-tidync}{}%
-Sumner, Michael. 2020. \emph{{tidync}: A Tidy Approach to 'NetCDF' Data Exploration and Extraction}. \url{https://CRAN.R-project.org/package=tidync}.
-
-\leavevmode\hypertarget{ref-sutherland2000orca}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-tidync}{}}%
+Sumner, Michael. 2020. \emph{\pkg{tidync}: A Tidy Approach to 'NetCDF' Data Exploration and Extraction}. \url{https://CRAN.R-project.org/package=tidync}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-sutherland2000orca}{}}%
 Sutherland, Peter, Anthony Rossini, Thomas Lumley, Nicholas Lewin-Koh, Julie Dickerson, Zach Cox, and Dianne Cook.
 2000.
-{``{Orca}: A Visualization Toolkit for High-Dimensional Data.''} \emph{Journal of Computational and Graphical Statistics} 9 (3): 509--29.
+{``\pkg{Orca}: A Visualization Toolkit for High-Dimensional Data.''} \emph{Journal of Computational and Graphical Statistics} 9 (3): 509--29.
 \url{https://www.tandfonline.com/doi/abs/10.1080/10618600.2000.10474896}.
 
-\leavevmode\hypertarget{ref-wang2020calendar}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-wang2020calendar}{}}%
 Wang, Earo, Dianne Cook, and Rob J Hyndman. 2020. {``Calendar-Based Graphics for Visualizing People's Daily Schedules.''} \emph{Journal of Computational and Graphical Statistics} 29 (3): 490--502.
 
-\leavevmode\hypertarget{ref-tidydata}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-tidydata}{}}%
 Wickham, Hadley. 2014. {``Tidy Data.''} \emph{Journal of Statistical Software} 59 (10): 1--23. \url{https://doi.org/10.18637/jss.v059.i10}.
 
-\leavevmode\hypertarget{ref-ggplot2}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-ggplot2}{}}%
 ---------. 2016. \emph{Ggplot2: Elegant Graphics for Data Analysis}. Springer-Verlag New York. \url{https://ggplot2.tidyverse.org}.
 
-\leavevmode\hypertarget{ref-tidyverse}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-tidyverse}{}}%
 Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy D'Agostino McGowan, Romain François, Garrett Grolemund, et al.
@@
 
-\leavevmode\hypertarget{ref-dplyr}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-dplyr}{}}%
 Wickham, Hadley, Romain François, Lionel Henry, and Kirill Müller. 2022. \emph{Dplyr: A Grammar of Data Manipulation}. \url{https://CRAN.R-project.org/package=dplyr}.
 
-\leavevmode\hypertarget{ref-Wickham2012-yr}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Wickham2012-yr}{}}%
 Wickham, Hadley, Heike Hofmann, Charlotte Wickham, and Dianne Cook. 2012. {``Glyph-Maps for Visually Exploring Temporal Patterns in Climate Data and Models.''} \emph{Environmetrics} 23 (5): 382--93.
 
-\leavevmode\hypertarget{ref-knitr}{}%
-Xie, Yihui. 2015. \emph{Dynamic Documents with {R} and {knitr}}. 2nd ed. Boca Raton, Florida: Chapman; Hall/CRC. \url{https://yihui.name/knitr/}.
-
-\leavevmode\hypertarget{ref-rmarkdown}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-knitr}{}}%
+Xie, Yihui. 2015. \emph{Dynamic Documents with \proglang{R} and \pkg{knitr}}. 2nd ed. Boca Raton, Florida: Chapman; Hall/CRC. \url{https://yihui.name/knitr/}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-rmarkdown}{}}%
 Xie, Yihui, J. J. Allaire, and Garrett Grolemund. 2018. \emph{R Markdown: The Definitive Guide}. Boca Raton, Florida: Chapman; Hall/CRC. \url{https://bookdown.org/yihui/rmarkdown}.
 
-\leavevmode\hypertarget{ref-xie2014reactive}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-xie2014reactive}{}}%
 Xie, Yihui, Heike Hofmann, and Xiaoyue Cheng. 2014. {``{Reactive Programming for Interactive Graphics}.''} \emph{Statistical Science} 29 (2): 201--13. \url{https://doi.org/10.1214/14-STS477}.

### 2022-04-28 1-intro.Rmd
address patricia's comment in intro

@@
 The SpatioTemporal CRAN task view [@ctvspatiotemporal] gathers information about R packages designed for spatio-temporal data and it has a section on *Representing data* that lists existing spatio-temporal data representations used in \proglang{R}.
-Among them, @spacetime proposes four spatio-temporal layouts: full grid layout, sparse grid layout, irregular layout, and trajectory.
-These layouts are implemented in the \pkg{spacetime} [@spacetime] package via the underlying spatial and temporal classes \pkg{sp} [@sp] and \pkg{xts} [@xts].
-The \pkg{stars} [@stars] package has been a recent implementation to handle both raster and vector data using spatio-temporal arrays.
-It interfaces with \pkg{sf} [@sf], which adopts the simple features representation of spatial data, and \pkg{tidyverse} [@tidyverse], a suite of tools to wrangle and visualise data in \proglang{R}.
-
-**The paragraph above needs a bit more flesh**
+Among them, @spacetime summarises spatio-temporal data into three forms: time-wide, space-wide, and long formats.
+The associated package \pkg{spacetime} [@spacetime] implements four spatio-temporal layouts (full grid, sparse grid, irregular, and trajectory) to handle different space and time combinations.
+The \pkg{stars} [@stars] package has been a recent implementation to use dense arrays to represent spatio-temporal cubes.
+It also interfaces with \pkg{sf} [@sf], a package commonly used for wrangling spatial data, and the \pkg{tidyverse} [@tidyverse] suite for general data wrangling and visualisation in \proglang{R}.
 
 Still, the data representation for spatio-temporal data can be further improved and there are two reasons for this.
-Firstly, the raw data sourced in the wild is less often presented in any one of the representations above (Which representations - in the previous paragraph there are more than two?), and fitting the raw data into a data object can sometimes be difficult.
+Firstly, the raw data sourced in the wild is less often presented in any one of the layouts above, and fitting the raw data into a data object can sometimes be difficult.
 More often, spatio-temporal data are collected in separate 2D tables and analysts need to assemble them into a whole piece before exploring the data.
-Examples of components of spatio-temporal data can be 1) map data recording the shape of a collection of areas of interest; 2) geo-scientific data recording the longitude and latitude coordinates of locations in the area along with other metadata related to the location, and; 3) temporal variables indexed by both location and time.
+Examples of components of spatio-temporal data can be 1) areal data recording the shape of a collection of areas of interest; 2) geostatistical data recording the longitude and latitude coordinates of locations, potentially also with other metadata related to the location, and; 3) temporal data of each location across time.
 
-**I would aligned the examples of spatio-temporal data with those discuss in the st books. See Paula Moraga for example**
+<!-- **I would aligned the examples of spatio-temporal data with those discuss in the st books. See Paula Moraga for example** -->
 
@@
 In the package, a new data structure, also called \code{cubble}, is proposed to organise spatial and temporal variables as two forms of a single data object so that they can be wrangled separately or combined while being kept synchronised.
-Among the four spacetime layouts in @spacetime, \pkg{cubble} can be applied to full di layout, sparse grid layout, or irregular layout, but not trajectory, which is outside the scope of this work.
+Among the four spacetime layouts in @spacetime, \pkg{cubble} can be applied to full grid, sparse grid, or irregular, but not trajectory, which is outside the scope of this work.
 The software is available from the Comprehensive R Archive Network (CRAN) at https://CRAN.R-project.org/package=cubble.

### 2022-04-28 arxiv/cubble-arxiv.tex
address patricia's comment in intro

@@
 Among them, E.
-Pebesma (\protect\hyperlink{ref-spacetime}{2012}) proposes four spatio-temporal layouts: full grid layout, sparse grid layout, irregular layout, and trajectory.
-These layouts are implemented in the \pkg{spacetime} (\protect\hyperlink{ref-spacetime}{E.
-Pebesma 2012}) package via the underlying spatial and temporal classes \pkg{sp} (\protect\hyperlink{ref-sp}{E.
-Pebesma and Bivand 2005}) and \pkg{xts} (\protect\hyperlink{ref-xts}{Ryan and Ulrich 2020}).
+Pebesma (\protect\hyperlink{ref-spacetime}{2012}) summarises spatio-temporal data into three forms: time-wide, space-wide, and long formats.
+The associated package \pkg{spacetime} (\protect\hyperlink{ref-spacetime}{E.
+Pebesma 2012}) implements four spatio-temporal layouts (full grid, sparse grid, irregular, and trajectory) to handle different space and time combinations.
 The \pkg{stars} (\protect\hyperlink{ref-stars}{E.
-Pebesma 2021}) package has been a recent implementation to handle both raster and vector data using spatio-temporal arrays.
-It interfaces with \pkg{sf} (\protect\hyperlink{ref-sf}{E.
+Pebesma 2021}) package has been a recent implementation to use dense arrays to represent spatio-temporal cubes.
+It also interfaces with \pkg{sf} (\protect\hyperlink{ref-sf}{E.
 J.
-Pebesma 2018}), which adopts the simple features representation of spatial data, and \pkg{tidyverse} (\protect\hyperlink{ref-tidyverse}{Wickham et al.
-2019}), a suite of tools to wrangle and visualise data in \proglang{R}.
-
-\textbf{The paragraph above needs a bit more flesh}
+Pebesma 2018}), a package commonly used for wrangling spatial data, and the \pkg{tidyverse} (\protect\hyperlink{ref-tidyverse}{Wickham et al.
+2019}) suite for general data wrangling and visualisation in \proglang{R}.
 
 Still, the data representation for spatio-temporal data can be further improved and there are two reasons for this.
-Firstly, the raw data sourced in the wild is less often presented in any one of the representations above (Which representations - in the previous paragraph there are more than two?), and fitting the raw data into a data object can sometimes be difficult.
+Firstly, the raw data sourced in the wild is less often presented in any one of the layouts above, and fitting the raw data into a data object can sometimes be difficult.
 More often, spatio-temporal data are collected in separate 2D tables and analysts need to assemble them into a whole piece before exploring the data.
-Examples of components of spatio-temporal data can be 1) map data recording the shape of a collection of areas of interest; 2) geo-scientific data recording the longitude and latitude coordinates of locations in the area along with other metadata related to the location, and; 3) temporal variables indexed by both location and time.
-
-\textbf{I would aligned the examples of spatio-temporal data with those discuss in the st books. See Paula Moraga for example}
+Examples of components of spatio-temporal data can be 1) areal data recording the shape of a collection of areas of interest; 2) geostatistical data recording the longitude and latitude coordinates of locations, potentially also with other metadata related to the location, and; 3) temporal data of each location across time.
 
@@
 Among the four spacetime layouts in E.
-Pebesma (\protect\hyperlink{ref-spacetime}{2012}), \pkg{cubble} can be applied to full di layout, sparse grid layout, or irregular layout, but not trajectory, which is outside the scope of this work.
+Pebesma (\protect\hyperlink{ref-spacetime}{2012}), \pkg{cubble} can be applied to full grid, sparse grid, or irregular, but not trajectory, which is outside the scope of this work.
 The software is available from the Comprehensive R Archive Network (CRAN) at \url{https://CRAN.R-project.org/package=cubble}.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-sp}{}}%
-Pebesma, Edzer, and Roger S Bivand. 2005. {``S Classes and Methods for Spatial Data: The \pkg{sp} Package.''} \emph{R News} 5 (2): 9--13.
-
 \leavevmode\vadjust pre{\hypertarget{ref-ncdf4}{}}%
 Pierce, David. 2019. \emph{\pkg{ncdf4}: Interface to Unidata netCDF (Version 4 or Earlier) Format Data Files}. \url{https://CRAN.R-project.org/package=ncdf4}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-xts}{}}%
-Ryan, Jeffrey A., and Joshua M. Ulrich. 2020. \emph{\pkg{xts}: eXtensible Time Series}. \url{https://CRAN.R-project.org/package=xts}.
 

### 2022-04-28 2-the-cubble-pkg.Rmd
misc small changes
address ursula on nest_by

--- 2022-04-26 2-the-cubble-pkg.Rmd
@@
 
-Each core function is introduced accompanied with a short example and the data `climate_flat` is used throughout this section.
+Each core \pkg{cubble} function is introduced and accompanied with a short example using the data `climate_flat`  throughout this section.
 `climate_flat` contains five weather stations in Australia with spatial information of each station: station id, latitude, longitude, elevation, station name, World Meteorology Organisation ID and also daily temporal information of date, maximum and minimum temperature and precipitation for 2020.
@@
 
-\textcolor{blue}{Ursula: I feel like this doesn't get the message accross, the main point seems to be that with cubble we can switch and never have to add information back in "by-hand" (with some join function), maybe make the point without refering to specific functions first and then explain the example?}
+Some readers may question why a new data structure is needed rather than using \code{dplyr::nest_by()} on the combined data to directly create a list-column.
+The reason is that cubble is specifically designed to utilize the spatio-temporal structure when arranging variables from different observational units into a single object.
+And it enables easy pivoting between purely spatial, purely temporal, or unfolded into a combined form.
 
-Some readers may raise the question why a new data structure rather than using \code{dplyr::nest_by()} on the combined data to directly create a list-column.
+<!-- Some readers may raise the question why a new data structure rather than using \code{dplyr::nest_by()} on the combined data to directly create a list-column.
 While \code{nest_by()} and its inverse \code{unnest()} can mimic the nested form in cubble, cubble can be seen as an attempt to arrange variables from different observational units into a single object, specifically for spatio-temporal data.
-While `nest_by()` can recreate the nested form, it cannot create a clean slate for temporal variables as used in the long form.
-For temporal variables, carrying the duplicated spatial variables in the combined form is not convenient to work with.
-This is because these spatial variables will be lost after \code{summarise()} and one will need to subset the spatial variables again from the original data and make a left join, which tends to create unnecessary intermediate products in the workflow.
-
-```{r eval = FALSE, echo = TRUE}
-# operation on the long form w/o cubble
-# cubble workflow
-dt |>
-  as_cubble(...) |>
-  face_temporal() |>
-  group_by() |>
-  summarise() |>
-  unfold(...)
-
-# tidyverse workflow
-temporal <- dt |>
-  group_by(...) |>
-  summarise(...)
-
-spatial <- dt |> select(....)
-
-spatial |> left_join(temporal)
-```
-
-<!-- ```{r} -->
-<!-- library(dplyr) -->
-<!-- vars <- c(id, lat, long, elev, name, wmo_id) -->
-<!-- left_join(cubble::stations, cubble::climate) |>  -->
-<!--   group_by(id, lat, long, elev, name, wmo_id) |>  -->
-<!--   summarise(tmax_mean = mean(tmax)) -->
-
-<!-- cubble::climate |>  -->
-<!--   mutate(month = lubridate::month(date)) |>  -->
-<!--   group_by(id, month) |>  -->
-<!--   summarise(tmax_mean = mean(tmax)) |>  -->
-<!--   ungroup() |>  -->
-<!--   left_join(cubble::stations) |>  -->
-<!--   nest_by(id, lat, long, elev, name, wmo_id) -->
-
-<!-- ``` --
-
-<!-- ``` r -->
-<!-- library(cubble) -->
-<!-- library(tidyverse) -->
-<!-- left_join(cubble::stations, cubble::climate) |> -->
-<!--   dplyr::nest_by(id, lat,  long,  elev, name, wmo_id) |> -->
-<!--   tidyr::unnest() |> -->
-<!--   group_by(........, geometry) -->
-<!-- ``` -->
+The benefit of arranging everything in a single object is that there is no need to manually split and join variables in the data.
+For various spatial, temporal, or spatio-temporal tasks, the data format required can alternate among a pure spatial, pure temporal, or a combined format.
+Analysts will need to manually split the combined data into a spatial or temporal one, or join them into a combined one.
+\pkg{Cubble} simplifies this switch of format with a nested and long form and allowing spatial variables to be unfolded into the long form for a combined data all in a single data object.
+-->
 

### 2022-04-28 5-conclusion.Rmd
elaborate on future direction

@@
 In the spatial aspect, the simple feature representation wraps spatial coordinates in a list column, for some complex spatial operations, while sometimes, analysts may want to unpack them into longitude and latitude columns to work from.
-Another extension to \pkg{cubble} is to build a smooth transition between the coordinate columns and simple feature geometry column.
+A smooth transition between the coordinate and simple feature representation can be added to \pkg{cubble}.
+\pkg{cubble} can also be extended to handle sites with moving coordinates and this would involve constructing a \code{sp} column for locations, similar to \code{ts} for time, and a form locations can be pivoted into, like the long form for temporal variables.
+In the multivariate aspect, \pkg{cubble} can also be extended with interface to more high dimensional visualisation methods, i.e., the tour method, to understand variable importance or comparing location similarities.
 

### 2022-04-28 arxiv/cubble-arxiv.tex
elaborate on future direction

@@
 
-\textcolor{blue}{Ursula: I feel like this doesn't get the message accross, the main point seems to be that with cubble we can switch and never have to add information back in "by-hand" (with some join function), maybe make the point without refering to specific functions first and then explain the example?}
-
-Some readers may raise the question why a new data structure rather than using \code{dplyr::nest_by()} on the combined data to directly create a list-column.
-While \code{nest_by()} and its inverse \code{unnest()} can mimic the nested form in cubble, cubble can be seen as an attempt to arrange variables from different observational units into a single object, specifically for spatio-temporal data.
-While \texttt{nest\_by()} can recreate the nested form, it cannot create a clean slate for temporal variables as used in the long form.
-For temporal variables, carrying the duplicated spatial variables in the combined form is not convenient to work with.
-This is because these spatial variables will be lost after \code{summarise()} and one will need to subset the spatial variables again from the original data and make a left join, which tends to create unnecessary intermediate products in the workflow.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# operation on the long form w/o cubble}
-\CommentTok{\# cubble workflow}
-\NormalTok{dt }\SpecialCharTok{|\textgreater{}}
-  \FunctionTok{as\_cubble}\NormalTok{(...) }\SpecialCharTok{|\textgreater{}}
-  \FunctionTok{face\_temporal}\NormalTok{() }\SpecialCharTok{|\textgreater{}}
-  \FunctionTok{group\_by}\NormalTok{() }\SpecialCharTok{|\textgreater{}}
-  \FunctionTok{summarise}\NormalTok{() }\SpecialCharTok{|\textgreater{}}
-  \FunctionTok{unfold}\NormalTok{(...)}
-
-\CommentTok{\# tidyverse workflow}
-\NormalTok{temporal }\OtherTok{\textless{}{-}}\NormalTok{ dt }\SpecialCharTok{|\textgreater{}}
-  \FunctionTok{group\_by}\NormalTok{(...) }\SpecialCharTok{|\textgreater{}}
-  \FunctionTok{summarise}\NormalTok{(...)}
-
-\NormalTok{spatial }\OtherTok{\textless{}{-}}\NormalTok{ dt }\SpecialCharTok{|\textgreater{}} \FunctionTok{select}\NormalTok{(....)}
-
-\NormalTok{spatial }\SpecialCharTok{|\textgreater{}} \FunctionTok{left\_join}\NormalTok{(temporal)}
-\end{Highlighting}
-\end{Shaded}
+Some readers may question why a new data structure is needed rather than using \code{dplyr::nest_by()} on the combined data to directly create a list-column.
+The reason is that cubble is specifically designed to utilize the spatio-temporal structure when arranging variables from different observational units into a single object.
+And it enables easy pivoting between purely spatial, purely temporal, or unfolded into a combined form.
 
@@
 ## Dimension:     XY
-## Bounding box:  xmin: 142.3643 ymin: -38.67876 xmax: 147.386 ymax: -36.39269
+## Bounding box:  xmin: 142.3535 ymin: -38.67876 xmax: 147.3909 ymax: -36.39269
 ## Geodetic CRS:  WGS 84
 ##     lga_code_2018            lga state_code_2016 state_name_2016 areasqkm_2018
-## 1         20110     Alpine (S)               2        Victoria     4788.1568
-## 2         20260    Ararat (RC)               2        Victoria     4211.1171
-## 3         20570   Ballarat (C)               2        Victoria      739.0321
-## 4         20660    Banyule (C)               2        Victoria       62.5402
-## 5         20740 Bass Coast (S)               2        Victoria      865.8095
+## 132         20110     Alpine (S)               2        Victoria     4788.1568
+## 133         20260    Ararat (RC)               2        Victoria     4211.1171
+## 134         20570   Ballarat (C)               2        Victoria      739.0321
+## 135         20660    Banyule (C)               2        Victoria       62.5402
+## 136         20740 Bass Coast (S)               2        Victoria      865.8095
 ##     cent_long  cent_lat                       geometry
-## 1  146.9742 -36.85357 MULTIPOLYGON (((146.7258 -3...
-## 2  142.8432 -37.47271 MULTIPOLYGON (((143.1807 -3...
-## 3  143.7815 -37.49286 MULTIPOLYGON (((143.6622 -3...
-## 4  145.0851 -37.73043 MULTIPOLYGON (((145.1357 -3...
-## 5  145.5581 -38.50730 MULTIPOLYGON (((145.5207 -3...
+## 132  146.9742 -36.85357 MULTIPOLYGON (((146.7258 -3...
+## 133  142.8432 -37.47271 MULTIPOLYGON (((143.1807 -3...
+## 134  143.7815 -37.49286 MULTIPOLYGON (((143.6622 -3...
+## 135  145.0851 -37.73043 MULTIPOLYGON (((145.1357 -3...
+## 136  145.5581 -38.50730 MULTIPOLYGON (((145.5207 -3...
 \end{verbatim}

### 2022-04-29 2-0-cube.Rmd
update Figure 1

@@
 The most common data cube using the three axes (time, latitude, longitude)  could be considered as taking snapshots of the space and stacking on time.
-Here we will define the spatio-temporal cube slightly differently with the three axes (time, site, variable), as illustrated in Figure \ref{fig:cube-def}.
+Here we will define the spatio-temporal cube slightly differently with the three axes (time, site, variable), as illustrated in the leftmost cube in Figure \ref{fig:cube-def}.
 The time axis is the same in both versions, while the site axis now captures both latitude and longitude.
@@
 This notion is adopted to avoid using hypercubes when describing multivariate spatio-temporal data and is the conceptual framework behind \pkg{cubble}.
+With this conceptual model, operations on spatio-temporal data can be mapped to operations on the cube and the rest of Figure \ref{fig:cube-def} show examples of slicing on site, time, and variable.
 
@@
 
-```{r cube-def, echo = FALSE, fig.align="center", out.width = "50%", fig.cap = "An illustration of the conceptualised spatio-temporal cube with three axes: time, site, and variable."}
-knitr::include_graphics(here::here("figures/cube-def.png"))
+```{r cube-def, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "An illustration of the conceptual spatio-temporal cube with different slicing on time, site, and variable.
+For each axis, the slicing can be on a single value or a set of values."}
+knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.007.png"))
 ```

### 2022-04-29 5-conclusion.Rmd
add highlight of examples in conclusion

@@
 The data structure is also flexible to the techniques and packages analysts use to analyse the data, for example, in the matching example in section 4.3, users are free to use algorithms from another package to cluster stations.
+
+The data structure and software are demonstrated with extensive examples.
+These include creating and coercing wild-caught data with potential mismatch on sites, handling hierarchical data, matching series spatially and temporally, reproducing ERA5 results from NetCDF data, and, lastly fitting into the interactive graphic pipeline of \pkg{plotly} and \pkg{leaflet}.
 

### 2022-04-29 arxiv/cubble-arxiv.tex
add highlight of examples in conclusion

@@
 The most common data cube using the three axes (time, latitude, longitude) could be considered as taking snapshots of the space and stacking on time.
-Here we will define the spatio-temporal cube slightly differently with the three axes (time, site, variable), as illustrated in Figure \ref{fig:cube-def}.
+Here we will define the spatio-temporal cube slightly differently with the three axes (time, site, variable), as illustrated in the leftmost cube in Figure \ref{fig:cube-def}.
 The time axis is the same in both versions, while the site axis now captures both latitude and longitude.
@@
 
-{\centering \includegraphics[width=0.5\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/cube-def}
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.007}
 
@@
 
-\caption{An illustration of the conceptualised spatio-temporal cube with three axes: time, site, and variable.}\label{fig:cube-def}
+\caption{An illustration of the conceptual spatio-temporal cube with different slicing on time, site, and variable.
+For each axis, the slicing can be on a single value or a set of values.}\label{fig:cube-def}
 \end{figure}
@@
 
+The data structure and software are demonstrated with extensive examples.
+These include creating and coercing wild-caught data with potential mismatch on sites, handling hierarchical data, matching series spatially and temporally, reproducing ERA5 results from NetCDF data, and, lastly fitting into the interactive graphic pipeline of \pkg{plotly} and \pkg{leaflet}.
+
 Future directions of the package involves responding to changes in the tidyverse packages that \pkg{cubble} imports, in particular, \pkg{tibble}, \pkg{tidyr}, and \pkg{dplyr}.

### 2022-04-30 arxiv/cubble-arxiv.Rmd
Updating abstract

@@
 However for some studies, it is important to analyse different aspects of the spatio-temporal data simultaneouly, like for instance, temporal trends of multiple variables across locations.
-In order to facilitate the study of different sections or combinations of spatio-temporal data, we introduce a new data structure, \code{cubble}, with a suite of functions enabling easy slicing and dicing on the different components spatio-temporal components, for visual and numerical explorations while facilitating data wrangling for modelling.
-The proposed \code{cubble} structure ensures that all the components of the data are easy to access and manipulate while providing flexibility for data analysis.
-The \code{cubble} structure and the tools provided in the \code{cubble} R package equip users with the capability to handle hierarchical spatial and temporal structures.
-The tools are illustrated with examples of Australian climate data, merging climate and river level data sources, reproducing a climate reanalysis (ERA5), and creating interactive graphics.
+In order to facilitate the study of different portions or combinations of spatio-temporal data, we introduce a new data structure, \code{cubble}, with a suite of functions enabling easy slicing and dicing on the different components spatio-temporal components.
+The proposed \code{cubble} structure ensures that all the components of the data are easy to access and manipulate while providing flexibility for data analysis.In addition,   \code{cubble} facilitates visual and numerical explorations of the data while easing data wrangling and modelling.
+The \code{cubble} structure and the functions provided in the \code{cubble} R package equip users with the capability to handle hierarchical spatial and temporal structures.
+The \code{cubble} structure and the tools implemented in the package are illustrated with different examples of Australian climate data.
 keywords:

### 2022-04-30 arxiv/cubble-arxiv.tex
Updating abstract

@@
 However for some studies, it is important to analyse different aspects of the spatio-temporal data simultaneouly, like for instance, temporal trends of multiple variables across locations.
-In order to facilitate the study of different sections or combinations of spatio-temporal data, we introduce a new data structure, \code{cubble}, with a suite of functions enabling easy slicing and dicing on the different components spatio-temporal components, for visual and numerical explorations while facilitating data wrangling for modelling.
-The proposed \code{cubble} structure ensures that all the components of the data are easy to access and manipulate while providing flexibility for data analysis.
-The \code{cubble} structure and the tools provided in the \code{cubble} R package equip users with the capability to handle hierarchical spatial and temporal structures.
-The tools are illustrated with examples of Australian climate data, merging climate and river level data sources, reproducing a climate reanalysis (ERA5), and creating interactive graphics.
+In order to facilitate the study of different portions or combinations of spatio-temporal data, we introduce a new data structure, \code{cubble}, with a suite of functions enabling easy slicing and dicing on the different components spatio-temporal components.
+The proposed \code{cubble} structure ensures that all the components of the data are easy to access and manipulate while providing flexibility for data analysis.In addition, \code{cubble} facilitates visual and numerical explorations of the data while easing data wrangling and modelling.
+The \code{cubble} structure and the functions provided in the \code{cubble} R package equip users with the capability to handle hierarchical spatial and temporal structures.
+The \code{cubble} structure and the tools implemented in the package are illustrated with different examples of Australian climate data.
 \end{abstract}
@@
 The SpatioTemporal CRAN task view (\protect\hyperlink{ref-ctvspatiotemporal}{Edzer Pebesma 2022}) gathers information about R packages designed for spatio-temporal data and it has a section on \emph{Representing data} that lists existing spatio-temporal data representations used in \proglang{R}.
-Among them, E.
-Pebesma (\protect\hyperlink{ref-spacetime}{2012}) summarises spatio-temporal data into three forms: time-wide, space-wide, and long formats.
+Among them, \protect\hyperlink{ref-spacetime}{E.
+Pebesma} (\protect\hyperlink{ref-spacetime}{2012}) summarises spatio-temporal data into three forms: time-wide, space-wide, and long formats.
 The associated package \pkg{spacetime} (\protect\hyperlink{ref-spacetime}{E.
@@
 In the package, a new data structure, also called \code{cubble}, is proposed to organise spatial and temporal variables as two forms of a single data object so that they can be wrangled separately or combined while being kept synchronised.
-Among the four spacetime layouts in E.
-Pebesma (\protect\hyperlink{ref-spacetime}{2012}), \pkg{cubble} can be applied to full grid, sparse grid, or irregular, but not trajectory, which is outside the scope of this work.
+Among the four spacetime layouts in \protect\hyperlink{ref-spacetime}{E.
+Pebesma} (\protect\hyperlink{ref-spacetime}{2012}), \pkg{cubble} can be applied to full grid, sparse grid, or irregular, but not trajectory, which is outside the scope of this work.
 The software is available from the Comprehensive R Archive Network (CRAN) at \url{https://CRAN.R-project.org/package=cubble}.
@@
 Spatio-temporal data can be conceptualised using a cubical data model with three axes, typically time, latitude and longitude.
-This abstraction can be useful for generalising operations and visualisation purposes: Lu, Appel, and Pebesma (\protect\hyperlink{ref-lu_multidimensional_2018}{2018}) shows how array operations (select, scale, reduce, rearrange, and compute) can be mapped onto the cube; Bach et al.
-(\protect\hyperlink{ref-bach_review_2014}{2014}) reviews the temporal data visualisation based on space-time cube operations.
-Notice that the term space-time cube in their article ``does not need to involve spatial data'', but refers to ``an abstract 2D substrate that is used to visualize data at a specific time''.
-Despite its main focus being on temporal data, the mindset of abstracting out data representation to construct visualisation still applies to our spatio-temporal data manipulation and visualisation.
+This abstraction can be useful for generalising operations and visualisation purposes: \protect\hyperlink{ref-lu_multidimensional_2018}{Lu, Appel, and Pebesma} (\protect\hyperlink{ref-lu_multidimensional_2018}{2018}) shows how array operations (select, scale, reduce, rearrange, and compute) can be mapped onto the cube; \protect\hyperlink{ref-bach_review_2014}{Bach et al.} (\protect\hyperlink{ref-bach_review_2014}{2014}) reviews the temporal data visualisation based on space-time cube operations.
+Notice that the term space-time cube in their article ``does not need to involve spatial data,'' but refers to ``an abstract 2D substrate that is used to visualize data at a specific time.'' Despite its main focus being on temporal data, the mindset of abstracting out data representation to construct visualisation still applies to our spatio-temporal data manipulation and visualisation.
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.007}
+{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/diagram-keynotes/diagram-keynotes.007}
 
@@
 \begin{Highlighting}[]
-\NormalTok{cubble\_nested }\SpecialCharTok{|\textgreater{}}
+\NormalTok{cubble\_nested }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
   \FunctionTok{mutate}\NormalTok{(}\AttributeTok{rain\_day =} \FunctionTok{sum}\NormalTok{(ts}\SpecialCharTok{$}\NormalTok{prcp }\SpecialCharTok{!=} \DecValTok{0}\NormalTok{))}
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.008}
+{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/diagram-keynotes/diagram-keynotes.008}
 
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.003}
+{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/diagram-keynotes/diagram-keynotes.003}
 
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/illu-matching} \caption{An illustration of temporal matching in cubble.
+\includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/illu-matching} \caption{An illustration of temporal matching in cubble.
 Three highest peaks are identified in each series and intervals are constructed on series \code{A}.
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.004}
+{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/diagram-keynotes/diagram-keynotes.004}
 
@@
 In \proglang{R}, \pkg{GGally} implements glyph maps through the \code{glyphs()} function.
-The function constructs a data frame with calculated position (\code{gx}, \code{gy}, \code{gid}) of each point on the time series using linear algebra (Equation 1 and 2 in Wickham et al.
-(\protect\hyperlink{ref-Wickham2012-yr}{2012})).
+The function constructs a data frame with calculated position (\code{gx}, \code{gy}, \code{gid}) of each point on the time series using linear algebra (Equation 1 and 2 in \protect\hyperlink{ref-Wickham2012-yr}{Wickham et al.} (\protect\hyperlink{ref-Wickham2012-yr}{2012})).
 The data can then be piped into \code{ggplot} to create the glyph map as:
@@
 \begin{Highlighting}[]
-\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|\textgreater{}}
-  \FunctionTok{face\_temporal}\NormalTok{() }\SpecialCharTok{|\textgreater{}}
+\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
+  \FunctionTok{face\_temporal}\NormalTok{() }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
   \FunctionTok{filter}\NormalTok{(lubridate}\SpecialCharTok{::}\FunctionTok{year}\NormalTok{(date) }\SpecialCharTok{\%in\%} \FunctionTok{c}\NormalTok{(}\DecValTok{1971}\SpecialCharTok{:}\DecValTok{1975}\NormalTok{, }\DecValTok{2016}\SpecialCharTok{:}\DecValTok{2020}\NormalTok{))}
@@
 \begin{Highlighting}[]
-\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|\textgreater{}}
+\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
   \FunctionTok{group\_by}\NormalTok{(}\AttributeTok{month =}\NormalTok{ lubridate}\SpecialCharTok{::}\FunctionTok{month}\NormalTok{(date),}
          \AttributeTok{group =} \FunctionTok{as.factor}\NormalTok{(}\FunctionTok{ifelse}\NormalTok{(lubridate}\SpecialCharTok{::}\FunctionTok{year}\NormalTok{(date) }\SpecialCharTok{\textgreater{}} \DecValTok{2015}\NormalTok{,}
-                                  \StringTok{"2016 \textasciitilde{} 2020"}\NormalTok{, }\StringTok{"1971 \textasciitilde{} 1975"}\NormalTok{))) }\SpecialCharTok{|\textgreater{}}
+                                  \StringTok{"2016 \textasciitilde{} 2020"}\NormalTok{, }\StringTok{"1971 \textasciitilde{} 1975"}\NormalTok{))) }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
   \FunctionTok{summarise}\NormalTok{(}\AttributeTok{tmax =} \FunctionTok{mean}\NormalTok{(tmax, }\AttributeTok{na.rm =} \ConstantTok{TRUE}\NormalTok{))}
@@
 \begin{Highlighting}[]
-\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|\textgreater{}} \FunctionTok{face\_spatial}\NormalTok{() }\SpecialCharTok{|\textgreater{}} \FunctionTok{filter}\NormalTok{(}\FunctionTok{nrow}\NormalTok{(ts) }\SpecialCharTok{==} \DecValTok{24}\NormalTok{)}
+\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{face\_spatial}\NormalTok{() }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{filter}\NormalTok{(}\FunctionTok{nrow}\NormalTok{(ts) }\SpecialCharTok{==} \DecValTok{24}\NormalTok{)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|\textgreater{}} \FunctionTok{face\_temporal}\NormalTok{() }\SpecialCharTok{|\textgreater{}} \FunctionTok{unfold}\NormalTok{(latitude, longitude)}
+\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{face\_temporal}\NormalTok{() }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{unfold}\NormalTok{(latitude, longitude)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{nsw\_vic }\OtherTok{\textless{}{-}}\NormalTok{ ozmaps}\SpecialCharTok{::}\NormalTok{abs\_ste }\SpecialCharTok{|\textgreater{}}
+\NormalTok{nsw\_vic }\OtherTok{\textless{}{-}}\NormalTok{ ozmaps}\SpecialCharTok{::}\NormalTok{abs\_ste }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
   \FunctionTok{filter}\NormalTok{(NAME }\SpecialCharTok{\%in\%} \FunctionTok{c}\NormalTok{(}\StringTok{"Victoria"}\NormalTok{, }\StringTok{"New South Wales"}\NormalTok{))}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth,height=0.7\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/basic-manip} \caption{A glyph map of the average maximum temperature by month of Victoria and New South Wales weather stations in Australia.
+\includegraphics[width=1\linewidth,height=0.7\textheight]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/basic-manip} \caption{A glyph map of the average maximum temperature by month of Victoria and New South Wales weather stations in Australia.
 On the top left corner is an insetted plot of station Cobar highlighted in the black box.}\label{fig:basic-manip}
@@
 \begin{Highlighting}[]
-\NormalTok{station\_nested }\OtherTok{\textless{}{-}}\NormalTok{ weatherdata}\SpecialCharTok{::}\NormalTok{climate\_full }\SpecialCharTok{|\textgreater{}}
+\NormalTok{station\_nested }\OtherTok{\textless{}{-}}\NormalTok{ weatherdata}\SpecialCharTok{::}\NormalTok{climate\_full }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
   \FunctionTok{mutate}\NormalTok{(}\AttributeTok{cluster =}\NormalTok{ ...)}
@@
 \begin{Highlighting}[]
-\NormalTok{cluster\_nested }\OtherTok{\textless{}{-}}\NormalTok{ station\_nested }\SpecialCharTok{|\textgreater{}} \FunctionTok{switch\_key}\NormalTok{(cluster)}
+\NormalTok{cluster\_nested }\OtherTok{\textless{}{-}}\NormalTok{ station\_nested }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{switch\_key}\NormalTok{(cluster)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{cluster\_nested }\OtherTok{\textless{}{-}}\NormalTok{ cluster\_nested }\SpecialCharTok{|\textgreater{}} \FunctionTok{get\_centroid}\NormalTok{()}
+\NormalTok{cluster\_nested }\OtherTok{\textless{}{-}}\NormalTok{ cluster\_nested }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{get\_centroid}\NormalTok{()}
 \end{Highlighting}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/basic-agg} \caption{Profile of aggregated precipitation from 639 weather stations in Australia.
+\includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/basic-agg} \caption{Profile of aggregated precipitation from 639 weather stations in Australia.
 Subplot A shows the glyph map of the weekly averaged precipitation of each cluster.
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/matching-map} \caption{Location of weather stations and river gauges in Victoria, Australia.}\label{fig:matching-map}
+\includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/matching-map} \caption{Location of weather stations and river gauges in Victoria, Australia.}\label{fig:matching-map}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/matching} \caption{Matched weather stations and river gauges on the map (A) and across time (B).
+\includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/matching} \caption{Matched weather stations and river gauges on the map (A) and across time (B).
 Precipitation and water level have been standardised between 0 and 1 to be displayed on the same scale.
@@
 
-Figure \ref{fig:netcdf} reproduces the ERA5 data row of Figure 19 in Hersbach et al.
-(\protect\hyperlink{ref-hersbach2020era5}{2020}).
+Figure \ref{fig:netcdf} reproduces the ERA5 data row of Figure 19 in \protect\hyperlink{ref-hersbach2020era5}{Hersbach et al.} (\protect\hyperlink{ref-hersbach2020era5}{2020}).
 It shows the southern polar vortex splits into two on 2002-09-26 and further splits into four on 2002-10-04 in the stratosphere.
-Readers interested in the analysis of this figure can refer to Hersbach et al.
-(\protect\hyperlink{ref-hersbach2020era5}{2020}), Simmons et al.
-(\protect\hyperlink{ref-simmons2020global}{2020}), and Simmons et al.
-(\protect\hyperlink{ref-simmons2005ecmwf}{2005}) for more details.
+Readers interested in the analysis of this figure can refer to \protect\hyperlink{ref-hersbach2020era5}{Hersbach et al.} (\protect\hyperlink{ref-hersbach2020era5}{2020}), \protect\hyperlink{ref-simmons2020global}{Simmons et al.} (\protect\hyperlink{ref-simmons2020global}{2020}), and \protect\hyperlink{ref-simmons2005ecmwf}{Simmons et al.} (\protect\hyperlink{ref-simmons2005ecmwf}{2005}) for more details.
 
 \begin{figure}
-\includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/netcdf} \caption{A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020).}\label{fig:netcdf}
+\includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/netcdf} \caption{A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020).}\label{fig:netcdf}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/linking} \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/linking-north} \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/linking-lower} \caption{Exploring temperature variation using linking of a map and seasonal display.
+\includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/linking} \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/linking-north} \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/linking-lower} \caption{Exploring temperature variation using linking of a map and seasonal display.
 Each row is a screen dump of the process.
@@
 
-{\centering \includegraphics[width=0.45\linewidth,height=0.25\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/popup-mid}
+{\centering \includegraphics[width=0.45\linewidth,height=0.25\textheight]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/popup-mid}
 
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.005}
+{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/diagram-keynotes/diagram-keynotes.005}
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-michna2013rnetcdf}{}}%
-Michna, Pavel, and Milton Woods. 2013. {``\pkg{RNetCDF}: A Package for Reading and Writing NetCDF Datasets.''} \emph{The R Journal} 5 (2): 29--36.
-
-\leavevmode\vadjust pre{\hypertarget{ref-rnetcdf}{}}%
----------. 2021. \emph{\pkg{RNetCDF}: Interface to 'NetCDF' Datasets}. \url{https://CRAN.R-project.org/package=RNetCDF}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-spacetime}{}}%
-Pebesma, Edzer. 2012. {``\pkg{spacetime}: Spatio-Temporal Data in r.''} \emph{Journal of Statistical Software} 51 (7): 1--30. \url{https://doi.org/10.18637/jss.v051.i07}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-stars}{}}%
----------. 2021. \emph{\pkg{stars}: Spatiotemporal Arrays, Raster and Vector Data Cubes}. \url{https://CRAN.R-project.org/package=stars}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-sf}{}}%
-Pebesma, Edzer J. 2018. {``Simple Features for \proglang{R}: Standardized Support for Spatial Vector Data.''} \emph{R Journal} 10 (1): 439.
-
-\leavevmode\vadjust pre{\hypertarget{ref-ncdf4}{}}%
-Pierce, David. 2019. \emph{\pkg{ncdf4}: Interface to Unidata netCDF (Version 4 or Earlier) Format Data Files}. \url{https://CRAN.R-project.org/package=ncdf4}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-simmons2005ecmwf}{}}%
+\leavevmode\hypertarget{ref-michna2013rnetcdf}{}%
+Michna, Pavel, and Milton Woods. 2013. {``{RNetCDF}: A Package for Reading and Writing NetCDF Datasets.''} \emph{The R Journal} 5 (2): 29--36.
+
+\leavevmode\hypertarget{ref-rnetcdf}{}%
+---------. 2021. \emph{{RNetCDF}: Interface to 'NetCDF' Datasets}. \url{https://CRAN.R-project.org/package=RNetCDF}.
+
+\leavevmode\hypertarget{ref-spacetime}{}%
+Pebesma, Edzer. 2012. {``{spacetime}: Spatio-Temporal Data in r.''} \emph{Journal of Statistical Software} 51 (7): 1--30. \url{https://doi.org/10.18637/jss.v051.i07}.
+
+\leavevmode\hypertarget{ref-stars}{}%
+---------. 2021. \emph{{stars}: Spatiotemporal Arrays, Raster and Vector Data Cubes}. \url{https://CRAN.R-project.org/package=stars}.
+
+\leavevmode\hypertarget{ref-sf}{}%
+Pebesma, Edzer J. 2018. {``Simple Features for {R}: Standardized Support for Spatial Vector Data.''} \emph{R Journal} 10 (1): 439.
+
+\leavevmode\hypertarget{ref-ncdf4}{}%
+Pierce, David. 2019. \emph{{ncdf4}: Interface to Unidata netCDF (Version 4 or Earlier) Format Data Files}. \url{https://CRAN.R-project.org/package=ncdf4}.
+
+\leavevmode\hypertarget{ref-simmons2005ecmwf}{}%
 Simmons, Adrian, Mariano Hortal, Graeme Kelly, Anthony McNally, Agathe Untch, and Sakari Uppala.

### 2022-04-30 1-intro.Rmd
Updates intro section

@@
 The associated package \pkg{spacetime} [@spacetime] implements four spatio-temporal layouts (full grid, sparse grid, irregular, and trajectory) to handle different space and time combinations.
-The \pkg{stars} [@stars] package has been a recent implementation to use dense arrays to represent spatio-temporal cubes.
+The \pkg{stars} [@stars] package has a new implementation to use dense arrays to represent spatio-temporal cubes.
 It also interfaces with \pkg{sf} [@sf], a package commonly used for wrangling spatial data, and the \pkg{tidyverse} [@tidyverse] suite for general data wrangling and visualisation in \proglang{R}.
 
-Still, the data representation for spatio-temporal data can be further improved and there are two reasons for this.
+Still, the data representation for spatio-temporal data can be further extended and there are two reasons for this.
 Firstly, the raw data sourced in the wild is less often presented in any one of the layouts above, and fitting the raw data into a data object can sometimes be difficult.
 More often, spatio-temporal data are collected in separate 2D tables and analysts need to assemble them into a whole piece before exploring the data.
-Examples of components of spatio-temporal data can be 1) areal data recording the shape of a collection of areas of interest; 2) geostatistical data recording the longitude and latitude coordinates of locations, potentially also with other metadata related to the location, and; 3) temporal data of each location across time.
+Examples of components of spatio-temporal data can be 1) areal data recording the shape of a collection of areas of interest; 2) geostatistical data storing the longitude and latitude coordinates of locations, typically also with other metadata related to the location, and; 3) temporal data of each location across time.
 
@@
 The other reason is about tidy data concepts [@tidydata] and how they should be applied to spatio-temporal data.
-Tidy data should be structured into 1) one row per observation, 2) one column per variable, and 3) one type of data per table.
-The long form data is preferred over wide data given the downstream software i.e.
-\pkg{dplyr} [@dplyr] and \pkg{ggplot2} [@ggplot2] for data wrangling and visualisation.
-However, the long form can be inefficient to store feature geometries, especially for large multipolygons for hourly, daily or sub-daily periods over years, which are extensively collected and handled in time series analysis.
+According to the tidy data principles, data should be structured into 1) one row per observation, 2) one column per variable, and 3) one type of data per table.
+The long form data is preferred over wide data form given the downstream software such as \pkg{dplyr} [@dplyr] and \pkg{ggplot2} [@ggplot2] for data wrangling and visualisation.
+However, the long form can be inefficient to store feature geometries, especially for large multipolygons for hourly, daily or sub-daily periods over years, which are extensively collected and handled, for example in time series analysis.
 This poses the question of how to arrange spatial and temporal variables in a way that would make data wrangling, visualizing and analysing spatio-temporal data easier.
@@
 The software is available from the Comprehensive R Archive Network (CRAN) at https://CRAN.R-project.org/package=cubble.
-**At this point what is cubble is adding with respect the other packages it is not  clear to me**
 
@@
 Section \@ref(cubble) presents the main design and functionality of \pkg{cubble}.
-Section \@ref(others) explains how cubble deals with more advanced considerations, including data with hierarchical structure, data matching, how cubble fits with existing static and interactive visualisation tools, and spatio-temporal data transformation.
+Section \@ref(others) explains how cubble deals with more advanced considerations, including data with hierarchical structure, data matching and how cubble fits with existing static and interactive visualisation tools.
+Moreover we also illustrate how \pkg{cubble} deals with spatio-temporal data transformations.
 Section \@ref(examples) uses Australian weather station data and river level data as examples to demonstrate the use of \pkg{cubble}.
 An example of how \pkg{cubble} handles NetCDF data is also provided.
-Section \@ref(conclude) concludes the paper.
+Section \@ref(conclude) discuss the paper contributions and future directions.
 

### 2022-04-30 arxiv/cubble-arxiv.tex
Updates intro section

@@
 The \pkg{stars} (\protect\hyperlink{ref-stars}{E.
-Pebesma 2021}) package has been a recent implementation to use dense arrays to represent spatio-temporal cubes.
+Pebesma 2021}) package has a new implementation to use dense arrays to represent spatio-temporal cubes.
 It also interfaces with \pkg{sf} (\protect\hyperlink{ref-sf}{E.
@@
 
-Still, the data representation for spatio-temporal data can be further improved and there are two reasons for this.
+Still, the data representation for spatio-temporal data can be further extended and there are two reasons for this.
 Firstly, the raw data sourced in the wild is less often presented in any one of the layouts above, and fitting the raw data into a data object can sometimes be difficult.
 More often, spatio-temporal data are collected in separate 2D tables and analysts need to assemble them into a whole piece before exploring the data.
-Examples of components of spatio-temporal data can be 1) areal data recording the shape of a collection of areas of interest; 2) geostatistical data recording the longitude and latitude coordinates of locations, potentially also with other metadata related to the location, and; 3) temporal data of each location across time.
+Examples of components of spatio-temporal data can be 1) areal data recording the shape of a collection of areas of interest; 2) geostatistical data storing the longitude and latitude coordinates of locations, typically also with other metadata related to the location, and; 3) temporal data of each location across time.
 
 The other reason is about tidy data concepts (\protect\hyperlink{ref-tidydata}{Wickham 2014}) and how they should be applied to spatio-temporal data.
-Tidy data should be structured into 1) one row per observation, 2) one column per variable, and 3) one type of data per table.
-The long form data is preferred over wide data given the downstream software i.e.~\pkg{dplyr} (\protect\hyperlink{ref-dplyr}{Wickham et al.
+According to the tidy data principles, data should be structured into 1) one row per observation, 2) one column per variable, and 3) one type of data per table.
+The long form data is preferred over wide data form given the downstream software such as \pkg{dplyr} (\protect\hyperlink{ref-dplyr}{Wickham et al.
 2022}) and \pkg{ggplot2} (\protect\hyperlink{ref-ggplot2}{Wickham 2016}) for data wrangling and visualisation.
-However, the long form can be inefficient to store feature geometries, especially for large multipolygons for hourly, daily or sub-daily periods over years, which are extensively collected and handled in time series analysis.
+However, the long form can be inefficient to store feature geometries, especially for large multipolygons for hourly, daily or sub-daily periods over years, which are extensively collected and handled, for example in time series analysis.
 This poses the question of how to arrange spatial and temporal variables in a way that would make data wrangling, visualizing and analysing spatio-temporal data easier.
@@
 The software is available from the Comprehensive R Archive Network (CRAN) at \url{https://CRAN.R-project.org/package=cubble}.
-\textbf{At this point what is cubble is adding with respect the other packages it is not clear to me}
 
@@
 Section \ref{cubble} presents the main design and functionality of \pkg{cubble}.
-Section \ref{others} explains how cubble deals with more advanced considerations, including data with hierarchical structure, data matching, how cubble fits with existing static and interactive visualisation tools, and spatio-temporal data transformation.
+Section \ref{others} explains how cubble deals with more advanced considerations, including data with hierarchical structure, data matching and how cubble fits with existing static and interactive visualisation tools.
+Moreover we also illustrate how \pkg{cubble} deals with spatio-temporal data transformations.
 Section \ref{examples} uses Australian weather station data and river level data as examples to demonstrate the use of \pkg{cubble}.
 An example of how \pkg{cubble} handles NetCDF data is also provided.
-Section \ref{conclude} concludes the paper.
+Section \ref{conclude} discuss the paper contributions and future directions.
 

### 2022-04-30 2-0-cube.Rmd
Updating 2-0-cubble section

@@
 
-Spatio-temporal data can be conceptualised using a cubical data model with three axes, typically time, latitude and longitude.
+Spatio-temporal data can be conceptualised using a cubical data model with three axes which typically are, time, latitude and longitude.
 This abstraction can be useful for generalising operations and visualisation purposes: @lu_multidimensional_2018 shows how array operations (select, scale, reduce, rearrange, and compute) can be mapped onto the cube; @bach_review_2014 reviews the temporal data visualisation based on space-time cube operations.
 Notice that the term space-time cube in their article "does not need to involve spatial data", but refers to "an abstract 2D substrate that is used to visualize data at a specific time".
-Despite its main focus being on temporal data, the mindset of abstracting out data representation to construct visualisation still applies to our spatio-temporal data manipulation and visualisation.
+Despite its main focus being on temporal data, the mindset of abstracting out data representation to construct visualisations, still applies to our spatio-temporal data manipulation and visualisation approaches.
 
-The most common data cube using the three axes (time, latitude, longitude)  could be considered as taking snapshots of the space and stacking on time.
-Here we will define the spatio-temporal cube slightly differently with the three axes (time, site, variable), as illustrated in the leftmost cube in Figure \ref{fig:cube-def}.
+The most common data cube using the three axes (time, latitude, longitude) could be considered as taking snapshots of the space and stacking them across time.
+Here we define the spatio-temporal cube slightly differently by considering the three axes to be, time, site and  variable/s.
+That is illustrated in the leftmost cube in Figure \ref{fig:cube-def}.
 The time axis is the same in both versions, while the site axis now captures both latitude and longitude.
@@
 
-While a cubical data model is neat to view the spatio-temporal data conceptually, a 3D data representation, or an array,  may not be convenient for data wrangling.
-There are two reasons for this: 1) While arrays are efficient for the computation on numerical values, spatio-temporal data contains more than just numerical variables.
+While a cubical data model is neat to view spatio-temporal data conceptually, a 3D data representation, or an array,  may not be convenient for data wrangling.
+There are two reasons for this: 1) While arrays are efficient for the computation on numerical values, spatio-temporal data, in general, contains more than just numerical variables.
 For example, wrangling on character strings and specific datetime classes is common for spatio-temporal data but may not be easily done with array objects.
-2) Creating new variables in an array is more complex than in a data frame since it now needs to be defined in a 2D space, rather than a 1D vector, to be added into the array.
+2) Creating new variables in an array is more complex than in a data frame since it would need to be defined in a 2D space, rather than a 1D vector, to be added into the array.
 

### 2022-04-30 arxiv/cubble-arxiv.tex
Updating 2-0-cubble section

@@
 
-Spatio-temporal data can be conceptualised using a cubical data model with three axes, typically time, latitude and longitude.
+Spatio-temporal data can be conceptualised using a cubical data model with three axes which typically are, time, latitude and longitude.
 This abstraction can be useful for generalising operations and visualisation purposes: \protect\hyperlink{ref-lu_multidimensional_2018}{Lu, Appel, and Pebesma} (\protect\hyperlink{ref-lu_multidimensional_2018}{2018}) shows how array operations (select, scale, reduce, rearrange, and compute) can be mapped onto the cube; \protect\hyperlink{ref-bach_review_2014}{Bach et al.} (\protect\hyperlink{ref-bach_review_2014}{2014}) reviews the temporal data visualisation based on space-time cube operations.
-Notice that the term space-time cube in their article ``does not need to involve spatial data,'' but refers to ``an abstract 2D substrate that is used to visualize data at a specific time.'' Despite its main focus being on temporal data, the mindset of abstracting out data representation to construct visualisation still applies to our spatio-temporal data manipulation and visualisation.
-
-The most common data cube using the three axes (time, latitude, longitude) could be considered as taking snapshots of the space and stacking on time.
-Here we will define the spatio-temporal cube slightly differently with the three axes (time, site, variable), as illustrated in the leftmost cube in Figure \ref{fig:cube-def}.
+Notice that the term space-time cube in their article ``does not need to involve spatial data,'' but refers to ``an abstract 2D substrate that is used to visualize data at a specific time.'' Despite its main focus being on temporal data, the mindset of abstracting out data representation to construct visualisations, still applies to our spatio-temporal data manipulation and visualisation approaches.
+
+The most common data cube using the three axes (time, latitude, longitude) could be considered as taking snapshots of the space and stacking them across time.
+Here we define the spatio-temporal cube slightly differently by considering the three axes to be, time, site and variable/s.
+That is illustrated in the leftmost cube in Figure \ref{fig:cube-def}.
 The time axis is the same in both versions, while the site axis now captures both latitude and longitude.
@@
 
-While a cubical data model is neat to view the spatio-temporal data conceptually, a 3D data representation, or an array, may not be convenient for data wrangling.
-There are two reasons for this: 1) While arrays are efficient for the computation on numerical values, spatio-temporal data contains more than just numerical variables.
+While a cubical data model is neat to view spatio-temporal data conceptually, a 3D data representation, or an array, may not be convenient for data wrangling.
+There are two reasons for this: 1) While arrays are efficient for the computation on numerical values, spatio-temporal data, in general, contains more than just numerical variables.
 For example, wrangling on character strings and specific datetime classes is common for spatio-temporal data but may not be easily done with array objects.
-2) Creating new variables in an array is more complex than in a data frame since it now needs to be defined in a 2D space, rather than a 1D vector, to be added into the array.
+2) Creating new variables in an array is more complex than in a data frame since it would need to be defined in a 2D space, rather than a 1D vector, to be added into the array.
 

### 2022-04-30 2-the-cubble-pkg.Rmd
Updating 2-the-cubble-pkg section

@@
 Each core \pkg{cubble} function is introduced and accompanied with a short example using the data `climate_flat`  throughout this section.
-`climate_flat` contains five weather stations in Australia with spatial information of each station: station id, latitude, longitude, elevation, station name, World Meteorology Organisation ID and also daily temporal information of date, maximum and minimum temperature and precipitation for 2020.
+`climate_flat` contains data on five weather stations in Australia with spatial information of each station: station id, latitude, longitude, elevation, station name, World Meteorology Organisation ID and also daily temporal information, maximum and minimum temperature values and precipitation records for 2020.
 The first five rows of `climate_flat` are shown below:
@@
 The arguments \code{key} and \code{index} follow the wording in \pkg{tsibble} to describe the temporal order and multiple series while \code{coords} specifies the spatial location of each site.
-The code below creates a cubble out of \code{climate_flat} with \code{id} as the key, \code{date} as the index, and \code{c(long, lat)} as the coordinates:
+The code below creates a cubble out of \code{climate_flat} data with \code{id} as the key, \code{date} as the index, and \code{c(long, lat)} as the coordinates:
 
@@
 
-The cubble header provides some information about this data.
+The cubble header provides some information about the data.
 \code{id} is the variable name to identify each location and there are five unique locations.
-The bounding box is \code{[115.97, -32.94, 133.55, -12.42]}.
-The third row shows the name and type of all variables nested in the \code{ts} column and here it includes \code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.
+The bounding box is \code{[115.97, -32.94, 133.55, -12.42]} and provides information about the coordinates in the data set.
+The third row shows the name and type of all variables nested in the \code{ts} column.
+In this example, it includes \code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.
 
@@
 All the temporal variables are nested in a list column, hence it is also called the nested cubble.
-The rowwise structure makes it simpler to calculate on the list using the \code{mutate()} syntax, which is simpler than the \code{map()} when working with list column.
+The rowwise structure makes it simpler to operate on the list using the \code{mutate()} syntax, which is simpler than the \code{map()} when working with a list column.
 For example, calculating the number of raining days can be done by:
@@
 \pkg{cubble} also provides a long form, which expands the \code{ts} column and temporarily "hides" the spatial variables.
-The function \code{face_temporal()} is used to switch a nested cubble into a long one and the first row in Figure \ref{fig:face} illustrates this operation where the focus of the cube now changes from the site-variable face to the time-variable face.
+The function \code{face_temporal()} is used to switch from a nested cubble into a long one.
+The first row in Figure \ref{fig:face} illustrates this operation where the focus of the cube now changes from the site-variable face to the time-variable face.
 This code switches the cubble just created into its long form:
@@
 
-The first line in the header now shows the cubble is in its long form and the third line has been changed to show the name and type of spatial variables: \code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]}.
+The first line in the header now shows the cubble is in its long form and the third line has been changed to display the name and type of spatial variables: \code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]}.
 Unlike the nested form, the long cubble is built from class \code{groupped_df} where all the observations from the same sites form a group.
@@
 
-Wrangling spatio-temporal data can be an iterative process in the spatial and temporal dimensions.
+Wrangling spatio-temporal data can be seen as an iterative process in the spatial and temporal dimensions.
 Switching the focus back to the site-variable face can be accomplished by the function \code{face_spatial()}, which is the inverse of \code{face_temporal()}.
-The second row of Figure \ref{fig:face} illustrates the function and below is an example on the climate data:
+The second row of Figure \ref{fig:face} illustrates the function and below is an example using the climate data introduced earlier:
 
@@
 
-This function should generally be used in the last step of the analysis since it is a temporary operation, meaning these added spatial variables will disappear if switched to the nested form and then switched back:
+This function should generally be used in the last step of the analysis since it is a temporary operation, meaning that these added spatial variables will disappear, if switched to the nested form and then switched back:
 
@@
 The reason is that cubble is specifically designed to utilize the spatio-temporal structure when arranging variables from different observational units into a single object.
-And it enables easy pivoting between purely spatial, purely temporal, or unfolded into a combined form.
+Moreover, it enables easy pivoting between purely spatial, purely temporal, or unfolded into a combined form.
 
@@
 
-The \code{dplyr} package has provided many tools for data wrangling tasks and these operations are useful in the spatio-temporal context.
+The \code{dplyr} package has introduced many tools for data wrangling tasks and these operations are useful in the spatio-temporal context.
 \code{cubble} provides methods that support the following \code{dplyr} verbs in both the nested and long form:
@@
 
-In the nested cubble created, each element in the list-column \code{ts} is of \code{tbl_ts} class and operations available to the tsibble class are still valid within the cubble.
-For example, the code below calculates two features of the maximum temperature:
+When a nested cubble is created, each element in the list-column \code{ts} is of \code{tbl_ts} class and operations available to the tsibble class are still valid within the cubble.
+For example, the code below calculates two time series features (mean and variance) of the maximum temperature variable:
 
@@
 
-As a spatial data object, \code{sf} creates a simple feature geometry list-column (\code{sfc}) in the data frame to provide spatial operations on various geometry types (\code{POINT}, \code{LINESTRING}, \code{POLYGON}, \code{MULTIPOLYGON}, etc).
-When creating a cubble an \code{sf} object can be supplied as the spatial table.
-Once included in a cubble, methods for the \code{sfc} class can be applied in the nested form.
+As a spatial data object, \code{sf} creates a feature geometry list-column (\code{sfc}) in the data frame to provide spatial operations on various geometry types such as (\code{POINT}, \code{LINESTRING}, \code{POLYGON}, \code{MULTIPOLYGON}, etc).
+When creating a cubble a \code{sf} object can be supplied as the spatial table.
+Once that is included in a cubble, methods for the \code{sfc} class can be applied in the nested form inside the cubble object.
 An example of this is shown in \@ref(covid), which also handles the case when the site identifiers in the two tables do not match exactly.
-A spatial data object with \code{s2} vector can also be the input for the spatial table.
+A spatial data object with an \code{s2} vector can also be the input for the spatial table in \pkg{cubble}.
 

### 2022-04-30 arxiv/cubble-arxiv.tex
Updating 2-the-cubble-pkg section

@@
 Each core \pkg{cubble} function is introduced and accompanied with a short example using the data \texttt{climate\_flat} throughout this section.
-\texttt{climate\_flat} contains five weather stations in Australia with spatial information of each station: station id, latitude, longitude, elevation, station name, World Meteorology Organisation ID and also daily temporal information of date, maximum and minimum temperature and precipitation for 2020.
+\texttt{climate\_flat} contains data on five weather stations in Australia with spatial information of each station: station id, latitude, longitude, elevation, station name, World Meteorology Organisation ID and also daily temporal information, maximum and minimum temperature values and precipitation records for 2020.
 The first five rows of \texttt{climate\_flat} are shown below:
@@
 \pkg{cubble} also provides a long form, which expands the \code{ts} column and temporarily ``hides'' the spatial variables.
-The function \code{face_temporal()} is used to switch a nested cubble into a long one and the first row in Figure \ref{fig:face} illustrates this operation where the focus of the cube now changes from the site-variable face to the time-variable face.
+The function \code{face_temporal()} is used to switch from a nested cubble into a long one.
+The first row in Figure \ref{fig:face} illustrates this operation where the focus of the cube now changes from the site-variable face to the time-variable face.
 This code switches the cubble just created into its long form:
@@
 
-As a spatial data object, \code{sf} creates a simple feature geometry list-column (\code{sfc}) in the data frame to provide spatial operations on various geometry types (\code{POINT}, \code{LINESTRING}, \code{POLYGON}, \code{MULTIPOLYGON}, etc).
-When creating a cubble an \code{sf} object can be supplied as the spatial table.
-Once included in a cubble, methods for the \code{sfc} class can be applied in the nested form.
+As a spatial data object, \code{sf} creates a feature geometry list-column (\code{sfc}) in the data frame to provide spatial operations on various geometry types such as (\code{POINT}, \code{LINESTRING}, \code{POLYGON}, \code{MULTIPOLYGON}, etc).
+When creating a cubble a \code{sf} object can be supplied as the spatial table.
+Once that is included in a cubble, methods for the \code{sfc} class can be applied in the nested form inside the cubble object.
 An example of this is shown in \ref{covid}, which also handles the case when the site identifiers in the two tables do not match exactly.
-A spatial data object with \code{s2} vector can also be the input for the spatial table.
+A spatial data object with an \code{s2} vector can also be the input for the spatial table in \pkg{cubble}.
 

### 2022-04-30 3-advanced.Rmd
Advanced section-morework needed

@@
 
-Spatial locations can have grouping structures either inherent in the data or obtained by clustering.
-Rather than analysing variables in the site level, summarised variables in the cluster level can give a crisper picture of local areas.
-In cubble, \code{switch_key()} can be used to create a new level of grouping of spatial locations by specifying a clustering variable.
-Figure \ref{fig:illu-hier} illustrates the relationship of cubbles at station and cluster level, in both the long and nested form.
-By specifying \code{cluster_nested <- station_nested \%>\% switch_key(key = cluster)}, the cubble redefines the cubble key from the \code{id} in \code{station_nested} to \code{cluster} in \code{cluster_nested}.
-All the spatial variables variant to \code{cluster} are now nested into a \code{.val} column and cluster level variables can be computed in the same fashion as station level variables in \code{station_nested}.
+Spatial locations can have grouping structures either inherent to the data or obtained by clustering.
+In those cases, rather than analysing variables at the site level, it might be of interest to summarised variables at a cluster or group level, as that can give a crisper picture of local areas.
+In cubble, \code{switch_key()} can be used to create a new grouping level of spatial locations by specifying a clustering variable.
+Figure \ref{fig:illu-hier} illustrates the relationship of cubbles at station and cluster level, in both the long and nested forms.
+By specifying \code{cluster_nested <- station_nested \%>\% switch_key(key = cluster)}, the cubble redefines the cubble key from the \code{id} in \code{station_nested} to \code{cluster} into \code{cluster_nested}.
+All the spatial variables in \code{cluster} are now nested into a \code{.val} column and cluster level variables can be computed in the same fashion as station level variables in \code{station_nested}.
 
@@
 
-One task that may interest analysts in spatio-temporal data is to find how similar the time series from nearby sites are.
+One task that may interest analysts in spatio-temporal data is to find how similar time series, from nearby sites, are.
 This problem can be seen as a matching problem [@stuart2010matching; @mcintosh2018using] that pairs up similar time series in nearby locations or a data fusing exercise that merges data collected from different sources [@cocchi2019data].
@@
 The algorithm first matches the two data sources spatially by computing the pairwise distance on latitude and longitude.
-Pairs that pass the spatial matching are then matched temporally through computing the number of matched peaks within a fixed length window.
-Figure \ref{fig:illu-matching} illustrates this temporal matching in more details.
-Given two series \code{A} and \code{a}, three peaks have been picked in each series.
-An interval, with default length of five, is constructed for each peak in series \code{A} and the peaks in series \code{a} are tested against whether they fall into any of the intervals.
+After that, pairs that pass the spatial matching are then matched temporally by computing the number of matched peaks within a fixed length moving window.
+Figure \ref{fig:illu-matching} illustrates this temporal matching in more detail.
+Given two series \code{A} and \code{a}, three peaks have been identified in each series.
+An interval, with default length of five, is constructed for each peak in series \code{A}, while the peaks in series \code{a} are tested against whether they fall into any of the intervals.
 In this illustration, there are two matches for these two series.
@@
 
-The linking structure of two forms in cubble fits naturally with the interactive graphic pipeline discussed in the literature [@buja1988elements; @buja1996interactive;  @sutherland2000orca; @xie2014reactive; @cheng2016enabling].
+The linking structure of two forms in cubble fits naturally with the interactive graphic frameworks discussed in the literature [@buja1988elements; @buja1996interactive;  @sutherland2000orca; @xie2014reactive; @cheng2016enabling].
 Diagram \ref{fig:illu-interactive} illustrates how linking works from the map to the time series in cubble.

### 2022-04-30 arxiv/cubble-arxiv.tex
Advanced section-morework needed

@@
 
-One task that may interest analysts in spatio-temporal data is to find how similar the time series from nearby sites are.
+One task that may interest analysts in spatio-temporal data is to find how similar time series, from nearby sites, are.
 This problem can be seen as a matching problem (\protect\hyperlink{ref-stuart2010matching}{Stuart 2010}; \protect\hyperlink{ref-mcintosh2018using}{McIntosh et al.
@@
 
-The linking structure of two forms in cubble fits naturally with the interactive graphic pipeline discussed in the literature (\protect\hyperlink{ref-buja1988elements}{Buja, Asimov, and Hurley 1988}; \protect\hyperlink{ref-buja1996interactive}{Buja, Cook, and Swayne 1996}; \protect\hyperlink{ref-sutherland2000orca}{Sutherland et al.
+The linking structure of two forms in cubble fits naturally with the interactive graphic frameworks discussed in the literature (\protect\hyperlink{ref-buja1988elements}{Buja, Asimov, and Hurley 1988}; \protect\hyperlink{ref-buja1996interactive}{Buja, Cook, and Swayne 1996}; \protect\hyperlink{ref-sutherland2000orca}{Sutherland et al.
 2000}; \protect\hyperlink{ref-xie2014reactive}{Xie, Hofmann, and Cheng 2014}; \protect\hyperlink{ref-cheng2016enabling}{Cheng, Cook, and Hofmann 2016}).

### 2022-04-30 4-1-basic.Rmd
Updating conclusion

@@
 
-The Victoria State Government in Australia provides daily COVID case numbers by date, source, and local government area (LGA).
+The Australian Victoria State Governmentprovides daily COVID case numbers providing information about the source, and local government area (LGA) of the recorded cases.
 This data can be used to visualise COVID spread when combined with map information on LGA, available from the Australian Bureau of Statistics.
-The first five rows of both data are printed below:
+The first five rows these data sets are printed below:
 
@@
 
-Cubble can be created with separate spatial and temporal tables in a list, with other arguments (\code{key}, \code{index}, and \code{coords}) introduced in section \@ref(create).
+A \code{cubble} object can be created from separate spatial and temporal tables in a list, with other arguments (\code{key}, \code{index}, and \code{coords}) introduced in section \@ref(create).
 \code{as_cubble()} will automatically check the matching of sites in both tables and emit a warning if a location has missing spatial or temporal information.

### 2022-04-30 5-conclusion.Rmd
Updating conclusion

--- 2022-04-27 5-conclusion.Rmd
@@
 
-This paper describes an \proglang{R} package \pkg{cubble} for organising, manipulating and visualising spatio-temporal data.
-A new data structure, \code{cubble}, with the nested and long form, is proposed to connect the time invariant and varying variables in the spatio-temporal data.
+This paper presents an \proglang{R} package \pkg{cubble} for organising, manipulating and visualising spatio-temporal data.
+The package introduces a new data structure for spatio-temporal data, \code{cubble}, that connects the time invariant and varying variables and that allows the user to work with a nested and long form of the data.
 This design frees the data analysts from spending time on organising variables of different observational units.
-The data structure is also flexible to the techniques and packages analysts use to analyse the data, for example, in the matching example in section 4.3, users are free to use algorithms from another package to cluster stations.
+The data structure and the capabilities introduced in this package can be used and combined with existing spatio-temporal R packages such as \pkg{sf}, data wrangling packages such as \pkg{dplyr}, and visualization packages such as \pkg{ggplot2}, \pkg{plotly} and \pkg{leaflet}.
+
+The data structure and functions are demonstrated with extensive examples.
+These include creating and coercing wild-caught data with potential mismatch on sites, handling hierarchical data, matching time series spatially and temporally, as well as reproducing ERA5 results from NetCDF data.
+Visualization of the \code{cubble} objects and derivatives is presented via interactive graphic pipelines using \pkg{plotly} and \pkg{leaflet}.
 
@@
 In the spatial aspect, the simple feature representation wraps spatial coordinates in a list column, for some complex spatial operations, while sometimes, analysts may want to unpack them into longitude and latitude columns to work from.
-Another extension to \pkg{cubble} is to build a smooth transition between the coordinate columns and simple feature geometry column.
+A smooth transition between the coordinate and simple feature representation can be added to \pkg{cubble}.
+\pkg{cubble} can also be extended to handle sites with moving coordinates.
+
+The goal of this work is to add capabilities into the spatio-temporal practitioners toolbox that facilitates their work within the tidy data framework.
+
+<!-- and this would involve constructing a \code{sp} column for locations, similar to \code{ts} for time, \textcolor{green}{and a form locations can be pivoted into, like the long form for temporal variables}.
+The multivariate aspect, \pkg{cubble} can also be extended with interface to more high dimensional visualisation methods, i.e., the tour method, to understand variable importance or comparing location similarities.
+-->
 

### 2022-04-30 arxiv/cubble-arxiv.tex
Updating conclusion

@@
 
-Cubble can be created with separate spatial and temporal tables in a list, with other arguments (\code{key}, \code{index}, and \code{coords}) introduced in section \ref{create}.
+A \code{cubble} object can be created from separate spatial and temporal tables in a list, with other arguments (\code{key}, \code{index}, and \code{coords}) introduced in section \ref{create}.
 \code{as_cubble()} will automatically check the matching of sites in both tables and emit a warning if a location has missing spatial or temporal information.
@@
 
-This paper describes an \proglang{R} package \pkg{cubble} for organising, manipulating and visualising spatio-temporal data.
-A new data structure, \code{cubble}, with the nested and long form, is proposed to connect the time invariant and varying variables in the spatio-temporal data.
+This paper presents an \proglang{R} package \pkg{cubble} for organising, manipulating and visualising spatio-temporal data.
+The package introduces a new data structure for spatio-temporal data, \code{cubble}, that connects the time invariant and varying variables and that allows the user to work with a nested and long form of the data.
 This design frees the data analysts from spending time on organising variables of different observational units.
-The data structure is also flexible to the techniques and packages analysts use to analyse the data, for example, in the matching example in section 4.3, users are free to use algorithms from another package to cluster stations.
-
-The data structure and software are demonstrated with extensive examples.
-These include creating and coercing wild-caught data with potential mismatch on sites, handling hierarchical data, matching series spatially and temporally, reproducing ERA5 results from NetCDF data, and, lastly fitting into the interactive graphic pipeline of \pkg{plotly} and \pkg{leaflet}.
+The data structure and the capabilities introduced in this package can be used and combined with existing spatio-temporal R packages such as \pkg{sf}, data wrangling packages such as \pkg{dplyr}, and visualization packages such as \pkg{ggplot2}, \pkg{plotly} and \pkg{leaflet}.
+
+The data structure and functions are demonstrated with extensive examples.
+These include creating and coercing wild-caught data with potential mismatch on sites, handling hierarchical data, matching time series spatially and temporally, as well as reproducing ERA5 results from NetCDF data.
+Visualization of the \code{cubble} objects and derivatives is presented via interactive graphic pipelines using \pkg{plotly} and \pkg{leaflet}.
 
@@
 A smooth transition between the coordinate and simple feature representation can be added to \pkg{cubble}.
-\pkg{cubble} can also be extended to handle sites with moving coordinates and this would involve constructing a \code{sp} column for locations, similar to \code{ts} for time, and a form locations can be pivoted into, like the long form for temporal variables.
-In the multivariate aspect, \pkg{cubble} can also be extended with interface to more high dimensional visualisation methods, i.e., the tour method, to understand variable importance or comparing location similarities.
+\pkg{cubble} can also be extended to handle sites with moving coordinates.
+
+The goal of this work is to add capabilities into the spatio-temporal practitioners toolbox that facilitates their work within the tidy data framework.
 

### 2022-04-30 3-advanced.Rmd
small changes

@@
 Figure \ref{fig:illu-hier} illustrates the relationship of cubbles at station and cluster level, in both the long and nested forms.
-By specifying \code{cluster_nested <- station_nested \%>\% switch_key(key = cluster)}, the cubble redefines the cubble key from the \code{id} in \code{station_nested} to \code{cluster} into \code{cluster_nested}.
-All the spatial variables in \code{cluster} are now nested into a \code{.val} column and cluster level variables can be computed in the same fashion as station level variables in \code{station_nested}.
+By specifying \code{cluster_nested <- station_nested \%>\% switch_key(key = cluster)}, the cubble redefines the cubble key from the \code{id} column in \code{station_nested} to the \code{cluster} column in \code{cluster_nested}.
+All the spatial variables variant to the \code{cluster} column are now nested into a \code{.val} column and cluster level variables can be computed in the same fashion as station level variables in \code{station_nested}.
 
@@
 
-The linking structure of two forms in cubble fits naturally with the interactive graphic frameworks discussed in the literature [@buja1988elements; @buja1996interactive;  @sutherland2000orca; @xie2014reactive; @cheng2016enabling].
+The linking structure of two forms in cubble fits naturally with the interactive graphic pipeline discussed in the literature [@buja1988elements; @buja1996interactive;  @sutherland2000orca; @xie2014reactive; @cheng2016enabling].
 Diagram \ref{fig:illu-interactive} illustrates how linking works from the map to the time series in cubble.

### 2022-04-30 4-1-basic.Rmd
small changes

@@
 
-The Australian Victoria State Governmentprovides daily COVID case numbers providing information about the source, and local government area (LGA) of the recorded cases.
+The Victoria State Government in Australia provides daily COVID information about the source, and local government area (LGA) of the recorded cases.
 This data can be used to visualise COVID spread when combined with map information on LGA, available from the Australian Bureau of Statistics.

### 2022-04-30 5-conclusion.Rmd
small changes

@@
 The package introduces a new data structure for spatio-temporal data, \code{cubble}, that connects the time invariant and varying variables and that allows the user to work with a nested and long form of the data.
-This design frees the data analysts from spending time on organising variables of different observational units.
+TThe goal of this work is to add capabilities into the spatio-temporal practitioners toolbox that facilitates their work within the tidy data framework.
 The data structure and the capabilities introduced in this package can be used and combined with existing spatio-temporal R packages such as \pkg{sf}, data wrangling packages such as \pkg{dplyr}, and visualization packages such as \pkg{ggplot2}, \pkg{plotly} and \pkg{leaflet}.
@@
 
-Future directions of the package involves responding to changes in the tidyverse packages that \pkg{cubble} imports, in particular, \pkg{tibble}, \pkg{tidyr}, and \pkg{dplyr}.
-In the spatial aspect, the simple feature representation wraps spatial coordinates in a list column, for some complex spatial operations, while sometimes, analysts may want to unpack them into longitude and latitude columns to work from.
-A smooth transition between the coordinate and simple feature representation can be added to \pkg{cubble}.
-\pkg{cubble} can also be extended to handle sites with moving coordinates.
-
-The goal of this work is to add capabilities into the spatio-temporal practitioners toolbox that facilitates their work within the tidy data framework.
-
-<!-- and this would involve constructing a \code{sp} column for locations, similar to \code{ts} for time, \textcolor{green}{and a form locations can be pivoted into, like the long form for temporal variables}.
-The multivariate aspect, \pkg{cubble} can also be extended with interface to more high dimensional visualisation methods, i.e., the tour method, to understand variable importance or comparing location similarities.
--->
+Future directions of the package involves handling sites with moving coordinates.
+This would involve constructing a list-column for location coordinates and a form these locations can be pivoted into, like the long form for temporal variables.
+In the multivariate aspect, \pkg{cubble} can also be extended with interface to more high dimensional visualisation methods, i.e., the tour method, to understand variable importance or comparing location similarities.
 

### 2022-04-30 arxiv/cubble-arxiv.Rmd
small changes

@@
 In order to facilitate the study of different portions or combinations of spatio-temporal data, we introduce a new data structure, \code{cubble}, with a suite of functions enabling easy slicing and dicing on the different components spatio-temporal components.
-The proposed \code{cubble} structure ensures that all the components of the data are easy to access and manipulate while providing flexibility for data analysis.In addition,   \code{cubble} facilitates visual and numerical explorations of the data while easing data wrangling and modelling.
+The proposed \code{cubble} structure ensures that all the components of the data are easy to access and manipulate while providing flexibility for data analysis.
+In addition, \code{cubble} facilitates visual and numerical explorations of the data while easing data wrangling and modelling.
 The \code{cubble} structure and the functions provided in the \code{cubble} R package equip users with the capability to handle hierarchical spatial and temporal structures.

### 2022-04-30 arxiv/cubble-arxiv.tex
small changes

@@
 In order to facilitate the study of different portions or combinations of spatio-temporal data, we introduce a new data structure, \code{cubble}, with a suite of functions enabling easy slicing and dicing on the different components spatio-temporal components.
-The proposed \code{cubble} structure ensures that all the components of the data are easy to access and manipulate while providing flexibility for data analysis.In addition, \code{cubble} facilitates visual and numerical explorations of the data while easing data wrangling and modelling.
+The proposed \code{cubble} structure ensures that all the components of the data are easy to access and manipulate while providing flexibility for data analysis.
+In addition, \code{cubble} facilitates visual and numerical explorations of the data while easing data wrangling and modelling.
 The \code{cubble} structure and the functions provided in the \code{cubble} R package equip users with the capability to handle hierarchical spatial and temporal structures.
@@
 The SpatioTemporal CRAN task view (\protect\hyperlink{ref-ctvspatiotemporal}{Edzer Pebesma 2022}) gathers information about R packages designed for spatio-temporal data and it has a section on \emph{Representing data} that lists existing spatio-temporal data representations used in \proglang{R}.
-Among them, \protect\hyperlink{ref-spacetime}{E.
-Pebesma} (\protect\hyperlink{ref-spacetime}{2012}) summarises spatio-temporal data into three forms: time-wide, space-wide, and long formats.
+Among them, E.
+Pebesma (\protect\hyperlink{ref-spacetime}{2012}) summarises spatio-temporal data into three forms: time-wide, space-wide, and long formats.
 The associated package \pkg{spacetime} (\protect\hyperlink{ref-spacetime}{E.
@@
 In the package, a new data structure, also called \code{cubble}, is proposed to organise spatial and temporal variables as two forms of a single data object so that they can be wrangled separately or combined while being kept synchronised.
-Among the four spacetime layouts in \protect\hyperlink{ref-spacetime}{E.
-Pebesma} (\protect\hyperlink{ref-spacetime}{2012}), \pkg{cubble} can be applied to full grid, sparse grid, or irregular, but not trajectory, which is outside the scope of this work.
+Among the four spacetime layouts in E.
+Pebesma (\protect\hyperlink{ref-spacetime}{2012}), \pkg{cubble} can be applied to full grid, sparse grid, or irregular, but not trajectory, which is outside the scope of this work.
 The software is available from the Comprehensive R Archive Network (CRAN) at \url{https://CRAN.R-project.org/package=cubble}.
@@
 Spatio-temporal data can be conceptualised using a cubical data model with three axes which typically are, time, latitude and longitude.
-This abstraction can be useful for generalising operations and visualisation purposes: \protect\hyperlink{ref-lu_multidimensional_2018}{Lu, Appel, and Pebesma} (\protect\hyperlink{ref-lu_multidimensional_2018}{2018}) shows how array operations (select, scale, reduce, rearrange, and compute) can be mapped onto the cube; \protect\hyperlink{ref-bach_review_2014}{Bach et al.} (\protect\hyperlink{ref-bach_review_2014}{2014}) reviews the temporal data visualisation based on space-time cube operations.
-Notice that the term space-time cube in their article ``does not need to involve spatial data,'' but refers to ``an abstract 2D substrate that is used to visualize data at a specific time.'' Despite its main focus being on temporal data, the mindset of abstracting out data representation to construct visualisations, still applies to our spatio-temporal data manipulation and visualisation approaches.
+This abstraction can be useful for generalising operations and visualisation purposes: Lu, Appel, and Pebesma (\protect\hyperlink{ref-lu_multidimensional_2018}{2018}) shows how array operations (select, scale, reduce, rearrange, and compute) can be mapped onto the cube; Bach et al.
+(\protect\hyperlink{ref-bach_review_2014}{2014}) reviews the temporal data visualisation based on space-time cube operations.
+Notice that the term space-time cube in their article ``does not need to involve spatial data'', but refers to ``an abstract 2D substrate that is used to visualize data at a specific time''.
+Despite its main focus being on temporal data, the mindset of abstracting out data representation to construct visualisations, still applies to our spatio-temporal data manipulation and visualisation approaches.
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/diagram-keynotes/diagram-keynotes.007}
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.007}
 
@@
 
-The linking structure of two forms in cubble fits naturally with the interactive graphic frameworks discussed in the literature (\protect\hyperlink{ref-buja1988elements}{Buja, Asimov, and Hurley 1988}; \protect\hyperlink{ref-buja1996interactive}{Buja, Cook, and Swayne 1996}; \protect\hyperlink{ref-sutherland2000orca}{Sutherland et al.
+The linking structure of two forms in cubble fits naturally with the interactive graphic pipeline discussed in the literature (\protect\hyperlink{ref-buja1988elements}{Buja, Asimov, and Hurley 1988}; \protect\hyperlink{ref-buja1996interactive}{Buja, Cook, and Swayne 1996}; \protect\hyperlink{ref-sutherland2000orca}{Sutherland et al.
 2000}; \protect\hyperlink{ref-xie2014reactive}{Xie, Hofmann, and Cheng 2014}; \protect\hyperlink{ref-cheng2016enabling}{Cheng, Cook, and Hofmann 2016}).
@@
 In \proglang{R}, \pkg{GGally} implements glyph maps through the \code{glyphs()} function.
-The function constructs a data frame with calculated position (\code{gx}, \code{gy}, \code{gid}) of each point on the time series using linear algebra (Equation 1 and 2 in \protect\hyperlink{ref-Wickham2012-yr}{Wickham et al.} (\protect\hyperlink{ref-Wickham2012-yr}{2012})).
+The function constructs a data frame with calculated position (\code{gx}, \code{gy}, \code{gid}) of each point on the time series using linear algebra (Equation 1 and 2 in Wickham et al.
+(\protect\hyperlink{ref-Wickham2012-yr}{2012})).
 The data can then be piped into \code{ggplot} to create the glyph map as:
@@
 
-Future directions of the package involves responding to changes in the tidyverse packages that \pkg{cubble} imports, in particular, \pkg{tibble}, \pkg{tidyr}, and \pkg{dplyr}.
-In the spatial aspect, the simple feature representation wraps spatial coordinates in a list column, for some complex spatial operations, while sometimes, analysts may want to unpack them into longitude and latitude columns to work from.
-A smooth transition between the coordinate and simple feature representation can be added to \pkg{cubble}.
-\pkg{cubble} can also be extended to handle sites with moving coordinates.
-
-The goal of this work is to add capabilities into the spatio-temporal practitioners toolbox that facilitates their work within the tidy data framework.
+Future directions of the package involves handling sites with moving coordinates.
+This would involve constructing a list-column for location coordinates and a form these locations can be pivoted into, like the long form for temporal variables.
+In the multivariate aspect, \pkg{cubble} can also be extended with interface to more high dimensional visualisation methods, i.e., the tour method, to understand variable importance or comparing location similarities.
 
@@
 
-\leavevmode\hypertarget{ref-michna2013rnetcdf}{}%
-Michna, Pavel, and Milton Woods. 2013. {``{RNetCDF}: A Package for Reading and Writing NetCDF Datasets.''} \emph{The R Journal} 5 (2): 29--36.
-
-\leavevmode\hypertarget{ref-rnetcdf}{}%
----------. 2021. \emph{{RNetCDF}: Interface to 'NetCDF' Datasets}. \url{https://CRAN.R-project.org/package=RNetCDF}.
-
-\leavevmode\hypertarget{ref-spacetime}{}%
-Pebesma, Edzer. 2012. {``{spacetime}: Spatio-Temporal Data in r.''} \emph{Journal of Statistical Software} 51 (7): 1--30. \url{https://doi.org/10.18637/jss.v051.i07}.
-
-\leavevmode\hypertarget{ref-stars}{}%
----------. 2021. \emph{{stars}: Spatiotemporal Arrays, Raster and Vector Data Cubes}. \url{https://CRAN.R-project.org/package=stars}.
-
-\leavevmode\hypertarget{ref-sf}{}%
-Pebesma, Edzer J. 2018. {``Simple Features for {R}: Standardized Support for Spatial Vector Data.''} \emph{R Journal} 10 (1): 439.
-
-\leavevmode\hypertarget{ref-ncdf4}{}%
-Pierce, David. 2019. \emph{{ncdf4}: Interface to Unidata netCDF (Version 4 or Earlier) Format Data Files}. \url{https://CRAN.R-project.org/package=ncdf4}.
-
-\leavevmode\hypertarget{ref-simmons2005ecmwf}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-michna2013rnetcdf}{}}%
+Michna, Pavel, and Milton Woods. 2013. {``\pkg{RNetCDF}: A Package for Reading and Writing NetCDF Datasets.''} \emph{The R Journal} 5 (2): 29--36.
+
+\leavevmode\vadjust pre{\hypertarget{ref-rnetcdf}{}}%
+---------. 2021. \emph{\pkg{RNetCDF}: Interface to 'NetCDF' Datasets}. \url{https://CRAN.R-project.org/package=RNetCDF}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-spacetime}{}}%
+Pebesma, Edzer. 2012. {``\pkg{spacetime}: Spatio-Temporal Data in r.''} \emph{Journal of Statistical Software} 51 (7): 1--30. \url{https://doi.org/10.18637/jss.v051.i07}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-stars}{}}%
+---------. 2021. \emph{\pkg{stars}: Spatiotemporal Arrays, Raster and Vector Data Cubes}. \url{https://CRAN.R-project.org/package=stars}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-sf}{}}%
+Pebesma, Edzer J. 2018. {``Simple Features for \proglang{R}: Standardized Support for Spatial Vector Data.''} \emph{R Journal} 10 (1): 439.
+
+\leavevmode\vadjust pre{\hypertarget{ref-ncdf4}{}}%
+Pierce, David. 2019. \emph{\pkg{ncdf4}: Interface to Unidata netCDF (Version 4 or Earlier) Format Data Files}. \url{https://CRAN.R-project.org/package=ncdf4}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-simmons2005ecmwf}{}}%
 Simmons, Adrian, Mariano Hortal, Graeme Kelly, Anthony McNally, Agathe Untch, and Sakari Uppala.

### 2022-05-06 2-the-cubble-pkg.Rmd
changes made with the team

@@
 
-This section will first introduce the core functions in the cubble class: \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and \code{unfold()} in subsection \@ref(create) - \@ref(unfold).
-The next subsection addresses the necessity of creating a new class.
-Subsection \@ref(compact) will then show the compatibility of cubble with existing packages in spatial and temporal analysis.
+Here we introduce the core functionalities of cubble.
+The main functions, \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and \code{unfold()} are described in Section \@ref(create) - \@ref(unfold).
+Section \@ref(tidyverse) explains the need to create a new class.
+Section \@ref(compact) will then discusses how cubble relates to existing packages for spatial and temporal data analysis.
 
-Each core \pkg{cubble} function is introduced and accompanied with a short example using the data `climate_flat`  throughout this section.
-`climate_flat` contains data on five weather stations in Australia with spatial information of each station: station id, latitude, longitude, elevation, station name, World Meteorology Organisation ID and also daily temporal information, maximum and minimum temperature values and precipitation records for 2020.
+The dataset `climate_flat` is used to illustrate \pkg{cubble} functionality.
+This is a subset from the ORIGIN and contains the spatial variables, station id, latitude, longitude, elevation, station name, World Meteorology Organisation ID, in additional to daily temporal information,  maximum and minimum temperature values and precipitation records for 2020.
 The first five rows of `climate_flat` are shown below:
@@
 
-Spatio-temporal data can come in various formats and shapes and a cubble can be created from various types of raw data.
-This includes tibble and its variates in multiple tables (an example is provided in section \@ref(covid)), and netCDF (detailed in section \@ref(netcdf)).
-The function \code{as_cubble()} is used to create a cubble with three additional inputs: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order  (longitude, latitude).
+A cubble can be created from various common spatio-temporal data formats, including basic R objects like tibble, tsibble, sf.
+Section \@ref(covid) describes converting multiple tables into a cubble object and Section \@ref(netcdf) illustrate how to convert a netCDF object.
+
+The function \code{as_cubble()} is used to create a cubble with three arguments: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order  (longitude, latitude).
 The arguments \code{key} and \code{index} follow the wording in \pkg{tsibble} to describe the temporal order and multiple series while \code{coords} specifies the spatial location of each site.
-The code below creates a cubble out of \code{climate_flat} data with \code{id} as the key, \code{date} as the index, and \code{c(long, lat)} as the coordinates:
+The code below creates a cubble out of \code{climate_flat} (a single tibble) with \code{id} as the key, \code{date} as the index, and \code{c(long, lat)} as the coordinates:
 
@@
 
-The cubble header provides some information about the data.
+Printing a cubble provides some information about the data.
 \code{id} is the variable name to identify each location and there are five unique locations.
@@
 
-The resulting cubble is built from a \code{rowwise_df} class where each row forms a group.
+A cubble object is an extension of a \code{rowwise_df} object where each row forms a group.
 All the temporal variables are nested in a list column, hence it is also called the nested cubble.
-The rowwise structure makes it simpler to operate on the list using the \code{mutate()} syntax, which is simpler than the \code{map()} when working with a list column.
-For example, calculating the number of raining days can be done by:
+The rowwise structure makes it simpler to operate on the list using the \code{mutate()} syntax, which is simpler than the \code{purr::map()} when working with a list column.
+For example, calculating the number of rainy days can be done by:
 
@@
 The nested form can be used for those operations where the output is only indexed by the spatial identifier (\code{key}), but becomes inadequate when outputs need both a spatial and a temporal identifier (\code{key} and \code{index}).
-\pkg{cubble} also provides a long form, which expands the \code{ts} column and temporarily "hides" the spatial variables.
-The function \code{face_temporal()} is used to switch from a nested cubble into a long one.
+The \pkg{cubble} class also provides a long form, which expands the \code{ts} column and temporarily "hides" the spatial variables.
+The function \code{face_temporal()} is used to switch from the nested cubble into the long one.
 The first row in Figure \ref{fig:face} illustrates this operation where the focus of the cube now changes from the site-variable face to the time-variable face.
@@
 
-The first line in the header now shows the cubble is in its long form and the third line has been changed to display the name and type of spatial variables: \code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]}.
-Unlike the nested form, the long cubble is built from class \code{groupped_df} where all the observations from the same sites form a group.
+The first line in printed cubble now shows the cubble in its long form and the third line has been changed to display the name and type of spatial variables: \code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]}.
+Unlike the nested form, the long cubble is built from a \code{grouped_df} object where all the observations from the same sites form a group.
 
@@
 Switching the focus back to the site-variable face can be accomplished by the function \code{face_spatial()}, which is the inverse of \code{face_temporal()}.
-The second row of Figure \ref{fig:face} illustrates the function and below is an example using the climate data introduced earlier:
+The second row of Figure \ref{fig:face} illustrates the function, which is used as follows:
 
@@
 Sometimes, analysts may need to apply some variable transformation that involves both the spatial and temporal variables.
-An example of this is the transformation of temporal variables into the spatial dimension in glyph maps, which will be elaborated in section \ref{st_transformation}.
-This type of operation can be seen as flattening, or *unfolding*, the cube into a 2D data frame.
+An example of this is the transformation of temporal variables into the spatial dimension in glyph maps [ref].
+(Glyph maps will be explained in Section \ref{st_transformation}.) This type of operation can be seen as flattening, or *unfolding*, the cube into a 2D data frame.
 Here the function \code{unfold()} moves the spatial variables \code{long} and \code{lat} into the long cubble:
@@
 <!-- https://www.jstatsoft.org/style#my-latex-paper-does-not-compile-when-there-is-jss-markup-in-section-titles-what-should-i-do -->
-\subsection[Why not just use]{Why not just use \code{dplyr::nest\_by()}?}
+
+## Why not just use the existing tidyverse functions {#tidyverse}
+
+<!-- \subsection[Why not just use]{Why not just use \code{dplyr::nest\_by()}?} -->
 

### 2022-05-06 arxiv/cubble-arxiv.tex
changes made with the team

@@
 
-This section will first introduce the core functions in the cubble class: \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and \code{unfold()} in subsection \ref{create} - \ref{unfold}.
-The next subsection addresses the necessity of creating a new class.
-Subsection \ref{compact} will then show the compatibility of cubble with existing packages in spatial and temporal analysis.
-
-Each core \pkg{cubble} function is introduced and accompanied with a short example using the data \texttt{climate\_flat} throughout this section.
-\texttt{climate\_flat} contains data on five weather stations in Australia with spatial information of each station: station id, latitude, longitude, elevation, station name, World Meteorology Organisation ID and also daily temporal information, maximum and minimum temperature values and precipitation records for 2020.
+Here we introduce the core functionalities of cubble.
+The main functions, \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and \code{unfold()} are described in Section \ref{create} - \ref{unfold}.
+Section \ref{tidyverse} explains the need to create a new class.
+Section \ref{compact} will then discusses how cubble relates to existing packages for spatial and temporal data analysis.
+
+The dataset \texttt{climate\_flat} is used to illustrate \pkg{cubble} functionality.
+This is a subset from the ORIGIN and contains the spatial variables, station id, latitude, longitude, elevation, station name, World Meteorology Organisation ID, in additional to daily temporal information, maximum and minimum temperature values and precipitation records for 2020.
 The first five rows of \texttt{climate\_flat} are shown below:
@@
 
-Spatio-temporal data can come in various formats and shapes and a cubble can be created from various types of raw data.
-This includes tibble and its variates in multiple tables (an example is provided in section \ref{covid}), and netCDF (detailed in section \ref{netcdf}).
-The function \code{as_cubble()} is used to create a cubble with three additional inputs: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order (longitude, latitude).
+A cubble can be created from various common spatio-temporal data formats, including basic R objects like tibble, tsibble, sf.
+Section \ref{covid} describes converting multiple tables into a cubble object and Section \ref{netcdf} illustrate how to convert a netCDF object.
+
+The function \code{as_cubble()} is used to create a cubble with three arguments: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order (longitude, latitude).
 The arguments \code{key} and \code{index} follow the wording in \pkg{tsibble} to describe the temporal order and multiple series while \code{coords} specifies the spatial location of each site.
-The code below creates a cubble out of \code{climate_flat} data with \code{id} as the key, \code{date} as the index, and \code{c(long, lat)} as the coordinates:
+The code below creates a cubble out of \code{climate_flat} (a single tibble) with \code{id} as the key, \code{date} as the index, and \code{c(long, lat)} as the coordinates:
 
@@
 The nested form can be used for those operations where the output is only indexed by the spatial identifier (\code{key}), but becomes inadequate when outputs need both a spatial and a temporal identifier (\code{key} and \code{index}).
-\pkg{cubble} also provides a long form, which expands the \code{ts} column and temporarily ``hides'' the spatial variables.
-The function \code{face_temporal()} is used to switch from a nested cubble into a long one.
+The \pkg{cubble} class also provides a long form, which expands the \code{ts} column and temporarily ``hides'' the spatial variables.
+The function \code{face_temporal()} is used to switch from the nested cubble into the long one.
 The first row in Figure \ref{fig:face} illustrates this operation where the focus of the cube now changes from the site-variable face to the time-variable face.
@@
 Sometimes, analysts may need to apply some variable transformation that involves both the spatial and temporal variables.
-An example of this is the transformation of temporal variables into the spatial dimension in glyph maps, which will be elaborated in section \ref{st_transformation}.
-This type of operation can be seen as flattening, or \emph{unfolding}, the cube into a 2D data frame.
+An example of this is the transformation of temporal variables into the spatial dimension in glyph maps {[}ref{]}.
+(Glyph maps will be explained in Section \ref{st_transformation}.) This type of operation can be seen as flattening, or \emph{unfolding}, the cube into a 2D data frame.
 Here the function \code{unfold()} moves the spatial variables \code{long} and \code{lat} into the long cubble:
@@
 
-\subsection[Why not just use]{Why not just use \code{dplyr::nest\_by()}?}
+\hypertarget{tidyverse}{%
+\subsection{Why not just use the existing tidyverse functions}\label{tidyverse}}
 
@@
 
-A \code{cubble} object can be created from separate spatial and temporal tables in a list, with other arguments (\code{key}, \code{index}, and \code{coords}) introduced in section \ref{create}.
+A \code{cubble} object can be created from separate spatial and temporal tables in a list, with other arguments (\code{key}, \code{index}, and \code{coords}) introduced in Section \ref{create}.
 \code{as_cubble()} will automatically check the matching of sites in both tables and emit a warning if a location has missing spatial or temporal information.

### 2022-05-25 2-0-cube.Rmd
section 2 tweaked

@@
 
-While a cubical data model is neat to view spatio-temporal data conceptually, a 3D data representation, or an array,  may not be convenient for data wrangling.
-There are two reasons for this: 1) While arrays are efficient for the computation on numerical values, spatio-temporal data, in general, contains more than just numerical variables.
-For example, wrangling on character strings and specific datetime classes is common for spatio-temporal data but may not be easily done with array objects.
-2) Creating new variables in an array is more complex than in a data frame since it would need to be defined in a 2D space, rather than a 1D vector, to be added into the array.
+While the data cube model is conceptually convenient for spatio-temporal data, a 3D data array is not sufficiently rich for data wrangling, for several reasons.
+Although arrays can be efficient for the computation on numerical values, spatio-temporal data typically includes various types of variables.
+For example, character strings and specific datetime classes are common.
+In addition, it will be generally useful to be able to create new variables which is trickier to manage in an array.
+Thus for convenient wrangling, we have opted for a list object as the `cubble` structure.
 

### 2022-05-25 jss/cubble-jss.tex
section 2 tweaked

@@
 
-Spatio-temporal data has a spatial component referring to the location of each observation, a temporal component that is recorded at regular or irregular time intervals, and may have multiple variables measured at each spatial and temporal value.
-The analysis of spatio-temporal data can fix the time to explore the spatial distribution, fix the spatial location to explore temporal trends, or dynamically explore the space and time simultaneously.
-For these tasks to be carried out easily, there is a need for a spatio-temporal data representation that underpins various operations.
-
-The \emph{Representing data} section of the SpatioTemproal CRAN task view \citep{ctvspatiotemporal} lists existing spatio-temporal data representation.
-Among them, \citet{spacetime} proposes four layouts: full grid layout, sparse grid layout, irregular layout, and trajectory.
-These layouts are implemented in the \pkg{spacetime} \citep{spacetime} package with the underlying spatial and temporal class: \pkg{sp} \citep{sp} and \pkg{xts} \citep{xts}.
-The \pkg{stars} \citep{stars} package has been a recent implementation to handle both raster and vector spatio-temporal data using array.
-It depends on the newer spatial package \pkg{sf} \citep{sf} and interfaces with the widely-used tidyverse verbs.
-
-{[}till here{]} One issue with wrangling spatio-temporal data is that the raw data does not naturally structured in any one of the representation above.
-Also, in data collection, spatio-temporal data is less often collected as a whole piece but fragmented spatial and temporal data for analysts to ensemble.
-For example, map data records the shape of a collection of area of interest is provided by official government website; geo-scientific data records the longitude and latitude coordinate of locations in the area along with other meta data relates to the location; and temporal variables indexed by both location and time.
-For spatio-temporal data analysis, merging all these sources can be challenging and joining these tables can go wrong because of the slightest unmatched of the linking variable from different sources.
-
-Another issue is how the concept of tidy data \citep{tidydata} should be applied to spatio-temporal data.
-The tidy data concept advocates the data to be shaped into 1) one row per observation; 2) one column per variable; and 3) one type of data per table.
-The long form data is preferred over wide data given the downstream software i.e.~\pkg{dplyr} and \pkg{ggplot2} for data wrangling and visualisation.
-However, the long form can be inefficient to store feature geometries, especially for large multipolygon over long time period.
-This poses the question on how to arrange spatial and temporal variables that makes analysing spatio-temporal data easier.
-
-This paper presents a new \proglang{R} package, \pkg{cubble}, to {[}overcome the problem{]} with organising variables in different levels when working with spatio-temporal data.
-The data structure focuses on connecting to the upstream data collection and is suitable for temporal data collected at the fixed location over time.
-Among the four spacetime layouts in \citet{spacetime}, it can be full grid layout, sparse grid layout, or irregular layout, but trajectory is not within the scope of this work.
-In the package a new data structure, also called \code{cubble}, is proposed to organise spatial and temporal variables as two forms of a single data object so that they can be wrangled separately while synchronised as a whole piece.
-With this new data object, analysts can spend less time writing codes to organise spatial and temporal variables and focus more on the exploratory data analysis itself.
-The software is available from the Comprehensive R Archive Network (CRAN) at {[}CRAN link{]}.
-
-The rest of the paper is divided as follows: Section \ref{cube} discusses the spatio-temporal cube to conceptualise spatio-temporal data.
+Spatio-temporal data has a spatial component referring to the location of each observation and a temporal component that is recorded at regular or irregular time intervals.
+It may also include multiple variables measured at each spatial and temporal values.
+With spatio-temporal data, one can fix the time to explore the spatial features of the data, fix the spatial location/s to explore temporal aspects, or dynamically explore the space and time simultaneously.
+In order to computationally explore the spatial, temporal and spatio-temporal faces of such data, the data needs to be stored and represented under a specific data object that allows the user to query, group and dissect all the data faces.
+
+The SpatioTemporal CRAN task view \citep{ctvspatiotemporal} gathers information about R packages designed for spatio-temporal data and it has a section on \emph{Representing data} that lists existing spatio-temporal data representations used in \proglang{R}.
+Among them, \citet{spacetime} summarises spatio-temporal data into three forms: time-wide, space-wide, and long formats.
+The associated package \pkg{spacetime} \citep{spacetime} implements four spatio-temporal layouts (full grid, sparse grid, irregular, and trajectory) to handle different space and time combinations.
+The \pkg{stars} \citep{stars} package has a new implementation to use dense arrays to represent spatio-temporal cubes.
+It also interfaces with \pkg{sf} \citep{sf}, a package commonly used for wrangling spatial data, and the \pkg{tidyverse} \citep{tidyverse} suite for general data wrangling and visualisation in \proglang{R}.
+
+Still, the data representation for spatio-temporal data can be further extended and there are two reasons for this.
+Firstly, the raw data sourced in the wild is less often presented in any one of the layouts above, and fitting the raw data into a data object can sometimes be difficult.
+More often, spatio-temporal data are collected in separate 2D tables and analysts need to assemble them into a whole piece before exploring the data.
+Examples of components of spatio-temporal data can be 1) areal data recording the shape of a collection of areas of interest; 2) geostatistical data storing the longitude and latitude coordinates of locations, typically also with other metadata related to the location, and; 3) temporal data of each location across time.
+
+The other reason is about tidy data concepts \citep{tidydata} and how they should be applied to spatio-temporal data.
+According to the tidy data principles, data should be structured into 1) one row per observation, 2) one column per variable, and 3) one type of data per table.
+The long form data is preferred over wide data form given the downstream software such as \pkg{dplyr} \citep{dplyr} and \pkg{ggplot2} \citep{ggplot2} for data wrangling and visualisation.
+However, the long form can be inefficient to store feature geometries, especially for large multipolygons for hourly, daily or sub-daily periods over years, which are extensively collected and handled, for example in time series analysis.
+This poses the question of how to arrange spatial and temporal variables in a way that would make data wrangling, visualizing and analysing spatio-temporal data easier.
+
+This paper presents a new \proglang{R} package, \pkg{cubble} which addresses the two issues mentioned above.
+In the package, a new data structure, also called \code{cubble}, is proposed to organise spatial and temporal variables as two forms of a single data object so that they can be wrangled separately or combined while being kept synchronised.
+Among the four spacetime layouts in \citet{spacetime}, \pkg{cubble} can be applied to full grid, sparse grid, or irregular, but not trajectory, which is outside the scope of this work.
+The software is available from the Comprehensive R Archive Network (CRAN) at \url{https://CRAN.R-project.org/package=cubble}.
+
+The rest of the paper is organized as follows: Section \ref{cube} introduces the proposed cube structure as a way to conceptualise multivariate spatio-temporal data.
 Section \ref{cubble} presents the main design and functionality of \pkg{cubble}.
-Section \ref{others} explains how cubble deals with more advanced considerations including data with hierarchical structure, data matching, how cubble fits with existing static and interactive visualisation tools, and spatio-temporal data transformation.
-Section \ref{examples} uses Australia weather station data and river level data as examples to demonstrate the use of \pkg{cubble}.
-An example on how \pkg{cubble} handles NetCDF data is also provided.
-Section \ref{conclude} concludes the paper.
+Section \ref{others} explains how cubble deals with more advanced considerations, including data with hierarchical structure, data matching and how cubble fits with existing static and interactive visualisation tools.
+Moreover we also illustrate how \pkg{cubble} deals with spatio-temporal data transformations.
+Section \ref{examples} uses Australian weather station data and river level data as examples to demonstrate the use of \pkg{cubble}.
+An example of how \pkg{cubble} handles NetCDF data is also provided.
+Section \ref{conclude} discuss the paper contributions and future directions.
 
@@
 
-Spatio-temporal data can be conceptualised using a cubical data model with reference to variables, time and space.
-This naturally motivates the usage of array to represent and store spatio-temporal data, as evident by satellite imageries, large climate models, among others.
-The work by \citet{lu_multidimensional_2018} has discussed the suitability of this representation through array algebra and practical spatio-temporal tasks.
-\textcolor{magenta}{SZ: is this too direct?}
-
-The cubical conceptual framework provides a generalisation to the operation can be done on the data for visualisation.
-A paper by \citet{bach_review_2014} has reviewed the temporal data visualisation based on space-time cube operations.
-Notice that the term space-time cube in their article ``does not need to involve spatial data'', but referring to ``an abstract 2D substrate that is used to visualize data at a specific time''.
-Despite its main focus is on temporal data, the mindset of abstracting out data representation to construct visualisation is applicable to our spatio-temporal data manipulation and visualisation.
+Spatio-temporal data can be conceptualised using a cubical data model with three axes which typically are, time, latitude and longitude.
+This abstraction can be useful for generalising operations and visualisation purposes: \citet{lu_multidimensional_2018} shows how array operations (select, scale, reduce, rearrange, and compute) can be mapped onto the cube; \citet{bach_review_2014} reviews the temporal data visualisation based on space-time cube operations.
+Notice that the term space-time cube in their article ``does not need to involve spatial data'', but refers to ``an abstract 2D substrate that is used to visualize data at a specific time''.
+Despite its main focus being on temporal data, the mindset of abstracting out data representation to construct visualisations, still applies to our spatio-temporal data manipulation and visualisation approaches.
+
+The most common data cube using the three axes (time, latitude, longitude) could be considered as taking snapshots of the space and stacking them across time.
+Here we define the spatio-temporal cube slightly differently by considering the three axes to be, time, site and variable/s.
+That is illustrated in the leftmost cube in Figure \ref{fig:cube-def}.
+The time axis is the same in both versions, while the site axis now captures both latitude and longitude.
+Finally, variables are stacked on this space-time canvas, with one observation per site and time point.
+This notion is adopted to avoid using hypercubes when describing multivariate spatio-temporal data and is the conceptual framework behind \pkg{cubble}.
+With this conceptual model, operations on spatio-temporal data can be mapped to operations on the cube and the rest of Figure \ref{fig:cube-def} show examples of slicing on site, time, and variable.
+
+While the data cube model is conceptually convenient for spatio-temporal data, a 3D data array is not sufficiently rich for data wrangling, for several reasons.
+Although arrays can be efficient for the computation on numerical values, spatio-temporal data typically includes various types of variables.
+For example, character strings and specific datetime classes are common.
+In addition, it will be generally useful to be able to create new variables which is trickier to manage in an array.
+Thus for convenient wrangling, we have opted for a list object as the \texttt{cubble} structure.
+
+\begin{CodeChunk}
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{../figures/diagram-keynotes/diagram-keynotes.007}
+
+}
+
+\caption[An illustration of the conceptual spatio-temporal cube with different slicing on time, site, and variable]{An illustration of the conceptual spatio-temporal cube with different slicing on time, site, and variable.
+For each axis, the slicing can be on a single value or a set of values.}\label{fig:cube-def}
+\end{figure}
+\end{CodeChunk}
 
@@
 
-This section will first introduce the core functions in the cubble class: \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and \code{unfold()} from subsection \ref{create} to \ref{unfold}.
-The next subsection addresses why bother a new class before subsection \ref{compact} finishing on the compatibility of cubble with existing packages in spatial and temporal analysis.
-
-Each core function is introduced with a short example accompanied and the data \texttt{climate\_flat} is used throughout this section.
-\texttt{climate\_flat} contains five weather stations in Australia with spatial information of each station: station id, latitude, longitude, elevation, station name and World Meteorology Organisation ID and also daily temporal information of date, maximum and minimum temperature and precipitation for 2020.
-Below prints out the first five rows of \texttt{climate\_flat}:
+Here we introduce the core functionalities of cubble.
+The main functions, \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and \code{unfold()} are described in Section \ref{create} - \ref{unfold}.
+Section \ref{tidyverse} explains the need to create a new class.
+Section \ref{compact} will then discusses how cubble relates to existing packages for spatial and temporal data analysis.
+
+The dataset \texttt{climate\_flat} is used to illustrate \pkg{cubble} functionality.
+This is a subset from the ORIGIN and contains the spatial variables, station id, latitude, longitude, elevation, station name, World Meteorology Organisation ID, in additional to daily temporal information, maximum and minimum temperature values and precipitation records for 2020.
+The first five rows of \texttt{climate\_flat} are shown below:
 
@@
 \hypertarget{create}{%
-\subsection{Create a cubble in the nested form}\label{create}}
-
-As described in the introduction, spatio-temporal data can come in various formats and shapes.
-A cubble can be created from various raw data, including tibble and its variates multiple tables (an example provided in section \ref{covid}), and netcdf (detailed in section \ref{netcdf}).
-The function \code{as_cubble()} is used to create a cubble with three additional components: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order of longitude and latitude.
-Argument \code{key} and \code{index} follow the wording in \pkg{tsibble} to define temporal order and multiple series while \code{coords} is a space specific argument to locate each site in two columns.
-The code below creates a cubble out of \code{climate_flat} with \code{id} as the key, \code{date} as the index, and \code{c(long, lat)} as the coordinates:
+\subsection{Create a cubble}\label{create}}
+
+A cubble can be created from various common spatio-temporal data formats, including basic R objects like tibble, tsibble, sf.
+Section \ref{covid} describes converting multiple tables into a cubble object and Section \ref{netcdf} illustrate how to convert a netCDF object.
+
+The function \code{as_cubble()} is used to create a cubble with three arguments: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order (longitude, latitude).
+The arguments \code{key} and \code{index} follow the wording in \pkg{tsibble} to describe the temporal order and multiple series while \code{coords} specifies the spatial location of each site.
+The code below creates a cubble out of \code{climate_flat} (a single tibble) with \code{id} as the key, \code{date} as the index, and \code{c(long, lat)} as the coordinates:
 
@@
 
-The cubble header provides some information about this data.
+Printing a cubble provides some information about the data.
 \code{id} is the variable name to identify each location and there are five unique locations.
-The bounding box is \code{[115.97, -32.94, 133.55, -12.42]} and the names of variable nested in the \code{ts} column, along with its type, are \code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.
-
-The created cubble is built from a \code{rowwise_df} class where each row forms a group and all the temporal variables are nested in a list column, hence it is also called the nested cubble.
-The rowwise structure makes it simpler to calculate on the list using \code{mutate()} and avoid the \code{map()} when working with list column.
-For example calculating the number of raining day can be done by:
-
-\begin{CodeChunk}
+The bounding box is \code{[115.97, -32.94, 133.55, -12.42]} and provides information about the coordinates in the data set.
+The third row shows the name and type of all variables nested in the \code{ts} column.
+In this example, it includes \code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.
+
+A cubble object is an extension of a \code{rowwise_df} object where each row forms a group.
+All the temporal variables are nested in a list column, hence it is also called the nested cubble.
+The rowwise structure makes it simpler to operate on the list using the \code{mutate()} syntax, which is simpler than the \code{purr::map()} when working with a list column.
+For example, calculating the number of rainy days can be done by:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> cubble_nested |>
++   mutate(rain_day = sum(ts$prcp != 0))
+\end{CodeInput}
 \begin{CodeOutput}
@@
 
-\hypertarget{switch-from-the-nested-cubble-into-the-long-form}{%
-\subsection{Switch from the nested cubble into the long form}\label{switch-from-the-nested-cubble-into-the-long-form}}
-
-The nested format is convenient for those operations whose output does not contain a time dimension.
-For those outputs that are cross-identified by the spatial and temporal identifier, a long cubble needs to be used.
-The function \code{face_temporal()} switches the cubble from the nested form to the long form and the first row in Figure \ref{fig:face} sketches the function in a diagram.
-This code shows how to switch the cubble just created into its long form:
+\hypertarget{change-focus-by-facing-the-time-variables}{%
+\subsection{Change focus by facing the time-variables}\label{change-focus-by-facing-the-time-variables}}
+
+The nested form can be used for those operations where the output is only indexed by the spatial identifier (\code{key}), but becomes inadequate when outputs need both a spatial and a temporal identifier (\code{key} and \code{index}).
+The \pkg{cubble} class also provides a long form, which expands the \code{ts} column and temporarily ``hides'' the spatial variables.
+The function \code{face_temporal()} is used to switch from the nested cubble into the long one.
+The first row in Figure \ref{fig:face} illustrates this operation where the focus of the cube now changes from the site-variable face to the time-variable face.
+This code switches the cubble just created into its long form:
 
@@
 
-Notice that the third line in the header now shows the name and type of spatial variables: \code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]} and these variables are now stored as a \code{spatial} attribute of the data:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> attr(cubble_long, "spatial")
-\end{CodeInput}
-\begin{CodeOutput}
-# A tibble: 5 x 6
-  id            lat  long  elev name           wmo_id
-  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl>
-1 ASN00009021 -31.9  116.  15.4 perth airport   94610
-2 ASN00010311 -31.9  117. 179   york            94623
-3 ASN00010614 -32.9  117. 338   narrogin        94627
-4 ASN00014015 -12.4  131.  30.4 darwin airport  94120
-5 ASN00015131 -17.6  134. 220   elliott         94236
-\end{CodeOutput}
-\end{CodeChunk}
-
-The long form is built form class \code{groupped_df} where all the observations from the same sites forms a group.
-
-\begin{CodeChunk}
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.008}
+The first line in printed cubble now shows the cubble in its long form and the third line has been changed to display the name and type of spatial variables: \code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]}.
+Unlike the nested form, the long cubble is built from a \code{grouped_df} object where all the observations from the same sites form a group.
+
+\begin{CodeChunk}
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{../figures/diagram-keynotes/diagram-keynotes.008}
 
@@
 In the first row, \code{face\_temporal} switches a cubble from the nested form into the long form and the focus has switched from the spatial aspect (the side face) to the temporal aspect (the front face).
-In the second row, \code{face\_spatial} switches a cubble back to the nested form from te long form and shift focus back to the spatial aspect.}\label{fig:face}
-\end{figure}
-\end{CodeChunk}
-
-\hypertarget{switch-a-long-cubble-back-to-the-nested-form}{%
-\subsection{Switch a long cubble back to the nested form}\label{switch-a-long-cubble-back-to-the-nested-form}}
-
-Manipulation on the spatial and temporal dimension can be an iterative process and analysts may need to go back and forth between the nested and long cubble.
-\code{face_spatial()}, the inverse of \code{face_temporal()}, switches the long cubble back to the nested form, as sketched in the second row in Figure \ref{fig:face}:
+In the second row, \code{face\_spatial} switches a cubble back to the nested form from the long form and shifts focus back to the spatial aspect.}\label{fig:face}
+\end{figure}
+\end{CodeChunk}
+
+\hypertarget{change-focus-back-to-the-site-variable-face}{%
+\subsection{Change focus back to the site-variable face}\label{change-focus-back-to-the-site-variable-face}}
+
+Wrangling spatio-temporal data can be seen as an iterative process in the spatial and temporal dimensions.
+Switching the focus back to the site-variable face can be accomplished by the function \code{face_spatial()}, which is the inverse of \code{face_temporal()}.
+The second row of Figure \ref{fig:face} illustrates the function, which is used as follows:
 
@@
 \end{CodeOutput}
+\begin{CodeInput}
+R> identical(cubble_nested, cubble_back)
+\end{CodeInput}
+\begin{CodeOutput}
+[1] TRUE
+\end{CodeOutput}
 \end{CodeChunk}
@@
 \hypertarget{unfold}{%
-\subsection{Flatten spatial variables into the long cubble}\label{unfold}}
-
-Sometimes, analysts may need to apply some variable transformation that involves both the spatial and temporal variable.
-An example of this is the transformation of temporal variables into the spatial dimension in glyph maps, which will be elaborated in section \ref{st_transformation}.
-This type of operations can be seen as flattening, or \emph{unfolding}, the cube into a 2D data frame.
+\subsection{Unfold spatial variables into the long cubble}\label{unfold}}
+
+Sometimes, analysts may need to apply some variable transformation that involves both the spatial and temporal variables.
+An example of this is the transformation of temporal variables into the spatial dimension in glyph maps {[}ref{]}.
+(Glyph maps will be explained in Section \ref{st_transformation}.) This type of operation can be seen as flattening, or \emph{unfolding}, the cube into a 2D data frame.
 Here the function \code{unfold()} moves the spatial variables \code{long} and \code{lat} into the long cubble:
@@
 
-This function should generally be used in the last step of the analysis since it is a temporary operation, meaning these added spatial variables are not stored in the long form and will disappear if switched to the nested form and then switched back:
+This function should generally be used in the last step of the analysis since it is a temporary operation, meaning that these added spatial variables will disappear, if switched to the nested form and then switched back:
 
@@
 
-\subsection[Why not just use]{Why not just use \code{dplyr::nest\_by()}}
-
-It may naturally raise the question that why bother a new data structure since the nested form can already be created by the \pkg{dplyr} function \code{nest_by()}?
-While \code{nest_by()} and \code{unnest()} can mimic the nested form in cubble, cubble can be seen as an attempt to arrange variables from different observational units into a single object, specifically for spatio-temporal data.
-While \texttt{nest\_by()} solves this problem on the spatial side, it doesn't solve the temporal side.
-For temporal data, a combined representation with spatial variables is useful for spatio-temporal operations but its duplication can be a burden to operate on.
-Also, spatial variables are dropped when summarised and you will need to either specified all of them in the \code{group_by()} before \code{summarise}, or add a joining after \code{summarise()}.
+\hypertarget{tidyverse}{%
+\subsection{Why not just use the existing tidyverse functions}\label{tidyverse}}
+
+Some readers may question why a new data structure is needed rather than using \code{dplyr::nest_by()} on the combined data to directly create a list-column.
+The reason is that cubble is specifically designed to utilize the spatio-temporal structure when arranging variables from different observational units into a single object.
+Moreover, it enables easy pivoting between purely spatial, purely temporal, or unfolded into a combined form.
 
@@
 
-In the nested cubble created, each element in the list-column \code{ts} is of \code{tbl_ts} class and operations available to the tsibble class is still valid under cubble.
-For example, the code below calculates two features of the maximum temperature:
+When a nested cubble is created, each element in the list-column \code{ts} is of \code{tbl_ts} class and operations available to the tsibble class are still valid within the cubble.
+For example, the code below calculates two time series features (mean and variance) of the maximum temperature variable:
 
@@
 
-As a spatial data object, \code{sf} creates a simple feature geometry list-column (\code{sfc}) in the data frame to provide spatial operations on various geometry types (\code{POINT}, \code{LINESTRING}, \code{POLYGON}, \code{MULTIPOLYGON}, etc).
-These spatial operations are also valuable for spatio-temporal data analysis, but an \code{sf} object \emph{usually} does not contain temporal variables.
-This means \code{sf} cannot be directly cast into a \code{cubble}, however, \code{cubble} does support \code{sfc} columns in the nested form and spatial operations applied to the \code{sfc} column in \code{sf} can still be applied to the \code{sfc} column in a cubble.
-The following example shows how to create an \code{sfc} column of \code{POINT} type from latitude and longitude in cubble.
-Then \code{sf::st_within} is used to add the state \code{MULTIPOLYGON} of each weather station before a coordinate transformation is made.
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> library(sf)
-R> # create a cubble
-R> cb <- climate_flat |>
-+   cubble::as_cubble(key = id, index = date, coords = c(long, lat))
-R> 
-R> aus <- ozmaps::abs_ste
-R> 
-R> dt <- cb |>
-+   mutate(
-+     # create `sfc` column based on long and lat
-+     ll = st_sfc(
-+       purrr::map2(long, lat, ~st_point(c(.x, .y))),
-+       crs = st_crs(aus)),
-+ 
-+     # append state multi-polygon based on the `sfc` created
-+     state = aus$geometry[st_within(ll, aus, sparse = FALSE)],
-+ 
-+     # adopt a different projection: lambert conformal conic (EPSG:3112)
-+     state = st_transform(state, crs = "EPSG:3112")
-+     )
-R> 
-R> dt
-\end{CodeInput}
-\begin{CodeOutput}
-# cubble:   id [5]: nested form
-# bbox:     [115.97, -32.94, 133.55, -12.42]
-# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
-  id            lat  long  elev name   wmo_id ts                        ll
-  <chr>       <dbl> <dbl> <dbl> <chr>   <dbl> <list>           <POINT [°]>
-1 ASN00009021 -31.9  116.  15.4 perth~  94610 <tibble> (115.9764 -31.9275)
-2 ASN00010311 -31.9  117. 179   york    94623 <tibble>  (116.765 -31.8997)
-3 ASN00010614 -32.9  117. 338   narro~  94627 <tibble> (117.1797 -32.9342)
-4 ASN00014015 -12.4  131.  30.4 darwi~  94120 <tibble> (130.8925 -12.4239)
-5 ASN00015131 -17.6  134. 220   ellio~  94236 <tibble> (133.5407 -17.5521)
-# ... with 1 more variable: state <MULTIPOLYGON [m]>
-\end{CodeOutput}
-\end{CodeChunk}
-
-An \pkg{s2} \code{lnglat} vector can similarly be created as an \code{sfc} in cubble before using any \code{s2}-prefixed function:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> library(s2)
-R> # Western Australia map
-R> wa <- ozmaps::abs_ste |> filter(NAME == "Western Australia")
-R> 
-R> # mutate a `s2_lnglat` vector on `cb` created in the last chunk
-R> cb |>
-+   mutate(ll = s2_lnglat(long, lat)) |>
-+   filter(s2_within(ll, wa))
-\end{CodeInput}
-\begin{CodeOutput}
-# cubble:   id [3]: nested form
-# bbox:     [115.97, -32.94, 117.18, -31.89]
-# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
-  id            lat  long  elev name          wmo_id ts                 ll      
-  <chr>       <dbl> <dbl> <dbl> <chr>          <dbl> <list>             <s2_lng>
-1 ASN00009021 -31.9  116.  15.4 perth airport  94610 <tibble [366 x 4]> (115.97~
-2 ASN00010311 -31.9  117. 179   york           94623 <tibble [366 x 4]> (116.76~
-3 ASN00010614 -32.9  117. 338   narrogin       94627 <tibble [366 x 4]> (117.17~
-\end{CodeOutput}
-\end{CodeChunk}
+As a spatial data object, \code{sf} creates a feature geometry list-column (\code{sfc}) in the data frame to provide spatial operations on various geometry types such as (\code{POINT}, \code{LINESTRING}, \code{POLYGON}, \code{MULTIPOLYGON}, etc).
+When creating a cubble a \code{sf} object can be supplied as the spatial table.
+Once that is included in a cubble, methods for the \code{sfc} class can be applied in the nested form inside the cubble object.
+An example of this is shown in \ref{covid}, which also handles the case when the site identifiers in the two tables do not match exactly.
+A spatial data object with an \code{s2} vector can also be the input for the spatial table in \pkg{cubble}.
 
@@
 
-NetCDF data has two main components: \emph{dimension} for defining the spatio-temporal grid (longitude, latitude, and time) and \emph{variable} that populates the defined grid.
-Attributes are usually associated with dimension and variable in the NetCDF format data and a \href{http://cfconventions.org/}{metadata convention for climate and forecast} has been designed to standardise the format of the attributes.
-A few packages in R exists for manipulating NetCDF data and this includes a high-level R interface: \pkg{ncdf4} \citep{ncdf4}, a low-level interface that calls C-interface: \pkg{RNetCDF} \citep{rnetcdf, michna2013rnetcdf}, and a tidyverse implementation: \pkg{tidync} \citep{tidync}.
+NetCDF data is another format that is commonly used for storing spatio-temporal data.
+It has two main components: \emph{dimension} for defining the spatio-temporal grid (longitude, latitude, and time) and \emph{variable} that populates the defined grid.
+Attributes are usually associated with dimensions and variables in the NetCDF format data and a \href{http://cfconventions.org/}{metadata convention for climate and forecast} has been designed to standardise the format of the attributes.
+A few packages in R exist for manipulating NetCDF data and these include a high-level R interface: \pkg{ncdf4} \citep{ncdf4}, a low-level interface that calls a C-interface: \pkg{RNetCDF} \citep{rnetcdf, michna2013rnetcdf}, and a tidyverse implementation: \pkg{tidync} \citep{tidync}.
 
@@
 
-The memory limit with NetCDF data in cubble depends on the number of longitude grid points times the number of latitude grid points times the number of time grid points times the number of variables.
-Cubble can handle slightly more than 300 x 300 (longitude x longitude) grid points for 3 variables in one year.
-The spatial grid can be reduced to trade for longer time periods and more variables.
+The memory limit with NetCDF data in cubble depends on the number of longitude grid points \(\times\) the number of latitude grid points \(\times\) the number of time grid points \(\times\) the number of variables.
+Cubble can handle slightly more than 300 \(\times\) 300 (longitude \(\times\) longitude) grid points for three daily variables in one year.
 A 300 by 300 spatial grid can be a bounding box of {[}100, -80, 180, 0{]} at 0.25 degree resolution or a global bounding box {[}-180, -90, 180, -90{]} at 1 degree resolution.
-Subsetting longitude and latitude grids is available through \code{long_range} and \code{lat_range} if the NetCDF file has finer resolution than needed.
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> # Assume my_ncdf has bounding box of [-180, -90, 180, -90]
+The spatial grid can also be subsetted to trade for longer time periods and more variables through \code{long_range} and \code{lat_range} if the NetCDF file has finer resolution than needed:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> # Assume my_ncdf has a bounding box of [-180, -90, 180, -90]
 R> # at 0.25 degree resolution and subset it to have
@@
 
-Spatial locations can have grouping structure either inherent to the data or formed by clustering.
-Rather than analysing variables in the site level, summarised variables in the cluster level can give a crisper picture of local areas.
-In cubble, \code{switch_key()} can be used to create a new level of grouping of spatial locations by specifying a clustering variable.
-Figure \ref{fig:illu-hier} illustrates the relationship of cubbles at station and cluster level, in both the long and nested form.
-By specifying \code{cluster_nested <- station_nested \%>\% switch_key(key = cluster)}, the cubble re-defines the cubble key from \code{id} in \code{station_nested} to \code{cluster} in \code{cluster_nested}.
-All the spatial variables variant to \code{cluster} are now nested into a \code{.val} column and cluster level variables can be computed in the same fashion as station level variables in \code{station_nested}.
-
-\begin{CodeChunk}
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.003}
+Spatial locations can have grouping structures either inherent to the data or obtained by clustering.
+In those cases, rather than analysing variables at the site level, it might be of interest to summarised variables at a cluster or group level, as that can give a crisper picture of local areas.
+In cubble, \code{switch_key()} can be used to create a new grouping level of spatial locations by specifying a clustering variable.
+Figure \ref{fig:illu-hier} illustrates the relationship of cubbles at station and cluster level, in both the long and nested forms.
+By specifying \code{cluster_nested <- station_nested \%>\% switch_key(key = cluster)}, the cubble redefines the cubble key from the \code{id} column in \code{station_nested} to the \code{cluster} column in \code{cluster_nested}.
+All the spatial variables variant to the \code{cluster} column are now nested into a \code{.val} column and cluster level variables can be computed in the same fashion as station level variables in \code{station_nested}.
+
+\begin{CodeChunk}
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{../figures/diagram-keynotes/diagram-keynotes.003}
 
@@
 
-\hypertarget{data-fusion-and-matching}{%
-\subsection{Data fusion and matching}\label{data-fusion-and-matching}}
-
-One task that may interest analysts in spatio-temporal data is to find how similar the time series from nearby sites are.
-This can be seen as a matching problem \citep{stuart2010matching, mcintosh2018using} that pairs up similar similar time series in nearby locations or a data fusing exercise that merges data collected from different sources \citep{cocchi2019data}.
+\hypertarget{matching}{%
+\subsection{Data fusion and matching}\label{matching}}
+
+One task that may interest analysts in spatio-temporal data is to find how similar time series, from nearby sites, are.
+This problem can be seen as a matching problem \citep{stuart2010matching, mcintosh2018using} that pairs up similar time series in nearby locations or a data fusing exercise that merges data collected from different sources \citep{cocchi2019data}.
 \code{match_sites()} in \pkg{cubble} provides a simple algorithm for this task.
-The algorithm first matches the two data sources spatially through computing the pairwise distance on latitude and longitude.
-Pairs that pass the spatial matching are then matched temporally through computing the number of matched peaks within a fixed length window.
-Figure \ref{fig:illu-matching} illustrates this temporal matching in more details.
-Given two series \code{A} and \code{a}, 3 peaks have been picked in each series.
-An interval, with default length of 5, is constructed for each peak in series \code{A} and the peaks in series \code{a} are tested against whether they fall into the any of the intervals.
-In this illustration, there are 2 matches for these two series.
+The algorithm first matches the two data sources spatially by computing the pairwise distance on latitude and longitude.
+After that, pairs that pass the spatial matching are then matched temporally by computing the number of matched peaks within a fixed length moving window.
+Figure \ref{fig:illu-matching} illustrates this temporal matching in more detail.
+Given two series \code{A} and \code{a}, three peaks have been identified in each series.
+An interval, with default length of five, is constructed for each peak in series \code{A}, while the peaks in series \code{a} are tested against whether they fall into any of the intervals.
+In this illustration, there are two matches for these two series.
 Several arguments are available in \code{match_sites()} to fine-tune the matching:
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/illu-matching}
+{\centering \includegraphics[width=1\linewidth]{../figures/illu-matching}
 
@@
 
-Cubble fits in naturally with the interactive graphic pipeline discussed in the literature \citep{buja1988elements, buja1996interactive, sutherland2000orca, xie2014reactive, cheng2016enabling}.
+The linking structure of two forms in cubble fits naturally with the interactive graphic pipeline discussed in the literature \citep{buja1988elements, buja1996interactive, sutherland2000orca, xie2014reactive, cheng2016enabling}.
 Diagram \ref{fig:illu-interactive} illustrates how linking works from the map to the time series in cubble.
@@
 All the observations that share the same \code{id} are then activated and this includes other points in the same time series in the long cubble and the corresponding observation of site in the nested cubble.
-These activated observations will then being reflected in the updated plots and Diagram \ref{fig:illu-interactive-2} in the Appendix illustrates this process.
-
-\begin{CodeChunk}
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.004}
+These activated observations will then be reflected in the updated plots and Diagram \ref{fig:illu-interactive-2} in the Appendix illustrates this process.
+
+\begin{CodeChunk}
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{../figures/diagram-keynotes/diagram-keynotes.004}
 
@@
 
-Spatio-temporal transformation is a useful technique to extract information form spatio-temporal data.
-Glyph maps \citep{Wickham2012-yr} transform the time coordinates into the space coordinates to plot the time series of different locations on the map.
-Calendar plots \citep{wang2020calendar} reconstructs time into calendar-based grid to discover weekday and weekend pattern.
+Different types of transformations can be useful to extract new information from spatio-temporal data.
+Glyph maps \citep{Wickham2012-yr} transform the time coordinates into space coordinates to plot the time series of different locations on the map.
+Calendar plots \citep{wang2020calendar} reconstruct time into a calendar-based grid to discover weekday and weekend pattern.
 Projection, or linear combination, of variables summarises multivariate information into lower dimension to further digest.
@@
 In \proglang{R}, \pkg{GGally} implements glyph maps through the \code{glyphs()} function.
-The function constructs a data frame with calculated position (\code{gx}.
-\code{gy}.
-\code{gid}) of each point on the time series using linear algebra (Equation 1 and 2 in \citet{Wickham2012-yr}).
+The function constructs a data frame with calculated position (\code{gx}, \code{gy}, \code{gid}) of each point on the time series using linear algebra (Equation 1 and 2 in \citet{Wickham2012-yr}).
 The data can then be piped into \code{ggplot} to create the glyph map as:
@@
 
-This example illustrate casting different sources into a cubble object.
-Victoria government of Australia provides daily COVID case number by date, source, and local government area (LGA).
-This data can be useful for visualising COVID spread when combined with LGA map information, which is available from Australian Bureau of Statistics.
-Variable \code{covid} and \code{lga} show the first 5 rows of these two data:
+The Victoria State Government in Australia provides daily COVID information about the source, and local government area (LGA) of the recorded cases.
+This data can be used to visualise COVID spread when combined with map information on LGA, available from the Australian Bureau of Statistics.
+The first five rows these data sets are printed below:
 
@@
 Dimension:     XY
-Bounding box:  xmin: 142.3643 ymin: -38.67876 xmax: 147.386 ymax: -36.39269
+Bounding box:  xmin: 142.3535 ymin: -38.67876 xmax: 147.3909 ymax: -36.39269
 Geodetic CRS:  WGS 84
     lga_code_2018            lga state_code_2016 state_name_2016 areasqkm_2018
-1         20110     Alpine (S)               2        Victoria     4788.1568
-2         20260    Ararat (RC)               2        Victoria     4211.1171
-3         20570   Ballarat (C)               2        Victoria      739.0321
-4         20660    Banyule (C)               2        Victoria       62.5402
-5         20740 Bass Coast (S)               2        Victoria      865.8095
+132         20110     Alpine (S)               2        Victoria     4788.1568
+133         20260    Ararat (RC)               2        Victoria     4211.1171
+134         20570   Ballarat (C)               2        Victoria      739.0321
+135         20660    Banyule (C)               2        Victoria       62.5402
+136         20740 Bass Coast (S)               2        Victoria      865.8095
     cent_long  cent_lat                       geometry
-1  146.9742 -36.85357 MULTIPOLYGON (((146.7258 -3...
-2  142.8432 -37.47271 MULTIPOLYGON (((143.1807 -3...
-3  143.7815 -37.49286 MULTIPOLYGON (((143.6622 -3...
-4  145.0851 -37.73043 MULTIPOLYGON (((145.1357 -3...
-5  145.5581 -38.50730 MULTIPOLYGON (((145.5207 -3...
-\end{CodeOutput}
-\end{CodeChunk}
-
-To cast these two data into a cubble object, supply them as a list with name \code{spatial} and \code{temporal} and other arguments (\code{key}, \code{index}, and \code{coords}).
-\code{as_cubble()} will automatically check the match of spatial and temporal information in the two datasets and warnings will emit if a location has missing spatial or temporal information.
+132  146.9742 -36.85357 MULTIPOLYGON (((146.7258 -3...
+133  142.8432 -37.47271 MULTIPOLYGON (((143.1807 -3...
+134  143.7815 -37.49286 MULTIPOLYGON (((143.6622 -3...
+135  145.0851 -37.73043 MULTIPOLYGON (((145.1357 -3...
+136  145.5581 -38.50730 MULTIPOLYGON (((145.5207 -3...
+\end{CodeOutput}
+\end{CodeChunk}
+
+A \code{cubble} object can be created from separate spatial and temporal tables in a list, with other arguments (\code{key}, \code{index}, and \code{coords}) introduced in section \ref{create}.
+\code{as_cubble()} will automatically check the matching of sites in both tables and emit a warning if a location has missing spatial or temporal information.
 
@@
 $others$temporal
-[1] "Interstate"       "Overseas"         "Queenscliffe (B)" "Unknown"
+[1] "Interstate" "Overseas"   "Unknown"
 
 $others$spatial
-character(0)
-\end{CodeOutput}
-\end{CodeChunk}
-
-\pkg{cubble} will make an attempt to pair the unmatched sites as well as showing unpaired locations. These can be useful information to help clean up the data.
+[1] "No usual address (Vic.)"               
+[2] "Migratory - Offshore - Shipping (Vic.)"
+\end{CodeOutput}
+\end{CodeChunk}
+
+\pkg{cubble} will attempt to pair the unmatched sites as well as those that cannot be paired. These can be helpful to clean up the data before creating the cubble again:
 
@@
 R> covid <- covid %>% filter(!lga %in% pair$others$temporal)
-\end{CodeInput}
-\end{CodeChunk}
-
-After cleaning, you can cast the data into cubble with the same syntax without warnings:
-
-\begin{CodeChunk}
-\begin{CodeInput}
+R> 
 R> cb <- as_cubble(data = list(spatial = lga, temporal = covid),
@@
 
-\hypertarget{australia-historical-maximum-temperature}{%
-\subsection{Australia historical maximum temperature}\label{australia-historical-maximum-temperature}}
-
-Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world.
-The dataset \code{weatherdata::historical_tmax} extracts the maximum temperature for 236 Australian stations from GHCN with starting from year 1969.
+\hypertarget{australian-historical-maximum-temperature}{%
+\subsection{Australian historical maximum temperature}\label{australian-historical-maximum-temperature}}
+
+The Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world.
+The dataset \code{weatherdata::historical_tmax} extracts the maximum temperature for 236 Australian stations from the GHCN starting from year 1969.
 \code{weatherdata::historical_tmax} is already in a cubble, with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates.
-This example compares the maximum temperature in two periods: 1971 - 1975 and 2016 - 2020 for stations in Victoria and New South Wales.
-
-Stations in the two states can be subsetted by the station number: Australia GHCN station number starts with ``ASN00'' and followed by the \href{http://www.bom.gov.au/climate/cdo/about/site-num.shtml}{Bureau of Meteorology (BOM) station number}, where the 2nd and 3rd digit (7th and 8th in the GHCN number) define the state of the station.
-New South Wales stations start from 46 to 75 and Victoria stations follow from 76 to 90.
-Filtering Victoria and New South Wales stations is an operation in the spatial dimension and hence uses the nested form:
+This example compares the maximum temperature in two periods: 1971-1975 and 2016-2020 for stations in Victoria and New South Wales.
+
+Stations in the two states can be subsetted by the station number: the Australian GHCN station numbers start with ``ASN00'' and are followed by the \href{http://www.bom.gov.au/climate/cdo/about/site-num.shtml}{Bureau of Meteorology (BOM) station number}.
+The second and third digits of the BOM station number (7th and 8th in the GHCN number) define the state of the station; the range 46-75 corresponds to New South Wales stations and the range 76-90 corresponds to Victorian stations.
+Filtering Victoria and New South Wales stations is a spatial operation and hence uses the nested form:
 
@@
 
-A few stations don't have records during 1971 - 1975 and further investigation shows that while the first and last year of each series is recorded, the missing years in this period is not reported.
-These stations are filtered out by examining whether the summarised time series has a total of 24 months.
+A few stations do not have records during the time period 1971-1975 and further investigation shows that while the first and last year of each series is recorded, the missing years in this period are not reported.
+These stations are filtered out by examining whether the summarised time series has 24 months.
 The long cubble needs to be switched to the nested form for this spatial operation using \code{face_spatial()}:
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.7\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/basic-manip}
+{\centering \includegraphics[width=1\linewidth,height=0.7\textheight]{../figures/basic-manip}
 
@@
 
-In the previous example, there has already been some overlapping of the glyphs for a few stations near (151E, 34S) and (152E, 33S) and this will be a problem when mapping more stations in the national level.
-Aggregation can be helpful to group series into clusters before visualising the cluster with glyph map.
-This example shows how to organise data at both level with \code{switch_key()}.
-
-\code{weatherdata::climate_full}, also extracted from GHCN, records daily precipitation and maximum/ minimum temperature for 639 stations in Australia from 2016 to 2020.
-A simple kmean algorithm based on the distance matrix is used to create 20 clusters.
-This creates \code{station_nested} as a station level nested cubble with a cluster column indicating the group each station belongs to.
-More complex algorithms can be used for other problem, as long as there is a mapping from each station to a cluster.
+In the previous example, there has already been some overlapping of the glyphs for a few stations near (151E, 34S) and (152E, 33S), which is a problem when mapping more stations at the national level.
+Aggregation can be helpful in grouping series into clusters before visualising the clusters with a glyph map.
+This example shows how to organise data at both levels with \code{switch_key()}.
+
+\code{weatherdata::climate_full}, also extracted from the GHCN, records daily precipitation and maximum/minimum temperature for 639 stations in Australia from 2016 to 2020.
+A simple \(k\)-means algorithm based on the distance matrix is used to create 20 clusters.
+The dataset \code{station_nested} is a nested cubble with a cluster column indicating the group to which each station belongs.
+More advanced clustering algorithms can be used for other applications, as long as there is a mapping from each station to a cluster.
 
@@
 
-Long form cubble at both levels can be accessed through stretching the nested form and with access to both station and cluster level cubbles, various plots can be made to understand the cluster.
-Figure \ref{fig:basic-agg} shows two example plots that can be made with this data: subplot A is a glyph map made with the cluster level cubble in the long form and subplot B inspects the station membership of each cluster using the station level cubble in the nested form.
-
-\begin{CodeChunk}
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/basic-agg}
+Long form cubble at both levels can be accessed through stretching the nested form and with access to both station and cluster-level cubbles, various plots can be made to understand the cluster.
+Figure \ref{fig:basic-agg} shows two example plots that can be made with this data.
+Subplot A is a glyph map made with the cluster level cubble in the long form and subplot B inspects the station membership of each cluster using the station level cubble in the nested form.
+
+\begin{CodeChunk}
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{../figures/basic-agg}
 
@@
 
-Bureau of Meteorology collects \href{http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full}{water data} from river gauges and this includes variables: electrical conductivity, turbidity, water course discharge, water course level, and water temperature.
-In particular, water level will interactive with precipitation from the climate data since rainfall will raise the water level in the river.
-Figure \ref{fig:matching-map} gives the location of available weather station and water gauges in Victoria.
-
-\begin{CodeChunk}
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/matching-map}
+The Bureau of Meteorology collects \href{http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full}{water data} from river gauges.
+The collected variables include: electrical conductivity, turbidity, watercourse discharge, watercourse level, and water temperature.
+In particular, water level will interact with precipitation from the climate data since rainfall will raise the water level in the river.
+Figure \ref{fig:matching-map} gives the location of available weather stations and water gauges in Victoria.
+
+\begin{CodeChunk}
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{../figures/matching-map}
 
@@
 
-From the map, a few water gauges and weather stations are close to each other and the fluctuation of the water level could be matched up with precipitation measured by the climate station.
-As introduced in Section 3.2, \code{match_sites()} can be used to match one source of data with another source in a cubble.
+From the map, one can see that a few water gauges and weather stations are close to each other and the fluctuation of the water level could be matched up with the precipitation measured by the nearby climate station.
+As introduced in Section \ref{matching}, \code{match_sites()} can be used to match one source of data with another source in a cubble.
 Here \code{Water_course_level} in \code{river} will be matched to \code{prcp} in \code{climate} in 2020.
 The two datasets need to be specified as the first two arguments and the variable to match can be specified in \code{temporal_by} using the \code{by} syntax in \code{join}.
-\code{temporal_independent} controls the variable used to construct the interval and the goal here is to see if precipitation will be reflected by the water level in the river.
-This puts precipitation \code{prcp}, as the independent.
-Given there is one year worth of data, the number of peak (\code{temporal_n_highest}) to consider is slightly raised from a default 20 to 30 and \code{temporal_min_match} is raised accordingly To return all the pairs of the match, \code{temporal_min_match} can be set to 0.
+\code{temporal_independent} controls the variable used to construct the interval and the goal here is to see if the water level in the river will reflect precipitation.
+This puts precipitation \code{prcp}, as the independent variable.
+Given there is one year's worth of data, the number of peaks (\code{temporal_n_highest}) to consider is slightly raised from a default of 20 to 30 and \code{temporal_min_match} is raised accordingly.
+To return all the pairs of the match, \code{temporal_min_match} can be set to 0.
 
@@
 
-Figure \ref{fig:matching} plots the matched pairs on the map or to view the matched series.
-There are four pairs of matches, which all locates in the middle Victoria and the concurrent increase of precipitation and water level can be observed.
-
-\begin{CodeChunk}
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/matching}
+Figure \ref{fig:matching} plots the matched pairs on the map.
+There are four pairs of matches, all located in the middle of Victoria and the expected concurrent increase in precipitation and water level can be observed.
+
+\begin{CodeChunk}
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{../figures/matching}
 
@@
 
-ERA5 data \citep{hersbach2020era5} is the latest reanalysis of global atmosphere, land surface, and ocean waves from 1950 onwards and is available in the NetCDF format from European Centre for Medium-Range Weather Forecasts (ECMWF).
-The data can be directly downloaded from \href{https://cds.climate.copernicus.eu/cdsapp\#!/dataset/reanalysis-era5-pressure-levels?tab=overview}{Copernicus Climate Data Store (CDS)} website or programmatically via an R package \pkg{ecmwfr} \citep{ecwmfr}.
-The \code{era5-pressure} data contains variable \emph{specific humidity} and \emph{geopotential} on the 10 hPa pressure level on four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04.
+The ERA5 dataset \citep{hersbach2020era5} is the latest reanalysis of global atmosphere, land surface, and ocean waves from 1950 onwards and is available in the NetCDF format from the European Centre for Medium-Range Weather Forecasts (ECMWF).
+The data can be directly downloaded from \href{https://cds.climate.copernicus.eu/cdsapp\#!/dataset/reanalysis-era5-pressure-levels?tab=overview}{Copernicus Climate Data Store (CDS)} website or programmatically via the R package \pkg{ecmwfr} \citep{ecwmfr}.
+The \code{era5-pressure} data contains the \emph{specific humidity} and \emph{geopotential} variables on the 10 hPa pressure level on four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04.
 Once downloaded, the data can be read into a cubble as:
@@
 It shows the southern polar vortex splits into two on 2002-09-26 and further splits into four on 2002-10-04 in the stratosphere.
-Readers interested in the analysis of this figure can refer to \citet{hersbach2020era5}, \citet{simmons2020global} and \citet{simmons2005ecmwf} for more details.
-
-\begin{CodeChunk}
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/netcdf}
+Readers interested in the analysis of this figure can refer to \citet{hersbach2020era5}, \citet{simmons2020global}, and \citet{simmons2005ecmwf} for more details.
+
+\begin{CodeChunk}
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{../figures/netcdf}
 
@@
 
-\hypertarget{interative-graphic}{%
-\subsection{Interative graphic}\label{interative-graphic}}
-
-With spatio-temporal data, users may wish to make plots to learn the spatial distribution of a variable, or to find patterns, such as trend or seasonality, in the time series.
-Combining this two types of plot with interactivity let users to link between points on the map and the corresponding time series to explore the spatial and temporal dimension of the data simultaneously.
-Below is an example that describes the process of building an interactive graphic with \pkg{cubble} and \pkg{crosstalk} The example explores the variation of monthly temperature range with \code{weatherdata::climate_full} data.
-
-The temperature range is calculated as the difference between \code{tmax} and \code{tmin} and its monthly average over 2016 - 2020 is taken before calculating the variance.
+\hypertarget{interactive-graphic}{%
+\subsection{Interactive graphic}\label{interactive-graphic}}
+
+With spatio-temporal data, users may wish to make plots to learn the spatial distribution of a variable or to find patterns, such as trend or seasonality, in the time series.
+Combining these two types of plots with interactivity lets users link between points on the map and the corresponding time series to explore the spatial and temporal dimensions of the data simultaneously.
+Below is an example that describes the process of building an interactive graphic with \pkg{cubble} and \pkg{crosstalk}.
+The example explores the variation of monthly temperature range with \code{weatherdata::climate_full} data.
+
+The temperature range is calculated as the difference between \code{tmax} and \code{tmin} and its monthly average over 2016-2020 is taken before calculating the variance.
 A \code{SharedData} object is constructed for each form of the cubble and the same \code{group} argument ensures the cross-linking of the two forms via the common \code{id} column.
-The spatial map and time series plot are then made with each \code{SharedData} objects separately.
-In this example, stations on the Australia map, made from the nested form, are coloured by the calculated variance and a ribbon band is constructed using the long form cubble to show the maximum and minimum temperature of each station across month.
-With a different dataset, users are free to calculate any per station measure in the nested form or to make any time-wise summary of the data in the long form to customise the spatial or temporal view.
+The spatial map and time series plot are then made with each \code{SharedData} object separately.
+In this example, stations on the Australia map, made from the nested form, are coloured by the calculated variance.
+A ribbon band is constructed using the long form cubble to show each station's maximum and minimum temperature across month.
+With a different dataset, users can calculate other per station measure in the nested form or make other time-wise summary of the data in the long form to customise the spatial or temporal view.
 The cross-linking between the two plots is always safeguarded by the shared \code{id} column embedded in the cubble structure.
-Below is the pseudo code that outlines the process to construct an interactive graphic described above:
+Below is the pseudo-code that outlines the process to construct an interactive graphic described above:
 
@@
 In Figure \ref{fig:interactive-linking}, the first row shows the initial view of the interactive graphic.
-On the map, most regions in Australia have low variance of temperature range while the north-west coastline, bottom of South Australia, and Victoria stands out with larger monthly changes.
+Most regions in Australia have low variance of temperature range, while the north-west coastline, bottom of South Australia, and Victoria stand out with larger monthly changes.
 In the second row, Mount Elizabeth is selected on the map given its high variance colour on the initial map and this links to the ribbon on the right.
-The third row the lowest temperature in August and this corresponds to Thredbo AWS in the Victoria and New South Wales border.
+The third row selects the lowest temperature in August and this corresponds to Thredbo AWS on the Victoria and New South Wales border.
 Another station in the Tasmania island is selected on the map to cross compare with Thredbo AWS.
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/linking} \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/linking-north} \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/linking-lower}
+{\centering \includegraphics[width=1\linewidth,height=0.23\textheight]{../figures/linking} \includegraphics[width=1\linewidth,height=0.23\textheight]{../figures/linking-north} \includegraphics[width=1\linewidth,height=0.23\textheight]{../figures/linking-lower}
 
@@
 
-Figure \ref{fig:interactive-popup} shows Figure \ref{fig:interactive-linking} made made with leaflet and popups \citep{leafpop}.
-
-\begin{CodeChunk}
-\begin{figure}
-
-{\centering \includegraphics[width=0.45\linewidth,height=0.25\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/popup-mid}
-
-}
-
-\caption[Same as Figure 11 with thetemperature variation shown as a popup in the leaflet map]{Same as Figure 11 with thetemperature variation shown as a popup in the leaflet map.}\label{fig:interactive-popup}
+Figure \ref{fig:interactive-popup} shows Figure \ref{fig:interactive-linking} made with leaflet and popups \citep{leafpop}.
+
+\begin{CodeChunk}
+\begin{figure}
+
+{\centering \includegraphics[width=0.45\linewidth,height=0.25\textheight]{../figures/popup-mid}
+
+}
+
+\caption[Same as Figure 11 with the temperature variation shown as a popup in the leaflet map]{Same as Figure 11 with the temperature variation shown as a popup in the leaflet map.}\label{fig:interactive-popup}
 \end{figure}
@@
 
-This paper describes an \proglang{R} package \pkg{cubble} for manipulating and visualising spatio-temporal data.
-A new data structure, \code{cubble} that builds from the \code{rowwise_df} and \code{grouped_df} class in the tidyverse ecosystem, is proposed to connect the time invariant and varying variables in the spatio-temporal data.
-This design frees the data analysts from spending time on organising variables of different observational units.
-The data structure is also flexible to the techniques and packages analysts use to analyse the data, for example, in the matching example in section 4.3, users are free to use algorithms from another package to cluster stations.
-
-Further development and maintenance of the package involves responding to changes in the tidyverse packages that \pkg{cubble} imports, in particular, \pkg{tibble}, \pkg{tidyr}, and \pkg{dplyr}.
-In the spatial aspect, the simple feature representation wraps spatial coordinates in a list column, for some complex spatial operations, while sometimes, analysts may want to unpack them into longitude and latitude columns to work from.
-Another extension to \pkg{cubble} is to build a smooth transition between the coordinate columns and simple feature geometry column.
+This paper presents an \proglang{R} package \pkg{cubble} for organising, manipulating and visualising spatio-temporal data.
+The package introduces a new data structure for spatio-temporal data, \code{cubble}, that connects the time invariant and varying variables and that allows the user to work with a nested and long form of the data.
+TThe goal of this work is to add capabilities into the spatio-temporal practitioners toolbox that facilitates their work within the tidy data framework.
+The data structure and the capabilities introduced in this package can be used and combined with existing spatio-temporal R packages such as \pkg{sf}, data wrangling packages such as \pkg{dplyr}, and visualization packages such as \pkg{ggplot2}, \pkg{plotly} and \pkg{leaflet}.
+
+The data structure and functions are demonstrated with extensive examples.
+These include creating and coercing wild-caught data with potential mismatch on sites, handling hierarchical data, matching time series spatially and temporally, as well as reproducing ERA5 results from NetCDF data.
+Visualization of the \code{cubble} objects and derivatives is presented via interactive graphic pipelines using \pkg{plotly} and \pkg{leaflet}.
+
+Future directions of the package involves handling sites with moving coordinates.
+This would involve constructing a list-column for location coordinates and a form these locations can be pivoted into, like the long form for temporal variables.
+In the multivariate aspect, \pkg{cubble} can also be extended with interface to more high dimensional visualisation methods, i.e., the tour method, to understand variable importance or comparing location similarities.
 
@@
 
-This work is funded by the Commonwealth Scientific and Industrial Research Organisation (CSIRO) Data61 Scholarship and started while Nicolas Langrené was affiliated with CSIRO's Data61.
+This work is funded by a Commonwealth Scientific and Industrial Research Organisation (CSIRO) Data61 Scholarship and started while Nicolas Langrené was affiliated with CSIRO's Data61.
 The article is created using \pkg{knitr} \citep{knitr} and \pkg{rmarkdown} \citep{rmarkdown} in R.
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.005}
+{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{../figures/diagram-keynotes/diagram-keynotes.005}
 

### 2022-05-25 2-the-cubble-pkg.Rmd
start of section 3

@@
 
-Here we introduce the core functionalities of cubble.
-The main functions, \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and \code{unfold()} are described in Section \@ref(create) - \@ref(unfold).
-Section \@ref(tidyverse) explains the need to create a new class.
-Section \@ref(compact) will then discusses how cubble relates to existing packages for spatial and temporal data analysis.
+The main functions in the cubble package are \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and \code{unfold()}.
+The following sections explain their role, why the new cubble structure is needed and how the package relates to existing packages for spatial and temporal data analysis.
 

### 2022-05-25 jss/cubble-jss.tex
start of section 3

@@
 
-Here we introduce the core functionalities of cubble.
-The main functions, \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and \code{unfold()} are described in Section \ref{create} - \ref{unfold}.
-Section \ref{tidyverse} explains the need to create a new class.
-Section \ref{compact} will then discusses how cubble relates to existing packages for spatial and temporal data analysis.
+The main functions in the cubble package are \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and \code{unfold()}.
+The following sections explain their role, why the new cubble structure is needed and how the package relates to existing packages for spatial and temporal data analysis.
 

### 2022-05-31 jss/cubble-jss.Rmd
also knit the jss version
match jss abstract with the arxiv version

@@
 abstract: >
-  Spatio-temporal data refer to measurements taken across space and time.
-In practice, spatio-temporal data can be decomposed into a spatial and temporal component: at one time, we would select a spatial location and inspect the temporal trend; at other time, we might select one or multiple time value(s) and explore the spatial distribution.
-Ideally, we could make multiple maps and multiple time series to explore these together, however, doing all of these actions is complicated when data arrive fragmented in multiple objects.
-To make it easy to do all these tasks, ideally spatial and temporal variables are in a single data object that we can slice and dice in different ways to conduct different visualisations.
-This work suggests a new data structure, \code{cubble}, that uses a nested form and a long form to organise spatial and temporal variables in a single data object.
-The new data structure ensures that data in the two forms are synchronised while been flexible to explore the two aspect of spatio-temporal data.
-It also provides capability for handling data with hierarchical structure, matching data from multiple sources, constructing interactive graphics, and performing spatio-temporal transformation.
-The paper will demonstrate \pkg{cubble} with examples using Australian climate weather stations, river level data, and climate reanalysis (ERA5) data.
+    Multivariate spatio-temporal data refers to multiple measurements taken across space and time.
+For many analyses, spatial and time components can be separately studied: for example, to explore the temporal trend of one variable for a single spatial location, or to model the spatial distribution of one variable at a given time.
+However for some studies, it is important to analyse different aspects of the spatio-temporal data simultaneouly, like for instance, temporal trends of multiple variables across locations.
+In order to facilitate the study of different portions or combinations of spatio-temporal data, we introduce a new data structure, \code{cubble}, with a suite of functions enabling easy slicing and dicing on the different components spatio-temporal components.
+The proposed \code{cubble} structure ensures that all the components of the data are easy to access and manipulate while providing flexibility for data analysis.
+In addition, \code{cubble} facilitates visual and numerical explorations of the data while easing data wrangling and modelling.
+The \code{cubble} structure and the functions provided in the \code{cubble} R package equip users with the capability to handle hierarchical spatial and temporal structures.
+The \code{cubble} structure and the tools implemented in the package are illustrated with different examples of Australian climate data.
 keywords:
@@
   \usepackage{xcolor}
+  \newcommand{\sherry}[1]{\textcolor{magenta}{#1}}
 output: 

### 2022-05-31 jss/cubble-jss.tex
match jss abstract with the arxiv version

@@
 \Abstract{
-Spatio-temporal data refer to measurements taken across space and time.
-In practice, spatio-temporal data can be decomposed into a spatial and temporal component: at one time, we would select a spatial location and inspect the temporal trend; at other time, we might select one or multiple time value(s) and explore the spatial distribution.
-Ideally, we could make multiple maps and multiple time series to explore these together, however, doing all of these actions is complicated when data arrive fragmented in multiple objects.
-To make it easy to do all these tasks, ideally spatial and temporal variables are in a single data object that we can slice and dice in different ways to conduct different visualisations.
-This work suggests a new data structure, \code{cubble}, that uses a nested form and a long form to organise spatial and temporal variables in a single data object.
-The new data structure ensures that data in the two forms are synchronised while been flexible to explore the two aspect of spatio-temporal data.
-It also provides capability for handling data with hierarchical structure, matching data from multiple sources, constructing interactive graphics, and performing spatio-temporal transformation.
-The paper will demonstrate \pkg{cubble} with examples using Australian climate weather stations, river level data, and climate reanalysis (ERA5) data.
+Multivariate spatio-temporal data refers to multiple measurements taken across space and time.
+For many analyses, spatial and time components can be separately studied: for example, to explore the temporal trend of one variable for a single spatial location, or to model the spatial distribution of one variable at a given time.
+However for some studies, it is important to analyse different aspects of the spatio-temporal data simultaneouly, like for instance, temporal trends of multiple variables across locations.
+In order to facilitate the study of different portions or combinations of spatio-temporal data, we introduce a new data structure, \code{cubble}, with a suite of functions enabling easy slicing and dicing on the different components spatio-temporal components.
+The proposed \code{cubble} structure ensures that all the components of the data are easy to access and manipulate while providing flexibility for data analysis.
+In addition, \code{cubble} facilitates visual and numerical explorations of the data while easing data wrangling and modelling.
+The \code{cubble} structure and the functions provided in the \code{cubble} R package equip users with the capability to handle hierarchical spatial and temporal structures.
+The \code{cubble} structure and the tools implemented in the package are illustrated with different examples of Australian climate data.
 }

### 2022-05-31 2-the-cubble-pkg.Rmd
fix code out of text

@@
 
-Some readers may question why a new data structure is needed rather than using \code{dplyr::nest_by()} on the combined data to directly create a list-column.
+Some readers may question why a new data structure is needed rather than directly creating a list-column  on the combined data using \code{dplyr::nest_by()}.
 The reason is that cubble is specifically designed to utilize the spatio-temporal structure when arranging variables from different observational units into a single object.
@@
 # add station-based features in the nested form.
-dt |> mutate(fabletools::features(ts, tmax, list(tmax_mean = mean, tmax_var = var)))
+dt |> 
+  mutate(
+    fabletools::features(
+      ts, tmax, list(tmax_mean = mean, tmax_var = var)
+      )
+  )
 ```

### 2022-05-31 4-1-basic.Rmd
fix code out of text

@@
 ```{r echo = TRUE, warning = TRUE, message= TRUE}
-cb <- as_cubble(list(spatial = lga, temporal = covid),
-                key = lga, index = date, coords = c(cent_long, cent_lat))
+cb <- as_cubble(
+  list(spatial = lga, temporal = covid),
+  key = lga, index = date, coords = c(cent_long, cent_lat)
+  )
 ```
@@
 ```{r echo = TRUE}
-pair <- as_cubble(list(spatial = lga, temporal = covid),
+pair <- as_cubble(
+  list(spatial = lga, temporal = covid),
   key = lga, index = date, coords = c(cent_long, cent_lat),
-                  output = "unmatch")
+  output = "unmatch"
+  )
 

### 2022-05-31 jss/cubble-jss.tex
fix code out of text

@@
 
-The main functions in the cubble package are \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and \code{unfold()}.
+The main functions in the cubble package are \code{as_cubble(), face_spatial(), face_temporal(), and, unfold()}.
+
+The main functions in the cubble package are \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and, \code{unfold()}.
 The following sections explain their role, why the new cubble structure is needed and how the package relates to existing packages for spatial and temporal data analysis.
@@
   <chr>    <dbl> <dbl> <dbl> <chr>  <dbl> <date>     <dbl> <dbl> <dbl>
-1 ASN00009021 -31.9  116.  15.4 perth airpo~  94610 2020-01-01     0  31.9  15.3
-2 ASN00009021 -31.9  116.  15.4 perth airpo~  94610 2020-01-02     0  24.9  16.4
-3 ASN00009021 -31.9  116.  15.4 perth airpo~  94610 2020-01-03     6  23.2  13  
-4 ASN00009021 -31.9  116.  15.4 perth airpo~  94610 2020-01-04     0  28.4  12.4
-5 ASN00009021 -31.9  116.  15.4 perth airpo~  94610 2020-01-05     0  35.3  11.6
+1 ASN0000~ -31.9  116.  15.4 pert~  94610 2020-01-01     0  31.9  15.3
+2 ASN0000~ -31.9  116.  15.4 pert~  94610 2020-01-02     0  24.9  16.4
+3 ASN0000~ -31.9  116.  15.4 pert~  94610 2020-01-03     6  23.2  13  
+4 ASN0000~ -31.9  116.  15.4 pert~  94610 2020-01-04     0  28.4  12.4
+5 ASN0000~ -31.9  116.  15.4 pert~  94610 2020-01-05     0  35.3  11.6
 # ... with 1,825 more rows
@@
   <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>  
-1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tibble [366 x 4]>
-2 ASN00010311 -31.9  117. 179   york            94623 <tibble [366 x 4]>
-3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble [366 x 4]>
-4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble [366 x 4]>
-5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble [366 x 4]>
+1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tibble>
+2 ASN00010311 -31.9  117. 179   york            94623 <tibble>
+3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble>
+4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble>
+5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble>
 \end{CodeOutput}
@@
 3 ASN00010614 -32.9  117. 338   narrogin       94627 <tibble>       90
-4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble>      106
+4 ASN00014015 -12.4  131.  30.4 darwin airpo~  94120 <tibble>      106
 5 ASN00015131 -17.6  134. 220   elliott        94236 <tibble>       63
@@
 # bbox:    [115.97, -32.94, 133.55, -12.42]
-# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]
+# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id
+#   [dbl]
   id          date        prcp  tmax  tmin
@@
 # bbox:    [115.97, -32.94, 133.55, -12.42]
-# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]
+# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id
+#   [dbl]
   id          date        prcp  tmax  tmin  long   lat
@@
 
-Some readers may question why a new data structure is needed rather than using \code{dplyr::nest_by()} on the combined data to directly create a list-column.
+Some readers may question why a new data structure is needed rather than directly creating a list-column on the combined data using \code{dplyr::nest_by()}.
 The reason is that cubble is specifically designed to utilize the spatio-temporal structure when arranging variables from different observational units into a single object.
@@
   <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>  
-1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tbl_ts [366 x 4]>
-2 ASN00010311 -31.9  117. 179   york            94623 <tbl_ts [366 x 4]>
-3 ASN00010614 -32.9  117. 338   narrogin        94627 <tbl_ts [366 x 4]>
-4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tbl_ts [366 x 4]>
-5 ASN00015131 -17.6  134. 220   elliott         94236 <tbl_ts [366 x 4]>
+1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tbl_ts>
+2 ASN00010311 -31.9  117. 179   york            94623 <tbl_ts>
+3 ASN00010614 -32.9  117. 338   narrogin        94627 <tbl_ts>
+4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tbl_ts>
+5 ASN00015131 -17.6  134. 220   elliott         94236 <tbl_ts>
 \end{CodeOutput}
@@
 R> # add station-based features in the nested form.
-R> dt |> mutate(fabletools::features(ts, tmax, list(tmax_mean = mean, tmax_var = var)))
+R> dt |> 
++   mutate(
++     fabletools::features(
++       ts, tmax, list(tmax_mean = mean, tmax_var = var)
++       )
++   )
 \end{CodeInput}
@@
   <chr>     <dbl> <dbl> <dbl> <chr>  <dbl> <list>       <dbl>    <dbl>
-1 ASN00009021 -31.9  116.  15.4 perth airport  94610 <tbl_ts>      25.7    38.6 
-2 ASN00010311 -31.9  117. 179   york           94623 <tbl_ts>      26.2    51.1 
-3 ASN00010614 -32.9  117. 338   narrogin       94627 <tbl_ts>      23.7    45.4 
-4 ASN00014015 -12.4  131.  30.4 darwin airpo~  94120 <tbl_ts>      33.1     3.02
-5 ASN00015131 -17.6  134. 220   elliott        94236 <tbl_ts>      34.6    24.7 
+1 ASN00009~ -31.9  116.  15.4 pert~  94610 <tbl_ts>      25.7    38.6 
+2 ASN00010~ -31.9  117. 179   york   94623 <tbl_ts>      26.2    51.1 
+3 ASN00010~ -32.9  117. 338   narr~  94627 <tbl_ts>      23.7    45.4 
+4 ASN00014~ -12.4  131.  30.4 darw~  94120 <tbl_ts>      33.1     3.02
+5 ASN00015~ -17.6  134. 220   elli~  94236 <tbl_ts>      34.6    24.7 
 \end{CodeOutput}
@@
 Geodetic CRS:  WGS 84
-    lga_code_2018            lga state_code_2016 state_name_2016 areasqkm_2018
-132         20110     Alpine (S)               2        Victoria     4788.1568
-133         20260    Ararat (RC)               2        Victoria     4211.1171
-134         20570   Ballarat (C)               2        Victoria      739.0321
-135         20660    Banyule (C)               2        Victoria       62.5402
-136         20740 Bass Coast (S)               2        Victoria      865.8095
-    cent_long  cent_lat                       geometry
-132  146.9742 -36.85357 MULTIPOLYGON (((146.7258 -3...
-133  142.8432 -37.47271 MULTIPOLYGON (((143.1807 -3...
-134  143.7815 -37.49286 MULTIPOLYGON (((143.6622 -3...
-135  145.0851 -37.73043 MULTIPOLYGON (((145.1357 -3...
-136  145.5581 -38.50730 MULTIPOLYGON (((145.5207 -3...
+    lga_code_2018            lga state_code_2016 state_name_2016
+132         20110     Alpine (S)               2        Victoria
+133         20260    Ararat (RC)               2        Victoria
+134         20570   Ballarat (C)               2        Victoria
+135         20660    Banyule (C)               2        Victoria
+136         20740 Bass Coast (S)               2        Victoria
+    areasqkm_2018 cent_long  cent_lat                       geometry
+132     4788.1568  146.9742 -36.85357 MULTIPOLYGON (((146.7258 -3...
+133     4211.1171  142.8432 -37.47271 MULTIPOLYGON (((143.1807 -3...
+134      739.0321  143.7815 -37.49286 MULTIPOLYGON (((143.6622 -3...
+135       62.5402  145.0851 -37.73043 MULTIPOLYGON (((145.1357 -3...
+136      865.8095  145.5581 -38.50730 MULTIPOLYGON (((145.5207 -3...
 \end{CodeOutput}
@@
 \begin{CodeInput}
-R> cb <- as_cubble(list(spatial = lga, temporal = covid),
-+                 key = lga, index = date, coords = c(cent_long, cent_lat))
-\end{CodeInput}
-\begin{CodeOutput}
-! Some sites in the temporal table don't have corresponding spatial information
-\end{CodeOutput}
-\begin{CodeOutput}
-! Some sites in the spatial table don't have corresponding temporal information
+R> cb <- as_cubble(
++   list(spatial = lga, temporal = covid),
++   key = lga, index = date, coords = c(cent_long, cent_lat)
++   )
+\end{CodeInput}
+\begin{CodeOutput}
+! Some sites in the temporal table don't have spatial information
+\end{CodeOutput}
+\begin{CodeOutput}
+! Some sites in the spatial table don't have temporal information
 \end{CodeOutput}
@@
 \begin{CodeInput}
-R> pair <- as_cubble(list(spatial = lga, temporal = covid),
+R> pair <- as_cubble(
++   list(spatial = lga, temporal = covid),
 +   key = lga, index = date, coords = c(cent_long, cent_lat),
-+                   output = "unmatch")
++   output = "unmatch"
++   )
 R> 
@@
   <chr>       <chr>     <dbl> <dbl> <chr> <dbl> <int> <list>     <int>
-1 405234      SEVEN CREEKS @ D/S~ -36.9  146. river  6.15     5 <tibble>      21
-2 404207      HOLLAND CREEK @ KE~ -36.6  146. river  8.54    10 <tibble>      21
-3 ASN00082042 strathbogie         -36.8  146. clim~  6.15     5 <tibble>      21
-4 ASN00082170 benalla airport     -36.6  146. clim~  8.54    10 <tibble>      21
-5 230200      MARIBYRNONG RIVER ~ -37.7  145. river  6.17     6 <tibble>      19
-6 ASN00086038 essendon airport    -37.7  145. clim~  6.17     6 <tibble>      19
-7 406213      CAMPASPE RIVER @ R~ -37.0  145. river  1.84     1 <tibble>      18
+1 405234      SEVEN CR~ -36.9  146. river  6.15     5 <tibble>      21
+2 404207      HOLLAND ~ -36.6  146. river  8.54    10 <tibble>      21
+3 ASN00082042 strathbo~ -36.8  146. clim~  6.15     5 <tibble>      21
+4 ASN00082170 benalla ~ -36.6  146. clim~  8.54    10 <tibble>      21
+5 230200      MARIBYRN~ -37.7  145. river  6.17     6 <tibble>      19
+6 ASN00086038 essendon~ -37.7  145. clim~  6.17     6 <tibble>      19
+7 406213      CAMPASPE~ -37.0  145. river  1.84     1 <tibble>      18
 8 ASN00088051 redesdale -37.0  145. clim~  1.84     1 <tibble>      18

### 2022-05-31 3-advanced.Rmd
fix reference placeholder

@@
 Different types of transformations can be useful to extract new information from spatio-temporal data.
-Glyph maps [@Wickham2012-yr] transform the time coordinates into space coordinates to plot the time series of different locations on the map.
+Glyph maps, first introduced in Section \ref{unfold}, transform the time coordinates into space coordinates to plot the time series of different locations on the map.
 Calendar plots [@wang2020calendar] reconstruct time into a calendar-based grid to discover weekday and weekend pattern.

### 2022-05-31 jss/cubble-jss.tex
fix reference placeholder

@@
 
-The main functions in the cubble package are \code{as_cubble(), face_spatial(), face_temporal(), and, unfold()}.
-
-The main functions in the cubble package are \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and, \code{unfold()}.
+The main functions in the cubble package are \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and \code{unfold()}.
 The following sections explain their role, why the new cubble structure is needed and how the package relates to existing packages for spatial and temporal data analysis.
@@
 Sometimes, analysts may need to apply some variable transformation that involves both the spatial and temporal variables.
-An example of this is the transformation of temporal variables into the spatial dimension in glyph maps {[}ref{]}.
+An example of this is the transformation of temporal variables into the spatial dimension in glyph maps \citep{Wickham2012-yr}.
 (Glyph maps will be explained in Section \ref{st_transformation}.) This type of operation can be seen as flattening, or \emph{unfolding}, the cube into a 2D data frame.
@@
 Different types of transformations can be useful to extract new information from spatio-temporal data.
-Glyph maps \citep{Wickham2012-yr} transform the time coordinates into space coordinates to plot the time series of different locations on the map.
+Glyph maps, first introduced in Section \ref{unfold}, transform the time coordinates into space coordinates to plot the time series of different locations on the map.
 Calendar plots \citep{wang2020calendar} reconstruct time into a calendar-based grid to discover weekday and weekend pattern.

### 2022-06-01 jss/cubble-jss.Rmd
fix code out of text
modified title

@@
 title:
-  formatted: "\\pkg{cubble}: An R Package for Structuring Spatio-temporal Data"
+  formatted: "\\pkg{cubble}: An R Package for Organizing and Wrangling Multivariate Spatio-temporal Data"
   # If you use tex in the formatted title, alsoge supply version without
-  plain:     "cubble: An R Package for Structuring Spatio-temporal Data"
+  plain:     "cubble: An R Package for Organizing and Wrangling Multivariate Spatio-temporal Data"
   # For running headers, if needed
@@
 options(prompt = "R> ", continue = "+ ",
-        tibble.print_max = 5, tibble.print_min = 5)
+        tibble.print_max = 5, tibble.print_min = 5,
+        width = 70)
 ```

### 2022-06-01 jss/cubble-jss.tex
modified title

--- 2022-05-31 jss/cubble-jss.tex
@@
 }
-\title{\pkg{cubble}: An R Package for Structuring Spatio-temporal Data}
+\title{\pkg{cubble}: An R Package for Organizing and Wrangling Multivariate Spatio-temporal Data}
 
 \Plainauthor{H. Sherry Zhang, Dianne Cook, Ursula Laa, Nicolas Langrené, Patricia Menéndez}
-\Plaintitle{cubble: An R Package for Structuring Spatio-temporal Data}
+\Plaintitle{cubble: An R Package for Organizing and Wrangling Multivariate Spatio-temporal Data}
 
@@
 \begin{CodeInput}
-R> cb <- as_cubble(list(spatial = lga, temporal = covid),
-+                 key = lga, index = date, coords = c(cent_long, cent_lat))
+R> cb <- as_cubble(
++   list(spatial = lga, temporal = covid),
++   key = lga, index = date, coords = c(cent_long, cent_lat)
++   )
 \end{CodeInput}

### 2022-06-02 jss/cubble-jss.Rmd
rearranging line breaks of authors to fit into one page

@@
     address: 21 Chancellors Walk, Clayton VIC 3800 Australia
-    # location: Melbourne, Australia
     email: dicook@monash.edu
@@
   - name: Nicolas Langrené
-    affiliation: 'BNU-HKBU United International College \AND'
+    affiliation: 'BNU-HKBU United International College'
     address: 2000 Jintong Road, Tangjiawan, Zhuhai, Guangdong Province, China
@@
   - name: Patricia Menéndez
-    affiliation: 'Monash University \AND'
+    affiliation: 'Monash University'
     address: 21 Chancellors Walk, Clayton VIC 3800 Australia
@@
 title:
-  formatted: "\\pkg{cubble}: An R Package for Organizing and Wrangling Multivariate Spatio-temporal Data"
+  formatted: "\\pkg{cubble}: An R Package for Structuring Spatio-temporal Data"
   # If you use tex in the formatted title, alsoge supply version without
-  plain:     "cubble: An R Package for Organizing and Wrangling Multivariate Spatio-temporal Data"
+  plain:     "cubble: An R Package for Structuring Spatio-temporal Data"
   # For running headers, if needed

### 2022-06-02 2-0-cube.Rmd
a paragraph on cubble overview

@@
 Finally, variables are stacked on this space-time canvas, with one observation per site and time point.
-This notion is adopted to avoid using hypercubes when describing multivariate spatio-temporal data and is the conceptual framework behind \pkg{cubble}.
+This notion is adopted to avoid using hypercubes when describing multivariate spatio-temporal data and is the conceptual framework behind the \pkg{cubble} objects.
 With this conceptual model, operations on spatio-temporal data can be mapped to operations on the cube and the rest of Figure \ref{fig:cube-def} show examples of slicing on site, time, and variable.

### 2022-06-02 2-the-cubble-pkg.Rmd
fix reference placeholder
a paragraph on cubble overview

@@
 # The cubble package {#cubble}
+
+A cubble is a data object that arranges spatio-temporal information in a spatial and temporal tidy way.
+It uses two forms to achieve this: the nested form presents the spatial variables, including the feature geometry, and nests the temporal information for spatial wrangling.
+The long form expands the temporal information and stashes the spatial variables for temporal wrangling.
+A benefit of cubble is that it interfaces directly with existing spatial and temporal data wrangling packages.
+For example, operations on feature geometry from the \code{sf} package and temporal operations, i.e., temporal ordering, duplication check.
+and gap filling, from the \code{tsibble} package.
 
@@
 Sometimes, analysts may need to apply some variable transformation that involves both the spatial and temporal variables.
-An example of this is the transformation of temporal variables into the spatial dimension in glyph maps [ref].
+An example of this is the transformation of temporal variables into the spatial dimension in glyph maps [@Wickham2012-yr].
 (Glyph maps will be explained in Section \ref{st_transformation}.) This type of operation can be seen as flattening, or *unfolding*, the cube into a 2D data frame.

### 2022-06-02 jss/cubble-jss.tex
a paragraph on cubble overview

--- 2022-05-31 jss/cubble-jss.tex
@@
 H.
-Sherry Zhang\\Monash University \And Dianne Cook\\Monash University \AND Ursula Laa\\University of Natural Resources and Life Sciences \AND Nicolas Langrené\\BNU-HKBU United International College \AND Patricia Menéndez\\Monash University \AND
+Sherry Zhang\\Monash University \And Dianne Cook\\Monash University \AND Ursula Laa\\University of Natural Resources and Life Sciences \AND Nicolas Langrené\\BNU-HKBU United International College \And Patricia Menéndez\\Monash University
 }
@@
 \section{The cubble package}\label{cubble}}
+
+A cubble is a data object that arranges spatio-temporal information in a spatial and temporal tidy way.
+It uses two forms to achieve this: the nested form presents the spatial variables, including the feature geometry, and nests the temporal information for spatial wrangling.
+The long form expands the temporal information and stashes the spatial variables for temporal wrangling.
+A benefit of cubble is that it interfaces directly with existing spatial and temporal data wrangling packages.
+For example, operations on feature geometry from the \code{sf} package and temporal operations, i.e., temporal ordering, duplication check.
+and gap filling, from the \code{tsibble} package.
 

### 2022-06-02 2-the-cubble-pkg.Rmd
another try

@@
 
-A cubble is a data object that arranges spatio-temporal information in a spatial and temporal tidy way.
+A cubble object is an S3 object [@advr] built on top of the \code{tibble} class to organise spatio-temporal data.
+Since spatial and temporal variables are usually recorded in different observational units, the \code{cubble} class uses an attribute, **form**,  to arrange the data into a spatial or temporal tidy way.
+The **form** attributes can take a value of either "nested" or "long".
+The nested cubble is a subclass of \code{rowwise_df}.
+It arranges each spatial site in a row and uses list columns to store feature geometry and temporal information.
+The long cubble, instead, is a subclass of \code{grouped_df}, which expands the temporal information into the long form and stores the spatial information in a unique attribute to the long cubble, **spatial**.
+
+<!-- A cubble is a data object that arranges spatio-temporal information in a spatial and temporal tidy way.
 It uses two forms to achieve this: the nested form presents the spatial variables, including the feature geometry, and nests the temporal information for spatial wrangling.
@@
 and gap filling, from the \code{tsibble} package.
+-->
 
@@
 Attributes are usually associated with dimensions and variables in the NetCDF format data and a [metadata convention for climate and forecast](http://cfconventions.org/) has been designed to standardise the format of the attributes.
-A few packages in R exist for manipulating NetCDF data and these include a high-level R interface: \pkg{ncdf4} [@ncdf4], a low-level interface that calls a C-interface: \pkg{RNetCDF} [@rnetcdf; @michna2013rnetcdf], and a tidyverse implementation: \pkg{tidync} [@tidync].
+A few packages in R exist for manipulating NetCDF data and these include a high-level R interface: \pkg{ncdf4} [@ncdf4], a low-level interface that calls a C-interface: \pkg{RNetCDF} [@rnetcdf], and a tidyverse implementation: \pkg{tidync} [@tidync].
 

### 2022-06-02 jss/cubble-jss.tex
another try

@@
 
-A cubble is a data object that arranges spatio-temporal information in a spatial and temporal tidy way.
-It uses two forms to achieve this: the nested form presents the spatial variables, including the feature geometry, and nests the temporal information for spatial wrangling.
-The long form expands the temporal information and stashes the spatial variables for temporal wrangling.
-A benefit of cubble is that it interfaces directly with existing spatial and temporal data wrangling packages.
-For example, operations on feature geometry from the \code{sf} package and temporal operations, i.e., temporal ordering, duplication check.
-and gap filling, from the \code{tsibble} package.
+A cubble object is an S3 object \citep{advr} built on top of the \code{tibble} class to organise spatio-temporal data.
+Since spatial and temporal variables are usually recorded in different observational units, the \code{cubble} class uses an attribute, \textbf{form}, to arrange the data into a spatial or temporal tidy way.
+The \textbf{form} attributes can take a value of either ``nested'' or ``long''.
+The nested cubble is a subclass of \code{rowwise_df}.
+It arranges each spatial site in a row and uses list columns to store feature geometry and temporal information.
+The long cubble, instead, is a subclass of \code{grouped_df}, which expands the temporal information into the long form and stores the spatial information in a unique attribute to the long cubble, \textbf{spatial}.
 
@@
 Attributes are usually associated with dimensions and variables in the NetCDF format data and a \href{http://cfconventions.org/}{metadata convention for climate and forecast} has been designed to standardise the format of the attributes.
-A few packages in R exist for manipulating NetCDF data and these include a high-level R interface: \pkg{ncdf4} \citep{ncdf4}, a low-level interface that calls a C-interface: \pkg{RNetCDF} \citep{rnetcdf, michna2013rnetcdf}, and a tidyverse implementation: \pkg{tidync} \citep{tidync}.
+A few packages in R exist for manipulating NetCDF data and these include a high-level R interface: \pkg{ncdf4} \citep{ncdf4}, a low-level interface that calls a C-interface: \pkg{RNetCDF} \citep{rnetcdf}, and a tidyverse implementation: \pkg{tidync} \citep{tidync}.
 

### 2022-06-08 2-0-cube.Rmd
tweakinf first few sections

@@
 
-The most common data cube using the three axes (time, latitude, longitude) could be considered as taking snapshots of the space and stacking them across time.
-Here we define the spatio-temporal cube slightly differently by considering the three axes to be, time, site and  variable/s.
-That is illustrated in the leftmost cube in Figure \ref{fig:cube-def}.
+The most common space-time cube uses the three axes, time, latitude, longitude, and can be considered stacking space across time.
+Ours is a multivariate spatio-temporal cube with the three axes defined to be time, site and variables, as illustrated in the leftmost column of Figure \ref{fig:cube-def}.
 The time axis is the same in both versions, while the site axis now captures both latitude and longitude.
@@
 In addition, it will be generally useful to be able to create new variables which is trickier to manage in an array.
-Thus for convenient wrangling, we have opted for a list object as the `cubble` structure.
+Thus for convenient wrangling, we have opted to create a special `cubble` class.
 

### 2022-06-08 2-the-cubble-pkg.Rmd
tweakinf first few sections

@@
 
-A cubble object is an S3 object [@advr] built on top of the \code{tibble} class to organise spatio-temporal data.
-Since spatial and temporal variables are usually recorded in different observational units, the \code{cubble} class uses an attribute, **form**,  to arrange the data into a spatial or temporal tidy way.
-The **form** attributes can take a value of either "nested" or "long".
+A \code{cubble} object is an S3 class [@advr] built on the \code{tibble} class, specifically to organise spatio-temporal data.
+The \code{cubble} class uses an attribute "form",  to arrange the spatial or temporal data components tidily.
+The form attribute can take a value of either "nested" or "long".
 The nested cubble is a subclass of \code{rowwise_df}.
-It arranges each spatial site in a row and uses list columns to store feature geometry and temporal information.
-The long cubble, instead, is a subclass of \code{grouped_df}, which expands the temporal information into the long form and stores the spatial information in a unique attribute to the long cubble, **spatial**.
+It arranges each spatial site in a row, and uses list columns to store the feature geometry and the temporal information.
+The long cubble is a subclass of \code{grouped_df}, which expands the temporal information into the long form and stores the spatial information in a "spatial" attribute.
 
@@
 
-The main functions in the cubble package are \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and \code{unfold()}.
+The main functions in the package are \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and \code{unfold()}.
 The following sections explain their role, why the new cubble structure is needed and how the package relates to existing packages for spatial and temporal data analysis.

### 2022-06-08 jss/cubble-jss.tex
tweakinf first few sections

@@
 
-The most common data cube using the three axes (time, latitude, longitude) could be considered as taking snapshots of the space and stacking them across time.
-Here we define the spatio-temporal cube slightly differently by considering the three axes to be, time, site and variable/s.
-That is illustrated in the leftmost cube in Figure \ref{fig:cube-def}.
+The most common space-time cube uses the three axes, time, latitude, longitude, and can be considered stacking space across time.
+Ours is a multivariate spatio-temporal cube with the three axes defined to be time, site and variables, as illustrated in the leftmost column of Figure \ref{fig:cube-def}.
 The time axis is the same in both versions, while the site axis now captures both latitude and longitude.
@@
 In addition, it will be generally useful to be able to create new variables which is trickier to manage in an array.
-Thus for convenient wrangling, we have opted for a list object as the \texttt{cubble} structure.
+Thus for convenient wrangling, we have opted to create a special \texttt{cubble} class.
 
@@
 
-A cubble object is an S3 object \citep{advr} built on top of the \code{tibble} class to organise spatio-temporal data.
-Since spatial and temporal variables are usually recorded in different observational units, the \code{cubble} class uses an attribute, \textbf{form}, to arrange the data into a spatial or temporal tidy way.
-The \textbf{form} attributes can take a value of either ``nested'' or ``long''.
+A \code{cubble} object is an S3 class \citep{advr} built on the \code{tibble} class, specifically to organise spatio-temporal data.
+The \code{cubble} class uses an attribute ``form'', to arrange the spatial or temporal data components tidily.
+The form attribute can take a value of either ``nested'' or ``long''.
 The nested cubble is a subclass of \code{rowwise_df}.
-It arranges each spatial site in a row and uses list columns to store feature geometry and temporal information.
-The long cubble, instead, is a subclass of \code{grouped_df}, which expands the temporal information into the long form and stores the spatial information in a unique attribute to the long cubble, \textbf{spatial}.
-
-The main functions in the cubble package are \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and \code{unfold()}.
+It arranges each spatial site in a row, and uses list columns to store the feature geometry and the temporal information.
+The long cubble is a subclass of \code{grouped_df}, which expands the temporal information into the long form and stores the spatial information in a ``spatial'' attribute.
+
+The main functions in the package are \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and \code{unfold()}.
 The following sections explain their role, why the new cubble structure is needed and how the package relates to existing packages for spatial and temporal data analysis.
@@
   <chr>       <dbl> <dbl> <dbl> <chr>         <dbl> <date>     <dbl> <dbl> <dbl>
-1 ASN0000~ -31.9  116.  15.4 pert~  94610 2020-01-01     0  31.9  15.3
-2 ASN0000~ -31.9  116.  15.4 pert~  94610 2020-01-02     0  24.9  16.4
-3 ASN0000~ -31.9  116.  15.4 pert~  94610 2020-01-03     6  23.2  13  
-4 ASN0000~ -31.9  116.  15.4 pert~  94610 2020-01-04     0  28.4  12.4
-5 ASN0000~ -31.9  116.  15.4 pert~  94610 2020-01-05     0  35.3  11.6
+1 ASN00009021 -31.9  116.  15.4 perth airpo~  94610 2020-01-01     0  31.9  15.3
+2 ASN00009021 -31.9  116.  15.4 perth airpo~  94610 2020-01-02     0  24.9  16.4
+3 ASN00009021 -31.9  116.  15.4 perth airpo~  94610 2020-01-03     6  23.2  13  
+4 ASN00009021 -31.9  116.  15.4 perth airpo~  94610 2020-01-04     0  28.4  12.4
+5 ASN00009021 -31.9  116.  15.4 perth airpo~  94610 2020-01-05     0  35.3  11.6
 # ... with 1,825 more rows
@@
   <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>            
-1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tibble>
-2 ASN00010311 -31.9  117. 179   york            94623 <tibble>
-3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble>
-4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble>
-5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble>
+1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tibble [366 x 4]>
+2 ASN00010311 -31.9  117. 179   york            94623 <tibble [366 x 4]>
+3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble [366 x 4]>
+4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble [366 x 4]>
+5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble [366 x 4]>
 \end{CodeOutput}
@@
 3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble>       90
-4 ASN00014015 -12.4  131.  30.4 darwin airpo~  94120 <tibble>      106
+4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble>      106
 5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble>       63
@@
 # bbox:    [115.97, -32.94, 133.55, -12.42]
-# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id
-#   [dbl]
+# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]
   id          date        prcp  tmax  tmin
@@
 # bbox:    [115.97, -32.94, 133.55, -12.42]
-# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id
-#   [dbl]
+# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]
   id          date        prcp  tmax  tmin  long   lat
@@
   <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>            
-1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tbl_ts>
-2 ASN00010311 -31.9  117. 179   york            94623 <tbl_ts>
-3 ASN00010614 -32.9  117. 338   narrogin        94627 <tbl_ts>
-4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tbl_ts>
-5 ASN00015131 -17.6  134. 220   elliott         94236 <tbl_ts>
+1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tbl_ts [366 x 4]>
+2 ASN00010311 -31.9  117. 179   york            94623 <tbl_ts [366 x 4]>
+3 ASN00010614 -32.9  117. 338   narrogin        94627 <tbl_ts [366 x 4]>
+4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tbl_ts [366 x 4]>
+5 ASN00015131 -17.6  134. 220   elliott         94236 <tbl_ts [366 x 4]>
 \end{CodeOutput}
@@
 Geodetic CRS:  WGS 84
-    lga_code_2018            lga state_code_2016 state_name_2016
-132         20110     Alpine (S)               2        Victoria
-133         20260    Ararat (RC)               2        Victoria
-134         20570   Ballarat (C)               2        Victoria
-135         20660    Banyule (C)               2        Victoria
-136         20740 Bass Coast (S)               2        Victoria
-    areasqkm_2018 cent_long  cent_lat                       geometry
-132     4788.1568  146.9742 -36.85357 MULTIPOLYGON (((146.7258 -3...
-133     4211.1171  142.8432 -37.47271 MULTIPOLYGON (((143.1807 -3...
-134      739.0321  143.7815 -37.49286 MULTIPOLYGON (((143.6622 -3...
-135       62.5402  145.0851 -37.73043 MULTIPOLYGON (((145.1357 -3...
-136      865.8095  145.5581 -38.50730 MULTIPOLYGON (((145.5207 -3...
+    lga_code_2018            lga state_code_2016 state_name_2016 areasqkm_2018
+132         20110     Alpine (S)               2        Victoria     4788.1568
+133         20260    Ararat (RC)               2        Victoria     4211.1171
+134         20570   Ballarat (C)               2        Victoria      739.0321
+135         20660    Banyule (C)               2        Victoria       62.5402
+136         20740 Bass Coast (S)               2        Victoria      865.8095
+    cent_long  cent_lat                       geometry
+132  146.9742 -36.85357 MULTIPOLYGON (((146.7258 -3...
+133  142.8432 -37.47271 MULTIPOLYGON (((143.1807 -3...
+134  143.7815 -37.49286 MULTIPOLYGON (((143.6622 -3...
+135  145.0851 -37.73043 MULTIPOLYGON (((145.1357 -3...
+136  145.5581 -38.50730 MULTIPOLYGON (((145.5207 -3...
 \end{CodeOutput}
@@
 \begin{CodeOutput}
-! Some sites in the temporal table don't have spatial information
-\end{CodeOutput}
-\begin{CodeOutput}
-! Some sites in the spatial table don't have temporal information
+! Some sites in the temporal table don't have corresponding spatial information
+\end{CodeOutput}
+\begin{CodeOutput}
+! Some sites in the spatial table don't have corresponding temporal information
 \end{CodeOutput}
@@
   <chr>       <chr>               <dbl> <dbl> <chr> <dbl> <int> <list>     <int>
-1 405234      SEVEN CR~ -36.9  146. river  6.15     5 <tibble>      21
-2 404207      HOLLAND ~ -36.6  146. river  8.54    10 <tibble>      21
-3 ASN00082042 strathbo~ -36.8  146. clim~  6.15     5 <tibble>      21
-4 ASN00082170 benalla ~ -36.6  146. clim~  8.54    10 <tibble>      21
-5 230200      MARIBYRN~ -37.7  145. river  6.17     6 <tibble>      19
-6 ASN00086038 essendon~ -37.7  145. clim~  6.17     6 <tibble>      19
-7 406213      CAMPASPE~ -37.0  145. river  1.84     1 <tibble>      18
+1 405234      SEVEN CREEKS @ D/S~ -36.9  146. river  6.15     5 <tibble>      21
+2 404207      HOLLAND CREEK @ KE~ -36.6  146. river  8.54    10 <tibble>      21
+3 ASN00082042 strathbogie         -36.8  146. clim~  6.15     5 <tibble>      21
+4 ASN00082170 benalla airport     -36.6  146. clim~  8.54    10 <tibble>      21
+5 230200      MARIBYRNONG RIVER ~ -37.7  145. river  6.17     6 <tibble>      19
+6 ASN00086038 essendon airport    -37.7  145. clim~  6.17     6 <tibble>      19
+7 406213      CAMPASPE RIVER @ R~ -37.0  145. river  1.84     1 <tibble>      18
 8 ASN00088051 redesdale           -37.0  145. clim~  1.84     1 <tibble>      18

### 2022-06-08 2-the-cubble-pkg.Rmd
up to section 4

@@
 Some readers may question why a new data structure is needed rather than directly creating a list-column  on the combined data using \code{dplyr::nest_by()}.
-The reason is that cubble is specifically designed to utilize the spatio-temporal structure when arranging variables from different observational units into a single object.
+The reason is that cubble is specifically designed to utilize the spatio-temporal structure when arranging observations in a single object.
 Moreover, it enables easy pivoting between purely spatial, purely temporal, or unfolded into a combined form.
@@
 
-This section will demonstrate how the \code{cubble} class interacts with existing packages commonly used in spatial and temporal analysis, specifically, \code{dplyr}, \code{tsibble}, \code{sf} (\code{s2}), and \code{netcdf4}.
+The \code{cubble} package leverages tools available in existing packages used for spatial and temporal analysis, specifically, \code{dplyr}, \code{tsibble}, \code{sf} (\code{s2}), and \code{netcdf4}, as explained here.
 
@@
 
-The \code{dplyr} package has introduced many tools for data wrangling tasks and these operations are useful in the spatio-temporal context.
-\code{cubble} provides methods that support the following \code{dplyr} verbs in both the nested and long form:
-
- > `mutate`, `filter`, `summarise`, `select`, `arrange`, `rename`, `left_join`, and the slice family (`slice_head`, `slice_tail`, `slice_sample`, `slice_min`, `slice_max`)
+The \code{dplyr} package has many tools for wrangling tidy data,  many of which are useful in the spatio-temporal analysis.
+The \code{cubble} package provides \code{methods} that support the use of the following \code{dplyr} operations on both the nested and long forms: `mutate`, `filter`, `summarise`, `select`, `arrange`, `rename`, `left_join`, and the slice family (`slice_head`, `slice_tail`, `slice_sample`, `slice_min`, `slice_max`)
 
@@
 
-\code{tsibble} is a temporal data structure that uses \code{index} and \code{key} to identify the time and different series.
-\code{cubble} can be seen as following the same vein as \code{tsibble} for spatio-temporal data.
-This makes it easy to cast a \code{tsibble} into a \code{cubble} as only the \code{coords} argument needs to be supplied:
+A \code{tsibble} is a \code{tibble} where the \code{index} and \code{key} components are used to store temporal and strata information, that makes working with temporal data cognitively efficient.
+A \code{cubble} can use a \code{tsibble} for storing the temporal information, and effectively utilize the specialist time series operations in the \code{tsibble} package.
+It is easy to cast an existing \code{tsibble} into a \code{cubble} simply by supplying the \code{coords}:
 
 ```{r echo = TRUE}
-# example with a tsibble created from climate_flat
-raw <- climate_flat |> tsibble::as_tsibble(key = id, index = date)
-dt <-  raw |> cubble::as_cubble(coords = c(long, lat))
-dt
+climate_flat_ts <- climate_flat |> 
+  tsibble::as_tsibble(key = id, index = date)
+climate_flat_cb <-  climate_flat_ts |> 
+  cubble::as_cubble(coords = c(long, lat))
+climate_flat_cb
 ```
 
-When a nested cubble is created, each element in the list-column \code{ts} is of \code{tbl_ts} class and operations available to the tsibble class are still valid within the cubble.
-For example, the code below calculates two time series features (mean and variance) of the maximum temperature variable:
+When a nested cubble is created, each element in the list-column \code{ts} is a \code{tsibble} class (labelled \code{tbl_ts})  and operations available to the \code{tsibble} class are still valid on this element.
+For example, the code below calculates two time series features (mean and variance) of maximum temperature, utilizing the tsibble structure in the cubble:
 
@@
 # add station-based features in the nested form.
-dt |> 
+climate_flat_cb |> 
   mutate(
@@
 
-As a spatial data object, \code{sf} creates a feature geometry list-column (\code{sfc}) in the data frame to provide spatial operations on various geometry types such as (\code{POINT}, \code{LINESTRING}, \code{POLYGON}, \code{MULTIPOLYGON}, etc).
-When creating a cubble a \code{sf} object can be supplied as the spatial table.
-Once that is included in a cubble, methods for the \code{sfc} class can be applied in the nested form inside the cubble object.
-An example of this is shown in \@ref(covid), which also handles the case when the site identifiers in the two tables do not match exactly.
-A spatial data object with an \code{s2} vector can also be the input for the spatial table in \pkg{cubble}.
+The \code{sf} is a spatial data object, containing a feature geometry list-column (\code{sfc}) in the form of various geometry types (\code{POINT}, \code{LINESTRING}, \code{POLYGON}, \code{MULTIPOLYGON}).
+The \code{sf} package provides functions that operate efficiently on this spatial information.
+A cubble can store the spatial information as an \code{sf} object.
+Methods for the \code{sfc} class can be applied in the nested form of the cubble object.
+An illustration is in Section \@ref(covid).
+A spatial data object with an \code{s2} vector can also be used to store the spatial information in a \code{cubble}.
 
@@
 
-NetCDF data is another format that is commonly used for storing spatio-temporal data.
+NetCDF data is another format commonly used for storing spatio-temporal data.
 It has two main components: *dimension* for defining the spatio-temporal grid (longitude, latitude, and time) and *variable* that populates the defined grid.
-Attributes are usually associated with dimensions and variables in the NetCDF format data and a [metadata convention for climate and forecast](http://cfconventions.org/) has been designed to standardise the format of the attributes.
+Attributes  can be associated with dimensions or variables.
+Because there can be many different styles of representing this information there is a metadata convention to standardise the format of the attributes [@REF(http://cfconventions.org/)].
 A few packages in R exist for manipulating NetCDF data and these include a high-level R interface: \pkg{ncdf4} [@ncdf4], a low-level interface that calls a C-interface: \pkg{RNetCDF} [@rnetcdf], and a tidyverse implementation: \pkg{tidync} [@tidync].
@@
 
-The memory limit with NetCDF data in cubble depends on the number of longitude grid points $\times$ the number of latitude grid points $\times$ the number of time grid points $\times$ the number of variables.
-Cubble can handle slightly more than 300 $\times$ 300 (longitude $\times$ longitude) grid points for three daily variables in one year.
+NetCDF data can be quite large, and it is sometimes best to subset the data when converting to a cubble.
+We would recommend reducing to about 300 $\times$ 300 grid points for three daily variables in one year.
 A 300 by 300 spatial grid can be a bounding box of [100, -80, 180, 0] at 0.25 degree resolution or a global bounding box [-180, -90, 180, -90] at 1 degree resolution.
-The spatial grid can also be subsetted to trade for longer time periods and more variables through \code{long_range} and \code{lat_range} if the NetCDF file has finer resolution than needed:
+The size of spatial grid can be reduced if longer time periods or more variables are needed, through \code{long_range} and \code{lat_range}:
 

### 2022-06-08 jss/cubble-jss.tex
up to section 4

@@
 
-The rest of the paper is organized as follows: Section \ref{cube} introduces the proposed cube structure as a way to conceptualise multivariate spatio-temporal data.
-Section \ref{cubble} presents the main design and functionality of \pkg{cubble}.
-Section \ref{others} explains how cubble deals with more advanced considerations, including data with hierarchical structure, data matching and how cubble fits with existing static and interactive visualisation tools.
+The rest of the paper is organized as follows: Section @ref(cube) introduces the proposed cube structure as a way to conceptualise multivariate spatio-temporal data.
+Section @ref(cubble) presents the main design and functionality of \pkg{cubble}.
+Section @ref(others) explains how cubble deals with more advanced considerations, including data with hierarchical structure, data matching and how cubble fits with existing static and interactive visualisation tools.
 Moreover we also illustrate how \pkg{cubble} deals with spatio-temporal data transformations.
-Section \ref{examples} uses Australian weather station data and river level data as examples to demonstrate the use of \pkg{cubble}.
+Section @ref(examples) uses Australian weather station data and river level data as examples to demonstrate the use of \pkg{cubble}.
 An example of how \pkg{cubble} handles NetCDF data is also provided.
-Section \ref{conclude} discuss the paper contributions and future directions.
+Section @ref(conclude) discuss the paper contributions and future directions.
 
@@
 \caption[An illustration of the conceptual spatio-temporal cube with different slicing on time, site, and variable]{An illustration of the conceptual spatio-temporal cube with different slicing on time, site, and variable.
-For each axis, the slicing can be on a single value or a set of values.}\label{fig:cube-def}
+For each axis, the slicing can be on a single value or a set of values.}(\#fig:cube-def)
 \end{figure}
@@
 A cubble can be created from various common spatio-temporal data formats, including basic R objects like tibble, tsibble, sf.
-Section \ref{covid} describes converting multiple tables into a cubble object and Section \ref{netcdf} illustrate how to convert a netCDF object.
+Section @ref(covid) describes converting multiple tables into a cubble object and Section @ref(netcdf) illustrate how to convert a netCDF object.
 
@@
 In the first row, \code{face\_temporal} switches a cubble from the nested form into the long form and the focus has switched from the spatial aspect (the side face) to the temporal aspect (the front face).
-In the second row, \code{face\_spatial} switches a cubble back to the nested form from the long form and shifts focus back to the spatial aspect.}\label{fig:face}
+In the second row, \code{face\_spatial} switches a cubble back to the nested form from the long form and shifts focus back to the spatial aspect.}(\#fig:face)
 \end{figure}
@@
 Some readers may question why a new data structure is needed rather than directly creating a list-column on the combined data using \code{dplyr::nest_by()}.
-The reason is that cubble is specifically designed to utilize the spatio-temporal structure when arranging variables from different observational units into a single object.
+The reason is that cubble is specifically designed to utilize the spatio-temporal structure when arranging observations in a single object.
 Moreover, it enables easy pivoting between purely spatial, purely temporal, or unfolded into a combined form.
@@
 
-The \code{dplyr} package has introduced many tools for data wrangling tasks and these operations are useful in the spatio-temporal context.
-\code{cubble} provides methods that support the following \code{dplyr} verbs in both the nested and long form:
-
-\begin{quote}
-\texttt{mutate}, \texttt{filter}, \texttt{summarise}, \texttt{select}, \texttt{arrange}, \texttt{rename}, \texttt{left\_join}, and the slice family (\texttt{slice\_head}, \texttt{slice\_tail}, \texttt{slice\_sample}, \texttt{slice\_min}, \texttt{slice\_max})
-\end{quote}
+The \code{dplyr} package has many tools for wrangling tidy data, many of which are useful in the spatio-temporal analysis.
+The \code{cubble} package provides \code{methods} that support the use of the following \code{dplyr} operations on both the nested and long forms: \texttt{mutate}, \texttt{filter}, \texttt{summarise}, \texttt{select}, \texttt{arrange}, \texttt{rename}, \texttt{left\_join}, and the slice family (\texttt{slice\_head}, \texttt{slice\_tail}, \texttt{slice\_sample}, \texttt{slice\_min}, \texttt{slice\_max})
 
@@
 
-\code{tsibble} is a temporal data structure that uses \code{index} and \code{key} to identify the time and different series.
-\code{cubble} can be seen as following the same vein as \code{tsibble} for spatio-temporal data.
-This makes it easy to cast a \code{tsibble} into a \code{cubble} as only the \code{coords} argument needs to be supplied:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> # example with a tsibble created from climate_flat
-R> raw <- climate_flat |> tsibble::as_tsibble(key = id, index = date)
-R> dt <-  raw |> cubble::as_cubble(coords = c(long, lat))
-R> dt
+A \code{tsibble} is a \code{tibble} where the \code{index} and \code{key} components are used to store temporal and strata information, that makes working with temporal data cognitively efficient.
+A \code{cubble} can use a \code{tsibble} for storing the temporal information, and effectively utilize the specialist time series operations in the \code{tsibble} package.
+It is easy to cast an existing \code{tsibble} into a \code{cubble} simply by supplying the \code{coords}:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> climate_flat_ts <- climate_flat |> 
++   tsibble::as_tsibble(key = id, index = date)
+R> climate_flat_cb <-  climate_flat_ts |> 
++   cubble::as_cubble(coords = c(long, lat))
+R> climate_flat_cb
 \end{CodeInput}
@@
   <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>  
-1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tbl_ts [366 x 4]>
-2 ASN00010311 -31.9  117. 179   york            94623 <tbl_ts [366 x 4]>
-3 ASN00010614 -32.9  117. 338   narrogin        94627 <tbl_ts [366 x 4]>
-4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tbl_ts [366 x 4]>
-5 ASN00015131 -17.6  134. 220   elliott         94236 <tbl_ts [366 x 4]>
-\end{CodeOutput}
-\end{CodeChunk}
-
-When a nested cubble is created, each element in the list-column \code{ts} is of \code{tbl_ts} class and operations available to the tsibble class are still valid within the cubble.
-For example, the code below calculates two time series features (mean and variance) of the maximum temperature variable:
+1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tbl_ts>
+2 ASN00010311 -31.9  117. 179   york            94623 <tbl_ts>
+3 ASN00010614 -32.9  117. 338   narrogin        94627 <tbl_ts>
+4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tbl_ts>
+5 ASN00015131 -17.6  134. 220   elliott         94236 <tbl_ts>
+\end{CodeOutput}
+\end{CodeChunk}
+
+When a nested cubble is created, each element in the list-column \code{ts} is a \code{tsibble} class (labelled \code{tbl_ts}) and operations available to the \code{tsibble} class are still valid on this element.
+For example, the code below calculates two time series features (mean and variance) of maximum temperature, utilizing the tsibble structure in the cubble:
 
@@
 R> # add station-based features in the nested form.
-R> dt |> 
+R> climate_flat_cb |> 
 +   mutate(
@@
 
-As a spatial data object, \code{sf} creates a feature geometry list-column (\code{sfc}) in the data frame to provide spatial operations on various geometry types such as (\code{POINT}, \code{LINESTRING}, \code{POLYGON}, \code{MULTIPOLYGON}, etc).
-When creating a cubble a \code{sf} object can be supplied as the spatial table.
-Once that is included in a cubble, methods for the \code{sfc} class can be applied in the nested form inside the cubble object.
-An example of this is shown in \ref{covid}, which also handles the case when the site identifiers in the two tables do not match exactly.
-A spatial data object with an \code{s2} vector can also be the input for the spatial table in \pkg{cubble}.
+The \code{sf} is a spatial data object, containing a feature geometry list-column (\code{sfc}) in the form of various geometry types (\code{POINT}, \code{LINESTRING}, \code{POLYGON}, \code{MULTIPOLYGON}).
+The \code{sf} package provides functions that operate efficiently on this spatial information.
+A cubble can store the spatial information as an \code{sf} object.
+Methods for the \code{sfc} class can be applied in the nested form of the cubble object.
+An illustration is in Section @ref(covid).
+A spatial data object with an \code{s2} vector can also be used to store the spatial information in a \code{cubble}.
 
@@
 
-NetCDF data is another format that is commonly used for storing spatio-temporal data.
+NetCDF data is another format commonly used for storing spatio-temporal data.
 It has two main components: \emph{dimension} for defining the spatio-temporal grid (longitude, latitude, and time) and \emph{variable} that populates the defined grid.
-Attributes are usually associated with dimensions and variables in the NetCDF format data and a \href{http://cfconventions.org/}{metadata convention for climate and forecast} has been designed to standardise the format of the attributes.
+Attributes can be associated with dimensions or variables.
+Because there can be many different styles of representing this information there is a metadata convention to standardise the format of the attributes \citep[\url{http://cfconventions.org/})]{REF}.
 A few packages in R exist for manipulating NetCDF data and these include a high-level R interface: \pkg{ncdf4} \citep{ncdf4}, a low-level interface that calls a C-interface: \pkg{RNetCDF} \citep{rnetcdf}, and a tidyverse implementation: \pkg{tidync} \citep{tidync}.
@@
 
-The memory limit with NetCDF data in cubble depends on the number of longitude grid points \(\times\) the number of latitude grid points \(\times\) the number of time grid points \(\times\) the number of variables.
-Cubble can handle slightly more than 300 \(\times\) 300 (longitude \(\times\) longitude) grid points for three daily variables in one year.
+NetCDF data can be quite large, and it is sometimes best to subset the data when converting to a cubble.
+We would recommend reducing to about 300 \(\times\) 300 grid points for three daily variables in one year.
 A 300 by 300 spatial grid can be a bounding box of {[}100, -80, 180, 0{]} at 0.25 degree resolution or a global bounding box {[}-180, -90, 180, -90{]} at 1 degree resolution.
-The spatial grid can also be subsetted to trade for longer time periods and more variables through \code{long_range} and \code{lat_range} if the NetCDF file has finer resolution than needed:
+The size of spatial grid can be reduced if longer time periods or more variables are needed, through \code{long_range} and \code{lat_range}:
 
@@
 \caption[An illustration of the original and cluster level cubble in the nested form long form for hierarchical structure data]{An illustration of the original and cluster level cubble in the nested form long form for hierarchical structure data.
-\code{switch\_key()} changes the station level cubble into a cluster level cubble and both can be stretched into the long form.}\label{fig:illu-hier}
+\code{switch\_key()} changes the station level cubble into a cluster level cubble and both can be stretched into the long form.}(\#fig:illu-hier)
 \end{figure}
@@
 Three highest peaks are identified in each series and intervals are constructed on series \code{A}.
-Two peaks in series \code{a} fall into the intervals and hence the two series are considered to have two matches.}\label{fig:illu-matching}
+Two peaks in series \code{a} fall into the intervals and hence the two series are considered to have two matches.}(\#fig:illu-matching)
 \end{figure}
@@
 When a station is selected on the map (left), the corresponding row in the nested cubble will be activated.
-This will link to all the rows with the same id in the long cubble (middle) and update the line plot (right).}\label{fig:illu-interactive}
+This will link to all the rows with the same id in the long cubble (middle) and update the line plot (right).}(\#fig:illu-interactive)
 \end{figure}
@@
 
-A \code{cubble} object can be created from separate spatial and temporal tables in a list, with other arguments (\code{key}, \code{index}, and \code{coords}) introduced in section \ref{create}.
+A \code{cubble} object can be created from separate spatial and temporal tables in a list, with other arguments (\code{key}, \code{index}, and \code{coords}) introduced in section @ref(create).
 \code{as_cubble()} will automatically check the matching of sites in both tables and emit a warning if a location has missing spatial or temporal information.
@@
 \caption[A glyph map of the average maximum temperature by month of Victoria and New South Wales weather stations in Australia]{A glyph map of the average maximum temperature by month of Victoria and New South Wales weather stations in Australia.
-On the top left corner is an insetted plot of station Cobar highlighted in the black box.}\label{fig:basic-manip}
+On the top left corner is an insetted plot of station Cobar highlighted in the black box.}(\#fig:basic-manip)
 \end{figure}
@@
 The group number is printed in the middle of the y minor axis and can be used as a reference line to read the magnitude.
-Subplot B shows the station membership of each cluster.}\label{fig:basic-agg}
+Subplot B shows the station membership of each cluster.}(\#fig:basic-agg)
 \end{figure}
@@
 
-\caption[Location of weather stations and river gauges in Victoria, Australia]{Location of weather stations and river gauges in Victoria, Australia.}\label{fig:matching-map}
+\caption[Location of weather stations and river gauges in Victoria, Australia]{Location of weather stations and river gauges in Victoria, Australia.}(\#fig:matching-map)
 \end{figure}
@@
 From the map, one can see that a few water gauges and weather stations are close to each other and the fluctuation of the water level could be matched up with the precipitation measured by the nearby climate station.
-As introduced in Section \ref{matching}, \code{match_sites()} can be used to match one source of data with another source in a cubble.
+As introduced in Section @ref(matching), \code{match_sites()} can be used to match one source of data with another source in a cubble.
 Here \code{Water_course_level} in \code{river} will be matched to \code{prcp} in \code{climate} in 2020.
@@
 Precipitation and water level have been standardised between 0 and 1 to be displayed on the same scale.
-The water level reflects the increase in precipitation.}\label{fig:matching}
+The water level reflects the increase in precipitation.}(\#fig:matching)
 \end{figure}
@@
 
-\caption[A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020)]{A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020).}\label{fig:netcdf}
+\caption[A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020)]{A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020).}(\#fig:netcdf)
 \end{figure}
@@
 The bottom row first selects the lowest temperature in August in the seasonal display.
-A location in the Tasmania Island is then selected to compare the temperature variation with Thredbo AWS.}\label{fig:interactive-linking}
+A location in the Tasmania Island is then selected to compare the temperature variation with Thredbo AWS.}(\#fig:interactive-linking)
 \end{figure}
@@
 
-\caption[Same as Figure 11 with the temperature variation shown as a popup in the leaflet map]{Same as Figure 11 with the temperature variation shown as a popup in the leaflet map.}\label{fig:interactive-popup}
+\caption[Same as Figure 11 with the temperature variation shown as a popup in the leaflet map]{Same as Figure 11 with the temperature variation shown as a popup in the leaflet map.}(\#fig:interactive-popup)
 \end{figure}
@@
 This will link to all the rows with the same id in the long cubble and the row in the nested cubble with the same id (middle).
-Both plots will be updated with the full line selected and the point highlighted on the map (right).}\label{fig:illu-interactive-2}
+Both plots will be updated with the full line selected and the point highlighted on the map (right).}(\#fig:illu-interactive-2)
 \end{figure}

### 2022-06-08 1-intro.Rmd
proofread

@@
 With spatio-temporal data, one can fix the time to explore the spatial features of the data, fix the spatial location/s to explore temporal aspects, or dynamically explore the space and time simultaneously.
-<!-- Underlying all these tasks, there is a data representation that enables the spatio-temporal data to be sliced and diced differently.  -->
+
 In order to computationally explore the spatial, temporal and spatio-temporal faces of such data, the data needs to be stored and represented under a specific data object that allows the user to query, group and dissect all the data faces.
@@
 Examples of components of spatio-temporal data can be 1) areal data recording the shape of a collection of areas of interest; 2) geostatistical data storing the longitude and latitude coordinates of locations, typically also with other metadata related to the location, and; 3) temporal data of each location across time.
-
-<!-- **I would aligned the examples of spatio-temporal data with those discuss in the st books. See Paula Moraga for example** -->
-
-<!-- For spatio-temporal data analysis, merging all these sources can be challenging and joining these tables can go wrong because of the slightest unmatched of the linking variable from different sources.
--->
 
@@
 Section \@ref(conclude) discuss the paper contributions and future directions.
-
-<!-- there is a strong need for such tools given the ubiquitous spatial and temporal data collected in the society, i.e., climate observations from weather stations, air quality variables from air monitoring stations.
-The demand for such tools can also come from spatial data analysts wishing to incorporate temporal data in their analysis or vice versa.
--->
-
-<!-- This type of data can naturally be considered as a data cube with three axes: spatial identifier, temporal identifier, and variables.
-In this cubic formation, space only occupies one axis, rather than two, to avoid hypercubes for multivariate spatio-temporal data.
-While conceptually the data format is relatively simple, the data objects coming to the analysts can vary in forms in the analysis.
-Figure \ref{fig:illu-input} shows three examples of wild spatio-temporal data: 1) a single table that includes both the spatial and temporal variables; 2) separate tables for spatial and temporal variables; 3) array-based data, i.e.
-NetCDF (Network Common Data Form) data, which is commonly used in earth science, e.g.
-climatology, meteorology, and oceanography, among others, to organise multivariate data in a spatio-temporal grid.
--->
 

### 2022-06-08 2-the-cubble-pkg.Rmd
proofread

@@
 The form attribute can take a value of either "nested" or "long".
-The nested cubble is a subclass of \code{rowwise_df}.
+The nested cubble is a subclass of rowwise tibble (\code{rowwise_df}).
 It arranges each spatial site in a row, and uses list columns to store the feature geometry and the temporal information.
-The long cubble is a subclass of \code{grouped_df}, which expands the temporal information into the long form and stores the spatial information in a "spatial" attribute.
+The long cubble is a subclass of grouped tibble (\code{grouped_df}), which expands the temporal information into the long form and stores the spatial information in a "spatial" attribute.
 
@@
 The main functions in the package are \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and \code{unfold()}.
-The following sections explain their role, why the new cubble structure is needed and how the package relates to existing packages for spatial and temporal data analysis.
+The following sections explain their roles, why the new cubble structure is needed and how the package relates to existing packages for spatial and temporal data analysis.
 
 The dataset `climate_flat` is used to illustrate \pkg{cubble} functionality.
-This is a subset from the ORIGIN and contains the spatial variables, station id, latitude, longitude, elevation, station name, World Meteorology Organisation ID, in additional to daily temporal information,  maximum and minimum temperature values and precipitation records for 2020.
+This is a subset from National Oceanic and Atmospheric Administration (NOAA) Global Historical Climatology Network Daily (GHCND) data.
+It contains spatial variables, station ID, latitude, longitude, elevation, station name, World Meteorology Organisation ID, in additional to daily temporal information,  maximum and minimum temperature values and precipitation records for year 2020.
 The first five rows of `climate_flat` are shown below:
@@
 
-A cubble can be created from various common spatio-temporal data formats, including basic R objects like tibble, tsibble, sf.
-Section \@ref(covid) describes converting multiple tables into a cubble object and Section \@ref(netcdf) illustrate how to convert a netCDF object.
+\sherry{A cubble can be created from various common spatio-temporal data formats, including basic R objects like tibble, tsibble, sf}.
+Section \@ref(covid) describes converting multiple tables into a cubble object and Section \@ref(netcdf) illustrates how to convert a netCDF object.
 
@@
 
-A cubble object is an extension of a \code{rowwise_df} object where each row forms a group.
+A cubble object is a subclass of the \code{rowwise_df} class where each row forms a group.
 All the temporal variables are nested in a list column, hence it is also called the nested cubble.
@@
 
-The first line in printed cubble now shows the cubble in its long form and the third line has been changed to display the name and type of spatial variables: \code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]}.
-Unlike the nested form, the long cubble is built from a \code{grouped_df} object where all the observations from the same sites form a group.
+The first line in the printed cubble now shows it in the long form and the third line has been changed to display the name and type of spatial variables: \code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]}.
+Unlike the nested form, the long cubble is built from a \code{grouped_df} class where all the observations from the same site form a group.
 
@@
 The \code{dplyr} package has many tools for wrangling tidy data,  many of which are useful in the spatio-temporal analysis.
-The \code{cubble} package provides \code{methods} that support the use of the following \code{dplyr} operations on both the nested and long forms: `mutate`, `filter`, `summarise`, `select`, `arrange`, `rename`, `left_join`, and the slice family (`slice_head`, `slice_tail`, `slice_sample`, `slice_min`, `slice_max`)
+The \code{cubble} package provides \code{methods} that support the use of the following \code{dplyr} operations on both the nested and long forms: `mutate`, `filter`, `summarise`, `select`, `arrange`, `rename`, `left_join`, and the slice family (`slice_*`).
 
@@
 
-### netcdf {#netcdf}
+### netCDF {#netcdf}
 
@@
 Attributes  can be associated with dimensions or variables.
-Because there can be many different styles of representing this information there is a metadata convention to standardise the format of the attributes [@REF(http://cfconventions.org/)].
+Because there can be many different styles of representing this information there is a metadata convention [@hassell_data_2017] to standardise the format of the attributes.
 A few packages in R exist for manipulating NetCDF data and these include a high-level R interface: \pkg{ncdf4} [@ncdf4], a low-level interface that calls a C-interface: \pkg{RNetCDF} [@rnetcdf], and a tidyverse implementation: \pkg{tidync} [@tidync].

### 2022-06-08 jss/cubble-jss.tex
proofread

--- 2022-06-02 jss/cubble-jss.tex
@@
 With spatio-temporal data, one can fix the time to explore the spatial features of the data, fix the spatial location/s to explore temporal aspects, or dynamically explore the space and time simultaneously.
+
 In order to computationally explore the spatial, temporal and spatio-temporal faces of such data, the data needs to be stored and represented under a specific data object that allows the user to query, group and dissect all the data faces.
@@
 
-A cubble object is an S3 object \citep{advr} built on top of the \code{tibble} class to organise spatio-temporal data.
-Since spatial and temporal variables are usually recorded in different observational units, the \code{cubble} class uses an attribute, \textbf{form}, to arrange the data into a spatial or temporal tidy way.
-The \textbf{form} attributes can take a value of either ``nested'' or ``long''.
-The nested cubble is a subclass of \code{rowwise_df}.
-It arranges each spatial site in a row and uses list columns to store feature geometry and temporal information.
-The long cubble, instead, is a subclass of \code{grouped_df}, which expands the temporal information into the long form and stores the spatial information in a unique attribute to the long cubble, \textbf{spatial}.
-
-The main functions in the cubble package are \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and \code{unfold()}.
-The following sections explain their role, why the new cubble structure is needed and how the package relates to existing packages for spatial and temporal data analysis.
+A \code{cubble} object is an S3 class \citep{advr} built on the \code{tibble} class, specifically to organise spatio-temporal data.
+The \code{cubble} class uses an attribute ``form'', to arrange the spatial or temporal data components tidily.
+The form attribute can take a value of either ``nested'' or ``long''.
+The nested cubble is a subclass of rowwise tibble (\code{rowwise_df}).
+It arranges each spatial site in a row, and uses list columns to store the feature geometry and the temporal information.
+The long cubble is a subclass of grouped tibble (\code{grouped_df}), which expands the temporal information into the long form and stores the spatial information in a ``spatial'' attribute.
+
+The main functions in the package are \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and \code{unfold()}.
+The following sections explain their roles, why the new cubble structure is needed and how the package relates to existing packages for spatial and temporal data analysis.
 
 The dataset \texttt{climate\_flat} is used to illustrate \pkg{cubble} functionality.
-This is a subset from the ORIGIN and contains the spatial variables, station id, latitude, longitude, elevation, station name, World Meteorology Organisation ID, in additional to daily temporal information, maximum and minimum temperature values and precipitation records for 2020.
+This is a subset from National Oceanic and Atmospheric Administration (NOAA) Global Historical Climatology Network Daily (GHCND) data.
+It contains spatial variables, station ID, latitude, longitude, elevation, station name, World Meteorology Organisation ID, in additional to daily temporal information, maximum and minimum temperature values and precipitation records for year 2020.
 The first five rows of \texttt{climate\_flat} are shown below:
@@
 
-A cubble can be created from various common spatio-temporal data formats, including basic R objects like tibble, tsibble, sf.
-Section \ref{covid} describes converting multiple tables into a cubble object and Section \ref{netcdf} illustrate how to convert a netCDF object.
+\textcolor{magenta}{A cubble can be created from various common spatio-temporal data formats, including basic R objects like tibble, tsibble, sf}.
+Section \ref{covid} describes converting multiple tables into a cubble object and Section \ref{netcdf} illustrates how to convert a netCDF object.
 
@@
 
-The \code{dplyr} package has introduced many tools for data wrangling tasks and these operations are useful in the spatio-temporal context.
-\code{cubble} provides methods that support the following \code{dplyr} verbs in both the nested and long form:
-
-\begin{quote}
-\texttt{mutate}, \texttt{filter}, \texttt{summarise}, \texttt{select}, \texttt{arrange}, \texttt{rename}, \texttt{left\_join}, and the slice family (\texttt{slice\_head}, \texttt{slice\_tail}, \texttt{slice\_sample}, \texttt{slice\_min}, \texttt{slice\_max})
-\end{quote}
+The \code{dplyr} package has many tools for wrangling tidy data, many of which are useful in the spatio-temporal analysis.
+The \code{cubble} package provides \code{methods} that support the use of the following \code{dplyr} operations on both the nested and long forms: \texttt{mutate}, \texttt{filter}, \texttt{summarise}, \texttt{select}, \texttt{arrange}, \texttt{rename}, \texttt{left\_join}, and the slice family (\texttt{slice\_*}).
 
@@
 
-When a nested cubble is created, each element in the list-column \code{ts} is of \code{tbl_ts} class and operations available to the tsibble class are still valid within the cubble.
-For example, the code below calculates two time series features (mean and variance) of the maximum temperature variable:
+When a nested cubble is created, each element in the list-column \code{ts} is a \code{tsibble} class (labelled \code{tbl_ts}) and operations available to the \code{tsibble} class are still valid on this element.
+For example, the code below calculates two time series features (mean and variance) of maximum temperature, utilizing the tsibble structure in the cubble:
 
@@
 
-As a spatial data object, \code{sf} creates a feature geometry list-column (\code{sfc}) in the data frame to provide spatial operations on various geometry types such as (\code{POINT}, \code{LINESTRING}, \code{POLYGON}, \code{MULTIPOLYGON}, etc).
-When creating a cubble a \code{sf} object can be supplied as the spatial table.
-Once that is included in a cubble, methods for the \code{sfc} class can be applied in the nested form inside the cubble object.
-An example of this is shown in \ref{covid}, which also handles the case when the site identifiers in the two tables do not match exactly.
-A spatial data object with an \code{s2} vector can also be the input for the spatial table in \pkg{cubble}.
+The \code{sf} is a spatial data object, containing a feature geometry list-column (\code{sfc}) in the form of various geometry types (\code{POINT}, \code{LINESTRING}, \code{POLYGON}, \code{MULTIPOLYGON}).
+The \code{sf} package provides functions that operate efficiently on this spatial information.
+A cubble can store the spatial information as an \code{sf} object.
+Methods for the \code{sfc} class can be applied in the nested form of the cubble object.
+An illustration is in Section \ref{covid}.
+A spatial data object with an \code{s2} vector can also be used to store the spatial information in a \code{cubble}.
 
 \hypertarget{netcdf}{%
-\subsubsection{netcdf}\label{netcdf}}
-
-NetCDF data is another format that is commonly used for storing spatio-temporal data.
+\subsubsection{netCDF}\label{netcdf}}
+
+NetCDF data is another format commonly used for storing spatio-temporal data.
 It has two main components: \emph{dimension} for defining the spatio-temporal grid (longitude, latitude, and time) and \emph{variable} that populates the defined grid.
-Attributes are usually associated with dimensions and variables in the NetCDF format data and a \href{http://cfconventions.org/}{metadata convention for climate and forecast} has been designed to standardise the format of the attributes.
+Attributes can be associated with dimensions or variables.
+Because there can be many different styles of representing this information there is a metadata convention \citep{hassell_data_2017} to standardise the format of the attributes.
 A few packages in R exist for manipulating NetCDF data and these include a high-level R interface: \pkg{ncdf4} \citep{ncdf4}, a low-level interface that calls a C-interface: \pkg{RNetCDF} \citep{rnetcdf}, and a tidyverse implementation: \pkg{tidync} \citep{tidync}.

### 2022-06-13 arxiv/cubble-arxiv.Rmd
Removed extra word in abstract

@@
 However for some studies, it is important to analyse different aspects of the spatio-temporal data simultaneouly, like for instance, temporal trends of multiple variables across locations.
-In order to facilitate the study of different portions or combinations of spatio-temporal data, we introduce a new data structure, \code{cubble}, with a suite of functions enabling easy slicing and dicing on the different components spatio-temporal components.
+In order to facilitate the study of different portions or combinations of spatio-temporal data, we introduce a new data structure, \code{cubble}, with a suite of functions enabling easy slicing and dicing on the different spatio-temporal components.
 The proposed \code{cubble} structure ensures that all the components of the data are easy to access and manipulate while providing flexibility for data analysis.

### 2022-06-13 arxiv/cubble-arxiv.tex
Removed extra word in abstract

--- 2022-04-30 arxiv/cubble-arxiv.tex
@@
 However for some studies, it is important to analyse different aspects of the spatio-temporal data simultaneouly, like for instance, temporal trends of multiple variables across locations.
-In order to facilitate the study of different portions or combinations of spatio-temporal data, we introduce a new data structure, \code{cubble}, with a suite of functions enabling easy slicing and dicing on the different components spatio-temporal components.
-The proposed \code{cubble} structure ensures that all the components of the data are easy to access and manipulate while providing flexibility for data analysis.In addition, \code{cubble} facilitates visual and numerical explorations of the data while easing data wrangling and modelling.
+In order to facilitate the study of different portions or combinations of spatio-temporal data, we introduce a new data structure, \code{cubble}, with a suite of functions enabling easy slicing and dicing on the different spatio-temporal components.
+The proposed \code{cubble} structure ensures that all the components of the data are easy to access and manipulate while providing flexibility for data analysis.
+In addition, \code{cubble} facilitates visual and numerical explorations of the data while easing data wrangling and modelling.
 The \code{cubble} structure and the functions provided in the \code{cubble} R package equip users with the capability to handle hierarchical spatial and temporal structures.
@@
 
-The most common data cube using the three axes (time, latitude, longitude) could be considered as taking snapshots of the space and stacking them across time.
-Here we define the spatio-temporal cube slightly differently by considering the three axes to be, time, site and variable/s.
-That is illustrated in the leftmost cube in Figure \ref{fig:cube-def}.
+The most common space-time cube uses the three axes, time, latitude, longitude, and can be considered stacking space across time.
+Ours is a multivariate spatio-temporal cube with the three axes defined to be time, site and variables, as illustrated in the leftmost column of Figure \ref{fig:cube-def}.
 The time axis is the same in both versions, while the site axis now captures both latitude and longitude.
 Finally, variables are stacked on this space-time canvas, with one observation per site and time point.
-This notion is adopted to avoid using hypercubes when describing multivariate spatio-temporal data and is the conceptual framework behind \pkg{cubble}.
+This notion is adopted to avoid using hypercubes when describing multivariate spatio-temporal data and is the conceptual framework behind the \pkg{cubble} objects.
 With this conceptual model, operations on spatio-temporal data can be mapped to operations on the cube and the rest of Figure \ref{fig:cube-def} show examples of slicing on site, time, and variable.
 
-While a cubical data model is neat to view spatio-temporal data conceptually, a 3D data representation, or an array, may not be convenient for data wrangling.
-There are two reasons for this: 1) While arrays are efficient for the computation on numerical values, spatio-temporal data, in general, contains more than just numerical variables.
-For example, wrangling on character strings and specific datetime classes is common for spatio-temporal data but may not be easily done with array objects.
-2) Creating new variables in an array is more complex than in a data frame since it would need to be defined in a 2D space, rather than a 1D vector, to be added into the array.
+While the data cube model is conceptually convenient for spatio-temporal data, a 3D data array is not sufficiently rich for data wrangling, for several reasons.
+Although arrays can be efficient for the computation on numerical values, spatio-temporal data typically includes various types of variables.
+For example, character strings and specific datetime classes are common.
+In addition, it will be generally useful to be able to create new variables which is trickier to manage in an array.
+Thus for convenient wrangling, we have opted to create a special \texttt{cubble} class.
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/diagram-keynotes/diagram-keynotes.007}
+{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/diagram-keynotes/diagram-keynotes.007}
 
@@
 
-This section will first introduce the core functions in the cubble class: \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and \code{unfold()} in subsection \ref{create} - \ref{unfold}.
-The next subsection addresses the necessity of creating a new class.
-Subsection \ref{compact} will then show the compatibility of cubble with existing packages in spatial and temporal analysis.
-
-Each core \pkg{cubble} function is introduced and accompanied with a short example using the data \texttt{climate\_flat} throughout this section.
-\texttt{climate\_flat} contains data on five weather stations in Australia with spatial information of each station: station id, latitude, longitude, elevation, station name, World Meteorology Organisation ID and also daily temporal information, maximum and minimum temperature values and precipitation records for 2020.
+A \code{cubble} object is an S3 class (\protect\hyperlink{ref-advr}{Wickham 2019}) built on the \code{tibble} class, specifically to organise spatio-temporal data.
+The \code{cubble} class uses an attribute ``form,'' to arrange the spatial or temporal data components tidily.
+The form attribute can take a value of either ``nested'' or ``long.'' The nested cubble is a subclass of rowwise tibble (\code{rowwise_df}).
+It arranges each spatial site in a row, and uses list columns to store the feature geometry and the temporal information.
+The long cubble is a subclass of grouped tibble (\code{grouped_df}), which expands the temporal information into the long form and stores the spatial information in a ``spatial'' attribute.
+
+The main functions in the package are \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and \code{unfold()}.
+The following sections explain their roles, why the new cubble structure is needed and how the package relates to existing packages for spatial and temporal data analysis.
+
+The dataset \texttt{climate\_flat} is used to illustrate \pkg{cubble} functionality.
+This is a subset from National Oceanic and Atmospheric Administration (NOAA) Global Historical Climatology Network Daily (GHCND) data.
+It contains spatial variables, station ID, latitude, longitude, elevation, station name, World Meteorology Organisation ID, in additional to daily temporal information, maximum and minimum temperature values and precipitation records for year 2020.
 The first five rows of \texttt{climate\_flat} are shown below:
@@
 
-Spatio-temporal data can come in various formats and shapes and a cubble can be created from various types of raw data.
-This includes tibble and its variates in multiple tables (an example is provided in section \ref{covid}), and netCDF (detailed in section \ref{netcdf}).
-The function \code{as_cubble()} is used to create a cubble with three additional inputs: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order (longitude, latitude).
+\textcolor{magenta}{A cubble can be created from various common spatio-temporal data formats, including basic R objects like tibble, tsibble, sf}.
+Section \ref{covid} describes converting multiple tables into a cubble object and Section \ref{netcdf} illustrates how to convert a netCDF object.
+
+The function \code{as_cubble()} is used to create a cubble with three arguments: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order (longitude, latitude).
 The arguments \code{key} and \code{index} follow the wording in \pkg{tsibble} to describe the temporal order and multiple series while \code{coords} specifies the spatial location of each site.
-The code below creates a cubble out of \code{climate_flat} data with \code{id} as the key, \code{date} as the index, and \code{c(long, lat)} as the coordinates:
+The code below creates a cubble out of \code{climate_flat} (a single tibble) with \code{id} as the key, \code{date} as the index, and \code{c(long, lat)} as the coordinates:
 
@@
 
-The resulting cubble is built from a \code{rowwise_df} class where each row forms a group.
+A cubble object is a subclass of the \code{rowwise_df} class where each row forms a group.
 All the temporal variables are nested in a list column, hence it is also called the nested cubble.
-The rowwise structure makes it simpler to operate on the list using the \code{mutate()} syntax, which is simpler than the \code{map()} when working with a list column.
-For example, calculating the number of raining days can be done by:
+The rowwise structure makes it simpler to operate on the list using the \code{mutate()} syntax, which is simpler than the \code{purr::map()} when working with a list column.
+For example, calculating the number of rainy days can be done by:
 
@@
 
-The first line in the header now shows the cubble is in its long form and the third line has been changed to display the name and type of spatial variables: \code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]}.
-Unlike the nested form, the long cubble is built from class \code{groupped_df} where all the observations from the same sites form a group.
+The first line in the printed cubble now shows it in the long form and the third line has been changed to display the name and type of spatial variables: \code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]}.
+Unlike the nested form, the long cubble is built from a \code{grouped_df} class where all the observations from the same site form a group.
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/diagram-keynotes/diagram-keynotes.008}
+{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/diagram-keynotes/diagram-keynotes.008}
 
@@
 Sometimes, analysts may need to apply some variable transformation that involves both the spatial and temporal variables.
-An example of this is the transformation of temporal variables into the spatial dimension in glyph maps, which will be elaborated in section \ref{st_transformation}.
-This type of operation can be seen as flattening, or \emph{unfolding}, the cube into a 2D data frame.
+An example of this is the transformation of temporal variables into the spatial dimension in glyph maps (\protect\hyperlink{ref-Wickham2012-yr}{Wickham et al.
+2012}).
+(Glyph maps will be explained in Section \ref{st_transformation}.) This type of operation can be seen as flattening, or \emph{unfolding}, the cube into a 2D data frame.
 Here the function \code{unfold()} moves the spatial variables \code{long} and \code{lat} into the long cubble:
@@
 
-\subsection[Why not just use]{Why not just use \code{dplyr::nest\_by()}?}
-
-Some readers may question why a new data structure is needed rather than using \code{dplyr::nest_by()} on the combined data to directly create a list-column.
-The reason is that cubble is specifically designed to utilize the spatio-temporal structure when arranging variables from different observational units into a single object.
+\hypertarget{tidyverse}{%
+\subsection{Why not just use the existing tidyverse functions}\label{tidyverse}}
+
+Some readers may question why a new data structure is needed rather than directly creating a list-column on the combined data using \code{dplyr::nest_by()}.
+The reason is that cubble is specifically designed to utilize the spatio-temporal structure when arranging observations in a single object.
 Moreover, it enables easy pivoting between purely spatial, purely temporal, or unfolded into a combined form.
@@
 
-\code{tsibble} is a temporal data structure that uses \code{index} and \code{key} to identify the time and different series.
-\code{cubble} can be seen as following the same vein as \code{tsibble} for spatio-temporal data.
-This makes it easy to cast a \code{tsibble} into a \code{cubble} as only the \code{coords} argument needs to be supplied:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# example with a tsibble created from climate\_flat}
-\NormalTok{raw }\OtherTok{\textless{}{-}}\NormalTok{ climate\_flat }\SpecialCharTok{|}\ErrorTok{\textgreater{}}\NormalTok{ tsibble}\SpecialCharTok{::}\FunctionTok{as\_tsibble}\NormalTok{(}\AttributeTok{key =}\NormalTok{ id, }\AttributeTok{index =}\NormalTok{ date)}
-\NormalTok{dt }\OtherTok{\textless{}{-}}\NormalTok{  raw }\SpecialCharTok{|}\ErrorTok{\textgreater{}}\NormalTok{ cubble}\SpecialCharTok{::}\FunctionTok{as\_cubble}\NormalTok{(}\AttributeTok{coords =} \FunctionTok{c}\NormalTok{(long, lat))}
-\NormalTok{dt}
+A \code{tsibble} is a \code{tibble} where the \code{index} and \code{key} components are used to store temporal and strata information, that makes working with temporal data cognitively efficient.
+A \code{cubble} can use a \code{tsibble} for storing the temporal information, and effectively utilize the specialist time series operations in the \code{tsibble} package.
+It is easy to cast an existing \code{tsibble} into a \code{cubble} simply by supplying the \code{coords}:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{climate\_flat\_ts }\OtherTok{\textless{}{-}}\NormalTok{ climate\_flat }\SpecialCharTok{|}\ErrorTok{\textgreater{}} 
+\NormalTok{  tsibble}\SpecialCharTok{::}\FunctionTok{as\_tsibble}\NormalTok{(}\AttributeTok{key =}\NormalTok{ id, }\AttributeTok{index =}\NormalTok{ date)}
+\NormalTok{climate\_flat\_cb }\OtherTok{\textless{}{-}}\NormalTok{  climate\_flat\_ts }\SpecialCharTok{|}\ErrorTok{\textgreater{}} 
+\NormalTok{  cubble}\SpecialCharTok{::}\FunctionTok{as\_cubble}\NormalTok{(}\AttributeTok{coords =} \FunctionTok{c}\NormalTok{(long, lat))}
+\NormalTok{climate\_flat\_cb}
 \end{Highlighting}
@@
 \CommentTok{\# add station{-}based features in the nested form.}
-\NormalTok{dt }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{mutate}\NormalTok{(fabletools}\SpecialCharTok{::}\FunctionTok{features}\NormalTok{(ts, tmax, }\FunctionTok{list}\NormalTok{(}\AttributeTok{tmax\_mean =}\NormalTok{ mean, }\AttributeTok{tmax\_var =}\NormalTok{ var)))}
+\NormalTok{climate\_flat\_cb }\SpecialCharTok{|}\ErrorTok{\textgreater{}} 
+  \FunctionTok{mutate}\NormalTok{(}
+\NormalTok{    fabletools}\SpecialCharTok{::}\FunctionTok{features}\NormalTok{(}
+\NormalTok{      ts, tmax, }\FunctionTok{list}\NormalTok{(}\AttributeTok{tmax\_mean =}\NormalTok{ mean, }\AttributeTok{tmax\_var =}\NormalTok{ var)}
+\NormalTok{      )}
+\NormalTok{  )}
 \end{Highlighting}
@@
 
-As a spatial data object, \code{sf} creates a feature geometry list-column (\code{sfc}) in the data frame to provide spatial operations on various geometry types such as (\code{POINT}, \code{LINESTRING}, \code{POLYGON}, \code{MULTIPOLYGON}, etc).
-When creating a cubble a \code{sf} object can be supplied as the spatial table.
-Once that is included in a cubble, methods for the \code{sfc} class can be applied in the nested form inside the cubble object.
-An example of this is shown in \ref{covid}, which also handles the case when the site identifiers in the two tables do not match exactly.
-A spatial data object with an \code{s2} vector can also be the input for the spatial table in \pkg{cubble}.
+The \code{sf} is a spatial data object, containing a feature geometry list-column (\code{sfc}) in the form of various geometry types (\code{POINT}, \code{LINESTRING}, \code{POLYGON}, \code{MULTIPOLYGON}).
+The \code{sf} package provides functions that operate efficiently on this spatial information.
+A cubble can store the spatial information as an \code{sf} object.
+Methods for the \code{sfc} class can be applied in the nested form of the cubble object.
+An illustration is in Section \ref{covid}.
+A spatial data object with an \code{s2} vector can also be used to store the spatial information in a \code{cubble}.
 
 \hypertarget{netcdf}{%
-\subsubsection{netcdf}\label{netcdf}}
-
-NetCDF data is another format that is commonly used for storing spatio-temporal data.
+\subsubsection{netCDF}\label{netcdf}}
+
+NetCDF data is another format commonly used for storing spatio-temporal data.
 It has two main components: \emph{dimension} for defining the spatio-temporal grid (longitude, latitude, and time) and \emph{variable} that populates the defined grid.
-Attributes are usually associated with dimensions and variables in the NetCDF format data and a \href{http://cfconventions.org/}{metadata convention for climate and forecast} has been designed to standardise the format of the attributes.
-A few packages in R exist for manipulating NetCDF data and these include a high-level R interface: \pkg{ncdf4} (\protect\hyperlink{ref-ncdf4}{Pierce 2019}), a low-level interface that calls a C-interface: \pkg{RNetCDF} (\protect\hyperlink{ref-rnetcdf}{Michna and Woods 2021}, \protect\hyperlink{ref-michna2013rnetcdf}{2013}), and a tidyverse implementation: \pkg{tidync} (\protect\hyperlink{ref-tidync}{Sumner 2020}).
+Attributes can be associated with dimensions or variables.
+Because there can be many different styles of representing this information there is a metadata convention (\protect\hyperlink{ref-hassell_data_2017}{Hassell et al.
+2017}) to standardise the format of the attributes.
+A few packages in R exist for manipulating NetCDF data and these include a high-level R interface: \pkg{ncdf4} (\protect\hyperlink{ref-ncdf4}{Pierce 2019}), a low-level interface that calls a C-interface: \pkg{RNetCDF} (\protect\hyperlink{ref-rnetcdf}{Michna and Woods 2021}), and a tidyverse implementation: \pkg{tidync} (\protect\hyperlink{ref-tidync}{Sumner 2020}).
 
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/diagram-keynotes/diagram-keynotes.003}
+{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/diagram-keynotes/diagram-keynotes.003}
 
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/illu-matching} \caption{An illustration of temporal matching in cubble.
+\includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/illu-matching} \caption{An illustration of temporal matching in cubble.
 Three highest peaks are identified in each series and intervals are constructed on series \code{A}.
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/diagram-keynotes/diagram-keynotes.004}
+{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/diagram-keynotes/diagram-keynotes.004}
 
@@
 Different types of transformations can be useful to extract new information from spatio-temporal data.
-Glyph maps (\protect\hyperlink{ref-Wickham2012-yr}{Wickham et al.
-2012}) transform the time coordinates into space coordinates to plot the time series of different locations on the map.
+Glyph maps, first introduced in Section \ref{unfold}, transform the time coordinates into space coordinates to plot the time series of different locations on the map.
 Calendar plots (\protect\hyperlink{ref-wang2020calendar}{Wang, Cook, and Hyndman 2020}) reconstruct time into a calendar-based grid to discover weekday and weekend pattern.
@@
 \begin{Highlighting}[]
-\NormalTok{cb }\OtherTok{\textless{}{-}} \FunctionTok{as\_cubble}\NormalTok{(}\FunctionTok{list}\NormalTok{(}\AttributeTok{spatial =}\NormalTok{ lga, }\AttributeTok{temporal =}\NormalTok{ covid),}
-                \AttributeTok{key =}\NormalTok{ lga, }\AttributeTok{index =}\NormalTok{ date, }\AttributeTok{coords =} \FunctionTok{c}\NormalTok{(cent\_long, cent\_lat))}
+\NormalTok{cb }\OtherTok{\textless{}{-}} \FunctionTok{as\_cubble}\NormalTok{(}
+  \FunctionTok{list}\NormalTok{(}\AttributeTok{spatial =}\NormalTok{ lga, }\AttributeTok{temporal =}\NormalTok{ covid),}
+  \AttributeTok{key =}\NormalTok{ lga, }\AttributeTok{index =}\NormalTok{ date, }\AttributeTok{coords =} \FunctionTok{c}\NormalTok{(cent\_long, cent\_lat)}
+\NormalTok{  )}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{pair }\OtherTok{\textless{}{-}} \FunctionTok{as\_cubble}\NormalTok{(}\FunctionTok{list}\NormalTok{(}\AttributeTok{spatial =}\NormalTok{ lga, }\AttributeTok{temporal =}\NormalTok{ covid),}
+\NormalTok{pair }\OtherTok{\textless{}{-}} \FunctionTok{as\_cubble}\NormalTok{(}
+  \FunctionTok{list}\NormalTok{(}\AttributeTok{spatial =}\NormalTok{ lga, }\AttributeTok{temporal =}\NormalTok{ covid),}
   \AttributeTok{key =}\NormalTok{ lga, }\AttributeTok{index =}\NormalTok{ date, }\AttributeTok{coords =} \FunctionTok{c}\NormalTok{(cent\_long, cent\_lat),}
-                  \AttributeTok{output =} \StringTok{"unmatch"}\NormalTok{)}
+  \AttributeTok{output =} \StringTok{"unmatch"}
+\NormalTok{  )}
 
@@
 \begin{figure}
-\includegraphics[width=1\linewidth,height=0.7\textheight]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/basic-manip} \caption{A glyph map of the average maximum temperature by month of Victoria and New South Wales weather stations in Australia.
+\includegraphics[width=1\linewidth,height=0.7\textheight]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/basic-manip} \caption{A glyph map of the average maximum temperature by month of Victoria and New South Wales weather stations in Australia.
 On the top left corner is an insetted plot of station Cobar highlighted in the black box.}\label{fig:basic-manip}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/basic-agg} \caption{Profile of aggregated precipitation from 639 weather stations in Australia.
+\includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/basic-agg} \caption{Profile of aggregated precipitation from 639 weather stations in Australia.
 Subplot A shows the glyph map of the weekly averaged precipitation of each cluster.
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/matching-map} \caption{Location of weather stations and river gauges in Victoria, Australia.}\label{fig:matching-map}
+\includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/matching-map} \caption{Location of weather stations and river gauges in Victoria, Australia.}\label{fig:matching-map}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/matching} \caption{Matched weather stations and river gauges on the map (A) and across time (B).
+\includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/matching} \caption{Matched weather stations and river gauges on the map (A) and across time (B).
 Precipitation and water level have been standardised between 0 and 1 to be displayed on the same scale.
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/netcdf} \caption{A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020).}\label{fig:netcdf}
+\includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/netcdf} \caption{A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020).}\label{fig:netcdf}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/linking} \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/linking-north} \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/linking-lower} \caption{Exploring temperature variation using linking of a map and seasonal display.
+\includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/linking} \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/linking-north} \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/linking-lower} \caption{Exploring temperature variation using linking of a map and seasonal display.
 Each row is a screen dump of the process.
@@
 
-{\centering \includegraphics[width=0.45\linewidth,height=0.25\textheight]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/popup-mid}
+{\centering \includegraphics[width=0.45\linewidth,height=0.25\textheight]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/popup-mid}
 
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/diagram-keynotes/diagram-keynotes.005}
+{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/diagram-keynotes/diagram-keynotes.005}
 
@@
 
+\leavevmode\hypertarget{ref-hassell_data_2017}{}%
+Hassell, David, Jonathan Gregory, Jon Blower, Bryan N.
+Lawrence, and Karl E.
+Taylor.
+2017.
+{``A Data Model of the {Climate} and {Forecast} Metadata Conventions ({CF}-1.6) with a Software Implementation (Cf-Python V2.1).''} \emph{Geoscientific Model Development} 10 (12): 4619--46.
+\url{https://doi.org/10.5194/gmd-10-4619-2017}.
+
 \leavevmode\hypertarget{ref-hersbach2020era5}{}%
@@
 
-\leavevmode\hypertarget{ref-michna2013rnetcdf}{}%
-Michna, Pavel, and Milton Woods. 2013. {``{RNetCDF}: A Package for Reading and Writing NetCDF Datasets.''} \emph{The R Journal} 5 (2): 29--36.
-
 \leavevmode\hypertarget{ref-rnetcdf}{}%
----------. 2021. \emph{{RNetCDF}: Interface to 'NetCDF' Datasets}. \url{https://CRAN.R-project.org/package=RNetCDF}.
+Michna, Pavel, and Milton Woods. 2021. \emph{{RNetCDF}: Interface to 'NetCDF' Datasets}. \url{https://CRAN.R-project.org/package=RNetCDF}.
 
@@
 \leavevmode\hypertarget{ref-tidync}{}%
-Sumner, Michael. 2020. \emph{{tidync}: A Tidy Approach to 'NetCDF' Data Exploration and Extraction}. \url{https://CRAN.R-project.org/package=tidync}.
+Sumner, Michael. 2020. \emph{{tidync}: A Tidy Approach to NetCDF Data Exploration and Extraction}. \url{https://CRAN.R-project.org/package=tidync}.
 
@@
 \leavevmode\hypertarget{ref-ggplot2}{}%
----------. 2016. \emph{Ggplot2: Elegant Graphics for Data Analysis}. Springer-Verlag New York. \url{https://ggplot2.tidyverse.org}.
+---------. 2016. \emph{{ggplot2}: Elegant Graphics for Data Analysis}. Springer-Verlag New York. \url{https://ggplot2.tidyverse.org}.
+
+\leavevmode\hypertarget{ref-advr}{}%
+---------. 2019. \emph{Advanced r}. CRC press. \url{https://adv-r.hadley.nz/rcpp.html}.
 
@@
 \leavevmode\hypertarget{ref-dplyr}{}%
-Wickham, Hadley, Romain François, Lionel Henry, and Kirill Müller. 2022. \emph{Dplyr: A Grammar of Data Manipulation}. \url{https://CRAN.R-project.org/package=dplyr}.
+Wickham, Hadley, Romain François, Lionel Henry, and Kirill Müller. 2022. \emph{{dplyr}: A Grammar of Data Manipulation}. \url{https://CRAN.R-project.org/package=dplyr}.
 
 \leavevmode\hypertarget{ref-Wickham2012-yr}{}%
-Wickham, Hadley, Heike Hofmann, Charlotte Wickham, and Dianne Cook. 2012. {``Glyph-Maps for Visually Exploring Temporal Patterns in Climate Data and Models.''} \emph{Environmetrics} 23 (5): 382--93.
+Wickham, Hadley, Heike Hofmann, Charlotte Wickham, and Dianne Cook.
+2012.
+{``Glyph-Maps for Visually Exploring Temporal Patterns in Climate Data and Models.''} \emph{Environmetrics} 23 (5): 382--93.
+\url{https://doi.org/10.1002/env.2152}.
 

### 2022-06-13 4-1-basic.Rmd
Udpate section 5.2

@@
 The Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world.
-The dataset \code{weatherdata::historical_tmax} extracts the maximum temperature for `r nrow(weatherdata::historical_tmax)` Australian stations from the GHCN starting from year `r range(weatherdata::historical_tmax$first_year)[2]`.
-\code{weatherdata::historical_tmax} is already in a cubble, with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates.
-This example compares the maximum temperature in two periods: 1971-1975 and 2016-2020 for stations in Victoria and New South Wales.
+The **cubble** dataset \code{weatherdata::historical_tmax} extracts the maximum temperature for `r nrow(weatherdata::historical_tmax)` Australian stations from the GHCN starting from year `r range(weatherdata::historical_tmax$first_year)[2]` and provides information also on the latitude, longitude and elevation of each of the  stations.
+This **cubble** data  is already cast into a \code{cubble} data object, with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates.
+The goal of this example is to compare the monthly average maximum temperature  between two periods, 1971-1975 and 2016-2020, for stations in Victoria and New South Wales.
 
-Stations in the two states can be subsetted by the station number: the Australian GHCN station numbers start with "ASN00" and are followed by the [Bureau of Meteorology (BOM) station number](http://www.bom.gov.au/climate/cdo/about/site-num.shtml).
-The second and third digits of the BOM station number (7th and 8th in the GHCN number) define the state of the station; the range 46-75 corresponds to New South Wales stations and the range 76-90 corresponds to Victorian stations.
-Filtering Victoria and New South Wales stations is a spatial operation and hence uses the nested form:
+First of all, the stations across the two states need to be subsetted from the original \code{weatherdata::historical_tmax} data set.
+For that, the country and the station identifiers, stored within the 11 digits of the \code{id} variable entries, need to be retrieved.
+The country code is located in the first 5 digits (Australia is represented by "ASN00") while the next 6 digits encode the station information following the [Australian Bureau of Meteorology (BOM)](http://www.bom.gov.au/climate/cdo/about/site-num.shtml) coding protocols.
+According to those, New South Wales stations correspond to entries in the range 46-75 and the Victorian stations are stored under 76-90.
+Filtering Victoria and New South Wales stations is a spatial operation and hence uses the \code{cubble} nested form:
 
@@
 
-Filtering for the time periods 1971-1975 and 2016-2020 is a time operation and the nested cubble needs to be switched to the long cubble form by \code{stretch()}:
+and extracting the time periods 1971-1975 and 2016-2020 is a time operation and therefore the nested cubble needs to be switched to the long cubble form by \code{stretch()}:
 
@@
 
-A monthly average is used for both periods to smooth the maximum temperature, which is another time operation:
+A monthly maximum average temperature is then calculated for both periods where \code{cubble} easily allows to compute this time operation across all the stations:
 
@@
 
-A few stations do not have records during the time period 1971-1975 and further investigation shows that while the first and last year of each series is recorded, the missing years in this period are not reported.
-These stations are filtered out by examining whether the summarised time series has 24 months.
-The long cubble needs to be switched to the nested form for this spatial operation using \code{face_spatial()}:
+Observing the results of the operation above show that a few stations do not have records during the time period 1971-1975.
+Further investigation shows that while the first and last year of each series for those stations are recorded, the years in between are not.
+Hence,  these stations are  located and filtered out by examining whether the summarised time series has 24 months.
+Here, the long cubble needs to be switched to the nested form to carry out this spatial operation using \code{face_spatial()}:
 
@@
 
-Lastly, to create a glyph map, both the major (\code{longitude}, \code{latitude}) and minor (\code{month}, \code{tmax}) coordinates need to be in the same table.
-Spatial variables can be moved to the long form with \code{unfold()}:
+Lastly, to create a glyph map where the series of each station are overlayed on a map of Australia, both the major (\code{longitude}, \code{latitude}) and minor (\code{month}, \code{tmax}) coordinates defining the bounding box for the plot need to be on the same table.
+For that, spatial variables can be moved to the long form with \code{unfold()}:
 
@@
 
-\code{tmax} can then be supplied to \code{geom_glyph()} for the glyph map in Figure \ref{fig:basic-manip} with a station inset on the top left corner:
+the variable storing the maximum monthly average temperatures \code{tmax} can then be supplied to \code{geom_glyph()} for the glyph map in Figure \ref{fig:basic-manip} with a station inset on the top left corner:
 

### 2022-06-13 arxiv/cubble-arxiv.tex
Udpate section 5.2

@@
 The Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world.
-The dataset \code{weatherdata::historical_tmax} extracts the maximum temperature for 236 Australian stations from the GHCN starting from year 1969.
-\code{weatherdata::historical_tmax} is already in a cubble, with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates.
-This example compares the maximum temperature in two periods: 1971-1975 and 2016-2020 for stations in Victoria and New South Wales.
-
-Stations in the two states can be subsetted by the station number: the Australian GHCN station numbers start with ``ASN00'' and are followed by the \href{http://www.bom.gov.au/climate/cdo/about/site-num.shtml}{Bureau of Meteorology (BOM) station number}.
-The second and third digits of the BOM station number (7th and 8th in the GHCN number) define the state of the station; the range 46-75 corresponds to New South Wales stations and the range 76-90 corresponds to Victorian stations.
-Filtering Victoria and New South Wales stations is a spatial operation and hence uses the nested form:
+The \textbf{cubble} dataset \code{weatherdata::historical_tmax} extracts the maximum temperature for 236 Australian stations from the GHCN starting from year 1969 and provides information also on the latitude, longitude and elevation of each of the stations.
+This \textbf{cubble} data is already cast into a \code{cubble} data object, with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates.
+The goal of this example is to compare the monthly average maximum temperature between two periods, 1971-1975 and 2016-2020, for stations in Victoria and New South Wales.
+
+First of all, the stations across the two states need to be subsetted from the original \code{weatherdata::historical_tmax} data set.
+For that, the country and the station identifiers, stored within the 11 digits of the \code{id} variable entries, need to be retrieved.
+The country code is located in the first 5 digits (Australia is represented by ``ASN00'') while the next 6 digits encode the station information following the \href{http://www.bom.gov.au/climate/cdo/about/site-num.shtml}{Australian Bureau of Meteorology (BOM)} coding protocols.
+According to those, New South Wales stations correspond to entries in the range 46-75 and the Victorian stations are stored under 76-90.
+Filtering Victoria and New South Wales stations is a spatial operation and hence uses the \code{cubble} nested form:
 
@@
 
-A few stations do not have records during the time period 1971-1975 and further investigation shows that while the first and last year of each series is recorded, the missing years in this period are not reported.
-These stations are filtered out by examining whether the summarised time series has 24 months.
-The long cubble needs to be switched to the nested form for this spatial operation using \code{face_spatial()}:
+Observing the results of the operation above show that a few stations do not have records during the time period 1971-1975.
+Further investigation shows that while the first and last year of each series for those stations are recorded, the years in between are not.
+Hence, these stations are located and filtered out by examining whether the summarised time series has 24 months.
+Here, the long cubble needs to be switched to the nested form to carry out this spatial operation using \code{face_spatial()}:
 

### 2022-06-13 4-4-netcdf.Rmd
First update of seciton 5.5

@@
 
-The ERA5 dataset [@hersbach2020era5] is the latest reanalysis of global atmosphere, land surface, and ocean waves from 1950 onwards and is available in the NetCDF format from the European Centre for Medium-Range Weather Forecasts (ECMWF).
-The data can be directly downloaded from [Copernicus Climate Data Store (CDS)](https://cds.climate.copernicus.eu/cdsapp#!/dataset/reanalysis-era5-pressure-levels?tab=overview) website or programmatically via the R package \pkg{ecmwfr} [@ecwmfr].
-The \code{era5-pressure} data contains the *specific humidity* and *geopotential* variables on the 10 hPa pressure level on four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04.
-Once downloaded, the data can be read into a cubble as:
+The ERA5 dataset [@hersbach2020era5] provides hourly estimates across the Earth for atmospheric, land and oceanic climate variables.
+The data is available in the NetCDF format from the European Centre for Medium-Range Weather Forecasts (ECMWF) and it can be directly downloaded from [Copernicus Climate Data Store (CDS)](https://cds.climate.copernicus.eu/cdsapp#!/dataset/reanalysis-era5-pressure-levels?tab=overview) website or programmatically via the R package \pkg{ecmwfr} [@ecwmfr].
+In this example we focus on the \code{era5-pressure} data which provides hourly data on pressure levels from 1970 to present.
+Here we specifially look at the information  contained on the variables *specific humidity* measuring the mass of water vapour per kilogram of moist air  and the *geopotential* variable recording the gravitational potential energy of a unit mass, at a particular location, relative to mean sea level.
+Time-wise we restrict our analysis to four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04.
+Once downloaded, the data can be read into a \code{cubble} as:
 

### 2022-06-13 arxiv/cubble-arxiv.tex
First update of seciton 5.5

@@
 The ERA5 dataset (\protect\hyperlink{ref-hersbach2020era5}{Hersbach et al.
-2020}) is the latest reanalysis of global atmosphere, land surface, and ocean waves from 1950 onwards and is available in the NetCDF format from the European Centre for Medium-Range Weather Forecasts (ECMWF).
-The data can be directly downloaded from \href{https://cds.climate.copernicus.eu/cdsapp\#!/dataset/reanalysis-era5-pressure-levels?tab=overview}{Copernicus Climate Data Store (CDS)} website or programmatically via the R package \pkg{ecmwfr} (\protect\hyperlink{ref-ecwmfr}{Hufkens, Stauffer, and Campitelli 2019}).
-The \code{era5-pressure} data contains the \emph{specific humidity} and \emph{geopotential} variables on the 10 hPa pressure level on four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04.
-Once downloaded, the data can be read into a cubble as:
+2020}) provides hourly estimates across the Earth for atmospheric, land and oceanic climate variables.
+The data is available in the NetCDF format from the European Centre for Medium-Range Weather Forecasts (ECMWF) and it can be directly downloaded from \href{https://cds.climate.copernicus.eu/cdsapp\#!/dataset/reanalysis-era5-pressure-levels?tab=overview}{Copernicus Climate Data Store (CDS)} website or programmatically via the R package \pkg{ecmwfr} (\protect\hyperlink{ref-ecwmfr}{Hufkens, Stauffer, and Campitelli 2019}).
+In this example we focus on the \code{era5-pressure} data which provides hourly data on pressure levels from 1970 to present.
+Here we specifially look at the information contained on the variables \emph{specific humidity} measuring the mass of water vapour per kilogram of moist air and the \emph{geopotential} variable recording the gravitational potential energy of a unit mass, at a particular location, relative to mean sea level.
+Time-wise we restrict our analysis to four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04.
+Once downloaded, the data can be read into a \code{cubble} as:
 

### 2022-06-13 4-4-netcdf.Rmd
Add motivation example 5.5

@@
 In this example we focus on the \code{era5-pressure} data which provides hourly data on pressure levels from 1970 to present.
-Here we specifially look at the information  contained on the variables *specific humidity* measuring the mass of water vapour per kilogram of moist air  and the *geopotential* variable recording the gravitational potential energy of a unit mass, at a particular location, relative to mean sea level.
-Time-wise we restrict our analysis to four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04.
+Here we specifically look at the information  contained on the variables *specific humidity* measuring the mass of water vapor per kilogram of moist air  and the *geopotential* variable recording the gravitational potential energy of a unit mass, at a particular location, relative to mean sea level.
+The goal of this example is to reproduce the break-up of the southern polar vortex that happened in late September and early October in 2022 in the stratosphere in the Southern Hemisphere.
+Therefore, we filter data corresponding to the Southern Hemisphere and restrict our analysis to four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04.
 Once downloaded, the data can be read into a \code{cubble} as:

### 2022-06-13 arxiv/cubble-arxiv.tex
Add motivation example 5.5

@@
 In this example we focus on the \code{era5-pressure} data which provides hourly data on pressure levels from 1970 to present.
-Here we specifially look at the information contained on the variables \emph{specific humidity} measuring the mass of water vapour per kilogram of moist air and the \emph{geopotential} variable recording the gravitational potential energy of a unit mass, at a particular location, relative to mean sea level.
-Time-wise we restrict our analysis to four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04.
+Here we specifially look at the information contained on the variables \emph{specific humidity} measuring the mass of water vapour per kilogram of moist air and the \emph{geopotential} variable recording the gravitational potential energy of a unit mass, at a particular location, relative to mean sea level.The goal of this example is to reproduce the break-up of the southern polar vortex that happened in late September and early October in 2022 in the stratosphere.
+Therefore, we filter data corresponding to the Southern Hemisphere and restrict our analysis to four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04.
 Once downloaded, the data can be read into a \code{cubble} as:

### 2022-06-13 jss/cubble-jss.tex
Add motivation example 5.5

@@
 ]{jss}
-
-%% recommended packages
-\usepackage{orcidlink,thumbpdf,lmodern}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{../figures/diagram-keynotes/diagram-keynotes.007}
+{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/diagram-keynotes/diagram-keynotes.007}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{../figures/diagram-keynotes/diagram-keynotes.008}
+{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/diagram-keynotes/diagram-keynotes.008}
 
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{../figures/diagram-keynotes/diagram-keynotes.003}
+{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/diagram-keynotes/diagram-keynotes.003}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{../figures/illu-matching}
+{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/illu-matching}
 
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{../figures/diagram-keynotes/diagram-keynotes.004}
+{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/diagram-keynotes/diagram-keynotes.004}
 
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.7\textheight]{../figures/basic-manip}
+{\centering \includegraphics[width=1\linewidth,height=0.7\textheight]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/basic-manip}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{../figures/basic-agg}
+{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/basic-agg}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{../figures/matching-map}
+{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/matching-map}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{../figures/matching}
+{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/matching}
 
@@
 
-The ERA5 dataset \citep{hersbach2020era5} is the latest reanalysis of global atmosphere, land surface, and ocean waves from 1950 onwards and is available in the NetCDF format from the European Centre for Medium-Range Weather Forecasts (ECMWF).
-The data can be directly downloaded from \href{https://cds.climate.copernicus.eu/cdsapp\#!/dataset/reanalysis-era5-pressure-levels?tab=overview}{Copernicus Climate Data Store (CDS)} website or programmatically via the R package \pkg{ecmwfr} \citep{ecwmfr}.
-The \code{era5-pressure} data contains the \emph{specific humidity} and \emph{geopotential} variables on the 10 hPa pressure level on four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04.
-Once downloaded, the data can be read into a cubble as:
+The ERA5 dataset \citep{hersbach2020era5} provides hourly estimates across the Earth for atmospheric, land and oceanic climate variables.
+The data is available in the NetCDF format from the European Centre for Medium-Range Weather Forecasts (ECMWF) and it can be directly downloaded from \href{https://cds.climate.copernicus.eu/cdsapp\#!/dataset/reanalysis-era5-pressure-levels?tab=overview}{Copernicus Climate Data Store (CDS)} website or programmatically via the R package \pkg{ecmwfr} \citep{ecwmfr}.
+In this example we focus on the \code{era5-pressure} data which provides hourly data on pressure levels from 1970 to present.
+Here we specifically look at the information contained on the variables \emph{specific humidity} measuring the mass of water vapor per kilogram of moist air and the \emph{geopotential} variable recording the gravitational potential energy of a unit mass, at a particular location, relative to mean sea level.
+The goal of this example is to reproduce the break-up of the southern polar vortex that happened in late September and early October in 2022 in the stratosphere in the Southern Hemisphere.
+Therefore, we filter data corresponding to the Southern Hemisphere and restrict our analysis to four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04.
+Once downloaded, the data can be read into a \code{cubble} as:
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{../figures/netcdf}
+{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/netcdf}
 
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.23\textheight]{../figures/linking} \includegraphics[width=1\linewidth,height=0.23\textheight]{../figures/linking-north} \includegraphics[width=1\linewidth,height=0.23\textheight]{../figures/linking-lower}
+{\centering \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/linking} \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/linking-north} \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/linking-lower}
 
@@
 
-{\centering \includegraphics[width=0.45\linewidth,height=0.25\textheight]{../figures/popup-mid}
+{\centering \includegraphics[width=0.45\linewidth,height=0.25\textheight]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/popup-mid}
 
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{../figures/diagram-keynotes/diagram-keynotes.005}
+{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/diagram-keynotes/diagram-keynotes.005}
 

### 2022-06-13 3-advanced.Rmd
section 4.1

@@
 
-Spatial locations can have grouping structures either inherent to the data or obtained by clustering.
-In those cases, rather than analysing variables at the site level, it might be of interest to summarised variables at a cluster or group level, as that can give a crisper picture of local areas.
-In cubble, \code{switch_key()} can be used to create a new grouping level of spatial locations by specifying a clustering variable.
-Figure \ref{fig:illu-hier} illustrates the relationship of cubbles at station and cluster level, in both the long and nested forms.
-By specifying \code{cluster_nested <- station_nested \%>\% switch_key(key = cluster)}, the cubble redefines the cubble key from the \code{id} column in \code{station_nested} to the \code{cluster} column in \code{cluster_nested}.
-All the spatial variables variant to the \code{cluster} column are now nested into a \code{.val} column and cluster level variables can be computed in the same fashion as station level variables in \code{station_nested}.
+Spatial locations can have grouping structures either inherent to the data (e.g.
+state within country) or obtained during the analysis (e.g.
+cluster id).
+In this case, it can be useful to summarise variables at various levels of the hierarchy.
+The function \code{switch_key()} can be used to change the grouping level of spatial locations.
+The diagram in Figure \ref{fig:illu-hier} shows how this function can be used to switch the grouping from station ids to cluster ids.
+The result can also be stretched into long form.
+By specifying
 
-```{r illu-hier, echo = FALSE, fig.align="center",out.height="40%", out.width = "100%", fig.cap = "An illustration of the original and cluster level cubble in the nested form long form for hierarchical structure data.
-\\code{switch\\_key()} changes the station level cubble into a cluster level cubble and both can be stretched into the long form."}
+\code{cluster_nested <- station_nested \%>\% switch_key(key = cluster)},
+
+the cubble redefines the cubble key from the \code{id} column in \code{station_nested} to the \code{cluster} column in \code{cluster_nested}.
+All the spatial variables belonging to the \code{cluster} column are now nested into a \code{.val} column which allows for summarizing based on cluster.
+
+```{r illu-hier, echo = FALSE, fig.align="center",out.height="40%", out.width = "100%", fig.cap = "Hierarchical spatial structure can be handled using \\code{switch\\_key()}, to create summaries based on any level.
+Here the switch is between the station id and a cluster id.
+Once the change is made the data can be stretched into the long form."}
 knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.003.png"))

### 2022-06-13 jss/cubble-jss.tex
section 4.1

@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/diagram-keynotes/diagram-keynotes.007}
+{\centering \includegraphics[width=1\linewidth]{../figures/diagram-keynotes/diagram-keynotes.007}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/diagram-keynotes/diagram-keynotes.008}
+{\centering \includegraphics[width=1\linewidth]{../figures/diagram-keynotes/diagram-keynotes.008}
 
@@
 
-Spatial locations can have grouping structures either inherent to the data or obtained by clustering.
-In those cases, rather than analysing variables at the site level, it might be of interest to summarised variables at a cluster or group level, as that can give a crisper picture of local areas.
-In cubble, \code{switch_key()} can be used to create a new grouping level of spatial locations by specifying a clustering variable.
-Figure \ref{fig:illu-hier} illustrates the relationship of cubbles at station and cluster level, in both the long and nested forms.
-By specifying \code{cluster_nested <- station_nested \%>\% switch_key(key = cluster)}, the cubble redefines the cubble key from the \code{id} column in \code{station_nested} to the \code{cluster} column in \code{cluster_nested}.
-All the spatial variables variant to the \code{cluster} column are now nested into a \code{.val} column and cluster level variables can be computed in the same fashion as station level variables in \code{station_nested}.
-
-\begin{CodeChunk}
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/diagram-keynotes/diagram-keynotes.003}
-
-}
-
-\caption[An illustration of the original and cluster level cubble in the nested form long form for hierarchical structure data]{An illustration of the original and cluster level cubble in the nested form long form for hierarchical structure data.
-\code{switch\_key()} changes the station level cubble into a cluster level cubble and both can be stretched into the long form.}\label{fig:illu-hier}
+Spatial locations can have grouping structures either inherent to the data (e.g.~state within country) or obtained during the analysis (e.g.~cluster id).
+In this case, it can be useful to summarise variables at various levels of the hierarchy.
+The function \code{switch_key()} can be used to change the grouping level of spatial locations.
+The diagram in Figure \ref{fig:illu-hier} shows how this function can be used to switch the grouping from station ids to cluster ids.
+The result can also be stretched into long form.
+By specifying
+
+\code{cluster_nested <- station_nested \%>\% switch_key(key = cluster)},
+
+the cubble redefines the cubble key from the \code{id} column in \code{station_nested} to the \code{cluster} column in \code{cluster_nested}.
+All the spatial variables belonging to the \code{cluster} column are now nested into a \code{.val} column which allows for summarizing based on cluster.
+
+\begin{CodeChunk}
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{../figures/diagram-keynotes/diagram-keynotes.003}
+
+}
+
+\caption[Hierarchical spatial structure can be handled using \code{switch\_key()}, to create summaries based on any level]{Hierarchical spatial structure can be handled using \code{switch\_key()}, to create summaries based on any level.
+Here the switch is between the station id and a cluster id.
+Once the change is made the data can be stretched into the long form.}\label{fig:illu-hier}
 \end{figure}
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/illu-matching}
+{\centering \includegraphics[width=1\linewidth]{../figures/illu-matching}
 
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/diagram-keynotes/diagram-keynotes.004}
+{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{../figures/diagram-keynotes/diagram-keynotes.004}
 
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.7\textheight]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/basic-manip}
+{\centering \includegraphics[width=1\linewidth,height=0.7\textheight]{../figures/basic-manip}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/basic-agg}
+{\centering \includegraphics[width=1\linewidth]{../figures/basic-agg}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/matching-map}
+{\centering \includegraphics[width=1\linewidth]{../figures/matching-map}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/matching}
+{\centering \includegraphics[width=1\linewidth]{../figures/matching}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/netcdf}
+{\centering \includegraphics[width=1\linewidth]{../figures/netcdf}
 
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/linking} \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/linking-north} \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/linking-lower}
+{\centering \includegraphics[width=1\linewidth,height=0.23\textheight]{../figures/linking} \includegraphics[width=1\linewidth,height=0.23\textheight]{../figures/linking-north} \includegraphics[width=1\linewidth,height=0.23\textheight]{../figures/linking-lower}
 
@@
 
-{\centering \includegraphics[width=0.45\linewidth,height=0.25\textheight]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/popup-mid}
+{\centering \includegraphics[width=0.45\linewidth,height=0.25\textheight]{../figures/popup-mid}
 
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/diagram-keynotes/diagram-keynotes.005}
+{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{../figures/diagram-keynotes/diagram-keynotes.005}
 

### 2022-06-13 3-advanced.Rmd
section 4.2

@@
 
-One task that may interest analysts in spatio-temporal data is to find how similar time series, from nearby sites, are.
-This problem can be seen as a matching problem [@stuart2010matching; @mcintosh2018using] that pairs up similar time series in nearby locations or a data fusing exercise that merges data collected from different sources [@cocchi2019data].
-\code{match_sites()} in \pkg{cubble} provides a simple algorithm for this task.
-The algorithm first matches the two data sources spatially by computing the pairwise distance on latitude and longitude.
-After that, pairs that pass the spatial matching are then matched temporally by computing the number of matched peaks within a fixed length moving window.
-Figure \ref{fig:illu-matching} illustrates this temporal matching in more detail.
-Given two series \code{A} and \code{a}, three peaks have been identified in each series.
-An interval, with default length of five, is constructed for each peak in series \code{A}, while the peaks in series \code{a} are tested against whether they fall into any of the intervals.
-In this illustration, there are two matches for these two series.
-Several arguments are available in \code{match_sites()} to fine-tune the matching:
+One task that may interest spatio-temporal analysts is matching data collected at nearby but not exactly the same sites, for example, weather station measured rainfall and river levels.
+This can be considered to be a matching problem [@stuart2010matching; @mcintosh2018using] to pair similar time series from nearby locations, or even a data fusion exercise that merges data collected from different sources [@cocchi2019data].
+The function \code{match_sites()} in \pkg{cubble} provides a simple algorithm for this task.
+The algorithm first matches spatially by computing the pairwise distance on latitude and longitude.
+Then it matches temporally by computing the number of matched peaks within a fixed length moving window.
+Figure \ref{fig:illu-matching} illustrates this temporal matching.
+In the two series, \code{A} and \code{a}, three peaks have been identified in each.
+An interval, of fixed length, is constructed for each peak in series \code{A}, while the peaks in series \code{a} are tested against whether they fall into any of the intervals.
+Here two out of three peaks match.
+Options for \code{match_sites()} are:
 

### 2022-06-13 jss/cubble-jss.tex
section 4.2

@@
 
-One task that may interest analysts in spatio-temporal data is to find how similar time series, from nearby sites, are.
-This problem can be seen as a matching problem \citep{stuart2010matching, mcintosh2018using} that pairs up similar time series in nearby locations or a data fusing exercise that merges data collected from different sources \citep{cocchi2019data}.
-\code{match_sites()} in \pkg{cubble} provides a simple algorithm for this task.
-The algorithm first matches the two data sources spatially by computing the pairwise distance on latitude and longitude.
-After that, pairs that pass the spatial matching are then matched temporally by computing the number of matched peaks within a fixed length moving window.
-Figure \ref{fig:illu-matching} illustrates this temporal matching in more detail.
-Given two series \code{A} and \code{a}, three peaks have been identified in each series.
-An interval, with default length of five, is constructed for each peak in series \code{A}, while the peaks in series \code{a} are tested against whether they fall into any of the intervals.
-In this illustration, there are two matches for these two series.
-Several arguments are available in \code{match_sites()} to fine-tune the matching:
+One task that may interest spatio-temporal analysts is matching data collected at nearby but not exactly the same sites, for example, weather station measured rainfall and river levels.
+This can be considered to be a matching problem \citep{stuart2010matching, mcintosh2018using} to pair similar time series from nearby locations, or even a data fusion exercise that merges data collected from different sources \citep{cocchi2019data}.
+The function \code{match_sites()} in \pkg{cubble} provides a simple algorithm for this task.
+The algorithm first matches spatially by computing the pairwise distance on latitude and longitude.
+Then it matches temporally by computing the number of matched peaks within a fixed length moving window.
+Figure \ref{fig:illu-matching} illustrates this temporal matching.
+In the two series, \code{A} and \code{a}, three peaks have been identified in each.
+An interval, of fixed length, is constructed for each peak in series \code{A}, while the peaks in series \code{a} are tested against whether they fall into any of the intervals.
+Here two out of three peaks match.
+Options for \code{match_sites()} are:
 

### 2022-06-13 3-advanced.Rmd
section 4.3

@@
 
-The linking structure of two forms in cubble fits naturally with the interactive graphic pipeline discussed in the literature [@buja1988elements; @buja1996interactive;  @sutherland2000orca; @xie2014reactive; @cheng2016enabling].
-Diagram \ref{fig:illu-interactive} illustrates how linking works from the map to the time series in cubble.
-The map and time series plot is associated with the nested or long cubble, respectively, and when a user action is captured on the map, the site will be activated in the nested cubble (left).
-The nested cubble will communicate to the long cubble to activate all the observations with the same \code{id} (middle).
-The long cubble will then highlight the activated series in the time series plot (right).
+The cubble workflow fits works well with an interactive graphics pipeline (e.g.
+@buja1988elements, @buja1996interactive,  @sutherland2000orca, @xie2014reactive, @cheng2016enabling) that is available in R with the package crosstalk [@crosstalk].
+Figure \ref{fig:illu-interactive} illustrates how linking can be achieved between a map and multiple time series using a cubble.
+The map (produced from the nested form) and time series (produced from the long form) are both shared crosstalk object.
+When a user makes a selection on the map, the site is highlighted (left).
+This activates a row in the nested cubble object, which is then communicated to the long cubble -- all the observations with the same \code{id} (middle) will be selected.
+The long cubble will then highlight the corresponding series in the time series plot (right).
 
-The linking is also available from the time series plot to the map.
-The selection(s) on the time series is through selecting the point(s) on the time series and once a point is selected, it will be activated in the long cubble.
-All the observations that share the same \code{id} are then activated and this includes other points in the same time series in the long cubble and the corresponding observation of site in the nested cubble.
-These activated observations will then be reflected in the updated plots and Diagram \ref{fig:illu-interactive-2} in the Appendix illustrates this process.
+Linking is also available starting from the time series plot, by selecting points.
+This will be activate rows having the same \code{id} in the long cubble.
+The corresponding rows in the nested cubble are activated, and highlighted the map.
+(An illustration can be found in the appendix.) Note that this type of linking, both from the map or the time series, is what @CS07 would call categorical variable linking, where station id is the categorical variable.
+<!--\ref{fig:illu-interactive-2} in the Appendix illustrates this process.-->
 

### 2022-06-13 jss/cubble-jss.tex
section 4.3

@@
 
-The linking structure of two forms in cubble fits naturally with the interactive graphic pipeline discussed in the literature \citep{buja1988elements, buja1996interactive, sutherland2000orca, xie2014reactive, cheng2016enabling}.
-Diagram \ref{fig:illu-interactive} illustrates how linking works from the map to the time series in cubble.
-The map and time series plot is associated with the nested or long cubble, respectively, and when a user action is captured on the map, the site will be activated in the nested cubble (left).
-The nested cubble will communicate to the long cubble to activate all the observations with the same \code{id} (middle).
-The long cubble will then highlight the activated series in the time series plot (right).
-
-The linking is also available from the time series plot to the map.
-The selection(s) on the time series is through selecting the point(s) on the time series and once a point is selected, it will be activated in the long cubble.
-All the observations that share the same \code{id} are then activated and this includes other points in the same time series in the long cubble and the corresponding observation of site in the nested cubble.
-These activated observations will then be reflected in the updated plots and Diagram \ref{fig:illu-interactive-2} in the Appendix illustrates this process.
+The cubble workflow fits works well with an interactive graphics pipeline (e.g.
+\citet{buja1988elements}, \citet{buja1996interactive}, \citet{sutherland2000orca}, \citet{xie2014reactive}, \citet{cheng2016enabling}) that is available in R with the package crosstalk \citep{crosstalk}.
+Figure \ref{fig:illu-interactive} illustrates how linking can be achieved between a map and multiple time series using a cubble.
+The map (produced from the nested form) and time series (produced from the long form) are both shared crosstalk object.
+When a user makes a selection on the map, the site is highlighted (left).
+This activates a row in the nested cubble object, which is then communicated to the long cubble -- all the observations with the same \code{id} (middle) will be selected.
+The long cubble will then highlight the corresponding series in the time series plot (right).
+
+Linking is also available starting from the time series plot, by selecting points.
+This will be activate rows having the same \code{id} in the long cubble.
+The corresponding rows in the nested cubble are activated, and highlighted the map.
+(An illustration can be found in the appendix.) Note that this type of linking, both from the map or the time series, is what \citet{CS07} would call categorical variable linking, where station id is the categorical variable.
 

### 2022-06-13 2-the-cubble-pkg.Rmd
section 4.4

@@
 An example of this is the transformation of temporal variables into the spatial dimension in glyph maps [@Wickham2012-yr].
-(Glyph maps will be explained in Section \ref{st_transformation}.) This type of operation can be seen as flattening, or *unfolding*, the cube into a 2D data frame.
+(How to make glyph maps will be explained in Section \ref{st_transformation}, and are illustrated in the second example.) This type of operation can be seen as flattening, or *unfolding*, the cube into a 2D data frame.
 Here the function \code{unfold()} moves the spatial variables \code{long} and \code{lat} into the long cubble:

### 2022-06-13 3-advanced.Rmd
section 4.4

@@
 
-Different types of transformations can be useful to extract new information from spatio-temporal data.
-Glyph maps, first introduced in Section \ref{unfold}, transform the time coordinates into space coordinates to plot the time series of different locations on the map.
-Calendar plots [@wang2020calendar] reconstruct time into a calendar-based grid to discover weekday and weekend pattern.
-Projection, or linear combination, of variables summarises multivariate information into lower dimension to further digest.
-This section elaborates on the glyph map.
+Spatio-temporal data lends itself to a range of transformations.
+Glyph maps (Section \ref{unfold}) transform the measured variable and time coordinates into microplots at the spatial locations.
+Calendar plots [@wang2020calendar] deconstruct time to produce plots of variables in a calendar format.
+Summarising multiple variables is commonly done using projections, or linear combinations.
+Here we elaborate on the transformations made to produce a glyph map.
 
-In \proglang{R}, \pkg{GGally} implements glyph maps through the \code{glyphs()} function.
-The function constructs a data frame with calculated position (\code{gx}, \code{gy}, \code{gid}) of each point on the time series using linear algebra (Equation 1 and 2 in @Wickham2012-yr).
+The \proglang{R} package \pkg{GGally} implements glyph maps through the \code{glyphs()} function.
+The function constructs a data frame with calculated position (\code{gx}, \code{gy}, \code{gid}) of each point on the time series using linear algebra (Equations 1 and 2 in @Wickham2012-yr).
 The data can then be piped into \code{ggplot} to create the glyph map as:
@@
 
-A reimplementation of the glyph map as a ggproto, \code{GeomGlyph}, has been made in the \pkg{cubble} package and now the glyph map can be created with \code{geom_glyph()}:
+A new implementation of the glyph map as a ggproto, \code{GeomGlyph}, has been made in the \pkg{cubble} package so that a glyph map can be created with \code{geom_glyph()}:
 
@@
 
+An example using a glyph map is shown in Section \ref{precip}.
+
 Some useful controls over the glyph map are also available in the \code{geom_glyph()} implementation.
-Polar glyph map can be specified as a parameter \code{polar = TRUE} in the \code{geom_glyph()}, along with \code{width} and \code{height} in either absolute or relative value.
-Global and local scale can be controlled by the parameter \code{global_rescale} which default to \code{TRUE} for global scaling.
-Reference box and line can be added with separate \code{geom_glyph_box()} and \code{geom_glyph_line()}.
+Polar coordinate glyph maps are specified using \code{polar = TRUE}, and parameters \code{width} and \code{height} can be specified in either absolute or relative value.
+Global and local scale is specified with \code{global_rescale}, which defaults to \code{TRUE}.
+Reference boxes and lines can be added with separate \code{geom_glyph_box()} and \code{geom_glyph_line()} lines.

### 2022-06-13 jss/cubble-jss.tex
section 4.4

@@
 An example of this is the transformation of temporal variables into the spatial dimension in glyph maps \citep{Wickham2012-yr}.
-(Glyph maps will be explained in Section \ref{st_transformation}.) This type of operation can be seen as flattening, or \emph{unfolding}, the cube into a 2D data frame.
+(How to make glyph maps will be explained in Section \ref{st_transformation}, and are illustrated in the second example.) This type of operation can be seen as flattening, or \emph{unfolding}, the cube into a 2D data frame.
 Here the function \code{unfold()} moves the spatial variables \code{long} and \code{lat} into the long cubble:
@@
 
-Different types of transformations can be useful to extract new information from spatio-temporal data.
-Glyph maps, first introduced in Section \ref{unfold}, transform the time coordinates into space coordinates to plot the time series of different locations on the map.
-Calendar plots \citep{wang2020calendar} reconstruct time into a calendar-based grid to discover weekday and weekend pattern.
-Projection, or linear combination, of variables summarises multivariate information into lower dimension to further digest.
-This section elaborates on the glyph map.
-
-In \proglang{R}, \pkg{GGally} implements glyph maps through the \code{glyphs()} function.
-The function constructs a data frame with calculated position (\code{gx}, \code{gy}, \code{gid}) of each point on the time series using linear algebra (Equation 1 and 2 in \citet{Wickham2012-yr}).
+Spatio-temporal data lends itself to a range of transformations.
+Glyph maps (Section \ref{unfold}) transform the measured variable and time coordinates into microplots at the spatial locations.
+Calendar plots \citep{wang2020calendar} deconstruct time to produce plots of variables in a calendar format.
+Summarising multiple variables is commonly done using projections, or linear combinations.
+Here we elaborate on the transformations made to produce a glyph map.
+
+The \proglang{R} package \pkg{GGally} implements glyph maps through the \code{glyphs()} function.
+The function constructs a data frame with calculated position (\code{gx}, \code{gy}, \code{gid}) of each point on the time series using linear algebra (Equations 1 and 2 in \citet{Wickham2012-yr}).
 The data can then be piped into \code{ggplot} to create the glyph map as:
@@
 
+An example using a glyph map is shown in Section \ref{precip}.
+
 Some useful controls over the glyph map are also available in the \code{geom_glyph()} implementation.
-Polar glyph map can be specified as a parameter \code{polar = TRUE} in the \code{geom_glyph()}, along with \code{width} and \code{height} in either absolute or relative value.
-Global and local scale can be controlled by the parameter \code{global_rescale} which default to \code{TRUE} for global scaling.
-Reference box and line can be added with separate \code{geom_glyph_box()} and \code{geom_glyph_line()}.
+Polar coordinate glyph maps are specified using \code{polar = TRUE}, and parameters \code{width} and \code{height} can be specified in either absolute or relative value.
+Global and local scale is specified with \code{global_rescale}, which defaults to \code{TRUE}.
+Reference boxes and lines can be added with separate \code{geom_glyph_box()} and \code{geom_glyph_line()} lines.
 
@@
 
-\hypertarget{australia-precipitation-pattern-in-2020}{%
-\subsection{Australia precipitation pattern in 2020}\label{australia-precipitation-pattern-in-2020}}
+\hypertarget{precip}{%
+\subsection{Australia precipitation pattern in 2020}\label{precip}}
 

### 2022-06-14 5-conclusion.Rmd
appendix as a single document

@@
 The source code for reproducing this paper can be found at: https://github.com/huizezhang-sherry/paper-cubble.
-
-# Appendix
-
-```{r illu-interactive-2, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "An illustration of the data model under interactive graphics with cubble.
-When a point on the time series is selected, the corresponding row in the long cubble will be activated.
-This will link to all the rows with the same id in the long cubble and the row in the nested cubble with the same id (middle).
-Both plots will be updated with the full line selected and the point highlighted on the map (right)."}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.005.png"))
-```
-
-\newpage

### 2022-06-14 jss/appendix.Rmd
appendix as a single document

---
title: "appendix to cubble: An R Package for Organizing and Wrangling Multivariate Spatio-temporal Data"
author: "Sherry Zhang"
date: '2022-06-14'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Appendix

Figure \ref{fig:illu-interactive-2} illustrates the linking pipeline from a point selection on the time series to the map.

```{r illu-interactive-2, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "An illustration of the data model under interactive graphics with cubble. When a point on the time series is selected, the corresponding row in the long cubble will be activated. This will link to all the rows with the same id in the long cubble and the row in the nested cubble with the same id (middle). Both plots will be updated with the full line selected and the point highlighted on the map (right)."}
knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.005.png"))
```

\newpage

### 2022-06-14 jss/cubble-jss.tex
appendix as a single document

@@
 \begin{CodeOutput}
-! Some sites in the temporal table don't have corresponding spatial information
-\end{CodeOutput}
-\begin{CodeOutput}
-! Some sites in the spatial table don't have corresponding temporal information
+! Some sites in the temporal table don't have spatial information
+\end{CodeOutput}
+\begin{CodeOutput}
+! Some sites in the spatial table don't have temporal information
 \end{CodeOutput}
@@
 The Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world.
-The \textbf{cubble} dataset \code{weatherdata::historical_tmax} extracts the maximum temperature for 236 Australian stations from the GHCN starting from year 1969 and provides information also on the latitude, longitude and elevation of each of the stations.
+The \textbf{cubble} dataset \code{weatherdata::historical_tmax} extracts the maximum temperature for 204 Australian stations from the GHCN starting from year 1969 and provides information also on the latitude, longitude and elevation of each of the stations.
 This \textbf{cubble} data is already cast into a \code{cubble} data object, with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates.
@@
 
-\code{weatherdata::climate_full}, also extracted from the GHCN, records daily precipitation and maximum/minimum temperature for 639 stations in Australia from 2016 to 2020.
+\code{weatherdata::climate_full}, also extracted from the GHCN, records daily precipitation and maximum/minimum temperature for 640 stations in Australia from 2016 to 2020.
 A simple \(k\)-means algorithm based on the distance matrix is used to create 20 clusters.
@@
 
-\hypertarget{appendix}{%
-\section{Appendix}\label{appendix}}
-
-\begin{CodeChunk}
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{../figures/diagram-keynotes/diagram-keynotes.005}
-
-}
-
-\caption[An illustration of the data model under interactive graphics with cubble]{An illustration of the data model under interactive graphics with cubble.
-When a point on the time series is selected, the corresponding row in the long cubble will be activated.
-This will link to all the rows with the same id in the long cubble and the row in the nested cubble with the same id (middle).
-Both plots will be updated with the full line selected and the point highlighted on the map (right).}\label{fig:illu-interactive-2}
-\end{figure}
-\end{CodeChunk}
-
-\newpage
-
 \bibliography{../references.bib}

### 2022-06-14 jss/cubble-jss.tex
fix intext code width

--- 2022-06-13 jss/cubble-jss.tex
@@
 
-First of all, the stations across the two states need to be subsetted from the original \code{weatherdata::historical_tmax} data set.
+First of all, the stations across the two states need to be subsetted from the original data set \code{weatherdata::historical_tmax}.
 For that, the country and the station identifiers, stored within the 11 digits of the \code{id} variable entries, need to be retrieved.

### 2022-06-14 jss/cubble-jss.tex
cleared cache

--- 2022-06-14 jss/cubble-jss.tex
@@
 The Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world.
-The \textbf{cubble} dataset \code{weatherdata::historical_tmax} extracts the maximum temperature for 204 Australian stations from the GHCN starting from year 1969 and provides information also on the latitude, longitude and elevation of each of the stations.
+The \textbf{cubble} dataset \code{weatherdata::historical_tmax} extracts the maximum temperature for 236 Australian stations from the GHCN starting from year 1969 and provides information also on the latitude, longitude and elevation of each of the stations.
 This \textbf{cubble} data is already cast into a \code{cubble} data object, with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates.
@@
 
-\code{weatherdata::climate_full}, also extracted from the GHCN, records daily precipitation and maximum/minimum temperature for 640 stations in Australia from 2016 to 2020.
+\code{weatherdata::climate_full}, also extracted from the GHCN, records daily precipitation and maximum/minimum temperature for 639 stations in Australia from 2016 to 2020.
 A simple \(k\)-means algorithm based on the distance matrix is used to create 20 clusters.

### 2022-06-14 3-advanced.Rmd
glyph map tweaking

@@
 
-One task that may interest spatio-temporal analysts is matching data collected at nearby but not exactly the same sites, for example, weather station measured rainfall and river levels.
+One task that may interest spatio-temporal analysts is combining data collected at nearby but not exactly the same sites, for example, weather station measured rainfall and river levels.
 This can be considered to be a matching problem [@stuart2010matching; @mcintosh2018using] to pair similar time series from nearby locations, or even a data fusion exercise that merges data collected from different sources [@cocchi2019data].
@@
 Figure \ref{fig:illu-interactive} illustrates how linking can be achieved between a map and multiple time series using a cubble.
-The map (produced from the nested form) and time series (produced from the long form) are both shared crosstalk object.
+The map (produced from the nested form) and time series (produced from the long form) are both shared crosstalk objects.
 When a user makes a selection on the map, the site is highlighted (left).

### 2022-06-14 4-1-basic.Rmd
fix intext code width
glyph map tweaking

@@
 
-First of all, the stations across the two states need to be subsetted from the original \code{weatherdata::historical_tmax} data set.
+First of all, the stations across the two states need to be subsetted from the original  data set \code{weatherdata::historical_tmax}.
 For that, the country and the station identifiers, stored within the 11 digits of the \code{id} variable entries, need to be retrieved.
@@
 
-```{r basic-manip, out.width="100%", out.height="70%", fig.cap="A glyph map of the average maximum temperature by month of Victoria and New South Wales weather stations in Australia.
-On the top left corner is an insetted plot of station Cobar highlighted in the black box."}
+```{r basic-manip, out.width="70%", fig.cap="A glyph map of the average maximum temperature by month for parts of Victoria and New South Wales weather stations in Australia for two five year periods (1971-1975, 2016-2020).
+An inset (top left) showing detail is the station Cobar.
+The mothly trend is similar at all locations (low in the winter, high in the summer), and small increased temperatures, particularly in late summer can be seen at most stations."}
 knitr::include_graphics(here::here("figures/basic-manip.png"))

### 2022-06-14 jss/cubble-jss.tex
glyph map tweaking

@@
 
-One task that may interest spatio-temporal analysts is matching data collected at nearby but not exactly the same sites, for example, weather station measured rainfall and river levels.
+One task that may interest spatio-temporal analysts is combining data collected at nearby but not exactly the same sites, for example, weather station measured rainfall and river levels.
 This can be considered to be a matching problem \citep{stuart2010matching, mcintosh2018using} to pair similar time series from nearby locations, or even a data fusion exercise that merges data collected from different sources \citep{cocchi2019data}.
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.7\textheight]{../figures/basic-manip}
-
-}
-
-\caption[A glyph map of the average maximum temperature by month of Victoria and New South Wales weather stations in Australia]{A glyph map of the average maximum temperature by month of Victoria and New South Wales weather stations in Australia.
-On the top left corner is an insetted plot of station Cobar highlighted in the black box.}\label{fig:basic-manip}
+{\centering \includegraphics[width=0.7\linewidth]{../figures/basic-manip}
+
+}
+
+\caption[A glyph map of the average maximum temperature by month of Victoria and New South Wales weather stations in Australia for two five year periods (1971-1975, 2016-2020)]{A glyph map of the average maximum temperature by month of Victoria and New South Wales weather stations in Australia for two five year periods (1971-1975, 2016-2020).
+An inset (top left) showing detail is the station Cobar.
+The mothly trend is similar at all locations (low in the winter, high in the summer), and small increased temperatures, particularly in late summer can be seen at most stations.}\label{fig:basic-manip}
 \end{figure}

### 2022-06-14 scripts/basic-manip.R

@@ -1,4 +1,5 @@
 library(tidyverse)
+library(ggthemes)
 tmax <- weatherdata::historical_tmax |>
   # filter out VIC and NSW stations
   filter(between(stringr::str_sub(id, 7, 8), 46, 90)) |> 
@@ -24,15 +25,17 @@ tmax <- weatherdata::historical_tmax |>
 nsw_vic <- ozmaps::abs_ste |> filter(NAME %in% c("Victoria", "New South Wales"))
 
 p1 <- ggplot() + 
-  geom_sf(data = nsw_vic, fill = "transparent", color = "grey", linetype = "dotted") + 
+  geom_sf(data = nsw_vic, fill = "grey95", color = "white") + 
   geom_glyph(data = tmax, 
              aes(x_major = longitude, x_minor = month, 
                  y_major = latitude, y_minor = tmax,
-                 group = interaction(id, group), color = group),
+                 group = interaction(id, group), 
+                 color = group),
              width = 1, height = 0.5) +
-  scale_color_brewer(palette = "Dark2") + 
+  scale_color_brewer("", palette = "Dark2") + 
   theme_bw() + 
   coord_sf(xlim = c(141, 154), ylim = c(-36, -29)) + 
+  theme_map() +
   theme(legend.position = "bottom") +
   labs(x = "Longitude", y = "Latitude")
 
@@ -45,8 +48,10 @@ tmax |> filter(id == "ASN00048027") |>
              color = group)) +
   geom_line(size = 1.5) +
   scale_color_brewer(palette = "Dark2", guide = "none") +
-  scale_x_continuous(breaks = seq(1, 12, 1)) +
-  labs(x = "Month", y  = "Temperature", title = "ASN00048027: Cobar") +
+  scale_x_continuous(breaks = seq(1, 12, 1), 
+                     labels = c("J", "F", "M", "A", "M", "J", 
+                                "J", "A", "S", "O", "N", "D")) +
+  labs(x = "", y  = "Temp (C)", title = "ASN00048027: Cobar") +
   theme_bw() +
   theme(
     aspect.ratio = 0.3,
@@ -74,4 +79,5 @@ p1 +
   ggimg::geom_point_img(data = single,
                         aes(x = 143.7, y = -30, img = img), size = 6)
 
-ggsave(filename = "figures/basic-manip.png", width = 10, height = 10)
+ggsave(filename = "figures/basic-manip.png", width = 6, 
+       height = 5)

### 2022-06-14 4-1-basic.Rmd
still tinkering with glyph map

@@
 
-```{r basic-manip, out.width="70%", fig.cap="A glyph map of the average maximum temperature by month for parts of Victoria and New South Wales weather stations in Australia for two five year periods (1971-1975, 2016-2020).
-An inset (top left) showing detail is the station Cobar.
+```{r basic-manip, out.width="90%", fig.cap="A glyph map of the average maximum temperature by month for parts of Victoria and New South Wales weather stations in Australia for two five year periods (1971-1975, 2016-2020).
+An inset (top left) showing detail is the station Cobar (highlighted by black rectangle).
 The mothly trend is similar at all locations (low in the winter, high in the summer), and small increased temperatures, particularly in late summer can be seen at most stations."}

### 2022-06-14 jss/cubble-jss.tex
still tinkering with glyph map

@@
 
-{\centering \includegraphics[width=0.7\linewidth]{../figures/basic-manip}
-
-}
-
-\caption[A glyph map of the average maximum temperature by month of Victoria and New South Wales weather stations in Australia for two five year periods (1971-1975, 2016-2020)]{A glyph map of the average maximum temperature by month of Victoria and New South Wales weather stations in Australia for two five year periods (1971-1975, 2016-2020).
-An inset (top left) showing detail is the station Cobar.
+{\centering \includegraphics[width=0.9\linewidth]{../figures/basic-manip}
+
+}
+
+\caption[A glyph map of the average maximum temperature by month for parts of Victoria and New South Wales weather stations in Australia for two five year periods (1971-1975, 2016-2020)]{A glyph map of the average maximum temperature by month for parts of Victoria and New South Wales weather stations in Australia for two five year periods (1971-1975, 2016-2020).
+An inset (top left) showing detail is the station Cobar (highlighted by black rectangle).
 The mothly trend is similar at all locations (low in the winter, high in the summer), and small increased temperatures, particularly in late summer can be seen at most stations.}\label{fig:basic-manip}

### 2022-06-14 4-2-agg.Rmd
section 4.4

@@
 
-## Australia precipitation pattern in 2020
+## Australian precipitation pattern in 2020 {#precip}
 
-In the previous example, there has already been some overlapping of the glyphs for a few stations near (151E, 34S) and (152E, 33S), which is a problem when mapping more stations at the national level.
+In the previous example, some overlapping of the glyphs occurred for a few nearby stations such as the pairs (151E, 34S) and (152E, 33S).
+This is a problem when mapping more stations at the national level.
 Aggregation can be helpful in grouping series into clusters before visualising the clusters with a glyph map.
-This example shows how to organise data at both levels with \code{switch_key()}.
+This new example shows how to organise data at both levels with \code{switch_key()}.
 
 \code{weatherdata::climate_full}, also extracted from the GHCN, records daily precipitation and maximum/minimum temperature for `r nrow(weatherdata::climate_full)` stations in Australia from 2016 to 2020.
-A simple $k$-means algorithm based on the distance matrix is used to create 20 clusters.
+A simple $k$-means algorithm based on the distance matrix between stations is used to create 20 clusters.
 The dataset \code{station_nested} is a nested cubble with a cluster column indicating the group to which each station belongs.
-More advanced clustering algorithms can be used for other applications, as long as there is a mapping from each station to a cluster.
+More advanced clustering algorithms can be used as well, as long as they provide a mapping from each station to a cluster.
 
@@
 
-Long form cubble at both levels can be accessed through stretching the nested form and with access to both station and cluster-level cubbles, various plots can be made to understand the cluster.
+Long form cubble at both levels can be accessed through stretching the nested form.
+With access to both station and cluster-level cubbles, various plots can be made to understand the cluster.
 Figure \ref{fig:basic-agg} shows two example plots that can be made with this data.
@@
 
-```{r basic-agg, out.width="100%", fig.cap="Profile of aggregated precipitation from 639 weather stations in Australia.
+```{r basic-agg, out.width="100%", fig.cap="Profile of aggregated precipitation at 639 weather stations in Australia.
 Subplot A shows the glyph map of the weekly averaged precipitation of each cluster.

### 2022-06-14 jss/cubble-jss.tex
sizing of images

@@
 \hypertarget{precip}{%
-\subsection{Australia precipitation pattern in 2020}\label{precip}}
-
-In the previous example, there has already been some overlapping of the glyphs for a few stations near (151E, 34S) and (152E, 33S), which is a problem when mapping more stations at the national level.
+\subsection{Australian precipitation pattern in 2020}\label{precip}}
+
+In the previous example, some overlapping of the glyphs occurred for a few nearby stations such as the pairs (151E, 34S) and (152E, 33S).
+This is a problem when mapping more stations at the national level.
 Aggregation can be helpful in grouping series into clusters before visualising the clusters with a glyph map.
-This example shows how to organise data at both levels with \code{switch_key()}.
+This new example shows how to organise data at both levels with \code{switch_key()}.
 
 \code{weatherdata::climate_full}, also extracted from the GHCN, records daily precipitation and maximum/minimum temperature for 639 stations in Australia from 2016 to 2020.
-A simple \(k\)-means algorithm based on the distance matrix is used to create 20 clusters.
+A simple \(k\)-means algorithm based on the distance matrix between stations is used to create 20 clusters.
 The dataset \code{station_nested} is a nested cubble with a cluster column indicating the group to which each station belongs.
-More advanced clustering algorithms can be used for other applications, as long as there is a mapping from each station to a cluster.
+More advanced clustering algorithms can be used as well, as long as they provide a mapping from each station to a cluster.
 
@@
 
-\caption[Profile of aggregated precipitation from 639 weather stations in Australia]{Profile of aggregated precipitation from 639 weather stations in Australia.
+\caption[Profile of aggregated precipitation at 639 weather stations in Australia]{Profile of aggregated precipitation at 639 weather stations in Australia.
 Subplot A shows the glyph map of the weekly averaged precipitation of each cluster.

### 2022-06-14 4-5-interactive.Rmd
Update 4-5-interactive.Rmd
5.6 clean-up

@@
 
-## Interactive graphic
+## Interactive graphics
 
-With spatio-temporal data, users may wish to make plots to learn the spatial distribution of a variable or to find patterns, such as trend or seasonality, in the time series.
+When dealing with spatio-temporal data, users may wish to make plots to learn the spatial distribution of a variable or to find patterns, such as trend or seasonality, in the time series.
 Combining these two types of plots with interactivity lets users link between points on the map and the corresponding time series to explore the spatial and temporal dimensions of the data simultaneously.
 Below is an example that describes the process of building an interactive graphic with \pkg{cubble} and \pkg{crosstalk}.
-The example explores the variation of monthly temperature range with \code{weatherdata::climate_full} data.
+This example explores the variation of monthly temperature range within the \code{weatherdata::climate_full} data.
 
@@
 In this example, stations on the Australia map, made from the nested form, are coloured by the calculated variance.
-A ribbon band is constructed using the long form cubble to show each station's maximum and minimum temperature across month.
-With a different dataset, users can calculate other per station measure in the nested form or make other time-wise summary of the data in the long form to customise the spatial or temporal view.
+A ribbon band is constructed using the long form cubble to show each station's maximum and minimum temperature over time.
+In the case of a different dataset, users could calculate other station-dependent measures in the nested form or make other time-wise summary of the data in the long form to customise the spatial or temporal view.
 The cross-linking between the two plots is always safeguarded by the shared \code{id} column embedded in the cubble structure.
-Below is the pseudo-code that outlines the process to construct an interactive graphic described above:
+Below is the pseudo-code that outlines the process to construct the interactive graphic described above:
 
@@
 In Figure \ref{fig:interactive-linking}, the first row shows the initial view of the interactive graphic.
-Most regions in Australia have low variance of temperature range, while the north-west coastline, bottom of South Australia, and Victoria stand out with larger monthly changes.
-In the second row, Mount Elizabeth is selected on the map given its high variance colour on the initial map and this links to the ribbon on the right.
-The third row selects the lowest temperature in August and this corresponds to Thredbo AWS on the Victoria and New South Wales border.
+Most regions in Australia have a low variance of temperature range, while the North-West coastline, the bottom of South Australia, and Victoria stand out with larger monthly changes.
+In the second row, the Mount Elizabeth station, which shows a high variance colour on the initial map, is selected on the map and this produces the ribbon on the right.
+In the third row, the lowest temperature in August is selected, which highlights the corresponding Thredbo AWS station on the left map.
 Another station in the Tasmania island is selected on the map to cross compare with Thredbo AWS.
@@
 The top row shows all locations and all temperature profiles.
-Selecting a location with high variance on the map produces the plot in the second row.
-The maximum and minimum temperature is shown using a ribbon.
-The bottom row first selects the lowest temperature in August in the seasonal display.
-A location in the Tasmania Island is then selected to compare the temperature variation with Thredbo AWS.", fig.show='hold'}
+Selecting a particular location on the map (here Mount Elizabeth) produces the plot in the second row.
+The maximum and minimum temperatures are shown using a ribbon.
+The bottom row first selects the lowest temperature in August in the seasonal display, which highlights the corresponding station on the map (Thredbo AWS).
+Another  station, located in the Tasmania Island, is then selected to compare its temperature variation with Thredbo AWS.", fig.show='hold'}
 knitr::include_graphics(here::here("figures/linking.png"))
@@
 
-This plot can also be made using \pkg{cubble} and \pkg{leaflet} where the temperature range is displayed as a small subplot upon clicking on the map.
-The procedure involves first creating the popup plots from the long form cubble as a vector and then adding these plots to a leaflet map created from the nested cubble, with \code{leafpop::addPopupGraphs()}:
+This plot can also be made using \pkg{cubble} and \pkg{leaflet}, in which case the temperature range is displayed as a small subplot upon clicking on the map.
+This procedure involves first creating the popup plots from the long form cubble as a vector and then adding these plots to a leaflet map created from the nested cubble, with \code{leafpop::addPopupGraphs()}:
 
@@
 
-Figure \ref{fig:interactive-popup} shows Figure \ref{fig:interactive-linking} made with leaflet and popups [@leafpop].
+Figure \ref{fig:interactive-popup} shows the same information as Figure \ref{fig:interactive-linking} but using leaflet and popups [@leafpop].
 
-```{r interactive-popup, fig.align='center', echo = FALSE, out.width="45%", out.height="25%", fig.retina = 2, dpi = 300, fig.cap = "Same as Figure 11 with the temperature variation shown as a popup in the leaflet map.", fig.show='hold'}
+```{r interactive-popup, fig.align='center', echo = FALSE, out.width="45%", out.height="25%", fig.retina = 2, dpi = 300, fig.cap = "Same as Figure 11 except the temperature variation is now shown as a popup in the leaflet map.", fig.show='hold'}
 knitr::include_graphics(here::here("figures/popup-mid.png"))

### 2022-06-14 4-1-basic.Rmd
Update 5.1

@@
 
-The Victoria State Government in Australia provides daily COVID information about the source, and local government area (LGA) of the recorded cases.
-This data can be used to visualise COVID spread when combined with map information on LGA, available from the Australian Bureau of Statistics.
+The Victoria State Government in Australia provides daily COVID information about the source and local government area (LGA) of recorded cases.
+This data can be used to visualise COVID spread when combined with map information available from the Australian Bureau of Statistics.
+Here we will use the data to demonstrate how to match sites between these two tables with \pkg{cubble} in the case of a few mismatches.
 The first five rows these data sets are printed below:
@@
 
-A \code{cubble} object can be created from separate spatial and temporal tables in a list, with other arguments (\code{key}, \code{index}, and \code{coords}) introduced in section \@ref(create).
-\code{as_cubble()} will automatically check the matching of sites in both tables and emit a warning if a location has missing spatial or temporal information.
+A \code{cubble} object can be created from separate spatial and temporal tables in a list, the other arguments (\code{key}, \code{index}, and \code{coords}) are described in section \@ref(create).
+The function \code{as_cubble()} will automatically try to match the sites in both tables and show a warning message in case of missing information (mismatch between the sites in the two tables).
 
@@
 
-Here, some locations have been detected with this issue and you can use \code{output = "unmatch"} to check on these locations:
+In the output above we see that this is an issue here, and we proceed to use \code{output = "unmatch"} to check which locations are concerned.
+Note that \pkg{cubble} will attempt to pair the unmatched sites, these are shown separately from the ones that cannot be paired.
 
@@
 
-\pkg{cubble} will attempt to pair the unmatched sites as well as those that cannot be paired. These can be helpful to clean up the data before creating the cubble again:
+This information is helpful when cleaning up the data before creating the cubble again:
 

### 2022-06-14 jss/cubble-jss.tex
Update 5.1

@@
 
-{\centering \includegraphics[width=1\linewidth]{../figures/diagram-keynotes/diagram-keynotes.007}
+{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-cubble/figures/diagram-keynotes/diagram-keynotes.007}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{../figures/diagram-keynotes/diagram-keynotes.008}
+{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-cubble/figures/diagram-keynotes/diagram-keynotes.008}
 
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{../figures/diagram-keynotes/diagram-keynotes.003}
+{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/ursulalaa/papers/paper-cubble/figures/diagram-keynotes/diagram-keynotes.003}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{../figures/illu-matching}
+{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-cubble/figures/illu-matching}
 
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{../figures/diagram-keynotes/diagram-keynotes.004}
+{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/ursulalaa/papers/paper-cubble/figures/diagram-keynotes/diagram-keynotes.004}
 
@@
 
-The Victoria State Government in Australia provides daily COVID information about the source, and local government area (LGA) of the recorded cases.
-This data can be used to visualise COVID spread when combined with map information on LGA, available from the Australian Bureau of Statistics.
+The Victoria State Government in Australia provides daily COVID information about the source and local government area (LGA) of recorded cases.
+This data can be used to visualise COVID spread when combined with map information available from the Australian Bureau of Statistics.
 The first five rows these data sets are printed below:
@@
 
-A \code{cubble} object can be created from separate spatial and temporal tables in a list, with other arguments (\code{key}, \code{index}, and \code{coords}) introduced in section \ref{create}.
-\code{as_cubble()} will automatically check the matching of sites in both tables and emit a warning if a location has missing spatial or temporal information.
+A \code{cubble} object can be created from separate spatial and temporal tables in a list, the other arguments (\code{key}, \code{index}, and \code{coords}) are described in section \ref{create}.
+The function \code{as_cubble()} will automatically try to match the sites in both tables and show a warning message in case of missing information (mismatch between the sites in the two tables).
 
@@
 
-{\centering \includegraphics[width=0.9\linewidth]{../figures/basic-manip}
+{\centering \includegraphics[width=0.9\linewidth]{/Users/ursulalaa/papers/paper-cubble/figures/basic-manip}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{../figures/basic-agg}
+{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-cubble/figures/basic-agg}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{../figures/matching-map}
+{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-cubble/figures/matching-map}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{../figures/matching}
+{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-cubble/figures/matching}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{../figures/netcdf}
+{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-cubble/figures/netcdf}
 
@@
 
-\hypertarget{interactive-graphic}{%
-\subsection{Interactive graphic}\label{interactive-graphic}}
-
-With spatio-temporal data, users may wish to make plots to learn the spatial distribution of a variable or to find patterns, such as trend or seasonality, in the time series.
+\hypertarget{interactive-graphics-1}{%
+\subsection{Interactive graphics}\label{interactive-graphics-1}}
+
+When dealing with spatio-temporal data, users may wish to make plots to learn the spatial distribution of a variable or to find patterns, such as trend or seasonality, in the time series.
 Combining these two types of plots with interactivity lets users link between points on the map and the corresponding time series to explore the spatial and temporal dimensions of the data simultaneously.
 Below is an example that describes the process of building an interactive graphic with \pkg{cubble} and \pkg{crosstalk}.
-The example explores the variation of monthly temperature range with \code{weatherdata::climate_full} data.
+This example explores the variation of monthly temperature range within the \code{weatherdata::climate_full} data.
 
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.23\textheight]{../figures/linking} \includegraphics[width=1\linewidth,height=0.23\textheight]{../figures/linking-north} \includegraphics[width=1\linewidth,height=0.23\textheight]{../figures/linking-lower}
+{\centering \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/ursulalaa/papers/paper-cubble/figures/linking} \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/ursulalaa/papers/paper-cubble/figures/linking-north} \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/ursulalaa/papers/paper-cubble/figures/linking-lower}
 
@@
 The top row shows all locations and all temperature profiles.
-Selecting a location with high variance on the map produces the plot in the second row.
-The maximum and minimum temperature is shown using a ribbon.
-The bottom row first selects the lowest temperature in August in the seasonal display.
-A location in the Tasmania Island is then selected to compare the temperature variation with Thredbo AWS.}\label{fig:interactive-linking}
+Selecting a particular location on the map (here Mount Elizabeth) produces the plot in the second row.
+The maximum and minimum temperatures are shown using a ribbon.
+The bottom row first selects the lowest temperature in August in the seasonal display, which highlights the corresponding station on the map (Thredbo AWS).
+Another  station, located in the Tasmania Island, is then selected to compare its temperature variation with Thredbo AWS.}\label{fig:interactive-linking}
 \end{figure}
@@
 
-Figure \ref{fig:interactive-popup} shows Figure \ref{fig:interactive-linking} made with leaflet and popups \citep{leafpop}.
+Figure \ref{fig:interactive-popup} shows the same information as Figure \ref{fig:interactive-linking} but using leaflet and popups \citep{leafpop}.
 
@@
 
-{\centering \includegraphics[width=0.45\linewidth,height=0.25\textheight]{../figures/popup-mid}
-
-}
-
-\caption[Same as Figure 11 with the temperature variation shown as a popup in the leaflet map]{Same as Figure 11 with the temperature variation shown as a popup in the leaflet map.}\label{fig:interactive-popup}
+{\centering \includegraphics[width=0.45\linewidth,height=0.25\textheight]{/Users/ursulalaa/papers/paper-cubble/figures/popup-mid}
+
+}
+
+\caption[Same as Figure 11 except the temperature variation is now shown as a popup in the leaflet map]{Same as Figure 11 except the temperature variation is now shown as a popup in the leaflet map.}\label{fig:interactive-popup}
 \end{figure}

### 2022-06-14 4-3-water-level.Rmd
Section 5.4

@@
 The collected variables include: electrical conductivity, turbidity, watercourse discharge, watercourse level, and water temperature.
-In particular, water level will interact with precipitation from the climate data since rainfall will raise the water level in the river.
-Figure \ref{fig:matching-map} gives the location of available weather stations and water gauges in Victoria.
+We expect that the water level should be related to the precipitation available from climate data, since rainfall will raise the water level in the river.
+Figure \ref{fig:matching-map} shows the location of available weather stations and water gauges in Victoria, and we will try to match water gauges with nearby weather stations, taking into account both spatial and temporal information from 2020.
 
@@
 
-From the map, one can see that a few water gauges and weather stations are close to each other and the fluctuation of the water level could be matched up with the precipitation measured by the nearby climate station.
-As introduced in Section \@ref(matching), \code{match_sites()} can be used to match one source of data with another source in a cubble.
-Here \code{Water_course_level} in \code{river} will be matched to \code{prcp} in \code{climate} in 2020.
-The two datasets need to be specified as the first two arguments and the variable to match can be specified in \code{temporal_by} using the \code{by} syntax in \code{join}.
-\code{temporal_independent} controls the variable used to construct the interval and the goal here is to see if the water level in the river will reflect precipitation.
-This puts precipitation \code{prcp}, as the independent variable.
-Given there is one year's worth of data, the number of peaks (\code{temporal_n_highest}) to consider is slightly raised from a default of 20 to 30 and \code{temporal_min_match} is raised accordingly.
-To return all the pairs of the match, \code{temporal_min_match} can be set to 0.
+As introduced in Section \@ref(matching), \code{match_sites()} can be used for this task, and we choose \code{Water_course_level} in \code{river} and \code{prcp} in \code{climate} for the temporal matching.
+We first pass in the two datasets, followed by the variables used for the temporal matching, specified in \code{temporal_by} (using the \code{by} syntax from \code{dplyr::join}).
+With \code{temporal_independent} we select precipitation (\code{prcp}) as the independent variable, and the goal is to see if the water level in the river reflects what is observed in precipitation (for nearby stations).
+Since we consider a full year of data, we set the number of peaks considered (\code{temporal_n_highest}) to 30 (slightly above the default value of 20), and \code{temporal_min_match} is raised accordingly.
 
@@
 
-The output from matching is also a cubble, with additional column \code{dist} and \code{group} produced from spatial matching and \code{n_match} from temporal matching.
+The output from this function is also a cubble, with additional columns \code{dist} and \code{group} produced from spatial matching, and \code{n_match} from temporal matching.
 
@@
 
-Figure \ref{fig:matching} plots the matched pairs on the map.
-There are four pairs of matches, all located in the middle of Victoria and the expected concurrent increase in precipitation and water level can be observed.
+Figure \ref{fig:matching} (A) shows the four matched pairs on the map.
+For these the expected concurrent increase in precipitation and water level can be seen clearly when comparing these variables on a standardised scale, see Figure \ref{fig:matching} (B).
 
@@
 Precipitation and water level have been standardised between 0 and 1 to be displayed on the same scale.
-The water level reflects the increase in precipitation."}
+The water level reflects the increase in precipitation.
+The numbers (1, 5, 6, 10) indicate the group index derived from spatial matching, only those that were selectd by temporal matching are shown here."}
 knitr::include_graphics(here::here("figures/matching.png"))

### 2022-06-14 jss/cubble-jss.tex
Section 5.4

@@
 This data can be used to visualise COVID spread when combined with map information available from the Australian Bureau of Statistics.
+Here we will use the data to demonstrate how to match sites between these two tables with \pkg{cubble} in the case of a few mismatches.
 The first five rows these data sets are printed below:
@@
 
-From the map, one can see that a few water gauges and weather stations are close to each other and the fluctuation of the water level could be matched up with the precipitation measured by the nearby climate station.
-As introduced in Section \ref{matching}, \code{match_sites()} can be used to match one source of data with another source in a cubble.
-Here \code{Water_course_level} in \code{river} will be matched to \code{prcp} in \code{climate} in 2020.
-The two datasets need to be specified as the first two arguments and the variable to match can be specified in \code{temporal_by} using the \code{by} syntax in \code{join}.
-\code{temporal_independent} controls the variable used to construct the interval and the goal here is to see if the water level in the river will reflect precipitation.
-This puts precipitation \code{prcp}, as the independent variable.
-Given there is one year's worth of data, the number of peaks (\code{temporal_n_highest}) to consider is slightly raised from a default of 20 to 30 and \code{temporal_min_match} is raised accordingly.
-To return all the pairs of the match, \code{temporal_min_match} can be set to 0.
+As introduced in Section \ref{matching}, \code{match_sites()} can be used for this task, and we choose \code{Water_course_level} in \code{river} and \code{prcp} in \code{climate} for the temporal matching.
+We first pass in the two datasets, followed by the variables used for the temporal matching, specified in \code{temporal_by} (using the \code{by} syntax from \code{dplyr::join}).
+With \code{temporal_independent} we select precipitation (\code{prcp}) as the independent variable, and the goal is to see if the water level in the river reflects what is observed in precipitation (for nearby stations).
+Since we consider a full year of data, we set the number of peaks considered (\code{temporal_n_highest}) to 30 (slightly above the default value of 20), and \code{temporal_min_match} is raised accordingly.
 
@@
 Precipitation and water level have been standardised between 0 and 1 to be displayed on the same scale.
-The water level reflects the increase in precipitation.}\label{fig:matching}
+The water level reflects the increase in precipitation.
+The numbers (1, 5, 6, 10) indicate the group index derived from spatial matching, only those that were selectd by temporal matching are shown here.}\label{fig:matching}
 \end{figure}

### 2022-06-14 4-2-agg.Rmd
evaluate script inside rmd: section 5.2

@@
 
-```{r echo = TRUE, eval = FALSE}
-station_nested <- weatherdata::climate_full |>
-  mutate(cluster = ...)
+```{r eval = FALSE, echo = TRUE}
+station_nested <- weatherdata::climate_full |> mutate(cluster = ...)
 ```
@@
 
-```{r basic-agg, out.width="100%", fig.cap="Profile of aggregated precipitation at 639 weather stations in Australia.
+```{r eval = FALSE}
+coords <- cbind(prcp_aus$long, prcp_aus$lat)
+dist_raw <- geosphere::distm(coords, coords)
+
+station_long <- cubble::prcp_aus |>
+  face_temporal(ts) |>
+  mutate(wk = lubridate::week(date)) |>
+  group_by(wk) |>
+  summarise(prcp = sum(prcp, na.rm = TRUE)) |>
+  unfold(cluster)
+
+set.seed(123)
+station_nested <- station_long |>
+  face_spatial() |>
+  strip_rowwise() |>
+  mutate(cluster = kmeans(dist_raw,centers = 20, nstart = 500)$cluster)
+save(station_nested, file = here::here("data/station_nested.rda"))
+```
+
+```{r basic-agg, fig.height = 7, fig.width = 15, fig.cap="Profile of aggregated precipitation at 639 weather stations in Australia.
 Subplot A shows the glyph map of the weekly averaged precipitation of each cluster.
@@
 Subplot B shows the station membership of each cluster."}
-knitr::include_graphics(here::here("figures/basic-agg.png"))
+load(here::here("data/station_nested.rda"))
+cluster_nested <- station_nested |>
+  switch_key(cluster) |> 
+  get_centroid()
+
+cluster_long <- cluster_nested |> 
+  face_temporal() |> 
+  group_by(wk) |>
+  summarise(prcp = mean(prcp, na.rm = TRUE)) |> 
+  unfold(cent_long, cent_lat)
+
+state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)
+
+ggplot_smooth <- cluster_long |> 
+  ggplot() +
+  geom_smooth(aes(x = wk, y = prcp, group = cluster), span = 0.4)
+
+smoother <- layer_data(ggplot_smooth) |> 
+  left_join(cluster_long |> select(cluster, cent_long, cent_lat), by = c("group" = "cluster"))
+
+p1 <- ggplot(data = smoother, 
+             aes(x_minor = x, y_minor = y, x_major = cent_long, y_major = cent_lat)) + 
+  geom_sf(data = state_map, inherit.aes = FALSE,
+          color = "grey80", alpha = 0.4, linetype = 3) + 
+  geom_text(data = cluster_nested, 
+            aes(x = cent_long, y = cent_lat, label = cluster), inherit.aes = FALSE) +
+  geom_glyph(height = 2, width = 4) + 
+  theme_void()
+
+p2 <- ggplot() + 
+  geom_sf(data = state_map, inherit.aes = FALSE,
+          color = "grey80", alpha = 0.4, linetype = 3) + 
+  geom_point(data = station_nested, aes(x = long, y = lat), size = 0.5) +
+  ggforce::geom_mark_hull(data = cluster_nested |> tidyr::unnest(hull),
+                          expand = 0, radius = 0,
+                          aes(x = long, y = lat, group = cluster)) + 
+  theme_void()
+
+(p1 | p2) + plot_annotation(tag_levels = 'A')
 ```

### 2022-06-14 jss/cubble-jss.tex
evaluate script inside rmd: section 5.2

--- 2022-06-14 jss/cubble-jss.tex
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-cubble/figures/diagram-keynotes/diagram-keynotes.007}
+{\centering \includegraphics[width=1\linewidth]{../figures/diagram-keynotes/diagram-keynotes.007}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-cubble/figures/diagram-keynotes/diagram-keynotes.008}
+{\centering \includegraphics[width=1\linewidth]{../figures/diagram-keynotes/diagram-keynotes.008}
 
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/ursulalaa/papers/paper-cubble/figures/diagram-keynotes/diagram-keynotes.003}
+{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{../figures/diagram-keynotes/diagram-keynotes.003}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-cubble/figures/illu-matching}
+{\centering \includegraphics[width=1\linewidth]{../figures/illu-matching}
 
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/ursulalaa/papers/paper-cubble/figures/diagram-keynotes/diagram-keynotes.004}
+{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{../figures/diagram-keynotes/diagram-keynotes.004}
 
@@
 
-{\centering \includegraphics[width=0.9\linewidth]{/Users/ursulalaa/papers/paper-cubble/figures/basic-manip}
+{\centering \includegraphics[width=0.9\linewidth]{../figures/basic-manip}
 
@@
 
-\code{weatherdata::climate_full}, also extracted from the GHCN, records daily precipitation and maximum/minimum temperature for 639 stations in Australia from 2016 to 2020.
+\code{weatherdata::climate_full}, also extracted from the GHCN, records daily precipitation and maximum/minimum temperature for 640 stations in Australia from 2016 to 2020.
 A simple \(k\)-means algorithm based on the distance matrix between stations is used to create 20 clusters.
@@
 \begin{CodeInput}
-R> station_nested <- weatherdata::climate_full |>
-+   mutate(cluster = ...)
+R> station_nested <- weatherdata::climate_full |> mutate(cluster = ...)
 \end{CodeInput}
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-cubble/figures/basic-agg}
+{\centering \includegraphics{figures/basic-agg-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-cubble/figures/matching-map}
+{\centering \includegraphics[width=1\linewidth]{../figures/matching-map}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeOutput}
-# cubble:   id [8]: nested form
-# bbox:     [144.52, -37.73, 146.06, -36.55]
-# temporal: date [date], matched_var [dbl]
-  id          name        lat  long type   dist group ts       n_match
-  <chr>       <chr>     <dbl> <dbl> <chr> <dbl> <int> <list>     <int>
-1 405234      SEVEN CR~ -36.9  146. river  6.15     5 <tibble>      21
-2 404207      HOLLAND ~ -36.6  146. river  8.54    10 <tibble>      21
-3 ASN00082042 strathbo~ -36.8  146. clim~  6.15     5 <tibble>      21
-4 ASN00082170 benalla ~ -36.6  146. clim~  8.54    10 <tibble>      21
-5 230200      MARIBYRN~ -37.7  145. river  6.17     6 <tibble>      19
-6 ASN00086038 essendon~ -37.7  145. clim~  6.17     6 <tibble>      19
-7 406213      CAMPASPE~ -37.0  145. river  1.84     1 <tibble>      18
-8 ASN00088051 redesdale -37.0  145. clim~  1.84     1 <tibble>      18
-\end{CodeOutput}
-\end{CodeChunk}
-
 Figure \ref{fig:matching} plots the matched pairs on the map.
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-cubble/figures/matching}
+{\centering \includegraphics[width=1\linewidth]{../figures/matching}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-cubble/figures/netcdf}
+{\centering \includegraphics[width=1\linewidth]{../figures/netcdf}
 
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/ursulalaa/papers/paper-cubble/figures/linking} \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/ursulalaa/papers/paper-cubble/figures/linking-north} \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/ursulalaa/papers/paper-cubble/figures/linking-lower}
+{\centering \includegraphics[width=1\linewidth,height=0.23\textheight]{../figures/linking} \includegraphics[width=1\linewidth,height=0.23\textheight]{../figures/linking-north} \includegraphics[width=1\linewidth,height=0.23\textheight]{../figures/linking-lower}
 
@@
 
-{\centering \includegraphics[width=0.45\linewidth,height=0.25\textheight]{/Users/ursulalaa/papers/paper-cubble/figures/popup-mid}
+{\centering \includegraphics[width=0.45\linewidth,height=0.25\textheight]{../figures/popup-mid}
 

### 2022-06-14 4-3-water-level.Rmd

@@
 
-```{r}
+```{r eval = FALSE}
 climate <- weatherdata::climate_full |>
@@
 
-```{r echo = TRUE}
+```{r echo = TRUE,  eval = FALSE}
 res <- match_sites(
@@
 
-```{r}
+```{r,  eval = FALSE}
 res |> print(n = 8)

### 2022-06-14 jss/cubble-jss.tex

@@
 #   [dbl]
-  id          date        prcp  tmax  tmin
-  <chr>       <date>     <dbl> <dbl> <dbl>
-1 ASN00009021 2020-01-01     0  31.9  15.3
-2 ASN00009021 2020-01-02     0  24.9  16.4
-3 ASN00009021 2020-01-03     6  23.2  13  
-4 ASN00009021 2020-01-04     0  28.4  12.4
-5 ASN00009021 2020-01-05     0  35.3  11.6
+  id          date        prcp  tmax  tmin  long   lat
+  <chr>       <date>     <dbl> <dbl> <dbl> <dbl> <dbl>
+1 ASN00009021 2020-01-01     0  31.9  15.3  116. -31.9
+2 ASN00009021 2020-01-02     0  24.9  16.4  116. -31.9
+3 ASN00009021 2020-01-03     6  23.2  13    116. -31.9
+4 ASN00009021 2020-01-04     0  28.4  12.4  116. -31.9
+5 ASN00009021 2020-01-05     0  35.3  11.6  116. -31.9
 # ... with 1,825 more rows
@@
 \end{CodeOutput}
+\begin{CodeOutput}
+Joining, by = "lga"
+\end{CodeOutput}
 \end{CodeChunk}
@@
 
-The output from matching is also a cubble, with additional column \code{dist} and \code{group} produced from spatial matching and \code{n_match} from temporal matching.
-
-Figure \ref{fig:matching} plots the matched pairs on the map.
-There are four pairs of matches, all located in the middle of Victoria and the expected concurrent increase in precipitation and water level can be observed.
+The output from this function is also a cubble, with additional columns \code{dist} and \code{group} produced from spatial matching, and \code{n_match} from temporal matching.
+
+Figure \ref{fig:matching} (A) shows the four matched pairs on the map.
+For these the expected concurrent increase in precipitation and water level can be seen clearly when comparing these variables on a standardised scale, see Figure \ref{fig:matching} (B).
 

### 2022-06-14 4-3-water-level.Rmd
evaluate script inside rmd section 5.3

--- 2022-06-14 4-3-water-level.Rmd
@@
 
-```{r matching-map, out.width="100%", fig.cap="Location of weather stations and river gauges in Victoria, Australia."}
-knitr::include_graphics(here::here("figures/matching-map.png"))
+```{r matching-map, fig.height = 7, fig.width=10, fig.cap="Location of weather stations and river gauges in Victoria, Australia."}
+vic_map <- rmapshaper::ms_simplify(ozmaps::abs_ste |> filter(NAME == "Victoria"))  
+ggplot() + 
+  geom_sf(data = vic_map, fill = "transparent", color = "grey") + 
+  geom_point(data = dplyr::bind_rows(river, climate), 
+             aes(x = long, y = lat, color = type)) + 
+  ggplot2::theme_bw() +
+  ggplot2::theme(legend.position = "bottom") +
+  ggplot2::labs(x = "Longitude", y = "Latitude") + 
+  scale_color_brewer(palette = "Dark2") 
 ```
@@
 ```{r}
-res |> print(n = 8)
+res |> print()
 ```
@@
 
-```{r matching, out.width="100%", fig.cap="Matched weather stations and river gauges on the map (A) and across time (B).
+```{r matching, fig.height = 5, fig.width = 10, fig.cap="Matched weather stations and river gauges on the map (A) and across time (B).
 Precipitation and water level have been standardised between 0 and 1 to be displayed on the same scale.
@@
 The numbers (1, 5, 6, 10) indicate the group index derived from spatial matching, only those that were selectd by temporal matching are shown here."}
-knitr::include_graphics(here::here("figures/matching.png"))
+p1 <-ggplot() + 
+  geom_sf(data = vic_map, fill = "transparent", color = "grey") + 
+  geom_point(data = res, aes(x = long, y = lat, color = type)) +
+  ggrepel::geom_label_repel(
+    data = res |> filter(type == "river"), aes(x = long, y = lat, label = group)) +
+  scale_color_brewer(palette = "Dark2")  + 
+  ggplot2::theme_bw() +
+  ggplot2::theme(legend.position = "bottom") +
+  ggplot2::labs(x = "Longitude", y = "Latitude")
+
+res_long <- res |>
+  face_temporal(ts) |>
+  unfold(group, type) |>
+  rename(prcp = matched_var) |> 
+  mutate(prcp = (prcp - min(prcp, na.rm = TRUE))/ (max(prcp, na.rm = TRUE) - min(prcp, na.rm = TRUE)))
+
+p2 <- res_long |> 
+  ggplot(aes(x = date, y = prcp, color = type, group = id)) +
+  geom_line(alpha = 0.8) +
+  facet_wrap(vars(group)) +
+  scale_color_brewer(palette = "Dark2", guide = "none") +
+  theme_bw() +
+  scale_x_date(date_labels = "%b") + 
+  labs(x = "Date", y = "Precipitation/ water level")
+
+(p1 | p2) + 
+  patchwork::plot_layout(guides = "collect") + 
+  plot_annotation(tag_levels = 'A')&
+  theme(legend.position = "bottom")
 ```

### 2022-06-14 jss/cubble-jss.Rmd
proofread
evaluate script inside rmd: section 5.2
evaluate script inside rmd section 5.3

@@
   cache = TRUE,
-  fig.path = "figures/",
-  fig.height = 10
+  fig.path = "figures/"
 )
@@
 library(cubble)
-library(dplyr)
+library(tidyverse)
+library(patchwork)
 ```

### 2022-06-14 scripts/basic-agg.R

@@ -1,59 +1,0 @@
-library(tidyverse)
-library(cubble)
-library(patchwork)
-coords <- cbind(prcp_aus$long, prcp_aus$lat)
-dist_raw <- geosphere::distm(coords, coords)
-
-station_long <- prcp_aus |> 
-  face_temporal(ts) |> 
-  mutate(wk = lubridate::week(date)) |> 
-  group_by(wk) |> 
-  summarise(prcp = sum(prcp, na.rm = TRUE)) |> 
-  unfold(cluster)
-
-set.seed(123)
-station_nested <- station_long |> 
-  face_spatial() |> 
-  strip_rowwise() |> 
-  mutate(cluster = kmeans(dist_raw,centers = 20, nstart = 500)$cluster)
-#save(station_nested, file = here::here("data/station_nested.rda"))
-
-cluster_nested <- station_nested |>
-  switch_key(cluster) |> 
-  get_centroid()
-
-cluster_long <- cluster_nested |> 
-  face_temporal(ts) |> 
-  group_by(wk) |>
-  summarise(prcp = mean(prcp, na.rm = TRUE)) |> 
-  unfold(cent_long, cent_lat)
-
-state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)
-
-ggplot_smooth <- cluster_long |> 
-  ggplot() +
-  geom_smooth(aes(x = wk, y = prcp, group = cluster), span = 0.4) 
-
-smoother <- layer_data(ggplot_smooth) |> 
-  left_join(cluster_long |> select(cluster, cent_long, cent_lat), by = c("group" = "cluster"))
-
-p1 <- ggplot(data = smoother, 
-             aes(x_minor = x, y_minor = y, x_major = cent_long, y_major = cent_lat)) + 
-  geom_sf(data = state_map, inherit.aes = FALSE,
-          color = "grey80", alpha = 0.4, linetype = 3) + 
-  geom_text(data = cluster_nested, 
-            aes(x = cent_long, y = cent_lat, label = cluster), inherit.aes = FALSE) +
-  geom_glyph(height = 2, width = 4) + 
-  theme_void()
-
-p3 <- ggplot() + 
-  geom_sf(data = state_map, inherit.aes = FALSE,
-          color = "grey80", alpha = 0.4, linetype = 3) + 
-  geom_point(data = station_nested, aes(x = long, y = lat), size = 0.5) +
[…]

### 2022-06-14 scripts/matching.R

@@ -1,61 +1,0 @@
-library(tidyverse)
-library(cubble)
-library(patchwork)
-
-climate <- prcp_aus |>
-  filter(between(stringr::str_sub(id, 7, 8), 76, 90)) |>
-  face_temporal() |>
-  filter(lubridate::year(date) == 2020) |>
-  face_spatial() |>
-  mutate(type = "climate")
-
-river <- river |> mutate(type = "river")
-
-vic_map <- rmapshaper::ms_simplify(ozmaps::abs_ste |> filter(NAME == "Victoria"))  
-ggplot() + 
-  geom_sf(data = vic_map, fill = "transparent", color = "grey") + 
-  geom_point(data = dplyr::bind_rows(river, climate), 
-             aes(x = long, y = lat, color = type)) + 
-  ggplot2::theme_bw() +
-  ggplot2::theme(legend.position = "bottom") +
-  ggplot2::labs(x = "Longitude", y = "Latitude") + 
-  scale_color_brewer(palette = "Dark2") 
-ggsave(filename = "figures/matching-map.png", width = 10, height = 5)  
-
-res <- match_sites(river, climate,
-                   temporal_by = c("Water_course_level" = "prcp"),
-                   temporal_independent = "prcp",  
-                   temporal_n_highest = 30,
-                   temporal_min_match = 15)
-
-p1 <-ggplot() + 
-  geom_sf(data = vic_map, fill = "transparent", color = "grey") + 
-  geom_point(data = res, aes(x = long, y = lat, color = type)) +
-  ggrepel::geom_label_repel(
-    data = res |> filter(type == "river"), aes(x = long, y = lat, label = group)) +
-  scale_color_brewer(palette = "Dark2")  + 
-  ggplot2::theme_bw() +
-  ggplot2::theme(legend.position = "bottom") +
-  ggplot2::labs(x = "Longitude", y = "Latitude")
-
-res_long <- res |>
-  face_temporal(ts) |>
-  unfold(group, type) |>
-  rename(prcp = matched_var) |> 
-  mutate(prcp = (prcp - min(prcp, na.rm = TRUE))/ (max(prcp, na.rm = TRUE) - min(prcp, na.rm = TRUE))) 
-
-p2 <- res_long |> 
-  ggplot(aes(x = date, y = prcp, color = type, group = id)) +
-  geom_line(alpha = 0.8) +
-  facet_wrap(vars(group)) +
[…]

### 2022-06-15 2-the-cubble-pkg.Rmd
working on applications

@@
 
-This function should generally be used in the last step of the analysis since it is a temporary operation, meaning that these added spatial variables will disappear, if switched to the nested form and then switched back:
+This function should generally be used in the last step of the analysis since it is a temporary operation, meaning that these added spatial variables will disappear if switched to the nested form and then switched back:
 

### 2022-06-15 4-1-basic.Rmd
working on applications

@@
 # Examples {#examples}
+
+The six examples here are chosen to illustrate creating a cubble for exploring COVID data where id's need to be matched, using spatial transformations to make a glyph map of seasonal temperature changes over years, aggregating information spatially to explore precipitation patterns, matching river level data and weather station records for analysis of water supply, reading netCDF format data to reproduce a climate reanalysis plot, and the workflow to create complex interactive linked plots.
 

### 2022-06-15 jss/cubble-jss.tex
evaluate script inside rmd section 5.3
working on applications

@@
 #   [dbl]
-  id          date        prcp  tmax  tmin  long   lat
-  <chr>       <date>     <dbl> <dbl> <dbl> <dbl> <dbl>
-1 ASN00009021 2020-01-01     0  31.9  15.3  116. -31.9
-2 ASN00009021 2020-01-02     0  24.9  16.4  116. -31.9
-3 ASN00009021 2020-01-03     6  23.2  13    116. -31.9
-4 ASN00009021 2020-01-04     0  28.4  12.4  116. -31.9
-5 ASN00009021 2020-01-05     0  35.3  11.6  116. -31.9
+  id          date        prcp  tmax  tmin
+  <chr>       <date>     <dbl> <dbl> <dbl>
+1 ASN00009021 2020-01-01     0  31.9  15.3
+2 ASN00009021 2020-01-02     0  24.9  16.4
+3 ASN00009021 2020-01-03     6  23.2  13  
+4 ASN00009021 2020-01-04     0  28.4  12.4
+5 ASN00009021 2020-01-05     0  35.3  11.6
 # ... with 1,825 more rows
@@
 \section{Examples}\label{examples}}
+
+The six examples here are chosen to illustrate creating a cubble for exploring COVID data where id's need to be matched, using spatial transformations to make a glyph map of seasonal temperature changes over years, aggregating information spatially to explore precipitation patterns, matching river level data and weather station records for analysis of water supply, reading netCDF format data to reproduce a climate reanalysis plot, and the workflow to create complex interactive linked plots.
 
@@
 ! Use argument `output = "unmatch"` to check on the unmatched key
-\end{CodeOutput}
-\begin{CodeOutput}
-Joining, by = "lga"
 \end{CodeOutput}
@@
 
-\code{weatherdata::climate_full}, also extracted from the GHCN, records daily precipitation and maximum/minimum temperature for 640 stations in Australia from 2016 to 2020.
+\code{weatherdata::climate_full}, also extracted from the GHCN, records daily precipitation and maximum/minimum temperature for 639 stations in Australia from 2016 to 2020.
 A simple \(k\)-means algorithm based on the distance matrix between stations is used to create 20 clusters.
@@
 
-{\centering \includegraphics[width=1\linewidth]{../figures/matching-map}
+{\centering \includegraphics{figures/matching-map-1}
 
@@
 
+\begin{CodeChunk}
+\begin{CodeOutput}
+# cubble:   id [8]: nested form
+# bbox:     [144.52, -37.73, 146.06, -36.55]
+# temporal: date [date], matched_var [dbl]
+  id          name        lat  long type   dist group ts       n_match
+  <chr>       <chr>     <dbl> <dbl> <chr> <dbl> <int> <list>     <int>
+1 405234      SEVEN CR~ -36.9  146. river  6.15     5 <tibble>      21
+2 404207      HOLLAND ~ -36.6  146. river  8.54    10 <tibble>      21
+3 ASN00082042 strathbo~ -36.8  146. clim~  6.15     5 <tibble>      21
+4 ASN00082170 benalla ~ -36.6  146. clim~  8.54    10 <tibble>      21
+5 230200      MARIBYRN~ -37.7  145. river  6.17     6 <tibble>      19
+# ... with 3 more rows
+\end{CodeOutput}
+\end{CodeChunk}
+
 Figure \ref{fig:matching} (A) shows the four matched pairs on the map.
@@
 
-{\centering \includegraphics[width=1\linewidth]{../figures/matching}
+{\centering \includegraphics{figures/matching-1}
 

### 2022-06-15 4-1-basic.Rmd
eval script 5.1

@@
 The **cubble** dataset \code{weatherdata::historical_tmax} extracts the maximum temperature for `r nrow(weatherdata::historical_tmax)` Australian stations from the GHCN starting from year `r range(weatherdata::historical_tmax$first_year)[2]` and provides information also on the latitude, longitude and elevation of each of the  stations.
-This **cubble** data  is already cast into a \code{cubble} data object, with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates.
+This **cubble** data  is already cast into a \code{cubble} data object, with \code{id} as the key, \code{date} as the index, and \code{c(long, lat)} as the coordinates.
 The goal of this example is to compare the monthly average maximum temperature  between two periods, 1971-1975 and 2016-2020, for stations in Victoria and New South Wales.
@@
 
-```{r eval = FALSE, echo = TRUE}
+```{r echo = TRUE}
 tmax <- weatherdata::historical_tmax |>
@@
 
-and extracting the time periods 1971-1975 and 2016-2020 is a time operation and therefore the nested cubble needs to be switched to the long cubble form by \code{stretch()}:
+and extracting the time periods 1971-1975 and 2016-2020 is a time operation and therefore the nested cubble needs to be switched to the long cubble form by \code{face_temporal()}:
 
-```{r eval=  FALSE, echo = TRUE}
+```{r echo = TRUE}
 tmax <- tmax |>
@@
 
-```{r eval=  FALSE, echo = TRUE}
+```{r echo = TRUE}
 tmax <- tmax |>
@@
 
-```{r eval = FALSE, echo = TRUE}
+```{r echo = TRUE}
 tmax <- tmax |> face_spatial() |> filter(nrow(ts) == 24)
@@
 
-Lastly, to create a glyph map where the series of each station are overlayed on a map of Australia, both the major (\code{longitude}, \code{latitude}) and minor (\code{month}, \code{tmax}) coordinates defining the bounding box for the plot need to be on the same table.
+Lastly, to create a glyph map where the series of each station are overlayed on a map of Australia, both the major (\code{long}, \code{lat}) and minor (\code{month}, \code{tmax}) coordinates defining the bounding box for the plot need to be on the same table.
 For that, spatial variables can be moved to the long form with \code{unfold()}:
 
-```{r eval = FALSE, echo = TRUE}  
-tmax <- tmax |> face_temporal() |> unfold(latitude, longitude)
+```{r echo = TRUE}  
+tmax <- tmax |> face_temporal() |> unfold(long, lat)
 ```
@@
 
-```{r basic-manip, out.width="90%", fig.cap="A glyph map of the average maximum temperature by month for parts of Victoria and New South Wales weather stations in Australia for two five year periods (1971-1975, 2016-2020).
+```{r create-inset, eval = FALSE}
+
+tmax |> filter(id == "ASN00048027") |>
+  ggplot(aes(x = month,
+             y = tmax,
+             color = group)) +
+  geom_line(size = 1.5) +
+  scale_color_brewer(palette = "Dark2", guide = "none") +
+  scale_x_continuous(breaks = seq(1, 12, 1), 
+                     labels = c("J", "F", "M", "A", "M", "J", 
+                                "J", "A", "S", "O", "N", "D")) +
+  labs(x = "", y  = "Temp (C)", title = "ASN00048027: Cobar") +
+  theme_bw() +
+  theme(
+    aspect.ratio = 0.3,
+    axis.text = element_text(size = 20),
+    title =  element_text(size = 20)
+  )
+
+#ggsave(filename = here::here("figures/basic-manip-inset.png"), width = 7)
+```
+
+```{r basic-manip, fig.height = 7, fig.width = 10, fig.cap="A glyph map of the average maximum temperature by month for parts of Victoria and New South Wales weather stations in Australia for two five year periods (1971-1975, 2016-2020).
 An inset (top left) showing detail is the station Cobar (highlighted by black rectangle).
 The mothly trend is similar at all locations (low in the winter, high in the summer), and small increased temperatures, particularly in late summer can be seen at most stations."}
-knitr::include_graphics(here::here("figures/basic-manip.png"))
+nsw_vic <- ozmaps::abs_ste |> filter(NAME %in% c("Victoria", "New South Wales"))
+
+p1 <- ggplot() + 
+  geom_sf(data = nsw_vic, fill = "grey95", color = "white") + 
+  geom_glyph(data = tmax, 
+             aes(x_major = long, x_minor = month, 
+                 y_major = lat, y_minor = tmax,
+                 group = interaction(id, group), 
+                 color = group),
+             width = 1, height = 0.5) +
+  scale_color_brewer("", palette = "Dark2") + 
+  theme_bw() + 
+  coord_sf(xlim = c(141, 154), ylim = c(-36, -29)) + 
+  ggthemes::theme_map() +
+  theme(legend.position = "bottom") +
+  labs(x = "Longitude", y = "Latitude")
+
+box_df <- tmax |> face_spatial() |> filter(id == "ASN00048027")
+single <- tibble::tibble(img = here::here("figures/basic-manip-inset.png"))
+
+p1 +
+  geom_rect(data = box_df,
+            aes(xmin = long - 0.6, xmax = long + 0.6,
+                ymin = lat - 0.12, ymax = lat + 0.35),
+            fill = "transparent", color = "black") +
+  ggimg::geom_point_img(
+    data = single, aes(x = 143.7, y = -30, img = img), size = 6)
+
 ```

### 2022-06-15 jss/cubble-jss.tex
eval script 5.1

@@
 ! Use argument `output = "unmatch"` to check on the unmatched key
+\end{CodeOutput}
+\begin{CodeOutput}
+Joining, by = "lga"
 \end{CodeOutput}
@@
 The Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world.
-The \textbf{cubble} dataset \code{weatherdata::historical_tmax} extracts the maximum temperature for 236 Australian stations from the GHCN starting from year 1969 and provides information also on the latitude, longitude and elevation of each of the stations.
-This \textbf{cubble} data is already cast into a \code{cubble} data object, with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates.
+The \textbf{cubble} dataset \code{weatherdata::historical_tmax} extracts the maximum temperature for 204 Australian stations from the GHCN starting from year 1969 and provides information also on the latitude, longitude and elevation of each of the stations.
+This \textbf{cubble} data is already cast into a \code{cubble} data object, with \code{id} as the key, \code{date} as the index, and \code{c(long, lat)} as the coordinates.
 The goal of this example is to compare the monthly average maximum temperature between two periods, 1971-1975 and 2016-2020, for stations in Victoria and New South Wales.
@@
 
-and extracting the time periods 1971-1975 and 2016-2020 is a time operation and therefore the nested cubble needs to be switched to the long cubble form by \code{stretch()}:
+and extracting the time periods 1971-1975 and 2016-2020 is a time operation and therefore the nested cubble needs to be switched to the long cubble form by \code{face_temporal()}:
 
@@
 
-Lastly, to create a glyph map where the series of each station are overlayed on a map of Australia, both the major (\code{longitude}, \code{latitude}) and minor (\code{month}, \code{tmax}) coordinates defining the bounding box for the plot need to be on the same table.
+Lastly, to create a glyph map where the series of each station are overlayed on a map of Australia, both the major (\code{long}, \code{lat}) and minor (\code{month}, \code{tmax}) coordinates defining the bounding box for the plot need to be on the same table.
 For that, spatial variables can be moved to the long form with \code{unfold()}:
@@
 \begin{CodeInput}
-R> tmax <- tmax |> face_temporal() |> unfold(latitude, longitude)
+R> tmax <- tmax |> face_temporal() |> unfold(long, lat)
 \end{CodeInput}
@@
 
-{\centering \includegraphics[width=0.9\linewidth]{../figures/basic-manip}
+{\centering \includegraphics{figures/basic-manip-1}
 

### 2022-06-15 4-1-basic.Rmd
covid -> COVID, data set -> dataset (to "data" if possible)

@@
 
-## Victoria covid data {#covid}
+## Victoria COVID data {#covid}
 
@@
 Here we will use the data to demonstrate how to match sites between these two tables with \pkg{cubble} in the case of a few mismatches.
-The first five rows these data sets are printed below:
+The first five rows the data are printed below:
 
@@
 
-First of all, the stations across the two states need to be subsetted from the original  data set \code{weatherdata::historical_tmax}.
+First of all, the stations across the two states need to be subsetted from the original data  \code{weatherdata::historical_tmax}.
 For that, the country and the station identifiers, stored within the 11 digits of the \code{id} variable entries, need to be retrieved.

### 2022-06-15 jss/cubble-jss.tex
covid -> COVID, data set -> dataset (to "data" if possible)

@@
 
-The dataset \texttt{climate\_flat} is used to illustrate \pkg{cubble} functionality.
+The data \code{climate_flat} is used to illustrate \pkg{cubble} functionality.
 This is a subset from National Oceanic and Atmospheric Administration (NOAA) Global Historical Climatology Network Daily (GHCND) data.
 It contains spatial variables, station ID, latitude, longitude, elevation, station name, World Meteorology Organisation ID, in additional to daily temporal information, maximum and minimum temperature values and precipitation records for year 2020.
-The first five rows of \texttt{climate\_flat} are shown below:
+The first five rows of \code{climate_flat} are shown below:
 
@@
 \code{id} is the variable name to identify each location and there are five unique locations.
-The bounding box is \code{[115.97, -32.94, 133.55, -12.42]} and provides information about the coordinates in the data set.
-The third row shows the name and type of all variables nested in the \code{ts} column.
+The bounding box is \code{[115.97, -32.94, 133.55, -12.42]} and provides information about the coordinates in the data The third row shows the name and type of all variables nested in the \code{ts} column.
 In this example, it includes \code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.
@@
 \hypertarget{covid}{%
-\subsection{Victoria covid data}\label{covid}}
+\subsection{Victoria COVID data}\label{covid}}
 
@@
 
-First of all, the stations across the two states need to be subsetted from the original data set \code{weatherdata::historical_tmax}.
+First of all, the stations across the two states need to be subsetted from the original data \code{weatherdata::historical_tmax}.
 For that, the country and the station identifiers, stored within the 11 digits of the \code{id} variable entries, need to be retrieved.
@@
 A simple \(k\)-means algorithm based on the distance matrix between stations is used to create 20 clusters.
-The dataset \code{station_nested} is a nested cubble with a cluster column indicating the group to which each station belongs.
+The data \code{station_nested} is a nested cubble with a cluster column indicating the group to which each station belongs.
 More advanced clustering algorithms can be used as well, as long as they provide a mapping from each station to a cluster.
@@
 
-The ERA5 dataset \citep{hersbach2020era5} provides hourly estimates across the Earth for atmospheric, land and oceanic climate variables.
+The ERA5 data \citep{hersbach2020era5} provides hourly estimates across the Earth for atmospheric, land and oceanic climate variables.
 The data is available in the NetCDF format from the European Centre for Medium-Range Weather Forecasts (ECMWF) and it can be directly downloaded from \href{https://cds.climate.copernicus.eu/cdsapp\#!/dataset/reanalysis-era5-pressure-levels?tab=overview}{Copernicus Climate Data Store (CDS)} website or programmatically via the R package \pkg{ecmwfr} \citep{ecwmfr}.

### 2022-06-15 4-1-basic.Rmd
section -> Section

@@
 
-A \code{cubble} object can be created from separate spatial and temporal tables in a list, the other arguments (\code{key}, \code{index}, and \code{coords}) are described in section \@ref(create).
+A \code{cubble} object can be created from separate spatial and temporal tables in a list, the other arguments (\code{key}, \code{index}, and \code{coords}) are described in Section \@ref(create).
 The function \code{as_cubble()} will automatically try to match the sites in both tables and show a warning message in case of missing information (mismatch between the sites in the two tables).

### 2022-06-15 4-1-basic.Rmd
section 4.1 revised

--- 2022-06-15 4-1-basic.Rmd
@@
 
-## Victoria covid data {#covid}
+## Victoria COVID spatio-temporal incidence and spread {#covid}
 
-The Victoria State Government in Australia provides daily COVID information about the source and local government area (LGA) of recorded cases.
-This data can be used to visualise COVID spread when combined with map information available from the Australian Bureau of Statistics.
-Here we will use the data to demonstrate how to match sites between these two tables with \pkg{cubble} in the case of a few mismatches.
-The first five rows these data sets are printed below:
+Since the start of the pandemic, the Victoria State Government in Australia has provided daily COVID counts by local government area (LGA).
+This data can be used to visualise COVID incidence and spread spatially, when combined with map polygon data available from the Australian Bureau of Statistics.
+These different sources need to be combined for the analysis, by matching the LGA names.
+Here is how to do this with \pkg{cubble}, including how to handle mismatches arising from different names of the same LGAs in the two tables.
+The COVID data is a `csv` and looks like:
 
@@
 
-```{r echo = TRUE}
+```{r}
 covid |> head(5)
+```
+
+and the spatial polygons are an ESRI shapefile that looks like:
+
+```{r}
 lga |> head(5)
@@
 
-A \code{cubble} object can be created from separate spatial and temporal tables in a list, the other arguments (\code{key}, \code{index}, and \code{coords}) are described in section \@ref(create).
-The function \code{as_cubble()} will automatically try to match the sites in both tables and show a warning message in case of missing information (mismatch between the sites in the two tables).
+The function \code{as_cubble()} is used to create a \code{cubble} object from the two spatial and temporal tables, and requires specifying the arguments (\code{key}, \code{index}, and \code{coords}) (as described in Section \@ref(create)).
+It will automatically try to match the sites in both tables and will show a warning message when there are mismatches, as shown below:
 
@@
 
-In the output above we see that this is an issue here, and we proceed to use \code{output = "unmatch"} to check which locations are concerned.
-Note that \pkg{cubble} will attempt to pair the unmatched sites, these are shown separately from the ones that cannot be paired.
+It can be seen that there are two-way mismatches -- LGAs in the COVID data that don't match with LGAs in the spatial polygon data, and vice versa.
+The mismatches can be identified by using the \code{output = "unmatch"} argument.
+<!--Note that \pkg{cubble} will attempt to pair the unmatched sites, these are shown separately from the ones that cannot be paired.-->
 
@@
 
-This information is helpful when cleaning up the data before creating the cubble again:
+With this information both tables can be fixed, to create the desired cubble, as follows:
 
@@
 The Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world.
-The **cubble** dataset \code{weatherdata::historical_tmax} extracts the maximum temperature for `r nrow(weatherdata::historical_tmax)` Australian stations from the GHCN starting from year `r range(weatherdata::historical_tmax$first_year)[2]` and provides information also on the latitude, longitude and elevation of each of the  stations.
-This **cubble** data  is already cast into a \code{cubble} data object, with \code{id} as the key, \code{date} as the index, and \code{c(long, lat)} as the coordinates.
+The **cubble** example dataset \code{weatherdata::historical_tmax} is a subset containing the records of maximum temperature for `r nrow(weatherdata::historical_tmax)` Australian stations `r range(weatherdata::historical_tmax$first_year)[2]` through `r range(weatherdata::historical_tmax$first_year)[2]`.
+and provides information also on the latitude, longitude and elevation of each of the  stations.
+This **cubble** data  is already cast into a \code{cubble} data object, with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates.
 The goal of this example is to compare the monthly average maximum temperature  between two periods, 1971-1975 and 2016-2020, for stations in Victoria and New South Wales.
@@
 
-```{r echo = TRUE}
+```{r eval = FALSE, echo = TRUE}
 tmax <- weatherdata::historical_tmax |>
@@
 
-and extracting the time periods 1971-1975 and 2016-2020 is a time operation and therefore the nested cubble needs to be switched to the long cubble form by \code{face_temporal()}:
+and extracting the time periods 1971-1975 and 2016-2020 is a time operation and therefore the nested cubble needs to be switched to the long cubble form by \code{stretch()}:
 
-```{r echo = TRUE}
+```{r eval=  FALSE, echo = TRUE}
 tmax <- tmax |>
@@
 
-```{r echo = TRUE}
+```{r eval=  FALSE, echo = TRUE}
 tmax <- tmax |>
@@
 
-```{r echo = TRUE}
+```{r eval = FALSE, echo = TRUE}
 tmax <- tmax |> face_spatial() |> filter(nrow(ts) == 24)
@@
 
-Lastly, to create a glyph map where the series of each station are overlayed on a map of Australia, both the major (\code{long}, \code{lat}) and minor (\code{month}, \code{tmax}) coordinates defining the bounding box for the plot need to be on the same table.
+Lastly, to create a glyph map where the series of each station are overlayed on a map of Australia, both the major (\code{longitude}, \code{latitude}) and minor (\code{month}, \code{tmax}) coordinates defining the bounding box for the plot need to be on the same table.
 For that, spatial variables can be moved to the long form with \code{unfold()}:
 
-```{r echo = TRUE}  
-tmax <- tmax |> face_temporal() |> unfold(long, lat)
+```{r eval = FALSE, echo = TRUE}  
+tmax <- tmax |> face_temporal() |> unfold(latitude, longitude)
 ```
@@
 
-```{r create-inset, eval = FALSE}
-
-tmax |> filter(id == "ASN00048027") |>
-  ggplot(aes(x = month,
-             y = tmax,
-             color = group)) +
-  geom_line(size = 1.5) +
-  scale_color_brewer(palette = "Dark2", guide = "none") +
-  scale_x_continuous(breaks = seq(1, 12, 1), 
-                     labels = c("J", "F", "M", "A", "M", "J", 
-                                "J", "A", "S", "O", "N", "D")) +
-  labs(x = "", y  = "Temp (C)", title = "ASN00048027: Cobar") +
-  theme_bw() +
-  theme(
-    aspect.ratio = 0.3,
-    axis.text = element_text(size = 20),
-    title =  element_text(size = 20)
-  )
-
-#ggsave(filename = here::here("figures/basic-manip-inset.png"), width = 7)
-```
-
-```{r basic-manip, fig.height = 7, fig.width = 10, fig.cap="A glyph map of the average maximum temperature by month for parts of Victoria and New South Wales weather stations in Australia for two five year periods (1971-1975, 2016-2020).
+```{r basic-manip, out.width="90%", fig.cap="A glyph map of the average maximum temperature by month for parts of Victoria and New South Wales weather stations in Australia for two five year periods (1971-1975, 2016-2020).
 An inset (top left) showing detail is the station Cobar (highlighted by black rectangle).
 The mothly trend is similar at all locations (low in the winter, high in the summer), and small increased temperatures, particularly in late summer can be seen at most stations."}
-nsw_vic <- ozmaps::abs_ste |> filter(NAME %in% c("Victoria", "New South Wales"))
-
-p1 <- ggplot() + 
-  geom_sf(data = nsw_vic, fill = "grey95", color = "white") + 
-  geom_glyph(data = tmax, 
-             aes(x_major = long, x_minor = month, 
-                 y_major = lat, y_minor = tmax,
-                 group = interaction(id, group), 
-                 color = group),
-             width = 1, height = 0.5) +
-  scale_color_brewer("", palette = "Dark2") + 
-  theme_bw() + 
-  coord_sf(xlim = c(141, 154), ylim = c(-36, -29)) + 
-  ggthemes::theme_map() +
-  theme(legend.position = "bottom") +
-  labs(x = "Longitude", y = "Latitude")
-
-box_df <- tmax |> face_spatial() |> filter(id == "ASN00048027")
-single <- tibble::tibble(img = here::here("figures/basic-manip-inset.png"))
-
-p1 +
-  geom_rect(data = box_df,
-            aes(xmin = long - 0.6, xmax = long + 0.6,
-                ymin = lat - 0.12, ymax = lat + 0.35),
-            fill = "transparent", color = "black") +
-  ggimg::geom_point_img(
-    data = single, aes(x = 143.7, y = -30, img = img), size = 6)
-
+knitr::include_graphics(here::here("figures/basic-manip.png"))
 ```

### 2022-06-15 jss/cubble-jss.tex
section 4.1 revised

--- 2022-06-15 jss/cubble-jss.tex
@@
 \hypertarget{covid}{%
-\subsection{Victoria covid data}\label{covid}}
-
-The Victoria State Government in Australia provides daily COVID information about the source and local government area (LGA) of recorded cases.
-This data can be used to visualise COVID spread when combined with map information available from the Australian Bureau of Statistics.
-Here we will use the data to demonstrate how to match sites between these two tables with \pkg{cubble} in the case of a few mismatches.
-The first five rows these data sets are printed below:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> covid |> head(5)
-\end{CodeInput}
+\subsection{Victoria COVID spatiotemporal incidence and spread}\label{covid}}
+
+Since the start of the pandemic, the Victoria State Government in Australia has provided daily COVID counts by local government area (LGA).
+This data can be used to visualise COVID incidence and spread spatially, when combined with map polygon data available from the Australian Bureau of Statistics.
+These different sources need to be combined for the analysis, by matching the LGA names.
+Here is how to do this with \pkg{cubble}, including how to handle mismatches arising from different names of same LGAs in the two tables.
+The COVID data is a \texttt{csv} and looks like:
+
+\begin{CodeChunk}
 \begin{CodeOutput}
@@
 \end{CodeOutput}
-\begin{CodeInput}
-R> lga |> head(5)
-\end{CodeInput}
+\end{CodeChunk}
+
+and the spatial polygons are an ESRI shapefile that looks like:
+
+\begin{CodeChunk}
 \begin{CodeOutput}
@@
 
-A \code{cubble} object can be created from separate spatial and temporal tables in a list, the other arguments (\code{key}, \code{index}, and \code{coords}) are described in section \ref{create}.
-The function \code{as_cubble()} will automatically try to match the sites in both tables and show a warning message in case of missing information (mismatch between the sites in the two tables).
+The function \code{as_cubble()} is used to create a \code{cubble} object from the two spatial and temporal tables, and requires specifying the arguments (\code{key}, \code{index}, and \code{coords}) (as described in Section \ref{create}).
+It will automatically try to match the sites in both tables and will show a warning message when there are mismatches, as shown below:
 
@@
 
-In the output above we see that this is an issue here, and we proceed to use \code{output = "unmatch"} to check which locations are concerned.
-Note that \pkg{cubble} will attempt to pair the unmatched sites, these are shown separately from the ones that cannot be paired.
+It can be seen that there are two-way mismatches -- LGAs in the COVID data that don't match with LGAs in the spatial polygon data, and vice versa.
+The mismatches can be identified by using the \code{output = "unmatch"} argument.
 

### 2022-06-15 2-the-cubble-pkg.Rmd

@@
 
-The dataset `climate_flat` is used to illustrate \pkg{cubble} functionality.
+The data \code{climate_flat} is used to illustrate \pkg{cubble} functionality.
 This is a subset from National Oceanic and Atmospheric Administration (NOAA) Global Historical Climatology Network Daily (GHCND) data.
 It contains spatial variables, station ID, latitude, longitude, elevation, station name, World Meteorology Organisation ID, in additional to daily temporal information,  maximum and minimum temperature values and precipitation records for year 2020.
-The first five rows of `climate_flat` are shown below:
+The first five rows of \code{climate_flat} are shown below:
 

### 2022-06-15 4-1-basic.Rmd

--- 2022-06-15 4-1-basic.Rmd
@@
 
-## Victoria COVID data {#covid}
+## Victoria COVID spatio-temporal incidence and spread {#covid}
 
-The Victoria State Government in Australia provides daily COVID information about the source and local government area (LGA) of recorded cases.
-This data can be used to visualise COVID spread when combined with map information available from the Australian Bureau of Statistics.
-Here we will use the data to demonstrate how to match sites between these two tables with \pkg{cubble} in the case of a few mismatches.
-The first five rows the data are printed below:
+Since the start of the pandemic, the Victoria State Government in Australia has provided daily COVID counts by local government area (LGA).
+This data can be used to visualise COVID incidence and spread spatially, when combined with map polygon data available from the Australian Bureau of Statistics.
+These different sources need to be combined for the analysis, by matching the LGA names.
+Here is how to do this with \pkg{cubble}, including how to handle mismatches arising from different names of the same LGAs in the two tables.
+The COVID data is a `csv` and looks like:
 
@@
 
-A \code{cubble} object can be created from separate spatial and temporal tables in a list, the other arguments (\code{key}, \code{index}, and \code{coords}) are described in Section \@ref(create).
-The function \code{as_cubble()} will automatically try to match the sites in both tables and show a warning message in case of missing information (mismatch between the sites in the two tables).
+The function \code{as_cubble()} is used to create a \code{cubble} object from the two spatial and temporal tables, and requires specifying the arguments (\code{key}, \code{index}, and \code{coords}) (as described in Section \@ref(create)).
+It will automatically try to match the sites in both tables and will show a warning message when there are mismatches, as shown below:
 

### 2022-06-15 4-1-basic.Rmd
change maps to have lightgrey fill

@@
 
-```{r}
+```{r echo = TRUE}
 covid |> head(5)
@@
 
-```{r}
+```{r echo = TRUE}
 lga |> head(5)
@@
 The Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world.
-The **cubble** example dataset \code{weatherdata::historical_tmax} is a subset containing the records of maximum temperature for `r nrow(weatherdata::historical_tmax)` Australian stations `r range(weatherdata::historical_tmax$first_year)[2]` through `r range(weatherdata::historical_tmax$first_year)[2]`.
+The data \code{weatherdata::historical_tmax} is a subset containing the records of maximum temperature for `r nrow(weatherdata::historical_tmax)` Australian stations `r range(weatherdata::historical_tmax$first_year)[2]` through `r range(weatherdata::historical_tmax$first_year)[2]`.
 and provides information also on the latitude, longitude and elevation of each of the  stations.
@@
 
-```{r basic-manip, fig.height = 7, fig.width = 10, fig.cap="A glyph map of the average maximum temperature by month for parts of Victoria and New South Wales weather stations in Australia for two five year periods (1971-1975, 2016-2020).
+```{r basic-manip, fig.height = 10, fig.width = 10, fig.cap="A glyph map of the average maximum temperature by month for parts of Victoria and New South Wales weather stations in Australia for two five year periods (1971-1975, 2016-2020).
 An inset (top left) showing detail is the station Cobar (highlighted by black rectangle).
@@
   theme_bw() + 
-  coord_sf(xlim = c(141, 154), ylim = c(-36, -29)) + 
+  coord_sf(xlim = c(141, 154), ylim = c(-39, -28.5)) + 
   ggthemes::theme_map() +

### 2022-06-15 4-3-water-level.Rmd
change maps to have lightgrey fill

@@
 ggplot() + 
-  geom_sf(data = vic_map, fill = "transparent", color = "grey") + 
+  geom_sf(data = vic_map, fill = "grey95", color = "white") + 
   geom_point(data = dplyr::bind_rows(river, climate), 
              aes(x = long, y = lat, color = type)) + 
-  ggplot2::theme_bw() +
-  ggplot2::theme(legend.position = "bottom") +
-  ggplot2::labs(x = "Longitude", y = "Latitude") + 
+  ggthemes::theme_map() +
+  theme(legend.position = "bottom") +
+  labs(x = "Longitude", y = "Latitude") + 
   scale_color_brewer(palette = "Dark2") 
@@
 p1 <-ggplot() + 
-  geom_sf(data = vic_map, fill = "transparent", color = "grey") + 
+  geom_sf(data = vic_map, fill = "grey95", color = "white") + 
   geom_point(data = res, aes(x = long, y = lat, color = type)) +
@@
   scale_color_brewer(palette = "Dark2")  + 
-  ggplot2::theme_bw() +
+  ggthemes::theme_map() +
   ggplot2::theme(legend.position = "bottom") +

### 2022-06-15 jss/cubble-jss.tex
change maps to have lightgrey fill

--- 2022-06-15 jss/cubble-jss.tex
@@
 
-\textcolor{magenta}{A cubble can be created from various common spatio-temporal data formats, including basic R objects like tibble, tsibble, sf}.
-Section \ref{covid} describes converting multiple tables into a cubble object and Section \ref{netcdf} illustrates how to convert a netCDF object.
-
 The function \code{as_cubble()} is used to create a cubble with three arguments: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order (longitude, latitude).
@@
 \end{CodeChunk}
+
+A cubble can be created from various common spatio-temporal data formats, including basic R objects like tibble, tsibble, sf.
+Section \ref{covid} describes converting multiple tables into a cubble object and Section \ref{netcdf} illustrates how to convert a netCDF object.
 
@@
 \hypertarget{covid}{%
-\subsection{Victoria COVID data}\label{covid}}
-
-The Victoria State Government in Australia provides daily COVID information about the source and local government area (LGA) of recorded cases.
-This data can be used to visualise COVID spread when combined with map information available from the Australian Bureau of Statistics.
-Here we will use the data to demonstrate how to match sites between these two tables with \pkg{cubble} in the case of a few mismatches.
-The first five rows the data are printed below:
+\subsection{Victoria COVID spatio-temporal incidence and spread}\label{covid}}
+
+Since the start of the pandemic, the Victoria State Government in Australia has provided daily COVID counts by local government area (LGA).
+This data can be used to visualise COVID incidence and spread spatially, when combined with map polygon data available from the Australian Bureau of Statistics.
+These different sources need to be combined for the analysis, by matching the LGA names.
+Here is how to do this with \pkg{cubble}, including how to handle mismatches arising from different names of the same LGAs in the two tables.
+The COVID data is a \texttt{csv} and looks like:
 
@@
 \end{CodeOutput}
+\end{CodeChunk}
+
+and the spatial polygons are an ESRI shapefile that looks like:
+
+\begin{CodeChunk}
 \begin{CodeInput}
@@
 The Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world.
-The \textbf{cubble} dataset \code{weatherdata::historical_tmax} extracts the maximum temperature for 204 Australian stations from the GHCN starting from year 1969 and provides information also on the latitude, longitude and elevation of each of the stations.
-This \textbf{cubble} data is already cast into a \code{cubble} data object, with \code{id} as the key, \code{date} as the index, and \code{c(long, lat)} as the coordinates.
+The data \code{weatherdata::historical_tmax} is a subset containing the records of maximum temperature for 204 Australian stations 1969 through 1969.
+and provides information also on the latitude, longitude and elevation of each of the stations.
+This \textbf{cubble} data is already cast into a \code{cubble} data object, with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates.
 The goal of this example is to compare the monthly average maximum temperature between two periods, 1971-1975 and 2016-2020, for stations in Victoria and New South Wales.

### 2022-06-16 4-1-basic.Rmd
Update section 4-1

@@
 
-The six examples here are chosen to illustrate creating a cubble for exploring COVID data where id's need to be matched, using spatial transformations to make a glyph map of seasonal temperature changes over years, aggregating information spatially to explore precipitation patterns, matching river level data and weather station records for analysis of water supply, reading netCDF format data to reproduce a climate reanalysis plot, and the workflow to create complex interactive linked plots.
+The six examples here are chosen to illustrate creating a cubble for exploring COVID data where location identities need to be matched, using spatial transformations to make a glyph map of seasonal temperature changes over years, aggregating information spatially to explore precipitation patterns, matching river level data and weather station records for analysis of water supply, reading netCDF format data to reproduce a climate reanalysis plot, and the workflow to create complex interactive linked plots.
 
@@
 The Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world.
-The data \code{weatherdata::historical_tmax} is a subset containing the records of maximum temperature for `r nrow(weatherdata::historical_tmax)` Australian stations `r range(weatherdata::historical_tmax$first_year)[2]` through `r range(weatherdata::historical_tmax$first_year)[2]`.
-and provides information also on the latitude, longitude and elevation of each of the  stations.
-This **cubble** data  is already cast into a \code{cubble} data object, with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates.
+The data \code{weatherdata::historical_tmax} is a subset containing the records of maximum temperature for `r nrow(weatherdata::historical_tmax)` Australian stations `r from range(weatherdata::historical_tmax$first_year)[1]` through `r range(weatherdata::historical_tmax$first_year)[2]` and provides information also on the latitude, longitude and elevation of each of the  stations.
+This data  is already cast into a \code{cubble} data object (please see Appendix X for the code to extract  the data from XX and prepare the data into \code{cubble} form), with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates.
 The goal of this example is to compare the monthly average maximum temperature  between two periods, 1971-1975 and 2016-2020, for stations in Victoria and New South Wales.

### 2022-06-16 4-1-basic.Rmd
Udate legend Fig 6

@@
 ## Australian historical maximum temperature
-
 The Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world.
-The data \code{weatherdata::historical_tmax} is a subset containing the records of maximum temperature for `r nrow(weatherdata::historical_tmax)` Australian stations `r from range(weatherdata::historical_tmax$first_year)[1]` through `r range(weatherdata::historical_tmax$first_year)[2]` and provides information also on the latitude, longitude and elevation of each of the  stations.
+The data \code{weatherdata::historical_tmax} is a subset containing the records of maximum temperature for `r nrow(weatherdata::historical_tmax)` Australian stations from `r range(weatherdata::historical_tmax$first_year)[1]` through `r range(weatherdata::historical_tmax$first_year)[2]` and provides information also on the latitude, longitude and elevation of each of the  stations.
 This data  is already cast into a \code{cubble} data object (please see Appendix X for the code to extract  the data from XX and prepare the data into \code{cubble} form), with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates.
@@
 
-```{r basic-manip, fig.height = 10, fig.width = 10, fig.cap="A glyph map of the average maximum temperature by month for parts of Victoria and New South Wales weather stations in Australia for two five year periods (1971-1975, 2016-2020).
-An inset (top left) showing detail is the station Cobar (highlighted by black rectangle).
-The mothly trend is similar at all locations (low in the winter, high in the summer), and small increased temperatures, particularly in late summer can be seen at most stations."}
+```{r basic-manip, fig.height = 10, fig.width = 10, fig.cap="A glyph map of the monthly maximum average temperature for weather stations in  Victoria and New South Wales for the periods (1971-1975, 2016-2020).
+The corresponding average time series for the Cobar station are display on the top left corner.
+From the glyph map we can observe that the monthly trend is similar for all locations (low in the winter, high in the summer), and small increased temperatures, particularly in late summer can be seen at most stations."}
 nsw_vic <- ozmaps::abs_ste |> filter(NAME %in% c("Victoria", "New South Wales"))

### 2022-06-16 jss/cubble-jss.tex
Udate legend Fig 6

@@
 A cubble can be created from various common spatio-temporal data formats, including basic R objects like tibble, tsibble, sf.
-Section \ref{covid} describes converting multiple tables into a cubble object and Section \ref{netcdf} illustrates how to convert a netCDF object.
+Section @ref(covid) describes converting multiple tables into a cubble object and Section @ref(netcdf) illustrates how to convert a netCDF object.
 
@@
 Methods for the \code{sfc} class can be applied in the nested form of the cubble object.
-An illustration is in Section \ref{covid}.
+An illustration is in Section @ref(covid).
 A spatial data object with an \code{s2} vector can also be used to store the spatial information in a \code{cubble}.
@@
 Here the switch is between the station id and a cluster id.
-Once the change is made the data can be stretched into the long form.}\label{fig:illu-hier}
+Once the change is made the data can be stretched into the long form.}(\#fig:illu-hier)
 \end{figure}
@@
 
-The function \code{as_cubble()} is used to create a \code{cubble} object from the two spatial and temporal tables, and requires specifying the arguments (\code{key}, \code{index}, and \code{coords}) (as described in Section \ref{create}).
+The function \code{as_cubble()} is used to create a \code{cubble} object from the two spatial and temporal tables, and requires specifying the arguments (\code{key}, \code{index}, and \code{coords}) (as described in Section @ref(create)).
 It will automatically try to match the sites in both tables and will show a warning message when there are mismatches, as shown below:
@@
 The Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world.
-The data \code{weatherdata::historical_tmax} is a subset containing the records of maximum temperature for 204 Australian stations 1969 through 1969.
-and provides information also on the latitude, longitude and elevation of each of the stations.
-This \textbf{cubble} data is already cast into a \code{cubble} data object, with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates.
+The data \code{weatherdata::historical_tmax} is a subset containing the records of maximum temperature for 204 Australian stations from 1859 through 1969 and provides information also on the latitude, longitude and elevation of each of the stations.
+This data is already cast into a \code{cubble} data object (please see Appendix X for the code to extract the data from XX and prepare the data into \code{cubble} form), with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates.
 The goal of this example is to compare the monthly average maximum temperature between two periods, 1971-1975 and 2016-2020, for stations in Victoria and New South Wales.
@@
 
-\caption[A glyph map of the average maximum temperature by month for parts of Victoria and New South Wales weather stations in Australia for two five year periods (1971-1975, 2016-2020)]{A glyph map of the average maximum temperature by month for parts of Victoria and New South Wales weather stations in Australia for two five year periods (1971-1975, 2016-2020).
-An inset (top left) showing detail is the station Cobar (highlighted by black rectangle).
-The mothly trend is similar at all locations (low in the winter, high in the summer), and small increased temperatures, particularly in late summer can be seen at most stations.}\label{fig:basic-manip}
+\caption[A glyph map of the monthly maximum average temperature for weather stations in  Victoria and New South Wales for the periods (1971-1975, 2016-2020)]{A glyph map of the monthly maximum average temperature for weather stations in  Victoria and New South Wales for the periods (1971-1975, 2016-2020).
+The corresponding average time series for the Cobar station are display on the top left corner.
+From the glyph map we can observe that the monthly trend is similar for all locations (low in the winter, high in the summer), and small increased temperatures, particularly in late summer can be seen at most stations.}(\#fig:basic-manip)
 \end{figure}
@@
 
-As introduced in Section \ref{matching}, \code{match_sites()} can be used for this task, and we choose \code{Water_course_level} in \code{river} and \code{prcp} in \code{climate} for the temporal matching.
+As introduced in Section @ref(matching), \code{match_sites()} can be used for this task, and we choose \code{Water_course_level} in \code{river} and \code{prcp} in \code{climate} for the temporal matching.
 We first pass in the two datasets, followed by the variables used for the temporal matching, specified in \code{temporal_by} (using the \code{by} syntax from \code{dplyr::join}).
@@
 The water level reflects the increase in precipitation.
-The numbers (1, 5, 6, 10) indicate the group index derived from spatial matching, only those that were selectd by temporal matching are shown here.}\label{fig:matching}
+The numbers (1, 5, 6, 10) indicate the group index derived from spatial matching, only those that were selectd by temporal matching are shown here.}(\#fig:matching)
 \end{figure}
@@
 The bottom row first selects the lowest temperature in August in the seasonal display, which highlights the corresponding station on the map (Thredbo AWS).
-Another  station, located in the Tasmania Island, is then selected to compare its temperature variation with Thredbo AWS.}\label{fig:interactive-linking}
+Another  station, located in the Tasmania Island, is then selected to compare its temperature variation with Thredbo AWS.}(\#fig:interactive-linking)
 \end{figure}
@@
 
-\caption[Same as Figure 11 except the temperature variation is now shown as a popup in the leaflet map]{Same as Figure 11 except the temperature variation is now shown as a popup in the leaflet map.}\label{fig:interactive-popup}
+\caption[Same as Figure 11 except the temperature variation is now shown as a popup in the leaflet map]{Same as Figure 11 except the temperature variation is now shown as a popup in the leaflet map.}(\#fig:interactive-popup)
 \end{figure}

### 2022-06-16 jss/cubble-jss.tex
remove .aux and other knit associated files

--- 2022-06-15 jss/cubble-jss.tex
@@
 
-\caption[A glyph map of the average maximum temperature by month for parts of Victoria and New South Wales weather stations in Australia for two five year periods (1971-1975, 2016-2020)]{A glyph map of the average maximum temperature by month for parts of Victoria and New South Wales weather stations in Australia for two five year periods (1971-1975, 2016-2020).
-An inset (top left) showing detail is the station Cobar (highlighted by black rectangle).
-The mothly trend is similar at all locations (low in the winter, high in the summer), and small increased temperatures, particularly in late summer can be seen at most stations.}\label{fig:basic-manip}
+\caption[A glyph map of the monthly maximum average temperature for weather stations in  Victoria and New South Wales for the periods (1971-1975, 2016-2020)]{A glyph map of the monthly maximum average temperature for weather stations in  Victoria and New South Wales for the periods (1971-1975, 2016-2020).
+The corresponding average time series for the Cobar station are display on the top left corner.
+From the glyph map we can observe that the monthly trend is similar for all locations (low in the winter, high in the summer), and small increased temperatures, particularly in late summer can be seen at most stations.}\label{fig:basic-manip}
 \end{figure}

### 2022-06-16 jss/appendix.Rmd
appendix with script to create historical_tmax

@@
 ---
-title: "appendix to cubble: An R Package for Organizing and Wrangling Multivariate Spatio-temporal Data"
+title: "Appendix to cubble: An R Package for Organizing and Wrangling Multivariate Spatio-temporal Data"
 author: "Sherry Zhang"
 date: '2022-06-14'
-output: pdf_document
+output: 
+  pdf_document:
+    number_section: true
 ---
@@
 
-# Appendix
+# Linking pipeline
 
@@
 
-\newpage
+# Script to obtain example data
+
+The script below presents the codes required to obtain the data, `historical_tmax`, used in the example of Section 5.2 *Australian historical maximum temperature*.
+The function `rnoaa::meteo_pull_monitors()` may take a while to query a large number of stations in the first time.
+A copy of the data is provided in the data folder of the paper repository at: https://github.com/huizezhang-sherry/paper-cubble.
+
+```{r eval = FALSE}
+library(tidyverse)
+library(cubble)
+all_stations <- rnoaa::ghcnd_stations() |>
+  filter(str_starts(id, "ASN")) |> # Australian stations start wiht "ASN"
+  filter(last_year >= 2020) |>
+  mutate(wmo_id = as.numeric(wmo_id), name = str_to_lower(name)) |>
+  select(-state, -gsn_flag) |>
+  select(id, longitude, latitude, elevation, name,
+         wmo_id, element, first_year, last_year) |>
+  rename(long = longitude, lat = latitude, elev = elevation) |> 
+  filter(element == "TMAX", first_year < 1970, !is.na(wmo_id))
+
+raw_tmax <- all_stations |>
+  rowwise() |>
+  mutate(ts = list(rnoaa::meteo_pull_monitors(
+    monitors = id, var = "TMAX",
+    date_min = glue::glue("{first_year}-01-01"),
+    date_max = glue::glue("{last_year}-12-31")
+    ) |>
+      select(-id)
+    )
+  )
+
+historical_tmax <- raw_tmax |>
+  select(-element) |>
+  unnest(ts) |>
+  mutate(tmax = tmax/10) |>
+  filter(lubridate::year(date) %in% c(1971: 1975, 2016:2020)) |> 
+  as_cubble(index = date, key = id, coords = c(long, lat))
+
+save(historical_tmax, file = here::here("data/historical_tmax.rda"))
+```

### 2022-06-16 jss/cubble-jss.tex
edit in the main text accordingly

@@
 \begin{CodeInput}
-R> cubble_nested <- climate_flat |>
+R> cubble_nested <- climate_flat %>%
 +   as_cubble(key = id, index = date, coords = c(long, lat))
@@
 \begin{CodeInput}
-R> cubble_nested |>
+R> cubble_nested %>%
 +   mutate(rain_day = sum(ts$prcp != 0))
@@
 \begin{CodeInput}
-R> cubble_long <- cubble_nested |> face_temporal()
+R> cubble_long <- cubble_nested %>% face_temporal()
 R> cubble_long
@@
 \begin{CodeInput}
-R> cubble_back <- cubble_long |> face_spatial()
+R> cubble_back <- cubble_long %>% face_spatial()
 R> cubble_back
@@
 \begin{CodeInput}
-R> cubble_unfold <- cubble_long |> unfold(long, lat)
+R> cubble_unfold <- cubble_long %>% unfold(long, lat)
 R> cubble_unfold
@@
 \begin{CodeInput}
-R> cubble_unfold |> face_spatial() |> face_temporal()
+R> cubble_unfold %>% face_spatial() %>% face_temporal()
 \end{CodeInput}
@@
 \begin{CodeInput}
-R> climate_flat_ts <- climate_flat |> 
+R> climate_flat_ts <- climate_flat %>% 
 +   tsibble::as_tsibble(key = id, index = date)
-R> climate_flat_cb <-  climate_flat_ts |> 
+R> climate_flat_cb <-  climate_flat_ts %>% 
 +   cubble::as_cubble(coords = c(long, lat))
@@
 R> # add station-based features in the nested form.
-R> climate_flat_cb |> 
+R> climate_flat_cb %>% 
 +   mutate(
@@
 \begin{CodeInput}
-R> covid |> head(5)
+R> covid %>% head(5)
 \end{CodeInput}
@@
 \begin{CodeInput}
-R> lga |> head(5)
+R> lga %>% head(5)
 \end{CodeInput}
@@
 The Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world.
-The data \code{weatherdata::historical_tmax} is a subset containing the records of maximum temperature for 204 Australian stations from 1859 through 1969 and provides information also on the latitude, longitude and elevation of each of the stations.
-This data is already cast into a \code{cubble} data object (please see Appendix X for the code to extract the data from XX and prepare the data into \code{cubble} form), with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates.
+The data \code{historical_tmax} is a subset containing the records of maximum temperature for 204 Australian stations from 1859 through 1969 and provides information also on the latitude, longitude and elevation of each of the stations.
+This data is already cast into a \code{cubble} data object (please see Appendix X for the code to extract the data from XX and prepare the data into \code{cubble} form), with \code{id} as the key, \code{date} as the index, and \code{c(long, lat)} as the coordinates.
 The goal of this example is to compare the monthly average maximum temperature between two periods, 1971-1975 and 2016-2020, for stations in Victoria and New South Wales.
 
-First of all, the stations across the two states need to be subsetted from the original data \code{weatherdata::historical_tmax}.
+First of all, the stations across the two states need to be subsetted from the original data \code{historical_tmax}.
 For that, the country and the station identifiers, stored within the 11 digits of the \code{id} variable entries, need to be retrieved.
@@
 \begin{CodeInput}
-R> tmax <- weatherdata::historical_tmax |>
+R> tmax <- historical_tmax %>%
 +   filter(between(stringr::str_sub(id, 7, 8), 46, 90))
@@
 
-and extracting the time periods 1971-1975 and 2016-2020 is a time operation and therefore the nested cubble needs to be switched to the long cubble form by \code{face_temporal()}:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> tmax <- tmax |>
-+   face_temporal() |>
-+   filter(lubridate::year(date) %in% c(1971:1975, 2016:2020))
-\end{CodeInput}
-\end{CodeChunk}
-
-A monthly maximum average temperature is then calculated for both periods where \code{cubble} easily allows to compute this time operation across all the stations:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> tmax <- tmax |>
-+   group_by(month = lubridate::month(date),
+A monthly maximum average temperature is then calculated for both periods where \code{cubble} easily allows to compute this time operation across all the stations.
+This is a time operation and therefore the nested cubble needs to be first switched to the long cubble form by \code{face_temporal()}:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> tmax <- tmax %>%
++   face_temporal() %>% 
++   group_by(
++     month = lubridate::month(date),
 +     group = as.factor(ifelse(lubridate::year(date) > 2015,
-+                                   "2016 ~ 2020", "1971 ~ 1975"))) |>
++                              "2016 ~ 2020", "1971 ~ 1975"))) %>%
 +   summarise(tmax = mean(tmax, na.rm = TRUE))
@@
 \begin{CodeInput}
-R> tmax <- tmax |> face_spatial() |> filter(nrow(ts) == 24)
+R> tmax <- tmax %>% face_spatial() %>% filter(nrow(ts) == 24)
 \end{CodeInput}
@@
 \begin{CodeInput}
-R> tmax <- tmax |> face_temporal() |> unfold(long, lat)
+R> tmax <- tmax %>% face_temporal() %>% unfold(long, lat)
 \end{CodeInput}
@@
 \begin{CodeInput}
-R> nsw_vic <- ozmaps::abs_ste |>
+R> nsw_vic <- ozmaps::abs_ste %>%
 +   filter(NAME %in% c("Victoria", "New South Wales"))
@@
 +   geom_glyph(data = tmax,
-+              aes(x_major = longitude, x_minor = month,
-+                  y_major = latitude, y_minor = tmax,
++              aes(x_major = long, x_minor = month,
++                  y_major = lat, y_minor = tmax,
 +                  group = interaction(id, group), color = group),
@@
 \begin{CodeInput}
-R> station_nested <- weatherdata::climate_full |> mutate(cluster = ...)
+R> station_nested <- weatherdata::climate_full %>% mutate(cluster = ...)
 \end{CodeInput}
@@
 \begin{CodeInput}
-R> cluster_nested <- station_nested |> switch_key(cluster)
+R> cluster_nested <- station_nested %>% switch_key(cluster)
 \end{CodeInput}
@@
 \begin{CodeInput}
-R> cluster_nested <- cluster_nested |> get_centroid()
+R> cluster_nested <- cluster_nested %>% get_centroid()
 \end{CodeInput}
@@
 \CommentTok{\# data pre{-}processing}
-\NormalTok{clean }\OtherTok{\textless{}{-}}\NormalTok{ weatherdata}\SpecialCharTok{::}\NormalTok{climate\_full }\SpecialCharTok{|}\ErrorTok{\textgreater{}}\NormalTok{ ...}
+\NormalTok{clean }\OtherTok{\textless{}{-}}\NormalTok{ weatherdata}\SpecialCharTok{::}\NormalTok{climate\_full }\SpecialCharTok{\%\textgreater{}\%}\NormalTok{ ...}
 
 \CommentTok{\# created SharedData instance for crosstalk}
-\NormalTok{nested }\OtherTok{\textless{}{-}}\NormalTok{ clean }\SpecialCharTok{|}\ErrorTok{\textgreater{}}\NormalTok{ SharedData}\SpecialCharTok{$}\FunctionTok{new}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{id, }\AttributeTok{group =} \StringTok{"cubble"}\NormalTok{)}
-\NormalTok{long }\OtherTok{\textless{}{-}} \FunctionTok{face\_temporal}\NormalTok{(clean) }\SpecialCharTok{|}\ErrorTok{\textgreater{}}\NormalTok{ SharedData}\SpecialCharTok{$}\FunctionTok{new}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{id, }\AttributeTok{group =} \StringTok{"cubble"}\NormalTok{)}
+\NormalTok{nested }\OtherTok{\textless{}{-}}\NormalTok{ clean }\SpecialCharTok{\%\textgreater{}\%}\NormalTok{ SharedData}\SpecialCharTok{$}\FunctionTok{new}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{id, }\AttributeTok{group =} \StringTok{"cubble"}\NormalTok{)}
+\NormalTok{long }\OtherTok{\textless{}{-}} \FunctionTok{face\_temporal}\NormalTok{(clean) }\SpecialCharTok{\%\textgreater{}\%}\NormalTok{ SharedData}\SpecialCharTok{$}\FunctionTok{new}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{id, }\AttributeTok{group =} \StringTok{"cubble"}\NormalTok{)}
 
 \CommentTok{\# create the spatial and temporal view each with a ShareData instance}
-\NormalTok{p1 }\OtherTok{\textless{}{-}}\NormalTok{ nested }\SpecialCharTok{|}\ErrorTok{\textgreater{}}\NormalTok{ ...}
-\NormalTok{p2 }\OtherTok{\textless{}{-}}\NormalTok{ long }\SpecialCharTok{|}\ErrorTok{\textgreater{}}\NormalTok{ ...}
+\NormalTok{p1 }\OtherTok{\textless{}{-}}\NormalTok{ nested }\SpecialCharTok{\%\textgreater{}\%}\NormalTok{ ...}
+\NormalTok{p2 }\OtherTok{\textless{}{-}}\NormalTok{ long }\SpecialCharTok{\%\textgreater{}\%}\NormalTok{ ...}
 
@@
 \CommentTok{\# data pre{-}processing}
-\NormalTok{clean }\OtherTok{\textless{}{-}}\NormalTok{ weatherdata}\SpecialCharTok{::}\NormalTok{climate\_full }\SpecialCharTok{|}\ErrorTok{\textgreater{}}\NormalTok{ ...}
+\NormalTok{clean }\OtherTok{\textless{}{-}}\NormalTok{ weatherdata}\SpecialCharTok{::}\NormalTok{climate\_full }\SpecialCharTok{\%\textgreater{}\%}\NormalTok{ ...}
 
@@
 \NormalTok{p }\OtherTok{\textless{}{-}} \FunctionTok{map}\NormalTok{(}\DecValTok{1}\SpecialCharTok{:}\FunctionTok{length}\NormalTok{(df\_id), }\ControlFlowTok{function}\NormalTok{(i)\{}
-\NormalTok{  dt }\OtherTok{\textless{}{-}}\NormalTok{ clean }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{filter}\NormalTok{(id }\SpecialCharTok{==}\NormalTok{ df\_id[i])}
-  \FunctionTok{ggplot}\NormalTok{(dt) }\SpecialCharTok{|}\ErrorTok{\textgreater{}}\NormalTok{ ...}
+\NormalTok{  dt }\OtherTok{\textless{}{-}}\NormalTok{ clean }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(id }\SpecialCharTok{==}\NormalTok{ df\_id[i])}
+  \FunctionTok{ggplot}\NormalTok{(dt) }\SpecialCharTok{\%\textgreater{}\%}\NormalTok{ ...}
 \NormalTok{\})}
@@
 \NormalTok{nested }\OtherTok{\textless{}{-}} \FunctionTok{face\_spatial}\NormalTok{(clean)}
-\FunctionTok{leaflet}\NormalTok{(nested) }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
-  \FunctionTok{addTiles}\NormalTok{() }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
-  \FunctionTok{addCircleMarkers}\NormalTok{(}\AttributeTok{group =} \StringTok{"a"}\NormalTok{, ...) }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
+\FunctionTok{leaflet}\NormalTok{(nested) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{addTiles}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{addCircleMarkers}\NormalTok{(}\AttributeTok{group =} \StringTok{"a"}\NormalTok{, ...) }\SpecialCharTok{\%\textgreater{}\%}
 \NormalTok{  leafpop}\SpecialCharTok{::}\FunctionTok{addPopupGraphs}\NormalTok{(}\AttributeTok{graph =}\NormalTok{ p, ...)}

### 2022-06-16 4-1-basic.Rmd
use magritte pipe throughout

@@
 ```{r echo = TRUE}
-covid |> head(5)
+covid %>% head(5)
 ```
@@
 ```{r echo = TRUE}
-lga |> head(5)
+lga %>% head(5)
 ```
@@
 ```{r echo = TRUE}
-tmax <- weatherdata::historical_tmax |>
+tmax <- weatherdata::historical_tmax %>%
   filter(between(stringr::str_sub(id, 7, 8), 46, 90))
@@
 ```{r echo = TRUE}
-tmax <- tmax |>
-  face_temporal() |>
+tmax <- tmax %>%
+  face_temporal() %>%
   filter(lubridate::year(date) %in% c(1971:1975, 2016:2020))
@@
 ```{r echo = TRUE}
-tmax <- tmax |>
+tmax <- tmax %>%
   group_by(month = lubridate::month(date),
          group = as.factor(ifelse(lubridate::year(date) > 2015,
-                                  "2016 ~ 2020", "1971 ~ 1975"))) |>
+                                  "2016 ~ 2020", "1971 ~ 1975"))) %>%
   summarise(tmax = mean(tmax, na.rm = TRUE))
@@
 ```{r echo = TRUE}
-tmax <- tmax |> face_spatial() |> filter(nrow(ts) == 24)
+tmax <- tmax %>% face_spatial() %>% filter(nrow(ts) == 24)
 ```
@@
 ```{r echo = TRUE}  
-tmax <- tmax |> face_temporal() |> unfold(long, lat)
+tmax <- tmax %>% face_temporal() %>% unfold(long, lat)
 ```
@@
 ```{r eval = FALSE, echo = TRUE}
-nsw_vic <- ozmaps::abs_ste |>
+nsw_vic <- ozmaps::abs_ste %>%
   filter(NAME %in% c("Victoria", "New South Wales"))
@@
 
-tmax |> filter(id == "ASN00048027") |>
+tmax %>% filter(id == "ASN00048027") %>%
   ggplot(aes(x = month,
@@
 From the glyph map we can observe that the monthly trend is similar for all locations (low in the winter, high in the summer), and small increased temperatures, particularly in late summer can be seen at most stations."}
-nsw_vic <- ozmaps::abs_ste |> filter(NAME %in% c("Victoria", "New South Wales"))
+nsw_vic <- ozmaps::abs_ste %>% filter(NAME %in% c("Victoria", "New South Wales"))
 
@@
 
-box_df <- tmax |> face_spatial() |> filter(id == "ASN00048027")
+box_df <- tmax %>% face_spatial() %>% filter(id == "ASN00048027")
 single <- tibble::tibble(img = here::here("figures/basic-manip-inset.png"))

### 2022-06-16 4-2-agg.Rmd
use magritte pipe throughout

--- 2022-06-14 4-2-agg.Rmd
@@
 A simple $k$-means algorithm based on the distance matrix between stations is used to create 20 clusters.
-The dataset \code{station_nested} is a nested cubble with a cluster column indicating the group to which each station belongs.
+The data \code{station_nested} is a nested cubble with a cluster column indicating the group to which each station belongs.
 More advanced clustering algorithms can be used as well, as long as they provide a mapping from each station to a cluster.
@@
 
-```{r echo = TRUE, eval = FALSE}
-station_nested <- weatherdata::climate_full |>
-  mutate(cluster = ...)
+```{r eval = FALSE, echo = TRUE}
+station_nested <- weatherdata::climate_full %>% mutate(cluster = ...)
 ```
@@
 ```{r echo = TRUE}
-cluster_nested <- station_nested |> switch_key(cluster)
+cluster_nested <- station_nested %>% switch_key(cluster)
 ```
@@
 ```{r eval = FALSE, echo = TRUE}
-cluster_nested <- cluster_nested |> get_centroid()
+cluster_nested <- cluster_nested %>% get_centroid()
 ```
@@
 
-```{r basic-agg, out.width="100%", fig.cap="Profile of aggregated precipitation at 639 weather stations in Australia.
+```{r eval = FALSE}
+coords <- cbind(prcp_aus$long, prcp_aus$lat)
+dist_raw <- geosphere::distm(coords, coords)
+
+station_long <- cubble::prcp_aus %>%
+  face_temporal(ts) %>%
+  mutate(wk = lubridate::week(date)) %>%
+  group_by(wk) %>%
+  summarise(prcp = sum(prcp, na.rm = TRUE)) %>%
+  unfold(cluster)
+
+set.seed(123)
+station_nested <- station_long %>%
+  face_spatial() %>%
+  strip_rowwise() %>%
+  mutate(cluster = kmeans(dist_raw,centers = 20, nstart = 500)$cluster)
+save(station_nested, file = here::here("data/station_nested.rda"))
+```
+
+```{r basic-agg, fig.height = 7, fig.width = 15, fig.cap="Profile of aggregated precipitation at 639 weather stations in Australia.
 Subplot A shows the glyph map of the weekly averaged precipitation of each cluster.
@@
 Subplot B shows the station membership of each cluster."}
-knitr::include_graphics(here::here("figures/basic-agg.png"))
+load(here::here("data/station_nested.rda"))
+cluster_nested <- station_nested %>%
+  switch_key(cluster) %>% 
+  get_centroid()
+
+cluster_long <- cluster_nested %>% 
+  face_temporal() %>% 
+  group_by(wk) %>%
+  summarise(prcp = mean(prcp, na.rm = TRUE)) %>% 
+  unfold(cent_long, cent_lat)
+
+state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)
+
+ggplot_smooth <- cluster_long %>% 
+  ggplot() +
+  geom_smooth(aes(x = wk, y = prcp, group = cluster), span = 0.4)
+
+smoother <- layer_data(ggplot_smooth) %>% 
+  left_join(cluster_long %>% select(cluster, cent_long, cent_lat), by = c("group" = "cluster"))
+
+p1 <- ggplot(data = smoother, 
+             aes(x_minor = x, y_minor = y, x_major = cent_long, y_major = cent_lat)) + 
+  geom_sf(data = state_map, inherit.aes = FALSE,
+          color = "grey80", alpha = 0.4, linetype = 3) + 
+  geom_text(data = cluster_nested, 
+            aes(x = cent_long, y = cent_lat, label = cluster), inherit.aes = FALSE) +
+  geom_glyph(height = 2, width = 4) + 
+  theme_void()
+
+p2 <- ggplot() + 
+  geom_sf(data = state_map, inherit.aes = FALSE,
+          color = "grey80", alpha = 0.4, linetype = 3) + 
+  geom_point(data = station_nested, aes(x = long, y = lat), size = 0.5) +
+  ggforce::geom_mark_hull(data = cluster_nested %>% tidyr::unnest(hull),
+                          expand = 0, radius = 0,
+                          aes(x = long, y = lat, group = cluster)) + 
+  theme_void()
+
+(p1 | p2) + plot_annotation(tag_levels = 'A')
 ```

### 2022-06-16 4-3-water-level.Rmd
use magritte pipe throughout

--- 2022-06-14 4-3-water-level.Rmd
@@
 
-```{r eval = FALSE}
-climate <- weatherdata::climate_full |>
-  filter(between(stringr::str_sub(id, 7, 8), 76, 90)) |>
-  face_temporal() |>
-  filter(lubridate::year(date) == 2020) |>
-  face_spatial() |>
+```{r}
+climate <- weatherdata::climate_full %>%
+  filter(between(stringr::str_sub(id, 7, 8), 76, 90)) %>%
+  face_temporal() %>%
+  filter(lubridate::year(date) == 2020) %>%
+  face_spatial() %>%
   mutate(type = "climate")
 
-river <- river |> mutate(type = "river")
+river <- river %>% mutate(type = "river")
 ```
 
-```{r matching-map, out.width="100%", fig.cap="Location of weather stations and river gauges in Victoria, Australia."}
-knitr::include_graphics(here::here("figures/matching-map.png"))
+```{r matching-map, fig.height = 7, fig.width=10, fig.cap="Location of weather stations and river gauges in Victoria, Australia."}
+vic_map <- rmapshaper::ms_simplify(ozmaps::abs_ste %>% filter(NAME == "Victoria"))  
+ggplot() + 
+  geom_sf(data = vic_map, fill = "grey95", color = "white") + 
+  geom_point(data = dplyr::bind_rows(river, climate), 
+             aes(x = long, y = lat, color = type)) + 
+  ggthemes::theme_map() +
+  theme(legend.position = "bottom") +
+  labs(x = "Longitude", y = "Latitude") + 
+  scale_color_brewer(palette = "Dark2") 
 ```
@@
 
-```{r echo = TRUE,  eval = FALSE}
+```{r echo = TRUE}
 res <- match_sites(
@@
 
-```{r,  eval = FALSE}
-res |> print(n = 8)
+```{r}
+res %>% print()
 ```
@@
 The numbers (1, 5, 6, 10) indicate the group index derived from spatial matching, only those that were selectd by temporal matching are shown here."}
-knitr::include_graphics(here::here("figures/matching.png"))
+p1 <-ggplot() + 
+  geom_sf(data = vic_map, fill = "grey95", color = "white") + 
+  geom_point(data = res, aes(x = long, y = lat, color = type)) +
+  ggrepel::geom_label_repel(
+    data = res %>% filter(type == "river"), aes(x = long, y = lat, label = group)) +
+  scale_color_brewer(palette = "Dark2")  + 
+  ggthemes::theme_map() +
+  ggplot2::theme(legend.position = "bottom") +
+  ggplot2::labs(x = "Longitude", y = "Latitude")
+
+res_long <- res %>%
+  face_temporal(ts) %>%
+  unfold(group, type) %>%
+  rename(prcp = matched_var) %>% 
+  mutate(prcp = (prcp - min(prcp, na.rm = TRUE))/ (max(prcp, na.rm = TRUE) - min(prcp, na.rm = TRUE)))
+
+p2 <- res_long %>% 
+  ggplot(aes(x = date, y = prcp, color = type, group = id)) +
+  geom_line(alpha = 0.8) +
+  facet_wrap(vars(group)) +
+  scale_color_brewer(palette = "Dark2", guide = "none") +
+  theme_bw() +
+  scale_x_date(date_labels = "%b") + 
+  labs(x = "Date", y = "Precipitation/ water level")
+
+(p1 | p2) + 
+  patchwork::plot_layout(guides = "collect") + 
+  plot_annotation(tag_levels = 'A')&
+  theme(legend.position = "bottom")
 ```

### 2022-06-16 4-5-interactive.Rmd
use magritte pipe throughout

@@
 # data pre-processing
-clean <- weatherdata::climate_full |> ...
+clean <- weatherdata::climate_full %>% ...
 
 # created SharedData instance for crosstalk
-nested <- clean |> SharedData$new(~id, group = "cubble")
-long <- face_temporal(clean) |> SharedData$new(~id, group = "cubble")
+nested <- clean %>% SharedData$new(~id, group = "cubble")
+long <- face_temporal(clean) %>% SharedData$new(~id, group = "cubble")
 
 # create the spatial and temporal view each with a ShareData instance
-p1 <- nested |> ...
-p2 <- long |> ...
+p1 <- nested %>% ...
+p2 <- long %>% ...
 
@@
 # data pre-processing
-clean <- weatherdata::climate_full |> ...
+clean <- weatherdata::climate_full %>% ...
 
@@
 p <- map(1:length(df_id), function(i){
-  dt <- clean |> filter(id == df_id[i])
-  ggplot(dt) |> ...
+  dt <- clean %>% filter(id == df_id[i])
+  ggplot(dt) %>% ...
 })

### 2022-06-16 jss/appendix.Rmd
use magritte pipe throughout

--- 2022-06-14 jss/appendix.Rmd
@@
 
-\newpage
+# Script to obtain example data
+
+The script below presents the codes required to obtain the data, `historical_tmax`, used in the example of Section 5.2 *Australian historical maximum temperature*.
+The function `rnoaa::meteo_pull_monitors()` may take a while to query a large number of stations in the first time.
+A copy of the data is provided in the data folder of the paper repository at: https://github.com/huizezhang-sherry/paper-cubble.
+
+```{r eval = FALSE}
+library(tidyverse)
+library(cubble)
+all_stations <- rnoaa::ghcnd_stations() %>%
+  filter(str_starts(id, "ASN")) %>% # Australian stations start wiht "ASN"
+  filter(last_year >= 2020) %>%
+  mutate(wmo_id = as.numeric(wmo_id), name = str_to_lower(name)) %>%
+  select(-state, -gsn_flag) %>%
+  select(id, longitude, latitude, elevation, name,
+         wmo_id, element, first_year, last_year) %>%
+  rename(long = longitude, lat = latitude, elev = elevation) %>% 
+  filter(element == "TMAX", first_year < 1970, !is.na(wmo_id))
+
+raw_tmax <- all_stations %>%
+  rowwise() %>%
+  mutate(ts = list(rnoaa::meteo_pull_monitors(
+    monitors = id, var = "TMAX",
+    date_min = glue::glue("{first_year}-01-01"),
+    date_max = glue::glue("{last_year}-12-31")
+    ) %>%
+      select(-id)
+    )
+  )
+
+historical_tmax <- raw_tmax %>%
+  select(-element) %>%
+  unnest(ts) %>%
+  mutate(tmax = tmax/10) %>%
+  filter(lubridate::year(date) %in% c(1971: 1975, 2016:2020)) %>% 
+  as_cubble(index = date, key = id, coords = c(long, lat))
+
+save(historical_tmax, file = here::here("data/historical_tmax.rda"))
+```

### 2022-06-16 jss/cubble-jss.tex
use magritte pipe throughout

--- 2022-06-16 jss/cubble-jss.tex
@@
 \begin{CodeInput}
-R> tmax <- weatherdata::historical_tmax |>
+R> tmax <- weatherdata::historical_tmax %>%
 +   filter(between(stringr::str_sub(id, 7, 8), 46, 90))
@@
 \begin{CodeInput}
-R> tmax <- tmax |>
-+   face_temporal() |>
+R> tmax <- tmax %>%
++   face_temporal() %>%
 +   filter(lubridate::year(date) %in% c(1971:1975, 2016:2020))
@@
 \begin{CodeInput}
-R> tmax <- tmax |>
+R> tmax <- tmax %>%
 +   group_by(month = lubridate::month(date),
 +          group = as.factor(ifelse(lubridate::year(date) > 2015,
-+                                   "2016 ~ 2020", "1971 ~ 1975"))) |>
++                                   "2016 ~ 2020", "1971 ~ 1975"))) %>%
 +   summarise(tmax = mean(tmax, na.rm = TRUE))

### 2022-06-16 scripts/illu-interactive.R

@@ -4,14 +4,14 @@ library(ggplot2)
 library(tsibble)
 
 set.seed(123)
-dt <- prcp_aus |> slice_sample(n = 50)
+dt <- prcp_aus %>% slice_sample(n = 50)
 
 ###########################
 # from map to ts
 state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)
 plot_map(state_map) +
   geom_point(data = dt, aes(x = long, y = lat)) + 
-  geom_point(data = dt |> filter(id == "ASN00003057"), aes(x = long, y = lat), 
+  geom_point(data = dt %>% filter(id == "ASN00003057"), aes(x = long, y = lat), 
              color = "red", size =5) 
 
 ggsave(filename = "figures/keynotes-figures/illu-highlighted-map.png")
@@ -21,16 +21,16 @@ plot_map(state_map) +
   
 ggsave(filename = "figures/keynotes-figures/illu-default-map.png")
 
-long <- dt |> 
-  face_temporal() |> 
-  mutate(month = lubridate::month(date)) |> 
-  group_by(month) |> 
-  summarise(tmax = mean(tmax, na.rm = TRUE)) |> 
+long <- dt %>% 
+  face_temporal() %>% 
+  mutate(month = lubridate::month(date)) %>% 
+  group_by(month) %>% 
+  summarise(tmax = mean(tmax, na.rm = TRUE)) %>% 
   mutate(dummy_date = as.Date(glue::glue("2021-{month}-01")))
 
-highlight_line <- long |> filter(id == "ASN00003057")
-#highlight_point <- highlight_line |> filter(month == 9)
-long |> 
+highlight_line <- long %>% filter(id == "ASN00003057")
+#highlight_point <- highlight_line %>% filter(month == 9)
+long %>% 
   ggplot(aes(x = dummy_date, y = tmax, group = id)) +
   #geom_point(size = 0.3) + 
   geom_line() + 
@@ -41,7 +41,7 @@ long |>
   
 ggsave(filename = "figures/keynotes-figures/illu-highlighted-long.png", width = 5, height = 3)
 
-long |> 
+long %>% 
   ggplot(aes(x = dummy_date, y = tmax, group = id)) +
   geom_line() + 
   scale_x_date(date_labels = "%b") + 
@@ -52,9 +52,9 @@ ggsave(filename = "figures/keynotes-figures/illu-default-long.png", width = 5, h
 ###########################
 # from ts to map
 
[…]

### 2022-06-16 scripts/illu-matching.R

@@ -10,7 +10,7 @@ dt <- tibble(
       4, rnorm(5, sd = 0.5), 6, rnorm(4, sd = 0.5)
     )
   )
-) |> mutate(val = ifelse(val < 0, 0, val))
+) %>% mutate(val = ifelse(val < 0, 0, val))
 
 circle <- tibble(
   x = c(6, 14, 23, 7, 21, 27),
@@ -21,15 +21,15 @@ circle <- tibble(
 ) 
 
 errorbar <- bind_rows(
-  circle |> filter(id == "A"),
-  circle |> filter(id == "A") |> mutate(id = "a")
-) |>
+  circle %>% filter(id == "A"),
+  circle %>% filter(id == "A") %>% mutate(id = "a")
+) %>%
   mutate(id = factor(id, c("A", "a")))
 
 ggplot() +
   geom_rect(data = errorbar, aes(xmin = x, xmax = xend, ymin = 0, ymax = 10), fill = "grey90") +
   geom_line(data = dt, aes(x = date, y = val, group = id), color = "black") +
-  geom_vline(data = circle |> filter(id == "A"), 
+  geom_vline(data = circle %>% filter(id == "A"), 
              aes(xintercept = x), linetype = "longdash", color = "grey10", lwd = 0.2) +
   geom_point(data = circle, aes(x = x, y = y, color = match), size = 3) +
   facet_wrap(vars(id), nrow = 2) +

### 2022-06-16 scripts/interactive-linking.R

@@ -3,29 +3,29 @@ library(tidyverse)
 library(crosstalk)
 library(plotly)
 
-clean <- weatherdata::climate_full |>
+clean <- weatherdata::climate_full %>%
   filter(!id %in% c("ASN00067033", "ASN00072091", "ASN00059040", 
-                    "ASN00097053", "ASN00040856", "ASN00015548")) |>
-  face_temporal() |> 
-  mutate(month = lubridate::month(date)) |>
-  group_by(month) |>
+                    "ASN00097053", "ASN00040856", "ASN00015548")) %>%
+  face_temporal() %>% 
+  mutate(month = lubridate::month(date)) %>%
+  group_by(month) %>%
   summarise(tmax = mean(tmax, na.rm = TRUE),
             tmin = mean(tmin, na.rm = TRUE),
-            diff = mean(tmax - tmin, na.rm = TRUE)) |> 
-  ungroup(month) |>
-  mutate(dummy_date = as.Date(glue::glue("2021-{month}-01"))) |> 
-  face_spatial() |> 
-  mutate(temp_diff_var = ts |> pull(diff) |> var(na.rm = TRUE))
+            diff = mean(tmax - tmin, na.rm = TRUE)) %>% 
+  ungroup(month) %>%
+  mutate(dummy_date = as.Date(glue::glue("2021-{month}-01"))) %>% 
+  face_spatial() %>% 
+  mutate(temp_diff_var = ts %>% pull(diff) %>% var(na.rm = TRUE))
   
 
-nested <- clean |> 
+nested <- clean %>% 
   SharedData$new(~id, group = "cubble")
 
-long <- clean |> face_temporal() |> unfold(temp_diff_var) |> arrange(temp_diff_var) |> 
+long <- clean %>% face_temporal() %>% unfold(temp_diff_var) %>% arrange(temp_diff_var) %>% 
   SharedData$new(~id, group = "cubble")
 
 state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)
-p1 <- nested |> 
+p1 <- nested %>% 
   ggplot() +
   geom_sf(data = state_map, aes(geometry = geometry), fill = "transparent", color = "grey50") +
   geom_point(aes(x = long, y = lat,color = temp_diff_var, label = name)) +
@@ -38,7 +38,7 @@ p1 <- nested |>
   theme(panel.grid.major = element_blank(),
         legend.position = "none")
 
[…]

### 2022-06-16 scripts/interactive-popup.R

@@ -3,26 +3,26 @@ library(tidyverse)
 library(leaflet)
 library(leafpop)
 
-clean <-  weatherdata::climate_full |>
+clean <-  weatherdata::climate_full %>%
   filter(!id %in% c("ASN00067033", "ASN00072091", "ASN00059040", 
-                    "ASN00097053", "ASN00040856", "ASN00015548")) |>
-  face_temporal() |>
-  mutate(month = lubridate::month(date)) |> 
-  group_by(month) |> 
+                    "ASN00097053", "ASN00040856", "ASN00015548")) %>%
+  face_temporal() %>%
+  mutate(month = lubridate::month(date)) %>% 
+  group_by(month) %>% 
   summarise(tmax = mean(tmax, na.rm = TRUE),
             tmin = mean(tmin, na.rm = TRUE),
-            diff = mean(tmax - tmin, na.rm = TRUE)) |> 
-  ungroup(month) |> 
-  face_spatial() |>
-  mutate(temp_diff_var = ts |> pull(diff) |> var(na.rm = TRUE)) |> 
-  face_temporal() |> 
-  mutate(dummy_date = as.Date(glue::glue("2021-{month}-01"))) |> 
+            diff = mean(tmax - tmin, na.rm = TRUE)) %>% 
+  ungroup(month) %>% 
+  face_spatial() %>%
+  mutate(temp_diff_var = ts %>% pull(diff) %>% var(na.rm = TRUE)) %>% 
+  face_temporal() %>% 
+  mutate(dummy_date = as.Date(glue::glue("2021-{month}-01"))) %>% 
   unfold(name)
   
-df_id <- clean$id |> unique()
+df_id <- clean$id %>% unique()
 p <- purrr::map(1:length(df_id), function(i){
-  dt <- clean |> filter(id == df_id[i])
-  dt |> 
+  dt <- clean %>% filter(id == df_id[i])
+  dt %>% 
     ggplot(aes(x = dummy_date, ymin = tmin, ymax = tmax, group = id)) + 
     geom_ribbon(color = "grey50", fill = "grey50", alpha = 0.5) + 
     theme_bw() + 
@@ -40,9 +40,9 @@ pal <- colorNumeric(
     c2 = 40, l2= 85, c1 = 20, l1 = 30),
   domain = domain)
 
-out <- leaflet(nested) |> 
-  addTiles() |> 
[…]

### 2022-06-16 scripts/netcdf.R

@@ -29,20 +29,20 @@ wf_request(
 raw <- ncdf4::nc_open(here::here("data/era5-pressure.nc"))
 dt <- as_cubble(raw, vars = c("q", "z"))
 
-date <- c("2002-09-22", "2002-09-26", "2002-09-30", "2002-10-04") |> as.Date()
-res <- dt |> 
-  face_temporal() |> 
-  filter(lubridate::date(time) %in% date) |>
-  unfold(long, lat) |> 
+date <- c("2002-09-22", "2002-09-26", "2002-09-30", "2002-10-04") %>% as.Date()
+res <- dt %>% 
+  face_temporal() %>% 
+  filter(lubridate::date(time) %in% date) %>%
+  unfold(long, lat) %>% 
   mutate(q =  q* 10^6)
   
 box = c(xmin = -180, ymin = -90, xmax = 180, ymax = -5)
 sf::sf_use_s2(FALSE)
-country <- rnaturalearth::ne_coastline("small", returnclass = "sf") |> 
-  sf::st_crop(box) |> 
+country <- rnaturalearth::ne_coastline("small", returnclass = "sf") %>% 
+  sf::st_crop(box) %>% 
   sf::st_cast("MULTILINESTRING")
 
-res |> 
+res %>% 
   ggplot() +
   geom_point(aes(x = long, y = lat, color = q)) +
   geom_contour(data = res, aes(x = long, y = lat, z = z),

### 2022-06-16 4-1-basic.Rmd
also do for climate_full

@@
 ## Australian historical maximum temperature
+
+```{r}
+load(here::here("data/historical_tmax.rda"))
+```
+
 The Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world.
-The data \code{weatherdata::historical_tmax} is a subset containing the records of maximum temperature for `r nrow(weatherdata::historical_tmax)` Australian stations from `r range(weatherdata::historical_tmax$first_year)[1]` through `r range(weatherdata::historical_tmax$first_year)[2]` and provides information also on the latitude, longitude and elevation of each of the  stations.
+The data \code{weatherdata::historical_tmax} is a subset containing the records of maximum temperature for `r nrow(historical_tmax)` Australian stations from `r range(historical_tmax$first_year)[1]` through `r range(historical_tmax$first_year)[2]` and provides information also on the latitude, longitude and elevation of each of the  stations.
 This data  is already cast into a \code{cubble} data object (please see Appendix X for the code to extract  the data from XX and prepare the data into \code{cubble} form), with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates.
@@
 
-First of all, the stations across the two states need to be subsetted from the original data  \code{weatherdata::historical_tmax}.
+First of all, the stations across the two states need to be subsetted from the original data  \code{historical_tmax}.
 For that, the country and the station identifiers, stored within the 11 digits of the \code{id} variable entries, need to be retrieved.
@@
 ```{r echo = TRUE}
-tmax <- weatherdata::historical_tmax %>%
+tmax <- historical_tmax %>%
   filter(between(stringr::str_sub(id, 7, 8), 46, 90))
@@
 
-and extracting the time periods 1971-1975 and 2016-2020 is a time operation and therefore the nested cubble needs to be switched to the long cubble form by \code{face_temporal()}:
+A monthly maximum average temperature is then calculated for both periods where \code{cubble} easily allows to compute this time operation across all the stations.
+This is a temporal operation and therefore the nested cubble needs to be switched to the long cubble form by \code{face_temporal()}:
 
@@
   face_temporal() %>% 
-  filter(lubridate::year(date) %in% c(1971:1975, 2016:2020))
-```
-
-A monthly maximum average temperature is then calculated for both periods where \code{cubble} easily allows to compute this time operation across all the stations:
-
-```{r echo = TRUE}
-tmax <- tmax %>%
   group_by(month = lubridate::month(date),
@@
   geom_glyph(data = tmax,
-             aes(x_major = longitude, x_minor = month,
-                 y_major = latitude, y_minor = tmax,
+             aes(x_major = long, x_minor = month,
+                 y_major = lat, y_minor = tmax,
                  group = interaction(id, group), color = group),

### 2022-06-16 4-2-agg.Rmd
also do for climate_full

@@
 
+```{r}
+load(here::here("data/station_nested.rda"))
+load(here::here("data/climate_full.rda"))
+```
+
 In the previous example, some overlapping of the glyphs occurred for a few nearby stations such as the pairs (151E, 34S) and (152E, 33S).
@@
 
-\code{weatherdata::climate_full}, also extracted from the GHCN, records daily precipitation and maximum/minimum temperature for `r nrow(weatherdata::climate_full)` stations in Australia from 2016 to 2020.
+The data \code{climate_full}, also extracted from the GHCN, records daily precipitation and maximum/minimum temperature for `r nrow(climate_full)` stations in Australia from 2016 to 2020.
 A simple $k$-means algorithm based on the distance matrix between stations is used to create 20 clusters.
@@
 
-```{r}
-load(here::here("data/station_nested.rda"))
-```
-
 ```{r eval = FALSE, echo = TRUE}
-station_nested <- weatherdata::climate_full %>% mutate(cluster = ...)
+station_nested <- climate_full %>% mutate(cluster = ...)
 ```

### 2022-06-16 4-3-water-level.Rmd
also do for climate_full

@@
 ```{r}
-climate <- weatherdata::climate_full %>%
+climate <- climate_full %>%
   filter(between(stringr::str_sub(id, 7, 8), 76, 90)) %>%

### 2022-06-16 4-5-interactive.Rmd
also do for climate_full

@@
 Below is an example that describes the process of building an interactive graphic with \pkg{cubble} and \pkg{crosstalk}.
-This example explores the variation of monthly temperature range within the \code{weatherdata::climate_full} data.
+This example explores the variation of monthly temperature range within the \code{climate_full} data.
 
@@
 # data pre-processing
-clean <- weatherdata::climate_full %>% ...
+clean <- climate_full %>% ...
 

### 2022-06-16 jss/appendix.Rmd
also do for climate_full

@@
          wmo_id, element, first_year, last_year) %>%
-  rename(long = longitude, lat = latitude, elev = elevation) %>% 
+  rename(long = longitude, lat = latitude, elev = elevation)
+
+tmax_stations <- all_stations %>% 
   filter(element == "TMAX", first_year < 1970, !is.na(wmo_id))
@@
 ```
+
+The data `climate_full` used in Section 5.3, 5.4, and 5.6 can be obtained in a similar fashion with a change of variable and date parameter in `rnoaa::meteo_pull_monitors()`.
+The full script is provided below and a copy of the data is also available in the data folder of the paper repository linked above.
+
+```{r eval = FALSE}
+# all the Australian stations have all of the three PRCP, TMAX, and TMIN recorded
+aus_stations <- all_stations %>%
+  filter(element %in% c("PRCP", "TMAX", "TMIN")) %>%
+  nest(element: last_year) %>%
+  rowwise() %>%
+  filter(nrow(data) == 3) %>%
+  select(-data)
+
+aus_climate_raw <- aus_stations %>%
+  rowwise() %>%
+  mutate(ts = list(
+    rnoaa::meteo_pull_monitors(
+      monitors = id, var = c("PRCP", "TMAX", "TMIN"),
+      date_min = "2016-01-01", date_max = "2020-12-31"
+      ) %>%
+      select(-id)
+    )
+  )
+
+climate_full <- aus_climate_raw %>%
+  unnest(ts) %>%
+  mutate(tmax = tmax/10, tmin = tmin/10) %>%
+  cubble::as_cubble(key = id, index = date, coords = c(long, lat))
+
+save(climate_full, file = here::here("data/climate_full.rda"))
+```

### 2022-06-16 jss/cubble-jss.tex
also do for climate_full

@@
 The Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world.
-The data \code{weatherdata::historical_tmax} is a subset containing the records of maximum temperature for 204 Australian stations from 1859 through 1969 and provides information also on the latitude, longitude and elevation of each of the stations.
+The data \code{weatherdata::historical_tmax} is a subset containing the records of maximum temperature for 236 Australian stations from 1859 through 1969 and provides information also on the latitude, longitude and elevation of each of the stations.
 This data is already cast into a \code{cubble} data object (please see Appendix X for the code to extract the data from XX and prepare the data into \code{cubble} form), with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates.
@@
 
-First of all, the stations across the two states need to be subsetted from the original data \code{weatherdata::historical_tmax}.
+First of all, the stations across the two states need to be subsetted from the original data \code{historical_tmax}.
 For that, the country and the station identifiers, stored within the 11 digits of the \code{id} variable entries, need to be retrieved.
@@
 \begin{CodeInput}
-R> tmax <- weatherdata::historical_tmax %>%
+R> tmax <- historical_tmax %>%
 +   filter(between(stringr::str_sub(id, 7, 8), 46, 90))
@@
 +   face_temporal() %>% 
-+   filter(lubridate::year(date) %in% c(1971:1975, 2016:2020))
-\end{CodeInput}
-\end{CodeChunk}
-
-A monthly maximum average temperature is then calculated for both periods where \code{cubble} easily allows to compute this time operation across all the stations:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> tmax <- tmax %>%
 +   group_by(month = lubridate::month(date),
@@
 
-\code{weatherdata::climate_full}, also extracted from the GHCN, records daily precipitation and maximum/minimum temperature for 640 stations in Australia from 2016 to 2020.
+The data \code{climate_full}, also extracted from the GHCN, records daily precipitation and maximum/minimum temperature for 640 stations in Australia from 2016 to 2020.
 A simple \(k\)-means algorithm based on the distance matrix between stations is used to create 20 clusters.
@@
 \begin{CodeInput}
-R> station_nested <- weatherdata::climate_full %>% mutate(cluster = ...)
+R> station_nested <- climate_full %>% mutate(cluster = ...)
 \end{CodeInput}
@@
 \CommentTok{\# data pre{-}processing}
-\NormalTok{clean }\OtherTok{\textless{}{-}}\NormalTok{ weatherdata}\SpecialCharTok{::}\NormalTok{climate\_full }\SpecialCharTok{\%\textgreater{}\%}\NormalTok{ ...}
+\NormalTok{clean }\OtherTok{\textless{}{-}}\NormalTok{ climate\_full }\SpecialCharTok{\%\textgreater{}\%}\NormalTok{ ...}
 

### 2022-06-16 scripts/interactive-linking.R

@@ -3,7 +3,7 @@ library(tidyverse)
 library(crosstalk)
 library(plotly)
 
-clean <- weatherdata::climate_full %>%
+clean <- climate_full %>%
   filter(!id %in% c("ASN00067033", "ASN00072091", "ASN00059040", 
                     "ASN00097053", "ASN00040856", "ASN00015548")) %>%
   face_temporal() %>% 

### 2022-06-16 scripts/interactive-popup.R

@@ -3,7 +3,7 @@ library(tidyverse)
 library(leaflet)
 library(leafpop)
 
-clean <-  weatherdata::climate_full %>%
+clean <-  climate_full %>%
   filter(!id %in% c("ASN00067033", "ASN00072091", "ASN00059040", 
                     "ASN00097053", "ASN00040856", "ASN00015548")) %>%
   face_temporal() %>%

### 2022-06-17 jss/cubble-jss.tex
figure 5 modification and caption

@@
 
-\caption[An illustration of the data model under interactive graphics with cubble]{An illustration of the data model under interactive graphics with cubble.
-The line plot and the map is made separately with the long and nested cubble.
-When a station is selected on the map (left), the corresponding row in the nested cubble will be activated.
-This will link to all the rows with the same id in the long cubble (middle) and update the line plot (right).}\label{fig:illu-interactive}
+\caption[Linking between multiple plots]{Linking between multiple plots.
+The line plots and the map are constructed from shared crosstalk objects (long and nested cubbles).
+When a station is selected on the map (a), the corresponding row in the nested cubble will be activated.
+This will link to all the rows with the same id in the long cubble (b) and update the line plot (c).}(\#fig:illu-interactive)
 \end{figure}
@@
 The corresponding average time series for the Cobar station are display on the top left corner.
-From the glyph map we can observe that the monthly trend is similar for all locations (low in the winter, high in the summer), and small increased temperatures, particularly in late summer can be seen at most stations.}\label{fig:basic-manip}
+From the glyph map we can observe that the monthly trend is similar for all locations (low in the winter, high in the summer), and small increased temperatures, particularly in late summer can be seen at most stations.}(\#fig:basic-manip)
 \end{figure}

### 2022-06-17 2-0-cube.Rmd
sort out  keynote figures

@@
 For each axis, the slicing can be on a single value or a set of values."}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.007.png"))
+knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.001.png"))
 ```

### 2022-06-17 2-the-cubble-pkg.Rmd
move 3.1 first para to the end of the section
sort out  keynote figures

@@
 
-\sherry{A cubble can be created from various common spatio-temporal data formats, including basic R objects like tibble, tsibble, sf}.
-Section \@ref(covid) describes converting multiple tables into a cubble object and Section \@ref(netcdf) illustrates how to convert a netCDF object.
-
 The function \code{as_cubble()} is used to create a cubble with three arguments: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order  (longitude, latitude).
@@
 
+A cubble can be created from various common spatio-temporal data formats, including basic R objects like tibble, tsibble, sf.
+Section \@ref(covid) describes converting multiple tables into a cubble object and Section \@ref(netcdf) illustrates how to convert a netCDF object.
+
 ## Change focus by facing the time-variables
@@
 In the second row, \\code{face\\_spatial} switches a cubble back to the nested form from the long form and shifts focus back to the spatial aspect."}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.008.png"))
+knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.001.png"))
 ```

### 2022-06-17 3-advanced.Rmd
sort out  keynote figures

@@
 Once the change is made the data can be stretched into the long form."}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.003.png"))
+knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.002.png"))
 ```
@@
 This will link to all the rows with the same id in the long cubble (middle) and update the line plot (right)."}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.004.png"))
+knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.003.png"))
 ```

### 2022-06-17 jss/appendix.Rmd
sort out  keynote figures

@@
 
-Figure \ref{fig:illu-interactive-2} illustrates the linking pipeline from a point selection on the time series to the map.
+<!-- Figure \ref{fig:illu-interactive-2} illustrates the linking pipeline from a point selection on the time series to the map. -->
 
@@
 Both plots will be updated with the full line selected and the point highlighted on the map (right)."}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.005.png"))
+knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.004.png"))
 ```
@@
 
-The script below presents the codes required to obtain the data, `historical_tmax`, used in the example of Section 5.2 *Australian historical maximum temperature*.
+## Historical maximum temperature
+
+The script below presents the codes required to obtain the data `historical_tmax` used in the example of Section 5.2 *Australian historical maximum temperature*.
 The function `rnoaa::meteo_pull_monitors()` may take a while to query a large number of stations in the first time.
@@
 
-The data `climate_full` used in Section 5.3, 5.4, and 5.6 can be obtained in a similar fashion with a change of variable and date parameter in `rnoaa::meteo_pull_monitors()`.
+## Australian 2016-2020 climate data
+
+The data `climate_full` used in the example in Section 5.3, 5.4, and 5.6 of the main paper can be obtained in a similar fashion with a slight change on the variable and date parameter in `rnoaa::meteo_pull_monitors()`.
 The full script is provided below and a copy of the data is also available in the data folder of the paper repository linked above.

### 2022-06-17 jss/cubble-jss.tex
sort out  keynote figures

--- 2022-06-16 jss/cubble-jss.tex
@@
 
-{\centering \includegraphics[width=1\linewidth]{../figures/diagram-keynotes/diagram-keynotes.007}
+{\centering \includegraphics[width=1\linewidth]{../figures/diagram-keynotes/diagram-keynotes.001}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{../figures/diagram-keynotes/diagram-keynotes.008}
+{\centering \includegraphics[width=1\linewidth]{../figures/diagram-keynotes/diagram-keynotes.001}
 
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{../figures/diagram-keynotes/diagram-keynotes.003}
+{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{../figures/diagram-keynotes/diagram-keynotes.002}
 
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{../figures/diagram-keynotes/diagram-keynotes.004}
+{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{../figures/diagram-keynotes/diagram-keynotes.003}
 

### 2022-06-17 1-intro.Rmd

@@
 Section \@ref(conclude) discuss the paper contributions and future directions.
-
-```{r illu-input, echo = FALSE, out.width = "100%", fig.cap="An illustration of different spatio-temporal data layouts: (1) a single table for both spatial and temporal variables; (2) separate tables for spatial and temporal variables; and (3) an array data structure that populates variables into a gridded longitude-latitude-time (hyper)cube.", fig.align='center'}
-#knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.001.png"))
-```
 

### 2022-06-17 3-advanced.Rmd

@@
 
-```{r illu-interactive, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "An illustration of the data model under interactive graphics with cubble.
-The line plot and the map is made separately with the long and nested cubble.
-When a station is selected on the map (left), the corresponding row in the nested cubble will be activated.
-This will link to all the rows with the same id in the long cubble (middle) and update the line plot (right)."}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.003.png"))
+```{r illu-interactive, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "Linking between multiple plots.
+The line plots and the map are constructed from shared crosstalk objects (long and nested cubbles).
+When a station is selected on the map (a), the corresponding row in the nested cubble will be activated.
+This will link to all the rows with the same id in the long cubble (b) and update the line plot (c)."}
+knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.004.png"))
 ```

### 2022-06-17 2-the-cubble-pkg.Rmd
add back the first keynote diagram

@@
 In the second row, \\code{face\\_spatial} switches a cubble back to the nested form from the long form and shifts focus back to the spatial aspect."}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.001.png"))
+knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.002.png"))
 ```

### 2022-06-17 3-advanced.Rmd
add back the first keynote diagram

@@
 Once the change is made the data can be stretched into the long form."}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.002.png"))
+knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.003.png"))
 ```

### 2022-06-17 jss/appendix.Rmd
add back the first keynote diagram

@@
 Both plots will be updated with the full line selected and the point highlighted on the map (right)."}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.004.png"))
+knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.005.png"))
 ```

### 2022-06-17 jss/cubble-jss.tex
add back the first keynote diagram

--- 2022-06-17 jss/cubble-jss.tex
@@
 
-{\centering \includegraphics[width=1\linewidth]{../figures/diagram-keynotes/diagram-keynotes.001}
+{\centering \includegraphics[width=1\linewidth]{../figures/diagram-keynotes/diagram-keynotes.002}
 
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{../figures/diagram-keynotes/diagram-keynotes.002}
+{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{../figures/diagram-keynotes/diagram-keynotes.003}
 
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{../figures/diagram-keynotes/diagram-keynotes.003}
-
-}
-
-\caption[An illustration of the data model under interactive graphics with cubble]{An illustration of the data model under interactive graphics with cubble.
-The line plot and the map is made separately with the long and nested cubble.
-When a station is selected on the map (left), the corresponding row in the nested cubble will be activated.
-This will link to all the rows with the same id in the long cubble (middle) and update the line plot (right).}\label{fig:illu-interactive}
+{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{../figures/diagram-keynotes/diagram-keynotes.004}
+
+}
+
+\caption[Linking between multiple plots]{Linking between multiple plots.
+The line plots and the map are constructed from shared crosstalk objects (long and nested cubbles).
+When a station is selected on the map (a), the corresponding row in the nested cubble will be activated.
+This will link to all the rows with the same id in the long cubble (b) and update the line plot (c).}\label{fig:illu-interactive}
 \end{figure}

### 2022-06-17 jss/appendix.Rmd
appendix

@@
 ---
-title: "Appendix to cubble: An R Package for Organizing and Wrangling Multivariate Spatio-temporal Data"
-author: "Sherry Zhang"
+title: "Appendix to 'cubble: An R Package for Organizing and Wrangling Multivariate Spatio-temporal Data'"
+author: "Sherry Zhang, Dianne Cook, Ursula Laa, Nicolas Langrené, Patricia Menéndez"
 date: '2022-06-14'
@@
 
-# Linking pipeline
+This is the supplementary material for the main paper, containing additional information about the data sets used, and another illustration of linking between multiple plots.
+
+# Additional illustration on multiple linked plots
 
@@
 
-```{r illu-interactive-2, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "An illustration of the data model under interactive graphics with cubble.
-When a point on the time series is selected, the corresponding row in the long cubble will be activated.
-This will link to all the rows with the same id in the long cubble and the row in the nested cubble with the same id (middle).
-Both plots will be updated with the full line selected and the point highlighted on the map (right)."}
+This figure is a supplement to Section 4.3 of the main paper, illustrating how linking from the time series plot to the map is achieved.
+
+```{r illu-interactive-2, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "Linking between multiple plots.
+The line plots and the map are constructed from shared crosstalk objects (long and nested cubbles).
+When a point on the time series is selected, the corresponding row in the long cubble will be activated (a).
+This will link to all the rows with the same id in the long cubble and the row in the nested cubble with the same id (b).
+Both plots will be updated with the full line selected and the point highlighted on the map (c)."}
 knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.005.png"))
@@
 
-# Script to obtain example data
+# Scripts for creating the example data
 

### 2022-06-17 4-1-basic.Rmd
weatherdata sentences cleaned up

@@
 The Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world.
-The data \code{weatherdata::historical_tmax} is a subset containing the records of maximum temperature for `r nrow(historical_tmax)` Australian stations from `r range(historical_tmax$first_year)[1]` through `r range(historical_tmax$first_year)[2]` and provides information also on the latitude, longitude and elevation of each of the  stations.
-This data  is already cast into a \code{cubble} data object (please see Appendix X for the code to extract  the data from XX and prepare the data into \code{cubble} form), with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates.
-The goal of this example is to compare the monthly average maximum temperature  between two periods, 1971-1975 and 2016-2020, for stations in Victoria and New South Wales.
+The data \code{historical_tmax} is a subset containing the records of maximum temperature for `r nrow(historical_tmax)` Australian stations from `r range(historical_tmax$first_year)[1]` through `r range(historical_tmax$first_year)[2]` and provides information also on the latitude, longitude and elevation of each of the  stations, extracted using the `rnoaa` package.
+<!-- This data  is already cast into a \code{cubble} data object (please see Appendix X for the code to extract  the data from XX and prepare the data into \code{cubble} form), with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates.--> The goal of this example is to compare the monthly average maximum temperature  between two periods, 1971-1975 and 2016-2020, for stations in Victoria and New South Wales.
 

### 2022-06-17 jss/cubble-jss.tex
appendix
weatherdata sentences cleaned up

@@
 The Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world.
-The data \code{weatherdata::historical_tmax} is a subset containing the records of maximum temperature for 236 Australian stations from 1859 through 1969 and provides information also on the latitude, longitude and elevation of each of the stations.
-This data is already cast into a \code{cubble} data object (please see Appendix X for the code to extract the data from XX and prepare the data into \code{cubble} form), with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates.
+The data \code{historical_tmax} is a subset containing the records of maximum temperature for 236 Australian stations from 1859 through 1969 and provides information also on the latitude, longitude and elevation of each of the stations, extracted using the \texttt{rnoaa} package.
 The goal of this example is to compare the monthly average maximum temperature between two periods, 1971-1975 and 2016-2020, for stations in Victoria and New South Wales.

### 2022-06-17 2-the-cubble-pkg.Rmd
Fix typos section 3

@@
 Printing a cubble provides some information about the data.
-\code{id} is the variable name to identify each location and there are five unique locations.
-The bounding box is \code{[115.97, -32.94, 133.55, -12.42]} and provides information about the coordinates in the data The third row shows the name and type of all variables nested in the \code{ts} column.
+Here \code{id} is the variable name to identify each location and there are five unique locations.
+The bounding box is \code{[115.97, -32.94, 133.55, -12.42]} and provides information about the coordinates in the data.
+The third row shows the name and type of all variables nested in the \code{ts} column.
 In this example, it includes \code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.
@@
 
-A cubble can be created from various common spatio-temporal data formats, including basic R objects like tibble, tsibble, sf.
+A cubble can be created from various common spatio-temporal data formats, including basic R objects like tibble, tsibble and sf.
 Section \@ref(covid) describes converting multiple tables into a cubble object and Section \@ref(netcdf) illustrates how to convert a netCDF object.

### 2022-06-17 jss/cubble-jss.tex
Fix typos section 3

@@
 
-{\centering \includegraphics[width=1\linewidth]{../figures/diagram-keynotes/diagram-keynotes.001}
+{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/diagram-keynotes/diagram-keynotes.001}
 
@@
 
-A cubble can be created from various common spatio-temporal data formats, including basic R objects like tibble, tsibble, sf.
+A cubble can be created from various common spatio-temporal data formats, including basic R objects like tibble, tsibble and sf.
 Section \ref{covid} describes converting multiple tables into a cubble object and Section \ref{netcdf} illustrates how to convert a netCDF object.
@@
 
-{\centering \includegraphics[width=1\linewidth]{../figures/diagram-keynotes/diagram-keynotes.002}
+{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/diagram-keynotes/diagram-keynotes.002}
 

### 2022-06-17 4-1-basic.Rmd
Move example 5.3 to appendix + update appendix

@@
 
-The six examples here are chosen to illustrate creating a cubble for exploring COVID data where location identities need to be matched, using spatial transformations to make a glyph map of seasonal temperature changes over years, aggregating information spatially to explore precipitation patterns, matching river level data and weather station records for analysis of water supply, reading netCDF format data to reproduce a climate reanalysis plot, and the workflow to create complex interactive linked plots.
+The five examples here together with an additional in the Appendix are chosen to illustrate creating a cubble for exploring COVID data where location identities need to be matched, using spatial transformations to make a glyph map of seasonal temperature changes over years, aggregating information spatially to explore precipitation patterns, matching river level data and weather station records for analysis of water supply, reading netCDF format data to reproduce a climate reanalysis plot, and the workflow to create complex interactive linked plots.
 
@@
 
-## Australian historical maximum temperature
+## Australian historical maximum temperature {#historicaltmax}
 
@@
 ```
+It is worth to mentioned that some overlapping of the glyphs occurred for a few nearby stations as it can be seen from Figure  \ref{fig:basic-manip}.
+This is a common issue when mapping multiple stations in a geographical area.
+This issue can be mitigated in some instances by aggregating data to group the data into a larger areas which might be useful to learn about clusters in the data as well as for visualizing those clusters in a glyph map.
+An continuing discussion of Example \@ref(historicaltmax) discussing all those issues and can be found in the Appendix.

### 2022-06-17 4-3-water-level.Rmd
Move example 5.3 to appendix + update appendix

@@
 ```{r}
+load(here::here("data/climate_full.rda"))
+
 climate <- climate_full %>%

### 2022-06-17 jss/appendix.Rmd
Move example 5.3 to appendix + update appendix

@@
 knitr::opts_chunk$set(echo = TRUE)
+library(tidyverse)
+library(cubble)
+library(patchwork)
 ```
 
-This is the supplementary material for the main paper, containing additional information about the data sets used, and another illustration of linking between multiple plots.
+This is the supplementary material for the main paper, containing
+an extended example following example 5.2 to highlight how **cubble** can be used to deal with data that has a hierarchical structure.
+It also describes in detailed the process to create linking plots.
+Furthermore, this appendix contains additional information about the sources of the data sets use in the paper and the necessary code for extracting and preparing the data as used in the main text with the goal to ensure reproducibility.
+
+# Extension of example 5.2: Australian precipitation pattern in 2020 {#precip}
+
+```{r , echo = FALSE}
+load(here::here("data/station_nested.rda"))
+load(here::here("data/climate_full.rda"))
+```
+
+In the previous example, some overlapping of the glyphs occurred for a few nearby stations such as the pairs (151E, 34S) and (152E, 33S).
+This is a problem when mapping more stations at the national level.
+Aggregation can be helpful in grouping series into clusters before visualising the clusters with a glyph map.
+This new example shows how to organise data at both levels with `switch_key()`.
+
+The data `climate_full`, also extracted from the GHCN, records daily precipitation and maximum/minimum temperature for `r nrow(climate_full)` stations in Australia from 2016 to 2020.
+A simple $k$-means algorithm based on the distance matrix between stations is used to create 20 clusters.
+The data `station_nested` is a nested cubble with a cluster column indicating the group to which each station belongs.
+More advanced clustering algorithms can be used as well, as long as they provide a mapping from each station to a cluster.
+
+```{r eval = FALSE, echo = TRUE}
+station_nested <- climate_full %>% mutate(cluster = ...)
+```
+
+To create a group-level cubble, use `switch_key()` with the new key variable, `cluster`:
+
+```{r echo = TRUE}
+cluster_nested <- station_nested %>% switch_key(cluster)
+```
+
+With the group-level cubble, `get_centroid()` is useful to compute the centroid of each cluster, which will be used as the major axis for the glyph map later:
+
+```{r eval = FALSE, echo = TRUE}
+cluster_nested <- cluster_nested %>% get_centroid()
+```
+
+Long form cubble at both levels can be accessed through stretching the nested form.
+With access to both station and cluster-level cubbles, various plots can be made to understand the cluster.
+Figure \ref{fig:basic-agg} shows two example plots that can be made with this data.
+Subplot A is a glyph map made with the cluster level cubble in the long form and subplot B inspects the station membership of each cluster using the station level cubble in the nested form.
+
+```{r eval = FALSE}
+coords <- cbind(prcp_aus$long, prcp_aus$lat)
+dist_raw <- geosphere::distm(coords, coords)
+
+station_long <- cubble::prcp_aus %>%
+  face_temporal(ts) %>%
+  mutate(wk = lubridate::week(date)) %>%
+  group_by(wk) %>%
+  summarise(prcp = sum(prcp, na.rm = TRUE)) %>%
+  unfold(cluster)
+
+set.seed(123)
+station_nested <- station_long %>%
+  face_spatial() %>%
+  strip_rowwise() %>%
+  mutate(cluster = kmeans(dist_raw,centers = 20, nstart = 500)$cluster)
+save(station_nested, file = here::here("data/station_nested.rda"))
+```
+
+```{r basic-agg, fig.height = 7, fig.width = 15, fig.cap="Profile of aggregated precipitation at 639 weather stations in Australia.
+Subplot A shows the glyph map of the weekly averaged precipitation of each cluster.
+The group number is printed in the middle of the y minor axis and can be used as a reference line to read the magnitude.
+Subplot B shows the station membership of each cluster."}
+load(here::here("data/station_nested.rda"))
+cluster_nested <- station_nested %>%
+  switch_key(cluster) %>% 
+  get_centroid()
+
+cluster_long <- cluster_nested %>% 
+  face_temporal() %>% 
+  group_by(wk) %>%
+  summarise(prcp = mean(prcp, na.rm = TRUE)) %>% 
+  unfold(cent_long, cent_lat)
+
+state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)
+
+ggplot_smooth <- cluster_long %>% 
+  ggplot() +
+  geom_smooth(aes(x = wk, y = prcp, group = cluster), span = 0.4)
+
+smoother <- layer_data(ggplot_smooth) %>% 
+  left_join(cluster_long %>% select(cluster, cent_long, cent_lat), by = c("group" = "cluster"))
+
+p1 <- ggplot(data = smoother, 
+             aes(x_minor = x, y_minor = y, x_major = cent_long, y_major = cent_lat)) + 
+  geom_sf(data = state_map, inherit.aes = FALSE,
+          color = "grey80", alpha = 0.4, linetype = 3) + 
+  geom_text(data = cluster_nested, 
+            aes(x = cent_long, y = cent_lat, label = cluster), inherit.aes = FALSE) +
+  geom_glyph(height = 2, width = 4) + 
+  theme_void()
+
+p2 <- ggplot() + 
+  geom_sf(data = state_map, inherit.aes = FALSE,
+          color = "grey80", alpha = 0.4, linetype = 3) + 
+  geom_point(data = station_nested, aes(x = long, y = lat), size = 0.5) +
+  ggforce::geom_mark_hull(data = cluster_nested %>% tidyr::unnest(hull),
+                          expand = 0, radius = 0,
+                          aes(x = long, y = lat, group = cluster)) + 
+  theme_void()
+
+(p1 | p2) + plot_annotation(tag_levels = 'A')
+```
 
@@
 ```
-
-# Scripts for creating the example data
-
-## Historical maximum temperature
-
-The script below presents the codes required to obtain the data `historical_tmax` used in the example of Section 5.2 *Australian historical maximum temperature*.
-The function `rnoaa::meteo_pull_monitors()` may take a while to query a large number of stations in the first time.
-A copy of the data is provided in the data folder of the paper repository at: https://github.com/huizezhang-sherry/paper-cubble.
-
-```{r eval = FALSE}
-library(tidyverse)
-library(cubble)
-all_stations <- rnoaa::ghcnd_stations() %>%
-  filter(str_starts(id, "ASN")) %>% # Australian stations start wiht "ASN"
-  filter(last_year >= 2020) %>%
-  mutate(wmo_id = as.numeric(wmo_id), name = str_to_lower(name)) %>%
-  select(-state, -gsn_flag) %>%
-  select(id, longitude, latitude, elevation, name,
-         wmo_id, element, first_year, last_year) %>%
-  rename(long = longitude, lat = latitude, elev = elevation)
-
-tmax_stations <- all_stations %>% 
-  filter(element == "TMAX", first_year < 1970, !is.na(wmo_id))
-
-raw_tmax <- all_stations %>%
-  rowwise() %>%
-  mutate(ts = list(rnoaa::meteo_pull_monitors(
-    monitors = id, var = "TMAX",
-    date_min = glue::glue("{first_year}-01-01"),
-    date_max = glue::glue("{last_year}-12-31")
-    ) %>%
-      select(-id)
-    )
-  )
-
-historical_tmax <- raw_tmax %>%
-  select(-element) %>%
-  unnest(ts) %>%
-  mutate(tmax = tmax/10) %>%
-  filter(lubridate::year(date) %in% c(1971: 1975, 2016:2020)) %>% 
-  as_cubble(index = date, key = id, coords = c(long, lat))
-
-save(historical_tmax, file = here::here("data/historical_tmax.rda"))
-```
-
-## Australian 2016-2020 climate data
-
-The data `climate_full` used in the example in Section 5.3, 5.4, and 5.6 of the main paper can be obtained in a similar fashion with a slight change on the variable and date parameter in `rnoaa::meteo_pull_monitors()`.
-The full script is provided below and a copy of the data is also available in the data folder of the paper repository linked above.
-
-```{r eval = FALSE}
-# all the Australian stations have all of the three PRCP, TMAX, and TMIN recorded
-aus_stations <- all_stations %>%
-  filter(element %in% c("PRCP", "TMAX", "TMIN")) %>%
-  nest(element: last_year) %>%
-  rowwise() %>%
-  filter(nrow(data) == 3) %>%
-  select(-data)
-
-aus_climate_raw <- aus_stations %>%
-  rowwise() %>%
-  mutate(ts = list(
-    rnoaa::meteo_pull_monitors(
-      monitors = id, var = c("PRCP", "TMAX", "TMIN"),
-      date_min = "2016-01-01", date_max = "2020-12-31"
-      ) %>%
-      select(-id)
-    )
-  )
-
-climate_full <- aus_climate_raw %>%
-  unnest(ts) %>%
-  mutate(tmax = tmax/10, tmin = tmin/10) %>%
-  cubble::as_cubble(key = id, index = date, coords = c(long, lat))
-
-save(climate_full, file = here::here("data/climate_full.rda"))
-```

### 2022-06-17 jss/cubble-jss.tex
Move example 5.3 to appendix + update appendix

@@
 
-\hypertarget{australian-historical-maximum-temperature}{%
-\subsection{Australian historical maximum temperature}\label{australian-historical-maximum-temperature}}
+\hypertarget{historicaltmax}{%
+\subsection{Australian historical maximum temperature}\label{historicaltmax}}
 
@@
 
-\hypertarget{precip}{%
-\subsection{Australian precipitation pattern in 2020}\label{precip}}
-
-In the previous example, some overlapping of the glyphs occurred for a few nearby stations such as the pairs (151E, 34S) and (152E, 33S).
-This is a problem when mapping more stations at the national level.
-Aggregation can be helpful in grouping series into clusters before visualising the clusters with a glyph map.
-This new example shows how to organise data at both levels with \code{switch_key()}.
-
-The data \code{climate_full}, also extracted from the GHCN, records daily precipitation and maximum/minimum temperature for 640 stations in Australia from 2016 to 2020.
-A simple \(k\)-means algorithm based on the distance matrix between stations is used to create 20 clusters.
-The data \code{station_nested} is a nested cubble with a cluster column indicating the group to which each station belongs.
-More advanced clustering algorithms can be used as well, as long as they provide a mapping from each station to a cluster.
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> station_nested <- climate_full %>% mutate(cluster = ...)
-\end{CodeInput}
-\end{CodeChunk}
-
-To create a group-level cubble, use \code{switch_key()} with the new key variable, \code{cluster}:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> cluster_nested <- station_nested %>% switch_key(cluster)
-\end{CodeInput}
-\end{CodeChunk}
-
-With the group-level cubble, \code{get_centroid()} is useful to compute the centroid of each cluster, which will be used as the major axis for the glyph map later:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> cluster_nested <- cluster_nested %>% get_centroid()
-\end{CodeInput}
-\end{CodeChunk}
-
-Long form cubble at both levels can be accessed through stretching the nested form.
-With access to both station and cluster-level cubbles, various plots can be made to understand the cluster.
-Figure \ref{fig:basic-agg} shows two example plots that can be made with this data.
-Subplot A is a glyph map made with the cluster level cubble in the long form and subplot B inspects the station membership of each cluster using the station level cubble in the nested form.
-
-\begin{CodeChunk}
-\begin{figure}
-
-{\centering \includegraphics{figures/basic-agg-1}
-
-}
-
-\caption[Profile of aggregated precipitation at 639 weather stations in Australia]{Profile of aggregated precipitation at 639 weather stations in Australia.
-Subplot A shows the glyph map of the weekly averaged precipitation of each cluster.
-The group number is printed in the middle of the y minor axis and can be used as a reference line to read the magnitude.
-Subplot B shows the station membership of each cluster.}\label{fig:basic-agg}
-\end{figure}
-\end{CodeChunk}
+It is worth to mentioned that some overlapping of the glyphs occurred for a few nearby stations as it can be seen from Figure \ref{fig:basic-manip}.
+This is a common issue when mapping multiple stations in a geographical area.
+This issue can be mitigated in some instances by aggregating data to group the data into a larger areas which might be useful to learn about clusters in the data as well as for visualizing those clusters in a glyph map.
+An continuing discussion of Example \ref{historicaltmax} discussing all those issues and can be found in the Appendix.
 

### 2022-06-17 jss/appendix.Rmd
fix on the display of code in the appendix

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
 library(tidyverse)
@@
 
-```{r eval = FALSE}
+```{r eval = FALSE, echo = FALSE}
 coords <- cbind(prcp_aus$long, prcp_aus$lat)
@@
 
-```{r basic-agg, fig.height = 7, fig.width = 15, fig.cap="Profile of aggregated precipitation at 639 weather stations in Australia.
+```{r basic-agg, fig.height = 7, fig.width = 15, echo = FALSE, fig.cap="Profile of aggregated precipitation at 639 weather stations in Australia.
 Subplot A shows the glyph map of the weekly averaged precipitation of each cluster.

### 2022-06-17 jss/appendix.Rmd
Add scripts for examples

@@
 ```
+
+# Scripts for creating the example data
+
+This section contains the codes for extracting the data for the examples discussed in the main manuscript.
+
+## Historical maximum temperature
+
+The script below presents the codes required to obtain the data `historical_tmax` used in the example of Section 5.2 *Australian historical maximum temperature*.
+The function `rnoaa::meteo_pull_monitors()` may take a while to query a large number of stations in the first time.
+A copy of the data is provided in the data folder of the paper repository at: https://github.com/huizezhang-sherry/paper-cubble.
+
+```{r eval = FALSE}
+library(tidyverse)
+library(cubble)
+all_stations <- rnoaa::ghcnd_stations() %>%
+  filter(str_starts(id, "ASN")) %>% # Australian stations start wiht "ASN"
+  filter(last_year >= 2020) %>%
+  mutate(wmo_id = as.numeric(wmo_id), name = str_to_lower(name)) %>%
+  select(-state, -gsn_flag) %>%
+  select(id, longitude, latitude, elevation, name,
+         wmo_id, element, first_year, last_year) %>%
+  rename(long = longitude, lat = latitude, elev = elevation)
+
+tmax_stations <- all_stations %>% 
+  filter(element == "TMAX", first_year < 1970, !is.na(wmo_id))
+
+raw_tmax <- all_stations %>%
+  rowwise() %>%
+  mutate(ts = list(rnoaa::meteo_pull_monitors(
+    monitors = id, var = "TMAX",
+    date_min = glue::glue("{first_year}-01-01"),
+    date_max = glue::glue("{last_year}-12-31")
+    ) %>%
+      select(-id)
+    )
+  )
+
+historical_tmax <- raw_tmax %>%
+  select(-element) %>%
+  unnest(ts) %>%
+  mutate(tmax = tmax/10) %>%
+  filter(lubridate::year(date) %in% c(1971: 1975, 2016:2020)) %>% 
+  as_cubble(index = date, key = id, coords = c(long, lat))
+
+save(historical_tmax, file = here::here("data/historical_tmax.rda"))
+```
+
+## Australian 2016-2020 climate data
+
+The data `climate_full` used in the example in Section 5.3, 5.4, and 5.6 of the main paper can be obtained in a similar fashion with a slight change on the variable and date parameter in `rnoaa::meteo_pull_monitors()`.
+The full script is provided below and a copy of the data is also available in the data folder of the paper repository linked above.
+
+```{r eval = FALSE}
+# all the Australian stations have all of the three PRCP, TMAX, and TMIN recorded
+aus_stations <- all_stations %>%
+  filter(element %in% c("PRCP", "TMAX", "TMIN")) %>%
+  nest(element: last_year) %>%
+  rowwise() %>%
+  filter(nrow(data) == 3) %>%
+  select(-data)
+
+aus_climate_raw <- aus_stations %>%
+  rowwise() %>%
+  mutate(ts = list(
+    rnoaa::meteo_pull_monitors(
+      monitors = id, var = c("PRCP", "TMAX", "TMIN"),
+      date_min = "2016-01-01", date_max = "2020-12-31"
+      ) %>%
+      select(-id)
+    )
+  )
+
+climate_full <- aus_climate_raw %>%
+  unnest(ts) %>%
+  mutate(tmax = tmax/10, tmin = tmin/10) %>%
+  cubble::as_cubble(key = id, index = date, coords = c(long, lat))
+
+save(climate_full, file = here::here("data/climate_full.rda"))
+```

### 2022-06-19 4-4-netcdf.Rmd
update figure 9 caption

@@
 
-The ERA5 dataset [@hersbach2020era5] provides hourly estimates across the Earth for atmospheric, land and oceanic climate variables.
+The ERA5 data [@hersbach2020era5] provides hourly estimates across the Earth for atmospheric, land and oceanic climate variables.
 The data is available in the NetCDF format from the European Centre for Medium-Range Weather Forecasts (ECMWF) and it can be directly downloaded from [Copernicus Climate Data Store (CDS)](https://cds.climate.copernicus.eu/cdsapp#!/dataset/reanalysis-era5-pressure-levels?tab=overview) website or programmatically via the R package \pkg{ecmwfr} [@ecwmfr].
@@
 Here we specifically look at the information  contained on the variables *specific humidity* measuring the mass of water vapor per kilogram of moist air  and the *geopotential* variable recording the gravitational potential energy of a unit mass, at a particular location, relative to mean sea level.
-The goal of this example is to reproduce the break-up of the southern polar vortex that happened in late September and early October in 2022 in the stratosphere in the Southern Hemisphere.
+The goal of this example is to reproduce the break-up of the southern polar vortex that happened in late September and early October in 2002 in the stratosphere in the Southern Hemisphere.
 Therefore, we filter data corresponding to the Southern Hemisphere and restrict our analysis to four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04.
@@
 
-```{r netcdf, out.width="100%", fig.cap = "A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020)."}
+```{r netcdf, out.width="100%", fig.cap = "A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020) to illustrate the break-up of sourthern polar vortex in late September and early October 2002.
+The polar vortex, signalled by the high speicfic humidity, splits into two on 2002-09-26 and furthers split into four on 2002-10-04."}
 knitr::include_graphics(here::here("figures/netcdf.png"))

### 2022-06-19 jss/cubble-jss.tex
update figure 9 caption

@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/diagram-keynotes/diagram-keynotes.001}
+{\centering \includegraphics[width=1\linewidth]{../figures/diagram-keynotes/diagram-keynotes.001}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/diagram-keynotes/diagram-keynotes.002}
+{\centering \includegraphics[width=1\linewidth]{../figures/diagram-keynotes/diagram-keynotes.002}
 
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/diagram-keynotes/diagram-keynotes.003}
+{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{../figures/diagram-keynotes/diagram-keynotes.003}
 
@@
 Here we specifically look at the information contained on the variables \emph{specific humidity} measuring the mass of water vapor per kilogram of moist air and the \emph{geopotential} variable recording the gravitational potential energy of a unit mass, at a particular location, relative to mean sea level.
-The goal of this example is to reproduce the break-up of the southern polar vortex that happened in late September and early October in 2022 in the stratosphere in the Southern Hemisphere.
+The goal of this example is to reproduce the break-up of the southern polar vortex that happened in late September and early October in 2002 in the stratosphere in the Southern Hemisphere.
 Therefore, we filter data corresponding to the Southern Hemisphere and restrict our analysis to four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04.
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/netcdf}
-
-}
-
-\caption[A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020)]{A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020).}\label{fig:netcdf}
+{\centering \includegraphics[width=1\linewidth]{../figures/netcdf}
+
+}
+
+\caption[A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020) to illustrate the break-up of sourthern polar vortex in late September and early October 2002]{A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020) to illustrate the break-up of sourthern polar vortex in late September and early October 2002.
+The polar vortex, signalled by the high speicfic humidity, splits into two on 2002-09-26 and furthers split into four on 2002-10-04.}\label{fig:netcdf}
 \end{figure}

### 2022-06-19 scripts/netcdf.R

@@ -44,7 +44,9 @@ country <- rnaturalearth::ne_coastline("small", returnclass = "sf") %>%
 
 res %>% 
   ggplot() +
+  # q for specific humidity
   geom_point(aes(x = long, y = lat, color = q)) +
+  # z for geopotential
   geom_contour(data = res, aes(x = long, y = lat, z = z),
                color = "grey20", binwidth = 4000, size = 0.5) +
   geom_sf(data = country , alpha = 0.5, fill = "transparent", color = "lightgreen") +

### 2022-06-19 jss/appendix.Rmd

@@
 output:
-  pdf_document:
+  bookdown::pdf_document2:
+    toc: false
     number_section: true
+
 ---
@@
 an extended example following example 5.2 to highlight how **cubble** can be used to deal with data that has a hierarchical structure.
-It also describes in detailed the process to create linking plots.
-Furthermore, this appendix contains additional information about the sources of the data sets use in the paper and the necessary code for extracting and preparing the data as used in the main text with the goal to ensure reproducibility.
+It also describes the process to create linking plots.
+Furthermore, this appendix contains additional information about the data sources and the necessary code for extracting and preparing the data sets used in the paper with the goal to ensure reproducibility.
 
@@
 
-In the previous example, some overlapping of the glyphs occurred for a few nearby stations such as the pairs (151E, 34S) and (152E, 33S).
-This is a problem when mapping more stations at the national level.
-Aggregation can be helpful in grouping series into clusters before visualising the clusters with a glyph map.
-This new example shows how to organise data at both levels with `switch_key()`.
+In  example 5.2, some overlapping of the glyphs occurred for a number of stations on the right hand side of the map in Figure 6.
+This is a problem when mapping time series or other glyphs corresponding to locations that are geographically closed on the map.
+In some cases it is better to display such information at an aggregated level by grouping the data adequately before exhibiting the information on a figure.
+The example below shows how can  spatio-temporal data be organised at different hierarchical levels so that information can be grouped both temporal and spatially using  `switch_key()`.
+The goal of this example is to highlight how easy is to move across the data hierarchy using **cubble**.
 
@@
 
-The data `climate_full` used in the example in Section 5.3, 5.4, and 5.6 of the main paper can be obtained in a similar fashion with a slight change on the variable and date parameter in `rnoaa::meteo_pull_monitors()`.
-The full script is provided below and a copy of the data is also available in the data folder of the paper repository linked above.
+The data `climate_full`, used in the examples in Sections 5.3, 5.4, and 5.5 of the main paper and in \ref{#precip} here, can be obtained in a similar fashion with a slight change on the selected variable and date parameter in `rnoaa::meteo_pull_monitors()` as shown below.
+The full script is provided below and a copy of the data is also available in the data folder of the paper GitHub repository linked above.
 

### 2022-06-21 4-1-basic.Rmd
still working on 5.2

@@
 
-The five examples here together with an additional in the Appendix are chosen to illustrate creating a cubble for exploring COVID data where location identities need to be matched, using spatial transformations to make a glyph map of seasonal temperature changes over years, aggregating information spatially to explore precipitation patterns, matching river level data and weather station records for analysis of water supply, reading netCDF format data to reproduce a climate reanalysis plot, and the workflow to create complex interactive linked plots.
+The five examples here are chosen to illustrate these aspects of using cubble: creating a cubble from two COVID data tables with the complication of differing location names, using spatial transformations to make a glyph map of seasonal temperature changes over years, aggregating information spatially to explore precipitation patterns, matching river level data and weather station records for analysis of water supply, reading netCDF format data to reproduce a climate reanalysis plot, and the workflow to create complex interactive linked plots.
+(There is an additional example and figures in the Appendix, and more examples in the package vignettes.)
 
@@
 The Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world.
-The data \code{historical_tmax} is a subset containing the records of maximum temperature for `r nrow(historical_tmax)` Australian stations from `r range(historical_tmax$first_year)[1]` through `r range(historical_tmax$first_year)[2]` and provides information also on the latitude, longitude and elevation of each of the  stations, extracted using the `rnoaa` package.
-<!-- This data  is already cast into a \code{cubble} data object (please see Appendix X for the code to extract  the data from XX and prepare the data into \code{cubble} form), with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates.--> The goal of this example is to compare the monthly average maximum temperature  between two periods, 1971-1975 and 2016-2020, for stations in Victoria and New South Wales.
+The data used here (\code{historical_tmax}) is a subset extracted using the `rnoaa` package, containing the records of maximum temperature for `r nrow(historical_tmax)` Australian stations from `r range(historical_tmax$first_year)[1]` through `r range(historical_tmax$first_year)[2]` and provides information also on the latitude, longitude and elevation of each of the  stations.
+<!-- This data  is already cast into a \code{cubble} data object (please see Appendix X for the code to extract  the data from XX and prepare the data into \code{cubble} form), with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates.--> The goal of this example is to compare the monthly average maximum temperature between two periods, 1971-1975 and 2016-2020, for stations in Victoria and New South Wales, using a glyph map.
 
-First of all, the stations across the two states need to be subsetted from the original data  \code{historical_tmax}.
-For that, the country and the station identifiers, stored within the 11 digits of the \code{id} variable entries, need to be retrieved.
-The country code is located in the first 5 digits (Australia is represented by "ASN00") while the next 6 digits encode the station information following the [Australian Bureau of Meteorology (BOM)](http://www.bom.gov.au/climate/cdo/about/site-num.shtml) coding protocols.
-According to those, New South Wales stations correspond to entries in the range 46-75 and the Victorian stations are stored under 76-90.
-Filtering Victoria and New South Wales stations is a spatial operation and hence uses the \code{cubble} nested form:
+First, the stations need to be filtered to only those corresponding to NSW and Victoria, by using the country and the station identifiers, stored within the 11 digits of the \code{id} variable entries.
+The country code is in the first 5 digits (Australia is represented by "ASN00") and the next 6 digits encode the station following the [Australian Bureau of Meteorology (BOM)](http://www.bom.gov.au/climate/cdo/about/site-num.shtml) coding protocols.
+New South Wales stations correspond to entries in the range 46-75 and the Victorian stations to 76-90.
+Filtering Victoria and New South Wales stations is a *spatial operation* and hence uses the \code{cubble} nested form:
 
@@
 
-A monthly maximum average temperature is then calculated for both periods where \code{cubble} easily allows to compute this time operation across all the stations.
-This is a temporal operation and therefore the nested cubble needs to be switched to the long cubble form by \code{face_temporal()}:
+Next, a monthly maximum average temperature is  calculated for both periods.
+This is a *temporal operation* requiring a switch into the long cubble form, and creating a new indicator for the two time periods, before the calculation:
 
@@
   group_by(month = lubridate::month(date),
-         group = as.factor(ifelse(lubridate::year(date) > 2015,
+         group = as.factor(
+           ifelse(lubridate::year(date) > 2015,
            "2016 ~ 2020", "1971 ~ 1975"))) %>%
@@
 
-the variable storing the maximum monthly average temperatures \code{tmax} can then be supplied to \code{geom_glyph()} for the glyph map in Figure \ref{fig:basic-manip} with a station inset on the top left corner:
+The variable \code{tmax} is used by \code{geom_glyph()} to create the glyph map in Figure \ref{fig:basic-manip}. (An insset of one station inset is shown in the top left corner).
 
@@
 
-```{r basic-manip, fig.height = 10, fig.width = 10, fig.cap="A glyph map of the monthly maximum average temperature for weather stations in  Victoria and New South Wales for the periods (1971-1975, 2016-2020).
+```{r basic-manip, out.width="70%", fig.height = 8, fig.width = 8, fig.cap="A glyph map of the monthly maximum average temperature for weather stations in  Victoria and New South Wales for the periods (1971-1975, 2016-2020).
 The corresponding average time series for the Cobar station are display on the top left corner.

### 2022-06-21 jss/cubble-jss.tex
still working on 5.2

@@
 The Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world.
-The data \code{historical_tmax} is a subset containing the records of maximum temperature for 236 Australian stations from 1859 through 1969 and provides information also on the latitude, longitude and elevation of each of the stations, extracted using the \texttt{rnoaa} package.
-The goal of this example is to compare the monthly average maximum temperature between two periods, 1971-1975 and 2016-2020, for stations in Victoria and New South Wales.
-
-First of all, the stations across the two states need to be subsetted from the original data \code{historical_tmax}.
-For that, the country and the station identifiers, stored within the 11 digits of the \code{id} variable entries, need to be retrieved.
-The country code is located in the first 5 digits (Australia is represented by ``ASN00'') while the next 6 digits encode the station information following the \href{http://www.bom.gov.au/climate/cdo/about/site-num.shtml}{Australian Bureau of Meteorology (BOM)} coding protocols.
-According to those, New South Wales stations correspond to entries in the range 46-75 and the Victorian stations are stored under 76-90.
-Filtering Victoria and New South Wales stations is a spatial operation and hence uses the \code{cubble} nested form:
+The data used here (\code{historical_tmax}) is a subset extracted using the \texttt{rnoaa} package, containing the records of maximum temperature for 236 Australian stations from 1859 through 1969 and provides information also on the latitude, longitude and elevation of each of the stations.
+The goal of this example is to compare the monthly average maximum temperature between two periods, 1971-1975 and 2016-2020, for stations in Victoria and New South Wales, using a glyph map.
+
+First, the stations need to be filtered to only those corresponding to NSW and Victoria, by using the country and the station identifiers, stored within the 11 digits of the \code{id} variable entries.
+The country code is in the first 5 digits (Australia is represented by ``ASN00'') and the next 6 digits encode the station following the \href{http://www.bom.gov.au/climate/cdo/about/site-num.shtml}{Australian Bureau of Meteorology (BOM)} coding protocols.
+New South Wales stations correspond to entries in the range 46-75 and the Victorian stations to 76-90.
+Filtering Victoria and New South Wales stations is a \emph{spatial operation} and hence uses the \code{cubble} nested form:
 
@@
 
-A monthly maximum average temperature is then calculated for both periods where \code{cubble} easily allows to compute this time operation across all the stations.
-This is a temporal operation and therefore the nested cubble needs to be switched to the long cubble form by \code{face_temporal()}:
+Next, a monthly maximum average temperature is calculated for both periods.
+This is a \emph{temporal operation} requiring a switch into the long cubble form, and creating a new indicator for the two time periods, before the calculation:
 
@@
 +   group_by(month = lubridate::month(date),
-+          group = as.factor(ifelse(lubridate::year(date) > 2015,
++          group = as.factor(
++            ifelse(lubridate::year(date) > 2015,
 +            "2016 ~ 2020", "1971 ~ 1975"))) %>%
@@
 
-{\centering \includegraphics{figures/basic-manip-1}
+{\centering \includegraphics[width=0.7\linewidth]{figures/basic-manip-1}
 

### 2022-06-21 4-1-basic.Rmd
done with section 5.2

@@
 
-First, the stations need to be filtered to only those corresponding to NSW and Victoria, by using the country and the station identifiers, stored within the 11 digits of the \code{id} variable entries.
+First, the stations need to be filtered to those in NSW and Victoria, by using the station identifiers, stored within the 11 digits of the \code{id} variable entries.
 The country code is in the first 5 digits (Australia is represented by "ASN00") and the next 6 digits encode the station following the [Australian Bureau of Meteorology (BOM)](http://www.bom.gov.au/climate/cdo/about/site-num.shtml) coding protocols.
@@
 
-Observing the results of the operation above show that a few stations do not have records during the time period 1971-1975.
-Further investigation shows that while the first and last year of each series for those stations are recorded, the years in between are not.
-Hence,  these stations are  located and filtered out by examining whether the summarised time series has 24 months.
-Here, the long cubble needs to be switched to the nested form to carry out this spatial operation using \code{face_spatial()}:
+A quick check on the number of observations for each location is made, revealing that there are several with less than 24 -- these lack temperature values for some months.
+These stations are removed, by switching to a long cubble to operate on the spatial component over time, and then back to the nested cubble (to make the glyph map):
 
 ```{r echo = TRUE}
-tmax <- tmax %>% face_spatial() %>% filter(nrow(ts) == 24)
+tmax %>% 
+  face_spatial() %>% 
+  mutate(n = nrow(ts)) %>%
+  arrange(n) %>%
+  pull(n) %>% 
+  head(10)
+
+tmax <- tmax %>% 
+  face_spatial() %>% 
+  filter(nrow(ts) == 24) %>%
+  face_temporal()
 ```
 
-Lastly, to create a glyph map where the series of each station are overlayed on a map of Australia, both the major (\code{long}, \code{lat}) and minor (\code{month}, \code{tmax}) coordinates defining the bounding box for the plot need to be on the same table.
-For that, spatial variables can be moved to the long form with \code{unfold()}:
-
-```{r echo = TRUE}  
-tmax <- tmax %>% face_temporal() %>% unfold(long, lat)
-```
-
-The variable \code{tmax} is used by \code{geom_glyph()} to create the glyph map in Figure \ref{fig:basic-manip}. (An insset of one station inset is shown in the top left corner).
+To create a glyph map of the monthly series overlaid on the map (Figure \ref{fig:basic-manip}), requires that the spatial variables be unfolded with the temporal variables.
+Both of the major (\code{long}, \code{lat}) and minor (\code{month}, \code{tmax}) coordinates need to be on the same table.
+The \code{geom_glyph()} function does both the transformation and the plotting.
+(An inset of the glyph for one station is shown in the top left corner.)
 
@@
   geom_sf(data = nsw_vic,
-          fill = "transparent", color = "grey", linetype = "dotted") +
+          fill = "transparent", color = "grey",
+          linetype = "dotted") +
   geom_glyph(data = tmax,
@@
   geom_line(size = 1.5) +
-  scale_color_brewer(palette = "Dark2", guide = "none") +
+  scale_color_brewer(palette = "Dark2", 
+                     guide = "none") +
   scale_x_continuous(breaks = seq(1, 12, 1), 
-                     labels = c("J", "F", "M", "A", "M", "J", 
-                                "J", "A", "S", "O", "N", "D")) +
-  labs(x = "", y  = "Temp (C)", title = "ASN00048027: Cobar") +
+                     labels = c("J", "F", "M", "A",
+                                "M", "J", "J", "A",
+                                "S", "O", "N", "D")) +
+  labs(x = "", y  = "Temp (C)", 
+       title = "ASN00048027: Cobar") +
   theme_bw() +
@@
 From the glyph map we can observe that the monthly trend is similar for all locations (low in the winter, high in the summer), and small increased temperatures, particularly in late summer can be seen at most stations."}
-nsw_vic <- ozmaps::abs_ste %>% filter(NAME %in% c("Victoria", "New South Wales"))
+tmax <- tmax %>% unfold(long, lat)
+
+nsw_vic <- ozmaps::abs_ste %>% 
+  filter(NAME %in% c("Victoria", "New South Wales"))
 
 p1 <- ggplot() + 
-  geom_sf(data = nsw_vic, fill = "grey95", color = "white") + 
+  geom_sf(data = nsw_vic, fill = "grey95", 
+          color = "white") + 
   geom_glyph(data = tmax, 
@@
   ggthemes::theme_map() +
-  theme(legend.position = "bottom") +
-  labs(x = "Longitude", y = "Latitude")
+  theme(legend.position = "bottom") #+
+  #labs(x = "Longitude", y = "Latitude")
 
-box_df <- tmax %>% face_spatial() %>% filter(id == "ASN00048027")
+box_df <- tmax %>% 
+  face_spatial() %>% 
+  filter(id == "ASN00048027")
 single <- tibble::tibble(img = here::here("figures/basic-manip-inset.png"))
@@
 ```
-It is worth to mentioned that some overlapping of the glyphs occurred for a few nearby stations as it can be seen from Figure  \ref{fig:basic-manip}.
-This is a common issue when mapping multiple stations in a geographical area.
-This issue can be mitigated in some instances by aggregating data to group the data into a larger areas which might be useful to learn about clusters in the data as well as for visualizing those clusters in a glyph map.
-An continuing discussion of Example \@ref(historicaltmax) discussing all those issues and can be found in the Appendix.
+
+Glyph maps work well to explore temporal patterns across spatial locations, particularly when the spatial locations are gridded.
+Here they are irregularly spaced, which can result in overlapping glyphs obscurring each other.
+To fix this one could aggregate data from nearby stations.
+An example is included in the Appendix.

### 2022-06-21 jss/cubble-jss.tex
done with section 5.2

@@
 
-First, the stations need to be filtered to only those corresponding to NSW and Victoria, by using the country and the station identifiers, stored within the 11 digits of the \code{id} variable entries.
+First, the stations need to be filtered to those in NSW and Victoria, by using the station identifiers, stored within the 11 digits of the \code{id} variable entries.
 The country code is in the first 5 digits (Australia is represented by ``ASN00'') and the next 6 digits encode the station following the \href{http://www.bom.gov.au/climate/cdo/about/site-num.shtml}{Australian Bureau of Meteorology (BOM)} coding protocols.
@@
 
-Observing the results of the operation above show that a few stations do not have records during the time period 1971-1975.
-Further investigation shows that while the first and last year of each series for those stations are recorded, the years in between are not.
-Hence, these stations are located and filtered out by examining whether the summarised time series has 24 months.
-Here, the long cubble needs to be switched to the nested form to carry out this spatial operation using \code{face_spatial()}:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> tmax <- tmax %>% face_spatial() %>% filter(nrow(ts) == 24)
-\end{CodeInput}
-\end{CodeChunk}
-
-Lastly, to create a glyph map where the series of each station are overlayed on a map of Australia, both the major (\code{long}, \code{lat}) and minor (\code{month}, \code{tmax}) coordinates defining the bounding box for the plot need to be on the same table.
-For that, spatial variables can be moved to the long form with \code{unfold()}:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> tmax <- tmax %>% face_temporal() %>% unfold(long, lat)
-\end{CodeInput}
-\end{CodeChunk}
-
-The variable \code{tmax} is used by \code{geom_glyph()} to create the glyph map in Figure \ref{fig:basic-manip}. (An insset of one station inset is shown in the top left corner).
+A quick check on the number of observations for each location is made, revealing that there are several with less than 24 -- these lack temperature values for some months.
+These stations are removed, by switching to a long cubble to operate on the spatial component over time, and then back to the nested cubble (to make the glyph map):
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> tmax %>% 
++   face_spatial() %>% 
++   mutate(n = nrow(ts)) %>%
++   arrange(n) %>%
++   pull(n) %>% 
++   head(10)
+\end{CodeInput}
+\begin{CodeOutput}
+ [1] 12 12 12 13 19 19 20 24 24 24
+\end{CodeOutput}
+\begin{CodeInput}
+R> tmax <- tmax %>% 
++   face_spatial() %>% 
++   filter(nrow(ts) == 24) %>%
++   face_temporal()
+\end{CodeInput}
+\end{CodeChunk}
+
+To create a glyph map of the monthly series overlaid on the map (Figure \ref{fig:basic-manip}), requires that the spatial variables be unfolded with the temporal variables.
+Both of the major (\code{long}, \code{lat}) and minor (\code{month}, \code{tmax}) coordinates need to be on the same table.
+The \code{geom_glyph()} function does both the transformation and the plotting.
+(An inset of the glyph for one station is shown in the top left corner.)
 
@@
 +   geom_sf(data = nsw_vic,
-+           fill = "transparent", color = "grey", linetype = "dotted") +
++           fill = "transparent", color = "grey",
++           linetype = "dotted") +
 +   geom_glyph(data = tmax,
@@
 
-It is worth to mentioned that some overlapping of the glyphs occurred for a few nearby stations as it can be seen from Figure \ref{fig:basic-manip}.
-This is a common issue when mapping multiple stations in a geographical area.
-This issue can be mitigated in some instances by aggregating data to group the data into a larger areas which might be useful to learn about clusters in the data as well as for visualizing those clusters in a glyph map.
-An continuing discussion of Example \ref{historicaltmax} discussing all those issues and can be found in the Appendix.
+Glyph maps work well to explore temporal patterns across spatial locations, particularly when the spatial locations are gridded.
+Here they are irregularly spaced, which can result in overlapping glyphs obscurring each other.
+To fix this one could aggregate data from nearby stations.
+An example is included in the Appendix.
 

### 2022-06-21 4-1-basic.Rmd
tweaking plot style of glyph map

@@
 
-```{r basic-manip, out.width="70%", fig.height = 8, fig.width = 8, fig.cap="A glyph map of the monthly maximum average temperature for weather stations in  Victoria and New South Wales for the periods (1971-1975, 2016-2020).
+```{r basic-manip, out.width="50%", fig.height = 6, fig.width = 5, fig.cap="A glyph map of the monthly maximum average temperature for weather stations in  Victoria and New South Wales for the periods (1971-1975, 2016-2020).
 The corresponding average time series for the Cobar station are display on the top left corner.
@@
   geom_sf(data = nsw_vic, fill = "grey95", 
-          color = "white") + 
+          color = "white", size=2) + 
   geom_glyph(data = tmax, 
@@
                  color = group),
-             width = 1, height = 0.5) +
+             width = 1.2, height = 1, alpha=0.5, 
+             size=1.1) +
   scale_color_brewer("", palette = "Dark2") + 

### 2022-06-22 4-3-water-level.Rmd
section 5.3

@@
 
-## River level data in Victoria water gauges
+## River levels and rainfall in Victoria
 
-The Bureau of Meteorology collects [water data](http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full) from river gauges.
+River level and rainfall data for the same areas should have some similarity.
+Here we examine the river gauge data from the Bureau of Meteorology [@REF] in relation to weather station rainfall from NOAA's climate data.
+Figure \ref{fig:matching-map} shows the location of available weather stations and water gauges in Victoria.
+The goal is to match water gauges with nearby weather stations, spatially and temporally, using 2020 measurements.
+
+<!--The Bureau of Meteorology collects [water data](http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full) from river gauges.
 The collected variables include: electrical conductivity, turbidity, watercourse discharge, watercourse level, and water temperature.
 We expect that the water level should be related to the precipitation available from climate data, since rainfall will raise the water level in the river.
-Figure \ref{fig:matching-map} shows the location of available weather stations and water gauges in Victoria, and we will try to match water gauges with nearby weather stations, taking into account both spatial and temporal information from 2020.
+Figure \ref{fig:matching-map} shows the location of available weather stations and water gauges in Victoria, and we will try to match water gauges with nearby weather stations, taking into account both spatial and temporal information from 2020.-->
 
@@
 ```{r matching-map, fig.height = 7, fig.width=10, fig.cap="Location of weather stations and river gauges in Victoria, Australia."}
-vic_map <- rmapshaper::ms_simplify(ozmaps::abs_ste %>% filter(NAME == "Victoria"))  
+vic_map <- rmapshaper::ms_simplify(ozmaps::abs_ste %>%
+               filter(NAME == "Victoria"))
+
 ggplot() + 
@@
 
-As introduced in Section \@ref(matching), \code{match_sites()} can be used for this task, and we choose \code{Water_course_level} in \code{river} and \code{prcp} in \code{climate} for the temporal matching.
-We first pass in the two datasets, followed by the variables used for the temporal matching, specified in \code{temporal_by} (using the \code{by} syntax from \code{dplyr::join}).
+The \code{match_sites()} function is used.
+\code{Water_course_level} from \code{river} and \code{prcp} from \code{climate} is used for the temporal matching.
+The peaks are compared, as explained in Section \@ref(sec:matching).
+A value of 30 is used for the maximum number of peaks to match, which corresponds to about 2 per month.
+<!--We first pass in the two datasets, followed by the variables used for the temporal matching, specified in \code{temporal_by} (using the \code{by} syntax from \code{dplyr::join}).
 With \code{temporal_independent} we select precipitation (\code{prcp}) as the independent variable, and the goal is to see if the water level in the river reflects what is observed in precipitation (for nearby stations).
-Since we consider a full year of data, we set the number of peaks considered (\code{temporal_n_highest}) to 30 (slightly above the default value of 20), and \code{temporal_min_match} is raised accordingly.
+Since we consider a full year of data, we set the number of peaks considered (\code{temporal_n_highest}) to 30 (slightly above the default value of 20), and \code{temporal_min_match} is raised accordingly.-->
 
@@
 
-The output from this function is also a cubble, with additional columns \code{dist} and \code{group} produced from spatial matching, and \code{n_match} from temporal matching.
+This function returns a cubble, with additional columns \code{dist} and \code{group} summarising spatial matching, and \code{n_match} for temporal matching.
 
@@
 
-Figure \ref{fig:matching} (A) shows the four matched pairs on the map.
-For these the expected concurrent increase in precipitation and water level can be seen clearly when comparing these variables on a standardised scale, see Figure \ref{fig:matching} (B).
+Figure \ref{fig:matching} shows four matched pairs on the map (A) and standardised data as time series (B). The expected concurrent increase in precipitation and water level can be seen clearly.
 
-```{r matching, fig.height = 5, fig.width = 10, fig.cap="Matched weather stations and river gauges on the map (A) and across time (B).
+```{r matching, out.width="100%", fig.height = 5, fig.width = 10, fig.cap="Matched weather stations and river gauges on the map (A) and across time (B).
 Precipitation and water level have been standardised between 0 and 1 to be displayed on the same scale.

### 2022-06-22 jss/cubble-jss.tex
tweaking plot style of glyph map
section 5.3

@@
 
-{\centering \includegraphics[width=0.7\linewidth]{figures/basic-manip-1}
+{\centering \includegraphics[width=0.5\linewidth]{figures/basic-manip-1}
 
@@
 
-\hypertarget{river-level-data-in-victoria-water-gauges}{%
-\subsection{River level data in Victoria water gauges}\label{river-level-data-in-victoria-water-gauges}}
-
-The Bureau of Meteorology collects \href{http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full}{water data} from river gauges.
-The collected variables include: electrical conductivity, turbidity, watercourse discharge, watercourse level, and water temperature.
-We expect that the water level should be related to the precipitation available from climate data, since rainfall will raise the water level in the river.
-Figure \ref{fig:matching-map} shows the location of available weather stations and water gauges in Victoria, and we will try to match water gauges with nearby weather stations, taking into account both spatial and temporal information from 2020.
+\hypertarget{river-levels-and-rainfall-in-victoria}{%
+\subsection{River levels and rainfall in Victoria}\label{river-levels-and-rainfall-in-victoria}}
+
+River level and rainfall data for the same areas should have some similarity.
+Here we examine the river gauge data from the Bureau of Meteorology \citep{REF} in relation to weather station rainfall from NOAA's climate data.
+Figure \ref{fig:matching-map} shows the location of available weather stations and water gauges in Victoria.
+The goal is to match water gauges with nearby weather stations, spatially and temporally, using 2020 measurements.
 
@@
 
-As introduced in Section \ref{matching}, \code{match_sites()} can be used for this task, and we choose \code{Water_course_level} in \code{river} and \code{prcp} in \code{climate} for the temporal matching.
-We first pass in the two datasets, followed by the variables used for the temporal matching, specified in \code{temporal_by} (using the \code{by} syntax from \code{dplyr::join}).
-With \code{temporal_independent} we select precipitation (\code{prcp}) as the independent variable, and the goal is to see if the water level in the river reflects what is observed in precipitation (for nearby stations).
-Since we consider a full year of data, we set the number of peaks considered (\code{temporal_n_highest}) to 30 (slightly above the default value of 20), and \code{temporal_min_match} is raised accordingly.
+The \code{match_sites()} function is used.
+\code{Water_course_level} from \code{river} and \code{prcp} from \code{climate} is used for the temporal matching.
+The peaks are compared, as explained in Section \ref{sec:matching}.
+A value of 30 is used for the maximum number of peaks to match, which corresponds to about 2 per month.
 
@@
 
-Figure \ref{fig:matching} (A) shows the four matched pairs on the map.
-For these the expected concurrent increase in precipitation and water level can be seen clearly when comparing these variables on a standardised scale, see Figure \ref{fig:matching} (B).
+Figure \ref{fig:matching} shows four matched pairs on the map (A) and standardised data as time series (B). The expected concurrent increase in precipitation and water level can be seen clearly.
 
@@
 
-{\centering \includegraphics{figures/matching-1}
+{\centering \includegraphics[width=1\linewidth]{figures/matching-1}
 

### 2022-06-22 4-4-netcdf.Rmd
section 5.4

@@
 
+Figure \ref{fig:netcdf} reproduces the ERA5 data row of Figure 19 in @hersbach2020era5.
+Here we explain how this would be done using cubble.
+The plots show that the southern polar vortex splits into two on 2002-09-26 and further splits into four on 2002-10-04.
+Further explanation of why this is interesting can be found in the figure source, and also @simmons2020global, @simmons2005ecmwf.
+
 The ERA5 data [@hersbach2020era5] provides hourly estimates across the Earth for atmospheric, land and oceanic climate variables.
-The data is available in the NetCDF format from the European Centre for Medium-Range Weather Forecasts (ECMWF) and it can be directly downloaded from [Copernicus Climate Data Store (CDS)](https://cds.climate.copernicus.eu/cdsapp#!/dataset/reanalysis-era5-pressure-levels?tab=overview) website or programmatically via the R package \pkg{ecmwfr} [@ecwmfr].
-In this example we focus on the \code{era5-pressure} data which provides hourly data on pressure levels from 1970 to present.
-Here we specifically look at the information  contained on the variables *specific humidity* measuring the mass of water vapor per kilogram of moist air  and the *geopotential* variable recording the gravitational potential energy of a unit mass, at a particular location, relative to mean sea level.
-The goal of this example is to reproduce the break-up of the southern polar vortex that happened in late September and early October in 2002 in the stratosphere in the Southern Hemisphere.
-Therefore, we filter data corresponding to the Southern Hemisphere and restrict our analysis to four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04.
-Once downloaded, the data can be read into a \code{cubble} as:
+The data is available in the NetCDF format from the European Centre for Medium-Range Weather Forecasts (ECMWF).
+It can be directly downloaded from [Copernicus Climate Data Store (CDS)](https://cds.climate.copernicus.eu/cdsapp#!/dataset/reanalysis-era5-pressure-levels?tab=overview) website or via the R package \pkg{ecmwfr} [@ecwmfr].
+
+For the reproduction, we focus on the \code{era5-pressure} data, hourly pressure levels from 1970 to present, with the *specific humidity* and *geopotential*.
+<!--Here we specifically look at the information  contained on the variables *specific humidity* measuring the mass of water vapor per kilogram of moist air  and the *geopotential* variable recording the gravitational potential energy of a unit mass, at a particular location, relative to mean sea level.-->
+
+<!--The goal of this example is to reproduce the break-up of the southern polar vortex that happened in late September and early October in 2002 in the stratosphere in the Southern Hemisphere.
+Therefore, we filter data corresponding to the Southern Hemisphere and restrict our analysis to four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04.--> The downloaded netcdf data is read into R using the \code{ncdf4} package, and converted to a \code{cubble}:
 
@@
 
-Figure \ref{fig:netcdf} reproduces the ERA5 data row of Figure 19 in @hersbach2020era5.
+<!-- Figure \ref{fig:netcdf} reproduces the ERA5 data row of Figure 19 in @hersbach2020era5.
 It shows the southern polar vortex splits into two on 2002-09-26 and further splits into four on 2002-10-04 in the stratosphere.
-Readers interested in the analysis of this figure can refer to @hersbach2020era5, @simmons2020global, and @simmons2005ecmwf for more details.
+Readers interested in the analysis of this figure can refer to @hersbach2020era5, @simmons2020global, and @simmons2005ecmwf for more details.-->
+
+Creating the plot requires making transformations on time, unfolding the data for computing the statistic of interest, which is plotted directly as a contour plot with \code{ggplot}.
+Code is provided to accomplish this in the supplementary material.
 

### 2022-06-22 jss/cubble-jss.tex
section 5.4

@@
 
+Figure \ref{fig:netcdf} reproduces the ERA5 data row of Figure 19 in \citet{hersbach2020era5}.
+Here we explain how this would be done using cubble.
+The plots show that the southern polar vortex splits into two on 2002-09-26 and further splits into four on 2002-10-04.
+Further explanation of why this is interesting can be found in the figure source, and also \citet{simmons2020global}, \citet{simmons2005ecmwf}.
+
 The ERA5 data \citep{hersbach2020era5} provides hourly estimates across the Earth for atmospheric, land and oceanic climate variables.
-The data is available in the NetCDF format from the European Centre for Medium-Range Weather Forecasts (ECMWF) and it can be directly downloaded from \href{https://cds.climate.copernicus.eu/cdsapp\#!/dataset/reanalysis-era5-pressure-levels?tab=overview}{Copernicus Climate Data Store (CDS)} website or programmatically via the R package \pkg{ecmwfr} \citep{ecwmfr}.
-In this example we focus on the \code{era5-pressure} data which provides hourly data on pressure levels from 1970 to present.
-Here we specifically look at the information contained on the variables \emph{specific humidity} measuring the mass of water vapor per kilogram of moist air and the \emph{geopotential} variable recording the gravitational potential energy of a unit mass, at a particular location, relative to mean sea level.
-The goal of this example is to reproduce the break-up of the southern polar vortex that happened in late September and early October in 2002 in the stratosphere in the Southern Hemisphere.
-Therefore, we filter data corresponding to the Southern Hemisphere and restrict our analysis to four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04.
-Once downloaded, the data can be read into a \code{cubble} as:
+The data is available in the NetCDF format from the European Centre for Medium-Range Weather Forecasts (ECMWF).
+It can be directly downloaded from \href{https://cds.climate.copernicus.eu/cdsapp\#!/dataset/reanalysis-era5-pressure-levels?tab=overview}{Copernicus Climate Data Store (CDS)} website or via the R package \pkg{ecmwfr} \citep{ecwmfr}.
+
+For the reproduction, we focus on the \code{era5-pressure} data, hourly pressure levels from 1970 to present, with the \emph{specific humidity} and \emph{geopotential}.
+
+The downloaded netcdf data is read into R using the \code{ncdf4} package, and converted to a \code{cubble}:
 
@@
 
-Figure \ref{fig:netcdf} reproduces the ERA5 data row of Figure 19 in \citet{hersbach2020era5}.
-It shows the southern polar vortex splits into two on 2002-09-26 and further splits into four on 2002-10-04 in the stratosphere.
-Readers interested in the analysis of this figure can refer to \citet{hersbach2020era5}, \citet{simmons2020global}, and \citet{simmons2005ecmwf} for more details.
+Creating the plot requires making transformations on time, unfolding the data for computing the statistic of interest, which is plotted directly as a contour plot with \code{ggplot}.
+Code is provided to accomplish this in the supplementary material.
 

### 2022-06-22 4-5-interactive.Rmd
appendix

@@
 
-When dealing with spatio-temporal data, users may wish to make plots to learn the spatial distribution of a variable or to find patterns, such as trend or seasonality, in the time series.
-Combining these two types of plots with interactivity lets users link between points on the map and the corresponding time series to explore the spatial and temporal dimensions of the data simultaneously.
-Below is an example that describes the process of building an interactive graphic with \pkg{cubble} and \pkg{crosstalk}.
-This example explores the variation of monthly temperature range within the \code{climate_full} data.
+Interactive graphics can be useful to make it possible to look at the data in a multiple of ways.
+This is especially important for spatiotemporal data, where we would like to interactively connect spatial and temporal displays.
+This example describes the process of building an interactive display with \pkg{cubble} and \pkg{crosstalk}.
+It explores the variation of monthly temperature range within the \code{climate_full} data.
+
+<!--When dealing with spatio-temporal data, users may wish to make plots to learn the spatial distribution of a variable or to find patterns, such as trend or seasonality, in the time series.
+Combining these two types of plots with interactivity lets users link between points on the map and the corresponding time series to explore the spatial and temporal dimensions of the data simultaneously.-->
 
@@
 ```
-
-This plot can also be made using \pkg{cubble} and \pkg{leaflet}, in which case the temperature range is displayed as a small subplot upon clicking on the map.
-This procedure involves first creating the popup plots from the long form cubble as a vector and then adding these plots to a leaflet map created from the nested cubble, with \code{leafpop::addPopupGraphs()}:
-
-``` r
-# data pre-processing
-clean <- climate_full %>% ...
-
-# use the long form to create subplots for each station
-df_id <- unique(clean$id)
-p <- map(1:length(df_id), function(i){
-  dt <- clean %>% filter(id == df_id[i])
-  ggplot(dt) %>% ...
-})
-
-# create nested form leaflet map with temperature band as subplots
-nested <- face_spatial(clean)
-leaflet(nested) |>
-  addTiles() |>
-  addCircleMarkers(group = "a", ...) |>
-  leafpop::addPopupGraphs(graph = p, ...)
-
-```
-
-Figure \ref{fig:interactive-popup} shows the same information as Figure \ref{fig:interactive-linking} but using leaflet and popups [@leafpop].
-
-```{r interactive-popup, fig.align='center', echo = FALSE, out.width="45%", out.height="25%", fig.retina = 2, dpi = 300, fig.cap = "Same as Figure 11 except the temperature variation is now shown as a popup in the leaflet map.", fig.show='hold'}
-knitr::include_graphics(here::here("figures/popup-mid.png"))
-```

### 2022-06-22 jss/appendix.Rmd
appendix

@@
     number_section: true
-
 ---
@@
 
+# Extension to section 5.5: Interactive graphics
+
+Figure XXX in this section can also be made using \pkg{cubble} and \pkg{leaflet}, in which case the temperature range is displayed as a small subplot upon clicking on the map.
+This procedure involves first creating the popup plots from the long form cubble as a vector and then adding these plots to a leaflet map created from the nested cubble, with \code{leafpop::addPopupGraphs()}:
+
+``` r
+# data pre-processing
+clean <- climate_full %>% ...
+
+# use the long form to create subplots for each station
+df_id <- unique(clean$id)
+p <- map(1:length(df_id), function(i){
+  dt <- clean %>% filter(id == df_id[i])
+  ggplot(dt) %>% ...
+})
+
+# create nested form leaflet map with temperature band as subplots
+nested <- face_spatial(clean)
+leaflet(nested) |>
+  addTiles() |>
+  addCircleMarkers(group = "a", ...) |>
+  leafpop::addPopupGraphs(graph = p, ...)
+
+```
+
+Figure \ref{fig:interactive-popup} shows the same information as Figure XXX but using leaflet and popups [@leafpop].
+
+```{r interactive-popup, fig.align='center', echo = FALSE, out.width="45%", out.height="25%", fig.retina = 2, dpi = 300, fig.cap = "Same as Figure 11 except the temperature variation is now shown as a popup in the leaflet map.", fig.show='hold'}
+knitr::include_graphics(here::here("figures/popup-mid.png"))
+```
+
 # Additional illustration on multiple linked plots

### 2022-06-22 jss/cubble-jss.tex
appendix

@@
 
-When dealing with spatio-temporal data, users may wish to make plots to learn the spatial distribution of a variable or to find patterns, such as trend or seasonality, in the time series.
-Combining these two types of plots with interactivity lets users link between points on the map and the corresponding time series to explore the spatial and temporal dimensions of the data simultaneously.
-Below is an example that describes the process of building an interactive graphic with \pkg{cubble} and \pkg{crosstalk}.
-This example explores the variation of monthly temperature range within the \code{climate_full} data.
+Interactive graphics can be useful to make it possible to look at the data in a multiple of ways.
+This is especially important for spatiotemporal data, where we would like to interactively connect spatial and temporal displays.
+This example describes the process of building an interactive display with \pkg{cubble} and \pkg{crosstalk}.
+It explores the variation of monthly temperature range within the \code{climate_full} data.
 

### 2022-06-22 jss/appendix.Rmd
knitted appendix

@@
 
-Figure XXX in this section can also be made using \pkg{cubble} and \pkg{leaflet}, in which case the temperature range is displayed as a small subplot upon clicking on the map.
+Figure XXX in this section can also be made using `cubble` and `leaflet`, in which case the temperature range is displayed as a small subplot upon clicking on the map.
 This procedure involves first creating the popup plots from the long form cubble as a vector and then adding these plots to a leaflet map created from the nested cubble, with \code{leafpop::addPopupGraphs()}:

### 2022-06-22 jss/appendix.Rmd
knitted appendix

@@
 Figure XXX in this section can also be made using `cubble` and `leaflet`, in which case the temperature range is displayed as a small subplot upon clicking on the map.
-This procedure involves first creating the popup plots from the long form cubble as a vector and then adding these plots to a leaflet map created from the nested cubble, with \code{leafpop::addPopupGraphs()}:
+This procedure involves first creating the popup plots from the long form cubble as a vector and then adding these plots to a leaflet map created from the nested cubble, with `leafpop::addPopupGraphs()`:
 

### 2022-06-22 4-5-interactive.Rmd
di is done with entire paper

@@
 
-Interactive graphics can be useful to make it possible to look at the data in a multiple of ways.
-This is especially important for spatiotemporal data, where we would like to interactively connect spatial and temporal displays.
-This example describes the process of building an interactive display with \pkg{cubble} and \pkg{crosstalk}.
-It explores the variation of monthly temperature range within the \code{climate_full} data.
+Interactive graphics can be useful because they make it possible to look at the data in a multiple of ways on-the-fly.
+This is especially important for spatio-temporal data, where we would like to interactively connect spatial and temporal displays.
+This example describes the process of using \pkg{cubble} with \pkg{crosstalk} to build an interactive display connecting a map of Australia, with ribbon plots of temperature range observed at the stations.
+The purpose is to explore the variation of monthly temperature range over the country.
+Figure \ref{fig:interactive-linking} shows three snapshots of the interactivity.
 
@@
 
-The temperature range is calculated as the difference between \code{tmax} and \code{tmin} and its monthly average over 2016-2020 is taken before calculating the variance.
+The key steps are to convert both the nested and long forms of the data into shared objects, and to plot these sid-by-side. The two are linked by the station identifier.
+
+<!--The temperature range is calculated as the difference between \code{tmax} and \code{tmin} and its monthly average over 2016-2020 is taken before calculating the variance.
 A \code{SharedData} object is constructed for each form of the cubble and the same \code{group} argument ensures the cross-linking of the two forms via the common \code{id} column.
@@
 The cross-linking between the two plots is always safeguarded by the shared \code{id} column embedded in the cubble structure.
-Below is the pseudo-code that outlines the process to construct the interactive graphic described above:
+Below is the pseudo-code that outlines the process to construct the interactive graphic described above:-->
 
@@
 
-In Figure \ref{fig:interactive-linking}, the first row shows the initial view of the interactive graphic.
+Plot (a) shows the initial state of the interactive display: all locations are shown as dots on the map, coloured by temperature range, and the right plot shows the ribbons representing maximum to minimum for all stations.
+In plot (b) the "mount elizabeth" station, which shows a high variance colour on the initial map, is selected on the map and this produces the ribbon on the right.
+In plot (c) the lowest temperature in August is selected, which is "thredbo" station on the left map.
+It was surprising to us that this was not a station in Tasmania, so for comparison a station in Tasmania is selected on the map to show in relation to thredbo.
+We can see that thredbo has a bigger winter dip in temperature, and although Tasmania is cold generally, it's temperatures are more consistent.
+
+<!-- In Figure \ref{fig:interactive-linking}, the first row shows the initial view of the interactive graphic.
 Most regions in Australia have a low variance of temperature range, while the North-West coastline, the bottom of South Australia, and Victoria stand out with larger monthly changes.
 In the second row, the Mount Elizabeth station, which shows a high variance colour on the initial map, is selected on the map and this produces the ribbon on the right.
-In the third row, the lowest temperature in August is selected, which highlights the corresponding Thredbo AWS station on the left map.
-Another station in the Tasmania island is selected on the map to cross compare with Thredbo AWS.
+In the third row, the lowest temperature in August is selected, which highlights the corresponding thredbo station on the left map.
+-->
 

### 2022-06-22 5-conclusion.Rmd
di is done with entire paper

@@
 The package introduces a new data structure for spatio-temporal data, \code{cubble}, that connects the time invariant and varying variables and that allows the user to work with a nested and long form of the data.
-TThe goal of this work is to add capabilities into the spatio-temporal practitioners toolbox that facilitates their work within the tidy data framework.
+This work adds capabilities into the spatio-temporal practitioners toolbox to integrate it with a tidy data framework.
 The data structure and the capabilities introduced in this package can be used and combined with existing spatio-temporal R packages such as \pkg{sf}, data wrangling packages such as \pkg{dplyr}, and visualization packages such as \pkg{ggplot2}, \pkg{plotly} and \pkg{leaflet}.
 
-The data structure and functions are demonstrated with extensive examples.
-These include creating and coercing wild-caught data with potential mismatch on sites, handling hierarchical data, matching time series spatially and temporally, as well as reproducing ERA5 results from NetCDF data.
-Visualization of the \code{cubble} objects and derivatives is presented via interactive graphic pipelines using \pkg{plotly} and \pkg{leaflet}.
+Numerous examples are provided.
+These include creating and coercing data with mismatched names, handling hierarchical data, matching time series spatially and temporally, reproducing ERA5 plots from NetCDF data.
+Visualization of the \code{cubble} objects can be done with interactive graphic pipelines using \pkg{crosstalk}, \pkg{plotly} and \pkg{leaflet}.
 
-Future directions of the package involves handling sites with moving coordinates.
-This would involve constructing a list-column for location coordinates and a form these locations can be pivoted into, like the long form for temporal variables.
-In the multivariate aspect, \pkg{cubble} can also be extended with interface to more high dimensional visualisation methods, i.e., the tour method, to understand variable importance or comparing location similarities.
-
-\newpage
+There are several possible future directions for the work.
+The data structure only described fixed spatial sites, and it could be useful to provide tools to accommodate moving coordinates as might be encountered in animal movement data.
+It could be achieved with a list-column for the location coordinates, and an additional form that these locations can be pivoted into, like the long form for temporal variables.
+For multiple measured variables, \pkg{cubble} could be integrated with dimension reduction methods, and dynamic graphics for multiple dimensions such as a tour [@REF].
 

### 2022-06-22 jss/cubble-jss.tex
di is done with entire paper

@@
 
-Interactive graphics can be useful to make it possible to look at the data in a multiple of ways.
-This is especially important for spatiotemporal data, where we would like to interactively connect spatial and temporal displays.
-This example describes the process of building an interactive display with \pkg{cubble} and \pkg{crosstalk}.
-It explores the variation of monthly temperature range within the \code{climate_full} data.
-
-The temperature range is calculated as the difference between \code{tmax} and \code{tmin} and its monthly average over 2016-2020 is taken before calculating the variance.
-A \code{SharedData} object is constructed for each form of the cubble and the same \code{group} argument ensures the cross-linking of the two forms via the common \code{id} column.
-The spatial map and time series plot are then made with each \code{SharedData} object separately.
-In this example, stations on the Australia map, made from the nested form, are coloured by the calculated variance.
-A ribbon band is constructed using the long form cubble to show each station's maximum and minimum temperature over time.
-In the case of a different dataset, users could calculate other station-dependent measures in the nested form or make other time-wise summary of the data in the long form to customise the spatial or temporal view.
-The cross-linking between the two plots is always safeguarded by the shared \code{id} column embedded in the cubble structure.
-Below is the pseudo-code that outlines the process to construct the interactive graphic described above:
+Interactive graphics can be useful because they make it possible to look at the data in a multiple of ways on-the-fly.
+This is especially important for spatio-temporal data, where we would like to interactively connect spatial and temporal displays.
+This example describes the process of using \pkg{cubble} with \pkg{crosstalk} to build an interactive display connecting a map of Australia, with ribbon plots of temperature range observed at the stations.
+The purpose is to explore the variation of monthly temperature range over the country.
+Figure \ref{fig:interactive-linking} shows three snapshots of the interactivity.
+
+The key steps are to convert both the nested and long forms of the data into shared objects, and to plot these sid-by-side. The two are linked by the station identifier.
 
@@
 
-In Figure \ref{fig:interactive-linking}, the first row shows the initial view of the interactive graphic.
-Most regions in Australia have a low variance of temperature range, while the North-West coastline, the bottom of South Australia, and Victoria stand out with larger monthly changes.
-In the second row, the Mount Elizabeth station, which shows a high variance colour on the initial map, is selected on the map and this produces the ribbon on the right.
-In the third row, the lowest temperature in August is selected, which highlights the corresponding Thredbo AWS station on the left map.
-Another station in the Tasmania island is selected on the map to cross compare with Thredbo AWS.
+Plot (a) shows the initial state of the interactive display: all locations are shown as dots on the map, coloured by temperature range, and the right plot shows the ribbons representing maximum to minimum for all stations.
+In plot (b) the ``mount elizabeth'' station, which shows a high variance colour on the initial map, is selected on the map and this produces the ribbon on the right.
+In plot (c) the lowest temperature in August is selected, which is ``thredbo'' station on the left map.
+It was surprising to us that this was not a station in Tasmania, so for comparison a station in Tasmania is selected on the map to show in relation to thredbo.
+We can see that thredbo has a bigger winter dip in temperature, and although Tasmania is cold generally, it's temperatures are more consistent.
 
@@
 
-This plot can also be made using \pkg{cubble} and \pkg{leaflet}, in which case the temperature range is displayed as a small subplot upon clicking on the map.
-This procedure involves first creating the popup plots from the long form cubble as a vector and then adding these plots to a leaflet map created from the nested cubble, with \code{leafpop::addPopupGraphs()}:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# data pre{-}processing}
-\NormalTok{clean }\OtherTok{\textless{}{-}}\NormalTok{ climate\_full }\SpecialCharTok{\%\textgreater{}\%}\NormalTok{ ...}
-
-\CommentTok{\# use the long form to create subplots for each station}
-\NormalTok{df\_id }\OtherTok{\textless{}{-}} \FunctionTok{unique}\NormalTok{(clean}\SpecialCharTok{$}\NormalTok{id)}
-\NormalTok{p }\OtherTok{\textless{}{-}} \FunctionTok{map}\NormalTok{(}\DecValTok{1}\SpecialCharTok{:}\FunctionTok{length}\NormalTok{(df\_id), }\ControlFlowTok{function}\NormalTok{(i)\{}
-\NormalTok{  dt }\OtherTok{\textless{}{-}}\NormalTok{ clean }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(id }\SpecialCharTok{==}\NormalTok{ df\_id[i])}
-  \FunctionTok{ggplot}\NormalTok{(dt) }\SpecialCharTok{\%\textgreater{}\%}\NormalTok{ ...}
-\NormalTok{\})}
-
-\CommentTok{\# create nested form leaflet map with temperature band as subplots}
-\NormalTok{nested }\OtherTok{\textless{}{-}} \FunctionTok{face\_spatial}\NormalTok{(clean)}
-\FunctionTok{leaflet}\NormalTok{(nested) }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
-  \FunctionTok{addTiles}\NormalTok{() }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
-  \FunctionTok{addCircleMarkers}\NormalTok{(}\AttributeTok{group =} \StringTok{"a"}\NormalTok{, ...) }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
-\NormalTok{  leafpop}\SpecialCharTok{::}\FunctionTok{addPopupGraphs}\NormalTok{(}\AttributeTok{graph =}\NormalTok{ p, ...)}
-\end{Highlighting}
-\end{Shaded}
-
-Figure \ref{fig:interactive-popup} shows the same information as Figure \ref{fig:interactive-linking} but using leaflet and popups \citep{leafpop}.
-
-\begin{CodeChunk}
-\begin{figure}
-
-{\centering \includegraphics[width=0.45\linewidth,height=0.25\textheight]{../figures/popup-mid}
-
-}
-
-\caption[Same as Figure 11 except the temperature variation is now shown as a popup in the leaflet map]{Same as Figure 11 except the temperature variation is now shown as a popup in the leaflet map.}\label{fig:interactive-popup}
-\end{figure}
-\end{CodeChunk}
-
 \hypertarget{conclude}{%
@@
 The package introduces a new data structure for spatio-temporal data, \code{cubble}, that connects the time invariant and varying variables and that allows the user to work with a nested and long form of the data.
-TThe goal of this work is to add capabilities into the spatio-temporal practitioners toolbox that facilitates their work within the tidy data framework.
+This work adds capabilities into the spatio-temporal practitioners toolbox to integrate it with a tidy data framework.
 The data structure and the capabilities introduced in this package can be used and combined with existing spatio-temporal R packages such as \pkg{sf}, data wrangling packages such as \pkg{dplyr}, and visualization packages such as \pkg{ggplot2}, \pkg{plotly} and \pkg{leaflet}.
 
-The data structure and functions are demonstrated with extensive examples.
-These include creating and coercing wild-caught data with potential mismatch on sites, handling hierarchical data, matching time series spatially and temporally, as well as reproducing ERA5 results from NetCDF data.
-Visualization of the \code{cubble} objects and derivatives is presented via interactive graphic pipelines using \pkg{plotly} and \pkg{leaflet}.
-
-Future directions of the package involves handling sites with moving coordinates.
-This would involve constructing a list-column for location coordinates and a form these locations can be pivoted into, like the long form for temporal variables.
-In the multivariate aspect, \pkg{cubble} can also be extended with interface to more high dimensional visualisation methods, i.e., the tour method, to understand variable importance or comparing location similarities.
-
-\newpage
+Numerous examples are provided.
+These include creating and coercing data with mismatched names, handling hierarchical data, matching time series spatially and temporally, reproducing ERA5 plots from NetCDF data.
+Visualization of the \code{cubble} objects can be done with interactive graphic pipelines using \pkg{crosstalk}, \pkg{plotly} and \pkg{leaflet}.
+
+There are several possible future directions for the work.
+The data structure only described fixed spatial sites, and it could be useful to provide tools to accommodate moving coordinates as might be encountered in animal movement data.
+It could be achieved with a list-column for the location coordinates, and an additional form that these locations can be pivoted into, like the long form for temporal variables.
+For multiple measured variables, \pkg{cubble} could be integrated with dimension reduction methods, and dynamic graphics for multiple dimensions such as a tour \citep{REF}.
 

### 2022-06-22 4-1-basic.Rmd
Update section 5.1

@@
 
-The five examples here are chosen to illustrate these aspects of using cubble: creating a cubble from two COVID data tables with the complication of differing location names, using spatial transformations to make a glyph map of seasonal temperature changes over years, aggregating information spatially to explore precipitation patterns, matching river level data and weather station records for analysis of water supply, reading netCDF format data to reproduce a climate reanalysis plot, and the workflow to create complex interactive linked plots.
+The five examples here are chosen to illustrate these aspects of using cubble: creating a \code{cubble} from two COVID data tables with the complication of differing location names, using spatial transformations to make a glyph map of seasonal temperature changes over years, aggregating information spatially to explore precipitation patterns, matching river level data and weather station records for analysis of water supply, reading netCDF format data to reproduce a climate reanalysis plot, and the workflow to create complex interactive linked plots.
 (There is an additional example and figures in the Appendix, and more examples in the package vignettes.)
@@
 Here is how to do this with \pkg{cubble}, including how to handle mismatches arising from different names of the same LGAs in the two tables.
-The COVID data is a `csv` and looks like:
+The COVID data is store in a `csv` file and looks like:
 
@@
 
-and the spatial polygons are an ESRI shapefile that looks like:
+and the spatial polygons are an ESRI shapefile as follows:
 
@@
 
-The function \code{as_cubble()} is used to create a \code{cubble} object from the two spatial and temporal tables, and requires specifying the arguments (\code{key}, \code{index}, and \code{coords}) (as described in Section \@ref(create)).
-It will automatically try to match the sites in both tables and will show a warning message when there are mismatches, as shown below:
+The function \code{as_cubble()} is used to create a \code{cubble} object from the two spatial and temporal tables, and requires specifying the arguments \code{key}, \code{index}, and \code{coords}, (as described in Section \@ref(create)).
+It will automatically try to match the sites in both tables using the location names and will show a warning message when there are mismatches, as shown below:
 
@@
 
-It can be seen that there are two-way mismatches -- LGAs in the COVID data that don't match with LGAs in the spatial polygon data, and vice versa.
-The mismatches can be identified by using the \code{output = "unmatch"} argument.
-<!--Note that \pkg{cubble} will attempt to pair the unmatched sites, these are shown separately from the ones that cannot be paired.-->
+It can be seen that there are two-way mismatches -- LGAs in the COVID data that do not  match with LGAs names in the spatial polygon data, and vice versa.
+The mismatches can be identified by using the \code{output = "unmatch"} argument in the \code{as_cubble()} functiton: <!--Note that \pkg{cubble} will attempt to pair the unmatched sites, these are shown separately from the ones that cannot be paired.-->
 
@@
 
-With this information both tables can be fixed, to create the desired cubble, as follows:
+With this information both tables can be fixed, to create the desired \code{cubble} object, as follows:
 

### 2022-06-22 jss/cubble-jss.tex
Update section 5.1

@@
 Here is how to do this with \pkg{cubble}, including how to handle mismatches arising from different names of the same LGAs in the two tables.
-The COVID data is a \texttt{csv} and looks like:
+The COVID data is store in a \texttt{csv} file and looks like:
 
@@
 
-The function \code{as_cubble()} is used to create a \code{cubble} object from the two spatial and temporal tables, and requires specifying the arguments (\code{key}, \code{index}, and \code{coords}) (as described in Section \ref{create}).
-It will automatically try to match the sites in both tables and will show a warning message when there are mismatches, as shown below:
+The function \code{as_cubble()} is used to create a \code{cubble} object from the two spatial and temporal tables, and requires specifying the arguments \code{key}, \code{index}, and \code{coords}, (as described in Section \ref{create}).
+It will automatically try to match the sites in both tables using the location names and will show a warning message when there are mismatches, as shown below:
 
@@
 \end{CodeOutput}
-\end{CodeChunk}
-
-It can be seen that there are two-way mismatches -- LGAs in the COVID data that don't match with LGAs in the spatial polygon data, and vice versa.
-The mismatches can be identified by using the \code{output = "unmatch"} argument.
+\begin{CodeOutput}
+Joining, by = "lga"
+\end{CodeOutput}
+\end{CodeChunk}
+
+It can be seen that there are two-way mismatches -- LGAs in the COVID data that do not match with LGAs names in the spatial polygon data, and vice versa.
+The mismatches can be identified by using the \code{output = "unmatch"} argument in the \code{as_cubble()} functiton:
 

### 2022-06-22 4-1-basic.Rmd
Update section 5.2

@@
 
-The function \code{as_cubble()} is used to create a \code{cubble} object from the two spatial and temporal tables, and requires specifying the arguments \code{key}, \code{index}, and \code{coords}, (as described in Section \@ref(create)).
+The function \code{as_cubble()} is used to create a \code{cubble} object from the two spatial and temporal tables, and requires specifying the arguments \code{key}, \code{index}, and \code{coords} (as described in Section \@ref(create)).
 It will automatically try to match the sites in both tables using the location names and will show a warning message when there are mismatches, as shown below:
@@
 It can be seen that there are two-way mismatches -- LGAs in the COVID data that do not  match with LGAs names in the spatial polygon data, and vice versa.
-The mismatches can be identified by using the \code{output = "unmatch"} argument in the \code{as_cubble()} functiton: <!--Note that \pkg{cubble} will attempt to pair the unmatched sites, these are shown separately from the ones that cannot be paired.-->
+The mismatches can be identified by using the \code{output = "unmatch"} argument in the \code{as_cubble()} function: <!--Note that \pkg{cubble} will attempt to pair the unmatched sites, these are shown separately from the ones that cannot be paired.-->
 
@@
 
-First, the stations need to be filtered to those in NSW and Victoria, by using the station identifiers, stored within the 11 digits of the \code{id} variable entries.
+First, the stations need to be filtered to those in NSW and Victoria by using the station identifiers, stored within the 11 digits of the \code{id} variable entries.
 The country code is in the first 5 digits (Australia is represented by "ASN00") and the next 6 digits encode the station following the [Australian Bureau of Meteorology (BOM)](http://www.bom.gov.au/climate/cdo/about/site-num.shtml) coding protocols.
@@
 
-Next, a monthly maximum average temperature is  calculated for both periods.
-This is a *temporal operation* requiring a switch into the long cubble form, and creating a new indicator for the two time periods, before the calculation:
+Next, the monthly maximum average temperature is  calculated for both periods.
+This is a *temporal operation* requiring a switch into the long \code{cubble} form using the \code{face_temporal()} function.
+In addition, a new indicator for the two time periods of interest is created before the calculation of monthly averages:
 
@@
 
-A quick check on the number of observations for each location is made, revealing that there are several with less than 24 -- these lack temperature values for some months.
-These stations are removed, by switching to a long cubble to operate on the spatial component over time, and then back to the nested cubble (to make the glyph map):
+A quick check on the number of observations for each location is made, revealing that there are several with less than 24 observations -- these stations lack temperature values for some months.
+In this example, those stations are removed by switching to a long \code{cubble} to operate on the spatial component over time, and then, move back into the nested \code{cubble} (to make the glyph map):
 
@@
 
-To create a glyph map of the monthly series overlaid on the map (Figure \ref{fig:basic-manip}), requires that the spatial variables be unfolded with the temporal variables.
-Both of the major (\code{long}, \code{lat}) and minor (\code{month}, \code{tmax}) coordinates need to be on the same table.
+In order to create a glyph map displaying the monthly series  (Figure \ref{fig:basic-manip}), the spatial variables need to be unfolded with the temporal variables.
+The reason being that the major (\code{long}, \code{lat}) and minor (\code{month}, \code{tmax}) coordinates need to be on the same table to create the glyph map.
 The \code{geom_glyph()} function does both the transformation and the plotting.
-(An inset of the glyph for one station is shown in the top left corner.)
+\textcolor{green}{(An inset of the glyph for one station is shown in the top left corner.)}
 
@@
 Glyph maps work well to explore temporal patterns across spatial locations, particularly when the spatial locations are gridded.
-Here they are irregularly spaced, which can result in overlapping glyphs obscurring each other.
-To fix this one could aggregate data from nearby stations.
-An example is included in the Appendix.
+In this example, they are irregularly spaced, which can result in overlapping glyphs obscuring each other.
+To fix this, one could aggregate data from nearby stations.
+An example of this using **cubble** is included in the Appendix.

### 2022-06-22 jss/cubble-jss.tex
Update section 5.2

@@
 
-The function \code{as_cubble()} is used to create a \code{cubble} object from the two spatial and temporal tables, and requires specifying the arguments \code{key}, \code{index}, and \code{coords}, (as described in Section \ref{create}).
+The function \code{as_cubble()} is used to create a \code{cubble} object from the two spatial and temporal tables, and requires specifying the arguments \code{key}, \code{index}, and \code{coords} (as described in Section \ref{create}).
 It will automatically try to match the sites in both tables using the location names and will show a warning message when there are mismatches, as shown below:
@@
 It can be seen that there are two-way mismatches -- LGAs in the COVID data that do not match with LGAs names in the spatial polygon data, and vice versa.
-The mismatches can be identified by using the \code{output = "unmatch"} argument in the \code{as_cubble()} functiton:
+The mismatches can be identified by using the \code{output = "unmatch"} argument in the \code{as_cubble()} function:
 
@@
 
-First, the stations need to be filtered to those in NSW and Victoria, by using the station identifiers, stored within the 11 digits of the \code{id} variable entries.
+First, the stations need to be filtered to those in NSW and Victoria by using the station identifiers, stored within the 11 digits of the \code{id} variable entries.
 The country code is in the first 5 digits (Australia is represented by ``ASN00'') and the next 6 digits encode the station following the \href{http://www.bom.gov.au/climate/cdo/about/site-num.shtml}{Australian Bureau of Meteorology (BOM)} coding protocols.
@@
 
-Next, a monthly maximum average temperature is calculated for both periods.
-This is a \emph{temporal operation} requiring a switch into the long cubble form, and creating a new indicator for the two time periods, before the calculation:
+Next, the monthly maximum average temperature is calculated for both periods.
+This is a \emph{temporal operation} requiring a switch into the long \code{cubble} form using the \code{face_temporal()} function.
+In addition, a new indicator for the two time periods of interest is created before the calculation of monthly averages:
 
@@
 
-To create a glyph map of the monthly series overlaid on the map (Figure \ref{fig:basic-manip}), requires that the spatial variables be unfolded with the temporal variables.
-Both of the major (\code{long}, \code{lat}) and minor (\code{month}, \code{tmax}) coordinates need to be on the same table.
+In order to create a glyph map displaying the monthly series (Figure \ref{fig:basic-manip}), the spatial variables need to be unfolded with the temporal variables.
+The reason being that the major (\code{long}, \code{lat}) and minor (\code{month}, \code{tmax}) coordinates need to be on the same table to create the glyph map.
 The \code{geom_glyph()} function does both the transformation and the plotting.
-(An inset of the glyph for one station is shown in the top left corner.)
+\textcolor{green}{(An inset of the glyph for one station is shown in the top left corner.)}
 
@@
 Glyph maps work well to explore temporal patterns across spatial locations, particularly when the spatial locations are gridded.
-Here they are irregularly spaced, which can result in overlapping glyphs obscurring each other.
-To fix this one could aggregate data from nearby stations.
-An example is included in the Appendix.
+In this example, they are irregularly spaced, which can result in overlapping glyphs obscuring each other.
+To fix this, one could aggregate data from nearby stations.
+An example of this using \textbf{cubble} is included in the Appendix.
 

### 2022-06-22 4-3-water-level.Rmd
Udpate section 5.3

@@
 River level and rainfall data for the same areas should have some similarity.
-Here we examine the river gauge data from the Bureau of Meteorology [@REF] in relation to weather station rainfall from NOAA's climate data.
+Here we examine the river gauge data (\code{Water_course_level}) from the Bureau of Meteorology [@REF] in relation to weather station rainfall from NOAA's climate data [@REF] (\code{climate}).
 Figure \ref{fig:matching-map} shows the location of available weather stations and water gauges in Victoria.
-The goal is to match water gauges with nearby weather stations, spatially and temporally, using 2020 measurements.
-
+The goal is to match water gauges with nearby weather stations, spatially and temporally, using 2020 measurements using
+\code{match_sites()} function.
 <!--The Bureau of Meteorology collects [water data](http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full) from river gauges.
@@
 
-The \code{match_sites()} function is used.
-\code{Water_course_level} from \code{river} and \code{prcp} from \code{climate} is used for the temporal matching.
-The peaks are compared, as explained in Section \@ref(sec:matching).
-A value of 30 is used for the maximum number of peaks to match, which corresponds to about 2 per month.
+This function requires passing the major and minor data sets used for matching, in this case those are \code{river} and \code{climate}.
+The variables used for the temporal matching  are \code{Water_course_level} from the \code{river} data set and \code{prcp} in the climate data set.
+The rest of the arguments, as explained in Section \@ref(sec:matching), correspond to the maximum and minimum number of peaks in the time series to be matched.
+In this example those are set to be a maximum of 30 and a minimum of 15 (approximately 2 and 1 per month).
+
 <!--We first pass in the two datasets, followed by the variables used for the temporal matching, specified in \code{temporal_by} (using the \code{by} syntax from \code{dplyr::join}).
@@
 
-This function returns a cubble, with additional columns \code{dist} and \code{group} summarising spatial matching, and \code{n_match} for temporal matching.
+This function returns a \code{cubble} object, with additional columns: \code{dist} storing the distance between matched stations, \code{group} summarising spatial matching, and \code{n_match} showing the temporal matching.
 

### 2022-06-22 jss/cubble-jss.tex
Udpate section 5.3

@@
 River level and rainfall data for the same areas should have some similarity.
-Here we examine the river gauge data from the Bureau of Meteorology \citep{REF} in relation to weather station rainfall from NOAA's climate data.
+Here we examine the river gauge data (\code{Water_course_level}) from the Bureau of Meteorology \citep{REF} in relation to weather station rainfall from NOAA's climate data \citep{REF} (\code{climate}).
 Figure \ref{fig:matching-map} shows the location of available weather stations and water gauges in Victoria.
-The goal is to match water gauges with nearby weather stations, spatially and temporally, using 2020 measurements.
+The goal is to match water gauges with nearby weather stations, spatially and temporally, using 2020 measurements using
+\code{match_sites()} function.
 
@@
 
-The \code{match_sites()} function is used.
-\code{Water_course_level} from \code{river} and \code{prcp} from \code{climate} is used for the temporal matching.
-The peaks are compared, as explained in Section \ref{sec:matching}.
-A value of 30 is used for the maximum number of peaks to match, which corresponds to about 2 per month.
+This function requires passing the major and minor data sets used for matching, in this case those are \code{river} and \code{climate}.
+The variables used for the temporal matching are \code{Water_course_level} from the \code{river} data set and \code{prcp} in the climate data set.
+The rest of the arguments, as explained in Section \ref{sec:matching}, correspond to the maximum and minimum number of peaks in the time series to be matched.
+In this example those are set to be a maximum of 30 and a minimum of 15 (approximately 2 and 1 per month).
 

### 2022-06-22 5-conclusion.Rmd
Update conclusion

@@
 This work adds capabilities into the spatio-temporal practitioners toolbox to integrate it with a tidy data framework.
-The data structure and the capabilities introduced in this package can be used and combined with existing spatio-temporal R packages such as \pkg{sf}, data wrangling packages such as \pkg{dplyr}, and visualization packages such as \pkg{ggplot2}, \pkg{plotly} and \pkg{leaflet}.
+The data structure and functions introduced in this package can be used and combined with existing spatio-temporal R packages such as \pkg{sf}, data wrangling packages such as \pkg{dplyr}, and visualization packages such as \pkg{ggplot2}, \pkg{plotly} and \pkg{leaflet}.
 
-Numerous examples are provided.
+Numerous examples are provided in the main text, appendix and package vignettes.
 These include creating and coercing data with mismatched names, handling hierarchical data, matching time series spatially and temporally, reproducing ERA5 plots from NetCDF data.
@@
 The data structure only described fixed spatial sites, and it could be useful to provide tools to accommodate moving coordinates as might be encountered in animal movement data.
-It could be achieved with a list-column for the location coordinates, and an additional form that these locations can be pivoted into, like the long form for temporal variables.
+That could be achieved with a list-column for the location coordinates, and an additional form that these locations can be pivoted into, like the long form for temporal variables.
 For multiple measured variables, \pkg{cubble} could be integrated with dimension reduction methods, and dynamic graphics for multiple dimensions such as a tour [@REF].

### 2022-06-22 jss/cubble-jss.tex
Update conclusion

@@
 This work adds capabilities into the spatio-temporal practitioners toolbox to integrate it with a tidy data framework.
-The data structure and the capabilities introduced in this package can be used and combined with existing spatio-temporal R packages such as \pkg{sf}, data wrangling packages such as \pkg{dplyr}, and visualization packages such as \pkg{ggplot2}, \pkg{plotly} and \pkg{leaflet}.
-
-Numerous examples are provided.
+The data structure and functions introduced in this package can be used and combined with existing spatio-temporal R packages such as \pkg{sf}, data wrangling packages such as \pkg{dplyr}, and visualization packages such as \pkg{ggplot2}, \pkg{plotly} and \pkg{leaflet}.
+
+Numerous examples are provided in the main text, appendix and package vignettes.
 These include creating and coercing data with mismatched names, handling hierarchical data, matching time series spatially and temporally, reproducing ERA5 plots from NetCDF data.
@@
 The data structure only described fixed spatial sites, and it could be useful to provide tools to accommodate moving coordinates as might be encountered in animal movement data.
-It could be achieved with a list-column for the location coordinates, and an additional form that these locations can be pivoted into, like the long form for temporal variables.
+That could be achieved with a list-column for the location coordinates, and an additional form that these locations can be pivoted into, like the long form for temporal variables.
 For multiple measured variables, \pkg{cubble} could be integrated with dimension reduction methods, and dynamic graphics for multiple dimensions such as a tour \citep{REF}.

### 2022-06-24 4-1-basic.Rmd
fix citation on websites

@@
 First, the stations need to be filtered to those in NSW and Victoria by using the station identifiers, stored within the 11 digits of the \code{id} variable entries.
-The country code is in the first 5 digits (Australia is represented by "ASN00") and the next 6 digits encode the station following the [Australian Bureau of Meteorology (BOM)](http://www.bom.gov.au/climate/cdo/about/site-num.shtml) coding protocols.
+The country code is in the first 5 digits (Australia is represented by "ASN00") and the next 6 digits encode the station following the Australian Bureau of Meteorology (BOM) [@bom] coding protocols.
 New South Wales stations correspond to entries in the range 46-75 and the Victorian stations to 76-90.
 Filtering Victoria and New South Wales stations is a *spatial operation* and hence uses the \code{cubble} nested form:
+
+<!-- BOM link: http://www.bom.gov.au/climate/cdo/about/site-num.shtml -->
 

### 2022-06-24 4-3-water-level.Rmd
fix citation on websites

@@
 River level and rainfall data for the same areas should have some similarity.
-Here we examine the river gauge data (\code{Water_course_level}) from the Bureau of Meteorology [@REF] in relation to weather station rainfall from NOAA's climate data [@REF] (\code{climate}).
+Here we examine the river gauge data (\code{Water_course_level}) from the Bureau of Meteorology [@bom] in relation to weather station rainfall from NOAA's climate data [@noaa] (\code{climate}).
 Figure \ref{fig:matching-map} shows the location of available weather stations and water gauges in Victoria.
@@
 The variables used for the temporal matching  are \code{Water_course_level} from the \code{river} data set and \code{prcp} in the climate data set.
-The rest of the arguments, as explained in Section \@ref(sec:matching), correspond to the maximum and minimum number of peaks in the time series to be matched.
+The rest of the arguments, as explained in Section \@ref(matching), correspond to the maximum and minimum number of peaks in the time series to be matched.
 In this example those are set to be a maximum of 30 and a minimum of 15 (approximately 2 and 1 per month).

### 2022-06-24 4-4-netcdf.Rmd
fix citation on websites

@@
 
-Figure \ref{fig:netcdf} reproduces the ERA5 data row of Figure 19 in @hersbach2020era5.
+Figure \ref{fig:netcdf} reproduces the ERA5 data [@hersbach2020era5] row of Figure 19 in [@hersbach2020era5].
 Here we explain how this would be done using cubble.
 The plots show that the southern polar vortex splits into two on 2002-09-26 and further splits into four on 2002-10-04.
-Further explanation of why this is interesting can be found in the figure source, and also @simmons2020global, @simmons2005ecmwf.
+Further explanation of why this is interesting can be found in the figure source, and also [@simmons2020global], [@simmons2005ecmwf].
 
@@
 The data is available in the NetCDF format from the European Centre for Medium-Range Weather Forecasts (ECMWF).
-It can be directly downloaded from [Copernicus Climate Data Store (CDS)](https://cds.climate.copernicus.eu/cdsapp#!/dataset/reanalysis-era5-pressure-levels?tab=overview) website or via the R package \pkg{ecmwfr} [@ecwmfr].
+It can be directly downloaded from Copernicus Climate Data Store (CDS) [@cds] website or via the R package \pkg{ecmwfr} [@ecwmfr].
 

### 2022-06-24 5-conclusion.Rmd
fix citation on websites

@@
 That could be achieved with a list-column for the location coordinates, and an additional form that these locations can be pivoted into, like the long form for temporal variables.
-For multiple measured variables, \pkg{cubble} could be integrated with dimension reduction methods, and dynamic graphics for multiple dimensions such as a tour [@REF].
+For multiple measured variables, \pkg{cubble} could be integrated with dimension reduction methods, and dynamic graphics for multiple dimensions such as a tour [@wickham_tourr_2011].
 

### 2022-06-24 jss/cubble-jss.tex
fix citation on websites

@@
 First, the stations need to be filtered to those in NSW and Victoria by using the station identifiers, stored within the 11 digits of the \code{id} variable entries.
-The country code is in the first 5 digits (Australia is represented by ``ASN00'') and the next 6 digits encode the station following the \href{http://www.bom.gov.au/climate/cdo/about/site-num.shtml}{Australian Bureau of Meteorology (BOM)} coding protocols.
+The country code is in the first 5 digits (Australia is represented by ``ASN00'') and the next 6 digits encode the station following the Australian Bureau of Meteorology (BOM) \citep{bom} coding protocols.
 New South Wales stations correspond to entries in the range 46-75 and the Victorian stations to 76-90.
@@
 River level and rainfall data for the same areas should have some similarity.
-Here we examine the river gauge data (\code{Water_course_level}) from the Bureau of Meteorology \citep{REF} in relation to weather station rainfall from NOAA's climate data \citep{REF} (\code{climate}).
+Here we examine the river gauge data (\code{Water_course_level}) from the Bureau of Meteorology \citep{bom} in relation to weather station rainfall from NOAA's climate data \citep{noaa} (\code{climate}).
 Figure \ref{fig:matching-map} shows the location of available weather stations and water gauges in Victoria.
@@
 The variables used for the temporal matching are \code{Water_course_level} from the \code{river} data set and \code{prcp} in the climate data set.
-The rest of the arguments, as explained in Section \ref{sec:matching}, correspond to the maximum and minimum number of peaks in the time series to be matched.
+The rest of the arguments, as explained in Section \ref{matching}, correspond to the maximum and minimum number of peaks in the time series to be matched.
 In this example those are set to be a maximum of 30 and a minimum of 15 (approximately 2 and 1 per month).
@@
 
-Figure \ref{fig:netcdf} reproduces the ERA5 data row of Figure 19 in \citet{hersbach2020era5}.
+Figure \ref{fig:netcdf} reproduces the ERA5 data \citep{hersbach2020era5} row of Figure 19 in \citep{hersbach2020era5}.
 Here we explain how this would be done using cubble.
 The plots show that the southern polar vortex splits into two on 2002-09-26 and further splits into four on 2002-10-04.
-Further explanation of why this is interesting can be found in the figure source, and also \citet{simmons2020global}, \citet{simmons2005ecmwf}.
+Further explanation of why this is interesting can be found in the figure source, and also \citep{simmons2020global}, \citep{simmons2005ecmwf}.
 
@@
 The data is available in the NetCDF format from the European Centre for Medium-Range Weather Forecasts (ECMWF).
-It can be directly downloaded from \href{https://cds.climate.copernicus.eu/cdsapp\#!/dataset/reanalysis-era5-pressure-levels?tab=overview}{Copernicus Climate Data Store (CDS)} website or via the R package \pkg{ecmwfr} \citep{ecwmfr}.
+It can be directly downloaded from Copernicus Climate Data Store (CDS) \citep{cds} website or via the R package \pkg{ecmwfr} \citep{ecwmfr}.
 
@@
 That could be achieved with a list-column for the location coordinates, and an additional form that these locations can be pivoted into, like the long form for temporal variables.
-For multiple measured variables, \pkg{cubble} could be integrated with dimension reduction methods, and dynamic graphics for multiple dimensions such as a tour \citep{REF}.
+For multiple measured variables, \pkg{cubble} could be integrated with dimension reduction methods, and dynamic graphics for multiple dimensions such as a tour \citep{wickham_tourr_2011}.
 

### 2022-06-24 4-3-water-level.Rmd
fix figure 7

@@
 Here we examine the river gauge data (\code{Water_course_level}) from the Bureau of Meteorology [@bom] in relation to weather station rainfall from NOAA's climate data [@noaa] (\code{climate}).
-Figure \ref{fig:matching-map} shows the location of available weather stations and water gauges in Victoria.
-The goal is to match water gauges with nearby weather stations, spatially and temporally, using 2020 measurements using
-\code{match_sites()} function.
+The goal is to match water gauges with nearby weather stations, spatially and temporally, using 2020 measurements using \code{match_sites()} function.
 <!--The Bureau of Meteorology collects [water data](http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full) from river gauges.
@@
 
-```{r matching-map, fig.height = 7, fig.width=10, fig.cap="Location of weather stations and river gauges in Victoria, Australia."}
-vic_map <- rmapshaper::ms_simplify(ozmaps::abs_ste %>%
-               filter(NAME == "Victoria"))
-
-ggplot() + 
-  geom_sf(data = vic_map, fill = "grey95", color = "white") + 
-  geom_point(data = dplyr::bind_rows(river, climate), 
-             aes(x = long, y = lat, color = type)) + 
-  ggthemes::theme_map() +
-  theme(legend.position = "bottom") +
-  labs(x = "Longitude", y = "Latitude") + 
-  scale_color_brewer(palette = "Dark2") 
-```
-
 This function requires passing the major and minor data sets used for matching, in this case those are \code{river} and \code{climate}.
@@
   temporal_n_highest = 30,
-  temporal_min_match = 15
+  temporal_min_match = 15, 
 )
@@
 
-Figure \ref{fig:matching} shows four matched pairs on the map (A) and standardised data as time series (B). The expected concurrent increase in precipitation and water level can be seen clearly.
+Figure \ref{fig:matching} shows four matched pairs on the map (a) and standardised data as time series (b). The expected concurrent increase in precipitation and water level can be seen clearly.
 
-```{r matching, out.width="100%", fig.height = 5, fig.width = 10, fig.cap="Matched weather stations and river gauges on the map (A) and across time (B).
+```{r matching, out.width="100%", fig.height = 5, fig.width = 10, fig.cap="Weather stations and river gauges with matched pairs labelled on the map (a) and plotted across time (b).
 Precipitation and water level have been standardised between 0 and 1 to be displayed on the same scale.
@@
 The numbers (1, 5, 6, 10) indicate the group index derived from spatial matching, only those that were selectd by temporal matching are shown here."}
+vic_map <- ozmaps::abs_ste %>% 
+  filter(NAME == "Victoria") %>% 
+  rmapshaper::ms_simplify()
+
 p1 <-ggplot() + 
   geom_sf(data = vic_map, fill = "grey95", color = "white") + 
+  geom_point(data = dplyr::bind_rows(river, climate), 
+             aes(x = long, y = lat, color = type), alpha = 0.2, fill = 0.2) +
   geom_point(data = res, aes(x = long, y = lat, color = type)) +
@@
   patchwork::plot_layout(guides = "collect") + 
-  plot_annotation(tag_levels = 'A')&
+  plot_annotation(tag_levels = 'a')&
   theme(legend.position = "bottom")

### 2022-06-24 jss/cubble-jss.tex
fix figure 7

@@
 Here we examine the river gauge data (\code{Water_course_level}) from the Bureau of Meteorology \citep{bom} in relation to weather station rainfall from NOAA's climate data \citep{noaa} (\code{climate}).
-Figure \ref{fig:matching-map} shows the location of available weather stations and water gauges in Victoria.
-The goal is to match water gauges with nearby weather stations, spatially and temporally, using 2020 measurements using
-\code{match_sites()} function.
-
-\begin{CodeChunk}
-\begin{figure}
-
-{\centering \includegraphics{figures/matching-map-1}
-
-}
-
-\caption[Location of weather stations and river gauges in Victoria, Australia]{Location of weather stations and river gauges in Victoria, Australia.}\label{fig:matching-map}
-\end{figure}
-\end{CodeChunk}
+The goal is to match water gauges with nearby weather stations, spatially and temporally, using 2020 measurements using \code{match_sites()} function.
 
@@
 +   temporal_n_highest = 30,
-+   temporal_min_match = 15
++   temporal_min_match = 15, 
 + )
@@
 
-Figure \ref{fig:matching} shows four matched pairs on the map (A) and standardised data as time series (B). The expected concurrent increase in precipitation and water level can be seen clearly.
+Figure \ref{fig:matching} shows four matched pairs on the map (a) and standardised data as time series (b). The expected concurrent increase in precipitation and water level can be seen clearly.
 
@@
 
-\caption[Matched weather stations and river gauges on the map (A) and across time (B)]{Matched weather stations and river gauges on the map (A) and across time (B).
+\caption[Weather stations and river gauges with matched pairs labelled on the map (a) and plotted across time (b)]{Weather stations and river gauges with matched pairs labelled on the map (a) and plotted across time (b).
 Precipitation and water level have been standardised between 0 and 1 to be displayed on the same scale.

### 2022-06-24 4-1-basic.Rmd
fix figure guide size

@@
   ggthemes::theme_map() +
-  theme(legend.position = "bottom") #+
-  #labs(x = "Longitude", y = "Latitude")
+  theme(legend.position = "bottom",
+        legend.text = element_text(size = 15)) #+
 
@@
   ggimg::geom_point_img(
-    data = single, aes(x = 143.7, y = -30, img = img), size = 6)
+    data = single, aes(x = 143.7, y = -30, img = img), size = 6) + 
+  guides(color = guide_legend(override.aes = list(size=5)))
 

### 2022-06-24 4-5-interactive.Rmd
add labels for subplots in the interactive graphic

@@
 Each row is a screen dump of the process.
-The top row shows all locations and all temperature profiles.
-Selecting a particular location on the map (here Mount Elizabeth) produces the plot in the second row.
-The maximum and minimum temperatures are shown using a ribbon.
-The bottom row first selects the lowest temperature in August in the seasonal display, which highlights the corresponding station on the map (Thredbo AWS).
-Another  station, located in the Tasmania Island, is then selected to compare its temperature variation with Thredbo AWS.", fig.show='hold'}
+The top row shows all locations (a) and all temperature profiles (d).
+Selecting a particular location (Mount Elizabeth) on the map (b) produces the plot in the second row.
+The maximum and minimum temperatures are shown using a ribbon (e).
+The bottom row first selects the lowest temperature in August in the seasonal display (f), which highlights the corresponding station on the map (Thredbo AWS).
+Another station, located in the Tasmania Island, is then selected (c) to compare its temperature variation with Thredbo AWS.", fig.show='hold'}
 knitr::include_graphics(here::here("figures/linking.png"))

### 2022-06-24 jss/appendix.Rmd
add labels for subplots in the interactive graphic

@@
 
-Figure XXX in this section can also be made using `cubble` and `leaflet`, in which case the temperature range is displayed as a small subplot upon clicking on the map.
+Figure 9 in this section can also be made using `cubble` and `leaflet`, in which case the temperature range is displayed as a small subplot upon clicking on the map.
 This procedure involves first creating the popup plots from the long form cubble as a vector and then adding these plots to a leaflet map created from the nested cubble, with `leafpop::addPopupGraphs()`:
@@
 
-Figure \ref{fig:interactive-popup} shows the same information as Figure XXX but using leaflet and popups [@leafpop].
+Figure \ref{fig:interactive-popup} shows the same information as Figure 9 but using leaflet and popups.
 

### 2022-06-24 jss/cubble-jss.tex
add labels for subplots in the interactive graphic

@@
 Each row is a screen dump of the process.
-The top row shows all locations and all temperature profiles.
-Selecting a particular location on the map (here Mount Elizabeth) produces the plot in the second row.
-The maximum and minimum temperatures are shown using a ribbon.
-The bottom row first selects the lowest temperature in August in the seasonal display, which highlights the corresponding station on the map (Thredbo AWS).
-Another  station, located in the Tasmania Island, is then selected to compare its temperature variation with Thredbo AWS.}\label{fig:interactive-linking}
+The top row shows all locations (a) and all temperature profiles (d).
+Selecting a particular location (Mount Elizabeth) on the map (b) produces the plot in the second row.
+The maximum and minimum temperatures are shown using a ribbon (e).
+The bottom row first selects the lowest temperature in August in the seasonal display (f), which highlights the corresponding station on the map (Thredbo AWS).
+Another station, located in the Tasmania Island, is then selected (c) to compare its temperature variation with Thredbo AWS.}\label{fig:interactive-linking}
 \end{figure}

### 2022-06-24 5-1-covid.Rmd
better naming of the chapters

--- 2022-06-22 4-1-basic.Rmd
@@
 ---
-title: "4-examples"
+title: "5-1-covid"
 output: html_document
@@
 ```
-
-## Australian historical maximum temperature {#historicaltmax}
-
-```{r}
-load(here::here("data/historical_tmax.rda"))
-```
-
-The Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world.
-The data used here (\code{historical_tmax}) is a subset extracted using the `rnoaa` package, containing the records of maximum temperature for `r nrow(historical_tmax)` Australian stations from `r range(historical_tmax$first_year)[1]` through `r range(historical_tmax$first_year)[2]` and provides information also on the latitude, longitude and elevation of each of the  stations.
-<!-- This data  is already cast into a \code{cubble} data object (please see Appendix X for the code to extract  the data from XX and prepare the data into \code{cubble} form), with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates.--> The goal of this example is to compare the monthly average maximum temperature between two periods, 1971-1975 and 2016-2020, for stations in Victoria and New South Wales, using a glyph map.
-
-First, the stations need to be filtered to those in NSW and Victoria by using the station identifiers, stored within the 11 digits of the \code{id} variable entries.
-The country code is in the first 5 digits (Australia is represented by "ASN00") and the next 6 digits encode the station following the [Australian Bureau of Meteorology (BOM)](http://www.bom.gov.au/climate/cdo/about/site-num.shtml) coding protocols.
-New South Wales stations correspond to entries in the range 46-75 and the Victorian stations to 76-90.
-Filtering Victoria and New South Wales stations is a *spatial operation* and hence uses the \code{cubble} nested form:
-
-```{r echo = TRUE}
-tmax <- historical_tmax %>%
-  filter(between(stringr::str_sub(id, 7, 8), 46, 90))
-```
-
-Next, the monthly maximum average temperature is  calculated for both periods.
-This is a *temporal operation* requiring a switch into the long \code{cubble} form using the \code{face_temporal()} function.
-In addition, a new indicator for the two time periods of interest is created before the calculation of monthly averages:
-
-```{r echo = TRUE}
-tmax <- tmax %>%
-  face_temporal() %>% 
-  group_by(month = lubridate::month(date),
-         group = as.factor(
-           ifelse(lubridate::year(date) > 2015,
-           "2016 ~ 2020", "1971 ~ 1975"))) %>%
-  summarise(tmax = mean(tmax, na.rm = TRUE))
-```
-
-A quick check on the number of observations for each location is made, revealing that there are several with less than 24 observations -- these stations lack temperature values for some months.
-In this example, those stations are removed by switching to a long \code{cubble} to operate on the spatial component over time, and then, move back into the nested \code{cubble} (to make the glyph map):
-
-```{r echo = TRUE}
-tmax %>% 
-  face_spatial() %>% 
-  mutate(n = nrow(ts)) %>%
-  arrange(n) %>%
-  pull(n) %>% 
-  head(10)
-
-tmax <- tmax %>% 
-  face_spatial() %>% 
-  filter(nrow(ts) == 24) %>%
-  face_temporal()
-```
-
-In order to create a glyph map displaying the monthly series  (Figure \ref{fig:basic-manip}), the spatial variables need to be unfolded with the temporal variables.
-The reason being that the major (\code{long}, \code{lat}) and minor (\code{month}, \code{tmax}) coordinates need to be on the same table to create the glyph map.
-The \code{geom_glyph()} function does both the transformation and the plotting.
-\textcolor{green}{(An inset of the glyph for one station is shown in the top left corner.)}
-
-```{r eval = FALSE, echo = TRUE}
-nsw_vic <- ozmaps::abs_ste %>%
-  filter(NAME %in% c("Victoria", "New South Wales"))
-
-ggplot() +
-  geom_sf(data = nsw_vic,
-          fill = "transparent", color = "grey",
-          linetype = "dotted") +
-  geom_glyph(data = tmax,
-             aes(x_major = long, x_minor = month,
-                 y_major = lat, y_minor = tmax,
-                 group = interaction(id, group), color = group),
-             width = 1, height = 0.5) +
-  ...
-```
-
-```{r create-inset, eval = FALSE}
-
-tmax %>% filter(id == "ASN00048027") %>%
-  ggplot(aes(x = month,
-             y = tmax,
-             color = group)) +
-  geom_line(size = 1.5) +
-  scale_color_brewer(palette = "Dark2", 
-                     guide = "none") +
-  scale_x_continuous(breaks = seq(1, 12, 1), 
-                     labels = c("J", "F", "M", "A",
-                                "M", "J", "J", "A",
-                                "S", "O", "N", "D")) +
-  labs(x = "", y  = "Temp (C)", 
-       title = "ASN00048027: Cobar") +
-  theme_bw() +
-  theme(
-    aspect.ratio = 0.3,
-    axis.text = element_text(size = 20),
-    title =  element_text(size = 20)
-  )
-
-#ggsave(filename = here::here("figures/basic-manip-inset.png"), width = 7)
-```
-
-```{r basic-manip, out.width="50%", fig.height = 6, fig.width = 5, fig.cap="A glyph map of the monthly maximum average temperature for weather stations in  Victoria and New South Wales for the periods (1971-1975, 2016-2020).
-The corresponding average time series for the Cobar station are display on the top left corner.
-From the glyph map we can observe that the monthly trend is similar for all locations (low in the winter, high in the summer), and small increased temperatures, particularly in late summer can be seen at most stations."}
-tmax <- tmax %>% unfold(long, lat)
-
-nsw_vic <- ozmaps::abs_ste %>% 
-  filter(NAME %in% c("Victoria", "New South Wales"))
-
-p1 <- ggplot() + 
-  geom_sf(data = nsw_vic, fill = "grey95", 
-          color = "white", size=2) + 
-  geom_glyph(data = tmax, 
-             aes(x_major = long, x_minor = month, 
-                 y_major = lat, y_minor = tmax,
-                 group = interaction(id, group), 
-                 color = group),
-             width = 1.2, height = 1, alpha=0.5, 
-             size=1.1) +
-  scale_color_brewer("", palette = "Dark2") + 
-  theme_bw() + 
-  coord_sf(xlim = c(141, 154), ylim = c(-39, -28.5)) + 
-  ggthemes::theme_map() +
-  theme(legend.position = "bottom") #+
-  #labs(x = "Longitude", y = "Latitude")
-
-box_df <- tmax %>% 
-  face_spatial() %>% 
-  filter(id == "ASN00048027")
-single <- tibble::tibble(img = here::here("figures/basic-manip-inset.png"))
-
-p1 +
-  geom_rect(data = box_df,
-            aes(xmin = long - 0.6, xmax = long + 0.6,
-                ymin = lat - 0.12, ymax = lat + 0.35),
-            fill = "transparent", color = "black") +
-  ggimg::geom_point_img(
-    data = single, aes(x = 143.7, y = -30, img = img), size = 6)
-
-```
-
-Glyph maps work well to explore temporal patterns across spatial locations, particularly when the spatial locations are gridded.
-In this example, they are irregularly spaced, which can result in overlapping glyphs obscuring each other.
-To fix this, one could aggregate data from nearby stations.
-An example of this using **cubble** is included in the Appendix.

### 2022-06-24 5-2-aus-hist-max-temp.Rmd
better naming of the chapters

--- 2022-06-24 4-1-basic.Rmd
@@
 ---
-title: "4-examples"
+title: "5-2-aus-hist-max-temp"
 output: html_document
 ---
-
-# Examples {#examples}
-
-The five examples here are chosen to illustrate these aspects of using cubble: creating a \code{cubble} from two COVID data tables with the complication of differing location names, using spatial transformations to make a glyph map of seasonal temperature changes over years, aggregating information spatially to explore precipitation patterns, matching river level data and weather station records for analysis of water supply, reading netCDF format data to reproduce a climate reanalysis plot, and the workflow to create complex interactive linked plots.
-(There is an additional example and figures in the Appendix, and more examples in the package vignettes.)
-
-## Victoria COVID spatio-temporal incidence and spread {#covid}
-
-Since the start of the pandemic, the Victoria State Government in Australia has provided daily COVID counts by local government area (LGA).
-This data can be used to visualise COVID incidence and spread spatially, when combined with map polygon data available from the Australian Bureau of Statistics.
-These different sources need to be combined for the analysis, by matching the LGA names.
-Here is how to do this with \pkg{cubble}, including how to handle mismatches arising from different names of the same LGAs in the two tables.
-The COVID data is store in a `csv` file and looks like:
-
-```{r eval = FALSE}
-library(tidyverse)
-#https://www.coronavirus.vic.gov.au/victorian-coronavirus-covid-19-data
-raw <- read_csv(here::here("data/NCOV_COVID_Cases_by_LGA_Source_20220324.csv"),
-                  col_names = c("date", "postcode", "source", "lga"), skip = 1)
-
-covid <- raw %>%
-  count(date, lga, source, postcode) %>%
-  filter(date >= as.Date("2022-01-01")) %>%
-  filter(source == "Contact with a confirmed case") %>%
-  group_by(date, lga, source) %>%
-  count() %>%
-  ungroup(date) %>%
-  mutate(roll_mean = zoo::rollmean(n, k = 14, fill = NA)) %>%
-  tsibble::as_tsibble(key = lga, index = date) %>%
-  tsibble::fill_gaps(.full = TRUE)
-save(covid, file = here::here("data/covid.rda"))
-```
-
-```{r}
-load(here::here("data/covid.rda"))
-
-lga <- strayr::read_absmap("lga2018") %>%
-  rename(lga = lga_name_2018) %>%
-  filter(state_name_2016 == "Victoria") 
-```
-
-```{r echo = TRUE}
-covid %>% head(5)
-```
-
-and the spatial polygons are an ESRI shapefile as follows:
-
-```{r echo = TRUE}
-lga %>% head(5)
-```
-
-The function \code{as_cubble()} is used to create a \code{cubble} object from the two spatial and temporal tables, and requires specifying the arguments \code{key}, \code{index}, and \code{coords} (as described in Section \@ref(create)).
-It will automatically try to match the sites in both tables using the location names and will show a warning message when there are mismatches, as shown below:
-
-```{r echo = TRUE, warning = TRUE, message= TRUE}
-cb <- as_cubble(
-  list(spatial = lga, temporal = covid),
-  key = lga, index = date, coords = c(cent_long, cent_lat)
-  )
-```
-
-It can be seen that there are two-way mismatches -- LGAs in the COVID data that do not  match with LGAs names in the spatial polygon data, and vice versa.
-The mismatches can be identified by using the \code{output = "unmatch"} argument in the \code{as_cubble()} function: <!--Note that \pkg{cubble} will attempt to pair the unmatched sites, these are shown separately from the ones that cannot be paired.-->
-
-```{r echo = TRUE}
-pair <- as_cubble(
-  list(spatial = lga, temporal = covid),
-  key = lga, index = date, coords = c(cent_long, cent_lat),
-  output = "unmatch"
-  )
-
-pair
-```
-
-With this information both tables can be fixed, to create the desired \code{cubble} object, as follows:
-
-```{r echo = TRUE}
-lga <- lga %>%
-  mutate(lga = ifelse(lga == "Kingston (C) (Vic.)", "Kingston (C)", lga),
-         lga = ifelse(lga == "Latrobe (C) (Vic.)", "Latrobe (C)", lga)) %>%
-  filter(!lga %in% pair$others$spatial)
-
-covid <- covid %>% filter(!lga %in% pair$others$temporal)
-
-cb <- as_cubble(data = list(spatial = lga, temporal = covid),
-                key = lga, index = date, coords = c(cent_long, cent_lat))
-```
 

### 2022-06-24 5-3-river-level-rainfall.Rmd
better naming of the chapters

--- 2022-06-24 4-3-water-level.Rmd
@@
   ggthemes::theme_map() +
-  ggplot2::theme(legend.position = "bottom") +
-  ggplot2::labs(x = "Longitude", y = "Latitude")
+  ggplot2::theme(legend.position = "bottom",
+                 legend.text = element_text(size = 15)) +
+  ggplot2::labs(x = "Longitude", y = "Latitude") + 
+  guides(color = guide_legend(override.aes = list(size=5)))
 

### 2022-06-24 jss/cubble-jss.Rmd
edit in the main text accordingly
Move example 5.3 to appendix + update appendix
better naming of the chapters

@@
 
-```{r child = here::here("2-0-cube.Rmd")}
+```{r child = here::here("2-cube.Rmd")}
 
@@
 
-```{r child = here::here("2-the-cubble-pkg.Rmd")}
+```{r child = here::here("3-the-cubble-pkg.Rmd")}
 
@@
 
-```{r child = here::here("3-advanced.Rmd")}
+```{r child = here::here("4-other-features.Rmd")}
 
@@
 
-```{r child = here::here("4-1-basic.Rmd")}
+```{r child = here::here("5-1-covid.Rmd")}
 
@@
 
-```{r child = here::here("4-2-agg.Rmd")}
+```{r child = here::here("5-2-aus-hist-max-temp.Rmd")}
 
@@
 
-```{r child = here::here("4-3-water-level.Rmd")}
+```{r child = here::here("5-3-river-level-rainfall.Rmd")}
 
@@
 
-```{r child = here::here("4-4-netcdf.Rmd")}
+```{r child = here::here("5-4-era5.Rmd")}
 
@@
 
-```{r child = here::here("4-5-interactive.Rmd")}
+```{r child = here::here("5-5-interactive-graphics.Rmd")}
 
@@
 
-```{r child = here::here("5-conclusion.Rmd")}
+```{r child = here::here("6-conclusion.Rmd")}
 

### 2022-06-24 reference-hack.Rmd
better naming of the chapters

---
title: "reference-hack"
output: html_document
---

The `rticles::jss_article` template used during preparation (https://github.com/rstudio/rticles/tree/c085e125d874c11bb474ec53f4b754e179f8a4e9) has the behavior that a reference prints out all the author names when it is used for the first time. See a standalone example [here](https://github.com/huizezhang-sherry/jss).

Here a hack is used to overcome this temporarily

```{r}
tex <- readLines(here::here("references.bib"))

b <- tex %>% stringr::str_extract("\\@(article|techreport|misc|Manual|Book)\\{.+,")
b <- b[!is.na(b)]
ref_head <- b %>% stringr::str_replace("\\@(article|techreport|misc|Manual|Book)\\{", "") %>% 
  stringr::str_replace(",", "")

res <- paste0("@", ref_head, collapse = ", ")
```

Here I reference `r res`

### 2022-06-24 2-the-cubble-pkg.R

@@ -1,0 +1,73 @@
+## -----------------------------------------------------------------------------
+cubble::climate_flat
+
+
+## ----echo = TRUE--------------------------------------------------------------
+cubble_nested <- climate_flat |>
+  as_cubble(key = id, index = date, coords = c(long, lat))
+cubble_nested
+
+
+## ----echo = TRUE--------------------------------------------------------------
+cubble_nested |>
+  mutate(rain_day = sum(ts$prcp != 0))
+
+
+## ----echo = TRUE--------------------------------------------------------------
+cubble_long <- cubble_nested |> face_temporal()
+cubble_long
+
+
+## ----face, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "An illustration of function \\code{face\\_temporal} and \\code{face\\_spatial} in cubble. In the first row, \\code{face\\_temporal} switches a cubble from the nested form into the long form and the focus has switched from the spatial aspect (the side face) to the temporal aspect (the front face). In the second row, \\code{face\\_spatial} switches a cubble back to the nested form from the long form and shifts focus back to the spatial aspect."----
+knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.002.png"))
+
+
+## ----echo = TRUE--------------------------------------------------------------
+cubble_back <- cubble_long |> face_spatial()
+cubble_back
+identical(cubble_nested, cubble_back)
+
+
+## ----echo = TRUE--------------------------------------------------------------
+cubble_unfold <- cubble_long |> unfold(long, lat)
+cubble_unfold
+
+
+## ----echo = TRUE--------------------------------------------------------------
+cubble_unfold |> face_spatial() |> face_temporal()
+
+
+## ----echo = TRUE--------------------------------------------------------------
+climate_flat_ts <- climate_flat |> 
+  tsibble::as_tsibble(key = id, index = date)
+climate_flat_cb <-  climate_flat_ts |> 
+  cubble::as_cubble(coords = c(long, lat))
+climate_flat_cb
+
+
+## ----echo = TRUE--------------------------------------------------------------
+# add station-based features in the nested form.
+climate_flat_cb |> 
+  mutate(
+    fabletools::features(
+      ts, tmax, list(tmax_mean = mean, tmax_var = var)
+      )
+  )
+
+
+## ----echo = TRUE, eval = FALSE------------------------------------------------
[…]

### 2022-06-24 3-the-cubble-pkg.Rmd
better naming of the chapters
create separate files for reproduciblity

--- 2022-06-17 2-the-cubble-pkg.Rmd
@@
 
-```{r face, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "An illustration of function \\code{face\\_temporal} and \\code{face\\_spatial} in cubble.
+```{r face, echo = FALSE, fig.align="center", out.height = "35%", out.width = "100%", fig.cap = "An illustration of function \\code{face\\_temporal} and \\code{face\\_spatial} in cubble.
 In the first row, \\code{face\\_temporal} switches a cubble from the nested form into the long form and the focus has switched from the spatial aspect (the side face) to the temporal aspect (the front face).
@@
 cubble_unfold
-```
-
-This function should generally be used in the last step of the analysis since it is a temporary operation, meaning that these added spatial variables will disappear if switched to the nested form and then switched back:
-
-```{r echo = TRUE}
-cubble_unfold |> face_spatial() |> face_temporal()
 ```
@@
 ```{r eval = FALSE, echo = TRUE}
-# Assume my_ncdf has a bounding box of [-180, -90, 180, -90]
+# Assume my_ncdf has a bounding box of [-180, -90, 180, 90]
 # at 0.25 degree resolution and subset it to have

### 2022-06-24 4-other-features.Rmd
create separate files for reproduciblity

--- 2022-06-17 3-advanced.Rmd
@@
 
-```{r illu-interactive, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "Linking between multiple plots.
+```{r illu-interactive, echo = FALSE, fig.align="center", out.height = "35%", out.width = "100%", fig.cap = "Linking between multiple plots.
 The line plots and the map are constructed from shared crosstalk objects (long and nested cubbles).
@@
 
-An example using a glyph map is shown in Section \ref{precip}.
+An example using a glyph map is shown in Section \@ref(historicaltmax).
 

### 2022-06-24 5-2-aus-hist-max-temp.Rmd
create separate files for reproduciblity

@@
 The \code{geom_glyph()} function does both the transformation and the plotting.
-\textcolor{green}{(An inset of the glyph for one station is shown in the top left corner.)}
 
@@
 
-```{r create-inset, eval = FALSE}
-
-tmax %>% filter(id == "ASN00048027") %>%
-  ggplot(aes(x = month,
-             y = tmax,
-             color = group)) +
-  geom_line(size = 1.5) +
-  scale_color_brewer(palette = "Dark2", 
-                     guide = "none") +
-  scale_x_continuous(breaks = seq(1, 12, 1), 
+```{r}
+cobar <- tmax %>% filter(id == "ASN00048027") %>%
+  ggplot(aes(x = month, y = tmax, color = group)) +
+  geom_line() +
+  scale_color_brewer(palette = "Dark2", guide = "none") +
+  scale_x_continuous(
+    breaks = seq(1, 12, 1), 
     labels = c("J", "F", "M", "A",
                "M", "J", "J", "A",
-                                "S", "O", "N", "D")) +
-  labs(x = "", y  = "Temp (C)", 
-       title = "ASN00048027: Cobar") +
+               "S", "O", "N", "D")
+    ) +
+  labs(x = "", y  = "Temp (C)") +
   theme_bw() + 
-  theme(
-    aspect.ratio = 0.3,
-    axis.text = element_text(size = 20),
-    title =  element_text(size = 20)
-  )
+  theme(aspect.ratio = 0.3)
 
-#ggsave(filename = here::here("figures/basic-manip-inset.png"), width = 7)
 ```
 
-```{r basic-manip, out.width="50%", fig.height = 6, fig.width = 5, fig.cap="A glyph map of the monthly maximum average temperature for weather stations in  Victoria and New South Wales for the periods (1971-1975, 2016-2020).
+```{r basic-manip, fig.height=6, fig.cap="A glyph map of the monthly maximum average temperature for weather stations in  Victoria and New South Wales for the periods (1971-1975, 2016-2020).
 The corresponding average time series for the Cobar station are display on the top left corner.
-From the glyph map we can observe that the monthly trend is similar for all locations (low in the winter, high in the summer), and small increased temperatures, particularly in late summer can be seen at most stations."}
+From the glyph map we can observe that the monthly trend is similar for all locations (low in the winter, high in the summer), and small increased temperatures, particularly in late summer can be seen at most stations in New South Wales."}
 tmax <- tmax %>% unfold(long, lat)
-
-nsw_vic <- ozmaps::abs_ste %>% 
-  filter(NAME %in% c("Victoria", "New South Wales"))
+box_df <- tmax %>% face_spatial() %>% filter(id == "ASN00048027")
+nsw_vic <- ozmaps::abs_ste %>% filter(NAME %in% c("Victoria","New South Wales"))
 
@@
           color = "white", size=2) + 
-  geom_glyph(data = tmax, 
+  geom_glyph(
+    data = tmax, 
     aes(x_major = long, x_minor = month, 
         y_major = lat, y_minor = tmax,
-                 group = interaction(id, group), 
-                 color = group),
-             width = 1.2, height = 1, alpha=0.5, 
-             size=1.1) +
-  scale_color_brewer("", palette = "Dark2") + 
-  theme_bw() + 
-  coord_sf(xlim = c(141, 154), ylim = c(-39, -28.5)) + 
-  ggthemes::theme_map() +
-  theme(legend.position = "bottom",
-        legend.text = element_text(size = 15)) #+
-
-box_df <- tmax %>% 
-  face_spatial() %>% 
-  filter(id == "ASN00048027")
-single <- tibble::tibble(img = here::here("figures/basic-manip-inset.png"))
-
-p1 +
-  geom_rect(data = box_df,
+        group = interaction(id, group), color = group),
+    width = 1.2, height = 0.8) +
+    geom_rect(
+      data = box_df,
       aes(xmin = long - 0.6, xmax = long + 0.6,
@@
       fill = "transparent", color = "black") +
-  ggimg::geom_point_img(
-    data = single, aes(x = 143.7, y = -30, img = img), size = 6) + 
-  guides(color = guide_legend(override.aes = list(size=5)))
+  scale_color_brewer("", palette = "Dark2") + 
+  coord_sf(xlim = c(141, 154), ylim = c(-39, -28.5)) + 
+  ggthemes::theme_map() +
+  theme(legend.position = "bottom", legend.text = element_text(size = 15)) + 
+  guides(color = guide_legend(override.aes = list(size=2)))
+
+(cobar / p1)  + 
+  patchwork::plot_layout(heights = c(1,5)) + 
+  plot_annotation(tag_levels = 'a')
 

### 2022-06-24 5-3-river-level-rainfall.Rmd
create separate files for reproduciblity

--- 2022-06-24 4-3-water-level.Rmd
@@
   ggthemes::theme_map() +
-  ggplot2::theme(legend.position = "bottom") +
-  ggplot2::labs(x = "Longitude", y = "Latitude")
+  ggplot2::theme(legend.position = "bottom",
+                 legend.text = element_text(size = 15),
+                 legend.title = element_text(size = 15)) +
+  ggplot2::labs(x = "Longitude", y = "Latitude") + 
+  guides(color = guide_legend(override.aes = list(size=5)))
 

### 2022-06-24 5-4-era5.Rmd
create separate files for reproduciblity

--- 2022-06-24 4-4-netcdf.Rmd
@@
 
-```{r eval = FALSE, echo = TRUE}
+```{r echo = TRUE}
 raw <- ncdf4::nc_open(here::here("data/era5-pressure.nc"))
-dt <- as_cubble(raw, vars = c("q", "z"))
+dt <- as_cubble(
+  raw, vars = c("q", "z"),
+  long_range = seq(-180, 180, 1), lat_range = seq(-90, 90, 1))
 ```
@@
 
-```{r netcdf, out.width="100%", fig.cap = "A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020) to illustrate the break-up of sourthern polar vortex in late September and early October 2002.
+```{r eval = FALSE}
+# prepare data/era5-pressure.nc
+library(ecmwfr)
+my_request <- list(
+  product_type = "reanalysis",
+  format = "netcdf",
+  pressure_level = "10",
+  year = "2002",
+  month = c("09", "10"),
+  day = c("04", "22", "26", "30"),
+  time = "00:00",
+  area = c(-5, -180, -90, 180),
+  variable = c("specific_humidity", 'geopotential'),
+  dataset_short_name = "reanalysis-era5-pressure-levels",
+  target = "era5-pressure.nc"
+)
+# 
+# # here you need to create an account on Climate Data Store 
+# # at https://cds.climate.copernicus.eu/cdsapp#!/home) to obtain the user ID and API key
+# # then set up the with ecmwfr::wf_set_key(user, key, service = "cds")
+# # the user below is the same `user` in the setup line
+wf_request(
+  user = "113339",
+  request = my_request,
+  transfer = TRUE,
+  path = here::here("data/"))
+```
+
+```{r netcdf,  fig.height = 4, fig.width = 10, fig.cap = "A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020) to illustrate the break-up of sourthern polar vortex in late September and early October 2002.
 The polar vortex, signalled by the high speicfic humidity, splits into two on 2002-09-26 and furthers split into four on 2002-10-04."}
-knitr::include_graphics(here::here("figures/netcdf.png"))
+
+date <- c("2002-09-22", "2002-09-26", "2002-09-30", "2002-10-04") %>% as.Date()
+res <- dt %>% 
+  face_temporal() %>% 
+  filter(lubridate::date(time) %in% date) %>%
+  unfold(long, lat) %>% 
+  mutate(q =  q* 10^6)
+
+box = c(xmin = -180, ymin = -90, xmax = 180, ymax = -5)
+sf::sf_use_s2(FALSE)
+country <- rnaturalearth::ne_coastline("small", returnclass = "sf") %>% 
+  sf::st_crop(box) %>% 
+  sf::st_cast("MULTILINESTRING")
+
+res %>% 
+  ggplot() +
+  # q for specific humidity
+  geom_point(aes(x = long, y = lat, color = q)) +
+  # z for geopotential
+  geom_contour(data = res, aes(x = long, y = lat, z = z),
+               color = "grey20", binwidth = 4000, size = 0.5) +
+  geom_sf(data = country , alpha = 0.5, fill = "transparent", color = "lightgreen") +
+  coord_sf(
+    crs = "+proj=stere +lat_0=-90 +lon_0=-180 +k=1 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs ", default_crs = sf::st_crs(4326), label_graticule = "EW") + 
+  theme_bw() + 
+  theme(legend.position = "bottom",
+        axis.ticks = element_blank(),
+        panel.border = element_blank(),
+        axis.text = element_blank(), 
+        panel.grid = element_blank()) + 
+  facet_wrap(vars(as.Date(time)), nrow = 1, shrink = TRUE)+  
+  colorspace::scale_color_continuous_sequential("Purple-Yellow", name = "Specific humidity") + 
+  labs(x ="", y = "")
+
 ```

### 2022-06-24 5-5-interactive-graphics.Rmd
create separate files for reproduciblity

--- 2022-06-22 4-5-interactive.Rmd
@@
 
+```{r eval = FALSE}
+# This chunk generates the interactive graphic and save it as figures/linking.html, where the three screencasts in the figure comes from.
+library(crosstalk)
+library(plotly)
+
+clean <- climate_full %>%
+  filter(!id %in% c("ASN00067033", "ASN00072091", "ASN00059040", 
+                    "ASN00097053", "ASN00040856", "ASN00015548")) %>%
+  face_temporal() %>% 
+  mutate(month = lubridate::month(date)) %>%
+  group_by(month) %>%
+  summarise(
+    tmax = mean(tmax, na.rm = TRUE),
+    tmin = mean(tmin, na.rm = TRUE),
+    diff = mean(tmax - tmin, na.rm = TRUE)
+    ) %>% 
+  ungroup(month) %>%
+  mutate(dummy_date = as.Date(glue::glue("2021-{month}-01"))) %>% 
+  face_spatial() %>% 
+  mutate(temp_diff_var = ts %>% pull(diff) %>% var(na.rm = TRUE))
+
+nested <- clean %>% 
+  SharedData$new(~id, group = "cubble")
+
+long <- clean %>% 
+  face_temporal() %>% 
+  unfold(temp_diff_var) %>% 
+  arrange(temp_diff_var) %>% 
+  SharedData$new(~id, group = "cubble")
+
+state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)
+
+p1 <- nested %>% 
+  ggplot() +
+  geom_sf(data = state_map, fill = "grey95", color = "white") +
+  geom_point(aes(x = long, y = lat,color = temp_diff_var, label = name)) +
+  colorspace::scale_color_continuous_sequential(
+    "Rocket",  n_interp = 7, cmax = 90, rev = TRUE, 
+    c2 = 40, l2= 85, c1 = 20, l1 = 30, name = "Var. temp. diff.") +
+  labs(x = "Longitude", y = "Latitude") + 
+  coord_sf() + 
+  theme_bw() + 
+  theme(
+    panel.grid.major = element_blank(),
+    legend.position = "none"
+    )
+
+p2 <- long %>% 
+  ggplot(aes(x = dummy_date, group = id, 
+             fill = temp_diff_var, color = temp_diff_var)) +
+  geom_ribbon(aes(ymin = tmin, ymax = tmax), size = 0.1, alpha = 0.3) + 
+  geom_line(aes(y = tmin), alpha = 0.1) + 
+  geom_line(aes(y = tmax), alpha = 0.1) + 
+  geom_point(aes(y = tmax), size = 0.1) + 
+  geom_point(aes(y = tmin), size = 0.1) + 
+  colorspace::scale_fill_continuous_sequential(
+    "Rocket",  n_interp = 7, cmax = 90, rev = TRUE, 
+    c2 = 40, l2= 85, c1 = 20, l1 = 30, name = "Var. temp. diff.") +
+  colorspace::scale_colour_continuous_sequential(
+    "Rocket",  n_interp = 7, cmax = 90, rev = TRUE, 
+    c2 = 40, l2= 85, c1 = 20, l1 = 30, name = "Var. temp. diff.") +
+  scale_x_date(date_labels = "%b", date_breaks = "1 month") + 
+  labs(x = "Month", y = "Temperature") + 
+  theme_bw() + 
+  theme(
+    panel.grid.major = element_blank(),
+    legend.position = "bottom"
+    )
+
+out <- bscols(
+  ggplotly(p1, width = 900, height = 700, tooltip = "label") %>%
+    highlight(
+      on = "plotly_selected", 
+      off = "plotly_deselect", 
+      opacityDim = 0.1),
+  ggplotly(p2, width = 1200, height = 700, tooltip = "label") %>% 
+    highlight(
+      on = "plotly_selected", 
+      off = "plotly_deselect", 
+      opacityDim = 0.012),
+  widths = c(5, 5)
+)
+
+htmltools::save_html(out, file = "figures/linking.html")
+
+```
+
 ```{r interactive-linking, echo = FALSE, out.width="100%", out.height="23%", fig.retina = 2, dpi = 300, fig.cap = "Exploring temperature variation using linking of a map and seasonal display.
@@
 knitr::include_graphics(here::here("figures/linking-lower.png"))
-#knitr::include_graphics(here::here("figures/linking-south.png"))
 ```

### 2022-06-24 jss/cubble-jss.tex
better naming of the chapters
create separate files for reproduciblity

--- 2022-06-24 jss/cubble-jss.tex
@@
 
-{\centering \includegraphics[width=1\linewidth]{../figures/diagram-keynotes/diagram-keynotes.002}
+{\centering \includegraphics[width=1\linewidth,height=0.35\textheight]{../figures/diagram-keynotes/diagram-keynotes.002}
 
@@
 
-This function should generally be used in the last step of the analysis since it is a temporary operation, meaning that these added spatial variables will disappear if switched to the nested form and then switched back:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> cubble_unfold |> face_spatial() |> face_temporal()
-\end{CodeInput}
-\begin{CodeOutput}
-# cubble:  date, id [5]: long form
-# bbox:    [115.97, -32.94, 133.55, -12.42]
-# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id
-#   [dbl]
-  id          date        prcp  tmax  tmin  long   lat
-  <chr>       <date>     <dbl> <dbl> <dbl> <dbl> <dbl>
-1 ASN00009021 2020-01-01     0  31.9  15.3  116. -31.9
-2 ASN00009021 2020-01-02     0  24.9  16.4  116. -31.9
-3 ASN00009021 2020-01-03     6  23.2  13    116. -31.9
-4 ASN00009021 2020-01-04     0  28.4  12.4  116. -31.9
-5 ASN00009021 2020-01-05     0  35.3  11.6  116. -31.9
-# ... with 1,825 more rows
-\end{CodeOutput}
-\end{CodeChunk}
-
 \hypertarget{tidyverse}{%
@@
 \begin{CodeInput}
-R> # Assume my_ncdf has a bounding box of [-180, -90, 180, -90]
+R> # Assume my_ncdf has a bounding box of [-180, -90, 180, 90]
 R> # at 0.25 degree resolution and subset it to have
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{../figures/diagram-keynotes/diagram-keynotes.004}
+{\centering \includegraphics[width=1\linewidth,height=0.35\textheight]{../figures/diagram-keynotes/diagram-keynotes.004}
 
@@
 
-An example using a glyph map is shown in Section \ref{precip}.
+An example using a glyph map is shown in Section \ref{historicaltmax}.
 
@@
 
-{\centering \includegraphics[width=0.5\linewidth]{figures/basic-manip-1}
+{\centering \includegraphics{figures/basic-manip-1}
 
@@
 The corresponding average time series for the Cobar station are display on the top left corner.
-From the glyph map we can observe that the monthly trend is similar for all locations (low in the winter, high in the summer), and small increased temperatures, particularly in late summer can be seen at most stations.}\label{fig:basic-manip}
+From the glyph map we can observe that the monthly trend is similar for all locations (low in the winter, high in the summer), and small increased temperatures, particularly in late summer can be seen at most stations in New South Wales.}\label{fig:basic-manip}
 \end{figure}
@@
 R> raw <- ncdf4::nc_open(here::here("data/era5-pressure.nc"))
-R> dt <- as_cubble(raw, vars = c("q", "z"))
+R> dt <- as_cubble(
++   raw, vars = c("q", "z"),
++   long_range = seq(-180, 180, 1), lat_range = seq(-90, 90, 1))
 \end{CodeInput}
@@
 
-{\centering \includegraphics[width=1\linewidth]{../figures/netcdf}
+{\centering \includegraphics{figures/netcdf-1}
 

### 2022-06-24 2-the-cubble-pkg.R

@@ -1,73 +1,0 @@
-## -----------------------------------------------------------------------------
-cubble::climate_flat
-
-
-## ----echo = TRUE--------------------------------------------------------------
-cubble_nested <- climate_flat |>
-  as_cubble(key = id, index = date, coords = c(long, lat))
-cubble_nested
-
-
-## ----echo = TRUE--------------------------------------------------------------
-cubble_nested |>
-  mutate(rain_day = sum(ts$prcp != 0))
-
-
-## ----echo = TRUE--------------------------------------------------------------
-cubble_long <- cubble_nested |> face_temporal()
-cubble_long
-
-
-## ----face, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "An illustration of function \\code{face\\_temporal} and \\code{face\\_spatial} in cubble. In the first row, \\code{face\\_temporal} switches a cubble from the nested form into the long form and the focus has switched from the spatial aspect (the side face) to the temporal aspect (the front face). In the second row, \\code{face\\_spatial} switches a cubble back to the nested form from the long form and shifts focus back to the spatial aspect."----
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.002.png"))
-
-
-## ----echo = TRUE--------------------------------------------------------------
-cubble_back <- cubble_long |> face_spatial()
-cubble_back
-identical(cubble_nested, cubble_back)
-
-
-## ----echo = TRUE--------------------------------------------------------------
-cubble_unfold <- cubble_long |> unfold(long, lat)
-cubble_unfold
-
-
-## ----echo = TRUE--------------------------------------------------------------
-cubble_unfold |> face_spatial() |> face_temporal()
-
-
-## ----echo = TRUE--------------------------------------------------------------
-climate_flat_ts <- climate_flat |> 
-  tsibble::as_tsibble(key = id, index = date)
-climate_flat_cb <-  climate_flat_ts |> 
-  cubble::as_cubble(coords = c(long, lat))
-climate_flat_cb
-
-
-## ----echo = TRUE--------------------------------------------------------------
-# add station-based features in the nested form.
-climate_flat_cb |> 
-  mutate(
-    fabletools::features(
-      ts, tmax, list(tmax_mean = mean, tmax_var = var)
-      )
-  )
-
-
-## ----echo = TRUE, eval = FALSE------------------------------------------------
[…]

### 2022-06-24 reproducible/3-the-cubble-pkg.R

@@ -1,0 +1,69 @@
+## -----------------------------------------------------------------------------
+cubble::climate_flat
+
+
+## ----echo = TRUE--------------------------------------------------------------
+cubble_nested <- climate_flat |>
+  as_cubble(key = id, index = date, coords = c(long, lat))
+cubble_nested
+
+
+## ----echo = TRUE--------------------------------------------------------------
+cubble_nested |>
+  mutate(rain_day = sum(ts$prcp != 0))
+
+
+## ----echo = TRUE--------------------------------------------------------------
+cubble_long <- cubble_nested |> face_temporal()
+cubble_long
+
+
+## ----face, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "An illustration of function \\code{face\\_temporal} and \\code{face\\_spatial} in cubble. In the first row, \\code{face\\_temporal} switches a cubble from the nested form into the long form and the focus has switched from the spatial aspect (the side face) to the temporal aspect (the front face). In the second row, \\code{face\\_spatial} switches a cubble back to the nested form from the long form and shifts focus back to the spatial aspect."----
+knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.002.png"))
+
+
+## ----echo = TRUE--------------------------------------------------------------
+cubble_back <- cubble_long |> face_spatial()
+cubble_back
+identical(cubble_nested, cubble_back)
+
+
+## ----echo = TRUE--------------------------------------------------------------
+cubble_unfold <- cubble_long |> unfold(long, lat)
+cubble_unfold
+
+
+## ----echo = TRUE--------------------------------------------------------------
+climate_flat_ts <- climate_flat |> 
+  tsibble::as_tsibble(key = id, index = date)
+climate_flat_cb <-  climate_flat_ts |> 
+  cubble::as_cubble(coords = c(long, lat))
+climate_flat_cb
+
+
+## ----echo = TRUE--------------------------------------------------------------
+# add station-based features in the nested form.
+climate_flat_cb |> 
+  mutate(
+    fabletools::features(
+      ts, tmax, list(tmax_mean = mean, tmax_var = var)
+      )
+  )
+
+
+## ----echo = TRUE, eval = FALSE------------------------------------------------
+## # read in the .nc file as a ncdf4 class
+## raw <- ncdf4::nc_open(here::here("data/era5-pressure.nc"))
+## 
[…]

### 2022-06-24 reproducible/5-1-covid.R

@@ -1,0 +1,62 @@
+## ----eval = FALSE-------------------------------------------------------------
+## library(tidyverse)
+## #https://www.coronavirus.vic.gov.au/victorian-coronavirus-covid-19-data
+## raw <- read_csv(here::here("data/NCOV_COVID_Cases_by_LGA_Source_20220324.csv"),
+##                   col_names = c("date", "postcode", "source", "lga"), skip = 1)
+## 
+## covid <- raw %>%
+##   count(date, lga, source, postcode) %>%
+##   filter(date >= as.Date("2022-01-01")) %>%
+##   filter(source == "Contact with a confirmed case") %>%
+##   group_by(date, lga, source) %>%
+##   count() %>%
+##   ungroup(date) %>%
+##   mutate(roll_mean = zoo::rollmean(n, k = 14, fill = NA)) %>%
+##   tsibble::as_tsibble(key = lga, index = date) %>%
+##   tsibble::fill_gaps(.full = TRUE)
+## save(covid, file = here::here("data/covid.rda"))
+
+
+## -----------------------------------------------------------------------------
+load(here::here("data/covid.rda"))
+
+lga <- strayr::read_absmap("lga2018") %>%
+  rename(lga = lga_name_2018) %>%
+  filter(state_name_2016 == "Victoria") 
+
+
+## ----echo = TRUE--------------------------------------------------------------
+covid %>% head(5)
+
+
+## ----echo = TRUE--------------------------------------------------------------
+lga %>% head(5)
+
+
+## ----echo = TRUE, warning = TRUE, message= TRUE-------------------------------
+cb <- as_cubble(
+  list(spatial = lga, temporal = covid),
+  key = lga, index = date, coords = c(cent_long, cent_lat)
+  )
+
+
+## ----echo = TRUE--------------------------------------------------------------
+pair <- as_cubble(
+  list(spatial = lga, temporal = covid),
+  key = lga, index = date, coords = c(cent_long, cent_lat),
+  output = "unmatch"
+  )
+
+pair
+
+
+## ----echo = TRUE--------------------------------------------------------------
+lga <- lga %>%
+  mutate(lga = ifelse(lga == "Kingston (C) (Vic.)", "Kingston (C)", lga),
+         lga = ifelse(lga == "Latrobe (C) (Vic.)", "Latrobe (C)", lga)) %>%
+  filter(!lga %in% pair$others$spatial)
+
[…]

### 2022-06-24 reproducible/5-2-aus-hist-max-temp.R

@@ -1,0 +1,96 @@
+## -----------------------------------------------------------------------------
+load(here::here("data/historical_tmax.rda"))
+
+
+## ----echo = TRUE--------------------------------------------------------------
+tmax <- historical_tmax %>%
+  filter(between(stringr::str_sub(id, 7, 8), 46, 90))
+
+
+## ----echo = TRUE--------------------------------------------------------------
+tmax <- tmax %>%
+  face_temporal() %>% 
+  group_by(month = lubridate::month(date),
+         group = as.factor(
+           ifelse(lubridate::year(date) > 2015,
+           "2016 ~ 2020", "1971 ~ 1975"))) %>%
+  summarise(tmax = mean(tmax, na.rm = TRUE))
+
+
+## ----echo = TRUE--------------------------------------------------------------
+tmax %>% 
+  face_spatial() %>% 
+  mutate(n = nrow(ts)) %>%
+  arrange(n) %>%
+  pull(n) %>% 
+  head(10)
+
+tmax <- tmax %>% 
+  face_spatial() %>% 
+  filter(nrow(ts) == 24) %>%
+  face_temporal()
+
+
+## ----eval = FALSE, echo = TRUE------------------------------------------------
+## nsw_vic <- ozmaps::abs_ste %>%
+##   filter(NAME %in% c("Victoria", "New South Wales"))
+## 
+## ggplot() +
+##   geom_sf(data = nsw_vic,
+##           fill = "transparent", color = "grey",
+##           linetype = "dotted") +
+##   geom_glyph(data = tmax,
+##              aes(x_major = long, x_minor = month,
+##                  y_major = lat, y_minor = tmax,
+##                  group = interaction(id, group), color = group),
+##              width = 1, height = 0.5) +
+##   ...
+
+
+## -----------------------------------------------------------------------------
+cobar <- tmax %>% filter(id == "ASN00048027") %>%
+  ggplot(aes(x = month, y = tmax, color = group)) +
+  geom_line() +
+  scale_color_brewer(palette = "Dark2", guide = "none") +
+  scale_x_continuous(
+    breaks = seq(1, 12, 1), 
+    labels = c("J", "F", "M", "A",
+               "M", "J", "J", "A",
+               "S", "O", "N", "D")
+    ) +
+  labs(x = "", y  = "Temp (C)") +
+  theme_bw() + 
+  theme(aspect.ratio = 0.3)
+
+
+
[…]

### 2022-06-24 reproducible/5-3-river-level-rainfall.R

@@ -1,0 +1,67 @@
+## -----------------------------------------------------------------------------
+load(here::here("data/climate_full.rda"))
+
+climate <- climate_full %>%
+  filter(between(stringr::str_sub(id, 7, 8), 76, 90)) %>%
+  face_temporal() %>%
+  filter(lubridate::year(date) == 2020) %>%
+  face_spatial() %>%
+  mutate(type = "climate")
+
+river <- river %>% mutate(type = "river")
+
+
+## ----echo = TRUE--------------------------------------------------------------
+res <- match_sites(
+  river, climate,
+  temporal_by = c("Water_course_level" = "prcp"),
+  temporal_independent = "prcp",  
+  temporal_n_highest = 30,
+  temporal_min_match = 15, 
+)
+
+
+## -----------------------------------------------------------------------------
+res %>% print()
+
+
+## ----matching, out.width="100%", fig.height = 5, fig.width = 10, fig.cap="Weather stations and river gauges with matched pairs labelled on the map (a) and plotted across time (b). Precipitation and water level have been standardised between 0 and 1 to be displayed on the same scale. The water level reflects the increase in precipitation. The numbers (1, 5, 6, 10) indicate the group index derived from spatial matching, only those that were selectd by temporal matching are shown here."----
+vic_map <- ozmaps::abs_ste %>% 
+  filter(NAME == "Victoria") %>% 
+  rmapshaper::ms_simplify()  
+
+p1 <-ggplot() + 
+  geom_sf(data = vic_map, fill = "grey95", color = "white") + 
+  geom_point(data = dplyr::bind_rows(river, climate), 
+             aes(x = long, y = lat, color = type), alpha = 0.2, fill = 0.2) +
+  geom_point(data = res, aes(x = long, y = lat, color = type)) +
+  ggrepel::geom_label_repel(
+    data = res %>% filter(type == "river"), aes(x = long, y = lat, label = group)) +
+  scale_color_brewer(palette = "Dark2")  + 
+  ggthemes::theme_map() +
+  ggplot2::theme(legend.position = "bottom",
+                 legend.text = element_text(size = 15),
+                 legend.title = element_text(size = 15)) +
+  ggplot2::labs(x = "Longitude", y = "Latitude") + 
[…]

### 2022-06-24 reproducible/5-4-era5.R

@@ -1,0 +1,71 @@
+## ----echo = TRUE--------------------------------------------------------------
+raw <- ncdf4::nc_open(here::here("data/era5-pressure.nc"))
+dt <- as_cubble(
+  raw, vars = c("q", "z"),
+  long_range = seq(-180, 180, 1), lat_range = seq(-90, 90, 1))
+
+
+## ----eval = FALSE-------------------------------------------------------------
+## # prepare data/era5-pressure.nc
+## library(ecmwfr)
+## my_request <- list(
+##   product_type = "reanalysis",
+##   format = "netcdf",
+##   pressure_level = "10",
+##   year = "2002",
+##   month = c("09", "10"),
+##   day = c("04", "22", "26", "30"),
+##   time = "00:00",
+##   area = c(-5, -180, -90, 180),
+##   variable = c("specific_humidity", 'geopotential'),
+##   dataset_short_name = "reanalysis-era5-pressure-levels",
+##   target = "era5-pressure.nc"
+## )
+## #
+## # # here you need to create an account on Climate Data Store
+## # # at https://cds.climate.copernicus.eu/cdsapp#!/home) to obtain the user ID and API key
+## # # then set up the with ecmwfr::wf_set_key(user, key, service = "cds")
+## # # the user below is the same `user` in the setup line
+## wf_request(
+##   user = "113339",
+##   request = my_request,
+##   transfer = TRUE,
+##   path = here::here("data/"))
+
+
+## ----netcdf,  fig.height = 4, fig.width = 10, fig.cap = "A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020) to illustrate the break-up of sourthern polar vortex in late September and early October 2002. The polar vortex, signalled by the high speicfic humidity, splits into two on 2002-09-26 and furthers split into four on 2002-10-04."----
+
+date <- c("2002-09-22", "2002-09-26", "2002-09-30", "2002-10-04") %>% as.Date()
+res <- dt %>% 
+  face_temporal() %>% 
+  filter(lubridate::date(time) %in% date) %>%
+  unfold(long, lat) %>% 
+  mutate(q =  q* 10^6)
+
+box = c(xmin = -180, ymin = -90, xmax = 180, ymax = -5)
+sf::sf_use_s2(FALSE)
+country <- rnaturalearth::ne_coastline("small", returnclass = "sf") %>% 
[…]

### 2022-06-24 reproducible/5-5-interactive-graphics.R

@@ -1,0 +1,92 @@
+## ----eval = FALSE-------------------------------------------------------------
+## # This chunk generates the interactive graphic and save it as figures/linking.html, where the three screencasts in the figure comes from.
+## library(crosstalk)
+## library(plotly)
+## 
+## clean <- climate_full %>%
+##   filter(!id %in% c("ASN00067033", "ASN00072091", "ASN00059040",
+##                     "ASN00097053", "ASN00040856", "ASN00015548")) %>%
+##   face_temporal() %>%
+##   mutate(month = lubridate::month(date)) %>%
+##   group_by(month) %>%
+##   summarise(
+##     tmax = mean(tmax, na.rm = TRUE),
+##     tmin = mean(tmin, na.rm = TRUE),
+##     diff = mean(tmax - tmin, na.rm = TRUE)
+##     ) %>%
+##   ungroup(month) %>%
+##   mutate(dummy_date = as.Date(glue::glue("2021-{month}-01"))) %>%
+##   face_spatial() %>%
+##   mutate(temp_diff_var = ts %>% pull(diff) %>% var(na.rm = TRUE))
+## 
+## nested <- clean %>%
+##   SharedData$new(~id, group = "cubble")
+## 
+## long <- clean %>%
+##   face_temporal() %>%
+##   unfold(temp_diff_var) %>%
+##   arrange(temp_diff_var) %>%
+##   SharedData$new(~id, group = "cubble")
+## 
+## state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)
+## 
+## p1 <- nested %>%
+##   ggplot() +
+##   geom_sf(data = state_map, fill = "grey95", color = "white") +
+##   geom_point(aes(x = long, y = lat,color = temp_diff_var, label = name)) +
+##   colorspace::scale_color_continuous_sequential(
+##     "Rocket",  n_interp = 7, cmax = 90, rev = TRUE,
+##     c2 = 40, l2= 85, c1 = 20, l1 = 30, name = "Var. temp. diff.") +
+##   labs(x = "Longitude", y = "Latitude") +
+##   coord_sf() +
+##   theme_bw() +
+##   theme(
+##     panel.grid.major = element_blank(),
+##     legend.position = "none"
+##     )
+## 
+## p2 <- long %>%
+##   ggplot(aes(x = dummy_date, group = id,
[…]

### 2022-06-24 scripts/interactive-linking.R

@@ -1,68 +1,0 @@
-library(cubble)
-library(tidyverse)
-library(crosstalk)
-library(plotly)
-
-clean <- climate_full %>%
-  filter(!id %in% c("ASN00067033", "ASN00072091", "ASN00059040", 
-                    "ASN00097053", "ASN00040856", "ASN00015548")) %>%
-  face_temporal() %>% 
-  mutate(month = lubridate::month(date)) %>%
-  group_by(month) %>%
-  summarise(tmax = mean(tmax, na.rm = TRUE),
-            tmin = mean(tmin, na.rm = TRUE),
-            diff = mean(tmax - tmin, na.rm = TRUE)) %>% 
-  ungroup(month) %>%
-  mutate(dummy_date = as.Date(glue::glue("2021-{month}-01"))) %>% 
-  face_spatial() %>% 
-  mutate(temp_diff_var = ts %>% pull(diff) %>% var(na.rm = TRUE))
-  
-
-nested <- clean %>% 
-  SharedData$new(~id, group = "cubble")
-
-long <- clean %>% face_temporal() %>% unfold(temp_diff_var) %>% arrange(temp_diff_var) %>% 
-  SharedData$new(~id, group = "cubble")
-
-state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)
-p1 <- nested %>% 
-  ggplot() +
-  geom_sf(data = state_map, aes(geometry = geometry), fill = "transparent", color = "grey50") +
-  geom_point(aes(x = long, y = lat,color = temp_diff_var, label = name)) +
-  colorspace::scale_color_continuous_sequential(
-    "Rocket",  n_interp = 7, cmax = 90, rev = TRUE, c2 = 40, l2= 85, c1 = 20, l1 = 30,
-    name = "Var. temp. diff.") +
-  labs(x = "Longitude", y = "Latitude") + 
-  coord_sf() + 
-  theme_bw() + 
-  theme(panel.grid.major = element_blank(),
-        legend.position = "none")
-
-p2 <- long %>% 
-  ggplot(aes(x = dummy_date, group = id, fill = temp_diff_var, color = temp_diff_var)) +
-  geom_ribbon(aes(ymin = tmin, ymax = tmax), size = 0.1, alpha = 0.3) + 
-  geom_line(aes(y = tmin), alpha = 0.1) + 
-  geom_line(aes(y = tmax), alpha = 0.1) + 
[…]

### 2022-06-24 scripts/netcdf.R

@@ -1,64 +1,0 @@
-library(tidyverse)
-library(cubble)
-library(ecmwfr)
-
-my_request <- list(
-  product_type = "reanalysis",
-  format = "netcdf",
-  pressure_level = "10",
-  year = "2002",
-  month = c("09", "10"),
-  day = c("04", "22", "26", "30"),
-  time = "00:00",
-  area = c(-5, -180, -90, 180),
-  variable = c("specific_humidity", 'geopotential'),
-  dataset_short_name = "reanalysis-era5-pressure-levels",
-  target = "era5-pressure.nc"
-)
-
-# here you need to create an account on Climate Data Store 
-# at https://cds.climate.copernicus.eu/cdsapp#!/home) to obtain the user ID and API key
-# then set up the with ecmwfr::wf_set_key(user, key, service = "cds")
-# the user below is the same `user` in the setup line
-wf_request(
-  user = "113339",
-  request = my_request,
-  transfer = TRUE,
-  path = "~/Documents/research/paper-cubble/data/")
-
-raw <- ncdf4::nc_open(here::here("data/era5-pressure.nc"))
-dt <- as_cubble(raw, vars = c("q", "z"))
-
-date <- c("2002-09-22", "2002-09-26", "2002-09-30", "2002-10-04") %>% as.Date()
-res <- dt %>% 
-  face_temporal() %>% 
-  filter(lubridate::date(time) %in% date) %>%
-  unfold(long, lat) %>% 
-  mutate(q =  q* 10^6)
-  
-box = c(xmin = -180, ymin = -90, xmax = 180, ymax = -5)
-sf::sf_use_s2(FALSE)
-country <- rnaturalearth::ne_coastline("small", returnclass = "sf") %>% 
-  sf::st_crop(box) %>% 
-  sf::st_cast("MULTILINESTRING")
-
-res %>% 
-  ggplot() +
-  # q for specific humidity
-  geom_point(aes(x = long, y = lat, color = q)) +
-  # z for geopotential
-  geom_contour(data = res, aes(x = long, y = lat, z = z),
-               color = "grey20", binwidth = 4000, size = 0.5) +
-  geom_sf(data = country , alpha = 0.5, fill = "transparent", color = "lightgreen") +
[…]

### 2022-06-26 cubble-jss.R
generate script all in one

@@ -1,0 +1,549 @@
+## ----setup, echo = FALSE------------------------------------------------------
+knitr::opts_chunk$set(
+  echo = FALSE,
+  warning = FALSE,
+  message = FALSE,
+  cache = TRUE,
+  fig.path = "figures/"
+)
+
+options(prompt = "R> ", continue = "+ ",
+        tibble.print_max = 5, tibble.print_min = 5,
+        width = 70)
+
+
+## ----echo = FALSE-------------------------------------------------------------
+library(cubble)
+library(tidyverse)
+library(patchwork)
+
+
+## ----child = here::here("1-intro.Rmd")----------------------------------------
+
+
+
+
+## ----child = here::here("2-cube.Rmd")-----------------------------------------
+
+## ----cube-def, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "An illustration of the conceptual spatio-temporal cube with different slicing on time, site, and variable. For each axis, the slicing can be on a single value or a set of values."----
+knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.001.png"))
+
+
+
+## ----child = here::here("3-the-cubble-pkg.Rmd")-------------------------------
+
+## -----------------------------------------------------------------------------
+cubble::climate_flat
+
+
+## ----echo = TRUE--------------------------------------------------------------
+cubble_nested <- climate_flat |>
+  as_cubble(key = id, index = date, coords = c(long, lat))
+cubble_nested
+
+
+## ----echo = TRUE--------------------------------------------------------------
+cubble_nested |>
+  mutate(rain_day = sum(ts$prcp != 0))
+
+
+## ----echo = TRUE--------------------------------------------------------------
+cubble_long <- cubble_nested |> face_temporal()
+cubble_long
+
+
[…]

### 2022-06-26 reproducible/3-the-cubble-pkg.R

@@ -1,69 +1,0 @@
-## -----------------------------------------------------------------------------
-cubble::climate_flat
-
-
-## ----echo = TRUE--------------------------------------------------------------
-cubble_nested <- climate_flat |>
-  as_cubble(key = id, index = date, coords = c(long, lat))
-cubble_nested
-
-
-## ----echo = TRUE--------------------------------------------------------------
-cubble_nested |>
-  mutate(rain_day = sum(ts$prcp != 0))
-
-
-## ----echo = TRUE--------------------------------------------------------------
-cubble_long <- cubble_nested |> face_temporal()
-cubble_long
-
-
-## ----face, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "An illustration of function \\code{face\\_temporal} and \\code{face\\_spatial} in cubble. In the first row, \\code{face\\_temporal} switches a cubble from the nested form into the long form and the focus has switched from the spatial aspect (the side face) to the temporal aspect (the front face). In the second row, \\code{face\\_spatial} switches a cubble back to the nested form from the long form and shifts focus back to the spatial aspect."----
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.002.png"))
-
-
-## ----echo = TRUE--------------------------------------------------------------
-cubble_back <- cubble_long |> face_spatial()
-cubble_back
-identical(cubble_nested, cubble_back)
-
-
-## ----echo = TRUE--------------------------------------------------------------
-cubble_unfold <- cubble_long |> unfold(long, lat)
-cubble_unfold
-
-
-## ----echo = TRUE--------------------------------------------------------------
-climate_flat_ts <- climate_flat |> 
-  tsibble::as_tsibble(key = id, index = date)
-climate_flat_cb <-  climate_flat_ts |> 
-  cubble::as_cubble(coords = c(long, lat))
-climate_flat_cb
-
-
-## ----echo = TRUE--------------------------------------------------------------
-# add station-based features in the nested form.
-climate_flat_cb |> 
-  mutate(
-    fabletools::features(
-      ts, tmax, list(tmax_mean = mean, tmax_var = var)
-      )
-  )
-
-
-## ----echo = TRUE, eval = FALSE------------------------------------------------
-## # read in the .nc file as a ncdf4 class
-## raw <- ncdf4::nc_open(here::here("data/era5-pressure.nc"))
-## 
[…]

### 2022-06-26 reproducible/5-1-covid.R

@@ -1,62 +1,0 @@
-## ----eval = FALSE-------------------------------------------------------------
-## library(tidyverse)
-## #https://www.coronavirus.vic.gov.au/victorian-coronavirus-covid-19-data
-## raw <- read_csv(here::here("data/NCOV_COVID_Cases_by_LGA_Source_20220324.csv"),
-##                   col_names = c("date", "postcode", "source", "lga"), skip = 1)
-## 
-## covid <- raw %>%
-##   count(date, lga, source, postcode) %>%
-##   filter(date >= as.Date("2022-01-01")) %>%
-##   filter(source == "Contact with a confirmed case") %>%
-##   group_by(date, lga, source) %>%
-##   count() %>%
-##   ungroup(date) %>%
-##   mutate(roll_mean = zoo::rollmean(n, k = 14, fill = NA)) %>%
-##   tsibble::as_tsibble(key = lga, index = date) %>%
-##   tsibble::fill_gaps(.full = TRUE)
-## save(covid, file = here::here("data/covid.rda"))
-
-
-## -----------------------------------------------------------------------------
-load(here::here("data/covid.rda"))
-
-lga <- strayr::read_absmap("lga2018") %>%
-  rename(lga = lga_name_2018) %>%
-  filter(state_name_2016 == "Victoria") 
-
-
-## ----echo = TRUE--------------------------------------------------------------
-covid %>% head(5)
-
-
-## ----echo = TRUE--------------------------------------------------------------
-lga %>% head(5)
-
-
-## ----echo = TRUE, warning = TRUE, message= TRUE-------------------------------
-cb <- as_cubble(
-  list(spatial = lga, temporal = covid),
-  key = lga, index = date, coords = c(cent_long, cent_lat)
-  )
-
-
-## ----echo = TRUE--------------------------------------------------------------
-pair <- as_cubble(
-  list(spatial = lga, temporal = covid),
-  key = lga, index = date, coords = c(cent_long, cent_lat),
-  output = "unmatch"
-  )
-
-pair
-
-
-## ----echo = TRUE--------------------------------------------------------------
-lga <- lga %>%
-  mutate(lga = ifelse(lga == "Kingston (C) (Vic.)", "Kingston (C)", lga),
-         lga = ifelse(lga == "Latrobe (C) (Vic.)", "Latrobe (C)", lga)) %>%
-  filter(!lga %in% pair$others$spatial)
-
[…]

### 2022-06-26 jss/appendix.Rmd
script on appendix

@@
 library(patchwork)
+library(leaflet)
+library(leafpop)
 ```
@@
 
-```{r echo = TRUE}
+```{r eval = FALSE, echo = TRUE}
 cluster_nested <- station_nested %>% switch_key(cluster)
@@
 ```{r eval = FALSE, echo = FALSE}
+# The kmeans clustering in station_nested data can takes some time to run, 
+# This chunk runs the algorithm and save the result as staiton_nested for further use
 coords <- cbind(prcp_aus$long, prcp_aus$lat)
@@
 
-Figure 9 in this section can also be made using `cubble` and `leaflet`, in which case the temperature range is displayed as a small subplot upon clicking on the map.
+Figure \ref{fig:interactive-popup} in this section can also be made using `cubble` and `leaflet`, in which case the temperature range is displayed as a small subplot upon clicking on the map.
 This procedure involves first creating the popup plots from the long form cubble as a vector and then adding these plots to a leaflet map created from the nested cubble, with `leafpop::addPopupGraphs()`:
@@
 
-Figure \ref{fig:interactive-popup} shows the same information as Figure 9 but using leaflet and popups.
+Figure \ref{fig:interactive-popup} shows the same information as Figure 9 but using leaflet and popups [@leafpop].
+
+```{r eval = FALSE, echo = FALSE}
+clean <-  climate_full %>%
+  filter(!id %in% c("ASN00067033", "ASN00072091", "ASN00059040", 
+                    "ASN00097053", "ASN00040856", "ASN00015548")) %>%
+  face_temporal() %>%
+  mutate(month = lubridate::month(date)) %>% 
+  group_by(month) %>% 
+  summarise(tmax = mean(tmax, na.rm = TRUE),
+            tmin = mean(tmin, na.rm = TRUE),
+            diff = mean(tmax - tmin, na.rm = TRUE)) %>% 
+  ungroup(month) %>% 
+  face_spatial() %>%
+  mutate(temp_diff_var = ts %>% pull(diff) %>% var(na.rm = TRUE)) %>% 
+  face_temporal() %>% 
+  mutate(dummy_date = as.Date(glue::glue("2021-{month}-01"))) %>% 
+  unfold(name)
+
+df_id <- clean$id %>% unique()
+p <- purrr::map(1:length(df_id), function(i){
+  dt <- clean %>% filter(id == df_id[i])
+  dt %>% 
+    ggplot(aes(x = dummy_date, ymin = tmin, ymax = tmax, group = id)) + 
+    geom_ribbon(color = "grey50", fill = "grey50", alpha = 0.5) + 
+    theme_bw() + 
+    scale_x_date(date_labels = "%b", date_breaks = "1 month") + 
+    labs(x = "Month", y = "Temperature") + 
+    ylim(-10, 50) + 
+    ggtitle(paste0(stringr::str_to_title(unique(dt$name))))
+})
+
+nested <- face_spatial(clean) 
+domain <- nested$temp_diff_var
+pal <- colorNumeric(
+  colorspace::sequential_hcl(
+    "Rocket",  n = 7, cmax = 90, rev = TRUE, 
+    c2 = 40, l2= 85, c1 = 20, l1 = 30),
+  domain = domain)
+
+leaflet(nested) %>% 
+  addTiles() %>% 
+  addCircleMarkers(color = ~pal(domain), radius = 0.1, 
+                   fillOpacity = 1, opacity = 1, 
+                   popup = popupGraph(p, width = 300, height =200)) 
+```
 
 ```{r interactive-popup, fig.align='center', echo = FALSE, out.width="45%", out.height="25%", fig.retina = 2, dpi = 300, fig.cap = "Same as Figure 11 except the temperature variation is now shown as a popup in the leaflet map.", fig.show='hold'}
-knitr::include_graphics(here::here("figures/popup-mid.png"))
+knitr::include_graphics(here::here("figures/popup.png"))
 ```

### 2022-06-26 figures/scripts/illu-interactive.R

@@ -1,0 +1,67 @@
+library(cubble)
+library(dplyr)
+library(ggplot2)
+library(tsibble)
+
+set.seed(123)
+dt <- prcp_aus %>% slice_sample(n = 50)
+
+###########################
+# from map to ts
+state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)
+plot_map(state_map) +
+  geom_point(data = dt, aes(x = long, y = lat)) + 
+  geom_point(data = dt %>% filter(id == "ASN00003057"), aes(x = long, y = lat), 
+             color = "red", size =5) 
+
+ggsave(filename = "figures/keynotes-figures/illu-highlighted-map.png")
+
+plot_map(state_map) +
+  geom_point(data = dt, aes(x = long, y = lat)) 
+  
+ggsave(filename = "figures/keynotes-figures/illu-default-map.png")
+
+long <- dt %>% 
+  face_temporal() %>% 
+  mutate(month = lubridate::month(date)) %>% 
+  group_by(month) %>% 
+  summarise(tmax = mean(tmax, na.rm = TRUE)) %>% 
+  mutate(dummy_date = as.Date(glue::glue("2021-{month}-01")))
+
+highlight_line <- long %>% filter(id == "ASN00003057")
+#highlight_point <- highlight_line %>% filter(month == 9)
+long %>% 
+  ggplot(aes(x = dummy_date, y = tmax, group = id)) +
+  #geom_point(size = 0.3) + 
+  geom_line() + 
+  geom_line(data = highlight_line, color = "red", size = 1.5) + 
+  #geom_point(data = highlight_point, color = "red") + 
+  scale_x_date(date_labels = "%b") + 
+  xlab("Month")
+  
+ggsave(filename = "figures/keynotes-figures/illu-highlighted-long.png", width = 5, height = 3)
+
+long %>% 
+  ggplot(aes(x = dummy_date, y = tmax, group = id)) +
+  geom_line() + 
+  scale_x_date(date_labels = "%b") + 
+  xlab("Month")
+
+ggsave(filename = "figures/keynotes-figures/illu-default-long.png", width = 5, height = 3)
+
+###########################
+# from ts to map
+
+highlight_line <- long %>% filter(id == "ASN00003057")
+highlight_point <- highlight_line %>% filter(month == 9)
+long %>% 
+  ggplot(aes(x = dummy_date, y = tmax, group = id)) +
[…]

### 2022-06-26 figures/scripts/illu-matching.R

@@ -1,0 +1,44 @@
+library(tidyverse)
+set.seed(123)
+dt <- tibble(
+  id = factor(c(rep("A", 31), rep("a", 31)), levels = c("A", "a")),
+  date = rep(1:31, 2),
+  val = c(
+    c(rnorm(5), 10, rnorm(7), 5, rnorm(8), 7, rnorm(8)),
+    c(
+      rnorm(6, sd = 0.5), 7, rnorm(7, sd = 0.5), rnorm(6, sd = 0.5),
+      4, rnorm(5, sd = 0.5), 6, rnorm(4, sd = 0.5)
+    )
+  )
+) %>% mutate(val = ifelse(val < 0, 0, val))
+
+circle <- tibble(
+  x = c(6, 14, 23, 7, 21, 27),
+  y = c(10, 5, 7, 7, 4, 6),
+  id = factor(c(rep("A", 3), rep("a", 3)), levels = c("A", "a")),
+  xend = x + 5,
+  match = factor(c("yes", "no", rep("yes", 2), "no", "yes"), levels = c("yes", "no"))
+) 
+
+errorbar <- bind_rows(
+  circle %>% filter(id == "A"),
+  circle %>% filter(id == "A") %>% mutate(id = "a")
+) %>%
+  mutate(id = factor(id, c("A", "a")))
+
+ggplot() +
+  geom_rect(data = errorbar, aes(xmin = x, xmax = xend, ymin = 0, ymax = 10), fill = "grey90") +
+  geom_line(data = dt, aes(x = date, y = val, group = id), color = "black") +
+  geom_vline(data = circle %>% filter(id == "A"), 
+             aes(xintercept = x), linetype = "longdash", color = "grey10", lwd = 0.2) +
+  geom_point(data = circle, aes(x = x, y = y, color = match), size = 3) +
+  facet_wrap(vars(id), nrow = 2) +
+  scale_y_continuous(breaks = seq(0, 10, 2)) +
+  scale_x_continuous(breaks = seq(1, 31, 1)) +
+  scale_color_brewer(palette = "Dark2") + 
+  theme_bw() +
+  theme(panel.grid = element_blank(),
+        legend.position = "bottom") +
+  labs(x = "Time", y = "Value")
+
+ggsave(filename = "figures/illu-matching.png", width = 10, height = 5)

### 2022-06-26 scripts/illu-interactive.R

@@ -1,67 +1,0 @@
-library(cubble)
-library(dplyr)
-library(ggplot2)
-library(tsibble)
-
-set.seed(123)
-dt <- prcp_aus %>% slice_sample(n = 50)
-
-###########################
-# from map to ts
-state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)
-plot_map(state_map) +
-  geom_point(data = dt, aes(x = long, y = lat)) + 
-  geom_point(data = dt %>% filter(id == "ASN00003057"), aes(x = long, y = lat), 
-             color = "red", size =5) 
-
-ggsave(filename = "figures/keynotes-figures/illu-highlighted-map.png")
-
-plot_map(state_map) +
-  geom_point(data = dt, aes(x = long, y = lat)) 
-  
-ggsave(filename = "figures/keynotes-figures/illu-default-map.png")
-
-long <- dt %>% 
-  face_temporal() %>% 
-  mutate(month = lubridate::month(date)) %>% 
-  group_by(month) %>% 
-  summarise(tmax = mean(tmax, na.rm = TRUE)) %>% 
-  mutate(dummy_date = as.Date(glue::glue("2021-{month}-01")))
-
-highlight_line <- long %>% filter(id == "ASN00003057")
-#highlight_point <- highlight_line %>% filter(month == 9)
-long %>% 
-  ggplot(aes(x = dummy_date, y = tmax, group = id)) +
-  #geom_point(size = 0.3) + 
-  geom_line() + 
-  geom_line(data = highlight_line, color = "red", size = 1.5) + 
-  #geom_point(data = highlight_point, color = "red") + 
-  scale_x_date(date_labels = "%b") + 
-  xlab("Month")
-  
-ggsave(filename = "figures/keynotes-figures/illu-highlighted-long.png", width = 5, height = 3)
-
-long %>% 
-  ggplot(aes(x = dummy_date, y = tmax, group = id)) +
-  geom_line() + 
-  scale_x_date(date_labels = "%b") + 
-  xlab("Month")
-
-ggsave(filename = "figures/keynotes-figures/illu-default-long.png", width = 5, height = 3)
-
-###########################
-# from ts to map
-
-highlight_line <- long %>% filter(id == "ASN00003057")
-highlight_point <- highlight_line %>% filter(month == 9)
-long %>% 
-  ggplot(aes(x = dummy_date, y = tmax, group = id)) +
[…]

### 2022-06-26 scripts/illu-matching.R

@@ -1,44 +1,0 @@
-library(tidyverse)
-set.seed(123)
-dt <- tibble(
-  id = factor(c(rep("A", 31), rep("a", 31)), levels = c("A", "a")),
-  date = rep(1:31, 2),
-  val = c(
-    c(rnorm(5), 10, rnorm(7), 5, rnorm(8), 7, rnorm(8)),
-    c(
-      rnorm(6, sd = 0.5), 7, rnorm(7, sd = 0.5), rnorm(6, sd = 0.5),
-      4, rnorm(5, sd = 0.5), 6, rnorm(4, sd = 0.5)
-    )
-  )
-) %>% mutate(val = ifelse(val < 0, 0, val))
-
-circle <- tibble(
-  x = c(6, 14, 23, 7, 21, 27),
-  y = c(10, 5, 7, 7, 4, 6),
-  id = factor(c(rep("A", 3), rep("a", 3)), levels = c("A", "a")),
-  xend = x + 5,
-  match = factor(c("yes", "no", rep("yes", 2), "no", "yes"), levels = c("yes", "no"))
-) 
-
-errorbar <- bind_rows(
-  circle %>% filter(id == "A"),
-  circle %>% filter(id == "A") %>% mutate(id = "a")
-) %>%
-  mutate(id = factor(id, c("A", "a")))
-
-ggplot() +
-  geom_rect(data = errorbar, aes(xmin = x, xmax = xend, ymin = 0, ymax = 10), fill = "grey90") +
-  geom_line(data = dt, aes(x = date, y = val, group = id), color = "black") +
-  geom_vline(data = circle %>% filter(id == "A"), 
-             aes(xintercept = x), linetype = "longdash", color = "grey10", lwd = 0.2) +
-  geom_point(data = circle, aes(x = x, y = y, color = match), size = 3) +
-  facet_wrap(vars(id), nrow = 2) +
-  scale_y_continuous(breaks = seq(0, 10, 2)) +
-  scale_x_continuous(breaks = seq(1, 31, 1)) +
-  scale_color_brewer(palette = "Dark2") + 
-  theme_bw() +
-  theme(panel.grid = element_blank(),
-        legend.position = "bottom") +
-  labs(x = "Time", y = "Value")
-
-ggsave(filename = "figures/illu-matching.png", width = 10, height = 5)

### 2022-06-26 scripts/interactive-popup.R

@@ -1,48 +1,0 @@
-library(cubble)
-library(tidyverse)
-library(leaflet)
-library(leafpop)
-
-clean <-  climate_full %>%
-  filter(!id %in% c("ASN00067033", "ASN00072091", "ASN00059040", 
-                    "ASN00097053", "ASN00040856", "ASN00015548")) %>%
-  face_temporal() %>%
-  mutate(month = lubridate::month(date)) %>% 
-  group_by(month) %>% 
-  summarise(tmax = mean(tmax, na.rm = TRUE),
-            tmin = mean(tmin, na.rm = TRUE),
-            diff = mean(tmax - tmin, na.rm = TRUE)) %>% 
-  ungroup(month) %>% 
-  face_spatial() %>%
-  mutate(temp_diff_var = ts %>% pull(diff) %>% var(na.rm = TRUE)) %>% 
-  face_temporal() %>% 
-  mutate(dummy_date = as.Date(glue::glue("2021-{month}-01"))) %>% 
-  unfold(name)
-  
-df_id <- clean$id %>% unique()
-p <- purrr::map(1:length(df_id), function(i){
-  dt <- clean %>% filter(id == df_id[i])
-  dt %>% 
-    ggplot(aes(x = dummy_date, ymin = tmin, ymax = tmax, group = id)) + 
-    geom_ribbon(color = "grey50", fill = "grey50", alpha = 0.5) + 
-    theme_bw() + 
-    scale_x_date(date_labels = "%b", date_breaks = "1 month") + 
-    labs(x = "Month", y = "Temperature") + 
-    ylim(-10, 50) + 
-    ggtitle(paste0(stringr::str_to_title(unique(dt$name))))
-})
-
-nested <- face_spatial(clean) 
-domain <- nested$temp_diff_var
-pal <- colorNumeric(
-  colorspace::sequential_hcl(
-    "Rocket",  n = 7, cmax = 90, rev = TRUE, 
-    c2 = 40, l2= 85, c1 = 20, l1 = 30),
-  domain = domain)
-
-out <- leaflet(nested) %>% 
-  addTiles() %>% 
-  addCircleMarkers(color = ~pal(domain), radius = 0.1, fillOpacity = 1, opacity = 1) %>% 
-  addPopupGraphs(graph = p, group = "a", width = 300, height =200)
-
-htmlwidgets::saveWidget(out, file = "figures/popup.html")

### 2022-06-27 5-2-aus-hist-max-temp.Rmd
reordering plots

@@
 
-```{r basic-manip, fig.height=6, fig.cap="A glyph map of the monthly maximum average temperature for weather stations in  Victoria and New South Wales for the periods (1971-1975, 2016-2020).
+```{r basic-manip, out.width="100%", fig.height = 5, fig.width = 10, fig.cap="A glyph map of the monthly maximum average temperature for weather stations in  Victoria and New South Wales for the periods (1971-1975, 2016-2020).
 The corresponding average time series for the Cobar station are display on the top left corner.
@@
   ggthemes::theme_map() +
-  theme(legend.position = "bottom", legend.text = element_text(size = 15)) + 
+  theme(legend.position = "bottom", legend.text = element_text(size = 10)) + 
   guides(color = guide_legend(override.aes = list(size=2)))
 
-(cobar / p1)  + 
-  patchwork::plot_layout(heights = c(1,5)) + 
+(p1 | cobar)  + 
+  patchwork::plot_layout(width = c(1,1)) + 
   plot_annotation(tag_levels = 'a')

### 2022-06-27 5-4-era5.Rmd
reordering plots

--- 2022-06-24 4-4-netcdf.Rmd
@@
 
-```{r eval = FALSE, echo = TRUE}
+```{r echo = TRUE}
 raw <- ncdf4::nc_open(here::here("data/era5-pressure.nc"))
-dt <- as_cubble(raw, vars = c("q", "z"))
+dt <- as_cubble(
+  raw, vars = c("q", "z"),
+  long_range = seq(-180, 180, 1), lat_range = seq(-88, 88, 1))
 ```
@@
 
-```{r netcdf, out.width="100%", fig.cap = "A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020) to illustrate the break-up of sourthern polar vortex in late September and early October 2002.
+```{r eval = FALSE}
+# prepare data/era5-pressure.nc
+library(ecmwfr)
+my_request <- list(
+  product_type = "reanalysis",
+  format = "netcdf",
+  pressure_level = "10",
+  year = "2002",
+  month = c("09", "10"),
+  day = c("04", "22", "26", "30"),
+  time = "00:00",
+  area = c(-5, -180, -90, 180),
+  variable = c("specific_humidity", 'geopotential'),
+  dataset_short_name = "reanalysis-era5-pressure-levels",
+  target = "era5-pressure.nc"
+)
+# 
+# # here you need to create an account on Climate Data Store 
+# # at https://cds.climate.copernicus.eu/cdsapp#!/home) to obtain the user ID and API key
+# # then set up the with ecmwfr::wf_set_key(user, key, service = "cds")
+# # the user below is the same `user` in the setup line
+wf_request(
+  user = "113339",
+  request = my_request,
+  transfer = TRUE,
+  path = here::here("data/"))
+```
+
+```{r netcdf, out.width="100%", fig.height = 4, fig.width = 10, fig.cap = "A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020) to illustrate the break-up of sourthern polar vortex in late September and early October 2002.
 The polar vortex, signalled by the high speicfic humidity, splits into two on 2002-09-26 and furthers split into four on 2002-10-04."}
-knitr::include_graphics(here::here("figures/netcdf.png"))
+date <- c("2002-09-22", "2002-09-26", "2002-09-30", "2002-10-04") %>% as.Date()
+res <- dt %>% 
+  face_temporal() %>% 
+  filter(lubridate::date(time) %in% date) %>%
+  unfold(long, lat) %>% 
+  mutate(q =  q* 10^6)
+
+box = c(xmin = -180, ymin = -90, xmax = 180, ymax = -5)
+sf::sf_use_s2(FALSE)
+country <- rnaturalearth::ne_coastline("small", returnclass = "sf") %>% 
+  sf::st_crop(box) %>% 
+  sf::st_cast("MULTILINESTRING")
+
+res %>% 
+  ggplot() +
+  # q for specific humidity
+  geom_point(aes(x = long, y = lat, color = q), size = 2) + 
+  #geom_tile(aes(x = long, y = lat, fill = q), color = "transparent") +
+  # z for geopotential
+  geom_contour(data = res, aes(x = long, y = lat, z = z),
+               color = "grey20", binwidth = 4000, size = 0.5) +
+  geom_sf(data = country , alpha = 0.5, fill = "transparent", color = "lightgreen") +
+  coord_sf(
+    crs = "+proj=stere +lat_0=-90 +lon_0=-180 +k=1 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs ", default_crs = sf::st_crs(4326), label_graticule = "EW") + 
+  theme_bw() + 
+  theme(legend.position = "bottom",
+        axis.ticks = element_blank(),
+        panel.border = element_blank(),
+        axis.text = element_blank(), 
+        panel.grid = element_blank()) + 
+  facet_wrap(vars(as.Date(time)), nrow = 1, shrink = TRUE)+  
+  colorspace::scale_color_continuous_sequential("Purple-Yellow", name = "Specific humidity") + 
+  theme(legend.text = element_text(size = 13),
+        legend.title = element_text(size = 13)) + 
+  labs(x ="", y = "")
 ```

### 2022-06-27 4-other-features.Rmd
misc small fixes

@@
 The map (produced from the nested form) and time series (produced from the long form) are both shared crosstalk objects.
-When a user makes a selection on the map, the site is highlighted (left).
-This activates a row in the nested cubble object, which is then communicated to the long cubble -- all the observations with the same \code{id} (middle) will be selected.
-The long cubble will then highlight the corresponding series in the time series plot (right).
+When a user makes a selection on the map, the site is highlighted (a).
+This activates a row in the nested cubble object, which is then communicated to the long cubble -- all the observations with the same \code{id} (b) will be selected.
+The long cubble will then highlight the corresponding series in the time series plot (c).
 

### 2022-06-27 jss/cubble-jss.tex
misc small fixes

@@
 
+The \texttt{rticles::jss\_article} template used during preparation (\url{https://github.com/rstudio/rticles/tree/c085e125d874c11bb474ec53f4b754e179f8a4e9}) has the behavior that a reference prints out all the author names when it is used for the first time.
+See a standalone example \href{https://github.com/huizezhang-sherry/jss}{here}.
+
+Here a hack is used to overcome this temporarily
+
+Here I reference \citet{ctvspatiotemporal}, \citet{bach_review_2014}, \citet{lu_multidimensional_2018}, \citet{tidyverse}, \citet{tibble}, \citet{sf}, \citet{stars}, \citet{spacetime}, \citet{dplyr}, \citet{ggplot2}, \citet{tsibble}, \citet{hassell_data_2017}, \citet{tidydata}, \citet{leaflet}, \citet{crosstalk}, \citet{cheng2016enabling}, \citet{buja1996interactive}, \citet{xie2014reactive}, \citet{sutherland2000orca}, \citet{buja1988elements}, \citet{Wickham2012-yr}, \citet{hersbach2020era5}, \citet{tidync}, \citet{rnetcdf}, \citet{ncdf4}, \citet{ecwmfr}, \citet{simmons2020global}, \citet{simmons2005ecmwf}, \citet{mcintosh2018using}, \citet{stuart2010matching}, \citet{castanedo2013review}, \citet{cocchi2019data}, \citet{knitr}, \citet{rmarkdown}, \citet{wang2020calendar}, \citet{advr}, \citet{CS07Author}=``Cook, D.
+and Swayne, D.'',, \citet{bom}, \citet{noaa}, \citet{cds}, \citet{wickham_tourr_2011}
+
+\newpage
+
 \hypertarget{introduction}{%
@@
 
-{\centering \includegraphics{figures/basic-manip-1}
+{\centering \includegraphics[width=1\linewidth]{figures/basic-manip-1}
 
@@
 +   raw, vars = c("q", "z"),
-+   long_range = seq(-180, 180, 1), lat_range = seq(-90, 90, 1))
++   long_range = seq(-180, 180, 1), lat_range = seq(-88, 88, 1))
 \end{CodeInput}
@@
 
-{\centering \includegraphics{figures/netcdf-1}
+{\centering \includegraphics[width=1\linewidth]{figures/netcdf-1}
 
@@
 This work is funded by a Commonwealth Scientific and Industrial Research Organisation (CSIRO) Data61 Scholarship and started while Nicolas Langrené was affiliated with CSIRO's Data61.
-The article is created using \pkg{knitr} \citep{knitr} and \pkg{rmarkdown} \citep{rmarkdown} in R.
+The article is created using \pkg{knitr} \citep{knitr} and \pkg{rmarkdown} \citep{rmarkdown} in R using the \texttt{rticles::jss\_article} template.
 The source code for reproducing this paper can be found at: \url{https://github.com/huizezhang-sherry/paper-cubble}.

### 2022-06-27 cubble-jss.R

@@ -18,6 +18,19 @@ library(tidyverse)
 library(patchwork)
 
 
+## ----child = here::here("reference-hack.Rmd")---------------------------------
+
+## -----------------------------------------------------------------------------
+
+tex <- readLines(here::here("references.bib"))
+b <- tex %>% stringr::str_extract("\\@(article|techreport|misc|Manual|Book)\\{.+,")
+b <- b[!is.na(b)]
+ref_head <- b %>% stringr::str_replace("\\@(article|techreport|misc|Manual|Book)\\{", "") %>% 
+  stringr::str_replace(",", "") 
+res <- paste0("@", ref_head, collapse = ", ")
+
+
+
 ## ----child = here::here("1-intro.Rmd")----------------------------------------
 
 
@@ -269,7 +282,7 @@ cobar <- tmax %>% filter(id == "ASN00048027") %>%
 
 
 
-## ----basic-manip, fig.height=6, fig.cap="A glyph map of the monthly maximum average temperature for weather stations in  Victoria and New South Wales for the periods (1971-1975, 2016-2020). The corresponding average time series for the Cobar station are display on the top left corner. From the glyph map we can observe that the monthly trend is similar for all locations (low in the winter, high in the summer), and small increased temperatures, particularly in late summer can be seen at most stations in New South Wales."----
+## ----basic-manip, out.width="100%", fig.height = 5, fig.width = 10, fig.cap="A glyph map of the monthly maximum average temperature for weather stations in  Victoria and New South Wales for the periods (1971-1975, 2016-2020). The corresponding average time series for the Cobar station are display on the top left corner. From the glyph map we can observe that the monthly trend is similar for all locations (low in the winter, high in the summer), and small increased temperatures, particularly in late summer can be seen at most stations in New South Wales."----
 tmax <- tmax %>% unfold(long, lat)
 box_df <- tmax %>% face_spatial() %>% filter(id == "ASN00048027")
 nsw_vic <- ozmaps::abs_ste %>% filter(NAME %in% c("Victoria","New South Wales"))
@@ -291,11 +304,11 @@ p1 <- ggplot() +
   scale_color_brewer("", palette = "Dark2") + 
   coord_sf(xlim = c(141, 154), ylim = c(-39, -28.5)) + 
[…]

### 2022-06-27 code/appendix.Rmd
clean up repo

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(cubble)
library(patchwork)
library(leaflet)
library(leafpop)

```
```{r echo = FALSE}
load(here::here("data/station_nested.rda"))
load(here::here("data/climate_full.rda"))

```
```{r eval = FALSE}
## # The kmeans clustering in station_nested data can takes some time to run,
## # This chunk runs the algorithm and save the result as staiton_nested for further use
## coords <- cbind(prcp_aus$long, prcp_aus$lat)
## dist_raw <- geosphere::distm(coords, coords)
## 
## station_long <- cubble::prcp_aus %>%
##   face_temporal(ts) %>%
##   mutate(wk = lubridate::week(date)) %>%
##   group_by(wk) %>%
##   summarise(prcp = sum(prcp, na.rm = TRUE)) %>%
##   unfold(cluster)
## 
## set.seed(123)
## station_nested <- station_long %>%
##   face_spatial() %>%
##   strip_rowwise() %>%
##   mutate(cluster = kmeans(dist_raw,centers = 20, nstart = 500)$cluster)
## save(station_nested, file = here::here("data/station_nested.rda"))

```
```{r basic-agg, fig.height = 7, fig.width = 15, fig.cap="Profile of aggregated precipitation at 639 weather stations in Australia. Subplot A shows the glyph map of the weekly averaged precipitation of each cluster. The group number is printed in the middle of the y minor axis and can be used as a reference line to read the magnitude. Subplot B shows the station membership of each cluster."}
load(here::here("data/station_nested.rda"))
cluster_nested <- station_nested %>%
  switch_key(cluster) %>% 
  get_centroid()

cluster_long <- cluster_nested %>% 
  face_temporal() %>% 
  group_by(wk) %>%
  summarise(prcp = mean(prcp, na.rm = TRUE)) %>% 
  unfold(cent_long, cent_lat)

state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)

ggplot_smooth <- cluster_long %>% 
  ggplot() +
  geom_smooth(aes(x = wk, y = prcp, group = cluster), span = 0.4)

smoother <- layer_data(ggplot_smooth) %>% 
  left_join(cluster_long %>% select(cluster, cent_long, cent_lat), by = c("group" = "cluster"))

p1 <- ggplot(data = smoother, 
             aes(x_minor = x, y_minor = y, x_major = cent_long, y_major = cent_lat)) + 
  geom_sf(data = state_map, inherit.aes = FALSE,
          color = "grey80", alpha = 0.4, linetype = 3) + 
  geom_text(data = cluster_nested, 
            aes(x = cent_long, y = cent_lat, label = cluster), inherit.aes = FALSE) +
  geom_glyph(height = 2, width = 4) + 
  theme_void()

p2 <- ggplot() + 
  geom_sf(data = state_map, inherit.aes = FALSE,
          color = "grey80", alpha = 0.4, linetype = 3) + 
  geom_point(data = station_nested, aes(x = long, y = lat), size = 0.5) +
  ggforce::geom_mark_hull(data = cluster_nested %>% tidyr::unnest(hull),
                          expand = 0, radius = 0,
                          aes(x = long, y = lat, group = cluster)) + 
  theme_void()

(p1 | p2) + plot_annotation(tag_levels = 'A')

```
```{r eval = FALSE}
## clean <-  climate_full %>%
##   filter(!id %in% c("ASN00067033", "ASN00072091", "ASN00059040",
##                     "ASN00097053", "ASN00040856", "ASN00015548")) %>%
##   face_temporal() %>%
##   mutate(month = lubridate::month(date)) %>%
##   group_by(month) %>%
##   summarise(tmax = mean(tmax, na.rm = TRUE),
##             tmin = mean(tmin, na.rm = TRUE),
##             diff = mean(tmax - tmin, na.rm = TRUE)) %>%
##   ungroup(month) %>%
##   face_spatial() %>%
##   mutate(temp_diff_var = ts %>% pull(diff) %>% var(na.rm = TRUE)) %>%
##   face_temporal() %>%
##   mutate(dummy_date = as.Date(glue::glue("2021-{month}-01"))) %>%
##   unfold(name)
## 
## df_id <- clean$id %>% unique()
## p <- purrr::map(1:length(df_id), function(i){
##   dt <- clean %>% filter(id == df_id[i])
##   dt %>%
##     ggplot(aes(x = dummy_date, ymin = tmin, ymax = tmax, group = id)) +
##     geom_ribbon(color = "grey50", fill = "grey50", alpha = 0.5) +
##     theme_bw() +
##     scale_x_date(date_labels = "%b", date_breaks = "1 month") +
##     labs(x = "Month", y = "Temperature") +
##     ylim(-10, 50) +
##     ggtitle(paste0(stringr::str_to_title(unique(dt$name))))
## })
## 
## nested <- face_spatial(clean)
## domain <- nested$temp_diff_var
## pal <- colorNumeric(
##   colorspace::sequential_hcl(
##     "Rocket",  n = 7, cmax = 90, rev = TRUE,
##     c2 = 40, l2= 85, c1 = 20, l1 = 30),
##   domain = domain)
## 
## leaflet(nested) %>%
##   addTiles() %>%
##   addCircleMarkers(color = ~pal(domain), radius = 0.1,
##                    fillOpacity = 1, opacity = 1,
##                    popup = popupGraph(p, width = 300, height =200))

```
```{r interactive-popup, fig.align='center', echo = FALSE, out.width="45%", out.height="25%", fig.retina = 2, dpi = 300, fig.cap = "Same as Figure 11 except the temperature variation is now shown as a popup in the leaflet map.", fig.show='hold'}
knitr::include_graphics(here::here("figures/popup.png"))

```
```{r eval = FALSE}
## library(tidyverse)
## library(cubble)
## all_stations <- rnoaa::ghcnd_stations() %>%
##   filter(str_starts(id, "ASN")) %>% # Australian stations start wiht "ASN"
##   filter(last_year >= 2020) %>%
##   mutate(wmo_id = as.numeric(wmo_id), name = str_to_lower(name)) %>%
##   select(-state, -gsn_flag) %>%
##   select(id, longitude, latitude, elevation, name,
##          wmo_id, element, first_year, last_year) %>%
##   rename(long = longitude, lat = latitude, elev = elevation)
## 
## tmax_stations <- all_stations %>%
##   filter(element == "TMAX", first_year < 1970, !is.na(wmo_id))
## 
## raw_tmax <- all_stations %>%
##   rowwise() %>%
##   mutate(ts = list(rnoaa::meteo_pull_monitors(
##     monitors = id, var = "TMAX",
##     date_min = glue::glue("{first_year}-01-01"),
##     date_max = glue::glue("{last_year}-12-31")
##     ) %>%
##       select(-id)
##     )
##   )
## 
## historical_tmax <- raw_tmax %>%
##   select(-element) %>%
##   unnest(ts) %>%
##   mutate(tmax = tmax/10) %>%
##   filter(lubridate::year(date) %in% c(1971: 1975, 2016:2020)) %>%
##   as_cubble(index = date, key = id, coords = c(long, lat))
## 
## save(historical_tmax, file = here::here("data/historical_tmax.rda"))

```
```{r eval = FALSE}
## # all the Australian stations have all of the three PRCP, TMAX, and TMIN recorded
## aus_stations <- all_stations %>%
##   filter(element %in% c("PRCP", "TMAX", "TMIN")) %>%
##   nest(element: last_year) %>%
##   rowwise() %>%
##   filter(nrow(data) == 3) %>%
##   select(-data)
## 
## aus_climate_raw <- aus_stations %>%
##   rowwise() %>%
##   mutate(ts = list(
##     rnoaa::meteo_pull_monitors(
##       monitors = id, var = c("PRCP", "TMAX", "TMIN"),
##       date_min = "2016-01-01", date_max = "2020-12-31"
##       ) %>%
##       select(-id)
##     )
##   )
## 
## climate_full <- aus_climate_raw %>%
##   unnest(ts) %>%
##   mutate(tmax = tmax/10, tmin = tmin/10) %>%
##   cubble::as_cubble(key = id, index = date, coords = c(long, lat))
## 
## save(climate_full, file = here::here("data/climate_full.rda"))
```

### 2022-06-27 code/main.Rmd
clean up repo

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  cache = TRUE,
  fig.path = here::here("figures/")
)

options(prompt = "R> ", continue = "+ ",
        tibble.print_max = 5, tibble.print_min = 5,
        width = 70)

```
```{r echo = FALSE}
library(cubble)
library(tidyverse)
library(patchwork)

```
```{r }
cubble::climate_flat

```
```{r echo = TRUE}
cubble_nested <- climate_flat |>
  as_cubble(key = id, index = date, coords = c(long, lat))
cubble_nested

```
```{r echo = TRUE}
cubble_nested |>
  mutate(rain_day = sum(ts$prcp != 0))

```
```{r echo = TRUE}
cubble_long <- cubble_nested |> face_temporal()
cubble_long

```
```{r echo = TRUE}
cubble_back <- cubble_long |> face_spatial()
cubble_back
identical(cubble_nested, cubble_back)

```
```{r echo = TRUE}
cubble_unfold <- cubble_long |> unfold(long, lat)
cubble_unfold

```
```{r echo = TRUE}
climate_flat_ts <- climate_flat |> 
  tsibble::as_tsibble(key = id, index = date)
climate_flat_cb <-  climate_flat_ts |> 
  cubble::as_cubble(coords = c(long, lat))
climate_flat_cb

```
```{r echo = TRUE}
# add station-based features in the nested form.
climate_flat_cb |> 
  mutate(
    fabletools::features(
      ts, tmax, list(tmax_mean = mean, tmax_var = var)
      )
  )

```
```{r eval = FALSE}
## library(tidyverse)
## #https://www.coronavirus.vic.gov.au/victorian-coronavirus-covid-19-data
## raw <- read_csv(here::here("data/NCOV_COVID_Cases_by_LGA_Source_20220324.csv"),
##                   col_names = c("date", "postcode", "source", "lga"), skip = 1)
## 
## covid <- raw %>%
##   count(date, lga, source, postcode) %>%
##   filter(date >= as.Date("2022-01-01")) %>%
##   filter(source == "Contact with a confirmed case") %>%
##   group_by(date, lga, source) %>%
##   count() %>%
##   ungroup(date) %>%
##   mutate(roll_mean = zoo::rollmean(n, k = 14, fill = NA)) %>%
##   tsibble::as_tsibble(key = lga, index = date) %>%
##   tsibble::fill_gaps(.full = TRUE)
## save(covid, file = here::here("data/covid.rda"))

```
```{r }
load(here::here("data/covid.rda"))

lga <- strayr::read_absmap("lga2018") %>%
  rename(lga = lga_name_2018) %>%
  filter(state_name_2016 == "Victoria")

```
```{r echo = TRUE}
covid %>% head(5)

```
```{r echo = TRUE}
lga %>% head(5)

```
```{r echo = TRUE, warning = TRUE, message= TRUE}
cb <- as_cubble(
  list(spatial = lga, temporal = covid),
  key = lga, index = date, coords = c(cent_long, cent_lat)
  )

```
```{r echo = TRUE}
pair <- as_cubble(
  list(spatial = lga, temporal = covid),
  key = lga, index = date, coords = c(cent_long, cent_lat),
  output = "unmatch"
  )

pair

```
```{r echo = TRUE}
lga <- lga %>%
  mutate(lga = ifelse(lga == "Kingston (C) (Vic.)", "Kingston (C)", lga),
         lga = ifelse(lga == "Latrobe (C) (Vic.)", "Latrobe (C)", lga)) %>%
  filter(!lga %in% pair$others$spatial)

covid <- covid %>% filter(!lga %in% pair$others$temporal)

cb <- as_cubble(data = list(spatial = lga, temporal = covid),
                key = lga, index = date, coords = c(cent_long, cent_lat))

```
```{r }
load(here::here("data/historical_tmax.rda"))

```
```{r echo = TRUE}
tmax <- historical_tmax %>%
  filter(between(stringr::str_sub(id, 7, 8), 46, 90))

```
```{r echo = TRUE}
tmax <- tmax %>%
  face_temporal() %>% 
  group_by(month = lubridate::month(date),
         group = as.factor(
           ifelse(lubridate::year(date) > 2015,
           "2016 ~ 2020", "1971 ~ 1975"))) %>%
  summarise(tmax = mean(tmax, na.rm = TRUE))

```
```{r echo = TRUE}
tmax %>% 
  face_spatial() %>% 
  mutate(n = nrow(ts)) %>%
  arrange(n) %>%
  pull(n) %>% 
  head(10)

tmax <- tmax %>% 
  face_spatial() %>% 
  filter(nrow(ts) == 24) %>%
  face_temporal()

```
```{r }
cobar <- tmax %>% filter(id == "ASN00048027") %>%
  ggplot(aes(x = month, y = tmax, color = group)) +
  geom_line() +
  scale_color_brewer(palette = "Dark2", guide = "none") +
  scale_x_continuous(
    breaks = seq(1, 12, 1), 
    labels = c("J", "F", "M", "A",
               "M", "J", "J", "A",
               "S", "O", "N", "D")
    ) +
  labs(x = "", y  = "Temp (C)") +
  theme_bw() + 
  theme(aspect.ratio = 0.3)

```
```{r basic-manip, out.width="100%", fig.height = 5, fig.width = 10, fig.cap="A glyph map of the monthly maximum average temperature for weather stations in  Victoria and New South Wales for the periods (1971-1975, 2016-2020). The corresponding average time series for the Cobar station are display on the top left corner. From the glyph map we can observe that the monthly trend is similar for all locations (low in the winter, high in the summer), and small increased temperatures, particularly in late summer can be seen at most stations in New South Wales."}
tmax <- tmax %>% unfold(long, lat)
box_df <- tmax %>% face_spatial() %>% filter(id == "ASN00048027")
nsw_vic <- ozmaps::abs_ste %>% filter(NAME %in% c("Victoria","New South Wales"))

p1 <- ggplot() + 
  geom_sf(data = nsw_vic, fill = "grey95", 
          color = "white", size=2) + 
  geom_glyph(
    data = tmax, 
    aes(x_major = long, x_minor = month, 
        y_major = lat, y_minor = tmax,
        group = interaction(id, group), color = group),
    width = 1.2, height = 0.8) +
    geom_rect(
      data = box_df,
      aes(xmin = long - 0.6, xmax = long + 0.6,
          ymin = lat - 0.12, ymax = lat + 0.35),
      fill = "transparent", color = "black") +
  scale_color_brewer("", palette = "Dark2") + 
  coord_sf(xlim = c(141, 154), ylim = c(-39, -28.5)) + 
  ggthemes::theme_map() +
  theme(legend.position = "bottom", legend.text = element_text(size = 10)) + 
  guides(color = guide_legend(override.aes = list(size=2)))

(p1 | cobar)  + 
  patchwork::plot_layout(width = c(1,1)) + 
  plot_annotation(tag_levels = 'a')

```
```{r }
load(here::here("data/climate_full.rda"))

climate <- climate_full %>%
  filter(between(stringr::str_sub(id, 7, 8), 76, 90)) %>%
  face_temporal() %>%
  filter(lubridate::year(date) == 2020) %>%
  face_spatial() %>%
  mutate(type = "climate")

river <- river %>% mutate(type = "river")

```
```{r echo = TRUE}
res <- match_sites(
  river, climate,
  temporal_by = c("Water_course_level" = "prcp"),
  temporal_independent = "prcp",  
  temporal_n_highest = 30,
  temporal_min_match = 15, 
)

```
```{r }
res %>% print()

```
```{r matching, out.width="100%", fig.height = 5, fig.width = 10, fig.cap="Weather stations and river gauges with matched pairs labelled on the map (a) and plotted across time (b). Precipitation and water level have been standardised between 0 and 1 to be displayed on the same scale. The water level reflects the increase in precipitation. The numbers (1, 5, 6, 10) indicate the group index derived from spatial matching, only those that were selectd by temporal matching are shown here."}
vic_map <- ozmaps::abs_ste %>% 
  filter(NAME == "Victoria") %>% 
  rmapshaper::ms_simplify()

p1 <-ggplot() + 
  geom_sf(data = vic_map, fill = "grey95", color = "white") + 
  geom_point(data = dplyr::bind_rows(river, climate), 
             aes(x = long, y = lat, color = type), alpha = 0.2, fill = 0.2) +
  geom_point(data = res, aes(x = long, y = lat, color = type)) +
  ggrepel::geom_label_repel(
    data = res %>% filter(type == "river"), aes(x = long, y = lat, label = group)) +
  scale_color_brewer(palette = "Dark2")  + 
  ggthemes::theme_map() +
  ggplot2::theme(legend.position = "bottom",
                 legend.text = element_text(size = 15),
                 legend.title = element_text(size = 15)) +
  ggplot2::labs(x = "Longitude", y = "Latitude") + 
  guides(color = guide_legend(override.aes = list(size=5)))

res_long <- res %>%
  face_temporal(ts) %>%
  unfold(group, type) %>%
  rename(prcp = matched_var) %>% 
  mutate(prcp = (prcp - min(prcp, na.rm = TRUE))/ (max(prcp, na.rm = TRUE) - min(prcp, na.rm = TRUE)))

p2 <- res_long %>% 
  ggplot(aes(x = date, y = prcp, color = type, group = id)) +
  geom_line(alpha = 0.8) +
  facet_wrap(vars(group)) +
  scale_color_brewer(palette = "Dark2", guide = "none") +
  theme_bw() +
  scale_x_date(date_labels = "%b") + 
  labs(x = "Date", y = "Precipitation/ water level")

(p1 | p2) + 
  patchwork::plot_layout(guides = "collect") + 
  plot_annotation(tag_levels = 'a')&
  theme(legend.position = "bottom")

```
```{r echo = TRUE}
raw <- ncdf4::nc_open(here::here("data/era5-pressure.nc"))
dt <- as_cubble(
  raw, vars = c("q", "z"),
  long_range = seq(-180, 180, 1), lat_range = seq(-88, 88, 1))

```
```{r eval = FALSE}
## # prepare data/era5-pressure.nc
## library(ecmwfr)
## my_request <- list(
##   product_type = "reanalysis",
##   format = "netcdf",
##   pressure_level = "10",
##   year = "2002",
##   month = c("09", "10"),
##   day = c("04", "22", "26", "30"),
##   time = "00:00",
##   area = c(-5, -180, -90, 180),
##   variable = c("specific_humidity", 'geopotential'),
##   dataset_short_name = "reanalysis-era5-pressure-levels",
##   target = "era5-pressure.nc"
## )
## #
## # # here you need to create an account on Climate Data Store
## # # at https://cds.climate.copernicus.eu/cdsapp#!/home) to obtain the user ID and API key
## # # then set up the with ecmwfr::wf_set_key(user, key, service = "cds")
## # # the user below is the same `user` in the setup line
## wf_request(
##   user = "113339",
##   request = my_request,
##   transfer = TRUE,
##   path = here::here("data/"))

```
```{r netcdf, out.width="100%", fig.height = 4, fig.width = 10, fig.cap = "A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020) to illustrate the break-up of sourthern polar vortex in late September and early October 2002. The polar vortex, signalled by the high speicfic humidity, splits into two on 2002-09-26 and furthers split into four on 2002-10-04."}
date <- c("2002-09-22", "2002-09-26", "2002-09-30", "2002-10-04") %>% as.Date()
res <- dt %>% 
  face_temporal() %>% 
  filter(lubridate::date(time) %in% date) %>%
  unfold(long, lat) %>% 
  mutate(q =  q* 10^6)

box = c(xmin = -180, ymin = -90, xmax = 180, ymax = -5)
sf::sf_use_s2(FALSE)
country <- rnaturalearth::ne_coastline("small", returnclass = "sf") %>% 
  sf::st_crop(box) %>% 
  sf::st_cast("MULTILINESTRING")

res %>% 
  ggplot() +
  # q for specific humidity
  geom_point(aes(x = long, y = lat, color = q), size = 2) + 
  #geom_tile(aes(x = long, y = lat, fill = q), color = "transparent") +
  # z for geopotential
  geom_contour(data = res, aes(x = long, y = lat, z = z),
               color = "grey20", binwidth = 4000, size = 0.5) +
  geom_sf(data = country , alpha = 0.5, fill = "transparent", color = "lightgreen") +
  coord_sf(
    crs = "+proj=stere +lat_0=-90 +lon_0=-180 +k=1 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs ", default_crs = sf::st_crs(4326), label_graticule = "EW") + 
  theme_bw() + 
  theme(legend.position = "bottom",
        axis.ticks = element_blank(),
        panel.border = element_blank(),
        axis.text = element_blank(), 
        panel.grid = element_blank()) + 
  facet_wrap(vars(as.Date(time)), nrow = 1, shrink = TRUE)+  
  colorspace::scale_color_continuous_sequential("Purple-Yellow", name = "Specific humidity") + 
  theme(legend.text = element_text(size = 13),
        legend.title = element_text(size = 13)) + 
  labs(x ="", y = "")

```
```{r eval = FALSE}
## # This chunk generates the interactive graphic and save it as figures/linking.html, where the three screencasts in the figure comes from.
## library(crosstalk)
## library(plotly)
## 
## clean <- climate_full %>%
##   filter(!id %in% c("ASN00067033", "ASN00072091", "ASN00059040",
##                     "ASN00097053", "ASN00040856", "ASN00015548")) %>%
##   face_temporal() %>%
##   mutate(month = lubridate::month(date)) %>%
##   group_by(month) %>%
##   summarise(
##     tmax = mean(tmax, na.rm = TRUE),
##     tmin = mean(tmin, na.rm = TRUE),
##     diff = mean(tmax - tmin, na.rm = TRUE)
##     ) %>%
##   ungroup(month) %>%
##   mutate(dummy_date = as.Date(glue::glue("2021-{month}-01"))) %>%
##   face_spatial() %>%
##   mutate(temp_diff_var = ts %>% pull(diff) %>% var(na.rm = TRUE))
## 
## nested <- clean %>%
##   SharedData$new(~id, group = "cubble")
## 
## long <- clean %>%
##   face_temporal() %>%
##   unfold(temp_diff_var) %>%
##   arrange(temp_diff_var) %>%
##   SharedData$new(~id, group = "cubble")
## 
## state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)
## 
## p1 <- nested %>%
##   ggplot() +
##   geom_sf(data = state_map, fill = "grey95", color = "white") +
##   geom_point(aes(x = long, y = lat,color = temp_diff_var, label = name)) +
##   colorspace::scale_color_continuous_sequential(
##     "Rocket",  n_interp = 7, cmax = 90, rev = TRUE,
##     c2 = 40, l2= 85, c1 = 20, l1 = 30, name = "Var. temp. diff.") +
##   labs(x = "Longitude", y = "Latitude") +
##   coord_sf() +
##   theme_bw() +
##   theme(
##     panel.grid.major = element_blank(),
##     legend.position = "none"
##     )
## 
## p2 <- long %>%
##   ggplot(aes(x = dummy_date, group = id,
##              fill = temp_diff_var, color = temp_diff_var)) +
##   geom_ribbon(aes(ymin = tmin, ymax = tmax), size = 0.1, alpha = 0.3) +
##   geom_line(aes(y = tmin), alpha = 0.1) +
##   geom_line(aes(y = tmax), alpha = 0.1) +
##   geom_point(aes(y = tmax), size = 0.1) +
##   geom_point(aes(y = tmin), size = 0.1) +
##   colorspace::scale_fill_continuous_sequential(
##     "Rocket",  n_interp = 7, cmax = 90, rev = TRUE,
##     c2 = 40, l2= 85, c1 = 20, l1 = 30, name = "Var. temp. diff.") +
##   colorspace::scale_colour_continuous_sequential(
##     "Rocket",  n_interp = 7, cmax = 90, rev = TRUE,
##     c2 = 40, l2= 85, c1 = 20, l1 = 30, name = "Var. temp. diff.") +
##   scale_x_date(date_labels = "%b", date_breaks = "1 month") +
##   labs(x = "Month", y = "Temperature") +
##   theme_bw() +
##   theme(
##     panel.grid.major = element_blank(),
##     legend.position = "bottom"
##     )
## 
## out <- bscols(
##   ggplotly(p1, width = 900, height = 700, tooltip = "label") %>%
##     highlight(
##       on = "plotly_selected",
##       off = "plotly_deselect",
##       opacityDim = 0.1),
##   ggplotly(p2, width = 1200, height = 700, tooltip = "label") %>%
##     highlight(
##       on = "plotly_selected",
##       off = "plotly_deselect",
##       opacityDim = 0.012),
##   widths = c(5, 5)
## )
## 
## htmltools::save_html(out, file = "figures/linking.html")
##

```
```{r interactive-linking, echo = FALSE, out.width="100%", out.height="23%", fig.retina = 2, dpi = 300, fig.cap = "Exploring temperature variation using linking of a map and seasonal display. Each row is a screen dump of the process. The top row shows all locations and all temperature profiles. Selecting a particular location on the map (here Mount Elizabeth) produces the plot in the second row. The maximum and minimum temperatures are shown using a ribbon. The bottom row first selects the lowest temperature in August in the seasonal display, which highlights the corresponding station on the map (Thredbo AWS). Another  station, located in the Tasmania Island, is then selected to compare its temperature variation with Thredbo AWS.", fig.show='hold'}
knitr::include_graphics(here::here("figures/linking.png"))
knitr::include_graphics(here::here("figures/linking-north.png"))
knitr::include_graphics(here::here("figures/linking-lower.png"))

```
```{r}
sessioninfo::session_info()
```

### 2022-06-27 jss/appendix.Rmd
knit
clean up repo

@@
 
-Figure \ref{fig:interactive-popup} shows the same information as Figure 9 but using leaflet and popups [@leafpop].
+Figure \ref{fig:interactive-popup} shows the same information as Figure 9 but using leaflet and popups.
 
@@
 
-```{r illu-interactive-2, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "Linking between multiple plots.
+```{r illu-interactive-2, echo = FALSE, fig.align="center", out.height = "35%", out.width = "100%", fig.cap = "Linking between multiple plots.
 The line plots and the map are constructed from shared crosstalk objects (long and nested cubbles).

### 2022-06-27 jss/main.Rmd
clean up repo

--- 2022-06-24 jss/cubble-jss.Rmd
@@
   cache = TRUE,
-  fig.path = "figures/"
+  fig.path = here::here("figures/")
 )
@@
 library(patchwork)
+```
+
+\newpage
+
+```{r child = here::here("jss/reference-hack.Rmd")}
+
 ```

### 2022-06-27 jss/main.tex
clean up repo

--- 2022-06-27 jss/cubble-jss.tex
@@
 
-{\centering \includegraphics[width=1\linewidth]{figures/basic-manip-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/research/paper-cubble/figures/basic-manip-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{figures/matching-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/research/paper-cubble/figures/matching-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{figures/netcdf-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/research/paper-cubble/figures/netcdf-1}
 

### 2022-06-27 jss/reference-hack.Rmd
clean up repo

--- 2022-06-24 reference-hack.Rmd
@@
 ```{r}
+
 tex <- readLines(here::here("references.bib"))
-
 b <- tex %>% stringr::str_extract("\\@(article|techreport|misc|Manual|Book)\\{.+,")
@@
   stringr::str_replace(",", "") 
-
 res <- paste0("@", ref_head, collapse = ", ")

### 2022-06-27 appendix.R

@@ -1,205 +1,0 @@
-## ----setup, include=FALSE-----------------------------------------------------
-knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
-library(tidyverse)
-library(cubble)
-library(patchwork)
-library(leaflet)
-library(leafpop)
-
-
-## ---- echo = FALSE------------------------------------------------------------
-load(here::here("data/station_nested.rda"))
-load(here::here("data/climate_full.rda"))
-
-
-## ----eval = FALSE, echo = TRUE------------------------------------------------
-## station_nested <- climate_full %>% mutate(cluster = ...)
-
-
-## ----eval = FALSE, echo = TRUE------------------------------------------------
-## cluster_nested <- station_nested %>% switch_key(cluster)
-
-
-## ----eval = FALSE, echo = TRUE------------------------------------------------
-## cluster_nested <- cluster_nested %>% get_centroid()
-
-
-## ----eval = FALSE, echo = FALSE-----------------------------------------------
-## # The kmeans clustering in station_nested data can takes some time to run,
-## # This chunk runs the algorithm and save the result as staiton_nested for further use
-## coords <- cbind(prcp_aus$long, prcp_aus$lat)
-## dist_raw <- geosphere::distm(coords, coords)
-## 
-## station_long <- cubble::prcp_aus %>%
-##   face_temporal(ts) %>%
-##   mutate(wk = lubridate::week(date)) %>%
-##   group_by(wk) %>%
-##   summarise(prcp = sum(prcp, na.rm = TRUE)) %>%
-##   unfold(cluster)
-## 
-## set.seed(123)
-## station_nested <- station_long %>%
-##   face_spatial() %>%
-##   strip_rowwise() %>%
-##   mutate(cluster = kmeans(dist_raw,centers = 20, nstart = 500)$cluster)
-## save(station_nested, file = here::here("data/station_nested.rda"))
-
-
[…]

### 2022-06-27 code/appendix.R

@@ -1,0 +1,187 @@
+## ----setup, include=FALSE-------------------------------------------
+knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
+library(tidyverse)
+library(cubble)
+library(patchwork)
+library(leaflet)
+library(leafpop)
+
+
+## ---- echo = FALSE--------------------------------------------------
+load(here::here("data/station_nested.rda"))
+load(here::here("data/climate_full.rda"))
+
+## ----eval = FALSE-------------------------------------
+## # The kmeans clustering in station_nested data can takes some time to run,
+## # This chunk runs the algorithm and save the result as staiton_nested for further use
+## coords <- cbind(prcp_aus$long, prcp_aus$lat)
+## dist_raw <- geosphere::distm(coords, coords)
+## 
+## station_long <- cubble::prcp_aus %>%
+##   face_temporal(ts) %>%
+##   mutate(wk = lubridate::week(date)) %>%
+##   group_by(wk) %>%
+##   summarise(prcp = sum(prcp, na.rm = TRUE)) %>%
+##   unfold(cluster)
+## 
+## set.seed(123)
+## station_nested <- station_long %>%
+##   face_spatial() %>%
+##   strip_rowwise() %>%
+##   mutate(cluster = kmeans(dist_raw,centers = 20, nstart = 500)$cluster)
+## save(station_nested, file = here::here("data/station_nested.rda"))
+
+
+## ----basic-agg, fig.height = 7, fig.width = 15, fig.cap="Profile of aggregated precipitation at 639 weather stations in Australia. Subplot A shows the glyph map of the weekly averaged precipitation of each cluster. The group number is printed in the middle of the y minor axis and can be used as a reference line to read the magnitude. Subplot B shows the station membership of each cluster."----
+load(here::here("data/station_nested.rda"))
+cluster_nested <- station_nested %>%
+  switch_key(cluster) %>% 
+  get_centroid()
+
+cluster_long <- cluster_nested %>% 
+  face_temporal() %>% 
+  group_by(wk) %>%
+  summarise(prcp = mean(prcp, na.rm = TRUE)) %>% 
+  unfold(cent_long, cent_lat)
+
[…]

### 2022-06-27 code/main.R

@@ -1,0 +1,447 @@
+## ----setup, echo = FALSE--------------------------------------------
+knitr::opts_chunk$set(
+  echo = TRUE,
+  warning = FALSE,
+  message = FALSE,
+  cache = TRUE,
+  fig.path = here::here("figures/")
+)
+
+options(prompt = "R> ", continue = "+ ",
+        tibble.print_max = 5, tibble.print_min = 5,
+        width = 70)
+
+
+## ----echo = FALSE---------------------------------------------------
+library(cubble)
+library(tidyverse)
+library(patchwork)
+
+## -------------------------------------------------------------------
+cubble::climate_flat
+
+
+## ----echo = TRUE----------------------------------------------------
+cubble_nested <- climate_flat |>
+  as_cubble(key = id, index = date, coords = c(long, lat))
+cubble_nested
+
+
+## ----echo = TRUE----------------------------------------------------
+cubble_nested |>
+  mutate(rain_day = sum(ts$prcp != 0))
+
+
+## ----echo = TRUE----------------------------------------------------
+cubble_long <- cubble_nested |> face_temporal()
+cubble_long
+
+
+## ----echo = TRUE----------------------------------------------------
+cubble_back <- cubble_long |> face_spatial()
+cubble_back
+identical(cubble_nested, cubble_back)
+
+
+## ----echo = TRUE----------------------------------------------------
+cubble_unfold <- cubble_long |> unfold(long, lat)
+cubble_unfold
+
+
+## ----echo = TRUE----------------------------------------------------
+climate_flat_ts <- climate_flat |> 
+  tsibble::as_tsibble(key = id, index = date)
+climate_flat_cb <-  climate_flat_ts |> 
+  cubble::as_cubble(coords = c(long, lat))
+climate_flat_cb
+
+
+## ----echo = TRUE----------------------------------------------------
+# add station-based features in the nested form.
+climate_flat_cb |> 
+  mutate(
+    fabletools::features(
+      ts, tmax, list(tmax_mean = mean, tmax_var = var)
+      )
+  )
+
+
+## ----eval = FALSE---------------------------------------------------
+## library(tidyverse)
+## #https://www.coronavirus.vic.gov.au/victorian-coronavirus-covid-19-data
[…]

### 2022-06-27 cubble-jss.R

@@ -1,563 +1,0 @@
-## ----setup, echo = FALSE------------------------------------------------------
-knitr::opts_chunk$set(
-  echo = FALSE,
-  warning = FALSE,
-  message = FALSE,
-  cache = TRUE,
-  fig.path = "figures/"
-)
-
-options(prompt = "R> ", continue = "+ ",
-        tibble.print_max = 5, tibble.print_min = 5,
-        width = 70)
-
-
-## ----echo = FALSE-------------------------------------------------------------
-library(cubble)
-library(tidyverse)
-library(patchwork)
-
-
-## ----child = here::here("reference-hack.Rmd")---------------------------------
-
-## -----------------------------------------------------------------------------
-
-tex <- readLines(here::here("references.bib"))
-b <- tex %>% stringr::str_extract("\\@(article|techreport|misc|Manual|Book)\\{.+,")
-b <- b[!is.na(b)]
-ref_head <- b %>% stringr::str_replace("\\@(article|techreport|misc|Manual|Book)\\{", "") %>% 
-  stringr::str_replace(",", "") 
-res <- paste0("@", ref_head, collapse = ", ")
-
-
-
-## ----child = here::here("1-intro.Rmd")----------------------------------------
-
-
-
-
-## ----child = here::here("2-cube.Rmd")-----------------------------------------
-
-## ----cube-def, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "An illustration of the conceptual spatio-temporal cube with different slicing on time, site, and variable. For each axis, the slicing can be on a single value or a set of values."----
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.001.png"))
-
-
-
-## ----child = here::here("3-the-cubble-pkg.Rmd")-------------------------------
-
-## -----------------------------------------------------------------------------
-cubble::climate_flat
-
-
-## ----echo = TRUE--------------------------------------------------------------
-cubble_nested <- climate_flat |>
-  as_cubble(key = id, index = date, coords = c(long, lat))
-cubble_nested
-
-
-## ----echo = TRUE--------------------------------------------------------------
-cubble_nested |>
-  mutate(rain_day = sum(ts$prcp != 0))
-
-
-## ----echo = TRUE--------------------------------------------------------------
-cubble_long <- cubble_nested |> face_temporal()
-cubble_long
-
-
[…]

### 2022-06-27 1-intro.Rmd
visualisation -> visualization

@@
 The \pkg{stars} [@stars] package has a new implementation to use dense arrays to represent spatio-temporal cubes.
-It also interfaces with \pkg{sf} [@sf], a package commonly used for wrangling spatial data, and the \pkg{tidyverse} [@tidyverse] suite for general data wrangling and visualisation in \proglang{R}.
+It also interfaces with \pkg{sf} [@sf], a package commonly used for wrangling spatial data, and the \pkg{tidyverse} [@tidyverse] suite for general data wrangling and visualization in \proglang{R}.
 
@@
 According to the tidy data principles, data should be structured into 1) one row per observation, 2) one column per variable, and 3) one type of data per table.
-The long form data is preferred over wide data form given the downstream software such as \pkg{dplyr} [@dplyr] and \pkg{ggplot2} [@ggplot2] for data wrangling and visualisation.
+The long form data is preferred over wide data form given the downstream software such as \pkg{dplyr} [@dplyr] and \pkg{ggplot2} [@ggplot2] for data wrangling and visualization.
 However, the long form can be inefficient to store feature geometries, especially for large multipolygons for hourly, daily or sub-daily periods over years, which are extensively collected and handled, for example in time series analysis.
-This poses the question of how to arrange spatial and temporal variables in a way that would make data wrangling, visualizing and analysing spatio-temporal data easier.
+This poses the question of how to arrange spatial and temporal variables in a way that would make data wrangling, visualizing and analyzing spatio-temporal data easier.
 
 This paper presents a new \proglang{R} package, \pkg{cubble} which addresses the two issues mentioned above.
-In the package, a new data structure, also called \code{cubble}, is proposed to organise spatial and temporal variables as two forms of a single data object so that they can be wrangled separately or combined while being kept synchronised.
+In the package, a new data structure, also called \code{cubble}, is proposed to organize spatial and temporal variables as two forms of a single data object so that they can be wrangled separately or combined while being kept synchronized.
 Among the four spacetime layouts in @spacetime, \pkg{cubble} can be applied to full grid, sparse grid, or irregular, but not trajectory, which is outside the scope of this work.
@@
 
-The rest of the paper is organized as follows: Section \@ref(cube) introduces the proposed cube structure as a way to conceptualise multivariate spatio-temporal data.
+The rest of the paper is organized as follows: Section \@ref(cube) introduces the proposed cube structure as a way to conceptualize multivariate spatio-temporal data.
 Section \@ref(cubble) presents the main design and functionality of \pkg{cubble}.
-Section \@ref(others) explains how cubble deals with more advanced considerations, including data with hierarchical structure, data matching and how cubble fits with existing static and interactive visualisation tools.
+Section \@ref(others) explains how cubble deals with more advanced considerations, including data with hierarchical structure, data matching and how cubble fits with existing static and interactive visualization tools.
 Moreover we also illustrate how \pkg{cubble} deals with spatio-temporal data transformations.

### 2022-06-27 2-cube.Rmd
visualisation -> visualization

--- 2022-06-17 2-0-cube.Rmd
@@
 
-Spatio-temporal data can be conceptualised using a cubical data model with three axes which typically are, time, latitude and longitude.
-This abstraction can be useful for generalising operations and visualisation purposes: @lu_multidimensional_2018 shows how array operations (select, scale, reduce, rearrange, and compute) can be mapped onto the cube; @bach_review_2014 reviews the temporal data visualisation based on space-time cube operations.
+Spatio-temporal data can be conceptualized using a cubical data model with three axes which typically are, time, latitude and longitude.
+This abstraction can be useful for generalizing operations and visualization purposes: @lu_multidimensional_2018 shows how array operations (select, scale, reduce, rearrange, and compute) can be mapped onto the cube; @bach_review_2014 reviews the temporal data visualization based on space-time cube operations.
 Notice that the term space-time cube in their article "does not need to involve spatial data", but refers to "an abstract 2D substrate that is used to visualize data at a specific time".
-Despite its main focus being on temporal data, the mindset of abstracting out data representation to construct visualisations, still applies to our spatio-temporal data manipulation and visualisation approaches.
+Despite its main focus being on temporal data, the mindset of abstracting out data representation to construct visualizations, still applies to our spatio-temporal data manipulation and visualization approaches.
 
@@
 Finally, variables are stacked on this space-time canvas, with one observation per site and time point.
-This notion is adopted to avoid using hypercubes when describing multivariate spatio-temporal data and is the conceptual framework behind the \pkg{cubble} objects.
+This notion is adopted to avoid using hyper-cubes when describing multivariate spatio-temporal data and is the conceptual framework behind the \pkg{cubble} objects.
 With this conceptual model, operations on spatio-temporal data can be mapped to operations on the cube and the rest of Figure \ref{fig:cube-def} show examples of slicing on site, time, and variable.

### 2022-06-27 3-the-cubble-pkg.Rmd
visualisation -> visualization

@@
 
-A \code{cubble} object is an S3 class [@advr] built on the \code{tibble} class, specifically to organise spatio-temporal data.
+A \code{cubble} object is an S3 class [@advr] built on the \code{tibble} class, specifically to organize spatio-temporal data.
 The \code{cubble} class uses an attribute "form",  to arrange the spatial or temporal data components tidily.
@@
 Attributes  can be associated with dimensions or variables.
-Because there can be many different styles of representing this information there is a metadata convention [@hassell_data_2017] to standardise the format of the attributes.
+Because there can be many different styles of representing this information there is a metadata convention [@hassell_data_2017] to standardize the format of the attributes.
 A few packages in R exist for manipulating NetCDF data and these include a high-level R interface: \pkg{ncdf4} [@ncdf4], a low-level interface that calls a C-interface: \pkg{RNetCDF} [@rnetcdf], and a tidyverse implementation: \pkg{tidync} [@tidync].

### 2022-06-27 4-other-features.Rmd
visualisation -> visualization

@@
 cluster id).
-In this case, it can be useful to summarise variables at various levels of the hierarchy.
+In this case, it can be useful to summarize variables at various levels of the hierarchy.
 The function \code{switch_key()} can be used to change the grouping level of spatial locations.
@@
 Calendar plots [@wang2020calendar] deconstruct time to produce plots of variables in a calendar format.
-Summarising multiple variables is commonly done using projections, or linear combinations.
+Summarizing multiple variables is commonly done using projections, or linear combinations.
 Here we elaborate on the transformations made to produce a glyph map.

### 2022-06-27 5-1-covid.Rmd
visualisation -> visualization

@@
 Since the start of the pandemic, the Victoria State Government in Australia has provided daily COVID counts by local government area (LGA).
-This data can be used to visualise COVID incidence and spread spatially, when combined with map polygon data available from the Australian Bureau of Statistics.
+This data can be used to visualize COVID incidence and spread spatially, when combined with map polygon data available from the Australian Bureau of Statistics.
 These different sources need to be combined for the analysis, by matching the LGA names.

### 2022-06-27 5-3-river-level-rainfall.Rmd
visualisation -> visualization

@@
 
-This function returns a \code{cubble} object, with additional columns: \code{dist} storing the distance between matched stations, \code{group} summarising spatial matching, and \code{n_match} showing the temporal matching.
+This function returns a \code{cubble} object, with additional columns: \code{dist} storing the distance between matched stations, \code{group} summarizing spatial matching, and \code{n_match} showing the temporal matching.
 
@@
 
-Figure \ref{fig:matching} shows four matched pairs on the map (a) and standardised data as time series (b). The expected concurrent increase in precipitation and water level can be seen clearly.
+Figure \ref{fig:matching} shows four matched pairs on the map (a) and standardized data as time series (b). The expected concurrent increase in precipitation and water level can be seen clearly.
 

### 2022-06-27 5-4-era5.Rmd
visualisation -> visualization

@@
 <!--The goal of this example is to reproduce the break-up of the southern polar vortex that happened in late September and early October in 2002 in the stratosphere in the Southern Hemisphere.
-Therefore, we filter data corresponding to the Southern Hemisphere and restrict our analysis to four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04.--> The downloaded netcdf data is read into R using the \code{ncdf4} package, and converted to a \code{cubble}:
+Therefore, we filter data corresponding to the Southern Hemisphere and restrict our analysis to four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04.--> The downloaded netCDF data is read into R using the \code{ncdf4} package, and converted to a \code{cubble}:
 

### 2022-06-27 5-5-interactive-graphics.Rmd
visualisation -> visualization

@@
 
-The key steps are to convert both the nested and long forms of the data into shared objects, and to plot these sid-by-side. The two are linked by the station identifier.
+The key steps are to convert both the nested and long forms of the data into shared objects, and to plot these side-by-side. The two are linked by the station identifier.
 
@@
 Plot (a) shows the initial state of the interactive display: all locations are shown as dots on the map, coloured by temperature range, and the right plot shows the ribbons representing maximum to minimum for all stations.
-In plot (b) the "mount elizabeth" station, which shows a high variance colour on the initial map, is selected on the map and this produces the ribbon on the right.
-In plot (c) the lowest temperature in August is selected, which is "thredbo" station on the left map.
-It was surprising to us that this was not a station in Tasmania, so for comparison a station in Tasmania is selected on the map to show in relation to thredbo.
-We can see that thredbo has a bigger winter dip in temperature, and although Tasmania is cold generally, it's temperatures are more consistent.
+In plot (b) the "Mount Elizabeth" station, which shows a high variance colour on the initial map, is selected on the map and this produces the ribbon on the right.
+In plot (c) the lowest temperature in August is selected, which is "Thredbo" station on the left map.
+It was surprising to us that this was not a station in Tasmania, so for comparison a station in Tasmania is selected on the map to show in relation to Thredbo.
+We can see that Thredbo has a bigger winter dip in temperature, and although Tasmania is cold generally, it's temperatures are more consistent.
 

### 2022-06-27 6-conclusion.Rmd
misc small fixes
visualisation -> visualization

--- 2022-06-24 5-conclusion.Rmd
@@
 
-This paper presents an \proglang{R} package \pkg{cubble} for organising, manipulating and visualising spatio-temporal data.
+This paper presents an \proglang{R} package \pkg{cubble} for organizing, manipulating and visualizing spatio-temporal data.
 The package introduces a new data structure for spatio-temporal data, \code{cubble}, that connects the time invariant and varying variables and that allows the user to work with a nested and long form of the data.
@@
 Numerous examples are provided in the main text, appendix and package vignettes.
-These include creating and coercing data with mismatched names, handling hierarchical data, matching time series spatially and temporally, reproducing ERA5 plots from NetCDF data.
+These include creating and coercing data with mismatched names, handling hierarchical data, matching time series spatially and temporally, reproducing ERA5 plots from netCDF data.
 Visualization of the \code{cubble} objects can be done with interactive graphic pipelines using \pkg{crosstalk}, \pkg{plotly} and \pkg{leaflet}.
@@
 This work is funded by a Commonwealth Scientific and Industrial Research Organisation (CSIRO) Data61 Scholarship and started while Nicolas Langrené was affiliated with CSIRO's Data61.
-The article is created using \pkg{knitr} [@knitr] and \pkg{rmarkdown} [@rmarkdown] in R.
+The article is created using \pkg{knitr} [@knitr] and \pkg{rmarkdown} [@rmarkdown] in R using the `rticles::jss_article` template.
 The source code for reproducing this paper can be found at: https://github.com/huizezhang-sherry/paper-cubble.

### 2022-06-27 jss/main.Rmd
visualisation -> visualization

@@
 For many analyses, spatial and time components can be separately studied: for example, to explore the temporal trend of one variable for a single spatial location, or to model the spatial distribution of one variable at a given time.
-However for some studies, it is important to analyse different aspects of the spatio-temporal data simultaneouly, like for instance, temporal trends of multiple variables across locations.
+However for some studies, it is important to analyze different aspects of the spatio-temporal data simultaneously, like for instance, temporal trends of multiple variables across locations.
 In order to facilitate the study of different portions or combinations of spatio-temporal data, we introduce a new data structure, \code{cubble}, with a suite of functions enabling easy slicing and dicing on the different components spatio-temporal components.

### 2022-06-27 jss/main.tex
visualisation -> visualization

@@
 The \pkg{stars} \citep{stars} package has a new implementation to use dense arrays to represent spatio-temporal cubes.
-It also interfaces with \pkg{sf} \citep{sf}, a package commonly used for wrangling spatial data, and the \pkg{tidyverse} \citep{tidyverse} suite for general data wrangling and visualisation in \proglang{R}.
+It also interfaces with \pkg{sf} \citep{sf}, a package commonly used for wrangling spatial data, and the \pkg{tidyverse} \citep{tidyverse} suite for general data wrangling and visualization in \proglang{R}.
 
@@
 According to the tidy data principles, data should be structured into 1) one row per observation, 2) one column per variable, and 3) one type of data per table.
-The long form data is preferred over wide data form given the downstream software such as \pkg{dplyr} \citep{dplyr} and \pkg{ggplot2} \citep{ggplot2} for data wrangling and visualisation.
+The long form data is preferred over wide data form given the downstream software such as \pkg{dplyr} \citep{dplyr} and \pkg{ggplot2} \citep{ggplot2} for data wrangling and visualization.
 However, the long form can be inefficient to store feature geometries, especially for large multipolygons for hourly, daily or sub-daily periods over years, which are extensively collected and handled, for example in time series analysis.
-This poses the question of how to arrange spatial and temporal variables in a way that would make data wrangling, visualizing and analysing spatio-temporal data easier.
+This poses the question of how to arrange spatial and temporal variables in a way that would make data wrangling, visualizing and analyzing spatio-temporal data easier.
 
 This paper presents a new \proglang{R} package, \pkg{cubble} which addresses the two issues mentioned above.
-In the package, a new data structure, also called \code{cubble}, is proposed to organise spatial and temporal variables as two forms of a single data object so that they can be wrangled separately or combined while being kept synchronised.
+In the package, a new data structure, also called \code{cubble}, is proposed to organize spatial and temporal variables as two forms of a single data object so that they can be wrangled separately or combined while being kept synchronized.
 Among the four spacetime layouts in \citet{spacetime}, \pkg{cubble} can be applied to full grid, sparse grid, or irregular, but not trajectory, which is outside the scope of this work.
@@
 
-The rest of the paper is organized as follows: Section \ref{cube} introduces the proposed cube structure as a way to conceptualise multivariate spatio-temporal data.
+The rest of the paper is organized as follows: Section \ref{cube} introduces the proposed cube structure as a way to conceptualize multivariate spatio-temporal data.
 Section \ref{cubble} presents the main design and functionality of \pkg{cubble}.
-Section \ref{others} explains how cubble deals with more advanced considerations, including data with hierarchical structure, data matching and how cubble fits with existing static and interactive visualisation tools.
+Section \ref{others} explains how cubble deals with more advanced considerations, including data with hierarchical structure, data matching and how cubble fits with existing static and interactive visualization tools.
 Moreover we also illustrate how \pkg{cubble} deals with spatio-temporal data transformations.
@@
 
-Spatio-temporal data can be conceptualised using a cubical data model with three axes which typically are, time, latitude and longitude.
-This abstraction can be useful for generalising operations and visualisation purposes: \citet{lu_multidimensional_2018} shows how array operations (select, scale, reduce, rearrange, and compute) can be mapped onto the cube; \citet{bach_review_2014} reviews the temporal data visualisation based on space-time cube operations.
+Spatio-temporal data can be conceptualized using a cubical data model with three axes which typically are, time, latitude and longitude.
+This abstraction can be useful for generalizing operations and visualization purposes: \citet{lu_multidimensional_2018} shows how array operations (select, scale, reduce, rearrange, and compute) can be mapped onto the cube; \citet{bach_review_2014} reviews the temporal data visualization based on space-time cube operations.
 Notice that the term space-time cube in their article ``does not need to involve spatial data'', but refers to ``an abstract 2D substrate that is used to visualize data at a specific time''.
-Despite its main focus being on temporal data, the mindset of abstracting out data representation to construct visualisations, still applies to our spatio-temporal data manipulation and visualisation approaches.
+Despite its main focus being on temporal data, the mindset of abstracting out data representation to construct visualizations, still applies to our spatio-temporal data manipulation and visualization approaches.
 
@@
 
-A \code{cubble} object is an S3 class \citep{advr} built on the \code{tibble} class, specifically to organise spatio-temporal data.
+A \code{cubble} object is an S3 class \citep{advr} built on the \code{tibble} class, specifically to organize spatio-temporal data.
 The \code{cubble} class uses an attribute ``form'', to arrange the spatial or temporal data components tidily.
@@
 Attributes can be associated with dimensions or variables.
-Because there can be many different styles of representing this information there is a metadata convention \citep{hassell_data_2017} to standardise the format of the attributes.
+Because there can be many different styles of representing this information there is a metadata convention \citep{hassell_data_2017} to standardize the format of the attributes.
 A few packages in R exist for manipulating NetCDF data and these include a high-level R interface: \pkg{ncdf4} \citep{ncdf4}, a low-level interface that calls a C-interface: \pkg{RNetCDF} \citep{rnetcdf}, and a tidyverse implementation: \pkg{tidync} \citep{tidync}.
@@
 Spatial locations can have grouping structures either inherent to the data (e.g.~state within country) or obtained during the analysis (e.g.~cluster id).
-In this case, it can be useful to summarise variables at various levels of the hierarchy.
+In this case, it can be useful to summarize variables at various levels of the hierarchy.
 The function \code{switch_key()} can be used to change the grouping level of spatial locations.
@@
 Calendar plots \citep{wang2020calendar} deconstruct time to produce plots of variables in a calendar format.
-Summarising multiple variables is commonly done using projections, or linear combinations.
+Summarizing multiple variables is commonly done using projections, or linear combinations.
 Here we elaborate on the transformations made to produce a glyph map.
@@
 
-The downloaded netcdf data is read into R using the \code{ncdf4} package, and converted to a \code{cubble}:
+The downloaded netCDF data is read into R using the \code{ncdf4} package, and converted to a \code{cubble}:
 
@@
 Plot (a) shows the initial state of the interactive display: all locations are shown as dots on the map, coloured by temperature range, and the right plot shows the ribbons representing maximum to minimum for all stations.
-In plot (b) the ``mount elizabeth'' station, which shows a high variance colour on the initial map, is selected on the map and this produces the ribbon on the right.
-In plot (c) the lowest temperature in August is selected, which is ``thredbo'' station on the left map.
-It was surprising to us that this was not a station in Tasmania, so for comparison a station in Tasmania is selected on the map to show in relation to thredbo.
-We can see that thredbo has a bigger winter dip in temperature, and although Tasmania is cold generally, it's temperatures are more consistent.
+In plot (b) the ``Mount Elizabeth'' station, which shows a high variance colour on the initial map, is selected on the map and this produces the ribbon on the right.
+In plot (c) the lowest temperature in August is selected, which is ``Thredbo'' station on the left map.
+It was surprising to us that this was not a station in Tasmania, so for comparison a station in Tasmania is selected on the map to show in relation to Thredbo.
+We can see that Thredbo has a bigger winter dip in temperature, and although Tasmania is cold generally, it's temperatures are more consistent.
 

### 2022-07-09 3-the-cubble-pkg.Rmd
remove code comments in the manuscript

@@
 ```{r echo = TRUE}
-# add station-based features in the nested form.
 climate_flat_cb |> 
@@
 ```{r echo = TRUE, eval = FALSE}
-# read in the .nc file as a ncdf4 class
 raw <- ncdf4::nc_open(here::here("data/era5-pressure.nc"))
-
-# convert the variable q and z in the ncdf4 into a cubble
 dt <- as_cubble(raw, vars = c("q", "z"))
@@
 ```{r eval = FALSE, echo = TRUE}
-# Assume my_ncdf has a bounding box of [-180, -90, 180, 90]
-# at 0.25 degree resolution and subset it to have
-# 1 degree resolution:
-dt <- as_cubble(my_ncdf, vars = c("q", "z"),
-                long_range = seq(-180, 180, 1),
-                lat_range = seq(-90, 90, 1))
+dt <- as_cubble(
+  my_ncdf, vars = c("q", "z"),
+  long_range = seq(-180, 180, 1),lat_range = seq(-90, 90, 1)
+  )
 ```

### 2022-07-09 5-5-interactive-graphics.Rmd
remove code comments in the manuscript

@@
 ``` r
-# data pre-processing
 clean <- climate_full %>% ...
 
-# created SharedData instance for crosstalk
 nested <- clean %>% SharedData$new(~id, group = "cubble")
@@
 
-# create the spatial and temporal view each with a ShareData instance
 p1 <- nested %>% ...
@@
 
-# Combine p1 and p2
 crosstalk::bscols(plotly::ggplotly(p1), plotly::ggplotly(p2), ...)

### 2022-07-09 jss/appendix.Rmd
remove code comments in the manuscript

@@
 ``` r
-# data pre-processing
 clean <- climate_full %>% ...
 
-# use the long form to create subplots for each station
 df_id <- unique(clean$id)
@@
 
-# create nested form leaflet map with temperature band as subplots
 nested <- face_spatial(clean)
@@
 ```{r eval = FALSE}
-# all the Australian stations have all of the three PRCP, TMAX, and TMIN recorded
 aus_stations <- all_stations %>%

### 2022-07-09 jss/appendix.Rmd
use library("foo") style as suggested by JSS

@@
 knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
-library(tidyverse)
-library(cubble)
-library(patchwork)
-library(leaflet)
-library(leafpop)
+library("tidyverse")
+library("cubble")
+library("patchwork")
+library("leaflet")
+library("leafpop")
 ```

### 2022-07-09 4-other-features.Rmd
use library("foo") style as suggested by JSS
use "e.g.," instead of "e.g."

@@
 
-Spatial locations can have grouping structures either inherent to the data (e.g.
-state within country) or obtained during the analysis (e.g.
-cluster id).
+Spatial locations can have grouping structures either inherent to the data (e.g., state within country) or obtained during the analysis (e.g., cluster id).
 In this case, it can be useful to summarize variables at various levels of the hierarchy.
@@
 
-The cubble workflow fits works well with an interactive graphics pipeline (e.g.
-@buja1988elements, @buja1996interactive,  @sutherland2000orca, @xie2014reactive, @cheng2016enabling) that is available in R with the package crosstalk [@crosstalk].
+The cubble workflow fits works well with an interactive graphics pipeline (e.g., @buja1988elements, @buja1996interactive,  @sutherland2000orca, @xie2014reactive, @cheng2016enabling) that is available in R with the package crosstalk [@crosstalk].
 Figure \ref{fig:illu-interactive} illustrates how linking can be achieved between a map and multiple time series using a cubble.
@@
 ```{r eval = FALSE, echo = TRUE}
-library(ggplot2)
+library("ggplot2")
 gly <- glyphs(data,

### 2022-07-09 4-other-features.Rmd
bullet point punctuation

@@
 
- - \code{spatial_n_keep}: the number of spatial match for each site to keep
- - \code{spatial_dist_max}: the maximum distance allowed for a matched pair
- - \code{temporal_n_highest}: the number of peaks used - 3 in the example above
- - \code{temporal_window}: the length of the interval - 5 in the example above
- - \code{temporal_min_match}: the minimum number of matched peaks for a valid matched pair
+ - \code{spatial_n_keep}: the number of spatial match for each site to keep;
+ - \code{spatial_dist_max}: the maximum distance allowed for a matched pair;
+ - \code{temporal_n_highest}: the number of peaks used - 3 in the example above;
+ - \code{temporal_window}: the length of the interval - 5 in the example above; and
+ - \code{temporal_min_match}: the minimum number of matched peaks for a valid matched pair.
 

### 2022-07-09 5-1-covid.Rmd
use the new |> as suggested by JSS

@@
 
-covid <- raw %>%
-  count(date, lga, source, postcode) %>%
-  filter(date >= as.Date("2022-01-01")) %>%
-  filter(source == "Contact with a confirmed case") %>%
-  group_by(date, lga, source) %>%
-  count() %>%
-  ungroup(date) %>%
-  mutate(roll_mean = zoo::rollmean(n, k = 14, fill = NA)) %>%
-  tsibble::as_tsibble(key = lga, index = date) %>%
+covid <- raw |>
+  count(date, lga, source, postcode) |>
+  filter(date >= as.Date("2022-01-01")) |>
+  filter(source == "Contact with a confirmed case") |>
+  group_by(date, lga, source) |>
+  count() |>
+  ungroup(date) |>
+  mutate(roll_mean = zoo::rollmean(n, k = 14, fill = NA)) |>
+  tsibble::as_tsibble(key = lga, index = date) |>
   tsibble::fill_gaps(.full = TRUE)
@@
 
-lga <- strayr::read_absmap("lga2018") %>%
-  rename(lga = lga_name_2018) %>%
+lga <- strayr::read_absmap("lga2018") |>
+  rename(lga = lga_name_2018) |>
   filter(state_name_2016 == "Victoria") 
@@
 ```{r echo = TRUE}
-covid %>% head(5)
+covid |> head(5)
 ```
@@
 ```{r echo = TRUE}
-lga %>% head(5)
+lga |> head(5)
 ```
@@
 ```{r echo = TRUE}
-lga <- lga %>%
+lga <- lga |>
   mutate(lga = ifelse(lga == "Kingston (C) (Vic.)", "Kingston (C)", lga),
-         lga = ifelse(lga == "Latrobe (C) (Vic.)", "Latrobe (C)", lga)) %>%
+         lga = ifelse(lga == "Latrobe (C) (Vic.)", "Latrobe (C)", lga)) |>
   filter(!lga %in% pair$others$spatial)
 
-covid <- covid %>% filter(!lga %in% pair$others$temporal)
+covid <- covid |> filter(!lga %in% pair$others$temporal)
 

### 2022-07-09 5-2-aus-hist-max-temp.Rmd
use the new |> as suggested by JSS

@@
 ```{r echo = TRUE}
-tmax <- historical_tmax %>%
+tmax <- historical_tmax |>
   filter(between(stringr::str_sub(id, 7, 8), 46, 90))
@@
 ```{r echo = TRUE}
-tmax <- tmax %>%
-  face_temporal() %>% 
+tmax <- tmax |>
+  face_temporal() |> 
   group_by(month = lubridate::month(date),
@@
            ifelse(lubridate::year(date) > 2015,
-           "2016 ~ 2020", "1971 ~ 1975"))) %>%
+           "2016 ~ 2020", "1971 ~ 1975"))) |>
   summarise(tmax = mean(tmax, na.rm = TRUE))
@@
 ```{r echo = TRUE}
-tmax %>% 
-  face_spatial() %>% 
-  mutate(n = nrow(ts)) %>%
-  arrange(n) %>%
-  pull(n) %>% 
+tmax |> 
+  face_spatial() |> 
+  mutate(n = nrow(ts)) |>
+  arrange(n) |>
+  pull(n) |> 
   head(10)
 
-tmax <- tmax %>% 
-  face_spatial() %>% 
-  filter(nrow(ts) == 24) %>%
+tmax <- tmax |> 
+  face_spatial() |> 
+  filter(nrow(ts) == 24) |>
   face_temporal()
@@
 ```{r eval = FALSE, echo = TRUE}
-nsw_vic <- ozmaps::abs_ste %>%
+nsw_vic <- ozmaps::abs_ste |>
   filter(NAME %in% c("Victoria", "New South Wales"))
@@
 ```{r}
-cobar <- tmax %>% filter(id == "ASN00048027") %>%
+cobar <- tmax |> filter(id == "ASN00048027") |>
   ggplot(aes(x = month, y = tmax, color = group)) +
@@
 From the glyph map we can observe that the monthly trend is similar for all locations (low in the winter, high in the summer), and small increased temperatures, particularly in late summer can be seen at most stations in New South Wales."}
-tmax <- tmax %>% unfold(long, lat)
-box_df <- tmax %>% face_spatial() %>% filter(id == "ASN00048027")
-nsw_vic <- ozmaps::abs_ste %>% filter(NAME %in% c("Victoria","New South Wales"))
+tmax <- tmax |> unfold(long, lat)
+box_df <- tmax |> face_spatial() |> filter(id == "ASN00048027")
+nsw_vic <- ozmaps::abs_ste |> filter(NAME %in% c("Victoria","New South Wales"))
 

### 2022-07-09 5-3-river-level-rainfall.Rmd
use the new |> as suggested by JSS

@@
 
-climate <- climate_full %>%
-  filter(between(stringr::str_sub(id, 7, 8), 76, 90)) %>%
-  face_temporal() %>%
-  filter(lubridate::year(date) == 2020) %>%
-  face_spatial() %>%
+climate <- climate_full |>
+  filter(between(stringr::str_sub(id, 7, 8), 76, 90)) |>
+  face_temporal() |>
+  filter(lubridate::year(date) == 2020) |>
+  face_spatial() |>
   mutate(type = "climate")
 
-river <- river %>% mutate(type = "river")
+river <- river |> mutate(type = "river")
 ```
@@
 ```{r}
-res %>% print()
+res |> print()
 ```
@@
 The numbers (1, 5, 6, 10) indicate the group index derived from spatial matching, only those that were selectd by temporal matching are shown here."}
-vic_map <- ozmaps::abs_ste %>% 
-  filter(NAME == "Victoria") %>% 
+vic_map <- ozmaps::abs_ste |> 
+  filter(NAME == "Victoria") |> 
   rmapshaper::ms_simplify()
@@
   ggrepel::geom_label_repel(
-    data = res %>% filter(type == "river"), aes(x = long, y = lat, label = group)) +
+    data = res |> filter(type == "river"), aes(x = long, y = lat, label = group)) +
   scale_color_brewer(palette = "Dark2")  + 
@@
 
-res_long <- res %>%
-  face_temporal(ts) %>%
-  unfold(group, type) %>%
-  rename(prcp = matched_var) %>% 
+res_long <- res |>
+  face_temporal(ts) |>
+  unfold(group, type) |>
+  rename(prcp = matched_var) |> 
   mutate(prcp = (prcp - min(prcp, na.rm = TRUE))/ (max(prcp, na.rm = TRUE) - min(prcp, na.rm = TRUE)))
 
-p2 <- res_long %>% 
+p2 <- res_long |> 
   ggplot(aes(x = date, y = prcp, color = type, group = id)) +

### 2022-07-09 5-4-era5.Rmd
use the new |> as suggested by JSS

@@
 The polar vortex, signalled by the high speicfic humidity, splits into two on 2002-09-26 and furthers split into four on 2002-10-04."}
-date <- c("2002-09-22", "2002-09-26", "2002-09-30", "2002-10-04") %>% as.Date()
-res <- dt %>% 
-  face_temporal() %>% 
-  filter(lubridate::date(time) %in% date) %>%
-  unfold(long, lat) %>% 
+date <- c("2002-09-22", "2002-09-26", "2002-09-30", "2002-10-04") |> as.Date()
+res <- dt |> 
+  face_temporal() |> 
+  filter(lubridate::date(time) %in% date) |>
+  unfold(long, lat) |> 
   mutate(q =  q* 10^6)
@@
 sf::sf_use_s2(FALSE)
-country <- rnaturalearth::ne_coastline("small", returnclass = "sf") %>% 
-  sf::st_crop(box) %>% 
+country <- rnaturalearth::ne_coastline("small", returnclass = "sf") |> 
+  sf::st_crop(box) |> 
   sf::st_cast("MULTILINESTRING")
 
-res %>% 
+res |> 
   ggplot() +

### 2022-07-09 5-5-interactive-graphics.Rmd
use library("foo") style as suggested by JSS
use the new |> as suggested by JSS

@@
 ``` r
-clean <- climate_full %>% ...
+clean <- climate_full |> ...
 
-nested <- clean %>% SharedData$new(~id, group = "cubble")
-long <- face_temporal(clean) %>% SharedData$new(~id, group = "cubble")
+nested <- clean |> SharedData$new(~id, group = "cubble")
+long <- face_temporal(clean) |> SharedData$new(~id, group = "cubble")
 
-p1 <- nested %>% ...
-p2 <- long %>% ...
+p1 <- nested |> ...
+p2 <- long |> ...
 
@@
 # This chunk generates the interactive graphic and save it as figures/linking.html, where the three screencasts in the figure comes from.
-library(crosstalk)
-library(plotly)
+library("crosstalk")
+library("plotly")
 
-clean <- climate_full %>%
+clean <- climate_full |>
   filter(!id %in% c("ASN00067033", "ASN00072091", "ASN00059040", 
-                    "ASN00097053", "ASN00040856", "ASN00015548")) %>%
-  face_temporal() %>% 
-  mutate(month = lubridate::month(date)) %>%
-  group_by(month) %>%
+                    "ASN00097053", "ASN00040856", "ASN00015548")) |>
+  face_temporal() |> 
+  mutate(month = lubridate::month(date)) |>
+  group_by(month) |>
   summarise(
@@
     diff = mean(tmax - tmin, na.rm = TRUE)
-    ) %>% 
-  ungroup(month) %>%
-  mutate(dummy_date = as.Date(glue::glue("2021-{month}-01"))) %>% 
-  face_spatial() %>% 
-  mutate(temp_diff_var = ts %>% pull(diff) %>% var(na.rm = TRUE))
+    ) |> 
+  ungroup(month) |>
+  mutate(dummy_date = as.Date(glue::glue("2021-{month}-01"))) |> 
+  face_spatial() |> 
+  mutate(temp_diff_var = ts |> pull(diff) |> var(na.rm = TRUE))
 
-nested <- clean %>% 
+nested <- clean |> 
   SharedData$new(~id, group = "cubble")
 
-long <- clean %>% 
-  face_temporal() %>% 
-  unfold(temp_diff_var) %>% 
-  arrange(temp_diff_var) %>% 
+long <- clean |> 
+  face_temporal() |> 
+  unfold(temp_diff_var) |> 
+  arrange(temp_diff_var) |> 
   SharedData$new(~id, group = "cubble")
@@
 
-p1 <- nested %>% 
+p1 <- nested |> 
   ggplot() +
@@
 
-p2 <- long %>% 
+p2 <- long |> 
   ggplot(aes(x = dummy_date, group = id, 
@@
 out <- bscols(
-  ggplotly(p1, width = 900, height = 700, tooltip = "label") %>%
+  ggplotly(p1, width = 900, height = 700, tooltip = "label") |>
     highlight(
@@
       opacityDim = 0.1),
-  ggplotly(p2, width = 1200, height = 700, tooltip = "label") %>% 
+  ggplotly(p2, width = 1200, height = 700, tooltip = "label") |> 
     highlight(

### 2022-07-09 jss/appendix.Rmd
use the new |> as suggested by JSS

@@
 ```{r eval = FALSE, echo = TRUE}
-station_nested <- climate_full %>% mutate(cluster = ...)
+station_nested <- climate_full |> mutate(cluster = ...)
 ```
@@
 ```{r eval = FALSE, echo = TRUE}
-cluster_nested <- station_nested %>% switch_key(cluster)
+cluster_nested <- station_nested |> switch_key(cluster)
 ```
@@
 
-station_long <- cubble::prcp_aus %>%
-  face_temporal(ts) %>%
-  mutate(wk = lubridate::week(date)) %>%
-  group_by(wk) %>%
-  summarise(prcp = sum(prcp, na.rm = TRUE)) %>%
+station_long <- cubble::prcp_aus |>
+  face_temporal(ts) |>
+  mutate(wk = lubridate::week(date)) |>
+  group_by(wk) |>
+  summarise(prcp = sum(prcp, na.rm = TRUE)) |>
   unfold(cluster)
@@
 set.seed(123)
-station_nested <- station_long %>%
-  face_spatial() %>%
-  strip_rowwise() %>%
+station_nested <- station_long |>
+  face_spatial() |>
+  strip_rowwise() |>
   mutate(cluster = kmeans(dist_raw,centers = 20, nstart = 500)$cluster)
@@
 load(here::here("data/station_nested.rda"))
-cluster_nested <- station_nested %>%
-  switch_key(cluster) %>% 
+cluster_nested <- station_nested |>
+  switch_key(cluster) |> 
   get_centroid()
 
-cluster_long <- cluster_nested %>% 
-  face_temporal() %>% 
-  group_by(wk) %>%
-  summarise(prcp = mean(prcp, na.rm = TRUE)) %>% 
+cluster_long <- cluster_nested |> 
+  face_temporal() |> 
+  group_by(wk) |>
+  summarise(prcp = mean(prcp, na.rm = TRUE)) |> 
   unfold(cent_long, cent_lat)
@@
 
-ggplot_smooth <- cluster_long %>% 
+ggplot_smooth <- cluster_long |> 
   ggplot() +
@@
 
-smoother <- layer_data(ggplot_smooth) %>% 
-  left_join(cluster_long %>% select(cluster, cent_long, cent_lat), by = c("group" = "cluster"))
+smoother <- layer_data(ggplot_smooth) |> 
+  left_join(cluster_long |> select(cluster, cent_long, cent_lat), by = c("group" = "cluster"))
 
@@
   geom_point(data = station_nested, aes(x = long, y = lat), size = 0.5) +
-  ggforce::geom_mark_hull(data = cluster_nested %>% tidyr::unnest(hull),
+  ggforce::geom_mark_hull(data = cluster_nested |> tidyr::unnest(hull),
                           expand = 0, radius = 0,
@@
 ``` r
-clean <- climate_full %>% ...
+clean <- climate_full |> ...
 
@@
 ```{r eval = FALSE, echo = FALSE}
-clean <-  climate_full %>%
+clean <-  climate_full |>
   filter(!id %in% c("ASN00067033", "ASN00072091", "ASN00059040", 
-                    "ASN00097053", "ASN00040856", "ASN00015548")) %>%
-  face_temporal() %>%
-  mutate(month = lubridate::month(date)) %>% 
-  group_by(month) %>% 
+                    "ASN00097053", "ASN00040856", "ASN00015548")) |>
+  face_temporal() |>
+  mutate(month = lubridate::month(date)) |> 
+  group_by(month) |> 
   summarise(tmax = mean(tmax, na.rm = TRUE),
             tmin = mean(tmin, na.rm = TRUE),
-            diff = mean(tmax - tmin, na.rm = TRUE)) %>% 
-  ungroup(month) %>% 
-  face_spatial() %>%
-  mutate(temp_diff_var = ts %>% pull(diff) %>% var(na.rm = TRUE)) %>% 
-  face_temporal() %>% 
-  mutate(dummy_date = as.Date(glue::glue("2021-{month}-01"))) %>% 
+            diff = mean(tmax - tmin, na.rm = TRUE)) |> 
+  ungroup(month) |> 
+  face_spatial() |>
+  mutate(temp_diff_var = ts |> pull(diff) |> var(na.rm = TRUE)) |> 
+  face_temporal() |> 
+  mutate(dummy_date = as.Date(glue::glue("2021-{month}-01"))) |> 
   unfold(name)
 
-df_id <- clean$id %>% unique()
+df_id <- clean$id |> unique()
 p <- purrr::map(1:length(df_id), function(i){
-  dt <- clean %>% filter(id == df_id[i])
-  dt %>% 
+  dt <- clean |> filter(id == df_id[i])
+  dt |> 
     ggplot(aes(x = dummy_date, ymin = tmin, ymax = tmax, group = id)) + 
@@
 
-leaflet(nested) %>% 
-  addTiles() %>% 
+leaflet(nested) |> 
+  addTiles() |> 
   addCircleMarkers(color = ~pal(domain), radius = 0.1, 
@@
 library(cubble)
-all_stations <- rnoaa::ghcnd_stations() %>%
-  filter(str_starts(id, "ASN")) %>% # Australian stations start wiht "ASN"
-  filter(last_year >= 2020) %>%
-  mutate(wmo_id = as.numeric(wmo_id), name = str_to_lower(name)) %>%
-  select(-state, -gsn_flag) %>%
+all_stations <- rnoaa::ghcnd_stations() |>
+  filter(str_starts(id, "ASN")) |> # Australian stations start wiht "ASN"
+  filter(last_year >= 2020) |>
+  mutate(wmo_id = as.numeric(wmo_id), name = str_to_lower(name)) |>
+  select(-state, -gsn_flag) |>
   select(id, longitude, latitude, elevation, name,
-         wmo_id, element, first_year, last_year) %>%
+         wmo_id, element, first_year, last_year) |>
   rename(long = longitude, lat = latitude, elev = elevation)
 
-tmax_stations <- all_stations %>% 
+tmax_stations <- all_stations |> 
   filter(element == "TMAX", first_year < 1970, !is.na(wmo_id))
 
-raw_tmax <- all_stations %>%
-  rowwise() %>%
+raw_tmax <- all_stations |>
+  rowwise() |>
   mutate(ts = list(rnoaa::meteo_pull_monitors(
@@
     date_max = glue::glue("{last_year}-12-31")
-    ) %>%
+    ) |>
       select(-id)
@@
 
-historical_tmax <- raw_tmax %>%
-  select(-element) %>%
-  unnest(ts) %>%
-  mutate(tmax = tmax/10) %>%
-  filter(lubridate::year(date) %in% c(1971: 1975, 2016:2020)) %>% 
+historical_tmax <- raw_tmax |>
+  select(-element) |>
+  unnest(ts) |>
+  mutate(tmax = tmax/10) |>
+  filter(lubridate::year(date) %in% c(1971: 1975, 2016:2020)) |> 
   as_cubble(index = date, key = id, coords = c(long, lat))
@@
 ```{r eval = FALSE}
-aus_stations <- all_stations %>%
-  filter(element %in% c("PRCP", "TMAX", "TMIN")) %>%
-  nest(element: last_year) %>%
-  rowwise() %>%
-  filter(nrow(data) == 3) %>%
+aus_stations <- all_stations |>
+  filter(element %in% c("PRCP", "TMAX", "TMIN")) |>
+  nest(element: last_year) |>
+  rowwise() |>
+  filter(nrow(data) == 3) |>
   select(-data)
 
-aus_climate_raw <- aus_stations %>%
-  rowwise() %>%
+aus_climate_raw <- aus_stations |>
+  rowwise() |>
   mutate(ts = list(
@@
       date_min = "2016-01-01", date_max = "2020-12-31"
-      ) %>%
+      ) |>
       select(-id)
@@
 
-climate_full <- aus_climate_raw %>%
-  unnest(ts) %>%
-  mutate(tmax = tmax/10, tmin = tmin/10) %>%
+climate_full <- aus_climate_raw |>
+  unnest(ts) |>
+  mutate(tmax = tmax/10, tmin = tmin/10) |>
   cubble::as_cubble(key = id, index = date, coords = c(long, lat))

### 2022-07-09 1-intro.Rmd
check on \proglang, \pkg, \code being used throughout the paper

@@
 
-The SpatioTemporal CRAN task view [@ctvspatiotemporal] gathers information about R packages designed for spatio-temporal data and it has a section on *Representing data* that lists existing spatio-temporal data representations used in \proglang{R}.
+The SpatioTemporal CRAN task view [@ctvspatiotemporal] gathers information about \proglang{R} packages designed for spatio-temporal data and it has a section on *Representing data* that lists existing spatio-temporal data representations used in \proglang{R}.
 Among them, @spacetime summarises spatio-temporal data into three forms: time-wide, space-wide, and long formats.
@@
 Among the four spacetime layouts in @spacetime, \pkg{cubble} can be applied to full grid, sparse grid, or irregular, but not trajectory, which is outside the scope of this work.
-The software is available from the Comprehensive R Archive Network (CRAN) at https://CRAN.R-project.org/package=cubble.
+The software is available from the Comprehensive \proglang{R} Archive Network (CRAN) at https://CRAN.R-project.org/package=cubble.
 

### 2022-07-09 3-the-cubble-pkg.Rmd
check on \proglang, \pkg, \code being used throughout the paper

@@
 
-A cubble can be created from various common spatio-temporal data formats, including basic R objects like tibble, tsibble and sf.
+A cubble can be created from various common spatio-temporal data formats, including basic \proglang{R} objects like tibble, tsibble and sf.
 Section \@ref(covid) describes converting multiple tables into a cubble object and Section \@ref(netcdf) illustrates how to convert a netCDF object.
@@
 Because there can be many different styles of representing this information there is a metadata convention [@hassell_data_2017] to standardize the format of the attributes.
-A few packages in R exist for manipulating NetCDF data and these include a high-level R interface: \pkg{ncdf4} [@ncdf4], a low-level interface that calls a C-interface: \pkg{RNetCDF} [@rnetcdf], and a tidyverse implementation: \pkg{tidync} [@tidync].
+A few packages in \proglang{R} exist for manipulating NetCDF data and these include a high-level R interface: \pkg{ncdf4} [@ncdf4], a low-level interface that calls a C-interface: \pkg{RNetCDF} [@rnetcdf], and a tidyverse implementation: \pkg{tidync} [@tidync].
 

### 2022-07-09 4-other-features.Rmd
check on \proglang, \pkg, \code being used throughout the paper

@@
 
-The cubble workflow fits works well with an interactive graphics pipeline (e.g., @buja1988elements, @buja1996interactive,  @sutherland2000orca, @xie2014reactive, @cheng2016enabling) that is available in R with the package crosstalk [@crosstalk].
+The cubble workflow fits works well with an interactive graphics pipeline (e.g., @buja1988elements, @buja1996interactive,  @sutherland2000orca, @xie2014reactive, @cheng2016enabling) that is available in \proglang{R} with the package crosstalk [@crosstalk].
 Figure \ref{fig:illu-interactive} illustrates how linking can be achieved between a map and multiple time series using a cubble.

### 2022-07-09 5-2-aus-hist-max-temp.Rmd
check on \proglang, \pkg, \code being used throughout the paper

@@
 The Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world.
-The data used here (\code{historical_tmax}) is a subset extracted using the `rnoaa` package, containing the records of maximum temperature for `r nrow(historical_tmax)` Australian stations from `r range(historical_tmax$first_year)[1]` through `r range(historical_tmax$first_year)[2]` and provides information also on the latitude, longitude and elevation of each of the  stations.
+The data used here (\code{historical_tmax}) is a subset extracted using the \pkg{rnoaa}[@rnoaa] package, containing the records of maximum temperature for `r nrow(historical_tmax)` Australian stations from `r range(historical_tmax$first_year)[1]` through `r range(historical_tmax$first_year)[2]` and provides information also on the latitude, longitude and elevation of each of the  stations.
 <!-- This data  is already cast into a \code{cubble} data object (please see Appendix X for the code to extract  the data from XX and prepare the data into \code{cubble} form), with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates.--> The goal of this example is to compare the monthly average maximum temperature between two periods, 1971-1975 and 2016-2020, for stations in Victoria and New South Wales, using a glyph map.

### 2022-07-09 5-4-era5.Rmd
check on \proglang, \pkg, \code being used throughout the paper

@@
 The data is available in the NetCDF format from the European Centre for Medium-Range Weather Forecasts (ECMWF).
-It can be directly downloaded from Copernicus Climate Data Store (CDS) [@cds] website or via the R package \pkg{ecmwfr} [@ecwmfr].
+It can be directly downloaded from Copernicus Climate Data Store (CDS) [@cds] website or via the \proglang{R} package \pkg{ecmwfr} [@ecwmfr].
 
@@
 <!--The goal of this example is to reproduce the break-up of the southern polar vortex that happened in late September and early October in 2002 in the stratosphere in the Southern Hemisphere.
-Therefore, we filter data corresponding to the Southern Hemisphere and restrict our analysis to four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04.--> The downloaded netCDF data is read into R using the \code{ncdf4} package, and converted to a \code{cubble}:
+Therefore, we filter data corresponding to the Southern Hemisphere and restrict our analysis to four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04.--> The downloaded netCDF data is read into \proglang{R} using the \code{ncdf4} package, and converted to a \code{cubble}:
 

### 2022-07-09 6-conclusion.Rmd
check on \proglang, \pkg, \code being used throughout the paper

@@
 This work adds capabilities into the spatio-temporal practitioners toolbox to integrate it with a tidy data framework.
-The data structure and functions introduced in this package can be used and combined with existing spatio-temporal R packages such as \pkg{sf}, data wrangling packages such as \pkg{dplyr}, and visualization packages such as \pkg{ggplot2}, \pkg{plotly} and \pkg{leaflet}.
+The data structure and functions introduced in this package can be used and combined with existing spatio-temporal \proglang{R} packages such as \pkg{sf}, data wrangling packages such as \pkg{dplyr}, and visualization packages such as \pkg{ggplot2}, \pkg{plotly} and \pkg{leaflet}.
 
@@
 This work is funded by a Commonwealth Scientific and Industrial Research Organisation (CSIRO) Data61 Scholarship and started while Nicolas Langrené was affiliated with CSIRO's Data61.
-The article is created using \pkg{knitr} [@knitr] and \pkg{rmarkdown} [@rmarkdown] in R using the `rticles::jss_article` template.
+The article is created using \pkg{knitr} [@knitr] and \pkg{rmarkdown} [@rmarkdown] in R using the \code{rticles::jss_article} template.
 The source code for reproducing this paper can be found at: https://github.com/huizezhang-sherry/paper-cubble.

### 2022-07-09 jss/appendix.Rmd
check on \proglang, \pkg, \code being used throughout the paper

@@
 In some cases it is better to display such information at an aggregated level by grouping the data adequately before exhibiting the information on a figure.
-The example below shows how can  spatio-temporal data be organised at different hierarchical levels so that information can be grouped both temporal and spatially using  `switch_key()`.
+The example below shows how can  spatio-temporal data be organised at different hierarchical levels so that information can be grouped both temporal and spatially using  \code{switch_key()}.
 The goal of this example is to highlight how easy is to move across the data hierarchy using **cubble**.

### 2022-07-09 jss/main.Rmd
check on \proglang, \pkg, \code being used throughout the paper

@@
 title:
-  formatted: "\\pkg{cubble}: An R Package for Organizing and Wrangling Multivariate Spatio-temporal Data"
+  formatted: "\\pkg{cubble}: An \proglang{R} Package for Organizing and Wrangling Multivariate Spatio-temporal Data"
   # If you use tex in the formatted title, alsoge supply version without
@@
 In addition, \code{cubble} facilitates visual and numerical explorations of the data while easing data wrangling and modelling.
-The \code{cubble} structure and the functions provided in the \code{cubble} R package equip users with the capability to handle hierarchical spatial and temporal structures.
+The \code{cubble} structure and the functions provided in the \code{cubble} \proglang{R} package equip users with the capability to handle hierarchical spatial and temporal structures.
 The \code{cubble} structure and the tools implemented in the package are illustrated with different examples of Australian climate data.

### 2022-07-09 jss/main.Rmd
knit

@@
 title:
-  formatted: "\\pkg{cubble}: An \proglang{R} Package for Organizing and Wrangling Multivariate Spatio-temporal Data"
+  formatted: "\\pkg{cubble}: An \\proglang{R} Package for Organizing and Wrangling Multivariate Spatio-temporal Data"
   # If you use tex in the formatted title, alsoge supply version without

### 2022-07-09 jss/main.tex
knit

@@
 }
-\title{\pkg{cubble}: An R Package for Organizing and Wrangling Multivariate Spatio-temporal Data}
+\title{\pkg{cubble}: An \proglang{R} Package for Organizing and Wrangling Multivariate Spatio-temporal Data}
 
@@
 
-The SpatioTemporal CRAN task view \citep{ctvspatiotemporal} gathers information about R packages designed for spatio-temporal data and it has a section on \emph{Representing data} that lists existing spatio-temporal data representations used in \proglang{R}.
+The SpatioTemporal CRAN task view \citep{ctvspatiotemporal} gathers information about \proglang{R} packages designed for spatio-temporal data and it has a section on \emph{Representing data} that lists existing spatio-temporal data representations used in \proglang{R}.
 Among them, \citet{spacetime} summarises spatio-temporal data into three forms: time-wide, space-wide, and long formats.
@@
 Among the four spacetime layouts in \citet{spacetime}, \pkg{cubble} can be applied to full grid, sparse grid, or irregular, but not trajectory, which is outside the scope of this work.
-The software is available from the Comprehensive R Archive Network (CRAN) at \url{https://CRAN.R-project.org/package=cubble}.
+The software is available from the Comprehensive \proglang{R} Archive Network (CRAN) at \url{https://CRAN.R-project.org/package=cubble}.
 
@@
 
-A cubble can be created from various common spatio-temporal data formats, including basic R objects like tibble, tsibble and sf.
+A cubble can be created from various common spatio-temporal data formats, including basic \proglang{R} objects like tibble, tsibble and sf.
 Section \ref{covid} describes converting multiple tables into a cubble object and Section \ref{netcdf} illustrates how to convert a netCDF object.
@@
 \begin{CodeInput}
-R> # add station-based features in the nested form.
 R> climate_flat_cb |> 
@@
 Because there can be many different styles of representing this information there is a metadata convention \citep{hassell_data_2017} to standardize the format of the attributes.
-A few packages in R exist for manipulating NetCDF data and these include a high-level R interface: \pkg{ncdf4} \citep{ncdf4}, a low-level interface that calls a C-interface: \pkg{RNetCDF} \citep{rnetcdf}, and a tidyverse implementation: \pkg{tidync} \citep{tidync}.
+A few packages in \proglang{R} exist for manipulating NetCDF data and these include a high-level R interface: \pkg{ncdf4} \citep{ncdf4}, a low-level interface that calls a C-interface: \pkg{RNetCDF} \citep{rnetcdf}, and a tidyverse implementation: \pkg{tidync} \citep{tidync}.
 
@@
 \begin{CodeInput}
-R> # read in the .nc file as a ncdf4 class
 R> raw <- ncdf4::nc_open(here::here("data/era5-pressure.nc"))
-R> 
-R> # convert the variable q and z in the ncdf4 into a cubble
 R> dt <- as_cubble(raw, vars = c("q", "z"))
@@
 \begin{CodeInput}
-R> # Assume my_ncdf has a bounding box of [-180, -90, 180, 90]
-R> # at 0.25 degree resolution and subset it to have
-R> # 1 degree resolution:
-R> dt <- as_cubble(my_ncdf, vars = c("q", "z"),
-+                 long_range = seq(-180, 180, 1),
-+                 lat_range = seq(-90, 90, 1))
+R> dt <- as_cubble(
++   my_ncdf, vars = c("q", "z"),
++   long_range = seq(-180, 180, 1),lat_range = seq(-90, 90, 1)
++   )
 \end{CodeInput}
@@
 
-Spatial locations can have grouping structures either inherent to the data (e.g.~state within country) or obtained during the analysis (e.g.~cluster id).
+Spatial locations can have grouping structures either inherent to the data (e.g., state within country) or obtained during the analysis (e.g., cluster id).
 In this case, it can be useful to summarize variables at various levels of the hierarchy.
@@
 \item
-  \code{spatial_n_keep}: the number of spatial match for each site to keep
+  \code{spatial_n_keep}: the number of spatial match for each site to keep;
 \item
-  \code{spatial_dist_max}: the maximum distance allowed for a matched pair
+  \code{spatial_dist_max}: the maximum distance allowed for a matched pair;
 \item
-  \code{temporal_n_highest}: the number of peaks used - 3 in the example above
+  \code{temporal_n_highest}: the number of peaks used - 3 in the example above;
 \item
-  \code{temporal_window}: the length of the interval - 5 in the example above
+  \code{temporal_window}: the length of the interval - 5 in the example above; and
 \item
-  \code{temporal_min_match}: the minimum number of matched peaks for a valid matched pair
+  \code{temporal_min_match}: the minimum number of matched peaks for a valid matched pair.
 \end{itemize}
@@
 
-The cubble workflow fits works well with an interactive graphics pipeline (e.g.
-\citet{buja1988elements}, \citet{buja1996interactive}, \citet{sutherland2000orca}, \citet{xie2014reactive}, \citet{cheng2016enabling}) that is available in R with the package crosstalk \citep{crosstalk}.
+The cubble workflow fits works well with an interactive graphics pipeline (e.g., \citet{buja1988elements}, \citet{buja1996interactive}, \citet{sutherland2000orca}, \citet{xie2014reactive}, \citet{cheng2016enabling}) that is available in \proglang{R} with the package crosstalk \citep{crosstalk}.
 Figure \ref{fig:illu-interactive} illustrates how linking can be achieved between a map and multiple time series using a cubble.
@@
 \begin{CodeInput}
-R> library(ggplot2)
+R> library("ggplot2")
 R> gly <- glyphs(data,
@@
 Here is how to do this with \pkg{cubble}, including how to handle mismatches arising from different names of the same LGAs in the two tables.
-The COVID data is store in a \texttt{csv} file and looks like:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> covid %>% head(5)
+The COVID data is store in a csv file and looks like:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> covid |> head(5)
 \end{CodeInput}
@@
 \begin{CodeInput}
-R> lga %>% head(5)
+R> lga |> head(5)
 \end{CodeInput}
@@
 \begin{CodeInput}
-R> lga <- lga %>%
+R> lga <- lga |>
 +   mutate(lga = ifelse(lga == "Kingston (C) (Vic.)", "Kingston (C)", lga),
-+          lga = ifelse(lga == "Latrobe (C) (Vic.)", "Latrobe (C)", lga)) %>%
++          lga = ifelse(lga == "Latrobe (C) (Vic.)", "Latrobe (C)", lga)) |>
 +   filter(!lga %in% pair$others$spatial)
 R> 
-R> covid <- covid %>% filter(!lga %in% pair$others$temporal)
+R> covid <- covid |> filter(!lga %in% pair$others$temporal)
 R> 
@@
 The Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world.
-The data used here (\code{historical_tmax}) is a subset extracted using the \texttt{rnoaa} package, containing the records of maximum temperature for 236 Australian stations from 1859 through 1969 and provides information also on the latitude, longitude and elevation of each of the stations.
+The data used here (\code{historical_tmax}) is a subset extracted using the \pkg{rnoaa}\citep{rnoaa} package, containing the records of maximum temperature for 236 Australian stations from 1859 through 1969 and provides information also on the latitude, longitude and elevation of each of the stations.
 The goal of this example is to compare the monthly average maximum temperature between two periods, 1971-1975 and 2016-2020, for stations in Victoria and New South Wales, using a glyph map.
@@
 \begin{CodeInput}
-R> tmax <- historical_tmax %>%
+R> tmax <- historical_tmax |>
 +   filter(between(stringr::str_sub(id, 7, 8), 46, 90))
@@
 \begin{CodeInput}
-R> tmax <- tmax %>%
-+   face_temporal() %>% 
+R> tmax <- tmax |>
++   face_temporal() |> 
 +   group_by(month = lubridate::month(date),
@@
 +            ifelse(lubridate::year(date) > 2015,
-+            "2016 ~ 2020", "1971 ~ 1975"))) %>%
++            "2016 ~ 2020", "1971 ~ 1975"))) |>
 +   summarise(tmax = mean(tmax, na.rm = TRUE))
@@
 \begin{CodeInput}
-R> tmax %>% 
-+   face_spatial() %>% 
-+   mutate(n = nrow(ts)) %>%
-+   arrange(n) %>%
-+   pull(n) %>% 
+R> tmax |> 
++   face_spatial() |> 
++   mutate(n = nrow(ts)) |>
++   arrange(n) |>
++   pull(n) |> 
 +   head(10)
@@
 \begin{CodeInput}
-R> tmax <- tmax %>% 
-+   face_spatial() %>% 
-+   filter(nrow(ts) == 24) %>%
+R> tmax <- tmax |> 
++   face_spatial() |> 
++   filter(nrow(ts) == 24) |>
 +   face_temporal()
@@
 \begin{CodeInput}
-R> nsw_vic <- ozmaps::abs_ste %>%
+R> nsw_vic <- ozmaps::abs_ste |>
 +   filter(NAME %in% c("Victoria", "New South Wales"))
@@
 The data is available in the NetCDF format from the European Centre for Medium-Range Weather Forecasts (ECMWF).
-It can be directly downloaded from Copernicus Climate Data Store (CDS) \citep{cds} website or via the R package \pkg{ecmwfr} \citep{ecwmfr}.
+It can be directly downloaded from Copernicus Climate Data Store (CDS) \citep{cds} website or via the \proglang{R} package \pkg{ecmwfr} \citep{ecwmfr}.
 
@@
 
-The downloaded netCDF data is read into R using the \code{ncdf4} package, and converted to a \code{cubble}:
+The downloaded netCDF data is read into \proglang{R} using the \code{ncdf4} package, and converted to a \code{cubble}:
 
@@
 \begin{Highlighting}[]
-\CommentTok{\# data pre{-}processing}
-\NormalTok{clean }\OtherTok{\textless{}{-}}\NormalTok{ climate\_full }\SpecialCharTok{\%\textgreater{}\%}\NormalTok{ ...}
-
-\CommentTok{\# created SharedData instance for crosstalk}
-\NormalTok{nested }\OtherTok{\textless{}{-}}\NormalTok{ clean }\SpecialCharTok{\%\textgreater{}\%}\NormalTok{ SharedData}\SpecialCharTok{$}\FunctionTok{new}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{id, }\AttributeTok{group =} \StringTok{"cubble"}\NormalTok{)}
-\NormalTok{long }\OtherTok{\textless{}{-}} \FunctionTok{face\_temporal}\NormalTok{(clean) }\SpecialCharTok{\%\textgreater{}\%}\NormalTok{ SharedData}\SpecialCharTok{$}\FunctionTok{new}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{id, }\AttributeTok{group =} \StringTok{"cubble"}\NormalTok{)}
-
-\CommentTok{\# create the spatial and temporal view each with a ShareData instance}
-\NormalTok{p1 }\OtherTok{\textless{}{-}}\NormalTok{ nested }\SpecialCharTok{\%\textgreater{}\%}\NormalTok{ ...}
-\NormalTok{p2 }\OtherTok{\textless{}{-}}\NormalTok{ long }\SpecialCharTok{\%\textgreater{}\%}\NormalTok{ ...}
-
-\CommentTok{\# Combine p1 and p2}
+\NormalTok{clean }\OtherTok{\textless{}{-}}\NormalTok{ climate\_full }\SpecialCharTok{|}\ErrorTok{\textgreater{}}\NormalTok{ ...}
+
+\NormalTok{nested }\OtherTok{\textless{}{-}}\NormalTok{ clean }\SpecialCharTok{|}\ErrorTok{\textgreater{}}\NormalTok{ SharedData}\SpecialCharTok{$}\FunctionTok{new}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{id, }\AttributeTok{group =} \StringTok{"cubble"}\NormalTok{)}
+\NormalTok{long }\OtherTok{\textless{}{-}} \FunctionTok{face\_temporal}\NormalTok{(clean) }\SpecialCharTok{|}\ErrorTok{\textgreater{}}\NormalTok{ SharedData}\SpecialCharTok{$}\FunctionTok{new}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{id, }\AttributeTok{group =} \StringTok{"cubble"}\NormalTok{)}
+
+\NormalTok{p1 }\OtherTok{\textless{}{-}}\NormalTok{ nested }\SpecialCharTok{|}\ErrorTok{\textgreater{}}\NormalTok{ ...}
+\NormalTok{p2 }\OtherTok{\textless{}{-}}\NormalTok{ long }\SpecialCharTok{|}\ErrorTok{\textgreater{}}\NormalTok{ ...}
+
 \NormalTok{crosstalk}\SpecialCharTok{::}\FunctionTok{bscols}\NormalTok{(plotly}\SpecialCharTok{::}\FunctionTok{ggplotly}\NormalTok{(p1), plotly}\SpecialCharTok{::}\FunctionTok{ggplotly}\NormalTok{(p2), ...)}
@@
 This work is funded by a Commonwealth Scientific and Industrial Research Organisation (CSIRO) Data61 Scholarship and started while Nicolas Langrené was affiliated with CSIRO's Data61.
-The article is created using \pkg{knitr} \citep{knitr} and \pkg{rmarkdown} \citep{rmarkdown} in R using the \texttt{rticles::jss\_article} template.
+The article is created using \pkg{knitr} \citep{knitr} and \pkg{rmarkdown} \citep{rmarkdown} in R using the \code{rticles::jss_article} template.
 The source code for reproducing this paper can be found at: \url{https://github.com/huizezhang-sherry/paper-cubble}.

### 2022-07-09 1-intro.Rmd
quick fixes on code markup \pkg{}, \code{}

@@
 
-This paper presents a new \proglang{R} package, \pkg{cubble} which addresses the two issues mentioned above.
-In the package, a new data structure, also called \code{cubble}, is proposed to organize spatial and temporal variables as two forms of a single data object so that they can be wrangled separately or combined while being kept synchronized.
-Among the four spacetime layouts in @spacetime, \pkg{cubble} can be applied to full grid, sparse grid, or irregular, but not trajectory, which is outside the scope of this work.
+This paper presents a new \proglang{R} package, \pkg{cubble}, which addresses the two issues mentioned above.
+In the package, a new data structure, also called cubble, is proposed to organize spatial and temporal variables as two forms of a single data object so that they can be wrangled separately or combined while being kept synchronized.
+Among the four spacetime layouts in @spacetime, cubble can be applied to full grid, sparse grid, or irregular, but not trajectory, which is outside the scope of this work.
 The software is available from the Comprehensive \proglang{R} Archive Network (CRAN) at https://CRAN.R-project.org/package=cubble.
@@
 The rest of the paper is organized as follows: Section \@ref(cube) introduces the proposed cube structure as a way to conceptualize multivariate spatio-temporal data.
-Section \@ref(cubble) presents the main design and functionality of \pkg{cubble}.
-Section \@ref(others) explains how cubble deals with more advanced considerations, including data with hierarchical structure, data matching and how cubble fits with existing static and interactive visualization tools.
-Moreover we also illustrate how \pkg{cubble} deals with spatio-temporal data transformations.
-Section \@ref(examples) uses Australian weather station data and river level data as examples to demonstrate the use of \pkg{cubble}.
-An example of how \pkg{cubble} handles NetCDF data is also provided.
+Section \@ref(cubble) presents the main design and functionality of the \pkg{cubble} package.
+Section \@ref(others) explains how the \pkg{cubble} package deals with more advanced considerations, including data with hierarchical structure, data matching and how it fits with existing static and interactive visualization tools.
+Moreover we also illustrate how the \pkg{cubble} package deals with spatio-temporal data transformations.
+Section \@ref(examples) uses Australian weather station data and river level data as examples to demonstrate the use of the \pkg{cubble} package.
+An example of how cubble handles NetCDF data is also provided.
 Section \@ref(conclude) discuss the paper contributions and future directions.

### 2022-07-09 2-cube.Rmd
quick fixes on code markup \pkg{}, \code{}

@@
 Finally, variables are stacked on this space-time canvas, with one observation per site and time point.
-This notion is adopted to avoid using hyper-cubes when describing multivariate spatio-temporal data and is the conceptual framework behind the \pkg{cubble} objects.
+This notion is adopted to avoid using hyper-cubes when describing multivariate spatio-temporal data and is the conceptual framework behind the \code{cubble} objects.
 With this conceptual model, operations on spatio-temporal data can be mapped to operations on the cube and the rest of Figure \ref{fig:cube-def} show examples of slicing on site, time, and variable.
@@
 In addition, it will be generally useful to be able to create new variables which is trickier to manage in an array.
-Thus for convenient wrangling, we have opted to create a special `cubble` class.
+Thus for convenient wrangling, we have opted to create a special \code{cubble} class.
 

### 2022-07-09 3-the-cubble-pkg.Rmd
quick fixes on code markup \pkg{}, \code{}

@@
 The function \code{as_cubble()} is used to create a cubble with three arguments: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order  (longitude, latitude).
-The arguments \code{key} and \code{index} follow the wording in \pkg{tsibble} to describe the temporal order and multiple series while \code{coords} specifies the spatial location of each site.
+The arguments \code{key} and \code{index} follow the wording in the \pkg{tsibble} package to describe the temporal order and multiple series while \code{coords} specifies the spatial location of each site.
 The code below creates a cubble out of \code{climate_flat} (a single tibble) with \code{id} as the key, \code{date} as the index, and \code{c(long, lat)} as the coordinates:
@@
 
-A cubble object is a subclass of the \code{rowwise_df} class where each row forms a group.
+A \code{cubble} object is a subclass of the \code{rowwise_df} class where each row forms a group.
 All the temporal variables are nested in a list column, hence it is also called the nested cubble.
@@
 
-A cubble can be created from various common spatio-temporal data formats, including basic \proglang{R} objects like tibble, tsibble and sf.
-Section \@ref(covid) describes converting multiple tables into a cubble object and Section \@ref(netcdf) illustrates how to convert a netCDF object.
+A cubble can be created from various common spatio-temporal data formats, including basic \proglang{R} objects like \code{tibble}, \code{tsibble}, and \code{sf}.
+Section \@ref(covid) describes converting multiple tables into a \code{cubble} object and Section \@ref(netcdf) illustrates how to convert a netCDF object.
 
@@
 The nested form can be used for those operations where the output is only indexed by the spatial identifier (\code{key}), but becomes inadequate when outputs need both a spatial and a temporal identifier (\code{key} and \code{index}).
-The \pkg{cubble} class also provides a long form, which expands the \code{ts} column and temporarily "hides" the spatial variables.
+The \code{cubble} class also provides a long form, which expands the \code{ts} column and temporarily "hides" the spatial variables.
 The function \code{face_temporal()} is used to switch from the nested cubble into the long one.
@@
 
-The \code{cubble} package leverages tools available in existing packages used for spatial and temporal analysis, specifically, \code{dplyr}, \code{tsibble}, \code{sf} (\code{s2}), and \code{netcdf4}, as explained here.
+The \pkg{cubble} package leverages tools available in existing packages used for spatial and temporal analysis, specifically, \pkg{dplyr}, \pkg{tsibble}, \pkg{sf} (\pkg{s2}), and \pkg{netcdf4}, as explained here.
 
@@
 
-A \code{tsibble} is a \code{tibble} where the \code{index} and \code{key} components are used to store temporal and strata information, that makes working with temporal data cognitively efficient.
-A \code{cubble} can use a \code{tsibble} for storing the temporal information, and effectively utilize the specialist time series operations in the \code{tsibble} package.
-It is easy to cast an existing \code{tsibble} into a \code{cubble} simply by supplying the \code{coords}:
+A \code{tsibble} object is a tibble where the \code{index} and \code{key} components are used to store temporal and strata information, that makes working with temporal data cognitively efficient.
+A \code{cubble} object can use a tsibble for storing the temporal information, and effectively utilize the specialist time series operations in the \pkg{tsibble} package.
+It is easy to cast an existing \code{tsibble} object into a cubble simply by supplying the \code{coords}:
 
@@
 
-When a nested cubble is created, each element in the list-column \code{ts} is a \code{tsibble} class (labelled \code{tbl_ts})  and operations available to the \code{tsibble} class are still valid on this element.
+When a nested cubble is created, each element in the list-column \code{ts} has a \code{tsibble} class (labelled \code{tbl_ts})  and operations available to the \code{tsibble} class are still valid on this element.
 For example, the code below calculates two time series features (mean and variance) of maximum temperature, utilizing the tsibble structure in the cubble:
@@
 
-The \code{sf} is a spatial data object, containing a feature geometry list-column (\code{sfc}) in the form of various geometry types (\code{POINT}, \code{LINESTRING}, \code{POLYGON}, \code{MULTIPOLYGON}).
-The \code{sf} package provides functions that operate efficiently on this spatial information.
+An \code{sf} object is a spatial data object, containing a feature geometry list-column (\code{sfc}) in the form of various geometry types (\code{POINT}, \code{LINESTRING}, \code{POLYGON}, \code{MULTIPOLYGON}).
+The \pkg{sf} package provides functions that operate efficiently on this spatial information.
 A cubble can store the spatial information as an \code{sf} object.
@@
 An illustration is in Section \@ref(covid).
-A spatial data object with an \code{s2} vector can also be used to store the spatial information in a \code{cubble}.
+The spatial information can also be stored as an \code{s2} in a cubble object.
 
@@
 
-Cubble provides an \code{as_cubble()} method to coerce the \code{ncdf4} class from the \pkg{ncdf4} package into a \code{cubble}.
+Cubble provides an \code{as_cubble()} method to coerce the \code{ncdf4} class from the \pkg{ncdf4} package into a \code{cubble} object.
 It maps each combination of longitude and latitude into an \code{id} as the \code{key}:

### 2022-07-09 4-other-features.Rmd
quick fixes on code markup \pkg{}, \code{}

--- 2022-07-09 4-other-features.Rmd
@@
 
-Spatial locations can have grouping structures either inherent to the data (e.g., state within country) or obtained during the analysis (e.g., cluster id).
+Spatial locations can have grouping structures either inherent to the data e.g., state within country or obtained during the analysis e.g., cluster id.
 In this case, it can be useful to summarize variables at various levels of the hierarchy.
@@
 
-\code{cluster_nested <- station_nested \%>\% switch_key(key = cluster)},
+```{r echo = TRUE, eval = FALSE}
+cluster_nested <- station_nested %>% switch_key(key = cluster)
+```
 
@@
 This can be considered to be a matching problem [@stuart2010matching; @mcintosh2018using] to pair similar time series from nearby locations, or even a data fusion exercise that merges data collected from different sources [@cocchi2019data].
-The function \code{match_sites()} in \pkg{cubble} provides a simple algorithm for this task.
+The function \code{match_sites()} in the \pkg{cubble} package provides a simple algorithm for this task.
 The algorithm first matches spatially by computing the pairwise distance on latitude and longitude.
@@
 Figure \ref{fig:illu-matching} illustrates this temporal matching.
-In the two series, \code{A} and \code{a}, three peaks have been identified in each.
-An interval, of fixed length, is constructed for each peak in series \code{A}, while the peaks in series \code{a} are tested against whether they fall into any of the intervals.
+In the two series, A and a, three peaks have been identified in each.
+An interval, of fixed length, is constructed for each peak in series A, while the peaks in series a are tested against whether they fall into any of the intervals.
 Here two out of three peaks match.
@@
 
-The cubble workflow fits works well with an interactive graphics pipeline (e.g., @buja1988elements, @buja1996interactive,  @sutherland2000orca, @xie2014reactive, @cheng2016enabling) that is available in R with the package crosstalk [@crosstalk].
+The cubble workflow fits works well with an interactive graphics pipeline (e.g., @buja1988elements, @buja1996interactive,  @sutherland2000orca, @xie2014reactive, @cheng2016enabling) that is available in \proglang{R} with the package \pkg{crosstalk} [@crosstalk].
 Figure \ref{fig:illu-interactive} illustrates how linking can be achieved between a map and multiple time series using a cubble.
-The map (produced from the nested form) and time series (produced from the long form) are both shared crosstalk objects.
+The map (produced from the nested form) and time series (produced from the long form) are both shared \code{crosstalk} objects.
 When a user makes a selection on the map, the site is highlighted (a).
-This activates a row in the nested cubble object, which is then communicated to the long cubble -- all the observations with the same \code{id} (b) will be selected.
+This activates a row in the nested \code{cubble} object, which is then communicated to the long cubble -- all the observations with the same id (b) will be selected.
 The long cubble will then highlight the corresponding series in the time series plot (c).
@@
 Linking is also available starting from the time series plot, by selecting points.
-This will be activate rows having the same \code{id} in the long cubble.
+This will be activate rows having the same id in the long cubble.
 The corresponding rows in the nested cubble are activated, and highlighted the map.
@@
 ```{r illu-interactive, echo = FALSE, fig.align="center", out.height = "35%", out.width = "100%", fig.cap = "Linking between multiple plots.
-The line plots and the map are constructed from shared crosstalk objects (long and nested cubbles).
+The line plots and the map are constructed from shared \\code{crosstalk} objects (long and nested cubbles).
 When a station is selected on the map (a), the corresponding row in the nested cubble will be activated.
@@
 
-The \proglang{R} package \pkg{GGally} implements glyph maps through the \code{glyphs()} function.
+The package \pkg{GGally} [@ggally] has implemented glyph maps through the \code{glyphs()} function.
 The function constructs a data frame with calculated position (\code{gx}, \code{gy}, \code{gid}) of each point on the time series using linear algebra (Equations 1 and 2 in @Wickham2012-yr).
-The data can then be piped into \code{ggplot} to create the glyph map as:
+The data can then be piped into ggplot to create the glyph map as:
 

### 2022-07-09 5-1-covid.Rmd
check on \proglang, \pkg, \code being used throughout the paper
quick fixes on code markup \pkg{}, \code{}

@@
 
-The five examples here are chosen to illustrate these aspects of using cubble: creating a \code{cubble} from two COVID data tables with the complication of differing location names, using spatial transformations to make a glyph map of seasonal temperature changes over years, aggregating information spatially to explore precipitation patterns, matching river level data and weather station records for analysis of water supply, reading netCDF format data to reproduce a climate reanalysis plot, and the workflow to create complex interactive linked plots.
+The five examples here are chosen to illustrate these aspects of using cubble: creating a \code{cubble} object from two COVID data tables with the complication of differing location names, using spatial transformations to make a glyph map of seasonal temperature changes over years, aggregating information spatially to explore precipitation patterns, matching river level data and weather station records for analysis of water supply, reading netCDF format data to reproduce a climate reanalysis plot, and the workflow to create complex interactive linked plots.
 (There is an additional example and figures in the Appendix, and more examples in the package vignettes.)
@@
 These different sources need to be combined for the analysis, by matching the LGA names.
-Here is how to do this with \pkg{cubble}, including how to handle mismatches arising from different names of the same LGAs in the two tables.
-The COVID data is store in a `csv` file and looks like:
+Here is how to do this with the \pkg{cubble} package, including how to handle mismatches arising from different names of the same LGAs in the two tables.
+The COVID data is store in a csv file and looks like:
 

### 2022-07-09 5-2-aus-hist-max-temp.Rmd
quick fixes on code markup \pkg{}, \code{}

--- 2022-07-09 5-2-aus-hist-max-temp.Rmd
@@
 The Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world.
-The data used here (\code{historical_tmax}) is a subset extracted using the `rnoaa` package, containing the records of maximum temperature for `r nrow(historical_tmax)` Australian stations from `r range(historical_tmax$first_year)[1]` through `r range(historical_tmax$first_year)[2]` and provides information also on the latitude, longitude and elevation of each of the  stations.
+The data used here (\code{historical_tmax}) is a subset extracted using the package \pkg{rnoaa} [@rnoaa], containing the records of maximum temperature for `r nrow(historical_tmax)` Australian stations from `r range(historical_tmax$first_year)[1]` through `r range(historical_tmax$first_year)[2]` and provides information also on the latitude, longitude and elevation of each of the  stations.
 <!-- This data  is already cast into a \code{cubble} data object (please see Appendix X for the code to extract  the data from XX and prepare the data into \code{cubble} form), with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates.--> The goal of this example is to compare the monthly average maximum temperature between two periods, 1971-1975 and 2016-2020, for stations in Victoria and New South Wales, using a glyph map.
@@
 New South Wales stations correspond to entries in the range 46-75 and the Victorian stations to 76-90.
-Filtering Victoria and New South Wales stations is a *spatial operation* and hence uses the \code{cubble} nested form:
+Filtering Victoria and New South Wales stations is a *spatial operation* and hence uses the \code{cubble} object in the nested form:
 
@@
 Next, the monthly maximum average temperature is  calculated for both periods.
-This is a *temporal operation* requiring a switch into the long \code{cubble} form using the \code{face_temporal()} function.
+This is a *temporal operation* requiring a switch into the long cubble form using the \code{face_temporal()} function.
 In addition, a new indicator for the two time periods of interest is created before the calculation of monthly averages:
@@
 To fix this, one could aggregate data from nearby stations.
-An example of this using **cubble** is included in the Appendix.
+An example of this using is included in the Appendix.

### 2022-07-09 5-4-era5.Rmd
quick fixes on code markup \pkg{}, \code{}

--- 2022-07-09 5-4-era5.Rmd
@@
 The data is available in the NetCDF format from the European Centre for Medium-Range Weather Forecasts (ECMWF).
-It can be directly downloaded from Copernicus Climate Data Store (CDS) [@cds] website or via the R package \pkg{ecmwfr} [@ecwmfr].
+It can be directly downloaded from Copernicus Climate Data Store (CDS) [@cds] website or via the package \pkg{ecmwfr} [@ecwmfr].
 
@@
 <!--The goal of this example is to reproduce the break-up of the southern polar vortex that happened in late September and early October in 2002 in the stratosphere in the Southern Hemisphere.
-Therefore, we filter data corresponding to the Southern Hemisphere and restrict our analysis to four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04.--> The downloaded netCDF data is read into R using the \code{ncdf4} package, and converted to a \code{cubble}:
+Therefore, we filter data corresponding to the Southern Hemisphere and restrict our analysis to four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04.--> The downloaded netCDF data is read into \proglang{R} using the \pkg{ncdf4} package, and converted to a \code{cubble} object:
 
@@
 
-Creating the plot requires making transformations on time, unfolding the data for computing the statistic of interest, which is plotted directly as a contour plot with \code{ggplot}.
+Creating the plot requires making transformations on time, unfolding the data for computing the statistic of interest, which is plotted directly as a contour plot with ggplot.
 Code is provided to accomplish this in the supplementary material.

### 2022-07-09 6-conclusion.Rmd
quick fixes on code markup \pkg{}, \code{}

--- 2022-06-27 6-conclusion.Rmd
@@
 This work adds capabilities into the spatio-temporal practitioners toolbox to integrate it with a tidy data framework.
-The data structure and functions introduced in this package can be used and combined with existing spatio-temporal R packages such as \pkg{sf}, data wrangling packages such as \pkg{dplyr}, and visualization packages such as \pkg{ggplot2}, \pkg{plotly} and \pkg{leaflet}.
+The data structure and functions introduced in this package can be used and combined with existing spatial data analysis packages such as \pkg{sf}, data wrangling packages such as \pkg{dplyr}, and visualization packages such as \pkg{ggplot2}, \pkg{plotly} and \pkg{leaflet}.
 
@@
 That could be achieved with a list-column for the location coordinates, and an additional form that these locations can be pivoted into, like the long form for temporal variables.
-For multiple measured variables, \pkg{cubble} could be integrated with dimension reduction methods, and dynamic graphics for multiple dimensions such as a tour [@wickham_tourr_2011].
+For multiple measured variables, the \pkg{cubble} package could be integrated with dimension reduction methods, and dynamic graphics for multiple dimensions such as a tour [@wickham_tourr_2011].
 
@@
 This work is funded by a Commonwealth Scientific and Industrial Research Organisation (CSIRO) Data61 Scholarship and started while Nicolas Langrené was affiliated with CSIRO's Data61.
-The article is created using \pkg{knitr} [@knitr] and \pkg{rmarkdown} [@rmarkdown] in R using the `rticles::jss_article` template.
+The article is created using the package \pkg{knitr} [@knitr] and \pkg{rmarkdown} [@rmarkdown] in \proglang{R} with the \code{rticles::jss_article} template.
 The source code for reproducing this paper can be found at: https://github.com/huizezhang-sherry/paper-cubble.

### 2022-07-09 jss/main.Rmd
quick fixes on code markup \pkg{}, \code{}

@@
 In addition, \code{cubble} facilitates visual and numerical explorations of the data while easing data wrangling and modelling.
-The \code{cubble} structure and the functions provided in the \code{cubble} \proglang{R} package equip users with the capability to handle hierarchical spatial and temporal structures.
+The \code{cubble} structure and the functions provided in the \pkg{cubble} \proglang{R} package equip users with the capability to handle hierarchical spatial and temporal structures.
 The \code{cubble} structure and the tools implemented in the package are illustrated with different examples of Australian climate data.

### 2022-07-09 jss/main.tex
quick fixes on code markup \pkg{}, \code{}

@@
 
-This paper presents a new \proglang{R} package, \pkg{cubble} which addresses the two issues mentioned above.
-In the package, a new data structure, also called \code{cubble}, is proposed to organize spatial and temporal variables as two forms of a single data object so that they can be wrangled separately or combined while being kept synchronized.
-Among the four spacetime layouts in \citet{spacetime}, \pkg{cubble} can be applied to full grid, sparse grid, or irregular, but not trajectory, which is outside the scope of this work.
+This paper presents a new \proglang{R} package, \pkg{cubble}, which addresses the two issues mentioned above.
+In the package, a new data structure, also called cubble, is proposed to organize spatial and temporal variables as two forms of a single data object so that they can be wrangled separately or combined while being kept synchronized.
+Among the four spacetime layouts in \citet{spacetime}, cubble can be applied to full grid, sparse grid, or irregular, but not trajectory, which is outside the scope of this work.
 The software is available from the Comprehensive \proglang{R} Archive Network (CRAN) at \url{https://CRAN.R-project.org/package=cubble}.
@@
 The rest of the paper is organized as follows: Section \ref{cube} introduces the proposed cube structure as a way to conceptualize multivariate spatio-temporal data.
-Section \ref{cubble} presents the main design and functionality of \pkg{cubble}.
-Section \ref{others} explains how cubble deals with more advanced considerations, including data with hierarchical structure, data matching and how cubble fits with existing static and interactive visualization tools.
-Moreover we also illustrate how \pkg{cubble} deals with spatio-temporal data transformations.
-Section \ref{examples} uses Australian weather station data and river level data as examples to demonstrate the use of \pkg{cubble}.
-An example of how \pkg{cubble} handles NetCDF data is also provided.
+Section \ref{cubble} presents the main design and functionality of the \pkg{cubble} package.
+Section \ref{others} explains how the \pkg{cubble} package deals with more advanced considerations, including data with hierarchical structure, data matching and how it fits with existing static and interactive visualization tools.
+Moreover we also illustrate how the \pkg{cubble} package deals with spatio-temporal data transformations.
+Section \ref{examples} uses Australian weather station data and river level data as examples to demonstrate the use of the \pkg{cubble} package.
+An example of how cubble handles NetCDF data is also provided.
 Section \ref{conclude} discuss the paper contributions and future directions.
@@
 In addition, it will be generally useful to be able to create new variables which is trickier to manage in an array.
-Thus for convenient wrangling, we have opted to create a special \texttt{cubble} class.
+Thus for convenient wrangling, we have opted to create a special \code{cubble} class.
 
@@
 The function \code{as_cubble()} is used to create a cubble with three arguments: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order (longitude, latitude).
-The arguments \code{key} and \code{index} follow the wording in \pkg{tsibble} to describe the temporal order and multiple series while \code{coords} specifies the spatial location of each site.
+The arguments \code{key} and \code{index} follow the wording in the \pkg{tsibble} package to describe the temporal order and multiple series while \code{coords} specifies the spatial location of each site.
 The code below creates a cubble out of \code{climate_flat} (a single tibble) with \code{id} as the key, \code{date} as the index, and \code{c(long, lat)} as the coordinates:
@@
 
-A cubble can be created from various common spatio-temporal data formats, including basic \proglang{R} objects like tibble, tsibble and sf.
-Section \ref{covid} describes converting multiple tables into a cubble object and Section \ref{netcdf} illustrates how to convert a netCDF object.
+A cubble can be created from various common spatio-temporal data formats, including basic \proglang{R} objects like \code{tibble}, \code{tsibble}, and \code{sf}.
+Section \ref{covid} describes converting multiple tables into a \code{cubble} object and Section \ref{netcdf} illustrates how to convert a netCDF object.
 
@@
 The nested form can be used for those operations where the output is only indexed by the spatial identifier (\code{key}), but becomes inadequate when outputs need both a spatial and a temporal identifier (\code{key} and \code{index}).
-The \pkg{cubble} class also provides a long form, which expands the \code{ts} column and temporarily ``hides'' the spatial variables.
+The \code{cubble} class also provides a long form, which expands the \code{ts} column and temporarily ``hides'' the spatial variables.
 The function \code{face_temporal()} is used to switch from the nested cubble into the long one.
@@
 
-When a nested cubble is created, each element in the list-column \code{ts} is a \code{tsibble} class (labelled \code{tbl_ts}) and operations available to the \code{tsibble} class are still valid on this element.
+When a nested cubble is created, each element in the list-column \code{ts} has a \code{tsibble} class (labelled \code{tbl_ts}) and operations available to the \code{tsibble} class are still valid on this element.
 For example, the code below calculates two time series features (mean and variance) of maximum temperature, utilizing the tsibble structure in the cubble:
@@
 An illustration is in Section \ref{covid}.
-A spatial data object with an \code{s2} vector can also be used to store the spatial information in a \code{cubble}.
+The spatial information can also be stored as an \code{s2} in a cubble object.
 
@@
 
-\code{cluster_nested <- station_nested \%>\% switch_key(key = cluster)},
+\begin{CodeChunk}
+\begin{CodeInput}
+R> cluster_nested <- station_nested %>% switch_key(key = cluster)
+\end{CodeInput}
+\end{CodeChunk}
 
@@
 This can be considered to be a matching problem \citep{stuart2010matching, mcintosh2018using} to pair similar time series from nearby locations, or even a data fusion exercise that merges data collected from different sources \citep{cocchi2019data}.
-The function \code{match_sites()} in \pkg{cubble} provides a simple algorithm for this task.
+The function \code{match_sites()} in the \pkg{cubble} package provides a simple algorithm for this task.
 The algorithm first matches spatially by computing the pairwise distance on latitude and longitude.
@@
 
-The cubble workflow fits works well with an interactive graphics pipeline (e.g., \citet{buja1988elements}, \citet{buja1996interactive}, \citet{sutherland2000orca}, \citet{xie2014reactive}, \citet{cheng2016enabling}) that is available in \proglang{R} with the package crosstalk \citep{crosstalk}.
+The cubble workflow fits works well with an interactive graphics pipeline (e.g., \citet{buja1988elements}, \citet{buja1996interactive}, \citet{sutherland2000orca}, \citet{xie2014reactive}, \citet{cheng2016enabling}) that is available in \proglang{R} with the package \pkg{crosstalk} \citep{crosstalk}.
 Figure \ref{fig:illu-interactive} illustrates how linking can be achieved between a map and multiple time series using a cubble.
-The map (produced from the nested form) and time series (produced from the long form) are both shared crosstalk objects.
+The map (produced from the nested form) and time series (produced from the long form) are both shared \code{crosstalk} objects.
 When a user makes a selection on the map, the site is highlighted (a).
-This activates a row in the nested cubble object, which is then communicated to the long cubble -- all the observations with the same \code{id} (b) will be selected.
+This activates a row in the nested \code{cubble} object, which is then communicated to the long cubble -- all the observations with the same id (b) will be selected.
 The long cubble will then highlight the corresponding series in the time series plot (c).
@@
 \caption[Linking between multiple plots]{Linking between multiple plots.
-The line plots and the map are constructed from shared crosstalk objects (long and nested cubbles).
+The line plots and the map are constructed from shared \code{crosstalk} objects (long and nested cubbles).
 When a station is selected on the map (a), the corresponding row in the nested cubble will be activated.
@@
 
-The \proglang{R} package \pkg{GGally} implements glyph maps through the \code{glyphs()} function.
+The package \pkg{GGally} \citep{ggally} has implemented glyph maps through the \code{glyphs()} function.
 The function constructs a data frame with calculated position (\code{gx}, \code{gy}, \code{gid}) of each point on the time series using linear algebra (Equations 1 and 2 in \citet{Wickham2012-yr}).
-The data can then be piped into \code{ggplot} to create the glyph map as:
+The data can then be piped into ggplot to create the glyph map as:
 
@@
 These different sources need to be combined for the analysis, by matching the LGA names.
-Here is how to do this with \pkg{cubble}, including how to handle mismatches arising from different names of the same LGAs in the two tables.
+Here is how to do this with the \pkg{cubble} package, including how to handle mismatches arising from different names of the same LGAs in the two tables.
 The COVID data is store in a csv file and looks like:
@@
 The Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world.
-The data used here (\code{historical_tmax}) is a subset extracted using the \pkg{rnoaa}\citep{rnoaa} package, containing the records of maximum temperature for 236 Australian stations from 1859 through 1969 and provides information also on the latitude, longitude and elevation of each of the stations.
+The data used here (\code{historical_tmax}) is a subset extracted using the package \pkg{rnoaa} \citep{rnoaa}, containing the records of maximum temperature for 236 Australian stations from 1859 through 1969 and provides information also on the latitude, longitude and elevation of each of the stations.
 The goal of this example is to compare the monthly average maximum temperature between two periods, 1971-1975 and 2016-2020, for stations in Victoria and New South Wales, using a glyph map.
@@
 New South Wales stations correspond to entries in the range 46-75 and the Victorian stations to 76-90.
-Filtering Victoria and New South Wales stations is a \emph{spatial operation} and hence uses the \code{cubble} nested form:
+Filtering Victoria and New South Wales stations is a \emph{spatial operation} and hence uses the \code{cubble} object in the nested form:
 
@@
 Next, the monthly maximum average temperature is calculated for both periods.
-This is a \emph{temporal operation} requiring a switch into the long \code{cubble} form using the \code{face_temporal()} function.
+This is a \emph{temporal operation} requiring a switch into the long cubble form using the \code{face_temporal()} function.
 In addition, a new indicator for the two time periods of interest is created before the calculation of monthly averages:
@@
 To fix this, one could aggregate data from nearby stations.
-An example of this using \textbf{cubble} is included in the Appendix.
+An example of this using is included in the Appendix.
 
@@
 The data is available in the NetCDF format from the European Centre for Medium-Range Weather Forecasts (ECMWF).
-It can be directly downloaded from Copernicus Climate Data Store (CDS) \citep{cds} website or via the \proglang{R} package \pkg{ecmwfr} \citep{ecwmfr}.
+It can be directly downloaded from Copernicus Climate Data Store (CDS) \citep{cds} website or via the package \pkg{ecmwfr} \citep{ecwmfr}.
 
@@
 
-The downloaded netCDF data is read into \proglang{R} using the \code{ncdf4} package, and converted to a \code{cubble}:
+The downloaded netCDF data is read into \proglang{R} using the \pkg{ncdf4} package, and converted to a \code{cubble} object:
 
@@
 
-The key steps are to convert both the nested and long forms of the data into shared objects, and to plot these side-by-side. The two are linked by the station identifier.
+The key steps are to convert both the nested and long forms of the data into shared \code{crosstalk} objects, and to plot these side-by-side. The two are linked by the station identifier.
 
@@
 This work adds capabilities into the spatio-temporal practitioners toolbox to integrate it with a tidy data framework.
-The data structure and functions introduced in this package can be used and combined with existing spatio-temporal \proglang{R} packages such as \pkg{sf}, data wrangling packages such as \pkg{dplyr}, and visualization packages such as \pkg{ggplot2}, \pkg{plotly} and \pkg{leaflet}.
+The data structure and functions introduced in this package can be used and combined with existing spatial data analysis packages such as \pkg{sf}, data wrangling packages such as \pkg{dplyr}, and visualization packages such as \pkg{ggplot2}, \pkg{plotly} and \pkg{leaflet}.
 
@@
 That could be achieved with a list-column for the location coordinates, and an additional form that these locations can be pivoted into, like the long form for temporal variables.
-For multiple measured variables, \pkg{cubble} could be integrated with dimension reduction methods, and dynamic graphics for multiple dimensions such as a tour \citep{wickham_tourr_2011}.
+For multiple measured variables, the \pkg{cubble} package could be integrated with dimension reduction methods, and dynamic graphics for multiple dimensions such as a tour \citep{wickham_tourr_2011}.
 
@@
 This work is funded by a Commonwealth Scientific and Industrial Research Organisation (CSIRO) Data61 Scholarship and started while Nicolas Langrené was affiliated with CSIRO's Data61.
-The article is created using \pkg{knitr} \citep{knitr} and \pkg{rmarkdown} \citep{rmarkdown} in R using the \code{rticles::jss_article} template.
+The article is created using the package \pkg{knitr} \citep{knitr} and \pkg{rmarkdown} \citep{rmarkdown} in \proglang{R} with the \code{rticles::jss_article} template.
 The source code for reproducing this paper can be found at: \url{https://github.com/huizezhang-sherry/paper-cubble}.

### 2022-07-09 1-intro.Rmd
fix abbreviation + more fix on markups

@@
 
-The SpatioTemporal CRAN task view [@ctvspatiotemporal] gathers information about \proglang{R} packages designed for spatio-temporal data and it has a section on *Representing data* that lists existing spatio-temporal data representations used in \proglang{R}.
+The Comprehensive \proglang{R} Archive Network (CRAN) task view SpatioTemporal [@ctvspatiotemporal] gathers information about \proglang{R} packages designed for spatio-temporal data and it has a section on *Representing data* that lists existing spatio-temporal data representations used in \proglang{R}.
 Among them, @spacetime summarises spatio-temporal data into three forms: time-wide, space-wide, and long formats.
 The associated package \pkg{spacetime} [@spacetime] implements four spatio-temporal layouts (full grid, sparse grid, irregular, and trajectory) to handle different space and time combinations.
-The \pkg{stars} [@stars] package has a new implementation to use dense arrays to represent spatio-temporal cubes.
-It also interfaces with \pkg{sf} [@sf], a package commonly used for wrangling spatial data, and the \pkg{tidyverse} [@tidyverse] suite for general data wrangling and visualization in \proglang{R}.
+The package \pkg{stars} [@stars] has a new implementation to use dense arrays to represent spatio-temporal cubes.
+It also interfaces with the package \pkg{sf} [@sf], commonly used for wrangling spatial data, and the \pkg{tidyverse} [@tidyverse] suite for general data wrangling and visualization in \proglang{R}.
 
@@
 According to the tidy data principles, data should be structured into 1) one row per observation, 2) one column per variable, and 3) one type of data per table.
-The long form data is preferred over wide data form given the downstream software such as \pkg{dplyr} [@dplyr] and \pkg{ggplot2} [@ggplot2] for data wrangling and visualization.
+The long form data is preferred over wide data form given the downstream packages such as \pkg{dplyr} [@dplyr] and \pkg{ggplot2} [@ggplot2] for data wrangling and visualization.
 However, the long form can be inefficient to store feature geometries, especially for large multipolygons for hourly, daily or sub-daily periods over years, which are extensively collected and handled, for example in time series analysis.
@@
 Section \@ref(cubble) presents the main design and functionality of the \pkg{cubble} package.
-Section \@ref(others) explains how the \pkg{cubble} package deals with more advanced considerations, including data with hierarchical structure, data matching and how it fits with existing static and interactive visualization tools.
+Section \@ref(others) explains how the \pkg{cubble} package deals with more advanced considerations, including data with hierarchical structure, data matching and how the package fits with existing static and interactive visualization tools.
 Moreover we also illustrate how the \pkg{cubble} package deals with spatio-temporal data transformations.
-Section \@ref(examples) uses Australian weather station data and river level data as examples to demonstrate the use of the \pkg{cubble} package.
-An example of how cubble handles NetCDF data is also provided.
+Section \@ref(examples) uses Australian weather station data and river level data as examples to demonstrate the use of the package.
+An example of how the \pkg{cubble} package handles Network Common Data Form (NetCDF) data is also provided.
 Section \@ref(conclude) discuss the paper contributions and future directions.

### 2022-07-09 3-the-cubble-pkg.Rmd
fix abbreviation + more fix on markups

@@
 The main functions in the package are \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and \code{unfold()}.
-The following sections explain their roles, why the new cubble structure is needed and how the package relates to existing packages for spatial and temporal data analysis.
+The following sections explain their roles, why the new \code{cubble} structure is needed and how the package relates to existing packages for spatial and temporal data analysis.
 
-The data \code{climate_flat} is used to illustrate \pkg{cubble} functionality.
-This is a subset from National Oceanic and Atmospheric Administration (NOAA) Global Historical Climatology Network Daily (GHCND) data.
-It contains spatial variables, station ID, latitude, longitude, elevation, station name, World Meteorology Organisation ID, in additional to daily temporal information,  maximum and minimum temperature values and precipitation records for year 2020.
-The first five rows of \code{climate_flat} are shown below:
+The data \code{climate_flat} is used to illustrate functionality in the \pkg{cubble} package.
+This is a subset from National Oceanic and Atmospheric Administration (NOAA) [@noaa] Global Historical Climatology Network (GHCN) Daily data.
+It contains spatial variables, station id, latitude, longitude, elevation, station name, world meteorology organisation id, in additional to daily temporal information,  maximum and minimum temperature values and precipitation records for year 2020.
+The first five rows of the data are shown below:
 
@@
 
-The function \code{as_cubble()} is used to create a cubble with three arguments: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order  (longitude, latitude).
+The function \code{as_cubble()} is used to create a \code{cubble} object with three arguments: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order (longitude, latitude).
 The arguments \code{key} and \code{index} follow the wording in the \pkg{tsibble} package to describe the temporal order and multiple series while \code{coords} specifies the spatial location of each site.
-The code below creates a cubble out of \code{climate_flat} (a single tibble) with \code{id} as the key, \code{date} as the index, and \code{c(long, lat)} as the coordinates:
+The code below creates a \code{cubble} object out of \code{climate_flat} (a single tibble) with \code{id} as the key, \code{date} as the index, and \code{c(long, lat)} as the coordinates:
 
@@
 
-Printing a cubble provides some information about the data.
+Printing a \code{cubble} object provides some information about the data.
 Here \code{id} is the variable name to identify each location and there are five unique locations.
@@
 A \code{cubble} object is a subclass of the \code{rowwise_df} class where each row forms a group.
-All the temporal variables are nested in a list column, hence it is also called the nested cubble.
+All the temporal variables are nested in a list column, hence it is also called the nested \code{cubble}.
 The rowwise structure makes it simpler to operate on the list using the \code{mutate()} syntax, which is simpler than the \code{purr::map()} when working with a list column.
@@
 
-A cubble can be created from various common spatio-temporal data formats, including basic \proglang{R} objects like \code{tibble}, \code{tsibble}, and \code{sf}.
-Section \@ref(covid) describes converting multiple tables into a \code{cubble} object and Section \@ref(netcdf) illustrates how to convert a netCDF object.
+A \code{cubble} object can be created from various common spatio-temporal data formats, including basic \proglang{R} objects like \code{tibble}, \code{tsibble}, and \code{sf}.
+Section \@ref(covid) describes converting multiple tables into a \code{cubble} object and Section \@ref(netcdf) illustrates how to convert a NetCDF object.
 
@@
 The \code{cubble} class also provides a long form, which expands the \code{ts} column and temporarily "hides" the spatial variables.
-The function \code{face_temporal()} is used to switch from the nested cubble into the long one.
+The function \code{face_temporal()} is used to switch from the nested \code{cubble} into the long one.
 The first row in Figure \ref{fig:face} illustrates this operation where the focus of the cube now changes from the site-variable face to the time-variable face.
-This code switches the cubble just created into its long form:
+This code switches the \code{cubble} object just created into its long form:
 
@@
 
-The first line in the printed cubble now shows it in the long form and the third line has been changed to display the name and type of spatial variables: \code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]}.
-Unlike the nested form, the long cubble is built from a \code{grouped_df} class where all the observations from the same site form a group.
+The first line in the header now shows it in the long form and the third line has been changed to display the name and type of spatial variables: \code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]}.
+Unlike the nested form, the long \code{cubble} is built from a \code{grouped_df} class where all the observations from the same site form a group.
 
-```{r face, echo = FALSE, fig.align="center", out.height = "35%", out.width = "100%", fig.cap = "An illustration of function \\code{face\\_temporal} and \\code{face\\_spatial} in cubble.
-In the first row, \\code{face\\_temporal} switches a cubble from the nested form into the long form and the focus has switched from the spatial aspect (the side face) to the temporal aspect (the front face).
-In the second row, \\code{face\\_spatial} switches a cubble back to the nested form from the long form and shifts focus back to the spatial aspect."}
+```{r face, echo = FALSE, fig.align="center", out.height = "35%", out.width = "100%", fig.cap = "An illustration of function \\code{face\\_temporal} and \\code{face\\_spatial}.
+In the first row, \\code{face\\_temporal} switches a \\code{cubble} object from the nested form into the long form and the focus has switched from the spatial aspect (the side face) to the temporal aspect (the front face).
+In the second row, \\code{face\\_spatial} switches a \\code{cubble} object back to the nested form from the long form and shifts focus back to the spatial aspect."}
 knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.002.png"))
@@
 
-## Unfold spatial variables into the long cubble {#unfold}
+## Unfold spatial variables into the long form {#unfold}
 
@@
 Some readers may question why a new data structure is needed rather than directly creating a list-column  on the combined data using \code{dplyr::nest_by()}.
-The reason is that cubble is specifically designed to utilize the spatio-temporal structure when arranging observations in a single object.
+The reason is that the \code{cubble} object is specifically designed to utilize the spatio-temporal structure when arranging observations in a single object.
 Moreover, it enables easy pivoting between purely spatial, purely temporal, or unfolded into a combined form.
@@
 
-The \code{dplyr} package has many tools for wrangling tidy data,  many of which are useful in the spatio-temporal analysis.
-The \code{cubble} package provides \code{methods} that support the use of the following \code{dplyr} operations on both the nested and long forms: `mutate`, `filter`, `summarise`, `select`, `arrange`, `rename`, `left_join`, and the slice family (`slice_*`).
+The \pkg{dplyr} package has many tools for wrangling tidy data,  many of which are useful in the spatio-temporal analysis.
+The \pkg{cubble} package provides methods that support the use of the following operations in the \pkg{dplyr} package on both the nested and long forms: `mutate`, `filter`, `summarise`, `select`, `arrange`, `rename`, `left_join`, and the slice family (`slice_*`).
 
@@
 
-A \code{tsibble} object is a tibble where the \code{index} and \code{key} components are used to store temporal and strata information, that makes working with temporal data cognitively efficient.
-A \code{cubble} object can use a tsibble for storing the temporal information, and effectively utilize the specialist time series operations in the \pkg{tsibble} package.
-It is easy to cast an existing \code{tsibble} object into a cubble simply by supplying the \code{coords}:
+A \code{tsibble} object is a \code{tibble} where the \code{index} and \code{key} components are used to store temporal and strata information, that makes working with temporal data cognitively efficient.
+A \code{cubble} object can use a \code{tsibble} to store the temporal information, and effectively utilize the specialist time series operations in the \pkg{tsibble} package.
+It is easy to cast an existing \code{tsibble} object into a \code{cubble} simply by supplying the argument \code{coords}:
 
@@
 
-When a nested cubble is created, each element in the list-column \code{ts} has a \code{tsibble} class (labelled \code{tbl_ts})  and operations available to the \code{tsibble} class are still valid on this element.
-For example, the code below calculates two time series features (mean and variance) of maximum temperature, utilizing the tsibble structure in the cubble:
+When a nested \code{cubble} is created, each element in the list-column \code{ts} has a \code{tsibble} class (labelled \code{tbl_ts})  and operations available to the \code{tsibble} class are still valid on this element.
+For example, the code below calculates two time series features (mean and variance) of maximum temperature, utilizing the \code{tsibble} structure in the \code{cubble} object:
 
@@
 An \code{sf} object is a spatial data object, containing a feature geometry list-column (\code{sfc}) in the form of various geometry types (\code{POINT}, \code{LINESTRING}, \code{POLYGON}, \code{MULTIPOLYGON}).
-The \pkg{sf} package provides functions that operate efficiently on this spatial information.
-A cubble can store the spatial information as an \code{sf} object.
+The package \pkg{sf} provides functions that operate efficiently on this spatial information.
+A \code{cubble} object can store the spatial information as an \code{sf} object.
 Methods for the \code{sfc} class can be applied in the nested form of the cubble object.
 An illustration is in Section \@ref(covid).
-The spatial information can also be stored as an \code{s2} in a cubble object.
+The spatial information can also be stored as an \code{s2} vector in a \code{cubble} object.
 
@@
 
-NetCDF data is another format commonly used for storing spatio-temporal data.
+The NetCDF data is another format commonly used for storing spatio-temporal data.
 It has two main components: *dimension* for defining the spatio-temporal grid (longitude, latitude, and time) and *variable* that populates the defined grid.
@@
 
-Cubble provides an \code{as_cubble()} method to coerce the \code{ncdf4} class from the \pkg{ncdf4} package into a \code{cubble} object.
+The \pkg{cubble} package provides an \code{as_cubble()} method to coerce the \code{ncdf4} class from the \pkg{ncdf4} package into a \code{cubble} object.
 It maps each combination of longitude and latitude into an \code{id} as the \code{key}:
@@
 
-NetCDF data can be quite large, and it is sometimes best to subset the data when converting to a cubble.
+Sometimes NetCDF data can be quite large, and it is best to subset the data when converting to a \code{cubble}.
 We would recommend reducing to about 300 $\times$ 300 grid points for three daily variables in one year.
 A 300 by 300 spatial grid can be a bounding box of [100, -80, 180, 0] at 0.25 degree resolution or a global bounding box [-180, -90, 180, -90] at 1 degree resolution.
-The size of spatial grid can be reduced if longer time periods or more variables are needed, through \code{long_range} and \code{lat_range}:
+The size of spatial grid can be reduced if longer time periods or more variables are needed, through the arguments \code{long_range} and \code{lat_range}:
 
@@
   my_ncdf, vars = c("q", "z"),
-  long_range = seq(-180, 180, 1),lat_range = seq(-90, 90, 1)
+  long_range = seq(-180, 180, 1), lat_range = seq(-90, 90, 1)
   )

### 2022-07-09 4-other-features.Rmd
fix abbreviation + more fix on markups

@@
 The result can also be stretched into long form.
-By specifying
-
-```{r echo = TRUE, eval = FALSE}
-cluster_nested <- station_nested %>% switch_key(key = cluster)
-```
-
-the cubble redefines the cubble key from the \code{id} column in \code{station_nested} to the \code{cluster} column in \code{cluster_nested}.
+By specifying \code{cluster_nested <- station_nested \%>\% switch_key(key = cluster)}, the \code{cubble} objects redefines its \code{key} from the \code{id} column in \code{station_nested} to the \code{cluster} column in \code{cluster_nested}.
 All the spatial variables belonging to the \code{cluster} column are now nested into a \code{.val} column which allows for summarizing based on cluster.
@@
 Figure \ref{fig:illu-matching} illustrates this temporal matching.
-In the two series, A and a, three peaks have been identified in each.
-An interval, of fixed length, is constructed for each peak in series A, while the peaks in series a are tested against whether they fall into any of the intervals.
+In the two series, *A* and *a*, three peaks have been identified in each.
+An interval, of fixed length, is constructed for each peak in series *A*, while the peaks in series *a* are tested against whether they fall into any of the intervals.
 Here two out of three peaks match.
@@
 The cubble workflow fits works well with an interactive graphics pipeline (e.g., @buja1988elements, @buja1996interactive,  @sutherland2000orca, @xie2014reactive, @cheng2016enabling) that is available in \proglang{R} with the package \pkg{crosstalk} [@crosstalk].
-Figure \ref{fig:illu-interactive} illustrates how linking can be achieved between a map and multiple time series using a cubble.
+Figure \ref{fig:illu-interactive} illustrates how linking can be achieved between a map and multiple time series in a \code{cubble} object.
 The map (produced from the nested form) and time series (produced from the long form) are both shared \code{crosstalk} objects.
 When a user makes a selection on the map, the site is highlighted (a).
-This activates a row in the nested \code{cubble} object, which is then communicated to the long cubble -- all the observations with the same id (b) will be selected.
-The long cubble will then highlight the corresponding series in the time series plot (c).
+This activates a row in the nested \code{cubble} object, which is then communicated to the long \code{cubble} -- all the observations with the same id (b) will be selected.
+The long \code{cubble} will then highlight the corresponding series in the time series plot (c).
 
 Linking is also available starting from the time series plot, by selecting points.
-This will be activate rows having the same id in the long cubble.
-The corresponding rows in the nested cubble are activated, and highlighted the map.
+This will be activate rows having the same id in the long \code{cubble} The corresponding rows in the nested \code{cubble} are activated, and highlighted the map.
 (An illustration can be found in the appendix.) Note that this type of linking, both from the map or the time series, is what @CS07 would call categorical variable linking, where station id is the categorical variable.
@@
 ```{r illu-interactive, echo = FALSE, fig.align="center", out.height = "35%", out.width = "100%", fig.cap = "Linking between multiple plots.
-The line plots and the map are constructed from shared \\code{crosstalk} objects (long and nested cubbles).
-When a station is selected on the map (a), the corresponding row in the nested cubble will be activated.
-This will link to all the rows with the same id in the long cubble (b) and update the line plot (c)."}
+The line plots and the map are constructed from shared \\code{crosstalk} objects (long and nested \\code{cubble}).
+When a station is selected on the map (a), the corresponding row in the nested \\code{cubbles} will be activated.
+This will link to all the rows with the same id in the long \\code{cubbles} (b) and update the line plot (c)."}
 knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.004.png"))
@@
 The package \pkg{GGally} [@ggally] has implemented glyph maps through the \code{glyphs()} function.
-The function constructs a data frame with calculated position (\code{gx}, \code{gy}, \code{gid}) of each point on the time series using linear algebra (Equations 1 and 2 in @Wickham2012-yr).
+The function constructs a \code{data.frame} with calculated position (\code{gx}, \code{gy}, \code{gid}) of each point on the time series using linear algebra (Equations 1 and 2 in @Wickham2012-yr).
 The data can then be piped into ggplot to create the glyph map as:
@@
 Some useful controls over the glyph map are also available in the \code{geom_glyph()} implementation.
-Polar coordinate glyph maps are specified using \code{polar = TRUE}, and parameters \code{width} and \code{height} can be specified in either absolute or relative value.
+Polar coordinate glyph maps are specified using \code{polar = TRUE}, and arguments \code{width} and \code{height} can be specified in either absolute or relative value.
 Global and local scale is specified with \code{global_rescale}, which defaults to \code{TRUE}.

### 2022-07-09 5-1-covid.Rmd
fix abbreviation + more fix on markups

@@
 
-The five examples here are chosen to illustrate these aspects of using cubble: creating a \code{cubble} object from two COVID data tables with the complication of differing location names, using spatial transformations to make a glyph map of seasonal temperature changes over years, aggregating information spatially to explore precipitation patterns, matching river level data and weather station records for analysis of water supply, reading netCDF format data to reproduce a climate reanalysis plot, and the workflow to create complex interactive linked plots.
+The five examples here are chosen to illustrate these aspects of the \pkg{cubble} package: creating a \code{cubble} object from two coronavirus (COVID) data tables with the complication of differing location names, using spatial transformations to make a glyph map of seasonal temperature changes over years, aggregating information spatially to explore precipitation patterns, matching river level data and weather station records for analysis of water supply, reading NetCDF format data to reproduce a climate reanalysis plot, and the workflow to create complex interactive linked plots.
 (There is an additional example and figures in the Appendix, and more examples in the package vignettes.)

### 2022-07-09 5-2-aus-hist-max-temp.Rmd
fix abbreviation + more fix on markups

@@
 
-The Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world.
+The GHCN provides daily climate measures from stations across the world.
 The data used here (\code{historical_tmax}) is a subset extracted using the package \pkg{rnoaa} [@rnoaa], containing the records of maximum temperature for `r nrow(historical_tmax)` Australian stations from `r range(historical_tmax$first_year)[1]` through `r range(historical_tmax$first_year)[2]` and provides information also on the latitude, longitude and elevation of each of the  stations.
-<!-- This data  is already cast into a \code{cubble} data object (please see Appendix X for the code to extract  the data from XX and prepare the data into \code{cubble} form), with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates.--> The goal of this example is to compare the monthly average maximum temperature between two periods, 1971-1975 and 2016-2020, for stations in Victoria and New South Wales, using a glyph map.
+<!-- This data  is already cast into a \code{cubble} data object (please see Appendix X for the code to extract  the data from XX and prepare the data into \code{cubble} form), with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates.--> The goal of this example is to compare the monthly average maximum temperature between two periods, 1971-1975 and 2016-2020, for stations in Victoria and New South Wales (NSW), using a glyph map.
 
-First, the stations need to be filtered to those in NSW and Victoria by using the station identifiers, stored within the 11 digits of the \code{id} variable entries.
+First, the stations need to be filtered to those in Victoria and NSW by using the station identifiers, stored within the 11 digits of the \code{id} variable entries.
 The country code is in the first 5 digits (Australia is represented by "ASN00") and the next 6 digits encode the station following the Australian Bureau of Meteorology (BOM) [@bom] coding protocols.
-New South Wales stations correspond to entries in the range 46-75 and the Victorian stations to 76-90.
+The NSW stations correspond to entries in the range 46-75 and the Victorian stations to 76-90.
 Filtering Victoria and New South Wales stations is a *spatial operation* and hence uses the \code{cubble} object in the nested form:
@@
 Next, the monthly maximum average temperature is  calculated for both periods.
-This is a *temporal operation* requiring a switch into the long cubble form using the \code{face_temporal()} function.
+This is a *temporal operation* requiring a switch into the long \code{cubble} form using the \code{face_temporal()} function.
 In addition, a new indicator for the two time periods of interest is created before the calculation of monthly averages:
@@
 
-```{r basic-manip, out.width="100%", fig.height = 5, fig.width = 10, fig.cap="A glyph map of the monthly maximum average temperature for weather stations in  Victoria and New South Wales for the periods (1971-1975, 2016-2020).
-The corresponding average time series for the Cobar station are display on the top left corner.
-From the glyph map we can observe that the monthly trend is similar for all locations (low in the winter, high in the summer), and small increased temperatures, particularly in late summer can be seen at most stations in New South Wales."}
+```{r basic-manip, out.width="100%", fig.height = 5, fig.width = 10, fig.cap="A glyph map of the monthly maximum average temperature for weather stations in  Victoria and New South Wales (NSW) for the periods (1971-1975, 2016-2020).
+The corresponding average time series for the cobar station are display on the top left corner.
+From the glyph map we can observe that the monthly trend is similar for all locations (low in the winter, high in the summer), and small increased temperatures, particularly in late summer can be seen at most stations in NSW."}
 tmax <- tmax |> unfold(long, lat)

### 2022-07-09 5-3-river-level-rainfall.Rmd
fix abbreviation + more fix on markups

@@
 River level and rainfall data for the same areas should have some similarity.
-Here we examine the river gauge data (\code{Water_course_level}) from the Bureau of Meteorology [@bom] in relation to weather station rainfall from NOAA's climate data [@noaa] (\code{climate}).
+Here we examine the river gauge data (\code{Water_course_level}) from the Bureau of Meteorology [@bom] in relation to weather station rainfall from NOAA's climate data (\code{climate}).
 The goal is to match water gauges with nearby weather stations, spatially and temporally, using 2020 measurements using \code{match_sites()} function.

### 2022-07-09 5-4-era5.Rmd
fix abbreviation + more fix on markups

@@
 
-Figure \ref{fig:netcdf} reproduces the ERA5 data [@hersbach2020era5] row of Figure 19 in [@hersbach2020era5].
-Here we explain how this would be done using cubble.
+Figure \ref{fig:netcdf} reproduces the ERA5 data row of Figure 19 in [@hersbach2020era5].
+Here we explain how this would be done using in the \pkg{cubble} package.
 The plots show that the southern polar vortex splits into two on 2002-09-26 and further splits into four on 2002-10-04.
@@
 The data is available in the NetCDF format from the European Centre for Medium-Range Weather Forecasts (ECMWF).
-It can be directly downloaded from Copernicus Climate Data Store (CDS) [@cds] website or via the package \pkg{ecmwfr} [@ecwmfr].
+It can be directly downloaded from Copernicus Climate Data Store (CDS) [@cds] website or via the \pkg{ecmwfr} package [@ecwmfr].
 
@@
 <!--The goal of this example is to reproduce the break-up of the southern polar vortex that happened in late September and early October in 2002 in the stratosphere in the Southern Hemisphere.
-Therefore, we filter data corresponding to the Southern Hemisphere and restrict our analysis to four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04.--> The downloaded netCDF data is read into \proglang{R} using the \pkg{ncdf4} package, and converted to a \code{cubble} object:
+Therefore, we filter data corresponding to the Southern Hemisphere and restrict our analysis to four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04.--> The downloaded NetCDF data is read into \proglang{R} using the \pkg{ncdf4} package, and converted to a \code{cubble} object:
 

### 2022-07-09 5-5-interactive-graphics.Rmd
quick fixes on code markup \pkg{}, \code{}
fix abbreviation + more fix on markups

@@
 This is especially important for spatio-temporal data, where we would like to interactively connect spatial and temporal displays.
-This example describes the process of using \pkg{cubble} with \pkg{crosstalk} to build an interactive display connecting a map of Australia, with ribbon plots of temperature range observed at the stations.
+This example describes the process of using the \pkg{cubble} package with the \pkg{crosstalk} package to build an interactive display connecting a map of Australia, with ribbon plots of temperature range observed at the stations.
 The purpose is to explore the variation of monthly temperature range over the country.
@@
 The maximum and minimum temperatures are shown using a ribbon.
-The bottom row first selects the lowest temperature in August in the seasonal display, which highlights the corresponding station on the map (Thredbo AWS).
-Another  station, located in the Tasmania Island, is then selected to compare its temperature variation with Thredbo AWS.", fig.show='hold'}
+The bottom row first selects the lowest temperature in August in the seasonal display, which highlights the corresponding station on the map (Thredbo).
+Another  station, located in the Tasmania Island, is then selected to compare its temperature variation with the Thredbo station.", fig.show='hold'}
 knitr::include_graphics(here::here("figures/linking.png"))

### 2022-07-09 6-conclusion.Rmd
fix abbreviation + more fix on markups

@@
 Numerous examples are provided in the main text, appendix and package vignettes.
-These include creating and coercing data with mismatched names, handling hierarchical data, matching time series spatially and temporally, reproducing ERA5 plots from netCDF data.
+These include creating and coercing data with mismatched names, handling hierarchical data, matching time series spatially and temporally, reproducing ERA5 plots from NetCDF data.
 Visualization of the \code{cubble} objects can be done with interactive graphic pipelines using \pkg{crosstalk}, \pkg{plotly} and \pkg{leaflet}.

### 2022-07-09 jss/main.tex
fix abbreviation + more fix on markups

@@
 Here I reference \citet{ctvspatiotemporal}, \citet{bach_review_2014}, \citet{lu_multidimensional_2018}, \citet{tidyverse}, \citet{tibble}, \citet{sf}, \citet{stars}, \citet{spacetime}, \citet{dplyr}, \citet{ggplot2}, \citet{tsibble}, \citet{hassell_data_2017}, \citet{tidydata}, \citet{leaflet}, \citet{crosstalk}, \citet{cheng2016enabling}, \citet{buja1996interactive}, \citet{xie2014reactive}, \citet{sutherland2000orca}, \citet{buja1988elements}, \citet{Wickham2012-yr}, \citet{hersbach2020era5}, \citet{tidync}, \citet{rnetcdf}, \citet{ncdf4}, \citet{ecwmfr}, \citet{simmons2020global}, \citet{simmons2005ecmwf}, \citet{mcintosh2018using}, \citet{stuart2010matching}, \citet{castanedo2013review}, \citet{cocchi2019data}, \citet{knitr}, \citet{rmarkdown}, \citet{wang2020calendar}, \citet{advr}, \citet{CS07Author}=``Cook, D.
-and Swayne, D.'',, \citet{bom}, \citet{noaa}, \citet{cds}, \citet{wickham_tourr_2011}
+and Swayne, D.'',, \citet{bom}, \citet{noaa}, \citet{cds}, \citet{wickham_tourr_2011}, \citet{rnoaa}, \citet{ggally}
 
@@
 
-The SpatioTemporal CRAN task view \citep{ctvspatiotemporal} gathers information about \proglang{R} packages designed for spatio-temporal data and it has a section on \emph{Representing data} that lists existing spatio-temporal data representations used in \proglang{R}.
+The Comprehensive \proglang{R} Archive Network (CRAN) task view SpatioTemporal \citep{ctvspatiotemporal} gathers information about \proglang{R} packages designed for spatio-temporal data and it has a section on \emph{Representing data} that lists existing spatio-temporal data representations used in \proglang{R}.
 Among them, \citet{spacetime} summarises spatio-temporal data into three forms: time-wide, space-wide, and long formats.
 The associated package \pkg{spacetime} \citep{spacetime} implements four spatio-temporal layouts (full grid, sparse grid, irregular, and trajectory) to handle different space and time combinations.
-The \pkg{stars} \citep{stars} package has a new implementation to use dense arrays to represent spatio-temporal cubes.
-It also interfaces with \pkg{sf} \citep{sf}, a package commonly used for wrangling spatial data, and the \pkg{tidyverse} \citep{tidyverse} suite for general data wrangling and visualization in \proglang{R}.
+The package \pkg{stars} \citep{stars} has a new implementation to use dense arrays to represent spatio-temporal cubes.
+It also interfaces with the package \pkg{sf} \citep{sf}, commonly used for wrangling spatial data, and the \pkg{tidyverse} \citep{tidyverse} suite for general data wrangling and visualization in \proglang{R}.
 
@@
 According to the tidy data principles, data should be structured into 1) one row per observation, 2) one column per variable, and 3) one type of data per table.
-The long form data is preferred over wide data form given the downstream software such as \pkg{dplyr} \citep{dplyr} and \pkg{ggplot2} \citep{ggplot2} for data wrangling and visualization.
+The long form data is preferred over wide data form given the downstream packages such as \pkg{dplyr} \citep{dplyr} and \pkg{ggplot2} \citep{ggplot2} for data wrangling and visualization.
 However, the long form can be inefficient to store feature geometries, especially for large multipolygons for hourly, daily or sub-daily periods over years, which are extensively collected and handled, for example in time series analysis.
@@
 Section \ref{cubble} presents the main design and functionality of the \pkg{cubble} package.
-Section \ref{others} explains how the \pkg{cubble} package deals with more advanced considerations, including data with hierarchical structure, data matching and how it fits with existing static and interactive visualization tools.
+Section \ref{others} explains how the \pkg{cubble} package deals with more advanced considerations, including data with hierarchical structure, data matching and how the package fits with existing static and interactive visualization tools.
 Moreover we also illustrate how the \pkg{cubble} package deals with spatio-temporal data transformations.
-Section \ref{examples} uses Australian weather station data and river level data as examples to demonstrate the use of the \pkg{cubble} package.
-An example of how cubble handles NetCDF data is also provided.
+Section \ref{examples} uses Australian weather station data and river level data as examples to demonstrate the use of the package.
+An example of how the \pkg{cubble} package handles Network Common Data Form (NetCDF) data is also provided.
 Section \ref{conclude} discuss the paper contributions and future directions.
@@
 The main functions in the package are \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and \code{unfold()}.
-The following sections explain their roles, why the new cubble structure is needed and how the package relates to existing packages for spatial and temporal data analysis.
-
-The data \code{climate_flat} is used to illustrate \pkg{cubble} functionality.
-This is a subset from National Oceanic and Atmospheric Administration (NOAA) Global Historical Climatology Network Daily (GHCND) data.
-It contains spatial variables, station ID, latitude, longitude, elevation, station name, World Meteorology Organisation ID, in additional to daily temporal information, maximum and minimum temperature values and precipitation records for year 2020.
-The first five rows of \code{climate_flat} are shown below:
+The following sections explain their roles, why the new \code{cubble} structure is needed and how the package relates to existing packages for spatial and temporal data analysis.
+
+The data \code{climate_flat} is used to illustrate functionality in the \pkg{cubble} package.
+This is a subset from National Oceanic and Atmospheric Administration (NOAA) \citep{noaa} Global Historical Climatology Network (GHCN) Daily data.
+It contains spatial variables, station id, latitude, longitude, elevation, station name, world meteorology organisation id, in additional to daily temporal information, maximum and minimum temperature values and precipitation records for year 2020.
+The first five rows of the data are shown below:
 
@@
 
-The function \code{as_cubble()} is used to create a cubble with three arguments: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order (longitude, latitude).
+The function \code{as_cubble()} is used to create a \code{cubble} object with three arguments: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order (longitude, latitude).
 The arguments \code{key} and \code{index} follow the wording in the \pkg{tsibble} package to describe the temporal order and multiple series while \code{coords} specifies the spatial location of each site.
-The code below creates a cubble out of \code{climate_flat} (a single tibble) with \code{id} as the key, \code{date} as the index, and \code{c(long, lat)} as the coordinates:
+The code below creates a \code{cubble} object out of \code{climate_flat} (a single tibble) with \code{id} as the key, \code{date} as the index, and \code{c(long, lat)} as the coordinates:
 
@@
 
-A cubble can be created from various common spatio-temporal data formats, including basic \proglang{R} objects like \code{tibble}, \code{tsibble}, and \code{sf}.
-Section \ref{covid} describes converting multiple tables into a \code{cubble} object and Section \ref{netcdf} illustrates how to convert a netCDF object.
+A \code{cubble} object can be created from various common spatio-temporal data formats, including basic \proglang{R} objects like \code{tibble}, \code{tsibble}, and \code{sf}.
+Section \ref{covid} describes converting multiple tables into a \code{cubble} object and Section \ref{netcdf} illustrates how to convert a NetCDF object.
 
@@
 The \code{cubble} class also provides a long form, which expands the \code{ts} column and temporarily ``hides'' the spatial variables.
-The function \code{face_temporal()} is used to switch from the nested cubble into the long one.
+The function \code{face_temporal()} is used to switch from the nested \code{cubble} into the long one.
 The first row in Figure \ref{fig:face} illustrates this operation where the focus of the cube now changes from the site-variable face to the time-variable face.
-This code switches the cubble just created into its long form:
+This code switches the \code{cubble} object just created into its long form:
 
@@
 
-The first line in the printed cubble now shows it in the long form and the third line has been changed to display the name and type of spatial variables: \code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]}.
-Unlike the nested form, the long cubble is built from a \code{grouped_df} class where all the observations from the same site form a group.
+The first line in the header now shows it in the long form and the third line has been changed to display the name and type of spatial variables: \code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]}.
+Unlike the nested form, the long \code{cubble} is built from a \code{grouped_df} class where all the observations from the same site form a group.
 
@@
 
-\caption{An illustration of function \code{face\_temporal} and \code{face\_spatial} in cubble.
-In the first row, \code{face\_temporal} switches a cubble from the nested form into the long form and the focus has switched from the spatial aspect (the side face) to the temporal aspect (the front face).
-In the second row, \code{face\_spatial} switches a cubble back to the nested form from the long form and shifts focus back to the spatial aspect.}\label{fig:face}
+\caption{An illustration of function \code{face\_temporal} and \code{face\_spatial}.
+In the first row, \code{face\_temporal} switches a \code{cubble} object from the nested form into the long form and the focus has switched from the spatial aspect (the side face) to the temporal aspect (the front face).
+In the second row, \code{face\_spatial} switches a \code{cubble} object back to the nested form from the long form and shifts focus back to the spatial aspect.}\label{fig:face}
 \end{figure}
@@
 \hypertarget{unfold}{%
-\subsection{Unfold spatial variables into the long cubble}\label{unfold}}
+\subsection{Unfold spatial variables into the long form}\label{unfold}}
 
@@
 Some readers may question why a new data structure is needed rather than directly creating a list-column on the combined data using \code{dplyr::nest_by()}.
-The reason is that cubble is specifically designed to utilize the spatio-temporal structure when arranging observations in a single object.
+The reason is that the \code{cubble} object is specifically designed to utilize the spatio-temporal structure when arranging observations in a single object.
 Moreover, it enables easy pivoting between purely spatial, purely temporal, or unfolded into a combined form.
@@
 
-The \code{dplyr} package has many tools for wrangling tidy data, many of which are useful in the spatio-temporal analysis.
-The \code{cubble} package provides \code{methods} that support the use of the following \code{dplyr} operations on both the nested and long forms: \texttt{mutate}, \texttt{filter}, \texttt{summarise}, \texttt{select}, \texttt{arrange}, \texttt{rename}, \texttt{left\_join}, and the slice family (\texttt{slice\_*}).
+The \pkg{dplyr} package has many tools for wrangling tidy data, many of which are useful in the spatio-temporal analysis.
+The \pkg{cubble} package provides methods that support the use of the following operations in the \pkg{dplyr} package on both the nested and long forms: \texttt{mutate}, \texttt{filter}, \texttt{summarise}, \texttt{select}, \texttt{arrange}, \texttt{rename}, \texttt{left\_join}, and the slice family (\texttt{slice\_*}).
 
@@
 
-When a nested cubble is created, each element in the list-column \code{ts} has a \code{tsibble} class (labelled \code{tbl_ts}) and operations available to the \code{tsibble} class are still valid on this element.
-For example, the code below calculates two time series features (mean and variance) of maximum temperature, utilizing the tsibble structure in the cubble:
+When a nested \code{cubble} is created, each element in the list-column \code{ts} has a \code{tsibble} class (labelled \code{tbl_ts}) and operations available to the \code{tsibble} class are still valid on this element.
+For example, the code below calculates two time series features (mean and variance) of maximum temperature, utilizing the \code{tsibble} structure in the \code{cubble} object:
 
@@
 An \code{sf} object is a spatial data object, containing a feature geometry list-column (\code{sfc}) in the form of various geometry types (\code{POINT}, \code{LINESTRING}, \code{POLYGON}, \code{MULTIPOLYGON}).
-The \pkg{sf} package provides functions that operate efficiently on this spatial information.
-A cubble can store the spatial information as an \code{sf} object.
+The package \pkg{sf} provides functions that operate efficiently on this spatial information.
+A \code{cubble} object can store the spatial information as an \code{sf} object.
 Methods for the \code{sfc} class can be applied in the nested form of the cubble object.
 An illustration is in Section \ref{covid}.
-The spatial information can also be stored as an \code{s2} in a cubble object.
+The spatial information can also be stored as an \code{s2} vector in a \code{cubble} object.
 
@@
 
-NetCDF data is another format commonly used for storing spatio-temporal data.
+The NetCDF data is another format commonly used for storing spatio-temporal data.
 It has two main components: \emph{dimension} for defining the spatio-temporal grid (longitude, latitude, and time) and \emph{variable} that populates the defined grid.
@@
 
-NetCDF data can be quite large, and it is sometimes best to subset the data when converting to a cubble.
+Sometimes NetCDF data can be quite large, and it is best to subset the data when converting to a \code{cubble}.
 We would recommend reducing to about 300 \(\times\) 300 grid points for three daily variables in one year.
 A 300 by 300 spatial grid can be a bounding box of {[}100, -80, 180, 0{]} at 0.25 degree resolution or a global bounding box {[}-180, -90, 180, -90{]} at 1 degree resolution.
-The size of spatial grid can be reduced if longer time periods or more variables are needed, through \code{long_range} and \code{lat_range}:
+The size of spatial grid can be reduced if longer time periods or more variables are needed, through the arguments \code{long_range} and \code{lat_range}:
 
@@
 +   my_ncdf, vars = c("q", "z"),
-+   long_range = seq(-180, 180, 1),lat_range = seq(-90, 90, 1)
++   long_range = seq(-180, 180, 1), lat_range = seq(-90, 90, 1)
 +   )
@@
 The result can also be stretched into long form.
-By specifying
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> cluster_nested <- station_nested %>% switch_key(key = cluster)
-\end{CodeInput}
-\end{CodeChunk}
-
-the cubble redefines the cubble key from the \code{id} column in \code{station_nested} to the \code{cluster} column in \code{cluster_nested}.
+By specifying \code{cluster_nested <- station_nested \%>\% switch_key(key = cluster)}, the \code{cubble} objects redefines its \code{key} from the \code{id} column in \code{station_nested} to the \code{cluster} column in \code{cluster_nested}.
 All the spatial variables belonging to the \code{cluster} column are now nested into a \code{.val} column which allows for summarizing based on cluster.
@@
 Figure \ref{fig:illu-matching} illustrates this temporal matching.
-In the two series, A and a, three peaks have been identified in each.
-An interval, of fixed length, is constructed for each peak in series A, while the peaks in series a are tested against whether they fall into any of the intervals.
+In the two series, \emph{A} and \emph{a}, three peaks have been identified in each.
+An interval, of fixed length, is constructed for each peak in series \emph{A}, while the peaks in series \emph{a} are tested against whether they fall into any of the intervals.
 Here two out of three peaks match.
@@
 The cubble workflow fits works well with an interactive graphics pipeline (e.g., \citet{buja1988elements}, \citet{buja1996interactive}, \citet{sutherland2000orca}, \citet{xie2014reactive}, \citet{cheng2016enabling}) that is available in \proglang{R} with the package \pkg{crosstalk} \citep{crosstalk}.
-Figure \ref{fig:illu-interactive} illustrates how linking can be achieved between a map and multiple time series using a cubble.
+Figure \ref{fig:illu-interactive} illustrates how linking can be achieved between a map and multiple time series in a \code{cubble} object.
 The map (produced from the nested form) and time series (produced from the long form) are both shared \code{crosstalk} objects.
 When a user makes a selection on the map, the site is highlighted (a).
-This activates a row in the nested \code{cubble} object, which is then communicated to the long cubble -- all the observations with the same id (b) will be selected.
-The long cubble will then highlight the corresponding series in the time series plot (c).
+This activates a row in the nested \code{cubble} object, which is then communicated to the long \code{cubble} -- all the observations with the same id (b) will be selected.
+The long \code{cubble} will then highlight the corresponding series in the time series plot (c).
 
 Linking is also available starting from the time series plot, by selecting points.
-This will be activate rows having the same id in the long cubble.
-The corresponding rows in the nested cubble are activated, and highlighted the map.
+This will be activate rows having the same id in the long \code{cubble} The corresponding rows in the nested \code{cubble} are activated, and highlighted the map.
 (An illustration can be found in the appendix.) Note that this type of linking, both from the map or the time series, is what \citet{CS07} would call categorical variable linking, where station id is the categorical variable.
@@
 
-\caption[Linking between multiple plots]{Linking between multiple plots.
-The line plots and the map are constructed from shared \code{crosstalk} objects (long and nested cubbles).
-When a station is selected on the map (a), the corresponding row in the nested cubble will be activated.
-This will link to all the rows with the same id in the long cubble (b) and update the line plot (c).}\label{fig:illu-interactive}
+\caption{Linking between multiple plots.
+The line plots and the map are constructed from shared \code{crosstalk} objects (long and nested \code{cubble}).
+When a station is selected on the map (a), the corresponding row in the nested \code{cubbles} will be activated.
+This will link to all the rows with the same id in the long \code{cubbles} (b) and update the line plot (c).}\label{fig:illu-interactive}
 \end{figure}
@@
 The package \pkg{GGally} \citep{ggally} has implemented glyph maps through the \code{glyphs()} function.
-The function constructs a data frame with calculated position (\code{gx}, \code{gy}, \code{gid}) of each point on the time series using linear algebra (Equations 1 and 2 in \citet{Wickham2012-yr}).
+The function constructs a \code{data.frame} with calculated position (\code{gx}, \code{gy}, \code{gid}) of each point on the time series using linear algebra (Equations 1 and 2 in \citet{Wickham2012-yr}).
 The data can then be piped into ggplot to create the glyph map as:
@@
 
-The Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world.
+The GHCN provides daily climate measures from stations across the world.
 The data used here (\code{historical_tmax}) is a subset extracted using the package \pkg{rnoaa} \citep{rnoaa}, containing the records of maximum temperature for 236 Australian stations from 1859 through 1969 and provides information also on the latitude, longitude and elevation of each of the stations.
-The goal of this example is to compare the monthly average maximum temperature between two periods, 1971-1975 and 2016-2020, for stations in Victoria and New South Wales, using a glyph map.
-
-First, the stations need to be filtered to those in NSW and Victoria by using the station identifiers, stored within the 11 digits of the \code{id} variable entries.
+The goal of this example is to compare the monthly average maximum temperature between two periods, 1971-1975 and 2016-2020, for stations in Victoria and New South Wales (NSW), using a glyph map.
+
+First, the stations need to be filtered to those in Victoria and NSW by using the station identifiers, stored within the 11 digits of the \code{id} variable entries.
 The country code is in the first 5 digits (Australia is represented by ``ASN00'') and the next 6 digits encode the station following the Australian Bureau of Meteorology (BOM) \citep{bom} coding protocols.
-New South Wales stations correspond to entries in the range 46-75 and the Victorian stations to 76-90.
+The NSW stations correspond to entries in the range 46-75 and the Victorian stations to 76-90.
 Filtering Victoria and New South Wales stations is a \emph{spatial operation} and hence uses the \code{cubble} object in the nested form:
@@
 Next, the monthly maximum average temperature is calculated for both periods.
-This is a \emph{temporal operation} requiring a switch into the long cubble form using the \code{face_temporal()} function.
+This is a \emph{temporal operation} requiring a switch into the long \code{cubble} form using the \code{face_temporal()} function.
 In addition, a new indicator for the two time periods of interest is created before the calculation of monthly averages:
@@
 
-\caption[A glyph map of the monthly maximum average temperature for weather stations in  Victoria and New South Wales for the periods (1971-1975, 2016-2020)]{A glyph map of the monthly maximum average temperature for weather stations in  Victoria and New South Wales for the periods (1971-1975, 2016-2020).
-The corresponding average time series for the Cobar station are display on the top left corner.
-From the glyph map we can observe that the monthly trend is similar for all locations (low in the winter, high in the summer), and small increased temperatures, particularly in late summer can be seen at most stations in New South Wales.}\label{fig:basic-manip}
+\caption[A glyph map of the monthly maximum average temperature for weather stations in  Victoria and New South Wales (NSW) for the periods (1971-1975, 2016-2020)]{A glyph map of the monthly maximum average temperature for weather stations in  Victoria and New South Wales (NSW) for the periods (1971-1975, 2016-2020).
+The corresponding average time series for the cobar station are display on the top left corner.
+From the glyph map we can observe that the monthly trend is similar for all locations (low in the winter, high in the summer), and small increased temperatures, particularly in late summer can be seen at most stations in NSW.}\label{fig:basic-manip}
 \end{figure}
@@
 River level and rainfall data for the same areas should have some similarity.
-Here we examine the river gauge data (\code{Water_course_level}) from the Bureau of Meteorology \citep{bom} in relation to weather station rainfall from NOAA's climate data \citep{noaa} (\code{climate}).
+Here we examine the river gauge data (\code{Water_course_level}) from the Bureau of Meteorology \citep{bom} in relation to weather station rainfall from NOAA's climate data (\code{climate}).
 The goal is to match water gauges with nearby weather stations, spatially and temporally, using 2020 measurements using \code{match_sites()} function.
@@
 
-Figure \ref{fig:netcdf} reproduces the ERA5 data \citep{hersbach2020era5} row of Figure 19 in \citep{hersbach2020era5}.
-Here we explain how this would be done using cubble.
+Figure \ref{fig:netcdf} reproduces the ERA5 data row of Figure 19 in \citep{hersbach2020era5}.
+Here we explain how this would be done using in the \pkg{cubble} package.
 The plots show that the southern polar vortex splits into two on 2002-09-26 and further splits into four on 2002-10-04.
@@
 The data is available in the NetCDF format from the European Centre for Medium-Range Weather Forecasts (ECMWF).
-It can be directly downloaded from Copernicus Climate Data Store (CDS) \citep{cds} website or via the package \pkg{ecmwfr} \citep{ecwmfr}.
+It can be directly downloaded from Copernicus Climate Data Store (CDS) \citep{cds} website or via the \pkg{ecmwfr} package \citep{ecwmfr}.
 
@@
 
-The downloaded netCDF data is read into \proglang{R} using the \pkg{ncdf4} package, and converted to a \code{cubble} object:
+The downloaded NetCDF data is read into \proglang{R} using the \pkg{ncdf4} package, and converted to a \code{cubble} object:
 
@@
 The maximum and minimum temperatures are shown using a ribbon.
-The bottom row first selects the lowest temperature in August in the seasonal display, which highlights the corresponding station on the map (Thredbo AWS).
-Another  station, located in the Tasmania Island, is then selected to compare its temperature variation with Thredbo AWS.}\label{fig:interactive-linking}
+The bottom row first selects the lowest temperature in August in the seasonal display, which highlights the corresponding station on the map (Thredbo).
+Another  station, located in the Tasmania Island, is then selected to compare its temperature variation with the Thredbo station.}\label{fig:interactive-linking}
 \end{figure}

### 2022-07-09 jss/reference-hack.Rmd
fix abbreviation + more fix on markups

--- 2022-06-24 reference-hack.Rmd
@@
 tex <- readLines(here::here("references.bib"))
-
 b <- tex %>% stringr::str_extract("\\@(article|techreport|misc|Manual|Book)\\{.+,")

### 2022-07-11 1-intro.Rmd
more markups

@@
 This paper presents a new \proglang{R} package, \pkg{cubble}, which addresses the two issues mentioned above.
-In the package, a new data structure, also called cubble, is proposed to organize spatial and temporal variables as two forms of a single data object so that they can be wrangled separately or combined while being kept synchronized.
-Among the four spacetime layouts in @spacetime, cubble can be applied to full grid, sparse grid, or irregular, but not trajectory, which is outside the scope of this work.
+In the package, a new class, also called cubble, is proposed to organize spatial and temporal variables as two forms of a single data object so that they can be wrangled separately or combined while being kept synchronized.
+Among the four spacetime layouts in @spacetime, the \code{cubble} class can be applied to full grid, sparse grid, or irregular, but not trajectory, which is outside the scope of this work.
 The software is available from the Comprehensive \proglang{R} Archive Network (CRAN) at https://CRAN.R-project.org/package=cubble.

### 2022-07-11 2-cube.Rmd
more markups

@@
 Finally, variables are stacked on this space-time canvas, with one observation per site and time point.
-This notion is adopted to avoid using hyper-cubes when describing multivariate spatio-temporal data and is the conceptual framework behind the \code{cubble} objects.
+This notion is adopted to avoid using hyper-cubes when describing multivariate spatio-temporal data and is the conceptual framework behind the \code{cubble} class.
 With this conceptual model, operations on spatio-temporal data can be mapped to operations on the cube and the rest of Figure \ref{fig:cube-def} show examples of slicing on site, time, and variable.

### 2022-07-11 3-the-cubble-pkg.Rmd
more markups

@@
 
-A \code{cubble} object is an S3 class [@advr] built on the \code{tibble} class, specifically to organize spatio-temporal data.
+The \code{cubble} class is an S3 class [@advr] built on the \code{tibble} class, specifically to organize spatio-temporal data.
 The \code{cubble} class uses an attribute "form",  to arrange the spatial or temporal data components tidily.
 The form attribute can take a value of either "nested" or "long".
-The nested cubble is a subclass of rowwise tibble (\code{rowwise_df}).
+The nested \code{cubble} is a subclass of rowwise \code{tibble} (\code{rowwise_df}).
 It arranges each spatial site in a row, and uses list columns to store the feature geometry and the temporal information.
-The long cubble is a subclass of grouped tibble (\code{grouped_df}), which expands the temporal information into the long form and stores the spatial information in a "spatial" attribute.
+The long \code{cubble} is a subclass of grouped \code{tibble} (\code{grouped_df}), which expands the temporal information into the long form and stores the spatial information in a "spatial" attribute.
 
@@
 The main functions in the package are \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and \code{unfold()}.
-The following sections explain their roles, why the new \code{cubble} structure is needed and how the package relates to existing packages for spatial and temporal data analysis.
+The following sections explain their roles, why the new \code{cubble} class is needed and how the package relates to existing packages for spatial and temporal data analysis.
 
@@
 The arguments \code{key} and \code{index} follow the wording in the \pkg{tsibble} package to describe the temporal order and multiple series while \code{coords} specifies the spatial location of each site.
-The code below creates a \code{cubble} object out of \code{climate_flat} (a single tibble) with \code{id} as the key, \code{date} as the index, and \code{c(long, lat)} as the coordinates:
+The code below creates a \code{cubble} object out of \code{climate_flat} (a single \code{tibble}) with \code{id} as the key, \code{date} as the index, and \code{c(long, lat)} as the coordinates:
 
@@
 
-A \code{cubble} object is a subclass of the \code{rowwise_df} class where each row forms a group.
+The created \code{cubble} is a subclass of the \code{rowwise_df} class where each row forms a group.
 All the temporal variables are nested in a list column, hence it is also called the nested \code{cubble}.
@@
 
-A \code{cubble} object can be created from various common spatio-temporal data formats, including basic \proglang{R} objects like \code{tibble}, \code{tsibble}, and \code{sf}.
+A \code{cubble} class can be created from various common spatio-temporal data formats, including \code{tibble}, \code{tsibble}, and \code{sf} when both spatial and tmeporal information are available.
 Section \@ref(covid) describes converting multiple tables into a \code{cubble} object and Section \@ref(netcdf) illustrates how to convert a NetCDF object.
@@
 (How to make glyph maps will be explained in Section \ref{st_transformation}, and are illustrated in the second example.) This type of operation can be seen as flattening, or *unfolding*, the cube into a 2D data frame.
-Here the function \code{unfold()} moves the spatial variables \code{long} and \code{lat} into the long cubble:
+Here the function \code{unfold()} moves the spatial variables \code{long} and \code{lat} into the long \code{cubble}:
 
@@
 
-The \pkg{cubble} package leverages tools available in existing packages used for spatial and temporal analysis, specifically, \pkg{dplyr}, \pkg{tsibble}, \pkg{sf} (\pkg{s2}), and \pkg{netcdf4}, as explained here.
+The \pkg{cubble} package leverages tools available in existing packages used for spatial and temporal analysis, specifically, \pkg{dplyr}, \pkg{tsibble}, \pkg{sf} (\pkg{s2}), and \pkg{ncdf4}, as explained here.
 
-### dplyr
+\subsubsection[]{\code{dplyr}}
 
@@
 
-### tsibble
+\subsubsection[]{\code{tsibble}}
 
-A \code{tsibble} object is a \code{tibble} where the \code{index} and \code{key} components are used to store temporal and strata information, that makes working with temporal data cognitively efficient.
-A \code{cubble} object can use a \code{tsibble} to store the temporal information, and effectively utilize the specialist time series operations in the \pkg{tsibble} package.
-It is easy to cast an existing \code{tsibble} object into a \code{cubble} simply by supplying the argument \code{coords}:
+The \code{tsibble} class is a subclass of \code{tibble} where the \code{index} and \code{key} components are used to store temporal and strata information, that makes working with temporal data cognitively efficient.
+A \code{cubble} object can use the \code{tsibble} class to store the temporal information, and effectively utilize the specialist time series operations in the \pkg{tsibble} package.
+A \code{tsibble} object can also be casted into a \code{cubble} object through supplying the coordinate information in the argument `coords`:
 
@@
 
-When a nested \code{cubble} is created, each element in the list-column \code{ts} has a \code{tsibble} class (labelled \code{tbl_ts})  and operations available to the \code{tsibble} class are still valid on this element.
-For example, the code below calculates two time series features (mean and variance) of maximum temperature, utilizing the \code{tsibble} structure in the \code{cubble} object:
+When a nested \code{cubble} is created, each element in the list-column \code{ts} is in the \code{tsibble} class (labelled \code{tbl_ts})  and operations available to the \code{tsibble} class are still valid on these elements.
+For example, the code below calculates two time series features (mean and variance) of maximum temperature, utilizing the \code{tsibble} syntax in the \code{cubble} object:
 
@@
 
-### sf and s2
+\subsubsection[]{\code{sf} (\code{s2})}
 
-An \code{sf} object is a spatial data object, containing a feature geometry list-column (\code{sfc}) in the form of various geometry types (\code{POINT}, \code{LINESTRING}, \code{POLYGON}, \code{MULTIPOLYGON}).
+The \code{sf} class is also a subclass of \code{tibble} with a specialized feature geometry list-column (\code{sfc}) to store different geometry types (\code{POINT}, \code{LINESTRING}, \code{POLYGON}, \code{MULTIPOLYGON}, etc).
 The package \pkg{sf} provides functions that operate efficiently on this spatial information.
-A \code{cubble} object can store the spatial information as an \code{sf} object.
+A \code{cubble} object can store spatial information in  the \code{sf} class.
 Methods for the \code{sfc} class can be applied in the nested form of the cubble object.
@@
 
-### netCDF {#netcdf}
+\subsubsection[]{\code{ncdf4}}\label{netcdf}
 
@@
 
-Sometimes NetCDF data can be quite large, and it is best to subset the data when converting to a \code{cubble}.
+Sometimes NetCDF data can be quite large, and it is best to subset the data when converting to a \code{cubble} object.
 We would recommend reducing to about 300 $\times$ 300 grid points for three daily variables in one year.

### 2022-07-11 4-other-features.Rmd
more markups

@@
 The result can also be stretched into long form.
-By specifying \code{cluster_nested <- station_nested \%>\% switch_key(key = cluster)}, the \code{cubble} objects redefines its \code{key} from the \code{id} column in \code{station_nested} to the \code{cluster} column in \code{cluster_nested}.
+By specifying \code{cluster_nested <- station_nested \%>\% switch_key(key = cluster)}, the \code{cubble} object redefines its \code{key} from the \code{id} column in \code{station_nested} to the \code{cluster} column in \code{cluster_nested}.
 All the spatial variables belonging to the \code{cluster} column are now nested into a \code{.val} column which allows for summarizing based on cluster.
@@
 
-```{r illu-matching, out.width="100%", fig.cap="An illustration of temporal matching in cubble.
+```{r illu-matching, out.width="100%", fig.cap="An illustration of temporal matching in the \\pkg{cubble} package.
 Three highest peaks are identified in each series and intervals are constructed on series \\code{A}.
@@
 
-The cubble workflow fits works well with an interactive graphics pipeline (e.g., @buja1988elements, @buja1996interactive,  @sutherland2000orca, @xie2014reactive, @cheng2016enabling) that is available in \proglang{R} with the package \pkg{crosstalk} [@crosstalk].
+The workflow with the \code{cubble} class fits works well with an interactive graphics pipeline (e.g., @buja1988elements, @buja1996interactive,  @sutherland2000orca, @xie2014reactive, @cheng2016enabling) that is available in \proglang{R} with the package \pkg{crosstalk} [@crosstalk].
 Figure \ref{fig:illu-interactive} illustrates how linking can be achieved between a map and multiple time series in a \code{cubble} object.
@@
 When a user makes a selection on the map, the site is highlighted (a).
-This activates a row in the nested \code{cubble} object, which is then communicated to the long \code{cubble} -- all the observations with the same id (b) will be selected.
+This activates a row in the nested \code{cubble}, which is then communicated to the long \code{cubble} -- all the observations with the same id (b) will be selected.
 The long \code{cubble} will then highlight the corresponding series in the time series plot (c).
@@
 The line plots and the map are constructed from shared \\code{crosstalk} objects (long and nested \\code{cubble}).
-When a station is selected on the map (a), the corresponding row in the nested \\code{cubbles} will be activated.
-This will link to all the rows with the same id in the long \\code{cubbles} (b) and update the line plot (c)."}
+When a station is selected on the map (a), the corresponding row in the nested \\code{cubble} will be activated.
+This will link to all the rows with the same id in the long \\code{cubble} (b) and update the line plot (c)."}
 knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.004.png"))

### 2022-07-11 5-1-covid.Rmd
more markups

--- 2022-07-09 5-1-covid.Rmd
@@
 
-The five examples here are chosen to illustrate these aspects of using cubble: creating a \code{cubble} object from two COVID data tables with the complication of differing location names, using spatial transformations to make a glyph map of seasonal temperature changes over years, aggregating information spatially to explore precipitation patterns, matching river level data and weather station records for analysis of water supply, reading netCDF format data to reproduce a climate reanalysis plot, and the workflow to create complex interactive linked plots.
+The five examples here are chosen to illustrate these aspects of the \pkg{cubble} package: creating a \code{cubble} object from two Coronavirus (COVID) data tables with the complication of differing location names, using spatial transformations to make a glyph map of seasonal temperature changes over years, aggregating information spatially to explore precipitation patterns, matching river level data and weather station records for analysis of water supply, reading NetCDF format data to reproduce a climate reanalysis plot, and the workflow to create complex interactive linked plots.
 (There is an additional example and figures in the Appendix, and more examples in the package vignettes.)

### 2022-07-11 5-2-aus-hist-max-temp.Rmd
more markups

@@
 The NSW stations correspond to entries in the range 46-75 and the Victorian stations to 76-90.
-Filtering Victoria and New South Wales stations is a *spatial operation* and hence uses the \code{cubble} object in the nested form:
+Filtering Victoria and NSW stations is a *spatial operation* and hence uses the nested \code{cubble}:
 
@@
 Next, the monthly maximum average temperature is  calculated for both periods.
-This is a *temporal operation* requiring a switch into the long \code{cubble} form using the \code{face_temporal()} function.
+This is a *temporal operation* requiring a switch into the long \code{cubble} using the \code{face_temporal()} function.
 In addition, a new indicator for the two time periods of interest is created before the calculation of monthly averages:

### 2022-07-11 5-4-era5.Rmd
more markups

@@
 It can be directly downloaded from Copernicus Climate Data Store (CDS) [@cds] website or via the \pkg{ecmwfr} package [@ecwmfr].
+For the reproduction, we focus on the \code{era5-pressure} data, hourly pressure levels from 1970 to present, with the *specific humidity* and *geopotential*.
+The downloaded NetCDF data is read into \proglang{R} using the \pkg{ncdf4} package, and converted to a \code{cubble} object:
 
-For the reproduction, we focus on the \code{era5-pressure} data, hourly pressure levels from 1970 to present, with the *specific humidity* and *geopotential*.
 <!--Here we specifically look at the information  contained on the variables *specific humidity* measuring the mass of water vapor per kilogram of moist air  and the *geopotential* variable recording the gravitational potential energy of a unit mass, at a particular location, relative to mean sea level.-->
@@
 <!--The goal of this example is to reproduce the break-up of the southern polar vortex that happened in late September and early October in 2002 in the stratosphere in the Southern Hemisphere.
-Therefore, we filter data corresponding to the Southern Hemisphere and restrict our analysis to four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04.--> The downloaded NetCDF data is read into \proglang{R} using the \pkg{ncdf4} package, and converted to a \code{cubble} object:
-
+Therefore, we filter data corresponding to the Southern Hemisphere and restrict our analysis to four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04.-->
 ```{r echo = TRUE}

### 2022-07-11 6-conclusion.Rmd
more markups

@@
 This paper presents an \proglang{R} package \pkg{cubble} for organizing, manipulating and visualizing spatio-temporal data.
-The package introduces a new data structure for spatio-temporal data, \code{cubble}, that connects the time invariant and varying variables and that allows the user to work with a nested and long form of the data.
+The package introduces a new data class for spatio-temporal data, \code{cubble}, that connects the time invariant and varying variables and that allows the user to work with a nested and long form of the data.
 This work adds capabilities into the spatio-temporal practitioners toolbox to integrate it with a tidy data framework.

### 2022-07-11 jss/main.Rmd
more markups

@@
 However for some studies, it is important to analyze different aspects of the spatio-temporal data simultaneously, like for instance, temporal trends of multiple variables across locations.
-In order to facilitate the study of different portions or combinations of spatio-temporal data, we introduce a new data structure, \code{cubble}, with a suite of functions enabling easy slicing and dicing on the different components spatio-temporal components.
-The proposed \code{cubble} structure ensures that all the components of the data are easy to access and manipulate while providing flexibility for data analysis.
-In addition, \code{cubble} facilitates visual and numerical explorations of the data while easing data wrangling and modelling.
-The \code{cubble} structure and the functions provided in the \pkg{cubble} \proglang{R} package equip users with the capability to handle hierarchical spatial and temporal structures.
-The \code{cubble} structure and the tools implemented in the package are illustrated with different examples of Australian climate data.
+In order to facilitate the study of different portions or combinations of spatio-temporal data, we introduce a new class, \code{cubble}, with a suite of functions enabling easy slicing and dicing on the different components spatio-temporal components.
+The proposed \code{cubble} class ensures that all the components of the data are easy to access and manipulate while providing flexibility for data analysis.
+In addition, the \pkg{cubble} package facilitates visual and numerical explorations of the data while easing data wrangling and modelling.
+The \code{cubble} class and the functions provided in the \pkg{cubble} \proglang{R} package equip users with the capability to handle hierarchical spatial and temporal structures.
+The \code{cubble} class and the tools implemented in the package are illustrated with different examples of Australian climate data.
 keywords:

### 2022-07-11 jss/main.tex
more markups

@@
 However for some studies, it is important to analyze different aspects of the spatio-temporal data simultaneously, like for instance, temporal trends of multiple variables across locations.
-In order to facilitate the study of different portions or combinations of spatio-temporal data, we introduce a new data structure, \code{cubble}, with a suite of functions enabling easy slicing and dicing on the different components spatio-temporal components.
-The proposed \code{cubble} structure ensures that all the components of the data are easy to access and manipulate while providing flexibility for data analysis.
-In addition, \code{cubble} facilitates visual and numerical explorations of the data while easing data wrangling and modelling.
-The \code{cubble} structure and the functions provided in the \pkg{cubble} \proglang{R} package equip users with the capability to handle hierarchical spatial and temporal structures.
-The \code{cubble} structure and the tools implemented in the package are illustrated with different examples of Australian climate data.
+In order to facilitate the study of different portions or combinations of spatio-temporal data, we introduce a new class, \code{cubble}, with a suite of functions enabling easy slicing and dicing on the different components spatio-temporal components.
+The proposed \code{cubble} class ensures that all the components of the data are easy to access and manipulate while providing flexibility for data analysis.
+In addition, the \pkg{cubble} package facilitates visual and numerical explorations of the data while easing data wrangling and modelling.
+The \code{cubble} class and the functions provided in the \pkg{cubble} \proglang{R} package equip users with the capability to handle hierarchical spatial and temporal structures.
+The \code{cubble} class and the tools implemented in the package are illustrated with different examples of Australian climate data.
 }
@@
 This paper presents a new \proglang{R} package, \pkg{cubble}, which addresses the two issues mentioned above.
-In the package, a new data structure, also called cubble, is proposed to organize spatial and temporal variables as two forms of a single data object so that they can be wrangled separately or combined while being kept synchronized.
-Among the four spacetime layouts in \citet{spacetime}, cubble can be applied to full grid, sparse grid, or irregular, but not trajectory, which is outside the scope of this work.
+In the package, a new class, also called cubble, is proposed to organize spatial and temporal variables as two forms of a single data object so that they can be wrangled separately or combined while being kept synchronized.
+Among the four spacetime layouts in \citet{spacetime}, the \code{cubble} class can be applied to full grid, sparse grid, or irregular, but not trajectory, which is outside the scope of this work.
 The software is available from the Comprehensive \proglang{R} Archive Network (CRAN) at \url{https://CRAN.R-project.org/package=cubble}.
@@
 
-A \code{cubble} object is an S3 class \citep{advr} built on the \code{tibble} class, specifically to organize spatio-temporal data.
+The \code{cubble} class is an S3 class \citep{advr} built on the \code{tibble} class, specifically to organize spatio-temporal data.
 The \code{cubble} class uses an attribute ``form'', to arrange the spatial or temporal data components tidily.
 The form attribute can take a value of either ``nested'' or ``long''.
-The nested cubble is a subclass of rowwise tibble (\code{rowwise_df}).
+The nested \code{cubble} is a subclass of rowwise \code{tibble} (\code{rowwise_df}).
 It arranges each spatial site in a row, and uses list columns to store the feature geometry and the temporal information.
-The long cubble is a subclass of grouped tibble (\code{grouped_df}), which expands the temporal information into the long form and stores the spatial information in a ``spatial'' attribute.
+The long \code{cubble} is a subclass of grouped \code{tibble} (\code{grouped_df}), which expands the temporal information into the long form and stores the spatial information in a ``spatial'' attribute.
 
 The main functions in the package are \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and \code{unfold()}.
-The following sections explain their roles, why the new \code{cubble} structure is needed and how the package relates to existing packages for spatial and temporal data analysis.
+The following sections explain their roles, why the new \code{cubble} class is needed and how the package relates to existing packages for spatial and temporal data analysis.
 
@@
 
-A \code{cubble} object can be created from various common spatio-temporal data formats, including basic \proglang{R} objects like \code{tibble}, \code{tsibble}, and \code{sf}.
+A \code{cubble} class can be created from various common spatio-temporal data formats, including \code{tibble}, \code{tsibble}, and \code{sf} when both spatial and tmeporal information are available.
 Section \ref{covid} describes converting multiple tables into a \code{cubble} object and Section \ref{netcdf} illustrates how to convert a NetCDF object.
@@
 (How to make glyph maps will be explained in Section \ref{st_transformation}, and are illustrated in the second example.) This type of operation can be seen as flattening, or \emph{unfolding}, the cube into a 2D data frame.
-Here the function \code{unfold()} moves the spatial variables \code{long} and \code{lat} into the long cubble:
+Here the function \code{unfold()} moves the spatial variables \code{long} and \code{lat} into the long \code{cubble}:
 
@@
 
-The \pkg{cubble} package leverages tools available in existing packages used for spatial and temporal analysis, specifically, \pkg{dplyr}, \pkg{tsibble}, \pkg{sf} (\pkg{s2}), and \pkg{netcdf4}, as explained here.
-
-\hypertarget{dplyr}{%
-\subsubsection{dplyr}\label{dplyr}}
+The \pkg{cubble} package leverages tools available in existing packages used for spatial and temporal analysis, specifically, \pkg{dplyr}, \pkg{tsibble}, \pkg{sf} (\pkg{s2}), and \pkg{ncdf4}, as explained here.
+
+\subsubsection[]{\code{dplyr}}
 
@@
 
-\hypertarget{tsibble}{%
-\subsubsection{tsibble}\label{tsibble}}
-
-A \code{tsibble} object is a \code{tibble} where the \code{index} and \code{key} components are used to store temporal and strata information, that makes working with temporal data cognitively efficient.
-A \code{cubble} object can use a \code{tsibble} to store the temporal information, and effectively utilize the specialist time series operations in the \pkg{tsibble} package.
-It is easy to cast an existing \code{tsibble} object into a \code{cubble} simply by supplying the argument \code{coords}:
+\subsubsection[]{\code{tsibble}}
+
+The \code{tsibble} class is a subclass of \code{tibble} where the \code{index} and \code{key} components are used to store temporal and strata information, that makes working with temporal data cognitively efficient.
+A \code{cubble} object can use the \code{tsibble} class to store the temporal information, and effectively utilize the specialist time series operations in the \pkg{tsibble} package.
+A \code{tsibble} object can also be casted into a \code{cubble} object through supplying the coordinate information in the argument \texttt{coords}:
 
@@
 
-When a nested \code{cubble} is created, each element in the list-column \code{ts} has a \code{tsibble} class (labelled \code{tbl_ts}) and operations available to the \code{tsibble} class are still valid on this element.
-For example, the code below calculates two time series features (mean and variance) of maximum temperature, utilizing the \code{tsibble} structure in the \code{cubble} object:
+When a nested \code{cubble} is created, each element in the list-column \code{ts} is in the \code{tsibble} class (labelled \code{tbl_ts}) and operations available to the \code{tsibble} class are still valid on these elements.
+For example, the code below calculates two time series features (mean and variance) of maximum temperature, utilizing the \code{tsibble} syntax in the \code{cubble} object:
 
@@
 
-\hypertarget{sf-and-s2}{%
-\subsubsection{sf and s2}\label{sf-and-s2}}
-
-An \code{sf} object is a spatial data object, containing a feature geometry list-column (\code{sfc}) in the form of various geometry types (\code{POINT}, \code{LINESTRING}, \code{POLYGON}, \code{MULTIPOLYGON}).
+\subsubsection[]{\code{sf} (\code{s2})}
+
+The \code{sf} class is also a subclass of \code{tibble} with a specialized feature geometry list-column (\code{sfc}) to store different geometry types (\code{POINT}, \code{LINESTRING}, \code{POLYGON}, \code{MULTIPOLYGON}, etc).
 The package \pkg{sf} provides functions that operate efficiently on this spatial information.
-A \code{cubble} object can store the spatial information as an \code{sf} object.
+A \code{cubble} object can store spatial information in the \code{sf} class.
 Methods for the \code{sfc} class can be applied in the nested form of the cubble object.
@@
 
-\hypertarget{netcdf}{%
-\subsubsection{netCDF}\label{netcdf}}
+\subsubsection[]{\code{ncdf4}}\label{netcdf}
 
@@
 
-Sometimes NetCDF data can be quite large, and it is best to subset the data when converting to a \code{cubble}.
+Sometimes NetCDF data can be quite large, and it is best to subset the data when converting to a \code{cubble} object.
 We would recommend reducing to about 300 \(\times\) 300 grid points for three daily variables in one year.
@@
 
-\caption{An illustration of temporal matching in cubble.
+\caption{An illustration of temporal matching in the \pkg{cubble} package.
 Three highest peaks are identified in each series and intervals are constructed on series \code{A}.
@@
 
-The cubble workflow fits works well with an interactive graphics pipeline (e.g., \citet{buja1988elements}, \citet{buja1996interactive}, \citet{sutherland2000orca}, \citet{xie2014reactive}, \citet{cheng2016enabling}) that is available in \proglang{R} with the package \pkg{crosstalk} \citep{crosstalk}.
+The workflow with the \code{cubble} class fits works well with an interactive graphics pipeline (e.g., \citet{buja1988elements}, \citet{buja1996interactive}, \citet{sutherland2000orca}, \citet{xie2014reactive}, \citet{cheng2016enabling}) that is available in \proglang{R} with the package \pkg{crosstalk} \citep{crosstalk}.
 Figure \ref{fig:illu-interactive} illustrates how linking can be achieved between a map and multiple time series in a \code{cubble} object.
@@
 The line plots and the map are constructed from shared \code{crosstalk} objects (long and nested \code{cubble}).
-When a station is selected on the map (a), the corresponding row in the nested \code{cubbles} will be activated.
-This will link to all the rows with the same id in the long \code{cubbles} (b) and update the line plot (c).}\label{fig:illu-interactive}
+When a station is selected on the map (a), the corresponding row in the nested \code{cubble} will be activated.
+This will link to all the rows with the same id in the long \code{cubble} (b) and update the line plot (c).}\label{fig:illu-interactive}
 \end{figure}
@@
 
-The five examples here are chosen to illustrate these aspects of the \pkg{cubble} package: creating a \code{cubble} object from two coronavirus (COVID) data tables with the complication of differing location names, using spatial transformations to make a glyph map of seasonal temperature changes over years, aggregating information spatially to explore precipitation patterns, matching river level data and weather station records for analysis of water supply, reading NetCDF format data to reproduce a climate reanalysis plot, and the workflow to create complex interactive linked plots.
+The five examples here are chosen to illustrate these aspects of the \pkg{cubble} package: creating a \code{cubble} object from two Coronavirus (COVID) data tables with the complication of differing location names, using spatial transformations to make a glyph map of seasonal temperature changes over years, aggregating information spatially to explore precipitation patterns, matching river level data and weather station records for analysis of water supply, reading NetCDF format data to reproduce a climate reanalysis plot, and the workflow to create complex interactive linked plots.
 (There is an additional example and figures in the Appendix, and more examples in the package vignettes.)
@@
 The NSW stations correspond to entries in the range 46-75 and the Victorian stations to 76-90.
-Filtering Victoria and New South Wales stations is a \emph{spatial operation} and hence uses the \code{cubble} object in the nested form:
+Filtering Victoria and NSW stations is a \emph{spatial operation} and hence uses the nested \code{cubble}:
 
@@
 Next, the monthly maximum average temperature is calculated for both periods.
-This is a \emph{temporal operation} requiring a switch into the long \code{cubble} form using the \code{face_temporal()} function.
+This is a \emph{temporal operation} requiring a switch into the long \code{cubble} using the \code{face_temporal()} function.
 In addition, a new indicator for the two time periods of interest is created before the calculation of monthly averages:
@@
 It can be directly downloaded from Copernicus Climate Data Store (CDS) \citep{cds} website or via the \pkg{ecmwfr} package \citep{ecwmfr}.
-
 For the reproduction, we focus on the \code{era5-pressure} data, hourly pressure levels from 1970 to present, with the \emph{specific humidity} and \emph{geopotential}.
-
 The downloaded NetCDF data is read into \proglang{R} using the \pkg{ncdf4} package, and converted to a \code{cubble} object:

### 2022-07-11 code/reproducible-script.Rmd
all the codes in a single script

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  cache = TRUE,
  fig.path = here::here("figures/")
)

options(prompt = "R> ", continue = "+ ",
        tibble.print_max = 5, tibble.print_min = 5,
        width = 70)

```
```{r echo = FALSE}
library(cubble)
library(tidyverse)
library(patchwork)

```
```{r }
cubble::climate_flat

```
```{r echo = TRUE}
cubble_nested <- climate_flat |>
  as_cubble(key = id, index = date, coords = c(long, lat))
cubble_nested

```
```{r echo = TRUE}
cubble_nested |>
  mutate(rain_day = sum(ts$prcp != 0))

```
```{r echo = TRUE}
cubble_long <- cubble_nested |> face_temporal()
cubble_long

```
```{r echo = TRUE}
cubble_back <- cubble_long |> face_spatial()
cubble_back
identical(cubble_nested, cubble_back)

```
```{r echo = TRUE}
cubble_unfold <- cubble_long |> unfold(long, lat)
cubble_unfold

```
```{r echo = TRUE}
climate_flat_ts <- climate_flat |> 
  tsibble::as_tsibble(key = id, index = date)
climate_flat_cb <-  climate_flat_ts |> 
  cubble::as_cubble(coords = c(long, lat))
climate_flat_cb

```
```{r echo = TRUE}
climate_flat_cb |> 
  mutate(
    fabletools::features(
      ts, tmax, list(tmax_mean = mean, tmax_var = var)
      )
  )

```
```{r eval = FALSE}
## library(tidyverse)
## #https://www.coronavirus.vic.gov.au/victorian-coronavirus-covid-19-data
## raw <- read_csv(here::here("data/NCOV_COVID_Cases_by_LGA_Source_20220324.csv"),
##                   col_names = c("date", "postcode", "source", "lga"), skip = 1)
## 
## covid <- raw |>
##   count(date, lga, source, postcode) |>
##   filter(date >= as.Date("2022-01-01")) |>
##   filter(source == "Contact with a confirmed case") |>
##   group_by(date, lga, source) |>
##   count() |>
##   ungroup(date) |>
##   mutate(roll_mean = zoo::rollmean(n, k = 14, fill = NA)) |>
##   tsibble::as_tsibble(key = lga, index = date) |>
##   tsibble::fill_gaps(.full = TRUE)
## save(covid, file = here::here("data/covid.rda"))

```
```{r }
load(here::here("data/covid.rda"))

lga <- strayr::read_absmap("lga2018") |>
  rename(lga = lga_name_2018) |>
  filter(state_name_2016 == "Victoria")

```
```{r echo = TRUE}
covid |> head(5)

```
```{r echo = TRUE}
lga |> head(5)

```
```{r echo = TRUE, warning = TRUE, message= TRUE}
cb <- as_cubble(
  list(spatial = lga, temporal = covid),
  key = lga, index = date, coords = c(cent_long, cent_lat)
  )

```
```{r echo = TRUE}
pair <- as_cubble(
  list(spatial = lga, temporal = covid),
  key = lga, index = date, coords = c(cent_long, cent_lat),
  output = "unmatch"
  )

pair

```
```{r echo = TRUE}
lga <- lga |>
  mutate(lga = ifelse(lga == "Kingston (C) (Vic.)", "Kingston (C)", lga),
         lga = ifelse(lga == "Latrobe (C) (Vic.)", "Latrobe (C)", lga)) |>
  filter(!lga %in% pair$others$spatial)

covid <- covid |> filter(!lga %in% pair$others$temporal)

cb <- as_cubble(data = list(spatial = lga, temporal = covid),
                key = lga, index = date, coords = c(cent_long, cent_lat))

```
```{r }
load(here::here("data/historical_tmax.rda"))

```
```{r echo = TRUE}
tmax <- historical_tmax |>
  filter(between(stringr::str_sub(id, 7, 8), 46, 90))

```
```{r echo = TRUE}
tmax <- tmax |>
  face_temporal() |> 
  group_by(month = lubridate::month(date),
         group = as.factor(
           ifelse(lubridate::year(date) > 2015,
           "2016 ~ 2020", "1971 ~ 1975"))) |>
  summarise(tmax = mean(tmax, na.rm = TRUE))

```
```{r echo = TRUE}
tmax |> 
  face_spatial() |> 
  mutate(n = nrow(ts)) |>
  arrange(n) |>
  pull(n) |> 
  head(10)

tmax <- tmax |> 
  face_spatial() |> 
  filter(nrow(ts) == 24) |>
  face_temporal()

```
```{r }
cobar <- tmax |> filter(id == "ASN00048027") |>
  ggplot(aes(x = month, y = tmax, color = group)) +
  geom_line() +
  scale_color_brewer(palette = "Dark2", guide = "none") +
  scale_x_continuous(
    breaks = seq(1, 12, 1), 
    labels = c("J", "F", "M", "A",
               "M", "J", "J", "A",
               "S", "O", "N", "D")
    ) +
  labs(x = "", y  = "Temp (C)") +
  theme_bw() + 
  theme(aspect.ratio = 0.3)

```
```{r basic-manip, out.width="100%", fig.height = 5, fig.width = 10, fig.cap="A glyph map of the monthly maximum average temperature for weather stations in  Victoria and New South Wales (NSW) for the periods (1971-1975, 2016-2020). The corresponding average time series for the cobar station are display on the top left corner. From the glyph map we can observe that the monthly trend is similar for all locations (low in the winter, high in the summer), and small increased temperatures, particularly in late summer can be seen at most stations in NSW."}
tmax <- tmax |> unfold(long, lat)
box_df <- tmax |> face_spatial() |> filter(id == "ASN00048027")
nsw_vic <- ozmaps::abs_ste |> filter(NAME %in% c("Victoria","New South Wales"))

p1 <- ggplot() + 
  geom_sf(data = nsw_vic, fill = "grey95", 
          color = "white", size=2) + 
  geom_glyph(
    data = tmax, 
    aes(x_major = long, x_minor = month, 
        y_major = lat, y_minor = tmax,
        group = interaction(id, group), color = group),
    width = 1.2, height = 0.8) +
    geom_rect(
      data = box_df,
      aes(xmin = long - 0.6, xmax = long + 0.6,
          ymin = lat - 0.12, ymax = lat + 0.35),
      fill = "transparent", color = "black") +
  scale_color_brewer("", palette = "Dark2") + 
  coord_sf(xlim = c(141, 154), ylim = c(-39, -28.5)) + 
  ggthemes::theme_map() +
  theme(legend.position = "bottom", legend.text = element_text(size = 10)) + 
  guides(color = guide_legend(override.aes = list(size=2)))

(p1 | cobar)  + 
  patchwork::plot_layout(width = c(1,1)) + 
  plot_annotation(tag_levels = 'a')

```
```{r }
load(here::here("data/climate_full.rda"))

climate <- climate_full |>
  filter(between(stringr::str_sub(id, 7, 8), 76, 90)) |>
  face_temporal() |>
  filter(lubridate::year(date) == 2020) |>
  face_spatial() |>
  mutate(type = "climate")

river <- river |> mutate(type = "river")

```
```{r echo = TRUE}
res <- match_sites(
  river, climate,
  temporal_by = c("Water_course_level" = "prcp"),
  temporal_independent = "prcp",  
  temporal_n_highest = 30,
  temporal_min_match = 15, 
)

```
```{r }
res |> print()

```
```{r matching, out.width="100%", fig.height = 5, fig.width = 10, fig.cap="Weather stations and river gauges with matched pairs labelled on the map (a) and plotted across time (b). Precipitation and water level have been standardised between 0 and 1 to be displayed on the same scale. The water level reflects the increase in precipitation. The numbers (1, 5, 6, 10) indicate the group index derived from spatial matching, only those that were selectd by temporal matching are shown here."}
vic_map <- ozmaps::abs_ste |> 
  filter(NAME == "Victoria") |> 
  rmapshaper::ms_simplify()

p1 <-ggplot() + 
  geom_sf(data = vic_map, fill = "grey95", color = "white") + 
  geom_point(data = dplyr::bind_rows(river, climate), 
             aes(x = long, y = lat, color = type), alpha = 0.2, fill = 0.2) +
  geom_point(data = res, aes(x = long, y = lat, color = type)) +
  ggrepel::geom_label_repel(
    data = res |> filter(type == "river"), aes(x = long, y = lat, label = group)) +
  scale_color_brewer(palette = "Dark2")  + 
  ggthemes::theme_map() +
  ggplot2::theme(legend.position = "bottom",
                 legend.text = element_text(size = 15),
                 legend.title = element_text(size = 15)) +
  ggplot2::labs(x = "Longitude", y = "Latitude") + 
  guides(color = guide_legend(override.aes = list(size=5)))

res_long <- res |>
  face_temporal(ts) |>
  unfold(group, type) |>
  rename(prcp = matched_var) |> 
  mutate(prcp = (prcp - min(prcp, na.rm = TRUE))/ (max(prcp, na.rm = TRUE) - min(prcp, na.rm = TRUE)))

p2 <- res_long |> 
  ggplot(aes(x = date, y = prcp, color = type, group = id)) +
  geom_line(alpha = 0.8) +
  facet_wrap(vars(group)) +
  scale_color_brewer(palette = "Dark2", guide = "none") +
  theme_bw() +
  scale_x_date(date_labels = "%b") + 
  labs(x = "Date", y = "Precipitation/ water level")

(p1 | p2) + 
  patchwork::plot_layout(guides = "collect") + 
  plot_annotation(tag_levels = 'a')&
  theme(legend.position = "bottom")

```
```{r echo = TRUE}
raw <- ncdf4::nc_open(here::here("data/era5-pressure.nc"))
dt <- as_cubble(
  raw, vars = c("q", "z"),
  long_range = seq(-180, 180, 1), lat_range = seq(-88, 88, 1))

```
```{r eval = FALSE}
## # prepare data/era5-pressure.nc
## library(ecmwfr)
## my_request <- list(
##   product_type = "reanalysis",
##   format = "netcdf",
##   pressure_level = "10",
##   year = "2002",
##   month = c("09", "10"),
##   day = c("04", "22", "26", "30"),
##   time = "00:00",
##   area = c(-5, -180, -90, 180),
##   variable = c("specific_humidity", 'geopotential'),
##   dataset_short_name = "reanalysis-era5-pressure-levels",
##   target = "era5-pressure.nc"
## )
## #
## # # here you need to create an account on Climate Data Store
## # # at https://cds.climate.copernicus.eu/cdsapp#!/home) to obtain the user ID and API key
## # # then set up the with ecmwfr::wf_set_key(user, key, service = "cds")
## # # the user below is the same `user` in the setup line
## wf_request(
##   user = "113339",
##   request = my_request,
##   transfer = TRUE,
##   path = here::here("data/"))

```
```{r netcdf, out.width="100%", fig.height = 4, fig.width = 10, fig.cap = "A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020) to illustrate the break-up of sourthern polar vortex in late September and early October 2002. The polar vortex, signalled by the high speicfic humidity, splits into two on 2002-09-26 and furthers split into four on 2002-10-04."}
date <- c("2002-09-22", "2002-09-26", "2002-09-30", "2002-10-04") |> as.Date()
res <- dt |> 
  face_temporal() |> 
  filter(lubridate::date(time) %in% date) |>
  unfold(long, lat) |> 
  mutate(q =  q* 10^6)

box = c(xmin = -180, ymin = -90, xmax = 180, ymax = -5)
sf::sf_use_s2(FALSE)
country <- rnaturalearth::ne_coastline("small", returnclass = "sf") |> 
  sf::st_crop(box) |> 
  sf::st_cast("MULTILINESTRING")

res |> 
  ggplot() +
  # q for specific humidity
  geom_point(aes(x = long, y = lat, color = q), size = 2) + 
  #geom_tile(aes(x = long, y = lat, fill = q), color = "transparent") +
  # z for geopotential
  geom_contour(data = res, aes(x = long, y = lat, z = z),
               color = "grey20", binwidth = 4000, size = 0.5) +
  geom_sf(data = country , alpha = 0.5, fill = "transparent", color = "lightgreen") +
  coord_sf(
    crs = "+proj=stere +lat_0=-90 +lon_0=-180 +k=1 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs ", default_crs = sf::st_crs(4326), label_graticule = "EW") + 
  theme_bw() + 
  theme(legend.position = "bottom",
        axis.ticks = element_blank(),
        panel.border = element_blank(),
        axis.text = element_blank(), 
        panel.grid = element_blank()) + 
  facet_wrap(vars(as.Date(time)), nrow = 1, shrink = TRUE)+  
  colorspace::scale_color_continuous_sequential("Purple-Yellow", name = "Specific humidity") + 
  theme(legend.text = element_text(size = 13),
        legend.title = element_text(size = 13)) + 
  labs(x ="", y = "")

```
```{r eval = FALSE}
## # This chunk generates the interactive graphic and save it as figures/linking.html, where the three screencasts in the figure comes from.
## library("crosstalk")
## library("plotly")
## 
## clean <- climate_full |>
##   filter(!id %in% c("ASN00067033", "ASN00072091", "ASN00059040",
##                     "ASN00097053", "ASN00040856", "ASN00015548")) |>
##   face_temporal() |>
##   mutate(month = lubridate::month(date)) |>
##   group_by(month) |>
##   summarise(
##     tmax = mean(tmax, na.rm = TRUE),
##     tmin = mean(tmin, na.rm = TRUE),
##     diff = mean(tmax - tmin, na.rm = TRUE)
##     ) |>
##   ungroup(month) |>
##   mutate(dummy_date = as.Date(glue::glue("2021-{month}-01"))) |>
##   face_spatial() |>
##   mutate(temp_diff_var = ts |> pull(diff) |> var(na.rm = TRUE))
## 
## nested <- clean |>
##   SharedData$new(~id, group = "cubble")
## 
## long <- clean |>
##   face_temporal() |>
##   unfold(temp_diff_var) |>
##   arrange(temp_diff_var) |>
##   SharedData$new(~id, group = "cubble")
## 
## state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)
## 
## p1 <- nested |>
##   ggplot() +
##   geom_sf(data = state_map, fill = "grey95", color = "white") +
##   geom_point(aes(x = long, y = lat,color = temp_diff_var, label = name)) +
##   colorspace::scale_color_continuous_sequential(
##     "Rocket",  n_interp = 7, cmax = 90, rev = TRUE,
##     c2 = 40, l2= 85, c1 = 20, l1 = 30, name = "Var. temp. diff.") +
##   labs(x = "Longitude", y = "Latitude") +
##   coord_sf() +
##   theme_bw() +
##   theme(
##     panel.grid.major = element_blank(),
##     legend.position = "none"
##     )
## 
## p2 <- long |>
##   ggplot(aes(x = dummy_date, group = id,
##              fill = temp_diff_var, color = temp_diff_var)) +
##   geom_ribbon(aes(ymin = tmin, ymax = tmax), size = 0.1, alpha = 0.3) +
##   geom_line(aes(y = tmin), alpha = 0.1) +
##   geom_line(aes(y = tmax), alpha = 0.1) +
##   geom_point(aes(y = tmax), size = 0.1) +
##   geom_point(aes(y = tmin), size = 0.1) +
##   colorspace::scale_fill_continuous_sequential(
##     "Rocket",  n_interp = 7, cmax = 90, rev = TRUE,
##     c2 = 40, l2= 85, c1 = 20, l1 = 30, name = "Var. temp. diff.") +
##   colorspace::scale_colour_continuous_sequential(
##     "Rocket",  n_interp = 7, cmax = 90, rev = TRUE,
##     c2 = 40, l2= 85, c1 = 20, l1 = 30, name = "Var. temp. diff.") +
##   scale_x_date(date_labels = "%b", date_breaks = "1 month") +
##   labs(x = "Month", y = "Temperature") +
##   theme_bw() +
##   theme(
##     panel.grid.major = element_blank(),
##     legend.position = "bottom"
##     )
## 
## out <- bscols(
##   ggplotly(p1, width = 900, height = 700, tooltip = "label") |>
##     highlight(
##       on = "plotly_selected",
##       off = "plotly_deselect",
##       opacityDim = 0.1),
##   ggplotly(p2, width = 1200, height = 700, tooltip = "label") |>
##     highlight(
##       on = "plotly_selected",
##       off = "plotly_deselect",
##       opacityDim = 0.012),
##   widths = c(5, 5)
## )
## 
## htmltools::save_html(out, file = "figures/linking.html")
##

library("leaflet")
library("leafpop")

```
```{r echo = FALSE}
load(here::here("data/station_nested.rda"))
load(here::here("data/climate_full.rda"))

```
```{r eval = FALSE}
## # The kmeans clustering in station_nested data can takes some time to run,
## # This chunk runs the algorithm and save the result as staiton_nested for further use
## coords <- cbind(prcp_aus$long, prcp_aus$lat)
## dist_raw <- geosphere::distm(coords, coords)
## 
## station_long <- cubble::prcp_aus |>
##   face_temporal(ts) |>
##   mutate(wk = lubridate::week(date)) |>
##   group_by(wk) |>
##   summarise(prcp = sum(prcp, na.rm = TRUE)) |>
##   unfold(cluster)
## 
## set.seed(123)
## station_nested <- station_long |>
##   face_spatial() |>
##   strip_rowwise() |>
##   mutate(cluster = kmeans(dist_raw,centers = 20, nstart = 500)$cluster)
## save(station_nested, file = here::here("data/station_nested.rda"))

```
```{r basic-agg, fig.height = 7, fig.width = 15, fig.cap="Profile of aggregated precipitation at 639 weather stations in Australia. Subplot A shows the glyph map of the weekly averaged precipitation of each cluster. The group number is printed in the middle of the y minor axis and can be used as a reference line to read the magnitude. Subplot B shows the station membership of each cluster."}
load(here::here("data/station_nested.rda"))
cluster_nested <- station_nested |>
  switch_key(cluster) |> 
  get_centroid()

cluster_long <- cluster_nested |> 
  face_temporal() |> 
  group_by(wk) |>
  summarise(prcp = mean(prcp, na.rm = TRUE)) |> 
  unfold(cent_long, cent_lat)

state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)

ggplot_smooth <- cluster_long |> 
  ggplot() +
  geom_smooth(aes(x = wk, y = prcp, group = cluster), span = 0.4)

smoother <- layer_data(ggplot_smooth) |> 
  left_join(cluster_long |> select(cluster, cent_long, cent_lat), by = c("group" = "cluster"))

p1 <- ggplot(data = smoother, 
             aes(x_minor = x, y_minor = y, x_major = cent_long, y_major = cent_lat)) + 
  geom_sf(data = state_map, inherit.aes = FALSE,
          color = "grey80", alpha = 0.4, linetype = 3) + 
  geom_text(data = cluster_nested, 
            aes(x = cent_long, y = cent_lat, label = cluster), inherit.aes = FALSE) +
  geom_glyph(height = 2, width = 4) + 
  theme_void()

p2 <- ggplot() + 
  geom_sf(data = state_map, inherit.aes = FALSE,
          color = "grey80", alpha = 0.4, linetype = 3) + 
  geom_point(data = station_nested, aes(x = long, y = lat), size = 0.5) +
  ggforce::geom_mark_hull(data = cluster_nested |> tidyr::unnest(hull),
                          expand = 0, radius = 0,
                          aes(x = long, y = lat, group = cluster)) + 
  theme_void()

(p1 | p2) + plot_annotation(tag_levels = 'A')

```
```{r eval = FALSE }
## clean <-  climate_full |>
##   filter(!id %in% c("ASN00067033", "ASN00072091", "ASN00059040",
##                     "ASN00097053", "ASN00040856", "ASN00015548")) |>
##   face_temporal() |>
##   mutate(month = lubridate::month(date)) |>
##   group_by(month) |>
##   summarise(tmax = mean(tmax, na.rm = TRUE),
##             tmin = mean(tmin, na.rm = TRUE),
##             diff = mean(tmax - tmin, na.rm = TRUE)) |>
##   ungroup(month) |>
##   face_spatial() |>
##   mutate(temp_diff_var = ts |> pull(diff) |> var(na.rm = TRUE)) |>
##   face_temporal() |>
##   mutate(dummy_date = as.Date(glue::glue("2021-{month}-01"))) |>
##   unfold(name)
## 
## df_id <- clean$id |> unique()
## p <- purrr::map(1:length(df_id), function(i){
##   dt <- clean |> filter(id == df_id[i])
##   dt |>
##     ggplot(aes(x = dummy_date, ymin = tmin, ymax = tmax, group = id)) +
##     geom_ribbon(color = "grey50", fill = "grey50", alpha = 0.5) +
##     theme_bw() +
##     scale_x_date(date_labels = "%b", date_breaks = "1 month") +
##     labs(x = "Month", y = "Temperature") +
##     ylim(-10, 50) +
##     ggtitle(paste0(stringr::str_to_title(unique(dt$name))))
## })
## 
## nested <- face_spatial(clean)
## domain <- nested$temp_diff_var
## pal <- colorNumeric(
##   colorspace::sequential_hcl(
##     "Rocket",  n = 7, cmax = 90, rev = TRUE,
##     c2 = 40, l2= 85, c1 = 20, l1 = 30),
##   domain = domain)
## 
## leaflet(nested) |>
##   addTiles() |>
##   addCircleMarkers(color = ~pal(domain), radius = 0.1,
##                    fillOpacity = 1, opacity = 1,
##                    popup = popupGraph(p, width = 300, height =200))

```
```{r interactive-popup, fig.align='center', echo = FALSE, out.width="45%", out.height="25%", fig.retina = 2, dpi = 300, fig.cap = "Same as Figure 11 except the temperature variation is now shown as a popup in the leaflet map.", fig.show='hold'}
knitr::include_graphics(here::here("figures/popup.png"))

```
```{r eval = FALSE}
## library(tidyverse)
## library(cubble)
## all_stations <- rnoaa::ghcnd_stations() |>
##   filter(str_starts(id, "ASN")) |> # Australian stations start wiht "ASN"
##   filter(last_year >= 2020) |>
##   mutate(wmo_id = as.numeric(wmo_id), name = str_to_lower(name)) |>
##   select(-state, -gsn_flag) |>
##   select(id, longitude, latitude, elevation, name,
##          wmo_id, element, first_year, last_year) |>
##   rename(long = longitude, lat = latitude, elev = elevation)
## 
## tmax_stations <- all_stations |>
##   filter(element == "TMAX", first_year < 1970, !is.na(wmo_id))
## 
## raw_tmax <- all_stations |>
##   rowwise() |>
##   mutate(ts = list(rnoaa::meteo_pull_monitors(
##     monitors = id, var = "TMAX",
##     date_min = glue::glue("{first_year}-01-01"),
##     date_max = glue::glue("{last_year}-12-31")
##     ) |>
##       select(-id)
##     )
##   )
## 
## historical_tmax <- raw_tmax |>
##   select(-element) |>
##   unnest(ts) |>
##   mutate(tmax = tmax/10) |>
##   filter(lubridate::year(date) %in% c(1971: 1975, 2016:2020)) |>
##   as_cubble(index = date, key = id, coords = c(long, lat))
## 
## save(historical_tmax, file = here::here("data/historical_tmax.rda"))

```
```{r eval = FALSE}
## aus_stations <- all_stations |>
##   filter(element %in% c("PRCP", "TMAX", "TMIN")) |>
##   nest(element: last_year) |>
##   rowwise() |>
##   filter(nrow(data) == 3) |>
##   select(-data)
## 
## aus_climate_raw <- aus_stations |>
##   rowwise() |>
##   mutate(ts = list(
##     rnoaa::meteo_pull_monitors(
##       monitors = id, var = c("PRCP", "TMAX", "TMIN"),
##       date_min = "2016-01-01", date_max = "2020-12-31"
##       ) |>
##       select(-id)
##     )
##   )
## 
## climate_full <- aus_climate_raw |>
##   unnest(ts) |>
##   mutate(tmax = tmax/10, tmin = tmin/10) |>
##   cubble::as_cubble(key = id, index = date, coords = c(long, lat))
## 
## save(climate_full, file = here::here("data/climate_full.rda"))
```

### 2022-07-11 code/appendix.R

@@ -1,57 +1,70 @@
-## ----setup, include=FALSE-------------------------------------------
+## ----setup, include=FALSE-----------------------------------------------------
 knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
-library(tidyverse)
-library(cubble)
-library(patchwork)
-library(leaflet)
-library(leafpop)
+library("tidyverse")
+library("cubble")
+library("patchwork")
+library("leaflet")
+library("leafpop")
 
 
-## ---- echo = FALSE--------------------------------------------------
+## ---- echo = FALSE------------------------------------------------------------
 load(here::here("data/station_nested.rda"))
 load(here::here("data/climate_full.rda"))
 
-## ----eval = FALSE-------------------------------------
+
+## ----eval = FALSE, echo = TRUE------------------------------------------------
+## station_nested <- climate_full |> mutate(cluster = ...)
+
+
+## ----eval = FALSE, echo = TRUE------------------------------------------------
+## cluster_nested <- station_nested |> switch_key(cluster)
+
+
+## ----eval = FALSE, echo = TRUE------------------------------------------------
+## cluster_nested <- cluster_nested |> get_centroid()
+
+
+## ----eval = FALSE, echo = FALSE-----------------------------------------------
 ## # The kmeans clustering in station_nested data can takes some time to run,
 ## # This chunk runs the algorithm and save the result as staiton_nested for further use
 ## coords <- cbind(prcp_aus$long, prcp_aus$lat)
 ## dist_raw <- geosphere::distm(coords, coords)
 ## 
-## station_long <- cubble::prcp_aus %>%
-##   face_temporal(ts) %>%
-##   mutate(wk = lubridate::week(date)) %>%
-##   group_by(wk) %>%
-##   summarise(prcp = sum(prcp, na.rm = TRUE)) %>%
+## station_long <- cubble::prcp_aus |>
+##   face_temporal(ts) |>
+##   mutate(wk = lubridate::week(date)) |>
+##   group_by(wk) |>
+##   summarise(prcp = sum(prcp, na.rm = TRUE)) |>
 ##   unfold(cluster)
 ## 
 ## set.seed(123)
-## station_nested <- station_long %>%
-##   face_spatial() %>%
[…]

### 2022-07-11 code/gen-scripts-from-rmd.R

@@ -1,0 +1,7 @@
+knitr::purl(input = here::here("jss/main.Rmd"), output = here::here("code/reproducible-script.R"))
+knitr::purl(input = here::here("jss/appendix.Rmd"), output = here::here("code/appendix.R"))
+# manually merge the two scripts into one
+# remove chunks for illustration, but keep codes for deriving data
+# remove "child = here::here("5-1-covid.Rmd")" and alike
+# change in line 5 to echo = TRUE to show all the codes, also line 446, 467, 509
+knitr::spin(here::here("code/reproducible-script.R"), knit = FALSE)

### 2022-07-11 code/main.R

@@ -1,447 +1,0 @@
-## ----setup, echo = FALSE--------------------------------------------
-knitr::opts_chunk$set(
-  echo = TRUE,
-  warning = FALSE,
-  message = FALSE,
-  cache = TRUE,
-  fig.path = here::here("figures/")
-)
-
-options(prompt = "R> ", continue = "+ ",
-        tibble.print_max = 5, tibble.print_min = 5,
-        width = 70)
-
-
-## ----echo = FALSE---------------------------------------------------
-library(cubble)
-library(tidyverse)
-library(patchwork)
-
-## -------------------------------------------------------------------
-cubble::climate_flat
-
-
-## ----echo = TRUE----------------------------------------------------
-cubble_nested <- climate_flat |>
-  as_cubble(key = id, index = date, coords = c(long, lat))
-cubble_nested
-
-
-## ----echo = TRUE----------------------------------------------------
-cubble_nested |>
-  mutate(rain_day = sum(ts$prcp != 0))
-
-
-## ----echo = TRUE----------------------------------------------------
-cubble_long <- cubble_nested |> face_temporal()
-cubble_long
-
-
-## ----echo = TRUE----------------------------------------------------
-cubble_back <- cubble_long |> face_spatial()
-cubble_back
-identical(cubble_nested, cubble_back)
-
-
-## ----echo = TRUE----------------------------------------------------
-cubble_unfold <- cubble_long |> unfold(long, lat)
-cubble_unfold
-
-
-## ----echo = TRUE----------------------------------------------------
-climate_flat_ts <- climate_flat |> 
-  tsibble::as_tsibble(key = id, index = date)
-climate_flat_cb <-  climate_flat_ts |> 
-  cubble::as_cubble(coords = c(long, lat))
-climate_flat_cb
-
-
-## ----echo = TRUE----------------------------------------------------
-# add station-based features in the nested form.
-climate_flat_cb |> 
-  mutate(
-    fabletools::features(
-      ts, tmax, list(tmax_mean = mean, tmax_var = var)
-      )
-  )
-
-
-## ----eval = FALSE---------------------------------------------------
-## library(tidyverse)
-## #https://www.coronavirus.vic.gov.au/victorian-coronavirus-covid-19-data
[…]

### 2022-07-11 code/reproducible-script.R

@@ -1,0 +1,619 @@
+## ----setup, echo = FALSE------------------------------------------------------
+knitr::opts_chunk$set(
+  echo = TRUE,
+  warning = FALSE,
+  message = FALSE,
+  cache = TRUE,
+  fig.path = here::here("figures/")
+)
+
+options(prompt = "R> ", continue = "+ ",
+        tibble.print_max = 5, tibble.print_min = 5,
+        width = 70)
+
+
+## ----echo = FALSE-------------------------------------------------------------
+library(cubble)
+library(tidyverse)
+library(patchwork)
+
+## -----------------------------------------------------------------------------
+cubble::climate_flat
+
+
+## ----echo = TRUE--------------------------------------------------------------
+cubble_nested <- climate_flat |>
+  as_cubble(key = id, index = date, coords = c(long, lat))
+cubble_nested
+
+
+## ----echo = TRUE--------------------------------------------------------------
+cubble_nested |>
+  mutate(rain_day = sum(ts$prcp != 0))
+
+
+## ----echo = TRUE--------------------------------------------------------------
+cubble_long <- cubble_nested |> face_temporal()
+cubble_long
+
+
+## ----echo = TRUE--------------------------------------------------------------
+cubble_back <- cubble_long |> face_spatial()
+cubble_back
+identical(cubble_nested, cubble_back)
+
+
+## ----echo = TRUE--------------------------------------------------------------
+cubble_unfold <- cubble_long |> unfold(long, lat)
+cubble_unfold
+
+
+## ----echo = TRUE--------------------------------------------------------------
+climate_flat_ts <- climate_flat |> 
+  tsibble::as_tsibble(key = id, index = date)
+climate_flat_cb <-  climate_flat_ts |> 
+  cubble::as_cubble(coords = c(long, lat))
+climate_flat_cb
+
+
+## ----echo = TRUE--------------------------------------------------------------
+climate_flat_cb |> 
+  mutate(
+    fabletools::features(
+      ts, tmax, list(tmax_mean = mean, tmax_var = var)
+      )
+  )
+
+
+## ----eval = FALSE-------------------------------------------------------------
+## library(tidyverse)
+## #https://www.coronavirus.vic.gov.au/victorian-coronavirus-covid-19-data
[…]

### 2022-07-20 4-other-features.Rmd
minor fix in the reference

@@
 
-The workflow with the \code{cubble} class fits works well with an interactive graphics pipeline (e.g., @buja1988elements, @buja1996interactive,  @sutherland2000orca, @xie2014reactive, @cheng2016enabling) that is available in \proglang{R} with the package \pkg{crosstalk} [@crosstalk].
+The workflow with the \code{cubble} class fits works well with an interactive graphics pipeline (e.g., @buja1988elements; @buja1996interactive;  @sutherland2000orca; @xie2014reactive; @cheng2016enabling) that is available in \proglang{R} with the package \pkg{crosstalk} [@crosstalk].
 Figure \ref{fig:illu-interactive} illustrates how linking can be achieved between a map and multiple time series in a \code{cubble} object.

### 2022-07-20 5-4-era5.Rmd
minor fix in the reference

@@
 
-Figure \ref{fig:netcdf} reproduces the ERA5 data row of Figure 19 in [@hersbach2020era5].
+Figure \ref{fig:netcdf} reproduces the ERA5 data row of Figure 19 in @hersbach2020era5.
 Here we explain how this would be done using in the \pkg{cubble} package.
 The plots show that the southern polar vortex splits into two on 2002-09-26 and further splits into four on 2002-10-04.
-Further explanation of why this is interesting can be found in the figure source, and also [@simmons2020global], [@simmons2005ecmwf].
+Further explanation of why this is interesting can be found in the figure source, and also in @simmons2020global and @simmons2005ecmwf.
 

### 2022-07-20 jss/main.tex
generate final submitted article
minor fix in the reference

@@
 \documentclass[
-]{jss}
+  shortnames]{jss}
 
@@
 
-\newpage
-
-The \texttt{rticles::jss\_article} template used during preparation (\url{https://github.com/rstudio/rticles/tree/c085e125d874c11bb474ec53f4b754e179f8a4e9}) has the behavior that a reference prints out all the author names when it is used for the first time.
-See a standalone example \href{https://github.com/huizezhang-sherry/jss}{here}.
-
-Here a hack is used to overcome this temporarily
-
-Here I reference \citet{ctvspatiotemporal}, \citet{bach_review_2014}, \citet{lu_multidimensional_2018}, \citet{tidyverse}, \citet{tibble}, \citet{sf}, \citet{stars}, \citet{spacetime}, \citet{dplyr}, \citet{ggplot2}, \citet{tsibble}, \citet{hassell_data_2017}, \citet{tidydata}, \citet{leaflet}, \citet{crosstalk}, \citet{cheng2016enabling}, \citet{buja1996interactive}, \citet{xie2014reactive}, \citet{sutherland2000orca}, \citet{buja1988elements}, \citet{Wickham2012-yr}, \citet{hersbach2020era5}, \citet{tidync}, \citet{rnetcdf}, \citet{ncdf4}, \citet{ecwmfr}, \citet{simmons2020global}, \citet{simmons2005ecmwf}, \citet{mcintosh2018using}, \citet{stuart2010matching}, \citet{castanedo2013review}, \citet{cocchi2019data}, \citet{knitr}, \citet{rmarkdown}, \citet{wang2020calendar}, \citet{advr}, \citet{CS07Author}=``Cook, D.
-and Swayne, D.'',, \citet{bom}, \citet{noaa}, \citet{cds}, \citet{wickham_tourr_2011}, \citet{rnoaa}, \citet{ggally}
-
-\newpage
-
 \hypertarget{introduction}{%
@@
 R> climate_flat_cb |> 
-+   mutate(
-+     fabletools::features(
++   mutate(fabletools::features(
 +     ts, tmax, list(tmax_mean = mean, tmax_var = var)
-+       )
-+   )
++     ))
 \end{CodeInput}
@@
 
-The workflow with the \code{cubble} class fits works well with an interactive graphics pipeline (e.g., \citet{buja1988elements}, \citet{buja1996interactive}, \citet{sutherland2000orca}, \citet{xie2014reactive}, \citet{cheng2016enabling}) that is available in \proglang{R} with the package \pkg{crosstalk} \citep{crosstalk}.
+The workflow with the \code{cubble} class fits works well with an interactive graphics pipeline (e.g., \citet{buja1988elements}; \citet{buja1996interactive}; \citet{sutherland2000orca}; \citet{xie2014reactive}; \citet{cheng2016enabling}) that is available in \proglang{R} with the package \pkg{crosstalk} \citep{crosstalk}.
 Figure \ref{fig:illu-interactive} illustrates how linking can be achieved between a map and multiple time series in a \code{cubble} object.
@@
 
-Figure \ref{fig:netcdf} reproduces the ERA5 data row of Figure 19 in \citep{hersbach2020era5}.
+Figure \ref{fig:netcdf} reproduces the ERA5 data row of Figure 19 in \citet{hersbach2020era5}.
 Here we explain how this would be done using in the \pkg{cubble} package.
 The plots show that the southern polar vortex splits into two on 2002-09-26 and further splits into four on 2002-10-04.
-Further explanation of why this is interesting can be found in the figure source, and also \citep{simmons2020global}, \citep{simmons2005ecmwf}.
+Further explanation of why this is interesting can be found in the figure source, and also in \citet{simmons2020global} and \citet{simmons2005ecmwf}.
 
@@
 \begin{Highlighting}[]
-\NormalTok{clean }\OtherTok{\textless{}{-}}\NormalTok{ climate\_full }\SpecialCharTok{|}\ErrorTok{\textgreater{}}\NormalTok{ ...}
-
-\NormalTok{nested }\OtherTok{\textless{}{-}}\NormalTok{ clean }\SpecialCharTok{|}\ErrorTok{\textgreater{}}\NormalTok{ SharedData}\SpecialCharTok{$}\FunctionTok{new}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{id, }\AttributeTok{group =} \StringTok{"cubble"}\NormalTok{)}
-\NormalTok{long }\OtherTok{\textless{}{-}} \FunctionTok{face\_temporal}\NormalTok{(clean) }\SpecialCharTok{|}\ErrorTok{\textgreater{}}\NormalTok{ SharedData}\SpecialCharTok{$}\FunctionTok{new}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{id, }\AttributeTok{group =} \StringTok{"cubble"}\NormalTok{)}
-
-\NormalTok{p1 }\OtherTok{\textless{}{-}}\NormalTok{ nested }\SpecialCharTok{|}\ErrorTok{\textgreater{}}\NormalTok{ ...}
-\NormalTok{p2 }\OtherTok{\textless{}{-}}\NormalTok{ long }\SpecialCharTok{|}\ErrorTok{\textgreater{}}\NormalTok{ ...}
+\NormalTok{clean }\OtherTok{\textless{}{-}}\NormalTok{ climate\_full }\SpecialCharTok{|\textgreater{}}\NormalTok{ ...}
+
+\NormalTok{nested }\OtherTok{\textless{}{-}}\NormalTok{ clean }\SpecialCharTok{|\textgreater{}}\NormalTok{ SharedData}\SpecialCharTok{$}\FunctionTok{new}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{id, }\AttributeTok{group =} \StringTok{"cubble"}\NormalTok{)}
+\NormalTok{long }\OtherTok{\textless{}{-}} \FunctionTok{face\_temporal}\NormalTok{(clean) }\SpecialCharTok{|\textgreater{}}\NormalTok{ SharedData}\SpecialCharTok{$}\FunctionTok{new}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{id, }\AttributeTok{group =} \StringTok{"cubble"}\NormalTok{)}
+
+\NormalTok{p1 }\OtherTok{\textless{}{-}}\NormalTok{ nested }\SpecialCharTok{|\textgreater{}}\NormalTok{ ...}
+\NormalTok{p2 }\OtherTok{\textless{}{-}}\NormalTok{ long }\SpecialCharTok{|\textgreater{}}\NormalTok{ ...}
 

### 2022-11-13 3-the-cubble-pkg.Rmd
generate final submitted article
update from submission

@@
 
-```{r face, echo = FALSE, fig.align="center", out.height = "35%", out.width = "100%", fig.cap = "An illustration of function \\code{face\\_temporal} and \\code{face\\_spatial}.
+```{r face, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "An illustration of function \\code{face\\_temporal} and \\code{face\\_spatial}.
 In the first row, \\code{face\\_temporal} switches a \\code{cubble} object from the nested form into the long form and the focus has switched from the spatial aspect (the side face) to the temporal aspect (the front face).
@@
 climate_flat_cb |> 
-  mutate(
-    fabletools::features(
+  mutate(fabletools::features(
     ts, tmax, list(tmax_mean = mean, tmax_var = var)
-      )
-  )
+    ))
 ```
@@
 ```{r echo = TRUE, eval = FALSE}
-raw <- ncdf4::nc_open(here::here("data/era5-pressure.nc"))
-dt <- as_cubble(raw, vars = c("q", "z"))
+dt <- as_cubble(my_ncdf, vars = c("q", "z"))
 ```

### 2022-11-13 5-4-era5.Rmd
update from submission

@@
 For the reproduction, we focus on the \code{era5-pressure} data, hourly pressure levels from 1970 to present, with the *specific humidity* and *geopotential*.
-The downloaded NetCDF data is read into \proglang{R} using the \pkg{ncdf4} package, and converted to a \code{cubble} object:
+The downloaded NetCDF data (`raw`) is read into \proglang{R} using the \pkg{ncdf4} package, and converted to a \code{cubble} object:
 
@@
 Therefore, we filter data corresponding to the Southern Hemisphere and restrict our analysis to four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04.-->
+```{r echo = FALSE, eval = TRUE}
+raw <- ncdf4::nc_open(here::here("data/era5-pressure.nc"))
+```
+
 ```{r echo = TRUE}
-raw <- ncdf4::nc_open(here::here("data/era5-pressure.nc"))
 dt <- as_cubble(
@@
   time = "00:00",
-  area = c(-5, -180, -90, 180),
+  area = c(-15, -180, -90, 180),
   variable = c("specific_humidity", 'geopotential'),
@@
 The polar vortex, signalled by the high speicfic humidity, splits into two on 2002-09-26 and furthers split into four on 2002-10-04."}
+library(sf)
 date <- c("2002-09-22", "2002-09-26", "2002-09-30", "2002-10-04") |> as.Date()
@@
 
-box = c(xmin = -180, ymin = -90, xmax = 180, ymax = -5)
+con <- rnaturalearth::ne_coastline("small", returnclass = "sf")
+box <- st_bbox(c(xmin = -180, ymin = -90, xmax = 180, ymax = -15), crs = st_crs(con)) 
 sf::sf_use_s2(FALSE)
-country <- rnaturalearth::ne_coastline("small", returnclass = "sf") |> 
+country <- con %>% st_geometry() %>% 
   sf::st_crop(box) |> 
@@
   geom_contour(data = res, aes(x = long, y = lat, z = z),
-               color = "grey20", binwidth = 4000, size = 0.5) +
+               color = "grey20", binwidth = 4000, linewidth = 0.5) +
   geom_sf(data = country , alpha = 0.5, fill = "transparent", color = "lightgreen") +
   coord_sf(
-    crs = "+proj=stere +lat_0=-90 +lon_0=-180 +k=1 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs ", default_crs = sf::st_crs(4326), label_graticule = "EW") + 
+    crs = "+proj=stere +lat_0=-90 +lon_0=-180 +k=1 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs ", default_crs = sf::st_crs(4326), clip = "off") +
   theme_bw() +
@@
         panel.grid = element_blank()) +
-  facet_wrap(vars(as.Date(time)), nrow = 1, shrink = TRUE)+  
+  facet_wrap(vars(as.Date(time)), nrow = 1)+  
   colorspace::scale_color_continuous_sequential("Purple-Yellow", name = "Specific humidity") + 

### 2022-11-13 jss/main.Rmd
minor fix in the reference
update from submission

@@
 documentclass: jss
+classoption: shortnames
 author:
@@
         tibble.print_max = 5, tibble.print_min = 5,
-        width = 70)
+        width = 70, styler.cache_root = "styler-perm")
 ```
@@
 library(patchwork)
-```
-
-\newpage
-
-```{r child = here::here("jss/reference-hack.Rmd")}
-
 ```

### 2022-11-13 jss/main.tex
update from submission

@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.35\textheight]{../figures/diagram-keynotes/diagram-keynotes.002}
+{\centering \includegraphics[width=1\linewidth]{../figures/diagram-keynotes/diagram-keynotes.002}
 
@@
 \begin{CodeInput}
-R> raw <- ncdf4::nc_open(here::here("data/era5-pressure.nc"))
-R> dt <- as_cubble(raw, vars = c("q", "z"))
+R> dt <- as_cubble(my_ncdf, vars = c("q", "z"))
 \end{CodeInput}
@@
 For the reproduction, we focus on the \code{era5-pressure} data, hourly pressure levels from 1970 to present, with the \emph{specific humidity} and \emph{geopotential}.
-The downloaded NetCDF data is read into \proglang{R} using the \pkg{ncdf4} package, and converted to a \code{cubble} object:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> raw <- ncdf4::nc_open(here::here("data/era5-pressure.nc"))
+The downloaded NetCDF data (\texttt{raw}) is read into \proglang{R} using the \pkg{ncdf4} package, and converted to a \code{cubble} object:
+
+\begin{CodeChunk}
+\begin{CodeInput}
 R> dt <- as_cubble(

### 2022-11-13 5-2-aus-hist-max-temp.Rmd
more updates

@@
   face_temporal() |> 
-  group_by(month = lubridate::month(date),
+  mutate(month = lubridate::month(date),
          group = as.factor(
@@
            "2016 ~ 2020", "1971 ~ 1975"))) |>
+  group_by(month, group) |> 
   summarise(tmax = mean(tmax, na.rm = TRUE))
@@
   geom_sf(data = nsw_vic, fill = "grey95", 
-          color = "white", size=2) + 
+          color = "white") + 
   geom_glyph(

### 2022-11-13 jss/main.tex
more updates

@@
 +   face_temporal() |> 
-+   group_by(month = lubridate::month(date),
++   mutate(month = lubridate::month(date),
 +          group = as.factor(
@@
 +            "2016 ~ 2020", "1971 ~ 1975"))) |>
++   group_by(month, group) |> 
 +   summarise(tmax = mean(tmax, na.rm = TRUE))
@@
 For the reproduction, we focus on the \code{era5-pressure} data, hourly pressure levels from 1970 to present, with the \emph{specific humidity} and \emph{geopotential}.
-The downloaded NetCDF data (\texttt{raw}) is read into \proglang{R} using the \pkg{ncdf4} package, and converted to a \code{cubble} object:
+The downloaded NetCDF data (\texttt{raw}) is first converted to a \code{cubble} object:
 

### 2023-03-04 3-the-cubble-pkg.Rmd
language edit'

@@
 The long \code{cubble} is a subclass of grouped \code{tibble} (\code{grouped_df}), which expands the temporal information into the long form and stores the spatial information in a "spatial" attribute.
+The main functions in the package are \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and \code{unfold()}.
 
@@
 -->
-
-The main functions in the package are \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and \code{unfold()}.
-The following sections explain their roles, why the new \code{cubble} class is needed and how the package relates to existing packages for spatial and temporal data analysis.
 
@@
 The function \code{as_cubble()} is used to create a \code{cubble} object with three arguments: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order (longitude, latitude).
-The arguments \code{key} and \code{index} follow the wording in the \pkg{tsibble} package to describe the temporal order and multiple series while \code{coords} specifies the spatial location of each site.
+The arguments \code{key} and \code{index} follow the conventions in the \pkg{tsibble} package to describe the temporal order and multiple series while \code{coords} specifies the spatial location of each site.
 The code below creates a \code{cubble} object out of \code{climate_flat} (a single \code{tibble}) with \code{id} as the key, \code{date} as the index, and \code{c(long, lat)} as the coordinates:
@@
 
-A \code{cubble} class can be created from various common spatio-temporal data formats, including \code{tibble}, \code{tsibble}, and \code{sf} when both spatial and tmeporal information are available.
+A \code{cubble} class can be created from various common spatio-temporal data formats, including \code{tibble}, \code{tsibble}, and \code{sf} when both spatial and temporal information are available.
 Section \@ref(covid) describes converting multiple tables into a \code{cubble} object and Section \@ref(netcdf) illustrates how to convert a NetCDF object.

### 2023-03-04 4-other-features.Rmd
language edit'

@@
 
-The workflow with the \code{cubble} class fits works well with an interactive graphics pipeline (e.g., @buja1988elements; @buja1996interactive;  @sutherland2000orca; @xie2014reactive; @cheng2016enabling) that is available in \proglang{R} with the package \pkg{crosstalk} [@crosstalk].
+The workflow with the \code{cubble} class works well with an interactive graphics pipeline (e.g., @buja1988elements; @buja1996interactive;  @sutherland2000orca; @xie2014reactive; @cheng2016enabling) that is available in \proglang{R} with the package \pkg{crosstalk} [@crosstalk].
 Figure \ref{fig:illu-interactive} illustrates how linking can be achieved between a map and multiple time series in a \code{cubble} object.
@@
 Linking is also available starting from the time series plot, by selecting points.
-This will be activate rows having the same id in the long \code{cubble} The corresponding rows in the nested \code{cubble} are activated, and highlighted the map.
+This will activate rows having the same id in the long \code{cubble}.
+The corresponding rows in the nested \code{cubble} are activated, and highlighted in the map.
 (An illustration can be found in the appendix.) Note that this type of linking, both from the map or the time series, is what @CS07 would call categorical variable linking, where station id is the categorical variable.
@@
 Spatio-temporal data lends itself to a range of transformations.
-Glyph maps (Section \ref{unfold}) transform the measured variable and time coordinates into microplots at the spatial locations.
+Glyph maps [@Wickham2012-yr] transform the measured variable and time coordinates into microplots at the spatial locations.
 Calendar plots [@wang2020calendar] deconstruct time to produce plots of variables in a calendar format.

### 2023-03-04 5-5-interactive-graphics.Rmd
language edit'

@@
 
-Interactive graphics can be useful because they make it possible to look at the data in a multiple of ways on-the-fly.
+Interactive graphics can be useful because they make it possible to look at the data in multiple ways on-the-fly.
 This is especially important for spatio-temporal data, where we would like to interactively connect spatial and temporal displays.
@@
 It was surprising to us that this was not a station in Tasmania, so for comparison a station in Tasmania is selected on the map to show in relation to Thredbo.
-We can see that Thredbo has a bigger winter dip in temperature, and although Tasmania is cold generally, it's temperatures are more consistent.
+We can see that Thredbo has a bigger winter dip in temperature, and although Tasmania is cold generally, it's temperatures are more constant
 

### 2023-03-04 code/reproducible-script.Rmd
language edit'

@@
 ```{r netcdf, out.width="100%", fig.height = 4, fig.width = 10, fig.cap = "A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020) to illustrate the break-up of sourthern polar vortex in late September and early October 2002.
-The polar vortex, signalled by the high speicfic humidity, splits into two on 2002-09-26 and furthers split into four on 2002-10-04."}
+The polar vortex, signalled by the high specific humidity, splits into two on 2002-09-26 and further splits into four on 2002-10-04."}
 date <- c("2002-09-22", "2002-09-26", "2002-09-30", "2002-10-04") |> as.Date()

### 2023-03-04 jss/main.Rmd
language edit'

@@
 However for some studies, it is important to analyze different aspects of the spatio-temporal data simultaneously, like for instance, temporal trends of multiple variables across locations.
-In order to facilitate the study of different portions or combinations of spatio-temporal data, we introduce a new class, \code{cubble}, with a suite of functions enabling easy slicing and dicing on the different components spatio-temporal components.
+In order to facilitate the study of different portions or combinations of spatio-temporal data, we introduce a new class, \code{cubble}, with a suite of functions enabling easy slicing and dicing on different spatio-temporal components.
 The proposed \code{cubble} class ensures that all the components of the data are easy to access and manipulate while providing flexibility for data analysis.

### 2023-03-04 jss/main.tex
language edit'

@@
 The long \code{cubble} is a subclass of grouped \code{tibble} (\code{grouped_df}), which expands the temporal information into the long form and stores the spatial information in a ``spatial'' attribute.
-
 The main functions in the package are \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and \code{unfold()}.
-The following sections explain their roles, why the new \code{cubble} class is needed and how the package relates to existing packages for spatial and temporal data analysis.
 
@@
 
-A \code{cubble} class can be created from various common spatio-temporal data formats, including \code{tibble}, \code{tsibble}, and \code{sf} when both spatial and tmeporal information are available.
+A \code{cubble} class can be created from various common spatio-temporal data formats, including \code{tibble}, \code{tsibble}, and \code{sf} when both spatial and temporal information are available.
 Section \ref{covid} describes converting multiple tables into a \code{cubble} object and Section \ref{netcdf} illustrates how to convert a NetCDF object.
@@
 
-The workflow with the \code{cubble} class fits works well with an interactive graphics pipeline (e.g., \citet{buja1988elements}; \citet{buja1996interactive}; \citet{sutherland2000orca}; \citet{xie2014reactive}; \citet{cheng2016enabling}) that is available in \proglang{R} with the package \pkg{crosstalk} \citep{crosstalk}.
+The workflow with the \code{cubble} class works well with an interactive graphics pipeline (e.g., \citet{buja1988elements}; \citet{buja1996interactive}; \citet{sutherland2000orca}; \citet{xie2014reactive}; \citet{cheng2016enabling}) that is available in \proglang{R} with the package \pkg{crosstalk} \citep{crosstalk}.
 Figure \ref{fig:illu-interactive} illustrates how linking can be achieved between a map and multiple time series in a \code{cubble} object.
@@
 Linking is also available starting from the time series plot, by selecting points.
-This will be activate rows having the same id in the long \code{cubble} The corresponding rows in the nested \code{cubble} are activated, and highlighted the map.
+This will activate rows having the same id in the long \code{cubble}.
+The corresponding rows in the nested \code{cubble} are activated, and highlighted in the map.
 (An illustration can be found in the appendix.) Note that this type of linking, both from the map or the time series, is what \citet{CS07} would call categorical variable linking, where station id is the categorical variable.
@@
 Spatio-temporal data lends itself to a range of transformations.
-Glyph maps (Section \ref{unfold}) transform the measured variable and time coordinates into microplots at the spatial locations.
+Glyph maps \citep{Wickham2012-yr} transform the measured variable and time coordinates into microplots at the spatial locations.
 Calendar plots \citep{wang2020calendar} deconstruct time to produce plots of variables in a calendar format.
@@
 To fix this, one could aggregate data from nearby stations.
-An example of this using is included in the Appendix.
+An example of this is included in the Appendix.
 
@@
 Here we examine the river gauge data (\code{Water_course_level}) from the Bureau of Meteorology \citep{bom} in relation to weather station rainfall from NOAA's climate data (\code{climate}).
-The goal is to match water gauges with nearby weather stations, spatially and temporally, using 2020 measurements using \code{match_sites()} function.
+The goal is to match water gauges with nearby weather stations, spatially and temporally, using the \code{match_sites()} function.
 
@@
 \caption[A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020) to illustrate the break-up of sourthern polar vortex in late September and early October 2002]{A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020) to illustrate the break-up of sourthern polar vortex in late September and early October 2002.
-The polar vortex, signalled by the high speicfic humidity, splits into two on 2002-09-26 and furthers split into four on 2002-10-04.}\label{fig:netcdf}
+The polar vortex, signalled by the high specific humidity, splits into two on 2002-09-26 and further splits into four on 2002-10-04.}\label{fig:netcdf}
 \end{figure}

### 2023-03-04 code/reproducible-script.R

@@ -310,7 +310,7 @@ dt <- as_cubble(
 ##   path = here::here("data/"))
 
 
-## ----netcdf, out.width="100%", fig.height = 4, fig.width = 10, fig.cap = "A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020) to illustrate the break-up of sourthern polar vortex in late September and early October 2002. The polar vortex, signalled by the high speicfic humidity, splits into two on 2002-09-26 and furthers split into four on 2002-10-04."----
+## ----netcdf, out.width="100%", fig.height = 4, fig.width = 10, fig.cap = "A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020) to illustrate the break-up of sourthern polar vortex in late September and early October 2002. The polar vortex, signalled by the high specific humidity, splits into two on 2002-09-26 and further splits into four on 2002-10-04."----
 date <- c("2002-09-22", "2002-09-26", "2002-09-30", "2002-10-04") |> as.Date()
 res <- dt |> 
   face_temporal() |> 

### 2023-03-04 review/response-B.Rmd
add review and response

---
title: "Response"
author: "H. Sherry Zhang, Dianne Cook, Ursula Laa, Nicolas Langrené, Patricia Menéndez"
date: '2023-03-04'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Response to B-review_1_1

## The Paper

The package and the proposed data structures could certainly constitute a valuable base for the spatio- temporal statistics tooling. Unfortunately the paper does not clearly describe the data structure, nor the associated functionality. I had to read the package vignettes, run the attached code and even look into the package’s code in order to fully understand the data structure.

**Re: **

Given the small size of the package, the paper feels a bit on the heavy side. Most of the code in the paper is not self-contained which makes it rather difficult to follow. Examples at the end of the paper don’t elucidate the inner workings of the package but rather advertise the functionality of other, mostly visualization, packages.

**Re: **

More fundamentally, the design of the package has an immediately apparent drawback. The cubble class is used to represent two incompatible data types - long and nested. Internally, the two data types are distinguished by the internal attribute “form”. In other words, the authors use an attribute of an object to mimic the functionality of a sub-class.

**Re: **

Relatedly, the rationale for the parallel naming convention long/nested and temporal/spatial is not entirely clear. It seems to me that the semantically unambiguous temporal/spatial could be used throughout without ambiguity, thus resolving the terminological redundancy.

**Re: **

My recommendation would be a combination of the following:

  * Rework the definitional parts (sections 2 and 3) by following more closely the “design” vignette. **Re: **
  * Describe at a glance the core functions which operate on the class and their motivation. **Re: **
  * Rework section 4 by making it more concise and provide links to sections in supplementary material or dedicated vignettes. **Re: **
  * Rework examples section 5 into “Applications”. Briefly describe the applications and provide and refer to dedicated and self-contained vignettes. **Re: **

## The Package

The package is generally badly documented. The meaning of the arguments is almost never clear from the documentation alone. For example the documentation of as_cubble states:

* What objects are supported as input data?
* What is the assumption of the “nestedness” of the input data. 
* What is the accepted data type of key, index and coords?

**Re: **

## Further suggestion

Consider not throwing when face_spacial is applied to the spatial object. This would allow for generic code where the “form” of the input the data is not known or does not matter.

**Re: **

It would be useful to print a few rows of the spatial attribute of the “long” object.

**Re: **

Section 3.6 is missing `tidyr` package which is especially relevant here given its nesting and unnesting operations https://tidyr.tidyverse.org/articles/nest.html

**Re: **

More comments inlined in the pdf.

  - on p4 "The following sections explain their roles, why the new cubble class is needed
and how the package relates to existing packages for spatial and temporal data analysis." The paper is a bit on the verbose side. Such comments could be safely removed. **Re:** sentence removed.

  - on p5 section 3.1, "The arguments key and index follow the wording in the tsibble package to describe the temporal order and multiple series while coords specifies the spatial location of each site.", Unclear what `key` and `index` mean. Isn't key a spatial identifier and index a temporal one? **Re: **

  - on p5 section 3.1, the word "wording" is highlighted in the sentence "The arguments key and index follow the wording in the tsibble package to ...", conventions. **Re:** *wording* changed to *conventions*

  - on p6 section 3.2, 'The cubble class also provides a long form, which expands the ts column and temporarily “hides” the spatial variables.', confusing: `hides` means storing it as an attribute. **Re: **

  - on p6 section 3.2, I think it would be a good idea to clearly describe the two data structures in an earlier section. By this point it's still not clear what those are and what the meaning of key and index is. **Re: **

  - on p6, section 3.2, yet another reason to call this data structure "cubble_temporal" instead of the semantically unobvious "long". **Re: **

  - on p8 section 3.5, Isn't the root reason the memory efficiency? That is, the spatial component cannot be represented as part of the nested tibble without repetition. **Re: **

  - on p8 section 3.6, The missing elephant is tidyr which has nesting and unnesting operations  https://tidyr.tidyverse.org/articles/nest.html. **Re: **

# Response to B-review_2_1

## The Manuscript

The manuscript is in general clearly structured. However, an important reference to the R-package sftime is missing that also deals with the representation of spatiotemporal data. It might be beneficial to provide illustrative examples in the manuscript that clearly compare between a cubble representation and representations of existing packages. Which data sets/structures cannot (or with a greater effort) be represented with existing R-packages such as e.g. stars and sftime?

**Re: **

For spatial data, the coordinate reference system is essential metadata information. In one example of the manuscript, the input data does have information on the CRS but it is not discussed in the manuscript how that is handled within cubble. How would coordinate transformations be handled with cubble?

**Re: **

The use-case of temporal matching based on features of the time series appears an interesting, but also a very specific one. How could this be generalized to a more generic approach?

**Re: **

Additionally, several typos and language issues arise and limit the readability of the manuscript. Some examples are:

- doubled/missing words/none correct sentences: 
  - “*components* spatio-temporal *components*” 
  - “... fits works ...” 
  - “... be activate rows ...” 
  - “... highlighted *in* the ...” 
  - “. . . using 2020 measurements using match_sites() function.” 
  - “An example of this using is included in the Appendix” 
  - “. . . the data in -a- multiple -of- ways on-the-fly” 
  - **Re**: the mentioned examples have been changed to: "... on different spatio-temporal components", "The workflow ... works well with ...", "... will activate rows having ...", "... highlighted in the map", "... using the match_sites() function.", "An example of this is included in the Appendix", and "... the data in multiple ways on-the-fly".

- Surprising references: in Section 4.4 it says “. . . Glyph maps (Section 3.4)”, but Section 3.4 says “. . . glyph maps will be explained in Section 4.4”. **Re:** the reference to Section 3.4 in Section 4.4 has been replaced with the literature (Wickham et al. 2012).

- Should “. . . it’s temperatures are more *consistent*” rather be “. . . it’s temperatures are more *constant*”? **Re: ** changed to "more constant"

- Typos:
  - “. . . spatial and *tmeporal* information are available.” 
  - The polar vortex, signalled by the high *speicfic* humidity, splits into two on 2002-09-26 and *further-s- split_s_* into four on 2002-10-04. 
  - the data in *-a- multiple -of-* ways on-the-fly. 
  - **Re: ** typo corrected to "temporal", "specific", "further splits", and "multiple ways"

## The package

I have been a bit puzzled by the “print” of a nested cubble in its temporal face. The row # temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl] appears at first sight (also in comparison with the print of the spatial face), as if the temporal domain is given by all those variables. To me, a notion such as

`# temporal: date [date]`

`# variables: prcp [dbl], tmax [dbl], tmin [dbl]`

would have been more intuitive. Possibly also including the temporal range (as for the spatial domain its bbox).

**Re: **

In the manuscript and in the reproducible-script.R, a data set `historical_tmax` is introduced. In the package, the corresponding data set seems to be `tmax_hist`.

**Re: **

The manuscript cannot be reproduced, as the reproducible-script.R (part of the submission to JSS) only uses a subset, but does not provide a separate result file that would allow to compare the reproduced outputs with the desired output for the subset.

**Re: **

### 2023-03-25 3-the-cubble-pkg.Rmd
some outlines on comparing between cubble/ sftime/stars

@@
 ```
+
+## Comparison with existing packages
+
+The development on spatio-temporal data structure in R has witnessed great changes from the initial \pkg{spacetime} [@spacetime] package.
+New packages, such as \pkg{stars} [@stars] and \pkg{sftime} [@sftime], have [...].
+
+Below compare the additional benefit of cubble on top of the two packages.
+
+The `stars` package creates a `stars` object from an array and a dimension object.
+In many cases, i.e.
+satellite images, this is convenient, however, there are also many spatio-temporal data are formatted in 2D data frames.
+
+This requires converting the variables into an array format.
+
+Consider the following spatial and temporal data.
+
+The temporal data contains two variables are recorded at two stations (`id`) from 2023 March 21st to 2023 March 25th.
+Notice is that id2 only records from 2023 March 22nd and variable 2 has an `NA` value on  2023 March 23rd at id2.
+
+```{r echo = FALSE}
+library(stars)
+spatial <- tibble(
+  id = c(1,2),
+  long = c(130, 140),
+  lat = c(-38, -44),
+  name = c("station 1", "station 2"))
+
+spatial
+
+temporal <- tibble(
+  id = c(rep(1, each = 5), rep(2, each = 4)),
+  time = c(seq(as.Date("2023-03-21"), as.Date("2023-03-25"), 1),
+           seq(as.Date("2023-03-22"), as.Date("2023-03-25"), 1)),
+  var1 = c(3, 6, 8, 2, 5, 7, 9, 14, 5),
+  var2 = c(1, 3, 4, 2, 6, 7, NA, 5, 8)
+)
+temporal
+```
+
+The following code can be used to create the dimension object:
+
+```{r echo = TRUE}
+spatial_sf <- spatial %>% sf::st_as_sf(coords = c("long", "lat"), crs = 4326) 
+station_vec <- st_geometry(spatial_sf)
+time_vec <- unique(temporal$time)
+var_vec <- spatial_sf$name
+d <- st_dimensions(station = station_vec, time = time_vec)
+d
+```
+
+However, the issue comes when casting the variables into an array of the same dimension:
+
+```{r echo = TRUE, eval = FALSE}
+temporal$var1 %>% matrix(byrow = TRUE, nrow = 2, ncol = 5) %>% array(dim = dim(d))
+```
+
+Due to the missingness on 2023 March 21st for id2, the array reuses the first value of the variable to fill in the dimension (the missingness at (2,1) is not flagged and 3 is reused to fill in the dimension).
+
+This causes an issue when multiple stations with different recording periods present in the data, commonly found in the weather data.
+
+The \pkg{sftime} package doesn't have this issue.
+
+Consider the climate data used in the \pkg{cubble} package, \code{climate_aus}.
+It contains daily precipitation, minimum and maximum temperature for 639 weather stations across Australia in 2020.
+A `sftime` object can be created from the spatial and temporal aspect of the data as follows:
+
+```{r echo = TRUE}
+library(sftime)
+
+climate_spatial <- climate_aus %>% select(id:wmo_id)
+climate_temporal <- climate_aus %>% face_temporal() %>% as_tibble()
+
+data_sftime <- left_join(climate_spatial, climate_temporal,  
+                         by = "id", multiple = "all") %>% 
+  st_as_sftime(coords = c("long", "lat"))
+data_sftime
+```
+
+An issue here is the object size created of `data_sftime`. whichis `r `object.size(data_sftime)/object.size(climate_aus)` times the size of \code{climate_aus}
+
+```{r echo = TRUE}
+object.size(data_sftime)
+object.size(climate_aus)
+```
+
+This is the initial issue raised from this paper where a conventional left join of long temporal data with spatial object can create the large spatio-temporal data by repeating the spatial geometry at every time stamp.

### 2023-03-25 5-2-aus-hist-max-temp.Rmd
language edit'
some outlines on comparing between cubble/ sftime/stars

@@
 tmax <- historical_tmax |>
-  filter(between(stringr::str_sub(id, 7, 8), 46, 90))
+  filter(between(as.numeric(stringr::str_sub(id, 7, 8)), 46, 90))
 ```
@@
 To fix this, one could aggregate data from nearby stations.
-An example of this using is included in the Appendix.
+An example of this is included in the Appendix.

### 2023-03-25 5-3-river-level-rainfall.Rmd
language edit'
some outlines on comparing between cubble/ sftime/stars

@@
 Here we examine the river gauge data (\code{Water_course_level}) from the Bureau of Meteorology [@bom] in relation to weather station rainfall from NOAA's climate data (\code{climate}).
-The goal is to match water gauges with nearby weather stations, spatially and temporally, using 2020 measurements using \code{match_sites()} function.
+The goal is to match water gauges with nearby weather stations, spatially and temporally, using the \code{match_sites()} function.
 <!--The Bureau of Meteorology collects [water data](http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full) from river gauges.
@@
 climate <- climate_full |>
-  filter(between(stringr::str_sub(id, 7, 8), 76, 90)) |>
+  filter(between(as.numeric(stringr::str_sub(id, 7, 8)), 76, 90)) |>
   face_temporal() |>

### 2023-03-25 jss/main.tex
some outlines on comparing between cubble/ sftime/stars

@@
 
+\hypertarget{comparison-with-existing-packages}{%
+\subsection{Comparison with existing packages}\label{comparison-with-existing-packages}}
+
+The development on spatio-temporal data structure in R has witnessed great changes from the initial \pkg{spacetime} \citep{spacetime} package.
+New packages, such as \pkg{stars} \citep{stars} and \pkg{sftime} \citep{sftime}, have {[}\ldots{]}.
+
+Below compare the additional benefit of cubble on top of the two packages.
+
+The \texttt{stars} package creates a \texttt{stars} object from an array and a dimension object.
+In many cases, i.e.~satellite images, this is convenient, however, there are also many spatio-temporal data are formatted in 2D data frames.
+
+This requires converting the variables into an array format.
+
+Consider the following spatial and temporal data.
+
+The temporal data contains two variables are recorded at two stations (\texttt{id}) from 2023 March 21st to 2023 March 25th.
+Notice is that id2 only records from 2023 March 22nd and variable 2 has an \texttt{NA} value on 2023 March 23rd at id2.
+
+\begin{CodeChunk}
+\begin{CodeOutput}
+# A tibble: 2 x 4
+     id  long   lat name     
+  <dbl> <dbl> <dbl> <chr>    
+1     1   130   -38 station 1
+2     2   140   -44 station 2
+\end{CodeOutput}
+\begin{CodeOutput}
+# A tibble: 9 x 4
+     id time        var1  var2
+  <dbl> <date>     <dbl> <dbl>
+1     1 2023-03-21     3     1
+2     1 2023-03-22     6     3
+3     1 2023-03-23     8     4
+4     1 2023-03-24     2     2
+5     1 2023-03-25     5     6
+# ... with 4 more rows
+\end{CodeOutput}
+\end{CodeChunk}
+
+The following code can be used to create the dimension object:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> spatial_sf <- spatial %>% sf::st_as_sf(coords = c("long", "lat"), crs = 4326) 
+R> station_vec <- st_geometry(spatial_sf)
+R> time_vec <- unique(temporal$time)
+R> var_vec <- spatial_sf$name
+R> d <- st_dimensions(station = station_vec, time = time_vec)
+R> d
+\end{CodeInput}
+\begin{CodeOutput}
+        from to     offset  delta refsys point
+station    1  2         NA     NA WGS 84  TRUE
+time       1  5 2023-03-21 1 days   Date FALSE
+                                  values
+station POINT (130 -38), POINT (140 -44)
+time                                NULL
+\end{CodeOutput}
+\end{CodeChunk}
+
+However, the issue comes when casting the variables into an array of the same dimension:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> temporal$var1 %>% matrix(byrow = TRUE, nrow = 2, ncol = 5) %>% array(dim = dim(d))
+\end{CodeInput}
+\end{CodeChunk}
+
+Due to the missingness on 2023 March 21st for id2, the array reuses the first value of the variable to fill in the dimension (the missingness at (2,1) is not flagged and 3 is reused to fill in the dimension).
+
+This causes an issue when multiple stations with different recording periods present in the data, commonly found in the weather data.
+
+The \pkg{sftime} package doesn't have this issue.
+
+Consider the climate data used in the \pkg{cubble} package, \code{climate_aus}.
+It contains daily precipitation, minimum and maximum temperature for 639 weather stations across Australia in 2020.
+A \texttt{sftime} object can be created from the spatial and temporal aspect of the data as follows:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> library(sftime)
+R> 
+R> climate_spatial <- climate_aus %>% select(id:wmo_id)
+R> climate_temporal <- climate_aus %>% face_temporal() %>% as_tibble()
+R> 
+R> data_sftime <- left_join(climate_spatial, climate_temporal,  
++                          by = "id", multiple = "all") %>% 
++   st_as_sftime(coords = c("long", "lat"))
+R> data_sftime
+\end{CodeInput}
+\begin{CodeOutput}
+Spatiotemporal feature collection with 230760 features and 7 fields
+Geometry type: POINT
+Dimension:     XY
+Bounding box:  xmin: 113.5319 ymin: -43.6578 xmax: 153.6358 ymax: -10.0511
+CRS:           NA
+Time column with class: 'Date'.
+Ranging from 2020-01-01 to 2020-12-31.
+# A tibble: 230,760 x 9
+  id           elev name         wmo_id date        prcp  tmax  tmin
+* <chr>       <dbl> <chr>         <dbl> <date>     <dbl> <dbl> <dbl>
+1 ASN00001006   3.8 wyndham aero  95214 2020-01-01   164  38.3  25.3
+2 ASN00001006   3.8 wyndham aero  95214 2020-01-02     0  40.6  30.5
+3 ASN00001006   3.8 wyndham aero  95214 2020-01-03    16  39.7  27.2
+4 ASN00001006   3.8 wyndham aero  95214 2020-01-04     0  38.2  27.3
+5 ASN00001006   3.8 wyndham aero  95214 2020-01-05     2  39.3  26.7
+# ... with 230,755 more rows, and 1 more variable: geometry <POINT>
+\end{CodeOutput}
+\end{CodeChunk}
+
+An issue here is the object size created of \texttt{data\_sftime}. whichis \texttt{r}object.size(data\_sftime)/object.size(climate\_aus)` times the size of \code{climate_aus}
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> object.size(data_sftime)
+\end{CodeInput}
+\begin{CodeOutput}
+114549848 bytes
+\end{CodeOutput}
+\begin{CodeInput}
+R> object.size(climate_aus)
+\end{CodeInput}
+\begin{CodeOutput}
+8641608 bytes
+\end{CodeOutput}
+\end{CodeChunk}
+
+This is the initial issue raised from this paper where a conventional left join of long temporal data with spatial object can create the large spatio-temporal data by repeating the spatial geometry at every time stamp.
+
 \hypertarget{others}{%
@@
 \end{CodeOutput}
-\begin{CodeOutput}
-Joining, by = "lga"
-\end{CodeOutput}
 \end{CodeChunk}
@@
 $paired
-# A tibble: 2 x 2
+# A tibble: 80 x 2
   spatial        temporal      
   <chr>          <chr>         
-1 Kingston (C) (Vic.) Kingston (C)
-2 Latrobe (C) (Vic.)  Latrobe (C)
+1 Alpine (S)     Alpine (S)    
+2 Ararat (RC)    Ararat (RC)   
+3 Ballarat (C)   Ballarat (C)  
+4 Banyule (C)    Banyule (C)   
+5 Bass Coast (S) Bass Coast (S)
+# ... with 75 more rows
 
@@
 R> tmax <- historical_tmax |>
-+   filter(between(stringr::str_sub(id, 7, 8), 46, 90))
++   filter(between(as.numeric(stringr::str_sub(id, 7, 8)), 46, 90))
 \end{CodeInput}
@@
 \begin{CodeOutput}
- [1] 12 12 12 13 19 19 20 24 24 24
+ [1] 24 24 12 24 24 24 24 24 24 24
 \end{CodeOutput}
@@
   <chr>       <chr>     <dbl> <dbl> <chr> <dbl> <int> <list>     <int>
-1 405234      SEVEN CR~ -36.9  146. river  6.15     5 <tibble>      21
+1 230200      MARIBYRN~ -37.7  145. river  6.17     6 <tibble>      19
 2 404207      HOLLAND ~ -36.6  146. river  8.54    10 <tibble>      21
-3 ASN00082042 strathbo~ -36.8  146. clim~  6.15     5 <tibble>      21
-4 ASN00082170 benalla ~ -36.6  146. clim~  8.54    10 <tibble>      21
-5 230200      MARIBYRN~ -37.7  145. river  6.17     6 <tibble>      19
+3 405234      SEVEN CR~ -36.9  146. river  6.15     5 <tibble>      21
+4 406213      CAMPASPE~ -37.0  145. river  1.84     1 <tibble>      18
+5 ASN00082042 strathbo~ -36.8  146. clim~  6.15     5 <tibble>      21
 # ... with 3 more rows

### 2023-05-22 review/response-B.Rmd
fill in response to reviewers

@@
 ---
-title: "Response"
+title: "Response to reviewers article 4780"
+subtitle: "cubble: An R Package for Organizing and Wrangling Multivariate Spatio-temporal Data"
 author: "H. Sherry Zhang, Dianne Cook, Ursula Laa, Nicolas Langrené, Patricia Menéndez"
@@
 
-**Re: **
+**\textcolor{purple}{Re: We thank the reviewer for the support on our proposed data structure and the suggestion to improve the paper and the package documentation.}**
 
@@
 
-**Re: **
+**\textcolor{purple}{Re: We select the code and output relevant to demonstrate the cubble package in the examples while leaving the rest in the github repository, where users can look up for more details: https://github.com/huizezhang-sherry/paper-cubble.
+[on the visualisation packages: xxx]}**
 
@@
 
-**Re: **
+**\textcolor{purple}{Re: [TODO] }**
 
@@
 
-**Re: **
+**\textcolor{purple}{Re: xxx}**
 
@@
 
-  * Rework the definitional parts (sections 2 and 3) by following more closely the “design” vignette. **Re: **
-  * Describe at a glance the core functions which operate on the class and their motivation. **Re: **
-  * Rework section 4 by making it more concise and provide links to sections in supplementary material or dedicated vignettes. **Re: **
-  * Rework examples section 5 into “Applications”. Briefly describe the applications and provide and refer to dedicated and self-contained vignettes. **Re: **
+  * Rework the definitional parts (sections 2 and 3) by following more closely the “design” vignette. **\textcolor{purple}{Re: [TODO: follow the design vignette]}**
+  * Describe at a glance the core functions which operate on the class and their motivation.
+**\textcolor{purple}{Re: Methods available for a cubble object is now printed in xxx, before introducing the main functionality: face\_temporal(), face\_spatial(), and unfold().}**
+  * Rework section 4 by making it more concise and provide links to sections in supplementary material or dedicated vignettes. **\textcolor{purple}{Re: xxx}**
+  * Rework examples section 5 into “Applications”. Briefly describe the applications and provide and refer to dedicated and self-contained vignettes. **\textcolor{purple}{Re: xxx}**
 
@@
 
-**Re: **
+**\textcolor{purple}{Re: [TODO: add documentation on these! ]}**
 
@@
 
-Consider not throwing when face_spacial is applied to the spatial object. This would allow for generic code where the “form” of the input the data is not known or does not matter.
+Consider not throwing an error when face_spatial is applied to the spatial object. This would allow for generic code where the “form” of the input the data is not known or does not matter.
 
-**Re: **
+**\textcolor{purple}{Re: Adopted.
+Now when applying face\_spatial to a nested cubble (a spatial object), the cubble will be printed as it is with a message signalling it is already in the nested form.
+see the reprex below.
+Same for face\_temporal on a long cubble}**
+
+``` r
+library(cubble)
+nested <- climate_flat |>
+  as_cubble(key = id, index = date, coords = c(long, lat))
+class(nested) 
+#> [1] "cubble_df"  "rowwise_df" "tbl_df"     "tbl"        "data.frame"
+face_spatial(nested) 
+#> i The cubble is already in the nested form
+#> # cubble:   id [5]: nested form
+#> # bbox:     [115.9764, -32.9342, 133.5407, -12.4239]
+#> # temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
+#>   id            lat  long  elev name           wmo_id ts                
+#>   <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>            
+#> 1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tibble [366 × 4]>
+#> 2 ASN00010311 -31.9  117. 179   york            94623 <tibble [366 × 4]>
+#> 3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble [366 × 4]>
+#> 4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble [366 × 4]>
+#> 5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble [366 × 4]>
+```
+
+<sup>Created on 2023-05-22 with [reprex v2.0.2](https://reprex.tidyverse.org)</sup>
 
@@
 
-**Re: **
+**\textcolor{purple}{Re: xxx. we could have a function to print }**
 
@@
   - on p4 "The following sections explain their roles, why the new cubble class is needed
-and how the package relates to existing packages for spatial and temporal data analysis." The paper is a bit on the verbose side. Such comments could be safely removed. **Re:** sentence removed.
+and how the package relates to existing packages for spatial and temporal data analysis." The paper is a bit on the verbose side.
+Such comments could be safely removed.
+**\textcolor{purple}{Re: sentence removed.}**
 
@@
 Isn't key a spatial identifier and index a temporal one?
-**Re: **
+**\textcolor{purple}{Re: [TODO] Yes, key is the spatial identifier and index is the temporal one.
+They are now explained in the text as well as referenced to the tsibble package.}**
 
   - on p5 section 3.1, the word "wording" is highlighted in the sentence "The arguments key and index follow the wording in the tsibble package to ...", conventions.
-**Re:** *wording* changed to *conventions*
+**\textcolor{purple}{Re: *wording* changed to *conventions*}**
 
   - on p6 section 3.2, 'The cubble class also provides a long form, which expands the ts column and temporarily “hides” the spatial variables.', confusing: `hides` means storing it as an attribute.
-**Re: **
+**\textcolor{purple}{Re: The sentence is now changed to ..., which expands the ts column and stores the spatial variables as an attribute}**
 
@@
 By this point it's still not clear what those are and what the meaning of key and index is.
-**Re: **
+**\textcolor{purple}{Re: addressed in bullet point "on p5 section 3.1"}**
 
-  - on p6, section 3.2, yet another reason to call this data structure "cubble_temporal" instead of the semantically unobvious "long". **Re: **
+  - on p6, section 3.2, yet another reason to call this data structure "cubble_temporal" instead of the semantically unobvious "long".
+**\textcolor{purple}{Re: see paragraph 4 under section The paper }**
 
-  - on p8 section 3.5, Isn't the root reason the memory efficiency? That is, the spatial component cannot be represented as part of the nested tibble without repetition. **Re: **
+  - on p8 section 3.5, Isn't the root reason the memory efficiency?
+That is, the spatial component cannot be represented as part of the nested tibble without repetition.
+**\textcolor{purple}{Re: [TODO - also compare with sftime] added.}**
 
-  - on p8 section 3.6, The missing elephant is tidyr which has nesting and unnesting operations  https://tidyr.tidyverse.org/articles/nest.html. **Re: **
+  - on p8 section 3.6, The missing elephant is tidyr which has nesting and unnesting operations  https://tidyr.tidyverse.org/articles/nest.html.
+**\textcolor{purple}{Re: addressed under section Further suggestion}**
 
@@
 
-**Re: **
+**\textcolor{purple}{Re: [TODO] The section 3.5 are extended to also discuss why not just use existing data structure.
+This includes the comparision with sftime on memory efficiency, with stars on representing spatio-temporal data in an array format with illustrative examples}**
 
@@
 
-**Re: **
+**\textcolor{purple}{Re: We understand spatial data is, strictly speaking, not uniquely identified if CRS information is not provided, while many wild data still do not contain this crucial piece of information.
+In cubble, if the spatial data is from an sf object, the cubble will retain the sf class and users are welcomed to use functions in sf to perform coordinate transformations.
+If not, the function cubble::add\_geometry\_column() can be used to construct the geometry column from long and lat, and the resulting cubble will have the sf class.
+if CRS is missing, cubble::add\_geometry\_column() will by default assume OGC:CRS84 (WGS84).}**
 
@@
 - doubled/missing words/none correct sentences: 
-  - “*components* spatio-temporal *components*” 
-  - “... fits works ...” 
-  - “... be activate rows ...” 
-  - “... highlighted *in* the ...” 
-  - “. . . using 2020 measurements using match_sites() function.” 
-  - “An example of this using is included in the Appendix” 
-  - “. . . the data in -a- multiple -of- ways on-the-fly” 
-  - **Re**: the mentioned examples have been changed to: "...
-on different spatio-temporal components", "The workflow ...
-works well with ...", "...
-will activate rows having ...", "...
-highlighted in the map", "...
-using the match_sites() function.", "An example of this is included in the Appendix", and "...
-the data in multiple ways on-the-fly".
+  - “*components* spatio-temporal *components*” **\textcolor{purple}{Re: to "... spatio-temporal components"}**
+  - “... fits works ...”  **\textcolor{purple}{Re: to ".. works well with ..."}**
+  - “... be activate rows ...” **\textcolor{purple}{Re: to "... will activate rows ..."}**
+  - “... highlighted *in* the ...” **\textcolor{purple}{Re: to "... highlighted on the map"}**
+  - “. . . using 2020 measurements using match_sites() function.”  **\textcolor{purple}{Re: to "... using the match\_sites() function"}**
+  - “An example of this using is included in the Appendix” **\textcolor{purple}{Re: to "... this use ..."}**
+  - “. . . the data in -a- multiple -of- ways on-the-fly”  **\textcolor{purple}{Re: to "... in multiple ways ..."}**
 
@@
 glyph maps will be explained in Section 4.4”.
-**Re:** the reference to Section 3.4 in Section 4.4 has been replaced with the literature (Wickham et al.
-2012).
+**\textcolor{purple}{Re: the looped reference in section 3.3 and 3.4 is replaced with (Wickham et al.
+2012)}.**
 
-- Should “. . . it’s temperatures are more *consistent*” rather be “. . . it’s temperatures are more *constant*”? **Re: ** changed to "more constant"
+- Should “. . . it’s temperatures are more *consistent*” rather be “. . . it’s temperatures are more *constant*”? **\textcolor{purple}{Re: to "more constant"}**
 
 - Typos:
-  - “. . . spatial and *tmeporal* information are available.” 
+  - “. . . spatial and *tmeporal* information are available.” **\textcolor{purple}{Re: to "temporal"}**
   - The polar vortex, signalled by the high *speicfic* humidity, splits into two on 2002-09-26 and *further-s- split_s_* into four on 2002-10-04.
-  - the data in *-a- multiple -of-* ways on-the-fly. 
-  - **Re: ** typo corrected to "temporal", "specific", "further splits", and "multiple ways"
+**\textcolor{purple}{Re: to "specific", "further splits"}**
+  - the data in *-a- multiple -of-* ways on-the-fly. **\textcolor{purple}{Re:  to "multiple ways"}**
 
@@
 
-**Re: **
+**\textcolor{purple}{Re: [TODO] use tsibble functions to print temporal range in the long form, something looks like:}**
+
+```r
+## nested:
+
+# cubble:   key: id [5], index: date, nested form
+# extent:   [115.9764, -32.9342, 133.5407, -12.4239]
+# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
+
+## long:
+# cubble:   key: id [5], index: date, nested form
+# extent:   from/to/by
+# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]
+```
 
@@
 
-**Re: **
+**\textcolor{purple}{Re: [TODO] The two datasets are related but not the same.
+The data historical\_tmax contains daily weather stations data for the two periods: 1971-1975 and 2016-2020 and amounts to 3.2MB, which is not suitable to include in the package given the size restriction.
+The tmax\_hist summarises the daily data into monthly to produce the dataset suitable to include in the package.
+The documentation to tmax\_hist now include more details on this.
+}**
 
@@
 
-**Re: **
+**\textcolor{purple}{Re: A result file in html has been submitted this time.}**
+
+section cubble object under 3. the cubble package, on 1) the conceptual framework, 2)  the nested and long form as per the design vig
+
+print of methods
+
+section 4/5: work on the text to make it more concise
+
+connnecting dat ais gneerally a difficult task from different sources in spatial temporal data, point it to the future work
+
+In the end talk to the users to check the vignette for the full code for examples. match titles.

### 2023-05-22 1-intro.Rmd
combine section 2 and 3; rearrange section 3 with bullet points

@@
 
-The rest of the paper is organized as follows: Section \@ref(cube) introduces the proposed cube structure as a way to conceptualize multivariate spatio-temporal data.
-Section \@ref(cubble) presents the main design and functionality of the \pkg{cubble} package.
+The rest of the paper is organized as follows: Section \@ref(cubble) presents the main design and functionality of the \pkg{cubble} package.
 Section \@ref(others) explains how the \pkg{cubble} package deals with more advanced considerations, including data with hierarchical structure, data matching and how the package fits with existing static and interactive visualization tools.

### 2023-05-22 3-the-cubble-pkg.Rmd
fill in response to reviewers
combine section 2 and 3; rearrange section 3 with bullet points

--- 2023-03-04 3-the-cubble-pkg.Rmd
@@
 # The cubble package {#cubble}
+
+## The cubble object
+
+*the conceptual framework + nested and long form as per the design vignette*
+
+  - diagram, 
+  - two forms: long/ nested - spatial/ temporal
+  - attributes
+  - headers
+
+<!-- Spatio-temporal data can be conceptualized using a cubical data model with three axes which typically are, time, latitude and longitude.
+This abstraction can be useful for generalizing operations and visualization purposes: @lu_multidimensional_2018 shows how array operations (select, scale, reduce, rearrange, and compute) can be mapped onto the cube; @bach_review_2014 reviews the temporal data visualization based on space-time cube operations.
+Notice that the term space-time cube in their article "does not need to involve spatial data", but refers to "an abstract 2D substrate that is used to visualize data at a specific time".
+Despite its main focus being on temporal data, the mindset of abstracting out data representation to construct visualizations, still applies to our spatio-temporal data manipulation and visualization approaches.
+-->
+
+<!-- The most common space-time cube uses the three axes, time, latitude, longitude, and can be considered stacking space across time.
+Ours is a multivariate spatio-temporal cube with the three axes defined to be time, site and variables, as illustrated in the leftmost column of Figure \ref{fig:cube-def}.
+The time axis is the same in both versions, while the site axis now captures both latitude and longitude.
+Finally, variables are stacked on this space-time canvas, with one observation per site and time point.
+This notion is adopted to avoid using hyper-cubes when describing multivariate spatio-temporal data and is the conceptual framework behind the \code{cubble} class.
+With this conceptual model, operations on spatio-temporal data can be mapped to operations on the cube and the rest of Figure \ref{fig:cube-def} show examples of slicing on site, time, and variable.
+-->
 
@@
 The main functions in the package are \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and \code{unfold()}.
-
-<!-- A cubble is a data object that arranges spatio-temporal information in a spatial and temporal tidy way.
-It uses two forms to achieve this: the nested form presents the spatial variables, including the feature geometry, and nests the temporal information for spatial wrangling.
-The long form expands the temporal information and stashes the spatial variables for temporal wrangling.
-A benefit of cubble is that it interfaces directly with existing spatial and temporal data wrangling packages.
-For example, operations on feature geometry from the \code{sf} package and temporal operations, i.e., temporal ordering, duplication check.
-and gap filling, from the \code{tsibble} package.
--->
 
@@
 
-## Create a cubble {#create}
+## Casting into the cubble class {#create}
+
+  - key, index, coords
+  - import different data: an expand data, spatial + temporal, netcdf, stars, sftime
 
@@
 
-## Change focus by facing the time-variables
+*netcdf*
+
+The NetCDF data is another format commonly used for storing spatio-temporal data.
+It has two main components: *dimension* for defining the spatio-temporal grid (longitude, latitude, and time) and *variable* that populates the defined grid.
+Attributes  can be associated with dimensions or variables.
+Because there can be many different styles of representing this information there is a metadata convention [@hassell_data_2017] to standardize the format of the attributes.
+A few packages in \proglang{R} exist for manipulating NetCDF data and these include a high-level R interface: \pkg{ncdf4} [@ncdf4], a low-level interface that calls a C-interface: \pkg{RNetCDF} [@rnetcdf], and a tidyverse implementation: \pkg{tidync} [@tidync].
+
+The \pkg{cubble} package provides an \code{as_cubble()} method to coerce the \code{ncdf4} class from the \pkg{ncdf4} package into a \code{cubble} object.
+It maps each combination of longitude and latitude into an \code{id} as the \code{key}:
+
+```{r echo = TRUE, eval = FALSE}
+dt <- as_cubble(my_ncdf, vars = c("q", "z"))
+```
+
+Sometimes NetCDF data can be quite large, and it is best to subset the data when converting to a \code{cubble} object.
+We would recommend reducing to about 300 $\times$ 300 grid points for three daily variables in one year.
+A 300 by 300 spatial grid can be a bounding box of [100, -80, 180, 0] at 0.25 degree resolution or a global bounding box [-180, -90, 180, -90] at 1 degree resolution.
+The size of spatial grid can be reduced if longer time periods or more variables are needed, through the arguments \code{long_range} and \code{lat_range}:
+
+```{r eval = FALSE, echo = TRUE}
+dt <- as_cubble(
+  my_ncdf, vars = c("q", "z"),
+  long_range = seq(-180, 180, 1), lat_range = seq(-90, 90, 1)
+  )
+```
+
+## Functions and methods in cubble
+
+*Print all the methods with methods(class = "cubble_df")*
+
+### cubble specific methods
+
+The three cubble specific functions are \code{face_temporal()}, \code{face_spatial()}, and \code{unfold}.
 
 The nested form can be used for those operations where the output is only indexed by the spatial identifier (\code{key}), but becomes inadequate when outputs need both a spatial and a temporal identifier (\code{key} and \code{index}).
-The \code{cubble} class also provides a long form, which expands the \code{ts} column and temporarily "hides" the spatial variables.
+The \code{cubble} class also provides a long form, which expands the \code{ts} column and stores the spatial variables as an attribute.
 The function \code{face_temporal()} is used to switch from the nested \code{cubble} into the long one.
@@
 
-## Change focus back to the site-variable face
-
 Wrangling spatio-temporal data can be seen as an iterative process in the spatial and temporal dimensions.
@@
 ```
-
-## Unfold spatial variables into the long form {#unfold}
 
@@
 
-<!-- https://www.jstatsoft.org/style#my-latex-paper-does-not-compile-when-there-is-jss-markup-in-section-titles-what-should-i-do -->
-
-## Why not just use the existing tidyverse functions {#tidyverse}
-
-<!-- \subsection[Why not just use]{Why not just use \code{dplyr::nest\_by()}?} -->
-
-Some readers may question why a new data structure is needed rather than directly creating a list-column  on the combined data using \code{dplyr::nest_by()}.
-The reason is that the \code{cubble} object is specifically designed to utilize the spatio-temporal structure when arranging observations in a single object.
-Moreover, it enables easy pivoting between purely spatial, purely temporal, or unfolded into a combined form.
-
-<!-- Some readers may raise the question why a new data structure rather than using \code{dplyr::nest_by()} on the combined data to directly create a list-column.
-While \code{nest_by()} and its inverse \code{unnest()} can mimic the nested form in cubble, cubble can be seen as an attempt to arrange variables from different observational units into a single object, specifically for spatio-temporal data.
-The benefit of arranging everything in a single object is that there is no need to manually split and join variables in the data.
-For various spatial, temporal, or spatio-temporal tasks, the data format required can alternate among a pure spatial, pure temporal, or a combined format.
-Analysts will need to manually split the combined data into a spatial or temporal one, or join them into a combined one.
-\pkg{Cubble} simplifies this switch of format with a nested and long form and allowing spatial variables to be unfolded into the long form for a combined data all in a single data object.
--->
-
-## Compatibility with existing packages {#compact}
-
-The \pkg{cubble} package leverages tools available in existing packages used for spatial and temporal analysis, specifically, \pkg{dplyr}, \pkg{tsibble}, \pkg{sf} (\pkg{s2}), and \pkg{ncdf4}, as explained here.
-
-\subsubsection[]{\code{dplyr}}
+### dplyr methods
 
@@
 The \pkg{cubble} package provides methods that support the use of the following operations in the \pkg{dplyr} package on both the nested and long forms: `mutate`, `filter`, `summarise`, `select`, `arrange`, `rename`, `left_join`, and the slice family (`slice_*`).
+
+## Compatibility with other packages
 
@@
 
+*add an example on changing CRS?*
+
 The \code{sf} class is also a subclass of \code{tibble} with a specialized feature geometry list-column (\code{sfc}) to store different geometry types (\code{POINT}, \code{LINESTRING}, \code{POLYGON}, \code{MULTIPOLYGON}, etc).
@@
 
-<!-- These spatial operations are also valuable for spatio-temporal data analysis, but an \code{sf} object *usually* does not contain temporal variables.
-This means \code{sf} cannot be directly cast into a \code{cubble}, however, \code{cubble} does support \code{sfc} columns in the nested form and spatial operations applied to the \code{sfc} column in \code{sf} can still be applied to the \code{sfc} column in a cubble.
-The following example shows how to create an \code{sfc} column of \code{POINT} type from latitude and longitude in cubble.
-Then \code{sf::st_within} is used to add the state \code{MULTIPOLYGON} of each weather station before a coordinate transformation is made.
+<!-- https://www.jstatsoft.org/style#my-latex-paper-does-not-compile-when-there-is-jss-markup-in-section-titles-what-should-i-do -->
+
+## Comparison to other spatio-temporal classes {#tidyverse}
+
+Some readers may question why a new data structure is needed rather than directly creating a list-column  on the combined data using \code{dplyr::nest_by()}.
+The reason is that the \code{cubble} object is specifically designed to utilize the spatio-temporal structure when arranging observations in a single object.
+Moreover, it enables easy pivoting between purely spatial, purely temporal, or unfolded into a combined form.
+
+From the initial \pkg{spacetime} [@spacetime] package.
+New packages, such as \pkg{stars} [@stars] and \pkg{sftime} [@sftime], have [...].
+This section compares and contrast the cubble data structure with other existing alternative, namely \pkg{stars} and \pkg{sftime}.
+
+The `stars` package creates a `stars` object from an array and a dimension object.
+In many cases, i.e.
+satellite images, this is convenient, however, there are also many spatio-temporal data are formatted in 2D data frames.
+
+create a dimension object, populate the data into the array
+
+- array in \proglang{R} allows for character, 
+- specific datetime class
+
+*make the example simpler*
+
+  - `temporal$var1 %>% matrix(byrow = TRUE, nrow = 2, ncol = 5)`
+
+This happens when different stations have different start date and require other software to reshape the data correctly.
+
+<!-- The temporal data contains two variables are recorded at two stations (`id`) from 2023 March 21st to 2023 March 25th.
+Notice is that id2 only records from 2023 March 22nd and variable 2 has an `NA` value on  2023 March 23rd at id2.
 -->
 
-<!-- ```{r echo = TRUE} -->
-<!-- library(sf) -->
-<!-- # create a cubble -->
-<!-- cb <- climate_flat |> -->
-<!--   cubble::as_cubble(key = id, index = date, coords = c(long, lat)) -->
+<!-- ```{r echo = FALSE} -->
+<!-- library(stars) -->
+<!-- spatial <- tibble( -->
+<!--   id = c(1,2), -->
+<!--   long = c(130, 140), -->
+<!--   lat = c(-38, -44), -->
+<!--   name = c("station 1", "station 2"))  -->
 
-<!-- aus <- ozmaps::abs_ste -->
+<!-- spatial -->
 
-<!-- dt <- cb |> -->
-<!--   mutate( -->
-<!--     # create `sfc` column based on long and lat -->
-<!--     ll = st_sfc( -->
-<!--       purrr::map2(long, lat, ~st_point(c(.x, .y))), -->
-<!--       crs = st_crs(aus)), -->
-
-<!--     # append state multi-polygon based on the `sfc` created -->
-<!--     state = aus$geometry[st_within(ll, aus, sparse = FALSE)], -->
-
-<!--     # adopt a different projection: lambert conformal conic (EPSG:3112) -->
-<!--     state = st_transform(state, crs = "EPSG:3112") -->
+<!-- temporal <- tibble( -->
+<!--   id = c(rep(1, each = 5), rep(2, each = 4)), -->
+<!--   time = c(seq(as.Date("2023-03-21"), as.Date("2023-03-25"), 1), -->
+<!--            seq(as.Date("2023-03-22"), as.Date("2023-03-25"), 1)), -->
+<!--   var1 = c(3, 6, 8, 2, 5, 7, 9, 14, 5), -->
+<!--   var2 = c(1, 3, 4, 2, 6, 7, NA, 5, 8) -->
 <!-- ) -->
-
-<!-- dt -->
+<!-- temporal -->
 <!-- ``` -->
 
-<!-- An \pkg{s2} \code{lnglat} vector can similarly be created as an \code{sfc} in cubble before using any \code{s2}-prefixed function: -->
+<!-- The following code can be used to create the dimension object: -->
 
 <!-- ```{r echo = TRUE} -->
-<!-- library(s2) -->
-<!-- # Western Australia map -->
-<!-- wa <- ozmaps::abs_ste |> filter(NAME == "Western Australia") -->
-
-<!-- # mutate a `s2_lnglat` vector on `cb` created in the last chunk -->
-<!-- cb |> -->
-<!--   mutate(ll = s2_lnglat(long, lat)) |> -->
-<!--   filter(s2_within(ll, wa)) -->
+<!-- spatial_sf <- spatial %>% sf::st_as_sf(coords = c("long", "lat"), crs = 4326)  -->
+<!-- station_vec <- st_geometry(spatial_sf) -->
+<!-- time_vec <- unique(temporal$time) -->
+<!-- var_vec <- spatial_sf$name -->
+<!-- d <- st_dimensions(station = station_vec, time = time_vec) -->
+<!-- d -->
 <!-- ``` -->
 
-\subsubsection[]{\code{ncdf4}}\label{netcdf}
+<!-- However, the issue comes when casting the variables into an array of the same dimension: -->
 
-The NetCDF data is another format commonly used for storing spatio-temporal data.
-It has two main components: *dimension* for defining the spatio-temporal grid (longitude, latitude, and time) and *variable* that populates the defined grid.
-Attributes  can be associated with dimensions or variables.
-Because there can be many different styles of representing this information there is a metadata convention [@hassell_data_2017] to standardize the format of the attributes.
-A few packages in \proglang{R} exist for manipulating NetCDF data and these include a high-level R interface: \pkg{ncdf4} [@ncdf4], a low-level interface that calls a C-interface: \pkg{RNetCDF} [@rnetcdf], and a tidyverse implementation: \pkg{tidync} [@tidync].
+<!-- ```{r echo = TRUE, eval = FALSE} -->
+<!-- temporal$var1 %>% matrix(byrow = TRUE, nrow = 2, ncol = 5) %>% array(dim = dim(d)) -->
+<!-- ``` -->
 
-The \pkg{cubble} package provides an \code{as_cubble()} method to coerce the \code{ncdf4} class from the \pkg{ncdf4} package into a \code{cubble} object.
-It maps each combination of longitude and latitude into an \code{id} as the \code{key}:
+<!-- Due to the missingness on 2023 March 21st for id2, the array reuses the first value of the variable to fill in the dimension (the missingness at (2,1) is not flagged and 3 is reused to fill in the dimension).
+-->
 
-```{r echo = TRUE, eval = FALSE}
-dt <- as_cubble(my_ncdf, vars = c("q", "z"))
+<!-- This causes an issue when multiple stations with different recording periods present in the data, commonly found in the weather data.  -->
+
+The approach taken by the \pkg{sftime} package is to combine the spatial and temporal variables together into a joint table. [for sf objects with daily observation, memory efficiency]
+
+Consider the climate data used in the \pkg{cubble} package, \code{climate_aus}.
+It contains daily precipitation, minimum and maximum temperature for 639 weather stations across Australia in 2020.
+The converted sftime object is `r library(cubble); object.size(sftime::st_as_sftime(climate_aus))/object.size(climate_aus)` times the size of the cubble object:
+
+```{r echo = TRUE}
+climate_aus %>% sftime::st_as_sftime() %>% object.size()
+climate_aus %>% object.size()
 ```
-
-Sometimes NetCDF data can be quite large, and it is best to subset the data when converting to a \code{cubble} object.
-We would recommend reducing to about 300 $\times$ 300 grid points for three daily variables in one year.
-A 300 by 300 spatial grid can be a bounding box of [100, -80, 180, 0] at 0.25 degree resolution or a global bounding box [-180, -90, 180, -90] at 1 degree resolution.
-The size of spatial grid can be reduced if longer time periods or more variables are needed, through the arguments \code{long_range} and \code{lat_range}:
-
-```{r eval = FALSE, echo = TRUE}
-dt <- as_cubble(
-  my_ncdf, vars = c("q", "z"),
-  long_range = seq(-180, 180, 1), lat_range = seq(-90, 90, 1)
-  )
-```

### 2023-05-22 jss/main.tex
combine section 2 and 3; rearrange section 3 with bullet points

--- 2023-03-04 jss/main.tex
@@
 
-The rest of the paper is organized as follows: Section \ref{cube} introduces the proposed cube structure as a way to conceptualize multivariate spatio-temporal data.
-Section \ref{cubble} presents the main design and functionality of the \pkg{cubble} package.
+The rest of the paper is organized as follows: Section \ref{cubble} presents the main design and functionality of the \pkg{cubble} package.
 Section \ref{others} explains how the \pkg{cubble} package deals with more advanced considerations, including data with hierarchical structure, data matching and how the package fits with existing static and interactive visualization tools.
@@
 
-\hypertarget{cube}{%
-\section{Conceptual framework: spatio-temporal cube}\label{cube}}
-
-Spatio-temporal data can be conceptualized using a cubical data model with three axes which typically are, time, latitude and longitude.
-This abstraction can be useful for generalizing operations and visualization purposes: \citet{lu_multidimensional_2018} shows how array operations (select, scale, reduce, rearrange, and compute) can be mapped onto the cube; \citet{bach_review_2014} reviews the temporal data visualization based on space-time cube operations.
-Notice that the term space-time cube in their article ``does not need to involve spatial data'', but refers to ``an abstract 2D substrate that is used to visualize data at a specific time''.
-Despite its main focus being on temporal data, the mindset of abstracting out data representation to construct visualizations, still applies to our spatio-temporal data manipulation and visualization approaches.
-
-The most common space-time cube uses the three axes, time, latitude, longitude, and can be considered stacking space across time.
-Ours is a multivariate spatio-temporal cube with the three axes defined to be time, site and variables, as illustrated in the leftmost column of Figure \ref{fig:cube-def}.
-The time axis is the same in both versions, while the site axis now captures both latitude and longitude.
-Finally, variables are stacked on this space-time canvas, with one observation per site and time point.
-This notion is adopted to avoid using hyper-cubes when describing multivariate spatio-temporal data and is the conceptual framework behind the \code{cubble} class.
-With this conceptual model, operations on spatio-temporal data can be mapped to operations on the cube and the rest of Figure \ref{fig:cube-def} show examples of slicing on site, time, and variable.
-
-While the data cube model is conceptually convenient for spatio-temporal data, a 3D data array is not sufficiently rich for data wrangling, for several reasons.
-Although arrays can be efficient for the computation on numerical values, spatio-temporal data typically includes various types of variables.
-For example, character strings and specific datetime classes are common.
-In addition, it will be generally useful to be able to create new variables which is trickier to manage in an array.
-Thus for convenient wrangling, we have opted to create a special \code{cubble} class.
-
-\begin{CodeChunk}
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{../figures/diagram-keynotes/diagram-keynotes.001}
-
-}
-
-\caption[An illustration of the conceptual spatio-temporal cube with different slicing on time, site, and variable]{An illustration of the conceptual spatio-temporal cube with different slicing on time, site, and variable.
-For each axis, the slicing can be on a single value or a set of values.}\label{fig:cube-def}
-\end{figure}
-\end{CodeChunk}
-
 \hypertarget{cubble}{%
 \section{The cubble package}\label{cubble}}
+
+\hypertarget{the-cubble-object}{%
+\subsection{The cubble object}\label{the-cubble-object}}
+
+\emph{the conceptual framework + nested and long form as per the design vignette}
+
+\begin{itemize}
+\tightlist
+\item
+  diagram,
+\item
+  two forms: long/ nested - spatial/ temporal
+\item
+  attributes
+\item
+  headers
+\end{itemize}
 
@@
 \hypertarget{create}{%
-\subsection{Create a cubble}\label{create}}
+\subsection{Casting into the cubble class}\label{create}}
+
+\begin{itemize}
+\tightlist
+\item
+  key, index, coords
+\item
+  import different data: an expand data, spatial + temporal, netcdf, stars, sftime
+\end{itemize}
 
@@
 # cubble:   id [5]: nested form
-# bbox:     [115.97, -32.94, 133.55, -12.42]
+# bbox:     [115.9764, -32.9342, 133.5407, -12.4239]
 # temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
@@
 # cubble:   id [5]: nested form
-# bbox:     [115.97, -32.94, 133.55, -12.42]
+# bbox:     [115.9764, -32.9342, 133.5407, -12.4239]
 # temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
-  id            lat  long  elev name          wmo_id ts       rain_day
+  id            lat  long  elev name           wmo_id ts       rain_~1
   <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>     <int>
@@
 3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble>      90
-4 ASN00014015 -12.4  131.  30.4 darwin airpo~  94120 <tibble>      106
+4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble>     106
 5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble>      63
+# ... with abbreviated variable name 1: rain_day
 \end{CodeOutput}
@@
 
-\hypertarget{change-focus-by-facing-the-time-variables}{%
-\subsection{Change focus by facing the time-variables}\label{change-focus-by-facing-the-time-variables}}
+\emph{netcdf}
+
+The NetCDF data is another format commonly used for storing spatio-temporal data.
+It has two main components: \emph{dimension} for defining the spatio-temporal grid (longitude, latitude, and time) and \emph{variable} that populates the defined grid.
+Attributes can be associated with dimensions or variables.
+Because there can be many different styles of representing this information there is a metadata convention \citep{hassell_data_2017} to standardize the format of the attributes.
+A few packages in \proglang{R} exist for manipulating NetCDF data and these include a high-level R interface: \pkg{ncdf4} \citep{ncdf4}, a low-level interface that calls a C-interface: \pkg{RNetCDF} \citep{rnetcdf}, and a tidyverse implementation: \pkg{tidync} \citep{tidync}.
+
+The \pkg{cubble} package provides an \code{as_cubble()} method to coerce the \code{ncdf4} class from the \pkg{ncdf4} package into a \code{cubble} object.
+It maps each combination of longitude and latitude into an \code{id} as the \code{key}:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> dt <- as_cubble(my_ncdf, vars = c("q", "z"))
+\end{CodeInput}
+\end{CodeChunk}
+
+Sometimes NetCDF data can be quite large, and it is best to subset the data when converting to a \code{cubble} object.
+We would recommend reducing to about 300 \(\times\) 300 grid points for three daily variables in one year.
+A 300 by 300 spatial grid can be a bounding box of {[}100, -80, 180, 0{]} at 0.25 degree resolution or a global bounding box {[}-180, -90, 180, -90{]} at 1 degree resolution.
+The size of spatial grid can be reduced if longer time periods or more variables are needed, through the arguments \code{long_range} and \code{lat_range}:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> dt <- as_cubble(
++   my_ncdf, vars = c("q", "z"),
++   long_range = seq(-180, 180, 1), lat_range = seq(-90, 90, 1)
++   )
+\end{CodeInput}
+\end{CodeChunk}
+
+\hypertarget{functions-and-methods-in-cubble}{%
+\subsection{Functions and methods in cubble}\label{functions-and-methods-in-cubble}}
+
+\emph{Print all the methods with methods(class = ``cubble\_df'')}
+
+\hypertarget{cubble-specific-methods}{%
+\subsubsection{cubble specific methods}\label{cubble-specific-methods}}
+
+The three cubble specific functions are \code{face_temporal()}, \code{face_spatial()}, and \code{unfold}.
 
 The nested form can be used for those operations where the output is only indexed by the spatial identifier (\code{key}), but becomes inadequate when outputs need both a spatial and a temporal identifier (\code{key} and \code{index}).
-The \code{cubble} class also provides a long form, which expands the \code{ts} column and temporarily ``hides'' the spatial variables.
+The \code{cubble} class also provides a long form, which expands the \code{ts} column and stores the spatial variables as an attribute.
 The function \code{face_temporal()} is used to switch from the nested \code{cubble} into the long one.
@@
 # cubble:  date, id [5]: long form
-# bbox:    [115.97, -32.94, 133.55, -12.42]
+# bbox:    [115.9764, -32.9342, 133.5407, -12.4239]
 # spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id
@@
 
-\hypertarget{change-focus-back-to-the-site-variable-face}{%
-\subsection{Change focus back to the site-variable face}\label{change-focus-back-to-the-site-variable-face}}
-
 Wrangling spatio-temporal data can be seen as an iterative process in the spatial and temporal dimensions.
@@
 
-\hypertarget{unfold}{%
-\subsection{Unfold spatial variables into the long form}\label{unfold}}
-
 Sometimes, analysts may need to apply some variable transformation that involves both the spatial and temporal variables.
@@
 
-\hypertarget{tidyverse}{%
-\subsection{Why not just use the existing tidyverse functions}\label{tidyverse}}
-
-Some readers may question why a new data structure is needed rather than directly creating a list-column on the combined data using \code{dplyr::nest_by()}.
-The reason is that the \code{cubble} object is specifically designed to utilize the spatio-temporal structure when arranging observations in a single object.
-Moreover, it enables easy pivoting between purely spatial, purely temporal, or unfolded into a combined form.
-
-\hypertarget{compact}{%
-\subsection{Compatibility with existing packages}\label{compact}}
-
-The \pkg{cubble} package leverages tools available in existing packages used for spatial and temporal analysis, specifically, \pkg{dplyr}, \pkg{tsibble}, \pkg{sf} (\pkg{s2}), and \pkg{ncdf4}, as explained here.
-
-\subsubsection[]{\code{dplyr}}
+\hypertarget{dplyr-methods}{%
+\subsubsection{dplyr methods}\label{dplyr-methods}}
 
@@
 The \pkg{cubble} package provides methods that support the use of the following operations in the \pkg{dplyr} package on both the nested and long forms: \texttt{mutate}, \texttt{filter}, \texttt{summarise}, \texttt{select}, \texttt{arrange}, \texttt{rename}, \texttt{left\_join}, and the slice family (\texttt{slice\_*}).
+
+\hypertarget{compatibility-with-other-packages}{%
+\subsection{Compatibility with other packages}\label{compatibility-with-other-packages}}
 
@@
 # cubble:   id [5]: nested form
-# bbox:     [115.97, -32.94, 133.55, -12.42]
+# bbox:     [115.9764, -32.9342, 133.5407, -12.4239]
 # temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
-  id          lat  long  elev name  wmo_id ts       tmax_mean tmax_var
+  id            lat  long  elev name   wmo_id ts       tmax_~1 tmax_~2
   <chr>       <dbl> <dbl> <dbl> <chr>   <dbl> <list>     <dbl>   <dbl>
-1 ASN00009~ -31.9  116.  15.4 pert~  94610 <tbl_ts>      25.7    38.6 
-2 ASN00010~ -31.9  117. 179   york   94623 <tbl_ts>      26.2    51.1 
-3 ASN00010~ -32.9  117. 338   narr~  94627 <tbl_ts>      23.7    45.4 
-4 ASN00014~ -12.4  131.  30.4 darw~  94120 <tbl_ts>      33.1     3.02
-5 ASN00015~ -17.6  134. 220   elli~  94236 <tbl_ts>      34.6    24.7 
+1 ASN00009021 -31.9  116.  15.4 perth~  94610 <tbl_ts>    25.7   38.6 
+2 ASN00010311 -31.9  117. 179   york    94623 <tbl_ts>    26.2   51.1 
+3 ASN00010614 -32.9  117. 338   narro~  94627 <tbl_ts>    23.7   45.4 
+4 ASN00014015 -12.4  131.  30.4 darwi~  94120 <tbl_ts>    33.1    3.02
+5 ASN00015131 -17.6  134. 220   ellio~  94236 <tbl_ts>    34.6   24.7 
+# ... with abbreviated variable names 1: tmax_mean, 2: tmax_var
 \end{CodeOutput}
@@
 \subsubsection[]{\code{sf} (\code{s2})}
+
+\emph{add an example on changing CRS?}
 
@@
 
-\subsubsection[]{\code{ncdf4}}\label{netcdf}
-
-The NetCDF data is another format commonly used for storing spatio-temporal data.
-It has two main components: \emph{dimension} for defining the spatio-temporal grid (longitude, latitude, and time) and \emph{variable} that populates the defined grid.
-Attributes can be associated with dimensions or variables.
-Because there can be many different styles of representing this information there is a metadata convention \citep{hassell_data_2017} to standardize the format of the attributes.
-A few packages in \proglang{R} exist for manipulating NetCDF data and these include a high-level R interface: \pkg{ncdf4} \citep{ncdf4}, a low-level interface that calls a C-interface: \pkg{RNetCDF} \citep{rnetcdf}, and a tidyverse implementation: \pkg{tidync} \citep{tidync}.
-
-The \pkg{cubble} package provides an \code{as_cubble()} method to coerce the \code{ncdf4} class from the \pkg{ncdf4} package into a \code{cubble} object.
-It maps each combination of longitude and latitude into an \code{id} as the \code{key}:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> dt <- as_cubble(my_ncdf, vars = c("q", "z"))
-\end{CodeInput}
-\end{CodeChunk}
-
-Sometimes NetCDF data can be quite large, and it is best to subset the data when converting to a \code{cubble} object.
-We would recommend reducing to about 300 \(\times\) 300 grid points for three daily variables in one year.
-A 300 by 300 spatial grid can be a bounding box of {[}100, -80, 180, 0{]} at 0.25 degree resolution or a global bounding box {[}-180, -90, 180, -90{]} at 1 degree resolution.
-The size of spatial grid can be reduced if longer time periods or more variables are needed, through the arguments \code{long_range} and \code{lat_range}:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> dt <- as_cubble(
-+   my_ncdf, vars = c("q", "z"),
-+   long_range = seq(-180, 180, 1), lat_range = seq(-90, 90, 1)
-+   )
-\end{CodeInput}
+\hypertarget{tidyverse}{%
+\subsection{Comparison to other spatio-temporal classes}\label{tidyverse}}
+
+Some readers may question why a new data structure is needed rather than directly creating a list-column on the combined data using \code{dplyr::nest_by()}.
+The reason is that the \code{cubble} object is specifically designed to utilize the spatio-temporal structure when arranging observations in a single object.
+Moreover, it enables easy pivoting between purely spatial, purely temporal, or unfolded into a combined form.
+
+From the initial \pkg{spacetime} \citep{spacetime} package.
+New packages, such as \pkg{stars} \citep{stars} and \pkg{sftime} \citep{sftime}, have {[}\ldots{]}.
+This section compares and contrast the cubble data structure with other existing alternative, namely \pkg{stars} and \pkg{sftime}.
+
+The \texttt{stars} package creates a \texttt{stars} object from an array and a dimension object.
+In many cases, i.e.~satellite images, this is convenient, however, there are also many spatio-temporal data are formatted in 2D data frames.
+
+create a dimension object, populate the data into the array
+
+\begin{itemize}
+\tightlist
+\item
+  array in \proglang{R} allows for character,
+\item
+  specific datetime class
+\end{itemize}
+
+\emph{make the example simpler}
+
+\begin{itemize}
+\tightlist
+\item
+  \texttt{temporal\$var1\ \%\textgreater{}\%\ matrix(byrow\ =\ TRUE,\ nrow\ =\ 2,\ ncol\ =\ 5)}
+\end{itemize}
+
+This happens when different stations have different start date and require other software to reshape the data correctly.
+
+The approach taken by the \pkg{sftime} package is to combine the spatial and temporal variables together into a joint table. {[}for sf objects with daily observation, memory efficiency{]}
+
+Consider the climate data used in the \pkg{cubble} package, \code{climate_aus}.
+It contains daily precipitation, minimum and maximum temperature for 639 weather stations across Australia in 2020.
+The converted sftime object is 13.6830016 times the size of the cubble object:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> climate_aus %>% sftime::st_as_sftime() %>% object.size()
+\end{CodeInput}
+\begin{CodeOutput}
+118243136 bytes
+\end{CodeOutput}
+\begin{CodeInput}
+R> climate_aus %>% object.size()
+\end{CodeInput}
+\begin{CodeOutput}
+8641608 bytes
+\end{CodeOutput}
 \end{CodeChunk}
@@
 This will activate rows having the same id in the long \code{cubble}.
-The corresponding rows in the nested \code{cubble} are activated, and highlighted in the map.
+The corresponding rows in the nested \code{cubble} are activated, and highlighted on the map.
 (An illustration can be found in the appendix.) Note that this type of linking, both from the map or the time series, is what \citet{CS07} would call categorical variable linking, where station id is the categorical variable.
@@
 \hypertarget{examples}{%
-\section{Examples}\label{examples}}
+\section{Applications}\label{examples}}
 
@@
 To fix this, one could aggregate data from nearby stations.
-An example of this is included in the Appendix.
+An example of this use is included in the Appendix.
 
@@
 # cubble:   id [8]: nested form
-# bbox:     [144.52, -37.73, 146.06, -36.55]
+# bbox:     [144.5203, -37.727706, 146.059205, -36.5523]
 # temporal: date [date], matched_var [dbl]

### 2023-05-29 5-1-covid.Rmd
fill in response to reviewers
update the covid example with the latest code

@@
 
-# Examples {#examples}
+# Applications {#examples}
 
@@
 
-```{r}
+```{r covid-1}
 load(here::here("data/covid.rda"))
-
 lga <- strayr::read_absmap("lga2018") |>
   rename(lga = lga_name_2018) |>
-  filter(state_name_2016 == "Victoria") 
+  filter(state_name_2016 == "Victoria") |>
+  select(lga, geometry)
 ```
 
-```{r echo = TRUE}
+```{r covid-2, echo = TRUE}
 covid |> head(5)
@@
 
-```{r echo = TRUE}
+```{r covid-3, echo = TRUE}
 lga |> head(5)
@@
 
-The function \code{as_cubble()} is used to create a \code{cubble} object from the two spatial and temporal tables, and requires specifying the arguments \code{key}, \code{index}, and \code{coords} (as described in Section \@ref(create)).
+The function \code{make_cubble()} is used to create a \code{cubble} object from the two spatial and temporal tables, and requires specifying the arguments \code{key}, \code{index}, and \code{coords} (as described in Section \@ref(create)).
 It will automatically try to match the sites in both tables using the location names and will show a warning message when there are mismatches, as shown below:
 
-```{r echo = TRUE, warning = TRUE, message= TRUE}
-cb <- as_cubble(
-  list(spatial = lga, temporal = covid),
-  key = lga, index = date, coords = c(cent_long, cent_lat)
-  )
+```{r covid-4, echo = TRUE, warning = TRUE, message= TRUE}
+cb <- make_cubble(spatial = lga, temporal = covid)
 ```
@@
 It can be seen that there are two-way mismatches -- LGAs in the COVID data that do not  match with LGAs names in the spatial polygon data, and vice versa.
-The mismatches can be identified by using the \code{output = "unmatch"} argument in the \code{as_cubble()} function: <!--Note that \pkg{cubble} will attempt to pair the unmatched sites, these are shown separately from the ones that cannot be paired.-->
+The mismatches can be identified by using the function \code{check_key}:
 
-```{r echo = TRUE}
-pair <- as_cubble(
-  list(spatial = lga, temporal = covid),
-  key = lga, index = date, coords = c(cent_long, cent_lat),
-  output = "unmatch"
-  )
-
-pair
+```{r covid-5, echo = TRUE}
+(check_res <- check_key(spatial = lga, temporal = covid))
 ```
@@
 
-```{r echo = TRUE}
-lga <- lga |>
+```{r covid-6, echo = TRUE}
+lga2 <- lga |>
   mutate(lga = ifelse(lga == "Kingston (C) (Vic.)", "Kingston (C)", lga),
          lga = ifelse(lga == "Latrobe (C) (Vic.)", "Latrobe (C)", lga)) |>
-  filter(!lga %in% pair$others$spatial)
+  filter(!lga %in% check_res$others$spatial)
 
-covid <- covid |> filter(!lga %in% pair$others$temporal)
+covid2 <- covid |> filter(!lga %in% check_res$others$temporal)
 
-cb <- as_cubble(data = list(spatial = lga, temporal = covid),
-                key = lga, index = date, coords = c(cent_long, cent_lat))
+(cb <- make_cubble(spatial = lga2, temporal = covid2))
 ```

### 2023-05-29 3-the-cubble-pkg.Rmd
update code and rearrange subsection

@@
 ## The cubble object
-
-*the conceptual framework + nested and long form as per the design vignette*
-
-  - diagram, 
-  - two forms: long/ nested - spatial/ temporal
-  - attributes
-  - headers
 
@@
 
+*the conceptual framework + nested and long form as per the design vignette*
+
+  - diagram, 
+  - two forms: long/ nested - spatial/ temporal
+  - attributes
+  - headers
+
 The \code{cubble} class is an S3 class [@advr] built on the \code{tibble} class, specifically to organize spatio-temporal data.
@@
 The long \code{cubble} is a subclass of grouped \code{tibble} (\code{grouped_df}), which expands the temporal information into the long form and stores the spatial information in a "spatial" attribute.
-The main functions in the package are \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and \code{unfold()}.
-
-The data \code{climate_flat} is used to illustrate functionality in the \pkg{cubble} package.
-This is a subset from National Oceanic and Atmospheric Administration (NOAA) [@noaa] Global Historical Climatology Network (GHCN) Daily data.
-It contains spatial variables, station id, latitude, longitude, elevation, station name, world meteorology organisation id, in additional to daily temporal information,  maximum and minimum temperature values and precipitation records for year 2020.
-The first five rows of the data are shown below:
 
 ```{r}
-cubble::climate_flat
+cb_nested <- climate_mel
 ```
 
-## Casting into the cubble class {#create}
-
-  - key, index, coords
-  - import different data: an expand data, spatial + temporal, netcdf, stars, sftime
-
-The function \code{as_cubble()} is used to create a \code{cubble} object with three arguments: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order (longitude, latitude).
-The arguments \code{key} and \code{index} follow the conventions in the \pkg{tsibble} package to describe the temporal order and multiple series while \code{coords} specifies the spatial location of each site.
-The code below creates a \code{cubble} object out of \code{climate_flat} (a single \code{tibble}) with \code{id} as the key, \code{date} as the index, and \code{c(long, lat)} as the coordinates:
-
 ```{r echo = TRUE}
-cubble_nested <- climate_flat |>
-  as_cubble(key = id, index = date, coords = c(long, lat))
-cubble_nested
+cb_nested
+class(cb_nested)
 ```
@@
 
-The created \code{cubble} is a subclass of the \code{rowwise_df} class where each row forms a group.
-All the temporal variables are nested in a list column, hence it is also called the nested \code{cubble}.
-The rowwise structure makes it simpler to operate on the list using the \code{mutate()} syntax, which is simpler than the \code{purr::map()} when working with a list column.
-For example, calculating the number of rainy days can be done by:
+```{r}
+cb_long <- face_temporal(climate_mel)
+```
 
 ```{r echo = TRUE}
-cubble_nested |>
-  mutate(rain_day = sum(ts$prcp != 0))
+cb_long
+class(cb_long)
+```
+
+<!-- The data \code{climate_flat} is used to illustrate functionality in the \pkg{cubble} package.
+This is a subset from National Oceanic and Atmospheric Administration (NOAA) [@noaa] Global Historical Climatology Network (GHCN) Daily data.
+It contains spatial variables, station id, latitude, longitude, elevation, station name, world meteorology organisation id, in additional to daily temporal information,  maximum and minimum temperature values and precipitation records for year 2020.
+The first five rows of the data are shown below: -->
+
+## Creation and coercion {#create}
+
+### Create from separate spatial and temporal table
+
+```{r echo = TRUE}
+make_cubble(
+  spatial = stations, temporal = meteo,
+  key = id, index = date, coords = c(long, lat)
+  )
+```
+
+```{r echo = TRUE, eval = FALSE}
+make_cubble(spatial = stations_sf, temporal = meteo, 
+            key = id, index = date)
+
+make_cubble(spatial = stations, temporal = meteo_ts, 
+            coords = c(long, lat))
+
+make_cubble(spatial = stations_sf, temporal = meteo_ts)
+```
+
+### Coerce from other objects
+
+  - import different data: an expand data, spatial + temporal, netcdf, stars, sftime
+
+The function \code{as_cubble()} is used to create a \code{cubble} object with three arguments: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order (longitude, latitude).
+The arguments \code{key} and \code{index} follow the conventions in the \pkg{tsibble} package to describe the temporal order and multiple series while \code{coords} specifies the spatial location of each site.
+The code below creates a \code{cubble} object out of \code{climate_flat} (a single \code{tibble}) with \code{id} as the key, \code{date} as the index, and \code{c(long, lat)} as the coordinates:
+
+```{r echo = TRUE}
+climate_flat |> as_cubble(key = id, index = date, coords = c(long, lat))
 ```
@@
 
+*stars*
+
+*sftimes*
+
 ## Functions and methods in cubble
@@
 *Print all the methods with methods(class = "cubble_df")*
+
+### dplyr methods
+
+The \pkg{dplyr} package has many tools for wrangling tidy data,  many of which are useful in the spatio-temporal analysis.
+The \pkg{cubble} package provides methods that support the use of the following operations in the \pkg{dplyr} package on both the nested and long forms: `mutate`, `filter`, `summarise`, `select`, `arrange`, `rename`, `left_join`, and the slice family (`slice_*`).
+
+The created \code{cubble} is a subclass of the \code{rowwise_df} class where each row forms a group.
+All the temporal variables are nested in a list column, hence it is also called the nested \code{cubble}.
+The rowwise structure makes it simpler to operate on the list using the \code{mutate()} syntax, which is simpler than the \code{purr::map()} when working with a list column.
+For example, calculating the number of rainy days can be done by:
+
+```{r echo = TRUE}
+cb_nested |> mutate(rain_day = sum(ts$prcp != 0))
+```
 
@@
 ```{r echo = TRUE}
-cubble_long <- cubble_nested |> face_temporal()
-cubble_long
+(cb_long <- cb_nested |> face_temporal())
 ```
@@
 ```{r echo = TRUE}
-cubble_back <- cubble_long |> face_spatial()
-cubble_back
-identical(cubble_nested, cubble_back)
+(cb_back <- cb_long |> face_spatial())
+#identical(cb_nested, cb_back)
 ```
@@
 ```{r echo = TRUE}
-cubble_unfold <- cubble_long |> unfold(long, lat)
-cubble_unfold
+cb_long |> unfold(long, lat)
 ```
-
-### dplyr methods
-
-The \pkg{dplyr} package has many tools for wrangling tidy data,  many of which are useful in the spatio-temporal analysis.
-The \pkg{cubble} package provides methods that support the use of the following operations in the \pkg{dplyr} package on both the nested and long forms: `mutate`, `filter`, `summarise`, `select`, `arrange`, `rename`, `left_join`, and the slice family (`slice_*`).
 
@@
 ```{r echo = TRUE}
-climate_flat_ts <- climate_flat |> 
-  tsibble::as_tsibble(key = id, index = date)
-climate_flat_cb <-  climate_flat_ts |> 
-  cubble::as_cubble(coords = c(long, lat))
-climate_flat_cb
+cb <- make_cubble(
+  spatial = stations, temporal = meteo_ts, 
+  coords = c(long, lat))
 ```
@@
 ```{r echo = TRUE}
-climate_flat_cb |> 
-  mutate(fabletools::features(
-    ts, tmax, list(tmax_mean = mean, tmax_var = var)
-    ))
+cb %>% face_temporal() %>% tsibble::has_gaps()
 ```
@@
 
-<!-- https://www.jstatsoft.org/style#my-latex-paper-does-not-compile-when-there-is-jss-markup-in-section-titles-what-should-i-do -->
+```{r echo = TRUE}
+cb <- make_cubble(
+  spatial = stations_sf, temporal = meteo, 
+  key = id, index = date)
+class(cb)
+#cb %>% sf::st_transform(crs = "EPSG:3857")
+```
 
@@
 It contains daily precipitation, minimum and maximum temperature for 639 weather stations across Australia in 2020.
-The converted sftime object is `r library(cubble); object.size(sftime::st_as_sftime(climate_aus))/object.size(climate_aus)` times the size of the cubble object:
+The converted sftime object is `r library(cubble); #object.size(sftime::st_as_sftime(climate_aus))/object.size(climate_aus)` times the size of the cubble object:
 
 ```{r echo = TRUE}
-climate_aus %>% sftime::st_as_sftime() %>% object.size()
+#climate_aus %>% sftime::st_as_sftime() %>% object.size()
 climate_aus %>% object.size()

### 2023-05-29 code/reproducible-script.R
update data

@@ -556,64 +556,64 @@ knitr::include_graphics(here::here("figures/popup.png"))
 
 
 ## ----eval = FALSE-------------------------------------------------------------
-## library(tidyverse)
-## library(cubble)
-## all_stations <- rnoaa::ghcnd_stations() |>
-##   filter(str_starts(id, "ASN")) |> # Australian stations start wiht "ASN"
-##   filter(last_year >= 2020) |>
-##   mutate(wmo_id = as.numeric(wmo_id), name = str_to_lower(name)) |>
-##   select(-state, -gsn_flag) |>
-##   select(id, longitude, latitude, elevation, name,
-##          wmo_id, element, first_year, last_year) |>
-##   rename(long = longitude, lat = latitude, elev = elevation)
-## 
-## tmax_stations <- all_stations |>
-##   filter(element == "TMAX", first_year < 1970, !is.na(wmo_id))
-## 
-## raw_tmax <- all_stations |>
-##   rowwise() |>
-##   mutate(ts = list(rnoaa::meteo_pull_monitors(
-##     monitors = id, var = "TMAX",
-##     date_min = glue::glue("{first_year}-01-01"),
-##     date_max = glue::glue("{last_year}-12-31")
-##     ) |>
-##       select(-id)
-##     )
-##   )
-## 
-## historical_tmax <- raw_tmax |>
-##   select(-element) |>
-##   unnest(ts) |>
-##   mutate(tmax = tmax/10) |>
-##   filter(lubridate::year(date) %in% c(1971: 1975, 2016:2020)) |>
-##   as_cubble(index = date, key = id, coords = c(long, lat))
-## 
-## save(historical_tmax, file = here::here("data/historical_tmax.rda"))
+# library(tidyverse)
+# library(cubble)
+# all_stations <- rnoaa::ghcnd_stations() |>
+#   filter(str_starts(id, "ASN")) |> # Australian stations start wiht "ASN"
+#   filter(last_year >= 2020) |>
+#   mutate(wmo_id = as.numeric(wmo_id), name = str_to_lower(name)) |>
+#   select(-state, -gsn_flag) |>
+#   select(id, longitude, latitude, elevation, name,
+#          wmo_id, element, first_year, last_year) |>
+#   rename(long = longitude, lat = latitude, elev = elevation)
+# 
+# tmax_stations <- all_stations |>
[…]

### 2023-05-29 5-2-aus-hist-max-temp.Rmd
fill in response to reviewers
knit

--- 2022-11-13 5-2-aus-hist-max-temp.Rmd
@@
 ```{r}
+
 load(here::here("data/historical_tmax.rda"))
@@
 ```{r echo = TRUE}
+
 tmax <- historical_tmax |>
-  filter(between(stringr::str_sub(id, 7, 8), 46, 90))
+  filter(between(as.numeric(stringr::str_sub(id, 7, 8)), 46, 90))
 ```
@@
   face_temporal() |> 
-  mutate(month = lubridate::month(date),
-         group = as.factor(
-           ifelse(lubridate::year(date) > 2015,
-           "2016 ~ 2020", "1971 ~ 1975"))) |>
-  group_by(month, group) |> 
+  group_by(
+    yearmonth = tsibble::make_yearmonth(
+      year = ifelse(lubridate::year(date) > 2015, 2016, 1971),
+      month = lubridate::month(date))
+    )|>
   summarise(tmax = mean(tmax, na.rm = TRUE))
@@
 To fix this, one could aggregate data from nearby stations.
-An example of this using is included in the Appendix.
+An example of this use is included in the Appendix.

### 2023-05-29 jss/main.tex
knit

@@
 The long \code{cubble} is a subclass of grouped \code{tibble} (\code{grouped_df}), which expands the temporal information into the long form and stores the spatial information in a ``spatial'' attribute.
-The main functions in the package are \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and \code{unfold()}.
-
-The data \code{climate_flat} is used to illustrate functionality in the \pkg{cubble} package.
-This is a subset from National Oceanic and Atmospheric Administration (NOAA) \citep{noaa} Global Historical Climatology Network (GHCN) Daily data.
-It contains spatial variables, station id, latitude, longitude, elevation, station name, world meteorology organisation id, in additional to daily temporal information, maximum and minimum temperature values and precipitation records for year 2020.
-The first five rows of the data are shown below:
-
-\begin{CodeChunk}
-\begin{CodeOutput}
-# A tibble: 1,830 x 10
-  id         lat  long  elev name  wmo_id date        prcp  tmax  tmin
-  <chr>    <dbl> <dbl> <dbl> <chr>  <dbl> <date>     <dbl> <dbl> <dbl>
-1 ASN0000~ -31.9  116.  15.4 pert~  94610 2020-01-01     0  31.9  15.3
-2 ASN0000~ -31.9  116.  15.4 pert~  94610 2020-01-02     0  24.9  16.4
-3 ASN0000~ -31.9  116.  15.4 pert~  94610 2020-01-03     6  23.2  13  
-4 ASN0000~ -31.9  116.  15.4 pert~  94610 2020-01-04     0  28.4  12.4
-5 ASN0000~ -31.9  116.  15.4 pert~  94610 2020-01-05     0  35.3  11.6
-# ... with 1,825 more rows
-\end{CodeOutput}
-\end{CodeChunk}
-
-\hypertarget{create}{%
-\subsection{Casting into the cubble class}\label{create}}
-
-\begin{itemize}
-\tightlist
-\item
-  key, index, coords
-\item
-  import different data: an expand data, spatial + temporal, netcdf, stars, sftime
-\end{itemize}
-
-The function \code{as_cubble()} is used to create a \code{cubble} object with three arguments: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order (longitude, latitude).
-The arguments \code{key} and \code{index} follow the conventions in the \pkg{tsibble} package to describe the temporal order and multiple series while \code{coords} specifies the spatial location of each site.
-The code below creates a \code{cubble} object out of \code{climate_flat} (a single \code{tibble}) with \code{id} as the key, \code{date} as the index, and \code{c(long, lat)} as the coordinates:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> cubble_nested <- climate_flat |>
-+   as_cubble(key = id, index = date, coords = c(long, lat))
-R> cubble_nested
-\end{CodeInput}
-\begin{CodeOutput}
-# cubble:   id [5]: nested form
-# bbox:     [115.9764, -32.9342, 133.5407, -12.4239]
-# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
-  id            lat  long  elev name           wmo_id ts      
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> cb_nested
+\end{CodeInput}
+\begin{CodeOutput}
+# cubble:   key: id [3], index: date, nested form
+# extent:   [144.8321, -37.98, 145.0964, -37.6655]
+# temporal: prcp [dbl], tmax [dbl], tmin [dbl]
+  id           long   lat  elev name              wmo_id ts      
   <chr>       <dbl> <dbl> <dbl> <chr>              <dbl> <list>  
-1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tibble>
-2 ASN00010311 -31.9  117. 179   york            94623 <tibble>
-3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble>
-4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble>
-5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble>
+1 ASN00086038  145. -37.7  78.4 essendon airport   95866 <tibble>
+2 ASN00086077  145. -38.0  12.1 moorabbin airport  94870 <tibble>
+3 ASN00086282  145. -37.7 113.  melbourne airport  94866 <tibble>
+\end{CodeOutput}
+\begin{CodeInput}
+R> class(cb_nested)
+\end{CodeInput}
+\begin{CodeOutput}
+[1] "spatial_cubble_df" "cubble_df"         "rowwise_df"       
+[4] "tbl_df"            "tbl"               "data.frame"       
 \end{CodeOutput}
@@
 
-The created \code{cubble} is a subclass of the \code{rowwise_df} class where each row forms a group.
-All the temporal variables are nested in a list column, hence it is also called the nested \code{cubble}.
-The rowwise structure makes it simpler to operate on the list using the \code{mutate()} syntax, which is simpler than the \code{purr::map()} when working with a list column.
-For example, calculating the number of rainy days can be done by:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> cubble_nested |>
-+   mutate(rain_day = sum(ts$prcp != 0))
-\end{CodeInput}
-\begin{CodeOutput}
-# cubble:   id [5]: nested form
-# bbox:     [115.9764, -32.9342, 133.5407, -12.4239]
-# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
-  id            lat  long  elev name           wmo_id ts       rain_~1
-  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>     <int>
-1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tibble>     104
-2 ASN00010311 -31.9  117. 179   york            94623 <tibble>      89
-3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble>      90
-4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble>     106
-5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble>      63
-# ... with abbreviated variable name 1: rain_day
+\begin{CodeChunk}
+\begin{CodeInput}
+R> cb_long
+\end{CodeInput}
+\begin{CodeOutput}
+# cubble:  key: id [3], index: date, long form
+# extent:  2020-01-01 -- 2020-01-10 [1D], no gaps
+# spatial: long [dbl], lat [dbl], elev [dbl], name [chr], wmo_id
+#   [dbl]
+  id          date        prcp  tmax  tmin
+  <chr>       <date>     <dbl> <dbl> <dbl>
+1 ASN00086038 2020-01-01     0  26.8  11  
+2 ASN00086038 2020-01-02     0  26.3  12.2
+3 ASN00086038 2020-01-03     0  34.5  12.7
+4 ASN00086038 2020-01-04     0  29.3  18.8
+5 ASN00086038 2020-01-05    18  16.1  12.5
+# i 25 more rows
+\end{CodeOutput}
+\begin{CodeInput}
+R> class(cb_long)
+\end{CodeInput}
+\begin{CodeOutput}
+[1] "temporal_cubble_df" "cubble_df"          "tbl_df"            
+[4] "tbl"                "data.frame"        
+\end{CodeOutput}
+\end{CodeChunk}
+
+\hypertarget{create}{%
+\subsection{Creation and coercion}\label{create}}
+
+\hypertarget{create-from-separate-spatial-and-temporal-table}{%
+\subsubsection{Create from separate spatial and temporal table}\label{create-from-separate-spatial-and-temporal-table}}
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> make_cubble(
++   spatial = stations, temporal = meteo,
++   key = id, index = date, coords = c(long, lat)
++   )
+\end{CodeInput}
+\begin{CodeOutput}
+# cubble:   key: id [3], index: date, nested form
+# extent:   [144.8321, -37.98, 145.0964, -37.6655]
+# temporal: prcp [dbl], tmax [dbl], tmin [dbl]
+  id           long   lat  elev name              wmo_id ts      
+  <chr>       <dbl> <dbl> <dbl> <chr>              <dbl> <list>  
+1 ASN00086038  145. -37.7  78.4 essendon airport   95866 <tibble>
+2 ASN00086077  145. -38.0  12.1 moorabbin airport  94870 <tibble>
+3 ASN00086282  145. -37.7 113.  melbourne airport  94866 <tibble>
+\end{CodeOutput}
+\end{CodeChunk}
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> make_cubble(
++   spatial = stations_sf, temporal = meteo, key = id, index = date
++   )
+R> 
+R> make_cubble(
++   spatial = stations, temporal = meteo_ts, coords = c(long, lat)
++   )
+R> 
+R> make_cubble(spatial = stations_sf, temporal = meteo_ts)
+\end{CodeInput}
+\end{CodeChunk}
+
+\hypertarget{coerce-from-other-objects}{%
+\subsubsection{Coerce from other objects}\label{coerce-from-other-objects}}
+
+\begin{itemize}
+\tightlist
+\item
+  import different data: an expand data, spatial + temporal, netcdf, stars, sftime
+\end{itemize}
+
+The function \code{as_cubble()} is used to create a \code{cubble} object with three arguments: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order (longitude, latitude).
+The arguments \code{key} and \code{index} follow the conventions in the \pkg{tsibble} package to describe the temporal order and multiple series while \code{coords} specifies the spatial location of each site.
+The code below creates a \code{cubble} object out of \code{climate_flat} (a single \code{tibble}) with \code{id} as the key, \code{date} as the index, and \code{c(long, lat)} as the coordinates:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> climate_flat |> as_cubble(key = id, index = date, coords = c(long, lat))
+\end{CodeInput}
+\begin{CodeOutput}
+# cubble:   key: id [3], index: date, nested form
+# extent:   [144.8321, -37.98, 145.0964, -37.6655]
+# temporal: prcp [dbl], tmax [dbl], tmin [dbl]
+  id           long   lat  elev name              wmo_id ts      
+  <chr>       <dbl> <dbl> <dbl> <chr>              <dbl> <list>  
+1 ASN00086038  145. -37.7  78.4 essendon airport   95866 <tibble>
+2 ASN00086077  145. -38.0  12.1 moorabbin airport  94870 <tibble>
+3 ASN00086282  145. -37.7 113.  melbourne airport  94866 <tibble>
 \end{CodeOutput}
@@
 
+\emph{stars}
+
+\emph{sftimes}
+
 \hypertarget{functions-and-methods-in-cubble}{%
@@
 \emph{Print all the methods with methods(class = ``cubble\_df'')}
+
+\hypertarget{dplyr-methods}{%
+\subsubsection{dplyr methods}\label{dplyr-methods}}
+
+The \pkg{dplyr} package has many tools for wrangling tidy data, many of which are useful in the spatio-temporal analysis.
+The \pkg{cubble} package provides methods that support the use of the following operations in the \pkg{dplyr} package on both the nested and long forms: \texttt{mutate}, \texttt{filter}, \texttt{summarise}, \texttt{select}, \texttt{arrange}, \texttt{rename}, \texttt{left\_join}, and the slice family (\texttt{slice\_*}).
+
+The created \code{cubble} is a subclass of the \code{rowwise_df} class where each row forms a group.
+All the temporal variables are nested in a list column, hence it is also called the nested \code{cubble}.
+The rowwise structure makes it simpler to operate on the list using the \code{mutate()} syntax, which is simpler than the \code{purr::map()} when working with a list column.
+For example, calculating the number of rainy days can be done by:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> cb_nested |> mutate(rain_day = sum(ts$prcp != 0))
+\end{CodeInput}
+\begin{CodeOutput}
+# cubble:   key: id [3], index: date, nested form
+# extent:   [144.8321, -37.98, 145.0964, -37.6655]
+# temporal: prcp [dbl], tmax [dbl], tmin [dbl]
+  id           long   lat  elev name          wmo_id rain_day ts      
+  <chr>       <dbl> <dbl> <dbl> <chr>          <dbl>    <int> <list>  
+1 ASN00086038  145. -37.7  78.4 essendon air~  95866        3 <tibble>
+2 ASN00086077  145. -38.0  12.1 moorabbin ai~  94870        3 <tibble>
+3 ASN00086282  145. -37.7 113.  melbourne ai~  94866        3 <tibble>
+\end{CodeOutput}
+\end{CodeChunk}
 
@@
 \begin{CodeInput}
-R> cubble_long <- cubble_nested |> face_temporal()
-R> cubble_long
-\end{CodeInput}
-\begin{CodeOutput}
-# cubble:  date, id [5]: long form
-# bbox:    [115.9764, -32.9342, 133.5407, -12.4239]
-# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id
+R> (cb_long <- cb_nested |> face_temporal())
+\end{CodeInput}
+\begin{CodeOutput}
+# cubble:  key: id [3], index: date, long form
+# extent:  2020-01-01 -- 2020-01-10 [1D], no gaps
+# spatial: long [dbl], lat [dbl], elev [dbl], name [chr], wmo_id
 #   [dbl]
@@
   <chr>       <date>     <dbl> <dbl> <dbl>
-1 ASN00009021 2020-01-01     0  31.9  15.3
-2 ASN00009021 2020-01-02     0  24.9  16.4
-3 ASN00009021 2020-01-03     6  23.2  13  
-4 ASN00009021 2020-01-04     0  28.4  12.4
-5 ASN00009021 2020-01-05     0  35.3  11.6
-# ... with 1,825 more rows
+1 ASN00086038 2020-01-01     0  26.8  11  
+2 ASN00086038 2020-01-02     0  26.3  12.2
+3 ASN00086038 2020-01-03     0  34.5  12.7
+4 ASN00086038 2020-01-04     0  29.3  18.8
+5 ASN00086038 2020-01-05    18  16.1  12.5
+# i 25 more rows
 \end{CodeOutput}
@@
 \begin{CodeInput}
-R> cubble_back <- cubble_long |> face_spatial()
-R> cubble_back
-\end{CodeInput}
-\begin{CodeOutput}
-# cubble:   id [5]: nested form
-# bbox:     [115.9764, -32.9342, 133.5407, -12.4239]
-# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
-  id            lat  long  elev name           wmo_id ts      
+R> cb_back <- cb_long |> face_spatial()
+R> cb_back
+\end{CodeInput}
+\begin{CodeOutput}
+# cubble:   key: id [3], index: date, nested form
+# extent:   [144.8321, -37.98, 145.0964, -37.6655]
+# temporal: prcp [dbl], tmax [dbl], tmin [dbl]
+  id           long   lat  elev name              wmo_id ts      
   <chr>       <dbl> <dbl> <dbl> <chr>              <dbl> <list>  
-1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tibble>
-2 ASN00010311 -31.9  117. 179   york            94623 <tibble>
-3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble>
-4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble>
-5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble>
-\end{CodeOutput}
-\begin{CodeInput}
-R> identical(cubble_nested, cubble_back)
-\end{CodeInput}
-\begin{CodeOutput}
-[1] TRUE
+1 ASN00086038  145. -37.7  78.4 essendon airport   95866 <tibble>
+2 ASN00086077  145. -38.0  12.1 moorabbin airport  94870 <tibble>
+3 ASN00086282  145. -37.7 113.  melbourne airport  94866 <tibble>
+\end{CodeOutput}
+\begin{CodeInput}
+R> identical(cb_nested, cb_back)
+\end{CodeInput}
+\begin{CodeOutput}
+[1] FALSE
 \end{CodeOutput}
@@
 \begin{CodeInput}
-R> cubble_unfold <- cubble_long |> unfold(long, lat)
-R> cubble_unfold
-\end{CodeInput}
-\begin{CodeOutput}
-# cubble:  date, id [5]: long form
-# bbox:    [115.9764, -32.9342, 133.5407, -12.4239]
-# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id
+R> cb_long |> unfold(long, lat)
+\end{CodeInput}
+\begin{CodeOutput}
+# cubble:  key: id [3], index: date, long form
+# extent:  2020-01-01 -- 2020-01-10 [1D], no gaps
+# spatial: long [dbl], lat [dbl], elev [dbl], name [chr], wmo_id
 #   [dbl]
@@
   <chr>       <date>     <dbl> <dbl> <dbl> <dbl> <dbl>
-1 ASN00009021 2020-01-01     0  31.9  15.3  116. -31.9
-2 ASN00009021 2020-01-02     0  24.9  16.4  116. -31.9
-3 ASN00009021 2020-01-03     6  23.2  13    116. -31.9
-4 ASN00009021 2020-01-04     0  28.4  12.4  116. -31.9
-5 ASN00009021 2020-01-05     0  35.3  11.6  116. -31.9
-# ... with 1,825 more rows
-\end{CodeOutput}
-\end{CodeChunk}
-
-\hypertarget{dplyr-methods}{%
-\subsubsection{dplyr methods}\label{dplyr-methods}}
-
-The \pkg{dplyr} package has many tools for wrangling tidy data, many of which are useful in the spatio-temporal analysis.
-The \pkg{cubble} package provides methods that support the use of the following operations in the \pkg{dplyr} package on both the nested and long forms: \texttt{mutate}, \texttt{filter}, \texttt{summarise}, \texttt{select}, \texttt{arrange}, \texttt{rename}, \texttt{left\_join}, and the slice family (\texttt{slice\_*}).
+1 ASN00086038 2020-01-01     0  26.8  11    145. -37.7
+2 ASN00086038 2020-01-02     0  26.3  12.2  145. -37.7
+3 ASN00086038 2020-01-03     0  34.5  12.7  145. -37.7
+4 ASN00086038 2020-01-04     0  29.3  18.8  145. -37.7
+5 ASN00086038 2020-01-05    18  16.1  12.5  145. -37.7
+# i 25 more rows
+\end{CodeOutput}
+\end{CodeChunk}
 
@@
 \begin{CodeInput}
-R> climate_flat_ts <- climate_flat |> 
-+   tsibble::as_tsibble(key = id, index = date)
-R> climate_flat_cb <-  climate_flat_ts |> 
-+   cubble::as_cubble(coords = c(long, lat))
-R> climate_flat_cb
-\end{CodeInput}
-\begin{CodeOutput}
-# cubble:   id [5]: nested form
-# bbox:     [115.9764, -32.9342, 133.5407, -12.4239]
-# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
-  id            lat  long  elev name           wmo_id ts      
-  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>  
-1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tbl_ts>
-2 ASN00010311 -31.9  117. 179   york            94623 <tbl_ts>
-3 ASN00010614 -32.9  117. 338   narrogin        94627 <tbl_ts>
-4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tbl_ts>
-5 ASN00015131 -17.6  134. 220   elliott         94236 <tbl_ts>
-\end{CodeOutput}
+R> cb <- make_cubble(
++   spatial = stations, temporal = meteo_ts, 
++   coords = c(long, lat))
+\end{CodeInput}
 \end{CodeChunk}
@@
 \begin{CodeInput}
-R> climate_flat_cb |> 
-+   mutate(fabletools::features(
-+     ts, tmax, list(tmax_mean = mean, tmax_var = var)
-+     ))
-\end{CodeInput}
-\begin{CodeOutput}
-# cubble:   id [5]: nested form
-# bbox:     [115.9764, -32.9342, 133.5407, -12.4239]
-# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
-  id            lat  long  elev name   wmo_id ts       tmax_~1 tmax_~2
-  <chr>       <dbl> <dbl> <dbl> <chr>   <dbl> <list>     <dbl>   <dbl>
-1 ASN00009021 -31.9  116.  15.4 perth~  94610 <tbl_ts>    25.7   38.6 
-2 ASN00010311 -31.9  117. 179   york    94623 <tbl_ts>    26.2   51.1 
-3 ASN00010614 -32.9  117. 338   narro~  94627 <tbl_ts>    23.7   45.4 
-4 ASN00014015 -12.4  131.  30.4 darwi~  94120 <tbl_ts>    33.1    3.02
-5 ASN00015131 -17.6  134. 220   ellio~  94236 <tbl_ts>    34.6   24.7 
-# ... with abbreviated variable names 1: tmax_mean, 2: tmax_var
+R> cb %>% face_temporal() %>% tsibble::has_gaps()
+\end{CodeInput}
+\begin{CodeOutput}
+# A tibble: 3 x 2
+  id          .gaps
+  <chr>       <lgl>
+1 ASN00086038 FALSE
+2 ASN00086077 FALSE
+3 ASN00086282 FALSE
 \end{CodeOutput}
@@
 
+\begin{CodeChunk}
+\begin{CodeInput}
+R> cb <- make_cubble(
++   spatial = stations_sf, temporal = meteo, 
++   key = id, index = date)
+R> class(cb)
+\end{CodeInput}
+\begin{CodeOutput}
+[1] "spatial_cubble_df" "cubble_df"         "sf"               
+[4] "tbl_df"            "tbl"               "data.frame"       
+\end{CodeOutput}
+\begin{CodeInput}
+R> #cb %>% sf::st_transform(crs = "EPSG:3857")
+\end{CodeInput}
+\end{CodeChunk}
+
 \hypertarget{tidyverse}{%
@@
 It contains daily precipitation, minimum and maximum temperature for 639 weather stations across Australia in 2020.
-The converted sftime object is 13.6830016 times the size of the cubble object:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> climate_aus %>% sftime::st_as_sftime() %>% object.size()
-\end{CodeInput}
-\begin{CodeOutput}
-118243136 bytes
-\end{CodeOutput}
-\begin{CodeInput}
+The converted sftime object is times the size of the cubble object:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> #climate_aus %>% sftime::st_as_sftime() %>% object.size()
 R> climate_aus %>% object.size()
@@
 \begin{CodeOutput}
-8641608 bytes
+8523040 bytes
 \end{CodeOutput}
@@
 \begin{CodeOutput}
-Simple feature collection with 5 features and 7 fields
+Simple feature collection with 5 features and 1 field
 Geometry type: MULTIPOLYGON
@@
 Geodetic CRS:  WGS 84
-    lga_code_2018            lga state_code_2016 state_name_2016
-132         20110     Alpine (S)               2        Victoria
-133         20260    Ararat (RC)               2        Victoria
-134         20570   Ballarat (C)               2        Victoria
-135         20660    Banyule (C)               2        Victoria
-136         20740 Bass Coast (S)               2        Victoria
-    areasqkm_2018 cent_long  cent_lat                       geometry
-132     4788.1568  146.9742 -36.85357 MULTIPOLYGON (((146.7258 -3...
-133     4211.1171  142.8432 -37.47271 MULTIPOLYGON (((143.1807 -3...
-134      739.0321  143.7815 -37.49286 MULTIPOLYGON (((143.6622 -3...
-135       62.5402  145.0851 -37.73043 MULTIPOLYGON (((145.1357 -3...
-136      865.8095  145.5581 -38.50730 MULTIPOLYGON (((145.5207 -3...
-\end{CodeOutput}
-\end{CodeChunk}
-
-The function \code{as_cubble()} is used to create a \code{cubble} object from the two spatial and temporal tables, and requires specifying the arguments \code{key}, \code{index}, and \code{coords} (as described in Section \ref{create}).
+               lga                       geometry
+132     Alpine (S) MULTIPOLYGON (((146.7258 -3...
+133    Ararat (RC) MULTIPOLYGON (((143.1807 -3...
+134   Ballarat (C) MULTIPOLYGON (((143.6622 -3...
+135    Banyule (C) MULTIPOLYGON (((145.1357 -3...
+136 Bass Coast (S) MULTIPOLYGON (((145.5207 -3...
+\end{CodeOutput}
+\end{CodeChunk}
+
+The function \code{make_cubble()} is used to create a \code{cubble} object from the two spatial and temporal tables, and requires specifying the arguments \code{key}, \code{index}, and \code{coords} (as described in Section \ref{create}).
 It will automatically try to match the sites in both tables using the location names and will show a warning message when there are mismatches, as shown below:
@@
 \begin{CodeInput}
-R> cb <- as_cubble(
-+   list(spatial = lga, temporal = covid),
-+   key = lga, index = date, coords = c(cent_long, cent_lat)
-+   )
-\end{CodeInput}
+R> cb <- make_cubble(spatial = lga, temporal = covid)
+\end{CodeInput}
+\begin{CodeOutput}
+Warning: st_centroid assumes attributes are constant over geometries
+\end{CodeOutput}
+\begin{CodeOutput}
+! Some sites in the spatial table don't have temporal information
+\end{CodeOutput}
 \begin{CodeOutput}
@@
 \begin{CodeOutput}
-! Some sites in the spatial table don't have temporal information
-\end{CodeOutput}
-\begin{CodeOutput}
-! Use argument `output = "unmatch"` to check on the unmatched key
+! Use `check_key()` to check on the unmatched key
+The cubble is created only with sites having both spatial and temporal information
 \end{CodeOutput}
@@
 It can be seen that there are two-way mismatches -- LGAs in the COVID data that do not match with LGAs names in the spatial polygon data, and vice versa.
-The mismatches can be identified by using the \code{output = "unmatch"} argument in the \code{as_cubble()} function:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> pair <- as_cubble(
-+   list(spatial = lga, temporal = covid),
-+   key = lga, index = date, coords = c(cent_long, cent_lat),
-+   output = "unmatch"
-+   )
-R> 
-R> pair
+The mismatches can be identified by using the function \code{check_key}:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> (check_res <- check_key(spatial = lga, temporal = covid))
 \end{CodeInput}
@@
 $paired
-# A tibble: 80 x 2
+# A tibble: 78 x 2
   spatial        temporal      
@@
 5 Bass Coast (S) Bass Coast (S)
-# ... with 75 more rows
+# i 73 more rows
+
+$potential_pairs
+# A tibble: 2 x 2
+  spatial             temporal    
+  <chr>               <chr>       
+1 Kingston (C) (Vic.) Kingston (C)
+2 Latrobe (C) (Vic.)  Latrobe (C)
 
 $others
-$others$temporal
-[1] "Interstate" "Overseas"   "Unknown"
-
 $others$spatial
@@
 [2] "Migratory - Offshore - Shipping (Vic.)"
+
+$others$temporal
+[1] "Interstate" "Overseas"   "Unknown"   
 \end{CodeOutput}
@@
 \begin{CodeInput}
-R> lga <- lga |>
+R> lga2 <- lga |>
 +   mutate(lga = ifelse(lga == "Kingston (C) (Vic.)", "Kingston (C)", lga),
 +          lga = ifelse(lga == "Latrobe (C) (Vic.)", "Latrobe (C)", lga)) |>
-+   filter(!lga %in% pair$others$spatial)
++   filter(!lga %in% check_res$others$spatial)
 R> 
-R> covid <- covid |> filter(!lga %in% pair$others$temporal)
+R> covid2 <- covid |> filter(!lga %in% check_res$others$temporal)
 R> 
-R> cb <- as_cubble(data = list(spatial = lga, temporal = covid),
-+                 key = lga, index = date, coords = c(cent_long, cent_lat))
-\end{CodeInput}
+R> (cb <- make_cubble(spatial = lga2, temporal = covid2))
+\end{CodeInput}
+\begin{CodeOutput}
+# cubble:   key: lga [80], index: date, nested form, [sf]
+# extent:   [140.961682, -39.1591895, 149.976291, -33.9804256]
+# temporal: source [chr], n [int], roll_mean [dbl]
+  lga             long   lat                       geometry ts        
+  <chr>          <dbl> <dbl>             <MULTIPOLYGON [°]> <list>    
+1 Alpine (S)      147. -36.9 (((146.7258 -36.45922, 146.80~ <gropd_ts>
+2 Ararat (RC)     143. -37.5 (((143.1807 -37.73152, 143.17~ <gropd_ts>
+3 Ballarat (C)    144. -37.5 (((143.6622 -37.57241, 143.66~ <gropd_ts>
+4 Banyule (C)     145. -37.7 (((145.1357 -37.74091, 145.13~ <gropd_ts>
+5 Bass Coast (S)  146. -38.5 (((145.5207 -38.30667, 145.54~ <gropd_ts>
+# i 75 more rows
+\end{CodeOutput}
 \end{CodeChunk}
@@
 The GHCN provides daily climate measures from stations across the world.
-The data used here (\code{historical_tmax}) is a subset extracted using the package \pkg{rnoaa} \citep{rnoaa}, containing the records of maximum temperature for 236 Australian stations from 1859 through 1969 and provides information also on the latitude, longitude and elevation of each of the stations.
+The data used here (\code{historical_tmax}) is a subset extracted using the package \pkg{rnoaa} \citep{rnoaa}, containing the records of maximum temperature for 237 Australian stations from \ensuremath{\infty{}} through \ensuremath{-\infty{}} and provides information also on the latitude, longitude and elevation of each of the stations.
 The goal of this example is to compare the monthly average maximum temperature between two periods, 1971-1975 and 2016-2020, for stations in Victoria and New South Wales (NSW), using a glyph map.
@@
 +   face_temporal() |> 
-+   mutate(month = lubridate::month(date),
-+          group = as.factor(
-+            ifelse(lubridate::year(date) > 2015,
-+            "2016 ~ 2020", "1971 ~ 1975"))) |>
-+   group_by(month, group) |> 
++   group_by(
++     yearmonth = tsibble::make_yearmonth(
++       year = ifelse(lubridate::year(date) > 2015, 2016, 1971),
++       month = lubridate::month(date))
++     )|>
 +   summarise(tmax = mean(tmax, na.rm = TRUE))
@@
 \begin{CodeOutput}
- [1] 24 24 12 24 24 24 24 24 24 24
+ [1] 12 12 12 13 19 19 20 24 24 24
 \end{CodeOutput}
@@
 
-\begin{CodeChunk}
-\begin{CodeOutput}
-# cubble:   id [8]: nested form
-# bbox:     [144.5203, -37.727706, 146.059205, -36.5523]
-# temporal: date [date], matched_var [dbl]
-  id          name        lat  long type   dist group ts       n_match
-  <chr>       <chr>     <dbl> <dbl> <chr> <dbl> <int> <list>     <int>
-1 230200      MARIBYRN~ -37.7  145. river  6.17     6 <tibble>      19
-2 404207      HOLLAND ~ -36.6  146. river  8.54    10 <tibble>      21
-3 405234      SEVEN CR~ -36.9  146. river  6.15     5 <tibble>      21
-4 406213      CAMPASPE~ -37.0  145. river  1.84     1 <tibble>      18
-5 ASN00082042 strathbo~ -36.8  146. clim~  6.15     5 <tibble>      21
-# ... with 3 more rows
-\end{CodeOutput}
-\end{CodeChunk}
-
 Figure \ref{fig:matching} shows four matched pairs on the map (a) and standardized data as time series (b). The expected concurrent increase in precipitation and water level can be seen clearly.
-
-\begin{CodeChunk}
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/research/paper-cubble/figures/matching-1}
-
-}
-
-\caption[Weather stations and river gauges with matched pairs labelled on the map (a) and plotted across time (b)]{Weather stations and river gauges with matched pairs labelled on the map (a) and plotted across time (b).
-Precipitation and water level have been standardised between 0 and 1 to be displayed on the same scale.
-The water level reflects the increase in precipitation.
-The numbers (1, 5, 6, 10) indicate the group index derived from spatial matching, only those that were selectd by temporal matching are shown here.}\label{fig:matching}
-\end{figure}
-\end{CodeChunk}
 

### 2023-05-29 review/B-response.Rmd
rename review files

---
title: "Response to reviewers article 4780"
subtitle: "cubble: An R Package for Organizing and Wrangling Multivariate Spatio-temporal Data"
author: "H. Sherry Zhang, Dianne Cook, Ursula Laa, Nicolas Langrené, Patricia Menéndez"
date: '2023-03-04'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Response to B-review_1_1

## The Paper

The package and the proposed data structures could certainly constitute a valuable base for the spatio- temporal statistics tooling. Unfortunately the paper does not clearly describe the data structure, nor the associated functionality. I had to read the package vignettes, run the attached code and even look into the package’s code in order to fully understand the data structure.

**\textcolor{purple}{Re: We thank the reviewer for the support on our proposed data structure and the suggestion to improve the paper and the package documentation.}**

Given the small size of the package, the paper feels a bit on the heavy side. Most of the code in the paper is not self-contained which makes it rather difficult to follow. Examples at the end of the paper don’t elucidate the inner workings of the package but rather advertise the functionality of other, mostly visualization, packages.

**\textcolor{purple}{Re: We select the code and output relevant to demonstrate the cubble package in the examples while leaving the rest in the github repository, where users can look up for more details: https://github.com/huizezhang-sherry/paper-cubble. [on the visualisation packages: xxx]}. After addressing all the issues raised by the reviewers, the paper is now xxx pages, including 4 pages of reference and required affiliation.**

More fundamentally, the design of the package has an immediately apparent drawback. The cubble class is used to represent two incompatible data types - long and nested. Internally, the two data types are distinguished by the internal attribute “form”. In other words, the authors use an attribute of an object to mimic the functionality of a sub-class.

**\textcolor{purple}{Re: [TODO] }**

A more natural implementation would be start with two distinct data structures, both sub-classes of the abstract cubble_df class, c("spatial_cubble_df", "cubble_df") and c("spatial_temporal_df", "cubble_df"). Such a generic implementation would allow for natural extension of the functionality trough dedicated methods. Currently an ever-present check on “form” attribute is necessary throughout the code base. To make matters worse, as_cubble currently produces yet another data structure - a list of paired matches, as can be seen from example 5.1 in the paper.

**\textcolor{purple}{Re: Now as\_cubble will consistantly produce a cubble object, or error messages if it fails. When there are unmatched pairs, as\_cubble will create with cubble with sites that have both spatial and temporal information and instruct users to use check\_key() to check on key matching. Below is a much simplified example to show the current workflow:}**

``` r
library(cubble)
library(dplyr)
#> 
#> Attaching package: 'dplyr'
#> The following objects are masked from 'package:stats':
#> 
#>     filter, lag
#> The following objects are masked from 'package:base':
#> 
#>     intersect, setdiff, setequal, union
three_sites <-  stations %>% head(3)
five_ts <- climate %>% filter(lubridate::month(date) == 1)
cb <- as_cubble(
  list(spatial = three_sites, temporal = five_ts),
  key = id, index = date, coords = c(long, lat)
)
#> ! Some sites in the temporal table don't have spatial information
#> ! Use `check_key()` to check on the unmatched key
#> The cubble is created only with sites having both spatial and temporal information
check_key(spatial = three_sites, temporal = five_ts)
#> $paired
#> # A tibble: 3 × 2
#>   spatial     temporal   
#>   <chr>       <chr>      
#> 1 ASN00009021 ASN00009021
#> 2 ASN00010311 ASN00010311
#> 3 ASN00010614 ASN00010614
#> 
#> $potential_pairs
#> # A tibble: 0 × 0
#> 
#> $others
#> $others$temporal
#> [1] "ASN00014015" "ASN00015131"
#> 
#> $others$spatial
#> character(0)
```

<sup>Created on 2023-05-23 with [reprex v2.0.2](https://reprex.tidyverse.org)</sup>

Relatedly, the rationale for the parallel naming convention long/nested and temporal/spatial is not entirely clear. It seems to me that the semantically unambiguous temporal/spatial could be used throughout without ambiguity, thus resolving the terminological redundancy.

**\textcolor{purple}{Re: The naming of long/nested comes from the tidy data concept where the "spatial" form has the nested temporal information and the "temporal" form uses the long form data from the tidy data. We understand it could be more intuitive to refer to them as spaital/ temporal and we add them as a "common name" to the form in the definitional section (now section 2). This naming is also used in the paper when it is more straightforward to use spatial/temporal than nested/long.}**

My recommendation would be a combination of the following:

  * Rework the definitional parts (sections 2 and 3) by following more closely the “design” vignette. **\textcolor{purple}{Re: [TODO: follow the design vignette]}**
  * Describe at a glance the core functions which operate on the class and their motivation. **\textcolor{purple}{Re: Methods available for a cubble object is now printed in xxx, before introducing the main functionality: face\_temporal(), face\_spatial(), and unfold().}**
  * Rework section 4 by making it more concise and provide links to sections in supplementary material or dedicated vignettes. **\textcolor{purple}{Re: }**
  * Rework examples section 5 into “Applications”. Briefly describe the applications and provide and refer to dedicated and self-contained vignettes. **\textcolor{purple}{Re: xxx}**

## The Package

The package is generally badly documented. The meaning of the arguments is almost never clear from the documentation alone. For example the documentation of as_cubble states:

**\textcolor{purple}{Re: Documentation of the package has been improved with a focus on clarifying the input arguemnt types, providing examples, [more] }**

* What objects are supported as input data? **\textcolor{purple}{Re: the objects supported are 1) list with two elements named spatial and temporal, 2) data frame/ tibble, 3) netcdf objects, 4) stars objects, and 5) sftime objects. }**
* What is the assumption of the “nestedness” of the input data. **\textcolor{purple}{Re: }**
* What is the accepted data type of key, index and coords? **\textcolor{purple}{Re: Key supports character and factor. Index supports base R classes Date, POSIXlt, POSIXct and tsibble's yearmonth, yearweek, and yearquarter class. Coords support both projected/ unprojected coordinates in columns (can be omitted if created from sf and its subclass). }**

## Further suggestion

Consider not throwing an error when face_spatial is applied to the spatial object. This would allow for generic code where the “form” of the input the data is not known or does not matter.

**\textcolor{purple}{Re: Adopted. Now when applying face\_spatial to a nested cubble (a spatial object), the cubble will be printed as it is with a message signalling it is already in the nested form. see the reprex below. Same for face\_temporal on a long cubble}**

``` r
library(cubble)
nested <- climate_flat |>
  as_cubble(key = id, index = date, coords = c(long, lat))
class(nested) 
#> [1] "cubble_df"  "rowwise_df" "tbl_df"     "tbl"        "data.frame"
face_spatial(nested) 
#> i The cubble is already in the nested form
#> # cubble:   id [5]: nested form
#> # bbox:     [115.9764, -32.9342, 133.5407, -12.4239]
#> # temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
#>   id            lat  long  elev name           wmo_id ts                
#>   <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>            
#> 1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tibble [366 × 4]>
#> 2 ASN00010311 -31.9  117. 179   york            94623 <tibble [366 × 4]>
#> 3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble [366 × 4]>
#> 4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble [366 × 4]>
#> 5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble [366 × 4]>
```

<sup>Created on 2023-05-22 with [reprex v2.0.2](https://reprex.tidyverse.org)</sup>

It would be useful to print a few rows of the spatial attribute of the “long” object.

**\textcolor{purple}{Re: xxx. we could have a function to print }**

Section 3.6 is missing `tidyr` package which is especially relevant here given its nesting and unnesting operations https://tidyr.tidyverse.org/articles/nest.html

**\textcolor{purple}{Re: xxx}**

More comments inlined in the pdf.

  - on p4 "The following sections explain their roles, why the new cubble class is needed
and how the package relates to existing packages for spatial and temporal data analysis." The paper is a bit on the verbose side. Such comments could be safely removed. **\textcolor{purple}{Re: sentence removed.}**

  - on p5 section 3.1, "The arguments key and index follow the wording in the tsibble package to describe the temporal order and multiple series while coords specifies the spatial location of each site.", Unclear what `key` and `index` mean. Isn't key a spatial identifier and index a temporal one? **\textcolor{purple}{Re: [TODO] Yes, key is the spatial identifier and index is the temporal one. They are now explained in the text as well as referenced to the tsibble package.}**

  - on p5 section 3.1, the word "wording" is highlighted in the sentence "The arguments key and index follow the wording in the tsibble package to ...", conventions. **\textcolor{purple}{Re: *wording* changed to *conventions*}**

  - on p6 section 3.2, 'The cubble class also provides a long form, which expands the ts column and temporarily “hides” the spatial variables.', confusing: `hides` means storing it as an attribute. **\textcolor{purple}{Re: The sentence is now changed to ..., which expands the ts column and stores the spatial variables as an attribute}**

  - on p6 section 3.2, I think it would be a good idea to clearly describe the two data structures in an earlier section. By this point it's still not clear what those are and what the meaning of key and index is. **\textcolor{purple}{Re: addressed in bullet point "on p5 section 3.1"}**

  - on p6, section 3.2, yet another reason to call this data structure "cubble_temporal" instead of the semantically unobvious "long". **\textcolor{purple}{Re: see paragraph 4 under section The paper }**

  - on p8 section 3.5, Isn't the root reason the memory efficiency? That is, the spatial component cannot be represented as part of the nested tibble without repetition. **\textcolor{purple}{Re: [TODO - also compare with sftime] added.}**

  - on p8 section 3.6, The missing elephant is tidyr which has nesting and unnesting operations  https://tidyr.tidyverse.org/articles/nest.html. **\textcolor{purple}{Re: addressed under section Further suggestion}**

# Response to B-review_2_1

## The Manuscript

The manuscript is in general clearly structured. However, an important reference to the R-package sftime is missing that also deals with the representation of spatiotemporal data. It might be beneficial to provide illustrative examples in the manuscript that clearly compare between a cubble representation and representations of existing packages. Which data sets/structures cannot (or with a greater effort) be represented with existing R-packages such as e.g. stars and sftime?

**\textcolor{purple}{Re: [TODO] The section 3.5 are extended to also discuss why not just use existing data structure. This includes the comparision with sftime on memory efficiency, with stars on representing spatio-temporal data in an array format with illustrative examples}**

For spatial data, the coordinate reference system is essential metadata information. In one example of the manuscript, the input data does have information on the CRS but it is not discussed in the manuscript how that is handled within cubble. How would coordinate transformations be handled with cubble?

**\textcolor{purple}{Re: We understand spatial data is, strictly speaking, not uniquely identified if CRS information is not provided, while many wild data still do not contain this crucial piece of information. In cubble, if the spatial data is from an sf object, the cubble will retain the sf class and users are welcomed to use functions in sf to perform coordinate transformations. If not, the function cubble::add\_geometry\_column() can be used to construct the geometry column from long and lat, and the resulting cubble will have the sf class. if CRS is missing, cubble::add\_geometry\_column() will by default assume OGC:CRS84 (WGS84).}**

The use-case of temporal matching based on features of the time series appears an interesting, but also a very specific one. How could this be generalized to a more generic approach?

**\textcolor{purple}{Re: xxx}**

Additionally, several typos and language issues arise and limit the readability of the manuscript. Some examples are:

- doubled/missing words/none correct sentences: 
  - “*components* spatio-temporal *components*” **\textcolor{purple}{Re: to "... spatio-temporal components"}**
  - “... fits works ...”  **\textcolor{purple}{Re: to ".. works well with ..."}**
  - “... be activate rows ...” **\textcolor{purple}{Re: to "... will activate rows ..."}**
  - “... highlighted *in* the ...” **\textcolor{purple}{Re: to "... highlighted on the map"}**
  - “. . . using 2020 measurements using match_sites() function.”  **\textcolor{purple}{Re: to "... using the match\_sites() function"}**
  - “An example of this using is included in the Appendix” **\textcolor{purple}{Re: to "... this use ..."}**
  - “. . . the data in -a- multiple -of- ways on-the-fly”  **\textcolor{purple}{Re: to "... in multiple ways ..."}**

- Surprising references: in Section 4.4 it says “. . . Glyph maps (Section 3.4)”, but Section 3.4 says “. . . glyph maps will be explained in Section 4.4”. **\textcolor{purple}{Re: the looped reference in section 3.3 and 3.4 is replaced with (Wickham et al. 2012)}.**

- Should “. . . it’s temperatures are more *consistent*” rather be “. . . it’s temperatures are more *constant*”? **\textcolor{purple}{Re: to "more constant"}**

- Typos:
  - “. . . spatial and *tmeporal* information are available.” **\textcolor{purple}{Re: to "temporal"}**
  - The polar vortex, signalled by the high *speicfic* humidity, splits into two on 2002-09-26 and *further-s- split_s_* into four on 2002-10-04. **\textcolor{purple}{Re: to "specific", "further splits"}**
  - the data in *-a- multiple -of-* ways on-the-fly. **\textcolor{purple}{Re:  to "multiple ways"}**

## The package

I have been a bit puzzled by the “print” of a nested cubble in its temporal face. The row # temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl] appears at first sight (also in comparison with the print of the spatial face), as if the temporal domain is given by all those variables. To me, a notion such as

`# temporal: date [date]`

`# variables: prcp [dbl], tmax [dbl], tmin [dbl]`

would have been more intuitive. Possibly also including the temporal range (as for the spatial domain its bbox).

**\textcolor{purple}{Re: Thanks reviewers for the suggestion and this has now been incorporated in the package. The second line in the header, when printed in the long (temporal) form will now show the start and the end date, the date interval, and whether there are gaps in the data. See the reprex example below: }**

``` r
library(cubble)
face_temporal(climate_aus)
#> # cubble:  key: id [639], index: date, long form
#> # extent:  2020-01-01 -- 2020-12-31 [1D], has gaps!
#> # spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]
#>    id          date        prcp  tmax  tmin
#>    <chr>       <date>     <dbl> <dbl> <dbl>
#>  1 ASN00001006 2020-01-01   164  38.3  25.3
#>  2 ASN00001006 2020-01-02     0  40.6  30.5
#>  3 ASN00001006 2020-01-03    16  39.7  27.2
#>  4 ASN00001006 2020-01-04     0  38.2  27.3
#>  5 ASN00001006 2020-01-05     2  39.3  26.7
#>  6 ASN00001006 2020-01-06    60  32.9  25.6
#>  7 ASN00001006 2020-01-07   146  34.1  25.5
#>  8 ASN00001006 2020-01-08    40  36.6  26.2
#>  9 ASN00001006 2020-01-09     0  38.2  27.6
#> 10 ASN00001006 2020-01-10     0  38.9  29.7
#> # … with 230,750 more rows
```

<sup>Created on 2023-05-23 with [reprex v2.0.2](https://reprex.tidyverse.org)</sup>

In the manuscript and in the reproducible-script.R, a data set `historical_tmax` is introduced. In the package, the corresponding data set seems to be `tmax_hist`.

**\textcolor{purple}{Re: [TODO] The two datasets are related but not the same. The data historical\_tmax contains daily weather stations data for the two periods: 1971-1975 and 2016-2020 and amounts to 3.2MB, which is not suitable to include in the package given the size restriction. The tmax\_hist summarises the daily data into monthly to produce the dataset suitable to include in the package. The documentation to tmax\_hist now include more details on this. }**

The manuscript cannot be reproduced, as the reproducible-script.R (part of the submission to JSS) only uses a subset, but does not provide a separate result file that would allow to compare the reproduced outputs with the desired output for the subset.

**\textcolor{purple}{Re: A result file in html has been submitted this time.}**

section cubble object under 3. the cubble package, on 1) the conceptual framework, 2)  the nested and long form as per the design vig

print of methods

section 4/5: work on the text to make it more concise

connnecting dat ais gneerally a difficult task from different sources in spatial temporal data, point it to the future work

In the end talk to the users to check the vignette for the full code for examples. match titles.

### 2023-05-29 4-3-matching.Rmd
rename files

--- 2023-03-25 5-3-river-level-rainfall.Rmd
@@
 ```{r}
+
 load(here::here("data/climate_full.rda"))
@@
 
-river <- river |> mutate(type = "river")
+river <- cubble::river %>% unnest(ts) %>% as_cubble(key = station, index = date, coords = c(long, lat)) |> mutate(type = "river")
 ```
@@
 
-```{r echo = TRUE}
+```{r echo = TRUE, eval = FALSE}
 res <- match_sites(
@@
 
-```{r}
+```{r eval = FALSE}
 res |> print()
@@
 
-```{r matching, out.width="100%", fig.height = 5, fig.width = 10, fig.cap="Weather stations and river gauges with matched pairs labelled on the map (a) and plotted across time (b).
+```{r matching, eval = FALSE, out.width="100%", fig.height = 5, fig.width = 10, fig.cap="Weather stations and river gauges with matched pairs labelled on the map (a) and plotted across time (b).
 Precipitation and water level have been standardised between 0 and 1 to be displayed on the same scale.

### 2023-05-29 jss/main.tex
rename files

@@
 \begin{CodeInput}
-R> make_cubble(
-+   spatial = stations_sf, temporal = meteo, key = id, index = date
-+   )
+R> make_cubble(spatial = stations_sf, temporal = meteo, 
++             key = id, index = date)
 R> 
-R> make_cubble(
-+   spatial = stations, temporal = meteo_ts, coords = c(long, lat)
-+   )
+R> make_cubble(spatial = stations, temporal = meteo_ts, 
++             coords = c(long, lat))
 R> 
@@
 \begin{CodeInput}
-R> cb_back <- cb_long |> face_spatial()
-R> cb_back
+R> (cb_back <- cb_long |> face_spatial())
 \end{CodeInput}
@@
 \begin{CodeInput}
-R> identical(cb_nested, cb_back)
-\end{CodeInput}
-\begin{CodeOutput}
-[1] FALSE
-\end{CodeOutput}
+R> #identical(cb_nested, cb_back)
+\end{CodeInput}
 \end{CodeChunk}

### 2023-05-30 2-cubble.Rmd
subsection coercing from foreign objects

--- 2023-05-29 3-the-cubble-pkg.Rmd
@@
 
+The arguments \code{key} and \code{index} follow the conventions in the \pkg{tsibble} package to describe the temporal order and multiple series while \code{coords} specifies the spatial location of each site.
+
 ```{r}
@@
 
-### Create from separate spatial and temporal table
+### Creating from separate spatial and temporal table
 
@@
             coords = c(long, lat))
+```
 
+```{r echo = TRUE}
 make_cubble(spatial = stations_sf, temporal = meteo_ts)
@@
 
-### Coerce from other objects
+### Coercing from foregin objects
 
-  - import different data: an expand data, spatial + temporal, netcdf, stars, sftime
+Spatio-temporal data in other foreign objects can be coerced into cubble with the function \code{as_cubble()}.
+This includes casting from a \code{tibble} or \code{data.frame} with both spatial and temporal information, a netCDF object, a \code{stars} object [@stars], and a \code{sftime} object [@sftime].
+The two examples below shows the casting from a tibble and a netcdf object to a cubble object.
 
-The function \code{as_cubble()} is used to create a \code{cubble} object with three arguments: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order (longitude, latitude).
-The arguments \code{key} and \code{index} follow the conventions in the \pkg{tsibble} package to describe the temporal order and multiple series while \code{coords} specifies the spatial location of each site.
-The code below creates a \code{cubble} object out of \code{climate_flat} (a single \code{tibble}) with \code{id} as the key, \code{date} as the index, and \code{c(long, lat)} as the coordinates:
+The dataset \code{climate_flat} joins the spatial data, \code{stations}, with the temporal data, \code{meteo}, into a single tibble object and it can be coerced into a cubble using:
 
@@
 
-A \code{cubble} class can be created from various common spatio-temporal data formats, including \code{tibble}, \code{tsibble}, and \code{sf} when both spatial and temporal information are available.
-Section \@ref(covid) describes converting multiple tables into a \code{cubble} object and Section \@ref(netcdf) illustrates how to convert a NetCDF object.
+The NetCDF data is another format commonly used for storing spatio-temporal data.
+In \proglang{R}, packages for wrangling NetCDF data include a high-level R interface: \pkg{ncdf4} [@ncdf4], a low-level interface that calls a C-interface: \pkg{RNetCDF} [@rnetcdf], and a tidyverse implementation: \pkg{tidync} [@tidync].
+The code below casts a netCDF object in the \code{ncdf4} class into a cubble object:
 
-*netcdf*
+```{r echo = TRUE}
+path <- system.file("ncdf/era5-pressure.nc", package = "cubble")
+raw <- ncdf4::nc_open(path)
+as_cubble(raw)
+```
 
-The NetCDF data is another format commonly used for storing spatio-temporal data.
-It has two main components: *dimension* for defining the spatio-temporal grid (longitude, latitude, and time) and *variable* that populates the defined grid.
-Attributes  can be associated with dimensions or variables.
-Because there can be many different styles of representing this information there is a metadata convention [@hassell_data_2017] to standardize the format of the attributes.
-A few packages in \proglang{R} exist for manipulating NetCDF data and these include a high-level R interface: \pkg{ncdf4} [@ncdf4], a low-level interface that calls a C-interface: \pkg{RNetCDF} [@rnetcdf], and a tidyverse implementation: \pkg{tidync} [@tidync].
-
-The \pkg{cubble} package provides an \code{as_cubble()} method to coerce the \code{ncdf4} class from the \pkg{ncdf4} package into a \code{cubble} object.
-It maps each combination of longitude and latitude into an \code{id} as the \code{key}:
+Sometimes, one may want to read in a subset of the netcdf data and the argument \code{vars}, \code{long_range} and \code{lat_range} can be used to subset on the variable and the grid resolution:
 
 ```{r echo = TRUE, eval = FALSE}
-dt <- as_cubble(my_ncdf, vars = c("q", "z"))
-```
-
-Sometimes NetCDF data can be quite large, and it is best to subset the data when converting to a \code{cubble} object.
-We would recommend reducing to about 300 $\times$ 300 grid points for three daily variables in one year.
-A 300 by 300 spatial grid can be a bounding box of [100, -80, 180, 0] at 0.25 degree resolution or a global bounding box [-180, -90, 180, -90] at 1 degree resolution.
-The size of spatial grid can be reduced if longer time periods or more variables are needed, through the arguments \code{long_range} and \code{lat_range}:
-
-```{r eval = FALSE, echo = TRUE}
-dt <- as_cubble(
-  my_ncdf, vars = c("q", "z"),
+as_cubble(
+  raw, vars = "q",
   long_range = seq(-180, 180, 1), lat_range = seq(-90, 90, 1)
@@
 
-*stars*
-
-*sftimes*
+We would recommend reducing to about 300 $\times$ 300 grid points for three daily variables in one year.
+A 300 by 300 spatial grid can be a bounding box of [100, -80, 180, 0] at 0.25 degree resolution or a global bounding box [-180, -90, 180, -90] at 1 degree resolution.
 

### 2023-05-30 jss/main.tex
subsection coercing from foreign objects

@@
 
+The arguments \code{key} and \code{index} follow the conventions in the \pkg{tsibble} package to describe the temporal order and multiple series while \code{coords} specifies the spatial location of each site.
+
 \begin{CodeChunk}
@@
 
-\hypertarget{create-from-separate-spatial-and-temporal-table}{%
-\subsubsection{Create from separate spatial and temporal table}\label{create-from-separate-spatial-and-temporal-table}}
+\hypertarget{creating-from-separate-spatial-and-temporal-table}{%
+\subsubsection{Creating from separate spatial and temporal table}\label{creating-from-separate-spatial-and-temporal-table}}
 
@@
 +             coords = c(long, lat))
-R> 
+\end{CodeInput}
+\end{CodeChunk}
+
+\begin{CodeChunk}
+\begin{CodeInput}
 R> make_cubble(spatial = stations_sf, temporal = meteo_ts)
 \end{CodeInput}
-\end{CodeChunk}
-
-\hypertarget{coerce-from-other-objects}{%
-\subsubsection{Coerce from other objects}\label{coerce-from-other-objects}}
-
-\begin{itemize}
-\tightlist
-\item
-  import different data: an expand data, spatial + temporal, netcdf, stars, sftime
-\end{itemize}
-
-The function \code{as_cubble()} is used to create a \code{cubble} object with three arguments: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order (longitude, latitude).
-The arguments \code{key} and \code{index} follow the conventions in the \pkg{tsibble} package to describe the temporal order and multiple series while \code{coords} specifies the spatial location of each site.
-The code below creates a \code{cubble} object out of \code{climate_flat} (a single \code{tibble}) with \code{id} as the key, \code{date} as the index, and \code{c(long, lat)} as the coordinates:
+\begin{CodeOutput}
+# cubble:   key: id [3], index: date, nested form, [sf]
+# extent:   [144.8321, -37.98, 145.0964, -37.6655]
+# temporal: prcp [dbl], tmax [dbl], tmin [dbl]
+  id           long   lat  elev name  wmo_id            geometry
+  <chr>       <dbl> <dbl> <dbl> <chr>  <dbl>         <POINT [°]>
+1 ASN00086038  145. -37.7  78.4 esse~  95866 (144.9066 -37.7276)
+2 ASN00086077  145. -38.0  12.1 moor~  94870   (145.0964 -37.98)
+3 ASN00086282  145. -37.7 113.  melb~  94866 (144.8321 -37.6655)
+# i 1 more variable: ts <list>
+\end{CodeOutput}
+\end{CodeChunk}
+
+\hypertarget{coercing-from-foregin-objects}{%
+\subsubsection{Coercing from foregin objects}\label{coercing-from-foregin-objects}}
+
+Spatio-temporal data in other foreign objects can be coerced into cubble with the function \code{as_cubble()}.
+This includes casting from a \code{tibble} or \code{data.frame} with both spatial and temporal information, a netCDF object, a \code{stars} object \citep{stars}, and a \code{sftime} object \citep{sftime}.
+The two examples below shows the casting from a tibble and a netcdf object to a cubble object.
+
+The dataset \code{climate_flat} joins the spatial data, \code{stations}, with the temporal data, \code{meteo}, into a single tibble object and it can be coerced into a cubble using:
 
@@
 
-A \code{cubble} class can be created from various common spatio-temporal data formats, including \code{tibble}, \code{tsibble}, and \code{sf} when both spatial and temporal information are available.
-Section \ref{covid} describes converting multiple tables into a \code{cubble} object and Section \ref{netcdf} illustrates how to convert a NetCDF object.
-
-\emph{netcdf}
-
 The NetCDF data is another format commonly used for storing spatio-temporal data.
-It has two main components: \emph{dimension} for defining the spatio-temporal grid (longitude, latitude, and time) and \emph{variable} that populates the defined grid.
-Attributes can be associated with dimensions or variables.
-Because there can be many different styles of representing this information there is a metadata convention \citep{hassell_data_2017} to standardize the format of the attributes.
-A few packages in \proglang{R} exist for manipulating NetCDF data and these include a high-level R interface: \pkg{ncdf4} \citep{ncdf4}, a low-level interface that calls a C-interface: \pkg{RNetCDF} \citep{rnetcdf}, and a tidyverse implementation: \pkg{tidync} \citep{tidync}.
-
-The \pkg{cubble} package provides an \code{as_cubble()} method to coerce the \code{ncdf4} class from the \pkg{ncdf4} package into a \code{cubble} object.
-It maps each combination of longitude and latitude into an \code{id} as the \code{key}:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> dt <- as_cubble(my_ncdf, vars = c("q", "z"))
-\end{CodeInput}
-\end{CodeChunk}
-
-Sometimes NetCDF data can be quite large, and it is best to subset the data when converting to a \code{cubble} object.
+In \proglang{R}, packages for wrangling NetCDF data include a high-level R interface: \pkg{ncdf4} \citep{ncdf4}, a low-level interface that calls a C-interface: \pkg{RNetCDF} \citep{rnetcdf}, and a tidyverse implementation: \pkg{tidync} \citep{tidync}.
+The code below casts a netCDF object in the \code{ncdf4} class into a cubble object:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> path <- system.file("ncdf/era5-pressure.nc", package = "cubble")
+R> raw <- ncdf4::nc_open(path)
+R> as_cubble(raw)
+\end{CodeInput}
+\begin{CodeOutput}
+# cubble:   key: id [26565], index: time, nested form
+# extent:   [113, -53, 153, -12]
+# temporal: q [dbl], z [dbl]
+     id  long   lat ts              
+  <int> <dbl> <dbl> <list>          
+1     1  113    -12 <tibble [6 x 3]>
+2     2  113.   -12 <tibble [6 x 3]>
+3     3  114.   -12 <tibble [6 x 3]>
+4     4  114.   -12 <tibble [6 x 3]>
+5     5  114    -12 <tibble [6 x 3]>
+# i 26,560 more rows
+\end{CodeOutput}
+\end{CodeChunk}
+
+Sometimes, one may want to read in a subset of the netcdf data and the argument \code{vars}, \code{long_range} and \code{lat_range} can be used to subset on the variable and the grid resolution:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> as_cubble(
++   raw, vars = "q",
++   long_range = seq(-180, 180, 1), lat_range = seq(-90, 90, 1)
++   )
+\end{CodeInput}
+\end{CodeChunk}
+
 We would recommend reducing to about 300 \(\times\) 300 grid points for three daily variables in one year.
 A 300 by 300 spatial grid can be a bounding box of {[}100, -80, 180, 0{]} at 0.25 degree resolution or a global bounding box {[}-180, -90, 180, -90{]} at 1 degree resolution.
-The size of spatial grid can be reduced if longer time periods or more variables are needed, through the arguments \code{long_range} and \code{lat_range}:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> dt <- as_cubble(
-+   my_ncdf, vars = c("q", "z"),
-+   long_range = seq(-180, 180, 1), lat_range = seq(-90, 90, 1)
-+   )
-\end{CodeInput}
-\end{CodeChunk}
-
-\emph{stars}
-
-\emph{sftimes}
 

### 2023-05-30 2-cubble.Rmd
subsection 2.2 done

@@
 
-## The cubble object
+## The cubble object {#object}
 
@@
 
-### Creating from separate spatial and temporal table
+### Creating from separate spatial and temporal tables
+
+Spatio-temporal data may arrives in separate tables for analysts.
+For example, in climate data, analysts may initially receive station data containing geographic location information, variables recorded and their recording periods.
+They can then query the temporal variables using the stations of interest to obtain the corresponding temporal data.
+Alternatively, analyses may begin as purely spatial or temporal, and analysts may obtain additional temporal or spatial data to expand the result to spatio-temporal.
+
+The function \code{make_cubble()} composes a \code{cubble} object from a spatial table (\code{spatial}) and a temporal table (\code{temporal}) with the three attributes introduced in the subsection \@ref(object):  \code{key}, \code{index}, and \code{coords}.
+The following code creates the nested \code{cubble} object shown previously:
+
+```{r echo = TRUE, eval = FALSE}
+make_cubble(spatial = stations, temporal = meteo,
+            key = id, index = date, coords = c(long, lat))
+```
+
+The \code{key} and \code{index} argument comes from the same arguments from the \code{tsibble} objects and \code{coords} derives from the geometry column in an \code{sf} object.
+Hence, the corresponding argument can be safely omitted, if the spatial data is an \code{sf} object, i.e.
+\code{stations_sf}, or the temporal data is a \code{tsibble} object, i.e.
+\code{meteo_ts}.
+The \code{sf} and \code{tsibble} class from the input will be carried over to the \code{cubble} object:
 
 ```{r echo = TRUE}
-make_cubble(
-  spatial = stations, temporal = meteo,
-  key = id, index = date, coords = c(long, lat)
-  )
-```
-
-```{r echo = TRUE, eval = FALSE}
-make_cubble(spatial = stations_sf, temporal = meteo, 
-            key = id, index = date)
-
-make_cubble(spatial = stations, temporal = meteo_ts, 
-            coords = c(long, lat))
-```
-
-```{r echo = TRUE}
-make_cubble(spatial = stations_sf, temporal = meteo_ts)
+(res <- make_cubble(spatial = stations_sf, temporal = meteo_ts))
+class(res)
+class(res$ts[[1]])
 ```
@@
 Spatio-temporal data in other foreign objects can be coerced into cubble with the function \code{as_cubble()}.
-This includes casting from a \code{tibble} or \code{data.frame} with both spatial and temporal information, a netCDF object, a \code{stars} object [@stars], and a \code{sftime} object [@sftime].
-The two examples below shows the casting from a tibble and a netcdf object to a cubble object.
+This includes casting from a \code{tibble} or \code{data.frame} with both spatial and temporal information, a NetCDF object, a \code{stars} object [@stars], and a \code{sftime} object [@sftime].
+The two examples below shows the casting from a tibble and a Netcdf object to a cubble object.
 
@@
 
-```{r echo = TRUE}
+```{r echo = TRUE, eval = FALSE}
 climate_flat |> as_cubble(key = id, index = date, coords = c(long, lat))
@@
 In \proglang{R}, packages for wrangling NetCDF data include a high-level R interface: \pkg{ncdf4} [@ncdf4], a low-level interface that calls a C-interface: \pkg{RNetCDF} [@rnetcdf], and a tidyverse implementation: \pkg{tidync} [@tidync].
-The code below casts a netCDF object in the \code{ncdf4} class into a cubble object:
+The code below casts a NetCDF object in the \code{ncdf4} class into a cubble object:
 
@@
 
-Sometimes, one may want to read in a subset of the netcdf data and the argument \code{vars}, \code{long_range} and \code{lat_range} can be used to subset on the variable and the grid resolution:
+Sometimes, one may want to read in a subset of the NetCDF data and the argument \code{vars}, \code{long_range} and \code{lat_range} can be used to subset on the variable and the grid resolution:
 
 ```{r echo = TRUE, eval = FALSE}
-as_cubble(
-  raw, vars = "q",
-  long_range = seq(-180, 180, 1), lat_range = seq(-90, 90, 1)
-  )
+as_cubble(raw, vars = "q",
+          long_range = seq(-180, 180, 1), lat_range = seq(-90, 90, 1))
 ```

### 2023-05-30 jss/main.tex
subsection 2.2 done

@@
 
-\hypertarget{the-cubble-object}{%
-\subsection{The cubble object}\label{the-cubble-object}}
+\hypertarget{object}{%
+\subsection{The cubble object}\label{object}}
 
@@
 
-\hypertarget{creating-from-separate-spatial-and-temporal-table}{%
-\subsubsection{Creating from separate spatial and temporal table}\label{creating-from-separate-spatial-and-temporal-table}}
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> make_cubble(
-+   spatial = stations, temporal = meteo,
-+   key = id, index = date, coords = c(long, lat)
-+   )
-\end{CodeInput}
-\begin{CodeOutput}
-# cubble:   key: id [3], index: date, nested form
-# extent:   [144.8321, -37.98, 145.0964, -37.6655]
-# temporal: prcp [dbl], tmax [dbl], tmin [dbl]
-  id           long   lat  elev name              wmo_id ts      
-  <chr>       <dbl> <dbl> <dbl> <chr>              <dbl> <list>  
-1 ASN00086038  145. -37.7  78.4 essendon airport   95866 <tibble>
-2 ASN00086077  145. -38.0  12.1 moorabbin airport  94870 <tibble>
-3 ASN00086282  145. -37.7 113.  melbourne airport  94866 <tibble>
-\end{CodeOutput}
-\end{CodeChunk}
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> make_cubble(spatial = stations_sf, temporal = meteo, 
-+             key = id, index = date)
-R> 
-R> make_cubble(spatial = stations, temporal = meteo_ts, 
-+             coords = c(long, lat))
-\end{CodeInput}
-\end{CodeChunk}
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> make_cubble(spatial = stations_sf, temporal = meteo_ts)
+\hypertarget{creating-from-separate-spatial-and-temporal-tables}{%
+\subsubsection{Creating from separate spatial and temporal tables}\label{creating-from-separate-spatial-and-temporal-tables}}
+
+Spatio-temporal data may arrives in separate tables for analysts.
+For example, in climate data, analysts may initially receive station data containing geographic location information, variables recorded and their recording periods.
+They can then query the temporal variables using the stations of interest to obtain the corresponding temporal data.
+Alternatively, analyses may begin as purely spatial or temporal, and analysts may obtain additional temporal or spatial data to expand the result to spatio-temporal.
+
+The function \code{make_cubble()} composes a \code{cubble} object from a spatial table (\code{spatial}) and a temporal table (\code{temporal}) with the three attributes introduced in the subsection \ref{object}: \code{key}, \code{index}, and \code{coords}.
+The following code creates the nested \code{cubble} object shown previously:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> make_cubble(spatial = stations, temporal = meteo,
++             key = id, index = date, coords = c(long, lat))
+\end{CodeInput}
+\end{CodeChunk}
+
+The \code{key} and \code{index} argument comes from the same arguments from the \code{tsibble} objects and \code{coords} derives from the geometry column in an \code{sf} object.
+Hence, the corresponding argument can be safely omitted, if the spatial data is an \code{sf} object, i.e.~\code{stations_sf}, or the temporal data is a \code{tsibble} object, i.e.~\code{meteo_ts}.
+The \code{sf} and \code{tsibble} class from the input will be carried over to the \code{cubble} object:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> (res <- make_cubble(spatial = stations_sf, temporal = meteo_ts))
 \end{CodeInput}
@@
 \end{CodeOutput}
+\begin{CodeInput}
+R> class(res)
+\end{CodeInput}
+\begin{CodeOutput}
+[1] "spatial_cubble_df" "cubble_df"         "sf"               
+[4] "tbl_df"            "tbl"               "data.frame"       
+\end{CodeOutput}
+\begin{CodeInput}
+R> class(res$ts[[1]])
+\end{CodeInput}
+\begin{CodeOutput}
+[1] "tbl_ts"     "tbl_df"     "tbl"        "data.frame"
+\end{CodeOutput}
 \end{CodeChunk}
@@
 Spatio-temporal data in other foreign objects can be coerced into cubble with the function \code{as_cubble()}.
-This includes casting from a \code{tibble} or \code{data.frame} with both spatial and temporal information, a netCDF object, a \code{stars} object \citep{stars}, and a \code{sftime} object \citep{sftime}.
-The two examples below shows the casting from a tibble and a netcdf object to a cubble object.
+This includes casting from a \code{tibble} or \code{data.frame} with both spatial and temporal information, a NetCDF object, a \code{stars} object \citep{stars}, and a \code{sftime} object \citep{sftime}.
+The two examples below shows the casting from a tibble and a Netcdf object to a cubble object.
 
@@
 \end{CodeInput}
-\begin{CodeOutput}
-# cubble:   key: id [3], index: date, nested form
-# extent:   [144.8321, -37.98, 145.0964, -37.6655]
-# temporal: prcp [dbl], tmax [dbl], tmin [dbl]
-  id           long   lat  elev name              wmo_id ts      
-  <chr>       <dbl> <dbl> <dbl> <chr>              <dbl> <list>  
-1 ASN00086038  145. -37.7  78.4 essendon airport   95866 <tibble>
-2 ASN00086077  145. -38.0  12.1 moorabbin airport  94870 <tibble>
-3 ASN00086282  145. -37.7 113.  melbourne airport  94866 <tibble>
-\end{CodeOutput}
 \end{CodeChunk}
@@
 In \proglang{R}, packages for wrangling NetCDF data include a high-level R interface: \pkg{ncdf4} \citep{ncdf4}, a low-level interface that calls a C-interface: \pkg{RNetCDF} \citep{rnetcdf}, and a tidyverse implementation: \pkg{tidync} \citep{tidync}.
-The code below casts a netCDF object in the \code{ncdf4} class into a cubble object:
+The code below casts a NetCDF object in the \code{ncdf4} class into a cubble object:
 
@@
 
-Sometimes, one may want to read in a subset of the netcdf data and the argument \code{vars}, \code{long_range} and \code{lat_range} can be used to subset on the variable and the grid resolution:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> as_cubble(
-+   raw, vars = "q",
-+   long_range = seq(-180, 180, 1), lat_range = seq(-90, 90, 1)
-+   )
+Sometimes, one may want to read in a subset of the NetCDF data and the argument \code{vars}, \code{long_range} and \code{lat_range} can be used to subset on the variable and the grid resolution:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> as_cubble(raw, vars = "q",
++           long_range = seq(-180, 180, 1), lat_range = seq(-90, 90, 1))
 \end{CodeInput}

### 2023-05-31 review/B-response.Rmd
update response

@@
 **\textcolor{purple}{Re: We select the code and output relevant to demonstrate the cubble package in the examples while leaving the rest in the github repository, where users can look up for more details: https://github.com/huizezhang-sherry/paper-cubble.
-[on the visualisation packages: xxx]}.
-After addressing all the issues raised by the reviewers, the paper is now xxx pages, including 4 pages of reference and required affiliation.**
+[on the visualisation packages: xxx].
+After addressing all the issues raised by the reviewers, the paper is now xxx pages, including 4 pages of reference and required affiliation.}**
 
@@
 In other words, the authors use an attribute of an object to mimic the functionality of a sub-class.
-
-**\textcolor{purple}{Re: [TODO] }**
-
 A more natural implementation would be start with two distinct data structures, both sub-classes of the abstract cubble_df class, c("spatial_cubble_df", "cubble_df") and c("spatial_temporal_df", "cubble_df").
 Such a generic implementation would allow for natural extension of the functionality trough dedicated methods.
+
+**\textcolor{purple}{Re: We thanks reviewer for the suggestion on the refactorand now: }**
+
+  * **\textcolor{purple}{a nested/spatial cubble has class c("spatial\_cubble\_df", "cubble\_df")}**
+  * **\textcolor{purple}{a long/temporal cubble has class c("temporal\_cubble\_df", "cubble\_df")}**
+
+**\textcolor{purple}{As a result, S3 methods are now written separately for the two classes unless it doesn't differentiate between the form. For example:}**
+
+  * **\textcolor{purple}{face\_temporal() will do different things depending on whether it is a nested/long cubble:}**
+
+```{r eval = FALSE}
+face_temporal <- function(...){
+  UseMethod("face_temporal")
+}
+
+face_temporal.spatial_cubble_df <- function(...){
+
+}
+
+face_temporal.temporal_cubble_df <- function(...){
+  # if already in the temporal face, return the cubble as it is
+}
+```
+
+  * **\textcolor{purple}{key\_vars() will anyway extract the key variable of a cubble object despite of the form:}**
+
+```{r eval = FALSE}
+key_vars.cubble_df <- function(){
+  ...
+}
+```
+
 Currently an ever-present check on “form” attribute is necessary throughout the code base.
@@
 
-**\textcolor{purple}{Re: Now as\_cubble will consistantly produce a cubble object, or error messages if it fails.
-When there are unmatched pairs, as\_cubble will create with cubble with sites that have both spatial and temporal information and instruct users to use check\_key() to check on key matching.
-Below is a much simplified example to show the current workflow:}**
-
-``` r
-library(cubble)
-library(dplyr)
-#> 
-#> Attaching package: 'dplyr'
-#> The following objects are masked from 'package:stats':
-#> 
-#>     filter, lag
-#> The following objects are masked from 'package:base':
-#> 
-#>     intersect, setdiff, setequal, union
-three_sites <-  stations %>% head(3)
-five_ts <- climate %>% filter(lubridate::month(date) == 1)
-cb <- as_cubble(
-  list(spatial = three_sites, temporal = five_ts),
-  key = id, index = date, coords = c(long, lat)
-)
-#> ! Some sites in the temporal table don't have spatial information
-#> ! Use `check_key()` to check on the unmatched key
-#> The cubble is created only with sites having both spatial and temporal information
-check_key(spatial = three_sites, temporal = five_ts)
-#> $paired
-#> # A tibble: 3 × 2
-#>   spatial     temporal   
-#>   <chr>       <chr>      
-#> 1 ASN00009021 ASN00009021
-#> 2 ASN00010311 ASN00010311
-#> 3 ASN00010614 ASN00010614
-#> 
-#> $potential_pairs
-#> # A tibble: 0 × 0
-#> 
-#> $others
-#> $others$temporal
-#> [1] "ASN00014015" "ASN00015131"
-#> 
-#> $others$spatial
-#> character(0)
-```
-
-<sup>Created on 2023-05-23 with [reprex v2.0.2](https://reprex.tidyverse.org)</sup>
+**\textcolor{purple}{Re: as\_cubble() now can coerce four types of foreign objects into a cubble object: 1) a single joint table, 2) NetCDF data, 3) a stars object, 4) a sftime object.
+The code related to output matching pairs has been moved to a separate function check\_key() and reflected in the current example 5.1}**
 
@@
 
-  * Rework the definitional parts (sections 2 and 3) by following more closely the “design” vignette. **\textcolor{purple}{Re: [TODO: follow the design vignette]}**
-  * Describe at a glance the core functions which operate on the class and their motivation.
-**\textcolor{purple}{Re: Methods available for a cubble object is now printed in xxx, before introducing the main functionality: face\_temporal(), face\_spatial(), and unfold().}**
+  * Rework the definitional parts (sections 2 and 3) by following more closely the “design” vignette.
+**\textcolor{purple}{Re: Section 2 and 3 have now been combined into a section 2 and the following subsections are includeded: }**
+
+      > **\textcolor{purple}{2.1 introduce the cubble class}**
+
+      > **\textcolor{purple}{2.2 create and coerce foreign objects into a cubble object}**
+
+      > **\textcolor{purple}{2.3 functions and methods in cubble (dplyr and cubble functions)}**
+
+      > **\textcolor{purple}{2.4 compatibility with tsibble and sf}**
+
+      > **\textcolor{purple}{2.5 comparision with other existing spatio-temporal class: stars and sftime}**
+
+  * Describe at a glance the core functions which operate on the class and their motivation. **\textcolor{purple}{Re: A method at glance has been printed at the beginning of subsection 2.3 }**
   * Rework section 4 by making it more concise and provide links to sections in supplementary material or dedicated vignettes. **\textcolor{purple}{Re: }**
@@
 **\textcolor{purple}{Re: Adopted.
-Now when applying face\_spatial to a nested cubble (a spatial object), the cubble will be printed as it is with a message signalling it is already in the nested form.
+Now when applying face\_spatial to a nested cubble, the cubble will be printed as it is with a message signalling it is already in the nested form.
 see the reprex below.
-Same for face\_temporal on a long cubble}**
+Same for face\_temporal on a long cubble object}**
 
@@
 
-**\textcolor{purple}{Re: xxx. we could have a function to print }**
+**\textcolor{purple}{Re: the spatial attribute can be checked with function spaital(<cubbble\_obj>)}**
 
@@
 That is, the spatial component cannot be represented as part of the nested tibble without repetition.
-**\textcolor{purple}{Re: [TODO - also compare with sftime] added.}**
+**\textcolor{purple}{Re: this has now been incorporated in subsection 2.5 to compare with a sftime object}**
 
   - on p8 section 3.6, The missing elephant is tidyr which has nesting and unnesting operations  https://tidyr.tidyverse.org/articles/nest.html.
-**\textcolor{purple}{Re: addressed under section Further suggestion}**
+**\textcolor{purple}{Re: addressed under section Further suggestion in this response}**
 
@@
 
-**\textcolor{purple}{Re: [TODO] The section 3.5 are extended to also discuss why not just use existing data structure.
-This includes the comparision with sftime on memory efficiency, with stars on representing spatio-temporal data in an array format with illustrative examples}**
+**\textcolor{purple}{Re: The comparison with other existing spatio-temporal class, i.e.
+stars and sftime, has been addressed in subsection 2.5 (after combining the original section 2 and 3 into section 2)}**
 
@@
 **\textcolor{purple}{Re: [TODO] The two datasets are related but not the same.
-The data historical\_tmax contains daily weather stations data for the two periods: 1971-1975 and 2016-2020 and amounts to 3.2MB, which is not suitable to include in the package given the size restriction.
+The data historical\_tmax contains daily weather stations data for the two periods: 1971-1975 and 2016-2020 and amounts to 3.2MB, which is not suitable to include in the package given its size.
 The tmax\_hist summarises the daily data into monthly to produce the dataset suitable to include in the package.
@@
 **\textcolor{purple}{Re: A result file in html has been submitted this time.}**
-
-section cubble object under 3. the cubble package, on 1) the conceptual framework, 2)  the nested and long form as per the design vig
-
-print of methods
-
-section 4/5: work on the text to make it more concise
-
-connnecting dat ais gneerally a difficult task from different sources in spatial temporal data, point it to the future work
-
-In the end talk to the users to check the vignette for the full code for examples. match titles.

### 2023-05-31 2-cubble.Rmd
fix gramma

@@
 
-Spatio-temporal data may arrives in separate tables for analysts.
+Spatio-temporal data may arrive in separate tables for analysts.
 For example, in climate data, analysts may initially receive station data containing geographic location information, variables recorded and their recording periods.
@@
 
-The \code{key} and \code{index} argument comes from the same arguments from the \code{tsibble} objects and \code{coords} derives from the geometry column in an \code{sf} object.
+The \code{key} and \code{index} arguments corresponds to the \code{tsibble} counterpart and \code{coords} derives from the geometry column in an \code{sf} object.
 Hence, the corresponding argument can be safely omitted, if the spatial data is an \code{sf} object, i.e.
@@
 
-Spatio-temporal data in other foreign objects can be coerced into cubble with the function \code{as_cubble()}.
+Spatio-temporal data in other foreign objects can be coerced into a \code{cubble} object with the function \code{as_cubble()}.
 This includes casting from a \code{tibble} or \code{data.frame} with both spatial and temporal information, a NetCDF object, a \code{stars} object [@stars], and a \code{sftime} object [@sftime].
-The two examples below shows the casting from a tibble and a Netcdf object to a cubble object.
+The two examples below show the casting from a tibble and a NetCDF object to a cubble object.
 

### 2023-05-31 jss/main.tex
fix gramma

@@
 
-The \code{key} and \code{index} argument comes from the same arguments from the \code{tsibble} objects and \code{coords} derives from the geometry column in an \code{sf} object.
+The \code{key} and \code{index} arguments corresponds to the \code{tsibble} counterpart and \code{coords} derives from the geometry column in an \code{sf} object.
 Hence, the corresponding argument can be safely omitted, if the spatial data is an \code{sf} object, i.e.~\code{stations_sf}, or the temporal data is a \code{tsibble} object, i.e.~\code{meteo_ts}.
@@
 
-Spatio-temporal data in other foreign objects can be coerced into cubble with the function \code{as_cubble()}.
+Spatio-temporal data in other foreign objects can be coerced into a \code{cubble} object with the function \code{as_cubble()}.
 This includes casting from a \code{tibble} or \code{data.frame} with both spatial and temporal information, a NetCDF object, a \code{stars} object \citep{stars}, and a \code{sftime} object \citep{sftime}.
-The two examples below shows the casting from a tibble and a Netcdf object to a cubble object.
+The two examples below show the casting from a tibble and a NetCDF object to a cubble object.
 

### 2023-05-31 2-cubble.Rmd
work on subsection 2.3 functionalities

@@
 In this example, it includes \code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.
+
+The first line in the header now shows it in the long form and the third line has been changed to display the name and type of spatial variables: \code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]}.
+Unlike the nested form, the long \code{cubble} is built from a \code{grouped_df} class where all the observations from the same site form a group.
+
+The nested form can be used for those operations where the output is only indexed by the spatial identifier (\code{key}), but becomes inadequate when outputs need both a spatial and a temporal identifier (\code{key} and \code{index}).
+The \code{cubble} class also provides a long form, which expands the \code{ts} column and stores the spatial variables as an attribute.
+The function \code{face_temporal()} is used to switch from the nested \code{cubble} into the long one.
+The first row in Figure  illustrates this operation where the focus of the cube now changes from the site-variable face to the time-variable face.
 
@@
 
-## Functions and methods in cubble
+## Functionalities
 
-*Print all the methods with methods(class = "cubble_df")*
+Table \@ref(tab:funs) summarises functions implemented in the \pkg{cubble} package.
 
-### dplyr methods
+Table: (\#tab:funs) Functions implemented in the \pkg{cubble} package
+
+|Category | Details |  Functions|
+|---|---|--------------|
+|base R | subset | `[`, `[<-`, `names<-`, `[[`, `[[<-`| 
+|tidyverse | dplyr |`dplyr_row_slice`, `dplyr_col_modify`, `dplyr_reconstruct`, `select`, `mutate`, `arrange`, `filter`, `summarise`, `slice`, `group_by`, `ungroup`|
+|tidyverse| tidyr| | 
+|tidyverse| pillar| `type_sum` | 
+|sf| | | 
+|tsibble| | | 
+|cubble | basics| `face_temporal`, `face_spatial`, `unfold`|
+| | attributes| `key`, `key_vars`, `key_data`, `index`, `index_var`, `coords`, `spatial`|
+| |hierarchy|  `switch_key`|
+| |matching |   `match_sites`, `match_spatial`, `match_temporal`| 
+| |glyph map | `geom_glyph`, `geom_glyph_box`, `geom_glyph_line`|
+
+### dplyr and tidyr
 
@@
 
-The created \code{cubble} is a subclass of the \code{rowwise_df} class where each row forms a group.
-All the temporal variables are nested in a list column, hence it is also called the nested \code{cubble}.
-The rowwise structure makes it simpler to operate on the list using the \code{mutate()} syntax, which is simpler than the \code{purr::map()} when working with a list column.
-For example, calculating the number of rainy days can be done by:
+The toy example below demonstrates using dplyr verbs on a nested cubble: 
+```{r echo = TRUE}
+cb_nested %>% 
+  mutate(avg_temp = mean(ts$tmax, na.rm = TRUE)) %>% 
+  select(-elev) %>% 
+  filter(avg_temp > 26) %>% 
+  arrange(-avg_temp)
+```
+
+### cubble
+
+The three basic cubble functions, \code{face_temporal()}, \code{face_spatial()}, and \code{unfold()}, are introduced in this section before more advanced functionalities in Section \@ref(others).
+
+The pair of verbs, \code{face_temporal()} and \code{face_spatial()}, pivot the cubble object between the spatial and temporal face of the multivariate spatio-temporal cube, as illustrated in Figure \ref{fig:face}.
+The code below uses the function \code{face_temporal()} and \code{face_spatial()} to transform between a nested cubble (\code{cb_nested}) and a long one (\code{cb_long}), introduced earlier in subsection \@ref(object):
 
 ```{r echo = TRUE}
-cb_nested |> mutate(rain_day = sum(ts$prcp != 0))
+identical(face_temporal(cb_nested), cb_long)
+identical(face_spatial(cb_long), cb_nested)
 ```
 
-### cubble specific methods
-
-The three cubble specific functions are \code{face_temporal()}, \code{face_spatial()}, and \code{unfold}.
-
-The nested form can be used for those operations where the output is only indexed by the spatial identifier (\code{key}), but becomes inadequate when outputs need both a spatial and a temporal identifier (\code{key} and \code{index}).
-The \code{cubble} class also provides a long form, which expands the \code{ts} column and stores the spatial variables as an attribute.
-The function \code{face_temporal()} is used to switch from the nested \code{cubble} into the long one.
-The first row in Figure \ref{fig:face} illustrates this operation where the focus of the cube now changes from the site-variable face to the time-variable face.
-This code switches the \code{cubble} object just created into its long form:
+The pair of verbs are exact inverse and apply both functions on a cubble object will result in the object itself:
 
 ```{r echo = TRUE}
-(cb_long <- cb_nested |> face_temporal())
+identical(face_spatial(face_temporal(cb_nested)), cb_nested)
+identical(face_temporal(face_spatial(cb_long)), cb_long)
 ```
 
-The first line in the header now shows it in the long form and the third line has been changed to display the name and type of spatial variables: \code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]}.
-Unlike the nested form, the long \code{cubble} is built from a \code{grouped_df} class where all the observations from the same site form a group.
-
-```{r face, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "An illustration of function \\code{face\\_temporal} and \\code{face\\_spatial}.
-In the first row, \\code{face\\_temporal} switches a \\code{cubble} object from the nested form into the long form and the focus has switched from the spatial aspect (the side face) to the temporal aspect (the front face).
-In the second row, \\code{face\\_spatial} switches a \\code{cubble} object back to the nested form from the long form and shifts focus back to the spatial aspect."}
+```{r face, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "An illustration of the function \\code{face\\_temporal()} and \\code{face\\_spatial()}: \\code{face\\_temporal()} converts a spatial cubble (neseted form) into a temporal cubble (long form) to focus on the temporal variables.
+Conversely, \\code{face\\_spatial()} transforms a temporal cubble into a spatial one to emphasize spatial variables."}
 knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.002.png"))
-```
-
-Wrangling spatio-temporal data can be seen as an iterative process in the spatial and temporal dimensions.
-Switching the focus back to the site-variable face can be accomplished by the function \code{face_spatial()}, which is the inverse of \code{face_temporal()}.
-The second row of Figure \ref{fig:face} illustrates the function, which is used as follows:
-
-```{r echo = TRUE}
-(cb_back <- cb_long |> face_spatial())
-#identical(cb_nested, cb_back)
 ```

### 2023-05-31 jss/main.tex
work on subsection 2.3 functionalities

@@
 
+The first line in the header now shows it in the long form and the third line has been changed to display the name and type of spatial variables: \code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]}.
+Unlike the nested form, the long \code{cubble} is built from a \code{grouped_df} class where all the observations from the same site form a group.
+
+The nested form can be used for those operations where the output is only indexed by the spatial identifier (\code{key}), but becomes inadequate when outputs need both a spatial and a temporal identifier (\code{key} and \code{index}).
+The \code{cubble} class also provides a long form, which expands the \code{ts} column and stores the spatial variables as an attribute.
+The function \code{face_temporal()} is used to switch from the nested \code{cubble} into the long one.
+The first row in Figure illustrates this operation where the focus of the cube now changes from the site-variable face to the time-variable face.
+
 \begin{CodeChunk}
@@
 
-\hypertarget{functions-and-methods-in-cubble}{%
-\subsection{Functions and methods in cubble}\label{functions-and-methods-in-cubble}}
-
-\emph{Print all the methods with methods(class = ``cubble\_df'')}
-
-\hypertarget{dplyr-methods}{%
-\subsubsection{dplyr methods}\label{dplyr-methods}}
+\hypertarget{functionalities}{%
+\subsection{Functionalities}\label{functionalities}}
+
+Table \ref{tab:funs} summarises functions implemented in the \pkg{cubble} package.
+
+\begin{longtable}[]{@{}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1500}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1500}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.7000}}@{}}
+\caption{\label{tab:funs} Functions implemented in the \pkg{cubble} package}\tabularnewline
+\toprule()
+\begin{minipage}[b]{\linewidth}\raggedright
+Category
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Details
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Functions
+\end{minipage} \\
+\midrule()
+\endfirsthead
+\toprule()
+\begin{minipage}[b]{\linewidth}\raggedright
+Category
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Details
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Functions
+\end{minipage} \\
+\midrule()
+\endhead
+base R & subset & \texttt{{[}}, \texttt{{[}\textless{}-}, \texttt{names\textless{}-}, \texttt{{[}{[}}, \texttt{{[}{[}\textless{}-} \\
+tidyverse & dplyr & \texttt{dplyr\_row\_slice}, \texttt{dplyr\_col\_modify}, \texttt{dplyr\_reconstruct}, \texttt{select}, \texttt{mutate}, \texttt{arrange}, \texttt{filter}, \texttt{summarise}, \texttt{slice}, \texttt{group\_by}, \texttt{ungroup} \\
+tidyverse & tidyr & \\
+tidyverse & pillar & \texttt{type\_sum} \\
+sf & & \\
+tsibble & & \\
+cubble & basics & \texttt{face\_temporal}, \texttt{face\_spatial}, \texttt{unfold} \\
+& attributes & \texttt{key}, \texttt{key\_vars}, \texttt{key\_data}, \texttt{index}, \texttt{index\_var}, \texttt{coords}, \texttt{spatial} \\
+& hierarchy & \texttt{switch\_key} \\
+& matching & \texttt{match\_sites}, \texttt{match\_spatial}, \texttt{match\_temporal} \\
+& glyph map & \texttt{geom\_glyph}, \texttt{geom\_glyph\_box}, \texttt{geom\_glyph\_line} \\
+\bottomrule()
+\end{longtable}
+
+\hypertarget{dplyr-and-tidyr}{%
+\subsubsection{dplyr and tidyr}\label{dplyr-and-tidyr}}
 
@@
 
-The created \code{cubble} is a subclass of the \code{rowwise_df} class where each row forms a group.
-All the temporal variables are nested in a list column, hence it is also called the nested \code{cubble}.
-The rowwise structure makes it simpler to operate on the list using the \code{mutate()} syntax, which is simpler than the \code{purr::map()} when working with a list column.
-For example, calculating the number of rainy days can be done by:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> cb_nested |> mutate(rain_day = sum(ts$prcp != 0))
-\end{CodeInput}
-\begin{CodeOutput}
-# cubble:   key: id [3], index: date, nested form
-# extent:   [144.8321, -37.98, 145.0964, -37.6655]
+The toy example below demonstrates using dplyr verbs on a nested cubble:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> cb_nested %>% 
++   mutate(avg_temp = mean(ts$tmax, na.rm = TRUE)) %>% 
++   select(-elev) %>% 
++   filter(avg_temp > 26) %>% 
++   arrange(-avg_temp)
+\end{CodeInput}
+\begin{CodeOutput}
+# cubble:   key: id [2], index: date, nested form
+# extent:   [144.8321, -37.7276, 144.9066, -37.6655]
 # temporal: prcp [dbl], tmax [dbl], tmin [dbl]
-  id           long   lat  elev name          wmo_id rain_day ts      
-  <chr>       <dbl> <dbl> <dbl> <chr>          <dbl>    <int> <list>  
-1 ASN00086038  145. -37.7  78.4 essendon air~  95866        3 <tibble>
-2 ASN00086077  145. -38.0  12.1 moorabbin ai~  94870        3 <tibble>
-3 ASN00086282  145. -37.7 113.  melbourne ai~  94866        3 <tibble>
-\end{CodeOutput}
-\end{CodeChunk}
-
-\hypertarget{cubble-specific-methods}{%
-\subsubsection{cubble specific methods}\label{cubble-specific-methods}}
-
-The three cubble specific functions are \code{face_temporal()}, \code{face_spatial()}, and \code{unfold}.
-
-The nested form can be used for those operations where the output is only indexed by the spatial identifier (\code{key}), but becomes inadequate when outputs need both a spatial and a temporal identifier (\code{key} and \code{index}).
-The \code{cubble} class also provides a long form, which expands the \code{ts} column and stores the spatial variables as an attribute.
-The function \code{face_temporal()} is used to switch from the nested \code{cubble} into the long one.
-The first row in Figure \ref{fig:face} illustrates this operation where the focus of the cube now changes from the site-variable face to the time-variable face.
-This code switches the \code{cubble} object just created into its long form:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> (cb_long <- cb_nested |> face_temporal())
-\end{CodeInput}
-\begin{CodeOutput}
-# cubble:  key: id [3], index: date, long form
-# extent:  2020-01-01 -- 2020-01-10 [1D], no gaps
-# spatial: long [dbl], lat [dbl], elev [dbl], name [chr], wmo_id
-#   [dbl]
-  id          date        prcp  tmax  tmin
-  <chr>       <date>     <dbl> <dbl> <dbl>
-1 ASN00086038 2020-01-01     0  26.8  11  
-2 ASN00086038 2020-01-02     0  26.3  12.2
-3 ASN00086038 2020-01-03     0  34.5  12.7
-4 ASN00086038 2020-01-04     0  29.3  18.8
-5 ASN00086038 2020-01-05    18  16.1  12.5
-# i 25 more rows
-\end{CodeOutput}
-\end{CodeChunk}
-
-The first line in the header now shows it in the long form and the third line has been changed to display the name and type of spatial variables: \code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]}.
-Unlike the nested form, the long \code{cubble} is built from a \code{grouped_df} class where all the observations from the same site form a group.
+  id           long   lat name              wmo_id avg_temp ts      
+  <chr>       <dbl> <dbl> <chr>              <dbl>    <dbl> <list>  
+1 ASN00086282  145. -37.7 melbourne airport  94866     26.6 <tibble>
+2 ASN00086038  145. -37.7 essendon airport   95866     26.5 <tibble>
+\end{CodeOutput}
+\end{CodeChunk}
+
+\hypertarget{cubble-1}{%
+\subsubsection{cubble}\label{cubble-1}}
+
+The three basic cubble functions, \code{face_temporal()}, \code{face_spatial()}, and \code{unfold()}, are introduced in this section before more advanced functionalities in Section \ref{others}.
+
+The pair of verbs, \code{face_temporal()} and \code{face_spatial()}, pivot the cubble object between the spatial and temporal face of the multivariate spatio-temporal cube, as illustrated in Figure \ref{fig:face}.
+The code below uses the function \code{face_temporal()} and \code{face_spatial()} to transform between a nested cubble (\code{cb_nested}) and a long one (\code{cb_long}), introduced earlier in subsection \ref{object}:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> identical(face_temporal(cb_nested), cb_long)
+\end{CodeInput}
+\begin{CodeOutput}
+[1] TRUE
+\end{CodeOutput}
+\begin{CodeInput}
+R> identical(face_spatial(cb_long), cb_nested)
+\end{CodeInput}
+\begin{CodeOutput}
+[1] TRUE
+\end{CodeOutput}
+\end{CodeChunk}
+
+The pair of verbs are exact inverse and apply both functions on a cubble object will result in the object itself:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> identical(face_spatial(face_temporal(cb_nested)), cb_nested)
+\end{CodeInput}
+\begin{CodeOutput}
+[1] TRUE
+\end{CodeOutput}
+\begin{CodeInput}
+R> identical(face_temporal(face_spatial(cb_long)), cb_long)
+\end{CodeInput}
+\begin{CodeOutput}
+[1] TRUE
+\end{CodeOutput}
+\end{CodeChunk}
 
@@
 
-\caption{An illustration of function \code{face\_temporal} and \code{face\_spatial}.
-In the first row, \code{face\_temporal} switches a \code{cubble} object from the nested form into the long form and the focus has switched from the spatial aspect (the side face) to the temporal aspect (the front face).
-In the second row, \code{face\_spatial} switches a \code{cubble} object back to the nested form from the long form and shifts focus back to the spatial aspect.}\label{fig:face}
+\caption{An illustration of the function \code{face\_temporal()} and \code{face\_spatial()}: \code{face\_temporal()} converts a spatial cubble (neseted form) into a temporal cubble (long form) to focus on the temporal variables.
+Conversely, \code{face\_spatial()} transforms a temporal cubble into a spatial one to emphasize spatial variables.}\label{fig:face}
 \end{figure}
-\end{CodeChunk}
-
-Wrangling spatio-temporal data can be seen as an iterative process in the spatial and temporal dimensions.
-Switching the focus back to the site-variable face can be accomplished by the function \code{face_spatial()}, which is the inverse of \code{face_temporal()}.
-The second row of Figure \ref{fig:face} illustrates the function, which is used as follows:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> (cb_back <- cb_long |> face_spatial())
-\end{CodeInput}
-\begin{CodeOutput}
-# cubble:   key: id [3], index: date, nested form
-# extent:   [144.8321, -37.98, 145.0964, -37.6655]
-# temporal: prcp [dbl], tmax [dbl], tmin [dbl]
-  id           long   lat  elev name              wmo_id ts      
-  <chr>       <dbl> <dbl> <dbl> <chr>              <dbl> <list>  
-1 ASN00086038  145. -37.7  78.4 essendon airport   95866 <tibble>
-2 ASN00086077  145. -38.0  12.1 moorabbin airport  94870 <tibble>
-3 ASN00086282  145. -37.7 113.  melbourne airport  94866 <tibble>
-\end{CodeOutput}
-\begin{CodeInput}
-R> #identical(cb_nested, cb_back)
-\end{CodeInput}
 \end{CodeChunk}

### 2023-06-01 2-cubble.Rmd
section 2.4 compatibility with tsibble and sf

@@
 
-The \code{key} and \code{index} arguments corresponds to the \code{tsibble} counterpart and \code{coords} derives from the geometry column in an \code{sf} object.
-Hence, the corresponding argument can be safely omitted, if the spatial data is an \code{sf} object, i.e.
-\code{stations_sf}, or the temporal data is a \code{tsibble} object, i.e.
-\code{meteo_ts}.
-The \code{sf} and \code{tsibble} class from the input will be carried over to the \code{cubble} object:
-
-```{r echo = TRUE}
-(res <- make_cubble(spatial = stations_sf, temporal = meteo_ts))
-class(res)
-class(res$ts[[1]])
-```
+[also talk about the by argument with example]
 
@@
 |Category | Details |  Functions|
-|---|---|--------------|
+|--|---|----------------|
 |base R | subset | `[`, `[<-`, `names<-`, `[[`, `[[<-`| 
 |tidyverse | dplyr |`dplyr_row_slice`, `dplyr_col_modify`, `dplyr_reconstruct`, `select`, `mutate`, `arrange`, `filter`, `summarise`, `slice`, `group_by`, `ungroup`|
-|tidyverse| tidyr| | 
-|tidyverse| pillar| `type_sum` | 
+|| tidyr| | 
+|| pillar| `type_sum` | 
 |sf| | | 
@@
 
-\subsubsection[]{\code{tsibble}}
+Analysts often have their own preferred spatial or temporal data structure, which they may wish to keep for spatio-temporal analysis.
+For example, the \code{tbl_ts} class from the \code{tsibble} package [@tsibble] is commonly used in time series forecasting and similarly, the \code{sf} class [@sf] is often used in spatial data science.
+In \code{cubble}, analysts can combine these two structures together by allowing the spatial component to also be an \code{sf} object and the temporal component to also be a \code{tsibble} object.
 
-The \code{tsibble} class is a subclass of \code{tibble} where the \code{index} and \code{key} components are used to store temporal and strata information, that makes working with temporal data cognitively efficient.
-A \code{cubble} object can use the \code{tsibble} class to store the temporal information, and effectively utilize the specialist time series operations in the \pkg{tsibble} package.
-A \code{tsibble} object can also be casted into a \code{cubble} object through supplying the coordinate information in the argument `coords`:
+### tsibble
+
+The \code{key} and \code{index} arguments in a \code{cubble} object corresponds to the \code{tsibble} counterparts and they can be safely omitted, if the temporal component is a \code{tsibble} object, i.e.
+\code{meteo_ts} in the example below.
+The \code{tsibble} class from the input will be carried over to the \code{cubble} object:
 
 ```{r echo = TRUE}
-cb <- make_cubble(
-  spatial = stations, temporal = meteo_ts, 
-  coords = c(long, lat))
+res <- make_cubble(
+  spatial = stations, temporal = meteo_ts, coords = c(long, lat))
+(res_long <- face_temporal(res))
+class(res_long)
 ```
 
-When a nested \code{cubble} is created, each element in the list-column \code{ts} is in the \code{tsibble} class (labelled \code{tbl_ts})  and operations available to the \code{tsibble} class are still valid on these elements.
-For example, the code below calculates two time series features (mean and variance) of maximum temperature, utilizing the \code{tsibble} syntax in the \code{cubble} object:
+The long cubble shows \code{[tsibble]} in the header to indicate the object also being in a \code{tbl_ts} class (\code{tsibble}).
+Methods applies to the \code{tbl_ts} class can also be applied to the temporal cubble objects, for example, checking whether the data contain temporal gaps:
 
 ```{r echo = TRUE}
-cb %>% face_temporal() %>% tsibble::has_gaps()
+res_long %>% tsibble::has_gaps()
 ```
 
-\subsubsection[]{\code{sf} (\code{s2})}
-
-*add an example on changing CRS?*
-
-The \code{sf} class is also a subclass of \code{tibble} with a specialized feature geometry list-column (\code{sfc}) to store different geometry types (\code{POINT}, \code{LINESTRING}, \code{POLYGON}, \code{MULTIPOLYGON}, etc).
-The package \pkg{sf} provides functions that operate efficiently on this spatial information.
-A \code{cubble} object can store spatial information in  the \code{sf} class.
-Methods for the \code{sfc} class can be applied in the nested form of the cubble object.
-An illustration is in Section \@ref(covid).
-The spatial information can also be stored as an \code{s2} vector in a \code{cubble} object.
+An existing \code{cubble} object can promote its temporal component to a \code{tsibble} object by applying \code{make_temporal_tsibble()}.
+The following code illustrates this with the object \code{cb_long} created in section \@ref(create) and the promoted cubble object is equal to the cubble object originally created from a \code{tsibble} object:
 
 ```{r echo = TRUE}
-cb <- make_cubble(
+cb_long2 <- cb_long %>% make_temporal_tsibble() 
+all.equal(cb_long2, res_long)
+```
+
+### sf
+
+Similarly, an \code{sf} object can be supplied as the spatial component to create a \code{cubble} object, with the \code{coords} argument being omitted.
+This opens up the possibility to represent fixed area with polygons or multipolygons (see Applications \@ref(covid)) and the \code{coords} argument will be calculated as the centroids of the (multi)polygons.
+The \code{[sf]} print in the cubble header suggest an spatial component being also a \code{sf} object:
+
+```{r echo = TRUE}
+(res_nested <- make_cubble(
   spatial = stations_sf, temporal = meteo, 
-  key = id, index = date)
-class(cb)
-#cb %>% sf::st_transform(crs = "EPSG:3857")
+  key = id, index = date))
+class(res_nested)
+```
+
+The \code{sf} functions applicable to a \code{cubble} object have been listed in Table \@ref(tab:funs) and the following code shows how to perform coordinate transformation with \code{st_transform} on a cubble object:
+
+```{r echo =TRUE}
+#res_nested %>% sf::st_transform(crs = "EPSG:3857")
+```
+
+The counterpart to promote the spatial component in an existing \code{cubble} to be an \code{sf} object is \code{make_spatial_sf()}:
+
+```{r echo = TRUE}
+cb_nested2 <- cb_nested %>% make_spatial_sf() 
+all.equal(cb_nested2, res_nested)
 ```

### 2023-06-01 jss/main.Rmd
combine section 2 and 3; rearrange section 3 with bullet points
rename files
section 2.4 compatibility with tsibble and sf

@@
         tibble.print_max = 5, tibble.print_min = 5,
-        width = 70, styler.cache_root = "styler-perm")
+        width = 90, styler.cache_root = "styler-perm")
 ```
@@
 
-```{r child = here::here("2-cube.Rmd")}
+```{r child = here::here("2-cubble.Rmd")}
 
@@
 
-```{r child = here::here("3-the-cubble-pkg.Rmd")}
+```{r child = here::here("3-advanced.Rmd")}
 
@@
 
-```{r child = here::here("4-other-features.Rmd")}
+```{r child = here::here("4-1-covid.Rmd")}
 
@@
 
-```{r child = here::here("5-1-covid.Rmd")}
+```{r child = here::here("4-2-glyph.Rmd")}
 
@@
 
-```{r child = here::here("5-2-aus-hist-max-temp.Rmd")}
+```{r child = here::here("4-3-matching.Rmd")}
 
@@
 
-```{r child = here::here("5-3-river-level-rainfall.Rmd")}
+```{r child = here::here("4-4-netcdf.Rmd")}
 
@@
 
-```{r child = here::here("5-4-era5.Rmd")}
+```{r child = here::here("4-5-interactive.Rmd")}
 
@@
 
-```{r child = here::here("5-5-interactive-graphics.Rmd")}
+```{r child = here::here("5-conclusion.Rmd")}
 
 ```
-
-```{r child = here::here("6-conclusion.Rmd")}
-
-```

### 2023-06-01 jss/main.tex
section 2.4 compatibility with tsibble and sf

@@
 
-The \code{key} and \code{index} arguments corresponds to the \code{tsibble} counterpart and \code{coords} derives from the geometry column in an \code{sf} object.
-Hence, the corresponding argument can be safely omitted, if the spatial data is an \code{sf} object, i.e.~\code{stations_sf}, or the temporal data is a \code{tsibble} object, i.e.~\code{meteo_ts}.
-The \code{sf} and \code{tsibble} class from the input will be carried over to the \code{cubble} object:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> (res <- make_cubble(spatial = stations_sf, temporal = meteo_ts))
-\end{CodeInput}
-\begin{CodeOutput}
-# cubble:   key: id [3], index: date, nested form, [sf]
-# extent:   [144.8321, -37.98, 145.0964, -37.6655]
-# temporal: prcp [dbl], tmax [dbl], tmin [dbl]
-  id           long   lat  elev name  wmo_id            geometry
-  <chr>       <dbl> <dbl> <dbl> <chr>  <dbl>         <POINT [°]>
-1 ASN00086038  145. -37.7  78.4 esse~  95866 (144.9066 -37.7276)
-2 ASN00086077  145. -38.0  12.1 moor~  94870   (145.0964 -37.98)
-3 ASN00086282  145. -37.7 113.  melb~  94866 (144.8321 -37.6655)
-# i 1 more variable: ts <list>
-\end{CodeOutput}
-\begin{CodeInput}
-R> class(res)
-\end{CodeInput}
-\begin{CodeOutput}
-[1] "spatial_cubble_df" "cubble_df"         "sf"               
-[4] "tbl_df"            "tbl"               "data.frame"       
-\end{CodeOutput}
-\begin{CodeInput}
-R> class(res$ts[[1]])
-\end{CodeInput}
-\begin{CodeOutput}
-[1] "tbl_ts"     "tbl_df"     "tbl"        "data.frame"
-\end{CodeOutput}
-\end{CodeChunk}
+{[}also talk about the by argument with example{]}
 
@@
 \begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1500}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1500}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.7000}}@{}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.0952}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1429}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.7619}}@{}}
 \caption{\label{tab:funs} Functions implemented in the \pkg{cubble} package}\tabularnewline
@@
 tidyverse & dplyr & \texttt{dplyr\_row\_slice}, \texttt{dplyr\_col\_modify}, \texttt{dplyr\_reconstruct}, \texttt{select}, \texttt{mutate}, \texttt{arrange}, \texttt{filter}, \texttt{summarise}, \texttt{slice}, \texttt{group\_by}, \texttt{ungroup} \\
-tidyverse & tidyr & \\
-tidyverse & pillar & \texttt{type\_sum} \\
+& tidyr & \\
+& pillar & \texttt{type\_sum} \\
 sf & & \\
@@
 \begin{CodeOutput}
-[1] TRUE
+[1] FALSE
 \end{CodeOutput}
@@
 
-\subsubsection[]{\code{tsibble}}
-
-The \code{tsibble} class is a subclass of \code{tibble} where the \code{index} and \code{key} components are used to store temporal and strata information, that makes working with temporal data cognitively efficient.
-A \code{cubble} object can use the \code{tsibble} class to store the temporal information, and effectively utilize the specialist time series operations in the \pkg{tsibble} package.
-A \code{tsibble} object can also be casted into a \code{cubble} object through supplying the coordinate information in the argument \texttt{coords}:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> cb <- make_cubble(
-+   spatial = stations, temporal = meteo_ts, 
-+   coords = c(long, lat))
-\end{CodeInput}
-\end{CodeChunk}
-
-When a nested \code{cubble} is created, each element in the list-column \code{ts} is in the \code{tsibble} class (labelled \code{tbl_ts}) and operations available to the \code{tsibble} class are still valid on these elements.
-For example, the code below calculates two time series features (mean and variance) of maximum temperature, utilizing the \code{tsibble} syntax in the \code{cubble} object:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> cb %>% face_temporal() %>% tsibble::has_gaps()
+Analysts often have their own preferred spatial or temporal data structure, which they may wish to keep for spatio-temporal analysis.
+For example, the \code{tbl_ts} class from the \code{tsibble} package \citep{tsibble} is commonly used in time series forecasting and similarly, the \code{sf} class \citep{sf} is often used in spatial data science.
+In \code{cubble}, analysts can combine these two structures together by allowing the spatial component to also be an \code{sf} object and the temporal component to also be a \code{tsibble} object.
+
+\hypertarget{tsibble}{%
+\subsubsection{tsibble}\label{tsibble}}
+
+The \code{key} and \code{index} arguments in a \code{cubble} object corresponds to the \code{tsibble} counterparts and they can be safely omitted, if the temporal component is a \code{tsibble} object, i.e.~\code{meteo_ts} in the example below.
+The \code{tsibble} class from the input will be carried over to the \code{cubble} object:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> res <- make_cubble(
++   spatial = stations, temporal = meteo_ts, coords = c(long, lat))
+R> (res_long <- face_temporal(res))
+\end{CodeInput}
+\begin{CodeOutput}
+# cubble:  key: id [3], index: date, long form, [tsibble]
+# extent:  2020-01-01 -- 2020-01-10 [1D], no gaps
+# spatial: long [dbl], lat [dbl], elev [dbl], name [chr], wmo_id
+#   [dbl]
+  id          date        prcp  tmax  tmin
+  <chr>       <date>     <dbl> <dbl> <dbl>
+1 ASN00086038 2020-01-01     0  26.8  11  
+2 ASN00086038 2020-01-02     0  26.3  12.2
+3 ASN00086038 2020-01-03     0  34.5  12.7
+4 ASN00086038 2020-01-04     0  29.3  18.8
+5 ASN00086038 2020-01-05    18  16.1  12.5
+# i 25 more rows
+\end{CodeOutput}
+\begin{CodeInput}
+R> class(res_long)
+\end{CodeInput}
+\begin{CodeOutput}
+[1] "temporal_cubble_df" "cubble_df"          "tbl_ts"            
+[4] "tbl_df"             "tbl"                "data.frame"        
+\end{CodeOutput}
+\end{CodeChunk}
+
+The long cubble shows \code{[tsibble]} in the header to indicate the object also being in a \code{tbl_ts} class (\code{tsibble}).
+Methods applies to the \code{tbl_ts} class can also be applied to the temporal cubble objects, for example, checking whether the data contain temporal gaps:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> res_long %>% tsibble::has_gaps()
 \end{CodeInput}
@@
 
-\subsubsection[]{\code{sf} (\code{s2})}
-
-\emph{add an example on changing CRS?}
-
-The \code{sf} class is also a subclass of \code{tibble} with a specialized feature geometry list-column (\code{sfc}) to store different geometry types (\code{POINT}, \code{LINESTRING}, \code{POLYGON}, \code{MULTIPOLYGON}, etc).
-The package \pkg{sf} provides functions that operate efficiently on this spatial information.
-A \code{cubble} object can store spatial information in the \code{sf} class.
-Methods for the \code{sfc} class can be applied in the nested form of the cubble object.
-An illustration is in Section \ref{covid}.
-The spatial information can also be stored as an \code{s2} vector in a \code{cubble} object.
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> cb <- make_cubble(
+An existing \code{cubble} object can promote its temporal component to a \code{tsibble} object by applying \code{make_temporal_tsibble()}.
+The following code illustrates this with the object \code{cb_long} created in section \ref{create} and the promoted cubble object is equal to the cubble object originally created from a \code{tsibble} object:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> cb_long2 <- cb_long %>% make_temporal_tsibble() 
+R> all.equal(cb_long2, res_long)
+\end{CodeInput}
+\begin{CodeOutput}
+[1] TRUE
+\end{CodeOutput}
+\end{CodeChunk}
+
+\hypertarget{sf}{%
+\subsubsection{sf}\label{sf}}
+
+Similarly, an \code{sf} object can be supplied as the spatial component to create a \code{cubble} object, with the \code{coords} argument being omitted.
+This opens up the possibility to represent fixed area with polygons or multipolygons (see Applications \ref{covid}) and the \code{coords} argument will be calculated as the centroids of the (multi)polygons.
+The \code{[sf]} print in the cubble header suggest an spatial component being also a \code{sf} object:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> (res_nested <- make_cubble(
 +   spatial = stations_sf, temporal = meteo, 
-+   key = id, index = date)
-R> class(cb)
++   key = id, index = date))
+\end{CodeInput}
+\begin{CodeOutput}
+# cubble:   key: id [3], index: date, nested form, [sf]
+# extent:   [144.8321, -37.98, 145.0964, -37.6655]
+# temporal: prcp [dbl], tmax [dbl], tmin [dbl]
+  id           long   lat  elev name  wmo_id            geometry
+  <chr>       <dbl> <dbl> <dbl> <chr>  <dbl>         <POINT [°]>
+1 ASN00086038  145. -37.7  78.4 esse~  95866 (144.9066 -37.7276)
+2 ASN00086077  145. -38.0  12.1 moor~  94870   (145.0964 -37.98)
+3 ASN00086282  145. -37.7 113.  melb~  94866 (144.8321 -37.6655)
+# i 1 more variable: ts <list>
+\end{CodeOutput}
+\begin{CodeInput}
+R> class(res_nested)
 \end{CodeInput}
@@
 [1] "spatial_cubble_df" "cubble_df"         "sf"               
-[4] "tbl_df"            "tbl"               "data.frame"       
-\end{CodeOutput}
-\begin{CodeInput}
-R> #cb %>% sf::st_transform(crs = "EPSG:3857")
-\end{CodeInput}
+[4] "rowwise_df"        "tbl_df"            "tbl"              
+[7] "data.frame"       
+\end{CodeOutput}
+\end{CodeChunk}
+
+The \code{sf} functions applicable to a \code{cubble} object have been listed in Table \ref{tab:funs} and the following code shows how to perform coordinate transformation with \code{st_transform} on a cubble object:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> #res_nested %>% sf::st_transform(crs = "EPSG:3857")
+\end{CodeInput}
+\end{CodeChunk}
+
+The counterpart to promote the spatial component in an existing \code{cubble} to be an \code{sf} object is \code{make_spatial_sf()}:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> cb_nested2 <- cb_nested %>% make_spatial_sf() 
+R> all.equal(cb_nested2, res_nested)
+\end{CodeInput}
+\begin{CodeOutput}
+[1] TRUE
+\end{CodeOutput}
 \end{CodeChunk}

### 2023-06-02 2-cubble.Rmd
section 2.1 getting into shape

@@
 ## The cubble object {#object}
+
+<!-- Spatio-temporal data -->
+
+<!-- remote sensing, spatio-temporal cube: longitude, latitude, and time -->
+
+<!-- trajectory: each observation has unique coordinate (spatial) and temporal information -->
+
+<!-- the spacetime paper classifies spatio-temporal data into four different categories based on different spatial and temporal characteristics.  -->
+
+<!-- Analysts to be clear about the spatio-temporal layouts and choose the corresponding data structure for analysis. -->
 
@@
 
-*the conceptual framework + nested and long form as per the design vignette*
+[conceptual framework] The \code{cubble} class tackles the spatio-temporal data collected at unique fixed locations and allowed for irregularity in the temporal dimension, i.e.
+irregular time steps and missingness.
 
-  - diagram, 
-  - two forms: long/ nested - spatial/ temporal
-  - attributes
-  - headers
+The \code{cubble} class (\code{cubble_df}) is an S3 class [@advr] built on the \code{tibble} class, specifically to organize spatio-temporal data. 
+It uses three attributes to character the spatio-temporal data:
 
-The \code{cubble} class is an S3 class [@advr] built on the \code{tibble} class, specifically to organize spatio-temporal data.
-The \code{cubble} class uses an attribute "form",  to arrange the spatial or temporal data components tidily.
-The form attribute can take a value of either "nested" or "long".
-The nested \code{cubble} is a subclass of rowwise \code{tibble} (\code{rowwise_df}).
-It arranges each spatial site in a row, and uses list columns to store the feature geometry and the temporal information.
-The long \code{cubble} is a subclass of grouped \code{tibble} (\code{grouped_df}), which expands the temporal information into the long form and stores the spatial information in a "spatial" attribute.
+  * \code{key} identifies the spatial locations 
+  * \code{date} identifies the temporal order, and 
+  * \code{coords} identifies the longitude and latitude coordinate associated with the location.
 
-The arguments \code{key} and \code{index} follow the conventions in the \pkg{tsibble} package to describe the temporal order and multiple series while \code{coords} specifies the spatial location of each site.
+The base class \code{cubble_df} has two specialised class \code{spatial_cubble_df} and \code{temporal_cubble_df} to arrange the spatio-temporal data in two forms, as illustrated in Figure \ref{fig:class}.
+
+### nested cubble
+
+A nested cubble (in the spatial form) has class \code{c("spatial_cubble_df", "cubble_df")} and is built from a rowwise tibble (\code{rowwise_df}).
+The nested cubble arranges spatial variables in columns and nests temporal variables in a specialised \code{ts} column.
+The \code{rowwise_df} class is used to work better with the \code{ts} list column to compute per station summary.
+
+Below prints an nested cubble object using the weather station data collected in three airports.
+The data is a subset of a larger data \code{climate_aus} collected from National Oceanic and Atmospheric Administration (NOAA) Global Historical Climatology Network Daily (GHCND).
+The data include spatial variables such as station ID, longitude, latitude, elevation, station name, world meteorology organisation ID, in additional to daily temporal information, maximum and minimum temperature values and precipitation for the year 2020.
+The airport data is a subset of three airport stations in Melbourne for the initial 10 days of the year (the creation of a cubble object will be explained in Section \@ref(create)):
 
@@
 
-Printing a \code{cubble} object provides some information about the data.
-Here \code{id} is the variable name to identify each location and there are five unique locations.
-The bounding box is \code{[115.97, -32.94, 133.55, -12.42]} and provides information about the coordinates in the data.
-The third row shows the name and type of all variables nested in the \code{ts} column.
-In this example, it includes \code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.
+A nested cubble header prints the key and index variable name and its form.
+The spatial extent is given by the bounding box with CRS information (missing here).
+The nested cubble header also shows the name of temporal variables nested in the \code{ts} column and their variable types: \code{prcp [dbl], tmax [dbl], tmin [dbl]}.
 
-The first line in the header now shows it in the long form and the third line has been changed to display the name and type of spatial variables: \code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]}.
-Unlike the nested form, the long \code{cubble} is built from a \code{grouped_df} class where all the observations from the same site form a group.
+### long cubble
 
-The nested form can be used for those operations where the output is only indexed by the spatial identifier (\code{key}), but becomes inadequate when outputs need both a spatial and a temporal identifier (\code{key} and \code{index}).
-The \code{cubble} class also provides a long form, which expands the \code{ts} column and stores the spatial variables as an attribute.
-The function \code{face_temporal()} is used to switch from the nested \code{cubble} into the long one.
-The first row in Figure  illustrates this operation where the focus of the cube now changes from the site-variable face to the time-variable face.
+A long cubble (in the temporal form) has class \code{c("temporal_cubble_df", "cubble_df")}.
+It is built from a grouped tibble (\code{grouped_df}), which expands the temporal variables into the long form and stores the spatial variables as a data attribute.
 
@@
 
-<!-- The data \code{climate_flat} is used to illustrate functionality in the \pkg{cubble} package.
-This is a subset from National Oceanic and Atmospheric Administration (NOAA) [@noaa] Global Historical Climatology Network (GHCN) Daily data.
-It contains spatial variables, station id, latitude, longitude, elevation, station name, world meteorology organisation id, in additional to daily temporal information,  maximum and minimum temperature values and precipitation records for year 2020.
-The first five rows of the data are shown below: -->
+In the header of a long cubble object, the extent now prints the temporal range, the interval, and whether there is gaps in the data.
+It also prints the available spatial variables, as oppose to the temporal variables in the nested cubble.
+The spatial component can be printed with \code{spatial()}:
+
+```{r}
+spatial(cb_long)
+```
+
+<!-- The data \code{climate_flat} is used to illustrate functionality in the \pkg{cubble} package.  The first five rows of the data are shown below: -->
 
@@
 ```{r echo = TRUE}
-res <- make_cubble(
+ts_nested <- make_cubble(
   spatial = stations, temporal = meteo_ts, coords = c(long, lat))
-(res_long <- face_temporal(res))
-class(res_long)
+(ts_long <- face_temporal(ts_nested))
+class(ts_long)
 ```
@@
 ```{r echo = TRUE}
-res_long %>% tsibble::has_gaps()
+ts_long %>% tsibble::has_gaps()
 ```
@@
 ```{r echo = TRUE}
-cb_long2 <- cb_long %>% make_temporal_tsibble() 
-all.equal(cb_long2, res_long)
+ts_long2 <- cb_long %>% make_temporal_tsibble() 
+identical(ts_long2, ts_long)
 ```
@@
 ```{r echo = TRUE}
-(res_nested <- make_cubble(
+(sf_nested <- make_cubble(
   spatial = stations_sf, temporal = meteo, 
   key = id, index = date))
-class(res_nested)
+class(sf_nested)
 ```
@@
 ```{r echo =TRUE}
-#res_nested %>% sf::st_transform(crs = "EPSG:3857")
+#sf_nested %>% sf::st_transform(crs = "EPSG:3857")
 ```
@@
 ```{r echo = TRUE}
-cb_nested2 <- cb_nested %>% make_spatial_sf() 
-all.equal(cb_nested2, res_nested)
+sf_nested <- cb_nested %>% make_spatial_sf() 
+all.equal(sf_nested, sf_nested)
 ```

### 2023-06-02 jss/main.tex
section 2.1 getting into shape

@@
 
-\emph{the conceptual framework + nested and long form as per the design vignette}
+{[}conceptual framework{]} The \code{cubble} class tackles the spatio-temporal data collected at unique fixed locations and allowed for irregularity in the temporal dimension, i.e.~irregular time steps and missingness.
+
+The \code{cubble} class (\code{cubble_df}) is an S3 class \citep{advr} built on the \code{tibble} class, specifically to organize spatio-temporal data.
+It uses three attributes to character the spatio-temporal data:
 
@@
 \item
-  diagram,
+  \code{key} identifies the spatial locations
 \item
-  two forms: long/ nested - spatial/ temporal
+  \code{date} identifies the temporal order, and
 \item
-  attributes
-\item
-  headers
+  \code{coords} identifies the longitude and latitude coordinate associated with the location.
 \end{itemize}
 
-The \code{cubble} class is an S3 class \citep{advr} built on the \code{tibble} class, specifically to organize spatio-temporal data.
-The \code{cubble} class uses an attribute ``form'', to arrange the spatial or temporal data components tidily.
-The form attribute can take a value of either ``nested'' or ``long''.
-The nested \code{cubble} is a subclass of rowwise \code{tibble} (\code{rowwise_df}).
-It arranges each spatial site in a row, and uses list columns to store the feature geometry and the temporal information.
-The long \code{cubble} is a subclass of grouped \code{tibble} (\code{grouped_df}), which expands the temporal information into the long form and stores the spatial information in a ``spatial'' attribute.
-
-The arguments \code{key} and \code{index} follow the conventions in the \pkg{tsibble} package to describe the temporal order and multiple series while \code{coords} specifies the spatial location of each site.
+The base class \code{cubble_df} has two specialised class \code{spatial_cubble_df} and \code{temporal_cubble_df} to arrange the spatio-temporal data in two forms, as illustrated in Figure \ref{fig:class}.
+
+\hypertarget{nested-cubble}{%
+\subsubsection{nested cubble}\label{nested-cubble}}
+
+A nested cubble (in the spatial form) has class \code{c("spatial_cubble_df", "cubble_df")} and is built from a rowwise tibble (\code{rowwise_df}).
+The nested cubble arranges spatial variables in columns and nests temporal variables in a specialised \code{ts} column.
+The \code{rowwise_df} class is used to work better with the \code{ts} list column to compute per station summary.
+
+Below prints an nested cubble object using the weather station data collected in three airports.
+The data is a subset of a larger data \code{climate_aus} collected from National Oceanic and Atmospheric Administration (NOAA) Global Historical Climatology Network Daily (GHCND).
+The data include spatial variables such as station ID, longitude, latitude, elevation, station name, world meteorology organisation ID, in additional to daily temporal information, maximum and minimum temperature values and precipitation for the year 2020.
+The airport data is a subset of three airport stations in Melbourne for the initial 10 days of the year (the creation of a cubble object will be explained in Section \ref{create}):
 
@@
 
-Printing a \code{cubble} object provides some information about the data.
-Here \code{id} is the variable name to identify each location and there are five unique locations.
-The bounding box is \code{[115.97, -32.94, 133.55, -12.42]} and provides information about the coordinates in the data.
-The third row shows the name and type of all variables nested in the \code{ts} column.
-In this example, it includes \code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.
-
-The first line in the header now shows it in the long form and the third line has been changed to display the name and type of spatial variables: \code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]}.
-Unlike the nested form, the long \code{cubble} is built from a \code{grouped_df} class where all the observations from the same site form a group.
-
-The nested form can be used for those operations where the output is only indexed by the spatial identifier (\code{key}), but becomes inadequate when outputs need both a spatial and a temporal identifier (\code{key} and \code{index}).
-The \code{cubble} class also provides a long form, which expands the \code{ts} column and stores the spatial variables as an attribute.
-The function \code{face_temporal()} is used to switch from the nested \code{cubble} into the long one.
-The first row in Figure illustrates this operation where the focus of the cube now changes from the site-variable face to the time-variable face.
+A nested cubble header prints the key and index variable name and its form.
+The spatial extent is given by the bounding box with CRS information (missing here).
+The nested cubble header also shows the name of temporal variables nested in the \code{ts} column and their variable types: \code{prcp [dbl], tmax [dbl], tmin [dbl]}.
+
+\hypertarget{long-cubble}{%
+\subsubsection{long cubble}\label{long-cubble}}
+
+A long cubble (in the temporal form) has class \code{c("temporal_cubble_df", "cubble_df")}.
+It is built from a grouped tibble (\code{grouped_df}), which expands the temporal variables into the long form and stores the spatial variables as a data attribute.
 
@@
 [4] "tbl"                "data.frame"        
+\end{CodeOutput}
+\end{CodeChunk}
+
+In the header of a long cubble object, the extent now prints the temporal range, the interval, and whether there is gaps in the data.
+It also prints the available spatial variables, as oppose to the temporal variables in the nested cubble.
+The spatial component can be printed with \code{spatial()}:
+
+\begin{CodeChunk}
+\begin{CodeOutput}
+# A tibble: 3 x 6
+# Rowwise: 
+  id           long   lat  elev name              wmo_id
+  <chr>       <dbl> <dbl> <dbl> <chr>              <dbl>
+1 ASN00086038  145. -37.7  78.4 essendon airport   95866
+2 ASN00086077  145. -38.0  12.1 moorabbin airport  94870
+3 ASN00086282  145. -37.7 113.  melbourne airport  94866
 \end{CodeOutput}
@@
   <chr>       <dbl> <dbl> <chr>              <dbl>    <dbl> <list>           
-1 ASN00086282  145. -37.7 melbourne airport  94866     26.6 <tibble>
-2 ASN00086038  145. -37.7 essendon airport   95866     26.5 <tibble>
+1 ASN00086282  145. -37.7 melbourne airport  94866     26.6 <tibble [10 x 4]>
+2 ASN00086038  145. -37.7 essendon airport   95866     26.5 <tibble [10 x 4]>
 \end{CodeOutput}
@@
 # extent:  2020-01-01 -- 2020-01-10 [1D], no gaps
-# spatial: long [dbl], lat [dbl], elev [dbl], name [chr], wmo_id
-#   [dbl]
+# spatial: long [dbl], lat [dbl], elev [dbl], name [chr], wmo_id [dbl]
   id          date        prcp  tmax  tmin  long   lat
@@
 \begin{CodeInput}
-R> res <- make_cubble(
+R> ts_nested <- make_cubble(
 +   spatial = stations, temporal = meteo_ts, coords = c(long, lat))
-R> (res_long <- face_temporal(res))
+R> (ts_long <- face_temporal(ts_nested))
 \end{CodeInput}
@@
 # extent:  2020-01-01 -- 2020-01-10 [1D], no gaps
-# spatial: long [dbl], lat [dbl], elev [dbl], name [chr], wmo_id
-#   [dbl]
+# spatial: long [dbl], lat [dbl], elev [dbl], name [chr], wmo_id [dbl]
   id          date        prcp  tmax  tmin
@@
 \begin{CodeInput}
-R> class(res_long)
+R> class(ts_long)
 \end{CodeInput}
@@
 \begin{CodeInput}
-R> res_long %>% tsibble::has_gaps()
+R> ts_long %>% tsibble::has_gaps()
 \end{CodeInput}
@@
 \begin{CodeInput}
-R> cb_long2 <- cb_long %>% make_temporal_tsibble() 
-R> all.equal(cb_long2, res_long)
+R> ts_long2 <- cb_long %>% make_temporal_tsibble() 
+R> identical(ts_long2, ts_long)
 \end{CodeInput}
@@
 \begin{CodeInput}
-R> (res_nested <- make_cubble(
+R> (sf_nested <- make_cubble(
 +   spatial = stations_sf, temporal = meteo, 
@@
 # temporal: prcp [dbl], tmax [dbl], tmin [dbl]
-  id           long   lat  elev name  wmo_id            geometry
-  <chr>       <dbl> <dbl> <dbl> <chr>  <dbl>         <POINT [°]>
-1 ASN00086038  145. -37.7  78.4 esse~  95866 (144.9066 -37.7276)
-2 ASN00086077  145. -38.0  12.1 moor~  94870   (145.0964 -37.98)
-3 ASN00086282  145. -37.7 113.  melb~  94866 (144.8321 -37.6655)
-# i 1 more variable: ts <list>
-\end{CodeOutput}
-\begin{CodeInput}
-R> class(res_nested)
+  id           long   lat  elev name   wmo_id            geometry ts      
+  <chr>       <dbl> <dbl> <dbl> <chr>   <dbl>         <POINT [°]> <list>  
+1 ASN00086038  145. -37.7  78.4 essen~  95866 (144.9066 -37.7276) <tibble>
+2 ASN00086077  145. -38.0  12.1 moora~  94870   (145.0964 -37.98) <tibble>
+3 ASN00086282  145. -37.7 113.  melbo~  94866 (144.8321 -37.6655) <tibble>
+\end{CodeOutput}
+\begin{CodeInput}
+R> class(sf_nested)
 \end{CodeInput}
@@
 \begin{CodeInput}
-R> #res_nested %>% sf::st_transform(crs = "EPSG:3857")
+R> #sf_nested %>% sf::st_transform(crs = "EPSG:3857")
 \end{CodeInput}
@@
 \begin{CodeInput}
-R> cb_nested2 <- cb_nested %>% make_spatial_sf() 
-R> all.equal(cb_nested2, res_nested)
+R> sf_nested <- cb_nested %>% make_spatial_sf() 
+R> all.equal(sf_nested, sf_nested)
 \end{CodeInput}

### 2023-06-03 2-cubble.Rmd
section 2.5 2.5  Comparison to other spatio-temporal classes

@@
 
+```{r class, fig.align="center", out.width = "100%", fig.cap = "Illustration of a cubble object in the long form and nested form, along with the associated attributes."}
+knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.007.png"))
+```
+
 ### nested cubble
@@
 The nested cubble arranges spatial variables in columns and nests temporal variables in a specialised \code{ts} column.
-The \code{rowwise_df} class is used to work better with the \code{ts} list column to compute per station summary.
+The \code{rowwise_df} class is imposed since it is helpful to manipulate with list column, i.e.
+compute per station summary of temporal variables, and you may remove it with \code{strip_rowwise()}.
 
@@
 
-<!-- The data \code{climate_flat} is used to illustrate functionality in the \pkg{cubble} package.  The first five rows of the data are shown below: -->
-
 ## Creation and coercion {#create}
@@
 ```
-
-[also talk about the by argument with example]
 
@@
 
-Some readers may question why a new data structure is needed rather than directly creating a list-column  on the combined data using \code{dplyr::nest_by()}.
-The reason is that the \code{cubble} object is specifically designed to utilize the spatio-temporal structure when arranging observations in a single object.
-Moreover, it enables easy pivoting between purely spatial, purely temporal, or unfolded into a combined form.
+Given the vast spatio-temporal data structure available  in \proglang{R} to represent spatio-temporal data, this section compares and contrast the \pkg{cubble} with other existing alternative, namely \pkg{stars} and \pkg{sftime}.
 
-From the initial \pkg{spacetime} [@spacetime] package.
-New packages, such as \pkg{stars} [@stars] and \pkg{sftime} [@sftime], have [...].
-This section compares and contrast the cubble data structure with other existing alternative, namely \pkg{stars} and \pkg{sftime}.
+The \pkg{stars} package [@stars] uses an array structure, as oppose to tibble in the \code{cubble} package, to represent multivariate spatio-temporal data.
+While both packages support vector and raster data, it is a matter of choice on which structure to use given the application.
+Analysts working on satellite imageries may prefer the array structure while others originally working with spatio-temporal data in the 2D data frames may find \pkg{cubble} easier to adopt from their existing computing workflow.
 
-The `stars` package creates a `stars` object from an array and a dimension object.
-In many cases, i.e.
-satellite images, this is convenient, however, there are also many spatio-temporal data are formatted in 2D data frames.
-
-create a dimension object, populate the data into the array
-
-- array in \proglang{R} allows for character, 
-- specific datetime class
-
-*make the example simpler*
-
-  - `temporal$var1 %>% matrix(byrow = TRUE, nrow = 2, ncol = 5)`
-
-This happens when different stations have different start date and require other software to reshape the data correctly.
-
-<!-- The temporal data contains two variables are recorded at two stations (`id`) from 2023 March 21st to 2023 March 25th.
-Notice is that id2 only records from 2023 March 22nd and variable 2 has an `NA` value on  2023 March 23rd at id2.
--->
-
-<!-- ```{r echo = FALSE} -->
-<!-- library(stars) -->
-<!-- spatial <- tibble( -->
-<!--   id = c(1,2), -->
-<!--   long = c(130, 140), -->
-<!--   lat = c(-38, -44), -->
-<!--   name = c("station 1", "station 2"))  -->
-
-<!-- spatial -->
-
-<!-- temporal <- tibble( -->
-<!--   id = c(rep(1, each = 5), rep(2, each = 4)), -->
-<!--   time = c(seq(as.Date("2023-03-21"), as.Date("2023-03-25"), 1), -->
-<!--            seq(as.Date("2023-03-22"), as.Date("2023-03-25"), 1)), -->
-<!--   var1 = c(3, 6, 8, 2, 5, 7, 9, 14, 5), -->
-<!--   var2 = c(1, 3, 4, 2, 6, 7, NA, 5, 8) -->
-<!-- ) -->
-<!-- temporal -->
-<!-- ``` -->
-
-<!-- The following code can be used to create the dimension object: -->
-
-<!-- ```{r echo = TRUE} -->
-<!-- spatial_sf <- spatial %>% sf::st_as_sf(coords = c("long", "lat"), crs = 4326)  -->
-<!-- station_vec <- st_geometry(spatial_sf) -->
-<!-- time_vec <- unique(temporal$time) -->
-<!-- var_vec <- spatial_sf$name -->
-<!-- d <- st_dimensions(station = station_vec, time = time_vec) -->
-<!-- d -->
-<!-- ``` -->
-
-<!-- However, the issue comes when casting the variables into an array of the same dimension: -->
-
-<!-- ```{r echo = TRUE, eval = FALSE} -->
-<!-- temporal$var1 %>% matrix(byrow = TRUE, nrow = 2, ncol = 5) %>% array(dim = dim(d)) -->
-<!-- ``` -->
-
-<!-- Due to the missingness on 2023 March 21st for id2, the array reuses the first value of the variable to fill in the dimension (the missingness at (2,1) is not flagged and 3 is reused to fill in the dimension).
--->
-
-<!-- This causes an issue when multiple stations with different recording periods present in the data, commonly found in the weather data.  -->
-
-The approach taken by the \pkg{sftime} package is to combine the spatial and temporal variables together into a joint table. [for sf objects with daily observation, memory efficiency]
-
-Consider the climate data used in the \pkg{cubble} package, \code{climate_aus}.
-It contains daily precipitation, minimum and maximum temperature for 639 weather stations across Australia in 2020.
-The converted sftime object is `r library(cubble); #object.size(sftime::st_as_sftime(climate_aus))/object.size(climate_aus)` times the size of the cubble object:
-
-```{r echo = TRUE}
-#climate_aus %>% sftime::st_as_sftime() %>% object.size()
-climate_aus %>% object.size()
-```
+The \pkg{sftime} package [@sftime] also builds from a tibble object, with a main focus to handle irregular spatio-temporal data.
+While \pkg{sftime} can also handle full space-time grids and sparse space-time layouts, \pkg{cubble} can be thought of as simplifying the repetition in spatial variables with two linked tibbles.
+This simplification gives memory efficiency when data is observed frequent, i.e.
+daily or subdaily, or the spatial geometry is expensive to repeat, i.e.
+polygons or multipolygons.
+Consider the \code{climate_aus} data in the \pkg{cubble} package with `r nrow(cubble::climate_aus)` stations observed daily in a single year 2020.
+The created \code{sftime} object is `r library(cubble); round(object.size(sftime::st_as_sftime(climate_aus))/object.size(climate_aus), 2)` times larger than the corresponding \code{cubble} object (`r round(object.size(sftime::st_as_sftime(climate_aus))/1e6, 2)` MB vs.
+`r round(object.size(cubble::climate_aus)/1e6, 2)` MB).

### 2023-06-03 jss/main.tex
section 2.5 2.5  Comparison to other spatio-temporal classes

@@
 
+\begin{CodeChunk}
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{../figures/diagram-keynotes/diagram-keynotes.007}
+
+}
+
+\caption[Illustration of a cubble object in the long form and nested form, along with the associated attributes]{Illustration of a cubble object in the long form and nested form, along with the associated attributes.}\label{fig:class}
+\end{figure}
+\end{CodeChunk}
+
 \hypertarget{nested-cubble}{%
@@
 The nested cubble arranges spatial variables in columns and nests temporal variables in a specialised \code{ts} column.
-The \code{rowwise_df} class is used to work better with the \code{ts} list column to compute per station summary.
+The \code{rowwise_df} class is imposed since it is helpful to manipulate with list column, i.e.~compute per station summary of temporal variables, and you may remove it with \code{strip_rowwise()}.
 
@@
 \end{CodeChunk}
-
-{[}also talk about the by argument with example{]}
 
@@
 
-Some readers may question why a new data structure is needed rather than directly creating a list-column on the combined data using \code{dplyr::nest_by()}.
-The reason is that the \code{cubble} object is specifically designed to utilize the spatio-temporal structure when arranging observations in a single object.
-Moreover, it enables easy pivoting between purely spatial, purely temporal, or unfolded into a combined form.
-
-From the initial \pkg{spacetime} \citep{spacetime} package.
-New packages, such as \pkg{stars} \citep{stars} and \pkg{sftime} \citep{sftime}, have {[}\ldots{]}.
-This section compares and contrast the cubble data structure with other existing alternative, namely \pkg{stars} and \pkg{sftime}.
-
-The \texttt{stars} package creates a \texttt{stars} object from an array and a dimension object.
-In many cases, i.e.~satellite images, this is convenient, however, there are also many spatio-temporal data are formatted in 2D data frames.
-
-create a dimension object, populate the data into the array
-
-\begin{itemize}
-\tightlist
-\item
-  array in \proglang{R} allows for character,
-\item
-  specific datetime class
-\end{itemize}
-
-\emph{make the example simpler}
-
-\begin{itemize}
-\tightlist
-\item
-  \texttt{temporal\$var1\ \%\textgreater{}\%\ matrix(byrow\ =\ TRUE,\ nrow\ =\ 2,\ ncol\ =\ 5)}
-\end{itemize}
-
-This happens when different stations have different start date and require other software to reshape the data correctly.
-
-The approach taken by the \pkg{sftime} package is to combine the spatial and temporal variables together into a joint table. {[}for sf objects with daily observation, memory efficiency{]}
-
-Consider the climate data used in the \pkg{cubble} package, \code{climate_aus}.
-It contains daily precipitation, minimum and maximum temperature for 639 weather stations across Australia in 2020.
-The converted sftime object is times the size of the cubble object:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> #climate_aus %>% sftime::st_as_sftime() %>% object.size()
-R> climate_aus %>% object.size()
-\end{CodeInput}
-\begin{CodeOutput}
-8523040 bytes
-\end{CodeOutput}
-\end{CodeChunk}
+Given the vast spatio-temporal data structure available in \proglang{R} to represent spatio-temporal data, this section compares and contrast the \pkg{cubble} with other existing alternative, namely \pkg{stars} and \pkg{sftime}.
+
+The \pkg{stars} package \citep{stars} uses an array structure, as oppose to tibble in the \code{cubble} package, to represent multivariate spatio-temporal data.
+While both packages support vector and raster data, it is a matter of choice on which structure to use given the application.
+Analysts working on satellite imageries may prefer the array structure while others originally working with spatio-temporal data in the 2D data frames may find \pkg{cubble} easier to adopt from their existing computing workflow.
+
+The \pkg{sftime} package \citep{sftime} also builds from a tibble object, with a main focus to handle irregular spatio-temporal data.
+While \pkg{sftime} can also handle full space-time grids and sparse space-time layouts, \pkg{cubble} can be thought of as simplifying the repetition in spatial variables with two linked tibbles.
+This simplification gives memory efficiency when data is observed frequent, i.e.~daily or subdaily, or the spatial geometry is expensive to repeat, i.e.~polygons or multipolygons.
+Consider the \code{climate_aus} data in the \pkg{cubble} package with 639 stations observed daily in a single year 2020.
+The created \code{sftime} object is 13.87 times larger than the corresponding \code{cubble} object (118.24 MB vs.~8.52 MB).
 

### 2023-06-03 review/B-response.Rmd
update response

@@
 **\textcolor{purple}{Re: The comparison with other existing spatio-temporal class, i.e.
-stars and sftime, has been addressed in subsection 2.5 (after combining the original section 2 and 3 into section 2)}**
+stars and sftime, has been addressed in Section 2.5 (after combining the original section 2 and 3 into section 2)}.**
+
+  * **\textcolor{purple}{The stars object uses an array structure underneath while cubble is built from a tibble.
+It is a matter of choice from the analysts on which structure they are more comfortable to use.
+We believe spatio-temporal analysts who are more comfortable processing data with data frame (tidyverse) will benefit from cubble}**
+
+  * **\textcolor{purple}{With sftime, the main benefit cubble provides is memory efficiency and this efficiency comes with the assumption that the spatial sites are fixed in cubble (full space-time grids and sparse space-time layouts).
+This is beneficial when spatial variables are duplicated because of long temporal observational frequency: daily or subdaily or when the spatial geoemtry (polygons or multipolygons) are expensive to duplicate.
+We include a quick comparison of object size between an sftime object and the same data in a cubble object in the paper.}**
 

### 2023-06-03 4-1-covid.Rmd
go through the covid data example

---
title: "5-1-covid"
output: html_document
---

# Applications {#examples}

The five examples here are chosen to illustrate these aspects of the \pkg{cubble} package: creating a \code{cubble} object from two Coronavirus (COVID) data tables with the complication of differing location names, using spatial transformations to make a glyph map of seasonal temperature changes over years, matching river level data and weather station records for analysis of water supply, reading NetCDF format data to reproduce a climate reanalysis plot, and the workflow to create complex interactive linked plots. (The Appendix provides another example on aggregating information spatially to explore precipitation patterns.)

## Victoria COVID spatio-temporal incidence and spread {#covid}

Since the start of the pandemic, the Victoria State Government in Australia has provided daily COVID counts by local government area (LGA). This data can be used to visualize COVID incidence and spread spatially, when combined with map polygon data available from the Australian Bureau of Statistics. These different sources need to be combined for the analysis, by matching the LGA names. Both the COVID count data (\code{covid}) and the LGA information (\code{lga}) are available in the \pkg{cubble} package with the \code{covid} data stored as a \code{tsibble} object, the \code{lga} data as an \code{sf} object.

The \code{make_cubble()} function is used to create a \code{cubble} object from the two spatial and temporal tables as described in Section \@ref(create). The \code{by} argument allows for mismatch of key names from the two tables in the \code{*_join()} \code{by} syntax:

```{r covid-4, echo = TRUE, warning = TRUE, message= TRUE}
cb <- make_cubble(lga, covid, by = c("lga_name_2018" = "lga"))
```

The warning message reports on the mismatches of location from both sides: there are LGAs in the COVID data that do not match with LGAs names in the spatial polygon data, and vice versa. The \code{make_cubble()} function prompts the use of \code{check_key()} to identify the mismatches:

```{r covid-5, echo = TRUE}
(check_res <- check_key(
  spatial = lga, temporal = covid, 
  by = c("lga_name_2018" = "lga")
))
```

The check result is a list with three elements: 1) matched pairs from both tables, 2) those pairs that can be potentially paired, and 3) others. Here the main mismatch comes from different encodings of two LGAs: Kingston and Latrobe. Analysts can then modify the input spatial and temporal data accordingly and create the cubble again:

```{r covid-6, echo = TRUE}
lga2 <- lga |>
  rename(lga = lga_name_2018) |> 
  mutate(lga = ifelse(lga == "Kingston (C) (Vic.)", "Kingston (C)", lga),
         lga = ifelse(lga == "Latrobe (C) (Vic.)", "Latrobe (C)", lga)) |>
  filter(!lga %in% check_res$others$spatial)

covid2 <- covid |> filter(!lga %in% check_res$others$temporal)

(cb <- make_cubble(spatial = lga2, temporal = covid2))
```

### 2023-06-03 2-cubble.Rmd
misc fix

@@
 
-The toy example below demonstrates using dplyr verbs on a nested cubble: 
-```{r echo = TRUE}
+The toy example below demonstrates the using of a collection of dplyr functions (\code{mutate}, \code{select}, \code{filter}, and \code{arrange}) on a nested cubble object \code{cb_nested}:
+
+```{r echo = TRUE, eval = FALSE}
 cb_nested %>% 
@@
 
-```{r echo = TRUE}
+```{r cubble-fun, echo = TRUE}
 identical(face_temporal(cb_nested), cb_long)
@@
 
-```{r echo = TRUE}
+```{r cubble-fun2, echo = TRUE}
 identical(face_spatial(face_temporal(cb_nested)), cb_nested)

### 2023-06-03 4-2-glyph.Rmd
misc fix

--- 2023-03-25 5-2-aus-hist-max-temp.Rmd
@@
   face_temporal() |> 
-  mutate(month = lubridate::month(date),
-         group = as.factor(
-           ifelse(lubridate::year(date) > 2015,
-           "2016 ~ 2020", "1971 ~ 1975"))) |>
-  group_by(month, group) |> 
-  summarise(tmax = mean(tmax, na.rm = TRUE))
+  group_by(
+    yearmonth = tsibble::make_yearmonth(
+      year = ifelse(lubridate::year(date) > 2015, 2016, 1971),
+      month = lubridate::month(date))
+      )|>
+  summarise(tmax = mean(tmax, na.rm = TRUE)) %>% 
+  mutate(group = as.factor(lubridate::year(yearmonth)),
+         month = lubridate::month(yearmonth))
 ```
@@
 A quick check on the number of observations for each location is made, revealing that there are several with less than 24 observations -- these stations lack temperature values for some months.
-In this example, those stations are removed by switching to a long \code{cubble} to operate on the spatial component over time, and then, move back into the nested \code{cubble} (to make the glyph map):
+In this example, those stations are removed by switching to the nested \code{cubble} to operate on the spatial component over time, and then, move back into the long \code{cubble} (to make the glyph map):
 
 ```{r echo = TRUE}
-tmax |> 
-  face_spatial() |> 
-  mutate(n = nrow(ts)) |>
-  arrange(n) |>
-  pull(n) |> 
-  head(10)
-
 tmax <- tmax |> 
@@
 ```{r eval = FALSE, echo = TRUE}
-nsw_vic <- ozmaps::abs_ste |>
-  filter(NAME %in% c("Victoria", "New South Wales"))
-
 ggplot() +
-  geom_sf(data = nsw_vic,
-          fill = "transparent", color = "grey",
-          linetype = "dotted") +
+  geom_sf(data = MAP_DATA, ...) +
   geom_glyph(data = tmax,
@@
 To fix this, one could aggregate data from nearby stations.
-An example of this is included in the Appendix.
+An example of this use is included in the Appendix.

### 2023-06-03 4-4-netcdf.Rmd
misc fix

--- 2022-06-22 4-4-netcdf.Rmd
@@
 Figure \ref{fig:netcdf} reproduces the ERA5 data row of Figure 19 in @hersbach2020era5.
-Here we explain how this would be done using cubble.
+Here we explain how this would be done using in the \pkg{cubble} package.
 The plots show that the southern polar vortex splits into two on 2002-09-26 and further splits into four on 2002-10-04.
-Further explanation of why this is interesting can be found in the figure source, and also @simmons2020global, @simmons2005ecmwf.
+Further explanation of why this is interesting can be found in the figure source, and also in @simmons2020global and @simmons2005ecmwf.
 
@@
 The data is available in the NetCDF format from the European Centre for Medium-Range Weather Forecasts (ECMWF).
-It can be directly downloaded from [Copernicus Climate Data Store (CDS)](https://cds.climate.copernicus.eu/cdsapp#!/dataset/reanalysis-era5-pressure-levels?tab=overview) website or via the R package \pkg{ecmwfr} [@ecwmfr].
+It can be directly downloaded from Copernicus Climate Data Store (CDS) [@cds] website or via the \pkg{ecmwfr} package [@ecwmfr].
+For the reproduction, we focus on the \code{era5-pressure} data, hourly pressure levels from 1970 to present, with the *specific humidity* and *geopotential*.
+The downloaded NetCDF data (`raw`) is first converted to a \code{cubble} object:
 
-For the reproduction, we focus on the \code{era5-pressure} data, hourly pressure levels from 1970 to present, with the *specific humidity* and *geopotential*.
 <!--Here we specifically look at the information  contained on the variables *specific humidity* measuring the mass of water vapor per kilogram of moist air  and the *geopotential* variable recording the gravitational potential energy of a unit mass, at a particular location, relative to mean sea level.-->
@@
 <!--The goal of this example is to reproduce the break-up of the southern polar vortex that happened in late September and early October in 2002 in the stratosphere in the Southern Hemisphere.
-Therefore, we filter data corresponding to the Southern Hemisphere and restrict our analysis to four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04.--> The downloaded netcdf data is read into R using the \code{ncdf4} package, and converted to a \code{cubble}:
+Therefore, we filter data corresponding to the Southern Hemisphere and restrict our analysis to four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04.-->
+```{r echo = FALSE, eval = TRUE}
+raw <- ncdf4::nc_open(here::here("data/era5-pressure.nc"))
+```
 
-```{r eval = FALSE, echo = TRUE}
-raw <- ncdf4::nc_open(here::here("data/era5-pressure.nc"))
-dt <- as_cubble(raw, vars = c("q", "z"))
+```{r echo = TRUE}
+dt <- as_cubble(
+  raw, vars = c("q", "z"),
+  long_range = seq(-180, 180, 1), lat_range = seq(-88, 88, 1))
 ```
@@
 
-Creating the plot requires making transformations on time, unfolding the data for computing the statistic of interest, which is plotted directly as a contour plot with \code{ggplot}.
+Creating the plot requires making transformations on time, unfolding the data for computing the statistic of interest, which is plotted directly as a contour plot with ggplot.
 Code is provided to accomplish this in the supplementary material.
 
-```{r netcdf, out.width="100%", fig.cap = "A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020) to illustrate the break-up of sourthern polar vortex in late September and early October 2002.
-The polar vortex, signalled by the high speicfic humidity, splits into two on 2002-09-26 and furthers split into four on 2002-10-04."}
-knitr::include_graphics(here::here("figures/netcdf.png"))
+```{r eval = FALSE}
+# prepare data/era5-pressure.nc
+library(ecmwfr)
+my_request <- list(
+  product_type = "reanalysis",
+  format = "netcdf",
+  pressure_level = "10",
+  year = "2002",
+  month = c("09", "10"),
+  day = c("04", "22", "26", "30"),
+  time = "00:00",
+  area = c(-15, -180, -90, 180),
+  variable = c("specific_humidity", 'geopotential'),
+  dataset_short_name = "reanalysis-era5-pressure-levels",
+  target = "era5-pressure.nc"
+)
+# 
+# # here you need to create an account on Climate Data Store 
+# # at https://cds.climate.copernicus.eu/cdsapp#!/home) to obtain the user ID and API key
+# # then set up the with ecmwfr::wf_set_key(user, key, service = "cds")
+# # the user below is the same `user` in the setup line
+wf_request(
+  user = "113339",
+  request = my_request,
+  transfer = TRUE,
+  path = here::here("data/"))
 ```
+
+```{r netcdf, eval =FALSE, out.width="100%", fig.height = 4, fig.width = 10, fig.cap = "A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020) to illustrate the break-up of sourthern polar vortex in late September and early October 2002.
+The polar vortex, signalled by the high specific humidity, splits into two on 2002-09-26 and further splits into four on 2002-10-04."}
+library(sf)
+date <- c("2002-09-22", "2002-09-26", "2002-09-30", "2002-10-04") |> as.Date()
+res <- dt |> 
+  face_temporal() |> 
+  filter(lubridate::date(time) %in% date) |>
+  unfold(long, lat) |> 
+  mutate(q =  q* 10^6)
+
+con <- rnaturalearth::ne_coastline("small", returnclass = "sf")
+box <- st_bbox(c(xmin = -180, ymin = -90, xmax = 180, ymax = -15), crs = st_crs(con)) 
+sf::sf_use_s2(FALSE)
+country <- con %>% st_geometry() %>% 
+  sf::st_crop(box) |> 
+  sf::st_cast("MULTILINESTRING")
+
+res |> 
+  ggplot() +
+  # q for specific humidity
+  geom_point(aes(x = long, y = lat, color = q), size = 2) + 
+  #geom_tile(aes(x = long, y = lat, fill = q), color = "transparent") +
+  # z for geopotential
+  geom_contour(data = res, aes(x = long, y = lat, z = z),
+               color = "grey20", binwidth = 4000, linewidth = 0.5) +
+  geom_sf(data = country , alpha = 0.5, fill = "transparent", color = "lightgreen") +
+  coord_sf(
+    crs = "+proj=stere +lat_0=-90 +lon_0=-180 +k=1 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs ", default_crs = sf::st_crs(4326), clip = "off") +
+  theme_bw() +
+  theme(legend.position = "bottom",
+        axis.ticks = element_blank(),
+        panel.border = element_blank(),
+        axis.text = element_blank(),
+        panel.grid = element_blank()) +
+  facet_wrap(vars(as.Date(time)), nrow = 1)+  
+  colorspace::scale_color_continuous_sequential("Purple-Yellow", name = "Specific humidity") + 
+  theme(legend.text = element_text(size = 13),
+        legend.title = element_text(size = 13)) + 
+  labs(x ="", y = "")
+```

### 2023-06-03 jss/main.tex
misc fix

@@
   <chr>       <dbl> <dbl> <dbl> <chr>              <dbl> <list>           
-1 ASN00086038  145. -37.7  78.4 essendon airport   95866 <tibble>
-2 ASN00086077  145. -38.0  12.1 moorabbin airport  94870 <tibble>
-3 ASN00086282  145. -37.7 113.  melbourne airport  94866 <tibble>
+1 ASN00086038  145. -37.7  78.4 essendon airport   95866 <tibble [10 x 4]>
+2 ASN00086077  145. -38.0  12.1 moorabbin airport  94870 <tibble [10 x 4]>
+3 ASN00086282  145. -37.7 113.  melbourne airport  94866 <tibble [10 x 4]>
 \end{CodeOutput}
@@
 
-The toy example below demonstrates using dplyr verbs on a nested cubble:
+The toy example below demonstrates the using of a collection of dplyr functions (\code{mutate}, \code{select}, \code{filter}, and \code{arrange}) on a nested cubble object \code{cb_nested}:
 
@@
 \end{CodeInput}
-\begin{CodeOutput}
-# cubble:   key: id [2], index: date, nested form
-# extent:   [144.8321, -37.7276, 144.9066, -37.6655]
-# temporal: prcp [dbl], tmax [dbl], tmin [dbl]
-  id           long   lat name              wmo_id avg_temp ts               
-  <chr>       <dbl> <dbl> <chr>              <dbl>    <dbl> <list>           
-1 ASN00086282  145. -37.7 melbourne airport  94866     26.6 <tibble [10 x 4]>
-2 ASN00086038  145. -37.7 essendon airport   95866     26.5 <tibble [10 x 4]>
-\end{CodeOutput}
 \end{CodeChunk}
@@
 \begin{CodeOutput}
-[1] FALSE
+[1] TRUE
 \end{CodeOutput}
@@
 
-The five examples here are chosen to illustrate these aspects of the \pkg{cubble} package: creating a \code{cubble} object from two Coronavirus (COVID) data tables with the complication of differing location names, using spatial transformations to make a glyph map of seasonal temperature changes over years, aggregating information spatially to explore precipitation patterns, matching river level data and weather station records for analysis of water supply, reading NetCDF format data to reproduce a climate reanalysis plot, and the workflow to create complex interactive linked plots.
-(There is an additional example and figures in the Appendix, and more examples in the package vignettes.)
+The five examples here are chosen to illustrate these aspects of the \pkg{cubble} package: creating a \code{cubble} object from two Coronavirus (COVID) data tables with the complication of differing location names, using spatial transformations to make a glyph map of seasonal temperature changes over years, matching river level data and weather station records for analysis of water supply, reading NetCDF format data to reproduce a climate reanalysis plot, and the workflow to create complex interactive linked plots.
+(The Appendix provides another example on aggregating information spatially to explore precipitation patterns.)
 
@@
 These different sources need to be combined for the analysis, by matching the LGA names.
-Here is how to do this with the \pkg{cubble} package, including how to handle mismatches arising from different names of the same LGAs in the two tables.
-The COVID data is store in a csv file and looks like:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> covid |> head(5)
-\end{CodeInput}
-\begin{CodeOutput}
-# A tsibble: 5 x 5 [1D]
-# Key:       lga [1]
-# Groups:    lga, source [1]
-  date       lga        source                            n roll_mean
-  <date>     <chr>      <chr>                         <int>     <dbl>
-1 2022-01-01 Alpine (S) Contact with a confirmed case     1        NA
-2 2022-01-02 Alpine (S) Contact with a confirmed case     2        NA
-3 2022-01-03 Alpine (S) Contact with a confirmed case     4        NA
-4 2022-01-04 Alpine (S) Contact with a confirmed case     4        NA
-5 2022-01-05 Alpine (S) Contact with a confirmed case     2        NA
-\end{CodeOutput}
-\end{CodeChunk}
-
-and the spatial polygons are an ESRI shapefile as follows:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> lga |> head(5)
-\end{CodeInput}
-\begin{CodeOutput}
-Simple feature collection with 5 features and 1 field
-Geometry type: MULTIPOLYGON
-Dimension:     XY
-Bounding box:  xmin: 142.3535 ymin: -38.67876 xmax: 147.3909 ymax: -36.39269
-Geodetic CRS:  WGS 84
-               lga                       geometry
-132     Alpine (S) MULTIPOLYGON (((146.7258 -3...
-133    Ararat (RC) MULTIPOLYGON (((143.1807 -3...
-134   Ballarat (C) MULTIPOLYGON (((143.6622 -3...
-135    Banyule (C) MULTIPOLYGON (((145.1357 -3...
-136 Bass Coast (S) MULTIPOLYGON (((145.5207 -3...
-\end{CodeOutput}
-\end{CodeChunk}
-
-The function \code{make_cubble()} is used to create a \code{cubble} object from the two spatial and temporal tables, and requires specifying the arguments \code{key}, \code{index}, and \code{coords} (as described in Section \ref{create}).
-It will automatically try to match the sites in both tables using the location names and will show a warning message when there are mismatches, as shown below:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> cb <- make_cubble(spatial = lga, temporal = covid)
+Both the COVID count data (\code{covid}) and the LGA information (\code{lga}) are available in the \pkg{cubble} package with the \code{covid} data stored as a \code{tsibble} object, the \code{lga} data as an \code{sf} object.
+
+The \code{make_cubble()} function is used to create a \code{cubble} object from the two spatial and temporal tables as described in Section \ref{create}.
+The \code{by} argument allows for mismatch of key names from the two tables in the \code{*_join()} \code{by} syntax:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> cb <- make_cubble(lga, covid, by = c("lga_name_2018" = "lga"))
 \end{CodeInput}
@@
 
-It can be seen that there are two-way mismatches -- LGAs in the COVID data that do not match with LGAs names in the spatial polygon data, and vice versa.
-The mismatches can be identified by using the function \code{check_key}:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> (check_res <- check_key(spatial = lga, temporal = covid))
+The warning message reports on the mismatches of location from both sides: there are LGAs in the COVID data that do not match with LGAs names in the spatial polygon data, and vice versa.
+The \code{make_cubble()} function prompts the use of \code{check_key()} to identify the mismatches:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> (check_res <- check_key(
++   spatial = lga, temporal = covid, 
++   by = c("lga_name_2018" = "lga")
++ ))
 \end{CodeInput}
@@
 
-With this information both tables can be fixed, to create the desired \code{cubble} object, as follows:
+The check result is a list with three elements: 1) matched pairs from both tables, 2) those pairs that can be potentially paired, and 3) others.
+Here the main mismatch comes from different encodings of two LGAs: Kingston and Latrobe.
+Analysts can then modify the input spatial and temporal data accordingly and create the cubble again:
 
@@
 R> lga2 <- lga |>
++   rename(lga = lga_name_2018) |> 
 +   mutate(lga = ifelse(lga == "Kingston (C) (Vic.)", "Kingston (C)", lga),
@@
   <chr>          <dbl> <dbl>                         <MULTIPOLYGON [°]> <list>  
-1 Alpine (S)      147. -36.9 (((146.7258 -36.45922, 146.80~ <gropd_ts>
-2 Ararat (RC)     143. -37.5 (((143.1807 -37.73152, 143.17~ <gropd_ts>
-3 Ballarat (C)    144. -37.5 (((143.6622 -37.57241, 143.66~ <gropd_ts>
-4 Banyule (C)     145. -37.7 (((145.1357 -37.74091, 145.13~ <gropd_ts>
-5 Bass Coast (S)  146. -38.5 (((145.5207 -38.30667, 145.54~ <gropd_ts>
+1 Alpine (S)      147. -36.9 (((146.7258 -36.45922, 146.8033 -36.45139~ <tbl_ts>
+2 Ararat (RC)     143. -37.5 (((143.1807 -37.73152, 143.1793 -37.73242~ <tbl_ts>
+3 Ballarat (C)    144. -37.5 (((143.6622 -37.57241, 143.6686 -37.53844~ <tbl_ts>
+4 Banyule (C)     145. -37.7 (((145.1357 -37.74091, 145.1331 -37.74281~ <tbl_ts>
+5 Bass Coast (S)  146. -38.5 (((145.5207 -38.30667, 145.5418 -38.30938~ <tbl_ts>
 # i 75 more rows
@@
 +       )|>
-+   summarise(tmax = mean(tmax, na.rm = TRUE))
++   summarise(tmax = mean(tmax, na.rm = TRUE)) %>% 
++   mutate(group = as.factor(lubridate::year(yearmonth)),
++          month = lubridate::month(yearmonth))
 \end{CodeInput}
@@
 A quick check on the number of observations for each location is made, revealing that there are several with less than 24 observations -- these stations lack temperature values for some months.
-In this example, those stations are removed by switching to a long \code{cubble} to operate on the spatial component over time, and then, move back into the nested \code{cubble} (to make the glyph map):
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> tmax |> 
-+   face_spatial() |> 
-+   mutate(n = nrow(ts)) |>
-+   arrange(n) |>
-+   pull(n) |> 
-+   head(10)
-\end{CodeInput}
-\begin{CodeOutput}
- [1] 12 12 12 13 19 19 20 24 24 24
-\end{CodeOutput}
+In this example, those stations are removed by switching to the nested \code{cubble} to operate on the spatial component over time, and then, move back into the long \code{cubble} (to make the glyph map):
+
+\begin{CodeChunk}
 \begin{CodeInput}
@@
 \begin{CodeInput}
-R> nsw_vic <- ozmaps::abs_ste |>
-+   filter(NAME %in% c("Victoria", "New South Wales"))
-R> 
 R> ggplot() +
-+   geom_sf(data = nsw_vic,
-+           fill = "transparent", color = "grey",
-+           linetype = "dotted") +
++   geom_sf(data = MAP_DATA, ...) +
 +   geom_glyph(data = tmax,
@@
 Code is provided to accomplish this in the supplementary material.
-
-\begin{CodeChunk}
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/research/paper-cubble/figures/netcdf-1}
-
-}
-
-\caption[A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020) to illustrate the break-up of sourthern polar vortex in late September and early October 2002]{A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020) to illustrate the break-up of sourthern polar vortex in late September and early October 2002.
-The polar vortex, signalled by the high specific humidity, splits into two on 2002-09-26 and further splits into four on 2002-10-04.}\label{fig:netcdf}
-\end{figure}
-\end{CodeChunk}
 

### 2023-06-08 2-cubble.Rmd
update as per the pkg dev

@@
 
-A nested cubble (in the spatial form) has class \code{c("spatial_cubble_df", "cubble_df")} and is built from a rowwise tibble (\code{rowwise_df}).
-The nested cubble arranges spatial variables in columns and nests temporal variables in a specialised \code{ts} column.
-The \code{rowwise_df} class is imposed since it is helpful to manipulate with list column, i.e.
-compute per station summary of temporal variables, and you may remove it with \code{strip_rowwise()}.
-
+A nested cubble (in the spatial form) has class \code{c("spatial_cubble_df", "cubble_df")}.
+It arranges spatial variables in columns and nests temporal variables in a specialised \code{ts} column.
 Below prints an nested cubble object using the weather station data collected in three airports.
@@
 
-A long cubble (in the temporal form) has class \code{c("temporal_cubble_df", "cubble_df")}.
-It is built from a grouped tibble (\code{grouped_df}), which expands the temporal variables into the long form and stores the spatial variables as a data attribute.
+A long cubble (in the temporal form) has class \code{c("temporal_cubble_df", "cubble_df")}, which expands the temporal variables into the long form and stores the spatial variables as a data attribute.
 
@@
 
-## Functionalities
+## Functions and methods
 
-Table \@ref(tab:funs) summarises functions implemented in the \pkg{cubble} package.
+Table \@ref(tab:funs) summarises functions implemented in the \pkg{cubble} package and Table \@ref(tab:methods) details the methods implemented for each of the three cubble classes: \code{cubble\_df}, \code{spatial\_cubble\_df}, and \code{temporal\_cubble\_df}.
 
@@
 
-|Category | Details |  Functions|
-|--|---|----------------|
-|base R | subset | `[`, `[<-`, `names<-`, `[[`, `[[<-`| 
-|tidyverse | dplyr |`dplyr_row_slice`, `dplyr_col_modify`, `dplyr_reconstruct`, `select`, `mutate`, `arrange`, `filter`, `summarise`, `slice`, `group_by`, `ungroup`|
-|| tidyr| | 
-|| pillar| `type_sum` | 
-|sf| | | 
-|tsibble| | | 
-|cubble | basics| `face_temporal`, `face_spatial`, `unfold`|
-| | attributes| `key`, `key_vars`, `key_data`, `index`, `index_var`, `coords`, `spatial`|
-| |hierarchy|  `switch_key`|
-| |matching |   `match_sites`, `match_spatial`, `match_temporal`| 
-| |glyph map | `geom_glyph`, `geom_glyph_box`, `geom_glyph_line`|
+|Category |   Functions|
+|---|----------------|
+|base R | `[`, `[<-`, `names<-`, `[[`, `[[<-`| 
+|tidyverse |`dplyr_row_slice`, `dplyr_col_modify`, `dplyr_reconstruct`, `select`, `mutate`, `arrange`, `filter`, `summarise`, `select`, `slice`, `rowwise`, `rename`, `bind_cols`, `group_by`, `ungroup`, `slice_*`, `*_join`, `relocate`, `type_sum` |
+|cubble | `cubble`, `make_cubble`, `new_spatial_cubble`, `new_temporal_cubble`, `make_spatial_sf`, `make_temporal_tsibble`, `face_temporal`, `face_spatial`, `unfold`, `key`, `key_vars`, `key_data`, `index`, `index_var`, `coords`, `spatial`, `fill_gaps`, `scan_gaps`, `switch_key`, `match_sites`, `match_spatial`, `match_temporal`, `geom_glyph`, `geom_glyph_box`, `geom_glyph_line`|
+
+Table: (\#tab:methods) Methods implemented for the three \code{cubble} classes: \code{cubble\_df}, \code{spatial\_cubble\_df}, and \code{temporal\_cubble\_df}
+
+|Class | Methods | 
+| ----- | ----------------|
+|`cubble_df` | `[, dplyr_col_modify, key_data, key_vars, key, names<-, print, tbl_sum`|
+|`spatial_cubble_df` | `dplyr_reconstruct, dplyr_row_slice, face_spatial, face_temporal, arrange, rename, rowwise, group_by, ungroup, select, spatial, summarise, tbl_sum, unfold, update_cubble`|
+|`temporal_cubble_df` | `arrange, dplyr_reconstruct, dplyr_row_slice, face_spatial, face_temporal, fill_gaps, group_by, ungroup,  rename, rowwise, scan_gaps, select, spatial, summarise, tbl_sum, unfold, update_cubble`|
+
+<!-- methods(class = "cubble_df") |> paste0(collapse = ", ") |> stringr::str_remove_all(".cubble_df") -->
+<!-- methods(class = "spatial_cubble_df") |> paste0(collapse = ", ") |> stringr::str_remove_all(".spatial_cubble_df") -->
+<!-- methods(class = "temporal_cubble_df") |> paste0(collapse = ", ") |> stringr::str_remove_all(".temporal_cubble_df") -->
 
@@
 ```{r echo = TRUE}
-ts_long %>% tsibble::has_gaps()
+#ts_long %>% has_gaps()
 ```
@@
 
-```{r echo =TRUE}
-#sf_nested %>% sf::st_transform(crs = "EPSG:3857")
+```{r echo =TRUE, message=FALSE}
+sf_nested %>% sf::st_transform(crs = "EPSG:3857")
 ```
@@
 Consider the \code{climate_aus} data in the \pkg{cubble} package with `r nrow(cubble::climate_aus)` stations observed daily in a single year 2020.
-The created \code{sftime} object is `r library(cubble); round(object.size(sftime::st_as_sftime(climate_aus))/object.size(climate_aus), 2)` times larger than the corresponding \code{cubble} object (`r round(object.size(sftime::st_as_sftime(climate_aus))/1e6, 2)` MB vs.
+The created \code{sftime} object is approximately `r library("cubble"); round(object.size(sftime::st_as_sftime(climate_aus))/object.size(climate_aus))` times larger than the corresponding \code{cubble} object (`r round(object.size(sftime::st_as_sftime(climate_aus))/1e6, 2)` MB vs.
 `r round(object.size(cubble::climate_aus)/1e6, 2)` MB).

### 2023-06-08 4-2-glyph.Rmd
update as per the pkg dev

@@
 ```{r}
-
 load(here::here("data/historical_tmax.rda"))
@@
   face_spatial() |> 
+  rowwise() %>% 
   filter(nrow(ts) == 24) |>

### 2023-06-08 jss/main.Rmd
section 2.1 getting into shape
update as per the pkg dev

@@
   \usepackage{array}
-  \usepackage{xcolor}
-  \newcommand{\sherry}[1]{\textcolor{magenta}{#1}}
 output: 
@@
         tibble.print_max = 5, tibble.print_min = 5,
-        width = 90, styler.cache_root = "styler-perm")
+        width = 80, styler.cache_root = "styler-perm")
 ```
@@
 library(patchwork)
+library(tsibble)
+library(sf)
 ```

### 2023-06-08 jss/main.tex
update as per the pkg dev

@@
 
-\usepackage{amsmath} \usepackage{array} \usepackage{xcolor} \newcommand{\sherry}[1]{\textcolor{magenta}{#1}}
+\usepackage{amsmath} \usepackage{array}
 
@@
 
-A nested cubble (in the spatial form) has class \code{c("spatial_cubble_df", "cubble_df")} and is built from a rowwise tibble (\code{rowwise_df}).
-The nested cubble arranges spatial variables in columns and nests temporal variables in a specialised \code{ts} column.
-The \code{rowwise_df} class is imposed since it is helpful to manipulate with list column, i.e.~compute per station summary of temporal variables, and you may remove it with \code{strip_rowwise()}.
-
+A nested cubble (in the spatial form) has class \code{c("spatial_cubble_df", "cubble_df")}.
+It arranges spatial variables in columns and nests temporal variables in a specialised \code{ts} column.
 Below prints an nested cubble object using the weather station data collected in three airports.
@@
 # cubble:   key: id [3], index: date, nested form
-# extent:   [144.8321, -37.98, 145.0964, -37.6655]
-# temporal: prcp [dbl], tmax [dbl], tmin [dbl]
+# spatial:  [144.8321, -37.98, 145.0964, -37.6655], Missing CRS!
+# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
   id           long   lat  elev name              wmo_id ts               
@@
 \begin{CodeOutput}
-[1] "spatial_cubble_df" "cubble_df"         "rowwise_df"       
-[4] "tbl_df"            "tbl"               "data.frame"       
+[1] "spatial_cubble_df" "cubble_df"         "tbl_df"           
+[4] "tbl"               "data.frame"       
 \end{CodeOutput}
@@
 # cubble:   key: id [3], index: date, long form
-# extent:  2020-01-01 -- 2020-01-10 [1D], no gaps
+# temporal: 2020-01-01 -- 2020-01-10 [1D], no gaps
 # spatial:  long [dbl], lat [dbl], elev [dbl], name [chr], wmo_id [dbl]
@@
 # A tibble: 3 x 6
-# Rowwise: 
   id           long   lat  elev name              wmo_id
@@
 # cubble:   key: id [26565], index: time, nested form
-# extent:   [113, -53, 153, -12]
-# temporal: q [dbl], z [dbl]
+# spatial:  [113, -53, 153, -12], Missing CRS!
+# temporal: time [dttm], q [dbl], z [dbl]
      id  long   lat ts              
@@
 
-\hypertarget{functionalities}{%
-\subsection{Functionalities}\label{functionalities}}
-
-Table \ref{tab:funs} summarises functions implemented in the \pkg{cubble} package.
+\hypertarget{functions-and-methods}{%
+\subsection{Functions and methods}\label{functions-and-methods}}
+
+Table \ref{tab:funs} summarises functions implemented in the \pkg{cubble} package and Table \ref{tab:methods} details the methods implemented for each of the three cubble classes: \code{cubble\_df}, \code{spatial\_cubble\_df}, and \code{temporal\_cubble\_df}.
 
 \begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.0952}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1429}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.7619}}@{}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.1579}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.8421}}@{}}
 \caption{\label{tab:funs} Functions implemented in the \pkg{cubble} package}\tabularnewline
@@
 Category
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Details
 \end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
@@
 \end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Details
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
 Functions
@@
 \endhead
-base R & subset & \texttt{{[}}, \texttt{{[}\textless{}-}, \texttt{names\textless{}-}, \texttt{{[}{[}}, \texttt{{[}{[}\textless{}-} \\
-tidyverse & dplyr & \texttt{dplyr\_row\_slice}, \texttt{dplyr\_col\_modify}, \texttt{dplyr\_reconstruct}, \texttt{select}, \texttt{mutate}, \texttt{arrange}, \texttt{filter}, \texttt{summarise}, \texttt{slice}, \texttt{group\_by}, \texttt{ungroup} \\
-& tidyr & \\
-& pillar & \texttt{type\_sum} \\
-sf & & \\
-tsibble & & \\
-cubble & basics & \texttt{face\_temporal}, \texttt{face\_spatial}, \texttt{unfold} \\
-& attributes & \texttt{key}, \texttt{key\_vars}, \texttt{key\_data}, \texttt{index}, \texttt{index\_var}, \texttt{coords}, \texttt{spatial} \\
-& hierarchy & \texttt{switch\_key} \\
-& matching & \texttt{match\_sites}, \texttt{match\_spatial}, \texttt{match\_temporal} \\
-& glyph map & \texttt{geom\_glyph}, \texttt{geom\_glyph\_box}, \texttt{geom\_glyph\_line} \\
+base R & \texttt{{[}}, \texttt{{[}\textless{}-}, \texttt{names\textless{}-}, \texttt{{[}{[}}, \texttt{{[}{[}\textless{}-} \\
+tidyverse & \texttt{dplyr\_row\_slice}, \texttt{dplyr\_col\_modify}, \texttt{dplyr\_reconstruct}, \texttt{select}, \texttt{mutate}, \texttt{arrange}, \texttt{filter}, \texttt{summarise}, \texttt{select}, \texttt{slice}, \texttt{rowwise}, \texttt{rename}, \texttt{bind\_cols}, \texttt{group\_by}, \texttt{ungroup}, \texttt{slice\_*}, \texttt{*\_join}, \texttt{relocate}, \texttt{type\_sum} \\
+cubble & \texttt{cubble}, \texttt{make\_cubble}, \texttt{new\_spatial\_cubble}, \texttt{new\_temporal\_cubble}, \texttt{make\_spatial\_sf}, \texttt{make\_temporal\_tsibble}, \texttt{face\_temporal}, \texttt{face\_spatial}, \texttt{unfold}, \texttt{key}, \texttt{key\_vars}, \texttt{key\_data}, \texttt{index}, \texttt{index\_var}, \texttt{coords}, \texttt{spatial}, \texttt{fill\_gaps}, \texttt{scan\_gaps}, \texttt{switch\_key}, \texttt{match\_sites}, \texttt{match\_spatial}, \texttt{match\_temporal}, \texttt{geom\_glyph}, \texttt{geom\_glyph\_box}, \texttt{geom\_glyph\_line} \\
+\bottomrule()
+\end{longtable}
+
+\begin{longtable}[]{@{}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.2381}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.7619}}@{}}
+\caption{\label{tab:methods} Methods implemented for the three \code{cubble} classes: \code{cubble\_df}, \code{spatial\_cubble\_df}, and \code{temporal\_cubble\_df}}\tabularnewline
+\toprule()
+\begin{minipage}[b]{\linewidth}\raggedright
+Class
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Methods
+\end{minipage} \\
+\midrule()
+\endfirsthead
+\toprule()
+\begin{minipage}[b]{\linewidth}\raggedright
+Class
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Methods
+\end{minipage} \\
+\midrule()
+\endhead
+\texttt{cubble\_df} & \texttt{{[},\ dplyr\_col\_modify,\ key\_data,\ key\_vars,\ key,\ names\textless{}-,\ print,\ tbl\_sum} \\
+\texttt{spatial\_cubble\_df} & \texttt{dplyr\_reconstruct,\ dplyr\_row\_slice,\ face\_spatial,\ face\_temporal,\ arrange,\ rename,\ rowwise,\ group\_by,\ ungroup,\ select,\ spatial,\ summarise,\ tbl\_sum,\ unfold,\ update\_cubble} \\
+\texttt{temporal\_cubble\_df} & \texttt{arrange,\ dplyr\_reconstruct,\ dplyr\_row\_slice,\ face\_spatial,\ face\_temporal,\ fill\_gaps,\ group\_by,\ ungroup,\ \ rename,\ rowwise,\ scan\_gaps,\ select,\ spatial,\ summarise,\ tbl\_sum,\ unfold,\ update\_cubble} \\
 \bottomrule()
@@
 # cubble:   key: id [3], index: date, long form, [tsibble]
-# extent:  2020-01-01 -- 2020-01-10 [1D], no gaps
+# temporal: 2020-01-01 -- 2020-01-10 [1D], no gaps
 # spatial:  long [dbl], lat [dbl], elev [dbl], name [chr], wmo_id [dbl]
@@
 \begin{CodeInput}
-R> ts_long %>% tsibble::has_gaps()
-\end{CodeInput}
-\begin{CodeOutput}
-# A tibble: 3 x 2
-  id          .gaps
-  <chr>       <lgl>
-1 ASN00086038 FALSE
-2 ASN00086077 FALSE
-3 ASN00086282 FALSE
-\end{CodeOutput}
+R> #ts_long %>% has_gaps()
+\end{CodeInput}
 \end{CodeChunk}
@@
 # cubble:   key: id [3], index: date, nested form, [sf]
-# extent:   [144.8321, -37.98, 145.0964, -37.6655]
-# temporal: prcp [dbl], tmax [dbl], tmin [dbl]
-  id           long   lat  elev name   wmo_id            geometry ts      
-  <chr>       <dbl> <dbl> <dbl> <chr>   <dbl>         <POINT [°]> <list>  
-1 ASN00086038  145. -37.7  78.4 essen~  95866 (144.9066 -37.7276) <tibble>
-2 ASN00086077  145. -38.0  12.1 moora~  94870   (145.0964 -37.98) <tibble>
-3 ASN00086282  145. -37.7 113.  melbo~  94866 (144.8321 -37.6655) <tibble>
+# spatial:  [144.8321, -37.98, 145.0964, -37.6655], WGS 84
+# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
+  id           elev name   wmo_id  long   lat            geometry ts      
+  <chr>       <dbl> <chr>   <dbl> <dbl> <dbl>         <POINT [°]> <list>  
+1 ASN00086038  78.4 essen~  95866  145. -37.7 (144.9066 -37.7276) <tibble>
+2 ASN00086077  12.1 moora~  94870  145. -38.0   (145.0964 -37.98) <tibble>
+3 ASN00086282 113.  melbo~  94866  145. -37.7 (144.8321 -37.6655) <tibble>
 \end{CodeOutput}
@@
 [1] "spatial_cubble_df" "cubble_df"         "sf"               
-[4] "rowwise_df"        "tbl_df"            "tbl"              
-[7] "data.frame"       
+[4] "tbl_df"            "tbl"               "data.frame"       
 \end{CodeOutput}
@@
 \begin{CodeInput}
-R> #sf_nested %>% sf::st_transform(crs = "EPSG:3857")
-\end{CodeInput}
+R> sf_nested %>% sf::st_transform(crs = "EPSG:3857")
+\end{CodeInput}
+\begin{CodeOutput}
+# cubble:   key: id [3], index: date, nested form, [sf]
+# spatial:  [16122635.6225205, -4576600.8687746, 16152057.3639371,
+#   -4532279.35567565], WGS 84
+# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
+  id           elev name   wmo_id  long   lat            geometry ts      
+  <chr>       <dbl> <chr>   <dbl> <dbl> <dbl>         <POINT [°]> <list>  
+1 ASN00086038  78.4 essen~  95866  145. -37.7 (16130929 -4541016) <tibble>
+2 ASN00086077  12.1 moora~  94870  145. -38.0 (16152057 -4576601) <tibble>
+3 ASN00086282 113.  melbo~  94866  145. -37.7 (16122636 -4532279) <tibble>
+\end{CodeOutput}
 \end{CodeChunk}
@@
 Consider the \code{climate_aus} data in the \pkg{cubble} package with 639 stations observed daily in a single year 2020.
-The created \code{sftime} object is 13.87 times larger than the corresponding \code{cubble} object (118.24 MB vs.~8.52 MB).
+The created \code{sftime} object is approximately 14 times larger than the corresponding \code{cubble} object (118.24 MB vs.~8.43 MB).
 
@@
 # cubble:   key: lga [80], index: date, nested form, [sf]
-# extent:   [140.961682, -39.1591895, 149.976291, -33.9804256]
-# temporal: source [chr], n [int], roll_mean [dbl]
+# spatial:  [140.961682, -39.1591895, 149.976291, -33.9804256], WGS 84
+# temporal: date [date], n [dbl], avg_7day [dbl]
   lga             long   lat                                   geometry ts      
@@
 +   face_spatial() |> 
++   rowwise() %>% 
 +   filter(nrow(ts) == 24) |>

### 2023-06-10 3-advanced.Rmd
fix citation on websites
matching code update and misc

@@
 
-Spatial locations can have grouping structures either inherent to the data (e.g.
-state within country) or obtained during the analysis (e.g.
-cluster id).
-In this case, it can be useful to summarise variables at various levels of the hierarchy.
+Spatial locations can have grouping structures either inherent to the data e.g., state within country or obtained during the analysis e.g., cluster id.
+In this case, it can be useful to summarize variables at various levels of the hierarchy.
 The function \code{switch_key()} can be used to change the grouping level of spatial locations.
@@
 The result can also be stretched into long form.
-By specifying
-
-\code{cluster_nested <- station_nested \%>\% switch_key(key = cluster)},
-
-the cubble redefines the cubble key from the \code{id} column in \code{station_nested} to the \code{cluster} column in \code{cluster_nested}.
+By specifying \code{cluster_nested <- station_nested \%>\% switch_key(key = cluster)}, the \code{cubble} object redefines its \code{key} from the \code{id} column in \code{station_nested} to the \code{cluster} column in \code{cluster_nested}.
 All the spatial variables belonging to the \code{cluster} column are now nested into a \code{.val} column which allows for summarizing based on cluster.
@@
 Once the change is made the data can be stretched into the long form."}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.003.png"))
+#knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.003.png"))
 ```
@@
 Figure \ref{fig:illu-matching} illustrates this temporal matching.
-In the two series, \code{A} and \code{a}, three peaks have been identified in each.
-An interval, of fixed length, is constructed for each peak in series \code{A}, while the peaks in series \code{a} are tested against whether they fall into any of the intervals.
+In the two series, *A* and *a*, three peaks have been identified in each.
+An interval, of fixed length, is constructed for each peak in series *A*, while the peaks in series *a* are tested against whether they fall into any of the intervals.
 Here two out of three peaks match.
@@
 
-The cubble workflow fits works well with an interactive graphics pipeline (e.g.
-@buja1988elements, @buja1996interactive,  @sutherland2000orca, @xie2014reactive, @cheng2016enabling) that is available in R with the package crosstalk [@crosstalk].
-Figure \ref{fig:illu-interactive} illustrates how linking can be achieved between a map and multiple time series using a cubble.
-The map (produced from the nested form) and time series (produced from the long form) are both shared crosstalk objects.
-When a user makes a selection on the map, the site is highlighted (left).
-This activates a row in the nested cubble object, which is then communicated to the long cubble -- all the observations with the same \code{id} (middle) will be selected.
-The long cubble will then highlight the corresponding series in the time series plot (right).
+The workflow with the \code{cubble} class works well with an interactive graphics pipeline (e.g., @buja1988elements; @buja1996interactive;  @sutherland2000orca; @xie2014reactive; @cheng2016enabling) that is available in \proglang{R} with the package \pkg{crosstalk} [@crosstalk].
+Figure \ref{fig:illu-interactive} illustrates how linking can be achieved between a map and multiple time series in a \code{cubble} object.
+The map (produced from the nested form) and time series (produced from the long form) are both shared \code{crosstalk} objects.
+When a user makes a selection on the map, the site is highlighted (a).
+This activates a row in the nested \code{cubble}, which is then communicated to the long \code{cubble} -- all the observations with the same id (b) will be selected.
+The long \code{cubble} will then highlight the corresponding series in the time series plot (c).
 
 Linking is also available starting from the time series plot, by selecting points.
-This will be activate rows having the same \code{id} in the long cubble.
-The corresponding rows in the nested cubble are activated, and highlighted the map.
+This will activate rows having the same id in the long \code{cubble}.
+The corresponding rows in the nested \code{cubble} are activated, and highlighted on the map.
 (An illustration can be found in the appendix.) Note that this type of linking, both from the map or the time series, is what @CS07 would call categorical variable linking, where station id is the categorical variable.
@@
 
-```{r illu-interactive, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "Linking between multiple plots.
-The line plots and the map are constructed from shared crosstalk objects (long and nested cubbles).
-When a station is selected on the map (a), the corresponding row in the nested cubble will be activated.
-This will link to all the rows with the same id in the long cubble (b) and update the line plot (c)."}
+```{r illu-interactive, echo = FALSE, fig.align="center", out.height = "35%", out.width = "100%", fig.cap = "Linking between multiple plots.
+The line plots and the map are constructed from shared \\code{crosstalk} objects (long and nested \\code{cubble}).
+When a station is selected on the map (a), the corresponding row in the nested \\code{cubble} will be activated.
+This will link to all the rows with the same id in the long \\code{cubble} (b) and update the line plot (c)."}
 knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.004.png"))
@@
 Spatio-temporal data lends itself to a range of transformations.
-Glyph maps (Section \ref{unfold}) transform the measured variable and time coordinates into microplots at the spatial locations.
+Glyph maps [@Wickham2012-yr] transform the measured variable and time coordinates into microplots at the spatial locations.
 Calendar plots [@wang2020calendar] deconstruct time to produce plots of variables in a calendar format.
-Summarising multiple variables is commonly done using projections, or linear combinations.
+Summarizing multiple variables is commonly done using projections, or linear combinations.
 Here we elaborate on the transformations made to produce a glyph map.
 
-The \proglang{R} package \pkg{GGally} implements glyph maps through the \code{glyphs()} function.
-The function constructs a data frame with calculated position (\code{gx}, \code{gy}, \code{gid}) of each point on the time series using linear algebra (Equations 1 and 2 in @Wickham2012-yr).
-The data can then be piped into \code{ggplot} to create the glyph map as:
+The package \pkg{GGally} [@ggally] has implemented glyph maps through the \code{glyphs()} function.
+The function constructs a \code{data.frame} with calculated position (\code{gx}, \code{gy}, \code{gid}) of each point on the time series using linear algebra (Equations 1 and 2 in @Wickham2012-yr).
+The data can then be piped into ggplot to create the glyph map as:
 
 ```{r eval = FALSE, echo = TRUE}
-library(ggplot2)
+library("ggplot2")
 gly <- glyphs(data,
@@
 
-An example using a glyph map is shown in Section \ref{precip}.
+An example using a glyph map is shown in Section \@ref(historicaltmax).
 
 Some useful controls over the glyph map are also available in the \code{geom_glyph()} implementation.
-Polar coordinate glyph maps are specified using \code{polar = TRUE}, and parameters \code{width} and \code{height} can be specified in either absolute or relative value.
+Polar coordinate glyph maps are specified using \code{polar = TRUE}, and arguments \code{width} and \code{height} can be specified in either absolute or relative value.
 Global and local scale is specified with \code{global_rescale}, which defaults to \code{TRUE}.

### 2023-06-10 4-3-matching.Rmd
matching code update and misc

--- 2023-03-25 5-3-river-level-rainfall.Rmd
@@
 ```{r}
-load(here::here("data/climate_full.rda"))
-
-climate <- climate_full |>
+climate_vic <- climate_aus |>
   filter(between(as.numeric(stringr::str_sub(id, 7, 8)), 76, 90)) |>
-  face_temporal() |>
-  filter(lubridate::year(date) == 2020) |>
-  face_spatial() |>
   mutate(type = "climate")
 
-river <- river |> mutate(type = "river")
+river <- cubble::river %>% mutate(type = "river")
+
+vic_map <- ozmaps::abs_ste %>%  filter(NAME == "Victoria")
+
+# ggplot() + 
+#   geom_sf(data = vic_map) + 
+#   geom_point(data = dplyr::bind_rows(river, climate_vic), 
+#              aes(x = long, y = lat, color = type)) + 
+#   theme_bw() +
+#   theme(legend.position = "bottom") +
+#   labs(x = "Longitude", y = "Latitude") + 
+#   scale_color_brewer(palette = "Dark2")
 ```
@@
 
+```{r echo = TRUE,  tibble.print_max = 10,   tibble.print_min = 10}
+(res_sp <- match_spatial(climate_vic, river, spatial_n_group = 10))
+```
+
+```{r echo = FALSE}
+res_sp <- res_sp %>% 
+  filter(!group %in% c(5, 8)) %>% 
+  tidyr::pivot_longer(from:to, names_to = "order", values_to = "id") %>% 
+  select(-order)
+res_sp <- bind_rows(climate_vic, river %>% rename(id = station)) %>% 
+  inner_join(res_sp) 
+```
+
+```{r}
+res_sp
+```
+
+```{r echo = TRUE}
+(res_tm <- res_sp %>% 
+  match_temporal(
+    match_id = group,
+    temporal_by = c("prcp" = "Water_course_level"),
+    temporal_n_highest = 30,
+    temporal_min_match = 15))
+```
+
+Or use \code{return_cubble = TRUE} to return a cubble object:
+
+```{r}
+(res_tm <- res_sp %>% 
+  match_temporal(
+    match_id = group,
+    temporal_by = c("prcp" = "Water_course_level"),
+    return_cubble = TRUE, 
+    temporal_n_highest = 30,
+    temporal_min_match = 15))
+```
+
 <!--We first pass in the two datasets, followed by the variables used for the temporal matching, specified in \code{temporal_by} (using the \code{by} syntax from \code{dplyr::join}).
@@
 
-```{r echo = TRUE}
-res <- match_sites(
-  river, climate,
-  temporal_by = c("Water_course_level" = "prcp"),
-  temporal_independent = "prcp",  
-  temporal_n_highest = 30,
-  temporal_min_match = 15, 
-)
-```
-
 This function returns a \code{cubble} object, with additional columns: \code{dist} storing the distance between matched stations, \code{group} summarizing spatial matching, and \code{n_match} showing the temporal matching.
-
-```{r}
-res |> print()
-```
 
@@
 The numbers (1, 5, 6, 10) indicate the group index derived from spatial matching, only those that were selectd by temporal matching are shown here."}
+res_tm <- res_tm %>%  filter(group %in% c(1, 7, 6, 9))
+
+res_tm_long <- res_tm %>%  
+  face_temporal() %>%  
+  unfold(group, type) %>%  
+  rename(prcp = matched) %>%  
+  group_by(group, type) %>%
+  mutate(prcp = (prcp - min(prcp, na.rm = TRUE))/ (max(prcp, na.rm = TRUE) - min(prcp, na.rm = TRUE)))
+
 vic_map <- ozmaps::abs_ste |> 
@@
   geom_sf(data = vic_map, fill = "grey95", color = "white") + 
-  geom_point(data = dplyr::bind_rows(river, climate), 
+  geom_point(data = dplyr::bind_rows(river, climate_vic), 
              aes(x = long, y = lat, color = type), alpha = 0.2, fill = 0.2) +
-  geom_point(data = res, aes(x = long, y = lat, color = type)) +
+  geom_point(data = res_tm, aes(x = long, y = lat, color = type)) +
   ggrepel::geom_label_repel(
-    data = res |> filter(type == "river"), aes(x = long, y = lat, label = group)) +
+    data = res_tm |> filter(type == "river"), aes(x = long, y = lat, label = group)) +
   scale_color_brewer(palette = "Dark2")  + 
@@
 
-res_long <- res |>
-  face_temporal(ts) |>
-  unfold(group, type) |>
-  rename(prcp = matched_var) |> 
-  mutate(prcp = (prcp - min(prcp, na.rm = TRUE))/ (max(prcp, na.rm = TRUE) - min(prcp, na.rm = TRUE)))
-
-p2 <- res_long |> 
-  ggplot(aes(x = date, y = prcp, color = type, group = id)) +
-  geom_line(alpha = 0.8) +
+p2 <- res_tm_long %>%  
+  ggplot(aes(x = date, y = prcp, group = type,color = type)) + 
+  geom_line() + 
   facet_wrap(vars(group)) + 
@@
   theme_bw() + 
+  labs(x=  "date") + 
   scale_x_date(date_labels = "%b") + 
-  labs(x = "Date", y = "Precipitation/ water level")
+  labs(x = "Week", y = "Precipitation/ water level")
 

### 2023-06-10 4-4-netcdf.Rmd
matching code update and misc

@@
 
-```{r netcdf, eval =FALSE, out.width="100%", fig.height = 4, fig.width = 10, fig.cap = "A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020) to illustrate the break-up of sourthern polar vortex in late September and early October 2002.
+```{r netcdf, out.width="100%", fig.height = 4, fig.width = 10, fig.cap = "A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020) to illustrate the break-up of sourthern polar vortex in late September and early October 2002.
 The polar vortex, signalled by the high specific humidity, splits into two on 2002-09-26 and further splits into four on 2002-10-04."}

### 2023-06-10 jss/main.tex
matching code update and misc

@@
 \begin{CodeInput}
-R> #ts_long %>% has_gaps()
-\end{CodeInput}
+R> ts_long %>% has_gaps()
+\end{CodeInput}
+\begin{CodeOutput}
+# A tibble: 3 x 2
+  id          .gaps
+  <chr>       <lgl>
+1 ASN00086038 FALSE
+2 ASN00086077 FALSE
+3 ASN00086282 FALSE
+\end{CodeOutput}
 \end{CodeChunk}
@@
 Consider the \code{climate_aus} data in the \pkg{cubble} package with 639 stations observed daily in a single year 2020.
-The created \code{sftime} object is approximately 14 times larger than the corresponding \code{cubble} object (118.24 MB vs.~8.43 MB).
+The created \code{sftime} object is approximately 14 times larger than the corresponding \code{cubble} object (118.24 MB vs.~8.52 MB).
 
@@
 All the spatial variables belonging to the \code{cluster} column are now nested into a \code{.val} column which allows for summarizing based on cluster.
-
-\begin{CodeChunk}
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{../figures/diagram-keynotes/diagram-keynotes.003}
-
-}
-
-\caption[Hierarchical spatial structure can be handled using \code{switch\_key()}, to create summaries based on any level]{Hierarchical spatial structure can be handled using \code{switch\_key()}, to create summaries based on any level.
-Here the switch is between the station id and a cluster id.
-Once the change is made the data can be stretched into the long form.}\label{fig:illu-hier}
-\end{figure}
-\end{CodeChunk}
 
@@
 \begin{CodeInput}
-R> res <- match_sites(
-+   river, climate,
-+   temporal_by = c("Water_course_level" = "prcp"),
-+   temporal_independent = "prcp",  
+R> (res_sp <- match_spatial(climate_vic, river, spatial_n_group = 10))
+\end{CodeInput}
+\begin{CodeOutput}
+# A tibble: 10 x 4
+  from        to      dist group
+  <chr>       <chr>    [m] <int>
+1 ASN00088051 406213 1838.     1
+2 ASN00084145 222201 2185.     2
+3 ASN00085072 226027 3282.     3
+4 ASN00080015 406704 4034.     4
+5 ASN00085298 226027 4207.     5
+# i 5 more rows
+\end{CodeOutput}
+\end{CodeChunk}
+
+\begin{CodeChunk}
+\begin{CodeOutput}
+# cubble:   key: id [16], index: date, nested form
+# spatial:  [144.5203, -38.144913, 148.4667, -36.128657], Missing CRS!
+# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
+  id           long   lat  elev name           wmo_id ts       type   dist group
+  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>   <chr>   [m] <int>
+1 ASN00080015  145. -36.2  96   echuca  aerod~  94861 <tibble> clim~ 4034.     4
+2 ASN00080091  145. -36.3 105   kyabram dpi     95833 <tibble> clim~ 7460.    10
+3 ASN00082042  146. -36.8 502   strathbogie     95843 <tibble> clim~ 6153.     6
+4 ASN00084145  148. -37.7  62.7 orbost          95918 <tibble> clim~ 2185.     2
+5 ASN00085072  147. -38.1   4.6 east sale air~  94907 <tibble> clim~ 3282.     3
+# i 11 more rows
+\end{CodeOutput}
+\end{CodeChunk}
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> (res_tm <- res_sp %>% 
++   match_temporal(
++     match_id = group,
++     temporal_by = c("prcp" = "Water_course_level"),
 +     temporal_n_highest = 30,
-+   temporal_min_match = 15, 
-+ )
-\end{CodeInput}
++     temporal_min_match = 15))
+\end{CodeInput}
+\begin{CodeOutput}
+# A tibble: 8 x 2
+  group n_matches
+  <int>     <dbl>
+1     1        30
+2     7        26
+3     6        23
+4     9        21
+5     4        20
+# i 3 more rows
+\end{CodeOutput}
+\end{CodeChunk}
+
+Or use \code{return_cubble = TRUE} to return a cubble object:
+
+\begin{CodeChunk}
+\begin{CodeOutput}
+# cubble:   key: id [16], index: date, nested form
+# spatial:  [144.5203, -38.144913, 148.4667, -36.128657], Missing CRS!
+# temporal: date [date], matched [dbl]
+  id          long   lat  elev name  wmo_id type   dist group ts       n_matches
+  <chr>      <dbl> <dbl> <dbl> <chr>  <dbl> <chr>   [m] <int> <list>       <dbl>
+1 ASN000880~  145. -37.0 290   rede~  94859 clim~ 1838.     1 <tibble>        30
+2 406213      145. -37.0  NA   CAMP~     NA river 1838.     1 <tibble>        30
+3 ASN000841~  148. -37.7  62.7 orbo~  95918 clim~ 2185.     2 <tibble>         5
+4 222201      148. -37.7  NA   SNOW~     NA river 2185.     2 <tibble>         5
+5 ASN000850~  147. -38.1   4.6 east~  94907 clim~ 3282.     3 <tibble>        14
+# i 11 more rows
+\end{CodeOutput}
 \end{CodeChunk}
@@
 Figure \ref{fig:matching} shows four matched pairs on the map (a) and standardized data as time series (b). The expected concurrent increase in precipitation and water level can be seen clearly.
+
+\begin{CodeChunk}
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/research/paper-cubble/figures/matching-1}
+
+}
+
+\caption[Weather stations and river gauges with matched pairs labelled on the map (a) and plotted across time (b)]{Weather stations and river gauges with matched pairs labelled on the map (a) and plotted across time (b).
+Precipitation and water level have been standardised between 0 and 1 to be displayed on the same scale.
+The water level reflects the increase in precipitation.
+The numbers (1, 5, 6, 10) indicate the group index derived from spatial matching, only those that were selectd by temporal matching are shown here.}\label{fig:matching}
+\end{figure}
+\end{CodeChunk}
 
@@
 Code is provided to accomplish this in the supplementary material.
+
+\begin{CodeChunk}
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/research/paper-cubble/figures/netcdf-1}
+
+}
+
+\caption[A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020) to illustrate the break-up of sourthern polar vortex in late September and early October 2002]{A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020) to illustrate the break-up of sourthern polar vortex in late September and early October 2002.
+The polar vortex, signalled by the high specific humidity, splits into two on 2002-09-26 and further splits into four on 2002-10-04.}\label{fig:netcdf}
+\end{figure}
+\end{CodeChunk}
 

### 2023-06-14 2-cubble.Rmd
matching code update and misc
update the cubble class intro as the pkg vignette

@@
 
-[conceptual framework] The \code{cubble} class tackles the spatio-temporal data collected at unique fixed locations and allowed for irregularity in the temporal dimension, i.e.
-irregular time steps and missingness.
+[conceptual framework] Spatio-temporal data comes various spatial and temporal characteristics and requires different data structures to wrangle: climate weather stations are recorded at fixed point location but with potential temporal data quality issue (missingness on the day); GPS data tracks unique point locations at different timestamps; satellite imageries captures snapshots of landscape at selected time.
+The type of spatio-temporal data cubble tackles are those collected at unique fixed locations while allowed for irregularity in the temporal dimension, like the weather station data.
+In the four layouts presented by the spacetime paper [@spacetime], cubble handles full space-time and sparse space-time layouts.
 
-The \code{cubble} class (\code{cubble_df}) is an S3 class [@advr] built on the \code{tibble} class, specifically to organize spatio-temporal data. 
-It uses three attributes to character the spatio-temporal data:
+The cubble class is an S3 class, built on tibble, to pivot spatio-temporal data into a nested (spatial) form and a long (temporal) form. It has two subclasses:
 
-  * \code{key} identifies the spatial locations 
-  * \code{date} identifies the temporal order, and 
-  * \code{coords} identifies the longitude and latitude coordinate associated with the location.
+-   a nested/ spatial cubble has class `c("spatial_cubble_df", "cubble_df")`
+-   a long/ temporal cubble has class `c("temporal_cubble_df", "cubble_df")`
 
-The base class \code{cubble_df} has two specialised class \code{spatial_cubble_df} and \code{temporal_cubble_df} to arrange the spatio-temporal data in two forms, as illustrated in Figure \ref{fig:class}.
-
-```{r class, fig.align="center", out.width = "100%", fig.cap = "Illustration of a cubble object in the long form and nested form, along with the associated attributes."}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.007.png"))
-```
-
-### nested cubble
-
-A nested cubble (in the spatial form) has class \code{c("spatial_cubble_df", "cubble_df")}.
-It arranges spatial variables in columns and nests temporal variables in a specialised \code{ts} column.
-Below prints an nested cubble object using the weather station data collected in three airports.
-The data is a subset of a larger data \code{climate_aus} collected from National Oceanic and Atmospheric Administration (NOAA) Global Historical Climatology Network Daily (GHCND).
-The data include spatial variables such as station ID, longitude, latitude, elevation, station name, world meteorology organisation ID, in additional to daily temporal information, maximum and minimum temperature values and precipitation for the year 2020.
-The airport data is a subset of three airport stations in Melbourne for the initial 10 days of the year (the creation of a cubble object will be explained in Section \@ref(create)):
+A nested cubble arranges spatial variables in columns and nests temporal variables in a specialised `ts` column.
+Below prints an nested cubble object using the weather station data collected in three airports (the creation of a cubble object will be explained in Section \@ref(create)).
+This toy data is a subset of a larger data \code{climate_aus} collected from Global Historical Climatology Network Daily (GHCND).
+The three airport stations in Melbourne are recorded with station metadata: station ID, longitude, latitude, elevation, station name, world meteorology organisation ID.
+The temporal variables are precipitation, maximum and minimum temperature, which can be read from the cubble header:
 
@@
 cb_nested
-class(cb_nested)
 ```
 
-A nested cubble header prints the key and index variable name and its form.
+<!-- A nested cubble header prints the key and index variable name and its form.
 The spatial extent is given by the bounding box with CRS information (missing here).
 The nested cubble header also shows the name of temporal variables nested in the \code{ts} column and their variable types: \code{prcp [dbl], tmax [dbl], tmin [dbl]}.
+-->
 
-### long cubble
-
-A long cubble (in the temporal form) has class \code{c("temporal_cubble_df", "cubble_df")}, which expands the temporal variables into the long form and stores the spatial variables as a data attribute.
+A long cubble expands the temporal variables into the long form and stores the spatial variables as a data attribute.
+In the header of a long cubble object, the extent now prints the temporal range, the interval, and whether there is gaps in the data.
+It also prints the available spatial variables, as oppose to the temporal variables in the nested cubble.
 
@@
 cb_long
-class(cb_long)
 ```
 
-In the header of a long cubble object, the extent now prints the temporal range, the interval, and whether there is gaps in the data.
-It also prints the available spatial variables, as oppose to the temporal variables in the nested cubble.
-The spatial component can be printed with \code{spatial()}:
+### The cubble attributes
 
-```{r}
+```{r class, fig.align="center", out.width = "100%", fig.cap = "Illustration of a cubble object in the long form and nested form, along with the associated attributes."}
+#knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.004.png"))
+```
+
+A cubble object inherits the attributes from tibble (and its subclasses): `class`, `row.names`, and `names`, in addition to three specialised attributes:
+
+-   `key`: the spatial identifier
+-   `index`: the temporal identifier
+-   `coords`: a pair of ordered coordinates associated with the location
+
+Readers known the `key` and `index` attributes from the `tsibble` package would already be familiar the two arguments.
+In cubble, the `key` attribute identifies the row in the nested cubble, and together with the `index` argument, identifies the row in the long cubble.
+Currently, cubble only supports one variable as the key and the accepted temporal class for index includes the base R class `Date`, `POSIXlt`, `POSIXct` and tsibble's `yearmonth`, `yearweek`, and `yearquarter` class.
+
+The `coords` attribute takes an ordered pair of coordinate.
+It can be a unprojected pair of longitude and latitude, or a projected easting and northing values.
+Under the hood, the `sf` package is used to calculate the bounding box, shown in the header of a nested cubble, and other spatial operations.
+
+The long cubble has a special attribute `spatial` to store the spatial variables: all the variables in the nested cubble, except for the `ts` column.
+The shortcut function are available to extract components in the attributes, for example, \code{spatial()} for extracting the tibble object for the spatial variables:
+
+```{r echo = TRUE}
 spatial(cb_long)
@@
 
-### Coercing from foregin objects
+### Coercing from foreign objects
 
@@
 
-Table: (\#tab:funs) Functions implemented in the \pkg{cubble} package
+| Category  | Functions |
+|--------|--------------------------------------------------------|
+| base R    | `[`, `[[<-`, `names<-` |
+| tidyverse | `dplyr_row_slice`, `dplyr_col_modify`, `dplyr_reconstruct`, `select`, `mutate`, `arrange`, `filter`, `group_by`, `ungroup`, `summarise`, `select`, `slice`, `rowwise`, `rename`, `bind_rows`, `bind_cols`, `relocate`, `type_sum`, the slice family (`slice_head`, `slice_tail`, `slice_max`, `slice_min`, `slice_sample`) and the join family (`left_join`, `right_join`, `inner_join`, `full_join`, `anti_join`, `semi_join`) |
+| cubble    | `as_cubble`, `cubble`, `make_cubble`, `check_key` `face_temporal`, `face_spatial`, `unfold`, `key`, `key_vars`, `key_data`, `index`, `index_var`, `coords`, `spatial`, `match_sites`, `match_spatial`, `match_temporal`, `geom_glyph`, `geom_glyph_box`, `geom_glyph_line`, `make_spatial_sf`, `make_temporal_tsibble`, `fill_gaps`, `scan_gaps`                                                                                |
 
-|Category |   Functions|
-|---|----------------|
-|base R | `[`, `[<-`, `names<-`, `[[`, `[[<-`| 
-|tidyverse |`dplyr_row_slice`, `dplyr_col_modify`, `dplyr_reconstruct`, `select`, `mutate`, `arrange`, `filter`, `summarise`, `select`, `slice`, `rowwise`, `rename`, `bind_cols`, `group_by`, `ungroup`, `slice_*`, `*_join`, `relocate`, `type_sum` |
-|cubble | `cubble`, `make_cubble`, `new_spatial_cubble`, `new_temporal_cubble`, `make_spatial_sf`, `make_temporal_tsibble`, `face_temporal`, `face_spatial`, `unfold`, `key`, `key_vars`, `key_data`, `index`, `index_var`, `coords`, `spatial`, `fill_gaps`, `scan_gaps`, `switch_key`, `match_sites`, `match_spatial`, `match_temporal`, `geom_glyph`, `geom_glyph_box`, `geom_glyph_line`|
+: (#tab:funs) Functions implemented in the \pkg{cubble} package
 
-Table: (\#tab:methods) Methods implemented for the three \code{cubble} classes: \code{cubble\_df}, \code{spatial\_cubble\_df}, and \code{temporal\_cubble\_df}
+| Class                | Methods|
+|------------------|------------------------------------------------------|
+| `cubble_df`          | `[[<-, dplyr_col_modify, key_data, key_vars, key, print`                                                                                                                                                                                     |
+| `spatial_cubble_df`  | `[, names<-,  tbl_sum, dplyr_reconstruct, dplyr_row_slice, face_spatial, face_temporal, unfold, arrange, rename, rowwise, group_by, ungroup, select, spatial, summarise, unfold, update_cubble`                                              |
+| `temporal_cubble_df` | `[, names<-, tbl_sum, arrange, dplyr_reconstruct, dplyr_row_slice, face_spatial, face_temporal, unfold, fill_gaps, group_by, ungroup,  rename, rowwise, scan_gaps, select, spatial, summarise, tbl_sum, bind_rows, bind_cols, update_cubble` |
 
-|Class | Methods | 
-| ----- | ----------------|
-|`cubble_df` | `[, dplyr_col_modify, key_data, key_vars, key, names<-, print, tbl_sum`|
-|`spatial_cubble_df` | `dplyr_reconstruct, dplyr_row_slice, face_spatial, face_temporal, arrange, rename, rowwise, group_by, ungroup, select, spatial, summarise, tbl_sum, unfold, update_cubble`|
-|`temporal_cubble_df` | `arrange, dplyr_reconstruct, dplyr_row_slice, face_spatial, face_temporal, fill_gaps, group_by, ungroup,  rename, rowwise, scan_gaps, select, spatial, summarise, tbl_sum, unfold, update_cubble`|
+: (#tab:methods) Methods implemented for the three \code{cubble} classes: \code{cubble\_df}, \code{spatial\_cubble\_df}, and \code{temporal\_cubble\_df}
 
 <!-- methods(class = "cubble_df") |> paste0(collapse = ", ") |> stringr::str_remove_all(".cubble_df") -->
+
 <!-- methods(class = "spatial_cubble_df") |> paste0(collapse = ", ") |> stringr::str_remove_all(".spatial_cubble_df") -->
+
 <!-- methods(class = "temporal_cubble_df") |> paste0(collapse = ", ") |> stringr::str_remove_all(".temporal_cubble_df") -->
 
-### dplyr and tidyr
+### dplyr
 
@@
 Conversely, \\code{face\\_spatial()} transforms a temporal cubble into a spatial one to emphasize spatial variables."}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.002.png"))
+knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.001.png"))
 ```
@@
 ```{r echo = TRUE}
-#ts_long %>% has_gaps()
+ts_long %>% has_gaps()
 ```

### 2023-06-14 3-advanced.Rmd
other update as goes

@@
 
-## Hierarchical structure
-
-Spatial locations can have grouping structures either inherent to the data e.g., state within country or obtained during the analysis e.g., cluster id.
-In this case, it can be useful to summarize variables at various levels of the hierarchy.
-The function \code{switch_key()} can be used to change the grouping level of spatial locations.
-The diagram in Figure \ref{fig:illu-hier} shows how this function can be used to switch the grouping from station ids to cluster ids.
-The result can also be stretched into long form.
-By specifying \code{cluster_nested <- station_nested \%>\% switch_key(key = cluster)}, the \code{cubble} object redefines its \code{key} from the \code{id} column in \code{station_nested} to the \code{cluster} column in \code{cluster_nested}.
-All the spatial variables belonging to the \code{cluster} column are now nested into a \code{.val} column which allows for summarizing based on cluster.
-
-```{r illu-hier, echo = FALSE, fig.align="center",out.height="40%", out.width = "100%", fig.cap = "Hierarchical spatial structure can be handled using \\code{switch\\_key()}, to create summaries based on any level.
-Here the switch is between the station id and a cluster id.
-Once the change is made the data can be stretched into the long form."}
-#knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.003.png"))
-```
-
 ## Data fusion and matching {#matching}
@@
 The function \code{match_sites()} in the \pkg{cubble} package provides a simple algorithm for this task.
-The algorithm first matches spatially by computing the pairwise distance on latitude and longitude.
+
+Spatial matching based on distance.
+
+````
+match_spatial(df1, df2, ...)
+````
+
+Allow one-to-many output, control the number of group and number of match within each group with arguments:
+
+Temporal matching, apply on spatially matched group
+
+````
+match_temporal(<SPATIAL_RES>, group_id, match_id, temporal_by = c("xxx" = "xxx"))
+````
+
+Different matching functions can be tuned with `temporal_match_fn`, which takes a list of matched time series and additional arguments.
+By Default, the matching function calculates the number of peaks one series falls into another within a temporal window: `match_peak`
+
+<!-- The algorithm first matches spatially by computing the pairwise distance on latitude and longitude.
 Then it matches temporally by computing the number of matched peaks within a fixed length moving window.
@@
 Here two out of three peaks match.
-Options for \code{match_sites()} are:
-
-```{r illu-matching, out.width="100%", fig.cap="An illustration of temporal matching in the \\pkg{cubble} package.
-Three highest peaks are identified in each series and intervals are constructed on series \\code{A}.
-Two peaks in series \\code{a} fall into the intervals and hence the two series are considered to have two matches."}
-knitr::include_graphics(here::here("figures/illu-matching.png"))
-```
-
- - \code{spatial_n_keep}: the number of spatial match for each site to keep;
- - \code{spatial_dist_max}: the maximum distance allowed for a matched pair;
- - \code{temporal_n_highest}: the number of peaks used - 3 in the example above;
- - \code{temporal_window}: the length of the interval - 5 in the example above; and
- - \code{temporal_min_match}: the minimum number of matched peaks for a valid matched pair.
+Options for \code{match_sites()} are: -->
 
 ## Interactive graphics
-
-<!-- Interactive graphics can listen to users' actions on the plot to provide additional information that facilitates data exploration.
-This is a useful technique for spatio-temporal data since users can zoom or pan the map to view the local and global structure of the map; use tooltips or popups to query more information about a graphic element; or highlight points to explore its linked views in other plots.
-In the R community, many implementations have been developed to connect \proglang{R} to \proglang{javascript} to create interactive graphics.
-In relation to spatio, temporal, and spatio-temporal data, the general purpose packages \pkg{plotly} [@plotly] and \pkg{leaflet} [@leaflet] realise various interactive actions through their corresponding javascript libraries.
-\pkg{crosstalk} [@crosstalk] and \pkg{tsibbletalk} [@tsibbletalk] implement brushed linking between htmlwidgets.
-\pkg{ggiraph} [@ggiraph] enables tooltip, self-linking, and customised actions specified through its own \proglang{javascript}.
--->
-
-<!-- While many graphic implementations present worked examples to illustrate the usage of the package, few documents the underlying pipeline that transforms the raw data step-by-step to the final view on the screen.
-There have been some early work in building the data pipeline for (interactive) graphics [@buja1988elements; @buja1996interactive;  @sutherland2000orca] and more recent discussions include @wickham2009plumbing, @xie2014reactive, and @cheng2016enabling.
--->
 
@@
 (An illustration can be found in the appendix.) Note that this type of linking, both from the map or the time series, is what @CS07 would call categorical variable linking, where station id is the categorical variable.
-<!--\ref{fig:illu-interactive-2} in the Appendix illustrates this process.-->
 
@@
 This will link to all the rows with the same id in the long \\code{cubble} (b) and update the line plot (c)."}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.004.png"))
+knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.002.png"))
 ```
@@
 Here we elaborate on the transformations made to produce a glyph map.
-
 The package \pkg{GGally} [@ggally] has implemented glyph maps through the \code{glyphs()} function.
 The function constructs a \code{data.frame} with calculated position (\code{gx}, \code{gy}, \code{gid}) of each point on the time series using linear algebra (Equations 1 and 2 in @Wickham2012-yr).
-The data can then be piped into ggplot to create the glyph map as:
-
-```{r eval = FALSE, echo = TRUE}
-library("ggplot2")
-gly <- glyphs(data,
-              x_major = ..., x_minor = ...,
-              y_major = ..., y_minor = ..., ...)
-
-ggplot(gly, aes(gx, gy, group = gid)) +
-  geom_path()
-```
-
 A new implementation of the glyph map as a ggproto, \code{GeomGlyph}, has been made in the \pkg{cubble} package so that a glyph map can be created with \code{geom_glyph()}:
@@
 ```{r eval = FALSE, echo = TRUE}
-ggplot(data = data) +
+data |>
+  ggplot() +
   geom_glyph(aes(x_major = ..., x_minor = ...,
@@
 
-An example using a glyph map is shown in Section \@ref(historicaltmax).
+Some useful controls over the glyph map are also available in the \code{geom_glyph()} implementation
 
-Some useful controls over the glyph map are also available in the \code{geom_glyph()} implementation.
-Polar coordinate glyph maps are specified using \code{polar = TRUE}, and arguments \code{width} and \code{height} can be specified in either absolute or relative value.
-Global and local scale is specified with \code{global_rescale}, which defaults to \code{TRUE}.
-Reference boxes and lines can be added with separate \code{geom_glyph_box()} and \code{geom_glyph_line()} lines.
+  * polar coordinate glyph maps are specified using \code{polar = TRUE}, 
+  * arguments \code{width} and \code{height} can be specified in either absolute or relative value, 
+  * global and local scale is specified with \code{global_rescale}, which defaults to \code{TRUE}, and 
+  *reference boxes and lines can be added with geoms: \code{geom_glyph_box()} and \code{geom_glyph_line()}.

### 2023-06-14 4-3-matching.Rmd
other update as goes

@@
   mutate(type = "climate")
-
 river <- cubble::river %>% mutate(type = "river") 
-
 vic_map <- ozmaps::abs_ste %>%  filter(NAME == "Victoria")
-
-# ggplot() + 
-#   geom_sf(data = vic_map) + 
-#   geom_point(data = dplyr::bind_rows(river, climate_vic), 
-#              aes(x = long, y = lat, color = type)) + 
-#   theme_bw() +
-#   theme(legend.position = "bottom") +
-#   labs(x = "Longitude", y = "Latitude") + 
-#   scale_color_brewer(palette = "Dark2")
 ```
@@
 ```{r echo = FALSE}
-res_sp <- res_sp %>% 
-  filter(!group %in% c(5, 8)) %>% 
-  tidyr::pivot_longer(from:to, names_to = "order", values_to = "id") %>% 
-  select(-order)
-res_sp <- bind_rows(climate_vic, river %>% rename(id = station)) %>% 
-  inner_join(res_sp) 
-```
-
-```{r}
-res_sp
+res_sp <- match_spatial(climate_vic, river, spatial_n_group = 10, return_cubble = TRUE)
+(res_sp2 <- res_sp[-c(5, 8)] %>% bind_rows())
 ```
@@
 ```{r echo = TRUE}
-(res_tm <- res_sp %>% 
+(res_tm <- res_sp2 %>% 
   match_temporal(
-    match_id = group,
+    match_id = group, data_id = type,
     temporal_by = c("prcp" = "Water_course_level"),
@@
 ```{r}
-(res_tm <- res_sp %>% 
+(res_tm <- res_sp2 %>% 
   match_temporal(
-    match_id = group,
+    match_id = group, data_id = type,
     temporal_by = c("prcp" = "Water_course_level"),
@@
     temporal_n_highest = 30,
-    temporal_min_match = 15))
+    temporal_min_match = 15) %>% 
+   bind_rows())
 ```
@@
   unfold(group, type) %>%  
-  rename(prcp = matched) %>%  
   group_by(group, type) %>%
-  mutate(prcp = (prcp - min(prcp, na.rm = TRUE))/ (max(prcp, na.rm = TRUE) - min(prcp, na.rm = TRUE)))
+  mutate(matched = (matched - min(matched, na.rm = TRUE))/ (max(matched, na.rm = TRUE) - min(matched, na.rm = TRUE)))
 
@@
   geom_point(data = dplyr::bind_rows(river, climate_vic), 
-             aes(x = long, y = lat, color = type), alpha = 0.2, fill = 0.2) +
-  geom_point(data = res_tm, aes(x = long, y = lat, color = type)) +
+             aes(x = long, y = lat, color = type), 
+             alpha = 0.2, fill = 0.2) +
+  geom_point(data = res_tm %>% as_tibble(), 
+             aes(x = long, y = lat, color = type)) +
   ggrepel::geom_label_repel(
-    data = res_tm |> filter(type == "river"), aes(x = long, y = lat, label = group)) +
+    data = res_tm |> filter(type == "river") %>% as_tibble(), 
+    aes(x = long, y = lat, label = group)) +
   scale_color_brewer(palette = "Dark2")  + 
@@
 p2 <- res_tm_long %>%  
-  ggplot(aes(x = date, y = prcp, group = type,color = type)) + 
+  ggplot(aes(x = date, y = matched, group = type,color = type)) + 
   geom_line() + 

### 2023-06-14 jss/main.tex
other update as goes

@@
 
-{[}conceptual framework{]} The \code{cubble} class tackles the spatio-temporal data collected at unique fixed locations and allowed for irregularity in the temporal dimension, i.e.~irregular time steps and missingness.
-
-The \code{cubble} class (\code{cubble_df}) is an S3 class \citep{advr} built on the \code{tibble} class, specifically to organize spatio-temporal data.
-It uses three attributes to character the spatio-temporal data:
+{[}conceptual framework{]} Spatio-temporal data comes various spatial and temporal characteristics and requires different data structures to wrangle: climate weather stations are recorded at fixed point location but with potential temporal data quality issue (missingness on the day); GPS data tracks unique point locations at different timestamps; satellite imageries captures snapshots of landscape at selected time.
+The type of spatio-temporal data cubble tackles are those collected at unique fixed locations while allowed for irregularity in the temporal dimension, like the weather station data.
+In the four layouts presented by the spacetime paper \citep{spacetime}, cubble handles full space-time and sparse space-time layouts.
+
+The cubble class is an S3 class, built on tibble, to pivot spatio-temporal data into a nested (spatial) form and a long (temporal) form. It has two subclasses:
 
@@
 \item
-  \code{key} identifies the spatial locations
+  a nested/ spatial cubble has class \texttt{c("spatial\_cubble\_df",\ "cubble\_df")}
 \item
-  \code{date} identifies the temporal order, and
-\item
-  \code{coords} identifies the longitude and latitude coordinate associated with the location.
+  a long/ temporal cubble has class \texttt{c("temporal\_cubble\_df",\ "cubble\_df")}
 \end{itemize}
 
-The base class \code{cubble_df} has two specialised class \code{spatial_cubble_df} and \code{temporal_cubble_df} to arrange the spatio-temporal data in two forms, as illustrated in Figure \ref{fig:class}.
-
-\begin{CodeChunk}
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{../figures/diagram-keynotes/diagram-keynotes.007}
-
-}
-
-\caption[Illustration of a cubble object in the long form and nested form, along with the associated attributes]{Illustration of a cubble object in the long form and nested form, along with the associated attributes.}\label{fig:class}
-\end{figure}
-\end{CodeChunk}
-
-\hypertarget{nested-cubble}{%
-\subsubsection{nested cubble}\label{nested-cubble}}
-
-A nested cubble (in the spatial form) has class \code{c("spatial_cubble_df", "cubble_df")}.
-It arranges spatial variables in columns and nests temporal variables in a specialised \code{ts} column.
-Below prints an nested cubble object using the weather station data collected in three airports.
-The data is a subset of a larger data \code{climate_aus} collected from National Oceanic and Atmospheric Administration (NOAA) Global Historical Climatology Network Daily (GHCND).
-The data include spatial variables such as station ID, longitude, latitude, elevation, station name, world meteorology organisation ID, in additional to daily temporal information, maximum and minimum temperature values and precipitation for the year 2020.
-The airport data is a subset of three airport stations in Melbourne for the initial 10 days of the year (the creation of a cubble object will be explained in Section \ref{create}):
+A nested cubble arranges spatial variables in columns and nests temporal variables in a specialised \texttt{ts} column.
+Below prints an nested cubble object using the weather station data collected in three airports (the creation of a cubble object will be explained in Section \ref{create}).
+This toy data is a subset of a larger data \code{climate_aus} collected from Global Historical Climatology Network Daily (GHCND).
+The three airport stations in Melbourne are recorded with station metadata: station ID, longitude, latitude, elevation, station name, world meteorology organisation ID.
+The temporal variables are precipitation, maximum and minimum temperature, which can be read from the cubble header:
 
@@
 \end{CodeOutput}
-\begin{CodeInput}
-R> class(cb_nested)
-\end{CodeInput}
-\begin{CodeOutput}
-[1] "spatial_cubble_df" "cubble_df"         "tbl_df"           
-[4] "tbl"               "data.frame"       
-\end{CodeOutput}
-\end{CodeChunk}
-
-A nested cubble header prints the key and index variable name and its form.
-The spatial extent is given by the bounding box with CRS information (missing here).
-The nested cubble header also shows the name of temporal variables nested in the \code{ts} column and their variable types: \code{prcp [dbl], tmax [dbl], tmin [dbl]}.
-
-\hypertarget{long-cubble}{%
-\subsubsection{long cubble}\label{long-cubble}}
-
-A long cubble (in the temporal form) has class \code{c("temporal_cubble_df", "cubble_df")}, which expands the temporal variables into the long form and stores the spatial variables as a data attribute.
+\end{CodeChunk}
+
+A long cubble expands the temporal variables into the long form and stores the spatial variables as a data attribute.
+In the header of a long cubble object, the extent now prints the temporal range, the interval, and whether there is gaps in the data.
+It also prints the available spatial variables, as oppose to the temporal variables in the nested cubble.
 
@@
 \end{CodeOutput}
-\begin{CodeInput}
-R> class(cb_long)
-\end{CodeInput}
-\begin{CodeOutput}
-[1] "temporal_cubble_df" "cubble_df"          "tbl_df"            
-[4] "tbl"                "data.frame"        
-\end{CodeOutput}
-\end{CodeChunk}
-
-In the header of a long cubble object, the extent now prints the temporal range, the interval, and whether there is gaps in the data.
-It also prints the available spatial variables, as oppose to the temporal variables in the nested cubble.
-The spatial component can be printed with \code{spatial()}:
-
-\begin{CodeChunk}
+\end{CodeChunk}
+
+\hypertarget{the-cubble-attributes}{%
+\subsubsection{The cubble attributes}\label{the-cubble-attributes}}
+
+A cubble object inherits the attributes from tibble (and its subclasses): \texttt{class}, \texttt{row.names}, and \texttt{names}, in addition to three specialised attributes:
+
+\begin{itemize}
+\tightlist
+\item
+  \texttt{key}: the spatial identifier
+\item
+  \texttt{index}: the temporal identifier
+\item
+  \texttt{coords}: a pair of ordered coordinates associated with the location
+\end{itemize}
+
+Readers known the \texttt{key} and \texttt{index} attributes from the \texttt{tsibble} package would already be familiar the two arguments.
+In cubble, the \texttt{key} attribute identifies the row in the nested cubble, and together with the \texttt{index} argument, identifies the row in the long cubble.
+Currently, cubble only supports one variable as the key and the accepted temporal class for index includes the base R class \texttt{Date}, \texttt{POSIXlt}, \texttt{POSIXct} and tsibble's \texttt{yearmonth}, \texttt{yearweek}, and \texttt{yearquarter} class.
+
+The \texttt{coords} attribute takes an ordered pair of coordinate.
+It can be a unprojected pair of longitude and latitude, or a projected easting and northing values.
+Under the hood, the \texttt{sf} package is used to calculate the bounding box, shown in the header of a nested cubble, and other spatial operations.
+
+The long cubble has a special attribute \texttt{spatial} to store the spatial variables: all the variables in the nested cubble, except for the \texttt{ts} column.
+The shortcut function are available to extract components in the attributes, for example, \code{spatial()} for extracting the tibble object for the spatial variables:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> spatial(cb_long)
+\end{CodeInput}
 \begin{CodeOutput}
@@
 
-\hypertarget{coercing-from-foregin-objects}{%
-\subsubsection{Coercing from foregin objects}\label{coercing-from-foregin-objects}}
+\hypertarget{coercing-from-foreign-objects}{%
+\subsubsection{Coercing from foreign objects}\label{coercing-from-foreign-objects}}
 
@@
 \begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.1579}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.8421}}@{}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.1250}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.8750}}@{}}
 \caption{\label{tab:funs} Functions implemented in the \pkg{cubble} package}\tabularnewline
@@
 \endhead
-base R & \texttt{{[}}, \texttt{{[}\textless{}-}, \texttt{names\textless{}-}, \texttt{{[}{[}}, \texttt{{[}{[}\textless{}-} \\
-tidyverse & \texttt{dplyr\_row\_slice}, \texttt{dplyr\_col\_modify}, \texttt{dplyr\_reconstruct}, \texttt{select}, \texttt{mutate}, \texttt{arrange}, \texttt{filter}, \texttt{summarise}, \texttt{select}, \texttt{slice}, \texttt{rowwise}, \texttt{rename}, \texttt{bind\_cols}, \texttt{group\_by}, \texttt{ungroup}, \texttt{slice\_*}, \texttt{*\_join}, \texttt{relocate}, \texttt{type\_sum} \\
-cubble & \texttt{cubble}, \texttt{make\_cubble}, \texttt{new\_spatial\_cubble}, \texttt{new\_temporal\_cubble}, \texttt{make\_spatial\_sf}, \texttt{make\_temporal\_tsibble}, \texttt{face\_temporal}, \texttt{face\_spatial}, \texttt{unfold}, \texttt{key}, \texttt{key\_vars}, \texttt{key\_data}, \texttt{index}, \texttt{index\_var}, \texttt{coords}, \texttt{spatial}, \texttt{fill\_gaps}, \texttt{scan\_gaps}, \texttt{switch\_key}, \texttt{match\_sites}, \texttt{match\_spatial}, \texttt{match\_temporal}, \texttt{geom\_glyph}, \texttt{geom\_glyph\_box}, \texttt{geom\_glyph\_line} \\
+base R & \texttt{{[}}, \texttt{{[}{[}\textless{}-}, \texttt{names\textless{}-} \\
+tidyverse & \texttt{dplyr\_row\_slice}, \texttt{dplyr\_col\_modify}, \texttt{dplyr\_reconstruct}, \texttt{select}, \texttt{mutate}, \texttt{arrange}, \texttt{filter}, \texttt{group\_by}, \texttt{ungroup}, \texttt{summarise}, \texttt{select}, \texttt{slice}, \texttt{rowwise}, \texttt{rename}, \texttt{bind\_rows}, \texttt{bind\_cols}, \texttt{relocate}, \texttt{type\_sum}, the slice family (\texttt{slice\_head}, \texttt{slice\_tail}, \texttt{slice\_max}, \texttt{slice\_min}, \texttt{slice\_sample}) and the join family (\texttt{left\_join}, \texttt{right\_join}, \texttt{inner\_join}, \texttt{full\_join}, \texttt{anti\_join}, \texttt{semi\_join}) \\
+cubble & \texttt{as\_cubble}, \texttt{cubble}, \texttt{make\_cubble}, \texttt{check\_key} \texttt{face\_temporal}, \texttt{face\_spatial}, \texttt{unfold}, \texttt{key}, \texttt{key\_vars}, \texttt{key\_data}, \texttt{index}, \texttt{index\_var}, \texttt{coords}, \texttt{spatial}, \texttt{match\_sites}, \texttt{match\_spatial}, \texttt{match\_temporal}, \texttt{geom\_glyph}, \texttt{geom\_glyph\_box}, \texttt{geom\_glyph\_line}, \texttt{make\_spatial\_sf}, \texttt{make\_temporal\_tsibble}, \texttt{fill\_gaps}, \texttt{scan\_gaps} \\
 \bottomrule()
@@
 \begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.2381}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.7619}}@{}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.2500}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.7500}}@{}}
 \caption{\label{tab:methods} Methods implemented for the three \code{cubble} classes: \code{cubble\_df}, \code{spatial\_cubble\_df}, and \code{temporal\_cubble\_df}}\tabularnewline
@@
 \endhead
-\texttt{cubble\_df} & \texttt{{[},\ dplyr\_col\_modify,\ key\_data,\ key\_vars,\ key,\ names\textless{}-,\ print,\ tbl\_sum} \\
-\texttt{spatial\_cubble\_df} & \texttt{dplyr\_reconstruct,\ dplyr\_row\_slice,\ face\_spatial,\ face\_temporal,\ arrange,\ rename,\ rowwise,\ group\_by,\ ungroup,\ select,\ spatial,\ summarise,\ tbl\_sum,\ unfold,\ update\_cubble} \\
-\texttt{temporal\_cubble\_df} & \texttt{arrange,\ dplyr\_reconstruct,\ dplyr\_row\_slice,\ face\_spatial,\ face\_temporal,\ fill\_gaps,\ group\_by,\ ungroup,\ \ rename,\ rowwise,\ scan\_gaps,\ select,\ spatial,\ summarise,\ tbl\_sum,\ unfold,\ update\_cubble} \\
+\texttt{cubble\_df} & \texttt{{[}{[}\textless{}-,\ dplyr\_col\_modify,\ key\_data,\ key\_vars,\ key,\ print} \\
+\texttt{spatial\_cubble\_df} & \texttt{{[},\ names\textless{}-,\ \ tbl\_sum,\ dplyr\_reconstruct,\ dplyr\_row\_slice,\ face\_spatial,\ face\_temporal,\ unfold,\ arrange,\ rename,\ rowwise,\ group\_by,\ ungroup,\ select,\ spatial,\ summarise,\ unfold,\ update\_cubble} \\
+\texttt{temporal\_cubble\_df} & \texttt{{[},\ names\textless{}-,\ tbl\_sum,\ arrange,\ dplyr\_reconstruct,\ dplyr\_row\_slice,\ face\_spatial,\ face\_temporal,\ unfold,\ fill\_gaps,\ group\_by,\ ungroup,\ \ rename,\ rowwise,\ scan\_gaps,\ select,\ spatial,\ summarise,\ tbl\_sum,\ bind\_rows,\ bind\_cols,\ update\_cubble} \\
 \bottomrule()
@@
 
-\hypertarget{dplyr-and-tidyr}{%
-\subsubsection{dplyr and tidyr}\label{dplyr-and-tidyr}}
+\hypertarget{dplyr}{%
+\subsubsection{dplyr}\label{dplyr}}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{../figures/diagram-keynotes/diagram-keynotes.002}
+{\centering \includegraphics[width=1\linewidth]{../figures/diagram-keynotes/diagram-keynotes.001}
 
@@
 
-\hypertarget{hierarchical-structure}{%
-\subsection{Hierarchical structure}\label{hierarchical-structure}}
-
-Spatial locations can have grouping structures either inherent to the data e.g., state within country or obtained during the analysis e.g., cluster id.
-In this case, it can be useful to summarize variables at various levels of the hierarchy.
-The function \code{switch_key()} can be used to change the grouping level of spatial locations.
-The diagram in Figure \ref{fig:illu-hier} shows how this function can be used to switch the grouping from station ids to cluster ids.
-The result can also be stretched into long form.
-By specifying \code{cluster_nested <- station_nested \%>\% switch_key(key = cluster)}, the \code{cubble} object redefines its \code{key} from the \code{id} column in \code{station_nested} to the \code{cluster} column in \code{cluster_nested}.
-All the spatial variables belonging to the \code{cluster} column are now nested into a \code{.val} column which allows for summarizing based on cluster.
-
 \hypertarget{matching}{%
@@
 The function \code{match_sites()} in the \pkg{cubble} package provides a simple algorithm for this task.
-The algorithm first matches spatially by computing the pairwise distance on latitude and longitude.
-Then it matches temporally by computing the number of matched peaks within a fixed length moving window.
-Figure \ref{fig:illu-matching} illustrates this temporal matching.
-In the two series, \emph{A} and \emph{a}, three peaks have been identified in each.
-An interval, of fixed length, is constructed for each peak in series \emph{A}, while the peaks in series \emph{a} are tested against whether they fall into any of the intervals.
-Here two out of three peaks match.
-Options for \code{match_sites()} are:
-
-\begin{CodeChunk}
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{../figures/illu-matching}
-
-}
-
-\caption{An illustration of temporal matching in the \pkg{cubble} package.
-Three highest peaks are identified in each series and intervals are constructed on series \code{A}.
-Two peaks in series \code{a} fall into the intervals and hence the two series are considered to have two matches.}\label{fig:illu-matching}
-\end{figure}
-\end{CodeChunk}
-
-\begin{itemize}
-\tightlist
-\item
-  \code{spatial_n_keep}: the number of spatial match for each site to keep;
-\item
-  \code{spatial_dist_max}: the maximum distance allowed for a matched pair;
-\item
-  \code{temporal_n_highest}: the number of peaks used - 3 in the example above;
-\item
-  \code{temporal_window}: the length of the interval - 5 in the example above; and
-\item
-  \code{temporal_min_match}: the minimum number of matched peaks for a valid matched pair.
-\end{itemize}
+
+Spatial matching based on distance.
+
+\begin{verbatim}
+match_spatial(df1, df2, ...)
+\end{verbatim}
+
+Allow one-to-many output, control the number of group and number of match within each group with arguments:
+
+Temporal matching, apply on spatially matched group
+
+\begin{verbatim}
+match_temporal(<SPATIAL_RES>, group_id, match_id, temporal_by = c("xxx" = "xxx"))
+\end{verbatim}
+
+Different matching functions can be tuned with \texttt{temporal\_match\_fn}, which takes a list of matched time series and additional arguments.
+By Default, the matching function calculates the number of peaks one series falls into another within a temporal window: \texttt{match\_peak}
 
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.35\textheight]{../figures/diagram-keynotes/diagram-keynotes.004}
+{\centering \includegraphics[width=1\linewidth,height=0.35\textheight]{../figures/diagram-keynotes/diagram-keynotes.002}
 
@@
 Here we elaborate on the transformations made to produce a glyph map.
-
 The package \pkg{GGally} \citep{ggally} has implemented glyph maps through the \code{glyphs()} function.
 The function constructs a \code{data.frame} with calculated position (\code{gx}, \code{gy}, \code{gid}) of each point on the time series using linear algebra (Equations 1 and 2 in \citet{Wickham2012-yr}).
-The data can then be piped into ggplot to create the glyph map as:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> library("ggplot2")
-R> gly <- glyphs(data,
-+               x_major = ..., x_minor = ...,
-+               y_major = ..., y_minor = ..., ...)
-R> 
-R> ggplot(gly, aes(gx, gy, group = gid)) +
-+   geom_path()
-\end{CodeInput}
-\end{CodeChunk}
-
 A new implementation of the glyph map as a ggproto, \code{GeomGlyph}, has been made in the \pkg{cubble} package so that a glyph map can be created with \code{geom_glyph()}:
@@
 \begin{CodeInput}
-R> ggplot(data = data) +
+R> data |>
++   ggplot() +
 +   geom_glyph(aes(x_major = ..., x_minor = ...,
@@
 
-An example using a glyph map is shown in Section \ref{historicaltmax}.
-
-Some useful controls over the glyph map are also available in the \code{geom_glyph()} implementation.
-Polar coordinate glyph maps are specified using \code{polar = TRUE}, and arguments \code{width} and \code{height} can be specified in either absolute or relative value.
-Global and local scale is specified with \code{global_rescale}, which defaults to \code{TRUE}.
-Reference boxes and lines can be added with separate \code{geom_glyph_box()} and \code{geom_glyph_line()} lines.
+Some useful controls over the glyph map are also available in the \code{geom_glyph()} implementation
+
+\begin{itemize}
+\tightlist
+\item
+  polar coordinate glyph maps are specified using \code{polar = TRUE},
+\item
+  arguments \code{width} and \code{height} can be specified in either absolute or relative value,
+\item
+  global and local scale is specified with \code{global_rescale}, which defaults to \code{TRUE}, and
+  *reference boxes and lines can be added with geoms: \code{geom_glyph_box()} and \code{geom_glyph_line()}.
+\end{itemize}
 
@@
 \begin{CodeOutput}
-# cubble:   key: id [16], index: date, nested form
-# spatial:  [144.5203, -38.144913, 148.4667, -36.128657], Missing CRS!
+# cubble:   key: id [16], index: date, nested form, [sf]
+# spatial:  [144.5203, -38.144913, 148.4667, -36.128657], WGS 84
 # temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
-  id           long   lat  elev name           wmo_id ts       type   dist group
-  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>   <chr>   [m] <int>
-1 ASN00080015  145. -36.2  96   echuca  aerod~  94861 <tibble> clim~ 4034.     4
-2 ASN00080091  145. -36.3 105   kyabram dpi     95833 <tibble> clim~ 7460.    10
-3 ASN00082042  146. -36.8 502   strathbogie     95843 <tibble> clim~ 6153.     6
-4 ASN00084145  148. -37.7  62.7 orbost          95918 <tibble> clim~ 2185.     2
-5 ASN00085072  147. -38.1   4.6 east sale air~  94907 <tibble> clim~ 3282.     3
+  id      long   lat  elev name  wmo_id ts       type              geometry
+  <chr>  <dbl> <dbl> <dbl> <chr>  <dbl> <list>   <chr>          <POINT [°]>
+1 ASN00~  145. -37.0 290   rede~  94859 <tibble> clim~  (144.5203 -37.0194)
+2 406213  145. -37.0  NA   CAMP~     NA <tibble> river (144.5403 -37.01512)
+3 ASN00~  148. -37.7  62.7 orbo~  95918 <tibble> clim~  (148.4667 -37.6922)
+4 222201  148. -37.7  NA   SNOW~     NA <tibble> river  (148.451 -37.70739)
+5 ASN00~  147. -38.1   4.6 east~  94907 <tibble> clim~  (147.1322 -38.1156)
 # i 11 more rows
-\end{CodeOutput}
-\end{CodeChunk}
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> (res_tm <- res_sp %>% 
+# i 2 more variables: group <int>, dist [m]
+\end{CodeOutput}
+\end{CodeChunk}
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> (res_tm <- res_sp2 %>% 
 +   match_temporal(
-+     match_id = group,
++     match_id = group, data_id = type,
 +     temporal_by = c("prcp" = "Water_course_level"),
@@
 # A tibble: 8 x 2
-  group n_matches
+  group match_res
   <int>     <dbl>
 1     1        30
-2     7        26
-3     6        23
-4     9        21
-5     4        20
+2     2         5
+3     3        14
+4     4        20
+5     6        23
 # i 3 more rows
@@
 \begin{CodeOutput}
-# cubble:   key: id [16], index: date, nested form
-# spatial:  [144.5203, -38.144913, 148.4667, -36.128657], Missing CRS!
+# cubble:   key: id [16], index: date, nested form, [sf]
+# spatial:  [144.5203, -38.144913, 148.4667, -36.128657], WGS 84
 # temporal: date [date], matched [dbl]
-  id          long   lat  elev name  wmo_id type   dist group ts       n_matches
-  <chr>      <dbl> <dbl> <dbl> <chr>  <dbl> <chr>   [m] <int> <list>       <dbl>
-1 ASN000880~  145. -37.0 290   rede~  94859 clim~ 1838.     1 <tibble>        30
-2 406213      145. -37.0  NA   CAMP~     NA river 1838.     1 <tibble>        30
-3 ASN000841~  148. -37.7  62.7 orbo~  95918 clim~ 2185.     2 <tibble>         5
-4 222201      148. -37.7  NA   SNOW~     NA river 2185.     2 <tibble>         5
-5 ASN000850~  147. -38.1   4.6 east~  94907 clim~ 3282.     3 <tibble>        14
+  id         long   lat  elev name  wmo_id type              geometry group
+  <chr>     <dbl> <dbl> <dbl> <chr>  <dbl> <chr>          <POINT [°]> <int>
+1 ASN00088~  145. -37.0 290   rede~  94859 clim~  (144.5203 -37.0194)     1
+2 406213     145. -37.0  NA   CAMP~     NA river (144.5403 -37.01512)     1
+3 ASN00084~  148. -37.7  62.7 orbo~  95918 clim~  (148.4667 -37.6922)     2
+4 222201     148. -37.7  NA   SNOW~     NA river  (148.451 -37.70739)     2
+5 ASN00085~  147. -38.1   4.6 east~  94907 clim~  (147.1322 -38.1156)     3
 # i 11 more rows
+# i 3 more variables: dist [m], ts <list>, match_res <dbl>
 \end{CodeOutput}

### 2023-06-15 3-advanced.Rmd
text on matching

@@
 This can be considered to be a matching problem [@stuart2010matching; @mcintosh2018using] to pair similar time series from nearby locations, or even a data fusion exercise that merges data collected from different sources [@cocchi2019data].
-The function \code{match_sites()} in the \pkg{cubble} package provides a simple algorithm for this task.
 
-Spatial matching based on distance.
+In cubble, data can be matched spatially or temporarily with `match_spatial()` and `match_temporal()`.
+The function `match_spatial()` calculates the spatial distance between observations in two cubbles.
+Different distances are available with projected or unprojected coordinate reference system.
+Analysts can subset the number of matched group to output with argument `spatial_n_group` (by default 4 groups) and the number of match per group with argument `spatial_n_group` (by default 1, that is, one-to-one matching).
+The syntax to use `match_spatial()` is:
 
 ````
-match_spatial(df1, df2, ...)
+match_spatial(<cubble_obj1>, <cubble_obj2>, ...)
 ````
 
-Allow one-to-many output, control the number of group and number of match within each group with arguments:
+The function `match_temporal()` calculates the time series similarity between spatially matched groups.
+Two identifiers needs to be specified on the variable separates the each matched group (`match_id`) and the variable separates the two sources (`data_id`).
+The argument `temporal_by` uses the `by` syntax from dplyr `*_join` to specify the temporal matching variable.
 
-Temporal matching, apply on spatially matched group
+  The similarity score between two time series in the spatially matched group is calculated by a matching function, which analysts can customise.
+The matching function should take two time series in a list and output a single numerical score, which allows for interfacing with existing time series distance calculation implementation.
+By default, cubble implements a simple peak matching algorithm (`match_peak`) that counts the number of peaks in two time series that fall within a specified temporal window.
+The syntax to use `match_temporal()` is:
 
 ````
-match_temporal(<SPATIAL_RES>, group_id, match_id, temporal_by = c("xxx" = "xxx"))
+match_temporal(
+  <cubble_obj_from_match_spatial>, 
+  data_id = ... , match_id = ..., 
+  temporal_by = c("..." = "...")
+)
 ````
-
-Different matching functions can be tuned with `temporal_match_fn`, which takes a list of matched time series and additional arguments.
-By Default, the matching function calculates the number of peaks one series falls into another within a temporal window: `match_peak`
-
-<!-- The algorithm first matches spatially by computing the pairwise distance on latitude and longitude.
-Then it matches temporally by computing the number of matched peaks within a fixed length moving window.
-Figure \ref{fig:illu-matching} illustrates this temporal matching.
-In the two series, *A* and *a*, three peaks have been identified in each.
-An interval, of fixed length, is constructed for each peak in series *A*, while the peaks in series *a* are tested against whether they fall into any of the intervals.
-Here two out of three peaks match.
-Options for \code{match_sites()} are: -->
 

### 2023-06-15 4-3-matching.Rmd
text on matching

@@
 
-River level and rainfall data for the same areas should have some similarity.
-Here we examine the river gauge data (\code{Water_course_level}) from the Bureau of Meteorology [@bom] in relation to weather station rainfall from NOAA's climate data (\code{climate}).
-The goal is to match water gauges with nearby weather stations, spatially and temporally, using the \code{match_sites()} function.
-<!--The Bureau of Meteorology collects [water data](http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full) from river gauges.
-The collected variables include: electrical conductivity, turbidity, watercourse discharge, watercourse level, and water temperature.
-We expect that the water level should be related to the precipitation available from climate data, since rainfall will raise the water level in the river.
-Figure \ref{fig:matching-map} shows the location of available weather stations and water gauges in Victoria, and we will try to match water gauges with nearby weather stations, taking into account both spatial and temporal information from 2020.-->
+Bureau of Meteorology collects [water data](http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full) from river gauges and this includes variables: electrical conductivity, turbidity, water course discharge, water course level, and water temperature.
+In particular, water level will interactive with precipitation from the climate data since rainfall will raise the water level in the river.
+Both `climate_vic` and `river` are cubble objects and we can get a summary of the 10 closest pairs:
 
@@
 river <- cubble::river %>% mutate(type = "river") 
-vic_map <- ozmaps::abs_ste %>%  filter(NAME == "Victoria")
-```
-
-This function requires passing the major and minor data sets used for matching, in this case those are \code{river} and \code{climate}.
-The variables used for the temporal matching  are \code{Water_course_level} from the \code{river} data set and \code{prcp} in the climate data set.
-The rest of the arguments, as explained in Section \@ref(matching), correspond to the maximum and minimum number of peaks in the time series to be matched.
-In this example those are set to be a maximum of 30 and a minimum of 15 (approximately 2 and 1 per month).
-
-```{r echo = TRUE,  tibble.print_max = 10,   tibble.print_min = 10}
-(res_sp <- match_spatial(climate_vic, river, spatial_n_group = 10))
-```
-
-```{r echo = FALSE}
-res_sp <- match_spatial(climate_vic, river, spatial_n_group = 10, return_cubble = TRUE)
-(res_sp2 <- res_sp[-c(5, 8)] %>% bind_rows())
 ```
@@
 ```{r echo = TRUE}
-(res_tm <- res_sp2 %>% 
-  match_temporal(
-    match_id = group, data_id = type,
-    temporal_by = c("prcp" = "Water_course_level"),
-    temporal_n_highest = 30,
-    temporal_min_match = 15))
+res_sp <- match_spatial(climate_vic, river, spatial_n_group = 10)
+print(res_sp, n = 20)
 ```
 
-Or use \code{return_cubble = TRUE} to return a cubble object:
+The result can also be returned as cubble objects with argument `return_cubble = TRUE`.
+The output is a list where each element is a paired cubble object and you may consider combining all the results into a single cubble with `bind_rows()`.
+Care needs to be taken on in the case when a site is close to two stations since by construction, cubble require unique rows in the nested form.
+From the summary table above, river station `226027` is matched to more than one weather station: `ASN00085072` (group 3) and `ASN00085298` (group 5).
+(Similarly river station `230200` is matched in group 7 and 8).
+One can either deselect one pair before binding the results, or take the list and work with the `purrr::map` syntax:
 
-```{r}
-(res_tm <- res_sp2 %>% 
-  match_temporal(
-    match_id = group, data_id = type,
-    temporal_by = c("prcp" = "Water_course_level"),
-    return_cubble = TRUE, 
-    temporal_n_highest = 30,
-    temporal_min_match = 15) %>% 
-   bind_rows())
+```{r echo = TRUE}
+res_sp <- match_spatial(
+  climate_vic, river, 
+  spatial_n_group = 10, return_cubble = TRUE)
+str(res_sp, max.level = 0)
+(res_sp <- res_sp[-c(5, 8)] %>% bind_rows())
 ```
 
-<!--We first pass in the two datasets, followed by the variables used for the temporal matching, specified in \code{temporal_by} (using the \code{by} syntax from \code{dplyr::join}).
-With \code{temporal_independent} we select precipitation (\code{prcp}) as the independent variable, and the goal is to see if the water level in the river reflects what is observed in precipitation (for nearby stations).
-Since we consider a full year of data, we set the number of peaks considered (\code{temporal_n_highest}) to 30 (slightly above the default value of 20), and \code{temporal_min_match} is raised accordingly.-->
+For temporal matching, the variable water level (`Water_course_level`) from the river data will be matched to precipitation (`prcp`) in the weather station data.
+The variable identifying each matched group is `group` and the variable identifying the two datasets is `type`, and similarly, the cubble output can be returned with the argument `return_cubble = TRUE`.
+Here we select the four pairs with the highest number of matching peaks:
 
-This function returns a \code{cubble} object, with additional columns: \code{dist} storing the distance between matched stations, \code{group} summarizing spatial matching, and \code{n_match} showing the temporal matching.
+```{r echo = TRUE}
+res_tm <- res_sp %>% 
+  match_temporal(
+    data_id = type, match_id = group,
+    temporal_by = c("prcp" = "Water_course_level"),
+    return_cubble = TRUE)
+(res_tm <- res_tm %>% bind_rows() %>% filter(group %in% c(1, 7, 6, 9)))
+```
 
-Figure \ref{fig:matching} shows four matched pairs on the map (a) and standardized data as time series (b). The expected concurrent increase in precipitation and water level can be seen clearly.
+Figure \ref{fig:matching} shows four matched pairs on the map (a) and standardized data as time series (b) with concurrent increasing.
 
@@
 The water level reflects the increase in precipitation.
-The numbers (1, 5, 6, 10) indicate the group index derived from spatial matching, only those that were selectd by temporal matching are shown here."}
+The numbers (1, 7, 6, 9) indicate the group index derived from spatial matching, only those that were selectd by temporal matching are shown here."}
 res_tm <- res_tm %>%  filter(group %in% c(1, 7, 6, 9))
@@
   group_by(group, type) %>%
-  mutate(matched = (matched - min(matched, na.rm = TRUE))/ (max(matched, na.rm = TRUE) - min(matched, na.rm = TRUE)))
+  mutate(matched = (matched - min(matched, na.rm = TRUE))/ 
+           (max(matched, na.rm = TRUE) - min(matched, na.rm = TRUE)))
 

### 2023-06-15 jss/main.tex
text on matching

@@
 This can be considered to be a matching problem \citep{stuart2010matching, mcintosh2018using} to pair similar time series from nearby locations, or even a data fusion exercise that merges data collected from different sources \citep{cocchi2019data}.
-The function \code{match_sites()} in the \pkg{cubble} package provides a simple algorithm for this task.
-
-Spatial matching based on distance.
+
+In cubble, data can be matched spatially or temporarily with \texttt{match\_spatial()} and \texttt{match\_temporal()}.
+The function \texttt{match\_spatial()} calculates the spatial distance between observations in two cubbles.
+Different distances are available with projected or unprojected coordinate reference system.
+Analysts can subset the number of matched group to output with argument \texttt{spatial\_n\_group} (by default 4 groups) and the number of match per group with argument \texttt{spatial\_n\_group} (by default 1, that is, one-to-one matching).
+The syntax to use \texttt{match\_spatial()} is:
 
 \begin{verbatim}
-match_spatial(df1, df2, ...)
+match_spatial(<cubble_obj1>, <cubble_obj2>, ...)
 \end{verbatim}
 
-Allow one-to-many output, control the number of group and number of match within each group with arguments:
-
-Temporal matching, apply on spatially matched group
+The function \texttt{match\_temporal()} calculates the time series similarity between spatially matched groups.
+Two identifiers needs to be specified on the variable separates the each matched group (\texttt{match\_id}) and the variable separates the two sources (\texttt{data\_id}).
+The argument \texttt{temporal\_by} uses the \texttt{by} syntax from dplyr \texttt{*\_join} to specify the temporal matching variable.
+
+The similarity score between two time series in the spatially matched group is calculated by a matching function, which analysts can customise.
+The matching function should take two time series in a list and output a single numerical score, which allows for interfacing with existing time series distance calculation implementation.
+By default, cubble implements a simple peak matching algorithm (\texttt{match\_peak}) that counts the number of peaks in two time series that fall within a specified temporal window.
+The syntax to use \texttt{match\_temporal()} is:
 
 \begin{verbatim}
-match_temporal(<SPATIAL_RES>, group_id, match_id, temporal_by = c("xxx" = "xxx"))
+match_temporal(
+  <cubble_obj_from_match_spatial>, 
+  data_id = ... , match_id = ..., 
+  temporal_by = c("..." = "...")
+)
 \end{verbatim}
-
-Different matching functions can be tuned with \texttt{temporal\_match\_fn}, which takes a list of matched time series and additional arguments.
-By Default, the matching function calculates the number of peaks one series falls into another within a temporal window: \texttt{match\_peak}
 
@@
 
-River level and rainfall data for the same areas should have some similarity.
-Here we examine the river gauge data (\code{Water_course_level}) from the Bureau of Meteorology \citep{bom} in relation to weather station rainfall from NOAA's climate data (\code{climate}).
-The goal is to match water gauges with nearby weather stations, spatially and temporally, using the \code{match_sites()} function.
-
-This function requires passing the major and minor data sets used for matching, in this case those are \code{river} and \code{climate}.
-The variables used for the temporal matching are \code{Water_course_level} from the \code{river} data set and \code{prcp} in the climate data set.
-The rest of the arguments, as explained in Section \ref{matching}, correspond to the maximum and minimum number of peaks in the time series to be matched.
-In this example those are set to be a maximum of 30 and a minimum of 15 (approximately 2 and 1 per month).
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> (res_sp <- match_spatial(climate_vic, river, spatial_n_group = 10))
+Bureau of Meteorology collects \href{http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full}{water data} from river gauges and this includes variables: electrical conductivity, turbidity, water course discharge, water course level, and water temperature.
+In particular, water level will interactive with precipitation from the climate data since rainfall will raise the water level in the river.
+Both \texttt{climate\_vic} and \texttt{river} are cubble objects and we can get a summary of the 10 closest pairs:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> res_sp <- match_spatial(climate_vic, river, spatial_n_group = 10)
+R> print(res_sp, n = 20)
 \end{CodeInput}
@@
  5 ASN00085298 226027 4207.     5
-# i 5 more rows
-\end{CodeOutput}
-\end{CodeChunk}
-
-\begin{CodeChunk}
+ 6 ASN00082042 405234 6153.     6
+ 7 ASN00086038 230200 6167.     7
+ 8 ASN00086282 230200 6928.     8
+ 9 ASN00085279 224217 7431.     9
+10 ASN00080091 406756 7460.    10
+\end{CodeOutput}
+\end{CodeChunk}
+
+The result can also be returned as cubble objects with argument \texttt{return\_cubble\ =\ TRUE}.
+The output is a list where each element is a paired cubble object and you may consider combining all the results into a single cubble with \texttt{bind\_rows()}.
+Care needs to be taken on in the case when a site is close to two stations since by construction, cubble require unique rows in the nested form.
+From the summary table above, river station \texttt{226027} is matched to more than one weather station: \texttt{ASN00085072} (group 3) and \texttt{ASN00085298} (group 5).
+(Similarly river station \texttt{230200} is matched in group 7 and 8).
+One can either deselect one pair before binding the results, or take the list and work with the \texttt{purrr::map} syntax:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> res_sp <- match_spatial(
++   climate_vic, river, 
++   spatial_n_group = 10, return_cubble = TRUE)
+R> str(res_sp, max.level = 0)
+\end{CodeInput}
+\begin{CodeOutput}
+List of 10
+\end{CodeOutput}
+\begin{CodeInput}
+R> (res_sp <- res_sp[-c(5, 8)] %>% bind_rows())
+\end{CodeInput}
 \begin{CodeOutput}
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> (res_tm <- res_sp2 %>% 
+For temporal matching, the variable water level (\texttt{Water\_course\_level}) from the river data will be matched to precipitation (\texttt{prcp}) in the weather station data.
+The variable identifying each matched group is \texttt{group} and the variable identifying the two datasets is \texttt{type}, and similarly, the cubble output can be returned with the argument \texttt{return\_cubble\ =\ TRUE}.
+Here we select the four pairs with the highest number of matching peaks:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> res_tm <- res_sp %>% 
 +   match_temporal(
-+     match_id = group, data_id = type,
++     data_id = type, match_id = group,
 +     temporal_by = c("prcp" = "Water_course_level"),
-+     temporal_n_highest = 30,
-+     temporal_min_match = 15))
-\end{CodeInput}
-\begin{CodeOutput}
-# A tibble: 8 x 2
-  group match_res
-  <int>     <dbl>
-1     1        30
-2     2         5
-3     3        14
-4     4        20
-5     6        23
-# i 3 more rows
-\end{CodeOutput}
-\end{CodeChunk}
-
-Or use \code{return_cubble = TRUE} to return a cubble object:
-
-\begin{CodeChunk}
-\begin{CodeOutput}
-# cubble:   key: id [16], index: date, nested form, [sf]
-# spatial:  [144.5203, -38.144913, 148.4667, -36.128657], WGS 84
++     return_cubble = TRUE)
+R> (res_tm <- res_tm %>% bind_rows() %>% filter(group %in% c(1, 7, 6, 9)))
+\end{CodeInput}
+\begin{CodeOutput}
+# cubble:   key: id [8], index: date, nested form, [sf]
+# spatial:  [144.5203, -37.8817, 147.572223, -36.8472], WGS 84
 # temporal: date [date], matched [dbl]
@@
 2 406213     145. -37.0  NA   CAMP~     NA river (144.5403 -37.01512)     1
-3 ASN00084~  148. -37.7  62.7 orbo~  95918 clim~  (148.4667 -37.6922)     2
-4 222201     148. -37.7  NA   SNOW~     NA river  (148.451 -37.70739)     2
-5 ASN00085~  147. -38.1   4.6 east~  94907 clim~  (147.1322 -38.1156)     3
-# i 11 more rows
+3 ASN00082~  146. -36.8 502   stra~  95843 clim~  (145.7308 -36.8472)     6
+4 405234     146. -36.9  NA   SEVE~     NA river (145.6828 -36.88701)     6
+5 ASN00086~  145. -37.7  78.4 esse~  95866 clim~  (144.9066 -37.7276)     7
+# i 3 more rows
 # i 3 more variables: dist [m], ts <list>, match_res <dbl>
@@
 
-This function returns a \code{cubble} object, with additional columns: \code{dist} storing the distance between matched stations, \code{group} summarizing spatial matching, and \code{n_match} showing the temporal matching.
-
-Figure \ref{fig:matching} shows four matched pairs on the map (a) and standardized data as time series (b). The expected concurrent increase in precipitation and water level can be seen clearly.
+Figure \ref{fig:matching} shows four matched pairs on the map (a) and standardized data as time series (b) with concurrent increasing.
 
@@
 The water level reflects the increase in precipitation.
-The numbers (1, 5, 6, 10) indicate the group index derived from spatial matching, only those that were selectd by temporal matching are shown here.}\label{fig:matching}
+The numbers (1, 7, 6, 9) indicate the group index derived from spatial matching, only those that were selectd by temporal matching are shown here.}\label{fig:matching}
 \end{figure}

### 2023-06-18 2-cubble.Rmd
text update as per pkg vignettes

@@
 
-[conceptual framework] Spatio-temporal data comes various spatial and temporal characteristics and requires different data structures to wrangle: climate weather stations are recorded at fixed point location but with potential temporal data quality issue (missingness on the day); GPS data tracks unique point locations at different timestamps; satellite imageries captures snapshots of landscape at selected time.
-The type of spatio-temporal data cubble tackles are those collected at unique fixed locations while allowed for irregularity in the temporal dimension, like the weather station data.
-In the four layouts presented by the spacetime paper [@spacetime], cubble handles full space-time and sparse space-time layouts.
+[conceptual framework]
 
-The cubble class is an S3 class, built on tibble, to pivot spatio-temporal data into a nested (spatial) form and a long (temporal) form. It has two subclasses:
+Spatio-temporal data can encompass data with various spatial and temporal characteristics and different data require different structures for wrangling and analysis: climate weather stations record data at fixed point location but may suffer from potential temporal data quality issue such as missing data for certain days.
+GPS data tracks unique point locations at different timestamps and is represented as trajectories.
+Satellite imageries capture snapshots of landscapes at selected time and is commonly structured as raster data.
+The spatio-temporal data cubble address are those collected at unique fixed locations while allowing for irregularity in the temporal dimension, such as the weather station data.
+This corresponds to the full space-time and sparse space-time layouts in the spacetime paper [@spacetime].
 
--   a nested/ spatial cubble has class `c("spatial_cubble_df", "cubble_df")`
--   a long/ temporal cubble has class `c("temporal_cubble_df", "cubble_df")`
+The cubble class is an S3 class built on tibble that allows the spatio-temporal data to be wrangled in two forms:  a nested/ spatialform and a long/temporal form. It consists of two subclasses:
 
-A nested cubble arranges spatial variables in columns and nests temporal variables in a specialised `ts` column.
-Below prints an nested cubble object using the weather station data collected in three airports (the creation of a cubble object will be explained in Section \@ref(create)).
+-   a nested/ spatial cubble is represented by the class `c("spatial_cubble_df", "cubble_df")`
+-   a long/ temporal cubble is represented by the class `c("temporal_cubble_df", "cubble_df")`
+
+In a nested cubble, spaital variables are organised as columns and temporal variables are nested within a specialised `ts` column.
+The nested cubble object printed below contains weather records in three airport stations (the creation of a cubble object will be explained in Section \@ref(create)).
 This toy data is a subset of a larger data \code{climate_aus} collected from Global Historical Climatology Network Daily (GHCND).
-The three airport stations in Melbourne are recorded with station metadata: station ID, longitude, latitude, elevation, station name, world meteorology organisation ID.
-The temporal variables are precipitation, maximum and minimum temperature, which can be read from the cubble header:
+It records three airport stations located in Melbourne, Australia and includes spatial variables such as station ID, longitude, latitude, elevation, station name, World Meteorology Organisation ID.
+The dataset contains temporal variables including precipitation, maximum and minimum temperature, which can be read from the cubble header:
 
@@
 
-A long cubble expands the temporal variables into the long form and stores the spatial variables as a data attribute.
-In the header of a long cubble object, the extent now prints the temporal range, the interval, and whether there is gaps in the data.
-It also prints the available spatial variables, as oppose to the temporal variables in the nested cubble.
+In a long cubble, temporal variables are expanded in the long form and spatial variables are stored as a data attribute:
 
@@
 
-A cubble object inherits the attributes from tibble (and its subclasses): `class`, `row.names`, and `names`, in addition to three specialised attributes:
+A cubble object inherits the attributes from tibble (and its subclasses): `class`, `row.names`, and `names`. Additionally, it has three specialised attributes:
 
@@
 
-Readers known the `key` and `index` attributes from the `tsibble` package would already be familiar the two arguments.
-In cubble, the `key` attribute identifies the row in the nested cubble, and together with the `index` argument, identifies the row in the long cubble.
-Currently, cubble only supports one variable as the key and the accepted temporal class for index includes the base R class `Date`, `POSIXlt`, `POSIXct` and tsibble's `yearmonth`, `yearweek`, and `yearquarter` class.
+Readers who are familiar with the `key` and `index` attributes from the `tsibble` package would already understand the two arguments.
+In cubble, the `key` attribute identifies the row in the nested cubble, and when combined with the `index` argument, it identifies the row in the long cubble.
+Currently, cubble only supports one variable as the key, and the accepted temporal classes for index includes the base R classes `Date`, `POSIXlt`, `POSIXct`, as well as tsibble's `yearmonth`, `yearweek`, and `yearquarter` classes.
 
-The `coords` attribute takes an ordered pair of coordinate.
-It can be a unprojected pair of longitude and latitude, or a projected easting and northing values.
-Under the hood, the `sf` package is used to calculate the bounding box, shown in the header of a nested cubble, and other spatial operations.
+The `coords` attribute represents an ordered pair of coordinates.
+It can be either an unprojected pair of longitude and latitude, or a projected easting and northing value.
+The `sf` package is used under teh hood to calculate the bounding box, displayed in the header of a nested cubble.
 
-The long cubble has a special attribute `spatial` to store the spatial variables: all the variables in the nested cubble, except for the `ts` column.
-The shortcut function are available to extract components in the attributes, for example, \code{spatial()} for extracting the tibble object for the spatial variables:
+The long cubble has a special attribute called `spatial` to store the spatial variables, which includes all the variables from the nested cubble, except for the `ts` column.
+Shortcut functions are available to extract attributes, for example, \code{spatial()} for extracting spatial variables from the long cubble:
 
@@
 
-Spatio-temporal data may arrive in separate tables for analysts.
-For example, in climate data, analysts may initially receive station data containing geographic location information, variables recorded and their recording periods.
-They can then query the temporal variables using the stations of interest to obtain the corresponding temporal data.
+In many cases, spatio-temporal data arrive in separate tables for analysis.
+For example, in climate data, analysts may initially receive station data containing geographic location information, recorded variables and their recording periods.
+They can then query the temporal variables using the stations of interest to obtain the relevant temporal data.
 Alternatively, analyses may begin as purely spatial or temporal, and analysts may obtain additional temporal or spatial data to expand the result to spatio-temporal.
 
-The function \code{make_cubble()} composes a \code{cubble} object from a spatial table (\code{spatial}) and a temporal table (\code{temporal}) with the three attributes introduced in the subsection \@ref(object): \code{key}, \code{index}, and \code{coords}.
+The function \code{make_cubble()} composes a \code{cubble} object from a spatial table (\code{spatial}) and a temporal table (\code{temporal}), along with three attributes introduced in the subsection \@ref(object): \code{key}, \code{index}, and \code{coords}.
 The following code creates the nested \code{cubble} object shown previously:
@@
 This includes casting from a \code{tibble} or \code{data.frame} with both spatial and temporal information, a NetCDF object, a \code{stars} object [@stars], and a \code{sftime} object [@sftime].
-The two examples below show the casting from a tibble and a NetCDF object to a cubble object.
-
-The dataset \code{climate_flat} joins the spatial data, \code{stations}, with the temporal data, \code{meteo}, into a single tibble object and it can be coerced into a cubble using:
+The two examples below show the casting from a tibble and a NetCDF object.
+The dataset \code{climate_flat} combines the spatial data, \code{stations}, with the temporal data, \code{meteo}, into a single tibble object.
+It can be coerced into a cubble using:
 
@@
 
-The NetCDF data is another format commonly used for storing spatio-temporal data.
-In \proglang{R}, packages for wrangling NetCDF data include a high-level R interface: \pkg{ncdf4} [@ncdf4], a low-level interface that calls a C-interface: \pkg{RNetCDF} [@rnetcdf], and a tidyverse implementation: \pkg{tidync} [@tidync].
-The code below casts a NetCDF object in the \code{ncdf4} class into a cubble object:
+In \proglang{R}, there are several packages available for wrangling NetCDF data, including \code{ncdf4}, \code{RNetCDF}, and \code{tidync}.
+The code below converts a NetCDF object of class \code{ncdf4} into a cubble object:
 
@@
 
-Sometimes, one may want to read in a subset of the NetCDF data and the argument \code{vars}, \code{long_range} and \code{lat_range} can be used to subset on the variable and the grid resolution:
+Sometimes, analysts may choose to read onlny a subset of the NetCDF data.
+In such cases, the arguments \code{vars}, \code{long_range} and \code{lat_range} can be used to subset the data based on variable and the grid resolution:
 
@@
 
-We would recommend reducing to about 300 $\times$ 300 grid points for three daily variables in one year.
-A 300 by 300 spatial grid can be a bounding box of [100, -80, 180, 0] at 0.25 degree resolution or a global bounding box [-180, -90, 180, -90] at 1 degree resolution.
-
 ## Functions and methods
 
-Table \@ref(tab:funs) summarises functions implemented in the \pkg{cubble} package and Table \@ref(tab:methods) details the methods implemented for each of the three cubble classes: \code{cubble\_df}, \code{spatial\_cubble\_df}, and \code{temporal\_cubble\_df}.
+Table \@ref(tab:funs) summarises the functions implemented in the \pkg{cubble} package and Table \@ref(tab:methods) details the methods implemented for each of the three cubble classes.
+The \code{cubble_df} class handles methods that behave consistently in both nested (\code{spatial_cubble_df}) and long cubble (\code{temproal_cubble_df}).
+When the method has distinct behavior for the nested cubble  and temporal cubble, it is implemented separately for each subclass.
 
@@
 | tidyverse | `dplyr_row_slice`, `dplyr_col_modify`, `dplyr_reconstruct`, `select`, `mutate`, `arrange`, `filter`, `group_by`, `ungroup`, `summarise`, `select`, `slice`, `rowwise`, `rename`, `bind_rows`, `bind_cols`, `relocate`, `type_sum`, the slice family (`slice_head`, `slice_tail`, `slice_max`, `slice_min`, `slice_sample`) and the join family (`left_join`, `right_join`, `inner_join`, `full_join`, `anti_join`, `semi_join`) |
-| cubble    | `as_cubble`, `cubble`, `make_cubble`, `check_key` `face_temporal`, `face_spatial`, `unfold`, `key`, `key_vars`, `key_data`, `index`, `index_var`, `coords`, `spatial`, `match_sites`, `match_spatial`, `match_temporal`, `geom_glyph`, `geom_glyph_box`, `geom_glyph_line`, `make_spatial_sf`, `make_temporal_tsibble`, `fill_gaps`, `scan_gaps`                                                                                |
+| cubble    | `as_cubble`, `cubble`, `make_cubble`, `check_key` `face_temporal`, `face_spatial`, `unfold`, `key`, `key_vars`, `key_data`, `index`, `index_var`, `coords`, `spatial`, `match_sites`, `match_spatial`, `match_temporal`, `geom_glyph`, `geom_glyph_box`, `geom_glyph_line`, `make_spatial_sf`, `make_temporal_tsibble`, `fill_gaps`, and `scan_gaps`                                                                                |
 
@@
 
-: (#tab:methods) Methods implemented for the three \code{cubble} classes: \code{cubble\_df}, \code{spatial\_cubble\_df}, and \code{temporal\_cubble\_df}
+: (#tab:methods) Methods implemented for the three \code{cubble} classes.
 
-<!-- methods(class = "cubble_df") |> paste0(collapse = ", ") |> stringr::str_remove_all(".cubble_df") -->
-
-<!-- methods(class = "spatial_cubble_df") |> paste0(collapse = ", ") |> stringr::str_remove_all(".spatial_cubble_df") -->
-
-<!-- methods(class = "temporal_cubble_df") |> paste0(collapse = ", ") |> stringr::str_remove_all(".temporal_cubble_df") -->
-
-### dplyr
-
-The \pkg{dplyr} package has many tools for wrangling tidy data, many of which are useful in the spatio-temporal analysis.
-The \pkg{cubble} package provides methods that support the use of the following operations in the \pkg{dplyr} package on both the nested and long forms: `mutate`, `filter`, `summarise`, `select`, `arrange`, `rename`, `left_join`, and the slice family (`slice_*`).
-
-The toy example below demonstrates the using of a collection of dplyr functions (\code{mutate}, \code{select}, \code{filter}, and \code{arrange}) on a nested cubble object \code{cb_nested}:
-
-```{r echo = TRUE, eval = FALSE}
-cb_nested %>% 
-  mutate(avg_temp = mean(ts$tmax, na.rm = TRUE)) %>% 
-  select(-elev) %>% 
-  filter(avg_temp > 26) %>% 
-  arrange(-avg_temp)
-```
-
-### cubble
-
-The three basic cubble functions, \code{face_temporal()}, \code{face_spatial()}, and \code{unfold()}, are introduced in this section before more advanced functionalities in Section \@ref(others).
-
-The pair of verbs, \code{face_temporal()} and \code{face_spatial()}, pivot the cubble object between the spatial and temporal face of the multivariate spatio-temporal cube, as illustrated in Figure \ref{fig:face}.
+Specifically, the pair of verbs, \code{face_temporal()} and \code{face_spatial()}, pivot the cubble object between the spatial and temporal face of the multivariate spatio-temporal cube, as illustrated in Figure \ref{fig:face}.
 The code below uses the function \code{face_temporal()} and \code{face_spatial()} to transform between a nested cubble (\code{cb_nested}) and a long one (\code{cb_long}), introduced earlier in subsection \@ref(object):
@@
 
-Sometimes, analysts may need to apply some variable transformation that involves both the spatial and temporal variables.
-An example of this is the transformation of temporal variables into the spatial dimension in glyph maps [@Wickham2012-yr].
-(How to make glyph maps will be explained in Section \ref{st_transformation}, and are illustrated in the second example.) This type of operation can be seen as flattening, or *unfolding*, the cube into a 2D data frame.
-Here the function \code{unfold()} moves the spatial variables \code{long} and \code{lat} into the long \code{cubble}:
+## Compatibility with tsibble and sf
 
-```{r echo = TRUE}
-cb_long |> unfold(long, lat)
-```
+Analysts often have their preferred spatial or temporal data structure that they prefer to use for spatio-temporal analysis.
+For example, the \code{tbl_ts} class from the \code{tsibble} package [@tsibble] is commonly used in time series forecasting and similarly, the \code{sf} class [@sf] is frequently used in spatial data science.
+In \code{cubble}, analysts have the flexibility to combine these two structures together by allowing the spatial component to also be an \code{sf} object and the temporal component to also be a \code{tsibble} object.
 
-## Compatibility with other packages
-
-Analysts often have their own preferred spatial or temporal data structure, which they may wish to keep for spatio-temporal analysis.
-For example, the \code{tbl_ts} class from the \code{tsibble} package [@tsibble] is commonly used in time series forecasting and similarly, the \code{sf} class [@sf] is often used in spatial data science.
-In \code{cubble}, analysts can combine these two structures together by allowing the spatial component to also be an \code{sf} object and the temporal component to also be a \code{tsibble} object.
-
-### tsibble
+### Using a tsibble object as the temporal component
 
@@
 
-### sf
+### Using an sf object as the spatial component
 
@@
 ```{r echo = TRUE}
-sf_nested <- cb_nested %>% make_spatial_sf() 
-all.equal(sf_nested, sf_nested)
+cb_nested %>% make_spatial_sf() 
 ```
@@
 
-Given the vast spatio-temporal data structure available in \proglang{R} to represent spatio-temporal data, this section compares and contrast the \pkg{cubble} with other existing alternative, namely \pkg{stars} and \pkg{sftime}.
+In \proglang{R}, there are other existing spatio-temporal data structure and this section compares and contrasts \pkg{cubble} with other existing alternative, specifically \pkg{stars} and \pkg{sftime}.
+The \pkg{stars} package [@stars] uses an array structure, as oppose to tibble, to represent multivariate spatio-temporal data.
+While both \pkg{stars} and \pkg{cubble} support vector and raster data, it is a matter of choice on which structure to use given the application.
+Analysts working on satellite imageries may prefer the array structure in \pkg{stars}, while others originally working with spatio-temporal data in 2D data frames may find \pkg{cubble} easier to adopt from their existing computing workflow.
 
-The \pkg{stars} package [@stars] uses an array structure, as oppose to tibble in the \code{cubble} package, to represent multivariate spatio-temporal data.
-While both packages support vector and raster data, it is a matter of choice on which structure to use given the application.
-Analysts working on satellite imageries may prefer the array structure while others originally working with spatio-temporal data in the 2D data frames may find \pkg{cubble} easier to adopt from their existing computing workflow.
-
-The \pkg{sftime} package [@sftime] also builds from a tibble object, with a main focus to handle irregular spatio-temporal data.
-While \pkg{sftime} can also handle full space-time grids and sparse space-time layouts, \pkg{cubble} can be thought of as simplifying the repetition in spatial variables with two linked tibbles.
-This simplification gives memory efficiency when data is observed frequent, i.e.
-daily or subdaily, or the spatial geometry is expensive to repeat, i.e.
+The \pkg{sftime} package [@sftime] also builds from a tibble object and its focus is on handling irregular spatio-temporal data.
+This means \pkg{sftime} can also handle full space-time grids and sparse space-time layouts represented in \pkg{cubble}, but \pkg{cubble} uses nesting to avoid storing spatial variables repetitively at each timestamp.
+This provides memory efficiency when data is observed frequent, i.e.
+daily or sub-daily, or the spatial geometry is expensive to repeat, i.e.
 polygons or multipolygons.

### 2023-06-18 3-advanced.Rmd
text update as per pkg vignettes

@@
 
-One task that may interest spatio-temporal analysts is combining data collected at nearby but not exactly the same sites, for example, weather station measured rainfall and river levels.
-This can be considered to be a matching problem [@stuart2010matching; @mcintosh2018using] to pair similar time series from nearby locations, or even a data fusion exercise that merges data collected from different sources [@cocchi2019data].
-
-In cubble, data can be matched spatially or temporarily with `match_spatial()` and `match_temporal()`.
-The function `match_spatial()` calculates the spatial distance between observations in two cubbles.
-Different distances are available with projected or unprojected coordinate reference system.
-Analysts can subset the number of matched group to output with argument `spatial_n_group` (by default 4 groups) and the number of match per group with argument `spatial_n_group` (by default 1, that is, one-to-one matching).
+One common task when working with spatio-temporal data is to match nearby sites.
+For example, we may want to verify the location of an old list of stations with current stations, or we may want to match the data from different data sources.
+In cubble, spatial and temporal matching are performed using the functions `match_spatial()` and `match_temporal()`.
+The `match_spatial()` function calculates the spatial distance between observations in two cubble objects.
+Various distance measures are available (check `sf::st_distance`).
+Analysts can specify the number of matched groups to output using the `spatial_n_group` argument (default to 4 groups) and the number of matches per group using the `spatial_n_group` argument (default to 1, one-to-one matching).
 The syntax to use `match_spatial()` is:
@@
 
-The function `match_temporal()` calculates the time series similarity between spatially matched groups.
-Two identifiers needs to be specified on the variable separates the each matched group (`match_id`) and the variable separates the two sources (`data_id`).
-The argument `temporal_by` uses the `by` syntax from dplyr `*_join` to specify the temporal matching variable.
+The function `match_temporal()` calculates the similarity between time series within spatially matched groups.
+Two identifiers are required: one for separating each matched group (`match_id`) and one for separating the two data sources (`data_id`).
+The argument `temporal_by` uses the `by` syntax from dplyr's `*_join` to specify the temporal variables to match.
 
-  The similarity score between two time series in the spatially matched group is calculated by a matching function, which analysts can customise.
-The matching function should take two time series in a list and output a single numerical score, which allows for interfacing with existing time series distance calculation implementation.
+The similarity score between two time series is calculated using a matching function, which can be customised by the analysts.
+The matching function takes two time series as a list and returns a single numerical score.
+This allows for flexibility in using existing time series distance calculation implementation.
 By default, cubble implements a simple peak matching algorithm (`match_peak`) that counts the number of peaks in two time series that fall within a specified temporal window.
-The syntax to use `match_temporal()` is:
+The syntax to use `match_temporal()` is
 
@@
   <cubble_obj_from_match_spatial>, 
-  data_id = ... , match_id = ..., 
+  data_id = , match_id = , 
   temporal_by = c("..." = "...")
@@
 
-Spatio-temporal data lends itself to a range of transformations.
-Glyph maps [@Wickham2012-yr] transform the measured variable and time coordinates into microplots at the spatial locations.
-Calendar plots [@wang2020calendar] deconstruct time to produce plots of variables in a calendar format.
-Summarizing multiple variables is commonly done using projections, or linear combinations.
-Here we elaborate on the transformations made to produce a glyph map.
-The package \pkg{GGally} [@ggally] has implemented glyph maps through the \code{glyphs()} function.
-The function constructs a \code{data.frame} with calculated position (\code{gx}, \code{gy}, \code{gid}) of each point on the time series using linear algebra (Equations 1 and 2 in @Wickham2012-yr).
-A new implementation of the glyph map as a ggproto, \code{GeomGlyph}, has been made in the \pkg{cubble} package so that a glyph map can be created with \code{geom_glyph()}:
+Sometimes, we wish to communicate spatial and temporal information collectively through visualisation.
+This can be achieved through several graphical displays: one can make faceted maps across time, creating map animations, or constructing interactive graphics to link between map and time series plot.
+The concept of glyph maps was initially proposed in @Wickham2012-yr.
+The underlying idea is to transform the temporal coordinates into spatial coordinates so that time series plot can be displayed on the map.
 
-```{r eval = FALSE, echo = TRUE}
+The package `GGally` initially implement the glyph map.
+It uses `glyphs()` to calculate the axis transformation and then uses `geom_polygon()` to draw the map.
+In cubble, a ggproto implementation `geom_glyph()` performs the linear algebra internally as data transformation .
+The `geom_glyph()` requires four aesthetics: `x_major`, `y_major`, `x_minor`, and `y_minor`.
+The major axes are the outer spatial coordinates and the minor axes are the inner/ temporal coordinates:
+
+```
 data |> 
@@
 
-Some useful controls over the glyph map are also available in the \code{geom_glyph()} implementation
+Some useful controls over the glyph map includes:
 
-  * polar coordinate glyph maps are specified using \code{polar = TRUE}, 
-  * arguments \code{width} and \code{height} can be specified in either absolute or relative value, 
-  * global and local scale is specified with \code{global_rescale}, which defaults to \code{TRUE}, and 
-  *reference boxes and lines can be added with geoms: \code{geom_glyph_box()} and \code{geom_glyph_line()}.
+  * polar coordinate glyph maps with \code{polar = TRUE}, 
+  * adjust glyph size with arguments \code{width} and \code{height},
+  * glyph scale to fixed (\code{global_rescale} defaults to \code{TRUE}) or free, and 
+  * reference boxes and lines with \code{geom_glyph_box()} and \code{geom_glyph_line()}.

### 2023-06-18 4-3-matching.Rmd
text update as per pkg vignettes

@@
 
-Bureau of Meteorology collects [water data](http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full) from river gauges and this includes variables: electrical conductivity, turbidity, water course discharge, water course level, and water temperature.
-In particular, water level will interactive with precipitation from the climate data since rainfall will raise the water level in the river.
-Both `climate_vic` and `river` are cubble objects and we can get a summary of the 10 closest pairs:
+One common task when working with spatio-temporal data is to match nearby sites.
+For example, we may want to verify the location of an old list of stations with current stations, or we may want to match the data from different data sources.
+In this example, we will introduce the spatial and temporal matching in cubble using an example on matching river level data with precipitation in Victoria, Australia.
+
+The water level data collected by the Bureau of Meteorology, can be compared with the precipitation since rainfall can directly impact water level in river.
+Both `climate_vic` and `river` are cubble objects, and we can obtain a summary of the 10 closest pairs between them:
 
@@
 
-The result can also be returned as cubble objects with argument `return_cubble = TRUE`.
-The output is a list where each element is a paired cubble object and you may consider combining all the results into a single cubble with `bind_rows()`.
-Care needs to be taken on in the case when a site is close to two stations since by construction, cubble require unique rows in the nested form.
-From the summary table above, river station `226027` is matched to more than one weather station: `ASN00085072` (group 3) and `ASN00085298` (group 5).
-(Similarly river station `230200` is matched in group 7 and 8).
-One can either deselect one pair before binding the results, or take the list and work with the `purrr::map` syntax:
+The result can also be returned as cubble objects by setting the argument `return_cubble = TRUE`.
+The output is be a list where each element is a paired cubble object.
+To combine all the results into a single cubble, you can use `bind_rows()`.
+In the case when a site in the second cubble (the `river` data here) is matched to two stations in the first cubble (`climate_vic` here), the binding may not be successful since cubble requires unique rows in the nested form.
+In the summary table above, the river station `226027` is matched to more than one weather station: `ASN00085072` (group 3) and `ASN00085298` (group 5).
+Similarly, the river station `230200` is matched in group 7 and 8).
+In such cases, you can either deselect one pair before combining, or work with the list output with the `purrr::map` syntax:
 
@@
 
-For temporal matching, the variable water level (`Water_course_level`) from the river data will be matched to precipitation (`prcp`) in the weather station data.
-The variable identifying each matched group is `group` and the variable identifying the two datasets is `type`, and similarly, the cubble output can be returned with the argument `return_cubble = TRUE`.
-Here we select the four pairs with the highest number of matching peaks:
+For temporal matching, we match the variable `Water_course_level` from the river data to `prcp` in the weather station data.
+The variable `group` and `types` identify the matching group and the two datasets:
+
+```{r echo = TRUE}
+(res_tm <- res_sp %>% 
+  match_temporal(
+    data_id = type, match_id = group,
+    temporal_by = c("prcp" = "Water_course_level")))
+```
+
+Similarly, the cubble output can be returned using the argument `return_cubble = TRUE`. Here we select the four pairs with the highest number of matching peaks:
 

### 2023-06-18 4-5-interactive.Rmd
text update as per pkg vignettes

--- 2022-06-22 4-5-interactive.Rmd
@@
 
-Interactive graphics can be useful because they make it possible to look at the data in a multiple of ways on-the-fly.
+Interactive graphics can be useful because they make it possible to look at the data in multiple ways on-the-fly.
 This is especially important for spatio-temporal data, where we would like to interactively connect spatial and temporal displays.
-This example describes the process of using \pkg{cubble} with \pkg{crosstalk} to build an interactive display connecting a map of Australia, with ribbon plots of temperature range observed at the stations.
+This example describes the process of using the \pkg{cubble} package with the \pkg{crosstalk} package to build an interactive display connecting a map of Australia, with ribbon plots of temperature range observed at the stations.
 The purpose is to explore the variation of monthly temperature range over the country.
@@
 
-<!--When dealing with spatio-temporal data, users may wish to make plots to learn the spatial distribution of a variable or to find patterns, such as trend or seasonality, in the time series.
-Combining these two types of plots with interactivity lets users link between points on the map and the corresponding time series to explore the spatial and temporal dimensions of the data simultaneously.-->
-
-The key steps are to convert both the nested and long forms of the data into shared objects, and to plot these sid-by-side. The two are linked by the station identifier.
-
-<!--The temperature range is calculated as the difference between \code{tmax} and \code{tmin} and its monthly average over 2016-2020 is taken before calculating the variance.
-A \code{SharedData} object is constructed for each form of the cubble and the same \code{group} argument ensures the cross-linking of the two forms via the common \code{id} column.
-The spatial map and time series plot are then made with each \code{SharedData} object separately.
-In this example, stations on the Australia map, made from the nested form, are coloured by the calculated variance.
-A ribbon band is constructed using the long form cubble to show each station's maximum and minimum temperature over time.
-In the case of a different dataset, users could calculate other station-dependent measures in the nested form or make other time-wise summary of the data in the long form to customise the spatial or temporal view.
-The cross-linking between the two plots is always safeguarded by the shared \code{id} column embedded in the cubble structure.
-Below is the pseudo-code that outlines the process to construct the interactive graphic described above:-->
+The key steps are to convert both the nested and long forms of the data into shared \code{crosstalk} objects, and to plot these side-by-side. The two are linked by the station identifier.
 
 ``` r
-# data pre-processing
-clean <- climate_full %>% ...
+clean <- climate_full |> ...
 
-# created SharedData instance for crosstalk
-nested <- clean %>% SharedData$new(~id, group = "cubble")
-long <- face_temporal(clean) %>% SharedData$new(~id, group = "cubble")
+nested <- clean |> SharedData$new(~id, group = "cubble")
+long <- face_temporal(clean) |> SharedData$new(~id, group = "cubble")
 
-# create the spatial and temporal view each with a ShareData instance
-p1 <- nested %>% ...
-p2 <- long %>% ...
+p1 <- nested |> ...
+p2 <- long |> ...
 
-# Combine p1 and p2
 crosstalk::bscols(plotly::ggplotly(p1), plotly::ggplotly(p2), ...)
@@
 Plot (a) shows the initial state of the interactive display: all locations are shown as dots on the map, coloured by temperature range, and the right plot shows the ribbons representing maximum to minimum for all stations.
-In plot (b) the "mount elizabeth" station, which shows a high variance colour on the initial map, is selected on the map and this produces the ribbon on the right.
-In plot (c) the lowest temperature in August is selected, which is "thredbo" station on the left map.
-It was surprising to us that this was not a station in Tasmania, so for comparison a station in Tasmania is selected on the map to show in relation to thredbo.
-We can see that thredbo has a bigger winter dip in temperature, and although Tasmania is cold generally, it's temperatures are more consistent.
+In plot (b) the "Mount Elizabeth" station, which shows a high variance colour on the initial map, is selected on the map and this produces the ribbon on the right.
+In plot (c) the lowest temperature in August is selected, which is "Thredbo" station on the left map.
+It was surprising to us that this was not a station in Tasmania, so for comparison a station in Tasmania is selected on the map to show in relation to Thredbo.
+We can see that Thredbo has a bigger winter dip in temperature, and although Tasmania is cold generally, it's temperatures are more constant
 
-<!-- In Figure \ref{fig:interactive-linking}, the first row shows the initial view of the interactive graphic.
-Most regions in Australia have a low variance of temperature range, while the North-West coastline, the bottom of South Australia, and Victoria stand out with larger monthly changes.
-In the second row, the Mount Elizabeth station, which shows a high variance colour on the initial map, is selected on the map and this produces the ribbon on the right.
-In the third row, the lowest temperature in August is selected, which highlights the corresponding thredbo station on the left map.
--->
+```{r eval = FALSE}
+# This chunk generates the interactive graphic and save it as figures/linking.html, where the three screencasts in the figure comes from.
+library("crosstalk")
+library("plotly")
+
+clean <- climate_full |>
+  filter(!id %in% c("ASN00067033", "ASN00072091", "ASN00059040", 
+                    "ASN00097053", "ASN00040856", "ASN00015548")) |>
+  face_temporal() |> 
+  mutate(month = lubridate::month(date)) |>
+  group_by(month) |>
+  summarise(
+    tmax = mean(tmax, na.rm = TRUE),
+    tmin = mean(tmin, na.rm = TRUE),
+    diff = mean(tmax - tmin, na.rm = TRUE)
+    ) |> 
+  ungroup(month) |>
+  mutate(dummy_date = as.Date(glue::glue("2021-{month}-01"))) |> 
+  face_spatial() |> 
+  mutate(temp_diff_var = ts |> pull(diff) |> var(na.rm = TRUE))
+
+nested <- clean |> 
+  SharedData$new(~id, group = "cubble")
+
+long <- clean |> 
+  face_temporal() |> 
+  unfold(temp_diff_var) |> 
+  arrange(temp_diff_var) |> 
+  SharedData$new(~id, group = "cubble")
+
+state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)
+
+p1 <- nested |> 
+  ggplot() +
+  geom_sf(data = state_map, fill = "grey95", color = "white") +
+  geom_point(aes(x = long, y = lat,color = temp_diff_var, label = name)) +
+  colorspace::scale_color_continuous_sequential(
+    "Rocket",  n_interp = 7, cmax = 90, rev = TRUE, 
+    c2 = 40, l2= 85, c1 = 20, l1 = 30, name = "Var. temp. diff.") +
+  labs(x = "Longitude", y = "Latitude") + 
+  coord_sf() + 
+  theme_bw() + 
+  theme(
+    panel.grid.major = element_blank(),
+    legend.position = "none"
+    )
+
+p2 <- long |> 
+  ggplot(aes(x = dummy_date, group = id, 
+             fill = temp_diff_var, color = temp_diff_var)) +
+  geom_ribbon(aes(ymin = tmin, ymax = tmax), size = 0.1, alpha = 0.3) + 
+  geom_line(aes(y = tmin), alpha = 0.1) + 
+  geom_line(aes(y = tmax), alpha = 0.1) + 
+  geom_point(aes(y = tmax), size = 0.1) + 
+  geom_point(aes(y = tmin), size = 0.1) + 
+  colorspace::scale_fill_continuous_sequential(
+    "Rocket",  n_interp = 7, cmax = 90, rev = TRUE, 
+    c2 = 40, l2= 85, c1 = 20, l1 = 30, name = "Var. temp. diff.") +
+  colorspace::scale_colour_continuous_sequential(
+    "Rocket",  n_interp = 7, cmax = 90, rev = TRUE, 
+    c2 = 40, l2= 85, c1 = 20, l1 = 30, name = "Var. temp. diff.") +
+  scale_x_date(date_labels = "%b", date_breaks = "1 month") + 
+  labs(x = "Month", y = "Temperature") + 
+  theme_bw() + 
+  theme(
+    panel.grid.major = element_blank(),
+    legend.position = "bottom"
+    )
+
+out <- bscols(
+  ggplotly(p1, width = 900, height = 700, tooltip = "label") |>
+    highlight(
+      on = "plotly_selected", 
+      off = "plotly_deselect", 
+      opacityDim = 0.1),
+  ggplotly(p2, width = 1200, height = 700, tooltip = "label") |> 
+    highlight(
+      on = "plotly_selected", 
+      off = "plotly_deselect", 
+      opacityDim = 0.012),
+  widths = c(5, 5)
+)
+
+htmltools::save_html(out, file = "figures/linking.html")
+
+```
 

### 2023-06-18 jss/main.tex
text update as per pkg vignettes

@@
 
-{[}conceptual framework{]} Spatio-temporal data comes various spatial and temporal characteristics and requires different data structures to wrangle: climate weather stations are recorded at fixed point location but with potential temporal data quality issue (missingness on the day); GPS data tracks unique point locations at different timestamps; satellite imageries captures snapshots of landscape at selected time.
-The type of spatio-temporal data cubble tackles are those collected at unique fixed locations while allowed for irregularity in the temporal dimension, like the weather station data.
-In the four layouts presented by the spacetime paper \citep{spacetime}, cubble handles full space-time and sparse space-time layouts.
-
-The cubble class is an S3 class, built on tibble, to pivot spatio-temporal data into a nested (spatial) form and a long (temporal) form. It has two subclasses:
+{[}conceptual framework{]}
+
+Spatio-temporal data can encompass data with various spatial and temporal characteristics and different data require different structures for wrangling and analysis: climate weather stations record data at fixed point location but may suffer from potential temporal data quality issue such as missing data for certain days.
+GPS data tracks unique point locations at different timestamps and is represented as trajectories.
+Satellite imageries capture snapshots of landscapes at selected time and is commonly structured as raster data.
+The spatio-temporal data cubble address are those collected at unique fixed locations while allowing for irregularity in the temporal dimension, such as the weather station data.
+This corresponds to the full space-time and sparse space-time layouts in the spacetime paper \citep{spacetime}.
+
+The cubble class is an S3 class built on tibble that allows the spatio-temporal data to be wrangled in two forms: a nested/ spatialform and a long/temporal form. It consists of two subclasses:
 
@@
 \item
-  a nested/ spatial cubble has class \texttt{c("spatial\_cubble\_df",\ "cubble\_df")}
+  a nested/ spatial cubble is represented by the class \texttt{c("spatial\_cubble\_df",\ "cubble\_df")}
 \item
-  a long/ temporal cubble has class \texttt{c("temporal\_cubble\_df",\ "cubble\_df")}
+  a long/ temporal cubble is represented by the class \texttt{c("temporal\_cubble\_df",\ "cubble\_df")}
 \end{itemize}
 
-A nested cubble arranges spatial variables in columns and nests temporal variables in a specialised \texttt{ts} column.
-Below prints an nested cubble object using the weather station data collected in three airports (the creation of a cubble object will be explained in Section \ref{create}).
+In a nested cubble, spaital variables are organised as columns and temporal variables are nested within a specialised \texttt{ts} column.
+The nested cubble object printed below contains weather records in three airport stations (the creation of a cubble object will be explained in Section \ref{create}).
 This toy data is a subset of a larger data \code{climate_aus} collected from Global Historical Climatology Network Daily (GHCND).
-The three airport stations in Melbourne are recorded with station metadata: station ID, longitude, latitude, elevation, station name, world meteorology organisation ID.
-The temporal variables are precipitation, maximum and minimum temperature, which can be read from the cubble header:
+It records three airport stations located in Melbourne, Australia and includes spatial variables such as station ID, longitude, latitude, elevation, station name, World Meteorology Organisation ID.
+The dataset contains temporal variables including precipitation, maximum and minimum temperature, which can be read from the cubble header:
 
@@
 
-A cubble object inherits the attributes from tibble (and its subclasses): \texttt{class}, \texttt{row.names}, and \texttt{names}, in addition to three specialised attributes:
+A cubble object inherits the attributes from tibble (and its subclasses): \texttt{class}, \texttt{row.names}, and \texttt{names}. Additionally, it has three specialised attributes:
 
@@
 
-Readers known the \texttt{key} and \texttt{index} attributes from the \texttt{tsibble} package would already be familiar the two arguments.
-In cubble, the \texttt{key} attribute identifies the row in the nested cubble, and together with the \texttt{index} argument, identifies the row in the long cubble.
-Currently, cubble only supports one variable as the key and the accepted temporal class for index includes the base R class \texttt{Date}, \texttt{POSIXlt}, \texttt{POSIXct} and tsibble's \texttt{yearmonth}, \texttt{yearweek}, and \texttt{yearquarter} class.
-
-The \texttt{coords} attribute takes an ordered pair of coordinate.
-It can be a unprojected pair of longitude and latitude, or a projected easting and northing values.
-Under the hood, the \texttt{sf} package is used to calculate the bounding box, shown in the header of a nested cubble, and other spatial operations.
-
-The long cubble has a special attribute \texttt{spatial} to store the spatial variables: all the variables in the nested cubble, except for the \texttt{ts} column.
-The shortcut function are available to extract components in the attributes, for example, \code{spatial()} for extracting the tibble object for the spatial variables:
+Readers who are familiar with the \texttt{key} and \texttt{index} attributes from the \texttt{tsibble} package would already understand the two arguments.
+In cubble, the \texttt{key} attribute identifies the row in the nested cubble, and when combined with the \texttt{index} argument, it identifies the row in the long cubble.
+Currently, cubble only supports one variable as the key, and the accepted temporal classes for index includes the base R classes \texttt{Date}, \texttt{POSIXlt}, \texttt{POSIXct}, as well as tsibble's \texttt{yearmonth}, \texttt{yearweek}, and \texttt{yearquarter} classes.
+
+The \texttt{coords} attribute represents an ordered pair of coordinates.
+It can be either an unprojected pair of longitude and latitude, or a projected easting and northing value.
+The \texttt{sf} package is used under teh hood to calculate the bounding box, displayed in the header of a nested cubble.
+
+The long cubble has a special attribute called \texttt{spatial} to store the spatial variables, which includes all the variables from the nested cubble, except for the \texttt{ts} column.
+Shortcut functions are available to extract attributes, for example, \code{spatial()} for extracting spatial variables from the long cubble:
 
@@
 
-Spatio-temporal data may arrive in separate tables for analysts.
-For example, in climate data, analysts may initially receive station data containing geographic location information, variables recorded and their recording periods.
-They can then query the temporal variables using the stations of interest to obtain the corresponding temporal data.
+In many cases, spatio-temporal data arrive in separate tables for analysis.
+For example, in climate data, analysts may initially receive station data containing geographic location information, recorded variables and their recording periods.
+They can then query the temporal variables using the stations of interest to obtain the relevant temporal data.
 Alternatively, analyses may begin as purely spatial or temporal, and analysts may obtain additional temporal or spatial data to expand the result to spatio-temporal.
 
-The function \code{make_cubble()} composes a \code{cubble} object from a spatial table (\code{spatial}) and a temporal table (\code{temporal}) with the three attributes introduced in the subsection \ref{object}: \code{key}, \code{index}, and \code{coords}.
+The function \code{make_cubble()} composes a \code{cubble} object from a spatial table (\code{spatial}) and a temporal table (\code{temporal}), along with three attributes introduced in the subsection \ref{object}: \code{key}, \code{index}, and \code{coords}.
 The following code creates the nested \code{cubble} object shown previously:
@@
 This includes casting from a \code{tibble} or \code{data.frame} with both spatial and temporal information, a NetCDF object, a \code{stars} object \citep{stars}, and a \code{sftime} object \citep{sftime}.
-The two examples below show the casting from a tibble and a NetCDF object to a cubble object.
-
-The dataset \code{climate_flat} joins the spatial data, \code{stations}, with the temporal data, \code{meteo}, into a single tibble object and it can be coerced into a cubble using:
+The two examples below show the casting from a tibble and a NetCDF object.
+The dataset \code{climate_flat} combines the spatial data, \code{stations}, with the temporal data, \code{meteo}, into a single tibble object.
+It can be coerced into a cubble using:
 
@@
 
-The NetCDF data is another format commonly used for storing spatio-temporal data.
-In \proglang{R}, packages for wrangling NetCDF data include a high-level R interface: \pkg{ncdf4} \citep{ncdf4}, a low-level interface that calls a C-interface: \pkg{RNetCDF} \citep{rnetcdf}, and a tidyverse implementation: \pkg{tidync} \citep{tidync}.
-The code below casts a NetCDF object in the \code{ncdf4} class into a cubble object:
+In \proglang{R}, there are several packages available for wrangling NetCDF data, including \code{ncdf4}, \code{RNetCDF}, and \code{tidync}.
+The code below converts a NetCDF object of class \code{ncdf4} into a cubble object:
 
@@
 
-We would recommend reducing to about 300 \(\times\) 300 grid points for three daily variables in one year.
-A 300 by 300 spatial grid can be a bounding box of {[}100, -80, 180, 0{]} at 0.25 degree resolution or a global bounding box {[}-180, -90, 180, -90{]} at 1 degree resolution.
-
 \hypertarget{functions-and-methods}{%
@@
 
-Table \ref{tab:funs} summarises functions implemented in the \pkg{cubble} package and Table \ref{tab:methods} details the methods implemented for each of the three cubble classes: \code{cubble\_df}, \code{spatial\_cubble\_df}, and \code{temporal\_cubble\_df}.
+Table \ref{tab:funs} summarises the functions implemented in the \pkg{cubble} package and Table \ref{tab:methods} details the methods implemented for each of the three cubble classes.
+The \code{cubble_df} class handles methods that behave consistently in both nested (\code{spatial_cubble_df}) and long cubble (\code{temproal_cubble_df}).
+When the method has distinct behavior for the nested cubble and temporal cubble, it is implemented separately for each subclass.
 
@@
 tidyverse & \texttt{dplyr\_row\_slice}, \texttt{dplyr\_col\_modify}, \texttt{dplyr\_reconstruct}, \texttt{select}, \texttt{mutate}, \texttt{arrange}, \texttt{filter}, \texttt{group\_by}, \texttt{ungroup}, \texttt{summarise}, \texttt{select}, \texttt{slice}, \texttt{rowwise}, \texttt{rename}, \texttt{bind\_rows}, \texttt{bind\_cols}, \texttt{relocate}, \texttt{type\_sum}, the slice family (\texttt{slice\_head}, \texttt{slice\_tail}, \texttt{slice\_max}, \texttt{slice\_min}, \texttt{slice\_sample}) and the join family (\texttt{left\_join}, \texttt{right\_join}, \texttt{inner\_join}, \texttt{full\_join}, \texttt{anti\_join}, \texttt{semi\_join}) \\
-cubble & \texttt{as\_cubble}, \texttt{cubble}, \texttt{make\_cubble}, \texttt{check\_key} \texttt{face\_temporal}, \texttt{face\_spatial}, \texttt{unfold}, \texttt{key}, \texttt{key\_vars}, \texttt{key\_data}, \texttt{index}, \texttt{index\_var}, \texttt{coords}, \texttt{spatial}, \texttt{match\_sites}, \texttt{match\_spatial}, \texttt{match\_temporal}, \texttt{geom\_glyph}, \texttt{geom\_glyph\_box}, \texttt{geom\_glyph\_line}, \texttt{make\_spatial\_sf}, \texttt{make\_temporal\_tsibble}, \texttt{fill\_gaps}, \texttt{scan\_gaps} \\
+cubble & \texttt{as\_cubble}, \texttt{cubble}, \texttt{make\_cubble}, \texttt{check\_key} \texttt{face\_temporal}, \texttt{face\_spatial}, \texttt{unfold}, \texttt{key}, \texttt{key\_vars}, \texttt{key\_data}, \texttt{index}, \texttt{index\_var}, \texttt{coords}, \texttt{spatial}, \texttt{match\_sites}, \texttt{match\_spatial}, \texttt{match\_temporal}, \texttt{geom\_glyph}, \texttt{geom\_glyph\_box}, \texttt{geom\_glyph\_line}, \texttt{make\_spatial\_sf}, \texttt{make\_temporal\_tsibble}, \texttt{fill\_gaps}, and \texttt{scan\_gaps} \\
 \bottomrule()
@@
   >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.7500}}@{}}
-\caption{\label{tab:methods} Methods implemented for the three \code{cubble} classes: \code{cubble\_df}, \code{spatial\_cubble\_df}, and \code{temporal\_cubble\_df}}\tabularnewline
+\caption{\label{tab:methods} Methods implemented for the three \code{cubble} classes.}\tabularnewline
 \toprule()
@@
 
-\hypertarget{dplyr}{%
-\subsubsection{dplyr}\label{dplyr}}
-
-The \pkg{dplyr} package has many tools for wrangling tidy data, many of which are useful in the spatio-temporal analysis.
-The \pkg{cubble} package provides methods that support the use of the following operations in the \pkg{dplyr} package on both the nested and long forms: \texttt{mutate}, \texttt{filter}, \texttt{summarise}, \texttt{select}, \texttt{arrange}, \texttt{rename}, \texttt{left\_join}, and the slice family (\texttt{slice\_*}).
-
-The toy example below demonstrates the using of a collection of dplyr functions (\code{mutate}, \code{select}, \code{filter}, and \code{arrange}) on a nested cubble object \code{cb_nested}:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> cb_nested %>% 
-+   mutate(avg_temp = mean(ts$tmax, na.rm = TRUE)) %>% 
-+   select(-elev) %>% 
-+   filter(avg_temp > 26) %>% 
-+   arrange(-avg_temp)
-\end{CodeInput}
-\end{CodeChunk}
-
-\hypertarget{cubble-1}{%
-\subsubsection{cubble}\label{cubble-1}}
-
-The three basic cubble functions, \code{face_temporal()}, \code{face_spatial()}, and \code{unfold()}, are introduced in this section before more advanced functionalities in Section \ref{others}.
-
-The pair of verbs, \code{face_temporal()} and \code{face_spatial()}, pivot the cubble object between the spatial and temporal face of the multivariate spatio-temporal cube, as illustrated in Figure \ref{fig:face}.
+Specifically, the pair of verbs, \code{face_temporal()} and \code{face_spatial()}, pivot the cubble object between the spatial and temporal face of the multivariate spatio-temporal cube, as illustrated in Figure \ref{fig:face}.
 The code below uses the function \code{face_temporal()} and \code{face_spatial()} to transform between a nested cubble (\code{cb_nested}) and a long one (\code{cb_long}), introduced earlier in subsection \ref{object}:
@@
 
-Sometimes, analysts may need to apply some variable transformation that involves both the spatial and temporal variables.
-An example of this is the transformation of temporal variables into the spatial dimension in glyph maps \citep{Wickham2012-yr}.
-(How to make glyph maps will be explained in Section \ref{st_transformation}, and are illustrated in the second example.) This type of operation can be seen as flattening, or \emph{unfolding}, the cube into a 2D data frame.
-Here the function \code{unfold()} moves the spatial variables \code{long} and \code{lat} into the long \code{cubble}:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> cb_long |> unfold(long, lat)
-\end{CodeInput}
-\begin{CodeOutput}
-# cubble:   key: id [3], index: date, long form
-# temporal: 2020-01-01 -- 2020-01-10 [1D], no gaps
-# spatial:  long [dbl], lat [dbl], elev [dbl], name [chr], wmo_id [dbl]
-  id          date        prcp  tmax  tmin  long   lat
-  <chr>       <date>     <dbl> <dbl> <dbl> <dbl> <dbl>
-1 ASN00086038 2020-01-01     0  26.8  11    145. -37.7
-2 ASN00086038 2020-01-02     0  26.3  12.2  145. -37.7
-3 ASN00086038 2020-01-03     0  34.5  12.7  145. -37.7
-4 ASN00086038 2020-01-04     0  29.3  18.8  145. -37.7
-5 ASN00086038 2020-01-05    18  16.1  12.5  145. -37.7
-# i 25 more rows
-\end{CodeOutput}
-\end{CodeChunk}
-
-\hypertarget{compatibility-with-other-packages}{%
-\subsection{Compatibility with other packages}\label{compatibility-with-other-packages}}
-
-Analysts often have their own preferred spatial or temporal data structure, which they may wish to keep for spatio-temporal analysis.
-For example, the \code{tbl_ts} class from the \code{tsibble} package \citep{tsibble} is commonly used in time series forecasting and similarly, the \code{sf} class \citep{sf} is often used in spatial data science.
-In \code{cubble}, analysts can combine these two structures together by allowing the spatial component to also be an \code{sf} object and the temporal component to also be a \code{tsibble} object.
-
-\hypertarget{tsibble}{%
-\subsubsection{tsibble}\label{tsibble}}
+\hypertarget{compatibility-with-tsibble-and-sf}{%
+\subsection{Compatibility with tsibble and sf}\label{compatibility-with-tsibble-and-sf}}
+
+Analysts often have their preferred spatial or temporal data structure that they prefer to use for spatio-temporal analysis.
+For example, the \code{tbl_ts} class from the \code{tsibble} package \citep{tsibble} is commonly used in time series forecasting and similarly, the \code{sf} class \citep{sf} is frequently used in spatial data science.
+In \code{cubble}, analysts have the flexibility to combine these two structures together by allowing the spatial component to also be an \code{sf} object and the temporal component to also be a \code{tsibble} object.
+
+\hypertarget{using-a-tsibble-for-the-temporal-component}{%
+\subsubsection{Using a tsibble for the temporal component}\label{using-a-tsibble-for-the-temporal-component}}
 
@@
 
-\hypertarget{sf}{%
-\subsubsection{sf}\label{sf}}
+\hypertarget{using-sf-for-the-spatial-component}{%
+\subsubsection{Using sf for the spatial component}\label{using-sf-for-the-spatial-component}}
 
@@
 \begin{CodeInput}
-R> sf_nested <- cb_nested %>% make_spatial_sf() 
-R> all.equal(sf_nested, sf_nested)
-\end{CodeInput}
-\begin{CodeOutput}
-[1] TRUE
+R> cb_nested %>% make_spatial_sf() 
+\end{CodeInput}
+\begin{CodeOutput}
+# cubble:   key: id [3], index: date, nested form, [sf]
+# spatial:  [144.8321, -37.98, 145.0964, -37.6655], WGS 84
+# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
+  id           long   lat  elev name   wmo_id ts                  geometry
+  <chr>       <dbl> <dbl> <dbl> <chr>   <dbl> <list>           <POINT [°]>
+1 ASN00086038  145. -37.7  78.4 essen~  95866 <tibble> (144.9066 -37.7276)
+2 ASN00086077  145. -38.0  12.1 moora~  94870 <tibble>   (145.0964 -37.98)
+3 ASN00086282  145. -37.7 113.  melbo~  94866 <tibble> (144.8321 -37.6655)
 \end{CodeOutput}
@@
 
-Given the vast spatio-temporal data structure available in \proglang{R} to represent spatio-temporal data, this section compares and contrast the \pkg{cubble} with other existing alternative, namely \pkg{stars} and \pkg{sftime}.
-
-The \pkg{stars} package \citep{stars} uses an array structure, as oppose to tibble in the \code{cubble} package, to represent multivariate spatio-temporal data.
-While both packages support vector and raster data, it is a matter of choice on which structure to use given the application.
-Analysts working on satellite imageries may prefer the array structure while others originally working with spatio-temporal data in the 2D data frames may find \pkg{cubble} easier to adopt from their existing computing workflow.
-
-The \pkg{sftime} package \citep{sftime} also builds from a tibble object, with a main focus to handle irregular spatio-temporal data.
-While \pkg{sftime} can also handle full space-time grids and sparse space-time layouts, \pkg{cubble} can be thought of as simplifying the repetition in spatial variables with two linked tibbles.
-This simplification gives memory efficiency when data is observed frequent, i.e.~daily or subdaily, or the spatial geometry is expensive to repeat, i.e.~polygons or multipolygons.
+In \proglang{R}, there are other existing spatio-temporal data structure and this section compares and contrasts \pkg{cubble} with other existing alternative, specifically \pkg{stars} and \pkg{sftime}.
+The \pkg{stars} package \citep{stars} uses an array structure, as oppose to tibble, to represent multivariate spatio-temporal data.
+While both \pkg{stars} and \pkg{cubble} support vector and raster data, it is a matter of choice on which structure to use given the application.
+Analysts working on satellite imageries may prefer the array structure in \pkg{stars}, while others originally working with spatio-temporal data in 2D data frames may find \pkg{cubble} easier to adopt from their existing computing workflow.
+
+The \pkg{sftime} package \citep{sftime} also builds from a tibble object and its focus is on handling irregular spatio-temporal data.
+This means \pkg{sftime} can also handle full space-time grids and sparse space-time layouts represented in \pkg{cubble}, but \pkg{cubble} uses nesting to avoid storing spatial variables repetitively at each timestamp.
+This provides memory efficiency when data is observed frequent, i.e.~daily or sub-daily, or the spatial geometry is expensive to repeat, i.e.~polygons or multipolygons.
 Consider the \code{climate_aus} data in the \pkg{cubble} package with 639 stations observed daily in a single year 2020.
@@
 
-One task that may interest spatio-temporal analysts is combining data collected at nearby but not exactly the same sites, for example, weather station measured rainfall and river levels.
-This can be considered to be a matching problem \citep{stuart2010matching, mcintosh2018using} to pair similar time series from nearby locations, or even a data fusion exercise that merges data collected from different sources \citep{cocchi2019data}.
-
-In cubble, data can be matched spatially or temporarily with \texttt{match\_spatial()} and \texttt{match\_temporal()}.
-The function \texttt{match\_spatial()} calculates the spatial distance between observations in two cubbles.
-Different distances are available with projected or unprojected coordinate reference system.
-Analysts can subset the number of matched group to output with argument \texttt{spatial\_n\_group} (by default 4 groups) and the number of match per group with argument \texttt{spatial\_n\_group} (by default 1, that is, one-to-one matching).
+One common task when working with spatio-temporal data is to match nearby sites.
+For example, we may want to verify the location of an old list of stations with current stations, or we may want to match the data from different data sources.
+In cubble, spatial and temporal matching are performed using the functions \texttt{match\_spatial()} and \texttt{match\_temporal()}.
+The \texttt{match\_spatial()} function calculates the spatial distance between observations in two cubble objects.
+Various distance measures are available (check \texttt{sf::st\_distance}).
+Analysts can specify the number of matched groups to output using the \texttt{spatial\_n\_group} argument (default to 4 groups) and the number of matches per group using the \texttt{spatial\_n\_group}argument (default to 1, one-to-one matching).
 The syntax to use \texttt{match\_spatial()} is:
@@
 
-The function \texttt{match\_temporal()} calculates the time series similarity between spatially matched groups.
-Two identifiers needs to be specified on the variable separates the each matched group (\texttt{match\_id}) and the variable separates the two sources (\texttt{data\_id}).
-The argument \texttt{temporal\_by} uses the \texttt{by} syntax from dplyr \texttt{*\_join} to specify the temporal matching variable.
-
-The similarity score between two time series in the spatially matched group is calculated by a matching function, which analysts can customise.
-The matching function should take two time series in a list and output a single numerical score, which allows for interfacing with existing time series distance calculation implementation.
+The function \texttt{match\_temporal()} calculates the similarity between time series within spatially matched groups.
+Two identifiers are required: one for separating each matched group (\texttt{match\_id}) and one for separating the two data sources (\texttt{data\_id}).
+The argument \texttt{temporal\_by} uses the \texttt{by} syntax from dplyr's \texttt{*\_join} to specify the temporal variables to match.
+
+The similarity score between two time series is calculated using a matching function, which can be customised by the analysts.
+The matching function takes two time series as a list and returns a single numerical score.
+This allows for flexibility in using existing time series distance calculation implementation.
 By default, cubble implements a simple peak matching algorithm (\texttt{match\_peak}) that counts the number of peaks in two time series that fall within a specified temporal window.
-The syntax to use \texttt{match\_temporal()} is:
+The syntax to use \texttt{match\_temporal()} is
 
@@
 
-Spatio-temporal data lends itself to a range of transformations.
-Glyph maps \citep{Wickham2012-yr} transform the measured variable and time coordinates into microplots at the spatial locations.
-Calendar plots \citep{wang2020calendar} deconstruct time to produce plots of variables in a calendar format.
-Summarizing multiple variables is commonly done using projections, or linear combinations.
-Here we elaborate on the transformations made to produce a glyph map.
-The package \pkg{GGally} \citep{ggally} has implemented glyph maps through the \code{glyphs()} function.
-The function constructs a \code{data.frame} with calculated position (\code{gx}, \code{gy}, \code{gid}) of each point on the time series using linear algebra (Equations 1 and 2 in \citet{Wickham2012-yr}).
-A new implementation of the glyph map as a ggproto, \code{GeomGlyph}, has been made in the \pkg{cubble} package so that a glyph map can be created with \code{geom_glyph()}:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> data |>
-+   ggplot() +
-+   geom_glyph(aes(x_major = ..., x_minor = ...,
-+                  y_major = ..., y_minor = ...))
-\end{CodeInput}
-\end{CodeChunk}
-
-Some useful controls over the glyph map are also available in the \code{geom_glyph()} implementation
+Sometimes, we wish to communicate spatial and temporal information collectively through visualisation.
+This can be achieved through several graphical displays: one can make faceted maps across time, creating map animations, or constructing interactive graphics to link between map and time series plot.
+The concept of glyph maps was initially proposed in \citet{Wickham2012-yr}.
+The underlying idea is to transform the temporal coordinates into spatial coordinates so that time series plot can be displayed on the map.
+
+The package \texttt{GGally} initially implement the glyph map.
+It uses \texttt{glyphs()} to calculate the axis transformation and then uses \texttt{geom\_polygon()} to draw the map.
+In cubble, a ggproto implementation \texttt{geom\_glyph()} performs the linear algebra internally as data transformation .
+The \texttt{geom\_glyph()} requires four aesthetics: \texttt{x\_major}, \texttt{y\_major}, \texttt{x\_minor}, and \texttt{y\_minor}.
+The major axes are the outer spatial coordinates and the minor axes are the inner/ temporal coordinates:
+
+\begin{verbatim}
+data |> 
+  ggplot() +
+  geom_glyph(aes(x_major = ..., x_minor = ..., 
+                 y_major = ..., y_minor = ...))
+\end{verbatim}
+
+Some useful controls over the glyph map includes:
 
@@
 \item
-  polar coordinate glyph maps are specified using \code{polar = TRUE},
+  polar coordinate glyph maps with \code{polar = TRUE},
 \item
-  arguments \code{width} and \code{height} can be specified in either absolute or relative value,
+  adjust glyph size with arguments \code{width} and \code{height},
 \item
-  global and local scale is specified with \code{global_rescale}, which defaults to \code{TRUE}, and
-  *reference boxes and lines can be added with geoms: \code{geom_glyph_box()} and \code{geom_glyph_line()}.
+  glyph scale to fixed (\code{global_rescale} defaults to \code{TRUE}) or free, and
+\item
+  reference boxes and lines with \code{geom_glyph_box()} and \code{geom_glyph_line()}.
 \end{itemize}
@@
 The five examples here are chosen to illustrate these aspects of the \pkg{cubble} package: creating a \code{cubble} object from two Coronavirus (COVID) data tables with the complication of differing location names, using spatial transformations to make a glyph map of seasonal temperature changes over years, matching river level data and weather station records for analysis of water supply, reading NetCDF format data to reproduce a climate reanalysis plot, and the workflow to create complex interactive linked plots.
-(The Appendix provides another example on aggregating information spatially to explore precipitation patterns.)
 
@@
 
-Bureau of Meteorology collects \href{http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full}{water data} from river gauges and this includes variables: electrical conductivity, turbidity, water course discharge, water course level, and water temperature.
-In particular, water level will interactive with precipitation from the climate data since rainfall will raise the water level in the river.
-Both \texttt{climate\_vic} and \texttt{river} are cubble objects and we can get a summary of the 10 closest pairs:
+One common task when working with spatio-temporal data is to match nearby sites.
+For example, we may want to verify the location of an old list of stations with current stations, or we may want to match the data from different data sources.
+In this example, we will introduce the spatial and temporal matching in cubble using an example on matching river level data with precipitation in Victoria, Australia.
+
+The water level data collected by the Bureau of Meteorology, can be compared with the precipitation since rainfall can directly impact water level in river.
+Both \texttt{climate\_vic} and \texttt{river} are cubble objects, and we can obtain a summary of the 10 closest pairs between them:
 
@@
 
-The result can also be returned as cubble objects with argument \texttt{return\_cubble\ =\ TRUE}.
-The output is a list where each element is a paired cubble object and you may consider combining all the results into a single cubble with \texttt{bind\_rows()}.
-Care needs to be taken on in the case when a site is close to two stations since by construction, cubble require unique rows in the nested form.
-From the summary table above, river station \texttt{226027} is matched to more than one weather station: \texttt{ASN00085072} (group 3) and \texttt{ASN00085298} (group 5).
-(Similarly river station \texttt{230200} is matched in group 7 and 8).
-One can either deselect one pair before binding the results, or take the list and work with the \texttt{purrr::map} syntax:
+The result can also be returned as cubble objects by setting the argument \texttt{return\_cubble\ =\ TRUE}.
+The output is be a list where each element is a paired cubble object.
+To combine all the results into a single cubble, you can use \texttt{bind\_rows()}.
+In the case when a site in the second cubble (the \texttt{river} data here) is matched to two stations in the first cubble (\texttt{climate\_vic} here), the binding may not be successful since cubble requires unique rows in the nested form.
+In the summary table above, the river station \texttt{226027} is matched to more than one weather station: \texttt{ASN00085072} (group 3) and \texttt{ASN00085298} (group 5).
+Similarly, the river station \texttt{230200} is matched in group 7 and 8).
+In such cases, you can either deselect one pair before combining, or work with the list output with the \texttt{purrr::map} syntax:
 
@@
 
-For temporal matching, the variable water level (\texttt{Water\_course\_level}) from the river data will be matched to precipitation (\texttt{prcp}) in the weather station data.
-The variable identifying each matched group is \texttt{group} and the variable identifying the two datasets is \texttt{type}, and similarly, the cubble output can be returned with the argument \texttt{return\_cubble\ =\ TRUE}.
-Here we select the four pairs with the highest number of matching peaks:
+For temporal matching, we match the variable \texttt{Water\_course\_level} from the river data to \texttt{prcp} in the weather station data.
+The variable \texttt{group} and \texttt{types} identify the matching group and the two datasets:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> (res_tm <- res_sp %>% 
++   match_temporal(
++     data_id = type, match_id = group,
++     temporal_by = c("prcp" = "Water_course_level")))
+\end{CodeInput}
+\begin{CodeOutput}
+# A tibble: 8 x 2
+  group match_res
+  <int>     <dbl>
+1     1        30
+2     2         5
+3     3        14
+4     4        20
+5     6        23
+# i 3 more rows
+\end{CodeOutput}
+\end{CodeChunk}
+
+Similarly, the cubble output can be returned using the argument \texttt{return\_cubble\ =\ TRUE}. Here we select the four pairs with the highest number of matching peaks:
 

### 2023-06-18 review/B-response.Rmd
go through the response to reviewers

--- 2023-05-31 review/B-response.Rmd
@@
 author: "H. Sherry Zhang, Dianne Cook, Ursula Laa, Nicolas Langrené, Patricia Menéndez"
-date: '2023-03-04'
+date: "`r Sys.Date()`"
 output: pdf_document
@@
 
+We provide an overall summary of the changes made to the paper and the package before addressing reviewers' comments point-by-point.
+
+# Summary of Main Changes
+
+## The Paper:
+
+Section 2 have been reworked to combine the original section 2 and 3.
+It now focuses more on the inner work of the package: the cubble class and its attributes, object creation and coercion, functions and methods implemented, compatibility with `tsibble` and `sf`, and comparison with other existing spatio-temporal packages (`stars` and `sftime`).
+
+In the comparison section (Section 2.5), we compare `cubble` with `stars` and `sftime`.
+With `stars`, we believe spatio-temporal analysts who are more comfortable processing data with data frame (tidyverse) will benefit from cubble, given `stars` is built from arrays.
+With `sftime`, we argue cubble is more efficient in memory since it doesn't repeat spatial variables at each time stamp.
+We give an example to compare the object size of the same data stored as an `sftime` object and a `cubble` object (118MB vs.
+9MB for 639 climate weather stations observed daily in 2020).
+
+Other sections of the paper have been edited for conciseness and they match with the package vignettes.
+
+## The Package:
+
+The cubble class has been refactored as suggested by reviewer 1. Now the cubble class has two sub-classes:
+
+  * nested/ spatial cubble: `c("spatial_cubble_df", "cubble_df")`, and
+  * long/ temporal cubble: `c("temporal_cubble_df", "cubble_df")`.
+
+All the S3 methods have been modified accordingly.
+Methods behave consistently in the nested and long cubble are implemented in the `cubble_df` class, otherwise, separately in the `spatial_cubble_df` and `temporal_cubble_df` class.
+
+The package documentation have been substantially improved, specifically, [...]. A new version of the package (v0.3.0) is now available on CRAN.
+
+In the point-by-point response to reviewers' comments below, the original comments are in black and our responses are in red.
+
 # Response to B-review_1_1
@@
 ## The Paper
+
+The paper describes a new data structure cubble and associated workflows. In a nutshell cubble subsumes two data structures:
+
+  * nested which is indexed by spatial location and keeps the time-series data in a list column 
+  long which is indexed by time and keeps the spatial data in a hidden “spatial” attribute.
+
+The functionality of the package consists of
+
+  * data transformation functions: face_temporal, face_spatial and unfold
+  * data matching (merging) functions: match_sites, match_spatial and match_temporal
 
@@
 
-**\textcolor{purple}{Re: We thank the reviewer for the support on our proposed data structure and the suggestion to improve the paper and the package documentation.}**
+**\textcolor{purple}{We thank the reviewer for the support on our proposed data structure.
+We have reworked on section 2 to describe the class and its associated functionality.
+Additionally, the text in the paper has been aligned with the package vignette.
+}**
 
@@
 
-**\textcolor{purple}{Re: We select the code and output relevant to demonstrate the cubble package in the examples while leaving the rest in the github repository, where users can look up for more details: https://github.com/huizezhang-sherry/paper-cubble.
-[on the visualisation packages: xxx].
-After addressing all the issues raised by the reviewers, the paper is now xxx pages, including 4 pages of reference and required affiliation.}**
+**\textcolor{purple}{We have selected and displayed the codes and outputs relevant to demonstrate the cubble functionality in the examples.
+The full script is available in the reproducible submitted reproducible script or the paper GitHub repository: https://github.com/huizezhang-sherry/paper-cubble.}**
 
@@
 A more natural implementation would be start with two distinct data structures, both sub-classes of the abstract cubble_df class, c("spatial_cubble_df", "cubble_df") and c("spatial_temporal_df", "cubble_df").
-Such a generic implementation would allow for natural extension of the functionality trough dedicated methods.
-
-**\textcolor{purple}{Re: We thanks reviewer for the suggestion on the refactorand now: }**
-
-  * **\textcolor{purple}{a nested/spatial cubble has class c("spatial\_cubble\_df", "cubble\_df")}**
-  * **\textcolor{purple}{a long/temporal cubble has class c("temporal\_cubble\_df", "cubble\_df")}**
-
-**\textcolor{purple}{As a result, S3 methods are now written separately for the two classes unless it doesn't differentiate between the form. For example:}**
-
-  * **\textcolor{purple}{face\_temporal() will do different things depending on whether it is a nested/long cubble:}**
-
-```{r eval = FALSE}
-face_temporal <- function(...){
-  UseMethod("face_temporal")
-}
-
-face_temporal.spatial_cubble_df <- function(...){
-
-}
-
-face_temporal.temporal_cubble_df <- function(...){
-  # if already in the temporal face, return the cubble as it is
-}
-```
-
-  * **\textcolor{purple}{key\_vars() will anyway extract the key variable of a cubble object despite of the form:}**
-
-```{r eval = FALSE}
-key_vars.cubble_df <- function(){
-  ...
-}
-```
-
+Such a generic implementation would allow for natural extension of the functionality through dedicated methods.
 Currently an ever-present check on “form” attribute is necessary throughout the code base.
@@
 
-**\textcolor{purple}{Re: as\_cubble() now can coerce four types of foreign objects into a cubble object: 1) a single joint table, 2) NetCDF data, 3) a stars object, 4) a sftime object.
-The code related to output matching pairs has been moved to a separate function check\_key() and reflected in the current example 5.1}**
+**\textcolor{purple}{The suggested class implemented has been adopted, see The Package section in the Summary of Main Changes.
+The codes on checking key values from multiple tables have been moved to a separate function check\_key() and reflected in the example 5.1 now.
+The function as\_cubble() now returns nested cubbles.}**
 
@@
 
-**\textcolor{purple}{Re: The naming of long/nested comes from the tidy data concept where the "spatial" form has the nested temporal information and the "temporal" form uses the long form data from the tidy data.
-We understand it could be more intuitive to refer to them as spaital/ temporal and we add them as a "common name" to the form in the definitional section (now section 2).
-This naming is also used in the paper when it is more straightforward to use spatial/temporal than nested/long.}**
+**\textcolor{purple}{The naming of long/nested comes from the tidy data concept where the spatial cubble has the nested temporal information and the temporal cubble uses the long form data from the tidy data.
+We understand it could be more intuitive to refer to them as spatial/temporal cubble and we adopt the spatial/temporal cubble when it is more straightforward to use.}**
 
+As cubble is a new package and no reverse dependencies yet exist, I would suggest rewriting the class dispatch mechanism before pursuing with the paper publication.
 My recommendation would be a combination of the following:
@@
   * Rework the definitional parts (sections 2 and 3) by following more closely the “design” vignette.
-**\textcolor{purple}{Re: Section 2 and 3 have now been combined into a section 2 and the following subsections are includeded: }**
+**\textcolor{purple}{Adopted, see The Paper section in the Summary of Main Changes of this response}**
 
-      > **\textcolor{purple}{2.1 introduce the cubble class}**
-
-      > **\textcolor{purple}{2.2 create and coerce foreign objects into a cubble object}**
-
-      > **\textcolor{purple}{2.3 functions and methods in cubble (dplyr and cubble functions)}**
-
-      > **\textcolor{purple}{2.4 compatibility with tsibble and sf}**
-
-      > **\textcolor{purple}{2.5 comparision with other existing spatio-temporal class: stars and sftime}**
-
-  * Describe at a glance the core functions which operate on the class and their motivation. **\textcolor{purple}{Re: A method at glance has been printed at the beginning of subsection 2.3 }**
-  * Rework section 4 by making it more concise and provide links to sections in supplementary material or dedicated vignettes. **\textcolor{purple}{Re: }**
-  * Rework examples section 5 into “Applications”. Briefly describe the applications and provide and refer to dedicated and self-contained vignettes. **\textcolor{purple}{Re: xxx}**
+  * Describe at a glance the core functions which operate on the class and their motivation. **\textcolor{purple}{Functions and methods at a glance is now available at the start of Section 2.3.}**
+  * Rework section 4 by making it more concise and provide links to sections in supplementary material or dedicated vignettes. **\textcolor{purple}{Reworked. Text has been edited for conciseness}**
+  * Rework examples section 5 into “Applications”.
+Briefly describe the applications and provide and refer to dedicated and self-contained vignettes.
+**\textcolor{purple}{Adopted, see the last point.}**
 
@@
 
-**\textcolor{purple}{Re: Documentation of the package has been improved with a focus on clarifying the input arguemnt types, providing examples, [more] }**
+````
+    data: the object to be created or tested as cubble
+     key: the spatial identifier
+   index: the time identifier
+  coords: the coordinates that characterise the spatial dimension
+     ...: a list object to create new cubble
+````
 
   * What objects are supported as input data? 
-**\textcolor{purple}{Re: the objects supported are 1) list with two elements named spatial and temporal, 2) data frame/ tibble, 3) netcdf objects, 4) stars objects, and 5) sftime objects.
-}**
-* What is the assumption of the “nestedness” of the input data. **\textcolor{purple}{Re: }**
+  * What is the assumption of the “nestedness” of the input data. 
   * What is the accepted data type of key, index and coords?
-**\textcolor{purple}{Re: Key supports character and factor.
-Index supports base R classes Date, POSIXlt, POSIXct and tsibble's yearmonth, yearweek, and yearquarter class.
-Coords support both projected/ unprojected coordinates in columns (can be omitted if created from sf and its subclass).
-}**
+
+The user would need to visit the documentation of tsibble in order to understand the meaning of the key and index. But the relationship to tsibble is not even mentioned in the docs.
+The documentation of the key and index is not perfect in tsible either, but sheds some more light on their role:
+
+````
+     key: Variable(s) that uniquely determine time indices. ‘NULL’ for
+          empty key, and ‘c()’ for multiple variables. It works with
+          tidy selector (e.g. ‘dplyr::starts_with()’).
+   index: A variable to specify the time index variable.
+
+````
+
+The doc states:
+
+````
+    The constructor for the cubble class
+````
+
+If it’s the constructor then the convention is to name it cubble().
+as_xyz is the converter which translates between different data types.
+For example, if the proper sub-classing would be used, as_spatial_cubble and as_temporal_cubble would be more standard names rather than the face_spatial and face_temporal respectively.
+
+**\textcolor{purple}{Documentation of the package has been substantially improved with a focus on clarifying the input arguemnt types.
+For the specific example of as\_cubble(), the mentioned arguments now state:  }**
+
+````
+    data: an object to be converted into an cubble object. 
+          Currently support objects of classes `tibble`, `ncdf4`, `stars`, and `sftime`. 
+     key: a character (or symbol), the spatial identifier.
+   index: a character (or symbol), the temporal identifier. Currently support base R 
+          classes Date, POSIXlt, POSIXct and tsibble's yearmonth, yearweek, and yearquarter
+          class.
+  coords: a vector of character (or symbol) of length two, in the order of longitude
+          first and then latitude, the argument can be omitted if created from an sf 
+          and its subclasses. In case the sf geometry column is not POINT, coords will 
+          be the centroid coordinates.
+
+````
+
+**\textcolor{purple}{TODO: more here}**
 
@@
 
-**\textcolor{purple}{Re: Adopted.
+**\textcolor{purple}{Adopted.
 Now when applying face\_spatial to a nested cubble, the cubble will be printed as it is with a message signalling it is already in the nested form.
 see the reprex below.
-Same for face\_temporal on a long cubble object}**
+Same for face\_temporal on a long cubble object.}**
 
@@
 library(cubble)
-nested <- climate_flat |>
-  as_cubble(key = id, index = date, coords = c(long, lat))
-class(nested) 
-#> [1] "cubble_df"  "rowwise_df" "tbl_df"     "tbl"        "data.frame"
-face_spatial(nested) 
-#> i The cubble is already in the nested form
-#> # cubble:   id [5]: nested form
-#> # bbox:     [115.9764, -32.9342, 133.5407, -12.4239]
-#> # temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
-#>   id            lat  long  elev name           wmo_id ts                
-#>   <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>            
-#> 1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tibble [366 × 4]>
-#> 2 ASN00010311 -31.9  117. 179   york            94623 <tibble [366 × 4]>
-#> 3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble [366 × 4]>
-#> 4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble [366 × 4]>
-#> 5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble [366 × 4]>
+class(climate_mel)
+#> [1] "spatial_cubble_df" "cubble_df"         "tbl_df"           
+#> [4] "tbl"               "data.frame"
+invisible(face_spatial(climate_mel))
+#> The cubble is already in the nested form
 ```
 
-<sup>Created on 2023-05-22 with [reprex v2.0.2](https://reprex.tidyverse.org)</sup>
+<sup>Created on 2023-06-18 with [reprex v2.0.2](https://reprex.tidyverse.org)</sup>
 
@@
 
-**\textcolor{purple}{Re: the spatial attribute can be checked with function spaital(<cubbble\_obj>)}**
+**\textcolor{purple}{the spatial attribute can be checked with function spaital(<cubbble\_obj>)}**
 
@@
 
-**\textcolor{purple}{Re: xxx}**
+**\textcolor{purple}{TODO}**
 
@@
 Such comments could be safely removed.
-**\textcolor{purple}{Re: sentence removed.}**
+**\textcolor{purple}{sentence removed.}**
 
@@
 Isn't key a spatial identifier and index a temporal one?
-**\textcolor{purple}{Re: [TODO] Yes, key is the spatial identifier and index is the temporal one.
-They are now explained in the text as well as referenced to the tsibble package.}**
+**\textcolor{purple}{The `key` and `index`, along with other attributes are explained under heading *The cubble attributes* in Section 2.1.}**
 
   - on p5 section 3.1, the word "wording" is highlighted in the sentence "The arguments key and index follow the wording in the tsibble package to ...", conventions.
-**\textcolor{purple}{Re: *wording* changed to *conventions*}**
+**\textcolor{purple}{The setence has been rewrite}**
 
   - on p6 section 3.2, 'The cubble class also provides a long form, which expands the ts column and temporarily “hides” the spatial variables.', confusing: `hides` means storing it as an attribute.
-**\textcolor{purple}{Re: The sentence is now changed to ..., which expands the ts column and stores the spatial variables as an attribute}**
+**\textcolor{purple}{The sentence is now on p4 section 2.1 before the printing of object cb\_long as "In a long cubble, temporal variables are expanded in the long form and spatial variables are stored as a data attribute."}**
 
@@
 By this point it's still not clear what those are and what the meaning of key and index is.
-**\textcolor{purple}{Re: addressed in bullet point "on p5 section 3.1"}**
+**\textcolor{purple}{The description of the two data structures is now at section 2.1}**
 
   - on p6, section 3.2, yet another reason to call this data structure "cubble_temporal" instead of the semantically unobvious "long".
-**\textcolor{purple}{Re: see paragraph 4 under section The paper }**
+**\textcolor{purple}{see paragraph 4 under section The paper in this response.}**
 
@@
 That is, the spatial component cannot be represented as part of the nested tibble without repetition.
-**\textcolor{purple}{Re: this has now been incorporated in subsection 2.5 to compare with a sftime object}**
+**\textcolor{purple}{this has now been incorporated in section 2.5.}**
 
   - on p8 section 3.6, The missing elephant is tidyr which has nesting and unnesting operations  https://tidyr.tidyverse.org/articles/nest.html.
-**\textcolor{purple}{Re: addressed under section Further suggestion in this response}**
+**\textcolor{purple}{addressed in Further suggestion in this response}**
 
@@
 
-**\textcolor{purple}{Re: The comparison with other existing spatio-temporal class, i.e.
-stars and sftime, has been addressed in subsection 2.5 (after combining the original section 2 and 3 into section 2)}**
+**\textcolor{purple}{A new section, Section 2.5, has been dedicated to compare cubble with existing spatio-temporal class, specifically, stars and sftime.
+See The Paper section in the Summary of Main Changes of this response}.**
 
@@
 
-**\textcolor{purple}{Re: We understand spatial data is, strictly speaking, not uniquely identified if CRS information is not provided, while many wild data still do not contain this crucial piece of information.
-In cubble, if the spatial data is from an sf object, the cubble will retain the sf class and users are welcomed to use functions in sf to perform coordinate transformations.
-If not, the function cubble::add\_geometry\_column() can be used to construct the geometry column from long and lat, and the resulting cubble will have the sf class.
-if CRS is missing, cubble::add\_geometry\_column() will by default assume OGC:CRS84 (WGS84).}**
+**\textcolor{purple}{The coordinate reference system is handled, under the hood, by the sf::st\_tranform().
+In the case where a cubble object is created from an sf object, the sf class will be retained.
+When it is not, users can promote the cubble object to inlcude include the sf class with make\_spatial\_sf(): }**
 
@@
 
-**\textcolor{purple}{Re: xxx}**
+**\textcolor{purple}{In spatial matching, several distance calculations are now made available under sf::st\_distance.
+Temporal matching now accepts an argument temporal\_match\_fn to allow user specified functions for matching.}**
 
@@
 - doubled/missing words/none correct sentences: 
-  - “*components* spatio-temporal *components*” **\textcolor{purple}{Re: to "... spatio-temporal components"}**
-  - “... fits works ...”  **\textcolor{purple}{Re: to ".. works well with ..."}**
-  - “... be activate rows ...” **\textcolor{purple}{Re: to "... will activate rows ..."}**
-  - “... highlighted *in* the ...” **\textcolor{purple}{Re: to "... highlighted on the map"}**
-  - “. . . using 2020 measurements using match_sites() function.”  **\textcolor{purple}{Re: to "... using the match\_sites() function"}**
-  - “An example of this using is included in the Appendix” **\textcolor{purple}{Re: to "... this use ..."}**
-  - “. . . the data in -a- multiple -of- ways on-the-fly”  **\textcolor{purple}{Re: to "... in multiple ways ..."}**
+  - “*components* spatio-temporal *components*” **\textcolor{purple}{to "... spatio-temporal components"}**
+  - “... fits works ...”  **\textcolor{purple}{to ".. works well with ..."}**
+  - “... be activate rows ...” **\textcolor{purple}{to "... will activate rows ..."}**
+  - “... highlighted *in* the ...” **\textcolor{purple}{to "... highlighted on the map"}**
+  - “. . . using 2020 measurements using match_sites() function.”  **\textcolor{purple}{to "... using the match\_sites() function"}**
+  - “An example of this using is included in the Appendix” **\textcolor{purple}{to "... this use ..."}**
+  - “. . . the data in -a- multiple -of- ways on-the-fly”  **\textcolor{purple}{to "... in multiple ways ..."}**
 
@@
 glyph maps will be explained in Section 4.4”.
-**\textcolor{purple}{Re: the looped reference in section 3.3 and 3.4 is replaced with (Wickham et al.
+**\textcolor{purple}{the looped reference in section 3.3 and 3.4 is replaced with (Wickham et al.
 2012)}.**
 
-- Should “. . . it’s temperatures are more *consistent*” rather be “. . . it’s temperatures are more *constant*”? **\textcolor{purple}{Re: to "more constant"}**
+- Should “. . . it’s temperatures are more *consistent*” rather be “. . . it’s temperatures are more *constant*”? **\textcolor{purple}{to "more constant"}**
 
 - Typos:
-  - “. . . spatial and *tmeporal* information are available.” **\textcolor{purple}{Re: to "temporal"}**
+  - “. . . spatial and *tmeporal* information are available.” **\textcolor{purple}{to "temporal"}**
   - The polar vortex, signalled by the high *speicfic* humidity, splits into two on 2002-09-26 and *further-s- split_s_* into four on 2002-10-04.
-**\textcolor{purple}{Re: to "specific", "further splits"}**
-  - the data in *-a- multiple -of-* ways on-the-fly. **\textcolor{purple}{Re:  to "multiple ways"}**
+**\textcolor{purple}{to "specific", "further splits"}**
+  - the data in *-a- multiple -of-* ways on-the-fly. **\textcolor{purple}{ to "multiple ways"}**
 
@@
 
-**\textcolor{purple}{Re: Thanks reviewers for the suggestion and this has now been incorporated in the package.
-The second line in the header, when printed in the long (temporal) form will now show the start and the end date, the date interval, and whether there are gaps in the data.
-See the reprex example below: }**
+**\textcolor{purple}{For temporal cubble, the header now prints the temporal meta data: start date, end date, interval, and whether there is gaps in the data. }**
 
-``` r
-library(cubble)
-face_temporal(climate_aus)
-#> # cubble:  key: id [639], index: date, long form
-#> # extent:  2020-01-01 -- 2020-12-31 [1D], has gaps!
-#> # spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]
-#>    id          date        prcp  tmax  tmin
-#>    <chr>       <date>     <dbl> <dbl> <dbl>
-#>  1 ASN00001006 2020-01-01   164  38.3  25.3
-#>  2 ASN00001006 2020-01-02     0  40.6  30.5
-#>  3 ASN00001006 2020-01-03    16  39.7  27.2
-#>  4 ASN00001006 2020-01-04     0  38.2  27.3
-#>  5 ASN00001006 2020-01-05     2  39.3  26.7
-#>  6 ASN00001006 2020-01-06    60  32.9  25.6
-#>  7 ASN00001006 2020-01-07   146  34.1  25.5
-#>  8 ASN00001006 2020-01-08    40  36.6  26.2
-#>  9 ASN00001006 2020-01-09     0  38.2  27.6
-#> 10 ASN00001006 2020-01-10     0  38.9  29.7
-#> # … with 230,750 more rows
-```
-
-<sup>Created on 2023-05-23 with [reprex v2.0.2](https://reprex.tidyverse.org)</sup>
+````
+#> # cubble:   key: id [3], index: date, long form
+#> # temporal: 2020-01-01 -- 2020-01-10 [1D], no gaps
+#> # spatial:  long [dbl], lat [dbl], elev [dbl], name [chr], wmo_id [dbl]
+````
 
@@
 
-**\textcolor{purple}{Re: [TODO] The two datasets are related but not the same.
-The data historical\_tmax contains daily weather stations data for the two periods: 1971-1975 and 2016-2020 and amounts to 3.2MB, which is not suitable to include in the package given its size.
-The tmax\_hist summarises the daily data into monthly to produce the dataset suitable to include in the package.
-The documentation to tmax\_hist now include more details on this.
-}**
+**\textcolor{purple}{The data historical\_tmax contains daily weather stations data for the two periods: 1971-1975 and 2016-2020 and amounts to 3.2MB, which is not suitable to include in the package given its size.
+The tmax\_hist data, which was included in the package, is now removed, given its similarity to the climate\_aus data, which is more useful.
+We provide the script to generate historical\_tmax in the Appendix and a copy of the data is available in the paper GitHub repository: https://github.com/huizezhang-sherry/paper-cubble.}**
 
@@
 
-**\textcolor{purple}{Re: A result file in html has been submitted this time.}**
+**\textcolor{purple}{A result file in html is submitted this time for result comparison.}**

### 2023-06-18 review/B-response.Rmd
response to reviewers done

@@
 
-The cubble class has been refactored as suggested by reviewer 1. Now the cubble class has two sub-classes:
+The cubble class has been re-factored as suggested by reviewer 1. Now the cubble class has two sub-classes:
 
@@
 
-The package documentation have been substantially improved, specifically, [...]. A new version of the package (v0.3.0) is now available on CRAN.
+Other noticeable changes made to the package include:
+
+  - The syntax for creating a cubble from separate spatial and temporal component has been changed from `as_cubble(list(spatial = ..., temporal = ...), ...)` to `make_cubble(spatial = ..., temporal = ..., ...)`.
+  - The new added toy data for demonstration: `stations` and `stations_sf` (the spatial component of three airport weather stations and it as an sf object), `meteo` and `meteo_ts` (the temporal component of the three weather stations and it as a tsibble object).
+  - A new version of the package (v0.3.0) is now available on CRAN.
 
@@
 
-**\textcolor{purple}{The suggested class implemented has been adopted, see The Package section in the Summary of Main Changes.
+**\textcolor{purple}{The suggested class implemented has been adopted, see The Package section in the Summary of Main Changes in this reponse.
 The codes on checking key values from multiple tables have been moved to a separate function check\_key() and reflected in the example 5.1 now.
@@
 The user would need to visit the documentation of tsibble in order to understand the meaning of the key and index. But the relationship to tsibble is not even mentioned in the docs.
-The documentation of the key and index is not perfect in tsible either, but sheds some more light on their role:
+The documentation of the key and index is not perfect in tsibble either, but sheds some more light on their role:
 
@@
 ````
-
 The doc states:
@@
 **\textcolor{purple}{Documentation of the package has been substantially improved with a focus on clarifying the input arguemnt types.
-For the specific example of as\_cubble(), the mentioned arguments now state:  }**
+For the specific example of as\_cubble(), the documentation now states:}**
 
@@
 
-**\textcolor{purple}{TODO: more here}**
+**\textcolor{purple}{cubble() is the class constructor and the documentation now states "Create a cubble object".
+as\_cubble() is the coercion function and documented as "Coerce foreign objects into a cubble object".}**
 
@@
 
-**\textcolor{purple}{the spatial attribute can be checked with function spaital(<cubbble\_obj>)}**
+**\textcolor{purple}{The spatial attribute can be checked with function spaital(<cubbble\_obj>).}**
 
@@
 
-**\textcolor{purple}{TODO}**
+**\textcolor{purple}{Cubble internally uses tidyr::nest() in the construction of a nested/spatial cubble (hence the name).
+We now mention the tidyr::nest() under the cubble attributes in Section 2.1 as an implication of rows being uniquely identified in the nested cubble.}**
 
@@
 
-  - on p6, section 3.2, yet another reason to call this data structure "cubble_temporal" instead of the semantically unobvious "long".
+  - on p6, section 3.2, yet another reason to call this data structure "cubble_temporal" instead of the semantically obvious "long".
 **\textcolor{purple}{see paragraph 4 under section The paper in this response.}**

### 2023-06-18 2-cubble.Rmd
section 2 and 3 done

@@
 
-<!-- Spatio-temporal data -->
-
-<!-- remote sensing, spatio-temporal cube: longitude, latitude, and time -->
-
-<!-- trajectory: each observation has unique coordinate (spatial) and temporal information -->
-
-<!-- the spacetime paper classifies spatio-temporal data into four different categories based on different spatial and temporal characteristics.  -->
-
-<!-- Analysts to be clear about the spatio-temporal layouts and choose the corresponding data structure for analysis. -->
-
-<!-- Spatio-temporal data can be conceptualized using a cubical data model with three axes which typically are, time, latitude and longitude.
-This abstraction can be useful for generalizing operations and visualization purposes: @lu_multidimensional_2018 shows how array operations (select, scale, reduce, rearrange, and compute) can be mapped onto the cube; @bach_review_2014 reviews the temporal data visualization based on space-time cube operations.
-Notice that the term space-time cube in their article "does not need to involve spatial data", but refers to "an abstract 2D substrate that is used to visualize data at a specific time".
-Despite its main focus being on temporal data, the mindset of abstracting out data representation to construct visualizations, still applies to our spatio-temporal data manipulation and visualization approaches.
--->
-
-<!-- The most common space-time cube uses the three axes, time, latitude, longitude, and can be considered stacking space across time.
-Ours is a multivariate spatio-temporal cube with the three axes defined to be time, site and variables, as illustrated in the leftmost column of Figure \ref{fig:cube-def}.
-The time axis is the same in both versions, while the site axis now captures both latitude and longitude.
-Finally, variables are stacked on this space-time canvas, with one observation per site and time point.
-This notion is adopted to avoid using hyper-cubes when describing multivariate spatio-temporal data and is the conceptual framework behind the \code{cubble} class.
-With this conceptual model, operations on spatio-temporal data can be mapped to operations on the cube and the rest of Figure \ref{fig:cube-def} show examples of slicing on site, time, and variable.
--->
-
-[conceptual framework]
-
 Spatio-temporal data can encompass data with various spatial and temporal characteristics and different data require different structures for wrangling and analysis: climate weather stations record data at fixed point location but may suffer from potential temporal data quality issue such as missing data for certain days.
@@
 The spatio-temporal data cubble address are those collected at unique fixed locations while allowing for irregularity in the temporal dimension, such as the weather station data.
-This corresponds to the full space-time and sparse space-time layouts in the spacetime paper [@spacetime].
+This corresponds to the full space-time and sparse space-time layouts described in the spacetime paper [@spacetime] and reproduced in Figure  \ref{fig:layout}.
+
+```{r layout, echo = FALSE, fig.cap="The full space-time and sparse space-time layouts handled by the cubble package.
+Reproduced from Figure 1 in the spacetime paper \\citep{spacetime}.", fig.width = 8, fig.height=4, fig.align='center'}
+library(tidyverse)
+library(patchwork)
+p1 <- tibble(
+  id = 1:12,
+  x = rep(1:4, each = 3),
+  y = rep(1:3, 4)
+) %>% 
+  ggplot(aes(x = as.factor(x) ,y = as.factor(y))) + 
+  geom_point() + 
+  geom_text(aes(y = y - 0.2, label = id)) + 
+  theme_bw() + 
+  theme(aspect.ratio = 1) + 
+  scale_x_discrete(labels = c("1st", "2nd", "3rd", "4th")) + 
+  scale_y_discrete(labels = c("1st", "2nd", "3rd", "4th")) + 
+  labs(x = "Time points", y = "Spatial features", title = "STF: full grid layout")
+
+p2 <- tibble(
+  id = 1:7,
+  x = c(1, 1, 1, 2, 2, 3, 4),
+  y = c(1, 2, 3, 2, 3, 1, 2),
+  label = c("1[1,1]", "2[1,2]", "3[1,3]", "4[2,2]", "5[3,2]", "6[1,3]", "7[2,4]")
+) %>% 
+  ggplot(aes(x = as.factor(x) ,y = as.factor(y))) + 
+  geom_point() + 
+  geom_text(aes(y = y - 0.2, label = label)) + 
+  theme_bw() + 
+  theme(aspect.ratio = 1) + 
+  scale_x_discrete(labels = c("1st", "2nd", "3rd", "4th")) + 
+  scale_y_discrete(labels = c("1st", "2nd", "3rd", "4th")) + 
+  labs(x = "Time points", y = "Spatial features", title = "STS: sparse grid layout")
+
+p1 | p2
+```
 
@@
 
--   a nested/ spatial cubble is represented by the class `c("spatial_cubble_df", "cubble_df")`
--   a long/ temporal cubble is represented by the class `c("temporal_cubble_df", "cubble_df")`
+-   a nested/ spatial cubble with class `c("spatial_cubble_df", "cubble_df")`
+-   a long/ temporal cubble with class `c("temporal_cubble_df", "cubble_df")`
 
@@
 
-A cubble object inherits the attributes from tibble (and its subclasses): `class`, `row.names`, and `names`. Additionally, it has three specialised attributes:
-
--   `key`: the spatial identifier
--   `index`: the temporal identifier
--   `coords`: a pair of ordered coordinates associated with the location
-
+A cubble object inherits the attributes from tibble (and its subclasses): `class`, `row.names`, and `names`.
+Additionally, it has three specialised attributes: `key`, `index`, and `coords`.
 Readers who are familiar with the `key` and `index` attributes from the `tsibble` package would already understand the two arguments.
-In cubble, the `key` attribute identifies the row in the nested cubble, and when combined with the `index` argument, it identifies the row in the long cubble.
+In cubble, the `key` attribute identifies the row in the nested cubble (given the use of `tidyr::nest()` for nesting), and when combined with the `index` argument, it identifies the row in the long cubble.
 Currently, cubble only supports one variable as the key, and the accepted temporal classes for index includes the base R classes `Date`, `POSIXlt`, `POSIXct`, as well as tsibble's `yearmonth`, `yearweek`, and `yearquarter` classes.
@@
 It can be either an unprojected pair of longitude and latitude, or a projected easting and northing value.
-The `sf` package is used under teh hood to calculate the bounding box, displayed in the header of a nested cubble.
+The `sf` package is used under the hood to calculate the bounding box, displayed in the header of a nested cubble.
 

### 2023-06-18 3-advanced.Rmd
section 2 and 3 done

@@
 The workflow with the \code{cubble} class works well with an interactive graphics pipeline (e.g., @buja1988elements; @buja1996interactive;  @sutherland2000orca; @xie2014reactive; @cheng2016enabling) that is available in \proglang{R} with the package \pkg{crosstalk} [@crosstalk].
-Figure \ref{fig:illu-interactive} illustrates how linking can be achieved between a map and multiple time series in a \code{cubble} object.
-The map (produced from the nested form) and time series (produced from the long form) are both shared \code{crosstalk} objects.
-When a user makes a selection on the map, the site is highlighted (a).
-This activates a row in the nested \code{cubble}, which is then communicated to the long \code{cubble} -- all the observations with the same id (b) will be selected.
-The long \code{cubble} will then highlight the corresponding series in the time series plot (c).
-
-Linking is also available starting from the time series plot, by selecting points.
-This will activate rows having the same id in the long \code{cubble}.
-The corresponding rows in the nested \code{cubble} are activated, and highlighted on the map.
-(An illustration can be found in the appendix.) Note that this type of linking, both from the map or the time series, is what @CS07 would call categorical variable linking, where station id is the categorical variable.
+Figure \ref{fig:illu-interactive} illustrates the linking between a map and multiple time series in a \code{cubble} object.
+Both the map, generated from the nested cubble, and the time series plot, generated from the long cubble, are shared crosstalk objects.
+When a user selects a location on the map (a), the corresponding site is highlighted.
+This selection activates a row in the nested \code{cubble}, which is then communicated to the long \code{cubble}, resulting in teh selection of all observations with the same ID in (b).
+Consequently the long \code{cubble} highlights the corresponding series in the time series plot (c).
+Linking can also be initiated from the time series plot, by selecting points on the time series.
+This action activates rows with the same ID in the long \code{cubble} and the row in the nested \code{cubble}, which will be highlighted on the map.
+Please refer to the appendix for the corresponding illustration.
 

### 2023-06-18 jss/main.tex
section 2 and 3 done

@@
 
-{[}conceptual framework{]}
-
 Spatio-temporal data can encompass data with various spatial and temporal characteristics and different data require different structures for wrangling and analysis: climate weather stations record data at fixed point location but may suffer from potential temporal data quality issue such as missing data for certain days.
@@
 The spatio-temporal data cubble address are those collected at unique fixed locations while allowing for irregularity in the temporal dimension, such as the weather station data.
-This corresponds to the full space-time and sparse space-time layouts in the spacetime paper \citep{spacetime}.
+This corresponds to the full space-time and sparse space-time layouts described in the spacetime paper \citep{spacetime} and reproduced in Figure \ref{fig:layout}.
+
+\begin{CodeChunk}
+\begin{figure}
+
+{\centering \includegraphics{/Users/hzha400/Documents/research/paper-cubble/figures/layout-1}
+
+}
+
+\caption[The full space-time and sparse space-time layouts handled by the cubble package]{The full space-time and sparse space-time layouts handled by the cubble package.
+Reproduced from Figure 1 in the spacetime paper \citep{spacetime}.}\label{fig:layout}
+\end{figure}
+\end{CodeChunk}
 
@@
 \item
-  a nested/ spatial cubble is represented by the class \texttt{c("spatial\_cubble\_df",\ "cubble\_df")}
+  a nested/ spatial cubble with class \texttt{c("spatial\_cubble\_df",\ "cubble\_df")}
 \item
-  a long/ temporal cubble is represented by the class \texttt{c("temporal\_cubble\_df",\ "cubble\_df")}
+  a long/ temporal cubble with class \texttt{c("temporal\_cubble\_df",\ "cubble\_df")}
 \end{itemize}
@@
 
-A cubble object inherits the attributes from tibble (and its subclasses): \texttt{class}, \texttt{row.names}, and \texttt{names}. Additionally, it has three specialised attributes:
-
-\begin{itemize}
-\tightlist
-\item
-  \texttt{key}: the spatial identifier
-\item
-  \texttt{index}: the temporal identifier
-\item
-  \texttt{coords}: a pair of ordered coordinates associated with the location
-\end{itemize}
-
+A cubble object inherits the attributes from tibble (and its subclasses): \texttt{class}, \texttt{row.names}, and \texttt{names}.
+Additionally, it has three specialised attributes: \texttt{key}, \texttt{index}, and \texttt{coords}.
 Readers who are familiar with the \texttt{key} and \texttt{index} attributes from the \texttt{tsibble} package would already understand the two arguments.
-In cubble, the \texttt{key} attribute identifies the row in the nested cubble, and when combined with the \texttt{index} argument, it identifies the row in the long cubble.
+In cubble, the \texttt{key} attribute identifies the row in the nested cubble (given the use of \texttt{tidyr::nest()} for nesting), and when combined with the \texttt{index} argument, it identifies the row in the long cubble.
 Currently, cubble only supports one variable as the key, and the accepted temporal classes for index includes the base R classes \texttt{Date}, \texttt{POSIXlt}, \texttt{POSIXct}, as well as tsibble's \texttt{yearmonth}, \texttt{yearweek}, and \texttt{yearquarter} classes.
@@
 It can be either an unprojected pair of longitude and latitude, or a projected easting and northing value.
-The \texttt{sf} package is used under teh hood to calculate the bounding box, displayed in the header of a nested cubble.
+The \texttt{sf} package is used under the hood to calculate the bounding box, displayed in the header of a nested cubble.
 
@@
 
-\hypertarget{using-a-tsibble-for-the-temporal-component}{%
-\subsubsection{Using a tsibble for the temporal component}\label{using-a-tsibble-for-the-temporal-component}}
+\hypertarget{using-a-tsibble-object-as-the-temporal-component}{%
+\subsubsection{Using a tsibble object as the temporal component}\label{using-a-tsibble-object-as-the-temporal-component}}
 
@@
 
-\hypertarget{using-sf-for-the-spatial-component}{%
-\subsubsection{Using sf for the spatial component}\label{using-sf-for-the-spatial-component}}
+\hypertarget{using-an-sf-object-as-the-spatial-component}{%
+\subsubsection{Using an sf object as the spatial component}\label{using-an-sf-object-as-the-spatial-component}}
 
@@
 Various distance measures are available (check \texttt{sf::st\_distance}).
-Analysts can specify the number of matched groups to output using the \texttt{spatial\_n\_group} argument (default to 4 groups) and the number of matches per group using the \texttt{spatial\_n\_group}argument (default to 1, one-to-one matching).
+Analysts can specify the number of matched groups to output using the \texttt{spatial\_n\_group} argument (default to 4 groups) and the number of matches per group using the \texttt{spatial\_n\_group} argument (default to 1, one-to-one matching).
 The syntax to use \texttt{match\_spatial()} is:
@@
 The workflow with the \code{cubble} class works well with an interactive graphics pipeline (e.g., \citet{buja1988elements}; \citet{buja1996interactive}; \citet{sutherland2000orca}; \citet{xie2014reactive}; \citet{cheng2016enabling}) that is available in \proglang{R} with the package \pkg{crosstalk} \citep{crosstalk}.
-Figure \ref{fig:illu-interactive} illustrates how linking can be achieved between a map and multiple time series in a \code{cubble} object.
-The map (produced from the nested form) and time series (produced from the long form) are both shared \code{crosstalk} objects.
-When a user makes a selection on the map, the site is highlighted (a).
-This activates a row in the nested \code{cubble}, which is then communicated to the long \code{cubble} -- all the observations with the same id (b) will be selected.
-The long \code{cubble} will then highlight the corresponding series in the time series plot (c).
-
-Linking is also available starting from the time series plot, by selecting points.
-This will activate rows having the same id in the long \code{cubble}.
-The corresponding rows in the nested \code{cubble} are activated, and highlighted on the map.
-(An illustration can be found in the appendix.) Note that this type of linking, both from the map or the time series, is what \citet{CS07} would call categorical variable linking, where station id is the categorical variable.
+Figure \ref{fig:illu-interactive} illustrates the linking between a map and multiple time series in a \code{cubble} object.
+Both the map, generated from the nested cubble, and the time series plot, generated from the long cubble, are shared crosstalk objects.
+When a user selects a location on the map (a), the corresponding site is highlighted.
+This selection activates a row in the nested \code{cubble}, which is then communicated to the long \code{cubble}, resulting in teh selection of all observations with the same ID in (b).
+Consequently the long \code{cubble} highlights the corresponding series in the time series plot (c).
+Linking can also be initiated from the time series plot, by selecting points on the time series.
+This action activates rows with the same ID in the long \code{cubble} and the row in the nested \code{cubble}, which will be highlighted on the map.
+Please refer to the appendix for the corresponding illustration.
 

### 2023-06-19 review/B-response.Rmd
re-formatted response

@@
 ---
-title: "Response to reviewers article 4780"
+title: "Response to Reviewers - Article 4780"
 subtitle: "cubble: An R Package for Organizing and Wrangling Multivariate Spatio-temporal Data"
@@
 
-We provide an overall summary of the changes made to the paper and the package before addressing reviewers' comments point-by-point.
+We provide an overall summary of the changes made to the paper and the package and then we make a point-by-point explanation of how we addressed the reviewers' comments.
 
-# Summary of Main Changes
+# Summary of the changes
 
-## The Paper:
+## Paper
 
@@
 
-## The Package:
+## Software
 
@@
 
-In the point-by-point response to reviewers' comments below, the original comments are in black and our responses are in red.
+\bigskip
+\bigskip
+\hrule
+
+\bigskip
+\bigskip
+
+Reviewers' comments are in black and \textcolor{purple}{our responses are in red}.
 
@@
 
-## The Paper
+## Paper
 
@@
 
-## The Package
+## Software
 
@@
 
-## Further suggestion
+## Further suggestions
 
@@
 
-## The Manuscript
+## Paper
 
@@
 
-## The package
+## Software
 

### 2023-06-20 2-cubble.Rmd
more wording updates

@@
 
-The cubble class is an S3 class built on tibble that allows the spatio-temporal data to be wrangled in two forms:  a nested/ spatialform and a long/temporal form. It consists of two subclasses:
+The cubble class is an S3 class built on tibble that allows the spatio-temporal data to be wrangled in two forms:  a nested/ spatial form and a long/temporal form. It consists of two subclasses:
 

### 2023-06-20 4-1-covid.Rmd
text update as per pkg vignettes
more wording updates

@@
 
-The five examples here are chosen to illustrate these aspects of the \pkg{cubble} package: creating a \code{cubble} object from two Coronavirus (COVID) data tables with the complication of differing location names, using spatial transformations to make a glyph map of seasonal temperature changes over years, matching river level data and weather station records for analysis of water supply, reading NetCDF format data to reproduce a climate reanalysis plot, and the workflow to create complex interactive linked plots.
-(The Appendix provides another example on aggregating information spatially to explore precipitation patterns.)
+Five examples are chosen to illustrate different aspects of the \pkg{cubble} package: creating a \code{cubble} object from two Coronavirus (COVID) data tables with the complication of differing location names, using spatial transformations to make a glyph map of seasonal temperature changes, matching river level data and weather station records for analysis of water supply, reading NetCDF format data to reproduce a climate reanalysis plot, and the workflow to create complex interactive linked plots.
 
@@
 
-Since the start of the pandemic, the Victoria State Government in Australia has provided daily COVID counts by local government area (LGA).
-This data can be used to visualize COVID incidence and spread spatially, when combined with map polygon data available from the Australian Bureau of Statistics.
-These different sources need to be combined for the analysis, by matching the LGA names.
-Both the COVID count data (\code{covid}) and the LGA information (\code{lga}) are available in the \pkg{cubble} package with the \code{covid} data stored as a \code{tsibble} object, the \code{lga} data as an \code{sf} object.
+Since the start of the pandemic, the Victoria State Government in Australia has been providing daily COVID counts per Logal Government Area (LGA).
+This data can be combined with map polygon data, available from the Australian Bureau of Statistics (ABS), to visualize COVID incidence and spread.
+In the \pkg{cubble} package, the COVID count data (\code{covid}) and the LGA information (\code{lga}) are available as a \code{tsibble} object and an \code{sf} object respectively.
 
-The \code{make_cubble()} function is used to create a \code{cubble} object from the two spatial and temporal tables as described in Section \@ref(create).
-The \code{by} argument allows for mismatch of key names from the two tables in the \code{*_join()} \code{by} syntax:
+A cubble object can be created from separate spatial and temporal component using the \code{make_cubble()} function, introduced in Section \@ref(create). The \code{by} argument
+is used to specify the joining variable from the two component using the \code{*_join()} \code{by} syntax:
 
@@
 
-The warning message reports on the mismatches of location from both sides: there are LGAs in the COVID data that do not match with LGAs names in the spatial polygon data, and vice versa.
-The \code{make_cubble()} function prompts the use of \code{check_key()} to identify the mismatches:
+The warning message suggests the slight difference of LGA encoding used by Victoria government and ABS and prompts analysts to use the function \code{check_key()} to identify the mismatches:
 
@@
 
-The check result is a list with three elements: 1) matched pairs from both tables, 2) those pairs that can be potentially paired, and 3) others.
-Here the main mismatch comes from different encodings of two LGAs: Kingston and Latrobe.
-Analysts can then modify the input spatial and temporal data accordingly and create the cubble again:
+The result of the \code{check_key()} function is a list containing three elements: 1) matched keys from both tables, 2) potentially pairable keys, and 3) others.
+Here, the main mismatch arises from the two LGAs: Kingston and Latrobe (Kingston is a LGA in both Victoria and South Australia and Latrobe is a LGA in both Victoria and Tasmania).
+Analysts can modify the input spatial and temporal data accordingly and recreate the cubble object:
 

### 2023-06-20 4-2-glyph.Rmd
more wording updates

@@
 
-p1 <- ggplot() + 
+p1 <- tmax %>% 
+  ggplot(aes(x_major = long, x_minor = month, 
+             y_major = lat, y_minor = tmax,
+             group = interaction(id, group))) + 
   geom_sf(data = nsw_vic, fill = "grey95", 
-          color = "white") + 
-  geom_glyph(
-    data = tmax, 
-    aes(x_major = long, x_minor = month, 
-        y_major = lat, y_minor = tmax,
-        group = interaction(id, group), color = group),
-    width = 1.2, height = 0.8) +
-    geom_rect(
-      data = box_df,
-      aes(xmin = long - 0.6, xmax = long + 0.6,
-          ymin = lat - 0.12, ymax = lat + 0.35),
-      fill = "transparent", color = "black") +
+          color = "white", inherit.aes = FALSE) + 
+  geom_glyph_box(data = tmax, width = 0.8, height = 0.3) +
+  geom_glyph(data = tmax, aes(color = group), width = 0.8, height = 0.3) +
+    # geom_rect(
+    #   data = box_df,
+    #   aes(xmin = long - 0.6, xmax = long + 0.6,
+    #       ymin = lat - 0.12, ymax = lat + 0.35),
+    #   fill = "transparent", color = "black") +
   scale_color_brewer("", palette = "Dark2") + 
@@
 In this example, they are irregularly spaced, which can result in overlapping glyphs obscuring each other.
-To fix this, one could aggregate data from nearby stations.
-An example of this use is included in the Appendix.

### 2023-06-20 4-4-netcdf.Rmd
more wording updates

@@
 
-Figure \ref{fig:netcdf} reproduces the ERA5 data row of Figure 19 in @hersbach2020era5.
-Here we explain how this would be done using in the \pkg{cubble} package.
-The plots show that the southern polar vortex splits into two on 2002-09-26 and further splits into four on 2002-10-04.
+The ERA5 data [@hersbach2020era5] provides hourly estimates across the Earth for atmospheric, land and oceanic climate variables.
+The data is available in the NetCDF format from the European Centre for Medium-Range Weather Forecasts (ECMWF) and can be downloaded from Copernicus Climate Data Store (CDS) [@cds] website or via the \pkg{ecmwfr} package [@ecwmfr].
+This example reproduces the ERA5 data row of Figure 19 in @hersbach2020era5 to show the southern polar vortex splitting into two on 2002-09-26 and further splitting into four on 2002-10-04.
 Further explanation of why this is interesting can be found in the figure source, and also in @simmons2020global and @simmons2005ecmwf.
 
-The ERA5 data [@hersbach2020era5] provides hourly estimates across the Earth for atmospheric, land and oceanic climate variables.
-The data is available in the NetCDF format from the European Centre for Medium-Range Weather Forecasts (ECMWF).
-It can be directly downloaded from Copernicus Climate Data Store (CDS) [@cds] website or via the \pkg{ecmwfr} package [@ecwmfr].
 For the reproduction, we focus on the \code{era5-pressure} data, hourly pressure levels from 1970 to present, with the *specific humidity* and *geopotential*.
@@
 Creating the plot requires making transformations on time, unfolding the data for computing the statistic of interest, which is plotted directly as a contour plot with ggplot.
-Code is provided to accomplish this in the supplementary material.
 

### 2023-06-20 jss/main.tex
more wording updates

@@
 
-The cubble class is an S3 class built on tibble that allows the spatio-temporal data to be wrangled in two forms: a nested/ spatialform and a long/temporal form. It consists of two subclasses:
+The cubble class is an S3 class built on tibble that allows the spatio-temporal data to be wrangled in two forms: a nested/ spatial form and a long/temporal form. It consists of two subclasses:
 
@@
 
-The five examples here are chosen to illustrate these aspects of the \pkg{cubble} package: creating a \code{cubble} object from two Coronavirus (COVID) data tables with the complication of differing location names, using spatial transformations to make a glyph map of seasonal temperature changes over years, matching river level data and weather station records for analysis of water supply, reading NetCDF format data to reproduce a climate reanalysis plot, and the workflow to create complex interactive linked plots.
+Five examples are chosen to illustrate different aspects of the \pkg{cubble} package: creating a \code{cubble} object from two Coronavirus (COVID) data tables with the complication of differing location names, using spatial transformations to make a glyph map of seasonal temperature changes, matching river level data and weather station records for analysis of water supply, reading NetCDF format data to reproduce a climate reanalysis plot, and the workflow to create complex interactive linked plots.
 
@@
 
-Since the start of the pandemic, the Victoria State Government in Australia has provided daily COVID counts by local government area (LGA).
-This data can be used to visualize COVID incidence and spread spatially, when combined with map polygon data available from the Australian Bureau of Statistics.
-These different sources need to be combined for the analysis, by matching the LGA names.
-Both the COVID count data (\code{covid}) and the LGA information (\code{lga}) are available in the \pkg{cubble} package with the \code{covid} data stored as a \code{tsibble} object, the \code{lga} data as an \code{sf} object.
-
-The \code{make_cubble()} function is used to create a \code{cubble} object from the two spatial and temporal tables as described in Section \ref{create}.
-The \code{by} argument allows for mismatch of key names from the two tables in the \code{*_join()} \code{by} syntax:
+Since the start of the pandemic, the Victoria State Government in Australia has been providing daily COVID counts per Logal Government Area (LGA).
+This data can be combined with map polygon data, available from the Australian Bureau of Statistics (ABS), to visualize COVID incidence and spread.
+In the \pkg{cubble} package, the COVID count data (\code{covid}) and the LGA information (\code{lga}) are available as a \code{tsibble} object and an \code{sf} object respectively.
+
+A cubble object can be created from separate spatial and temporal component using the \code{make_cubble()} function, introduced in Section \ref{create}. The \code{by} argument
+is used to specify the joining variable from the two component using the \code{*_join()} \code{by} syntax:
 
@@
 In this example, they are irregularly spaced, which can result in overlapping glyphs obscuring each other.
-To fix this, one could aggregate data from nearby stations.
-An example of this use is included in the Appendix.
 
@@
 
-Figure \ref{fig:netcdf} reproduces the ERA5 data row of Figure 19 in \citet{hersbach2020era5}.
-Here we explain how this would be done using in the \pkg{cubble} package.
-The plots show that the southern polar vortex splits into two on 2002-09-26 and further splits into four on 2002-10-04.
+The ERA5 data \citep{hersbach2020era5} provides hourly estimates across the Earth for atmospheric, land and oceanic climate variables.
+The data is available in the NetCDF format from the European Centre for Medium-Range Weather Forecasts (ECMWF) and can be downloaded from Copernicus Climate Data Store (CDS) \citep{cds} website or via the \pkg{ecmwfr} package \citep{ecwmfr}.
+This example reproduces the ERA5 data row of Figure 19 in \citet{hersbach2020era5} to show the southern polar vortex splitting into two on 2002-09-26 and further splitting into four on 2002-10-04.
 Further explanation of why this is interesting can be found in the figure source, and also in \citet{simmons2020global} and \citet{simmons2005ecmwf}.
 
-The ERA5 data \citep{hersbach2020era5} provides hourly estimates across the Earth for atmospheric, land and oceanic climate variables.
-The data is available in the NetCDF format from the European Centre for Medium-Range Weather Forecasts (ECMWF).
-It can be directly downloaded from Copernicus Climate Data Store (CDS) \citep{cds} website or via the \pkg{ecmwfr} package \citep{ecwmfr}.
 For the reproduction, we focus on the \code{era5-pressure} data, hourly pressure levels from 1970 to present, with the \emph{specific humidity} and \emph{geopotential}.
@@
 
-\hypertarget{interactive-graphics-1}{%
-\subsection{Interactive graphics}\label{interactive-graphics-1}}
+\hypertarget{australian-temperature-range}{%
+\subsection{Australian temperature range}\label{australian-temperature-range}}
 

### 2023-06-20 review/B-response.Rmd
changes after discussing with supervisors

@@
 
-Section 2 have been reworked to combine the original section 2 and 3.
+Section 2 has been reworked to combine the original Section 2 and 3.
 It now focuses more on the inner work of the package: the cubble class and its attributes, object creation and coercion, functions and methods implemented, compatibility with `tsibble` and `sf`, and comparison with other existing spatio-temporal packages (`stars` and `sftime`).
@@
 
-  * nested/ spatial cubble: `c("spatial_cubble_df", "cubble_df")`, and
-  * long/ temporal cubble: `c("temporal_cubble_df", "cubble_df")`.
+-   nested/ spatial cubble: `c("spatial_cubble_df", "cubble_df")`, and
+-   long/ temporal cubble: `c("temporal_cubble_df", "cubble_df")`.
 
 All the S3 methods have been modified accordingly.
-Methods behave consistently in the nested and long cubble are implemented in the `cubble_df` class, otherwise, separately in the `spatial_cubble_df` and `temporal_cubble_df` class.
+Methods that behave consistently in the nested and long cubble are implemented in the `cubble_df` class, otherwise, separately in the `spatial_cubble_df` and `temporal_cubble_df` classes.
 
@@
 
-  - The syntax for creating a cubble from separate spatial and temporal component has been changed from `as_cubble(list(spatial = ..., temporal = ...), ...)` to `make_cubble(spatial = ..., temporal = ..., ...)`.
-  - The new added toy data for demonstration: `stations` and `stations_sf` (the spatial component of three airport weather stations and it as an sf object), `meteo` and `meteo_ts` (the temporal component of the three weather stations and it as a tsibble object).
+-   The syntax for creating a cubble from separate spatial and temporal components has been changed from `as_cubble(list(spatial = ..., temporal = ...), ...)` to `make_cubble(spatial = ..., temporal = ..., ...)`.
+-   The key checking in creating cubble from separate spatial and temporal components has been moved to a separate function: `check_key()`.
+-   The newly added toy data for demonstration: `stations` and `stations_sf` (the spatial component of three airport weather stations and it as an sf object), `meteo` and `meteo_ts` (the temporal component of the three weather stations and it as a tsibble object).
 -   A new version of the package (v0.3.0) is now available on CRAN.
 
+```{=tex}
 \bigskip
@@
 \hrule
-
+```
+```{=tex}
 \bigskip
 \bigskip
+```
+Reviewers' comments are in black and **\textcolor{purple}{our responses are in red}**.
 
-Reviewers' comments are in black and \textcolor{purple}{our responses are in red}.
-
-# Response to B-review_1_1
+# Response to B-review_1\_1
 
@@
 
-  * nested which is indexed by spatial location and keeps the time-series data in a list column 
-  long which is indexed by time and keeps the spatial data in a hidden “spatial” attribute.
+-   nested which is indexed by spatial location and keeps the time-series data in a list column long which is indexed by time and keeps the spatial data in a hidden "spatial" attribute.
 
@@
 
-  * data transformation functions: face_temporal, face_spatial and unfold
-  * data matching (merging) functions: match_sites, match_spatial and match_temporal
+-   data transformation functions: face_temporal, face_spatial and unfold
+-   data matching (merging) functions: match_sites, match_spatial and match_temporal
 
@@
 Unfortunately the paper does not clearly describe the data structure, nor the associated functionality.
-I had to read the package vignettes, run the attached code and even look into the package’s code in order to fully understand the data structure.
+I had to read the package vignettes, run the attached code and even look into the package's code in order to fully understand the data structure.
 
-**\textcolor{purple}{We thank the reviewer for the support on our proposed data structure.
-We have reworked on section 2 to describe the class and its associated functionality.
-Additionally, the text in the paper has been aligned with the package vignette.
-}**
+**\textcolor{purple}{We have reworked Section 2 to describe the cubble class and its associated functionality. Additionally, the text in the paper has been aligned with the package vignette. }**
 
@@
 Most of the code in the paper is not self-contained which makes it rather difficult to follow.
-Examples at the end of the paper don’t elucidate the inner workings of the package but rather advertise the functionality of other, mostly visualization, packages.
+Examples at the end of the paper don't elucidate the inner workings of the package but rather advertise the functionality of other, mostly visualization, packages.
 
-**\textcolor{purple}{We have selected and displayed the codes and outputs relevant to demonstrate the cubble functionality in the examples.
-The full script is available in the reproducible submitted reproducible script or the paper GitHub repository: https://github.com/huizezhang-sherry/paper-cubble.}**
+**\textcolor{purple}{The data structure proposed in the paper is an infrastructure development to carry the spatio-temporal data analysis and visualisation is a crucial part of the analysis that delivers the results in visuals.
+We add a paragraph at the beginning of the now Section 4 to explain the purpose of these examples.}**
 
@@
 The cubble class is used to represent two incompatible data types - long and nested.
-Internally, the two data types are distinguished by the internal attribute “form”.
+Internally, the two data types are distinguished by the internal attribute "form".
 In other words, the authors use an attribute of an object to mimic the functionality of a sub-class.
@@
 Such a generic implementation would allow for natural extension of the functionality through dedicated methods.
-Currently an ever-present check on “form” attribute is necessary throughout the code base.
+Currently an ever-present check on "form" attribute is necessary throughout the code base.
 To make matters worse, as_cubble currently produces yet another data structure - a list of paired matches, as can be seen from example 5.1 in the paper.
 
-**\textcolor{purple}{The suggested class implemented has been adopted, see The Package section in the Summary of Main Changes in this reponse.
-The codes on checking key values from multiple tables have been moved to a separate function check\_key() and reflected in the example 5.1 now.
-The function as\_cubble() now returns nested cubbles.}**
+**\textcolor{purple}{Your suggestion has been adopted, see the Software section in the summary of the changes for details. The code in the example (now Section 4.1) is now simpler.}**
 
@@
 It seems to me that the semantically unambiguous temporal/spatial could be used throughout without ambiguity, thus resolving the terminological redundancy.
-
-**\textcolor{purple}{The naming of long/nested comes from the tidy data concept where the spatial cubble has the nested temporal information and the temporal cubble uses the long form data from the tidy data.
-We understand it could be more intuitive to refer to them as spatial/temporal cubble and we adopt the spatial/temporal cubble when it is more straightforward to use.}**
+**\textcolor{purple}{We have replaced nested/long cubble with spatial/temporal cubble.}**
 
@@
 
-  * Rework the definitional parts (sections 2 and 3) by following more closely the “design” vignette.
-**\textcolor{purple}{Adopted, see The Paper section in the Summary of Main Changes of this response}**
+-   Rework the definitional parts (sections 2 and 3) by following more closely the "design" vignette. **\textcolor{purple}{Done.}**
 
-  * Describe at a glance the core functions which operate on the class and their motivation. **\textcolor{purple}{Functions and methods at a glance is now available at the start of Section 2.3.}**
-  * Rework section 4 by making it more concise and provide links to sections in supplementary material or dedicated vignettes. **\textcolor{purple}{Reworked. Text has been edited for conciseness}**
-  * Rework examples section 5 into “Applications”.
-Briefly describe the applications and provide and refer to dedicated and self-contained vignettes.
-**\textcolor{purple}{Adopted, see the last point.}**
+-   Describe at a glance the core functions which operate on the class and their motivation. **\textcolor{purple}{Done in Section 2.3.}**
+
+-   Rework section 4 by making it more concise and provide links to sections in supplementary material or dedicated vignettes. **\textcolor{purple}{Done.}**
+
+-   Rework examples section 5 into "Applications". Briefly describe the applications and provide and refer to dedicated and self-contained vignettes. **\textcolor{purple}{Done.}**
 
@@
 
-````
         data: the object to be created or tested as cubble
@@
          ...: a list object to create new cubble
-````
 
-  * What objects are supported as input data? 
-  * What is the assumption of the “nestedness” of the input data. 
-  * What is the accepted data type of key, index and coords?
+-   What objects are supported as input data?
+-   What is the assumption of the "nestedness" of the input data.
+-   What is the accepted data type of key, index and coords?
 
-The user would need to visit the documentation of tsibble in order to understand the meaning of the key and index. But the relationship to tsibble is not even mentioned in the docs.
+The user would need to visit the documentation of tsibble in order to understand the meaning of the key and index.
+But the relationship to tsibble is not even mentioned in the docs.
 The documentation of the key and index is not perfect in tsibble either, but sheds some more light on their role:
 
-````
          key: Variable(s) that uniquely determine time indices. ‘NULL’ for
@@
 
-````
 The doc states:
 
-````
        The constructor for the cubble class
-````
 
-If it’s the constructor then the convention is to name it cubble().
+If it's the constructor then the convention is to name it cubble().
 as_xyz is the converter which translates between different data types.
@@
 
-**\textcolor{purple}{Documentation of the package has been substantially improved with a focus on clarifying the input arguemnt types.
+**\textcolor{purple}{Documentation of the package has been substantially improved with a focus on clarifying the input argument types.
 For the specific example of as\_cubble(), the documentation now states:}**
 
-````
-    data: an object to be converted into an cubble object. 
-          Currently support objects of classes `tibble`, `ncdf4`, `stars`, and `sftime`. 
-     key: a character (or symbol), the spatial identifier.
-   index: a character (or symbol), the temporal identifier. Currently support base R 
-          classes Date, POSIXlt, POSIXct and tsibble's yearmonth, yearweek, and yearquarter
-          class.
-  coords: a vector of character (or symbol) of length two, in the order of longitude
-          first and then latitude, the argument can be omitted if created from an sf 
-          and its subclasses. In case the sf geometry column is not POINT, coords will 
-          be the centroid coordinates.
-
-````
-
-**\textcolor{purple}{cubble() is the class constructor and the documentation now states "Create a cubble object".
-as\_cubble() is the coercion function and documented as "Coerce foreign objects into a cubble object".}**
+  - **\textcolor{purple}{data: an object to be converted into a cubble object. Currently support objects of classes tibble, ncdf4, stars, and sftime.}**
+  - **\textcolor{purple}{key: a character (or symbol), the spatial identifier.}**
+  - **\textcolor{purple}{index: a character (or symbol), the temporal identifier. Currently support base R classes Date, POSIXlt, POSIXct, and tsibble's yearmonth, yearweek, and yearquarter class.}**
+  - **\textcolor{purple}{coords: a vector of character (or symbol) of length two, in the order of longitude first and then latitude, the argument can be omitted if created from an sf and its subclasses.
+In case the sf geometry column is not POINT, coords will be the centroid coordinates.}**
 
@@
 
-Consider not throwing an error when face_spatial is applied to the spatial object. This would allow for generic code where the “form” of the input the data is not known or does not matter.
+Consider not throwing an error when face_spatial is applied to the spatial object. This would allow for generic code where the "form" of the input the data is not known or does not matter.
 
 **\textcolor{purple}{Adopted.
-Now when applying face\_spatial to a nested cubble, the cubble will be printed as it is with a message signalling it is already in the nested form.
+Now when applying face\_spatial to a spatial cubble, the cubble will be printed as it is with a message signalling it is already in the nested form.
 see the reprex below.
-Same for face\_temporal on a long cubble object.}**
+Same for face\_temporal on a temporal cubble object.}**
 
@@
 
-It would be useful to print a few rows of the spatial attribute of the “long” object.
+It would be useful to print a few rows of the spatial attribute of the "long" object.
+**\textcolor{purple}{It doesn't make sense to print spatial rows if it is a subclass temporal\_cubble\_df.
+But they do get printed in the subclass spatial\_cubble\_df and you can extract the spatial variables with spatial() as explained in Section 2.1.
+}**
 
-**\textcolor{purple}{The spatial attribute can be checked with function spaital(<cubbble\_obj>).}**
+Section 3.6 is missing `tidyr` package which is especially relevant here given its nesting and unnesting operations <https://tidyr.tidyverse.org/articles/nest.html> **\textcolor{purple}{The tidyr functions are not applicable to a cubble.
+The idea of tidy data is underlying the cubble class as explained in Section 2.1.}**
 
-Section 3.6 is missing `tidyr` package which is especially relevant here given its nesting and unnesting operations https://tidyr.tidyverse.org/articles/nest.html
+More comments inlined in the pdf. **\textcolor{purple}{Done with the section rework.}**
 
-**\textcolor{purple}{Cubble internally uses tidyr::nest() in the construction of a nested/spatial cubble (hence the name).
-We now mention the tidyr::nest() under the cubble attributes in Section 2.1 as an implication of rows being uniquely identified in the nested cubble.}**
-
-More comments inlined in the pdf.
-
-  - on p4 "The following sections explain their roles, why the new cubble class is needed
-and how the package relates to existing packages for spatial and temporal data analysis." The paper is a bit on the verbose side.
-Such comments could be safely removed.
-**\textcolor{purple}{sentence removed.}**
-
-  - on p5 section 3.1, "The arguments key and index follow the wording in the tsibble package to describe the temporal order and multiple series while coords specifies the spatial location of each site.", Unclear what `key` and `index` mean.
-Isn't key a spatial identifier and index a temporal one?
-**\textcolor{purple}{The `key` and `index`, along with other attributes are explained under heading *The cubble attributes* in Section 2.1.}**
-
-  - on p5 section 3.1, the word "wording" is highlighted in the sentence "The arguments key and index follow the wording in the tsibble package to ...", conventions.
-**\textcolor{purple}{The setence has been rewrite}**
-
-  - on p6 section 3.2, 'The cubble class also provides a long form, which expands the ts column and temporarily “hides” the spatial variables.', confusing: `hides` means storing it as an attribute.
-**\textcolor{purple}{The sentence is now on p4 section 2.1 before the printing of object cb\_long as "In a long cubble, temporal variables are expanded in the long form and spatial variables are stored as a data attribute."}**
-
-  - on p6 section 3.2, I think it would be a good idea to clearly describe the two data structures in an earlier section.
-By this point it's still not clear what those are and what the meaning of key and index is.
-**\textcolor{purple}{The description of the two data structures is now at section 2.1}**
-
-  - on p6, section 3.2, yet another reason to call this data structure "cubble_temporal" instead of the semantically obvious "long".
-**\textcolor{purple}{see paragraph 4 under section The paper in this response.}**
-
-  - on p8 section 3.5, Isn't the root reason the memory efficiency?
-That is, the spatial component cannot be represented as part of the nested tibble without repetition.
-**\textcolor{purple}{this has now been incorporated in section 2.5.}**
-
-  - on p8 section 3.6, The missing elephant is tidyr which has nesting and unnesting operations  https://tidyr.tidyverse.org/articles/nest.html.
-**\textcolor{purple}{addressed in Further suggestion in this response}**
-
-# Response to B-review_2_1
+# Response to B-review_2\_1
 
@@
 
-**\textcolor{purple}{A new section, Section 2.5, has been dedicated to compare cubble with existing spatio-temporal class, specifically, stars and sftime.
-See The Paper section in the Summary of Main Changes of this response}.**
+**\textcolor{purple}{A new section, Section 2.5, has been dedicated to comparing cubble with existing spatio-temporal classes, specifically, stars and sftime.
+See the Paper section in the summary of the changes.}**
 
@@
 In the case where a cubble object is created from an sf object, the sf class will be retained.
-When it is not, users can promote the cubble object to inlcude include the sf class with make\_spatial\_sf(): }**
+When it is not, users can promote the cubble object to include the sf class with make\_spatial\_sf()}**
 
@@
 **\textcolor{purple}{In spatial matching, several distance calculations are now made available under sf::st\_distance.
-Temporal matching now accepts an argument temporal\_match\_fn to allow user specified functions for matching.}**
+Temporal matching now accepts an argument temporal\_match\_fn to allow user-specified functions for matching.}**
 
@@
 -   doubled/missing words/none correct sentences:
-  - “*components* spatio-temporal *components*” **\textcolor{purple}{to "... spatio-temporal components"}**
-  - “... fits works ...”  **\textcolor{purple}{to ".. works well with ..."}**
-  - “... be activate rows ...” **\textcolor{purple}{to "... will activate rows ..."}**
-  - “... highlighted *in* the ...” **\textcolor{purple}{to "... highlighted on the map"}**
-  - “. . . using 2020 measurements using match_sites() function.”  **\textcolor{purple}{to "... using the match\_sites() function"}**
-  - “An example of this using is included in the Appendix” **\textcolor{purple}{to "... this use ..."}**
-  - “. . . the data in -a- multiple -of- ways on-the-fly”  **\textcolor{purple}{to "... in multiple ways ..."}**
 
-- Surprising references: in Section 4.4 it says “.
+    -   "*components* spatio-temporal *components*" **\textcolor{purple}{to "... spatio-temporal components"}**
+    -   "... fits works ..." **\textcolor{purple}{to ".. works well with ..."}**
+    -   "... be activate rows ..." **\textcolor{purple}{to "... will activate rows ..."}**
+    -   "... highlighted *in* the ..." **\textcolor{purple}{to "... highlighted on the map"}**
+    -   ". . . using 2020 measurements using match_sites() function." \**textcolor{purple}{to "... using the match\_sites() function"}**
+    -   "An example of this using is included in the Appendix" **\textcolor{purple}{to "... this use ..."}**
+    -   ". . . the data in -a- multiple -of- ways on-the-fly" **\textcolor{purple}{to "... in multiple ways ..."}**
+
+-   Surprising references: in Section 4.4 it says ".
 .
 .
-Glyph maps (Section 3.4)”, but Section 3.4 says “.
+Glyph maps (Section 3.4)", but Section 3.4 says ".
 .
 .
-glyph maps will be explained in Section 4.4”.
+glyph maps will be explained in Section 4.4".
 **\textcolor{purple}{the looped reference in section 3.3 and 3.4 is replaced with (Wickham et al.
@@
 
-- Should “. . . it’s temperatures are more *consistent*” rather be “. . . it’s temperatures are more *constant*”? **\textcolor{purple}{to "more constant"}**
+-   Should ". . . it's temperatures are more *consistent*" rather be ". . . it's temperatures are more *constant*"? **\textcolor{purple}{to "more constant"}**
 
 -   Typos:
-  - “. . . spatial and *tmeporal* information are available.” **\textcolor{purple}{to "temporal"}**
-  - The polar vortex, signalled by the high *speicfic* humidity, splits into two on 2002-09-26 and *further-s- split_s_* into four on 2002-10-04.
+
+    -   ". . . spatial and *tmeporal* information are available." **\textcolor{purple}{to "temporal"}**
+    -   The polar vortex, signalled by the high *speicfic* humidity, splits into two on 2002-09-26 and *further-s- split_s\_* into four on 2002-10-04.
 **\textcolor{purple}{to "specific", "further splits"}**
@@
 
-I have been a bit puzzled by the “print” of a nested cubble in its temporal face.
-The row # temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl] appears at first sight (also in comparison with the print of the spatial face), as if the temporal domain is given by all those variables.
+I have been a bit puzzled by the "print" of a nested cubble in its temporal face.
+The row \# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl] appears at first sight (also in comparison with the print of the spatial face), as if the temporal domain is given by all those variables.
 To me, a notion such as
@@
 
-**\textcolor{purple}{For temporal cubble, the header now prints the temporal meta data: start date, end date, interval, and whether there is gaps in the data. }**
-
-````
-#> # cubble:   key: id [3], index: date, long form
-#> # temporal: 2020-01-01 -- 2020-01-10 [1D], no gaps
-#> # spatial:  long [dbl], lat [dbl], elev [dbl], name [chr], wmo_id [dbl]
-````
+**\textcolor{purple}{For temporal cubble, the header now prints the temporal meta data: start date, end date, interval, and whether there are gaps in the data, e.g.
+"temporal: 2020-01-01 -- 2020-01-10 [1D], no gaps"}**
 
@@
 
-**\textcolor{purple}{The data historical\_tmax contains daily weather stations data for the two periods: 1971-1975 and 2016-2020 and amounts to 3.2MB, which is not suitable to include in the package given its size.
+**\textcolor{purple}{The data historical\_tmax contains daily weather station data for the two periods: 1971-1975 and 2016-2020 and amounts to 3.2MB, which is not suitable to include in the package given its size.
 The tmax\_hist data, which was included in the package, is now removed, given its similarity to the climate\_aus data, which is more useful.
@@
 The manuscript cannot be reproduced, as the reproducible-script.R (part of the submission to JSS) only uses a subset, but does not provide a separate result file that would allow to compare the reproduced outputs with the desired output for the subset.
-
-**\textcolor{purple}{A result file in html is submitted this time for result comparison.}**
+**\textcolor{purple}{A result file in HTML is submitted this time for result comparison.}**

### 2023-06-20 review/B-response.Rmd
response revised

@@
 
-Section 2 has been reworked to combine the original Section 2 and 3.
-It now focuses more on the inner work of the package: the cubble class and its attributes, object creation and coercion, functions and methods implemented, compatibility with `tsibble` and `sf`, and comparison with other existing spatio-temporal packages (`stars` and `sftime`).
+Section 2 has been reworked to combine the original Sections 2 and 3.
+It now focuses more on the inner workings of the package: the cubble class and its attributes, object creation and coercion, functions and methods implemented, compatibility with `tsibble` and `sf`, and comparison with other existing spatio-temporal packages (`stars` and `sftime`).
 
 In the comparison section (Section 2.5), we compare `cubble` with `stars` and `sftime`.
-With `stars`, we believe spatio-temporal analysts who are more comfortable processing data with data frame (tidyverse) will benefit from cubble, given `stars` is built from arrays.
-With `sftime`, we argue cubble is more efficient in memory since it doesn't repeat spatial variables at each time stamp.
+With `stars`, we believe spatio-temporal analysts who are more comfortable processing data with tidy tools operating on `tibble`s from tidyverse will find `cubble` useful, especially since `stars` is built from arrays.
+With `sftime`, we argue that `cubble` is more efficient in memory since it doesn't repeat spatial variables at each time stamp.
 We give an example to compare the object size of the same data stored as an `sftime` object and a `cubble` object (118MB vs.
@@
 
-Other sections of the paper have been edited for conciseness and they match with the package vignettes.
+Other sections of the paper have been changed for conciseness and to better match the numerous package vignettes.
 
@@
 
-The cubble class has been re-factored as suggested by reviewer 1. Now the cubble class has two sub-classes:
+The `cubble` class has been re-factored as suggested by reviewer 1. Now the `cubble` class has two sub-classes, instead of forms:
 
@@
 All the S3 methods have been modified accordingly.
-Methods that behave consistently in the nested and long cubble are implemented in the `cubble_df` class, otherwise, separately in the `spatial_cubble_df` and `temporal_cubble_df` classes.
+Methods that behave consistently in the nested and long cubble are implemented in the `cubble_df` class, or if appropriate, separately in the `spatial_cubble_df` and `temporal_cubble_df` classes.
 
-Other noticeable changes made to the package include:
+Other substantial changes made to the package include:
 
--   The syntax for creating a cubble from separate spatial and temporal components has been changed from `as_cubble(list(spatial = ..., temporal = ...), ...)` to `make_cubble(spatial = ..., temporal = ..., ...)`.
--   The key checking in creating cubble from separate spatial and temporal components has been moved to a separate function: `check_key()`.
--   The newly added toy data for demonstration: `stations` and `stations_sf` (the spatial component of three airport weather stations and it as an sf object), `meteo` and `meteo_ts` (the temporal component of the three weather stations and it as a tsibble object).
+-   The syntax for creating a `cubble` from separate spatial and temporal components has been changed from `as_cubble(list(spatial = ..., temporal = ...), ...)` to `make_cubble(spatial = ..., temporal = ..., ...)`.
+-   The key checking in creating `cubble` from separate spatial and temporal components has been moved into a separate function: `check_key()`.
+-   The newly added toy data for demonstration: `stations` and `stations_sf` (the spatial component of three airport weather stations and it as an `sf` object), `meteo` and `meteo_ts` (the temporal component of the three weather stations and it as a `tsibble` object).
 -   A new version of the package (v0.3.0) is now available on CRAN.
@@
 ```
+
 Reviewers' comments are in black and **\textcolor{purple}{our responses are in red}**.
@@
 
-**\textcolor{purple}{We have reworked Section 2 to describe the cubble class and its associated functionality. Additionally, the text in the paper has been aligned with the package vignette. }**
+**\textcolor{purple}{We have reworked Section 2 to more completely describe the cubble class and its associated functionality.
+Additionally, the text in the paper has been aligned with the package vignette.}**
 
@@
 
-**\textcolor{purple}{The data structure proposed in the paper is an infrastructure development to carry the spatio-temporal data analysis and visualisation is a crucial part of the analysis that delivers the results in visuals.
-We add a paragraph at the beginning of the now Section 4 to explain the purpose of these examples.}**
+**\textcolor{purple}{The data structure proposed in the paper is an infrastructure development to make the the analysis of spatio-temporal data simpler and to better construct visualisation.
+The examples illustrate how the infrastructure can be used as part of routine analyses.
+We added a paragraph at the beginning of Section 4 to explain this purpose.}**
 
@@
 
-**\textcolor{purple}{Your suggestion has been adopted, see the Software section in the summary of the changes for details. The code in the example (now Section 4.1) is now simpler.}**
+**\textcolor{purple}{Your suggestion has been adopted, see the Software section in the summary of the changes for details. The code in the example (now Section 4.1) is simplified.}**
 
@@
 
-**\textcolor{purple}{Adopted.
+**\textcolor{purple}{Done.
 Now when applying face\_spatial to a spatial cubble, the cubble will be printed as it is with a message signalling it is already in the nested form.
-see the reprex below.
+See the reprex below.
 Same for face\_temporal on a temporal cubble object.}**
@@
 It would be useful to print a few rows of the spatial attribute of the "long" object.
-**\textcolor{purple}{It doesn't make sense to print spatial rows if it is a subclass temporal\_cubble\_df.
-But they do get printed in the subclass spatial\_cubble\_df and you can extract the spatial variables with spatial() as explained in Section 2.1.
-}**
+**\textcolor{purple}{When is is subclass spatial\_cubble\_df these rows are printed, but it could be misleading to do so if it is a subclass temporal\_cubble\_df.
+Users can extract the rows of the spatial variables with the spatial() function, as explained in Section 2.1.}**
 
-Section 3.6 is missing `tidyr` package which is especially relevant here given its nesting and unnesting operations <https://tidyr.tidyverse.org/articles/nest.html> **\textcolor{purple}{The tidyr functions are not applicable to a cubble.
-The idea of tidy data is underlying the cubble class as explained in Section 2.1.}**
+Section 3.6 is missing `tidyr` package which is especially relevant here given its nesting and unnesting operations <https://tidyr.tidyverse.org/articles/nest.html> **\textcolor{purple}{The idea of tidy data is underlies the cubble class (as explained in Section 2.1) but tidyr functions are not applicable to a cubble.
+This is the reason that tidyr was not included in the original section 3.6.
+The section has been replaced by Tables 1 and 2.}**
 
-More comments inlined in the pdf. **\textcolor{purple}{Done with the section rework.}**
+More comments inlined in the pdf. **\textcolor{purple}{Done with the Section restructuring.}**
 
@@
 
-    -   "*components* spatio-temporal *components*" **\textcolor{purple}{to "... spatio-temporal components"}**
-    -   "... fits works ..." **\textcolor{purple}{to ".. works well with ..."}**
-    -   "... be activate rows ..." **\textcolor{purple}{to "... will activate rows ..."}**
-    -   "... highlighted *in* the ..." **\textcolor{purple}{to "... highlighted on the map"}**
-    -   ". . . using 2020 measurements using match_sites() function." \**textcolor{purple}{to "... using the match\_sites() function"}**
-    -   "An example of this using is included in the Appendix" **\textcolor{purple}{to "... this use ..."}**
-    -   ". . . the data in -a- multiple -of- ways on-the-fly" **\textcolor{purple}{to "... in multiple ways ..."}**
+    -   "*components* spatio-temporal *components*" **\textcolor{purple}{Fixed.}**
+    -   "... fits works ..." **\textcolor{purple}{Fixed.}**
+    -   "... be activate rows ..." **\textcolor{purple}{Fixed.}**
+    -   "... highlighted *in* the ..." **\textcolor{purple}{Fixed.}**
+    -   ". . . using 2020 measurements using match_sites() function." **\textcolor{purple}{Fixed.}**
+    -   "An example of this using is included in the Appendix" **\textcolor{purple}{Fixed.}**
+    -   ". . . the data in -a- multiple -of- ways on-the-fly" **\textcolor{purple}{Fixed.}**
 
--   Surprising references: in Section 4.4 it says ".
-.
-.
-Glyph maps (Section 3.4)", but Section 3.4 says ".
-.
-.
-glyph maps will be explained in Section 4.4".
-**\textcolor{purple}{the looped reference in section 3.3 and 3.4 is replaced with (Wickham et al.
-2012)}.**
+-   Surprising references: in Section 4.4 it says ". . . Glyph maps (Section 3.4)", but Section 3.4 says ". . . glyph maps will be explained in Section 4.4". **\textcolor{purple}{Fixed.}**
 
--   Should ". . . it's temperatures are more *consistent*" rather be ". . . it's temperatures are more *constant*"? **\textcolor{purple}{to "more constant"}**
+-   Should ". . . it's temperatures are more *consistent*" rather be ". . . it's temperatures are more *constant*"? **\textcolor{purple}{Changed.}**
 
@@
 
-    -   ". . . spatial and *tmeporal* information are available." **\textcolor{purple}{to "temporal"}**
-    -   The polar vortex, signalled by the high *speicfic* humidity, splits into two on 2002-09-26 and *further-s- split_s\_* into four on 2002-10-04.
-**\textcolor{purple}{to "specific", "further splits"}**
-    -   the data in *-a- multiple -of-* ways on-the-fly. **\textcolor{purple}{ to "multiple ways"}**
+    -   ". . . spatial and *tmeporal* information are available." **\textcolor{purple}{Fixed.}**
+    -   The polar vortex, signalled by the high *speicfic* humidity, splits into two on 2002-09-26 and *further-s- split_s\_* into four on 2002-10-04. **\textcolor{purple}{Fixed.}**
+    -   the data in *-a- multiple -of-* ways on-the-fly. **\textcolor{purple}{Fixed.}**
 
@@
 The tmax\_hist data, which was included in the package, is now removed, given its similarity to the climate\_aus data, which is more useful.
-We provide the script to generate historical\_tmax in the Appendix and a copy of the data is available in the paper GitHub repository: https://github.com/huizezhang-sherry/paper-cubble.}**
+The scripts to generate historical\_tmax are provided with the paper.}**
 
 The manuscript cannot be reproduced, as the reproducible-script.R (part of the submission to JSS) only uses a subset, but does not provide a separate result file that would allow to compare the reproduced outputs with the desired output for the subset.
-**\textcolor{purple}{A result file in HTML is submitted this time for result comparison.}**
+**\textcolor{purple}{A result file in HTML is submitted along with the reproducible-script.R file.}**

### 2023-06-20 2-cubble.Rmd
nested/long all to spatial/temporal cubble

--- 2023-06-18 2-cubble.Rmd
@@
 
-The cubble class is an S3 class built on tibble that allows the spatio-temporal data to be wrangled in two forms:  a nested/ spatialform and a long/temporal form. It consists of two subclasses:
+The cubble class is an S3 class built on tibble that allows the spatio-temporal data to be wrangled in two forms:  a nested form (spatial cubble) and a long form (temporal cubble).
+It consists of two subclasses:
 
--   a nested/ spatial cubble with class `c("spatial_cubble_df", "cubble_df")`
--   a long/ temporal cubble with class `c("temporal_cubble_df", "cubble_df")`
+-   a spatial cubble with class `c("spatial_cubble_df", "cubble_df")`
+-   a emporal cubble with class `c("temporal_cubble_df", "cubble_df")`
 
-In a nested cubble, spaital variables are organised as columns and temporal variables are nested within a specialised `ts` column.
-The nested cubble object printed below contains weather records in three airport stations (the creation of a cubble object will be explained in Section \@ref(create)).
+In a spatial cubble, spaital variables are organised as columns and temporal variables are nested within a specialised `ts` column.
+The cubble object printed below contains weather records in three airport stations (the creation of a cubble object will be explained in Section \@ref(create)).
 This toy data is a subset of a larger data \code{climate_aus} collected from Global Historical Climatology Network Daily (GHCND).
@@
 
-In a long cubble, temporal variables are expanded in the long form and spatial variables are stored as a data attribute:
+In a temporal cubble, temporal variables are expanded in the long form and spatial variables are stored as a data attribute:
 
@@
 Readers who are familiar with the `key` and `index` attributes from the `tsibble` package would already understand the two arguments.
-In cubble, the `key` attribute identifies the row in the nested cubble (given the use of `tidyr::nest()` for nesting), and when combined with the `index` argument, it identifies the row in the long cubble.
+In cubble, the `key` attribute identifies the row in the spatial cubble (given the use of `tidyr::nest()` for nesting), and when combined with the `index` argument, it identifies the row in the temporal cubble.
 Currently, cubble only supports one variable as the key, and the accepted temporal classes for index includes the base R classes `Date`, `POSIXlt`, `POSIXct`, as well as tsibble's `yearmonth`, `yearweek`, and `yearquarter` classes.
@@
 It can be either an unprojected pair of longitude and latitude, or a projected easting and northing value.
-The `sf` package is used under the hood to calculate the bounding box, displayed in the header of a nested cubble.
+The `sf` package is used under the hood to calculate the bounding box, displayed in the header of a spatial cubble.
 
-The long cubble has a special attribute called `spatial` to store the spatial variables, which includes all the variables from the nested cubble, except for the `ts` column.
-Shortcut functions are available to extract attributes, for example, \code{spatial()} for extracting spatial variables from the long cubble:
+The temporal cubble has a special attribute called `spatial` to store the spatial variables.
+Shortcut functions are available to extract attributes, for example, \code{spatial()} for extracting spatial variables from the temporal cubble:
 
@@
 The function \code{make_cubble()} composes a \code{cubble} object from a spatial table (\code{spatial}) and a temporal table (\code{temporal}), along with three attributes introduced in the subsection \@ref(object): \code{key}, \code{index}, and \code{coords}.
-The following code creates the nested \code{cubble} object shown previously:
+The following code creates the spatial \code{cubble} object shown previously:
 
@@
 Table \@ref(tab:funs) summarises the functions implemented in the \pkg{cubble} package and Table \@ref(tab:methods) details the methods implemented for each of the three cubble classes.
-The \code{cubble_df} class handles methods that behave consistently in both nested (\code{spatial_cubble_df}) and long cubble (\code{temproal_cubble_df}).
-When the method has distinct behavior for the nested cubble  and temporal cubble, it is implemented separately for each subclass.
+The \code{cubble_df} class handles methods that behave consistently in both spatial (\code{spatial_cubble_df}) and temporal cubble (\code{temporal_cubble_df}).
+When the method has distinct behavior for the spatial cubble and temporal cubble, it is implemented separately for each subclass.
 
@@
 Specifically, the pair of verbs, \code{face_temporal()} and \code{face_spatial()}, pivot the cubble object between the spatial and temporal face of the multivariate spatio-temporal cube, as illustrated in Figure \ref{fig:face}.
-The code below uses the function \code{face_temporal()} and \code{face_spatial()} to transform between a nested cubble (\code{cb_nested}) and a long one (\code{cb_long}), introduced earlier in subsection \@ref(object):
+The code below uses the function \code{face_temporal()} and \code{face_spatial()} to transform between a spatial cubble (\code{cb_nested}) and a temporal one (\code{cb_long}), introduced earlier in subsection \@ref(object):
 
@@
 
-The long cubble shows \code{[tsibble]} in the header to indicate the object also being in a \code{tbl_ts} class (\code{tsibble}).
+The temporal cubble shows \code{[tsibble]} in the header to indicate the object also being in a \code{tbl_ts} class (\code{tsibble}).
 Methods applies to the \code{tbl_ts} class can also be applied to the temporal cubble objects, for example, checking whether the data contain temporal gaps:

### 2023-06-20 3-advanced.Rmd
nested/long all to spatial/temporal cubble

@@
 Figure \ref{fig:illu-interactive} illustrates the linking between a map and multiple time series in a \code{cubble} object.
-Both the map, generated from the nested cubble, and the time series plot, generated from the long cubble, are shared crosstalk objects.
+Both the map, generated from the spatial cubble, and the time series plot, generated from the temporal cubble, are shared crosstalk objects.
 When a user selects a location on the map (a), the corresponding site is highlighted.
-This selection activates a row in the nested \code{cubble}, which is then communicated to the long \code{cubble}, resulting in teh selection of all observations with the same ID in (b).
-Consequently the long \code{cubble} highlights the corresponding series in the time series plot (c).
+This selection activates a row in the spatial \code{cubble}, which is then communicated to the temporal \code{cubble}, resulting in the selection of all observations with the same ID in (b).
+Consequently the temporal \code{cubble} highlights the corresponding series in the time series plot (c).
 Linking can also be initiated from the time series plot, by selecting points on the time series.
-This action activates rows with the same ID in the long \code{cubble} and the row in the nested \code{cubble}, which will be highlighted on the map.
+This action activates rows with the same ID in the temporal \code{cubble} and the row in the spatial \code{cubble}, which will be highlighted on the map.
 Please refer to the appendix for the corresponding illustration.
@@
 ```{r illu-interactive, echo = FALSE, fig.align="center", out.height = "35%", out.width = "100%", fig.cap = "Linking between multiple plots.
-The line plots and the map are constructed from shared \\code{crosstalk} objects (long and nested \\code{cubble}).
-When a station is selected on the map (a), the corresponding row in the nested \\code{cubble} will be activated.
-This will link to all the rows with the same id in the long \\code{cubble} (b) and update the line plot (c)."}
+The line plots and the map are constructed from shared \\code{crosstalk} objects.
+When a station is selected on the map (a), the corresponding row in the spatial \\code{cubble} will be activated.
+This will link to all the rows with the same id in the temporal \\code{cubble} (b) and update the line plot (c)."}
 knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.002.png"))

### 2023-06-20 4-2-glyph.Rmd
nested/long all to spatial/temporal cubble

@@
 The NSW stations correspond to entries in the range 46-75 and the Victorian stations to 76-90.
-Filtering Victoria and NSW stations is a *spatial operation* and hence uses the nested \code{cubble}:
+Filtering Victoria and NSW stations is a *spatial operation* and hence uses the spatial \code{cubble}:
 
@@
 Next, the monthly maximum average temperature is  calculated for both periods.
-This is a *temporal operation* requiring a switch into the long \code{cubble} using the \code{face_temporal()} function.
+This is a *temporal operation* requiring a switch into the temporal \code{cubble} using the \code{face_temporal()} function.
 In addition, a new indicator for the two time periods of interest is created before the calculation of monthly averages:
@@
 A quick check on the number of observations for each location is made, revealing that there are several with less than 24 observations -- these stations lack temperature values for some months.
-In this example, those stations are removed by switching to the nested \code{cubble} to operate on the spatial component over time, and then, move back into the long \code{cubble} (to make the glyph map):
+In this example, those stations are removed by switching to the spatial \code{cubble} to operate on the spatial component over time, and then, move back into the temporal \code{cubble} (to make the glyph map):
 

### 2023-06-20 jss/main.tex
nested/long all to spatial/temporal cubble

@@
 
-The cubble class is an S3 class built on tibble that allows the spatio-temporal data to be wrangled in two forms: a nested/ spatial form and a long/temporal form. It consists of two subclasses:
+The cubble class is an S3 class built on tibble that allows the spatio-temporal data to be wrangled in two forms: a nested form (spatial cubble) and a long form (temporal cubble).
+It consists of two subclasses:
 
@@
 \item
-  a nested/ spatial cubble with class \texttt{c("spatial\_cubble\_df",\ "cubble\_df")}
+  a spatial cubble with class \texttt{c("spatial\_cubble\_df",\ "cubble\_df")}
 \item
-  a long/ temporal cubble with class \texttt{c("temporal\_cubble\_df",\ "cubble\_df")}
+  a emporal cubble with class \texttt{c("temporal\_cubble\_df",\ "cubble\_df")}
 \end{itemize}
 
-In a nested cubble, spaital variables are organised as columns and temporal variables are nested within a specialised \texttt{ts} column.
-The nested cubble object printed below contains weather records in three airport stations (the creation of a cubble object will be explained in Section \ref{create}).
+In a spatial cubble, spaital variables are organised as columns and temporal variables are nested within a specialised \texttt{ts} column.
+The cubble object printed below contains weather records in three airport stations (the creation of a cubble object will be explained in Section \ref{create}).
 This toy data is a subset of a larger data \code{climate_aus} collected from Global Historical Climatology Network Daily (GHCND).
@@
 Readers who are familiar with the \texttt{key} and \texttt{index} attributes from the \texttt{tsibble} package would already understand the two arguments.
-In cubble, the \texttt{key} attribute identifies the row in the nested cubble (given the use of \texttt{tidyr::nest()} for nesting), and when combined with the \texttt{index} argument, it identifies the row in the long cubble.
+In cubble, the \texttt{key} attribute identifies the row in the spatial cubble (given the use of \texttt{tidyr::nest()} for nesting), and when combined with the \texttt{index} argument, it identifies the row in the temporal cubble.
 Currently, cubble only supports one variable as the key, and the accepted temporal classes for index includes the base R classes \texttt{Date}, \texttt{POSIXlt}, \texttt{POSIXct}, as well as tsibble's \texttt{yearmonth}, \texttt{yearweek}, and \texttt{yearquarter} classes.
@@
 It can be either an unprojected pair of longitude and latitude, or a projected easting and northing value.
-The \texttt{sf} package is used under the hood to calculate the bounding box, displayed in the header of a nested cubble.
-
-The long cubble has a special attribute called \texttt{spatial} to store the spatial variables, which includes all the variables from the nested cubble, except for the \texttt{ts} column.
-Shortcut functions are available to extract attributes, for example, \code{spatial()} for extracting spatial variables from the long cubble:
+The \texttt{sf} package is used under the hood to calculate the bounding box, displayed in the header of a spatial cubble.
+
+The temporal cubble has a special attribute called \texttt{spatial} to store the spatial variables.
+Shortcut functions are available to extract attributes, for example, \code{spatial()} for extracting spatial variables from the temporal cubble:
 
@@
 The function \code{make_cubble()} composes a \code{cubble} object from a spatial table (\code{spatial}) and a temporal table (\code{temporal}), along with three attributes introduced in the subsection \ref{object}: \code{key}, \code{index}, and \code{coords}.
-The following code creates the nested \code{cubble} object shown previously:
+The following code creates the spatial \code{cubble} object shown previously:
 
@@
 Table \ref{tab:funs} summarises the functions implemented in the \pkg{cubble} package and Table \ref{tab:methods} details the methods implemented for each of the three cubble classes.
-The \code{cubble_df} class handles methods that behave consistently in both nested (\code{spatial_cubble_df}) and long cubble (\code{temproal_cubble_df}).
-When the method has distinct behavior for the nested cubble and temporal cubble, it is implemented separately for each subclass.
+The \code{cubble_df} class handles methods that behave consistently in both spatial (\code{spatial_cubble_df}) and temporal cubble (\code{temporal_cubble_df}).
+When the method has distinct behavior for the spatial cubble and temporal cubble, it is implemented separately for each subclass.
 
@@
 Specifically, the pair of verbs, \code{face_temporal()} and \code{face_spatial()}, pivot the cubble object between the spatial and temporal face of the multivariate spatio-temporal cube, as illustrated in Figure \ref{fig:face}.
-The code below uses the function \code{face_temporal()} and \code{face_spatial()} to transform between a nested cubble (\code{cb_nested}) and a long one (\code{cb_long}), introduced earlier in subsection \ref{object}:
+The code below uses the function \code{face_temporal()} and \code{face_spatial()} to transform between a spatial cubble (\code{cb_nested}) and a temporal one (\code{cb_long}), introduced earlier in subsection \ref{object}:
 
@@
 \caption{Linking between multiple plots.
-The line plots and the map are constructed from shared \code{crosstalk} objects (long and nested \code{cubble}).
-When a station is selected on the map (a), the corresponding row in the nested \code{cubble} will be activated.
-This will link to all the rows with the same id in the long \code{cubble} (b) and update the line plot (c).}\label{fig:illu-interactive}
+The line plots and the map are constructed from shared \code{crosstalk} objects.
+When a station is selected on the map (a), the corresponding row in the spatial \code{cubble} will be activated.
+This will link to all the rows with the same id in the temporal \code{cubble} (b) and update the line plot (c).}\label{fig:illu-interactive}
 \end{figure}
@@
 The NSW stations correspond to entries in the range 46-75 and the Victorian stations to 76-90.
-Filtering Victoria and NSW stations is a \emph{spatial operation} and hence uses the nested \code{cubble}:
+Filtering Victoria and NSW stations is a \emph{spatial operation} and hence uses the spatial \code{cubble}:
 
@@
 Next, the monthly maximum average temperature is calculated for both periods.
-This is a \emph{temporal operation} requiring a switch into the long \code{cubble} using the \code{face_temporal()} function.
+This is a \emph{temporal operation} requiring a switch into the temporal \code{cubble} using the \code{face_temporal()} function.
 In addition, a new indicator for the two time periods of interest is created before the calculation of monthly averages:

### 2023-06-20 2-cubble.Rmd
edit on the ERA5 example

@@
 
-### Creating from separate spatial and temporal tables
-
 In many cases, spatio-temporal data arrive in separate tables for analysis.
@@
 
-### Coercing from foreign objects
-
 Spatio-temporal data in other foreign objects can be coerced into a \code{cubble} object with the function \code{as_cubble()}.
@@
 
-```{r echo = TRUE, eval = FALSE}
+```{r echo = TRUE}
 climate_flat |> as_cubble(key = id, index = date, coords = c(long, lat))
-```
-
-In \proglang{R}, there are several packages available for wrangling NetCDF data, including \code{ncdf4}, \code{RNetCDF}, and \code{tidync}.
-The code below converts a NetCDF object of class \code{ncdf4} into a cubble object:
-
-```{r echo = TRUE}
-path <- system.file("ncdf/era5-pressure.nc", package = "cubble")
-raw <- ncdf4::nc_open(path)
-as_cubble(raw)
-```
-
-Sometimes, analysts may choose to read onlny a subset of the NetCDF data.
-In such cases, the arguments \code{vars}, \code{long_range} and \code{lat_range} can be used to subset the data based on variable and the grid resolution:
-
-```{r echo = TRUE, eval = FALSE}
-as_cubble(raw, vars = "q",
-          long_range = seq(-180, 180, 1), lat_range = seq(-90, 90, 1))
 ```

### 2023-06-20 4-4-netcdf.Rmd
edit on the ERA5 example

@@
 
-The ERA5 data [@hersbach2020era5] provides hourly estimates across the Earth for atmospheric, land and oceanic climate variables.
-The data is available in the NetCDF format from the European Centre for Medium-Range Weather Forecasts (ECMWF) and can be downloaded from Copernicus Climate Data Store (CDS) [@cds] website or via the \pkg{ecmwfr} package [@ecwmfr].
-This example reproduces the ERA5 data row of Figure 19 in @hersbach2020era5 to show the southern polar vortex splitting into two on 2002-09-26 and further splitting into four on 2002-10-04.
+The ERA5 reanalysis [@hersbach2020era5] provides hourly estimates of atmospheric, land and oceanic climate variables on a global scale.
+The hourly pressure level data in the NetCDF format can be downloaded from Copernicus Climate Data Store (CDS) [@cds] or via the \pkg{ecmwfr} package [@ecwmfr].
+This example reproduces the row created from ERA5 reanalysis data shown in Figure 19 by @hersbach2020era5.
+The plot shows the southern polar vortex splitting into two on 2002-09-26, and further splitting into four on 2002-10-04.
 Further explanation of why this is interesting can be found in the figure source, and also in @simmons2020global and @simmons2005ecmwf.
 
-For the reproduction, we focus on the \code{era5-pressure} data, hourly pressure levels from 1970 to present, with the *specific humidity* and *geopotential*.
-The downloaded NetCDF data (`raw`) is first converted to a \code{cubble} object:
+To work with NetCDF data in R, several packages are available, including \code{ncdf4}, \code{RNetCDF}, and \code{tidync}.
+The following code converts a NetCDF object of class \code{ncdf4} into a cubble object:
 
-<!--Here we specifically look at the information  contained on the variables *specific humidity* measuring the mass of water vapor per kilogram of moist air  and the *geopotential* variable recording the gravitational potential energy of a unit mass, at a particular location, relative to mean sea level.-->
-
-<!--The goal of this example is to reproduce the break-up of the southern polar vortex that happened in late September and early October in 2002 in the stratosphere in the Southern Hemisphere.
-Therefore, we filter data corresponding to the Southern Hemisphere and restrict our analysis to four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04.-->
-```{r echo = FALSE, eval = TRUE}
+```{r echo = TRUE}
 raw <- ncdf4::nc_open(here::here("data/era5-pressure.nc"))
@@
 
+Analysts can extract a subset of the NetCDF data with the arguments \code{vars}, \code{long_range} and \code{lat_range}.
+In this example, the variables q (specific humidity) and z (geopotential) are selected and the coordinates are subsetted to every degree in longitude and latitude:
+
 ```{r echo = TRUE}
-dt <- as_cubble(
-  raw, vars = c("q", "z"),
-  long_range = seq(-180, 180, 1), lat_range = seq(-88, 88, 1))
+(dt <- as_cubble(raw, vars = c("q", "z"),
+                 long_range = seq(-180, 180, 1), lat_range = seq(-88, -15, 1)))
 ```
 
-<!-- Figure \ref{fig:netcdf} reproduces the ERA5 data row of Figure 19 in @hersbach2020era5.
-It shows the southern polar vortex splits into two on 2002-09-26 and further splits into four on 2002-10-04 in the stratosphere.
-Readers interested in the analysis of this figure can refer to @hersbach2020era5, @simmons2020global, and @simmons2005ecmwf for more details.-->
+Once the NetCDF data is coerced into a cubble object, subsequent analysis can be conducted to filter on the date of interest, scale the variable specific humidity and create visualisation in ggplot as in Figure \ref{fig:netcdf}.
 
-Creating the plot requires making transformations on time, unfolding the data for computing the statistic of interest, which is plotted directly as a contour plot with ggplot.
+```{r netcdf, out.width="100%", fig.height = 4, fig.width = 10, fig.cap = "A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020) to illustrate the break-up of sourthern polar vortex in late September and early October 2002.
+The polar vortex, signalled by the high specific humidity, splits into two on 2002-09-26 and further splits into four on 2002-10-04."}
+date <- c("2002-09-22", "2002-09-26", "2002-09-30", "2002-10-04") |> as.Date()
+res <- dt |> 
+  face_temporal() |> 
+  filter(lubridate::date(time) %in% date) |>
+  unfold(long, lat) |> 
+  mutate(q = q* 10^6)
+
+con <- rnaturalearth::ne_coastline("small", returnclass = "sf")
+box <- st_bbox(c(xmin = -180, ymin = -90, xmax = 180, ymax = -15), crs = st_crs(con)) 
+sf::sf_use_s2(FALSE)
+country <- con %>% st_geometry() %>% 
+  sf::st_crop(box) |> 
+  sf::st_cast("MULTILINESTRING")
+
+res |> 
+  ggplot() +
+  # q for specific humidity
+  geom_point(aes(x = long, y = lat, color = q)) + 
+  #geom_tile(aes(x = long, y = lat, fill = q), color = "transparent") +
+  # z for geopotential
+  geom_contour(data = res, aes(x = long, y = lat, z = z),
+               color = "grey20", binwidth = 4000, linewidth = 0.5) +
+  geom_sf(data = country , alpha = 0.5, fill = "transparent", color = "lightgreen") +
+  coord_sf(
+    crs = "+proj=stere +lat_0=-90 +lon_0=-180 +k=1 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs", default_crs = sf::st_crs(4326), clip = "off") +
+  theme_bw() +
+  theme(legend.position = "bottom",
+        axis.ticks = element_blank(),
+        panel.border = element_blank(),
+        axis.text = element_blank(),
+        panel.grid = element_blank()) +
+  facet_wrap(vars(as.Date(time)), nrow = 1)+  
+  colorspace::scale_color_continuous_sequential("Purple-Yellow", name = "Specific humidity") + 
+  theme(legend.text = element_text(size = 13),
+        legend.title = element_text(size = 13)) + 
+  labs(x ="", y = "")
+```
 
@@
 ```
-
-```{r netcdf, out.width="100%", fig.height = 4, fig.width = 10, fig.cap = "A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020) to illustrate the break-up of sourthern polar vortex in late September and early October 2002.
-The polar vortex, signalled by the high specific humidity, splits into two on 2002-09-26 and further splits into four on 2002-10-04."}
-library(sf)
-date <- c("2002-09-22", "2002-09-26", "2002-09-30", "2002-10-04") |> as.Date()
-res <- dt |> 
-  face_temporal() |> 
-  filter(lubridate::date(time) %in% date) |>
-  unfold(long, lat) |> 
-  mutate(q =  q* 10^6)
-
-con <- rnaturalearth::ne_coastline("small", returnclass = "sf")
-box <- st_bbox(c(xmin = -180, ymin = -90, xmax = 180, ymax = -15), crs = st_crs(con)) 
-sf::sf_use_s2(FALSE)
-country <- con %>% st_geometry() %>% 
-  sf::st_crop(box) |> 
-  sf::st_cast("MULTILINESTRING")
-
-res |> 
-  ggplot() +
-  # q for specific humidity
-  geom_point(aes(x = long, y = lat, color = q), size = 2) + 
-  #geom_tile(aes(x = long, y = lat, fill = q), color = "transparent") +
-  # z for geopotential
-  geom_contour(data = res, aes(x = long, y = lat, z = z),
-               color = "grey20", binwidth = 4000, linewidth = 0.5) +
-  geom_sf(data = country , alpha = 0.5, fill = "transparent", color = "lightgreen") +
-  coord_sf(
-    crs = "+proj=stere +lat_0=-90 +lon_0=-180 +k=1 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs ", default_crs = sf::st_crs(4326), clip = "off") +
-  theme_bw() +
-  theme(legend.position = "bottom",
-        axis.ticks = element_blank(),
-        panel.border = element_blank(),
-        axis.text = element_blank(),
-        panel.grid = element_blank()) +
-  facet_wrap(vars(as.Date(time)), nrow = 1)+  
-  colorspace::scale_color_continuous_sequential("Purple-Yellow", name = "Specific humidity") + 
-  theme(legend.text = element_text(size = 13),
-        legend.title = element_text(size = 13)) + 
-  labs(x ="", y = "")
-```

### 2023-06-20 jss/main.tex
edit on the ERA5 example

@@
 
-\hypertarget{creating-from-separate-spatial-and-temporal-tables}{%
-\subsubsection{Creating from separate spatial and temporal tables}\label{creating-from-separate-spatial-and-temporal-tables}}
-
 In many cases, spatio-temporal data arrive in separate tables for analysis.
@@
 \end{CodeChunk}
-
-\hypertarget{coercing-from-foreign-objects}{%
-\subsubsection{Coercing from foreign objects}\label{coercing-from-foreign-objects}}
 
@@
 \end{CodeInput}
-\end{CodeChunk}
-
-In \proglang{R}, there are several packages available for wrangling NetCDF data, including \code{ncdf4}, \code{RNetCDF}, and \code{tidync}.
-The code below converts a NetCDF object of class \code{ncdf4} into a cubble object:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> path <- system.file("ncdf/era5-pressure.nc", package = "cubble")
-R> raw <- ncdf4::nc_open(path)
-R> as_cubble(raw)
-\end{CodeInput}
-\begin{CodeOutput}
-# cubble:   key: id [26565], index: time, nested form
-# spatial:  [113, -53, 153, -12], Missing CRS!
-# temporal: time [dttm], q [dbl], z [dbl]
-     id  long   lat ts              
-  <int> <dbl> <dbl> <list>          
-1     1  113    -12 <tibble [6 x 3]>
-2     2  113.   -12 <tibble [6 x 3]>
-3     3  114.   -12 <tibble [6 x 3]>
-4     4  114.   -12 <tibble [6 x 3]>
-5     5  114    -12 <tibble [6 x 3]>
-# i 26,560 more rows
-\end{CodeOutput}
-\end{CodeChunk}
-
-Sometimes, analysts may choose to read onlny a subset of the NetCDF data.
-In such cases, the arguments \code{vars}, \code{long_range} and \code{lat_range} can be used to subset the data based on variable and the grid resolution:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> as_cubble(raw, vars = "q",
-+           long_range = seq(-180, 180, 1), lat_range = seq(-90, 90, 1))
-\end{CodeInput}
+\begin{CodeOutput}
+# cubble:   key: id [3], index: date, nested form
+# spatial:  [144.8321, -37.98, 145.0964, -37.6655], Missing CRS!
+# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
+  id           long   lat  elev name              wmo_id ts               
+  <chr>       <dbl> <dbl> <dbl> <chr>              <dbl> <list>           
+1 ASN00086038  145. -37.7  78.4 essendon airport   95866 <tibble [10 x 4]>
+2 ASN00086077  145. -38.0  12.1 moorabbin airport  94870 <tibble [10 x 4]>
+3 ASN00086282  145. -37.7 113.  melbourne airport  94866 <tibble [10 x 4]>
+\end{CodeOutput}
 \end{CodeChunk}
@@
 ! Use `check_key()` to check on the unmatched key
-The cubble is created only with sites having both spatial and temporal information
+The cubble is created only with sites having both spatial and
+temporal information
 \end{CodeOutput}
@@
 
-The ERA5 data \citep{hersbach2020era5} provides hourly estimates across the Earth for atmospheric, land and oceanic climate variables.
-The data is available in the NetCDF format from the European Centre for Medium-Range Weather Forecasts (ECMWF) and can be downloaded from Copernicus Climate Data Store (CDS) \citep{cds} website or via the \pkg{ecmwfr} package \citep{ecwmfr}.
-This example reproduces the ERA5 data row of Figure 19 in \citet{hersbach2020era5} to show the southern polar vortex splitting into two on 2002-09-26 and further splitting into four on 2002-10-04.
+The ERA5 reanalysis \citep{hersbach2020era5} provides hourly estimates of atmospheric, land and oceanic climate variables on a global scale.
+The hourly pressure level data in the NetCDF format can be downloaded from Copernicus Climate Data Store (CDS) \citep{cds} or via the \pkg{ecmwfr} package \citep{ecwmfr}.
+This example reproduces the row created from ERA5 reanalysis data shown in Figure 19 by \citet{hersbach2020era5}.
+The plot shows the southern polar vortex splitting into two on 2002-09-26, and further splitting into four on 2002-10-04.
 Further explanation of why this is interesting can be found in the figure source, and also in \citet{simmons2020global} and \citet{simmons2005ecmwf}.
 
-For the reproduction, we focus on the \code{era5-pressure} data, hourly pressure levels from 1970 to present, with the \emph{specific humidity} and \emph{geopotential}.
-The downloaded NetCDF data (\texttt{raw}) is first converted to a \code{cubble} object:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> dt <- as_cubble(
-+   raw, vars = c("q", "z"),
-+   long_range = seq(-180, 180, 1), lat_range = seq(-88, 88, 1))
-\end{CodeInput}
-\end{CodeChunk}
-
-Creating the plot requires making transformations on time, unfolding the data for computing the statistic of interest, which is plotted directly as a contour plot with ggplot.
+To work with NetCDF data in R, several packages are available, including \code{ncdf4}, \code{RNetCDF}, and \code{tidync}.
+The following code converts a NetCDF object of class \code{ncdf4} into a cubble object:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> raw <- ncdf4::nc_open(here::here("data/era5-pressure.nc"))
+\end{CodeInput}
+\end{CodeChunk}
+
+Analysts can extract a subset of the NetCDF data with the arguments \code{vars}, \code{long_range} and \code{lat_range}.
+In this example, the variables q (specific humidity) and z (geopotential) are selected and the coordinates are subsetted to every degree in longitude and latitude:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> (dt <- as_cubble(raw, vars = c("q", "z"),
++                  long_range = seq(-180, 180, 1), lat_range = seq(-88, -15, 1)))
+\end{CodeInput}
+\begin{CodeOutput}
+# cubble:   key: id [26640], index: time, nested form
+# spatial:  [-180, -88, 179, -15], Missing CRS!
+# temporal: time [date], q [dbl], z [dbl]
+     id  long   lat ts              
+  <int> <dbl> <dbl> <list>          
+1     1  -180   -15 <tibble [8 x 3]>
+2     2  -179   -15 <tibble [8 x 3]>
+3     3  -178   -15 <tibble [8 x 3]>
+4     4  -177   -15 <tibble [8 x 3]>
+5     5  -176   -15 <tibble [8 x 3]>
+# i 26,635 more rows
+\end{CodeOutput}
+\end{CodeChunk}
+
+Once the NetCDF data is coerced into a cubble object, subsequent analysis can be conducted to filter on the date of interest, scale the variable specific humidity and create visualisation in ggplot as in Figure \ref{fig:netcdf}.
 

### 2023-06-21 main.Rmd
compare array with cubble, different formats spatial and temporal data are stored
%>% to |>
merge all the Rmds into a single oneg

@@
   - name: Dianne Cook
-    affiliation: 'Monash University \AND'
+    affiliation: Monash University
     address: 21 Chancellors Walk, Clayton VIC 3800 Australia
-    # location: Melbourne, Australia
     email: dicook@monash.edu
   - name: Ursula Laa
-    affiliation: 'University of Natural Resources and Life Sciences \AND'
+    affiliation: 'Universität für Bodenkultur Wiens \AND'
     address: Gregor-Mendel-Straße 33, 1180 Wien, Austria
@@
   - name: Nicolas Langrené
-    affiliation: 'BNU-HKBU United International College \AND'
+    affiliation: BNU-HKBU United International College
     address: 2000 Jintong Road, Tangjiawan, Zhuhai, Guangdong Province, China
@@
   - name: Patricia Menéndez
-    affiliation: 'Monash University \AND'
+    affiliation: Monash University
     address: 21 Chancellors Walk, Clayton VIC 3800 Australia
@@
 title:
-  formatted: "\\pkg{cubble}: An R Package for Structuring Spatio-temporal Data"
+  formatted: "\\pkg{cubble}: An \\proglang{R} Package for Organizing and Wrangling Multivariate Spatio-temporal Data"
   # If you use tex in the formatted title, alsoge supply version without
-  plain:     "cubble: An R Package for Structuring Spatio-temporal Data"
+  plain:     "cubble: An R Package for Organizing and Wrangling Multivariate Spatio-temporal Data"
   # For running headers, if needed
@@
 abstract: >
-  Spatio-temporal data refer to measurements taken across space and time.
-In practice, spatio-temporal data can be decomposed into a spatial and temporal component: at one time, we would select a spatial location and inspect the temporal trend; at other time, we might select one or multiple time value(s) and explore the spatial distribution.
-Ideally, we could make multiple maps and multiple time series to explore these together, however, doing all of these actions is complicated when data arrive fragmented in multiple objects.
-To make it easy to do all these tasks, ideally spatial and temporal variables are in a single data object that we can slice and dice in different ways to conduct different visualisations.
-This work suggests a new data structure, \code{cubble}, that uses a nested form and a long form to organise spatial and temporal variables in a single data object.
-The new data structure ensures that data in the two forms are synchronised while been flexible to explore the two aspect of spatio-temporal data.
-It also provides capability for handling data with hierarchical structure, matching data from multiple sources, constructing interactive graphics, and performing spatio-temporal transformation.
-The paper will demonstrate \pkg{cubble} with examples using Australian climate weather stations, river level data, and climate reanalysis (ERA5) data.
+     Multivariate spatio-temporal data refers to multiple measurements taken across space and time.
+For many analyses, spatial and time components can be separately studied: for example, to explore the temporal trend of one variable for a single spatial location, or to model the spatial distribution of one variable at a given time.
+However for some studies, it is important to analyze different aspects of the spatio-temporal data simultaneously, like for instance, temporal trends of multiple variables across locations.
+In order to facilitate the study of different portions or combinations of spatio-temporal data, we introduce a new class, \code{cubble}, with a suite of functions enabling easy slicing and dicing on different spatio-temporal components.
+The proposed \code{cubble} class ensures that all the components of the data are easy to access and manipulate while providing flexibility for data analysis.
+In addition, the \pkg{cubble} package facilitates visual and numerical explorations of the data while easing data wrangling and modelling.
+The \code{cubble} class and the functions provided in the \pkg{cubble} \proglang{R} package equip users with the capability to handle hierarchical spatial and temporal structures.
+The \code{cubble} class and the tools implemented in the package are illustrated with different examples of Australian climate data.
 keywords:
@@
   \usepackage{array}
-  \usepackage{xcolor}
 output: rticles::jss_article
@@
   cache = TRUE,
-  fig.path = "figures/",
-  fig.height = 10
+  fig.path = here::here("figures/")
 )
+
 options(prompt = "R> ", continue = "+ ",
-        tibble.print_max = 5, tibble.print_min = 5)
+        tibble.print_max = 5, tibble.print_min = 5,
+        width = 80, styler.cache_root = "styler-perm")
 ```
@@
 library(cubble)
-library(forcats)
-library(stringr)
-library(dplyr)
-library(lubridate)
-library(ggplot2)
-library(GGally)
-library(plotly)
-library(crosstalk)
+library(tidyverse)
 library(patchwork)
+library(tsibble)
+library(sf)
 ```
 
-\newpage
+# Introduction
 
-```{r child = "1-intro.Rmd"}
+Spatio-temporal data has a spatial component referring to the location of each observation and a temporal component that is recorded at regular or irregular time intervals.
+It may also include multiple variables measured at each spatial and temporal values.
+With spatio-temporal data, one can fix the time to explore the spatial features of the data, fix the spatial location/s to explore temporal aspects, or dynamically explore the space and time simultaneously.
+
+In order to computationally explore the spatial, temporal and spatio-temporal faces of such data, the data needs to be stored and represented under a specific data object that allows the user to query, group and dissect all the data faces.
+
+The Comprehensive \proglang{R} Archive Network (CRAN) task view SpatioTemporal [@ctvspatiotemporal] gathers information about \proglang{R} packages designed for spatio-temporal data and it has a section on *Representing data* that lists existing spatio-temporal data representations used in \proglang{R}.
+Among them, @spacetime summarises spatio-temporal data into three forms: time-wide, space-wide, and long formats.
+The associated package \pkg{spacetime} [@spacetime] implements four spatio-temporal layouts (full grid, sparse grid, irregular, and trajectory) to handle different space and time combinations.
+The package \pkg{stars} [@stars] has a new implementation to use dense arrays to represent spatio-temporal cubes.
+It also interfaces with the package \pkg{sf} [@sf], commonly used for wrangling spatial data, and the \pkg{tidyverse} [@tidyverse] suite for general data wrangling and visualization in \proglang{R}.
+
+Still, the data representation for spatio-temporal data can be further extended and there are two reasons for this.
+Firstly, the raw data sourced in the wild is less often presented in any one of the layouts above, and fitting the raw data into a data object can sometimes be difficult.
+More often, spatio-temporal data are collected in separate 2D tables and analysts need to assemble them into a whole piece before exploring the data.
+Examples of components of spatio-temporal data can be 1) areal data recording the shape of a collection of areas of interest; 2) geostatistical data storing the longitude and latitude coordinates of locations, typically also with other metadata related to the location, and; 3) temporal data of each location across time.
+
+The other reason is about tidy data concepts [@tidydata] and how they should be applied to spatio-temporal data.
+According to the tidy data principles, data should be structured into 1) one row per observation, 2) one column per variable, and 3) one type of data per table.
+The long form data is preferred over wide data form given the downstream packages such as \pkg{dplyr} [@dplyr] and \pkg{ggplot2} [@ggplot2] for data wrangling and visualization.
+However, the long form can be inefficient to store feature geometries, especially for large multipolygons for hourly, daily or sub-daily periods over years, which are extensively collected and handled, for example in time series analysis.
+This poses the question of how to arrange spatial and temporal variables in a way that would make data wrangling, visualizing and analyzing spatio-temporal data easier.
+
+This paper presents a new \proglang{R} package, \pkg{cubble}, which addresses the two issues mentioned above.
+In the package, a new class, also called cubble, is proposed to organize spatial and temporal variables as two forms of a single data object so that they can be wrangled separately or combined while being kept synchronized.
+Among the four spacetime layouts in @spacetime, the \code{cubble} class can be applied to full grid, sparse grid, or irregular, but not trajectory, which is outside the scope of this work.
+The software is available from the Comprehensive \proglang{R} Archive Network (CRAN) at https://CRAN.R-project.org/package=cubble.
+
+The rest of the paper is organized as follows: Section \@ref(cubble) presents the main design and functionality of the \pkg{cubble} package.
+Section \@ref(others) explains how the \pkg{cubble} package deals with more advanced considerations, including data with hierarchical structure, data matching and how the package fits with existing static and interactive visualization tools.
+Moreover we also illustrate how the \pkg{cubble} package deals with spatio-temporal data transformations.
+Section \@ref(examples) uses Australian weather station data and river level data as examples to demonstrate the use of the package.
+An example of how the \pkg{cubble} package handles Network Common Data Form (NetCDF) data is also provided.
+Section \@ref(conclude) discuss the paper contributions and future directions.
+
+# The cubble package {#cubble}
+
+## The cubble object {#object}
+
+Spatio-temporal data can encompass data with various spatial and temporal characteristics and different data require different structures for wrangling and analysis: climate weather stations record data at fixed point location but may suffer from potential temporal data quality issue such as missing data for certain days.
+GPS data tracks unique point locations at different timestamps and is represented as trajectories.
+Satellite imageries capture snapshots of landscapes at selected time and is commonly structured as raster data.
+The spatio-temporal data cubble address are those collected at unique fixed locations while allowing for irregularity in the temporal dimension, such as the weather station data.
+This corresponds to the full space-time and sparse space-time layouts described in the spacetime paper [@spacetime] and reproduced in Figure  \ref{fig:layout}.
+
+```{r layout, echo = FALSE, fig.cap="The full space-time and sparse space-time layouts handled by the cubble package.
+Reproduced from Figure 1 in the spacetime paper \\citep{spacetime}.", fig.width = 8, fig.height=4, fig.align='center'}
+library(tidyverse)
+library(patchwork)
+p1 <- tibble(
+  id = 1:12,
+  x = rep(1:4, each = 3),
+  y = rep(1:3, 4)
+) %>% 
+  ggplot(aes(x = as.factor(x) ,y = as.factor(y))) + 
+  geom_point() + 
+  geom_text(aes(y = y - 0.2, label = id)) + 
+  theme_bw() + 
+  theme(aspect.ratio = 1) + 
+  scale_x_discrete(labels = c("1st", "2nd", "3rd", "4th")) + 
+  scale_y_discrete(labels = c("1st", "2nd", "3rd", "4th")) + 
+  labs(x = "Time points", y = "Spatial features", title = "STF: full grid layout")
+
+p2 <- tibble(
+  id = 1:7,
+  x = c(1, 1, 1, 2, 2, 3, 4),
+  y = c(1, 2, 3, 2, 3, 1, 2),
+  label = c("1[1,1]", "2[1,2]", "3[1,3]", "4[2,2]", "5[3,2]", "6[1,3]", "7[2,4]")
+) %>% 
+  ggplot(aes(x = as.factor(x) ,y = as.factor(y))) + 
+  geom_point() + 
+  geom_text(aes(y = y - 0.2, label = label)) + 
+  theme_bw() + 
+  theme(aspect.ratio = 1) + 
+  scale_x_discrete(labels = c("1st", "2nd", "3rd", "4th")) + 
+  scale_y_discrete(labels = c("1st", "2nd", "3rd", "4th")) + 
+  labs(x = "Time points", y = "Spatial features", title = "STS: sparse grid layout")
+
+p1 | p2
+```
+
+The cubble class is an S3 class built on tibble that allows the spatio-temporal data to be wrangled in two forms:  a nested form (spatial cubble) and a long form (temporal cubble).
+It consists of two subclasses:
+
+-   a spatial cubble with class `c("spatial_cubble_df", "cubble_df")`
+-   a emporal cubble with class `c("temporal_cubble_df", "cubble_df")`
+
+In a spatial cubble, spaital variables are organised as columns and temporal variables are nested within a specialised `ts` column.
+The cubble object printed below contains weather records in three airport stations (the creation of a cubble object will be explained in Section \@ref(create)).
+This toy data is a subset of a larger data \code{climate_aus} collected from Global Historical Climatology Network Daily (GHCND).
+It records three airport stations located in Melbourne, Australia and includes spatial variables such as station ID, longitude, latitude, elevation, station name, World Meteorology Organisation ID.
+The dataset contains temporal variables including precipitation, maximum and minimum temperature, which can be read from the cubble header:
+
+```{r}
+cb_nested <- climate_mel
+```
+
+```{r echo = TRUE}
+cb_nested
+```
+
+<!-- A nested cubble header prints the key and index variable name and its form.
+The spatial extent is given by the bounding box with CRS information (missing here).
+The nested cubble header also shows the name of temporal variables nested in the \code{ts} column and their variable types: \code{prcp [dbl], tmax [dbl], tmin [dbl]}.
+-->
+
+In a temporal cubble, temporal variables are expanded in the long form and spatial variables are stored as a data attribute:
+
+```{r}
+cb_long <- face_temporal(climate_mel)
+```
+
+```{r echo = TRUE}
+cb_long
+```
+
+### The cubble attributes
+
+```{r class, fig.align="center", out.width = "100%", fig.cap = "Illustration of a cubble object in the long form and nested form, along with the associated attributes."}
+#knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.004.png"))
+```
+
+A cubble object inherits the attributes from tibble (and its subclasses): `class`, `row.names`, and `names`.
+Additionally, it has three specialised attributes: `key`, `index`, and `coords`.
+Readers who are familiar with the `key` and `index` attributes from the `tsibble` package would already understand the two arguments.
+In cubble, the `key` attribute identifies the row in the spatial cubble (given the use of `tidyr::nest()` for nesting), and when combined with the `index` argument, it identifies the row in the temporal cubble.
+Currently, cubble only supports one variable as the key, and the accepted temporal classes for index includes the base R classes `Date`, `POSIXlt`, `POSIXct`, as well as tsibble's `yearmonth`, `yearweek`, and `yearquarter` classes.
+
+The `coords` attribute represents an ordered pair of coordinates.
+It can be either an unprojected pair of longitude and latitude, or a projected easting and northing value.
+The `sf` package is used under the hood to calculate the bounding box, displayed in the header of a spatial cubble.
+
+The temporal cubble has a special attribute called `spatial` to store the spatial variables.
+Shortcut functions are available to extract attributes, for example, \code{spatial()} for extracting spatial variables from the temporal cubble:
+
+```{r echo = TRUE}
+spatial(cb_long)
+```
+
+## Creation and coercion {#create}
+
+In many cases, spatio-temporal data arrive in separate tables for analysis.
+For example, in climate data, analysts may initially receive station data containing geographic location information, recorded variables and their recording periods.
+They can then query the temporal variables using the stations of interest to obtain the relevant temporal data.
+Alternatively, analyses may begin as purely spatial or temporal, and analysts may obtain additional temporal or spatial data to expand the result to spatio-temporal.
+
+The function \code{make_cubble()} composes a \code{cubble} object from a spatial table (\code{spatial}) and a temporal table (\code{temporal}), along with three attributes introduced in the subsection \@ref(object): \code{key}, \code{index}, and \code{coords}.
+The following code creates the spatial \code{cubble} object shown previously:
+
+```{r echo = TRUE, eval = FALSE}
+make_cubble(spatial = stations, temporal = meteo,
+            key = id, index = date, coords = c(long, lat))
+```
+
+Spatio-temporal data in other foreign objects can be coerced into a \code{cubble} object with the function \code{as_cubble()}.
+This includes casting from a \code{tibble} or \code{data.frame} with both spatial and temporal information, a NetCDF object, a \code{stars} object [@stars], and a \code{sftime} object [@sftime].
+The two examples below show the casting from a tibble and a NetCDF object.
+The dataset \code{climate_flat} combines the spatial data, \code{stations}, with the temporal data, \code{meteo}, into a single tibble object.
+It can be coerced into a cubble using:
+
+```{r echo = TRUE}
+climate_flat |> as_cubble(key = id, index = date, coords = c(long, lat))
+```
+
+## Functions and methods
+
+Table \@ref(tab:funs) summarises the functions implemented in the \pkg{cubble} package and Table \@ref(tab:methods) details the methods implemented for each of the three cubble classes.
+The \code{cubble_df} class handles methods that behave consistently in both spatial (\code{spatial_cubble_df}) and temporal cubble (\code{temporal_cubble_df}).
+When the method has distinct behavior for the spatial cubble and temporal cubble, it is implemented separately for each subclass.
+
+| Category  | Functions |
+|--------|--------------------------------------------------------|
+| base R    | `[`, `[[<-`, `names<-` |
+| tidyverse | `dplyr_row_slice`, `dplyr_col_modify`, `dplyr_reconstruct`, `select`, `mutate`, `arrange`, `filter`, `group_by`, `ungroup`, `summarise`, `select`, `slice`, `rowwise`, `rename`, `bind_rows`, `bind_cols`, `relocate`, `type_sum`, the slice family (`slice_head`, `slice_tail`, `slice_max`, `slice_min`, `slice_sample`) and the join family (`left_join`, `right_join`, `inner_join`, `full_join`, `anti_join`, `semi_join`) |
+| cubble    | `as_cubble`, `cubble`, `make_cubble`, `check_key` `face_temporal`, `face_spatial`, `unfold`, `key`, `key_vars`, `key_data`, `index`, `index_var`, `coords`, `spatial`, `match_sites`, `match_spatial`, `match_temporal`, `geom_glyph`, `geom_glyph_box`, `geom_glyph_line`, `make_spatial_sf`, `make_temporal_tsibble`, `fill_gaps`, and `scan_gaps`                                                                                |
+
+: (#tab:funs) Functions implemented in the \pkg{cubble} package
+
+| Class                | Methods|
+|------------------|------------------------------------------------------|
+| `cubble_df`          | `[[<-, dplyr_col_modify, key_data, key_vars, key, print`                                                                                                                                                                                     |
+| `spatial_cubble_df`  | `[, names<-,  tbl_sum, dplyr_reconstruct, dplyr_row_slice, face_spatial, face_temporal, unfold, arrange, rename, rowwise, group_by, ungroup, select, spatial, summarise, unfold, update_cubble`                                              |
+| `temporal_cubble_df` | `[, names<-, tbl_sum, arrange, dplyr_reconstruct, dplyr_row_slice, face_spatial, face_temporal, unfold, fill_gaps, group_by, ungroup,  rename, rowwise, scan_gaps, select, spatial, summarise, tbl_sum, bind_rows, bind_cols, update_cubble` |
+
+: (#tab:methods) Methods implemented for the three \code{cubble} classes.
+
+Specifically, the pair of verbs, \code{face_temporal()} and \code{face_spatial()}, pivot the cubble object between the spatial and temporal face of the multivariate spatio-temporal cube, as illustrated in Figure \ref{fig:face}.
+The code below uses the function \code{face_temporal()} and \code{face_spatial()} to transform between a spatial cubble (\code{cb_nested}) and a temporal one (\code{cb_long}), introduced earlier in subsection \@ref(object):
+
+```{r cubble-fun, echo = TRUE}
+identical(face_temporal(cb_nested), cb_long)
+identical(face_spatial(cb_long), cb_nested)
+```
+
+The pair of verbs are exact inverse and apply both functions on a cubble object will result in the object itself:
+
+```{r cubble-fun2, echo = TRUE}
+identical(face_spatial(face_temporal(cb_nested)), cb_nested)
+identical(face_temporal(face_spatial(cb_long)), cb_long)
+```
+
+```{r face, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "An illustration of the function \\code{face\\_temporal()} and \\code{face\\_spatial()}: \\code{face\\_temporal()} converts a spatial cubble (neseted form) into a temporal cubble (long form) to focus on the temporal variables.
+Conversely, \\code{face\\_spatial()} transforms a temporal cubble into a spatial one to emphasize spatial variables."}
+knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.001.png"))
+```
+
+## Compatibility with tsibble and sf
+
+Analysts often have their preferred spatial or temporal data structure that they prefer to use for spatio-temporal analysis.
+For example, the \code{tbl_ts} class from the \code{tsibble} package [@tsibble] is commonly used in time series forecasting and similarly, the \code{sf} class [@sf] is frequently used in spatial data science.
+In \code{cubble}, analysts have the flexibility to combine these two structures together by allowing the spatial component to also be an \code{sf} object and the temporal component to also be a \code{tsibble} object.
+
+### Using a tsibble object as the temporal component
+
+The \code{key} and \code{index} arguments in a \code{cubble} object corresponds to the \code{tsibble} counterparts and they can be safely omitted, if the temporal component is a \code{tsibble} object, i.e.
+\code{meteo_ts} in the example below.
+The \code{tsibble} class from the input will be carried over to the \code{cubble} object:
+
+```{r echo = TRUE}
+ts_nested <- make_cubble(
+  spatial = stations, temporal = meteo_ts, coords = c(long, lat))
+(ts_long <- face_temporal(ts_nested))
+class(ts_long)
+```
+
+The temporal cubble shows \code{[tsibble]} in the header to indicate the object also being in a \code{tbl_ts} class (\code{tsibble}).
+Methods applies to the \code{tbl_ts} class can also be applied to the temporal cubble objects, for example, checking whether the data contain temporal gaps:
+
+```{r echo = TRUE}
+ts_long %>% has_gaps()
+```
+
+An existing \code{cubble} object can promote its temporal component to a \code{tsibble} object by applying \code{make_temporal_tsibble()}.
+The following code illustrates this with the object \code{cb_long} created in section \@ref(create) and the promoted cubble object is equal to the cubble object originally created from a \code{tsibble} object:
+
+```{r echo = TRUE}
+ts_long2 <- cb_long %>% make_temporal_tsibble() 
+identical(ts_long2, ts_long)
+```
+
+### Using an sf object as the spatial component
+
+Similarly, an \code{sf} object can be supplied as the spatial component to create a \code{cubble} object, with the \code{coords} argument being omitted.
+This opens up the possibility to represent fixed area with polygons or multipolygons (see Applications \@ref(covid)) and the \code{coords} argument will be calculated as the centroids of the (multi)polygons.
+The \code{[sf]} print in the cubble header suggest an spatial component being also a \code{sf} object:
+
+```{r echo = TRUE}
+(sf_nested <- make_cubble(
+  spatial = stations_sf, temporal = meteo, 
+  key = id, index = date))
+class(sf_nested)
+```
+
+The \code{sf} functions applicable to a \code{cubble} object have been listed in Table \@ref(tab:funs) and the following code shows how to perform coordinate transformation with \code{st_transform} on a cubble object:
+
+```{r echo =TRUE, message=FALSE}
+sf_nested %>% sf::st_transform(crs = "EPSG:3857")
+```
+
+The counterpart to promote the spatial component in an existing \code{cubble} to be an \code{sf} object is \code{make_spatial_sf()}:
+
+```{r echo = TRUE}
+cb_nested %>% make_spatial_sf() 
+```
+
+## Comparison to other spatio-temporal classes {#tidyverse}
+
+In \proglang{R}, there are other existing spatio-temporal data structure and this section compares and contrasts \pkg{cubble} with other existing alternative, specifically \pkg{stars} and \pkg{sftime}.
+The \pkg{stars} package [@stars] uses an array structure, as oppose to tibble, to represent multivariate spatio-temporal data.
+While both \pkg{stars} and \pkg{cubble} support vector and raster data, it is a matter of choice on which structure to use given the application.
+Analysts working on satellite imageries may prefer the array structure in \pkg{stars}, while others originally working with spatio-temporal data in 2D data frames may find \pkg{cubble} easier to adopt from their existing computing workflow.
+
+The \pkg{sftime} package [@sftime] also builds from a tibble object and its focus is on handling irregular spatio-temporal data.
+This means \pkg{sftime} can also handle full space-time grids and sparse space-time layouts represented in \pkg{cubble}, but \pkg{cubble} uses nesting to avoid storing spatial variables repetitively at each timestamp.
+This provides memory efficiency when data is observed frequent, i.e.
+daily or sub-daily, or the spatial geometry is expensive to repeat, i.e.
+polygons or multipolygons.
+Consider the \code{climate_aus} data in the \pkg{cubble} package with `r nrow(cubble::climate_aus)` stations observed daily in a single year 2020.
+The created \code{sftime} object is approximately `r library("cubble"); round(object.size(sftime::st_as_sftime(climate_aus))/object.size(climate_aus))` times larger than the corresponding \code{cubble} object (`r round(object.size(sftime::st_as_sftime(climate_aus))/1e6, 2)` MB vs.
+`r round(object.size(cubble::climate_aus)/1e6, 2)` MB).
+
+# Other features and considerations {#others}
+
+## Data fusion and matching {#matching}
+
+One common task when working with spatio-temporal data is to match nearby sites.
+For example, we may want to verify the location of an old list of stations with current stations, or we may want to match the data from different data sources.
+In cubble, spatial and temporal matching are performed using the functions `match_spatial()` and `match_temporal()`.
+The `match_spatial()` function calculates the spatial distance between observations in two cubble objects.
+Various distance measures are available (check `sf::st_distance`).
+Analysts can specify the number of matched groups to output using the `spatial_n_group` argument (default to 4 groups) and the number of matches per group using the `spatial_n_group` argument (default to 1, one-to-one matching).
+The syntax to use `match_spatial()` is:
+
+````
+match_spatial(<cubble_obj1>, <cubble_obj2>, ...)
+````
+
+The function `match_temporal()` calculates the similarity between time series within spatially matched groups.
+Two identifiers are required: one for separating each matched group (`match_id`) and one for separating the two data sources (`data_id`).
+The argument `temporal_by` uses the `by` syntax from dplyr's `*_join` to specify the temporal variables to match.
+
+The similarity score between two time series is calculated using a matching function, which can be customised by the analysts.
+The matching function takes two time series as a list and returns a single numerical score.
+This allows for flexibility in using existing time series distance calculation implementation.
+By default, cubble implements a simple peak matching algorithm (`match_peak`) that counts the number of peaks in two time series that fall within a specified temporal window.
+The syntax to use `match_temporal()` is
+
+````
+match_temporal(
+  <cubble_obj_from_match_spatial>, 
+  data_id = , match_id = , 
+  temporal_by = c("..." = "...")
+)
+````
+
+## Interactive graphics
+
+The workflow with the \code{cubble} class works well with an interactive graphics pipeline (e.g., @buja1988elements; @buja1996interactive;  @sutherland2000orca; @xie2014reactive; @cheng2016enabling) that is available in \proglang{R} with the package \pkg{crosstalk} [@crosstalk].
+Figure \ref{fig:illu-interactive} illustrates the linking between a map and multiple time series in a \code{cubble} object.
+Both the map, generated from the spatial cubble, and the time series plot, generated from the temporal cubble, are shared crosstalk objects.
+When a user selects a location on the map (a), the corresponding site is highlighted.
+This selection activates a row in the spatial \code{cubble}, which is then communicated to the temporal \code{cubble}, resulting in the selection of all observations with the same ID in (b).
+Consequently the temporal \code{cubble} highlights the corresponding series in the time series plot (c).
+Linking can also be initiated from the time series plot, by selecting points on the time series.
+This action activates rows with the same ID in the temporal \code{cubble} and the row in the spatial \code{cubble}, which will be highlighted on the map.
+Please refer to the appendix for the corresponding illustration.
+
+```{r illu-interactive, echo = FALSE, fig.align="center", out.height = "35%", out.width = "100%", fig.cap = "Linking between multiple plots.
+The line plots and the map are constructed from shared \\code{crosstalk} objects.
+When a station is selected on the map (a), the corresponding row in the spatial \\code{cubble} will be activated.
+This will link to all the rows with the same id in the temporal \\code{cubble} (b) and update the line plot (c)."}
+knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.002.png"))
+```
+
+## Spatio-temporal transformations {#st_transformation}
+
+Sometimes, we wish to communicate spatial and temporal information collectively through visualisation.
+This can be achieved through several graphical displays: one can make faceted maps across time, creating map animations, or constructing interactive graphics to link between map and time series plot.
+The concept of glyph maps was initially proposed in @Wickham2012-yr.
+The underlying idea is to transform the temporal coordinates into spatial coordinates so that time series plot can be displayed on the map.
+
+The package `GGally` initially implement the glyph map.
+It uses `glyphs()` to calculate the axis transformation and then uses `geom_polygon()` to draw the map.
+In cubble, a ggproto implementation `geom_glyph()` performs the linear algebra internally as data transformation .
+The `geom_glyph()` requires four aesthetics: `x_major`, `y_major`, `x_minor`, and `y_minor`.
+The major axes are the outer spatial coordinates and the minor axes are the inner/ temporal coordinates:
+
+```
+data |> 
+  ggplot() +
+  geom_glyph(aes(x_major = ..., x_minor = ..., 
+                 y_major = ..., y_minor = ...))
+```
+
+Some useful controls over the glyph map includes:
+
+  * polar coordinate glyph maps with \code{polar = TRUE}, 
+  * adjust glyph size with arguments \code{width} and \code{height},
+  * glyph scale to fixed (\code{global_rescale} defaults to \code{TRUE}) or free, and 
+  * reference boxes and lines with \code{geom_glyph_box()} and \code{geom_glyph_line()}.
+
+# Applications {#examples}
+
+Five examples are chosen to illustrate different aspects of the \pkg{cubble} package: creating a \code{cubble} object from two Coronavirus (COVID) data tables with the complication of differing location names, using spatial transformations to make a glyph map of seasonal temperature changes, matching river level data and weather station records for analysis of water supply, reading NetCDF format data to reproduce a climate reanalysis plot, and the workflow to create complex interactive linked plots.
+
+## Victoria COVID spatio-temporal incidence and spread {#covid}
+
+Since the start of the pandemic, the Victoria State Government in Australia has been providing daily COVID counts per Logal Government Area (LGA).
+This data can be combined with map polygon data, available from the Australian Bureau of Statistics (ABS), to visualize COVID incidence and spread.
+In the \pkg{cubble} package, the COVID count data (\code{covid}) and the LGA information (\code{lga}) are available as a \code{tsibble} object and an \code{sf} object respectively.
+
+A cubble object can be created from separate spatial and temporal component using the \code{make_cubble()} function, introduced in Section \@ref(create). The \code{by} argument
+is used to specify the joining variable from the two component using the \code{*_join()} \code{by} syntax:
+
+```{r covid-4, echo = TRUE, warning = TRUE, message= TRUE}
+cb <- make_cubble(lga, covid, by = c("lga_name_2018" = "lga"))
+```
+
+The warning message suggests the slight difference of LGA encoding used by Victoria government and ABS and prompts analysts to use the function \code{check_key()} to identify the mismatches:
+
+```{r covid-5, echo = TRUE}
+(check_res <- check_key(
+  spatial = lga, temporal = covid, 
+  by = c("lga_name_2018" = "lga")
+))
+```
+
+The result of the \code{check_key()} function is a list containing three elements: 1) matched keys from both tables, 2) potentially pairable keys, and 3) others.
+Here, the main mismatch arises from the two LGAs: Kingston and Latrobe (Kingston is a LGA in both Victoria and South Australia and Latrobe is a LGA in both Victoria and Tasmania).
+Analysts can modify the input spatial and temporal data accordingly and recreate the cubble object:
+
+```{r covid-6, echo = TRUE}
+lga2 <- lga |>
+  rename(lga = lga_name_2018) |> 
+  mutate(lga = ifelse(lga == "Kingston (C) (Vic.)", "Kingston (C)", lga),
+         lga = ifelse(lga == "Latrobe (C) (Vic.)", "Latrobe (C)", lga)) |>
+  filter(!lga %in% check_res$others$spatial)
+
+covid2 <- covid |> filter(!lga %in% check_res$others$temporal)
+
+(cb <- make_cubble(spatial = lga2, temporal = covid2))
+```
+
+## Australian historical maximum temperature {#historicaltmax}
+
+```{r}
+load(here::here("data/historical_tmax.rda"))
+```
+
+The GHCN provides daily climate measures from stations across the world.
+The data used here (\code{historical_tmax}) is a subset extracted using the package \pkg{rnoaa} [@rnoaa], containing the records of maximum temperature for `r nrow(historical_tmax)` Australian stations from `r range(historical_tmax$first_year)[1]` through `r range(historical_tmax$first_year)[2]` and provides information also on the latitude, longitude and elevation of each of the  stations.
+<!-- This data  is already cast into a \code{cubble} data object (please see Appendix X for the code to extract  the data from XX and prepare the data into \code{cubble} form), with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates.--> The goal of this example is to compare the monthly average maximum temperature between two periods, 1971-1975 and 2016-2020, for stations in Victoria and New South Wales (NSW), using a glyph map.
+
+First, the stations need to be filtered to those in Victoria and NSW by using the station identifiers, stored within the 11 digits of the \code{id} variable entries.
+The country code is in the first 5 digits (Australia is represented by "ASN00") and the next 6 digits encode the station following the Australian Bureau of Meteorology (BOM) [@bom] coding protocols.
+The NSW stations correspond to entries in the range 46-75 and the Victorian stations to 76-90.
+Filtering Victoria and NSW stations is a *spatial operation* and hence uses the spatial \code{cubble}:
+
+<!-- BOM link: http://www.bom.gov.au/climate/cdo/about/site-num.shtml -->
+
+```{r echo = TRUE}
+tmax <- historical_tmax |>
+  filter(between(as.numeric(stringr::str_sub(id, 7, 8)), 46, 90))
+```
+
+Next, the monthly maximum average temperature is  calculated for both periods.
+This is a *temporal operation* requiring a switch into the temporal \code{cubble} using the \code{face_temporal()} function.
+In addition, a new indicator for the two time periods of interest is created before the calculation of monthly averages:
+
+```{r echo = TRUE}
+tmax <- tmax |>
+  face_temporal() |> 
+  group_by(
+    yearmonth = tsibble::make_yearmonth(
+      year = ifelse(lubridate::year(date) > 2015, 2016, 1971),
+      month = lubridate::month(date))
+      )|>
+  summarise(tmax = mean(tmax, na.rm = TRUE)) %>% 
+  mutate(group = as.factor(lubridate::year(yearmonth)),
+         month = lubridate::month(yearmonth))
+```
+
+A quick check on the number of observations for each location is made, revealing that there are several with less than 24 observations -- these stations lack temperature values for some months.
+In this example, those stations are removed by switching to the spatial \code{cubble} to operate on the spatial component over time, and then, move back into the temporal \code{cubble} (to make the glyph map):
+
+```{r echo = TRUE}
+tmax <- tmax |> 
+  face_spatial() |> 
+  rowwise() %>% 
+  filter(nrow(ts) == 24) |>
+  face_temporal()
+```
+
+In order to create a glyph map displaying the monthly series  (Figure \ref{fig:basic-manip}), the spatial variables need to be unfolded with the temporal variables.
+The reason being that the major (\code{long}, \code{lat}) and minor (\code{month}, \code{tmax}) coordinates need to be on the same table to create the glyph map.
+The \code{geom_glyph()} function does both the transformation and the plotting.
+
+```{r eval = FALSE, echo = TRUE}
+ggplot() +
+  geom_sf(data = MAP_DATA, ...) +
+  geom_glyph(data = tmax,
+             aes(x_major = long, x_minor = month,
+                 y_major = lat, y_minor = tmax,
+                 group = interaction(id, group), color = group),
+             width = 1, height = 0.5) + 
+  ...
+```
+
+```{r}
+cobar <- tmax |> filter(id == "ASN00048027") |>
+  ggplot(aes(x = month, y = tmax, color = group)) +
+  geom_line() +
+  scale_color_brewer(palette = "Dark2", guide = "none") +
+  scale_x_continuous(
+    breaks = seq(1, 12, 1),
+    labels = c("J", "F", "M", "A",
+               "M", "J", "J", "A",
+               "S", "O", "N", "D")
+    ) +
+  labs(x = "", y  = "Temp (C)") +
+  theme_bw() + 
+  theme(aspect.ratio = 0.3)
 
@@
 
-```{r child = "2-the-cubble-pkg.Rmd"}
+```{r basic-manip, out.width="100%", fig.height = 5, fig.width = 10, fig.cap="A glyph map of the monthly maximum average temperature for weather stations in  Victoria and New South Wales (NSW) for the periods (1971-1975, 2016-2020).
+The corresponding average time series for the cobar station are display on the top left corner.
+From the glyph map we can observe that the monthly trend is similar for all locations (low in the winter, high in the summer), and small increased temperatures, particularly in late summer can be seen at most stations in NSW."}
+tmax <- tmax |> unfold(long, lat)
+box_df <- tmax |> face_spatial() |> filter(id == "ASN00048027")
+nsw_vic <- ozmaps::abs_ste |> filter(NAME %in% c("Victoria","New South Wales"))
+
+p1 <- tmax %>% 
+  ggplot(aes(x_major = long, x_minor = month, 
+             y_major = lat, y_minor = tmax,
+             group = interaction(id, group))) + 
+  geom_sf(data = nsw_vic, fill = "grey95", 
+          color = "white", inherit.aes = FALSE) + 
+  geom_glyph_box(data = tmax, width = 0.8, height = 0.3) +
+  geom_glyph(data = tmax, aes(color = group), width = 0.8, height = 0.3) +
+    # geom_rect(
+    #   data = box_df,
+    #   aes(xmin = long - 0.6, xmax = long + 0.6,
+    #       ymin = lat - 0.12, ymax = lat + 0.35),
+    #   fill = "transparent", color = "black") +
+  scale_color_brewer("", palette = "Dark2") + 
+  coord_sf(xlim = c(141, 154), ylim = c(-39, -28.5)) + 
+  ggthemes::theme_map() +
+  theme(legend.position = "bottom", legend.text = element_text(size = 10)) + 
+  guides(color = guide_legend(override.aes = list(size=2)))
+
+(p1 | cobar)  + 
+  patchwork::plot_layout(width = c(1,1)) + 
+  plot_annotation(tag_levels = 'a')
 
@@
 
-```{r child = "3-advanced.Rmd"}
+Glyph maps work well to explore temporal patterns across spatial locations, particularly when the spatial locations are gridded.
+In this example, they are irregularly spaced, which can result in overlapping glyphs obscuring each other.
+
+## River levels and rainfall in Victoria
+
+One common task when working with spatio-temporal data is to match nearby sites.
+For example, we may want to verify the location of an old list of stations with current stations, or we may want to match the data from different data sources.
+In this example, we will introduce the spatial and temporal matching in cubble using an example on matching river level data with precipitation in Victoria, Australia.
+
+The water level data collected by the Bureau of Meteorology, can be compared with the precipitation since rainfall can directly impact water level in river.
+Both `climate_vic` and `river` are cubble objects, and we can obtain a summary of the 10 closest pairs between them:
+
+```{r}
+climate_vic <- climate_aus |>
+  filter(between(as.numeric(stringr::str_sub(id, 7, 8)), 76, 90)) |>
+  mutate(type = "climate")
+river <- cubble::river %>% mutate(type = "river") 
+```
+
+```{r echo = TRUE}
+res_sp <- match_spatial(climate_vic, river, spatial_n_group = 10)
+print(res_sp, n = 20)
+```
+
+The result can also be returned as cubble objects by setting the argument `return_cubble = TRUE`.
+The output is be a list where each element is a paired cubble object.
+To combine all the results into a single cubble, you can use `bind_rows()`.
+In the case when a site in the second cubble (the `river` data here) is matched to two stations in the first cubble (`climate_vic` here), the binding may not be successful since cubble requires unique rows in the nested form.
+In the summary table above, the river station `226027` is matched to more than one weather station: `ASN00085072` (group 3) and `ASN00085298` (group 5).
+Similarly, the river station `230200` is matched in group 7 and 8).
+In such cases, you can either deselect one pair before combining, or work with the list output with the `purrr::map` syntax:
+
+```{r echo = TRUE}
+res_sp <- match_spatial(
+  climate_vic, river, 
+  spatial_n_group = 10, return_cubble = TRUE)
+str(res_sp, max.level = 0)
+(res_sp <- res_sp[-c(5, 8)] %>% bind_rows())
+```
+
+For temporal matching, we match the variable `Water_course_level` from the river data to `prcp` in the weather station data.
+The variable `group` and `types` identify the matching group and the two datasets:
+
+```{r echo = TRUE}
+(res_tm <- res_sp %>% 
+  match_temporal(
+    data_id = type, match_id = group,
+    temporal_by = c("prcp" = "Water_course_level")))
+```
+
+Similarly, the cubble output can be returned using the argument `return_cubble = TRUE`. Here we select the four pairs with the highest number of matching peaks:
+
+```{r echo = TRUE}
+res_tm <- res_sp %>% 
+  match_temporal(
+    data_id = type, match_id = group,
+    temporal_by = c("prcp" = "Water_course_level"),
+    return_cubble = TRUE)
+(res_tm <- res_tm %>% bind_rows() %>% filter(group %in% c(1, 7, 6, 9)))
+```
+
+Figure \ref{fig:matching} shows four matched pairs on the map (a) and standardized data as time series (b) with concurrent increasing.
+
+```{r matching, out.width="100%", fig.height = 5, fig.width = 10, fig.cap="Weather stations and river gauges with matched pairs labelled on the map (a) and plotted across time (b).
+Precipitation and water level have been standardised between 0 and 1 to be displayed on the same scale.
+The water level reflects the increase in precipitation.
+The numbers (1, 7, 6, 9) indicate the group index derived from spatial matching, only those that were selectd by temporal matching are shown here."}
+res_tm <- res_tm %>%  filter(group %in% c(1, 7, 6, 9))
+
+res_tm_long <- res_tm %>%  
+  face_temporal() %>%  
+  unfold(group, type) %>%  
+  group_by(group, type) %>%
+  mutate(matched = (matched - min(matched, na.rm = TRUE))/ 
+           (max(matched, na.rm = TRUE) - min(matched, na.rm = TRUE)))
+
+vic_map <- ozmaps::abs_ste |> 
+  filter(NAME == "Victoria") |> 
+  rmapshaper::ms_simplify()
+
+p1 <-ggplot() + 
+  geom_sf(data = vic_map, fill = "grey95", color = "white") + 
+  geom_point(data = dplyr::bind_rows(river, climate_vic), 
+             aes(x = long, y = lat, color = type), 
+             alpha = 0.2, fill = 0.2) +
+  geom_point(data = res_tm %>% as_tibble(), 
+             aes(x = long, y = lat, color = type)) +
+  ggrepel::geom_label_repel(
+    data = res_tm |> filter(type == "river") %>% as_tibble(), 
+    aes(x = long, y = lat, label = group)) +
+  scale_color_brewer(palette = "Dark2")  + 
+  ggthemes::theme_map() +
+  ggplot2::theme(legend.position = "bottom",
+                 legend.text = element_text(size = 15),
+                 legend.title = element_text(size = 15)) +
+  ggplot2::labs(x = "Longitude", y = "Latitude") + 
+  guides(color = guide_legend(override.aes = list(size=5)))
+
+p2 <- res_tm_long %>%  
+  ggplot(aes(x = date, y = matched, group = type,color = type)) + 
+  geom_line() + 
+  facet_wrap(vars(group)) + 
+  scale_color_brewer(palette = "Dark2", guide = "none") + 
+  theme_bw() + 
+  labs(x=  "date") + 
+  scale_x_date(date_labels = "%b") + 
+  labs(x = "Week", y = "Precipitation/ water level")
+
+(p1 | p2) + 
+  patchwork::plot_layout(guides = "collect") + 
+  plot_annotation(tag_levels = 'a')&
+  theme(legend.position = "bottom") 
+```
+
+## ERA5: climate reanalysis data
+
+The ERA5 reanalysis [@hersbach2020era5] provides hourly estimates of atmospheric, land and oceanic climate variables on a global scale.
+The hourly pressure level data in the NetCDF format can be downloaded from Copernicus Climate Data Store (CDS) [@cds] or via the \pkg{ecmwfr} package [@ecwmfr].
+This example reproduces the row created from ERA5 reanalysis data shown in Figure 19 by @hersbach2020era5.
+The plot shows the southern polar vortex splitting into two on 2002-09-26, and further splitting into four on 2002-10-04.
+Further explanation of why this is interesting can be found in the figure source, and also in @simmons2020global and @simmons2005ecmwf.
+
+To work with NetCDF data in R, several packages are available, including \code{ncdf4}, \code{RNetCDF}, and \code{tidync}.
+The following code converts a NetCDF object of class \code{ncdf4} into a cubble object:
+
+```{r echo = TRUE}
+raw <- ncdf4::nc_open(here::here("data/era5-pressure.nc"))
+```
+
+Analysts can extract a subset of the NetCDF data with the arguments \code{vars}, \code{long_range} and \code{lat_range}.
+In this example, the variables q (specific humidity) and z (geopotential) are selected and the coordinates are subsetted to every degree in longitude and latitude:
+
+```{r echo = TRUE}
+(dt <- as_cubble(raw, vars = c("q", "z"),
+                 long_range = seq(-180, 180, 1), lat_range = seq(-88, -15, 1)))
+```
+
+Once the NetCDF data is coerced into a cubble object, subsequent analysis can be conducted to filter on the date of interest, scale the variable specific humidity and create visualisation in ggplot as in Figure \ref{fig:netcdf}.
+
+```{r netcdf, out.width="100%", fig.height = 4, fig.width = 10, fig.cap = "A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020) to illustrate the break-up of sourthern polar vortex in late September and early October 2002.
+The polar vortex, signalled by the high specific humidity, splits into two on 2002-09-26 and further splits into four on 2002-10-04."}
+date <- c("2002-09-22", "2002-09-26", "2002-09-30", "2002-10-04") |> as.Date()
+res <- dt |> 
+  face_temporal() |> 
+  filter(lubridate::date(time) %in% date) |>
+  unfold(long, lat) |> 
+  mutate(q = q* 10^6)
+
+con <- rnaturalearth::ne_coastline("small", returnclass = "sf")
+box <- st_bbox(c(xmin = -180, ymin = -90, xmax = 180, ymax = -15), crs = st_crs(con)) 
+sf::sf_use_s2(FALSE)
+country <- con %>% st_geometry() %>% 
+  sf::st_crop(box) |> 
+  sf::st_cast("MULTILINESTRING")
+
+res |> 
+  ggplot() +
+  # q for specific humidity
+  geom_point(aes(x = long, y = lat, color = q)) + 
+  #geom_tile(aes(x = long, y = lat, fill = q), color = "transparent") +
+  # z for geopotential
+  geom_contour(data = res, aes(x = long, y = lat, z = z),
+               color = "grey20", binwidth = 4000, linewidth = 0.5) +
+  geom_sf(data = country , alpha = 0.5, fill = "transparent", color = "lightgreen") +
+  coord_sf(
+    crs = "+proj=stere +lat_0=-90 +lon_0=-180 +k=1 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs", default_crs = sf::st_crs(4326), clip = "off") +
+  theme_bw() +
+  theme(legend.position = "bottom",
+        axis.ticks = element_blank(),
+        panel.border = element_blank(),
+        axis.text = element_blank(),
+        panel.grid = element_blank()) +
+  facet_wrap(vars(as.Date(time)), nrow = 1)+  
+  colorspace::scale_color_continuous_sequential("Purple-Yellow", name = "Specific humidity") + 
+  theme(legend.text = element_text(size = 13),
+        legend.title = element_text(size = 13)) + 
+  labs(x ="", y = "")
+```
+
+```{r eval = FALSE}
+# prepare data/era5-pressure.nc
+library(ecmwfr)
+my_request <- list(
+  product_type = "reanalysis",
+  format = "netcdf",
+  pressure_level = "10",
+  year = "2002",
+  month = c("09", "10"),
+  day = c("04", "22", "26", "30"),
+  time = "00:00",
+  area = c(-15, -180, -90, 180),
+  variable = c("specific_humidity", 'geopotential'),
+  dataset_short_name = "reanalysis-era5-pressure-levels",
+  target = "era5-pressure.nc"
+)
+# 
+# # here you need to create an account on Climate Data Store 
+# # at https://cds.climate.copernicus.eu/cdsapp#!/home) to obtain the user ID and API key
+# # then set up the with ecmwfr::wf_set_key(user, key, service = "cds")
+# # the user below is the same `user` in the setup line
+wf_request(
+  user = "113339",
+  request = my_request,
+  transfer = TRUE,
+  path = here::here("data/"))
+```
+
+## Australian temperature range
+
+Interactive graphics can be useful because they make it possible to look at the data in multiple ways on-the-fly.
+This is especially important for spatio-temporal data, where we would like to interactively connect spatial and temporal displays.
+This example describes the process of using the \pkg{cubble} package with the \pkg{crosstalk} package to build an interactive display connecting a map of Australia, with ribbon plots of temperature range observed at the stations.
+The purpose is to explore the variation of monthly temperature range over the country.
+Figure \ref{fig:interactive-linking} shows three snapshots of the interactivity.
+
+The key steps are to convert both the nested and long forms of the data into shared \code{crosstalk} objects, and to plot these side-by-side. The two are linked by the station identifier.
+
+``` r
+clean <- climate_full |> ...
+
+nested <- clean |> SharedData$new(~id, group = "cubble")
+long <- face_temporal(clean) |> SharedData$new(~id, group = "cubble")
+
+p1 <- nested |> ...
+p2 <- long |> ...
+
+crosstalk::bscols(plotly::ggplotly(p1), plotly::ggplotly(p2), ...)
 
@@
 
-```{r child = "4-1-basic.Rmd"}
+Plot (a) shows the initial state of the interactive display: all locations are shown as dots on the map, coloured by temperature range, and the right plot shows the ribbons representing maximum to minimum for all stations.
+In plot (b) the "Mount Elizabeth" station, which shows a high variance colour on the initial map, is selected on the map and this produces the ribbon on the right.
+In plot (c) the lowest temperature in August is selected, which is "Thredbo" station on the left map.
+It was surprising to us that this was not a station in Tasmania, so for comparison a station in Tasmania is selected on the map to show in relation to Thredbo.
+We can see that Thredbo has a bigger winter dip in temperature, and although Tasmania is cold generally, it's temperatures are more constant
+
+```{r eval = FALSE}
+# This chunk generates the interactive graphic and save it as figures/linking.html, where the three screencasts in the figure comes from.
+library("crosstalk")
+library("plotly")
+
+clean <- climate_full |>
+  filter(!id %in% c("ASN00067033", "ASN00072091", "ASN00059040", 
+                    "ASN00097053", "ASN00040856", "ASN00015548")) |>
+  face_temporal() |> 
+  mutate(month = lubridate::month(date)) |>
+  group_by(month) |>
+  summarise(
+    tmax = mean(tmax, na.rm = TRUE),
+    tmin = mean(tmin, na.rm = TRUE),
+    diff = mean(tmax - tmin, na.rm = TRUE)
+    ) |> 
+  ungroup(month) |>
+  mutate(dummy_date = as.Date(glue::glue("2021-{month}-01"))) |> 
+  face_spatial() |> 
+  mutate(temp_diff_var = ts |> pull(diff) |> var(na.rm = TRUE))
+
+nested <- clean |> 
+  SharedData$new(~id, group = "cubble")
+
+long <- clean |> 
+  face_temporal() |> 
+  unfold(temp_diff_var) |> 
+  arrange(temp_diff_var) |> 
+  SharedData$new(~id, group = "cubble")
+
+state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)
+
+p1 <- nested |> 
+  ggplot() +
+  geom_sf(data = state_map, fill = "grey95", color = "white") +
+  geom_point(aes(x = long, y = lat,color = temp_diff_var, label = name)) +
+  colorspace::scale_color_continuous_sequential(
+    "Rocket",  n_interp = 7, cmax = 90, rev = TRUE, 
+    c2 = 40, l2= 85, c1 = 20, l1 = 30, name = "Var. temp. diff.") +
+  labs(x = "Longitude", y = "Latitude") + 
+  coord_sf() + 
+  theme_bw() + 
+  theme(
+    panel.grid.major = element_blank(),
+    legend.position = "none"
+    )
+
+p2 <- long |> 
+  ggplot(aes(x = dummy_date, group = id, 
+             fill = temp_diff_var, color = temp_diff_var)) +
+  geom_ribbon(aes(ymin = tmin, ymax = tmax), size = 0.1, alpha = 0.3) + 
+  geom_line(aes(y = tmin), alpha = 0.1) + 
+  geom_line(aes(y = tmax), alpha = 0.1) + 
+  geom_point(aes(y = tmax), size = 0.1) + 
+  geom_point(aes(y = tmin), size = 0.1) + 
+  colorspace::scale_fill_continuous_sequential(
+    "Rocket",  n_interp = 7, cmax = 90, rev = TRUE, 
+    c2 = 40, l2= 85, c1 = 20, l1 = 30, name = "Var. temp. diff.") +
+  colorspace::scale_colour_continuous_sequential(
+    "Rocket",  n_interp = 7, cmax = 90, rev = TRUE, 
+    c2 = 40, l2= 85, c1 = 20, l1 = 30, name = "Var. temp. diff.") +
+  scale_x_date(date_labels = "%b", date_breaks = "1 month") + 
+  labs(x = "Month", y = "Temperature") + 
+  theme_bw() + 
+  theme(
+    panel.grid.major = element_blank(),
+    legend.position = "bottom"
+    )
+
+out <- bscols(
+  ggplotly(p1, width = 900, height = 700, tooltip = "label") |>
+    highlight(
+      on = "plotly_selected", 
+      off = "plotly_deselect", 
+      opacityDim = 0.1),
+  ggplotly(p2, width = 1200, height = 700, tooltip = "label") |> 
+    highlight(
+      on = "plotly_selected", 
+      off = "plotly_deselect", 
+      opacityDim = 0.012),
+  widths = c(5, 5)
+)
+
+htmltools::save_html(out, file = "figures/linking.html")
 
@@
 
-```{r child = "4-2-agg.Rmd"}
-
+```{r interactive-linking, echo = FALSE, out.width="100%", out.height="23%", fig.retina = 2, dpi = 300, fig.cap = "Exploring temperature variation using linking of a map and seasonal display.
+Each row is a screen dump of the process.
+The top row shows all locations and all temperature profiles.
+Selecting a particular location on the map (here Mount Elizabeth) produces the plot in the second row.
+The maximum and minimum temperatures are shown using a ribbon.
+The bottom row first selects the lowest temperature in August in the seasonal display, which highlights the corresponding station on the map (Thredbo).
+Another  station, located in the Tasmania Island, is then selected to compare its temperature variation with the Thredbo station.", fig.show='hold'}
+knitr::include_graphics(here::here("figures/linking.png"))
+knitr::include_graphics(here::here("figures/linking-north.png"))
+knitr::include_graphics(here::here("figures/linking-lower.png"))
 ```
 
-```{r child = "4-3-water-level.Rmd"}
+# Conclusion {#conclude}
 
-```
+This paper presents an \proglang{R} package \pkg{cubble} for organizing, manipulating and visualizing spatio-temporal data.
+The package introduces a new data class for spatio-temporal data, \code{cubble}, that connects the time invariant and varying variables and that allows the user to work with a nested and long form of the data.
+This work adds capabilities into the spatio-temporal practitioners toolbox to integrate it with a tidy data framework.
+The data structure and functions introduced in this package can be used and combined with existing spatial data analysis packages such as \pkg{sf}, data wrangling packages such as \pkg{dplyr}, and visualization packages such as \pkg{ggplot2}, \pkg{plotly} and \pkg{leaflet}.
 
-```{r child = "4-4-netcdf.Rmd"}
+Numerous examples are provided in the main text, appendix and package vignettes.
+These include creating and coercing data with mismatched names, handling hierarchical data, matching time series spatially and temporally, reproducing ERA5 plots from NetCDF data.
+Visualization of the \code{cubble} objects can be done with interactive graphic pipelines using \pkg{crosstalk}, \pkg{plotly} and \pkg{leaflet}.
 
-```
+There are several possible future directions for the work.
+The data structure only described fixed spatial sites, and it could be useful to provide tools to accommodate moving coordinates as might be encountered in animal movement data.
+That could be achieved with a list-column for the location coordinates, and an additional form that these locations can be pivoted into, like the long form for temporal variables.
+For multiple measured variables, the \pkg{cubble} package could be integrated with dimension reduction methods, and dynamic graphics for multiple dimensions such as a tour [@wickham_tourr_2011].
 
-```{r child = "4-5-interactive.Rmd"}
+# Acknowledgement
 
-```
-
-```{r child = "5-conclusion.Rmd"}
-
-```
+This work is funded by a Commonwealth Scientific and Industrial Research Organisation (CSIRO) Data61 Scholarship and started while Nicolas Langrené was affiliated with CSIRO's Data61.
+The article is created using the package \pkg{knitr} [@knitr] and \pkg{rmarkdown} [@rmarkdown] in \proglang{R} with the \code{rticles::jss_article} template.
+The source code for reproducing this paper can be found at: https://github.com/huizezhang-sherry/paper-cubble.

### 2023-06-21 main.tex
merge all the Rmds into a single oneg

@@
 \author{
-H. Sherry Zhang\\Monash University \And Dianne Cook\\Monash University
-\AND Ursula Laa\\University of Natural Resources and Life Sciences
-\AND Nicolas Langrené\\BNU-HKBU United International College
-\AND Patricia Menéndez\\Monash University \AND
+H. Sherry Zhang\\Monash University \And Dianne Cook\\Monash
+University \And Ursula Laa\\Universität für Bodenkultur Wiens
+\AND Nicolas Langrené\\BNU-HKBU United International
+College \And Patricia Menéndez\\Monash University
 }
-\title{\pkg{cubble}: An R Package for Structuring Spatio-temporal Data}
+\title{\pkg{cubble}: An \proglang{R} Package for Organizing and
+Wrangling Multivariate Spatio-temporal Data}
 
@@
 Langrené, Patricia Menéndez}
-\Plaintitle{cubble: An R Package for Structuring Spatio-temporal Data}
+\Plaintitle{cubble: An R Package for Organizing and Wrangling
+Multivariate Spatio-temporal Data}
 
 \Abstract{
-Spatio-temporal data refer to measurements taken across space and time.
-In practice, spatio-temporal data can be decomposed into a spatial and
-temporal component: at one time, we would select a spatial location and
-inspect the temporal trend; at other time, we might select one or
-multiple time value(s) and explore the spatial distribution. Ideally, we
-could make multiple maps and multiple time series to explore these
-together, however, doing all of these actions is complicated when data
-arrive fragmented in multiple objects. To make it easy to do all these
-tasks, ideally spatial and temporal variables are in a single data
-object that we can slice and dice in different ways to conduct different
-visualisations. This work suggests a new data structure, \code{cubble},
-that uses a nested form and a long form to organise spatial and temporal
-variables in a single data object. The new data structure ensures that
-data in the two forms are synchronised while been flexible to explore
-the two aspect of spatio-temporal data. It also provides capability for
-handling data with hierarchical structure, matching data from multiple
-sources, constructing interactive graphics, and performing
-spatio-temporal transformation. The paper will demonstrate \pkg{cubble}
-with examples using Australian climate weather stations, river level
-data, and climate reanalysis (ERA5) data.
+Multivariate spatio-temporal data refers to multiple measurements taken
+across space and time. For many analyses, spatial and time components
+can be separately studied: for example, to explore the temporal trend of
+one variable for a single spatial location, or to model the spatial
+distribution of one variable at a given time. However for some studies,
+it is important to analyze different aspects of the spatio-temporal data
+simultaneously, like for instance, temporal trends of multiple variables
+across locations. In order to facilitate the study of different portions
+or combinations of spatio-temporal data, we introduce a new class,
+\code{cubble}, with a suite of functions enabling easy slicing and
+dicing on different spatio-temporal components. The proposed
+\code{cubble} class ensures that all the components of the data are easy
+to access and manipulate while providing flexibility for data analysis.
+In addition, the \pkg{cubble} package facilitates visual and numerical
+explorations of the data while easing data wrangling and modelling. The
+\code{cubble} class and the functions provided in the \pkg{cubble}
+\proglang{R} package equip users with the capability to handle
+hierarchical spatial and temporal structures. The \code{cubble} class
+and the tools implemented in the package are illustrated with different
+examples of Australian climate data.
 }
@@
       Ursula Laa\\
-    University of Natural Resources and Life Sciences\\
+    Universität für Bodenkultur Wiens\\
     Gregor-Mendel-Straße 33, 1180 Wien, Austria\\
@@
 
-\usepackage{amsmath} \usepackage{array} \usepackage{xcolor}
+% From pandoc table feature
+\usepackage{longtable,booktabs,array}
+\usepackage{calc} % for calculating minipage widths
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
+
+\usepackage{amsmath} \usepackage{array}
 
 \begin{document}
-
-\newpage
 
@@
 
-Spatio-temporal data involves variables observed in the space across
-time and this paper focuses on spatio-temporal vector data with fixed
-locations and regular time intervals. This type of data can naturally be
-considered as a data cube with three axes: spatial identifier, temporal
-identifier, and variables. In this cubic formation, space only occupies
-one axis, rather than two, to avoid hypercubes for multivariate
-spatio-temporal data. While conceptually the data format is relatively
-simple, the data objects coming to the analysts can vary in forms in the
-analysis. Figure \ref{fig:illu-input} shows three examples of wild
-spatio-temporal data: 1) a single table that includes both the spatial
-and temporal variables; 2) separate tables for spatial and temporal
-variables; 3) array-based data, i.e.~NetCDF (Network Common Data Form)
-data, which is commonly used in earth science, e.g.~climatology,
-meteorology, and oceanography, among others, to organise multivariate
-data in a spatio-temporal grid.
+Spatio-temporal data has a spatial component referring to the location
+of each observation and a temporal component that is recorded at regular
+or irregular time intervals. It may also include multiple variables
+measured at each spatial and temporal values. With spatio-temporal data,
+one can fix the time to explore the spatial features of the data, fix
+the spatial location/s to explore temporal aspects, or dynamically
+explore the space and time simultaneously.
+
+In order to computationally explore the spatial, temporal and
+spatio-temporal faces of such data, the data needs to be stored and
+represented under a specific data object that allows the user to query,
+group and dissect all the data faces.
+
+The Comprehensive \proglang{R} Archive Network (CRAN) task view
+SpatioTemporal \citep{ctvspatiotemporal} gathers information about
+\proglang{R} packages designed for spatio-temporal data and it has a
+section on \emph{Representing data} that lists existing spatio-temporal
+data representations used in \proglang{R}. Among them, \citet{spacetime}
+summarises spatio-temporal data into three forms: time-wide, space-wide,
+and long formats. The associated package \pkg{spacetime}
+\citep{spacetime} implements four spatio-temporal layouts (full grid,
+sparse grid, irregular, and trajectory) to handle different space and
+time combinations. The package \pkg{stars} \citep{stars} has a new
+implementation to use dense arrays to represent spatio-temporal cubes.
+It also interfaces with the package \pkg{sf} \citep{sf}, commonly used
+for wrangling spatial data, and the \pkg{tidyverse} \citep{tidyverse}
+suite for general data wrangling and visualization in \proglang{R}.
+
+Still, the data representation for spatio-temporal data can be further
+extended and there are two reasons for this. Firstly, the raw data
+sourced in the wild is less often presented in any one of the layouts
+above, and fitting the raw data into a data object can sometimes be
+difficult. More often, spatio-temporal data are collected in separate 2D
+tables and analysts need to assemble them into a whole piece before
+exploring the data. Examples of components of spatio-temporal data can
+be 1) areal data recording the shape of a collection of areas of
+interest; 2) geostatistical data storing the longitude and latitude
+coordinates of locations, typically also with other metadata related to
+the location, and; 3) temporal data of each location across time.
+
+The other reason is about tidy data concepts \citep{tidydata} and how
+they should be applied to spatio-temporal data. According to the tidy
+data principles, data should be structured into 1) one row per
+observation, 2) one column per variable, and 3) one type of data per
+table. The long form data is preferred over wide data form given the
+downstream packages such as \pkg{dplyr} \citep{dplyr} and \pkg{ggplot2}
+\citep{ggplot2} for data wrangling and visualization. However, the long
+form can be inefficient to store feature geometries, especially for
+large multipolygons for hourly, daily or sub-daily periods over years,
+which are extensively collected and handled, for example in time series
+analysis. This poses the question of how to arrange spatial and temporal
+variables in a way that would make data wrangling, visualizing and
+analyzing spatio-temporal data easier.
+
+This paper presents a new \proglang{R} package, \pkg{cubble}, which
+addresses the two issues mentioned above. In the package, a new class,
+also called cubble, is proposed to organize spatial and temporal
+variables as two forms of a single data object so that they can be
+wrangled separately or combined while being kept synchronized. Among the
+four spacetime layouts in \citet{spacetime}, the \code{cubble} class can
+be applied to full grid, sparse grid, or irregular, but not trajectory,
+which is outside the scope of this work. The software is available from
+the Comprehensive \proglang{R} Archive Network (CRAN) at
+\url{https://CRAN.R-project.org/package=cubble}.
+
+The rest of the paper is organized as follows: Section @ref(cubble)
+presents the main design and functionality of the \pkg{cubble} package.
+Section @ref(others) explains how the \pkg{cubble} package deals with
+more advanced considerations, including data with hierarchical
+structure, data matching and how the package fits with existing static
+and interactive visualization tools. Moreover we also illustrate how the
+\pkg{cubble} package deals with spatio-temporal data transformations.
+Section @ref(examples) uses Australian weather station data and river
+level data as examples to demonstrate the use of the package. An example
+of how the \pkg{cubble} package handles Network Common Data Form
+(NetCDF) data is also provided. Section @ref(conclude) discuss the paper
+contributions and future directions.
+
+\hypertarget{cubble}{%
+\section{The cubble package}\label{cubble}}
+
+\hypertarget{object}{%
+\subsection{The cubble object}\label{object}}
+
+Spatio-temporal data can encompass data with various spatial and
+temporal characteristics and different data require different structures
+for wrangling and analysis: climate weather stations record data at
+fixed point location but may suffer from potential temporal data quality
+issue such as missing data for certain days. GPS data tracks unique
+point locations at different timestamps and is represented as
+trajectories. Satellite imageries capture snapshots of landscapes at
+selected time and is commonly structured as raster data. The
+spatio-temporal data cubble address are those collected at unique fixed
+locations while allowing for irregularity in the temporal dimension,
+such as the weather station data. This corresponds to the full
+space-time and sparse space-time layouts described in the spacetime
+paper \citep{spacetime} and reproduced in Figure \ref{fig:layout}.
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.001}
+{\centering \includegraphics{/Users/hzha400/Documents/research/paper-cubble/figures/layout-1}
 
@@
 
-\caption[An illustration of different spatio-temporal data layouts]{An illustration of different spatio-temporal data layouts: (1) a single table for both spatial and temporal variables; (2) separate tables for spatial and temporal variables; and (3) an array data structure that populates variables into a gridded longitude-latitude-time (hyper)cube.}\label{fig:illu-input}
+\caption[The full space-time and sparse space-time layouts handled by the cubble package]{The full space-time and sparse space-time layouts handled by the cubble package.
+Reproduced from Figure 1 in the spacetime paper \citep{spacetime}.}(\#fig:layout)
 \end{figure}
@@
 
-While incorporating both spatial and temporal information enriches the
-data in the analysis, spatio-temporal data can sometimes be hard to work
-with given its two observational units in the variables. The single
-table format in Figure \ref{fig:illu-input} is convenient for
-spatio-temporal operations that involve both types of variables,
-however, it makes duplicates of spatial variables and is not ideal when
-only looking at the spatial aspect of the data. The two tables format
-allows separate operations on spatial and temporal variables while
-analysts need to work with two data objects simultaneously and join them
-together into the single table format for spatio-temporal operations.
-Neither of the two formats is desirable at the same time for spatial,
-temporal, and spatio-temporal operations. This issue implies analysts
-need to constantly think about the data format and arrange variables
-into the proper shape before the actual operations, which can be a
-laborious step when switching between the two formats back and forth
-several times during the analysis.
-
-In the \proglang{R} community, many data objects have been proposed to
-work with specific spatio-temporal tasks, while few can freely explore
-the data. \pkg{spacetime} \citep{spacetime} is an attempt for such a
-data object, but the spatial and temporal classes it is built on have
-now been gradually replaced by more recent implementations. The
-\pkg{stars} \citep{stars} package uses an array to wrangle
-spatio-temporal data, but the dense array it uses can quickly exhaust
-the memory for data in the case of irregular spatial grid. While there
-has not yet been a data object that makes it easy to wrangle
-spatio-temporal data, there is a strong need for such tools given the
-ubiquitous spatial and temporal data collected in the society, i.e.,
-climate observations from weather stations, air quality variables from
-air monitoring stations. The demand for such tools can also come from
-spatial data analysts wishing to incorporate temporal data in their
-analysis or vice versa.
-
-This paper presents a new \proglang{R} package, \pkg{cubble}, to
-overcome the problem with organising variables in different levels when
-working with spatio-temporal data. In the package a new data structure,
-also called \code{cubble}, is proposed to organise spatial and temporal
-variables as two forms of a single data object so that they can be
-wrangled separately while synchronised as a whole piece. With this new
-data object, analysts can spend less time writing codes to organise
-spatial and temporal variables and focus more on the exploratory data
-analysis itself. The software is available from the Comprehensive R
-Archive Network (CRAN) at {[}CRAN link{]}.
-
-The rest of the paper is divided as follows: Section 2 presents the main
-design and functionality of \pkg{cubble}. Section 3 explains how cubble
-deals with more advanced considerations including data with hierarchical
-structure, data matching, how cubble fits with existing static and
-interactive visualisation tools, and spatio-temporal data
-transformation. Section 4 uses Australia weather station data and river
-level data as examples to demonstrate the use of \pkg{cubble}. An
-example on how \pkg{cubble} handles NetCDF data is also provided.
-Section 5 concludes the paper.
-
-\hypertarget{conceptual-framework-spatio-temporal-cube}{%
-\section{Conceptual framework: spatio-temporal
-cube}\label{conceptual-framework-spatio-temporal-cube}}
-
-Multivariate spatio-temporal data can be conceptualised using a cubical
-data model with reference to variables, time and space. This naturally
-motivates the usage of array to represent and store spatio-temporal
-data, as evident by satellite imageries, large climate models, among
-others. The work by \citet{lu_multidimensional_2018} has discussed the
-suitability of this representation through array algebra and practical
-spatio-temporal tasks.
-
-The cubical conceptual framework provides a generalisation to the
-operation can be done on the data for visualisation. A paper by
-\citet{bach_review_2014} has reviewed the temporal data visualisation
-based on space-time cube operations. Notice that the term space-time
-cube in their article ``does not need to involve spatial data'', but
-referring to ``an abstract 2D substrate that is used to visualize data
-at a specific time''. Despite its main focus is on temporal data, the
-mindset of abstracting out data representation to construct
-visualisation is applicable to our spatio-temporal data manipulation and
-visualisation.
-
-\hypertarget{existing-work-and-new-challenges}{%
-\section{Existing work and new
-challenges}\label{existing-work-and-new-challenges}}
-
-To represent spatio-temporal data, \citet{spacetime} proposes four
-layouts: full grid layout (STF), sparse grid layout (STS), irregular
-layout (STI), and trajectory (STT). These layouts are implemented in the
-\pkg{spacetime} package with underlying class \pkg{sp} \citep{sp} and
-\pkg{xts} \citep{xts} for handling spatial and temporal variables. While
-these layouts and classes are still being used and well-maintained
-today, they are not compatible with the \pkg{tidyverse} ecosystem
-\citep{tidyverse}, which has a huge influence on data analysis in
-\proglang{R}, including in the spatial and temporal communities. The
-package \pkg{sf} \citep{sf} uses data frame objects to represent spatial
-data with an additional \code{geometry} column dedicating to feature
-geometries \citep{jr_herring_opengis_2011} (e.g.~POINT, MULTIPOINT,
-POLYGON, MULTIPOLYGON etc). For temporal data, \pkg{tsibble}
-\citep{tsibble} builds on top of the \pkg{tibble} \citep{tibble} object
-and explicitly includes date and time in columns as variables, rather
-than data attributes.
-
-Underlying tidyverse is the concept of tidy data \citep{tidydata}, which
-considers how variables should be structured for analysis. In the three
-formats (time-wide, space-wide, and long) summarised in the
-\citet{spacetime} paper, only the long form is considered as tidy. The
-main advantages of it over the time-wide and space-wide format is that
-the latter treat either time or location as variable names (Section 3.1
-in \citet{tidydata}) rather than values of variables, making them
-difficult to be wrangled. One problem with the long format is that it
-can be inefficient to store feature geometries, especially for large
-MULTIPOLYGON over long time period. This motivates a new tidyverse
-compatible data object to work with spatio-temporal data.
-
-\textbf{two challenages: multiple data sources + large spatio-temporal
-object} In additional to the shift into more tidyverse workflow for
-working with data, spatio-temporal data have some unique challenges. The
-first one being the split of variables in multiple tables. Often the
-time, spatio-temporal data is not presented in a single data object, but
-split into fragmented pieces from different sources. For example, map
-data recorded the shape of a collection of area of interest is provided
-by the government website; the geo-scientific data is recorded with
-longitude and latitude coordinate of locations in the area; and temporal
-variables indexed by both location and time need to be further queried
-using location data. For spatio-temporal data analysis, merging all
-these sources can be challenging and joining these tables can go wrong
-because of the slightest unmatched of the linking variable from
-different sources.
-
-Another challenge when analysing spatio-temporal data is on the simple
-feature representation of spatial data. While the simple feature is able
-to handle large and complex spatial geometry and perform precise spatial
-operations, data analysts usually do not have a large expectation on the
-map accuracy and would prefer a simplified map with quick rendering for
-visualisation. Also, the feature geometry column in \pkg{sf} wraps the
-spatial coordinates in a list column, sometimes, analysts may like to
-unpack them into longitude and latitude to work with. A smooth
-transition with the coordinate columns and and simple feature geometry
-column would satisfy analysts from both ends.
-
-With some of the aforementioned issues recognised, the \pkg{stars}
-package has been a new implementation to handle spatio-temporal data
-conceptually as data cubes and implemented as arrays. The package
-depends on the \pkg{sf} package and has allowed more access to tidyverse
-verbs in the workflow. However, for vector spatio-temporal data, major
-difficulties to work with \pkg{stars} is first to cast the variables
-from multiple tables into a star object and then keep the operations in
-the high-dimensional cube. Given that most table mentioned before
-collect variables in a 2D table, analysts could benefit from a structure
-to allow them work in 2D tables.
-
-\hypertarget{the-cubble-package}{%
-\section{The cubble package}\label{the-cubble-package}}
-
-This section will present two formats that cubble uses to arrange
-spatio-temporal data. The main cubble functions will then be introduced
-to illustrate how to work with these two formats in short examples.
-Lastly, a subsection will be dedicated to how existing packages, in
-spatial and temporal analysis, fit in with cubble.
-
-In cubble, a single data object can be arranged into two forms: nested
-form and long form. Figure \ref{fig:illu-cubble} sketches the two forms
-with the associated attributes. The decision on which form to use in
-data wrangling is output-oriented, meaning analysts need to first think
-about whether the output of an operation is identified only by the
-spatial identifier, or a combination of spatial and temporal identifier.
-The nested cubble is suitable for working with operations that are only
-identified by the site and this type of operation can be a pure
-manipulation of time invariant variables, or an operation that
-summarises time varying variables into site. Underneath the nested form,
-a cubble is built from a \code{rowwise_df} class where each site forms a
-separate group. This structure simplifies the calculation that involves
-temporal variables by avoiding the use of \code{purrr::map()} syntax
-when working with list-column.
-
-For those operations whose output involves both a spatial and temporal
-identifier, long form should be used. The long form uses a
-\code{grouped_df} class to forms all the time of a site as a group. Time
-invariant variables are stored separately as an special attribute of the
-long cubble. This design avoids repeating the spatial variables at each
-time stamp while not dropping information from spatial variables.
-
-\begin{CodeChunk}
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.002}
-
-}
-
-\caption{An illustration of the nested and long form in cubble.
-The nested form defines each site in a row and nests the time varying variables into a single column \code{ts}.
-The long form cubble uses \code{id} and \code{t} to identify each row and stores the time invariant variables as an attribute, \code{spatial}.}\label{fig:illu-cubble}
-\end{figure}
-\end{CodeChunk}
-
-\hypertarget{create-a-cubble-in-the-nested-form}{%
-\subsection{Create a cubble in the nested
-form}\label{create-a-cubble-in-the-nested-form}}
-
-To use functions in the \texttt{cubble} package, an analyst will first
-need to turn the data object into a \code{cubble} class.
-\code{as_cubble()} does this through supplying the three key components:
-\code{key} as the spatial identifier; \code{index} as the temporal
-identifier; and a vector of \code{coords} in the order of longitude and
-latitude. The use of \code{key} and \code{index} follows the design in
-the \pkg{tsibble} package and the cubble created by default is in the
-nested form.
-
-Before the formal examples in Section \ref{examples}, each function
-introduced in this section will be accompanied by a small example. The
-data used for the small examples is a subset of a larger climate weather
-station data used in the formal examples and has been simplified to
-include only five weather stations. It contains spatial information of
-each station: station id, latitude, longitude, elevation, station name
-and World Meteorology Organisation ID and also daily temporal
-information of date, maximum and minimum temperature and precipitation
-for 2020. \code{climate_flat} stores this data in format 1) in Figure
-\ref{fig:illu-input} and the code below creates a cubble out of
-\code{climate_flat} with \code{id} as the key, \code{date} as the index,
-and \code{c(long, lat)} as the coordinates:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> cubble_nested <- cubble::climate_flat |>
-+   as_cubble(key = id, index = date, coords = c(long, lat))
-R> cubble_nested
-\end{CodeInput}
-\begin{CodeOutput}
-# cubble:   id [5]: nested form
-# bbox:     [115.97, -32.94, 133.55, -12.42]
-# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
-  id            lat  long  elev name           wmo_id ts                
-  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>            
-1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tibble [366 x 4]>
-2 ASN00010311 -31.9  117. 179   york            94623 <tibble [366 x 4]>
-3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble [366 x 4]>
-4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble [366 x 4]>
-5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble [366 x 4]>
-\end{CodeOutput}
-\end{CodeChunk}
-
-There are a few information in the cubble header: the name of the
-\code{key} variable: \code{id} and its unique number: 5, the bounding
-box: \code{[115.97, -32.94, 133.55, -12.42]} , and also the name of
-variable nested in the \code{ts} column with its type:
-\code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.
-
-\hypertarget{stretch-a-nested-cubble-into-the-long-form}{%
-\subsection{Stretch a nested cubble into the long
-form}\label{stretch-a-nested-cubble-into-the-long-form}}
-
-The nested format is convenient for those operations whose output does
-not contain a time dimension. For those outputs that are
-cross-identified by the spatial and temporal identifier, a long cubble
-needs to be used. The function \code{stretch()} is designed to switch
-the cubble from the nested form to the long form. This code shows how to
-switch the nested cubble just created into its long form:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> cubble_long <- cubble_nested |> face_temporal()
-R> cubble_long
-\end{CodeInput}
-\begin{CodeOutput}
-# cubble:  date, id [5]: long form
-# bbox:    [115.97, -32.94, 133.55, -12.42]
-# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]
-  id          date        prcp  tmax  tmin
-  <chr>       <date>     <dbl> <dbl> <dbl>
-1 ASN00009021 2020-01-01     0  31.9  15.3
-2 ASN00009021 2020-01-02     0  24.9  16.4
-3 ASN00009021 2020-01-03     6  23.2  13  
-4 ASN00009021 2020-01-04     0  28.4  12.4
-5 ASN00009021 2020-01-05     0  35.3  11.6
-# ... with 1,825 more rows
-\end{CodeOutput}
-\end{CodeChunk}
-
-Notice that the third line in the header now shows the name and type of
-spatial variables:
-\code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]} and
-these variables are now stored as a \code{spatial} attribute of the
-data:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> attr(cubble_long, "spatial")
-\end{CodeInput}
-\begin{CodeOutput}
-# A tibble: 5 x 6
-  id            lat  long  elev name           wmo_id
-  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl>
-1 ASN00009021 -31.9  116.  15.4 perth airport   94610
-2 ASN00010311 -31.9  117. 179   york            94623
-3 ASN00010614 -32.9  117. 338   narrogin        94627
-4 ASN00014015 -12.4  131.  30.4 darwin airport  94120
-5 ASN00015131 -17.6  134. 220   elliott         94236
-\end{CodeOutput}
-\end{CodeChunk}
-
-\hypertarget{tamp-a-long-cubble-back-to-the-nested-form}{%
-\subsection{Tamp a long cubble back to the nested
-form}\label{tamp-a-long-cubble-back-to-the-nested-form}}
-
-Manipulation on the spatial and temporal dimension can be an iterative
-process and analysts may need to go back and forth between the nested
-and long cubble. \code{tamp()}, an inverse of \code{stretch()}, switches
-a long cubble to a nested cubble:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> cubble_back <- cubble_long |> face_spatial()
-R> cubble_back
-\end{CodeInput}
-\begin{CodeOutput}
-# cubble:   id [5]: nested form
-# bbox:     [115.97, -32.94, 133.55, -12.42]
-# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
-  id            lat  long  elev name           wmo_id ts                
-  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>            
-1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tibble [366 x 4]>
-2 ASN00010311 -31.9  117. 179   york            94623 <tibble [366 x 4]>
-3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble [366 x 4]>
-4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble [366 x 4]>
-5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble [366 x 4]>
-\end{CodeOutput}
-\end{CodeChunk}
-
-\hypertarget{migrate-spatial-variables-to-a-long-cubble}{%
-\subsection{Migrate spatial variables to a long
-cubble}\label{migrate-spatial-variables-to-a-long-cubble}}
-
-Sometimes, analysts may need to apply some variable transformation that
-involves both the spatial and temporal variable. An example of this is
-the transformation of temporal variables into the spatial dimension in
-glyph maps (which will be elaborated in section
-\ref{st_transformation}). Cubble allows this operation through
-\code{unfold()}, which moves the specified spatial variables into the
-long cubble:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> cubble_mig <- cubble_long |> unfold(long, lat)
-R> cubble_mig
-\end{CodeInput}
-\begin{CodeOutput}
-# cubble:  date, id [5]: long form
-# bbox:    [115.97, -32.94, 133.55, -12.42]
-# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]
-  id          date        prcp  tmax  tmin  long   lat
-  <chr>       <date>     <dbl> <dbl> <dbl> <dbl> <dbl>
-1 ASN00009021 2020-01-01     0  31.9  15.3  116. -31.9
-2 ASN00009021 2020-01-02     0  24.9  16.4  116. -31.9
-3 ASN00009021 2020-01-03     6  23.2  13    116. -31.9
-4 ASN00009021 2020-01-04     0  28.4  12.4  116. -31.9
-5 ASN00009021 2020-01-05     0  35.3  11.6  116. -31.9
-# ... with 1,825 more rows
-\end{CodeOutput}
-\end{CodeChunk}
-
-This function should generally be used in the last step of the analysis
-since it is a temporary operation, meaning these added spatial variables
-are not stored in the long form and will disappear if switched to the
-nested form and then switched back:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> cubble_mig |> face_spatial() |> face_temporal()
-\end{CodeInput}
-\begin{CodeOutput}
-# cubble:  date, id [5]: long form
-# bbox:    [115.97, -32.94, 133.55, -12.42]
-# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]
-  id          date        prcp  tmax  tmin
-  <chr>       <date>     <dbl> <dbl> <dbl>
-1 ASN00009021 2020-01-01     0  31.9  15.3
-2 ASN00009021 2020-01-02     0  24.9  16.4
-3 ASN00009021 2020-01-03     6  23.2  13  
-4 ASN00009021 2020-01-04     0  28.4  12.4
-5 ASN00009021 2020-01-05     0  35.3  11.6
-# ... with 1,825 more rows
-\end{CodeOutput}
-\end{CodeChunk}
-
-\hypertarget{compatibility-with-existing-packages}{%
-\subsection{Compatibility with existing
-packages}\label{compatibility-with-existing-packages}}
-
-The previous four subsections have introduced operations specific to the
-\code{cubble} class and this section will demonstrate how the
-\code{cubble} class interacts with existing packages commonly used in
-spatial and temporal analysis, specifically, \code{dplyr},
-\code{tsibble}, \code{sf} (\code{s2}), and \code{netcdf4}.
-
-\hypertarget{dplyr}{%
-\subsubsection{dplyr}\label{dplyr}}
-
-The \code{dplyr} package has provided many tools for data wrangling
-tasks and these operations are useful in the spatio-temporal context.
-\code{cubble} provides methods that support the following \code{dplyr}
-verbs in both the nested and long form:
-
-\begin{quote}
-\texttt{mutate}, \texttt{filter}, \texttt{summarise}, \texttt{select},
-\texttt{arrange}, \texttt{rename}, \texttt{left\_join}, and the slice
-family (\texttt{slice\_head}, \texttt{slice\_tail},
-\texttt{slice\_sample}, \texttt{slice\_min}, \texttt{slice\_max})
-\end{quote}
-
-\hypertarget{tsibble}{%
-\subsubsection{tsibble}\label{tsibble}}
-
-\code{tsibble} is a temporal data structure that uses \code{index} and
-\code{key} to identify the time and different series. \code{cubble} can
-be seen as a natural extension of \code{tsibble} for spatio-temporal
-data with an additional coordinates component and using two forms to
-arrange variables. This makes it easy to cast a \code{tsibble} into a
-\code{cubble} as only the \code{coords} argument needs to be supplied:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> # example with a tsibble created from climate_flat
-R> raw <- climate_flat |>
-+   tsibble::as_tsibble(key = id, index = date)
-R> 
-R> dt <-  raw |>
-+   cubble::as_cubble(coords = c(long, lat))
-R> dt
-\end{CodeInput}
-\begin{CodeOutput}
-# cubble:   id [5]: nested form
-# bbox:     [115.97, -32.94, 133.55, -12.42]
-# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
-  id            lat  long  elev name           wmo_id ts                
-  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>            
-1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tbl_ts [366 x 4]>
-2 ASN00010311 -31.9  117. 179   york            94623 <tbl_ts [366 x 4]>
-3 ASN00010614 -32.9  117. 338   narrogin        94627 <tbl_ts [366 x 4]>
-4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tbl_ts [366 x 4]>
-5 ASN00015131 -17.6  134. 220   elliott         94236 <tbl_ts [366 x 4]>
-\end{CodeOutput}
-\end{CodeChunk}
-
-In the nested cubble created, each element in the list-column \code{ts}
-is of \code{tbl_ts} class and operations available to the tsibble class
-is still valid under cubble. For example, the code below calculates two
-features of the maximum temperature:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> # add station-based features in the nested form.
-R> dt |>
-+   mutate(fabletools::features(ts, tmax, list(tmax_mean = mean, tmax_var = var)))
-\end{CodeInput}
-\begin{CodeOutput}
-# cubble:   id [5]: nested form
-# bbox:     [115.97, -32.94, 133.55, -12.42]
-# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
-  id            lat  long  elev name          wmo_id ts       tmax_mean tmax_var
-  <chr>       <dbl> <dbl> <dbl> <chr>          <dbl> <list>       <dbl>    <dbl>
-1 ASN00009021 -31.9  116.  15.4 perth airport  94610 <tbl_ts>      25.7    38.6 
-2 ASN00010311 -31.9  117. 179   york           94623 <tbl_ts>      26.2    51.1 
-3 ASN00010614 -32.9  117. 338   narrogin       94627 <tbl_ts>      23.7    45.4 
-4 ASN00014015 -12.4  131.  30.4 darwin airpo~  94120 <tbl_ts>      33.1     3.02
-5 ASN00015131 -17.6  134. 220   elliott        94236 <tbl_ts>      34.6    24.7 
-\end{CodeOutput}
-\end{CodeChunk}
-
-\hypertarget{sf-and-s2}{%
-\subsubsection{sf and s2}\label{sf-and-s2}}
-
-As a spatial data object, \code{sf} creates a simple feature geometry
-list-column (\code{sfc}) in the data frame to provide spatial operations
-on various geometry types (\code{POINT}, \code{LINESTRING},
-\code{POLYGON}, \code{MULTIPOLYGON}, etc). These spatial operations are
-also valuable for spatio-temporal data analysis, but an \code{sf} object
-\emph{usually} does not contain temporal variables. This means \code{sf}
-cannot be directly cast into a \code{cubble}, however, \code{cubble}
-does support \code{sfc} columns in the nested form and spatial
-operations applied to the \code{sfc} column in \code{sf} can still be
-applied to the \code{sfc} column in a cubble. The following example
-shows how to create an \code{sfc} column of \code{POINT} type from
-latitude and longitude in cubble. Then \code{sf::st_within} is used to
-add the state \code{MULTIPOLYGON} of each weather station before a
-coordinate transformation is made.
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> library(sf)
-R> # create a cubble
-R> cb <- climate_flat |>
-+   cubble::as_cubble(key = id, index = date, coords = c(long, lat))
-R> 
-R> aus <- ozmaps::abs_ste
-R> 
-R> dt <- cb |>
-+   mutate(
-+     # create `sfc` column based on long and lat
-+     ll = st_sfc(
-+       purrr::map2(long, lat, ~st_point(c(.x, .y))),
-+       crs = st_crs(aus)),
-+ 
-+     # append state multi-polygon based on the `sfc` created
-+     state = aus$geometry[st_within(ll, aus, sparse = FALSE)],
-+ 
-+     # adopt a different projection: lambert conformal conic (EPSG:3112)
-+     state = st_transform(state, crs = "EPSG:3112")
-+     )
-R> 
-R> dt
-\end{CodeInput}
-\begin{CodeOutput}
-# cubble:   id [5]: nested form
-# bbox:     [115.97, -32.94, 133.55, -12.42]
-# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
-  id            lat  long  elev name   wmo_id ts                        ll
-  <chr>       <dbl> <dbl> <dbl> <chr>   <dbl> <list>           <POINT [°]>
-1 ASN00009021 -31.9  116.  15.4 perth~  94610 <tibble> (115.9764 -31.9275)
-2 ASN00010311 -31.9  117. 179   york    94623 <tibble>  (116.765 -31.8997)
-3 ASN00010614 -32.9  117. 338   narro~  94627 <tibble> (117.1797 -32.9342)
-4 ASN00014015 -12.4  131.  30.4 darwi~  94120 <tibble> (130.8925 -12.4239)
-5 ASN00015131 -17.6  134. 220   ellio~  94236 <tibble> (133.5407 -17.5521)
-# ... with 1 more variable: state <MULTIPOLYGON [m]>
-\end{CodeOutput}
-\end{CodeChunk}
-
-An \pkg{s2} \code{lnglat} vector can similarly be created as an
-\code{sfc} in cubble before using any \code{s2}-prefixed function:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> library(s2)
-R> # Western Australia map
-R> wa <- ozmaps::abs_ste |> filter(NAME == "Western Australia")
-R> 
-R> # mutate a `s2_lnglat` vector on `cb` created in the last chunk
-R> cb |>
-+   mutate(ll = s2_lnglat(long, lat)) |>
-+   filter(s2_within(ll, wa))
-\end{CodeInput}
-\begin{CodeOutput}
-# cubble:   id [3]: nested form
-# bbox:     [115.97, -32.94, 117.18, -31.89]
-# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
-  id            lat  long  elev name          wmo_id ts                 ll      
-  <chr>       <dbl> <dbl> <dbl> <chr>          <dbl> <list>             <s2_lng>
-1 ASN00009021 -31.9  116.  15.4 perth airport  94610 <tibble [366 x 4]> (115.97~
-2 ASN00010311 -31.9  117. 179   york           94623 <tibble [366 x 4]> (116.76~
-3 ASN00010614 -32.9  117. 338   narrogin       94627 <tibble [366 x 4]> (117.17~
-\end{CodeOutput}
-\end{CodeChunk}
-
-\hypertarget{netcdf}{%
-\subsubsection{netcdf}\label{netcdf}}
-
-NetCDF data has two main components: \emph{dimension} for defining the
-spatio-temporal grid (longitude, latitude, and time) and \emph{variable}
-that populates the defined grid. Attributes are usually associated with
-dimension and variable in the NetCDF format data and a
-\href{http://cfconventions.org/}{metadata convention for climate and
-forecast} has been designed to standardise the format of the attributes.
-A few packages in R exists for manipulating NetCDF data and this
-includes a high-level R interface: \pkg{ncdf4} \citep{ncdf4}, a
-low-level interface that calls C-interface: \pkg{RNetCDF}
-\citep{rnetcdf, michna2013rnetcdf}, and a tidyverse implementation:
-\pkg{tidync} \citep{tidync}.
-
-Cubble provides an \code{as_cubble()} method to coerce the \code{ncdf4}
-class from the \pkg{ncdf4} package into a \code{cubble}. It maps each
-combination of longitude and latitude into an \code{id} as the
-\code{key}:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> # read in the .nc file as a ncdf4 class
-R> raw <- ncdf4::nc_open(here::here("data/era5-pressure.nc"))
-R> 
-R> # convert the variable q and z in the ncdf4 into a cubble
-R> dt <- as_cubble(raw, vars = c("q", "z"))
-\end{CodeInput}
-\end{CodeChunk}
-
-The memory limit with NetCDF data in cubble depends on the number of
-longitude grid points times the number of latitude grid points times the
-number of time grid points times the number of variables. Cubble can
-handle slightly more than 300 x 300 (longitude x longitude) grid points
-for 3 variables in one year. The spatial grid can be reduced to trade
-for longer time periods and more variables. A 300 by 300 spatial grid
-can be a bounding box of {[}100, -80, 180, 0{]} at 0.25 degree
-resolution or a global bounding box {[}-180, -90, 180, -90{]} at 1
-degree resolution. Subsetting longitude and latitude grids is available
-through \code{long_range} and \code{lat_range} if the NetCDF file has
-finer resolution than needed.
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> # Assume my_ncdf has bounding box of [-180, -90, 180, -90]
-R> # at 0.25 degree resolution and subset it to have
-R> # 1 degree resolution:
-R> dt <- as_cubble(my_ncdf, vars = c("q", "z"),
-+                 long_range = seq(-180, 180, 1),
-+                 lat_range = seq(-90, 90, 1))
-\end{CodeInput}
-\end{CodeChunk}
-
-\hypertarget{other-features-and-considerations}{%
-\section{Other features and
-considerations}\label{other-features-and-considerations}}
-
-\hypertarget{hierarchical-structure}{%
-\subsection{Hierarchical structure}\label{hierarchical-structure}}
-
-Spatial locations can have grouping structure either inherent to the
-data or formed by clustering. Rather than analysing variables in the
-site level, summarised variables in the cluster level can give a crisper
-picture of local areas. In cubble, \code{switch_key()} can be used to
-create a new level of grouping of spatial locations by specifying a
-clustering variable. Figure \ref{fig:illu-hier} illustrates the
-relationship of cubbles at station and cluster level, in both the long
-and nested form. By specifying
-\code{cluster_nested <- station_nested \%>\% switch_key(key = cluster)},
-the cubble re-defines the cubble key from \code{id} in
-\code{station_nested} to \code{cluster} in \code{cluster_nested}. All
-the spatial variables variant to \code{cluster} are now nested into a
-\code{.val} column and cluster level variables can be computed in the
-same fashion as station level variables in \code{station_nested}.
-
-\begin{CodeChunk}
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.003}
-
-}
-
-\caption[An illustration of the original and cluster level cubble in the nested form long form for hierarchical structure data]{An illustration of the original and cluster level cubble in the nested form long form for hierarchical structure data.
-\code{switch\_key()} changes the station level cubble into a cluster level cubble and both can be stretched into the long form.}\label{fig:illu-hier}
-\end{figure}
-\end{CodeChunk}
-
-\hypertarget{data-fusion-and-matching}{%
-\subsection{Data fusion and matching}\label{data-fusion-and-matching}}
-
-One task that may interest analysts in spatio-temporal data is to find
-how similar the time series from nearby sites are. This can be seen as a
-matching problem \citep{stuart2010matching, mcintosh2018using} that
-pairs up similar similar time series in nearby locations or a data
-fusing exercise that merges data collected from different sources
-\citep{cocchi2019data}. \code{match_sites()} in \pkg{cubble} provides a
-simple algorithm for this task. The algorithm first matches the two data
-sources spatially through computing the pairwise distance on latitude
-and longitude. Pairs that pass the spatial matching are then matched
-temporally through computing the number of matched peaks within a fixed
-length window. Figure \ref{fig:illu-matching} illustrates this temporal
-matching in more details. Given two series \code{A} and \code{a}, 3
-peaks have been picked in each series. An interval, with default length
-of 5, is constructed for each peak in series \code{A} and the peaks in
-series \code{a} are tested against whether they fall into the any of the
-intervals. In this illustration, there are 2 matches for these two
-series. Several arguments are available in \code{match_sites()} to
-fine-tune the matching:
-
-\begin{CodeChunk}
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{figures/illu-matching}
-
-}
-
-\caption{An illustration of temporal matching in cubble.
-Three highest peaks are identified in each series and intervals are constructed on series \code{A}.
-Two peaks in series \code{a} fall into the intervals and hence the two series are considered to have 2 matches.}\label{fig:illu-matching}
-\end{figure}
-\end{CodeChunk}
+The cubble class is an S3 class built on tibble that allows the
+spatio-temporal data to be wrangled in two forms: a nested form (spatial
+cubble) and a long form (temporal cubble). It consists of two
+subclasses:
 
@@
 \item
-  \code{spatial_n_keep}: the number of spatial match for each site to
-  keep
+  a spatial cubble with class
+  \texttt{c("spatial\_cubble\_df",\ "cubble\_df")}
 \item
-  \code{spatial_dist_max}: the maximum distance allowed for a matched
-  pair
-\item
-  \code{temporal_n_highest}: the number of peak used - 3 in the example
-  above
-\item
-  \code{temporal_window}: the length of the interval - 5 in the example
-  above
-\item
-  \code{temporal_min_match}: the minimum number of matched peak for a
-  valid matched pair
+  a emporal cubble with class
+  \texttt{c("temporal\_cubble\_df",\ "cubble\_df")}
 \end{itemize}
+
+In a spatial cubble, spaital variables are organised as columns and
+temporal variables are nested within a specialised \texttt{ts} column.
+The cubble object printed below contains weather records in three
+airport stations (the creation of a cubble object will be explained in
+Section @ref(create)). This toy data is a subset of a larger data
+\code{climate_aus} collected from Global Historical Climatology Network
+Daily (GHCND). It records three airport stations located in Melbourne,
+Australia and includes spatial variables such as station ID, longitude,
+latitude, elevation, station name, World Meteorology Organisation ID.
+The dataset contains temporal variables including precipitation, maximum
+and minimum temperature, which can be read from the cubble header:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> cb_nested
+\end{CodeInput}
+\begin{CodeOutput}
+# cubble:   key: id [3], index: date, nested form
+# spatial:  [144.8321, -37.98, 145.0964, -37.6655], Missing CRS!
+# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
+  id           long   lat  elev name              wmo_id ts               
+  <chr>       <dbl> <dbl> <dbl> <chr>              <dbl> <list>           
+1 ASN00086038  145. -37.7  78.4 essendon airport   95866 <tibble [10 x 4]>
+2 ASN00086077  145. -38.0  12.1 moorabbin airport  94870 <tibble [10 x 4]>
+3 ASN00086282  145. -37.7 113.  melbourne airport  94866 <tibble [10 x 4]>
+\end{CodeOutput}
+\end{CodeChunk}
+
+In a temporal cubble, temporal variables are expanded in the long form
+and spatial variables are stored as a data attribute:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> cb_long
+\end{CodeInput}
+\begin{CodeOutput}
+# cubble:   key: id [3], index: date, long form
+# temporal: 2020-01-01 -- 2020-01-10 [1D], no gaps
+# spatial:  long [dbl], lat [dbl], elev [dbl], name [chr], wmo_id [dbl]
+  id          date        prcp  tmax  tmin
+  <chr>       <date>     <dbl> <dbl> <dbl>
+1 ASN00086038 2020-01-01     0  26.8  11  
+2 ASN00086038 2020-01-02     0  26.3  12.2
+3 ASN00086038 2020-01-03     0  34.5  12.7
+4 ASN00086038 2020-01-04     0  29.3  18.8
+5 ASN00086038 2020-01-05    18  16.1  12.5
+# i 25 more rows
+\end{CodeOutput}
+\end{CodeChunk}
+
+\hypertarget{the-cubble-attributes}{%
+\subsubsection{The cubble attributes}\label{the-cubble-attributes}}
+
+A cubble object inherits the attributes from tibble (and its
+subclasses): \texttt{class}, \texttt{row.names}, and \texttt{names}.
+Additionally, it has three specialised attributes: \texttt{key},
+\texttt{index}, and \texttt{coords}. Readers who are familiar with the
+\texttt{key} and \texttt{index} attributes from the \texttt{tsibble}
+package would already understand the two arguments. In cubble, the
+\texttt{key} attribute identifies the row in the spatial cubble (given
+the use of \texttt{tidyr::nest()} for nesting), and when combined with
+the \texttt{index} argument, it identifies the row in the temporal
+cubble. Currently, cubble only supports one variable as the key, and the
+accepted temporal classes for index includes the base R classes
+\texttt{Date}, \texttt{POSIXlt}, \texttt{POSIXct}, as well as tsibble's
+\texttt{yearmonth}, \texttt{yearweek}, and \texttt{yearquarter} classes.
+
+The \texttt{coords} attribute represents an ordered pair of coordinates.
+It can be either an unprojected pair of longitude and latitude, or a
+projected easting and northing value. The \texttt{sf} package is used
+under the hood to calculate the bounding box, displayed in the header of
+a spatial cubble.
+
+The temporal cubble has a special attribute called \texttt{spatial} to
+store the spatial variables. Shortcut functions are available to extract
+attributes, for example, \code{spatial()} for extracting spatial
+variables from the temporal cubble:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> spatial(cb_long)
+\end{CodeInput}
+\begin{CodeOutput}
+# A tibble: 3 x 6
+  id           long   lat  elev name              wmo_id
+  <chr>       <dbl> <dbl> <dbl> <chr>              <dbl>
+1 ASN00086038  145. -37.7  78.4 essendon airport   95866
+2 ASN00086077  145. -38.0  12.1 moorabbin airport  94870
+3 ASN00086282  145. -37.7 113.  melbourne airport  94866
+\end{CodeOutput}
+\end{CodeChunk}
+
+\hypertarget{create}{%
+\subsection{Creation and coercion}\label{create}}
+
+In many cases, spatio-temporal data arrive in separate tables for
+analysis. For example, in climate data, analysts may initially receive
+station data containing geographic location information, recorded
+variables and their recording periods. They can then query the temporal
+variables using the stations of interest to obtain the relevant temporal
+data. Alternatively, analyses may begin as purely spatial or temporal,
+and analysts may obtain additional temporal or spatial data to expand
+the result to spatio-temporal.
+
+The function \code{make_cubble()} composes a \code{cubble} object from a
+spatial table (\code{spatial}) and a temporal table (\code{temporal}),
+along with three attributes introduced in the subsection @ref(object):
+\code{key}, \code{index}, and \code{coords}. The following code creates
+the spatial \code{cubble} object shown previously:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> make_cubble(spatial = stations, temporal = meteo,
++             key = id, index = date, coords = c(long, lat))
+\end{CodeInput}
+\end{CodeChunk}
+
+Spatio-temporal data in other foreign objects can be coerced into a
+\code{cubble} object with the function \code{as_cubble()}. This includes
+casting from a \code{tibble} or \code{data.frame} with both spatial and
+temporal information, a NetCDF object, a \code{stars} object
+\citep{stars}, and a \code{sftime} object \citep{sftime}. The two
+examples below show the casting from a tibble and a NetCDF object. The
+dataset \code{climate_flat} combines the spatial data, \code{stations},
+with the temporal data, \code{meteo}, into a single tibble object. It
+can be coerced into a cubble using:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> climate_flat |> as_cubble(key = id, index = date, coords = c(long, lat))
+\end{CodeInput}
+\begin{CodeOutput}
+# cubble:   key: id [3], index: date, nested form
+# spatial:  [144.8321, -37.98, 145.0964, -37.6655], Missing CRS!
+# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
+  id           long   lat  elev name              wmo_id ts               
+  <chr>       <dbl> <dbl> <dbl> <chr>              <dbl> <list>           
+1 ASN00086038  145. -37.7  78.4 essendon airport   95866 <tibble [10 x 4]>
+2 ASN00086077  145. -38.0  12.1 moorabbin airport  94870 <tibble [10 x 4]>
+3 ASN00086282  145. -37.7 113.  melbourne airport  94866 <tibble [10 x 4]>
+\end{CodeOutput}
+\end{CodeChunk}
+
+\hypertarget{functions-and-methods}{%
+\subsection{Functions and methods}\label{functions-and-methods}}
+
+Table @ref(tab:funs) summarises the functions implemented in the
+\pkg{cubble} package and Table @ref(tab:methods) details the methods
+implemented for each of the three cubble classes. The \code{cubble_df}
+class handles methods that behave consistently in both spatial
+(\code{spatial_cubble_df}) and temporal cubble
+(\code{temporal_cubble_df}). When the method has distinct behavior for
+the spatial cubble and temporal cubble, it is implemented separately for
+each subclass.
+
+\begin{longtable}[]{@{}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.1250}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.8750}}@{}}
+\caption{(\#tab:funs) Functions implemented in the \pkg{cubble}
+package}\tabularnewline
+\toprule()
+\begin{minipage}[b]{\linewidth}\raggedright
+Category
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Functions
+\end{minipage} \\
+\midrule()
+\endfirsthead
+\toprule()
+\begin{minipage}[b]{\linewidth}\raggedright
+Category
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Functions
+\end{minipage} \\
+\midrule()
+\endhead
+base R & \texttt{{[}}, \texttt{{[}{[}\textless{}-},
+\texttt{names\textless{}-} \\
+tidyverse & \texttt{dplyr\_row\_slice}, \texttt{dplyr\_col\_modify},
+\texttt{dplyr\_reconstruct}, \texttt{select}, \texttt{mutate},
+\texttt{arrange}, \texttt{filter}, \texttt{group\_by}, \texttt{ungroup},
+\texttt{summarise}, \texttt{select}, \texttt{slice}, \texttt{rowwise},
+\texttt{rename}, \texttt{bind\_rows}, \texttt{bind\_cols},
+\texttt{relocate}, \texttt{type\_sum}, the slice family
+(\texttt{slice\_head}, \texttt{slice\_tail}, \texttt{slice\_max},
+\texttt{slice\_min}, \texttt{slice\_sample}) and the join family
+(\texttt{left\_join}, \texttt{right\_join}, \texttt{inner\_join},
+\texttt{full\_join}, \texttt{anti\_join}, \texttt{semi\_join}) \\
+cubble & \texttt{as\_cubble}, \texttt{cubble}, \texttt{make\_cubble},
+\texttt{check\_key} \texttt{face\_temporal}, \texttt{face\_spatial},
+\texttt{unfold}, \texttt{key}, \texttt{key\_vars}, \texttt{key\_data},
+\texttt{index}, \texttt{index\_var}, \texttt{coords}, \texttt{spatial},
+\texttt{match\_sites}, \texttt{match\_spatial},
+\texttt{match\_temporal}, \texttt{geom\_glyph},
+\texttt{geom\_glyph\_box}, \texttt{geom\_glyph\_line},
+\texttt{make\_spatial\_sf}, \texttt{make\_temporal\_tsibble},
+\texttt{fill\_gaps}, and \texttt{scan\_gaps} \\
+\bottomrule()
+\end{longtable}
+
+\begin{longtable}[]{@{}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.2500}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.7500}}@{}}
+\caption{(\#tab:methods) Methods implemented for the three \code{cubble}
+classes.}\tabularnewline
+\toprule()
+\begin{minipage}[b]{\linewidth}\raggedright
+Class
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Methods
+\end{minipage} \\
+\midrule()
+\endfirsthead
+\toprule()
+\begin{minipage}[b]{\linewidth}\raggedright
+Class
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Methods
+\end{minipage} \\
+\midrule()
+\endhead
+\texttt{cubble\_df} &
+\texttt{{[}{[}\textless{}-,\ dplyr\_col\_modify,\ key\_data,\ key\_vars,\ key,\ print} \\
+\texttt{spatial\_cubble\_df} &
+\texttt{{[},\ names\textless{}-,\ \ tbl\_sum,\ dplyr\_reconstruct,\ dplyr\_row\_slice,\ face\_spatial,\ face\_temporal,\ unfold,\ arrange,\ rename,\ rowwise,\ group\_by,\ ungroup,\ select,\ spatial,\ summarise,\ unfold,\ update\_cubble} \\
+\texttt{temporal\_cubble\_df} &
+\texttt{{[},\ names\textless{}-,\ tbl\_sum,\ arrange,\ dplyr\_reconstruct,\ dplyr\_row\_slice,\ face\_spatial,\ face\_temporal,\ unfold,\ fill\_gaps,\ group\_by,\ ungroup,\ \ rename,\ rowwise,\ scan\_gaps,\ select,\ spatial,\ summarise,\ tbl\_sum,\ bind\_rows,\ bind\_cols,\ update\_cubble} \\
+\bottomrule()
+\end{longtable}
+
+Specifically, the pair of verbs, \code{face_temporal()} and
+\code{face_spatial()}, pivot the cubble object between the spatial and
+temporal face of the multivariate spatio-temporal cube, as illustrated
+in Figure \ref{fig:face}. The code below uses the function
+\code{face_temporal()} and \code{face_spatial()} to transform between a
+spatial cubble (\code{cb_nested}) and a temporal one (\code{cb_long}),
+introduced earlier in subsection @ref(object):
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> identical(face_temporal(cb_nested), cb_long)
+\end{CodeInput}
+\begin{CodeOutput}
+[1] TRUE
+\end{CodeOutput}
+\begin{CodeInput}
+R> identical(face_spatial(cb_long), cb_nested)
+\end{CodeInput}
+\begin{CodeOutput}
+[1] TRUE
+\end{CodeOutput}
+\end{CodeChunk}
+
+The pair of verbs are exact inverse and apply both functions on a cubble
+object will result in the object itself:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> identical(face_spatial(face_temporal(cb_nested)), cb_nested)
+\end{CodeInput}
+\begin{CodeOutput}
+[1] TRUE
+\end{CodeOutput}
+\begin{CodeInput}
+R> identical(face_temporal(face_spatial(cb_long)), cb_long)
+\end{CodeInput}
+\begin{CodeOutput}
+[1] TRUE
+\end{CodeOutput}
+\end{CodeChunk}
+
+\begin{CodeChunk}
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{figures/diagram-keynotes/diagram-keynotes.001}
+
+}
+
+\caption{An illustration of the function \code{face\_temporal()} and \code{face\_spatial()}: \code{face\_temporal()} converts a spatial cubble (neseted form) into a temporal cubble (long form) to focus on the temporal variables.
+Conversely, \code{face\_spatial()} transforms a temporal cubble into a spatial one to emphasize spatial variables.}(\#fig:face)
+\end{figure}
+\end{CodeChunk}
+
+\hypertarget{compatibility-with-tsibble-and-sf}{%
+\subsection{Compatibility with tsibble and
+sf}\label{compatibility-with-tsibble-and-sf}}
+
+Analysts often have their preferred spatial or temporal data structure
+that they prefer to use for spatio-temporal analysis. For example, the
+\code{tbl_ts} class from the \code{tsibble} package \citep{tsibble} is
+commonly used in time series forecasting and similarly, the \code{sf}
+class \citep{sf} is frequently used in spatial data science. In
+\code{cubble}, analysts have the flexibility to combine these two
+structures together by allowing the spatial component to also be an
+\code{sf} object and the temporal component to also be a \code{tsibble}
+object.
+
+\hypertarget{using-a-tsibble-object-as-the-temporal-component}{%
+\subsubsection{Using a tsibble object as the temporal
+component}\label{using-a-tsibble-object-as-the-temporal-component}}
+
+The \code{key} and \code{index} arguments in a \code{cubble} object
+corresponds to the \code{tsibble} counterparts and they can be safely
+omitted, if the temporal component is a \code{tsibble} object,
+i.e.~\code{meteo_ts} in the example below. The \code{tsibble} class from
+the input will be carried over to the \code{cubble} object:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> ts_nested <- make_cubble(
++   spatial = stations, temporal = meteo_ts, coords = c(long, lat))
+R> (ts_long <- face_temporal(ts_nested))
+\end{CodeInput}
+\begin{CodeOutput}
+# cubble:   key: id [3], index: date, long form, [tsibble]
+# temporal: 2020-01-01 -- 2020-01-10 [1D], no gaps
+# spatial:  long [dbl], lat [dbl], elev [dbl], name [chr], wmo_id [dbl]
+  id          date        prcp  tmax  tmin
+  <chr>       <date>     <dbl> <dbl> <dbl>
+1 ASN00086038 2020-01-01     0  26.8  11  
+2 ASN00086038 2020-01-02     0  26.3  12.2
+3 ASN00086038 2020-01-03     0  34.5  12.7
+4 ASN00086038 2020-01-04     0  29.3  18.8
+5 ASN00086038 2020-01-05    18  16.1  12.5
+# i 25 more rows
+\end{CodeOutput}
+\begin{CodeInput}
+R> class(ts_long)
+\end{CodeInput}
+\begin{CodeOutput}
+[1] "temporal_cubble_df" "cubble_df"          "tbl_ts"            
+[4] "tbl_df"             "tbl"                "data.frame"        
+\end{CodeOutput}
+\end{CodeChunk}
+
+The temporal cubble shows \code{[tsibble]} in the header to indicate the
+object also being in a \code{tbl_ts} class (\code{tsibble}). Methods
+applies to the \code{tbl_ts} class can also be applied to the temporal
+cubble objects, for example, checking whether the data contain temporal
+gaps:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> ts_long %>% has_gaps()
+\end{CodeInput}
+\begin{CodeOutput}
+# A tibble: 3 x 2
+  id          .gaps
+  <chr>       <lgl>
+1 ASN00086038 FALSE
+2 ASN00086077 FALSE
+3 ASN00086282 FALSE
+\end{CodeOutput}
+\end{CodeChunk}
+
+An existing \code{cubble} object can promote its temporal component to a
+\code{tsibble} object by applying \code{make_temporal_tsibble()}. The
+following code illustrates this with the object \code{cb_long} created
+in section @ref(create) and the promoted cubble object is equal to the
+cubble object originally created from a \code{tsibble} object:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> ts_long2 <- cb_long %>% make_temporal_tsibble() 
+R> identical(ts_long2, ts_long)
+\end{CodeInput}
+\begin{CodeOutput}
+[1] TRUE
+\end{CodeOutput}
+\end{CodeChunk}
+
+\hypertarget{using-an-sf-object-as-the-spatial-component}{%
+\subsubsection{Using an sf object as the spatial
+component}\label{using-an-sf-object-as-the-spatial-component}}
+
+Similarly, an \code{sf} object can be supplied as the spatial component
+to create a \code{cubble} object, with the \code{coords} argument being
+omitted. This opens up the possibility to represent fixed area with
+polygons or multipolygons (see Applications @ref(covid)) and the
+\code{coords} argument will be calculated as the centroids of the
+(multi)polygons. The \code{[sf]} print in the cubble header suggest an
+spatial component being also a \code{sf} object:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> (sf_nested <- make_cubble(
++   spatial = stations_sf, temporal = meteo, 
++   key = id, index = date))
+\end{CodeInput}
+\begin{CodeOutput}
+# cubble:   key: id [3], index: date, nested form, [sf]
+# spatial:  [144.8321, -37.98, 145.0964, -37.6655], WGS 84
+# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
+  id           elev name   wmo_id  long   lat            geometry ts      
+  <chr>       <dbl> <chr>   <dbl> <dbl> <dbl>         <POINT [°]> <list>  
+1 ASN00086038  78.4 essen~  95866  145. -37.7 (144.9066 -37.7276) <tibble>
+2 ASN00086077  12.1 moora~  94870  145. -38.0   (145.0964 -37.98) <tibble>
+3 ASN00086282 113.  melbo~  94866  145. -37.7 (144.8321 -37.6655) <tibble>
+\end{CodeOutput}
+\begin{CodeInput}
+R> class(sf_nested)
+\end{CodeInput}
+\begin{CodeOutput}
+[1] "spatial_cubble_df" "cubble_df"         "sf"               
+[4] "tbl_df"            "tbl"               "data.frame"       
+\end{CodeOutput}
+\end{CodeChunk}
+
+The \code{sf} functions applicable to a \code{cubble} object have been
+listed in Table @ref(tab:funs) and the following code shows how to
+perform coordinate transformation with \code{st_transform} on a cubble
+object:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> sf_nested %>% sf::st_transform(crs = "EPSG:3857")
+\end{CodeInput}
+\begin{CodeOutput}
+# cubble:   key: id [3], index: date, nested form, [sf]
+# spatial:  [16122635.6225205, -4576600.8687746, 16152057.3639371,
+#   -4532279.35567565], WGS 84
+# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
+  id           elev name   wmo_id  long   lat            geometry ts      
+  <chr>       <dbl> <chr>   <dbl> <dbl> <dbl>         <POINT [°]> <list>  
+1 ASN00086038  78.4 essen~  95866  145. -37.7 (16130929 -4541016) <tibble>
+2 ASN00086077  12.1 moora~  94870  145. -38.0 (16152057 -4576601) <tibble>
+3 ASN00086282 113.  melbo~  94866  145. -37.7 (16122636 -4532279) <tibble>
+\end{CodeOutput}
+\end{CodeChunk}
+
+The counterpart to promote the spatial component in an existing
+\code{cubble} to be an \code{sf} object is \code{make_spatial_sf()}:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> cb_nested %>% make_spatial_sf() 
+\end{CodeInput}
+\begin{CodeOutput}
+# cubble:   key: id [3], index: date, nested form, [sf]
+# spatial:  [144.8321, -37.98, 145.0964, -37.6655], WGS 84
+# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
+  id           long   lat  elev name   wmo_id ts                  geometry
+  <chr>       <dbl> <dbl> <dbl> <chr>   <dbl> <list>           <POINT [°]>
+1 ASN00086038  145. -37.7  78.4 essen~  95866 <tibble> (144.9066 -37.7276)
+2 ASN00086077  145. -38.0  12.1 moora~  94870 <tibble>   (145.0964 -37.98)
+3 ASN00086282  145. -37.7 113.  melbo~  94866 <tibble> (144.8321 -37.6655)
+\end{CodeOutput}
+\end{CodeChunk}
+
+\hypertarget{tidyverse}{%
+\subsection{Comparison to other spatio-temporal
+classes}\label{tidyverse}}
+
+In \proglang{R}, there are other existing spatio-temporal data structure
+and this section compares and contrasts \pkg{cubble} with other existing
+alternative, specifically \pkg{stars} and \pkg{sftime}. The \pkg{stars}
+package \citep{stars} uses an array structure, as oppose to tibble, to
+represent multivariate spatio-temporal data. While both \pkg{stars} and
+\pkg{cubble} support vector and raster data, it is a matter of choice on
+which structure to use given the application. Analysts working on
+satellite imageries may prefer the array structure in \pkg{stars}, while
+others originally working with spatio-temporal data in 2D data frames
+may find \pkg{cubble} easier to adopt from their existing computing
+workflow.
+
+The \pkg{sftime} package \citep{sftime} also builds from a tibble object
+and its focus is on handling irregular spatio-temporal data. This means
+\pkg{sftime} can also handle full space-time grids and sparse space-time
+layouts represented in \pkg{cubble}, but \pkg{cubble} uses nesting to
+avoid storing spatial variables repetitively at each timestamp. This
+provides memory efficiency when data is observed frequent, i.e.~daily or
+sub-daily, or the spatial geometry is expensive to repeat, i.e.~polygons
+or multipolygons. Consider the \code{climate_aus} data in the
+\pkg{cubble} package with 639 stations observed daily in a single year
+2020. The created \code{sftime} object is approximately 14 times larger
+than the corresponding \code{cubble} object (118.24 MB vs.~8.52 MB).
+
+\hypertarget{others}{%
+\section{Other features and considerations}\label{others}}
+
+\hypertarget{matching}{%
+\subsection{Data fusion and matching}\label{matching}}
+
+One common task when working with spatio-temporal data is to match
+nearby sites. For example, we may want to verify the location of an old
+list of stations with current stations, or we may want to match the data
+from different data sources. In cubble, spatial and temporal matching
+are performed using the functions \texttt{match\_spatial()} and
+\texttt{match\_temporal()}. The \texttt{match\_spatial()} function
+calculates the spatial distance between observations in two cubble
+objects. Various distance measures are available (check
+\texttt{sf::st\_distance}). Analysts can specify the number of matched
+groups to output using the \texttt{spatial\_n\_group} argument (default
+to 4 groups) and the number of matches per group using the
+\texttt{spatial\_n\_group} argument (default to 1, one-to-one matching).
+The syntax to use \texttt{match\_spatial()} is:
+
+\begin{verbatim}
+match_spatial(<cubble_obj1>, <cubble_obj2>, ...)
+\end{verbatim}
+
+The function \texttt{match\_temporal()} calculates the similarity
+between time series within spatially matched groups. Two identifiers are
+required: one for separating each matched group (\texttt{match\_id}) and
+one for separating the two data sources (\texttt{data\_id}). The
+argument \texttt{temporal\_by} uses the \texttt{by} syntax from dplyr's
+\texttt{*\_join} to specify the temporal variables to match.
+
+The similarity score between two time series is calculated using a
+matching function, which can be customised by the analysts. The matching
+function takes two time series as a list and returns a single numerical
+score. This allows for flexibility in using existing time series
+distance calculation implementation. By default, cubble implements a
+simple peak matching algorithm (\texttt{match\_peak}) that counts the
+number of peaks in two time series that fall within a specified temporal
+window. The syntax to use \texttt{match\_temporal()} is
+
+\begin{verbatim}
+match_temporal(
+  <cubble_obj_from_match_spatial>, 
+  data_id = , match_id = , 
+  temporal_by = c("..." = "...")
+)
+\end{verbatim}
 
@@
 
-Cubble fits in naturally with the interactive graphic pipeline discussed
-in the literature
-\citep{buja1988elements, buja1996interactive, sutherland2000orca, xie2014reactive, cheng2016enabling}.
-Diagram \ref{fig:illu-interactive} illustrates how linking works from
-the map to the time series in cubble. The map and time series plot is
-associated with the nested or long cubble, respectively, and when a user
-action is captured on the map, the site will be activated in the nested
-cubble (left). The nested cubble will communicate to the long cubble to
-activate all the observations with the same \code{id} (middle). The long
-cubble will then highlight the activated series in the time series plot
-(right).
-
-The linking is also available from the time series plot to the map. The
-selection(s) on the time series is through selecting the point(s) on the
-time series and once a point is selected, it will be activated in the
-long cubble. All the observations that share the same \code{id} are then
-activated and this includes other points in the same time series in the
-long cubble and the corresponding observation of site in the nested
-cubble. These activated observations will then being reflected in the
-updated plots and Diagram \ref{fig:illu-interactive-2} in the Appendix
-illustrates this process.
+The workflow with the \code{cubble} class works well with an interactive
+graphics pipeline (e.g., \citet{buja1988elements};
+\citet{buja1996interactive}; \citet{sutherland2000orca};
+\citet{xie2014reactive}; \citet{cheng2016enabling}) that is available in
+\proglang{R} with the package \pkg{crosstalk} \citep{crosstalk}. Figure
+\ref{fig:illu-interactive} illustrates the linking between a map and
+multiple time series in a \code{cubble} object. Both the map, generated
+from the spatial cubble, and the time series plot, generated from the
+temporal cubble, are shared crosstalk objects. When a user selects a
+location on the map (a), the corresponding site is highlighted. This
+selection activates a row in the spatial \code{cubble}, which is then
+communicated to the temporal \code{cubble}, resulting in the selection
+of all observations with the same ID in (b). Consequently the temporal
+\code{cubble} highlights the corresponding series in the time series
+plot (c). Linking can also be initiated from the time series plot, by
+selecting points on the time series. This action activates rows with the
+same ID in the temporal \code{cubble} and the row in the spatial
+\code{cubble}, which will be highlighted on the map. Please refer to the
+appendix for the corresponding illustration.
 
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.004}
+{\centering \includegraphics[width=1\linewidth,height=0.35\textheight]{figures/diagram-keynotes/diagram-keynotes.002}
 
@@
 
-\caption[An illustration of the data model under interactive graphics with cubble]{An illustration of the data model under interactive graphics with cubble.
-The line plot and the map is made separately with the long and nested cubble.
-When a station is selected on the map (left), the corresponding row in the nested cubble will be activated.
-This will link to all the rows with the same id in the long cubble (middle) and update the line plot (right).}\label{fig:illu-interactive}
+\caption{Linking between multiple plots.
+The line plots and the map are constructed from shared \code{crosstalk} objects.
+When a station is selected on the map (a), the corresponding row in the spatial \code{cubble} will be activated.
+This will link to all the rows with the same id in the temporal \code{cubble} (b) and update the line plot (c).}(\#fig:illu-interactive)
 \end{figure}
@@
 
-Spatio-temporal transformation is a useful technique to extract
-information form spatio-temporal data. Glyph maps \citep{Wickham2012-yr}
-transform the time coordinates into the space coordinates to plot the
-time series of different locations on the map. Calendar plots
-\citep{wang2020calendar} reconstructs time into calendar-based grid to
-discover weekday and weekend pattern. Projection, or linear combination,
-of variables summarises multivariate information into lower dimension to
-further digest. This section elaborates on the glyph map.
-
-In \proglang{R}, \pkg{GGally} implements glyph maps through the
-\code{glyphs()} function. The function constructs a data frame with
-calculated position (\code{gx}. \code{gy}. \code{gid}) of each point on
-the time series using linear algebra (Equation 1 and 2 in
-\citet{Wickham2012-yr}). The data can then be piped into \code{ggplot}
-to create the glyph map as:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> library(ggplot2)
-R> gly <- glyphs(data, 
-+               x_major = ..., x_minor = ..., 
-+               y_major = ..., y_minor = ..., ...)
+Sometimes, we wish to communicate spatial and temporal information
+collectively through visualisation. This can be achieved through several
+graphical displays: one can make faceted maps across time, creating map
+animations, or constructing interactive graphics to link between map and
+time series plot. The concept of glyph maps was initially proposed in
+\citet{Wickham2012-yr}. The underlying idea is to transform the temporal
+coordinates into spatial coordinates so that time series plot can be
+displayed on the map.
+
+The package \texttt{GGally} initially implement the glyph map. It uses
+\texttt{glyphs()} to calculate the axis transformation and then uses
+\texttt{geom\_polygon()} to draw the map. In cubble, a ggproto
+implementation \texttt{geom\_glyph()} performs the linear algebra
+internally as data transformation . The \texttt{geom\_glyph()} requires
+four aesthetics: \texttt{x\_major}, \texttt{y\_major},
+\texttt{x\_minor}, and \texttt{y\_minor}. The major axes are the outer
+spatial coordinates and the minor axes are the inner/ temporal
+coordinates:
+
+\begin{verbatim}
+data |> 
+  ggplot() +
+  geom_glyph(aes(x_major = ..., x_minor = ..., 
+                 y_major = ..., y_minor = ...))
+\end{verbatim}
+
+Some useful controls over the glyph map includes:
+
+\begin{itemize}
+\tightlist
+\item
+  polar coordinate glyph maps with \code{polar = TRUE},
+\item
+  adjust glyph size with arguments \code{width} and \code{height},
+\item
+  glyph scale to fixed (\code{global_rescale} defaults to \code{TRUE})
+  or free, and
+\item
+  reference boxes and lines with \code{geom_glyph_box()} and
+  \code{geom_glyph_line()}.
+\end{itemize}
+
+\hypertarget{examples}{%
+\section{Applications}\label{examples}}
+
+Five examples are chosen to illustrate different aspects of the
+\pkg{cubble} package: creating a \code{cubble} object from two
+Coronavirus (COVID) data tables with the complication of differing
+location names, using spatial transformations to make a glyph map of
+seasonal temperature changes, matching river level data and weather
+station records for analysis of water supply, reading NetCDF format data
+to reproduce a climate reanalysis plot, and the workflow to create
+complex interactive linked plots.
+
+\hypertarget{covid}{%
+\subsection{Victoria COVID spatio-temporal incidence and
+spread}\label{covid}}
+
+Since the start of the pandemic, the Victoria State Government in
+Australia has been providing daily COVID counts per Logal Government
+Area (LGA). This data can be combined with map polygon data, available
+from the Australian Bureau of Statistics (ABS), to visualize COVID
+incidence and spread. In the \pkg{cubble} package, the COVID count data
+(\code{covid}) and the LGA information (\code{lga}) are available as a
+\code{tsibble} object and an \code{sf} object respectively.
+
+A cubble object can be created from separate spatial and temporal
+component using the \code{make_cubble()} function, introduced in Section
+@ref(create). The \code{by} argument is used to specify the joining
+variable from the two component using the \code{*_join()} \code{by}
+syntax:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> cb <- make_cubble(lga, covid, by = c("lga_name_2018" = "lga"))
+\end{CodeInput}
+\begin{CodeOutput}
+Warning: st_centroid assumes attributes are constant over geometries
+\end{CodeOutput}
+\begin{CodeOutput}
+! Some sites in the spatial table don't have temporal information
+\end{CodeOutput}
+\begin{CodeOutput}
+! Some sites in the temporal table don't have spatial information
+\end{CodeOutput}
+\begin{CodeOutput}
+! Use `check_key()` to check on the unmatched key
+The cubble is created only with sites having both spatial and
+temporal information
+\end{CodeOutput}
+\end{CodeChunk}
+
+The warning message suggests the slight difference of LGA encoding used
+by Victoria government and ABS and prompts analysts to use the function
+\code{check_key()} to identify the mismatches:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> (check_res <- check_key(
++   spatial = lga, temporal = covid, 
++   by = c("lga_name_2018" = "lga")
++ ))
+\end{CodeInput}
+\begin{CodeOutput}
+$paired
+# A tibble: 78 x 2
+  spatial        temporal      
+  <chr>          <chr>         
+1 Alpine (S)     Alpine (S)    
+2 Ararat (RC)    Ararat (RC)   
+3 Ballarat (C)   Ballarat (C)  
+4 Banyule (C)    Banyule (C)   
+5 Bass Coast (S) Bass Coast (S)
+# i 73 more rows
+
+$potential_pairs
+# A tibble: 2 x 2
+  spatial             temporal    
+  <chr>               <chr>       
+1 Kingston (C) (Vic.) Kingston (C)
+2 Latrobe (C) (Vic.)  Latrobe (C)
+
+$others
+$others$spatial
+[1] "No usual address (Vic.)"               
+[2] "Migratory - Offshore - Shipping (Vic.)"
+
+$others$temporal
+[1] "Interstate" "Overseas"   "Unknown"   
+\end{CodeOutput}
+\end{CodeChunk}
+
+The result of the \code{check_key()} function is a list containing three
+elements: 1) matched keys from both tables, 2) potentially pairable
+keys, and 3) others. Here, the main mismatch arises from the two LGAs:
+Kingston and Latrobe (Kingston is a LGA in both Victoria and South
+Australia and Latrobe is a LGA in both Victoria and Tasmania). Analysts
+can modify the input spatial and temporal data accordingly and recreate
+the cubble object:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> lga2 <- lga |>
++   rename(lga = lga_name_2018) |> 
++   mutate(lga = ifelse(lga == "Kingston (C) (Vic.)", "Kingston (C)", lga),
++          lga = ifelse(lga == "Latrobe (C) (Vic.)", "Latrobe (C)", lga)) |>
++   filter(!lga %in% check_res$others$spatial) 
 R>   
-R> ggplot(gly, aes(gx, gy, group = gid)) + 
-+   geom_path() 
-\end{CodeInput}
-\end{CodeChunk}
-
-A re-implementation of the glyph map as a ggproto, \code{GeomGlyph}, has
-been made in the \pkg{cubble} package and now the glyph map can be
-created with \code{geom_glyph()}:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> ggplot(data = data) +
-+   geom_glyph(aes(x_major = ..., x_minor = ..., 
-+                  y_major = ..., y_minor = ...))
-\end{CodeInput}
-\end{CodeChunk}
-
-Some useful controls over the glyph map is also available in the
-\code{geom_glyph()} implementation. Polar glyph map can be specify as a
-parameter, \code{polar = TRUE}. in the \code{geom_glyph()}, along with
-\code{width} and \code{height} in either absolute or relative value.
-Global and local scale can be controlled by the parameter
-\code{global_rescale}. which default to \code{TRUE} for global scaling.
-Reference box and line can be added with separate
-\code{geom_glyph_box()} and \code{geom_glyph_line()}.
-
-\hypertarget{examples}{%
-\section{Examples}\label{examples}}
-
-\hypertarget{australia-historical-maximum-temperature}{%
-\subsection{Australia historical maximum
-temperature}\label{australia-historical-maximum-temperature}}
-
-Global Historical Climatology Network (GHCN) provides daily climate
-measures from stations across the world. The dataset
-\code{weatherdata::historical_tmax} extracts the maximum temperature for
-236 Australian stations from GHCN with starting from year 1969.
-\code{weatherdata::historical_tmax} is already in a cubble, with
-\code{id} as the key, \code{date} as the index, and
-\code{c(longitude, latitude)} as the coordinates. This example compares
-the maximum temperature in two periods: 1971 - 1975 and 2016 - 2020 for
-stations in Victoria and New South Wales.
-
-Stations in the two states can be subsetted by the station number:
-Australia GHCN station number starts with ``ASN00'' and followed by the
-\href{http://www.bom.gov.au/climate/cdo/about/site-num.shtml}{Bureau of
-Meteorology (BOM) station number}, where the 2nd and 3rd digit (7th and
-8th in the GHCN number) define the state of the station. New South Wales
-stations start from 46 to 75 and Victoria stations follow from 76 to 90.
-Filtering Victoria and New South Wales stations is an operation in the
-spatial dimension and hence uses the nested form:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> tmax <- weatherdata::historical_tmax |>
-+   filter(between(stringr::str_sub(id, 7, 8), 46, 90))
-\end{CodeInput}
-\end{CodeChunk}
-
-Filtering for the period 1971 \textasciitilde{} 1975 and 2016
-\textasciitilde{} 2020 is an operation on the time dimension and the
-nested cubble needs to be switched to the long cubble by
-\code{stretch()}:
+R> covid2 <- covid |> filter(!lga %in% check_res$others$temporal)
+R> 
+R> (cb <- make_cubble(spatial = lga2, temporal = covid2))
+\end{CodeInput}
+\begin{CodeOutput}
+# cubble:   key: lga [80], index: date, nested form, [sf]
+# spatial:  [140.961682, -39.1591895, 149.976291, -33.9804256], WGS 84
+# temporal: date [date], n [dbl], avg_7day [dbl]
+  lga             long   lat                                   geometry ts      
+  <chr>          <dbl> <dbl>                         <MULTIPOLYGON [°]> <list>  
+1 Alpine (S)      147. -36.9 (((146.7258 -36.45922, 146.8033 -36.45139~ <tbl_ts>
+2 Ararat (RC)     143. -37.5 (((143.1807 -37.73152, 143.1793 -37.73242~ <tbl_ts>
+3 Ballarat (C)    144. -37.5 (((143.6622 -37.57241, 143.6686 -37.53844~ <tbl_ts>
+4 Banyule (C)     145. -37.7 (((145.1357 -37.74091, 145.1331 -37.74281~ <tbl_ts>
+5 Bass Coast (S)  146. -38.5 (((145.5207 -38.30667, 145.5418 -38.30938~ <tbl_ts>
+# i 75 more rows
+\end{CodeOutput}
+\end{CodeChunk}
+
+\hypertarget{historicaltmax}{%
+\subsection{Australian historical maximum
+temperature}\label{historicaltmax}}
+
+The GHCN provides daily climate measures from stations across the world.
+The data used here (\code{historical_tmax}) is a subset extracted using
+the package \pkg{rnoaa} \citep{rnoaa}, containing the records of maximum
+temperature for 237 Australian stations from \ensuremath{\infty{}}
+through \ensuremath{-\infty{}} and provides information also on the
+latitude, longitude and elevation of each of the stations. The goal of
+this example is to compare the monthly average maximum temperature
+between two periods, 1971-1975 and 2016-2020, for stations in Victoria
+and New South Wales (NSW), using a glyph map.
+
+First, the stations need to be filtered to those in Victoria and NSW by
+using the station identifiers, stored within the 11 digits of the
+\code{id} variable entries. The country code is in the first 5 digits
+(Australia is represented by ``ASN00'') and the next 6 digits encode the
+station following the Australian Bureau of Meteorology (BOM) \citep{bom}
+coding protocols. The NSW stations correspond to entries in the range
+46-75 and the Victorian stations to 76-90. Filtering Victoria and NSW
+stations is a \emph{spatial operation} and hence uses the spatial
+\code{cubble}:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> tmax <- historical_tmax |>
++   filter(between(as.numeric(stringr::str_sub(id, 7, 8)), 46, 90))
+\end{CodeInput}
+\end{CodeChunk}
+
+Next, the monthly maximum average temperature is calculated for both
+periods. This is a \emph{temporal operation} requiring a switch into the
+temporal \code{cubble} using the \code{face_temporal()} function. In
+addition, a new indicator for the two time periods of interest is
+created before the calculation of monthly averages:
 
@@
 +   face_temporal() |> 
-+   filter(lubridate::year(date) %in% c(1971:1975, 2016:2020)) 
-\end{CodeInput}
-\end{CodeChunk}
-
-A monthly average is used for both periods to smooth the maximum
-temperature and it is again an operation on the time dimension:
++   group_by(
++     yearmonth = tsibble::make_yearmonth(
++       year = ifelse(lubridate::year(date) > 2015, 2016, 1971),
++       month = lubridate::month(date))
++       )|>
++   summarise(tmax = mean(tmax, na.rm = TRUE)) %>% 
++   mutate(group = as.factor(lubridate::year(yearmonth)),
++          month = lubridate::month(yearmonth))
+\end{CodeInput}
+\end{CodeChunk}
+
+A quick check on the number of observations for each location is made,
+revealing that there are several with less than 24 observations -- these
+stations lack temperature values for some months. In this example, those
+stations are removed by switching to the spatial \code{cubble} to
+operate on the spatial component over time, and then, move back into the
+temporal \code{cubble} (to make the glyph map):
 
@@
 R> tmax <- tmax |> 
-+   group_by(month = lubridate::month(date), 
-+          group = as.factor(ifelse(lubridate::year(date) > 2015, 
-+                                   "2016 ~ 2020", "1971 ~ 1975"))) |>
-+   summarise(tmax = mean(tmax, na.rm = TRUE))
-\end{CodeInput}
-\end{CodeChunk}
-
-A few stations don't have records during 1971 - 1975 and further
-investigation shows that while the first and last year of each series is
-recorded, the missing years in this period is not reported. These
-stations are filtered out by examining whether the summarised time
-series has a total of 24 months. The long cubble needs to be switched to
-the nested form for this spatial operation using \code{face_spatial()}:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> tmax <- tmax |> face_spatial() |> filter(nrow(ts) == 24) 
-\end{CodeInput}
-\end{CodeChunk}
-
-Lastly, to create a glyph map, both the major (\code{longitude},
-\code{latitude}) and minor (\code{month}, \code{tmax}) coordinates need
-to be in the same table. Spatial variables can be moved to the long form
-with \code{migrate()}:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> tmax <- tmax |> face_temporal() |> unfold(latitude, longitude)
-\end{CodeInput}
-\end{CodeChunk}
-
-\code{tmax} can then be supplied to \code{geom_glyph()} for the glyph
-map in Figure \ref{fig:basic-manip} with a station inset on the top left
-corner:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> nsw_vic <- ozmaps::abs_ste |> 
-+   filter(NAME %in% c("Victoria", "New South Wales"))
-R> 
++   face_spatial() |> 
++   rowwise() %>% 
++   filter(nrow(ts) == 24) |>
++   face_temporal()
+\end{CodeInput}
+\end{CodeChunk}
+
+In order to create a glyph map displaying the monthly series (Figure
+\ref{fig:basic-manip}), the spatial variables need to be unfolded with
+the temporal variables. The reason being that the major (\code{long},
+\code{lat}) and minor (\code{month}, \code{tmax}) coordinates need to be
+on the same table to create the glyph map. The \code{geom_glyph()}
+function does both the transformation and the plotting.
+
+\begin{CodeChunk}
+\begin{CodeInput}
 R> ggplot() +
-+   geom_sf(data = nsw_vic, 
-+           fill = "transparent", color = "grey", linetype = "dotted") + 
++   geom_sf(data = MAP_DATA, ...) +
 +   geom_glyph(data = tmax,
-+              aes(x_major = longitude, x_minor = month, 
-+                  y_major = latitude, y_minor = tmax,
++              aes(x_major = long, x_minor = month,
++                  y_major = lat, y_minor = tmax,
 +                  group = interaction(id, group), color = group),
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.7\textheight]{figures/basic-manip}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/research/paper-cubble/figures/basic-manip-1}
 
@@
 
-\caption[A glyph map of the average maximum temperature by month of Victoria and New South Wales weather stations in Australia]{A glyph map of the average maximum temperature by month of Victoria and New South Wales weather stations in Australia.
-On the top left corner is an insetted plot of station Cobar highlighted in the black box.}\label{fig:basic-manip}
+\caption[A glyph map of the monthly maximum average temperature for weather stations in  Victoria and New South Wales (NSW) for the periods (1971-1975, 2016-2020)]{A glyph map of the monthly maximum average temperature for weather stations in  Victoria and New South Wales (NSW) for the periods (1971-1975, 2016-2020).
+The corresponding average time series for the cobar station are display on the top left corner.
+From the glyph map we can observe that the monthly trend is similar for all locations (low in the winter, high in the summer), and small increased temperatures, particularly in late summer can be seen at most stations in NSW.}(\#fig:basic-manip)
 \end{figure}
@@
 
-\hypertarget{australia-precipitation-pattern-in-2020}{%
-\subsection{Australia precipitation pattern in
-2020}\label{australia-precipitation-pattern-in-2020}}
-
-In the previous example, there has already been some overlapping of the
-glyphs for a few stations near (151E, 34S) and (152E, 33S) and this will
-be a problem when mapping more stations in the national level.
-Aggregation can be helpful to group series into clusters before
-visualising the cluster with glyph map. This example shows how to
-organise data at both level with \code{switch_key()}.
-
-\code{weatherdata::climate_full}, also extracted from GHCN, records
-daily precipitation and maximum/ minimum temperature for 639 stations in
-Australia from 2016 to 2020. A simple kmean algorithm based on the
-distance matrix is used to create 20 clusters. This creates
-\code{station_nested} as a station level nested cubble with a cluster
-column indicating the group each station belongs to. More complex
-algorithms can be used for other problem, as long as there is a mapping
-from each station to a cluster.
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> station_nested <- weatherdata::climate_full |> 
-+   mutate(cluster = ...)
-\end{CodeInput}
-\end{CodeChunk}
-
-To create a group level cubble, use \code{switch_key()} with the new key
-variable, \code{cluster}:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> cluster_nested <- station_nested |> switch_key(cluster) 
-\end{CodeInput}
-\end{CodeChunk}
-
-With the group level cubble, \code{get_centroid()} is useful to compute
-the centroid of each cluster, which will be used as the major axis for
-the glyph map later:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> cluster_nested <- cluster_nested |> get_centroid()
-\end{CodeInput}
-\end{CodeChunk}
-
-Long form cubble at both levels can be accessed through stretching the
-nested form and with access to both station and cluster level cubbles,
-various plots can be made to understand the cluster. Figure
-\ref{fig:basic-agg} shows two example plots that can be made with this
-data: subplot A is a glyph map made with the cluster level cubble in the
-long form and subplot B inspects the station membership of each cluster
-using the station level cubble in the nested form.
+Glyph maps work well to explore temporal patterns across spatial
+locations, particularly when the spatial locations are gridded. In this
+example, they are irregularly spaced, which can result in overlapping
+glyphs obscuring each other.
+
+\hypertarget{river-levels-and-rainfall-in-victoria}{%
+\subsection{River levels and rainfall in
+Victoria}\label{river-levels-and-rainfall-in-victoria}}
+
+One common task when working with spatio-temporal data is to match
+nearby sites. For example, we may want to verify the location of an old
+list of stations with current stations, or we may want to match the data
+from different data sources. In this example, we will introduce the
+spatial and temporal matching in cubble using an example on matching
+river level data with precipitation in Victoria, Australia.
+
+The water level data collected by the Bureau of Meteorology, can be
+compared with the precipitation since rainfall can directly impact water
+level in river. Both \texttt{climate\_vic} and \texttt{river} are cubble
+objects, and we can obtain a summary of the 10 closest pairs between
+them:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> res_sp <- match_spatial(climate_vic, river, spatial_n_group = 10)
+R> print(res_sp, n = 20)
+\end{CodeInput}
+\begin{CodeOutput}
+# A tibble: 10 x 4
+   from        to      dist group
+   <chr>       <chr>    [m] <int>
+ 1 ASN00088051 406213 1838.     1
+ 2 ASN00084145 222201 2185.     2
+ 3 ASN00085072 226027 3282.     3
+ 4 ASN00080015 406704 4034.     4
+ 5 ASN00085298 226027 4207.     5
+ 6 ASN00082042 405234 6153.     6
+ 7 ASN00086038 230200 6167.     7
+ 8 ASN00086282 230200 6928.     8
+ 9 ASN00085279 224217 7431.     9
+10 ASN00080091 406756 7460.    10
+\end{CodeOutput}
+\end{CodeChunk}
+
+The result can also be returned as cubble objects by setting the
+argument \texttt{return\_cubble\ =\ TRUE}. The output is be a list where
+each element is a paired cubble object. To combine all the results into
+a single cubble, you can use \texttt{bind\_rows()}. In the case when a
+site in the second cubble (the \texttt{river} data here) is matched to
+two stations in the first cubble (\texttt{climate\_vic} here), the
+binding may not be successful since cubble requires unique rows in the
+nested form. In the summary table above, the river station
+\texttt{226027} is matched to more than one weather station:
+\texttt{ASN00085072} (group 3) and \texttt{ASN00085298} (group 5).
+Similarly, the river station \texttt{230200} is matched in group 7 and
+8). In such cases, you can either deselect one pair before combining, or
+work with the list output with the \texttt{purrr::map} syntax:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> res_sp <- match_spatial(
++   climate_vic, river, 
++   spatial_n_group = 10, return_cubble = TRUE)
+R> str(res_sp, max.level = 0)
+\end{CodeInput}
+\begin{CodeOutput}
+List of 10
+\end{CodeOutput}
+\begin{CodeInput}
+R> (res_sp <- res_sp[-c(5, 8)] %>% bind_rows())
+\end{CodeInput}
+\begin{CodeOutput}
+# cubble:   key: id [16], index: date, nested form, [sf]
+# spatial:  [144.5203, -38.144913, 148.4667, -36.128657], WGS 84
+# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
+  id      long   lat  elev name  wmo_id ts       type              geometry
+  <chr>  <dbl> <dbl> <dbl> <chr>  <dbl> <list>   <chr>          <POINT [°]>
+1 ASN00~  145. -37.0 290   rede~  94859 <tibble> clim~  (144.5203 -37.0194)
+2 406213  145. -37.0  NA   CAMP~     NA <tibble> river (144.5403 -37.01512)
+3 ASN00~  148. -37.7  62.7 orbo~  95918 <tibble> clim~  (148.4667 -37.6922)
+4 222201  148. -37.7  NA   SNOW~     NA <tibble> river  (148.451 -37.70739)
+5 ASN00~  147. -38.1   4.6 east~  94907 <tibble> clim~  (147.1322 -38.1156)
+# i 11 more rows
+# i 2 more variables: group <int>, dist [m]
+\end{CodeOutput}
+\end{CodeChunk}
+
+For temporal matching, we match the variable
+\texttt{Water\_course\_level} from the river data to \texttt{prcp} in
+the weather station data. The variable \texttt{group} and \texttt{types}
+identify the matching group and the two datasets:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> (res_tm <- res_sp %>% 
++   match_temporal(
++     data_id = type, match_id = group,
++     temporal_by = c("prcp" = "Water_course_level")))
+\end{CodeInput}
+\begin{CodeOutput}
+# A tibble: 8 x 2
+  group match_res
+  <int>     <dbl>
+1     1        30
+2     2         5
+3     3        14
+4     4        20
+5     6        23
+# i 3 more rows
+\end{CodeOutput}
+\end{CodeChunk}
+
+Similarly, the cubble output can be returned using the argument
+\texttt{return\_cubble\ =\ TRUE}. Here we select the four pairs with the
+highest number of matching peaks:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> res_tm <- res_sp %>% 
++   match_temporal(
++     data_id = type, match_id = group,
++     temporal_by = c("prcp" = "Water_course_level"),
++     return_cubble = TRUE)
+R> (res_tm <- res_tm %>% bind_rows() %>% filter(group %in% c(1, 7, 6, 9)))
+\end{CodeInput}
+\begin{CodeOutput}
+# cubble:   key: id [8], index: date, nested form, [sf]
+# spatial:  [144.5203, -37.8817, 147.572223, -36.8472], WGS 84
+# temporal: date [date], matched [dbl]
+  id         long   lat  elev name  wmo_id type              geometry group
+  <chr>     <dbl> <dbl> <dbl> <chr>  <dbl> <chr>          <POINT [°]> <int>
+1 ASN00088~  145. -37.0 290   rede~  94859 clim~  (144.5203 -37.0194)     1
+2 406213     145. -37.0  NA   CAMP~     NA river (144.5403 -37.01512)     1
+3 ASN00082~  146. -36.8 502   stra~  95843 clim~  (145.7308 -36.8472)     6
+4 405234     146. -36.9  NA   SEVE~     NA river (145.6828 -36.88701)     6
+5 ASN00086~  145. -37.7  78.4 esse~  95866 clim~  (144.9066 -37.7276)     7
+# i 3 more rows
+# i 3 more variables: dist [m], ts <list>, match_res <dbl>
+\end{CodeOutput}
+\end{CodeChunk}
+
+Figure \ref{fig:matching} shows four matched pairs on the map (a) and
+standardized data as time series (b) with concurrent increasing.
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{figures/basic-agg}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/research/paper-cubble/figures/matching-1}
 
@@
 
-\caption[Profile of aggregated precipitation from 639 weather stations in Australia]{Profile of aggregated precipitation from 639 weather stations in Australia.
-Subplot A shows the glyph map of the weekly averaged precipitation of each cluster.
-The group number of printed in the middle of y minor axis and can be used as a reference line to read the magnitude.
-Subplot B shows the station membership of each cluster.}\label{fig:basic-agg}
-\end{figure}
-\end{CodeChunk}
-
-\hypertarget{river-level-data-in-victoria-water-gauges}{%
-\subsection{River level data in Victoria water
-gauges}\label{river-level-data-in-victoria-water-gauges}}
-
-Bureau of Meteorology collects
-\href{http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full}{water
-data} from river gauges and this includes variables: electrical
-conductivity, turbidity, water course discharge, water course level, and
-water temperature. In particular, water level will interactive with
-precipitation from the climate data since rainfall will raise the water
-level in the river. Figure \ref{fig:matching-map} gives the location of
-available weather station and water gauges in Victoria.
-
-\begin{CodeChunk}
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{figures/matching-map}
-
-}
-
-\caption[Location of weather stations and river gauges in Victoria, Australia]{Location of weather stations and river gauges in Victoria, Australia.}\label{fig:matching-map}
-\end{figure}
-\end{CodeChunk}
-
-From the map, a few water gauges and weather stations are close to each
-other and the fluctuation of the water level could be matched up with
-precipitation measured by the climate station. As introduced in Section
-3.2, \code{match_sites()} can be used to match one source of data with
-another source in a cubble. Here \code{Water_course_level} in
-\code{river} will be matched to \code{prcp} in \code{climate} in 2020.
-The two datasets need to be specified as the first two arguments and the
-variable to match can be specified in \code{temporal_by} using the
-\code{by} syntax in \code{join}. \code{temporal_independent} controls
-the variable used to construct the interval and the goal here is to see
-if precipitation will be reflected by the water level in the river. This
-puts precipitation \code{prcp}, as the independent. Given there is one
-year worth of data, the number of peak (\code{temporal_n_highest}) to
-consider is slightly raised from a default 20 to 30 and
-\code{temporal_min_match} is raised accordingly To return all the pairs
-of the match, \code{temporal_min_match} can be set to 0.
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> res <- match_sites(
-+   river, climate,
-+   temporal_by = c("Water_course_level" = "prcp"),
-+   temporal_independent = "prcp",  
-+   temporal_n_highest = 30,
-+   temporal_min_match = 15
-+ )
-\end{CodeInput}
-\end{CodeChunk}
-
-The output from matching is also a cubble, with additional column
-\code{dist} and \code{group} produced from spatial matching and
-\code{n_match} from temporal matching.
-
-\begin{CodeChunk}
-\begin{CodeOutput}
-# cubble:   id [8]: nested form
-# bbox:     [144.52, -37.73, 146.06, -36.55]
-# temporal: date [date], matched_var [dbl]
-  id          name                  lat  long type   dist group ts       n_match
-  <chr>       <chr>               <dbl> <dbl> <chr> <dbl> <int> <list>     <int>
-1 405234      SEVEN CREEKS @ D/S~ -36.9  146. river  6.15     5 <tibble>      21
-2 404207      HOLLAND CREEK @ KE~ -36.6  146. river  8.54    10 <tibble>      21
-3 ASN00082042 strathbogie         -36.8  146. clim~  6.15     5 <tibble>      21
-4 ASN00082170 benalla airport     -36.6  146. clim~  8.54    10 <tibble>      21
-5 230200      MARIBYRNONG RIVER ~ -37.7  145. river  6.17     6 <tibble>      19
-6 ASN00086038 essendon airport    -37.7  145. clim~  6.17     6 <tibble>      19
-7 406213      CAMPASPE RIVER @ R~ -37.0  145. river  1.84     1 <tibble>      18
-8 ASN00088051 redesdale           -37.0  145. clim~  1.84     1 <tibble>      18
-\end{CodeOutput}
-\end{CodeChunk}
-
-Figure \ref{fig:matching} plots the matched pairs on the map or to view
-the matched series. There are four pairs of matches, which all locates
-in the middle Victoria and the concurrent increase of precipitation and
-water level can be observed.
-
-\begin{CodeChunk}
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{figures/matching}
-
-}
-
-\caption[Matched weather stations and river gauges on the map (A) and across time (B)]{Matched weather stations and river gauges on the map (A) and across time (B).
+\caption[Weather stations and river gauges with matched pairs labelled on the map (a) and plotted across time (b)]{Weather stations and river gauges with matched pairs labelled on the map (a) and plotted across time (b).
 Precipitation and water level have been standardised between 0 and 1 to be displayed on the same scale.
-The increases in the precipitation is reflected by the water level.}\label{fig:matching}
+The water level reflects the increase in precipitation.
+The numbers (1, 7, 6, 9) indicate the group index derived from spatial matching, only those that were selectd by temporal matching are shown here.}(\#fig:matching)
 \end{figure}
@@
 
-ERA5 data \citep{hersbach2020era5} is the latest reanalysis of global
-atmosphere, land surface, and ocean waves from 1950 onwards and is
-available in the NetCDF format from European Centre for Medium-Range
-Weather Forecasts (ECMWF). The data can be directly downloaded from
-\href{https://cds.climate.copernicus.eu/cdsapp\#!/dataset/reanalysis-era5-pressure-levels?tab=overview}{Copernicus
-Climate Data Store (CDS)} website or programmatically via an R package
-\pkg{ecmwfr} \citep{ecwmfr}. The \code{era5-pressure} data contains
-variable \emph{specific humidity} and \emph{geopotential} on the 10 hPa
-pressure level on four dates: 2002-09-22, 2002-09-26, 2002-09-30, and
-2002-10-04. Once downloaded, the data can be read into a cubble as:
+The ERA5 reanalysis \citep{hersbach2020era5} provides hourly estimates
+of atmospheric, land and oceanic climate variables on a global scale.
+The hourly pressure level data in the NetCDF format can be downloaded
+from Copernicus Climate Data Store (CDS) \citep{cds} or via the
+\pkg{ecmwfr} package \citep{ecwmfr}. This example reproduces the row
+created from ERA5 reanalysis data shown in Figure 19 by
+\citet{hersbach2020era5}. The plot shows the southern polar vortex
+splitting into two on 2002-09-26, and further splitting into four on
+2002-10-04. Further explanation of why this is interesting can be found
+in the figure source, and also in \citet{simmons2020global} and
+\citet{simmons2005ecmwf}.
+
+To work with NetCDF data in R, several packages are available, including
+\code{ncdf4}, \code{RNetCDF}, and \code{tidync}. The following code
+converts a NetCDF object of class \code{ncdf4} into a cubble object:
 
@@
 R> raw <- ncdf4::nc_open(here::here("data/era5-pressure.nc"))
-R> dt <- as_cubble(raw, vars = c("q", "z"))
-\end{CodeInput}
-\end{CodeChunk}
-
-Figure \ref{fig:netcdf} reproduces the ERA5 data row of Figure 19 in
-\citet{hersbach2020era5}. It shows the southern polar vortex splits into
-two on 2002-09-26 and further splits into four on 2002-10-04 in the
-stratosphere. Readers interested in the analysis of this figure can
-refer to \citet{hersbach2020era5}, \citet{simmons2020global} and
-\citet{simmons2005ecmwf} for more details.
+\end{CodeInput}
+\end{CodeChunk}
+
+Analysts can extract a subset of the NetCDF data with the arguments
+\code{vars}, \code{long_range} and \code{lat_range}. In this example,
+the variables q (specific humidity) and z (geopotential) are selected
+and the coordinates are subsetted to every degree in longitude and
+latitude:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> (dt <- as_cubble(raw, vars = c("q", "z"),
++                  long_range = seq(-180, 180, 1), lat_range = seq(-88, -15, 1)))
+\end{CodeInput}
+\begin{CodeOutput}
+# cubble:   key: id [26640], index: time, nested form
+# spatial:  [-180, -88, 179, -15], Missing CRS!
+# temporal: time [date], q [dbl], z [dbl]
+     id  long   lat ts              
+  <int> <dbl> <dbl> <list>          
+1     1  -180   -15 <tibble [8 x 3]>
+2     2  -179   -15 <tibble [8 x 3]>
+3     3  -178   -15 <tibble [8 x 3]>
+4     4  -177   -15 <tibble [8 x 3]>
+5     5  -176   -15 <tibble [8 x 3]>
+# i 26,635 more rows
+\end{CodeOutput}
+\end{CodeChunk}
+
+Once the NetCDF data is coerced into a cubble object, subsequent
+analysis can be conducted to filter on the date of interest, scale the
+variable specific humidity and create visualisation in ggplot as in
+Figure \ref{fig:netcdf}.
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/netcdf}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/research/paper-cubble/figures/netcdf-1}
 
@@
 
-\caption[A reproduction of the second row (ERA5 data) of Figure 19 in  Hersbach et al (2020)]{A reproduction of the second row (ERA5 data) of Figure 19 in  Hersbach et al (2020).}\label{fig:netcdf}
+\caption[A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020) to illustrate the break-up of sourthern polar vortex in late September and early October 2002]{A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020) to illustrate the break-up of sourthern polar vortex in late September and early October 2002.
+The polar vortex, signalled by the high specific humidity, splits into two on 2002-09-26 and further splits into four on 2002-10-04.}(\#fig:netcdf)
 \end{figure}
@@
 
-\hypertarget{interative-graphic}{%
-\subsection{Interative graphic}\label{interative-graphic}}
-
-With spatio-temporal data, users may wish to make plots to learn the
-spatial distribution of a variable, or to find patterns, such as trend
-or seasonality, in the time series. Combining this two types of plot
-with interactivity let users to link between points on the map and the
-corresponding time series to explore the spatial and temporal dimension
-of the data simultaneously. Below is an example that describes the
-process of building an interactive graphic with \pkg{cubble} and
-\pkg{crosstalk} The example explores the variation of monthly
-temperature range with \code{weatherdata::climate_full} data.
-
-The temperature range is calculated as the difference between
-\code{tmax} and \code{tmin} and its monthly average over 2016 - 2020 is
-taken before calculating the variance. A \code{SharedData} object is
-constructed for each form of the cubble and the same \code{group}
-argument ensures the cross-linking of the two forms via the common
-\code{id} column. The spatial map and time series plot are then made
-with each \code{SharedData} objects separately. In this example,
-stations on the Australia map, made from the nested form, are coloured
-by the calculated variance and a ribbon band is constructed using the
-long form cubble to show the maximum and minimum temperature of each
-station across month. With a different dataset, users are free to
-calculate any per station measure in the nested form or to make any
-time-wise summary of the data in the long form to customise the spatial
-or temporal view. The cross-linking between the two plots is always
-safeguarded by the shared \code{id} column embedded in the cubble
-structure. Below is the pseudo code that outlines the process to
-construct an interactive graphic described above:
+\hypertarget{australian-temperature-range}{%
+\subsection{Australian temperature
+range}\label{australian-temperature-range}}
+
+Interactive graphics can be useful because they make it possible to look
+at the data in multiple ways on-the-fly. This is especially important
+for spatio-temporal data, where we would like to interactively connect
+spatial and temporal displays. This example describes the process of
+using the \pkg{cubble} package with the \pkg{crosstalk} package to build
+an interactive display connecting a map of Australia, with ribbon plots
+of temperature range observed at the stations. The purpose is to explore
+the variation of monthly temperature range over the country. Figure
+\ref{fig:interactive-linking} shows three snapshots of the
+interactivity.
+
+The key steps are to convert both the nested and long forms of the data
+into shared \code{crosstalk} objects, and to plot these side-by-side.
+The two are linked by the station identifier.
 
@@
 \begin{Highlighting}[]
-\CommentTok{\# data pre{-}processing}
-\NormalTok{clean }\OtherTok{\textless{}{-}}\NormalTok{ weatherdata}\SpecialCharTok{::}\NormalTok{climate\_full }\SpecialCharTok{|\textgreater{}}\NormalTok{ ...}
-
-\CommentTok{\# created SharedData instance for crosstalk}
+\NormalTok{clean }\OtherTok{\textless{}{-}}\NormalTok{ climate\_full }\SpecialCharTok{|\textgreater{}}\NormalTok{ ...}
+
 \NormalTok{nested }\OtherTok{\textless{}{-}}\NormalTok{ clean }\SpecialCharTok{|\textgreater{}}\NormalTok{ SharedData}\SpecialCharTok{$}\FunctionTok{new}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{id, }\AttributeTok{group =} \StringTok{"cubble"}\NormalTok{)}
@@
 
-\CommentTok{\# create the spatial and temporal view each with a ShareData instance}
 \NormalTok{p1 }\OtherTok{\textless{}{-}}\NormalTok{ nested }\SpecialCharTok{|\textgreater{}}\NormalTok{ ...}
@@
 
-\CommentTok{\# Combine p1 and p2}
 \NormalTok{crosstalk}\SpecialCharTok{::}\FunctionTok{bscols}\NormalTok{(plotly}\SpecialCharTok{::}\FunctionTok{ggplotly}\NormalTok{(p1), plotly}\SpecialCharTok{::}\FunctionTok{ggplotly}\NormalTok{(p2), ...)}
@@
 
-In Figure \ref{fig:interactive-linking}, the first row shows the initial
-view of the interactive graphic. On the map, most regions in Australia
-have low variance of temperature range while the north-west coastline,
-bottom of South Australia, and Victoria stands out with larger monthly
-changes. In the second row, Mount Elizabeth is selected on the map given
-its high variance colour on the initial map and this links to the ribbon
-on the right. The third row the lowest temperature in August and this
-corresponds to Thredbo AWS in the Victoria and New South Wales border.
-Another station in the Tasmania island is selected on the map to cross
-compare with Thredbo AWS.
+Plot (a) shows the initial state of the interactive display: all
+locations are shown as dots on the map, coloured by temperature range,
+and the right plot shows the ribbons representing maximum to minimum for
+all stations. In plot (b) the ``Mount Elizabeth'' station, which shows a
+high variance colour on the initial map, is selected on the map and this
+produces the ribbon on the right. In plot (c) the lowest temperature in
+August is selected, which is ``Thredbo'' station on the left map. It was
+surprising to us that this was not a station in Tasmania, so for
+comparison a station in Tasmania is selected on the map to show in
+relation to Thredbo. We can see that Thredbo has a bigger winter dip in
+temperature, and although Tasmania is cold generally, it's temperatures
+are more constant
 
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/linking} \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/linking-north} \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/linking-lower}
+{\centering \includegraphics[width=1\linewidth,height=0.23\textheight]{figures/linking} \includegraphics[width=1\linewidth,height=0.23\textheight]{figures/linking-north} \includegraphics[width=1\linewidth,height=0.23\textheight]{figures/linking-lower}
 
@@
 The top row shows all locations and all temperature profiles.
-Selecting a location with high variance on the map produces the plot in the second row.
-The maximum nad minimum temperature is shown using a ribbon.
-The bottom row first selects the lowest temperature in Auguest in the seasonal display.
-A location in the Tasmania Island is then selected to compare the temperature variation with Thredbo AWS.}\label{fig:interactive-linking}
+Selecting a particular location on the map (here Mount Elizabeth) produces the plot in the second row.
+The maximum and minimum temperatures are shown using a ribbon.
+The bottom row first selects the lowest temperature in August in the seasonal display, which highlights the corresponding station on the map (Thredbo).
+Another  station, located in the Tasmania Island, is then selected to compare its temperature variation with the Thredbo station.}(\#fig:interactive-linking)
 \end{figure}
@@
 
-This plot can also be made using \pkg{cubble} and \pkg{leaflet} where
-the temperature range can be displayed as a small subplot upon clicking
-on the map. This would require first creating the popup plots from the
-long form cubble as a vector and then add these plots to a leaflet map
-created from the nested cubble, with \code{leafpop::addPopupGraphs()}:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# data pre{-}processing}
-\NormalTok{clean }\OtherTok{\textless{}{-}}\NormalTok{ weatherdata}\SpecialCharTok{::}\NormalTok{climate\_full }\SpecialCharTok{|\textgreater{}}\NormalTok{ ...}
-
-\CommentTok{\# use the long form to create subplots for each station}
-\NormalTok{df\_id }\OtherTok{\textless{}{-}} \FunctionTok{unique}\NormalTok{(clean}\SpecialCharTok{$}\NormalTok{id)}
-\NormalTok{p }\OtherTok{\textless{}{-}} \FunctionTok{map}\NormalTok{(}\DecValTok{1}\SpecialCharTok{:}\FunctionTok{length}\NormalTok{(df\_id), }\ControlFlowTok{function}\NormalTok{(i)\{}
-\NormalTok{  dt }\OtherTok{\textless{}{-}}\NormalTok{ clean }\SpecialCharTok{|\textgreater{}} \FunctionTok{filter}\NormalTok{(id }\SpecialCharTok{==}\NormalTok{ df\_id[i])}
-  \FunctionTok{ggplot}\NormalTok{(dt) }\SpecialCharTok{|\textgreater{}}\NormalTok{ ...}
-\NormalTok{\})}
-
-\CommentTok{\# create nested form leaflet map with temperature band as subplots}
-\NormalTok{nested }\OtherTok{\textless{}{-}} \FunctionTok{face\_spatial}\NormalTok{(clean)}
-\FunctionTok{leaflet}\NormalTok{(nested) }\SpecialCharTok{|\textgreater{}}
-  \FunctionTok{addTiles}\NormalTok{() }\SpecialCharTok{|\textgreater{}}
-  \FunctionTok{addCircleMarkers}\NormalTok{(}\AttributeTok{group =} \StringTok{"a"}\NormalTok{, ...) }\SpecialCharTok{|\textgreater{}}
-\NormalTok{  leafpop}\SpecialCharTok{::}\FunctionTok{addPopupGraphs}\NormalTok{(}\AttributeTok{graph =}\NormalTok{ p, ...)}
-\end{Highlighting}
-\end{Shaded}
-
-Figure \ref{fig:interactive-popup} shows Figure
-\ref{fig:interactive-linking} made made with leaflet and popups
-\citep{leafpop}.
-
-\begin{CodeChunk}
-\begin{figure}
-
-{\centering \includegraphics[width=0.45\linewidth,height=0.25\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/popup-mid}
-
-}
-
-\caption[Same as Figure 11 with thetemperature variation shown as a popup in the leaflet map]{Same as Figure 11 with thetemperature variation shown as a popup in the leaflet map.}\label{fig:interactive-popup}
-\end{figure}
-\end{CodeChunk}
-
-\hypertarget{conclusion}{%
-\section{Conclusion}\label{conclusion}}
-
-This paper describes an \proglang{R} package \pkg{cubble} for
-manipulating and visualising spatio-temporal data. A new data structure,
-\code{cubble} that builds from the \code{rowwise_df} and
-\code{grouped_df} class in the tidyverse ecosystem, is proposed to
-connect the time invariant and varying variables in the spatio-temporal
-data. This design frees the data analysts from spending time on
-organising variables of different observational units. The data
-structure is also flexible to the techniques and packages analysts use
-to analyse the data, for example, in the matching example in section
-4.3, users are free to use algorithms from another package to cluster
-stations.
-
-Further development and maintenance of the package involves responding
-to changes in the tidyverse packages that \pkg{cubble} imports, in
-particular, \pkg{tibble}, \pkg{tidyr}, and \pkg{dplyr}. Another area for
-further development is to extend \pkg{cubble} to other spatial objects
-other than points.
-
-\newpage
+\hypertarget{conclude}{%
+\section{Conclusion}\label{conclude}}
+
+This paper presents an \proglang{R} package \pkg{cubble} for organizing,
+manipulating and visualizing spatio-temporal data. The package
+introduces a new data class for spatio-temporal data, \code{cubble},
+that connects the time invariant and varying variables and that allows
+the user to work with a nested and long form of the data. This work adds
+capabilities into the spatio-temporal practitioners toolbox to integrate
+it with a tidy data framework. The data structure and functions
+introduced in this package can be used and combined with existing
+spatial data analysis packages such as \pkg{sf}, data wrangling packages
+such as \pkg{dplyr}, and visualization packages such as \pkg{ggplot2},
+\pkg{plotly} and \pkg{leaflet}.
+
+Numerous examples are provided in the main text, appendix and package
+vignettes. These include creating and coercing data with mismatched
+names, handling hierarchical data, matching time series spatially and
+temporally, reproducing ERA5 plots from NetCDF data. Visualization of
+the \code{cubble} objects can be done with interactive graphic pipelines
+using \pkg{crosstalk}, \pkg{plotly} and \pkg{leaflet}.
+
+There are several possible future directions for the work. The data
+structure only described fixed spatial sites, and it could be useful to
+provide tools to accommodate moving coordinates as might be encountered
+in animal movement data. That could be achieved with a list-column for
+the location coordinates, and an additional form that these locations
+can be pivoted into, like the long form for temporal variables. For
+multiple measured variables, the \pkg{cubble} package could be
+integrated with dimension reduction methods, and dynamic graphics for
+multiple dimensions such as a tour \citep{wickham_tourr_2011}.
 
@@
 
-This work is funded by the Commonwealth Scientific and Industrial
-Research Organisation (CSIRO) Data61 Scholarship and started while
-Nicolas Langrené was affiliated with CSIRO's Data61. The article is
-created using \pkg{knitr} \citep{knitr} and \pkg{rmarkdown}
-\citep{rmarkdown} in R. The source code for reproducing this paper can
-be found at: \url{https://github.com/huizezhang-sherry/paper-cubble}.
-
-\hypertarget{appendix}{%
-\section{Appendix}\label{appendix}}
-
-\begin{CodeChunk}
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.005}
-
-}
-
-\caption[An illustration of the data model under interactive graphics with cubble]{An illustration of the data model under interactive graphics with cubble.
-When a point on the time series is selected, the corresponding row in the long cubble will be activated.
-This will link to all the rows with the same id in the long cubble and the row in the nested cubble with the smae id (middle).
-Both plots will be updated with the full line selected and the point highlighted on the map (right).}\label{fig:illu-interactive-2}
-\end{figure}
-\end{CodeChunk}
-
-\newpage
+This work is funded by a Commonwealth Scientific and Industrial Research
+Organisation (CSIRO) Data61 Scholarship and started while Nicolas
+Langrené was affiliated with CSIRO's Data61. The article is created
+using the package \pkg{knitr} \citep{knitr} and \pkg{rmarkdown}
+\citep{rmarkdown} in \proglang{R} with the \code{rticles::jss_article}
+template. The source code for reproducing this paper can be found at:
+\url{https://github.com/huizezhang-sherry/paper-cubble}.
 

### 2023-06-21 main.Rmd
update conclusion

@@
   - name: Ursula Laa
-    affiliation: 'Universität für Bodenkultur Wiens \AND'
+    affiliation: 'University of  Natural `\\`{=latex} Resources and Life Sciences \AND'
     address: Gregor-Mendel-Straße 33, 1180 Wien, Austria
@@
   - name: Nicolas Langrené
-    affiliation: BNU-HKBU United International College
+    affiliation: 'BNU-HKBU `\\`{=latex} United International College'
     address: 2000 Jintong Road, Tangjiawan, Zhuhai, Guangdong Province, China
@@
   \usepackage{array}
-output: rticles::jss_article
+output: 
+  bookdown::pdf_book:
+    base_format: rticles::jss_article
 bibliography: references.bib
 ---
-
+<!-- Universität für Bodenkultur Wiens \AND -->
 ```{r setup, echo = FALSE}
@@
 
-This paper presents an \proglang{R} package \pkg{cubble} for organizing, manipulating and visualizing spatio-temporal data.
-The package introduces a new data class for spatio-temporal data, \code{cubble}, that connects the time invariant and varying variables and that allows the user to work with a nested and long form of the data.
-This work adds capabilities into the spatio-temporal practitioners toolbox to integrate it with a tidy data framework.
-The data structure and functions introduced in this package can be used and combined with existing spatial data analysis packages such as \pkg{sf}, data wrangling packages such as \pkg{dplyr}, and visualization packages such as \pkg{ggplot2}, \pkg{plotly} and \pkg{leaflet}.
+This paper presents the \proglang{R} package \pkg{cubble} for organizing, wrangling and visualizing spatio-temporal data.
+The package introduces a new data class, \code{cubble}, consisting of two subclass, spatial cubble and a temporal cubble, to organise spatio-temporal data in two different formats within the tidy data framework.
+The data structure and functions introduced in this package can be used and combined with existing spatial and temporal data analysis packages such as \pkg{sf} and \pkg{tsibble}, data wrangling packages such as \pkg{dplyr}, and visualization packages such as \pkg{ggplot2}, \pkg{plotly}, and \pkg{leaflet}.
 
-Numerous examples are provided in the main text, appendix and package vignettes.
-These include creating and coercing data with mismatched names, handling hierarchical data, matching time series spatially and temporally, reproducing ERA5 plots from NetCDF data.
-Visualization of the \code{cubble} objects can be done with interactive graphic pipelines using \pkg{crosstalk}, \pkg{plotly} and \pkg{leaflet}.
-
-There are several possible future directions for the work.
-The data structure only described fixed spatial sites, and it could be useful to provide tools to accommodate moving coordinates as might be encountered in animal movement data.
-That could be achieved with a list-column for the location coordinates, and an additional form that these locations can be pivoted into, like the long form for temporal variables.
-For multiple measured variables, the \pkg{cubble} package could be integrated with dimension reduction methods, and dynamic graphics for multiple dimensions such as a tour [@wickham_tourr_2011].
+The paper includes numerous examples to illustrate the utility of \pkg{cubble} as a data structure for spatio-temporal analysis.
+These examples cover different tasks of a typical analysis workflow: handling data with spatial and temporal misalignment, matching data from multiple sources, and creating both static and interactive spatio-temporal visualisation.
+For future directions, other commonly-used spatial or temporal data structures can be integrated into \pkg{cubble} to extend analysts' familiar spatial and temporal toolkit to spatio-temporal.
 

### 2023-06-21 main.tex
update conclusion

@@
 \author{
-H. Sherry Zhang\\Monash University \And Dianne Cook\\Monash
-University \And Ursula Laa\\Universität für Bodenkultur Wiens
-\AND Nicolas Langrené\\BNU-HKBU United International
-College \And Patricia Menéndez\\Monash University
+H.
+Sherry Zhang\\Monash University \And Dianne Cook\\Monash University \And Ursula Laa\\University of Natural \\ Resources and Life Sciences \AND Nicolas Langrené\\BNU-HKBU \\ United International College \And Patricia Menéndez\\Monash University
 }
-\title{\pkg{cubble}: An \proglang{R} Package for Organizing and
-Wrangling Multivariate Spatio-temporal Data}
-
-\Plainauthor{H. Sherry Zhang, Dianne Cook, Ursula Laa, Nicolas
-Langrené, Patricia Menéndez}
-\Plaintitle{cubble: An R Package for Organizing and Wrangling
-Multivariate Spatio-temporal Data}
+\title{\pkg{cubble}: An \proglang{R} Package for Organizing and Wrangling Multivariate Spatio-temporal Data}
+
+\Plainauthor{H. Sherry Zhang, Dianne Cook, Ursula Laa, Nicolas Langrené, Patricia Menéndez}
+\Plaintitle{cubble: An R Package for Organizing and Wrangling Multivariate Spatio-temporal Data}
 
 \Abstract{
-Multivariate spatio-temporal data refers to multiple measurements taken
-across space and time. For many analyses, spatial and time components
-can be separately studied: for example, to explore the temporal trend of
-one variable for a single spatial location, or to model the spatial
-distribution of one variable at a given time. However for some studies,
-it is important to analyze different aspects of the spatio-temporal data
-simultaneously, like for instance, temporal trends of multiple variables
-across locations. In order to facilitate the study of different portions
-or combinations of spatio-temporal data, we introduce a new class,
-\code{cubble}, with a suite of functions enabling easy slicing and
-dicing on different spatio-temporal components. The proposed
-\code{cubble} class ensures that all the components of the data are easy
-to access and manipulate while providing flexibility for data analysis.
-In addition, the \pkg{cubble} package facilitates visual and numerical
-explorations of the data while easing data wrangling and modelling. The
-\code{cubble} class and the functions provided in the \pkg{cubble}
-\proglang{R} package equip users with the capability to handle
-hierarchical spatial and temporal structures. The \code{cubble} class
-and the tools implemented in the package are illustrated with different
-examples of Australian climate data.
+Multivariate spatio-temporal data refers to multiple measurements taken across space and time.
+For many analyses, spatial and time components can be separately studied: for example, to explore the temporal trend of one variable for a single spatial location, or to model the spatial distribution of one variable at a given time.
+However for some studies, it is important to analyze different aspects of the spatio-temporal data simultaneously, like for instance, temporal trends of multiple variables across locations.
+In order to facilitate the study of different portions or combinations of spatio-temporal data, we introduce a new class, \code{cubble}, with a suite of functions enabling easy slicing and dicing on different spatio-temporal components.
+The proposed \code{cubble} class ensures that all the components of the data are easy to access and manipulate while providing flexibility for data analysis.
+In addition, the \pkg{cubble} package facilitates visual and numerical explorations of the data while easing data wrangling and modelling.
+The \code{cubble} class and the functions provided in the \pkg{cubble} \proglang{R} package equip users with the capability to handle hierarchical spatial and temporal structures.
+The \code{cubble} class and the tools implemented in the package are illustrated with different examples of Australian climate data.
 }
 
-\Keywords{spatial, temporal, spatio
-temporal, \proglang{R}, environmental data, exploratory data analysis}
-\Plainkeywords{spatial, temporal, spatio temporal, R, environmental
-data, exploratory data analysis}
+\Keywords{spatial, temporal, spatio temporal, \proglang{R}, environmental data, exploratory data analysis}
+\Plainkeywords{spatial, temporal, spatio temporal, R, environmental data, exploratory data analysis}
 
@@
       Ursula Laa\\
-    Universität für Bodenkultur Wiens\\
+    University of Natural \\ Resources and Life Sciences\\
     Gregor-Mendel-Straße 33, 1180 Wien, Austria\\
@@
       Nicolas Langrené\\
-    BNU-HKBU United International College\\
+    BNU-HKBU \\ United International College\\
     2000 Jintong Road, Tangjiawan, Zhuhai, Guangdong Province, China\\
@@
 
-Spatio-temporal data has a spatial component referring to the location
-of each observation and a temporal component that is recorded at regular
-or irregular time intervals. It may also include multiple variables
-measured at each spatial and temporal values. With spatio-temporal data,
-one can fix the time to explore the spatial features of the data, fix
-the spatial location/s to explore temporal aspects, or dynamically
-explore the space and time simultaneously.
-
-In order to computationally explore the spatial, temporal and
-spatio-temporal faces of such data, the data needs to be stored and
-represented under a specific data object that allows the user to query,
-group and dissect all the data faces.
-
-The Comprehensive \proglang{R} Archive Network (CRAN) task view
-SpatioTemporal \citep{ctvspatiotemporal} gathers information about
-\proglang{R} packages designed for spatio-temporal data and it has a
-section on \emph{Representing data} that lists existing spatio-temporal
-data representations used in \proglang{R}. Among them, \citet{spacetime}
-summarises spatio-temporal data into three forms: time-wide, space-wide,
-and long formats. The associated package \pkg{spacetime}
-\citep{spacetime} implements four spatio-temporal layouts (full grid,
-sparse grid, irregular, and trajectory) to handle different space and
-time combinations. The package \pkg{stars} \citep{stars} has a new
-implementation to use dense arrays to represent spatio-temporal cubes.
-It also interfaces with the package \pkg{sf} \citep{sf}, commonly used
-for wrangling spatial data, and the \pkg{tidyverse} \citep{tidyverse}
-suite for general data wrangling and visualization in \proglang{R}.
-
-Still, the data representation for spatio-temporal data can be further
-extended and there are two reasons for this. Firstly, the raw data
-sourced in the wild is less often presented in any one of the layouts
-above, and fitting the raw data into a data object can sometimes be
-difficult. More often, spatio-temporal data are collected in separate 2D
-tables and analysts need to assemble them into a whole piece before
-exploring the data. Examples of components of spatio-temporal data can
-be 1) areal data recording the shape of a collection of areas of
-interest; 2) geostatistical data storing the longitude and latitude
-coordinates of locations, typically also with other metadata related to
-the location, and; 3) temporal data of each location across time.
-
-The other reason is about tidy data concepts \citep{tidydata} and how
-they should be applied to spatio-temporal data. According to the tidy
-data principles, data should be structured into 1) one row per
-observation, 2) one column per variable, and 3) one type of data per
-table. The long form data is preferred over wide data form given the
-downstream packages such as \pkg{dplyr} \citep{dplyr} and \pkg{ggplot2}
-\citep{ggplot2} for data wrangling and visualization. However, the long
-form can be inefficient to store feature geometries, especially for
-large multipolygons for hourly, daily or sub-daily periods over years,
-which are extensively collected and handled, for example in time series
-analysis. This poses the question of how to arrange spatial and temporal
-variables in a way that would make data wrangling, visualizing and
-analyzing spatio-temporal data easier.
-
-This paper presents a new \proglang{R} package, \pkg{cubble}, which
-addresses the two issues mentioned above. In the package, a new class,
-also called cubble, is proposed to organize spatial and temporal
-variables as two forms of a single data object so that they can be
-wrangled separately or combined while being kept synchronized. Among the
-four spacetime layouts in \citet{spacetime}, the \code{cubble} class can
-be applied to full grid, sparse grid, or irregular, but not trajectory,
-which is outside the scope of this work. The software is available from
-the Comprehensive \proglang{R} Archive Network (CRAN) at
-\url{https://CRAN.R-project.org/package=cubble}.
-
-The rest of the paper is organized as follows: Section @ref(cubble)
-presents the main design and functionality of the \pkg{cubble} package.
-Section @ref(others) explains how the \pkg{cubble} package deals with
-more advanced considerations, including data with hierarchical
-structure, data matching and how the package fits with existing static
-and interactive visualization tools. Moreover we also illustrate how the
-\pkg{cubble} package deals with spatio-temporal data transformations.
-Section @ref(examples) uses Australian weather station data and river
-level data as examples to demonstrate the use of the package. An example
-of how the \pkg{cubble} package handles Network Common Data Form
-(NetCDF) data is also provided. Section @ref(conclude) discuss the paper
-contributions and future directions.
+Spatio-temporal data has a spatial component referring to the location of each observation and a temporal component that is recorded at regular or irregular time intervals.
+It may also include multiple variables measured at each spatial and temporal values.
+With spatio-temporal data, one can fix the time to explore the spatial features of the data, fix the spatial location/s to explore temporal aspects, or dynamically explore the space and time simultaneously.
+
+In order to computationally explore the spatial, temporal and spatio-temporal faces of such data, the data needs to be stored and represented under a specific data object that allows the user to query, group and dissect all the data faces.
+
+The Comprehensive \proglang{R} Archive Network (CRAN) task view SpatioTemporal \citep{ctvspatiotemporal} gathers information about \proglang{R} packages designed for spatio-temporal data and it has a section on \emph{Representing data} that lists existing spatio-temporal data representations used in \proglang{R}.
+Among them, \citet{spacetime} summarises spatio-temporal data into three forms: time-wide, space-wide, and long formats.
+The associated package \pkg{spacetime} \citep{spacetime} implements four spatio-temporal layouts (full grid, sparse grid, irregular, and trajectory) to handle different space and time combinations.
+The package \pkg{stars} \citep{stars} has a new implementation to use dense arrays to represent spatio-temporal cubes.
+It also interfaces with the package \pkg{sf} \citep{sf}, commonly used for wrangling spatial data, and the \pkg{tidyverse} \citep{tidyverse} suite for general data wrangling and visualization in \proglang{R}.
+
+Still, the data representation for spatio-temporal data can be further extended and there are two reasons for this.
+Firstly, the raw data sourced in the wild is less often presented in any one of the layouts above, and fitting the raw data into a data object can sometimes be difficult.
+More often, spatio-temporal data are collected in separate 2D tables and analysts need to assemble them into a whole piece before exploring the data.
+Examples of components of spatio-temporal data can be 1) areal data recording the shape of a collection of areas of interest; 2) geostatistical data storing the longitude and latitude coordinates of locations, typically also with other metadata related to the location, and; 3) temporal data of each location across time.
+
+The other reason is about tidy data concepts \citep{tidydata} and how they should be applied to spatio-temporal data.
+According to the tidy data principles, data should be structured into 1) one row per observation, 2) one column per variable, and 3) one type of data per table.
+The long form data is preferred over wide data form given the downstream packages such as \pkg{dplyr} \citep{dplyr} and \pkg{ggplot2} \citep{ggplot2} for data wrangling and visualization.
+However, the long form can be inefficient to store feature geometries, especially for large multipolygons for hourly, daily or sub-daily periods over years, which are extensively collected and handled, for example in time series analysis.
+This poses the question of how to arrange spatial and temporal variables in a way that would make data wrangling, visualizing and analyzing spatio-temporal data easier.
+
+This paper presents a new \proglang{R} package, \pkg{cubble}, which addresses the two issues mentioned above.
+In the package, a new class, also called cubble, is proposed to organize spatial and temporal variables as two forms of a single data object so that they can be wrangled separately or combined while being kept synchronized.
+Among the four spacetime layouts in \citet{spacetime}, the \code{cubble} class can be applied to full grid, sparse grid, or irregular, but not trajectory, which is outside the scope of this work.
+The software is available from the Comprehensive \proglang{R} Archive Network (CRAN) at \url{https://CRAN.R-project.org/package=cubble}.
+
+The rest of the paper is organized as follows: Section \ref{cubble} presents the main design and functionality of the \pkg{cubble} package.
+Section \ref{others} explains how the \pkg{cubble} package deals with more advanced considerations, including data with hierarchical structure, data matching and how the package fits with existing static and interactive visualization tools.
+Moreover we also illustrate how the \pkg{cubble} package deals with spatio-temporal data transformations.
+Section \ref{examples} uses Australian weather station data and river level data as examples to demonstrate the use of the package.
+An example of how the \pkg{cubble} package handles Network Common Data Form (NetCDF) data is also provided.
+Section \ref{conclude} discuss the paper contributions and future directions.
 
@@
 
-Spatio-temporal data can encompass data with various spatial and
-temporal characteristics and different data require different structures
-for wrangling and analysis: climate weather stations record data at
-fixed point location but may suffer from potential temporal data quality
-issue such as missing data for certain days. GPS data tracks unique
-point locations at different timestamps and is represented as
-trajectories. Satellite imageries capture snapshots of landscapes at
-selected time and is commonly structured as raster data. The
-spatio-temporal data cubble address are those collected at unique fixed
-locations while allowing for irregularity in the temporal dimension,
-such as the weather station data. This corresponds to the full
-space-time and sparse space-time layouts described in the spacetime
-paper \citep{spacetime} and reproduced in Figure \ref{fig:layout}.
+Spatio-temporal data can encompass data with various spatial and temporal characteristics and different data require different structures for wrangling and analysis: climate weather stations record data at fixed point location but may suffer from potential temporal data quality issue such as missing data for certain days.
+GPS data tracks unique point locations at different timestamps and is represented as trajectories.
+Satellite imageries capture snapshots of landscapes at selected time and is commonly structured as raster data.
+The spatio-temporal data cubble address are those collected at unique fixed locations while allowing for irregularity in the temporal dimension, such as the weather station data.
+This corresponds to the full space-time and sparse space-time layouts described in the spacetime paper \citep{spacetime} and reproduced in Figure \ref{fig:layout}.
 
@@
 \caption[The full space-time and sparse space-time layouts handled by the cubble package]{The full space-time and sparse space-time layouts handled by the cubble package.
-Reproduced from Figure 1 in the spacetime paper \citep{spacetime}.}(\#fig:layout)
+Reproduced from Figure 1 in the spacetime paper \citep{spacetime}.}\label{fig:layout}
 \end{figure}
@@
 
-The cubble class is an S3 class built on tibble that allows the
-spatio-temporal data to be wrangled in two forms: a nested form (spatial
-cubble) and a long form (temporal cubble). It consists of two
-subclasses:
+The cubble class is an S3 class built on tibble that allows the spatio-temporal data to be wrangled in two forms: a nested form (spatial cubble) and a long form (temporal cubble).
+It consists of two subclasses:
 
@@
 \item
-  a spatial cubble with class
-  \texttt{c("spatial\_cubble\_df",\ "cubble\_df")}
+  a spatial cubble with class \texttt{c("spatial\_cubble\_df",\ "cubble\_df")}
 \item
-  a emporal cubble with class
-  \texttt{c("temporal\_cubble\_df",\ "cubble\_df")}
+  a emporal cubble with class \texttt{c("temporal\_cubble\_df",\ "cubble\_df")}
 \end{itemize}
 
-In a spatial cubble, spaital variables are organised as columns and
-temporal variables are nested within a specialised \texttt{ts} column.
-The cubble object printed below contains weather records in three
-airport stations (the creation of a cubble object will be explained in
-Section @ref(create)). This toy data is a subset of a larger data
-\code{climate_aus} collected from Global Historical Climatology Network
-Daily (GHCND). It records three airport stations located in Melbourne,
-Australia and includes spatial variables such as station ID, longitude,
-latitude, elevation, station name, World Meteorology Organisation ID.
-The dataset contains temporal variables including precipitation, maximum
-and minimum temperature, which can be read from the cubble header:
+In a spatial cubble, spaital variables are organised as columns and temporal variables are nested within a specialised \texttt{ts} column.
+The cubble object printed below contains weather records in three airport stations (the creation of a cubble object will be explained in Section \ref{create}).
+This toy data is a subset of a larger data \code{climate_aus} collected from Global Historical Climatology Network Daily (GHCND).
+It records three airport stations located in Melbourne, Australia and includes spatial variables such as station ID, longitude, latitude, elevation, station name, World Meteorology Organisation ID.
+The dataset contains temporal variables including precipitation, maximum and minimum temperature, which can be read from the cubble header:
 
@@
 
-In a temporal cubble, temporal variables are expanded in the long form
-and spatial variables are stored as a data attribute:
+In a temporal cubble, temporal variables are expanded in the long form and spatial variables are stored as a data attribute:
 
@@
 
-A cubble object inherits the attributes from tibble (and its
-subclasses): \texttt{class}, \texttt{row.names}, and \texttt{names}.
-Additionally, it has three specialised attributes: \texttt{key},
-\texttt{index}, and \texttt{coords}. Readers who are familiar with the
-\texttt{key} and \texttt{index} attributes from the \texttt{tsibble}
-package would already understand the two arguments. In cubble, the
-\texttt{key} attribute identifies the row in the spatial cubble (given
-the use of \texttt{tidyr::nest()} for nesting), and when combined with
-the \texttt{index} argument, it identifies the row in the temporal
-cubble. Currently, cubble only supports one variable as the key, and the
-accepted temporal classes for index includes the base R classes
-\texttt{Date}, \texttt{POSIXlt}, \texttt{POSIXct}, as well as tsibble's
-\texttt{yearmonth}, \texttt{yearweek}, and \texttt{yearquarter} classes.
+A cubble object inherits the attributes from tibble (and its subclasses): \texttt{class}, \texttt{row.names}, and \texttt{names}.
+Additionally, it has three specialised attributes: \texttt{key}, \texttt{index}, and \texttt{coords}.
+Readers who are familiar with the \texttt{key} and \texttt{index} attributes from the \texttt{tsibble} package would already understand the two arguments.
+In cubble, the \texttt{key} attribute identifies the row in the spatial cubble (given the use of \texttt{tidyr::nest()} for nesting), and when combined with the \texttt{index} argument, it identifies the row in the temporal cubble.
+Currently, cubble only supports one variable as the key, and the accepted temporal classes for index includes the base R classes \texttt{Date}, \texttt{POSIXlt}, \texttt{POSIXct}, as well as tsibble's \texttt{yearmonth}, \texttt{yearweek}, and \texttt{yearquarter} classes.
 
 The \texttt{coords} attribute represents an ordered pair of coordinates.
-It can be either an unprojected pair of longitude and latitude, or a
-projected easting and northing value. The \texttt{sf} package is used
-under the hood to calculate the bounding box, displayed in the header of
-a spatial cubble.
-
-The temporal cubble has a special attribute called \texttt{spatial} to
-store the spatial variables. Shortcut functions are available to extract
-attributes, for example, \code{spatial()} for extracting spatial
-variables from the temporal cubble:
+It can be either an unprojected pair of longitude and latitude, or a projected easting and northing value.
+The \texttt{sf} package is used under the hood to calculate the bounding box, displayed in the header of a spatial cubble.
+
+The temporal cubble has a special attribute called \texttt{spatial} to store the spatial variables.
+Shortcut functions are available to extract attributes, for example, \code{spatial()} for extracting spatial variables from the temporal cubble:
 
@@
 
-In many cases, spatio-temporal data arrive in separate tables for
-analysis. For example, in climate data, analysts may initially receive
-station data containing geographic location information, recorded
-variables and their recording periods. They can then query the temporal
-variables using the stations of interest to obtain the relevant temporal
-data. Alternatively, analyses may begin as purely spatial or temporal,
-and analysts may obtain additional temporal or spatial data to expand
-the result to spatio-temporal.
-
-The function \code{make_cubble()} composes a \code{cubble} object from a
-spatial table (\code{spatial}) and a temporal table (\code{temporal}),
-along with three attributes introduced in the subsection @ref(object):
-\code{key}, \code{index}, and \code{coords}. The following code creates
-the spatial \code{cubble} object shown previously:
+In many cases, spatio-temporal data arrive in separate tables for analysis.
+For example, in climate data, analysts may initially receive station data containing geographic location information, recorded variables and their recording periods.
+They can then query the temporal variables using the stations of interest to obtain the relevant temporal data.
+Alternatively, analyses may begin as purely spatial or temporal, and analysts may obtain additional temporal or spatial data to expand the result to spatio-temporal.
+
+The function \code{make_cubble()} composes a \code{cubble} object from a spatial table (\code{spatial}) and a temporal table (\code{temporal}), along with three attributes introduced in the subsection \ref{object}: \code{key}, \code{index}, and \code{coords}.
+The following code creates the spatial \code{cubble} object shown previously:
 
@@
 
-Spatio-temporal data in other foreign objects can be coerced into a
-\code{cubble} object with the function \code{as_cubble()}. This includes
-casting from a \code{tibble} or \code{data.frame} with both spatial and
-temporal information, a NetCDF object, a \code{stars} object
-\citep{stars}, and a \code{sftime} object \citep{sftime}. The two
-examples below show the casting from a tibble and a NetCDF object. The
-dataset \code{climate_flat} combines the spatial data, \code{stations},
-with the temporal data, \code{meteo}, into a single tibble object. It
-can be coerced into a cubble using:
+Spatio-temporal data in other foreign objects can be coerced into a \code{cubble} object with the function \code{as_cubble()}.
+This includes casting from a \code{tibble} or \code{data.frame} with both spatial and temporal information, a NetCDF object, a \code{stars} object \citep{stars}, and a \code{sftime} object \citep{sftime}.
+The two examples below show the casting from a tibble and a NetCDF object.
+The dataset \code{climate_flat} combines the spatial data, \code{stations}, with the temporal data, \code{meteo}, into a single tibble object.
+It can be coerced into a cubble using:
 
@@
 
-Table @ref(tab:funs) summarises the functions implemented in the
-\pkg{cubble} package and Table @ref(tab:methods) details the methods
-implemented for each of the three cubble classes. The \code{cubble_df}
-class handles methods that behave consistently in both spatial
-(\code{spatial_cubble_df}) and temporal cubble
-(\code{temporal_cubble_df}). When the method has distinct behavior for
-the spatial cubble and temporal cubble, it is implemented separately for
-each subclass.
+Table \ref{tab:funs} summarises the functions implemented in the \pkg{cubble} package and Table \ref{tab:methods} details the methods implemented for each of the three cubble classes.
+The \code{cubble_df} class handles methods that behave consistently in both spatial (\code{spatial_cubble_df}) and temporal cubble (\code{temporal_cubble_df}).
+When the method has distinct behavior for the spatial cubble and temporal cubble, it is implemented separately for each subclass.
 
@@
   >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.8750}}@{}}
-\caption{(\#tab:funs) Functions implemented in the \pkg{cubble}
-package}\tabularnewline
+\caption{\label{tab:funs} Functions implemented in the \pkg{cubble} package}\tabularnewline
 \toprule()
@@
 \endhead
-base R & \texttt{{[}}, \texttt{{[}{[}\textless{}-},
-\texttt{names\textless{}-} \\
-tidyverse & \texttt{dplyr\_row\_slice}, \texttt{dplyr\_col\_modify},
-\texttt{dplyr\_reconstruct}, \texttt{select}, \texttt{mutate},
-\texttt{arrange}, \texttt{filter}, \texttt{group\_by}, \texttt{ungroup},
-\texttt{summarise}, \texttt{select}, \texttt{slice}, \texttt{rowwise},
-\texttt{rename}, \texttt{bind\_rows}, \texttt{bind\_cols},
-\texttt{relocate}, \texttt{type\_sum}, the slice family
-(\texttt{slice\_head}, \texttt{slice\_tail}, \texttt{slice\_max},
-\texttt{slice\_min}, \texttt{slice\_sample}) and the join family
-(\texttt{left\_join}, \texttt{right\_join}, \texttt{inner\_join},
-\texttt{full\_join}, \texttt{anti\_join}, \texttt{semi\_join}) \\
-cubble & \texttt{as\_cubble}, \texttt{cubble}, \texttt{make\_cubble},
-\texttt{check\_key} \texttt{face\_temporal}, \texttt{face\_spatial},
-\texttt{unfold}, \texttt{key}, \texttt{key\_vars}, \texttt{key\_data},
-\texttt{index}, \texttt{index\_var}, \texttt{coords}, \texttt{spatial},
-\texttt{match\_sites}, \texttt{match\_spatial},
-\texttt{match\_temporal}, \texttt{geom\_glyph},
-\texttt{geom\_glyph\_box}, \texttt{geom\_glyph\_line},
-\texttt{make\_spatial\_sf}, \texttt{make\_temporal\_tsibble},
-\texttt{fill\_gaps}, and \texttt{scan\_gaps} \\
+base R & \texttt{{[}}, \texttt{{[}{[}\textless{}-}, \texttt{names\textless{}-} \\
+tidyverse & \texttt{dplyr\_row\_slice}, \texttt{dplyr\_col\_modify}, \texttt{dplyr\_reconstruct}, \texttt{select}, \texttt{mutate}, \texttt{arrange}, \texttt{filter}, \texttt{group\_by}, \texttt{ungroup}, \texttt{summarise}, \texttt{select}, \texttt{slice}, \texttt{rowwise}, \texttt{rename}, \texttt{bind\_rows}, \texttt{bind\_cols}, \texttt{relocate}, \texttt{type\_sum}, the slice family (\texttt{slice\_head}, \texttt{slice\_tail}, \texttt{slice\_max}, \texttt{slice\_min}, \texttt{slice\_sample}) and the join family (\texttt{left\_join}, \texttt{right\_join}, \texttt{inner\_join}, \texttt{full\_join}, \texttt{anti\_join}, \texttt{semi\_join}) \\
+cubble & \texttt{as\_cubble}, \texttt{cubble}, \texttt{make\_cubble}, \texttt{check\_key} \texttt{face\_temporal}, \texttt{face\_spatial}, \texttt{unfold}, \texttt{key}, \texttt{key\_vars}, \texttt{key\_data}, \texttt{index}, \texttt{index\_var}, \texttt{coords}, \texttt{spatial}, \texttt{match\_sites}, \texttt{match\_spatial}, \texttt{match\_temporal}, \texttt{geom\_glyph}, \texttt{geom\_glyph\_box}, \texttt{geom\_glyph\_line}, \texttt{make\_spatial\_sf}, \texttt{make\_temporal\_tsibble}, \texttt{fill\_gaps}, and \texttt{scan\_gaps} \\
 \bottomrule()
@@
   >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.7500}}@{}}
-\caption{(\#tab:methods) Methods implemented for the three \code{cubble}
-classes.}\tabularnewline
+\caption{\label{tab:methods} Methods implemented for the three \code{cubble} classes.}\tabularnewline
 \toprule()
@@
 \endhead
-\texttt{cubble\_df} &
-\texttt{{[}{[}\textless{}-,\ dplyr\_col\_modify,\ key\_data,\ key\_vars,\ key,\ print} \\
-\texttt{spatial\_cubble\_df} &
-\texttt{{[},\ names\textless{}-,\ \ tbl\_sum,\ dplyr\_reconstruct,\ dplyr\_row\_slice,\ face\_spatial,\ face\_temporal,\ unfold,\ arrange,\ rename,\ rowwise,\ group\_by,\ ungroup,\ select,\ spatial,\ summarise,\ unfold,\ update\_cubble} \\
-\texttt{temporal\_cubble\_df} &
-\texttt{{[},\ names\textless{}-,\ tbl\_sum,\ arrange,\ dplyr\_reconstruct,\ dplyr\_row\_slice,\ face\_spatial,\ face\_temporal,\ unfold,\ fill\_gaps,\ group\_by,\ ungroup,\ \ rename,\ rowwise,\ scan\_gaps,\ select,\ spatial,\ summarise,\ tbl\_sum,\ bind\_rows,\ bind\_cols,\ update\_cubble} \\
+\texttt{cubble\_df} & \texttt{{[}{[}\textless{}-,\ dplyr\_col\_modify,\ key\_data,\ key\_vars,\ key,\ print} \\
+\texttt{spatial\_cubble\_df} & \texttt{{[},\ names\textless{}-,\ \ tbl\_sum,\ dplyr\_reconstruct,\ dplyr\_row\_slice,\ face\_spatial,\ face\_temporal,\ unfold,\ arrange,\ rename,\ rowwise,\ group\_by,\ ungroup,\ select,\ spatial,\ summarise,\ unfold,\ update\_cubble} \\
+\texttt{temporal\_cubble\_df} & \texttt{{[},\ names\textless{}-,\ tbl\_sum,\ arrange,\ dplyr\_reconstruct,\ dplyr\_row\_slice,\ face\_spatial,\ face\_temporal,\ unfold,\ fill\_gaps,\ group\_by,\ ungroup,\ \ rename,\ rowwise,\ scan\_gaps,\ select,\ spatial,\ summarise,\ tbl\_sum,\ bind\_rows,\ bind\_cols,\ update\_cubble} \\
 \bottomrule()
@@
 
-Specifically, the pair of verbs, \code{face_temporal()} and
-\code{face_spatial()}, pivot the cubble object between the spatial and
-temporal face of the multivariate spatio-temporal cube, as illustrated
-in Figure \ref{fig:face}. The code below uses the function
-\code{face_temporal()} and \code{face_spatial()} to transform between a
-spatial cubble (\code{cb_nested}) and a temporal one (\code{cb_long}),
-introduced earlier in subsection @ref(object):
+Specifically, the pair of verbs, \code{face_temporal()} and \code{face_spatial()}, pivot the cubble object between the spatial and temporal face of the multivariate spatio-temporal cube, as illustrated in Figure \ref{fig:face}.
+The code below uses the function \code{face_temporal()} and \code{face_spatial()} to transform between a spatial cubble (\code{cb_nested}) and a temporal one (\code{cb_long}), introduced earlier in subsection \ref{object}:
 
@@
 
-The pair of verbs are exact inverse and apply both functions on a cubble
-object will result in the object itself:
+The pair of verbs are exact inverse and apply both functions on a cubble object will result in the object itself:
 
@@
 \caption{An illustration of the function \code{face\_temporal()} and \code{face\_spatial()}: \code{face\_temporal()} converts a spatial cubble (neseted form) into a temporal cubble (long form) to focus on the temporal variables.
-Conversely, \code{face\_spatial()} transforms a temporal cubble into a spatial one to emphasize spatial variables.}(\#fig:face)
+Conversely, \code{face\_spatial()} transforms a temporal cubble into a spatial one to emphasize spatial variables.}\label{fig:face}
 \end{figure}
@@
 \hypertarget{compatibility-with-tsibble-and-sf}{%
-\subsection{Compatibility with tsibble and
-sf}\label{compatibility-with-tsibble-and-sf}}
-
-Analysts often have their preferred spatial or temporal data structure
-that they prefer to use for spatio-temporal analysis. For example, the
-\code{tbl_ts} class from the \code{tsibble} package \citep{tsibble} is
-commonly used in time series forecasting and similarly, the \code{sf}
-class \citep{sf} is frequently used in spatial data science. In
-\code{cubble}, analysts have the flexibility to combine these two
-structures together by allowing the spatial component to also be an
-\code{sf} object and the temporal component to also be a \code{tsibble}
-object.
+\subsection{Compatibility with tsibble and sf}\label{compatibility-with-tsibble-and-sf}}
+
+Analysts often have their preferred spatial or temporal data structure that they prefer to use for spatio-temporal analysis.
+For example, the \code{tbl_ts} class from the \code{tsibble} package \citep{tsibble} is commonly used in time series forecasting and similarly, the \code{sf} class \citep{sf} is frequently used in spatial data science.
+In \code{cubble}, analysts have the flexibility to combine these two structures together by allowing the spatial component to also be an \code{sf} object and the temporal component to also be a \code{tsibble} object.
 
 \hypertarget{using-a-tsibble-object-as-the-temporal-component}{%
-\subsubsection{Using a tsibble object as the temporal
-component}\label{using-a-tsibble-object-as-the-temporal-component}}
-
-The \code{key} and \code{index} arguments in a \code{cubble} object
-corresponds to the \code{tsibble} counterparts and they can be safely
-omitted, if the temporal component is a \code{tsibble} object,
-i.e.~\code{meteo_ts} in the example below. The \code{tsibble} class from
-the input will be carried over to the \code{cubble} object:
+\subsubsection{Using a tsibble object as the temporal component}\label{using-a-tsibble-object-as-the-temporal-component}}
+
+The \code{key} and \code{index} arguments in a \code{cubble} object corresponds to the \code{tsibble} counterparts and they can be safely omitted, if the temporal component is a \code{tsibble} object, i.e.~\code{meteo_ts} in the example below.
+The \code{tsibble} class from the input will be carried over to the \code{cubble} object:
 
@@
 
-The temporal cubble shows \code{[tsibble]} in the header to indicate the
-object also being in a \code{tbl_ts} class (\code{tsibble}). Methods
-applies to the \code{tbl_ts} class can also be applied to the temporal
-cubble objects, for example, checking whether the data contain temporal
-gaps:
+The temporal cubble shows \code{[tsibble]} in the header to indicate the object also being in a \code{tbl_ts} class (\code{tsibble}).
+Methods applies to the \code{tbl_ts} class can also be applied to the temporal cubble objects, for example, checking whether the data contain temporal gaps:
 
@@
 
-An existing \code{cubble} object can promote its temporal component to a
-\code{tsibble} object by applying \code{make_temporal_tsibble()}. The
-following code illustrates this with the object \code{cb_long} created
-in section @ref(create) and the promoted cubble object is equal to the
-cubble object originally created from a \code{tsibble} object:
+An existing \code{cubble} object can promote its temporal component to a \code{tsibble} object by applying \code{make_temporal_tsibble()}.
+The following code illustrates this with the object \code{cb_long} created in section \ref{create} and the promoted cubble object is equal to the cubble object originally created from a \code{tsibble} object:
 
@@
 \hypertarget{using-an-sf-object-as-the-spatial-component}{%
-\subsubsection{Using an sf object as the spatial
-component}\label{using-an-sf-object-as-the-spatial-component}}
-
-Similarly, an \code{sf} object can be supplied as the spatial component
-to create a \code{cubble} object, with the \code{coords} argument being
-omitted. This opens up the possibility to represent fixed area with
-polygons or multipolygons (see Applications @ref(covid)) and the
-\code{coords} argument will be calculated as the centroids of the
-(multi)polygons. The \code{[sf]} print in the cubble header suggest an
-spatial component being also a \code{sf} object:
+\subsubsection{Using an sf object as the spatial component}\label{using-an-sf-object-as-the-spatial-component}}
+
+Similarly, an \code{sf} object can be supplied as the spatial component to create a \code{cubble} object, with the \code{coords} argument being omitted.
+This opens up the possibility to represent fixed area with polygons or multipolygons (see Applications \ref{covid}) and the \code{coords} argument will be calculated as the centroids of the (multi)polygons.
+The \code{[sf]} print in the cubble header suggest an spatial component being also a \code{sf} object:
 
@@
 
-The \code{sf} functions applicable to a \code{cubble} object have been
-listed in Table @ref(tab:funs) and the following code shows how to
-perform coordinate transformation with \code{st_transform} on a cubble
-object:
+The \code{sf} functions applicable to a \code{cubble} object have been listed in Table \ref{tab:funs} and the following code shows how to perform coordinate transformation with \code{st_transform} on a cubble object:
 
@@
 
-The counterpart to promote the spatial component in an existing
-\code{cubble} to be an \code{sf} object is \code{make_spatial_sf()}:
+The counterpart to promote the spatial component in an existing \code{cubble} to be an \code{sf} object is \code{make_spatial_sf()}:
 
@@
 \hypertarget{tidyverse}{%
-\subsection{Comparison to other spatio-temporal
-classes}\label{tidyverse}}
-
-In \proglang{R}, there are other existing spatio-temporal data structure
-and this section compares and contrasts \pkg{cubble} with other existing
-alternative, specifically \pkg{stars} and \pkg{sftime}. The \pkg{stars}
-package \citep{stars} uses an array structure, as oppose to tibble, to
-represent multivariate spatio-temporal data. While both \pkg{stars} and
-\pkg{cubble} support vector and raster data, it is a matter of choice on
-which structure to use given the application. Analysts working on
-satellite imageries may prefer the array structure in \pkg{stars}, while
-others originally working with spatio-temporal data in 2D data frames
-may find \pkg{cubble} easier to adopt from their existing computing
-workflow.
-
-The \pkg{sftime} package \citep{sftime} also builds from a tibble object
-and its focus is on handling irregular spatio-temporal data. This means
-\pkg{sftime} can also handle full space-time grids and sparse space-time
-layouts represented in \pkg{cubble}, but \pkg{cubble} uses nesting to
-avoid storing spatial variables repetitively at each timestamp. This
-provides memory efficiency when data is observed frequent, i.e.~daily or
-sub-daily, or the spatial geometry is expensive to repeat, i.e.~polygons
-or multipolygons. Consider the \code{climate_aus} data in the
-\pkg{cubble} package with 639 stations observed daily in a single year
-2020. The created \code{sftime} object is approximately 14 times larger
-than the corresponding \code{cubble} object (118.24 MB vs.~8.52 MB).
+\subsection{Comparison to other spatio-temporal classes}\label{tidyverse}}
+
+In \proglang{R}, there are other existing spatio-temporal data structure and this section compares and contrasts \pkg{cubble} with other existing alternative, specifically \pkg{stars} and \pkg{sftime}.
+The \pkg{stars} package \citep{stars} uses an array structure, as oppose to tibble, to represent multivariate spatio-temporal data.
+While both \pkg{stars} and \pkg{cubble} support vector and raster data, it is a matter of choice on which structure to use given the application.
+Analysts working on satellite imageries may prefer the array structure in \pkg{stars}, while others originally working with spatio-temporal data in 2D data frames may find \pkg{cubble} easier to adopt from their existing computing workflow.
+
+The \pkg{sftime} package \citep{sftime} also builds from a tibble object and its focus is on handling irregular spatio-temporal data.
+This means \pkg{sftime} can also handle full space-time grids and sparse space-time layouts represented in \pkg{cubble}, but \pkg{cubble} uses nesting to avoid storing spatial variables repetitively at each timestamp.
+This provides memory efficiency when data is observed frequent, i.e.~daily or sub-daily, or the spatial geometry is expensive to repeat, i.e.~polygons or multipolygons.
+Consider the \code{climate_aus} data in the \pkg{cubble} package with 639 stations observed daily in a single year 2020.
+The created \code{sftime} object is approximately 14 times larger than the corresponding \code{cubble} object (118.24 MB vs.~8.52 MB).
 
@@
 
-One common task when working with spatio-temporal data is to match
-nearby sites. For example, we may want to verify the location of an old
-list of stations with current stations, or we may want to match the data
-from different data sources. In cubble, spatial and temporal matching
-are performed using the functions \texttt{match\_spatial()} and
-\texttt{match\_temporal()}. The \texttt{match\_spatial()} function
-calculates the spatial distance between observations in two cubble
-objects. Various distance measures are available (check
-\texttt{sf::st\_distance}). Analysts can specify the number of matched
-groups to output using the \texttt{spatial\_n\_group} argument (default
-to 4 groups) and the number of matches per group using the
-\texttt{spatial\_n\_group} argument (default to 1, one-to-one matching).
+One common task when working with spatio-temporal data is to match nearby sites.
+For example, we may want to verify the location of an old list of stations with current stations, or we may want to match the data from different data sources.
+In cubble, spatial and temporal matching are performed using the functions \texttt{match\_spatial()} and \texttt{match\_temporal()}.
+The \texttt{match\_spatial()} function calculates the spatial distance between observations in two cubble objects.
+Various distance measures are available (check \texttt{sf::st\_distance}).
+Analysts can specify the number of matched groups to output using the \texttt{spatial\_n\_group} argument (default to 4 groups) and the number of matches per group using the \texttt{spatial\_n\_group} argument (default to 1, one-to-one matching).
 The syntax to use \texttt{match\_spatial()} is:
@@
 
-The function \texttt{match\_temporal()} calculates the similarity
-between time series within spatially matched groups. Two identifiers are
-required: one for separating each matched group (\texttt{match\_id}) and
-one for separating the two data sources (\texttt{data\_id}). The
-argument \texttt{temporal\_by} uses the \texttt{by} syntax from dplyr's
-\texttt{*\_join} to specify the temporal variables to match.
-
-The similarity score between two time series is calculated using a
-matching function, which can be customised by the analysts. The matching
-function takes two time series as a list and returns a single numerical
-score. This allows for flexibility in using existing time series
-distance calculation implementation. By default, cubble implements a
-simple peak matching algorithm (\texttt{match\_peak}) that counts the
-number of peaks in two time series that fall within a specified temporal
-window. The syntax to use \texttt{match\_temporal()} is
+The function \texttt{match\_temporal()} calculates the similarity between time series within spatially matched groups.
+Two identifiers are required: one for separating each matched group (\texttt{match\_id}) and one for separating the two data sources (\texttt{data\_id}).
+The argument \texttt{temporal\_by} uses the \texttt{by} syntax from dplyr's \texttt{*\_join} to specify the temporal variables to match.
+
+The similarity score between two time series is calculated using a matching function, which can be customised by the analysts.
+The matching function takes two time series as a list and returns a single numerical score.
+This allows for flexibility in using existing time series distance calculation implementation.
+By default, cubble implements a simple peak matching algorithm (\texttt{match\_peak}) that counts the number of peaks in two time series that fall within a specified temporal window.
+The syntax to use \texttt{match\_temporal()} is
 
@@
 
-The workflow with the \code{cubble} class works well with an interactive
-graphics pipeline (e.g., \citet{buja1988elements};
-\citet{buja1996interactive}; \citet{sutherland2000orca};
-\citet{xie2014reactive}; \citet{cheng2016enabling}) that is available in
-\proglang{R} with the package \pkg{crosstalk} \citep{crosstalk}. Figure
-\ref{fig:illu-interactive} illustrates the linking between a map and
-multiple time series in a \code{cubble} object. Both the map, generated
-from the spatial cubble, and the time series plot, generated from the
-temporal cubble, are shared crosstalk objects. When a user selects a
-location on the map (a), the corresponding site is highlighted. This
-selection activates a row in the spatial \code{cubble}, which is then
-communicated to the temporal \code{cubble}, resulting in the selection
-of all observations with the same ID in (b). Consequently the temporal
-\code{cubble} highlights the corresponding series in the time series
-plot (c). Linking can also be initiated from the time series plot, by
-selecting points on the time series. This action activates rows with the
-same ID in the temporal \code{cubble} and the row in the spatial
-\code{cubble}, which will be highlighted on the map. Please refer to the
-appendix for the corresponding illustration.
+The workflow with the \code{cubble} class works well with an interactive graphics pipeline (e.g., \citet{buja1988elements}; \citet{buja1996interactive}; \citet{sutherland2000orca}; \citet{xie2014reactive}; \citet{cheng2016enabling}) that is available in \proglang{R} with the package \pkg{crosstalk} \citep{crosstalk}.
+Figure \ref{fig:illu-interactive} illustrates the linking between a map and multiple time series in a \code{cubble} object.
+Both the map, generated from the spatial cubble, and the time series plot, generated from the temporal cubble, are shared crosstalk objects.
+When a user selects a location on the map (a), the corresponding site is highlighted.
+This selection activates a row in the spatial \code{cubble}, which is then communicated to the temporal \code{cubble}, resulting in the selection of all observations with the same ID in (b).
+Consequently the temporal \code{cubble} highlights the corresponding series in the time series plot (c).
+Linking can also be initiated from the time series plot, by selecting points on the time series.
+This action activates rows with the same ID in the temporal \code{cubble} and the row in the spatial \code{cubble}, which will be highlighted on the map.
+Please refer to the appendix for the corresponding illustration.
 
@@
 When a station is selected on the map (a), the corresponding row in the spatial \code{cubble} will be activated.
-This will link to all the rows with the same id in the temporal \code{cubble} (b) and update the line plot (c).}(\#fig:illu-interactive)
+This will link to all the rows with the same id in the temporal \code{cubble} (b) and update the line plot (c).}\label{fig:illu-interactive}
 \end{figure}
@@
 
-Sometimes, we wish to communicate spatial and temporal information
-collectively through visualisation. This can be achieved through several
-graphical displays: one can make faceted maps across time, creating map
-animations, or constructing interactive graphics to link between map and
-time series plot. The concept of glyph maps was initially proposed in
-\citet{Wickham2012-yr}. The underlying idea is to transform the temporal
-coordinates into spatial coordinates so that time series plot can be
-displayed on the map.
-
-The package \texttt{GGally} initially implement the glyph map. It uses
-\texttt{glyphs()} to calculate the axis transformation and then uses
-\texttt{geom\_polygon()} to draw the map. In cubble, a ggproto
-implementation \texttt{geom\_glyph()} performs the linear algebra
-internally as data transformation . The \texttt{geom\_glyph()} requires
-four aesthetics: \texttt{x\_major}, \texttt{y\_major},
-\texttt{x\_minor}, and \texttt{y\_minor}. The major axes are the outer
-spatial coordinates and the minor axes are the inner/ temporal
-coordinates:
+Sometimes, we wish to communicate spatial and temporal information collectively through visualisation.
+This can be achieved through several graphical displays: one can make faceted maps across time, creating map animations, or constructing interactive graphics to link between map and time series plot.
+The concept of glyph maps was initially proposed in \citet{Wickham2012-yr}.
+The underlying idea is to transform the temporal coordinates into spatial coordinates so that time series plot can be displayed on the map.
+
+The package \texttt{GGally} initially implement the glyph map.
+It uses \texttt{glyphs()} to calculate the axis transformation and then uses \texttt{geom\_polygon()} to draw the map.
+In cubble, a ggproto implementation \texttt{geom\_glyph()} performs the linear algebra internally as data transformation .
+The \texttt{geom\_glyph()} requires four aesthetics: \texttt{x\_major}, \texttt{y\_major}, \texttt{x\_minor}, and \texttt{y\_minor}.
+The major axes are the outer spatial coordinates and the minor axes are the inner/ temporal coordinates:
 
@@
 \item
-  glyph scale to fixed (\code{global_rescale} defaults to \code{TRUE})
-  or free, and
+  glyph scale to fixed (\code{global_rescale} defaults to \code{TRUE}) or free, and
 \item
-  reference boxes and lines with \code{geom_glyph_box()} and
-  \code{geom_glyph_line()}.
+  reference boxes and lines with \code{geom_glyph_box()} and \code{geom_glyph_line()}.
 \end{itemize}
@@
 
-Five examples are chosen to illustrate different aspects of the
-\pkg{cubble} package: creating a \code{cubble} object from two
-Coronavirus (COVID) data tables with the complication of differing
-location names, using spatial transformations to make a glyph map of
-seasonal temperature changes, matching river level data and weather
-station records for analysis of water supply, reading NetCDF format data
-to reproduce a climate reanalysis plot, and the workflow to create
-complex interactive linked plots.
+Five examples are chosen to illustrate different aspects of the \pkg{cubble} package: creating a \code{cubble} object from two Coronavirus (COVID) data tables with the complication of differing location names, using spatial transformations to make a glyph map of seasonal temperature changes, matching river level data and weather station records for analysis of water supply, reading NetCDF format data to reproduce a climate reanalysis plot, and the workflow to create complex interactive linked plots.
 
 \hypertarget{covid}{%
-\subsection{Victoria COVID spatio-temporal incidence and
-spread}\label{covid}}
-
-Since the start of the pandemic, the Victoria State Government in
-Australia has been providing daily COVID counts per Logal Government
-Area (LGA). This data can be combined with map polygon data, available
-from the Australian Bureau of Statistics (ABS), to visualize COVID
-incidence and spread. In the \pkg{cubble} package, the COVID count data
-(\code{covid}) and the LGA information (\code{lga}) are available as a
-\code{tsibble} object and an \code{sf} object respectively.
-
-A cubble object can be created from separate spatial and temporal
-component using the \code{make_cubble()} function, introduced in Section
-@ref(create). The \code{by} argument is used to specify the joining
-variable from the two component using the \code{*_join()} \code{by}
-syntax:
+\subsection{Victoria COVID spatio-temporal incidence and spread}\label{covid}}
+
+Since the start of the pandemic, the Victoria State Government in Australia has been providing daily COVID counts per Logal Government Area (LGA).
+This data can be combined with map polygon data, available from the Australian Bureau of Statistics (ABS), to visualize COVID incidence and spread.
+In the \pkg{cubble} package, the COVID count data (\code{covid}) and the LGA information (\code{lga}) are available as a \code{tsibble} object and an \code{sf} object respectively.
+
+A cubble object can be created from separate spatial and temporal component using the \code{make_cubble()} function, introduced in Section \ref{create}. The \code{by} argument
+is used to specify the joining variable from the two component using the \code{*_join()} \code{by} syntax:
 
@@
 
-The warning message suggests the slight difference of LGA encoding used
-by Victoria government and ABS and prompts analysts to use the function
-\code{check_key()} to identify the mismatches:
+The warning message suggests the slight difference of LGA encoding used by Victoria government and ABS and prompts analysts to use the function \code{check_key()} to identify the mismatches:
 
@@
 
-The result of the \code{check_key()} function is a list containing three
-elements: 1) matched keys from both tables, 2) potentially pairable
-keys, and 3) others. Here, the main mismatch arises from the two LGAs:
-Kingston and Latrobe (Kingston is a LGA in both Victoria and South
-Australia and Latrobe is a LGA in both Victoria and Tasmania). Analysts
-can modify the input spatial and temporal data accordingly and recreate
-the cubble object:
+The result of the \code{check_key()} function is a list containing three elements: 1) matched keys from both tables, 2) potentially pairable keys, and 3) others.
+Here, the main mismatch arises from the two LGAs: Kingston and Latrobe (Kingston is a LGA in both Victoria and South Australia and Latrobe is a LGA in both Victoria and Tasmania).
+Analysts can modify the input spatial and temporal data accordingly and recreate the cubble object:
 
@@
 \hypertarget{historicaltmax}{%
-\subsection{Australian historical maximum
-temperature}\label{historicaltmax}}
+\subsection{Australian historical maximum temperature}\label{historicaltmax}}
 
 The GHCN provides daily climate measures from stations across the world.
-The data used here (\code{historical_tmax}) is a subset extracted using
-the package \pkg{rnoaa} \citep{rnoaa}, containing the records of maximum
-temperature for 237 Australian stations from \ensuremath{\infty{}}
-through \ensuremath{-\infty{}} and provides information also on the
-latitude, longitude and elevation of each of the stations. The goal of
-this example is to compare the monthly average maximum temperature
-between two periods, 1971-1975 and 2016-2020, for stations in Victoria
-and New South Wales (NSW), using a glyph map.
-
-First, the stations need to be filtered to those in Victoria and NSW by
-using the station identifiers, stored within the 11 digits of the
-\code{id} variable entries. The country code is in the first 5 digits
-(Australia is represented by ``ASN00'') and the next 6 digits encode the
-station following the Australian Bureau of Meteorology (BOM) \citep{bom}
-coding protocols. The NSW stations correspond to entries in the range
-46-75 and the Victorian stations to 76-90. Filtering Victoria and NSW
-stations is a \emph{spatial operation} and hence uses the spatial
-\code{cubble}:
+The data used here (\code{historical_tmax}) is a subset extracted using the package \pkg{rnoaa} \citep{rnoaa}, containing the records of maximum temperature for 237 Australian stations from \ensuremath{\infty{}} through \ensuremath{-\infty{}} and provides information also on the latitude, longitude and elevation of each of the stations.
+The goal of this example is to compare the monthly average maximum temperature between two periods, 1971-1975 and 2016-2020, for stations in Victoria and New South Wales (NSW), using a glyph map.
+
+First, the stations need to be filtered to those in Victoria and NSW by using the station identifiers, stored within the 11 digits of the \code{id} variable entries.
+The country code is in the first 5 digits (Australia is represented by ``ASN00'') and the next 6 digits encode the station following the Australian Bureau of Meteorology (BOM) \citep{bom} coding protocols.
+The NSW stations correspond to entries in the range 46-75 and the Victorian stations to 76-90.
+Filtering Victoria and NSW stations is a \emph{spatial operation} and hence uses the spatial \code{cubble}:
 
@@
 
-Next, the monthly maximum average temperature is calculated for both
-periods. This is a \emph{temporal operation} requiring a switch into the
-temporal \code{cubble} using the \code{face_temporal()} function. In
-addition, a new indicator for the two time periods of interest is
-created before the calculation of monthly averages:
+Next, the monthly maximum average temperature is calculated for both periods.
+This is a \emph{temporal operation} requiring a switch into the temporal \code{cubble} using the \code{face_temporal()} function.
+In addition, a new indicator for the two time periods of interest is created before the calculation of monthly averages:
 
@@
 
-A quick check on the number of observations for each location is made,
-revealing that there are several with less than 24 observations -- these
-stations lack temperature values for some months. In this example, those
-stations are removed by switching to the spatial \code{cubble} to
-operate on the spatial component over time, and then, move back into the
-temporal \code{cubble} (to make the glyph map):
+A quick check on the number of observations for each location is made, revealing that there are several with less than 24 observations -- these stations lack temperature values for some months.
+In this example, those stations are removed by switching to the spatial \code{cubble} to operate on the spatial component over time, and then, move back into the temporal \code{cubble} (to make the glyph map):
 
@@
 
-In order to create a glyph map displaying the monthly series (Figure
-\ref{fig:basic-manip}), the spatial variables need to be unfolded with
-the temporal variables. The reason being that the major (\code{long},
-\code{lat}) and minor (\code{month}, \code{tmax}) coordinates need to be
-on the same table to create the glyph map. The \code{geom_glyph()}
-function does both the transformation and the plotting.
+In order to create a glyph map displaying the monthly series (Figure \ref{fig:basic-manip}), the spatial variables need to be unfolded with the temporal variables.
+The reason being that the major (\code{long}, \code{lat}) and minor (\code{month}, \code{tmax}) coordinates need to be on the same table to create the glyph map.
+The \code{geom_glyph()} function does both the transformation and the plotting.
 
@@
 The corresponding average time series for the cobar station are display on the top left corner.
-From the glyph map we can observe that the monthly trend is similar for all locations (low in the winter, high in the summer), and small increased temperatures, particularly in late summer can be seen at most stations in NSW.}(\#fig:basic-manip)
+From the glyph map we can observe that the monthly trend is similar for all locations (low in the winter, high in the summer), and small increased temperatures, particularly in late summer can be seen at most stations in NSW.}\label{fig:basic-manip}
 \end{figure}
@@
 
-Glyph maps work well to explore temporal patterns across spatial
-locations, particularly when the spatial locations are gridded. In this
-example, they are irregularly spaced, which can result in overlapping
-glyphs obscuring each other.
+Glyph maps work well to explore temporal patterns across spatial locations, particularly when the spatial locations are gridded.
+In this example, they are irregularly spaced, which can result in overlapping glyphs obscuring each other.
 
 \hypertarget{river-levels-and-rainfall-in-victoria}{%
-\subsection{River levels and rainfall in
-Victoria}\label{river-levels-and-rainfall-in-victoria}}
-
-One common task when working with spatio-temporal data is to match
-nearby sites. For example, we may want to verify the location of an old
-list of stations with current stations, or we may want to match the data
-from different data sources. In this example, we will introduce the
-spatial and temporal matching in cubble using an example on matching
-river level data with precipitation in Victoria, Australia.
-
-The water level data collected by the Bureau of Meteorology, can be
-compared with the precipitation since rainfall can directly impact water
-level in river. Both \texttt{climate\_vic} and \texttt{river} are cubble
-objects, and we can obtain a summary of the 10 closest pairs between
-them:
+\subsection{River levels and rainfall in Victoria}\label{river-levels-and-rainfall-in-victoria}}
+
+One common task when working with spatio-temporal data is to match nearby sites.
+For example, we may want to verify the location of an old list of stations with current stations, or we may want to match the data from different data sources.
+In this example, we will introduce the spatial and temporal matching in cubble using an example on matching river level data with precipitation in Victoria, Australia.
+
+The water level data collected by the Bureau of Meteorology, can be compared with the precipitation since rainfall can directly impact water level in river.
+Both \texttt{climate\_vic} and \texttt{river} are cubble objects, and we can obtain a summary of the 10 closest pairs between them:
 
@@
 
-The result can also be returned as cubble objects by setting the
-argument \texttt{return\_cubble\ =\ TRUE}. The output is be a list where
-each element is a paired cubble object. To combine all the results into
-a single cubble, you can use \texttt{bind\_rows()}. In the case when a
-site in the second cubble (the \texttt{river} data here) is matched to
-two stations in the first cubble (\texttt{climate\_vic} here), the
-binding may not be successful since cubble requires unique rows in the
-nested form. In the summary table above, the river station
-\texttt{226027} is matched to more than one weather station:
-\texttt{ASN00085072} (group 3) and \texttt{ASN00085298} (group 5).
-Similarly, the river station \texttt{230200} is matched in group 7 and
-8). In such cases, you can either deselect one pair before combining, or
-work with the list output with the \texttt{purrr::map} syntax:
+The result can also be returned as cubble objects by setting the argument \texttt{return\_cubble\ =\ TRUE}.
+The output is be a list where each element is a paired cubble object.
+To combine all the results into a single cubble, you can use \texttt{bind\_rows()}.
+In the case when a site in the second cubble (the \texttt{river} data here) is matched to two stations in the first cubble (\texttt{climate\_vic} here), the binding may not be successful since cubble requires unique rows in the nested form.
+In the summary table above, the river station \texttt{226027} is matched to more than one weather station: \texttt{ASN00085072} (group 3) and \texttt{ASN00085298} (group 5).
+Similarly, the river station \texttt{230200} is matched in group 7 and 8).
+In such cases, you can either deselect one pair before combining, or work with the list output with the \texttt{purrr::map} syntax:
 
@@
 
-For temporal matching, we match the variable
-\texttt{Water\_course\_level} from the river data to \texttt{prcp} in
-the weather station data. The variable \texttt{group} and \texttt{types}
-identify the matching group and the two datasets:
+For temporal matching, we match the variable \texttt{Water\_course\_level} from the river data to \texttt{prcp} in the weather station data.
+The variable \texttt{group} and \texttt{types} identify the matching group and the two datasets:
 
@@
 
-Similarly, the cubble output can be returned using the argument
-\texttt{return\_cubble\ =\ TRUE}. Here we select the four pairs with the
-highest number of matching peaks:
+Similarly, the cubble output can be returned using the argument \texttt{return\_cubble\ =\ TRUE}. Here we select the four pairs with the highest number of matching peaks:
 
@@
 
-Figure \ref{fig:matching} shows four matched pairs on the map (a) and
-standardized data as time series (b) with concurrent increasing.
+Figure \ref{fig:matching} shows four matched pairs on the map (a) and standardized data as time series (b) with concurrent increasing.
 
@@
 The water level reflects the increase in precipitation.
-The numbers (1, 7, 6, 9) indicate the group index derived from spatial matching, only those that were selectd by temporal matching are shown here.}(\#fig:matching)
+The numbers (1, 7, 6, 9) indicate the group index derived from spatial matching, only those that were selectd by temporal matching are shown here.}\label{fig:matching}
 \end{figure}
@@
 \hypertarget{era5-climate-reanalysis-data}{%
-\subsection{ERA5: climate reanalysis
-data}\label{era5-climate-reanalysis-data}}
-
-The ERA5 reanalysis \citep{hersbach2020era5} provides hourly estimates
-of atmospheric, land and oceanic climate variables on a global scale.
-The hourly pressure level data in the NetCDF format can be downloaded
-from Copernicus Climate Data Store (CDS) \citep{cds} or via the
-\pkg{ecmwfr} package \citep{ecwmfr}. This example reproduces the row
-created from ERA5 reanalysis data shown in Figure 19 by
-\citet{hersbach2020era5}. The plot shows the southern polar vortex
-splitting into two on 2002-09-26, and further splitting into four on
-2002-10-04. Further explanation of why this is interesting can be found
-in the figure source, and also in \citet{simmons2020global} and
-\citet{simmons2005ecmwf}.
-
-To work with NetCDF data in R, several packages are available, including
-\code{ncdf4}, \code{RNetCDF}, and \code{tidync}. The following code
-converts a NetCDF object of class \code{ncdf4} into a cubble object:
+\subsection{ERA5: climate reanalysis data}\label{era5-climate-reanalysis-data}}
+
+The ERA5 reanalysis \citep{hersbach2020era5} provides hourly estimates of atmospheric, land and oceanic climate variables on a global scale.
+The hourly pressure level data in the NetCDF format can be downloaded from Copernicus Climate Data Store (CDS) \citep{cds} or via the \pkg{ecmwfr} package \citep{ecwmfr}.
+This example reproduces the row created from ERA5 reanalysis data shown in Figure 19 by \citet{hersbach2020era5}.
+The plot shows the southern polar vortex splitting into two on 2002-09-26, and further splitting into four on 2002-10-04.
+Further explanation of why this is interesting can be found in the figure source, and also in \citet{simmons2020global} and \citet{simmons2005ecmwf}.
+
+To work with NetCDF data in R, several packages are available, including \code{ncdf4}, \code{RNetCDF}, and \code{tidync}.
+The following code converts a NetCDF object of class \code{ncdf4} into a cubble object:
 
@@
 
-Analysts can extract a subset of the NetCDF data with the arguments
-\code{vars}, \code{long_range} and \code{lat_range}. In this example,
-the variables q (specific humidity) and z (geopotential) are selected
-and the coordinates are subsetted to every degree in longitude and
-latitude:
+Analysts can extract a subset of the NetCDF data with the arguments \code{vars}, \code{long_range} and \code{lat_range}.
+In this example, the variables q (specific humidity) and z (geopotential) are selected and the coordinates are subsetted to every degree in longitude and latitude:
 
@@
 
-Once the NetCDF data is coerced into a cubble object, subsequent
-analysis can be conducted to filter on the date of interest, scale the
-variable specific humidity and create visualisation in ggplot as in
-Figure \ref{fig:netcdf}.
+Once the NetCDF data is coerced into a cubble object, subsequent analysis can be conducted to filter on the date of interest, scale the variable specific humidity and create visualisation in ggplot as in Figure \ref{fig:netcdf}.
 
@@
 \caption[A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020) to illustrate the break-up of sourthern polar vortex in late September and early October 2002]{A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020) to illustrate the break-up of sourthern polar vortex in late September and early October 2002.
-The polar vortex, signalled by the high specific humidity, splits into two on 2002-09-26 and further splits into four on 2002-10-04.}(\#fig:netcdf)
+The polar vortex, signalled by the high specific humidity, splits into two on 2002-09-26 and further splits into four on 2002-10-04.}\label{fig:netcdf}
 \end{figure}
@@
 \hypertarget{australian-temperature-range}{%
-\subsection{Australian temperature
-range}\label{australian-temperature-range}}
-
-Interactive graphics can be useful because they make it possible to look
-at the data in multiple ways on-the-fly. This is especially important
-for spatio-temporal data, where we would like to interactively connect
-spatial and temporal displays. This example describes the process of
-using the \pkg{cubble} package with the \pkg{crosstalk} package to build
-an interactive display connecting a map of Australia, with ribbon plots
-of temperature range observed at the stations. The purpose is to explore
-the variation of monthly temperature range over the country. Figure
-\ref{fig:interactive-linking} shows three snapshots of the
-interactivity.
-
-The key steps are to convert both the nested and long forms of the data
-into shared \code{crosstalk} objects, and to plot these side-by-side.
-The two are linked by the station identifier.
+\subsection{Australian temperature range}\label{australian-temperature-range}}
+
+Interactive graphics can be useful because they make it possible to look at the data in multiple ways on-the-fly.
+This is especially important for spatio-temporal data, where we would like to interactively connect spatial and temporal displays.
+This example describes the process of using the \pkg{cubble} package with the \pkg{crosstalk} package to build an interactive display connecting a map of Australia, with ribbon plots of temperature range observed at the stations.
+The purpose is to explore the variation of monthly temperature range over the country.
+Figure \ref{fig:interactive-linking} shows three snapshots of the interactivity.
+
+The key steps are to convert both the nested and long forms of the data into shared \code{crosstalk} objects, and to plot these side-by-side. The two are linked by the station identifier.
 
@@
 
-Plot (a) shows the initial state of the interactive display: all
-locations are shown as dots on the map, coloured by temperature range,
-and the right plot shows the ribbons representing maximum to minimum for
-all stations. In plot (b) the ``Mount Elizabeth'' station, which shows a
-high variance colour on the initial map, is selected on the map and this
-produces the ribbon on the right. In plot (c) the lowest temperature in
-August is selected, which is ``Thredbo'' station on the left map. It was
-surprising to us that this was not a station in Tasmania, so for
-comparison a station in Tasmania is selected on the map to show in
-relation to Thredbo. We can see that Thredbo has a bigger winter dip in
-temperature, and although Tasmania is cold generally, it's temperatures
-are more constant
+Plot (a) shows the initial state of the interactive display: all locations are shown as dots on the map, coloured by temperature range, and the right plot shows the ribbons representing maximum to minimum for all stations.
+In plot (b) the ``Mount Elizabeth'' station, which shows a high variance colour on the initial map, is selected on the map and this produces the ribbon on the right.
+In plot (c) the lowest temperature in August is selected, which is ``Thredbo'' station on the left map.
+It was surprising to us that this was not a station in Tasmania, so for comparison a station in Tasmania is selected on the map to show in relation to Thredbo.
+We can see that Thredbo has a bigger winter dip in temperature, and although Tasmania is cold generally, it's temperatures are more constant
 
@@
 The bottom row first selects the lowest temperature in August in the seasonal display, which highlights the corresponding station on the map (Thredbo).
-Another  station, located in the Tasmania Island, is then selected to compare its temperature variation with the Thredbo station.}(\#fig:interactive-linking)
+Another  station, located in the Tasmania Island, is then selected to compare its temperature variation with the Thredbo station.}\label{fig:interactive-linking}
 \end{figure}
@@
 
-This paper presents an \proglang{R} package \pkg{cubble} for organizing,
-manipulating and visualizing spatio-temporal data. The package
-introduces a new data class for spatio-temporal data, \code{cubble},
-that connects the time invariant and varying variables and that allows
-the user to work with a nested and long form of the data. This work adds
-capabilities into the spatio-temporal practitioners toolbox to integrate
-it with a tidy data framework. The data structure and functions
-introduced in this package can be used and combined with existing
-spatial data analysis packages such as \pkg{sf}, data wrangling packages
-such as \pkg{dplyr}, and visualization packages such as \pkg{ggplot2},
-\pkg{plotly} and \pkg{leaflet}.
-
-Numerous examples are provided in the main text, appendix and package
-vignettes. These include creating and coercing data with mismatched
-names, handling hierarchical data, matching time series spatially and
-temporally, reproducing ERA5 plots from NetCDF data. Visualization of
-the \code{cubble} objects can be done with interactive graphic pipelines
-using \pkg{crosstalk}, \pkg{plotly} and \pkg{leaflet}.
-
-There are several possible future directions for the work. The data
-structure only described fixed spatial sites, and it could be useful to
-provide tools to accommodate moving coordinates as might be encountered
-in animal movement data. That could be achieved with a list-column for
-the location coordinates, and an additional form that these locations
-can be pivoted into, like the long form for temporal variables. For
-multiple measured variables, the \pkg{cubble} package could be
-integrated with dimension reduction methods, and dynamic graphics for
-multiple dimensions such as a tour \citep{wickham_tourr_2011}.
+This paper presents the \proglang{R} package \pkg{cubble} for organizing, wrangling and visualizing spatio-temporal data.
+The package introduces a new data class, \code{cubble}, consisting of two subclass, spatial cubble and a temporal cubble, to organise spatio-temporal data in two different formats within the tidy data framework.
+The data structure and functions introduced in this package can be used and combined with existing spatial and temporal data analysis packages such as \pkg{sf} and \pkg{tsibble}, data wrangling packages such as \pkg{dplyr}, and visualization packages such as \pkg{ggplot2}, \pkg{plotly}, and \pkg{leaflet}.
+
+The paper includes numerous examples to illustrate the utility of \pkg{cubble} as a data structure for spatio-temporal analysis.
+These examples cover different tasks of a typical analysis workflow: handling data with spatial and temporal misalignment, matching data from multiple sources, and creating both static and interactive spatio-temporal visualisation.
+For future directions, other commonly-used spatial or temporal data structures can be integrated into \pkg{cubble} to extend analysts' familiar spatial and temporal toolkit to spatio-temporal.
 
@@
 
-This work is funded by a Commonwealth Scientific and Industrial Research
-Organisation (CSIRO) Data61 Scholarship and started while Nicolas
-Langrené was affiliated with CSIRO's Data61. The article is created
-using the package \pkg{knitr} \citep{knitr} and \pkg{rmarkdown}
-\citep{rmarkdown} in \proglang{R} with the \code{rticles::jss_article}
-template. The source code for reproducing this paper can be found at:
-\url{https://github.com/huizezhang-sherry/paper-cubble}.
+This work is funded by a Commonwealth Scientific and Industrial Research Organisation (CSIRO) Data61 Scholarship and started while Nicolas Langrené was affiliated with CSIRO's Data61.
+The article is created using the package \pkg{knitr} \citep{knitr} and \pkg{rmarkdown} \citep{rmarkdown} in \proglang{R} with the \code{rticles::jss_article} template.
+The source code for reproducing this paper can be found at: \url{https://github.com/huizezhang-sherry/paper-cubble}.
 

### 2023-06-24 main.Rmd
update introduction

@@
 In addition, the \pkg{cubble} package facilitates visual and numerical explorations of the data while easing data wrangling and modelling.
-The \code{cubble} class and the functions provided in the \pkg{cubble} \proglang{R} package equip users with the capability to handle hierarchical spatial and temporal structures.
 The \code{cubble} class and the tools implemented in the package are illustrated with different examples of Australian climate data.
@@
 ---
-<!-- Universität für Bodenkultur Wiens \AND -->
+
 ```{r setup, echo = FALSE}
@@
 The Comprehensive \proglang{R} Archive Network (CRAN) task view SpatioTemporal [@ctvspatiotemporal] gathers information about \proglang{R} packages designed for spatio-temporal data and it has a section on *Representing data* that lists existing spatio-temporal data representations used in \proglang{R}.
-Among them, @spacetime summarises spatio-temporal data into three forms: time-wide, space-wide, and long formats.
-The associated package \pkg{spacetime} [@spacetime] implements four spatio-temporal layouts (full grid, sparse grid, irregular, and trajectory) to handle different space and time combinations.
-The package \pkg{stars} [@stars] has a new implementation to use dense arrays to represent spatio-temporal cubes.
-It also interfaces with the package \pkg{sf} [@sf], commonly used for wrangling spatial data, and the \pkg{tidyverse} [@tidyverse] suite for general data wrangling and visualization in \proglang{R}.
+Among them, the \pkg{spacetime} package [@spacetime] implements four S4 classes to handle spatio-temporal data with different spatio-temporal layouts (full grid, sparse grid, irregular, and trajectory).
+The \pkg{stars} package [@stars] implements an S3 class built from dense arrays.
 
-Still, the data representation for spatio-temporal data can be further extended and there are two reasons for this.
-Firstly, the raw data sourced in the wild is less often presented in any one of the layouts above, and fitting the raw data into a data object can sometimes be difficult.
-More often, spatio-temporal data are collected in separate 2D tables and analysts need to assemble them into a whole piece before exploring the data.
-Examples of components of spatio-temporal data can be 1) areal data recording the shape of a collection of areas of interest; 2) geostatistical data storing the longitude and latitude coordinates of locations, typically also with other metadata related to the location, and; 3) temporal data of each location across time.
+Still, these two implementations are not necessarily easy to work in analysis for analysts with a tidyverse mindset.
+In tidyverse, data are in tables and the tidy data concept [@tidydata] prescribes three principles on how data should be organised for easier analysis as 1) one observation a row, 2) one variable a column, and 3) one type of observation a table.
+The third principle of tidy data is particularly relevant for spatio-temporal data since spatial and temporal data are naturally observed at different units: the location and location at different times.
+While the tidyverse suite implements data wrangling and visualisation tools operated on a single table, there has not been many tools for handling relational data for spatio-temporal analysis.
+This motivates a new design to organise spatio-temporal data in a way that would make data wrangling, visualizing and analyzing easier.
 
-The other reason is about tidy data concepts [@tidydata] and how they should be applied to spatio-temporal data.
-According to the tidy data principles, data should be structured into 1) one row per observation, 2) one column per variable, and 3) one type of data per table.
-The long form data is preferred over wide data form given the downstream packages such as \pkg{dplyr} [@dplyr] and \pkg{ggplot2} [@ggplot2] for data wrangling and visualization.
-However, the long form can be inefficient to store feature geometries, especially for large multipolygons for hourly, daily or sub-daily periods over years, which are extensively collected and handled, for example in time series analysis.
-This poses the question of how to arrange spatial and temporal variables in a way that would make data wrangling, visualizing and analyzing spatio-temporal data easier.
-
-This paper presents a new \proglang{R} package, \pkg{cubble}, which addresses the two issues mentioned above.
-In the package, a new class, also called cubble, is proposed to organize spatial and temporal variables as two forms of a single data object so that they can be wrangled separately or combined while being kept synchronized.
-Among the four spacetime layouts in @spacetime, the \code{cubble} class can be applied to full grid, sparse grid, or irregular, but not trajectory, which is outside the scope of this work.
+This paper presents a new \proglang{R} package, \pkg{cubble}, which implements a new cubble class to organize spatial and temporal variables as two forms of a single data object so that they can be wrangled separately or combined, while being kept synchronized.
+Among the four spacetime layouts in @spacetime, the \code{cubble} class can handle the full grid layout and the sparse grid layout.
 The software is available from the Comprehensive \proglang{R} Archive Network (CRAN) at https://CRAN.R-project.org/package=cubble.
@@
 The rest of the paper is organized as follows: Section \@ref(cubble) presents the main design and functionality of the \pkg{cubble} package.
-Section \@ref(others) explains how the \pkg{cubble} package deals with more advanced considerations, including data with hierarchical structure, data matching and how the package fits with existing static and interactive visualization tools.
+Section \@ref(others) explains how the \pkg{cubble} package deals with more advanced considerations, including data matching and how the package fits with existing static and interactive visualization tools.
 Moreover we also illustrate how the \pkg{cubble} package deals with spatio-temporal data transformations.
-Section \@ref(examples) uses Australian weather station data and river level data as examples to demonstrate the use of the package.
+Section \@ref(examples) uses primarily Australian weather station data as examples to demonstrate the use of the package.
 An example of how the \pkg{cubble} package handles Network Common Data Form (NetCDF) data is also provided.
@@
 The spatio-temporal data cubble address are those collected at unique fixed locations while allowing for irregularity in the temporal dimension, such as the weather station data.
-This corresponds to the full space-time and sparse space-time layouts described in the spacetime paper [@spacetime] and reproduced in Figure  \ref{fig:layout}.
-
-```{r layout, echo = FALSE, fig.cap="The full space-time and sparse space-time layouts handled by the cubble package.
-Reproduced from Figure 1 in the spacetime paper \\citep{spacetime}.", fig.width = 8, fig.height=4, fig.align='center'}
-library(tidyverse)
-library(patchwork)
-p1 <- tibble(
-  id = 1:12,
-  x = rep(1:4, each = 3),
-  y = rep(1:3, 4)
-) %>% 
-  ggplot(aes(x = as.factor(x) ,y = as.factor(y))) + 
-  geom_point() + 
-  geom_text(aes(y = y - 0.2, label = id)) + 
-  theme_bw() + 
-  theme(aspect.ratio = 1) + 
-  scale_x_discrete(labels = c("1st", "2nd", "3rd", "4th")) + 
-  scale_y_discrete(labels = c("1st", "2nd", "3rd", "4th")) + 
-  labs(x = "Time points", y = "Spatial features", title = "STF: full grid layout")
-
-p2 <- tibble(
-  id = 1:7,
-  x = c(1, 1, 1, 2, 2, 3, 4),
-  y = c(1, 2, 3, 2, 3, 1, 2),
-  label = c("1[1,1]", "2[1,2]", "3[1,3]", "4[2,2]", "5[3,2]", "6[1,3]", "7[2,4]")
-) %>% 
-  ggplot(aes(x = as.factor(x) ,y = as.factor(y))) + 
-  geom_point() + 
-  geom_text(aes(y = y - 0.2, label = label)) + 
-  theme_bw() + 
-  theme(aspect.ratio = 1) + 
-  scale_x_discrete(labels = c("1st", "2nd", "3rd", "4th")) + 
-  scale_y_discrete(labels = c("1st", "2nd", "3rd", "4th")) + 
-  labs(x = "Time points", y = "Spatial features", title = "STS: sparse grid layout")
-
-p1 | p2
-```
 
@@
 -   a spatial cubble with class `c("spatial_cubble_df", "cubble_df")`
--   a emporal cubble with class `c("temporal_cubble_df", "cubble_df")`
+-   a temporal cubble with class `c("temporal_cubble_df", "cubble_df")`
 
-In a spatial cubble, spaital variables are organised as columns and temporal variables are nested within a specialised `ts` column.
+In a spatial cubble, spatial variables are organised as columns and temporal variables are nested within a specialised `ts` column.
 The cubble object printed below contains weather records in three airport stations (the creation of a cubble object will be explained in Section \@ref(create)).
@@
 
-Since the start of the pandemic, the Victoria State Government in Australia has been providing daily COVID counts per Logal Government Area (LGA).
+Since the start of the pandemic, the Victoria State Government in Australia has been providing daily COVID counts per Local Government Area (LGA).
 This data can be combined with map polygon data, available from the Australian Bureau of Statistics (ABS), to visualize COVID incidence and spread.
@@
 
-The result of the \code{check_key()} function is a list containing three elements: 1) matched keys from both tables, 2) potentially pairable keys, and 3) others.
+The result of the \code{check_key()} function is a list containing three elements: 1) matched keys from both tables, 2) potentially paired keys, and 3) others.
 Here, the main mismatch arises from the two LGAs: Kingston and Latrobe (Kingston is a LGA in both Victoria and South Australia and Latrobe is a LGA in both Victoria and Tasmania).
@@
 
-Glyph maps work well to explore temporal patterns across spatial locations, particularly when the spatial locations are gridded.
+Glyph maps work well to explore temporal patterns across spatial locations, particularly when the spatial locations are grided.
 In this example, they are irregularly spaced, which can result in overlapping glyphs obscuring each other.

### 2023-06-24 main.tex
update introduction

@@
 The Comprehensive \proglang{R} Archive Network (CRAN) task view SpatioTemporal \citep{ctvspatiotemporal} gathers information about \proglang{R} packages designed for spatio-temporal data and it has a section on \emph{Representing data} that lists existing spatio-temporal data representations used in \proglang{R}.
-Among them, \citet{spacetime} summarises spatio-temporal data into three forms: time-wide, space-wide, and long formats.
-The associated package \pkg{spacetime} \citep{spacetime} implements four spatio-temporal layouts (full grid, sparse grid, irregular, and trajectory) to handle different space and time combinations.
-The package \pkg{stars} \citep{stars} has a new implementation to use dense arrays to represent spatio-temporal cubes.
-It also interfaces with the package \pkg{sf} \citep{sf}, commonly used for wrangling spatial data, and the \pkg{tidyverse} \citep{tidyverse} suite for general data wrangling and visualization in \proglang{R}.
-
-Still, the data representation for spatio-temporal data can be further extended and there are two reasons for this.
-Firstly, the raw data sourced in the wild is less often presented in any one of the layouts above, and fitting the raw data into a data object can sometimes be difficult.
-More often, spatio-temporal data are collected in separate 2D tables and analysts need to assemble them into a whole piece before exploring the data.
-Examples of components of spatio-temporal data can be 1) areal data recording the shape of a collection of areas of interest; 2) geostatistical data storing the longitude and latitude coordinates of locations, typically also with other metadata related to the location, and; 3) temporal data of each location across time.
-
-The other reason is about tidy data concepts \citep{tidydata} and how they should be applied to spatio-temporal data.
-According to the tidy data principles, data should be structured into 1) one row per observation, 2) one column per variable, and 3) one type of data per table.
-The long form data is preferred over wide data form given the downstream packages such as \pkg{dplyr} \citep{dplyr} and \pkg{ggplot2} \citep{ggplot2} for data wrangling and visualization.
-However, the long form can be inefficient to store feature geometries, especially for large multipolygons for hourly, daily or sub-daily periods over years, which are extensively collected and handled, for example in time series analysis.
-This poses the question of how to arrange spatial and temporal variables in a way that would make data wrangling, visualizing and analyzing spatio-temporal data easier.
-
-This paper presents a new \proglang{R} package, \pkg{cubble}, which addresses the two issues mentioned above.
-In the package, a new class, also called cubble, is proposed to organize spatial and temporal variables as two forms of a single data object so that they can be wrangled separately or combined while being kept synchronized.
-Among the four spacetime layouts in \citet{spacetime}, the \code{cubble} class can be applied to full grid, sparse grid, or irregular, but not trajectory, which is outside the scope of this work.
+Among them, the \pkg{spacetime} package \citep{spacetime} implements four S4 classes to handle spatio-temporal data with different spatio-temporal layouts (full grid, sparse grid, irregular, and trajectory).
+The \pkg{stars} package \citep{stars} implements an S3 class built from dense arrays.
+
+Still, these two implementations are not necessarily easy to work in analysis for analysts with a tidyverse mindset.
+In tidyverse, data are in tables and the tidy data concept \citep{tidydata} prescribes three principles on how data should be organised for easier analysis as 1) one observation a row, 2) one variable a column, and 3) one type of observation a table.
+The third principle of tidy data is particularly relevant for spatio-temporal data since spatial and temporal data are naturally observed at different units: the location and location at different times.
+While the tidyverse suite implements data wrangling and visualisation tools operated on a single table, there has not been many tools for handling relational data for spatio-temporal analysis.
+This motivates a new design to organise spatio-temporal data in a way that would make data wrangling, visualizing and analyzing easier.
+
+This paper presents a new \proglang{R} package, \pkg{cubble}, which implements a new cubble class to organize spatial and temporal variables as two forms of a single data object so that they can be wrangled separately or combined, while being kept synchronized.
+Among the four spacetime layouts in \citet{spacetime}, the \code{cubble} class can handle the full grid layout and the sparse grid layout.
 The software is available from the Comprehensive \proglang{R} Archive Network (CRAN) at \url{https://CRAN.R-project.org/package=cubble}.
@@
 The rest of the paper is organized as follows: Section \ref{cubble} presents the main design and functionality of the \pkg{cubble} package.
-Section \ref{others} explains how the \pkg{cubble} package deals with more advanced considerations, including data with hierarchical structure, data matching and how the package fits with existing static and interactive visualization tools.
+Section \ref{others} explains how the \pkg{cubble} package deals with more advanced considerations, including data matching and how the package fits with existing static and interactive visualization tools.
 Moreover we also illustrate how the \pkg{cubble} package deals with spatio-temporal data transformations.
-Section \ref{examples} uses Australian weather station data and river level data as examples to demonstrate the use of the package.
+Section \ref{examples} uses primarily Australian weather station data as examples to demonstrate the use of the package.
 An example of how the \pkg{cubble} package handles Network Common Data Form (NetCDF) data is also provided.
@@
 The spatio-temporal data cubble address are those collected at unique fixed locations while allowing for irregularity in the temporal dimension, such as the weather station data.
-This corresponds to the full space-time and sparse space-time layouts described in the spacetime paper \citep{spacetime} and reproduced in Figure \ref{fig:layout}.
-
-\begin{CodeChunk}
-\begin{figure}
-
-{\centering \includegraphics{/Users/hzha400/Documents/research/paper-cubble/figures/layout-1}
-
-}
-
-\caption[The full space-time and sparse space-time layouts handled by the cubble package]{The full space-time and sparse space-time layouts handled by the cubble package.
-Reproduced from Figure 1 in the spacetime paper \citep{spacetime}.}\label{fig:layout}
-\end{figure}
-\end{CodeChunk}
 
@@
 \item
-  a emporal cubble with class \texttt{c("temporal\_cubble\_df",\ "cubble\_df")}
+  a temporal cubble with class \texttt{c("temporal\_cubble\_df",\ "cubble\_df")}
 \end{itemize}
 
-In a spatial cubble, spaital variables are organised as columns and temporal variables are nested within a specialised \texttt{ts} column.
+In a spatial cubble, spatial variables are organised as columns and temporal variables are nested within a specialised \texttt{ts} column.
 The cubble object printed below contains weather records in three airport stations (the creation of a cubble object will be explained in Section \ref{create}).

### 2023-06-26 main.Rmd
polish up the glyph map example

@@
 
-```{r}
-load(here::here("data/historical_tmax.rda"))
+The Global Historical Climatology Network (GHCN) provides daily climate measures for stations worldwide.
+In the \pkg{cubble} package, the dataset \code{historical_tmax} contains daily maximum temperature data for `r nrow(historical_tmax)` stations in Australia, covering two periods: 1971-1975 and 2016-2020.
+The objective of this example is to compare the changes in maximum temperature between these two periods using a glyph map.
+
+To prevent overlapping of weather stations on teh map, we subset the stations based on minimum distance of 50km between them.
+This can be done by first promoting the spatial cubble to also be an sf object with \code{make_spatial_sf()}, calculating the distance matrix using the sf function \code{st_distance()}, and filtering the stations with dplyr's \code{filter()}:
+
+```{r echo = TRUE}
+a <- historical_tmax %>% make_spatial_sf() %>% st_distance()
+a[upper.tri(a, diag = TRUE)] <- 1e6
+
+(tmax <- historical_tmax |> 
+  filter(rowSums(a < units::as_units(50, "km")) == 0))
 ```
 
-The GHCN provides daily climate measures from stations across the world.
-The data used here (\code{historical_tmax}) is a subset extracted using the package \pkg{rnoaa} [@rnoaa], containing the records of maximum temperature for `r nrow(historical_tmax)` Australian stations from `r range(historical_tmax$first_year)[1]` through `r range(historical_tmax$first_year)[2]` and provides information also on the latitude, longitude and elevation of each of the  stations.
-<!-- This data  is already cast into a \code{cubble} data object (please see Appendix X for the code to extract  the data from XX and prepare the data into \code{cubble} form), with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates.--> The goal of this example is to compare the monthly average maximum temperature between two periods, 1971-1975 and 2016-2020, for stations in Victoria and New South Wales (NSW), using a glyph map.
-
-First, the stations need to be filtered to those in Victoria and NSW by using the station identifiers, stored within the 11 digits of the \code{id} variable entries.
-The country code is in the first 5 digits (Australia is represented by "ASN00") and the next 6 digits encode the station following the Australian Bureau of Meteorology (BOM) [@bom] coding protocols.
-The NSW stations correspond to entries in the range 46-75 and the Victorian stations to 76-90.
-Filtering Victoria and NSW stations is a *spatial operation* and hence uses the spatial \code{cubble}:
-
-<!-- BOM link: http://www.bom.gov.au/climate/cdo/about/site-num.shtml -->
+Next, daily maximum temperature is averaged into monthly series for each periods in the temporal cubble.
+The last step with \code{unfold()} moves the two coordinate columns \code{long, lat} into the temporal cubble, preparing the data for the glyph map:
 
 ```{r echo = TRUE}
-tmax <- historical_tmax |>
-  filter(between(as.numeric(stringr::str_sub(id, 7, 8)), 46, 90))
-```
-
-Next, the monthly maximum average temperature is  calculated for both periods.
-This is a *temporal operation* requiring a switch into the temporal \code{cubble} using the \code{face_temporal()} function.
-In addition, a new indicator for the two time periods of interest is created before the calculation of monthly averages:
-
-```{r echo = TRUE}
-tmax <- tmax |>
+(tmax <- tmax |>
   face_temporal() |> 
@@
   mutate(group = as.factor(lubridate::year(yearmonth)),
-         month = lubridate::month(yearmonth))
+         month = lubridate::month(yearmonth)) %>% 
+  unfold(long, lat))
 ```
@@
   face_spatial() |> 
-  rowwise() %>% 
+  rowwise() |>
   filter(nrow(ts) == 24) |>
@@
 
-In order to create a glyph map displaying the monthly series  (Figure \ref{fig:basic-manip}), the spatial variables need to be unfolded with the temporal variables.
-The reason being that the major (\code{long}, \code{lat}) and minor (\code{month}, \code{tmax}) coordinates need to be on the same table to create the glyph map.
-The \code{geom_glyph()} function does both the transformation and the plotting.
+The following code creates the glyph map (a) in Figure \ref{fig:basic-manip} with additional ggplot2 code for highlighting the single station, Cobar and styling:
 
 ```{r eval = FALSE, echo = TRUE}
-ggplot() +
-  geom_sf(data = MAP_DATA, ...) +
-  geom_glyph(data = tmax,
-             aes(x_major = long, x_minor = month,
+nsw_vic <- ozmaps::abs_ste |> 
+  filter(NAME %in% c("Victoria","New South Wales"))
+
+tmax %>% 
+  ggplot(aes(x_major = long, x_minor = month, 
              y_major = lat, y_minor = tmax,
-                 group = interaction(id, group), color = group),
-             width = 1, height = 0.5) + 
+             group = interaction(id, group))) + 
+  geom_sf(data =  nsw_vic, ...,  inherit.aes = FALSE) + 
+  geom_glyph_box(width = 0.8, height = 0.3) + 
+  geom_glyph(aes(color = group), width = 0.8, height = 0.3) +
   ...
@@
 
-```{r}
-cobar <- tmax |> filter(id == "ASN00048027") |>
+```{r glyphmap, out.width="100%", fig.width=15, fig.height=10, fig.cap="Comparison of Average maximum temperature between 1971-1975 and 2016-2020 for 54 stations in Victoria and New South Wales, Australia.
+(a) and (b): the monthly temperature series for the two periods in a glyph map and for a single station Cobar, highlighted in orange in the glyph map.
+(c) and (d): the difference series between the two periods (2016s minus 1971s) in a glyph map and for station Cobar.
+The grey horizontal line marks zero difference.
+The glyph map displaying the difference series (c) reveals more pronounced changes between the two periods, with many inland locations in New South Wales show an increased temperature in late summer (Jan-Feb) in recent years."}
+tmax2 <- tmax %>% 
+  pivot_wider(id_cols = -yearmonth, names_from = group, values_from = tmax) %>% 
+  mutate(diff = `2016` - `1971`) %>% 
+  mutate(x = long + cubble:::rescale11(month) * 0.8/2, 
+         y = lat + last(cubble:::rescale11(c(diff, 0)) * 0.3/2))
+
+box_df <- tmax |> filter(id == "ASN00048027")
+nsw_vic <- ozmaps::abs_ste |> filter(NAME %in% c("Victoria","New South Wales"))
+
+cobar1 <-  tmax |> filter(id == "ASN00048027") |>
   ggplot(aes(x = month, y = tmax, color = group)) +
@@
   theme_bw() + 
-  theme(aspect.ratio = 0.3)
+  theme(aspect.ratio = 0.3,panel.grid.minor = element_blank())
 
-```
-
-```{r basic-manip, out.width="100%", fig.height = 5, fig.width = 10, fig.cap="A glyph map of the monthly maximum average temperature for weather stations in  Victoria and New South Wales (NSW) for the periods (1971-1975, 2016-2020).
-The corresponding average time series for the cobar station are display on the top left corner.
-From the glyph map we can observe that the monthly trend is similar for all locations (low in the winter, high in the summer), and small increased temperatures, particularly in late summer can be seen at most stations in NSW."}
-tmax <- tmax |> unfold(long, lat)
-box_df <- tmax |> face_spatial() |> filter(id == "ASN00048027")
-nsw_vic <- ozmaps::abs_ste |> filter(NAME %in% c("Victoria","New South Wales"))
+cobar2 <- tmax2 |> filter(id == "ASN00048027") |>
+  ggplot(aes(x = month, y = diff)) +
+  geom_hline(yintercept = 0, color = "grey", linewidth = 0.8) + 
+  geom_line() +
+  scale_color_brewer(palette = "Dark2", guide = "none") +
+  scale_x_continuous(
+    breaks = seq(1, 12, 1),
+    labels = c("J", "F", "M", "A",
+               "M", "J", "J", "A",
+               "S", "O", "N", "D")
+  ) +
+  labs(x = "", y  = "Temp. diff. (C)") +
+  theme_bw() + 
+  theme(aspect.ratio = 0.3,panel.grid.minor = element_blank())
 
@@
              group = interaction(id, group))) + 
-  geom_sf(data = nsw_vic, fill = "grey95", 
-          color = "white", inherit.aes = FALSE) + 
-  geom_glyph_box(data = tmax, width = 0.8, height = 0.3) +
-  geom_glyph(data = tmax, aes(color = group), width = 0.8, height = 0.3) +
-    # geom_rect(
-    #   data = box_df,
-    #   aes(xmin = long - 0.6, xmax = long + 0.6,
-    #       ymin = lat - 0.12, ymax = lat + 0.35),
-    #   fill = "transparent", color = "black") +
+  geom_sf(data =  nsw_vic, color = "white", 
+          fill = "grey95", inherit.aes = FALSE) + 
+  geom_glyph_box(width = 0.8, height = 0.3) + 
+  geom_glyph(aes(color = group), width = 0.8, height = 0.3) +
+  geom_rect(data = box_df,
+            aes(xmin = long - 0.6, xmax = long + 0.6,
+                ymin = lat - 0.2, ymax = lat + 0.2),
+            fill = "transparent", color = "orange", inherit.aes = FALSE) +
   scale_color_brewer("", palette = "Dark2") + 
-  coord_sf(xlim = c(141, 154), ylim = c(-39, -28.5)) + 
+  coord_sf(xlim = c(141.2, 153.4), ylim = c(-39.2, -28.9)) + 
   ggthemes::theme_map() +
@@
 
-(p1 | cobar)  + 
-  patchwork::plot_layout(width = c(1,1)) + 
-  plot_annotation(tag_levels = 'a')
+p2 <- tmax2 %>% 
+  ggplot(aes(x_major = long, x_minor = month, 
+             y_major = lat, y_minor = diff)) + 
+  geom_sf(data =  nsw_vic, color = "white", 
+          fill = "grey95", inherit.aes = FALSE) + 
+  geom_path(aes(x = x, y = y, group = id), color = "grey", linewidth = 0.8) + 
+  geom_glyph(width = 0.8, height = 0.3) +
+  geom_rect(data = box_df,
+    aes(xmin = long - 0.6, xmax = long + 0.6,
+        ymin = lat - 0.2, ymax = lat + 0.2),
+    fill = "transparent", color = "orange", inherit.aes = FALSE) +
+  scale_color_brewer("", palette = "Dark2") + 
+  coord_sf(xlim = c(141.2, 153.4), ylim = c(-39.2, -28.9)) + 
+  ggthemes::theme_map() +
+  theme(legend.position = "bottom", legend.text = element_text(size = 10)) + 
+  guides(color = guide_legend(override.aes = list(size=2)))
 
+((p1/cobar1) | (p2/cobar2)) + 
+  patchwork::plot_layout(guides = "collect") + 
+  plot_annotation(tag_levels = 'a')  &
+  theme(legend.position='bottom')
 ```
-
-Glyph maps work well to explore temporal patterns across spatial locations, particularly when the spatial locations are grided.
-In this example, they are irregularly spaced, which can result in overlapping glyphs obscuring each other.
 

### 2023-06-26 main.tex
polish up the glyph map example

@@
 
-The GHCN provides daily climate measures from stations across the world.
-The data used here (\code{historical_tmax}) is a subset extracted using the package \pkg{rnoaa} \citep{rnoaa}, containing the records of maximum temperature for 237 Australian stations from \ensuremath{\infty{}} through \ensuremath{-\infty{}} and provides information also on the latitude, longitude and elevation of each of the stations.
-The goal of this example is to compare the monthly average maximum temperature between two periods, 1971-1975 and 2016-2020, for stations in Victoria and New South Wales (NSW), using a glyph map.
-
-First, the stations need to be filtered to those in Victoria and NSW by using the station identifiers, stored within the 11 digits of the \code{id} variable entries.
-The country code is in the first 5 digits (Australia is represented by ``ASN00'') and the next 6 digits encode the station following the Australian Bureau of Meteorology (BOM) \citep{bom} coding protocols.
-The NSW stations correspond to entries in the range 46-75 and the Victorian stations to 76-90.
-Filtering Victoria and NSW stations is a \emph{spatial operation} and hence uses the spatial \code{cubble}:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> tmax <- historical_tmax |>
-+   filter(between(as.numeric(stringr::str_sub(id, 7, 8)), 46, 90))
-\end{CodeInput}
-\end{CodeChunk}
-
-Next, the monthly maximum average temperature is calculated for both periods.
-This is a \emph{temporal operation} requiring a switch into the temporal \code{cubble} using the \code{face_temporal()} function.
-In addition, a new indicator for the two time periods of interest is created before the calculation of monthly averages:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> tmax <- tmax |>
+The Global Historical Climatology Network (GHCN) provides daily climate measures for stations worldwide.
+In the \pkg{cubble} package, the dataset \code{historical_tmax} contains daily maximum temperature data for 75 stations in Australia, covering two periods: 1971-1975 and 2016-2020.
+The objective of this example is to compare the changes in maximum temperature between these two periods using a glyph map.
+
+To prevent overlapping of weather stations on teh map, we subset the stations based on minimum distance of 50km between them.
+This can be done by first promoting the spatial cubble to also be an sf object with \code{make_spatial_sf()}, calculating the distance matrix using the sf function \code{st_distance()}, and filtering the stations with dplyr's \code{filter()}:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> a <- historical_tmax %>% make_spatial_sf() %>% st_distance()
+R> a[upper.tri(a, diag = TRUE)] <- 1e6
+R> 
+R> (tmax <- historical_tmax |> 
++   filter(rowSums(a < units::as_units(50, "km")) == 0))
+\end{CodeInput}
+\begin{CodeOutput}
+# cubble:   key: id [54], index: date, nested form
+# spatial:  [141.2652, -39.1297, 153.3633, -28.9786], Missing CRS!
+# temporal: date [date], tmax [dbl]
+  id           long   lat  elev name                     wmo_id ts      
+  <chr>       <dbl> <dbl> <dbl> <chr>                     <dbl> <list>  
+1 ASN00047016  141. -34.0    43 lake victoria storage     94692 <tibble>
+2 ASN00047019  142. -32.4    61 menindee post office      94694 <tibble>
+3 ASN00048015  147. -30.0   115 brewarrina hospital       95512 <tibble>
+4 ASN00048027  146. -31.5   260 cobar mo                  94711 <tibble>
+5 ASN00048031  149. -29.5   145 collarenebri (albert st)  95520 <tibble>
+# i 49 more rows
+\end{CodeOutput}
+\end{CodeChunk}
+
+Next, daily maximum temperature is averaged into monthly series for each periods in the temporal cubble.
+The last step with \code{unfold()} moves the two coordinate columns \code{long, lat} into the temporal cubble, preparing the data for the glyph map:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> (tmax <- tmax |>
 +   face_temporal() |> 
@@
 +   mutate(group = as.factor(lubridate::year(yearmonth)),
-+          month = lubridate::month(yearmonth))
-\end{CodeInput}
++          month = lubridate::month(yearmonth)) %>% 
++   unfold(long, lat))
+\end{CodeInput}
+\begin{CodeOutput}
+# cubble:   key: id [54], index: yearmonth, long form
+# temporal: 1971 Jan -- 2016 Dec [1M], has gaps!
+# spatial:  long [dbl], lat [dbl], elev [dbl], name [chr], wmo_id [dbl]
+  yearmonth id           tmax group month  long   lat
+      <mth> <chr>       <dbl> <fct> <dbl> <dbl> <dbl>
+1  1971 Jan ASN00047016  31.1 1971      1  141. -34.0
+2  1971 Jan ASN00047019  33.1 1971      1  142. -32.4
+3  1971 Jan ASN00048015  33.9 1971      1  147. -30.0
+4  1971 Jan ASN00048027  32.5 1971      1  146. -31.5
+5  1971 Jan ASN00048031  33.3 1971      1  149. -29.5
+# i 1,276 more rows
+\end{CodeOutput}
 \end{CodeChunk}
@@
 +   face_spatial() |> 
-+   rowwise() %>% 
++   rowwise() |>
 +   filter(nrow(ts) == 24) |>
@@
 
-In order to create a glyph map displaying the monthly series (Figure \ref{fig:basic-manip}), the spatial variables need to be unfolded with the temporal variables.
-The reason being that the major (\code{long}, \code{lat}) and minor (\code{month}, \code{tmax}) coordinates need to be on the same table to create the glyph map.
-The \code{geom_glyph()} function does both the transformation and the plotting.
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> ggplot() +
-+   geom_sf(data = MAP_DATA, ...) +
-+   geom_glyph(data = tmax,
-+              aes(x_major = long, x_minor = month,
+The following code creates the glyph map (a) in Figure \ref{fig:basic-manip} with additional ggplot2 code for highlighting the single station, Cobar and styling:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> nsw_vic <- ozmaps::abs_ste |> 
++   filter(NAME %in% c("Victoria","New South Wales"))
+R> 
+R> tmax %>% 
++   ggplot(aes(x_major = long, x_minor = month, 
 +              y_major = lat, y_minor = tmax,
-+                  group = interaction(id, group), color = group),
-+              width = 1, height = 0.5) + 
++              group = interaction(id, group))) + 
++   geom_sf(data =  nsw_vic, ...,  inherit.aes = FALSE) + 
++   geom_glyph_box(width = 0.8, height = 0.3) + 
++   geom_glyph(aes(color = group), width = 0.8, height = 0.3) +
 +   ...
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/research/paper-cubble/figures/basic-manip-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/research/paper-cubble/figures/glyphmap-1}
 
@@
 
-\caption[A glyph map of the monthly maximum average temperature for weather stations in  Victoria and New South Wales (NSW) for the periods (1971-1975, 2016-2020)]{A glyph map of the monthly maximum average temperature for weather stations in  Victoria and New South Wales (NSW) for the periods (1971-1975, 2016-2020).
-The corresponding average time series for the cobar station are display on the top left corner.
-From the glyph map we can observe that the monthly trend is similar for all locations (low in the winter, high in the summer), and small increased temperatures, particularly in late summer can be seen at most stations in NSW.}\label{fig:basic-manip}
+\caption[Comparison of Average maximum temperature between 1971-1975 and 2016-2020 for 54 stations in Victoria and New South Wales, Australia]{Comparison of Average maximum temperature between 1971-1975 and 2016-2020 for 54 stations in Victoria and New South Wales, Australia.
+(a) and (b): the monthly temperature series for the two periods in a glyph map and for a single station Cobar, highlighted in orange in the glyph map.
+(c) and (d): the difference series between the two periods (2016s minus 1971s) in a glyph map and for station Cobar.
+The grey horizontal line marks zero difference.
+The glyph map displaying the difference series (c) reveals more pronounced changes between the two periods, with many inland locations in New South Wales show an increased temperature in late summer (Jan-Feb) in recent years.}\label{fig:glyphmap}
 \end{figure}
 \end{CodeChunk}
-
-Glyph maps work well to explore temporal patterns across spatial locations, particularly when the spatial locations are grided.
-In this example, they are irregularly spaced, which can result in overlapping glyphs obscuring each other.
 

### 2023-06-26 review/B-response.Rmd
update response

@@
 
-**\textcolor{purple}{The data historical\_tmax contains daily weather station data for the two periods: 1971-1975 and 2016-2020 and amounts to 3.2MB, which is not suitable to include in the package given its size.
-The tmax\_hist data, which was included in the package, is now removed, given its similarity to the climate\_aus data, which is more useful.
-The scripts to generate historical\_tmax are provided with the paper.}**
+**\textcolor{purple}{The data historical\_tmax used in the paper now replaces the tmax\_hist data in the package.}**
 

### 2023-06-26 main.Rmd
small fix

@@
 
-To prevent overlapping of weather stations on teh map, we subset the stations based on minimum distance of 50km between them.
+To prevent overlapping of weather stations on the map, we subset the stations based on minimum distance of 50km between them.
 This can be done by first promoting the spatial cubble to also be an sf object with \code{make_spatial_sf()}, calculating the distance matrix using the sf function \code{st_distance()}, and filtering the stations with dplyr's \code{filter()}:
@@
 
-The following code creates the glyph map (a) in Figure \ref{fig:basic-manip} with additional ggplot2 code for highlighting the single station, Cobar and styling:
+The following code creates the glyph map (a) in Figure \ref{fig:glyphmap} with additional ggplot2 code for highlighting the single station, Cobar and styling:
 
-```{r eval = FALSE, echo = TRUE}
+````
 nsw_vic <- ozmaps::abs_ste |> 
@@
   ...
-```
+````
 
-```{r glyphmap, out.width="100%", fig.width=15, fig.height=10, fig.cap="Comparison of Average maximum temperature between 1971-1975 and 2016-2020 for 54 stations in Victoria and New South Wales, Australia.
+```{r glyphmap, out.width="100%", fig.width=15, fig.height=10, fig.cap="Comparison of average maximum temperature between 1971-1975 and 2016-2020 for 54 stations in Victoria and New South Wales, Australia.
 (a) and (b): the monthly temperature series for the two periods in a glyph map and for a single station Cobar, highlighted in orange in the glyph map.

### 2023-06-26 main.tex
small fix

@@
 
-The following code creates the glyph map (a) in Figure \ref{fig:basic-manip} with additional ggplot2 code for highlighting the single station, Cobar and styling:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> nsw_vic <- ozmaps::abs_ste |> 
-+   filter(NAME %in% c("Victoria","New South Wales"))
-R> 
-R> tmax %>% 
-+   ggplot(aes(x_major = long, x_minor = month, 
-+              y_major = lat, y_minor = tmax,
-+              group = interaction(id, group))) + 
-+   geom_sf(data =  nsw_vic, ...,  inherit.aes = FALSE) + 
-+   geom_glyph_box(width = 0.8, height = 0.3) + 
-+   geom_glyph(aes(color = group), width = 0.8, height = 0.3) +
-+   ...
-\end{CodeInput}
-\end{CodeChunk}
+The following code creates the glyph map (a) in Figure \ref{fig:glyphmap} with additional ggplot2 code for highlighting the single station, Cobar and styling:
+
+\begin{verbatim}
+nsw_vic <- ozmaps::abs_ste |> 
+  filter(NAME %in% c("Victoria","New South Wales"))
+
+tmax %>% 
+  ggplot(aes(x_major = long, x_minor = month, 
+             y_major = lat, y_minor = tmax,
+             group = interaction(id, group))) + 
+  geom_sf(data =  nsw_vic, ...,  inherit.aes = FALSE) + 
+  geom_glyph_box(width = 0.8, height = 0.3) + 
+  geom_glyph(aes(color = group), width = 0.8, height = 0.3) +
+  ...
+\end{verbatim}
 
@@
 
-\caption[Comparison of Average maximum temperature between 1971-1975 and 2016-2020 for 54 stations in Victoria and New South Wales, Australia]{Comparison of Average maximum temperature between 1971-1975 and 2016-2020 for 54 stations in Victoria and New South Wales, Australia.
+\caption[Comparison of average maximum temperature between 1971-1975 and 2016-2020 for 54 stations in Victoria and New South Wales, Australia]{Comparison of average maximum temperature between 1971-1975 and 2016-2020 for 54 stations in Victoria and New South Wales, Australia.
 (a) and (b): the monthly temperature series for the two periods in a glyph map and for a single station Cobar, highlighted in orange in the glyph map.

### 2023-06-27 cubble.Rmd
rename main.Rmd -> cubble.Rmd, use |> pipe

--- 2023-06-26 main.Rmd
@@
 ```{r echo = TRUE}
-ts_long %>% has_gaps()
+ts_long |> has_gaps()
 ```
@@
 ```{r echo = TRUE}
-ts_long2 <- cb_long %>% make_temporal_tsibble() 
+ts_long2 <- cb_long |> make_temporal_tsibble() 
 identical(ts_long2, ts_long)
@@
 ```{r echo =TRUE, message=FALSE}
-sf_nested %>% sf::st_transform(crs = "EPSG:3857")
+sf_nested |> sf::st_transform(crs = "EPSG:3857")
 ```
@@
 ```{r echo = TRUE}
-cb_nested %>% make_spatial_sf() 
+cb_nested |> make_spatial_sf() 
 ```
@@
 ```{r echo = TRUE}
-a <- historical_tmax %>% make_spatial_sf() %>% st_distance()
+a <- historical_tmax |> make_spatial_sf() |> st_distance()
 a[upper.tri(a, diag = TRUE)] <- 1e6
@@
   )|>
-  summarise(tmax = mean(tmax, na.rm = TRUE)) %>% 
+  summarise(tmax = mean(tmax, na.rm = TRUE)) |> 
   mutate(group = as.factor(lubridate::year(yearmonth)),
-         month = lubridate::month(yearmonth)) %>% 
+         month = lubridate::month(yearmonth)) |> 
   unfold(long, lat))
@@
 
-tmax %>% 
+tmax |> 
   ggplot(aes(x_major = long, x_minor = month, 
@@
 The glyph map displaying the difference series (c) reveals more pronounced changes between the two periods, with many inland locations in New South Wales show an increased temperature in late summer (Jan-Feb) in recent years."}
-tmax2 <- tmax %>% 
-  pivot_wider(id_cols = -yearmonth, names_from = group, values_from = tmax) %>% 
-  mutate(diff = `2016` - `1971`) %>% 
+tmax2 <- tmax |> 
+  pivot_wider(id_cols = -yearmonth, names_from = group, values_from = tmax) |> 
+  mutate(diff = `2016` - `1971`) |> 
   mutate(x = long + cubble:::rescale11(month) * 0.8/2, 
@@
 
-p1 <- tmax %>% 
+p1 <- tmax |> 
   ggplot(aes(x_major = long, x_minor = month, 
@@
 
-p2 <- tmax2 %>% 
+p2 <- tmax2 |> 
   ggplot(aes(x_major = long, x_minor = month, 
@@
   mutate(type = "climate")
-river <- cubble::river %>% mutate(type = "river") 
+river <- cubble::river |> mutate(type = "river") 
 ```
@@
 str(res_sp, max.level = 0)
-(res_sp <- res_sp[-c(5, 8)] %>% bind_rows())
+(res_sp <- res_sp[-c(5, 8)] |> bind_rows())
 ```
@@
 ```{r echo = TRUE}
-(res_tm <- res_sp %>% 
+(res_tm <- res_sp |> 
   match_temporal(
@@
 ```{r echo = TRUE}
-res_tm <- res_sp %>% 
+res_tm <- res_sp |> 
   match_temporal(
@@
     return_cubble = TRUE)
-(res_tm <- res_tm %>% bind_rows() %>% filter(group %in% c(1, 7, 6, 9)))
+(res_tm <- res_tm |> bind_rows() |> filter(group %in% c(1, 7, 6, 9)))
 ```
@@
 The numbers (1, 7, 6, 9) indicate the group index derived from spatial matching, only those that were selectd by temporal matching are shown here."}
-res_tm <- res_tm %>%  filter(group %in% c(1, 7, 6, 9))
+res_tm <- res_tm |>  filter(group %in% c(1, 7, 6, 9))
 
-res_tm_long <- res_tm %>%  
-  face_temporal() %>%  
-  unfold(group, type) %>%  
-  group_by(group, type) %>%
+res_tm_long <- res_tm |>  
+  face_temporal() |>  
+  unfold(group, type) |>  
+  group_by(group, type) |>
   mutate(matched = (matched - min(matched, na.rm = TRUE))/ 
@@
              alpha = 0.2, fill = 0.2) +
-  geom_point(data = res_tm %>% as_tibble(), 
+  geom_point(data = res_tm |> as_tibble(), 
              aes(x = long, y = lat, color = type)) +
   ggrepel::geom_label_repel(
-    data = res_tm |> filter(type == "river") %>% as_tibble(), 
+    data = res_tm |> filter(type == "river") |> as_tibble(), 
     aes(x = long, y = lat, label = group)) +
@@
 
-p2 <- res_tm_long %>%  
+p2 <- res_tm_long |>  
   ggplot(aes(x = date, y = matched, group = type,color = type)) + 
@@
 sf::sf_use_s2(FALSE)
-country <- con %>% st_geometry() %>% 
+country <- con |> st_geometry() |> 
   sf::st_crop(box) |> 

### 2023-06-27 cubble.tex
rename main.Rmd -> cubble.Rmd, use |> pipe

--- 2023-06-26 main.tex
@@
 \begin{CodeInput}
-R> ts_long %>% has_gaps()
+R> ts_long |> has_gaps()
 \end{CodeInput}
@@
 \begin{CodeInput}
-R> ts_long2 <- cb_long %>% make_temporal_tsibble() 
+R> ts_long2 <- cb_long |> make_temporal_tsibble() 
 R> identical(ts_long2, ts_long)
@@
 \begin{CodeInput}
-R> sf_nested %>% sf::st_transform(crs = "EPSG:3857")
+R> sf_nested |> sf::st_transform(crs = "EPSG:3857")
 \end{CodeInput}
@@
 \begin{CodeInput}
-R> cb_nested %>% make_spatial_sf() 
+R> cb_nested |> make_spatial_sf() 
 \end{CodeInput}
@@
 $others$spatial
-[1] "No usual address (Vic.)"               
-[2] "Migratory - Offshore - Shipping (Vic.)"
+character(0)
 
@@
 # cubble:   key: lga [80], index: date, nested form, [sf]
-# spatial:  [140.961682, -39.1591895, 149.976291, -33.9804256], WGS 84
+# spatial:  [140.961682, -39.1339581, 149.976291, -33.9960517], WGS 84
 # temporal: date [date], n [dbl], avg_7day [dbl]
   lga             long   lat                                   geometry ts      
-  <chr>          <dbl> <dbl>                         <MULTIPOLYGON [°]> <list>  
-1 Alpine (S)      147. -36.9 (((146.7258 -36.45922, 146.8033 -36.45139~ <tbl_ts>
-2 Ararat (RC)     143. -37.5 (((143.1807 -37.73152, 143.1793 -37.73242~ <tbl_ts>
-3 Ballarat (C)    144. -37.5 (((143.6622 -37.57241, 143.6686 -37.53844~ <tbl_ts>
-4 Banyule (C)     145. -37.7 (((145.1357 -37.74091, 145.1331 -37.74281~ <tbl_ts>
-5 Bass Coast (S)  146. -38.5 (((145.5207 -38.30667, 145.5418 -38.30938~ <tbl_ts>
+  <chr>          <dbl> <dbl>                             <GEOMETRY [°]> <list>  
+1 Alpine (S)      147. -36.9 POLYGON ((146.7258 -36.45922, 146.7198 -3~ <tbl_ts>
+2 Ararat (RC)     143. -37.5 POLYGON ((143.1807 -37.73152, 143.0609 -3~ <tbl_ts>
+3 Ballarat (C)    144. -37.5 POLYGON ((143.6622 -37.57241, 143.68 -37.~ <tbl_ts>
+4 Banyule (C)     145. -37.7 POLYGON ((145.1357 -37.74091, 145.1437 -3~ <tbl_ts>
+5 Bass Coast (S)  146. -38.5 MULTIPOLYGON (((145.5207 -38.30667, 145.5~ <tbl_ts>
 # i 75 more rows
@@
 \begin{CodeInput}
-R> a <- historical_tmax %>% make_spatial_sf() %>% st_distance()
+R> a <- historical_tmax |> make_spatial_sf() |> st_distance()
 R> a[upper.tri(a, diag = TRUE)] <- 1e6
@@
 +   )|>
-+   summarise(tmax = mean(tmax, na.rm = TRUE)) %>% 
++   summarise(tmax = mean(tmax, na.rm = TRUE)) |> 
 +   mutate(group = as.factor(lubridate::year(yearmonth)),
-+          month = lubridate::month(yearmonth)) %>% 
++          month = lubridate::month(yearmonth)) |> 
 +   unfold(long, lat))
@@
 \begin{CodeInput}
-R> (res_sp <- res_sp[-c(5, 8)] %>% bind_rows())
+R> (res_sp <- res_sp[-c(5, 8)] |> bind_rows())
 \end{CodeInput}
@@
 \begin{CodeInput}
-R> (res_tm <- res_sp %>% 
+R> (res_tm <- res_sp |> 
 +   match_temporal(
@@
 \begin{CodeInput}
-R> res_tm <- res_sp %>% 
+R> res_tm <- res_sp |> 
 +   match_temporal(
@@
 +     return_cubble = TRUE)
-R> (res_tm <- res_tm %>% bind_rows() %>% filter(group %in% c(1, 7, 6, 9)))
+R> (res_tm <- res_tm |> bind_rows() |> filter(group %in% c(1, 7, 6, 9)))
 \end{CodeInput}

### 2023-06-27 cubble.Rmd
update table captions

@@
 
-: (#tab:funs) Functions implemented in the \pkg{cubble} package
+: (#tab:funs) An overview of functions implemented in the \pkg{cubble} package, categorised into base R, tidyverse, and cubble functions.
 
@@
 
-: (#tab:methods) Methods implemented for the three \code{cubble} classes.
+: (#tab:methods) An overview of the methods implemented in the three \code{cubble} classes.
+Methods are implemented in the \code{cubble\_df} class when they behave consistent across the two subclasses; otherwise, they are implemented separately in each subclass.
 
@@
   mutate(lga = ifelse(lga == "Kingston (C) (Vic.)", "Kingston (C)", lga),
-         lga = ifelse(lga == "Latrobe (C) (Vic.)", "Latrobe (C)", lga)) |>
-  filter(!lga %in% check_res$others$spatial)
+         lga = ifelse(lga == "Latrobe (C) (Vic.)", "Latrobe (C)", lga))
 

### 2023-06-27 cubble.tex
update table captions

@@
   >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.8750}}@{}}
-\caption{\label{tab:funs} Functions implemented in the \pkg{cubble} package}\tabularnewline
+\caption{\label{tab:funs} An overview of functions implemented in the \pkg{cubble} package, categorised into base R, tidyverse, and cubble functions.}\tabularnewline
 \toprule()
@@
   >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.7500}}@{}}
-\caption{\label{tab:methods} Methods implemented for the three \code{cubble} classes.}\tabularnewline
+\caption{\label{tab:methods} An overview of the methods implemented in the three \code{cubble} classes.
+Methods are implemented in the \code{cubble\_df} class when they behave consistent across the two subclasses; otherwise, they are implemented separately in each subclass.}\tabularnewline
 \toprule()
@@
 +   mutate(lga = ifelse(lga == "Kingston (C) (Vic.)", "Kingston (C)", lga),
-+          lga = ifelse(lga == "Latrobe (C) (Vic.)", "Latrobe (C)", lga)) |>
-+   filter(!lga %in% check_res$others$spatial) 
++          lga = ifelse(lga == "Latrobe (C) (Vic.)", "Latrobe (C)", lga))
 R>   

### 2023-06-27 cubble.Rmd
clean up reproducibility script

@@
 ```{r echo = FALSE}
-library(cubble)
-library(tidyverse)
-library(patchwork)
-library(tsibble)
-library(sf)
+library("cubble")
+library("dplyr")
+library("ggplot2")
+library("patchwork")
+library("tsibble")
+library("sf")
+library("lubridate")
+library("ozmaps")
+library("ggthemes")
+library("ggrepel")
+library("ncdf4")
+library("rnaturalearth")
+library("colorspace")
+library("units")
 ```
@@
 climate_vic <- climate_aus |>
-  filter(between(as.numeric(stringr::str_sub(id, 7, 8)), 76, 90)) |>
+  filter(between(as.numeric(substr(id, 7, 8)), 76, 90)) |>
   mutate(type = "climate")
@@
   ggthemes::theme_map() +
-  ggplot2::theme(legend.position = "bottom",
+  theme(legend.position = "bottom",
         legend.text = element_text(size = 15),
         legend.title = element_text(size = 15)) +
-  ggplot2::labs(x = "Longitude", y = "Latitude") + 
+  labs(x = "Longitude", y = "Latitude") + 
   guides(color = guide_legend(override.aes = list(size=5)))
@@
 box <- st_bbox(c(xmin = -180, ymin = -90, xmax = 180, ymax = -15), crs = st_crs(con)) 
-sf::sf_use_s2(FALSE)
+sf_use_s2(FALSE)
 country <- con |> st_geometry() |> 
-  sf::st_crop(box) |> 
-  sf::st_cast("MULTILINESTRING")
+  st_crop(box) |> 
+  st_cast("MULTILINESTRING")
 
@@
   coord_sf(
-    crs = "+proj=stere +lat_0=-90 +lon_0=-180 +k=1 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs", default_crs = sf::st_crs(4326), clip = "off") +
+    crs = "+proj=stere +lat_0=-90 +lon_0=-180 +k=1 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs", default_crs = st_crs(4326), clip = "off") +
   theme_bw() +
@@
   labs(x ="", y = "")
-```
-
-```{r eval = FALSE}
-# prepare data/era5-pressure.nc
-library(ecmwfr)
-my_request <- list(
-  product_type = "reanalysis",
-  format = "netcdf",
-  pressure_level = "10",
-  year = "2002",
-  month = c("09", "10"),
-  day = c("04", "22", "26", "30"),
-  time = "00:00",
-  area = c(-15, -180, -90, 180),
-  variable = c("specific_humidity", 'geopotential'),
-  dataset_short_name = "reanalysis-era5-pressure-levels",
-  target = "era5-pressure.nc"
-)
-# 
-# # here you need to create an account on Climate Data Store 
-# # at https://cds.climate.copernicus.eu/cdsapp#!/home) to obtain the user ID and API key
-# # then set up the with ecmwfr::wf_set_key(user, key, service = "cds")
-# # the user below is the same `user` in the setup line
-wf_request(
-  user = "113339",
-  request = my_request,
-  transfer = TRUE,
-  path = here::here("data/"))
 ```
@@
 It was surprising to us that this was not a station in Tasmania, so for comparison a station in Tasmania is selected on the map to show in relation to Thredbo.
-We can see that Thredbo has a bigger winter dip in temperature, and although Tasmania is cold generally, it's temperatures are more constant
-
-```{r eval = FALSE}
-# This chunk generates the interactive graphic and save it as figures/linking.html, where the three screencasts in the figure comes from.
-library("crosstalk")
-library("plotly")
-
-clean <- climate_full |>
-  filter(!id %in% c("ASN00067033", "ASN00072091", "ASN00059040", 
-                    "ASN00097053", "ASN00040856", "ASN00015548")) |>
-  face_temporal() |> 
-  mutate(month = lubridate::month(date)) |>
-  group_by(month) |>
-  summarise(
-    tmax = mean(tmax, na.rm = TRUE),
-    tmin = mean(tmin, na.rm = TRUE),
-    diff = mean(tmax - tmin, na.rm = TRUE)
-    ) |> 
-  ungroup(month) |>
-  mutate(dummy_date = as.Date(glue::glue("2021-{month}-01"))) |> 
-  face_spatial() |> 
-  mutate(temp_diff_var = ts |> pull(diff) |> var(na.rm = TRUE))
-
-nested <- clean |> 
-  SharedData$new(~id, group = "cubble")
-
-long <- clean |> 
-  face_temporal() |> 
-  unfold(temp_diff_var) |> 
-  arrange(temp_diff_var) |> 
-  SharedData$new(~id, group = "cubble")
-
-state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)
-
-p1 <- nested |> 
-  ggplot() +
-  geom_sf(data = state_map, fill = "grey95", color = "white") +
-  geom_point(aes(x = long, y = lat,color = temp_diff_var, label = name)) +
-  colorspace::scale_color_continuous_sequential(
-    "Rocket",  n_interp = 7, cmax = 90, rev = TRUE, 
-    c2 = 40, l2= 85, c1 = 20, l1 = 30, name = "Var. temp. diff.") +
-  labs(x = "Longitude", y = "Latitude") + 
-  coord_sf() + 
-  theme_bw() + 
-  theme(
-    panel.grid.major = element_blank(),
-    legend.position = "none"
-    )
-
-p2 <- long |> 
-  ggplot(aes(x = dummy_date, group = id, 
-             fill = temp_diff_var, color = temp_diff_var)) +
-  geom_ribbon(aes(ymin = tmin, ymax = tmax), size = 0.1, alpha = 0.3) + 
-  geom_line(aes(y = tmin), alpha = 0.1) + 
-  geom_line(aes(y = tmax), alpha = 0.1) + 
-  geom_point(aes(y = tmax), size = 0.1) + 
-  geom_point(aes(y = tmin), size = 0.1) + 
-  colorspace::scale_fill_continuous_sequential(
-    "Rocket",  n_interp = 7, cmax = 90, rev = TRUE, 
-    c2 = 40, l2= 85, c1 = 20, l1 = 30, name = "Var. temp. diff.") +
-  colorspace::scale_colour_continuous_sequential(
-    "Rocket",  n_interp = 7, cmax = 90, rev = TRUE, 
-    c2 = 40, l2= 85, c1 = 20, l1 = 30, name = "Var. temp. diff.") +
-  scale_x_date(date_labels = "%b", date_breaks = "1 month") + 
-  labs(x = "Month", y = "Temperature") + 
-  theme_bw() + 
-  theme(
-    panel.grid.major = element_blank(),
-    legend.position = "bottom"
-    )
-
-out <- bscols(
-  ggplotly(p1, width = 900, height = 700, tooltip = "label") |>
-    highlight(
-      on = "plotly_selected", 
-      off = "plotly_deselect", 
-      opacityDim = 0.1),
-  ggplotly(p2, width = 1200, height = 700, tooltip = "label") |> 
-    highlight(
-      on = "plotly_selected", 
-      off = "plotly_deselect", 
-      opacityDim = 0.012),
-  widths = c(5, 5)
-)
-
-htmltools::save_html(out, file = "figures/linking.html")
-
-```
+We can see that Thredbo has a bigger winter dip in temperature, and although Tasmania is cold generally, it's temperatures are more constant.
 

### 2023-06-27 code/appendix.R

@@ -1,204 +1,0 @@
-## ----setup, include=FALSE-----------------------------------------------------
-knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
-library("tidyverse")
-library("cubble")
-library("patchwork")
-library("leaflet")
-library("leafpop")
-
-
-## ---- echo = FALSE------------------------------------------------------------
-load(here::here("data/station_nested.rda"))
-load(here::here("data/climate_full.rda"))
-
-
-## ----eval = FALSE, echo = TRUE------------------------------------------------
-## station_nested <- climate_full |> mutate(cluster = ...)
-
-
-## ----eval = FALSE, echo = TRUE------------------------------------------------
-## cluster_nested <- station_nested |> switch_key(cluster)
-
-
-## ----eval = FALSE, echo = TRUE------------------------------------------------
-## cluster_nested <- cluster_nested |> get_centroid()
-
-
-## ----eval = FALSE, echo = FALSE-----------------------------------------------
-## # The kmeans clustering in station_nested data can takes some time to run,
-## # This chunk runs the algorithm and save the result as staiton_nested for further use
-## coords <- cbind(prcp_aus$long, prcp_aus$lat)
-## dist_raw <- geosphere::distm(coords, coords)
-## 
-## station_long <- cubble::prcp_aus |>
-##   face_temporal(ts) |>
-##   mutate(wk = lubridate::week(date)) |>
-##   group_by(wk) |>
-##   summarise(prcp = sum(prcp, na.rm = TRUE)) |>
-##   unfold(cluster)
-## 
-## set.seed(123)
-## station_nested <- station_long |>
-##   face_spatial() |>
-##   strip_rowwise() |>
-##   mutate(cluster = kmeans(dist_raw,centers = 20, nstart = 500)$cluster)
-## save(station_nested, file = here::here("data/station_nested.rda"))
-
-
[…]

### 2023-06-27 code/download-era5.R

@@ -1,0 +1,25 @@
+# prepare data/era5-pressure.nc
+library(ecmwfr)
+my_request <- list(
+  product_type = "reanalysis",
+  format = "netcdf",
+  pressure_level = "10",
+  year = "2002",
+  month = c("09", "10"),
+  day = c("04", "22", "26", "30"),
+  time = "00:00",
+  area = c(-15, -180, -90, 180),
+  variable = c("specific_humidity", 'geopotential'),
+  dataset_short_name = "reanalysis-era5-pressure-levels",
+  target = "era5-pressure.nc"
+)
+# 
+# # here you need to create an account on Climate Data Store 
+# # at https://cds.climate.copernicus.eu/cdsapp#!/home) to obtain the user ID and API key
+# # then set up the with ecmwfr::wf_set_key(user, key, service = "cds")
+# # the user below is the same `user` in the setup line
+wf_request(
+  user = "113339",
+  request = my_request,
+  transfer = TRUE,
+  path = here::here("data/"))

### 2023-06-27 code/gen-scripts-from-rmd.R

@@ -1,7 +1,0 @@
-knitr::purl(input = here::here("jss/main.Rmd"), output = here::here("code/reproducible-script.R"))
-knitr::purl(input = here::here("jss/appendix.Rmd"), output = here::here("code/appendix.R"))
-# manually merge the two scripts into one
-# remove chunks for illustration, but keep codes for deriving data
-# remove "child = here::here("5-1-covid.Rmd")" and alike
-# change in line 5 to echo = TRUE to show all the codes, also line 446, 467, 509
-knitr::spin(here::here("code/reproducible-script.R"), knit = FALSE)

### 2023-06-27 code/interactive-graphic.R

@@ -1,0 +1,83 @@
+# This chunk generates the interactive graphic and save it as figures/linking.html, where the three screencasts in the figure comes from.
+library("crosstalk")
+library("plotly")
+
+clean <- climate_full |>
+  filter(!id %in% c("ASN00067033", "ASN00072091", "ASN00059040", 
+                    "ASN00097053", "ASN00040856", "ASN00015548")) |>
+  face_temporal() |> 
+  mutate(month = lubridate::month(date)) |>
+  group_by(month) |>
+  summarise(
+    tmax = mean(tmax, na.rm = TRUE),
+    tmin = mean(tmin, na.rm = TRUE),
+    diff = mean(tmax - tmin, na.rm = TRUE)
+  ) |> 
+  ungroup(month) |>
+  mutate(dummy_date = as.Date(glue::glue("2021-{month}-01"))) |> 
+  face_spatial() |> 
+  mutate(temp_diff_var = ts |> pull(diff) |> var(na.rm = TRUE))
+
+nested <- clean |> 
+  SharedData$new(~id, group = "cubble")
+
+long <- clean |> 
+  face_temporal() |> 
+  unfold(temp_diff_var) |> 
+  arrange(temp_diff_var) |> 
+  SharedData$new(~id, group = "cubble")
+
+state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)
+
+p1 <- nested |> 
+  ggplot() +
+  geom_sf(data = state_map, fill = "grey95", color = "white") +
+  geom_point(aes(x = long, y = lat,color = temp_diff_var, label = name)) +
+  colorspace::scale_color_continuous_sequential(
+    "Rocket",  n_interp = 7, cmax = 90, rev = TRUE, 
+    c2 = 40, l2= 85, c1 = 20, l1 = 30, name = "Var. temp. diff.") +
+  labs(x = "Longitude", y = "Latitude") + 
+  coord_sf() + 
+  theme_bw() + 
+  theme(
+    panel.grid.major = element_blank(),
+    legend.position = "none"
+  )
+
+p2 <- long |> 
+  ggplot(aes(x = dummy_date, group = id, 
+             fill = temp_diff_var, color = temp_diff_var)) +
+  geom_ribbon(aes(ymin = tmin, ymax = tmax), size = 0.1, alpha = 0.3) + 
+  geom_line(aes(y = tmin), alpha = 0.1) + 
[…]

### 2023-06-27 code/reproducible-script.R

@@ -1,619 +1,0 @@
-## ----setup, echo = FALSE------------------------------------------------------
-knitr::opts_chunk$set(
-  echo = TRUE,
-  warning = FALSE,
-  message = FALSE,
-  cache = TRUE,
-  fig.path = here::here("figures/")
-)
-
-options(prompt = "R> ", continue = "+ ",
-        tibble.print_max = 5, tibble.print_min = 5,
-        width = 70)
-
-
-## ----echo = FALSE-------------------------------------------------------------
-library(cubble)
-library(tidyverse)
-library(patchwork)
-
-## -----------------------------------------------------------------------------
-cubble::climate_flat
-
-
-## ----echo = TRUE--------------------------------------------------------------
-cubble_nested <- climate_flat |>
-  as_cubble(key = id, index = date, coords = c(long, lat))
-cubble_nested
-
-
-## ----echo = TRUE--------------------------------------------------------------
-cubble_nested |>
-  mutate(rain_day = sum(ts$prcp != 0))
-
-
-## ----echo = TRUE--------------------------------------------------------------
-cubble_long <- cubble_nested |> face_temporal()
-cubble_long
-
-
-## ----echo = TRUE--------------------------------------------------------------
-cubble_back <- cubble_long |> face_spatial()
-cubble_back
-identical(cubble_nested, cubble_back)
-
-
-## ----echo = TRUE--------------------------------------------------------------
-cubble_unfold <- cubble_long |> unfold(long, lat)
-cubble_unfold
-
-
-## ----echo = TRUE--------------------------------------------------------------
-climate_flat_ts <- climate_flat |> 
-  tsibble::as_tsibble(key = id, index = date)
-climate_flat_cb <-  climate_flat_ts |> 
-  cubble::as_cubble(coords = c(long, lat))
-climate_flat_cb
-
-
-## ----echo = TRUE--------------------------------------------------------------
-climate_flat_cb |> 
-  mutate(
-    fabletools::features(
-      ts, tmax, list(tmax_mean = mean, tmax_var = var)
-      )
-  )
-
-
-## ----eval = FALSE-------------------------------------------------------------
-## library(tidyverse)
-## #https://www.coronavirus.vic.gov.au/victorian-coronavirus-covid-19-data
[…]

### 2023-06-27 install-instructions.R

@@ -1,0 +1,16 @@
+# An R version of 4.2 is needed to run the code and it can be downloaded from https://cran.r-project.org/
+
+# The following R packages, and their dependencies, are needed to run the reproducible code:  
+
+packages <- c(
+	"cubble", "dplyr", "stringr", "ggplot2", "tibble", "knitr", 
+	"patchwork", "tsibble", "here", "lubridate", "ozmaps", 
+	"ggthemes", "ggrepel", "ncdf4", "sf", "rnaturalearth", 
+	"rnaturalearthdata", "colorspace", "units"
+)
+
+# They can be installed via:
+
+install.packages(packages, repos = "https://cloud.r-project.org/")
+
+# A cubble version >=0.3.0 is required for reproducing the paper and a dplyr version >= 1.0.8 is required for the cubble package.

### 2023-06-27 reproducible-script.R

@@ -1,0 +1,382 @@
+## ----setup, echo = FALSE------------------------------------------------------
+knitr::opts_chunk$set(
+  echo = FALSE,
+  warning = FALSE,
+  message = FALSE,
+  cache = TRUE,
+  fig.path = here::here("figures/")
+)
+
+options(prompt = "R> ", continue = "+ ",
+        tibble.print_max = 5, tibble.print_min = 5,
+        width = 80, styler.cache_root = "styler-perm")
+
+
+## ----echo = FALSE-------------------------------------------------------------
+library("cubble")
+library("dplyr")
+library("ggplot2")
+library("patchwork")
+library("tsibble")
+library("sf")
+library("lubridate")
+library("ozmaps")
+library("ggthemes")
+library("ggrepel")
+library("ncdf4")
+library("rnaturalearth")
+library("colorspace")
+library("units")
+
+
+## -----------------------------------------------------------------------------
+cb_nested <- climate_mel
+
+
+## ----echo = TRUE--------------------------------------------------------------
+cb_nested
+
+
+## -----------------------------------------------------------------------------
+cb_long <- face_temporal(climate_mel)
+
+
+## ----echo = TRUE--------------------------------------------------------------
+cb_long
+
+
+## ----class, fig.align="center", out.width = "100%", fig.cap = "Illustration of a cubble object in the long form and nested form, along with the associated attributes."----
+#knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.004.png"))
+
+
+## ----echo = TRUE--------------------------------------------------------------
+spatial(cb_long)
+
+
+## ----echo = TRUE, eval = FALSE------------------------------------------------
+## make_cubble(spatial = stations, temporal = meteo,
+##             key = id, index = date, coords = c(long, lat))
+
+
+## ----echo = TRUE--------------------------------------------------------------
+climate_flat |> as_cubble(key = id, index = date, coords = c(long, lat))
+
+
+## ----cubble-fun, echo = TRUE--------------------------------------------------
+identical(face_temporal(cb_nested), cb_long)
+identical(face_spatial(cb_long), cb_nested)
+
+
[…]

### 2023-06-27 cubble.Rmd
proofread section 2.1

@@
 
-Spatio-temporal data can encompass data with various spatial and temporal characteristics and different data require different structures for wrangling and analysis: climate weather stations record data at fixed point location but may suffer from potential temporal data quality issue such as missing data for certain days.
+Spatio-temporal data can encompass data with various spatial and temporal characteristics, requiring different structures for wrangling and analysis.
+For example, climate weather stations typically store station metadata in one table and the climate time series in another.
 GPS data tracks unique point locations at different timestamps and is represented as trajectories.
 Satellite imageries capture snapshots of landscapes at selected time and is commonly structured as raster data.
-The spatio-temporal data cubble address are those collected at unique fixed locations while allowing for irregularity in the temporal dimension, such as the weather station data.
+In this paper, we propose the \pkg{cubble} package to organise spatio-temporal data collected at unique fixed locations while allowing for irregularities in the temporal dimension, such as weather station data.
 
-The cubble class is an S3 class built on tibble that allows the spatio-temporal data to be wrangled in two forms:  a nested form (spatial cubble) and a long form (temporal cubble).
-It consists of two subclasses:
+The cubble class is an S3 class built on tibble that allows the spatio-temporal data to be wrangled in two forms (subclasses):
 
@@
 In a spatial cubble, spatial variables are organised as columns and temporal variables are nested within a specialised `ts` column.
-The cubble object printed below contains weather records in three airport stations (the creation of a cubble object will be explained in Section \@ref(create)).
-This toy data is a subset of a larger data \code{climate_aus} collected from Global Historical Climatology Network Daily (GHCND).
-It records three airport stations located in Melbourne, Australia and includes spatial variables such as station ID, longitude, latitude, elevation, station name, World Meteorology Organisation ID.
-The dataset contains temporal variables including precipitation, maximum and minimum temperature, which can be read from the cubble header:
+The spatial cubble object, \code{cb_nested}, printed below contains weather records of three airport stations from Global Historical Climatology Network Daily (GHCND).
+The spatial cubble is convenient for wrangling the spatial variables:
 
@@
 
-<!-- A nested cubble header prints the key and index variable name and its form.
-The spatial extent is given by the bounding box with CRS information (missing here).
-The nested cubble header also shows the name of temporal variables nested in the \code{ts} column and their variable types: \code{prcp [dbl], tmax [dbl], tmin [dbl]}.
--->
-
-In a temporal cubble, temporal variables are expanded in the long form and spatial variables are stored as a data attribute:
+In a temporal cubble, temporal variables are expanded in the long form and spatial variables are stored as a data attribute.
+The temporal cubble object, \code{cb_long}, contains the same spatio-temporal data as the spatial cubble object, \code{cb_nested}, but in a structure that is easier for temporal analysis:
 
@@
 Readers who are familiar with the `key` and `index` attributes from the `tsibble` package would already understand the two arguments.
-In cubble, the `key` attribute identifies the row in the spatial cubble (given the use of `tidyr::nest()` for nesting), and when combined with the `index` argument, it identifies the row in the temporal cubble.
+In cubble, the `key` attribute identifies the row in the spatial cubble (given the internal use of `tidyr::nest()` for nesting), and when combined with the `index` argument, it identifies the row in the temporal cubble.
 Currently, cubble only supports one variable as the key, and the accepted temporal classes for index includes the base R classes `Date`, `POSIXlt`, `POSIXct`, as well as tsibble's `yearmonth`, `yearweek`, and `yearquarter` classes.
-
 The `coords` attribute represents an ordered pair of coordinates.
@@
 
-In many cases, spatio-temporal data arrive in separate tables for analysis.
+<!-- In many cases, spatio-temporal data arrive in separate tables for analysis.
 For example, in climate data, analysts may initially receive station data containing geographic location information, recorded variables and their recording periods.
@@
 Alternatively, analyses may begin as purely spatial or temporal, and analysts may obtain additional temporal or spatial data to expand the result to spatio-temporal.
+-->
 
-The function \code{make_cubble()} composes a \code{cubble} object from a spatial table (\code{spatial}) and a temporal table (\code{temporal}), along with three attributes introduced in the subsection \@ref(object): \code{key}, \code{index}, and \code{coords}.
-The following code creates the spatial \code{cubble} object shown previously:
+The function \code{make_cubble()} composes a spatial cubble object from a spatial table (\code{spatial}) and a temporal table (\code{temporal}), along with three attributes introduced in the subsection \@ref(object): \code{key}, \code{index}, and \code{coords}.
+The following code creates the spatial cubble, \code{cb_nested}, from its spatial component, \code{stations} and temporal component \code{meteo}:
 
-```{r echo = TRUE, eval = FALSE}
+```{r echo = TRUE}
 make_cubble(spatial = stations, temporal = meteo,
@@
 Spatio-temporal data in other foreign objects can be coerced into a \code{cubble} object with the function \code{as_cubble()}.
-This includes casting from a \code{tibble} or \code{data.frame} with both spatial and temporal information, a NetCDF object, a \code{stars} object [@stars], and a \code{sftime} object [@sftime].
-The two examples below show the casting from a tibble and a NetCDF object.
-The dataset \code{climate_flat} combines the spatial data, \code{stations}, with the temporal data, \code{meteo}, into a single tibble object.
-It can be coerced into a cubble using:
+This includes casting from a joined \code{tibble} or \code{data.frame}, a NetCDF object, a \code{stars} object [@stars], and a \code{sftime} object [@sftime].
+In the example below, the spatial cubble object is created from \code{climate_flat}, which combines the spatial data, \code{stations}, with the temporal data, \code{meteo}, into a single tibble object:
 

### 2023-06-27 cubble.tex
proofread section 2.1

@@
 
-Spatio-temporal data can encompass data with various spatial and temporal characteristics and different data require different structures for wrangling and analysis: climate weather stations record data at fixed point location but may suffer from potential temporal data quality issue such as missing data for certain days.
+Spatio-temporal data can encompass data with various spatial and temporal characteristics, requiring different structures for wrangling and analysis.
+For example, climate weather stations typically store station metadata in one table and the climate time series in another.
 GPS data tracks unique point locations at different timestamps and is represented as trajectories.
 Satellite imageries capture snapshots of landscapes at selected time and is commonly structured as raster data.
-The spatio-temporal data cubble address are those collected at unique fixed locations while allowing for irregularity in the temporal dimension, such as the weather station data.
-
-The cubble class is an S3 class built on tibble that allows the spatio-temporal data to be wrangled in two forms: a nested form (spatial cubble) and a long form (temporal cubble).
-It consists of two subclasses:
+In this paper, we propose the \pkg{cubble} package to organise spatio-temporal data collected at unique fixed locations while allowing for irregularities in the temporal dimension, such as weather station data.
+
+The cubble class is an S3 class built on tibble that allows the spatio-temporal data to be wrangled in two forms (subclasses):
 
@@
 In a spatial cubble, spatial variables are organised as columns and temporal variables are nested within a specialised \texttt{ts} column.
-The cubble object printed below contains weather records in three airport stations (the creation of a cubble object will be explained in Section \ref{create}).
-This toy data is a subset of a larger data \code{climate_aus} collected from Global Historical Climatology Network Daily (GHCND).
-It records three airport stations located in Melbourne, Australia and includes spatial variables such as station ID, longitude, latitude, elevation, station name, World Meteorology Organisation ID.
-The dataset contains temporal variables including precipitation, maximum and minimum temperature, which can be read from the cubble header:
+The spatial cubble object, \code{cb_nested}, printed below contains weather records of three airport stations from Global Historical Climatology Network Daily (GHCND).
+The spatial cubble is convenient for wrangling the spatial variables:
 
@@
 
-In a temporal cubble, temporal variables are expanded in the long form and spatial variables are stored as a data attribute:
+In a temporal cubble, temporal variables are expanded in the long form and spatial variables are stored as a data attribute.
+The temporal cubble object, \code{cb_long}, contains the same spatio-temporal data as the spatial cubble object, \code{cb_nested}, but in a structure that is easier for temporal analysis:
 
@@
 Readers who are familiar with the \texttt{key} and \texttt{index} attributes from the \texttt{tsibble} package would already understand the two arguments.
-In cubble, the \texttt{key} attribute identifies the row in the spatial cubble (given the use of \texttt{tidyr::nest()} for nesting), and when combined with the \texttt{index} argument, it identifies the row in the temporal cubble.
+In cubble, the \texttt{key} attribute identifies the row in the spatial cubble (given the internal use of \texttt{tidyr::nest()} for nesting), and when combined with the \texttt{index} argument, it identifies the row in the temporal cubble.
 Currently, cubble only supports one variable as the key, and the accepted temporal classes for index includes the base R classes \texttt{Date}, \texttt{POSIXlt}, \texttt{POSIXct}, as well as tsibble's \texttt{yearmonth}, \texttt{yearweek}, and \texttt{yearquarter} classes.
-
 The \texttt{coords} attribute represents an ordered pair of coordinates.
@@
 
-In many cases, spatio-temporal data arrive in separate tables for analysis.
-For example, in climate data, analysts may initially receive station data containing geographic location information, recorded variables and their recording periods.
-They can then query the temporal variables using the stations of interest to obtain the relevant temporal data.
-Alternatively, analyses may begin as purely spatial or temporal, and analysts may obtain additional temporal or spatial data to expand the result to spatio-temporal.
-
-The function \code{make_cubble()} composes a \code{cubble} object from a spatial table (\code{spatial}) and a temporal table (\code{temporal}), along with three attributes introduced in the subsection \ref{object}: \code{key}, \code{index}, and \code{coords}.
-The following code creates the spatial \code{cubble} object shown previously:
+The function \code{make_cubble()} composes a spatial cubble object from a spatial table (\code{spatial}) and a temporal table (\code{temporal}), along with three attributes introduced in the subsection \ref{object}: \code{key}, \code{index}, and \code{coords}.
+The following code creates the spatial cubble, \code{cb_nested}, from its spatial component, \code{stations} and temporal component \code{meteo}:
 
@@
 \end{CodeInput}
+\begin{CodeOutput}
+# cubble:   key: id [3], index: date, nested form
+# spatial:  [144.8321, -37.98, 145.0964, -37.6655], Missing CRS!
+# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
+  id           long   lat  elev name              wmo_id ts               
+  <chr>       <dbl> <dbl> <dbl> <chr>              <dbl> <list>           
+1 ASN00086038  145. -37.7  78.4 essendon airport   95866 <tibble [10 x 4]>
+2 ASN00086077  145. -38.0  12.1 moorabbin airport  94870 <tibble [10 x 4]>
+3 ASN00086282  145. -37.7 113.  melbourne airport  94866 <tibble [10 x 4]>
+\end{CodeOutput}
 \end{CodeChunk}
@@
 Spatio-temporal data in other foreign objects can be coerced into a \code{cubble} object with the function \code{as_cubble()}.
-This includes casting from a \code{tibble} or \code{data.frame} with both spatial and temporal information, a NetCDF object, a \code{stars} object \citep{stars}, and a \code{sftime} object \citep{sftime}.
-The two examples below show the casting from a tibble and a NetCDF object.
-The dataset \code{climate_flat} combines the spatial data, \code{stations}, with the temporal data, \code{meteo}, into a single tibble object.
-It can be coerced into a cubble using:
+This includes casting from a joined \code{tibble} or \code{data.frame}, a NetCDF object, a \code{stars} object \citep{stars}, and a \code{sftime} object \citep{sftime}.
+In the example below, the spatial cubble object is created from \code{climate_flat}, which combines the spatial data, \code{stations}, with the temporal data, \code{meteo}, into a single tibble object:
 
@@
 It was surprising to us that this was not a station in Tasmania, so for comparison a station in Tasmania is selected on the map to show in relation to Thredbo.
-We can see that Thredbo has a bigger winter dip in temperature, and although Tasmania is cold generally, it's temperatures are more constant
+We can see that Thredbo has a bigger winter dip in temperature, and although Tasmania is cold generally, it's temperatures are more constant.
 

### 2023-06-28 cubble.Rmd
proofread Section 2

@@
 The function \code{make_cubble()} composes a spatial cubble object from a spatial table (\code{spatial}) and a temporal table (\code{temporal}), along with three attributes introduced in the subsection \@ref(object): \code{key}, \code{index}, and \code{coords}.
-The following code creates the spatial cubble, \code{cb_nested}, from its spatial component, \code{stations} and temporal component \code{meteo}:
+The following code creates a spatial cubble from its spatial component, \code{stations} and temporal component \code{meteo}:
 
@@
 
-Spatio-temporal data in other foreign objects can be coerced into a \code{cubble} object with the function \code{as_cubble()}.
-This includes casting from a joined \code{tibble} or \code{data.frame}, a NetCDF object, a \code{stars} object [@stars], and a \code{sftime} object [@sftime].
-In the example below, the spatial cubble object is created from \code{climate_flat}, which combines the spatial data, \code{stations}, with the temporal data, \code{meteo}, into a single tibble object:
+Other foreign spatio-temporal objects in \proglang{R} can be coerced into a \code{cubble} object with the function \code{as_cubble()}.
+This includes a joined \code{tibble} or \code{data.frame}, a NetCDF object, a \code{stars} object [@stars], and a \code{sftime} object [@sftime].
+In the example below, the spatial cubble object is created from \code{climate_flat}, which combines the previous \code{stations} and \code{meteo} into a single tibble object:
 
@@
 Table \@ref(tab:funs) summarises the functions implemented in the \pkg{cubble} package and Table \@ref(tab:methods) details the methods implemented for each of the three cubble classes.
-The \code{cubble_df} class handles methods that behave consistently in both spatial (\code{spatial_cubble_df}) and temporal cubble (\code{temporal_cubble_df}).
-When the method has distinct behavior for the spatial cubble and temporal cubble, it is implemented separately for each subclass.
+The \code{cubble_df} class handles methods that behave consistently in both spatial and temporal cubble.
+When the method has distinct behavior, it is implemented separately in \code{spatial_cubble_df} and \code{temporal_cubble_df}.
 
@@
 | tidyverse | `dplyr_row_slice`, `dplyr_col_modify`, `dplyr_reconstruct`, `select`, `mutate`, `arrange`, `filter`, `group_by`, `ungroup`, `summarise`, `select`, `slice`, `rowwise`, `rename`, `bind_rows`, `bind_cols`, `relocate`, `type_sum`, the slice family (`slice_head`, `slice_tail`, `slice_max`, `slice_min`, `slice_sample`) and the join family (`left_join`, `right_join`, `inner_join`, `full_join`, `anti_join`, `semi_join`) |
-| cubble    | `as_cubble`, `cubble`, `make_cubble`, `check_key` `face_temporal`, `face_spatial`, `unfold`, `key`, `key_vars`, `key_data`, `index`, `index_var`, `coords`, `spatial`, `match_sites`, `match_spatial`, `match_temporal`, `geom_glyph`, `geom_glyph_box`, `geom_glyph_line`, `make_spatial_sf`, `make_temporal_tsibble`, `fill_gaps`, and `scan_gaps`                                                                                |
+| cubble    | `as_cubble`, `cubble`, `make_cubble`, `check_key`, `face_temporal`, `face_spatial`, `unfold`, `key`, `key_vars`, `key_data`, `index`, `index_var`, `coords`, `spatial`, `match_sites`, `match_spatial`, `match_temporal`, `geom_glyph`, `geom_glyph_box`, `geom_glyph_line`, `make_spatial_sf`, `make_temporal_tsibble`, `fill_gaps`, and `scan_gaps`                                                                                |
 
@@
 : (#tab:methods) An overview of the methods implemented in the three \code{cubble} classes.
-Methods are implemented in the \code{cubble\_df} class when they behave consistent across the two subclasses; otherwise, they are implemented separately in each subclass.
+Methods are implemented in the \code{cubble\_df} class when they behave consistent across the spatial and temporal cubble; otherwise, they are implemented separately.
 
-Specifically, the pair of verbs, \code{face_temporal()} and \code{face_spatial()}, pivot the cubble object between the spatial and temporal face of the multivariate spatio-temporal cube, as illustrated in Figure \ref{fig:face}.
-The code below uses the function \code{face_temporal()} and \code{face_spatial()} to transform between a spatial cubble (\code{cb_nested}) and a temporal one (\code{cb_long}), introduced earlier in subsection \@ref(object):
+The pair of cubble verbs, \code{face_temporal()} and \code{face_spatial()}, pivots the cubble object between its two forms, as illustrated in Figure \ref{fig:face}.
+The code below connects the a spatial cubble (\code{cb_nested}) and a temporal temporal (\code{cb_long}) introduced in Section \@ref(object) with \code{face_temporal()} and \code{face_spatial()}:
 
@@
 
-The pair of verbs are exact inverse and apply both functions on a cubble object will result in the object itself:
+Both verbs are the exact inverse of each other and apply both functions on a cubble object will result in the object itself:
 
@@
 
-```{r face, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "An illustration of the function \\code{face\\_temporal()} and \\code{face\\_spatial()}: \\code{face\\_temporal()} converts a spatial cubble (neseted form) into a temporal cubble (long form) to focus on the temporal variables.
-Conversely, \\code{face\\_spatial()} transforms a temporal cubble into a spatial one to emphasize spatial variables."}
+```{r face, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "An illustration of the function \\code{face\\_temporal()} and \\code{face\\_spatial()}: \\code{face\\_temporal()} converts a spatial cubble into a temporal cubbl to focus on the temporal variables.
+Conversely, \\code{face\\_spatial()} transforms a temporal cubble into a spatial one to for focus on the spatial variables."}
 knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.001.png"))
@@
 
-Analysts often have their preferred spatial or temporal data structure that they prefer to use for spatio-temporal analysis.
-For example, the \code{tbl_ts} class from the \code{tsibble} package [@tsibble] is commonly used in time series forecasting and similarly, the \code{sf} class [@sf] is frequently used in spatial data science.
-In \code{cubble}, analysts have the flexibility to combine these two structures together by allowing the spatial component to also be an \code{sf} object and the temporal component to also be a \code{tsibble} object.
+Analysts often have their preferred spatial or temporal data structure for spatial or temporal analysis, which they may wish to continue using for spatio-temporal analysis.
+With \code{cubble}, analysts can incorporate the \code{tsibble} class in a temporal cubble and the \code{sf} class in a spatial cubble.
 
@@
 \code{meteo_ts} in the example below.
-The \code{tsibble} class from the input will be carried over to the \code{cubble} object:
+The \code{tsibble} class (\code{tbl_ts}) from the input will be carried over to the temporal cubble, indicated by the \code{[tsibble]} in the header and in the object class:
 
@@
 
-The temporal cubble shows \code{[tsibble]} in the header to indicate the object also being in a \code{tbl_ts} class (\code{tsibble}).
 Methods applies to the \code{tbl_ts} class can also be applied to the temporal cubble objects, for example, checking whether the data contain temporal gaps:
@@
 
-An existing \code{cubble} object can promote its temporal component to a \code{tsibble} object by applying \code{make_temporal_tsibble()}.
-The following code illustrates this with the object \code{cb_long} created in section \@ref(create) and the promoted cubble object is equal to the cubble object originally created from a \code{tsibble} object:
+A created temporal cubble can promote its temporal component to a \code{tsibble} object using \code{make_temporal_tsibble()}.
+See the code example below using the \code{cb_long} object created in Section \@ref(create):
 
 ```{r echo = TRUE}
-ts_long2 <- cb_long |> make_temporal_tsibble() 
-identical(ts_long2, ts_long)
+cb_long |> make_temporal_tsibble() 
 ```
@@
 
-Similarly, an \code{sf} object can be supplied as the spatial component to create a \code{cubble} object, with the \code{coords} argument being omitted.
-This opens up the possibility to represent fixed area with polygons or multipolygons (see Applications \@ref(covid)) and the \code{coords} argument will be calculated as the centroids of the (multi)polygons.
-The \code{[sf]} print in the cubble header suggest an spatial component being also a \code{sf} object:
+Similarly, the spatial component  of a cubble object can be an \code{sf} object and if the \code{coords} argument is omitted, it will be calculated from the sf geometry.
+The sf status is signalled by the \code{[sf]} label in the cubble header:
 
@@
 
-The \code{sf} functions applicable to a \code{cubble} object have been listed in Table \@ref(tab:funs) and the following code shows how to perform coordinate transformation with \code{st_transform} on a cubble object:
+This allows applying functions from the \code{sf} package to a cubble object, for example, to handle coordinate transformation with \code{st_transform}:
 
@@
 
-The counterpart to promote the spatial component in an existing \code{cubble} to be an \code{sf} object is \code{make_spatial_sf()}:
+The spatial component of a created \code{cubble} can also be promoted into an \code{sf} object with \code{make_spatial_sf()}:
 

### 2023-06-28 cubble.tex
proofread Section 2

@@
 The function \code{make_cubble()} composes a spatial cubble object from a spatial table (\code{spatial}) and a temporal table (\code{temporal}), along with three attributes introduced in the subsection \ref{object}: \code{key}, \code{index}, and \code{coords}.
-The following code creates the spatial cubble, \code{cb_nested}, from its spatial component, \code{stations} and temporal component \code{meteo}:
+The following code creates a spatial cubble from its spatial component, \code{stations} and temporal component \code{meteo}:
 
@@
 
-Spatio-temporal data in other foreign objects can be coerced into a \code{cubble} object with the function \code{as_cubble()}.
-This includes casting from a joined \code{tibble} or \code{data.frame}, a NetCDF object, a \code{stars} object \citep{stars}, and a \code{sftime} object \citep{sftime}.
-In the example below, the spatial cubble object is created from \code{climate_flat}, which combines the spatial data, \code{stations}, with the temporal data, \code{meteo}, into a single tibble object:
+Other foreign spatio-temporal objects in \proglang{R} can be coerced into a \code{cubble} object with the function \code{as_cubble()}.
+This includes a joined \code{tibble} or \code{data.frame}, a NetCDF object, a \code{stars} object \citep{stars}, and a \code{sftime} object \citep{sftime}.
+In the example below, the spatial cubble object is created from \code{climate_flat}, which combines the previous \code{stations} and \code{meteo} into a single tibble object:
 
@@
 Table \ref{tab:funs} summarises the functions implemented in the \pkg{cubble} package and Table \ref{tab:methods} details the methods implemented for each of the three cubble classes.
-The \code{cubble_df} class handles methods that behave consistently in both spatial (\code{spatial_cubble_df}) and temporal cubble (\code{temporal_cubble_df}).
-When the method has distinct behavior for the spatial cubble and temporal cubble, it is implemented separately for each subclass.
+The \code{cubble_df} class handles methods that behave consistently in both spatial and temporal cubble.
+When the method has distinct behavior, it is implemented separately in \code{spatial_cubble_df} and \code{temporal_cubble_df}.
 
@@
 tidyverse & \texttt{dplyr\_row\_slice}, \texttt{dplyr\_col\_modify}, \texttt{dplyr\_reconstruct}, \texttt{select}, \texttt{mutate}, \texttt{arrange}, \texttt{filter}, \texttt{group\_by}, \texttt{ungroup}, \texttt{summarise}, \texttt{select}, \texttt{slice}, \texttt{rowwise}, \texttt{rename}, \texttt{bind\_rows}, \texttt{bind\_cols}, \texttt{relocate}, \texttt{type\_sum}, the slice family (\texttt{slice\_head}, \texttt{slice\_tail}, \texttt{slice\_max}, \texttt{slice\_min}, \texttt{slice\_sample}) and the join family (\texttt{left\_join}, \texttt{right\_join}, \texttt{inner\_join}, \texttt{full\_join}, \texttt{anti\_join}, \texttt{semi\_join}) \\
-cubble & \texttt{as\_cubble}, \texttt{cubble}, \texttt{make\_cubble}, \texttt{check\_key} \texttt{face\_temporal}, \texttt{face\_spatial}, \texttt{unfold}, \texttt{key}, \texttt{key\_vars}, \texttt{key\_data}, \texttt{index}, \texttt{index\_var}, \texttt{coords}, \texttt{spatial}, \texttt{match\_sites}, \texttt{match\_spatial}, \texttt{match\_temporal}, \texttt{geom\_glyph}, \texttt{geom\_glyph\_box}, \texttt{geom\_glyph\_line}, \texttt{make\_spatial\_sf}, \texttt{make\_temporal\_tsibble}, \texttt{fill\_gaps}, and \texttt{scan\_gaps} \\
+cubble & \texttt{as\_cubble}, \texttt{cubble}, \texttt{make\_cubble}, \texttt{check\_key}, \texttt{face\_temporal}, \texttt{face\_spatial}, \texttt{unfold}, \texttt{key}, \texttt{key\_vars}, \texttt{key\_data}, \texttt{index}, \texttt{index\_var}, \texttt{coords}, \texttt{spatial}, \texttt{match\_sites}, \texttt{match\_spatial}, \texttt{match\_temporal}, \texttt{geom\_glyph}, \texttt{geom\_glyph\_box}, \texttt{geom\_glyph\_line}, \texttt{make\_spatial\_sf}, \texttt{make\_temporal\_tsibble}, \texttt{fill\_gaps}, and \texttt{scan\_gaps} \\
 \bottomrule()
@@
 \caption{\label{tab:methods} An overview of the methods implemented in the three \code{cubble} classes.
-Methods are implemented in the \code{cubble\_df} class when they behave consistent across the two subclasses; otherwise, they are implemented separately in each subclass.}\tabularnewline
+Methods are implemented in the \code{cubble\_df} class when they behave consistent across the spatial and temporal cubble; otherwise, they are implemented separately.}\tabularnewline
 \toprule()
@@
 
-Specifically, the pair of verbs, \code{face_temporal()} and \code{face_spatial()}, pivot the cubble object between the spatial and temporal face of the multivariate spatio-temporal cube, as illustrated in Figure \ref{fig:face}.
-The code below uses the function \code{face_temporal()} and \code{face_spatial()} to transform between a spatial cubble (\code{cb_nested}) and a temporal one (\code{cb_long}), introduced earlier in subsection \ref{object}:
+The pair of cubble verbs, \code{face_temporal()} and \code{face_spatial()}, pivots the cubble object between its two forms, as illustrated in Figure \ref{fig:face}.
+The code below connects the a spatial cubble (\code{cb_nested}) and a temporal temporal (\code{cb_long}) introduced in Section \ref{object} with \code{face_temporal()} and \code{face_spatial()}:
 
@@
 
-\caption{An illustration of the function \code{face\_temporal()} and \code{face\_spatial()}: \code{face\_temporal()} converts a spatial cubble (neseted form) into a temporal cubble (long form) to focus on the temporal variables.
-Conversely, \code{face\_spatial()} transforms a temporal cubble into a spatial one to emphasize spatial variables.}\label{fig:face}
+\caption{An illustration of the function \code{face\_temporal()} and \code{face\_spatial()}: \code{face\_temporal()} converts a spatial cubble into a temporal cubbl to focus on the temporal variables.
+Conversely, \code{face\_spatial()} transforms a temporal cubble into a spatial one to for focus on the spatial variables.}\label{fig:face}
 \end{figure}
@@
 
-Analysts often have their preferred spatial or temporal data structure that they prefer to use for spatio-temporal analysis.
-For example, the \code{tbl_ts} class from the \code{tsibble} package \citep{tsibble} is commonly used in time series forecasting and similarly, the \code{sf} class \citep{sf} is frequently used in spatial data science.
-In \code{cubble}, analysts have the flexibility to combine these two structures together by allowing the spatial component to also be an \code{sf} object and the temporal component to also be a \code{tsibble} object.
+Analysts often have their preferred spatial or temporal data structure for spatial or temporal analysis, which they may wish to continue using for spatio-temporal analysis.
+With \code{cubble}, analysts can incorporate the \code{tsibble} class in a temporal cubble and the \code{sf} class in a spatial cubble.
 
@@
 The \code{key} and \code{index} arguments in a \code{cubble} object corresponds to the \code{tsibble} counterparts and they can be safely omitted, if the temporal component is a \code{tsibble} object, i.e.~\code{meteo_ts} in the example below.
-The \code{tsibble} class from the input will be carried over to the \code{cubble} object:
+The \code{tsibble} class (\code{tbl_ts}) from the input will be carried over to the temporal cubble, indicated by the \code{[tsibble]} in the header and in the object class:
 
@@
 
-An existing \code{cubble} object can promote its temporal component to a \code{tsibble} object by applying \code{make_temporal_tsibble()}.
-The following code illustrates this with the object \code{cb_long} created in section \ref{create} and the promoted cubble object is equal to the cubble object originally created from a \code{tsibble} object:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> ts_long2 <- cb_long |> make_temporal_tsibble() 
-R> identical(ts_long2, ts_long)
-\end{CodeInput}
-\begin{CodeOutput}
-[1] TRUE
+A created temporal cubble can promote its temporal component to a \code{tsibble} object using \code{make_temporal_tsibble()}.
+See the code example below using the \code{cb_long} object created in Section \ref{create}:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> cb_long |> make_temporal_tsibble() 
+\end{CodeInput}
+\begin{CodeOutput}
+# cubble:   key: id [3], index: date, long form, [tsibble]
+# temporal: 2020-01-01 -- 2020-01-10 [1D], no gaps
+# spatial:  long [dbl], lat [dbl], elev [dbl], name [chr], wmo_id [dbl]
+  id          date        prcp  tmax  tmin
+  <chr>       <date>     <dbl> <dbl> <dbl>
+1 ASN00086038 2020-01-01     0  26.8  11  
+2 ASN00086038 2020-01-02     0  26.3  12.2
+3 ASN00086038 2020-01-03     0  34.5  12.7
+4 ASN00086038 2020-01-04     0  29.3  18.8
+5 ASN00086038 2020-01-05    18  16.1  12.5
+# i 25 more rows
 \end{CodeOutput}
@@
 
-Similarly, an \code{sf} object can be supplied as the spatial component to create a \code{cubble} object, with the \code{coords} argument being omitted.
-This opens up the possibility to represent fixed area with polygons or multipolygons (see Applications \ref{covid}) and the \code{coords} argument will be calculated as the centroids of the (multi)polygons.
-The \code{[sf]} print in the cubble header suggest an spatial component being also a \code{sf} object:
+Similarly, the spatial component of a cubble object can be an \code{sf} object and if the \code{coords} argument is omitted, it will be calculated from the sf geometry.
+The sf status is signalled by the \code{[sf]} label in the cubble header:
 
@@
 
-The \code{sf} functions applicable to a \code{cubble} object have been listed in Table \ref{tab:funs} and the following code shows how to perform coordinate transformation with \code{st_transform} on a cubble object:
+This allows applying functions from the \code{sf} package to a cubble object, for example, to handle coordinate transformation with \code{st_transform}:
 

### 2023-06-28 cubble.Rmd
deal with large figures and reference

@@
 library("dplyr")
+library("tidyr")
 library("ggplot2")
@@
 Analysts often have their preferred spatial or temporal data structure for spatial or temporal analysis, which they may wish to continue using for spatio-temporal analysis.
-With \code{cubble}, analysts can incorporate the \code{tsibble} class in a temporal cubble and the \code{sf} class in a spatial cubble.
+With \code{cubble}, analysts can incorporate the \code{tsibble} class [@tsibble] in a temporal cubble and the \code{sf} class [@sf] in a spatial cubble.
 
@@
 
-The package `GGally` initially implement the glyph map.
-It uses `glyphs()` to calculate the axis transformation and then uses `geom_polygon()` to draw the map.
+The package `GGally` [@ggally] initially implements the glyph map, using `glyphs()` to calculate the axis transformation and then `geom_polygon()` to draw the map.
 In cubble, a ggproto implementation `geom_glyph()` performs the linear algebra internally as data transformation .
@@
 The ERA5 reanalysis [@hersbach2020era5] provides hourly estimates of atmospheric, land and oceanic climate variables on a global scale.
-The hourly pressure level data in the NetCDF format can be downloaded from Copernicus Climate Data Store (CDS) [@cds] or via the \pkg{ecmwfr} package [@ecwmfr].
+The hourly pressure level data in the NetCDF format can be downloaded from Copernicus Climate Data Store (CDS) or via the \pkg{ecmwfr} package [@ecwmfr].
 This example reproduces the row created from ERA5 reanalysis data shown in Figure 19 by @hersbach2020era5.
@@
 
-To work with NetCDF data in R, several packages are available, including \code{ncdf4}, \code{RNetCDF}, and \code{tidync}.
-The following code converts a NetCDF object of class \code{ncdf4} into a cubble object:
+ The following code converts a NetCDF object of class \code{ncdf4} [@ncdf4] into a cubble object:
 
@@
 ```{r echo = TRUE}
-(dt <- as_cubble(raw, vars = c("q", "z"),
+(dt <- as_cubble(
+  raw, vars = c("q", "z"),
   long_range = seq(-180, 180, 1), lat_range = seq(-88, -15, 1)))
@@
 ```{r netcdf, out.width="100%", fig.height = 4, fig.width = 10, fig.cap = "A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020) to illustrate the break-up of sourthern polar vortex in late September and early October 2002.
-The polar vortex, signalled by the high specific humidity, splits into two on 2002-09-26 and further splits into four on 2002-10-04."}
+The polar vortex, signalled by the high specific humidity, splits into two on 2002-09-26 and further splits into four on 2002-10-04.", dev = "png"}
 date <- c("2002-09-22", "2002-09-26", "2002-09-30", "2002-10-04") |> as.Date()

### 2023-06-28 cubble.tex
deal with large figures and reference

@@
 Analysts often have their preferred spatial or temporal data structure for spatial or temporal analysis, which they may wish to continue using for spatio-temporal analysis.
-With \code{cubble}, analysts can incorporate the \code{tsibble} class in a temporal cubble and the \code{sf} class in a spatial cubble.
+With \code{cubble}, analysts can incorporate the \code{tsibble} class \citep{tsibble} in a temporal cubble and the \code{sf} class \citep{sf} in a spatial cubble.
 
@@
 
-The package \texttt{GGally} initially implement the glyph map.
-It uses \texttt{glyphs()} to calculate the axis transformation and then uses \texttt{geom\_polygon()} to draw the map.
+The package \texttt{GGally} \citep{ggally} initially implements the glyph map, using \texttt{glyphs()} to calculate the axis transformation and then \texttt{geom\_polygon()} to draw the map.
 In cubble, a ggproto implementation \texttt{geom\_glyph()} performs the linear algebra internally as data transformation .
@@
 The ERA5 reanalysis \citep{hersbach2020era5} provides hourly estimates of atmospheric, land and oceanic climate variables on a global scale.
-The hourly pressure level data in the NetCDF format can be downloaded from Copernicus Climate Data Store (CDS) \citep{cds} or via the \pkg{ecmwfr} package \citep{ecwmfr}.
+The hourly pressure level data in the NetCDF format can be downloaded from Copernicus Climate Data Store (CDS) or via the \pkg{ecmwfr} package \citep{ecwmfr}.
 This example reproduces the row created from ERA5 reanalysis data shown in Figure 19 by \citet{hersbach2020era5}.
@@
 
-To work with NetCDF data in R, several packages are available, including \code{ncdf4}, \code{RNetCDF}, and \code{tidync}.
-The following code converts a NetCDF object of class \code{ncdf4} into a cubble object:
+The following code converts a NetCDF object of class \code{ncdf4} \citep{ncdf4} into a cubble object:
 
@@
 \begin{CodeInput}
-R> (dt <- as_cubble(raw, vars = c("q", "z"),
+R> (dt <- as_cubble(
++   raw, vars = c("q", "z"),
 +   long_range = seq(-180, 180, 1), lat_range = seq(-88, -15, 1)))

### 2023-06-28 cubble.Rmd
proofread Section 3

@@
 
-One common task when working with spatio-temporal data is to match nearby sites.
-For example, we may want to verify the location of an old list of stations with current stations, or we may want to match the data from different data sources.
-In cubble, spatial and temporal matching are performed using the functions `match_spatial()` and `match_temporal()`.
-The `match_spatial()` function calculates the spatial distance between observations in two cubble objects.
-Various distance measures are available (check `sf::st_distance`).
-Analysts can specify the number of matched groups to output using the `spatial_n_group` argument (default to 4 groups) and the number of matches per group using the `spatial_n_group` argument (default to 1, one-to-one matching).
-The syntax to use `match_spatial()` is:
+Matching time series from an old list of stations to a new list is a common task in spatio-temporal data analysis.
+In cubble, matching based on distance and time series feature can be performed using the functions `match_spatial()` and `match_temporal()`.
+The `match_spatial()` function finds the matched pairs in two cubble objects based on distance between sites:
 
@@
 
-The function `match_temporal()` calculates the similarity between time series within spatially matched groups.
-Two identifiers are required: one for separating each matched group (`match_id`) and one for separating the two data sources (`data_id`).
-The argument `temporal_by` uses the `by` syntax from dplyr's `*_join` to specify the temporal variables to match.
+Two arguments are available to control the outputs: the argument `spatial_n_group` specifies the number of paired groups to output and the argument `spatial_n_each` specifies the number of each for each item in the first cubble object (default to 1 for one-to-one matching).
 
-The similarity score between two time series is calculated using a matching function, which can be customised by the analysts.
-The matching function takes two time series as a list and returns a single numerical score.
-This allows for flexibility in using existing time series distance calculation implementation.
-By default, cubble implements a simple peak matching algorithm (`match_peak`) that counts the number of peaks in two time series that fall within a specified temporal window.
-The syntax to use `match_temporal()` is
+The function `match_temporal()` takes the outputs from spatial matching and calculates a similarity score of the time series between spatially matched pairs.
+The temporal matching requires two identifiers: one for separating each spatially matched group: `match_id` and one for separating the two data sources: `data_id`.
+Matching between different variables can be specified using the `temporal_by` argument, similar to the `by` syntax from dplyr's `*_join`.
 
@@
 match_temporal(
-  <cubble_obj_from_match_spatial>, 
-  data_id = , match_id = , 
+  <obj_from_match_spatial>, 
+  data_id = ... , match_id = ..., 
   temporal_by = c("..." = "...")
@@
 
+The similarity score between two time series is calculated using a matching function, which can be customised by the analysts based on the time series feature relevant to match.
+The matching function takes two time series as a list and returns a single numerical score.
+By default, cubble uses a simple peak matching algorithm (`match_peak`) to count the number of peaks in two time series that fall within a specified temporal window.
+
 ## Interactive graphics
 
-The workflow with the \code{cubble} class works well with an interactive graphics pipeline (e.g., @buja1988elements; @buja1996interactive;  @sutherland2000orca; @xie2014reactive; @cheng2016enabling) that is available in \proglang{R} with the package \pkg{crosstalk} [@crosstalk].
-Figure \ref{fig:illu-interactive} illustrates the linking between a map and multiple time series in a \code{cubble} object.
-Both the map, generated from the spatial cubble, and the time series plot, generated from the temporal cubble, are shared crosstalk objects.
+The cubble workflow works well with an interactive graphics pipeline (e.g., @buja1988elements; @buja1996interactive;  @sutherland2000orca; @xie2014reactive; @cheng2016enabling).
+This section describes the linking between a map and multiple time series in a \code{cubble} object using the package \pkg{crosstalk} [@crosstalk], as illustrated in Figure \ref{fig:illu-interactive}.
+The spatial and temporal cubble can be constructed into a shared crosstalk object to create linking between a map and a time series plot.
 When a user selects a location on the map (a), the corresponding site is highlighted.
-This selection activates a row in the spatial \code{cubble}, which is then communicated to the temporal \code{cubble}, resulting in the selection of all observations with the same ID in (b).
-Consequently the temporal \code{cubble} highlights the corresponding series in the time series plot (c).
-Linking can also be initiated from the time series plot, by selecting points on the time series.
-This action activates rows with the same ID in the temporal \code{cubble} and the row in the spatial \code{cubble}, which will be highlighted on the map.
-Please refer to the appendix for the corresponding illustration.
+This selection activates a row in the spatial cubble, which is then communicated to the temporal cubble, resulting in the selection of all observations with the same ID in (b).
+Consequently the temporal cubble highlights the corresponding series in the time series plot (c).
+Linking can also be initiated from the time series plot, by selecting points on the time series, which activates rows with the same ID in the temporal cubble and the corresponding row in the spatial cubble to highlight on the map.
 
@@
 This will link to all the rows with the same id in the temporal \\code{cubble} (b) and update the line plot (c)."}
-knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.002.png"))
+knitr::include_graphics(here::here("figures/diagram-keynotes/‎diagram-keynotes.‎002.png")) 
 ```
@@
 
-Sometimes, we wish to communicate spatial and temporal information collectively through visualisation.
-This can be achieved through several graphical displays: one can make faceted maps across time, creating map animations, or constructing interactive graphics to link between map and time series plot.
-The concept of glyph maps was initially proposed in @Wickham2012-yr.
-The underlying idea is to transform the temporal coordinates into spatial coordinates so that time series plot can be displayed on the map.
-
-The package `GGally` [@ggally] initially implements the glyph map, using `glyphs()` to calculate the axis transformation and then `geom_polygon()` to draw the map.
-In cubble, a ggproto implementation `geom_glyph()` performs the linear algebra internally as data transformation .
-The `geom_glyph()` requires four aesthetics: `x_major`, `y_major`, `x_minor`, and `y_minor`.
-The major axes are the outer spatial coordinates and the minor axes are the inner/ temporal coordinates:
+To communicate spatial and temporal information collectively in display,  a few visualisation options are available: one can make faceted maps across time, creating map animations, or constructing interactive graphics to link between map and time series plot.
+Faceted maps and animation, while commonly used, may not be helpful for comparing across time since eyes need to jump across multiple facets or frames to locate the space before preceiving the changes in time.
+The glyph map [@Wickham2012-yr] resolves this issue by imposing the time series onto the map as a glyph through coordinate transformation.
+The transformation uses linear algebra to convert the temporal coordinates (minor coordinates) into the spatial coordinates (major coordinates) and is implemented in the package `GGally` [@ggally] package.
+The \pkg{cubble} package provides a ggproto implementation, `geom_glyph()`, for glyph maps and it takes four required aesthetics: `x_major`, `y_major`, `x_minor`, and `y_minor`:
 
@@
 
-Some useful controls over the glyph map includes:
+Other useful controls over the glyph map includes:
 
@@
   * adjust glyph size with arguments \code{width} and \code{height},
-  * glyph scale to fixed (\code{global_rescale} defaults to \code{TRUE}) or free, and 
+  * transformation relative to the all series (\code{global_rescale} defaults to \code{TRUE}) or each single series, and 
   * reference boxes and lines with \code{geom_glyph_box()} and \code{geom_glyph_line()}.

### 2023-06-28 cubble.tex
proofread Section 3

@@
 
-One common task when working with spatio-temporal data is to match nearby sites.
-For example, we may want to verify the location of an old list of stations with current stations, or we may want to match the data from different data sources.
-In cubble, spatial and temporal matching are performed using the functions \texttt{match\_spatial()} and \texttt{match\_temporal()}.
-The \texttt{match\_spatial()} function calculates the spatial distance between observations in two cubble objects.
-Various distance measures are available (check \texttt{sf::st\_distance}).
-Analysts can specify the number of matched groups to output using the \texttt{spatial\_n\_group} argument (default to 4 groups) and the number of matches per group using the \texttt{spatial\_n\_group} argument (default to 1, one-to-one matching).
-The syntax to use \texttt{match\_spatial()} is:
+Matching time series from an old list of stations to a new list is a common task in spatio-temporal data analysis.
+In cubble, matching based on distance and time series feature can be performed using the functions \texttt{match\_spatial()} and \texttt{match\_temporal()}.
+The \texttt{match\_spatial()} function finds the matched pairs in two cubble objects based on distance between sites:
 
@@
 
-The function \texttt{match\_temporal()} calculates the similarity between time series within spatially matched groups.
-Two identifiers are required: one for separating each matched group (\texttt{match\_id}) and one for separating the two data sources (\texttt{data\_id}).
-The argument \texttt{temporal\_by} uses the \texttt{by} syntax from dplyr's \texttt{*\_join} to specify the temporal variables to match.
-
-The similarity score between two time series is calculated using a matching function, which can be customised by the analysts.
-The matching function takes two time series as a list and returns a single numerical score.
-This allows for flexibility in using existing time series distance calculation implementation.
-By default, cubble implements a simple peak matching algorithm (\texttt{match\_peak}) that counts the number of peaks in two time series that fall within a specified temporal window.
-The syntax to use \texttt{match\_temporal()} is
+Two arguments are available to control the outputs: the argument \texttt{spatial\_n\_group} specifies the number of paired groups to output and the argument \texttt{spatial\_n\_each} specifies the number of each for each item in the first cubble object (default to 1 for one-to-one matching).
+
+The function \texttt{match\_temporal()} takes the outputs from spatial matching and calculates a similarity score of the time series between spatially matched pairs.
+The temporal matching requires two identifiers: one for separating each spatially matched group: \texttt{match\_id} and one for separating the two data sources: \texttt{data\_id}.
+Matching between different variables can be specified using the \texttt{temporal\_by} argument, similar to the \texttt{by} syntax from dplyr's \texttt{*\_join}.
 
@@
 
+The similarity score between two time series is calculated using a matching function, which can be customised by the analysts based on the time series feature relevant to match.
+The matching function takes two time series as a list and returns a single numerical score.
+By default, cubble uses a simple peak matching algorithm (\texttt{match\_peak}) to count the number of peaks in two time series that fall within a specified temporal window.
+
 \hypertarget{interactive-graphics}{%
@@
 
-The workflow with the \code{cubble} class works well with an interactive graphics pipeline (e.g., \citet{buja1988elements}; \citet{buja1996interactive}; \citet{sutherland2000orca}; \citet{xie2014reactive}; \citet{cheng2016enabling}) that is available in \proglang{R} with the package \pkg{crosstalk} \citep{crosstalk}.
-Figure \ref{fig:illu-interactive} illustrates the linking between a map and multiple time series in a \code{cubble} object.
-Both the map, generated from the spatial cubble, and the time series plot, generated from the temporal cubble, are shared crosstalk objects.
+The cubble workflow works well with an interactive graphics pipeline (e.g., \citet{buja1988elements}; \citet{buja1996interactive}; \citet{sutherland2000orca}; \citet{xie2014reactive}; \citet{cheng2016enabling}).
+This section describes the linking between a map and multiple time series in a \code{cubble} object using the package \pkg{crosstalk} \citep{crosstalk}, as illustrated in Figure \ref{fig:illu-interactive}.
+The spatial and temporal cubble can be constructed into a shared crosstalk object to create linking between a map and a time series plot.
 When a user selects a location on the map (a), the corresponding site is highlighted.
-This selection activates a row in the spatial \code{cubble}, which is then communicated to the temporal \code{cubble}, resulting in the selection of all observations with the same ID in (b).
-Consequently the temporal \code{cubble} highlights the corresponding series in the time series plot (c).
-Linking can also be initiated from the time series plot, by selecting points on the time series.
-This action activates rows with the same ID in the temporal \code{cubble} and the row in the spatial \code{cubble}, which will be highlighted on the map.
-Please refer to the appendix for the corresponding illustration.
+This selection activates a row in the spatial cubble, which is then communicated to the temporal cubble, resulting in the selection of all observations with the same ID in (b).
+Consequently the temporal cubble highlights the corresponding series in the time series plot (c).
+Linking can also be initiated from the time series plot, by selecting points on the time series, which activates rows with the same ID in the temporal cubble and the corresponding row in the spatial cubble to highlight on the map.
 
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.35\textheight]{figures/diagram-keynotes/diagram-keynotes.002}
+{\centering \includegraphics[width=1\linewidth,height=0.35\textheight]{figures/diagram-keynotes/‎diagram-keynotes.‎002}
 
@@
 
-Sometimes, we wish to communicate spatial and temporal information collectively through visualisation.
-This can be achieved through several graphical displays: one can make faceted maps across time, creating map animations, or constructing interactive graphics to link between map and time series plot.
-The concept of glyph maps was initially proposed in \citet{Wickham2012-yr}.
-The underlying idea is to transform the temporal coordinates into spatial coordinates so that time series plot can be displayed on the map.
-
-The package \texttt{GGally} \citep{ggally} initially implements the glyph map, using \texttt{glyphs()} to calculate the axis transformation and then \texttt{geom\_polygon()} to draw the map.
-In cubble, a ggproto implementation \texttt{geom\_glyph()} performs the linear algebra internally as data transformation .
-The \texttt{geom\_glyph()} requires four aesthetics: \texttt{x\_major}, \texttt{y\_major}, \texttt{x\_minor}, and \texttt{y\_minor}.
-The major axes are the outer spatial coordinates and the minor axes are the inner/ temporal coordinates:
+To communicate spatial and temporal information collectively in display, a few visualisation options are available: one can make faceted maps across time, creating map animations, or constructing interactive graphics to link between map and time series plot.
+Faceted maps and animation, while commonly used, may not be helpful for comparing across time since eyes need to jump across multiple facets or frames to locate the space before preceiving the changes in time.
+The glyph map \citep{Wickham2012-yr} resolves this issue by imposing the time series onto the map as a glyph through coordinate transformation.
+The transformation uses linear algebra to convert the temporal coordinates (minor coordinates) into the spatial coordinates (major coordinates) and is implemented in the package \texttt{GGally} \citep{ggally} package.
+The \pkg{cubble} package provides a ggproto implementation, \texttt{geom\_glyph()}, for glyph maps and it takes four required aesthetics: \texttt{x\_major}, \texttt{y\_major}, \texttt{x\_minor}, and \texttt{y\_minor}:
 
@@
 \item
-  glyph scale to fixed (\code{global_rescale} defaults to \code{TRUE}) or free, and
+  transformation relative to the all series (\code{global_rescale} defaults to \code{TRUE}) or each single series, and
 \item

### 2023-06-28 cubble.Rmd
before section 4.3

@@
 This data can be combined with map polygon data, available from the Australian Bureau of Statistics (ABS), to visualize COVID incidence and spread.
-In the \pkg{cubble} package, the COVID count data (\code{covid}) and the LGA information (\code{lga}) are available as a \code{tsibble} object and an \code{sf} object respectively.
-
-A cubble object can be created from separate spatial and temporal component using the \code{make_cubble()} function, introduced in Section \@ref(create). The \code{by} argument
-is used to specify the joining variable from the two component using the \code{*_join()} \code{by} syntax:
+In the \pkg{cubble} package, the COVID count data (\code{covid}) and the LGA information (\code{lga}) are available as a \code{tsibble} object and an \code{sf} object, respectively.
+A cubble object can be created from these two component using \code{make_cubble()}.
+The \code{id} column containing LGA information has different names in the two data and can be specified with using the \code{by} argument:
 
@@
 
-The warning message suggests the slight difference of LGA encoding used by Victoria government and ABS and prompts analysts to use the function \code{check_key()} to identify the mismatches:
+The warning message suggests the slight difference of LGA encoding used by Victoria government and ABS and prompts analysts to check the mismatch using \code{check_key()}:
 
@@
 Here, the main mismatch arises from the two LGAs: Kingston and Latrobe (Kingston is a LGA in both Victoria and South Australia and Latrobe is a LGA in both Victoria and Tasmania).
-Analysts can modify the input spatial and temporal data accordingly and recreate the cubble object:
+Analysts can then reconcile the spatial and temporal data based on this check summary and recreate the cubble object:
 
@@
 The Global Historical Climatology Network (GHCN) provides daily climate measures for stations worldwide.
-In the \pkg{cubble} package, the dataset \code{historical_tmax} contains daily maximum temperature data for `r nrow(historical_tmax)` stations in Australia, covering two periods: 1971-1975 and 2016-2020.
-The objective of this example is to compare the changes in maximum temperature between these two periods using a glyph map.
+In the \pkg{cubble} package, the cubble object \code{historical_tmax} contains daily maximum temperature data for `r nrow(historical_tmax)` stations in Australia, covering two periods: 1971-1975 and 2016-2020.
+This example uses glyph maps to compare the changes in temperature between these two periods.
 
-To prevent overlapping of weather stations on the map, we subset the stations based on minimum distance of 50km between them.
-This can be done by first promoting the spatial cubble to also be an sf object with \code{make_spatial_sf()}, calculating the distance matrix using the sf function \code{st_distance()}, and filtering the stations with dplyr's \code{filter()}:
+To prevent overlapping of weather stations on the map, stations are selected to ensure a minimum distance of 50km.
+Distance between stations can be calcualted with \code{sf::st_distance()} after turning the spatial cubble to also be an sf object with \code{make_spatial_sf()}:
 
@@
 
-Next, daily maximum temperature is averaged into monthly series for each periods in the temporal cubble.
-The last step with \code{unfold()} moves the two coordinate columns \code{long, lat} into the temporal cubble, preparing the data for the glyph map:
+Daily maximum temperature is then averaged into monthly series for each periods in the temporal cubble.
+The last step with \code{unfold()} moves the two coordinate columns (\code{long, lat}) into the temporal cubble, preparing the data for the glyph map:
 
@@
 
-The following code creates the glyph map (a) in Figure \ref{fig:glyphmap} with additional ggplot2 code for highlighting the single station, Cobar and styling:
+The following code creates the glyph map (a) in Figure \ref{fig:glyphmap} (additional codes are needed for highlighting the single station, Cobar and styling) and the glyph map (c) is produced similarly with the difference series, rather than the two original series, being plotted.
 
@@
 
-One common task when working with spatio-temporal data is to match nearby sites.
-For example, we may want to verify the location of an old list of stations with current stations, or we may want to match the data from different data sources.
 In this example, we will introduce the spatial and temporal matching in cubble using an example on matching river level data with precipitation in Victoria, Australia.

### 2023-06-28 cubble.tex
before section 4.3

@@
 This data can be combined with map polygon data, available from the Australian Bureau of Statistics (ABS), to visualize COVID incidence and spread.
-In the \pkg{cubble} package, the COVID count data (\code{covid}) and the LGA information (\code{lga}) are available as a \code{tsibble} object and an \code{sf} object respectively.
-
-A cubble object can be created from separate spatial and temporal component using the \code{make_cubble()} function, introduced in Section \ref{create}. The \code{by} argument
-is used to specify the joining variable from the two component using the \code{*_join()} \code{by} syntax:
+In the \pkg{cubble} package, the COVID count data (\code{covid}) and the LGA information (\code{lga}) are available as a \code{tsibble} object and an \code{sf} object, respectively.
+A cubble object can be created from these two component using \code{make_cubble()}.
+The \code{id} column containing LGA information has different names in the two data and can be specified with using the \code{by} argument:
 
@@
 The Global Historical Climatology Network (GHCN) provides daily climate measures for stations worldwide.
-In the \pkg{cubble} package, the dataset \code{historical_tmax} contains daily maximum temperature data for 75 stations in Australia, covering two periods: 1971-1975 and 2016-2020.
-The objective of this example is to compare the changes in maximum temperature between these two periods using a glyph map.
-
-To prevent overlapping of weather stations on the map, we subset the stations based on minimum distance of 50km between them.
-This can be done by first promoting the spatial cubble to also be an sf object with \code{make_spatial_sf()}, calculating the distance matrix using the sf function \code{st_distance()}, and filtering the stations with dplyr's \code{filter()}:
+In the \pkg{cubble} package, the cubble object \code{historical_tmax} contains daily maximum temperature data for 75 stations in Australia, covering two periods: 1971-1975 and 2016-2020.
+This example uses glyph maps to compare the changes in temperature between these two periods.
+
+To prevent overlapping of weather stations on the map, stations are selected to ensure a minimum distance of 50km.
+Distance between stations can be calcualted with \code{sf::st_distance()} after turning the spatial cubble to also be an sf object with \code{make_spatial_sf()}:
 

### 2023-06-28 cubble.Rmd
proofread done

@@
 
-In this example, we will introduce the spatial and temporal matching in cubble using an example on matching river level data with precipitation in Victoria, Australia.
+The Bureau of Meteorology collects water level data and they can be matched with the precipitation data from the climate weather stations.
+The data `river` from the \pkg{cubble} package contains water course level data for 71 river gauges collected in Victoria, Australia.
+Victoria weather station data can be subsetted from the `climate_aus` data in \pkg{cubble}.
 
-The water level data collected by the Bureau of Meteorology, can be compared with the precipitation since rainfall can directly impact water level in river.
-Both `climate_vic` and `river` are cubble objects, and we can obtain a summary of the 10 closest pairs between them:
-
-```{r}
+```{r echo = TRUE}
 climate_vic <- climate_aus |>
@@
 
+Spatial match on the site location is first performed since rainfall can directly impact water level in the nearby river. With \code{match_spatial()}, we can obtain a summary of the 10 closest pairs:
+
 ```{r echo = TRUE}
@@
 
-The result can also be returned as cubble objects by setting the argument `return_cubble = TRUE`.
-The output is be a list where each element is a paired cubble object.
-To combine all the results into a single cubble, you can use `bind_rows()`.
-In the case when a site in the second cubble (the `river` data here) is matched to two stations in the first cubble (`climate_vic` here), the binding may not be successful since cubble requires unique rows in the nested form.
-In the summary table above, the river station `226027` is matched to more than one weather station: `ASN00085072` (group 3) and `ASN00085298` (group 5).
-Similarly, the river station `230200` is matched in group 7 and 8).
-In such cases, you can either deselect one pair before combining, or work with the list output with the `purrr::map` syntax:
+The result can also be returned as a list of matched cubbles, by setting the argument \code{return_cubble = TRUE}.
+All the results can be combined into a single cubble using \code{bind_rows()}, after excluding the two pairs where a river station is matched to more than one weather stations (river station `226027` is matched twice in group 3 and 5 and similarly for station `230200` in group 7 and 8).
 
@@
   spatial_n_group = 10, return_cubble = TRUE)
-str(res_sp, max.level = 0)
 (res_sp <- res_sp[-c(5, 8)] |> bind_rows())
@@
 
-For temporal matching, we match the variable `Water_course_level` from the river data to `prcp` in the weather station data.
-The variable `group` and `types` identify the matching group and the two datasets:
+To match the water level series with precipitation, the function \code{match_temporal()} is used with the variable \code{group} and \code{types} identifying the matching group and the two data sources:
 
@@
 
-Similarly, the cubble output can be returned using the argument `return_cubble = TRUE`. Here we select the four pairs with the highest number of matching peaks:
+Similarly, the cubble output can be returned using the argument `return_cubble = TRUE`.
+Here we select the four pairs with the highest number of matching peaks and show them on the map (a) and as standardized series (b) in Figure \ref{fig:matching}.
 
@@
 
-Figure \ref{fig:matching} shows four matched pairs on the map (a) and standardized data as time series (b) with concurrent increasing.
-
-```{r matching, out.width="100%", fig.height = 5, fig.width = 10, fig.cap="Weather stations and river gauges with matched pairs labelled on the map (a) and plotted across time (b).
-Precipitation and water level have been standardised between 0 and 1 to be displayed on the same scale.
-The water level reflects the increase in precipitation.
-The numbers (1, 7, 6, 9) indicate the group index derived from spatial matching, only those that were selectd by temporal matching are shown here."}
+```{r matching, out.width="100%", fig.height = 5, fig.width = 10, fig.cap="Matched weather stations and river gauges on the map (a) and across time (b).
+Precipitation and water level have been standardised between 0 and 1 to be displayed in the same scale in (b).
+The water level reflects the increase in precipitation."}
 res_tm <- res_tm |>  filter(group %in% c(1, 7, 6, 9))
@@
 
-The ERA5 reanalysis [@hersbach2020era5] provides hourly estimates of atmospheric, land and oceanic climate variables on a global scale.
-The hourly pressure level data in the NetCDF format can be downloaded from Copernicus Climate Data Store (CDS) or via the \pkg{ecmwfr} package [@ecwmfr].
-This example reproduces the row created from ERA5 reanalysis data shown in Figure 19 by @hersbach2020era5.
+The ERA5 reanalysis [@hersbach2020era5] provides hourly estimates of atmospheric, land and oceanic climate variables on a global scale and is available in the NetCDF format from Copernicus Climate Data Store (CDS).
+This example reproduces Figure 19 in the @hersbach2020era5 paper with cubble, in Figure \ref{fig:netcdf}.
 The plot shows the southern polar vortex splitting into two on 2002-09-26, and further splitting into four on 2002-10-04.
@@
 
- The following code converts a NetCDF object of class \code{ncdf4} [@ncdf4] into a cubble object:
+A \code{ncdf4} object [@ncdf4] can be converted into a cubble using \code{as_cubble()} and the NetCDF data can be subsetted with arguments \code{vars}, \code{long_range} and \code{lat_range}.
+In this example, the variables q (specific humidity) and z (geopotential) are read in and the coordinates are subsetted to every degree in longitude and latitude:
 
@@
 raw <- ncdf4::nc_open(here::here("data/era5-pressure.nc"))
-```
-
-Analysts can extract a subset of the NetCDF data with the arguments \code{vars}, \code{long_range} and \code{lat_range}.
-In this example, the variables q (specific humidity) and z (geopotential) are selected and the coordinates are subsetted to every degree in longitude and latitude:
-
-```{r echo = TRUE}
 (dt <- as_cubble(
@@
 
-Once the NetCDF data is coerced into a cubble object, subsequent analysis can be conducted to filter on the date of interest, scale the variable specific humidity and create visualisation in ggplot as in Figure \ref{fig:netcdf}.
+Once the NetCDF data is coerced into a cubble object, subsequent analysis can be conducted to filter on the date of interest, scale the variable specific humidity and create visualisation in ggplot to reproduce the ERA5 plot.
+A shnippet of code to create Figure \ref{fig:netcdf} is provided below with additional codes needed to style the plot
+
+````
+res <- dt |> 
+  face_temporal() |> 
+  filter(lubridate::date(time) %in% 
+           as.Date(c("2002-09-22", "2002-09-26",
+                     "2002-09-30", "2002-10-04"))) |>
+  unfold(long, lat) |> 
+  mutate(q = q* 10^6)
+
+con <- rnaturalearth::ne_coastline("small", returnclass = "sf")
+box <- st_bbox(c(xmin = -180, ymin = -90, xmax = 180, ymax = -15), crs = st_crs(con)) 
+country <- con |> 
+  st_geometry() |> 
+  st_crop(box) |> 
+  st_cast("MULTILINESTRING")
+
+res |> 
+  ggplot() +
+  geom_point(aes(x = long, y = lat, color = q)) + 
+  geom_contour(data = res, aes(x = long, y = lat, z = z), ...) +
+  geom_sf(data = country, ...) +
+  ...
+
+````
 
@@
 The polar vortex, signalled by the high specific humidity, splits into two on 2002-09-26 and further splits into four on 2002-10-04.", dev = "png"}
-date <- c("2002-09-22", "2002-09-26", "2002-09-30", "2002-10-04") |> as.Date()
 res <- dt |> 
   face_temporal() |> 
-  filter(lubridate::date(time) %in% date) |>
+  filter(lubridate::date(time) %in% 
+           as.Date(c("2002-09-22", "2002-09-26",
+                     "2002-09-30", "2002-10-04"))) |>
   unfold(long, lat) |> 
@@
 sf_use_s2(FALSE)
-country <- con |> st_geometry() |> 
+country <- con |> 
+  st_geometry() |> 
   st_crop(box) |> 
@@
 
-Interactive graphics can be useful because they make it possible to look at the data in multiple ways on-the-fly.
-This is especially important for spatio-temporal data, where we would like to interactively connect spatial and temporal displays.
-This example describes the process of using the \pkg{cubble} package with the \pkg{crosstalk} package to build an interactive display connecting a map of Australia, with ribbon plots of temperature range observed at the stations.
+Interactive graphics can be especially useful for spatio-temporal data because they make it possible to look at the data in multiple ways on-the-fly.
+The last example describes the process of using cubble with the \pkg{crosstalk} package to build an interactive display connecting a map of Australia, with ribbon plots of temperature range observed at the stations in 2020.
 The purpose is to explore the variation of monthly temperature range over the country.
+
+We will first summarise the daily data in \code{climate_aus} into monthly average and calculate the variance of the temperature difference, which will be used to color the temperature band later.
+
+````
+clean <- climate_aus |>
+  face_temporal() |> 
+  mutate(month = lubridate::month(date)) |>
+  group_by(month) |>
+  summarise(
+    tmax = mean(tmax, na.rm = TRUE),
+    tmin = mean(tmin, na.rm = TRUE),
+    diff = mean(tmax - tmin, na.rm = TRUE)
+  ) |> 
+  face_spatial() |> 
+  rowwise() |>
+  mutate(temp_diff_var = var(ts$diff, na.rm =TRUE))
+````
+
+The spatial and temporal cubble are then created into shared \code{crosstalk} objects, plotted as ggplots, and combined together using \code{crosstalk::bscols()}:
+
+````
+nested <- clean |> SharedData$new(~id, group = "cubble")
+
+long <- clean |> 
+  face_temporal() |> 
+  SharedData$new(~id, group = "cubble")
+
+p1 <- nested |> ggplot() + ...
+p2 <- long |> ggplot() + ...
+crosstalk::bscols(plotly::ggplotly(p1), plotly::ggplotly(p2), ...)
+````
+
 Figure \ref{fig:interactive-linking} shows three snapshots of the interactivity.
-
-The key steps are to convert both the nested and long forms of the data into shared \code{crosstalk} objects, and to plot these side-by-side. The two are linked by the station identifier.
-
-``` r
-clean <- climate_full |> ...
-
-nested <- clean |> SharedData$new(~id, group = "cubble")
-long <- face_temporal(clean) |> SharedData$new(~id, group = "cubble")
-
-p1 <- nested |> ...
-p2 <- long |> ...
-
-crosstalk::bscols(plotly::ggplotly(p1), plotly::ggplotly(p2), ...)
-
-```
-
-Plot (a) shows the initial state of the interactive display: all locations are shown as dots on the map, coloured by temperature range, and the right plot shows the ribbons representing maximum to minimum for all stations.
-In plot (b) the "Mount Elizabeth" station, which shows a high variance colour on the initial map, is selected on the map and this produces the ribbon on the right.
-In plot (c) the lowest temperature in August is selected, which is "Thredbo" station on the left map.
-It was surprising to us that this was not a station in Tasmania, so for comparison a station in Tasmania is selected on the map to show in relation to Thredbo.
-We can see that Thredbo has a bigger winter dip in temperature, and although Tasmania is cold generally, it's temperatures are more constant.
+Plot (a) shows the initial state of the interactive display: all locations are shown as dots on the map, coloured by the temperature range, and the right plot shows the ribbons representing maximum to minimum for all stations.
+In plot (b) the station shows a high variance on the initial map, the "Mount Elizabeth" station, is selected and this produces the ribbon on the right.
+In plot (c) the lowest temperature in August is selected on the left map and this corresponds to the "Thredbo" station in the mountain area in Victoria and New South Wales.
+This station is compared to a station in the Tasmania island, the southernmost island of the country, selected on the map.
 
@@
 This paper presents the \proglang{R} package \pkg{cubble} for organizing, wrangling and visualizing spatio-temporal data.
-The package introduces a new data class, \code{cubble}, consisting of two subclass, spatial cubble and a temporal cubble, to organise spatio-temporal data in two different formats within the tidy data framework.
-The data structure and functions introduced in this package can be used and combined with existing spatial and temporal data analysis packages such as \pkg{sf} and \pkg{tsibble}, data wrangling packages such as \pkg{dplyr}, and visualization packages such as \pkg{ggplot2}, \pkg{plotly}, and \pkg{leaflet}.
+The package introduces a new data structure, \code{cubble}, consisting of two subclass, spatial cubble and a temporal cubble, to organise spatio-temporal data in two different formats within the tidy data framework.
+The data structure and functions introduced in the package can be used and combined with existing tools for  data wrangling, spatial and temporal data analysis, and visualization.
 
-The paper includes numerous examples to illustrate the utility of \pkg{cubble} as a data structure for spatio-temporal analysis.
+The paper includes plenty examples to illustrate the utility of \pkg{cubble} as a data structure for spatio-temporal analysis.
 These examples cover different tasks of a typical analysis workflow: handling data with spatial and temporal misalignment, matching data from multiple sources, and creating both static and interactive spatio-temporal visualisation.
-For future directions, other commonly-used spatial or temporal data structures can be integrated into \pkg{cubble} to extend analysts' familiar spatial and temporal toolkit to spatio-temporal.
+For future directions, other commonly-used spatial or temporal data structures can be integrated into the \pkg{cubble} package to extend analysts' familiar spatial and temporal toolkit to spatio-temporal.
 

### 2023-06-28 cubble.tex
proofread done

@@
   }
-
-% Pandoc syntax highlighting
-\usepackage{color}
-\usepackage{fancyvrb}
-\newcommand{\VerbBar}{|}
-\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
-\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
-% Add ',fontsize=\small' for more characters per line
-\usepackage{framed}
-\definecolor{shadecolor}{RGB}{248,248,248}
-\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\BuiltInTok}[1]{#1}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
-\newcommand{\ExtensionTok}[1]{#1}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ImportTok}[1]{#1}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\NormalTok}[1]{#1}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\RegionMarkerTok}[1]{#1}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
 
@@
 
-In this example, we will introduce the spatial and temporal matching in cubble using an example on matching river level data with precipitation in Victoria, Australia.
-
-The water level data collected by the Bureau of Meteorology, can be compared with the precipitation since rainfall can directly impact water level in river.
-Both \texttt{climate\_vic} and \texttt{river} are cubble objects, and we can obtain a summary of the 10 closest pairs between them:
+The Bureau of Meteorology collects water level data and they can be matched with the precipitation data from the climate weather stations.
+The data \texttt{river} from the \pkg{cubble} package contains water course level data for 71 river gauges collected in Victoria, Australia.
+Victoria weather station data can be subsetted from the \texttt{climate\_aus} data in \pkg{cubble}.
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> climate_vic <- climate_aus |>
++   filter(between(as.numeric(substr(id, 7, 8)), 76, 90)) |>
++   mutate(type = "climate")
+R> river <- cubble::river |> mutate(type = "river") 
+\end{CodeInput}
+\end{CodeChunk}
+
+Spatial match on the site location is first performed since rainfall can directly impact water level in the nearby river. With \code{match_spatial()}, we can obtain a summary of the 10 closest pairs:
 
@@
 
-The result can also be returned as cubble objects by setting the argument \texttt{return\_cubble\ =\ TRUE}.
-The output is be a list where each element is a paired cubble object.
-To combine all the results into a single cubble, you can use \texttt{bind\_rows()}.
-In the case when a site in the second cubble (the \texttt{river} data here) is matched to two stations in the first cubble (\texttt{climate\_vic} here), the binding may not be successful since cubble requires unique rows in the nested form.
-In the summary table above, the river station \texttt{226027} is matched to more than one weather station: \texttt{ASN00085072} (group 3) and \texttt{ASN00085298} (group 5).
-Similarly, the river station \texttt{230200} is matched in group 7 and 8).
-In such cases, you can either deselect one pair before combining, or work with the list output with the \texttt{purrr::map} syntax:
+The result can also be returned as a list of matched cubbles, by setting the argument \code{return_cubble = TRUE}.
+All the results can be combined into a single cubble using \code{bind_rows()}, after excluding the two pairs where a river station is matched to more than one weather stations (river station \texttt{226027} is matched twice in group 3 and 5 and similarly for station \texttt{230200} in group 7 and 8).
 
@@
 +   spatial_n_group = 10, return_cubble = TRUE)
-R> str(res_sp, max.level = 0)
-\end{CodeInput}
-\begin{CodeOutput}
-List of 10
-\end{CodeOutput}
-\begin{CodeInput}
 R> (res_sp <- res_sp[-c(5, 8)] |> bind_rows())
@@
 
-For temporal matching, we match the variable \texttt{Water\_course\_level} from the river data to \texttt{prcp} in the weather station data.
-The variable \texttt{group} and \texttt{types} identify the matching group and the two datasets:
+To match the water level series with precipitation, the function \code{match_temporal()} is used with the variable \code{group} and \code{types} identifying the matching group and the two data sources:
 
@@
 
-Similarly, the cubble output can be returned using the argument \texttt{return\_cubble\ =\ TRUE}. Here we select the four pairs with the highest number of matching peaks:
+Similarly, the cubble output can be returned using the argument \texttt{return\_cubble\ =\ TRUE}.
+Here we select the four pairs with the highest number of matching peaks and show them on the map (a) and as standardized series (b) in Figure \ref{fig:matching}.
 
@@
 
-Figure \ref{fig:matching} shows four matched pairs on the map (a) and standardized data as time series (b) with concurrent increasing.
-
 \begin{CodeChunk}
@@
 
-\caption[Weather stations and river gauges with matched pairs labelled on the map (a) and plotted across time (b)]{Weather stations and river gauges with matched pairs labelled on the map (a) and plotted across time (b).
-Precipitation and water level have been standardised between 0 and 1 to be displayed on the same scale.
-The water level reflects the increase in precipitation.
-The numbers (1, 7, 6, 9) indicate the group index derived from spatial matching, only those that were selectd by temporal matching are shown here.}\label{fig:matching}
+\caption[Matched weather stations and river gauges on the map (a) and across time (b)]{Matched weather stations and river gauges on the map (a) and across time (b).
+Precipitation and water level have been standardised between 0 and 1 to be displayed in the same scale in (b).
+The water level reflects the increase in precipitation.}\label{fig:matching}
 \end{figure}
@@
 
-The ERA5 reanalysis \citep{hersbach2020era5} provides hourly estimates of atmospheric, land and oceanic climate variables on a global scale.
-The hourly pressure level data in the NetCDF format can be downloaded from Copernicus Climate Data Store (CDS) or via the \pkg{ecmwfr} package \citep{ecwmfr}.
-This example reproduces the row created from ERA5 reanalysis data shown in Figure 19 by \citet{hersbach2020era5}.
+The ERA5 reanalysis \citep{hersbach2020era5} provides hourly estimates of atmospheric, land and oceanic climate variables on a global scale and is available in the NetCDF format from Copernicus Climate Data Store (CDS).
+This example reproduces Figure 19 in the \citet{hersbach2020era5} paper with cubble, in Figure \ref{fig:netcdf}.
 The plot shows the southern polar vortex splitting into two on 2002-09-26, and further splitting into four on 2002-10-04.
@@
 
-The following code converts a NetCDF object of class \code{ncdf4} \citep{ncdf4} into a cubble object:
+A \code{ncdf4} object \citep{ncdf4} can be converted into a cubble using \code{as_cubble()} and the NetCDF data can be subsetted with arguments \code{vars}, \code{long_range} and \code{lat_range}.
+In this example, the variables q (specific humidity) and z (geopotential) are read in and the coordinates are subsetted to every degree in longitude and latitude:
 
@@
 R> raw <- ncdf4::nc_open(here::here("data/era5-pressure.nc"))
-\end{CodeInput}
-\end{CodeChunk}
-
-Analysts can extract a subset of the NetCDF data with the arguments \code{vars}, \code{long_range} and \code{lat_range}.
-In this example, the variables q (specific humidity) and z (geopotential) are selected and the coordinates are subsetted to every degree in longitude and latitude:
-
-\begin{CodeChunk}
-\begin{CodeInput}
 R> (dt <- as_cubble(
@@
 
-Once the NetCDF data is coerced into a cubble object, subsequent analysis can be conducted to filter on the date of interest, scale the variable specific humidity and create visualisation in ggplot as in Figure \ref{fig:netcdf}.
+Once the NetCDF data is coerced into a cubble object, subsequent analysis can be conducted to filter on the date of interest, scale the variable specific humidity and create visualisation in ggplot to reproduce the ERA5 plot.
+A shnippet of code to create Figure \ref{fig:netcdf} is provided below with additional codes needed to style the plot
+
+\begin{verbatim}
+res <- dt |> 
+  face_temporal() |> 
+  filter(lubridate::date(time) %in% 
+           as.Date(c("2002-09-22", "2002-09-26",
+                     "2002-09-30", "2002-10-04"))) |>
+  unfold(long, lat) |> 
+  mutate(q = q* 10^6)
+
+con <- rnaturalearth::ne_coastline("small", returnclass = "sf")
+box <- st_bbox(c(xmin = -180, ymin = -90, xmax = 180, ymax = -15), crs = st_crs(con)) 
+country <- con |> 
+  st_geometry() |> 
+  st_crop(box) |> 
+  st_cast("MULTILINESTRING")
+
+res |> 
+  ggplot() +
+  geom_point(aes(x = long, y = lat, color = q)) + 
+  geom_contour(data = res, aes(x = long, y = lat, z = z), ...) +
+  geom_sf(data = country, ...) +
+  ...
+\end{verbatim}
 
@@
 
-Interactive graphics can be useful because they make it possible to look at the data in multiple ways on-the-fly.
-This is especially important for spatio-temporal data, where we would like to interactively connect spatial and temporal displays.
-This example describes the process of using the \pkg{cubble} package with the \pkg{crosstalk} package to build an interactive display connecting a map of Australia, with ribbon plots of temperature range observed at the stations.
+Interactive graphics can be especially useful for spatio-temporal data because they make it possible to look at the data in multiple ways on-the-fly.
+The last example describes the process of using cubble with the \pkg{crosstalk} package to build an interactive display connecting a map of Australia, with ribbon plots of temperature range observed at the stations in 2020.
 The purpose is to explore the variation of monthly temperature range over the country.
+
+We will first summarise the daily data in \code{climate_aus} into monthly average and calculate the variance of the temperature difference, which will be used to color the temperature band later.
+
+\begin{verbatim}
+clean <- climate_aus |>
+  face_temporal() |> 
+  mutate(month = lubridate::month(date)) |>
+  group_by(month) |>
+  summarise(
+    tmax = mean(tmax, na.rm = TRUE),
+    tmin = mean(tmin, na.rm = TRUE),
+    diff = mean(tmax - tmin, na.rm = TRUE)
+  ) |> 
+  face_spatial() |> 
+  rowwise() |>
+  mutate(temp_diff_var = var(ts$diff, na.rm =TRUE))
+\end{verbatim}
+
+The spatial and temporal cubble are then created into shared \code{crosstalk} objects, plotted as ggplots, and combined together using \code{crosstalk::bscols()}:
+
+\begin{verbatim}
+nested <- clean |> SharedData$new(~id, group = "cubble")
+
+long <- clean |> 
+  face_temporal() |> 
+  SharedData$new(~id, group = "cubble")
+
+p1 <- nested |> ggplot() + ...
+p2 <- long |> ggplot() + ...
+crosstalk::bscols(plotly::ggplotly(p1), plotly::ggplotly(p2), ...)
+\end{verbatim}
+
 Figure \ref{fig:interactive-linking} shows three snapshots of the interactivity.
-
-The key steps are to convert both the nested and long forms of the data into shared \code{crosstalk} objects, and to plot these side-by-side. The two are linked by the station identifier.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{clean }\OtherTok{\textless{}{-}}\NormalTok{ climate\_full }\SpecialCharTok{|\textgreater{}}\NormalTok{ ...}
-
-\NormalTok{nested }\OtherTok{\textless{}{-}}\NormalTok{ clean }\SpecialCharTok{|\textgreater{}}\NormalTok{ SharedData}\SpecialCharTok{$}\FunctionTok{new}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{id, }\AttributeTok{group =} \StringTok{"cubble"}\NormalTok{)}
-\NormalTok{long }\OtherTok{\textless{}{-}} \FunctionTok{face\_temporal}\NormalTok{(clean) }\SpecialCharTok{|\textgreater{}}\NormalTok{ SharedData}\SpecialCharTok{$}\FunctionTok{new}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{id, }\AttributeTok{group =} \StringTok{"cubble"}\NormalTok{)}
-
-\NormalTok{p1 }\OtherTok{\textless{}{-}}\NormalTok{ nested }\SpecialCharTok{|\textgreater{}}\NormalTok{ ...}
-\NormalTok{p2 }\OtherTok{\textless{}{-}}\NormalTok{ long }\SpecialCharTok{|\textgreater{}}\NormalTok{ ...}
-
-\NormalTok{crosstalk}\SpecialCharTok{::}\FunctionTok{bscols}\NormalTok{(plotly}\SpecialCharTok{::}\FunctionTok{ggplotly}\NormalTok{(p1), plotly}\SpecialCharTok{::}\FunctionTok{ggplotly}\NormalTok{(p2), ...)}
-\end{Highlighting}
-\end{Shaded}
-
-Plot (a) shows the initial state of the interactive display: all locations are shown as dots on the map, coloured by temperature range, and the right plot shows the ribbons representing maximum to minimum for all stations.
-In plot (b) the ``Mount Elizabeth'' station, which shows a high variance colour on the initial map, is selected on the map and this produces the ribbon on the right.
-In plot (c) the lowest temperature in August is selected, which is ``Thredbo'' station on the left map.
-It was surprising to us that this was not a station in Tasmania, so for comparison a station in Tasmania is selected on the map to show in relation to Thredbo.
-We can see that Thredbo has a bigger winter dip in temperature, and although Tasmania is cold generally, it's temperatures are more constant.
+Plot (a) shows the initial state of the interactive display: all locations are shown as dots on the map, coloured by the temperature range, and the right plot shows the ribbons representing maximum to minimum for all stations.
+In plot (b) the station shows a high variance on the initial map, the ``Mount Elizabeth'' station, is selected and this produces the ribbon on the right.
+In plot (c) the lowest temperature in August is selected on the left map and this corresponds to the ``Thredbo'' station in the mountain area in Victoria and New South Wales.
+This station is compared to a station in the Tasmania island, the southernmost island of the country, selected on the map.
 
@@
 This paper presents the \proglang{R} package \pkg{cubble} for organizing, wrangling and visualizing spatio-temporal data.
-The package introduces a new data class, \code{cubble}, consisting of two subclass, spatial cubble and a temporal cubble, to organise spatio-temporal data in two different formats within the tidy data framework.
-The data structure and functions introduced in this package can be used and combined with existing spatial and temporal data analysis packages such as \pkg{sf} and \pkg{tsibble}, data wrangling packages such as \pkg{dplyr}, and visualization packages such as \pkg{ggplot2}, \pkg{plotly}, and \pkg{leaflet}.
-
-The paper includes numerous examples to illustrate the utility of \pkg{cubble} as a data structure for spatio-temporal analysis.
+The package introduces a new data structure, \code{cubble}, consisting of two subclass, spatial cubble and a temporal cubble, to organise spatio-temporal data in two different formats within the tidy data framework.
+The data structure and functions introduced in the package can be used and combined with existing tools for data wrangling, spatial and temporal data analysis, and visualization.
+
+The paper includes plenty examples to illustrate the utility of \pkg{cubble} as a data structure for spatio-temporal analysis.
 These examples cover different tasks of a typical analysis workflow: handling data with spatial and temporal misalignment, matching data from multiple sources, and creating both static and interactive spatio-temporal visualisation.
-For future directions, other commonly-used spatial or temporal data structures can be integrated into \pkg{cubble} to extend analysts' familiar spatial and temporal toolkit to spatio-temporal.
+For future directions, other commonly-used spatial or temporal data structures can be integrated into the \pkg{cubble} package to extend analysts' familiar spatial and temporal toolkit to spatio-temporal.
 

### 2023-06-28 code/interactive-graphic.R

@@ -2,7 +2,7 @@
 library("crosstalk")
 library("plotly")
 
-clean <- climate_full |>
+clean <- climate_aus |>
   filter(!id %in% c("ASN00067033", "ASN00072091", "ASN00059040", 
                     "ASN00097053", "ASN00040856", "ASN00015548")) |>
   face_temporal() |> 
@@ -13,7 +13,6 @@ clean <- climate_full |>
     tmin = mean(tmin, na.rm = TRUE),
     diff = mean(tmax - tmin, na.rm = TRUE)
   ) |> 
-  ungroup(month) |>
   mutate(dummy_date = as.Date(glue::glue("2021-{month}-01"))) |> 
   face_spatial() |> 
   mutate(temp_diff_var = ts |> pull(diff) |> var(na.rm = TRUE))

### 2023-06-28 reproducible-script.R

@@ -1,4 +1,4 @@
-## ----setup, echo = FALSE------------------------------------------------------
+## ----setup, echo = FALSE--------------------------------------------------------
 knitr::opts_chunk$set(
   echo = FALSE,
   warning = FALSE,
@@ -12,9 +12,10 @@ options(prompt = "R> ", continue = "+ ",
         width = 80, styler.cache_root = "styler-perm")
 
 
-## ----echo = FALSE-------------------------------------------------------------
+## ----echo = FALSE---------------------------------------------------------------
 library("cubble")
 library("dplyr")
+library("tidyr")
 library("ggplot2")
 library("patchwork")
 library("tsibble")
@@ -29,19 +30,19 @@ library("colorspace")
 library("units")
 
 
-## -----------------------------------------------------------------------------
+## -------------------------------------------------------------------------------
 cb_nested <- climate_mel
 
 
-## ----echo = TRUE--------------------------------------------------------------
+## ----echo = TRUE----------------------------------------------------------------
 cb_nested
 
 
-## -----------------------------------------------------------------------------
+## -------------------------------------------------------------------------------
 cb_long <- face_temporal(climate_mel)
 
 
-## ----echo = TRUE--------------------------------------------------------------
+## ----echo = TRUE----------------------------------------------------------------
 cb_long
 
 
@@ -49,80 +50,79 @@ cb_long
 #knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.004.png"))
 
 
-## ----echo = TRUE--------------------------------------------------------------
+## ----echo = TRUE----------------------------------------------------------------
 spatial(cb_long)
 
 
-## ----echo = TRUE, eval = FALSE------------------------------------------------
-## make_cubble(spatial = stations, temporal = meteo,
-##             key = id, index = date, coords = c(long, lat))
+## ----echo = TRUE----------------------------------------------------------------
+make_cubble(spatial = stations, temporal = meteo,
+            key = id, index = date, coords = c(long, lat))
 
 
-## ----echo = TRUE--------------------------------------------------------------
+## ----echo = TRUE----------------------------------------------------------------
 climate_flat |> as_cubble(key = id, index = date, coords = c(long, lat))
 
 
[…]

### 2023-06-28 cubble.Rmd
remove the dependency on github version of sftime

@@
 Consider the \code{climate_aus} data in the \pkg{cubble} package with `r nrow(cubble::climate_aus)` stations observed daily in a single year 2020.
-The created \code{sftime} object is approximately `r library("cubble"); round(object.size(sftime::st_as_sftime(climate_aus))/object.size(climate_aus))` times larger than the corresponding \code{cubble} object (`r round(object.size(sftime::st_as_sftime(climate_aus))/1e6, 2)` MB vs.
-`r round(object.size(cubble::climate_aus)/1e6, 2)` MB).
+The created \code{sftime} object is approximately 14 times larger than the corresponding \code{cubble} object (118 MB vs.
+8.5 MB).
+
+<!-- round(object.size(sftime::st_as_sftime(climate_aus))/object.size(climate_aus)) -->
+<!-- round(object.size(sftime::st_as_sftime(climate_aus))/1e6, 2) -->
+ <!-- round(object.size(cubble::climate_aus)/1e6, 2) -->
 

### 2023-06-28 cubble.tex
cache to false

@@
 \caption{\label{tab:funs} An overview of functions implemented in the \pkg{cubble} package, categorised into base R, tidyverse, and cubble functions.}\tabularnewline
-\toprule()
+\toprule\noalign{}
 \begin{minipage}[b]{\linewidth}\raggedright
@@
 \end{minipage} \\
-\midrule()
+\midrule\noalign{}
 \endfirsthead
-\toprule()
+\toprule\noalign{}
 \begin{minipage}[b]{\linewidth}\raggedright
@@
 \end{minipage} \\
-\midrule()
+\midrule\noalign{}
 \endhead
+\bottomrule\noalign{}
+\endlastfoot
 base R & \texttt{{[}}, \texttt{{[}{[}\textless{}-}, \texttt{names\textless{}-} \\
@@
 cubble & \texttt{as\_cubble}, \texttt{cubble}, \texttt{make\_cubble}, \texttt{check\_key}, \texttt{face\_temporal}, \texttt{face\_spatial}, \texttt{unfold}, \texttt{key}, \texttt{key\_vars}, \texttt{key\_data}, \texttt{index}, \texttt{index\_var}, \texttt{coords}, \texttt{spatial}, \texttt{match\_sites}, \texttt{match\_spatial}, \texttt{match\_temporal}, \texttt{geom\_glyph}, \texttt{geom\_glyph\_box}, \texttt{geom\_glyph\_line}, \texttt{make\_spatial\_sf}, \texttt{make\_temporal\_tsibble}, \texttt{fill\_gaps}, and \texttt{scan\_gaps} \\
-\bottomrule()
 \end{longtable}
@@
 Methods are implemented in the \code{cubble\_df} class when they behave consistent across the spatial and temporal cubble; otherwise, they are implemented separately.}\tabularnewline
-\toprule()
+\toprule\noalign{}
 \begin{minipage}[b]{\linewidth}\raggedright
@@
 \end{minipage} \\
-\midrule()
+\midrule\noalign{}
 \endhead
+\bottomrule\noalign{}
+\endlastfoot
 \texttt{cubble\_df} & \texttt{{[}{[}\textless{}-,\ dplyr\_col\_modify,\ key\_data,\ key\_vars,\ key,\ print} \\
@@
 \texttt{temporal\_cubble\_df} & \texttt{{[},\ names\textless{}-,\ tbl\_sum,\ arrange,\ dplyr\_reconstruct,\ dplyr\_row\_slice,\ face\_spatial,\ face\_temporal,\ unfold,\ fill\_gaps,\ group\_by,\ ungroup,\ \ rename,\ rowwise,\ scan\_gaps,\ select,\ spatial,\ summarise,\ tbl\_sum,\ bind\_rows,\ bind\_cols,\ update\_cubble} \\
-\bottomrule()
 \end{longtable}
@@
 Consider the \code{climate_aus} data in the \pkg{cubble} package with 639 stations observed daily in a single year 2020.
-The created \code{sftime} object is approximately 14 times larger than the corresponding \code{cubble} object (118.24 MB vs.~8.52 MB).
+The created \code{sftime} object is approximately 14 times larger than the corresponding \code{cubble} object (118 MB vs.~8.5 MB).
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/research/paper-cubble/figures/glyphmap-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/cookd/students_PhD/Sherry/paper-cubble/figures/glyphmap-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/research/paper-cubble/figures/matching-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/cookd/students_PhD/Sherry/paper-cubble/figures/matching-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/research/paper-cubble/figures/netcdf-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/cookd/students_PhD/Sherry/paper-cubble/figures/netcdf-1}
 

### 2023-06-29 cubble.Rmd
cache to false
Correct spelling

@@
   message = FALSE,
-  cache = TRUE,
+  cache = FALSE,
   fig.path = here::here("figures/")
@@
 To communicate spatial and temporal information collectively in display,  a few visualisation options are available: one can make faceted maps across time, creating map animations, or constructing interactive graphics to link between map and time series plot.
-Faceted maps and animation, while commonly used, may not be helpful for comparing across time since eyes need to jump across multiple facets or frames to locate the space before preceiving the changes in time.
+Faceted maps and animation, while commonly used, may not be helpful for comparing across time since eyes need to jump across multiple facets or frames to locate the space before perceiving the changes in time.
 The glyph map [@Wickham2012-yr] resolves this issue by imposing the time series onto the map as a glyph through coordinate transformation.
@@
 To prevent overlapping of weather stations on the map, stations are selected to ensure a minimum distance of 50km.
-Distance between stations can be calcualted with \code{sf::st_distance()} after turning the spatial cubble to also be an sf object with \code{make_spatial_sf()}:
+Distance between stations can be calculated with \code{sf::st_distance()} after turning the spatial cubble to also be an sf object with \code{make_spatial_sf()}:
 
@@
 Once the NetCDF data is coerced into a cubble object, subsequent analysis can be conducted to filter on the date of interest, scale the variable specific humidity and create visualisation in ggplot to reproduce the ERA5 plot.
-A shnippet of code to create Figure \ref{fig:netcdf} is provided below with additional codes needed to style the plot
+A snippet of code to create Figure \ref{fig:netcdf} is provided below with additional codes needed to style the plot
 

### 2023-06-29 cubble.tex
Correct spelling

--- 2023-06-28 cubble.tex
@@
 
-{\centering \includegraphics[width=1\linewidth]{figures/diagram-keynotes/diagram-keynotes.001}
+{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Desktop/paper-cubble/figures/diagram-keynotes/diagram-keynotes.001}
 
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.35\textheight]{figures/diagram-keynotes/‎diagram-keynotes.‎002}
+{\centering \includegraphics[width=1\linewidth,height=0.35\textheight]{/Users/pmen0008/Desktop/paper-cubble/figures/diagram-keynotes/‎diagram-keynotes.‎002}
 
@@
 To communicate spatial and temporal information collectively in display, a few visualisation options are available: one can make faceted maps across time, creating map animations, or constructing interactive graphics to link between map and time series plot.
-Faceted maps and animation, while commonly used, may not be helpful for comparing across time since eyes need to jump across multiple facets or frames to locate the space before preceiving the changes in time.
+Faceted maps and animation, while commonly used, may not be helpful for comparing across time since eyes need to jump across multiple facets or frames to locate the space before perceiving the changes in time.
 The glyph map \citep{Wickham2012-yr} resolves this issue by imposing the time series onto the map as a glyph through coordinate transformation.
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/research/paper-cubble/figures/glyphmap-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Desktop/paper-cubble/figures/glyphmap-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/research/paper-cubble/figures/matching-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Desktop/paper-cubble/figures/matching-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/research/paper-cubble/figures/netcdf-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Desktop/paper-cubble/figures/netcdf-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.23\textheight]{figures/linking} \includegraphics[width=1\linewidth,height=0.23\textheight]{figures/linking-north} \includegraphics[width=1\linewidth,height=0.23\textheight]{figures/linking-lower}
+{\centering \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/pmen0008/Desktop/paper-cubble/figures/linking} \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/pmen0008/Desktop/paper-cubble/figures/linking-north} \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/pmen0008/Desktop/paper-cubble/figures/linking-lower}
 

### 2023-06-29 cubble.Rmd
Fixing key words to be homogeneous with main text

@@
   # at least one keyword must be supplied
-  formatted: [spatial, temporal, spatio temporal, "\\proglang{R}", environmental data, exploratory data analysis]
-  plain:     [spatial, temporal, spatio temporal, R, environmental data, exploratory data analysis]
+  formatted: [spatial, temporal, spatio-temporal, "\\proglang{R}", environmental data, exploratory data analysis]
+  plain:     [spatial, temporal, spatio-temporal, R, environmental data, exploratory data analysis]
 preamble: >

### 2023-06-29 cubble.tex
Fixing key words to be homogeneous with main text

@@
 
-\Keywords{spatial, temporal, spatio temporal, \proglang{R}, environmental data, exploratory data analysis}
-\Plainkeywords{spatial, temporal, spatio temporal, R, environmental data, exploratory data analysis}
+\Keywords{spatial, temporal, spatio-temporal, \proglang{R}, environmental data, exploratory data analysis}
+\Plainkeywords{spatial, temporal, spatio-temporal, R, environmental data, exploratory data analysis}
 

### 2023-06-29 cubble.Rmd
Update 3rd paragraph intro

@@
 
-Still, these two implementations are not necessarily easy to work in analysis for analysts with a tidyverse mindset.
-In tidyverse, data are in tables and the tidy data concept [@tidydata] prescribes three principles on how data should be organised for easier analysis as 1) one observation a row, 2) one variable a column, and 3) one type of observation a table.
-The third principle of tidy data is particularly relevant for spatio-temporal data since spatial and temporal data are naturally observed at different units: the location and location at different times.
-While the tidyverse suite implements data wrangling and visualisation tools operated on a single table, there has not been many tools for handling relational data for spatio-temporal analysis.
+However, the data representation implemented in those packages might present certain challenges when applying the principles of tidy data [@tidydata] for data analysis.
+The concept of tidy data is based on three principles regarding how data should be organized in tables to facilitate easier analysis: 1) one observation a row, 2) one variable a column, and 3) one type of observation a table.
+The third principle of tidy data is particularly relevant for spatio-temporal data since these data are naturally observed at different units: the spatial locations and the temporal units.
+While the tidyverse suite of R packages implements data wrangling and visualization tools primarily focused on working with single tables, there are not many tools available for handling relational data specifically for spatio-temporal data.
 This motivates a new design to organise spatio-temporal data in a way that would make data wrangling, visualizing and analyzing easier.

### 2023-06-29 cubble.tex
Update 3rd paragraph intro

@@
 
-Still, these two implementations are not necessarily easy to work in analysis for analysts with a tidyverse mindset.
-In tidyverse, data are in tables and the tidy data concept \citep{tidydata} prescribes three principles on how data should be organised for easier analysis as 1) one observation a row, 2) one variable a column, and 3) one type of observation a table.
-The third principle of tidy data is particularly relevant for spatio-temporal data since spatial and temporal data are naturally observed at different units: the location and location at different times.
-While the tidyverse suite implements data wrangling and visualisation tools operated on a single table, there has not been many tools for handling relational data for spatio-temporal analysis.
+However, the data representation implemented in those packages might present certain challenges when applying the principles of tidy data \citep{tidydata} for data analysis.
+The concept of tidy data is based on three principles regarding how data should be organized in tables to facilitate easier analysis: 1) one observation a row, 2) one variable a column, and 3) one type of observation a table.
+The third principle of tidy data is particularly relevant for spatio-temporal data since these data are naturally observed at different units: the spatial locations and the temporal units.
+While the tidyverse suite of R packages implements data wrangling and visualization tools primarily focused on working with single tables, there are not many tools available for handling relational data specifically for spatio-temporal data.
 This motivates a new design to organise spatio-temporal data in a way that would make data wrangling, visualizing and analyzing easier.

### 2023-06-29 cubble.Rmd
Minor changes section 2.1

@@
 GPS data tracks unique point locations at different timestamps and is represented as trajectories.
-Satellite imageries capture snapshots of landscapes at selected time and is commonly structured as raster data.
+Satellite imageries capture snapshots of landscapes at selected times and is commonly structured as raster data.
 In this paper, we propose the \pkg{cubble} package to organise spatio-temporal data collected at unique fixed locations while allowing for irregularities in the temporal dimension, such as weather station data.
@@
 
-In a spatial cubble, spatial variables are organised as columns and temporal variables are nested within a specialised `ts` column.
-The spatial cubble object, \code{cb_nested}, printed below contains weather records of three airport stations from Global Historical Climatology Network Daily (GHCND).
-The spatial cubble is convenient for wrangling the spatial variables:
+In a spatial cubble object, spatial variables are organised as columns and temporal variables are nested within a specialised `ts` column.
+For example, the spatial cubble object, \code{cb_nested} printed below, contains weather records of three airport stations from the Global Historical Climatology Network Daily (GHCND) database.
+In this case, the spatial cubble is convenient for wrangling the spatial variables:
 
@@
 In cubble, the `key` attribute identifies the row in the spatial cubble (given the internal use of `tidyr::nest()` for nesting), and when combined with the `index` argument, it identifies the row in the temporal cubble.
-Currently, cubble only supports one variable as the key, and the accepted temporal classes for index includes the base R classes `Date`, `POSIXlt`, `POSIXct`, as well as tsibble's `yearmonth`, `yearweek`, and `yearquarter` classes.
-The `coords` attribute represents an ordered pair of coordinates.
-It can be either an unprojected pair of longitude and latitude, or a projected easting and northing value.
-The `sf` package is used under the hood to calculate the bounding box, displayed in the header of a spatial cubble.
+Currently, cubble only supports one variable as the key, and the accepted temporal classes for `index` includes the base R classes `Date`, `POSIXlt`, `POSIXct`, as well as tsibble's `yearmonth`, `yearweek`, and `yearquarter` classes.
+The `coords` attribute represents an ordered pair of coordinates that can be either an unprojected pair of longitude and latitude, or a projected easting and northing value.
+The `sf` package is used under the hood to calculate the spatial bounding box, displayed in the header of a spatial cubble.
 
 The temporal cubble has a special attribute called `spatial` to store the spatial variables.
-Shortcut functions are available to extract attributes, for example, \code{spatial()} for extracting spatial variables from the temporal cubble:
+Shortcut functions are available to extract attributes from the temporal cubble object, for example, \code{spatial()} for extracting spatial variables:
 

### 2023-06-29 cubble.tex
Minor changes section 2.1

@@
 
-In a spatial cubble, spatial variables are organised as columns and temporal variables are nested within a specialised \texttt{ts} column.
-The spatial cubble object, \code{cb_nested}, printed below contains weather records of three airport stations from Global Historical Climatology Network Daily (GHCND).
-The spatial cubble is convenient for wrangling the spatial variables:
+In a spatial cubble object, spatial variables are organised as columns and temporal variables are nested within a specialised \texttt{ts} column.
+For example, the spatial cubble object, \code{cb_nested} printed below, contains weather records of three airport stations from the Global Historical Climatology Network Daily (GHCND) database.
+In this case, the spatial cubble is convenient for wrangling the spatial variables:
 
@@
 In cubble, the \texttt{key} attribute identifies the row in the spatial cubble (given the internal use of \texttt{tidyr::nest()} for nesting), and when combined with the \texttt{index} argument, it identifies the row in the temporal cubble.
-Currently, cubble only supports one variable as the key, and the accepted temporal classes for index includes the base R classes \texttt{Date}, \texttt{POSIXlt}, \texttt{POSIXct}, as well as tsibble's \texttt{yearmonth}, \texttt{yearweek}, and \texttt{yearquarter} classes.
-The \texttt{coords} attribute represents an ordered pair of coordinates.
-It can be either an unprojected pair of longitude and latitude, or a projected easting and northing value.
-The \texttt{sf} package is used under the hood to calculate the bounding box, displayed in the header of a spatial cubble.
+Currently, cubble only supports one variable as the key, and the accepted temporal classes for \texttt{index} includes the base R classes \texttt{Date}, \texttt{POSIXlt}, \texttt{POSIXct}, as well as tsibble's \texttt{yearmonth}, \texttt{yearweek}, and \texttt{yearquarter} classes.
+The \texttt{coords} attribute represents an ordered pair of coordinates that can be either an unprojected pair of longitude and latitude, or a projected easting and northing value.
+The \texttt{sf} package is used under the hood to calculate the spatial bounding box, displayed in the header of a spatial cubble.
 
 The temporal cubble has a special attribute called \texttt{spatial} to store the spatial variables.
-Shortcut functions are available to extract attributes, for example, \code{spatial()} for extracting spatial variables from the temporal cubble:
+Shortcut functions are available to extract attributes from the temporal cubble object, for example, \code{spatial()} for extracting spatial variables:
 

### 2023-06-29 cubble.Rmd
Changes section 2.3

@@
 
-Table \@ref(tab:funs) summarises the functions implemented in the \pkg{cubble} package and Table \@ref(tab:methods) details the methods implemented for each of the three cubble classes.
-The \code{cubble_df} class handles methods that behave consistently in both spatial and temporal cubble.
-When the method has distinct behavior, it is implemented separately in \code{spatial_cubble_df} and \code{temporal_cubble_df}.
+The \pkg{cubble} package has several functions implemented for data wrangling and to facilitate data analysis as summarized in Table \@ref(tab:funs).
+In addition, for each of the three cubble classes there are number of methods implemented that facilitates the handling of the data as shown in Table \@ref(tab:methods).
+In particular, the \code{cubble_df} class handles methods that behave consistently in both spatial and temporal cubble.
+When the method has a distinct behavior on the spatial and temporal cubble, it is implemented separately in \code{spatial_cubble_df} and \code{temporal_cubble_df}.
 
@@
 
-The pair of cubble verbs, \code{face_temporal()} and \code{face_spatial()}, pivots the cubble object between its two forms, as illustrated in Figure \ref{fig:face}.
-The code below connects the a spatial cubble (\code{cb_nested}) and a temporal temporal (\code{cb_long}) introduced in Section \@ref(object) with \code{face_temporal()} and \code{face_spatial()}:
+The pair of cubble verbs, \code{face_temporal()} and \code{face_spatial()}, pivots the cubble object between its two forms or faces, as illustrated in Figure \ref{fig:face}.
+The code below connects a spatial cubble (\code{cb_nested}) and a temporal cubble (\code{cb_long}), introduced in Section \@ref(object), with \code{face_temporal()} and \code{face_spatial()}:
 

### 2023-06-29 cubble.tex
Changes section 2.3

@@
 
-Table \ref{tab:funs} summarises the functions implemented in the \pkg{cubble} package and Table \ref{tab:methods} details the methods implemented for each of the three cubble classes.
-The \code{cubble_df} class handles methods that behave consistently in both spatial and temporal cubble.
-When the method has distinct behavior, it is implemented separately in \code{spatial_cubble_df} and \code{temporal_cubble_df}.
+The \pkg{cubble} package has several functions implemented for data wrangling and to facilitate data analysis as summarized in Table \ref{tab:funs}.
+In addition, for each of the three cubble classes there are number of methods implemented that facilitates the handling of the data as shown in Table \ref{tab:methods}.
+In particular, the \code{cubble_df} class handles methods that behave consistently in both spatial and temporal cubble.
+When the method has a distinct behavior on the spatial and temporal cubble, it is implemented separately in \code{spatial_cubble_df} and \code{temporal_cubble_df}.
 
@@
 
-The pair of cubble verbs, \code{face_temporal()} and \code{face_spatial()}, pivots the cubble object between its two forms, as illustrated in Figure \ref{fig:face}.
-The code below connects the a spatial cubble (\code{cb_nested}) and a temporal temporal (\code{cb_long}) introduced in Section \ref{object} with \code{face_temporal()} and \code{face_spatial()}:
+The pair of cubble verbs, \code{face_temporal()} and \code{face_spatial()}, pivots the cubble object between its two forms or faces, as illustrated in Figure \ref{fig:face}.
+The code below connects a spatial cubble (\code{cb_nested}) and a temporal cubble (\code{cb_long}), introduced in Section \ref{object}, with \code{face_temporal()} and \code{face_spatial()}:
 

### 2023-06-29 cubble.Rmd
Minor changes section 2.4

@@
 
-```{r face, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "An illustration of the function \\code{face\\_temporal()} and \\code{face\\_spatial()}: \\code{face\\_temporal()} converts a spatial cubble into a temporal cubbl to focus on the temporal variables.
-Conversely, \\code{face\\_spatial()} transforms a temporal cubble into a spatial one to for focus on the spatial variables."}
+```{r face, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "An illustration of the functions \\code{face\\_temporal()} and \\code{face\\_spatial()}: \\code{face\\_temporal()} converts a spatial cubble into a temporal cubble to focus on the temporal variables.
+Conversely, \\code{face\\_spatial()} transforms a temporal cubble into a spatial one to focus on the spatial variables."}
 knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.001.png"))
@@
 
-Methods applies to the \code{tbl_ts} class can also be applied to the temporal cubble objects, for example, checking whether the data contain temporal gaps:
+Methods applied to the \code{tbl_ts} class can also be applied to the temporal cubble objects, for example, checking whether the data contain temporal gaps:
 

### 2023-06-29 cubble.tex
Minor changes section 2.4

@@
 
-\caption{An illustration of the function \code{face\_temporal()} and \code{face\_spatial()}: \code{face\_temporal()} converts a spatial cubble into a temporal cubbl to focus on the temporal variables.
-Conversely, \code{face\_spatial()} transforms a temporal cubble into a spatial one to for focus on the spatial variables.}\label{fig:face}
+\caption{An illustration of the functions \code{face\_temporal()} and \code{face\_spatial()}: \code{face\_temporal()} converts a spatial cubble into a temporal cubble to focus on the temporal variables.
+Conversely, \code{face\_spatial()} transforms a temporal cubble into a spatial one to focus on the spatial variables.}\label{fig:face}
 \end{figure}

### 2023-06-29 cubble.Rmd
changes section 3.2

@@
 
-In \proglang{R}, there are other existing spatio-temporal data structure and this section compares and contrasts \pkg{cubble} with other existing alternative, specifically \pkg{stars} and \pkg{sftime}.
+In \proglang{R}, there are other existing spatio-temporal data structure and this section compares and contrasts \pkg{cubble} with other existing alternatives, specifically \pkg{stars} and \pkg{sftime}.
 The \pkg{stars} package [@stars] uses an array structure, as oppose to tibble, to represent multivariate spatio-temporal data.
@@
 The \pkg{sftime} package [@sftime] also builds from a tibble object and its focus is on handling irregular spatio-temporal data.
-This means \pkg{sftime} can also handle full space-time grids and sparse space-time layouts represented in \pkg{cubble}, but \pkg{cubble} uses nesting to avoid storing spatial variables repetitively at each timestamp.
+This means \pkg{sftime} can also handle full space-time grids and sparse space-time layouts represented in \pkg{cubble}.
+However, \pkg{cubble} uses nesting to avoid storing spatial variables repetitively at each timestamp.
 This provides memory efficiency when data is observed frequent, i.e.
-daily or sub-daily, or the spatial geometry is expensive to repeat, i.e.
+daily or sub-daily, or the spatial geometry is computationally expensive to store repeatedly, i.e.
 polygons or multipolygons.
-Consider the \code{climate_aus} data in the \pkg{cubble} package with `r nrow(cubble::climate_aus)` stations observed daily in a single year 2020.
-The created \code{sftime} object is approximately 14 times larger than the corresponding \code{cubble} object (118 MB vs.
+Consider the \code{climate_aus} data in the \pkg{cubble} package with `r nrow(cubble::climate_aus)` stations observed daily throughout the year 2020.
+In that case, the \code{sftime} object is approximately 14 times larger than the corresponding \code{cubble} object (118 MB vs.
 8.5 MB).
@@
 Matching time series from an old list of stations to a new list is a common task in spatio-temporal data analysis.
-In cubble, matching based on distance and time series feature can be performed using the functions `match_spatial()` and `match_temporal()`.
+In cubble, matching based on distance and time series features can be performed using the functions `match_spatial()` and `match_temporal()`.
 The `match_spatial()` function finds the matched pairs in two cubble objects based on distance between sites:
@@
 
-The cubble workflow works well with an interactive graphics pipeline (e.g., @buja1988elements; @buja1996interactive;  @sutherland2000orca; @xie2014reactive; @cheng2016enabling).
+The cubble workflow can be easily incorporated into an interactive graphics pipeline (e.g., @buja1988elements; @buja1996interactive;  @sutherland2000orca; @xie2014reactive; @cheng2016enabling).
 This section describes the linking between a map and multiple time series in a \code{cubble} object using the package \pkg{crosstalk} [@crosstalk], as illustrated in Figure \ref{fig:illu-interactive}.
-The spatial and temporal cubble can be constructed into a shared crosstalk object to create linking between a map and a time series plot.
-When a user selects a location on the map (a), the corresponding site is highlighted.
-This selection activates a row in the spatial cubble, which is then communicated to the temporal cubble, resulting in the selection of all observations with the same ID in (b).
-Consequently the temporal cubble highlights the corresponding series in the time series plot (c).
-Linking can also be initiated from the time series plot, by selecting points on the time series, which activates rows with the same ID in the temporal cubble and the corresponding row in the spatial cubble to highlight on the map.
+The spatial and temporal cubble can be constructed as a shared crosstalk object to create a link between a map and a time series plot.
+For instance, when a user selects a location on the map as shown on panel (a), the corresponding site is highlighted.
+This selection activates a row in the spatial cubble, which is then connected to the temporal cubble, resulting in the selection of all observations with the same ID as depicted in panel (b).
+Consequently the temporal cubble highlights the corresponding series in the time series plot displayed in panel (c).
+The linking can also be initiated from the time series plot by selecting points on the time series graph.
+This action selects rows with the same ID in the temporal cubble and the corresponding row in the spatial cubble so that points can be highlight on the map.
 

### 2023-06-29 cubble.tex
changes section 3.2

@@
 
-In \proglang{R}, there are other existing spatio-temporal data structure and this section compares and contrasts \pkg{cubble} with other existing alternative, specifically \pkg{stars} and \pkg{sftime}.
+In \proglang{R}, there are other existing spatio-temporal data structure and this section compares and contrasts \pkg{cubble} with other existing alternatives, specifically \pkg{stars} and \pkg{sftime}.
 The \pkg{stars} package \citep{stars} uses an array structure, as oppose to tibble, to represent multivariate spatio-temporal data.
@@
 The \pkg{sftime} package \citep{sftime} also builds from a tibble object and its focus is on handling irregular spatio-temporal data.
-This means \pkg{sftime} can also handle full space-time grids and sparse space-time layouts represented in \pkg{cubble}, but \pkg{cubble} uses nesting to avoid storing spatial variables repetitively at each timestamp.
-This provides memory efficiency when data is observed frequent, i.e.~daily or sub-daily, or the spatial geometry is expensive to repeat, i.e.~polygons or multipolygons.
-Consider the \code{climate_aus} data in the \pkg{cubble} package with 639 stations observed daily in a single year 2020.
-The created \code{sftime} object is approximately 14 times larger than the corresponding \code{cubble} object (118 MB vs.~8.5 MB).
+This means \pkg{sftime} can also handle full space-time grids and sparse space-time layouts represented in \pkg{cubble}.
+However, \pkg{cubble} uses nesting to avoid storing spatial variables repetitively at each timestamp.
+This provides memory efficiency when data is observed frequent, i.e.~daily or sub-daily, or the spatial geometry is computationally expensive to store repeatedly, i.e.~polygons or multipolygons.
+Consider the \code{climate_aus} data in the \pkg{cubble} package with 639 stations observed daily throughout the year 2020.
+In that case, the \code{sftime} object is approximately 14 times larger than the corresponding \code{cubble} object (118 MB vs.~8.5 MB).
 
@@
 Matching time series from an old list of stations to a new list is a common task in spatio-temporal data analysis.
-In cubble, matching based on distance and time series feature can be performed using the functions \texttt{match\_spatial()} and \texttt{match\_temporal()}.
+In cubble, matching based on distance and time series features can be performed using the functions \texttt{match\_spatial()} and \texttt{match\_temporal()}.
 The \texttt{match\_spatial()} function finds the matched pairs in two cubble objects based on distance between sites:
@@
 
-The cubble workflow works well with an interactive graphics pipeline (e.g., \citet{buja1988elements}; \citet{buja1996interactive}; \citet{sutherland2000orca}; \citet{xie2014reactive}; \citet{cheng2016enabling}).
+The cubble workflow can be easily incorporated into an interactive graphics pipeline (e.g., \citet{buja1988elements}; \citet{buja1996interactive}; \citet{sutherland2000orca}; \citet{xie2014reactive}; \citet{cheng2016enabling}).
 This section describes the linking between a map and multiple time series in a \code{cubble} object using the package \pkg{crosstalk} \citep{crosstalk}, as illustrated in Figure \ref{fig:illu-interactive}.
-The spatial and temporal cubble can be constructed into a shared crosstalk object to create linking between a map and a time series plot.
-When a user selects a location on the map (a), the corresponding site is highlighted.
-This selection activates a row in the spatial cubble, which is then communicated to the temporal cubble, resulting in the selection of all observations with the same ID in (b).
-Consequently the temporal cubble highlights the corresponding series in the time series plot (c).
-Linking can also be initiated from the time series plot, by selecting points on the time series, which activates rows with the same ID in the temporal cubble and the corresponding row in the spatial cubble to highlight on the map.
+The spatial and temporal cubble can be constructed as a shared crosstalk object to create a link between a map and a time series plot.
+For instance, when a user selects a location on the map as shown on panel (a), the corresponding site is highlighted.
+This selection activates a row in the spatial cubble, which is then connected to the temporal cubble, resulting in the selection of all observations with the same ID as depicted in panel (b).
+Consequently the temporal cubble highlights the corresponding series in the time series plot displayed in panel (c).
+The linking can also be initiated from the time series plot by selecting points on the time series graph.
+This action selects rows with the same ID in the temporal cubble and the corresponding row in the spatial cubble so that points can be highlight on the map.
 

### 2023-06-29 cubble.Rmd
changes section 3.3

@@
 
-To communicate spatial and temporal information collectively in display,  a few visualisation options are available: one can make faceted maps across time, creating map animations, or constructing interactive graphics to link between map and time series plot.
-Faceted maps and animation, while commonly used, may not be helpful for comparing across time since eyes need to jump across multiple facets or frames to locate the space before perceiving the changes in time.
-The glyph map [@Wickham2012-yr] resolves this issue by imposing the time series onto the map as a glyph through coordinate transformation.
-The transformation uses linear algebra to convert the temporal coordinates (minor coordinates) into the spatial coordinates (major coordinates) and is implemented in the package `GGally` [@ggally] package.
-The \pkg{cubble} package provides a ggproto implementation, `geom_glyph()`, for glyph maps and it takes four required aesthetics: `x_major`, `y_major`, `x_minor`, and `y_minor`:
+Visualizing spatio-temporal data is important for exploring and understanding the data at hand, aiding in decision-making, and facilitating effective communication.
+To collectively visualize spatial and temporal information, several options are available such as:  faceted maps across time, map animations, or interactive graphics that link maps and time series plots among others.
+Faceted maps and spatio-temporal animations, while commonly used, may be difficult to compare across time and space since users need to inspect multiple facets or frames to find changes in time or space.
+The glyph map [@Wickham2012-yr] resolves this issue by imposing the time series onto the map as a glyph through a coordinate transformation.
+The transformation uses linear algebra to convert the temporal coordinates (minor coordinates) into the spatial coordinates (major coordinates) and is implemented in the package `GGally` [@ggally].
+The \pkg{cubble} package provides a ggproto implementation to create glyph maps, `geom_glyph()`  and it takes four required aestheticsf : `x_major`, `y_major`, `x_minor`, and `y_minor`:
 
@@
 
-Other useful controls over the glyph map includes:
+Other useful controls to modify the glyph map that can be include are:
 
-  * polar coordinate glyph maps with \code{polar = TRUE}, 
-  * adjust glyph size with arguments \code{width} and \code{height},
-  * transformation relative to the all series (\code{global_rescale} defaults to \code{TRUE}) or each single series, and 
-  * reference boxes and lines with \code{geom_glyph_box()} and \code{geom_glyph_line()}.
+  * the implementation of a polar coordinate glyph maps with \code{polar = TRUE}, 
+  * the adjustment of the glyph size arguments using \code{width} and \code{height},
+  * a transformation relative to the all series (\code{global_rescale} defaults to \code{TRUE}) or each single series, and 
+  * the XX of the reference boxes and lines with \code{geom_glyph_box()} and \code{geom_glyph_line()}.
 

### 2023-06-29 cubble.tex
changes section 3.3

@@
 
-To communicate spatial and temporal information collectively in display, a few visualisation options are available: one can make faceted maps across time, creating map animations, or constructing interactive graphics to link between map and time series plot.
-Faceted maps and animation, while commonly used, may not be helpful for comparing across time since eyes need to jump across multiple facets or frames to locate the space before perceiving the changes in time.
-The glyph map \citep{Wickham2012-yr} resolves this issue by imposing the time series onto the map as a glyph through coordinate transformation.
-The transformation uses linear algebra to convert the temporal coordinates (minor coordinates) into the spatial coordinates (major coordinates) and is implemented in the package \texttt{GGally} \citep{ggally} package.
-The \pkg{cubble} package provides a ggproto implementation, \texttt{geom\_glyph()}, for glyph maps and it takes four required aesthetics: \texttt{x\_major}, \texttt{y\_major}, \texttt{x\_minor}, and \texttt{y\_minor}:
+Visualizing spatio-temporal data is important for exploring and understanding the data at hand, aiding in decision-making, and facilitating effective communication.
+To collectively visualize spatial and temporal information, several options are available such as: faceted maps across time, map animations, or interactive graphics that link maps and time series plots among others.
+Faceted maps and spatio-temporal animations, while commonly used, may be difficult to compare across time and space since users need to inspect multiple facets or frames to find changes in time or space.
+The glyph map \citep{Wickham2012-yr} resolves this issue by imposing the time series onto the map as a glyph through a coordinate transformation.
+The transformation uses linear algebra to convert the temporal coordinates (minor coordinates) into the spatial coordinates (major coordinates) and is implemented in the package \texttt{GGally} \citep{ggally}.
+The \pkg{cubble} package provides a ggproto implementation to create glyph maps, \texttt{geom\_glyph()} and it takes four required aestheticsf : \texttt{x\_major}, \texttt{y\_major}, \texttt{x\_minor}, and \texttt{y\_minor}:
 
@@
 
-Other useful controls over the glyph map includes:
+Other useful controls to modify the glyph map that can be include are:
 
@@
 \item
-  polar coordinate glyph maps with \code{polar = TRUE},
+  the implementation of a polar coordinate glyph maps with \code{polar = TRUE},
 \item
-  adjust glyph size with arguments \code{width} and \code{height},
+  the adjustment of the glyph size arguments using \code{width} and \code{height},
 \item
-  transformation relative to the all series (\code{global_rescale} defaults to \code{TRUE}) or each single series, and
+  a transformation relative to the all series (\code{global_rescale} defaults to \code{TRUE}) or each single series, and
 \item
-  reference boxes and lines with \code{geom_glyph_box()} and \code{geom_glyph_line()}.
+  the XX of the reference boxes and lines with \code{geom_glyph_box()} and \code{geom_glyph_line()}.
 \end{itemize}

### 2023-06-29 cubble.Rmd
Changes section 4.1

@@
 
-Five examples are chosen to illustrate different aspects of the \pkg{cubble} package: creating a \code{cubble} object from two Coronavirus (COVID) data tables with the complication of differing location names, using spatial transformations to make a glyph map of seasonal temperature changes, matching river level data and weather station records for analysis of water supply, reading NetCDF format data to reproduce a climate reanalysis plot, and the workflow to create complex interactive linked plots.
+Five examples are chosen to illustrate different aspects of the \pkg{cubble} package: creating a \code{cubble} object from two Coronavirus (COVID) data tables with the challenge of having different location names, using spatial transformations to make a glyph map of seasonal temperature changes, matching river level data with weather station records to analyze water supply, reading NetCDF format data to replicate a climate reanalysis plot, and demonstrating the workflow to create complex interactive linked plots.
 
@@
 
-Since the start of the pandemic, the Victoria State Government in Australia has been providing daily COVID counts per Local Government Area (LGA).
-This data can be combined with map polygon data, available from the Australian Bureau of Statistics (ABS), to visualize COVID incidence and spread.
-In the \pkg{cubble} package, the COVID count data (\code{covid}) and the LGA information (\code{lga}) are available as a \code{tsibble} object and an \code{sf} object, respectively.
+Since the start of the COVID-19 pandemic, the Victoria State Government in Australia has been providing daily COVID-19 case counts per Local Government Area (LGA).
+This data can be combined with map polygon data, available from the Australian Bureau of Statistics (ABS), to visualize COVID-19 incidence and spread.
+In the \pkg{cubble} package, the COVID-19 count data (\code{covid}) and the LGA information (\code{lga}) are available as a \code{tsibble} object and an \code{sf} object respectively.
 A cubble object can be created from these two component using \code{make_cubble()}.
-The \code{id} column containing LGA information has different names in the two data and can be specified with using the \code{by} argument:
+The \code{id} column containing LGA information has different names in the two data sets and that can be specified via the \code{by} argument:
 
@@
 
-The warning message suggests the slight difference of LGA encoding used by Victoria government and ABS and prompts analysts to check the mismatch using \code{check_key()}:
+The difference in LGA naming between both data sets triggers cubble to issue a warning, alerting the user to this discrepancy.
+The warning message suggests there are some differences between the LGA encoding used by Victoria government and ABS and prompts analysts to check the mismatch using \code{check_key()}:
 
@@
 
-The result of the \code{check_key()} function is a list containing three elements: 1) matched keys from both tables, 2) potentially paired keys, and 3) others.
+The result of the \code{check_key()} function is a list containing three elements: 1) matched keys from both tables, 2) potentially paired keys, and 3) others (XXX what is others exactly?).
 Here, the main mismatch arises from the two LGAs: Kingston and Latrobe (Kingston is a LGA in both Victoria and South Australia and Latrobe is a LGA in both Victoria and Tasmania).

### 2023-06-30 cubble.Rmd
small changes suggested by Ursula and Patriciag

@@
 In a spatial cubble object, spatial variables are organised as columns and temporal variables are nested within a specialised `ts` column.
-For example, the spatial cubble object, \code{cb_nested} printed below, contains weather records of three airport stations from the Global Historical Climatology Network Daily (GHCND) database.
+For example, the spatial cubble object, \code{cb_nested} printed below, contains weather records of three airport stations from the Global Historical Climatology Network Daily (GHCND) database [@menne2012overview].
 In this case, the spatial cubble is convenient for wrangling the spatial variables:
@@
 
-A cubble object inherits the attributes from tibble (and its subclasses): `class`, `row.names`, and `names`.
-Additionally, it has three specialised attributes: `key`, `index`, and `coords`.
-Readers who are familiar with the `key` and `index` attributes from the `tsibble` package would already understand the two arguments.
-In cubble, the `key` attribute identifies the row in the spatial cubble (given the internal use of `tidyr::nest()` for nesting), and when combined with the `index` argument, it identifies the row in the temporal cubble.
-Currently, cubble only supports one variable as the key, and the accepted temporal classes for `index` includes the base R classes `Date`, `POSIXlt`, `POSIXct`, as well as tsibble's `yearmonth`, `yearweek`, and `yearquarter` classes.
-The `coords` attribute represents an ordered pair of coordinates that can be either an unprojected pair of longitude and latitude, or a projected easting and northing value.
-The `sf` package is used under the hood to calculate the spatial bounding box, displayed in the header of a spatial cubble.
+A cubble object inherits the attributes from tibble (and its subclasses): \code{class}, \code{row.names}, and \code{names}.
+Additionally, it has three specialised attributes: \code{key}, \code{index}, and \code{coords}.
+\code{key} and \code{index} are used as introduced in the \pkg{tsibble} package [@tsibble].
+In cubble, the \code{key} attribute identifies the row in the spatial cubble (given the internal use of \code{tidyr::nest()} for nesting), and when combined with the \code{index} argument, it identifies the row in the temporal cubble.
+Currently, cubble only supports one variable as the key, and the accepted temporal classes for \code{index} includes the base R classes \code{Date}, \code{POSIXlt}, \code{POSIXct}, as well as tsibble's \code{yearmonth}, \code{yearweek}, and \code{yearquarter} classes.
+The \code{coords} attribute represents an ordered pair of coordinates that can be either an unprojected pair of longitude and latitude, or a projected easting and northing value.
+The \pkg{sf} package [@sf] is used under the hood to calculate the spatial bounding box, displayed in the header of a spatial cubble.
 
-The temporal cubble has a special attribute called `spatial` to store the spatial variables.
+The temporal cubble has a special attribute called \code{spatial} to store the spatial variables.
 Shortcut functions are available to extract attributes from the temporal cubble object, for example, \code{spatial()} for extracting spatial variables:
@@
 
-Other foreign spatio-temporal objects in \proglang{R} can be coerced into a \code{cubble} object with the function \code{as_cubble()}.
+Other \proglang{R} spatio-temporal objects can be coerced into a \code{cubble} object with the function \code{as_cubble()}.
 This includes a joined \code{tibble} or \code{data.frame}, a NetCDF object, a \code{stars} object [@stars], and a \code{sftime} object [@sftime].
@@
 In particular, the \code{cubble_df} class handles methods that behave consistently in both spatial and temporal cubble.
-When the method has a distinct behavior on the spatial and temporal cubble, it is implemented separately in \code{spatial_cubble_df} and \code{temporal_cubble_df}.
+When the method works differently internally on the spatial and temporal cubble, it is implemented separately in \code{spatial_cubble_df} and \code{temporal_cubble_df}.
 
@@
 The pair of cubble verbs, \code{face_temporal()} and \code{face_spatial()}, pivots the cubble object between its two forms or faces, as illustrated in Figure \ref{fig:face}.
-The code below connects a spatial cubble (\code{cb_nested}) and a temporal cubble (\code{cb_long}), introduced in Section \@ref(object), with \code{face_temporal()} and \code{face_spatial()}:
+The code applies \code{face_temporal()} on the spatial cubble (\code{cb_nested}) introduced in Section \@ref(object) to get a temporal cubble:
 
 ```{r cubble-fun, echo = TRUE}
-identical(face_temporal(cb_nested), cb_long)
-identical(face_spatial(cb_long), cb_nested)
+face_temporal(cb_nested)
 ```
@@
 Analysts often have their preferred spatial or temporal data structure for spatial or temporal analysis, which they may wish to continue using for spatio-temporal analysis.
-With \code{cubble}, analysts can incorporate the \code{tsibble} class [@tsibble] in a temporal cubble and the \code{sf} class [@sf] in a spatial cubble.
+With \code{cubble}, analysts can incorporate the \code{tsibble} class in a temporal cubble and the \code{sf} class in a spatial cubble.
 
@@
 
-A created temporal cubble can promote its temporal component to a \code{tsibble} object using \code{make_temporal_tsibble()}.
-See the code example below using the \code{cb_long} object created in Section \@ref(create):
+The temporal component of a created temporal cubble can be turned to a \code{tsibble} object using \code{make_temporal_tsibble()}.
+See the code example below using the \code{cb_long} object, created in Section \@ref(create):
 
@@
 
-Matching time series from an old list of stations to a new list is a common task in spatio-temporal data analysis.
+Matching time series from two lists is a common task in spatio-temporal data analysis.
 In cubble, matching based on distance and time series features can be performed using the functions `match_spatial()` and `match_temporal()`.
@@
 
-Two arguments are available to control the outputs: the argument `spatial_n_group` specifies the number of paired groups to output and the argument `spatial_n_each` specifies the number of each for each item in the first cubble object (default to 1 for one-to-one matching).
+Two arguments are available to control the matching: the argument `spatial_n_group` specifies the number of paired groups to return and the argument `spatial_n_each` specifies the number of matches for each item in the first cubble object (default to 1 for one-to-one matching).
 
@@
 The Bureau of Meteorology collects water level data and they can be matched with the precipitation data from the climate weather stations.
-The data `river` from the \pkg{cubble} package contains water course level data for 71 river gauges collected in Victoria, Australia.
-Victoria weather station data can be subsetted from the `climate_aus` data in \pkg{cubble}.
+The data \code{river} from the \pkg{cubble} package contains water course level data for 71 river gauges collected in Victoria, Australia.
+Victoria weather station data can be subsetted from the \code{climate_aus} data in \pkg{cubble}.
 
@@
 The result can also be returned as a list of matched cubbles, by setting the argument \code{return_cubble = TRUE}.
-All the results can be combined into a single cubble using \code{bind_rows()}, after excluding the two pairs where a river station is matched to more than one weather stations (river station `226027` is matched twice in group 3 and 5 and similarly for station `230200` in group 7 and 8).
+All the results can be combined into a single cubble using \code{bind_rows()}, after excluding the two pairs where a river station is matched to more than one weather stations (river station 226027 is matched twice in group 3 and 5 and similarly for station 230200 in group 7 and 8).
 
@@
 
-Similarly, the cubble output can be returned using the argument `return_cubble = TRUE`.
+Similarly, the cubble output can be returned using the argument \code{return_cubble = TRUE}.
 Here we select the four pairs with the highest number of matching peaks and show them on the map (a) and as standardized series (b) in Figure \ref{fig:matching}.

### 2023-06-30 cubble.tex
small changes suggested by Ursula and Patriciag

@@
 In a spatial cubble object, spatial variables are organised as columns and temporal variables are nested within a specialised \texttt{ts} column.
-For example, the spatial cubble object, \code{cb_nested} printed below, contains weather records of three airport stations from the Global Historical Climatology Network Daily (GHCND) database.
+For example, the spatial cubble object, \code{cb_nested} printed below, contains weather records of three airport stations from the Global Historical Climatology Network Daily (GHCND) database \citep{menne2012overview}.
 In this case, the spatial cubble is convenient for wrangling the spatial variables:
@@
 
-A cubble object inherits the attributes from tibble (and its subclasses): \texttt{class}, \texttt{row.names}, and \texttt{names}.
-Additionally, it has three specialised attributes: \texttt{key}, \texttt{index}, and \texttt{coords}.
-Readers who are familiar with the \texttt{key} and \texttt{index} attributes from the \texttt{tsibble} package would already understand the two arguments.
-In cubble, the \texttt{key} attribute identifies the row in the spatial cubble (given the internal use of \texttt{tidyr::nest()} for nesting), and when combined with the \texttt{index} argument, it identifies the row in the temporal cubble.
-Currently, cubble only supports one variable as the key, and the accepted temporal classes for \texttt{index} includes the base R classes \texttt{Date}, \texttt{POSIXlt}, \texttt{POSIXct}, as well as tsibble's \texttt{yearmonth}, \texttt{yearweek}, and \texttt{yearquarter} classes.
-The \texttt{coords} attribute represents an ordered pair of coordinates that can be either an unprojected pair of longitude and latitude, or a projected easting and northing value.
-The \texttt{sf} package is used under the hood to calculate the spatial bounding box, displayed in the header of a spatial cubble.
-
-The temporal cubble has a special attribute called \texttt{spatial} to store the spatial variables.
+A cubble object inherits the attributes from tibble (and its subclasses): \code{class}, \code{row.names}, and \code{names}.
+Additionally, it has three specialised attributes: \code{key}, \code{index}, and \code{coords}.
+\code{key} and \code{index} are used as introduced in the \pkg{tsibble} package \citep{tsibble}.
+In cubble, the \code{key} attribute identifies the row in the spatial cubble (given the internal use of \code{tidyr::nest()} for nesting), and when combined with the \code{index} argument, it identifies the row in the temporal cubble.
+Currently, cubble only supports one variable as the key, and the accepted temporal classes for \code{index} includes the base R classes \code{Date}, \code{POSIXlt}, \code{POSIXct}, as well as tsibble's \code{yearmonth}, \code{yearweek}, and \code{yearquarter} classes.
+The \code{coords} attribute represents an ordered pair of coordinates that can be either an unprojected pair of longitude and latitude, or a projected easting and northing value.
+The \pkg{sf} package \citep{sf} is used under the hood to calculate the spatial bounding box, displayed in the header of a spatial cubble.
+
+The temporal cubble has a special attribute called \code{spatial} to store the spatial variables.
 Shortcut functions are available to extract attributes from the temporal cubble object, for example, \code{spatial()} for extracting spatial variables:
@@
 
-Other foreign spatio-temporal objects in \proglang{R} can be coerced into a \code{cubble} object with the function \code{as_cubble()}.
+Other \proglang{R} spatio-temporal objects can be coerced into a \code{cubble} object with the function \code{as_cubble()}.
 This includes a joined \code{tibble} or \code{data.frame}, a NetCDF object, a \code{stars} object \citep{stars}, and a \code{sftime} object \citep{sftime}.
@@
 The pair of cubble verbs, \code{face_temporal()} and \code{face_spatial()}, pivots the cubble object between its two forms or faces, as illustrated in Figure \ref{fig:face}.
-The code below connects a spatial cubble (\code{cb_nested}) and a temporal cubble (\code{cb_long}), introduced in Section \ref{object}, with \code{face_temporal()} and \code{face_spatial()}:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> identical(face_temporal(cb_nested), cb_long)
+The code applies \code{face_temporal()} on the spatial cubble (\code{cb_nested}) introduced in Section \ref{object} to get a temporal cubble:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> face_temporal(cb_nested)
+\end{CodeInput}
+\begin{CodeOutput}
+# cubble:   key: id [3], index: date, long form
+# temporal: 2020-01-01 -- 2020-01-10 [1D], no gaps
+# spatial:  long [dbl], lat [dbl], elev [dbl], name [chr], wmo_id [dbl]
+  id          date        prcp  tmax  tmin
+  <chr>       <date>     <dbl> <dbl> <dbl>
+1 ASN00086038 2020-01-01     0  26.8  11  
+2 ASN00086038 2020-01-02     0  26.3  12.2
+3 ASN00086038 2020-01-03     0  34.5  12.7
+4 ASN00086038 2020-01-04     0  29.3  18.8
+5 ASN00086038 2020-01-05    18  16.1  12.5
+# i 25 more rows
+\end{CodeOutput}
+\end{CodeChunk}
+
+Both verbs are the exact inverse of each other and apply both functions on a cubble object will result in the object itself:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> identical(face_spatial(face_temporal(cb_nested)), cb_nested)
 \end{CodeInput}
@@
 \begin{CodeInput}
-R> identical(face_spatial(cb_long), cb_nested)
+R> identical(face_temporal(face_spatial(cb_long)), cb_long)
 \end{CodeInput}
@@
 
-Both verbs are the exact inverse of each other and apply both functions on a cubble object will result in the object itself:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> identical(face_spatial(face_temporal(cb_nested)), cb_nested)
-\end{CodeInput}
-\begin{CodeOutput}
-[1] TRUE
-\end{CodeOutput}
-\begin{CodeInput}
-R> identical(face_temporal(face_spatial(cb_long)), cb_long)
-\end{CodeInput}
-\begin{CodeOutput}
-[1] TRUE
-\end{CodeOutput}
-\end{CodeChunk}
-
 \begin{CodeChunk}
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Desktop/paper-cubble/figures/diagram-keynotes/diagram-keynotes.001}
+{\centering \includegraphics[width=1\linewidth]{figures/diagram-keynotes/diagram-keynotes.001}
 
@@
 Analysts often have their preferred spatial or temporal data structure for spatial or temporal analysis, which they may wish to continue using for spatio-temporal analysis.
-With \code{cubble}, analysts can incorporate the \code{tsibble} class \citep{tsibble} in a temporal cubble and the \code{sf} class \citep{sf} in a spatial cubble.
+With \code{cubble}, analysts can incorporate the \code{tsibble} class in a temporal cubble and the \code{sf} class in a spatial cubble.
 
@@
 
-A created temporal cubble can promote its temporal component to a \code{tsibble} object using \code{make_temporal_tsibble()}.
-See the code example below using the \code{cb_long} object created in Section \ref{create}:
+The temporal component of a created temporal cubble can be turned to a \code{tsibble} object using \code{make_temporal_tsibble()}.
+See the code example below using the \code{cb_long} object, created in Section \ref{create}:
 
@@
 
-Matching time series from an old list of stations to a new list is a common task in spatio-temporal data analysis.
+Matching time series from two lists is a common task in spatio-temporal data analysis.
 In cubble, matching based on distance and time series features can be performed using the functions \texttt{match\_spatial()} and \texttt{match\_temporal()}.
@@
 
-Two arguments are available to control the outputs: the argument \texttt{spatial\_n\_group} specifies the number of paired groups to output and the argument \texttt{spatial\_n\_each} specifies the number of each for each item in the first cubble object (default to 1 for one-to-one matching).
+Two arguments are available to control the matching: the argument \texttt{spatial\_n\_group} specifies the number of paired groups to return and the argument \texttt{spatial\_n\_each} specifies the number of matches for each item in the first cubble object (default to 1 for one-to-one matching).
 
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.35\textheight]{/Users/pmen0008/Desktop/paper-cubble/figures/diagram-keynotes/‎diagram-keynotes.‎002}
+{\centering \includegraphics[width=1\linewidth,height=0.35\textheight]{figures/diagram-keynotes/‎diagram-keynotes.‎002}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Desktop/paper-cubble/figures/glyphmap-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/research/paper-cubble/figures/glyphmap-1}
 
@@
 The Bureau of Meteorology collects water level data and they can be matched with the precipitation data from the climate weather stations.
-The data \texttt{river} from the \pkg{cubble} package contains water course level data for 71 river gauges collected in Victoria, Australia.
-Victoria weather station data can be subsetted from the \texttt{climate\_aus} data in \pkg{cubble}.
+The data \code{river} from the \pkg{cubble} package contains water course level data for 71 river gauges collected in Victoria, Australia.
+Victoria weather station data can be subsetted from the \code{climate_aus} data in \pkg{cubble}.
 
@@
 The result can also be returned as a list of matched cubbles, by setting the argument \code{return_cubble = TRUE}.
-All the results can be combined into a single cubble using \code{bind_rows()}, after excluding the two pairs where a river station is matched to more than one weather stations (river station \texttt{226027} is matched twice in group 3 and 5 and similarly for station \texttt{230200} in group 7 and 8).
+All the results can be combined into a single cubble using \code{bind_rows()}, after excluding the two pairs where a river station is matched to more than one weather stations (river station 226027 is matched twice in group 3 and 5 and similarly for station 230200 in group 7 and 8).
 
@@
 
-Similarly, the cubble output can be returned using the argument \texttt{return\_cubble\ =\ TRUE}.
+Similarly, the cubble output can be returned using the argument \code{return_cubble = TRUE}.
 Here we select the four pairs with the highest number of matching peaks and show them on the map (a) and as standardized series (b) in Figure \ref{fig:matching}.
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Desktop/paper-cubble/figures/matching-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/research/paper-cubble/figures/matching-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Desktop/paper-cubble/figures/netcdf-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/research/paper-cubble/figures/netcdf-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/pmen0008/Desktop/paper-cubble/figures/linking} \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/pmen0008/Desktop/paper-cubble/figures/linking-north} \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/pmen0008/Desktop/paper-cubble/figures/linking-lower}
+{\centering \includegraphics[width=1\linewidth,height=0.23\textheight]{figures/linking} \includegraphics[width=1\linewidth,height=0.23\textheight]{figures/linking-north} \includegraphics[width=1\linewidth,height=0.23\textheight]{figures/linking-lower}
 

### 2023-07-02 cubble.Rmd
a paragraph to give an overview to Section 2

@@
 
+<!-- In this paper, we propose the \pkg{cubble} package to organise spatio-temporal data collected at unique fixed locations while allowing for irregularities in the temporal dimension, such as weather station data.
+-->
+
+<!-- Spatio-temporal data can encompass data with various spatial and temporal characteristics, requiring different structures for wrangling and analysis.
+For example, climate weather stations typically store station metadata in one table and the climate time series in another.
+GPS data tracks unique point locations at different timestamps and is represented as trajectories.
+Satellite imageries capture snapshots of landscapes at selected times and is commonly structured as raster data.
+-->
+
 The rest of the paper is organized as follows: Section \@ref(cubble) presents the main design and functionality of the \pkg{cubble} package.
@@
 
-## The cubble object {#object}
+The cubble class includes two subclasses: the spatial cubble and the temporal cubble, which can be pivoted back and forth to focus on the two aspects of the spatio-temporal data, as illustrated in Figure \ref{fig:face}.
+This section provides an overview of the \pkg{cubble} package, including the cubble class and its attributes, class creation and coercion, a summary of implemented functionality, the compatibility with other spatial and temporal packages (\pkg{sf} and \pkg{tsibble}), and a comparison with other spatio-temporal packages (\pkg{stars} and \pkg{sftime}).
 
-Spatio-temporal data can encompass data with various spatial and temporal characteristics, requiring different structures for wrangling and analysis.
-For example, climate weather stations typically store station metadata in one table and the climate time series in another.
-GPS data tracks unique point locations at different timestamps and is represented as trajectories.
-Satellite imageries capture snapshots of landscapes at selected times and is commonly structured as raster data.
-In this paper, we propose the \pkg{cubble} package to organise spatio-temporal data collected at unique fixed locations while allowing for irregularities in the temporal dimension, such as weather station data.
+## The cubble class {#object}
 

### 2023-07-02 cubble.tex
a paragraph to give an overview to Section 2

@@
 
+The cubble class includes two subclasses: the spatial cubble and the temporal cubble, which can be pivoted back and forth to focus on the two aspects of the spatio-temporal data, as illustrated in Figure \ref{fig:face}.
+This section provides an overview of the \pkg{cubble} package, including the cubble class and its attributes, class creation and coercion, a summary of implemented functionality, the compatibility with other spatial and temporal packages (\pkg{sf} and \pkg{tsibble}), and a comparison with other spatio-temporal packages (\pkg{stars} and \pkg{sftime}).
+
 \hypertarget{object}{%
-\subsection{The cubble object}\label{object}}
-
-Spatio-temporal data can encompass data with various spatial and temporal characteristics, requiring different structures for wrangling and analysis.
-For example, climate weather stations typically store station metadata in one table and the climate time series in another.
-GPS data tracks unique point locations at different timestamps and is represented as trajectories.
-Satellite imageries capture snapshots of landscapes at selected times and is commonly structured as raster data.
-In this paper, we propose the \pkg{cubble} package to organise spatio-temporal data collected at unique fixed locations while allowing for irregularities in the temporal dimension, such as weather station data.
+\subsection{The cubble class}\label{object}}
 

### 2023-07-02 cubble.Rmd
change cb_nested -> cb_spatial,cb_long -> cb_temporal, and alike

@@
 In a spatial cubble object, spatial variables are organised as columns and temporal variables are nested within a specialised `ts` column.
-For example, the spatial cubble object, \code{cb_nested} printed below, contains weather records of three airport stations from the Global Historical Climatology Network Daily (GHCND) database [@menne2012overview].
+For example, the spatial cubble object, \code{cb_spatial} printed below, contains weather records of three airport stations from the Global Historical Climatology Network Daily (GHCND) database [@menne2012overview].
 In this case, the spatial cubble is convenient for wrangling the spatial variables:
@@
 ```{r}
-cb_nested <- climate_mel
+cb_spatial <- climate_mel
 ```
@@
 ```{r echo = TRUE}
-cb_nested
+cb_spatial
 ```
@@
 In a temporal cubble, temporal variables are expanded in the long form and spatial variables are stored as a data attribute.
-The temporal cubble object, \code{cb_long}, contains the same spatio-temporal data as the spatial cubble object, \code{cb_nested}, but in a structure that is easier for temporal analysis:
+The temporal cubble object, \code{cb_temporal}, contains the same spatio-temporal data as the spatial cubble object, \code{cb_spatial}, but in a structure that is easier for temporal analysis:
 
 ```{r}
-cb_long <- face_temporal(climate_mel)
+cb_temporal <- face_temporal(climate_mel)
 ```
@@
 ```{r echo = TRUE}
-cb_long
+cb_temporal
 ```
@@
 ```{r echo = TRUE}
-spatial(cb_long)
+spatial(cb_temporal)
 ```
@@
 The pair of cubble verbs, \code{face_temporal()} and \code{face_spatial()}, pivots the cubble object between its two forms or faces, as illustrated in Figure \ref{fig:face}.
-The code applies \code{face_temporal()} on the spatial cubble (\code{cb_nested}) introduced in Section \@ref(object) to get a temporal cubble:
+The code applies \code{face_temporal()} on the spatial cubble (\code{cb_spatial}) introduced in Section \@ref(object) to get a temporal cubble:
 
 ```{r cubble-fun, echo = TRUE}
-face_temporal(cb_nested)
+face_temporal(cb_spatial)
 ```
@@
 ```{r cubble-fun2, echo = TRUE}
-identical(face_spatial(face_temporal(cb_nested)), cb_nested)
-identical(face_temporal(face_spatial(cb_long)), cb_long)
+identical(face_spatial(face_temporal(cb_spatial)), cb_spatial)
+identical(face_temporal(face_spatial(cb_temporal)), cb_temporal)
 ```
@@
 ```{r echo = TRUE}
-ts_nested <- make_cubble(
+ts_spatial <- make_cubble(
   spatial = stations, temporal = meteo_ts, coords = c(long, lat))
-(ts_long <- face_temporal(ts_nested))
-class(ts_long)
+(ts_temporal <- face_temporal(ts_spatial))
+class(ts_temporal)
 ```
@@
 ```{r echo = TRUE}
-ts_long |> has_gaps()
+ts_temporal |> has_gaps()
 ```
@@
 The temporal component of a created temporal cubble can be turned to a \code{tsibble} object using \code{make_temporal_tsibble()}.
-See the code example below using the \code{cb_long} object, created in Section \@ref(create):
+See the code example below using the \code{cb_temporal} object, created in Section \@ref(create):
 
 ```{r echo = TRUE}
-cb_long |> make_temporal_tsibble() 
+cb_temporal |> make_temporal_tsibble() 
 ```
@@
 ```{r echo = TRUE}
-(sf_nested <- make_cubble(
+(sf_spatial <- make_cubble(
   spatial = stations_sf, temporal = meteo, 
   key = id, index = date))
-class(sf_nested)
+class(sf_spatial)
 ```
@@
 ```{r echo =TRUE, message=FALSE}
-sf_nested |> sf::st_transform(crs = "EPSG:3857")
+sf_spatial |> sf::st_transform(crs = "EPSG:3857")
 ```
@@
 ```{r echo = TRUE}
-cb_nested |> make_spatial_sf() 
+cb_spatial |> make_spatial_sf() 
 ```
@@
 ````
-nested <- clean |> SharedData$new(~id, group = "cubble")
+sd_spatial <- clean |> SharedData$new(~id, group = "cubble")
 
-long <- clean |> 
+sd_temporal <- clean |> 
   face_temporal() |> 
@@
 
-p1 <- nested |> ggplot() + ...
-p2 <- long |> ggplot() + ...
+p1 <- sd_spatial |> ggplot() + ...
+p2 <- sd_temporal |> ggplot() + ...
 crosstalk::bscols(plotly::ggplotly(p1), plotly::ggplotly(p2), ...)

### 2023-07-02 cubble.Rmd
make it clearer the attributes apply to both cubble subclasses

@@
 
-A cubble object inherits the attributes from tibble (and its subclasses): \code{class}, \code{row.names}, and \code{names}.
-Additionally, it has three specialised attributes: \code{key}, \code{index}, and \code{coords}.
+Both cubble objects inherit the attributes \code{class}, \code{row.names}, and \code{names} from tibble, which originates from data frames.
+Additionally, both have three specialised attributes: \code{key}, \code{index}, and \code{coords}.
 \code{key} and \code{index} are used as introduced in the \pkg{tsibble} package [@tsibble].
@@
 The \code{coords} attribute represents an ordered pair of coordinates that can be either an unprojected pair of longitude and latitude, or a projected easting and northing value.
-The \pkg{sf} package [@sf] is used under the hood to calculate the spatial bounding box, displayed in the header of a spatial cubble.
-
-The temporal cubble has a special attribute called \code{spatial} to store the spatial variables.
+Moreover, temporal cubbles have a special attribute called \code{spatial} to store the spatial variables.
 Shortcut functions are available to extract attributes from the temporal cubble object, for example, \code{spatial()} for extracting spatial variables:

### 2023-07-02 cubble.Rmd
more explanation to open Section 2.2

@@
 
-Both cubble objects inherit the attributes \code{class}, \code{row.names}, and \code{names} from tibble, which originates from data frames.
+Both cubble objects inherit tibble's attributes (which originates from data frames): \code{class}, \code{row.names}, and \code{names}.
 Additionally, both have three specialised attributes: \code{key}, \code{index}, and \code{coords}.
@@
 
-<!-- In many cases, spatio-temporal data arrive in separate tables for analysis.
-For example, in climate data, analysts may initially receive station data containing geographic location information, recorded variables and their recording periods.
-They can then query the temporal variables using the stations of interest to obtain the relevant temporal data.
-Alternatively, analyses may begin as purely spatial or temporal, and analysts may obtain additional temporal or spatial data to expand the result to spatio-temporal.
--->
-
-The function \code{make_cubble()} composes a spatial cubble object from a spatial table (\code{spatial}) and a temporal table (\code{temporal}), along with three attributes introduced in the subsection \@ref(object): \code{key}, \code{index}, and \code{coords}.
+The spatial and temporal aspect of spatio-temporal data are often stored separately in the database.
+For climate data, analysts may initially receive station metadata and then query the time series based on the metadata.
+A (spatial) cubble object can be constructed from separate spatial and temporal tables using the function \code{make_cubble()}.
+The three attributes \code{key}, \code{index}, and \code{coords} need to be specified in creation.
 The following code creates a spatial cubble from its spatial component, \code{stations} and temporal component \code{meteo}:

### 2023-07-02 cubble.tex
more explanation to open Section 2.2

@@
 In a spatial cubble object, spatial variables are organised as columns and temporal variables are nested within a specialised \texttt{ts} column.
-For example, the spatial cubble object, \code{cb_nested} printed below, contains weather records of three airport stations from the Global Historical Climatology Network Daily (GHCND) database \citep{menne2012overview}.
+For example, the spatial cubble object, \code{cb_spatial} printed below, contains weather records of three airport stations from the Global Historical Climatology Network Daily (GHCND) database \citep{menne2012overview}.
 In this case, the spatial cubble is convenient for wrangling the spatial variables:
@@
 \begin{CodeInput}
-R> cb_nested
+R> cb_spatial
 \end{CodeInput}
@@
 In a temporal cubble, temporal variables are expanded in the long form and spatial variables are stored as a data attribute.
-The temporal cubble object, \code{cb_long}, contains the same spatio-temporal data as the spatial cubble object, \code{cb_nested}, but in a structure that is easier for temporal analysis:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> cb_long
+The temporal cubble object, \code{cb_temporal}, contains the same spatio-temporal data as the spatial cubble object, \code{cb_spatial}, but in a structure that is easier for temporal analysis:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> cb_temporal
 \end{CodeInput}
@@
 
-A cubble object inherits the attributes from tibble (and its subclasses): \code{class}, \code{row.names}, and \code{names}.
-Additionally, it has three specialised attributes: \code{key}, \code{index}, and \code{coords}.
+Both cubble objects inherit tibble's attributes (which originates from data frames): \code{class}, \code{row.names}, and \code{names}.
+Additionally, both have three specialised attributes: \code{key}, \code{index}, and \code{coords}.
 \code{key} and \code{index} are used as introduced in the \pkg{tsibble} package \citep{tsibble}.
@@
 The \code{coords} attribute represents an ordered pair of coordinates that can be either an unprojected pair of longitude and latitude, or a projected easting and northing value.
-The \pkg{sf} package \citep{sf} is used under the hood to calculate the spatial bounding box, displayed in the header of a spatial cubble.
-
-The temporal cubble has a special attribute called \code{spatial} to store the spatial variables.
+Moreover, temporal cubbles have a special attribute called \code{spatial} to store the spatial variables.
 Shortcut functions are available to extract attributes from the temporal cubble object, for example, \code{spatial()} for extracting spatial variables:
@@
 \begin{CodeInput}
-R> spatial(cb_long)
+R> spatial(cb_temporal)
 \end{CodeInput}
@@
 
-The function \code{make_cubble()} composes a spatial cubble object from a spatial table (\code{spatial}) and a temporal table (\code{temporal}), along with three attributes introduced in the subsection \ref{object}: \code{key}, \code{index}, and \code{coords}.
+The spatial and temporal aspect of spatio-temporal data are often stored separately in the database.
+For climate data, analysts may initially receive station metadata and then query the time series based on the metadata.
+A (spatial) cubble object can be constructed from separate spatial and temporal tables using the function \code{make_cubble()}.
+The three attributes \code{key}, \code{index}, and \code{coords} need to be specified in creation.
 The following code creates a spatial cubble from its spatial component, \code{stations} and temporal component \code{meteo}:
@@
 The pair of cubble verbs, \code{face_temporal()} and \code{face_spatial()}, pivots the cubble object between its two forms or faces, as illustrated in Figure \ref{fig:face}.
-The code applies \code{face_temporal()} on the spatial cubble (\code{cb_nested}) introduced in Section \ref{object} to get a temporal cubble:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> face_temporal(cb_nested)
+The code applies \code{face_temporal()} on the spatial cubble (\code{cb_spatial}) introduced in Section \ref{object} to get a temporal cubble:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> face_temporal(cb_spatial)
 \end{CodeInput}
@@
 \begin{CodeInput}
-R> identical(face_spatial(face_temporal(cb_nested)), cb_nested)
+R> identical(face_spatial(face_temporal(cb_spatial)), cb_spatial)
 \end{CodeInput}
@@
 \begin{CodeInput}
-R> identical(face_temporal(face_spatial(cb_long)), cb_long)
+R> identical(face_temporal(face_spatial(cb_temporal)), cb_temporal)
 \end{CodeInput}
@@
 \begin{CodeInput}
-R> ts_nested <- make_cubble(
+R> ts_spatial <- make_cubble(
 +   spatial = stations, temporal = meteo_ts, coords = c(long, lat))
-R> (ts_long <- face_temporal(ts_nested))
+R> (ts_temporal <- face_temporal(ts_spatial))
 \end{CodeInput}
@@
 \begin{CodeInput}
-R> class(ts_long)
+R> class(ts_temporal)
 \end{CodeInput}
@@
 \begin{CodeInput}
-R> ts_long |> has_gaps()
+R> ts_temporal |> has_gaps()
 \end{CodeInput}
@@
 The temporal component of a created temporal cubble can be turned to a \code{tsibble} object using \code{make_temporal_tsibble()}.
-See the code example below using the \code{cb_long} object, created in Section \ref{create}:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> cb_long |> make_temporal_tsibble() 
+See the code example below using the \code{cb_temporal} object, created in Section \ref{create}:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> cb_temporal |> make_temporal_tsibble() 
 \end{CodeInput}
@@
 \begin{CodeInput}
-R> (sf_nested <- make_cubble(
+R> (sf_spatial <- make_cubble(
 +   spatial = stations_sf, temporal = meteo, 
@@
 \begin{CodeInput}
-R> class(sf_nested)
+R> class(sf_spatial)
 \end{CodeInput}
@@
 \begin{CodeInput}
-R> sf_nested |> sf::st_transform(crs = "EPSG:3857")
+R> sf_spatial |> sf::st_transform(crs = "EPSG:3857")
 \end{CodeInput}
@@
 \begin{CodeInput}
-R> cb_nested |> make_spatial_sf() 
+R> cb_spatial |> make_spatial_sf() 
 \end{CodeInput}
@@
 \begin{verbatim}
-nested <- clean |> SharedData$new(~id, group = "cubble")
-
-long <- clean |> 
+sd_spatial <- clean |> SharedData$new(~id, group = "cubble")
+
+sd_temporal <- clean |> 
   face_temporal() |> 

### 2023-07-02 cubble.Rmd
introduce unfold and check_key

@@
 ```{r cubble-fun2, echo = TRUE}
-identical(face_spatial(face_temporal(cb_spatial)), cb_spatial)
-identical(face_temporal(face_spatial(cb_temporal)), cb_temporal)
+face_spatial(face_temporal(cb_spatial))
 ```
@@
 knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.001.png"))
+```
+
+To enable operations involve both spatial and temporal variables, the function \code{unfold} incorporates spatial variables into the temporal cubble.
+Below is an example to include the coordinate columns (\code{long} and \code{lat}) into \code{cb_temporal} to prepare the data for a glyph map transformation, which will be discussed in Section \ref{st_transformation}.
+
+```{r echo = TRUE}
+cb_temporal |> unfold(long, lat)
 ```
@@
 The difference in LGA naming between both data sets triggers cubble to issue a warning, alerting the user to this discrepancy.
-The warning message suggests there are some differences between the LGA encoding used by Victoria government and ABS and prompts analysts to check the mismatch using \code{check_key()}:
+The warning message suggests there are some differences between the LGA encoding used by Victoria government and ABS and prompts analysts to check the mismatch using \code{check_key()}, which takes the same inputs as \code{make_cubble}, but returns a summary of key matches between the spatial and temporal input data.
 
@@
 ```{r echo = TRUE}
-res_sp <- match_spatial(climate_vic, river, spatial_n_group = 10)
+res_sp <- match_spatial(df1 = climate_vic, df2 = river, 
+                        spatial_n_group = 10)
 print(res_sp, n = 20)
@@
 res_sp <- match_spatial(
-  climate_vic, river, 
+  df1 = climate_vic, df2 = river, 
   spatial_n_group = 10, return_cubble = TRUE)
@@
 ```{r echo = TRUE}
-(res_tm <- res_sp |> 
-  match_temporal(
+(res_tm <- match_temporal(data = res_sp,
                           data_id = type, match_id = group,
@@
 ```{r echo = TRUE}
-res_tm <- res_sp |> 
-  match_temporal(
+res_tm <- match_temporal(data =  res_sp,
                          data_id = type, match_id = group,

### 2023-07-02 cubble.Rmd
add a sentence  to highlight how cubble is used in each example

@@
 This data can be combined with map polygon data, available from the Australian Bureau of Statistics (ABS), to visualize COVID-19 incidence and spread.
-In the \pkg{cubble} package, the COVID-19 count data (\code{covid}) and the LGA information (\code{lga}) are available as a \code{tsibble} object and an \code{sf} object respectively.
-A cubble object can be created from these two component using \code{make_cubble()}.
-The \code{id} column containing LGA information has different names in the two data sets and that can be specified via the \code{by} argument:
+The COVID-19 count data (\code{covid}) and the LGA information (\code{lga}) are available in the \pkg{cubble} package as a \code{tsibble} object and an \code{sf} object, respectively.
+This example illustrates the creation of a cubble object using real-world data, where analysts may deal with slightly varied encoding of the spatial units used across different institutes.
+Cubble can flag this discrepancies when creating the object and guide analysts to further investigate this issue.
+
+The function \code{make_cubble()} is used to create a cubble object from separate spatial and temporal components.
+When the spatial identifier in the two data has different names, they can be handled by the \code{by} argument:
 
@@
 
-The result of the \code{check_key()} function is a list containing three elements: 1) matched keys from both tables, 2) potentially paired keys, and 3) others (XXX what is others exactly?).
+The result of the \code{check_key()} function is a list containing three elements: 1) matched keys from both tables, 2) potentially paired keys, and 3) others keys that can't be matched.
 Here, the main mismatch arises from the two LGAs: Kingston and Latrobe (Kingston is a LGA in both Victoria and South Australia and Latrobe is a LGA in both Victoria and Tasmania).
@@
 In the \pkg{cubble} package, the cubble object \code{historical_tmax} contains daily maximum temperature data for `r nrow(historical_tmax)` stations in Australia, covering two periods: 1971-1975 and 2016-2020.
-This example uses glyph maps to compare the changes in temperature between these two periods.
+This example uses dplyr verbs to wrangle a cubble object, which is pivoted between the spatial and temporal form during the analysis.
+Glyph maps are then created with ggplot2 to compare the changes in temperature between these two periods.
 
@@
 The data \code{river} from the \pkg{cubble} package contains water course level data for 71 river gauges collected in Victoria, Australia.
-Victoria weather station data can be subsetted from the \code{climate_aus} data in \pkg{cubble}.
+Victoria weather station data can be subsetted from the \code{climate_aus} data in the \pkg{cubble} package.
+The example demonstrates the use of matching function introduced in Section \ref{matching} to detect river gauges that can reflects precipitation in Victoria.
 
@@
 The ERA5 reanalysis [@hersbach2020era5] provides hourly estimates of atmospheric, land and oceanic climate variables on a global scale and is available in the NetCDF format from Copernicus Climate Data Store (CDS).
-This example reproduces Figure 19 in the @hersbach2020era5 paper with cubble, in Figure \ref{fig:netcdf}.
+This example demonstrates a case of analysing raster spatio-temporal data using cubble, replicating Figure 19 from the @hersbach2020era5 paper.
 The plot shows the southern polar vortex splitting into two on 2002-09-26, and further splitting into four on 2002-10-04.
@@
 con <- rnaturalearth::ne_coastline("small", returnclass = "sf")
-box <- st_bbox(c(xmin = -180, ymin = -90, xmax = 180, ymax = -15), crs = st_crs(con)) 
+box <- st_bbox(c(xmin = -180, ymin = -90, xmax = 180, ymax = -15), 
+               crs = st_crs(con)) 
 country <- con |> 

### 2023-07-02 cubble.tex
add a sentence  to highlight how cubble is used in each example

@@
 \begin{CodeInput}
-R> identical(face_spatial(face_temporal(cb_spatial)), cb_spatial)
-\end{CodeInput}
-\begin{CodeOutput}
-[1] TRUE
-\end{CodeOutput}
-\begin{CodeInput}
-R> identical(face_temporal(face_spatial(cb_temporal)), cb_temporal)
-\end{CodeInput}
-\begin{CodeOutput}
-[1] TRUE
+R> face_spatial(face_temporal(cb_spatial))
+\end{CodeInput}
+\begin{CodeOutput}
+# cubble:   key: id [3], index: date, nested form
+# spatial:  [144.8321, -37.98, 145.0964, -37.6655], Missing CRS!
+# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
+  id           long   lat  elev name              wmo_id ts               
+  <chr>       <dbl> <dbl> <dbl> <chr>              <dbl> <list>           
+1 ASN00086038  145. -37.7  78.4 essendon airport   95866 <tibble [10 x 4]>
+2 ASN00086077  145. -38.0  12.1 moorabbin airport  94870 <tibble [10 x 4]>
+3 ASN00086282  145. -37.7 113.  melbourne airport  94866 <tibble [10 x 4]>
 \end{CodeOutput}
@@
 \end{figure}
+\end{CodeChunk}
+
+To enable operations involve both spatial and temporal variables, the function \code{unfold} incorporates spatial variables into the temporal cubble.
+Below is an example to include the coordinate columns (\code{long} and \code{lat}) into \code{cb_temporal} to prepare the data for a glyph map transformation, which will be discussed in Section \ref{st_transformation}.
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> cb_temporal |> unfold(long, lat)
+\end{CodeInput}
+\begin{CodeOutput}
+# cubble:   key: id [3], index: date, long form
+# temporal: 2020-01-01 -- 2020-01-10 [1D], no gaps
+# spatial:  long [dbl], lat [dbl], elev [dbl], name [chr], wmo_id [dbl]
+  id          date        prcp  tmax  tmin  long   lat
+  <chr>       <date>     <dbl> <dbl> <dbl> <dbl> <dbl>
+1 ASN00086038 2020-01-01     0  26.8  11    145. -37.7
+2 ASN00086038 2020-01-02     0  26.3  12.2  145. -37.7
+3 ASN00086038 2020-01-03     0  34.5  12.7  145. -37.7
+4 ASN00086038 2020-01-04     0  29.3  18.8  145. -37.7
+5 ASN00086038 2020-01-05    18  16.1  12.5  145. -37.7
+# i 25 more rows
+\end{CodeOutput}
 \end{CodeChunk}
@@
 In the \pkg{cubble} package, the cubble object \code{historical_tmax} contains daily maximum temperature data for 75 stations in Australia, covering two periods: 1971-1975 and 2016-2020.
-This example uses glyph maps to compare the changes in temperature between these two periods.
+This example uses dplyr verbs to wrangle a cubble object, which is pivoted between the spatial and temporal form during the analysis.
+Glyph maps are then created with ggplot2 to compare the changes in temperature between these two periods.
 
@@
 \begin{CodeInput}
-R> res_sp <- match_spatial(climate_vic, river, spatial_n_group = 10)
+R> res_sp <- match_spatial(df1 = climate_vic, df2 = river, 
++                         spatial_n_group = 10)
 R> print(res_sp, n = 20)
@@
 R> res_sp <- match_spatial(
-+   climate_vic, river, 
++   df1 = climate_vic, df2 = river, 
 +   spatial_n_group = 10, return_cubble = TRUE)
@@
 \begin{CodeInput}
-R> (res_tm <- res_sp |> 
-+   match_temporal(
+R> (res_tm <- match_temporal(data = res_sp,
 +                           data_id = type, match_id = group,
@@
 \begin{CodeInput}
-R> res_tm <- res_sp |> 
-+   match_temporal(
+R> res_tm <- match_temporal(data =  res_sp,
 +                          data_id = type, match_id = group,
@@
 The ERA5 reanalysis \citep{hersbach2020era5} provides hourly estimates of atmospheric, land and oceanic climate variables on a global scale and is available in the NetCDF format from Copernicus Climate Data Store (CDS).
-This example reproduces Figure 19 in the \citet{hersbach2020era5} paper with cubble, in Figure \ref{fig:netcdf}.
+This example demonstrates a case of analysing raster spatio-temporal data using cubble, replicating Figure 19 from the \citet{hersbach2020era5} paper.
 The plot shows the southern polar vortex splitting into two on 2002-09-26, and further splitting into four on 2002-10-04.

### 2023-07-02 cubble.Rmd
check on code markup with \code{}g

@@
 Other \proglang{R} spatio-temporal objects can be coerced into a \code{cubble} object with the function \code{as_cubble()}.
-This includes a joined \code{tibble} or \code{data.frame}, a NetCDF object, a \code{stars} object [@stars], and a \code{sftime} object [@sftime].
+This includes a joined \code{tibble} or \code{data.frame} object, a NetCDF object, a \code{stars} object [@stars], and a \code{sftime} object [@sftime].
 In the example below, the spatial cubble object is created from \code{climate_flat}, which combines the previous \code{stations} and \code{meteo} into a single tibble object:
@@
 The pair of cubble verbs, \code{face_temporal()} and \code{face_spatial()}, pivots the cubble object between its two forms or faces, as illustrated in Figure \ref{fig:face}.
-The code applies \code{face_temporal()} on the spatial cubble (\code{cb_spatial}) introduced in Section \@ref(object) to get a temporal cubble:
+The code applies \code{face_temporal()} on the spatial cubble, \code{cb_spatial}, introduced in Section \@ref(object) to get a temporal cubble:
 
@@
 
-The \code{key} and \code{index} arguments in a \code{cubble} object corresponds to the \code{tsibble} counterparts and they can be safely omitted, if the temporal component is a \code{tsibble} object, i.e.
-\code{meteo_ts} in the example below.
+The \code{key} and \code{index} arguments in a \code{cubble} object corresponds to the tsibble counterparts and they can be safely omitted, if the temporal component is a tsibble object (\code{tbl_ts}).
 The \code{tsibble} class (\code{tbl_ts}) from the input will be carried over to the temporal cubble, indicated by the \code{[tsibble]} in the header and in the object class:
@@
 ```{r echo = TRUE}
+class(meteo_ts)
 ts_spatial <- make_cubble(
@@
 
-Methods applied to the \code{tbl_ts} class can also be applied to the temporal cubble objects, for example, checking whether the data contain temporal gaps:
+Methods applied to tsibble objects (\code{tbl_ts}) can also be applied to the temporal cubble objects, for example, checking whether the data contain temporal gaps:
 
@@
 
-The temporal component of a created temporal cubble can be turned to a \code{tsibble} object using \code{make_temporal_tsibble()}.
+The temporal component of a created temporal cubble can include class \code{tbl_ts} to also be a tsibble object \code{tsibble} object using \code{make_temporal_tsibble()}.
 See the code example below using the \code{cb_temporal} object, created in Section \@ref(create):
@@
 
-Similarly, the spatial component  of a cubble object can be an \code{sf} object and if the \code{coords} argument is omitted, it will be calculated from the sf geometry.
+Similarly, the spatial component  of a cubble object can be an \code{sf} object and if the \code{coords} argument is omitted, it will be calculated from the \code{sf} geometry.
 The sf status is signalled by the \code{[sf]} label in the cubble header:
@@
 
-This allows applying functions from the \code{sf} package to a cubble object, for example, to handle coordinate transformation with \code{st_transform}:
+This allows applying functions from the \code{sf} package to a cubble object, for example, to handle coordinate transformation with \code{st_transform()}:
 
@@
 
-The spatial component of a created \code{cubble} can also be promoted into an \code{sf} object with \code{make_spatial_sf()}:
+The spatial component of a created cubble can also be an \code{sf} object using \code{make_spatial_sf()}:
 
@@
 
-In \proglang{R}, there are other existing spatio-temporal data structure and this section compares and contrasts \pkg{cubble} with other existing alternatives, specifically \pkg{stars} and \pkg{sftime}.
+In \proglang{R}, there are other existing spatio-temporal data structures and this section compares and contrasts \pkg{cubble} with other existing alternatives, specifically \pkg{stars} and \pkg{sftime}.
 The \pkg{stars} package [@stars] uses an array structure, as oppose to tibble, to represent multivariate spatio-temporal data.
@@
 The difference in LGA naming between both data sets triggers cubble to issue a warning, alerting the user to this discrepancy.
-The warning message suggests there are some differences between the LGA encoding used by Victoria government and ABS and prompts analysts to check the mismatch using \code{check_key()}, which takes the same inputs as \code{make_cubble}, but returns a summary of key matches between the spatial and temporal input data.
+The warning message suggests there are some differences between the LGA encoding used by Victoria government and ABS and prompts analysts to check the mismatch using \code{check_key()}, which takes the same inputs as \code{make_cubble()}, but returns a summary of key matches between the spatial and temporal input data.
 
@@
 A quick check on the number of observations for each location is made, revealing that there are several with less than 24 observations -- these stations lack temperature values for some months.
-In this example, those stations are removed by switching to the spatial \code{cubble} to operate on the spatial component over time, and then, move back into the temporal \code{cubble} (to make the glyph map):
+In this example, those stations are removed by switching to the spatial cubble to operate on the spatial component over time, and then, move back into the temporal cubble (to make the glyph map):
 

### 2023-07-02 cubble.tex
check on code markup with \code{}g

@@
 Other \proglang{R} spatio-temporal objects can be coerced into a \code{cubble} object with the function \code{as_cubble()}.
-This includes a joined \code{tibble} or \code{data.frame}, a NetCDF object, a \code{stars} object \citep{stars}, and a \code{sftime} object \citep{sftime}.
+This includes a joined \code{tibble} or \code{data.frame} object, a NetCDF object, a \code{stars} object \citep{stars}, and a \code{sftime} object \citep{sftime}.
 In the example below, the spatial cubble object is created from \code{climate_flat}, which combines the previous \code{stations} and \code{meteo} into a single tibble object:
@@
 The pair of cubble verbs, \code{face_temporal()} and \code{face_spatial()}, pivots the cubble object between its two forms or faces, as illustrated in Figure \ref{fig:face}.
-The code applies \code{face_temporal()} on the spatial cubble (\code{cb_spatial}) introduced in Section \ref{object} to get a temporal cubble:
+The code applies \code{face_temporal()} on the spatial cubble, \code{cb_spatial}, introduced in Section \ref{object} to get a temporal cubble:
 
@@
 
-The \code{key} and \code{index} arguments in a \code{cubble} object corresponds to the \code{tsibble} counterparts and they can be safely omitted, if the temporal component is a \code{tsibble} object, i.e.~\code{meteo_ts} in the example below.
+The \code{key} and \code{index} arguments in a \code{cubble} object corresponds to the tsibble counterparts and they can be safely omitted, if the temporal component is a tsibble object (\code{tbl_ts}).
 The \code{tsibble} class (\code{tbl_ts}) from the input will be carried over to the temporal cubble, indicated by the \code{[tsibble]} in the header and in the object class:
@@
 \begin{CodeChunk}
+\begin{CodeInput}
+R> class(meteo_ts)
+\end{CodeInput}
+\begin{CodeOutput}
+[1] "tbl_ts"     "tbl_df"     "tbl"        "data.frame"
+\end{CodeOutput}
 \begin{CodeInput}
@@
 
-The temporal component of a created temporal cubble can be turned to a \code{tsibble} object using \code{make_temporal_tsibble()}.
+The temporal component of a created temporal cubble can include class \code{tbl_ts} to also be a tsibble object \code{tsibble} object using \code{make_temporal_tsibble()}.
 See the code example below using the \code{cb_temporal} object, created in Section \ref{create}:
@@
 
-Similarly, the spatial component of a cubble object can be an \code{sf} object and if the \code{coords} argument is omitted, it will be calculated from the sf geometry.
+Similarly, the spatial component of a cubble object can be an \code{sf} object and if the \code{coords} argument is omitted, it will be calculated from the \code{sf} geometry.
 The sf status is signalled by the \code{[sf]} label in the cubble header:
@@
 
-In \proglang{R}, there are other existing spatio-temporal data structure and this section compares and contrasts \pkg{cubble} with other existing alternatives, specifically \pkg{stars} and \pkg{sftime}.
+In \proglang{R}, there are other existing spatio-temporal data structures and this section compares and contrasts \pkg{cubble} with other existing alternatives, specifically \pkg{stars} and \pkg{sftime}.
 The \pkg{stars} package \citep{stars} uses an array structure, as oppose to tibble, to represent multivariate spatio-temporal data.

### 2023-07-03 cubble.Rmd
captions revised

@@
 : (#tab:methods) An overview of the methods implemented in the three \code{cubble} classes.
-Methods are implemented in the \code{cubble\_df} class when they behave consistent across the spatial and temporal cubble; otherwise, they are implemented separately.
+Methods are implemented in the \code{cubble\_df} class when they behave consistently across the spatial and temporal cubble; otherwise, they are implemented separately.
 
@@
 
-```{r face, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "An illustration of the functions \\code{face\\_temporal()} and \\code{face\\_spatial()}: \\code{face\\_temporal()} converts a spatial cubble into a temporal cubble to focus on the temporal variables.
-Conversely, \\code{face\\_spatial()} transforms a temporal cubble into a spatial one to focus on the spatial variables."}
+```{r face, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "Illustration of main functions.
+To focus on he temporal variables \\code{face\\_temporal()} converts a spatial cubble into a temporal cubble.
+To focus on the spatial variables \\code{face\\_spatial()} transforms a temporal cubble into a spatial cubble.
+This pivoting makes it easy to separately do spatial or temporal analysis."}
 knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.001.png"))
@@
 
-```{r illu-interactive, echo = FALSE, fig.align="center", out.height = "35%", out.width = "100%", fig.cap = "Linking between multiple plots.
-The line plots and the map are constructed from shared \\code{crosstalk} objects.
+```{r illu-interactive, echo = FALSE, fig.align="center", out.height = "35%", out.width = "100%", fig.cap = "Linking between multiple plots is made possible by shared \\code{crosstalk} objects.
 When a station is selected on the map (a), the corresponding row in the spatial \\code{cubble} will be activated.
-This will link to all the rows with the same id in the temporal \\code{cubble} (b) and update the line plot (c)."}
+This will activate the row with the same id in the temporal \\code{cubble} (b) to trigger an update of the line plot (c).
+The cubble package makes linking between spatial and temporal plots easy."}
 knitr::include_graphics(here::here("figures/diagram-keynotes/‎diagram-keynotes.‎002.png")) 
@@
 
-```{r glyphmap, out.width="100%", fig.width=15, fig.height=10, fig.cap="Comparison of average maximum temperature between 1971-1975 and 2016-2020 for 54 stations in Victoria and New South Wales, Australia.
-(a) and (b): the monthly temperature series for the two periods in a glyph map and for a single station Cobar, highlighted in orange in the glyph map.
-(c) and (d): the difference series between the two periods (2016s minus 1971s) in a glyph map and for station Cobar.
-The grey horizontal line marks zero difference.
-The glyph map displaying the difference series (c) reveals more pronounced changes between the two periods, with many inland locations in New South Wales show an increased temperature in late summer (Jan-Feb) in recent years."}
+```{r glyphmap, out.width="100%", fig.width=15, fig.height=10, fig.cap="Glyph maps comparing temperature change between 1971-1975 and 2016-2020 for 54 stations in Victoria and New South Wales, Australia.
+Overlaid line plots show monthly temperature (a) where a hint of late summer warming can be seen.
+Transforming to temperature differences (c) shows pronounced changes between the two periods.
+The horizontal guideline marks zero difference.
+One station, Cobar, is highlighted in the glyph maps and shown separately (b, d).
+Here the late summer (Jan-Feb) warming pattern, which is more prevalent at inland locations, is clear."}
 tmax2 <- tmax |> 
@@
 p1 <- tmax |> 
+  mutate(group = factor(group, levels = c("1971", "2016"), 
+     labels = c("1971-1975", "2016-2020"))) |>
   ggplot(aes(x_major = long, x_minor = month, 
@@
 
-```{r matching, out.width="100%", fig.height = 5, fig.width = 10, fig.cap="Matched weather stations and river gauges on the map (a) and across time (b).
+```{r matching, out.width="100%", fig.height = 5, fig.width = 10, fig.cap="Example of matching weather stations and river gauges.
+These four stations show on the map (a) and time (b) would be considered to be matching.
 Precipitation and water level have been standardised between 0 and 1 to be displayed in the same scale in (b).
-The water level reflects the increase in precipitation."}
+The peaks in the time series roughly match, and would reflect percipitation increasing water levels."}
 res_tm <- res_tm |>  filter(group %in% c(1, 7, 6, 9))
@@
 Once the NetCDF data is coerced into a cubble object, subsequent analysis can be conducted to filter on the date of interest, scale the variable specific humidity and create visualisation in ggplot to reproduce the ERA5 plot.
-A snippet of code to create Figure \ref{fig:netcdf} is provided below with additional codes needed to style the plot
+A snippet of code to create Figure \ref{fig:netcdf} is provided below with additional codes needed to style the plot.
 
@@
 
-```{r netcdf, out.width="100%", fig.height = 4, fig.width = 10, fig.cap = "A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020) to illustrate the break-up of sourthern polar vortex in late September and early October 2002.
-The polar vortex, signalled by the high specific humidity, splits into two on 2002-09-26 and further splits into four on 2002-10-04.", dev = "png"}
+```{r netcdf, out.width="100%", fig.height = 4, fig.width = 10, fig.cap = "An example illustrating that cubble can be used to readily reproduce common spatiotemporal analyses.
+This plot of ERA5 reanalysis (Fig.
+19, Hersbach et al, 2020) shows the break-up of the southern polar vortex in late September and early October 2002.
+The polar vortex, signalled by the high specific humidity, splits into two on 2002-09-26 and further splits into four on 2002-10-04.", dev = "png", fig.retina=4}
 res <- dt |> 
@@
 
-```{r interactive-linking, echo = FALSE, out.width="100%", out.height="23%", fig.retina = 2, dpi = 300, fig.cap = "Exploring temperature variation using linking of a map and seasonal display.
+```{r interactive-linking, echo = FALSE, out.width="100%", out.height="23%", fig.retina = 2, dpi = 300, fig.cap = "Illustration of using cubble for interactive graphics.
+Here we explore temperature variation by linking a map and a seasonal display.
 Each row is a screen dump of the process.

### 2023-07-03 cubble.tex
captions revised

@@
 \caption{\label{tab:methods} An overview of the methods implemented in the three \code{cubble} classes.
-Methods are implemented in the \code{cubble\_df} class when they behave consistent across the spatial and temporal cubble; otherwise, they are implemented separately.}\tabularnewline
-\toprule()
+Methods are implemented in the \code{cubble\_df} class when they behave consistently across the spatial and temporal cubble; otherwise, they are implemented separately.}\tabularnewline
+\toprule\noalign{}
 \begin{minipage}[b]{\linewidth}\raggedright
@@
 
-\caption{An illustration of the functions \code{face\_temporal()} and \code{face\_spatial()}: \code{face\_temporal()} converts a spatial cubble into a temporal cubble to focus on the temporal variables.
-Conversely, \code{face\_spatial()} transforms a temporal cubble into a spatial one to focus on the spatial variables.}\label{fig:face}
+\caption{Illustration of main functions.
+To focus on he temporal variables \code{face\_temporal()} converts a spatial cubble into a temporal cubble.
+To focus on the spatial variables \code{face\_spatial()} transforms a temporal cubble into a spatial cubble.
+This pivoting makes it easy to separately do spatial or temporal analysis.}\label{fig:face}
 \end{figure}
@@
 
-\caption{Linking between multiple plots.
-The line plots and the map are constructed from shared \code{crosstalk} objects.
+\caption{Linking between multiple plots is made possible by shared \code{crosstalk} objects.
 When a station is selected on the map (a), the corresponding row in the spatial \code{cubble} will be activated.
-This will link to all the rows with the same id in the temporal \code{cubble} (b) and update the line plot (c).}\label{fig:illu-interactive}
+This will activate the row with the same id in the temporal \code{cubble} (b) to trigger an update of the line plot (c).
+The cubble package makes linking between spatial and temporal plots easy.}\label{fig:illu-interactive}
 \end{figure}
@@
 
-\caption[Comparison of average maximum temperature between 1971-1975 and 2016-2020 for 54 stations in Victoria and New South Wales, Australia]{Comparison of average maximum temperature between 1971-1975 and 2016-2020 for 54 stations in Victoria and New South Wales, Australia.
-(a) and (b): the monthly temperature series for the two periods in a glyph map and for a single station Cobar, highlighted in orange in the glyph map.
-(c) and (d): the difference series between the two periods (2016s minus 1971s) in a glyph map and for station Cobar.
-The grey horizontal line marks zero difference.
-The glyph map displaying the difference series (c) reveals more pronounced changes between the two periods, with many inland locations in New South Wales show an increased temperature in late summer (Jan-Feb) in recent years.}\label{fig:glyphmap}
+\caption[Glyph maps comparing temperature change between 1971-1975 and 2016-2020 for 54 stations in Victoria and New South Wales, Australia]{Glyph maps comparing temperature change between 1971-1975 and 2016-2020 for 54 stations in Victoria and New South Wales, Australia.
+Overlaid line plots show monthly temperature (a) where a hint of late summer warming can be seen.
+Transforming to temperature differences (c) shows pronounced changes between the two periods.
+The horizontal guideline marks zero difference.
+One station, Cobar, is highlighted in the glyph maps and shown separately (b, d).
+Here the late summer (Jan-Feb) warming pattern, which is more prevalent at inland locations, is clear.}\label{fig:glyphmap}
 \end{figure}
@@
 
-\caption[Matched weather stations and river gauges on the map (a) and across time (b)]{Matched weather stations and river gauges on the map (a) and across time (b).
+\caption[Example of matching weather stations and river gauges]{Example of matching weather stations and river gauges.
+These four stations show on the map (a) and time (b) would be considered to be matching.
 Precipitation and water level have been standardised between 0 and 1 to be displayed in the same scale in (b).
-The water level reflects the increase in precipitation.}\label{fig:matching}
+The peaks in the time series roughly match, and would reflect percipitation increasing water levels.}\label{fig:matching}
 \end{figure}
@@
 
-\caption[A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020) to illustrate the break-up of sourthern polar vortex in late September and early October 2002]{A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020) to illustrate the break-up of sourthern polar vortex in late September and early October 2002.
+\caption[An example illustrating that cubble can be used to readily reproduce common spatiotemporal analyses]{An example illustrating that cubble can be used to readily reproduce common spatiotemporal analyses.
+This plot of ERA5 reanalysis (Fig.
+19, Hersbach et al, 2020) shows the break-up of the southern polar vortex in late September and early October 2002.
 The polar vortex, signalled by the high specific humidity, splits into two on 2002-09-26 and further splits into four on 2002-10-04.}\label{fig:netcdf}
@@
 
-\caption[Exploring temperature variation using linking of a map and seasonal display]{Exploring temperature variation using linking of a map and seasonal display.
+\caption[Illustration of using cubble for interactive graphics]{Illustration of using cubble for interactive graphics.
+Here we explore temperature variation by linking a map and a seasonal display.
 Each row is a screen dump of the process.

### 2023-07-03 cubble.Rmd
Changes section 3.1

@@
 
-## Data fusion and matching {#matching}
+## Spatial and temporal matching {#matching}
 
-Matching time series from two lists is a common task in spatio-temporal data analysis.
-In cubble, matching based on distance and time series features can be performed using the functions `match_spatial()` and `match_temporal()`.
-The `match_spatial()` function finds the matched pairs in two cubble objects based on distance between sites:
+Finding nearby locations across two spatio-temporal data sets measuring different processes is a common task in spatio-temporal data analysis.
+For example, we might want to related measurements obtained from weather stations and river gauge levels recorded at different locations and the goal is to find nearby river sensors that reflect the change in precipitation recorded by the weather station.
+
+This type of problem can be generalize to matched pairs of locations accross different data sets or more generality groups of locations
+
+or groups with a number of locations selected depending on the problem at hand. For example, we might want to find for a given weather station the closest river gauges to compare levels
+
+For example, we might want to find a group of close stations and then use score functions to describe time series features between the series.
+In cubble, matching locations based on distance can be performed using the function `match_spatial()` which finds  matched groups in two cubble objects by first fiding the closests pairs and.
+
+Two arguments are available to control the matching: the argument `spatial_n_group` specifies the number of paired groups to return and the argument `spatial_n_each` specifies the number of matches for each item in the first cubble object (default to 1 for one-to-one matching).
+
+and matching time series features for selected locations`match_temporal()`. The `match_spatial()` function :
 
@@
 ````
-
-Two arguments are available to control the matching: the argument `spatial_n_group` specifies the number of paired groups to return and the argument `spatial_n_each` specifies the number of matches for each item in the first cubble object (default to 1 for one-to-one matching).
 

### 2023-07-03 cubble.Rmd
proof-read abstarct and intro, small changes

@@
 For many analyses, spatial and time components can be separately studied: for example, to explore the temporal trend of one variable for a single spatial location, or to model the spatial distribution of one variable at a given time.
-However for some studies, it is important to analyze different aspects of the spatio-temporal data simultaneously, like for instance, temporal trends of multiple variables across locations.
+However for some studies, it is important to analyze different aspects of the spatio-temporal data simultaneously, for instance, temporal trends of multiple variables across locations.
 In order to facilitate the study of different portions or combinations of spatio-temporal data, we introduce a new class, \code{cubble}, with a suite of functions enabling easy slicing and dicing on different spatio-temporal components.
@@
 In addition, the \pkg{cubble} package facilitates visual and numerical explorations of the data while easing data wrangling and modelling.
-The \code{cubble} class and the tools implemented in the package are illustrated with different examples of Australian climate data.
+The \code{cubble} class and the tools implemented in the package are illustrated with examples from climate data analysis.
 keywords:
@@
 With spatio-temporal data, one can fix the time to explore the spatial features of the data, fix the spatial location/s to explore temporal aspects, or dynamically explore the space and time simultaneously.
-
-In order to computationally explore the spatial, temporal and spatio-temporal faces of such data, the data needs to be stored and represented under a specific data object that allows the user to query, group and dissect all the data faces.
+In order to computationally explore the spatial, temporal and spatio-temporal and multiple variable aspects of such data, it needs to be stored in a data object that allows the user to query, group and dissect all the different data faces.
 
@@
 
-This paper presents a new \proglang{R} package, \pkg{cubble}, which implements a new cubble class to organize spatial and temporal variables as two forms of a single data object so that they can be wrangled separately or combined, while being kept synchronized.
+This paper presents the \proglang{R} package, \pkg{cubble}, which implements a new cubble class to organize spatial and temporal variables as two forms of a single data object so that they can be wrangled separately or combined, while being kept synchronized.
 Among the four spacetime layouts in @spacetime, the \code{cubble} class can handle the full grid layout and the sparse grid layout.

### 2023-07-03 cubble.tex
proof-read abstarct and intro, small changes

@@
 In addition, the \pkg{cubble} package facilitates visual and numerical explorations of the data while easing data wrangling and modelling.
-The \code{cubble} class and the tools implemented in the package are illustrated with different examples of Australian climate data.
+The \code{cubble} class and the tools implemented in the package are illustrated with examples from climate data analysis.
 }
@@
 
-This paper presents a new \proglang{R} package, \pkg{cubble}, which implements a new cubble class to organize spatial and temporal variables as two forms of a single data object so that they can be wrangled separately or combined, while being kept synchronized.
+This paper presents the \proglang{R} package, \pkg{cubble}, which implements a new cubble class to organize spatial and temporal variables as two forms of a single data object so that they can be wrangled separately or combined, while being kept synchronized.
 Among the four spacetime layouts in \citet{spacetime}, the \code{cubble} class can handle the full grid layout and the sparse grid layout.
@@
 \hypertarget{matching}{%
-\subsection{Data fusion and matching}\label{matching}}
-
-Matching time series from two lists is a common task in spatio-temporal data analysis.
-In cubble, matching based on distance and time series features can be performed using the functions \texttt{match\_spatial()} and \texttt{match\_temporal()}.
-The \texttt{match\_spatial()} function finds the matched pairs in two cubble objects based on distance between sites:
+\subsection{Spatial and temporal matching}\label{matching}}
+
+Finding nearby locations across two spatio-temporal data sets measuring different processes is a common task in spatio-temporal data analysis.
+For example, we might want to related measurements obtained from weather stations and river gauge levels recorded at different locations and the goal is to find nearby river sensors that reflect the change in precipitation recorded by the weather station.
+
+This type of problem can be generalize to matched pairs of locations accross different data sets or more generality groups of locations
+
+or groups with a number of locations selected depending on the problem at hand. For example, we might want to find for a given weather station the closest river gauges to compare levels
+
+For example, we might want to find a group of close stations and then use score functions to describe time series features between the series.
+In cubble, matching locations based on distance can be performed using the function \texttt{match\_spatial()} which finds matched groups in two cubble objects by first fiding the closests pairs and.
+
+Two arguments are available to control the matching: the argument \texttt{spatial\_n\_group} specifies the number of paired groups to return and the argument \texttt{spatial\_n\_each} specifies the number of matches for each item in the first cubble object (default to 1 for one-to-one matching).
+
+and matching time series features for selected locations\texttt{match\_temporal()}. The \texttt{match\_spatial()} function :
 
@@
 \end{verbatim}
-
-Two arguments are available to control the matching: the argument \texttt{spatial\_n\_group} specifies the number of paired groups to return and the argument \texttt{spatial\_n\_each} specifies the number of matches for each item in the first cubble object (default to 1 for one-to-one matching).
 

### 2023-07-03 cubble.Rmd
section 2 proofed, small changes

@@
 Both cubble objects inherit tibble's attributes (which originates from data frames): \code{class}, \code{row.names}, and \code{names}.
-Additionally, both have three specialised attributes: \code{key}, \code{index}, and \code{coords}.
-\code{key} and \code{index} are used as introduced in the \pkg{tsibble} package [@tsibble].
+Additionally, both have three specialised attributes: \code{key}, \code{index}, and \code{coords}, where \code{key} and \code{index} are used as introduced in the \pkg{tsibble} package [@tsibble].
 In cubble, the \code{key} attribute identifies the row in the spatial cubble (given the internal use of \code{tidyr::nest()} for nesting), and when combined with the \code{index} argument, it identifies the row in the temporal cubble.
-Currently, cubble only supports one variable as the key, and the accepted temporal classes for \code{index} includes the base R classes \code{Date}, \code{POSIXlt}, \code{POSIXct}, as well as tsibble's \code{yearmonth}, \code{yearweek}, and \code{yearquarter} classes.
+Currently, cubble only supports one variable as the key.
+The accepted temporal classes for \code{index} includes the base R classes \code{Date}, \code{POSIXlt}, \code{POSIXct}, as well as tsibble's \code{yearmonth}, \code{yearweek}, and \code{yearquarter} classes.
 The \code{coords} attribute represents an ordered pair of coordinates that can be either an unprojected pair of longitude and latitude, or a projected easting and northing value.
@@
 A (spatial) cubble object can be constructed from separate spatial and temporal tables using the function \code{make_cubble()}.
-The three attributes \code{key}, \code{index}, and \code{coords} need to be specified in creation.
+The three attributes \code{key}, \code{index}, and \code{coords} need to be specified.
 The following code creates a spatial cubble from its spatial component, \code{stations} and temporal component \code{meteo}:
@@
 In \proglang{R}, there are other existing spatio-temporal data structures and this section compares and contrasts \pkg{cubble} with other existing alternatives, specifically \pkg{stars} and \pkg{sftime}.
-The \pkg{stars} package [@stars] uses an array structure, as oppose to tibble, to represent multivariate spatio-temporal data.
+The \pkg{stars} package [@stars] uses an array structure, as opposed to a tibble, to represent multivariate spatio-temporal data.
 While both \pkg{stars} and \pkg{cubble} support vector and raster data, it is a matter of choice on which structure to use given the application.
@@
 However, \pkg{cubble} uses nesting to avoid storing spatial variables repetitively at each timestamp.
-This provides memory efficiency when data is observed frequent, i.e.
+This provides memory efficiency when data is observed frequently, i.e.
 daily or sub-daily, or the spatial geometry is computationally expensive to store repeatedly, i.e.

### 2023-07-03 cubble.tex
section 2 proofed, small changes

@@
 Both cubble objects inherit tibble's attributes (which originates from data frames): \code{class}, \code{row.names}, and \code{names}.
-Additionally, both have three specialised attributes: \code{key}, \code{index}, and \code{coords}.
-\code{key} and \code{index} are used as introduced in the \pkg{tsibble} package \citep{tsibble}.
+Additionally, both have three specialised attributes: \code{key}, \code{index}, and \code{coords}, where \code{key} and \code{index} are used as introduced in the \pkg{tsibble} package \citep{tsibble}.
 In cubble, the \code{key} attribute identifies the row in the spatial cubble (given the internal use of \code{tidyr::nest()} for nesting), and when combined with the \code{index} argument, it identifies the row in the temporal cubble.
-Currently, cubble only supports one variable as the key, and the accepted temporal classes for \code{index} includes the base R classes \code{Date}, \code{POSIXlt}, \code{POSIXct}, as well as tsibble's \code{yearmonth}, \code{yearweek}, and \code{yearquarter} classes.
+Currently, cubble only supports one variable as the key.
+The accepted temporal classes for \code{index} includes the base R classes \code{Date}, \code{POSIXlt}, \code{POSIXct}, as well as tsibble's \code{yearmonth}, \code{yearweek}, and \code{yearquarter} classes.
 The \code{coords} attribute represents an ordered pair of coordinates that can be either an unprojected pair of longitude and latitude, or a projected easting and northing value.
@@
 In \proglang{R}, there are other existing spatio-temporal data structures and this section compares and contrasts \pkg{cubble} with other existing alternatives, specifically \pkg{stars} and \pkg{sftime}.
-The \pkg{stars} package \citep{stars} uses an array structure, as oppose to tibble, to represent multivariate spatio-temporal data.
+The \pkg{stars} package \citep{stars} uses an array structure, as opposed to a tibble, to represent multivariate spatio-temporal data.
 While both \pkg{stars} and \pkg{cubble} support vector and raster data, it is a matter of choice on which structure to use given the application.
@@
 However, \pkg{cubble} uses nesting to avoid storing spatial variables repetitively at each timestamp.
-This provides memory efficiency when data is observed frequent, i.e.~daily or sub-daily, or the spatial geometry is computationally expensive to store repeatedly, i.e.~polygons or multipolygons.
+This provides memory efficiency when data is observed frequently, i.e.~daily or sub-daily, or the spatial geometry is computationally expensive to store repeatedly, i.e.~polygons or multipolygons.
 Consider the \code{climate_aus} data in the \pkg{cubble} package with 639 stations observed daily throughout the year 2020.

### 2023-07-03 cubble.Rmd
improve description to the matching motivation

@@
 
-Finding nearby locations across two spatio-temporal data sets measuring different processes is a common task in spatio-temporal data analysis.
-For example, we might want to related measurements obtained from weather stations and river gauge levels recorded at different locations and the goal is to find nearby river sensors that reflect the change in precipitation recorded by the weather station.
+A common task in spatio-temporal data analysis is to identify related temporal series from geographically proximate locations.
+For example, we may want to associate weather stations with nearby river sensors to find the water level series that reflect changes in precipitation.
+This problem can be generalised by first matching locations across two datasets and then calculating the temporal features of the series.
+In cubble, locations from two datasets can be matched using the function `match_spatial()`.
+The function calculates the distance matrix of the two datasets and returns `spatial_n_group` groups with the smallest distances.
+For each group, analysts can include more location with the argument `spatial_n_each` (default to 1 for one-on-one matching).
 
-This type of problem can be generalize to matched pairs of locations accross different data sets or more generality groups of locations
-
-or groups with a number of locations selected depending on the problem at hand. For example, we might want to find for a given weather station the closest river gauges to compare levels
-
-For example, we might want to find a group of close stations and then use score functions to describe time series features between the series.
-In cubble, matching locations based on distance can be performed using the function `match_spatial()` which finds  matched groups in two cubble objects by first fiding the closests pairs and.
-
-Two arguments are available to control the matching: the argument `spatial_n_group` specifies the number of paired groups to return and the argument `spatial_n_each` specifies the number of matches for each item in the first cubble object (default to 1 for one-to-one matching).
-
-and matching time series features for selected locations`match_temporal()`. The `match_spatial()` function :
-
-````
-match_spatial(<cubble_obj1>, <cubble_obj2>, ...)
-````
-
-The function `match_temporal()` takes the outputs from spatial matching and calculates a similarity score of the time series between spatially matched pairs.
+The temporal matching then calculates a similarity score between the time series of spatially matched pairs using the function `match_temporal()`.
 The temporal matching requires two identifiers: one for separating each spatially matched group: `match_id` and one for separating the two data sources: `data_id`.

### 2023-07-03 cubble.tex
improve description to the matching motivation

@@
 \caption{\label{tab:funs} An overview of functions implemented in the \pkg{cubble} package, categorised into base R, tidyverse, and cubble functions.}\tabularnewline
-\toprule\noalign{}
+\toprule()
 \begin{minipage}[b]{\linewidth}\raggedright
@@
 \end{minipage} \\
-\midrule\noalign{}
+\midrule()
 \endfirsthead
-\toprule\noalign{}
+\toprule()
 \begin{minipage}[b]{\linewidth}\raggedright
@@
 \end{minipage} \\
-\midrule\noalign{}
+\midrule()
 \endhead
-\bottomrule\noalign{}
-\endlastfoot
 base R & \texttt{{[}}, \texttt{{[}{[}\textless{}-}, \texttt{names\textless{}-} \\
@@
 cubble & \texttt{as\_cubble}, \texttt{cubble}, \texttt{make\_cubble}, \texttt{check\_key}, \texttt{face\_temporal}, \texttt{face\_spatial}, \texttt{unfold}, \texttt{key}, \texttt{key\_vars}, \texttt{key\_data}, \texttt{index}, \texttt{index\_var}, \texttt{coords}, \texttt{spatial}, \texttt{match\_sites}, \texttt{match\_spatial}, \texttt{match\_temporal}, \texttt{geom\_glyph}, \texttt{geom\_glyph\_box}, \texttt{geom\_glyph\_line}, \texttt{make\_spatial\_sf}, \texttt{make\_temporal\_tsibble}, \texttt{fill\_gaps}, and \texttt{scan\_gaps} \\
+\bottomrule()
 \end{longtable}
@@
 Methods are implemented in the \code{cubble\_df} class when they behave consistently across the spatial and temporal cubble; otherwise, they are implemented separately.}\tabularnewline
-\toprule\noalign{}
+\toprule()
 \begin{minipage}[b]{\linewidth}\raggedright
@@
 \end{minipage} \\
-\midrule\noalign{}
+\midrule()
 \endhead
-\bottomrule\noalign{}
-\endlastfoot
 \texttt{cubble\_df} & \texttt{{[}{[}\textless{}-,\ dplyr\_col\_modify,\ key\_data,\ key\_vars,\ key,\ print} \\
@@
 \texttt{temporal\_cubble\_df} & \texttt{{[},\ names\textless{}-,\ tbl\_sum,\ arrange,\ dplyr\_reconstruct,\ dplyr\_row\_slice,\ face\_spatial,\ face\_temporal,\ unfold,\ fill\_gaps,\ group\_by,\ ungroup,\ \ rename,\ rowwise,\ scan\_gaps,\ select,\ spatial,\ summarise,\ tbl\_sum,\ bind\_rows,\ bind\_cols,\ update\_cubble} \\
+\bottomrule()
 \end{longtable}
@@
 
-Finding nearby locations across two spatio-temporal data sets measuring different processes is a common task in spatio-temporal data analysis.
-For example, we might want to related measurements obtained from weather stations and river gauge levels recorded at different locations and the goal is to find nearby river sensors that reflect the change in precipitation recorded by the weather station.
-
-This type of problem can be generalize to matched pairs of locations accross different data sets or more generality groups of locations
-
-or groups with a number of locations selected depending on the problem at hand. For example, we might want to find for a given weather station the closest river gauges to compare levels
-
-For example, we might want to find a group of close stations and then use score functions to describe time series features between the series.
-In cubble, matching locations based on distance can be performed using the function \texttt{match\_spatial()} which finds matched groups in two cubble objects by first fiding the closests pairs and.
-
-Two arguments are available to control the matching: the argument \texttt{spatial\_n\_group} specifies the number of paired groups to return and the argument \texttt{spatial\_n\_each} specifies the number of matches for each item in the first cubble object (default to 1 for one-to-one matching).
-
-and matching time series features for selected locations\texttt{match\_temporal()}. The \texttt{match\_spatial()} function :
-
-\begin{verbatim}
-match_spatial(<cubble_obj1>, <cubble_obj2>, ...)
-\end{verbatim}
-
-The function \texttt{match\_temporal()} takes the outputs from spatial matching and calculates a similarity score of the time series between spatially matched pairs.
+A common task in spatio-temporal data analysis is to identify related temporal series from geographically proximate locations.
+For example, we may want to associate weather stations with nearby river sensors to find the water level series that reflect changes in precipitation.
+This problem can be generalised by first matching locations across two datasets and then calculating the temporal features of the series.
+In cubble, locations from two datasets can be matched using the function \texttt{match\_spatial()}.
+The function calculates the distance matrix of the two datasets and returns \texttt{spatial\_n\_group} groups with the smallest distances.
+For each group, analysts can include more location with the argument \texttt{spatial\_n\_each} (default to 1 for one-on-one matching).
+
+The temporal matching then calculates a similarity score between the time series of spatially matched pairs using the function \texttt{match\_temporal()}.
 The temporal matching requires two identifiers: one for separating each spatially matched group: \texttt{match\_id} and one for separating the two data sources: \texttt{data\_id}.
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/cookd/students_PhD/Sherry/paper-cubble/figures/glyphmap-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/research/paper-cubble/figures/glyphmap-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/cookd/students_PhD/Sherry/paper-cubble/figures/matching-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/research/paper-cubble/figures/matching-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/cookd/students_PhD/Sherry/paper-cubble/figures/netcdf-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/research/paper-cubble/figures/netcdf-1}
 

### 2023-07-03 cubble.Rmd
highligh what cubble does in interactive graphics in Section 3.2

@@
 The cubble workflow can be easily incorporated into an interactive graphics pipeline (e.g., @buja1988elements; @buja1996interactive;  @sutherland2000orca; @xie2014reactive; @cheng2016enabling).
-This section describes the linking between a map and multiple time series in a \code{cubble} object using the package \pkg{crosstalk} [@crosstalk], as illustrated in Figure \ref{fig:illu-interactive}.
-The spatial and temporal cubble can be constructed as a shared crosstalk object to create a link between a map and a time series plot.
-For instance, when a user selects a location on the map as shown on panel (a), the corresponding site is highlighted.
+The cubble data structure simplifies data pre-processing and prepares the ingredient for interactive graphics.
+Specifically, the spatial and temporal cubble correspond to the spatial and temporal visualisation, such as a map or a time series plot, that can be linked.
+Figure \ref{fig:illu-interactive} illustrates the linking mechanism between a map and multiple time series in the \pkg{crosstalk} [@crosstalk] package.
+When a user selects a location on the map as shown on panel (a), the corresponding site is highlighted.
 This selection activates a row in the spatial cubble, which is then connected to the temporal cubble, resulting in the selection of all observations with the same ID as depicted in panel (b).

### 2023-07-03 cubble.tex
highligh what cubble does in interactive graphics in Section 3.2

@@
 The cubble workflow can be easily incorporated into an interactive graphics pipeline (e.g., \citet{buja1988elements}; \citet{buja1996interactive}; \citet{sutherland2000orca}; \citet{xie2014reactive}; \citet{cheng2016enabling}).
-This section describes the linking between a map and multiple time series in a \code{cubble} object using the package \pkg{crosstalk} \citep{crosstalk}, as illustrated in Figure \ref{fig:illu-interactive}.
-The spatial and temporal cubble can be constructed as a shared crosstalk object to create a link between a map and a time series plot.
-For instance, when a user selects a location on the map as shown on panel (a), the corresponding site is highlighted.
+The cubble data structure simplifies data pre-processing and prepares the ingredient for interactive graphics.
+Specifically, the spatial and temporal cubble correspond to the spatial and temporal visualisation, such as a map or a time series plot, that can be linked.
+Figure \ref{fig:illu-interactive} illustrates the linking mechanism between a map and multiple time series in the \pkg{crosstalk} \citep{crosstalk} package.
+When a user selects a location on the map as shown on panel (a), the corresponding site is highlighted.
 This selection activates a row in the spatial cubble, which is then connected to the temporal cubble, resulting in the selection of all observations with the same ID as depicted in panel (b).

### 2023-07-03 cubble.Rmd
section 3.1

@@
 
-A common task in spatio-temporal data analysis is to identify related temporal series from geographically proximate locations.
-For example, we may want to associate weather stations with nearby river sensors to find the water level series that reflect changes in precipitation.
-This problem can be generalised by first matching locations across two datasets and then calculating the temporal features of the series.
+A useful task in spatio-temporal data analysis is to combine related temporal series within a close geographic neighborhood.
+For example, we may want to examine data from weather stations with water flow records in nearby river sensors to understand how precipitation relates to water supply.
+This might be useful for better predicting potential for drought and flood.
+
+This problem can be tackled by first matching spatial locations across two datasets and then calculating the temporal features of the multiple series.
 In cubble, locations from two datasets can be matched using the function `match_spatial()`.
 The function calculates the distance matrix of the two datasets and returns `spatial_n_group` groups with the smallest distances.
-For each group, analysts can include more location with the argument `spatial_n_each` (default to 1 for one-on-one matching).
+For each group, analysts can include more locations with the argument `spatial_n_each` (default to 1 for one-on-one matching).
 
@@
 
-The similarity score between two time series is calculated using a matching function, which can be customised by the analysts based on the time series feature relevant to match.
-The matching function takes two time series as a list and returns a single numerical score.
-By default, cubble uses a simple peak matching algorithm (`match_peak`) to count the number of peaks in two time series that fall within a specified temporal window.
+The similarity score is computed by a matching function which can be customized to any desired time series feature.
+The matching function takes two time series as a list and returns a single numerical value.
+By default, cubble uses a simple peak matching algorithm (`match_peak`) to count the number of peaks in two time series that fall within a specified time window.
 

### 2023-07-03 cubble.tex
section 3.1

--- 2023-07-03 cubble.tex
@@
 
-Finding nearby locations across two spatio-temporal data sets measuring different processes is a common task in spatio-temporal data analysis.
-For example, we might want to related measurements obtained from weather stations and river gauge levels recorded at different locations and the goal is to find nearby river sensors that reflect the change in precipitation recorded by the weather station.
-
-This type of problem can be generalize to matched pairs of locations accross different data sets or more generality groups of locations
-
-or groups with a number of locations selected depending on the problem at hand. For example, we might want to find for a given weather station the closest river gauges to compare levels
-
-For example, we might want to find a group of close stations and then use score functions to describe time series features between the series.
-In cubble, matching locations based on distance can be performed using the function \texttt{match\_spatial()} which finds matched groups in two cubble objects by first fiding the closests pairs and.
-
-Two arguments are available to control the matching: the argument \texttt{spatial\_n\_group} specifies the number of paired groups to return and the argument \texttt{spatial\_n\_each} specifies the number of matches for each item in the first cubble object (default to 1 for one-to-one matching).
-
-and matching time series features for selected locations\texttt{match\_temporal()}. The \texttt{match\_spatial()} function :
-
-\begin{verbatim}
-match_spatial(<cubble_obj1>, <cubble_obj2>, ...)
-\end{verbatim}
-
-The function \texttt{match\_temporal()} takes the outputs from spatial matching and calculates a similarity score of the time series between spatially matched pairs.
+A useful task in spatio-temporal data analysis is to combine related temporal series within a close geographic neighborhood.
+For example, we may want to examine data from weather stations with water flow records in nearby river sensors to understand how precipitation relates to water supply.
+This might be useful for better predicting potential for drought and flood.
+
+This problem can be tackled by first matching spatial locations across two datasets and then calculating the temporal features of the multiple series.
+In cubble, locations from two datasets can be matched using the function \texttt{match\_spatial()}.
+The function calculates the distance matrix of the two datasets and returns \texttt{spatial\_n\_group} groups with the smallest distances.
+For each group, analysts can include more locations with the argument \texttt{spatial\_n\_each} (default to 1 for one-on-one matching).
+
+The temporal matching then calculates a similarity score between the time series of spatially matched pairs using the function \texttt{match\_temporal()}.
 The temporal matching requires two identifiers: one for separating each spatially matched group: \texttt{match\_id} and one for separating the two data sources: \texttt{data\_id}.
@@
 
-The similarity score between two time series is calculated using a matching function, which can be customised by the analysts based on the time series feature relevant to match.
-The matching function takes two time series as a list and returns a single numerical score.
-By default, cubble uses a simple peak matching algorithm (\texttt{match\_peak}) to count the number of peaks in two time series that fall within a specified temporal window.
+The similarity score is computed by a matching function which can be customized to any desired time series feature.
+The matching function takes two time series as a list and returns a single numerical value.
+By default, cubble uses a simple peak matching algorithm (\texttt{match\_peak}) to count the number of peaks in two time series that fall within a specified time window.
 

### 2023-07-03 cubble.Rmd
section 3.2

@@
 
-The cubble workflow can be easily incorporated into an interactive graphics pipeline (e.g., @buja1988elements; @buja1996interactive;  @sutherland2000orca; @xie2014reactive; @cheng2016enabling).
-The cubble data structure simplifies data pre-processing and prepares the ingredient for interactive graphics.
-Specifically, the spatial and temporal cubble correspond to the spatial and temporal visualisation, such as a map or a time series plot, that can be linked.
-Figure \ref{fig:illu-interactive} illustrates the linking mechanism between a map and multiple time series in the \pkg{crosstalk} [@crosstalk] package.
+The cubble workflow neatly allows building an interactive graphics pipeline (e.g., @buja1988elements; @buja1996interactive;  @sutherland2000orca; @xie2014reactive; @cheng2016enabling), simplifying the data pre-processing and preparing the ingredients for linked plots.
+Specifically, the spatial and temporal cubble correspond to the spatial and temporal visualisation, such as a map or a time series plot, that can be linked using functionality in the \pkg{crosstalk} [@crosstalk] package.
+
+Figure \ref{fig:illu-interactive} illustrates the linking mechanism between a map and multiple time series.
 When a user selects a location on the map as shown on panel (a), the corresponding site is highlighted.

### 2023-07-03 cubble.tex
section 3.2

@@
 
-The cubble workflow can be easily incorporated into an interactive graphics pipeline (e.g., \citet{buja1988elements}; \citet{buja1996interactive}; \citet{sutherland2000orca}; \citet{xie2014reactive}; \citet{cheng2016enabling}).
-The cubble data structure simplifies data pre-processing and prepares the ingredient for interactive graphics.
-Specifically, the spatial and temporal cubble correspond to the spatial and temporal visualisation, such as a map or a time series plot, that can be linked.
-Figure \ref{fig:illu-interactive} illustrates the linking mechanism between a map and multiple time series in the \pkg{crosstalk} \citep{crosstalk} package.
+The cubble workflow neatly allows building an interactive graphics pipeline (e.g., \citet{buja1988elements}; \citet{buja1996interactive}; \citet{sutherland2000orca}; \citet{xie2014reactive}; \citet{cheng2016enabling}), simplifying the data pre-processing and preparing the ingredients for linked plots.
+Specifically, the spatial and temporal cubble correspond to the spatial and temporal visualisation, such as a map or a time series plot, that can be linked using functionality in the \pkg{crosstalk} \citep{crosstalk} package.
+
+Figure \ref{fig:illu-interactive} illustrates the linking mechanism between a map and multiple time series.
 When a user selects a location on the map as shown on panel (a), the corresponding site is highlighted.

### 2023-07-03 cubble.Rmd
section 3.3

@@
 
-Visualizing spatio-temporal data is important for exploring and understanding the data at hand, aiding in decision-making, and facilitating effective communication.
-To collectively visualize spatial and temporal information, several options are available such as:  faceted maps across time, map animations, or interactive graphics that link maps and time series plots among others.
-Faceted maps and spatio-temporal animations, while commonly used, may be difficult to compare across time and space since users need to inspect multiple facets or frames to find changes in time or space.
-The glyph map [@Wickham2012-yr] resolves this issue by imposing the time series onto the map as a glyph through a coordinate transformation.
-The transformation uses linear algebra to convert the temporal coordinates (minor coordinates) into the spatial coordinates (major coordinates) and is implemented in the package `GGally` [@ggally].
-The \pkg{cubble} package provides a ggproto implementation to create glyph maps, `geom_glyph()`  and it takes four required aestheticsf : `x_major`, `y_major`, `x_minor`, and `y_minor`:
+Visualizing both space and time is important for exploring and understanding the data more completely, aiding in decision-making, and facilitating effective communication.
+Several approaches are common: facet maps across time, map animations, or interactive graphics that link maps and time series plots among others.
+Faceted maps and spatio-temporal animations focus on the spatial pattern making it difficult to assess temporal trends.
+The glyph map [@Wickham2012-yr] addresses this issue by plotting time series onto the map as a glyph.
+
+It is achieved through a coordinate transformation.
+The transformation uses linear algebra to convert the temporal coordinates (minor coordinates) into the spatial coordinates (major coordinates) and is implemented in the package \pkg{GGally} [@ggally].
+The \pkg{cubble} package provides a new \code{ggproto} implementation to create glyph maps, \code{geom_glyph()}  requiring four aesthetics : \code{x_major}, \code{y_major}, \code{x_minor}, and \code{y_minor}:
 

### 2023-07-03 cubble.tex
section 3.3

@@
 
-Visualizing spatio-temporal data is important for exploring and understanding the data at hand, aiding in decision-making, and facilitating effective communication.
-To collectively visualize spatial and temporal information, several options are available such as: faceted maps across time, map animations, or interactive graphics that link maps and time series plots among others.
-Faceted maps and spatio-temporal animations, while commonly used, may be difficult to compare across time and space since users need to inspect multiple facets or frames to find changes in time or space.
-The glyph map \citep{Wickham2012-yr} resolves this issue by imposing the time series onto the map as a glyph through a coordinate transformation.
-The transformation uses linear algebra to convert the temporal coordinates (minor coordinates) into the spatial coordinates (major coordinates) and is implemented in the package \texttt{GGally} \citep{ggally}.
-The \pkg{cubble} package provides a ggproto implementation to create glyph maps, \texttt{geom\_glyph()} and it takes four required aestheticsf : \texttt{x\_major}, \texttt{y\_major}, \texttt{x\_minor}, and \texttt{y\_minor}:
+Visualizing both space and time is important for exploring and understanding the data more completely, aiding in decision-making, and facilitating effective communication.
+Several approaches are common: facet maps across time, map animations, or interactive graphics that link maps and time series plots among others.
+Faceted maps and spatio-temporal animations focus on the spatial pattern making it difficult to assess temporal trends.
+The glyph map \citep{Wickham2012-yr} addresses this issue by plotting time series onto the map as a glyph.
+
+It is achieved through a coordinate transformation.
+The transformation uses linear algebra to convert the temporal coordinates (minor coordinates) into the spatial coordinates (major coordinates) and is implemented in the package \pkg{GGally} \citep{ggally}.
+The \pkg{cubble} package provides a new \code{ggproto} implementation to create glyph maps, \code{geom_glyph()} requiring four aesthetics : \code{x_major}, \code{y_major}, \code{x_minor}, and \code{y_minor}:
 

### 2023-07-03 cubble.Rmd
section 4.1

@@
 
-Five examples are chosen to illustrate different aspects of the \pkg{cubble} package: creating a \code{cubble} object from two Coronavirus (COVID) data tables with the challenge of having different location names, using spatial transformations to make a glyph map of seasonal temperature changes, matching river level data with weather station records to analyze water supply, reading NetCDF format data to replicate a climate reanalysis plot, and demonstrating the workflow to create complex interactive linked plots.
+Five examples are chosen to illustrate different aspects of the \pkg{cubble} package: creating a \code{cubble} object from two Coronavirus (COVID) data tables with the challenge of having different location names, using spatial transformations to make a glyph map of seasonal temperature changes, matching river level data with weather station records to analyze water supply, reading NetCDF format data to replicate a climate reanalysis plot, and demonstrating the workflow to create interactively linked plots.
 
@@
 
-Since the start of the COVID-19 pandemic, the Victoria State Government in Australia has been providing daily COVID-19 case counts per Local Government Area (LGA).
+Since the start of the COVID-19 pandemic, the Victoria State Government in Australia has been providing daily COVID-19 case counts per local government area (LGA).
 This data can be combined with map polygon data, available from the Australian Bureau of Statistics (ABS), to visualize COVID-19 incidence and spread.
 The COVID-19 count data (\code{covid}) and the LGA information (\code{lga}) are available in the \pkg{cubble} package as a \code{tsibble} object and an \code{sf} object, respectively.
-This example illustrates the creation of a cubble object using real-world data, where analysts may deal with slightly varied encoding of the spatial units used across different institutes.
-Cubble can flag this discrepancies when creating the object and guide analysts to further investigate this issue.
+As is common, the different agencies have some difference in text id's labelling the spatial regions, so this example illustrates how this can be caught with \pkg{cubble}, and fixed.
+Discrepancies are flagged when creating the cubble object, notifying analysts of something that needs checking.
 
-The function \code{make_cubble()} is used to create a cubble object from separate spatial and temporal components.
-When the spatial identifier in the two data has different names, they can be handled by the \code{by} argument:
+The \code{by} argument of the function \code{make_cubble()} is used to specify the spatial identifier in the two data sets:
 
@@
 
-The difference in LGA naming between both data sets triggers cubble to issue a warning, alerting the user to this discrepancy.
-The warning message suggests there are some differences between the LGA encoding used by Victoria government and ABS and prompts analysts to check the mismatch using \code{check_key()}, which takes the same inputs as \code{make_cubble()}, but returns a summary of key matches between the spatial and temporal input data.
+The difference in LGA naming between both data sets triggers a warning, alerting the user to this discrepancy.
+The warning message suggests there are some differences between the LGA encoding used by Victoria government and ABS.
+The mismatches can be checked using \code{check_key()}, which takes the same inputs as \code{make_cubble()}, but returns a summary of key matches between the spatial and temporal input data:
 
@@
 The result of the \code{check_key()} function is a list containing three elements: 1) matched keys from both tables, 2) potentially paired keys, and 3) others keys that can't be matched.
-Here, the main mismatch arises from the two LGAs: Kingston and Latrobe (Kingston is a LGA in both Victoria and South Australia and Latrobe is a LGA in both Victoria and Tasmania).
-Analysts can then reconcile the spatial and temporal data based on this check summary and recreate the cubble object:
+Here, the main mismatch arises from the two LGAs: Kingston and Latrobe (Kingston is an LGA in both Victoria and South Australia and Latrobe is an LGA in both Victoria and Tasmania).
+Analysts can then reconcile the spatial and temporal data based on this summary and recreate the cubble object:
 

### 2023-07-03 cubble.tex
section 4.1

@@
 
-Since the start of the COVID-19 pandemic, the Victoria State Government in Australia has been providing daily COVID-19 case counts per Local Government Area (LGA).
+Since the start of the COVID-19 pandemic, the Victoria State Government in Australia has been providing daily COVID-19 case counts per local government area (LGA).
 This data can be combined with map polygon data, available from the Australian Bureau of Statistics (ABS), to visualize COVID-19 incidence and spread.
 The COVID-19 count data (\code{covid}) and the LGA information (\code{lga}) are available in the \pkg{cubble} package as a \code{tsibble} object and an \code{sf} object, respectively.
-This example illustrates the creation of a cubble object using real-world data, where analysts may deal with slightly varied encoding of the spatial units used across different institutes.
-Cubble can flag this discrepancies when creating the object and guide analysts to further investigate this issue.
-
-The function \code{make_cubble()} is used to create a cubble object from separate spatial and temporal components.
-When the spatial identifier in the two data has different names, they can be handled by the \code{by} argument:
+As is common, the different agencies have some difference in text id's labelling the spatial regions, so this example illustrates how this can be caught with \pkg{cubble}, and fixed.
+Discrepancies are flagged when creating the cubble object, notifying analysts of something that needs checking.
+
+The \code{by} argument of the function \code{make_cubble()} is used to specify the spatial identifier in the two data sets:
 

### 2023-07-04 cubble.Rmd
Update section 3.1

@@
 A useful task in spatio-temporal data analysis is to combine related temporal series within a close geographic neighborhood.
-For example, we may want to examine data from weather stations with water flow records in nearby river sensors to understand how precipitation relates to water supply.
-This might be useful for better predicting potential for drought and flood.
+For example, we may want to examine data from weather stations with water flow records from nearby river sensors to understand how precipitation relates to river levels.
+This might be useful for predicting potential for droughts and floods.
 
-This problem can be tackled by first matching spatial locations across two datasets and then calculating the temporal features of the multiple series.
+Matching temporal data across different locations from different data sources could be done by initially identifying the corresponding spatial locations between the two data sets.
+Subsequently, a set of temporal features can be calculated for the series at the selected locations that can be used to matched the selected time series across locations.
 In cubble, locations from two datasets can be matched using the function `match_spatial()`.
-The function calculates the distance matrix of the two datasets and returns `spatial_n_group` groups with the smallest distances.
-For each group, analysts can include more locations with the argument `spatial_n_each` (default to 1 for one-on-one matching).
+The functi on calculates the distance matrix of the locations between the two data sets and returns  groups (`spatial_n_group`) with the smallest distances.
+For a given group, it is possible can include more locations with the argument `spatial_n_each` (default to 1 for one-on-one matching).
 
-The temporal matching then calculates a similarity score between the time series of spatially matched pairs using the function `match_temporal()`.
+For the temporal matching a similarity score between the time series of spatially matched pairs is computed using the function `match_temporal()`.
+The similarity score is computed by a matching function which can be customized to any desired time series feature.
+The function `match_temporal()` takes as argument two time series in the form of a list and returns a single numerical value.
+By default, cubble uses a simple peak matching algorithm (`match_peak`) to count the number of peaks in two time series that fall within a specified time window.
+
 The temporal matching requires two identifiers: one for separating each spatially matched group: `match_id` and one for separating the two data sources: `data_id`.
@@
 ````
-
-The similarity score is computed by a matching function which can be customized to any desired time series feature.
-The matching function takes two time series as a list and returns a single numerical value.
-By default, cubble uses a simple peak matching algorithm (`match_peak`) to count the number of peaks in two time series that fall within a specified time window.
 

### 2023-07-04 cubble.tex
Update section 3.1

@@
 A useful task in spatio-temporal data analysis is to combine related temporal series within a close geographic neighborhood.
-For example, we may want to examine data from weather stations with water flow records in nearby river sensors to understand how precipitation relates to water supply.
-This might be useful for better predicting potential for drought and flood.
-
-This problem can be tackled by first matching spatial locations across two datasets and then calculating the temporal features of the multiple series.
+For example, we may want to examine data from weather stations with water flow records from nearby river sensors to understand how precipitation relates to river levels.
+This might be useful for predicting potential for droughts and floods.
+
+Matching temporal data across different locations from different data sources could be done by initially identifying the corresponding spatial locations between the two data sets.
+Subsequently, a set of temporal features can be calculated for the series at the selected locations that can be used to matched the selected time series across locations.
 In cubble, locations from two datasets can be matched using the function \texttt{match\_spatial()}.
-The function calculates the distance matrix of the two datasets and returns \texttt{spatial\_n\_group} groups with the smallest distances.
-For each group, analysts can include more locations with the argument \texttt{spatial\_n\_each} (default to 1 for one-on-one matching).
-
-The temporal matching then calculates a similarity score between the time series of spatially matched pairs using the function \texttt{match\_temporal()}.
+The functi on calculates the distance matrix of the locations between the two data sets and returns groups (\texttt{spatial\_n\_group}) with the smallest distances.
+For a given group, it is possible can include more locations with the argument \texttt{spatial\_n\_each} (default to 1 for one-on-one matching).
+
+For the temporal matching a similarity score between the time series of spatially matched pairs is computed using the function \texttt{match\_temporal()}.
+The similarity score is computed by a matching function which can be customized to any desired time series feature.
+The function \texttt{match\_temporal()} takes as argument two time series in the form of a list and returns a single numerical value.
+By default, cubble uses a simple peak matching algorithm (\texttt{match\_peak}) to count the number of peaks in two time series that fall within a specified time window.
+
 The temporal matching requires two identifiers: one for separating each spatially matched group: \texttt{match\_id} and one for separating the two data sources: \texttt{data\_id}.
@@
 
-The similarity score is computed by a matching function which can be customized to any desired time series feature.
-The matching function takes two time series as a list and returns a single numerical value.
-By default, cubble uses a simple peak matching algorithm (\texttt{match\_peak}) to count the number of peaks in two time series that fall within a specified time window.
-
 \hypertarget{interactive-graphics}{%
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/cookd/students_PhD/Sherry/paper-cubble/figures/glyphmap-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Desktop/paper-cubble/figures/glyphmap-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/cookd/students_PhD/Sherry/paper-cubble/figures/matching-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Desktop/paper-cubble/figures/matching-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/cookd/students_PhD/Sherry/paper-cubble/figures/netcdf-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Desktop/paper-cubble/figures/netcdf-1}
 

### 2023-07-04 cubble.Rmd
Update section 4.2

@@
 
-Daily maximum temperature is then averaged into monthly series for each periods in the temporal cubble.
-The last step with \code{unfold()} moves the two coordinate columns (\code{long, lat}) into the temporal cubble, preparing the data for the glyph map:
+The daily maximum temperature is then averaged into monthly series for each period within the temporal cube.
+In the code above, the last step with \code{unfold()} moves the two coordinate columns (\code{long, lat}) into the temporal cubble, preparing the data for the construction of a glyph map:
 
@@
 
-A quick check on the number of observations for each location is made, revealing that there are several with less than 24 observations -- these stations lack temperature values for some months.
+The code below counts the number of observations for each location, revealing that there are several with less than 24 observations -- these stations lack temperature values for some months.
 In this example, those stations are removed by switching to the spatial cubble to operate on the spatial component over time, and then, move back into the temporal cubble (to make the glyph map):

### 2023-07-04 cubble.Rmd
Update section 4.4

@@
 
-The Bureau of Meteorology collects water level data and they can be matched with the precipitation data from the climate weather stations.
+The Bureau of Meteorology collects water level data that can be matched with precipitation data from climate weather stations.
 The data \code{river} from the \pkg{cubble} package contains water course level data for 71 river gauges collected in Victoria, Australia.
 Victoria weather station data can be subsetted from the \code{climate_aus} data in the \pkg{cubble} package.
-The example demonstrates the use of matching function introduced in Section \ref{matching} to detect river gauges that can reflects precipitation in Victoria.
+This example demonstrates the use of the matching function introduced in Section \ref{matching} to find river gauges that mirror changes in precipitation regimes captured in the climate weather stations in Victoria.
 
@@
 
-Spatial match on the site location is first performed since rainfall can directly impact water level in the nearby river. With \code{match_spatial()}, we can obtain a summary of the 10 closest pairs:
+The first step is to perform a spatial match on the site locations between both data sets.
+The rainfall recorded at a particular station can directly impact the water level in nearby rivers, which emphasizes the need to find related or matching locations.
+With \code{match_spatial()}, we can obtain a summary of the 10 closest pairs of weather stations and river gauges:
 
@@
 
-The result can also be returned as a list of matched cubbles, by setting the argument \code{return_cubble = TRUE}.
-All the results can be combined into a single cubble using \code{bind_rows()}, after excluding the two pairs where a river station is matched to more than one weather stations (river station 226027 is matched twice in group 3 and 5 and similarly for station 230200 in group 7 and 8).
+The matching or pairing results can also be returned as a list of matched cubbles, by setting the argument \code{return_cubble = TRUE}.
+After excluding the two pairs where a river station is matched to more than one weather stations (river station 226027 is matched twice in group 3 and 5 and similarly for station 230200 in group 7 and 8), all the results can be combined into a single cubble using \code{bind_rows()},
 
@@
 
-To match the water level series with precipitation, the function \code{match_temporal()} is used with the variable \code{group} and \code{types} identifying the matching group and the two data sources:
+To match the water level and precipitation time series across the matched locations, the function \code{match_temporal()} is used with the variables \code{group} and \code{types} identifying the matching group and the two data sources:
 
@@
 Similarly, the cubble output can be returned using the argument \code{return_cubble = TRUE}.
-Here we select the four pairs with the highest number of matching peaks and show them on the map (a) and as standardized series (b) in Figure \ref{fig:matching}.
+Here, we select the four pairs of time series (precipitation/water level) with the highest number of matching peaks and show them on the map (a).
+The time series of river levels is standardize for better comparison  in Figure \ref{fig:matching} panel (b).
 

### 2023-07-04 cubble.tex
Update section 4.4

@@
 Similarly, the cubble output can be returned using the argument \code{return_cubble = TRUE}.
-Here we select the four pairs with the highest number of matching peaks and show them on the map (a) and as standardized series (b) in Figure \ref{fig:matching}.
+Here, we select the four pairs of time series (precipitation/water level) with the highest number of matching peaks and show them on the map (a).
+The time series of river levels is standardize for better comparison in Figure \ref{fig:matching} panel (b).
 

### 2023-07-04 cubble.Rmd
Update section 4.5

@@
 
-Interactive graphics can be especially useful for spatio-temporal data because they make it possible to look at the data in multiple ways on-the-fly.
-The last example describes the process of using cubble with the \pkg{crosstalk} package to build an interactive display connecting a map of Australia, with ribbon plots of temperature range observed at the stations in 2020.
+Interactive graphics can be especially useful for spatio-temporal data because they make it possible to look at the data in multiple ways on-the-fly. This last example describes the process of using 
+\pkg{cubble} with the \pkg{crosstalk} package to build an interactive display connecting a map of Australia, with ribbon plots of temperature range observed at a group of stations in 2020.
 The purpose is to explore the variation of monthly temperature range over the country.
 
-We will first summarise the daily data in \code{climate_aus} into monthly average and calculate the variance of the temperature difference, which will be used to color the temperature band later.
+Firstly, we summarise the daily data in \code{climate_aus} into monthly averages and calculate the variance of the monthly averages differences between the minimum and maximum temperatures.
+This variance will be used to color the temperature band later.
 

### 2023-07-04 cubble.tex
Update section 4.5

@@
 
-Interactive graphics can be especially useful for spatio-temporal data because they make it possible to look at the data in multiple ways on-the-fly.
-The last example describes the process of using cubble with the \pkg{crosstalk} package to build an interactive display connecting a map of Australia, with ribbon plots of temperature range observed at the stations in 2020.
+Interactive graphics can be especially useful for spatio-temporal data because they make it possible to look at the data in multiple ways on-the-fly. This last example describes the process of using
+\pkg{cubble} with the \pkg{crosstalk} package to build an interactive display connecting a map of Australia, with ribbon plots of temperature range observed at a group of stations in 2020.
 The purpose is to explore the variation of monthly temperature range over the country.
 
-We will first summarise the daily data in \code{climate_aus} into monthly average and calculate the variance of the temperature difference, which will be used to color the temperature band later.
+Firstly, we summarise the daily data in \code{climate_aus} into monthly averages and calculate the variance of the monthly averages differences between the minimum and maximum temperatures.
+This variance will be used to color the temperature band later.
 

### 2023-07-04 cubble.Rmd
Update conclusion

@@
 This paper presents the \proglang{R} package \pkg{cubble} for organizing, wrangling and visualizing spatio-temporal data.
-The package introduces a new data structure, \code{cubble}, consisting of two subclass, spatial cubble and a temporal cubble, to organise spatio-temporal data in two different formats within the tidy data framework.
+The package introduces a new data structure, \code{cubble}, consisting of two subclasses, spatial cubble and a temporal cubble, to organise spatio-temporal data in two different formats within the tidy data framework.
 The data structure and functions introduced in the package can be used and combined with existing tools for  data wrangling, spatial and temporal data analysis, and visualization.
 
-The paper includes plenty examples to illustrate the utility of \pkg{cubble} as a data structure for spatio-temporal analysis.
-These examples cover different tasks of a typical analysis workflow: handling data with spatial and temporal misalignment, matching data from multiple sources, and creating both static and interactive spatio-temporal visualisation.
+The paper includes several examples to illustrate the utility of \pkg{cubble} as a data structure for spatio-temporal analysis.
+These examples cover different tasks of a typical data analysis workflow involving spatio-temporal data: handling data with spatial and temporal misalignment, matching data from multiple sources, and creating both static and interactive spatio-temporal visualisation.
 For future directions, other commonly-used spatial or temporal data structures can be integrated into the \pkg{cubble} package to extend analysts' familiar spatial and temporal toolkit to spatio-temporal.

### 2023-07-04 cubble.tex
Update conclusion

@@
 This paper presents the \proglang{R} package \pkg{cubble} for organizing, wrangling and visualizing spatio-temporal data.
-The package introduces a new data structure, \code{cubble}, consisting of two subclass, spatial cubble and a temporal cubble, to organise spatio-temporal data in two different formats within the tidy data framework.
+The package introduces a new data structure, \code{cubble}, consisting of two subclasses, spatial cubble and a temporal cubble, to organise spatio-temporal data in two different formats within the tidy data framework.
 The data structure and functions introduced in the package can be used and combined with existing tools for data wrangling, spatial and temporal data analysis, and visualization.
 
-The paper includes plenty examples to illustrate the utility of \pkg{cubble} as a data structure for spatio-temporal analysis.
-These examples cover different tasks of a typical analysis workflow: handling data with spatial and temporal misalignment, matching data from multiple sources, and creating both static and interactive spatio-temporal visualisation.
+The paper includes several examples to illustrate the utility of \pkg{cubble} as a data structure for spatio-temporal analysis.
+These examples cover different tasks of a typical data analysis workflow involving spatio-temporal data: handling data with spatial and temporal misalignment, matching data from multiple sources, and creating both static and interactive spatio-temporal visualisation.
 For future directions, other commonly-used spatial or temporal data structures can be integrated into the \pkg{cubble} package to extend analysts' familiar spatial and temporal toolkit to spatio-temporal.

### 2023-07-04 cubble.Rmd
use \code{} markup

@@
 
--   a spatial cubble with class `c("spatial_cubble_df", "cubble_df")`
--   a temporal cubble with class `c("temporal_cubble_df", "cubble_df")`
+-   a spatial cubble with class \code{c("spatial_cubble_df", "cubble_df")}
+-   a temporal cubble with class \code{c("temporal_cubble_df", "cubble_df")}
 
-In a spatial cubble object, spatial variables are organised as columns and temporal variables are nested within a specialised `ts` column.
+In a spatial cubble object, spatial variables are organised as columns and temporal variables are nested within a specialised \code{ts} column.
 For example, the spatial cubble object, \code{cb_spatial} printed below, contains weather records of three airport stations from the Global Historical Climatology Network Daily (GHCND) database [@menne2012overview].
@@
 Subsequently, a set of temporal features can be calculated for the series at the selected locations that can be used to matched the selected time series across locations.
-In cubble, locations from two datasets can be matched using the function `match_spatial()`.
-The functi on calculates the distance matrix of the locations between the two data sets and returns  groups (`spatial_n_group`) with the smallest distances.
-For a given group, it is possible can include more locations with the argument `spatial_n_each` (default to 1 for one-on-one matching).
+In cubble, locations from two datasets can be matched using the function \code{match_spatial()}.
+The functi on calculates the distance matrix of the locations between the two data sets and returns  groups (\code{spatial_n_group}) with the smallest distances.
+For a given group, it is possible can include more locations with the argument \code{spatial_n_each} (default to 1 for one-on-one matching).
 
-For the temporal matching a similarity score between the time series of spatially matched pairs is computed using the function `match_temporal()`.
+For the temporal matching a similarity score between the time series of spatially matched pairs is computed using the function \code{match_temporal()}.
 The similarity score is computed by a matching function which can be customized to any desired time series feature.
-The function `match_temporal()` takes as argument two time series in the form of a list and returns a single numerical value.
-By default, cubble uses a simple peak matching algorithm (`match_peak`) to count the number of peaks in two time series that fall within a specified time window.
+The function \code{match_temporal()} takes as argument two time series in the form of a list and returns a single numerical value.
+By default, cubble uses a simple peak matching algorithm (\code{match_peak}) to count the number of peaks in two time series that fall within a specified time window.
 
-The temporal matching requires two identifiers: one for separating each spatially matched group: `match_id` and one for separating the two data sources: `data_id`.
-Matching between different variables can be specified using the `temporal_by` argument, similar to the `by` syntax from dplyr's `*_join`.
+The temporal matching requires two identifiers: one for separating each spatially matched group: \code{match_id} and one for separating the two data sources: \code{data_id}.
+Matching between different variables can be specified using the \code{temporal_by} argument, similar to the \code{by} syntax from dplyr's \code{*_join}.
 

### 2023-07-04 cubble.Rmd
finally caption under table, phewwwwwwwwww

@@
 
-| Category  | Functions |
-|--------|--------------------------------------------------------|
-| base R    | `[`, `[[<-`, `names<-` |
-| tidyverse | `dplyr_row_slice`, `dplyr_col_modify`, `dplyr_reconstruct`, `select`, `mutate`, `arrange`, `filter`, `group_by`, `ungroup`, `summarise`, `select`, `slice`, `rowwise`, `rename`, `bind_rows`, `bind_cols`, `relocate`, `type_sum`, the slice family (`slice_head`, `slice_tail`, `slice_max`, `slice_min`, `slice_sample`) and the join family (`left_join`, `right_join`, `inner_join`, `full_join`, `anti_join`, `semi_join`) |
-| cubble    | `as_cubble`, `cubble`, `make_cubble`, `check_key`, `face_temporal`, `face_spatial`, `unfold`, `key`, `key_vars`, `key_data`, `index`, `index_var`, `coords`, `spatial`, `match_sites`, `match_spatial`, `match_temporal`, `geom_glyph`, `geom_glyph_box`, `geom_glyph_line`, `make_spatial_sf`, `make_temporal_tsibble`, `fill_gaps`, and `scan_gaps`                                                                                |
+\begin{table}
+\begin{tabular}{p{0.1\textwidth}p{0.8\textwidth}}
+\hline
+Category & Functions \\
+\hline
+base R & \texttt{{[}}, \texttt{{[}{[}\textless{}-}, \texttt{names\textless{}-} \\
+tidyverse & \texttt{dplyr\_row\_slice}, \texttt{dplyr\_col\_modify}, \texttt{dplyr\_reconstruct}, \texttt{select}, \texttt{mutate}, \texttt{arrange}, \texttt{filter}, \texttt{group\_by}, \texttt{ungroup}, \texttt{summarise}, \texttt{select}, \texttt{slice}, \texttt{rowwise}, \texttt{rename}, \texttt{bind\_rows}, \texttt{bind\_cols}, \texttt{relocate}, \texttt{type\_sum}, the slice family (\texttt{slice\_head}, \texttt{slice\_tail}, \texttt{slice\_max}, \texttt{slice\_min}, \texttt{slice\_sample}) and the join family (\texttt{left\_join}, \texttt{right\_join}, \texttt{inner\_join}, \texttt{full\_join}, \texttt{anti\_join}, \texttt{semi\_join}) \\
+cubble & \texttt{as\_cubble}, \texttt{cubble}, \texttt{make\_cubble}, \texttt{check\_key}, \texttt{face\_temporal}, \texttt{face\_spatial}, \texttt{unfold}, \texttt{key}, \texttt{key\_vars}, \texttt{key\_data}, \texttt{index}, \texttt{index\_var}, \texttt{coords}, \texttt{spatial}, \texttt{match\_sites}, \texttt{match\_spatial}, \texttt{match\_temporal}, \texttt{geom\_glyph}, \texttt{geom\_glyph\_box}, \texttt{geom\_glyph\_line}, \texttt{make\_spatial\_sf}, \texttt{make\_temporal\_tsibble}, \texttt{fill\_gaps}, and \texttt{scan\_gaps} \\
+\hline
+\end{tabular}
+\caption{An overview of functions implemented in the cubble package, categorised into base R, tidyverse, and cubble functions.}
+\label{tab:funs}
+\end{table}
 
-: (#tab:funs) An overview of functions implemented in the \pkg{cubble} package, categorised into base R, tidyverse, and cubble functions.
-
-| Class                | Methods|
-|------------------|------------------------------------------------------|
-| `cubble_df`          | `[[<-, dplyr_col_modify, key_data, key_vars, key, print`                                                                                                                                                                                     |
-| `spatial_cubble_df`  | `[, names<-,  tbl_sum, dplyr_reconstruct, dplyr_row_slice, face_spatial, face_temporal, unfold, arrange, rename, rowwise, group_by, ungroup, select, spatial, summarise, unfold, update_cubble`                                              |
-| `temporal_cubble_df` | `[, names<-, tbl_sum, arrange, dplyr_reconstruct, dplyr_row_slice, face_spatial, face_temporal, unfold, fill_gaps, group_by, ungroup,  rename, rowwise, scan_gaps, select, spatial, summarise, tbl_sum, bind_rows, bind_cols, update_cubble` |
-
-: (#tab:methods) An overview of the methods implemented in the three \code{cubble} classes.
-Methods are implemented in the \code{cubble\_df} class when they behave consistently across the spatial and temporal cubble; otherwise, they are implemented separately.
+\begin{table}
+\begin{tabular}{p{0.25\textwidth}p{0.65\textwidth}}
+\hline
+Class & Method \\
+\hline
+\texttt{cubble\_df} & \texttt{{[}{[}\textless{}-,\ dplyr\_col\_modify,\ key\_data,\ key\_vars,\ key,\ print} \\
+\texttt{spatial\_cubble\_df} & \texttt{{[},\ names\textless{}-,\ \ tbl\_sum,\ dplyr\_reconstruct,\ dplyr\_row\_slice,\ face\_spatial,\ face\_temporal,\ unfold,\ arrange,\ rename,\ rowwise,\ group\_by,\ ungroup,\ select,\ spatial,\ summarise,\ unfold,\ update\_cubble} \\
+\texttt{temporal\_cubble\_df} & \texttt{{[},\ names\textless{}-,\ tbl\_sum,\ arrange,\ dplyr\_reconstruct,\ dplyr\_row\_slice,\ face\_spatial,\ face\_temporal,\ unfold,\ fill\_gaps,\ group\_by,\ ungroup,\ \ rename,\ rowwise,\ scan\_gaps,\ select,\ spatial,\ summarise,\ tbl\_sum,\ bind\_rows,\ bind\_cols,\ update\_cubble} \\
+\hline
+\end{tabular}
+\caption{An overview of the methods implemented in the three \code{cubble} classes.
+Methods are implemented in the \code{cubble\_df} class when they behave consistently across the spatial and temporal cubble; otherwise, they are implemented separately.}
+\label{tab:methods} 
+\end{table}
 
@@
 In cubble, locations from two datasets can be matched using the function \code{match_spatial()}.
-The functi on calculates the distance matrix of the locations between the two data sets and returns  groups (\code{spatial_n_group}) with the smallest distances.
+The function on calculates the distance matrix of the locations between the two data sets and returns  groups (\code{spatial_n_group}) with the smallest distances.
 For a given group, it is possible can include more locations with the argument \code{spatial_n_each} (default to 1 for one-on-one matching).

### 2023-07-04 cubble.tex
finally caption under table, phewwwwwwwwww

@@
 \item
-  a spatial cubble with class \texttt{c("spatial\_cubble\_df",\ "cubble\_df")}
+  a spatial cubble with class \code{c("spatial_cubble_df", "cubble_df")}
 \item
-  a temporal cubble with class \texttt{c("temporal\_cubble\_df",\ "cubble\_df")}
+  a temporal cubble with class \code{c("temporal_cubble_df", "cubble_df")}
 \end{itemize}
 
-In a spatial cubble object, spatial variables are organised as columns and temporal variables are nested within a specialised \texttt{ts} column.
+In a spatial cubble object, spatial variables are organised as columns and temporal variables are nested within a specialised \code{ts} column.
 For example, the spatial cubble object, \code{cb_spatial} printed below, contains weather records of three airport stations from the Global Historical Climatology Network Daily (GHCND) database \citep{menne2012overview}.
@@
 
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.1250}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.8750}}@{}}
-\caption{\label{tab:funs} An overview of functions implemented in the \pkg{cubble} package, categorised into base R, tidyverse, and cubble functions.}\tabularnewline
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-Category
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Functions
-\end{minipage} \\
-\midrule()
-\endfirsthead
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-Category
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Functions
-\end{minipage} \\
-\midrule()
-\endhead
+\begin{table}
+\begin{tabular}{p{0.1\textwidth}p{0.8\textwidth}}
+\hline
+Category & Functions \\
+\hline
 base R & \texttt{{[}}, \texttt{{[}{[}\textless{}-}, \texttt{names\textless{}-} \\
@@
 cubble & \texttt{as\_cubble}, \texttt{cubble}, \texttt{make\_cubble}, \texttt{check\_key}, \texttt{face\_temporal}, \texttt{face\_spatial}, \texttt{unfold}, \texttt{key}, \texttt{key\_vars}, \texttt{key\_data}, \texttt{index}, \texttt{index\_var}, \texttt{coords}, \texttt{spatial}, \texttt{match\_sites}, \texttt{match\_spatial}, \texttt{match\_temporal}, \texttt{geom\_glyph}, \texttt{geom\_glyph\_box}, \texttt{geom\_glyph\_line}, \texttt{make\_spatial\_sf}, \texttt{make\_temporal\_tsibble}, \texttt{fill\_gaps}, and \texttt{scan\_gaps} \\
-\bottomrule()
-\end{longtable}
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.2500}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.7500}}@{}}
-\caption{\label{tab:methods} An overview of the methods implemented in the three \code{cubble} classes.
-Methods are implemented in the \code{cubble\_df} class when they behave consistently across the spatial and temporal cubble; otherwise, they are implemented separately.}\tabularnewline
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-Class
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Methods
-\end{minipage} \\
-\midrule()
-\endfirsthead
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-Class
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Methods
-\end{minipage} \\
-\midrule()
-\endhead
+\hline
+\end{tabular}
+\caption{An overview of functions implemented in the cubble package, categorised into base R, tidyverse, and cubble functions.}
+\label{tab:funs}
+\end{table}
+
+\begin{table}
+\begin{tabular}{p{0.25\textwidth}p{0.65\textwidth}}
+\hline
+Class & Method \\
+\hline
 \texttt{cubble\_df} & \texttt{{[}{[}\textless{}-,\ dplyr\_col\_modify,\ key\_data,\ key\_vars,\ key,\ print} \\
@@
 \texttt{temporal\_cubble\_df} & \texttt{{[},\ names\textless{}-,\ tbl\_sum,\ arrange,\ dplyr\_reconstruct,\ dplyr\_row\_slice,\ face\_spatial,\ face\_temporal,\ unfold,\ fill\_gaps,\ group\_by,\ ungroup,\ \ rename,\ rowwise,\ scan\_gaps,\ select,\ spatial,\ summarise,\ tbl\_sum,\ bind\_rows,\ bind\_cols,\ update\_cubble} \\
-\bottomrule()
-\end{longtable}
+\hline
+\end{tabular}
+\caption{An overview of the methods implemented in the three \code{cubble} classes.
+Methods are implemented in the \code{cubble\_df} class when they behave consistently across the spatial and temporal cubble; otherwise, they are implemented separately.}
+\label{tab:methods} 
+\end{table}
 
@@
 Subsequently, a set of temporal features can be calculated for the series at the selected locations that can be used to matched the selected time series across locations.
-In cubble, locations from two datasets can be matched using the function \texttt{match\_spatial()}.
-The functi on calculates the distance matrix of the locations between the two data sets and returns groups (\texttt{spatial\_n\_group}) with the smallest distances.
-For a given group, it is possible can include more locations with the argument \texttt{spatial\_n\_each} (default to 1 for one-on-one matching).
-
-For the temporal matching a similarity score between the time series of spatially matched pairs is computed using the function \texttt{match\_temporal()}.
+In cubble, locations from two datasets can be matched using the function \code{match_spatial()}.
+The function on calculates the distance matrix of the locations between the two data sets and returns groups (\code{spatial_n_group}) with the smallest distances.
+For a given group, it is possible can include more locations with the argument \code{spatial_n_each} (default to 1 for one-on-one matching).
+
+For the temporal matching a similarity score between the time series of spatially matched pairs is computed using the function \code{match_temporal()}.
 The similarity score is computed by a matching function which can be customized to any desired time series feature.
-The function \texttt{match\_temporal()} takes as argument two time series in the form of a list and returns a single numerical value.
-By default, cubble uses a simple peak matching algorithm (\texttt{match\_peak}) to count the number of peaks in two time series that fall within a specified time window.
-
-The temporal matching requires two identifiers: one for separating each spatially matched group: \texttt{match\_id} and one for separating the two data sources: \texttt{data\_id}.
-Matching between different variables can be specified using the \texttt{temporal\_by} argument, similar to the \texttt{by} syntax from dplyr's \texttt{*\_join}.
+The function \code{match_temporal()} takes as argument two time series in the form of a list and returns a single numerical value.
+By default, cubble uses a simple peak matching algorithm (\code{match_peak}) to count the number of peaks in two time series that fall within a specified time window.
+
+The temporal matching requires two identifiers: one for separating each spatially matched group: \code{match_id} and one for separating the two data sources: \code{data_id}.
+Matching between different variables can be specified using the \code{temporal_by} argument, similar to the \code{by} syntax from dplyr's \code{*_join}.
 

### 2023-07-04 cubble.Rmd
section 4.2, plus some unmarked \pkg

@@
 In the \pkg{cubble} package, the cubble object \code{historical_tmax} contains daily maximum temperature data for `r nrow(historical_tmax)` stations in Australia, covering two periods: 1971-1975 and 2016-2020.
-This example uses dplyr verbs to wrangle a cubble object, which is pivoted between the spatial and temporal form during the analysis.
-Glyph maps are then created with ggplot2 to compare the changes in temperature between these two periods.
+This example uses \pkg{dplyr} verbs to wrangle a cubble object, and  pivot between the spatial and temporal form for different parts of the analysis.
+The result is glyph maps to compare the changes in temperature between these two periods, created with \pkg{ggplot2}.
 
@@
 
-The following code creates the glyph map (a) in Figure \ref{fig:glyphmap} (additional codes are needed for highlighting the single station, Cobar and styling) and the glyph map (c) is produced similarly with the difference series, rather than the two original series, being plotted.
+The following code creates the glyph map (a) in Figure \ref{fig:glyphmap} (additional codes are needed for highlighting the single station, Cobar and styling) and the glyph map (c) is produced similarly after further processing the data.
 

### 2023-07-04 cubble.tex
section 4.2, plus some unmarked \pkg

@@
 In the \pkg{cubble} package, the cubble object \code{historical_tmax} contains daily maximum temperature data for 75 stations in Australia, covering two periods: 1971-1975 and 2016-2020.
-This example uses dplyr verbs to wrangle a cubble object, which is pivoted between the spatial and temporal form during the analysis.
-Glyph maps are then created with ggplot2 to compare the changes in temperature between these two periods.
+This example uses \pkg{dplyr} verbs to wrangle a cubble object, and pivot between the spatial and temporal form for different parts of the analysis.
+The result is glyph maps to compare the changes in temperature between these two periods, created with \pkg{ggplot2}.
 

### 2023-07-04 cubble.Rmd
section 4.3

@@
 
-The matching or pairing results can also be returned as a list of matched cubbles, by setting the argument \code{return_cubble = TRUE}.
+The results can also be returned as a list of matched cubbles, by setting the argument \code{return_cubble = TRUE}.
 After excluding the two pairs where a river station is matched to more than one weather stations (river station 226027 is matched twice in group 3 and 5 and similarly for station 230200 in group 7 and 8), all the results can be combined into a single cubble using \code{bind_rows()},
@@
 
-To match the water level and precipitation time series across the matched locations, the function \code{match_temporal()} is used with the variables \code{group} and \code{types} identifying the matching group and the two data sources:
+To match the water level and precipitation time series across the matched locations, the function \code{match_temporal()} is used with the variables \code{group} and \code{type} identifying the matching group and the two data sources:
 
@@
 Similarly, the cubble output can be returned using the argument \code{return_cubble = TRUE}.
-Here, we select the four pairs of time series (precipitation/water level) with the highest number of matching peaks and show them on the map (a).
-The time series of river levels is standardize for better comparison  in Figure \ref{fig:matching} panel (b).
+Here, we select the four pairs of time series (precipitation/water level) with the highest number of matching peaks and show them on the map (Figure \ref{fig:matching} a).
+The time series of river levels is standardized to make the comparison easier in panel (b).
 

### 2023-07-04 cubble.tex
section 4.3

@@
 Similarly, the cubble output can be returned using the argument \code{return_cubble = TRUE}.
-Here, we select the four pairs of time series (precipitation/water level) with the highest number of matching peaks and show them on the map (a).
-The time series of river levels is standardize for better comparison in Figure \ref{fig:matching} panel (b).
+Here, we select the four pairs of time series (precipitation/water level) with the highest number of matching peaks and show them on the map (Figure \ref{fig:matching} a).
+The time series of river levels is standardized to make the comparison easier in panel (b).
 

### 2023-07-05 cubble.Rmd
finalising conclusion

@@
 
-The paper includes several examples to illustrate the utility of \pkg{cubble} as a data structure for spatio-temporal analysis.
-These examples cover different tasks of a typical data analysis workflow involving spatio-temporal data: handling data with spatial and temporal misalignment, matching data from multiple sources, and creating both static and interactive spatio-temporal visualisation.
-For future directions, other commonly-used spatial or temporal data structures can be integrated into the \pkg{cubble} package to extend analysts' familiar spatial and temporal toolkit to spatio-temporal.
+The paper includes several examples to illustrate how \pkg{cubble} is useful for spatio-temporal analysis.
+These examples cover different tasks of a typical data analysis workflow: handling data with spatial and temporal misalignment, matching data from multiple sources, and creating both static and interactive graphics.
+In addition, a re-working of an existing climate reanalysis using cubble is explained.
+
+Possible future directions, other commonly-used spatial or temporal data structures can be integrated into the \pkg{cubble} package to extend analysts' familiar spatial and temporal toolkit to spatio-temporal.
 

### 2023-07-05 cubble.Rmd
conclusion

@@
 
-Possible future directions, other commonly-used spatial or temporal data structures can be integrated into the \pkg{cubble} package to extend analysts' familiar spatial and temporal toolkit to spatio-temporal.
+Possible future directions would be in two main directions: handling much larger data sets, and integrating smoothly with modeling.
+Spatio-temporal data can be huge, and the current recommendation is to first reduce the spatial and temporal components before constructing a cubble.
+Better support could be provided with additional pre-processing functions.
+Considerable effort has been exerted to provide tidy tools for models, with the \pkg{tidymodels} project.
+It provides a much better interface to the vast array of statistical and machine learning model architecture.
+A similar direction would be to extend cubble to provide a more unified interface to spatio-temporal modeling.
 

### 2023-07-05 cubble.tex
conclusion

@@
 
-The paper includes several examples to illustrate the utility of \pkg{cubble} as a data structure for spatio-temporal analysis.
-These examples cover different tasks of a typical data analysis workflow involving spatio-temporal data: handling data with spatial and temporal misalignment, matching data from multiple sources, and creating both static and interactive spatio-temporal visualisation.
-For future directions, other commonly-used spatial or temporal data structures can be integrated into the \pkg{cubble} package to extend analysts' familiar spatial and temporal toolkit to spatio-temporal.
+The paper includes several examples to illustrate how \pkg{cubble} is useful for spatio-temporal analysis.
+These examples cover different tasks of a typical data analysis workflow: handling data with spatial and temporal misalignment, matching data from multiple sources, and creating both static and interactive graphics.
+In addition, a re-working of an existing climate reanalysis using cubble is explained.
+
+Possible future directions would be in two main directions: handling much larger data sets, and integrating smoothly with modeling.
+Spatio-temporal data can be huge, and the current recommendation is to first reduce the spatial and temporal components before constructing a cubble.
+Better support could be provided with additional pre-processing functions.
+Considerable effort has been exerted to provide tidy tools for models, with the \pkg{tidymodels} project.
+It provides a much better interface to the vast array of statistical and machine learning model architecture.
+A similar direction would be to extend cubble to provide a more unified interface to spatio-temporal modeling.
 

### 2023-07-07 cubble.Rmd
finalise install instructions

@@
 
-```{r setup, echo = FALSE}
+```{r setup, echo = FALSE, purl= FALSE}
 knitr::opts_chunk$set(
@@
 library("units")
+library("here")
 ```
@@
 The software is available from the Comprehensive \proglang{R} Archive Network (CRAN) at https://CRAN.R-project.org/package=cubble.
-
-<!-- In this paper, we propose the \pkg{cubble} package to organise spatio-temporal data collected at unique fixed locations while allowing for irregularities in the temporal dimension, such as weather station data.
--->
-
-<!-- Spatio-temporal data can encompass data with various spatial and temporal characteristics, requiring different structures for wrangling and analysis.
-For example, climate weather stations typically store station metadata in one table and the climate time series in another.
-GPS data tracks unique point locations at different timestamps and is represented as trajectories.
-Satellite imageries capture snapshots of landscapes at selected times and is commonly structured as raster data.
--->
 
@@
 ### The cubble attributes
-
-```{r class, fig.align="center", out.width = "100%", fig.cap = "Illustration of a cubble object in the long form and nested form, along with the associated attributes."}
-#knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.004.png"))
-```
 
@@
 
-```{r face, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "Illustration of main functions.
+```{r face, echo = FALSE, purl= FALSE, fig.align="center", out.width = "100%", fig.cap = "Illustration of main functions.
 To focus on he temporal variables \\code{face\\_temporal()} converts a spatial cubble into a temporal cubble.
@@
 
-```{r illu-interactive, echo = FALSE, fig.align="center", out.height = "35%", out.width = "100%", fig.cap = "Linking between multiple plots is made possible by shared \\code{crosstalk} objects.
+```{r illu-interactive, echo = FALSE, purl= FALSE, fig.align="center", out.height = "35%", out.width = "100%", fig.cap = "Linking between multiple plots is made possible by shared \\code{crosstalk} objects.
 When a station is selected on the map (a), the corresponding row in the spatial \\code{cubble} will be activated.
@@
   filter(NAME == "Victoria") |>
-  rmapshaper::ms_simplify()
+  st_transform("EPSG:28356") |> 
+  sf::st_simplify(dTolerance = 3000)
 
@@
   scale_color_brewer(palette = "Dark2")  + 
+  coord_sf(crs = "EPSG:4283") + 
   ggthemes::theme_map() +
@@
   ggplot() +
-  # q for specific humidity
   geom_point(aes(x = long, y = lat, color = q)) + 
-  #geom_tile(aes(x = long, y = lat, fill = q), color = "transparent") +
-  # z for geopotential
   geom_contour(data = res, aes(x = long, y = lat, z = z),
@@
 The bottom row first selects the lowest temperature in August in the seasonal display, which highlights the corresponding station on the map (Thredbo).
-Another  station, located in the Tasmania Island, is then selected to compare its temperature variation with the Thredbo station.", fig.show='hold'}
+Another  station, located in the Tasmania Island, is then selected to compare its temperature variation with the Thredbo station.", fig.show='hold', purl=FALSE}
 knitr::include_graphics(here::here("figures/linking.png"))

### 2023-07-07 install-instructions.R

@@ -3,14 +3,12 @@
 # The following R packages, and their dependencies, are needed to run the reproducible code:  
 
 packages <- c(
-	"cubble", "dplyr", "stringr", "ggplot2", "tibble", "knitr", 
-	"patchwork", "tsibble", "here", "lubridate", "ozmaps", 
-	"ggthemes", "ggrepel", "ncdf4", "sf", "rnaturalearth", 
-	"rnaturalearthdata", "colorspace", "units"
+  "dplyr", "ggplot2", "tibble", "patchwork", "tsibble", "lubridate", "ggthemes", 
+  "ggrepel", "ncdf4", "colorspace", "units", "here", "sf", "cubble", 
+  "rnaturalearth", "rnaturalearthdata", "ozmaps"
 )
 
 # They can be installed via:
 
 install.packages(packages, repos = "https://cloud.r-project.org/")
-
 # A cubble version >=0.3.0 is required for reproducing the paper and a dplyr version >= 1.0.8 is required for the cubble package.

### 2023-07-07 reproducible-script.R

@@ -1,17 +1,3 @@
-## ----setup, echo = FALSE--------------------------------------------------------
-knitr::opts_chunk$set(
-  echo = FALSE,
-  warning = FALSE,
-  message = FALSE,
-  cache = TRUE,
-  fig.path = here::here("figures/")
-)
-
-options(prompt = "R> ", continue = "+ ",
-        tibble.print_max = 5, tibble.print_min = 5,
-        width = 80, styler.cache_root = "styler-perm")
-
-
 ## ----echo = FALSE---------------------------------------------------------------
 library("cubble")
 library("dplyr")
@@ -28,30 +14,27 @@ library("ncdf4")
 library("rnaturalearth")
 library("colorspace")
 library("units")
+library("here")
 
 
 ## -------------------------------------------------------------------------------
-cb_nested <- climate_mel
+cb_spatial <- climate_mel
 
 
 ## ----echo = TRUE----------------------------------------------------------------
-cb_nested
+cb_spatial
 
 
 ## -------------------------------------------------------------------------------
-cb_long <- face_temporal(climate_mel)
+cb_temporal <- face_temporal(climate_mel)
 
 
 ## ----echo = TRUE----------------------------------------------------------------
-cb_long
-
-
-## ----class, fig.align="center", out.width = "100%", fig.cap = "Illustration of a cubble object in the long form and nested form, along with the associated attributes."----
-#knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.004.png"))
+cb_temporal
 
 
 ## ----echo = TRUE----------------------------------------------------------------
-spatial(cb_long)
+spatial(cb_temporal)
 
 
 ## ----echo = TRUE----------------------------------------------------------------
@@ -64,51 +47,50 @@ climate_flat |> as_cubble(key = id, index = date, coords = c(long, lat))
 
 
 ## ----cubble-fun, echo = TRUE----------------------------------------------------
-identical(face_temporal(cb_nested), cb_long)
-identical(face_spatial(cb_long), cb_nested)
+face_temporal(cb_spatial)
 
 
 ## ----cubble-fun2, echo = TRUE---------------------------------------------------
-identical(face_spatial(face_temporal(cb_nested)), cb_nested)
-identical(face_temporal(face_spatial(cb_long)), cb_long)
+face_spatial(face_temporal(cb_spatial))
 
 
[…]

### 2023-07-11 cubble.Rmd
minor chananges with arxiv submission

@@
 The cubble package makes linking between spatial and temporal plots easy."}
-knitr::include_graphics(here::here("figures/diagram-keynotes/‎diagram-keynotes.‎002.png")) 
+knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.002.png")) 
 ```

### 2023-07-11 cubble.tex
minor chananges with arxiv submission

--- 2023-07-05 cubble.tex
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.35\textheight]{figures/diagram-keynotes/‎diagram-keynotes.‎002}
+{\centering \includegraphics[width=1\linewidth,height=0.35\textheight]{figures/diagram-keynotes/diagram-keynotes.002}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/cookd/students_PhD/Sherry/paper-cubble/figures/glyphmap-1}
+{\centering \includegraphics[width=1\linewidth]{figures/glyphmap-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/cookd/students_PhD/Sherry/paper-cubble/figures/matching-1}
+{\centering \includegraphics[width=1\linewidth]{figures/matching-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/cookd/students_PhD/Sherry/paper-cubble/figures/netcdf-1}
+{\centering \includegraphics[width=1\linewidth]{figures/netcdf-1}
 

### 2023-11-03 cubble.Rmd
maintain consistency with the thesis

@@
 
-Spatio-temporal data has a spatial component referring to the location of each observation and a temporal component that is recorded at regular or irregular time intervals.
+Spatio-temporal data [@bivand2008applied; @lovelace2019geocomputation; @pebesma2019spatial] has a spatial component referring to the location of each observation and a temporal component that is recorded at regular or irregular time intervals.
 It may also include multiple variables measured at each spatial and temporal values.
@@
 
-## The cubble class {#object}
+## The cubble class {#sec-cb-object}
 
@@
 Another  station, located in the Tasmania Island, is then selected to compare its temperature variation with the Thredbo station.", fig.show='hold', purl=FALSE}
-knitr::include_graphics(here::here("figures/linking.png"))
+knitr::include_graphics(here::here("figures/linking-all.png"))
 knitr::include_graphics(here::here("figures/linking-north.png"))

### 2024-01-02 cubble.Rmd
work on review round 2

--- 2023-07-11 cubble.Rmd
@@
   - name: Patricia Menéndez
-    affiliation: Monash University
-    address: 21 Chancellors Walk, Clayton VIC 3800 Australia
-    email:  patricia.menendez@monash.edu
+    affiliation: University of Melbourne
+    affiliation2: Monash University
+    address: 813 Swanston Street, Parkville VIC 3052, Australia
+    email:  Patricia.menendez@unimelb.edu.au
 title:
@@
 An example of how the \pkg{cubble} package handles Network Common Data Form (NetCDF) data is also provided.
-Section \@ref(conclude) discuss the paper contributions and future directions.
+Section \@ref(conclude) discusses the paper contributions and future directions.
 
@@
 ```{r echo = TRUE}
+stations
+meteo
 make_cubble(spatial = stations, temporal = meteo,
@@
 The \pkg{cubble} package has several functions implemented for data wrangling and to facilitate data analysis as summarized in Table \@ref(tab:funs).
-In addition, for each of the three cubble classes there are number of methods implemented that facilitates the handling of the data as shown in Table \@ref(tab:methods).
+In addition, for each of the three cubble classes there are a number of methods implemented that facilitates the handling of the data as shown in Table \@ref(tab:methods).
 In particular, the \code{cubble_df} class handles methods that behave consistently in both spatial and temporal cubble.
@@
 The pair of cubble verbs, \code{face_temporal()} and \code{face_spatial()}, pivots the cubble object between its two forms or faces, as illustrated in Figure \ref{fig:face}.
-The code applies \code{face_temporal()} on the spatial cubble, \code{cb_spatial}, introduced in Section \@ref(object) to get a temporal cubble:
+The code applies \code{face_temporal()} on the spatial cubble, \code{cb_spatial}, introduced in Section \@ref(cb-object) to get a temporal cubble:
 
@@
 ```{r face, echo = FALSE, purl= FALSE, fig.align="center", out.width = "100%", fig.cap = "Illustration of main functions.
-To focus on he temporal variables \\code{face\\_temporal()} converts a spatial cubble into a temporal cubble.
+To focus on the temporal variables \\code{face\\_temporal()} converts a spatial cubble into a temporal cubble.
 To focus on the spatial variables \\code{face\\_spatial()} transforms a temporal cubble into a spatial cubble.
@@
 The \pkg{stars} package [@stars] uses an array structure, as opposed to a tibble, to represent multivariate spatio-temporal data.
-While both \pkg{stars} and \pkg{cubble} support vector and raster data, it is a matter of choice on which structure to use given the application.
+While both \pkg{stars} and \pkg{cubble} support vector and raster data, it is a matter of choice which structure to use given the application.
 Analysts working on satellite imageries may prefer the array structure in \pkg{stars}, while others originally working with spatio-temporal data in 2D data frames may find \pkg{cubble} easier to adopt from their existing computing workflow.
@@
 Matching temporal data across different locations from different data sources could be done by initially identifying the corresponding spatial locations between the two data sets.
-Subsequently, a set of temporal features can be calculated for the series at the selected locations that can be used to matched the selected time series across locations.
+Subsequently, a set of temporal features can be calculated for the series at the selected locations that can be used to match the selected time series across locations.
 In cubble, locations from two datasets can be matched using the function \code{match_spatial()}.
-The function on calculates the distance matrix of the locations between the two data sets and returns  groups (\code{spatial_n_group}) with the smallest distances.
-For a given group, it is possible can include more locations with the argument \code{spatial_n_each} (default to 1 for one-on-one matching).
+The function calculates the distance matrix of the locations between the two data sets and returns  groups (\code{spatial_n_group}) with the smallest distances.
+For a given group, it is possible to include more locations with the argument \code{spatial_n_each} (default to 1 for one-on-one matching).
 
@@
 
-Five examples are chosen to illustrate different aspects of the \pkg{cubble} package: creating a \code{cubble} object from two Coronavirus (COVID) data tables with the challenge of having different location names, using spatial transformations to make a glyph map of seasonal temperature changes, matching river level data with weather station records to analyze water supply, reading NetCDF format data to replicate a climate reanalysis plot, and demonstrating the workflow to create interactively linked plots.
+Five examples are chosen to illustrate different aspects of the \pkg{cubble} package: (1) creating a \code{cubble} object from two Coronavirus (COVID) data tables with the challenge of having different location names, (2) using spatial transformations to make a glyph map of seasonal temperature changes, (3) matching river level data with weather station records to analyze water supply, (4) reading NetCDF format data to replicate a climate reanalysis plot, and (5) demonstrating the workflow to create interactively linked plots.
 
@@
 As is common, the different agencies have some difference in text id's labelling the spatial regions, so this example illustrates how this can be caught with \pkg{cubble}, and fixed.
-Discrepancies are flagged when creating the cubble object, notifying analysts of something that needs checking.
+Discrepancies are flagged when creating the cubble object (see warnings in the R output below), notifying analysts of something that needs checking.
 
@@
 
-The first step is to perform a spatial match on the site locations between both data sets.
+Ideal matches would exhibit a spatial proximity and also a temporal similarity, with a potential offset due to translating the rainfall into water level.
+This motivates the two step approach where a spatial match on the site locations is first performed between both data sets.
 The rainfall recorded at a particular station can directly impact the water level in nearby rivers, which emphasizes the need to find related or matching locations.
@@
                         spatial_n_group = 10)
-print(res_sp, n = 20)
+print(res_sp)
 ```
@@
 ```{r matching, out.width="100%", fig.height = 5, fig.width = 10, fig.cap="Example of matching weather stations and river gauges.
-These four stations show on the map (a) and time (b) would be considered to be matching.
+These four station pairs shown on the map (a) and as time series plots (b) would be considered to be matching.
 Precipitation and water level have been standardised between 0 and 1 to be displayed in the same scale in (b).
-The peaks in the time series roughly match, and would reflect percipitation increasing water levels."}
+The peaks in the time series roughly match, and would reflect precipitation increasing water levels."}
 res_tm <- res_tm |>  filter(group %in% c(1, 7, 6, 9))
@@
 The maximum and minimum temperatures are shown using a ribbon.
-The bottom row first selects the lowest temperature in August in the seasonal display, which highlights the corresponding station on the map (Thredbo).
+The bottom row first selects the lowest temperature in August in the seasonal display from the first row, which highlights the corresponding station on the map (Thredbo).
 Another  station, located in the Tasmania Island, is then selected to compare its temperature variation with the Thredbo station.", fig.show='hold', purl=FALSE}
@@
 Considerable effort has been exerted to provide tidy tools for models, with the \pkg{tidymodels} project.
-It provides a much better interface to the vast array of statistical and machine learning model architecture.
+It provides the infrastructural interface to build vast array of statistical and machine learning models.
 A similar direction would be to extend cubble to provide a more unified interface to spatio-temporal modeling.

### 2024-01-02 cubble.tex
work on review round 2

--- 2023-07-07 cubble.tex
@@
 H.
-Sherry Zhang\\Monash University \And Dianne Cook\\Monash University \And Ursula Laa\\University of Natural \\ Resources and Life Sciences \AND Nicolas Langrené\\BNU-HKBU \\ United International College \And Patricia Menéndez\\Monash University
+Sherry Zhang\\Monash University \And Dianne Cook\\Monash University \And Ursula Laa\\University of Natural \\ Resources and Life Sciences \AND Nicolas Langrené\\BNU-HKBU \\ United International College \And Patricia Menéndez\\University of Melbourne
 }
@@
     Monash University\\
-    21 Chancellors Walk, Clayton VIC 3800 Australia\\
-  E-mail: \href{mailto:patricia.menendez@monash.edu}{\nolinkurl{patricia.menendez@monash.edu}}\\
+    813 Swanston Street, Parkville VIC 3052, Australia\\
+  E-mail: \href{mailto:Patricia.menendez@unimelb.edu.au}{\nolinkurl{Patricia.menendez@unimelb.edu.au}}\\
 
@@
 
-Spatio-temporal data has a spatial component referring to the location of each observation and a temporal component that is recorded at regular or irregular time intervals.
+Spatio-temporal data \citep{bivand2008applied, lovelace2019geocomputation, pebesma2019spatial} has a spatial component referring to the location of each observation and a temporal component that is recorded at regular or irregular time intervals.
 It may also include multiple variables measured at each spatial and temporal values.
@@
 An example of how the \pkg{cubble} package handles Network Common Data Form (NetCDF) data is also provided.
-Section \ref{conclude} discuss the paper contributions and future directions.
+Section \ref{conclude} discusses the paper contributions and future directions.
 
@@
 
-\hypertarget{object}{%
-\subsection{The cubble class}\label{object}}
+\hypertarget{sec-cb-object}{%
+\subsection{The cubble class}\label{sec-cb-object}}
 
@@
 \begin{CodeInput}
+R> stations
+\end{CodeInput}
+\begin{CodeOutput}
+# A tibble: 3 x 6
+  id           long   lat  elev name              wmo_id
+  <chr>       <dbl> <dbl> <dbl> <chr>              <dbl>
+1 ASN00086038  145. -37.7  78.4 essendon airport   95866
+2 ASN00086077  145. -38.0  12.1 moorabbin airport  94870
+3 ASN00086282  145. -37.7 113.  melbourne airport  94866
+\end{CodeOutput}
+\begin{CodeInput}
+R> meteo
+\end{CodeInput}
+\begin{CodeOutput}
+# A tibble: 30 x 5
+  id          date        prcp  tmax  tmin
+  <chr>       <date>     <dbl> <dbl> <dbl>
+1 ASN00086038 2020-01-01     0  26.8  11  
+2 ASN00086038 2020-01-02     0  26.3  12.2
+3 ASN00086038 2020-01-03     0  34.5  12.7
+4 ASN00086038 2020-01-04     0  29.3  18.8
+5 ASN00086038 2020-01-05    18  16.1  12.5
+# i 25 more rows
+\end{CodeOutput}
+\begin{CodeInput}
 R> make_cubble(spatial = stations, temporal = meteo,
@@
 The \pkg{cubble} package has several functions implemented for data wrangling and to facilitate data analysis as summarized in Table \ref{tab:funs}.
-In addition, for each of the three cubble classes there are number of methods implemented that facilitates the handling of the data as shown in Table \ref{tab:methods}.
+In addition, for each of the three cubble classes there are a number of methods implemented that facilitates the handling of the data as shown in Table \ref{tab:methods}.
 In particular, the \code{cubble_df} class handles methods that behave consistently in both spatial and temporal cubble.
@@
 The pair of cubble verbs, \code{face_temporal()} and \code{face_spatial()}, pivots the cubble object between its two forms or faces, as illustrated in Figure \ref{fig:face}.
-The code applies \code{face_temporal()} on the spatial cubble, \code{cb_spatial}, introduced in Section \ref{object} to get a temporal cubble:
+The code applies \code{face_temporal()} on the spatial cubble, \code{cb_spatial}, introduced in Section \ref{cb-object} to get a temporal cubble:
 
@@
 \caption{Illustration of main functions.
-To focus on he temporal variables \code{face\_temporal()} converts a spatial cubble into a temporal cubble.
+To focus on the temporal variables \code{face\_temporal()} converts a spatial cubble into a temporal cubble.
 To focus on the spatial variables \code{face\_spatial()} transforms a temporal cubble into a spatial cubble.
@@
 The \pkg{stars} package \citep{stars} uses an array structure, as opposed to a tibble, to represent multivariate spatio-temporal data.
-While both \pkg{stars} and \pkg{cubble} support vector and raster data, it is a matter of choice on which structure to use given the application.
+While both \pkg{stars} and \pkg{cubble} support vector and raster data, it is a matter of choice which structure to use given the application.
 Analysts working on satellite imageries may prefer the array structure in \pkg{stars}, while others originally working with spatio-temporal data in 2D data frames may find \pkg{cubble} easier to adopt from their existing computing workflow.
@@
 Matching temporal data across different locations from different data sources could be done by initially identifying the corresponding spatial locations between the two data sets.
-Subsequently, a set of temporal features can be calculated for the series at the selected locations that can be used to matched the selected time series across locations.
+Subsequently, a set of temporal features can be calculated for the series at the selected locations that can be used to match the selected time series across locations.
 In cubble, locations from two datasets can be matched using the function \code{match_spatial()}.
-The function on calculates the distance matrix of the locations between the two data sets and returns groups (\code{spatial_n_group}) with the smallest distances.
-For a given group, it is possible can include more locations with the argument \code{spatial_n_each} (default to 1 for one-on-one matching).
+The function calculates the distance matrix of the locations between the two data sets and returns groups (\code{spatial_n_group}) with the smallest distances.
+For a given group, it is possible to include more locations with the argument \code{spatial_n_each} (default to 1 for one-on-one matching).
 
@@
 [1] "Interstate" "Overseas"   "Unknown"
+
+attr(,"class")
+[1] "key_tbl" "list"   
 \end{CodeOutput}
@@
 +                         spatial_n_group = 10)
-R> print(res_sp, n = 20)
+R> print(res_sp)
 \end{CodeInput}
@@
 5 ASN00085298 226027 4207.     5
- 6 ASN00082042 405234 6153.     6
- 7 ASN00086038 230200 6167.     7
- 8 ASN00086282 230200 6928.     8
- 9 ASN00085279 224217 7431.     9
-10 ASN00080091 406756 7460.    10
+# i 5 more rows
 \end{CodeOutput}
@@
 \caption[Example of matching weather stations and river gauges]{Example of matching weather stations and river gauges.
-These four stations show on the map (a) and time (b) would be considered to be matching.
+These four station pairs shown on the map (a) and as time series plots (b) would be considered to be matching.
 Precipitation and water level have been standardised between 0 and 1 to be displayed in the same scale in (b).
-The peaks in the time series roughly match, and would reflect percipitation increasing water levels.}\label{fig:matching}
+The peaks in the time series roughly match, and would reflect precipitation increasing water levels.}\label{fig:matching}
 \end{figure}
@@
 The maximum and minimum temperatures are shown using a ribbon.
-The bottom row first selects the lowest temperature in August in the seasonal display, which highlights the corresponding station on the map (Thredbo).
+The bottom row first selects the lowest temperature in August in the seasonal display from the first row, which highlights the corresponding station on the map (Thredbo).
 Another  station, located in the Tasmania Island, is then selected to compare its temperature variation with the Thredbo station.}\label{fig:interactive-linking}

### 2024-01-02 review/B-response-V2.Rmd
work on review round 2

---
title: "Response to Reviewers - Article 4780"
subtitle: "cubble: An R Package for Organizing and Wrangling Multivariate Spatio-temporal Data"
author: "H. Sherry Zhang, Dianne Cook, Ursula Laa, Nicolas Langrené, Patricia Menéndez"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We provide an overall summary of the changes made to the paper and the package and then we make a point-by-point explanation of how we addressed the reviewers' comments.

# Response to B-review_2\_2

### Sec. 1
  * Typo: “Section 5 discuss**es** the paper” **\textcolor{purple}{Fixed.}**

### Sec. 2.2
  * Consider to also show the print or str of the input data sets station and meteo to make clear how the connection is established between both tables. **\textcolor{purple}{The input data are now printed.}**
  * Typo on page 5: “there are **a** number of methods” **\textcolor{purple}{Fixed.}**

### Sec. 2.3
  * Fig. 1: The time dimension runs backwards in the data cubes, that might be irritating. **\textcolor{purple}{Fixed.}**
  * Typo in caption: “To focus on **t**he temporal” **\textcolor{purple}{Fixed.}**

### Sec. 2.5
  * Typo: “it is a matter of choice ~~on~~ which structure to use given the application” **\textcolor{purple}{Fixed.}**

### Sec. 3.1: 
  * Typo: “that can be used to match~~ed~~ the selected time series across locations. **\textcolor{purple}{Fixed.}**
  * Typo: “The function ~~on~~ calculates” **\textcolor{purple}{Fixed.}**
  * Typo: “it is possible to ~~can~~ include” **\textcolor{purple}{Fixed.}**
  * Is a matching based on index (date and time) only also possible in cubble?

### Sec. 4: 
  * The first paragraph is hard to read. It could possibly be restructure as a numbered list, or description listing, or simply adding roman numbers in the text, e.g.: “Five examples are chosen to illustrate different aspects of the cubble package: (i) creating a cubble object from two Coronavirus (COVID) data tables with the challenge of having different location names, (ii) using spatial transformations ... ” **\textcolor{purple}{Done.}**

  * It is nice to mention the challenges of all examples, as interested users of cubble could directly jump to the relevant example for their own use case.

### Sec. 4.1
  * I suggest to be even more explicit about warnings to make clear, that they are expected and not due to malfunctioning code: “Discrepancies are flagged (see warnings in the R output below) when creating the cubble object.” **\textcolor{purple}{Done.}**

### Sec. 4.3

At the very beginning of the example, please clearly motivate the two-step approach. **\textcolor{purple}{Added. The two-step approach is motivated because ideal matches require both spatial proximity and temporal similarity in trend.}** I thought I had understood it, but am again puzzled. What I understand is that you pick (i) the 10 closest pairs of met stations and river stations and then (ii) additionally require a temporal similarity (in terms of peaks). But how is the temporal offset treated and controlled in these cases? **\textcolor{purple}{see dot point 3 below}** I also see some logic in an approach where you (i) select 10 spatially closest met stations per river station (i.e. setting `spatial_n_each` = 10) and (ii) pick the highest data similarity between met and river stations within each group. Here, it might be possible, that close-by met and river stations are less related than apparently further apart stations due to local terrain and catchment boundaries. The motivation of the example is crucial and might raise awareness for unintended matches in individual user data. Please clearly describe what this example is meant to illustrate to the reader and potential user of cubble. **\textcolor{purple}{The proposed strategy to pick pairs has been considered, which motivates the arguemnt spaital\_n\_each.}**

  * Note that R> print(`res_sp`, n = 20) only shows ten rows as `res_sp` only has ten at that point in the script. I’d suggest to drop that argument n = 20 here. **\textcolor{purple}{Done.}**
  * After the use of `bind_rows()` the matched stations are stored in rows 2i-1 and 2i, i = 1, . . . , 8. I believe that this notion of how to write pairs in the cubble should be mentioned and explained.
   * How is temporal offset handled, due to delay in the river network from upstream precipitation. Can that be controlled during the temporal matching? Can that be illustrated in the figure? **\textcolor{purple}{This can be controlled by the arguemnt temporal\_window, default to 5. The figure shows 1 year worth of daily data and an offset of 5 days would be indifferentiable in the plot. }**
   * Fig. 4 caption:
      * Typo: “These four station **pairs** show**n** on the map (a) and **as** time **series plots** (b)” **\textcolor{purple}{Fixed.}**
       * Typo: “reflect ~~percipitation~~ precipitation” **\textcolor{purple}{Fixed.}**

### Sec. 4.5:
  * Fig. 6: “The bottom row first selects the lowest temperature in August in the seasonal display **from the first row**” **\textcolor{purple}{Fixed.}**

### Sec. 5:
  * “It provides a much better interface to the vast array of statistical and machine learning model architecture.” Much better than ..? **\textcolor{purple}{Changed to: It provides the infrastructural interface to build ...}**

### 2024-01-04 cubble.Rmd
review response and revision changes made

@@
 
-The Bureau of Meteorology collects water level data that can be matched with precipitation data from climate weather stations.
+This example illustrates spatio-temporal matching of water level data from stream sensors with precipitation data from climate weather stations.
 The data \code{river} from the \pkg{cubble} package contains water course level data for 71 river gauges collected in Victoria, Australia.
-Victoria weather station data can be subsetted from the \code{climate_aus} data in the \pkg{cubble} package.
+Victoria weather station data can be extracted from the \code{climate_aus} data in the \pkg{cubble} package.
 This example demonstrates the use of the matching function introduced in Section \ref{matching} to find river gauges that mirror changes in precipitation regimes captured in the climate weather stations in Victoria.
@@
 
-Ideal matches would exhibit a spatial proximity and also a temporal similarity, with a potential offset due to translating the rainfall into water level.
-This motivates the two step approach where a spatial match on the site locations is first performed between both data sets.
-The rainfall recorded at a particular station can directly impact the water level in nearby rivers, which emphasizes the need to find related or matching locations.
-With \code{match_spatial()}, we can obtain a summary of the 10 closest pairs of weather stations and river gauges:
+We assume that matches would be close spatially and have a similar temporal pattern, accounting for scale (volume of rainfall relative to volume of stream capacity) and temporal lags (time to take water to travel to location).
+This motivates a two-step approach: spatial match is first performed between both data sets, followed by a temporal pattern match on a reduced set of locations.
+
+With \code{match_spatial()}, we can obtain a summary of the closest pairs (here it is 10) of weather stations and river gauges to print or as a list of matched cubbles (if \code{return_cubble = TRUE}).
+The following shows the print results:
 
@@
 
-The results can also be returned as a list of matched cubbles, by setting the argument \code{return_cubble = TRUE}.
-After excluding the two pairs where a river station is matched to more than one weather stations (river station 226027 is matched twice in group 3 and 5 and similarly for station 230200 in group 7 and 8), all the results can be combined into a single cubble using \code{bind_rows()},
+Notice from the printed list that river station 226027 matches two weather stations.
+This is possible, and is why printing the matching results first is preferable.
+The user would need to decide how to handle the multiple matches, for example by removing one or aggregating results, because a cubble requires unique IDs.
+This example creates the list of matched cubbles after excluding the two pairs where a river station is matched to more than one weather station (river station 226027 is matched twice in group 3 and 5 and similarly for station 230200 in group 7 and 8).
+The \code{bind_rows} function is used to TODO
 
@@
 ```
+
+It is possible to keep multiple matches in order to find the best temporal match among the set using the \code{spatial_n_each} argument.
+In combination with \code{spatial_n_group} more complicated matched pairs can be provided to the temporal matching.
+For example, if  \code{spatial_n_each=4} and \code{spatial_n_group=2}, two groups each with the closest four spatial neighbors would be created.
+The temporal matching would follow on each group.
 
@@
 
-Possible future directions would be in two main directions: handling much larger data sets, and integrating smoothly with modeling.
-Spatio-temporal data can be huge, and the current recommendation is to first reduce the spatial and temporal components before constructing a cubble.
-Better support could be provided with additional pre-processing functions.
-Considerable effort has been exerted to provide tidy tools for models, with the \pkg{tidymodels} project.
-It provides the infrastructural interface to build vast array of statistical and machine learning models.
-A similar direction would be to extend cubble to provide a more unified interface to spatio-temporal modeling.
+Possible future improvements would be in two main directions: handling much larger data, and integrating smoothly with modeling.
+Our work has been designed to make it easier to focus on intricate space-time patterns, and it expects that the analyst first reduces the data if the spatio-temporal data is huge, before creating a cubble.
+Additional tools for pre-processing large data would be helpful.
+Because the cubble structure is based on tidy data principles, it potentially integrates well with tidy tools for models, as provided by the \pkg{tidymodels} project.
+Although that project focuses on providing an interface to the vast array of statistical and machine learning models, it would be useful to extend its capacity to provide a more unified interface to many spatio-temporal modeling software.
 

### 2024-01-04 cubble.tex
review response and revision changes made

@@
 
+It is possible to keep multiple matches in order to find the best temporal match among the set using the \code{spatial_n_each} argument.
+In combination with \code{spatial_n_group} more complicated matched pairs can be provided to the temporal matching.
+For example, if \code{spatial_n_each=4} and \code{spatial_n_group=2}, two groups each with the closest four spatial neighbors would be created.
+The temporal matching would follow on each group.
+
 To match the water level and precipitation time series across the matched locations, the function \code{match_temporal()} is used with the variables \code{group} and \code{type} identifying the matching group and the two data sources:

### 2024-01-04 review/B-response-V2.Rmd
review response and revision changes made

@@
   * Is a matching based on index (date and time) only also possible in cubble?
+**It can handle matching for any of the temporal objects on which cubble is based, see section 2.1.
+No changes made to the paper.**
 
@@
   * It is nice to mention the challenges of all examples, as interested users of cubble could directly jump to the relevant example for their own use case.
+**\textcolor{purple}{No change made, current list at top of section 4 should be sufficient for readers to jump to the example that is most relevant for them.}**
 
@@
 At the very beginning of the example, please clearly motivate the two-step approach.
-**\textcolor{purple}{Added.
-The two-step approach is motivated because ideal matches require both spatial proximity and temporal similarity in trend.}** I thought I had understood it, but am again puzzled.
+I thought I had understood it, but am again puzzled.
 What I understand is that you pick (i) the 10 closest pairs of met stations and river stations and then (ii) additionally require a temporal similarity (in terms of peaks).
 But how is the temporal offset treated and controlled in these cases?
-**\textcolor{purple}{see dot point 3 below}** I also see some logic in an approach where you (i) select 10 spatially closest met stations per river station (i.e.
+I also see some logic in an approach where you (i) select 10 spatially closest met stations per river station (i.e.
 setting `spatial_n_each` = 10) and (ii) pick the highest data similarity between met and river stations within each group.
@@
 Please clearly describe what this example is meant to illustrate to the reader and potential user of cubble.
-**\textcolor{purple}{The proposed strategy to pick pairs has been considered, which motivates the arguemnt spaital\_n\_each.}**
+**\textcolor{purple}{We have clarified the text in this section to make it clearer.
+Yes, you are right, that searches can accommodate a broader spatial neighborhood using the spatial\_n\_group argument, and multiple matches can be accommodated using the spatial\_n\_each argument.
+The revised text mentions these more complicated matching possibilities.}**
 
@@
   * “It provides a much better interface to the vast array of statistical and machine learning model architecture.” Much better than ..?
-**\textcolor{purple}{Changed to: It provides the infrastructural interface to build ...}**
+**\textcolor{purple}{Paragraph has been  revised, and is now much clearer.}**

### 2024-01-10 cubble.Rmd
revision done

@@
 This is possible, and is why printing the matching results first is preferable.
-The user would need to decide how to handle the multiple matches, for example by removing one or aggregating results, because a cubble requires unique IDs.
+It is possible to keep multiple matches in order to find the best temporal match among the set using the \code{spatial_n_each} argument.
+In combination with \code{spatial_n_group} more complicated matched pairs can be provided to the temporal matching.
+For example, if \code{spatial_n_each=4} and \code{spatial_n_group=2}, two groups each with the closest four spatial neighbors would be created.
+
+The argument \code{return_cubble = TRUE} can be used to return the match as a list of matched cubbles.
+This list is of length \code{saptial_n_group} and each cubble element has 2x\code{saptial_n_each} rows.
+The subsequent temporal matches can be then performed on each cubble element using the \code{purrr::map()} or \code{lapply()} syntax.
+
+In the case of matched pairs (\code{spatial_n_each = 1} as default here), the list can be converted to a single cubble with the \code{bind_rows()} function.
+The matched pair/group $i$ will be stored in row $2i-1$ and $2i$ in the augmented cubble.
+However, the user would need to decide how to handle the multiple matches, as observed in this example, for example by removing one or aggregating results, because a cubble requires unique IDs.
 This example creates the list of matched cubbles after excluding the two pairs where a river station is matched to more than one weather station (river station 226027 is matched twice in group 3 and 5 and similarly for station 230200 in group 7 and 8).
-The \code{bind_rows} function is used to TODO
+The temporal matching would follow on each pair/group $i$.
 
@@
 
-It is possible to keep multiple matches in order to find the best temporal match among the set using the \code{spatial_n_each} argument.
-In combination with \code{spatial_n_group} more complicated matched pairs can be provided to the temporal matching.
-For example, if  \code{spatial_n_each=4} and \code{spatial_n_group=2}, two groups each with the closest four spatial neighbors would be created.
-The temporal matching would follow on each group.
-
-To match the water level and precipitation time series across the matched locations, the function \code{match_temporal()} is used with the variables \code{group} and \code{type} identifying the matching group and the two data sources:
+To match the water level and precipitation time series across the matched locations, the function \code{match_temporal()} is used with the variables \code{group} and \code{type} identifying the matching group $i$, and the two data sources:
 

### 2024-01-10 cubble.tex
revision done

@@
 This is possible, and is why printing the matching results first is preferable.
-The user would need to decide how to handle the multiple matches, for example by removing one or aggregating results, because a cubble requires unique IDs.
+It is possible to keep multiple matches in order to find the best temporal match among the set using the \code{spatial_n_each} argument.
+In combination with \code{spatial_n_group} more complicated matched pairs can be provided to the temporal matching.
+For example, if \code{spatial_n_each=4} and \code{spatial_n_group=2}, two groups each with the closest four spatial neighbors would be created.
+
+The argument \code{return_cubble = TRUE} can be used to return the match as a list of matched cubbles.
+This list is of length \code{saptial_n_group} and each cubble element has 2x\code{saptial_n_each} rows.
+The subsequent temporal matches can be then performed on each cubble element using the \code{purrr::map()} or \code{lapply()} syntax.
+
+In the case of matched pairs (\code{spatial_n_each = 1} as default here), the list can be converted to a single cubble with the \code{bind_rows()} function.
+The matched pair/group \(i\) will be stored in row \(2i-1\) and \(2i\) in the augmented cubble.
+However, the user would need to decide how to handle the multiple matches, as observed in this example, for example by removing one or aggregating results, because a cubble requires unique IDs.
 This example creates the list of matched cubbles after excluding the two pairs where a river station is matched to more than one weather station (river station 226027 is matched twice in group 3 and 5 and similarly for station 230200 in group 7 and 8).
-The \code{bind_rows} function is used to TODO
+The temporal matching would follow on each pair/group \(i\).
 
@@
 
-It is possible to keep multiple matches in order to find the best temporal match among the set using the \code{spatial_n_each} argument.
-In combination with \code{spatial_n_group} more complicated matched pairs can be provided to the temporal matching.
-For example, if \code{spatial_n_each=4} and \code{spatial_n_group=2}, two groups each with the closest four spatial neighbors would be created.
-The temporal matching would follow on each group.
-
-To match the water level and precipitation time series across the matched locations, the function \code{match_temporal()} is used with the variables \code{group} and \code{type} identifying the matching group and the two data sources:
+To match the water level and precipitation time series across the matched locations, the function \code{match_temporal()} is used with the variables \code{group} and \code{type} identifying the matching group \(i\), and the two data sources:
 

### 2024-01-10 review/B-response-V2.Rmd
revision done

@@
   * Is a matching based on index (date and time) only also possible in cubble?
-**It can handle matching for any of the temporal objects on which cubble is based, see section 2.1.
-No changes made to the paper.**
+**\textcolor{purple}{It can handle matching for any of the temporal objects on which cubble is based, see section 2.1.
+No changes made to the paper.}**
 
@@
 I believe that this notion of how to write pairs in the cubble should be mentioned and explained.
+**\textcolor{purple}{We have expanded the text to explain the data structure of matched stations in both the list and cubble format (after the use of bind\_rows()).}**
    * How is temporal offset handled, due to delay in the river network from upstream precipitation.

### 2024-01-11 cubble.Rmd
modify affiliation and address for everyone

@@
     affiliation: Monash University
-    address: 21 Chancellors Walk, Clayton VIC 3800 Australia
+    address: Department of Econometrics and Business Statistics, Monash University, VIC 3800 Australia
     email: \email{huize.zhang@monash.edu}
@@
     affiliation: Monash University
-    address: 21 Chancellors Walk, Clayton VIC 3800 Australia
+    address: Department of Econometrics and Business Statistics, Monash University, VIC 3800 Australia
     email: dicook@monash.edu
@@
     affiliation: 'University of  Natural `\\`{=latex} Resources and Life Sciences \AND'
-    address: Gregor-Mendel-Straße 33, 1180 Wien, Austria
+    address: Institute of Statistics, University of  Natural Resources and Life Sciences, Peter-Jordan-Straße 82/I 1190 Vienna
     email:  ursula.laa@boku.ac.at
   - name: Nicolas Langrené
-    affiliation: 'BNU-HKBU `\\`{=latex} United International College'
-    address: 2000 Jintong Road, Tangjiawan, Zhuhai, Guangdong Province, China
+    affiliation: 'BNU-HKBU United International College'
+    address:  Guangdong Provincial Key Laboratory of Interdisciplinary Research and Application for Data Science, BNU-HKBU United International College, Zhuhai 519087 China
     email: nicolaslangrene@uic.edu.cn
@@
     affiliation2: Monash University
-    address: 813 Swanston Street, Parkville VIC 3052, Australia
+    address: School of Mathematics and Statistics, University of Melbourne, VIC 3052 Australia
     email:  Patricia.menendez@unimelb.edu.au
@@
 This work is funded by a Commonwealth Scientific and Industrial Research Organisation (CSIRO) Data61 Scholarship and started while Nicolas Langrené was affiliated with CSIRO's Data61.
+Nicolas Langrené acknowledges the partial support of the Guangdong Provincial Key Laboratory IRADS (2022B1212010006, R0400001-22) and the UIC Start-up Research Fund UICR0700041-22.
 The article is created using the package \pkg{knitr} [@knitr] and \pkg{rmarkdown} [@rmarkdown] in \proglang{R} with the \code{rticles::jss_article} template.

### 2024-01-11 cubble.tex
modify affiliation and address for everyone

@@
 H.
-Sherry Zhang\\Monash University \And Dianne Cook\\Monash University \And Ursula Laa\\University of Natural \\ Resources and Life Sciences \AND Nicolas Langrené\\BNU-HKBU \\ United International College \And Patricia Menéndez\\University of Melbourne
+Sherry Zhang\\Monash University \And Dianne Cook\\Monash University \And Ursula Laa\\University of Natural \\ Resources and Life Sciences \AND Nicolas Langrené\\BNU-HKBU United International College \And Patricia Menéndez\\University of Melbourne
 }
@@
     Monash University\\
-    21 Chancellors Walk, Clayton VIC 3800 Australia\\
+    Department of Econometrics and Business Statistics, Monash University, VIC 3800 Australia\\
   E-mail: \email{huize.zhang@monash.edu}\\
@@
     Monash University\\
-    21 Chancellors Walk, Clayton VIC 3800 Australia\\
+    Department of Econometrics and Business Statistics, Monash University, VIC 3800 Australia\\
   E-mail: \href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}\\
@@
     University of Natural \\ Resources and Life Sciences\\
-    Gregor-Mendel-Straße 33, 1180 Wien, Austria\\
+    Institute of Statistics, University of Natural Resources and Life Sciences, Peter-Jordan-Straße 82/I 1190 Vienna\\
   E-mail: \href{mailto:ursula.laa@boku.ac.at}{\nolinkurl{ursula.laa@boku.ac.at}}\\
@@
       Nicolas Langrené\\
-    BNU-HKBU \\ United International College\\
-    2000 Jintong Road, Tangjiawan, Zhuhai, Guangdong Province, China\\
+    BNU-HKBU United International College\\
+    Guangdong Provincial Key Laboratory of Interdisciplinary Research and Application for Data Science, BNU-HKBU United International College, Zhuhai 519087 China\\
   E-mail: \href{mailto:nicolaslangrene@uic.edu.cn}{\nolinkurl{nicolaslangrene@uic.edu.cn}}\\
@@
     Monash University\\
-    813 Swanston Street, Parkville VIC 3052, Australia\\
+    School of Mathematics and Statistics, University of Melbourne, VIC 3052 Australia\\
   E-mail: \href{mailto:Patricia.menendez@unimelb.edu.au}{\nolinkurl{Patricia.menendez@unimelb.edu.au}}\\
@@
 This work is funded by a Commonwealth Scientific and Industrial Research Organisation (CSIRO) Data61 Scholarship and started while Nicolas Langrené was affiliated with CSIRO's Data61.
+Nicolas Langrené acknowledges the partial support of the Guangdong Provincial Key Laboratory IRADS (2022B1212010006, R0400001-22) and the UIC Start-up Research Fund UICR0700041-22.
 The article is created using the package \pkg{knitr} \citep{knitr} and \pkg{rmarkdown} \citep{rmarkdown} in \proglang{R} with the \code{rticles::jss_article} template.

### 2024-01-11 review/B-response-V2.Rmd
respnse changed

@@
 
-We provide an overall summary of the changes made to the paper and the package and then we make a point-by-point explanation of how we addressed the reviewers' comments.
+We appreciate the helpful and careful reviews. This is a point-by-point explanation of how we modified the paper to address the comments.
 
@@
   * It is nice to mention the challenges of all examples, as interested users of cubble could directly jump to the relevant example for their own use case.
-**\textcolor{purple}{No change made, current list at top of section 4 should be sufficient for readers to jump to the example that is most relevant for them.}**
+**\textcolor{purple}{No change made, paragraph at top of section 4 should be sufficient for readers to jump to the example that is most relevant for them.}**
 

### 2024-01-16 cubble.Rmd
modify patricia's affiliation

--- 2024-01-10 cubble.Rmd
@@
     affiliation: 'University of Natural `\\`{=latex} Resources and Life Sciences \AND'
-    address: Gregor-Mendel-Straße 33, 1180 Wien, Austria
+    address: Institute of Statistics, University of  Natural Resources and Life Sciences, Peter-Jordan-Straße 82/I 1190 Vienna
     email:  ursula.laa@boku.ac.at
   - name: Nicolas Langrené
-    affiliation: 'BNU-HKBU `\\`{=latex} United International College'
-    address: 2000 Jintong Road, Tangjiawan, Zhuhai, Guangdong Province, China
+    affiliation: 'BNU-HKBU United International College'
+    address:  Guangdong Provincial Key Laboratory of Interdisciplinary Research and Application for Data Science, BNU-HKBU United International College, Zhuhai 519087 China
     email: nicolaslangrene@uic.edu.cn
   - name: Patricia Menéndez
-    affiliation: University of Melbourne
-    affiliation2: Monash University
-    address: 813 Swanston Street, Parkville VIC 3052, Australia
+    affiliation: 'University of Melbourne `\\`{=latex} Monash University'
+    address: School of Mathematics and Statistics, University of Melbourne, VIC 3052 Australia
     email:  Patricia.menendez@unimelb.edu.au

### 2024-01-16 cubble.tex
modify patricia's affiliation

--- 2024-01-11 cubble.tex
@@
 H.
-Sherry Zhang\\Monash University \And Dianne Cook\\Monash University \And Ursula Laa\\University of Natural \\ Resources and Life Sciences \AND Nicolas Langrené\\BNU-HKBU United International College \And Patricia Menéndez\\University of Melbourne
+Sherry Zhang\\Monash University \And Dianne Cook\\Monash University \And Ursula Laa\\University of Natural \\ Resources and Life Sciences \AND Nicolas Langrené\\BNU-HKBU United International College \And Patricia Menéndez\\University of Melbourne \\ Monash University
 }
@@
       Patricia Menéndez\\
-    Monash University\\
+    University of Melbourne \\ Monash University\\
     School of Mathematics and Statistics, University of Melbourne, VIC 3052 Australia\\
